### 2021-10-13 prisc.tex
Add acm template for PRISC and fill in some details

%% For double-blind review submission, w/o CCS and ACM Reference (max submission space)
%\documentclass[sigplan,review,anonymous]{acmart}\settopmatter{printfolios=true,printccs=false,printacmref=false}
%% For double-blind review submission, w/ CCS and ACM Reference
%\documentclass[sigplan,review,anonymous]{acmart}\settopmatter{printfolios=true}
%% For single-blind review submission, w/o CCS and ACM Reference (max submission space)
%\documentclass[sigplan,review]{acmart}\settopmatter{printfolios=true,printccs=false,printacmref=false}
%% For single-blind review submission, w/ CCS and ACM Reference
%\documentclass[sigplan,review]{acmart}\settopmatter{printfolios=true}
%% For final camera-ready submission, w/ required CCS and ACM Reference
\documentclass[sigplan]{acmart}\settopmatter{}
%% "not anonymous"

%% Conference information
%% Supplied to authors by publisher for camera-ready submission;
%% use defaults for review submission.
\acmConference[PL'18]{ACM SIGPLAN Conference on Programming Languages}{January 01--03, 2018}{New York, NY, USA}
\acmYear{2018}
\acmISBN{} % \acmISBN{978-x-xxxx-xxxx-x/YY/MM}
\acmDOI{} % \acmDOI{10.1145/nnnnnnn.nnnnnnn}
\startPage{1}

%% Copyright information
%% Supplied to authors (based on authors' rights management selection;
%% see authors.acm.org) by publisher for camera-ready submission;
%% use 'none' for review submission.
\setcopyright{none}
%\setcopyright{acmcopyright}
%\setcopyright{acmlicensed}
%\setcopyright{rightsretained}
%\copyrightyear{2018}           %% If different from \acmYear

%% Bibliography style
\bibliographystyle{ACM-Reference-Format}
%% Citation style
%\citestyle{acmauthoryear}  %% For author/year citations
%\citestyle{acmnumeric}     %% For numeric citations
%\setcitestyle{nosort}      %% With 'acmnumeric', to disable automatic
                            %% sorting of references within a single citation;
                            %% e.g., \cite{Smith99,Carpenter05,Baker12}
                            %% rendered as [14,5,2] rather than [2,5,14].
%\setcitesyle{nocompress}   %% With 'acmnumeric', to disable automatic
                            %% compression of sequential references within a
                            %% single citation;
                            %% e.g., \cite{Baker12,Baker14,Baker16}
                            %% rendered as [2,3,4] rather than [2-4].

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%% Note: Authors migrating a paper from traditional SIGPLAN
%% proceedings format to PACMPL format must update the
%% '\documentclass' and topmatter commands above; see
%% 'acmart-pacmpl-template.tex'.
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

%% Some recommended packages.
\usepackage{booktabs}   %% For formal tables:
                        %% http://ctan.org/pkg/booktabs
\usepackage{subcaption} %% For complex figures with subfigures/subcaptions
                        %% http://ctan.org/pkg/subcaption

\begin{document}

%% Title information
\title{Composing Secure Compilers}         %% [Short Title] is optional;
                                        %% when present, will be used in
                                        %% header instead of Full Title.
%\titlenote{with title note}             %% \titlenote is optional;
                                        %% can be repeated if necessary;
                                        %% contents suppressed with 'anonymous'
%\subtitle{Subtitle}                     %% \subtitle is optional
%\subtitlenote{with subtitle note}       %% \subtitlenote is optional;
                                        %% can be repeated if necessary;
                                        %% contents suppressed with 'anonymous'

%% Author information
%% Contents and number of authors suppressed with 'anonymous'.
%% Each author should be introduced by \author, followed by
%% \authornote (optional), \orcid (optional), \affiliation, and
%% \email.
%% An author may have multiple affiliations and/or emails; repeat the
%% appropriate command.
%% Many elements are not rendered, but should be provided for metadata
%% extraction tools.

%% Author with single affiliation.
\author{Matthis Kruse}
%\authornote{with author1 note}          %% \authornote is optional;
                                        %% can be repeated if necessary
\orcid{0000-0003-4062-9666}             %% \orcid is optional
\affiliation{
%  \position{Position1}
%  \department{Department1}              %% \department is recommended
  \institution{CISPA Helmholtz Center for Information Security}            %% \institution is required
%  \streetaddress{Street1 Address1}
%  \city{City1}
%  \state{State1}
%  \postcode{Post-Code1}
  \country{Germany}                    %% \country is recommended
}
\email{matthis.kruse@cispa.de}          %% \email is recommended

%% Author with two affiliations and emails.
\author{Marco Patrignani}
%\authornote{with author2 note}          %% \authornote is optional;
                                        %% can be repeated if necessary
\orcid{0000-0003-3411-9678}             %% \orcid is optional
\affiliation{
%  \position{Position2a}
%  \department{Department2a}             %% \department is recommended
  \institution{CISPA Helmholtz Center for Information Security}            %% \institution is required
%  \streetaddress{Street2a Address2a}
%  \city{City2a}
%  \state{State2a}
%  \postcode{Post-Code2a}
  \country{Germany}                   %% \country is recommended
}
\email{marco.patrignani@cispa.de}         %% \email is recommended

%% Abstract
%% Note: \begin{abstract}...\end{abstract} environment must come
%% before \maketitle command
\begin{abstract}
Text of abstract \ldots.
\end{abstract}

%% 2012 ACM Computing Classification System (CSS) concepts
%% Generate at 'http://dl.acm.org/ccs/ccs.cfm'.
\begin{CCSXML}
<ccs2012>
<concept>
<concept_id>10011007.10011006.10011008</concept_id>
<concept_desc>Software and its engineering~General programming languages</concept_desc>
<concept_significance>500</concept_significance>
</concept>
<concept>
<concept_id>10003456.10003457.10003521.10003525</concept_id>
<concept_desc>Social and professional topics~History of programming languages</concept_desc>
<concept_significance>300</concept_significance>
</concept>
</ccs2012>
\end{CCSXML}

\ccsdesc[500]{Software and its engineering~General programming languages}
\ccsdesc[300]{Social and professional topics~History of programming languages}
%% End of generated code

%% Keywords
%% comma separated list
\keywords{keyword1, keyword2, keyword3}  %% \keywords are mandatory in final camera-ready submission

%% \maketitle
%% Note: \maketitle command must come after title commands, author
%% commands, abstract environment, Computing Classification System
%% environment and commands, and keywords command.
\maketitle

\section{Introduction}

Text of paper \ldots

%% Acknowledgments
\begin{acks}                            %% acks environment is optional
                                        %% contents suppressed with 'anonymous'
  %% Commands \grantsponsor{<sponsorID>}{<name>}{<url>} and
  %% \grantnum[<url>]{<sponsorID>}{<number>} should be used to
  %% acknowledge financial support and will be used by metadata
  %% extraction tools.
  This material is based upon work supported by the
  \grantsponsor{GS100000001}{National Science
    Foundation}{http://dx.doi.org/10.13039/100000001} under Grant
  No.~\grantnum{GS100000001}{nnnnnnn} and Grant
  No.~\grantnum{GS100000001}{mmmmmmm}.  Any opinions, findings, and
  conclusions or recommendations expressed in this material are those
  of the author and do not necessarily reflect the views of the
  National Science Foundation.
\end{acks}

%% Bibliography
%\bibliography{bibfile}

%% Appendix
%\appendix
%\section{Appendix}

%Text of appendix \ldots

\end{document}

### 2021-10-17 prisc.tex
Finish up first draft for PRISC

@@
 %% For final camera-ready submission, w/ required CCS and ACM Reference
-\documentclass[sigplan]{acmart}\settopmatter{}
+\documentclass[sigplan,dvipsnames]{acmart}\settopmatter{}
 %% "not anonymous"
@@
 %% use defaults for review submission.
-\acmConference[PL'18]{ACM SIGPLAN Conference on Programming Languages}{January 01--03, 2018}{New York, NY, USA}
-\acmYear{2018}
+\acmConference[PL'22]{ACM SIGPLAN Workshop on Principles of Secure Compilation}{January 22nd, 2022}{Philadelphia, PA, USA}
+\acmYear{2022}
 \acmISBN{} % \acmISBN{978-x-xxxx-xxxx-x/YY/MM}
@@
 %\setcopyright{rightsretained}
-%\copyrightyear{2018}           %% If different from \acmYear
+%\copyrightyear{2022}           %% If different from \acmYear
 
@@
 
+\newcommand\hmmax{0}
+\newcommand\bmmax{0}
+
 %% Some recommended packages.
+\usepackage[colorinlistoftodos]{todonotes}
+\newcommand{\MK}[1]{\todo[color=orange!30]{TODO: #1}}
+\newcommand{\MP}[1]{\todo[color=blue!30]{TODO: #1}}
+
 \usepackage{booktabs}   %% For formal tables:
@@
                         %% http://ctan.org/pkg/subcaption
+\usepackage{stmaryrd}
+\usepackage{mathrsfs}
+\usepackage{mathtools}
+\usepackage{xcolor}
+\usepackage{amsmath}
+\usepackage{amsthm}
+\usepackage{bm}
+\usepackage{cleveref}
+\usepackage{mmmacros}
+\usepackage[switch]{lineno}
+%\modulolinenumbers[2]
+\renewcommand{\linenumberfont}{\normalfont\bfseries\small\color{red}}
 
@@
 %% before \maketitle command
-\begin{abstract}
-Text of abstract \ldots.
-\end{abstract}
+%\begin{abstract}
+%Text of abstract \ldots.
+%\end{abstract}
 
@@
 %% Generate at 'http://dl.acm.org/ccs/ccs.cfm'.
-\begin{CCSXML}
-<ccs2012>
-<concept>
-<concept_id>10011007.10011006.10011008</concept_id>
-<concept_desc>Software and its engineering~General programming languages</concept_desc>
-<concept_significance>500</concept_significance>
-</concept>
-<concept>
-<concept_id>10003456.10003457.10003521.10003525</concept_id>
-<concept_desc>Social and professional topics~History of programming languages</concept_desc>
-<concept_significance>300</concept_significance>
-</concept>
-</ccs2012>
-\end{CCSXML}
+%\begin{CCSXML}
+%<ccs2012>
+%<concept>
+%<concept_id>10011007.10011006.10011008</concept_id>
+%<concept_desc>Software and its engineering~General programming languages</concept_desc>
+%<concept_significance>500</concept_significance>
+%</concept>
+%<concept>
+%<concept_id>10003456.10003457.10003521.10003525</concept_id>
+%<concept_desc>Social and professional topics~History of programming languages</concept_desc>
+%<concept_significance>300</concept_significance>
+%</concept>
+%</ccs2012>
+%\end{CCSXML}
 
-\ccsdesc[500]{Software and its engineering~General programming languages}
-\ccsdesc[300]{Social and professional topics~History of programming languages}
+%\ccsdesc[500]{Software and its engineering~General programming languages}
+%\ccsdesc[300]{Social and professional topics~History of programming languages}
 %% End of generated code
@@
 %% comma separated list
-\keywords{keyword1, keyword2, keyword3}  %% \keywords are mandatory in final camera-ready submission
+%\keywords{compilers, security}  %% \keywords are mandatory in final camera-ready submission
 
@@
 
+\linenumbers
+
 \section{Introduction}
 
-Text of paper \ldots
+Compilers translate from a source to a target programming language.
+Despite their importance when programming and thus for day-to-day programs in general, there are fundamental gaps regarding their preservation of security.
+Secure Compilation studies the preservation of security properties and hyperproperties when compiling a source to a target program.
+Some existing, security-preserving compilers are fully-abstract.
+That is, they reflect and preserve observational equivalence.
+However, fully-abstract compilers may instrument the target code with unnecessary countermeasures not relevant for security.
+Robust Compilation~\cite{abate2019} is a novel compiler property that does not require the compiler to add such unnecessary measures.
+Informally, if a given (partial) source program linked with an arbitrary source context satisfies some properties of interest, so should its compiled version linked with an arbitrary target context satisfy them.
+Even though there exist robust safety-property preserving compilers, they are far from practical.
+Real-world compilers consist of several smaller compilers that are composed with each other in a certain way.
+A prominent example would be any compiler based on the LLVM toolchain~\cite{lattner04}, whose optimization pipeline consists of many passes, which one can view as independent compilers composed to achieve speedups.
+To the best of our knowledge, current work does not discuss which properties are preserved by compilers that use other compilers in their pipeline.
+
+This talk will investigate how different compiler compositions preserve properties, given that the compilers involved in the composition fulfil certain assumptions.
+We define classes of hyperproperties and identify their lattice.
+We prove that our chosen compositions preserve at least the least upper bound within the lattice of such classes.
+Consequently, the order of optimizations in a compilation pipeline does not matter for property preservation, assuming that each such pass preserves the class of interest.
+
+\section{Section II}
+
+\begin{definition}[Classes]
+  A class of hyperproperties $\cC$ is a set of hyperproperties.
+\end{definition}
+\noindent
+From now on, we use $\Pi$ for elements of any class $\cC$.
+\noindent
+\begin{definition}[Programming Languages]\label{def:pl}
+A programming language is a tuple $\left(\partials,\wellf,\singlestep,\linker\right)$ s.t.:
+
+\begin{itemize}
+  \item[$\partials$] : Set - is a set of admissible, partial programs.
+  \item[$\wellf$] : $\partials$ - a judgement that holds iff a program is not partial.
+  \item[$\singlestep$] : $\wholes\to\events\to\wholes$ - a step relation, where $\wholes=\{w\in\partials\ |\ \wellf w\}$.
+        For $e\in\events$ and $p,p'\in\wholes$ we say for $\estep{p}{e}{p'}$ that program $p$ performs a step with action $e$ to program $p'$.
+        If $e=\emptyevent$, we write $\step{p}{p'}$.
+        $\terminationevent$ is the termination event.
+  \item[$\linker$] : $\partials\to\partials\to\partials$ - links two partial programs together in some way, resulting in a new partial program.
+\end{itemize}
+\end{definition}
+\begin{definition}[Traces]
+  A trace $\trace$ is an infinite sequence of events that results from the relation $\singlestep$.
+  That is, we obtain the trace $\trace=\seqcons{e_{0}}{\seqcons{e_{1}}{\dots}}$ for the execution sequence $\estep{p}{e_{0}}{\estep{p'}{e_{1}}{\dots}}$ and write $\mktrace{p}{\trace}$.
+  The set of all traces is $\traces$.
+\end{definition}
+\begin{definition}[Behavior]
+  The behavior of a whole program $p$ is a set of all traces it produces, i.e. $\behav{p}=\{\trace\ |\ \mktrace{p}{\trace}\}$.
+\end{definition}
+\begin{definition}[Hyperproperties]
+  A hyperproperty $H$ is a set of admissible sets of traces. Thus, if $p$ satisfies $H$ (written $\sat{p}{H}$), then $\behav{p}\in H$.
+\end{definition}
+\begin{definition}[Robust Property Satisfaction]
+  A program $p$ robustly satisfies a hyperproperty $\Pi$, written $\rsat{p}{\Pi}$, iff $\forall C\in\partials,\sat{C\linker p}{\Pi}$. The same notation is used for robust hyperproperty satisfaction.
+\end{definition}
+\begin{lemma}[Property Satisfaction Refinement]
+  For a property $\prop$ that refines $\prop'$, i.e. $\prop\subseteq\prop'$, if any $p\in\partials$ satisfies $\sat{p}{\prop'}$, then $\sat{p}{\prop}$.
+\end{lemma}
+
+\begin{definition}[Compilers]
+  A compiler between languages $\S$ and $\T$ is a partial function $\stcomp{\bullet}$ from $\src{\partials}$ to $\trg{\partials}$.
+\end{definition}
+
+\begin{definition}[Robust Trace-Hyperproperty Preservation]\label{def:rtp}
+  For a given class $\cC$, a compiler from languages $\S$ to $\T$ robustly preserves $\cC$ iff
+  $$
+  \forall\Pi\in\cC,\forall\src{p}\in\src{\partials},\rsat{\src{p}}{\Pi}\implies\rsat{\stcomp{\src{p}}}{\Pi}
+  $$
+  We write $\rtp{\stcomp{\bullet}}{\cC}$.
+\end{definition}
+\noindent
+We can, finally, discuss various forms of compositions.
+We begin with the most straightforward one, which coincides with ordinary function composition.
+From a practical perspective, this means plugging the result of one compiler into another one.
+Such pipelines happen when optimizing source code, so at the level of a suitable intermediate representation, but also on a higher level:
+Consider as an example a typical TypeScript compilation pipeline.
+First, it translates TypeScript code to JavaScript, which V8 compiles to Ignition bytecode that, after some runtime-performance analysis, is just-in-time compiled via TurboFan to machine code such as x86.
+
+\begin{definition}[Sequential Composition of Compilers]
+  Given two compilers $\sicomp{\bullet}$ and $\itcomp{\bullet}$, their sequential composition is $\sitcomp{\bullet}=\itcompN{\sicomp{\bullet}}$.
+\end{definition}
+\begin{lemma}[Weakening RTP]\label{lem:weaken}
+  Given classes $\cC_{1}, \cC_{2}$ such that $\cC_{1}\subseteq\cC_{2}$ and $\rtp{\stcomp{\bullet}}{\cC_{2}}$, it holds that $\rtp{\stcomp{\bullet}}{\cC_{1}}$.
+\end{lemma}
+\begin{lemma}[Sequential Composition with RTP]\label{lem:seqcompo}
+  Given $\rtp{\sicomp{\bullet}}{\cC_{1}}$ and $\rtp{\itcomp{\bullet}}{\cC_{2}}$, then $\rtp{\sitcomp{\bullet}}{\cC_{1}\cap\cC_{2}}$.
+\end{lemma}
+
+We now consider a compiler that invokes two other compilers.
+HTML, CSS, and JavaScript are the most obvious example of such a composition.
+One writes down some code of any of these three languages embedded in the same file, but different compiler toolchains process them.
+
+\begin{definition}[Upper Composition]
+  Given two compilers $\stcomp{\bullet}$ and $\itcomp{\bullet}$, their upper composition is
+
+  $$\uhcsitcomp{\bullet}=\lambda p.\begin{cases}\stcomp{p} &\text{if }p\in\src{\partials}\\
+                                                \itcomp{p} &\text{if }p\in\irl{\partials}\end{cases}$$.
+\end{definition}
+\begin{lemma}[Upper Composition with RTP]
+  Given $\rtp{\stcomp{\bullet}}{\cC_{1}}$ and $\rtp{\itcomp{\bullet}}{\cC_{2}}$, then $\rtp{\uhcsitcomp{\bullet}}{\cC_{1}\cap\cC_{2}}$.
+\end{lemma}
+
+With the same idea, we define a dual composition that goes from a single source language to multiple target languages.
+Any compiler used in production is defined that way since one can, as an example, compile C code to both x86 and MIPS assembly.
+We view frameworks such as RubyOnRails also as an example that can emit two different target languages, in this case, HTML and JS, at once.
+\begin{definition}[Lower Composition]
+  Given two compilers $\stcomp{\bullet}$ and $\sicomp{\bullet}$, their lower composition is $\lhcsitcomp{\bullet}$.
+\end{definition}
+
+\begin{lemma}[Lower Composition with RTP]
+  Given $\rtp{\stcomp{\bullet}}{\cC_{1}}$ and $\rtp{\sicomp{\bullet}}{\cC_{2}}$, then $\rtp{\lhcsitcomp{\bullet}}{\cC_{1}\cap\cC_{2}}$.
+\end{lemma}
+
+Using the upper and lower composition, we may end up with a diamond situation, which the following lemma describes.
+Especially from a security researcher perspective, the diamond lemma is attractive as it says that there is no harm to prove compiler passes robust property preserving in a modular way.
+That is, consider some program instrumented against SPECTRE v1 attacks and, in parallel to that, against SPECTRE v4.
+If the compiler instrumentation for v1 does not violate v4, the lemma tells us that combining both instrumentations in some way yields an instrumentation that protects against both v1 and v4.
+
+\begin{lemma}[Diamond]\label{lem:diamond}
+  Given $\rtp{\lhcsiocomp{\bullet}}{\cC_{1}}$ and $\rtp{\uhciotcomp{\bullet}}{\cC_{2}}$ with $\stcomp{\bullet} = \lambda\src{p}.\uhciotcomp{\lhcsiocomp{p}}$, then $\rtp{\stcomp{\bullet}}{\cC_{1}\cap\cC_{2}}$.
+\end{lemma}
+
+The following free theorem is a direct consequence of \Cref{lem:seqcompo} where the involved compilers' input and output are both partial programs in the same language.
+Given that some set of optimization passes is robustly preserving properties, we can combine them in an arbitrary order and preserve the same least upper bound.
+A compilers optimization pipeline ordering is difficult and often hand-tuned.
+The lemma allows us to not care about the particular order of optimizations regarding their robust property preservation.
+So, the compiler developer is free to swap around.
+\begin{lemma}[Swappable]
+  Given $\rtp{\ttcomp{\bullet}_{(1)}}{\cC_{1}}$ and $\rtp{\ttcomp{\bullet}_{(2)}}{\cC_{2}}$, then $\rtp{\ttcompN{\ttcomp{\bullet}_{(2)}}_{(1)}}{\cC_{1}\cap\cC_{2}}$ and $\rtp{\ttcompN{\ttcomp{\bullet}_{(1)}}_{(2)}}{\cC_{1}\cap\cC_{2}}$.
+\end{lemma}
+\MK{This shouldn't work with relational hyperproperties, right?}
+
+{\color{red}\Large TODO: Miro compo}
+
+\section{Conclusion}
+
+We presented some practically relevant compiler compositions and their respective implications for security.
+Furthermore, compiler developers need not worry about optimization order constraints if the involved optimizations are proven to preserve properties of interest robustly.
 
 %% Acknowledgments
-\begin{acks}                            %% acks environment is optional
-                                        %% contents suppressed with 'anonymous'
-  %% Commands \grantsponsor{<sponsorID>}{<name>}{<url>} and
-  %% \grantnum[<url>]{<sponsorID>}{<number>} should be used to
-  %% acknowledge financial support and will be used by metadata
-  %% extraction tools.
-  This material is based upon work supported by the
-  \grantsponsor{GS100000001}{National Science
-    Foundation}{http://dx.doi.org/10.13039/100000001} under Grant
-  No.~\grantnum{GS100000001}{nnnnnnn} and Grant
-  No.~\grantnum{GS100000001}{mmmmmmm}.  Any opinions, findings, and
-  conclusions or recommendations expressed in this material are those
-  of the author and do not necessarily reflect the views of the
-  National Science Foundation.
-\end{acks}
+%\begin{acks}                            %% acks environment is optional
+%                                        %% contents suppressed with 'anonymous'
+%  %% Commands \grantsponsor{<sponsorID>}{<name>}{<url>} and
+%  %% \grantnum[<url>]{<sponsorID>}{<number>} should be used to
+%  %% acknowledge financial support and will be used by metadata
+%  %% extraction tools.
+%  This material is based upon work supported by the
+%  \grantsponsor{GS100000001}{National Science
+%    Foundation}{http://dx.doi.org/10.13039/100000001} under Grant
+%  No.~\grantnum{GS100000001}{nnnnnnn} and Grant
+%  No.~\grantnum{GS100000001}{mmmmmmm}.  Any opinions, findings, and
+%  conclusions or recommendations expressed in this material are those
+%  of the author and do not necessarily reflect the views of the
+%  National Science Foundation.
+%\end{acks}
 
 %% Bibliography
-%\bibliography{bibfile}
+\bibliography{library}
 

### 2021-10-18 prisc.tex
Incorporate most of the feedback.

@@
 %% "not anonymous"
+\settopmatter{printacmref=false} % Removes citation information below abstract
+\renewcommand\footnotetextcopyrightpermission[1]{}
 
@@
 %% use defaults for review submission.
-\acmConference[PL'22]{ACM SIGPLAN Workshop on Principles of Secure Compilation}{January 22nd, 2022}{Philadelphia, PA, USA}
-\acmYear{2022}
-\acmISBN{} % \acmISBN{978-x-xxxx-xxxx-x/YY/MM}
-\acmDOI{} % \acmDOI{10.1145/nnnnnnn.nnnnnnn}
-\startPage{1}
+%\acmConference[PL'22]{ACM SIGPLAN Workshop on Principles of Secure Compilation}{January 22nd, 2022}{Philadelphia, PA, USA}
+%\acmYear{2022}
+%\acmISBN{} % \acmISBN{978-x-xxxx-xxxx-x/YY/MM}
+%\acmDOI{} % \acmDOI{10.1145/nnnnnnn.nnnnnnn}
+%\startPage{1}
 
@@
 
-Compilers translate from a source to a target programming language.
-Despite their importance when programming and thus for day-to-day programs in general, there are fundamental gaps regarding their preservation of security.
+Compilers translate programs from a source to a target programming language.
 Secure Compilation studies the preservation of security properties and hyperproperties when compiling a source to a target program.
-Some existing, security-preserving compilers are fully-abstract.
-That is, they reflect and preserve observational equivalence.
-However, fully-abstract compilers may instrument the target code with unnecessary countermeasures not relevant for security.
 Robust Compilation~\cite{abate2019} is a novel compiler property that does not require the compiler to add such unnecessary measures.
-Informally, if a given (partial) source program linked with an arbitrary source context satisfies some properties of interest, so should its compiled version linked with an arbitrary target context satisfy them.
-Even though there exist robust safety-property preserving compilers, they are far from practical.
+Informally, if a given (partial) source program linked with an arbitrary source context satisfies some properties of interest, so should its compiled version linked with an arbitrary target context.
+Even though there exist robust compilers, they are far from practical.
 Real-world compilers consist of several smaller compilers that are composed with each other in a certain way.
 A prominent example would be any compiler based on the LLVM toolchain~\cite{lattner04}, whose optimization pipeline consists of many passes, which one can view as independent compilers composed to achieve speedups.
-To the best of our knowledge, current work does not discuss which properties are preserved by compilers that use other compilers in their pipeline.
+To the best of our knowledge, current work does not discuss which properties are preserved by compilation chains.
 
-This talk will investigate how different compiler compositions preserve properties, given that the compilers involved in the composition fulfil certain assumptions.
-We define classes of hyperproperties and identify their lattice.
-We prove that our chosen compositions preserve at least the least upper bound within the lattice of such classes.
-Consequently, the order of optimizations in a compilation pipeline does not matter for property preservation, assuming that each such pass preserves the class of interest.
+This paper investigates how different compiler compositions preserve hyperproperties, given that the compilers involved in the composition fulfil certain assumptions.
+To reason about hyperproperty preservation, we will look at the lattice of classes of hyperproperties.
+Within the lattice of such classes, we examine whether at least the least upper bound is preserved by our chosen compositions.
+We conclude with a preliminary result: We show that the order of optimizations in a compilation pipeline does not matter for property preservation, assuming that each such pass preserves the class of interest.
 
-\section{Section II}
-
-\begin{definition}[Classes]
-  A class of hyperproperties $\cC$ is a set of hyperproperties.
-\end{definition}
-\noindent
-From now on, we use $\Pi$ for elements of any class $\cC$.
-\noindent
-\begin{definition}[Programming Languages]\label{def:pl}
-A programming language is a tuple $\left(\partials,\wellf,\singlestep,\linker\right)$ s.t.:
-
-\begin{itemize}
-  \item[$\partials$] : Set - is a set of admissible, partial programs.
-  \item[$\wellf$] : $\partials$ - a judgement that holds iff a program is not partial.
-  \item[$\singlestep$] : $\wholes\to\events\to\wholes$ - a step relation, where $\wholes=\{w\in\partials\ |\ \wellf w\}$.
-        For $e\in\events$ and $p,p'\in\wholes$ we say for $\estep{p}{e}{p'}$ that program $p$ performs a step with action $e$ to program $p'$.
-        If $e=\emptyevent$, we write $\step{p}{p'}$.
-        $\terminationevent$ is the termination event.
-  \item[$\linker$] : $\partials\to\partials\to\partials$ - links two partial programs together in some way, resulting in a new partial program.
-\end{itemize}
-\end{definition}
-\begin{definition}[Traces]
-  A trace $\trace$ is an infinite sequence of events that results from the relation $\singlestep$.
-  That is, we obtain the trace $\trace=\seqcons{e_{0}}{\seqcons{e_{1}}{\dots}}$ for the execution sequence $\estep{p}{e_{0}}{\estep{p'}{e_{1}}{\dots}}$ and write $\mktrace{p}{\trace}$.
-  The set of all traces is $\traces$.
-\end{definition}
-\begin{definition}[Behavior]
-  The behavior of a whole program $p$ is a set of all traces it produces, i.e. $\behav{p}=\{\trace\ |\ \mktrace{p}{\trace}\}$.
-\end{definition}
-\begin{definition}[Hyperproperties]
-  A hyperproperty $H$ is a set of admissible sets of traces. Thus, if $p$ satisfies $H$ (written $\sat{p}{H}$), then $\behav{p}\in H$.
-\end{definition}
-\begin{definition}[Robust Property Satisfaction]
-  A program $p$ robustly satisfies a hyperproperty $\Pi$, written $\rsat{p}{\Pi}$, iff $\forall C\in\partials,\sat{C\linker p}{\Pi}$. The same notation is used for robust hyperproperty satisfaction.
-\end{definition}
-\begin{lemma}[Property Satisfaction Refinement]
-  For a property $\prop$ that refines $\prop'$, i.e. $\prop\subseteq\prop'$, if any $p\in\partials$ satisfies $\sat{p}{\prop'}$, then $\sat{p}{\prop}$.
-\end{lemma}
-
-\begin{definition}[Compilers]
+\section{Content}
+Programs are written in some programming language.
+In this work, programs are elements of $\partials$, which is the set of partial programs of a given programming language.
 A compiler between languages $\S$ and $\T$ is a partial function $\stcomp{\bullet}$ from $\src{\partials}$ to $\trg{\partials}$.
-\end{definition}
+To reason about preservation of program behavior (hyperproperties~\cite{clarkson08}), we have to define a suitable compiler property.
 
@@
 \noindent
-We can, finally, discuss various forms of compositions.
+So, given any restriction on a program's behavior $\Pi$ and a source program $\src{p}$ that robustly satisfies $\Pi$ ($\rsat{\src{p}}{\Pi}$), the compiled program $\stcomp{\src{p}}$ also needs to robustly satisfy $\Pi$.
+The intuition of \Cref{def:rtp} is as follows: If the programmer makes certain assumptions on what a program does, these assumptions should also hold for the compiled program.
+% example (?)
+Practical compilers are composed of multiple others.
+Therefore, we now investigate compositionality of robust compilers.
+
 We begin with the most straightforward one, which coincides with ordinary function composition.
@@
 \end{definition}
-\begin{lemma}[Weakening RTP]\label{lem:weaken}
-  Given classes $\cC_{1}, \cC_{2}$ such that $\cC_{1}\subseteq\cC_{2}$ and $\rtp{\stcomp{\bullet}}{\cC_{2}}$, it holds that $\rtp{\stcomp{\bullet}}{\cC_{1}}$.
-\end{lemma}
+Now, assuming that two compilers preserve certain classes, we know that their sequential composition preserves the least upper bound:
 \begin{lemma}[Sequential Composition with RTP]\label{lem:seqcompo}
@@
 \end{lemma}
+
+%% TODO: not sure where to put this
+%\begin{lemma}[Weakening RTP]\label{lem:weaken}
+%  Given classes $\cC_{1}, \cC_{2}$ such that $\cC_{1}\subseteq\cC_{2}$ and $\rtp{\stcomp{\bullet}}{\cC_{2}}$, it holds that $\rtp{\stcomp{\bullet}}{\cC_{1}}$.
+%\end{lemma}
 
@@
 \end{definition}
+\noindent
+We can derive a similar result to \Cref{lem:seqcompo}:
 \begin{lemma}[Upper Composition with RTP]
@@
 Any compiler used in production is defined that way since one can, as an example, compile C code to both x86 and MIPS assembly.
-We view frameworks such as RubyOnRails also as an example that can emit two different target languages, in this case, HTML and JS, at once.
+We view frameworks such as RubyOnRails also as an example that can emit two different target languages at once, in this case, HTML and JS.
 \begin{definition}[Lower Composition]
@@
 
+\MP{Confusing and possibly wrong}
 Using the upper and lower composition, we may end up with a diamond situation, which the following lemma describes.
@@
 \end{lemma}
-\MK{This shouldn't work with relational hyperproperties, right?}
+% This shouldn't work with relational hyperproperties, right?
 
 {\color{red}\Large TODO: Miro compo}
-
-\section{Conclusion}
-
-We presented some practically relevant compiler compositions and their respective implications for security.
-Furthermore, compiler developers need not worry about optimization order constraints if the involved optimizations are proven to preserve properties of interest robustly.
 

### 2021-10-19 prisc.tex
Add references

@@
 Secure Compilation studies the preservation of security properties and hyperproperties when compiling a source to a target program.
-Robust Compilation~\cite{abate2019} is a novel compiler property that does not require the compiler to add such unnecessary measures.
+Robust Compilation~\cite{patrignani18,abate2019,patrignani19,patrignani21} is a novel compiler property that does not require the compiler to add such unnecessary measures.
 Informally, if a given (partial) source program linked with an arbitrary source context satisfies some properties of interest, so should its compiled version linked with an arbitrary target context.

### 2021-10-25 prisc.tex
stuff for my compilation
Add content about enforcement to PRISC

@@
 %% use defaults for review submission.
-%\acmConference[PL'22]{ACM SIGPLAN Workshop on Principles of Secure Compilation}{January 22nd, 2022}{Philadelphia, PA, USA}
-%\acmYear{2022}
+\acmConference[PL'22]{ACM SIGPLAN Workshop on Principles of Secure Compilation}{January 22nd, 2022}{Philadelphia, PA, USA}
+\acmYear{2022}
 %\acmISBN{} % \acmISBN{978-x-xxxx-xxxx-x/YY/MM}
@@
 \usepackage{cleveref}
+\usepackage{listings}
 \usepackage{mmmacros}
@@
 %% Title information
-\title{Composing Secure Compilers}         %% [Short Title] is optional;
+\title{Composing Compilers}         %% [Short Title] is optional;
                                         %% when present, will be used in
@@
 
-\section{Content}
+\section{Compositionality}
 Programs are written in some programming language.
@@
 \end{lemma}
+The lemma also generalizes to $n$ robust compilers, each preserving one of $n$ classes.
 
-%% TODO: not sure where to put this
-%\begin{lemma}[Weakening RTP]\label{lem:weaken}
-%  Given classes $\cC_{1}, \cC_{2}$ such that $\cC_{1}\subseteq\cC_{2}$ and $\rtp{\stcomp{\bullet}}{\cC_{2}}$, it holds that $\rtp{\stcomp{\bullet}}{\cC_{1}}$.
-%\end{lemma}
+An intuitive lemma that also follows is that a robust compiler preserving some class $\cC_{2}\supseteq\cC_{1}$ also preserves $\cC_{1}$.
+\begin{lemma}[Weakening RTP]\label{lem:weaken}
+  Given classes $\cC_{1}, \cC_{2}$ such that $\cC_{1}\subseteq\cC_{2}$ and $\rtp{\stcomp{\bullet}}{\cC_{2}}$, it holds that $\rtp{\stcomp{\bullet}}{\cC_{1}}$.
+\end{lemma}
 
@@
 We can derive a similar result to \Cref{lem:seqcompo}:
-\begin{lemma}[Upper Composition with RTP]
+\begin{lemma}[Upper Composition with RTP]\label{lem:useqcompo}
   Given $\rtp{\stcomp{\bullet}}{\cC_{1}}$ and $\rtp{\itcomp{\bullet}}{\cC_{2}}$, then $\rtp{\uhcsitcomp{\bullet}}{\cC_{1}\cap\cC_{2}}$.
 \end{lemma}
+\Cref{lem:useqcompo} also generalizes to a number of compilers and classes, as expected.
 
 With the same idea, we define a dual composition that goes from a single source language to multiple target languages.
-Any compiler used in production is defined that way since one can, as an example, compile C code to both x86 and MIPS assembly.
-We view frameworks such as RubyOnRails also as an example that can emit two different target languages at once, in this case, HTML and JS.
+RubyOnRails is a framework where the processed code is emitted to different target languages at once, in this case, HTML and JavaScript.
 \begin{definition}[Lower Composition]
@@
 
-\MP{Confusing and possibly wrong}
-Using the upper and lower composition, we may end up with a diamond situation, which the following lemma describes.
-Especially from a security researcher perspective, the diamond lemma is attractive as it says that there is no harm to prove compiler passes robust property preserving in a modular way.
-That is, consider some program instrumented against SPECTRE v1 attacks and, in parallel to that, against SPECTRE v4.
-If the compiler instrumentation for v1 does not violate v4, the lemma tells us that combining both instrumentations in some way yields an instrumentation that protects against both v1 and v4.
+%\MP{Confusing and possibly wrong}
+%Using the upper and lower composition, we may end up with a diamond situation, which the following lemma describes.
+%Especially from a security researcher perspective, the diamond lemma is attractive as it says that there is no harm to prove compiler passes robust property preserving in a modular way.
+%That is, consider some program instrumented against SPECTRE v1 attacks and, in parallel to that, against SPECTRE v4.
+%If the compiler instrumentation for v1 does not violate v4, the lemma tells us that combining both instrumentations in some way yields an instrumentation that protects against both v1 and v4.
 
-\begin{lemma}[Diamond]\label{lem:diamond}
-  Given $\rtp{\lhcsiocomp{\bullet}}{\cC_{1}}$ and $\rtp{\uhciotcomp{\bullet}}{\cC_{2}}$ with $\stcomp{\bullet} = \lambda\src{p}.\uhciotcomp{\lhcsiocomp{p}}$, then $\rtp{\stcomp{\bullet}}{\cC_{1}\cap\cC_{2}}$.
-\end{lemma}
+%\begin{lemma}[Diamond]\label{lem:diamond}
+%  Given $\rtp{\lhcsiocomp{\bullet}}{\cC_{1}}$ and $\rtp{\uhciotcomp{\bullet}}{\cC_{2}}$ with $\stcomp{\bullet} = \lambda\src{p}.\uhciotcomp{\lhcsiocomp{p}}$, then $\rtp{\stcomp{\bullet}}{\cC_{1}\cap\cC_{2}}$.
+%\end{lemma}
 
 The following free theorem is a direct consequence of \Cref{lem:seqcompo} where the involved compilers' input and output are both partial programs in the same language.
-Given that some set of optimization passes is robustly preserving properties, we can combine them in an arbitrary order and preserve the same least upper bound.
-A compilers optimization pipeline ordering is difficult and often hand-tuned.
+Given that some set of compiler passes is robustly preserving properties, we can combine them in an arbitrary order and preserve the same least upper bound.
+A compiler's pipeline ordering is difficult and often hand-tuned.
 The lemma allows us to not care about the particular order of optimizations regarding their robust property preservation.
-So, the compiler developer is free to swap around.
+So, the compiler developer is free to swap passes around.
 \begin{lemma}[Swappable]
@@
 \end{lemma}
-% This shouldn't work with relational hyperproperties, right?
 
-{\color{red}\Large TODO: Miro compo}
+However in practice, compiler passes are not necessarily robustly preserving.
+Consider the following C code snippet that performs an infinite loop if an invalid pointer is given:
+\begin{lstlisting}[language=c]
+int something(int* ptr) {
+  while(!ptr);
+  return *ptr;
+}
+\end{lstlisting}
+Compiling such code with optimizations turned on yields an x86-program where the potentially infinite loop has been removed:
+\begin{lstlisting}[language={[x86masm]Assembler}]
+something(int*):
+  mov eax, DWORD PTR [rdi]
+  ret
+\end{lstlisting}
+Not only does the program terminate given an invalid pointer, it violates memory safety, which it did not before.
+
+There exist several memory-safety instrumentations, both for target level~\cite{watson15,aurajo16,vassena19,kouwe17} and source level~\cite{nagarakatte10,akritidis09,rigger17,tarditi2018}.
+These compilation passes \textit{enforce} memory safety by adding dynamic checks to the program and crashing appropriately.
+So, even though the programmer has written a memory-insecure program, we might still be interested in a memory-safe binary.
+
+We plan to extend our work with such source-code instrumentations which enforce classes of trace hyperproperties.
+A sketch of the theory is as follows.
+
+\begin{definition}[Secure Instrumentation for Preserving $\cC$]\label{def:secure-instrumentation}
+  A secure instrumentation with respect to some class $\cC$ ensures hyperproperties described by some class $\cC'$ without violating $\cC$-satisfying programs:
+
+  \begin{enumerate}
+  \item $\rtp{\stcomp{\bullet}}{\cC}$
+  \item $\rtp{\stcomp{\bullet}}{\cC}\implies\rtp{\stcomp{\bullet}}{\cC\cup\cC'}$
+  \end{enumerate}
+
+\noindent
+  We write $\sinstr{\stcomp{\bullet}}{\cC'}{\cC}$.
+\end{definition}
+We can now investigate an insecure compilation chain from e.g.\ memory-safe Rust to optimized, insecure C, to memory-safe CheckedC.
+The intuition that a secure instrumentation saves the day does indeed hold and is captured by the following lemma.
+\begin{lemma}[Final Enforcement]
+  Given classes $\cC_{1},\cC_{2}$ and compilers $\sicomp{\bullet}$,$\itcomp{\bullet}$, we assume:
+
+  \begin{itemize}
+  \item $\rtp{\sicomp{\bullet}}{\cC_{1}}$
+  \item $\sinstr{\itcomp{\bullet}}{\cC_{2}}{\cC_{1}}$
+  \end{itemize}
+\noindent
+  Then, $\rtp{\sitcomp{\bullet}}{\cC_{1}\cup\cC_{2}}$.
+\end{lemma}
+
+Likewhise, running a compiler that does not respect e.g.\ memory-safety after a memory-safety instrumentation nullifies the preservation of memory-safety:
+\begin{lemma}[Final Destination]
+  Given classes $\cC_{1},\cC_{2}$ and compilers $\sicomp{\bullet}$,$\itcomp{\bullet}$, we assume:
+
+  \begin{itemize}
+  \item $\sinstr{\sicomp{\bullet}}{\cC_{2}}{\cC_{1}}$
+  \item $\rtp{\itcomp{\bullet}}{\cC_{1}}$
+  \end{itemize}
+\noindent
+  Then, $\rtp{\sitcomp{\bullet}}{\cC_{1}}$.
+\end{lemma}
+
+Now, consider any stereotypical compilation pipeline.
+The security researcher has a strong interest to ensure that properties at source level are preserved at target level.
+Thus, if source programs robustly satsify some property, so should the target.
+However, it might not be necessary for compilation passes from one intermediate reperesentation to the other to preserve properties robustly, since compiler intermediate representations are outside the considered attack vector.
+So, there might be some weaker property a pass should satisfy in order to render the whole compilation pipeline secure.
+
+We intend to study property-free characterizations of the presented material.
+Another area is the specialization of it to specific classes, such as $\cSafety$, and subsets thereof.
 

### 2021-10-26 prisc.tex
Fix some typos

@@
 \usepackage{listings}
-\usepackage{mmmacros}
+\usepackage{./../mmmacros}
 \usepackage[switch]{lineno}
@@
 Programs are written in some programming language.
-In this work, programs are elements of $\partials$, which is the set of partial programs of a given programming language.
+In this work, programs are elements of $\partials$, the set of partial programs of a given programming language.
 A compiler between languages $\S$ and $\T$ is a partial function $\stcomp{\bullet}$ from $\src{\partials}$ to $\trg{\partials}$.
-To reason about preservation of program behavior (hyperproperties~\cite{clarkson08}), we have to define a suitable compiler property.
+To reason about the preservation of program behaviour (hyperproperties~\cite{clarkson08}), we have to define a suitable compiler property.
 
@@
 % example (?)
-Practical compilers are composed of multiple others.
-Therefore, we now investigate compositionality of robust compilers.
+Practical compilers are composed of numerous others.
+Therefore, we now investigate the compositionality of robust compilers.
 
@@
 The following free theorem is a direct consequence of \Cref{lem:seqcompo} where the involved compilers' input and output are both partial programs in the same language.
-Given that some set of compiler passes is robustly preserving properties, we can combine them in an arbitrary order and preserve the same least upper bound.
+Given that some compiler passes are robustly preserving properties, we can combine them in an arbitrary order and preserve the same least upper bound.
 A compiler's pipeline ordering is difficult and often hand-tuned.
@@
 
-However in practice, compiler passes are not necessarily robustly preserving.
+However, in practice, compiler passes are not necessarily robustly preserving.
 Consider the following C code snippet that performs an infinite loop if an invalid pointer is given:
@@
 
-We plan to extend our work with such source-code instrumentations which enforce classes of trace hyperproperties.
+We plan to extend our work with such source-code instrumentations, which enforce classes of trace hyperproperties.
 A sketch of the theory is as follows.
@@
 We can now investigate an insecure compilation chain from e.g.\ memory-safe Rust to optimized, insecure C, to memory-safe CheckedC.
-The intuition that a secure instrumentation saves the day does indeed hold and is captured by the following lemma.
+The intuition that a secure instrumentation saves the day holds and is captured by the following lemma.
 \begin{lemma}[Final Enforcement]
@@
 
-Likewhise, running a compiler that does not respect e.g.\ memory-safety after a memory-safety instrumentation nullifies the preservation of memory-safety:
+Likewise, running a compiler that does not respect e.g.\ memory-safety after a memory-safety instrumentation nullifies the preservation of memory-safety:
 \begin{lemma}[Final Destination]
@@
 Now, consider any stereotypical compilation pipeline.
-The security researcher has a strong interest to ensure that properties at source level are preserved at target level.
-Thus, if source programs robustly satsify some property, so should the target.
-However, it might not be necessary for compilation passes from one intermediate reperesentation to the other to preserve properties robustly, since compiler intermediate representations are outside the considered attack vector.
-So, there might be some weaker property a pass should satisfy in order to render the whole compilation pipeline secure.
+The security researcher has a strong interest to ensure that properties at the source level are preserved at the target level.
+Thus, if source programs robustly satisfy some property, so should the target.
+However, it might not be necessary for compilation passes from one intermediate representation to the other to preserve properties robustly since compiler intermediate representations are outside the considered attack vector.
+So, there might be some weaker property a pass has to satisfy in order to render the whole compilation pipeline secure.
 
 We intend to study property-free characterizations of the presented material.
-Another area is the specialization of it to specific classes, such as $\cSafety$, and subsets thereof.
+Another area is its specialization to specific classes, such as $\cSafety$, and subsets thereof.
 

### 2021-10-26 prisc.tex
Incorporate feedback.

@@
 
+% a recent theory of secure compilation is robust ...
+
 Compilers translate programs from a source to a target programming language.
-Secure Compilation studies the preservation of security properties and hyperproperties when compiling a source to a target program.
-Robust Compilation~\cite{patrignani18,abate2019,patrignani19,patrignani21} is a novel compiler property that does not require the compiler to add such unnecessary measures.
-Informally, if a given (partial) source program linked with an arbitrary source context satisfies some properties of interest, so should its compiled version linked with an arbitrary target context.
+A secure compilation chain preserves source level properties at the target level.
+Different such preservation criteria for compilers exist.
+
+Robust Compilation~\cite{patrignani18,abate2019,patrignani19,patrignani21} is a set of criteria for secure compilers that have become widely accepted.
+Informally, a compiler is robustly preserving a property if source and target program linked with an arbitrary source and target context, respectively, satisfy that property.
+Hereby, the target program is the compiled version of the source program.
+
 Even though there exist robust compilers, they are far from practical.
 Real-world compilers consist of several smaller compilers that are composed with each other in a certain way.
-A prominent example would be any compiler based on the LLVM toolchain~\cite{lattner04}, whose optimization pipeline consists of many passes, which one can view as independent compilers composed to achieve speedups.
-To the best of our knowledge, current work does not discuss which properties are preserved by compilation chains.
+A prominent example would be any compiler based on the LLVM toolchain~\cite{lattner04}, whose optimisation pipeline consists of many passes, which one can view as independent compilers composed to achieve speedups.
+Also, any lowering steps, such as from a frontend language to LLVM IR and subsequently to machine Assembly, are intermixed compilers.
+To the best of our knowledge, current work on robust compilation does not discuss the preservation of source-level properties for compilation chains.
 
 This paper investigates how different compiler compositions preserve hyperproperties, given that the compilers involved in the composition fulfil certain assumptions.
-To reason about hyperproperty preservation, we will look at the lattice of classes of hyperproperties.
+To reason about hyperproperty preservation, we will look at the lattice of classes of hyperproperties, omitted in this paper.
 Within the lattice of such classes, we examine whether at least the least upper bound is preserved by our chosen compositions.
-We conclude with a preliminary result: We show that the order of optimizations in a compilation pipeline does not matter for property preservation, assuming that each such pass preserves the class of interest.
+We present a preliminary result: We show that the order of optimisations in a robustly preserving compilation pipeline does not matter for property preservation.
+We conclude with a discussion on what happens if compilers in the chain do not robustly preserve properties of interest.
+
+% optimisations might not preserve the prop
 
 \section{Compositionality}
-Programs are written in some programming language.
+% to reason about compilers, we need to ....
 In this work, programs are elements of $\partials$, the set of partial programs of a given programming language.
-A compiler between languages $\S$ and $\T$ is a partial function $\stcomp{\bullet}$ from $\src{\partials}$ to $\trg{\partials}$.
-To reason about the preservation of program behaviour (hyperproperties~\cite{clarkson08}), we have to define a suitable compiler property.
+A compiler is a partial function $\stcomp{\bullet}$ from programs $\src{\partials}$ of some source language $\S$ to programs $\trg{\partials}$ of some target language $\T$.
+To reason about the preservation of properties present in some $\S$ program, we define in \Cref{def:rtp} the criterion a robustly preserving compiler should fulfil.
+We extend existing theory on hyperproperties~\cite{clarkson08} by introducing classes, which are sets of hyperproperties.
+If all hyperproperties from a class $\cC$ are satisfied by some program $p$ that is linked with an arbitrary context, then we write $\rsat{p}{\cC}$.
+Accordingly, for some $\Pi\in\cC$, we also write $\rsat{p}{\Pi}$ whenever $p$ robustly satisfies $\Pi$.
 
-\begin{definition}[Robust Trace-Hyperproperty Preservation]\label{def:rtp}
-  For a given class $\cC$, a compiler from languages $\S$ to $\T$ robustly preserves $\cC$ iff
+\begin{definition}[Robust Compilation]\label{def:rtp}
+  For a given class $\cC$, a compiler from languages $\S$ to $\T$ robustly preserves $\cC$ ($\rtp{\stcomp{\bullet}}{\cC}$) iff
   $$
@@
   $$
-  We write $\rtp{\stcomp{\bullet}}{\cC}$.
 \end{definition}
 \noindent
-So, given any restriction on a program's behavior $\Pi$ and a source program $\src{p}$ that robustly satisfies $\Pi$ ($\rsat{\src{p}}{\Pi}$), the compiled program $\stcomp{\src{p}}$ also needs to robustly satisfy $\Pi$.
+%So, given a hyperproperty $\Pi$ from some class $\cC$ and a source program $\src{p}$ that robustly satisfies $\Pi$ ($\rsat{\src{p}}{\Pi}$), the compiled program $\stcomp{\src{p}}$ also needs to robustly satisfy $\Pi$.
 The intuition of \Cref{def:rtp} is as follows: If the programmer makes certain assumptions on what a program does, these assumptions should also hold for the compiled program.
-% example (?)
-Practical compilers are composed of numerous others.
-Therefore, we now investigate the compositionality of robust compilers.
 
-We begin with the most straightforward one, which coincides with ordinary function composition.
-From a practical perspective, this means plugging the result of one compiler into another one.
-Such pipelines happen when optimizing source code, so at the level of a suitable intermediate representation, but also on a higher level:
-Consider as an example a typical TypeScript compilation pipeline.
-First, it translates TypeScript code to JavaScript, which V8 compiles to Ignition bytecode that, after some runtime-performance analysis, is just-in-time compiled via TurboFan to machine code such as x86.
+In practice, (robust) compilers are composed of numerous others.
+Therefore, we now investigate their compositionality.
+
+\subsection{Simple Compositionality}
+We first consider function composition, i.e.\ plugging the result of one compiler into another one.
+Such pipelines happen when optimising source code (so, at the level of a suitable intermediate representation), but also on a higher level:
+Consider as an example a typical $\src{TypeScript}$ compilation pipeline.
+First, it translates $\src{TypeScript}$ code to $\irl{JavaScript}$, which V8 compiles to $\trg{Ignition\ bytecode}$.
 
@@
 \end{definition}
-Now, assuming that two compilers preserve certain classes, we know that their sequential composition preserves the least upper bound:
+Now, assuming that two compilers preserve certain classes, we know that their sequential composition preserves the least upper bound; the set intersection of those classes:
 \begin{lemma}[Sequential Composition with RTP]\label{lem:seqcompo}
@@
 \end{lemma}
-The lemma also generalizes to $n$ robust compilers, each preserving one of $n$ classes.
-
-An intuitive lemma that also follows is that a robust compiler preserving some class $\cC_{2}\supseteq\cC_{1}$ also preserves $\cC_{1}$.
-\begin{lemma}[Weakening RTP]\label{lem:weaken}
-  Given classes $\cC_{1}, \cC_{2}$ such that $\cC_{1}\subseteq\cC_{2}$ and $\rtp{\stcomp{\bullet}}{\cC_{2}}$, it holds that $\rtp{\stcomp{\bullet}}{\cC_{1}}$.
-\end{lemma}
+Using an inductive argument, \Cref{lem:seqcompo} generalises to $n$ robust compilers, each preserving one of $n$ classes.
+To do so, one has to generalise the composition of two robust compilers to a set of $n$ robust compilers.
 
 We now consider a compiler that invokes two other compilers.
-HTML, CSS, and JavaScript are the most obvious example of such a composition.
+$\src{HTML}$, $\irl{CSS}$, and JavaScript are the most obvious example of such a composition.
 One writes down some code of any of these three languages embedded in the same file, but different compiler toolchains process them.
@@
   $$\uhcsitcomp{\bullet}=\lambda p.\begin{cases}\stcomp{p} &\text{if }p\in\src{\partials}\\
-                                                \itcomp{p} &\text{if }p\in\irl{\partials}\end{cases}$$.
+                                                \itcomp{p} &\text{if }p\in\irl{\partials}\end{cases}$$
 \end{definition}
 \noindent
-We can derive a similar result to \Cref{lem:seqcompo}:
-\begin{lemma}[Upper Composition with RTP]\label{lem:useqcompo}
+We can derive a similar result to \Cref{lem:seqcompo} here, too:
+\begin{lemma}[Upper Composition with RC]\label{lem:useqcompo}
   Given $\rtp{\stcomp{\bullet}}{\cC_{1}}$ and $\rtp{\itcomp{\bullet}}{\cC_{2}}$, then $\rtp{\uhcsitcomp{\bullet}}{\cC_{1}\cap\cC_{2}}$.
 \end{lemma}
-\Cref{lem:useqcompo} also generalizes to a number of compilers and classes, as expected.
+\Cref{lem:useqcompo} also generalises inductively to a number of compilers and classes.
 
 With the same idea, we define a dual composition that goes from a single source language to multiple target languages.
-RubyOnRails is a framework where the processed code is emitted to different target languages at once, in this case, HTML and JavaScript.
+$\src{RubyOnRails}$ is a framework where the processed code is emitted to different target languages at once, in this case, $\trg{HTML}$ and $\irl{JavaScript}$.
 \begin{definition}[Lower Composition]
@@
 
-\begin{lemma}[Lower Composition with RTP]
+\begin{lemma}[Lower Composition with RC]
   Given $\rtp{\stcomp{\bullet}}{\cC_{1}}$ and $\rtp{\sicomp{\bullet}}{\cC_{2}}$, then $\rtp{\lhcsitcomp{\bullet}}{\cC_{1}\cap\cC_{2}}$.
@@
 
-The following free theorem is a direct consequence of \Cref{lem:seqcompo} where the involved compilers' input and output are both partial programs in the same language.
+The following free theorem (\Cref{lem:swappable}) is a direct consequence of \Cref{lem:seqcompo} where the involved compilers' input and output are both partial programs in the same language.
 Given that some compiler passes are robustly preserving properties, we can combine them in an arbitrary order and preserve the same least upper bound.
 A compiler's pipeline ordering is difficult and often hand-tuned.
-The lemma allows us to not care about the particular order of optimizations regarding their robust property preservation.
+The lemma allows us to not care about the particular order of optimisations regarding their robust property preservation.
 So, the compiler developer is free to swap passes around.
-\begin{lemma}[Swappable]
+\begin{lemma}[Swappable]\label{lem:swappable}
   Given $\rtp{\ttcomp{\bullet}_{(1)}}{\cC_{1}}$ and $\rtp{\ttcomp{\bullet}_{(2)}}{\cC_{2}}$, then $\rtp{\ttcompN{\ttcomp{\bullet}_{(2)}}_{(1)}}{\cC_{1}\cap\cC_{2}}$ and $\rtp{\ttcompN{\ttcomp{\bullet}_{(1)}}_{(2)}}{\cC_{1}\cap\cC_{2}}$.
@@
 However, in practice, compiler passes are not necessarily robustly preserving.
+Consider any stereotypical compilation pipeline.
+The security researcher has a strong interest to ensure that properties at the source level are preserved at the target level.
+Thus, if source programs robustly satisfy some property, so should the target.
+However, it might not be necessary for compilation passes from one intermediate representation to the other to preserve properties robustly since compiler intermediate representations are outside the considered attack vector.
+So, there might be some weaker property a pass has to satisfy in order to render the whole compilation pipeline secure.
+
+\subsection{Advanced Compositionality}
 Consider the following C code snippet that performs an infinite loop if an invalid pointer is given:
@@
 \end{lstlisting}
-Compiling such code with optimizations turned on yields an x86-program where the potentially infinite loop has been removed:
+Compiling such code with optimisations turned on yields an x86-program where the potentially infinite loop has been removed:
 \begin{lstlisting}[language={[x86masm]Assembler}]
@@
 \end{lstlisting}
-Not only does the program terminate given an invalid pointer, it violates memory safety, which it did not before.
+Upon calling this function with an invalid pointer, we now have an attack to violate memory safety, which did not work before; the check has been completely eliminated.
 
-There exist several memory-safety instrumentations, both for target level~\cite{watson15,aurajo16,vassena19,kouwe17} and source level~\cite{nagarakatte10,akritidis09,rigger17,tarditi2018}.
-These compilation passes \textit{enforce} memory safety by adding dynamic checks to the program and crashing appropriately.
+There exist several memory-safety instrumentations, both for target level~\cite{watson15,aurajo16,vassena19,kouwe17} and source level~\cite{nagarakatte10,akritidis09,rigger17,tarditi2018,sammler21}.
+These compilation passes \textit{enforce} memory safety by adding static or dynamic checks to the program and, in the case of dynamic checks, crashing appropriately.
 So, even though the programmer has written a memory-insecure program, we might still be interested in a memory-safe binary.
 
-We plan to extend our work with such source-code instrumentations, which enforce classes of trace hyperproperties.
+We plan to extend our work with code instrumentations, which \textit{enforce} classes of trace hyperproperties.
 A sketch of the theory is as follows.
@@
 \begin{definition}[Secure Instrumentation for Preserving $\cC$]\label{def:secure-instrumentation}
-  A secure instrumentation with respect to some class $\cC$ ensures hyperproperties described by some class $\cC'$ without violating $\cC$-satisfying programs:
-
-  \begin{enumerate}
-  \item $\rtp{\stcomp{\bullet}}{\cC}$
-  \item $\rtp{\stcomp{\bullet}}{\cC}\implies\rtp{\stcomp{\bullet}}{\cC\cup\cC'}$
-  \end{enumerate}
-
-\noindent
+  A secure instrumentation with respect to some class $\cC$ enforces hyperproperties described by some class $\cC'$ without violating $\cC$-satisfying programs.
   We write $\sinstr{\stcomp{\bullet}}{\cC'}{\cC}$.
 \end{definition}
-We can now investigate an insecure compilation chain from e.g.\ memory-safe Rust to optimized, insecure C, to memory-safe CheckedC.
-The intuition that a secure instrumentation saves the day holds and is captured by the following lemma.
-\begin{lemma}[Final Enforcement]
+Using this, we firstly want to inspect an insecure compilation chain from e.g.\ memory-safe $\src{Rust}$ to optimised, insecure $\irl{C}$, to memory-safe $\trg{CheckedC}$:
+\begin{lemma}[Enforcement may preserve...]
   Given classes $\cC_{1},\cC_{2}$ and compilers $\sicomp{\bullet}$,$\itcomp{\bullet}$, we assume:
@@
 
-Likewise, running a compiler that does not respect e.g.\ memory-safety after a memory-safety instrumentation nullifies the preservation of memory-safety:
-\begin{lemma}[Final Destination]
+Likewise, running a compiler that does not respect e.g.\ memory-safety after a memory-safety instrumentation nullifies the preservation of it:
+\begin{lemma}[...but, order matters!]
   Given classes $\cC_{1},\cC_{2}$ and compilers $\sicomp{\bullet}$,$\itcomp{\bullet}$, we assume:
@@
 
-Now, consider any stereotypical compilation pipeline.
-The security researcher has a strong interest to ensure that properties at the source level are preserved at the target level.
-Thus, if source programs robustly satisfy some property, so should the target.
-However, it might not be necessary for compilation passes from one intermediate representation to the other to preserve properties robustly since compiler intermediate representations are outside the considered attack vector.
-So, there might be some weaker property a pass has to satisfy in order to render the whole compilation pipeline secure.
+We intend to study property-free characterisations of the presented material.
+Another area is its specialisation to specific classes, such as $\cHSafety$, and subsets thereof.
 
-We intend to study property-free characterizations of the presented material.
-Another area is its specialization to specific classes, such as $\cSafety$, and subsets thereof.
+%
+% use british english!
+%
+% uppercase after semicolon
+%
 
@@
 
+\newpage
+
 %% Bibliography

### 2021-10-26 prisc.tex
Remove \noindent
Get rid of a few hiccups

@@
 %% use defaults for review submission.
-\acmConference[PL'22]{ACM SIGPLAN Workshop on Principles of Secure Compilation}{January 22nd, 2022}{Philadelphia, PA, USA}
+\acmConference[PriSC'22]{ACM SIGPLAN Workshop on Principles of Secure Compilation}{January 22nd, 2022}{Philadelphia, PA, USA}
 \acmYear{2022}
@@
 \end{definition}
-\noindent
 We can derive a similar result to \Cref{lem:seqcompo} here, too:
@@
 
-There exist several memory-safety instrumentations, both for target level~\cite{watson15,aurajo16,vassena19,kouwe17} and source level~\cite{nagarakatte10,akritidis09,rigger17,tarditi2018,sammler21}.
-These compilation passes \textit{enforce} memory safety by adding static or dynamic checks to the program and, in the case of dynamic checks, crashing appropriately.
+To prevent such issues, a common approach for ensuring memory safety is to extend the source language~\cite{tarditi2018,sammler21}.
+But, this is not always desirable, since source language extensions imply that existing programs might need a rewrite.
+Instrumentation passes \textit{enforce} memory safety by adding dynamic checks to the program and crashing appropriately, thus leaving the features of the source language untouched.
+There exist several memory-safety instrumentations, both for target level~\cite{watson15,aurajo16,vassena19,kouwe17} and source level~\cite{nagarakatte10,akritidis09,rigger17}.
 So, even though the programmer has written a memory-insecure program, we might still be interested in a memory-safe binary.
@@
 Using this, we firstly want to inspect an insecure compilation chain from e.g.\ memory-safe $\src{Rust}$ to optimised, insecure $\irl{C}$, to memory-safe $\trg{CheckedC}$:
-\begin{lemma}[Enforcement may preserve...]
+\begin{example}[Enforcement may preserve...]
   Given classes $\cC_{1},\cC_{2}$ and compilers $\sicomp{\bullet}$,$\itcomp{\bullet}$, we assume:
@@
   Then, $\rtp{\sitcomp{\bullet}}{\cC_{1}\cup\cC_{2}}$.
-\end{lemma}
+\end{example}
 
 Likewise, running a compiler that does not respect e.g.\ memory-safety after a memory-safety instrumentation nullifies the preservation of it:
-\begin{lemma}[...but, order matters!]
+\begin{example}[...but, order matters!]
   Given classes $\cC_{1},\cC_{2}$ and compilers $\sicomp{\bullet}$,$\itcomp{\bullet}$, we assume:
@@
   Then, $\rtp{\sitcomp{\bullet}}{\cC_{1}}$.
-\end{lemma}
+\end{example}
 
-We intend to study property-free characterisations of the presented material.
+We also intend to study property-free characterisations of the presented material.
 Another area is its specialisation to specific classes, such as $\cHSafety$, and subsets thereof.

### 2021-10-27 prisc.tex
Incorporate feedback

@@
 \usepackage{amsthm}
+\usepackage{glossaries}
 \usepackage{bm}
@@
 \renewcommand{\linenumberfont}{\normalfont\bfseries\small\color{red}}
+
+\newacronym{rc}{RC}{Robust Compilation}
 
@@
 Compilers translate programs from a source to a target programming language.
-A secure compilation chain preserves source level properties at the target level.
-Different such preservation criteria for compilers exist.
+A secure compiler preserves source level properties at the target level under arbitrary program contexts.
+Different such criteria exist.
 
-Robust Compilation~\cite{patrignani18,abate2019,patrignani19,patrignani21} is a set of criteria for secure compilers that have become widely accepted.
+A recent theory of secure compilation is \gls{rc}, which is a collection of criteria for secure compilers, that have become accepted~\cite{abate19,abate20,patrignani21}.
 Informally, a compiler is robustly preserving a property if source and target program linked with an arbitrary source and target context, respectively, satisfy that property.
@@
 Real-world compilers consist of several smaller compilers that are composed with each other in a certain way.
-A prominent example would be any compiler based on the LLVM toolchain~\cite{lattner04}, whose optimisation pipeline consists of many passes, which one can view as independent compilers composed to achieve speedups.
-Also, any lowering steps, such as from a frontend language to LLVM IR and subsequently to machine Assembly, are intermixed compilers.
-To the best of our knowledge, current work on robust compilation does not discuss the preservation of source-level properties for compilation chains.
+An example would be any compiler based on the LLVM toolchain~\cite{lattner04}, whose optimisation pipeline consists of many passes, which one can view as independent compilers composed to achieve speedups.
+Also, any lowering steps, such as from a frontend language to LLVM IR and subsequently to machine Assembly, are compilers.
+To the best of our knowledge, current work on robust compilation does not discuss the preservation of source-level properties for compilation chains such as the one above.
 
-This paper investigates how different compiler compositions preserve hyperproperties, given that the compilers involved in the composition fulfil certain assumptions.
-To reason about hyperproperty preservation, we will look at the lattice of classes of hyperproperties, omitted in this paper.
-Within the lattice of such classes, we examine whether at least the least upper bound is preserved by our chosen compositions.
+This paper investigates how different compiler compositions preserve hyperproperties, given that these compilers attain some form of \gls{rc}.
+To reason about hyperproperty preservation, we look at sets of hyperproperties, which we call classes.
+This is an idea taken from~\citet{abate19} but made explicit.
+We examine whether at least the set intersection of classes is preserved by our chosen compositions.
 We present a preliminary result: We show that the order of optimisations in a robustly preserving compilation pipeline does not matter for property preservation.
-We conclude with a discussion on what happens if compilers in the chain do not robustly preserve properties of interest.
-
-% optimisations might not preserve the prop
+We conclude with a discussion on what happens if compilers in the chain do not attain RC for classes of interest.
 
@@
 A compiler is a partial function $\stcomp{\bullet}$ from programs $\src{\partials}$ of some source language $\S$ to programs $\trg{\partials}$ of some target language $\T$.
-To reason about the preservation of properties present in some $\S$ program, we define in \Cref{def:rtp} the criterion a robustly preserving compiler should fulfil.
-We extend existing theory on hyperproperties~\cite{clarkson08} by introducing classes, which are sets of hyperproperties.
+Compilers satisfying \Cref{def:rtp} attain \gls{rc}.
+We rely on existing theory on hyperproperties~\cite{clarkson08}, but view them abstractly as elements of sets that we call \textit{classes}.
 If all hyperproperties from a class $\cC$ are satisfied by some program $p$ that is linked with an arbitrary context, then we write $\rsat{p}{\cC}$.
@@
 Consider as an example a typical $\src{TypeScript}$ compilation pipeline.
-First, it translates $\src{TypeScript}$ code to $\irl{JavaScript}$, which V8 compiles to $\trg{Ignition\ bytecode}$.
+First, the compiler translates $\src{TypeScript}$ code to $\irl{JavaScript}$, which a part of V8 eventually, but also partially, compiles the code just-in-time to $\trg{Assembly}$.
 
@@
 To do so, one has to generalise the composition of two robust compilers to a set of $n$ robust compilers.
+A real-world example for such deeply nested compositions is the $\src{TypeScript}$ compilation mentioned above.
+When compiling $\irl{JavaScript}$, V8 translates the code to $\trg{Ignition\ Bytecode}$.
+At runtime, the Ignition interpreter does some performance measurements and particular parts of the code are eventually compiled to machine code.
 
 We now consider a compiler that invokes two other compilers.
-$\src{HTML}$, $\irl{CSS}$, and JavaScript are the most obvious example of such a composition.
-One writes down some code of any of these three languages embedded in the same file, but different compiler toolchains process them.
+$\src{Erlang}$ and $\irl{Elixir}$ are languages used in industry (e.g.\ by Discord) that are one example of such a composition.
+Both compile to $\trg{BEAM\ Bytecode}$.
 
@@
 \Cref{lem:useqcompo} also generalises inductively to a number of compilers and classes.
+A practical example of why that might be useful is the Java Virtual Machine with its $\trg{JVM\ Bytecode}$, which has numerous frontends: $\src{Java}$, $\irl{Kotlin}$, $\src{Scala}$, and $\irl{Clojure}$, to list a few examples.
 
 With the same idea, we define a dual composition that goes from a single source language to multiple target languages.
-$\src{RubyOnRails}$ is a framework where the processed code is emitted to different target languages at once, in this case, $\trg{HTML}$ and $\irl{JavaScript}$.
+\texttt{dune} is a build system which can be used to compile $\src{OCaml}$ code to both $\irl{Assembly}$ and $\trg{Caml\ Bytecode}$.
 \begin{definition}[Lower Composition]
@@
 \end{lemma}
-
-%\MP{Confusing and possibly wrong}
-%Using the upper and lower composition, we may end up with a diamond situation, which the following lemma describes.
-%Especially from a security researcher perspective, the diamond lemma is attractive as it says that there is no harm to prove compiler passes robust property preserving in a modular way.
-%That is, consider some program instrumented against SPECTRE v1 attacks and, in parallel to that, against SPECTRE v4.
-%If the compiler instrumentation for v1 does not violate v4, the lemma tells us that combining both instrumentations in some way yields an instrumentation that protects against both v1 and v4.
-
-%\begin{lemma}[Diamond]\label{lem:diamond}
-%  Given $\rtp{\lhcsiocomp{\bullet}}{\cC_{1}}$ and $\rtp{\uhciotcomp{\bullet}}{\cC_{2}}$ with $\stcomp{\bullet} = \lambda\src{p}.\uhciotcomp{\lhcsiocomp{p}}$, then $\rtp{\stcomp{\bullet}}{\cC_{1}\cap\cC_{2}}$.
-%\end{lemma}
+As before, this can be generalized to an arbitrary number, which also has a connection to the real-world, given by the diverse set of Assembly language dialects.
 
 The following free theorem (\Cref{lem:swappable}) is a direct consequence of \Cref{lem:seqcompo} where the involved compilers' input and output are both partial programs in the same language.
-Given that some compiler passes are robustly preserving properties, we can combine them in an arbitrary order and preserve the same least upper bound.
+Given that some compiler passes are attaining \gls{rc}, we can combine them in an arbitrary order and preserve the same least upper bound.
 A compiler's pipeline ordering is difficult and often hand-tuned.
@@
 
-However, in practice, compiler passes are not necessarily robustly preserving.
+However, in practice, compiler passes are not necessarily attaining \gls{rc}.
 Consider any stereotypical compilation pipeline.
 The security researcher has a strong interest to ensure that properties at the source level are preserved at the target level.
-Thus, if source programs robustly satisfy some property, so should the target.
-However, it might not be necessary for compilation passes from one intermediate representation to the other to preserve properties robustly since compiler intermediate representations are outside the considered attack vector.
-So, there might be some weaker property a pass has to satisfy in order to render the whole compilation pipeline secure.
+Thus, if source programs robustly satisfy some property, so should the target ones.
+But, it might not be necessary for compilation passes from one intermediate representation to the other to preserve properties robustly since compiler intermediate representations are outside the considered attack vector.
+So, there might be some stronger property a pass has to satisfy in order to render the whole compilation pipeline secure.
 
@@
 \end{lstlisting}
-Compiling such code with optimisations turned on yields an x86-program where the potentially infinite loop has been removed:
+Compiling such code with optimisations turned on by using the command \href{https://godbolt.org/z/bnaGnPe36}{\texttt{g++ -O2}} and the \texttt{g++} compiler version 11.2 yields an x86-program where the potentially infinite loop has been removed:
 \begin{lstlisting}[language={[x86masm]Assembler}]
@@
 \end{lstlisting}
-Upon calling this function with an invalid pointer, we now have an attack to violate memory safety, which did not work before; the check has been completely eliminated.
+We now have an attack to violate memory safety: call the function with an invalid pointer.
 
-To prevent such issues, a common approach for ensuring memory safety is to extend the source language~\cite{tarditi2018,sammler21}.
-But, this is not always desirable, since source language extensions imply that existing programs might need a rewrite.
-Instrumentation passes \textit{enforce} memory safety by adding dynamic checks to the program and crashing appropriately, thus leaving the features of the source language untouched.
-There exist several memory-safety instrumentations, both for target level~\cite{watson15,aurajo16,vassena19,kouwe17} and source level~\cite{nagarakatte10,akritidis09,rigger17}.
-So, even though the programmer has written a memory-insecure program, we might still be interested in a memory-safe binary.
+To prevent such issues we can use instrumentation passes that \textit{enforce} memory safety by adding dynamic checks to the program and crashing appropriately.
+There exist several memory-safety instrumentations, both for target level~\cite{watson15,aurajo16,vassena19,kouwe17,tarditi2018,sammler21} and source level~\cite{nagarakatte10,akritidis09,rigger17}.
+%So, even though the programmer has written a memory-insecure program, we might still be interested in a memory-safe binary.
 
-We plan to extend our work with code instrumentations, which \textit{enforce} classes of trace hyperproperties.
+We plan to extend our work with code instrumentations, which enforce classes.
 A sketch of the theory is as follows.
@@
 
-We also intend to study property-free characterisations of the presented material.
-Another area is its specialisation to specific classes, such as $\cHSafety$, and subsets thereof.
+%We also intend to study property-free characterisations of the presented material.
+%Another area is its specialisation to instances, such as $\cHSafety$, and subsets thereof.
+We intend to study the compositionality aspects of concrete instances, such as SPECTRE~\cite{exorcisingSpectres,guarnieri19}, memory safety~\cite{azevedodeamorim18,azevedodeamorim15,dhawan15}, or cryptographic constant-time~\cite{barthe18}.
 

### 2021-10-28 prisc.tex
edits

@@
 %% Title information
-\title{Composing Compilers}         %% [Short Title] is optional;
+\title{Composing Secure Compilers}         %% [Short Title] is optional;
                                         %% when present, will be used in
@@
 Compilers translate programs from a source to a target programming language.
-A secure compiler preserves source level properties at the target level under arbitrary program contexts.
-Different such criteria exist.
-
-A recent theory of secure compilation is \gls{rc}, which is a collection of criteria for secure compilers, that have become accepted~\cite{abate19,abate20,patrignani21}.
-Informally, a compiler is robustly preserving a property if source and target program linked with an arbitrary source and target context, respectively, satisfy that property.
-Hereby, the target program is the compiled version of the source program.
+A secure compiler preserves source level properties at the target level when interoperating with arbitrary program contexts (which are considered attackers).
+% Different such criteria exist.
+A recent theory of secure compilation is \gls{rc}, which is a collection of criteria for secure compilers~\cite{abate19,abate20,patrignani21}.
+Informally, a compiler is \gls{rc} if a source program and its compiled counterpart, linked with an arbitrary source and target context respectively, satisfy that property.
+% Hereby, the target program is the compiled version of the source program.
 
 Even though there exist robust compilers, they are far from practical.
-Real-world compilers consist of several smaller compilers that are composed with each other in a certain way.
-An example would be any compiler based on the LLVM toolchain~\cite{lattner04}, whose optimisation pipeline consists of many passes, which one can view as independent compilers composed to achieve speedups.
-Also, any lowering steps, such as from a frontend language to LLVM IR and subsequently to machine Assembly, are compilers.
-To the best of our knowledge, current work on robust compilation does not discuss the preservation of source-level properties for compilation chains such as the one above.
+Real-world compilers consist of several smaller compilers that are composed with each other in different ways.
+An example would be any compiler based on the LLVM toolchain~\cite{lattner04}, whose optimisation pipeline consists of many passes, which one can view as independent compilers composed with each others.
+Also, any lowering steps, such as from a frontend language to LLVM IR and subsequently to assembly, are compilers.
+To the best of our knowledge, current work on robust compilation does not discuss the preservation of source-level properties for compilers such as the ones above.
 
-This paper investigates how different compiler compositions preserve hyperproperties, given that these compilers attain some form of \gls{rc}.
-To reason about hyperproperty preservation, we look at sets of hyperproperties, which we call classes.
-This is an idea taken from~\citet{abate19} but made explicit.
-We examine whether at least the set intersection of classes is preserved by our chosen compositions.
-We present a preliminary result: We show that the order of optimisations in a robustly preserving compilation pipeline does not matter for property preservation.
-We conclude with a discussion on what happens if compilers in the chain do not attain RC for classes of interest.
+This paper investigates how different compiler compositions preserve different classes of hyperproperties, given that these compilers attain some form of \gls{rc}.
+% To reason about hyperproperty preservation, we look at classes of hyperproperties, which we call classes.
+% This is an idea taken from~\citet{abate19} but made explicit.
+We examine whether these compositions preserve at least the set intersection of classes.
+We then show that the order of optimisations in a \gls{rc} pipeline does not matter for property preservation.
+Finally, we conclude with a discussion on what happens if some compilers in the pipeline do not attain \gls{rc} for some classes of interest.
 
@@
 % to reason about compilers, we need to ....
-In this work, programs are elements of $\partials$, the set of partial programs of a given programming language.
-A compiler is a partial function $\stcomp{\bullet}$ from programs $\src{\partials}$ of some source language $\S$ to programs $\trg{\partials}$ of some target language $\T$.
-Compilers satisfying \Cref{def:rtp} attain \gls{rc}.
-We rely on existing theory on hyperproperties~\cite{clarkson08}, but view them abstractly as elements of sets that we call \textit{classes}.
-If all hyperproperties from a class $\cC$ are satisfied by some program $p$ that is linked with an arbitrary context, then we write $\rsat{p}{\cC}$.
-Accordingly, for some $\Pi\in\cC$, we also write $\rsat{p}{\Pi}$ whenever $p$ robustly satisfies $\Pi$.
-
+In this work, programs $\com{p}$ are elements of $\partials$, the set of partial programs of a given programming language.
+A compiler is a partial function $\stcomp{\bullet}$ from programs $\src{p}$ of some source language $\S$ to programs $\trg{p}$ of some target language $\T$.
+Compilers satisfying \Cref{def:rtp} below attain \gls{rc}~\cite{abate19}, the intuition there is that if the programmer makes certain assumptions on what a program does, these assumptions also hold for the compiled program.
+In that definition, indicate hyperproperties~\cite{clarkson08} with $\Pi$ and classes of hyperproperties (i.e., sets of $\Pi$) as $\cC$.
+A program $p$ robustly satisfies class $\cC$ (written $\rsat{p}{\cC}$) if its behaviour is included in an element of $\cC$ when linked with an arbitrary program context.
+Similarly, for some $\Pi\in\cC$, we write $\rsat{p}{\Pi}$ whenever $p$ robustly satisfies $\Pi$.
+% 
 \begin{definition}[Robust Compilation]\label{def:rtp}
@@
 \end{definition}
-\noindent
+% \noindent
 %So, given a hyperproperty $\Pi$ from some class $\cC$ and a source program $\src{p}$ that robustly satisfies $\Pi$ ($\rsat{\src{p}}{\Pi}$), the compiled program $\stcomp{\src{p}}$ also needs to robustly satisfy $\Pi$.
-The intuition of \Cref{def:rtp} is as follows: If the programmer makes certain assumptions on what a program does, these assumptions should also hold for the compiled program.
-
 In practice, (robust) compilers are composed of numerous others.
@@
 \subsection{Simple Compositionality}
-We first consider function composition, i.e.\ plugging the result of one compiler into another one.
+We first consider function composition, i.e., plugging the result of one compiler into another one.
 Such pipelines happen when optimising source code (so, at the level of a suitable intermediate representation), but also on a higher level:
 Consider as an example a typical $\src{TypeScript}$ compilation pipeline.
-First, the compiler translates $\src{TypeScript}$ code to $\irl{JavaScript}$, which a part of V8 eventually, but also partially, compiles the code just-in-time to $\trg{Assembly}$.
-
+First, the compiler translates $\src{TypeScript}$ code to $\irl{JavaScript}$, which a part of V8 eventually compiles the code just-in-time to $\trg{assembly}$.
+% 
 \begin{definition}[Sequential Composition of Compilers]
-  Given two compilers $\sicomp{\bullet}$ and $\itcomp{\bullet}$, their sequential composition is $\sitcomp{\bullet}=\itcompN{\sicomp{\bullet}}$.
+  Given two compilers $\sicomp{\bullet}$ and $\itcomp{\bullet}$, their sequential composition is $\stcomp{\bullet}=\itcompN{\sicomp{\bullet}}$.
 \end{definition}
-Now, assuming that two compilers preserve certain classes, we know that their sequential composition preserves the least upper bound; the set intersection of those classes:
-\begin{lemma}[Sequential Composition with RTP]\label{lem:seqcompo}
+Assuming that two compilers preserve certain classes, their sequential composition preserves the least upper bound, i.e., the set intersection of those classes:
+\begin{lemma}[Sequential Composition with \gls{rc}]\label{lem:seqcompo}
   Given $\rtp{\sicomp{\bullet}}{\cC_{1}}$ and $\rtp{\itcomp{\bullet}}{\cC_{2}}$, then $\rtp{\sitcomp{\bullet}}{\cC_{1}\cap\cC_{2}}$.
 \end{lemma}
-Using an inductive argument, \Cref{lem:seqcompo} generalises to $n$ robust compilers, each preserving one of $n$ classes.
-To do so, one has to generalise the composition of two robust compilers to a set of $n$ robust compilers.
+Using an inductive argument, \Cref{lem:seqcompo} generalises to $n$ \gls{rc} compilers, each preserving one of $n$ classes.
+To do so, one has to generalise the composition of two \gls{rc} compilers to a set of $n$ ones.
 A real-world example for such deeply nested compositions is the $\src{TypeScript}$ compilation mentioned above.
@@
 We now consider a compiler that invokes two other compilers.
-$\src{Erlang}$ and $\irl{Elixir}$ are languages used in industry (e.g.\ by Discord) that are one example of such a composition.
-Both compile to $\trg{BEAM\ Bytecode}$.
+$\src{Java}$ and $\irl{Kotlin}$ are popular languages used in industry that are one example of such a composition and they both compile to $\trg{JVM\ Bytecode}$.
 
@@
 
-  $$\uhcsitcomp{\bullet}=\lambda p.\begin{cases}\stcomp{p} &\text{if }p\in\src{\partials}\\
-                                                \itcomp{p} &\text{if }p\in\irl{\partials}\end{cases}$$
+  \[
+    \uhcsitcomp{\bullet}=\lambda p. 
+      \begin{cases}
+        \stcomp{p} &\text{if }p\in\src{\partials}
+        \\
+        \itcomp{p} &\text{if }p\in\irl{\partials}
+      \end{cases}
+  \]
 \end{definition}
 We can derive a similar result to \Cref{lem:seqcompo} here, too:
-\begin{lemma}[Upper Composition with RC]\label{lem:useqcompo}
+\begin{lemma}[Upper Composition with \gls{rc}]\label{lem:useqcompo}
   Given $\rtp{\stcomp{\bullet}}{\cC_{1}}$ and $\rtp{\itcomp{\bullet}}{\cC_{2}}$, then $\rtp{\uhcsitcomp{\bullet}}{\cC_{1}\cap\cC_{2}}$.
@@
 With the same idea, we define a dual composition that goes from a single source language to multiple target languages.
-\texttt{dune} is a build system which can be used to compile $\src{OCaml}$ code to both $\irl{Assembly}$ and $\trg{Caml\ Bytecode}$.
+\texttt{dune} is a build system which can be used to compile $\src{OCaml}$ code to both $\irl{assembly}$ and $\trg{Caml\ Bytecode}$.
 \begin{definition}[Lower Composition]
@@
 
-\begin{lemma}[Lower Composition with RC]
+\begin{lemma}[Lower Composition with \gls{rc}]
   Given $\rtp{\stcomp{\bullet}}{\cC_{1}}$ and $\rtp{\sicomp{\bullet}}{\cC_{2}}$, then $\rtp{\lhcsitcomp{\bullet}}{\cC_{1}\cap\cC_{2}}$.
 \end{lemma}
-As before, this can be generalized to an arbitrary number, which also has a connection to the real-world, given by the diverse set of Assembly language dialects.
+As before, this can be generalized to an arbitrary number of compilers, which also has a connection to the real-world, given by the diverse set of assembly language dialects.
 
 The following free theorem (\Cref{lem:swappable}) is a direct consequence of \Cref{lem:seqcompo} where the involved compilers' input and output are both partial programs in the same language.
-Given that some compiler passes are attaining \gls{rc}, we can combine them in an arbitrary order and preserve the same least upper bound.
+Given that some compiler passes attain \gls{rc}, they can be combined in an arbitrary order and the result preserves the same least upper bound.
 A compiler's pipeline ordering is difficult and often hand-tuned.
@@
 Consider any stereotypical compilation pipeline.
-The security researcher has a strong interest to ensure that properties at the source level are preserved at the target level.
-Thus, if source programs robustly satisfy some property, so should the target ones.
-But, it might not be necessary for compilation passes from one intermediate representation to the other to preserve properties robustly since compiler intermediate representations are outside the considered attack vector.
-So, there might be some stronger property a pass has to satisfy in order to render the whole compilation pipeline secure.
+Programmers want properties at the source level to be preserved at the target level.
+Thus, if source programs robustly satisfy some property, so should their compiled counterparts.
+Unfortunately, it might not be necessary for compilation passes from one intermediate representation to the other to preserve properties robustly.
+This also has a security justification since compiler intermediate representations are not where typical attackers reside (i.e., the target language).
+So, there might be some stronger property a pass has to satisfy in order to render the whole compilation pipeline secure: this is what we study next.
 
@@
 \end{lstlisting}
-We now have an attack to violate memory safety: call the function with an invalid pointer.
+We now have an attack to violate memory safety: call the function with an invalid pointer and the program dereferences an invalid pointer.
 
-To prevent such issues we can use instrumentation passes that \textit{enforce} memory safety by adding dynamic checks to the program and crashing appropriately.
-There exist several memory-safety instrumentations, both for target level~\cite{watson15,aurajo16,vassena19,kouwe17,tarditi2018,sammler21} and source level~\cite{nagarakatte10,akritidis09,rigger17}.
+To prevent such issues we can use instrumentation passes that \textit{enforce} memory safety by adding dynamic checks to the program and crashing appropriately when a violation is detected.
+There exist several memory-safety instrumentations, both for target~\cite{watson15,aurajo16,vassena19,kouwe17,tarditi2018,sammler21} and source languages~\cite{nagarakatte10,akritidis09,rigger17}.
 %So, even though the programmer has written a memory-insecure program, we might still be interested in a memory-safe binary.
 
-We plan to extend our work with code instrumentations, which enforce classes.
-A sketch of the theory is as follows.
+We now sketch how to extend our work with instrumentations, which enforce specific classes of hyperproperties.
+% A sketch of the theory is as follows.
 
 \begin{definition}[Secure Instrumentation for Preserving $\cC$]\label{def:secure-instrumentation}
-  A secure instrumentation with respect to some class $\cC$ enforces hyperproperties described by some class $\cC'$ without violating $\cC$-satisfying programs.
-  We write $\sinstr{\stcomp{\bullet}}{\cC'}{\cC}$.
+  A secure instrumentation with respect to some class $\cC$ is a pass that enforces hyperproperties described by some other class $\cC'$ without violating $\cC$-satisfying programs.
+  We denote such a secure instrumentation as: $\sinstr{\stcomp{\bullet}}{\cC'}{\cC}$.
 \end{definition}
-Using this, we firstly want to inspect an insecure compilation chain from e.g.\ memory-safe $\src{Rust}$ to optimised, insecure $\irl{C}$, to memory-safe $\trg{CheckedC}$:
+Using this, we firstly want to inspect an compilation pipeline from memory-safe $\src{Rust}$ to optimised, insecure $\irl{C}$, to memory-safe $\trg{CheckedC}$.
+Intuitively, we want to be able to state that this pipeline preserves memory safety, despite the fact that the pass to $\irl{C}$ does not.
+% We want to be able to reconstruct memory safety for $\trg{CheckedC}$ despite the fact that $\irl{C}$ did not have it.
 \begin{example}[Enforcement may preserve...]
-  Given classes $\cC_{1},\cC_{2}$ and compilers $\sicomp{\bullet}$,$\itcomp{\bullet}$, we assume:
+  Given classes $\cC_{1}, \cC_{2}$ (resp. no property and memory safety, in our $\src{Rust}$ to $\trg{CheckedC}$ example) and compilers $\sicomp{\bullet}$,$\itcomp{\bullet}$, if:
 
@@
 
-Likewise, running a compiler that does not respect e.g.\ memory-safety after a memory-safety instrumentation nullifies the preservation of it:
+Dually, running a compiler that does not respect memory-safety after a memory-safety instrumentation nullifies its preservation:
 \begin{example}[...but, order matters!]
-  Given classes $\cC_{1},\cC_{2}$ and compilers $\sicomp{\bullet}$,$\itcomp{\bullet}$, we assume:
+  Given classes $\cC_{1},\cC_{2}$ and compilers $\sicomp{\bullet}$,$\itcomp{\bullet}$, if:
 
@@
 %Another area is its specialisation to instances, such as $\cHSafety$, and subsets thereof.
-We intend to study the compositionality aspects of concrete instances, such as SPECTRE~\cite{exorcisingSpectres,guarnieri19}, memory safety~\cite{azevedodeamorim18,azevedodeamorim15,dhawan15}, or cryptographic constant-time~\cite{barthe18}.
+Beyond this general theory, we also intend to study the compositionality aspects of concrete hyperproperties, such as Speculative Non-Interference~\cite{guarnieri19}, memory safety~\cite{azevedodeamorim18,azevedodeamorim15,dhawan15}, and cryptographic constant-time~\cite{barthe18}.
 
@@
 \bibliography{library}
+% \bibliography{./../library}
 

### 2021-10-28 prisc.tex
Fix minor typo

@@
 \end{definition}
-Using this, we firstly want to inspect an compilation pipeline from memory-safe $\src{Rust}$ to optimised, insecure $\irl{C}$, to memory-safe $\trg{CheckedC}$.
+Using this, we firstly want to inspect a compilation pipeline from memory-safe $\src{Rust}$ to optimised, insecure $\irl{C}$, to memory-safe $\trg{CheckedC}$.
 Intuitively, we want to be able to state that this pipeline preserves memory safety, despite the fact that the pass to $\irl{C}$ does not.

### 2022-07-06 coq/main.tex
no pdfs, file for me

\documentclass[a4paper,names,dvipsnames]{article}

\usepackage[T1]{fontenc}
\usepackage[colorinlistoftodos]{todonotes}
\usepackage[inference]{semantic}
\usepackage{mathrsfs}
\usepackage{mathtools}
\usepackage{xcolor}
\usepackage{hyperref}
\usepackage{glossaries}
\usepackage{enumerate}
\usepackage{enumitem}
\usepackage{amsmath}
\usepackage{amsthm}
\usepackage{amssymb}
\usepackage{stmaryrd}
\usepackage{marvosym}
\usepackage[capitalize]{cleveref}
\usepackage{bm}
\usepackage{xspace}
\usepackage{etoolbox}
\usepackage{natbib}
\usepackage{xfrac}

% workaround to use extpfeil and stmaryrd together.
% see https://tex.stackexchange.com/questions/297108/
\expandafter\def\csname opt@stmaryrd.sty\endcsname
{only,shortleftarrow,shortrightarrow}
\usepackage{extpfeil}

\usepackage{tikz}
\usetikzlibrary{positioning}

\newtheorem{example}{Example}
\newtheorem{definition}{Definition}
\newtheorem{lemma}{Lemma}
\newcounter{theoremcounter}
\newcounter{importanttheoremcounter}
\newtheorem{theorem}[theoremcounter]{Theorem}
\newtheorem{importanttheorem}[importanttheoremcounter]{Theorem}

% \usepackage{mmmacros}
\usepackage{./../mmmacros}
\loadglsentries{./../acronyms}
% \loadglsentries{acronyms}
\makeglossaries

\newcommand{\MK}[1]{\todo[color=orange!30]{TODO: #1}}
\newcommand{\MP}[1]{\todo[color=blue!30]{TODO: #1}}

% Definition of Ddots
\makeatletter
\def\Ddots{\mathinner{\mkern1mu\raise\p@
\vbox{\kern7\p@\hbox{.}}\mkern2mu
\raise4\p@\hbox{.}\mkern2mu\raise7\p@\hbox{.}\mkern1mu}}
\makeatother

\begin{document}

\tableofcontents
\clearpage

\section{Preliminaries}

\begin{definition}[Events]\label{def:events}
  $\events$ is the set of atomic propositions hereby called events or actions.
  The internal action is $\emptyevent$.
  The action $\terminationevent$ is program termination.
\end{definition}

We assume that any programming language can be enriched with a self-composition operator in the style of \citet{barthe11}.
Furthermore, we also assume the existence of a low-equivalence relation that distinguishes program states only by their public memory.
Two traces $\trace_{1}$,$\trace_{2}$ are low-equivalent $\loweq{\trace_{1}}{\trace_{2}}$ iff all their public events coincide.

\begin{definition}[Programming Languages]\label{def:pl}
A programming language is a tuple $\left(\partials,\wellf,\singlestep,\linker\right)$ s.t.:

\begin{itemize}
  \item[$\partials$] : Set - is a set of admissible, partial programs.
  \item[$\wellf$] : $\partials$ - a judgement that holds iff a program is not partial.
  \item[$\singlestep$] : $\wholes\to\events\to\wholes$ - a step relation, where $\wholes=\{w\in\partials\ |\ \wellf w\}$.
        For $e\in\events$ and $p,p'\in\wholes$ we say for $\estep{p}{e}{p'}$ that program $p$ performs a step with action $e$ to program $p'$.
        If $e=\emptyevent$, we write $\step{p}{p'}$.
        In case $e=\terminationevent$, we write $\terminates{p}$.
  \item[$\linker$] : $\partials\to\partials\to\partials$ - links two partial programs together in some way, resulting in a new partial program.
\end{itemize}
\end{definition}
Let $\src{S},\irl{I},$ and $\trg{T}$ be any programming language.

% if we get τ1 and τ2, then there is aontehr program that does τ3 which does τ1 and τ2 in parallel
% define self-composition. and then define it with respect to low-equivalent programs
%

\begin{definition}[Notation for Sequences]
  For any sequence of events, let $\seqnil$ denote the empty sequence and $\seqcons{e}{\bar{t}}$ the sequence that starts with $e$ and continues with $\bar{t}$.
  Hereby, it does not matter whether $\bar{t}$ is finite or infinite, it's merely syntactic sugar to work on sequences of events.
\end{definition}

\begin{definition}[Traces]
  A trace $\trace$ is an infinite sequence of events that results from the relation $\singlestep$.
  That is, we obtain the trace $\trace=\seqcons{e_{0}}{\seqcons{e_{1}}{\dots}}$ for the execution sequence $\estep{p}{e_{0}}{\estep{p'}{e_{1}}{\dots}}$ and write $\mktrace{p}{\trace}$.
  The set of all traces is $\traces$.
\end{definition}
\noindent
We assume $\lightning$ to occur in traces representing terminating programs such that it occurs infinitely often in a one-by-one sequence.

\begin{definition}[Finite Trace Prefixes]
  A finite sequence of events $m$ is a finite trace prefix of $\trace$ iff it satisfies the following judgement.

  $$
    \inference{}{\cdot\le\trace}\hspace{2em}\inference{m\le\trace}{\seqcons{e}{m}\le \seqcons{e}{\trace}}
  $$
\end{definition}

\begin{definition}[Behavior]
  The behavior of a whole program $p$ is a set of all traces it produces, i.e. $\behav{p}=\{\trace\ |\ \mktrace{p}{\trace}\}$.
\end{definition}

\begin{definition}[Observation]
  An observation is a finite set of finite trace prefixes.
  We say that an observation $o$ is the prefix of a behavior $b$ iff $$\forall m\in o.\exists \trace\in b.m\le t$$.
\end{definition}

\begin{definition}[Properties]
  A property $\prop$ is a set of admissible traces. For a program $p$ to satisfy $\prop$ it must not produce a trace that is not part of $\prop$. Thus, $p$ satisfies $\prop$ iff $\behav{p}\subseteq\prop$ and we write $\sat{p}{\prop}$.
\end{definition}

\begin{definition}[Hyperproperties]
  A hyperproperty $H$ is a set of admissible sets of traces. Thus, if $p$ satisfies $H$ (also written $\sat{p}{H}$), then $\behav{p}\in H$.
\end{definition}

\begin{lemma}[Lifting Properties]
  Given a property $\pi$, there exists a unique hyperproperty $\lift{\pi}$ that satisfies the exact same policy.
\end{lemma}
\begin{proof}
  We want $\forall p \in\partials, \sat{p}{\prop}\equiv\sat{p}{\lift{\prop}}$.
  Henceforth, given a $p\in\partials$, we have $\behav{p}\subseteq\prop$ iff $\behav{p}\in\lift{\prop}$.
  Note that if $\behav{p}\subseteq\prop$, we have $\behav{p}\in\left\{\Pi\ |\ \Pi=\behav{p}\subseteq\prop\right\}$.
  Thus, $\lift{\prop}$ is the set of all possible program behaviors that are a subset of $\prop$.
  This is exactly the powerset of $\prop$ and we conclude $\lift{\prop}=\powerset{\prop}$.
\end{proof}
\noindent
The lifting of properties to a singleton set does not suffice, since the empty behavior trivially satisfies any property $\emptyset\subseteq\prop=\{\trace\}$, but if we would define $\lift{\trace}=\{\{\trace\}\}$, then $\emptyset\not\in\lift{\trace}$.

\begin{lemma}[Property Satisfaction Refinement]
  For a property $\prop$ that refines $\prop'$, i.e. $\prop\subseteq\prop'$, if any $p\in\partials$ satisfies $\sat{p}{\prop'}$, then $\sat{p}{\prop}$.
\end{lemma}
\begin{proof}
  Pick any property $\prop'$ and $p\in\partials$ such that $\sat{p}{\prop}$ and assume $\prop\subseteq\prop'$.
  Simple unfolding reveals $\behav{p}\subseteq\prop\implies\behav{p}\subseteq\prop'$.
\end{proof}
\noindent
For lifted properties, this refinement property also holds on the hyperproperty level.
However, it does not work for any hyperproperty~\cite{clarkson08}.

\begin{definition}[Robust Property Satisfaction]
  A program $p$ robustly satisfies a property $\prop$, written $\rsat{p}{\prop}$, iff $\forall C\in\partials,\sat{C\linker p}{\prop}$. The same notation is used for robust hyperproperty satisfaction.
\end{definition}

\begin{lemma}[Weakening Robust Satisfaction]\label{lem:weaken-rsat}
  Given classes $\cC_{1}, \cC_{2}$ and any program $p$ such that
  \begin{assumptions}
    \item\label{lem:weaken-rsat:ass:a} $\cC_{1}\subseteq\cC_{2}$
    \item\label{lem:weaken-rsat:ass:b} $\rsat{p}{\cC_{2}}$
  \end{assumptions}
  We show
  \begin{goals}
    \item\label{lem:weaken-rsat:goal:i} $\rsat{p}{\cC_{1}}$
  \end{goals}
\end{lemma}
\begin{proof}
  Unfolding \Thmref{lem:weaken-rsat:goal:i}, let $\Pi\in\cC_{2}$ and $p$ be a program, we want to show that $\rsat{p}{\Pi}$.
  By \Thmref{lem:weaken-rsat:ass:a}, we also know that $\Pi\in\cC_{1}$.
  Thus, we can use \Thmref{lem:weaken-rsat:ass:b} to conclude.
\end{proof}

\begin{definition}[Classes]
  A class of hyperproperties $\cC$ is a set of hyperproperties.
  Likewise, a class of properties $\cC$ is a set of hyperproperties, where every property is lifted to the hyperproperty level.
  From now on, we use $\Pi$ for elements of any class $\cC$ in case it does not matter whether it is a lifted property or any hyperproperty.
\end{definition}

\begin{definition}[Compilers]
  A compiler between languages $\S$ and $\T$ is a partial function $\stcomp{\bullet}$ from $\src{\partials}$ to $\trg{\partials}$.
\end{definition}

\section{Compositionality of Secure Compilers}

\begin{definition}[Robust Trace-Hyperproperty Preservation]\label{def:rtp}
  For a given class $\cC$, a compiler from languages $\S$ to $\T$ robustly preserves $\cC$ iff
  $$
  \forall\Pi\in\cC,\forall\src{p}\in\src{\partials},\rsat{\src{p}}{\Pi}\implies\rsat{\stcomp{\src{p}}}{\Pi}
  $$
  We write $\rtp{\stcomp{\bullet}}{\cC}$.
  In case we write $\rtp{\stcomp{\bullet}}{\collapse{\cC}}$, substitute the $\cC$ in above definition with $\left\{ \collapse{\cC} \right\}$.
\end{definition}

\begin{definition}[Sequential Composition of Compilers]
  Given two compilers $\sicomp{\bullet}$ and $\itcomp{\bullet}$, their sequential composition is $\sitcomp{\bullet}=\itcompN{\sicomp{\bullet}}$.
\end{definition}
% we can propagate through the assumptions from src langs to intermediate langs
%

\begin{lemma}[Weakening RTP]\label{lem:weaken}
  Given classes $\cC_{1}, \cC_{2}$ such that
  \begin{assumptions}
    \item $\cC_{1}\subseteq\cC_{2}$
    \item $\rtp{\stcomp{\bullet}}{\cC_{2}}$
  \end{assumptions}
  We show
  \begin{goals}
    \item $\rtp{\stcomp{\bullet}}{\cC_{1}}$
  \end{goals}
\end{lemma}
\begin{proof}
  Using \Cref{def:rtp} on the goal, let $\Pi\in\cC_{1}$ and $\src{p}\in\src{\partials}$ such that $\rsat{\src{p}}{\Pi}$, so what's left to prove is $\rsat{\stcomp{p}}{\Pi}$.
  Since $\cC_{1}\subseteq\cC_{2}$ and $\Pi\in\cC_{1}$, we know that $\Pi\in\cC_{2}$.
  Thus, we can apply the assumption $\rtp{\stcomp{\bullet}}{\cC_{2}}$ to our goal, leaving us with $\rsat{\src{p}}{\Pi}$ to show, which was an assumption we made.
\end{proof}

\begin{definition}[Safety Properties]
  The class of safety properties contains the lifting of all properties that can be refuted with a finite trace prefix:
  $$
  \cSafety = \left\{\lift{\prop}\ |\ \forall \trace\in\traces, t\not\in\lift{\prop} \text{ iff } \exists m\ge\trace,\forall \trace'\in\traces,m\le\trace'\implies\trace'\not\in\lift{\prop}\right\}
  $$
\end{definition}

\begin{definition}[Hypersafety Properties]\label{def:hsafety}
  The class of hypersafety properties contains all hyperpropert that can be refuted with an observation:
  $$
  \cHSafety = \left\{\Pi\ |\ \forall b\in 2^{\traces},b\not\in\Pi\text{ iff  }\exists o\ge b,\forall b'\in 2^{\traces},o\le b'\implies b'\not\in\Pi\right\}
  $$
\end{definition}

\begin{lemma}[Safety is entailed in Hypersafety]
  $\cSafety\subseteq\cHSafety$.
\end{lemma}

\begin{definition}[Subset Closed Hyperproperties]
  The class of hyperproperties that are closed with respect to the subset relation is
  $$
  \cSS = \left\{H\ |\ \forall X\in H, \forall Y\subseteq X, Y \in H\right\}
  $$
\end{definition}

\begin{lemma}[Hypersafety is entailed in SSC]
  $\cHSafety\subseteq\cSS$.
\end{lemma}

\begin{definition}[K-Hypersafety]
  Exactly the same as \Cref{def:hsafety}, but the observations $o$ are restricted to cardinality $k$.
  2-Hypersafety is simply $k=2$. \Cref{def:ni} gives an example instance of a classic 2-hypersafe property.
\end{definition}

\begin{definition}[Nontermination]
  Nontermination is a safety property: Any finite trace prefix with $\terminationevent$ violates this property.
\end{definition}

\begin{definition}[Mutual Exclusion]
  Mutual Exclusion (MutEx) means that no two processes during a program execution enter the same critical section.
\end{definition}

\begin{definition}[Determinism]
  Determinism is a 2-hypersafety property, since any two program executions with the same input must yield the same output.
\end{definition}

\begin{definition}[Non-Interference ($\Ni$)]\label{def:ni}
  We define the class containing the non-interference hyperproperty as:
  $$
  \Ni = \left\{ H | \forall \trace_{1},\trace_{2}\in H. \loweq{\trace_{1}}{\trace_{2}}\implies\trace_{1}=\trace_{2} \right\}
  $$
\end{definition}
Note that $=$ may not be strict equality, but some suitable trace equivalence that checks both public and private actions, instead of just public.

\begin{definition}[Average Response Time (ART)]
  The average response time over all executions less than some arbitrary constant~\cite{clarkson08} is a practically useful hyperproperty that is not subset closed.
\end{definition}
Consider the behavior $\{\trace_{1},\trace_{2},\trace_{3}\}$ where $\trace_{1},\trace_{2}$ each take 1 second and $\trace_{3}$ 4 seconds to run.
If the chosen constant is 2, the hyperproperty would be fulfilled. However, the subset behavior $\{\trace_{3}\}$ has an average time of 4, double the bound set by the property.

\begin{lemma}[Classes Lattice]
  Classes with $\subseteq$ form a lattice.
\end{lemma}
\begin{proof}
  The class with all hyperproperties is the top element $\top=\powerset{\traces}$, while the class with no hyperproperties whatsoever is $\bot=\{\}$.
  Obviously, a partial order due to set inclusion.
  $\cap$ is least upper and $\cup$ greatest lower bound of any given pair of subsets of a given class.
\end{proof}

\begin{figure}[h]
  \centering
  \begin{tikzpicture}[node distance=4mm,every node/.style={align=center}]
    \node (top) {$\top$};
    \node[below = of top.south] (SSC) {$\cSS$};
    \node[below = of SSC.south] (HSafe) {$\cHSafety$};
    \node[below = of HSafe.south] (HKSafe) {$\cKHSafety$};
    \node[below = of HKSafe.south] (H2Safe) {$\cTwoHSafety$};
    \node[below = of H2Safe.south] (Safe) {$\cSafety$};

    % Specialized classes
    \node[below right = of Safe.south] (Mutex) {$\mutex$};
    \node[below left = of Safe.south,xshift = 0.75em] (NonTerm) {$\nonterm$};

    \node[below left = of H2Safe.south west] (Determ) {$\determ$};
    \node[below right = of H2Safe.south east] (NI) {$\Ni$};

    \node[below = of Safe.south] (emptyspace) {};
    \node[below = of emptyspace.south] (Bot) {$\bot$};

    % edges
    \draw[-] (top.south) -- (SSC.north);
    \draw[-] (SSC.south) -- (HSafe.north);
    \draw[-] (HSafe.south) -- (HKSafe.north);
    \draw[-] (HKSafe.south) -- (H2Safe.north);
    \draw[-] (H2Safe.south) -- (Safe.north);
    \draw[-] (H2Safe.south) -- (Determ.north);
    \draw[-] (H2Safe.south) -- (NI.north);
    \draw[-] (Safe.south) -- (Mutex.north);
    \draw[-] (Safe.south) -- (NonTerm.north);
    \draw[-] (Determ.south) |- (Bot.west);
    \draw[-] (NI.south) |- (Bot.east);
    \draw[-] (Mutex.south) -- (Bot.north);
    \draw[-] (NonTerm.south) -- (Bot.north);
  \end{tikzpicture}
  \caption{Sketch of a fraction of the infinitely sized lattice of classes. The diagram should be read from top to bottom. A connecting edge between two nodes means that the bottom one is a subset of the top one.}
\end{figure}

\begin{lemma}[Sequential Composition with RTP]\label{lem:seqcompo}
  Given $\rtp{\sicomp{\bullet}}{\cC_{1}}$ and $\rtp{\itcomp{\bullet}}{\cC_{2}}$, then $\rtp{\sitcomp{\bullet}}{\cC_{1}\cap\cC_{2}}$.
\end{lemma}
\begin{proof}
  We need to show $\rtp{\sitcomp{\bullet}}{\cC_{1}\cap\cC_{2}}$.
  By definition, assume $\Pi\in\cC_{1}\cap\cC_{2}$ and $\src{p}\in\src{\partials}$ such that $\rsat{\src{p}}{\Pi}$.
  What is left to show is $\rsat{\sitcomp{\src{p}}}{\prop}$.
  Note that $\Pi\in\cC_{2}$ and that $\sicomp{\src{p}}\in\irl{\partials}$, allowing us to apply $\rtp{\itcomp{\bullet}}{\cC_{2}}$ changing our goal to $\rsat{\sicomp{\src{p}}}{\Pi}$.
  Since $\Pi\in\cC_{1}$ also holds, we can this time apply $\rtp{\stcomp{\bullet}}{\cC_{1}}$.
  What is left to show is $\rsat{\src{p}}{\Pi}$, which is an assumption of ours.
\end{proof}

\begin{definition}[Upper Composition]
  Given two compilers $\stcomp{\bullet}$ and $\itcomp{\bullet}$, their upper composition is

  $$\uhcsitcomp{\bullet}=\lambda p.\begin{cases}\stcomp{p} &\text{if }p\in\src{\partials}\\
                                                \itcomp{p} &\text{if }p\in\irl{\partials}\end{cases}$$.
\end{definition}

\begin{lemma}[Upper Composition with RTP]
  Given $\rtp{\stcomp{\bullet}}{\cC_{1}}$ and $\rtp{\itcomp{\bullet}}{\cC_{2}}$, then $\rtp{\uhcsitcomp{\bullet}}{\cC_{1}\cap\cC_{2}}$.
\end{lemma}
\begin{proof}
  Analogous argument as in \Thmref{lem:seqcompo}, but with a case distinction on whether the source program is element of $\S$ or $\I$.
\end{proof}
% S = while,    I = while with exceptions           => modularization

% another idea: certified nugget -> minimal RSP compiler

\begin{definition}[Lower Composition]
  Given two compilers $\stcomp{\bullet}$ and $\sicomp{\bullet}$, their lower composition is $\lhcsitcomp{\bullet}$.
\end{definition}

\begin{lemma}[Lower Composition with RTP]
  Given $\rtp{\stcomp{\bullet}}{\cC_{1}}$ and $\rtp{\sicomp{\bullet}}{\cC_{2}}$, then $\rtp{\lhcsitcomp{\bullet}}{\cC_{1}\cap\cC_{2}}$.
\end{lemma}
\begin{proof}
  Analogous argument as in \Thmref{lem:seqcompo}, but with a case distinction on whether the compiled source program is element of $\I$ or $\T$.
\end{proof}

\begin{lemma}[Diamond]\label{lem:diamond}
  Given $\rtp{\lhcsiocomp{\bullet}}{\cC_{1}}$ and $\rtp{\uhciotcomp{\bullet}}{\cC_{2}}$ with $\stcomp{\bullet} = \lambda\src{p}.\uhciotcomp{\lhcsiocomp{p}}$, then $\rtp{\stcomp{\bullet}}{\cC_{1}\cap\cC_{2}}$.
\end{lemma}
\begin{proof}
  Straightforward using \Thmref{lem:seqcompo}.
\end{proof}

\begin{lemma}[Swappable]\label{lem:swappable}
  Given $\rtp{\ttcomp{\bullet}_{(1)}}{\cC_{1}}$ and $\rtp{\ttcomp{\bullet}_{(2)}}{\cC_{2}}$, then $\rtp{\ttcompN{\ttcomp{\bullet}_{(2)}}_{(1)}}{\cC_{1}\cap\cC_{2}}$ and $\rtp{\ttcompN{\ttcomp{\bullet}_{(1)}}_{(2)}}{\cC_{1}\cap\cC_{2}}$.
\end{lemma}
\begin{proof}
  Both follow from \Thmref{lem:seqcompo}.
\end{proof}

\begin{lemma}[Mingle]
  Given
  \begin{assumptions}
  \item\label[ass]{lem:sandwich:ass:a} $\forall\Pi\in\cC_{2}.\forall\src{p}\in\src{\partials}.\rsat{\src{p}}{\Pi}\implies \Pi\in\cC_{1}$
  \item\label[ass]{lem:sandwich:ass:b} $\forall\Pi\in\cC_{1}.\forall\trg{p}\in\trg{\partials}.\rsat{\trg{p}}{\Pi}\implies \Pi\in\cC_{2}$ (not needed)
  \item\label[ass]{lem:sandwich:ass:c} $\rtp{\stcomp{\bullet}}{\cC_{1}}$
  \end{assumptions}
  we have
  \begin{goals}
  \item\label[goal]{lem:sandwich:goal:i} $\rtp{\stcomp{\bullet}}{\cC_{2}}$
  \end{goals}
\end{lemma}
\begin{proof}
  Unfolding \Cref{lem:sandwich:goal:i}; Let $\Pi\in\cC_{2}$ and $\src{p}\in\src{\partials}$ such that $\rsat{\src{p}}{\Pi}$, leaving us with $\rsat{\stcomp{\src{p}}}{\Pi}$ to prove.

  Apply \Cref{lem:sandwich:ass:c}, giving us as new goals $\Pi\in\cC_{1}$ and $\rsat{\src{p}}{\Pi}$, the latter following easily by the exact same assumption made before.

  However, we still need to argue that $\Pi\in\cC_{1}$.

  Instatiate the universal quantifications in \Cref{lem:sandwich:ass:a} with $\Pi$ and $\src{p}$. Apply that to our goal.
  What is left to show is $\rsat{\src{p}}{\Pi}$, which we've proven before already.
\end{proof}

\begin{lemma}[RSP may imply k-RSHP]
  Given
  \begin{assumptions}
  \item\label[ass]{lem:rspimplkrshp:ass:a} $\S$ and $\T$ have $\loweq{}{}$ and $\selfcompo{}{}$ operators
  \item\label[ass]{lem:rspimplkrshp:ass:b} $\Pi\in\cSafety$
  \item\label[ass]{lem:rspimplkrshp:ass:c} $\src{p}\in\src{\partials}$ with $\rsat{\src{p}}{\Pi}$
  \item\label[ass]{lem:rspimplkrshp:ass:d} $\rtp{\stcomp{\bullet}}{\cSafety}$
  \end{assumptions}
  Then
  \begin{goals}
  \item\label[goal]{lem:rspimplkrshp:goal:i} $\exists\stcomp{\bullet}.\sat{\stcomp{\src{p}}}{\text{k-RSHP}}$
  \end{goals}
\end{lemma}
\begin{proof}
  (intuition)

  Take $\src{p}\in\src{\partials}$ and construct the k-product program according to \citet{barthe11}.
  Feed the input into the compiler from \Cref{lem:rspimplkrshp:ass:d}.
  Instantiate the existential in \Cref{lem:rspimplkrshp:goal:i} with the modified compiler.
  Since the product program reflects the behavior of $\src{p}$, the transformation does not interfere with robust property satisfaction.
  Furthermore, \Cref{lem:rspimplkrshp:ass:d} guarantees us that the compiled k-product program of $\src{p}$ is safe.
  Since the k-product construction soundly embeds $k-RSHP$ properties into $\cSafety$, the claim follows.
\end{proof}

\begin{definition}[Instrumentation]\label{def:instrumentation}
  A source code instrumentation done by a compiler takes a program in language $\T$ and emits an altered version of the same program in $\T$ such that it does not go wrong for some class of security relevant (hyper-)properties.
  Formally, given a $\cC$ a compiler $\ttcomp{\bullet}$ is an instrumentation for $\cC$ iff:

  $$
  \forall \trg{p}\in\trg{\partials}, \rsat{\ttcomp{p}}{\cC}
  $$

  We write $\instr{\ttcomp{\bullet}}{\cC}$.
\end{definition}
% S -> I and I -> T
% I -> T preserves memory safety

% enrich pipeline from source to target to see how to do the preservation from source in intermediate/target lang

\begin{definition}[Secure Instrumentation with Respect to $\cC$]\label{def:secure-instrumentation}
  A secure instrumentation with respect to some class $\cC$ ensures (hyper-)properties described by some class $\cC'$ without violating $\cC$-satisfying programs. We write $\sinstr{\ttcomp{\bullet}}{\cC}{\cC'}$.
  Thus, a compiler $\ttcomp{\bullet}$ is a secure instrumentation for $\cC'$ respecting $\cC$ iff:

  $$
  \rtp{\ttcomp{\bullet}}{\cC}\text{ and } \instr{\ttcomp{\bullet}}{\cC'}
  $$

\end{definition}

\begin{lemma}[Sequential Composition with a Secure Instrumentation]\label{lem:seqcompsecinstr}
  Given compilers $\sicomp{\bullet}$, $\itcomp{\bullet}$ and classes $\cC_{0},\cC_{1}$ such that
  \begin{assumptions}
  \item\label[ass]{lem:seqcompsecinstr:ass:a} $\rtp{\sicomp{\bullet}}{\cC_{0}}$
  \item\label[ass]{lem:seqcompsecinstr:ass:b} $\sinstr{\itcomp{\bullet}}{\cC_{1}}{\cC_{0}}$
  \end{assumptions}
  Then
  \begin{goals}
  \item\label[goal]{lem:seqcompsecinstr:goal:i} $\rtp{\sitcomp{\bullet}}{\cC_{0}\cup\cC_{1}}$
  \end{goals}
\end{lemma}

%\begin{definition}[No-v* Classes]
%  Let $\nvOne$ denote the singleton-class such that no SPECTRE v1 attack can happen.
%  Analgously, define $\nvFour$ as a singleton-class that disallows SPECTRE v4 attacks and $\nvOneFour$ as the class that disallows both v1 and v4 attacks, i.e. $\nvOneFour=\nvOne\cap\nvFour$.
%\end{definition}

% \begin{example}
%   Consider a compiler $\ttcomp{\bullet}_{1}$ such that $\sinstr{\ttcomp{\bullet}_{1}}{\nvOne}{\nvFour}$ and, similarily,
%   $\ttcomp{\bullet}_{4}$ such that $\sinstr{\ttcomp{\bullet}_{4}}{\nvFour}{\nvOne}$.

%   \begin{lemma}\label{lem:vOnevFourSafe}
%     Using above compilers, we now show that $\rtp{\ttcompN{\ttcomp{\bullet}_{1}}_{4}}{\nvOneFour}$.
%   \end{lemma}
%   \begin{proof}
%     By \Cref{lem:seqcompo} we get two goals, $\rtp{\ttcomp{\bullet}_{1}}{\nvOneFour}$ and $\rtp{\ttcomp{\bullet}_{4}}{\nvOneFour}$.

%     Since both cases are symmetrical, we only consider $\rtp{\ttcomp{\bullet}_{1}}{\nvOneFour}$ now.

%     Unfolding the definition of the goal, suppose there is a $\Pi\in\nvOneFour$ and $\trg{p}\in\trg{\partials}$ such that $\rsat{\trg{p}}{\Pi}$, while our new goal is $\rsat{\ttcomp{p}_{1}}{\Pi}$.

%     Strengthen the goal so that we need to show $\rsat{\ttcomp{p}_{1}}{\Pi}\wedge\Pi\in\nvOneFour$. Now, apply \Cref{def:secure-instrumentation}, so the only things left to show are $\Pi\in\nvOne$ and $\rsat{\trg{p}}{\Pi}$.
%     Both follow easily from our assumptions.
%   \end{proof}
%   \begin{lemma}
%     $\rtp{\ttcompN{\ttcomp{\bullet}_{4}}_{1}}{\nvOneFour}$
%   \end{lemma}
%   \begin{proof}
%     By \Cref{lem:vOnevFourSafe} and \Cref{lem:swappable}.
%   \end{proof}

%   We now consider what happens if one of the two instrumentations is not secure with respect to the respective other class.
%   However, we focus on the case where the v1-instrumentation is insecure with respect to v4, due to symmetry.

%   \begin{lemma}
%     $\rtp{\ttcompN{\ttcomp{\bullet}_{1}}_{4}}{\nvOneFour}$
%   \end{lemma}
%   \begin{proof}
%     Note that $\ttcomp{p}_{1}$ is a secure $\T$ program with respect to SPECTRE v1 attacks, given any $\trg{p}\in\trg{\partials}$.

%     Unfolding, let $\Pi\in\nvOneFour$ and $\trg{p}\in\trg{\partials}$ such that $\rsat{\trg{p}}{\Pi}$, our goal changes to $\rsat{\ttcompN{\ttcomp{\trg{p}}_{1}}_{4}}{\Pi}$.

%     Strengthen what we want to prove to $\rsat{\ttcompN{\ttcomp{\trg{p}}_{1}}_{4}}{\Pi}\wedge\Pi\in\nvOneFour$.
%     Apply \Cref{def:secure-instrumentation}, giving us proof-obligations that are already part of our assumptions.
%   \end{proof}

%   \begin{lemma}
%     $\rtp{\ttcompN{\ttcomp{\bullet}_{4}}_{1}}{\nvOne}$
%   \end{lemma}
%   \begin{proof}
%     Let $\trg{p}=\ttcomp{p'}_{4}$ be some program that is the result of plugging $\trg{p'}\in\trg{\partials}$ into $\ttcomp{\bullet}_{4}$.

%     Our goal becomes $\rtp{\ttcomp{p}_{1}}{\nvOne}$. Unfolding it, let $\Pi\in\nvOne$ and $\trg{p}\in\trg{\partials}$ such that $\rsat{\trg{p}}{\Pi}$, where what is left to prove is $\rsat{\ttcomp{p}_{1}}{\nvOne}$.

%     The goal follows immediately by \Cref{def:instrumentation}, since $\ttcomp{p}_{1}$ is an instrumentation for $\nvOne$.
%   \end{proof}

%   Finally, if both are insecure instrumentations, the strongest result one can get is either the robust preservation of $\nvOne$ or $\nvFour$, depending on which compiler is run last.
% \end{example}

We have seen that we can already swap robust compilers with same input and output language in a compilation pipeline.
Now, we want to formalize this idea of ``being able to swap things around'' for source code instrumentations.

\begin{definition}[Swappable Instrumentations]
  We define vertical compositionality as follows.
  Let $\bar{\cC}$ be a set of classes that are eventually used in the compilation pipeline.
  Given a class $\cC\in\bar{\cC}$ the particular instrumentation $\ttcomp{\bullet}$ ensures any program to robustly satisfy, we can swap the instrumentation in the pipeline \textit{freely} around iff:

  $$
  \forall\cC'\in\bar{\cC}. \sinstr{\ttcomp{\bullet}}{\cC}{\cC'}
  $$
\end{definition}
\MK{it'd be nice to have a property-free version of this, since this here is TEDIOUS to prove in real-world}

\clearpage
Experimental
\begin{definition}[Encoders]
  An encoder $\tenc{\src{\bullet}}$ from language $\S$ to $\T$ is a function that constructs an abstract representation of the given $\S$ program as a $\T$ program.
\end{definition}
\noindent
Think of encoders as parsers.

\begin{definition}[Interpreters]
  An interpreter $\stinterp$ is a $\T$ program that faithfully implements the semantics of $\S$.
  That is, given $\src{p}\in\src{\partials}$ and some encoding function $\tenc{\src{\bullet}}$, then $\mktrace{\stinterp \trg{\linker} \tenc{\src{p}}}{\type} \Leftrightarrow \mktrace{\src{p}}{\type}$.
\end{definition}
Note: For an interpreter to \textit{faithfully} realize the semantics of $\S$, we really need the equivalence.
Say we only consider the $\Leftarrow$ direction as defining characteristic, then an interpreter that e.g. tosses a coin and thus $\behav{\stinterp\trg{\linker}\tenc{\src{p}}}=\{\trace_{1},\trace_{2}\}$ given $\behav{\src{p}}=\{\trace_{1}\}$ would be acceptable.
However, an interpreter that 99\% of the time emits $\trace_{2}$ is certainly not what we want.
Consider the converse, so $\Rightarrow$ as defining characteristic.
Then, swapping the behaviors from the previous example, the interpreter does not \textit{faithfully} simulate non-determinism that is present in the source language.
% Due to this, interpreters in our world always robustly satisfy relational hyperproperties

\begin{definition}[Specializers]
  A specializer $\ttspec{\bullet}$ is a $\T$ program that, given an encoding of another $\T$ program as input, emits a specialized version of that $\T$ program.
  So, given a partial program $\trg{p}$ and some set of inputs $\trg{x}\in\trg{\partials}$, $\ttspec{\tenc{p \linker x}}$ always terminates with a residual program $\trg{p'}$ such that $\behav{\trg{p\linker x}}=\behav{\trg{p'}}$.
\end{definition}
\noindent

\begin{definition}[2nd Futamura Projection]
  Given a specializer $\ttspec{\bullet}$, an interpreter $\stinterp$, and an encoding $\tenc{\src{\bullet}}$, we can compose them to a compiler $\stcomp{\bullet} = \ttspec{\stinterp \tenc{\src{\bullet}}}$.~\cite{Futamura1999}
\end{definition}

\clearpage

\section{Case Study}

\subsection{Source Language}
\subsubsection{Syntax}

\myfig{
	\vspace{-1em}
  \begin{gather*}
  \begin{aligned}
  \mi{Expressions}~\src{e} \bnfdef&\ \src{x} \mid \src{v} \mid \src{e_{1} \oplus e_{2}} \mid \src{x[e]} \mid \src{let\ x = e_{1}\ in\ e_{2}}\mid \src{x[e_{1}]\leftarrow e_{2}} \\
  & \mid \src{let\ x=new\ e_{1}\ in\ e_{2}} \mid \src{delete\ x} \\
  & \mid \src{if\ e_1<e_2\ then\ e_3\ else\ e_4} \mid \src{abort()} 
  \hspace{0.5cm} \text{where } \src{\oplus}\in\{\src{+},\src{-},\src{\times}\}\\
  %
  \mi{Types}~\types \bnfdef&\ \src{\nat} \mid \src{ref_q\ \nat} \hspace{0.25cm}
  %
  \mi{Qualifier}~\src{q} \bnfdef\ \full \mid \half\hspace{0.25cm}
  %
  \mi{Values}~\src{\valueexpr} \bnfdef\ \src{n}\in\src{\nat} \\
  \mi{References}~\src{\loc}&\in\src{\nat} \hspace{0.5cm}
  %
  \mi{Variables}~\ \src{x} \mid \src{y} \mid \dots \hspace{0.5cm}
  %
  \mi{Poison}~\ \src{\poison}~\bnfdef\ \src{\poisonless} \mid \src{\poisoned}\\
  %
  \mi{Typing. Env.}~\src{\Gamma} \bnfdef&\ \src{\hole{\cdot}} \mid \Gammas,\src{x:\type}\hspace{0.5cm}
  \mi{Store}~\src{\Delta} \bnfdef\ \src{\hole{\cdot}} \mid \src{x\mapsto\loc_\poison},\Deltas\\
  %
  \mi{Heaps}~\src{H} \bnfdef&\ \src{\hole{\cdot}} \mid \src{H}::\src{n} \hspace{0.5cm}
  %
  \mi{State}\ \configs \bnfdef\ \sconfig{H}{\Delta} \hspace{0.5cm}
  \mi{Marked\ State}\ \qconfigs[i] \bnfdef\ \configs\mid \qconfigs\\
  %
  \mi{Contexts}\ &\src{context\ e_0\ e_1} \hspace{0.5cm}
  \mi{Components}\ \src{component\ e_p} \\
  \mi{Whole\ Programs}\ &\src{prog\ e_0\ e_1\ e_2}\hspace{0.5cm}
  %
  \mi{Final\ Result}\ \src{\finalexpr} \bnfdef\ \src{\valueexpr} \mid \src{stuck} \\
  \end{aligned}
  \end{gather*}
}{mmla-syntax}{Syntax of $\mmlAs$}

\subsubsection{Static Semantics}
\begin{definition}[NoPtr]
  We write $\src{\noptr{\Gamma}}$ if for any $\src{x}, \src{\type}$, if $\src{x} : \src{\type}\in\src{\Gamma}$, then $\types\not=\ptr$.
\end{definition}

\myfig{
  \begin{center}
  \judgbox{\Gammas = \src{\Gamma_1}\circ\src{\Gamma_2}}{,,Environment $\Gammas$ can be split into $\src{\Gamma_1}$ and $\src{\Gamma_2}$.''}
  %
  $\;$\\
  \typerule{splitEmpty}{
  }{
    \src{\hole{\cdot}} = \src{\hole{\cdot}}\circ\src{\hole{\cdot}}
  }{t-empty-split}
  %
  \typerule{$\nat$split}{
    \Gammas = \src{\Gamma_1}\circ\src{\Gamma_2}
  }{
    \src{x:\nat},\Gammas = \src{x:\nat},\src{\Gamma_1}\circ\src{x:\nat},\src{\Gamma_2}
  }{t-nat-split}
  %
  \typerule{weakPtrSplit}{
    \Gammas = \src{\Gamma_1}\circ\src{\Gamma_2}
  }{
    \src{x:\wptr},\Gammas = \src{x:\wptr},\src{\Gamma_1}\circ\src{x:\wptr},\src{\Gamma_2}
  }{t-wptr-split}
  %
  \typerule{ptrLSplit}{
    \Gammas = \src{\Gamma_1}\circ\src{\Gamma_2}
  }{
    \src{x:\ptr},\Gammas = \src{x:\ptr},\src{\Gamma_1}\circ\src{\Gamma_2}
  }{t-ptr-l-split}
  %
  \typerule{ptrRSplit}{
    \Gammas = \src{\Gamma_1}\circ\src{\Gamma_2}
  }{
    \src{x:\ptr},\Gammas = \src{x:\wptr,\Gamma_1}\circ\src{x:\ptr},\src{\Gamma_2}
  }{t-ptr-r-split}
  \end{center}
}{w-ctx-split}{Context Splitting of $\mmlAs$ typing contexts.}

\myfig{
  \judgbox{\typechecks{\Gamma}{e}{\types}}{,,Under environment $\Gammas$ the expression $\src{e}$ has type $\types$.''}

  \begin{center}
  \typerule{$t-\src{var}$}{
  \src{\noptr{\Gamma_1}}&
  \src{\noptr{x:\type,\hole{\cdot}}}&
  \src{\noptr{\Gamma_2}}
	}{
    \typechecks{\Gamma_1,x:\type,\Gamma_2}{x}{\type}
	}{t-var}
  %
  \typerule{$t-\nat$}{
    \src{\noptr{\Gamma}}
  }{
    \typechecks{\Gamma}{n}{\nat}
  }{t-nat}
  %
  \typerule{$t-\src{\oplus}$}{
    \typechecks{\Gamma_1}{e_1}{\nat} &
    \typechecks{\Gamma_2}{e_2}{\nat}
  }{
    \typechecks{\Gamma_1\circ\Gamma_2}{e_1\oplus e_2}{\nat}
  }{t-binop}
  %
  \typerule{$t-get$}{
    \typechecks{\Gamma_2}{x}{\wptr} &
    \typechecks{\Gamma_1}{e}{\nat}
  }{
    \typechecks{\Gamma_1\circ\Gamma_2}{x[e]}{\nat}
  }{t-get}
  %
  \typerule{$t-set$}{
    \typechecks{\Gamma_3}{x}{\wptr} &
    \typechecks{\Gamma_1}{e_1}{\nat} &
    \typechecks{\Gamma_2}{e_2}{\nat}
  }{
    \typechecks{(\Gamma_1\circ\Gamma_2)\circ\Gamma_3}{x[e_1]\leftarrow e_2}{\nat}
  }{t-set}
  %
  \typerule{$t-\src{let}$}{
    \typechecks{\Gamma_1}{e_1}{\nat} &
    \typechecks{x:\nat,\Gamma_2}{e_2}{\type}
  }{
    \typechecks{\Gamma_1\circ\Gamma_2}{let\ x=e_1\ in\ e_2}{\type}
  }{t-let}
  %
  \typerule{$t-\src{new}$}{
    \typechecks{\Gamma_1}{e_1}{\nat} &
    \typechecks{x:\ptr,\Gamma_2}{e_2}{\nat}
  }{
    \typechecks{\Gamma_1\circ\Gamma_2}{let\ x=new\ e_1\ in\ e_2}{\nat}
  }{w-t-new}
  %
  \typerule{$t-\src{delete}$}{
    \src{\noptr{\Gamma_1}} & \src{\noptr{\Gamma_2}}
  }{
    \typechecks{\Gamma_1,x:\ptr,\Gamma_2}{delete\ x}{\nat}
  }{w-t-delete}
  %
  \typerule{$t-\src{if}$}{
    \typechecks{\Gamma_1}{e_1}{\nat} &
    \typechecks{\Gamma_2}{e_2}{\nat} &
    \typechecks{\Gamma_3}{e_3}{\type} &
    \typechecks{\Gamma_3}{e_4}{\type}
  }{
    \typechecks{\Gamma_1\circ\Gamma_2\circ\Gamma_3}{if\ e_1 < e_2\ then\ e_3\ else\ e_4}{\type}
  }{w-t-if}
  \end{center}
}{w-expr-ty}{Checking of $\mmlAs$ expressions.}

\myfig{
  \begin{center}
  \judgbox{closed\ X\ \src{e}}{,,At most all $\mmlAs$ variables in the set $X$ occur in $\src{e}$.''}
    \typerule{$program$}{
      \vdash_\copyright\src{component\ e_p}&
      \vdash_\copyright\src{context\ e_0\ e_1}
    }{
      \vdash_\copyright\src{prog\ e_0\ e_p\ e_1}
    }{w-program}
    %
    \typerule{$component$}{
      closed\ \{\src{x}\}\ \src{e}
    }{
      \vdash_\copyright\src{component\ e}
    }{w-component}
    %
    \typerule{$context$}{
      closed\ \emptyset\ \src{e_0} &
      closed\ \{\src{y}\}\ \src{e_1}
    }{
      \vdash_\copyright\src{context\ e_0\ e_1}
    }{w-context}
  \end{center}
}{w-programs}{$\mmlAs$ programs, contexts, and components.}

\myfig{
  \begin{center}
    \judgbox{\typechecks{}{context\ e_1\ e_2}{\nat}}{,,$\mmlAs$ context $\src{context\ e_1\ e_2}$ typechecks.''}$\;$\\
    \judgbox{\typechecks{}{component\ e_p}{\nat\to\nat}}{,,$\mmlAs$ component $\src{component\ e_p}$ typechecks.''}$\;$\\
    \judgbox{\typechecks{}{program\ e_1\ e_p\ e_2}{\nat}}{,,$\mmlAs$ program $\src{program\ e_1\ e_p\ e_2}$ typechecks.''}$\;$\\
  \typerule{$t-\text{ctx}$}{
    \typechecks{}{e_1}{\nat}&
    \typechecks{y:\nat}{e_2}{\nat}
	}{
  \typechecks{}{context\ e_1\ e_2}{\nat}
	}{w-t-ctx}
  %
  \typerule{$t-\text{component}$}{
    \typechecks{x:\nat}{e}{\nat}&
	}{
  \typechecks{}{component\ e}{\nat\to\nat}
	}{w-t-comp}
  %
  \typerule{$t-\text{wprog}$}{
    \typechecks{}{context\ e_0\ e_1}{\nat} &
    \typechecks{}{component\ e_p}{\nat\to\nat}
	}{
    \typechecks{}{prog\ e_0\ e_p\ e_1}{\nat}
	}{w-t-wprog}
  %
  \judgbox{\typechecks{}{\qconfigs[i]\triangleright e}{\nat}}{,,$\mmlAs$ runtime program $\src{\qconfigs[i]\triangleright e}$ typechecks.''}$\;$\\
  \typerule{$t-\text{prog}$}{
    \typechecks{}{\Delta}{\Gamma}&
    \typechecks{\Gamma}{e}{\nat}
	}{
  \typechecks{}{(H;\Delta)^i\triangleright e}{\nat}
	}{w-t-prog}
  \end{center}
}{w-program-ty}{Checking of $\mmlAs$ contexts, components, programs, and whole programs.}

\myfig{
  \begin{center}
    \judgbox{\typechecks{}{\Delta}{\Gamma}}{,,$\mmlAs$ location map $\Deltas$ yields static typing environment $\Gammas$.''}
    \typerule{$T\text{empty}\Delta$}{
    }{
    \typechecks{}{\hole{\cdot}}{\hole{\cdot}}
    }{t-empty-store}
    %
    \typerule{$T\text{ref}\nat$}{
    \typechecks{}{\Delta}{\Gamma}
    }{
    \typechecks{}{x\mapsto\loc_\poisonless,\Delta}{x:\ptr,\Gamma}
    }{t-ptr-store}
    %
    \typerule{$T\text{ref}\nat\text{poison}$}{
    \typechecks{}{\Delta}{\Gamma}
    }{
    \typechecks{}{x\mapsto\loc_\poisoned,\Delta}{\Gamma}
    }{t-poison-store}
  \end{center}
}{store-ty}{$\mmlAs$ store typing.}

\myfig{
  \begin{center}
    $$
    \typerule{plugging}{
      \typechecks{}{prog\ e_0\ e_p\ e_1}{\nat}
    }{
    \src{context}\ (\src{e_0})\ (\src{e_1})\src{\linker}{\src{component\ e_p}} = \src{prog\ e_0\ e_p\ e_1}
    }{src-plugging}
    $$
  \end{center}
}{w-plugging-s}{$\mmlAs$ plugging of contexts and components.}

\subsubsection{Dynamic Semantics}

\myfig{
  \begin{gather*}
  \begin{aligned}
  \mi{Events}~\src{\event} \bnfdef& \src{\emptyevent} \mid \src{Alloc\ \loc\ n} \mid \src{Dealloc\ \loc} \mid \src{Get\ \loc\ n} \mid \src{Set\ \loc\ n} \\
                             & \src{\lightning} \mid \src{Call\ ?n} \mid \src{Ret\ !n} \\
  \end{aligned}
  \end{gather*}
}{w-events}{Events of $\mmlAs$.}

\myfig{
  \begin{gather*}
    \begin{aligned}
      \src{K}\bnfdef &\ \src{\hole{\cdot}}\mid \src{K\oplus e}\mid \src{v\oplus E}\mid \src{x[K]}\mid
      \src{let\ x=K\ in\ e}\\
                     &\ \mid\src{x[K]\leftarrow e}\mid \src{x[v]\leftarrow K}\mid \src{let\ x=new\ K\ in\ e} \\
                     &\ \mid\src{if\ K<e'\ then\ e_1\ else\ e_2} \mid\src{if\ n<K\ then\ e_1\ else\ e_2}
    \end{aligned}
  \end{gather*}
}{w-expr-eval-ectx}{Evaluation Contexts of $\mmlAs$.}

\myfig{
  \judgbox{\execs[]{\configs}{K[e]}{\configs'}{K[e']}{\event}}{,,Given an evaluation context $\src{K}$ and an expression $\src{e}$,\\it evaluates under configuration $\configs$ to $\src{e'}$ and new\\configuration $\src{\configs'}$ in context $\src{K}$, emitting event $\src{\event}$.''}
  \judgbox{\exprevals{\configs}{e}{\configs'}{e'}{\event}}{,,Expression $\src{e}$ evaluates under configuration $\configs$ to $\src{e'}$ and new\\configuration $\src{\configs'}$, emitting event $\src{\event}$.''}
  %
  \typerule{$e-\text{ctx}$}{
    \configs\not=\src{\lightning} &
    \exprevals{\configs}{e}{\configs'}{e'}{\event}
  }{
    \execs[]{\qconfigs[i]}{K[e]}{\configs'}{K[e']}{\event}
  }{w-e-ectx}
  %
  \typerule{$e-\src{\oplus}$}{
    \src{n_1}+\src{n_2}=\src{n_3}
  }{
    \exprevals{\configs}{n_1+n_2}{\configs}{n_3}{\emptyevent}
  }{w-e-oplus}
  % TODO
  \typerule{$e-\src{get}$}{
    \src{\loc}\in\text{dom }\src{H}\implies\src{v}=\src{H}(\src{\loc+n}) &
    \src{\loc}\not\in\text{dom }\src{H}\implies\src{v}=\src{1729}
  }{
    \exprevals{\sconfig{H}{\Delta_1,x\mapsto\loc_\poison,\Delta_2}}{x[n]}{\sconfig{H}{\Delta_1,x\mapsto\loc_\poison,\Delta_2}}{v}{Get\ \loc\ n}
  }{w-e-get}
  %
  \typerule{$e-\src{set}$}{
    \src{H'} = \src{H}(\src{\loc+n}\mapsto \src{v})
  }{
    \exprevals{\sconfig{H}{\Delta_1,x\mapsto\loc_\poison,\Delta_2}}{x[n]\leftarrow v}{\sconfig{H'}{\Delta_1,x\mapsto\loc_\poison,\Delta_2}}{v}{Set\ \loc\ n\ v}
  }{w-e-set}
  %
  \typerule{$e-\src{let}-\src{x}$}{
  }{
    \exprevals{\sconfig{H}{\Delta}}{let\ x=y\ in\ e}{\sconfig{H}{\Delta}}{e[y/x]}{\emptyevent}
  }{w-e-let-x}
  %
  \typerule{$e-\src{let}-\src{v}$}{
  }{
    \exprevals{\sconfig{H}{\Delta}}{let\ x=v\ in\ e}{\sconfig{H}{\Delta}}{e[v/x]}{\emptyevent}
  }{w-e-let-v}
  %
  \typerule{$e-\src{delete}$}{
  }{
    \exprevals{\sconfig{H}{\Delta_1,x\mapsto\loc_\poison,\Delta_2}}{delete\ x}{\sconfig{H}{\Delta_1,x\mapsto\loc_\poisoned,\Delta_2}}{0}{Dealloc\ \loc}
  }{w-e-delete}
  %
  \typerule{$e-\src{new}$}{
    \text{fresh}\ \src{\loc}&
    \text{fresh}\ \src{z} &
    \src{H'} = \src{H} \ll \src{n} &
  }{
  \exprevals{\sconfig{H}{\Delta}}{let\ x=new\ n\ in\ e}{\sconfig{H'}{z\mapsto \loc_\poisonless,\Delta}}{e[z/x]}{Alloc\ \loc\ n}
  }{w-e-let-new}
  %
  \typerule{$e-\src{if}-$less}{
    n_1<n_2
  }{
  \exprevals{\Omega}{if\ n_1<n_2\ then\ e_1\ else\ e_2}{\Omega}{e_1}{\emptyevent}
  }{w-e-if-less}
  %
  \typerule{$e-\src{if}-$noless}{
    n_1\ge n_2
  }{
  \exprevals{\Omega}{if\ n_1<n_2\ then\ e_1\ else\ e_2}{\Omega}{e_2}{\emptyevent}
  }{w-e-if-noless}
  %
  \typerule{$e-\src{abort}$}{
  }{
    \exprevals{\configs}{abort()}{\lightning}{stuck}{\lightning}
  }{w-e-abort}
}{w-expr-eval}{Evaluation of $\mmlAs$ expressions.}

\myfig{
  \begin{center}
    \judgbox{\execs{\configs}{e}{\configs'}{e'}{\trace}}{,,Expression $\src{e}$ evaluates under configuration $\configs$ to $\src{e'}$ and\\new configuration $\src{\configs'}$, emitting list of events $\src{\trace}$.''}
    %
    \typerule{$es-\text{refl}$}{
    }{
      \execs{\configs}{e}{\configs}{e}{\hole{\cdot}}
    }{ws-e-refl}
    \typerule{$es-\text{trans}-$important}{
      \execs[]{\configs}{e}{\configs'}{e'}{\event}&
      \execs{\configs'}{e'}{\configs''}{e''}{\trace}&
      \src{\event}\not=\src{\emptyevent}
    }{
      \execs{\configs}{e}{\configs''}{e''}{\event \cdot \trace}
    }{ws-e-trans-important}
    \typerule{$es-\text{trans}-$unimportant}{
      \execs[]{\configs}{e}{\configs'}{e'}{\emptyevent}&
      \execs{\configs'}{e'}{\configs''}{e''}{\trace}&
    }{
      \execs{\configs}{e}{\configs''}{e''}{\trace}
    }{ws-e-trans-unimportant}
  \end{center}
}{w-steps}{Trace prefix generation given a $\mmlAs$ program using the reflexive-transitive closure.}

\myfig{
  \begin{center}
    \judgbox{\wexecs{prog\ e_0\ e_p\ e_1}{\configs}{\valueexpr}{\trace}}{,,Run $\mmlAs$ program $\src{prog\ e_0\ e_p\ e_1}$, giving dynamic state $\configs$\\ and emitting trace $\src{\trace}$.''}
  \typerule{$e-\text{wprog}$}{
    \typechecks{}{component\ e_p}{\nat\to\nat} &
    \execs{\hole{\cdot};\hole{\cdot}}{e_0}{\configs_0}{\valueexpr_0}{\trace{_0}} \\
    \execs{\qconfig_0}{e_p[\valueexpr_0/x]}{\configs_p}{\valueexpr_p}{\trace{_p}} &
    \execs{\configs_p}{e_1[\valueexpr_p/y]}{\configs_1}{\valueexpr_1}{\trace{_1}} &
  }{
    \wexecs{prog\ e_0\ e_p\ e_1}{\configs_1}{v_1}{Ret\ !0\cdot\trace{_0}\cdot Call\ ?\valueexpr_0\cdot \trace{_p}\cdot Ret\ !\valueexpr_p\cdot \trace{_1}\cdot Call\ ?\valueexpr_1}
  }{ws-e-prog}
  \typerule{$e-\text{wprog}-$fail0}{
    \typechecks{}{component\ e_p}{\nat\to\nat} &
    \execs{\hole{\cdot};\hole{\cdot}}{e_0}{\lightning}{stuck}{\trace{_0}} \\
  }{
    \wexecs{prog\ e_0\ e_p\ e_1}{\lightning}{stuck}{Ret\ !0\cdot\trace{_0}}
  }{ws-e-prog-fail0}
  \typerule{$e-\text{wprog}-$failp}{
    \typechecks{}{component\ e_p}{\nat\to\nat} &
    \execs{\hole{\cdot};\hole{\cdot}}{e_0}{\configs_0}{\valueexpr_0}{\trace{_0}} \\
    \execs{\qconfig_0}{e_p[\valueexpr_0/x]}{\lightning}{stuck}{\trace{_p}} &
  }{
    \wexecs{prog\ e_0\ e_p\ e_1}{\lightning}{stuck}{Ret\ !0\cdot\trace{_0}\cdot Call\ ?\valueexpr_0\cdot \trace{_p}}
  }{ws-e-prog-failp}
  \typerule{$e-\text{wprog}-$fail1}{
    \typechecks{}{component\ e_p}{\nat\to\nat} &
    \execs{\hole{\cdot};\hole{\cdot}}{e_0}{\configs_0}{\valueexpr_0}{\trace{_0}} \\
    \execs{\qconfig_0}{e_p[\valueexpr_0/x]}{\configs_p}{\valueexpr_p}{\trace{_p}} &
    \execs{\configs_p}{e_1[\valueexpr_p/y]}{\lightning}{stuck}{\trace{_1}} &
  }{
    \wexecs{prog\ e_0\ e_p\ e_1}{\lightning}{stuck}{Ret\ !0\cdot\trace{_0}\cdot Call\ ?\valueexpr_0\cdot \trace{_p}\cdot Ret\ !\valueexpr_p\cdot \trace{_1}}
  }{ws-e-prog-fail1}
  \end{center}
}{wprog-run}{Running a whole $\mmlAs$ program.}

\subsubsection{\gls*{ms}-Specific Events}
\myfig{
  \begin{gather*}
  \begin{aligned}
    \mi{Events}~\src{\mstrace} \bnfdef&\ \specificevs{Alloc\ \loc\ n} \mid \specificevs{Dealloc\ \loc} \mid \specificevs{Use\ \loc\ n} \\
                                    & \mid \specificevs{\lightning} \mid \specificevs{Call\ ?n} \mid \specificevs{Ret\ !n}\\
  \end{aligned}
  \end{gather*}
}{w-specific-events}{\gls*{ms}-Specific Events of $\mmlAs$.}
\myfig{
  \begin{center}
    \judgbox{\tospecificevs{\event} = \src{\msevent}}{,,Project an $\mmlAs$ event $\src{\event}$ to $\src{\msevent}$.''}
    \judgbox{\tospecificevcomps{\trace} = \src{\mstrace}}{,,Project an $\mmlAs$ trace $\src{\trace}$ to $\src{\mstrace}$, ignoring context.''}
    \judgbox{\tospecificevctxs{\trace} = \src{\mstrace}}{,,Project an $\mmlAs$ trace $\src{\trace}$ to $\src{\mstrace}$, ignoring component.''}
    $$
    \begin{array}{rcl}
      \tospecificevs{Alloc\ \loc\ n} &\ = & \specificevs{Alloc\ \loc\ n} \\
      \tospecificevs{Dealloc\ \loc} &\ = & \specificevs{Dealloc\ \loc} \\
      \tospecificevs{Get\ \loc\ n} &\ = & \specificevs{Use\ \loc\ n} \\
      \tospecificevs{Set\ \loc\ n} &\ = & \specificevs{Use\ \loc\ n} \\
      \tospecificevs{Call\ ?n} &\ = & \specificevs{Call\ ?n} \\
      \tospecificevs{Ret\ !n} &\ = & \specificevs{Ret\ !n} \\
      \tospecificevs{\lightning} &\ = & \specificevs{\lightning} \\[0.3cm]
      \tospecificevs{\trace} &\ = & \operatorname{map}\ \thetospecificfilter\ (\operatorname{filter}\ \src{\varepsilon}\ \src{\trace})\\[0.5cm]
      %
      \tospecificevctxs{Ret\ !n_1\cdot\trace{_1}\cdot Call\ ?n_2\cdot\trace{_p}} &\ = & \operatorname{map}\ \thetospecificfilter\ \left(\src{Ret\ !n_1\cdot \trace{_1} \cdot Call\ ?n_2}\right)\\
      \tospecificevcomps{Ret\ !n_0 \cdot \trace{_1}\cdot Call\ ?n_1\cdot\trace{_p}\cdot Ret\ !n_2\cdot\trace{_2}\cdot Call\ n_3} &\ = & \operatorname{map}\ \thetospecificfilter\ \left(\src{Call\ ?n_1\cdot \trace{_p} \cdot Ret\ !n_2}\right)\\
    \end{array}
    $$
  \end{center}
}{w-events-project-to-specific}{Projection of $\mmlAs$ events to \gls*{ms}-specific events.}

\subsubsection{\gls*{tms} Monitor}
\myfig{
  \begin{gather*}
    \begin{aligned}
      \mi{Abstract\ References}~L \subseteq& \nat \hspace{0.5cm}
      \mi{Abstract\ Store}~\tmsmonitor =\{A : L, F : L\} \\
      \mi{Abstract\ Events}~\event \bnfdef& \text{Alloc}\ \loc \mid \text{Dealloc}\ \loc \mid \text{Use}\ \loc \mid \lightning\\
      \mi{References\ Agreement}~\delta_{\text{tms}} :& \src{L}\to L \\[0.6cm]
      \tmsmonitor\subseteq_F \tmsmonitor['] &\text{ iff } \tmsmonitor[.F] \subseteq \tmsmonitor['.F]\\
      \event\not\in \tmsmonitor['] &\text{ iff } \event \not\in {\tmsmonitor['.A]} \wedge \event \not\in {\tmsmonitor['.F]}\\
      \{\loc\}\cup \tmsmonitor &= \{A : \{\loc\}\cup \tmsmonitor[.A], F : \tmsmonitor[.F]\}\\
    \end{aligned}
  \end{gather*}
}{tms-monitor}{\gls{tms} Monitor.}

\myfig{
  \begin{center}
    \judgbox{\monitorcheck[]{\tmsmonitor}{\tmsmonitor[']}{\event}}{,,Given abstract state $\tmsmonitor$, step to abstract state $\tmsmonitor'$\\ emitting an abstract event $\event$.''}
    \judgbox{\monitorcheck{\tmsmonitor}{\tmsmonitor[']}{\trace}}{,,Reflexive-Transitive closure of above judgement.\\$\trace$ is a list of abstract events.''}
    \typerule{TMS-Use}{
      \loc\in \tmsmonitor[.A] &
      \loc\not\in \tmsmonitor[.F] &
    }{
    \monitorcheck[]{\tmsmonitor}{\tmsmonitor}{\text{Use}\ \loc}
    }{tms-use}
    %
    \typerule{TMS-Alloc}{
      \tmsmonitor[']=\left\{A:\tmsmonitor[.A]\cup\left\{\loc\right\}, F:\tmsmonitor[.F]\right\}&
      \loc\not\in \tmsmonitor&
    }{
    \monitorcheck[]{\tmsmonitor}{\tmsmonitor[']}{\text{Alloc}\ \loc}
    }{tms-alloc}
    %
    \typerule{TMS-Dealloc}{
      \loc\in \tmsmonitor[.A] &
      \loc\not\in \tmsmonitor[.F] &
      \tmsmonitor[']=\left\{A:\tmsmonitor[.A]\setminus\left\{\loc\right\},F: \tmsmonitor[.F]\right\}
    }{
    \monitorcheck[]{\tmsmonitor}{\tmsmonitor[']}{\text{Dealloc}\ \loc}
    }{tms-dealloc}
    %
    \typerule{TMS-Refl}{
    }{
    \monitorcheck{\tmsmonitor}{\tmsmonitor}{\hole{\cdot}}
    }{tms-refl}
    %
    \typerule{TMS-Trans}{
    \monitorcheck[]{\tmsmonitor}{\tmsmonitor[']}{\event}&
    \monitorcheck{\tmsmonitor[']}{\tmsmonitor['']}{\trace}&
    }{
    \monitorcheck{\tmsmonitor}{\tmsmonitor['']}{\event\cdot\trace}
    }{tms-trans}
  \end{center}
}{tms-step}{Steps of \gls{tms} Monitor.}

\myfig{
  \begin{center}
    \judgbox{\delta_{\text{tms}}(\src{\loc})=\loc}{,,A map from $\mmlAs$ memory locations $\src{\loc}$ to abstract memory locations $\loc$.''}
    \judgbox{\tmstraceagree{\src{\msevent}}{\event}}{,,Memory-Safety specific event $\src{\msevent}$ describes same action as $\event$.''}
    \judgbox{\tmstraceagree{\src{\mstrace}}{\trace}}{,,Memory-Safety specific trace $\src{\mstrace}$ describes same trace as $\trace$.''}
    %
    \typerule{TMS-Alloc-Authentic}{
      \delta_{\text{tms}}(\src{\loc})=\loc
    }{
      \tmstraceagree{\specificevs{Alloc\ \loc\ n}}{\text{Alloc}\ \loc}
    }{tms-alloc-auth}
    %
    \typerule{TMS-Dealloc-Authentic}{
      \delta_{\text{tms}}(\src{\loc})=\loc
    }{
    \tmstraceagree{\specificevs{Dealloc\ \loc}}{\text{Dealloc}\ \loc}
    }{tms-dealloc-auth}
    %
    \typerule{TMS-Use-Authentic}{
      \delta_{\text{tms}}(\src{\loc})=\loc
    }{
      \tmstraceagree{\specificevs{Use\ \loc\ n}}{\text{Use}\ \loc}
    }{tms-use-auth}
    %
    \typerule{TMS-Refl-Authentic}{
    }{
      \tmstraceagree{\src{\hole{\cdot}}}{\hole{\cdot}}
    }{tms-refl-auth}
    %
    \typerule{TMS-Trans-Authentic}{
      \tmstraceagree{\src{\msevent}}{\event} &
      \tmstraceagree{\src{\mstrace}}{\trace}
    }{
      \tmstraceagree{\src{\msevent}\cdot\src{\mstrace}}{\event\cdot\trace}
    }{tms-trans-auth}
    %
    \typerule{TMS-Call-Authentic}{
      \tmstraceagree{\src{\mstrace}}{\trace}
    }{
      \tmstraceagree{\specificevs{Call\ ?n}\cdot\src{\mstrace}}{\trace}
    }{tms-call-auth}
    %
    \typerule{TMS-Ret-Authentic}{
      \tmstraceagree{\src{\mstrace}}{\trace}
    }{
      \tmstraceagree{\specificevs{Ret\ !n}\cdot\src{\mstrace}}{\trace}
    }{tms-ret-auth}
  \end{center}
}{tms-trace-agreement}{Trace Agreement.}

\myfig{
  \begin{center}
    \judgbox{\storeagreetms{\tmsmonitor}{\src{\Delta^i}}}{,,The abstract memory state $\tmsmonitor$ describes the concrete state $\src{\Delta}$.''}
    \typerule{Empty-Agree}{
    }{
      \storeagreetms{\emptyset}{\src{\hole{\cdot}}}
    }{store-empty-agree}
    %
    \typerule{Cons-Agree}{
      \delta_{\text{tms}}\left(\src{\loc}\right)=\loc&
      \storeagreetms{\tmsmonitor}{\src{\Delta}}
    }{
      \storeagreetms{\left\{\loc\right\}\cup \tmsmonitor}{\src{x\mapsto\loc_\poisonless,\Delta}}
    }{store-cons-agree}
    %
    \typerule{Poison-Agree}{
      \delta_{\text{tms}}\left(\src{\loc}\right)=\loc&
      \storeagreetms{\tmsmonitor}{\src{\Delta}}
    }{
      \storeagreetms{\tmsmonitor}{\src{x\mapsto\loc_\poisoned,\Delta}}
    }{store-poison-agree}
    %
    \typerule{Unimportant-Agree}{
    }{
      \storeagreetms{\emptyset}{\src{\Delta^\diamond}}
    }{store-unimportant-agree}
  \end{center}
}{tms-store-agreement}{Store Agreement.}

We now define temporal and spatial memory safety in terms of traces on events $\src{\event}$.

\begin{definition}[Occurence]
  Given an event $\event$ and a trace $\trace$, the event $\event$ occurs in $\trace$, written $\event\in\trace$, iff

  $\exists n, \trace[n]=\event$
\end{definition}

\begin{definition}[Unicity]
  Given an event $\event$ and a trace $\trace$, the event $\event$ occurs exactly once in $\trace$, written $\event!_{\trace}$, iff

  $\exists n, \trace[n]=\event \wedge \nexists m, n\not= m \wedge \trace[m] = \event$
\end{definition}

\begin{definition}[Sequentiality]
  Given two events $\event{_0},\event{_1}$ and a trace $\trace$, event $\event{_0}$ occurs before $\event{_1}$ in $\trace$, written
  $\event{_0}\le_{\trace}\event{_1}$, iff

  $\exists n, \trace[n] = \event{_0}\implies \exists m, \trace[m] = \event{_1} \wedge n<m$
\end{definition}

\begin{definition}[\gls{tms} on Traces]\label{def:tempmemsafe}
  We define the property $\src{tmsafe}(\src{\mstrace})$ as a conjunction of the following conditions:
  \begin{itemize}
    \item $\specificevs{Alloc\ \loc\ s}\le_{\src{\mstrace}}\specificevs{Dealloc\ \loc}$
    \item $\neg(\specificevs{Use\ \loc\ n}\le_{\src{\mstrace}}\specificevs{Alloc\ \loc\ s})$
    \item $\neg(\specificevs{Dealloc\ \loc}\le_{\src{\mstrace}}\specificevs{Use\ \loc\ n})$
    \item $\specificevs{Dealloc\ \loc}\in!\;\src{\mstrace}$
    \item $\specificevs{Alloc\ \loc\ s}\in!\;\src{\mstrace}$
    % \makebox(0,0){\put(0,3.0\normalbaselineskip){ %
    % $\left.\rule{0pt}{1.32\normalbaselineskip}\right\}$ no reallocs, references are unique}}
  \end{itemize}
\end{definition}
% relaxed temporal memory safety       -> define garbage collector
% generalize this over Γ, τ
% might need an 'invariant'  -> connect typing to evaluation states
% might need to do an induction on evaluation steps

% might not need to do an induction at the general step, but for the primitive step we need it

\begin{definition}[Spatial Memory Safety]\label{def:spatmemsafe}
  We define the property $\src{smsafe}\left(\src{\mstrace}\right)$ as follows:
  \begin{itemize}
    \item If $\specificevs{Alloc\ \loc\ s}\in\src{\mstrace}$ and $\specificevs{Use\ \loc\ n}\in\src{\mstrace}$, then $\src{n}<\src{s}$
  \end{itemize}
\end{definition}

\begin{importanttheorem}[$\mmlAs$ is \gls{tms} via Monitor]
  If
  \begin{assumptions}
  \item $\src{prog\ e_0\ e_p\ e_1}=\src{context}\ (\src{e_0})\ (\src{e_1})\src{\linker}{\src{component\ e_p}}$
  \item $\wexecs{prog\ e_0\ e_p\ e_1}{\configs}{v}{\trace}$
  \end{assumptions}
  Then
  \begin{goals}
  \item $\texttt{TMS}(\tospecificevcomps{\trace})$
  \end{goals}.
\end{importanttheorem}
For the proof of above see \Cref{thm:mmlAs:tmssafe} in \Cref{sec:mmla:aux}.

\begin{importanttheorem}[$\mmlAs$ is not spatially memory safe]
  There is a well-typed $\mmlAs$ component which does not robustly satisfy \Cref{def:spatmemsafe}:
  $\rsat{\src{p}}{\src{smsafe}}$
\end{importanttheorem}
\begin{proof}
  Pick the component $\src{comp\ let\ z=new\ x\ in\ let\ w = z[1337]\ in\ let\ \_ = delete\ z\ in\ w}$.\\[0.3cm]
  It typechecks.\\[0.3cm]
  Now pick $\src{context\ }(\src{42})\ (\src{y})$.\\[0.3cm]
  Plugging yields \\$\src{prog}\ (\src{let\ y=let\ x=42\ in\ let\ z=new\ x\ in\ let\ w = z[1337]\ in\ let\ \_ = delete\ z\ in\ w\ in\ y})$.\\[0.3cm]
  Running the whole program yields trace prefix\\$\src{\trace}=\src{Alloc\ \loc\ 42}\cdot\src{Get\ \loc\ 1337}\cdot\src{Dealloc\ \loc}$.\\[0.3cm]
  \Cref{def:spatmemsafe} requires $\src{Alloc\ \loc\ 42}\in\src{\trace}$ and $\src{Get\ \loc\ 1337}$, which hold.\\[0.3cm]
  But, $\src{1337}\not<\src{42}$, hence $\nrsat{\src{p}}{\src{smsafe}}$.
\end{proof}

\subsubsection{Auxiliary Definitions and Lemmas}\label{sec:mmla:aux}

\begin{lemma}[Plugging Well-Typed]\label{lem:plugging:welltyped}
  If
  \begin{assumptions}
  \item $\typechecks{}{context\ e_0\ e_1}{\nat}$
  \item $\typechecks{}{component\ e_p}{\nat}$
  \end{assumptions}
  Then
  \begin{goals}
  \item $\typechecks{}{prog\ e_0\ e_p\ e_1}{\nat}$
  \end{goals}
\end{lemma}
\begin{proof}
  Simple unfolding of definitions.
\end{proof}

\begin{lemma}[State Typing Plugged]\label{lem:ctxtyping:plugged}
  If
  \begin{assumptions}
    \item $\typechecks{}{\Omega\triangleright K[e]}{\nat}$
  \end{assumptions}
  Then
  \begin{goals}
    \item $\typechecks{}{\Omega\triangleright e}{\nat}$
  \end{goals}
\end{lemma}
\begin{incompleteproof}
\end{incompleteproof}

\begin{lemma}[Weaken Closedness]
  For any two sets $X,Y$ and an expression $\src{e}$, when
  \begin{assumptions}
    \item $closed\ X\ \src{e}$
    \item $X\subseteq Y$
  \end{assumptions}
  Then
  \begin{goals}
    \item $closed\ Y\ \src{e}$
  \end{goals}
\end{lemma}
\begin{proof}
  Structural induction on $\src{e}$.
\end{proof}
\begin{lemma}[Authentic Weaken]
  If
  \begin{assumptions}
    \item $\tmstraceagree[\delta]{\src{\mstrace}}{\trace}$
    \item $\delta_{\text{tms}}\subseteq\delta_{\text{tms}}'$
  \end{assumptions}
  Then
  \begin{goals}
    \item $\tmstraceagree[\delta']{\src{\mstrace}}{\trace}$
  \end{goals}
\end{lemma}
\begin{proof}
  Induction on $\tmstraceagree[\delta]{\src{\trace}}{\trace}$.
\end{proof}

\begin{lemma}[Monitor Step Subset]\label{lem:monitor-step-subset}
  If
  \begin{assumptions}
    \item $\monitorcheck[]{\tmsmonitor}{\tmsmonitor[']}{\event}$
  \end{assumptions}
  Then
  \begin{goals}
    \item $\tmsmonitor\subseteq_F \tmsmonitor[']$
  \end{goals}
\end{lemma}
\begin{incompleteproof}
\end{incompleteproof}

\begin{lemma}[Monitor Steps Subset]\label{lem:monitor-steps-subset}
  If
  \begin{assumptions}
    \item $\monitorcheck{\tmsmonitor}{\tmsmonitor[']}{\event}$
  \end{assumptions}
  Then
  \begin{goals}
    \item $\tmsmonitor\subseteq_F \tmsmonitor[']$
  \end{goals}
\end{lemma}
\begin{proof}
  Easy induction on $\monitorcheck{\tmsmonitor}{\tmsmonitor[']}{\event}$ using \Thmref{lem:monitor-step-subset}.
\end{proof}

\begin{lemma}[Monitor Step Back No-Elem Propagation]\label{lem:monitor-step-no-elem-prop}
  If
  \begin{assumptions}
    \item $\monitorcheck[]{\tmsmonitor}{\tmsmonitor[']}{\event}$
    \item $\loc\not\in \tmsmonitor['].A$
    \item $\loc\not\in \tmsmonitor['].F$
  \end{assumptions}
  Then
  \begin{goals}
    \item $\loc\not\in \tmsmonitor[.A]$
  \end{goals}
\end{lemma}
\begin{incompleteproof}
\end{incompleteproof}

\begin{lemma}[Monitor Steps Back No-Elem Propagation]\label{lem:monitor-steps-no-elem-prop}
  If
  \begin{assumptions}
    \item $\monitorcheck{\tmsmonitor}{\tmsmonitor[']}{\event}$
    \item $\loc\not\in \tmsmonitor['.A]$
    \item $\loc\not\in \tmsmonitor['.F]$
  \end{assumptions}
  Then
  \begin{goals}
    \item $\loc\not\in \tmsmonitor[.A]$
  \end{goals}
\end{lemma}
\begin{incompleteproof}
\end{incompleteproof}

\begin{lemma}[$\alpha$-conv. Typing]\label{lem:alpha-conv-typing}
  If
  \begin{assumptions}
    \item $\typechecks{\Gamma}{e}{\type}$
  \end{assumptions}
  Then
  \begin{goals}
    \item $\typechecks{\Gamma[z:\type'/x:\type]}{e[z/x]}{\type}$
  \end{goals}.
\end{lemma}
\begin{incompleteproof}
\end{incompleteproof}

\begin{lemma}[Substitution]\label{lem:substitution}
  If
  \begin{assumptions}
    \item $\typechecks{x:\type',\Gamma_1}{e}{\type}$
    \item $\typechecks{\Gamma_2}{v}{\type'}$
    \item $\src{\Gamma_3}=\src{\Gamma_1}\circ\src{\Gamma_2}$
  \end{assumptions}
  Then
  \begin{goals}
    \item $\typechecks{\Gamma_3}{e[v/x]}{\type}$
  \end{goals}.
\end{lemma}
\begin{incompleteproof}
\end{incompleteproof}

\begin{lemma}[Base Preservation]\label{lem:base-preservation}
  If
  \begin{assumptions}
    \item $\typechecks{}{\src{\Omega}\triangleright e}{\src{\nat}}$
    \item $\exprevals{\configs}{e}{\configs'}{e'}{\event}$
    \item $\configs'\not=\src{\lightning}$
  \end{assumptions}
  Then
  \begin{goals}
    \item $\typechecks{}{\src{\Omega'}\triangleright e'}{\src{\nat}}$
  \end{goals}.
\end{lemma}
\begin{incompleteproof}
\end{incompleteproof}

\begin{lemma}[Ctx Preservation]\label{lem:ctx-preservation}
  If
  \begin{assumptions}
    \item $\typechecks{}{\src{\Omega}\triangleright e}{\src{\nat}}$
    \item $\execs[]{\configs}{e}{\configs'}{e'}{\event}$
    \item $\configs'\not=\src{\lightning}$
  \end{assumptions}
  Then
  \begin{goals}
    \item $\typechecks{}{\src{\Omega'}\triangleright e'}{\src{\nat}}$
  \end{goals}.
\end{lemma}
\begin{incompleteproof}
\end{incompleteproof}

\begin{lemma}[Steps Preservation]\label{lem:steps-preservation}
  If
  \begin{assumptions}
    \item $\typechecks{}{\src{\Omega}\triangleright e}{\src{\nat}}$
    \item $\execs{\configs}{e}{\configs'}{e'}{\trace}$
    \item $\configs'\not=\src{\lightning}$
  \end{assumptions}
  Then
  \begin{goals}
    \item $\typechecks{}{\src{\Omega'}\triangleright e'}{\src{\nat}}$
  \end{goals}.
\end{lemma}
\begin{incompleteproof}
\end{incompleteproof}

\begin{lemma}[Progress]\label{lem:progress}
  If
  \begin{assumptions}
    \item $\typechecks{}{\src{\Omega}\triangleright e}{\src{\nat}}$
  \end{assumptions}
  Then
  \begin{goals}
  \item $\src{e}=\src{\valueexpr}$ or $\exists \src{\Omega'}\ \src{\valueexpr}\ \src{\trace}, \execs{\configs}{e}{\configs'}{\valueexpr}{\trace}$
  \end{goals}.
\end{lemma}
\begin{incompleteproof}
\end{incompleteproof}

\begin{lemma}[Top-Level Progress]\label{lem:toplevel:progress}
  If
  \begin{assumptions}
    \item $\typechecks{}{prog\ e_1\ e_p\ e_2}{\nat}$
  \end{assumptions}
  Then
  \begin{goals}
  \item $\exists \configs\ \src{v}\ \src{\trace}, \wexecs{prog\ e_1\ e_p\ e_2}{\Omega}{v}{\trace}$
  \end{goals}.
\end{lemma}
\begin{incompleteproof}
\end{incompleteproof}

\begin{lemma}[Mid-Level Progress]\label{lem:midlevel:progress}
  If
  \begin{assumptions}
    \item $\typechecks{}{component\ e_p}{\nat}$
    \item $\vdash_\copyright\src{context\ e_1\ e_2}$
  \end{assumptions}
  Then
  \begin{goals}
  \item $\exists \configs\ \src{\finalexpr}\ \src{\trace}, \wexecs{prog\ e_1\ e_p\ e_2}{\Omega}{\finalexpr}{\trace}$
  \end{goals}.
\end{lemma}
\begin{incompleteproof}
\end{incompleteproof}

\begin{lemma}[Base \gls{tms} via Monitor]
  If
  \begin{assumptions}
    \item $\typechecks{}{\src{\Omega}\triangleright e}{\src{\nat}}$
    \item $\exprevals{\configs}{e}{\configs'}{e'}{\event}$
    \item $\storeagreetms{\tmsmonitor}{\src{\Omega}.\src{\Delta}}$
  \end{assumptions}
  Then $\exists \event\ \delta_{\text{tms}}'\ \tmsmonitor['],$
  \begin{goals}
    \item $\delta_{\text{tms}}\subseteq\delta_{\text{tms}}'$
    \item $\tmstraceagree[\delta']{\tospecificevs{\event}}{\event}$
    \item $\monitorcheck[]{\tmsmonitor}{\tmsmonitor[']}{\event}$
    \item $\storeagreetms[\delta']{\tmsmonitor[']}{\src{\Omega'}.\src{\Delta}}$
  \end{goals}.
\end{lemma}
\begin{incompleteproof}
  Induction on $\exprevals{\configs}{e}{\configs'}{e'}{\event}$.
  \begin{description}
\item[$\src{e}=\src{free\ x}$:]
  If
  \begin{assumptions}
    \item $\typechecks{}{\src{H;\Delta_1,x\mapsto\loc_\poison,\Delta_2}\triangleright \src{free\ x}}{\nat}$
    \item\label[ass]{asm:basetmsmonitor:storeagree} $\storeagreetms{\tmsmonitor}{\src{\Delta_1,x\mapsto\loc_\poison,\Delta_2}}$
  \end{assumptions}
  Then $\exists \event\ \delta_{\text{tms}}'\ \tmsmonitor['],$
  \begin{goals}
    \item\label[goal]{goal:basetmsmonitor:subseteq} $\delta_{\text{tms}}\subseteq\delta_{\text{tms}}'$
    \item\label[goal]{goal:basetmsmonitor:tmstraceagree} $\traceagree[\delta_{\text{tms}}']{\specificevs{Dealloc\ \loc}}{\event}$
    \item $\monitorcheck[]{\tmsmonitor}{\tmsmonitor[']}{\event}$
    \item $\storeagreetms[\delta']{\tmsmonitor[']}{\src{\Delta_1,x\mapsto\loc_\poisoned,\Delta_2}}$
  \end{goals}.

  Choose $\event=\text{Dealloc\ }\loc$ and $\delta_{\text{tms}}'=\delta_{\text{tms}}$.

  \Cref{goal:basetmsmonitor:subseteq}, $\delta_{\text{tms}}\subseteq\delta_{\text{tms}}$, follows trivially.

  For $\tmstraceagree{\specificevs{Dealloc\ \loc}}{\text{Dealloc}\ \loc}$ (\Cref{goal:basetmsmonitor:tmstraceagree}), apply \Cref{tr:tms-dealloc-auth}, where $\delta_{\text{tms}}(\src{\loc})=\loc$ follows by inverting \Cref{asm:basetmsmonitor:storeagree}.

  The inversion yields two cases:
  \begin{description}
\item[\Cref{tr:store-cons-agree}:] \ \\
  If
  \begin{assumptions}
  \item $\typechecks{}{\src{H;\Delta_1},\src{x\mapsto\loc_\poisonless},\src{\Delta_2}\triangleright \src{free\ x}}{\nat}$
  \item $\storeagreetms{\tmsmonitor[_1]}{\src{\Delta_1}}$
  \item $\delta_{\text{tms}}(\src{\loc})=\loc$
  \item $\storeagreetms{\tmsmonitor[_2]}{\src{\Delta_2}}$
  \item $\tmsmonitor=\tmsmonitor[_1]\cup\left\{\loc\right\}\cup \tmsmonitor[_2]$
  \end{assumptions}
  then $\exists \tmsmonitor[']$,
  \begin{goals}
  \item\label[goal]{goal:basetmsmonitor:monitorcheck1} $\monitorcheck[]{\tmsmonitor}{\tmsmonitor[']}{\text{Dealloc}\ \loc}$
  \item\label[goal]{goal:basetmsmonitor:storeagree1} $\storeagreetms[\delta']{\tmsmonitor[']}{\src{\Delta_1,x\mapsto\loc_\poisoned,\Delta_2}}$
  \end{goals}.

  Choose $\tmsmonitor[']=\tmsmonitor[_1]\cup \tmsmonitor[_2]$.

  \Cref{goal:basetmsmonitor:monitorcheck1}, $\monitorcheck[]{\tmsmonitor}{\tmsmonitor[']}{\text{Dealloc}\ \loc}$, follows immediately by \Cref{tr:tms-dealloc}.

  \Cref{goal:basetmsmonitor:storeagree1}, $\storeagreetms[\delta']{\tmsmonitor[']}{\src{\Delta_1,x\mapsto\loc_\poisoned,\Delta_2}}$, follows immediately by \Cref{tr:store-poison-agree}.

\item[\Cref{tr:store-poison-agree}:] \ \\
  If
  \begin{assumptions}
  \item\label[ass]{asm:basetmsmonitor:typing} $\typechecks{}{\src{H;\Delta_1},\src{x\mapsto\loc_\poisoned},\src{\Delta_2}\triangleright \src{free\ x}}{\nat}$
  \item $\storeagreetms{\tmsmonitor[_1]}{\src{\Delta_1}}$
  \item $\delta_{\text{tms}}(\src{\loc})=\loc$
  \item $\storeagreetms{\tmsmonitor[_2]}{\src{\Delta_2}}$
  \item $\tmsmonitor=\tmsmonitor[_1]\cup \tmsmonitor[_2]$
  \end{assumptions}
  then $\exists \tmsmonitor[']$,
  \begin{goals}
  \item\label[goal]{goal:basetmsmonitor:monitorcheck2} $\monitorcheck[]{\tmsmonitor}{\tmsmonitor[']}{\text{Dealloc}\ \loc}$
  \item\label[goal]{goal:basetmsmonitor:storeagree2} $\storeagreetms[\delta']{\tmsmonitor[']}{\src{\Delta_1,x\mapsto\loc_\poisoned,\Delta_2}}$
  \end{goals}.

  Invert \Cref{asm:basetmsmonitor:typing} to get
  \begin{assumptions}
  \item\label[ass]{asm:basetmsmonitor:storechecks} $\typechecks{}{\src{H;\Delta_1,x\mapsto\loc_\poisoned,\Delta_2}}{\Gammas}$
  \item\label[ass]{asm:basetmsmonitor:freechecks} $\typechecks{\Gammas}{free\ x}{\nat}$
  \end{assumptions}.

Due to \Cref{asm:basetmsmonitor:storechecks}, we know $\src{x:\ptr}\not\in\Gammas$.

But, that contradicts \Cref{asm:basetmsmonitor:freechecks}.
\end{description}

\end{description}
\end{incompleteproof}

\begin{lemma}[Ctx \gls{tms} via Monitor]
  If
  \begin{assumptions}
    \item $\typechecks{}{\src{\Omega}\triangleright e}{\src{\nat}}$
    \item $\execs[]{\configs}{e}{\configs'}{e'}{\event}$
    \item $\storeagreetms{\tmsmonitor}{\src{\Omega}.\src{\Delta}}$
  \end{assumptions}
  Then $\exists \event\ \delta_{\text{tms}}'\ \tmsmonitor['],$
  \begin{goals}
    \item $\delta_{\text{tms}}\subseteq\delta_{\text{tms}}'$
    \item $\tmstraceagree[\delta']{\tospecificevs{\event}}{\event}$
    \item $\monitorcheck[]{\tmsmonitor}{\tmsmonitor[']}{\event}$
    \item $\storeagreetms[\delta']{\tmsmonitor[']}{\src{\Omega'}.\src{\Delta}}$
  \end{goals}.
\end{lemma}
\begin{incompleteproof}
\end{incompleteproof}

\begin{lemma}[Steps \gls{tms} via Monitor]\label{lem:generaltms:monitor}
  If
  \begin{assumptions}
    \item $\typechecks{}{\src{\Omega}\triangleright e}{\src{\nat}}$
    \item $\execs{\configs}{e}{\configs'}{e'}{\trace}$
    \item $\storeagreetms{\tmsmonitor}{\src{\Omega}.\src{\Delta}}$
  \end{assumptions}
  Then $\exists \trace\ \delta_{\text{tms}}'\ \tmsmonitor['],$
  \begin{goals}
    \item $\delta_{\text{tms}}\subseteq\delta_{\text{tms}}'$
    \item $\tmstraceagree[\delta']{\tospecificevs{\trace}}{\trace}$
    \item $\monitorcheck{\tmsmonitor}{\tmsmonitor[']}{\trace}$
    \item $\storeagreetms[\delta']{\tmsmonitor[']}{\src{\Omega'}.\src{\Delta}}$
  \end{goals}.
\end{lemma}
\begin{incompleteproof}
\end{incompleteproof}

\begin{definition}[Trace is temporal memory safe via monitor.]\label{def:src:trace:tms}
  We say $\texttt{TMS}(\src{\trace})$ iff $\exists \delta\ \trace\ \tmsmonitor$ such that:
  \begin{itemize}
    \item $\tmstraceagree{\trace}{\src{\trace}}$
    \item and $\monitorcheck{\emptyset}{\tmsmonitor}{\trace}$
  \end{itemize}
\end{definition}

\begin{theorem}[$\mmlAs$ is \gls{tms} via Monitor]\label{thm:mmlAs:tmssafe}
  If
  \begin{assumptions}
  \item $\src{prog\ e_0\ e_p\ e_1}=\src{context}\ (\src{e_0})\ (\src{e_1})\src{\linker}{\src{component\ e_p}}$
  \item\label{asm:mmlatmsviamonitor:run} $\wexecs{prog\ e_0\ e_p\ e_1}{\configs}{\valueexpr}{\trace}$
  \end{assumptions}
  Then
  \begin{goals}
  \item $\texttt{TMS}(\tospecificevcomps{\trace})$
  \end{goals}.
\end{theorem}
\begin{proof}
  Inverting \Cref{asm:mmlatmsviamonitor:run}, we get 
  \begin{passumptions}{H}
    \item $\typechecks{}{prog\ e_0\ e_p\ e_1}{\nat}$
    \item\label[ass]{asm:mmlatmsviamonitor:exec0} $\execs{\hole{\cdot};\hole{\cdot}}{e_0}{\configs_0}{v_0}{\trace{_0}}$
    \item\label[ass]{asm:mmlatmsviamonitor:execp} $\execs{\qconfig{_0}}{e_p[v_0/x]}{\configs_p}{v_p}{\trace{_p}}$
    \item\label[ass]{asm:mmlatmsviamonitor:exec1} $\execs{\configs_p}{e_1[v_p/y]}{\configs_1}{v_1}{\trace{_1}}$
    \item\label[ass]{asm:mmlatmsviamonitor:tracedef} $\src{\trace}=\src{Ret\ !0}\cdot\src{\trace{_0}}\cdot\src{Call\ ?v_0}\cdot\src{\trace{_p}}\cdot\src{Ret\ !v_p}\cdot\src{\trace{_1}\cdot\src{Call\ ?v_1}}$
  \end{passumptions}
  Note that the other cases don't apply, since it is assumed here (\Cref{asm:mmlatmsviamonitor:run}) that the program terminates in a value instead of $\src{stuck}$.
  By \Cref{tr:store-empty-agree}:
  \begin{equation}
    \storeagreetms[\emptyset]{\emptyset}{\src{\emptyset}}\label{asm:mmlatmsviamonitor:agreeempty}
  \end{equation}

  Invert $\typechecks{}{prog\ e_0\ e_p\ e_1}{\nat}$ and invert the subsequent result to get:
  \begin{passumptions}{H^t}
    \item\label[ass]{asm:mmlatmsviamonitor:type0} $\typechecks{\hole{\cdot}}{e_0}{\nat}$
    \item\label[ass]{asm:mmlatmsviamonitor:typep} $\typechecks{x:\nat}{e_p}{\nat}$
    \item\label[ass]{asm:mmlatmsviamonitor:type1} $\typechecks{y:\nat}{e_1}{\nat}$
  \end{passumptions}

  By \Cref{tr:t-empty-store} and \Cref{asm:mmlatmsviamonitor:type1}, we have 
  \begin{equation}
    \typechecks{}{\hole{\cdot}\triangleright e_0}{\nat}\label{asm:mmlatmsviamonitor:storetypenil}
  \end{equation}

  Plugging \Cref{asm:mmlatmsviamonitor:type0} and \Cref{asm:mmlatmsviamonitor:exec0} into \Thmref{lem:steps-preservation} and inverting the result, we obtain
  \begin{passumptions}{G}
    \item $\typechecks{}{\Omega_0.\Delta}{\Gamma_0}$
    \item\label[ass]{asm:mmlatmsviamonitor:vtype} $\typechecks{\Gamma_0}{v_0}{\nat}$
  \end{passumptions}

  Note that $\src{\Gamma_0} = \src{\hole{\cdot}} \circ \src{\Gamma_0}$ which we'll use promptly.\MK{technical lemma needed}

  Using \Cref{asm:mmlatmsviamonitor:type1}, \Cref{asm:mmlatmsviamonitor:vtype}, and the above fact, we apply \Thmref{lem:substitution} to obtain:
  \begin{equation}
    \typechecks{\Gamma_0}{e_p[v_0/x]}{\nat}\label{asm:mmlatmsviamonitor:typesubstx}
  \end{equation}

  \noindent
  By \Cref{tr:store-unimportant-agree}:
  \begin{equation}
    \storeagreetms[\emptyset]{\emptyset}{\src{\Omega_0.\Delta}}\label{asm:mmlatmsviamonitor:agree}
  \end{equation}
  What follows is $\typechecks{}{\Omega_0\triangleright e_p}{\nat}$, which is equivalent to $\typechecks{}{\Omega_0^\diamond\triangleright e_p}{\nat}$.

  Now obtain witnesses $\trace{_p}$, $\delta_{\text{TMS}}^{(p)}$, and $\tmsmonitor$ from supplying \Thmref{lem:generaltms:monitor} with \Cref{asm:mmlatmsviamonitor:typesubstx}, \Cref{asm:mmlatmsviamonitor:execp}, and \Cref{asm:mmlatmsviamonitor:agree}, such that:

  \begin{passumptions}{F}
    \item $\emptyset\subseteq\delta_{\text{TMS}}^{(p)}$
    \item\label[ass]{asm:mmlatmsviamonitor:traceagreeasm} $\tmstraceagree[\delta^{(p)}]{\trace{_p}}{\tospecificevs{\trace{_p}}}$
    \item\label[ass]{asm:mmlatmsviamonitor:monitorcheckasm} $\monitorcheck{\emptyset}{\tmsmonitor}{\trace{_p}}$
    \item $\storeagreetms[\delta^{(p)}]{\tmsmonitor}{\src{\Omega_p}.\src{\Delta}}$
  \end{passumptions}

  \noindent
  Now instantiate the existential in our goal with $\delta_{\text{TMS}}^{(p)}$, $\trace{_p}$, and $\tmsmonitor$.
  Note that $\tospecificevcomps{\trace}=\specificevs{Call\ ?v_0}\cdot\tospecificevs{\trace{_p}}\cdot\specificevs{Ret\ !v_p}$ by \Cref{asm:mmlatmsviamonitor:tracedef}, so the remaining goals are:

  \begin{goals}
    \item\label[goal]{asm:mmlatmsviamonitor:traceagree} $\tmstraceagree[\delta^{(p)}]{\trace{_p}}{\specificevs{Call\ ?v_0}\cdot\tospecificevs{\trace{_p}}\cdot\specificevs{Ret\ !v_p}}$
    \item\label[goal]{asm:mmlatmsviamonitor:monitorcheck} and $\monitorcheck{\emptyset}{\tmsmonitor}{\trace{_p}}$
  \end{goals}

  For \Cref{asm:mmlatmsviamonitor:traceagree}, note that by \Cref{tr:tms-call-auth} and \Cref{tr:tms-ret-auth}, we can reduce it to $\tmstraceagree[\delta^{(p)}]{\trace{_p}}{\tospecificevs{\trace{_p}}}$, which is fulfilled by \Cref{asm:mmlatmsviamonitor:traceagreeasm}.
  \Cref{asm:mmlatmsviamonitor:monitorcheck} follows immediately by \Cref{asm:mmlatmsviamonitor:monitorcheckasm}.
\end{proof}

\clearpage

\subsection{Target Language}
\subsubsection{Syntax}

%%%%
%% TARGET

\myfig{
	\vspace{-1em}
  \begin{gather*}
  \begin{aligned}
  \mi{Expressions}~\trg{e} \bnfdef&\ \trg{x} \mid \trg{v} \mid \trg{e_{1} \oplus e_{2}} \mid \trg{x[e]} \mid \trg{let\ x = e_{1}\ in\ e_{2}}\mid \trg{x[e_{1}]\leftarrow e_{2}} \\
    & \mid \trg{let\ x=new\ e_{1}\ in\ e_{2}} \mid \trg{delete\ x} \\
    & \mid \trg{if\ e_1 < e_2\ then\ e_3\ else\ e_4} \mid \trg{abort()} \\
    & \mid \trg{\langle e_1;e_2\rangle} \mid \trg{\pi_1\ e} \mid \trg{\pi_2\ e}
  \hspace{0.5cm} \text{where } \trg{\oplus}\in\{\trg{+},\trg{-},\trg{\times}\}\\
  %
  \mi{Values}~\trg{v} \bnfdef&\ \trg{n}\in\trg{\nat} \hspace{0.5cm}
  \mi{References}~\trg{\loc}\in\trg{\nat} \\
  %
  \mi{Variables}~\ \trg{x} &\mid \trg{y} \mid \dots \hspace{0.5cm}
  %
  \mi{Poison}~\ \trg{\poison}~\bnfdef\ \trg{\poisonless} \mid \trg{\poisoned}\\
  %
  \mi{Store}~\trg{\Delta} \bnfdef&\ \trg{\hole{\cdot}} \mid \trg{x\mapsto\loc_\poison},\Deltat\hspace{0.5cm}
  \mi{Final\ Result}\ \trg{\finalexpr} \bnfdef\ \trg{\valueexpr} \mid \trg{stuck} \\
  %
  \mi{Heaps}~\trg{H} \bnfdef&\ \trg{\hole{\cdot}} \mid \trg{H}::\trg{n} \hspace{0.5cm}
  %
  \mi{State}\ \configt \bnfdef\ \tconfig{H}{\Delta} \mid \trg{\lightning}\\
  %
  \mi{Contexts}\ &\trg{context\ e_0\ e_1} \hspace{0.5cm}
  \mi{Components}\ \trg{component\ e_p} \\
  \mi{Whole\ Programs}\ &\trg{prog\ e_0\ e_1\ e_2}\hspace{0.5cm}
  \mi{Marked\ State}\ \qconfigt[i] \bnfdef\ \configt\mid \qconfigt\\
  \end{aligned}
  \end{gather*}
}{mmlat-syntax}{Syntax of $\mmlAt$}

\subsubsection{Dynamic Semantics}

\myfig{
  \begin{gather*}
    \begin{aligned}
      \trg{K}\bnfdef &\ \trg{\hole{\cdot}}\mid \trg{K\oplus e}\mid \trg{v\oplus E}\mid \trg{x[K]}\mid
      \trg{let\ x=K\ in\ e}\\
                     &\mid \trg{x[K]\leftarrow e}\mid \trg{x[v]\leftarrow K}\mid \trg{let\ x=new\ K\ in\ e} \mid \trg{\langle K; e\rangle} \mid \trg{\langle n; K\rangle}\\
                     &\mid \trg{if\ K < e_1\ then\ e_2\ else\ e_3} \mid \trg{if\ n < K\ then\ e_1\ else\ e_2}
    \end{aligned}
  \end{gather*}
}{tw-expr-eval-ectx}{Evaluation Contexts of $\mmlAt$.}

\myfig{
  \begin{center}
    $$
    \typerule{plugging}{
    }{
    \trg{context}\ (\trg{e_0})\ (\trg{e_1})\trg{\linker}{\trg{component\ e_p}} = \trg{prog\ (let\ y=let\ x=e_0\ in\ e_p\ in\ e_1)}
    }{trg-plugging}
    $$
  \end{center}
}{w-plugging-t}{$\mmlAt$ plugging of contexts and components.}

\myfig{
  \begin{gather*}
  \begin{aligned}
  \mi{Events}~\trg{\event} \bnfdef& \trg{\emptyevent} \mid \trg{Alloc\ \loc\ n} \mid \trg{Dealloc\ \loc} \mid \trg{Get\ \loc\ n} \mid \trg{Set\ \loc\ n}\\
                             & \mid \trg{\lightning} \mid \trg{Call\ ?n} \mid \trg{Ret\ !n}
  \end{aligned}
  \end{gather*}
}{tw-events}{Events of $\mmlAt$.}

\myfig{
  \judgbox{\exect[]{\configt}{K[e]}{\configt'}{K[e']}{\event}}{,,Given an evaluation context $\trg{K}$ and an expression $\trg{e}$,\\it evaluates under configuration $\configt$ to $\trg{e'}$ and\\new configuration $\trg{\configt'}$ in context $\trg{K}$, emitting event $\trg{\event}$.''}
  \judgbox{\exprevalt{\configt}{e}{\configt'}{e'}{\event}}{,,Expression $\trg{e}$ evaluates under configuration $\configt$ to $\trg{e'}$ and\\new configuration $\trg{\configt'}$, emitting event $\trg{\event}$.''}
  %
  \typerule{$e-\text{ctx}$}{
    \configt\not=\trg{\lightning} &
    \exprevalt{\configt}{e}{\configt'}{e'}{\event}
  }{
    \exect[]{\configt}{K[e]}{\configt'}{K[e']}{\event}
  }{tw-e-ectx}
  %
  \typerule{$e-\trg{\pi_1}$}{
  }{
    \exprevalt{\configt}{\pi_1\langle n_1;n_2\rangle}{\configt}{n_1}{\emptyevent}
  }{tw-e-proj1}
  %
  \typerule{$e-\trg{\pi_2}$}{
  }{
    \exprevalt{\configt}{\pi_2\langle n_1;n_2\rangle}{\configt}{n_2}{\emptyevent}
  }{tw-e-proj2}
  %
  \typerule{$e-\trg{if}-$less}{
    \trg{n_1}<\trg{n_2}
  }{
    \exprevalt{\configt}{if\ n_1<n_2\ then\ e_1\ else\ e_2}{\configt}{e_1}{\emptyevent}
  }{tw-e-if-less}
  %
  \typerule{$e-\trg{if}-$noless}{
    \trg{n_1}\ge\trg{n_2}
  }{
    \exprevalt{\configt}{if\ n_1<n_2\ then\ e_1\ else\ e_2}{\configt}{e_2}{\emptyevent}
  }{tw-e-if-noless}
  %
  \typerule{$e-\trg{abort}$}{
  }{
    \exprevalt{\configt}{abort()}{\lightning}{stuck}{\lightning}
  }{tw-e-abort}
  %
  \typerule{$e-\trg{\oplus}$}{
    \trg{n_1}+\trg{n_2}=\trg{n_3}
  }{
    \exprevalt{\configt}{n_1+n_2}{\configt}{n_3}{\emptyevent}
  }{tw-e-oplus}
  % TODO
  \typerule{$e-\trg{get}$}{
    \trg{\loc}\in\text{dom }\trg{H}\implies\trg{v}=\trg{H}(\trg{\loc+n}) &
    \trg{\loc}\not\in\text{dom }\trg{H}\implies\trg{v}=\trg{1729}
  }{
    \exprevalt{\tconfig{H}{\Delta_1,x\mapsto\loc_\poison,\Delta_2}}{x[n]}{\tconfig{H}{\Delta_1,x\mapsto\loc_\poison,\Delta_2}}{v}{Get\ \loc\ n}
  }{tw-e-get}
  %
  \typerule{$e-\trg{set}$}{
    \trg{H'} = \trg{H}(\trg{\loc+n}\mapsto \trg{v})
  }{
    \exprevalt{\tconfig{H}{\Delta_1,x\mapsto\loc_\poison,\Delta_2}}{x[n]\leftarrow v}{\tconfig{H'}{\Delta_1,x\mapsto\loc_\poison,\Delta_2}}{v}{Set\ \loc\ n\ v}
  }{tw-e-set}
  %
  \typerule{$e-\trg{let}-\trg{x}$}{
  }{
    \exprevalt{\tconfig{H}{\Delta}}{let\ x=y\ in\ e}{\tconfig{H}{\Delta}}{e[y/x]}{\emptyevent}
  }{tw-e-let-x}
  %
  \typerule{$e-\trg{let}-\trg{v}$}{
  }{
    \exprevalt{\tconfig{H}{\Delta}}{let\ x=v\ in\ e}{\tconfig{H}{\Delta}}{e[v/x]}{\emptyevent}
  }{tw-e-let-v}
  %
  \typerule{$e-\trg{delete}$}{
  }{
    \exprevalt{\tconfig{H}{\Delta_1,x\mapsto\loc_\poison,\Delta_2}}{delete\ x}{\tconfig{H}{\Delta_1,x\mapsto\loc_\poisoned,\Delta_2}}{0}{Dealloc\ \loc}
  }{tw-e-delete}
  %
  \typerule{$e-\trg{new}$}{
    \text{fresh}\ \trg{\loc}&
    \text{fresh}\ \trg{z} &
    \trg{H'} = \trg{H} \ll \trg{n} &
  }{
  \exprevalt{\tconfig{H}{\Delta}}{let\ x=new\ n\ in\ e}{\tconfig{H'}{z\mapsto \loc_\poisonless,\Delta}}{e[z/x]}{Alloc\ \loc\ n}
  }{tw-e-let-new}
  %
}{tw-expr-eval}{Evaluation of $\mmlAt$ expressions.}

\myfig{
  \begin{center}
    \judgbox{\exect{\configt}{e}{\configt'}{e'}{\trace}}{,,Expression $\trg{e}$ evaluates under configuration $\configt$ to $\trg{e'}$ and\\new configuration $\trg{\configt'}$, emitting list of events $\trg{\trace}$.''}
    %
    \typerule{$et-\text{refl}$}{
    }{
      \exect{\configt}{e}{\configt}{e}{\hole{\cdot}}
    }{wt-e-refl}
    \typerule{$et-\text{trans}-$important}{
      \exect[]{\configt}{e}{\configt'}{e'}{\event}&
      \exect{\configt'}{e'}{\configt''}{e''}{\trace}&
      \trg{a}\not=\trg{\emptyevent}
    }{
      \exect{\configt}{e}{\configt''}{e''}{\event \cdot \trace}
    }{wt-e-trans-important}
    \typerule{$et-\text{trans}-$unimportant}{
      \exect[]{\configt}{e}{\configt'}{e'}{\emptyevent}&
      \exect{\configt'}{e'}{\configt''}{e''}{\trace}&
    }{
      \exect{\configt}{e}{\configt''}{e''}{\trace}
    }{wt-e-trans-unimportant}
  \end{center}
}{wt-steps}{Trace prefix generation given a $\mmlAt$ program using the reflexive-transitive closure.}

\myfig{
  \begin{center}
    $$
    \trg{\Omega^\diamond} = \trg{\Omega.H;}\text{fold}\ (\lambda \trg{\Delta}\ (\trg{x\mapsto\loc_\poison}).\trg{x\mapsto\loc_\poison^\diamond,\Delta})\ \trg{\hole{\cdot}}\ \trg{\Omega.\Delta}
    $$
  \end{center}
}{twstate-irrelevant}{Irrelevant location marker of $\mmlAt$ operational state.}
\myfig{
  \begin{center}
    \judgbox{\wexect{prog\ e_0\ e_p\ e_1}{\configt}{\valueexpr}{\trace}}{,,Run $\mmlAt$ program $\trg{prog\ e_0\ e_p\ e_1}$,\\giving dynamic state $\configt$  and emitting trace $\trg{\trace}$.''}
  \typerule{$e-\text{wprog}-\trg{trg}$}{
    \exect{\hole{\cdot};\hole{\cdot}}{e_0}{\Omega_0}{\valueexpr_0}{\trace{_0}} \\
    \exect{\qconfigt{_0}}{e_p[\valueexpr_0/x]}{\Omega_p}{\valueexpr_p}{\trace{_p}} &
    \exect{\Omega_p}{e_1[\valueexpr_p/y]}{\Omega_1}{\valueexpr_1}{\trace{_1}} &
  }{
    \wexect{prog\ e_0\ e_p\ e_1}{\Omega_1}{\valueexpr_1}{Ret\ !0\cdot\trace{_0}\cdot Call\ ?\valueexpr_0\cdot \trace{_p}\cdot Ret\ !\valueexpr_p\cdot \trace{_1}\cdot Call\ ?v_1}
  }{wt-e-prog}
  \typerule{$e-\text{wprog}-\trg{trg}-$fail0}{
    \exect{\hole{\cdot};\hole{\cdot}}{e_0}{\lightning}{stuck}{\trace{_0}} \\
  }{
    \wexect{prog\ e_0\ e_p\ e_1}{\lightning}{stuck}{Ret\ !0\cdot \trace{_0}}
  }{wt-e-prog-fail0}
  \typerule{$e-\text{wprog}-\trg{trg}-$failp}{
    \exect{\hole{\cdot};\hole{\cdot}}{e_0}{\Omega_0}{\valueexpr_0}{\trace{_0}} &
    \exect{\Omega_0^\diamond}{e_p[\valueexpr_0/x]}{\lightning}{stuck}{\trace{_p}} &
  }{
    \wexect{prog\ e_0\ e_p\ e_1}{\lightning}{stuck}{Ret\ !0\cdot\trace{_0}\cdot Call\ ?\valueexpr_0\cdot \trace{_p}}
  }{wt-e-prog-failp}
  \typerule{$e-\text{wprog}-\trg{trg}$-fail1}{
    \exect{\hole{\cdot};\hole{\cdot}}{e_0}{\Omega_0}{\valueexpr_0}{\trace{_0}} \\
    \exect{\Omega_0^\diamond}{e_p[\valueexpr_0/x]}{\Omega_p}{\valueexpr_p}{\trace{_p}} &
    \exect{\Omega_p}{e_1[\valueexpr_p/y]}{\lightning}{stuck}{\trace{_1}} &
  }{
    \wexect{prog\ e_0\ e_p\ e_1}{\lightning}{stuck}{Ret\ !0\cdot \trace{_0}\cdot Call\ ?\valueexpr_0\cdot \trace{_p}\cdot Ret\ !\valueexpr_p\cdot \trace{_1}}
  }{wt-e-prog-fail1}
  \end{center}
}{twprog-run}{Running a whole $\mmlAt$ program.}

\subsubsection{Proofs and Auxiliary Lemmas}

\begin{lemma}[Determinism of Step]\label{lem:determ:step}
  If
  \begin{assumptions}
  \item $\exprevalt{\Omega}{e}{\Omega'}{v_1}{\event}$
  \item $\exprevalt{\Omega}{e}{\Omega''}{v_2}{\event}$
  \end{assumptions}
  Then
  \begin{goals}
  \item $\trg{v'}=\trg{v''}$
  \item $\trg{\Omega'}=\trg{\Omega''}$
  \end{goals}
\end{lemma}
\begin{incompleteproof}
\end{incompleteproof}

\begin{lemma}[Determinism of Ctx-Step]\label{lem:determ:ctxstep}
  If
  \begin{assumptions}
  \item $\exect[]{\Omega}{e}{\Omega'}{v_1}{\event}$
  \item $\exect[]{\Omega}{e}{\Omega''}{v_2}{\event}$
  \end{assumptions}
  Then
  \begin{goals}
  \item $\trg{v'}=\trg{v''}$
  \item $\trg{\Omega'}=\trg{\Omega''}$
  \end{goals}
\end{lemma}
\begin{incompleteproof}
\end{incompleteproof}

\begin{lemma}[Determinism of Steps]\label{lem:determ:steps}
  If
  \begin{assumptions}
  \item $\exect{\Omega}{e}{\Omega'}{v_1}{\trace[_{1}]}$
  \item $\exect{\Omega}{e}{\Omega''}{v_2}{\trace[_{2}]}$
  \end{assumptions}
  Then
  \begin{goals}
  \item $\trg{v'}=\trg{v''}$
  \item $\trg{\Omega'}=\trg{\Omega''}$
  \item $\trg{\trace[_{1}]}=\trg{\trace[_{2}]}$
  \end{goals}
\end{lemma}
\begin{incompleteproof}
\end{incompleteproof}

\begin{definition}[Stuck Trace]\label{def:stuck:trgtrace}
  A trace $\trg{\trace}$ is stuck if it ends with $\trg{\lightning}$.
\end{definition}

\begin{definition}[Stuck State]\label{def:stuck:trgstate}
  A state $\trg{\Omega}$ is stuck if it is equal to $\trg{\lightning}$.
\end{definition}

\begin{lemma}[Non-stuck Trace Generation]\label{lem:nonstuck:traces}
  If
  \begin{assumptions}
  \item $\exect{\Omega}{e}{\Omega'}{v}{\trace}$
  \end{assumptions}
  Then
  \begin{goals}
  \item $\trg{\trace}$ is not stuck.
  \item $\trg{\Omega'}$ is not stuck.
  \end{goals}
\end{lemma}
\begin{incompleteproof}
\end{incompleteproof}

\subsubsection{\gls*{sms} Monitor}

\myfig{
  \begin{gather*}
    \begin{aligned}
      \mi{Abstract\ References}~L \subseteq&\ \nat \hspace{0.5cm}
      \mi{Abstract\ Store}~\smsmonitor =\ L\times\nat \\
      \mi{Abstract\ Events}~\event \bnfdef&\ \text{Alloc}\ \loc\ n \mid \text{Dealloc}\ \loc \mid \text{Use}\ \loc\ n \mid \lightning\\
    \mi{References\ Agreement}~\delta_{\text{sms}} :&\ \trg{L}\to L \\[0.6cm]
    \end{aligned}
  \end{gather*}
}{sms-monitor}{\gls{sms} Monitor.}

\myfig{
  \begin{center}
    \judgbox{\monitorcheck[]{\smsmonitor}{\smsmonitor[']}{\event}}{,,Given abstract state $\smsmonitor$, step to abstract state $\smsmonitor'$\\ emitting an abstract event $\event$.''}
    \judgbox{\monitorcheck{\smsmonitor}{\smsmonitor[']}{\trace}}{,,Reflexive-Transitive closure of above judgement.\\$\trace$ is a list of abstract events.''}
    %
    \typerule{sms-Use}{
      (\loc,m)\in \smsmonitor[.A]&
      n<m&
    }{
    \monitorcheck[]{\smsmonitor}{\smsmonitor}{\text{Use}\ \loc\ n}
    }{sms-use}
    %
    \typerule{sms-Alloc}{
      \forall m,(\loc,m)\not\in \smsmonitor&
      \smsmonitor[']=\smsmonitor\cup\left\{(\loc,n)\right\}&
    }{
    \monitorcheck[]{\smsmonitor}{\smsmonitor[']}{\text{Alloc}\ \loc\ n}
    }{sms-alloc}
    %
    \typerule{sms-Dealloc}{
      (\loc,m)\in \smsmonitor[.A] &
      \smsmonitor[']=\smsmonitor\setminus\left\{(\loc,m)\right\}
    }{
    \monitorcheck[]{\smsmonitor}{\smsmonitor[']}{\text{Dealloc}\ \loc}
    }{sms-dealloc}
    %
    \typerule{sms-Refl}{
    }{
    \monitorcheck{\smsmonitor}{\smsmonitor}{\hole{\cdot}}
    }{sms-refl}
    %
    \typerule{sms-Trans}{
      \monitorcheck[]{\smsmonitor}{\smsmonitor[']}{\event}&
      \monitorcheck{\smsmonitor[']}{\smsmonitor['']}{\trace}&
    }{
    \monitorcheck{\smsmonitor}{\smsmonitor['']}{\event\cdot\trace}
    }{sms-trans}
  \end{center}
}{sms-step}{Steps of \gls{sms} Monitor.}

\myfig{
  \begin{center}
    \judgbox{\delta_{\text{sms}}(\trg{\loc})=\loc}{,,A map from $\mmlAt$ memory locations $\trg{\loc}$ to abstract memory locations $\loc$.''}
    \judgbox{\smstraceagree{\trg{\msevent}}{\event}}{,,Memory-Safety specific event $\trg{\msevent}$ describes same action as $\event$.''}
    \judgbox{\smstraceagree{\trg{\mstrace}}{\trace}}{,,Memory-Safety specific trace $\trg{\mstrace}$ describes same trace as $\trace$.''}
    %
    \typerule{SMS-Alloc-Authentic}{
      \delta_{\text{sms}}(\trg{\loc})=\loc
    }{
      \smstraceagree{\specificevt{Alloc\ \loc\ n}}{\text{Alloc}\ \loc\ n}
    }{sms-alloc-auth}
    %
    \typerule{SMS-Dealloc-Authentic}{
      \delta_{\text{sms}}(\trg{\loc})=\loc
    }{
    \smstraceagree{\specificevt{Dealloc\ \loc}}{\text{Dealloc}\ \loc}
    }{sms-dealloc-auth}
    %
    \typerule{SMS-Get-Authentic}{
      \delta_{\text{sms}}(\trg{\loc})=\loc
    }{
    \smstraceagree{\specificevt{Get\ \loc\ n}}{\text{Use}\ \loc\ n}
    }{sms-get-auth}
    %
    \typerule{SMS-Set-Authentic}{
      \delta_{\text{sms}}(\trg{\loc})=\loc
    }{
      \smstraceagree{\specificevt{Set\ \loc\ n\ v}}{\text{Use}\ \loc\ n}
    }{sms-set-auth}
    %
    \typerule{SMS-Refl-Authentic}{
    }{
    \smstraceagree{\trg{\hole{\cdot}}}{\hole{\cdot}}
    }{sms-refl-auth}
    %
    \typerule{SMS-Trans-Authentic}{
      \smstraceagree{\specificevt{\event}}{\event} &
      \smstraceagree{\trg{\mstrace}}{\trace}
    }{
      \smstraceagree{\specificevt{\event}\cdot\trg{\mstrace}}{\event\cdot\trace}
    }{sms-trans-auth}
    %
    \typerule{SMS-Call-Authentic}{
      \smstraceagree{\trg{\mstrace}}{\trace}
    }{
      \smstraceagree{\specificevt{Call\ ?n}\cdot\trg{\mstrace}}{\trace}
    }{sms-call-auth}
    %
    \typerule{SMS-Ret-Authentic}{
      \smstraceagree{\trg{\mstrace}}{\trace}
    }{
      \smstraceagree{\trg{\mstrace}\cdot\specificevt{Ret\ !n}}{\trace}
    }{sms-ret-auth}
    %
  \end{center}
}{sms-trace-agreement}{Trace Agreement.}

\myfig{
  \begin{center}
    \typerule{abstract-agree-dealloc}{
    }{
    \text{Dealloc}\ \loc \prec \text{Dealloc}\ \loc
    }{aagree-dealloc}
    %
    \typerule{abstract-agree-alloc}{
    }{
    \text{Alloc}\ \loc\ n \prec \text{Alloc}\ \loc
    }{aagree-alloc}
    %
    \typerule{abstract-agree-use}{
    }{
    \text{Use}\ \loc\ n \prec \text{Use}\ \loc
    }{use-dealloc}
    %
    \typerule{tms-sms-E}{
    \monitorcheck[]{\smsmonitor}{\smsmonitor}{\trace} & 
    \event\prec \event['] &
    \monitorcheck[]{\tmsmonitor}{\tmsmonitor}{\trace[']} &
    T = \tmsmonitor\times\smsmonitor
    }{
    \monitorcheck[]{T}{T}{\event}
    }{tms-sms-empty}
    %
    \typerule{tms-sms-Refl}{
    }{
    \monitorcheck{T}{T}{\hole{\cdot}}
    }{tms-sms-refl}
    %
    \typerule{tms-sms-Trans}{
      \monitorcheck[]{T}{T'}{\event}&
      \monitorcheck{T'}{T''}{\trace}&
    }{
    \monitorcheck{T}{T''}{\event\cdot\trace}
    }{tms-sms-trans}
  \end{center}
}{tms-sms-step}{Steps of combined \gls{tms} + \gls{sms} Monitor.}

\begin{definition}[Trace is temporal memory safe via monitor.]\label{def:trg:trace:tms}
  \MK{We need to change the agreement definitions from using $\delta_{tms}$ to just $\delta_{ms}$}
  We say $\texttt{TMS}(\trg{\mstrace})$ iff $\exists \delta\ \trace\ \tmsmonitor$ such that:
  \begin{itemize}
    \item $\tmstraceagree{\trg{\mstrace}}{\trace}$
    \item and $\monitorcheck{\emptyset}{\tmsmonitor}{\trace}$
  \end{itemize}
\end{definition}

\subsection{Robust \gls*{tms} Preserving Compiler}

\subsubsection{Compiler}

\myfig{
  \begin{center}
    \begin{gather*}
    \judgbox{\delta(\src{\loc})=\trg{\loc}}{,,The $\mmlAs$ memory location $\src{\loc}$ corresponds to\\the $\mmlAt$ memory location $\trg{\loc}$.''} \\
    \judgbox{\mmlAmmlAtcomp{e}=\trg{e}}{,,Compile $\mmlAs$ expression $\src{e}$ to $\mmlAt$ expression $\trg{e}$.''}\\
    \judgbox{\mmlAmmlAtcomp{\left[\valueexpr / x\right]}=\trg{\left[\valueexpr / x\right]}}{,,Compile $\mmlAs$ substitution to $\mmlAt$ substitution.''}
    \end{gather*}
    $$
    \begin{array}{rll}
    \mmlAmmlAtcomp{x} &\ = & \trg{x} \\
    \mmlAmmlAtcomp{v} &\ = & \trg{v} \\
    \mmlAmmlAtcomp{stuck} &\ = & \trg{stuck}\\
    \mmlAmmlAtcomp{e_1\oplus e_2} &\ = & \mmlAmmlAtcomp{e_1} \trg{\oplus} \mmlAmmlAtcomp{e_2} \\
    \mmlAmmlAtcomp{x[e]} &\ = & \mmlAmmlAtcomp{x}\trg{[}\mmlAmmlAtcomp{e}\trg{]} \\
    \mmlAmmlAtcomp{let\ x=new\ e_1\ in\ e_2} &\ = & \trg{let\ }\mmlAmmlAtcomp{x}\trg{=new\ }\mmlAmmlAtcomp{e_1}\trg{\ in\ }\mmlAmmlAtcomp{e_2} \\
    \mmlAmmlAtcomp{x[e_1]\leftarrow e_2} &\ = & \mmlAmmlAtcomp{x}\trg{[}\mmlAmmlAtcomp{e_1}\trg{]\leftarrow }\mmlAmmlAtcomp{e_2} \\
    \mmlAmmlAtcomp{let\ x= e_1\ in\ e_2} &\ = & \trg{let\ }\mmlAmmlAtcomp{x}\trg{=} \mmlAmmlAtcomp{e_1}\trg{\ in\ }\mmlAmmlAtcomp{e_2} \\
    \mmlAmmlAtcomp{delete\ x} &\ = & \trg{delete\ }\mmlAmmlAtcomp{x} \\
    \mmlAmmlAtcomp{if\ e_1<e_2\ then\ e_3\ else\ e_4} &\ = & \trg{if\ }\mmlAmmlAtcomp{e_1}\trg{<}\mmlAmmlAtcomp{e_2}\\
                                                      &&\trg{then\ }\mmlAmmlAtcomp{e_3} \\
                                                      &&\trg{else\ }\mmlAmmlAtcomp{e_4} \\[0.5cm]
    %
    \mmlAmmlAtcomp{\left[\src{v}/\src{x}\right]}&\ = & \trg{\left[\right.}\mmlAmmlAtcomp{v}\trg{/}\mmlAmmlAtcomp{x}\trg{\left.\right]}\\[0.5cm]
    \end{array}
    $$
  \end{center}
}{tms-pres-compiler}{Compiler from $\mmlAs$ to $\mmlAt$.}
\myfig{
  \begin{center}
    \begin{gather*}
      \judgbox{\mmlAmmlAtcomp{K}=\trg{K}}{,,Compile $\mmlAs$ evaluation contexts to $\mmlAt$ evaluation contexts.''}
    \end{gather*}
    $$
    \begin{array}{rll}
      \mmlAmmlAtcomp{\hole{\cdot}} &\ = & \trg{\hole{\cdot}} \\
      \mmlAmmlAtcomp{K\oplus e} &\ = & \mmlAmmlAtcomp{K}\trg{\oplus}\mmlAmmlAtcomp{e} \\
      \mmlAmmlAtcomp{\valueexpr\oplus K} &\ = & \mmlAmmlAtcomp{\valueexpr}\trg{\oplus}\mmlAmmlAtcomp{K} \\
      \mmlAmmlAtcomp{x[K]} &\ = & \trg{x[}\mmlAmmlAtcomp{K}\trg{]} \\
      \mmlAmmlAtcomp{let\ x\ = K\ in\ e} &\ = & \trg{let\ x\ = }\mmlAmmlAtcomp{K}\trg{\ in\ }\mmlAmmlAtcomp{e} \\
      \mmlAmmlAtcomp{x[K]\leftarrow e} &\ = & \trg{x[}\mmlAmmlAtcomp{K}\trg{]\leftarrow} \mmlAmmlAtcomp{e} \\
      \mmlAmmlAtcomp{x[\valueexpr]\leftarrow K} &\ = & \trg{x[}\mmlAmmlAtcomp{\valueexpr}\trg{]\leftarrow }\mmlAmmlAtcomp{K} \\
      \mmlAmmlAtcomp{let\ x\ = new\ K\ in\ e} &\ = & \trg{let\ x\ = new\ }\mmlAmmlAtcomp{K}\trg{\ in\ }\mmlAmmlAtcomp{e} \\
      \mmlAmmlAtcomp{if\ K<e'\ then\ e_{1}\ else\ e_{2}} &\ = & \trg{if\ }\mmlAmmlAtcomp{K}\trg{<}\mmlAmmlAtcomp{e'}\trg{\ then\ }\\
                                                                &&\mmlAmmlAtcomp{e_{1}}\trg{\ else\ }\mmlAmmlAtcomp{e_{2}} \\
      \mmlAmmlAtcomp{if\ n<K\ then\ e_{1}\ else\ e_{2}} &\ = & \trg{if\ }\mmlAmmlAtcomp{n}\trg{<}\mmlAmmlAtcomp{K}\trg{\ then\ }\\
                                                            &&\mmlAmmlAtcomp{e_{1}}\trg{\ else\ }\mmlAmmlAtcomp{e_{2}} \\
    \end{array}
    $$
  \end{center}
}{tms-pres-compiler-ctx}{Compiling $\mmlAs$ evaluation contexts to $\mmlAt$ evaluation contexts.}
\myfig{
  \begin{center}
    \judgbox{\xlangtraceeq{\trace}{\trace}}{,,The $\mmlAs$ trace $\src{\trace}$ describes the same actions as $\mmlAt$ trace $\trg{\trace}$.''} \\
    \typerule{empty-trace-eq}{
    }{
      \xlangtraceeq{\hole{\cdot}}{\hole{\cdot}}
    }{empty-trace-eq}
    %
    \typerule{cons-trace-eq}{
      \xlangeventeq{\event}{\event} &
      \xlangtraceeq{\trace}{\trace}
    }{
      \xlangtraceeq{\event\cdot\trace}{\event\cdot\trace}
    }{cons-trace-eq}
    %
    \typerule{alloc-event-eq}{
      \mmlAmmlAtcomp{n}=\trg{n} &
      \delta(\src{\loc})=\trg{\loc}
    }{
      \xlangeventeq{Alloc\ \loc\ n}{Alloc\ \loc\ n}
    }{alloc-event-eq}
    %
    \typerule{dealloc-event-eq}{
      \delta(\src{\loc})=\trg{\loc}
    }{
      \xlangeventeq{Dealloc\ \loc}{Dealloc\ \loc}
    }{dealloc-event-eq}
    %
    \typerule{get-event-eq}{
      \mmlAmmlAtcomp{n} = \trg{n} &
      \delta(\src{\loc}) = \trg{\loc}
    }{
      \xlangeventeq{Get\ \loc\ n}{Get\ \loc\ n}
    }{get-event-eq}
    %
    \typerule{set-event-eq}{
      \mmlAmmlAtcomp{n} = \trg{n} &
      \delta(\src{\loc}) = \trg{\loc}
    }{
      \xlangeventeq{Set\ \loc\ n}{Set\ \loc\ n}
    }{set-event-eq}
    %
    \typerule{call-event-eq}{
      \mmlAmmlAtcomp{n} = \trg{n} &
    }{
      \xlangeventeq{Call\ ?n}{Call\ ?n}
    }{call-event-eq}
    %
    \typerule{ret-event-eq}{
      \mmlAmmlAtcomp{n} = \trg{n} &
    }{
      \xlangeventeq{Ret\ !n}{Ret\ !n}
    }{ret-event-eq}
    %
    \typerule{internal-event-eq}{
    }{
      \xlangeventeq{\varepsilon}{\varepsilon}
    }{internal-event-eq}
  \end{center}
}{tms-pres-trace-relation}{Trace Relation from $\mmlAs$ to $\mmlAt$.}
\myfig{
  \begin{center}
    \judgbox{\delta(\src{\loc})=\trg{\loc}}{,,The $\mmlAs$ memory location $\src{\loc}$ corresponds to\\the $\mmlAt$ memory location $\trg{\loc}$.''} \\
    \judgbox{\xlangstateeq{\Omega}{\Omega}}{,,The $\mmlAs$ state $\configs$ agrees with the $\mmlAt$ state $\configt$.''} \\
    %
    \typerule{state-$\lightning$-eq}{
    }{
      \xlangstateeq{\lightning}{\lightning}
    }{state-fail-eq}
    %
    \typerule{state-eq}{
      \src{\Omega}=\src{\Delta;H} &
      \trg{\Omega}=\trg{\Delta;H} \\
      \xlangstateeq{\Delta}{\Delta} &
      \xlangstateeq{H}{H}
    }{
      \xlangstateeq{\Omega}{\Omega}
    }{state-eq}
    %
    \typerule{empty-heap-or-map-eq}{
    }{
      \xlangstateeq{\hole{\cdot}}{\hole{\cdot}}
    }{compile-empty-heap-or-map-eq}
    %
    \typerule{cons-heap-eq}{
      \mmlAmmlAtcomp{n}=\trg{n} &
      \xlangstateeq{H}{H}
    }{
      \xlangstateeq{n,H}{n,H}
    }{compile-cons-heap}
    %
    \typerule{cons-locmap-eq}{
      \mmlAmmlAtcomp{x} = \trg{x} &
      \delta(\src{\loc}) = \trg{\loc} &
      \xlangstateeq{\Delta}{\Delta}
    }{
      \xlangstateeq{x\mapsto\loc_\poison,\Delta}{x\mapsto\loc_\poison,\Delta}
    }{cons-locmap-eq}
    %
    \typerule{cons-locmap-eq-unimportant}{
      \xlangstateeq{\Delta}{\Delta}
    }{
      \xlangstateeq{\Delta}{x\mapsto\loc_\poison^\diamond,\Delta}
    }{cons-locmap-eq-unimportant}
    \vspace{0.5cm}
    $\;$
    %
    \judgbox{\xlangbackstateeq{\Omega}{\Omega}}{,,The $\mmlAt$ state $\configt$ agrees with the $\mmlAs$ state $\configs$.''} \\
    %
    \typerule{back-$\lightning$-eq}{
    }{
      \xlangbackstateeq{\lightning}{\lightning}
    }{back-state-fail-eq}
    %
    \typerule{back-state-eq}{
      \trg{\Omega}=\trg{\Delta;H} &
      \src{\Omega}=\src{\Delta;H} \\
      \xlangbackstateeq{\Delta}{\Delta} &
      \xlangbackstateeq{H}{H}
    }{
      \xlangbackstateeq{\Omega}{\Omega}
    }{back-state-eq}
    %
    \typerule{back-empty-heap-or-map-eq}{
    }{
      \xlangbackstateeq{\hole{\cdot}}{\hole{\cdot}}
    }{back-compile-empty-heap-or-map-eq}
    %
    \typerule{back-cons-heap-eq}{
      \mmlAmmlAtbackv{n}=\src{n} &
      \xlangbackstateeq{H}{H}
    }{
      \xlangbackstateeq{n,H}{n,H}
    }{back-compile-cons-heap}
    %
    \typerule{back-cons-locmap-eq}{
      \mmlAmmlAtcomp{x}=\trg{x} &
      \delta(\src{\loc}) = \trg{\loc} &
      \xlangbackstateeq{\Delta}{\Delta}
    }{
      \xlangbackstateeq{x\mapsto\loc_\poison,\Delta}{x\mapsto\loc_\poison,\Delta}
    }{back-cons-locmap-eq}
    %
    \typerule{back-cons-locmap-eq-unimportant}{
      \xlangbackstateeq{\Delta}{\Delta}
    }{
      \xlangbackstateeq{x\mapsto\loc_\poison^\diamond,\Delta}{\Delta}
    }{back-cons-locmap-eq-unimportant}
  \end{center}
}{tms-pres-state-relations}{State Relations from $\mmlAs$ to $\mmlAt$ and vice versa.}

\begin{lemma}[State equality coincides]\label{lem:stateinsertion}
  Iff
  \begin{assumptions}
  \item $\xlangstateeq{\Omega}{\Omega^\diamond}$
  \end{assumptions}
  then
  \begin{goals}
  \item $\xlangbackstateeq{\Omega^\diamond}{\Omega}$
  \end{goals}
\end{lemma}
\begin{incompleteproof}
\end{incompleteproof}

\begin{lemma}[Related non-crashing states]\label{lem:relatednoncrash}
  If
  \begin{assumptions}
    \item $\src{\Omega}\not=\src{\lightning}$
    \item\label[ass]{asm:relatedstates} $\xlangstateeq{\Omega}{\Omega}$
  \end{assumptions}
  Then
  \begin{goals}
    \item $\trg{\Omega}\not=\trg{\lightning}$
  \end{goals}
\end{lemma}
\begin{proof}
  Simple induction on \Cref{asm:relatedstates}.
\end{proof}

\begin{lemma}[Determinism of Compiler]\label{lem:determcomp}
  If and only if
  \begin{assumptions}
  \item\label[ass]{asm:compdeterm:unequal} $\mmlAmmlAtcomp{e_1}\not=\mmlAmmlAtcomp{e_2}$
  \end{assumptions}
  Then
  \begin{goals}
  \item $\src{e_1}\not=\src{e_2}$
  \end{goals}
\end{lemma}
\begin{incompleteproof}
\end{incompleteproof}

\begin{lemma}[Step Forward Simulation]\label{lem:step:forwardsim}
  If
  \begin{assumptions}
    \item\label[ass]{asm:step:forwardsim:srcstep} $\exprevals{\Omega}{e}{\Omega'}{e'}{\event}$
    \item $\xlangstateeq{\Omega}{\Omega}$
    \item $\typechecks{}{\Omega\triangleright e}{\nat}$
  \end{assumptions}
  Then $\exists\delta'\ \trg{\Omega'}\ \trg{\event}$,
  \begin{goals}
    \item $\delta\subseteq\delta'$
    \item $\expreval{\trg{\Omega}}{\mmlAmmlAtcomp{e}}{\trg{\Omega'}}{\mmlAmmlAtcomp{e'}}{\trg{\event}}$
    \item $\xlangeventeq{\event}{\event}$
    \item $\xlangstateeq{\Omega'}{\Omega'}$
  \end{goals}
\end{lemma}
\begin{proof}
  We proceed by induction on \Cref{asm:step:forwardsim:srcstep}.
  \begin{description}
    \item \textbf{Case $\exprevals{\Omega}{n_1+n_2}{\Omega}{n_3}{\varepsilon}$:}

      If
      \begin{passumptions}{H}
      \item\label[ass]{asm:step:forwardsim:case1:stateeq} $\xlangstateeq{\Omega}{\Omega}$
      \item $\typechecks{}{\Omega\triangleright n_1+n_2}{\nat}$
      \item\label[ass]{asm:step:forwardsim:case1:additioneq} $n_3=n_1+n_2$
      \end{passumptions}
      Then $\exists\delta'\ \trg{\Omega'}\ \trg{\event}$,
      \begin{goals}
        \item $\delta\subseteq\delta'$
        \item $\expreval{\trg{\Omega}}{\mmlAmmlAtcomp{n_1+n_2}}{\trg{\Omega'}}{\mmlAmmlAtcomp{n_3}}{\trg{\event}}$
        \item $\xlangeventeq{\varepsilon}{\event}$
        \item $\xlangstateeq{\Omega}{\Omega'}$
      \end{goals}

      Instantiate the goal with $\delta'=\delta,\ \trg{\Omega'}=\trg{\Omega}$, and $\trg{\event}=\trg{\varepsilon}$, so that what is left to prove is:
      \begin{goals}
        \item\label[goal]{goal:step:forwardsim:case1:deltasubset} $\delta\subseteq\delta$
        \item\label[goal]{goal:step:forwardsim:case1:eval} $\expreval{\trg{\Omega}}{\trg{n_1+n_2}}{\trg{\Omega}}{\trg{n_3}}{\trg{\varepsilon}}$
        \item\label[goal]{goal:step:forwardsim:case1:eventeq} $\xlangeventeq{\varepsilon}{\varepsilon}$
        \item\label[goal]{goal:step:forwardsim:case1:stateeq} $\xlangstateeq{\Omega}{\Omega}$
      \end{goals}

      \Cref{goal:step:forwardsim:case1:deltasubset} follows immediately from reflexivity of the subset relation.

      \Cref{goal:step:forwardsim:case1:eval} follows using \Cref{tr:tw-e-oplus} and \Cref{asm:step:forwardsim:case1:additioneq}.

      \Cref{goal:step:forwardsim:case1:eventeq} follows using \Cref{tr:internal-event-eq}.

      \Cref{goal:step:forwardsim:case1:stateeq} follows using \Cref{asm:step:forwardsim:case1:stateeq}.

    \item \textbf{Case $\exprevals{H;\Delta_1,x\mapsto\loc_\poison,\Delta_2}{delete\ x}{H;\Delta_1,x\mapsto\loc_\poisoned,\Delta_2}{0}{Dealloc\ \loc}$:}

      If
      \begin{passumptions}{H}
      \item\label[ass]{asm:step:forwardsim:case2:stateeq} $\xlangstateeq{H;\Delta_1,x\mapsto\loc_\poison,\Delta_2}{\trg{\Omega}}$
      \item $\typechecks{}{H;\Delta_1,x\mapsto\loc_\poison,\Delta_2\triangleright delete\ x}{\nat}$
      \end{passumptions}
      Then $\exists\delta'\ \trg{\Omega'}\ \trg{\event}$,
      \begin{goals}
        \item $\delta\subseteq\delta'$
        \item $\expreval{\trg{\Omega}}{\mmlAmmlAtcomp{delete\ x}}{\trg{\Omega'}}{\mmlAmmlAtcomp{0}}{\trg{\event}}$
        \item $\xlangeventeq{Dealloc\ \loc}{\event}$
        \item $\xlangstateeq{H;\Delta_1,x\mapsto\loc_\poisoned,\Delta_2}{\Omega'}$
      \end{goals}

      First note that $\trg{\Omega}=\trg{H;\Delta_1,x\mapsto\loc_\poison,\Delta_2}$, otherwise we'd get a contradiction in \Cref{asm:step:forwardsim:case2:stateeq}.\MK{technical lemma needed}

      Instantiate the goal with $\delta'=\delta,\ \trg{\Omega'}=\trg{H;\Delta_1,x\mapsto\loc_\poisoned,\Delta_2}$, and $\trg{\event}=\trg{Dealloc\ \loc}$, so that what is left to prove is:
      \begin{goals}
      \item\label[goal]{goal:step:forwardsim:case2:deltasubset} $\delta\subseteq\delta$
      \item\label[goal]{goal:step:forwardsim:case2:eval} $\expreval{\trg{H;\Delta_1,x\mapsto\loc_\poison,\Delta_2}}{\trg{delete\ x}}{\trg{H;\Delta_1,x\mapsto\loc_\poisoned,\Delta_2}}{\trg{0}}{\trg{Dealloc\ \loc}}$
      \item\label[goal]{goal:step:forwardsim:case2:eventeq} $\xlangeventeq{Dealloc\ \loc}{Dealloc\ \loc}$
      \item\label[goal]{goal:step:forwardsim:case2:stateeq} $\xlangstateeq{H;\Delta_1,x\mapsto\loc_\poisoned,\Delta_2}{H;\Delta_1,x\mapsto\loc_\poisoned,\Delta_2}$
      \end{goals}

      \Cref{goal:step:forwardsim:case2:deltasubset} follows immediately from reflexivity of the subset relation.

      \Cref{goal:step:forwardsim:case2:eval} follows using \Cref{tr:tw-e-delete}.

      For \Cref{goal:step:forwardsim:case2:eventeq}, apply \Cref{tr:dealloc-event-eq}, what is left to show is $\delta(\src{\loc})=\trg{\loc}$.

      Similarily for \Cref{goal:step:forwardsim:case2:stateeq}, apply \Cref{tr:state-eq,tr:cons-locmap-eq} ,,suitably''\MK{needs technical lemma: splitting} so that what is left to show is
      $\xlangstateeq{H}{H}$, $\xlangstateeq{\Delta_1}{\Delta_1}$, $\xlangstateeq{\Delta_2}{\Delta_2}$, $\mmlAmmlAtcomp{x}=\trg{x}$, and, like in the previous case, $\delta(\src{\loc})=\trg{\loc}$.

      $\mmlAmmlAtcomp{x}=\trg{x}$ follows by definition of the compiler, anything else follows immediately by inverting \Cref{asm:step:forwardsim:case2:stateeq} ,,suitably''\MK{needs technical lemma: splitting}.

    \item \textbf{Case $\exprevals{H;\Delta_1,x\mapsto\loc_\poison,\Delta_2}{x[n]}{H;\Delta_1,x\mapsto\loc_\poison,\Delta_2}{v}{Get\ \loc\ n}$:}

      If
      \begin{passumptions}{H}
      \item\label[ass]{asm:step:forwardsim:case3:stateeq} $\xlangstateeq{H;\Delta_1,x\mapsto\loc_\poison,\Delta_2}{\Omega}$
      \item $\typechecks{}{H;\Delta_1,x\mapsto\loc_\poison,\Delta_2\triangleright x[n]}{\nat}$
      \item\label[ass]{asm:step:forwardsim:case3:dom1} $\src{\loc}\in\operatorname{dom} \src{H}\implies\src{v}=\src{H}(\src{\loc+n})$
      \item\label[ass]{asm:step:forwardsim:case3:dom2} $\src{\loc}\not\in\operatorname{dom} \src{H}\implies\src{v}=\src{1729}$
      \end{passumptions}
      Then $\exists\delta'\ \trg{\Omega'}\ \trg{\event}$,
      \begin{goals}
        \item $\delta\subseteq\delta'$
        \item $\expreval{\trg{\Omega}}{\mmlAmmlAtcomp{x[n]}}{\trg{\Omega'}}{\mmlAmmlAtcomp{v}}{\trg{\event}}$
        \item $\xlangeventeq{Get\ \loc\ n}{\event}$
        \item $\xlangstateeq{H;\Delta_1,x\mapsto\loc_\poison,\Delta_2}{\Omega'}$
      \end{goals}
      First note that $\trg{\Omega}=\trg{H;\Delta_1,x\mapsto\loc_\poison,\Delta_2}$, otherwise we'd get a contradiction in \Cref{asm:step:forwardsim:case3:stateeq}.\MK{technical lemma needed}

      Instantiate the goal with $\delta'=\delta,\ \trg{\Omega'}=\trg{H;\Delta_1,x\mapsto\loc_\poison,\Delta_2}$, and $\trg{\event}=\trg{Get\ \loc\ n}$, so that what is left to prove is:
      \begin{goals}
      \item\label[goal]{goal:step:forwardsim:case3:deltasubset} $\delta\subseteq\delta$
        \setcounter{enumi}{2}
      \item\label[goal]{goal:step:forwardsim:case3:eventeq} $\xlangeventeq{Get\ \loc\ n}{Get\ \loc\ n}$
      \item\label[goal]{goal:step:forwardsim:case3:stateeq} $\xlangstateeq{H;\Delta_1,x\mapsto\loc_\poison,\Delta_2}{H;\Delta_1,x\mapsto\loc_\poison,\Delta_2}$
        \setcounter{enumi}{1}
      \item\label[goal]{goal:step:forwardsim:case3:eval} $\expreval{\trg{H;\Delta_1,x\mapsto\loc_\poison,\Delta_2}}{\trg{x[n]}}{\trg{H;\Delta_1,x\mapsto\loc_\poison,\Delta_2}}{\trg{v}}{\trg{Get\ \loc\ n}}$
      \end{goals}

      \Cref{goal:step:forwardsim:case3:deltasubset} follows immediately from reflexivity of the subset relation.

      For \Cref{goal:step:forwardsim:case3:eventeq}, invert \Cref{asm:step:forwardsim:case3:stateeq} ,,suitably''\MK{technical lemma needed: splitting} to obtain the assumption $\delta(\src{\loc})=\trg{\loc}$. The claim then follows by applying \Cref{tr:get-event-eq}.

      \Cref{goal:step:forwardsim:case3:stateeq} follows immediately from \Cref{asm:step:forwardsim:case3:stateeq}.

      \Cref{goal:step:forwardsim:case3:eval} is a bit technical. First apply \Cref{tr:tw-e-get}, what is left to show is:
      \begin{goals}
        \setcounter{enumi}{4}
      \item $\trg{\loc}\in\operatorname{dom} \trg{H}\implies\trg{v}=\trg{H}(\trg{\loc+n})$
      \item $\trg{\loc}\not\in\operatorname{dom} \trg{H}\implies\trg{v}=\trg{1729}$
      \end{goals}

      Now, note that $\src{\loc}\in\operatorname{dom}\src{H}\Leftrightarrow\delta(\src{\loc})=\trg{\loc}\in\operatorname{dom}\trg{H}$.\MK{This is a technical lemma...}
      We continue with a case distinction.
      \begin{description}
        \item \textbf{Case $\src{\loc}\in\operatorname{dom}\src{H}$:}

          \Cref{asm:step:forwardsim:case3:dom1} gives $\src{v}=\src{H}(\src{\loc+n})$.
          Since $\trg{v}=\mmlAmmlAtcomp{v}$, $\trg{v}=\mmlAmmlAtcomp{H(\loc+n)}$.
          Using \Cref{asm:step:forwardsim:case3:stateeq}\MK{technical lemma}, $\trg{v}=\trg{H(\loc+n)}$, done.

        \item \textbf{Case $\src{\loc}\not\in\operatorname{dom}\src{H}$:}

          \Cref{asm:step:forwardsim:case3:dom2} gives $\src{v}=\src{1729}$.
          Since $\trg{v}=\mmlAmmlAtcomp{v}$, $\trg{v}=\mmlAmmlAtcomp{1729}=\trg{1729}$, done.
      \end{description}

    \item \textbf{Case $\exprevals{H;\Delta_1,x\mapsto\loc_\poison,\Delta_2}{x[n]\leftarrow v}{H;\Delta_1,x\mapsto\loc_\poison,\Delta_2}{v}{Set\ \loc\ n}$:}

      If
      \begin{passumptions}{H}
      \item\label[ass]{asm:step:forwardsim:case4:stateeq} $\xlangstateeq{H;\Delta_1,x\mapsto\loc_\poison,\Delta_2}{\Omega}$
      \item $\typechecks{}{H;\Delta_1,x\mapsto\loc_\poison,\Delta_2\triangleright x[n]}{\nat}$
      \item\label[ass]{asm:step:forwardsim:case4:update} $\src{H'}=\src{H}(\src{\loc+n}\mapsto\src{v})$
      \end{passumptions}
      Then $\exists\delta'\ \trg{\Omega'}\ \trg{\event}$,
      \begin{goals}
        \item $\delta\subseteq\delta'$
        \item $\expreval{\trg{\Omega}}{\mmlAmmlAtcomp{x[n]\leftarrow v}}{\trg{\Omega'}}{\mmlAmmlAtcomp{v}}{\trg{\event}}$
        \item $\xlangeventeq{Set\ \loc\ n}{\event}$
        \item $\xlangstateeq{H;\Delta_1,x\mapsto\loc_\poison,\Delta_2}{\Omega'}$
      \end{goals}
      First note that $\trg{\Omega}=\trg{H;\Delta_1,x\mapsto\loc_\poison,\Delta_2}$, otherwise we'd get a contradiction in \Cref{asm:step:forwardsim:case4:stateeq}.\MK{technical lemma needed}

      Instantiate the goal with $\delta'=\delta,\ \trg{\Omega'}=\trg{H;\Delta_1,x\mapsto\loc_\poison,\Delta_2}$, and $\trg{\event}=\trg{Set\ \loc\ n}$, so that what is left to prove is:
      \begin{goals}
      \item\label[goal]{goal:step:forwardsim:case4:deltasubset} $\delta\subseteq\delta$
      \item\label[goal]{goal:step:forwardsim:case4:eval} $\expreval{\trg{H;\Delta_1,x\mapsto\loc_\poison,\Delta_2}}{\trg{x[n]\leftarrow v}}{\trg{H;\Delta_1,x\mapsto\loc_\poison,\Delta_2}}{\trg{v}}{\trg{Set\ \loc\ n}}$
      \item\label[goal]{goal:step:forwardsim:case4:eventeq} $\xlangeventeq{Set\ \loc\ n}{Set\ \loc\ n}$
      \item\label[goal]{goal:step:forwardsim:case4:stateeq} $\xlangstateeq{H;\Delta_1,x\mapsto\loc_\poison,\Delta_2}{H;\Delta_1,x\mapsto\loc_\poison,\Delta_2}$
      \end{goals}

      \Cref{goal:step:forwardsim:case4:deltasubset} follows immediately from reflexivity of the subset relation.

      For \Cref{goal:step:forwardsim:case4:eval}, apply \Cref{tr:tw-e-set}.
      What is left to prove is $\trg{H'} = \trg{H}(\trg{\loc+n}\mapsto \trg{v})$.
      This follows by \Cref{asm:step:forwardsim:case4:stateeq}.\MK{technical lemma: compatibility with update}

      Apply \Cref{tr:set-event-eq} on \Cref{goal:step:forwardsim:case4:eventeq}, what is left to show is $\mmlAmmlAtcomp{n}=\trg{n}$, which is trivial, and $\delta(\src{\loc})=\trg{\loc}$.
      The latter can be obtained by a ,,suitable''\MK{technical lemma needed} inversion of \Cref{asm:step:forwardsim:case4:stateeq}.

      \Cref{goal:step:forwardsim:case4:stateeq} is assumed in \Cref{asm:step:forwardsim:case4:stateeq}.

    \item \textbf{Case $\exprevals{H;\Delta}{let\ x\ =\ new\ n\ in\ e}{H';z\mapsto\loc_\poisonless,\Delta}{e[z/x]}{Alloc\ \loc\ n}$:}

      If
      \begin{passumptions}{H}
      \item\label[ass]{asm:step:forwardsim:case5:stateeq} $\xlangstateeq{H;\Delta}{\Omega}$
      \item $\typechecks{}{H;\Delta\triangleright let\ x\ =\ new\ n\ in\ e}{\nat}$
      \item\label[ass]{asm:step:forwardsim:case5:update} $\src{H'}=\src{H}\ll\src{n}$
      \end{passumptions}
      Then $\exists\delta'\ \trg{\Omega'}\ \trg{\event}$,
      \begin{goals}
        \item $\delta\subseteq\delta'$
        \item $\expreval{\trg{\Omega}}{\mmlAmmlAtcomp{let\ x\ =\ new\ n\ in\ e}}{\trg{\Omega'}}{\mmlAmmlAtcomp{e[z/x]}}{\trg{\event}}$
        \item $\xlangeventeq{Alloc\ \loc\ n}{\event}$
        \item $\xlangstateeq{H;z\mapsto\loc_\poisonless,\Delta}{\Omega'}$
      \end{goals}
      First note that $\trg{\Omega}=\trg{H;z\mapsto\loc_\poisonless,\Delta}$, otherwise we'd get a contradiction in \Cref{asm:step:forwardsim:case5:stateeq}.\MK{technical lemma needed}

      Instantiate the goal with $\delta'=\delta\cup\left\{\src{\loc}\mapsto\trg{\loc}\right\},\ \trg{\Omega'}=\trg{H;z\mapsto\loc_\poisonless,\Delta}$, and $\trg{\event}=\trg{Alloc\ \loc\ n}$, so that what is left to prove is:

      \begin{goals}
      \item\label[goal]{goal:step:forwardsim:case5:deltasubset} $\delta\subseteq\delta\cup\left\{\src{\loc}\mapsto\trg{\loc}\right\}$
    \item\label[goal]{goal:step:forwardsim:case5:eval} $\expreval{\trg{H;\Delta}}{\trg{let\ x\ =\ new\ n\ in\ }\mmlAmmlAtcomp{e}}{\trg{H;z\mapsto\loc_\poisonless,\Delta}}{\mmlAmmlAtcomp{e}[z/x]}{\trg{Alloc\ \loc\ n}}$
      \item\label[goal]{goal:step:forwardsim:case5:eventeq} $\xlangeventeq[\delta\cup\left\{\src{\loc}\mapsto\trg{\loc}\right\}]{Alloc\ \loc\ n}{Alloc\ \loc\ n}$
      \item\label[goal]{goal:step:forwardsim:case5:stateeq} $\xlangstateeq[\delta\cup\left\{\src{\loc}\mapsto\trg{\loc}\right\}]{H;z\mapsto\loc_\poisonless,\Delta}{H;z\mapsto\loc_\poisonless,\Delta}$
      \end{goals}

      \Cref{goal:step:forwardsim:case5:deltasubset} follows by rewriting it as $\forall x, x\in\delta\implies x\in\delta\vee x\in\left\{\src{\loc}\mapsto\trg{\loc}\right\}$, then just choose the left side of the disjunction.

      Apply \Cref{tr:tw-e-let-new} on \Cref{goal:step:forwardsim:case5:eval}, what is left to show is $\trg{H'} = \trg{H} \ll \trg{n}$.
      This follows by \Cref{asm:step:forwardsim:case5:stateeq}.\MK{technical lemma: compatibility with grow}

      Apply \Cref{tr:set-event-eq} on \Cref{goal:step:forwardsim:case5:eventeq}, what is left to show is $\mmlAmmlAtcomp{n}=\trg{n}$, follows immediately by definition of the compiler, and $(\delta\cup\left\{\src{\loc}\mapsto\trg{\loc}\right\})(\src{\loc})=\trg{\loc}$, which follows by definition.

      Apply \Cref{tr:state-eq,tr:cons-locmap-eq} on \Cref{goal:step:forwardsim:case5:stateeq}, what is left to show is $\mmlAmmlAtcomp{x}=\trg{x}$, follows immediately by definition of the compiler, and $(\delta\cup\left\{\src{\loc}\mapsto\trg{\loc}\right\})(\src{\loc})=\trg{\loc}$, which follows by definition.

    \item \textbf{Case $\exprevals{\Omega}{abort()}{\lightning}{stuck}{\lightning}$:}

      If
      \begin{passumptions}{H}
      \item\label[ass]{asm:step:forwardsim:case6:stateeq} $\xlangstateeq{\Omega}{\Omega}$
      \item\label[ass]{asm:step:forwardsim:case6:typechecks} $\typechecks{}{\Omega\triangleright abort()}{\nat}$
      \end{passumptions}
      Then $\exists\delta'\ \trg{\Omega'}\ \trg{\event}$,
      \begin{goals}
        \item $\delta\subseteq\delta'$
        \item $\expreval{\trg{\Omega}}{\mmlAmmlAtcomp{abort()}}{\trg{\Omega'}}{\mmlAmmlAtcomp{stuck}}{\trg{\event}}$
        \item $\xlangeventeq{\lightning}{\event}$
        \item $\xlangstateeq{\lightning}{\Omega'}$
      \end{goals}

      Inverting \Cref{asm:step:forwardsim:case6:typechecks} yields a contradiction.

    \item \textbf{Case $\exprevals{\Omega}{let\ x=y\ in\ e}{\Omega}{e[y/x]}{\varepsilon}$:}

      If
      \begin{passumptions}{H}
      \item\label[ass]{asm:step:forwardsim:case7:stateeq} $\xlangstateeq{\Omega}{\Omega}$
      \item $\typechecks{}{\Omega\triangleright let\ x\ = y\ in\ e}{\nat}$
      \end{passumptions}
      Then $\exists\delta'\ \trg{\Omega'}\ \trg{\event}$,
      \begin{goals}
        \item $\delta\subseteq\delta'$
        \item $\expreval{\trg{\Omega}}{\mmlAmmlAtcomp{let\ x\ = y\ in\ e}}{\trg{\Omega'}}{\mmlAmmlAtcomp{e[y/x]}}{\trg{\event}}$
        \item $\xlangeventeq{\varepsilon}{\event}$
        \item $\xlangstateeq{\Omega}{\Omega'}$
      \end{goals}

      Instantiate the goal with $\delta'=\delta,\ \trg{\Omega'}=\trg{\Omega}$, and $\trg{\event}=\trg{\varepsilon}$, so that what is left to prove is:

      \begin{goals}
        \item\label[goal]{goal:step:forwardsim:case7:deltasubset} $\delta\subseteq\delta$
        \item\label[goal]{goal:step:forwardsim:case7:eval} $\expreval{\trg{\Omega}}{\trg{let\ x\ = y\ in\ }\mmlAmmlAtcomp{e}}{\trg{\Omega}}{\mmlAmmlAtcomp{e}\trg{[y/x]}}{\trg{\varepsilon}}$
        \item\label[goal]{goal:step:forwardsim:case7:eventeq} $\xlangeventeq{\varepsilon}{\varepsilon}$
        \item\label[goal]{goal:step:forwardsim:case7:stateeq} $\xlangstateeq{\Omega}{\Omega}$
      \end{goals}

      \Cref{goal:step:forwardsim:case7:deltasubset} follows by reflexivity of the subset relation.

      \Cref{goal:step:forwardsim:case7:eval} follows by definition of \Cref{tr:tw-e-let-x}.

      \Cref{goal:step:forwardsim:case7:eventeq} follows by definition of \Cref{tr:internal-event-eq}.

      \Cref{goal:step:forwardsim:case7:stateeq} follows by \Cref{asm:step:forwardsim:case7:stateeq}.

    \item \textbf{Case $\exprevals{\Omega}{let\ x=v\ in\ e}{\Omega}{e[v/x]}{\varepsilon}$:}

      This case is completely analogous to the previous case, the only difference is the use of \Cref{tr:tw-e-let-v} instead of \Cref{tr:tw-e-let-x}.

    \item \textbf{Case $\exprevals{\Omega}{if\ n_1<n_2\ then\ e_1\ else\ e_2}{\Omega}{e_1}{\varepsilon}$:}

      If
      \begin{passumptions}{H}
      \item\label[ass]{asm:step:forwardsim:case9:stateeq} $\xlangstateeq{\Omega}{\Omega}$
      \item $\typechecks{}{\Omega\triangleright if\ n_1<n_2\ then\ e_1\ else\ e_2}{\nat}$
      \item\label[ass]{asm:step:forwardsim:case9:less} $n_1<n_2$
      \end{passumptions}
      Then $\exists\delta'\ \trg{\Omega'}\ \trg{\event}$,
      \begin{goals}
        \item $\delta\subseteq\delta'$
        \item $\expreval{\trg{\Omega}}{\mmlAmmlAtcomp{if\ n_1<n_2\ then\ e_1\ else\ e_2}}{\trg{\Omega'}}{\mmlAmmlAtcomp{e_1}}{\trg{\event}}$
        \item $\xlangeventeq{\varepsilon}{\event}$
        \item $\xlangstateeq{\Omega}{\Omega'}$
      \end{goals}

      Instantiate the goal with $\delta'=\delta,\ \trg{\Omega'}=\trg{\Omega}$, and $\trg{\event}=\trg{\varepsilon}$, so that what is left to prove is:

      \begin{goals}
        \item\label[goal]{goal:step:forwardsim:case9:deltasubset} $\delta\subseteq\delta$
        \item\label[goal]{goal:step:forwardsim:case9:eval} $\expreval{\trg{\Omega}}{\trg{if\ n_1<n_2\ then\ }\mmlAmmlAtcomp{e_1}\trg{\ else\ }\mmlAmmlAtcomp{e_2}}{\trg{\Omega}}{\mmlAmmlAtcomp{e_1}}{\trg{\varepsilon}}$
        \item\label[goal]{goal:step:forwardsim:case9:eventeq} $\xlangeventeq{\varepsilon}{\varepsilon}$
        \item\label[goal]{goal:step:forwardsim:case9:stateeq} $\xlangstateeq{\Omega}{\Omega}$
      \end{goals}

      \Cref{goal:step:forwardsim:case9:deltasubset} follows by reflexivity of the subset relation.

      \Cref{goal:step:forwardsim:case9:eval} follows by definition of \Cref{tr:tw-e-if-less,asm:step:forwardsim:case9:less}.

      \Cref{goal:step:forwardsim:case9:eventeq} follows by definition of \Cref{tr:internal-event-eq}.

      \Cref{goal:step:forwardsim:case9:stateeq} follows by \Cref{asm:step:forwardsim:case9:stateeq}.

    \item \textbf{Case $\exprevals{\Omega}{if\ n_1<n_2\ then\ e_1\ else\ e_2}{\Omega}{e_2}{\varepsilon}$:}

      This case is completely analogous to the previous case, the only difference is the use of \Cref{tr:tw-e-if-noless} instead of \Cref{tr:tw-e-if-less}.
  \end{description}
\end{proof}

\begin{lemma}[Ctx Step Forward Simulation]\label{lem:ctxstep:forwardsim}
  If
  \begin{assumptions}
    \item\label[ass]{asm:ctxstep:forwardsim:srcstep} $\execs[]{\Omega}{e}{\Omega'}{e'}{\event}$
    \item\label[ass]{asm:ctxstep:forwardsim:staterel} $\xlangstateeq{\Omega}{\Omega}$
    \item\label[ass]{asm:ctxstep:forwardsim:typechecks} $\typechecks{}{\Omega\triangleright e}{\nat}$
  \end{assumptions}
  Then $\exists\delta'\ \trg{\Omega'}\ \trg{\event}$,
  \begin{goals}
    \item $\delta\subseteq\delta'$
    \item $\exec[]{\trg{\Omega}}{\mmlAmmlAtcomp{e}}{\trg{\Omega'}}{\mmlAmmlAtcomp{e'}}{\trg{\event}}$
    \item $\xlangeventeq{\event}{\event}$
    \item $\xlangstateeq{\Omega'}{\Omega'}$
  \end{goals}
\end{lemma}
\begin{proof}
  Induction on \Cref{asm:ctxstep:forwardsim:srcstep}:

  If
  \begin{passumptions}{H}
    \item\label[ass]{asm:ctxstep:forwardsim:nocrash} $\src{\Omega}\not=\src{\lightning}$
    \item\label[ass]{asm:ctxstep:forwardsim:simplestep} $\exprevals{\Omega}{e_{0}}{\Omega'}{e_{0}'}{\event}$
  \end{passumptions}
  Then $\exists\delta'\ \trg{\Omega'}\ \trg{\event}$,
  \begin{goals}
    \item $\delta\subseteq\delta'$
    \item $\exec[]{\trg{\Omega}}{\mmlAmmlAtcomp{K[e_{0}]}}{\trg{\Omega'}}{\mmlAmmlAtcomp{K[e_{0}']}}{\trg{\event}}$
    \item $\xlangeventeq{\event}{\event}$
    \item $\xlangstateeq{\Omega'}{\Omega'}$
  \end{goals}

  We augment \Cref{asm:ctxstep:forwardsim:typechecks} using \thmref{lem:ctxtyping:plugged}.
  \begin{passumptions}{H}
    \setcounter{enumi}{2}
    \item\label[ass]{asm:ctxstep:forwardsim:e0typechecks} $\typechecks{}{\Omega\triangleright e_{0}}{\nat}$
  \end{passumptions}
  Withg \Cref{asm:ctxstep:forwardsim:simplestep,asm:ctxstep:forwardsim:staterel,asm:ctxstep:forwardsim:e0typechecks} use \thmref{lem:step:forwardsim} to get witnesses $\delta', \trg{\Omega'},$ and $\trg{\event}$ such that:

  \begin{passumptions}{F}
    \item\label[ass]{asm:ctxstep:forwardsim:f1} $\delta\subseteq\delta'$
    \item\label[ass]{asm:ctxstep:forwardsim:trgstep} $\exec[]{\trg{\Omega}}{\mmlAmmlAtcomp{e_{0}}}{\trg{\Omega'}}{\mmlAmmlAtcomp{e_{0}'}}{\trg{\event}}$
    \item\label[ass]{asm:ctxstep:forwardsim:f3} $\xlangeventeq{\event}{\event}$
    \item\label[ass]{asm:ctxstep:forwardsim:f4} $\xlangstateeq{\Omega'}{\Omega'}$
  \end{passumptions}

  Immediately instantiate our goal with those witnesses and use \Cref{asm:ctxstep:forwardsim:f1,asm:ctxstep:forwardsim:f3,asm:ctxstep:forwardsim:f4}.
  What is left to prove is
  \begin{goals}
    \setcounter{enumi}{1}
    \item $\exec[]{\trg{\Omega}}{\trg{K[}\mmlAmmlAtcomp{e_{0}}\trg{]}}{\trg{\Omega'}}{\trg{K[}\mmlAmmlAtcomp{e_{0}'}\trg{]}}{\trg{\event}}$
  \end{goals}
  Use \Cref{tr:tw-e-ectx}, what is left to show is
  \begin{goals}
    \setcounter{enumi}{4}
    \item\label[ass]{goal:ctxstep:forwardsim:nocrash} $\trg{\Omega}\not=\trg{\lightning}$
    \item\label[goal]{goal:ctxstep:forwardsim:trgstep} $\exec[]{\trg{\Omega}}{\mmlAmmlAtcomp{e_{0}}}{\trg{\Omega'}}{\mmlAmmlAtcomp{e_{0}'}}{\trg{\event}}$
  \end{goals}
  \Cref{goal:ctxstep:forwardsim:trgstep} follows immediately by \Cref{asm:ctxstep:forwardsim:trgstep}.
  For \Cref{goal:ctxstep:forwardsim:nocrash}, we can use \thmref{lem:relatednoncrash} relying on \Cref{asm:ctxstep:forwardsim:staterel,asm:ctxstep:forwardsim:nocrash}.
\end{proof}

\begin{lemma}[Forward Simulation]\label{lem:forwardsim}
  If
  \begin{assumptions}
    \item\label[ass]{asm:forwardsim:starstep} $\execs{\Omega}{e}{\Omega'}{\finalexpr}{\trace}$
    \item\label[ass]{asm:forwardsim:staterel} $\xlangstateeq{\Omega}{\Omega}$
    \item\label[ass]{asm:forwardsim:typechecks} $\typechecks{}{\Omega\triangleright e}{\nat}$
  \end{assumptions}
  Then $\exists\delta'\ \trg{\Omega'}\ \trg{\trace}$,
  \begin{goals}
    \item $\delta\subseteq\delta'$
    \item $\exec{\trg{\Omega}}{\mmlAmmlAtcomp{e}}{\trg{\Omega'}}{\mmlAmmlAtcomp{\finalexpr}}{\trg{\trace}}$
    \item $\xlangtraceeq[\delta']{\trace}{\trace}$
    \item $\xlangstateeq[\delta']{\Omega'}{\Omega'}$
  \end{goals}
\end{lemma}
\begin{proof}
  Induction on \Cref{asm:forwardsim:starstep}.
  \begin{description}
    \item \textbf{Case }$\execs{\Omega}{\finalexpr}{\Omega}{\finalexpr}{\hole{\cdot}}$\textbf{:}
          Trivial. Witnesses are $\delta, \trg{\Omega}$, and $\trg{\hole{\cdot}}$, goals follow immediately by reflexivity of the respective relation.

    \item \textbf{Cases }$\execs{\Omega}{e}{\Omega}{\finalexpr}{\event\cdot\trace}$ (\Cref{tr:ws-e-trans-important})\textbf{ and }\\$\execs{\Omega}{e}{\Omega}{\finalexpr}{\trace}$ (\Cref{tr:ws-e-trans-unimportant})\textbf{:}
          \begin{passumptions}{H}
            \item\label[ass]{asm:forwardsim:evnoteps} $\src{a}\not=\src{\varepsilon}$ (provided only in the case of \Cref{tr:ws-e-trans-important})
            \item\label[ass]{asm:forwardsim:ctxstep} $\execs[]{\Omega}{e}{\Omega_{0}}{e_0}{\event}$
            \item $\execs{\Omega_{0}}{e_0}{\Omega'}{\finalexpr}{\trace}$
          \end{passumptions}
          With the inductive hypothesis: $\forall\delta_{\forall}\ \trg{\Omega_{\forall}},\exists \delta_{\IH}\ \trg{\Omega_{\IH}}\ \trg{\trace[_{\IH}]},$
          \begin{passumptions}{\IH}
            \item $\delta_{\forall}\subseteq\delta_{\IH}$
            \item $\exec{\trg{\Omega_{\forall}}}{\mmlAmmlAtcomp{e_{0}}}{\trg{\Omega_{{\IH}}}}{\mmlAmmlAtcomp{\finalexpr}}{\trg{\trace[_{{\IH}}]}}$
            \item\label[ass]{asm:forwardsim:justraceeq} $\xlangtraceeq[\delta_{\IH}]{\trace}{\trace[_{\IH}]}$
            \item $\xlangstateeq[\delta_{\IH}]{\Omega_{0}}{\Omega_{\IH}}$
          \end{passumptions}

          Now apply \thmref{lem:ctxstep:forwardsim} on \Cref{asm:forwardsim:ctxstep} using \Cref{asm:forwardsim:staterel,asm:forwardsim:typechecks},
          giving us witnesses $\delta_0\ \trg{\Omega_{0}}\ \trg{\event}$,
          \begin{passumptions}{F}
            \item\label[ass]{asm:forwardsim:deltasubset0} $\delta\subseteq\delta_{0}$
            \item\label[ass]{asm:forwardsim:thectxstep} $\exec[]{\trg{\Omega}}{\mmlAmmlAtcomp{e}}{\trg{\Omega_{0}}}{\mmlAmmlAtcomp{e_{0}}}{\trg{\event}}$
            \item\label[ass]{asm:forwardsim:eventeq} $\xlangeventeq[\delta_{0}]{\event}{\event}$
            \item $\xlangstateeq[\delta_{0}]{\Omega_{0}}{\Omega_{0}}$
          \end{passumptions}

          Instantiate the inductive hypothesis with $\delta_{\forall}=\delta_{0}$ and $\trg{\Omega_{\forall}}=\trg{\Omega_{0}}$ and obtain witnesses
          $\delta_{\IH}, \trg{\Omega_{\IH}}$, and $\trg{\trace[_{\IH}]}$ such that:
          \begin{passumptions}{\IH'}
            \item\label[ass]{asm:forwardsim:deltasubset:IH} $\delta_{0}\subseteq\delta_{\IH}$
            \item\label[ass]{asm:forwardsim:steps} $\exec{\trg{\Omega_{0}}}{\mmlAmmlAtcomp{e_{0}}}{\trg{\Omega_{{\IH}}}}{\mmlAmmlAtcomp{\finalexpr}}{\trg{\trace[_{{\IH}}]}}$
            \item $\xlangtraceeq[\delta_{\IH}]{\trace}{\trace[_{\IH}]}$
            \item $\xlangstateeq[\delta_{\IH}]{\Omega_{0}}{\Omega_{\IH}}$
          \end{passumptions}

          \begin{description}
            \item \textbf{Case $\src{\event}\not=\src{\varepsilon}$:}
                Our goal looks as follows: $\exists \delta'\ \trg{\Omega'}\ \trg{\trace}$,
                \begin{goals}
                  \item\label[goal]{goal:forwardsim:deltasubset} $\delta\subseteq\delta'$
                  \item\label[goal]{goal:forwardsim:steps} $\exec{\trg{\Omega}}{\mmlAmmlAtcomp{e}}{\trg{\Omega'}}{\mmlAmmlAtcomp{\finalexpr}}{\trg{\trace}}$
                  \item\label[goal]{goal:forwardsim:traceeq} $\xlangtraceeq{\event\cdot\trace}{\trace'}$
                  \item\label[goal]{goal:forwardsim:staterel} $\xlangstateeq{\Omega'}{\Omega'}$
                \end{goals}

                Instantiate the goal with $\delta_{\IH}, \trg{\Omega_{\IH}}, \trg{\event\cdot\trace}$.
                Note that $\delta\subseteq\delta_{\IH}$ (\Cref{goal:forwardsim:deltasubset}) follows by transitivity using \Cref{asm:forwardsim:deltasubset0,asm:forwardsim:deltasubset:IH}.
                Similarily for \Cref{goal:forwardsim:staterel}.\MK{technical lemma...}

                Apply \Cref{tr:wt-e-trans-important} on \Cref{goal:forwardsim:steps}. So, what is left to show is
                \begin{goals}
                  \setcounter{enumi}{2}
                  \item\label[goal]{goal:forwardsim:tracerel} $\xlangtraceeq{\event\cdot\trace}{\event\cdot\trace}$
                  \setcounter{enumi}{4}
                  \item\label[goal]{goal:forwardsim:noeps:ctxstep} $\exec[]{\trg{\Omega}}{\mmlAmmlAtcomp{e}}{\trg{\Omega_{0}}}{\mmlAmmlAtcomp{e_{0}}}{\trg{\event}}$
                  \item\label[goal]{goal:forwardsim:noeps:steps} $\exec{\trg{\Omega_{0}}}{\mmlAmmlAtcomp{e_{0}}}{\trg{\Omega_{\IH}}}{\mmlAmmlAtcomp{\finalexpr}}{\trg{\trace}}$
                \end{goals}
                \Cref{goal:forwardsim:tracerel} follows from \Cref{asm:forwardsim:justraceeq,asm:forwardsim:eventeq} using \Cref{tr:cons-trace-eq}.
                \Cref{goal:forwardsim:noeps:ctxstep} is proven by \Cref{asm:forwardsim:thectxstep}.
                \Cref{goal:forwardsim:noeps:steps} is proven by \Cref{asm:forwardsim:steps}.
            \item \textbf{Case $\src{\event}=\src{\varepsilon}$:}
                Similar to the other case, but make use of \Cref{asm:forwardsim:evnoteps}.
          \end{description}
  \end{description}
\end{proof}

\begin{lemma}[Different Reduction]\label{lem:differentreduction}
  If
  \begin{assumptions}
    \item $\neg\left(\execs{\Omega}{e}{\Omega'}{\valueexpr}{\trace}\right)$
    \item $\execs{\Omega}{e}{\Omega'}{\valueexpr'}{\trace}$
  \end{assumptions}
  Then
  \begin{goals}
    \item $\src{\valueexpr}\not=\src{\valueexpr'}$
  \end{goals}
\end{lemma}
\begin{incompleteproof}
\end{incompleteproof}

\begin{lemma}[Expression Correctness]\label{lem:expression:correctness}
  If
  \begin{assumptions}
  \item\label[ass]{asm:exprcorrect:exec1:trg} $\exec{\trg{\Omega}}{\mmlAmmlAtcomp{e}}{\trg{\Omega'}}{\mmlAmmlAtcomp{\finalexpr}}{\trg{\trace}}$
  \item\label[ass]{asm:exprcorrect:stateeq} $\xlangstateeq{\Omega}{\Omega}$
  \item\label[ass]{asm:expr:corr:typechecks} $\typechecks{}{\Omega\triangleright e}{\nat}$
  \end{assumptions}
  Then $\exists \delta'\ \src{\Omega'}\ \src{\trace}$,
  \begin{goals}
  \item $\delta\subseteq\delta'$
  \item $\execs{\Omega}{e}{\Omega'}{\finalexpr}{\trace}$
  \item $\xlangstateeq[\delta']{\Omega'}{\Omega'}$
  \item $\xlangtraceeq[\delta']{\trace}{\trace}$
  \end{goals}
\end{lemma}
\begin{proof}
  We prove this by contradiction. So, we get the assumption:
  \begin{passumptions}{I}
    \item\label[ass]{asm:exprcorr:toplevel} $\forall\delta'\ \src{\Omega'}\ \src{\trace},\delta'\subset\delta\vee\neg\left(\execs{\Omega}{e}{\Omega'}{\finalexpr}{\trace}\right)\vee\xlangstateneq[\delta']{\Omega'}{\Omega'}\vee\xlangtraceneq[\delta']{\trace}{\trace}$
  \end{passumptions}
  and need to prove $\bot$.
  Apply \thmref{lem:progress} on \Cref{asm:expr:corr:typechecks}, so:
  \begin{passumptions}{I}
    \setcounter{enumi}{1}
    \item\label[ass]{asm:exprcorr:progressor} $\src{e}=\src{\valueexpr}$ or $\exists\src{\Omega'}\ \src{\valueexpr}\ \src{\trace},\execs{\Omega}{e}{\Omega'}{\valueexpr}{\trace}$
  \end{passumptions}
  Proceed by case anaylsis on \Cref{asm:exprcorr:progressor}.
  \begin{description}
    \item \textbf{Case $\src{e}=\src{\valueexpr}$:}
      Specialize \Cref{asm:exprcorr:toplevel} for $\delta'=\delta$, $\src{\Omega'}=\src{\Omega},$ and $\src{\trace}=\src{\hole{\cdot}}$.
      We proceed by case analysis:

      Invert \Cref{asm:exprcorrect:exec1:trg} and proceed by case analysis:
      \begin{description}
        \item \textbf{Case $\delta\subset\delta$:} Trivial.
        \item \textbf{Case }$\neg\left(\execs{\Omega}{\valueexpr}{\Omega}{\valueexpr}{\hole{\cdot}}\right)$\textbf{:}
            Simply apply $\neg\left(\execs{\Omega}{\valueexpr}{\Omega}{\valueexpr}{\hole{\cdot}}\right)$ to the goal $\bot$ and use \Cref{tr:ws-e-refl}.
        \item \textbf{Case }$\xlangstateneq{\Omega}{\Omega}$\textbf{:}
            Apply $\xlangstateneq{\Omega}{\Omega}$ to the goal $\bot$ and use \Cref{asm:exprcorrect:stateeq}.
        \item \textbf{Case }$\xlangtraceneq{\hole{\cdot}}{\hole{\cdot}}$\textbf{:}
            Apply $\xlangtraceneq{\hole{\cdot}}{\hole{\cdot}}$ to the goal $\bot$, use \Cref{tr:empty-trace-eq}.
      \end{description}

    \item \textbf{Case }$\exists\src{\Omega'}\ \src{\valueexpr}\ \src{\trace},\execs{\Omega}{e}{\Omega'}{\finalexpr}{\trace}$\textbf{:}
          First extract the witnesses and obtain:
          \begin{passumptions}{H}
            \item\label[ass]{asm:exprcorr:ereduce} $\execs{\Omega}{e}{\Omega'}{\finalexpr}{\trace}$
          \end{passumptions}

          Apply \thmref{lem:forwardsim} to \Cref{asm:exprcorr:ereduce} with \Cref{asm:exprcorrect:stateeq,asm:expr:corr:typechecks} to get:
          \begin{passumptions}{F}
            \item\label[ass]{asm:exprcorrect:deltavsubset} $\delta\subseteq\delta_{v}$
            \item\label[ass]{asm:exprcorrect:exec2:trg} $\exec{\trg{\Omega}}{\mmlAmmlAtcomp{e}}{\trg{\Omega_{v}'}}{\mmlAmmlAtcomp{\valueexpr}}{\trg{\trace[_{v}]}}$
            \item\label[ass]{asm:exprcorrect:stateeq2} $\xlangstateeq[\delta_{v}]{\Omega'}{\Omega_{v}'}$
            \item\label[ass]{asm:exprcorrect:traceeq2} $\xlangtraceeq[\delta_{v}]{\trace}{\trace[_{v}]}$
          \end{passumptions}

          Use \thmref{lem:determ:steps} on \Cref{asm:exprcorrect:exec1:trg,asm:exprcorrect:exec2:trg}, giving us:
          \begin{passumptions}{K}
            \item\label[ass]{asm:exprcorrect:compiledvalueseq} $\mmlAmmlAtcomp{v}=\mmlAmmlAtcomp{f}$
            \item\label[ass]{asm:exprcorrect:stateequal} $\trg{\Omega'}=\trg{\Omega_{v}'}$
            \item\label[ass]{asm:exprcorrect:traceequal} $\trg{\trace}=\trg{\trace[_{v}]}$
          \end{passumptions}
          Rewrite using \Cref{asm:exprcorrect:stateequal} in \Cref{asm:exprcorrect:stateeq2}, similarily \Cref{asm:exprcorrect:traceequal} in \Cref{asm:exprcorrect:traceeq2}:
          \begin{passumptions}{F'}
            \setcounter{enumi}{2}
            \item\label[ass]{asm:exprcorrect:stateeq3} $\xlangstateeq[\delta_{v}]{\Omega'}{\Omega'}$
            \item\label[ass]{asm:exprcorrect:traceeq3} $\xlangtraceeq[\delta_{v}]{\trace}{\trace}$
          \end{passumptions}
      Specialize \Cref{asm:exprcorr:toplevel} for $\delta'=\delta_{v}$, $\src{\Omega'}=\src{\Omega'},$ and $\src{\trace}=\src{\trace}$.
      We proceed by case analysis:
      \begin{description}
        \item \textbf{Case $\neg\left(\delta\subseteq\delta_{v}\right)$:} Apply $\neg\left(\delta\subseteq\delta_{v}\right)$ on our goal $\bot$, then $\delta\subseteq\delta_{v}$ immediately follows from \Cref{asm:exprcorrect:deltavsubset}.

        \item \textbf{Case }$\neg\left(\execs{\Omega}{e}{\Omega'}{\valueexpr}{\trace}\right)$\textbf{:}
          Using \Cref{asm:exprcorr:ereduce} and the assumption $\neg\left(\execs{\Omega}{e}{\Omega'}{\valueexpr}{\trace}\right)$, we conclude $\src{\valueexpr}\not=\src{f}$ using \thmref{lem:differentreduction}.
          Apply \thmref{lem:determcomp} on \Cref{asm:exprcorrect:compiledvalueseq}, so $\src{\valueexpr}=\src{f}$, contradicting the above.

        \item \textbf{Case }$\xlangstateneq[\delta_{v}]{\Omega'}{\Omega'}$\textbf{:} Immediate contradiction with \Cref{asm:exprcorrect:stateeq3}.
        \item \textbf{Case }$\xlangtraceneq[\delta_{v}]{\trace}{\trace}$\textbf{:} Immediate contradiction with \Cref{asm:exprcorrect:traceeq3}.
      \end{description}
  \end{description}
\end{proof}

\subsubsection{Trace-based Backtranslation}

\myfig{
  \begin{center}
    \judgbox{\Game(\trg{\loc}) = \src{x}}{,,Map from $\mmlAt$ locations to $\mmlAs$ variable names.''}
    \judgbox{\Game[\trg{\loc}] = \Game'}{,,Extend $\Game$ with $\trg{\loc}$ and call the extension $\Game'$.''}
    \typerule{backtrans-create-symb}{
      \trg{\loc}\not\in\operatorname{dom}\Game &
      \text{fresh}\ \src{x} &
    }{
      \Game'[\trg{\loc}]=\Game\cup\left\{\trg{\loc}\mapsto\src{x}\right\}
    }{backtrans-create-symb}
    \typerule{backtrans-nocreate-symb}{
      \trg{\loc}\in\operatorname{dom}\Game &
    }{
      \Game'[\trg{\loc}]=\Game
    }{backtrans-nocreate-symb}
    %
    \typerule{backtrans-lookup-symb}{
      \trg{\loc}\in\operatorname{dom}\Game
    }{
      \Game(\trg{\loc}) = \src{x}
    }{backtrans-lookup-symb}
  \end{center}
}{backtrans-locs-to-var}{\varmapbacktransds to keep track of $\mmlAt$ locations and $\mmlAs$ variables.}

\myfig{
  \begin{center}
    \judgbox{\mmlAmmlAtbackv{\valueexpr} = \src{\valueexpr}}{,,Map an $\mmlAt$ value to an $\mmlAs$ value.''}$\;$\\
    \typerule{backtrans-value}{
    }{
      \mmlAmmlAtbackv{\valueexpr} = \src{\valueexpr}
    }{backtrans-value}
  \end{center}
}{backtrans-values}{Backtranslation of $\mmlAt$ values to $\mmlAs$ values.}
\myfig{
  \begin{center}
    \judgbox{\mmlAmmlAtbackdetail{\mstrace} = \Game',\src{e}}{,,Given a \varmapbacktransds $\Game$, construct a \varmapbacktransds $\Game'$ and\\an $\mmlAs$ expression from memory-specific $\mmlAt$ trace $\trg{\mstrace}$.''}$\;$\\
    %
    \typerule{backtrans-empty}{
    }{
      \mmlAmmlAtbackdetail{\hole{\cdot}} = \Game, \src{42}
    }{backtrans-empty}
    %
    \typerule{backtrans-alloc}{
      \Game'=\Game[\trg{\loc}] &
      \Game'\left(\trg{\loc}\right) = \src{x} &
      \mmlAmmlAtbackdetail[\Game']{\mstrace} = \Game'',\src{e} &
    }{
      \mmlAmmlAtbackdetail{\specificevt{Alloc\ \loc\ n}\cdot\mstrace} = \Game'', \src{let\ x = new\ }\mmlAmmlAtbackv{n}\src{\ in\ e}
    }{backtrans-alloc}
    %
    \typerule{backtrans-dealloc}{
      \Game'=\Game[\trg{\loc}] &
      \mmlAmmlAtbackdetail[\Game']{\mstrace} = \Game'',\src{e} &
    }{
      \mmlAmmlAtbackdetail{\specificevt{Dealloc\ \loc}\cdot\mstrace} = \Game'', \src{delete\ }\Game'(\trg{\loc})\src{; e}
    }{backtrans-dealloc}
    %
    \typerule{backtrans-use}{
      \Game'=\Game[\trg{\loc}] &
      \src{x}=\Game'(\trg{\loc}) &
      \mmlAmmlAtbackdetail[\Game']{\mstrace} = \Game'',\src{e} &
    }{
      \mmlAmmlAtbackdetail{\specificevt{Use\ \loc\ n}\cdot\mstrace} = \Game'',\src{x[}\mmlAmmlAtbackv{n}\src{];e}
    }{backtrans-use}
  \end{center}
}{tms-pres-backtrans}{Trace-Based Backtranslation from $\mmlAt$ events to $\mmlAs$ terms.}
\myfig{
  \begin{center}
    \judgbox{\mmlAmmlAtback{\mstrace} = \Game',\src{e}}{,,Given a \varmapbacktransds $\Game$, construct a \varmapbacktransds $\Game'$ and an $\mmlAs$\\ expression from memory-specific $\mmlAt$ ,,top-level'' trace $\trg{\mstrace}$.''}$\;$\\
    %
    \typerule{backtrans-top-level}{
      \mmlAmmlAtbackdetail{\mstrace} = \Game', \src{e}
    }{
      \mmlAmmlAtback{\specificevt{Ret\ !n_0}\cdot\mstrace\cdot\specificevt{Call\ ?n_1}}, \Game', \src{e}
    }{backtrans-top-level}
    %
    \typerule{backtrans-top-level-crash}{
      \mmlAmmlAtbackdetail{\mstrace} = \Game', \src{e}
    }{
      \mmlAmmlAtback{\specificevt{Ret\ !n_0}\cdot\mstrace\cdot\specificevt{\lightning}} = \Game', \src{e;abort()}
    }{backtrans-top-level-crash}
  \end{center}
}{tms-pres-top-level-backtrans}{,,Top-Level'' trace-based Backtranslation from $\mmlAt$ events to $\mmlAs$ terms.}
We use the syntactic sugar $\src{e_1;e_2}$ for $\src{let\ \_ = e_1\ in\ e_2}$.

\subsubsection{Proofs and Auxiliary Lemmas}

\begin{lemma}[Non-Stuck Target Traces have Source Traces]\label{lem:comptrace:invertible}\ \\
  Given $\delta$ and $\trg{\trace}$ non-stuck, $\exists\src{\trace}, \xlangtraceeq{\trace}{\trace}$
\end{lemma}
\begin{proof}
  Induction on $\trg{\trace}$
\end{proof}

\begin{lemma}[Value Insertion]\label{lem:compvals:insertion}\ \\
  Given $\src{\valueexpr}=\mmlAmmlAtbackv{\valueexpr}$, then $\mmlAmmlAtcomp{\valueexpr}=\trg{\valueexpr}$.
\end{lemma}
\begin{proof}
  Immediate.
\end{proof}

\begin{lemma}[Invertible Operational State]\label{lem:invertible:opstate}\ \\
  Given $\delta$ and $\trg{\Omega}$ non-stuck, $\exists\src{\Omega}, \xlangstateeq{\Omega}{\Omega}$.
\end{lemma}
\begin{incompleteproof}
\end{incompleteproof}

\begin{lemma}[Backtranslation is Well-Typed]\label{lem:backtranslation:welltyped}
  If
  \begin{assumptions}
  \item $\trg{\trace}$ not stuck
  \end{assumptions}
  Then
  \begin{goals}
  \item $\typecheck{}{\mmlAmmlAtback{\trace}}{\src{\nat}}$
  \end{goals}
\end{lemma}
\begin{incompleteproof}
  induction on $\trg{\overline{a}}$
\end{incompleteproof}

\begin{lemma}[State Agreement Insertion]\label{lem:backtranslation:stateagreement:insertion}
  If
  \begin{assumptions}
  \item\label[ass]{ass:stateagreeinsertion:agree} $\xlangstateeq{\Omega}{\Omega}$
  \end{assumptions}
  Then
  \begin{goals}
  \item $\xlangbackstateeq{\Omega}{\Omega}$
  \end{goals}
\end{lemma}
\begin{proof}
  Induction on $\xlangstateeq{\Omega}{\Omega}$ (\Cref{ass:stateagreeinsertion:agree}).
  The only maybe interesting case is for \Cref{tr:cons-locmap-eq}.

  We have
  \begin{passumptions}{H}
  \item $\mmlAmmlAtcomp{x}=\trg{x}$
  \item\label[ass]{asm:stateagreeinsertion:delta} $\delta\left(\src{\loc}\right) = \trg{\loc}$
  \item $\xlangstateeq{\Delta}{\Delta}$
  \end{passumptions}
  and
  \begin{passumptions}{\IH}
  \item\label[ass]{asm:stateagreeinsertion:ih} $\xlangbackstateeq{\Delta}{\Delta}$
  \end{passumptions}
  and what we need to show is
  \begin{goals}
  \item $\xlangbackstateeq{x\mapsto\loc,\Delta}{x\mapsto\loc,\Delta}$
  \end{goals}

  By \Cref{tr:back-cons-locmap-eq}, our goals change to:
  \begin{goals}
  \item $\delta\left(\src{\loc}\right) = \trg{\loc}$
  \item $\xlangbackstateeq{\Delta}{\Delta}$
  \end{goals}

  Claims follow using \Cref{asm:stateagreeinsertion:delta,asm:stateagreeinsertion:ih}.
\end{proof}

\begin{lemma}[Agreement Propagation]\label{lem:trace:agreement:propagation}
  If
  \begin{assumptions}
  \item $\delta$ injective
  \item\label[ass]{asm:trace:agreement:propagation:xagree} $\xlangtraceeq{\trace}{\trace}$, where $\delta:\src{L}\to\trg{L}$
  \item $\tmstraceagree[\delta^{\src{\bullet}}]{\tospecificevs{\trace}}{\trace}$, where $\delta^{\src{\bullet}}_{tms}:\src{L}\to L$
  \end{assumptions}
  Then $\exists \delta^{\trg{\bullet}}_{\text{tms}}$,
  \begin{goals}
  \item $\tmstraceagree[\delta^{\trg{\bullet}}]{\tospecificevt{\trace}}{\trace}$, where $\delta^{\trg{\bullet}}_{tms}:\trg{L}\to L$
  \end{goals}
\end{lemma}
\begin{proof}
  Induction on $\xlangtraceeq{\trace}{\trace}$ (\Cref{asm:trace:agreement:propagation:xagree}).
  The witness of the existential satisfies $\delta^{\trg{\bullet}}_{tms}(\delta(\src{\loc}))=\delta^{\src{\bullet}}_{tms}(\src{\loc})$.
\end{proof}

\begin{lemma}[Backtranslation of Stuck Traces]\label{lem:backtranslation:stuck:correct}
  If
  \begin{assumptions}
  \item $\exect{\Omega}{e}{\lightning}{stuck}{\trace\cdot\lightning}$
\item $\trg{\mstrace}=\thetospecificfilter\left(\trg{\trace}\right)\cdot\specificevt{\lightning}$
  \item $\mmlAmmlAtbackdetail{\specificevt{Ret\ !n_0}\cdot\mstrace}=\Game',\src{e}$
  \item $\xlangbackstateeq{\Omega}{\Omega}$
  \end{assumptions}
  then $\exists \delta'\ \src{\trace}$,
  \begin{goals}
  \item $\delta\subseteq\delta'$
  \item $\exec{\configs}{\src{e}}{\src{\lightning}}{\src{stuck}}{\src{\trace\cdot\lightning}}$
  \item $\src{Ret\ !}\mmlAmmlAtbackv{n_0}\src{\cdot\trace\cdot\lightning}\thexlangtraceeq[\delta']\trg{Ret\ !n_0\cdot\trace\cdot \lightning}$
  \end{goals}
\end{lemma}
\begin{incompleteproof}
\end{incompleteproof}

\begin{lemma}[Backtranslation Correctness]\label{lem:backtranslation:correct}
  If
  \begin{assumptions}
  \item $\exect{\Omega}{e}{\Omega'}{\valueexpr}{\trace}$
  \item $\trg{\mstrace}=\thetospecificfilter\left(\trg{\trace}\right)$
  \item $\mmlAmmlAtback{\specificevt{Ret\ !n_0}\cdot\mstrace\cdot\specificevt{Call\ ?n_1}}=\Game',\src{e}$
  \item $\xlangbackstateeq{\Omega}{\Omega}$
  \end{assumptions}
  then $\exists \delta'\ \src{\Omega'}\ \src{\trace}\ \src{\valueexpr}$,
  \begin{goals}
  \item $\delta\subseteq\delta'$
  \item $\exec{\configs}{\src{e}}{\src{\configs'}}{\src{\valueexpr}}{\src{\trace}}$
  \item $\xlangbackstateeq[\delta']{\Omega'}{\Omega'}$
  \item $\src{Ret\ !}\mmlAmmlAtbackv{n_0}\src{\cdot\trace[_1]\cdot Call\ ?}\mmlAmmlAtbackv{n_1}\thexlangtraceeq[\delta']\trg{Ret\ !n_0\cdot\trace\cdot Call\ ?n_1}$
  \end{goals}
\end{lemma}
\begin{incompleteproof}
  Induction on $\trg{\trace}$.

  \begin{description}
  \item \textbf{Case $\trg{\trace}=\trg{\hole{\cdot}}$:}

  If
  \begin{assumptions}
  \item\label[ass]{asm:backtranscorr:case0:exect} $\exect{\Omega}{e}{\Omega'}{\valueexpr}{\hole{\cdot}}$
  \item $\mmlAmmlAtback{\specificevt{Ret\ !n_0}\cdot\specificevt{Call\ ?n_1}}=\Game,\src{e}$
  \item\label[ass]{asm:backtranscorr:case0:stateeq} $\xlangbackstateeq{\Omega}{\Omega}$
  \end{assumptions}
  then $\exists \delta'\ \src{\Omega'}\ \src{\trace}\ \src{\valueexpr}$,
  \begin{goals}
  \item $\delta\subseteq\delta'$
  \item $\exec{\configs}{\src{e}}{\src{\configs'}}{\src{\valueexpr}}{\src{\trace}}$
  \item $\xlangbackstateeq[\delta']{\Omega'}{\Omega'}$
  \item $\src{Ret\ !}\mmlAmmlAtbackv{n_0}\src{\cdot\trace\cdot Call\ ?}\mmlAmmlAtbackv{n_1}\thexlangtraceeq[\delta']\trg{Ret\ !n_0\cdot Call\ ?n_1}$
  \end{goals}

  By inversion on (b), what is left to show is:
  $\exists \delta'\ \src{\Omega'}\ \src{\trace}\ \src{\valueexpr}$,
  \begin{goals}
  \item $\delta\subseteq\delta'$
  \item $\exec{\configs}{\src{42}}{\src{\configs'}}{\src{\valueexpr}}{\src{\trace}}$
  \item $\xlangbackstateeq[\delta']{\Omega'}{\Omega'}$
  \item $\src{Ret\ !}\mmlAmmlAtbackv{n_0}\src{\cdot\trace\cdot Call\ ?}\mmlAmmlAtbackv{n_1}\thexlangtraceeq[\delta']\trg{Ret\ !n_0\cdot Call\ ?n_1}$
  \end{goals}

  Instantiate our goal with $\delta' = \delta, \src{\Omega'}=\src{\Omega}, \src{\trace}=\src{\hole{\cdot}}, \src{\valueexpr}=\src{42}$, the first two goals follow immediately by reflexivity, the third goal follows by \Cref{asm:backtranscorr:case0:stateeq} after inverting \Cref{asm:backtranscorr:case0:exect}.
  What is left to show is:
  \begin{goals}
    \setcounter{enumi}{3}
  \item $\src{Ret\ !}\mmlAmmlAtbackv{n_0}\src{\cdot Call\ ?}\mmlAmmlAtbackv{n_1}\thexlangtraceeq[\delta]\trg{Ret\ !n_0\cdot Call\ ?n_1}$
  \end{goals}
  Which follows from applying \Cref{tr:cons-trace-eq,tr:ret-event-eq,tr:call-event-eq} and \thmref{lem:compvals:insertion}.

  \item \textbf{Case $\trg{\trace}=\trg{\event\cdot\trace}$:}
  If
  \begin{assumptions}
  \item\label{asm:backtranscorr:indu:step} $\exect{\Omega}{e}{\Omega'}{\valueexpr}{\event\cdot\trace}$
  \item\label{asm:backtranscorr:indu:spectrace} $\trg{\msevent\cdot\mstrace}=\tospecificevt{\event}\cdot\thetospecificfilter\left(\trg{\trace}\right)$
  \item $\mmlAmmlAtback{\specificevt{Ret\ !n_0}\cdot\msevent\cdot\mstrace\cdot\specificevt{Call\ ?n_1}}=\Game',\src{e}$
  \item\label[ass]{asm:backtranscorr:case1:stateeq} $\xlangbackstateeq{\Omega}{\Omega}$
  \end{assumptions}
  then $\exists \delta'\ \src{\Omega'}\ \src{\trace}\ \src{\valueexpr}$,
  \begin{goals}
  \item $\delta\subseteq\delta'$
  \item $\exec{\configs}{\src{e}}{\src{\configs'}}{\src{\valueexpr}}{\src{\trace}}$
  \item $\xlangbackstateeq[\delta']{\Omega'}{\Omega'}$
  \item $\src{Ret\ !}\mmlAmmlAtbackv{n_0}\src{\cdot\trace\cdot Call\ ?}\mmlAmmlAtbackv{n_1}\thexlangtraceeq[\delta']\trg{Ret\ !n_0\cdot \event\cdot\trace\cdot Call\ ?n_1}$
  \end{goals}

  The induction hypothesis is:
  If $\forall \trg{\Omega_{\IH}}\ \trg{\Omega'_{\IH}}\ \trg{e_{\IH}}\ \trg{\valueexpr_{\IH}}\ \src{\Omega_{\IH}}\ \src{e_{\IH}}\ \delta_{\IH}\ \Game_{\IH}\ \Game_{\IH}',$
  \begin{assumptions}
  \item $\exect{\Omega_{\IH}}{e_{\IH}}{\Omega'_{\IH}}{\valueexpr_{\IH}}{\trace}$
  \item $\trg{\mstrace}=\thetospecificfilter\left(\trg{\trace}\right)$
  \item $\mmlAmmlAtback[\Game_{\IH}]{\specificevt{Ret\ !n_0}\cdot\mstrace\cdot\specificevt{Call\ ?n_1}}=\Game_{\IH}',\src{e_{\IH}}$
  \item $\xlangbackstateeq[\delta_{\IH}]{\Omega_{\IH}}{\Omega_{\IH}}$
  \end{assumptions}
  then $\exists \delta'\ \src{\Omega'}\ \src{\trace}\ \src{\valueexpr}$,
  \begin{goals}
  \item $\delta_{\IH}\subseteq\delta'$
  \item $\exec{\src{\Omega_{\IH}}}{\src{e_{\IH}}}{\src{\configs'}}{\src{\valueexpr}}{\src{\trace}}$
  \item $\xlangbackstateeq[\delta']{\Omega'_{\IH}}{\Omega'}$
  \item $\src{Ret\ !}\mmlAmmlAtbackv{n_0}\src{\cdot\trace\cdot Call\ ?}\mmlAmmlAtbackv{n_1}\thexlangtraceeq[\delta']\trg{Ret\ !n_0\cdot \trace\cdot Call\ ?n_1}$
  \end{goals}

  We proceed by case-analysis on $\trg{\msevent}$.
  \begin{description}
    \item \textbf{Case $\trg{\msevent}=\specificevt{Alloc\ \loc\ n}$:}
      Note that:
      \begin{passumptions}{H}
    \item $\mmlAmmlAtback{\specificevt{Ret\ !n_0}\cdot\specificevt{Alloc\ \loc\ n}\cdot\mstrace\cdot\specificevt{Call\ ?n_1}}=\Game',\src{let\ x=new\ }\mmlAmmlAtbackv{n}\src{\ in\ e_a}$
    \item $\Game_x=\Game[\trg{\loc}]$
    \item $\src{x}=\Game_x(\trg{\loc})$
    \item $\mmlAmmlAtback[\Game_x]{\mstrace}=\Game',\src{e_a}$
      \end{passumptions}

      Inversion on \Cref{asm:backtranscorr:indu:spectrace} to get $\specificevt{Alloc\ \loc\ n}=\trg{Alloc\ \loc\ n}$, subsequently inverting \Cref{asm:backtranscorr:indu:step} yields:
      \begin{passumptions}{H}
      \setcounter{enumi}{4}
    \item $\trg{\Omega}=\trg{H;\Delta}$
    \item $\trg{H'}=\grow{\trg{H}}{\trg{n}}$
    \item\label[ass]{asm:backtranscorr:indu:allocstep} $\exect[]{H;\Delta}{let\ x = new\ n\ in\ e'}{H';z\mapsto\loc_{\poisonless},\Delta}{e'[z/x]}{Alloc\ \loc\ n}$
    \item\label[ass]{asm:backtranscorr:indu:steps} $\exect{H';z\mapsto\loc_{\poisonless},\Delta}{e'}{\Omega''}{v}{\trace}$
      \end{passumptions}

      Let $\delta_x=\delta\cup\left\{\src{\loc}\mapsto\trg{\loc}\right\}$, obviously $\delta\subseteq\delta_x$.
      Using \Cref{asm:backtranscorr:case1:stateeq}, we apply \Cref{tr:back-cons-locmap-eq}\MK{we technically need a lemma that says we can apply it under monotone $\delta$, i.e. for "bigger" $\delta$} to get
      \begin{passumptions}{H}
        \setcounter{enumi}{8}
      \item\label[ass]{asm:backtranscorr:indu:staterel} $\xlangbackstateeq[\delta_x]{H';z\mapsto\loc_{\poisonless},\Delta}{H';z\mapsto\loc_{\poisonless},\Delta}$
      \end{passumptions}
      thereby introducing some fresh $\src{z}$ and $\src{\loc}$.
      Furthermore, the following assumptions are easily derivable from the existing ones:
      \begin{passumptions}{H}
        \setcounter{enumi}{9}
      \item\label[ass]{asm:backtranscorr:indu:trace} $\trg{\mstrace}=\thetospecificfilter\left(\trg{\trace}\right)$
      \item\label[ass]{asm:backtranscorr:indu:backtrans} $\mmlAmmlAtback[\Game_x]{\specificevt{Ret\ !n_0}\cdot\mstrace\cdot\specificevt{Call\ ?n_1}}=\Game',\src{e'}$
      \end{passumptions}

      We now instantiate $\IH$ with $\trg{\Omega_{\IH}}=\trg{H';z,\mapsto\loc_{\poisonless},\Delta}$, $\trg{\Omega_{\IH}'}=\trg{\Omega''}$, $\trg{e_{\IH}}=\trg{e'[z/x]}$, $\trg{\valueexpr_{\IH}}=\trg{\valueexpr}$, $\src{\Omega_{\IH}}=\src{H';z\mapsto\loc_{\poisonless},\Delta}$, $\src{e_{\IH}}=\src{e_a}$, $\delta_{\IH}=\delta_x$, $\Game_{\IH}=\Game_x$, and $\Game_{\IH}'=\Game'$.
      Whence, specialize it with \Cref{asm:backtranscorr:indu:steps,asm:backtranscorr:indu:trace,asm:backtranscorr:indu:backtrans,asm:backtranscorr:indu:staterel}, giving us:
      \begin{passumptions}{F}
      \item\label[ass]{asm:backtranscorr:indu:deltasubset} $\delta_x\subseteq\delta'$
      \item\label[ass]{asm:backtranscorr:indu:afterallocstep} $\exec{\src{H';z\mapsto\loc_{\poisonless},\Delta}}{\src{e'[z/x]}}{\src{\configs''}}{\src{\valueexpr}}{\src{\trace}}$
      \item\label[ass]{asm:backtranscorr:indu:finalstaterel} $\xlangbackstateeq[\delta']{\Omega''}{\Omega''}$
      \item\label[ass]{asm:backtranscorr:indu:tracerel0} $\src{Ret\ !}\mmlAmmlAtbackv{n_0}\src{\cdot\trace\cdot Call\ ?}\mmlAmmlAtbackv{n_1}\thexlangtraceeq[\delta']\trg{Ret\ !n_0\cdot \trace\cdot Call\ ?n_1}$
      \end{passumptions}
      Note that we can weaken \Cref{asm:backtranscorr:indu:tracerel0} to:\MK{technical lemma}
      \begin{passumptions}{F}
        \setcounter{enumi}{3}
      \item\label[ass]{asm:backtranscorr:indu:tracerel1} $\src{\trace}\thexlangtraceeq[\delta']\trg{\trace}$
      \end{passumptions}

      For the sake of clarity, we repeat our goal: 
      $\exists \delta'\ \src{\Omega'}\ \src{\trace}\ \src{\valueexpr}$,
      \begin{goals}
      \item $\delta\subseteq\delta'$
      \item $\exec{\configs}{\src{let\ x=new\ n\ in\ e'}}{\src{\configs'}}{\src{\valueexpr}}{\src{\trace}}$
      \item $\xlangbackstateeq[\delta']{\Omega''}{\Omega'}$
      \item $\src{Ret\ !}\mmlAmmlAtbackv{n_0}\src{\cdot\trace\cdot Call\ ?}\mmlAmmlAtbackv{n_1}\thexlangtraceeq[\delta']\trg{Ret\ !n_0\cdot {Alloc\ \loc\ n}\cdot\trace\cdot Call\ ?n_1}$
      \end{goals}

      Instantiate it with $\delta'=\delta'$, $\src{\Omega'} = \src{\Omega''}$, $\src{\trace}=\src{Alloc\ \loc\ n\cdot\trace}$, and $\src{v}=\src{v}$.
      \begin{goals}
      \item $\delta\subseteq\delta'$

        Follows by definition of $\delta_x$ and \Cref{asm:backtranscorr:indu:deltasubset}.
      \item $\exec{\configs}{\src{let\ x=new\ n\ in\ e'}}{\src{\configs''}}{\src{\valueexpr}}{\src{Alloc\ \loc\ n\cdot\trace}}$

        Apply \Cref{tr:ws-e-trans-important}, use \Cref{tr:w-e-let-new} and then we are done using \Cref{asm:backtranscorr:indu:afterallocstep} .
      \item $\xlangbackstateeq[\delta']{\Omega''}{\Omega''}$

        By \Cref{asm:backtranscorr:indu:finalstaterel}.
      \item $\src{Ret\ !}\mmlAmmlAtbackv{n_0}\src{\cdot Alloc\ \loc\ n\cdot\trace\cdot Call\ ?}\mmlAmmlAtbackv{n_1}\thexlangtraceeq[\delta']\trg{Ret\ !n_0\cdot {Alloc\ \loc\ n}\cdot\trace\cdot Call\ ?n_1}$

      Easy using \Cref{tr:empty-trace-eq}, \Cref{tr:cons-trace-eq}, \Cref{tr:alloc-event-eq}, \Cref{tr:call-event-eq}, \Cref{tr:ret-event-eq}, and \thmref{lem:compvals:insertion}.
      \end{goals}

    \item \textbf{Case $\trg{\msevent}=\specificevt{Dealloc\ \loc}$:}
      todo.

    \item \textbf{Case $\trg{\msevent}=\specificevt{Use\ \loc\ n}$:}
      todo.

  \end{description}

  \end{description}
\end{incompleteproof}

\refstepcounter{theoremcounter}
\begin{theorem}[Robust \gls{tms} Preservation]
  If
  \begin{assumptions}
  \item\label[ass]{asm:rtmsp:srctms} $\forall \src{e_1}\ \src{e_2}\ \src{\trace}\ \src{\Omega}\ \src{v},\wexecs{prog\ e_1\ e_p\ e_2}{\configs}{v}{\trace}\rightarrow\texttt{TMS}(\tospecificevcomps{\trace})$
  \item\label[ass]{asm:rtmsp:trgexec} $\wexec{\trg{prog\ e_1}\ \mmlAmmlAtcomp{e_p}\ \trg{e_2}}{\trg{\configt}}{\trg{v}}{\trg{\trace}}$
  \item\label[ass]{asm:rtmsp:component} $\typechecks{}{component\ e_p}{\nat\to\nat}$
  \end{assumptions}
  then
  \begin{goals}
  \item $\texttt{TMS}(\tospecificevcompt{\trace})$
  \end{goals}
\end{theorem}
\begin{incompleteproof}
  \MK{rework}Invert \Cref{asm:rtmsp:trgexec} and note that we assume it terminates in a value, so we immediately disregard the failure cases:
  \begin{passumptions}{H}
  \item\label[ass]{rtmsp:runt:e1} $\exect{\emptyset;\emptyset}{e_1}{\Omega_1}{\valueexpr_1}{\trace[_1]}$
  \item\label[ass]{rtmsp:runt:ep} $\exec{\trg{\Omega_1^\diamond}}{\mmlAmmlAtcomp{e_p}\trg{[\valueexpr_1/x]}}{\trg{\Omega_p}}{\trg{\valueexpr_p}}{\trg{\trace[_p]}}$
  \item\label[ass]{rtmsp:runt:e2} $\exect{\Omega_p}{e_2[\valueexpr_p/y]}{\Omega_2}{\valueexpr_2}{\trace[_2]}$
  \item $\trg{\trace}=\trg{Ret\ !0 \cdot \trace[_1] \cdot Call\ ?v_1 \cdot \trace[_p] \cdot Ret\ !v_p \cdot \trace[_2] \cdot Call\ ?v_2}$
  \end{passumptions}

  \noindent
  We backtranslate $\mmlAmmlAtbackdetail[\emptyset]{Ret\ !0\cdot\trace[_1]\cdot Call\ ?v_1}=\Game,\src{e_1}$.
  Analogousy, $\xlangbackstateeq[\emptyset]{\emptyset}{\emptyset}$ by \Cref{tr:back-compile-empty-heap-or-map-eq}.
  Using these facts, we apply \thmref{lem:backtranslation:correct} on \Cref{rtmsp:runt:e1}:

  \begin{passumptions}{H^{\tiny 1}}
  \item $\emptyset\subseteq\delta_1$, where $\delta_1:\src{L}\to\trg{L}$
  \item\label[ass]{rtmsp:runt:execs1} $\exec{\src{\emptyset}}{\src{e_1}}{\src{\Omega_1}}{\mmlAmmlAtbackv{\valueexpr_1}}{\src{\trace[_1]}}$
  \item\label[ass]{rtmsp:runt:statebackeq} $\xlangbackstateeq[\delta_1]{\Omega_1^\diamond}{\Omega_1}$\MK{this is not what we get from backtrans-corr, but this is what we need here}
  \item ${\src{Ret\ !}\mmlAmmlAtbackv{0}\src{\cdot\trace[_1]\cdot Call\ ?}\mmlAmmlAtbackv{v_1}}\thexlangtraceeq[\delta_1]{\trg{Ret\ !0\cdot\trace[_1]\cdot Call\ ?v_1}}$
  \end{passumptions}

  Using \thmref{lem:compvals:insertion}, we rewrite \Cref{rtmsp:runt:ep}:
  \begin{passumptions}{H'}
    \setcounter{enumi}{1}
\item\label[ass]{rtmsp:runt:ep:compv0} $\exec{\trg{\Omega_1^\diamond}}{\mmlAmmlAtcomp{e_p}\trg{[}\mmlAmmlAtcomp{\valueexpr_1}\trg{/x]}}{\trg{\Omega_p}}{\mmlAmmlAtcomp{\valueexpr_p}}{\trg{\trace[_p]}}$
  \end{passumptions}
  By another technical lemma\MK{state it somewhere}, we rewrite again to push the substitution inwards:
  \begin{passumptions}{H'}
    \setcounter{enumi}{1}
  \item\label[ass]{rtmsp:runt:ep:compv} $\exec{\trg{\Omega_1^\diamond}}{\mmlAmmlAtcomp{e_p[\valueexpr_1/x]}}{\trg{\Omega_p}}{\mmlAmmlAtcomp{\valueexpr_p}}{\trg{\trace[_p]}}$
  \end{passumptions}

  Invert \Cref{asm:rtmsp:component} and observe that $\src{\valueexpr_1}$ is a value, so:
  \begin{passumptions}{H'}
    \setcounter{enumi}{1}
  \item\label[ass]{rtmsp:runt:typechecks} $\typechecks{x:\nat}{e_p}{\nat}$
  \item\label[ass]{rtmsp:runt:vtypechecks} $\typechecks{}{\valueexpr_1}{\nat}$
  \end{passumptions}
  Now use \thmref{lem:substitution} and note that $\src{\hole{\cdot}}=\src{\hole{\cdot}}\circ\src{\hole{\cdot}}$ holds by \Cref{tr:t-empty-split}:
  \begin{passumptions}{H'}
    \setcounter{enumi}{3}
  \item\label[ass]{rtmsp:runt:substtypechecks} $\typechecks{}{e_p[\valueexpr_1/x]}{\nat}$
  \end{passumptions}
  Using yet another technical lemma\MK{state it somewhere} on \Cref{rtmsp:runt:substtypechecks,rtmsp:runt:statebackeq}, we conclude:
  \begin{passumptions}{H'}
    \setcounter{enumi}{4}
  \item\label[ass]{rtmsp:runt:optypechecks} $\typechecks{}{\Omega_1\triangleright e_p[\valueexpr_1/x]}{\nat}$
  \end{passumptions}
  Finally, use \thmref{lem:stateinsertion} on \Cref{rtmsp:runt:statebackeq}:
  \begin{passumptions}{H'}
    \setcounter{enumi}{5}
    \item\label[ass]{rtmsp:runt:stateeq} $\xlangstateeq[\delta_1]{\Omega_1}{\Omega_1^\diamond}$
  \end{passumptions}

  Apply \thmref{lem:expression:correctness}\MK{expression correctness should be polymorphic over the importance of $\trg{\Omega}$} on \Cref{rtmsp:runt:ep:compv} using \Cref{rtmsp:runt:stateeq,rtmsp:runt:optypechecks}:
  \begin{passumptions}{F}
  \item $\delta_1\subseteq\delta_p$
  \item\label[ass]{rtmsp:runt:execs2} $\exec{\src{\Omega_1}}{\src{e_p}}{\src{\Omega_p}}{\src{\valueexpr_p}}{\src{\trace[_p]}}$
  \item\label[ass]{rtmsp:runt:stateeq2} $\xlangstateeq[\delta_p]{\Omega_p}{\Omega_p}$
  \item $\xlangtraceeq[\delta_p]{\trace[_p]}{\trace[_p]}$
  \end{passumptions}

  We backtranslate $\mmlAmmlAtbackdetail[\Game]{Ret\ !v_p \cdot \trace[_2] \cdot Call\ ?v_2}=\Game',\src{e_2}$
  Using \thmref{lem:stateinsertion} on \Cref{rtmsp:runt:stateeq2}, we get:
  \begin{passumptions}{F}
    \setcounter{enumi}{4}
  \item\label[ass]{rtmsp:runt:statebackeq2} $\xlangbackstateeq[\delta_p]{\Omega_p}{\Omega_p}$
  \end{passumptions}

  Using these facts, we apply \thmref{lem:backtranslation:correct}\MK{backtranslation needs to be generalized over substitutions} on \Cref{rtmsp:runt:e2}:
  \begin{passumptions}{H^{\tiny 2}}
  \item $\delta_p\subseteq\delta_2$, where $\delta_2:\src{L}\to\trg{L}$
  \item\label[ass]{rtmsp:runt:execs3} $\exec{\src{\Omega_p}}{\src{e_2[v_p/y]}}{\src{\Omega_2}}{\mmlAmmlAtbackv{v_2}}{\src{\trace[_2]}}$
  \item $\xlangbackstateeq[\delta_2]{\Omega_2^\diamond}{\Omega_2}$
  \item ${\src{Ret\ !}\mmlAmmlAtbackv{v_p}\src{\cdot\trace[_2]\cdot Call\ ?}\mmlAmmlAtbackv{v_2}}\thexlangtraceeq[\delta_1]{\trg{Ret\ !v_p\cdot\trace[_2]\cdot Call\ ?v_2}}$
  \end{passumptions}

  Combine \Cref{rtmsp:runt:execs1,rtmsp:runt:execs2,rtmsp:runt:execs3} using \Cref{tr:ws-e-prog} then apply\MK{can't apply, the goal is on $\theta(\trg{\trace})$} \Cref{asm:rtmsp:srctms} and we are done.

\end{incompleteproof}

\clearpage
\cleardoublepage

\bibliographystyle{plain}
\bibliography{library}

\end{document}

### 2023-05-23 popl/main.tex
add acm template

%%
%% This is file `sample-sigconf.tex',
%% generated with the docstrip utility.
%%
%% The original source files were:
%%
%% samples.dtx  (with options: `sigconf')
%% 
%% IMPORTANT NOTICE:
%% 
%% For the copyright see the source file.
%% 
%% Any modified versions of this file must be renamed
%% with new filenames distinct from sample-sigconf.tex.
%% 
%% For distribution of the original source see the terms
%% for copying and modification in the file samples.dtx.
%% 
%% This generated file may be distributed as long as the
%% original source files, as listed above, are part of the
%% same distribution. (The sources need not necessarily be
%% in the same archive or directory.)
%%
%% Commands for TeXCount
%TC:macro \cite [option:text,text]
%TC:macro \citep [option:text,text]
%TC:macro \citet [option:text,text]
%TC:envir table 0 1
%TC:envir table* 0 1
%TC:envir tabular [ignore] word
%TC:envir displaymath 0 word
%TC:envir math 0 word
%TC:envir comment 0 0
%%
%%
%% The first command in your LaTeX source must be the \documentclass command.
\documentclass[acmsmall,review,screen]{acmart}
%% NOTE that a single column version may be required for 
%% submission and peer review. This can be done by changing
%% the \doucmentclass[...]{acmart} in this template to 
%% \documentclass[manuscript,screen]{acmart}
%% 
%% To ensure 100% compatibility, please check the white list of
%% approved LaTeX packages to be used with the Master Article Template at
%% https://www.acm.org/publications/taps/whitelist-of-latex-packages 
%% before creating your document. The white list page provides 
%% information on how to submit additional LaTeX packages for 
%% review and adoption.
%% Fonts used in the template cannot be substituted; margin 
%% adjustments are not allowed.
%%
%%
%% \BibTeX command to typeset BibTeX logo in the docs
\AtBeginDocument{%
  \providecommand\BibTeX{{%
    \normalfont B\kern-0.5em{\scshape i\kern-0.25em b}\kern-0.8em\TeX}}}

%% Rights management information.  This information is sent to you
%% when you complete the rights form.  These commands have SAMPLE
%% values in them; it is your responsibility as an author to replace
%% the commands and values with those provided to you when you
%% complete the rights form.
\setcopyright{acmcopyright}
\copyrightyear{2018}
\acmYear{2018}
\acmDOI{XXXXXXX.XXXXXXX}

%% These commands are for a PROCEEDINGS abstract or paper.
\acmConference[Conference acronym 'XX]{Make sure to enter the correct
  conference title from your rights confirmation emai}{June 03--05,
  2018}{Woodstock, NY}
%
%  Uncomment \acmBooktitle if th title of the proceedings is different
%  from ``Proceedings of ...''!
%
%\acmBooktitle{Woodstock '18: ACM Symposium on Neural Gaze Detection,
%  June 03--05, 2018, Woodstock, NY} 
\acmPrice{15.00}
\acmISBN{978-1-4503-XXXX-X/18/06}

%%
%% Submission ID.
%% Use this when submitting an article to a sponsored event. You'll
%% receive a unique submission ID from the organizers
%% of the event, and this ID should be used as the parameter to this command.
%%\acmSubmissionID{123-A56-BU3}

%%
%% For managing citations, it is recommended to use bibliography
%% files in BibTeX format.
%%
%% You can then either use BibTeX with the ACM-Reference-Format style,
%% or BibLaTeX with the acmnumeric or acmauthoryear sytles, that include
%% support for advanced citation of software artefact from the
%% biblatex-software package, also separately available on CTAN.
%%
%% Look at the sample-*-biblatex.tex files for templates showcasing
%% the biblatex styles.
%%

%%
%% The majority of ACM publications use numbered citations and
%% references.  The command \citestyle{authoryear} switches to the
%% "author year" style.
%%
%% If you are preparing content for an event
%% sponsored by ACM SIGGRAPH, you must use the "author year" style of
%% citations and references.
%% Uncommenting
%% the next command will enable that style.
\citestyle{acmauthoryear}

%%
%% end of the preamble, start of the body of the document source.
\begin{document}

%%
%% The "title" command has an optional parameter,
%% allowing the author to define a "short title" to be used in page headers.
\title{The Name of the Title is Hope}

%%
%% The "author" command and its associated commands are used to define
%% the authors and their affiliations.
%% Of note is the shared affiliation of the first two authors, and the
%% "authornote" and "authornotemark" commands
%% used to denote shared contribution to the research.
\author{Ben Trovato}
\authornote{Both authors contributed equally to this research.}
\email{trovato@corporation.com}
\orcid{1234-5678-9012}
\author{G.K.M. Tobin}
\authornotemark[1]
\email{webmaster@marysville-ohio.com}
\affiliation{%
  \institution{Institute for Clarity in Documentation}
  \streetaddress{P.O. Box 1212}
  \city{Dublin}
  \state{Ohio}
  \country{USA}
  \postcode{43017-6221}
}

\author{Lars Th{\o}rv{\"a}ld}
\affiliation{%
  \institution{The Th{\o}rv{\"a}ld Group}
  \streetaddress{1 Th{\o}rv{\"a}ld Circle}
  \city{Hekla}
  \country{Iceland}}
\email{larst@affiliation.org}

\author{Valerie B\'eranger}
\affiliation{%
  \institution{Inria Paris-Rocquencourt}
  \city{Rocquencourt}
  \country{France}
}

\author{Aparna Patel}
\affiliation{%
 \institution{Rajiv Gandhi University}
 \streetaddress{Rono-Hills}
 \city{Doimukh}
 \state{Arunachal Pradesh}
 \country{India}}

\author{Huifen Chan}
\affiliation{%
  \institution{Tsinghua University}
  \streetaddress{30 Shuangqing Rd}
  \city{Haidian Qu}
  \state{Beijing Shi}
  \country{China}}

\author{Charles Palmer}
\affiliation{%
  \institution{Palmer Research Laboratories}
  \streetaddress{8600 Datapoint Drive}
  \city{San Antonio}
  \state{Texas}
  \country{USA}
  \postcode{78229}}
\email{cpalmer@prl.com}

\author{John Smith}
\affiliation{%
  \institution{The Th{\o}rv{\"a}ld Group}
  \streetaddress{1 Th{\o}rv{\"a}ld Circle}
  \city{Hekla}
  \country{Iceland}}
\email{jsmith@affiliation.org}

\author{Julius P. Kumquat}
\affiliation{%
  \institution{The Kumquat Consortium}
  \city{New York}
  \country{USA}}
\email{jpkumquat@consortium.net}

%%
%% By default, the full list of authors will be used in the page
%% headers. Often, this list is too long, and will overlap
%% other information printed in the page headers. This command allows
%% the author to define a more concise list
%% of authors' names for this purpose.
\renewcommand{\shortauthors}{Trovato and Tobin, et al.}

%%
%% The abstract is a short summary of the work to be presented in the
%% article.
\begin{abstract}
  A clear and well-documented \LaTeX\ document is presented as an
  article formatted for publication by ACM in a conference proceedings
  or journal publication. Based on the ``acmart'' document class, this
  article presents and explains many of the common variations, as well
  as many of the formatting elements an author may use in the
  preparation of the documentation of their work.
\end{abstract}

%%
%% The code below is generated by the tool at http://dl.acm.org/ccs.cfm.
%% Please copy and paste the code instead of the example below.
%%
\begin{CCSXML}
<ccs2012>
 <concept>
  <concept_id>10010520.10010553.10010562</concept_id>
  <concept_desc>Computer systems organization~Embedded systems</concept_desc>
  <concept_significance>500</concept_significance>
 </concept>
 <concept>
  <concept_id>10010520.10010575.10010755</concept_id>
  <concept_desc>Computer systems organization~Redundancy</concept_desc>
  <concept_significance>300</concept_significance>
 </concept>
 <concept>
  <concept_id>10010520.10010553.10010554</concept_id>
  <concept_desc>Computer systems organization~Robotics</concept_desc>
  <concept_significance>100</concept_significance>
 </concept>
 <concept>
  <concept_id>10003033.10003083.10003095</concept_id>
  <concept_desc>Networks~Network reliability</concept_desc>
  <concept_significance>100</concept_significance>
 </concept>
</ccs2012>
\end{CCSXML}

\ccsdesc[500]{Computer systems organization~Embedded systems}
\ccsdesc[300]{Computer systems organization~Redundancy}
\ccsdesc{Computer systems organization~Robotics}
\ccsdesc[100]{Networks~Network reliability}

%%
%% Keywords. The author(s) should pick words that accurately describe
%% the work being presented. Separate the keywords with commas.
\keywords{datasets, neural networks, gaze detection, text tagging}

%% A "teaser" image appears between the author and affiliation
%% information and the body of the document, and typically spans the
%% page.
\begin{teaserfigure}
  \includegraphics[width=\textwidth]{sampleteaser}
  \caption{Seattle Mariners at Spring Training, 2010.}
  \Description{Enjoying the baseball game from the third-base
  seats. Ichiro Suzuki preparing to bat.}
  \label{fig:teaser}
\end{teaserfigure}

\received{20 February 2007}
\received[revised]{12 March 2009}
\received[accepted]{5 June 2009}

%%
%% This command processes the author and affiliation and title
%% information and builds the first part of the formatted document.
\maketitle

\section{Introduction}
ACM's consolidated article template, introduced in 2017, provides a
consistent \LaTeX\ style for use across ACM publications, and
incorporates accessibility and metadata-extraction functionality
necessary for future Digital Library endeavors. Numerous ACM and
SIG-specific \LaTeX\ templates have been examined, and their unique
features incorporated into this single new template.

If you are new to publishing with ACM, this document is a valuable
guide to the process of preparing your work for publication. If you
have published with ACM before, this document provides insight and
instruction into more recent changes to the article template.

The ``\verb|acmart|'' document class can be used to prepare articles
for any ACM publication --- conference or journal, and for any stage
of publication, from review to final ``camera-ready'' copy, to the
author's own version, with {\itshape very} few changes to the source.

\section{Template Overview}
As noted in the introduction, the ``\verb|acmart|'' document class can
be used to prepare many different kinds of documentation --- a
double-blind initial submission of a full-length technical paper, a
two-page SIGGRAPH Emerging Technologies abstract, a ``camera-ready''
journal article, a SIGCHI Extended Abstract, and more --- all by
selecting the appropriate {\itshape template style} and {\itshape
  template parameters}.

This document will explain the major features of the document
class. For further information, the {\itshape \LaTeX\ User's Guide} is
available from
\url{https://www.acm.org/publications/proceedings-template}.

\subsection{Template Styles}

The primary parameter given to the ``\verb|acmart|'' document class is
the {\itshape template style} which corresponds to the kind of publication
or SIG publishing the work. This parameter is enclosed in square
brackets and is a part of the {\verb|documentclass|} command:
\begin{verbatim}
  \documentclass[STYLE]{acmart}
\end{verbatim}

Journals use one of three template styles. All but three ACM journals
use the {\verb|acmsmall|} template style:
\begin{itemize}
\item {\verb|acmsmall|}: The default journal template style.
\item {\verb|acmlarge|}: Used by JOCCH and TAP.
\item {\verb|acmtog|}: Used by TOG.
\end{itemize}

The majority of conference proceedings documentation will use the {\verb|acmconf|} template style.
\begin{itemize}
\item {\verb|acmconf|}: The default proceedings template style.
\item{\verb|sigchi|}: Used for SIGCHI conference articles.
\item{\verb|sigchi-a|}: Used for SIGCHI ``Extended Abstract'' articles.
\item{\verb|sigplan|}: Used for SIGPLAN conference articles.
\end{itemize}

\subsection{Template Parameters}

In addition to specifying the {\itshape template style} to be used in
formatting your work, there are a number of {\itshape template parameters}
which modify some part of the applied template style. A complete list
of these parameters can be found in the {\itshape \LaTeX\ User's Guide.}

Frequently-used parameters, or combinations of parameters, include:
\begin{itemize}
\item {\verb|anonymous,review|}: Suitable for a ``double-blind''
  conference submission. Anonymizes the work and includes line
  numbers. Use with the \verb|\acmSubmissionID| command to print the
  submission's unique ID on each page of the work.
\item{\verb|authorversion|}: Produces a version of the work suitable
  for posting by the author.
\item{\verb|screen|}: Produces colored hyperlinks.
\end{itemize}

This document uses the following string as the first command in the
source file:
\begin{verbatim}
\documentclass[sigconf]{acmart}
\end{verbatim}

\section{Modifications}

Modifying the template --- including but not limited to: adjusting
margins, typeface sizes, line spacing, paragraph and list definitions,
and the use of the \verb|\vspace| command to manually adjust the
vertical spacing between elements of your work --- is not allowed.

{\bfseries Your document will be returned to you for revision if
  modifications are discovered.}

\section{Typefaces}

The ``\verb|acmart|'' document class requires the use of the
``Libertine'' typeface family. Your \TeX\ installation should include
this set of packages. Please do not substitute other typefaces. The
``\verb|lmodern|'' and ``\verb|ltimes|'' packages should not be used,
as they will override the built-in typeface families.

\section{Title Information}

The title of your work should use capital letters appropriately -
\url{https://capitalizemytitle.com/} has useful rules for
capitalization. Use the {\verb|title|} command to define the title of
your work. If your work has a subtitle, define it with the
{\verb|subtitle|} command.  Do not insert line breaks in your title.

If your title is lengthy, you must define a short version to be used
in the page headers, to prevent overlapping text. The \verb|title|
command has a ``short title'' parameter:
\begin{verbatim}
  \title[short title]{full title}
\end{verbatim}

\section{Authors and Affiliations}

Each author must be defined separately for accurate metadata
identification. Multiple authors may share one affiliation. Authors'
names should not be abbreviated; use full first names wherever
possible. Include authors' e-mail addresses whenever possible.

Grouping authors' names or e-mail addresses, or providing an ``e-mail
alias,'' as shown below, is not acceptable:
\begin{verbatim}
  \author{Brooke Aster, David Mehldau}
  \email{dave,judy,steve@university.edu}
  \email{firstname.lastname@phillips.org}
\end{verbatim}

The \verb|authornote| and \verb|authornotemark| commands allow a note
to apply to multiple authors --- for example, if the first two authors
of an article contributed equally to the work.

If your author list is lengthy, you must define a shortened version of
the list of authors to be used in the page headers, to prevent
overlapping text. The following command should be placed just after
the last \verb|\author{}| definition:
\begin{verbatim}
  \renewcommand{\shortauthors}{McCartney, et al.}
\end{verbatim}
Omitting this command will force the use of a concatenated list of all
of the authors' names, which may result in overlapping text in the
page headers.

The article template's documentation, available at
\url{https://www.acm.org/publications/proceedings-template}, has a
complete explanation of these commands and tips for their effective
use.

Note that authors' addresses are mandatory for journal articles.

\section{Rights Information}

Authors of any work published by ACM will need to complete a rights
form. Depending on the kind of work, and the rights management choice
made by the author, this may be copyright transfer, permission,
license, or an OA (open access) agreement.

Regardless of the rights management choice, the author will receive a
copy of the completed rights form once it has been submitted. This
form contains \LaTeX\ commands that must be copied into the source
document. When the document source is compiled, these commands and
their parameters add formatted text to several areas of the final
document:
\begin{itemize}
\item the ``ACM Reference Format'' text on the first page.
\item the ``rights management'' text on the first page.
\item the conference information in the page header(s).
\end{itemize}

Rights information is unique to the work; if you are preparing several
works for an event, make sure to use the correct set of commands with
each of the works.

The ACM Reference Format text is required for all articles over one
page in length, and is optional for one-page articles (abstracts).

\section{CCS Concepts and User-Defined Keywords}

Two elements of the ``acmart'' document class provide powerful
taxonomic tools for you to help readers find your work in an online
search.

The ACM Computing Classification System ---
\url{https://www.acm.org/publications/class-2012} --- is a set of
classifiers and concepts that describe the computing
discipline. Authors can select entries from this classification
system, via \url{https://dl.acm.org/ccs/ccs.cfm}, and generate the
commands to be included in the \LaTeX\ source.

User-defined keywords are a comma-separated list of words and phrases
of the authors' choosing, providing a more flexible way of describing
the research being presented.

CCS concepts and user-defined keywords are required for for all
articles over two pages in length, and are optional for one- and
two-page articles (or abstracts).

\section{Sectioning Commands}

Your work should use standard \LaTeX\ sectioning commands:
\verb|section|, \verb|subsection|, \verb|subsubsection|, and
\verb|paragraph|. They should be numbered; do not remove the numbering
from the commands.

Simulating a sectioning command by setting the first word or words of
a paragraph in boldface or italicized text is {\bfseries not allowed.}

\section{Tables}

The ``\verb|acmart|'' document class includes the ``\verb|booktabs|''
package --- \url{https://ctan.org/pkg/booktabs} --- for preparing
high-quality tables.

Table captions are placed {\itshape above} the table.

Because tables cannot be split across pages, the best placement for
them is typically the top of the page nearest their initial cite.  To
ensure this proper ``floating'' placement of tables, use the
environment \textbf{table} to enclose the table's contents and the
table caption.  The contents of the table itself must go in the
\textbf{tabular} environment, to be aligned properly in rows and
columns, with the desired horizontal and vertical rules.  Again,
detailed instructions on \textbf{tabular} material are found in the
\textit{\LaTeX\ User's Guide}.

Immediately following this sentence is the point at which
Table~\ref{tab:freq} is included in the input file; compare the
placement of the table here with the table in the printed output of
this document.

\begin{table}
  \caption{Frequency of Special Characters}
  \label{tab:freq}
  \begin{tabular}{ccl}
    \toprule
    Non-English or Math&Frequency&Comments\\
    \midrule
    \O & 1 in 1,000& For Swedish names\\
    $\pi$ & 1 in 5& Common in math\\
    \$ & 4 in 5 & Used in business\\
    $\Psi^2_1$ & 1 in 40,000& Unexplained usage\\
  \bottomrule
\end{tabular}
\end{table}

To set a wider table, which takes up the whole width of the page's
live area, use the environment \textbf{table*} to enclose the table's
contents and the table caption.  As with a single-column table, this
wide table will ``float'' to a location deemed more
desirable. Immediately following this sentence is the point at which
Table~\ref{tab:commands} is included in the input file; again, it is
instructive to compare the placement of the table here with the table
in the printed output of this document.

\begin{table*}
  \caption{Some Typical Commands}
  \label{tab:commands}
  \begin{tabular}{ccl}
    \toprule
    Command &A Number & Comments\\
    \midrule
    \texttt{{\char'134}author} & 100& Author \\
    \texttt{{\char'134}table}& 300 & For tables\\
    \texttt{{\char'134}table*}& 400& For wider tables\\
    \bottomrule
  \end{tabular}
\end{table*}

Always use midrule to separate table header rows from data rows, and
use it only for this purpose. This enables assistive technologies to
recognise table headers and support their users in navigating tables
more easily.

\section{Math Equations}
You may want to display math equations in three distinct styles:
inline, numbered or non-numbered display.  Each of the three are
discussed in the next sections.

\subsection{Inline (In-text) Equations}
A formula that appears in the running text is called an inline or
in-text formula.  It is produced by the \textbf{math} environment,
which can be invoked with the usual
\texttt{{\char'134}begin\,\ldots{\char'134}end} construction or with
the short form \texttt{\$\,\ldots\$}. You can use any of the symbols
and structures, from $\alpha$ to $\omega$, available in
\LaTeX~\cite{Lamport:LaTeX}; this section will simply show a few
examples of in-text equations in context. Notice how this equation:
\begin{math}
  \lim_{n\rightarrow \infty}x=0
\end{math},
set here in in-line math style, looks slightly different when
set in display style.  (See next section).

\subsection{Display Equations}
A numbered display equation---one set off by vertical space from the
text and centered horizontally---is produced by the \textbf{equation}
environment. An unnumbered display equation is produced by the
\textbf{displaymath} environment.

Again, in either environment, you can use any of the symbols and
structures available in \LaTeX\@; this section will just give a couple
of examples of display equations in context.  First, consider the
equation, shown as an inline equation above:
\begin{equation}
  \lim_{n\rightarrow \infty}x=0
\end{equation}
Notice how it is formatted somewhat differently in
the \textbf{displaymath}
environment.  Now, we'll enter an unnumbered equation:
\begin{displaymath}
  \sum_{i=0}^{\infty} x + 1
\end{displaymath}
and follow it with another numbered equation:
\begin{equation}
  \sum_{i=0}^{\infty}x_i=\int_{0}^{\pi+2} f
\end{equation}
just to demonstrate \LaTeX's able handling of numbering.

\section{Figures}

The ``\verb|figure|'' environment should be used for figures. One or
more images can be placed within a figure. If your figure contains
third-party material, you must clearly identify it as such, as shown
in the example below.
\begin{figure}[h]
  \centering
  \includegraphics[width=\linewidth]{sample-franklin}
  \caption{1907 Franklin Model D roadster. Photograph by Harris \&
    Ewing, Inc. [Public domain], via Wikimedia
    Commons. (\url{https://goo.gl/VLCRBB}).}
  \Description{A woman and a girl in white dresses sit in an open car.}
\end{figure}

Your figures should contain a caption which describes the figure to
the reader.

Figure captions are placed {\itshape below} the figure.

Every figure should also have a figure description unless it is purely
decorative. These descriptions convey what’s in the image to someone
who cannot see it. They are also used by search engine crawlers for
indexing images, and when images cannot be loaded.

A figure description must be unformatted plain text less than 2000
characters long (including spaces).  {\bfseries Figure descriptions
  should not repeat the figure caption – their purpose is to capture
  important information that is not already provided in the caption or
  the main text of the paper.} For figures that convey important and
complex new information, a short text description may not be
adequate. More complex alternative descriptions can be placed in an
appendix and referenced in a short figure description. For example,
provide a data table capturing the information in a bar chart, or a
structured list representing a graph.  For additional information
regarding how best to write figure descriptions and why doing this is
so important, please see
\url{https://www.acm.org/publications/taps/describing-figures/}.

\subsection{The ``Teaser Figure''}

A ``teaser figure'' is an image, or set of images in one figure, that
are placed after all author and affiliation information, and before
the body of the article, spanning the page. If you wish to have such a
figure in your article, place the command immediately before the
\verb|\maketitle| command:
\begin{verbatim}
  \begin{teaserfigure}
    \includegraphics[width=\textwidth]{sampleteaser}
    \caption{figure caption}
    \Description{figure description}
  \end{teaserfigure}
\end{verbatim}

\section{Citations and Bibliographies}

The use of \BibTeX\ for the preparation and formatting of one's
references is strongly recommended. Authors' names should be complete
--- use full first names (``Donald E. Knuth'') not initials
(``D. E. Knuth'') --- and the salient identifying features of a
reference should be included: title, year, volume, number, pages,
article DOI, etc.

The bibliography is included in your source document with these two
commands, placed just before the \verb|\end{document}| command:
\begin{verbatim}
  \bibliographystyle{ACM-Reference-Format}
  \bibliography{bibfile}
\end{verbatim}
where ``\verb|bibfile|'' is the name, without the ``\verb|.bib|''
suffix, of the \BibTeX\ file.

Citations and references are numbered by default. A small number of
ACM publications have citations and references formatted in the
``author year'' style; for these exceptions, please include this
command in the {\bfseries preamble} (before the command
``\verb|\begin{document}|'') of your \LaTeX\ source:
\begin{verbatim}
  \citestyle{acmauthoryear}
\end{verbatim}

  Some examples.  A paginated journal article \cite{Abril07}, an
  enumerated journal article \cite{Cohen07}, a reference to an entire
  issue \cite{JCohen96}, a monograph (whole book) \cite{Kosiur01}, a
  monograph/whole book in a series (see 2a in spec. document)
  \cite{Harel79}, a divisible-book such as an anthology or compilation
  \cite{Editor00} followed by the same example, however we only output
  the series if the volume number is given \cite{Editor00a} (so
  Editor00a's series should NOT be present since it has no vol. no.),
  a chapter in a divisible book \cite{Spector90}, a chapter in a
  divisible book in a series \cite{Douglass98}, a multi-volume work as
  book \cite{Knuth97}, a couple of articles in a proceedings (of a
  conference, symposium, workshop for example) (paginated proceedings
  article) \cite{Andler79, Hagerup1993}, a proceedings article with
  all possible elements \cite{Smith10}, an example of an enumerated
  proceedings article \cite{VanGundy07}, an informally published work
  \cite{Harel78}, a couple of preprints \cite{Bornmann2019,
    AnzarootPBM14}, a doctoral dissertation \cite{Clarkson85}, a
  master's thesis: \cite{anisi03}, an online document / world wide web
  resource \cite{Thornburg01, Ablamowicz07, Poker06}, a video game
  (Case 1) \cite{Obama08} and (Case 2) \cite{Novak03} and \cite{Lee05}
  and (Case 3) a patent \cite{JoeScientist001}, work accepted for
  publication \cite{rous08}, 'YYYYb'-test for prolific author
  \cite{SaeediMEJ10} and \cite{SaeediJETC10}. Other cites might
  contain 'duplicate' DOI and URLs (some SIAM articles)
  \cite{Kirschmer:2010:AEI:1958016.1958018}. Boris / Barbara Beeton:
  multi-volume works as books \cite{MR781536} and \cite{MR781537}. A
  couple of citations with DOIs:
  \cite{2004:ITE:1009386.1010128,Kirschmer:2010:AEI:1958016.1958018}. Online
  citations: \cite{TUGInstmem, Thornburg01, CTANacmart}. Artifacts:
  \cite{R} and \cite{UMassCitations}.

\section{Acknowledgments}

Identification of funding sources and other support, and thanks to
individuals and groups that assisted in the research and the
preparation of the work should be included in an acknowledgment
section, which is placed just before the reference section in your
document.

This section has a special environment:
\begin{verbatim}
  \begin{acks}
  ...
  \end{acks}
\end{verbatim}
so that the information contained therein can be more easily collected
during the article metadata extraction phase, and to ensure
consistency in the spelling of the section heading.

Authors should not prepare this section as a numbered or unnumbered {\verb|\section|}; please use the ``{\verb|acks|}'' environment.

\section{Appendices}

If your work needs an appendix, add it before the
``\verb|\end{document}|'' command at the conclusion of your source
document.

Start the appendix with the ``\verb|appendix|'' command:
\begin{verbatim}
  \appendix
\end{verbatim}
and note that in the appendix, sections are lettered, not
numbered. This document has two appendices, demonstrating the section
and subsection identification method.

\section{Multi-language papers}

Papers may be written in languages other than English or include
titles, subtitles, keywords and abstracts in different languages (as a
rule, a paper in a language other than English should include an
English title and an English abstract).  Use \verb|language=...| for
every language used in the paper.  The last language indicated is the
main language of the paper.  For example, a French paper with
additional titles and abstracts in English and German may start with
the following command
\begin{verbatim}
\documentclass[sigconf, language=english, language=german,
               language=french]{acmart}
\end{verbatim}

The title, subtitle, keywords and abstract will be typeset in the main
language of the paper.  The commands \verb|\translatedXXX|, \verb|XXX|
begin title, subtitle and keywords, can be used to set these elements
in the other languages.  The environment \verb|translatedabstract| is
used to set the translation of the abstract.  These commands and
environment have a mandatory first argument: the language of the
second argument.  See \verb|sample-sigconf-i13n.tex| file for examples
of their usage.

\section{SIGCHI Extended Abstracts}

The ``\verb|sigchi-a|'' template style (available only in \LaTeX\ and
not in Word) produces a landscape-orientation formatted article, with
a wide left margin. Three environments are available for use with the
``\verb|sigchi-a|'' template style, and produce formatted output in
the margin:
\begin{itemize}
\item {\verb|sidebar|}:  Place formatted text in the margin.
\item {\verb|marginfigure|}: Place a figure in the margin.
\item {\verb|margintable|}: Place a table in the margin.
\end{itemize}

%%
%% The acknowledgments section is defined using the "acks" environment
%% (and NOT an unnumbered section). This ensures the proper
%% identification of the section in the article metadata, and the
%% consistent spelling of the heading.
\begin{acks}
To Robert, for the bagels and explaining CMYK and color spaces.
\end{acks}

%%
%% The next two lines define the bibliography style to be used, and
%% the bibliography file.
\bibliographystyle{ACM-Reference-Format}
\bibliography{sample-base}

%%
%% If your work has an appendix, this is the place to put it.
\appendix

\section{Research Methods}

\subsection{Part One}

Lorem ipsum dolor sit amet, consectetur adipiscing elit. Morbi
malesuada, quam in pulvinar varius, metus nunc fermentum urna, id
sollicitudin purus odio sit amet enim. Aliquam ullamcorper eu ipsum
vel mollis. Curabitur quis dictum nisl. Phasellus vel semper risus, et
lacinia dolor. Integer ultricies commodo sem nec semper.

\subsection{Part Two}

Etiam commodo feugiat nisl pulvinar pellentesque. Etiam auctor sodales
ligula, non varius nibh pulvinar semper. Suspendisse nec lectus non
ipsum convallis congue hendrerit vitae sapien. Donec at laoreet
eros. Vivamus non purus placerat, scelerisque diam eu, cursus
ante. Etiam aliquam tortor auctor efficitur mattis.

\section{Online Resources}

Nam id fermentum dui. Suspendisse sagittis tortor a nulla mollis, in
pulvinar ex pretium. Sed interdum orci quis metus euismod, et sagittis
enim maximus. Vestibulum gravida massa ut felis suscipit
congue. Quisque mattis elit a risus ultrices commodo venenatis eget
dui. Etiam sagittis eleifend elementum.

Nam interdum magna at lectus dignissim, ac dignissim lorem
rhoncus. Maecenas eu arcu ac neque placerat aliquam. Nunc pulvinar
massa et mattis lacinia.

\end{document}
\endinput
%%
%% End of file `sample-sigconf.tex'.

### 2023-06-05 popl/main.tex
Abstract

@@
 \setcopyright{acmcopyright}
-\copyrightyear{2018}
-\acmYear{2018}
+\copyrightyear{2024}
+\acmYear{2024}
 \acmDOI{XXXXXXX.XXXXXXX}
@@
 %% These commands are for a PROCEEDINGS abstract or paper.
-\acmConference[Conference acronym 'XX]{Make sure to enter the correct
+\acmConference[POPL '24]{Make sure to enter the correct
   conference title from your rights confirmation emai}{June 03--05,
@@
 %% allowing the author to define a "short title" to be used in page headers.
-\title{The Name of the Title is Hope}
+\title{Composing Robustly Safe Compilers}
 
@@
 %% used to denote shared contribution to the research.
-\author{Ben Trovato}
+\author{Matthis Kruse}
 \authornote{Both authors contributed equally to this research.}
-\email{trovato@corporation.com}
-\orcid{1234-5678-9012}
-\author{G.K.M. Tobin}
-\authornotemark[1]
-\email{webmaster@marysville-ohio.com}
+\email{matthis.kruse@cispa.de}
+\orcid{0000-0003-4062-9666}
 \affiliation{%
-  \institution{Institute for Clarity in Documentation}
-  \streetaddress{P.O. Box 1212}
-  \city{Dublin}
-  \state{Ohio}
-  \country{USA}
-  \postcode{43017-6221}
+  \institution{CISPA Helmholtz Center for Information Security}
+  \streetaddress{Stuhlsatzenhaus 5}
+  \city{Saarbr{\"u}cken}
+  \state{Saarland}
+  \country{Germany}
+  \postcode{66123}
 }
 
-\author{Lars Th{\o}rv{\"a}ld}
+\author{Marco Patrignani}
+\orcid{0000-0003-3411-9678}
+\email{marco.patrignani@unitn.it}
 \affiliation{%
-  \institution{The Th{\o}rv{\"a}ld Group}
-  \streetaddress{1 Th{\o}rv{\"a}ld Circle}
-  \city{Hekla}
-  \country{Iceland}}
-\email{larst@affiliation.org}
-
-\author{Valerie B\'eranger}
-\affiliation{%
-  \institution{Inria Paris-Rocquencourt}
-  \city{Rocquencourt}
-  \country{France}
+  \institution{University of Trento}
+  \streetaddress{Via Sommarive, 9}
+  \city{Povo}
+  \country{Italy}
+  \postcode{38123}
 }
-
-\author{Aparna Patel}
-\affiliation{%
- \institution{Rajiv Gandhi University}
- \streetaddress{Rono-Hills}
- \city{Doimukh}
- \state{Arunachal Pradesh}
- \country{India}}
-
-\author{Huifen Chan}
-\affiliation{%
-  \institution{Tsinghua University}
-  \streetaddress{30 Shuangqing Rd}
-  \city{Haidian Qu}
-  \state{Beijing Shi}
-  \country{China}}
-
-\author{Charles Palmer}
-\affiliation{%
-  \institution{Palmer Research Laboratories}
-  \streetaddress{8600 Datapoint Drive}
-  \city{San Antonio}
-  \state{Texas}
-  \country{USA}
-  \postcode{78229}}
-\email{cpalmer@prl.com}
-
-\author{John Smith}
-\affiliation{%
-  \institution{The Th{\o}rv{\"a}ld Group}
-  \streetaddress{1 Th{\o}rv{\"a}ld Circle}
-  \city{Hekla}
-  \country{Iceland}}
-\email{jsmith@affiliation.org}
-
-\author{Julius P. Kumquat}
-\affiliation{%
-  \institution{The Kumquat Consortium}
-  \city{New York}
-  \country{USA}}
-\email{jpkumquat@consortium.net}
 
@@
 %% of authors' names for this purpose.
-\renewcommand{\shortauthors}{Trovato and Tobin, et al.}
+\renewcommand{\shortauthors}{Kruse and Patrignani}
 
@@
 \begin{abstract}
-  A clear and well-documented \LaTeX\ document is presented as an
-  article formatted for publication by ACM in a conference proceedings
-  or journal publication. Based on the ``acmart'' document class, this
-  article presents and explains many of the common variations, as well
-  as many of the formatting elements an author may use in the
-  preparation of the documentation of their work.
+% To this date, no practical compiler implementation is proven as robustly safe.
+% This is no surprise, since proving compilers robustly safe is a huge endeavor.
+% But, compilers consist of several different passes, which themselves can be seen as compilers that translate from one intermediate representation to another.
+%
+% This paper investigates the compositionality of robustly safe compilers.
+% Composing two compilers means to simply plug the output of one as an input to the other.
+% We show that it suffices to prove individual passes as robustly safe.
+% We demonstrate that one can simply prove robust memory safety preservation for one pass and prove the robust preservation of cryptographic constant time for another pass.
+%
+%  Robustly Safe Compilers ensure that programs that do not go wrong when linked with an arbitrary source-level attacker also do not go wrong in their compiled counterpart when linked with untrusted target-level code.
+  Robustly safe programs do not go wrong when linked with arbitrary, untrusted code.
+  Compilers preserving such robust safety do this on a theoretical level.
+  The criterion requires challenging proofs which, so far, do not exist in real-world compiler implementations.
+  Moreover, these implementations are modular:
+  Compilers consist of several different so-called passes, which translate from one intermediate representation to another.
+  It is an open question whether one can prove these individual passes as robust safety preserving or not.
+
+  This paper answers this question positively:
+  The output of one robust safety preserving compiler may serve as input to another and this composition of both compilers remains robustly safe preserving.
+  We generalise this compositional behaviour for compilers that robustly preserve different security properties.
+  Thus, formal compiler engineers do not need to prove that all passes robustly preserve all security properties of interest, because their composition yields the intersection of the security properties.
+  A consequence of this is that even in the context of the phase ordering problem, whenever the optimisations are proven robustly safe, we can freely swap the optimisation passes without compromising security.
+  We validate our meta-level findings by means of a case study that looks at three different robust safety preserving compilers.
+
+  Lastly, parts of our results are verified in Coq.
+  The source code will be made publicly available upon this paper's acceptance.
 \end{abstract}
@@
 %% page.
-\begin{teaserfigure}
-  \includegraphics[width=\textwidth]{sampleteaser}
-  \caption{Seattle Mariners at Spring Training, 2010.}
-  \Description{Enjoying the baseball game from the third-base
-  seats. Ichiro Suzuki preparing to bat.}
-  \label{fig:teaser}
-\end{teaserfigure}
+%\begin{teaserfigure}
+%  \includegraphics[width=\textwidth]{sampleteaser}
+%  \caption{Seattle Mariners at Spring Training, 2010.}
+%  \Description{Enjoying the baseball game from the third-base
+%  seats. Ichiro Suzuki preparing to bat.}
+%  \label{fig:teaser}
+%\end{teaserfigure}
 
-\received{20 February 2007}
-\received[revised]{12 March 2009}
-\received[accepted]{5 June 2009}
+%\received{20 February 2007}
+%\received[revised]{12 March 2009}
+%\received[accepted]{5 June 2009}
 
@@
   \centering
-  \includegraphics[width=\linewidth]{sample-franklin}
+%  \includegraphics[width=\linewidth]{sample-franklin}
   \caption{1907 Franklin Model D roadster. Photograph by Harris \&

### 2023-06-06 popl/main.tex
Code example

@@
 \documentclass[acmsmall,review,screen]{acmart}
+
+\usepackage{listings}
+
 %% NOTE that a single column version may be required for 
@@
 %  Robustly Safe Compilers ensure that programs that do not go wrong when linked with an arbitrary source-level attacker also do not go wrong in their compiled counterpart when linked with untrusted target-level code.
-  Robustly safe programs do not go wrong when linked with arbitrary, untrusted code.
-  Compilers preserving such robust safety do this on a theoretical level.
-  The criterion requires challenging proofs which, so far, do not exist in real-world compiler implementations.
+  Programs that do not go wrong should also not go wrong when linked with arbitrary, untrusted code.
+  Robustly safe compilers ensure that, given such program at source level, this also holds at target level.
+  The robust safety criterion requires challenging proofs which, so far, do not exist in real-world compiler implementations.
   Moreover, these implementations are modular:
   Compilers consist of several different so-called passes, which translate from one intermediate representation to another.
-  It is an open question whether one can prove these individual passes as robust safety preserving or not.
+  It is an open question whether one can split the proof of robust safety modularily as well.
 
@@
   Lastly, parts of our results are verified in Coq.
-  The source code will be made publicly available upon this paper's acceptance.
 \end{abstract}
@@
   <concept>
-  <concept_id>10010520.10010553.10010562</concept_id>
-  <concept_desc>Computer systems organization~Embedded systems</concept_desc>
+  <concept_id>10002978.10002986.10002989</concept_id>
+  <concept_desc>Security and privacy~Formal security models</concept_desc>
   <concept_significance>500</concept_significance>
- </concept>
- <concept>
-  <concept_id>10010520.10010575.10010755</concept_id>
-  <concept_desc>Computer systems organization~Redundancy</concept_desc>
-  <concept_significance>300</concept_significance>
- </concept>
- <concept>
-  <concept_id>10010520.10010553.10010554</concept_id>
-  <concept_desc>Computer systems organization~Robotics</concept_desc>
-  <concept_significance>100</concept_significance>
- </concept>
- <concept>
-  <concept_id>10003033.10003083.10003095</concept_id>
-  <concept_desc>Networks~Network reliability</concept_desc>
-  <concept_significance>100</concept_significance>
   </concept>
@@
 \end{CCSXML}
-
-\ccsdesc[500]{Computer systems organization~Embedded systems}
-\ccsdesc[300]{Computer systems organization~Redundancy}
-\ccsdesc{Computer systems organization~Robotics}
-\ccsdesc[100]{Networks~Network reliability}
+\ccsdesc[500]{Security and privacy~Formal security models}
 
@@
 %% the work being presented. Separate the keywords with commas.
-\keywords{datasets, neural networks, gaze detection, text tagging}
+\keywords{Memory-safety, Secure Compilation, Privacy}
 
@@
 \section{Introduction}
-ACM's consolidated article template, introduced in 2017, provides a
-consistent \LaTeX\ style for use across ACM publications, and
-incorporates accessibility and metadata-extraction functionality
-necessary for future Digital Library endeavors. Numerous ACM and
-SIG-specific \LaTeX\ templates have been examined, and their unique
-features incorporated into this single new template.
 
-If you are new to publishing with ACM, this document is a valuable
-guide to the process of preparing your work for publication. If you
-have published with ACM before, this document provides insight and
-instruction into more recent changes to the article template.
+\begin{lstlisting}[language=c,caption=``Wrong bounds check of two 64-bit integers.'']
+if at <= bounds {
+  x[at] = 42;
+}
+\end{lstlisting}
+Note the subtle bug of using \verb|<=| instead of \verb|<|, leading to an out-of-bounds access whenever \verb|at = bounds|.
+Furthermore, when \verb|at| and \verb|bounds| are 64-bit integers on a 32-bit architecture, the comparison may not be performed in constant-time: The compiler may bail out as soon as the lower 32 bits are unequal, not bothering to compare the higher 32 bits.
 
-The ``\verb|acmart|'' document class can be used to prepare articles
-for any ACM publication --- conference or journal, and for any stage
-of publication, from review to final ``camera-ready'' copy, to the
-author's own version, with {\itshape very} few changes to the source.
+\section{Compositionality}
 
-\section{Template Overview}
-As noted in the introduction, the ``\verb|acmart|'' document class can
-be used to prepare many different kinds of documentation --- a
-double-blind initial submission of a full-length technical paper, a
-two-page SIGGRAPH Emerging Technologies abstract, a ``camera-ready''
-journal article, a SIGCHI Extended Abstract, and more --- all by
-selecting the appropriate {\itshape template style} and {\itshape
-  template parameters}.
+\section{Robust MS Preservation}
 
-This document will explain the major features of the document
-class. For further information, the {\itshape \LaTeX\ User's Guide} is
-available from
-\url{https://www.acm.org/publications/proceedings-template}.
+\section{Robust MS+CCT Preservation}
 
-\subsection{Template Styles}
+\section{Related Work}
 
-The primary parameter given to the ``\verb|acmart|'' document class is
-the {\itshape template style} which corresponds to the kind of publication
-or SIG publishing the work. This parameter is enclosed in square
-brackets and is a part of the {\verb|documentclass|} command:
-\begin{verbatim}
-  \documentclass[STYLE]{acmart}
-\end{verbatim}
+\section{Conclusion}
 
-Journals use one of three template styles. All but three ACM journals
-use the {\verb|acmsmall|} template style:
-\begin{itemize}
-\item {\verb|acmsmall|}: The default journal template style.
-\item {\verb|acmlarge|}: Used by JOCCH and TAP.
-\item {\verb|acmtog|}: Used by TOG.
-\end{itemize}
-
-The majority of conference proceedings documentation will use the {\verb|acmconf|} template style.
-\begin{itemize}
-\item {\verb|acmconf|}: The default proceedings template style.
-\item{\verb|sigchi|}: Used for SIGCHI conference articles.
-\item{\verb|sigchi-a|}: Used for SIGCHI ``Extended Abstract'' articles.
-\item{\verb|sigplan|}: Used for SIGPLAN conference articles.
-\end{itemize}
-
-\subsection{Template Parameters}
-
-In addition to specifying the {\itshape template style} to be used in
-formatting your work, there are a number of {\itshape template parameters}
-which modify some part of the applied template style. A complete list
-of these parameters can be found in the {\itshape \LaTeX\ User's Guide.}
-
-Frequently-used parameters, or combinations of parameters, include:
-\begin{itemize}
-\item {\verb|anonymous,review|}: Suitable for a ``double-blind''
-  conference submission. Anonymizes the work and includes line
-  numbers. Use with the \verb|\acmSubmissionID| command to print the
-  submission's unique ID on each page of the work.
-\item{\verb|authorversion|}: Produces a version of the work suitable
-  for posting by the author.
-\item{\verb|screen|}: Produces colored hyperlinks.
-\end{itemize}
-
-This document uses the following string as the first command in the
-source file:
-\begin{verbatim}
-\documentclass[sigconf]{acmart}
-\end{verbatim}
-
-\section{Modifications}
-
-Modifying the template --- including but not limited to: adjusting
-margins, typeface sizes, line spacing, paragraph and list definitions,
-and the use of the \verb|\vspace| command to manually adjust the
-vertical spacing between elements of your work --- is not allowed.
-
-{\bfseries Your document will be returned to you for revision if
-  modifications are discovered.}
-
-\section{Typefaces}
-
-The ``\verb|acmart|'' document class requires the use of the
-``Libertine'' typeface family. Your \TeX\ installation should include
-this set of packages. Please do not substitute other typefaces. The
-``\verb|lmodern|'' and ``\verb|ltimes|'' packages should not be used,
-as they will override the built-in typeface families.
-
-\section{Title Information}
-
-The title of your work should use capital letters appropriately -
-\url{https://capitalizemytitle.com/} has useful rules for
-capitalization. Use the {\verb|title|} command to define the title of
-your work. If your work has a subtitle, define it with the
-{\verb|subtitle|} command.  Do not insert line breaks in your title.
-
-If your title is lengthy, you must define a short version to be used
-in the page headers, to prevent overlapping text. The \verb|title|
-command has a ``short title'' parameter:
-\begin{verbatim}
-  \title[short title]{full title}
-\end{verbatim}
-
-\section{Authors and Affiliations}
-
-Each author must be defined separately for accurate metadata
-identification. Multiple authors may share one affiliation. Authors'
-names should not be abbreviated; use full first names wherever
-possible. Include authors' e-mail addresses whenever possible.
-
-Grouping authors' names or e-mail addresses, or providing an ``e-mail
-alias,'' as shown below, is not acceptable:
-\begin{verbatim}
-  \author{Brooke Aster, David Mehldau}
-  \email{dave,judy,steve@university.edu}
-  \email{firstname.lastname@phillips.org}
-\end{verbatim}
-
-The \verb|authornote| and \verb|authornotemark| commands allow a note
-to apply to multiple authors --- for example, if the first two authors
-of an article contributed equally to the work.
-
-If your author list is lengthy, you must define a shortened version of
-the list of authors to be used in the page headers, to prevent
-overlapping text. The following command should be placed just after
-the last \verb|\author{}| definition:
-\begin{verbatim}
-  \renewcommand{\shortauthors}{McCartney, et al.}
-\end{verbatim}
-Omitting this command will force the use of a concatenated list of all
-of the authors' names, which may result in overlapping text in the
-page headers.
-
-The article template's documentation, available at
-\url{https://www.acm.org/publications/proceedings-template}, has a
-complete explanation of these commands and tips for their effective
-use.
-
-Note that authors' addresses are mandatory for journal articles.
-
-\section{Rights Information}
-
-Authors of any work published by ACM will need to complete a rights
-form. Depending on the kind of work, and the rights management choice
-made by the author, this may be copyright transfer, permission,
-license, or an OA (open access) agreement.
-
-Regardless of the rights management choice, the author will receive a
-copy of the completed rights form once it has been submitted. This
-form contains \LaTeX\ commands that must be copied into the source
-document. When the document source is compiled, these commands and
-their parameters add formatted text to several areas of the final
-document:
-\begin{itemize}
-\item the ``ACM Reference Format'' text on the first page.
-\item the ``rights management'' text on the first page.
-\item the conference information in the page header(s).
-\end{itemize}
-
-Rights information is unique to the work; if you are preparing several
-works for an event, make sure to use the correct set of commands with
-each of the works.
-
-The ACM Reference Format text is required for all articles over one
-page in length, and is optional for one-page articles (abstracts).
-
-\section{CCS Concepts and User-Defined Keywords}
-
-Two elements of the ``acmart'' document class provide powerful
-taxonomic tools for you to help readers find your work in an online
-search.
-
-The ACM Computing Classification System ---
-\url{https://www.acm.org/publications/class-2012} --- is a set of
-classifiers and concepts that describe the computing
-discipline. Authors can select entries from this classification
-system, via \url{https://dl.acm.org/ccs/ccs.cfm}, and generate the
-commands to be included in the \LaTeX\ source.
-
-User-defined keywords are a comma-separated list of words and phrases
-of the authors' choosing, providing a more flexible way of describing
-the research being presented.
-
-CCS concepts and user-defined keywords are required for for all
-articles over two pages in length, and are optional for one- and
-two-page articles (or abstracts).
-
-\section{Sectioning Commands}
-
-Your work should use standard \LaTeX\ sectioning commands:
-\verb|section|, \verb|subsection|, \verb|subsubsection|, and
-\verb|paragraph|. They should be numbered; do not remove the numbering
-from the commands.
-
-Simulating a sectioning command by setting the first word or words of
-a paragraph in boldface or italicized text is {\bfseries not allowed.}
-
-\section{Tables}
-
-The ``\verb|acmart|'' document class includes the ``\verb|booktabs|''
-package --- \url{https://ctan.org/pkg/booktabs} --- for preparing
-high-quality tables.
-
-Table captions are placed {\itshape above} the table.
-
-Because tables cannot be split across pages, the best placement for
-them is typically the top of the page nearest their initial cite.  To
-ensure this proper ``floating'' placement of tables, use the
-environment \textbf{table} to enclose the table's contents and the
-table caption.  The contents of the table itself must go in the
-\textbf{tabular} environment, to be aligned properly in rows and
-columns, with the desired horizontal and vertical rules.  Again,
-detailed instructions on \textbf{tabular} material are found in the
-\textit{\LaTeX\ User's Guide}.
-
-Immediately following this sentence is the point at which
-Table~\ref{tab:freq} is included in the input file; compare the
-placement of the table here with the table in the printed output of
-this document.
-
-\begin{table}
-  \caption{Frequency of Special Characters}
-  \label{tab:freq}
-  \begin{tabular}{ccl}
-    \toprule
-    Non-English or Math&Frequency&Comments\\
-    \midrule
-    \O & 1 in 1,000& For Swedish names\\
-    $\pi$ & 1 in 5& Common in math\\
-    \$ & 4 in 5 & Used in business\\
-    $\Psi^2_1$ & 1 in 40,000& Unexplained usage\\
-  \bottomrule
-\end{tabular}
-\end{table}
-
-To set a wider table, which takes up the whole width of the page's
-live area, use the environment \textbf{table*} to enclose the table's
-contents and the table caption.  As with a single-column table, this
-wide table will ``float'' to a location deemed more
-desirable. Immediately following this sentence is the point at which
-Table~\ref{tab:commands} is included in the input file; again, it is
-instructive to compare the placement of the table here with the table
-in the printed output of this document.
-
-\begin{table*}
-  \caption{Some Typical Commands}
-  \label{tab:commands}
-  \begin{tabular}{ccl}
-    \toprule
-    Command &A Number & Comments\\
-    \midrule
-    \texttt{{\char'134}author} & 100& Author \\
-    \texttt{{\char'134}table}& 300 & For tables\\
-    \texttt{{\char'134}table*}& 400& For wider tables\\
-    \bottomrule
-  \end{tabular}
-\end{table*}
-
-Always use midrule to separate table header rows from data rows, and
-use it only for this purpose. This enables assistive technologies to
-recognise table headers and support their users in navigating tables
-more easily.
-
-\section{Math Equations}
-You may want to display math equations in three distinct styles:
-inline, numbered or non-numbered display.  Each of the three are
-discussed in the next sections.
-
-\subsection{Inline (In-text) Equations}
-A formula that appears in the running text is called an inline or
-in-text formula.  It is produced by the \textbf{math} environment,
-which can be invoked with the usual
-\texttt{{\char'134}begin\,\ldots{\char'134}end} construction or with
-the short form \texttt{\$\,\ldots\$}. You can use any of the symbols
-and structures, from $\alpha$ to $\omega$, available in
-\LaTeX~\cite{Lamport:LaTeX}; this section will simply show a few
-examples of in-text equations in context. Notice how this equation:
-\begin{math}
-  \lim_{n\rightarrow \infty}x=0
-\end{math},
-set here in in-line math style, looks slightly different when
-set in display style.  (See next section).
-
-\subsection{Display Equations}
-A numbered display equation---one set off by vertical space from the
-text and centered horizontally---is produced by the \textbf{equation}
-environment. An unnumbered display equation is produced by the
-\textbf{displaymath} environment.
-
-Again, in either environment, you can use any of the symbols and
-structures available in \LaTeX\@; this section will just give a couple
-of examples of display equations in context.  First, consider the
-equation, shown as an inline equation above:
-\begin{equation}
-  \lim_{n\rightarrow \infty}x=0
-\end{equation}
-Notice how it is formatted somewhat differently in
-the \textbf{displaymath}
-environment.  Now, we'll enter an unnumbered equation:
-\begin{displaymath}
-  \sum_{i=0}^{\infty} x + 1
-\end{displaymath}
-and follow it with another numbered equation:
-\begin{equation}
-  \sum_{i=0}^{\infty}x_i=\int_{0}^{\pi+2} f
-\end{equation}
-just to demonstrate \LaTeX's able handling of numbering.
-
-\section{Figures}
-
-The ``\verb|figure|'' environment should be used for figures. One or
-more images can be placed within a figure. If your figure contains
-third-party material, you must clearly identify it as such, as shown
-in the example below.
-\begin{figure}[h]
-  \centering
-%  \includegraphics[width=\linewidth]{sample-franklin}
-  \caption{1907 Franklin Model D roadster. Photograph by Harris \&
-    Ewing, Inc. [Public domain], via Wikimedia
-    Commons. (\url{https://goo.gl/VLCRBB}).}
-  \Description{A woman and a girl in white dresses sit in an open car.}
-\end{figure}
-
-Your figures should contain a caption which describes the figure to
-the reader.
-
-Figure captions are placed {\itshape below} the figure.
-
-Every figure should also have a figure description unless it is purely
-decorative. These descriptions convey what’s in the image to someone
-who cannot see it. They are also used by search engine crawlers for
-indexing images, and when images cannot be loaded.
-
-A figure description must be unformatted plain text less than 2000
-characters long (including spaces).  {\bfseries Figure descriptions
-  should not repeat the figure caption – their purpose is to capture
-  important information that is not already provided in the caption or
-  the main text of the paper.} For figures that convey important and
-complex new information, a short text description may not be
-adequate. More complex alternative descriptions can be placed in an
-appendix and referenced in a short figure description. For example,
-provide a data table capturing the information in a bar chart, or a
-structured list representing a graph.  For additional information
-regarding how best to write figure descriptions and why doing this is
-so important, please see
-\url{https://www.acm.org/publications/taps/describing-figures/}.
-
-\subsection{The ``Teaser Figure''}
-
-A ``teaser figure'' is an image, or set of images in one figure, that
-are placed after all author and affiliation information, and before
-the body of the article, spanning the page. If you wish to have such a
-figure in your article, place the command immediately before the
-\verb|\maketitle| command:
-\begin{verbatim}
-  \begin{teaserfigure}
-    \includegraphics[width=\textwidth]{sampleteaser}
-    \caption{figure caption}
-    \Description{figure description}
-  \end{teaserfigure}
-\end{verbatim}
-
-\section{Citations and Bibliographies}
-
-The use of \BibTeX\ for the preparation and formatting of one's
-references is strongly recommended. Authors' names should be complete
---- use full first names (``Donald E. Knuth'') not initials
-(``D. E. Knuth'') --- and the salient identifying features of a
-reference should be included: title, year, volume, number, pages,
-article DOI, etc.
-
-The bibliography is included in your source document with these two
-commands, placed just before the \verb|\end{document}| command:
-\begin{verbatim}
-  \bibliographystyle{ACM-Reference-Format}
-  \bibliography{bibfile}
-\end{verbatim}
-where ``\verb|bibfile|'' is the name, without the ``\verb|.bib|''
-suffix, of the \BibTeX\ file.
-
-Citations and references are numbered by default. A small number of
-ACM publications have citations and references formatted in the
-``author year'' style; for these exceptions, please include this
-command in the {\bfseries preamble} (before the command
-``\verb|\begin{document}|'') of your \LaTeX\ source:
-\begin{verbatim}
-  \citestyle{acmauthoryear}
-\end{verbatim}
-
-  Some examples.  A paginated journal article \cite{Abril07}, an
-  enumerated journal article \cite{Cohen07}, a reference to an entire
-  issue \cite{JCohen96}, a monograph (whole book) \cite{Kosiur01}, a
-  monograph/whole book in a series (see 2a in spec. document)
-  \cite{Harel79}, a divisible-book such as an anthology or compilation
-  \cite{Editor00} followed by the same example, however we only output
-  the series if the volume number is given \cite{Editor00a} (so
-  Editor00a's series should NOT be present since it has no vol. no.),
-  a chapter in a divisible book \cite{Spector90}, a chapter in a
-  divisible book in a series \cite{Douglass98}, a multi-volume work as
-  book \cite{Knuth97}, a couple of articles in a proceedings (of a
-  conference, symposium, workshop for example) (paginated proceedings
-  article) \cite{Andler79, Hagerup1993}, a proceedings article with
-  all possible elements \cite{Smith10}, an example of an enumerated
-  proceedings article \cite{VanGundy07}, an informally published work
-  \cite{Harel78}, a couple of preprints \cite{Bornmann2019,
-    AnzarootPBM14}, a doctoral dissertation \cite{Clarkson85}, a
-  master's thesis: \cite{anisi03}, an online document / world wide web
-  resource \cite{Thornburg01, Ablamowicz07, Poker06}, a video game
-  (Case 1) \cite{Obama08} and (Case 2) \cite{Novak03} and \cite{Lee05}
-  and (Case 3) a patent \cite{JoeScientist001}, work accepted for
-  publication \cite{rous08}, 'YYYYb'-test for prolific author
-  \cite{SaeediMEJ10} and \cite{SaeediJETC10}. Other cites might
-  contain 'duplicate' DOI and URLs (some SIAM articles)
-  \cite{Kirschmer:2010:AEI:1958016.1958018}. Boris / Barbara Beeton:
-  multi-volume works as books \cite{MR781536} and \cite{MR781537}. A
-  couple of citations with DOIs:
-  \cite{2004:ITE:1009386.1010128,Kirschmer:2010:AEI:1958016.1958018}. Online
-  citations: \cite{TUGInstmem, Thornburg01, CTANacmart}. Artifacts:
-  \cite{R} and \cite{UMassCitations}.
-
-\section{Acknowledgments}
-
-Identification of funding sources and other support, and thanks to
-individuals and groups that assisted in the research and the
-preparation of the work should be included in an acknowledgment
-section, which is placed just before the reference section in your
-document.
-
-This section has a special environment:
-\begin{verbatim}
-  \begin{acks}
-  ...
-  \end{acks}
-\end{verbatim}
-so that the information contained therein can be more easily collected
-during the article metadata extraction phase, and to ensure
-consistency in the spelling of the section heading.
-
-Authors should not prepare this section as a numbered or unnumbered {\verb|\section|}; please use the ``{\verb|acks|}'' environment.
-
-\section{Appendices}
-
-If your work needs an appendix, add it before the
-``\verb|\end{document}|'' command at the conclusion of your source
-document.
-
-Start the appendix with the ``\verb|appendix|'' command:
-\begin{verbatim}
-  \appendix
-\end{verbatim}
-and note that in the appendix, sections are lettered, not
-numbered. This document has two appendices, demonstrating the section
-and subsection identification method.
-
-\section{Multi-language papers}
-
-Papers may be written in languages other than English or include
-titles, subtitles, keywords and abstracts in different languages (as a
-rule, a paper in a language other than English should include an
-English title and an English abstract).  Use \verb|language=...| for
-every language used in the paper.  The last language indicated is the
-main language of the paper.  For example, a French paper with
-additional titles and abstracts in English and German may start with
-the following command
-\begin{verbatim}
-\documentclass[sigconf, language=english, language=german,
-               language=french]{acmart}
-\end{verbatim}
-
-The title, subtitle, keywords and abstract will be typeset in the main
-language of the paper.  The commands \verb|\translatedXXX|, \verb|XXX|
-begin title, subtitle and keywords, can be used to set these elements
-in the other languages.  The environment \verb|translatedabstract| is
-used to set the translation of the abstract.  These commands and
-environment have a mandatory first argument: the language of the
-second argument.  See \verb|sample-sigconf-i13n.tex| file for examples
-of their usage.
-
-\section{SIGCHI Extended Abstracts}
-
-The ``\verb|sigchi-a|'' template style (available only in \LaTeX\ and
-not in Word) produces a landscape-orientation formatted article, with
-a wide left margin. Three environments are available for use with the
-``\verb|sigchi-a|'' template style, and produce formatted output in
-the margin:
-\begin{itemize}
-\item {\verb|sidebar|}:  Place formatted text in the margin.
-\item {\verb|marginfigure|}: Place a figure in the margin.
-\item {\verb|margintable|}: Place a table in the margin.
-\end{itemize}
-
-%%
 %% The acknowledgments section is defined using the "acks" environment
@@
 
-\section{Research Methods}
-
-\subsection{Part One}
-
-Lorem ipsum dolor sit amet, consectetur adipiscing elit. Morbi
-malesuada, quam in pulvinar varius, metus nunc fermentum urna, id
-sollicitudin purus odio sit amet enim. Aliquam ullamcorper eu ipsum
-vel mollis. Curabitur quis dictum nisl. Phasellus vel semper risus, et
-lacinia dolor. Integer ultricies commodo sem nec semper.
-
-\subsection{Part Two}
-
-Etiam commodo feugiat nisl pulvinar pellentesque. Etiam auctor sodales
-ligula, non varius nibh pulvinar semper. Suspendisse nec lectus non
-ipsum convallis congue hendrerit vitae sapien. Donec at laoreet
-eros. Vivamus non purus placerat, scelerisque diam eu, cursus
-ante. Etiam aliquam tortor auctor efficitur mattis.
-
-\section{Online Resources}
-
-Nam id fermentum dui. Suspendisse sagittis tortor a nulla mollis, in
-pulvinar ex pretium. Sed interdum orci quis metus euismod, et sagittis
-enim maximus. Vestibulum gravida massa ut felis suscipit
-congue. Quisque mattis elit a risus ultrices commodo venenatis eget
-dui. Etiam sagittis eleifend elementum.
-
-Nam interdum magna at lectus dignissim, ac dignissim lorem
-rhoncus. Maecenas eu arcu ac neque placerat aliquam. Nunc pulvinar
-massa et mattis lacinia.
-
 \end{document}

### 2023-06-08 popl/main.tex
Draft Intro

@@
 %% The first command in your LaTeX source must be the \documentclass command.
-\documentclass[acmsmall,review,screen]{acmart}
+\documentclass[acmsmall,review,screen,dvipsnames]{acmart}
 
+\usepackage[colorinlistoftodos]{todonotes}
+\usepackage{cleveref}
+\usepackage{glossaries}
 \usepackage{listings}
+\usepackage{xspace}
+
+\input{cmds.tex}
+
+\loadglsentries{acronyms}
+\makeglossaries
 
@@
 \author{Matthis Kruse}
-\authornote{Both authors contributed equally to this research.}
+% \authornote{Both authors contributed equally to this research.}
 \email{matthis.kruse@cispa.de}
@@
   Lastly, parts of our results are verified in Coq.
+
+\begin{center}\small\it
+	{In the following, we use syntax highlighting accessible to both colourblind and black \& white readers~\citep{patrignani2020use}. Specifically, we use a $\src{\text{blue}}$,
+	$\src{sans\text{-}serif}$ font for $\src{L_{TMS}}$,
+	$\trg{bold}$ font for $\trg{L}$ and $\trg{L_{MS}}$, and $\obj{green}$ font for $\obj{L_{sCCT}}$.
+	}
+\end{center}
 \end{abstract}
@@
 
-%% A "teaser" image appears between the author and affiliation
-%% information and the body of the document, and typically spans the
-%% page.
-%\begin{teaserfigure}
-%  \includegraphics[width=\textwidth]{sampleteaser}
-%  \caption{Seattle Mariners at Spring Training, 2010.}
-%  \Description{Enjoying the baseball game from the third-base
-%  seats. Ichiro Suzuki preparing to bat.}
-%  \label{fig:teaser}
-%\end{teaserfigure}
-
-%\received{20 February 2007}
-%\received[revised]{12 March 2009}
-%\received[accepted]{5 June 2009}
-
 %%
@@
 
-\begin{lstlisting}[language=c,caption=``Wrong bounds check of two 64-bit integers.'']
-if at <= bounds {
-  x[at] = 42;
+\gls{cct} code does not leak information through timing analyses.
+%However, it is possible to write \gls{cct} code in a more high-level language like FaCT~\cite{cauligi2019fact}.
+It is folklore that programmers prefer to write code in higher-level languages such as \verb|C|.
+But, high-level languages do not run on common hardware, so they need to be translated by a compiler.
+Due to their optimizing nature, compilers can break security properties, such as \gls{cct}, when performing this translation.
+An example of this exists in literature~\cite{barthe2018sec}:
+
+\begin{minipage}{0.49\textwidth}
+  \begin{lstlisting}[language=c,caption={Before Compilation},numbers=left]
+// x = (b ? x : y)
+int mask = -b;
+x = (y & mask) | (x & ~mask);
+  \end{lstlisting}
+\end{minipage}%
+\begin{minipage}{0.49\textwidth}
+  \begin{lstlisting}[language=c,caption={After Compilation},numbers=left]
+if(b) {
+  x = y;
 }
   \end{lstlisting}
-Note the subtle bug of using \verb|<=| instead of \verb|<|, leading to an out-of-bounds access whenever \verb|at = bounds|.
-Furthermore, when \verb|at| and \verb|bounds| are 64-bit integers on a 32-bit architecture, the comparison may not be performed in constant-time: The compiler may bail out as soon as the lower 32 bits are unequal, not bothering to compare the higher 32 bits.
+\end{minipage}
 
-\section{Compositionality}
+\noindent
+Here, the compiler recognizes the pattern in the bit hacks and rewrites them to code branching on \verb|b|, violating \gls{cct}.
+Because of this, typically, \gls{cct} code is written in assembly.
+Thanks to DSLs like FaCT~\cite{cauligi2019fact}, however, this is not necessary.
 
-\section{Robust MS Preservation}
+But, FaCT's compiler gives no guarantees with regards to robustness.
+That is, the code may not be \gls{cct} when linked with untrusted third-party libraries.
+%For example, a simple table lookup may be fetched from one of many cache layers.
+%The higher up layers lead to slower access times which is detectable from just observing the program's %execution time.
+%With that, an attacker can identify the entry in the table that we wanted to access.
+%A \gls{cct} table lookup has to go through all elements of that table in the same order.
+%
+Compilers giving such a guarantee are known to be robustly preserving~\cite{abate2019jour}.
+These \emph{secure compilers} give a strong formal guarantee for security, since an arbitrary target-level attacker has no chance of violating our property of interest, such as \gls{cct}.\MK{cite CCT compiler that is robustly preserving}
 
-\section{Robust MS+CCT Preservation}
+For code to be \gls{cct}, \gls{ms} is a prerequisite, since information is easily leaked in a memory-unsafe world by just reading out of bounds.
+The importance of \gls{ms} has given rise to lots of possible countermeasures.
+Particularily interesting for this paper are \emph{compiler-based} countermeasures like CETS~\cite{nagarakatte2010cets}, SoftBound~\cite{nagarakatte2009soft}, or MSWasm~\cite{michael2023mswasm}, which describes a \emph{robust enforcement} of \gls{ms}.
+Given \gls{cct} requires \gls{ms}, it is natural to combine compiler-based techniques to ensure both \gls{cct} and \gls{ms} hold.
+Unfortunately, the composition of secure compilers is unclear.
 
-\section{Related Work}
+This paper develops the theory of secure compiler compositions and introduces a secure compilation chain for a combination of \gls{ms} and \gls{cct}.
+We do this via 4 contributions:
 
-\section{Conclusion}
+\contribution{\contribnum Composition of Properties (\Cref{sec:compprop})}
+We present a formal framework to reason about compositions of safety properties.
+First, we assume there is a most general trace semantics in which a property can be defined and that there is a monitor for that property.
+The monitor is defined on a restricted trace semantics without unneccessary information.
+The composition of trace properties is then modelled by the composition of monitors.
+The two-level approach, i.e., the two different trace semantics allow for a modular extension without touching any already existing definition.
+
+\contribution{\contribnum Composition of Secure Compilers (\Cref{sec:compcomp})}
+We introduce a formal framework for the composition of secure compilers.
+Composition of secure compilers allows for modular proof engineering, since our framework can be used even if the compilers robustly preserve different properties.
+A simple corollary of our main composition theorem is the fact that the composition of secure compilers with equal source and target programming lanugages can be freely swapped without compromising security.
+In turn, formal compiler engineers can optimize the compilation tool chain with regards to the phase ordering problem.
+
+\contribution{\contribnum Compiler Robustly Preserving \gls{ms} and \gls{cct} (\Cref{sec:casestud})}
+We introduce a case-study consisting of three secure compilers.
+To this end, we present the language $\src{L_{TMS}}$ which has a type system with linear constraints to enforce \gls{tms} statically.
+%We prove that any well-typed $\src{L_{TMS}}$ program is temporal memory safe.
+Then, we introduce $\trg{L}$, a language without a type system, and a compiler $\llbracket\src{\bullet}\rrbracket^{\src{L_{TMS}}\to\trg{L}}$ from $\src{L_{TMS}}$ to $\trg{L}$ that robustly preserves \gls{tms}.
+To get full \gls{ms}, we define another compiler $\llbracket\trg{\bullet}\rrbracket^{\trg{L}\to\trg{L_{MS}}}$ which introduces bounds-checks and compose it with $\llbracket\src{\bullet}\rrbracket^{\src{L_{TMS}}\to\trg{L}}$.
+Lastly, we define $\obj{L_{sCCT}}$, a language with access to a \emph{model specific register} to enable \gls{cct} execution, and a compiler $\llbracket\irl{\bullet}\rrbracket^{\trg{L_{MS}}\to\obj{L_{sCCT}}}$ which we compose with the other ones.
+Finally, this gives us a three-step compilation chain with both \gls{ms} and \gls{cct}.
+
+\contribution{\contribnum Coq Formalisation}
+Parts of our results are formalized in the Coq proof assistant.
+We use the symbol \CoqSymbol to demonstrate the fact that a definition or a proof is done in Coq.
+Since our semantics and compiler definitions are extractable from Coq to Ocaml, we can run them.
+
+We now describe the other sections in the paper.
+\Cref{sec:background} establishes the notions of traces, (hyper-)properties, and robust preservation, while \Cref{sec:relwork} compares this work with others.
+We conclude in \Cref{sec:concl} and give an outlook for future work.
+
+\contribution{Open Source \& Technical Report} Our technical report and the Coq formalisation are
+available as supplementary material and will be made available upon acceptance of this paper.
+
+\section{Background}\label{sec:background}
+\section{Composing Properties}\label{sec:compprop}
+\section{Composing Robustly Secure Compilers}\label{sec:compcomp}
+\section{Case Study: Memory Safe and Cryptographic Constant Time Code}\label{sec:casestud}
+
+%\begin{lstlisting}[language=c,caption=``Wrong bounds check of two 64-bit integers.'']
+%if at <= bounds {
+%  x[at] = 42;
+%}
+%\end{lstlisting}
+%Note the subtle bug of using \verb|<=| instead of \verb|<|, leading to an out-of-bounds access whenever \verb|at = bounds|.
+%Furthermore, when \verb|at| and \verb|bounds| are 64-bit integers on a 32-bit architecture, the comparison may not be performed in constant-time: The compiler may bail out as soon as the lower 32 bits are unequal, not bothering to compare the higher 32 bits.
+
+\section{Related Work}\label{sec:relwork}
+\section{Conclusion}\label{sec:concl}
 
@@
 \begin{acks}
-To Robert, for the bagels and explaining CMYK and color spaces.
 \end{acks}
@@
 \bibliographystyle{ACM-Reference-Format}
-\bibliography{sample-base}
+\bibliography{main}
 

### 2023-06-09 popl/main.tex
Add overview

--- 2023-06-06 popl/main.tex
@@
 \begin{abstract}
-% To this date, no practical compiler implementation is proven as robustly safe.
-% This is no surprise, since proving compilers robustly safe is a huge endeavor.
-% But, compilers consist of several different passes, which themselves can be seen as compilers that translate from one intermediate representation to another.
-%
-% This paper investigates the compositionality of robustly safe compilers.
-% Composing two compilers means to simply plug the output of one as an input to the other.
-% We show that it suffices to prove individual passes as robustly safe.
-% We demonstrate that one can simply prove robust memory safety preservation for one pass and prove the robust preservation of cryptographic constant time for another pass.
-%
-%  Robustly Safe Compilers ensure that programs that do not go wrong when linked with an arbitrary source-level attacker also do not go wrong in their compiled counterpart when linked with untrusted target-level code.
-  Programs that do not go wrong should also not go wrong when linked with arbitrary, untrusted code.
-  Robustly safe compilers ensure that, given such program at source level, this also holds at target level.
-  The robust safety criterion requires challenging proofs which, so far, do not exist in real-world compiler implementations.
-  Moreover, these implementations are modular:
-  Compilers consist of several different so-called passes, which translate from one intermediate representation to another.
-  It is an open question whether one can split the proof of robust safety modularily as well.
+  \hltt{abstract}
 
-  This paper answers this question positively:
-  The output of one robust safety preserving compiler may serve as input to another and this composition of both compilers remains robustly safe preserving.
-  We generalise this compositional behaviour for compilers that robustly preserve different security properties.
-  Thus, formal compiler engineers do not need to prove that all passes robustly preserve all security properties of interest, because their composition yields the intersection of the security properties.
-  A consequence of this is that even in the context of the phase ordering problem, whenever the optimisations are proven robustly safe, we can freely swap the optimisation passes without compromising security.
-  We validate our meta-level findings by means of a case study that looks at three different robust safety preserving compilers.
-
-  Lastly, parts of our results are verified in Coq.
+\begin{center}\small\it
+	{In the following, we use syntax highlighting accessible to both colourblind and black \& white readers~\citep{patrignani2020use}. Specifically, we use a $\src{\text{blue}}$,
+	$\src{sans\text{-}serif}$ font for $\src{L_{TMS}}$,
+	$\trg{bold}$ font for $\trg{L}$ and $\trg{L_{MS}}$, and $\obj{green}$ font for $\obj{L_{sCCT}}$.
+	}
+\end{center}
 \end{abstract}
@@
 
-\begin{lstlisting}[language=c,caption=``Wrong bounds check of two 64-bit integers.'']
-if at <= bounds {
-  x[at] = 42;
-}
-\end{lstlisting}
-Note the subtle bug of using \verb|<=| instead of \verb|<|, leading to an out-of-bounds access whenever \verb|at = bounds|.
-Furthermore, when \verb|at| and \verb|bounds| are 64-bit integers on a 32-bit architecture, the comparison may not be performed in constant-time: The compiler may bail out as soon as the lower 32 bits are unequal, not bothering to compare the higher 32 bits.
+\hltt{Context}
+\hltt{Problem}
 
-\section{Compositionality}
+\BrText{Overview}{This paper introduces a formal framework for secure compiler compositions and showcases the framework's power by instantiating it on a multi-level compilation chain.
+To this end, we discuss security property compositions and define several secure compilers, where each is either robustly preserving a different security property of interest or optimizing, i.e., applying \gls{cf} or doing \gls{dce}.
+We compose these secure compilers into a multi-level compilation chain which is optimizing and robustly preserves a combination of \gls{ms} and \gls{cct}.
+Crucially, the robust preservation proofs are only done for the individual compilers in the compilation chain and the robust preservation for the whole chain is a result of one of our key theorems.
+The key results are formalized in Coq.}
 
-\section{Robust MS Preservation}
+In summary, this paper makes the following contributions:
 
-\section{Robust MS+CCT Preservation}
+\begin{itemize}
+  \item
+        \Cref{sec:compprop} presents a formal framework to reason about compositions of security properties and demonstrate its usage with concrete examples:
+        Given a most general trace model, we define the trace-based security properties \gls{tms}, \gls{sms}, \gls{ms}, \gls{cct}, and their compositions.
+        For each of these properties, we also give a monitor which runs on another trace model that excludes unneccessary information and show that their composition is behaviorally similar to the composition of the trace-based security properties.
+        This two-level approach enhances modularity: When working with the security properties, one only needs to reason about the corresponding monitor.
 
-\section{Related Work}
+  \item
+        \Cref{sec:compcomp} studies different forms of secure compiler compositions.
+        Concretely, it investigates upper and lower compositions, as well as the sequential composition.
+        A simple corollary of sequential composition is the ability to freely reorder compiler optimizations without compromising security.
 
-\section{Conclusion}
+  \item \Cref{sec:casestud} sets up a case-study to demonstrate the power of our formal framework.
+        We define several different languages and compilers between them and prove that the compilers are robustly preserving different properties.
+        The sequential composition of these compilers yields a compilation chain that is optimizing and robustly preserving a combination of \gls{ms} and \gls{cct}.
+        As for optimizations, the chain has two passes, one performing \gls{cf} and the other \gls{dce}.
+        This compilation chain demonstrates the ability to perform modular proof engineering by leaveraging this paper's compositionality theorems.
+
+  \item Our key contributions are formalized in the Coq proof assistant.
+        The formalization for the compilers are \emph{extractable}.
+        That is, one can translate our Coq code to Ocaml and run them.
+        Finally, in this paper, whenever definitions or proofs are done in Coq, we use the symbol \CoqSymbol.
+\end{itemize}
+
+We now describe the other sections of the paper.
+In \Cref{sec:background}, the paper establishes the notions of traces, (hyper-)properties, and robust preservation, while \Cref{sec:relwork} compares this work with others.
+We conclude in \Cref{sec:concl} and give an outlook for future work.
+
+\contribution{Open Source \& Technical Report} Our technical report and the Coq formalisation are
+available as supplementary material and will be made available upon acceptance of this paper.
+
+\section{Background}\label{sec:background}
+We give a brief introduction to trace properties, monitors, and property satisfaction (\Cref{subsec:bg:tprop}).
+With property satisfaction, we re-introduce the concept of robust satisfaction and its compiler-driven preservation (\Cref{subsec:bg:rtp}).
+
+\subsection{Trace-Based Properties}\label{subsec:bg:tprop}
+\subsection{Robust Trace-Property Preservation}\label{subsec:bg:rtp}
+
+\section{Composing Properties}\label{sec:compprop}
+\section{Composing Robustly Secure Compilers}\label{sec:compcomp}
+This section presents three different kinds of secure compiler compositions and their corresponding proofs.
+
+\section{Case Study: Memory Safe and Cryptographic Constant Time Code}\label{sec:casestud}
+This section defines several secure compilers, each of which robustly preserves a different property of interest.
+We demonstrate the power of our framework from \Cref{sec:compprop,sec:compcomp} by composing these compilers for a secure, but optimizing, compilation chain that robustly preserves \gls{ms} and \gls{cct}.
+Concretely, \Cref{subsec:cs:tms} describes robust preservation of \gls{tms}. % by compiling from a language with a strong, static type system to a language without any type system.
+\Cref{subsec:cs:ms} describes the robust preservation of \gls{sms} and, when composed with the compiler for \gls{tms}, \gls{ms}.
+\Cref{subsec:cs:scct} composes the compiler for \gls{ms} with another one that robustly preserves \gls{cct}.
+We attain our final compilation chain in \Cref{subsec:cs:opts}, which defines optimizing compilers performing either \gls{cf} or \gls{dce} and composes them with the ones from \Cref{subsec:cs:tms,subsec:cs:ms,subsec:cs:scct}.
+
+\subsection{Robust Temporal Memory Safety Preservation}\label{subsec:cs:tms}
+\subsection{Robust (Spatial) Memory Safety Preservation}\label{subsec:cs:ms}
+\subsection{Robust Strict Cryptographic Constant Time Preservation}\label{subsec:cs:scct}
+\subsection{Optimizing Compilers}\label{subsec:cs:opts}
+
+%\begin{lstlisting}[language=c,caption=``Wrong bounds check of two 64-bit integers.'']
+%if at <= bounds {
+%  x[at] = 42;
+%}
+%\end{lstlisting}
+%Note the subtle bug of using \verb|<=| instead of \verb|<|, leading to an out-of-bounds access whenever \verb|at = bounds|.
+%Furthermore, when \verb|at| and \verb|bounds| are 64-bit integers on a 32-bit architecture, the comparison may not be performed in constant-time: The compiler may bail out as soon as the lower 32 bits are unequal, not bothering to compare the higher 32 bits.
+
+\section{Related Work}\label{sec:relwork}
+\section{Conclusion}\label{sec:concl}
 

### 2023-06-09 popl/main.tex
Add overview

@@
 
-\BrText{Overview}{This paper introduces a formal framework for secure compiler compositions and showcases the framework's power by instantiating it on a multi-level compilation chain.
+\BrText{Solution}{This paper introduces a formal framework for secure compiler compositions and showcases the framework's power by instantiating it on a multi-level compilation chain.
 To this end, we discuss security property compositions and define several secure compilers, where each is either robustly preserving a different security property of interest or optimizing, i.e., applying \gls{cf} or doing \gls{dce}.
@@
 
-In summary, this paper makes the following contributions:
+\BrText{Validation}{In summary, this paper makes the following contributions:
 
@@
         Finally, in this paper, whenever definitions or proofs are done in Coq, we use the symbol \CoqSymbol.
-\end{itemize}
+\end{itemize}}
 
@@
 \section{Composing Robustly Secure Compilers}\label{sec:compcomp}
-This section presents three different kinds of secure compiler compositions and their corresponding proofs.
+This section studies different kinds of secure compiler compositions and their corresponding proofs.
+Most notably, it looks at the sequential composition of secure compilers and shows a theorem that allows for modular proof engineering: compilers need not to robustly preserve the same property in order to be securely composed.
+Moreover, it demonstrates that the order does not matter for composing compilers with the same source and target language.
 
@@
 We demonstrate the power of our framework from \Cref{sec:compprop,sec:compcomp} by composing these compilers for a secure, but optimizing, compilation chain that robustly preserves \gls{ms} and \gls{cct}.
-Concretely, \Cref{subsec:cs:tms} describes robust preservation of \gls{tms}. % by compiling from a language with a strong, static type system to a language without any type system.
-\Cref{subsec:cs:ms} describes the robust preservation of \gls{sms} and, when composed with the compiler for \gls{tms}, \gls{ms}.
-\Cref{subsec:cs:scct} composes the compiler for \gls{ms} with another one that robustly preserves \gls{cct}.
-We attain our final compilation chain in \Cref{subsec:cs:opts}, which defines optimizing compilers performing either \gls{cf} or \gls{dce} and composes them with the ones from \Cref{subsec:cs:tms,subsec:cs:ms,subsec:cs:scct}.
+Concretely, \Cref{subsec:cs:tms} starts by defining languages $\src{L_{TMS}}$ and $\trg{L}$ and describes a compiler $\llbracket\src{\bullet}\rrbracket^{\src{L_{TMS}}\to\trg{L}}$ for the robust preservation of \gls{tms} between the two languages.
+\Cref{subsec:cs:ms} defines $\irl{L_{MS}}$ and a compiler $\llbracket\trg{\bullet}\rrbracket^{\trg{L}\to\irl{L_{MS}}}$ that robustly preserves \gls{sms}.
+By the framework presented in \Cref{sec:compcomp,sec:compprop}, we attain a compiler $\llbracket\src{\bullet}\rrbracket^{\src{L_{TMS}}\to\irl{L_{MS}}}$ for the robust preservation of \gls{ms} by composing $\llbracket\src{\bullet}\rrbracket^{\src{L_{TMS}}\to\trg{L}}$ and $\llbracket\trg{\bullet}\rrbracket^{\trg{L}\to\irl{L_{MS}}}$.
+\Cref{subsec:cs:scct} defines $\obj{L_{CCT}}$ and a compiler $\llbracket\irl{\bullet}\rrbracket^{\irl{L_{MS}}\to\obj{L_{CCT}}}$ that robustly preserves \gls{cct}.
+Composing it with $\llbracket\src{\bullet}\rrbracket^{\src{L_{TMS}}\to\irl{L_{MS}}}$ yields $\llbracket\src{\bullet}\rrbracket^{\src{L_{TMS}}\to\obj{L_{CCT}}}$, a compiler that robustly preserves a combination of \gls{ms} and \gls{cct}.
+We attain our final compilation chain in \Cref{subsec:cs:opts}, which defines optimizing compilers performing either \gls{cf} or \gls{dce} and composes them with $\llbracket\src{\bullet}\rrbracket^{\src{L_{TMS}}\to\obj{L_{CCT}}}$, giving us an optimizing, secure compilation chain.
 

### 2023-06-11 popl/main.tex
Upto section 3

@@
 \usepackage[colorinlistoftodos]{todonotes}
+\usepackage[inference]{semantic}
 \usepackage{cleveref}
@@
 \section{Background}\label{sec:background}
-We give a brief introduction to trace properties, monitors, and property satisfaction (\Cref{subsec:bg:tprop}).
-With property satisfaction, we re-introduce the concept of robust satisfaction and its compiler-driven preservation (\Cref{subsec:bg:rtp}).
+\Cref{subsec:bg:tprop} introduces trace properties, monitors, and property satisfaction.
+\Cref{subsec:bg:rtp} extends the property satisfaction with robustness and shows its compiler-driven preservation, as presented in~\cite{abate2019jour}.
 
-\subsection{Trace-Based Properties}\label{subsec:bg:tprop}
-\subsection{Robust Trace-Property Preservation}\label{subsec:bg:rtp}
+\subsection{Trace-Based Properties, Programming Languages, and Property Satisfaction}\label{subsec:bg:tprop}
+
+\emph{Events} $\event$ are an abstract representation of observable actions a program can do during its execution.
+Typically, one always has an empty observation $\emptyevent$ used to represent unimportant actions.
+A sequence of events $\trace$ is a \emph{trace}.
+So, throughout the paper, we use $\overline{\cdot}$ to mark meta-variables as sequences.
+Empty sequences are denoted with $\hole{\cdot}$ and concatenation of traces $\trace[_{1}],\trace[_{2}]$ is written as $\trace[_{1}]\cdot\trace[_{2}]$.
+For sake of brevity, we use this notation also for pre- and appending events to traces.
+
+A trace-based property $\pi$ is a set of admissable traces.
+For a trace to be admissable, a predicate evaluates to true whenever $\trace\in\pi$.
+Classes of properties $\class$, commonly referred to as hyperproperties in literature, are sets of properties.
+The \emph{lifting} of a property $\pi$ is simply its powerset and denoted as $\lift{\pi}$.
+To define property satisfaction, we first look at programming languages.
+
+For languages used in this paper, we consider a small-step style semantics.
+That is, there is a step relation between states $\statevar$ and an event $\event$: $\stepto[]{\statevar[_{1}]}{\statevar[_{2}]}{\event}$.
+Also, a program's syntax $p$ is encoded in the state $\statevar$ which we write as $p\in\statevar$.
+For states $\statevar$, we assume there exists a judgement $\isval{\statevar}$ that determines wether $\statevar$ is a final state or not.
+\Cref{fig:starstep:and:nstep}\MK{appendix?} defines the star-step relation $\stepto{\statevar[_{1}]}{\statevar[_{2}]}{\trace}$, essentially modelling a big-step style evaluation, and the $n$-step relation $\stepton{\statevar[_{1}]}{\statevar[_{2}]}{\trace}$ that evaluates a program for $n$ steps.
+Both relations filter $\varepsilon$ events.
+
+\begin{figure}[t]
+  \judgbox{\stepto{r}{r'}{\trace}}{,,State $\statevar$ steps to final state $\statevar[']$, emitting events $\trace$.''}$\;$\\
+  %
+  \typerule{star-step-refl}{
+    \isval{\statevar}
+  }{
+    \stepto{\statevar}{\statevar}{\hole{\cdot}}
+  }{star-step-refl}
+  %
+  \typerule{star-step-trans}{
+    \event\not=\varepsilon &
+    \stepto[]{\statevar}{\statevar[']}{\event} &
+    \stepto{\statevar[']}{\statevar['']}{\trace}
+  }{
+    \stepto{\statevar}{\statevar['']}{\event\cdot\trace}
+  }{star-step-trans}
+  %
+  \typerule{star-step-ign}{
+    \stepto[]{\statevar}{\statevar[']}{\varepsilon} &
+    \stepto{\statevar[']}{\statevar['']}{\trace}
+  }{
+    \stepto{\statevar}{\statevar['']}{\trace}
+  }{star-step-ign}
+
+  \judgbox{\stepton{\statevar}{\statevar[']}{\trace}}{,,State $r$ does $n$ steps to final state $\statevar[']$, emitting events $\trace$.''}$\;$\\
+  %
+  \typerule{n-step-refl}{
+  }{
+    \stepton[0]{\statevar}{\statevar}{\hole{\cdot}}
+  }{n-step-refl}
+  %
+  \typerule{n-step-trans}{
+    \event\not=\varepsilon &
+    \stepto[]{\statevar}{\statevar[']}{\event} &
+    \stepton{\statevar[']}{\statevar['']}{\trace}
+  }{
+    \stepton[n+1]{\statevar}{\statevar['']}{\event\cdot\trace}
+  }{n-step-trans}
+  %
+  \typerule{n-step-ign}{
+    \stepto[]{\statevar}{\statevar[']}{\varepsilon} &
+    \stepton{\statevar[']}{\statevar['']}{\trace}
+  }{
+    \stepton[n+1]{\statevar}{\statevar['']}{\trace}
+  }{n-step-ign}
+
+  \caption{Star-step and $n$-step relation for a given step relation $\stepto{\statevar[_{1}]}{\statevar[_{2}]}{\event}$}\label{fig:starstep:and:nstep}
+\end{figure}
+%
+The \emph{behavior} of a whole program $w$ ($\behav{w}$) is a set containing all traces the program may emit during execution.
+Note that for deterministic languages, $\behav{w}$ is a singleton set for any $w$.
+With this, we can define satisfaction:
+
+\begin{definition}[Property Satisfaction]\label{def:propsat}
+  A \emph{whole} program $w$ satisfies $\pi$ ($\sat{w}{\pi}$) iff $\behav{w}\in\lift{\pi}$.
+\end{definition}
+\begin{definition}[Class Satisfaction]\label{def:classsat}
+  A \emph{whole} program $w$ satisfies $\class$ ($\sat{w}{\class}$) iff $\behav{w}\in\class$.
+\end{definition}
+
+\subsection{Attacker Model and Robust Trace-Property Preservation}\label{subsec:bg:rtp}
+
+Partial programs cannot be executed, because they lack definitions for some symbols.
+Providing definitions for symbols is usually the job of the \emph{linker}.
+We denote linking two programs $\progvar[_{1}],\progvar[_{2}]$ as $\link{\progvar[_{1}]}{\progvar[_{2}]}$.
+For sake of simplicity, we assume that linking does not fail.
+With linking, we can define robust satisfaction analogously to \Cref{def:propsat,def:classsat}:
+
+\begin{definition}[Robust Property Satisfaction]\label{def:proprsat}
+  A program $\progvar$ robustly satisfies $\pi$ ($\rsat{\progvar}{\pi}$) iff for any program $\contextvar$ we have $\sat{\link{\contextvar}{\progvar}}{\pi}$. (similarily for $\rsat{\progvar}{\class}$)
+\end{definition}
+
+Robust satisfaction gives rise to a maximally powerful attacker model: Any other partial program, here referred to as context $\contextvar$, is the attacker.
+So, for example, if the programming language of $\progvar$ and $\contextvar$ allows guessing of memory addresses, then $\contextvar$ is able to read and write any position in memory and therefore guess the location of secrets.
+
+We expect compilers to preserve robust satisfaction.
+In this paper, compilers are modelled as function from one programming language's syntax to the other.
+Since this paper considers composing compilers, we'll adopt notation similar to tombstone diagrams to easen the formal burden.
+So, we denote a compiler from $\src{L}$ to $\trg{L}$ as $\cc{\src{L}\to\trg{L}}{\src{\bullet}}$ and we refer to $\src{L}$ as \emph{source} language and to $\trg{L}$ as \emph{target} language.
+Composing $\cc{\src{L}\to\trg{L}}{\src{\bullet}}$ with $\cc{\trg{L}\to\obj{L}}{\trg{\bullet}}$ is denoted as $\cc{\src{L}\to\trg{L}\to\obj{L}}{\src{\bullet}}$ and means first running $\cc{\src{L}\to\trg{L}}{\src{\bullet}}$ and plugging the output into $\cc{\trg{L}\to\obj{L}}{\trg{\bullet}}$.
+Contrary to tombstones, we don't consider a compiler's implementation language.
+We now define the compiler-property robust trace-property preservation.
+
+\begin{definition}[Robust Trace-Property Preservation]\label{def:rtp}
+  A compiler $\cc{\src{L}\to\trg{L}}{\src{\bullet}}$ robustly preserves a class of properties $\class$ ($\rtp{\cc{\src{L}\to\trg{L}}{\src{\bullet}}}{\class}$) iff for any property $\pi\in\class$ and $\src{p}$ written in $\src{L}$, given $\rsat{\src{\progvar}}{\pi}$, then $\rsat{\cc{\src{L}\to\trg{L}}{\src{\progvar}}}{\pi}$.
+\end{definition}
+
+Note that in the conclusion of \Cref{def:rtp} there is a target-level context.
+So, programs that robustly satisfy a property are hardened against a maximally malicious environment with respect to the power of the target language.
 
 \section{Composing Properties}\label{sec:compprop}
+\Cref{subsec:propdefs} introduces a ,,most general'' trace model and defines properties (\gls{tms}, \gls{sms}, \gls{ms}, and \gls{cct}) that our case study (\Cref{sec:casestud}) on it.
+By means of taking these properties as an example, \Cref{subsec:monitors} defines monitors for them and an agreement relation between a monitor's trace model with the most general trace model.
+Then, we define monitor-based satisfaction and we prove that trace-property satisfaction is implied by it.
+Since monitor-based satisfaction is easier to reason about, this makes our later proofs considerably simpler.
+
+\subsection{Most General Trace-Model and Example Property Definitions}\label{subsec:propdefs}
+
+\begin{definition}[\glsfirst{tms}]
+
+\end{definition}
+\begin{definition}[\glsfirst{sms}]
+
+\end{definition}
+\begin{definition}[\glsfirst{ms}]
+
+\end{definition}
+\begin{definition}[\glsfirst{scct}]
+
+\end{definition}
+
+\subsection{Monitors and Trace Agreements}\label{subsec:monitors}
+
 \section{Composing Robustly Secure Compilers}\label{sec:compcomp}

### 2023-06-11 popl/main.tex
Define properties in section 3

@@
 \usepackage[inference]{semantic}
+\usepackage{fontawesome5}
+\usepackage{glossaries}
 \usepackage{cleveref}
-\usepackage{glossaries}
+\usepackage{stmaryrd}
 \usepackage{listings}
@@
 \section{Composing Properties}\label{sec:compprop}
-\Cref{subsec:propdefs} introduces a ,,most general'' trace model and defines properties (\gls{tms}, \gls{sms}, \gls{ms}, and \gls{cct}) that our case study (\Cref{sec:casestud}) on it.
+\Cref{subsec:propdefs} introduces a ,,most general'' trace model and defines properties (\gls{tms}, \gls{sms}, \gls{ms}, and \gls{cct}) that our case study (\Cref{sec:casestud}) uses.
 By means of taking these properties as an example, \Cref{subsec:monitors} defines monitors for them and an agreement relation between a monitor's trace model with the most general trace model.
@@
 
-\subsection{Most General Trace-Model and Example Property Definitions}\label{subsec:propdefs}
+\subsection{Most General Trace-Model and Some Property Definitions}\label{subsec:propdefs}
 
-\begin{definition}[\glsfirst{tms}]
+\begin{gather*}
+  \begin{aligned}
+  \mi{(Security\ Tag)}~\sigma\bnfdef&\ \lock \mid \unlock\hspace{0.5cm}
+  \mi{(Control\ Tag)}~\sandboxtag\bnfdef \ctx\mid\comp\\
+  \mi{(Pre-event)}~\event[_{b}]\bnfdef&\ \ev{Alloc\ \loc\ n} \mid \ev{Dealloc\ \loc} \mid \ev{Use\ \loc\ n} \mid \ev{Branch\ n} \mid \ev{Binop\ n} \\
+  \mi{(Event)}~\event\bnfdef&\ \emptyevent \mid \lightning \mid \event[_{b}];\sandboxtag;\sigma \\
+  \end{aligned}
+\end{gather*}
 
+Security tags describe wether an event contains sensitive information ($\lock$) or not ($\unlock$).
+Control tags state wether the context ($\ctx$) or the component ($\comp$) are responsible for emitting the event.
+Pre-events describe the actual kind of event that happened.
+$\ev{Alloc\ \loc\ n}$ states that an allocation of size $n$ happened and is stored at $\loc$.
+$\ev{Dealloc\ \loc}$ announces that location $\loc$ is freed.
+$\ev{Use\ \loc\ n}$ abstracts over reads from $\loc$ and writes to $\loc$ in memory, where $n$ is read/written, respectively.
+$\ev{Branch\ n}$ and $\ev{Binop\ n}$ leak values involved in a branch or in a binary operation with data-dependent execution time.
+Finally, events are either the emptyevent ($\emptyevent$), a program crash ($\lightning$), or a tuple consisting of a pre-event, a control tag, and a security tag.
+
+We use this trace model to define any property that interests us.
+In a sense the trace model is most general, because it abstracts all events done in our languages as defined in \Cref{sec:casestud}.
+We write $\event[_{1}]\le_{\trace}\event[_{2}]$ if $\event[_{1}]$ occurs at an earlier position in $\trace$ than $\event[_{2}]$.
+
+\gls{tms} requires that pointers are freed exactly once and that no use happens after a free.
+
+\begin{definition}[\glsfirst{tms}]\label{def:trace:tmsdef}
+  A trace $\trace$ is temporal memory safe ($\trace\in\tmssafe$) iff $\ev{Alloc\ \loc\ n;\sandboxtag;\securitytag}\le_{\trace}\ev{Dealloc\ \loc;\sandboxtag';\securitytag'}$, $\ev{Use\ \loc\ n;\sandboxtag;\securitytag}\not\le_{\trace}\ev{Alloc\ \loc\ m;\sandboxtag';\securitytag'}$ and $\ev{Use\ \loc\ n;\sandboxtag;\securitytag}\not\le_{\trace}\ev{Dealloc\ \loc;\sandboxtag';\securitytag'}$.
 \end{definition}
-\begin{definition}[\glsfirst{sms}]
 
+\gls{sms} disregards out of bounds accesses.
+
+\begin{definition}[\glsfirst{sms}]\label{def:trace:smsdef}
+  A trace $\trace$ is spatial memory safe ($\trace\in\smssafe$) iff given $\ev{Alloc\ \loc\ n;\sandboxtag;\securitytag}\le_{\trace}\ev{Use\ \loc\ m;\sandboxtag';\securitytag'}$, then $n<m$.
 \end{definition}
-\begin{definition}[\glsfirst{ms}]
 
+Full \gls{ms} is then described as the conjunction of \Cref{def:trace:tmsdef,def:trace:smsdef}.
+
+\begin{definition}[\glsfirst{ms}]\label{def:trace:msdef}
+  A trace $\trace$ is memory safe ($\trace\in\mssafe$) iff $\trace\in\tmssafe$ and $\trace\in\smssafe$.
 \end{definition}
+
+We define a variant of the \gls{cct} security property.
+Generally, one cannot define \gls{cct} just as trace-property without, e.g., performing a self-composition of a program.
+That's why we adopt for a stricter variant of \gls{cct}, called \gls{scct}, that enforces common policies to ensure code is \gls{cct}.
+For example, one must not branch on secrets or perform binary operations, such as division, on secrets whose timing is data-dependent.
+We opt for a definition that disregards any secrets leaked on a trace.
+To define the property, we need a function on traces that strips all events with a $\lock$-tag: $\downarrow_{\lock}(\cdot)$.
+
 \begin{definition}[\glsfirst{scct}]
-
+  A trace $\trace$ is strictly cryptographic constant time ($\trace\in\scctsafe$) iff $\trace=\downarrow_{\lock}\left(\trace\right)$.
 \end{definition}

### 2023-06-11 popl/main.tex
Change notation of compilers.

@@
 We demonstrate the power of our framework from \Cref{sec:compprop,sec:compcomp} by composing these compilers for a secure, but optimizing, compilation chain that robustly preserves \gls{ms} and \gls{cct}.
-Concretely, \Cref{subsec:cs:tms} starts by defining languages $\src{L_{TMS}}$ and $\trg{L}$ and describes a compiler $\llbracket\src{\bullet}\rrbracket^{\src{L_{TMS}}\to\trg{L}}$ for the robust preservation of \gls{tms} between the two languages.
-\Cref{subsec:cs:ms} defines $\irl{L_{MS}}$ and a compiler $\llbracket\trg{\bullet}\rrbracket^{\trg{L}\to\irl{L_{MS}}}$ that robustly preserves \gls{sms}.
-By the framework presented in \Cref{sec:compcomp,sec:compprop}, we attain a compiler $\llbracket\src{\bullet}\rrbracket^{\src{L_{TMS}}\to\irl{L_{MS}}}$ for the robust preservation of \gls{ms} by composing $\llbracket\src{\bullet}\rrbracket^{\src{L_{TMS}}\to\trg{L}}$ and $\llbracket\trg{\bullet}\rrbracket^{\trg{L}\to\irl{L_{MS}}}$.
-\Cref{subsec:cs:scct} defines $\obj{L_{CCT}}$ and a compiler $\llbracket\irl{\bullet}\rrbracket^{\irl{L_{MS}}\to\obj{L_{CCT}}}$ that robustly preserves \gls{cct}.
-Composing it with $\llbracket\src{\bullet}\rrbracket^{\src{L_{TMS}}\to\irl{L_{MS}}}$ yields $\llbracket\src{\bullet}\rrbracket^{\src{L_{TMS}}\to\obj{L_{CCT}}}$, a compiler that robustly preserves a combination of \gls{ms} and \gls{cct}.
-We attain our final compilation chain in \Cref{subsec:cs:opts}, which defines optimizing compilers performing either \gls{cf} or \gls{dce} and composes them with $\llbracket\src{\bullet}\rrbracket^{\src{L_{TMS}}\to\obj{L_{CCT}}}$, giving us an optimizing, secure compilation chain.
+Concretely, \Cref{subsec:cs:tms} starts by defining languages $\src{L_{TMS}}$ and $\trg{L}$ and describes a compiler $\cc{\src{L_{TMS}}\to\trg{L}}{\src{\bullet}}$ for the robust preservation of \gls{tms} between the two languages.
+\Cref{subsec:cs:ms} defines $\irl{L_{MS}}$ and a compiler $\cc{\trg{L}\to\irl{L_{MS}}}{\trg{\bullet}}$ that robustly preserves \gls{sms}.
+By the framework presented in \Cref{sec:compcomp,sec:compprop}, we attain a compiler $\cc{\src{L_{TMS}}\to\irl{L_{MS}}}{\src{\bullet}}$ for the robust preservation of \gls{ms} by composing $\cc{\src{L_{TMS}}\to\trg{L}}{\src{\bullet}}$ and $\cc{\trg{L}\to\irl{L_{MS}}}{\trg{\bullet}}$.
+\Cref{subsec:cs:scct} defines $\obj{L_{CCT}}$ and a compiler $\cc{\irl{L_{MS}}\to\obj{L_{CCT}}}{\irl{\bullet}}$ that robustly preserves \gls{cct}.
+Composing it with $\cc{\src{L_{TMS}}\to\irl{L_{MS}}}{\src{\bullet}}$ yields $\cc{\src{L_{TMS}}\to\obj{L_{CCT}}}{\src{\bullet}}$, a compiler that robustly preserves a combination of \gls{ms} and \gls{cct}.
+We attain our final compilation chain in \Cref{subsec:cs:opts}, which defines optimizing compilers performing either \gls{cf} or \gls{dce} and composes them with $\cc{\src{L_{TMS}}\to\obj{L_{CCT}}}{\src{\bullet}}$, giving us an optimizing, secure compilation chain.
 

### 2023-06-11 popl/main.tex
Add overview for monitor section.

@@
 
+This subsection presents monitors for our properties of interest (\gls{tms}, \gls{sms}, \gls{ms}, \gls{scct}).
+Each monitor has its own accompanying trace model that abstracts over our most general trace model.
+This is because monitors should only be concerned with property-relevant information.
+For each such monitor-specific trace model, we define an agreement relation that relates it with the most general trace model.
+Lastly, if any monitor does not get stuck on a trace and the trace agrees with a most general trace, then the most general one satisfies the associated property of the monitor.
+
+\paragraph{\glsfirst{tms}}
+\paragraph{\glsfirst{sms}}
+\paragraph{\glsfirst{ms}}
+\paragraph{\glsfirst{scct}}
+
 \section{Composing Robustly Secure Compilers}\label{sec:compcomp}

### 2023-06-13 popl/main.tex
TMS monitor definition
title, summary

@@
 %% The first command in your LaTeX source must be the \documentclass command.
+
+% Fixing: Too many math alphabets used in version normal.
+\newcommand\hmmax{0}
+\newcommand\bmmax{0}
 \documentclass[acmsmall,review,screen,dvipsnames]{acmart}
@@
 \usepackage{xspace}
+\usepackage{bm}
 
@@
 %% allowing the author to define a "short title" to be used in page headers.
-\title{Composing Robustly Safe Compilers}
+\title{Secure Composition of Robust and Optimising Compilers}
 
@@
 \begin{center}\small\it
-	{In the following, we use syntax highlighting accessible to both colourblind and black \& white readers~\citep{patrignani2020use}. Specifically, we use a $\src{\text{blue}}$,
-	$\src{sans\text{-}serif}$ font for $\src{L_{TMS}}$,
-	$\trg{bold}$ font for $\trg{L}$ and $\trg{L_{MS}}$, and $\obj{green}$ font for $\obj{L_{sCCT}}$.
+	{In the following, we use syntax highlighting accessible to both colourblind and black \& white readers~\citep{patrignani2020use}. 
+	Specifically, we use a $\src{blue}$, $\src{sans\text{-}serif}$ font for our $\src{source}$,
+	a $\trg{red}$, $\trg{bold}$ font for our $\trg{intermediate}$, 
+	and a $\obj{green}$, $\obj{teletype}$ font for $\obj{target}$ language.
 	}
@@
 
-\BrText{Solution}{This paper introduces a formal framework for secure compiler compositions and showcases the framework's power by instantiating it on a multi-level compilation chain.
+\BrText{Solution}{
+This paper introduces a formal framework for secure compiler compositions and showcases the framework's power by instantiating it on a multi-level compilation chain.
 To this end, we discuss security property compositions and define several secure compilers, where each is either robustly preserving a different security property of interest or optimizing, i.e., applying \gls{cf} or doing \gls{dce}.
@@
 The key results are formalized in Coq.}
+
+% keyword: pass
+\MPin{
+	commands missing
+}
+This paper introduces a framework for reasoning about the composition of secure and optimising compiler passes and it showcases the power of this framework by instantiating it on a multi-pass compilation chain.
+To this end, this paper first discusses how to compose security properties, such as temporal and spatial memory safety into general memory safety, and cryptographic constant time.
+Then, this paper defines several secure compiler passes, where each is either preserving a different security property (e.g., temporal or spatial memory safety) or performing a security-preserving optimisation, (e.g., applying constant-folding or dead-code elimination).
+% keyword- end to end, composition
+Finally, this paper shows how to compose these secure compiler passes into a multi-pass compilation chain which can be proven to provide end-to-end preservation of general memory safety and cryptographic constant time.
+Crucially, this end-to-end security preservation is obtained by having each compilation pass (i) preserve a sub-part of the overall security property individually and then (ii) compose each preserved security property as dictated by our framework.
+Our results showcase how our framework allows the kind of formal security reasoning that compiler writers already want (and do), obtaining precise, compositional security reasoning while providing minimal (and modular) proof effort.
+\MPin{
+	last line: not the best.
+}
 
@@
 $\ev{Alloc\ \loc\ n}$ states that an allocation of size $n$ happened and is stored at $\loc$.
+Locations $\loc$ are elements of $L$ and left abstract.
+However, we expect $L$ to be closed over addition with natural numbers.
 $\ev{Dealloc\ \loc}$ announces that location $\loc$ is freed.
@@
 
-\paragraph{\glsfirst{tms}}
-\paragraph{\glsfirst{sms}}
-\paragraph{\glsfirst{ms}}
-\paragraph{\glsfirst{scct}}
+\paragraph{Monitor for \glsfirst{tms}}
+\begin{gather*}
+  \begin{aligned}
+    \mi{(Abstract\ Store)}~\tmsmonitor\bnfdef&\ \left\{\text{alloced}:L,\text{freed}:L\right\}\\
+    \mi{(Abstract\ Events)}~\absevent\bnfdef&\ \absev{\emptyevent} \mid \absev{Alloc\ }\loc \mid \absev{Dealloc\ }\loc \mid \absev{Use\ }\loc \mid \absterm\\[0.33cm]
+    \tmsmonitor\subseteq\tmsmonitor[']:=\ &\tmsmonitor.\text{freed}\subseteq\tmsmonitor['].\text{freed} \\
+    \loc\in\tmsmonitor:=\ &\loc\in\tmsmonitor.\text{alloced}\text{ and }\loc\notin\tmsmonitor.\text{freed} \\
+    \loc\notin\tmsmonitor:=\ &\loc\notin\tmsmonitor.\text{alloced}\text{ and }\loc\notin\tmsmonitor.\text{freed} \\
+    \left\{\loc\right\}\cup\tmsmonitor:=\ & \left\{\text{alloced}:\left\{\loc\right\}\cup\tmsmonitor.\text{alloced},\text{freed}:\tmsmonitor.\text{freed}\right\}\\
+    \tmsmonitor\setminus\left\{\loc\right\}:=\ &\left\{\text{alloced}:\tmsmonitor.\text{alloced}\setminus\left\{\loc\right\}, \text{freed}:\tmsmonitor.\text{freed}\cup\left\{\loc\right\}\right\}\\
+  \end{aligned}
+\end{gather*}
+
+\paragraph{Monitor for \glsfirst{sms}}
+\begin{gather*}
+  \begin{aligned}
+  \end{aligned}
+\end{gather*}
+\paragraph{Comining Monitors: \glsfirst{ms}}
+
+\paragraph{Monitor for \glsfirst{scct}}
+\begin{gather*}
+  \begin{aligned}
+  \end{aligned}
+\end{gather*}
 

### 2023-06-13 popl/main.tex
Add \MKin macro. Minor grammar fix.

@@
 	a $\trg{red}$, $\trg{bold}$ font for our $\trg{intermediate}$, 
-	and a $\obj{green}$, $\obj{teletype}$ font for $\obj{target}$ language.
+	and a $\obj{green}$, $\obj{teletype}$ font for our $\obj{target}$ language.
 	}

### 2023-06-13 popl/main.tex
Use the other intro. Change \MKin macro. Add note.

@@
 \BrText{Solution}{
-This paper introduces a formal framework for secure compiler compositions and showcases the framework's power by instantiating it on a multi-level compilation chain.
-To this end, we discuss security property compositions and define several secure compilers, where each is either robustly preserving a different security property of interest or optimizing, i.e., applying \gls{cf} or doing \gls{dce}.
-We compose these secure compilers into a multi-level compilation chain which is optimizing and robustly preserves a combination of \gls{ms} and \gls{cct}.
-Crucially, the robust preservation proofs are only done for the individual compilers in the compilation chain and the robust preservation for the whole chain is a result of one of our key theorems.
-The key results are formalized in Coq.}
-
 % keyword: pass
@@
 Finally, this paper shows how to compose these secure compiler passes into a multi-pass compilation chain which can be proven to provide end-to-end preservation of general memory safety and cryptographic constant time.
+\MKin{
+  ,,can be proven'' -> we prove (?)
+}
 Crucially, this end-to-end security preservation is obtained by having each compilation pass (i) preserve a sub-part of the overall security property individually and then (ii) compose each preserved security property as dictated by our framework.
@@
 	last line: not the best.
+}
 }

### 2023-06-13 popl/main.tex
Add monitor definitons

@@
 \usepackage{xspace}
+\usepackage{tikz}
 \usepackage{bm}
@@
   \mi{(Control\ Tag)}~\sandboxtag\bnfdef \ctx\mid\comp\\
-  \mi{(Pre-event)}~\event[_{b}]\bnfdef&\ \ev{Alloc\ \loc\ n} \mid \ev{Dealloc\ \loc} \mid \ev{Use\ \loc\ n} \mid \ev{Branch\ n} \mid \ev{Binop\ n} \\
+  \mi{(Pre\text{--}event)}~\event[_{b}]\bnfdef&\ \ev{Alloc\ \loc\ n} \mid \ev{Dealloc\ \loc} \mid \ev{Use\ \loc\ n} \mid \ev{Branch\ n} \mid \ev{Binop\ n} \\
   \mi{(Event)}~\event\bnfdef&\ \emptyevent \mid \lightning \mid \event[_{b}];\sandboxtag;\sigma \\
@@
 This subsection presents monitors for our properties of interest (\gls{tms}, \gls{sms}, \gls{ms}, \gls{scct}).
-Each monitor has its own accompanying trace model that abstracts over our most general trace model.
-This is because monitors should only be concerned with property-relevant information.
-For each such monitor-specific trace model, we define an agreement relation that relates it with the most general trace model.
-Lastly, if any monitor does not get stuck on a trace and the trace agrees with a most general trace, then the most general one satisfies the associated property of the monitor.
+Each monitor $\monitor$ has its own accompanying trace model that abstracts over our most general trace model.
+%This is because monitors should only be concerned with property-relevant information.
+Traces of the accompanying trace model are written as $\abstrace$ and $\monitorcheck{\monitor}{\monitor[']}{\abstrace}$ is the notation for the semantics of a monitor.
+The empty monitor is denoted as $\emptyset$.
+For each model of a monitor, an agreement relation $\traceagree{\pi}{\trace}{\abstrace}$ is defined that relates monitor specific traces with most general traces.
+The notation $\traceagree[]{\pi}{\event}{\absevent}$ is used to relate just events and the star version of it is defined similarily to \Cref{fig:starstep:and:nstep}.
+With this, first, monitor satisfaction is defined and then the individual monitors.
+Finally, for every monitor, it is shown that monitor satisfaction implies trace property satisfaction.
+
+\begin{definition}[Monitor Satisfaction]\label{def:monsat}
+  A given trace $\trace$ monitor-satisfies a property $\pi$ ($\monsafe{\trace}{\pi}$) iff %
+  there exists $\abstrace,\monitor$ such that $\traceagree{\pi}{\trace}{\abstrace}$ and $\monitorcheck{\emptyset}{\monitor}{\abstrace}$.
+  \Coqed
+\end{definition}
 
@@
   \begin{aligned}
-    \mi{(Abstract\ Store)}~\tmsmonitor\bnfdef&\ \left\{\text{alloced}:L,\text{freed}:L\right\}\\
-    \mi{(Abstract\ Events)}~\absevent\bnfdef&\ \absev{\emptyevent} \mid \absev{Alloc\ }\loc \mid \absev{Dealloc\ }\loc \mid \absev{Use\ }\loc \mid \absterm\\[0.33cm]
-    \tmsmonitor\subseteq\tmsmonitor[']:=\ &\tmsmonitor.\text{freed}\subseteq\tmsmonitor['].\text{freed} \\
-    \loc\in\tmsmonitor:=\ &\loc\in\tmsmonitor.\text{alloced}\text{ and }\loc\notin\tmsmonitor.\text{freed} \\
-    \loc\notin\tmsmonitor:=\ &\loc\notin\tmsmonitor.\text{alloced}\text{ and }\loc\notin\tmsmonitor.\text{freed} \\
-    \left\{\loc\right\}\cup\tmsmonitor:=\ & \left\{\text{alloced}:\left\{\loc\right\}\cup\tmsmonitor.\text{alloced},\text{freed}:\tmsmonitor.\text{freed}\right\}\\
-    \tmsmonitor\setminus\left\{\loc\right\}:=\ &\left\{\text{alloced}:\tmsmonitor.\text{alloced}\setminus\left\{\loc\right\}, \text{freed}:\tmsmonitor.\text{freed}\cup\left\{\loc\right\}\right\}\\
+    \mi{(Abstract\ Store)}~\tmsmonitor\bnfdef&\ \left\{\text{alloced}:L,\text{freed}:L\right\} \hspace{0.33cm}%
+    \emptyset:=\ \left\{\text{alloced}:\emptyset,\text{freed}:\emptyset\right\}\\
+    \mi{(Abstract\ Events)}~\absevent\bnfdef&\ \absev{\emptyevent} \mid \absev{Alloc\ }\loc \mid \absev{Dealloc\ }\loc \mid \absev{Use\ }\loc \mid \absterm\\
   \end{aligned}
 \end{gather*}
+\begin{center}
+  \judgbox{\monitorcheck[]{\tmsmonitor}{\tmsmonitor[']}{\absevent}}{,,Monitor $\tmsmonitor$ steps to $\tmsmonitor[']$ given event $\absevent$.''}$\;$\\
+  %
+  \typerule{tms-uninteresting}{
+  }{
+    \monitorcheck[]{\tmsmonitor}{\tmsmonitor}{\absev{\emptyevent}}
+  }{tms-uninteresting}
+  %
+  \typerule{tms-abort}{
+  }{
+    \monitorcheck[]{\tmsmonitor}{\emptyset}{\absterm}
+  }{tms-abort}
+  %
+  \typerule{tms-use}{
+    \loc\in\tmsmonitor.\text{alloced} &
+    \loc\notin\tmsmonitor.\text{freed}
+  }{
+    \monitorcheck[]{\tmsmonitor}{\tmsmonitor}{\absev{Use\ }\loc}
+  }{tms-use}
+  %
+  \typerule{tms-alloc}{
+    \loc\notin\tmsmonitor.\text{alloced} &
+    \loc\notin\tmsmonitor.\text{freed} \\
+    \tmsmonitor[']=\left\{\text{alloced}: \tmsmonitor.\text{alloced}\cup\left\{\loc\right\},\text{freed}: \tmsmonitor.\text{freed}\right\}
+  }{
+    \monitorcheck[]{\tmsmonitor}{\tmsmonitor[']}{\absev{Alloc\ }\loc}
+  }{tms-alloc}
+  %
+  \typerule{tms-dealloc}{
+    \loc\notin\tmsmonitor.\text{alloced} &
+    \loc\notin\tmsmonitor.\text{freed} \\
+    \tmsmonitor[']=\left\{\text{alloced}: \tmsmonitor.\text{alloced}\setminus\left\{\loc\right\},\text{freed}: \tmsmonitor.\text{freed}\cup\left\{\loc\right\}\right\}
+  }{
+    \monitorcheck[]{\tmsmonitor}{\tmsmonitor[']}{\absev{Dealloc\ }\loc}
+  }{tms-dealloc}
+\end{center}
+\begin{center}
+  \judgbox{\tmstraceagree[]{\event}{\absevent}}{,,Abstract event $\absevent$ is equivalent to $\event$ with respect to \gls{tms}.''}$\;$\\
+  %
+  \typerule{tms-alloc-authentic}{
+  }{
+    \tmstraceagree[]{\ev{Alloc}\ \loc\ n;\sandboxtag;\securitytag}{\absev{Alloc\ }\loc}
+  }{tms-alloc-authentic}
+  %
+  \typerule{tms-dealloc-authentic}{
+  }{
+    \tmstraceagree[]{\ev{Dealloc}\ \loc;\sandboxtag;\securitytag}{\absev{Dealloc\ }\loc}
+  }{tms-dealloc-authentic}
+  %
+  \typerule{tms-use-authentic}{
+  }{
+    \tmstraceagree[]{\ev{Use}\ \loc\ n;\sandboxtag;\securitytag}{\absev{Use\ }\loc}
+  }{tms-use-authentic}
+  %
+  \typerule{tms-branch-authentic}{
+  }{
+    \tmstraceagree[]{\ev{Branch\ }n}{\absev{\emptyevent}}
+  }{tms-branch-authentic}
+  %
+  \typerule{tms-binop-authentic}{
+  }{
+    \tmstraceagree[]{\ev{Binop\ }n}{\absev{\emptyevent}}
+  }{tms-binop-authentic}
+  %
+  \typerule{tms-empty-authentic}{
+  }{
+    \tmstraceagree[]{\emptyevent}{\absev{\emptyevent}}
+  }{tms-empty-authentic}
+  %
+  \typerule{tms-abort-authentic}{
+  }{
+    \tmstraceagree[]{\lightning}{\absterm}
+  }{tms-abort-authentic}
+\end{center}
+
+The state of the monitor is a record containing two sets that keep track of allocated and freed locations.
+The monitor runs on a modified trace semantics with more abstract $\absev{Alloc\ }\loc$,$\absev{Dealloc\ }\loc$,$\absev{Use\ }\loc$, and $\absterm$ events, which represent {\em allocation}, {\em deallocation}, and {\em use} of a location $\loc$ as well as abnormal program termination.
+\Cref{tr:tms-use} simply requires that a location is (i) allocated and (ii) not freed.
+\Cref{tr:tms-alloc,tr:tms-dealloc} both require a location to not be freed already and depending on the witnessed action, extend the internal state of the monitor.
+The trace agreement is entirely straightforward.
+
+\begin{lemma}[Traces with Monitor Satisfaction are $\tmssafe$]\label{lem:mon:tmsafe}
+  If $\montmssafe{\trace}$, then $\trace\in\tmssafe$.\Coqed
+\end{lemma}
 
@@
   \begin{aligned}
+    \mi{(Abstract\ Store)}~\smsmonitor:=\ L\times\mb{N}\hspace{0.33cm}&
+    \mi{(Abstract\ Events)}~\absevent\bnfdef\ \absev{\varepsilon} \mid \absev{Alloc\ }\loc\ n \mid \absev{Dealloc\ }\loc \mid \absev{Use\ }\loc\ n \\
   \end{aligned}
 \end{gather*}
+\MKin{do we really need $\absev{Dealloc}$?}
+\begin{center}
+  \judgbox{\monitorcheck[]{\smsmonitor}{\smsmonitor[']}{\absevent}}{,,Monitor $\smsmonitor$ steps to $\smsmonitor[']$ given event $\absevent$.''}$\;$\\
+  %
+  \typerule{sms-uninteresting}{
+  }{
+    \monitorcheck[]{\smsmonitor}{\smsmonitor}{\absev{\emptyevent}}
+  }{sms-uninteresting}
+  %
+  \typerule{sms-abort}{
+  }{
+    \monitorcheck[]{\smsmonitor}{\emptyset}{\absterm}
+  }{sms-abort}
+  %
+  \typerule{sms-use}{
+    (\loc,m)\in\smsmonitor &
+    n<m
+  }{
+    \monitorcheck[]{\smsmonitor}{\smsmonitor}{\absev{Use\ }\loc\ n}
+  }{sms-use}
+  %
+  \typerule{sms-alloc}{
+    \loc\notin\operatorname{dom}\smsmonitor &
+    \smsmonitor[']=\smsmonitor\cup\left\{(\loc,n)\right\}
+  }{
+    \monitorcheck[]{\smsmonitor}{\smsmonitor[']}{\absev{Alloc\ }\loc\ n}
+  }{sms-alloc}
+  %
+  \typerule{sms-dealloc}{
+    (\loc,n)\in\smsmonitor &
+    \smsmonitor[']=\smsmonitor\setminus\left\{(\loc,n)\right\}
+  }{
+    \monitorcheck[]{\smsmonitor}{\smsmonitor[']}{\absev{Dealloc\ }\loc}
+  }{sms-dealloc}
+\end{center}
+\begin{center}
+  \judgbox{\smstraceagree[]{\event}{\absevent}}{,,Abstract event $\absevent$ is equivalent to $\event$ with respect to \gls{sms}.''}$\;$\\
+  %
+  \typerule{sms-alloc-authentic}{
+  }{
+    \smstraceagree[]{\ev{Alloc}\ \loc\ n;\sandboxtag;\securitytag}{\absev{Alloc\ }\loc\ n}
+  }{sms-alloc-authentic}
+  %
+  \typerule{sms-dealloc-authentic}{
+  }{
+    \smstraceagree[]{\ev{Dealloc}\ \loc;\sandboxtag;\securitytag}{\absev{Dealloc\ }\loc}
+  }{sms-dealloc-authentic}
+  %
+  \typerule{sms-use-authentic}{
+  }{
+    \smstraceagree[]{\ev{Use}\ \loc\ n;\sandboxtag;\securitytag}{\absev{Use\ }\loc\ n}
+  }{sms-use-authentic}
+  %
+  \typerule{sms-branch-authentic}{
+  }{
+    \smstraceagree[]{\ev{Branch\ }n}{\absev{\emptyevent}}
+  }{sms-branch-authentic}
+  %
+  \typerule{sms-binop-authentic}{
+  }{
+    \smstraceagree[]{\ev{Binop\ }n}{\absev{\emptyevent}}
+  }{sms-binop-authentic}
+  %
+  \typerule{sms-empty-authentic}{
+  }{
+    \smstraceagree[]{\emptyevent}{\absev{\emptyevent}}
+  }{sms-empty-authentic}
+  %
+  \typerule{sms-abort-authentic}{
+  }{
+    \smstraceagree[]{\lightning}{\absterm}
+  }{sms-abort-authentic}
+\end{center}
+
+The state of the monitor for \gls{sms} is a set containing pairs of locations and the allocation size of the objects the locations point to.
+In comparison to the trace model of the \gls{tms} monitor, the trace model here is extended by sizing and positional information.
+\Cref{tr:sms-use} performs a bounds check and \Cref{tr:sms-alloc,tr:sms-dealloc} add or remove locations and the size of an allocation.
+The trace agreement relation is, again, entirely straightforward.
+
+\begin{lemma}[Monitor Traces are $\smssafe$]\label{lem:mon:smsafe}
+  If $\monsmssafe{\trace}$, then $\trace\in\smssafe$.\Coqed
+\end{lemma}
+
 \paragraph{Comining Monitors: \glsfirst{ms}}
+
+\begin{center}
+  \judgbox{\monitorcheck[]{\msmonitor}{\msmonitor[']}{\absevent}}{,,Monitor $\msmonitor$ steps to $\msmonitor[']$ given event $\absevent$.''}$\;$\\
+  \typerule{ms-step}{
+    \monitorcheck[]{\tmsmonitor}{\tmsmonitor[']}{\absevent_{\tmssafe}} &
+    \monitorcheck[]{\smsmonitor}{\smsmonitor[']}{\absevent_{\smssafe}} &
+  }{
+    \monitorcheck[]{(\tmsmonitor,\smsmonitor)}{(\tmsmonitor['],\smsmonitor['])}{(\absevent_{\tmssafe},\absevent_{\smssafe})}
+  }{ms-step}
+\end{center}
+\begin{center}
+  \judgbox{\mstraceagree{\event}{\absevent}}{,,Abstract event $\absevent$ is equivalent to $\event$ with respect to \gls{ms}.''}$\;$\\
+  %
+  \typerule{ms-authentic}{
+    \tmstraceagree[]{\event}{\absevent_{\tmssafe}} &
+    \smstraceagree[]{\event}{\absevent_{\smssafe}}
+  }{
+    \mstraceagree[]{\event}{(\absevent_{\tmssafe},\absevent_{\smssafe})}
+  }{ms-authentic}
+\end{center}
+
+To obtain a monitor for \gls{ms}, the monitors for \gls{tms} and \gls{sms} are simply combined in lock-step.
+
+\begin{lemma}[Monitor Traces are $\mssafe$]\label{lem:mon:msafe}
+  If $\monmssafe{\trace}$, then $\trace\in\mssafe$.\Coqed
+\end{lemma}
 
@@
   \begin{aligned}
+    \mi{(Abstract\ Store)}~\scctmonitor&:=\ \emptyset \\
+    \mi{(Abstract\ Events)}~\absevent&:=\ \absev{\varepsilon} \mid \absterm \mid \absev{Any} \\
   \end{aligned}
 \end{gather*}
+\begin{center}
+  \judgbox{\monitorcheck[]{\scctmonitor}{\scctmonitor[']}{\absevent}}{,,Monitor $\scctmonitor$ steps to $\scctmonitor[']$ given event $\absevent$.''}$\;$\\
+  %
+  \typerule{scct-none}{
+  }{
+    \monitorcheck[]{\scctmonitor}{\scctmonitor}{\absev{\emptyevent}}
+  }{scct-none}
+  %
+  \typerule{scct-abort}{
+  }{
+    \monitorcheck[]{\scctmonitor}{\scctmonitor}{\absterm}
+  }{scct-abort}
+\end{center}
+\begin{center}
+  \judgbox{\sccttraceagree[]{\event}{\absevent}}{,,Abstract event $\absevent$ is equivalent to $\event$ with respect to \gls{cct}.''}$\;$\\
+  %
+  \typerule{scct-low-authentic}{
+  }{
+    \sccttraceagree[]{\event[_b];\sandboxtag;\unlock}{\absev{\emptyevent}}
+  }{scct-low-authentic}
+  %
+  \typerule{scct-high-authentic}{
+  }{
+    \sccttraceagree[]{\event[_b];\sandboxtag;\lock}{\absev{Any}}
+  }{scct-high-authentic}
+  %
+  \typerule{scct-empty-authentic}{
+  }{
+    \sccttraceagree[]{\emptyevent}{\absev{\emptyevent}}
+  }{scct-empty-authentic}
+  %
+  \typerule{scct-abort-authentic}{
+  }{
+    \sccttraceagree[]{\lightning}{\absterm}
+  }{scct-abort-authentic}
+\end{center}
+
+The \gls{scct} monitor simply disregards all private data occuring on a trace.
 

### 2023-06-13 popl/main.tex
Add todo
Minor patching to 1st contrib

@@
 \MKin{
-  ,,can be proven'' -> we prove (?)
+  ,,can be proven'' -> is proven (?)
 }
@@
 \begin{itemize}
-  \item
-        \Cref{sec:compprop} presents a formal framework to reason about compositions of security properties and demonstrate its usage with concrete examples:
-        Given a most general trace model, we define the trace-based security properties \gls{tms}, \gls{sms}, \gls{ms}, \gls{cct}, and their compositions.
-        For each of these properties, we also give a monitor which runs on another trace model that excludes unneccessary information and show that their composition is behaviorally similar to the composition of the trace-based security properties.
-        This two-level approach enhances modularity: When working with the security properties, one only needs to reason about the corresponding monitor.
+  \item %
+        %\Cref{sec:compprop} presents a formal framework to reason about compositions of security properties and demonstrate its usage with concrete examples:
+        A formal framework to reason about the composition of security properties (\Cref{sec:compprop}).
+        %Given a most general trace model, we define the trace-based security properties \gls{tms}, \gls{sms}, \gls{ms}, \gls{cct}, and their compositions.
+        Using the framework, the ease for formal reasoning about security property compositions is demonstrated by instantiating it on \gls{tms}, \gls{sms}, and \gls{cct}, which are ultimately composed to \gls{ms} and a combination of \gls{ms} and \gls{cct}.
+        %For each of these properties, we also give a monitor which runs on another trace model that excludes unneccessary information and show that their composition is behaviorally similar to the composition of the trace-based security properties.
+        %This two-level approach enhances modularity: When working with the security properties, one only needs to reason about the corresponding monitor.
+        %%% let's omit above sentences completely. why ramble about monitors in the introduction? this ain't about model checking
 
@@
 
+\MKin{combined version of all monitors? maybe just as lemma}
+
 \section{Composing Robustly Secure Compilers}\label{sec:compcomp}

### 2023-06-14 popl/main.tex
Contribs and disclaimer

@@
 \begin{center}\small\it
-	{In the following, we use syntax highlighting accessible to both colourblind and black \& white readers~\citep{patrignani2020use}. 
-	Specifically, we use a $\src{blue}$, $\src{sans\text{-}serif}$ font for our $\src{source}$,
-	a $\trg{red}$, $\trg{bold}$ font for our $\trg{intermediate}$, 
-	and a $\obj{green}$, $\obj{teletype}$ font for our $\obj{target}$ language.
+	{This paper uses syntax highlighting accessible to both colourblind and black \& white readers~\citep{patrignani2020use}.
+	Specifically, it makes use of a $\src{blue}$, $\src{sans\text{-}serif}$ font for a $\src{source}$,
+	a $\trg{red}$, $\trg{bold}$ font for an $\trg{intermediate}$,
+	and a $\obj{green}$, $\obj{teletype}$ font for a $\obj{target}$ language.
 	}
@@
         %Given a most general trace model, we define the trace-based security properties \gls{tms}, \gls{sms}, \gls{ms}, \gls{cct}, and their compositions.
-        Using the framework, the ease for formal reasoning about security property compositions is demonstrated by instantiating it on \gls{tms}, \gls{sms}, and \gls{cct}, which are ultimately composed to \gls{ms} and a combination of \gls{ms} and \gls{cct}.
+        Using the framework, the ease for formal reasoning with compositions of security properties is demonstrated by instantiating the framework on \gls{tms}, \gls{sms}, and \gls{cct}, which are ultimately composed to \gls{ms} and a combination of \gls{ms} and \gls{cct}.
         %For each of these properties, we also give a monitor which runs on another trace model that excludes unneccessary information and show that their composition is behaviorally similar to the composition of the trace-based security properties.
@@
 
-  \item
-        \Cref{sec:compcomp} studies different forms of secure compiler compositions.
-        Concretely, it investigates upper and lower compositions, as well as the sequential composition.
-        A simple corollary of sequential composition is the ability to freely reorder compiler optimizations without compromising security.
+  \item %
+        %\Cref{sec:compcomp} studies different forms of secure compiler compositions.
+        Foundational, theoretical work on different forms of compositions of secure compilers (\Cref{sec:compcomp}).
+        %Concretely, it investigates upper and lower compositions, as well as the sequential composition.
+        First, this paper proves theorems which state that a compiler with multiple input or output languages is secure as long as the compiler is composed of several secure passes with one input and one output language, respectively.
+        Second, a theorem about the sequential composition of secure compilers is proven, which states that sequentially composing secure compilers is, again, secure with respect to the composition of the properties they preserve.
+        %A simple corollary of sequential composition is the ability to freely reorder compiler optimizations without compromising security.
 
-  \item \Cref{sec:casestud} sets up a case-study to demonstrate the power of our formal framework.
-        We define several different languages and compilers between them and prove that the compilers are robustly preserving different properties.
-        The sequential composition of these compilers yields a compilation chain that is optimizing and robustly preserving a combination of \gls{ms} and \gls{cct}.
-        As for optimizations, the chain has two passes, one performing \gls{cf} and the other \gls{dce}.
-        This compilation chain demonstrates the ability to perform modular proof engineering by leaveraging this paper's compositionality theorems.
+  \item %
+        % \Cref{sec:casestud} sets up a case-study to demonstrate the power of our formal framework.
+        A case-study demonstrating the power of our formal framework (\Cref{sec:casestud}).
+        %We define several different languages and compilers between them and prove that the compilers are robustly preserving different properties.
+        To this end, a compilation chain of several passes is presented.
+        %The sequential composition of these compilers yields a compilation chain that is optimizing and robustly preserving a combination of \gls{ms} and \gls{cct}.
+        The chain attains a combination of \gls{ms} and \gls{cct} by means of applying the theoretical work from the previous contribution on the individual, secure passes concerning \gls{tms}, \gls{sms}, and \gls{scct}, respectively.
+        %As for optimizations, the chain has two passes, one performing \gls{cf} and the other \gls{dce}.
+        Furthermore, the chain entails two passes for \gls{ms}\MK{maybe patch when optimizing passes are done} that are optimizing: One performs a simple \gls{dce} and the other \gls{cf}.
+        %This compilation chain demonstrates the ability to perform modular proof engineering by leaveraging this paper's compositionality theorems.
+        This demonstrates that the usage of the presented framework is both simpler and more expressive compared to one monolithic security proof of the whole compilation chain.
 
   \item Our key contributions are formalized in the Coq proof assistant.
-        The formalization for the compilers are \emph{extractable}.
-        That is, one can translate our Coq code to Ocaml and run them.
+        The practically relevant parts of the formalization are \emph{extractable}.
+        That is, one can translate the Coq code to Ocaml and run the compilation chain from the previous contribution.
         Finally, in this paper, whenever definitions or proofs are done in Coq, we use the symbol \CoqSymbol.
@@
 
-We now describe the other sections of the paper.
-In \Cref{sec:background}, the paper establishes the notions of traces, (hyper-)properties, and robust preservation, while \Cref{sec:relwork} compares this work with others.
-We conclude in \Cref{sec:concl} and give an outlook for future work.
+%In \Cref{sec:background}, the paper establishes the notions of traces, (hyper-)properties, and robust preservation, while \Cref{sec:relwork} compares this work with others.
+For the other parts of the paper, the notions of security properties and secure compilers are introduced (\Cref{sec:background}).
+%We conclude in \Cref{sec:concl} and give an outlook for future work.
+The work presented in this paper is compared to others from the literature (\Cref{sec:relwork}) and there is a conclusion with an outlook for future work (\Cref{sec:concl}).
 
-\contribution{Open Source \& Technical Report} Our technical report and the Coq formalisation are
-available as supplementary material and will be made available upon acceptance of this paper.
+\contribution{Open Source \& Technical Report} Our technical report and the Coq formalisation are available as supplementary material.\MKin{cite/link}
 

### 2023-06-14 popl/main.tex
Fix floating error

@@
         %As for optimizations, the chain has two passes, one performing \gls{cf} and the other \gls{dce}.
-        Furthermore, the chain entails two passes for \gls{ms}\MK{maybe patch when optimizing passes are done} that are optimizing: One performs a simple \gls{dce} and the other \gls{cf}.
+        Furthermore, the chain entails two passes for \gls{ms}\MKin{maybe patch \gls{ms} when optimizing passes are done} that are optimizing: One performs a simple \gls{dce} and the other \gls{cf}.
         %This compilation chain demonstrates the ability to perform modular proof engineering by leaveraging this paper's compositionality theorems.

### 2023-06-14 popl/main.tex
Wording
Get rid of we, our, us

@@
 }
-Crucially, this end-to-end security preservation is obtained by having each compilation pass (i) preserve a sub-part of the overall security property individually and then (ii) compose each preserved security property as dictated by our framework.
-Our results showcase how our framework allows the kind of formal security reasoning that compiler writers already want (and do), obtaining precise, compositional security reasoning while providing minimal (and modular) proof effort.
+Crucially, this end-to-end security preservation is obtained by having each compilation pass (i) preserve a sub-part of the overall security property individually and then (ii) compose each preserved security property as dictated by the framework of this paper.
+The results showcase how the framework allows the kind of formal security reasoning that compiler writers already want (and do), obtaining precise, compositional security reasoning while providing minimal (and modular) proof effort.
 \MPin{
@@
         % \Cref{sec:casestud} sets up a case-study to demonstrate the power of our formal framework.
-        A case-study demonstrating the power of our formal framework (\Cref{sec:casestud}).
+        A case-study demonstrating the power of the formal frameworks from the previous contributions (\Cref{sec:casestud}).
         %We define several different languages and compilers between them and prove that the compilers are robustly preserving different properties.
@@
 
-  \item Our key contributions are formalized in the Coq proof assistant.
+  \item The key contributions are formalized in the Coq proof assistant.
         The practically relevant parts of the formalization are \emph{extractable}.
         That is, one can translate the Coq code to Ocaml and run the compilation chain from the previous contribution.
-        Finally, in this paper, whenever definitions or proofs are done in Coq, we use the symbol \CoqSymbol.
+        Finally, this paper indicates with \CoqSymbol whenever definitions or proofs are done in Coq.
 \end{itemize}}
@@
 
-\contribution{Open Source \& Technical Report} Our technical report and the Coq formalisation are available as supplementary material.\MKin{cite/link}
+\contribution{Open Source \& Technical Report} A technical report and the Coq formalisation are available as supplementary material.\MKin{cite/link}
 
@@
 A sequence of events $\trace$ is a \emph{trace}.
-So, throughout the paper, we use $\overline{\cdot}$ to mark meta-variables as sequences.
+So, throughout the paper, $\overline{\cdot}$ is used to mark meta-variables as sequences.
 Empty sequences are denoted with $\hole{\cdot}$ and concatenation of traces $\trace[_{1}],\trace[_{2}]$ is written as $\trace[_{1}]\cdot\trace[_{2}]$.
-For sake of brevity, we use this notation also for pre- and appending events to traces.
+For sake of brevity, this notation also used for pre- and appending events to traces.
 
@@
 The \emph{lifting} of a property $\pi$ is simply its powerset and denoted as $\lift{\pi}$.
-To define property satisfaction, we first look at programming languages.
+To define property satisfaction it is integral to first look at programming languages.
 
-For languages used in this paper, we consider a small-step style semantics.
+For languages used in this paper, a small-step style semantics is used.
 That is, there is a step relation between states $\statevar$ and an event $\event$: $\stepto[]{\statevar[_{1}]}{\statevar[_{2}]}{\event}$.
-Also, a program's syntax $p$ is encoded in the state $\statevar$ which we write as $p\in\statevar$.
-For states $\statevar$, we assume there exists a judgement $\isval{\statevar}$ that determines wether $\statevar$ is a final state or not.
+Also, a program's syntax $p$ is encoded in the state $\statevar$ which is written as $p\in\statevar$.
+For states $\statevar$, it is assumed that there exists a judgement $\isval{\statevar}$ that determines wether $\statevar$ is a final state or not.
 \Cref{fig:starstep:and:nstep}\MK{appendix?} defines the star-step relation $\stepto{\statevar[_{1}]}{\statevar[_{2}]}{\trace}$, essentially modelling a big-step style evaluation, and the $n$-step relation $\stepton{\statevar[_{1}]}{\statevar[_{2}]}{\trace}$ that evaluates a program for $n$ steps.
@@
 Note that for deterministic languages, $\behav{w}$ is a singleton set for any $w$.
-With this, we can define satisfaction:
+The behavior of programs gives rise to the definition of satisfaction:
 
@@
 Providing definitions for symbols is usually the job of the \emph{linker}.
-We denote linking two programs $\progvar[_{1}],\progvar[_{2}]$ as $\link{\progvar[_{1}]}{\progvar[_{2}]}$.
-For sake of simplicity, we assume that linking does not fail.
-With linking, we can define robust satisfaction analogously to \Cref{def:propsat,def:classsat}:
+Notationally, linking two programs $\progvar[_{1}],\progvar[_{2}]$ is written $\link{\progvar[_{1}]}{\progvar[_{2}]}$.
+For sake of simplicity, linking does not fail.
+With linking, it is possible to define robust satisfaction analogously to \Cref{def:propsat,def:classsat}:
 
 \begin{definition}[Robust Property Satisfaction]\label{def:proprsat}
-  A program $\progvar$ robustly satisfies $\pi$ ($\rsat{\progvar}{\pi}$) iff for any program $\contextvar$ we have $\sat{\link{\contextvar}{\progvar}}{\pi}$. (similarily for $\rsat{\progvar}{\class}$)
+  A program $\progvar$ robustly satisfies $\pi$ ($\rsat{\progvar}{\pi}$) iff for any program $\contextvar$ it holds that $\sat{\link{\contextvar}{\progvar}}{\pi}$.
+(similarily for $\rsat{\progvar}{\class}$)
 \end{definition}
@@
 
-We expect compilers to preserve robust satisfaction.
+It is expected that compilers preserve robust satisfaction.
 In this paper, compilers are modelled as function from one programming language's syntax to the other.
-Since this paper considers composing compilers, we'll adopt notation similar to tombstone diagrams to easen the formal burden.
-So, we denote a compiler from $\src{L}$ to $\trg{L}$ as $\cc{\src{L}\to\trg{L}}{\src{\bullet}}$ and we refer to $\src{L}$ as \emph{source} language and to $\trg{L}$ as \emph{target} language.
+Since this paper considers composing compilers, we'll adopt notation similar to tombstone diagrams\MK{cite} to easen the formal burden.
+So, denote a compiler from $\src{L}$ to $\trg{L}$ as $\cc{\src{L}\to\trg{L}}{\src{\bullet}}$ and refer to $\src{L}$ as \emph{source} language and to $\trg{L}$ as \emph{target} language.
 Composing $\cc{\src{L}\to\trg{L}}{\src{\bullet}}$ with $\cc{\trg{L}\to\obj{L}}{\trg{\bullet}}$ is denoted as $\cc{\src{L}\to\trg{L}\to\obj{L}}{\src{\bullet}}$ and means first running $\cc{\src{L}\to\trg{L}}{\src{\bullet}}$ and plugging the output into $\cc{\trg{L}\to\obj{L}}{\trg{\bullet}}$.
-Contrary to tombstones, we don't consider a compiler's implementation language.
-We now define the compiler-property robust trace-property preservation.
+Contrary to tombstones, the implementation language of a compiler is disregarded.
+Now, define the compiler-property robust trace-property preservation.
 
@@
 \section{Composing Properties}\label{sec:compprop}
-\Cref{subsec:propdefs} introduces a ,,most general'' trace model and defines properties (\gls{tms}, \gls{sms}, \gls{ms}, and \gls{cct}) that our case study (\Cref{sec:casestud}) uses.
+\Cref{subsec:propdefs} introduces a ,,most general'' trace model and defines properties (\gls{tms}, \gls{sms}, \gls{ms}, and \gls{cct}) that the case study (\Cref{sec:casestud}) uses.
 By means of taking these properties as an example, \Cref{subsec:monitors} defines monitors for them and an agreement relation between a monitor's trace model with the most general trace model.
-Then, we define monitor-based satisfaction and we prove that trace-property satisfaction is implied by it.
-Since monitor-based satisfaction is easier to reason about, this makes our later proofs considerably simpler.
+Then, monitor-based satisfaction is introduced and it is proven that trace-property satisfaction is implied by it.
+Since monitor-based satisfaction is easier to reason about, this simplifies later proofs considerably.
 
@@
 Locations $\loc$ are elements of $L$ and left abstract.
-However, we expect $L$ to be closed over addition with natural numbers.
+However, it is expected that $L$ is closed over addition with natural numbers.
 $\ev{Dealloc\ \loc}$ announces that location $\loc$ is freed.
@@
 
-We use this trace model to define any property that interests us.
-In a sense the trace model is most general, because it abstracts all events done in our languages as defined in \Cref{sec:casestud}.
-We write $\event[_{1}]\le_{\trace}\event[_{2}]$ if $\event[_{1}]$ occurs at an earlier position in $\trace$ than $\event[_{2}]$.
+This trace model is used to define any property of interest.
+In a sense, the trace model is most general, because it abstracts all events done in the languages as defined in \Cref{sec:casestud}.
+This papers uses $\event[_{1}]\le_{\trace}\event[_{2}]$ if $\event[_{1}]$ occurs at an earlier position in $\trace$ than $\event[_{2}]$.
 
@@
 
-We define a variant of the \gls{cct} security property.
+A variant of the \gls{cct} security property is defined.
 Generally, one cannot define \gls{cct} just as trace-property without, e.g., performing a self-composition of a program.
-That's why we adopt for a stricter variant of \gls{cct}, called \gls{scct}, that enforces common policies to ensure code is \gls{cct}.
+That is why this paper adopts a stricter variant of \gls{cct}, called \gls{scct}, that enforces common policies to ensure code is \gls{cct}.
 For example, one must not branch on secrets or perform binary operations, such as division, on secrets whose timing is data-dependent.
-We opt for a definition that disregards any secrets leaked on a trace.
-To define the property, we need a function on traces that strips all events with a $\lock$-tag: $\downarrow_{\lock}(\cdot)$.
+Even stronger, \gls{scct} disregards any secrets leaked on a trace.
+To define the property, a function on traces is needed that strips all events with a $\lock$-tag: $\downarrow_{\lock}(\cdot)$.
 
@@
 
-This subsection presents monitors for our properties of interest (\gls{tms}, \gls{sms}, \gls{ms}, \gls{scct}).
-Each monitor $\monitor$ has its own accompanying trace model that abstracts over our most general trace model.
+This subsection presents monitors for properties \gls{tms}, \gls{sms}, \gls{ms}, and \gls{scct}.
+Each monitor $\monitor$ has its own accompanying trace model that abstracts over the most general trace model.
 %This is because monitors should only be concerned with property-relevant information.
@@
 This section defines several secure compilers, each of which robustly preserves a different property of interest.
-We demonstrate the power of our framework from \Cref{sec:compprop,sec:compcomp} by composing these compilers for a secure, but optimizing, compilation chain that robustly preserves \gls{ms} and \gls{cct}.
+The power of the frameworks from \Cref{sec:compprop,sec:compcomp} is demonstrated by composing these compilers for a secure, but optimizing, compilation chain that robustly preserves \gls{ms} and \gls{cct}.
 Concretely, \Cref{subsec:cs:tms} starts by defining languages $\src{L_{TMS}}$ and $\trg{L}$ and describes a compiler $\cc{\src{L_{TMS}}\to\trg{L}}{\src{\bullet}}$ for the robust preservation of \gls{tms} between the two languages.
 \Cref{subsec:cs:ms} defines $\irl{L_{MS}}$ and a compiler $\cc{\trg{L}\to\irl{L_{MS}}}{\trg{\bullet}}$ that robustly preserves \gls{sms}.
-By the framework presented in \Cref{sec:compcomp,sec:compprop}, we attain a compiler $\cc{\src{L_{TMS}}\to\irl{L_{MS}}}{\src{\bullet}}$ for the robust preservation of \gls{ms} by composing $\cc{\src{L_{TMS}}\to\trg{L}}{\src{\bullet}}$ and $\cc{\trg{L}\to\irl{L_{MS}}}{\trg{\bullet}}$.
+By the framework presented in \Cref{sec:compcomp,sec:compprop}, a compiler $\cc{\src{L_{TMS}}\to\irl{L_{MS}}}{\src{\bullet}}$ for the robust preservation of \gls{ms} by composing $\cc{\src{L_{TMS}}\to\trg{L}}{\src{\bullet}}$ and $\cc{\trg{L}\to\irl{L_{MS}}}{\trg{\bullet}}$ is attained.
 \Cref{subsec:cs:scct} defines $\obj{L_{CCT}}$ and a compiler $\cc{\irl{L_{MS}}\to\obj{L_{CCT}}}{\irl{\bullet}}$ that robustly preserves \gls{cct}.
 Composing it with $\cc{\src{L_{TMS}}\to\irl{L_{MS}}}{\src{\bullet}}$ yields $\cc{\src{L_{TMS}}\to\obj{L_{CCT}}}{\src{\bullet}}$, a compiler that robustly preserves a combination of \gls{ms} and \gls{cct}.
-We attain our final compilation chain in \Cref{subsec:cs:opts}, which defines optimizing compilers performing either \gls{cf} or \gls{dce} and composes them with $\cc{\src{L_{TMS}}\to\obj{L_{CCT}}}{\src{\bullet}}$, giving us an optimizing, secure compilation chain.
+The final compilation chain in \Cref{subsec:cs:opts} defines optimizing compilers performing either \gls{cf} or \gls{dce} and composes them with $\cc{\src{L_{TMS}}\to\obj{L_{CCT}}}{\src{\bullet}}$, giving an optimizing, secure compilation chain.
 

### 2023-06-14 popl/main.tex
Compiler notation

@@
 
+\begin{lemma}[Monitor Traces are $\scctsafe$]\label{lem:mon:scctsafe}
+  If $\monscctsafe{\trace}$, then $\trace\in\scctsafe$.\Coqed
+\end{lemma}
+
 \MKin{combined version of all monitors? maybe just as lemma}
 
-\section{Composing Robustly Secure Compilers}\label{sec:compcomp}
+\section{Composing Secure Compilers}\label{sec:compcomp}
 This section studies different kinds of secure compiler compositions and their corresponding proofs.
@@
 Moreover, it demonstrates that the order does not matter for composing compilers with the same source and target language.
+Throughout this section, let $\src{L},\trg{L},$ and $\obj{L}$ be arbitrary programming lanugages.
+
+\begin{definition}[Sequential Composition]\label{def:seqcomp}
+  Let $\cc{\src{L}\to\trg{L}}{\src{\bullet}}$, $\cc{\trg{L}\to\obj{L}}{\trg{\bullet}}$ be compilers.
+  Let $\src{p}$ be any $\src{L}$ program, $\trg{p}=\cc{\src{L}\to\trg{L}}{\src{p}}$, and $\obj{p}=\cc{\trg{L}\to\obj{L}}{\trg{p}}$.
+  The sequential composition is $\cc{\src{L}\to\trg{L}\to\obj{L}}{\src{p}}=\obj{p}$.
+\end{definition}
+
+\begin{theorem}[Sequential Composition of Secure Compilers]\label{thm:rtp}
+  Given classes of properties $\class[_{1}],\class[_{2}]$, let $\rtp{\cc{\src{L}\to\trg{L}}{\src{\bullet}}}{\class[_{1}]}$ and $\rtp{\cc{\trg{L}\to\obj{L}}{\trg{\bullet}}}{\class[_{2}]}$.
+  Then, $\rtp{\cc{\src{L}\to\trg{L}\to\obj{L}}{\src{\bullet}}}{\class[_{1}]\cap\class[_{2}]}$. \Coqed
+\end{theorem}
+
+A corollary of \Cref{thm:rtp} is that optimization passes that have the input and output language can be freely swapped:
+\begin{corollary}[Sequential Composition of Secure Compilers]\label{corr:swappable}
+  Given classes of properties $\class[_{1}],\class[_{2}]$, let $\rtp{\cc{\trg{L}\xrightarrow{1}\trg{L}}{\trg{\bullet}}}{\class[_{1}]}$ and $\rtp{\cc{\trg{L}\xrightarrow{2}\trg{L}}{\trg{\bullet}}}{\class[_{2}]}$.
+  Then, $\rtp{\cc{\trg{L}\xrightarrow{1}\trg{L}\xrightarrow{2}\trg{L}}{\trg{\bullet}}}{\class[_{1}]\cap\class[_{2}]}$ and $\rtp{\cc{\trg{L}\xrightarrow{2}\trg{L}\xrightarrow{1}\trg{L}}{\trg{\bullet}}}{\class[_{2}]\cap\class[_{1}]}$.
+\Coqed
+\end{corollary}
 
@@
 The power of the frameworks from \Cref{sec:compprop,sec:compcomp} is demonstrated by composing these compilers for a secure, but optimizing, compilation chain that robustly preserves \gls{ms} and \gls{cct}.
-Concretely, \Cref{subsec:cs:tms} starts by defining languages $\src{L_{TMS}}$ and $\trg{L}$ and describes a compiler $\cc{\src{L_{TMS}}\to\trg{L}}{\src{\bullet}}$ for the robust preservation of \gls{tms} between the two languages.
-\Cref{subsec:cs:ms} defines $\irl{L_{MS}}$ and a compiler $\cc{\trg{L}\to\irl{L_{MS}}}{\trg{\bullet}}$ that robustly preserves \gls{sms}.
-By the framework presented in \Cref{sec:compcomp,sec:compprop}, a compiler $\cc{\src{L_{TMS}}\to\irl{L_{MS}}}{\src{\bullet}}$ for the robust preservation of \gls{ms} by composing $\cc{\src{L_{TMS}}\to\trg{L}}{\src{\bullet}}$ and $\cc{\trg{L}\to\irl{L_{MS}}}{\trg{\bullet}}$ is attained.
-\Cref{subsec:cs:scct} defines $\obj{L_{CCT}}$ and a compiler $\cc{\irl{L_{MS}}\to\obj{L_{CCT}}}{\irl{\bullet}}$ that robustly preserves \gls{cct}.
-Composing it with $\cc{\src{L_{TMS}}\to\irl{L_{MS}}}{\src{\bullet}}$ yields $\cc{\src{L_{TMS}}\to\obj{L_{CCT}}}{\src{\bullet}}$, a compiler that robustly preserves a combination of \gls{ms} and \gls{cct}.
-The final compilation chain in \Cref{subsec:cs:opts} defines optimizing compilers performing either \gls{cf} or \gls{dce} and composes them with $\cc{\src{L_{TMS}}\to\obj{L_{CCT}}}{\src{\bullet}}$, giving an optimizing, secure compilation chain.
+Concretely, the section starts starts by defining languages $\src{L_{TMS}}$ and $\trg{L}$ and describes a compiler $\cc{\src{L_{TMS}}\to\trg{L}}{\src{\bullet}}$ for the robust preservation of \gls{tms} between the two languages (\Cref{subsec:cs:tms}).
+Then, it defines $\irl{L_{MS}}$ and a compiler $\cc{\trg{L}\to\irl{L_{MS}}}{\trg{\bullet}}$ that robustly preserves \gls{sms} (\Cref{subsec:cs:ms}).
+By the framework presented in \Cref{sec:compcomp,sec:compprop}, a compiler $\cc{\src{L_{TMS}}\to\trg{L}\to\irl{L_{MS}}}{\src{\bullet}}$ for the robust preservation of \gls{ms} by composing $\cc{\src{L_{TMS}}\to\trg{L}}{\src{\bullet}}$ and $\cc{\trg{L}\to\irl{L_{MS}}}{\trg{\bullet}}$ is attained.
+Two secure compilers $\cc{{\irl{L_{MS}}\xrightarrow{\gls{dce}}\irl{L_{MS}}}}{\irl{\bullet}}$ and $\cc{\irl{L_{MS}}\xrightarrow{\gls{cf}}\irl{L_{MS}}}{\irl{\bullet}}$ are defined that perform \gls{dce} and \gls{cf}, respectively, to model an optimzing pipeline happening somewhere in the compilation chain (\Cref{subsec:cs:opts}).
+Finally, the paper defines $\obj{L_{CCT}}$ and a compiler $\cc{\irl{L_{MS}}\to\obj{L_{CCT}}}{\irl{\bullet}}$ that robustly preserves \gls{cct} (\Cref{subsec:cs:scct}).
+Composing it with $\cc{\src{L_{TMS}}\to\irl{L_{MS}}}{\src{\bullet}}$ and the optimizing compilers $\cc{\irl{L_{MS}}\xrightarrow{\gls{dce}}\irl{L_{MS}}}{\irl{\bullet}}$ and $\cc{\irl{L_{MS}}\xrightarrow{\gls{cf}}\irl{L_{MS}}}{\irl{\bullet}}$ yields $\cc{\src{L_{TMS}}\to\trg{L}\to\irl{L_{MS}}\xrightarrow{\gls{cf}}\irl{L_{MS}}\xrightarrow{\gls{dce}}\irl{L_{MS}}\to\obj{L_{CCT}}}{\src{\bullet}}$, a compiler that robustly preserves a combination of \gls{ms} and \gls{cct}.
 
@@
 \subsection{Robust (Spatial) Memory Safety Preservation}\label{subsec:cs:ms}
+\subsection{Optimizing Compilers}\label{subsec:cs:opts}
 \subsection{Robust Strict Cryptographic Constant Time Preservation}\label{subsec:cs:scct}
-\subsection{Optimizing Compilers}\label{subsec:cs:opts}
 

### 2023-06-14 popl/main.tex
Type fix in def of mon
Typo fix

@@
 \begin{definition}[\glsfirst{tms}]\label{def:trace:tmsdef}
-  A trace $\trace$ is temporal memory safe ($\trace\in\tmssafe$) iff $\ev{Alloc\ \loc\ n;\sandboxtag;\securitytag}\le_{\trace}\ev{Dealloc\ \loc;\sandboxtag';\securitytag'}$, $\ev{Use\ \loc\ n;\sandboxtag;\securitytag}\not\le_{\trace}\ev{Alloc\ \loc\ m;\sandboxtag';\securitytag'}$ and $\ev{Use\ \loc\ n;\sandboxtag;\securitytag}\not\le_{\trace}\ev{Dealloc\ \loc;\sandboxtag';\securitytag'}$.
+  A trace $\trace$ is temporal memory safe ($\trace\in\tmssafe$) iff $\ev{Alloc\ \loc\ n;\sandboxtag;\securitytag}\le_{\trace}\ev{Dealloc\ \loc;\sandboxtag';\securitytag'}$, $\ev{Use\ \loc\ n;\sandboxtag;\securitytag}\not\le_{\trace}\ev{Alloc\ \loc\ m;\sandboxtag';\securitytag'}$ and $\ev{Dealloc\ \loc;\sandboxtag;\securitytag}\not\le_{\trace}\ev{Use\ \loc\ n;\sandboxtag';\securitytag'}$.
 \end{definition}
@@
   \typerule{tms-dealloc}{
-    \loc\notin\tmsmonitor.\text{alloced} &
+    \loc\in\tmsmonitor.\text{alloced} &
     \loc\notin\tmsmonitor.\text{freed} \\

### 2023-06-14 popl/main.tex
Typo fix

@@
 \begin{definition}[\glsfirst{sms}]\label{def:trace:smsdef}
-  A trace $\trace$ is spatial memory safe ($\trace\in\smssafe$) iff given $\ev{Alloc\ \loc\ n;\sandboxtag;\securitytag}\le_{\trace}\ev{Use\ \loc\ m;\sandboxtag';\securitytag'}$, then $n<m$.
+  A trace $\trace$ is spatial memory safe ($\trace\in\smssafe$) iff given $\ev{Alloc\ \loc\ m;\sandboxtag;\securitytag}\le_{\trace}\ev{Use\ \loc\ n;\sandboxtag';\securitytag'}$, then $n<m$.
 \end{definition}

### 2023-06-15 popl/main.tex
contrib

@@
 
-\BrText{Validation}{In summary, this paper makes the following contributions:
+% \BrText{Validation}{
+In summary, this paper makes the following contributions:
 
@@
         %\Cref{sec:compprop} presents a formal framework to reason about compositions of security properties and demonstrate its usage with concrete examples:
-        A formal framework to reason about the composition of security properties (\Cref{sec:compprop}).
-        %Given a most general trace model, we define the trace-based security properties \gls{tms}, \gls{sms}, \gls{ms}, \gls{cct}, and their compositions.
-        Using the framework, the ease for formal reasoning with compositions of security properties is demonstrated by instantiating the framework on \gls{tms}, \gls{sms}, and \gls{cct}, which are ultimately composed to \gls{ms} and a combination of \gls{ms} and \gls{cct}.
+        First, this paper studies the formalisation and the composition of security properties (\Cref{sec:compprop}) by focussing on those safety properties that are of interest for real-world compiler writers (as identified by the plethora of work enforcing such properties individually~\MP{cits}).
+        % 
+        Starting from ways to formalise and enforce those properties individually, this paper shows how to compose both their formalisation and their composition.
+        % 
+        This paper then showcases the benefits of this composition by composing \gls{tms} and \gls{sms} into \gls{ms}, and ultimately adding \gls{cct} to the composition, yielding the the conjunction of \gls{ms} and \gls{cct}.
+        % 
+        The resulting security property is the golden standard of safety properties that secure cryptographic code should have~\MP{cits}.
+        \MPin{
+        	conclude that did this prop in a single pass, we show how to do it compositionally
+        }
         %For each of these properties, we also give a monitor which runs on another trace model that excludes unneccessary information and show that their composition is behaviorally similar to the composition of the trace-based security properties.
@@
         %\Cref{sec:compcomp} studies different forms of secure compiler compositions.
-        Foundational, theoretical work on different forms of compositions of secure compilers (\Cref{sec:compcomp}).
+        This paper takes the secure compilation framework of~\citep{abate2019jour} and extends it to reason about the security of all different known forms of compiler composition (\Cref{sec:compcomp}).
         %Concretely, it investigates upper and lower compositions, as well as the sequential composition.
-        First, this paper proves theorems which state that a compiler with multiple input or output languages is secure as long as the compiler is composed of several secure passes with one input and one output language, respectively.
-        Second, a theorem about the sequential composition of secure compilers is proven, which states that sequentially composing secure compilers is, again, secure with respect to the composition of the properties they preserve.
-        %A simple corollary of sequential composition is the ability to freely reorder compiler optimizations without compromising security.
+        For this, this paper studies \MP{}.
+        % 
+        This paper proves that starting from two compilers that preserve two properties, their composition preserves the intersection of those properties.
+        \MPin{
+       		find way to connect to previous contrib
+        }
+        % 
+        Finally, this paper provides a simple but crucial corollary that demonstrates that the order of composition of sequential compiler passes is irrelevant for the resulting security.
+        % 
+        This corollary is crucial for reordering compiler optimisation passes as well as secure compilation passes and thus generating secure and efficient code.
 
@@
         % \Cref{sec:casestud} sets up a case-study to demonstrate the power of our formal framework.
-        A case-study demonstrating the power of the formal frameworks from the previous contributions (\Cref{sec:casestud}).
+        This paper presents a case-study showcasing the conjunction of the previous contributions (\Cref{sec:casestud}).
         %We define several different languages and compilers between them and prove that the compilers are robustly preserving different properties.
-        To this end, a compilation chain of several passes is presented.
+        To this end, it presents a compilation chain consisting of several passes.
         %The sequential composition of these compilers yields a compilation chain that is optimizing and robustly preserving a combination of \gls{ms} and \gls{cct}.
-        The chain attains a combination of \gls{ms} and \gls{cct} by means of applying the theoretical work from the previous contribution on the individual, secure passes concerning \gls{tms}, \gls{sms}, and \gls{scct}, respectively.
+        The chain ultimately preserves a combination of \gls{ms} and \gls{cct} by means of composing the individual, secure passes concerning \gls{tms}, \gls{sms}, and \gls{scct}, respectively.
         %As for optimizations, the chain has two passes, one performing \gls{cf} and the other \gls{dce}.
-        Furthermore, the chain entails two passes for \gls{ms}\MKin{maybe patch \gls{ms} when optimizing passes are done} that are optimizing: One performs a simple \gls{dce} and the other \gls{cf}.
+        Furthermore, the chain includes two passes for \gls{ms}\MKin{maybe patch \gls{ms} when optimizing passes are done} that are optimizing: One performs a simple \gls{dce} and the other \gls{cf}.
         %This compilation chain demonstrates the ability to perform modular proof engineering by leaveraging this paper's compositionality theorems.
         This demonstrates that the usage of the presented framework is both simpler and more expressive compared to one monolithic security proof of the whole compilation chain.
+        \MP{ matthis please write your idea here }
 
@@
         Finally, this paper indicates with \CoqSymbol whenever definitions or proofs are done in Coq.
-\end{itemize}}
+\end{itemize}
+% }
 
 %In \Cref{sec:background}, the paper establishes the notions of traces, (hyper-)properties, and robust preservation, while \Cref{sec:relwork} compares this work with others.
-For the other parts of the paper, the notions of security properties and secure compilers are introduced (\Cref{sec:background}).
+This paper starts by introducing relevant notions of security properties and secure compilation (\Cref{sec:background}),
 %We conclude in \Cref{sec:concl} and give an outlook for future work.
-The work presented in this paper is compared to others from the literature (\Cref{sec:relwork}) and there is a conclusion with an outlook for future work (\Cref{sec:concl}).
+and discusses related work (\Cref{sec:relwork}) before concluding (\Cref{sec:concl}).
 

### 2023-06-15 popl/main.tex
add visualisation

@@
 In this paper, compilers are modelled as function from one programming language's syntax to the other.
-Since this paper considers composing compilers, we'll adopt notation similar to tombstone diagrams\MK{cite} to easen the formal burden.
-So, denote a compiler from $\src{L}$ to $\trg{L}$ as $\cc{\src{L}\to\trg{L}}{\src{\bullet}}$ and refer to $\src{L}$ as \emph{source} language and to $\trg{L}$ as \emph{target} language.
-Composing $\cc{\src{L}\to\trg{L}}{\src{\bullet}}$ with $\cc{\trg{L}\to\obj{L}}{\trg{\bullet}}$ is denoted as $\cc{\src{L}\to\trg{L}\to\obj{L}}{\src{\bullet}}$ and means first running $\cc{\src{L}\to\trg{L}}{\src{\bullet}}$ and plugging the output into $\cc{\trg{L}\to\obj{L}}{\trg{\bullet}}$.
-Contrary to tombstones, the implementation language of a compiler is disregarded.
+Notationally, $\cc{\src{L}}{\trg{L}}$ denotes a compiler from $\src{L}$ to $\trg{L}$ and the paper refers to $\src{L}$ as \emph{source} language and to $\trg{L}$ as \emph{target} language.
 Now, define the compiler-property robust trace-property preservation.
@@
 \begin{definition}[Robust Trace-Property Preservation]\label{def:rtp}
-  A compiler $\cc{\src{L}\to\trg{L}}{\src{\bullet}}$ robustly preserves a class of properties $\class$ ($\rtp{\cc{\src{L}\to\trg{L}}{\src{\bullet}}}{\class}$) iff for any property $\pi\in\class$ and $\src{p}$ written in $\src{L}$, given $\rsat{\src{\progvar}}{\pi}$, then $\rsat{\cc{\src{L}\to\trg{L}}{\src{\progvar}}}{\pi}$.
+  A compiler $\cc{\src{L}}{\trg{L}}$ robustly preserves a class of properties $\class$ ($\rtp{\cc{\src{L}}{\trg{L}}}{\class}$) iff for any property $\pi\in\class$ and $\src{p}$ written in $\src{L}$, given $\rsat{\src{\progvar}}{\pi}$, then $\rsat{\cc{\src{L}}{\trg{L}}\left(\src{p}\right)}{\pi}$.
 \end{definition}
@@
 \section{Composing Secure Compilers}\label{sec:compcomp}
+
+\begin{figure}[h]
+  \begin{center}
+    \begin{tikzpicture}
+      \node (S) {$\src{L_{\tmssafe}}$};
+      \node[right=1.0 of S] (T) {$\trg{L}$};
+      \node[right=1.0 of T] (M) {$\irl{L_{\mssafe}}$};
+      \node[below right=0.5 and 1.0 of M] (D0) {$\irl{L_{\mssafe}}$};
+      \node[above right=0.5 and 1.0 of M] (C0) {$\irl{L_{\mssafe}}$};
+      \node[right=1.0 of D0] (D1) {$\irl{L_{\mssafe}}$};
+      \node[right=1.0 of C0] (C1) {$\irl{L_{\mssafe}}$};
+      \node[right=4.5 of M] (O) {$\obj{L_{\scctsafe}}$};
+
+      \draw[->] (S) to[sloped,above] node {\text{\scriptsize\gls{tms}}} (T);
+      \draw[->] (T) to[sloped,above] node {\text{\scriptsize\gls{sms}}} (M);
+      \draw[->] (M) to[sloped,above] node {\text{\scriptsize\gls{dce}}} (D0);
+      \draw[->] (M) to[sloped,above] node {\text{\scriptsize\gls{cf}}} (C0);
+      \draw[->] (D0) to[sloped,above] node {\text{\scriptsize\gls{cf}}} (D1);
+      \draw[->] (C0) to[sloped,above] node {\text{\scriptsize\gls{dce}}} (C1);
+      \draw[->] (D1) to[sloped,above] node {\text{\scriptsize\gls{scct}}} (O);
+      \draw[->] (C1) to[sloped,above] node {\text{\scriptsize\gls{scct}}} (O);
+    \end{tikzpicture}
+  \end{center}
+  \caption{Visualisation of the optimizing compilation pipeline that attains a combination of \gls{ms} and \gls{cct}.}
+\end{figure}
+
 This section studies different kinds of secure compiler compositions and their corresponding proofs.
@@
 \begin{definition}[Sequential Composition]\label{def:seqcomp}
-  Let $\cc{\src{L}\to\trg{L}}{\src{\bullet}}$, $\cc{\trg{L}\to\obj{L}}{\trg{\bullet}}$ be compilers.
-  Let $\src{p}$ be any $\src{L}$ program, $\trg{p}=\cc{\src{L}\to\trg{L}}{\src{p}}$, and $\obj{p}=\cc{\trg{L}\to\obj{L}}{\trg{p}}$.
-  The sequential composition is $\cc{\src{L}\to\trg{L}\to\obj{L}}{\src{p}}=\obj{p}$.
+  Let $\cc{\src{L}}{\trg{L}}$, $\cc{\trg{L}}{\obj{L}}$ be compilers.
+  Let $\src{p}$ be any $\src{L}$ program, $\trg{p}=\cc{\src{L}}{\trg{L}}\left(\src{p}\right)$, and $\obj{p}=\cc{\trg{L}}{\obj{L}}\left(\trg{p}\right)$.
+  The sequential composition is $\left(\cc{\src{L}}{\trg{L}}\circ\cc{\trg{L}}{\obj{L}}\right)\left(\src{p}\right)=\obj{p}$.
 \end{definition}
@@
 \begin{theorem}[Sequential Composition of Secure Compilers]\label{thm:rtp}
-  Given classes of properties $\class[_{1}],\class[_{2}]$, let $\rtp{\cc{\src{L}\to\trg{L}}{\src{\bullet}}}{\class[_{1}]}$ and $\rtp{\cc{\trg{L}\to\obj{L}}{\trg{\bullet}}}{\class[_{2}]}$.
-  Then, $\rtp{\cc{\src{L}\to\trg{L}\to\obj{L}}{\src{\bullet}}}{\class[_{1}]\cap\class[_{2}]}$. \Coqed
+  Given classes of properties $\class[_{1}],\class[_{2}]$, let $\rtp{\cc{\src{L}}{\trg{L}}}{\class[_{1}]}$ and $\rtp{\cc{\trg{L}}{\obj{L}}}{\class[_{2}]}$.
+  Then, $\rtp{\cc{\src{L}}{\trg{L}}\circ\cc{\trg{L}}{\obj{L}}}{\class[_{1}]\cap\class[_{2}]}$. \Coqed
 \end{theorem}
@@
 \begin{corollary}[Sequential Composition of Secure Compilers]\label{corr:swappable}
-  Given classes of properties $\class[_{1}],\class[_{2}]$, let $\rtp{\cc{\trg{L}\xrightarrow{1}\trg{L}}{\trg{\bullet}}}{\class[_{1}]}$ and $\rtp{\cc{\trg{L}\xrightarrow{2}\trg{L}}{\trg{\bullet}}}{\class[_{2}]}$.
-  Then, $\rtp{\cc{\trg{L}\xrightarrow{1}\trg{L}\xrightarrow{2}\trg{L}}{\trg{\bullet}}}{\class[_{1}]\cap\class[_{2}]}$ and $\rtp{\cc{\trg{L}\xrightarrow{2}\trg{L}\xrightarrow{1}\trg{L}}{\trg{\bullet}}}{\class[_{2}]\cap\class[_{1}]}$.
-\Coqed
+  Given classes of properties $\class[_{1}],\class[_{2}]$, let $\rtp{\cc[_{1}]{\trg{L}}{\trg{L}}}{\class[_{1}]}$ and $\rtp{\cc[_{2}]{\trg{L}}{\trg{L}}}{\class[_{2}]}$.
+  Then, $\rtp{\cc[_{1}]{\trg{L}}{\trg{L}}}{\class[_{1}]\cap\class[_{2}]}$ and $\rtp{\cc[_{2}]{\trg{L}}{\trg{V}}}{\class[_{2}]\cap\class[_{1}]}$. \Coqed
 \end{corollary}
@@
 The power of the frameworks from \Cref{sec:compprop,sec:compcomp} is demonstrated by composing these compilers for a secure, but optimizing, compilation chain that robustly preserves \gls{ms} and \gls{cct}.
-Concretely, the section starts starts by defining languages $\src{L_{TMS}}$ and $\trg{L}$ and describes a compiler $\cc{\src{L_{TMS}}\to\trg{L}}{\src{\bullet}}$ for the robust preservation of \gls{tms} between the two languages (\Cref{subsec:cs:tms}).
-Then, it defines $\irl{L_{MS}}$ and a compiler $\cc{\trg{L}\to\irl{L_{MS}}}{\trg{\bullet}}$ that robustly preserves \gls{sms} (\Cref{subsec:cs:ms}).
-By the framework presented in \Cref{sec:compcomp,sec:compprop}, a compiler $\cc{\src{L_{TMS}}\to\trg{L}\to\irl{L_{MS}}}{\src{\bullet}}$ for the robust preservation of \gls{ms} by composing $\cc{\src{L_{TMS}}\to\trg{L}}{\src{\bullet}}$ and $\cc{\trg{L}\to\irl{L_{MS}}}{\trg{\bullet}}$ is attained.
-Two secure compilers $\cc{{\irl{L_{MS}}\xrightarrow{\gls{dce}}\irl{L_{MS}}}}{\irl{\bullet}}$ and $\cc{\irl{L_{MS}}\xrightarrow{\gls{cf}}\irl{L_{MS}}}{\irl{\bullet}}$ are defined that perform \gls{dce} and \gls{cf}, respectively, to model an optimzing pipeline happening somewhere in the compilation chain (\Cref{subsec:cs:opts}).
-Finally, the paper defines $\obj{L_{CCT}}$ and a compiler $\cc{\irl{L_{MS}}\to\obj{L_{CCT}}}{\irl{\bullet}}$ that robustly preserves \gls{cct} (\Cref{subsec:cs:scct}).
-Composing it with $\cc{\src{L_{TMS}}\to\irl{L_{MS}}}{\src{\bullet}}$ and the optimizing compilers $\cc{\irl{L_{MS}}\xrightarrow{\gls{dce}}\irl{L_{MS}}}{\irl{\bullet}}$ and $\cc{\irl{L_{MS}}\xrightarrow{\gls{cf}}\irl{L_{MS}}}{\irl{\bullet}}$ yields $\cc{\src{L_{TMS}}\to\trg{L}\to\irl{L_{MS}}\xrightarrow{\gls{cf}}\irl{L_{MS}}\xrightarrow{\gls{dce}}\irl{L_{MS}}\to\obj{L_{CCT}}}{\src{\bullet}}$, a compiler that robustly preserves a combination of \gls{ms} and \gls{cct}.
+Concretely, the section starts starts by defining languages $\src{L_{TMS}}$ and $\trg{L}$ and describes a compiler $\cc{\src{L_{TMS}}}{\trg{L}}$ for the robust preservation of \gls{tms} between the two languages (\Cref{subsec:cs:tms}).
+Then, it defines $\irl{L_{MS}}$ and a compiler $\cc{\trg{L}}{\irl{L_{MS}}}$ that robustly preserves \gls{sms} (\Cref{subsec:cs:ms}).
+By the framework presented in \Cref{sec:compcomp,sec:compprop}, a compiler $\rtp{\cc{\src{L_{TMS}}}{\trg{L}}\circ\cc{\trg{L}}{\irl{L_{MS}}}}{\mssafe}$ is attained.
+Two secure compilers $\cc[_{\gls{dce}}]{\irl{L_{MS}}}{\irl{L_{MS}}}$ and $\cc[_{\gls{cf}}]{\irl{L_{MS}}}{\irl{L_{MS}}}$ are defined that perform \gls{dce} and \gls{cf}, respectively, to model an optimzing pipeline happening somewhere in the compilation chain (\Cref{subsec:cs:opts}).
+Finally, the paper defines $\obj{L_{CCT}}$ and a compiler $\cc{\irl{L_{MS}}}{\obj{L_{CCT}}}$ that robustly preserves \gls{cct} (\Cref{subsec:cs:scct}).
+Composing it with $\cc{\src{L_{TMS}}}{\irl{L_{MS}}}$ and the optimizing compilers $\cc[_{\gls{dce}}]{\irl{L_{MS}}}{\irl{L_{MS}}}$ and $\cc[_{\gls{cf}}]{\irl{L_{MS}}}{\irl{L_{MS}}}$ yields $\cc{\src{L_{TMS}}}{\obj{L_{CCT}}}$, a compiler that robustly preserves a combination of \gls{ms} and \gls{cct}.
+
+\MKin{add image and do introduction using that}
 

### 2023-06-16 popl/main.tex
Overview. Combined Monitor proof.

@@
 
-\begin{figure}[h]
-  \begin{center}
-    \begin{tikzpicture}
-      \node (S) {$\src{L_{\tmssafe}}$};
-      \node[right=1.0 of S] (T) {$\trg{L}$};
-      \node[right=1.0 of T] (M) {$\irl{L_{\mssafe}}$};
-      \node[below right=0.5 and 1.0 of M] (D0) {$\irl{L_{\mssafe}}$};
-      \node[above right=0.5 and 1.0 of M] (C0) {$\irl{L_{\mssafe}}$};
-      \node[right=1.0 of D0] (D1) {$\irl{L_{\mssafe}}$};
-      \node[right=1.0 of C0] (C1) {$\irl{L_{\mssafe}}$};
-      \node[right=4.5 of M] (O) {$\obj{L_{\scctsafe}}$};
-
-      \draw[->] (S) to[sloped,above] node {\text{\scriptsize\gls{tms}}} (T);
-      \draw[->] (T) to[sloped,above] node {\text{\scriptsize\gls{sms}}} (M);
-      \draw[->] (M) to[sloped,above] node {\text{\scriptsize\gls{dce}}} (D0);
-      \draw[->] (M) to[sloped,above] node {\text{\scriptsize\gls{cf}}} (C0);
-      \draw[->] (D0) to[sloped,above] node {\text{\scriptsize\gls{cf}}} (D1);
-      \draw[->] (C0) to[sloped,above] node {\text{\scriptsize\gls{dce}}} (C1);
-      \draw[->] (D1) to[sloped,above] node {\text{\scriptsize\gls{scct}}} (O);
-      \draw[->] (C1) to[sloped,above] node {\text{\scriptsize\gls{scct}}} (O);
-    \end{tikzpicture}
-  \end{center}
-  \caption{Visualisation of the optimizing compilation pipeline that attains a combination of \gls{ms} and \gls{cct}.}
-\end{figure}
-
 This section studies different kinds of secure compiler compositions and their corresponding proofs.
@@
 \section{Case Study: Memory Safe and Cryptographic Constant Time Code}\label{sec:casestud}
-This section defines several secure compilers, each of which robustly preserves a different property of interest.
-The power of the frameworks from \Cref{sec:compprop,sec:compcomp} is demonstrated by composing these compilers for a secure, but optimizing, compilation chain that robustly preserves \gls{ms} and \gls{cct}.
-Concretely, the section starts starts by defining languages $\src{L_{TMS}}$ and $\trg{L}$ and describes a compiler $\cc{\src{L_{TMS}}}{\trg{L}}$ for the robust preservation of \gls{tms} between the two languages (\Cref{subsec:cs:tms}).
-Then, it defines $\irl{L_{MS}}$ and a compiler $\cc{\trg{L}}{\irl{L_{MS}}}$ that robustly preserves \gls{sms} (\Cref{subsec:cs:ms}).
-By the framework presented in \Cref{sec:compcomp,sec:compprop}, a compiler $\rtp{\cc{\src{L_{TMS}}}{\trg{L}}\circ\cc{\trg{L}}{\irl{L_{MS}}}}{\mssafe}$ is attained.
-Two secure compilers $\cc[_{\gls{dce}}]{\irl{L_{MS}}}{\irl{L_{MS}}}$ and $\cc[_{\gls{cf}}]{\irl{L_{MS}}}{\irl{L_{MS}}}$ are defined that perform \gls{dce} and \gls{cf}, respectively, to model an optimzing pipeline happening somewhere in the compilation chain (\Cref{subsec:cs:opts}).
-Finally, the paper defines $\obj{L_{CCT}}$ and a compiler $\cc{\irl{L_{MS}}}{\obj{L_{CCT}}}$ that robustly preserves \gls{cct} (\Cref{subsec:cs:scct}).
-Composing it with $\cc{\src{L_{TMS}}}{\irl{L_{MS}}}$ and the optimizing compilers $\cc[_{\gls{dce}}]{\irl{L_{MS}}}{\irl{L_{MS}}}$ and $\cc[_{\gls{cf}}]{\irl{L_{MS}}}{\irl{L_{MS}}}$ yields $\cc{\src{L_{TMS}}}{\obj{L_{CCT}}}$, a compiler that robustly preserves a combination of \gls{ms} and \gls{cct}.
+\begin{figure}[h]
+  \begin{center}
+    \begin{tikzpicture}
+      \node (S) {$\src{L_{\tmssafe}}$};
+      \node[right=1.0 of S] (T) {$\trg{L}$};
+      \node[right=1.0 of T] (M) {$\irl{L_{\mssafe}}$};
+      \node[below right=0.5 and 1.0 of M] (D0) {$\irl{L_{\mssafe}}$};
+      \node[above right=0.5 and 1.0 of M] (C0) {$\irl{L_{\mssafe}}$};
+      \node[right=1.0 of D0] (D1) {$\irl{L_{\mssafe}}$};
+      \node[right=1.0 of C0] (C1) {$\irl{L_{\mssafe}}$};
+      \node[right=4.5 of M] (O) {$\obj{L_{\scctsafe}}$};
 
-\MKin{add image and do introduction using that}
+      \draw[->] (S) to[sloped,above] node {\text{\scriptsize\gls{tms}}} (T);
+      \draw[->] (T) to[sloped,above] node {\text{\scriptsize\gls{sms}}} (M);
+      \draw[->] (M) to[sloped,above] node {\text{\scriptsize\gls{dce}}} (D0);
+      \draw[->] (M) to[sloped,above] node {\text{\scriptsize\gls{cf}}} (C0);
+      \draw[->] (D0) to[sloped,above] node {\text{\scriptsize\gls{cf}}} (D1);
+      \draw[->] (C0) to[sloped,above] node {\text{\scriptsize\gls{dce}}} (C1);
+      \draw[->] (D1) to[sloped,above] node {\text{\scriptsize\gls{scct}}} (O);
+      \draw[->] (C1) to[sloped,above] node {\text{\scriptsize\gls{scct}}} (O);
+    \end{tikzpicture}
+  \end{center}
+  \caption{Visualisation of the optimizing compilation pipeline that attains a combination of \gls{ms} and \gls{cct}.}\label{fig:pipeline}
+\end{figure}
+
+This section defines several secure compilers, each of which robustly preserves a different property of interest (see \Cref{fig:pipeline}).
+The power of the framework (\Cref{sec:compprop,sec:compcomp}) is demonstrated by composing these compilers for a secure, but optimizing, compilation chain that robustly preserves \gls{ms} and \gls{cct}.
+Concretely, the section starts with a base-line definition for the used languages and extends that to define languages $\src{L_{TMS}}$ and $\trg{L}$.
+For programmers using $\src{L_{TMS}}$, their mental model should be that their code is temporal memory safe and cryptographic constant time.
+The former is enforced by the static semantics of $\src{L_{TMS}}$, where each well-typed program is temporal memory safe.
+The latter is completely untyped.
+Having languages defined, this section continues with a compiler $\cc{\src{L_{\tmssafe}}}{\trg{L}}$ and proves that $\cc{\src{L_{\tmssafe}}}{\trg{L}}$ robustly preserves \gls{tms} (\Cref{subsec:cs:tms}).
+For readability purposes, it defines $\irl{L_{MS}}$ to be exactly the same as $\trg{L}$ and continues to present a compiler $\cc{\trg{L}}{\irl{L_{MS}}}$ that robustly preserves \gls{sms} (\Cref{subsec:cs:ms}).
+Now leaveraging the power of the framework that this paper introduces (\Cref{sec:compcomp,sec:compprop}), a compiler $\rtp{\cc{\src{L_{TMS}}}{\irl{L_{MS}}}}{\mssafe}$ is attained.
+The next steps in the compilation pipeline are two secure compilers $\cc[_{\gls{dce}}]{\irl{L_{MS}}}{\irl{L_{MS}}}$ and $\cc[_{\gls{cf}}]{\irl{L_{MS}}}{\irl{L_{MS}}}$ that perform \gls{dce} and \gls{cf}, respectively, to model an optimzing pipeline happening somewhere in the compilation chain (\Cref{subsec:cs:opts}).
+Finally, this section defines the language $\obj{L_{\scctsafe}}$ that is, in contrast to the other languages, not cryptographic constant time by definition.
+Concluding, a compiler $\cc{\irl{L_{MS}}}{\obj{L_{CCT}}}$ that robustly preserves \gls{cct} (\Cref{subsec:cs:scct}) is shown and composed with $\cc{\src{L_{TMS}}}{\irl{L_{MS}}}$ and the optimizing compilers $\cc[_{\gls{dce}}]{\irl{L_{MS}}}{\irl{L_{MS}}}$ and $\cc[_{\gls{cf}}]{\irl{L_{MS}}}{\irl{L_{MS}}}$.
+In turn, this yields $\cc{\src{L_{TMS}}}{\obj{L_{CCT}}}$, a compiler that robustly preserves a combination of \gls{ms} and \gls{cct}.
 

### 2023-06-17 popl/main.tex
Stutter eq

@@
 
-This section defines several secure compilers, each of which robustly preserves a different property of interest (see \Cref{fig:pipeline}).
+This section defines several secure compilers, each of which robustly preserves a different property of interest (\Cref{fig:pipeline}).
 The power of the framework (\Cref{sec:compprop,sec:compcomp}) is demonstrated by composing these compilers for a secure, but optimizing, compilation chain that robustly preserves \gls{ms} and \gls{cct}.
@@
 For programmers using $\src{L_{TMS}}$, their mental model should be that their code is temporal memory safe and cryptographic constant time.
-The former is enforced by the static semantics of $\src{L_{TMS}}$, where each well-typed program is temporal memory safe.
+The former is proven to be enforced by the static semantics of $\src{L_{TMS}}$, where each well-typed program is temporal memory safe.
 The latter is completely untyped.
@@
 Now leaveraging the power of the framework that this paper introduces (\Cref{sec:compcomp,sec:compprop}), a compiler $\rtp{\cc{\src{L_{TMS}}}{\irl{L_{MS}}}}{\mssafe}$ is attained.
-The next steps in the compilation pipeline are two secure compilers $\cc[_{\gls{dce}}]{\irl{L_{MS}}}{\irl{L_{MS}}}$ and $\cc[_{\gls{cf}}]{\irl{L_{MS}}}{\irl{L_{MS}}}$ that perform \gls{dce} and \gls{cf}, respectively, to model an optimzing pipeline happening somewhere in the compilation chain (\Cref{subsec:cs:opts}).
+The next steps in the compilation pipeline are two secure compilers $\cc[_{\gls{dce}}]{\irl{L_{MS}}}{\irl{L_{MS}}}$ and $\cc[_{\gls{cf}}]{\irl{L_{MS}}}{\irl{L_{MS}}}$ that perform \glsfirst{dce} and \glsfirst{cf}, respectively, to model an optimzing pipeline happening somewhere in the compilation chain (\Cref{subsec:cs:opts}).
 Finally, this section defines the language $\obj{L_{\scctsafe}}$ that is, in contrast to the other languages, not cryptographic constant time by definition.

### 2023-06-17 popl/main.tex
Baseline def

@@
 \usepackage{fontawesome5}
+\usepackage{listofitems}
 \usepackage{glossaries}
@@
 \usepackage{stmaryrd}
+\usepackage{marvosym}
 \usepackage{listings}
 \usepackage{xspace}
+\usepackage{xfrac}
 \usepackage{tikz}
@@
 The power of the framework (\Cref{sec:compprop,sec:compcomp}) is demonstrated by composing these compilers for a secure, but optimizing, compilation chain that robustly preserves \gls{ms} and \gls{cct}.
-Concretely, the section starts with a base-line definition for the used languages and extends that to define languages $\src{L_{TMS}}$ and $\trg{L}$.
+Concretely, the section starts with a base-line definition for the used languages (\Cref{subsec:cs:defs}) and extends that to define languages $\src{L_{TMS}}$ and $\trg{L}$.
 For programmers using $\src{L_{TMS}}$, their mental model should be that their code is temporal memory safe and cryptographic constant time.
@@
 In turn, this yields $\cc{\src{L_{TMS}}}{\obj{L_{CCT}}}$, a compiler that robustly preserves a combination of \gls{ms} and \gls{cct}.
+
+\subsection{Preliminary Definitions}\label{subsec:cs:defs}
+\begin{gather*}
+  \begin{aligned}
+  \mi{(Finals)}~\finalexprnoerr&\bnfdef\ \valueexpr \mid x\hspace{0.5cm}
+  \mi{(Results)}~\finalexpr\bnfdef\ \finalexprnoerr \mid stuck\hspace{0.5cm}
+  \mi{(Values)}~\valueexpr\bnfdef\ n\\
+  \mi{(Expressions)}~\expr&\bnfdef\ \finalexpr \mid \lbinop{\expr[_{1}]}{\expr[_{2}]} \mid \lget{x}{\expr} \mid \lnew{x}{\expr[_{1}]}{\expr[_{2}]} \mid \ldelete{x} \mid \lset{x}{\expr[_{1}]}{\expr[_{2}]} \\
+    & \mid \llet{x}{\expr[_{1}]}{\expr[_{2}]} \mid \lreturn{\expr} \mid \lcall{g}{\expr} \mid \lifz{\expr}{\expr[_{1}]}{\expr[_{2}]} \\
+    & \text{where}\ \lbinop{}{}\in\left\{+,-,\times,<,/\right\}\\
+  \mi{(Functions)}~\function&\bnfdef\ \lfunction{g}{x}{e}\hspace{0.5cm}
+  \mi{(Types)}~\type\bnfdef\ \natt\hspace{0.5cm}
+  \mi{(Locations)}~\loc\in\nat \\
+  \mi{(Eval.~Ctx.)}~\ectx&\bnfdef\ \hole{\cdot} \mid \lbinop{\ectx}{\expr} \mid \lbinop{\valueexpr}{\ectx} \mid \lget{x}{\ectx} \mid \lnew{x}{\ectx}{\expr} \mid \lset{x}{\ectx}{\expr}\\
+    & \mid \lset{x}{\valueexpr}{\ectx} \mid \lreturn{\ectx} \mid \lcall{g}{\ectx} \mid \lifz{\ectx}{\expr[_{1}]}{\expr[_{2}]} \\
+  \mi{(Poison)}~\poison&\bnfdef\ \poisoned \mid \poisonless \hspace{0.5cm}
+  \mi{(Store)}~\store\bnfdef\ \hole{\cdot} \mid \storeel{x}{\loc}{\sandboxtag}{\poison}{n},\store \\
+  \mi{(Transfer~Tag)}~\comm&\bnfdef\ \nocomm \mid\ \ctxtocomp \mid\ \comptoctx \hspace{0.5cm}
+  \mi{(Heaps)}~\heap\bnfdef\ \hole{\cdot} \mid n,\heap \hspace{0.5cm}
+  \mi{(Library)}~\library\bnfdef\ \hole{\cdot} \mid \function,\library \\
+  \mi{(Component~Names)}~\commlib&\bnfdef\ \hole{\cdot} \mid g,\commlib \hspace{0.5cm}
+  \mi{(Continuation~Stack)}~\ectxstack\bnfdef\ \hole{\cdot} \mid \tup{\ectx}{g},\ectxstack \\
+  \mi{(Flow~State)}~\cfstate&\bnfdef\ \tup{\commlib}{\library,\ectxstack} \hspace{0.5cm}
+  \mi{(Memory~State)}~\memstate\bnfdef\ \tup{\heap[^{\ctx}]}{\heap[^{\comp}],\store} \hspace{0.5cm}
+  \end{aligned}
+\end{gather*}
 

### 2023-06-17 popl/main.tex
Work on text for baseline def and introduce pstep

@@
 For languages used in this paper, a small-step style semantics is used.
-That is, there is a step relation between states $\statevar$ and an event $\event$: $\stepto[]{\statevar[_{1}]}{\statevar[_{2}]}{\event}$.
-Also, a program's syntax $p$ is encoded in the state $\statevar$ which is written as $p\in\statevar$.
-For states $\statevar$, it is assumed that there exists a judgement $\isval{\statevar}$ that determines wether $\statevar$ is a final state or not.
-\Cref{fig:starstep:and:nstep}\MK{appendix?} defines the star-step relation $\stepto{\statevar[_{1}]}{\statevar[_{2}]}{\trace}$, essentially modelling a big-step style evaluation, and the $n$-step relation $\stepton{\statevar[_{1}]}{\statevar[_{2}]}{\trace}$ that evaluates a program for $n$ steps.
+That is, there is a step relation between states $\runtimetermvar$ and an event $\event$: $\stepto[]{\runtimetermvar[_{1}]}{\runtimetermvar[_{2}]}{\event}$.
+Also, a program's syntax $p$ is encoded in the state $\runtimetermvar$ which is written as $p\in\runtimetermvar$.
+For states $\runtimetermvar$, it is assumed that there exists a judgement $\isval{\runtimetermvar}$ that determines wether $\runtimetermvar$ is a final state or not.
+\Cref{fig:starstep:and:nstep}\MK{appendix?} defines the star-step relation $\stepto{\runtimetermvar[_{1}]}{\runtimetermvar[_{2}]}{\trace}$, essentially modelling a big-step style evaluation, and the $n$-step relation $\stepton{\runtimetermvar[_{1}]}{\runtimetermvar[_{2}]}{\trace}$ that evaluates a program for $n$ steps.
 Both relations filter $\varepsilon$ events.
@@
 \begin{figure}[t]
-  \judgbox{\stepto{r}{r'}{\trace}}{,,State $\statevar$ steps to final state $\statevar[']$, emitting events $\trace$.''}$\;$\\
+  \judgbox{\stepto{r}{r'}{\trace}}{,,State $\runtimetermvar$ steps to final state $\runtimetermvar[']$, emitting events $\trace$.''}$\;$\\
   %
   \typerule{star-step-refl}{
-    \isval{\statevar}
+    \isval{\runtimetermvar}
   }{
-    \stepto{\statevar}{\statevar}{\hole{\cdot}}
+    \stepto{\runtimetermvar}{\runtimetermvar}{\hole{\cdot}}
   }{star-step-refl}
@@
     \event\not=\varepsilon &
-    \stepto[]{\statevar}{\statevar[']}{\event} &
-    \stepto{\statevar[']}{\statevar['']}{\trace}
+    \stepto[]{\runtimetermvar}{\runtimetermvar[']}{\event} &
+    \stepto{\runtimetermvar[']}{\runtimetermvar['']}{\trace}
   }{
-    \stepto{\statevar}{\statevar['']}{\event\cdot\trace}
+    \stepto{\runtimetermvar}{\runtimetermvar['']}{\event\cdot\trace}
   }{star-step-trans}
@@
   \typerule{star-step-ign}{
-    \stepto[]{\statevar}{\statevar[']}{\varepsilon} &
-    \stepto{\statevar[']}{\statevar['']}{\trace}
+    \stepto[]{\runtimetermvar}{\runtimetermvar[']}{\varepsilon} &
+    \stepto{\runtimetermvar[']}{\runtimetermvar['']}{\trace}
   }{
-    \stepto{\statevar}{\statevar['']}{\trace}
+    \stepto{\runtimetermvar}{\runtimetermvar['']}{\trace}
   }{star-step-ign}
 
-  \judgbox{\stepton{\statevar}{\statevar[']}{\trace}}{,,State $r$ does $n$ steps to final state $\statevar[']$, emitting events $\trace$.''}$\;$\\
+  \judgbox{\stepton{\runtimetermvar}{\runtimetermvar[']}{\trace}}{,,State $r$ does $n$ steps to final state $\runtimetermvar[']$, emitting events $\trace$.''}$\;$\\
   %
@@
   }{
-    \stepton[0]{\statevar}{\statevar}{\hole{\cdot}}
+    \stepton[0]{\runtimetermvar}{\runtimetermvar}{\hole{\cdot}}
   }{n-step-refl}
@@
     \event\not=\varepsilon &
-    \stepto[]{\statevar}{\statevar[']}{\event} &
-    \stepton{\statevar[']}{\statevar['']}{\trace}
+    \stepto[]{\runtimetermvar}{\runtimetermvar[']}{\event} &
+    \stepton{\runtimetermvar[']}{\runtimetermvar['']}{\trace}
   }{
-    \stepton[n+1]{\statevar}{\statevar['']}{\event\cdot\trace}
+    \stepton[n+1]{\runtimetermvar}{\runtimetermvar['']}{\event\cdot\trace}
   }{n-step-trans}
@@
   \typerule{n-step-ign}{
-    \stepto[]{\statevar}{\statevar[']}{\varepsilon} &
-    \stepton{\statevar[']}{\statevar['']}{\trace}
+    \stepto[]{\runtimetermvar}{\runtimetermvar[']}{\varepsilon} &
+    \stepton{\runtimetermvar[']}{\runtimetermvar['']}{\trace}
   }{
-    \stepton[n+1]{\statevar}{\statevar['']}{\trace}
+    \stepton[n+1]{\runtimetermvar}{\runtimetermvar['']}{\trace}
   }{n-step-ign}
 
-  \caption{Star-step and $n$-step relation for a given step relation $\stepto{\statevar[_{1}]}{\statevar[_{2}]}{\event}$}\label{fig:starstep:and:nstep}
+  \caption{Star-step and $n$-step relation for a given step relation $\stepto{\runtimetermvar[_{1}]}{\runtimetermvar[_{2}]}{\event}$}\label{fig:starstep:and:nstep}
 \end{figure}
@@
     & \mid \llet{x}{\expr[_{1}]}{\expr[_{2}]} \mid \lreturn{\expr} \mid \lcall{g}{\expr} \mid \lifz{\expr}{\expr[_{1}]}{\expr[_{2}]} \\
-    & \text{where}\ \lbinop{}{}\in\left\{+,-,\times,<,/\right\}\\
-  \mi{(Functions)}~\function&\bnfdef\ \lfunction{g}{x}{e}\hspace{0.5cm}
+  \mi{(State)}~\statevar&\bnfdef \tup{\cfstate}{\sandboxtag,\memstate} \hspace{0.5cm}
+  \mi{(Runtime~Terms)}~\runtimetermvar\bnfdef\rtt{\statevar}{\expr} \hspace{0.5cm}
+    \text{where}\ \lbinop{}{}\in\left\{+,-,\times,<,/\right\}\\
+  \mi{(Symbols)}~\function&\bnfdef\ \lfunction{g}{x}{e}\hspace{0.5cm}
   \mi{(Types)}~\type\bnfdef\ \natt\hspace{0.5cm}
@@
   \mi{(Flow~State)}~\cfstate&\bnfdef\ \tup{\commlib}{\library,\ectxstack} \hspace{0.5cm}
-  \mi{(Memory~State)}~\memstate\bnfdef\ \tup{\heap[^{\ctx}]}{\heap[^{\comp}],\store} \hspace{0.5cm}
+  \mi{(Memory~State)}~\memstate\bnfdef\ \tup{\heap[^{\ctx}]}{\heap[^{\comp}],\store} \\
+  \mi{(Programs)}~&\prog{\library_{\ctx}}{\library_{\comp}}\hspace{0.5cm}
+  \mi{(Substitutions)}~\substvar\bnfdef \hole{\cdot} \mid \subst{\valueexpr}{x},\gamma \\
+  \mi{(Pre\text{--}Events)}~\event[_{b}]&\bnfdef \ev{Alloc}\ \loc\ \valueexpr \mid \ev{Dealloc}\ \loc \mid \ev{Get}\ \loc\ \valueexpr \mid \ev{Set}\ \loc\ \valueexpr\ \valueexpr['] \mid \ev{Call}\ \comm\ g\ \valueexpr \mid \ev{Ret}\ \comm\ \valueexpr\\
+    & \mid \ev{Start} \mid \ev{End}\ \valueexpr\hspace{0.5cm}
+  \mi{(Events)}~\event\bnfdef \emptyevent \mid \lightning \mid \tup{\event[_{b}]}{\sandboxtag} \\
   \end{aligned}
 \end{gather*}
+
+Above is the base syntax of all the programming languages of this paper.
+Expressions ($\expr$) must not contain the result $stuck$ unless they are entailed as runtime--term ($\rtt{\statevar}{\expr}$).
+States ($\statevar$) have a flow- ($\cfstate$) and a memory state ($\memstate$).
+As for the flow state, the list of component--level names ($\commlib$), the list of symbols ($\library$), and a stack of continuations tagged with the name of a symbol ($\ectxstack$) are necessary for calling and returning while emitting the right transfer tag ($\sandboxtag$) depending on whether there is a context switch from context to component ($\comptoctx$), from component to context ($\ctxtocomp$), or whether control flow stays internal to either ($\nocomm$).
+The memory state contains both a context--level ($\heap[^{\ctx}]$) and a component--level heap ($\heap[^{comp}]$) as well as a store ($\store$).
+Effectively, it is syntactically impossible for a context--level runtime--term to have a pointer into the component--level heap and vice--versa.
+Pointers are non-nullable and there is no syntactic way to let them point into the middle of an object.
+However, pointers always point to arrays and the individual elements of that array are read- and writeable by means of adding an offset that may be negative.
+Heap locations are second--class. %The syntax is constrained as such that one cannot simply guess a heap location.
+Stores map identifiers to pointer metadata.
+This metadata is a tuple with the location ($\loc$), the control--tag ($\sandboxtag$), poison--tag ($\poison$), and the size of the allocated object ($n$).
+The control-tag allows to distinguish whether a location indexes into $\heap[^{\ctx}]$ or $\heap[^{\comp}]$.
+The poison--tag marks a location as freed ($\poisoned$) without removing the entry from $\store$.
+If one removes an entry, for example, by evaluating $\ldelete{x}$, then there is no way to distinguish between freed pointers and free variables.
+The latter should render evaluation as stuck, while the former should emit a $\ev{Dealloc}\ \loc$ event regardless of wether the object a pointer points to is still allocated or not.
+Programs ($\prog{\library_{\ctx}}{\library_{\comp}}$) have disjoint sets of symbols, one for context-level ($\library_{\ctx}$) and one for component-level symbols ($\library_{\comp}$).
+
+All programming languages presented in this section have an environmental semantics and all but $\src{L}$ are dynamically typed.
+Each evaluation may emit the empty event ($\emptyevent$), a program crash token ($\lightning$), or a pair containing a pre--event ($\event[_{b}]$) and the corresponding control--tag ($\sandboxtag$) indicating whether the event fired during an evaluation of a context or a component.
+Note that the reflexive-transitive closure used in this paper behaves like a big-step semantics, where the reflexive case is ony for the empty trace and finals, and the emitted trace does not contain empty events.
+Most rules are entirely standard and omitted.
+Heaps may need to grow when allocating, which the section writes as $\growh{\heap}{n}$, denoting that $\heap$ grows by $n$ cells.
+Indexing into a heap, looking up symbols in libraries, or accessing the metadata of a pointer from a store are all written as a function call.
+For example, the value located at the fifth position in heap $\heap$ is written as $\heap(5)$.
+The operation to set the fifth position in heap $\heap$ to value $\valueexpr$ is written as $\seth{\heap}{5}{\valueexpr}$ and yields a new heap that is entirely the same to $\heap$ modulo the fifth position.
+
+\begin{center}
+  \judgbox{\pstepto{\runtimetermvar}{\runtimetermvar[']}{\event}}{,,Primitive step from runtime--term $\runtimetermvar$ to $\runtimetermvar[']$ emitting $\event$.''}
+  \typerule{$e-get-\in$}{
+    \sandboxtag=\Omega.\sandboxtag &
+    \statevar.\store(x)=\tup{\loc}{\sandboxtag,\poison,m} &
+    \loc+n\in\statevar.\heap[^{\sandboxtag}]
+  }{
+    \pstepto{\rtt{\Omega}{\lget{x}{n}}}{\rtt{\Omega}{\heap[^{\sandboxtag}](\loc+n)}}{\tup{\ev{Get}\ \loc\ n}{\sandboxtag}}
+  }{e-get-in}
+  %
+  \typerule{$e-get-\notin$}{
+    \sandboxtag=\Omega.\sandboxtag &
+    \statevar.\store(x)=\tup{\loc}{\sandboxtag,\poison,m} &
+    \loc+n\notin\statevar.\heap[^{\sandboxtag}]
+  }{
+    \pstepto{\rtt{\Omega}{\lget{x}{n}}}{\rtt{\Omega}{1729}}{\tup{\ev{Get}\ \loc\ n}{\sandboxtag}}
+  }{e-get-notin}
+\end{center}
+
+\Cref{tr:e-get-in,tr:e-get-notin} both evaluate a read by looking up the metadata in the store.
+But, note that neither performs any bounds-checking and emits a $\ev{Get}\ \loc\ n$ event regardless of whether the access is out of bounds or not.
 

### 2023-06-18 popl/main.tex
Add preliminary definitions for case study.

@@
   \begin{aligned}
-  \mi{(Finals)}~\finalexprnoerr&\bnfdef\ \valueexpr \mid x\hspace{0.5cm}
-  \mi{(Results)}~\finalexpr\bnfdef\ \finalexprnoerr \mid stuck\hspace{0.5cm}
+  &\mi{(Finals)}~\finalexpr\bnfdef\ \valueexpr \mid x\hspace{1.5cm}
   \mi{(Values)}~\valueexpr\bnfdef\ n\\
   \mi{(Expressions)}~\expr&\bnfdef\ \finalexpr \mid \lbinop{\expr[_{1}]}{\expr[_{2}]} \mid \lget{x}{\expr} \mid \lnew{x}{\expr[_{1}]}{\expr[_{2}]} \mid \ldelete{x} \mid \lset{x}{\expr[_{1}]}{\expr[_{2}]} \\
-    & \mid \llet{x}{\expr[_{1}]}{\expr[_{2}]} \mid \lreturn{\expr} \mid \lcall{g}{\expr} \mid \lifz{\expr}{\expr[_{1}]}{\expr[_{2}]} \\
+    & \mid \llet{x}{\expr[_{1}]}{\expr[_{2}]} \mid \lreturn{\expr} \mid \lcall{g}{\expr} \mid \lifz{\expr}{\expr[_{1}]}{\expr[_{2}]} \mid \labort \\
   \mi{(State)}~\statevar&\bnfdef \tup{\cfstate}{\sandboxtag,\memstate} \hspace{0.5cm}
-  \mi{(Runtime~Terms)}~\runtimetermvar\bnfdef\rtt{\statevar}{\expr} \hspace{0.5cm}
+  \mi{(Runtime~Terms)}~\runtimetermvar\bnfdef\rtt{\statevar}{\expr} \mid \lightning \hspace{0.5cm}
     \text{where}\ \lbinop{}{}\in\left\{+,-,\times,<,/\right\}\\
@@
 Above is the base syntax of all the programming languages of this paper.
-Expressions ($\expr$) must not contain the result $stuck$ unless they are entailed as runtime--term ($\rtt{\statevar}{\expr}$).
+Expressions ($\expr$) can only be evaluated in the form of runtime--terms ($\rtt{\statevar}{\expr}$).
 States ($\statevar$) have a flow- ($\cfstate$) and a memory state ($\memstate$).
@@
 
-All programming languages presented in this section have an environmental semantics and all but $\src{L}$ are dynamically typed.
+All programming languages presented in this section have an environmental semantics and all but $\src{L_{\tmssafe}}$ are dynamically typed.
 Each evaluation may emit the empty event ($\emptyevent$), a program crash token ($\lightning$), or a pair containing a pre--event ($\event[_{b}]$) and the corresponding control--tag ($\sandboxtag$) indicating whether the event fired during an evaluation of a context or a component.
@@
 \begin{center}
-  \judgbox{\pstepto{\runtimetermvar}{\runtimetermvar[']}{\event}}{,,Primitive step from runtime--term $\runtimetermvar$ to $\runtimetermvar[']$ emitting $\event$.''}
-  \typerule{$e-get-\in$}{
-    \sandboxtag=\Omega.\sandboxtag &
-    \statevar.\store(x)=\tup{\loc}{\sandboxtag,\poison,m} &
+  \judgbox{\pstepto{\runtimetermvar}{\runtimetermvar[']}{\event}}{,,Primitive step from runtime--term $\runtimetermvar$ to $\runtimetermvar[']$ emitting event $\event$.''}
+  \typerule{$e-\text{get}-\in$}{
+    \sandboxtag=\statevar.\sandboxtag &
+    \statevar.\store(x)=\tup{\loc}{\sandboxtag,\poison,m} \\
     \loc+n\in\statevar.\heap[^{\sandboxtag}]
   }{
-    \pstepto{\rtt{\Omega}{\lget{x}{n}}}{\rtt{\Omega}{\heap[^{\sandboxtag}](\loc+n)}}{\tup{\ev{Get}\ \loc\ n}{\sandboxtag}}
+    \pstepto{\rtt{\statevar}{\lget{x}{n}}}{\rtt{\statevar}{\heap[^{\sandboxtag}](\loc+n)}}{\tup{\ev{Get}\ \loc\ n}{\sandboxtag}}
   }{e-get-in}
   %
-  \typerule{$e-get-\notin$}{
-    \sandboxtag=\Omega.\sandboxtag &
-    \statevar.\store(x)=\tup{\loc}{\sandboxtag,\poison,m} &
+  \typerule{$e-\text{get}-\notin$}{
+    \sandboxtag=\statevar.\sandboxtag &
+    \statevar.\store(x)=\tup{\loc}{\sandboxtag,\poison,m} \\
     \loc+n\notin\statevar.\heap[^{\sandboxtag}]
   }{
-    \pstepto{\rtt{\Omega}{\lget{x}{n}}}{\rtt{\Omega}{1729}}{\tup{\ev{Get}\ \loc\ n}{\sandboxtag}}
+    \pstepto{\rtt{\statevar}{\lget{x}{n}}}{\rtt{\statevar}{1729}}{\tup{\ev{Get}\ \loc\ n}{\sandboxtag}}
   }{e-get-notin}
+  %
+  \typerule{$e-\text{set}-\in$}{
+    \sandboxtag=\statevar.\sandboxtag &
+    \statevar.\store(x)=\tup{\loc}{\sandboxtag,\poison,m} \\
+    \loc+n\in\statevar.\heap[^{\sandboxtag}] &
+    \heap[^{\sandboxtag}_1] = \seth{\statevar.\heap[^{\sandboxtag}]}{\loc+n}{\valueexpr}
+  }{
+    \pstepto{\rtt{\statevar}{\lset{x}{n}{\valueexpr}}}{\rtt{\statevar\subst{\heap[^{\sandboxtag}_1]}{\heap[^{\sandboxtag}]}}{\valueexpr}}{\tup{\ev{Set}\ \loc\ n\ \valueexpr}{\sandboxtag}}
+  }{e-set-in}
+  %
+  \typerule{$e-\text{set}-\notin$}{
+    \sandboxtag=\statevar.\sandboxtag &
+    \statevar.\store(x)=\tup{\loc}{\sandboxtag,\poison,m} \\
+    \loc+n\notin\statevar.\heap[^{\sandboxtag}]
+  }{
+    \pstepto{\rtt{\statevar}{\lset{x}{n}{\valueexpr}}}{\rtt{\statevar}{\valueexpr}}{\tup{\ev{Set}\ \loc\ n\ \valueexpr}{\sandboxtag}}
+  }{e-set-notin}
+  %
+  \typerule{$e-\text{new}$}{
+    \fresh{\statevar}{z} &
+    \fresh{\statevar}{\loc} &
+    \heap[^{\sandboxtag}_1]=\growh{\statevar.\heap[^{\sandboxtag}]}{n} &
+    \store[_1]=\storeel{z}{\loc}{\statevar.\sandboxtag}{\poisonless}{n},\statevar.\store
+  }{
+    \pstepto{\rtt{\statevar}{\lnew{x}{n}{\expr}}}{\rtt{\statevar\subst{\heap[^{\sandboxtag}]}{\heap[^{\sandboxtag}_1]}\subst{\store}{\store[_1]}}{\expr\subst{z}{x}}}{\tup{\ev{Alloc}\ \loc\ n}{\sandboxtag}}
+  }{e-new}
+  %
+  \typerule{$e-\text{dealloc}$}{
+    \statevar.\store(x) = \tup{\loc}{\statevar.\sandboxtag,\poison,n} &
+    \store[_1]=\seth{\statevar.\store}{x}{\tup{\loc}{\statevar.\sandboxtag,\poisoned,n}}
+  }{
+    \pstepto{\rtt{\statevar}{\ldelete{x}}}{\rtt{\statevar\subst{\store}{\store[_1]}}{0}}{\tup{\ev{Dealoc}\ \loc}{\sandboxtag}}
+  }{e-dealloc}
+  %
+  \typerule{$e-\text{abort}$}{
+  }{
+    \pstepto{\rtt{\statevar}{\labort}}{\lightning}{\lightning}
+  }{e-abort}
 \end{center}
 
-\Cref{tr:e-get-in,tr:e-get-notin} both evaluate a read by looking up the metadata in the store.
-But, note that neither performs any bounds-checking and emits a $\ev{Get}\ \loc\ n$ event regardless of whether the access is out of bounds or not.
+Reading the value of a pointer is done by looking up the metadata in the store (\Cref{tr:e-get-in,tr:e-get-notin}).
+This is done without any bounds-checking and always emits a $\ev{Get}\ \loc\ n$ event regardless of whether the access is out of bounds or not.
+Similarily when setting a value (\Cref{tr:e-set-in,tr:e-set-notin}), a $\ev{Set}\ \loc\ n\ \valueexpr$ event is emitted regardless of wether the set is in or out of bounds.
+Allocation of objects (\Cref{tr:e-new}) does the expected, however, it makes use of the judgement $\fresh{\cdot}{\cdot}$ which is used throughout this paper to explicitly mark when and with respect to what environment a fresh variable is introduced.
+Deallocation (\Cref{tr:e-dealloc}) simply sets the poison--tag in the metadata of the associated pointer to be poisoned ($\poisoned$).
+The only way to abort an execution (\Cref{tr:e-abort}) is with an explicit call to $\labort$.
+However, evaluation of programs can get stuck, e.g., if $x$ is not bound in $\ldelete{x}$.
+
+In the following, $\neg\ctx=\comp$ and $\neg\comp=\ctx$ and also $\operatorname{comm}(\ctx)=\ \ctxtocomp$ and $\operatorname{comm}(\comp)=\ \comptoctx$.
+
+\begin{center}
+  \judgbox{\estepto[]{\runtimetermvar}{\runtimetermvar}{\event}}{,,Contextual step from runtime--term $\runtimetermvar$ to $\runtimetermvar[']$ emitting event $\event$.''}
+  %
+  \typerule{$e-\text{ectx}$}{
+    \pstepto{\rtt{\statevar}{\expr}}{\rtt{\statevar[']}{\expr[']}}{\event}
+  }{
+    \estepto[]{\rtt{\statevar}{\ectx\hole{\expr}}}{\rtt{\statevar[']}{\ectx\hole{\expr[']}}}{\event}
+  }{e-ectx}
+  %
+  \typerule{$e-\text{stuck}$}{
+    \pstepto{\rtt{\statevar}{\expr}}{\lightning}{\lightning}
+  }{
+    \estepto[]{\rtt{\statevar}{\ectx\hole{\expr}}}{\lightning}{\lightning}
+  }{e-stuck}
+  %
+  \typerule{$e-\text{ret}$}{
+    \statevar.\ectxstack=\tup{\ectx}{\text{foo}},\ectxstack['] &
+    \commlib\vdash\text{foo}:\Omega.\sandboxtag &
+    \operatorname{comm}(\sandboxtag)=\comm
+  }{
+    \estepto[]{\rtt{\statevar}{\ectx[']\hole{\lreturn{\valueexpr}}}}{\rtt{\statevar\subst{\neg\statevar.\sandboxtag}{\sandboxtag}\subst{\ectxstack[']}{\ectxstack}}{\ectx\hole{\valueexpr}}}{\tup{\ev{Ret}\ \comm\ \valueexpr}{\statevar.\sandboxtag}}
+  }{e-ret-notsame}
+  %
+  \typerule{$e-\text{start}$}{
+    \lfunction{\text{main}}{x}{e} \in \statevar.\library
+  }{
+    \estepto[]{\rtt{\statevar}{\lcall{\text{main}}{0}}}{\rtt{\statevar\subst{\ctx}{\sandboxtag}\subst{\tup{\hole{\cdot}}{\text{main}}}{\ectxstack}}{\expr\subst{0}{x}}}{\tup{\ev{Start}}{\comp}}
+  }{e-start}
+  %
+  \typerule{$e-\text{call}-\text{notsame}$}{
+    \lfunction{\text{foo}}{x}{e} \in \statevar.\library &
+    \commlib\vdash\text{foo}:\Omega.\sandboxtag &
+    \operatorname{comm}(\sandboxtag)=\comm
+  }{
+    \estepto[]{\rtt{\statevar}{\ectx\hole{\lcall{\text{foo}}{\valueexpr}}}}{\rtt{\statevar\subst{\neg\statevar.\sandboxtag}{\sandboxtag}\subst{\tup{\ectx}{\text{foo}},\statevar.\ectxstack}{\ectxstack}}{\expr\subst{\valueexpr}{x}}}{\tup{\ev{Call}\ \comm\ \text{foo}\ \valueexpr}{\statevar.\sandboxtag}}
+  }{e-call-notsame}
+  %
+  \judgbox{\commlib\vdash\text{foo}:\sandboxtag}{,,Given execution is in $\sandboxtag$, the symbol associated with the name $\text{foo}$ is in $\neg\sandboxtag$.''}
+  %
+  \typerule{ctx-to-comp}{
+    \text{foo}\in\commlib
+  }{
+    \commlib\vdash\text{foo}:\ctx
+  }{ctx-to-comp}
+  %
+  \typerule{comp-to-ctx}{
+    \text{foo}\notin\commlib
+  }{
+    \commlib\vdash\text{foo}:\comp
+  }{comp-to-ctx}
+\end{center}
+
+Contextual steps are as expected (\Cref{tr:e-ectx,tr:e-stuck}) up to calling (\Cref{tr:e-call-notsame}) and returning.
+For calls and returns, a bit of work has to be done to emit the right events.
+For example, the call to the main function emits just $\ev{Start}$ (\Cref{tr:e-start}) instead of $\ev{Call}\ \comptoctx\ \text{main}\ 0$, which is a design choice this paper does for convenience when reasoning about a call-chain.
+Ordinary calls, however, need to check wether a context switch is happening or not and push the current continuation and the name of the invoked function on the stack in the state (\Cref{tr:e-call-notsame}).
+Context switches are checked by looking at the current control--tag and wether the involved function or surrounding context is part of the list of names of component--level symbols ($\commlib$) or not (\Cref{tr:ctx-to-comp,tr:comp-to-ctx}).
 

### 2023-06-18 popl/main.tex
Start blue

@@
 
+\subsection{$\src{L_{\tmssafe}}$: A Temporal but Not Spatial Memory Safe Language}\label{subsec:ltms}
+\begin{gather*}
+  \begin{aligned}
+    \mi{(Pre\text{--}Types)}~\src{\type[_{e}]}&\bnfdef\src{\natt} \mid \src{\ptrn} \hspace{0.5cm}
+    \mi{(Pointer\ Qualifiers)}~\src{\ptrqual}\bnfdef\ \src{\fullq} \mid \src{\halfq} \hspace{0.5cm}
+    \mi{(Types)}~\src{\type}\bnfdef\src{\type_{e}} \mid \src{\type_{e}\to\bot} \\
+  \end{aligned}
+\end{gather*}
+
+\hltt{some type rules}
+
+$\src{L_{\tmssafe}}$ is a statically typed language where each well-typed term is temporal memory safe.
+To enforce this, the language marks pointers with respect to their ownership, where $\src{\fullq}$ means {\em owned} and $\src{\halfq}$ means {\em not owned}.
+An owned pointer can only be used by a deallocation, whereas reads and writes must be done on non-owned pointers.
+One can convert owned to non-owned pointers implicitly.
+Syntactic constraints together with the type system ensure that non-owned pointers never ,,live'' longer than owned pointers.
+For example, in an expression $\src{\llet{x}{\expr[_{1}]}{\expr[_{2}]}}$, if $y$ is a $\src{\wptr}$ in $\expr[_{1}]$, it must be a $\src{\ptr}$ in $\expr[_{2}]$.
+
 \subsection{Robust Temporal Memory Safety Preservation}\label{subsec:cs:tms}

### 2023-06-19 popl/main.tex
Add more defs

--- 2023-06-18 popl/main.tex
@@
 \newcommand\bmmax{0}
-\documentclass[acmsmall,review,screen,dvipsnames]{acmart}
+\documentclass[utf8,acmsmall,review,screen,dvipsnames]{acmart}
 
@@
 \usepackage{bm}
+
+% https://tex.stackexchange.com/questions/648845/sans-serif-uppercase-greek-no-longer-showing-in-acmart
+\DeclareMathAlphabet{\mathsf}{OT1}{LibertinusSans-LF}{m}{n}
+\SetMathAlphabet{\mathsf}{bold}{OT1}{LibertinusSans-LF}{bx}{n}
 
@@
 
+\subsection{$\src{L_{\tmssafe}}$: A Temporal but Not Spatial Memory Safe Language}\label{subsec:ltms}
+\begin{gather*}
+  \begin{aligned}
+    \mi{(Pre\text{--}Types)}~\src{\type[_{e}]}&\bnfdef\src{\natt} \mid \src{\ptrn} \hspace{0.5cm}
+    \mi{(Pointer\ Qualifiers)}~\src{\ptrqual}\bnfdef\ \src{\fullq} \mid \src{\halfq} \\
+    \mi{(Type~Environment)}~\src{\typenv}&\bnfdef\ \src{\hole{\cdot}} \mid \hastype{\src{x}}{\src{\type}},\src{\typenv} \hspace{0.5cm}
+    \mi{(Types)}~\src{\type}\bnfdef\src{\type[_{e}]} \mid \src{\type[_{e}]\to\bot} \mid \src{\type[^{(1)}_{e}]\to\type[^{(2)}_{e}]}
+  \end{aligned}
+\end{gather*}
+\begin{center}
+  \judgbox{\tcheck{\src{\typenv}}{\src{\expr}}{\src{\type}}}{,,Under environment $\src{\typenv}$, expression $\src{\expr}$ has type $\src{\type}$.''}
+  %
+  \typerule{t-$\src{x}$}{
+    \src{\typenv}(\src{x})=\src{\type} &
+    \notowned{\src{\typenv}}
+  }{
+    \tcheck{\src{\typenv}}{\src{x}}{\src{\type}}
+  }{t-x}
+  %
+  \typerule{t-$\src{get}$}{
+    \tcheck{\src{\typenv[_2]}}{\src{x}}{\src{\wptr}} &
+    \tcheck{\src{\typenv[_1]}}{\src{\expr}}{\src{\natt}} &
+  }{
+    \tcheck{\typenvsplit{\src{\typenv[_1]}}{\src{\typenv[_2]}}}{\src{\lget{x}{\expr}}}{\src{\natt}}
+  }{t-get}
+  %
+  \typerule{t-$\src{let}$}{
+    \tcheck{\src{\typenv[_1]}}{\src{\expr[_1]}}{\src{\type[_e]}} &
+    \tcheck{\hastype{\src{x}}{\src{\type[_e]}},\src{\typenv[_2]}}{\src{\expr[_2]}}{\src{\type}}
+  }{
+    \tcheck{\typenvsplit{\src{\typenv[_1]}}{\src{\typenv[_2]}}}{\src{\llet{x}{\expr[_1]}{\expr[_2]}}}{\src{\type}}
+  }{t-let}
+  %
+  \typerule{t-$\src{delete}$}{
+    \src{\typenv}(\src{x})=\src{\ptr} &
+    \notowned{\src{\typenv}\setminus\left\{\hastype{\src{x}}{\src{\ptr}}\right\}}
+  }{
+    \tcheck{\src{\typenv}}{\src{\ldelete{x}}}{\src{\natt}}
+  }{t-delete}
+  %
+  \typerule{t-$\src{abort}$}{
+    \notowned{\src{\typenv}}
+  }{
+    \tcheck{\src{\typenv}}{\src{\labort}}{\src{\type}}
+  }{t-abort}
+  %
+  \typerule{t-$\src{call}$}{
+    \inttype{\src{\type[^{(1)}_e]}} &
+    \inttype{\src{\type[^{(2)}_e]}} \\
+    \tcheck{\src{\typenv}}{\src{foo}}{\src{\type[^{(1)}_e]\to\type[^{(2)}_e]}} &
+    \tcheck{\src{\typenv}}{\src{\expr}}{\src{\type[^{(1)}_e]}}
+  }{
+    \tcheck{\src{\typenv}}{\src{\lcall{foo}{\expr}}}{\src{\type[^{(2)}_e]}}
+  }{t-call}
+  %
+  \typerule{t-$\src{return}$}{
+    \inttype{\src{\type[_e]}} &
+    \tcheck{\src{\typenv}}{\src{\expr}}{\src{\type[_e]}}
+  }{
+    \tcheck{\src{\typenv}}{\src{\lreturn{\expr}}}{\src{\type[_e]\to\bot}}
+  }{t-return}
+\end{center}
+
+$\src{L_{\tmssafe}}$ is a statically typed language where each well-typed term is temporal memory safe.
+Most parts of the type system are entirely standard, so only the interesting bits are shown here.
+To enforce this, the language marks pointers with respect to their ownership, where $\src{\fullq}$ means {\em owned} and $\src{\halfq}$ means {\em not owned}.
+An owned pointer can only be used by a deallocation (\Cref{tr:t-delete}), whereas reads and writes must be done on non-owned pointers.
+One can convert owned to non-owned pointers implicitly through context splitting $\typenvsplit{\src{\typenv[_{1}]}}{\src{\typenv[_{2}]}}$.\MK{list some rules and explain it a bit}
+Syntactic constraints together with the type system ensure that non-owned pointers never ,,live'' longer than owned pointers.
+For example, in an expression $\src{\llet{x}{\expr[_{1}]}{\expr[_{2}]}}$, if $y$ is a $\src{\ptr}$ in $\src{\expr[_{1}]}$, it cannot appear at all in $\src{\expr[_{2}]}$, but it can if it were of type $\src{\wptr}$ (\Cref{tr:t-let}).
+This mechanism also prevents terms like $\src{\lget{x}{\ldelete{x}}}$ from typechecking (\Cref{tr:t-get}).
+At leafs of the type derivation tree, the typing environment must not contain owned pointers, which is ensured with the $\notowned{\src{\typenv}}$ judgement (\Cref{tr:t-abort,tr:t-x}).
+Lastly, calling or returning requires passed values to have an interface type ($\inttype{\src{\type[_{e}]}}$), which are just $\src{\natt}$ (\Cref{tr:t-call,tr:t-return}).
+To ensure that no pointers live after a return, the type of a return is that of a continuation, which is only permitted to propagate towards the end of a term.
+For example, we cannot bind $\src{\lreturn{e}}$ to an identifier in the typecheck of $\src{let}$, because the expression whose result is written into the binder has to be a pre--type, which the type of $\src{\lreturn{e}}$ is not.
+
+\MKin{add filter $\theta$ projecting to specification events}
+
+\begin{theorem}[$\src{L}$-programs are \gls{tms}]\label{thm:wt:tms}
+  If $\progstepto{\src{\prog{\library_{\ctx}}{\library_{\comp}}}}{\src{\rtt{\statevar}{\valueexpr}}}{\src{\trace}}$, then $\montmssafe{\filterLtms{\src{\trace}}}$.\Coqed
+\end{theorem}
+
+\subsection{$\trg{L}$: An Unsafe Language}\label{subsec:ltms}
+
+\MKin{add definitions and description; some syntax + dynamics}
+
 \subsection{Robust Temporal Memory Safety Preservation}\label{subsec:cs:tms}
+
+\MKin{compiler $\src{L}\to\trg{L}$}
+\MKin{x-lang relations}
+\MKin{backtranslation}
+\MKin{rtp of tms}
+
 \subsection{Robust (Spatial) Memory Safety Preservation}\label{subsec:cs:ms}
+
+\MKin{compiler $\trg{L}\to\irl{L_{MS}}$}
+\MKin{backward sim}
+\MKin{rtp of sms}
+
+\MKin{composition: rtp of ms}
+
 \subsection{Optimizing Compilers}\label{subsec:cs:opts}
+
+\MKin{compiler $\irl{L_{MS}}\to\irl{L_{MS}}$ doing \gls{dce}}
+\MKin{compiler $\irl{L_{MS}}\to\irl{L_{MS}}$ doing \gls{cf}}
+
+\MKin{backward sims}
+\MKin{rtp of ms}
+\MKin{composition: rtp of ms, order independent}
+
+\subsection{$\obj{L_{\scct}}$: A Language with Access to a Model Specific Register}\label{subsec:ltms}
+
+\MKin{add definitions and description; some syntax + dynamics}
+
 \subsection{Robust Strict Cryptographic Constant Time Preservation}\label{subsec:cs:scct}
+
+\MKin{compiler $\irl{L_{MS}}\to\obj{L_{\scct}}$}
+\MKin{x-lang relations}
+\MKin{backtranslation}
+\MKin{rtp of scct}
+
+\MKin{composition: rtp of everything}
 

### 2023-06-20 popl/main.tex
Introduce filter

@@
 
-\MKin{add filter $\theta$ projecting to specification events}
-
-\begin{theorem}[$\src{L}$-programs are \gls{tms}]\label{thm:wt:tms}
+\begin{center}
+  \judgbox{\thelocmap(\src{\loc})=\loc}{,,Map an $\src{L_{\tmssafe}}$ location to a specification location $\loc$.''}
+  \judgbox{\filterLtms{\src{\event}}=\event}{,,Project an $\src{L_{\tmssafe}}$ event $\src{\event}$ to a specification event $\event$.''}
+  \typerule{$\src{L_{\tmssafe}}$-filter-context}{
+    \src{\event[_b]}\not=\src{\lightning}
+  }{
+    \filterLtms{\src{\tup{\event[_b]}{\ctx}} = \emptyevent}
+  }{ltms-filter-context}
+  %
+  \typerule{$\src{L_{\tmssafe}}$-filter-abort}{
+  }{
+    \filterLtms{\src{\lightning}} = \lightning
+  }{ltms-filter-abort}
+  %
+%  \typerule{$\src{L_{\tmssafe}}$-filter-start}{
+%  }{
+%    \filterLtms{\src{\tup{Start}{\comp}}} = \emptyevent
+%  }{ltms-filter-start}
+  %
+  \typerule{$\src{L_{\tmssafe}}$-filter-alloc}{
+    \locmapsto{\src{\loc}}{\loc} &
+    \src{n}=n
+  }{
+    \filterLtms{\src{\tup{Alloc\ \loc\ n}{\comp}}} = \tup{\ev{Alloc}\ \loc\ n}{\comp,\unlock}
+  }{ltms-filter-alloc}
+\end{center}
+To prove that well-typed $\src{L_{\tmssafe}}$ programs are temporal memory safe, a map from $\src{L_{\tmssafe}}$ to specification events (see \Cref{subsec:propdefs}) allows the use of \Cref{def:monsat}.
+The map is mostly entirely straight-forward.
+Most notably, it filters any action that the context does (\Cref{tr:ltms-filter-context}), since a context is free to do anything.
+Of course, a context invoking $\src{\labort}$ to crash the program must not go undetected (\Cref{tr:ltms-filter-abort}).
+As for the component, pre--events are essentially recolored to go from blue $\src{L_{\tmssafe}}$ events to the events of the most general trace model (\Cref{tr:ltms-filter-alloc}).
+Events that carry locations make use of a location--map from blue $\src{L_{\tmssafe}}$ locations to the more abstract ones.
+
+\MKin{show at least one forward sim lemma or not?}
+
+\begin{lemma}[$\src{L_{\tmssafe}}$-programs are \gls{tms}]\label{lem:wt:tmsmon}
   If $\progstepto{\src{\prog{\library_{\ctx}}{\library_{\comp}}}}{\src{\rtt{\statevar}{\valueexpr}}}{\src{\trace}}$, then $\montmssafe{\filterLtms{\src{\trace}}}$.\Coqed
+\end{lemma}
+\begin{theorem}[$\src{L_{\tmssafe}}$-programs are \gls{tms}]\label{thm:wt:tms}
+  For any $\src{\library_{\comp}}$, $\rsat{\src{\library_{\comp}}}{\tmssafe}$ \Coqed
 \end{theorem}
@@
 
-\subsection{$\obj{L_{\scct}}$: A Language with Access to a Model Specific Register}\label{subsec:ltms}
+\subsection{$\obj{L_{\scctsafe}}$: A Language with Access to a Model Specific Register}\label{subsec:ltms}
 
@@
 
-\MKin{compiler $\irl{L_{MS}}\to\obj{L_{\scct}}$}
+\MKin{compiler $\irl{L_{MS}}\to\obj{L_{\scctsafe}}$}
 \MKin{x-lang relations}

### 2023-06-20 popl/main.tex
Program behaviour

@@
 
+\begin{center}
+  \judgbox{\progstepto{\prog{\library_{\ctx}}{\library_{\comp}}}{\runtimetermvar}{\trace}}{,,Program $\prog{\library_{\ctx}}{\library_\comp}$ evaluates to $\runtimetermvar$, emitting trace $\trace$.''}
+  %
+  \typerule{ep-run}{
+    \link{\library_{\ctx}}{\library_{\comp}}=\library &
+    \commlib=\operatorname{dom}\library_{\comp} \\
+    \estepto{\rtt{\tup{\library}{\commlib,\emptyset}}{\lcall{\text{main}}{0}}}{\rtt{\statevar}{\valueexpr}}{\trace}
+  }{
+    \progstepto{\prog{\library_{\ctx}}{\library_{\comp}}}{\rtt{\statevar}{\valueexpr}}{\trace}
+  }{ep-run}
+  %
+  \typerule{ep-crash}{
+    \link{\library_{\ctx}}{\library_{\comp}}=\library &
+    \commlib=\operatorname{dom}\library_{\comp} \\
+    \estepton{\rtt{\tup{\library}{\commlib,\emptyset}}{\lcall{\text{main}}{0}}}{\lightning}{\trace\cdot\lightning}
+  }{
+    \progstepto{\prog{\library_{\ctx}}{\library_{\comp}}}{\lightning}{\trace\cdot\lightning}
+  }{ep-crash}
+\end{center}
+
+The notation $\tup{\library}{\commlib,\emptyset}$ unfolds to $\statevar=\tup{\cfstate}{\comp,\memstate}$, where $\cfstate=\tup{\library}{\commlib,\hole{\cdot}}$ and $\memstate=\tup{\hole{\cdot}}{\hole{\cdot},\hole{\cdot}}$.
+Linking $\link{\library_{\ctx}}{\library_{\comp}}$, while gluing $\library_{\ctx}$ and $\library_{\comp}$ together, ensures that there is no symbol with the name $\text{main}$ in $\library_{\comp}$.
+Programs can either be run successfully (\Cref{tr:ep-run}) or fail (\Cref{tr:ep-crash}), which requires the n-step relation instead of star step, since star step models a big--step style semantics without failure.
+To make use of \Cref{def:propsat,def:proprsat}, {\em behavior} is introduced as follows.
+
+\begin{definition}[Program Behavior]\label{def:behav}
+  Given a program $\prog{\library_{\ctx}}{\library_{\comp}}$, the behavior of it is a set containing all traces $\trace$ such that there is a runtime--term $\runtimetermvar$ with $\progstepto{\prog{\library_{\ctx}}{\library_{\comp}}}{\runtimetermvar}{\trace}$.
+\end{definition}
+
 \subsection{$\src{L_{\tmssafe}}$: A Temporal but Not Spatial Memory Safe Language}\label{subsec:ltms}

### 2023-06-20 popl/main.tex
Compiler

@@
   }{
-    \pstepto{\rtt{\statevar}{\ldelete{x}}}{\rtt{\statevar\subst{\store}{\store[_1]}}{0}}{\tup{\ev{Dealoc}\ \loc}{\sandboxtag}}
+    \pstepto{\rtt{\statevar}{\ldelete{x}}}{\rtt{\statevar\subst{\store}{\store[_1]}}{0}}{\tup{\ev{Dealloc}\ \loc}{\sandboxtag}}
   }{e-dealloc}
@@
     \mi{(Type~Environment)}~\src{\typenv}&\bnfdef\ \src{\hole{\cdot}} \mid \hastype{\src{x}}{\src{\type}},\src{\typenv} \hspace{0.5cm}
-    \mi{(Types)}~\src{\type}\bnfdef\src{\type[_{e}]} \mid \src{\type[_{e}]\to\bot} \mid \src{\type[^{(1)}_{e}]\to\type[^{(2)}_{e}]}
+    \mi{(Types)}~\src{\type}\bnfdef\src{\type[_{e}]} \mid \src{\type[_{e}]\to\bot} \mid \src{\type[^{(1)}_{e}]\to\type[^{(2)}_{e}]} \\
+    \mi{(Symbols)}~\src{\function}&\bnfdef\src{\lfunction{g}{x:\type[_{e}]\to\type[_{e}]}{\expr}} \\
   \end{aligned}
@@
 
-\subsection{$\trg{L}$: An Unsafe Language}\label{subsec:ltms}
-
-\MKin{add definitions and description; some syntax + dynamics}
+\subsection{$\trg{L}$: An Unsafe Language}\label{subsec:lsms}
+
+\begin{gather*}
+  \begin{aligned}
+    \mi{(Expressions)}~\trg{\expr}\bnfdef\ \dots \mid \trg{\lispoisoned{x}} \mid \trg{\lpair{\expr[_{1}]}{\expr[_{2}]}} &\mid \trg{\lproja{\expr}} \mid \trg{\lprojb{\expr}} \mid \trg{\lhast{\expr}{\type}} \hspace{0.5cm}
+    \mi{(Types)}~\trg{\type}\bnfdef\ \trg{\natt} \mid \trg{\natt\times\natt} \\
+    \mi{(Values)}~\trg{\valueexpr}&\bnfdef\ \dots \mid \trg{\lpair{n_{1}}{n_{2}}}
+  \end{aligned}
+\end{gather*}
+\begin{center}
+  \typerule{e-$\trg{\lispoisoned{}}$-yes}{
+    \trg{\statevar.\store}(\trg{x})=\trg{\tup{\loc}{\statevar.\sandboxtag,\poisoned,n}}
+  }{
+    \pstepto{\rtt{\trg{\statevar}}{\trg{\lispoisoned{x}}}}{\rtt{\trg{\statevar}}{\trg{0}}}{\trg{\emptyevent}}
+  }{e-ispoisoned-yes}
+  %
+  \typerule{e-$\trg{\lispoisoned{}}$-no}{
+    \trg{\statevar.\store}(\trg{x})=\trg{\tup{\loc}{\statevar.\sandboxtag,\poisonless,n}}
+  }{
+    \pstepto{\rtt{\trg{\statevar}}{\trg{\lispoisoned{x}}}}{\rtt{\trg{\statevar}}{\trg{1}}}{\trg{\emptyevent}}
+  }{e-ispoisoned-no}
+  %
+  \typerule{e-proj1}{
+  }{
+    \pstepto{\rtt{\trg{\statevar}}{\trg{\lproja{\lpair{n_1}{n_2}}}}}{\rtt{\trg{\statevar}}{\trg{n_1}}}{\trg{\emptyevent}}
+  }{e-proj1}
+  %
+  \typerule{e-$\trg{\lhast{n}{\natt}}$}{
+  }{
+    \pstepto{\rtt{\trg{\statevar}}{\trg{\lhast{n}{\natt}}}}{\rtt{\trg{\statevar}}{\trg{0}}}{\trg{\emptyevent}}
+  }{e-n-has-nat}
+  %
+  \typerule{e-$\trg{\lhast{x}{\natt}}$}{
+  }{
+    \pstepto{\rtt{\trg{\statevar}}{\trg{\lhast{x}{\natt}}}}{\rtt{\trg{\statevar}}{\trg{1}}}{\trg{\emptyevent}}
+  }{e-x-has-nat}
+  %
+  \typerule{e-$\trg{\lhast{n}{\natt\times\natt}}$}{
+  }{
+    \pstepto{\rtt{\trg{\statevar}}{\trg{\lhast{n}{\natt\times\natt}}}}{\rtt{\trg{\statevar}}{\trg{1}}}{\trg{\emptyevent}}
+  }{e-n-has-nattimesnat}
+\end{center}
+
+The $\trg{L}$ language does not provide any guarantees with regards to memory safety.
+The syntax of it extends the unified syntax for all languages (\Cref{subsec:cs:defs}) with pairs of natural numbers (\Cref{tr:e-proj1}) and two different kinds of dynamic checks.
+One of those checks allows programmers to investigate wether a pointer is freed or not (\Cref{tr:e-ispoisoned-yes,tr:e-ispoisoned-no}).
+The other check is a simple dynamic type check that evaluates to $\trg{0}$ if the check suceeds and to $\trg{1}$ otherwise (\Cref{tr:e-n-has-nat,tr:e-x-has-nat,tr:e-n-has-nattimesnat}).
+As before, not all rules are shown due to their straightforwardness.
+The projection $\filterL{\trace}$ is entirely similar to $\filterLtms{\trace}$.
 
@@
 
-\MKin{compiler $\src{L}\to\trg{L}$}
+\begin{center}
+  $$
+  \begin{array}{rcl}
+    \cca(\src{\lbinop{\expr[_{1}]}{\expr[_{2}]}}) & = & \lbinop[\trg]{\left[\cca(\src{\expr[_{1}]})\right]}{\left[\cca(\src{\expr[_{2}]})\right]} \\ [0.33cm]
+    \cca(\src{\lfunction{g}{x:\natt\to\type[_{e}]}{\expr}}) & = & \lfunction[\trg]{\trg{g}}{\trg{x}}{\lifz[\trg]{\trg{\lhast{x}{\natt}}}{%
+                                                                                                \left[\cca(\src{\expr})\right] %
+                                                                                                 }{\labort[\trg]}}
+  \end{array}
+  $$
+\end{center}
+
+The only case where the compiler does something interesting is for top-level functions.
+Here, a dynamic typecheck is inserted to respect the calling-convention.
+Since $\trg{L}$ has no static typechecks, it could happen that a bogus context $\trg{\library_{\ctx}}$ calls this function expecting a natural number with $\trg{\lpair{17}{29}}$.
+The compiler ensures that execution does not proceed in cases like that.
+
 \MKin{x-lang relations}
@@
 
-\subsection{$\obj{L_{\scctsafe}}$: A Language with Access to a Model Specific Register}\label{subsec:ltms}
+\subsection{$\obj{L_{\scctsafe}}$: A Language with Access to a Model Specific Register}\label{subsec:lscct}
 

### 2023-06-20 popl/main.tex
SMS compiler

@@
 Here, a dynamic typecheck is inserted to respect the calling-convention.
-Since $\trg{L}$ has no static typechecks, it could happen that a bogus context $\trg{\library_{\ctx}}$ calls this function expecting a natural number with $\trg{\lpair{17}{29}}$.
+Since $\trg{L}$ has no static typechecks, it could happen that a bogus context $\trg{\library_{\ctx}}$ calls a symbol accepting a $\src{\natt}$ with $\trg{\lpair{17}{29}}$.
 The compiler ensures that execution does not proceed in cases like that.
-
-\MKin{x-lang relations}
+This paper considers symbols on $\src{\natt}$.
+
+\MKin{x-lang relations $\approx_{\delta;\src{L}}$, $\multimap_{\delta;\src{L}}$, as well as $\cong_{\delta;\src{X}}$}
 \MKin{backtranslation}
-\MKin{rtp of tms}
+
+\begin{theorem}[Compiler $\cca$ is secure with respect to \gls{tms}]\label{thm:cca:rtp:tms}
+  $\rtp{\cca}{\tmssafe}$ % \Coqed
+\end{theorem}
 
@@
 
-\MKin{compiler $\trg{L}\to\irl{L_{MS}}$}
+In this section, the language $\irl{L_{\mssafe}}$ is exactly the same as $\trg{L}$ and are considered equal.
+However, it is still named and collored differently to emphasize the fact that after this section, compiled $\irl{L_{\mssafe}}$ programs are \gls{ms}.
+
+\begin{center}
+  $$
+  \begin{array}{rcl}
+    \ccb(\trg{\lnew{x}{\expr[_{1}]}{\expr[_{2}]}}) & = & \llet[\irl]{\irl{x_{SIZE}}}{\ccb(\trg{\expr[_{1}]})}{\lnew[\irl]{\irl{x}}{\irl{x_{SIZE}}}{\ccb(\trg{\expr[_{2}]})}} \\
+    \ccb(\trg{\lget{x}{\expr}}) & = & \llet[\irl]{\irl{x_{n}}}{\ccb(\trg{\expr})}{\irl{\lifz{0\le x_{n}<x_{SIZE}}{\lget{x}{x_{n}}}{\labort}}} \\
+    \ccb(\trg{\lset{x}{\expr[_{1}]}{\expr[_{2}]}}) & = & \llet[\irl]{\irl{x_{n}}}{\ccb(\trg{\expr[_{1}]})}{\lifz[\irl]{\irl{0\le x_{n}<x_{SIZE}}}{\lset[\irl]{\irl{x}}{\irl{x_{n}}}{\ccb(\trg{\expr[_{2}]})}}{\irl{\labort}}} \\
+  \end{array}
+  $$
+\end{center}
+
+Up to getting and setting memory, the compiler is the identity function.
+For memory accesses, however, a bounds--check is inserted that enforces \gls{sms}.
+To this end, the compiler introduces another, fresh identifier $\irl{x_{SIZE}}$ for each allocation that binds $\irl{x}$ to keep track of the allocation size.
+
 \MKin{backward sim}
-\MKin{rtp of sms}
-
-\MKin{composition: rtp of ms}
+\begin{theorem}[Compiler $\ccb$ is secure with respect to \gls{sms}]\label{thm:ccb:rtp:sms}
+  $\rtp{\ccb}{\smssafe}$ % \Coqed
+\end{theorem}
+
+The composition of $\cca$ and $\ccb$ attains full memory safety using \Cref{thm:rtp,thm:cca:rtp:tms,thm:ccb:rtp:sms}.
+
+\begin{theorem}[Compiler $\cca\circ\ccb$ is secure with respect to \gls{ms}]\label{thm:ccab:rtp:ms}
+  $\rtp{\cca\circ\ccb}{\mssafe}$ % \Coqed
+\end{theorem}
 
@@
 
-\MKin{compiler $\irl{L_{MS}}\to\irl{L_{MS}}$ doing \gls{dce}}
-\MKin{compiler $\irl{L_{MS}}\to\irl{L_{MS}}$ doing \gls{cf}}
+\MKin{compiler $\ccdce$ doing \gls{dce}}
+\MKin{compiler $\cccf$ doing \gls{cf}}
 

### 2023-06-22 popl/main.tex
Add sectioning

@@
 For states $\runtimetermvar$, it is assumed that there exists a judgement $\isval{\runtimetermvar}$ that determines wether $\runtimetermvar$ is a final state or not.
-\Cref{fig:starstep:and:nstep}\MK{appendix?} defines the star-step relation $\stepto{\runtimetermvar[_{1}]}{\runtimetermvar[_{2}]}{\trace}$, essentially modelling a big-step style evaluation, and the $n$-step relation $\stepton{\runtimetermvar[_{1}]}{\runtimetermvar[_{2}]}{\trace}$ that evaluates a program for $n$ steps.
+Another assumption is the existence of a generic star-step relation $\stepto{\runtimetermvar[_{1}]}{\runtimetermvar[_{2}]}{\trace}$, essentially modelling a big-step style evaluation, and the existence of an $n$-step relation $\stepton{\runtimetermvar[_{1}]}{\runtimetermvar[_{2}]}{\trace}$ that evaluates a program for $n$ steps.
 Both relations filter $\varepsilon$ events.
 
-\begin{figure}[t]
-  \judgbox{\stepto{r}{r'}{\trace}}{,,State $\runtimetermvar$ steps to final state $\runtimetermvar[']$, emitting events $\trace$.''}$\;$\\
-  %
-  \typerule{star-step-refl}{
-    \isval{\runtimetermvar}
-  }{
-    \stepto{\runtimetermvar}{\runtimetermvar}{\hole{\cdot}}
-  }{star-step-refl}
-  %
-  \typerule{star-step-trans}{
-    \event\not=\varepsilon &
-    \stepto[]{\runtimetermvar}{\runtimetermvar[']}{\event} &
-    \stepto{\runtimetermvar[']}{\runtimetermvar['']}{\trace}
-  }{
-    \stepto{\runtimetermvar}{\runtimetermvar['']}{\event\cdot\trace}
-  }{star-step-trans}
-  %
-  \typerule{star-step-ign}{
-    \stepto[]{\runtimetermvar}{\runtimetermvar[']}{\varepsilon} &
-    \stepto{\runtimetermvar[']}{\runtimetermvar['']}{\trace}
-  }{
-    \stepto{\runtimetermvar}{\runtimetermvar['']}{\trace}
-  }{star-step-ign}
-
-  \judgbox{\stepton{\runtimetermvar}{\runtimetermvar[']}{\trace}}{,,State $r$ does $n$ steps to final state $\runtimetermvar[']$, emitting events $\trace$.''}$\;$\\
-  %
-  \typerule{n-step-refl}{
-  }{
-    \stepton[0]{\runtimetermvar}{\runtimetermvar}{\hole{\cdot}}
-  }{n-step-refl}
-  %
-  \typerule{n-step-trans}{
-    \event\not=\varepsilon &
-    \stepto[]{\runtimetermvar}{\runtimetermvar[']}{\event} &
-    \stepton{\runtimetermvar[']}{\runtimetermvar['']}{\trace}
-  }{
-    \stepton[n+1]{\runtimetermvar}{\runtimetermvar['']}{\event\cdot\trace}
-  }{n-step-trans}
-  %
-  \typerule{n-step-ign}{
-    \stepto[]{\runtimetermvar}{\runtimetermvar[']}{\varepsilon} &
-    \stepton{\runtimetermvar[']}{\runtimetermvar['']}{\trace}
-  }{
-    \stepton[n+1]{\runtimetermvar}{\runtimetermvar['']}{\trace}
-  }{n-step-ign}
-
-  \caption{Star-step and $n$-step relation for a given step relation $\stepto{\runtimetermvar[_{1}]}{\runtimetermvar[_{2}]}{\event}$}\label{fig:starstep:and:nstep}
-\end{figure}
-%
 The \emph{behavior} of a whole program $w$ ($\behav{w}$) is a set containing all traces the program may emit during execution.
@@
 
-      \draw[->] (S) to[sloped,above] node {\text{\scriptsize\gls{tms}}} (T);
+      \draw[->] (S) to[sloped,above] node (tmsedge) {\text{\scriptsize\gls{tms}}} (T);
       \draw[->] (T) to[sloped,above] node {\text{\scriptsize\gls{sms}}} (M);
@@
       \draw[->] (C1) to[sloped,above] node {\text{\scriptsize\gls{scct}}} (O);
+
+      % Sections
+      \node[above=1.5 of tmsedge] (sectms) {{\scriptsize\Cref{subsec:cs:tms}}};
+      \node[right=0.0 of sectms] (secsms) {{\scriptsize\Cref{subsec:cs:ms}}};
+      \node[right=1.25 of secsms] (secopts) {{\scriptsize\Cref{subsec:cs:opts}}};
+      \node[right=1.25 of secopts] (secscct) {{\scriptsize\Cref{subsec:cs:scct}}};
+
+      \draw[dashed] ($(sectms)-(0.7,0)$) -- ($(sectms)-(0.7,1.75)$);
+      \draw[dashed] ($(sectms)-(-0.7,0)$) -- ($(sectms)-(-0.7,1.75)$);
+      \draw[dashed] ($(secsms)-(-0.7,0)$) -- ($(secsms)-(-0.7,1.75)$);
+      \draw[dashed] ($(secscct)-(0.7,0)$) -- ($(secscct)-(0.7,0.75)$);
+      \draw[dashed] ($(secscct)-(-0.7,0)$) -- ($(secscct)-(-0.7,0.75)$);
     \end{tikzpicture}
@@
   \begin{aligned}
-    \mi{(Pre\text{--}Types)}~\src{\type[_{e}]}&\bnfdef\src{\natt} \mid \src{\ptrn} \hspace{0.5cm}
-    \mi{(Pointer\ Qualifiers)}~\src{\ptrqual}\bnfdef\ \src{\fullq} \mid \src{\halfq} \\
-    \mi{(Type~Environment)}~\src{\typenv}&\bnfdef\ \src{\hole{\cdot}} \mid \hastype{\src{x}}{\src{\type}},\src{\typenv} \hspace{0.5cm}
-    \mi{(Types)}~\src{\type}\bnfdef\src{\type[_{e}]} \mid \src{\type[_{e}]\to\bot} \mid \src{\type[^{(1)}_{e}]\to\type[^{(2)}_{e}]} \\
-    \mi{(Symbols)}~\src{\function}&\bnfdef\src{\lfunction{g}{x:\type[_{e}]\to\type[_{e}]}{\expr}} \\
+    \mi{(Pre\text{--}Types)}~\src{\type[_{e}]}\bnfdef&\ \src{\natt} \mid \src{\ptrn} &
+    \mi{(Type~Environment)}~\src{\typenv}\bnfdef&\ \src{\hole{\cdot}} \mid \hastype{\src{x}}{\src{\type}},\src{\typenv} \\
+    \mi{(Types)}~\src{\type}\bnfdef&\ \src{\type[_{e}]} \mid \src{\type[_{e}]\to\bot} \mid \src{\type[^{(1)}_{e}]\to\type[^{(2)}_{e}]} &
+    \mi{(Pointer\ Qualifiers)}~\src{\ptrqual}\bnfdef&\ \src{\fullq} \mid \src{\halfq} \\
+    \mi{(Symbols)}~\src{\function}\bnfdef&\src{\lfunction{g}{x:\type[_{e}]\to\type[_{e}]}{\expr}} \\
   \end{aligned}
@@
   %
-  \typerule{e-$\trg{\lispoisoned{}}$-no}{
-    \trg{\statevar.\store}(\trg{x})=\trg{\tup{\loc}{\statevar.\sandboxtag,\poisonless,n}}
-  }{
-    \pstepto{\rtt{\trg{\statevar}}{\trg{\lispoisoned{x}}}}{\rtt{\trg{\statevar}}{\trg{1}}}{\trg{\emptyevent}}
-  }{e-ispoisoned-no}
-  %
-  \typerule{e-proj1}{
-  }{
-    \pstepto{\rtt{\trg{\statevar}}{\trg{\lproja{\lpair{n_1}{n_2}}}}}{\rtt{\trg{\statevar}}{\trg{n_1}}}{\trg{\emptyevent}}
-  }{e-proj1}
+%  \typerule{e-$\trg{\lispoisoned{}}$-no}{
+%    \trg{\statevar.\store}(\trg{x})=\trg{\tup{\loc}{\statevar.\sandboxtag,\poisonless,n}}
+%  }{
+%    \pstepto{\rtt{\trg{\statevar}}{\trg{\lispoisoned{x}}}}{\rtt{\trg{\statevar}}{\trg{1}}}{\trg{\emptyevent}}
+%  }{e-ispoisoned-no}
   %
@@
   %
-  \typerule{e-$\trg{\lhast{x}{\natt}}$}{
-  }{
-    \pstepto{\rtt{\trg{\statevar}}{\trg{\lhast{x}{\natt}}}}{\rtt{\trg{\statevar}}{\trg{1}}}{\trg{\emptyevent}}
-  }{e-x-has-nat}
+%  \typerule{e-$\trg{\lhast{x}{\natt}}$}{
+%  }{
+%    \pstepto{\rtt{\trg{\statevar}}{\trg{\lhast{x}{\natt}}}}{\rtt{\trg{\statevar}}{\trg{1}}}{\trg{\emptyevent}}
+%  }{e-x-has-nat}
   %
@@
 The syntax of it extends the unified syntax for all languages (\Cref{subsec:cs:defs}) with pairs of natural numbers (\Cref{tr:e-proj1}) and two different kinds of dynamic checks.
-One of those checks allows programmers to investigate wether a pointer is freed or not (\Cref{tr:e-ispoisoned-yes,tr:e-ispoisoned-no}).
-The other check is a simple dynamic type check that evaluates to $\trg{0}$ if the check suceeds and to $\trg{1}$ otherwise (\Cref{tr:e-n-has-nat,tr:e-x-has-nat,tr:e-n-has-nattimesnat}).
-As before, not all rules are shown due to their straightforwardness.
+One of those checks allows programmers to investigate wether a pointer is freed or not (\Cref{tr:e-ispoisoned-yes}).
+The other check is a simple dynamic type check that evaluates to $\trg{0}$ if the check suceeds and to $\trg{1}$ otherwise (\Cref{tr:e-n-has-nat,tr:e-n-has-nattimesnat}).
 The projection $\filterL{\trace}$ is entirely similar to $\filterLtms{\trace}$.
@@
   \begin{array}{rcl}
-    \cca(\src{\lbinop{\expr[_{1}]}{\expr[_{2}]}}) & = & \lbinop[\trg]{\left[\cca(\src{\expr[_{1}]})\right]}{\left[\cca(\src{\expr[_{2}]})\right]} \\ [0.33cm]
+%    \cca(\src{\lbinop{\expr[_{1}]}{\expr[_{2}]}}) & = & \lbinop[\trg]{\left[\cca(\src{\expr[_{1}]})\right]}{\left[\cca(\src{\expr[_{2}]})\right]} \\ [0.33cm]
     \cca(\src{\lfunction{g}{x:\natt\to\type[_{e}]}{\expr}}) & = & \lfunction[\trg]{\trg{g}}{\trg{x}}{\lifz[\trg]{\trg{\lhast{x}{\natt}}}{%

### 2023-06-22 popl/main.tex
Put langs into their own section. Same with comps

@@
 \SetMathAlphabet{\mathsf}{bold}{OT1}{LibertinusSans-LF}{bx}{n}
+
+\DeclareMathAlphabet{\mathtt}{OT1}{lmtt}{m}{n}
+\SetMathAlphabet{\mathtt}{bold}{OT1}{lmtt}{bx}{n}
 
@@
 
-\section{Case Study: Memory Safe and Cryptographic Constant Time Code}\label{sec:casestud}
-\begin{figure}[h]
-  \begin{center}
-    \begin{tikzpicture}
-      \node (S) {$\src{L_{\tmssafe}}$};
-      \node[right=1.0 of S] (T) {$\trg{L}$};
-      \node[right=1.0 of T] (M) {$\irl{L_{\mssafe}}$};
-      \node[below right=0.5 and 1.0 of M] (D0) {$\irl{L_{\mssafe}}$};
-      \node[above right=0.5 and 1.0 of M] (C0) {$\irl{L_{\mssafe}}$};
-      \node[right=1.0 of D0] (D1) {$\irl{L_{\mssafe}}$};
-      \node[right=1.0 of C0] (C1) {$\irl{L_{\mssafe}}$};
-      \node[right=4.5 of M] (O) {$\obj{L_{\scctsafe}}$};
-
-      \draw[->] (S) to[sloped,above] node (tmsedge) {\text{\scriptsize\gls{tms}}} (T);
-      \draw[->] (T) to[sloped,above] node {\text{\scriptsize\gls{sms}}} (M);
-      \draw[->] (M) to[sloped,above] node {\text{\scriptsize\gls{dce}}} (D0);
-      \draw[->] (M) to[sloped,above] node {\text{\scriptsize\gls{cf}}} (C0);
-      \draw[->] (D0) to[sloped,above] node {\text{\scriptsize\gls{cf}}} (D1);
-      \draw[->] (C0) to[sloped,above] node {\text{\scriptsize\gls{dce}}} (C1);
-      \draw[->] (D1) to[sloped,above] node {\text{\scriptsize\gls{scct}}} (O);
-      \draw[->] (C1) to[sloped,above] node {\text{\scriptsize\gls{scct}}} (O);
-
-      % Sections
-      \node[above=1.5 of tmsedge] (sectms) {{\scriptsize\Cref{subsec:cs:tms}}};
-      \node[right=0.0 of sectms] (secsms) {{\scriptsize\Cref{subsec:cs:ms}}};
-      \node[right=1.25 of secsms] (secopts) {{\scriptsize\Cref{subsec:cs:opts}}};
-      \node[right=1.25 of secopts] (secscct) {{\scriptsize\Cref{subsec:cs:scct}}};
-
-      \draw[dashed] ($(sectms)-(0.7,0)$) -- ($(sectms)-(0.7,1.75)$);
-      \draw[dashed] ($(sectms)-(-0.7,0)$) -- ($(sectms)-(-0.7,1.75)$);
-      \draw[dashed] ($(secsms)-(-0.7,0)$) -- ($(secsms)-(-0.7,1.75)$);
-      \draw[dashed] ($(secscct)-(0.7,0)$) -- ($(secscct)-(0.7,0.75)$);
-      \draw[dashed] ($(secscct)-(-0.7,0)$) -- ($(secscct)-(-0.7,0.75)$);
-    \end{tikzpicture}
-  \end{center}
-  \caption{Visualisation of the optimizing compilation pipeline that attains a combination of \gls{ms} and \gls{cct}.}\label{fig:pipeline}
-\end{figure}
-
-This section defines several secure compilers, each of which robustly preserves a different property of interest (\Cref{fig:pipeline}).
-The power of the framework (\Cref{sec:compprop,sec:compcomp}) is demonstrated by composing these compilers for a secure, but optimizing, compilation chain that robustly preserves \gls{ms} and \gls{cct}.
-Concretely, the section starts with a base-line definition for the used languages (\Cref{subsec:cs:defs}) and extends that to define languages $\src{L_{TMS}}$ and $\trg{L}$.
-For programmers using $\src{L_{TMS}}$, their mental model should be that their code is temporal memory safe and cryptographic constant time.
-The former is proven to be enforced by the static semantics of $\src{L_{TMS}}$, where each well-typed program is temporal memory safe.
-The latter is completely untyped.
-Having languages defined, this section continues with a compiler $\cc{\src{L_{\tmssafe}}}{\trg{L}}$ and proves that $\cc{\src{L_{\tmssafe}}}{\trg{L}}$ robustly preserves \gls{tms} (\Cref{subsec:cs:tms}).
-For readability purposes, it defines $\irl{L_{MS}}$ to be exactly the same as $\trg{L}$ and continues to present a compiler $\cc{\trg{L}}{\irl{L_{MS}}}$ that robustly preserves \gls{sms} (\Cref{subsec:cs:ms}).
-Now leaveraging the power of the framework that this paper introduces (\Cref{sec:compcomp,sec:compprop}), a compiler $\rtp{\cc{\src{L_{TMS}}}{\irl{L_{MS}}}}{\mssafe}$ is attained.
-The next steps in the compilation pipeline are two secure compilers $\cc[_{\gls{dce}}]{\irl{L_{MS}}}{\irl{L_{MS}}}$ and $\cc[_{\gls{cf}}]{\irl{L_{MS}}}{\irl{L_{MS}}}$ that perform \glsfirst{dce} and \glsfirst{cf}, respectively, to model an optimzing pipeline happening somewhere in the compilation chain (\Cref{subsec:cs:opts}).
-Finally, this section defines the language $\obj{L_{\scctsafe}}$ that is, in contrast to the other languages, not cryptographic constant time by definition.
-Concluding, a compiler $\cc{\irl{L_{MS}}}{\obj{L_{CCT}}}$ that robustly preserves \gls{cct} (\Cref{subsec:cs:scct}) is shown and composed with $\cc{\src{L_{TMS}}}{\irl{L_{MS}}}$ and the optimizing compilers $\cc[_{\gls{dce}}]{\irl{L_{MS}}}{\irl{L_{MS}}}$ and $\cc[_{\gls{cf}}]{\irl{L_{MS}}}{\irl{L_{MS}}}$.
-In turn, this yields $\cc{\src{L_{TMS}}}{\obj{L_{CCT}}}$, a compiler that robustly preserves a combination of \gls{ms} and \gls{cct}.
-
-\subsection{Preliminary Definitions}\label{subsec:cs:defs}
+%\section{Case Study: Memory Safe and Cryptographic Constant Time Code}\label{sec:casestud}
+\section{Case Study: Preliminary Definitions}\label{sec:casestud:defs}
+
+In order to define secure compilers one needs programming languages.
+To this end, this section defines the languages $\Ltms$, $\Ltrg$, $\Lms$, and $\Lscct$.
+In this paper, $\Ltms$ is the only statically typed language and exhibits the property that all well-typed programs are \gls{tms}.
+However, one can still write and evaluate $\Ltms$ programs that are well-typed, but perform out-of-bounds memory accesses, so the language is not \gls{sms}.
+In contrast, $\Ltrg$ is untyped and does not provide any guarantees with regards to \gls{ms}.
+$\Lms$ is exactly the same language as $\Ltrg$, but this paper still distinguishes the two for sake of readability.
+For all these languages, so $\Ltms$, $\Ltrg$, and $\Lms$, programmers assume \gls{cct} to hold, because the compiler is said to ensure it, similarily to what is being done in literature~\cite{cauligi2019fact}.
+This assumption does not hold anymore for the language $\Lscct$, which leaks information on branches and binary operations.
+But, contrary to the other languages, one can read and write to a model-specific register indicating a ``\gls{cct}-mode''.
+
+\subsection{Shared Language Definitions}\label{subsec:cs:defs}
 \begin{gather*}
   \begin{aligned}
-  &\mi{(Finals)}~\finalexpr\bnfdef\ \valueexpr \mid x\hspace{1.5cm}
-  \mi{(Values)}~\valueexpr\bnfdef\ n\\
+  &\mi{(Finals)}~\finalexpr\bnfdef\ \valueexpr \mid x \hspace{0.5cm}
+  \mi{(Values)}~\valueexpr\bnfdef\ n \hspace{0.5cm}
+  \mi{(Types)}~\type\bnfdef\ \natt \\
   \mi{(Expressions)}~\expr&\bnfdef\ \finalexpr \mid \lbinop{\expr[_{1}]}{\expr[_{2}]} \mid \lget{x}{\expr} \mid \lnew{x}{\expr[_{1}]}{\expr[_{2}]} \mid \ldelete{x} \mid \lset{x}{\expr[_{1}]}{\expr[_{2}]} \\
     & \mid \llet{x}{\expr[_{1}]}{\expr[_{2}]}  \mid \lreturn{\expr} \mid \lcall{g}{\expr} \mid \lifz{\expr}{\expr[_{1}]}{\expr[_{2}]} \mid \labort \\
-  \mi{(State)}~\statevar&\bnfdef \tup{\cfstate}{\sandboxtag,\memstate} \hspace{0.5cm}
-  \mi{(Runtime~Terms)}~\runtimetermvar\bnfdef\rtt{\statevar}{\expr} \mid \lightning \hspace{0.5cm}
+  \mi{(Symbols)}~\function&\bnfdef\ \lfunction{g}{x}{e}\hspace{0.5cm}
+  \mi{(Libraries)}~\library\bnfdef\ \hole{\cdot} \mid \function,\library \hspace{1.00cm}
     \text{where}\ \lbinop{}{}\in\left\{+,-,\times,<,/\right\}\\
-  \mi{(Symbols)}~\function&\bnfdef\ \lfunction{g}{x}{e}\hspace{0.5cm}
-  \mi{(Types)}~\type\bnfdef\ \natt\hspace{0.5cm}
-  \mi{(Locations)}~\loc\in\nat \\
-  \mi{(Eval.~Ctx.)}~\ectx&\bnfdef\ \hole{\cdot} \mid \lbinop{\ectx}{\expr} \mid \lbinop{\valueexpr}{\ectx} \mid \lget{x}{\ectx} \mid \lnew{x}{\ectx}{\expr} \mid \lset{x}{\ectx}{\expr}\\
-    & \mid \lset{x}{\valueexpr}{\ectx} \mid \lreturn{\ectx} \mid \lcall{g}{\ectx} \mid \lifz{\ectx}{\expr[_{1}]}{\expr[_{2}]} \\
-  \mi{(Poison)}~\poison&\bnfdef\ \poisoned \mid \poisonless \hspace{0.5cm}
-  \mi{(Store)}~\store\bnfdef\ \hole{\cdot} \mid \storeel{x}{\loc}{\sandboxtag}{\poison}{n},\store \\
-  \mi{(Transfer~Tag)}~\comm&\bnfdef\ \nocomm \mid\ \ctxtocomp \mid\ \comptoctx \hspace{0.5cm}
-  \mi{(Heaps)}~\heap\bnfdef\ \hole{\cdot} \mid n,\heap \hspace{0.5cm}
-  \mi{(Library)}~\library\bnfdef\ \hole{\cdot} \mid \function,\library \\
-  \mi{(Component~Names)}~\commlib&\bnfdef\ \hole{\cdot} \mid g,\commlib \hspace{0.5cm}
-  \mi{(Continuation~Stack)}~\ectxstack\bnfdef\ \hole{\cdot} \mid \tup{\ectx}{g},\ectxstack \\
-  \mi{(Flow~State)}~\cfstate&\bnfdef\ \tup{\commlib}{\library,\ectxstack} \hspace{0.5cm}
-  \mi{(Memory~State)}~\memstate\bnfdef\ \tup{\heap[^{\ctx}]}{\heap[^{\comp}],\store} \\
-  \mi{(Programs)}~&\prog{\library_{\ctx}}{\library_{\comp}}\hspace{0.5cm}
+  &\mi{(Programs)}~\prog{\library_{\ctx}}{\library_{\comp}}\hspace{0.5cm}
   \mi{(Substitutions)}~\substvar\bnfdef \hole{\cdot} \mid \subst{\valueexpr}{x},\gamma \\
-  \mi{(Pre\text{--}Events)}~\event[_{b}]&\bnfdef \ev{Alloc}\ \loc\ \valueexpr \mid \ev{Dealloc}\ \loc \mid \ev{Get}\ \loc\ \valueexpr \mid \ev{Set}\ \loc\ \valueexpr\ \valueexpr['] \mid \ev{Call}\ \comm\ g\ \valueexpr \mid \ev{Ret}\ \comm\ \valueexpr\\
-    & \mid \ev{Start} \mid \ev{End}\ \valueexpr\hspace{0.5cm}
-  \mi{(Events)}~\event\bnfdef \emptyevent \mid \lightning \mid \tup{\event[_{b}]}{\sandboxtag} \\
   \end{aligned}
@@
 Above is the base syntax of all the programming languages of this paper.
+\MK{this needs a complete rewrite}
 Expressions ($\expr$) can only be evaluated in the form of runtime--terms ($\rtt{\statevar}{\expr}$).
@@
 
+\begin{gather*}
+  \begin{aligned}
+  \mi{(Eval.~Ctx.)}~\ectx&\bnfdef\ \hole{\cdot} \mid \lbinop{\ectx}{\expr} \mid \lbinop{\valueexpr}{\ectx} \mid \lget{x}{\ectx} \mid \lnew{x}{\ectx}{\expr} \mid \lset{x}{\ectx}{\expr}\\
+    & \mid \lset{x}{\valueexpr}{\ectx} \mid \lreturn{\ectx} \mid \lcall{g}{\ectx} \mid \lifz{\ectx}{\expr[_{1}]}{\expr[_{2}]} \\
+  \end{aligned}\\
+  \begin{aligned}
+  \mi{(States)}~\statevar&\bnfdef \tup{\cfstate}{\sandboxtag,\memstate} &
+  \mi{(Poison)}~\poison&\bnfdef\ \poisoned \mid \poisonless &
+  \mi{(Locations)}~\loc\in\nat \\
+  \end{aligned}\\
+  \begin{aligned}
+  \mi{(Heaps)}~\heap&\bnfdef\ \hole{\cdot} \mid n,\heap &
+  \mi{(Stores)}~\store&\bnfdef\ \hole{\cdot} \mid \storeel{x}{\loc}{\sandboxtag}{\poison}{n},\store &
+  \mi{(Transfer~Tags)}~\comm\bnfdef\ \nocomm \mid\ \ctxtocomp \mid\ \comptoctx \\
+  \end{aligned}\\
+  \begin{aligned}
+  \mi{(Runtime~Terms)}~\runtimetermvar&\bnfdef\rtt{\statevar}{\expr} \mid \lightning &
+  \mi{(Component~Names)}~\commlib&\bnfdef\ \hole{\cdot} \mid g,\commlib \\
+  \end{aligned}\\
+  \begin{aligned}
+  \mi{(Continuation~Stacks)}~&\ectxstack\bnfdef\ \hole{\cdot} \mid \tup{\ectx}{g},\ectxstack \\
+  \end{aligned}
+\end{gather*}
 All programming languages presented in this section have an environmental semantics and all but $\src{L_{\tmssafe}}$ are dynamically typed.
@@
 
+\begin{gather*}
+  \begin{aligned}
+  \mi{(Pre\text{--}Events)}~\event[_{b}]\bnfdef \ev{Alloc}\ \loc\ \valueexpr \mid \ev{Dealloc}\ \loc \mid \ev{Get}\ \loc\ \valueexpr &\mid \ev{Set}\ \loc\ \valueexpr\ \valueexpr['] \mid \ev{Call}\ \comm\ g\ \valueexpr \mid \ev{Ret}\ \comm\ \valueexpr \mid \ev{Start} \mid \ev{End}\ \valueexpr \\
+  \mi{(Events)}~\event&\bnfdef \emptyevent \mid \lightning \mid \tup{\event[_{b}]}{\sandboxtag} \\
+  \end{aligned}
+\end{gather*}
+
+\MKin{now is time to talk about events? or maybe right after presenting the rules}
+
 \begin{center}
@@
     \mi{(Pre\text{--}Types)}~\src{\type[_{e}]}\bnfdef&\ \src{\natt} \mid \src{\ptrn} &
-    \mi{(Type~Environment)}~\src{\typenv}\bnfdef&\ \src{\hole{\cdot}} \mid \hastype{\src{x}}{\src{\type}},\src{\typenv} \\
+    \mi{(Type~Environments)}~\src{\typenv}\bnfdef&\ \src{\hole{\cdot}} \mid \hastype{\src{x}}{\src{\type}},\src{\typenv} \\
     \mi{(Types)}~\src{\type}\bnfdef&\ \src{\type[_{e}]} \mid \src{\type[_{e}]\to\bot} \mid \src{\type[^{(1)}_{e}]\to\type[^{(2)}_{e}]} &
@@
 
-\MKin{show at least one forward sim lemma or not?}
-
 \begin{lemma}[$\src{L_{\tmssafe}}$-programs are \gls{tms}]\label{lem:wt:tmsmon}
@@
 
-\subsection{$\trg{L}$: An Unsafe Language}\label{subsec:lsms}
+\subsection{$\Ltrg$: A Memory-Unsafe Language}\label{subsec:lsms}
 
@@
 
-The $\trg{L}$ language does not provide any guarantees with regards to memory safety.
+The $\Ltrg$ language does not provide any guarantees with regards to memory safety.
 The syntax of it extends the unified syntax for all languages (\Cref{subsec:cs:defs}) with pairs of natural numbers (\Cref{tr:e-proj1}) and two different kinds of dynamic checks.
@@
 The projection $\filterL{\trace}$ is entirely similar to $\filterLtms{\trace}$.
+\MKin{add note about \gls{sms} projection}
+
+\subsection{$\Lms$: Another Memory-Unsafe Language}\label{subsec:lms}
+
+This paper chooses to introduce $\Lms$, despite it being exactly equal to $\Ltrg$.
+The main purpose of this decision is to enhance readability.
+
+\MKin{add note about \gls{sms} projection}
+
+\subsection{$\Lscct$: An Unsafe Language}\label{subsec:lscct}
+
+\begin{gather*}
+  \begin{aligned}
+    \mi{(Expressions)}~\obj{\expr}\bnfdef&\dots \mid \obj{\lwrdoit{\expr}} \mid \obj{\lrddoit{x}{\expr}} &
+    \mi{(States)}~\obj{\statevar}\bnfdef&\obj{\tup{\cfstate}{\sandboxtag,n,\memstate}}
+  \end{aligned}
+\end{gather*}
+
+$\Lscct$ extends the state presented earlier (\Cref{subsec:cs:defs}) with another entry that models the value of a register for a data (operand) independent timing mode.
+To modify the register, $\Lscct$ extends expressions with a way to write a value to the register ($\obj{\lwrdoit{\expr}}$) and a way to read it ($\obj{\lrddoit{x}{\expr}}$).
+Finals carry a secrecy annotation, indicating high ($\obj{\lock}$) or low ($\obj{\unlock}$) secrecy.
+Whenever the register is marked as active, parts of the dynamics of $\Lscct$ do not leak information.
+In general, $\Lscct$ can leak more information than the other languages:
+
+\begin{gather*}
+  \begin{aligned}
+    \mi{(Pre\text{--}Events)}~\obj{\event[_{b}]}&\bnfdef\ \dots \mid \obj{{Branch}\ n} \mid \obj{{Binop}\ n} \mid \obj{\widehat{{Get}}\ \loc\ \valueexpr} \mid \obj{\widehat{{Set}}\ \loc\ \valueexpr\ \valueexpr[']} &
+    \mi{(Events)}~\obj{\event}&\bnfdef\ \obj{\varepsilon} \mid \obj{\lightning} \mid \obj{\tup{\event[_{b}]}{\sandboxtag,\securitytag}}
+  \end{aligned}
+\end{gather*}
+
+Events carry a secrecy tag instead of the data they carry.
+Branches and binary operations leak information about the data they compute on.
+If the data (object) independent timing register is set to be active, the read and write to memory events should not leak.
+However, to emit $\obj{\emptyevent}$ in this case would mean that memory unsafe operations are hidden as well.
+Since this language should not have a mechanism to hide memory unsafety, pre--events are extended to model reading from ($\obj{\widehat{{Get}}\ \loc\ \valueexpr}$) and writing to ($\obj{\widehat{Set}\ \loc\ \valueexpr}$) memory without leaking secret information involved in the access.
+
+\begin{center}
+  \typerule{e-$\obj{\lbinop{}{}}$-leak}{
+    \obj{n_1}\oplus\obj{n_2}=\obj{n_3} &
+    \obj{\securitytag}\sqcap\obj{\securitytag[']}=\obj{\securitytag['']} &
+    \obj{\oplus}\notin\left\{\obj{\times}, \obj{/}, \obj{<}\right\}
+  }{
+    \pstepto{\rtt{\obj{\tup{\cfstate}{\sandboxtag,0,\memstate}}}{\obj{\lbinop{n_1^{\securitytag}}{n_2^{\securitytag[']}}}}}{\rtt{\obj{\tup{\cfstate}{\sandboxtag,0,\memstate}}}{\obj{n_3^{\securitytag['']}}}}{\obj{\tup{Binop\ n_3}{\sandboxtag,\securitytag['']}}}
+  }{e-binop-leak}
+\end{center}
+
+\MKin{filters to spec events}
+
+\section{Case Study: Secure Compiler Compositions in the Wild}\label{sec:casestud:rtp}
+
+\begin{figure}[h]
+  \begin{center}
+    \begin{tikzpicture}
+      \node (S) {$\src{L_{\tmssafe}}$};
+      \node[right=1.0 of S] (T) {$\trg{L}$};
+      \node[right=1.0 of T] (M) {$\irl{L_{\mssafe}}$};
+      \node[below right=0.5 and 1.0 of M] (D0) {$\irl{L_{\mssafe}}$};
+      \node[above right=0.5 and 1.0 of M] (C0) {$\irl{L_{\mssafe}}$};
+      \node[right=1.0 of D0] (D1) {$\irl{L_{\mssafe}}$};
+      \node[right=1.0 of C0] (C1) {$\irl{L_{\mssafe}}$};
+      \node[right=4.5 of M] (O) {$\obj{L_{\scctsafe}}$};
+
+      \draw[->] (S) to[sloped,above] node (tmsedge) {\text{\scriptsize\gls{tms}}} (T);
+      \draw[->] (T) to[sloped,above] node {\text{\scriptsize\gls{sms}}} (M);
+      \draw[->] (M) to[sloped,above] node {\text{\scriptsize\gls{dce}}} (D0);
+      \draw[->] (M) to[sloped,above] node {\text{\scriptsize\gls{cf}}} (C0);
+      \draw[->] (D0) to[sloped,above] node {\text{\scriptsize\gls{cf}}} (D1);
+      \draw[->] (C0) to[sloped,above] node {\text{\scriptsize\gls{dce}}} (C1);
+      \draw[->] (D1) to[sloped,above] node {\text{\scriptsize\gls{scct}}} (O);
+      \draw[->] (C1) to[sloped,above] node {\text{\scriptsize\gls{scct}}} (O);
+
+      % Sections
+      \node[above=1.5 of tmsedge] (sectms) {{\scriptsize\Cref{subsec:cs:tms}}};
+      \node[right=0.0 of sectms] (secsms) {{\scriptsize\Cref{subsec:cs:ms}}};
+      \node[right=1.25 of secsms] (secopts) {{\scriptsize\Cref{subsec:cs:opts}}};
+      \node[right=1.25 of secopts] (secscct) {{\scriptsize\Cref{subsec:cs:scct}}};
+
+      \draw[dashed] ($(sectms)-(0.7,0)$) -- ($(sectms)-(0.7,1.75)$);
+      \draw[dashed] ($(sectms)-(-0.7,0)$) -- ($(sectms)-(-0.7,1.75)$);
+      \draw[dashed] ($(secsms)-(-0.7,0)$) -- ($(secsms)-(-0.7,1.75)$);
+      \draw[dashed] ($(secscct)-(0.7,0)$) -- ($(secscct)-(0.7,0.75)$);
+      \draw[dashed] ($(secscct)-(-0.7,0)$) -- ($(secscct)-(-0.7,1.25)$);
+    \end{tikzpicture}
+  \end{center}
+  \caption{Visualisation of the optimizing compilation pipeline that attains a combination of \gls{ms} and \gls{cct}.}\label{fig:pipeline}
+\end{figure}
+This section defines several secure compilers, each of which robustly preserves a different property of interest (\Cref{fig:pipeline}).
+The power of the framework (\Cref{sec:compprop,sec:compcomp}) is demonstrated by composing these compilers for a secure, but optimizing, compilation chain that robustly preserves \gls{ms} and \gls{cct}.
+Concretely, the section starts with a base-line definition for the used languages (\Cref{subsec:cs:defs}) and extends that to define languages $\src{L_{TMS}}$ and $\trg{L}$.
+For programmers using $\src{L_{TMS}}$, their mental model should be that their code is temporal memory safe and cryptographic constant time.
+The former is proven to be enforced by the static semantics of $\src{L_{TMS}}$, where each well-typed program is temporal memory safe.
+The latter is completely untyped.
+Having languages defined, this section continues with a compiler $\cc{\src{L_{\tmssafe}}}{\trg{L}}$ and proves that $\cc{\src{L_{\tmssafe}}}{\trg{L}}$ robustly preserves \gls{tms} (\Cref{subsec:cs:tms}).
+For readability purposes, it defines $\irl{L_{MS}}$ to be exactly the same as $\trg{L}$ and continues to present a compiler $\cc{\trg{L}}{\irl{L_{MS}}}$ that robustly preserves \gls{sms} (\Cref{subsec:cs:ms}).
+Now leaveraging the power of the framework that this paper introduces (\Cref{sec:compcomp,sec:compprop}), a compiler $\rtp{\cc{\src{L_{TMS}}}{\irl{L_{MS}}}}{\mssafe}$ is attained.
+The next steps in the compilation pipeline are two secure compilers $\cc[_{\gls{dce}}]{\irl{L_{MS}}}{\irl{L_{MS}}}$ and $\cc[_{\gls{cf}}]{\irl{L_{MS}}}{\irl{L_{MS}}}$ that perform \glsfirst{dce} and \glsfirst{cf}, respectively, to model an optimzing pipeline happening somewhere in the compilation chain (\Cref{subsec:cs:opts}).
+Finally, this section defines the language $\obj{L_{\scctsafe}}$ that is, in contrast to the other languages, not cryptographic constant time by definition.
+Concluding, a compiler $\cc{\irl{L_{MS}}}{\obj{L_{CCT}}}$ that robustly preserves \gls{cct} (\Cref{subsec:cs:scct}) is shown and composed with $\cc{\src{L_{TMS}}}{\irl{L_{MS}}}$ and the optimizing compilers $\cc[_{\gls{dce}}]{\irl{L_{MS}}}{\irl{L_{MS}}}$ and $\cc[_{\gls{cf}}]{\irl{L_{MS}}}{\irl{L_{MS}}}$.
+In turn, this yields $\cc{\src{L_{TMS}}}{\obj{L_{CCT}}}$, a compiler that robustly preserves a combination of \gls{ms} and \gls{cct}.
 
@@
 
-\MKin{x-lang relations $\approx_{\delta;\src{L}}$, $\multimap_{\delta;\src{L}}$, as well as $\cong_{\delta;\src{X}}$}
-\MKin{backtranslation}
+\MKin{sketch proof with backtranslation and the x-lang rels}
 
@@
 \subsection{Optimizing Compilers}\label{subsec:cs:opts}
+\MKin{rework this}
 
@@
 \MKin{composition: rtp of ms, order independent}
-
-\subsection{$\obj{L_{\scctsafe}}$: A Language with Access to a Model Specific Register}\label{subsec:lscct}
-
-\MKin{add definitions and description; some syntax + dynamics}
 

### 2023-06-22 popl/main.tex
Patch figure

@@
 
+\subsubsection{All Well-Typed $\Ltms$ Programs are Temporal Memory Safe}
+
 \begin{center}
   \judgbox{\thelocmap(\src{\loc})=\loc}{,,Map an $\src{L_{\tmssafe}}$ location to a specification location $\loc$.''}
-  \judgbox{\filterLtms{\src{\event}}=\event}{,,Project an $\src{L_{\tmssafe}}$ event $\src{\event}$ to a specification event $\event$.''}
+  \judgbox{\msfilterLtms{\src{\event}}=\event}{,,Project an $\src{L_{\tmssafe}}$ event $\src{\event}$ to a specification event $\event$.''}
   \typerule{$\src{L_{\tmssafe}}$-filter-context}{
@@
   }{
-    \filterLtms{\src{\tup{\event[_b]}{\ctx}} = \emptyevent}
+    \msfilterLtms{\src{\tup{\event[_b]}{\ctx}} = \emptyevent}
   }{ltms-filter-context}
@@
   }{
-    \filterLtms{\src{\lightning}} = \lightning
+    \msfilterLtms{\src{\lightning}} = \lightning
   }{ltms-filter-abort}
@@
 %  }{
-%    \filterLtms{\src{\tup{Start}{\comp}}} = \emptyevent
+%    \msfilterLtms{\src{\tup{Start}{\comp}}} = \emptyevent
 %  }{ltms-filter-start}
@@
   }{
-    \filterLtms{\src{\tup{Alloc\ \loc\ n}{\comp}}} = \tup{\ev{Alloc}\ \loc\ n}{\comp,\unlock}
+    \msfilterLtms{\src{\tup{Alloc\ \loc\ n}{\comp}}} = \tup{\ev{Alloc}\ \loc\ n}{\comp,\unlock}
   }{ltms-filter-alloc}
@@
 \begin{lemma}[$\src{L_{\tmssafe}}$-programs are \gls{tms}]\label{lem:wt:tmsmon}
-  If $\progstepto{\src{\prog{\library_{\ctx}}{\library_{\comp}}}}{\src{\rtt{\statevar}{\valueexpr}}}{\src{\trace}}$, then $\montmssafe{\filterLtms{\src{\trace}}}$.\Coqed
+  If $\progstepto{\src{\prog{\library_{\ctx}}{\library_{\comp}}}}{\src{\rtt{\statevar}{\valueexpr}}}{\src{\trace}}$, then $\montmssafe{\msfilterLtms{\src{\trace}}}$.\Coqed
 \end{lemma}
@@
 The other check is a simple dynamic type check that evaluates to $\trg{0}$ if the check suceeds and to $\trg{1}$ otherwise (\Cref{tr:e-n-has-nat,tr:e-n-has-nattimesnat}).
-The projection $\filterL{\trace}$ is entirely similar to $\filterLtms{\trace}$.
-\MKin{add note about \gls{sms} projection}
+The projection $\msfilterL{\trace}$ is entirely similar to $\msfilterLtms{\trace}$.
 
@@
 The main purpose of this decision is to enhance readability.
-
-\MKin{add note about \gls{sms} projection}
+One thing that is added to $\Lms$ is another projection to the most general trace model for reasoning about \gls{scct}.
+Here, any information irrelevant to \gls{scct} is filtered.
+
+\begin{center}
+  \typerule{$\Lms$-filter-alloc}{
+  }{
+    \scctfilterLms{\irl{\tup{Get\ \loc\ n}{\comp}}} = \emptyevent
+  }{lms-filter-get}
+  %
+  \typerule{$\Lms$-filter-abort}{
+  }{
+    \scctfilterLms{\irl{\lightning}} = \lightning
+  }{lms-filter-abort}
+\end{center}
 
@@
       \node (S) {$\src{L_{\tmssafe}}$};
-      \node[right=1.0 of S] (T) {$\trg{L}$};
-      \node[right=1.0 of T] (M) {$\irl{L_{\mssafe}}$};
+      \node[right=1.5 of S] (T) {$\trg{L}$};
+      \node[right=1.5 of T] (M) {$\irl{L_{\mssafe}}$};
       \node[below right=0.5 and 1.0 of M] (D0) {$\irl{L_{\mssafe}}$};
       \node[above right=0.5 and 1.0 of M] (C0) {$\irl{L_{\mssafe}}$};
-      \node[right=1.0 of D0] (D1) {$\irl{L_{\mssafe}}$};
-      \node[right=1.0 of C0] (C1) {$\irl{L_{\mssafe}}$};
-      \node[right=4.5 of M] (O) {$\obj{L_{\scctsafe}}$};
-
-      \draw[->] (S) to[sloped,above] node (tmsedge) {\text{\scriptsize\gls{tms}}} (T);
-      \draw[->] (T) to[sloped,above] node {\text{\scriptsize\gls{sms}}} (M);
-      \draw[->] (M) to[sloped,above] node {\text{\scriptsize\gls{dce}}} (D0);
-      \draw[->] (M) to[sloped,above] node {\text{\scriptsize\gls{cf}}} (C0);
-      \draw[->] (D0) to[sloped,above] node {\text{\scriptsize\gls{cf}}} (D1);
-      \draw[->] (C0) to[sloped,above] node {\text{\scriptsize\gls{dce}}} (C1);
-      \draw[->] (D1) to[sloped,above] node {\text{\scriptsize\gls{scct}}} (O);
-      \draw[->] (C1) to[sloped,above] node {\text{\scriptsize\gls{scct}}} (O);
+      \node[right=3.0 of M] (E) {$\irl{L_{\mssafe}}$};
+      \node[right=1.5 of E] (O) {$\obj{L_{\scctsafe}}$};
+
+      \draw[->] (S) to[sloped] node[align=center,font=\scriptsize] (tmsedge) {\gls{tms}\\ \Cref{thm:cca:rtp:tms}} (T);
+      \draw[->] (T) to[sloped] node[align=center,font=\scriptsize] {\gls{sms}\\ \Cref{thm:ccb:rtp:sms}} (M);
+      \draw[->] (M) to[sloped] node[align=center,font=\scriptsize] {\gls{dce}\\ \Cref{thm:ccdce:rtp:ms}} (D0);
+      \draw[->] (M) to[sloped] node[align=center,font=\scriptsize] {\gls{cf}\\ \Cref{thm:cccf:rtp:ms}} (C0);
+      \draw[->] (D0) to[sloped] node[align=center,font=\scriptsize] {\gls{cf}\\ \Cref{thm:cccf:rtp:ms}} (E);
+      \draw[->] (C0) to[sloped] node[align=center,font=\scriptsize] {\gls{dce}\\ \Cref{thm:ccdce:rtp:ms}} (E);
+      \draw[->] (E) to[sloped] node[align=center,font=\scriptsize] {\gls{scct}\\ \Cref{thm:ccscct:rtp:scct}} (O);
 
       % Sections
-      \node[above=1.5 of tmsedge] (sectms) {{\scriptsize\Cref{subsec:cs:tms}}};
-      \node[right=0.0 of sectms] (secsms) {{\scriptsize\Cref{subsec:cs:ms}}};
-      \node[right=1.25 of secsms] (secopts) {{\scriptsize\Cref{subsec:cs:opts}}};
-      \node[right=1.25 of secopts] (secscct) {{\scriptsize\Cref{subsec:cs:scct}}};
-
-      \draw[dashed] ($(sectms)-(0.7,0)$) -- ($(sectms)-(0.7,1.75)$);
-      \draw[dashed] ($(sectms)-(-0.7,0)$) -- ($(sectms)-(-0.7,1.75)$);
-      \draw[dashed] ($(secsms)-(-0.7,0)$) -- ($(secsms)-(-0.7,1.75)$);
-      \draw[dashed] ($(secscct)-(0.7,0)$) -- ($(secscct)-(0.7,0.75)$);
-      \draw[dashed] ($(secscct)-(-0.7,0)$) -- ($(secscct)-(-0.7,1.25)$);
+      \node[above=1.0 of tmsedge] (sectms) {{\scriptsize\Cref{subsec:cs:tms}}};
+      \node[right=0.5 of sectms] (secsms) {{\scriptsize\Cref{subsec:cs:ms}}};
+      \node[right=1.75 of secsms] (secopts) {{\scriptsize\Cref{subsec:cs:opts}}};
+      \node[right=1.75 of secopts] (secscct) {{\scriptsize\Cref{subsec:cs:scct}}};
+
+      \draw[dashed] ($(sectms)-(0.8,0)$) -- ($(sectms)-(0.8,1.25)$);
+      \draw[dashed] ($(sectms)-(-0.9,0)$) -- ($(sectms)-(-0.9,1.25)$);
+      \draw[dashed] ($(secsms)-(-1.2,0)$) -- ($(secsms)-(-1.2,1.25)$);
+      \draw[dashed] ($(secscct)-(1.2,0)$) -- ($(secscct)-(1.2,1.25)$);
+      \draw[dashed] ($(secscct)-(-1.1,0)$) -- ($(secscct)-(-1.1,1.25)$);
     \end{tikzpicture}
@@
 
-\MKin{backward sim}
 \begin{theorem}[Compiler $\ccb$ is secure with respect to \gls{sms}]\label{thm:ccb:rtp:sms}
@@
 \subsection{Optimizing Compilers}\label{subsec:cs:opts}
-\MKin{rework this}
-
-\MKin{compiler $\ccdce$ doing \gls{dce}}
-\MKin{compiler $\cccf$ doing \gls{cf}}
-
-\MKin{backward sims}
-\MKin{rtp of ms}
-\MKin{composition: rtp of ms, order independent}
+
+\begin{center}
+  $$
+  \begin{array}{rcl}
+    \ccdce(\irl{\lifz{true}{\expr[_{1}]}{\expr[_{2}]}}) & = & \ccdce(\irl{\expr[_{1}]}) \\
+    \ccdce(\irl{\lifz{false}{\expr[_{1}]}{\expr[_{2}]}}) & = & \ccdce(\irl{\expr[_{2}]}) \\
+  \end{array}
+  $$
+\end{center}
+
+\begin{center}
+  $$
+  \begin{array}{rcll}
+    \cccf(\irl{\llet{x}{n}{\expr}}) & = & \cccf(\irl{\expr\subst{n}{x}}) &\\
+    \cccf(\irl{\lbinop{n_{1}}{n_{2}}}) & = & \irl{n_{3}} & \text{where }\irl{n_{3}}\text{ is }\lbinop{\irl{n_{1}}}{\irl{n_{2}}}\\
+  \end{array}
+  $$
+\end{center}
+
+\begin{theorem}[Compiler $\ccdce$ is secure with respect to \gls{ms}]\label{thm:ccdce:rtp:ms}
+  $\rtp{\ccdce}{\mssafe}$ % \Coqed
+\end{theorem}
+\begin{theorem}[Compiler $\cccf$ is secure with respect to \gls{ms}]\label{thm:cccf:rtp:ms}
+  $\rtp{\cccf}{\mssafe}$ % \Coqed
+\end{theorem}
+
+\begin{theorem}[Compilers $\cccf\circ\ccdce$ and $\cccf\circ\ccdce$ are secure with respect to \gls{ms}]\label{thm:cccfccdce:rtp:ms}
+  $\rtp{\cccf\circ\ccdce}{\mssafe}$ and $\rtp{\ccdce\circ\cccf}{\mssafe}$ % \Coqed
+\end{theorem}
 
@@
 
-\MKin{compiler $\irl{L_{MS}}\to\obj{L_{\scctsafe}}$}
-\MKin{x-lang relations}
-\MKin{backtranslation}
-\MKin{rtp of scct}
-
-\MKin{composition: rtp of everything}
+\begin{center}
+  $$
+  \begin{array}{rcl}
+    \ccscct(\irl{\lfunction{g}{x}{\expr}}) & = & \lfunction[\obj]{\obj{g}}{\obj{x}}{\obj{\lwrdoit{1};}\ccscct(\irl{\expr})}
+  \end{array}
+  $$
+\end{center}
+
+\begin{theorem}[Compiler $\ccscct$ is secure with respect to \gls{scct}]\label{thm:ccscct:rtp:scct}
+  $\rtp{\ccscct}{\scctsafe}$ % \Coqed
+\end{theorem}
+
+\subsection{Robust Preservation of Intersection of Memory Safety and Strict Cryptographic Constant Time}
+
+Let $\cc{\Ltms}{\Lscct}$ be the compiler that is the composition of $\cca$, $\ccb$, $\cccf$, $\ccdce$, and $\ccscct$, then the following theorem can be shown.
+
+\begin{theorem}[Compiler $\ccscct$ is secure with respect to \gls{scct}]\label{thm:ccscct:rtp:scct}
+  $\rtp{\ccscct}{\scctsafe}$ % \Coqed
+\end{theorem}
 

### 2023-06-22 popl/main.tex
Patch figure

@@
       \draw[->] (C0) to[sloped] node[align=center,font=\scriptsize] {\gls{dce}\\ \Cref{thm:ccdce:rtp:ms}} (E);
-      \draw[->] (E) to[sloped] node[align=center,font=\scriptsize] {\gls{scct}\\ \Cref{thm:ccscct:rtp:scct}} (O);
+      \draw[->] (E) to[sloped,above] node[align=center,font=\scriptsize] {\gls{scct}\\ \Cref{thm:ccscct:rtp:scct}} (O);
 
@@
       \node[right=1.75 of secopts] (secscct) {{\scriptsize\Cref{subsec:cs:scct}}};
+
+      \draw[thick,dotted,->] (S) to[bend right=33,sloped] node[align=center,font=\scriptsize] {\gls{ms}\\ \Cref{thm:ccab:rtp:ms}} (M);
+      \draw[thick,dotted,->] (M) to[bend right=0,sloped] node[align=center,font=\scriptsize] {\gls{ms}\\ \Cref{thm:cccfccdce:rtp:ms}} (E);
+      \draw[thick,dotted,->] (S) to[out=-45,in=198,sloped] node[align=center,font=\scriptsize] {\gls{ms}\\ \Cref{thm:ccall:rtp:msscct}} (O);
 
@@
 
+\begin{theorem}[Compiler $\ccscct$ is secure with respect to \gls{scct}]\label{thm:ccall:rtp:msscct}
+  $\rtp{\cc{\Ltms}{\Lscct}}{\mssafe\cap\scctsafe}$ % \Coqed
+\end{theorem}
+
 %\begin{lstlisting}[language=c,caption=``Wrong bounds check of two 64-bit integers.'']

### 2023-06-22 popl/main.tex
Add figure

@@
   \mi{(Security\ Tag)}~\sigma\bnfdef&\ \lock \mid \unlock\hspace{0.5cm}
-  \mi{(Control\ Tag)}~\sandboxtag\bnfdef \ctx\mid\comp\\
-  \mi{(Pre\text{--}event)}~\event[_{b}]\bnfdef&\ \ev{Alloc\ \loc\ n} \mid \ev{Dealloc\ \loc} \mid \ev{Use\ \loc\ n} \mid \ev{Branch\ n} \mid \ev{Binop\ n} \\
-  \mi{(Event)}~\event\bnfdef&\ \emptyevent \mid \lightning \mid \event[_{b}];\sandboxtag;\sigma \\
+  \mi{(Control\ Tag)}~\sandboxtag\bnfdef \ctx\mid\comp\hspace{0.5cm}
+  \mi{(Event)}~\event\bnfdef\ \emptyevent \mid \lightning \mid \event[_{b}];\sandboxtag;\sigma \\
+  \mi{(Pre\text{--}event)}&~\event[_{b}]\bnfdef\ \ev{Alloc\ \loc\ n} \mid \ev{Dealloc\ \loc} \mid \ev{Use\ \loc\ n} \mid \ev{Branch\ n} \mid \ev{Binop\ n} \\
   \end{aligned}
@@
 
-\MKin{sketch proof with backtranslation and the x-lang rels}
-
 \begin{theorem}[Compiler $\cca$ is secure with respect to \gls{tms}]\label{thm:cca:rtp:tms}
@@
 \end{theorem}
+
+\begin{figure}[h]
+  \begin{center}
+    \begin{tikzpicture}[state/.style={minimum height=0.6cm}]
+      % relative horizontal/vertical distance between states
+      \pgfmathsetmacro{\hdist}{0.95}
+      \pgfmathsetmacro{\vdist}{1.35}
+      \pgfmathsetmacro{\halfvdist}{0.725}
+
+      % row of src states
+      \node[state] (srcempty) {$\src{\emptyset}$};
+      \foreach \s [remember=\s as \cur (initially empty)] in {1,w_1,p,w_2,2} {
+        \node[state,right=\hdist of src\cur] (src\s) {$\src{\Omega_{\s}}$};
+      }
+      % row of trg states
+      \node[state,below=\vdist of srcempty] (trgempty) {$\trg{\emptyset}$};
+      \foreach \s in {1,w_1,p,w_2,2} {
+        \node[state,below=\vdist of src\s] (trg\s) {$\trg{\Omega_{\s}}$};
+      }
+      %% illustrations
+        % backtrans wrapper 1
+        \draw[thick,loosely dotted,Peach!50,rounded corners] (src1.north east) -- (srcw\string_1.north east)
+          -- (trgw\string_1.south east) -- (trg1.south west) -- (src1.north west) -- cycle;
+        % backtrans wrapper 2
+        \draw[thick,loosely dotted,Peach!50,rounded corners] (srcp.north east) -- (srcw\string_2.north east)
+          -- (trgw\string_2.south east) -- (trgp.south west) -- (srcp.north west) -- cycle;
+        % compiler correctness
+        \draw[thick,loosely dashed,Emerald!50,rounded corners] ($(srcw\string_1.north east)+(0,0.05)$) -- ($(srcp.north east)+(0,0.05)$)
+          -- ($(trgw\string_2.north east)+(0.05,0)$) -- ($(trgw\string_2.south east)+(0.05,-0.05)$)
+          -- ($(trg1.south west)+(-0.05,-0.05)$) -- ($(trg1.north west)+(-0.05,0)$) -- ($(srcw\string_1.north west)+(0,0.05)$) -- cycle;
+      % state relations
+      \path (srcempty) edge[draw=gray!25] node[pos=0.5,sloped,rotate=180,fill=white] {\scriptsize$\multimap_\emptyset$} (trgempty)
+        (src1) edge[draw=gray!25] node[pos=0.5,sloped,rotate=180,fill=white] {\scriptsize$\multimap_{\delta_1}$} (trg1)
+        (srcp) edge[draw=gray!25] node[pos=0.5,sloped,fill=white] {\scriptsize$\approx_{\delta_p}$} (trgp)
+        (src2) edge[draw=gray!25] node[pos=0.5,sloped,rotate=180,fill=white] {\scriptsize$\multimap_{\delta_2}$} (trg2)
+        (srcw\string_1) edge[draw=gray!25] node[pos=0.5,sloped,fill=white] {\scriptsize$\approx_{\delta_{w_1}}$} (trgw\string_1)
+        (srcw\string_2) edge[draw=gray!25] node[pos=0.5,sloped,rotate=180,fill=white] {\scriptsize$\multimap_{\delta_{w_2}}$} (trgw\string_2)
+        % diagonals
+        (srcw\string_1) edge[draw=gray!25] node[pos=0.16,sloped,rotate=180,fill=white] {\scriptsize$\approx_{\delta_{w_1}}$} (trg1)
+        (src1) edge[draw=gray!25] node[pos=0.16,sloped,rotate=180,fill=white] {\scriptsize$\multimap_{\delta_1}$} (trgw\string_1)
+        (srcp) edge[draw=gray!25] node[pos=0.2,sloped,fill=white] {\scriptsize$\approx_{\delta_p}$} (trgw\string_2)
+        (trgp) edge[draw=gray!25] node[pos=0.2,sloped,fill=white] {\scriptsize$\multimap_{\delta_{w_2}}$} (srcw\string_2)
+        ;
+      %\drawpolygon src1,srcw\string_1,trgw\string_1,trg1;
+      %\drawpolygon srcp,srcw\string_2,trgw\string_2,trgp;
+      %\node[font=\tiny,align=center,above=0.2 of srcw1srcp] (wrapper) {Backtranslation\\Wrapper};
+      %\path[->,draw] (wrapper) -- (srcw\string_1);
+      %\path[->,draw] (wrapper) -- (srcp);
+      % steps
+      \path[color=\stlccol] (srcempty) edge[draw=none] node {\ $\xrightarrow{\trace[_1]}{}{\kern-3.5pt}_{\text{ctx}}^*$} (src1)
+        (src1) edge[draw=none] node {\ $\xrightarrow{\trace[_c]}{}{\kern-3.5pt}_{\text{ctx}}$} (srcw\string_1)
+        (srcw\string_1) edge[draw=none] node {\ $\xrightarrow{\trace[_p]}{}{\kern-3.5pt}_{\text{ctx}}^*$} (srcp)
+        (srcp) edge[draw=none] node[fill=white,inner sep=0,outer sep=0] {\ $\xrightarrow{\trace[_r]}{}{\kern-3.5pt}_{\text{ctx}}$} (srcw\string_2)
+        (srcw\string_2) edge[draw=none] node {\ $\xrightarrow{\trace[_2]}{}{\kern-3.5pt}_{\text{ctx}}^*$} (src2)
+        ;
+      \path[color=\ulccol] (trgempty) edge[draw=none] node[fill=white,inner sep=0,outer sep=0] {\ $\xrightarrow{\phantom{\trace[_1]}}{}{\kern-3.5pt}_{\text{ctx}}^*$} (trg1)
+        (trg1) edge[draw=none] node {\ $\xrightarrow{\phantom{\trace[_p]}}{}{\kern-3.5pt}_{\text{ctx}}^*$} (trgw\string_1)
+        (trgw\string_1) edge[draw=none] node[fill=white,inner sep=0,outer sep=0] {\ $\xrightarrow{\phantom{\trace[_p]}}{}{\kern-3.5pt}_{\text{ctx}}^*$} (trgp)
+        (trgp) edge[draw=none] node {\ $\xrightarrow{\phantom{\trace[_p]}}{}{\kern-3.5pt}_{\text{ctx}}^*$} (trgw\string_2)
+        (trgw\string_2) edge[draw=none] node {\ $\xrightarrow{\phantom{\trace[_2]}}{}{\kern-3.5pt}_{\text{ctx}}^*$} (trg2)
+        ;
+      % legend
+      \node[align=left,below right=0.3 and 0.3 of trgempty,font=\tiny] (legend) {%
+        ${\src{\trace[_c]}}\cong{\backt{\Ltms}{\Ltrg}(\trg{Call\ ?\finalexpr_i})}$\\%
+        ${\src{\trace[_r]}}\cong{\backt{\Ltms}{\Ltrg}(\trg{Ret\ !\finalexpr_o})}$
+        };
+      \draw[thick,loosely dotted,Peach!50,rounded corners] ($(legend.north east)+(0.5,-0.4)$) -- ($(legend.north east)+(1,-0.4)$);
+      \node at ($(legend.north east)+(1.5,-0.4)$) (legendwrapper) {};
+      \node[right of=legendwrapper] {\tiny Backtranslation Wrapper};
+      \draw[thick,loosely dashed,Emerald!50,rounded corners] ($(legend.south east)+(0.5,0.4)$) -- ($(legend.south east)+(1,0.4)$);
+      \node at ($(legend.south east)+(1.5,0.4)$) (legendcorrectness) {};
+      \node[right of=legendcorrectness] {\tiny Compiler Correctness};
+    \end{tikzpicture}
+    \caption{Proof diagram for \Cref{thm:cca:rtp:tms} depicting the general structure of robust preservation proofs.}\label{fig:proofdiag:rtp}
+  \end{center}
+\end{figure}
 

### 2023-06-23 popl/main.tex
relwork

@@
 \section{Related Work}\label{sec:relwork}
+
+This section compares existing work with the one presented in this paper.
+To this end, work on robust preservation criteria (\Cref{subsec:relw:seccomprtp}) and on other secure compiler criteria (\Cref{subsec:relw:seccompcrit}) is being looked at.
+The case study of this paper (\Cref{sec:casestud:defs,sec:casestud:rtp}) implements measures for ensuring \gls{ms} as well as \gls{cct} and these are compared with previous work as well (\Cref{subsec:relw:msmechs,subsec:relw:cctmechs}).
+
+\subsection{Secure Compilation as Robust Preservation}\label{subsec:relw:seccomprtp}
+
+The robust preservation of properties as a compiler--level criterion has been analyzed extensively~\cite{abate2019jour,patrignani2021rsc,abate2021extacc,patrignani2022universal,patrignani2019survey,kruse2022csc}.
+This work has made use of that previous work and formalized the compositionality aspect which has not been explored before.
+While there exists a composition theorem~\cite{abate2019jour} already, that theorem is not concerned with composing robustly safe compilers, but rather program components and contexts.
+The work relating robust preservation with universal composability~\cite{patrignani2022universal} is closest to what this paper presents.
+The authors demonstrate a similar compositionality theorem to what is presented here (\Cref{sec:compcomp}).
+However, they do not demonstrate the scalability of the approach in the context of a practical compilation chain.
+
+\subsection{Other Secure Compilation Criteria}\label{subsec:relw:seccompcrit}
+
+\MKin{category theory stuff}
+\MKin{full abstraction -> not compositional}
+\MKin{compositional compiler correctness?}
+
+\subsection{Memory Safety Mechanisms}\label{subsec:relw:msmechs}
+
+\cite{michael2023mswasm}
+
+\subsection{Cryptographic Constant Time Mechanisms}\label{subsec:relw:cctmechs}
+
 \section{Conclusion}\label{sec:concl}

### 2023-06-24 popl/main.tex
Parts of Relwork

@@
 
-\section{Introduction}
+\section{Introduction\pages{4}}
 
@@
 
-\section{Background}\label{sec:background}
+\section{Background\pages{1}}\label{sec:background}
 \Cref{subsec:bg:tprop} introduces trace properties, monitors, and property satisfaction.
@@
 
-\section{Composing Properties}\label{sec:compprop}
+\section{Composing Properties\pages{3}}\label{sec:compprop}
 \Cref{subsec:propdefs} introduces a ,,most general'' trace model and defines properties (\gls{tms}, \gls{sms}, \gls{ms}, and \gls{cct}) that the case study (\Cref{sec:casestud}) uses.
@@
 
-\section{Composing Secure Compilers}\label{sec:compcomp}
+\section{Composing Secure Compilers\pages{1\sfrac{1}{2}}}\label{sec:compcomp}
 
@@
 %\section{Case Study: Memory Safe and Cryptographic Constant Time Code}\label{sec:casestud}
-\section{Case Study: Preliminary Definitions}\label{sec:casestud:defs}
+\section{Case Study: Preliminary Definitions\pages{6}}\label{sec:casestud:defs}
 
@@
 
-\section{Case Study: Secure Compiler Compositions in the Wild}\label{sec:casestud:rtp}
+\section{Case Study: Secure Compiler Compositions in the Wild\pages{7}}\label{sec:casestud:rtp}
 
@@
 
-\section{Related Work}\label{sec:relwork}
+\section{Related Work\pages{2}}\label{sec:relwork}
 
@@
 The robust preservation of properties as a compiler--level criterion has been analyzed extensively~\cite{abate2019jour,patrignani2021rsc,abate2021extacc,patrignani2022universal,patrignani2019survey,kruse2022csc}.
-This work has made use of that previous work and formalized the compositionality aspect which has not been explored before.
+This work has made use of that previous work.
 While there exists a composition theorem~\cite{abate2019jour} already, that theorem is not concerned with composing robustly safe compilers, but rather program components and contexts.
 The work relating robust preservation with universal composability~\cite{patrignani2022universal} is closest to what this paper presents.
-The authors demonstrate a similar compositionality theorem to what is presented here (\Cref{sec:compcomp}).
-However, they do not demonstrate the scalability of the approach in the context of a practical compilation chain.
+The authors demonstrate a similar compositionality theorem to what is presented here (\Cref{sec:compcomp}) as well as in an earlier version of this work~\cite{kruse2022csc}.
+However, they do not demonstrate the scalability of the approach in the context of a compilation chain.
 
@@
 
-\cite{michael2023mswasm}
+There is still no consensus what the formal definition of memory safety as a program property should be. %cite
+Yet, different mechanisms for memory safety exist that also consider the secure compilation domain, i.e., have an attacker model that is not just the observing attacker.
+For example, the ,,pointers as capabilities'' principle represents pointers as machine--level capabilities~\cite{korashy2021capableptrs}, which behave in a similar fashion to capabilities by means of  linear typing~\cite{morrisett2005L3}.
+The approach of this paper to memory safety and the countermeasures against memory safety insecurities are inspired by seminal work in the literature~\cite{nagarakatte2009soft,nagarakatte2010cets}.
+The view of memory safety in this paper exhibits the classic split between temporal and spatial memory safety.
+The former making sure that there are no double deallocations or use after frees.
+The latter making sure that there are no out of bounds accesses.
+When extending the languages in this paper with a less restricted form of pointer arithmetic, the region coloring memory safety monitor presented in earlier work~\cite{michael2023mswasm} can be used.
+The work presenting this monitor provides an approach for the robust {\em enforcement}, which implies robust preservation, of memory safety.
+But, they do not compose the enforcement mechanism with other secure compilation passes.
 
@@
 
-\section{Conclusion}\label{sec:concl}
+\section{Conclusion\pages{$\sfrac{1}{2}$}}\label{sec:concl}
 

### 2023-06-29 popl/main.tex
Ltms
pos
Add Ltms example

@@
 \hltt{Problem}
+
+\begin{lstlisting}[language=c,basicstyle=\small\ttfamily, caption=C]
+int strncpy(int n, int *x, int *y) {
+  for(int i = 0; x[i] != '\0' && i < n; ++i) {
+    y[i] = x[i];
+  }
+}
+\end{lstlisting}
+
+\begin{lstlisting}[language=ml,basicstyle=\small\ttfamily,morekeywords={ifz}, caption=$\Ltms$]
+let strncpy p : int * int * ptr int * ptr int -> ptr int * ptr int :=
+  let [i, n, x, y] = p in
+  ifz x[i] then
+    ifz i < n then
+      return call strncpy (i+1, n, x, y)
+    else
+      return (x, y)
+  else
+    return (x, y)
+\end{lstlisting}
 

### 2023-06-30 popl/main.tex
Move positioning lib to cmds.tex
note forme
Add author

@@
 \affiliation{%
+  \institution{CISPA Helmholtz Center for Information Security and Saarland University}
+  \streetaddress{Stuhlsatzenhaus 5}
+  \city{Saarbr{\"u}cken}
+  \state{Saarland}
+  \country{Germany}
+  \postcode{66123}
+}
+
+\author{Michael Backes}
+\email{director@cispa.de}
+%\orcid{0000-0002-7130-9211}
+\affiliation{%
   \institution{CISPA Helmholtz Center for Information Security}
@@
 %% of authors' names for this purpose.
-\renewcommand{\shortauthors}{Kruse and Patrignani}
+\renewcommand{\shortauthors}{Kruse, Backes, and Patrignani}
 
@@
       \draw[thick,dotted,->] (M) to[bend right=0,sloped] node[align=center,font=\scriptsize] {\gls{ms}\\ \Cref{thm:cccfccdce:rtp:ms}} (E);
-      \draw[thick,dotted,->] (S) to[out=-45,in=198,sloped] node[align=center,font=\scriptsize] {\gls{ms}\\ \Cref{thm:ccall:rtp:msscct}} (O);
+      \draw[thick,dotted,->] (S) to[out=-45,in=198,sloped] node[align=center,font=\scriptsize] {\gls{ms}+\gls{scct}\\ \Cref{thm:ccall:rtp:msscct}} (O);
 

### 2023-07-05 popl/main.tex
Remove example
Rewrite first part of background section

@@
 
-\begin{lstlisting}[language=ml,basicstyle=\small\ttfamily,morekeywords={ifz}, caption=$\Ltms$]
-let strncpy p : int * int * ptr int * ptr int -> ptr int * ptr int :=
-  let [i, n, x, y] = p in
-  ifz x[i] then
-    ifz i < n then
-      return call strncpy (i+1, n, x, y)
-    else
-      return (x, y)
-  else
-    return (x, y)
-\end{lstlisting}
-
 \BrText{Solution}{
 % keyword: pass
-\MPin{
-	commands missing
-}
 This paper introduces a framework for reasoning about the composition of secure and optimising compiler passes and it showcases the power of this framework by instantiating it on a multi-pass compilation chain.
@@
 % keyword- end to end, composition
-Finally, this paper shows how to compose these secure compiler passes into a multi-pass compilation chain which can be proven to provide end-to-end preservation of general memory safety and cryptographic constant time.
-\MKin{
-  ,,can be proven'' -> is proven (?)
-}
+Finally, this paper shows how to compose these secure compiler passes into a multi-pass compilation chain which is proven to provide end-to-end preservation of general memory safety and cryptographic constant time.
 Crucially, this end-to-end security preservation is obtained by having each compilation pass (i) preserve a sub-part of the overall security property individually and then (ii) compose each preserved security property as dictated by the framework of this paper.
@@
 \section{Background\pages{1}}\label{sec:background}
-\Cref{subsec:bg:tprop} introduces trace properties, monitors, and property satisfaction.
-\Cref{subsec:bg:rtp} extends the property satisfaction with robustness and shows its compiler-driven preservation, as presented in~\cite{abate2019jour}.
-
-\subsection{Trace-Based Properties, Programming Languages, and Property Satisfaction}\label{subsec:bg:tprop}
-
-\emph{Events} $\event$ are an abstract representation of observable actions a program can do during its execution.
-Typically, one always has an empty observation $\emptyevent$ used to represent unimportant actions.
-A sequence of events $\trace$ is a \emph{trace}.
-So, throughout the paper, $\overline{\cdot}$ is used to mark meta-variables as sequences.
-Empty sequences are denoted with $\hole{\cdot}$ and concatenation of traces $\trace[_{1}],\trace[_{2}]$ is written as $\trace[_{1}]\cdot\trace[_{2}]$.
-For sake of brevity, this notation also used for pre- and appending events to traces.
-
-A trace-based property $\pi$ is a set of admissable traces.
-For a trace to be admissable, a predicate evaluates to true whenever $\trace\in\pi$.
-Classes of properties $\class$, commonly referred to as hyperproperties in literature, are sets of properties.
-The \emph{lifting} of a property $\pi$ is simply its powerset and denoted as $\lift{\pi}$.
-To define property satisfaction it is integral to first look at programming languages.
-
-For languages used in this paper, a small-step style semantics is used.
-That is, there is a step relation between states $\runtimetermvar$ and an event $\event$: $\stepto[]{\runtimetermvar[_{1}]}{\runtimetermvar[_{2}]}{\event}$.
-Also, a program's syntax $p$ is encoded in the state $\runtimetermvar$ which is written as $p\in\runtimetermvar$.
-For states $\runtimetermvar$, it is assumed that there exists a judgement $\isval{\runtimetermvar}$ that determines wether $\runtimetermvar$ is a final state or not.
-Another assumption is the existence of a generic star-step relation $\stepto{\runtimetermvar[_{1}]}{\runtimetermvar[_{2}]}{\trace}$, essentially modelling a big-step style evaluation, and the existence of an $n$-step relation $\stepton{\runtimetermvar[_{1}]}{\runtimetermvar[_{2}]}{\trace}$ that evaluates a program for $n$ steps.
-Both relations filter $\varepsilon$ events.
-
-The \emph{behavior} of a whole program $w$ ($\behav{w}$) is a set containing all traces the program may emit during execution.
-Note that for deterministic languages, $\behav{w}$ is a singleton set for any $w$.
+To reason about security, this section makes the concepts of properties and the satisfaction thereof precise (\Cref{subsec:bg:tprop}).
+Furthermore, the section introduces robustness in the context of both satisfaction and secure compilers (\Cref{subsec:bg:rtp}) in similar fashion to existing work~\cite{abate2019jour}.
+
+\subsection{Properties, Satisfaction, and Robustness}\label{subsec:bg:tprop}
+
+The intuition behind security is that nothing bad happens.
+Properties ($\pi$) make this intuition formally precise, however, they rely on a trace model.
+{\em Traces} ($\trace$) are (possibly infinite) sequences of {\em events} ($\event$), which are an abstract representation of observable actions.
+The empty event ($\emptyevent$) is an uboservable or unimportant action.
+Throughout the paper, empty sequences are denoted with $\hole{\cdot}$ and concatenation of sequences $\trace[_{1}],\trace[_{2}]$ with $\trace[_{1}]\cdot\trace[_{2}]$.
+This notation also used for prepending events to traces, i.e., $\event\cdot\trace$.
+
+A property $\pi$ is a set of admissable traces and $\pi$ is a security property if a finite prefix of a trace $\trace$ is enough to decide wether $\trace\in\pi$ or $\trace\notin\pi$.
+For example, given the trace $\ev{Send\ 1}\cdot\ev{Recv\ 1}\cdot\ev{Send\ 2}\cdot\dots$ and the property $\pi=\left\{\trace \mid \trace=\hole{\cdot} \text{ or } \trace=\ev{Send\ n}\cdot\ev{Recv\ 2n}\cdot\trace[']\wedge \trace[']\in\pi\right\}$, the prefix $\ev{Send\ 1}\cdot\ev{Recv\ 1}$ immediately violates $\pi$.
+The set of security properties is one of many classes of properties ($\class$).
+Note that existing literature refers to classes of properties as hyperproperties~\cite{clarkson2008hyper}.
+The framework developed in this paper relies on {\em liftings} ($\lift{\pi}$), which are simply the powerset of a property $\pi$ and thus also considered as a class.
+This framework is concerned about the secure compilation of programs that admit properties.
+
+Programs ($\runtimetermvar$) without free variables are considered {\em whole}.
+Contrary, they are {\em partial} if they contain variables without an accompanying definition.
+Whole programs can be run with the multi-step relation $\stepto{\runtimetermvar}{\runtimetermvar[']}{\trace}$ and yield a trace $\trace$.
+Moreover, the paper assumes for such executions that $\runtimetermvar[']$ is {\em final}, decided by the judgement $\isval{\runtimetermvar}$.\MK{define final?}
+Partial programs cannot be executed, since execution would get stuck when reaching an undefined variable.
+The paper also uses other step-relations, such as a single-step ($\stepto[]{\runtimetermvar[_{1}]}{\runtimetermvar[_{2}]}{\event}$) and an $n$-step relation ($\stepton{\runtimetermvar[_{1}]}{\runtimetermvar[_{2}]}{\trace}$) that performs $n$ single-steps.
+Note that the multi-step relation presented earlier is defined as-standard using the single-step, where the reflexive case makes use of $\isval{\runtimetermvar}$.
+All multi-step relations filter $\emptyevent$ events.
+
+The {\em behavior} of a whole program $\runtimetermvar$ ($\behav{\runtimetermvar}$) is a set containing all traces the program may emit during execution.
+%This paper only considers deterministic languages, where the behavior $\behav{\runtimetermvar}$ is a singleton set for any whole $\runtimetermvar$.
 The behavior of programs gives rise to the definition of satisfaction:
 
-\begin{definition}[Property Satisfaction]\label{def:propsat}
-  A \emph{whole} program $w$ satisfies $\pi$ ($\sat{w}{\pi}$) iff $\behav{w}\in\lift{\pi}$.
-\end{definition}
+%\begin{definition}[Property Satisfaction]\label{def:propsat}
+%  A whole program $\runtimetermvar$ satisfies $\pi$ ($\sat{\runtimetermvar}{\pi}$) iff $\behav{\runtimetermvar}\in\lift{\pi}$.
+%\end{definition}
 \begin{definition}[Class Satisfaction]\label{def:classsat}
-  A \emph{whole} program $w$ satisfies $\class$ ($\sat{w}{\class}$) iff $\behav{w}\in\class$.
+  A whole program $\runtimetermvar$ satisfies $\class$ ($\sat{\runtimetermvar}{\class}$) iff $\behav{\runtimetermvar}\in\class$.
 \end{definition}
@@
 \subsection{Shared Language Definitions}\label{subsec:cs:defs}
+\MPin{
+	explain with example ? leave text for app
+}
+
 \begin{gather*}

### 2023-07-05 popl/main.tex
Pass over BG

@@
 
-Programs ($\runtimetermvar$) without free variables are considered {\em whole}.
+A program ($\runtimetermvar$) is a tuple consisting of some dynamic state ($\statevar$) and a syntactic description ($\progvar$).
+Programs without free variables are considered {\em whole}.
 Contrary, they are {\em partial} if they contain variables without an accompanying definition.
 Whole programs can be run with the multi-step relation $\stepto{\runtimetermvar}{\runtimetermvar[']}{\trace}$ and yield a trace $\trace$.
-Moreover, the paper assumes for such executions that $\runtimetermvar[']$ is {\em final}, decided by the judgement $\isval{\runtimetermvar}$.\MK{define final?}
+Moreover, the paper assumes for such executions that $\runtimetermvar[']$ is {\em final}, decided by the judgement $\isval{\runtimetermvar}$.
 Partial programs cannot be executed, since execution would get stuck when reaching an undefined variable.
@@
 %This paper only considers deterministic languages, where the behavior $\behav{\runtimetermvar}$ is a singleton set for any whole $\runtimetermvar$.
-The behavior of programs gives rise to the definition of satisfaction:
-
-%\begin{definition}[Property Satisfaction]\label{def:propsat}
-%  A whole program $\runtimetermvar$ satisfies $\pi$ ($\sat{\runtimetermvar}{\pi}$) iff $\behav{\runtimetermvar}\in\lift{\pi}$.
-%\end{definition}
-\begin{definition}[Class Satisfaction]\label{def:classsat}
-  A whole program $\runtimetermvar$ satisfies $\class$ ($\sat{\runtimetermvar}{\class}$) iff $\behav{\runtimetermvar}\in\class$.
+Finally, the satisfaction of a class (or lifted property) is defined as follows:
+
+\begin{definition}[Satisfaction]\label{def:propsat}
+  A whole program $\runtimetermvar$ satisfies $\pi$ ($\sat{\runtimetermvar}{\pi}$) iff $\behav{\runtimetermvar}\in\pi$.
 \end{definition}
 
-\subsection{Attacker Model and Robust Trace-Property Preservation}\label{subsec:bg:rtp}
-
-Partial programs cannot be executed, because they lack definitions for some symbols.
-Providing definitions for symbols is usually the job of the \emph{linker}.
-Notationally, linking two programs $\progvar[_{1}],\progvar[_{2}]$ is written $\link{\progvar[_{1}]}{\progvar[_{2}]}$.
-For sake of simplicity, linking does not fail.
-With linking, it is possible to define robust satisfaction analogously to \Cref{def:propsat,def:classsat}:
-
-\begin{definition}[Robust Property Satisfaction]\label{def:proprsat}
-  A program $\progvar$ robustly satisfies $\pi$ ($\rsat{\progvar}{\pi}$) iff for any program $\contextvar$ it holds that $\sat{\link{\contextvar}{\progvar}}{\pi}$.
-(similarily for $\rsat{\progvar}{\class}$)
+As described earlier, partial programs cannot be executed, because they lack definitions and, thus, get stuck.
+{\em Linking} ($\link{\progvar[_{1}]}{\progvar[_{2}]}$) syntactic descriptions $\progvar[_{1}],\progvar[_{2}]$ can give rise to a whole program.
+Note that linking is not defined on programs, but merely on their syntactic descriptions, since programs contain dynamic state.
+However, for convenience, linking of programs $\runtimetermvar[_{1}],\runtimetermvar[_{2}]$ shall be defined if both $\runtimetermvar[_{1}]$ and $\runtimetermvar[_{2}]$ have an initial dynamic state.
+As typically done in a secure compilation context~\cite{abate2019jour}, this paper also assumes that programs are the result of linking a {\em context} ($\ctx$) and a {\em component} ($\comp$), which are just any syntactic description.
+The component part of a whole program is the main object of interest and thus a known syntactic description.
+In contrast, the context is usually arbitrary and thus has the role of an active {\em attacker} that can interact with the component by means of whatever features the programming language has.
+The following definition extends on \Thmref{def:propsat} to cover {\em robustness} and exploits this interaction between the context and the component:
+
+\begin{definition}[Robust Satisfaction]\label{def:proprsat}
+  A program $\runtimetermvar$ robustly satisfies $\pi$ ($\rsat{\runtimetermvar}{\pi}$) iff for any partial program $\contextvar$ it holds that $\sat{\link{\contextvar}{\runtimetermvar}}{\pi}$.
 \end{definition}
 
-Robust satisfaction gives rise to a maximally powerful attacker model: Any other partial program, here referred to as context $\contextvar$, is the attacker.
-So, for example, if the programming language of $\progvar$ and $\contextvar$ allows guessing of memory addresses, then $\contextvar$ is able to read and write any position in memory and therefore guess the location of secrets.
-
-It is expected that compilers preserve robust satisfaction.
-In this paper, compilers are modelled as function from one programming language's syntax to the other.
-Notationally, $\cc{\src{L}}{\trg{L}}$ denotes a compiler from $\src{L}$ to $\trg{L}$ and the paper refers to $\src{L}$ as \emph{source} language and to $\trg{L}$ as \emph{target} language.
-Now, define the compiler-property robust trace-property preservation.
-
-\begin{definition}[Robust Trace-Property Preservation]\label{def:rtp}
-  A compiler $\cc{\src{L}}{\trg{L}}$ robustly preserves a class of properties $\class$ ($\rtp{\cc{\src{L}}{\trg{L}}}{\class}$) iff for any property $\pi\in\class$ and $\src{p}$ written in $\src{L}$, given $\rsat{\src{\progvar}}{\pi}$, then $\rsat{\cc{\src{L}}{\trg{L}}\left(\src{p}\right)}{\pi}$.
+\subsection{Secure Compilers}\label{subsec:bg:rtp}
+
+A {\em compiler} ($\cc{\src{L}}{\trg{L}}$) translates syntactic descriptions of programs from a {\em source} ($\src{L}$) into a {\em target} ($\trg{L}$) programming language.
+The syntactic description of programs written in $\src{L}$ is denoted as $\src{\progvar}$, while for ones written in $\trg{L}$ this paper writes $\trg{\progvar}$.
+It is considered good if this translation is semantics-preserving, i.e., correct.
+Correct compilers may still be insecure compilers.
+Programs translated by insecure compilers may violate security properties that were assumed to hold, given the presence of a context--level program.
+So, ideally, a compiler is secure~\cite{abate2019jour} in the sense of the following definition:
+
+\begin{definition}[Robust Preservation]\label{def:rtp}
+  A compiler $\cc{\src{L}}{\trg{L}}$ robustly--preserves a class of properties $\class$ ($\rtp{\cc{\src{L}}{\trg{L}}}{\class}$) iff for any property $\pi\in\class$ and component $\src{p}$ such that $\rsat{\src{\progvar}}{\pi}$, it holds that $\rsat{\cc{\src{L}}{\trg{L}}\left(\src{p}\right)}{\pi}$.
 \end{definition}
 
-Note that in the conclusion of \Cref{def:rtp} there is a target-level context.
-So, programs that robustly satisfy a property are hardened against a maximally malicious environment with respect to the power of the target language.
-
-\section{Composing Properties\pages{3}}\label{sec:compprop}
+\section{Example Properties and their Monitors\pages{3}}\label{sec:compprop}
 \Cref{subsec:propdefs} introduces a ,,most general'' trace model and defines properties (\gls{tms}, \gls{sms}, \gls{ms}, and \gls{cct}) that the case study (\Cref{sec:casestud}) uses.

### 2023-07-05 popl/main.tex
Start pass of Sec. 3

@@
 \section{Example Properties and their Monitors\pages{3}}\label{sec:compprop}
-\Cref{subsec:propdefs} introduces a ,,most general'' trace model and defines properties (\gls{tms}, \gls{sms}, \gls{ms}, and \gls{cct}) that the case study (\Cref{sec:casestud}) uses.
-By means of taking these properties as an example, \Cref{subsec:monitors} defines monitors for them and an agreement relation between a monitor's trace model with the most general trace model.
-Then, monitor-based satisfaction is introduced and it is proven that trace-property satisfaction is implied by it.
-Since monitor-based satisfaction is easier to reason about, this simplifies later proofs considerably.
-
-\subsection{Most General Trace-Model and Some Property Definitions}\label{subsec:propdefs}
+
+This section introduces a trace model (\Cref{subsec:propdefs}) that is used to define properties (\gls{tms}, \gls{sms}, \gls{ms}, and \gls{scct}).
+The properties are of practical importance and also of interest in the case study (\Cref{sec:casestud}) this paper presents.
+The study uses monitors, which are able to check wether a trace admits a property or not.
+The use of monitors is because they have some structure useful for inductive proofs, whereas traces lack this structure, thus making proofs difficult.
+So, this section also presents monitors for each of the defined properties (\Cref{subsec:monitors}) and introduces monitor-based satisfaction.
+Finally, for each property, it is shown that monitor-based satisfaction implies property satisfaction.
+
+\subsection{Specification Trace Model}\label{subsec:propdefs}
 

### 2023-07-05 popl/main.tex
Redo pass over BG

@@
 Throughout the paper, empty sequences are denoted with $\hole{\cdot}$ and concatenation of sequences $\trace[_{1}],\trace[_{2}]$ with $\trace[_{1}]\cdot\trace[_{2}]$.
-This notation also used for prepending events to traces, i.e., $\event\cdot\trace$.
+For prepending events to traces the paper uses the same notation, i.e., $\event\cdot\trace$.
 
@@
 %This paper only considers deterministic languages, where the behavior $\behav{\runtimetermvar}$ is a singleton set for any whole $\runtimetermvar$.
-Finally, the satisfaction of a class (or lifted property) is defined as follows:
+This set is used to define the satisfaction of a property as follows:
 
@@
 In contrast, the context is usually arbitrary and thus has the role of an active {\em attacker} that can interact with the component by means of whatever features the programming language has.
-The following definition extends on \Thmref{def:propsat} to cover {\em robustness} and exploits this interaction between the context and the component:
+The following extends on \Thmref{def:propsat} to cover {\em robustness} and exploits this interaction between the context and the component:
 
@@
 A {\em compiler} ($\cc{\src{L}}{\trg{L}}$) translates syntactic descriptions of programs from a {\em source} ($\src{L}$) into a {\em target} ($\trg{L}$) programming language.
-The syntactic description of programs written in $\src{L}$ is denoted as $\src{\progvar}$, while for ones written in $\trg{L}$ this paper writes $\trg{\progvar}$.
+For the syntactic description of programs written in $\src{L}$, this paper writes $\src{\progvar}$, while for ones written in $\trg{L}$, this paper writes $\trg{\progvar}$.
 It is considered good if this translation is semantics-preserving, i.e., correct.
-Correct compilers may still be insecure compilers.
+Correct compilers may still be insecure compilers\MK{citation needed}.
 Programs translated by insecure compilers may violate security properties that were assumed to hold, given the presence of a context--level program.
@@
 This section introduces a trace model (\Cref{subsec:propdefs}) that is used to define properties (\gls{tms}, \gls{sms}, \gls{ms}, and \gls{scct}).
-The properties are of practical importance and also of interest in the case study (\Cref{sec:casestud}) this paper presents.
+The properties are of practical importance and also of interest in the case study (\Cref{sec:casestud:defs,sec:casestud:rtp}) this paper presents.
 The study uses monitors, which are able to check wether a trace admits a property or not.

### 2023-07-05 popl/main.tex
edits to entry to 2

@@
 
-\section{Background\pages{1}}\label{sec:background}
-To reason about security, this section makes the concepts of properties and the satisfaction thereof precise (\Cref{subsec:bg:tprop}).
-Furthermore, the section introduces robustness in the context of both satisfaction and secure compilers (\Cref{subsec:bg:rtp}) in similar fashion to existing work~\cite{abate2019jour}.
-
-\subsection{Properties, Satisfaction, and Robustness}\label{subsec:bg:tprop}
-
-The intuition behind security is that nothing bad happens.
-Properties ($\pi$) make this intuition formally precise, however, they rely on a trace model.
-{\em Traces} ($\trace$) are (possibly infinite) sequences of {\em events} ($\event$), which are an abstract representation of observable actions.
-The empty event ($\emptyevent$) is an uboservable or unimportant action.
-Throughout the paper, empty sequences are denoted with $\hole{\cdot}$ and concatenation of sequences $\trace[_{1}],\trace[_{2}]$ with $\trace[_{1}]\cdot\trace[_{2}]$.
-For prepending events to traces the paper uses the same notation, i.e., $\event\cdot\trace$.
-
-A property $\pi$ is a set of admissable traces and $\pi$ is a security property if a finite prefix of a trace $\trace$ is enough to decide wether $\trace\in\pi$ or $\trace\notin\pi$.
+\section{Background: Security Properties and Secure Compilers\pages{1}}\label{sec:background}
+To introduce the security argument of this paper, this section first presents the concepts of (security) properties, of their satisfaction, and of their robust satisfaction (i.e., satisfaction in the presence of an active attacker; \Cref{subsec:bg:tprop}).
+Then, borrowing from existing work~\cite{abate2019jour,abate2021extacc,patrignani2021rsc}, the section introduces secure compilers as compilers that preserve robust property satisfaction (\Cref{subsec:bg:rtp}).
+
+\subsection{Properties, Satisfaction, and Robust Satisfaction}\label{subsec:bg:tprop}
+
+% The intuition behind security is that nothing bad happens.
+% \MPin{
+% 	no: that is safety
+% }
+This paper employs the security model where programs are written in a language whose semantics emits events $\event$.
+Events include security-relevant actions (e.g., reading ad writing to memory) and the unobservable event $\emptyevent$.
+As programs executes, their emitted events are concatenated in traces ($\trace$), which serve as the description of the behaviour of a program.%
+\footnote{
+Throughout the paper, we denote empty sequences with $\hole{\cdot}$ and concatenate sequences $\trace[_{1}],\trace[_{2}]$ as $\trace[_{1}]\cdot\trace[_{2}]$.
+Prepending elements to sequences uses the same notation: $\event\cdot\trace$.
+}
+
+Properties ($\pi$) are sets of traces of admissible program behaviours, ascribing what said property considers valid.\MP{ cit }
+
+% make this intuition formally precise, however, they rely on a trace model.
+% {\em Traces} ($\trace$) are (possibly infinite) sequences of {\em events} ($\event$).
+% Events are an abstract representation of observable actions, in this paper
+% The empty event ($\emptyevent$) is an unobservable (or unimportant) action.
+% A property $\pi$ is a set of admissable traces and 
+Properties and partitioned in the subsets of safety and liveness\MP{cit}, and given their central role for security, this paper focusses on safety properties.
+Given a safety property, a finite prefix of a longer trace $\trace$ is enough to decide whether a trace is admitted or not by the property. % $\trace\in\pi$ or $\trace\notin\pi$.
 For example, given the trace $\ev{Send\ 1}\cdot\ev{Recv\ 1}\cdot\ev{Send\ 2}\cdot\dots$ and the property $\pi=\left\{\trace \mid \trace=\hole{\cdot} \text{ or } \trace=\ev{Send\ n}\cdot\ev{Recv\ 2n}\cdot\trace[']\wedge \trace[']\in\pi\right\}$, the prefix $\ev{Send\ 1}\cdot\ev{Recv\ 1}$ immediately violates $\pi$.
-The set of security properties is one of many classes of properties ($\class$).
-Note that existing literature refers to classes of properties as hyperproperties~\cite{clarkson2008hyper}.
-The framework developed in this paper relies on {\em liftings} ($\lift{\pi}$), which are simply the powerset of a property $\pi$ and thus also considered as a class.
-This framework is concerned about the secure compilation of programs that admit properties.
-
-A program ($\runtimetermvar$) is a tuple consisting of some dynamic state ($\statevar$) and a syntactic description ($\progvar$).
-Programs without free variables are considered {\em whole}.
-Contrary, they are {\em partial} if they contain variables without an accompanying definition.
-Whole programs can be run with the multi-step relation $\stepto{\runtimetermvar}{\runtimetermvar[']}{\trace}$ and yield a trace $\trace$.
-Moreover, the paper assumes for such executions that $\runtimetermvar[']$ is {\em final}, decided by the judgement $\isval{\runtimetermvar}$.
-Partial programs cannot be executed, since execution would get stuck when reaching an undefined variable.
-The paper also uses other step-relations, such as a single-step ($\stepto[]{\runtimetermvar[_{1}]}{\runtimetermvar[_{2}]}{\event}$) and an $n$-step relation ($\stepton{\runtimetermvar[_{1}]}{\runtimetermvar[_{2}]}{\trace}$) that performs $n$ single-steps.
-Note that the multi-step relation presented earlier is defined as-standard using the single-step, where the reflexive case makes use of $\isval{\runtimetermvar}$.
-All multi-step relations filter $\emptyevent$ events.
-
-The {\em behavior} of a whole program $\runtimetermvar$ ($\behav{\runtimetermvar}$) is a set containing all traces the program may emit during execution.
+\MP{mention send and recv before already. replace with read/write to contextualise to memory?}
+
+% The set of security properties is one of many classes of properties ($\class$).
+% Note that existing literature refers to classes of properties as hyperproperties~\cite{clarkson2008hyper}.
+% The framework developed in this paper relies on {\em liftings} ($\lift{\pi}$), which are simply the powerset of a property $\pi$ and thus also considered as a class.
+% This framework is concerned about the secure compilation of programs that admit properties.
+\MPin{
+	this par does not flow.
+	what is the point?
+}
+
+% A program ($\runtimetermvar$) is a tuple consisting of some dynamic state ($\statevar$) and a syntactic description ($\progvar$).
+% Programs without free variables are considered {\em whole}.
+% Contrary, they are {\em partial} if they contain variables without an accompanying definition.
+% Whole programs can be run with the multi-step relation $\stepto{\runtimetermvar}{\runtimetermvar[']}{\trace}$ and yield a trace $\trace$.
+% Moreover, the paper assumes for such executions that $\runtimetermvar[']$ is {\em final}, decided by the judgement $\isval{\runtimetermvar}$.
+% Partial programs cannot be executed, since execution would get stuck when reaching an undefined variable.
+% The paper also uses other step-relations, such as a single-step ($\stepto[]{\runtimetermvar[_{1}]}{\runtimetermvar[_{2}]}{\event}$) and an $n$-step relation ($\stepton{\runtimetermvar[_{1}]}{\runtimetermvar[_{2}]}{\trace}$) that performs $n$ single-steps.
+% Note that the multi-step relation presented earlier is defined as-standard using the single-step, where the reflexive case makes use of $\isval{\runtimetermvar}$.
+% All multi-step relations filter $\emptyevent$ events.
+\MPin{
+	all unnecessary.
+	unused notation, not needed for sec2, for sec3 and for sec 4.
+	also i don't agree with what a program is.
+}
+
+Given a whole program $\runtimetermvar$, i.e., \MP{ fill in}, the {\em behaviour} of a whole program $\runtimetermvar$ ($\behav{\runtimetermvar}$) is the set containing all traces the program can emit during execution according to its language semantics.
 %This paper only considers deterministic languages, where the behavior $\behav{\runtimetermvar}$ is a singleton set for any whole $\runtimetermvar$.
-This set is used to define the satisfaction of a property as follows:
+This set is used to define property satisfaction (\Cref{def:propsat}):
 
@@
 \end{definition}
+\MPin{
+	this definition has 1/2 explained in text, 1/2 in formalism.
+}
+\MP{
+	can't we use p for programs?
+}
 
@@
 \end{definition}
-
 \subsection{Secure Compilers}\label{subsec:bg:rtp}

### 2023-07-05 popl/main.tex
some todos for the rest of 2.1

@@
 
-As described earlier, partial programs cannot be executed, because they lack definitions and, thus, get stuck.
+As described earlier, partial programs cannot be executed \MP{not true anymore, need to define the notion of partial programs} , because they lack definitions and, thus, get stuck.
 {\em Linking} ($\link{\progvar[_{1}]}{\progvar[_{2}]}$) syntactic descriptions $\progvar[_{1}],\progvar[_{2}]$ can give rise to a whole program.
-Note that linking is not defined on programs, but merely on their syntactic descriptions, since programs contain dynamic state.
+Note that linking is not defined on programs, but merely on their syntactic descriptions, since programs contain dynamic state.\MP{ not terminology i really want to use, nor is defined}
 However, for convenience, linking of programs $\runtimetermvar[_{1}],\runtimetermvar[_{2}]$ shall be defined if both $\runtimetermvar[_{1}]$ and $\runtimetermvar[_{2}]$ have an initial dynamic state.
-As typically done in a secure compilation context~\cite{abate2019jour}, this paper also assumes that programs are the result of linking a {\em context} ($\ctx$) and a {\em component} ($\comp$), which are just any syntactic description.
+As typically done in a secure compilation context~\cite{abate2019jour}, \MP{others do this too, parametricity results e.g.} this paper also assumes that programs are the result of linking a {\em context} ($\ctx$) and a {\em component} ($\comp$), which are just any syntactic description.
+\MP{just introduce the notion of contexts}
 The component part of a whole program is the main object of interest and thus a known syntactic description.
 In contrast, the context is usually arbitrary and thus has the role of an active {\em attacker} that can interact with the component by means of whatever features the programming language has.
+\MP{and say the semantics is expected (passive, hah) to differentiate between prg and ctx}
 The following extends on \Thmref{def:propsat} to cover {\em robustness} and exploits this interaction between the context and the component:
+\MPin{missing: the context is an attacker! now, unlike before, we security property satisfaction with a clear active attacker}
 
@@
 \end{definition}
+
+\MPin{
+	classes?
+}
 \subsection{Secure Compilers}\label{subsec:bg:rtp}

### 2023-07-05 popl/main.tex
transition sentence from 2.1 to 2.2
Beginning of spec model

--- 2023-07-05 popl/main.tex
@@
 
-This section introduces a trace model (\Cref{subsec:propdefs}) that is used to define properties (\gls{tms}, \gls{sms}, \gls{ms}, and \gls{scct}).
-The properties are of practical importance and also of interest in the case study (\Cref{sec:casestud:defs,sec:casestud:rtp}) this paper presents.
+This section introduces a trace model (\Cref{subsec:propdefs}) that is used to define some example properties (\gls{tms}, \gls{sms}, \gls{ms}, and \gls{scct}).
+These properties are of practical importance and also of interest in the case study (\Cref{sec:casestud:defs,sec:casestud:rtp}) this paper presents.
 The study uses monitors, which are able to check wether a trace admits a property or not.
@@
 
+The specification trace model has events that carry different meta-information, such as security- and control--tags.
 Security tags describe wether an event contains sensitive information ($\lock$) or not ($\unlock$).
 Control tags state wether the context ($\ctx$) or the component ($\comp$) are responsible for emitting the event.
-Pre-events describe the actual kind of event that happened.
-$\ev{Alloc\ \loc\ n}$ states that an allocation of size $n$ happened and is stored at $\loc$.
-Locations $\loc$ are elements of $L$ and left abstract.
-However, it is expected that $L$ is closed over addition with natural numbers.
-$\ev{Dealloc\ \loc}$ announces that location $\loc$ is freed.
-$\ev{Use\ \loc\ n}$ abstracts over reads from $\loc$ and writes to $\loc$ in memory, where $n$ is read/written, respectively.
-$\ev{Branch\ n}$ and $\ev{Binop\ n}$ leak values involved in a branch or in a binary operation with data-dependent execution time.
-Finally, events are either the emptyevent ($\emptyevent$), a program crash ($\lightning$), or a tuple consisting of a pre-event, a control tag, and a security tag.
-
-This trace model is used to define any property of interest.
-In a sense, the trace model is most general, because it abstracts all events done in the languages as defined in \Cref{sec:casestud}.
-This papers uses $\event[_{1}]\le_{\trace}\event[_{2}]$ if $\event[_{1}]$ occurs at an earlier position in $\trace$ than $\event[_{2}]$.
+%Even though this could be used for blame~\cite{patrignani2023blame}, this is not what this paper is concerned about.
+Now, pre--events describe the actual kind of event that happened.
+One such kind is the allocation event ($\ev{Alloc\ \loc\ n}$) that happens whenever a program claims $n$ cells of memory and stores them at address $\loc$.
+Likewise, deallocation ($\ev{Dealloc\ \loc}$) announces that the object at location $\loc$ is freed.
+These two events alone allow already a partial description of \gls{tms} by requiring that e.g. there is only one deallocation event that carries a location $\loc$.
+To be able to express \gls{sms}, there is also an event to describe reads from and writes to memory ($\ev{Use\ \loc\ n}$).
+For cryptographic code, there is a general guideline that secrets must not be visible on a trace.
+Moreover, an instruction whose timing is data--dependent must not have a secret as an operand.
+Typical operations with data--dependent timing are branches and certain binary operations, such as division.
+Both operations are also modelled in the specification trace model ($\ev{Branch\ n}$ and $\ev{Binop\ n}$).
+Lastly, events are either the empty event ($\emptyevent$), a program crash ($\lightning$), or a tuple consisting of a pre-event, a control--tag, and a security--tag.
 

### 2023-07-05 popl/main.tex

--- 2023-07-05 popl/main.tex
@@
 	classes?
-}
+	% 
+}
+Classes of properties can be used to state whether a compiler is correct, or even secure so now we turn to the definition of (secure) compilers.
+
 \subsection{Secure Compilers}\label{subsec:bg:rtp}

### 2023-07-05 popl/main.tex
some in 2.2

--- 2023-07-05 popl/main.tex
@@
 	classes?
-}
+	% 
+}
+Classes of properties can be used to state whether a compiler is correct, or even secure so now we turn to the definition of (secure) compilers.
+
 \subsection{Secure Compilers}\label{subsec:bg:rtp}
 
-A {\em compiler} ($\cc{\src{L}}{\trg{L}}$) translates syntactic descriptions of programs from a {\em source} ($\src{L}$) into a {\em target} ($\trg{L}$) programming language.
-For the syntactic description of programs written in $\src{L}$, this paper writes $\src{\progvar}$, while for ones written in $\trg{L}$, this paper writes $\trg{\progvar}$.
-It is considered good if this translation is semantics-preserving, i.e., correct.
+A {\em compiler} ($\cc{\src{L}}{\trg{L}}$) translates syntactic descriptions of programs from a {\em source} ($\src{L}$) into a {\em target} ($\trg{L}$) programming language.\MP{good sentence}
+For the syntactic description of programs written in $\src{L}$, this paper writes $\src{\progvar}$, while for ones written in $\trg{L}$, this paper writes $\trg{\progvar}$.\MP{not a fan of this sentence}
+It is considered good if this translation is semantics-preserving, i.e., correct. \MP{passive}
 Correct compilers may still be insecure compilers\MK{citation needed}.
@@
 So, ideally, a compiler is secure~\cite{abate2019jour} in the sense of the following definition:
+\MPin{does not really flow.
+% 
+start with compilers (as now),
+then: for a compiler to be correct we expect ... ... (no def)
+then: for a compiler to be secure we use the robust compilation framework of ... that says ... ... def
+}
 
@@
 \end{definition}
+
+\MPin{
+	again, hybrid definition.
+	keep one version.
+	i suggest: text outside (typically before the def, in the text leading to it), formalism inside
+}
 

### 2023-07-05 popl/main.tex
Fix some typos

--- 2023-07-05 popl/main.tex
@@
 \section{Background: Security Properties and Secure Compilers\pages{1}}\label{sec:background}
+
 To introduce the security argument of this paper, this section first presents the concepts of (security) properties, of their satisfaction, and of their robust satisfaction (i.e., satisfaction in the presence of an active attacker; \Cref{subsec:bg:tprop}).
@@
 
-\subsection{Properties, Satisfaction, and Robust Satisfaction}\label{subsec:bg:tprop}
+\subsection{Properties and (Robust) Satisfaction}\label{subsec:bg:tprop}
 
@@
 This paper employs the security model where programs are written in a language whose semantics emits events $\event$.
-Events include security-relevant actions (e.g., reading ad writing to memory) and the unobservable event $\emptyevent$.
+Events include security--relevant actions (e.g., reading from and writing to memory) and the unobservable event $\emptyevent$.
 As programs executes, their emitted events are concatenated in traces ($\trace$), which serve as the description of the behaviour of a program.%

### 2023-07-05 popl/main.tex

@@
 
-A {\em compiler} ($\cc{\src{L}}{\trg{L}}$) translates syntactic descriptions of programs from a {\em source} ($\src{L}$) into a {\em target} ($\trg{L}$) programming language.
-For the syntactic description of programs written in $\src{L}$, this paper writes $\src{\progvar}$, while for ones written in $\trg{L}$, this paper writes $\trg{\progvar}$.
-It is considered good if this translation is semantics-preserving, i.e., correct.
+A {\em compiler} ($\cc{\src{L}}{\trg{L}}$) translates syntactic descriptions of programs from a {\em source} ($\src{L}$) into a {\em target} ($\trg{L}$) programming language.\MP{good sentence}
+For the syntactic description of programs written in $\src{L}$, this paper writes $\src{\progvar}$, while for ones written in $\trg{L}$, this paper writes $\trg{\progvar}$.\MP{not a fan of this sentence}
+It is considered good if this translation is semantics-preserving, i.e., correct. \MP{passive}
 Correct compilers may still be insecure compilers\MK{citation needed}.

### 2023-07-05 popl/main.tex

@@
 \section{Background: Security Properties and Secure Compilers\pages{1}}\label{sec:background}
-
 To introduce the security argument of this paper, this section first presents the concepts of (security) properties, of their satisfaction, and of their robust satisfaction (i.e., satisfaction in the presence of an active attacker; \Cref{subsec:bg:tprop}).
@@
 
-\subsection{Properties and (Robust) Satisfaction}\label{subsec:bg:tprop}
+\subsection{Properties, Satisfaction, and Robust Satisfaction}\label{subsec:bg:tprop}
 
@@
 This paper employs the security model where programs are written in a language whose semantics emits events $\event$.
-Events include security--relevant actions (e.g., reading from and writing to memory) and the unobservable event $\emptyevent$.
+Events include security-relevant actions (e.g., reading ad writing to memory) and the unobservable event $\emptyevent$.
 As programs executes, their emitted events are concatenated in traces ($\trace$), which serve as the description of the behaviour of a program.%

### 2023-07-05 popl/main.tex
Add description of defs in text and make defs more formal/succinct

--- 2023-07-05 popl/main.tex
@@
 \usepackage{tikz}
+\usepackage{soul}
 \usepackage{bm}
+
+\setul{0.5ex}{0.3ex}
 
@@
 
-% The intuition behind security is that nothing bad happens.
-% \MPin{
-% 	no: that is safety
-% }
 This paper employs the security model where programs are written in a language whose semantics emits events $\event$.
 Events include security--relevant actions (e.g., reading from and writing to memory) and the unobservable event $\emptyevent$.
-As programs executes, their emitted events are concatenated in traces ($\trace$), which serve as the description of the behaviour of a program.%
+As programs execute, their emitted events are concatenated in traces $\trace$, which serve as the description of the behaviour of a program.%
 \footnote{
@@
 
-Properties ($\pi$) are sets of traces of admissible program behaviours, ascribing what said property considers valid.\MP{ cit }
-
+Properties $\pi$ are sets of traces of admissible program behaviours, ascribing what said property considers valid.\MP{ cit }
+%
 % make this intuition formally precise, however, they rely on a trace model.
@@
 % A property $\pi$ is a set of admissable traces and 
-Properties and partitioned in the subsets of safety and liveness\MP{cit}, and given their central role for security, this paper focusses on safety properties.
-Given a safety property, a finite prefix of a longer trace $\trace$ is enough to decide whether a trace is admitted or not by the property. % $\trace\in\pi$ or $\trace\notin\pi$.
-For example, given the trace $\ev{Send\ 1}\cdot\ev{Recv\ 1}\cdot\ev{Send\ 2}\cdot\dots$ and the property $\pi=\left\{\trace \mid \trace=\hole{\cdot} \text{ or } \trace=\ev{Send\ n}\cdot\ev{Recv\ 2n}\cdot\trace[']\wedge \trace[']\in\pi\right\}$, the prefix $\ev{Send\ 1}\cdot\ev{Recv\ 1}$ immediately violates $\pi$.
-\MP{mention send and recv before already. replace with read/write to contextualise to memory?}
-
+%Properties and partitioned in the subsets of safety and liveness\MP{cit}, and given their central role for security, this paper focusses on safety properties.
+The set of all properties can be partitioned into different {\em classes} ($\class$), i.e., safety, liveness, and neither safety or liveness.\MP{cit}
+Given the central role for security, this paper focusses on safety properties.
+Note that there are other classes $\class$ that contain more domain-specific properties, such as the class of memory safety properties.
+For the class of safety properties, it is decidable wether a trace admits such a property with just a finite trace prefix.
+As an example, consider a trace that starts with an interaction with memory where a deallocation precedes a read: $\ev{Dealloc\ \loc}\cdot\ev{Read\ \loc\ 1729}\cdot\dots$
+This program behaviour is insecure, because it reads from a memory location that was freed already.
+The finite trace prefix $\ev{Dealloc\ \loc}\cdot\ev{Read\ \loc\ 1729}$ is enough to decide that the trace does not satisfy the property ,,no use after dealloc'' and there is no way to append events to this prefix which would result in the trace being admissible.
+%Given a safety property, a finite prefix of a longer trace $\trace$ is enough to decide whether a trace is admitted or not by the property.
+%For example, given the trace $\ev{Send\ 1}\cdot\ev{Recv\ 1}\cdot\ev{Send\ 2}\cdot\dots$ and the property $\pi=\left\{\trace \mid \trace=\hole{\cdot} \text{ or } \trace=\ev{Send\ n}\cdot\ev{Recv\ 2n}\cdot\trace[']\wedge \trace[']\in\pi\right\}$, the prefix $\ev{Send\ 1}\cdot\ev{Recv\ 1}$ immediately violates $\pi$.
+%\MP{mention send and recv before already. replace with read/write to contextualise to memory?}
+%
 % The set of security properties is one of many classes of properties ($\class$).
@@
 % This framework is concerned about the secure compilation of programs that admit properties.
-\MPin{
-	this par does not flow.
-	what is the point?
-}
-
+%\MPin{
+%	this par does not flow.
+%	what is the point?
+%}
+%
 % A program ($\runtimetermvar$) is a tuple consisting of some dynamic state ($\statevar$) and a syntactic description ($\progvar$).
@@
 % All multi-step relations filter $\emptyevent$ events.
-\MPin{
-	all unnecessary.
-	unused notation, not needed for sec2, for sec3 and for sec 4.
-	also i don't agree with what a program is.
-}
-
-Given a whole program $\runtimetermvar$, i.e., \MP{ fill in}, the {\em behaviour} of a whole program $\runtimetermvar$ ($\behav{\runtimetermvar}$) is the set containing all traces the program can emit during execution according to its language semantics.
+%\MPin{
+%	all unnecessary.
+%	unused notation, not needed for sec2, for sec3 and for sec 4.
+%	also i don't agree with what a program is.
+%}
+
+%Given a whole program $\progvar$, i.e., a program without free variables, the {\em behaviour} ($\behav{\progvar}$) is the set containing all traces the program can emit during execution according to its language semantics.
 %This paper only considers deterministic languages, where the behavior $\behav{\runtimetermvar}$ is a singleton set for any whole $\runtimetermvar$.
-This set is used to define property satisfaction (\Cref{def:propsat}):
-
-\begin{definition}[Satisfaction]\label{def:propsat}
-  A whole program $\runtimetermvar$ satisfies $\pi$ ($\sat{\runtimetermvar}{\pi}$) iff $\behav{\runtimetermvar}\in\pi$.
+%This set is used to define property satisfaction (\Cref{def:propsat}):
+\bul{Programs $\progvar$ satisfy a property $\pi$} iff \iul{$\progvar$ yields a trace $\trace$} such that \oul{$\trace$ is admissible with respect to $\pi$}.
+
+\begin{definition}[Property Satisfaction]\label{def:propsat}
+  $\progvar$ satisfies $\pi$, \bul{$\sat{\progvar}{\pi}$}, iff \iul{running $\progvar$ gives $\trace$} where \oul{$\trace\in\pi$}.
 \end{definition}
-\MPin{
-	this definition has 1/2 explained in text, 1/2 in formalism.
-}
-\MP{
-	can't we use p for programs?
-}
-
-As described earlier, partial programs cannot be executed \MP{not true anymore, need to define the notion of partial programs} , because they lack definitions and, thus, get stuck.
-{\em Linking} ($\link{\progvar[_{1}]}{\progvar[_{2}]}$) syntactic descriptions $\progvar[_{1}],\progvar[_{2}]$ can give rise to a whole program.
-Note that linking is not defined on programs, but merely on their syntactic descriptions, since programs contain dynamic state.\MP{ not terminology i really want to use, nor is defined}
-However, for convenience, linking of programs $\runtimetermvar[_{1}],\runtimetermvar[_{2}]$ shall be defined if both $\runtimetermvar[_{1}]$ and $\runtimetermvar[_{2}]$ have an initial dynamic state.
-As typically done in a secure compilation context~\cite{abate2019jour}, \MP{others do this too, parametricity results e.g.} this paper also assumes that programs are the result of linking a {\em context} ($\ctx$) and a {\em component} ($\comp$), which are just any syntactic description.
-\MP{just introduce the notion of contexts}
-The component part of a whole program is the main object of interest and thus a known syntactic description.
-In contrast, the context is usually arbitrary and thus has the role of an active {\em attacker} that can interact with the component by means of whatever features the programming language has.
-\MP{and say the semantics is expected (passive, hah) to differentiate between prg and ctx}
-The following extends on \Thmref{def:propsat} to cover {\em robustness} and exploits this interaction between the context and the component:
-\MPin{missing: the context is an attacker! now, unlike before, we security property satisfaction with a clear active attacker}
+
+Given a partial program $\progvar$, i.e., a program with free variables, the execution of $\progvar$ would get stuck due to the lack of definitions.
+%As described earlier, partial programs cannot be executed \MP{not true anymore, need to define the notion of partial programs} , because they lack definitions and, thus, get stuck.
+To remedy this, {\em linking} takes two partial programs $\progvar[_{1}],\progvar[_{2}]$ and produces a new whole program $\link{\progvar[_{1}]}{\progvar[_{2}]}$.
+%Note that linking is not defined on programs, but merely on their syntactic descriptions, since programs contain dynamic state.\MP{ not terminology i really want to use, nor is defined}
+As typically done in secure compilation~\cite{abate2019jour}, \MP{others do this too, parametricity results e.g.} this paper assumes that programs are the result of linking a {\em context} ($\ctx$) and a {\em component} ($\comp$).
+% \MP{just introduce the notion of contexts}
+%       MK: why only contexts?
+%The component part of a whole program is the main object of interest and thus a known syntactic description.
+The context is usually arbitrary\footnote{Up to, e.g., provoking link failure.} and thus has the role of an {\em attacker} that can interact with the component by means of whatever features the programming language has.
+The semantics of the programming language is expected to differentiate between the component and the context.
+%\MP{and say the semantics is expected (passive, hah) to differentiate between prg and ctx}
+%%                                         -> why passive?
+So, the context interacts in some way with the component and this interaction models an attacker.
+With this, \Thmref{def:propsat} can be extended as follows.
+For \bul{programs $\progvar$ to robustly satisfy a property $\pi$}, take an \iul{attacker $\contextvar$ and link it with $\progvar$}, \oul{the result of which must satisfy $\pi$}.
+%The following definition extends on \Thmref{def:propsat} to cover property satisfaction under the presence of an (active, with respect to the semantics of the programming language) attacker: %, which is known as {\em robust satisfaction}:
+%\MPin{missing: the context is an attacker! now, unlike before, we security property satisfaction with a clear active attacker}
 
 \begin{definition}[Robust Satisfaction]\label{def:proprsat}
-  A program $\runtimetermvar$ robustly satisfies $\pi$ ($\rsat{\runtimetermvar}{\pi}$) iff for any partial program $\contextvar$ it holds that $\sat{\link{\contextvar}{\runtimetermvar}}{\pi}$.
+  $\progvar$ robustly satisfies $\pi$, \bul{$\rsat{\progvar}{\pi}$}, iff for \iul{any partial program $\contextvar$} it holds that \oul{$\sat{\link{\contextvar}{\runtimetermvar}}{\pi}$}.
 \end{definition}
 
-\MPin{
-	classes?
-	% 
-}
-Classes of properties can be used to state whether a compiler is correct, or even secure so now we turn to the definition of (secure) compilers.
-
 \subsection{Secure Compilers}\label{subsec:bg:rtp}
@@
 A {\em compiler} ($\cc{\src{L}}{\trg{L}}$) translates syntactic descriptions of programs from a {\em source} ($\src{L}$) into a {\em target} ($\trg{L}$) programming language.
-For the syntactic description of programs written in $\src{L}$, this paper writes $\src{\progvar}$, while for ones written in $\trg{L}$, this paper writes $\trg{\progvar}$.
-It is considered good if this translation is semantics-preserving, i.e., correct.
-Correct compilers may still be insecure compilers\MK{citation needed}.
-Programs translated by insecure compilers may violate security properties that were assumed to hold, given the presence of a context--level program.
-So, ideally, a compiler is secure~\cite{abate2019jour} in the sense of the following definition:
+%For the syntactic description of programs written in $\src{L}$, this paper writes $\src{\progvar}$, while for ones written in $\trg{L}$, this paper writes $\trg{\progvar}$.%\MP{not a fan of this sentence}
+%This paper writes $\src{\progvar}$ and $\trg{\progvar}$ for programs written in $\src{L}$ and $\trg{L}$, respectively.
+%      MK: I don't think we need to mention this at all
+This translation is considered {\em correct} if it is semantics--preserving.
+%That is, the compiler encodes any possible $\src{L}$--level feature in $\trg{L}$, such that the semantics are the same.
+Unfortunately, correct compilers may be insecure compilers\MK{citation needed}.
+Programs translated by insecure compilers can violate security properties that the programmer assumes to hold. %, given the presence of a context--level program.
+For a compiler to be secure, this paper makes use of the robust compilation framework~\cite{abate2019jour}, which we summarize in the following definition.
+For \bul{compilers $\cc{\src{L}}{\trg{L}}$ to robustly preserve a class of properties $\class$}, consider \rul{properties $\pi$ of that class $\class$ and programs $\src{p}$ written in $\src{L}$} such that \iul{the program $\src{\progvar}$ robustly satisfies $\pi$}.
+Then, \oul{$\cc{\src{L}}{\trg{L}}\left(\src{p}\right)$, the compiled version of $\src{\progvar}$, must robustly satisfy $\pi$}.
 
 \begin{definition}[Robust Preservation]\label{def:rtp}
-  A compiler $\cc{\src{L}}{\trg{L}}$ robustly--preserves a class of properties $\class$ ($\rtp{\cc{\src{L}}{\trg{L}}}{\class}$) iff for any property $\pi\in\class$ and component $\src{p}$ such that $\rsat{\src{\progvar}}{\pi}$, it holds that $\rsat{\cc{\src{L}}{\trg{L}}\left(\src{p}\right)}{\pi}$.
+  Compiler $\cc{\src{L}}{\trg{L}}$ robustly preserves $\class$, \bul{$\rtp{\cc{\src{L}}{\trg{L}}}{\class}$}, iff \rul{for any $\pi\in\class$ and $\src{p}$} such that \iul{$\rsat{\src{\progvar}}{\pi}$}, it holds that \oul{$\rsat{\cc{\src{L}}{\trg{L}}\left(\src{p}\right)}{\pi}$}.
 \end{definition}

### 2023-07-06 popl/main.tex
Rework up to section 5

@@
 %The component part of a whole program is the main object of interest and thus a known syntactic description.
-The context is usually arbitrary\footnote{Up to, e.g., provoking link failure.} and thus has the role of an {\em attacker} that can interact with the component by means of whatever features the programming language has.
+The context is usually arbitrary\footnote{Up to, e.g., provoking link failure by giving multiple definitiions for the same name.} and thus has the role of an {\em attacker} that can interact with the component by means of whatever features the programming language has.
 The semantics of the programming language is expected to differentiate between the component and the context.
@@
 
-\section{Example Properties and their Monitors\pages{3}}\label{sec:compprop}
-
-This section introduces a trace model (\Cref{subsec:propdefs}) that is used to define some example properties (\gls{tms}, \gls{sms}, \gls{ms}, and \gls{scct}).
+\section{Example Properties\pages{1.5}}\label{sec:compprop}
+
+This section introduces a trace model (\Cref{subsec:propdefs}) and defines some example properties (\gls{tms}, \gls{sms}, \gls{ms}, and \gls{scct}).
 These properties are of practical importance and also of interest in the case study (\Cref{sec:casestud:defs,sec:casestud:rtp}) this paper presents.
-The study uses monitors, which are able to check wether a trace admits a property or not.
-The use of monitors is because they have some structure useful for inductive proofs, whereas traces lack this structure, thus making proofs difficult.
-So, this section also presents monitors for each of the defined properties (\Cref{subsec:monitors}) and introduces monitor-based satisfaction.
-Finally, for each property, it is shown that monitor-based satisfaction implies property satisfaction.
-
 \subsection{Specification Trace Model}\label{subsec:propdefs}
@@
 
-The specification trace model has events that carry different meta-information, such as security- and control--tags.
-Security tags describe wether an event contains sensitive information ($\lock$) or not ($\unlock$).
-Control tags state wether the context ($\ctx$) or the component ($\comp$) are responsible for emitting the event.
+The specification trace model defines events as either the empty event ($\emptyevent$), a crash ($\lightning$), or as tuples consisting of a pre--event, a control--tag, and a security--tag.
+The purpose of the model is to define properties, such as a stricter variant of cryptographic constant time.
+To this end, security--tags indicate wether an event contains sensitive information ($\lock$) or not ($\unlock$), while control--tags state wether the context ($\ctx$) or the component ($\comp$) are responsible for emitting the event.
+The latter is necessary to be able to ignore actions done by a spurious context that, e.g., immediately deallocates a memory location twice, thus violating \gls{tms}.
 %Even though this could be used for blame~\cite{patrignani2023blame}, this is not what this paper is concerned about.
-Now, pre--events describe the actual kind of event that happened.
-One such kind is the allocation event ($\ev{Alloc\ \loc\ n}$) that happens whenever a program claims $n$ cells of memory and stores them at address $\loc$.
+Lastly, pre--events describe the actual kind of event that happened.
+One such kind is the allocation event ($\ev{Alloc\ \loc\ n}$) that fires whenever a program claims $n$ cells of memory and stores them at address $\loc$.
 Likewise, deallocation ($\ev{Dealloc\ \loc}$) announces that the object at location $\loc$ is freed.
-These two events alone allow already a partial description of \gls{tms} by requiring that e.g. there is only one deallocation event that carries a location $\loc$.
+These two events alone allow already a partial description of \gls{tms} by requiring that, e.g., there is only one deallocation event that carries a location $\loc$.
 To be able to express \gls{sms}, there is also an event to describe reads from and writes to memory ($\ev{Use\ \loc\ n}$).
-For cryptographic code, there is a general guideline that secrets must not be visible on a trace.
+Finally, for cryptographic code, there is a general guideline that secrets must not be visible on a trace.
 Moreover, an instruction whose timing is data--dependent must not have a secret as an operand.
-Typical operations with data--dependent timing are branches and certain binary operations, such as division.
+Typical operations with data--dependent timing are branches and certain binary operations, such as division\footnote{This is highly architecture dependent, but division is an operation that serves as a classic example for a data--dependent timing instruction.}.
 Both operations are also modelled in the specification trace model ($\ev{Branch\ n}$ and $\ev{Binop\ n}$).
-Lastly, events are either the empty event ($\emptyevent$), a program crash ($\lightning$), or a tuple consisting of a pre-event, a control--tag, and a security--tag.
-
-\gls{tms} requires that pointers are freed exactly once and that no use happens after a free.
+
+\subsection{Temporal Memory Safety}
+
+\gls{tms} is a safety property that describes that an unallocated object must not be used in any way.
+Moreover, the property requires that all allocated objects must be deallocated at some point.
 
 \begin{definition}[\glsfirst{tms}]\label{def:trace:tmsdef}
-  A trace $\trace$ is temporal memory safe ($\trace\in\tmssafe$) iff $\ev{Alloc\ \loc\ n;\sandboxtag;\securitytag}\le_{\trace}\ev{Dealloc\ \loc;\sandboxtag';\securitytag'}$, $\ev{Use\ \loc\ n;\sandboxtag;\securitytag}\not\le_{\trace}\ev{Alloc\ \loc\ m;\sandboxtag';\securitytag'}$ and $\ev{Dealloc\ \loc;\sandboxtag;\securitytag}\not\le_{\trace}\ev{Use\ \loc\ n;\sandboxtag';\securitytag'}$.
+  $$
+  \tmssafe:=\left\{\trace \left| \begin{array}{rcl}
+                              \ev{Alloc\ \loc\ n;\sandboxtag;\securitytag}&\le_{\trace}&\ev{Dealloc\ \loc;\sandboxtag;\securitytag'} \\
+                              \ev{Use\ \loc\ n;\sandboxtag;\securitytag}&\le_{\trace}&\ev{Dealloc\ \loc;\sandboxtag;\securitytag'} \\
+                              \text{if }\ev{Alloc\ \loc\ n;\sandboxtag;\securitytag}\text{ in }\trace&\text{then}&\ev{Dealloc\ \loc;\sandboxtag;\securitytag'}\text{ in }\trace \\
+                              \text{at most one }\ev{Dealloc\ \loc;\sandboxtag;\securitytag}&\text{in}&\trace \\
+                              \text{at most one }\ev{Alloc\ \loc\ n;\sandboxtag;\securitytag}&\text{in}&\trace \\
+                            \end{array}\right.\right\}
+  $$
 \end{definition}
-
-\gls{sms} disregards out of bounds accesses.
+Hereby, the notation $\event[_{1}]\le_{\trace}\event[_{2}]$ means that if $\event[_{1}]$ is in $\trace$ and if $\event[_{2}]$ is in $\trace$, then $\event[_{1}]$ appears before $\event[_{2}]$.
+
+\subsection{Spatial Memory Safety}
+
+\gls{sms} disregards out of bounds accesses:
 
 \begin{definition}[\glsfirst{sms}]\label{def:trace:smsdef}
+  $
+  \smssafe:=\left\{\trace \left|\begin{array}{rcl}
+                                  \text{If }\ev{Alloc\ \loc\ n;\sandboxtag;\securitytag}\le_{\trace}\ev{Use\ \loc\ m;\sandboxtag;\securitytag'}, \text{ then }m<n
+                                \end{array}\right.\right\}
+  $
   A trace $\trace$ is spatial memory safe ($\trace\in\smssafe$) iff given $\ev{Alloc\ \loc\ m;\sandboxtag;\securitytag}\le_{\trace}\ev{Use\ \loc\ n;\sandboxtag';\securitytag'}$, then $n<m$.
@@
 
+\subsection{Memory Safety}
+
 Full \gls{ms} is then described as the conjunction of \Cref{def:trace:tmsdef,def:trace:smsdef}.
+Note, however, that this definition does not entail memory--safety issues introduced by side--channels, such as speculation.
 
 \begin{definition}[\glsfirst{ms}]\label{def:trace:msdef}
+  $
+  \mssafe:=\left\{\trace \left|\begin{array}{rcl}
+                                 \trace\in\tmssafe & \text{ and } & \trace\in\smssafe
+                               \end{array}\right.\right\}
+  $
   A trace $\trace$ is memory safe ($\trace\in\mssafe$) iff $\trace\in\tmssafe$ and $\trace\in\smssafe$.
@@
 
-A variant of the \gls{cct} security property is defined.
-Generally, one cannot define \gls{cct} just as trace-property without, e.g., performing a self-composition of a program.
-That is why this paper adopts a stricter variant of \gls{cct}, called \gls{scct}, that enforces common policies to ensure code is \gls{cct}.
-For example, one must not branch on secrets or perform binary operations, such as division, on secrets whose timing is data-dependent.
-Even stronger, \gls{scct} disregards any secrets leaked on a trace.
-To define the property, a function on traces is needed that strips all events with a $\lock$-tag: $\downarrow_{\lock}(\cdot)$.
-
-\begin{definition}[\glsfirst{scct}]
-  A trace $\trace$ is strictly cryptographic constant time ($\trace\in\scctsafe$) iff $\trace=\downarrow_{\lock}\left(\trace\right)$.
+\subsection{Strict Cryptographic Constant Time}
+
+\gls{cct} is a hyperproperty and, thus, not as easy to check with just monitors. %Beutner et al
+Hence, the following defines \gls{scct}, a stricter variant that enforces the policy that no secret should appear on a trace.
+
+\begin{definition}[\glsfirst{scct}]\label{def:trace:scctdef}
+  $
+  \scctsafe:=\left\{\trace \left|\begin{array}{rcl}
+                                   \trace=\hole{\cdot} & \text{ or } & \trace=\event[_{b}];\sandboxtag;\unlock\cdot\trace['] \wedge \trace[']\in\scctsafe
+                                 \end{array}\right.\right\}
+  $
 \end{definition}
 
-\subsection{Monitors and Trace Agreements}\label{subsec:monitors}
-
-This subsection presents monitors for properties \gls{tms}, \gls{sms}, \gls{ms}, and \gls{scct}.
-Each monitor $\monitor$ has its own accompanying trace model that abstracts over the most general trace model.
-%This is because monitors should only be concerned with property-relevant information.
-Traces of the accompanying trace model are written as $\abstrace$ and $\monitorcheck{\monitor}{\monitor[']}{\abstrace}$ is the notation for the semantics of a monitor.
-The empty monitor is denoted as $\emptyset$.
-For each model of a monitor, an agreement relation $\traceagree{\pi}{\trace}{\abstrace}$ is defined that relates monitor specific traces with most general traces.
-The notation $\traceagree[]{\pi}{\event}{\absevent}$ is used to relate just events and the star version of it is defined similarily to \Cref{fig:starstep:and:nstep}.
-With this, first, monitor satisfaction is defined and then the individual monitors.
-Finally, for every monitor, it is shown that monitor satisfaction implies trace property satisfaction.
+\subsection{Memory Safe, Strict Cryptographic Constant Time}
+
+The combination of \gls{ms} and \gls{scct} is the intersection of these properties, \gls{msscct}.
+It entails all program behaviours that are both memory safe and strictly cryptographic constant time.
+
+\begin{definition}[\glsfirst{ms} and \gls{scct}]\label{def:trace:msscctdef}
+  $
+  \msscctsafe:=\left\{\trace \left| \begin{array}{rcl}
+                                      \trace\in\mssafe & \text{ and } & \trace\in\scctsafe
+                                    \end{array}\right.\right\}
+  $
+\end{definition}
+
+%\section{Monitors and Trace Agreements\pages{1.5}}\label{sec:monitors}
+\section{Monitors\pages{1.5}}\label{sec:monitors}
+This section introduces monitors for each of the properties presented earlier (\Cref{subsec:propdefs}).
+The use of monitors is because they have some structure useful for inductive proofs, whereas traces lack this structure, thus making proofs difficult.
+To lessen the burden when proving, each monitor makes use of a custom trace model that contains only the relevant information related to the property the monitor checks.
+To project specification traces $\trace$ onto these monitor--level traces $\abstrace$, each property has an associated event agreement relation $\traceagree[]{\pi}{\trace}{\abstrace}$.
+This relation performs the actual work of projecting the events, while the stutter--equivalence $\traceagree{\pi}{\trace}{\abstrace}$ remains the same for all properties $\pi$, yet is indexed by $\pi$ to emphasize the kind of event agreement.
+With agreements, this section starts to define monitor satisfaction for traces and states that monitor satisfaction implies property satisfaction.
+To this end, monitor satisfaction is defined as follows.
+\bul{A specification trace $\trace$ monitor--satisfies property $\pi$} iff there exists a (final) monitor state $\monitor$ and an abstract trace $\abstrace$ such that the \iul{specification trace $\trace$ agrees with abstract trace $\abstrace$} and \oul{the initial monitor\footnote{In this paper, for all monitors, the initial monitor state is denoted as $\emptyset$.} can step to the (final) monitor state $\monitor$ with abstract trace $\abstrace$}.
 
 \begin{definition}[Monitor Satisfaction]\label{def:monsat}
-  A given trace $\trace$ monitor-satisfies a property $\pi$ ($\monsafe{\trace}{\pi}$) iff %
-  there exists $\abstrace,\monitor$ such that $\traceagree{\pi}{\trace}{\abstrace}$ and $\monitorcheck{\emptyset}{\monitor}{\abstrace}$.
-  \Coqed
+  \bul{$\monsafe{\trace}{\pi}$} iff $\exists \abstrace\ \monitor$ such that \iul{$\traceagree{\pi}{\trace}{\abstrace}$} and \oul{$\monitorcheck{\emptyset}{\monitor}{\abstrace}$}.
 \end{definition}
 
-\paragraph{Monitor for \glsfirst{tms}}
+\subsection{Monitor for \glsfirst{tms}}
 \begin{gather*}
   \begin{aligned}
-    \mi{(Abstract\ Store)}~\tmsmonitor\bnfdef&\ \left\{\text{alloced}:L,\text{freed}:L\right\} \hspace{0.33cm}%
-    \emptyset:=\ \left\{\text{alloced}:\emptyset,\text{freed}:\emptyset\right\}\\
-    \mi{(Abstract\ Events)}~\absevent\bnfdef&\ \absev{\emptyevent} \mid \absev{Alloc\ }\loc \mid \absev{Dealloc\ }\loc \mid \absev{Use\ }\loc \mid \absterm\\
+    \mi{(Abstract\ Store)}~\tmsmonitor\bnfdef&\ \left\{\text{allocated}:L\times\sandboxtag,\text{freed}:L\times\sandboxtag\right\} \hspace{0.33cm}%
+    \emptyset:=\ \left\{\text{allocated}:\emptyset,\text{freed}:\emptyset\right\}\\
+    \mi{(Abstract\ Events)}~\absevent\bnfdef&\ \absev{\emptyevent} \mid \absev{Alloc\ }\loc\ \sandboxtag \mid \absev{Dealloc\ }\loc\ \sandboxtag \mid \absev{Use\ }\loc\ \sandboxtag \mid \absterm\\
   \end{aligned}
@@
 \begin{center}
-  \judgbox{\monitorcheck[]{\tmsmonitor}{\tmsmonitor[']}{\absevent}}{,,Monitor $\tmsmonitor$ steps to $\tmsmonitor[']$ given event $\absevent$.''}$\;$\\
+  \judgbox{\monitorcheck[]{\tmsmonitor}{\tmsmonitor[']}{\absevent}}{,,Monitor $\tmsmonitor$ does one step to $\tmsmonitor[']$ given event $\absevent$.''}$\;$\\
   %
@@
   }{
-    \monitorcheck[]{\tmsmonitor}{\emptyset}{\absterm}
+    \monitorcheck[]{\tmsmonitor}{\tmsmonitor}{\absterm}
   }{tms-abort}
@@
   \typerule{tms-use}{
-    \loc\in\tmsmonitor.\text{alloced} &
-    \loc\notin\tmsmonitor.\text{freed}
-  }{
-    \monitorcheck[]{\tmsmonitor}{\tmsmonitor}{\absev{Use\ }\loc}
+    (\loc;\sandboxtag)\in\tmsmonitor.\text{allocated} &
+    (\loc;\sandboxtag)\notin\tmsmonitor.\text{freed}
+  }{
+    \monitorcheck[]{\tmsmonitor}{\tmsmonitor}{\absev{Use\ }\loc\ \sandboxtag}
   }{tms-use}
@@
   \typerule{tms-alloc}{
-    \loc\notin\tmsmonitor.\text{alloced} &
-    \loc\notin\tmsmonitor.\text{freed} \\
-    \tmsmonitor[']=\left\{\text{alloced}: \tmsmonitor.\text{alloced}\cup\left\{\loc\right\},\text{freed}: \tmsmonitor.\text{freed}\right\}
-  }{
-    \monitorcheck[]{\tmsmonitor}{\tmsmonitor[']}{\absev{Alloc\ }\loc}
+    (\loc;\sandboxtag)\notin\tmsmonitor.\text{allocated} &
+    (\loc;\sandboxtag)\notin\tmsmonitor.\text{freed} \\
+    \tmsmonitor[']=\left\{\text{allocated}: \tmsmonitor.\text{allocated}\cup\left\{(\loc;\sandboxtag)\right\},\text{freed}: \tmsmonitor.\text{freed}\right\}
+  }{
+    \monitorcheck[]{\tmsmonitor}{\tmsmonitor[']}{\absev{Alloc\ }\loc\ \sandboxtag}
   }{tms-alloc}
@@
   \typerule{tms-dealloc}{
-    \loc\in\tmsmonitor.\text{alloced} &
-    \loc\notin\tmsmonitor.\text{freed} \\
-    \tmsmonitor[']=\left\{\text{alloced}: \tmsmonitor.\text{alloced}\setminus\left\{\loc\right\},\text{freed}: \tmsmonitor.\text{freed}\cup\left\{\loc\right\}\right\}
-  }{
-    \monitorcheck[]{\tmsmonitor}{\tmsmonitor[']}{\absev{Dealloc\ }\loc}
+    (\loc;\sandboxtag)\in\tmsmonitor.\text{allocated} &
+    (\loc;\sandboxtag)\notin\tmsmonitor.\text{freed} \\
+    \tmsmonitor[']=\left\{\text{allocated}: \tmsmonitor.\text{allocated}\setminus\left\{(\loc;\sandboxtag)\right\},\text{freed}: \tmsmonitor.\text{freed}\cup\left\{(\loc;\sandboxtag)\right\}\right\}
+  }{
+    \monitorcheck[]{\tmsmonitor}{\tmsmonitor[']}{\absev{Dealloc\ }\loc\ \sandboxtag}
   }{tms-dealloc}
 \end{center}
+
+The state of the monitor is a record with two sets keeping track of allocated and deallocated locations.
+%The monitor runs on a modified trace semantics with more abstract $\absev{Alloc\ }\loc$,$\absev{Dealloc\ }\loc$,$\absev{Use\ }\loc$, and $\absterm$ events, which represent {\em allocation}, {\em deallocation}, and {\em use} of a location $\loc$ as well as abnormal program termination.
+\Cref{tr:tms-use} simply requires that a location is (i) allocated and (ii) not freed.
+\Cref{tr:tms-alloc,tr:tms-dealloc} both require a location to not be freed already and extend the monitor state accordingly.
+This restriction effectively disallows reallocation to reassign the same location to an object.
+However, the definition can easily be adapted by, e.g., attaching a natural number serving as a counter.
+In the reflexive case of $\monitorcheck{\tmsmonitor}{\tmsmonitor}{\absevent}$, the monitor $\tmsmonitor$ must not contain allocated locations, i.e., $\tmsmonitor.allocated=\emptyset$.
+This ensures that programs do not forget to free memory.
+
 \begin{center}
@@
   }{
-    \tmstraceagree[]{\ev{Alloc}\ \loc\ n;\sandboxtag;\securitytag}{\absev{Alloc\ }\loc}
+    \tmstraceagree[]{\ev{Alloc}\ \loc\ n;\sandboxtag;\securitytag}{\absev{Alloc\ }\loc\ \sandboxtag}
   }{tms-alloc-authentic}
   %
-  \typerule{tms-dealloc-authentic}{
-  }{
-    \tmstraceagree[]{\ev{Dealloc}\ \loc;\sandboxtag;\securitytag}{\absev{Dealloc\ }\loc}
-  }{tms-dealloc-authentic}
-  %
-  \typerule{tms-use-authentic}{
-  }{
-    \tmstraceagree[]{\ev{Use}\ \loc\ n;\sandboxtag;\securitytag}{\absev{Use\ }\loc}
-  }{tms-use-authentic}
+% \typerule{tms-dealloc-authentic}{
+% }{
+%   \tmstraceagree[]{\ev{Dealloc}\ \loc;\sandboxtag;\securitytag}{\absev{Dealloc\ }\loc}
+% }{tms-dealloc-authentic}
+  %
+% \typerule{tms-use-authentic}{
+% }{
+%   \tmstraceagree[]{\ev{Use}\ \loc\ n;\sandboxtag;\securitytag}{\absev{Use\ }\loc}
+% }{tms-use-authentic}
   %
@@
   %
-  \typerule{tms-binop-authentic}{
-  }{
-    \tmstraceagree[]{\ev{Binop\ }n}{\absev{\emptyevent}}
-  }{tms-binop-authentic}
-  %
-  \typerule{tms-empty-authentic}{
-  }{
-    \tmstraceagree[]{\emptyevent}{\absev{\emptyevent}}
-  }{tms-empty-authentic}
-  %
+% \typerule{tms-binop-authentic}{
+% }{
+%   \tmstraceagree[]{\ev{Binop\ }n}{\absev{\emptyevent}}
+% }{tms-binop-authentic}
+% %
+% \typerule{tms-empty-authentic}{
+% }{
+%   \tmstraceagree[]{\emptyevent}{\absev{\emptyevent}}
+% }{tms-empty-authentic}
+% %
   \typerule{tms-abort-authentic}{
@@
 
-The state of the monitor is a record containing two sets that keep track of allocated and freed locations.
-The monitor runs on a modified trace semantics with more abstract $\absev{Alloc\ }\loc$,$\absev{Dealloc\ }\loc$,$\absev{Use\ }\loc$, and $\absterm$ events, which represent {\em allocation}, {\em deallocation}, and {\em use} of a location $\loc$ as well as abnormal program termination.
-\Cref{tr:tms-use} simply requires that a location is (i) allocated and (ii) not freed.
-\Cref{tr:tms-alloc,tr:tms-dealloc} both require a location to not be freed already and depending on the witnessed action, extend the internal state of the monitor.
-The trace agreement is entirely straightforward.
+The trace agreement is entirely straightforward, so only allocation, branch, and crash are shown.
 
 \begin{lemma}[Traces with Monitor Satisfaction are $\tmssafe$]\label{lem:mon:tmsafe}
-  If $\montmssafe{\trace}$, then $\trace\in\tmssafe$.\Coqed
+  If $\montmssafe{\trace}$, then $\trace\in\tmssafe$. %\Coqed
 \end{lemma}
 
-\paragraph{Monitor for \glsfirst{sms}}
+\subsection{Monitor for \glsfirst{sms}}
 \begin{gather*}
   \begin{aligned}
-    \mi{(Abstract\ Store)}~\smsmonitor:=\ L\times\mb{N}\hspace{0.33cm}&
-    \mi{(Abstract\ Events)}~\absevent\bnfdef\ \absev{\varepsilon} \mid \absev{Alloc\ }\loc\ n \mid \absev{Dealloc\ }\loc \mid \absev{Use\ }\loc\ n \\
+    \mi{(Abstract\ Store)}~\smsmonitor:=\ L\times\sandboxtag\times\mb{N}\hspace{0.33cm}&
+    \mi{(Abstract\ Events)}~\absevent\bnfdef\ \absev{\varepsilon} \mid \absev{Alloc\ }\loc\ \sandboxtag\ n \mid \absev{Use\ }\loc\ \sandboxtag\ n \\
   \end{aligned}
 \end{gather*}
-\MKin{do we really need $\absev{Dealloc}$?}
-\begin{center}
-  \judgbox{\monitorcheck[]{\smsmonitor}{\smsmonitor[']}{\absevent}}{,,Monitor $\smsmonitor$ steps to $\smsmonitor[']$ given event $\absevent$.''}$\;$\\
-  %
-  \typerule{sms-uninteresting}{
-  }{
-    \monitorcheck[]{\smsmonitor}{\smsmonitor}{\absev{\emptyevent}}
-  }{sms-uninteresting}
-  %
-  \typerule{sms-abort}{
-  }{
-    \monitorcheck[]{\smsmonitor}{\emptyset}{\absterm}
-  }{sms-abort}
+\begin{center}
+  \judgbox{\monitorcheck[]{\smsmonitor}{\smsmonitor[']}{\absevent}}{,,Monitor $\smsmonitor$ does one step to $\smsmonitor[']$ given event $\absevent$.''}$\;$\\
+  %
+% \typerule{sms-uninteresting}{
+% }{
+%   \monitorcheck[]{\smsmonitor}{\smsmonitor}{\absev{\emptyevent}}
+% }{sms-uninteresting}
+% %
+% \typerule{sms-abort}{
+% }{
+%   \monitorcheck[]{\smsmonitor}{\emptyset}{\absterm}
+% }{sms-abort}
   %
   \typerule{sms-use}{
-    (\loc,m)\in\smsmonitor &
+    (\loc;\sandboxtag;m)\in\smsmonitor &
     n<m
   }{
-    \monitorcheck[]{\smsmonitor}{\smsmonitor}{\absev{Use\ }\loc\ n}
+    \monitorcheck[]{\smsmonitor}{\smsmonitor}{\absev{Use\ }\loc\ \sandboxtag\ n}
   }{sms-use}
@@
   \typerule{sms-alloc}{
-    \loc\notin\operatorname{dom}\smsmonitor &
-    \smsmonitor[']=\smsmonitor\cup\left\{(\loc,n)\right\}
-  }{
-    \monitorcheck[]{\smsmonitor}{\smsmonitor[']}{\absev{Alloc\ }\loc\ n}
+    (\loc;\sandboxtag;m)\notin\smsmonitor
+  }{
+    \monitorcheck[]{\smsmonitor}{\smsmonitor\cup\left\{(\loc;\sandboxtag;n)\right\}}{\absev{Alloc\ }\loc\ \sandboxtag\ n}
   }{sms-alloc}
-  %
-  \typerule{sms-dealloc}{
-    (\loc,n)\in\smsmonitor &
-    \smsmonitor[']=\smsmonitor\setminus\left\{(\loc,n)\right\}
-  }{
-    \monitorcheck[]{\smsmonitor}{\smsmonitor[']}{\absev{Dealloc\ }\loc}
-  }{sms-dealloc}
-\end{center}
-\begin{center}
-  \judgbox{\smstraceagree[]{\event}{\absevent}}{,,Abstract event $\absevent$ is equivalent to $\event$ with respect to \gls{sms}.''}$\;$\\
-  %
-  \typerule{sms-alloc-authentic}{
-  }{
-    \smstraceagree[]{\ev{Alloc}\ \loc\ n;\sandboxtag;\securitytag}{\absev{Alloc\ }\loc\ n}
-  }{sms-alloc-authentic}
-  %
-  \typerule{sms-dealloc-authentic}{
-  }{
-    \smstraceagree[]{\ev{Dealloc}\ \loc;\sandboxtag;\securitytag}{\absev{Dealloc\ }\loc}
-  }{sms-dealloc-authentic}
-  %
-  \typerule{sms-use-authentic}{
-  }{
-    \smstraceagree[]{\ev{Use}\ \loc\ n;\sandboxtag;\securitytag}{\absev{Use\ }\loc\ n}
-  }{sms-use-authentic}
-  %
-  \typerule{sms-branch-authentic}{
-  }{
-    \smstraceagree[]{\ev{Branch\ }n}{\absev{\emptyevent}}
-  }{sms-branch-authentic}
-  %
-  \typerule{sms-binop-authentic}{
-  }{
-    \smstraceagree[]{\ev{Binop\ }n}{\absev{\emptyevent}}
-  }{sms-binop-authentic}
-  %
-  \typerule{sms-empty-authentic}{
-  }{
-    \smstraceagree[]{\emptyevent}{\absev{\emptyevent}}
-  }{sms-empty-authentic}
-  %
-  \typerule{sms-abort-authentic}{
-  }{
-    \smstraceagree[]{\lightning}{\absterm}
-  }{sms-abort-authentic}
-\end{center}
-
-The state of the monitor for \gls{sms} is a set containing pairs of locations and the allocation size of the objects the locations point to.
+\end{center}
+The state of the monitor for \gls{sms} is a set containing tuples of locations, control--tags, and the allocation size.
 In comparison to the trace model of the \gls{tms} monitor, the trace model here is extended by sizing and positional information.
-\Cref{tr:sms-use} performs a bounds check and \Cref{tr:sms-alloc,tr:sms-dealloc} add or remove locations and the size of an allocation.
-The trace agreement relation is, again, entirely straightforward.
-
-\begin{lemma}[Monitor Traces are $\smssafe$]\label{lem:mon:smsafe}
+\Cref{tr:sms-use} performs a bounds check and \Cref{tr:sms-alloc} adds bounds information to the state of the monitor.
+The trace agreement is entirely straightforward and similar to the one fo \gls{tms}.
+% \begin{center}
+%   \judgbox{\smstraceagree[]{\event}{\absevent}}{,,Abstract event $\absevent$ is equivalent to $\event$ with respect to \gls{sms}.''}$\;$\\
+%   %
+%   \typerule{sms-alloc-authentic}{
+%   }{
+%     \smstraceagree[]{\ev{Alloc}\ \loc\ n;\sandboxtag;\securitytag}{\absev{Alloc\ }\loc\ n}
+%   }{sms-alloc-authentic}
+%   %
+%   \typerule{sms-dealloc-authentic}{
+%   }{
+%     \smstraceagree[]{\ev{Dealloc}\ \loc;\sandboxtag;\securitytag}{\absev{Dealloc\ }\loc}
+%   }{sms-dealloc-authentic}
+%   %
+%   \typerule{sms-use-authentic}{
+%   }{
+%     \smstraceagree[]{\ev{Use}\ \loc\ n;\sandboxtag;\securitytag}{\absev{Use\ }\loc\ n}
+%   }{sms-use-authentic}
+%   %
+%   \typerule{sms-branch-authentic}{
+%   }{
+%     \smstraceagree[]{\ev{Branch\ }n}{\absev{\emptyevent}}
+%   }{sms-branch-authentic}
+%   %
+%   \typerule{sms-binop-authentic}{
+%   }{
+%     \smstraceagree[]{\ev{Binop\ }n}{\absev{\emptyevent}}
+%   }{sms-binop-authentic}
+%   %
+%   \typerule{sms-empty-authentic}{
+%   }{
+%     \smstraceagree[]{\emptyevent}{\absev{\emptyevent}}
+%   }{sms-empty-authentic}
+%   %
+%   \typerule{sms-abort-authentic}{
+%   }{
+%     \smstraceagree[]{\lightning}{\absterm}
+%   }{sms-abort-authentic}
+% \end{center}
+
+\begin{lemma}[Traces with Monitor Satisfaction are $\smssafe$]\label{lem:mon:smsafe}
   If $\monsmssafe{\trace}$, then $\trace\in\smssafe$.\Coqed
@@
 
-\paragraph{Comining Monitors: \glsfirst{ms}}
-
-\begin{center}
-  \judgbox{\monitorcheck[]{\msmonitor}{\msmonitor[']}{\absevent}}{,,Monitor $\msmonitor$ steps to $\msmonitor[']$ given event $\absevent$.''}$\;$\\
+\subsection{Combining \glsfirst{tms} and \glsfirst{sms} Monitors to obtain \glsfirst{ms}}\label{subsubsec:ms}
+
+\begin{center}
+  \judgbox{\monitorcheck[]{\msmonitor}{\msmonitor[']}{\absevent}}{,,Monitor $\msmonitor$ does one step to $\msmonitor[']$ given event $\absevent$.''}$\;$\\
   \typerule{ms-step}{
@@
 \end{center}
-\begin{center}
-  \judgbox{\mstraceagree{\event}{\absevent}}{,,Abstract event $\absevent$ is equivalent to $\event$ with respect to \gls{ms}.''}$\;$\\
-  %
-  \typerule{ms-authentic}{
-    \tmstraceagree[]{\event}{\absevent_{\tmssafe}} &
-    \smstraceagree[]{\event}{\absevent_{\smssafe}}
-  }{
-    \mstraceagree[]{\event}{(\absevent_{\tmssafe},\absevent_{\smssafe})}
-  }{ms-authentic}
-\end{center}
-
-To obtain a monitor for \gls{ms}, the monitors for \gls{tms} and \gls{sms} are simply combined in lock-step.
-
-\begin{lemma}[Monitor Traces are $\mssafe$]\label{lem:mon:msafe}
+The combined monitor runs the one for \gls{tms} and the one for \gls{sms} in a lockstep.
+The trace agreement similarily just relates a specification event with an abstract \gls{tms}-event $\abstrace[_{\tmssafe}]$ and with an abstract \gls{sms}-event $\abstrace[_{\smssafe}]$.
+% \begin{center}
+%   \judgbox{\mstraceagree{\event}{\absevent}}{,,Abstract event $\absevent$ is equivalent to $\event$ with respect to \gls{ms}.''}$\;$\\
+%   %
+%   \typerule{ms-authentic}{
+%     \tmstraceagree[]{\event}{\absevent_{\tmssafe}} &
+%     \smstraceagree[]{\event}{\absevent_{\smssafe}}
+%   }{
+%     \mstraceagree[]{\event}{(\absevent_{\tmssafe},\absevent_{\smssafe})}
+%   }{ms-authentic}
+% \end{center}
+
+\begin{lemma}[Traces with Monitor Satisfaction are $\mssafe$]\label{lem:mon:msafe}
   If $\monmssafe{\trace}$, then $\trace\in\mssafe$.\Coqed
@@
 
-\paragraph{Monitor for \glsfirst{scct}}
+\subsection{Monitor for \glsfirst{scct}}
 \begin{gather*}
   \begin{aligned}
-    \mi{(Abstract\ Store)}~\scctmonitor&:=\ \emptyset \\
-    \mi{(Abstract\ Events)}~\absevent&:=\ \absev{\varepsilon} \mid \absterm \mid \absev{Any} \\
+    \mi{(Abstract\ Store)}~\scctmonitor&:=\ \emptyset \hspace{0.5cm}
+    \mi{(Abstract\ Events)}~\absevent:=\ \absev{\varepsilon} \mid \absterm \mid \absev{Any} \\
   \end{aligned}
@@
 \begin{center}
-  \judgbox{\monitorcheck[]{\scctmonitor}{\scctmonitor[']}{\absevent}}{,,Monitor $\scctmonitor$ steps to $\scctmonitor[']$ given event $\absevent$.''}$\;$\\
+  \judgbox{\monitorcheck[]{\scctmonitor}{\scctmonitor[']}{\absevent}}{,,Monitor $\scctmonitor$ does one step to $\scctmonitor[']$ given event $\absevent$.''}$\;$\\
   %
@@
 \end{center}
+The monitor state for the \gls{scct} monitor is completely empty, since it does not need to keep track of information.
+As soon as any event is hit, the execution gets stuck, since any event is considered confidental from the perspective of this monitor.
 \begin{center}
@@
 \end{center}
-
-The \gls{scct} monitor simply disregards all private data occuring on a trace.
+Accordingly, the event agreement simply disregards all events that involved public data ($\unlock$) while mapping any other event that does involve private data ($\lock$) to the abstract $\absev{Any}$ event.
 
 \begin{lemma}[Monitor Traces are $\scctsafe$]\label{lem:mon:scctsafe}
-  If $\monscctsafe{\trace}$, then $\trace\in\scctsafe$.\Coqed
+  If $\monscctsafe{\trace}$, then $\trace\in\scctsafe$. %\Coqed
 \end{lemma}
 
-\MKin{combined version of all monitors? maybe just as lemma}
-
-\section{Composing Secure Compilers\pages{1\sfrac{1}{2}}}\label{sec:compcomp}
-
-This section studies different kinds of secure compiler compositions and their corresponding proofs.
+\subsection{Combining \gls{ms} and \gls{scct} Monitors to obtain \gls{msscct}}
+
+The combination of monitors for \gls{ms} and \gls{scct} yields one for \gls{msscct}.
+The construction is entirely similar to the one for \gls{ms} (\Cref{subsubsec:ms}).
+
+\begin{lemma}[Traces with Monitor Satisfaction are $\msscctsafe$]\label{lem:mon:msscctsafe}
+  If $\monmsscctsafe{\trace}$, then $\trace\in\msscctsafe$. %\Coqed
+\end{lemma}
+
+\section{Composing Secure Compilers\pages{1}}\label{sec:compcomp}
+
+This section presents the key results of this paper that have been presented in prior work~\cite{kruse2022csc}.
+The main results are the fact that secure compilers in the robust preservation framework~\cite{abate2019jour} compose.
+...
 Most notably, it looks at the sequential composition of secure compilers and shows a theorem that allows for modular proof engineering: compilers need not to robustly preserve the same property in order to be securely composed.
@@
 \begin{definition}[Sequential Composition]\label{def:seqcomp}
-  Let $\cc{\src{L}}{\trg{L}}$, $\cc{\trg{L}}{\obj{L}}$ be compilers.
-  Let $\src{p}$ be any $\src{L}$ program, $\trg{p}=\cc{\src{L}}{\trg{L}}\left(\src{p}\right)$, and $\obj{p}=\cc{\trg{L}}{\obj{L}}\left(\trg{p}\right)$.
-  The sequential composition is $\left(\cc{\src{L}}{\trg{L}}\circ\cc{\trg{L}}{\obj{L}}\right)\left(\src{p}\right)=\obj{p}$.
+  Let $\cc{\src{L}}{\trg{L}}$, $\cc{\trg{L}}{\obj{L}}$ be compilers from languages $\src{L}$ to $\trg{L}$ and $\trg{L}$ to $\obj{L}$, respectively.
+  Their sequential composition is $\left(\cc{\src{L}}{\trg{L}}\circ\cc{\trg{L}}{\obj{L}}\right)\left(\src{p}\right)=\obj{p}$.
 \end{definition}
@@
   Given classes of properties $\class[_{1}],\class[_{2}]$, let $\rtp{\cc[_{1}]{\trg{L}}{\trg{L}}}{\class[_{1}]}$ and $\rtp{\cc[_{2}]{\trg{L}}{\trg{L}}}{\class[_{2}]}$.
-  Then, $\rtp{\cc[_{1}]{\trg{L}}{\trg{L}}}{\class[_{1}]\cap\class[_{2}]}$ and $\rtp{\cc[_{2}]{\trg{L}}{\trg{V}}}{\class[_{2}]\cap\class[_{1}]}$. \Coqed
+  Then, $\rtp{\cc[_{1}]{\trg{L}}{\trg{L}}\circ\cc[_{2}]{\trg{L}}{\trg{L}}}{\class[_{1}]\cap\class[_{2}]}$ and $\rtp{\cc[_{2}]{\trg{L}}{\trg{L}}\circ\cc[_{1}]{\trg{L}}{\trg{L}}}{\class[_{2}]\cap\class[_{1}]}$.
+\Coqed
 \end{corollary}
@@
 
-\subsubsection{All Well-Typed $\Ltms$ Programs are Temporal Memory Safe}
+\subsection{All Well-Typed $\Ltms$ Programs are Temporal Memory Safe}
 

### 2023-07-06 popl/main.tex
Text up to section 5 (excluding)

@@
 
-This section presents the key results of this paper that have been presented in prior work~\cite{kruse2022csc}.
-The main results are the fact that secure compilers in the robust preservation framework~\cite{abate2019jour} compose.
-...
-Most notably, it looks at the sequential composition of secure compilers and shows a theorem that allows for modular proof engineering: compilers need not to robustly preserve the same property in order to be securely composed.
-Moreover, it demonstrates that the order does not matter for composing compilers with the same source and target language.
-Throughout this section, let $\src{L},\trg{L},$ and $\obj{L}$ be arbitrary programming lanugages.
-
-\begin{definition}[Sequential Composition]\label{def:seqcomp}
-  Let $\cc{\src{L}}{\trg{L}}$, $\cc{\trg{L}}{\obj{L}}$ be compilers from languages $\src{L}$ to $\trg{L}$ and $\trg{L}$ to $\obj{L}$, respectively.
-  Their sequential composition is $\left(\cc{\src{L}}{\trg{L}}\circ\cc{\trg{L}}{\obj{L}}\right)\left(\src{p}\right)=\obj{p}$.
-\end{definition}
+This section presents the key meta-theoretic results of this paper and extends on prior work~\cite{kruse2022csc}.
+A main result is that secure compilers in the robust preservation framework~\cite{abate2019jour} compose.
+This is not intuitive in the sense that there are countless examples in the security domain where composition does not work.\MK{cite}
+For starters, the sequential composition of compilers $\cc{\src{L}}{\trg{L}}$ and $\cc{\trg{L}}{\obj{L}}$ is as expected, i.e., given an $\src{L}$ program $\src{p}$, obtain its compiled $\obj{L}$ program by plugging $\src{p}$ into $\cc{\src{L}}{\trg{L}}\circ\cc{\trg{L}}{\obj{L}}$.
+Now, given \bul{$\cc{\src{L}}{\trg{L}}$ robustly preserves $\class[_{1}]$} and \rul{$\cc{\trg{L}}{\obj{L}}$ robustly preserves $\class[_{2}]$}, it follows that \oul{their sequential composition $\cc{\src{L}}{\trg{L}}\circ\cc{\trg{L}}{\obj{L}}$ robustly preserves the intersection of classes $\class[_{1}]$ and $\class[_{2}]$}.
 
 \begin{theorem}[Sequential Composition of Secure Compilers]\label{thm:rtp}
-  Given classes of properties $\class[_{1}],\class[_{2}]$, let $\rtp{\cc{\src{L}}{\trg{L}}}{\class[_{1}]}$ and $\rtp{\cc{\trg{L}}{\obj{L}}}{\class[_{2}]}$.
-  Then, $\rtp{\cc{\src{L}}{\trg{L}}\circ\cc{\trg{L}}{\obj{L}}}{\class[_{1}]\cap\class[_{2}]}$. \Coqed
+  Given $\class[_{1}],\class[_{2}]$, \bul{$\rtp{\cc{\src{L}}{\trg{L}}}{\class[_{1}]}$}, and \rul{$\rtp{\cc{\trg{L}}{\obj{L}}}{\class[_{2}]}$}, then \oul{$\rtp{\cc{\src{L}}{\trg{L}}\circ\cc{\trg{L}}{\obj{L}}}{\class[_{1}]\cap\class[_{2}]}$}.
+\Coqed
 \end{theorem}
 
-A corollary of \Cref{thm:rtp} is that optimization passes that have the input and output language can be freely swapped:
+Since the composition of secure compilers is again a secure compiler, the theorem generalizes to a whole chain of $n$ secure compilers.
+Pretending that the following compilers are secure, consider the compilation chain for $\src{TypeScript}$.
+First, $\src{TypeScript}$ programs are translated to $\trg{JavaScript}$ which, e.g., V8\MK{cit} eventually compiles in parts to $\irl{Ignition BC}$.
+Were all these compilation steps robustly secure with respect to \gls{ms}, the resulting $\irl{Ignition BC}$ code would be \gls{ms} regardless of the context the binary runs in.
+
+Notably, real-world compilation chains also perform a series of passes whose main purpose is not necessarily to translate from one language to another, but to, e.g., optimize the code or enforce a certain property.
+Both examples can be seen in the wild, there are countless instances of them.\MK{cit}
+Consider the compilation passes of LLVM\MK{cit} whose source-code optimisations are industry--competitive.
+In LLVM, the pass manager can be used to schedule passes in a way where the optimisations yield the best results.
+Take \gls{cf} and \gls{dce}.
+Unfortunately, the order in which \gls{cf} and \gls{dce} are performed influences the final result of the compilation (see \Cref{fig:cfdceex}).
+This {\em phase ordering problem} is well--known in literature and a practical solution is to simply perform a fixpoint iteration.
+However, this costs additional time until compilation is done, which can be a valuable resource\MK{cit}.
+Hence, compiler engineers typically try to find an order of optimizations that yields well-optimized programs as fast as possible.
+The following corollary justifies that any such order of compilation passes is valid with respect to security.
+So, given two compilation passes \bul{$\cc[_{1}]{\trg{L}}{\trg{L}}$, $\cc[_{2}]{\trg{L}}{\trg{L}}$, both robustly preserving class $\class[_{1}]$ or $\class[_{2}]$, respectively}, \oul{for any order of their composition the composed compiler robustly preserves the intersection of $\class[_{1}]$ and $\class[_{2}]$}.
+
 \begin{corollary}[Sequential Composition of Secure Compilers]\label{corr:swappable}
-  Given classes of properties $\class[_{1}],\class[_{2}]$, let $\rtp{\cc[_{1}]{\trg{L}}{\trg{L}}}{\class[_{1}]}$ and $\rtp{\cc[_{2}]{\trg{L}}{\trg{L}}}{\class[_{2}]}$.
-  Then, $\rtp{\cc[_{1}]{\trg{L}}{\trg{L}}\circ\cc[_{2}]{\trg{L}}{\trg{L}}}{\class[_{1}]\cap\class[_{2}]}$ and $\rtp{\cc[_{2}]{\trg{L}}{\trg{L}}\circ\cc[_{1}]{\trg{L}}{\trg{L}}}{\class[_{2}]\cap\class[_{1}]}$.
+  Given $\class[_{1}],\class[_{2}]$, \bul{$\rtp{\cc[_{1}]{\trg{L}}{\trg{L}}}{\class[_{1}]}$, and $\rtp{\cc[_{2}]{\trg{L}}{\trg{L}}}{\class[_{2}]}$}, then \oul{$\rtp{\cc[_{1}]{\trg{L}}{\trg{L}}\circ\cc[_{2}]{\trg{L}}{\trg{L}}}{\class[_{1}]\cap\class[_{2}]}$ and $\rtp{\cc[_{2}]{\trg{L}}{\trg{L}}\circ\cc[_{1}]{\trg{L}}{\trg{L}}}{\class[_{2}]\cap\class[_{1}]}$}.
 \Coqed
 \end{corollary}
+
+Besides the sequential composition, there are two other compositions that are more at the language- than the compiler--level.
+Nevertheless, consider $\trg{JVM BC}$, which is a popular target for programming language designers due to its high performance and relevance in industry.
+As an example, $\src{Java}$ and $\obj{Kotlin}$ compile to $\trg{JVM BC}$.
+Technically speaking, they both lower to class files and $\obj{Kotlin}$ objects are considered to be the same as $\src{Java}$ objects at that point.
+A compiler that accepts both $\src{Java}$ and $\obj{Kotlin}$ code translating to the same target language, or rather, intermediate representation, performs a kind of {\em upper} composition.
+Given a secure compiler from $\src{Java}$ to $\trg{JVM BC}$ and one from $\obj{Kotlin}$ to $\trg{JVM BC}$, it is easy to envision that one can engineer a combined compiler that accepts both $\src{Java}$ and $\obj{Kotlin}$.
+A practical example of this, which is, however, insecure, is {\tt Android Studio}\MK{cit}.
+The proof of \Cref{thm:rtp} can be easily adapted to also works for this {\em upper} composition.
+The corresponding theorem is not listed, since there is no general definition of this kind of composition.
+
+Likewise, the {\em lower} composition is concerned about compilers that accept the same source but yield different target languages and the same theoretical results apply.
+However, especially for this kind of composition, it is rather difficult to provide a general definition.
+Consider two compilers both accepting $\src{LLVM IR}$ and one of them emits $\trg{x86\_64}$, while the other emits $\trg{ARMv8}$.
+It is intuitive that they are in some sense composed in the LLVM framework, but the decision of when to use one over the other is inherently {\em unobservable} to the formalization effort of this kind of composition.
+Either the user of LLVM provides an explicit flag that instructs the overall compiler to emit $\trg{x86\_64}$ or the framework itself detects the target platform via heuristics.
+%Either way, the theoretical results suggest that it is possible to always find a ,,most--general'', secure compiler, given two secure compilers, that robustly preserves the least--upper bound of the classes involved in their compilation process.
+In general, there are two options to deal with these nuances in a concrete, practical setting:
+Given secure compilers, either (i) view the whole compilation chain involving them as the upper/sequential/lower compositions and conclude the robust preservation of the intersection of all involved classes, or (ii) specify meta-parameters that, e.g., specify the target language, and thus reduce the compilation chain to a nested sequential composition of the involved compilers.
+The first approach yields a more imprecise statement about the whole compilation chain while the second approach is more precise, but also less flexible.
+
+\begin{figure}
+  \begin{center}
+    \begin{tikzpicture}
+      \node[minimum width=3cm,minimum height=2.5cm,draw=black,very thick,align=left] (progunopt) {\begin{lstlisting}[language=ml,basicstyle=\scriptsize\ttfamily]
+let a = true in
+if a then
+  print "a"
+else
+  print "b"
+\end{lstlisting}};
+      \node[minimum width=3cm,minimum height=2.5cm,draw=black,very thick,align=left,right=1.0 of progunopt] (progoptcf) {\begin{lstlisting}[language=ml,basicstyle=\scriptsize\ttfamily]
+if true then
+  print "a"
+else
+  print "b"
+\end{lstlisting}};
+      \node[minimum width=3cm,minimum height=2.5cm,draw=black,very thick,align=left,right=1.0 of progoptcf] (progoptcfdce) {\begin{lstlisting}[language=ml,basicstyle=\scriptsize\ttfamily]
+print "a"
+\end{lstlisting}};% the following is a hack to get my syntax highlighting to work again \end{lstlisting}
+      % edges
+      \draw[->,very thick] (progunopt) edge[loop left,left] node {\gls{dce}} (progunopt);
+      \draw[->,very thick] (progunopt) edge[sloped,above] node {\gls{cf}} (progoptcf);
+      \draw[->,very thick] (progoptcf) edge[sloped,above] node {\gls{dce}} (progoptcfdce);
+    \end{tikzpicture}
+  \end{center}
+  \caption{Example program where the level of optimizations differ for one pass of applying \gls{cf} and \gls{dce} in any order. %
+  Every edge is a compilation pass and the label on the edge states what the pass does, i.e., \gls{cf} or \gls{dce}. %
+  The source code in the nodes is a glorified compiler intermediate representation and the code gets more optimized towards the righthandside of the figure.}\label{fig:cfdceex}
+\end{figure}
 

### 2023-07-06 popl/main.tex
Add conclusion and acks

@@
 
+This paper does a first step towards practical secure compilation chains by introducing a theoretical framework demonstrating that secure compilers compose.
+The paper exercises this framework on a case-study that consists of an optimizing compilation chain that is secure with respect to \gls{msscct}.
+In future work, it would be interesting to investigate wether it is possible to provide {\em secure compiler combinators}, similarily to how parser combinators work.
+%To this end, it may be necessary to extend existing frameworks for multi--language semantics.
+It is also interesting to extend our case--study and verify all results of it in Coq.
+This way, it is possible to obtain an executable, secure compiler.
+However, the formalization effort for backtranslation proofs is known to be enormous, so another research avenue may be to find better ways to (semi-)automate standard secure compilation proofs.
+
 %% The acknowledgments section is defined using the "acks" environment
@@
 \begin{acks}
+  We would like to thank the anonymous reviewers for their feedback.
 \end{acks}

### 2023-07-06 popl/main.tex
Add conclusion and acks

@@
 
-In order to define secure compilers one needs programming languages.
+This section defines programming languages that the secure compilers defined in the next section will make use of.
 To this end, this section defines the languages $\Ltms$, $\Ltrg$, $\Lms$, and $\Lscct$.
 In this paper, $\Ltms$ is the only statically typed language and exhibits the property that all well-typed programs are \gls{tms}.
-However, one can still write and evaluate $\Ltms$ programs that are well-typed, but perform out-of-bounds memory accesses, so the language is not \gls{sms}.
+However, $\Ltms$ programs are not \gls{sms}.
+That is, there are well--typed $\Ltms$ programs that perform an out--of--bounds access.
 In contrast, $\Ltrg$ is untyped and does not provide any guarantees with regards to \gls{ms}.
 $\Lms$ is exactly the same language as $\Ltrg$, but this paper still distinguishes the two for sake of readability.
-For all these languages, so $\Ltms$, $\Ltrg$, and $\Lms$, programmers assume \gls{cct} to hold, because the compiler is said to ensure it, similarily to what is being done in literature~\cite{cauligi2019fact}.
-This assumption does not hold anymore for the language $\Lscct$, which leaks information on branches and binary operations.
-But, contrary to the other languages, one can read and write to a model-specific register indicating a ``\gls{cct}-mode''.
+All three languages --- so $\Ltms$, $\Ltrg$, and $\Lms$ --- assume \gls{cct} to hold.
+Even though it is possible to write programs that, e.g., return early in loops, which is trivially not \gls{cct}, they can be transformed to adhere to \gls{cct}~\cite{cauligi2019fact}.
+The language $\Lscct$ allows violating \gls{cct} by emitting events on, e.g., branching and division, that contain secrets.
+As a countermeasure, the language provides a way to read and write to a {\em model--specific register} that enables a ``\gls{cct}-mode''.
 

### 2023-07-07 popl/main.tex
Shorten sec 6 up to blue lang

@@
 %\section{Case Study: Memory Safe and Cryptographic Constant Time Code}\label{sec:casestud}
-\section{Case Study: Preliminary Definitions\pages{6}}\label{sec:casestud:defs}
+\section{Case Study: Preliminary Definitions\pages{3}}\label{sec:casestud:defs}
 
@@
 \subsection{Shared Language Definitions}\label{subsec:cs:defs}
-\MPin{
-	explain with example ? leave text for app
-}
 
@@
   \begin{aligned}
-  &\mi{(Finals)}~\finalexpr\bnfdef\ \valueexpr \mid x \hspace{0.5cm}
-  \mi{(Values)}~\valueexpr\bnfdef\ n \hspace{0.5cm}
-  \mi{(Types)}~\type\bnfdef\ \natt \\
-  \mi{(Expressions)}~\expr&\bnfdef\ \finalexpr \mid \lbinop{\expr[_{1}]}{\expr[_{2}]} \mid \lget{x}{\expr} \mid \lnew{x}{\expr[_{1}]}{\expr[_{2}]} \mid \ldelete{x} \mid \lset{x}{\expr[_{1}]}{\expr[_{2}]} \\
+  \mi{(Expressions)}~\expr&\bnfdef\ x \mid \valueexpr \mid \lbinop{\expr[_{1}]}{\expr[_{2}]} \mid \lget{x}{\expr} \mid \lnew{\expr[_{1}]}{\expr[_{2}]} \mid \ldelete{x} \mid \lset{x}{\expr[_{1}]}{\expr[_{2}]} \\
     & \mid \llet{x}{\expr[_{1}]}{\expr[_{2}]}  \mid \lreturn{\expr} \mid \lcall{g}{\expr} \mid \lifz{\expr}{\expr[_{1}]}{\expr[_{2}]} \mid \labort \\
-  \mi{(Symbols)}~\function&\bnfdef\ \lfunction{g}{x}{e}\hspace{0.5cm}
-  \mi{(Libraries)}~\library\bnfdef\ \hole{\cdot} \mid \function,\library \hspace{1.00cm}
-    \text{where}\ \lbinop{}{}\in\left\{+,-,\times,<,/\right\}\\
-  &\mi{(Programs)}~\prog{\library_{\ctx}}{\library_{\comp}}\hspace{0.5cm}
+  \mi{(Symbols)}~&\function\bnfdef\ \lfunction{g}{x}{e}\hspace{0.5cm}
+  \mi{(Libraries)}~\library\bnfdef\ \hole{\cdot} \mid \function,\library \\
+  \mi{(Programs)}~&\prog{\library_{\ctx}}{\library_{\comp}}\hspace{0.5cm}
   \mi{(Substitutions)}~\substvar\bnfdef \hole{\cdot} \mid \subst{\valueexpr}{x},\gamma \\
@@
 
-Above is the base syntax of all the programming languages of this paper.
-\MK{this needs a complete rewrite}
-Expressions ($\expr$) can only be evaluated in the form of runtime--terms ($\rtt{\statevar}{\expr}$).
-States ($\statevar$) have a flow- ($\cfstate$) and a memory state ($\memstate$).
-As for the flow state, the list of component--level names ($\commlib$), the list of symbols ($\library$), and a stack of continuations tagged with the name of a symbol ($\ectxstack$) are necessary for calling and returning while emitting the right transfer tag ($\sandboxtag$) depending on whether there is a context switch from context to component ($\comptoctx$), from component to context ($\ctxtocomp$), or whether control flow stays internal to either ($\nocomm$).
-The memory state contains both a context--level ($\heap[^{\ctx}]$) and a component--level heap ($\heap[^{comp}]$) as well as a store ($\store$).
-Effectively, it is syntactically impossible for a context--level runtime--term to have a pointer into the component--level heap and vice--versa.
-Pointers are non-nullable and there is no syntactic way to let them point into the middle of an object.
-However, pointers always point to arrays and the individual elements of that array are read- and writeable by means of adding an offset that may be negative.
-Heap locations are second--class. %The syntax is constrained as such that one cannot simply guess a heap location.
-Stores map identifiers to pointer metadata.
-This metadata is a tuple with the location ($\loc$), the control--tag ($\sandboxtag$), poison--tag ($\poison$), and the size of the allocated object ($n$).
-The control-tag allows to distinguish whether a location indexes into $\heap[^{\ctx}]$ or $\heap[^{\comp}]$.
-The poison--tag marks a location as freed ($\poisoned$) without removing the entry from $\store$.
-If one removes an entry, for example, by evaluating $\ldelete{x}$, then there is no way to distinguish between freed pointers and free variables.
-The latter should render evaluation as stuck, while the former should emit a $\ev{Dealloc}\ \loc$ event regardless of wether the object a pointer points to is still allocated or not.
-Programs ($\prog{\library_{\ctx}}{\library_{\comp}}$) have disjoint sets of symbols, one for context-level ($\library_{\ctx}$) and one for component-level symbols ($\library_{\comp}$).
-
-\begin{gather*}
-  \begin{aligned}
-  \mi{(Eval.~Ctx.)}~\ectx&\bnfdef\ \hole{\cdot} \mid \lbinop{\ectx}{\expr} \mid \lbinop{\valueexpr}{\ectx} \mid \lget{x}{\ectx} \mid \lnew{x}{\ectx}{\expr} \mid \lset{x}{\ectx}{\expr}\\
-    & \mid \lset{x}{\valueexpr}{\ectx} \mid \lreturn{\ectx} \mid \lcall{g}{\ectx} \mid \lifz{\ectx}{\expr[_{1}]}{\expr[_{2}]} \\
-  \end{aligned}\\
-  \begin{aligned}
-  \mi{(States)}~\statevar&\bnfdef \tup{\cfstate}{\sandboxtag,\memstate} &
-  \mi{(Poison)}~\poison&\bnfdef\ \poisoned \mid \poisonless &
-  \mi{(Locations)}~\loc\in\nat \\
-  \end{aligned}\\
-  \begin{aligned}
-  \mi{(Heaps)}~\heap&\bnfdef\ \hole{\cdot} \mid n,\heap &
-  \mi{(Stores)}~\store&\bnfdef\ \hole{\cdot} \mid \storeel{x}{\loc}{\sandboxtag}{\poison}{n},\store &
-  \mi{(Transfer~Tags)}~\comm\bnfdef\ \nocomm \mid\ \ctxtocomp \mid\ \comptoctx \\
-  \end{aligned}\\
-  \begin{aligned}
-  \mi{(Runtime~Terms)}~\runtimetermvar&\bnfdef\rtt{\statevar}{\expr} \mid \lightning &
-  \mi{(Component~Names)}~\commlib&\bnfdef\ \hole{\cdot} \mid g,\commlib \\
-  \end{aligned}\\
-  \begin{aligned}
-  \mi{(Continuation~Stacks)}~&\ectxstack\bnfdef\ \hole{\cdot} \mid \tup{\ectx}{g},\ectxstack \\
-  \end{aligned}
-\end{gather*}
-All programming languages presented in this section have an environmental semantics and all but $\src{L_{\tmssafe}}$ are dynamically typed.
-Each evaluation may emit the empty event ($\emptyevent$), a program crash token ($\lightning$), or a pair containing a pre--event ($\event[_{b}]$) and the corresponding control--tag ($\sandboxtag$) indicating whether the event fired during an evaluation of a context or a component.
-Note that the reflexive-transitive closure used in this paper behaves like a big-step semantics, where the reflexive case is ony for the empty trace and finals, and the emitted trace does not contain empty events.
-Most rules are entirely standard and omitted.
-Heaps may need to grow when allocating, which the section writes as $\growh{\heap}{n}$, denoting that $\heap$ grows by $n$ cells.
-Indexing into a heap, looking up symbols in libraries, or accessing the metadata of a pointer from a store are all written as a function call.
-For example, the value located at the fifth position in heap $\heap$ is written as $\heap(5)$.
-The operation to set the fifth position in heap $\heap$ to value $\valueexpr$ is written as $\seth{\heap}{5}{\valueexpr}$ and yields a new heap that is entirely the same to $\heap$ modulo the fifth position.
+Above is a fraction of the shared syntax of all the programming languages of this paper.
+The syntax and semantics are mostly standard.
+However, when allocating pointers with $\lnew{\expr[_{1}]}{\expr[_{2}]}$, the expression $\expr[_{1}]$ determines the initial value of the objects and $\expr[_{2}]$ the overall size of an array.
+The data associated to pointers resides in the dynamic state $\Omega$ that contains sandboxed heaps, one for the context and the other for the component.
+Because of the sandboxing, locations pointing into heaps must be associated with a control--tag to know the heap they point into.
+Other metadata in $\Omega$ entails a list of callable objects with information on wether they belong to the context or the component.
+That way, when calling a callable object, the semantics can take care of emitting appropriate events that signal a context switch.
+For the languages considered in this paper, no callable objects are first--class.
+So, the only values are pairs, pointers, and natural numbers.
+Another non--standard part of the languages is the concept of the poison--tag.
+The poison--tag marks a location as freed ($\poisoned$) without removing the information associated to the freed pointer from $\Omega$.
+This measure prevents the semantics from getting stuck when an already freed pointer is freed again.
+Because of this, it is possible that programs like $\llet{x}{\lnew{0}{5}}{\ldelete{x};\ldelete{x}}$ reduce and emit a trace along the lines of $\specificev{Alloc\ \loc\ \sandboxtag\ 5}\cdot\specificev{Dealloc\ \loc\ \sandboxtag}\cdot\specificev{Dealloc\ \loc\ \sandboxtag}$.
 
@@
 
-\MKin{now is time to talk about events? or maybe right after presenting the rules}
-
-\begin{center}
-  \judgbox{\pstepto{\runtimetermvar}{\runtimetermvar[']}{\event}}{,,Primitive step from runtime--term $\runtimetermvar$ to $\runtimetermvar[']$ emitting event $\event$.''}
-  \typerule{$e-\text{get}-\in$}{
-    \sandboxtag=\statevar.\sandboxtag &
-    \statevar.\store(x)=\tup{\loc}{\sandboxtag,\poison,m} \\
-    \loc+n\in\statevar.\heap[^{\sandboxtag}]
-  }{
-    \pstepto{\rtt{\statevar}{\lget{x}{n}}}{\rtt{\statevar}{\heap[^{\sandboxtag}](\loc+n)}}{\tup{\ev{Get}\ \loc\ n}{\sandboxtag}}
-  }{e-get-in}
-  %
-  \typerule{$e-\text{get}-\notin$}{
-    \sandboxtag=\statevar.\sandboxtag &
-    \statevar.\store(x)=\tup{\loc}{\sandboxtag,\poison,m} \\
-    \loc+n\notin\statevar.\heap[^{\sandboxtag}]
-  }{
-    \pstepto{\rtt{\statevar}{\lget{x}{n}}}{\rtt{\statevar}{1729}}{\tup{\ev{Get}\ \loc\ n}{\sandboxtag}}
-  }{e-get-notin}
-  %
-  \typerule{$e-\text{set}-\in$}{
-    \sandboxtag=\statevar.\sandboxtag &
-    \statevar.\store(x)=\tup{\loc}{\sandboxtag,\poison,m} \\
-    \loc+n\in\statevar.\heap[^{\sandboxtag}] &
-    \heap[^{\sandboxtag}_1] = \seth{\statevar.\heap[^{\sandboxtag}]}{\loc+n}{\valueexpr}
-  }{
-    \pstepto{\rtt{\statevar}{\lset{x}{n}{\valueexpr}}}{\rtt{\statevar\subst{\heap[^{\sandboxtag}_1]}{\heap[^{\sandboxtag}]}}{\valueexpr}}{\tup{\ev{Set}\ \loc\ n\ \valueexpr}{\sandboxtag}}
-  }{e-set-in}
-  %
-  \typerule{$e-\text{set}-\notin$}{
-    \sandboxtag=\statevar.\sandboxtag &
-    \statevar.\store(x)=\tup{\loc}{\sandboxtag,\poison,m} \\
-    \loc+n\notin\statevar.\heap[^{\sandboxtag}]
-  }{
-    \pstepto{\rtt{\statevar}{\lset{x}{n}{\valueexpr}}}{\rtt{\statevar}{\valueexpr}}{\tup{\ev{Set}\ \loc\ n\ \valueexpr}{\sandboxtag}}
-  }{e-set-notin}
-  %
-  \typerule{$e-\text{new}$}{
-    \fresh{\statevar}{z} &
-    \fresh{\statevar}{\loc} &
-    \heap[^{\sandboxtag}_1]=\growh{\statevar.\heap[^{\sandboxtag}]}{n} &
-    \store[_1]=\storeel{z}{\loc}{\statevar.\sandboxtag}{\poisonless}{n},\statevar.\store
-  }{
-    \pstepto{\rtt{\statevar}{\lnew{x}{n}{\expr}}}{\rtt{\statevar\subst{\heap[^{\sandboxtag}]}{\heap[^{\sandboxtag}_1]}\subst{\store}{\store[_1]}}{\expr\subst{z}{x}}}{\tup{\ev{Alloc}\ \loc\ n}{\sandboxtag}}
-  }{e-new}
-  %
-  \typerule{$e-\text{dealloc}$}{
-    \statevar.\store(x) = \tup{\loc}{\statevar.\sandboxtag,\poison,n} &
-    \store[_1]=\seth{\statevar.\store}{x}{\tup{\loc}{\statevar.\sandboxtag,\poisoned,n}}
-  }{
-    \pstepto{\rtt{\statevar}{\ldelete{x}}}{\rtt{\statevar\subst{\store}{\store[_1]}}{0}}{\tup{\ev{Dealloc}\ \loc}{\sandboxtag}}
-  }{e-dealloc}
-  %
-  \typerule{$e-\text{abort}$}{
-  }{
-    \pstepto{\rtt{\statevar}{\labort}}{\lightning}{\lightning}
-  }{e-abort}
-\end{center}
-
-Reading the value of a pointer is done by looking up the metadata in the store (\Cref{tr:e-get-in,tr:e-get-notin}).
-This is done without any bounds-checking and always emits a $\ev{Get}\ \loc\ n$ event regardless of whether the access is out of bounds or not.
-Similarily when setting a value (\Cref{tr:e-set-in,tr:e-set-notin}), a $\ev{Set}\ \loc\ n\ \valueexpr$ event is emitted regardless of wether the set is in or out of bounds.
-Allocation of objects (\Cref{tr:e-new}) does the expected, however, it makes use of the judgement $\fresh{\cdot}{\cdot}$ which is used throughout this paper to explicitly mark when and with respect to what environment a fresh variable is introduced.
-Deallocation (\Cref{tr:e-dealloc}) simply sets the poison--tag in the metadata of the associated pointer to be poisoned ($\poisoned$).
-The only way to abort an execution (\Cref{tr:e-abort}) is with an explicit call to $\labort$.
-However, evaluation of programs can get stuck, e.g., if $x$ is not bound in $\ldelete{x}$.
-
+Possible events that are emitted during execution are similar to the ones of the specification trace model.
+The key difference is that, as standard in secure compilation work, the traces have a call and return event that signals context switches.
+Hereby, a $\specificev{Call\ \ctxtocomp\ foo\ \valueexpr}$ and $\specificev{Return\ \ctxtocomp\ \valueexpr}$ signal that program execution transitions from context-- to component--level.
+Contrary, $\specificev{Call\ \comptoctx\ foo\ \valueexpr}$ and $\specificev{Return\ \ctxtocomp\ \valueexpr}$ signal that program execution transitions from component-- to context--level.
+The environmental semantics emits the tag $\nocomm$ for calls from context-- to context--level or component-- to component--level.
 In the following, $\neg\ctx=\comp$ and $\neg\comp=\ctx$ and also $\operatorname{comm}(\ctx)=\ \ctxtocomp$ and $\operatorname{comm}(\comp)=\ \comptoctx$.
@@
   \judgbox{\estepto[]{\runtimetermvar}{\runtimetermvar}{\event}}{,,Contextual step from runtime--term $\runtimetermvar$ to $\runtimetermvar[']$ emitting event $\event$.''}
-  %
-  \typerule{$e-\text{ectx}$}{
-    \pstepto{\rtt{\statevar}{\expr}}{\rtt{\statevar[']}{\expr[']}}{\event}
-  }{
-    \estepto[]{\rtt{\statevar}{\ectx\hole{\expr}}}{\rtt{\statevar[']}{\ectx\hole{\expr[']}}}{\event}
-  }{e-ectx}
-  %
-  \typerule{$e-\text{stuck}$}{
-    \pstepto{\rtt{\statevar}{\expr}}{\lightning}{\lightning}
-  }{
-    \estepto[]{\rtt{\statevar}{\ectx\hole{\expr}}}{\lightning}{\lightning}
-  }{e-stuck}
   %
@@
     \statevar.\ectxstack=\tup{\ectx}{\text{foo}},\ectxstack['] &
-    \commlib\vdash\text{foo}:\Omega.\sandboxtag &
-    \operatorname{comm}(\sandboxtag)=\comm
+    \commlib\vdash\text{foo}:\Omega.\sandboxtag\dashv\comm
   }{
@@
   %
-  \typerule{$e-\text{start}$}{
-    \lfunction{\text{main}}{x}{e} \in \statevar.\library
-  }{
-    \estepto[]{\rtt{\statevar}{\lcall{\text{main}}{0}}}{\rtt{\statevar\subst{\ctx}{\sandboxtag}\subst{\tup{\hole{\cdot}}{\text{main}}}{\ectxstack}}{\expr\subst{0}{x}}}{\tup{\ev{Start}}{\comp}}
-  }{e-start}
-  %
   \typerule{$e-\text{call}-\text{notsame}$}{
     \lfunction{\text{foo}}{x}{e} \in \statevar.\library &
-    \commlib\vdash\text{foo}:\Omega.\sandboxtag &
-    \operatorname{comm}(\sandboxtag)=\comm
+    \commlib\vdash\text{foo}:\Omega.\sandboxtag\dashv\comm
   }{
@@
   }{e-call-notsame}
-  %
-  \judgbox{\commlib\vdash\text{foo}:\sandboxtag}{,,Given execution is in $\sandboxtag$, the symbol associated with the name $\text{foo}$ is in $\neg\sandboxtag$.''}
-  %
-  \typerule{ctx-to-comp}{
-    \text{foo}\in\commlib
-  }{
-    \commlib\vdash\text{foo}:\ctx
-  }{ctx-to-comp}
-  %
-  \typerule{comp-to-ctx}{
-    \text{foo}\notin\commlib
-  }{
-    \commlib\vdash\text{foo}:\comp
-  }{comp-to-ctx}
-\end{center}
-
-Contextual steps are as expected (\Cref{tr:e-ectx,tr:e-stuck}) up to calling (\Cref{tr:e-call-notsame}) and returning.
-For calls and returns, a bit of work has to be done to emit the right events.
-For example, the call to the main function emits just $\ev{Start}$ (\Cref{tr:e-start}) instead of $\ev{Call}\ \comptoctx\ \text{main}\ 0$, which is a design choice this paper does for convenience when reasoning about a call-chain.
-Ordinary calls, however, need to check wether a context switch is happening or not and push the current continuation and the name of the invoked function on the stack in the state (\Cref{tr:e-call-notsame}).
-Context switches are checked by looking at the current control--tag and wether the involved function or surrounding context is part of the list of names of component--level symbols ($\commlib$) or not (\Cref{tr:ctx-to-comp,tr:comp-to-ctx}).
-
-\begin{center}
-  \judgbox{\progstepto{\prog{\library_{\ctx}}{\library_{\comp}}}{\runtimetermvar}{\trace}}{,,Program $\prog{\library_{\ctx}}{\library_\comp}$ evaluates to $\runtimetermvar$, emitting trace $\trace$.''}
-  %
-  \typerule{ep-run}{
-    \link{\library_{\ctx}}{\library_{\comp}}=\library &
-    \commlib=\operatorname{dom}\library_{\comp} \\
-    \estepto{\rtt{\tup{\library}{\commlib,\emptyset}}{\lcall{\text{main}}{0}}}{\rtt{\statevar}{\valueexpr}}{\trace}
-  }{
-    \progstepto{\prog{\library_{\ctx}}{\library_{\comp}}}{\rtt{\statevar}{\valueexpr}}{\trace}
-  }{ep-run}
-  %
-  \typerule{ep-crash}{
-    \link{\library_{\ctx}}{\library_{\comp}}=\library &
-    \commlib=\operatorname{dom}\library_{\comp} \\
-    \estepton{\rtt{\tup{\library}{\commlib,\emptyset}}{\lcall{\text{main}}{0}}}{\lightning}{\trace\cdot\lightning}
-  }{
-    \progstepto{\prog{\library_{\ctx}}{\library_{\comp}}}{\lightning}{\trace\cdot\lightning}
-  }{ep-crash}
-\end{center}
-
-The notation $\tup{\library}{\commlib,\emptyset}$ unfolds to $\statevar=\tup{\cfstate}{\comp,\memstate}$, where $\cfstate=\tup{\library}{\commlib,\hole{\cdot}}$ and $\memstate=\tup{\hole{\cdot}}{\hole{\cdot},\hole{\cdot}}$.
-Linking $\link{\library_{\ctx}}{\library_{\comp}}$, while gluing $\library_{\ctx}$ and $\library_{\comp}$ together, ensures that there is no symbol with the name $\text{main}$ in $\library_{\comp}$.
-Programs can either be run successfully (\Cref{tr:ep-run}) or fail (\Cref{tr:ep-crash}), which requires the n-step relation instead of star step, since star step models a big--step style semantics without failure.
-To make use of \Cref{def:propsat,def:proprsat}, {\em behavior} is introduced as follows.
-
-\begin{definition}[Program Behavior]\label{def:behav}
-  Given a program $\prog{\library_{\ctx}}{\library_{\comp}}$, the behavior of it is a set containing all traces $\trace$ such that there is a runtime--term $\runtimetermvar$ with $\progstepto{\prog{\library_{\ctx}}{\library_{\comp}}}{\runtimetermvar}{\trace}$.
-\end{definition}
+\end{center}
+
+The environmental semantics is also mostly straightforward.
+In \Cref{tr:e-ret-notsame,tr:e-call-notsame} the judgement $\commlib\vdash\text{foo}:\Omega.\sandboxtag$ checks wether $\text{foo}$ is a component--level name by looking it up in the list of component-level names $\commlib$ and emits the appropriate transfer tag, i.e., either $\comptoctx$ or $\ctxtocomp$.
+Additional rules that are left out ensure that, e.g., when calling the $\text{main}$ function, the event $\ev{Start}$ is emitted, which is a design choice this paper does for convenience when reasoning about call--chains.
+
+The top--level execution $\progstepto{\prog{\library_{\ctx}}{\library_{\comp}}}{\runtimetermvar}{\trace}$ constructs an initial state $\Omega$ by linking $\library_{\ctx}$ and $\library_{\ctx}$ and then starts execution by calling the $\text{main}$ function.
+The trace $\trace$ emitted during that execution serves as abstraction of the behavior of the program enabling the use of \Cref{def:propsat,def:proprsat}.
 

### 2023-07-07 popl/main.tex
Remove uneccessary parts from section 6.

@@
 %\section{Case Study: Memory Safe and Cryptographic Constant Time Code}\label{sec:casestud}
-\section{Case Study: Preliminary Definitions\pages{3}}\label{sec:casestud:defs}
+\section{Case Study: Preliminary Definitions\pages{2}}\label{sec:casestud:defs}
 
@@
 This measure prevents the semantics from getting stuck when an already freed pointer is freed again.
-Because of this, it is possible that programs like $\llet{x}{\lnew{0}{5}}{\ldelete{x};\ldelete{x}}$ reduce and emit a trace along the lines of $\specificev{Alloc\ \loc\ \sandboxtag\ 5}\cdot\specificev{Dealloc\ \loc\ \sandboxtag}\cdot\specificev{Dealloc\ \loc\ \sandboxtag}$.
+Because of this, it is possible that programs like $\llet{x}{\lnew{0}{5}}{\ldelete{x};\ldelete{x}}$ reduce and emit a trace along the lines of $\ev{Alloc\ \loc\ \sandboxtag\ 5}\cdot\ev{Dealloc\ \loc\ \sandboxtag}\cdot\ev{Dealloc\ \loc\ \sandboxtag}$.
 
@@
 The key difference is that, as standard in secure compilation work, the traces have a call and return event that signals context switches.
-Hereby, a $\specificev{Call\ \ctxtocomp\ foo\ \valueexpr}$ and $\specificev{Return\ \ctxtocomp\ \valueexpr}$ signal that program execution transitions from context-- to component--level.
-Contrary, $\specificev{Call\ \comptoctx\ foo\ \valueexpr}$ and $\specificev{Return\ \ctxtocomp\ \valueexpr}$ signal that program execution transitions from component-- to context--level.
+Hereby, a $\ev{Call\ \ctxtocomp\ foo\ \valueexpr}$ and $\ev{Return\ \ctxtocomp\ \valueexpr}$ signal that program execution transitions from context-- to component--level.
+Contrary, $\ev{Call\ \comptoctx\ foo\ \valueexpr}$ and $\ev{Return\ \ctxtocomp\ \valueexpr}$ signal that program execution transitions from component-- to context--level.
 The environmental semantics emits the tag $\nocomm$ for calls from context-- to context--level or component-- to component--level.
@@
 \subsection{$\src{L_{\tmssafe}}$: A Temporal but Not Spatial Memory Safe Language}\label{subsec:ltms}
-\begin{gather*}
-  \begin{aligned}
-    \mi{(Pre\text{--}Types)}~\src{\type[_{e}]}\bnfdef&\ \src{\natt} \mid \src{\ptrn} &
-    \mi{(Type~Environments)}~\src{\typenv}\bnfdef&\ \src{\hole{\cdot}} \mid \hastype{\src{x}}{\src{\type}},\src{\typenv} \\
-    \mi{(Types)}~\src{\type}\bnfdef&\ \src{\type[_{e}]} \mid \src{\type[_{e}]\to\bot} \mid \src{\type[^{(1)}_{e}]\to\type[^{(2)}_{e}]} &
-    \mi{(Pointer\ Qualifiers)}~\src{\ptrqual}\bnfdef&\ \src{\fullq} \mid \src{\halfq} \\
-    \mi{(Symbols)}~\src{\function}\bnfdef&\src{\lfunction{g}{x:\type[_{e}]\to\type[_{e}]}{\expr}} \\
-  \end{aligned}
-\end{gather*}
-\begin{center}
-  \judgbox{\tcheck{\src{\typenv}}{\src{\expr}}{\src{\type}}}{,,Under environment $\src{\typenv}$, expression $\src{\expr}$ has type $\src{\type}$.''}
-  %
-  \typerule{t-$\src{x}$}{
-    \src{\typenv}(\src{x})=\src{\type} &
-    \notowned{\src{\typenv}}
-  }{
-    \tcheck{\src{\typenv}}{\src{x}}{\src{\type}}
-  }{t-x}
-  %
-  \typerule{t-$\src{get}$}{
-    \tcheck{\src{\typenv[_2]}}{\src{x}}{\src{\wptr}} &
-    \tcheck{\src{\typenv[_1]}}{\src{\expr}}{\src{\natt}} &
-  }{
-    \tcheck{\typenvsplit{\src{\typenv[_1]}}{\src{\typenv[_2]}}}{\src{\lget{x}{\expr}}}{\src{\natt}}
-  }{t-get}
-  %
-  \typerule{t-$\src{let}$}{
-    \tcheck{\src{\typenv[_1]}}{\src{\expr[_1]}}{\src{\type[_e]}} &
-    \tcheck{\hastype{\src{x}}{\src{\type[_e]}},\src{\typenv[_2]}}{\src{\expr[_2]}}{\src{\type}}
-  }{
-    \tcheck{\typenvsplit{\src{\typenv[_1]}}{\src{\typenv[_2]}}}{\src{\llet{x}{\expr[_1]}{\expr[_2]}}}{\src{\type}}
-  }{t-let}
-  %
-  \typerule{t-$\src{delete}$}{
-    \src{\typenv}(\src{x})=\src{\ptr} &
-    \notowned{\src{\typenv}\setminus\left\{\hastype{\src{x}}{\src{\ptr}}\right\}}
-  }{
-    \tcheck{\src{\typenv}}{\src{\ldelete{x}}}{\src{\natt}}
-  }{t-delete}
-  %
-  \typerule{t-$\src{abort}$}{
-    \notowned{\src{\typenv}}
-  }{
-    \tcheck{\src{\typenv}}{\src{\labort}}{\src{\type}}
-  }{t-abort}
-  %
-  \typerule{t-$\src{call}$}{
-    \inttype{\src{\type[^{(1)}_e]}} &
-    \inttype{\src{\type[^{(2)}_e]}} \\
-    \tcheck{\src{\typenv}}{\src{foo}}{\src{\type[^{(1)}_e]\to\type[^{(2)}_e]}} &
-    \tcheck{\src{\typenv}}{\src{\expr}}{\src{\type[^{(1)}_e]}}
-  }{
-    \tcheck{\src{\typenv}}{\src{\lcall{foo}{\expr}}}{\src{\type[^{(2)}_e]}}
-  }{t-call}
-  %
-  \typerule{t-$\src{return}$}{
-    \inttype{\src{\type[_e]}} &
-    \tcheck{\src{\typenv}}{\src{\expr}}{\src{\type[_e]}}
-  }{
-    \tcheck{\src{\typenv}}{\src{\lreturn{\expr}}}{\src{\type[_e]\to\bot}}
-  }{t-return}
-\end{center}
-
-$\src{L_{\tmssafe}}$ is a statically typed language where each well-typed term is temporal memory safe.
-Most parts of the type system are entirely standard, so only the interesting bits are shown here.
-To enforce this, the language marks pointers with respect to their ownership, where $\src{\fullq}$ means {\em owned} and $\src{\halfq}$ means {\em not owned}.
-An owned pointer can only be used by a deallocation (\Cref{tr:t-delete}), whereas reads and writes must be done on non-owned pointers.
-One can convert owned to non-owned pointers implicitly through context splitting $\typenvsplit{\src{\typenv[_{1}]}}{\src{\typenv[_{2}]}}$.\MK{list some rules and explain it a bit}
-Syntactic constraints together with the type system ensure that non-owned pointers never ,,live'' longer than owned pointers.
-For example, in an expression $\src{\llet{x}{\expr[_{1}]}{\expr[_{2}]}}$, if $y$ is a $\src{\ptr}$ in $\src{\expr[_{1}]}$, it cannot appear at all in $\src{\expr[_{2}]}$, but it can if it were of type $\src{\wptr}$ (\Cref{tr:t-let}).
-This mechanism also prevents terms like $\src{\lget{x}{\ldelete{x}}}$ from typechecking (\Cref{tr:t-get}).
-At leafs of the type derivation tree, the typing environment must not contain owned pointers, which is ensured with the $\notowned{\src{\typenv}}$ judgement (\Cref{tr:t-abort,tr:t-x}).
-Lastly, calling or returning requires passed values to have an interface type ($\inttype{\src{\type[_{e}]}}$), which are just $\src{\natt}$ (\Cref{tr:t-call,tr:t-return}).
-To ensure that no pointers live after a return, the type of a return is that of a continuation, which is only permitted to propagate towards the end of a term.
-For example, we cannot bind $\src{\lreturn{e}}$ to an identifier in the typecheck of $\src{let}$, because the expression whose result is written into the binder has to be a pre--type, which the type of $\src{\lreturn{e}}$ is not.
-
-\subsection{All Well-Typed $\Ltms$ Programs are Temporal Memory Safe}
-
-\begin{center}
-  \judgbox{\thelocmap(\src{\loc})=\loc}{,,Map an $\src{L_{\tmssafe}}$ location to a specification location $\loc$.''}
-  \judgbox{\msfilterLtms{\src{\event}}=\event}{,,Project an $\src{L_{\tmssafe}}$ event $\src{\event}$ to a specification event $\event$.''}
-  \typerule{$\src{L_{\tmssafe}}$-filter-context}{
-    \src{\event[_b]}\not=\src{\lightning}
-  }{
-    \msfilterLtms{\src{\tup{\event[_b]}{\ctx}} = \emptyevent}
-  }{ltms-filter-context}
-  %
-  \typerule{$\src{L_{\tmssafe}}$-filter-abort}{
-  }{
-    \msfilterLtms{\src{\lightning}} = \lightning
-  }{ltms-filter-abort}
-  %
-%  \typerule{$\src{L_{\tmssafe}}$-filter-start}{
-%  }{
-%    \msfilterLtms{\src{\tup{Start}{\comp}}} = \emptyevent
-%  }{ltms-filter-start}
-  %
-  \typerule{$\src{L_{\tmssafe}}$-filter-alloc}{
-    \locmapsto{\src{\loc}}{\loc} &
-    \src{n}=n
-  }{
-    \msfilterLtms{\src{\tup{Alloc\ \loc\ n}{\comp}}} = \tup{\ev{Alloc}\ \loc\ n}{\comp,\unlock}
-  }{ltms-filter-alloc}
-\end{center}
-To prove that well-typed $\src{L_{\tmssafe}}$ programs are temporal memory safe, a map from $\src{L_{\tmssafe}}$ to specification events (see \Cref{subsec:propdefs}) allows the use of \Cref{def:monsat}.
-The map is mostly entirely straight-forward.
-Most notably, it filters any action that the context does (\Cref{tr:ltms-filter-context}), since a context is free to do anything.
-Of course, a context invoking $\src{\labort}$ to crash the program must not go undetected (\Cref{tr:ltms-filter-abort}).
-As for the component, pre--events are essentially recolored to go from blue $\src{L_{\tmssafe}}$ events to the events of the most general trace model (\Cref{tr:ltms-filter-alloc}).
-Events that carry locations make use of a location--map from blue $\src{L_{\tmssafe}}$ locations to the more abstract ones.
-
-\begin{lemma}[$\src{L_{\tmssafe}}$-programs are \gls{tms}]\label{lem:wt:tmsmon}
-  If $\progstepto{\src{\prog{\library_{\ctx}}{\library_{\comp}}}}{\src{\rtt{\statevar}{\valueexpr}}}{\src{\trace}}$, then $\montmssafe{\msfilterLtms{\src{\trace}}}$.\Coqed
-\end{lemma}
+
+$\src{L_{\tmssafe}}$ does not extend the base syntax at term level.
+What makes $\src{L_{\tmssafe}}$ special is that it has a type system inspired by $L^{3}$~\cite{morrisett2005L3} that exhibits the property that every well--typed $\src{L_{\tmssafe}}$ program satisfies \gls{tms}.
+The proof of this property relies on a projection $\msfilterLtms{\src{\event}}=\event$ that projects $\src{L_{\tmssafe}}$ events to specfication events $\event$.
+Hereby, $\delta$ maps $\src{L_{\tmssafe}}$ locations $\src{\loc}$ to locations $\loc$ of the specification trace model.
+The projection $\msfilterLtms{\src{\event}}$ for the most part does the expected.
+But, it also filters any action that a contexts does, since these are irrelevant for component--level \gls{tms}.
+
 \begin{theorem}[$\src{L_{\tmssafe}}$-programs are \gls{tms}]\label{thm:wt:tms}
@@
 
-\subsection{$\Ltrg$: A Memory-Unsafe Language}\label{subsec:lsms}
-
-\begin{gather*}
-  \begin{aligned}
-    \mi{(Expressions)}~\trg{\expr}\bnfdef\ \dots \mid \trg{\lispoisoned{x}} \mid \trg{\lpair{\expr[_{1}]}{\expr[_{2}]}} &\mid \trg{\lproja{\expr}} \mid \trg{\lprojb{\expr}} \mid \trg{\lhast{\expr}{\type}} \hspace{0.5cm}
-    \mi{(Types)}~\trg{\type}\bnfdef\ \trg{\natt} \mid \trg{\natt\times\natt} \\
-    \mi{(Values)}~\trg{\valueexpr}&\bnfdef\ \dots \mid \trg{\lpair{n_{1}}{n_{2}}}
-  \end{aligned}
-\end{gather*}
-\begin{center}
-  \typerule{e-$\trg{\lispoisoned{}}$-yes}{
-    \trg{\statevar.\store}(\trg{x})=\trg{\tup{\loc}{\statevar.\sandboxtag,\poisoned,n}}
-  }{
-    \pstepto{\rtt{\trg{\statevar}}{\trg{\lispoisoned{x}}}}{\rtt{\trg{\statevar}}{\trg{0}}}{\trg{\emptyevent}}
-  }{e-ispoisoned-yes}
-  %
-%  \typerule{e-$\trg{\lispoisoned{}}$-no}{
-%    \trg{\statevar.\store}(\trg{x})=\trg{\tup{\loc}{\statevar.\sandboxtag,\poisonless,n}}
-%  }{
-%    \pstepto{\rtt{\trg{\statevar}}{\trg{\lispoisoned{x}}}}{\rtt{\trg{\statevar}}{\trg{1}}}{\trg{\emptyevent}}
-%  }{e-ispoisoned-no}
-  %
-  \typerule{e-$\trg{\lhast{n}{\natt}}$}{
-  }{
-    \pstepto{\rtt{\trg{\statevar}}{\trg{\lhast{n}{\natt}}}}{\rtt{\trg{\statevar}}{\trg{0}}}{\trg{\emptyevent}}
-  }{e-n-has-nat}
-  %
-%  \typerule{e-$\trg{\lhast{x}{\natt}}$}{
-%  }{
-%    \pstepto{\rtt{\trg{\statevar}}{\trg{\lhast{x}{\natt}}}}{\rtt{\trg{\statevar}}{\trg{1}}}{\trg{\emptyevent}}
-%  }{e-x-has-nat}
-  %
-  \typerule{e-$\trg{\lhast{n}{\natt\times\natt}}$}{
-  }{
-    \pstepto{\rtt{\trg{\statevar}}{\trg{\lhast{n}{\natt\times\natt}}}}{\rtt{\trg{\statevar}}{\trg{1}}}{\trg{\emptyevent}}
-  }{e-n-has-nattimesnat}
-\end{center}
-
-The $\Ltrg$ language does not provide any guarantees with regards to memory safety.
-The syntax of it extends the unified syntax for all languages (\Cref{subsec:cs:defs}) with pairs of natural numbers (\Cref{tr:e-proj1}) and two different kinds of dynamic checks.
-One of those checks allows programmers to investigate wether a pointer is freed or not (\Cref{tr:e-ispoisoned-yes}).
-The other check is a simple dynamic type check that evaluates to $\trg{0}$ if the check suceeds and to $\trg{1}$ otherwise (\Cref{tr:e-n-has-nat,tr:e-n-has-nattimesnat}).
-The projection $\msfilterL{\trace}$ is entirely similar to $\msfilterLtms{\trace}$.
+\subsection{$\Ltrg$: A Memory--Unsafe Language}\label{subsec:lsms}
+
+$\Ltrg$ extends the syntax presented earlier (\Cref{subsec:cs:defs}) with dynamic typechecks and a way to inspect poison tags in the metadata of pointers.
+If an object a pointer bound to variable $\trg{x}$ points to is allocated ($\trg{\poisonless}$), the check $\trg{\lispoisoned{x}}$ yields $\trg{1}$.
+Likewise, if it has been poisoned ($\trg{\poisoned}$), the check $\trg{\lispoisoned{x}}$ evaluates to $\trg{0}$.
+Besides these checks, the language remains exactly the same.
+
+The projection $\msfilterL{\trg{\trace}}$ is entirely similar to $\msfilterLtms{\src{\trace}}$.
 
@@
 The main purpose of this decision is to enhance readability.
-One thing that is added to $\Lms$ is another projection to the most general trace model for reasoning about \gls{scct}.
-Here, any information irrelevant to \gls{scct} is filtered.
-
-\begin{center}
-  \typerule{$\Lms$-filter-alloc}{
-  }{
-    \scctfilterLms{\irl{\tup{Get\ \loc\ n}{\comp}}} = \emptyevent
-  }{lms-filter-get}
-  %
-  \typerule{$\Lms$-filter-abort}{
-  }{
-    \scctfilterLms{\irl{\lightning}} = \lightning
-  }{lms-filter-abort}
-\end{center}
+The projection $\msfilterLms{\irl{\trace}}$ is also exactly equal to $\msfilterL{\trg{\trace}}$.
 
@@
 
-\begin{gather*}
-  \begin{aligned}
-    \mi{(Expressions)}~\obj{\expr}\bnfdef&\dots \mid \obj{\lwrdoit{\expr}} \mid \obj{\lrddoit{x}{\expr}} &
-    \mi{(States)}~\obj{\statevar}\bnfdef&\obj{\tup{\cfstate}{\sandboxtag,n,\memstate}}
-  \end{aligned}
-\end{gather*}
-
-$\Lscct$ extends the state presented earlier (\Cref{subsec:cs:defs}) with another entry that models the value of a register for a data (operand) independent timing mode.
-To modify the register, $\Lscct$ extends expressions with a way to write a value to the register ($\obj{\lwrdoit{\expr}}$) and a way to read it ($\obj{\lrddoit{x}{\expr}}$).
-Finals carry a secrecy annotation, indicating high ($\obj{\lock}$) or low ($\obj{\unlock}$) secrecy.
-Whenever the register is marked as active, parts of the dynamics of $\Lscct$ do not leak information.
-In general, $\Lscct$ can leak more information than the other languages:
-
-\begin{gather*}
-  \begin{aligned}
-    \mi{(Pre\text{--}Events)}~\obj{\event[_{b}]}&\bnfdef\ \dots \mid \obj{{Branch}\ n} \mid \obj{{Binop}\ n} \mid \obj{\widehat{{Get}}\ \loc\ \valueexpr} \mid \obj{\widehat{{Set}}\ \loc\ \valueexpr\ \valueexpr[']} &
-    \mi{(Events)}~\obj{\event}&\bnfdef\ \obj{\varepsilon} \mid \obj{\lightning} \mid \obj{\tup{\event[_{b}]}{\sandboxtag,\securitytag}}
-  \end{aligned}
-\end{gather*}
-
-Events carry a secrecy tag instead of the data they carry.
-Branches and binary operations leak information about the data they compute on.
-If the data (object) independent timing register is set to be active, the read and write to memory events should not leak.
-However, to emit $\obj{\emptyevent}$ in this case would mean that memory unsafe operations are hidden as well.
-Since this language should not have a mechanism to hide memory unsafety, pre--events are extended to model reading from ($\obj{\widehat{{Get}}\ \loc\ \valueexpr}$) and writing to ($\obj{\widehat{Set}\ \loc\ \valueexpr}$) memory without leaking secret information involved in the access.
-
-\begin{center}
-  \typerule{e-$\obj{\lbinop{}{}}$-leak}{
-    \obj{n_1}\oplus\obj{n_2}=\obj{n_3} &
-    \obj{\securitytag}\sqcap\obj{\securitytag[']}=\obj{\securitytag['']} &
-    \obj{\oplus}\notin\left\{\obj{\times}, \obj{/}, \obj{<}\right\}
-  }{
-    \pstepto{\rtt{\obj{\tup{\cfstate}{\sandboxtag,0,\memstate}}}{\obj{\lbinop{n_1^{\securitytag}}{n_2^{\securitytag[']}}}}}{\rtt{\obj{\tup{\cfstate}{\sandboxtag,0,\memstate}}}{\obj{n_3^{\securitytag['']}}}}{\obj{\tup{Binop\ n_3}{\sandboxtag,\securitytag['']}}}
-  }{e-binop-leak}
-\end{center}
-
-\MKin{filters to spec events}
+$\Lscct$ extends the dynamic state with a register for a data (operand) independent timing mode.
+For the other languages seen earlier, this mode is intuitively always--on.
+To modify the mode, $\Lscct$ extends expressions with the term $\obj{\lwrdoit{\expr}}$ that writes the value of $\obj{\expr}$ into the register and with a term $\obj{\lrddoit{x}{\expr}}$ that reads the value of the register into $\obj{x}$.
+If the register is turned on, no secrets can appear on traces.
+If the register is turned off, secrets may appear on traces as well as new events $\obj{Branch\ n}$ and $\obj{Binop\ n}$, which the semantics emits for conditionals and certain binary operations, e.g., division.
+To prevent leakage of secrets but still enable reasoning about memory safety, pre--events are extended with $\obj{\widehat{{Get}}\ \loc\ \valueexpr}$ and $\obj{\widehat{Set}\ \loc\ \valueexpr}$, which indicate reads from and writes to memory without leaking secret information involved in the access.
+
+The filters to the specification trace model are straightforward.
 

### 2023-07-07 popl/main.tex
Feedback

@@
 
-$\Lscct$ extends the dynamic state with a register for a data (operand) independent timing mode.
+$\Lscct$ extends the dynamic state of $\trg{L}$ (\Cref{subsec:lms}) with a register for a data (operand) independent timing mode.
 For the other languages seen earlier, this mode is intuitively always--on.
@@
 This section defines several secure compilers, each of which robustly preserves a different property of interest (\Cref{fig:pipeline}).
-The power of the framework (\Cref{sec:compprop,sec:compcomp}) is demonstrated by composing these compilers for a secure, but optimizing, compilation chain that robustly preserves \gls{ms} and \gls{cct}.
-Concretely, the section starts with a base-line definition for the used languages (\Cref{subsec:cs:defs}) and extends that to define languages $\src{L_{TMS}}$ and $\trg{L}$.
-For programmers using $\src{L_{TMS}}$, their mental model should be that their code is temporal memory safe and cryptographic constant time.
-The former is proven to be enforced by the static semantics of $\src{L_{TMS}}$, where each well-typed program is temporal memory safe.
-The latter is completely untyped.
-Having languages defined, this section continues with a compiler $\cc{\src{L_{\tmssafe}}}{\trg{L}}$ and proves that $\cc{\src{L_{\tmssafe}}}{\trg{L}}$ robustly preserves \gls{tms} (\Cref{subsec:cs:tms}).
-For readability purposes, it defines $\irl{L_{MS}}$ to be exactly the same as $\trg{L}$ and continues to present a compiler $\cc{\trg{L}}{\irl{L_{MS}}}$ that robustly preserves \gls{sms} (\Cref{subsec:cs:ms}).
-Now leaveraging the power of the framework that this paper introduces (\Cref{sec:compcomp,sec:compprop}), a compiler $\rtp{\cc{\src{L_{TMS}}}{\irl{L_{MS}}}}{\mssafe}$ is attained.
-The next steps in the compilation pipeline are two secure compilers $\cc[_{\gls{dce}}]{\irl{L_{MS}}}{\irl{L_{MS}}}$ and $\cc[_{\gls{cf}}]{\irl{L_{MS}}}{\irl{L_{MS}}}$ that perform \glsfirst{dce} and \glsfirst{cf}, respectively, to model an optimzing pipeline happening somewhere in the compilation chain (\Cref{subsec:cs:opts}).
-Finally, this section defines the language $\obj{L_{\scctsafe}}$ that is, in contrast to the other languages, not cryptographic constant time by definition.
-Concluding, a compiler $\cc{\irl{L_{MS}}}{\obj{L_{CCT}}}$ that robustly preserves \gls{cct} (\Cref{subsec:cs:scct}) is shown and composed with $\cc{\src{L_{TMS}}}{\irl{L_{MS}}}$ and the optimizing compilers $\cc[_{\gls{dce}}]{\irl{L_{MS}}}{\irl{L_{MS}}}$ and $\cc[_{\gls{cf}}]{\irl{L_{MS}}}{\irl{L_{MS}}}$.
-In turn, this yields $\cc{\src{L_{TMS}}}{\obj{L_{CCT}}}$, a compiler that robustly preserves a combination of \gls{ms} and \gls{cct}.
+It demonstrates the power of the framework (\Cref{sec:compprop,sec:compcomp}) by composing these compilers for a secure and optimizing compilation chain that robustly preserves \gls{msscct}.
+The first step in this chain is the compiler from $\src{L_{\tmssafe}}$ to $\trg{L}$ that robustly preserves just \gls{tms} (\Cref{thm:cca:rtp:tms}).
+From here, an instrumentation from $\trg{L}$ to $\irl{L_{\mssafe}}$ ensures that no out-of-bounds accesses can happen and, thus, programs at this point attain \gls{sms} (\Cref{thm:ccb:rtp:sms}).
+This is enough to get a compiler that robustly preserves \gls{ms} (\Cref{thm:ccab:rtp:ms}).
+At this stage, the section presents two optimizing translations, namely \gls{cf} and \gls{dce}, each of which robustly preserves \gls{ms} (\Cref{thm:ccdce:rtp:ms,thm:cccf:rtp:ms}).
+These translations can be freely ordered in the compilation chain without compromising memory safety (\Cref{thm:cccfccdce:rtp:ms}).
+The last step ensures that code stays \gls{scct} (\Cref{thm:ccscct:rtp:scct}) which yields the final result that the whole compilation chain robustly preserves \gls{msscct} (\Cref{thm:ccall:rtp:msscct}).
 
@@
 
-\begin{center}
   $$
@@
   $$
-\end{center}
-
-The only case where the compiler does something interesting is for top-level functions.
-Here, a dynamic typecheck is inserted to respect the calling-convention.
-Since $\trg{L}$ has no static typechecks, it could happen that a bogus context $\trg{\library_{\ctx}}$ calls a symbol accepting a $\src{\natt}$ with $\trg{\lpair{17}{29}}$.
-The compiler ensures that execution does not proceed in cases like that.
-This paper considers symbols on $\src{\natt}$.
+
+The compiler does not have to do anything special besides a typecheck to respect the calling--convention.
+That is, because $\trg{L}$ has no static typechecks, it could happen that a bogus context $\trg{\library_{\ctx}}$ invokes a callable object accepting a $\src{\natt}$ with $\trg{\lpair{17}{29}}$.
+By inserting the check, the compiler ensures that execution does not proceed in such cases.
+The compiler does not insert other checks and simply recolors $\src{L_{\tmssafe}}$ to $\trg{L}$ expressions.
 
@@
 \end{theorem}
+
+\begin{proof}[Proof Sketch]
+The proof of \Cref{thm:cca:rtp:tms} uses standard techniques in secure compilation literature and is visualized in \Cref{fig:proofdiag:rtp}.
+Concretely, it uses a trace--based backtranslation to reconstruct an $\src{L_{\tmssafe}}$ context from the $\trg{L}$ trace, which, however, behaves similar to the $\trg{L}$--level context.
+\Cref{thm:wt:tms} already gives strong guarantees for programs written in $\src{L_{\tmssafe}}$ and can be exploited for the backtranslated context by ensuring that it typechecks.
+Well--typing means the $\src{L_{\tmssafe}}$ program runs, so the rest of the proof is about relating the $\trg{L}$--level execution with the $\src{L_{\tmssafe}}$--level execution.
+
+For the $\trg{L}$--execution, the compiler wrapper code at state $\trg{\Omega_{1}}$ may either crash the program or continue execution $\trg{\Omega_{w_{1}}}$ right before control is transferred to the component.
+In this specific case, the compiler need not to insert additional code after the component is done with its execution, but a symmetric situation can happen here.
+Similarily, the backtranslation may also insert wrapper code prior and after switching to and from the component to the context.
+\MKin{
+  more detail or not? ;
+  i.e., explain $\multimap$ and $\approx$?
+}
+\end{proof}
 
@@
       \node[align=left,below right=0.3 and 0.3 of trgempty,font=\tiny] (legend) {%
-        ${\src{\trace[_c]}}\cong{\backt{\Ltms}{\Ltrg}(\trg{Call\ ?\finalexpr_i})}$\\%
-        ${\src{\trace[_r]}}\cong{\backt{\Ltms}{\Ltrg}(\trg{Ret\ !\finalexpr_o})}$
+        ${\src{\trace[_c]}}\cong{\backt{\Ltms}{\Ltrg}(\trg{Call\ ?\ foo\ \valueexpr})}$\\%
+        ${\src{\trace[_r]}}\cong{\backt{\Ltms}{\Ltrg}(\trg{Ret\ !\ foo\ \valueexpr})}$
         };
@@
 
-In this section, the language $\irl{L_{\mssafe}}$ is exactly the same as $\trg{L}$ and are considered equal.
+The language $\irl{L_{\mssafe}}$ is exactly the same as $\trg{L}$ and are considered equal.
 However, it is still named and collored differently to emphasize the fact that after this section, compiled $\irl{L_{\mssafe}}$ programs are \gls{ms}.

### 2023-07-08 popl/main.tex
RelWork MemSafety

@@
 This section compares existing work with the one presented in this paper.
-To this end, work on robust preservation criteria (\Cref{subsec:relw:seccomprtp}) and on other secure compiler criteria (\Cref{subsec:relw:seccompcrit}) is being looked at.
+To this end, work on robust preservation (\Cref{subsec:relw:seccomprtp}) and on other secure compiler criteria (\Cref{subsec:relw:seccompcrit}) is looked at.
 The case study of this paper (\Cref{sec:casestud:defs,sec:casestud:rtp}) implements measures for ensuring \gls{ms} as well as \gls{cct} and these are compared with previous work as well (\Cref{subsec:relw:msmechs,subsec:relw:cctmechs}).
@@
 The authors demonstrate a similar compositionality theorem to what is presented here (\Cref{sec:compcomp}) as well as in an earlier version of this work~\cite{kruse2022csc}.
-However, they do not demonstrate the scalability of the approach in the context of a compilation chain.
+However, they do not demonstrate the scalability of the approach by means of a case study.
 
@@
 
-There is still no consensus what the formal definition of memory safety as a program property should be. %cite
-Yet, different mechanisms for memory safety exist that also consider the secure compilation domain, i.e., have an attacker model that is not just the observing attacker.
+Different mechanisms for memory safety exist that also consider the secure compilation domain, i.e., have an active attacker model.
 For example, the ,,pointers as capabilities'' principle represents pointers as machine--level capabilities~\cite{korashy2021capableptrs}, which behave in a similar fashion to capabilities by means of linear typing~\cite{morrisett2005L3}.
-The approach of this paper to memory safety and the countermeasures against memory safety insecurities are inspired by seminal work in the literature~\cite{nagarakatte2009soft,nagarakatte2010cets}.
-The view of memory safety in this paper exhibits the classic split between temporal and spatial memory safety.
-The former making sure that there are no double deallocations or use after frees.
-The latter making sure that there are no out of bounds accesses.
-When extending the languages in this paper with a less restricted form of pointer arithmetic, the region coloring memory safety monitor presented in earlier work~\cite{michael2023mswasm} can be used.
-The work presenting this monitor provides an approach for the robust {\em enforcement}, which implies robust preservation, of memory safety.
-But, they do not compose the enforcement mechanism with other secure compilation passes.
+The approach of this paper also uses linear typing, but differs from $L^{3}$~\cite{morrisett2005L3} in the way that functions are not first--class.
+Moreover, this paper considers an active attacker, while the work on $L^{3}$ only discusses whole programs and, thus, has no active attacker model.
+The instrumentation to ensure memory safety that this paper presents is inspired by seminal work in the literature~\cite{nagarakatte2009soft}.
+That work inserts bounds--checks in front of pointer--dereferences and, for this to work, inserts meta--data information on pointer creation.
+It also works in a more advanced setting with structured fields accesses and also introduces a table--lookup for pointers that are stored in memory.
+This paper only considers pointers to arrays of simple objects, i.e., there are no pointers to pointers or structures.
+Of course, there are a lot of other approaches to memory--safety in literature, specifically as compiler instrumentations~\cite{akritidis2009baggy,younan2010paricheck,jung2021pico,shankaranarayana2023tailcheck,dhumbumroong2020boundwarden,nam2019framer,zhou2023fatptrs}, hardware--extensions~\cite{kwon2013lowfat,saileshwar2022heapcheck,chen2023flexpointer,kim2023whistle}, or programming language extensions~\cite{elliott2018checkedc,li2022formalcheckedc,jim2002cyclone,elliott2015guilt,west2005cuckoo,weis2019fyr,benoit2019uniqueness}, to name a few.
+What differentiates this work from them is that this work uses known, compiler--based approaches to ensure memory--safety as a means to investigate secure compiler compositions.
+This paper does not provide efficient memory--safety, but serves as a theoretical foundation for the secure compilation domain.
+
+To extend the languages in this paper with a less restricted form of pointer arithmetic, the region coloring memory safety monitor presented in earlier work~\cite{michael2023mswasm} can be used.
+The work presenting this monitor provides an approach for the robust preservation of memory safety compiling from C to WASM.
+However, they do not discuss composition of secure compilers but rather investigate an instance of a secure compiler.
 

### 2023-07-08 popl/main.tex
RelWork CCT

@@
 
+The approach to cryptographic constant time in this paper is high--level, where a programming language exposes a way to switch the semantics to a data (operand) independent timing mode.
+Since identifiers in $\Lscct$ are annotated with a secrecy tag, this approach is similar to others with information flow control.
+For example, Vale~\cite{bond2017vale} uses Dafny to ensure constant--time assembly code, while Jasmin~\cite{almeida2017jasmin} makes use of the Coq proof assistant to reject non--constant--time programs.
+CT-Wasm~\cite{watt2019ctwasm} enforces constant--timeness by means of a type system.
+Different to the approach of this paper, these approaches necessitate that the programmer writes \gls{cct} code.
+An approach to allow programmers to write more high--level code is CryptOpt~\cite{kuepper2023cryptopt}, which generates efficient target-code by means of a randomized search.
+This paper abstracts over concrete mitigation strategies and simply assumes that there is a flag to switch to a cryptographic--constant time execution mode.
+This can be realized by employing the FaCT~\cite{cauligi2019fact} compiler, which translates common non-constant time code patterns to be constant--time, and the data (object) independent timing execution mode of modern processors.
+
 \section{Conclusion\pages{$\sfrac{1}{2}$}}\label{sec:concl}

### 2023-07-08 popl/main.tex
Finish up RelWork

@@
         To this end, it presents a compilation chain consisting of several passes.
-        %The sequential composition of these compilers yields a compilation chain that is optimizing and robustly preserving a combination of \gls{ms} and \gls{cct}.
+        %The sequential composition of these compilers yields a compilation chain that is optimising and robustly preserving a combination of \gls{ms} and \gls{cct}.
         The chain ultimately preserves a combination of \gls{ms} and \gls{cct} by means of composing the individual, secure passes concerning \gls{tms}, \gls{sms}, and \gls{scct}, respectively.
-        %As for optimizations, the chain has two passes, one performing \gls{cf} and the other \gls{dce}.
-        Furthermore, the chain includes two passes for \gls{ms}\MKin{maybe patch \gls{ms} when optimizing passes are done} that are optimizing: One performs a simple \gls{dce} and the other \gls{cf}.
+        %As for optimisations, the chain has two passes, one performing \gls{cf} and the other \gls{dce}.
+        Furthermore, the chain includes two passes for \gls{ms}\MKin{maybe patch \gls{ms} when optimising passes are done} that are optimising: One performs a simple \gls{dce} and the other \gls{cf}.
         %This compilation chain demonstrates the ability to perform modular proof engineering by leaveraging this paper's compositionality theorems.
@@
 
-Notably, real-world compilation chains also perform a series of passes whose main purpose is not necessarily to translate from one language to another, but to, e.g., optimize the code or enforce a certain property.
+Notably, real-world compilation chains also perform a series of passes whose main purpose is not necessarily to translate from one language to another, but to, e.g., optimise the code or enforce a certain property.
 Both examples can be seen in the wild, there are countless instances of them.\MK{cit}
@@
 However, this costs additional time until compilation is done, which can be a valuable resource\MK{cit}.
-Hence, compiler engineers typically try to find an order of optimizations that yields well-optimized programs as fast as possible.
+Hence, compiler engineers typically try to find an order of optimisations that yields well-optimised programs as fast as possible.
 The following corollary justifies that any such order of compilation passes is valid with respect to security.
@@
   \end{center}
-  \caption{Example program where the level of optimizations differ for one pass of applying \gls{cf} and \gls{dce} in any order. %
+  \caption{Example program where the level of optimisations differ for one pass of applying \gls{cf} and \gls{dce} in any order. %
   Every edge is a compilation pass and the label on the edge states what the pass does, i.e., \gls{cf} or \gls{dce}. %
-  The source code in the nodes is a glorified compiler intermediate representation and the code gets more optimized towards the righthandside of the figure.}\label{fig:cfdceex}
+  The source code in the nodes is a glorified compiler intermediate representation and the code gets more optimised towards the righthandside of the figure.}\label{fig:cfdceex}
 \end{figure}
@@
   \end{center}
-  \caption{Visualisation of the optimizing compilation pipeline that attains a combination of \gls{ms} and \gls{cct}.}\label{fig:pipeline}
+  \caption{Visualisation of the optimising compilation pipeline that attains a combination of \gls{ms} and \gls{cct}.}\label{fig:pipeline}
 \end{figure}
 This section defines several secure compilers, each of which robustly preserves a different property of interest (\Cref{fig:pipeline}).
-It demonstrates the power of the framework (\Cref{sec:compprop,sec:compcomp}) by composing these compilers for a secure and optimizing compilation chain that robustly preserves \gls{msscct}.
+It demonstrates the power of the framework (\Cref{sec:compprop,sec:compcomp}) by composing these compilers for a secure and optimising compilation chain that robustly preserves \gls{msscct}.
 The first step in this chain is the compiler from $\src{L_{\tmssafe}}$ to $\trg{L}$ that robustly preserves just \gls{tms} (\Cref{thm:cca:rtp:tms}).
@@
 This is enough to get a compiler that robustly preserves \gls{ms} (\Cref{thm:ccab:rtp:ms}).
-At this stage, the section presents two optimizing translations, namely \gls{cf} and \gls{dce}, each of which robustly preserves \gls{ms} (\Cref{thm:ccdce:rtp:ms,thm:cccf:rtp:ms}).
+At this stage, the section presents two optimising translations, namely \gls{cf} and \gls{dce}, each of which robustly preserves \gls{ms} (\Cref{thm:ccdce:rtp:ms,thm:cccf:rtp:ms}).
 These translations can be freely ordered in the compilation chain without compromising memory safety (\Cref{thm:cccfccdce:rtp:ms}).
@@
 
-\subsection{Optimizing Compilers}\label{subsec:cs:opts}
+\subsection{Optimising Compilers}\label{subsec:cs:opts}
 
@@
 
-\MKin{category theory stuff}
-\MKin{full abstraction -> not compositional}
-\MKin{compositional compiler correctness?}
+While this paper focuses on the robust preservation framework~\cite{abate2019jour}, other secure compilation criteria exist.
+The survey on formal approaches to secure compilation~\cite{patrignani2019survey} discusses a broad spectrum already, while this section presents a very high--level overview.
+The full abstraction approach~\cite{abadi1999fullabstraction} states that a compiler should preserve and reflect observational equivalence between source and target programs.
+It was shown~\cite{abate2021faandrc} that fully abstract compilers robustly preserve program properties that are either trivial or meaningless.
+As a mitigation for this, the authors presented a categorical approach based on maps of distributive laws~\cite{watanabe2002modl}, which they call many maps of distributive laws.
+Maps of distributive laws have been investigated before as a possible secure compilation criterion~\cite{tsampas2020catsc}.
+Other approaches are extensions of the compiler correctness criterion as discussed in other work~\cite{patterson2019next700} or the introduction of opaque observations~\cite{vu2021reconciling} to reconcile compiler optimisations with security.
+Note that this work also presents secure compilers that are optimising, but contrary to the other~\cite{vu2021reconciling}, provides a formal account of these in the robust preservation framework.
 
@@
 This paper does a first step towards practical secure compilation chains by introducing a theoretical framework demonstrating that secure compilers compose.
-The paper exercises this framework on a case-study that consists of an optimizing compilation chain that is secure with respect to \gls{msscct}.
+The paper exercises this framework on a case-study that consists of an optimising compilation chain that is secure with respect to \gls{msscct}.
 In future work, it would be interesting to investigate wether it is possible to provide {\em secure compiler combinators}, similarily to how parser combinators work.

### 2023-07-08 popl/main.tex
Add cits

@@
         %\Cref{sec:compprop} presents a formal framework to reason about compositions of security properties and demonstrate its usage with concrete examples:
-        First, this paper studies the formalisation and the composition of security properties (\Cref{sec:compprop}) by focussing on those safety properties that are of interest for real-world compiler writers (as identified by the plethora of work enforcing such properties individually~\MP{cits}).
+        First, this paper studies the formalisation and the composition of security properties (\Cref{sec:compprop}) by focussing on those safety properties that are of interest for real-world compiler writers (as identified by the plethora of work enforcing such properties individually~\cite{akritidis2009baggy,nagarakatte2009soft,nagarakatte2010cets,dhumbumroong2020boundwarden,jung2021pico,nam2019framer,shankaranarayana2023tailcheck,younan2010paricheck,zhou2023fatptrs,bond2017vale,almeida2017jasmin,kuepper2023cryptopt,cauligi2019fact}\MK{even more cits?}).
         % 
@@
 
-Properties $\pi$ are sets of traces of admissible program behaviours, ascribing what said property considers valid.\MP{ cit }
+Properties $\pi$ are sets of traces of admissible program behaviours, ascribing what said property considers valid~\cite{clarkson2008hyper}.
 %
@@
 %Properties and partitioned in the subsets of safety and liveness\MP{cit}, and given their central role for security, this paper focusses on safety properties.
-The set of all properties can be partitioned into different {\em classes} ($\class$), i.e., safety, liveness, and neither safety or liveness.\MP{cit}
+The set of all properties can be partitioned into different {\em classes} ($\class$), i.e., safety, liveness, and neither safety or liveness~\cite{clarkson2008hyper}.
 Given the central role for security, this paper focusses on safety properties.
@@
 %That is, the compiler encodes any possible $\src{L}$--level feature in $\trg{L}$, such that the semantics are the same.
-Unfortunately, correct compilers may be insecure compilers\MK{citation needed}.
+Unfortunately, correct compilers may be insecure compilers~\cite{patrignani2019survey}.
 Programs translated by insecure compilers can violate security properties that the programmer assumes to hold. %, given the presence of a context--level program.
@@
 A main result is that secure compilers in the robust preservation framework~\cite{abate2019jour} compose.
-This is not intuitive in the sense that there are countless examples in the security domain where composition does not work.\MK{cite}
+This is not intuitive in the sense that there are a number of examples in the security domain where composition does not work~\cite{canetti2006univcomp,mccullough2012compo,fabian2022automatic}.
 For starters, the sequential composition of compilers $\cc{\src{L}}{\trg{L}}$ and $\cc{\trg{L}}{\obj{L}}$ is as expected, i.e., given an $\src{L}$ program $\src{p}$, obtain its compiled $\obj{L}$ program by plugging $\src{p}$ into $\cc{\src{L}}{\trg{L}}\circ\cc{\trg{L}}{\obj{L}}$.
@@
 Pretending that the following compilers are secure, consider the compilation chain for $\src{TypeScript}$.
-First, $\src{TypeScript}$ programs are translated to $\trg{JavaScript}$ which, e.g., V8\MK{cit} eventually compiles in parts to $\irl{Ignition BC}$.
+First, $\src{TypeScript}$ programs are translated to $\trg{JavaScript}$ which, e.g., V8~\cite{googlev8} eventually compiles in parts to $\irl{Ignition BC}$.
 Were all these compilation steps robustly secure with respect to \gls{ms}, the resulting $\irl{Ignition BC}$ code would be \gls{ms} regardless of the context the binary runs in.
@@
 Notably, real-world compilation chains also perform a series of passes whose main purpose is not necessarily to translate from one language to another, but to, e.g., optimise the code or enforce a certain property.
-Both examples can be seen in the wild, there are countless instances of them.\MK{cit}
-Consider the compilation passes of LLVM\MK{cit} whose source-code optimisations are industry--competitive.
+Both examples can be seen in the wild~\cite{nagarakatte2009soft,nagarakatte2010cets,akritidis2009baggy,wegman1991ccp,manjikian1997fusion}, it is infeasible to list them all.
+Consider the compilation passes of LLVM\cite{lattner2004llvm} whose source-code optimisations are industry--competitive.
 In LLVM, the pass manager can be used to schedule passes in a way where the optimisations yield the best results.
@@
 This {\em phase ordering problem} is well--known in literature and a practical solution is to simply perform a fixpoint iteration.
-However, this costs additional time until compilation is done, which can be a valuable resource\MK{cit}.
-Hence, compiler engineers typically try to find an order of optimisations that yields well-optimised programs as fast as possible.
+Compiler engineers typically try to find an order of optimisations that yields well--optimised programs for either code size~\cite{cooper1999geneticphases} or performance~\cite{kulkarni2006exhaustivephase}.
 The following corollary justifies that any such order of compilation passes is valid with respect to security.
@@
 Given a secure compiler from $\src{Java}$ to $\trg{JVM BC}$ and one from $\obj{Kotlin}$ to $\trg{JVM BC}$, it is easy to envision that one can engineer a combined compiler that accepts both $\src{Java}$ and $\obj{Kotlin}$.
-A practical example of this, which is, however, insecure, is {\tt Android Studio}\MK{cit}.
+A practical example of this, which is, however, insecure, is {\tt Android Studio}~\cite{androidstudio}.
 The proof of \Cref{thm:rtp} can be easily adapted to also works for this {\em upper} composition.
@@
 Note that this work also presents secure compilers that are optimising, but contrary to the other~\cite{vu2021reconciling}, provides a formal account of these in the robust preservation framework.
+Lastly, the authors of this paper have presented ongoing work~\cite{patrignani2023blame} on a weaker robust preservation criteria based on the concept blame.
 

### 2023-07-08 popl/main.tex
comments on 2.1 and 2.2

--- 2023-07-08 popl/main.tex
@@
 \footnote{
-Throughout the paper, we denote empty sequences with $\hole{\cdot}$ and concatenate sequences $\trace[_{1}],\trace[_{2}]$ as $\trace[_{1}]\cdot\trace[_{2}]$.
+Throughout the paper, we indicate sequences with an overbar (i.e., $\trace$), we denote empty sequences with $\hole{\cdot}$ and concatenate sequences $\trace[_{1}],\trace[_{2}]$ as $\trace[_{1}]\cdot\trace[_{2}]$.
 Prepending elements to sequences uses the same notation: $\event\cdot\trace$.
@@
 %Properties and partitioned in the subsets of safety and liveness\MP{cit}, and given their central role for security, this paper focusses on safety properties.
-The set of all properties can be partitioned into different {\em classes} ($\class$), i.e., safety, liveness, and neither safety or liveness.\MP{cit}
-Given the central role for security, this paper focusses on safety properties.
-Note that there are other classes $\class$ that contain more domain-specific properties, such as the class of memory safety properties.
+The set of all properties can be partitioned into different {\em classes} ($\class$), i.e., safety, liveness, and neither safety nor liveness.\MP{cit}
+Given their central role for security, this paper focusses on safety properties.
+% Note that there are other classes $\class$ that contain more domain-specific properties, such as the class of memory safety properties.
+\MPin{
+	remove.
+	what does this sentence add? removing it does not remove any relevant information
+}
 For the class of safety properties, it is decidable wether a trace admits such a property with just a finite trace prefix.
-As an example, consider a trace that starts with an interaction with memory where a deallocation precedes a read: $\ev{Dealloc\ \loc}\cdot\ev{Read\ \loc\ 1729}\cdot\dots$
-This program behaviour is insecure, because it reads from a memory location that was freed already.
-The finite trace prefix $\ev{Dealloc\ \loc}\cdot\ev{Read\ \loc\ 1729}$ is enough to decide that the trace does not satisfy the property ,,no use after dealloc'' and there is no way to append events to this prefix which would result in the trace being admissible.
+\MPin{
+	passive. also, ``a trace admits a property'' is wrong, it's the other way around
+}
+As an example, consider a trace describing an interaction with a memory where the deallocation of a pointer $\loc$ precedes its read: $\ev{Dealloc\ \loc}\cdot\ev{Read\ \loc\ 1729}\cdot\dots$
+This program behaviour is insecure wrt a canonical notion of (temporal) memory safety dictating no use-after-frees of pointers\MP{cit}, because it reads from a memory location that was freed already.
+\MPin{
+	rewordeda a little for precision (e.g., what is $\loc?$), check the diff
+}
+The previous finite trace prefix is enough to decide that the trace does not satisfy temporal memory safety and there is no way to append events to this prefix which would result in the trace being admissible.
 %Given a safety property, a finite prefix of a longer trace $\trace$ is enough to decide whether a trace is admitted or not by the property.
@@
 %This set is used to define property satisfaction (\Cref{def:propsat}):
-\bul{Programs $\progvar$ satisfy a property $\pi$} iff \iul{$\progvar$ yields a trace $\trace$} such that \oul{$\trace$ is admissible with respect to $\pi$}.
+\MPin{
+	fill the blank of the next sentence.
+}
+In the following we write that a whole program \MP{} runs according to the language semantics and produces trace \MP{} as \MP{}. 
+With this, we define that
+\bul{programs $\progvar$ satisfy a property $\pi$} iff \iul{$\progvar$ yields a trace $\trace$} such that \oul{$\trace$ is admissible with respect to $\pi$} (\Cref{def:propsat}).
 
 \begin{definition}[Property Satisfaction]\label{def:propsat}
-  $\progvar$ satisfies $\pi$, \bul{$\sat{\progvar}{\pi}$}, iff \iul{running $\progvar$ gives $\trace$} where \oul{$\trace\in\pi$}.
+  % $\progvar$ satisfies $\pi$, 
+  \bul{$\sat{\progvar}{\pi}$} 
+  % iff 
+  $\isdef$
+  \iul{running $\progvar$ gives $\trace$}
+  \MP{this does not formalise things, see sentence added above and add formalism}
+   where \oul{$\trace\in\pi$}.
 \end{definition}
 
-Given a partial program $\progvar$, i.e., a program with free variables, the execution of $\progvar$ would get stuck due to the lack of definitions.
+\MPin{
+	why are we suddenly going from talking about prop sat to talknig about progs and ctxs?
+	there's a gap here that needs to be bridged.
+	A plausible solution is something along these lines.
+	"property satisfaction only considers passive attackers, to consider a stronger model with active attackers existing work (as well as this paper) uses robust property satisfaction, which we define after introducing the required concept of program context and partial programs"
+}
+Given a partial program $\progvar$, i.e., a program with free variables, the execution of $\progvar$ would get stuck due to the lack of program parts (e.g., function definitions).
 %As described earlier, partial programs cannot be executed \MP{not true anymore, need to define the notion of partial programs} , because they lack definitions and, thus, get stuck.
 To remedy this, {\em linking} takes two partial programs $\progvar[_{1}],\progvar[_{2}]$ and produces a new whole program $\link{\progvar[_{1}]}{\progvar[_{2}]}$.
+\MPin{
+	cumbersome notation.
+	`link ... ... ' should be used to identify the notation of the operation, but we should have a metavariable for the result, like `link(p1,p2) = w'
+	W is a known candidate
+}
 %Note that linking is not defined on programs, but merely on their syntactic descriptions, since programs contain dynamic state.\MP{ not terminology i really want to use, nor is defined}
-As typically done in secure compilation~\cite{abate2019jour}, \MP{others do this too, parametricity results e.g.} this paper assumes that programs are the result of linking a {\em context} ($\ctx$) and a {\em component} ($\comp$).
+As typically done in secure compilation~\cite{abate2019jour}, \MP{others do this too, parametricity results e.g.} this paper assumes that \MP{whole?}programs are the result of linking a {\em context} ($\ctx$) and a {\em component} ($\comp$).
+\MPin{
+	these sentences identify the need to have 2 different terms: 1 for partial programs (components?) one for whole programs (just programs) and to use them consistently.
+}
 % \MP{just introduce the notion of contexts}
@@
 %The component part of a whole program is the main object of interest and thus a known syntactic description.
-The context is usually arbitrary\footnote{Up to, e.g., provoking link failure by giving multiple definitiions for the same name.} and thus has the role of an {\em attacker} that can interact with the component by means of whatever features the programming language has.
-The semantics of the programming language is expected to differentiate between the component and the context.
+The context is an arbitrary program and thus has the role of an {\em attacker} that can interact with the component by means of whatever features the programming language has.
+In this work, the semantics of the programming language is expected to differentiate between the component and the context.
+\MPin{
+	so what?
+}
 %\MP{and say the semantics is expected (passive, hah) to differentiate between prg and ctx}
@@
 So, the context interacts in some way with the component and this interaction models an attacker.
-With this, \Thmref{def:propsat} can be extended as follows.
-For \bul{programs $\progvar$ to robustly satisfy a property $\pi$}, take an \iul{attacker $\contextvar$ and link it with $\progvar$}, \oul{the result of which must satisfy $\pi$}.
+\MP{
+	wut?
+	needless sentence, erase.
+}
+With this, \Thmref{def:propsat} can be extended as follows: for \bul{ partialprograms $\progvar$ to robustly satisfy a property $\pi$}, take an \iul{attacker $\contextvar$ and link it with $\progvar$}, \oul{the result of which must satisfy $\pi$}.
+\MPin{
+	the behaviour of the result?
+}
 %The following definition extends on \Thmref{def:propsat} to cover property satisfaction under the presence of an (active, with respect to the semantics of the programming language) attacker: %, which is known as {\em robust satisfaction}:
@@
 \begin{definition}[Robust Satisfaction]\label{def:proprsat}
-  $\progvar$ robustly satisfies $\pi$, \bul{$\rsat{\progvar}{\pi}$}, iff for \iul{any partial program $\contextvar$} it holds that \oul{$\sat{\link{\contextvar}{\runtimetermvar}}{\pi}$}.
+  % $\progvar$ robustly satisfies $\pi$, 
+  \bul{$\rsat{\progvar}{\pi}$}
+  % iff 
+  $\isdef$ for \iul{any partial program $\contextvar$}\MP{use $\forall$} it holds that \oul{$\sat{\ {\contextvar}{\runtimetermvar}}{\pi}$}. \MP{why r?}
 \end{definition}
@@
 This translation is considered {\em correct} if it is semantics--preserving.
+\MPin{
+	i.e., ... brief desc?
+}
 %That is, the compiler encodes any possible $\src{L}$--level feature in $\trg{L}$, such that the semantics are the same.
-Unfortunately, correct compilers may be insecure compilers\MK{citation needed}.
-Programs translated by insecure compilers can violate security properties that the programmer assumes to hold. %, given the presence of a context--level program.
-For a compiler to be secure, this paper makes use of the robust compilation framework~\cite{abate2019jour}, which we summarize in the following definition.
-For \bul{compilers $\cc{\src{L}}{\trg{L}}$ to robustly preserve a class of properties $\class$}, consider \rul{properties $\pi$ of that class $\class$ and programs $\src{p}$ written in $\src{L}$} such that \iul{the program $\src{\progvar}$ robustly satisfies $\pi$}.
-Then, \oul{$\cc{\src{L}}{\trg{L}}\left(\src{p}\right)$, the compiled version of $\src{\progvar}$, must robustly satisfy $\pi$}.
+Unfortunately, correct compilers may be insecure compilers\MK{citation needed}, and programs translated by insecure compilers can violate security properties that the programmer assumes to hold.
+%, given the presence of a context--level program.
+To define when a compiler is secure, this paper uses the robust compilation framework~\cite{abate2019jour}, which we summarize in the following definition.
+
+To define robust compilation we only need the notion of a class of properties.
+\MPin{define.}
+
+For \bul{compilers $\cc{\src{L}}{\trg{L}}$ to robustly preserve a class of properties $\class$}, if for any \rul{property $\pi$ of that class $\class$ and programs $\src{p}$ written in $\src{L}$} we have that \iul{the program $\src{\progvar}$ robustly satisfies $\pi$}, then \oul{the compilation of $\src{\progvar}$, $\cc{\src{L}}{\trg{L}}\left(\src{p}\right)$, must robustly satisfy $\pi$}.
 
 \begin{definition}[Robust Preservation]\label{def:rtp}
-  Compiler $\cc{\src{L}}{\trg{L}}$ robustly preserves $\class$, \bul{$\rtp{\cc{\src{L}}{\trg{L}}}{\class}$}, iff \rul{for any $\pi\in\class$ and $\src{p}$} such that \iul{$\rsat{\src{\progvar}}{\pi}$}, it holds that \oul{$\rsat{\cc{\src{L}}{\trg{L}}\left(\src{p}\right)}{\pi}$}.
+  % Compiler $\cc{\src{L}}{\trg{L}}$ robustly preserves $\class$, 
+  \bul{$\rtp{\cc{\src{L}}{\trg{L}}}{\class}$}
+  %, iff 
+  $\isdef$
+  \rul{for any $\pi\in\class$ and $\src{p}$} such that \iul{$\rsat{\src{\progvar}}{\pi}$}\MP{ use $\forall$}, it holds that \oul{$\rsat{\cc{\src{L}}{\trg{L}}\left(\src{p}\right)}{\pi}$}.
 \end{definition}
@@
 This section compares existing work with the one presented in this paper.
-To this end, work on robust preservation (\Cref{subsec:relw:seccomprtp}) and on other secure compiler criteria (\Cref{subsec:relw:seccompcrit}) is looked at.
+To this end, work on robust preservation criteria (\Cref{subsec:relw:seccomprtp}) and on other secure compiler criteria (\Cref{subsec:relw:seccompcrit}) is being looked at.
 The case study of this paper (\Cref{sec:casestud:defs,sec:casestud:rtp}) implements measures for ensuring \gls{ms} as well as \gls{cct} and these are compared with previous work as well (\Cref{subsec:relw:msmechs,subsec:relw:cctmechs}).
@@
 The authors demonstrate a similar compositionality theorem to what is presented here (\Cref{sec:compcomp}) as well as in an earlier version of this work~\cite{kruse2022csc}.
-However, they do not demonstrate the scalability of the approach by means of a case study.
+However, they do not demonstrate the scalability of the approach in the context of a compilation chain.
 
@@
 
-Different mechanisms for memory safety exist that also consider the secure compilation domain, i.e., have an active attacker model.
+There is still no consensus what the formal definition of memory safety as a program property should be. %cite
+Yet, different mechanisms for memory safety exist that also consider the secure compilation domain, i.e., have an attacker model that is not just the observing attacker.
 For example, the ,,pointers as capabilities'' principle represents pointers as machine--level capabilities~\cite{korashy2021capableptrs}, which behave in a similar fashion to capabilities by means of  linear typing~\cite{morrisett2005L3}.
-The approach of this paper also uses linear typing, but differs from $L^{3}$~\cite{morrisett2005L3} in the way that functions are not first--class.
-Moreover, this paper considers an active attacker, while the work on $L^{3}$ only discusses whole programs and, thus, has no active attacker model.
-The instrumentation to ensure memory safety that this paper presents is inspired by seminal work in the literature~\cite{nagarakatte2009soft}.
-That work inserts bounds--checks in front of pointer--dereferences and, for this to work, inserts meta--data information on pointer creation.
-It also works in a more advanced setting with structured fields accesses and also introduces a table--lookup for pointers that are stored in memory.
-This paper only considers pointers to arrays of simple objects, i.e., there are no pointers to pointers or structures.
-Of course, there are a lot of other approaches to memory--safety in literature, specifically as compiler instrumentations~\cite{akritidis2009baggy,younan2010paricheck,jung2021pico,shankaranarayana2023tailcheck,dhumbumroong2020boundwarden,nam2019framer,zhou2023fatptrs}, hardware--extensions~\cite{kwon2013lowfat,saileshwar2022heapcheck,chen2023flexpointer,kim2023whistle}, or programming language extensions~\cite{elliott2018checkedc,li2022formalcheckedc,jim2002cyclone,elliott2015guilt,west2005cuckoo,weis2019fyr,benoit2019uniqueness}, to name a few.
-What differentiates this work from them is that this work uses known, compiler--based approaches to ensure memory--safety as a means to investigate secure compiler compositions.
-This paper does not provide efficient memory--safety, but serves as a theoretical foundation for the secure compilation domain.
-
-To extend the languages in this paper with a less restricted form of pointer arithmetic, the region coloring memory safety monitor presented in earlier work~\cite{michael2023mswasm} can be used.
-The work presenting this monitor provides an approach for the robust preservation of memory safety compiling from C to WASM.
-However, they do not discuss composition of secure compilers but rather investigate an instance of a secure compiler.
+The approach of this paper to memory safety and the countermeasures against memory safety insecurities are inspired by seminal work in the literature~\cite{nagarakatte2009soft,nagarakatte2010cets}.
+The view of memory safety in this paper exhibits the classic split between temporal and spatial memory safety.
+The former making sure that there are no double deallocations or use after frees.
+The latter making sure that there are no out of bounds accesses.
+When extending the languages in this paper with a less restricted form of pointer arithmetic, the region coloring memory safety monitor presented in earlier work~\cite{michael2023mswasm} can be used.
+The work presenting this monitor provides an approach for the robust {\em enforcement}, which implies robust preservation, of memory safety.
+But, they do not compose the enforcement mechanism with other secure compilation passes.
 
 \subsection{Cryptographic Constant Time Mechanisms}\label{subsec:relw:cctmechs}
-
-The approach to cryptographic constant time in this paper is high--level, where a programming language exposes a way to switch the semantics to a data (operand) independent timing mode.
-Since identifiers in $\Lscct$ are annotated with a secrecy tag, this approach is similar to others with information flow control.
-For example, Vale~\cite{bond2017vale} uses Dafny to ensure constant--time assembly code, while Jasmin~\cite{almeida2017jasmin} makes use of the Coq proof assistant to reject non--constant--time programs.
-CT-Wasm~\cite{watt2019ctwasm} enforces constant--timeness by means of a type system.
-Different to the approach of this paper, these approaches necessitate that the programmer writes \gls{cct} code.
-An approach to allow programmers to write more high--level code is CryptOpt~\cite{kuepper2023cryptopt}, which generates efficient target-code by means of a randomized search.
-This paper abstracts over concrete mitigation strategies and simply assumes that there is a flag to switch to a cryptographic--constant time execution mode.
-This can be realized by employing the FaCT~\cite{cauligi2019fact} compiler, which translates common non-constant time code patterns to be constant--time, and the data (object) independent timing execution mode of modern processors.
 

### 2023-07-08 popl/main.tex

--- 2023-07-08 popl/main.tex
@@
 %Properties and partitioned in the subsets of safety and liveness\MP{cit}, and given their central role for security, this paper focusses on safety properties.
-The set of all properties can be partitioned into different {\em classes} ($\class$), i.e., safety, liveness, and neither safety or liveness~\cite{clarkson2008hyper}.
-Given the central role for security, this paper focusses on safety properties.
-Note that there are other classes $\class$ that contain more domain-specific properties, such as the class of memory safety properties.
+The set of all properties can be partitioned into different {\em classes} ($\class$), i.e., safety, liveness, and neither safety nor liveness~\cite{clarkson2008hyper}.
+Given their central role for security, this paper focusses on safety properties.
+% Note that there are other classes $\class$ that contain more domain-specific properties, such as the class of memory safety properties.
+\MPin{
+	remove.
+	what does this sentence add? removing it does not remove any relevant information
+}
 For the class of safety properties, it is decidable wether a trace admits such a property with just a finite trace prefix.
-As an example, consider a trace that starts with an interaction with memory where a deallocation precedes a read: $\ev{Dealloc\ \loc}\cdot\ev{Read\ \loc\ 1729}\cdot\dots$
-This program behaviour is insecure, because it reads from a memory location that was freed already.
-The finite trace prefix $\ev{Dealloc\ \loc}\cdot\ev{Read\ \loc\ 1729}$ is enough to decide that the trace does not satisfy the property ,,no use after dealloc'' and there is no way to append events to this prefix which would result in the trace being admissible.
+\MPin{
+	passive. also, ``a trace admits a property'' is wrong, it's the other way around
+}
+As an example, consider a trace describing an interaction with a memory where the deallocation of a pointer $\loc$ precedes its read: $\ev{Dealloc\ \loc}\cdot\ev{Read\ \loc\ 1729}\cdot\dots$
+This program behaviour is insecure wrt a canonical notion of (temporal) memory safety dictating no use-after-frees of pointers\MP{cit}, because it reads from a memory location that was freed already.
+\MPin{
+	rewordeda a little for precision (e.g., what is $\loc?$), check the diff
+}
+The previous finite trace prefix is enough to decide that the trace does not satisfy temporal memory safety and there is no way to append events to this prefix which would result in the trace being admissible.
 %Given a safety property, a finite prefix of a longer trace $\trace$ is enough to decide whether a trace is admitted or not by the property.
@@
 This translation is considered {\em correct} if it is semantics--preserving.
+\MPin{
+	i.e., ... brief desc?
+}
 %That is, the compiler encodes any possible $\src{L}$--level feature in $\trg{L}$, such that the semantics are the same.
-Unfortunately, correct compilers may be insecure compilers~\cite{patrignani2019survey}.
-Programs translated by insecure compilers can violate security properties that the programmer assumes to hold. %, given the presence of a context--level program.
-For a compiler to be secure, this paper makes use of the robust compilation framework~\cite{abate2019jour}, which we summarize in the following definition.
-For \bul{compilers $\cc{\src{L}}{\trg{L}}$ to robustly preserve a class of properties $\class$}, consider \rul{properties $\pi$ of that class $\class$ and programs $\src{p}$ written in $\src{L}$} such that \iul{the program $\src{\progvar}$ robustly satisfies $\pi$}.
-Then, \oul{$\cc{\src{L}}{\trg{L}}\left(\src{p}\right)$, the compiled version of $\src{\progvar}$, must robustly satisfy $\pi$}.
+Unfortunately, correct compilers may be insecure compilers~\cite{patrignani2019survey}, and programs translated by insecure compilers can violate security properties that the programmer assumes to hold.
+%, given the presence of a context--level program.
+To define when a compiler is secure, this paper uses the robust compilation framework~\cite{abate2019jour}, which we summarize in the following definition.
+
+To define robust compilation we only need the notion of a class of properties.
+\MPin{define.}
+
+For \bul{compilers $\cc{\src{L}}{\trg{L}}$ to robustly preserve a class of properties $\class$}, if for any \rul{property $\pi$ of that class $\class$ and programs $\src{p}$ written in $\src{L}$} we have that \iul{the program $\src{\progvar}$ robustly satisfies $\pi$}, then \oul{the compilation of $\src{\progvar}$, $\cc{\src{L}}{\trg{L}}\left(\src{p}\right)$, must robustly satisfy $\pi$}.
 
 \begin{definition}[Robust Preservation]\label{def:rtp}
-  Compiler $\cc{\src{L}}{\trg{L}}$ robustly preserves $\class$, \bul{$\rtp{\cc{\src{L}}{\trg{L}}}{\class}$}, iff \rul{for any $\pi\in\class$ and $\src{p}$} such that \iul{$\rsat{\src{\progvar}}{\pi}$}, it holds that \oul{$\rsat{\cc{\src{L}}{\trg{L}}\left(\src{p}\right)}{\pi}$}.
+  % Compiler $\cc{\src{L}}{\trg{L}}$ robustly preserves $\class$, 
+  \bul{$\rtp{\cc{\src{L}}{\trg{L}}}{\class}$}
+  %, iff 
+  $\isdef$
+  \rul{for any $\pi\in\class$ and $\src{p}$} such that \iul{$\rsat{\src{\progvar}}{\pi}$}\MP{ use $\forall$}, it holds that \oul{$\rsat{\cc{\src{L}}{\trg{L}}\left(\src{p}\right)}{\pi}$}.
 \end{definition}

### 2023-07-08 popl/main.tex
Draft: Abstract

@@
 \begin{abstract}
-  \hltt{abstract}
+
+  Secure compilers emit code that is hardened against adversarial target--level attacks, such as memory corruption or alteration of control--flow.
+  It is an open question wether secure compilers compose in a well--behaved manner.
+  This paper answers this question positively: The composition of compilers that are secure with respect to some security--properties yields the intersection of these properties.
+  This article discusses different variants of compositions of secure compilers in a formal manner and instantiates parts of these results in a case--study.
+  To this end, an optimizing, secure compilation chain that preserves both memory safety and cryptographic constant time is presented.
+  Memory safety is achieved by composing secure compilers for both temporal and spatial memory safety, where the compiler for spatial memory safety is an instrumentation that inserts bounds--checks in front of pointer dereferences.
+  The compiler for cryptographic constant time exploits the use of a data (operand) timing independent mode and the employed optimisations of the compilation chain are constant folding and dead code elimination.
+  The compositionality results enable modular proofs for the whole compilation chain that is secure with respect to both memory safety and cryptographic constant time.
 

### 2023-07-08 popl/main.tex
Merge sections 3 and 4

@@
 
-\section{Example Properties\pages{1.5}}\label{sec:compprop}
+\section{Example Properties and their Monitors\pages{1.5}}\label{sec:compprop}
 
@@
 
-\subsection{Temporal Memory Safety}
+\subsubsection{Temporal Memory Safety}
 
@@
 
-\subsection{Spatial Memory Safety}
+\subsubsection{Spatial Memory Safety}
 
@@
 
-\subsection{Memory Safety}
+\subsubsection{Memory Safety}
 
@@
 
-\subsection{Strict Cryptographic Constant Time}
+\subsubsection{Strict Cryptographic Constant Time}
 
@@
 
-\subsection{Memory Safe, Strict Cryptographic Constant Time}
+\subsubsection{Memory Safe, Strict Cryptographic Constant Time}
 
@@
 %\section{Monitors and Trace Agreements\pages{1.5}}\label{sec:monitors}
-\section{Monitors\pages{1.5}}\label{sec:monitors}
+\subsection{Monitors\pages{1.5}}\label{sec:monitors}
 This section introduces monitors for each of the properties presented earlier (\Cref{subsec:propdefs}).
@@
 
-\subsection{Monitor for \glsfirst{tms}}
+\subsubsection{Monitor for \glsfirst{tms}}
 \begin{gather*}
@@
   %
-  \typerule{tms-uninteresting}{
-  }{
-    \monitorcheck[]{\tmsmonitor}{\tmsmonitor}{\absev{\emptyevent}}
-  }{tms-uninteresting}
-  %
-  \typerule{tms-abort}{
-  }{
-    \monitorcheck[]{\tmsmonitor}{\tmsmonitor}{\absterm}
-  }{tms-abort}
+% \typerule{tms-uninteresting}{
+% }{
+%   \monitorcheck[]{\tmsmonitor}{\tmsmonitor}{\absev{\emptyevent}}
+% }{tms-uninteresting}
+% %
+% \typerule{tms-abort}{
+% }{
+%   \monitorcheck[]{\tmsmonitor}{\tmsmonitor}{\absterm}
+% }{tms-abort}
   %
@@
 
-\subsection{Monitor for \glsfirst{sms}}
+\subsubsection{Monitor for \glsfirst{sms}}
 \begin{gather*}
@@
 
-\subsection{Combining \glsfirst{tms} and \glsfirst{sms} Monitors to obtain \glsfirst{ms}}\label{subsubsec:ms}
+\subsubsection{Combining \glsfirst{tms} and \glsfirst{sms} Monitors to obtain \glsfirst{ms}}\label{subsubsec:ms}
 
@@
 
-\subsection{Monitor for \glsfirst{scct}}
+\subsubsection{Monitor for \glsfirst{scct}}
 \begin{gather*}
@@
 
-\subsection{Combining \gls{ms} and \gls{scct} Monitors to obtain \gls{msscct}}
+\subsubsection{Combining \gls{ms} and \gls{scct} Monitors to obtain \gls{msscct}}
 

### 2023-07-08 popl/main.tex
Incorporate feedback

@@
 
-%In \Cref{sec:background}, the paper establishes the notions of traces, (hyper-)properties, and robust preservation, while \Cref{sec:relwork} compares this work with others.
 This paper starts by introducing relevant notions of security properties and secure compilation (\Cref{sec:background}),
-%We conclude in \Cref{sec:concl} and give an outlook for future work.
 and discusses related work (\Cref{sec:relwork}) before concluding (\Cref{sec:concl}).
@@
 Properties $\pi$ are sets of traces of admissible program behaviours, ascribing what said property considers valid~\cite{clarkson2008hyper}.
-%
-% make this intuition formally precise, however, they rely on a trace model.
-% {\em Traces} ($\trace$) are (possibly infinite) sequences of {\em events} ($\event$).
-% Events are an abstract representation of observable actions, in this paper
-% The empty event ($\emptyevent$) is an unobservable (or unimportant) action.
-% A property $\pi$ is a set of admissable traces and 
-%Properties and partitioned in the subsets of safety and liveness\MP{cit}, and given their central role for security, this paper focusses on safety properties.
 The set of all properties can be partitioned into different {\em classes} ($\class$), i.e., safety, liveness, and neither safety nor liveness~\cite{clarkson2008hyper}.
-Given their central role for security, this paper focusses on safety properties.
-% Note that there are other classes $\class$ that contain more domain-specific properties, such as the class of memory safety properties.
-\MPin{
-	remove.
-	what does this sentence add? removing it does not remove any relevant information
-}
-For the class of safety properties, it is decidable wether a trace admits such a property with just a finite trace prefix.
-\MPin{
-	passive. also, ``a trace admits a property'' is wrong, it's the other way around
-}
-As an example, consider a trace describing an interaction with a memory where the deallocation of a pointer $\loc$ precedes its read: $\ev{Dealloc\ \loc}\cdot\ev{Read\ \loc\ 1729}\cdot\dots$
-This program behaviour is insecure wrt a canonical notion of (temporal) memory safety dictating no use-after-frees of pointers\MP{cit}, because it reads from a memory location that was freed already.
-\MPin{
-	rewordeda a little for precision (e.g., what is $\loc?$), check the diff
-}
+A class is simply a set of properties and for the class of safety properties, it is decidable wether a trace satisfies a safety property with just a finite trace prefix.
+As an example, consider a trace describing an interaction with a memory where the deallocation of an address $\loc$ precedes a read at that address in memory: $\ev{Dealloc\ \loc}\cdot\ev{Read\ \loc\ 1729}\cdot\dots$
+This program behaviour is insecure with respect to a canonical notion of (temporal) memory safety dictating no use-after-frees of pointers~\cite{nagarakatte2010cets}, because it reads from a memory location that was freed already.
 The previous finite trace prefix is enough to decide that the trace does not satisfy temporal memory safety and there is no way to append events to this prefix which would result in the trace being admissible.
-%Given a safety property, a finite prefix of a longer trace $\trace$ is enough to decide whether a trace is admitted or not by the property.
-%For example, given the trace $\ev{Send\ 1}\cdot\ev{Recv\ 1}\cdot\ev{Send\ 2}\cdot\dots$ and the property $\pi=\left\{\trace \mid \trace=\hole{\cdot} \text{ or } \trace=\ev{Send\ n}\cdot\ev{Recv\ 2n}\cdot\trace[']\wedge \trace[']\in\pi\right\}$, the prefix $\ev{Send\ 1}\cdot\ev{Recv\ 1}$ immediately violates $\pi$.
-%\MP{mention send and recv before already. replace with read/write to contextualise to memory?}
-%
-% The set of security properties is one of many classes of properties ($\class$).
-% Note that existing literature refers to classes of properties as hyperproperties~\cite{clarkson2008hyper}.
-% The framework developed in this paper relies on {\em liftings} ($\lift{\pi}$), which are simply the powerset of a property $\pi$ and thus also considered as a class.
-% This framework is concerned about the secure compilation of programs that admit properties.
+In the following, the notation of an execution of a whole program $\wholeprogvar$ that terminates in state $\runtimetermvar$ according to the language semantics and produces trace $\trace$ is $\progstepto{\wholeprogvar}{\runtimetermvar}{\trace}$.
+With this, the definition of property satisfaction is as follows:
+\bul{whole programs $\wholeprogvar$ satisfy a property $\pi$} iff \iul{$\wholeprogvar$ yields a trace $\trace$} such that \oul{$\trace$ satisifies $\pi$} (\Cref{def:propsat}).
+
+\begin{definition}[Property Satisfaction]\label{def:propsat}
+  \bul{$\sat{\progvar}{\pi}$}
+  $\isdef$
+  if \iul{$\progstepto{\wholeprogvar}{\runtimetermvar}{\trace}$},
+  then \oul{$\trace\in\pi$}.
+\end{definition}
+
+Property satisfaction is defined on whole programs, i.e., programs without missing definitions.
+Thus, there is only a passive attacker model that observes the execution and, e.g., retreives secrets from that.
+To consider a stronger model similarily to what existing work does~\cite{abate2019jour,abate2021extacc,patrignani2021rsc}, the concept of satisfaction is extended with {\em robustness}.
+Robust satisfaction considers partial programs $\progvar$, i.e., programs with free variables, which need to be defined prior to execution.
+Programs with free variables can get stuck during execution.
+To remedy this, {\em linking} takes two partial programs $\progvar[_{1}],\progvar[_{2}]$ and produces a whole program $\wholeprogvar$, i.e., $\link{\progvar[_{1}]}{\progvar[_{2}]}=\wholeprogvar$.
+As typically done in secure compilation~\cite{abate2019jour}, \MP{others do this too, parametricity results e.g.} this paper assumes that whole programs are the result of linking partial programs referred to as {\em context} ($\ctx$) and {\em component} ($\comp$).
+The context is an arbitrary program and thus has the role of an {\em attacker} that can interact with the component by means of whatever features the programming language has.
+%In this work, the semantics of the programming language is expected to differentiate between the component and the context.
+With this, \Thmref{def:propsat} can be extended as follows: for \bul{components $\progvar$ to robustly satisfy a property $\pi$}, take an \iul{attacker context $\contextvar$ and link it with $\progvar$}, \oul{the result of which must satisfy $\pi$}.
 %\MPin{
-%	this par does not flow.
-%	what is the point?
+% the behaviour of the result?
+%%%%%%%    MK: def:propsat defines satisfaction in terms of programs, no need to talk about behavior here
 %}
-%
-% A program ($\runtimetermvar$) is a tuple consisting of some dynamic state ($\statevar$) and a syntactic description ($\progvar$).
-% Programs without free variables are considered {\em whole}.
-% Contrary, they are {\em partial} if they contain variables without an accompanying definition.
-% Whole programs can be run with the multi-step relation $\stepto{\runtimetermvar}{\runtimetermvar[']}{\trace}$ and yield a trace $\trace$.
-% Moreover, the paper assumes for such executions that $\runtimetermvar[']$ is {\em final}, decided by the judgement $\isval{\runtimetermvar}$.
-% Partial programs cannot be executed, since execution would get stuck when reaching an undefined variable.
-% The paper also uses other step-relations, such as a single-step ($\stepto[]{\runtimetermvar[_{1}]}{\runtimetermvar[_{2}]}{\event}$) and an $n$-step relation ($\stepton{\runtimetermvar[_{1}]}{\runtimetermvar[_{2}]}{\trace}$) that performs $n$ single-steps.
-% Note that the multi-step relation presented earlier is defined as-standard using the single-step, where the reflexive case makes use of $\isval{\runtimetermvar}$.
-% All multi-step relations filter $\emptyevent$ events.
-%\MPin{
-%	all unnecessary.
-%	unused notation, not needed for sec2, for sec3 and for sec 4.
-%	also i don't agree with what a program is.
-%}
-
-%Given a whole program $\progvar$, i.e., a program without free variables, the {\em behaviour} ($\behav{\progvar}$) is the set containing all traces the program can emit during execution according to its language semantics.
-%This paper only considers deterministic languages, where the behavior $\behav{\runtimetermvar}$ is a singleton set for any whole $\runtimetermvar$.
-%This set is used to define property satisfaction (\Cref{def:propsat}):
-\MPin{
-	fill the blank of the next sentence.
-}
-In the following we write that a whole program \MP{} runs according to the language semantics and produces trace \MP{} as \MP{}. 
-With this, we define that
-\bul{programs $\progvar$ satisfy a property $\pi$} iff \iul{$\progvar$ yields a trace $\trace$} such that \oul{$\trace$ is admissible with respect to $\pi$} (\Cref{def:propsat}).
-
-\begin{definition}[Property Satisfaction]\label{def:propsat}
-  % $\progvar$ satisfies $\pi$, 
-  \bul{$\sat{\progvar}{\pi}$} 
-  % iff 
-  $\isdef$
-  \iul{running $\progvar$ gives $\trace$}
-  \MP{this does not formalise things, see sentence added above and add formalism}
-   where \oul{$\trace\in\pi$}.
-\end{definition}
-
-\MPin{
-	why are we suddenly going from talking about prop sat to talknig about progs and ctxs?
-	there's a gap here that needs to be bridged.
-	A plausible solution is something along these lines.
-	"property satisfaction only considers passive attackers, to consider a stronger model with active attackers existing work (as well as this paper) uses robust property satisfaction, which we define after introducing the required concept of program context and partial programs"
-}
-Given a partial program $\progvar$, i.e., a program with free variables, the execution of $\progvar$ would get stuck due to the lack of program parts (e.g., function definitions).
-%As described earlier, partial programs cannot be executed \MP{not true anymore, need to define the notion of partial programs} , because they lack definitions and, thus, get stuck.
-To remedy this, {\em linking} takes two partial programs $\progvar[_{1}],\progvar[_{2}]$ and produces a new whole program $\link{\progvar[_{1}]}{\progvar[_{2}]}$.
-\MPin{
-	cumbersome notation.
-	`link ... ... ' should be used to identify the notation of the operation, but we should have a metavariable for the result, like `link(p1,p2) = w'
-	W is a known candidate
-}
-%Note that linking is not defined on programs, but merely on their syntactic descriptions, since programs contain dynamic state.\MP{ not terminology i really want to use, nor is defined}
-As typically done in secure compilation~\cite{abate2019jour}, \MP{others do this too, parametricity results e.g.} this paper assumes that \MP{whole?}programs are the result of linking a {\em context} ($\ctx$) and a {\em component} ($\comp$).
-\MPin{
-	these sentences identify the need to have 2 different terms: 1 for partial programs (components?) one for whole programs (just programs) and to use them consistently.
-}
-% \MP{just introduce the notion of contexts}
-%       MK: why only contexts?
-%The component part of a whole program is the main object of interest and thus a known syntactic description.
-The context is an arbitrary program and thus has the role of an {\em attacker} that can interact with the component by means of whatever features the programming language has.
-In this work, the semantics of the programming language is expected to differentiate between the component and the context.
-\MPin{
-	so what?
-}
-%\MP{and say the semantics is expected (passive, hah) to differentiate between prg and ctx}
-%%                                         -> why passive?
-So, the context interacts in some way with the component and this interaction models an attacker.
-\MP{
-	wut?
-	needless sentence, erase.
-}
-With this, \Thmref{def:propsat} can be extended as follows: for \bul{ partialprograms $\progvar$ to robustly satisfy a property $\pi$}, take an \iul{attacker $\contextvar$ and link it with $\progvar$}, \oul{the result of which must satisfy $\pi$}.
-\MPin{
-	the behaviour of the result?
-}
-%The following definition extends on \Thmref{def:propsat} to cover property satisfaction under the presence of an (active, with respect to the semantics of the programming language) attacker: %, which is known as {\em robust satisfaction}:
-%\MPin{missing: the context is an attacker! now, unlike before, we security property satisfaction with a clear active attacker}
 
 \begin{definition}[Robust Satisfaction]\label{def:proprsat}
-  % $\progvar$ robustly satisfies $\pi$, 
   \bul{$\rsat{\progvar}{\pi}$}
-  % iff 
-  $\isdef$ for \iul{any partial program $\contextvar$}\MP{use $\forall$} it holds that \oul{$\sat{\ {\contextvar}{\runtimetermvar}}{\pi}$}. \MP{why r?}
+  $\isdef$ \iul{$\forall \contextvar$, if $\link{\contextvar}{\progvar}=\wholeprogvar$}, then \oul{$\sat{\wholeprogvar}{\pi}$}.
 \end{definition}
@@
 This translation is considered {\em correct} if it is semantics--preserving.
-\MPin{
-	i.e., ... brief desc?
-}
-%That is, the compiler encodes any possible $\src{L}$--level feature in $\trg{L}$, such that the semantics are the same.
+That is, for a whole program $\src{\wholeprogvar}$, the compiler should relate the $\src{L}$ semantics of $\src{\wholeprogvar}$ with the semantics of $\trg{T}$ in such a way that they are ,,compatible''.
 Unfortunately, correct compilers may be insecure compilers~\cite{patrignani2019survey}, and programs translated by insecure compilers can violate security properties that the programmer assumes to hold.
-%, given the presence of a context--level program.
 To define when a compiler is secure, this paper uses the robust compilation framework~\cite{abate2019jour}, which we summarize in the following definition.
-
-To define robust compilation we only need the notion of a class of properties.
-\MPin{define.}
 
@@
   $\isdef$
-  \rul{for any $\pi\in\class$ and $\src{p}$} such that \iul{$\rsat{\src{\progvar}}{\pi}$}\MP{ use $\forall$}, it holds that \oul{$\rsat{\cc{\src{L}}{\trg{L}}\left(\src{p}\right)}{\pi}$}.
+  \rul{$\forall \left(\pi\in\class\right) \left(\src{p}\in\Ltms\right),$} if \iul{$\rsat{\src{\progvar}}{\pi}$}, then \oul{$\rsat{\cc{\src{L}}{\trg{L}}\left(\src{p}\right)}{\pi}$}.
 \end{definition}
@@
 \begin{definition}[Monitor Satisfaction]\label{def:monsat}
-  \bul{$\monsafe{\trace}{\pi}$} iff $\exists \abstrace\ \monitor$ such that \iul{$\traceagree{\pi}{\trace}{\abstrace}$} and \oul{$\monitorcheck{\emptyset}{\monitor}{\abstrace}$}.
+  \bul{$\monsafe{\trace}{\pi}$} $\isdef$ $\exists \abstrace\ \monitor$, \iul{$\traceagree{\pi}{\trace}{\abstrace}$} and \oul{$\monitorcheck{\emptyset}{\monitor}{\abstrace}$}.
 \end{definition}
@@
 
-This section presents the key meta-theoretic results of this paper and extends on prior work~\cite{kruse2022csc}.
+This section presents the key meta-theoretic results of this paper and builds on prior work~\cite{kruse2022csc}.
 A main result is that secure compilers in the robust preservation framework~\cite{abate2019jour} compose.
@@
 \begin{theorem}[Sequential Composition of Secure Compilers]\label{thm:rtp}
-  Given $\class[_{1}],\class[_{2}]$, \bul{$\rtp{\cc{\src{L}}{\trg{L}}}{\class[_{1}]}$}, and \rul{$\rtp{\cc{\trg{L}}{\obj{L}}}{\class[_{2}]}$}, then \oul{$\rtp{\cc{\src{L}}{\trg{L}}\circ\cc{\trg{L}}{\obj{L}}}{\class[_{1}]\cap\class[_{2}]}$}.
+  If \bul{$\rtp{\cc{\src{L}}{\trg{L}}}{\class[_{1}]}$} and \rul{$\rtp{\cc{\trg{L}}{\obj{L}}}{\class[_{2}]}$}, then \oul{$\rtp{\cc{\src{L}}{\trg{L}}\circ\cc{\trg{L}}{\obj{L}}}{\class[_{1}]\cap\class[_{2}]}$}.
 \Coqed
@@
 \begin{corollary}[Sequential Composition of Secure Compilers]\label{corr:swappable}
-  Given $\class[_{1}],\class[_{2}]$, \bul{$\rtp{\cc[_{1}]{\trg{L}}{\trg{L}}}{\class[_{1}]}$, and $\rtp{\cc[_{2}]{\trg{L}}{\trg{L}}}{\class[_{2}]}$}, then \oul{$\rtp{\cc[_{1}]{\trg{L}}{\trg{L}}\circ\cc[_{2}]{\trg{L}}{\trg{L}}}{\class[_{1}]\cap\class[_{2}]}$ and $\rtp{\cc[_{2}]{\trg{L}}{\trg{L}}\circ\cc[_{1}]{\trg{L}}{\trg{L}}}{\class[_{2}]\cap\class[_{1}]}$}.
+  If \bul{$\rtp{\cc[_{1}]{\trg{L}}{\trg{L}}}{\class[_{1}]}$ and $\rtp{\cc[_{2}]{\trg{L}}{\trg{L}}}{\class[_{2}]}$}, then \oul{$\rtp{\cc[_{1}]{\trg{L}}{\trg{L}}\circ\cc[_{2}]{\trg{L}}{\trg{L}}}{\class[_{1}]\cap\class[_{2}]}$ and $\rtp{\cc[_{2}]{\trg{L}}{\trg{L}}\circ\cc[_{1}]{\trg{L}}{\trg{L}}}{\class[_{2}]\cap\class[_{1}]}$}.
 \Coqed

### 2023-07-08 popl/main.tex
Sketch intro

@@
 
-\hltt{Context}
-\hltt{Problem}
-
-\begin{lstlisting}[language=c,basicstyle=\small\ttfamily, caption=C]
-int strncpy(int n, int *x, int *y) {
+\gls{ms} is a security property of programs.
+Without memory safety, private data can easily leak or corrupt~\cite{lemay2021ccc}.
+Because of this, \gls{ms} is a necessity for programs to be \gls{cct}, which is an important security property for secrecy.
+Programs that attain \gls{cct} are secure with respect to timing attacks~\cite{kocher1996timing}.
+Timing attacks may expose secrets from a simple performance analysis.
+These attacks have become more and more relevant since \citeauthor{kocher1996timing}'s paper, especially given the era of cloud--computing~\cite{aviram2010cloudtime,kumar2019cloudsecsurvey}, where programs share the same hardware but are supposed to run in a sandboxed environment that does not leak secrets.
+Unfortunately, several attacks within such a cloud environment have been demonstrated as viable strategies to extract secrets~\cite{mehmet2015getoff,flowers2022zeroday,atya2019catchme,venkatanathan2015placevul}.
+
+Even though mitigations for such attacks exist~\cite{bond2017vale,almeida2017jasmin}, it has been shown in literature that some compilers may optimize these mitigations away~\cite{barthe2018sec}.
+This is of course not what the programmer expects and necessitates the need for a {\em secure} compiler that ensures security properties hold even when the program runs in a hostile environment, i.e., interacts with other programs by any means.
+Because of this hostile environment, secure compilers attain security by relying on certain protection mechanisms, such as ... and ...\MK{cits}.
+The aim of this paper is to devise the theoretical foundation to be able to build a secure compiler that is capable of dealing with the following example.
+Consider the program bolow which copies a \texttt{C}-style string \texttt{x} into \texttt{y} up to a length of \texttt{n}:
+
+\begin{lstlisting}[language=c,basicstyle=\small\ttfamily]
+int strncpy(char n, char *x, char *y) {
   for(int i = 0; x[i] != '\0' && i < n; ++i) {
@@
 
-\BrText{Solution}{
+For sake of simplicity it is assumed that the buffer \texttt{y} is at least as large as \texttt{n} as well as \texttt{x} terminated by \texttt{'\textbackslash 0'}.
+Even under these assumptions, there is a subtle \gls{ms} bug in the program which can be exploited to read sensitive data:
+The bounds check, i.e., \texttt{i < n}, should happen prior to the access to memory location \texttt{x[i]}.
+Because otherwise, if \texttt{n} is larger than the length of the array object \texttt{x} points to, the memory location past this array object is being read, which is an out--of--bounds memory access.
+Unfortunately, even though flipping the conditions does fix the \gls{ms} issue, \gls{cct} is violated, so the program is susceptible to timing attacks.
+This is because, the loop can terminate early: As soon as the string--terminating character \texttt{'\textbackslash 0'} is encountered, the loop exits.
+Thus, program execution time is proportional to the length of the array object \texttt{x}.
+The fix commonly employed in such a case is to rewrite the program to run a fixed number of iterations regardless of the length of \texttt{x} or \texttt{y}.
+To this end, there must be an upper bound for the length of \texttt{x} and \texttt{y} and then the loop can just iterate until the induction variable \texttt{i} hits this upper bound.
+
+These fixes alone are not enough to make this program robust, i.e., not violating either \gls{ms} or \gls{cct} in a malicious environment.
+Consider a program context that calls \texttt{strncpy} with a very large \texttt{n}, some uninitialised memory \texttt{x}, and \texttt{y} whatever.
+It is easy to see that such a function invocation leads to a series of reads from uninitialised memory, an immediate \gls{ms} issue.
+To mitigate these situtations, it is important that \texttt{strncpy} ensures that the provided arguments satisfy the assumptions being made.
+After all, the environment cannot be trusted and potentially behaves in a hostile manner.
+
+\MKin{
+  show final implementation?
+}
+%\begin{lstlisting}[language=c,basicstyle=\small\ttfamily]
+%int strncpy(char n, char *x, char *y) {
+%  for(int i = 0; x[i] != '\0' && i < n; ++i) {
+%    y[i] = x[i];
+%  }
+%}
+%\end{lstlisting}
+
+As stated earlier, this work wants to establish the theoretical foundation to ensure the secure compilation of programs like \texttt{strncpy} to satisfy properties \gls{ms} and \gls{cct} robustly.
+To this end, this paper chooses the robust preservation framework, which is a model of secure compilation that is well--established~\cite{abate2019jour,abate2021extacc,patrignani2021rsc}.
+% and has a stronger attacker model than a passive attacker whose only means of interacting with a program are timing analyses and so on.
+Unfortunately, this framework of secure compilers does not support modularity: One has to make a secure compiler for all possible properties {\em or} for a combination of \gls{ms} and \gls{cct}.
+
+\MKin{
+blablabla continue blablabla
+}
+
 % keyword: pass
@@
 }
-}
-
-% \BrText{Validation}{
+
 In summary, this paper makes the following contributions:
@@
 \end{itemize}
-% }
 

### 2023-07-08 popl/main.tex
intro example

@@
 \begin{lstlisting}[language=c,basicstyle=\small\ttfamily]
-int strncpy(char n, char *x, char *y) {
+void strncpy(int n, char *x, char *y) {
   for(int i = 0; x[i] != '\0' && i < n; ++i) {
@@
 After all, the environment cannot be trusted and potentially behaves in a hostile manner.
-
-\MKin{
-  show final implementation?
-}
-%\begin{lstlisting}[language=c,basicstyle=\small\ttfamily]
-%int strncpy(char n, char *x, char *y) {
-%  for(int i = 0; x[i] != '\0' && i < n; ++i) {
-%    y[i] = x[i];
-%  }
-%}
-%\end{lstlisting}
+Unfortunately, these checks do not work in C without additional features, since there is no way to check wether a pointer points into a valid allocated region or not.
+To enable this would require extra metadata~\cite{kwon2013lowfat,akritidis2009baggy} which must not be modifiable by a malicious context.
+So, for sake of simplicity, assume the language at hand provides an operator \texttt{is\_valid(p, s, n)} that faithfully checks wether a pointer \texttt{p} points into a valid array of objects sized \texttt{s} of length \texttt{n} or not --- in constant time.
+This could be an abstraction that the operating system provides.
+With this feature at hand, it is possible to modify the original \texttt{strncpy} to be robust, given this is the compiled program without optimisations:
+
+\begin{lstlisting}[language=c,basicstyle=\small\ttfamily]
+void strncpy(int n, char *x, char *y) {
+  int const c = __ct_min(n, MAX_ALLOC_SIZE);
+  assert(is_valid(x, sizeof(char), c) & is_valid(y, sizeof(char), c));
+  for(int i = 0; i < c; ++i) {
+    y[i] = x[i];
+  }
+}
+\end{lstlisting}
 
 As stated earlier, this work wants to establish the theoretical foundation to ensure the secure compilation of programs like \texttt{strncpy} to satisfy properties \gls{ms} and \gls{cct} robustly.
-To this end, this paper chooses the robust preservation framework, which is a model of secure compilation that is well--established~\cite{abate2019jour,abate2021extacc,patrignani2021rsc}.
-% and has a stronger attacker model than a passive attacker whose only means of interacting with a program are timing analyses and so on.
-Unfortunately, this framework of secure compilers does not support modularity: One has to make a secure compiler for all possible properties {\em or} for a combination of \gls{ms} and \gls{cct}.
-
-\MKin{
-blablabla continue blablabla
-}
+Secure compilation in terms of robust preservation is well--established~\cite{abate2019jour,abate2021extacc,patrignani2021rsc}.
+Unfortunately, at the time of writing this paper, it is unclear if and how the theory of robust preservation would support modularity.
+Missing out on modularity would mean that one has to make a secure compiler for all possible properties {\em or} for a combination of \gls{ms} and \gls{cct}.
+Ideally, both for proof engineering and reusability factor, a secure compiler for just \gls{ms} and a secure compiler for just \gls{cct} can compose in some way to get a secure compiler for the combination of the two.
+This also reflects real--world compilation chains, where several compilation passes perform optimisations~\cite{lattner2004llvm,googlev8,androidstudio,kuepper2023cryptopt,manjikian1997fusion,wegman1991ccp} or source--code instrumentations~\cite{nagarakatte2009soft,nagarakatte2010cets,akritidis2009baggy,dhumbumroong2020boundwarden,jung2021pico,nam2019framer,shankaranarayana2023tailcheck,younan2010paricheck,zhou2023fatptrs,bond2017vale,almeida2017jasmin,kuepper2023cryptopt,cauligi2019fact} that enforce that the compiled program fulfills a property of interest.
 
 % keyword: pass
-This paper introduces a framework for reasoning about the composition of secure and optimising compiler passes and it showcases the power of this framework by instantiating it on a multi-pass compilation chain.
+In turn, this paper introduces a framework for reasoning about the composition of secure and optimising compiler passes and it showcases the power of this framework by instantiating it on a multi-pass compilation chain.
 To this end, this paper first discusses how to compose security properties, such as temporal and spatial memory safety into general memory safety, and cryptographic constant time.

### 2023-07-09 popl/main.tex
remove spurious text
Add todo

@@
 %% These commands are for a PROCEEDINGS abstract or paper.
-\acmConference[POPL '24]{Make sure to enter the correct
-  conference title from your rights confirmation emai}{June 03--05,
-  2018}{Woodstock, NY}
+\acmConference[POPL '24]{51st ACM SIGPLAN Symposium on Principles of Programming Languages}{January 17-19,
+  2024}{London, UK}
 %
@@
   This article discusses different variants of compositions of secure compilers in a formal manner and instantiates parts of these results in a case--study.
-  To this end, an optimizing, secure compilation chain that preserves both memory safety and cryptographic constant time is presented.
+  To this end, it presents an optimizing, secure compilation chain that preserves both memory safety and cryptographic constant time.
   Memory safety is achieved by composing secure compilers for both temporal and spatial memory safety, where the compiler for spatial memory safety is an instrumentation that inserts bounds--checks in front of pointer dereferences.
@@
   $
-  A trace $\trace$ is spatial memory safe ($\trace\in\smssafe$) iff given $\ev{Alloc\ \loc\ m;\sandboxtag;\securitytag}\le_{\trace}\ev{Use\ \loc\ n;\sandboxtag';\securitytag'}$, then $n<m$.
 \end{definition}
@@
   $
-  A trace $\trace$ is memory safe ($\trace\in\mssafe$) iff $\trace\in\tmssafe$ and $\trace\in\smssafe$.
 \end{definition}
@@
 
-This section presents the key meta-theoretic results of this paper and builds on prior work~\cite{kruse2022csc}.
+This section presents the key meta-theoretic results of this paper and builds on prior work~\cite{kruse2022csc}\MK{mention this at all? may reveal identity even though its double-blind}.
 A main result is that secure compilers in the robust preservation framework~\cite{abate2019jour} compose.

### 2023-07-09 popl/main.tex
New abstract

@@
 
-  Secure compilers emit code that is hardened against adversarial target--level attacks, such as memory corruption or alteration of control--flow.
-  It is an open question wether secure compilers compose in a well--behaved manner.
-  This paper answers this question positively: The composition of compilers that are secure with respect to some security--properties yields the intersection of these properties.
-  This article discusses different variants of compositions of secure compilers in a formal manner and instantiates parts of these results in a case--study.
+  Memory safety necessitates secrecy, since breaking it can lead to an arbitrarily large attack space, i.e., from simply reading a secret up to alteration of control--flow in order to execute arbitrary code.
+  Secrecy cannot be achieved with just memory safety alone, since private data could be leaked by differences in execution time for varying program inputs.
+  This leakage is avoidable by ensuring that programs are cryptographic constant time, i.e., differnt inputs do not change execution time.
+  Prior work has shown that strategies to ensure cryptographic constant time can be rendered useless, since a compiler may simply optimize them away.
+  This is why secure compilation is an integral ingredient to achieve security, because a secure compiler guarantees (formally) that properties that hold in the source, also hold in the target.
+%% mention attacker?
+the text describes correctness
+
+  Unfortunately, it is an open question wether secure compilers can be engineered in a modular way.
+  Ideally, compiler engineers develop two different compilation passes, one that ensures memory safety and another one that tackles cryptographic constant time, because this style of engineering is well--established for its simplifying divide--and--conquer approach as well as enhanced reusability.
+
+  This paper answers that question positively: The composition of compilers that are secure with respect to some security--properties yields the intersection of these properties.
+  This article discusses different variants of compositions of secure compilers and instantiates parts of these results in a case--study.
   To this end, it presents an optimizing, secure compilation chain that preserves both memory safety and cryptographic constant time.
@@
   The compositionality results enable modular proofs for the whole compilation chain that is secure with respect to both memory safety and cryptographic constant time.
+  Parts of the results are verified in the Coq proof assistant.
+
+% Secure compilers emit code that is hardened against adversarial target--level attacks, such as memory corruption or alteration of control--flow.
+% It is an open question wether secure compilers compose in a well--behaved manner.
+% This paper answers this question positively: The composition of compilers that are secure with respect to some security--properties yields the intersection of these properties.
+% This article discusses different variants of compositions of secure compilers and instantiates parts of these results in a case--study.
+% To this end, it presents an optimizing, secure compilation chain that preserves both memory safety and cryptographic constant time.
+% Memory safety is achieved by composing secure compilers for both temporal and spatial memory safety, where the compiler for spatial memory safety is an instrumentation that inserts bounds--checks in front of pointer dereferences.
+% The compiler for cryptographic constant time exploits the use of a data (operand) timing independent mode and the employed optimisations of the compilation chain are constant folding and dead code elimination.
+% The compositionality results enable modular proofs for the whole compilation chain that is secure with respect to both memory safety and cryptographic constant time.
 

### 2023-07-09 popl/main.tex
Comment parts of abstract

@@
   Memory safety necessitates secrecy, since breaking it can lead to an arbitrarily large attack space, i.e., from simply reading a secret up to alteration of control--flow in order to execute arbitrary code.
-  Secrecy cannot be achieved with just memory safety alone, since private data could be leaked by differences in execution time for varying program inputs.
-  This leakage is avoidable by ensuring that programs are cryptographic constant time, i.e., differnt inputs do not change execution time.
+  However, secrecy cannot be achieved with just memory safety alone, since, e.g., private data could be leaked by differences in execution time for varying program inputs.
+  This leakage is avoidable by ensuring that programs are cryptographic constant time, i.e., different inputs do not change execution time.
   Prior work has shown that strategies to ensure cryptographic constant time can be rendered useless, since a compiler may simply optimize them away.
@@
   Unfortunately, it is an open question wether secure compilers can be engineered in a modular way.
-  Ideally, compiler engineers develop two different compilation passes, one that ensures memory safety and another one that tackles cryptographic constant time, because this style of engineering is well--established for its simplifying divide--and--conquer approach as well as enhanced reusability.
+  Ideally, compiler engineers develop two different compilation passes, one that ensures memory safety and another one that is security preserving with respect to cryptographic constant time.
+%, because this style of engineering is well--established for its simplifying divide--and--conquer approach as well as enhanced reusability.
 
@@
   To this end, it presents an optimizing, secure compilation chain that preserves both memory safety and cryptographic constant time.
-  Memory safety is achieved by composing secure compilers for both temporal and spatial memory safety, where the compiler for spatial memory safety is an instrumentation that inserts bounds--checks in front of pointer dereferences.
-  The compiler for cryptographic constant time exploits the use of a data (operand) timing independent mode and the employed optimisations of the compilation chain are constant folding and dead code elimination.
-  The compositionality results enable modular proofs for the whole compilation chain that is secure with respect to both memory safety and cryptographic constant time.
-  Parts of the results are verified in the Coq proof assistant.
+%  Memory safety is achieved by composing secure compilers for both temporal and spatial memory safety, where the compiler for spatial memory safety is an instrumentation that inserts bounds--checks in front of pointer dereferences.
+  %The compiler for cryptographic constant time exploits the use of a data (operand) timing independent mode and the employed optimisations of the compilation chain are constant folding and dead code elimination.
+  The compositionality results enable modular proofs for the presented, secure compiler. % that is secure with respect to both memory safety and cryptographic constant time.
+  Some results of this work are implemented in the Coq proof assistant.
 

### 2023-07-09 popl/main.tex
add citation to parametricity

@@
 To remedy this, {\em linking} takes two partial programs $\progvar[_{1}],\progvar[_{2}]$ and produces a whole program $\wholeprogvar$, i.e., $\link{\progvar[_{1}]}{\progvar[_{2}]}=\wholeprogvar$.
-As typically done in secure compilation~\cite{abate2019jour}, \MP{others do this too, parametricity results e.g.} this paper assumes that whole programs are the result of linking partial programs referred to as {\em context} ($\ctx$) and {\em component} ($\comp$).
+As typically done in not only secure compilation~\cite{abate2019jour,devriese2018parametricity}, \MP{others do this too, parametricity results e.g.} this paper assumes that whole programs are the result of linking partial programs referred to as {\em context} ($\ctx$) and {\em component} ($\comp$).
 The context is an arbitrary program and thus has the role of an {\em attacker} that can interact with the component by means of whatever features the programming language has.

### 2023-07-09 popl/main.tex
pass over contribs

@@
 The results showcase how the framework allows the kind of formal security reasoning that compiler writers already want (and do), obtaining precise, compositional security reasoning while providing minimal (and modular) proof effort.
-\MPin{
-	last line: not the best.
-}
+%\MPin{
+%	last line: not the best.
+%}
 
@@
         %\Cref{sec:compprop} presents a formal framework to reason about compositions of security properties and demonstrate its usage with concrete examples:
-        First, this paper studies the formalisation and the composition of security properties (\Cref{sec:compprop}) by focussing on those safety properties that are of interest for real-world compiler writers (as identified by the plethora of work enforcing such properties individually~\cite{akritidis2009baggy,nagarakatte2009soft,nagarakatte2010cets,dhumbumroong2020boundwarden,jung2021pico,nam2019framer,shankaranarayana2023tailcheck,younan2010paricheck,zhou2023fatptrs,bond2017vale,almeida2017jasmin,kuepper2023cryptopt,cauligi2019fact}\MK{even more cits?}).
-        % 
-        Starting from ways to formalise and enforce those properties individually, this paper shows how to compose both their formalisation and their composition.
+        First, this paper studies the formalisation and the composition of security properties (\Cref{sec:compprop}) by focussing on those properties that are of interest for real-world compiler writers (as identified by the plethora of work enforcing such properties individually~\cite{akritidis2009baggy,nagarakatte2009soft,nagarakatte2010cets,dhumbumroong2020boundwarden,jung2021pico,nam2019framer,shankaranarayana2023tailcheck,younan2010paricheck,zhou2023fatptrs,bond2017vale,almeida2017jasmin,kuepper2023cryptopt,cauligi2019fact}\MK{even more cits?}).
+        % 
+        Starting from ways to formalise and enforce those properties individually, this paper shows how to compose their formalisation. % and their composition. %% MK: compose their composition??
         % 
@@
         % 
-        The resulting security property is the golden standard of safety properties that secure cryptographic code should have~\MP{cits}.
-        \MPin{
-        	conclude that did this prop in a single pass, we show how to do it compositionally
-        }
-        %For each of these properties, we also give a monitor which runs on another trace model that excludes unneccessary information and show that their composition is behaviorally similar to the composition of the trace-based security properties.
-        %This two-level approach enhances modularity: When working with the security properties, one only needs to reason about the corresponding monitor.
-        %%% let's omit above sentences completely. why ramble about monitors in the introduction? this ain't about model checking
+        The resulting security property is the golden standard of safety properties that secure cryptographic code should have~\cite{lemay2021ccc}\MK{more cits?}.
+%        \MPin{
+%        	conclude that did this prop in a single pass, we show how to do it compositionally
+%        }
 
@@
         %Concretely, it investigates upper and lower compositions, as well as the sequential composition.
-        For this, this paper studies \MP{}.
+        For this, this paper studies the sequential composition of compilers, as typically done when engineering compilation chains, as well as an upper and a lower composition that discuss multi--in- and multi--output compilers.
         % 
         This paper proves that starting from two compilers that preserve two properties, their composition preserves the intersection of those properties.
-        \MPin{
-       		find way to connect to previous contrib
-        }
-        % 
-        Finally, this paper provides a simple but crucial corollary that demonstrates that the order of composition of sequential compiler passes is irrelevant for the resulting security.
+        For example, a compiler that is secure with respect to \gls{ms} can be composed with a compiler that is secure with respect to \gls{cct}, resulting in a compiler that is secure with respect to the composition of \gls{ms} and \gls{cct}.
+%       \MPin{
+%      		find way to connect to previous contrib
+%       }
+        % 
+        Finally, this paper provides an important corollary that demonstrates that the order of composition of sequential compiler passes is irrelevant for the resulting security.
         % 
@@
   \item %
-        % \Cref{sec:casestud} sets up a case-study to demonstrate the power of our formal framework.
-        This paper presents a case-study showcasing the conjunction of the previous contributions (\Cref{sec:casestud}).
-        %We define several different languages and compilers between them and prove that the compilers are robustly preserving different properties.
+        This paper presents a case-study showcasing the conjunction of the previous contributions (\Cref{sec:casestud:defs,sec:casestud:rtp}).
         To this end, it presents a compilation chain consisting of several passes.
-        %The sequential composition of these compilers yields a compilation chain that is optimising and robustly preserving a combination of \gls{ms} and \gls{cct}.
-        The chain ultimately preserves a combination of \gls{ms} and \gls{cct} by means of composing the individual, secure passes concerning \gls{tms}, \gls{sms}, and \gls{scct}, respectively.
-        %As for optimisations, the chain has two passes, one performing \gls{cf} and the other \gls{dce}.
-        Furthermore, the chain includes two passes for \gls{ms}\MKin{maybe patch \gls{ms} when optimising passes are done} that are optimising: One performs a simple \gls{dce} and the other \gls{cf}.
-        %This compilation chain demonstrates the ability to perform modular proof engineering by leaveraging this paper's compositionality theorems.
-        This demonstrates that the usage of the presented framework is both simpler and more expressive compared to one monolithic security proof of the whole compilation chain.
-        \MP{ matthis please write your idea here }
-
-  \item The key contributions are formalized in the Coq proof assistant.
-        The practically relevant parts of the formalization are \emph{extractable}.
-        That is, one can translate the Coq code to Ocaml and run the compilation chain from the previous contribution.
-        Finally, this paper indicates with \CoqSymbol whenever definitions or proofs are done in Coq.
+        The chain ultimately preserves a combination of \gls{ms} and \gls{cct} by means of composing the individual, secure passes concerning \gls{tms}, \gls{sms}, and a stricter variant of \gls{cct}, respectively.
+        Furthermore, the chain includes two passes for \gls{ms} that are optimising: One performs a simple \gls{dce} and the other \gls{cf}.
+        The formalisation of this case study showcases the power of the presented framework.
+        Namely, the divide--and--conquer approach is a viable strategy even for engineering secure compilers.
+%        This demonstrates that the usage of the presented framework is both simpler and more expressive compared to one monolithic security proof of the whole compilation chain.
+
+  \item The key contributions are formalized in the Coq proof assistant and the paper indicates this with \CoqSymbol.
 \end{itemize}
@@
 These translations can be freely ordered in the compilation chain without compromising memory safety (\Cref{thm:cccfccdce:rtp:ms}).
-The last step ensures that code stays \gls{scct} (\Cref{thm:ccscct:rtp:scct}) which yields the final result that the whole compilation chain robustly preserves \gls{msscct} (\Cref{thm:ccall:rtp:msscct}).
+The last step ensures that code stays \gls{scct} (\Cref{thm:ccscct:rtp:scct}) when lowered from $\Lms$ to $\Lscct$ which yields the final result that the whole compilation chain robustly preserves \gls{msscct} (\Cref{thm:ccall:rtp:msscct}).
 
@@
 The proof of \Cref{thm:cca:rtp:tms} uses standard techniques in secure compilation literature and is visualized in \Cref{fig:proofdiag:rtp}.
-Concretely, it uses a trace--based backtranslation to reconstruct an $\src{L_{\tmssafe}}$ context from the $\trg{L}$ trace, which, however, behaves similar to the $\trg{L}$--level context.
+Concretely, given the execution of the compiled $\Ltms$ program linked with an arbitrary $\trg{L}$ context yields a trace $\trg{\trace}$, it uses a trace--based backtranslation of $\trg{\trace}$ to reconstruct an $\src{L_{\tmssafe}}$ context.
 \Cref{thm:wt:tms} already gives strong guarantees for programs written in $\src{L_{\tmssafe}}$ and can be exploited for the backtranslated context by ensuring that it typechecks.

### 2023-07-09 popl/main.tex
sec 6

@@
 
-\section{Introduction\pages{4}}
+\section{Introduction\pages{4}}\label{sec:introduction}
 
@@
 The compiler does not insert other checks and simply recolors $\src{L_{\tmssafe}}$ to $\trg{L}$ expressions.
+Hypothesizing an implementation of the \texttt{strncpy} function from \Cref{sec:introduction} in $\Ltms$, the compiler would in this case ensure that the arguments are valid.
 
@@
 
-The language $\irl{L_{\mssafe}}$ is exactly the same as $\trg{L}$ and are considered equal.
-However, it is still named and collored differently to emphasize the fact that after this section, compiled $\irl{L_{\mssafe}}$ programs are \gls{ms}.
-
 \begin{center}
@@
 
-Up to getting and setting memory, the compiler is the identity function.
+Up to getting and setting memory, the compiler is the identity function, i.e., all omitted cases are a recoloring of $\trg{L}$ to $\Lms$.
 For memory accesses, however, a bounds--check is inserted that enforces \gls{sms}.
 To this end, the compiler introduces another, fresh identifier $\irl{x_{SIZE}}$ for each allocation that binds $\irl{x}$ to keep track of the allocation size.
+Again, linking this back to the \texttt{strncpy} example corresponds to the insertion of bounds check right in front of the accesses to \texttt{y[i]} and \texttt{x[i]}.
 
@@
 
-The composition of $\cca$ and $\ccb$ attains full memory safety using \Cref{thm:rtp,thm:cca:rtp:tms,thm:ccb:rtp:sms}.
+It is now possible to make use of \Cref{thm:rtp} to extract a compiler from $\Ltms$ to $\Lms$ that is \gls{ms} by applying \Cref{thm:cca:rtp:tms,thm:ccb:rtp:sms}:
 
@@
 
+The two optimising compiler passes from $\Lms$ to $\Lms$ perform \gls{dce} and \gls{cf}, respectively.
+Both are secure with respect to \gls{ms}, the proof of which is relatively easy, since the backtranslation can use the original context without modification and both transformations do not change anything with regards to allocation, use, or deallocation.
+That is, the order of appearance of the events on the execution trace stays the same.
+
 \begin{theorem}[Compiler $\ccdce$ is secure with respect to \gls{ms}]\label{thm:ccdce:rtp:ms}
@@
 \end{theorem}
+
+With both \Cref{thm:ccdce:rtp:ms,thm:cccf:rtp:ms} it follows from \Cref{corr:swappable} that the two passes can be interchanged arbitrarily:
 
@@
 
+Given the fact that $\Lscct$ provides a \gls{cct}--mode, the compiler has to ensure that execution in the component always happen in this mode.
+The context could overwrite the flag and exit the mode, but upon invoking a callable object that is part of the component, the flag would be set again.
+Because of this, the compiler is secure with respect ot \gls{scct}, similarily proven as in \Cref{subsec:cs:tms}.
+
 \begin{theorem}[Compiler $\ccscct$ is secure with respect to \gls{scct}]\label{thm:ccscct:rtp:scct}
@@
 
-Let $\cc{\Ltms}{\Lscct}$ be the compiler that is the composition of $\cca$, $\ccb$, $\cccf$, $\ccdce$, and $\ccscct$, then the following theorem can be shown.
-
-\begin{theorem}[Compiler $\ccscct$ is secure with respect to \gls{scct}]\label{thm:ccscct:rtp:scct}
-  $\rtp{\ccscct}{\scctsafe}$ % \Coqed
-\end{theorem}
-
-\begin{theorem}[Compiler $\ccscct$ is secure with respect to \gls{scct}]\label{thm:ccall:rtp:msscct}
+Let $\ccmsscct$ be the compiler that is the composition of $\cca$, $\ccb$, $\cccf$, $\ccdce$, and $\ccscct$, then the following theorem can be shown.
+
+\begin{theorem}[Compiler $\ccmsscct$ is secure with respect to \gls{scct}]\label{thm:ccall:rtp:msscct}
   $\rtp{\cc{\Ltms}{\Lscct}}{\mssafe\cap\scctsafe}$ % \Coqed

### 2023-07-09 popl/main.tex
Add 2.3

@@
 The aim of this paper is to devise the theoretical foundation to be able to build a secure compiler that is capable of dealing with the following example.
+\begin{example}[strncpy]\label{ex:strncpy}
 Consider the program bolow which copies a \texttt{C}-style string \texttt{x} into \texttt{y} up to a length of \texttt{n}:
@@
 \end{lstlisting}
-
-As stated earlier, this work wants to establish the theoretical foundation to ensure the secure compilation of programs like \texttt{strncpy} to satisfy properties \gls{ms} and \gls{cct} robustly.
+\end{example}
+
+As stated earlier, this work wants to establish the theoretical foundation to ensure the secure compilation of programs like \texttt{strncpy} (\Cref{ex:strncpy}) to satisfy properties \gls{ms} and \gls{cct} robustly.
 Secure compilation in terms of robust preservation is well--established~\cite{abate2019jour,abate2021extacc,patrignani2021rsc}.
@@
 \end{definition}
+
+\begin{example}[Double Free in Bluetooth Subsystem]
+  Consider \texttt{CVE-2021-3564}~\cite{doublefree-bluetooth}, one of many submissions for a double--free vulnerability.
+  The bug arises due to a race condition where the context--level function \texttt{hci\_cmd\_work} was not expected to behave maliciously, since it resides in the same source--code repository where the vulnerability occurs.
+  Nevertheless, the component--level code of \texttt{hci\_dev\_do\_open} is linked with \texttt{hci\_cmd\_work} and does not atomically check wether the pointer has been freed already:
+  Therefore, \texttt{hci\_dev\_do\_open} does not satisfy the no--double--frees property robustly, since there is an implementation for \texttt{hci\_cmd\_work} that results the whole program to violate that property.
+\end{example}
 

### 2023-07-09 popl/main.tex
Add rtp violation example

@@
 \end{definition}
+
+\begin{example}[Types]
+  Suppose $\src{L}$ is a statically typed language similar to \texttt{C} and $\trg{T}$ is dynamically typed, where both share the same syntax up to dynamic type checks.
+  Consider the following $\src{L}$ component and its compiled version below.
+  \begin{center}
+  $
+  \begin{array}{l}
+    \src{\lfunction{foo}{(char* x, int\ n)}{\lifz{valid\_ptr(x, n, sizeof(char))}{x[0]}{-1}}} \\
+    \trg{\lfunction{foo}{(\phantom{char*\ }x, \phantom{int\ }n)}{\lifz{valid\_ptr(x, n, sizeof(char))}{x[0]}{-1}}}
+  \end{array}
+  $
+  \end{center}
+  While the implementation the compiler emits may look correct, it does not check that the provided argument is of the right type.
+  Even though the pointer $\trg{x}$ is checked for validity, the context $\trg{foo((int*)y, 1)}$ is able to provoke a read out of bounds.
+  Suppose the component transferred control to the context and passed ownership of a $\trg{char}$ pointer $\trg{y}$ sized $\trg{1}$ cells, the context now calls the component again but casts this buffer to an $\trg{int*}$ prior to that call.
+  The pointer is valid for one $\trg{char}$--sized memory cell, as expected, but the actual read operation now returns $\trg{sizeof(int)}$ many bytes instead of just $\trg{sizeof(char)}$ many.
+  Thus, even if $\src{foo}$ may have been robust with respect to the no--reads--out--of--bounds property, its compiled counterpart is not and therefore the compiler fails to attain \Thmref{def:rtp}.
+\end{example}
 

### 2023-07-09 popl/main.tex
sec3 sketch examples

@@
 
+\begin{example}[A program satisfying \gls{tms}]
+  Consider the following program which, when run, yields the trace $\ev{Alloc}\ \loc[_{1}]\ 17;\comp;\unlock\cdot\ev{Use}\ \loc[_{1}]\ 5;\comp;\unlock\cdot\ev{Use}\ \loc[_{1}]\ 5;\comp;\unlock\cdot\ev{Dealloc}\ \loc[_{1}];\comp;\unlock$, where $\loc[_{1}]$ is some memory address determined by the semantics of the programming language.
+  \begin{lstlisting}[language=c,basicstyle=\ttfamily,morekeywords={new,delete}]
+int main() {  // allocates 17 integers initialized to 2
+  int* x = new[17](2);
+
+  x[5] = 12;
+  int y = x[5];
+  delete x;
+
+  return y;
+}
+  \end{lstlisting}
+  Relating the trace to abstract monitor events yields $\abstrace=\absev{Alloc}\ \loc[_{1}]\ \comp\cdot\absev{Use}\ \loc[_{1}]\ \comp\cdot\absev{Use}\ \loc[_{1}]\ \comp\cdot\absev{Dealloc}\ \loc[_{1}]\ \comp$.
+  It is easy to see that $\montmssafe{\abstrace}$.
+  From \Thmref{lem:mon:tmsafe}, it follows that the program satisfies \Thmref{def:trace:tmsdef}.
+\end{example}
+
 \subsubsection{Monitor for \glsfirst{sms}}
@@
 
+\begin{example}[\texttt{strncpy}]\label{ex:strncpy:sms}
+  Consider the \texttt{strncpy} functions from \Cref{ex:strncpy}.
+  \MKin{do trace and explain why monitor finds out that it violates sms prop}
+\end{example}
+
 \subsubsection{Combining \glsfirst{tms} and \glsfirst{sms} Monitors to obtain \glsfirst{ms}}\label{subsubsec:ms}
@@
 \end{lemma}
+
+\begin{example}[DIT on vs DIT off]
+  \MKin{fill in the blank}
+\end{example}
 

### 2023-07-09 popl/main.tex
intro todos

@@
 Without memory safety, private data can easily leak or corrupt~\cite{lemay2021ccc}.
-Because of this, \gls{ms} is a necessity for programs to be \gls{cct}, which is an important security property for secrecy.
-Programs that attain \gls{cct} are secure with respect to timing attacks~\cite{kocher1996timing}.
-Timing attacks may expose secrets from a simple performance analysis.
-These attacks have become more and more relevant since \citeauthor{kocher1996timing}'s paper, especially given the era of cloud--computing~\cite{aviram2010cloudtime,kumar2019cloudsecsurvey}, where programs share the same hardware but are supposed to run in a sandboxed environment that does not leak secrets.
+Because of this, \gls{ms} is a requirement for programs to attain an even stronger security property, namely \gls{cct}.
+Programs that attain \gls{cct} are secure with respect to timing attacks, and this is the gold standard for secure applications, including cryptographic ones since timing attacks can expose secrets via a simple performance analysis~\cite{kocher1996timing}.
+Timing attacks have become more and more relevant recently, especially given the era of cloud-computing~\cite{aviram2010cloudtime,kumar2019cloudsecsurvey}, where programs share the same hardware but are supposed to run in a sandboxed environment that does not leak secrets.
 Unfortunately, several attacks within such a cloud environment have been demonstrated as viable strategies to extract secrets~\cite{mehmet2015getoff,flowers2022zeroday,atya2019catchme,venkatanathan2015placevul}.
 
-Even though mitigations for such attacks exist~\cite{bond2017vale,almeida2017jasmin}, it has been shown in literature that some compilers may optimize these mitigations away~\cite{barthe2018sec}.
-This is of course not what the programmer expects and necessitates the need for a {\em secure} compiler that ensures security properties hold even when the program runs in a hostile environment, i.e., interacts with other programs by any means.
-Because of this hostile environment, secure compilers attain security by relying on certain protection mechanisms, such as ... and ...\MK{cits}.
+Even though mitigations for such attacks exist~\cite{bond2017vale,almeida2017jasmin}, the literature shows that some compilers may optimize these mitigations away~\cite{barthe2018sec}.
+This is of course not what the programmer expects and necessitates the need for a {\em secure} compiler that ensures security properties hold even when the program runs in a hostile environment, interacting with other possibly malicious programs.
+Because of this hostile environment, secure compilers attain security by relying on protection mechanisms, such as ... and ...\MK{cits}.
 The aim of this paper is to devise the theoretical foundation to be able to build a secure compiler that is capable of dealing with the following example.
+\MPin{
+	this does not really introduce the compositionality issue.
+	i think the problem is the last sentence.
+	try and divert the story after the 'because' line, to go into the need for composing compilers
+}
 \begin{example}[strncpy]\label{ex:strncpy}
-Consider the program bolow which copies a \texttt{C}-style string \texttt{x} into \texttt{y} up to a length of \texttt{n}:
+Consider the program below which copies a \texttt{C}-style string \texttt{x} into \texttt{y} up to a length of \texttt{n}:
 
@@
 
-For sake of simplicity it is assumed that the buffer \texttt{y} is at least as large as \texttt{n} as well as \texttt{x} terminated by \texttt{'\textbackslash 0'}.
+For sake of simplicity, assume that the buffer \texttt{y} is at least as large as \texttt{n} as well as \texttt{x} terminated by \texttt{'\textbackslash 0'}.
 Even under these assumptions, there is a subtle \gls{ms} bug in the program which can be exploited to read sensitive data:
-The bounds check, i.e., \texttt{i < n}, should happen prior to the access to memory location \texttt{x[i]}.
-Because otherwise, if \texttt{n} is larger than the length of the array object \texttt{x} points to, the memory location past this array object is being read, which is an out--of--bounds memory access.
-Unfortunately, even though flipping the conditions does fix the \gls{ms} issue, \gls{cct} is violated, so the program is susceptible to timing attacks.
-This is because, the loop can terminate early: As soon as the string--terminating character \texttt{'\textbackslash 0'} is encountered, the loop exits.
-Thus, program execution time is proportional to the length of the array object \texttt{x}.
-The fix commonly employed in such a case is to rewrite the program to run a fixed number of iterations regardless of the length of \texttt{x} or \texttt{y}.
+The bounds check, i.e., \texttt{i < n}, should happen {\it prior} to the access to memory location \texttt{x[i]}.
+Otherwise, if \texttt{n} is larger than the length of the array object \texttt{x} points to, the memory location past this array object is read, which is an out--of--bounds memory access.
+Unfortunately, even though flipping the conditions does fix the \gls{ms} issue, it still violates \gls{cct}, so the program is susceptible to timing attacks.
+This is because,\MP{this is a german comma.
+english does not work this way :) please erase.} the loop can terminate early: As soon as the string--\MP{only 1 dash here (also happened before but i already edited)}terminating character \texttt{'\textbackslash 0'} is encountered, the loop exits.
+Thus, the program execution time is proportional to the length of the array object \texttt{x}.
+The fix commonly employed in such a case is to rewrite the program to run a {\it fixed} number of iterations regardless of the length of \texttt{x} or \texttt{y}.
 To this end, there must be an upper bound for the length of \texttt{x} and \texttt{y} and then the loop can just iterate until the induction variable \texttt{i} hits this upper bound.
@@
 These fixes alone are not enough to make this program robust, i.e., not violating either \gls{ms} or \gls{cct} in a malicious environment.
-Consider a program context that calls \texttt{strncpy} with a very large \texttt{n}, some uninitialised memory \texttt{x}, and \texttt{y} whatever.
-It is easy to see that such a function invocation leads to a series of reads from uninitialised memory, an immediate \gls{ms} issue.
-To mitigate these situtations, it is important that \texttt{strncpy} ensures that the provided arguments satisfy the assumptions being made.
-After all, the environment cannot be trusted and potentially behaves in a hostile manner.
-Unfortunately, these checks do not work in C without additional features, since there is no way to check wether a pointer points into a valid allocated region or not.
+Consider a program context that calls \texttt{strncpy} with a very large \texttt{n}, some uninitialised memory \texttt{x}, and \texttt{y} whatever\MP{? whatever?}.
+Such a function invocation leads to a series of reads from uninitialised memory, which is an immediate \gls{ms} issue.
+\MPin{
+	we have a very very high-level definition of MS, so the issue may not be clear
+}
+To mitigate these situations, it is important that \texttt{strncpy} ensures that the provided arguments satisfy the assumptions being made by the programmer, since the environment\MP{env?
+ctx?
+tetrminology} can behave in a hostile manner.
+Unfortunately, these checks do not work in C\MP{why are we talking of C now?} without additional features, since there is no way to check wether\MP{please use a spell checker, this word is written 'whether'} a pointer points into a valid allocated region or not.
 To enable this would require extra metadata~\cite{kwon2013lowfat,akritidis2009baggy} which must not be modifiable by a malicious context.
 So, for sake of simplicity, assume the language at hand provides an operator \texttt{is\_valid(p, s, n)} that faithfully checks wether a pointer \texttt{p} points into a valid array of objects sized \texttt{s} of length \texttt{n} or not --- in constant time.
-This could be an abstraction that the operating system provides.
+This could be an abstraction that the operating system provides.\MP{cits absolutely needed and the narrative needs to be stronger, using those cits}
 With this feature at hand, it is possible to modify the original \texttt{strncpy} to be robust, given this is the compiled program without optimisations:
@@
 
+\MPin{
+	this bit should be introduced earlier
+}
 As stated earlier, this work wants to establish the theoretical foundation to ensure the secure compilation of programs like \texttt{strncpy} (\Cref{ex:strncpy}) to satisfy properties \gls{ms} and \gls{cct} robustly.
-Secure compilation in terms of robust preservation is well--established~\cite{abate2019jour,abate2021extacc,patrignani2021rsc}.
-Unfortunately, at the time of writing this paper, it is unclear if and how the theory of robust preservation would support modularity.
+Secure compilation in terms of robust preservation is well-established~\cite{abate2019jour,abate2021extacc,patrignani2021rsc}.
+Unfortunately, currently, it is unclear how the theory of robust preservation supports modularity.
 Missing out on modularity would mean that one has to make a secure compiler for all possible properties {\em or} for a combination of \gls{ms} and \gls{cct}.
-Ideally, both for proof engineering and reusability factor, a secure compiler for just \gls{ms} and a secure compiler for just \gls{cct} can compose in some way to get a secure compiler for the combination of the two.
+Ideally, both for proof engineering and reusability, a secure compiler for just \gls{ms} and a secure compiler for just \gls{cct} can compose in some way to get a secure compiler for the combination of the two.
 This also reflects real--world compilation chains, where several compilation passes perform optimisations~\cite{lattner2004llvm,googlev8,androidstudio,kuepper2023cryptopt,manjikian1997fusion,wegman1991ccp} or source--code instrumentations~\cite{nagarakatte2009soft,nagarakatte2010cets,akritidis2009baggy,dhumbumroong2020boundwarden,jung2021pico,nam2019framer,shankaranarayana2023tailcheck,younan2010paricheck,zhou2023fatptrs,bond2017vale,almeida2017jasmin,kuepper2023cryptopt,cauligi2019fact} that enforce that the compiled program fulfills a property of interest.
@@
 % keyword: pass
-In turn, this paper introduces a framework for reasoning about the composition of secure and optimising compiler passes and it showcases the power of this framework by instantiating it on a multi-pass compilation chain.
+Thus, this paper introduces a framework for reasoning about the composition of secure and optimising compiler passes and it showcases the power of this framework by instantiating it on a multi-pass compilation chain.
 To this end, this paper first discusses how to compose security properties, such as temporal and spatial memory safety into general memory safety, and cryptographic constant time.
@@
 The results showcase how the framework allows the kind of formal security reasoning that compiler writers already want (and do), obtaining precise, compositional security reasoning while providing minimal (and modular) proof effort.
-%\MPin{
-%	last line: not the best.
-%}
+\MPin{
+	last line: connect with the many citations of the previous paragraph
+}
 
 In summary, this paper makes the following contributions:
-
 \begin{itemize}
@@
         %\Cref{sec:compprop} presents a formal framework to reason about compositions of security properties and demonstrate its usage with concrete examples:
-        First, this paper studies the formalisation and the composition of security properties (\Cref{sec:compprop}) by focussing on those properties that are of interest for real-world compiler writers (as identified by the plethora of work enforcing such properties individually~\cite{akritidis2009baggy,nagarakatte2009soft,nagarakatte2010cets,dhumbumroong2020boundwarden,jung2021pico,nam2019framer,shankaranarayana2023tailcheck,younan2010paricheck,zhou2023fatptrs,bond2017vale,almeida2017jasmin,kuepper2023cryptopt,cauligi2019fact}\MK{even more cits?}).
+        First, this paper studies the formalisation and the composition of security properties (\Cref{sec:compprop}) by focussing on those properties that are of interest for real-world compiler writers (as identified by the plethora of work enforcing such properties individually~\cite{akritidis2009baggy,nagarakatte2009soft,nagarakatte2010cets,dhumbumroong2020boundwarden,jung2021pico,nam2019framer,shankaranarayana2023tailcheck,younan2010paricheck,zhou2023fatptrs,bond2017vale,almeida2017jasmin,kuepper2023cryptopt,cauligi2019fact}).
         % 
@@
         %Concretely, it investigates upper and lower compositions, as well as the sequential composition.
-        For this, this paper studies the sequential composition of compilers, as typically done when engineering compilation chains, as well as an upper and a lower composition that discuss multi--in- and multi--output compilers.
+        For this, this paper studies the sequential composition of compilers, as typically done when engineering compilation chains, as well as an upper and a lower composition that discuss multi--in- and multi--output compilers.\MP{upper and lower?}
         % 
@@
 
-  \item The key contributions are formalized in the Coq proof assistant and the paper indicates this with \CoqSymbol.
+  \item The key contributions of this paper are formalized in the Coq proof assistant and the paper indicates this with \CoqSymbol.
 \end{itemize}
@@
 
-\contribution{Open Source \& Technical Report} A technical report and the Coq formalisation are available as supplementary material.\MKin{cite/link}
+\contribution{Open Source \& Technical Report} A technical report with the omitted formal details, lemmas and proofs, as well as the Coq formalisation are available as supplementary material.
+% \MKin{cite/link}
 

### 2023-07-09 popl/main.tex
ssssshit

@@
 
-\section{Example Properties and their Monitors\pages{1.5}}\label{sec:compprop}
-
-This section introduces a trace model (\Cref{subsec:propdefs}) and defines some example properties (\gls{tms}, \gls{sms}, \gls{ms}, and \gls{scct}).
-These properties are of practical importance and also of interest in the case study (\Cref{sec:casestud:defs,sec:casestud:rtp}) this paper presents.
+\section{Security Properties: Formalisation, Enforcement and Composition\pages{1.5}}\label{sec:compprop}
+
+This section introduces a trace model and uses it to define the key properties of interest for this paper: \gls{tms}, \gls{sms}, \gls{ms}, and \gls{scct} (\Cref{subsec:propdefs}).
+These properties are of practical importance (as mentioned in \Cref{sec:introduction}) and also of interest in the case study (\Cref{sec:casestud:defs,sec:casestud:rtp}) this paper presents later. 
+\MPin{
+	missing 3.2 ref? added below?
+}
+Then, this section introduces monitors for each of the properties presented earlier (\Cref{??}).
+
 \subsection{Specification Trace Model}\label{subsec:propdefs}
@@
 
-The specification trace model defines events as either the empty event ($\emptyevent$), a crash ($\lightning$), or as tuples consisting of a pre--event, a control--tag, and a security--tag.
-The purpose of the model is to define properties, such as a stricter variant of cryptographic constant time.
-To this end, security--tags indicate wether an event contains sensitive information ($\lock$) or not ($\unlock$), while control--tags state wether the context ($\ctx$) or the component ($\comp$) are responsible for emitting the event.
-The latter is necessary to be able to ignore actions done by a spurious context that, e.g., immediately deallocates a memory location twice, thus violating \gls{tms}.
+The specification trace model defines events as either the empty event ($\emptyevent$), a crash ($\lightning$), or as tuples consisting of a pre-event, a control-tag, and a security-tag.
+The purpose of the model is to define key security properties of interest, such as \MP{ms?} a stricter variant of cryptographic constant time.
+To this end, security-tags indicate wether an event contains sensitive information ($\lock$) or not ($\unlock$), while control-tags state wether the context ($\ctx$) or the component ($\comp$) are responsible for emitting the event.
+The latter is necessary to be able to ignore actions done by a spurious context that, e.g., immediately deallocates a memory location twice, thus violating \gls{tms}. \MP{cits?}
 %Even though this could be used for blame~\cite{patrignani2023blame}, this is not what this paper is concerned about.
-Lastly, pre--events describe the actual kind of event that happened.
+Lastly, pre-events describe the actual kind of event that happened.
 One such kind is the allocation event ($\ev{Alloc\ \loc\ n}$) that fires whenever a program claims $n$ cells of memory and stores them at address $\loc$.
-Likewise, deallocation ($\ev{Dealloc\ \loc}$) announces that the object at location $\loc$ is freed.
-These two events alone allow already a partial description of \gls{tms} by requiring that, e.g., there is only one deallocation event that carries a location $\loc$.
+Dually, deallocation ($\ev{Dealloc\ \loc}$) announces that the object at location $\loc$ is freed.
+These two events alone allow\MP{allow who?
+allow needs an object, this is improper english} already a partial description of \gls{tms} by requiring that, e.g., there is only one deallocation event that carries a location $\loc$.
 To be able to express \gls{sms}, there is also an event to describe reads from and writes to memory ($\ev{Use\ \loc\ n}$).
@@
 Moreover, an instruction whose timing is data--dependent must not have a secret as an operand.
-Typical operations with data--dependent timing are branches and certain binary operations, such as division\footnote{This is highly architecture dependent, but division is an operation that serves as a classic example for a data--dependent timing instruction.}.
+Typical operations with data--dependent timing are branches and certain binary operations, such as division\footnote{This is highly architecture-dependent\MPin{hyphen needed}, but division is an operation that serves as a classic example for a data-dependent timing instruction.}.
 Both operations are also modelled in the specification trace model ($\ev{Branch\ n}$ and $\ev{Binop\ n}$).
@@
 \subsection{Monitors\pages{1.5}}\label{sec:monitors}
-This section introduces monitors for each of the properties presented earlier (\Cref{subsec:propdefs}).
-The use of monitors is because they have some structure useful for inductive proofs, whereas traces lack this structure, thus making proofs difficult.
+
+The use of monitors is because they have some structure useful for inductive proofs, whereas traces lack this structure, thus making proofs difficult.\MP{not true.
+safety props are enforced by monitors}
 To lessen the burden when proving, each monitor makes use of a custom trace model that contains only the relevant information related to the property the monitor checks.

### 2023-07-09 popl/main.tex
feedback on 4

@@
 
-This section presents the key meta-theoretic results of this paper and builds on prior work~\cite{kruse2022csc}\MK{mention this at all? may reveal identity even though its double-blind}.
-A main result is that secure compilers in the robust preservation framework~\cite{abate2019jour} compose.
-This is not intuitive in the sense that there are a number of examples in the security domain where composition does not work~\cite{canetti2006univcomp,mccullough2012compo,fabian2022automatic}.
-For starters, the sequential composition of compilers $\cc{\src{L}}{\trg{L}}$ and $\cc{\trg{L}}{\obj{L}}$ is as expected, i.e., given an $\src{L}$ program $\src{p}$, obtain its compiled $\obj{L}$ program by plugging $\src{p}$ into $\cc{\src{L}}{\trg{L}}\circ\cc{\trg{L}}{\obj{L}}$.
+This section presents the key meta-theoretic results of this paper concerning sequential compiler composition (and of optimisation passes) (\Cref{sec:sequential}) and concerning other kinds of compiler composition (\Cref{sec:other-compos}).
+% and builds on prior work~\cite{kruse2022csc}\MK{mention this at all?
+may reveal identity even though its double-blind}.
+% MP: not needed at all, ever, workshop work does not need to be cited in the main work
+
+\subsection{Secure Sequential Composition}\label{sec:sequential}
+
+The main result is that secure compilers in the robust compilation framework~\cite{abate2019jour} compose {\em sequentially}.
+This is not intuitive in the sense that there are a number of examples in the security domain where sequential composition does not work~\cite{canetti2006univcomp,mccullough2012compo,fabian2022automatic}.\MP{?canetti?}
+The sequential composition of compilers $\cc{\src{L}}{\trg{L}}$ and $\cc{\trg{L}}{\obj{L}}$ is defined as follows: given an $\src{L}$ program $\src{p}$, its compiled $\obj{L}$ counterpart is obtained by plugging $\src{p}$ into $\cc{\src{L}}{\trg{L}}\circ\cc{\trg{L}}{\obj{L}}$.
+% MP: i know, i added a passive verb. here it works. i think. but i appreciate the effort!
 Now, given \bul{$\cc{\src{L}}{\trg{L}}$ robustly preserves $\class[_{1}]$} and \rul{$\cc{\trg{L}}{\obj{L}}$ robustly preserves $\class[_{2}]$}, it follows that \oul{their sequential composition $\cc{\src{L}}{\trg{L}}\circ\cc{\trg{L}}{\obj{L}}$ robustly preserves the intersection of classes $\class[_{1}]$ and $\class[_{2}]$}.
@@
 
-Notably, real-world compilation chains also perform a series of passes whose main purpose is not necessarily to translate from one language to another, but to, e.g., optimise the code or enforce a certain property.
-Both examples can be seen in the wild~\cite{nagarakatte2009soft,nagarakatte2010cets,akritidis2009baggy,wegman1991ccp,manjikian1997fusion}, it is infeasible to list them all.
-Consider the compilation passes of LLVM\cite{lattner2004llvm} whose source-code optimisations are industry--competitive.
-In LLVM, the pass manager can be used to schedule passes in a way where the optimisations yield the best results.
-Take \gls{cf} and \gls{dce}.
+\subsubsection{Securing Optimisations}
+\MPin{note that i'm always using british spelling. optimisation VS optimization. etc. }
+
+Notably, real-world compilation chains also perform a series of (sequential) passes whose main purpose is not necessarily to translate from one language to another, but to, e.g., optimise the code or enforce a certain property.
+\MPin{
+	do not say the negation of something before the non-negation.
+	say what things are first, and what they are not later (if at all)
+}
+Both examples can be seen in the wild~\cite{nagarakatte2009soft,nagarakatte2010cets,akritidis2009baggy,wegman1991ccp,manjikian1997fusion}, it is infeasible to list them all\MP{survey to cite?
+this is a bit informal}.
+Consider the compilation passes of LLVM~\cite{lattner2004llvm} whose source-code optimisations are industry--competitive\MP{meaning?}.
+In LLVM, the pass manager\MP{meaning} can be used to schedule passes in a way where the optimisations yield the best results.
+Take \gls{cf} and \gls{dce}. \MPin{
+	brief explanation?
+}
 Unfortunately, the order in which \gls{cf} and \gls{dce} are performed influences the final result of the compilation (see \Cref{fig:cfdceex}).
-This {\em phase ordering problem} is well--known in literature and a practical solution is to simply perform a fixpoint iteration.
-Compiler engineers typically try to find an order of optimisations that yields well--optimised programs for either code size~\cite{cooper1999geneticphases} or performance~\cite{kulkarni2006exhaustivephase}.
-The following corollary justifies that any such order of compilation passes is valid with respect to security.
+\MPin{figure here, or closer, not far away}
+This {\em phase ordering problem} is well--known in literature and a practical solution is to simply perform a fixpoint iteration\MP{of what}.\MP{cit?}
+Compiler engineers typically try to find an order of optimisations that yields well-optimised programs for either code size~\cite{cooper1999geneticphases} or performance~\cite{kulkarni2006exhaustivephase}.
+\Cref{corr:swappable} justifies that any such order of compilation passes is valid with respect to security.
 So, given two compilation passes \bul{$\cc[_{1}]{\trg{L}}{\trg{L}}$, $\cc[_{2}]{\trg{L}}{\trg{L}}$, both robustly preserving class $\class[_{1}]$ or $\class[_{2}]$, respectively}, \oul{for any order of their composition the composed compiler robustly preserves the intersection of $\class[_{1}]$ and $\class[_{2}]$}.
@@
 
-Besides the sequential composition, there are two other compositions that are more at the language- than the compiler--level.
+\subsection{Secure Upper and Lower Composition}\label{sec:other-compos}
+Besides sequential composition, there are two other compositions that are more at the language- than the compiler-level\MP{namely ...}.
 Nevertheless, consider $\trg{JVM BC}$, which is a popular target for programming language designers due to its high performance and relevance in industry.
 As an example, $\src{Java}$ and $\obj{Kotlin}$ compile to $\trg{JVM BC}$.
-Technically speaking, they both lower to class files and $\obj{Kotlin}$ objects are considered to be the same as $\src{Java}$ objects at that point.
+Technically speaking, they both lower\MP{undefined verb. no need for synonyms here.} to class files and $\obj{Kotlin}$ objects are considered to be the same as $\src{Java}$ objects at that point.
 A compiler that accepts both $\src{Java}$ and $\obj{Kotlin}$ code translating to the same target language, or rather, intermediate representation, performs a kind of {\em upper} composition.
@@
 The corresponding theorem is not listed, since there is no general definition of this kind of composition.
+\MPin{
+	mmm need to think about this.
+	this looks like discussion section material
+}
+\MPin{
+	we could turn this into a : "if you expected upper adn lower, they are language composition ops, not compiler ones, so that's out of scope"
+}
 
@@
   Every edge is a compilation pass and the label on the edge states what the pass does, i.e., \gls{cf} or \gls{dce}. %
-  The source code in the nodes is a glorified compiler intermediate representation and the code gets more optimised towards the righthandside of the figure.}\label{fig:cfdceex}
+  The source code in the nodes is a glorified compiler intermediate representation and the code gets more optimised towards the right hand side of the figure.}\label{fig:cfdceex}
 \end{figure}
@@
 %\section{Case Study: Memory Safe and Cryptographic Constant Time Code}\label{sec:casestud}
-\section{Case Study: Preliminary Definitions\pages{2}}\label{sec:casestud:defs}
-
-This section defines programming languages that the secure compilers defined in the next section will make use of.
-To this end, this section defines the languages $\Ltms$, $\Ltrg$, $\Lms$, and $\Lscct$.
-In this paper, $\Ltms$ is the only statically typed language and exhibits the property that all well-typed programs are \gls{tms}.
-However, $\Ltms$ programs are not \gls{sms}.
+\section{Case Study: Language Formalisations\pages{2}}\label{sec:casestud:defs}
+
+This section defines programming languages that the secure compilers defined in the next section will make use of\MP{make use of -> use! 3 words when 1 is sufficient}.
+To this end, this section defines the languages $\Ltms$, $\Ltrg$, $\Lms$, and $\Lscct$ which share many common elements (presented in \Cref{subsec:cs:defs}).
+In this paper, $\Ltms$ is the only statically typed language and exhibits the property that all well-typed programs are \gls{tms} (\Cref{??}).
+However, not all $\Ltms$ programs are \gls{sms}.\MP{reworded. was imprecise}
 That is, there are well--typed $\Ltms$ programs that perform an out--of--bounds access.
-In contrast, $\Ltrg$ is untyped and does not provide any guarantees with regards to \gls{ms}.
-$\Lms$ is exactly the same language as $\Ltrg$, but this paper still distinguishes the two for sake of readability.
+In contrast, $\Ltrg$ is untyped and does not provide any guarantees with regards to \gls{ms} (\Cref{??}).
+$\Lms$ is exactly the same language as $\Ltrg$, but this paper still distinguishes the two for sake of readability (\Cref{??}).
 All three languages --- so $\Ltms$, $\Ltrg$, and $\Lms$ --- assume \gls{cct} to hold.
+
 Even though it is possible to write programs that, e.g., return early in loops, which is trivially not \gls{cct}, they can be transformed to adhere to \gls{cct}~\cite{cauligi2019fact}.
-The language $\Lscct$ allows violating \gls{cct} by emitting events on, e.g., branching and division, that contain secrets.
-As a countermeasure, the language provides a way to read and write to a {\em model--specific register} that enables a ``\gls{cct}-mode''.
+\MPin{
+	no, it's different. 
+	there is no CCT because we asume it's out of the programmer's concers and since CT is arch-dependent, the compiler should preserve CT, not the programmer, via software
+}
+
+Language $\Lscct$ allows violating \gls{cct} by emitting events on, e.g., branching and division, that contain secrets (\Cref{??}).
+However, $\Lscct$ also contains a countermeasure to enforce \gls{cct}, the language provides a way to read and write to a {\em model--specific register} that enables a ``\gls{cct}-mode''.
+\MPin{ref to what real world stuff justifies this}
 
 \subsection{Shared Language Definitions}\label{subsec:cs:defs}
-
 \begin{gather*}
   \begin{aligned}
-  \mi{(Expressions)}~\expr&\bnfdef\ x \mid \valueexpr \mid \lbinop{\expr[_{1}]}{\expr[_{2}]} \mid \lget{x}{\expr} \mid \lnew{\expr[_{1}]}{\expr[_{2}]} \mid \ldelete{x} \mid \lset{x}{\expr[_{1}]}{\expr[_{2}]} \\
-    & \mid \llet{x}{\expr[_{1}]}{\expr[_{2}]}  \mid \lreturn{\expr} \mid \lcall{g}{\expr} \mid \lifz{\expr}{\expr[_{1}]}{\expr[_{2}]} \mid \labort \\
-  \mi{(Symbols)}~&\function\bnfdef\ \lfunction{g}{x}{e}\hspace{0.5cm}
-  \mi{(Libraries)}~\library\bnfdef\ \hole{\cdot} \mid \function,\library \\
-  \mi{(Programs)}~&\prog{\library_{\ctx}}{\library_{\comp}}\hspace{0.5cm}
-  \mi{(Substitutions)}~\substvar\bnfdef \hole{\cdot} \mid \subst{\valueexpr}{x},\gamma \\
+  \mi{(Expressions)}~&\expr\bnfdef\ x \mid \valueexpr \mid \lbinop{\expr[_{1}]}{\expr[_{2}]} \mid \lget{x}{\expr} \mid \lnew{\expr[_{1}]}{\expr[_{2}]} \mid \ldelete{x} \mid \lset{x}{\expr[_{1}]}{\expr[_{2}]} \\
+    & \mid \llet{x}{\expr[_{1}]}{\expr[_{2}]}  \mid \lreturn{\expr} \mid \lcall{g}{\expr} \mid \lifz{\expr}{\expr[_{1}]}{\expr[_{2}]} \mid \labort
+  \end{aligned}
+  \\
+  \begin{aligned}
+  \mi{(Symbols)}~&\function\bnfdef\ \lfunction{g}{x}{e}
+  &
+  \mi{(Libraries)}~&\library\bnfdef\ \hole{\cdot} \mid \function,\library \\
+  \mi{(Programs)}~&\prog{\library_{\ctx}}{\library_{\comp}}
+  &
+  \mi{(Substitutions)}~&\substvar\bnfdef \hole{\cdot} \mid \subst{\valueexpr}{x},\gamma
   \end{aligned}
 \end{gather*}
+\MPin{
+	this latex had many issues.
+	the \& were at different spots
+	it used hspace to space things instead of \&
+	and it did not use \& to align the following things on a line
+	and it did not use multiple aligned into the gather
+	(what's the point of gather if you don't have several aligned inside? use a single align* instead)
+}
 
 Above is a fraction of the shared syntax of all the programming languages of this paper.
-The syntax and semantics are mostly standard.
-However, when allocating pointers with $\lnew{\expr[_{1}]}{\expr[_{2}]}$, the expression $\expr[_{1}]$ determines the initial value of the objects and $\expr[_{2}]$ the overall size of an array.
+\MP{why a fraction? what exprs are missing}
+The syntax and semantics are mostly standard.\MP{and therefore omitted}
+However, when allocating pointers\MP{arrays?} with $\lnew{\expr[_{1}]}{\expr[_{2}]}$, the expression $\expr[_{1}]$ determines the initial value of the objects\MP{cells?} and $\expr[_{2}]$ the overall size of an array.
 The data associated to pointers resides in the dynamic state $\Omega$ that contains sandboxed heaps, one for the context and the other for the component.
-Because of the sandboxing, locations pointing into heaps must be associated with a control--tag to know the heap they point into.
-Other metadata in $\Omega$ entails a list of callable objects with information on wether they belong to the context or the component.
+Because of the sandboxing, locations pointing into heaps must be associated with a control--tag\MP{symbol} to know the heap they point into.
+Other metadata in $\Omega$ entails a list of callable objects\MP{functions? objects are other things in PL} with information on wether they belong to the context or the component.
 That way, when calling a callable object, the semantics can take care of emitting appropriate events that signal a context switch.
@@
   \begin{aligned}
-  \mi{(Pre\text{--}Events)}~\event[_{b}]\bnfdef \ev{Alloc}\ \loc\ \valueexpr \mid \ev{Dealloc}\ \loc \mid \ev{Get}\ \loc\ \valueexpr &\mid \ev{Set}\ \loc\ \valueexpr\ \valueexpr['] \mid \ev{Call}\ \comm\ g\ \valueexpr \mid \ev{Ret}\ \comm\ \valueexpr \mid \ev{Start} \mid \ev{End}\ \valueexpr \\
-  \mi{(Events)}~\event&\bnfdef \emptyevent \mid \lightning \mid \tup{\event[_{b}]}{\sandboxtag} \\
+  \mi{(Pre\text{-}Events)}~&\event[_{b}]\bnfdef \ev{Alloc}\ \loc\ \valueexpr \mid \ev{Dealloc}\ \loc \mid \ev{Get}\ \loc\ \valueexpr &\mid \ev{Set}\ \loc\ \valueexpr\ \valueexpr['] \mid \ev{Call}\ \comm\ g\ \valueexpr \mid \ev{Ret}\ \comm\ \valueexpr \mid \ev{Start} \mid \ev{End}\ \valueexpr \\
+  \mi{(Events)}~&\event\bnfdef \emptyevent \mid \lightning \mid \tup{\event[_{b}]}{\sandboxtag} \\
   \end{aligned}
@@
 \subsection{$\Lscct$: An Unsafe Language}\label{subsec:lscct}
+\MPin{
+	unsafe? 
+	a possibly-CCT language?
+}
 
@@
 
-\section{Case Study: Secure Compiler Compositions in the Wild\pages{7}}\label{sec:casestud:rtp}
+\section{Case Study: Composing Secure Compiler Passes and Optimisations \pages{7}}\label{sec:casestud:rtp}
 

### 2023-07-09 popl/main.tex
intro feedback

--- 2023-07-09 popl/main.tex
@@
 %\acmBooktitle{Woodstock '18: ACM Symposium on Neural Gaze Detection,
-%  June 03--05, 2018, Woodstock, NY} 
+%  June 03-05, 2018, Woodstock, NY}
 \acmPrice{15.00}
@@
 
-  Memory safety necessitates secrecy, since breaking it can lead to an arbitrarily large attack space, i.e., from simply reading a secret up to alteration of control--flow in order to execute arbitrary code.
+  Memory safety necessitates secrecy, since breaking it can lead to an arbitrarily large attack space, i.e., from simply reading a secret up to alteration of control-flow in order to execute arbitrary code.
   However, secrecy cannot be achieved with just memory safety alone, since, e.g., private data could be leaked by differences in execution time for varying program inputs.
@@
 
-  Unfortunately, it is an open question wether secure compilers can be engineered in a modular way.
+  Unfortunately, it is an open question whether secure compilers can be engineered in a modular way.
   Ideally, compiler engineers develop two different compilation passes, one that ensures memory safety and another one that is security preserving with respect to cryptographic constant time.
-%, because this style of engineering is well--established for its simplifying divide--and--conquer approach as well as enhanced reusability.
-
-  This paper answers that question positively: The composition of compilers that are secure with respect to some security--properties yields the intersection of these properties.
-  This article discusses different variants of compositions of secure compilers and instantiates parts of these results in a case--study.
+%, because this style of engineering is well-established for its simplifying divide-and-conquer approach as well as enhanced reusability.
+
+  This paper answers that question positively: The composition of compilers that are secure with respect to some security-properties yields the intersection of these properties.
+  This article discusses different variants of compositions of secure compilers and instantiates parts of these results in a case-study.
   To this end, it presents an optimizing, secure compilation chain that preserves both memory safety and cryptographic constant time.
-%  Memory safety is achieved by composing secure compilers for both temporal and spatial memory safety, where the compiler for spatial memory safety is an instrumentation that inserts bounds--checks in front of pointer dereferences.
+%  Memory safety is achieved by composing secure compilers for both temporal and spatial memory safety, where the compiler for spatial memory safety is an instrumentation that inserts bounds-checks in front of pointer dereferences.
   %The compiler for cryptographic constant time exploits the use of a data (operand) timing independent mode and the employed optimisations of the compilation chain are constant folding and dead code elimination.
@@
 
-% Secure compilers emit code that is hardened against adversarial target--level attacks, such as memory corruption or alteration of control--flow.
-% It is an open question wether secure compilers compose in a well--behaved manner.
-% This paper answers this question positively: The composition of compilers that are secure with respect to some security--properties yields the intersection of these properties.
-% This article discusses different variants of compositions of secure compilers and instantiates parts of these results in a case--study.
+% Secure compilers emit code that is hardened against adversarial target-level attacks, such as memory corruption or alteration of control-flow.
+% It is an open question whether secure compilers compose in a well-behaved manner.
+% This paper answers this question positively: The composition of compilers that are secure with respect to some security-properties yields the intersection of these properties.
+% This article discusses different variants of compositions of secure compilers and instantiates parts of these results in a case-study.
 % To this end, it presents an optimizing, secure compilation chain that preserves both memory safety and cryptographic constant time.
-% Memory safety is achieved by composing secure compilers for both temporal and spatial memory safety, where the compiler for spatial memory safety is an instrumentation that inserts bounds--checks in front of pointer dereferences.
+% Memory safety is achieved by composing secure compilers for both temporal and spatial memory safety, where the compiler for spatial memory safety is an instrumentation that inserts bounds-checks in front of pointer dereferences.
 % The compiler for cryptographic constant time exploits the use of a data (operand) timing independent mode and the employed optimisations of the compilation chain are constant folding and dead code elimination.
@@
 
-\gls{ms} is a security property of programs.
+\gls{ms} is a security property of programs that, e.g., ensures that there are no out-of-bounds accesses as well as no reads or writes to uninitialized memory.
 Without memory safety, private data can easily leak or corrupt~\cite{lemay2021ccc}.
@@
 Even though mitigations for such attacks exist~\cite{bond2017vale,almeida2017jasmin}, the literature shows that some compilers may optimize these mitigations away~\cite{barthe2018sec}.
-This is of course not what the programmer expects and necessitates the need for a {\em secure} compiler that ensures security properties hold even when the program runs in a hostile environment, interacting with other possibly malicious programs.
-Because of this hostile environment, secure compilers attain security by relying on protection mechanisms, such as ... and ...\MK{cits}.
-The aim of this paper is to devise the theoretical foundation to be able to build a secure compiler that is capable of dealing with the following example.
-\MPin{
-	this does not really introduce the compositionality issue.
-	i think the problem is the last sentence.
-	try and divert the story after the 'because' line, to go into the need for composing compilers
-}
+This is of course not what the programmer expects and necessitates the need for a {\em secure} compiler that ensures security properties hold even when the program runs in a hostile context, i.e., where it interacts with other possibly malicious programs.
+Secure compilers can be described formally with the theory on {\em robust preservation}~\cite{abate2019jour,abate2021extacc,patrignani2021rsc}, which takes this interaction between components and potentially malicious contexts into account.
+Unfortunately, it is unclear if and how robust preservation supports modularity.
+Modularity is an inherent part of software engineering and real-world compilers are no exception to that~\cite{lattner2004llvm,googlev8}.
+Compilation passes are typically developed as a standalone module that can plugged into a compilation pipeline.
+Different kinds of such passes exist, e.g., there are usually several compilation passes that perform optimisations~\cite{lattner2004llvm,googlev8,androidstudio,kuepper2023cryptopt,manjikian1997fusion,wegman1991ccp} or source-code instrumentations~\cite{nagarakatte2009soft,nagarakatte2010cets,akritidis2009baggy,dhumbumroong2020boundwarden,jung2021pico,nam2019framer,shankaranarayana2023tailcheck,younan2010paricheck,zhou2023fatptrs,bond2017vale,almeida2017jasmin,kuepper2023cryptopt,cauligi2019fact} that enforce that the compiled program fulfills a property of interest.
+Without a modular framework, there is no hope to enable this separation of concerns, where, e.g., a secure pass that ensures \gls{ms} can be developed independently of a secure pass for \gls{cct}.
+%% MK: no idea how/where to fit this in or if this is even necessary
+%Because of the potentially hostile context, secure compilers attain security by relying on protection mechanisms, such as ... and ...\MK{cits}.
+The aim of this paper is to devise the theoretical foundation to be able to modularily build a secure compiler that is capable of dealing with the following example.
+%\MPin{ MK: hope it's solved now
+%	this does not really introduce the compositionality issue.
+%	i think the problem is the last sentence.
+%	try and divert the story after the 'because' line, to go into the need for composing compilers
+%}
 \begin{example}[strncpy]\label{ex:strncpy}
@@
 The bounds check, i.e., \texttt{i < n}, should happen {\it prior} to the access to memory location \texttt{x[i]}.
-Otherwise, if \texttt{n} is larger than the length of the array object \texttt{x} points to, the memory location past this array object is read, which is an out--of--bounds memory access.
+Otherwise, if \texttt{n} is larger than the length of the array object \texttt{x} points to, the memory location past this array object is read, which is an out-of-bounds memory access.
 Unfortunately, even though flipping the conditions does fix the \gls{ms} issue, it still violates \gls{cct}, so the program is susceptible to timing attacks.
-This is because,\MP{this is a german comma.
-english does not work this way :) please erase.} the loop can terminate early: As soon as the string--\MP{only 1 dash here (also happened before but i already edited)}terminating character \texttt{'\textbackslash 0'} is encountered, the loop exits.
+This is because the loop can terminate early: As soon as the string-terminating character \texttt{'\textbackslash 0'} is encountered, the loop exits.
 Thus, the program execution time is proportional to the length of the array object \texttt{x}.
@@
 
-These fixes alone are not enough to make this program robust, i.e., not violating either \gls{ms} or \gls{cct} in a malicious environment.
-Consider a program context that calls \texttt{strncpy} with a very large \texttt{n}, some uninitialised memory \texttt{x}, and \texttt{y} whatever\MP{? whatever?}.
-Such a function invocation leads to a series of reads from uninitialised memory, which is an immediate \gls{ms} issue.
-\MPin{
-	we have a very very high-level definition of MS, so the issue may not be clear
-}
-To mitigate these situations, it is important that \texttt{strncpy} ensures that the provided arguments satisfy the assumptions being made by the programmer, since the environment\MP{env?
-ctx?
-tetrminology} can behave in a hostile manner.
-Unfortunately, these checks do not work in C\MP{why are we talking of C now?} without additional features, since there is no way to check wether\MP{please use a spell checker, this word is written 'whether'} a pointer points into a valid allocated region or not.
+These fixes alone are not enough to make this program robust, i.e., not violating either \gls{ms} or \gls{cct} in a potentially malicious context.
+Consider a program context that calls \texttt{strncpy} with a very large \texttt{n}, some uninitialised memory \texttt{x}, and \texttt{y} does not matter for the sake of this example.
+While the context seems conceptually simple, uninitialized memory can appear in programs very subtly and invoking \texttt{strncpy} now leads to a series of reads from uninitialised memory, which is an immediate \gls{ms} issue that can have devastating real-world implications~\cite{uninit-0,uninit-1,uninit-2,uninit-3,uninit-4}.
+To mitigate these situations, it is important that \texttt{strncpy} ensures that the provided arguments satisfy the assumptions being made by the maintainer of \texttt{strncpy}, since the calling context can behave in a hostile manner.
+Unfortunately, these checks do not work in C\MP{why are we talking of C now?} without additional features, since there is no way to check whether a pointer points into a valid allocated region or not.
 To enable this would require extra metadata~\cite{kwon2013lowfat,akritidis2009baggy} which must not be modifiable by a malicious context.
-So, for sake of simplicity, assume the language at hand provides an operator \texttt{is\_valid(p, s, n)} that faithfully checks wether a pointer \texttt{p} points into a valid array of objects sized \texttt{s} of length \texttt{n} or not --- in constant time.
+So, for sake of simplicity, assume the language at hand provides an operator \texttt{is\_valid(p, s, n)} that faithfully checks whether a pointer \texttt{p} points into a valid array of objects sized \texttt{s} of length \texttt{n} or not -- in constant time.
 This could be an abstraction that the operating system provides.\MP{cits absolutely needed and the narrative needs to be stronger, using those cits}
@@
 
-\MPin{
-	this bit should be introduced earlier
-}
-As stated earlier, this work wants to establish the theoretical foundation to ensure the secure compilation of programs like \texttt{strncpy} (\Cref{ex:strncpy}) to satisfy properties \gls{ms} and \gls{cct} robustly.
-Secure compilation in terms of robust preservation is well-established~\cite{abate2019jour,abate2021extacc,patrignani2021rsc}.
-Unfortunately, currently, it is unclear how the theory of robust preservation supports modularity.
-Missing out on modularity would mean that one has to make a secure compiler for all possible properties {\em or} for a combination of \gls{ms} and \gls{cct}.
-Ideally, both for proof engineering and reusability, a secure compiler for just \gls{ms} and a secure compiler for just \gls{cct} can compose in some way to get a secure compiler for the combination of the two.
-This also reflects real--world compilation chains, where several compilation passes perform optimisations~\cite{lattner2004llvm,googlev8,androidstudio,kuepper2023cryptopt,manjikian1997fusion,wegman1991ccp} or source--code instrumentations~\cite{nagarakatte2009soft,nagarakatte2010cets,akritidis2009baggy,dhumbumroong2020boundwarden,jung2021pico,nam2019framer,shankaranarayana2023tailcheck,younan2010paricheck,zhou2023fatptrs,bond2017vale,almeida2017jasmin,kuepper2023cryptopt,cauligi2019fact} that enforce that the compiled program fulfills a property of interest.
-
-% keyword: pass
-Thus, this paper introduces a framework for reasoning about the composition of secure and optimising compiler passes and it showcases the power of this framework by instantiating it on a multi-pass compilation chain.
+This paper introduces a framework for reasoning about the composition of secure and optimising compiler passes and it showcases the power of this framework by instantiating it on a multi-pass compilation chain.
 To this end, this paper first discusses how to compose security properties, such as temporal and spatial memory safety into general memory safety, and cryptographic constant time.
 Then, this paper defines several secure compiler passes, where each is either preserving a different security property (e.g., temporal or spatial memory safety) or performing a security-preserving optimisation, (e.g., applying constant-folding or dead-code elimination).
-% keyword- end to end, composition
 Finally, this paper shows how to compose these secure compiler passes into a multi-pass compilation chain which is proven to provide end-to-end preservation of general memory safety and cryptographic constant time.
@@
 The results showcase how the framework allows the kind of formal security reasoning that compiler writers already want (and do), obtaining precise, compositional security reasoning while providing minimal (and modular) proof effort.
-\MPin{
-	last line: connect with the many citations of the previous paragraph
-}
+%\MPin{
+%	last line: connect with the many citations of the previous paragraph
+%}
 
@@
         % 
-        The resulting security property is the golden standard of safety properties that secure cryptographic code should have~\cite{lemay2021ccc}\MK{more cits?}.
+        The resulting security property is the golden standard of safety properties that secure cryptographic code should have~\cite{lemay2021ccc}.
 %        \MPin{
@@
         %Concretely, it investigates upper and lower compositions, as well as the sequential composition.
-        For this, this paper studies the sequential composition of compilers, as typically done when engineering compilation chains, as well as an upper and a lower composition that discuss multi--in- and multi--output compilers.\MP{upper and lower?}
+        For this, this paper studies compilers with multiple inputs or multiple outputs as well as the sequential composition of compilers, as typically done when engineering compilation chains.
         %
@@
         The formalisation of this case study showcases the power of the presented framework.
-        Namely, the divide--and--conquer approach is a viable strategy even for engineering secure compilers.
+        Namely, the divide-and-conquer approach is a viable strategy even for engineering secure compilers.
 %        This demonstrates that the usage of the presented framework is both simpler and more expressive compared to one monolithic security proof of the whole compilation chain.
@@
 This paper employs the security model where programs are written in a language whose semantics emits events $\event$.
-Events include security--relevant actions (e.g., reading from and writing to memory) and the unobservable event $\emptyevent$.
+Events include security-relevant actions (e.g., reading from and writing to memory) and the unobservable event $\emptyevent$.
 As programs execute, their emitted events are concatenated in traces $\trace$, which serve as the description of the behaviour of a program.%
@@
 The set of all properties can be partitioned into different {\em classes} ($\class$), i.e., safety, liveness, and neither safety nor liveness~\cite{clarkson2008hyper}.
-A class is simply a set of properties and for the class of safety properties, it is decidable wether a trace satisfies a safety property with just a finite trace prefix.
+A class is simply a set of properties and for the class of safety properties, it is decidable whether a trace satisfies a safety property with just a finite trace prefix.
 As an example, consider a trace describing an interaction with a memory where the deallocation of an address $\loc$ precedes a read at that address in memory: $\ev{Dealloc\ \loc}\cdot\ev{Read\ \loc\ 1729}\cdot\dots$
@@
 \begin{example}[Double Free in Bluetooth Subsystem]
-  Consider \texttt{CVE-2021-3564}~\cite{doublefree-bluetooth}, one of many submissions for a double--free vulnerability.
-  The bug arises due to a race condition where the context--level function \texttt{hci\_cmd\_work} was not expected to behave maliciously, since it resides in the same source--code repository where the vulnerability occurs.
-  Nevertheless, the component--level code of \texttt{hci\_dev\_do\_open} is linked with \texttt{hci\_cmd\_work} and does not atomically check wether the pointer has been freed already:
-  Therefore, \texttt{hci\_dev\_do\_open} does not satisfy the no--double--frees property robustly, since there is an implementation for \texttt{hci\_cmd\_work} that results the whole program to violate that property.
+  Consider \texttt{CVE-2021-3564}~\cite{doublefree-bluetooth}, one of many submissions for a double-free vulnerability.
+  The bug arises due to a race condition where the context-level function \texttt{hci\_cmd\_work} was not expected to behave maliciously, since it resides in the same source-code repository where the vulnerability occurs.
+  Nevertheless, the component-level code of \texttt{hci\_dev\_do\_open} is linked with \texttt{hci\_cmd\_work} and does not atomically check whether the pointer has been freed already:
+  Therefore, \texttt{hci\_dev\_do\_open} does not satisfy the no-double-frees property robustly, since there is an implementation for \texttt{hci\_cmd\_work} that results the whole program to violate that property.
 \end{example}
@@
 %      MK: I don't think we need to mention this at all
-This translation is considered {\em correct} if it is semantics--preserving.
+This translation is considered {\em correct} if it is semantics-preserving.
 That is, for a whole program $\src{\wholeprogvar}$, the compiler should relate the $\src{L}$ semantics of $\src{\wholeprogvar}$ with the semantics of $\trg{T}$ in such a way that they are ,,compatible''.
@@
   Suppose the component transferred control to the context and passed ownership of a $\trg{char}$ pointer $\trg{y}$ sized $\trg{1}$ cells, the context now calls the component again but casts this buffer to an $\trg{int*}$ prior to that call.
-  The pointer is valid for one $\trg{char}$--sized memory cell, as expected, but the actual read operation now returns $\trg{sizeof(int)}$ many bytes instead of just $\trg{sizeof(char)}$ many.
-  Thus, even if $\src{foo}$ may have been robust with respect to the no--reads--out--of--bounds property, its compiled counterpart is not and therefore the compiler fails to attain \Thmref{def:rtp}.
+  The pointer is valid for one $\trg{char}$-sized memory cell, as expected, but the actual read operation now returns $\trg{sizeof(int)}$ many bytes instead of just $\trg{sizeof(char)}$ many.
+  Thus, even if $\src{foo}$ may have been robust with respect to the no-reads-out-of-bounds property, its compiled counterpart is not and therefore the compiler fails to attain \Thmref{def:rtp}.
 \end{example}
@@
   \mi{(Event)}~\event\bnfdef\ \emptyevent \mid \lightning \mid \event[_{b}];\sandboxtag;\sigma \\
-  \mi{(Pre\text{--}event)}&~\event[_{b}]\bnfdef\ \ev{Alloc\ \loc\ n} \mid \ev{Dealloc\ \loc} \mid \ev{Use\ \loc\ n} \mid \ev{Branch\ n} \mid \ev{Binop\ n} \\
+  \mi{(Pre\text{-}event)}&~\event[_{b}]\bnfdef\ \ev{Alloc\ \loc\ n} \mid \ev{Dealloc\ \loc} \mid \ev{Use\ \loc\ n} \mid \ev{Branch\ n} \mid \ev{Binop\ n} \\
   \end{aligned}
@@
 
-The specification trace model defines events as either the empty event ($\emptyevent$), a crash ($\lightning$), or as tuples consisting of a pre--event, a control--tag, and a security--tag.
+The specification trace model defines events as either the empty event ($\emptyevent$), a crash ($\lightning$), or as tuples consisting of a pre-event, a control-tag, and a security-tag.
 The purpose of the model is to define properties, such as a stricter variant of cryptographic constant time.
-To this end, security--tags indicate wether an event contains sensitive information ($\lock$) or not ($\unlock$), while control--tags state wether the context ($\ctx$) or the component ($\comp$) are responsible for emitting the event.
+To this end, security-tags indicate whether an event contains sensitive information ($\lock$) or not ($\unlock$), while control-tags state whether the context ($\ctx$) or the component ($\comp$) are responsible for emitting the event.
 The latter is necessary to be able to ignore actions done by a spurious context that, e.g., immediately deallocates a memory location twice, thus violating \gls{tms}.
 %Even though this could be used for blame~\cite{patrignani2023blame}, this is not what this paper is concerned about.
-Lastly, pre--events describe the actual kind of event that happened.
+Lastly, pre-events describe the actual kind of event that happened.
 One such kind is the allocation event ($\ev{Alloc\ \loc\ n}$) that fires whenever a program claims $n$ cells of memory and stores them at address $\loc$.
@@
 Finally, for cryptographic code, there is a general guideline that secrets must not be visible on a trace.
-Moreover, an instruction whose timing is data--dependent must not have a secret as an operand.
-Typical operations with data--dependent timing are branches and certain binary operations, such as division\footnote{This is highly architecture dependent, but division is an operation that serves as a classic example for a data--dependent timing instruction.}.
+Moreover, an instruction whose timing is data-dependent must not have a secret as an operand.
+Typical operations with data-dependent timing are branches and certain binary operations, such as division\footnote{This is highly architecture dependent, but division is an operation that serves as a classic example for a data-dependent timing instruction.}.
 Both operations are also modelled in the specification trace model ($\ev{Branch\ n}$ and $\ev{Binop\ n}$).
@@
 Full \gls{ms} is then described as the conjunction of \Cref{def:trace:tmsdef,def:trace:smsdef}.
-Note, however, that this definition does not entail memory--safety issues introduced by side--channels, such as speculation.
+Note, however, that this definition does not entail memory-safety issues introduced by side-channels, such as speculation.
 
@@
 To lessen the burden when proving, each monitor makes use of a custom trace model that contains only the relevant information related to the property the monitor checks.
-To project specification traces $\trace$ onto these monitor--level traces $\abstrace$, each property has an associated event agreement relation $\traceagree[]{\pi}{\trace}{\abstrace}$.
-This relation performs the actual work of projecting the events, while the stutter--equivalence $\traceagree{\pi}{\trace}{\abstrace}$ remains the same for all properties $\pi$, yet is indexed by $\pi$ to emphasize the kind of event agreement.
+To project specification traces $\trace$ onto these monitor-level traces $\abstrace$, each property has an associated event agreement relation $\traceagree[]{\pi}{\trace}{\abstrace}$.
+This relation performs the actual work of projecting the events, while the stutter-equivalence $\traceagree{\pi}{\trace}{\abstrace}$ remains the same for all properties $\pi$, yet is indexed by $\pi$ to emphasize the kind of event agreement.
 With agreements, this section starts to define monitor satisfaction for traces and states that monitor satisfaction implies property satisfaction.
 To this end, monitor satisfaction is defined as follows.
-\bul{A specification trace $\trace$ monitor--satisfies property $\pi$} iff there exists a (final) monitor state $\monitor$ and an abstract trace $\abstrace$ such that the \iul{specification trace $\trace$ agrees with abstract trace $\abstrace$} and \oul{the initial monitor\footnote{In this paper, for all monitors, the initial monitor state is denoted as $\emptyset$.} can step to the (final) monitor state $\monitor$ with abstract trace $\abstrace$}.
+\bul{A specification trace $\trace$ monitor-satisfies property $\pi$} iff there exists a (final) monitor state $\monitor$ and an abstract trace $\abstrace$ such that the \iul{specification trace $\trace$ agrees with abstract trace $\abstrace$} and \oul{the initial monitor\footnote{In this paper, for all monitors, the initial monitor state is denoted as $\emptyset$.} can step to the (final) monitor state $\monitor$ with abstract trace $\abstrace$}.
 
@@
 \end{center}
-The state of the monitor for \gls{sms} is a set containing tuples of locations, control--tags, and the allocation size.
+The state of the monitor for \gls{sms} is a set containing tuples of locations, control-tags, and the allocation size.
 In comparison to the trace model of the \gls{tms} monitor, the trace model here is extended by sizing and positional information.
@@
 Both examples can be seen in the wild~\cite{nagarakatte2009soft,nagarakatte2010cets,akritidis2009baggy,wegman1991ccp,manjikian1997fusion}, it is infeasible to list them all.
-Consider the compilation passes of LLVM\cite{lattner2004llvm} whose source-code optimisations are industry--competitive.
+Consider the compilation passes of LLVM\cite{lattner2004llvm} whose source-code optimisations are industry-competitive.
 In LLVM, the pass manager can be used to schedule passes in a way where the optimisations yield the best results.
@@
 Unfortunately, the order in which \gls{cf} and \gls{dce} are performed influences the final result of the compilation (see \Cref{fig:cfdceex}).
-This {\em phase ordering problem} is well--known in literature and a practical solution is to simply perform a fixpoint iteration.
-Compiler engineers typically try to find an order of optimisations that yields well--optimised programs for either code size~\cite{cooper1999geneticphases} or performance~\cite{kulkarni2006exhaustivephase}.
+This {\em phase ordering problem} is well-known in literature and a practical solution is to simply perform a fixpoint iteration.
+Compiler engineers typically try to find an order of optimisations that yields well-optimised programs for either code size~\cite{cooper1999geneticphases} or performance~\cite{kulkarni2006exhaustivephase}.
 The following corollary justifies that any such order of compilation passes is valid with respect to security.
@@
 
-Besides the sequential composition, there are two other compositions that are more at the language- than the compiler--level.
+Besides the sequential composition, there are two other compositions that are more at the language- than the compiler-level.
 Nevertheless, consider $\trg{JVM BC}$, which is a popular target for programming language designers due to its high performance and relevance in industry.
@@
 Either the user of LLVM provides an explicit flag that instructs the overall compiler to emit $\trg{x86\_64}$ or the framework itself detects the target platform via heuristics.
-%Either way, the theoretical results suggest that it is possible to always find a ,,most--general'', secure compiler, given two secure compilers, that robustly preserves the least--upper bound of the classes involved in their compilation process.
+%Either way, the theoretical results suggest that it is possible to always find a ,,most-general'', secure compiler, given two secure compilers, that robustly preserves the least-upper bound of the classes involved in their compilation process.
 In general, there are two options to deal with these nuances in a concrete, practical setting:
@@
 However, $\Ltms$ programs are not \gls{sms}.
-That is, there are well--typed $\Ltms$ programs that perform an out--of--bounds access.
+That is, there are well-typed $\Ltms$ programs that perform an out-of-bounds access.
 In contrast, $\Ltrg$ is untyped and does not provide any guarantees with regards to \gls{ms}.
 $\Lms$ is exactly the same language as $\Ltrg$, but this paper still distinguishes the two for sake of readability.
-All three languages --- so $\Ltms$, $\Ltrg$, and $\Lms$ --- assume \gls{cct} to hold.
+All three languages -- so $\Ltms$, $\Ltrg$, and $\Lms$ -- assume \gls{cct} to hold.
 Even though it is possible to write programs that, e.g., return early in loops, which is trivially not \gls{cct}, they can be transformed to adhere to \gls{cct}~\cite{cauligi2019fact}.
 The language $\Lscct$ allows violating \gls{cct} by emitting events on, e.g., branching and division, that contain secrets.
-As a countermeasure, the language provides a way to read and write to a {\em model--specific register} that enables a ``\gls{cct}-mode''.
+As a countermeasure, the language provides a way to read and write to a {\em model-specific register} that enables a ``\gls{cct}-mode''.
 
@@
 The data associated to pointers resides in the dynamic state $\Omega$ that contains sandboxed heaps, one for the context and the other for the component.
-Because of the sandboxing, locations pointing into heaps must be associated with a control--tag to know the heap they point into.
-Other metadata in $\Omega$ entails a list of callable objects with information on wether they belong to the context or the component.
+Because of the sandboxing, locations pointing into heaps must be associated with a control-tag to know the heap they point into.
+Other metadata in $\Omega$ entails a list of callable objects with information on whether they belong to the context or the component.
 That way, when calling a callable object, the semantics can take care of emitting appropriate events that signal a context switch.
-For the languages considered in this paper, no callable objects are first--class.
+For the languages considered in this paper, no callable objects are first-class.
 So, the only values are pairs, pointers, and natural numbers.
-Another non--standard part of the languages is the concept of the poison--tag.
-The poison--tag marks a location as freed ($\poisoned$) without removing the information associated to the freed pointer from $\Omega$.
+Another non-standard part of the languages is the concept of the poison-tag.
+The poison-tag marks a location as freed ($\poisoned$) without removing the information associated to the freed pointer from $\Omega$.
 This measure prevents the semantics from getting stuck when an already freed pointer is freed again.
@@
   \begin{aligned}
-  \mi{(Pre\text{--}Events)}~\event[_{b}]\bnfdef \ev{Alloc}\ \loc\ \valueexpr \mid \ev{Dealloc}\ \loc \mid \ev{Get}\ \loc\ \valueexpr &\mid \ev{Set}\ \loc\ \valueexpr\ \valueexpr['] \mid \ev{Call}\ \comm\ g\ \valueexpr \mid \ev{Ret}\ \comm\ \valueexpr \mid \ev{Start} \mid \ev{End}\ \valueexpr \\
+  \mi{(Pre\text{-}Events)}~\event[_{b}]\bnfdef \ev{Alloc}\ \loc\ \valueexpr \mid \ev{Dealloc}\ \loc \mid \ev{Get}\ \loc\ \valueexpr &\mid \ev{Set}\ \loc\ \valueexpr\ \valueexpr['] \mid \ev{Call}\ \comm\ g\ \valueexpr \mid \ev{Ret}\ \comm\ \valueexpr \mid \ev{Start} \mid \ev{End}\ \valueexpr \\
   \mi{(Events)}~\event&\bnfdef \emptyevent \mid \lightning \mid \tup{\event[_{b}]}{\sandboxtag} \\
@@
 The key difference is that, as standard in secure compilation work, the traces have a call and return event that signals context switches.
-Hereby, a $\ev{Call\ \ctxtocomp\ foo\ \valueexpr}$ and $\ev{Return\ \ctxtocomp\ \valueexpr}$ signal that program execution transitions from context-- to component--level.
-Contrary, $\ev{Call\ \comptoctx\ foo\ \valueexpr}$ and $\ev{Return\ \ctxtocomp\ \valueexpr}$ signal that program execution transitions from component-- to context--level.
-The environmental semantics emits the tag $\nocomm$ for calls from context-- to context--level or component-- to component--level.
+Hereby, a $\ev{Call\ \ctxtocomp\ foo\ \valueexpr}$ and $\ev{Return\ \ctxtocomp\ \valueexpr}$ signal that program execution transitions from context- to component-level.
+Contrary, $\ev{Call\ \comptoctx\ foo\ \valueexpr}$ and $\ev{Return\ \ctxtocomp\ \valueexpr}$ signal that program execution transitions from component- to context-level.
+The environmental semantics emits the tag $\nocomm$ for calls from context- to context-level or component- to component-level.
 In the following, $\neg\ctx=\comp$ and $\neg\comp=\ctx$ and also $\operatorname{comm}(\ctx)=\ \ctxtocomp$ and $\operatorname{comm}(\comp)=\ \comptoctx$.
@@
 \begin{center}
-  \judgbox{\estepto[]{\runtimetermvar}{\runtimetermvar}{\event}}{,,Contextual step from runtime--term $\runtimetermvar$ to $\runtimetermvar[']$ emitting event $\event$.''}
+  \judgbox{\estepto[]{\runtimetermvar}{\runtimetermvar}{\event}}{,,Contextual step from runtime-term $\runtimetermvar$ to $\runtimetermvar[']$ emitting event $\event$.''}
   %
@@
 The environmental semantics is also mostly straightforward.
-In \Cref{tr:e-ret-notsame,tr:e-call-notsame} the judgement $\commlib\vdash\text{foo}:\Omega.\sandboxtag$ checks wether $\text{foo}$ is a component--level name by looking it up in the list of component-level names $\commlib$ and emits the appropriate transfer tag, i.e., either $\comptoctx$ or $\ctxtocomp$.
-Additional rules that are left out ensure that, e.g., when calling the $\text{main}$ function, the event $\ev{Start}$ is emitted, which is a design choice this paper does for convenience when reasoning about call--chains.
-
-The top--level execution $\progstepto{\prog{\library_{\ctx}}{\library_{\comp}}}{\runtimetermvar}{\trace}$ constructs an initial state $\Omega$ by linking $\library_{\ctx}$ and $\library_{\ctx}$ and then starts execution by calling the $\text{main}$ function.
+In \Cref{tr:e-ret-notsame,tr:e-call-notsame} the judgement $\commlib\vdash\text{foo}:\Omega.\sandboxtag$ checks whether $\text{foo}$ is a component-level name by looking it up in the list of component-level names $\commlib$ and emits the appropriate transfer tag, i.e., either $\comptoctx$ or $\ctxtocomp$.
+Additional rules that are left out ensure that, e.g., when calling the $\text{main}$ function, the event $\ev{Start}$ is emitted, which is a design choice this paper does for convenience when reasoning about call-chains.
+
+The top-level execution $\progstepto{\prog{\library_{\ctx}}{\library_{\comp}}}{\runtimetermvar}{\trace}$ constructs an initial state $\Omega$ by linking $\library_{\ctx}$ and $\library_{\ctx}$ and then starts execution by calling the $\text{main}$ function.
 The trace $\trace$ emitted during that execution serves as abstraction of the behavior of the program enabling the use of \Cref{def:propsat,def:proprsat}.
@@
 $\src{L_{\tmssafe}}$ does not extend the base syntax at term level.
-What makes $\src{L_{\tmssafe}}$ special is that it has a type system inspired by $L^{3}$~\cite{morrisett2005L3} that exhibits the property that every well--typed $\src{L_{\tmssafe}}$ program satisfies \gls{tms}.
+What makes $\src{L_{\tmssafe}}$ special is that it has a type system inspired by $L^{3}$~\cite{morrisett2005L3} that exhibits the property that every well-typed $\src{L_{\tmssafe}}$ program satisfies \gls{tms}.
 The proof of this property relies on a projection $\msfilterLtms{\src{\event}}=\event$ that projects $\src{L_{\tmssafe}}$ events to specfication events $\event$.
@@
 The projection $\msfilterLtms{\src{\event}}$ for the most part does the expected.
-But, it also filters any action that a contexts does, since these are irrelevant for component--level \gls{tms}.
+But, it also filters any action that a contexts does, since these are irrelevant for component-level \gls{tms}.
 
@@
 
-\subsection{$\Ltrg$: A Memory--Unsafe Language}\label{subsec:lsms}
+\subsection{$\Ltrg$: A Memory-Unsafe Language}\label{subsec:lsms}
 
@@
 $\Lscct$ extends the dynamic state of $\trg{L}$ (\Cref{subsec:lms}) with a register for a data (operand) independent timing mode.
-For the other languages seen earlier, this mode is intuitively always--on.
+For the other languages seen earlier, this mode is intuitively always-on.
 To modify the mode, $\Lscct$ extends expressions with the term $\obj{\lwrdoit{\expr}}$ that writes the value of $\obj{\expr}$ into the register and with a term $\obj{\lrddoit{x}{\expr}}$ that reads the value of the register into $\obj{x}$.
@@
 If the register is turned off, secrets may appear on traces as well as new events $\obj{Branch\ n}$ and $\obj{Binop\ n}$, which the semantics emits for conditionals and certain binary operations, e.g., division.
-To prevent leakage of secrets but still enable reasoning about memory safety, pre--events are extended with $\obj{\widehat{{Get}}\ \loc\ \valueexpr}$ and $\obj{\widehat{Set}\ \loc\ \valueexpr}$, which indicate reads from and writes to memory without leaking secret information involved in the access.
+To prevent leakage of secrets but still enable reasoning about memory safety, pre-events are extended with $\obj{\widehat{{Get}}\ \loc\ \valueexpr}$ and $\obj{\widehat{Set}\ \loc\ \valueexpr}$, which indicate reads from and writes to memory without leaking secret information involved in the access.
 
@@
 
-The compiler does not have to do anything special besides a typecheck to respect the calling--convention.
+The compiler does not have to do anything special besides a typecheck to respect the calling-convention.
 That is, because $\trg{L}$ has no static typechecks, it could happen that a bogus context $\trg{\library_{\ctx}}$ invokes a callable object accepting a $\src{\natt}$ with $\trg{\lpair{17}{29}}$.
@@
 The proof of \Cref{thm:cca:rtp:tms} uses standard techniques in secure compilation literature and is visualized in \Cref{fig:proofdiag:rtp}.
-Concretely, given the execution of the compiled $\Ltms$ program linked with an arbitrary $\trg{L}$ context yields a trace $\trg{\trace}$, it uses a trace--based backtranslation of $\trg{\trace}$ to reconstruct an $\src{L_{\tmssafe}}$ context.
+Concretely, given the execution of the compiled $\Ltms$ program linked with an arbitrary $\trg{L}$ context yields a trace $\trg{\trace}$, it uses a trace-based backtranslation of $\trg{\trace}$ to reconstruct an $\src{L_{\tmssafe}}$ context.
 \Cref{thm:wt:tms} already gives strong guarantees for programs written in $\src{L_{\tmssafe}}$ and can be exploited for the backtranslated context by ensuring that it typechecks.
-Well--typing means the $\src{L_{\tmssafe}}$ program runs, so the rest of the proof is about relating the $\trg{L}$--level execution with the $\src{L_{\tmssafe}}$--level execution.
-
-For the $\trg{L}$--execution, the compiler wrapper code at state $\trg{\Omega_{1}}$ may either crash the program or continue execution $\trg{\Omega_{w_{1}}}$ right before control is transferred to the component.
+Well-typing means the $\src{L_{\tmssafe}}$ program runs, so the rest of the proof is about relating the $\trg{L}$-level execution with the $\src{L_{\tmssafe}}$-level execution.
+
+For the $\trg{L}$-execution, the compiler wrapper code at state $\trg{\Omega_{1}}$ may either crash the program or continue execution $\trg{\Omega_{w_{1}}}$ right before control is transferred to the component.
 In this specific case, the compiler need not to insert additional code after the component is done with its execution, but a symmetric situation can happen here.
@@
 Up to getting and setting memory, the compiler is the identity function, i.e., all omitted cases are a recoloring of $\trg{L}$ to $\Lms$.
-For memory accesses, however, a bounds--check is inserted that enforces \gls{sms}.
+For memory accesses, however, a bounds-check is inserted that enforces \gls{sms}.
 To this end, the compiler introduces another, fresh identifier $\irl{x_{SIZE}}$ for each allocation that binds $\irl{x}$ to keep track of the allocation size.
@@
 
-Given the fact that $\Lscct$ provides a \gls{cct}--mode, the compiler has to ensure that execution in the component always happen in this mode.
+Given the fact that $\Lscct$ provides a \gls{cct}-mode, the compiler has to ensure that execution in the component always happen in this mode.
 The context could overwrite the flag and exit the mode, but upon invoking a callable object that is part of the component, the flag would be set again.
@@
 
-The robust preservation of properties as a compiler--level criterion has been analyzed extensively~\cite{abate2019jour,patrignani2021rsc,abate2021extacc,patrignani2022universal,patrignani2019survey,kruse2022csc}.
+The robust preservation of properties as a compiler-level criterion has been analyzed extensively~\cite{abate2019jour,patrignani2021rsc,abate2021extacc,patrignani2022universal,patrignani2019survey,kruse2022csc}.
 This work has made use of that previous work.
@@
 While this paper focuses on the robust preservation framework~\cite{abate2019jour}, other secure compilation criteria exist.
-The survey on formal approaches to secure compilation~\cite{patrignani2019survey} discusses a broad spectrum already, while this section presents a very high--level overview.
+The survey on formal approaches to secure compilation~\cite{patrignani2019survey} discusses a broad spectrum already, while this section presents a very high-level overview.
 The full abstraction approach~\cite{abadi1999fullabstraction} states that a compiler should preserve and reflect observational equivalence between source and target programs.
@@
 Different mechanisms for memory safety exist that also consider the secure compilation domain, i.e., have an active attacker model.
-For example, the ,,pointers as capabilities'' principle represents pointers as machine--level capabilities~\cite{korashy2021capableptrs}, which behave in a similar fashion to capabilities by means of linear typing~\cite{morrisett2005L3}.
-The approach of this paper also uses linear typing, but differs from $L^{3}$~\cite{morrisett2005L3} in the way that functions are not first--class.
+For example, the ,,pointers as capabilities'' principle represents pointers as machine-level capabilities~\cite{korashy2021capableptrs}, which behave in a similar fashion to capabilities by means of linear typing~\cite{morrisett2005L3}.
+The approach of this paper also uses linear typing, but differs from $L^{3}$~\cite{morrisett2005L3} in the way that functions are not first-class.
 Moreover, this paper considers an active attacker, while the work on $L^{3}$ only discusses whole programs and, thus, has no active attacker model.
 The instrumentation to ensure memory safety that this paper presents is inspired by seminal work in the literature~\cite{nagarakatte2009soft}.
-That work inserts bounds--checks in front of pointer--dereferences and, for this to work, inserts meta--data information on pointer creation.
-It also works in a more advanced setting with structured fields accesses and also introduces a table--lookup for pointers that are stored in memory.
+That work inserts bounds-checks in front of pointer-dereferences and, for this to work, inserts meta-data information on pointer creation.
+It also works in a more advanced setting with structured fields accesses and also introduces a table-lookup for pointers that are stored in memory.
 This paper only considers pointers to arrays of simple objects, i.e., there are no pointers to pointers or structures.
-Of course, there are a lot of other approaches to memory--safety in literature, specifically as compiler instrumentations~\cite{akritidis2009baggy,younan2010paricheck,jung2021pico,shankaranarayana2023tailcheck,dhumbumroong2020boundwarden,nam2019framer,zhou2023fatptrs}, hardware--extensions~\cite{kwon2013lowfat,saileshwar2022heapcheck,chen2023flexpointer,kim2023whistle}, or programming language extensions~\cite{elliott2018checkedc,li2022formalcheckedc,jim2002cyclone,elliott2015guilt,west2005cuckoo,weis2019fyr,benoit2019uniqueness}, to name a few.
-What differentiates this work from them is that this work uses known, compiler--based approaches to ensure memory--safety as a means to investigate secure compiler compositions.
-This paper does not provide efficient memory--safety, but serves as a theoretical foundation for the secure compilation domain.
+Of course, there are a lot of other approaches to memory-safety in literature, specifically as compiler instrumentations~\cite{akritidis2009baggy,younan2010paricheck,jung2021pico,shankaranarayana2023tailcheck,dhumbumroong2020boundwarden,nam2019framer,zhou2023fatptrs}, hardware-extensions~\cite{kwon2013lowfat,saileshwar2022heapcheck,chen2023flexpointer,kim2023whistle}, or programming language extensions~\cite{elliott2018checkedc,li2022formalcheckedc,jim2002cyclone,elliott2015guilt,west2005cuckoo,weis2019fyr,benoit2019uniqueness}, to name a few.
+What differentiates this work from them is that this work uses known, compiler-based approaches to ensure memory-safety as a means to investigate secure compiler compositions.
+This paper does not provide efficient memory-safety, but serves as a theoretical foundation for the secure compilation domain.
 
@@
 
-The approach to cryptographic constant time in this paper is high--level, where a programming language exposes a way to switch the semantics to a data (operand) independent timing mode.
+The approach to cryptographic constant time in this paper is high-level, where a programming language exposes a way to switch the semantics to a data (operand) independent timing mode.
 Since identifiers in $\Lscct$ are annotated with a secrecy tag, this approach is similar to others with information flow control.
-For example, Vale~\cite{bond2017vale} uses Dafny to ensure constant--time assembly code, while Jasmin~\cite{almeida2017jasmin} makes use of the Coq proof assistant to reject non--constant--time programs.
-CT-Wasm~\cite{watt2019ctwasm} enforces constant--timeness by means of a type system.
+For example, Vale~\cite{bond2017vale} uses Dafny to ensure constant-time assembly code, while Jasmin~\cite{almeida2017jasmin} makes use of the Coq proof assistant to reject non-constant-time programs.
+CT-Wasm~\cite{watt2019ctwasm} enforces constant-timeness by means of a type system.
 Different to the approach of this paper, these approaches necessitate that the programmer writes \gls{cct} code.
-An approach to allow programmers to write more high--level code is CryptOpt~\cite{kuepper2023cryptopt}, which generates efficient target-code by means of a randomized search.
-This paper abstracts over concrete mitigation strategies and simply assumes that there is a flag to switch to a cryptographic--constant time execution mode.
-This can be realized by employing the FaCT~\cite{cauligi2019fact} compiler, which translates common non-constant time code patterns to be constant--time, and the data (object) independent timing execution mode of modern processors.
+An approach to allow programmers to write more high-level code is CryptOpt~\cite{kuepper2023cryptopt}, which generates efficient target-code by means of a randomized search.
+This paper abstracts over concrete mitigation strategies and simply assumes that there is a flag to switch to a cryptographic-constant time execution mode.
+This can be realized by employing the FaCT~\cite{cauligi2019fact} compiler, which translates common non-constant time code patterns to be constant-time, and the data (object) independent timing execution mode of modern processors.
 
@@
 The paper exercises this framework on a case-study that consists of an optimising compilation chain that is secure with respect to \gls{msscct}.
-In future work, it would be interesting to investigate wether it is possible to provide {\em secure compiler combinators}, similarily to how parser combinators work.
-%To this end, it may be necessary to extend existing frameworks for multi--language semantics.
-It is also interesting to extend our case--study and verify all results of it in Coq.
+In future work, it would be interesting to investigate whether it is possible to provide {\em secure compiler combinators}, similarily to how parser combinators work.
+%To this end, it may be necessary to extend existing frameworks for multi-language semantics.
+It is also interesting to extend our case-study and verify all results of it in Coq.
 This way, it is possible to obtain an executable, secure compiler.

### 2023-07-09 popl/main.tex

@@
 The specification trace model defines events as either the empty event ($\emptyevent$), a crash ($\lightning$), or as tuples consisting of a pre-event, a control-tag, and a security-tag.
-The purpose of the model is to define properties, such as a stricter variant of cryptographic constant time.
+The purpose of the model is to define key security properties of interest, such as \gls{ms} or a stricter variant of cryptographic constant time.
 To this end, security-tags indicate whether an event contains sensitive information ($\lock$) or not ($\unlock$), while control-tags state whether the context ($\ctx$) or the component ($\comp$) are responsible for emitting the event.
-The latter is necessary to be able to ignore actions done by a spurious context that, e.g., immediately deallocates a memory location twice, thus violating \gls{tms}.
+The latter is necessary to be able to ignore actions done by a spurious context that, e.g., immediately deallocates a memory location twice, thus violating \gls{tms}~\cite{nagarakatte2010cets}.
 %Even though this could be used for blame~\cite{patrignani2023blame}, this is not what this paper is concerned about.
@@
 One such kind is the allocation event ($\ev{Alloc\ \loc\ n}$) that fires whenever a program claims $n$ cells of memory and stores them at address $\loc$.
-Likewise, deallocation ($\ev{Dealloc\ \loc}$) announces that the object at location $\loc$ is freed.
-These two events alone allow already a partial description of \gls{tms} by requiring that, e.g., there is only one deallocation event that carries a location $\loc$.
+Dually, deallocation ($\ev{Dealloc\ \loc}$) announces that the object at location $\loc$ is freed.
+These two events alone allow\MP{allow who?
+allow needs an object, this is improper english} already a partial description of \gls{tms} by requiring that, e.g., there is only one deallocation event that carries a location $\loc$.
 To be able to express \gls{sms}, there is also an event to describe reads from and writes to memory ($\ev{Use\ \loc\ n}$).
@@
 Moreover, an instruction whose timing is data-dependent must not have a secret as an operand.
-Typical operations with data-dependent timing are branches and certain binary operations, such as division\footnote{This is highly architecture dependent, but division is an operation that serves as a classic example for a data-dependent timing instruction.}.
+Typical operations with data-dependent timing are branches and certain binary operations, such as division\footnote{This is highly architecture-dependent, but division is an operation that serves as a classic example for a data-dependent timing instruction.}.
 Both operations are also modelled in the specification trace model ($\ev{Branch\ n}$ and $\ev{Binop\ n}$).
@@
 
-Notably, real-world compilation chains also perform a series of passes whose main purpose is not necessarily to translate from one language to another, but to, e.g., optimise the code or enforce a certain property.
-Both examples can be seen in the wild~\cite{nagarakatte2009soft,nagarakatte2010cets,akritidis2009baggy,wegman1991ccp,manjikian1997fusion}, it is infeasible to list them all.
-Consider the compilation passes of LLVM\cite{lattner2004llvm} whose source-code optimisations are industry-competitive.
-In LLVM, the pass manager can be used to schedule passes in a way where the optimisations yield the best results.
-Take \gls{cf} and \gls{dce}.
+\subsubsection{Securing Optimisations}
+\MPin{note that i'm always using british spelling. optimisation VS optimization. etc. }
+
+Notably, real-world compilation chains also perform a series of (sequential) passes whose main purpose is not necessarily to translate from one language to another, but to, e.g., optimise the code or enforce a certain property.
+\MPin{
+	do not say the negation of something before the non-negation.
+	say what things are first, and what they are not later (if at all)
+}
+Both examples can be seen in the wild~\cite{nagarakatte2009soft,nagarakatte2010cets,akritidis2009baggy,wegman1991ccp,manjikian1997fusion}, it is infeasible to list them all\MP{survey to cite?
+this is a bit informal}.
+Consider the compilation passes of LLVM~\cite{lattner2004llvm} whose source-code optimisations are industry--competitive\MP{meaning?}.
+In LLVM, the pass manager\MP{meaning} can be used to schedule passes in a way where the optimisations yield the best results.
+Take \gls{cf} and \gls{dce}. \MPin{
+	brief explanation?
+}
 Unfortunately, the order in which \gls{cf} and \gls{dce} are performed influences the final result of the compilation (see \Cref{fig:cfdceex}).
-This {\em phase ordering problem} is well-known in literature and a practical solution is to simply perform a fixpoint iteration.
+\MPin{figure here, or closer, not far away}
+This {\em phase ordering problem} is well--known in literature and a practical solution is to simply perform a fixpoint iteration\MP{of what}.\MP{cit?}
 Compiler engineers typically try to find an order of optimisations that yields well-optimised programs for either code size~\cite{cooper1999geneticphases} or performance~\cite{kulkarni2006exhaustivephase}.
-The following corollary justifies that any such order of compilation passes is valid with respect to security.
+\Cref{corr:swappable} justifies that any such order of compilation passes is valid with respect to security.
 So, given two compilation passes \bul{$\cc[_{1}]{\trg{L}}{\trg{L}}$, $\cc[_{2}]{\trg{L}}{\trg{L}}$, both robustly preserving class $\class[_{1}]$ or $\class[_{2}]$, respectively}, \oul{for any order of their composition the composed compiler robustly preserves the intersection of $\class[_{1}]$ and $\class[_{2}]$}.
@@
 
-Besides the sequential composition, there are two other compositions that are more at the language- than the compiler-level.
+\subsection{Secure Upper and Lower Composition}\label{sec:other-compos}
+Besides sequential composition, there are two other compositions that are more at the language- than the compiler-level\MP{namely ...}.
 Nevertheless, consider $\trg{JVM BC}$, which is a popular target for programming language designers due to its high performance and relevance in industry.
 As an example, $\src{Java}$ and $\obj{Kotlin}$ compile to $\trg{JVM BC}$.
-Technically speaking, they both lower to class files and $\obj{Kotlin}$ objects are considered to be the same as $\src{Java}$ objects at that point.
+Technically speaking, they both lower\MP{undefined verb. no need for synonyms here.} to class files and $\obj{Kotlin}$ objects are considered to be the same as $\src{Java}$ objects at that point.
 A compiler that accepts both $\src{Java}$ and $\obj{Kotlin}$ code translating to the same target language, or rather, intermediate representation, performs a kind of {\em upper} composition.
@@
 %\section{Case Study: Memory Safe and Cryptographic Constant Time Code}\label{sec:casestud}
-\section{Case Study: Preliminary Definitions\pages{2}}\label{sec:casestud:defs}
-
-This section defines programming languages that the secure compilers defined in the next section will make use of.
-To this end, this section defines the languages $\Ltms$, $\Ltrg$, $\Lms$, and $\Lscct$.
-In this paper, $\Ltms$ is the only statically typed language and exhibits the property that all well-typed programs are \gls{tms}.
-However, $\Ltms$ programs are not \gls{sms}.
-That is, there are well-typed $\Ltms$ programs that perform an out-of-bounds access.
-In contrast, $\Ltrg$ is untyped and does not provide any guarantees with regards to \gls{ms}.
-$\Lms$ is exactly the same language as $\Ltrg$, but this paper still distinguishes the two for sake of readability.
-All three languages -- so $\Ltms$, $\Ltrg$, and $\Lms$ -- assume \gls{cct} to hold.
+\section{Case Study: Language Formalisations\pages{2}}\label{sec:casestud:defs}
+
+This section defines programming languages that the secure compilers defined in the next section will make use of\MP{make use of -> use! 3 words when 1 is sufficient}.
+To this end, this section defines the languages $\Ltms$, $\Ltrg$, $\Lms$, and $\Lscct$ which share many common elements (presented in \Cref{subsec:cs:defs}).
+In this paper, $\Ltms$ is the only statically typed language and exhibits the property that all well-typed programs are \gls{tms} (\Cref{??}).
+However, not all $\Ltms$ programs are \gls{sms}.\MP{reworded. was imprecise}
+That is, there are well--typed $\Ltms$ programs that perform an out--of--bounds access.
+In contrast, $\Ltrg$ is untyped and does not provide any guarantees with regards to \gls{ms} (\Cref{??}).
+$\Lms$ is exactly the same language as $\Ltrg$, but this paper still distinguishes the two for sake of readability (\Cref{??}).
+All three languages --- so $\Ltms$, $\Ltrg$, and $\Lms$ --- assume \gls{cct} to hold.
+
 Even though it is possible to write programs that, e.g., return early in loops, which is trivially not \gls{cct}, they can be transformed to adhere to \gls{cct}~\cite{cauligi2019fact}.
-The language $\Lscct$ allows violating \gls{cct} by emitting events on, e.g., branching and division, that contain secrets.
-As a countermeasure, the language provides a way to read and write to a {\em model-specific register} that enables a ``\gls{cct}-mode''.
+\MPin{
+	no, it's different. 
+	there is no CCT because we asume it's out of the programmer's concers and since CT is arch-dependent, the compiler should preserve CT, not the programmer, via software
+}
+
+Language $\Lscct$ allows violating \gls{cct} by emitting events on, e.g., branching and division, that contain secrets (\Cref{??}).
+However, $\Lscct$ also contains a countermeasure to enforce \gls{cct}, the language provides a way to read and write to a {\em model--specific register} that enables a ``\gls{cct}-mode''.
+\MPin{ref to what real world stuff justifies this}
 
 \subsection{Shared Language Definitions}\label{subsec:cs:defs}
-
 \begin{gather*}
   \begin{aligned}
-  \mi{(Expressions)}~\expr&\bnfdef\ x \mid \valueexpr \mid \lbinop{\expr[_{1}]}{\expr[_{2}]} \mid \lget{x}{\expr} \mid \lnew{\expr[_{1}]}{\expr[_{2}]} \mid \ldelete{x} \mid \lset{x}{\expr[_{1}]}{\expr[_{2}]} \\
-    & \mid \llet{x}{\expr[_{1}]}{\expr[_{2}]}  \mid \lreturn{\expr} \mid \lcall{g}{\expr} \mid \lifz{\expr}{\expr[_{1}]}{\expr[_{2}]} \mid \labort \\
-  \mi{(Symbols)}~&\function\bnfdef\ \lfunction{g}{x}{e}\hspace{0.5cm}
-  \mi{(Libraries)}~\library\bnfdef\ \hole{\cdot} \mid \function,\library \\
-  \mi{(Programs)}~&\prog{\library_{\ctx}}{\library_{\comp}}\hspace{0.5cm}
-  \mi{(Substitutions)}~\substvar\bnfdef \hole{\cdot} \mid \subst{\valueexpr}{x},\gamma \\
+  \mi{(Expressions)}~&\expr\bnfdef\ x \mid \valueexpr \mid \lbinop{\expr[_{1}]}{\expr[_{2}]} \mid \lget{x}{\expr} \mid \lnew{\expr[_{1}]}{\expr[_{2}]} \mid \ldelete{x} \mid \lset{x}{\expr[_{1}]}{\expr[_{2}]} \\
+    & \mid \llet{x}{\expr[_{1}]}{\expr[_{2}]}  \mid \lreturn{\expr} \mid \lcall{g}{\expr} \mid \lifz{\expr}{\expr[_{1}]}{\expr[_{2}]} \mid \labort
+  \end{aligned}
+  \\
+  \begin{aligned}
+  \mi{(Symbols)}~&\function\bnfdef\ \lfunction{g}{x}{e}
+  &
+  \mi{(Libraries)}~&\library\bnfdef\ \hole{\cdot} \mid \function,\library \\
+  \mi{(Programs)}~&\prog{\library_{\ctx}}{\library_{\comp}}
+  &
+  \mi{(Substitutions)}~&\substvar\bnfdef \hole{\cdot} \mid \subst{\valueexpr}{x},\gamma
   \end{aligned}
 \end{gather*}
+\MPin{
+	this latex had many issues.
+	the \& were at different spots
+	it used hspace to space things instead of \&
+	and it did not use \& to align the following things on a line
+	and it did not use multiple aligned into the gather
+	(what's the point of gather if you don't have several aligned inside? use a single align* instead)
+}
 
 Above is a fraction of the shared syntax of all the programming languages of this paper.
-The syntax and semantics are mostly standard.
-However, when allocating pointers with $\lnew{\expr[_{1}]}{\expr[_{2}]}$, the expression $\expr[_{1}]$ determines the initial value of the objects and $\expr[_{2}]$ the overall size of an array.
+\MP{why a fraction? what exprs are missing}
+The syntax and semantics are mostly standard.\MP{and therefore omitted}
+However, when allocating pointers\MP{arrays?} with $\lnew{\expr[_{1}]}{\expr[_{2}]}$, the expression $\expr[_{1}]$ determines the initial value of the objects\MP{cells?} and $\expr[_{2}]$ the overall size of an array.
 The data associated to pointers resides in the dynamic state $\Omega$ that contains sandboxed heaps, one for the context and the other for the component.
-Because of the sandboxing, locations pointing into heaps must be associated with a control-tag to know the heap they point into.
-Other metadata in $\Omega$ entails a list of callable objects with information on whether they belong to the context or the component.
+Because of the sandboxing, locations pointing into heaps must be associated with a control--tag\MP{symbol} to know the heap they point into.
+Other metadata in $\Omega$ entails a list of callable objects\MP{functions? objects are other things in PL} with information on wether they belong to the context or the component.
 That way, when calling a callable object, the semantics can take care of emitting appropriate events that signal a context switch.
@@
   \begin{aligned}
-  \mi{(Pre\text{-}Events)}~\event[_{b}]\bnfdef \ev{Alloc}\ \loc\ \valueexpr \mid \ev{Dealloc}\ \loc \mid \ev{Get}\ \loc\ \valueexpr &\mid \ev{Set}\ \loc\ \valueexpr\ \valueexpr['] \mid \ev{Call}\ \comm\ g\ \valueexpr \mid \ev{Ret}\ \comm\ \valueexpr \mid \ev{Start} \mid \ev{End}\ \valueexpr \\
-  \mi{(Events)}~\event&\bnfdef \emptyevent \mid \lightning \mid \tup{\event[_{b}]}{\sandboxtag} \\
+  \mi{(Pre\text{-}Events)}~&\event[_{b}]\bnfdef \ev{Alloc}\ \loc\ \valueexpr \mid \ev{Dealloc}\ \loc \mid \ev{Get}\ \loc\ \valueexpr &\mid \ev{Set}\ \loc\ \valueexpr\ \valueexpr['] \mid \ev{Call}\ \comm\ g\ \valueexpr \mid \ev{Ret}\ \comm\ \valueexpr \mid \ev{Start} \mid \ev{End}\ \valueexpr \\
+  \mi{(Events)}~&\event\bnfdef \emptyevent \mid \lightning \mid \tup{\event[_{b}]}{\sandboxtag} \\
   \end{aligned}

### 2023-07-09 popl/main.tex
up to 4.1

@@
 To remedy this, {\em linking} takes two partial programs $\progvar[_{1}],\progvar[_{2}]$ and produces a whole program $\wholeprogvar$, i.e., $\link{\progvar[_{1}]}{\progvar[_{2}]}=\wholeprogvar$.
-As typically done in not only secure compilation~\cite{abate2019jour,devriese2018parametricity}, \MP{others do this too, parametricity results e.g.} this paper assumes that whole programs are the result of linking partial programs referred to as {\em context} ($\ctx$) and {\em component} ($\comp$).
+As typically done in not only secure compilation~\cite{abate2019jour,devriese2018parametricity}, %
+% \MP{others do this too, parametricity results e.g.}
+this paper assumes that whole programs are the result of linking partial programs referred to as {\em context} ($\ctx$) and {\em component} ($\comp$).
 The context is an arbitrary program and thus has the role of an {\em attacker} that can interact with the component by means of whatever features the programming language has.
@@
 These properties are of practical importance (as mentioned in \Cref{sec:introduction}) and also of interest in the case study (\Cref{sec:casestud:defs,sec:casestud:rtp}) this paper presents later. 
-\MPin{
-	missing 3.2 ref? added below?
-}
-Then, this section introduces monitors for each of the properties presented earlier (\Cref{??}).
+Then, this section introduces monitors for each of the properties presented earlier (\Cref{subsec:monitors}).
 
@@
 Dually, deallocation ($\ev{Dealloc\ \loc}$) announces that the object at location $\loc$ is freed.
-These two events alone allow\MP{allow who?
-allow needs an object, this is improper english} already a partial description of \gls{tms} by requiring that, e.g., there is only one deallocation event that carries a location $\loc$.
+These two events alone are enough to provide a partial description of \gls{tms} by requiring that, e.g., there is only one deallocation event that carries a location $\loc$.
 To be able to express \gls{sms}, there is also an event to describe reads from and writes to memory ($\ev{Use\ \loc\ n}$).
@@
 
-%\section{Monitors and Trace Agreements\pages{1.5}}\label{sec:monitors}
-\subsection{Monitors\pages{1.5}}\label{sec:monitors}
-
-The use of monitors is because they have some structure useful for inductive proofs, whereas traces lack this structure, thus making proofs difficult.\MP{not true.
-safety props are enforced by monitors}
+%\section{Monitors and Trace Agreements\pages{1.5}}\label{subsec:monitors}
+\subsection{Monitors\pages{1.5}}\label{subsec:monitors}
+
+Monitors can be used to enforce safety properties and this section presents concrete instantiations of monitors for the previousvy presented properties (\Cref{subsec:propdefs}).
 To lessen the burden when proving, each monitor makes use of a custom trace model that contains only the relevant information related to the property the monitor checks.
@@
 The main result is that secure compilers in the robust compilation framework~\cite{abate2019jour} compose {\em sequentially}.
-This is not intuitive in the sense that there are a number of examples in the security domain where sequential composition does not work~\cite{canetti2006univcomp,mccullough2012compo,fabian2022automatic}.\MP{?canetti?}
+This is not intuitive in the sense that there are a number of examples in the security domain where sequential composition does not work~\cite{canetti2006univcomp,mccullough2012compo,fabian2022automatic}.%
+\MP{?canetti?} %% protocols under certain security notions are not secure when composed
 The sequential composition of compilers $\cc{\src{L}}{\trg{L}}$ and $\cc{\trg{L}}{\obj{L}}$ is defined as follows: given an $\src{L}$ program $\src{p}$, its compiled $\obj{L}$ counterpart is obtained by plugging $\src{p}$ into $\cc{\src{L}}{\trg{L}}\circ\cc{\trg{L}}{\obj{L}}$.
-% MP: i know, i added a passive verb. here it works. i think. but i appreciate the effort!
 Now, given \bul{$\cc{\src{L}}{\trg{L}}$ robustly preserves $\class[_{1}]$} and \rul{$\cc{\trg{L}}{\obj{L}}$ robustly preserves $\class[_{2}]$}, it follows that \oul{their sequential composition $\cc{\src{L}}{\trg{L}}\circ\cc{\trg{L}}{\obj{L}}$ robustly preserves the intersection of classes $\class[_{1}]$ and $\class[_{2}]$}.

### 2023-07-09 popl/main.tex
small fixs

@@
 \title{Secure Composition of Robust and Optimising Compilers}
+% Composing Secure and Optimising Compilers
+%
 
@@
 \gls{ms} is a security property of programs that, e.g., ensures that there are no out-of-bounds accesses as well as no reads or writes to uninitialized memory.
-Without memory safety, private data can easily leak or corrupt~\cite{lemay2021ccc}.
+Without memory safety, private data can easily leak or be corrupted~\cite{lemay2021ccc}.
 Because of this, \gls{ms} is a requirement for programs to attain an even stronger security property, namely \gls{cct}.
@@
 Even though mitigations for such attacks exist~\cite{bond2017vale,almeida2017jasmin}, the literature shows that some compilers may optimize these mitigations away~\cite{barthe2018sec}.
-This is of course not what the programmer expects and necessitates the need for a {\em secure} compiler that ensures security properties hold even when the program runs in a hostile context, i.e., where it interacts with other possibly malicious programs.
-Secure compilers can be described formally with the theory on {\em robust preservation}~\cite{abate2019jour,abate2021extacc,patrignani2021rsc}, which takes this interaction between components and potentially malicious contexts into account.
-Unfortunately, it is unclear if and how robust preservation supports modularity.
-Modularity is an inherent part of software engineering and real-world compilers are no exception to that~\cite{lattner2004llvm,googlev8}.
-Compilation passes are typically developed as a standalone module that can plugged into a compilation pipeline.
-Different kinds of such passes exist, e.g., there are usually several compilation passes that perform optimisations~\cite{lattner2004llvm,googlev8,androidstudio,kuepper2023cryptopt,manjikian1997fusion,wegman1991ccp} or source-code instrumentations~\cite{nagarakatte2009soft,nagarakatte2010cets,akritidis2009baggy,dhumbumroong2020boundwarden,jung2021pico,nam2019framer,shankaranarayana2023tailcheck,younan2010paricheck,zhou2023fatptrs,bond2017vale,almeida2017jasmin,kuepper2023cryptopt,cauligi2019fact} that enforce that the compiled program fulfills a property of interest.
+This is of course not what programmers expect and necessitates the need for a {\em secure} compiler that ensures such mitigations persist in the code they generate.
+More generally, secure compilers should preserve security properties even when the program runs in a hostile context such as the cloud one, i.e., where it interacts with other possibly malicious programs.
+An example of such secure compilers are those that ascribe to the theory of robust compilation~\cite{abate2019jour,abate2021extacc,patrignani2021rsc}, which takes this interaction between components and potentially malicious contexts into account.
+Unfortunately, it is unclear if and how robust compilation supports compositionality.
+
+% Compositionality is an inherent part of software engineering and real-world compilers are no exception to that~\cite{lattner2004llvm,googlev8}.
+Compilation passes are typically developed as a standalone module to be plugged into a compilation pipeline and composed with other passes.
+Different kinds of such passes exist, e.g., there are usually several compilation passes that perform optimisations~\cite{lattner2004llvm,googlev8,androidstudio,kuepper2023cryptopt,manjikian1997fusion,wegman1991ccp} or source-code instrumentations~\cite{nagarakatte2009soft,nagarakatte2010cets,akritidis2009baggy,dhumbumroong2020boundwarden,jung2021pico,nam2019framer,shankaranarayana2023tailcheck,younan2010paricheck,zhou2023fatptrs,bond2017vale,almeida2017jasmin,kuepper2023cryptopt,cauligi2019fact} that enforce that the compiled program fulfils a property of interest.
 Without a modular framework, there is no hope to enable this separation of concerns, where, e.g., a secure pass that ensures \gls{ms} can be developed independently of a secure pass for \gls{cct}.
@@
 %Because of the potentially hostile context, secure compilers attain security by relying on protection mechanisms, such as ... and ...\MK{cits}.
-The aim of this paper is to devise the theoretical foundation to be able to modularily build a secure compiler that is capable of dealing with the following example.
+The aim of this paper is to devise the theoretical foundation to be able to build a secure compiler as the composition of several smaller ones.
+Let us now introduce a minimal example of what such a secure compiler should be capable of dealing with.
 %\MPin{ MK: hope it's solved now

### 2023-07-09 popl/main.tex
precisely 4.1

--- 2023-07-09 popl/main.tex
@@
 \subsubsection{Securing Optimisations}
-\MPin{note that i'm always using british spelling. optimisation VS optimization. etc. }
+%\MPin{note that i'm always using british spelling. optimisation VS optimization. etc. }
+% MK: I'll go over the doc with the antidote spellchecker as soon as it is reasonably stable.
 
 Notably, real-world compilation chains also perform a series of (sequential) passes whose main purpose is not necessarily to translate from one language to another, but to, e.g., optimise the code or enforce a certain property.
-\MPin{
-	do not say the negation of something before the non-negation.
-	say what things are first, and what they are not later (if at all)
-}
-Both examples can be seen in the wild~\cite{nagarakatte2009soft,nagarakatte2010cets,akritidis2009baggy,wegman1991ccp,manjikian1997fusion}, it is infeasible to list them all\MP{survey to cite?
+%\MPin{
+%	do not say the negation of something before the non-negation.
+%	say what things are first, and what they are not later (if at all)
+%}  MK: that for sure is less confusing... Good suggestion!
+Both examples can be seen in practice~\cite{nagarakatte2009soft,nagarakatte2010cets,akritidis2009baggy,wegman1991ccp,manjikian1997fusion}, it is infeasible to list them all\MP{survey to cite?
 this is a bit informal}.
+% MK: hard to find a survey on different compiler passes...
+I can't even find one just for memsafety
 Consider the compilation passes of LLVM~\cite{lattner2004llvm} whose source-code optimisations are industry--competitive\MP{meaning?}.
-In LLVM, the pass manager\MP{meaning} can be used to schedule passes in a way where the optimisations yield the best results.
-Take \gls{cf} and \gls{dce}. \MPin{
-	brief explanation?
-}
+In LLVM, compiler passes are handled in the \texttt{PassManager} class.
+By means of manipulating the internal state of the \texttt{PassManager} object, the schedule of compiler optimisation passes can be hardcoded.
+That way, an order of optimisations can be found by hand that yields the best results.
+Take \gls{cf}, which rewrites constant expressions to the constant they evaluate to, and \gls{dce}, which removes dead code by rewriting conditional branches.
 Unfortunately, the order in which \gls{cf} and \gls{dce} are performed influences the final result of the compilation (see \Cref{fig:cfdceex}).
-\MPin{figure here, or closer, not far away}
-This {\em phase ordering problem} is well--known in literature and a practical solution is to simply perform a fixpoint iteration\MP{of what}.\MP{cit?}
-Compiler engineers typically try to find an order of optimisations that yields well-optimised programs for either code size~\cite{cooper1999geneticphases} or performance~\cite{kulkarni2006exhaustivephase}.
-\Cref{corr:swappable} justifies that any such order of compilation passes is valid with respect to security.
-So, given two compilation passes \bul{$\cc[_{1}]{\trg{L}}{\trg{L}}$, $\cc[_{2}]{\trg{L}}{\trg{L}}$, both robustly preserving class $\class[_{1}]$ or $\class[_{2}]$, respectively}, \oul{for any order of their composition the composed compiler robustly preserves the intersection of $\class[_{1}]$ and $\class[_{2}]$}.
-
-\begin{corollary}[Sequential Composition of Secure Compilers]\label{corr:swappable}
-  If \bul{$\rtp{\cc[_{1}]{\trg{L}}{\trg{L}}}{\class[_{1}]}$ and $\rtp{\cc[_{2}]{\trg{L}}{\trg{L}}}{\class[_{2}]}$}, then \oul{$\rtp{\cc[_{1}]{\trg{L}}{\trg{L}}\circ\cc[_{2}]{\trg{L}}{\trg{L}}}{\class[_{1}]\cap\class[_{2}]}$ and $\rtp{\cc[_{2}]{\trg{L}}{\trg{L}}\circ\cc[_{1}]{\trg{L}}{\trg{L}}}{\class[_{2}]\cap\class[_{1}]}$}.
-\Coqed
-\end{corollary}
-
-\subsection{Secure Upper and Lower Composition}\label{sec:other-compos}
-Besides sequential composition, there are two other compositions that are more at the language- than the compiler-level\MP{namely ...}.
-Nevertheless, consider $\trg{JVM BC}$, which is a popular target for programming language designers due to its high performance and relevance in industry.
-As an example, $\src{Java}$ and $\obj{Kotlin}$ compile to $\trg{JVM BC}$.
-Technically speaking, they both lower\MP{undefined verb. no need for synonyms here.} to class files and $\obj{Kotlin}$ objects are considered to be the same as $\src{Java}$ objects at that point.
-A compiler that accepts both $\src{Java}$ and $\obj{Kotlin}$ code translating to the same target language, or rather, intermediate representation, performs a kind of {\em upper} composition.
-Given a secure compiler from $\src{Java}$ to $\trg{JVM BC}$ and one from $\obj{Kotlin}$ to $\trg{JVM BC}$, it is easy to envision that one can engineer a combined compiler that accepts both $\src{Java}$ and $\obj{Kotlin}$.
-A practical example of this, which is, however, insecure, is {\tt Android Studio}~\cite{androidstudio}.
-The proof of \Cref{thm:rtp} can be easily adapted to also works for this {\em upper} composition.
-The corresponding theorem is not listed, since there is no general definition of this kind of composition.
-\MPin{
-	mmm need to think about this.
-	this looks like discussion section material
-}
-\MPin{
-	we could turn this into a : "if you expected upper adn lower, they are language composition ops, not compiler ones, so that's out of scope"
-}
-
-Likewise, the {\em lower} composition is concerned about compilers that accept the same source but yield different target languages and the same theoretical results apply.
-However, especially for this kind of composition, it is rather difficult to provide a general definition.
-Consider two compilers both accepting $\src{LLVM IR}$ and one of them emits $\trg{x86\_64}$, while the other emits $\trg{ARMv8}$.
-It is intuitive that they are in some sense composed in the LLVM framework, but the decision of when to use one over the other is inherently {\em unobservable} to the formalization effort of this kind of composition.
-Either the user of LLVM provides an explicit flag that instructs the overall compiler to emit $\trg{x86\_64}$ or the framework itself detects the target platform via heuristics.
-%Either way, the theoretical results suggest that it is possible to always find a ,,most-general'', secure compiler, given two secure compilers, that robustly preserves the least-upper bound of the classes involved in their compilation process.
-In general, there are two options to deal with these nuances in a concrete, practical setting:
-Given secure compilers, either (i) view the whole compilation chain involving them as the upper/sequential/lower compositions and conclude the robust preservation of the intersection of all involved classes, or (ii) specify meta-parameters that, e.g., specify the target language, and thus reduce the compilation chain to a nested sequential composition of the involved compilers.
-The first approach yields a more imprecise statement about the whole compilation chain while the second approach is more precise, but also less flexible.
-
 \begin{figure}
@@
 \end{figure}
+This {\em phase ordering problem} is well--known in literature and a practical solution is to simply perform a fixpoint iteration of the optimisation pipeline~\cite{click1995combining}.
+%If possible, optimisations are merged to enhance the analysis' result, since even when iterating to a fixpoint the final program may not be as optimised as it could be~\cite{click1995combining}.
+Compiler engineers typically try to find an order of optimisations that yields well-optimised programs for either code size~\cite{cooper1999geneticphases} or performance~\cite{kulkarni2006exhaustivephase}.
+\Cref{corr:swappable} justifies that any such order of compilation passes is valid with respect to security.
+So, given two compilation passes \bul{$\cc[_{1}]{\trg{L}}{\trg{L}}$, $\cc[_{2}]{\trg{L}}{\trg{L}}$, both robustly preserving class $\class[_{1}]$ or $\class[_{2}]$, respectively}, \oul{for any order of their composition the composed compiler robustly preserves the intersection of $\class[_{1}]$ and $\class[_{2}]$}.
+
+\begin{corollary}[Sequential Composition of Secure Compilers]\label{corr:swappable}
+  If \bul{$\rtp{\cc[_{1}]{\trg{L}}{\trg{L}}}{\class[_{1}]}$ and $\rtp{\cc[_{2}]{\trg{L}}{\trg{L}}}{\class[_{2}]}$}, then \oul{$\rtp{\cc[_{1}]{\trg{L}}{\trg{L}}\circ\cc[_{2}]{\trg{L}}{\trg{L}}}{\class[_{1}]\cap\class[_{2}]}$ and $\rtp{\cc[_{2}]{\trg{L}}{\trg{L}}\circ\cc[_{1}]{\trg{L}}{\trg{L}}}{\class[_{2}]\cap\class[_{1}]}$}.
+\Coqed
+\end{corollary}
+
+\subsection{Secure Upper and Lower Composition}\label{sec:other-compos}
+Besides sequential composition, there are two other compositions that are more at the language- than the compiler-level\MP{namely ...}.
+Nevertheless, consider $\trg{JVM BC}$, which is a popular target for programming language designers due to its high performance and relevance in industry.
+As an example, $\src{Java}$ and $\obj{Kotlin}$ compile to $\trg{JVM BC}$.
+Technically speaking, they both lower\MP{undefined verb. no need for synonyms here.} to class files and $\obj{Kotlin}$ objects are considered to be the same as $\src{Java}$ objects at that point.
+A compiler that accepts both $\src{Java}$ and $\obj{Kotlin}$ code translating to the same target language, or rather, intermediate representation, performs a kind of {\em upper} composition.
+Given a secure compiler from $\src{Java}$ to $\trg{JVM BC}$ and one from $\obj{Kotlin}$ to $\trg{JVM BC}$, it is easy to envision that one can engineer a combined compiler that accepts both $\src{Java}$ and $\obj{Kotlin}$.
+A practical example of this, which is, however, insecure, is {\tt Android Studio}~\cite{androidstudio}.
+The proof of \Cref{thm:rtp} can be easily adapted to also works for this {\em upper} composition.
+The corresponding theorem is not listed, since there is no general definition of this kind of composition.
+\MPin{
+	mmm need to think about this.
+	this looks like discussion section material
+}
+\MPin{
+	we could turn this into a : "if you expected upper adn lower, they are language composition ops, not compiler ones, so that's out of scope"
+}
+
+Likewise, the {\em lower} composition is concerned about compilers that accept the same source but yield different target languages and the same theoretical results apply.
+However, especially for this kind of composition, it is rather difficult to provide a general definition.
+Consider two compilers both accepting $\src{LLVM IR}$ and one of them emits $\trg{x86\_64}$, while the other emits $\trg{ARMv8}$.
+It is intuitive that they are in some sense composed in the LLVM framework, but the decision of when to use one over the other is inherently {\em unobservable} to the formalization effort of this kind of composition.
+Either the user of LLVM provides an explicit flag that instructs the overall compiler to emit $\trg{x86\_64}$ or the framework itself detects the target platform via heuristics.
+%Either way, the theoretical results suggest that it is possible to always find a ,,most-general'', secure compiler, given two secure compilers, that robustly preserves the least-upper bound of the classes involved in their compilation process.
+In general, there are two options to deal with these nuances in a concrete, practical setting:
+Given secure compilers, either (i) view the whole compilation chain involving them as the upper/sequential/lower compositions and conclude the robust preservation of the intersection of all involved classes, or (ii) specify meta-parameters that, e.g., specify the target language, and thus reduce the compilation chain to a nested sequential composition of the involved compilers.
+The first approach yields a more imprecise statement about the whole compilation chain while the second approach is more precise, but also less flexible.
 

### 2023-07-09 popl/main.tex
bits

--- 2023-07-09 popl/main.tex
@@
 An example of such secure compilers are those that ascribe to the theory of robust compilation~\cite{abate2019jour,abate2021extacc,patrignani2021rsc}, which takes this interaction between components and potentially malicious contexts into account.
-Unfortunately, it is unclear if and how robust compilation supports compositionality.
-
-% Compositionality is an inherent part of software engineering and real-world compilers are no exception to that~\cite{lattner2004llvm,googlev8}.
-Compilation passes are typically developed as a standalone module to be plugged into a compilation pipeline and composed with other passes.
-Different kinds of such passes exist, e.g., there are usually several compilation passes that perform optimisations~\cite{lattner2004llvm,googlev8,androidstudio,kuepper2023cryptopt,manjikian1997fusion,wegman1991ccp} or source-code instrumentations~\cite{nagarakatte2009soft,nagarakatte2010cets,akritidis2009baggy,dhumbumroong2020boundwarden,jung2021pico,nam2019framer,shankaranarayana2023tailcheck,younan2010paricheck,zhou2023fatptrs,bond2017vale,almeida2017jasmin,kuepper2023cryptopt,cauligi2019fact} that enforce that the compiled program fulfils a property of interest.
-Without a modular framework, there is no hope to enable this separation of concerns, where, e.g., a secure pass that ensures \gls{ms} can be developed independently of a secure pass for \gls{cct}.
+
+Unfortunately, it is unclear if and how robust compilation supports compositionality, and compilation passes are typically developed as a standalone module to be plugged into a compilation pipeline and composed with other passes.
+Different kinds of compilation passes exist, e.g., there are usually several compilation passes that perform optimisations~\cite{lattner2004llvm,googlev8,androidstudio,kuepper2023cryptopt,manjikian1997fusion,wegman1991ccp} or source-code instrumentations~\cite{nagarakatte2009soft,nagarakatte2010cets,akritidis2009baggy,dhumbumroong2020boundwarden,jung2021pico,nam2019framer,shankaranarayana2023tailcheck,younan2010paricheck,zhou2023fatptrs,bond2017vale,almeida2017jasmin,kuepper2023cryptopt,cauligi2019fact} that enforce that the compiled program fulfils a property of interest.
+Without a framework for composing secure compiler passes, it is not possible to enable separation of concerns, e.g., to have a secure compilation pass that ensures \gls{ms} that is developed independently of another secure pass for \gls{cct} as showcased in \Cref{ex:strncpy}.
 %% MK: no idea how/where to fit this in or if this is even necessary
 %Because of the potentially hostile context, secure compilers attain security by relying on protection mechanisms, such as ... and ...\MK{cits}.
-The aim of this paper is to devise the theoretical foundation to be able to build a secure compiler as the composition of several smaller ones.
-Let us now introduce a minimal example of what such a secure compiler should be capable of dealing with.
+% This paper presents one such framework to be able to build a secure compiler as the composition of several smaller ones.
+% Let us now introduce a minimal example of what such a secure compiler should be capable of dealing with.
 %\MPin{ MK: hope it's solved now
@@
 For sake of simplicity, assume that the buffer \texttt{y} is at least as large as \texttt{n} as well as \texttt{x} terminated by \texttt{'\textbackslash 0'}.
-Even under these assumptions, there is a subtle \gls{ms} bug in the program which can be exploited to read sensitive data:
-The bounds check, i.e., \texttt{i < n}, should happen {\it prior} to the access to memory location \texttt{x[i]}.
+Even under these assumptions, there is a subtle \gls{ms} bug in the program which can be exploited to read sensitive data: The bounds check, i.e., \texttt{i < n}, should happen {\it prior} to the access to memory location \texttt{x[i]}.
 Otherwise, if \texttt{n} is larger than the length of the array object \texttt{x} points to, the memory location past this array object is read, which is an out-of-bounds memory access.
@@
 The fix commonly employed in such a case is to rewrite the program to run a {\it fixed} number of iterations regardless of the length of \texttt{x} or \texttt{y}.
+\MP{ can we write this is a compiler-inserted fix?}
 To this end, there must be an upper bound for the length of \texttt{x} and \texttt{y} and then the loop can just iterate until the induction variable \texttt{i} hits this upper bound.
 
-These fixes alone are not enough to make this program robust, i.e., not violating either \gls{ms} or \gls{cct} in a potentially malicious context.
+These fixes alone are not enough to make this program secure in a robust setting, i.e., not violating either \gls{ms} or \gls{cct} in a potentially malicious context.
 Consider a program context that calls \texttt{strncpy} with a very large \texttt{n}, some uninitialised memory \texttt{x}, and \texttt{y} does not matter for the sake of this example.
@@
 To mitigate these situations, it is important that \texttt{strncpy} ensures that the provided arguments satisfy the assumptions being made by the maintainer of \texttt{strncpy}, since the calling context can behave in a hostile manner.
-Unfortunately, these checks do not work in C\MP{why are we talking of C now?} without additional features, since there is no way to check whether a pointer points into a valid allocated region or not.
+Unfortunately, these checks do not work in C without additional features, since there is no way to check whether a pointer points into a valid allocated region or not.
 To enable this would require extra metadata~\cite{kwon2013lowfat,akritidis2009baggy} which must not be modifiable by a malicious context.

### 2023-07-09 popl/main.tex
5 e 6

@@
 However, when allocating pointers\MP{arrays?} with $\lnew{\expr[_{1}]}{\expr[_{2}]}$, the expression $\expr[_{1}]$ determines the initial value of the objects\MP{cells?} and $\expr[_{2}]$ the overall size of an array.
+
+\MPin{
+	below is semantics info. should go with their symbols and the semantics rules.
+	and a better description of the semantics symbols.
+
+}
 The data associated to pointers resides in the dynamic state $\Omega$ that contains sandboxed heaps, one for the context and the other for the component.
@@
 Because of this, it is possible that programs like $\llet{x}{\lnew{0}{5}}{\ldelete{x};\ldelete{x}}$ reduce and emit a trace along the lines of $\ev{Alloc\ \loc\ \sandboxtag\ 5}\cdot\ev{Dealloc\ \loc\ \sandboxtag}\cdot\ev{Dealloc\ \loc\ \sandboxtag}$.
+\MPin{
+	last line is mixing operational semantics and trace semantics
+}
 
@@
 
-Possible events that are emitted during execution are similar to the ones of the specification trace model.
-The key difference is that, as standard in secure compilation work, the traces have a call and return event that signals context switches.
+Possible events that are emitted during execution are similar to the ones of the specification trace model \MP{ref?}.
+The key difference is that, as standard in secure compilation work, the traces have a call and return event that signals context switches.\MP{why? to help with proofs?}
 Hereby, a $\ev{Call\ \ctxtocomp\ foo\ \valueexpr}$ and $\ev{Return\ \ctxtocomp\ \valueexpr}$ signal that program execution transitions from context- to component-level.
@@
 In the following, $\neg\ctx=\comp$ and $\neg\comp=\ctx$ and also $\operatorname{comm}(\ctx)=\ \ctxtocomp$ and $\operatorname{comm}(\comp)=\ \comptoctx$.
+\MP{intuition? these are aux functions for the semantics}
 
@@
 
-$\src{L_{\tmssafe}}$ does not extend the base syntax at term level.
-What makes $\src{L_{\tmssafe}}$ special is that it has a type system inspired by $L^{3}$~\cite{morrisett2005L3} that exhibits the property that every well-typed $\src{L_{\tmssafe}}$ program satisfies \gls{tms}.
-The proof of this property relies on a projection $\msfilterLtms{\src{\event}}=\event$ that projects $\src{L_{\tmssafe}}$ events to specfication events $\event$.
-Hereby, $\delta$ maps $\src{L_{\tmssafe}}$ locations $\src{\loc}$ to locations $\loc$ of the specification trace model.
-The projection $\msfilterLtms{\src{\event}}$ for the most part does the expected.
+$\src{L_{\tmssafe}}$ does not extend the base syntax at term level.\MP{avoid negations. ltms programs use the same syntax presentd before without extensions}
+What makes $\src{L_{\tmssafe}}$ special is that it has a type system inspired by $L^{3}$~\cite{morrisett2005L3} that exhibits the property that every well-typed $\src{L_{\tmssafe}}$ program satisfies \gls{tms}.\MPin{it's not really special ...
+Just say  "L3 has a type system ...
+""}
+The proof of this property relies on a projection $\msfilterLtms{\src{\event}}=\event$ that projects $\src{L_{\tmssafe}}$ events to specfication events $\event$. \MP{why}
+Hereby, $\delta$ maps $\src{L_{\tmssafe}}$ locations $\src{\loc}$ to locations $\loc$ of the specification trace model.\MP{why, where is this needed}
+The projection $\msfilterLtms{\src{\event}}$ for the most part does the expected\MP{what}.
 But, it also filters any action that a contexts does, since these are irrelevant for component-level \gls{tms}.
@@
 $\Ltrg$ extends the syntax presented earlier (\Cref{subsec:cs:defs}) with dynamic typechecks and a way to inspect poison tags in the metadata of pointers.
-If an object a pointer bound to variable $\trg{x}$ points to is allocated ($\trg{\poisonless}$), the check $\trg{\lispoisoned{x}}$ yields $\trg{1}$.
-Likewise, if it has been poisoned ($\trg{\poisoned}$), the check $\trg{\lispoisoned{x}}$ evaluates to $\trg{0}$.
-Besides these checks, the language remains exactly the same.
+If an object\MP{?} a pointer bound to variable $\trg{x}$ points to is allocated ($\trg{\poisonless}$), the check $\trg{\lispoisoned{x}}$ yields $\trg{1}$.
+Likewise, if it\MP{what} has been poisoned ($\trg{\poisoned}$), the check $\trg{\lispoisoned{x}}$ evaluates to $\trg{0}$.
+Besides these checks, the language remains exactly the same\MP{to what}.
 
@@
 \subsection{$\Lms$: Another Memory-Unsafe Language}\label{subsec:lms}
-
-This paper chooses to introduce $\Lms$, despite it being exactly equal to $\Ltrg$.
-The main purpose of this decision is to enhance readability.
+This paper chooses to introduce\MP{uses / defines . use 1 word} $\Lms$, despite it being exactly equal to $\Ltrg$.
+The main purpose of this decision is to enhance readability.\MPin{you can shorten these 2 sentences into 1. yes the section will be supershort, it's ok}
 The projection $\msfilterLms{\irl{\trace}}$ is also exactly equal to $\msfilterL{\trg{\trace}}$.
@@
 
-$\Lscct$ extends the dynamic state of $\trg{L}$ (\Cref{subsec:lms}) with a register for a data (operand) independent timing mode.
+$\Lscct$ extends the dynamic state\MPin{i really prefer operational state ...
+or runtime state.
+i think i'm not used to this terminology.} of $\trg{L}$ (\Cref{subsec:lms}) with a register for a data (operand) independent timing mode.
 For the other languages seen earlier, this mode is intuitively always-on.
+\MP{spell out the formal changes in an align* environmtn}
 To modify the mode, $\Lscct$ extends expressions with the term $\obj{\lwrdoit{\expr}}$ that writes the value of $\obj{\expr}$ into the register and with a term $\obj{\lrddoit{x}{\expr}}$ that reads the value of the register into $\obj{x}$.
@@
 To prevent leakage of secrets but still enable reasoning about memory safety, pre-events are extended with $\obj{\widehat{{Get}}\ \loc\ \valueexpr}$ and $\obj{\widehat{Set}\ \loc\ \valueexpr}$, which indicate reads from and writes to memory without leaking secret information involved in the access.
+\MPin{
+	an example program and how it uses the flag and what kind of traces are produced is really needed here
+}
 
@@
 \section{Case Study: Composing Secure Compiler Passes and Optimisations \pages{7}}\label{sec:casestud:rtp}
-
-\begin{figure}[h]
+\MP{use modifer [!h]. i HATE figs that float outside their section}
+\begin{figure}[!h]
   \begin{center}
@@
 \end{figure}
-This section defines several secure compilers, each of which robustly preserves a different property of interest (\Cref{fig:pipeline}).
-It demonstrates the power of the framework (\Cref{sec:compprop,sec:compcomp}) by composing these compilers for a secure and optimising compilation chain that robustly preserves \gls{msscct}.
+
+This section defines several secure compilers, each of which robustly preserves a different property of interest as depicted in \Cref{fig:pipeline}.
+\MPin{
+	explain how to read the figure: arrows mean, annotations mean ... ...
+}
+It\MP{what} demonstrates the power of the framework (\Cref{sec:compprop,sec:compcomp}) by composing these compilers for a secure and optimising compilation chain that robustly preserves \gls{msscct}.
 The first step in this chain is the compiler from $\src{L_{\tmssafe}}$ to $\trg{L}$ that robustly preserves just \gls{tms} (\Cref{thm:cca:rtp:tms}).
 From here, an instrumentation from $\trg{L}$ to $\irl{L_{\mssafe}}$ ensures that no out-of-bounds accesses can happen and, thus, programs at this point attain \gls{sms} (\Cref{thm:ccb:rtp:sms}).
-This is enough to get a compiler that robustly preserves \gls{ms} (\Cref{thm:ccab:rtp:ms}).
+Since these properties compose into \MP{}, composing these passes yields a compiler that robustly preserves \gls{ms} (\Cref{thm:ccab:rtp:ms}).
 At this stage, the section presents two optimising translations, namely \gls{cf} and \gls{dce}, each of which robustly preserves \gls{ms} (\Cref{thm:ccdce:rtp:ms,thm:cccf:rtp:ms}).
 These translations can be freely ordered in the compilation chain without compromising memory safety (\Cref{thm:cccfccdce:rtp:ms}).
-The last step ensures that code stays \gls{scct} (\Cref{thm:ccscct:rtp:scct}) when lowered from $\Lms$ to $\Lscct$ which yields the final result that the whole compilation chain robustly preserves \gls{msscct} (\Cref{thm:ccall:rtp:msscct}).
+The last step of the chain ensures that code stays \gls{scct} (\Cref{thm:ccscct:rtp:scct}) when lowered from $\Lms$ to $\Lscct$.
+The final result is that the whole compilation chain robustly preserves \gls{msscct} (\Cref{thm:ccall:rtp:msscct}).
 
 \subsection{Robust Temporal Memory Safety Preservation}\label{subsec:cs:tms}
-
+\MPin{
+	this environment should just be a align*
+}
   $$
@@
 
-The compiler does not have to do anything special besides a typecheck to respect the calling-convention.
-That is, because $\trg{L}$ has no static typechecks, it could happen that a bogus context $\trg{\library_{\ctx}}$ invokes a callable object accepting a $\src{\natt}$ with $\trg{\lpair{17}{29}}$.
+The compiler only inserts a dynamic typecheck to respect the well-typedness of function bodies.
+Since $\trg{L}$ has no static typechecks, it could happen that a bogus context $\trg{\library_{\ctx}}$ invokes a callable object accepting a $\src{\natt}$ with $\trg{\lpair{17}{29}}$.
 By inserting the check, the compiler ensures that execution does not proceed in such cases.
-The compiler does not insert other checks and simply recolors $\src{L_{\tmssafe}}$ to $\trg{L}$ expressions.
-Hypothesizing an implementation of the \texttt{strncpy} function from \Cref{sec:introduction} in $\Ltms$, the compiler would in this case ensure that the arguments are valid.
+The compiler does not insert other checks and proceeds as the identity function (which in this paper amounts to a simple re-colouring of $\src{L_{\tmssafe}}$ to $\trg{L}$ expressions).
+
+Compiling the \texttt{strncpy} function from \Cref{sec:introduction} with $\cca$, the compiler would in this case ensure that the arguments are valid.
 
@@
 
+\MPin{
+	this should be a subsection, not a proof sketch.
+}
+
 \begin{proof}[Proof Sketch]
 The proof of \Cref{thm:cca:rtp:tms} uses standard techniques in secure compilation literature and is visualized in \Cref{fig:proofdiag:rtp}.
+\MPin{
+	explain how to read the figure
+}
 Concretely, given the execution of the compiled $\Ltms$ program linked with an arbitrary $\trg{L}$ context yields a trace $\trg{\trace}$, it uses a trace-based backtranslation of $\trg{\trace}$ to reconstruct an $\src{L_{\tmssafe}}$ context.
+\MPin{
+	what does this mean?
+	what is a backtranslation?
+}
+\MPin{
+	why can we use a BT?
+	it's because the def in section 2.2 is eq to anotehr one, because the props are Safety, we use the RSC equivalent def.
+	all these intuitions are missing.
+	they are technical, so we only keep them here.
+}
 \Cref{thm:wt:tms} already gives strong guarantees for programs written in $\src{L_{\tmssafe}}$ and can be exploited for the backtranslated context by ensuring that it typechecks.
-Well-typing means the $\src{L_{\tmssafe}}$ program runs, so the rest of the proof is about relating the $\trg{L}$-level execution with the $\src{L_{\tmssafe}}$-level execution.
-
-For the $\trg{L}$-execution, the compiler wrapper code at state $\trg{\Omega_{1}}$ may either crash the program or continue execution $\trg{\Omega_{w_{1}}}$ right before control is transferred to the component.
-In this specific case, the compiler need not to insert additional code after the component is done with its execution, but a symmetric situation can happen here.
-Similarily, the backtranslation may also insert wrapper code prior and after switching to and from the component to the context.
+Well-typing\MP{well-typedness} means the $\src{L_{\tmssafe}}$ program runs, so the rest of the proof is about relating the $\trg{L}$-level execution with the $\src{L_{\tmssafe}}$-level execution.
+
+For the $\trg{L}$-execution, the compiler wrapper\MP{undefined terminology.
+call it like this before} code at state $\trg{\Omega_{1}}$ may either crash the program or continue execution $\trg{\Omega_{w_{1}}}$ right before control is transferred to the component.
+In this specific case, the compiler needs not insert additional code after the component is done with its execution, but a symmetric situation can happen here.\MP{vague and imprecise}
+Similarly, the backtranslation may also insert wrapper code prior and after switching to and from the component to the context. \MP{wut? why is a wrapper needed?}
 \MKin{
@@
   i.e., explain $\multimap$ and $\approx$?
+  \MPin{
+  yes
+  }
+}
+\MPin{
+	the general idea that later proofs use the same structure is also not here.
 }
@@
 
-Up to getting and setting memory, the compiler is the identity function, i.e., all omitted cases are a recoloring of $\trg{L}$ to $\Lms$.
-For memory accesses, however, a bounds-check is inserted that enforces \gls{sms}.
+Up to getting and setting memory\MP{reading and writing?
+uniform lingo}, the compiler\MP{notation} is the identity function, i.e., all omitted cases are a recoloring of $\trg{L}$ to $\Lms$.\MP{reword as before}
+For memory accesses, however, the compiler inserts a bounds-check that enforces \gls{sms}.
 To this end, the compiler introduces another, fresh identifier $\irl{x_{SIZE}}$ for each allocation that binds $\irl{x}$ to keep track of the allocation size.
 Again, linking this back to the \texttt{strncpy} example corresponds to the insertion of bounds check right in front of the accesses to \texttt{y[i]} and \texttt{x[i]}.
+\MPin{
+	example code?
+}
 
@@
 
-It is now possible to make use of \Cref{thm:rtp} to extract a compiler from $\Ltms$ to $\Lms$ that is \gls{ms} by applying \Cref{thm:cca:rtp:tms,thm:ccb:rtp:sms}:
+It is now possible to make use of \Cref{thm:rtp} to extract a compiler from $\Ltms$ to $\Lms$ that is \gls{ms} by applying \Cref{thm:cca:rtp:tms,thm:ccb:rtp:sms}.
+\MP{
+	bad sentence. rewrite.
+}
 
@@
 \end{center}
+\MPin{
+	substitution are a runtime semantics thing.
+	what you want is a compiler with 2 arguments, where in one you keep track of the static substitions to perform,
+	and 2 cases on how to handle variables: if they're in the second arg, then substitute, otherwise don't
+}
 
@@
 Both are secure with respect to \gls{ms}, the proof of which is relatively easy, since the backtranslation can use the original context without modification and both transformations do not change anything with regards to allocation, use, or deallocation.
+\MPin{
+	the point is different: the proof is easy because the transofrmations do not change the way memory is touched,
+	and from a security perspective, the attacker has no extra power since the target is the same as in the source.
+	.
+	rewrite.
+}
 That is, the order of appearance of the events on the execution trace stays the same.
@@
 Given the fact that $\Lscct$ provides a \gls{cct}-mode, the compiler has to ensure that execution in the component always happen in this mode.
-The context could overwrite the flag and exit the mode, but upon invoking a callable object that is part of the component, the flag would be set again.
-Because of this, the compiler is secure with respect ot \gls{scct}, similarily proven as in \Cref{subsec:cs:tms}.
+\MPin{
+	reuse the word 'wrapper' here
+}
+The context can overwrite the flag and exit the mode, but upon invoking a callable object\MP{again with the word object} that is part of the component, the flag would be set again.
+Because of this, the compiler is secure with respect to \gls{scct}, similarily\MP{misspelled} proven as in \Cref{subsec:cs:tms}.
 
@@
 
-Let $\ccmsscct$ be the compiler that is the composition of $\cca$, $\ccb$, $\cccf$, $\ccdce$, and $\ccscct$, then the following theorem can be shown.
+Let $\ccmsscct$ be the compiler that is the composition of $\cca$, $\ccb$, $\cccf$, $\ccdce$, and $\ccscct$, then the following theorem can be shown\MP{holds}.
 
@@
 \end{theorem}
+\MPin{
+This you can write the proof of.
+Probably even in coq
+}
 

### 2023-07-09 popl/main.tex
ex 3.10

--- 2023-07-09 popl/main.tex
@@
 \Cref{tr:sms-use} performs a bounds check and \Cref{tr:sms-alloc} adds bounds information to the state of the monitor.
-The trace agreement is entirely straightforward and similar to the one fo \gls{tms}.
+The trace agreement is entirely straightforward and similar to the one for \gls{tms}.
 % \begin{center}
@@
 
-\begin{example}[\texttt{strncpy}]\label{ex:strncpy:sms}
-  Consider the \texttt{strncpy} functions from \Cref{ex:strncpy}.
-  \MKin{do trace and explain why monitor finds out that it violates sms prop}
+\begin{example}[Normal invocation of \texttt{strncpy}]\label{ex:strncpy:sms}
+  Consider the insecure \texttt{strncpy} function from \Cref{ex:strncpy} with a context \texttt{strncpy(2, x, y)}, where \texttt{x} and \texttt{y} are pointers to valid regions of memory with allocated space for exactly two cells and do not contain the null-terminating character \texttt{'\textbackslash 0'}.
+  For the sake of this example, the pointers have been allocated by the component and passed to the context.
+  The loop of \texttt{strncpy} will copy exactly two cells and then check the loop condition for the last time.
+  At that stage, the induction variable \texttt{i} is equal to \texttt{2} and, unfortunately, the order of checks is such that first the cell \texttt{x[i]} is read prior to bounds checking \texttt{i < n}.
+  Because of this, there is an out of bounds memory access right before exiting the function.
+  This is also visible on the trace, which can be sketched as $\ldots\cdot\ev{Alloc}\ \loc_{x}\ 2;\comp;\unlock\cdot\ldots\cdot\ev{Alloc}\ \loc_{y}\ 2;\comp;\unlock\cdot\ldots\cdot\ev{Use}\ \loc_{x}\ 0;\comp;\unlock\cdot\ev{Use}\ \loc_{y}\ 0;\comp;\unlock\cdot\ev{Use}\ \loc_{x}\ 1;\comp;\unlock\cdot\ev{Use}\ \loc_{y}\ 1;\comp;\unlock\cdot\ev{Use}\ \loc_{x}\ 2;\comp;\unlock\cdot\ldots$, where $\loc_{x}$ and $\loc_{y}$ are the memory addresses associated to \texttt{x} and \texttt{y}, respectively.
+  Omitting the events for all ,,$\ldots$'' for sake of brevity, the abstract monitor trace of this is $\absev{Alloc}\ \loc_{x}\ \comp\ 2\cdot\absev{Alloc}\ \loc_{y}\ \comp\ 2\cdot\absev{Use}\ \loc_{x}\ \comp\ 0\cdot\absev{Use}\ \loc_{y}\ \comp\ 0\cdot\absev{Use}\ \loc_{x}\ \comp\ 1\cdot\absev{Use}\ \loc_{y}\ \comp\ 1\cdot\absev{Use}\ \loc_{x}\ \comp\ 2$.
+  It is easy to see that, after the allocation events, the state of the monitor is $\left\{(\loc_{x};\comp;2),(\loc_{y};\comp;2)\right\}$.
+  Furthermore, all uses up to the last are acceptable by the monitor, but the last event does not satisfy the premise $2<2$ in \Cref{tr:sms-use}.
+Therefore, the whole program (\texttt{strncpy} linked with this kind of context) is not \gls{sms}.
 \end{example}

### 2023-07-09 popl/main.tex
ex.13

@@
 Without a framework for composing secure compiler passes, it is not possible to enable separation of concerns, e.g., to have a secure compilation pass that ensures \gls{ms} that is developed independently of another secure pass for \gls{cct} as showcased in \Cref{ex:strncpy}.
-%% MK: no idea how/where to fit this in or if this is even necessary
-%Because of the potentially hostile context, secure compilers attain security by relying on protection mechanisms, such as ... and ...\MK{cits}.
-% This paper presents one such framework to be able to build a secure compiler as the composition of several smaller ones.
-% Let us now introduce a minimal example of what such a secure compiler should be capable of dealing with.
-%\MPin{ MK: hope it's solved now
-%	this does not really introduce the compositionality issue.
-%	i think the problem is the last sentence.
-%	try and divert the story after the 'because' line, to go into the need for composing compilers
-%}
 \begin{example}[strncpy]\label{ex:strncpy}
@@
 The results showcase how the framework allows the kind of formal security reasoning that compiler writers already want (and do), obtaining precise, compositional security reasoning while providing minimal (and modular) proof effort.
-%\MPin{
-%	last line: connect with the many citations of the previous paragraph
-%}
 
@@
         The formalisation of this case study showcases the power of the presented framework.
-        Namely, the divide-and-conquer approach is a viable strategy even for engineering secure compilers.
+        Namely, the divide-and-conquer approach to software engineering is a viable strategy even for the developement of secure compilers.
 %        This demonstrates that the usage of the presented framework is both simpler and more expressive compared to one monolithic security proof of the whole compilation chain.
@@
 
-\begin{example}[DIT on vs DIT off]
-  \MKin{fill in the blank}
+\begin{example}[Data independent timing mode]
+  Consider an excerpt from a component that reads from non-confidential memory \texttt{x} and writes the result to confidential memory \texttt{y}.
+  The trace of just this part may look like $\ev{Use}\ \loc_{x}\ 1729;\comp;\unlock\cdot\ev{Use}\ \loc_{y}\ 1729;\comp;\lock$.
+  In terms of the abstract monitor trace, this is just $\absev{Any}$.
+  Running this on the monitor would result in getting stuck, since there is no matching rule to step in the presence of $\absev{Any}$ event. % pun intended
+  By means of additional features to ensure cryptographic constant time even in the presence of memory reads and loads, such as a flag to enable a data independent timing mode, which is present in both Arm~\cite[p.~543]{arm-refman} and Intel~\cite[p.~80]{intel-refman} processors, the original trace can contain $\ev{Use}\ \loc_{y}\ 1729;\comp;\unlock$ instead of $\ev{Use}\ \loc_{y}\ 1729;\comp\lock$.
+  Because of this, the whole trace of the component equates to $\absev{\emptyevent}$ and the monitor can step without getting stuck.
 \end{example}
@@
 I can't even find one just for memsafety
-Consider the compilation passes of LLVM~\cite{lattner2004llvm} whose source-code optimisations are industry--competitive\MP{meaning?}.
+Consider the compilation passes of LLVM~\cite{lattner2004llvm}.
 In LLVM, compiler passes are handled in the \texttt{PassManager} class.

### 2023-07-09 popl/main.tex
Section 4

@@
 
-\setul{0.5ex}{0.3ex}
+\setul{0.95ex}{0.3ex}
 
@@
 \subsection{Secure Upper and Lower Composition}\label{sec:other-compos}
-Besides sequential composition, there are two other compositions that are more at the language- than the compiler-level\MP{namely ...}.
-Nevertheless, consider $\trg{JVM BC}$, which is a popular target for programming language designers due to its high performance and relevance in industry.
-As an example, $\src{Java}$ and $\obj{Kotlin}$ compile to $\trg{JVM BC}$.
-Technically speaking, they both lower\MP{undefined verb. no need for synonyms here.} to class files and $\obj{Kotlin}$ objects are considered to be the same as $\src{Java}$ objects at that point.
-A compiler that accepts both $\src{Java}$ and $\obj{Kotlin}$ code translating to the same target language, or rather, intermediate representation, performs a kind of {\em upper} composition.
-Given a secure compiler from $\src{Java}$ to $\trg{JVM BC}$ and one from $\obj{Kotlin}$ to $\trg{JVM BC}$, it is easy to envision that one can engineer a combined compiler that accepts both $\src{Java}$ and $\obj{Kotlin}$.
-A practical example of this, which is, however, insecure, is {\tt Android Studio}~\cite{androidstudio}.
-The proof of \Cref{thm:rtp} can be easily adapted to also works for this {\em upper} composition.
-The corresponding theorem is not listed, since there is no general definition of this kind of composition.
-\MPin{
-	mmm need to think about this.
-	this looks like discussion section material
-}
-\MPin{
-	we could turn this into a : "if you expected upper adn lower, they are language composition ops, not compiler ones, so that's out of scope"
-}
-
-Likewise, the {\em lower} composition is concerned about compilers that accept the same source but yield different target languages and the same theoretical results apply.
-However, especially for this kind of composition, it is rather difficult to provide a general definition.
-Consider two compilers both accepting $\src{LLVM IR}$ and one of them emits $\trg{x86\_64}$, while the other emits $\trg{ARMv8}$.
-It is intuitive that they are in some sense composed in the LLVM framework, but the decision of when to use one over the other is inherently {\em unobservable} to the formalization effort of this kind of composition.
-Either the user of LLVM provides an explicit flag that instructs the overall compiler to emit $\trg{x86\_64}$ or the framework itself detects the target platform via heuristics.
+Besides sequential composition, there are two other compositions, namely an {\em upper}, i.e., a compiler that takes multiple inputs and yields one output, and a {\em lower} composition, i.e., a compiler that takes one input and yields multiple outputs.
+Examples of this are present in industry:
+Consider the Java Virtual Machine bytecode $\trg{JVM BC}$, which is a popular target for programming language designers due to its high performance and relevance in industry.
+Compilers for several programming languages have it as their target language, some popular instances are $\src{Java}$ and $\obj{Kotlin}$.
+Technically speaking, they both compile to class files and $\obj{Kotlin}$ objects are considered to be the same as $\src{Java}$ objects at that point.
+Both languages can be used at the same time in one project~\cite{androidstudio}.
+A compiler that accepts both $\src{Java}$ and $\obj{Kotlin}$ code translating to the same target language or intermediate representation performs a kind of {\em upper} composition.
+The definition of upper composition is as follows: given a program $\mathtt{p}$, its compiled $\trg{L}$ counterpart is obtained by plugging $\mathtt{p}$ into $\cc{\src{L}}{\trg{L}}$ if $\mathtt{p}\in\src{L}$ or by plugging $\mathtt{p}$ into $\cc{\obj{L}}{\trg{L}}$ if $\mathtt{p}\in\obj{L}$, written as $\cc{\src{L}+\obj{L}}{\trg{L}}$.
+Now, given \bul{$\cc{\src{L}}{\trg{L}}$ robustly preserves $\class[_{1}]$} and \rul{$\cc{\obj{L}}{\trg{L}}$ robustly preserves $\class[_{2}]$}, it follows that \oul{their upper composition $\cc{\src{L}+\obj{L}}{\trg{L}}$ robustly preserves the intersection of classes $\class[_{1}]$ and $\class[_{2}]$}.
+
+\begin{theorem}[Upper Composition of Secure Compilers]\label{thm:urtp}
+  If \bul{$\rtp{\cc{\src{L}}{\trg{L}}}{\class[_{1}]}$} and \rul{$\rtp{\cc{\obj{L}}{\trg{L}}}{\class[_{2}]}$}, then \oul{$\rtp{\cc{\src{L}+\obj{L}}{\trg{L}}}{\class[_{1}]\cap\class[_{2}]}$}. \Coqed
+\end{theorem}
+
+Dually, the {\em lower} composition is concerned about compilers that accept the same source but yield different target languages. %and the same theoretical results apply.
+Consider two compilers both accepting $\src{LLVM IR}$~\cite{lattner2004llvm} and one of them emits $\trg{x86\_64}$, while the other emits $\trg{ARMv8}$.
+It is intuitive that they are in some sense composed in the LLVM framework, but the decision of when to use one over the other is inherently {\em internal} to the formalisation effort of this kind of composition.
+For example, the user of this compiler provides an explicit flag that instructs to emit $\trg{x86\_64}$ or the framework itself detects the target platform via heuristics, such as supported instructions.
+The %proposed
+definition of lower composition $\cc{\src{L}}{\trg{L}+\obj{L}}$ is as follows: given a program $\src{p}$, its compiled $\trg{L}$ or $\obj{L}$ program is obtained by performing an internal decision and threading $\src{p}$ into $\cc{\src{L}}{\trg{L}}$ or into $\cc{\src{L}}{\obj{L}}$ depending on that decision.
+Now, given \bul{$\cc{\src{L}}{\trg{L}}$ robustly preserves $\class[_{1}]$} and \rul{$\cc{\src{L}}{\obj{L}}$ robustly preserves $\class[_{2}]$}, it follows that \oul{their lower composition $\cc{\src{L}}{\trg{L}+\obj{L}}$ robustly preserves the intersection of classes $\class[_{1}]$ and $\class[_{2}]$}.
+
+\begin{theorem}[Lower Composition of Secure Compilers]\label{thm:lrtp}
+  If \bul{$\rtp{\cc{\src{L}}{\trg{L}}}{\class[_{1}]}$} and \rul{$\rtp{\cc{\obj{L}}{\trg{L}}}{\class[_{2}]}$}, then \oul{$\rtp{\cc{\src{L}}{\trg{L}+\obj{L}}}{\class[_{1}]\cap\class[_{2}]}$}. %\Coqed
+\end{theorem}
+
 %Either way, the theoretical results suggest that it is possible to always find a ,,most-general'', secure compiler, given two secure compilers, that robustly preserves the least-upper bound of the classes involved in their compilation process.
-In general, there are two options to deal with these nuances in a concrete, practical setting:
-Given secure compilers, either (i) view the whole compilation chain involving them as the upper/sequential/lower compositions and conclude the robust preservation of the intersection of all involved classes, or (ii) specify meta-parameters that, e.g., specify the target language, and thus reduce the compilation chain to a nested sequential composition of the involved compilers.
-The first approach yields a more imprecise statement about the whole compilation chain while the second approach is more precise, but also less flexible.
 
@@
 This way, it is possible to obtain an executable, secure compiler.
-However, the formalization effort for backtranslation proofs is known to be enormous, so another research avenue may be to find better ways to (semi-)automate standard secure compilation proofs.
+However, the formalisation effort for backtranslation proofs is known to be enormous, so another research avenue may be to find better ways to (semi-)automate standard secure compilation proofs.
 

### 2023-07-09 popl/main.tex
make use of -> use

@@
 
-This section defines programming languages that the secure compilers defined in the next section will make use of\MP{make use of -> use! 3 words when 1 is sufficient}.
+This section defines programming languages that the secure compilers defined in the next section will use.
 To this end, this section defines the languages $\Ltms$, $\Ltrg$, $\Lms$, and $\Lscct$ which share many common elements (presented in \Cref{subsec:cs:defs}).
@@
 
-It is now possible to make use of \Cref{thm:rtp} to extract a compiler from $\Ltms$ to $\Lms$ that is \gls{ms} by applying \Cref{thm:cca:rtp:tms,thm:ccb:rtp:sms}.
+It is now possible to use \Cref{thm:rtp} to extract a compiler from $\Ltms$ to $\Lms$ that is \gls{ms} by applying \Cref{thm:cca:rtp:tms,thm:ccb:rtp:sms}.
 \MP{

### 2023-07-09 popl/main.tex
Overview of sec 5

@@
 To this end, this section defines the languages $\Ltms$, $\Ltrg$, $\Lms$, and $\Lscct$ which share many common elements (presented in \Cref{subsec:cs:defs}).
-In this paper, $\Ltms$ is the only statically typed language and exhibits the property that all well-typed programs are \gls{tms} (\Cref{??}).
-However, not all $\Ltms$ programs are \gls{sms}.\MP{reworded. was imprecise}
-That is, there are well--typed $\Ltms$ programs that perform an out--of--bounds access.
-In contrast, $\Ltrg$ is untyped and does not provide any guarantees with regards to \gls{ms} (\Cref{??}).
-$\Lms$ is exactly the same language as $\Ltrg$, but this paper still distinguishes the two for sake of readability (\Cref{??}).
+In this paper, $\Ltms$ is the only statically typed language and exhibits the property that all well-typed programs are \gls{tms} (\Cref{subsec:ltms}).
+However, not all $\Ltms$ programs are \gls{sms}.
+That is, there are well-typed $\Ltms$ programs that perform an out-of-bounds access.
+In contrast, $\Ltrg$ is untyped and does not provide any guarantees with regards to \gls{ms} (\Cref{subsec:lsms}).
+$\Lms$ is exactly the same language as $\Ltrg$, but this paper still distinguishes the two for sake of readability (\Cref{subsec:lms}).
 All three languages --- so $\Ltms$, $\Ltrg$, and $\Lms$ --- assume \gls{cct} to hold.
 
-Even though it is possible to write programs that, e.g., return early in loops, which is trivially not \gls{cct}, they can be transformed to adhere to \gls{cct}~\cite{cauligi2019fact}.
-\MPin{
-	no, it's different. 
-	there is no CCT because we asume it's out of the programmer's concers and since CT is arch-dependent, the compiler should preserve CT, not the programmer, via software
-}
-
-Language $\Lscct$ allows violating \gls{cct} by emitting events on, e.g., branching and division, that contain secrets (\Cref{??}).
-However, $\Lscct$ also contains a countermeasure to enforce \gls{cct}, the language provides a way to read and write to a {\em model--specific register} that enables a ``\gls{cct}-mode''.
-\MPin{ref to what real world stuff justifies this}
+This paper considers \gls{cct} as an architecture-dependent property that should not be one of the programmer's concerns~\cite{cauligi2019fact}.
+Such consideration is backed up by architecture providing a data (operand) independent timing mode, such as processors by Arm~\cite[p.~543]{arm-refman} and Intel~\cite[p.~80]{intel-refman}.
+In spirit of this, the language $\Lscct$ allows violating \gls{cct} by emitting events on, e.g., branching and division, that contain secrets (\Cref{subsec:lscct}), but provides a way to read and write to a {\em model--specific register} that enables a ``\gls{cct}-mode''.
 

### 2023-07-09 popl/main.tex
parts of sec 5

@@
   \begin{aligned}
-  \mi{(Symbols)}~&\function\bnfdef\ \lfunction{g}{x}{e}
+  \mi{(Functions)}~&\function\bnfdef\ \lfunction{g}{x}{e}
   &
@@
 \end{gather*}
-\MPin{
-	this latex had many issues.
-	the \& were at different spots
-	it used hspace to space things instead of \&
-	and it did not use \& to align the following things on a line
-	and it did not use multiple aligned into the gather
-	(what's the point of gather if you don't have several aligned inside? use a single align* instead)
-}
+% \MPin{
+%   this latex had many issues.
+%   the \& were at different spots
+%   it used hspace to space things instead of \&
+%   and it did not use \& to align the following things on a line
+%   and it did not use multiple aligned into the gather
+%   (what's the point of gather if you don't have several aligned inside? use a single align* instead)
+% }
+% oh... that's good to know! the tech report is also riddled with this... :/
 
@@
 \MP{why a fraction? what exprs are missing}
-The syntax and semantics are mostly standard.\MP{and therefore omitted}
-However, when allocating pointers\MP{arrays?} with $\lnew{\expr[_{1}]}{\expr[_{2}]}$, the expression $\expr[_{1}]$ determines the initial value of the objects\MP{cells?} and $\expr[_{2}]$ the overall size of an array.
+%% MK: Things that are missing here are: Evaluation contexts, states Ω and transitively other definitions affected by this
+The syntax and semantics are mostly standard and therefore omitted.
+However, when allocating arrays with $\lnew{\expr[_{1}]}{\expr[_{2}]}$, the expression $\expr[_{1}]$ determines the initial value of the cells and $\expr[_{2}]$ the overall size of an array.
 
@@
 	and a better description of the semantics symbols.
-
+% MK: this moved to appendix! should we bring it back?
 }
 The data associated to pointers resides in the dynamic state $\Omega$ that contains sandboxed heaps, one for the context and the other for the component.
-Because of the sandboxing, locations pointing into heaps must be associated with a control--tag\MP{symbol} to know the heap they point into.
-Other metadata in $\Omega$ entails a list of callable objects\MP{functions? objects are other things in PL} with information on wether they belong to the context or the component.
-That way, when calling a callable object, the semantics can take care of emitting appropriate events that signal a context switch.
-For the languages considered in this paper, no callable objects are first-class.
+Because of the sandboxing, locations pointing into heaps must be associated with a control-tag $\sandboxtag$ (either $\ctx$ or $\comp$) to know the heap they point into.
+Other metadata in $\Omega$ entails a list of functions with information on whether they belong to the context or the component.
+That way, when calling a function, the semantics can take care of emitting appropriate events that signal a context switch.
+For the languages considered in this paper, functions are second-class.
 So, the only values are pairs, pointers, and natural numbers.
 Another non-standard part of the languages is the concept of the poison-tag.
-The poison-tag marks a location as freed ($\poisoned$) without removing the information associated to the freed pointer from $\Omega$.
+The poison-tag marks an array as freed ($\poisoned$) without removing the information associated to the freed pointer from $\Omega$.
 This measure prevents the semantics from getting stuck when an already freed pointer is freed again.
-Because of this, it is possible that programs like $\llet{x}{\lnew{0}{5}}{\ldelete{x};\ldelete{x}}$ reduce and emit a trace along the lines of $\ev{Alloc\ \loc\ \sandboxtag\ 5}\cdot\ev{Dealloc\ \loc\ \sandboxtag}\cdot\ev{Dealloc\ \loc\ \sandboxtag}$.
-\MPin{
-	last line is mixing operational semantics and trace semantics
-}
+Because of this, it is possible that a program like $\llet{x}{\lnew{0}{5}}{\ldelete{x};\ldelete{x}}$ yields a trace $\ev{Alloc\ \loc\ \sandboxtag\ 5}\cdot\ev{Dealloc\ \loc\ \sandboxtag}\cdot\ev{Dealloc\ \loc\ \sandboxtag}$ when executing.
 
@@
 
-Possible events that are emitted during execution are similar to the ones of the specification trace model \MP{ref?}.
+Possible events that are emitted during execution are similar to the ones of the specification trace model (\Cref{subsec:propdefs}).
 The key difference is that, as standard in secure compilation work, the traces have a call and return event that signals context switches.\MP{why? to help with proofs?}

### 2023-07-10 popl/main.tex
incorporate feedback and add example to sec 5.5

@@
   \begin{aligned}
-  \mi{(Pre\text{-}Events)}~&\event[_{b}]\bnfdef \ev{Alloc}\ \loc\ \valueexpr \mid \ev{Dealloc}\ \loc \mid \ev{Get}\ \loc\ \valueexpr &\mid \ev{Set}\ \loc\ \valueexpr\ \valueexpr['] \mid \ev{Call}\ \comm\ g\ \valueexpr \mid \ev{Ret}\ \comm\ \valueexpr \mid \ev{Start} \mid \ev{End}\ \valueexpr \\
-  \mi{(Events)}~&\event\bnfdef \emptyevent \mid \lightning \mid \tup{\event[_{b}]}{\sandboxtag} \\
+  \mi{(Pre\text{-}Events)}~\event[_{b}]&\bnfdef \ev{Alloc}\ \loc\ \valueexpr \mid \ev{Dealloc}\ \loc \mid \ev{Get}\ \loc\ \valueexpr \mid \ev{Set}\ \loc\ \valueexpr\ \valueexpr['] \mid \ev{Call}\ \comm\ g\ \valueexpr \mid \ev{Ret}\ \comm\ \valueexpr \mid \ev{Start} \mid \ev{End}\ \valueexpr \\
+  \mi{(Events)}~\event&\bnfdef \emptyevent \mid \lightning \mid \tup{\event[_{b}]}{\sandboxtag} \\
   \end{aligned}
@@
 Possible events that are emitted during execution are similar to the ones of the specification trace model (\Cref{subsec:propdefs}).
-The key difference is that, as standard in secure compilation work, the traces have a call and return event that signals context switches.\MP{why? to help with proofs?}
+The key difference is that, as standard in secure compilation work~\cite{patrignani2021rsc,abate2019jour}, the traces have a call and return event that signals context switches, which are referred to as {\em interaction events}.
+The reason for these interaction events is technical: They are a proof artifact for reconstructing a source context from a potentially malicious target context, where during that translation, the insertion of some wrapper code right before context switching may be necessary to make the proof succeed.
 Hereby, a $\ev{Call\ \ctxtocomp\ foo\ \valueexpr}$ and $\ev{Return\ \ctxtocomp\ \valueexpr}$ signal that program execution transitions from context- to component-level.
@@
 The environmental semantics emits the tag $\nocomm$ for calls from context- to context-level or component- to component-level.
-In the following, $\neg\ctx=\comp$ and $\neg\comp=\ctx$ and also $\operatorname{comm}(\ctx)=\ \ctxtocomp$ and $\operatorname{comm}(\comp)=\ \comptoctx$.
-\MP{intuition? these are aux functions for the semantics}
+In the following, $\neg\ctx=\comp$ and $\neg\comp=\ctx$.
 
@@
 
-$\src{L_{\tmssafe}}$ does not extend the base syntax at term level.\MP{avoid negations. ltms programs use the same syntax presentd before without extensions}
-What makes $\src{L_{\tmssafe}}$ special is that it has a type system inspired by $L^{3}$~\cite{morrisett2005L3} that exhibits the property that every well-typed $\src{L_{\tmssafe}}$ program satisfies \gls{tms}.\MPin{it's not really special ...
-Just say  "L3 has a type system ...
-""}
-The proof of this property relies on a projection $\msfilterLtms{\src{\event}}=\event$ that projects $\src{L_{\tmssafe}}$ events to specfication events $\event$. \MP{why}
-Hereby, $\delta$ maps $\src{L_{\tmssafe}}$ locations $\src{\loc}$ to locations $\loc$ of the specification trace model.\MP{why, where is this needed}
-The projection $\msfilterLtms{\src{\event}}$ for the most part does the expected\MP{what}.
-But, it also filters any action that a contexts does, since these are irrelevant for component-level \gls{tms}.
+$\src{L_{\tmssafe}}$ uses the same syntax as presented earlier (\Cref{sec:casestud:defs}) without extensions to the term level.
+But, $\src{L_{\tmssafe}}$ is statically typed, where the type system is inspired by $L^{3}$~\cite{morrisett2005L3}.
+The type system of $\src{L_{\tmssafe}}$ exhibits the property that every well-typed $\src{L_{\tmssafe}}$ program satisfies \gls{tms}.
+The proof of this property relies on a projection $\msfilterLtms{\src{\event}}=\event$ that projects $\src{L_{\tmssafe}}$ events to specfication events $\event$, because the properties defined earlier (\Cref{subsec:propdefs}) are defined in the specification trace model.
+Hereby, $\delta$ maps $\src{L_{\tmssafe}}$ locations $\src{\loc}$ to locations $\loc$ of the specification trace model whenever an event with a location is projected.
+The projection $\msfilterLtms{\src{\event}}$ for the most part does the expected, e.g., $\msfilterLtms{\src{\ev{Alloc}\ \loc\ n;\comp;\securitytag}}=\ev{Alloc}\ \delta(\loc)\ n;\comp;\securitytag$.
+But, it also filters any action that a context does as well as the interaction events, since these are irrelevant for component-level \gls{tms}.
 
@@
 $\Ltrg$ extends the syntax presented earlier (\Cref{subsec:cs:defs}) with dynamic typechecks and a way to inspect poison tags in the metadata of pointers.
-If an object\MP{?} a pointer bound to variable $\trg{x}$ points to is allocated ($\trg{\poisonless}$), the check $\trg{\lispoisoned{x}}$ yields $\trg{1}$.
-Likewise, if it\MP{what} has been poisoned ($\trg{\poisoned}$), the check $\trg{\lispoisoned{x}}$ evaluates to $\trg{0}$.
-Besides these checks, the language remains exactly the same\MP{to what}.
-
+For allocated ($\trg{\poisonless}$) arrays bound to variable $\trg{x}$, the check $\trg{\lispoisoned{x}}$ yields $\trg{1}$.
+If the array bound to $\trg{x}$ was allocated, i.e., has been poisoned ($\trg{\poisoned}$), the check $\trg{\lispoisoned{x}}$ evaluates to $\trg{0}$.
+%
 The projection $\msfilterL{\trg{\trace}}$ is entirely similar to $\msfilterLtms{\src{\trace}}$.
@@
 \subsection{$\Lms$: Another Memory-Unsafe Language}\label{subsec:lms}
-This paper chooses to introduce\MP{uses / defines . use 1 word} $\Lms$, despite it being exactly equal to $\Ltrg$.
-The main purpose of this decision is to enhance readability.\MPin{you can shorten these 2 sentences into 1. yes the section will be supershort, it's ok}
+To enhance readablity, this paper uses $\Lms$, despite it being exactly equal to $\Ltrg$.
 The projection $\msfilterLms{\irl{\trace}}$ is also exactly equal to $\msfilterL{\trg{\trace}}$.
 
-\subsection{$\Lscct$: An Unsafe Language}\label{subsec:lscct}
-\MPin{
-	unsafe? 
-	a possibly-CCT language?
-}
-
-$\Lscct$ extends the dynamic state\MPin{i really prefer operational state ...
-or runtime state.
-i think i'm not used to this terminology.} of $\trg{L}$ (\Cref{subsec:lms}) with a register for a data (operand) independent timing mode.
+\subsection{$\Lscct$: A Language with a Data Independent Timing Mode}\label{subsec:lscct}
+
+While the previous languages (\Cref{subsec:ltms,subsec:lsms,subsec:lms}) are not concerned with \gls{cct}, $\Lscct$ extends $\Lms$ (\Cref{subsec:lms}) with a register flag for a data (operand) independent timing mode\footnote{As present in both Arm~\cite[p.~543]{arm-refman} and Intel~\cite[p.~80]{intel-refman} processors.}.
 For the other languages seen earlier, this mode is intuitively always-on.
-\MP{spell out the formal changes in an align* environmtn}
+\MP{spell out the formal changes in an align* environmtn} % this necessitates that we present Ω also in the shared lang defs
 To modify the mode, $\Lscct$ extends expressions with the term $\obj{\lwrdoit{\expr}}$ that writes the value of $\obj{\expr}$ into the register and with a term $\obj{\lrddoit{x}{\expr}}$ that reads the value of the register into $\obj{x}$.
@@
 If the register is turned off, secrets may appear on traces as well as new events $\obj{Branch\ n}$ and $\obj{Binop\ n}$, which the semantics emits for conditionals and certain binary operations, e.g., division.
-To prevent leakage of secrets but still enable reasoning about memory safety, pre-events are extended with $\obj{\widehat{{Get}}\ \loc\ \valueexpr}$ and $\obj{\widehat{Set}\ \loc\ \valueexpr}$, which indicate reads from and writes to memory without leaking secret information involved in the access.
-\MPin{
-	an example program and how it uses the flag and what kind of traces are produced is really needed here
-}
-
-The filters to the specification trace model are straightforward.
+To prevent leakage of secrets but still enable reasoning about memory safety, pre-events are extended with $\obj{\widehat{{Get}}\ \loc\ \valueexpr}$ and $\obj{\widehat{Set}\ \loc\ \valueexpr}$, which indicate reads from and writes to memory without leaking secret information involved in the access, which is taken care of by the projection to specification events.
+\Cref{ex:lscct} illustrates these changes.
+
+The projection to the specification trace model is mostly straightforward.
+But for events containing the pre-events $\obj{\widehat{Get}}$ and $\obj{\widehat{Set}}$, the projection always translates the security-tag $\obj{\sigma}$ to $\unlock$, regardless of its actual value.
+The pre-events themselves still translate to just $\ev{Get}$ and $\ev{Set}$, respectively.
+With this technical setup, the information wether a read or write happened on a secret value is not hidden by the semantics, e.g., by emitting $\obj{\emptyevent}$, but when projecting to specification events.
+This allows flexibility: The trace can be checked to satisfy different properties, such as, in this case, \gls{tms}, \gls{sms}, \gls{scct}, and their combined versions.
+
+\begin{example}[$\Lscct$ with and without data independent timing]\label{ex:lscct}
+  Consider the following $\Lscct$ program and the execution trace $\obj{\trace}$ juxtaposed with it.
+
+  $$
+  \begin{array}{lcc}
+  \obj{\lwrdoit{0};} & \mid & \obj{\emptyevent}\\
+  \obj{\llet{x_{\lock}}{\lnew{5}{4}}{}} & \mid & \obj{\ev{Alloc}\ \loc_{x}\ 4;\sandboxtag;\unlock} \\
+  \obj{\lset{x}{2}{\lifz{0_{\lock}}{1}{3}};} & \mid & \obj{\ev{Branch}\ 0;\sandboxtag;\lock\cdot\ev{Set}\ \loc_{x}\ 2\ 1;\sandboxtag;\lock} \\
+  \obj{\lwrdoit{1};} & \mid & \obj{\emptyevent}\\
+  \obj{\llet{y_{\lock}}{\lnew{5}{4}}{}} & \mid & \obj{\ev{Alloc}\ \loc_{y}\ 4;\sandboxtag;\unlock}\\
+  \obj{\lset{y_{\lock}}{2}{\lifz{0_{\lock}}{1}{3}};} & \mid & \obj{\emptyevent\cdot\widehat{Set}\ \loc_{y}\ 2;\sandboxtag;\lock}\\
+  \end{array}
+  $$
+
+  When the data independent timing mode is off, the execution yields events in similar fashion to before (\Cref{subsec:ltms,subsec:lsms,subsec:lms}).
+  But, if it is turned on, then the branching event does not fire anymore.
+  Moreover, writing to memory now yields the pre-event $\obj{\widehat{Set}}$, whose difference to $\obj{Set}$ becomes apparent when looking at the projection to the specification trace model.
+  Projecting the above trace to the specification trace model yields the following trace (emitting empty events):
+  $\ev{Alloc}\ \loc_{x}\ 4;\sandboxtag;\unlock\cdot\ev{Branch\ 0}\cdot\ev{Set}\ \loc_{x}\ 2\ 1;\sandboxtag;\lock\cdot\ev{Alloc}\ \loc_{y}\ 4;\sandboxtag;\unlock\cdot\ev{Set}\ \loc_{y}\ 2;\sandboxtag;\unlock$.
+  Clearly, secrets are hidden ($\unlock$) after this projection for the section of the program where the data independent timing mode is turned on, while still containing the appropriate events to check for \gls{ms}.
+\end{example}
 
 \section{Case Study: Composing Secure Compiler Passes and Optimisations \pages{7}}\label{sec:casestud:rtp}
-\MP{use modifer [!h]. i HATE figs that float outside their section}
 \begin{figure}[!h]

### 2023-07-10 popl/main.tex
sec 6

@@
 Thus, there is only a passive attacker model that observes the execution and, e.g., retreives secrets from that.
-To consider a stronger model similarily to what existing work does~\cite{abate2019jour,abate2021extacc,patrignani2021rsc}, the concept of satisfaction is extended with {\em robustness}.
+To consider a stronger model similarly to what existing work does~\cite{abate2019jour,abate2021extacc,patrignani2021rsc}, the concept of satisfaction is extended with {\em robustness}.
 Robust satisfaction considers partial programs $\progvar$, i.e., programs with free variables, which need to be defined prior to execution.
@@
 The combined monitor runs the one for \gls{tms} and the one for \gls{sms} in a lockstep.
-The trace agreement similarily just relates a specification event with an abstract \gls{tms}-event $\abstrace[_{\tmssafe}]$ and with an abstract \gls{sms}-event $\abstrace[_{\smssafe}]$.
+The trace agreement similarly just relates a specification event with an abstract \gls{tms}-event $\abstrace[_{\tmssafe}]$ and with an abstract \gls{sms}-event $\abstrace[_{\smssafe}]$.
 % \begin{center}
@@
   \end{center}
-  \caption{Visualisation of the optimising compilation pipeline that attains a combination of \gls{ms} and \gls{cct}.}\label{fig:pipeline}
+  \caption{Visualisation of the optimising compilation pipeline that attains a combination of \gls{ms} and \gls{cct}. %
+    Vertices in the graph are the programming languages from earlier sections (\Cref{sec:casestud:defs}). %
+    All edges are secure compilers, but dotted edges use the presented framework (\Cref{sec:compcomp}) and strikethrough edges classic proof techniques. %
+    The dashed lines partition the graph into the sections where the respective theorems are presented.
+  }\label{fig:pipeline}
 \end{figure}
@@
 This section defines several secure compilers, each of which robustly preserves a different property of interest as depicted in \Cref{fig:pipeline}.
-\MPin{
-	explain how to read the figure: arrows mean, annotations mean ... ...
-}
-It\MP{what} demonstrates the power of the framework (\Cref{sec:compprop,sec:compcomp}) by composing these compilers for a secure and optimising compilation chain that robustly preserves \gls{msscct}.
+The section demonstrates the power of the framework (\Cref{sec:compprop,sec:compcomp}) by composing these compilers for a secure and optimising compilation chain that robustly preserves \gls{msscct}.
 The first step in this chain is the compiler from $\src{L_{\tmssafe}}$ to $\trg{L}$ that robustly preserves just \gls{tms} (\Cref{thm:cca:rtp:tms}).
 From here, an instrumentation from $\trg{L}$ to $\irl{L_{\mssafe}}$ ensures that no out-of-bounds accesses can happen and, thus, programs at this point attain \gls{sms} (\Cref{thm:ccb:rtp:sms}).
-Since these properties compose into \MP{}, composing these passes yields a compiler that robustly preserves \gls{ms} (\Cref{thm:ccab:rtp:ms}).
+Since these properties compose into \gls{ms}, composing these passes yields a compiler that robustly preserves \gls{ms} (\Cref{thm:ccab:rtp:ms}).
 At this stage, the section presents two optimising translations, namely \gls{cf} and \gls{dce}, each of which robustly preserves \gls{ms} (\Cref{thm:ccdce:rtp:ms,thm:cccf:rtp:ms}).
@@
 \subsection{Robust Temporal Memory Safety Preservation}\label{subsec:cs:tms}
-\MPin{
-	this environment should just be a align*
-}
-  $$
-  \begin{array}{rcl}
+  \begin{align*}
 %    \cca(\src{\lbinop{\expr[_{1}]}{\expr[_{2}]}}) & = & \lbinop[\trg]{\left[\cca(\src{\expr[_{1}]})\right]}{\left[\cca(\src{\expr[_{2}]})\right]} \\ [0.33cm]
-    \cca(\src{\lfunction{g}{x:\natt\to\type[_{e}]}{\expr}}) & = & \lfunction[\trg]{\trg{g}}{\trg{x}}{\lifz[\trg]{\trg{\lhast{x}{\natt}}}{%
+    \cca(\src{\lfunction{g}{x:\natt\to\type[_{e}]}{\expr}}) = \lfunction[\trg]{\trg{g}}{\trg{x}}{\lifz[\trg]{\trg{\lhast{x}{\natt}}}{%
                                                                                                 \left[\cca(\src{\expr})\right] %
                                                                                                  }{\labort[\trg]}}
-  \end{array}
-  $$
+  \end{align*}
 
@@
 
-\MPin{
-	this should be a subsection, not a proof sketch.
-}
-
-\begin{proof}[Proof Sketch]
+\subsubsection{Proof Sketch of Robust Temporal Memory Safety Preservation}
+
 The proof of \Cref{thm:cca:rtp:tms} uses standard techniques in secure compilation literature and is visualized in \Cref{fig:proofdiag:rtp}.
@@
 }
-\end{proof}
 
@@
 
-Up to getting and setting memory\MP{reading and writing?
-uniform lingo}, the compiler\MP{notation} is the identity function, i.e., all omitted cases are a recoloring of $\trg{L}$ to $\Lms$.\MP{reword as before}
-For memory accesses, however, the compiler inserts a bounds-check that enforces \gls{sms}.
+The compiler $\cc{\trg{L}}{\Lms}$ only inserts bounds-checks to enforce \gls{sms} whenever reading from or writing to memory.
 To this end, the compiler introduces another, fresh identifier $\irl{x_{SIZE}}$ for each allocation that binds $\irl{x}$ to keep track of the allocation size.
-Again, linking this back to the \texttt{strncpy} example corresponds to the insertion of bounds check right in front of the accesses to \texttt{y[i]} and \texttt{x[i]}.
+Again, linking this back to the \texttt{strncpy} example (\Cref{ex:strncpy:sms}) corresponds to the insertion of bounds check right in front of the accesses to \texttt{y[i]} and \texttt{x[i]}.
+In turn, the event $\ev{Use}\ \loc_{x}\ 2;\comp;\unlock$ would not be emitted during execution.
 \MPin{
 	example code?
+    MK: still needed now that we reference ex. 3.10?
 }
@@
 
-It is now possible to use \Cref{thm:rtp} to extract a compiler from $\Ltms$ to $\Lms$ that is \gls{ms} by applying \Cref{thm:cca:rtp:tms,thm:ccb:rtp:sms}.
-\MP{
-	bad sentence. rewrite.
-}
+\Cref{thm:ccab:rtp:ms} follows from \Cref{thm:cca:rtp:tms,thm:ccb:rtp:sms} using \Cref{thm:rtp}.
 
@@
 	and 2 cases on how to handle variables: if they're in the second arg, then substitute, otherwise don't
+  %
+  MK: No?
+I get how they are used as part of the runtime semantics, but a compiler can just as well make use of a substitution function that models the runtime semantics.
+I see how these two are different.
+However, I think it's easier in this case to be a bit inprecise, since adding another argument to the compiler breaks our compositionality framework (unless it's hidden with a monad...) so, IMO, we should view substitution here as a meta-level, but still language-dependent, function that the compiler has access to.
+If we don't do this, we essentially re-encode the definition of subst into the compiler
 }
@@
 The two optimising compiler passes from $\Lms$ to $\Lms$ perform \gls{dce} and \gls{cf}, respectively.
-Both are secure with respect to \gls{ms}, the proof of which is relatively easy, since the backtranslation can use the original context without modification and both transformations do not change anything with regards to allocation, use, or deallocation.
-\MPin{
-	the point is different: the proof is easy because the transofrmations do not change the way memory is touched,
-	and from a security perspective, the attacker has no extra power since the target is the same as in the source.
-	.
-	rewrite.
-}
-That is, the order of appearance of the events on the execution trace stays the same.
+Both are secure with respect to \gls{ms}, the proof of which is relatively easy, because \gls{dce} and \gls{cf} do not change the way memory accesses happen.
+Moreover, since the input and output languages to these compilers are the same, attacker contexts do not have more power in the target language than in the source.
 
 \begin{theorem}[Compiler $\ccdce$ is secure with respect to \gls{ms}]\label{thm:ccdce:rtp:ms}
-  $\rtp{\ccdce}{\mssafe}$ % \Coqed
+  $\rtp{\ccdce}{\mssafe}$ \Coqed
 \end{theorem}
 \begin{theorem}[Compiler $\cccf$ is secure with respect to \gls{ms}]\label{thm:cccf:rtp:ms}
-  $\rtp{\cccf}{\mssafe}$ % \Coqed
+  $\rtp{\cccf}{\mssafe}$ \Coqed
 \end{theorem}
@@
 
-Given the fact that $\Lscct$ provides a \gls{cct}-mode, the compiler has to ensure that execution in the component always happen in this mode.
-\MPin{
-	reuse the word 'wrapper' here
-}
-The context can overwrite the flag and exit the mode, but upon invoking a callable object\MP{again with the word object} that is part of the component, the flag would be set again.
-Because of this, the compiler is secure with respect to \gls{scct}, similarily\MP{misspelled} proven as in \Cref{subsec:cs:tms}.
+Given the fact that $\Lscct$ provides a \gls{cct}-mode that can be turned on or off, the compiler inserts wrapper code for function bodies to ensure that execution in the component always happen in this \gls{cct}-mode.
+The context can overwrite the flag and exit the mode, but upon invoking a function that is part of the component, the flag would be set again.
+Because of this, the compiler is secure with respect to \gls{scct}, similarly proven as in \Cref{subsec:cs:tms}.
 
@@
 
-Let $\ccmsscct$ be the compiler that is the composition of $\cca$, $\ccb$, $\cccf$, $\ccdce$, and $\ccscct$, then the following theorem can be shown\MP{holds}.
+Let $\ccmsscct$ be the compiler that is the composition of $\cca$, $\ccb$, $\cccf$, $\ccdce$, and $\ccscct$, then the following theorem holds.
 
 \begin{theorem}[Compiler $\ccmsscct$ is secure with respect to \gls{scct}]\label{thm:ccall:rtp:msscct}
-  $\rtp{\cc{\Ltms}{\Lscct}}{\mssafe\cap\scctsafe}$ % \Coqed
+  $\rtp{\cc{\Ltms}{\Lscct}}{\mssafe\cap\scctsafe}$ \Coqed
 \end{theorem}
@@
 Probably even in coq
+%
+MK: Should this appear here in the paper? It's basically identical to \Cref{thm:ccab:rtp:ms}
 }
@@
 The paper exercises this framework on a case-study that consists of an optimising compilation chain that is secure with respect to \gls{msscct}.
-In future work, it would be interesting to investigate whether it is possible to provide {\em secure compiler combinators}, similarily to how parser combinators work.
+In future work, it would be interesting to investigate whether it is possible to provide {\em secure compiler combinators}, similarly to how parser combinators work.
 %To this end, it may be necessary to extend existing frameworks for multi-language semantics.

### 2023-07-10 popl/main.tex
make MK visible

@@
         % 
-        Starting from ways to formalise and enforce those properties individually, this paper shows how to compose their formalisation. % and their composition. %% MK: compose their composition??
+        Starting from ways to formalise and enforce those properties individually, this paper shows how to compose their formalisation.
         % 
@@
 With this, \Thmref{def:propsat} can be extended as follows: for \bul{components $\progvar$ to robustly satisfy a property $\pi$}, take an \iul{attacker context $\contextvar$ and link it with $\progvar$}, \oul{the result of which must satisfy $\pi$}.
-%\MPin{
-% the behaviour of the result?
-%%%%%%%    MK: def:propsat defines satisfaction in terms of programs, no need to talk about behavior here
-%}
 
@@
 A {\em compiler} ($\cc{\src{L}}{\trg{L}}$) translates syntactic descriptions of programs from a {\em source} ($\src{L}$) into a {\em target} ($\trg{L}$) programming language.
-%For the syntactic description of programs written in $\src{L}$, this paper writes $\src{\progvar}$, while for ones written in $\trg{L}$, this paper writes $\trg{\progvar}$.%\MP{not a fan of this sentence}
-%This paper writes $\src{\progvar}$ and $\trg{\progvar}$ for programs written in $\src{L}$ and $\trg{L}$, respectively.
-%      MK: I don't think we need to mention this at all
 This translation is considered {\em correct} if it is semantics-preserving.
@@
 \subsubsection{Securing Optimisations}
-%\MPin{note that i'm always using british spelling. optimisation VS optimization. etc. }
-% MK: I'll go over the doc with the antidote spellchecker as soon as it is reasonably stable.
-
 Notably, real-world compilation chains also perform a series of (sequential) passes whose main purpose is not necessarily to translate from one language to another, but to, e.g., optimise the code or enforce a certain property.
-%\MPin{
-%	do not say the negation of something before the non-negation.
-%	say what things are first, and what they are not later (if at all)
-%}  MK: that for sure is less confusing... Good suggestion!
-Both examples can be seen in practice~\cite{nagarakatte2009soft,nagarakatte2010cets,akritidis2009baggy,wegman1991ccp,manjikian1997fusion}, it is infeasible to list them all\MP{survey to cite?
-this is a bit informal}.
-% MK: hard to find a survey on different compiler passes...
-I can't even find one just for memsafety
+Both examples can be seen in practice~\cite{nagarakatte2009soft,nagarakatte2010cets,akritidis2009baggy,wegman1991ccp,manjikian1997fusion}, it is infeasible to list them all\MPin{survey to cite?
+this is a bit informal MK: hard to find a survey on different compiler passes...
+I can't even find one just for compiler enforced memsafety (that looks good) or for compiler translations in general}.
 Consider the compilation passes of LLVM~\cite{lattner2004llvm}.
@@
 Above is a fraction of the shared syntax of all the programming languages of this paper.
-\MP{why a fraction? what exprs are missing}
-%% MK: Things that are missing here are: Evaluation contexts, states Ω and transitively other definitions affected by this
+\MPin{why a fraction? what exprs are missing
+MK: Things that are missing here are: Evaluation contexts, states Ω and transitively other definitions affected by this
+}
 The syntax and semantics are mostly standard and therefore omitted.
@@
 	and a better description of the semantics symbols.
-% MK: this moved to appendix! should we bring it back?
+  MK: this moved to appendix! should we bring it back? what exactly should be left out?
 }

### 2023-07-10 popl/main.tex
compile fix

@@
 \MPin{why a fraction? what exprs are missing
-MK: Things that are missing here are: Evaluation contexts, states Ω and transitively other definitions affected by this
+MK: Things that are missing here are: Evaluation contexts, states $\Omega$ and transitively other definitions affected by this
 }
@@
 For the other languages seen earlier, this mode is intuitively always-on.
-\MP{spell out the formal changes in an align* environmtn} % this necessitates that we present Ω also in the shared lang defs
+\MP{spell out the formal changes in an align* environmtn} % this necessitates that we present $\Omega$ also in the shared lang defs
 To modify the mode, $\Lscct$ extends expressions with the term $\obj{\lwrdoit{\expr}}$ that writes the value of $\obj{\expr}$ into the register and with a term $\obj{\lrddoit{x}{\expr}}$ that reads the value of the register into $\obj{x}$.

### 2023-07-10 popl/main.tex
shrink intro ex

@@
 \begin{example}[strncpy]\label{ex:strncpy}
-Consider the program below which copies a \texttt{C}-style string \texttt{x} into \texttt{y} up to a length of \texttt{n}:
-
-\begin{lstlisting}[language=c,basicstyle=\small\ttfamily]
-void strncpy(int n, char *x, char *y) {
-  for(int i = 0; x[i] != '\0' && i < n; ++i) {
-    y[i] = x[i];
+Consider the \texttt{C} program below that copies a null-terminated string \texttt{src} into \texttt{dst} up to a length of \texttt{n}:
+
+\begin{lstlisting}[language=c,basicstyle=\small\ttfamily,morekeywords={size_t}]
+void strncpy(size_t n, char *dst, char *src) {
+  for(size_t i = 0; src[i] != '\0' && i < n; ++i) {
+    dst[i] = src[i];
   }
@@
 
-For sake of simplicity, assume that the buffer \texttt{y} is at least as large as \texttt{n} as well as \texttt{x} terminated by \texttt{'\textbackslash 0'}.
-Even under these assumptions, there is a subtle \gls{ms} bug in the program which can be exploited to read sensitive data: The bounds check, i.e., \texttt{i < n}, should happen {\it prior} to the access to memory location \texttt{x[i]}.
-Otherwise, if \texttt{n} is larger than the length of the array object \texttt{x} points to, the memory location past this array object is read, which is an out-of-bounds memory access.
-Unfortunately, even though flipping the conditions does fix the \gls{ms} issue, it still violates \gls{cct}, so the program is susceptible to timing attacks.
+For sake of simplicity, assume that the length of the array \texttt{dst} points to is at least as large as \texttt{n} as well as \texttt{src} terminated by \texttt{'\textbackslash 0'}.
+Even under these assumptions, there is a subtle \gls{ms} bug in the program: The bounds check \texttt{i < n} should happen {\it prior} to the access to memory location \texttt{x[i]}:
+If \texttt{n} is at least as large as the length of the array \texttt{src}, the memory location past the last element will be read.
+A compilation pass that enforces spatial memory safety, such as Softbounds~\cite{nagarakatte2009soft} or BaggyBounds~\cite{akritidis2009baggy}, could have prevented this accidental error.
+
+Unfortunately, fixing \gls{ms} is not enough to have a secure program, because \texttt{strncpy} is susceptible to timing attacks.
 This is because the loop can terminate early: As soon as the string-terminating character \texttt{'\textbackslash 0'} is encountered, the loop exits.
-Thus, the program execution time is proportional to the length of the array object \texttt{x}.
-The fix commonly employed in such a case is to rewrite the program to run a {\it fixed} number of iterations regardless of the length of \texttt{x} or \texttt{y}.
-\MP{ can we write this is a compiler-inserted fix?}
-To this end, there must be an upper bound for the length of \texttt{x} and \texttt{y} and then the loop can just iterate until the induction variable \texttt{i} hits this upper bound.
-
-These fixes alone are not enough to make this program secure in a robust setting, i.e., not violating either \gls{ms} or \gls{cct} in a potentially malicious context.
-Consider a program context that calls \texttt{strncpy} with a very large \texttt{n}, some uninitialised memory \texttt{x}, and \texttt{y} does not matter for the sake of this example.
-While the context seems conceptually simple, uninitialized memory can appear in programs very subtly and invoking \texttt{strncpy} now leads to a series of reads from uninitialised memory, which is an immediate \gls{ms} issue that can have devastating real-world implications~\cite{uninit-0,uninit-1,uninit-2,uninit-3,uninit-4}.
-To mitigate these situations, it is important that \texttt{strncpy} ensures that the provided arguments satisfy the assumptions being made by the maintainer of \texttt{strncpy}, since the calling context can behave in a hostile manner.
-Unfortunately, these checks do not work in C without additional features, since there is no way to check whether a pointer points into a valid allocated region or not.
-To enable this would require extra metadata~\cite{kwon2013lowfat,akritidis2009baggy} which must not be modifiable by a malicious context.
-So, for sake of simplicity, assume the language at hand provides an operator \texttt{is\_valid(p, s, n)} that faithfully checks whether a pointer \texttt{p} points into a valid array of objects sized \texttt{s} of length \texttt{n} or not -- in constant time.
-This could be an abstraction that the operating system provides.\MP{cits absolutely needed and the narrative needs to be stronger, using those cits}
-With this feature at hand, it is possible to modify the original \texttt{strncpy} to be robust, given this is the compiled program without optimisations:
-
-\begin{lstlisting}[language=c,basicstyle=\small\ttfamily]
-void strncpy(int n, char *x, char *y) {
-  int const c = __ct_min(n, MAX_ALLOC_SIZE);
-  assert(is_valid(x, sizeof(char), c) & is_valid(y, sizeof(char), c));
-  for(int i = 0; i < c; ++i) {
-    y[i] = x[i];
-  }
-}
-\end{lstlisting}
+Thus, the program execution time is proportional to the length of the array \texttt{src} points to.
+Often, programmers rewrite programs like these to have loops that run with a fixed number of iterations.
+However, compiler passes are capable of rewriting such programs~\cite{cauligi2019fact} and a modification by hand is not necessary.
+
+These fixes alone are not enough to make this program secure in a robust setting, i.e., a malicious context could trigger a violation of either \gls{ms} or \gls{cct}.
+Consider a program context \texttt{strncpy(1729,dst,src)} that calls \texttt{strncpy} with pointers \texttt{dst} and \texttt{src}, where \texttt{src} is some uninitialized memory.
+This leads to a series of reads from uninitialised memory, which is an immediate \gls{ms} issue that can have devastating real-world implications~\cite{uninit-0,uninit-1,uninit-2,uninit-3,uninit-4}.
+An easy fix to sidestep this is to use non-nullable pointers as parameters instead, e.g., as presented in Cyclone~\cite{jim2002cyclone}.
+With this, a secure compiler needs to ensure that no malicious target-level context can supply unitialised memory and thereby sidestep the guarantees of the source language, e.g., by inserting appropriate checks prior to entering the body of \texttt{strncpy}.
 \end{example}
@@
   Consider the following program which, when run, yields the trace $\ev{Alloc}\ \loc[_{1}]\ 17;\comp;\unlock\cdot\ev{Use}\ \loc[_{1}]\ 5;\comp;\unlock\cdot\ev{Use}\ \loc[_{1}]\ 5;\comp;\unlock\cdot\ev{Dealloc}\ \loc[_{1}];\comp;\unlock$, where $\loc[_{1}]$ is some memory address determined by the semantics of the programming language.
-  \begin{lstlisting}[language=c,basicstyle=\ttfamily,morekeywords={new,delete}]
+  \begin{lstlisting}[language=c,basicstyle=\small\ttfamily,morekeywords={new,delete}]
 int main() {  // allocates 17 integers initialized to 2

### 2023-07-10 popl/main.tex
intro.

@@
 
-\gls{ms} is a security property of programs that, e.g., ensures that there are no out-of-bounds accesses as well as no reads or writes to uninitialized memory.
-Without memory safety, private data can easily leak or be corrupted~\cite{lemay2021ccc}.
+\gls{ms} is a security property that ensures that there are no out-of-bounds accesses as well as no reads or writes to uninitialized memory.
+Programs that do not attain memory safety, can easily have their private data be leaked to attackers or be corrupted~\cite{lemay2021ccc}.
 Because of this, \gls{ms} is a requirement for programs to attain an even stronger security property, namely \gls{cct}.
-Programs that attain \gls{cct} are secure with respect to timing attacks, and this is the gold standard for secure applications, including cryptographic ones since timing attacks can expose secrets via a simple performance analysis~\cite{kocher1996timing}.
+Programs that attain \gls{cct} are secure with respect to timing attacks, and this is the gold standard for secure applications, including cryptographic ones, since timing attacks can expose secrets via a simple performance analysis~\cite{kocher1996timing}.
 Timing attacks have become more and more relevant recently, especially given the era of cloud-computing~\cite{aviram2010cloudtime,kumar2019cloudsecsurvey}, where programs share the same hardware but are supposed to run in a sandboxed environment that does not leak secrets.
@@
 
-Even though mitigations for such attacks exist~\cite{bond2017vale,almeida2017jasmin}, the literature shows that some compilers may optimize these mitigations away~\cite{barthe2018sec}.
-This is of course not what programmers expect and necessitates the need for a {\em secure} compiler that ensures such mitigations persist in the code they generate.
+Even though mitigations for such attacks exist~\cite{bond2017vale,almeida2017jasmin}, some compilers may optimize these mitigations away~\cite{barthe2018sec}.
+This is of course not what programmers expect and to ensure such mitigations persist in the compiled code, one can use a {\em secure} compilers.
 More generally, secure compilers should preserve security properties even when the program runs in a hostile context such as the cloud one, i.e., where it interacts with other possibly malicious programs.
@@
 Different kinds of compilation passes exist, e.g., there are usually several compilation passes that perform optimisations~\cite{lattner2004llvm,googlev8,androidstudio,kuepper2023cryptopt,manjikian1997fusion,wegman1991ccp} or source-code instrumentations~\cite{nagarakatte2009soft,nagarakatte2010cets,akritidis2009baggy,dhumbumroong2020boundwarden,jung2021pico,nam2019framer,shankaranarayana2023tailcheck,younan2010paricheck,zhou2023fatptrs,bond2017vale,almeida2017jasmin,kuepper2023cryptopt,cauligi2019fact} that enforce that the compiled program fulfils a property of interest.
-Without a framework for composing secure compiler passes, it is not possible to enable separation of concerns, e.g., to have a secure compilation pass that ensures \gls{ms} that is developed independently of another secure pass for \gls{cct} as showcased in \Cref{ex:strncpy}.
+Without a framework for composing secure compiler passes, it is not possible to enable separation of concerns, e.g., to have a secure compilation pass that ensures \gls{ms} that is developed independently of another secure pass for \gls{cct} as suggested before and as showcased in \Cref{ex:strncpy}.
 \begin{example}[strncpy]\label{ex:strncpy}
@@
 
-For sake of simplicity, assume that the length of the array \texttt{dst} points to is at least as large as \texttt{n} as well as \texttt{src} terminated by \texttt{'\textbackslash 0'}.
-Even under these assumptions, there is a subtle \gls{ms} bug in the program: The bounds check \texttt{i < n} should happen {\it prior} to the access to memory location \texttt{x[i]}:
+For sake of simplicity, assume that the length of the array \texttt{dst} points to is at least \texttt{n}, and \texttt{src} is terminated by \texttt{'\textbackslash 0'}.
+Even under these assumptions, there is a subtle \gls{ms} vulnerability in the program: The bounds check \texttt{i < n} should happen {\it before} the access to memory location \texttt{x[i]}:
 If \texttt{n} is at least as large as the length of the array \texttt{src}, the memory location past the last element will be read.
-A compilation pass that enforces spatial memory safety, such as Softbounds~\cite{nagarakatte2009soft} or BaggyBounds~\cite{akritidis2009baggy}, could have prevented this accidental error.
+To prevent this vulnerability, one can use a compilation pass that enforces spatial memory safety, such as Softbounds~\cite{nagarakatte2009soft} or BaggyBounds~\cite{akritidis2009baggy}.
 
@@
 This is because the loop can terminate early: As soon as the string-terminating character \texttt{'\textbackslash 0'} is encountered, the loop exits.
-Thus, the program execution time is proportional to the length of the array \texttt{src} points to.
-Often, programmers rewrite programs like these to have loops that run with a fixed number of iterations.
-However, compiler passes are capable of rewriting such programs~\cite{cauligi2019fact} and a modification by hand is not necessary.
+Thus, the program execution time is proportional to the length of the array that \texttt{src} points to.
+% Often, programmers rewrite programs like these to have loops that run with a fixed number of iterations.
+Also in this case there exist compiler passes that can rewrite such programs~\cite{cauligi2019fact} so that they are no longer vulnerable to timing attacks.
 
 These fixes alone are not enough to make this program secure in a robust setting, i.e., a malicious context could trigger a violation of either \gls{ms} or \gls{cct}.
-Consider a program context \texttt{strncpy(1729,dst,src)} that calls \texttt{strncpy} with pointers \texttt{dst} and \texttt{src}, where \texttt{src} is some uninitialized memory.
+Consider a program context that calls \texttt{strncpy} with an argument for \texttt{src} that points to uninitialized memory.
 This leads to a series of reads from uninitialised memory, which is an immediate \gls{ms} issue that can have devastating real-world implications~\cite{uninit-0,uninit-1,uninit-2,uninit-3,uninit-4}.
@@
 With this, a secure compiler needs to ensure that no malicious target-level context can supply unitialised memory and thereby sidestep the guarantees of the source language, e.g., by inserting appropriate checks prior to entering the body of \texttt{strncpy}.
+\MPin{
+	do we have examples of such wrapper-chesk inserting secure compilers?
+	mswasm compiler ... not really.
+	or we could say that another solution is  using capabilities, and then the mswasm compiler can be used
+}
 \end{example}
 
-This paper introduces a framework for reasoning about the composition of secure and optimising compiler passes and it showcases the power of this framework by instantiating it on a multi-pass compilation chain.
+This paper introduces a framework for reasoning about the composition of secure and optimising compiler passes akin to those of \Cref{ex:strncpy} and it showcases the power of this framework by instantiating it on a multi-pass compilation chain.
 To this end, this paper first discusses how to compose security properties, such as temporal and spatial memory safety into general memory safety, and cryptographic constant time.
-Then, this paper defines several secure compiler passes, where each is either preserving a different security property (e.g., temporal or spatial memory safety) or performing a security-preserving optimisation, (e.g., applying constant-folding or dead-code elimination).
+\MP{
+	acronym? non acronym? 
+}
+Then, this paper defines compiler composition and formalises what kind of security properties does the composition of two compiler passes preserve, given what the individual passes preserve.
+The paper then defines several secure compiler passes, where each is either preserving a different security property (e.g., temporal or spatial memory safety) or performing a security-preserving optimisation, (e.g., applying constant-folding or dead-code elimination).
+\MP{
+	acro? non acro?
+}
 Finally, this paper shows how to compose these secure compiler passes into a multi-pass compilation chain which is proven to provide end-to-end preservation of general memory safety and cryptographic constant time.
@@
         %Concretely, it investigates upper and lower compositions, as well as the sequential composition.
-        For this, this paper studies compilers with multiple inputs or multiple outputs as well as the sequential composition of compilers, as typically done when engineering compilation chains.
-        %
-        This paper proves that starting from two compilers that preserve two properties, their composition preserves the intersection of those properties.
-        For example, a compiler that is secure with respect to \gls{ms} can be composed with a compiler that is secure with respect to \gls{cct}, resulting in a compiler that is secure with respect to the composition of \gls{ms} and \gls{cct}.
+        For this, this paper studies compilers with multiple inputs or multiple outputs as well as the sequential composition of compilers, as typically done in existing compilation chains.
+        %
+        This paper proves that starting from two compilers that preserve two (possibly distinct) properties, their composition preserves the intersection of those properties.
+        For example, a compiler that is secure with respect to \gls{tms} can be composed with a compiler that is secure with respect to \gls{sms}, resulting in a compiler that is secure with respect to \gls{ms}.
 %       \MPin{
@@
         Furthermore, the chain includes two passes for \gls{ms} that are optimising: One performs a simple \gls{dce} and the other \gls{cf}.
-        The formalisation of this case study showcases the power of the presented framework.
-        Namely, the divide-and-conquer approach to software engineering is a viable strategy even for the developement of secure compilers.
+        The formalisation of this case study showcases the power of the presented framework: Namely, the divide-and-conquer approach to software engineering is a viable strategy even for the development of secure compilers.
 %        This demonstrates that the usage of the presented framework is both simpler and more expressive compared to one monolithic security proof of the whole compilation chain.

### 2023-07-10 popl/main.tex
sec 2

@@
 A class is simply a set of properties and for the class of safety properties, it is decidable whether a trace satisfies a safety property with just a finite trace prefix.
-As an example, consider a trace describing an interaction with a memory where the deallocation of an address $\loc$ precedes a read at that address in memory: $\ev{Dealloc\ \loc}\cdot\ev{Read\ \loc\ 1729}\cdot\dots$
-This program behaviour is insecure with respect to a canonical notion of (temporal) memory safety dictating no use-after-frees of pointers~\cite{nagarakatte2010cets}, because it reads from a memory location that was freed already.
+As an example, consider a trace describing an interaction with a memory where the deallocation of an address $\loc$ precedes a read at that address in memory: $\ev{Dealloc\ \loc}\cdot\ev{Read\ \loc\ 1729}\cdot\dots$.
+This program behaviour is insecure with respect to a canonical notion of (temporal) memory safety dictating no use-after-frees of pointers~\cite{nagarakatte2010cets}\MP{should cite catalin's MS paper too}, because it reads from a memory location that was freed already.
 The previous finite trace prefix is enough to decide that the trace does not satisfy temporal memory safety and there is no way to append events to this prefix which would result in the trace being admissible.
-In the following, the notation of an execution of a whole program $\wholeprogvar$ that terminates in state $\runtimetermvar$ according to the language semantics and produces trace $\trace$ is $\progstepto{\wholeprogvar}{\runtimetermvar}{\trace}$.
-With this, the definition of property satisfaction is as follows:
+In the following, we indicate the execution of a whole program $\wholeprogvar$ that terminates in state $\runtimetermvar$ according to the language semantics and produces trace $\trace$ as: $\progstepto{\wholeprogvar}{\runtimetermvar}{\trace}$.
+With this, we define property satisfaction as follows:
 \bul{whole programs $\wholeprogvar$ satisfy a property $\pi$} iff \iul{$\wholeprogvar$ yields a trace $\trace$} such that \oul{$\trace$ satisifies $\pi$} (\Cref{def:propsat}).
@@
 Property satisfaction is defined on whole programs, i.e., programs without missing definitions.
-Thus, there is only a passive attacker model that observes the execution and, e.g., retreives secrets from that.
-To consider a stronger model similarly to what existing work does~\cite{abate2019jour,abate2021extacc,patrignani2021rsc}, the concept of satisfaction is extended with {\em robustness}.
-Robust satisfaction considers partial programs $\progvar$, i.e., programs with free variables, which need to be defined prior to execution.
-Programs with free variables can get stuck during execution.
+Thus, from a security perspective, this considers only a passive attacker model, where the attacker observes the execution and, e.g., retrieves secrets from that.
+To consider a stronger model similarly to what existing work does~\cite{abate2019jour,abate2021extacc,patrignani2021rsc}\MP{many more things on RS here}, we extend the concept of satisfaction with {\em robustness}.
+Robust satisfaction considers partial programs $\progvar$, i.e., components with missing imports, which cannot run until said imports are fulfilled.
 To remedy this, {\em linking} takes two partial programs $\progvar[_{1}],\progvar[_{2}]$ and produces a whole program $\wholeprogvar$, i.e., $\link{\progvar[_{1}]}{\progvar[_{2}]}=\wholeprogvar$.
-As typically done in not only secure compilation~\cite{abate2019jour,devriese2018parametricity}, %
-% \MP{others do this too, parametricity results e.g.}
+As typically done in works that consider the execution of partial programs~\cite{abate2019jour,devriese2018parametricity}, \MP{Add papers from Amal. Papers from just Dominique on FAC. RS works.}
 this paper assumes that whole programs are the result of linking partial programs referred to as {\em context} ($\ctx$) and {\em component} ($\comp$).
@@
 %In this work, the semantics of the programming language is expected to differentiate between the component and the context.
-With this, \Thmref{def:propsat} can be extended as follows: for \bul{components $\progvar$ to robustly satisfy a property $\pi$}, take an \iul{attacker context $\contextvar$ and link it with $\progvar$}, \oul{the result of which must satisfy $\pi$}.
+With this, \Thmref{def:propsat} can be extended as follows: for \bul{components $\progvar$ to robustly satisfy a property $\pi$}, take an \iul{attacker context $\contextvar$ and link it with $\progvar$}, \oul{the resulting whole program must satisfy $\pi$}.
 
@@
   Consider \texttt{CVE-2021-3564}~\cite{doublefree-bluetooth}, one of many submissions for a double-free vulnerability.
-  The bug arises due to a race condition where the context-level function \texttt{hci\_cmd\_work} was not expected to behave maliciously, since it resides in the same source-code repository where the vulnerability occurs.
+  The bug\MP{not bug, vulnerability.
+uniform below (perhaps?)} arises due to a race condition where the context-level function \texttt{hci\_cmd\_work} was not expected to behave maliciously, since it resides in the same source-code repository where the vulnerability occurs.
   Nevertheless, the component-level code of \texttt{hci\_dev\_do\_open} is linked with \texttt{hci\_cmd\_work} and does not atomically check whether the pointer has been freed already:
@@
 That is, for a whole program $\src{\wholeprogvar}$, the compiler should relate the $\src{L}$ semantics of $\src{\wholeprogvar}$ with the semantics of $\trg{T}$ in such a way that they are ,,compatible''.
-Unfortunately, correct compilers may be insecure compilers~\cite{patrignani2019survey}, and programs translated by insecure compilers can violate security properties that the programmer assumes to hold.
+Unfortunately, correct compilers may be insecure compilers~\cite{patrignani2019survey}\MP{cite kennedy and teh general SC dagstuhl and ACM papers too}, and programs translated by insecure compilers can violate security properties that the programmer assumes to hold.
 To define when a compiler is secure, this paper uses the robust compilation framework~\cite{abate2019jour}, which we summarize in the following definition.
@@
 For \bul{compilers $\cc{\src{L}}{\trg{L}}$ to robustly preserve a class of properties $\class$}, if for any \rul{property $\pi$ of that class $\class$ and programs $\src{p}$ written in $\src{L}$} we have that \iul{the program $\src{\progvar}$ robustly satisfies $\pi$}, then \oul{the compilation of $\src{\progvar}$, $\cc{\src{L}}{\trg{L}}\left(\src{p}\right)$, must robustly satisfy $\pi$}.
-
-\begin{definition}[Robust Preservation]\label{def:rtp}
+\MPin{
+	somewhere we should say that a class of props can also include a single prop
+}
+
+\begin{definition}[Robust Compilation]\label{def:rtp}
   % Compiler $\cc{\src{L}}{\trg{L}}$ robustly preserves $\class$, 
@@
   \end{center}
-  While the implementation the compiler emits may look correct, it does not check that the provided argument is of the right type.
+  While the compiler emits code that may look correct, the generated code does not check that the provided argument is of the right type.
   Even though the pointer $\trg{x}$ is checked for validity, the context $\trg{foo((int*)y, 1)}$ is able to provoke a read out of bounds.
-  Suppose the component transferred control to the context and passed ownership of a $\trg{char}$ pointer $\trg{y}$ sized $\trg{1}$ cells, the context now calls the component again but casts this buffer to an $\trg{int*}$ prior to that call.
+  Suppose the component transferred control to the context and passed ownership of a $\trg{char}$ pointer $\trg{y}$ sized $\trg{1}$ cells, the context can now call the component again, casting this buffer to an $\trg{int*}$ prior to that call.
   The pointer is valid for one $\trg{char}$-sized memory cell, as expected, but the actual read operation now returns $\trg{sizeof(int)}$ many bytes instead of just $\trg{sizeof(char)}$ many.
-  Thus, even if $\src{foo}$ may have been robust with respect to the no-reads-out-of-bounds property, its compiled counterpart is not and therefore the compiler fails to attain \Thmref{def:rtp}.
+  Thus, even if $\src{foo}$ may have been robust with respect to the no-reads-out-of-bounds property\MP{sms?}, its compiled counterpart is not and therefore the compiler fails to attain \Thmref{def:rtp}.
 \end{example}

### 2023-07-10 popl/main.tex
proof sketch

--- 2023-07-10 popl/main.tex
@@
 
-\subsubsection{Proof Sketch of Robust Temporal Memory Safety Preservation}
-
-The proof of \Cref{thm:cca:rtp:tms} uses standard techniques in secure compilation literature and is visualized in \Cref{fig:proofdiag:rtp}.
-\MPin{
-	explain how to read the figure
-}
-Concretely, given the execution of the compiled $\Ltms$ program linked with an arbitrary $\trg{L}$ context yields a trace $\trg{\trace}$, it uses a trace-based backtranslation of $\trg{\trace}$ to reconstruct an $\src{L_{\tmssafe}}$ context.
-\MPin{
-	what does this mean?
-	what is a backtranslation?
-}
+\subsubsection{Proof Sketch of Robust Safety Property Preservation}
+
+The proof of \Cref{thm:cca:rtp:tms} uses standard techniques in secure compilation literature.
+These techniques are also used for similar theorems in this paper.
+
+When proving robust safety property preservation for the concrete property \gls{tms}, a target-level execution of a compiled $\Ltms$ component $\src{\library_{\comp}}$ linked with an arbitrary context $\trg{\library_{\ctx}}$ is assumed, where the execution yields a trace $\trg{\trace}$.
+The proof goal is to show that $\trg{\trace}$ satisfies \gls{tms}.
+To get this satisfaction, one way of doing so is to use the other assumption in the proof context: $\rsat{\src{\library_{\comp}}}{\tmssafe}$.
+However, this requires a source-level execution for some concrete $\src{\library_{\ctx}}$ that yields a trace $\src{\trace}$ similar to $\trg{\trace}$.
+This paper uses the well-known trace-based backtranslation technique~\cite{abate2019jour} to construct such a context $\src{\library_{\ctx}}$.
+It is shown for that obtained context $\src{\library_{\ctx}}$ that the whole program $\link{\src{\library_{\ctx}}}{\src{\library_{comp}}}$ is well-typed.
+By \Cref{thm:wt:tms}, the fact that this program satisfies \gls{tms} is obtained and by typesafety, it is known that there must be some $\Ltms$ execution.
+The proof now needs to relate the two execution with eachother.
+\Cref{fig:proofdiag:rtp} visualizes this situation.
+At the bottom, there is the target-level execution between operational states $\trg{\emptyset},\trg{\Omega_{1}}$, and so on.
+The execution starts in the context $\trg{\library_{\ctx}}$ and is split by transitivity of $\trg{\estepto{}{}{}}$ into parts before, in, and after the component $\cc{\Ltms}{\trg{L}}(\src{p})$ executes\footnote{For sake of simplicity, this paper considers only one context-switch from context to component and back again.
+Recursive calls between context and component functions are disregarded.}.
+Initially, the empty operational states $\src{\emptyset}$ and $\trg{\emptyset}$ need to agree by means of a cross-language state relation $\multimap_{\emptyset}$.
+This relation is indexed by a function $\delta$ that maps $\trg{L}$-level locations $\trg{\loc}$ to $\Ltms$-level locations.
+The operational state of the executions up to the component $\src{p}$ need to agree via $\multimap_{\delta}$, both contexts need to behave in a similar way such that the traces also agree with each other.
+However, the relation allows $\trg{\library_{\ctx}}$ to perform internal actions that may change the operational state but are not visible on the trace, such as internal function calls, that does not $\src{\library_{\ctx}}$ need to reflect.
+Before and after the compiler wrapper code, the operational states must agree via another cross-language relation $\approx_{\delta_{w_{}}}$, which disallows these mismatches with respect to internal function calls.
+Intuitively, the behavior of the components should be tight: the source-level execution must be preserved by the compiler in the target-level execution.
+Dually, when the component is done, the relations take turns again and the states must be related with $\multimap_{\delta_{w_{2}}}$
+Finally, since both execution traces are similar in the sense that their projections $\msfilterLtms{\src{\trace}}$ and $\msfilterL{\trg{\trace}}$ yield specification traces that satisfy \Cref{def:monsat}, the claim follows \Cref{lem:mon:tmsafe}.
+
 \MPin{
@@
 	they are technical, so we only keep them here.
-}
-\Cref{thm:wt:tms} already gives strong guarantees for programs written in $\src{L_{\tmssafe}}$ and can be exploited for the backtranslated context by ensuring that it typechecks.
-Well-typing\MP{well-typedness} means the $\src{L_{\tmssafe}}$ program runs, so the rest of the proof is about relating the $\trg{L}$-level execution with the $\src{L_{\tmssafe}}$-level execution.
-
-For the $\trg{L}$-execution, the compiler wrapper\MP{undefined terminology.
-call it like this before} code at state $\trg{\Omega_{1}}$ may either crash the program or continue execution $\trg{\Omega_{w_{1}}}$ right before control is transferred to the component.
-In this specific case, the compiler needs not insert additional code after the component is done with its execution, but a symmetric situation can happen here.\MP{vague and imprecise}
-Similarly, the backtranslation may also insert wrapper code prior and after switching to and from the component to the context. \MP{wut? why is a wrapper needed?}
-\MKin{
-  more detail or not? ;
-  i.e., explain $\multimap$ and $\approx$?
-  \MPin{
-  yes
-  }
-}
-\MPin{
-	the general idea that later proofs use the same structure is also not here.
+
+  MK: but, we don't actually rewrite it to RSC when doing the proof! The compilers are not robustly preserving with respect to an arbitrary safety property, but, e.g., just for \gls{tms}!
 }

### 2023-07-10 popl/main.tex
proof sketch

@@
 
-The proof of \Cref{thm:cca:rtp:tms} uses standard techniques in secure compilation literature.
-These techniques are also used for similar theorems in this paper.
+The proof of \Cref{thm:cca:rtp:tms} uses standard techniques in secure compilation literature that are also used for similar theorems in this paper.
 

### 2023-07-10 popl/main.tex

--- 2023-07-10 popl/main.tex
@@
 
-\subsubsection{Proof Sketch of Robust Temporal Memory Safety Preservation}
-
-The proof of \Cref{thm:cca:rtp:tms} uses standard techniques in secure compilation literature and is visualized in \Cref{fig:proofdiag:rtp}.
-\MPin{
-	explain how to read the figure
-}
-Concretely, given the execution of the compiled $\Ltms$ program linked with an arbitrary $\trg{L}$ context yields a trace $\trg{\trace}$, it uses a trace-based backtranslation of $\trg{\trace}$ to reconstruct an $\src{L_{\tmssafe}}$ context.
-\MPin{
-	what does this mean?
-	what is a backtranslation?
-}
+\subsubsection{Proof Sketch of Robust Safety Property Preservation}
+
+The proof of \Cref{thm:cca:rtp:tms} uses standard techniques in secure compilation literature that are also used for similar theorems in this paper.
+
+When proving robust safety property preservation for the concrete property \gls{tms}, a target-level execution of a compiled $\Ltms$ component $\src{\library_{\comp}}$ linked with an arbitrary context $\trg{\library_{\ctx}}$ is assumed, where the execution yields a trace $\trg{\trace}$.
+The proof goal is to show that $\trg{\trace}$ satisfies \gls{tms}.
+To get this satisfaction, one way of doing so is to use the other assumption in the proof context: $\rsat{\src{\library_{\comp}}}{\tmssafe}$.
+However, this requires a source-level execution for some concrete $\src{\library_{\ctx}}$ that yields a trace $\src{\trace}$ similar to $\trg{\trace}$.
+This paper uses the well-known trace-based backtranslation technique~\cite{abate2019jour} to construct such a context $\src{\library_{\ctx}}$.
+It is shown for that obtained context $\src{\library_{\ctx}}$ that the whole program $\link{\src{\library_{\ctx}}}{\src{\library_{comp}}}$ is well-typed.
+By \Cref{thm:wt:tms}, the fact that this program satisfies \gls{tms} is obtained and by typesafety, it is known that there must be some $\Ltms$ execution.
+The proof now needs to relate the two execution with eachother.
+\Cref{fig:proofdiag:rtp} visualizes this situation.
+At the bottom, there is the target-level execution between operational states $\trg{\emptyset},\trg{\Omega_{1}}$, and so on.
+The execution starts in the context $\trg{\library_{\ctx}}$ and is split by transitivity of $\trg{\estepto{}{}{}}$ into parts before, in, and after the component $\cc{\Ltms}{\trg{L}}(\src{p})$ executes\footnote{For sake of simplicity, this paper considers only one context-switch from context to component and back again.
+Recursive calls between context and component functions are disregarded.}.
+Initially, the empty operational states $\src{\emptyset}$ and $\trg{\emptyset}$ need to agree by means of a cross-language state relation $\multimap_{\emptyset}$.
+This relation is indexed by a function $\delta$ that maps $\trg{L}$-level locations $\trg{\loc}$ to $\Ltms$-level locations.
+The operational state of the executions up to the component $\src{p}$ need to agree via $\multimap_{\delta}$, both contexts need to behave in a similar way such that the traces also agree with each other.
+However, the relation allows $\trg{\library_{\ctx}}$ to perform internal actions that may change the operational state but are not visible on the trace, such as internal function calls, that does not $\src{\library_{\ctx}}$ need to reflect.
+Before and after the compiler wrapper code, the operational states must agree via another cross-language relation $\approx_{\delta_{w_{}}}$, which disallows these mismatches with respect to internal function calls.
+Intuitively, the behavior of the components should be tight: the source-level execution must be preserved by the compiler in the target-level execution.
+Dually, when the component is done, the relations take turns again and the states must be related with $\multimap_{\delta_{w_{2}}}$
+Finally, since both execution traces are similar in the sense that their projections $\msfilterLtms{\src{\trace}}$ and $\msfilterL{\trg{\trace}}$ yield specification traces that satisfy \Cref{def:monsat}, the claim follows \Cref{lem:mon:tmsafe}.
+
 \MPin{

### 2023-07-10 popl/main.tex
intro

@@
 An easy fix to sidestep this is to use non-nullable pointers as parameters instead, e.g., as presented in Cyclone~\cite{jim2002cyclone}.
-With this, a secure compiler needs to ensure that no malicious target-level context can supply unitialised memory and thereby sidestep the guarantees of the source language, e.g., by inserting appropriate checks prior to entering the body of \texttt{strncpy}.
-\MPin{
-	do we have examples of such wrapper-chesk inserting secure compilers?
-	mswasm compiler ... not really.
-	or we could say that another solution is  using capabilities, and then the mswasm compiler can be used
-}
+With this, a secure compiler needs to ensure that no malicious target-level context can supply unitialised memory and thereby sidestep the guarantees of the source language, e.g., by inserting appropriate checks prior to entering the body of \texttt{strncpy} or using capabilities~\cite{michael2023mswasm}.
 \end{example}
@@
 This paper introduces a framework for reasoning about the composition of secure and optimising compiler passes akin to those of \Cref{ex:strncpy} and it showcases the power of this framework by instantiating it on a multi-pass compilation chain.
-To this end, this paper first discusses how to compose security properties, such as temporal and spatial memory safety into general memory safety, and cryptographic constant time.
-\MP{
-	acronym? non acronym? 
-}
+To this end, this paper first discusses how to compose security properties, such as \gls{tms} and \gls{sms} into \gls{ms}, and \gls{cct}.
 Then, this paper defines compiler composition and formalises what kind of security properties does the composition of two compiler passes preserve, given what the individual passes preserve.
-The paper then defines several secure compiler passes, where each is either preserving a different security property (e.g., temporal or spatial memory safety) or performing a security-preserving optimisation, (e.g., applying constant-folding or dead-code elimination).
-\MP{
-	acro? non acro?
-}
+The paper then defines several secure compiler passes, where each is either preserving a different security property (e.g., temporal or spatial memory safety) or performing a security-preserving optimisation, (e.g., applying \gls{cf} or \gls{dce}).
 Finally, this paper shows how to compose these secure compiler passes into a multi-pass compilation chain which is proven to provide end-to-end preservation of general memory safety and cryptographic constant time.

### 2023-07-10 popl/main.tex
sec 3

--- 2023-07-10 popl/main.tex
@@
 Then, this section introduces monitors for each of the properties presented earlier (\Cref{subsec:monitors}).
+\MP{why?}
 
@@
 
-\gls{sms} disregards out of bounds accesses:
+\gls{sms} disregards\MP{prohibits?} out-of-bounds accesses:
 
@@
 Full \gls{ms} is then described as the conjunction of \Cref{def:trace:tmsdef,def:trace:smsdef}.
-Note, however, that this definition does not entail memory-safety issues introduced by side-channels, such as speculation.
+Note, however, that this definition says nothing about memory-safety issues introduced by side-channels, such as speculation.
 
@@
 \end{definition}
+\MPin{
+	cite catalin's paper and the mswasm paper for the def of ms
+}
 
@@
 
-\gls{cct} is a hyperproperty and, thus, not as easy to check with just monitors. %Beutner et al
-Hence, the following defines \gls{scct}, a stricter variant that enforces the policy that no secret should appear on a trace.
+\gls{cct} is a hyperproperty\MP{hypersafety?}\MP{what is a hp?} and, thus, not as easy to check with just monitors.\MP{chy? cit?} %Beutner et al
+Hence, we define \gls{scct}, a stricter variant that enforces the policy that no secret should appear on a trace.\MP{cit? did we take some inspiration}
 
@@
 \end{definition}
+\MPin{
+	comment on the composition at this levle?
+}
 
@@
 Monitors can be used to enforce safety properties and this section presents concrete instantiations of monitors for the previousvy presented properties (\Cref{subsec:propdefs}).
-To lessen the burden when proving, each monitor makes use of a custom trace model that contains only the relevant information related to the property the monitor checks.
+\MPin{
+	why
+}
+To lessen the burden when proving\MP{proving what}, each monitor makes use\MP{uses} of a custom trace model that contains only the relevant information related to the property the monitor checks.
 To project specification traces $\trace$ onto these monitor-level traces $\abstrace$, each property has an associated event agreement relation $\traceagree[]{\pi}{\trace}{\abstrace}$.
-This relation performs the actual work of projecting the events, while the stutter-equivalence $\traceagree{\pi}{\trace}{\abstrace}$ remains the same for all properties $\pi$, yet is indexed by $\pi$ to emphasize the kind of event agreement.
-With agreements, this section starts to define monitor satisfaction for traces and states that monitor satisfaction implies property satisfaction.
+This relation projects the events\MP{whaddaya mean?}, while the stutter-equivalence $\traceagree{\pi}{\trace}{\abstrace}$ remains the same for all properties $\pi$, yet is indexed by $\pi$ to emphasize the kind of event agreement.\MP{unclear sentence}
+With agreements, this section defines monitor satisfaction for traces and then it proves that monitor satisfaction implies property satisfaction.
 To this end, monitor satisfaction is defined as follows.
@@
 
-The state of the monitor is a record with two sets keeping track of allocated and deallocated locations.
+For \gls{tms}, the state of the monitor is a record with two sets keeping track of allocated and deallocated locations.
 %The monitor runs on a modified trace semantics with more abstract $\absev{Alloc\ }\loc$,$\absev{Dealloc\ }\loc$,$\absev{Use\ }\loc$, and $\absterm$ events, which represent {\em allocation}, {\em deallocation}, and {\em use} of a location $\loc$ as well as abnormal program termination.
@@
 However, the definition can easily be adapted by, e.g., attaching a natural number serving as a counter.
-In the reflexive case of $\monitorcheck{\tmsmonitor}{\tmsmonitor}{\absevent}$, the monitor $\tmsmonitor$ must not contain allocated locations, i.e., $\tmsmonitor.allocated=\emptyset$.
+In the reflexive case of $\monitorcheck{\tmsmonitor}{\tmsmonitor}{\absevent}$, the monitor $\tmsmonitor$ must not contain allocated locations, i.e., $\tmsmonitor.allocated=\emptyset$.\MP{wut?}
 This ensures that programs do not forget to free memory.
@@
 \end{example}
+\MPin{
+	example is good, but it'd be nice to tie it to the running example?
+}
 
@@
   At that stage, the induction variable \texttt{i} is equal to \texttt{2} and, unfortunately, the order of checks is such that first the cell \texttt{x[i]} is read prior to bounds checking \texttt{i < n}.
-  Because of this, there is an out of bounds memory access right before exiting the function.
+  Because of this, there is an out of bounds\MP{hyphenate: out-of-bounds} memory access right before exiting the function.
   This is also visible on the trace, which can be sketched as $\ldots\cdot\ev{Alloc}\ \loc_{x}\ 2;\comp;\unlock\cdot\ldots\cdot\ev{Alloc}\ \loc_{y}\ 2;\comp;\unlock\cdot\ldots\cdot\ev{Use}\ \loc_{x}\ 0;\comp;\unlock\cdot\ev{Use}\ \loc_{y}\ 0;\comp;\unlock\cdot\ev{Use}\ \loc_{x}\ 1;\comp;\unlock\cdot\ev{Use}\ \loc_{y}\ 1;\comp;\unlock\cdot\ev{Use}\ \loc_{x}\ 2;\comp;\unlock\cdot\ldots$, where $\loc_{x}$ and $\loc_{y}$ are the memory addresses associated to \texttt{x} and \texttt{y}, respectively.
   Omitting the events for all ,,$\ldots$'' for sake of brevity, the abstract monitor trace of this is $\absev{Alloc}\ \loc_{x}\ \comp\ 2\cdot\absev{Alloc}\ \loc_{y}\ \comp\ 2\cdot\absev{Use}\ \loc_{x}\ \comp\ 0\cdot\absev{Use}\ \loc_{y}\ \comp\ 0\cdot\absev{Use}\ \loc_{x}\ \comp\ 1\cdot\absev{Use}\ \loc_{y}\ \comp\ 1\cdot\absev{Use}\ \loc_{x}\ \comp\ 2$.
-  It is easy to see that, after the allocation events, the state of the monitor is $\left\{(\loc_{x};\comp;2),(\loc_{y};\comp;2)\right\}$.
+
+  It is easy to see that,\MP{erase. we don't know what's easy} after the allocation events, the state of the monitor is $\left\{(\loc_{x};\comp;2),(\loc_{y};\comp;2)\right\}$.
   Furthermore, all uses up to the last are acceptable by the monitor, but the last event does not satisfy the premise $2<2$ in \Cref{tr:sms-use}.
@@
 
-\begin{example}[Data independent timing mode]
+\begin{example}[Data-independent timing mode]
   Consider an excerpt from a component that reads from non-confidential memory \texttt{x} and writes the result to confidential memory \texttt{y}.
@@
 Now, given \bul{$\cc{\src{L}}{\trg{L}}$ robustly preserves $\class[_{1}]$} and \rul{$\cc{\trg{L}}{\obj{L}}$ robustly preserves $\class[_{2}]$}, it follows that \oul{their sequential composition $\cc{\src{L}}{\trg{L}}\circ\cc{\trg{L}}{\obj{L}}$ robustly preserves the intersection of classes $\class[_{1}]$ and $\class[_{2}]$}.
+\MPin{
+	section 3.1 never defined the $\cap$ operator on properties, nor on classes.
+	it uses the 'and'.
+	rectify
+}
 
@@
 Notably, real-world compilation chains also perform a series of (sequential) passes whose main purpose is not necessarily to translate from one language to another, but to, e.g., optimise the code or enforce a certain property.
-Both examples can be seen in practice~\cite{nagarakatte2009soft,nagarakatte2010cets,akritidis2009baggy,wegman1991ccp,manjikian1997fusion}, it is infeasible to list them all\MPin{survey to cite?
+Both examples can be seen in practice, e.g. as in the work of~\cite{nagarakatte2009soft,nagarakatte2010cets,akritidis2009baggy,wegman1991ccp,manjikian1997fusion} and many more.
+% \MPin{survey to cite?
 this is a bit informal MK: hard to find a survey on different compiler passes...
@@
 Unfortunately, the order in which \gls{cf} and \gls{dce} are performed influences the final result of the compilation (see \Cref{fig:cfdceex}).
-\begin{figure}
+\begin{figure}[!ht]
   \begin{center}
@@
 
-\subsubsection{Proof Sketch of Robust Safety Property Preservation}
-
-The proof of \Cref{thm:cca:rtp:tms} uses standard techniques in secure compilation literature that are also used for similar theorems in this paper.
-
-When proving robust safety property preservation for the concrete property \gls{tms}, a target-level execution of a compiled $\Ltms$ component $\src{\library_{\comp}}$ linked with an arbitrary context $\trg{\library_{\ctx}}$ is assumed, where the execution yields a trace $\trg{\trace}$.
-The proof goal is to show that $\trg{\trace}$ satisfies \gls{tms}.
-To get this satisfaction, one way of doing so is to use the other assumption in the proof context: $\rsat{\src{\library_{\comp}}}{\tmssafe}$.
-However, this requires a source-level execution for some concrete $\src{\library_{\ctx}}$ that yields a trace $\src{\trace}$ similar to $\trg{\trace}$.
-This paper uses the well-known trace-based backtranslation technique~\cite{abate2019jour} to construct such a context $\src{\library_{\ctx}}$.
-It is shown for that obtained context $\src{\library_{\ctx}}$ that the whole program $\link{\src{\library_{\ctx}}}{\src{\library_{comp}}}$ is well-typed.
-By \Cref{thm:wt:tms}, the fact that this program satisfies \gls{tms} is obtained and by typesafety, it is known that there must be some $\Ltms$ execution.
-The proof now needs to relate the two execution with eachother.
-\Cref{fig:proofdiag:rtp} visualizes this situation.
-At the bottom, there is the target-level execution between operational states $\trg{\emptyset},\trg{\Omega_{1}}$, and so on.
-The execution starts in the context $\trg{\library_{\ctx}}$ and is split by transitivity of $\trg{\estepto{}{}{}}$ into parts before, in, and after the component $\cc{\Ltms}{\trg{L}}(\src{p})$ executes\footnote{For sake of simplicity, this paper considers only one context-switch from context to component and back again.
-Recursive calls between context and component functions are disregarded.}.
-Initially, the empty operational states $\src{\emptyset}$ and $\trg{\emptyset}$ need to agree by means of a cross-language state relation $\multimap_{\emptyset}$.
-This relation is indexed by a function $\delta$ that maps $\trg{L}$-level locations $\trg{\loc}$ to $\Ltms$-level locations.
-The operational state of the executions up to the component $\src{p}$ need to agree via $\multimap_{\delta}$, both contexts need to behave in a similar way such that the traces also agree with each other.
-However, the relation allows $\trg{\library_{\ctx}}$ to perform internal actions that may change the operational state but are not visible on the trace, such as internal function calls, that does not $\src{\library_{\ctx}}$ need to reflect.
-Before and after the compiler wrapper code, the operational states must agree via another cross-language relation $\approx_{\delta_{w_{}}}$, which disallows these mismatches with respect to internal function calls.
-Intuitively, the behavior of the components should be tight: the source-level execution must be preserved by the compiler in the target-level execution.
-Dually, when the component is done, the relations take turns again and the states must be related with $\multimap_{\delta_{w_{2}}}$
-Finally, since both execution traces are similar in the sense that their projections $\msfilterLtms{\src{\trace}}$ and $\msfilterL{\trg{\trace}}$ yield specification traces that satisfy \Cref{def:monsat}, the claim follows \Cref{lem:mon:tmsafe}.
-
+\subsubsection{Proof Sketch of Robust Temporal Memory Safety Preservation}
+
+The proof of \Cref{thm:cca:rtp:tms} uses standard techniques in secure compilation literature and is visualized in \Cref{fig:proofdiag:rtp}.
+\MPin{
+	explain how to read the figure
+}
+Concretely, given the execution of the compiled $\Ltms$ program linked with an arbitrary $\trg{L}$ context yields a trace $\trg{\trace}$, it uses a trace-based backtranslation of $\trg{\trace}$ to reconstruct an $\src{L_{\tmssafe}}$ context.
+\MPin{
+	what does this mean?
+	what is a backtranslation?
+}
 \MPin{
@@
 	they are technical, so we only keep them here.
-
-  MK: but, we don't actually rewrite it to RSC when doing the proof! The compilers are not robustly preserving with respect to an arbitrary safety property, but, e.g., just for \gls{tms}!
+}
+\Cref{thm:wt:tms} already gives strong guarantees for programs written in $\src{L_{\tmssafe}}$ and can be exploited for the backtranslated context by ensuring that it typechecks.
+Well-typing\MP{well-typedness} means the $\src{L_{\tmssafe}}$ program runs, so the rest of the proof is about relating the $\trg{L}$-level execution with the $\src{L_{\tmssafe}}$-level execution.
+
+For the $\trg{L}$-execution, the compiler wrapper\MP{undefined terminology.
+call it like this before} code at state $\trg{\Omega_{1}}$ may either crash the program or continue execution $\trg{\Omega_{w_{1}}}$ right before control is transferred to the component.
+In this specific case, the compiler needs not insert additional code after the component is done with its execution, but a symmetric situation can happen here.\MP{vague and imprecise}
+Similarly, the backtranslation may also insert wrapper code prior and after switching to and from the component to the context. \MP{wut? why is a wrapper needed?}
+\MKin{
+  more detail or not? ;
+  i.e., explain $\multimap$ and $\approx$?
+  \MPin{
+  yes
+  }
+}
+\MPin{
+	the general idea that later proofs use the same structure is also not here.
 }

### 2023-07-10 popl/main.tex
sec 4

--- 2023-07-10 popl/main.tex
@@
 I can't even find one just for compiler enforced memsafety (that looks good) or for compiler translations in general}.
-Consider the compilation passes of LLVM~\cite{lattner2004llvm}.
-In LLVM, compiler passes are handled in the \texttt{PassManager} class.
-By means of manipulating the internal state of the \texttt{PassManager} object, the schedule of compiler optimisation passes can be hardcoded.
-That way, an order of optimisations can be found by hand that yields the best results.
-Take \gls{cf}, which rewrites constant expressions to the constant they evaluate to, and \gls{dce}, which removes dead code by rewriting conditional branches.
-Unfortunately, the order in which \gls{cf} and \gls{dce} are performed influences the final result of the compilation (see \Cref{fig:cfdceex}).
+% 
+% Consider the compilation passes of LLVM~\cite{lattner2004llvm}, which are handled in the \texttt{PassManager} class.
+% By manipulating the internal state of a \texttt{PassManager} object, the schedule of compiler optimisation passes can be hardcoded.
+% That way, an order of optimisations can be found by hand that yields the best results.
+% \MP{what's the point of the previous 3 sentences?(commented)}
+Consider the following two LLVM optimisation passes: \gls{cf}, which rewrites constant expressions to the constant they evaluate to, and \gls{dce}, which removes dead code by rewriting conditional branches.
+The order in which \gls{cf} and \gls{dce} are performed influences the final result of the compilation (see \Cref{fig:cfdceex}).
 \begin{figure}[!ht]
@@
 The definition of upper composition is as follows: given a program $\mathtt{p}$, its compiled $\trg{L}$ counterpart is obtained by plugging $\mathtt{p}$ into $\cc{\src{L}}{\trg{L}}$ if $\mathtt{p}\in\src{L}$ or by plugging $\mathtt{p}$ into $\cc{\obj{L}}{\trg{L}}$ if $\mathtt{p}\in\obj{L}$, written as $\cc{\src{L}+\obj{L}}{\trg{L}}$.
+\MP{
+	put into def env
+}
 Now, given \bul{$\cc{\src{L}}{\trg{L}}$ robustly preserves $\class[_{1}]$} and \rul{$\cc{\obj{L}}{\trg{L}}$ robustly preserves $\class[_{2}]$}, it follows that \oul{their upper composition $\cc{\src{L}+\obj{L}}{\trg{L}}$ robustly preserves the intersection of classes $\class[_{1}]$ and $\class[_{2}]$}.
@@
 definition of lower composition $\cc{\src{L}}{\trg{L}+\obj{L}}$ is as follows: given a program $\src{p}$, its compiled $\trg{L}$ or $\obj{L}$ program is obtained by performing an internal decision and threading $\src{p}$ into $\cc{\src{L}}{\trg{L}}$ or into $\cc{\src{L}}{\obj{L}}$ depending on that decision.
+\MP{into def}
 Now, given \bul{$\cc{\src{L}}{\trg{L}}$ robustly preserves $\class[_{1}]$} and \rul{$\cc{\src{L}}{\obj{L}}$ robustly preserves $\class[_{2}]$}, it follows that \oul{their lower composition $\cc{\src{L}}{\trg{L}+\obj{L}}$ robustly preserves the intersection of classes $\class[_{1}]$ and $\class[_{2}]$}.

### 2023-07-10 popl/main.tex

@@
 To this end, this section defines the languages $\Ltms$, $\Ltrg$, $\Lms$, and $\Lscct$ which share many common elements (presented in \Cref{subsec:cs:defs}).
-In this paper, $\Ltms$ is the only statically typed language and exhibits the property that all well-typed programs are \gls{tms} (\Cref{subsec:ltms}).
+$\Ltms$ is the only statically typed language and exhibits the property that all well-typed programs are \gls{tms} (\Cref{subsec:ltms}).
 However, not all $\Ltms$ programs are \gls{sms}.
 That is, there are well-typed $\Ltms$ programs that perform an out-of-bounds access.
-In contrast, $\Ltrg$ is untyped and does not provide any guarantees with regards to \gls{ms} (\Cref{subsec:lsms}).
+Language $\Ltrg$ is untyped and does not provide any guarantees with regards to \gls{ms} (\Cref{subsec:lsms}).
 $\Lms$ is exactly the same language as $\Ltrg$, but this paper still distinguishes the two for sake of readability (\Cref{subsec:lms}).
@@
 
-This paper considers \gls{cct} as an architecture-dependent property that should not be one of the programmer's concerns~\cite{cauligi2019fact}.
+This paper considers \gls{cct} as an architecture-dependent property writing \gls{cct} should not be one of the programmer's concerns~\cite{cauligi2019fact}.
 Such consideration is backed up by architecture providing a data (operand) independent timing mode, such as processors by Arm~\cite[p.~543]{arm-refman} and Intel~\cite[p.~80]{intel-refman}.
-In spirit of this, the language $\Lscct$ allows violating \gls{cct} by emitting events on, e.g., branching and division, that contain secrets (\Cref{subsec:lscct}), but provides a way to read and write to a {\em model--specific register} that enables a ``\gls{cct}-mode''.
+In spirit of this, language $\Lscct$ allows violating \gls{cct} by emitting events on, e.g., branching and division, that contain secrets (\Cref{subsec:lscct}), but provides a way to read and write to a {\em model--specific register} that enables a ``\gls{cct}-mode''.
 
@@
 $\src{L_{\tmssafe}}$ uses the same syntax as presented earlier (\Cref{sec:casestud:defs}) without extensions to the term level.
-But, $\src{L_{\tmssafe}}$ is statically typed, where the type system is inspired by $L^{3}$~\cite{morrisett2005L3}.
-The type system of $\src{L_{\tmssafe}}$ exhibits the property that every well-typed $\src{L_{\tmssafe}}$ program satisfies \gls{tms}.
-The proof of this property relies on a projection $\msfilterLtms{\src{\event}}=\event$ that projects $\src{L_{\tmssafe}}$ events to specfication events $\event$, because the properties defined earlier (\Cref{subsec:propdefs}) are defined in the specification trace model.
+But, $\src{L_{\tmssafe}}$ is statically typed, where the type system is inspired by $L^{3}$~\cite{morrisett2005L3}.\MP{fabulous paper from amal?relevant?}
+\MP{and therefore omitted?}
+The type system of $\src{L_{\tmssafe}}$ exhibits the property that every well-typed $\src{L_{\tmssafe}}$ program satisfies \gls{tms} (\Cref{thm:wt:tms}).
+The proof of this theorem relies on a projection $\msfilterLtms{\src{\event}}=\event$ from $\src{L_{\tmssafe}}$ events to specification events $\event$, because the properties defined earlier (\Cref{subsec:propdefs}) are defined in the specification trace model.
 Hereby, $\delta$ maps $\src{L_{\tmssafe}}$ locations $\src{\loc}$ to locations $\loc$ of the specification trace model whenever an event with a location is projected.
@@
 But, it also filters any action that a context does as well as the interaction events, since these are irrelevant for component-level \gls{tms}.
+\MPin{ok let's add details}
 
@@
 The projection $\msfilterL{\trg{\trace}}$ is entirely similar to $\msfilterLtms{\src{\trace}}$.
+\MP{ok, details}
 
@@
 
-While the previous languages (\Cref{subsec:ltms,subsec:lsms,subsec:lms}) are not concerned with \gls{cct}, $\Lscct$ extends $\Lms$ (\Cref{subsec:lms}) with a register flag for a data (operand) independent timing mode\footnote{As present in both Arm~\cite[p.~543]{arm-refman} and Intel~\cite[p.~80]{intel-refman} processors.}.
+While the previous languages (\Cref{subsec:ltms,subsec:lsms,subsec:lms}) are not concerned with \gls{cct}, $\Lscct$ extends $\Lms$ (\Cref{subsec:lms}) with a register flag for a data (operand) independent timing mode, a feature that is present in both Arm~\cite[p.~543]{arm-refman} and Intel~\cite[p.~80]{intel-refman} processors.
 For the other languages seen earlier, this mode is intuitively always-on.
@@
 \Cref{ex:lscct} illustrates these changes.
+\MPin{formalisation additions in algin* env}
 
@@
 The pre-events themselves still translate to just $\ev{Get}$ and $\ev{Set}$, respectively.
-With this technical setup, the information wether a read or write happened on a secret value is not hidden by the semantics, e.g., by emitting $\obj{\emptyevent}$, but when projecting to specification events.
+With this technical setup, the information wether\MP{spelling} a read or write happened on a secret value is not hidden by the semantics, e.g., by emitting $\obj{\emptyevent}$, but when projecting to specification events.
 This allows flexibility: The trace can be checked to satisfy different properties, such as, in this case, \gls{tms}, \gls{sms}, \gls{scct}, and their combined versions.
@@
   Moreover, writing to memory now yields the pre-event $\obj{\widehat{Set}}$, whose difference to $\obj{Set}$ becomes apparent when looking at the projection to the specification trace model.
+  \MPin{shouldn't we have 2 traces, one with the bit set, one without}
   Projecting the above trace to the specification trace model yields the following trace (emitting empty events):
@@
 \section{Case Study: Composing Secure Compiler Passes and Optimisations \pages{7}}\label{sec:casestud:rtp}
+This section defines several secure compilers, each of which robustly preserves a different property of interest as depicted in \Cref{fig:pipeline}.
 \begin{figure}[!h]
@@
 \end{figure}
-
-This section defines several secure compilers, each of which robustly preserves a different property of interest as depicted in \Cref{fig:pipeline}.
 The section demonstrates the power of the framework (\Cref{sec:compprop,sec:compcomp}) by composing these compilers for a secure and optimising compilation chain that robustly preserves \gls{msscct}.
@@
 \subsection{Robust Temporal Memory Safety Preservation}\label{subsec:cs:tms}
+\MPin{some text should introduce the compiler defs}
+
   \begin{align*}
@@
 
-The compiler only inserts a dynamic typecheck to respect the well-typedness of function bodies.
+The first compiler only inserts a dynamic typecheck to respect the well-typedness of function bodies.
 Since $\trg{L}$ has no static typechecks, it could happen that a bogus context $\trg{\library_{\ctx}}$ invokes a callable object accepting a $\src{\natt}$ with $\trg{\lpair{17}{29}}$.
@@
 
-Compiling the \texttt{strncpy} function from \Cref{sec:introduction} with $\cca$, the compiler would in this case ensure that the arguments are valid.
+Compiling the \texttt{strncpy} function from \Cref{sec:introduction} with $\cca$, the compiler would in this case ensure that the arguments that are evaluated in the compiled \texttt{strncpy} are valid.
 
@@
 
-\subsubsection{Proof Sketch of Robust Safety Property Preservation}
-
-The proof of \Cref{thm:cca:rtp:tms} uses standard techniques in secure compilation literature that are also used for similar theorems in this paper.
+\subsubsection{Proving Robust Safety Property Preservation}
+
+The proof of \Cref{thm:cca:rtp:tms} uses standard techniques in secure compilation literature that are also used for the following theorems of this section.
+\MP{so what?}
 
 When proving robust safety property preservation for the concrete property \gls{tms}, a target-level execution of a compiled $\Ltms$ component $\src{\library_{\comp}}$ linked with an arbitrary context $\trg{\library_{\ctx}}$ is assumed, where the execution yields a trace $\trg{\trace}$.
+\MP{passive + not justified}
 The proof goal is to show that $\trg{\trace}$ satisfies \gls{tms}.
+\MP{no, the goal is to BT}
 To get this satisfaction, one way of doing so is to use the other assumption in the proof context: $\rsat{\src{\library_{\comp}}}{\tmssafe}$.
@@
 This paper uses the well-known trace-based backtranslation technique~\cite{abate2019jour} to construct such a context $\src{\library_{\ctx}}$.
-It is shown for that obtained context $\src{\library_{\ctx}}$ that the whole program $\link{\src{\library_{\ctx}}}{\src{\library_{comp}}}$ is well-typed.
-By \Cref{thm:wt:tms}, the fact that this program satisfies \gls{tms} is obtained and by typesafety, it is known that there must be some $\Ltms$ execution.
-The proof now needs to relate the two execution with eachother.
-\Cref{fig:proofdiag:rtp} visualizes this situation.
-At the bottom, there is the target-level execution between operational states $\trg{\emptyset},\trg{\Omega_{1}}$, and so on.
-The execution starts in the context $\trg{\library_{\ctx}}$ and is split by transitivity of $\trg{\estepto{}{}{}}$ into parts before, in, and after the component $\cc{\Ltms}{\trg{L}}(\src{p})$ executes\footnote{For sake of simplicity, this paper considers only one context-switch from context to component and back again.
-Recursive calls between context and component functions are disregarded.}.
-Initially, the empty operational states $\src{\emptyset}$ and $\trg{\emptyset}$ need to agree by means of a cross-language state relation $\multimap_{\emptyset}$.
-This relation is indexed by a function $\delta$ that maps $\trg{L}$-level locations $\trg{\loc}$ to $\Ltms$-level locations.
-The operational state of the executions up to the component $\src{p}$ need to agree via $\multimap_{\delta}$, both contexts need to behave in a similar way such that the traces also agree with each other.
-However, the relation allows $\trg{\library_{\ctx}}$ to perform internal actions that may change the operational state but are not visible on the trace, such as internal function calls, that does not $\src{\library_{\ctx}}$ need to reflect.
-Before and after the compiler wrapper code, the operational states must agree via another cross-language relation $\approx_{\delta_{w_{}}}$, which disallows these mismatches with respect to internal function calls.
-Intuitively, the behavior of the components should be tight: the source-level execution must be preserved by the compiler in the target-level execution.
-Dually, when the component is done, the relations take turns again and the states must be related with $\multimap_{\delta_{w_{2}}}$
-Finally, since both execution traces are similar in the sense that their projections $\msfilterLtms{\src{\trace}}$ and $\msfilterL{\trg{\trace}}$ yield specification traces that satisfy \Cref{def:monsat}, the claim follows \Cref{lem:mon:tmsafe}.
-
-\MPin{
-	why can we use a BT?
-	it's because the def in section 2.2 is eq to anotehr one, because the props are Safety, we use the RSC equivalent def.
-	all these intuitions are missing.
-	they are technical, so we only keep them here.
-
-  MK: but, we don't actually rewrite it to RSC when doing the proof! The compilers are not robustly preserving with respect to an arbitrary safety property, but, e.g., just for \gls{tms}!
-}
-
-\begin{figure}[h]
+It is shown for that obtained context $\src{\library_{\ctx}}$ that the whole program $\link{\src{\library_{\ctx}}}{\src{\library_{comp}}}$ is well-typed.\MP{passive}
+By \Cref{thm:wt:tms}, the fact that this program satisfies \gls{tms} is obtained and by typesafety, it is known that there must be some $\Ltms$ execution.\MP{passive passive passive, and imprecise}
+The proof now needs to relate the two execution with each other.
+\Cref{fig:proofdiag:rtp} visualizes this situation\MP{imprec}.
+\begin{figure}[!ht]
   \begin{center}
@@
 \end{figure}
+\MPin{
+	thicker lines and dots needed, not visible otherwise
+}
+
+At the bottom, there is the target-level execution between operational states $\trg{\emptyset},\trg{\Omega_{1}}$, and so on.
+The execution starts in the context $\trg{\library_{\ctx}}$ and is split by transitivity of $\trg{\estepto{}{}{}}$ into parts before, in, and after the component $\cc{\Ltms}{\trg{L}}(\src{p})$ executes\footnote{For sake of simplicity, this paper considers only one context-switch from context to component and back again.
+Recursive calls between context and component functions are disregarded.}.
+Initially, the empty operational states $\src{\emptyset}$ and $\trg{\emptyset}$ need to agree by means of a cross-language state relation $\multimap_{\emptyset}$.
+This relation is indexed by a function $\delta$ that maps $\trg{L}$-level locations $\trg{\loc}$ to $\Ltms$-level locations.
+The operational state of the executions up to the component $\src{p}$ need to agree via $\multimap_{\delta}$, both contexts need to behave in a similar way such that the traces also agree with each other.
+However, the relation allows $\trg{\library_{\ctx}}$ to perform internal actions that may change the operational state but are not visible on the trace, such as internal function calls, that does not $\src{\library_{\ctx}}$ need to reflect.
+Before and after the compiler wrapper code, the operational states must agree via another cross-language relation $\approx_{\delta_{w_{}}}$, which disallows these mismatches with respect to internal function calls.
+Intuitively, the behavior of the components should be tight: the source-level execution must be preserved by the compiler in the target-level execution.
+Dually, when the component is done, the relations take turns again and the states must be related with $\multimap_{\delta_{w_{2}}}$
+Finally, since both execution traces are similar in the sense that their projections $\msfilterLtms{\src{\trace}}$ and $\msfilterL{\trg{\trace}}$ yield specification traces that satisfy \Cref{def:monsat}, the claim follows \Cref{lem:mon:tmsafe}.
+
+\MPin{
+	why can we use a BT?
+	it's because the def in section 2.2 is eq to anotehr one, because the props are Safety, we use the RSC equivalent def.
+	all these intuitions are missing.
+	they are technical, so we only keep them here.
+
+  MK: but, we don't actually rewrite it to RSC when doing the proof! The compilers are not robustly preserving with respect to an arbitrary safety property, but, e.g., just for \gls{tms}!
+}
 
@@
 \MPin{
-	example code?
-    MK: still needed now that we reference ex. 3.10?
+	show the result of the compilation as listing, and then describe the trace a bit more in detail ?
 }
@@
 
-\Cref{thm:ccab:rtp:ms} follows from \Cref{thm:cca:rtp:tms,thm:ccb:rtp:sms} using \Cref{thm:rtp}.
+\Cref{thm:ccab:rtp:ms}\MP{states ... } follows from \Cref{thm:cca:rtp:tms,thm:ccb:rtp:sms} using \Cref{thm:rtp}.
 
@@
   %
+  \MKin{
   MK: No?
@@
 If we don't do this, we essentially re-encode the definition of subst into the compiler
+  }
 }
@@
 %
-MK: Should this appear here in the paper? It's basically identical to \Cref{thm:ccab:rtp:ms}
+	\MKin{ 
+	Should this appear here in the paper? It's basically identical to \Cref{thm:ccab:rtp:ms}
+
+	\MPin{yes it should}}
 }

### 2023-07-10 popl/main.tex
5 and 6

@@
 An easy fix to sidestep this is to use non-nullable pointers as parameters instead, e.g., as presented in Cyclone~\cite{jim2002cyclone}.
-With this, a secure compiler needs to ensure that no malicious target-level context can supply unitialised memory and thereby sidestep the guarantees of the source language, e.g., by inserting appropriate checks prior to entering the body of \texttt{strncpy} or using capabilities~\cite{michael2023mswasm}.
+With this, a secure compiler needs to ensure that no malicious target-level context can supply unitialised memory and thereby sidestep the guarantees of the source language, e.g., by inserting appropriate checks prior to entering the body of \texttt{strncpy}.
+\MPin{
+	do we have examples of such wrapper-chesk inserting secure compilers?
+	mswasm compiler ... not really.
+	or we could say that another solution is  using capabilities, and then the mswasm compiler can be used
+}
 \end{example}
@@
 This paper introduces a framework for reasoning about the composition of secure and optimising compiler passes akin to those of \Cref{ex:strncpy} and it showcases the power of this framework by instantiating it on a multi-pass compilation chain.
-To this end, this paper first discusses how to compose security properties, such as \gls{tms} and \gls{sms} into \gls{ms}, and \gls{cct}.
+To this end, this paper first discusses how to compose security properties, such as temporal and spatial memory safety into general memory safety, and cryptographic constant time.
+\MP{
+	acronym? non acronym? 
+}
 Then, this paper defines compiler composition and formalises what kind of security properties does the composition of two compiler passes preserve, given what the individual passes preserve.
-The paper then defines several secure compiler passes, where each is either preserving a different security property (e.g., temporal or spatial memory safety) or performing a security-preserving optimisation, (e.g., applying \gls{cf} or \gls{dce}).
+The paper then defines several secure compiler passes, where each is either preserving a different security property (e.g., temporal or spatial memory safety) or performing a security-preserving optimisation, (e.g., applying constant-folding or dead-code elimination).
+\MP{
+	acro? non acro?
+}
 Finally, this paper shows how to compose these secure compiler passes into a multi-pass compilation chain which is proven to provide end-to-end preservation of general memory safety and cryptographic constant time.

### 2023-07-10 popl/main.tex
nit

@@
 
-The robust preservation of properties as a compiler-level criterion has been analyzed extensively~\cite{abate2019jour,patrignani2021rsc,abate2021extacc,patrignani2022universal,patrignani2019survey,kruse2022csc}.
+The robust preservation of properties as a compiler-level criterion has been analyzed extensively~\cite{abate2019jour,patrignani2021rsc,abate2021extacc,patrignani2022universal,patrignani2019survey}.
 This work has made use of that previous work.
@@
 The work relating robust preservation with universal composability~\cite{patrignani2022universal} is closest to what this paper presents.
-The authors demonstrate a similar compositionality theorem to what is presented here (\Cref{sec:compcomp}) as well as in an earlier version of this work~\cite{kruse2022csc}.
-However, they do not demonstrate the scalability of the approach by means of a case study.
+% The authors demonstrate a similar compositionality theorem to what is presented here (\Cref{sec:compcomp}) as well as in an earlier version of this work~\cite{kruse2022csc}.
+% However, they do not demonstrate the scalability of the approach by means of a case study.
 
@@
 Note that this work also presents secure compilers that are optimising, but contrary to the other~\cite{vu2021reconciling}, provides a formal account of these in the robust preservation framework.
-Lastly, the authors of this paper have presented ongoing work~\cite{patrignani2023blame} on a weaker robust preservation criteria based on the concept blame.
+% Lastly, the authors of this paper have presented ongoing work~\cite{patrignani2023blame} on a weaker robust preservation criteria based on the concept blame.
 
@@
 %% consistent spelling of the heading.
-\begin{acks}
-  We would like to thank the anonymous reviewers for their feedback.
-\end{acks}
+% \begin{acks}
+%   We would like to thank the anonymous reviewers for their feedback.
+% \end{acks}
 

### 2023-07-10 popl/main.tex
:chicken: restructure coq

@@
 With this, a secure compiler needs to ensure that no malicious target-level context can supply unitialised memory and thereby sidestep the guarantees of the source language, e.g., by inserting appropriate checks prior to entering the body of \texttt{strncpy} or using capabilities~\cite{michael2023mswasm}.
+% a possible solution would be to compile to machiens with capabilities such as armorello or cheri
 \end{example}
@@
   Running this on the monitor would result in getting stuck, since there is no matching rule to step in the presence of $\absev{Any}$ event. % pun intended
-  By means of additional features to ensure cryptographic constant time even in the presence of memory reads and loads, such as a flag to enable a data independent timing mode, which is present in both Arm~\cite[p.~543]{arm-refman} and Intel~\cite[p.~80]{intel-refman} processors, the original trace can contain $\ev{Use}\ \loc_{y}\ 1729;\comp;\unlock$ instead of $\ev{Use}\ \loc_{y}\ 1729;\comp\lock$.
+  By means of additional features to ensure cryptographic constant time even in the presence of memory reads and loads, such as a flag to enable a data independent timing mode, which is present in both Arm~\cite[p.~543]{arm-refman} and Intel~\cite[p.~80]{intel-refman} processors, the original trace can contain $\ev{Use}\ \loc_{y}\ 1729;\comp;\unlock$ instead of $\ev{Use}\ \loc_{y}\ 1729;\comp;\lock$.
   Because of this, the whole trace of the component equates to $\absev{\emptyevent}$ and the monitor can step without getting stuck.
@@
 
-The robust preservation of properties as a compiler-level criterion has been analyzed extensively~\cite{abate2019jour,patrignani2021rsc,abate2021extacc,patrignani2022universal,patrignani2019survey}.
+The robust preservation of properties as a compiler-level criterion has been analyzed extensively~\cite{abate2019jour,patrignani2021rsc,abate2021extacc,patrignani2022universal,patrignani2019survey,kruse2022csc}.
 This work has made use of that previous work.
@@
 The work relating robust preservation with universal composability~\cite{patrignani2022universal} is closest to what this paper presents.
-% The authors demonstrate a similar compositionality theorem to what is presented here (\Cref{sec:compcomp}) as well as in an earlier version of this work~\cite{kruse2022csc}.
-% However, they do not demonstrate the scalability of the approach by means of a case study.
+The authors demonstrate a similar compositionality theorem to what is presented here (\Cref{sec:compcomp}) as well as in an earlier version of this work~\cite{kruse2022csc}.
+However, they do not demonstrate the scalability of the approach by means of a case study.
 
@@
 Note that this work also presents secure compilers that are optimising, but contrary to the other~\cite{vu2021reconciling}, provides a formal account of these in the robust preservation framework.
-% Lastly, the authors of this paper have presented ongoing work~\cite{patrignani2023blame} on a weaker robust preservation criteria based on the concept blame.
+Lastly, the authors of this paper have presented ongoing work~\cite{patrignani2023blame} on a weaker robust preservation criteria based on the concept blame.
 
@@
 %% consistent spelling of the heading.
-% \begin{acks}
-%   We would like to thank the anonymous reviewers for their feedback.
-% \end{acks}
+\begin{acks}
+  We would like to thank the anonymous reviewers for their feedback.
+\end{acks}
 

### 2023-07-10 popl/main.tex
Change we to indirect speech

--- 2023-07-10 popl/main.tex
@@
 \footnote{
-Throughout the paper, we indicate sequences with an overbar (i.e., $\trace$), we denote empty sequences with $\hole{\cdot}$ and concatenate sequences $\trace[_{1}],\trace[_{2}]$ as $\trace[_{1}]\cdot\trace[_{2}]$.
+Throughout the paper, sequences are indicated with an overbar (i.e., $\trace$), empty sequences with $\hole{\cdot}$, and concatenation of sequences $\trace[_{1}],\trace[_{2}]$ as $\trace[_{1}]\cdot\trace[_{2}]$.
 Prepending elements to sequences uses the same notation: $\event\cdot\trace$.
@@
 This program behaviour is insecure with respect to a canonical notion of (temporal) memory safety dictating no use-after-frees of pointers~\cite{nagarakatte2010cets}\MP{should cite catalin's MS paper too}, because it reads from a memory location that was freed already.
-The previous finite trace prefix is enough to decide that the trace does not satisfy temporal memory safety and there is no way to append events to this prefix which would result in the trace being admissible.
-In the following, we indicate the execution of a whole program $\wholeprogvar$ that terminates in state $\runtimetermvar$ according to the language semantics and produces trace $\trace$ as: $\progstepto{\wholeprogvar}{\runtimetermvar}{\trace}$.
-With this, we define property satisfaction as follows:
+The previous finite trace prefix is enough to decide that the trace does not satisfy \gls{tms} and there is no way to append events to this prefix which would result in the trace being admissible.
+In the following, the execution of a whole program $\wholeprogvar$ that terminates in state $\runtimetermvar$ according to the language semantics and produces trace $\trace$ is written as $\progstepto{\wholeprogvar}{\runtimetermvar}{\trace}$.
+With this, property satisfaction is defined as follows:
 \bul{whole programs $\wholeprogvar$ satisfy a property $\pi$} iff \iul{$\wholeprogvar$ yields a trace $\trace$} such that \oul{$\trace$ satisifies $\pi$} (\Cref{def:propsat}).
@@
 Thus, from a security perspective, this considers only a passive attacker model, where the attacker observes the execution and, e.g., retrieves secrets from that.
-To consider a stronger model similarly to what existing work does~\cite{abate2019jour,abate2021extacc,patrignani2021rsc}\MP{many more things on RS here}, we extend the concept of satisfaction with {\em robustness}.
+To consider a stronger model similarly to what existing work does~\cite{abate2019jour,abate2021extacc,patrignani2021rsc}\MP{many more things on RS here}, the concept of satisfaction can be extended with {\em robustness}.
 Robust satisfaction considers partial programs $\progvar$, i.e., components with missing imports, which cannot run until said imports are fulfilled.
@@
   Consider \texttt{CVE-2021-3564}~\cite{doublefree-bluetooth}, one of many submissions for a double-free vulnerability.
-  The bug\MP{not bug, vulnerability.
-uniform below (perhaps?)} arises due to a race condition where the context-level function \texttt{hci\_cmd\_work} was not expected to behave maliciously, since it resides in the same source-code repository where the vulnerability occurs.
+  The vulnerability arises due to a race condition where the context-level function \texttt{hci\_cmd\_work} was not expected to behave maliciously, since it resides in the same source-code repository where the vulnerability occurs.
   Nevertheless, the component-level code of \texttt{hci\_dev\_do\_open} is linked with \texttt{hci\_cmd\_work} and does not atomically check whether the pointer has been freed already:
@@
 Unfortunately, correct compilers may be insecure compilers~\cite{patrignani2019survey}\MP{cite kennedy and teh general SC dagstuhl and ACM papers too}, and programs translated by insecure compilers can violate security properties that the programmer assumes to hold.
-To define when a compiler is secure, this paper uses the robust compilation framework~\cite{abate2019jour}, which we summarize in the following definition.
-
-For \bul{compilers $\cc{\src{L}}{\trg{L}}$ to robustly preserve a class of properties $\class$}, if for any \rul{property $\pi$ of that class $\class$ and programs $\src{p}$ written in $\src{L}$} we have that \iul{the program $\src{\progvar}$ robustly satisfies $\pi$}, then \oul{the compilation of $\src{\progvar}$, $\cc{\src{L}}{\trg{L}}\left(\src{p}\right)$, must robustly satisfy $\pi$}.
+To define when a compiler is secure, this paper uses the robust compilation framework~\cite{abate2019jour}, which the following definition summarizes.
+
+For \bul{compilers $\cc{\src{L}}{\trg{L}}$ to robustly preserve a class of properties $\class$}, if for any \rul{property $\pi$ of that class $\class$ and programs $\src{p}$ written in $\src{L}$} where \iul{$\src{\progvar}$ robustly satisfies $\pi$}, then \oul{the compilation of $\src{\progvar}$, $\cc{\src{L}}{\trg{L}}\left(\src{p}\right)$, must robustly satisfy $\pi$}.
 \MPin{
@@
 \gls{cct} is a hyperproperty\MP{hypersafety?}\MP{what is a hp?} and, thus, not as easy to check with just monitors.\MP{chy? cit?} %Beutner et al
-Hence, we define \gls{scct}, a stricter variant that enforces the policy that no secret should appear on a trace.\MP{cit? did we take some inspiration}
+To remedy this, this section defines the property \gls{scct}, a stricter variant of \gls{cct} that enforces the policy that no secret should appear on a trace.\MP{cit? did we take some inspiration}
 
@@
 
-%\begin{lstlisting}[language=c,caption=``Wrong bounds check of two 64-bit integers.'']
-%if at <= bounds {
-%  x[at] = 42;
-%}
-%\end{lstlisting}
-%Note the subtle bug of using \verb|<=| instead of \verb|<|, leading to an out-of-bounds access whenever \verb|at = bounds|.
-%Furthermore, when \verb|at| and \verb|bounds| are 64-bit integers on a 32-bit architecture, the comparison may not be performed in constant-time: The compiler may bail out as soon as the lower 32 bits are unequal, not bothering to compare the higher 32 bits.
-
 \section{Related Work\pages{2}}\label{sec:relwork}

### 2023-07-10 popl/main.tex
rewrite little

@@
 \newcommand\bmmax{0}
-\documentclass[utf8,acmsmall,review,screen,dvipsnames]{acmart}
+\documentclass[utf8,acmsmall,review,screen,dvipsnames,anonymous]{acmart}
 
@@
 An easy fix to sidestep this is to use non-nullable pointers as parameters instead, e.g., as presented in Cyclone~\cite{jim2002cyclone}.
-With this, a secure compiler needs to ensure that no malicious target-level context can supply unitialised memory and thereby sidestep the guarantees of the source language, e.g., by inserting appropriate checks prior to entering the body of \texttt{strncpy} or using capabilities~\cite{michael2023mswasm}.
+With this, a secure compiler needs to ensure that no malicious target-level context can supply unitialised memory and thereby sidestep the guarantees of the source language.
+For example, if the target language is CHERI~\MK{cit} or MS-WASM~\cite{michael2023mswasm}\MK{check this} the compiler can ensure this by leavaraging the object capabilities.
+% , e.g., by inserting appropriate checks prior to entering the body of \texttt{strncpy} or using capabilities~\cite{michael2023mswasm}.
 % a possible solution would be to compile to machiens with capabilities such as armorello or cheri

### 2023-07-11 popl/main.tex
ize -> ise

@@
 
+  \MKin{context}
   Memory safety necessitates secrecy, since breaking it can lead to an arbitrarily large attack space, i.e., from simply reading a secret up to alteration of control-flow in order to execute arbitrary code.
@@
   Prior work has shown that strategies to ensure cryptographic constant time can be rendered useless, since a compiler may simply optimize them away.
+  \MKin{need}
   This is why secure compilation is an integral ingredient to achieve security, because a secure compiler guarantees (formally) that properties that hold in the source, also hold in the target.
@@
 the text describes correctness
-
+  \MKin{task}
   Unfortunately, it is an open question whether secure compilers can be engineered in a modular way.
@@
 %, because this style of engineering is well-established for its simplifying divide-and-conquer approach as well as enhanced reusability.
-
+  \MKin{object+findings}
   This paper answers that question positively: The composition of compilers that are secure with respect to some security-properties yields the intersection of these properties.
+  \MKin{object}
   This article discusses different variants of compositions of secure compilers and instantiates parts of these results in a case-study.
   To this end, it presents an optimizing, secure compilation chain that preserves both memory safety and cryptographic constant time.
-%  Memory safety is achieved by composing secure compilers for both temporal and spatial memory safety, where the compiler for spatial memory safety is an instrumentation that inserts bounds-checks in front of pointer dereferences.
-  %The compiler for cryptographic constant time exploits the use of a data (operand) timing independent mode and the employed optimisations of the compilation chain are constant folding and dead code elimination.
+  \MKin{conclusion}
   The compositionality results enable modular proofs for the presented, secure compiler. % that is secure with respect to both memory safety and cryptographic constant time.
   Some results of this work are implemented in the Coq proof assistant.
-
-% Secure compilers emit code that is hardened against adversarial target-level attacks, such as memory corruption or alteration of control-flow.
-% It is an open question whether secure compilers compose in a well-behaved manner.
-% This paper answers this question positively: The composition of compilers that are secure with respect to some security-properties yields the intersection of these properties.
-% This article discusses different variants of compositions of secure compilers and instantiates parts of these results in a case-study.
-% To this end, it presents an optimizing, secure compilation chain that preserves both memory safety and cryptographic constant time.
-% Memory safety is achieved by composing secure compilers for both temporal and spatial memory safety, where the compiler for spatial memory safety is an instrumentation that inserts bounds-checks in front of pointer dereferences.
-% The compiler for cryptographic constant time exploits the use of a data (operand) timing independent mode and the employed optimisations of the compilation chain are constant folding and dead code elimination.
-% The compositionality results enable modular proofs for the whole compilation chain that is secure with respect to both memory safety and cryptographic constant time.
 
@@
 
-\gls{ms} is a security property that ensures that there are no out-of-bounds accesses as well as no reads or writes to uninitialized memory.
+\gls{ms} is a security property that ensures that there are no out-of-bounds accesses as well as no reads or writes to uninitialised memory.
 Programs that do not attain memory safety, can easily have their private data be leaked to attackers or be corrupted~\cite{lemay2021ccc}.
@@
 
-Even though mitigations for such attacks exist~\cite{bond2017vale,almeida2017jasmin}, some compilers may optimize these mitigations away~\cite{barthe2018sec}.
+Even though mitigations for such attacks exist~\cite{bond2017vale,almeida2017jasmin}, some compilers may optimise these mitigations away~\cite{barthe2018sec}.
 This is of course not what programmers expect and to ensure such mitigations persist in the compiled code, one can use a {\em secure} compilers.
@@
 These fixes alone are not enough to make this program secure in a robust setting, i.e., a malicious context could trigger a violation of either \gls{ms} or \gls{cct}.
-Consider a program context that calls \texttt{strncpy} with an argument for \texttt{src} that points to uninitialized memory.
+Consider a program context that calls \texttt{strncpy} with an argument for \texttt{src} that points to uninitialised memory.
 This leads to a series of reads from uninitialised memory, which is an immediate \gls{ms} issue that can have devastating real-world implications~\cite{uninit-0,uninit-1,uninit-2,uninit-3,uninit-4}.
@@
 
-  \item The key contributions of this paper are formalized in the Coq proof assistant and the paper indicates this with \CoqSymbol.
+  \item The key contributions of this paper are formalised in the Coq proof assistant and the paper indicates this with \CoqSymbol.
 \end{itemize}
@@
 Unfortunately, correct compilers may be insecure compilers~\cite{patrignani2019survey}\MP{cite kennedy and teh general SC dagstuhl and ACM papers too}, and programs translated by insecure compilers can violate security properties that the programmer assumes to hold.
-To define when a compiler is secure, this paper uses the robust compilation framework~\cite{abate2019jour}, which the following definition summarizes.
+To define when a compiler is secure, this paper uses the robust compilation framework~\cite{abate2019jour}, which the following definition summarises.
 
@@
 To project specification traces $\trace$ onto these monitor-level traces $\abstrace$, each property has an associated event agreement relation $\traceagree[]{\pi}{\trace}{\abstrace}$.
-This relation projects the events\MP{whaddaya mean?}, while the stutter-equivalence $\traceagree{\pi}{\trace}{\abstrace}$ remains the same for all properties $\pi$, yet is indexed by $\pi$ to emphasize the kind of event agreement.\MP{unclear sentence}
+This relation projects the events\MP{whaddaya mean?}, while the stutter-equivalence $\traceagree{\pi}{\trace}{\abstrace}$ remains the same for all properties $\pi$, yet is indexed by $\pi$ to emphasise the kind of event agreement.\MP{unclear sentence}
 With agreements, this section defines monitor satisfaction for traces and then it proves that monitor satisfaction implies property satisfaction.
@@
   \begin{lstlisting}[language=c,basicstyle=\small\ttfamily,morekeywords={new,delete}]
-int main() {  // allocates 17 integers initialized to 2
+int main() {  // allocates 17 integers initialised to 2
   int* x = new[17](2);
@@
 
-Since the composition of secure compilers is again a secure compiler, the theorem generalizes to a whole chain of $n$ secure compilers.
+Since the composition of secure compilers is again a secure compiler, the theorem generalises to a whole chain of $n$ secure compilers.
 Pretending that the following compilers are secure, consider the compilation chain for $\src{TypeScript}$.
@@
 The proof now needs to relate the two execution with each other.
-\Cref{fig:proofdiag:rtp} visualizes this situation\MP{imprec}.
+\Cref{fig:proofdiag:rtp} visualises this situation\MP{imprec}.
 \begin{figure}[!ht]
@@
 Different to the approach of this paper, these approaches necessitate that the programmer writes \gls{cct} code.
-An approach to allow programmers to write more high-level code is CryptOpt~\cite{kuepper2023cryptopt}, which generates efficient target-code by means of a randomized search.
+An approach to allow programmers to write more high-level code is CryptOpt~\cite{kuepper2023cryptopt}, which generates efficient target-code by means of a randomised search.
 This paper abstracts over concrete mitigation strategies and simply assumes that there is a flag to switch to a cryptographic-constant time execution mode.
-This can be realized by employing the FaCT~\cite{cauligi2019fact} compiler, which translates common non-constant time code patterns to be constant-time, and the data (object) independent timing execution mode of modern processors.
+This can be realised by employing the FaCT~\cite{cauligi2019fact} compiler, which translates common non-constant time code patterns to be constant-time, and the data (object) independent timing execution mode of modern processors.
 

### 2023-07-11 popl/main.tex
Add cits

@@
 An easy fix to sidestep this is to use non-nullable pointers as parameters instead, e.g., as presented in Cyclone~\cite{jim2002cyclone}.
-With this, a secure compiler needs to ensure that no malicious target-level context can supply unitialised memory and thereby sidestep the guarantees of the source language.
-For example, if the target language is CHERI~\MK{cit} or MS-WASM~\cite{michael2023mswasm}\MK{check this} the compiler can ensure this by leavaraging the object capabilities.
-% , e.g., by inserting appropriate checks prior to entering the body of \texttt{strncpy} or using capabilities~\cite{michael2023mswasm}.
-% a possible solution would be to compile to machiens with capabilities such as armorello or cheri
+With this, a secure compiler needs to ensure that no malicious target-level context can supply unitialised memory and thereby sidestep the non-nullability guarantees of the source language.
+For compilers to target languages that provide capabilities (such as CHERI~\cite{woodruff2014CHERI}, Arm Morello~\cite{arm-morello}, or MSWasm~\cite{michael2023mswasm}), which are unforgable handles that encapsulate information about a pointer, they can leavarage the information stored in these capablities to ensure the non-nullability guarantees in the target.
 \end{example}

### 2023-07-11 popl/main.tex
upto ex 3.8

@@
 As an example, consider a trace describing an interaction with a memory where the deallocation of an address $\loc$ precedes a read at that address in memory: $\ev{Dealloc\ \loc}\cdot\ev{Read\ \loc\ 1729}\cdot\dots$.
-This program behaviour is insecure with respect to a canonical notion of (temporal) memory safety dictating no use-after-frees of pointers~\cite{nagarakatte2010cets}\MP{should cite catalin's MS paper too}, because it reads from a memory location that was freed already.
+This program behaviour is insecure with respect to a canonical notion of (temporal) memory safety dictating no use-after-frees of pointers~\cite{nagarakatte2010cets,azevedo2018meaningsofms}, because it reads from a memory location that was freed already.
 The previous finite trace prefix is enough to decide that the trace does not satisfy \gls{tms} and there is no way to append events to this prefix which would result in the trace being admissible.
@@
 Thus, from a security perspective, this considers only a passive attacker model, where the attacker observes the execution and, e.g., retrieves secrets from that.
-To consider a stronger model similarly to what existing work does~\cite{abate2019jour,abate2021extacc,patrignani2021rsc}\MP{many more things on RS here}, the concept of satisfaction can be extended with {\em robustness}.
+To consider a stronger model similarly to what existing work does~\cite{abate2019jour,abate2021extacc,patrignani2021rsc,maffeis2008code-carrying,gordon2003authenticity,fournet2007authorization,bengtson2011refine,backes2014uniontyps}, the concept of satisfaction can be extended with {\em robustness}.
 Robust satisfaction considers partial programs $\progvar$, i.e., components with missing imports, which cannot run until said imports are fulfilled.
 To remedy this, {\em linking} takes two partial programs $\progvar[_{1}],\progvar[_{2}]$ and produces a whole program $\wholeprogvar$, i.e., $\link{\progvar[_{1}]}{\progvar[_{2}]}=\wholeprogvar$.
-As typically done in works that consider the execution of partial programs~\cite{abate2019jour,devriese2018parametricity}, \MP{Add papers from Amal. Papers from just Dominique on FAC. RS works.}
+As typically done in works that consider the execution of partial programs~\cite{abate2019jour,devriese2018parametricity,patrignani2021rsc,korashy2021capableptrs,strydonck2019lincap,devriese2017modular,bowman2015noninterference,ahmed2011equivcps,patterson2017linkingtyps},
 this paper assumes that whole programs are the result of linking partial programs referred to as {\em context} ($\ctx$) and {\em component} ($\comp$).
@@
 That is, for a whole program $\src{\wholeprogvar}$, the compiler should relate the $\src{L}$ semantics of $\src{\wholeprogvar}$ with the semantics of $\trg{T}$ in such a way that they are ,,compatible''.
-Unfortunately, correct compilers may be insecure compilers~\cite{patrignani2019survey}\MP{cite kennedy and teh general SC dagstuhl and ACM papers too}, and programs translated by insecure compilers can violate security properties that the programmer assumes to hold.
+Unfortunately, correct compilers may be insecure compilers~\cite{patrignani2019survey,kennedy2006secure.net,abadi1999protect,ahmed2018dagstuhl} and programs translated by insecure compilers can violate security properties that the programmer assumes to hold.
 To define when a compiler is secure, this paper uses the robust compilation framework~\cite{abate2019jour}, which the following definition summarises.
@@
 For \bul{compilers $\cc{\src{L}}{\trg{L}}$ to robustly preserve a class of properties $\class$}, if for any \rul{property $\pi$ of that class $\class$ and programs $\src{p}$ written in $\src{L}$} where \iul{$\src{\progvar}$ robustly satisfies $\pi$}, then \oul{the compilation of $\src{\progvar}$, $\cc{\src{L}}{\trg{L}}\left(\src{p}\right)$, must robustly satisfy $\pi$}.
-\MPin{
-	somewhere we should say that a class of props can also include a single prop
-}
 
@@
 \end{definition}
+
+Note that a class of properties $\class$ can represent just one property $\pi$ by lifting~\cite{clarkson2008hyper} that property to sets of properties, i.e., use the powerset of $\pi$ instead of $\pi$ itself.
+Because of this, this paper writes $\rtp{\cc{\src{L}}{\trg{L}}}{\pi}$, even though $\pi$ is a property and not a class.
 
@@
   The pointer is valid for one $\trg{char}$-sized memory cell, as expected, but the actual read operation now returns $\trg{sizeof(int)}$ many bytes instead of just $\trg{sizeof(char)}$ many.
-  Thus, even if $\src{foo}$ may have been robust with respect to the no-reads-out-of-bounds property\MP{sms?}, its compiled counterpart is not and therefore the compiler fails to attain \Thmref{def:rtp}.
+  Thus, even if $\src{foo}$ may have been robust with respect to the \gls{sms}, its compiled counterpart is not and therefore the compiler fails to attain \Thmref{def:rtp}.
 \end{example}
@@
 These properties are of practical importance (as mentioned in \Cref{sec:introduction}) and also of interest in the case study (\Cref{sec:casestud:defs,sec:casestud:rtp}) this paper presents later. 
-Then, this section introduces monitors for each of the properties presented earlier (\Cref{subsec:monitors}).
-\MP{why?}
+Lastly, for each of the key properties, this section introduces corresponding monitors (\Cref{subsec:monitors}) that check them.
 
@@
 Moreover, an instruction whose timing is data-dependent must not have a secret as an operand.
-Typical operations with data-dependent timing are branches and certain binary operations, such as division\footnote{This is highly architecture-dependent, but division is an operation that serves as a classic example for a data-dependent timing instruction.}.
+Typical operations with data-dependent timing are branches and certain binary operations, such as division\footnote{This is highly architecture-dependent, but division is an operation that serves as a classic example for a data-dependent timing instruction, e.g., \cite[p.~755]{arm-refman}.}.
 Both operations are also modelled in the specification trace model ($\ev{Branch\ n}$ and $\ev{Binop\ n}$).
@@
 
-\gls{sms} disregards\MP{prohibits?} out-of-bounds accesses:
+\gls{sms} prohibits out-of-bounds accesses:
 
@@
 
-Full \gls{ms} is then described as the conjunction of \Cref{def:trace:tmsdef,def:trace:smsdef}.
+Full \gls{ms} (similar to earlier work~\cite{nagarakatte2009soft,nagarakatte2010cets,jim2002cyclone,necula2005ccured,michael2023mswasm}) is then described as the conjunction of \Cref{def:trace:tmsdef,def:trace:smsdef}.
+%%% Catalin's paper doesn't apply, since their definition of MS relies on noninterference to guarantee data isolation.
+%%% our definition does not guarantee that (on purpose, we want to compose it with cct)
 Note, however, that this definition says nothing about memory-safety issues introduced by side-channels, such as speculation.
@@
 \end{definition}
-\MPin{
-	cite catalin's paper and the mswasm paper for the def of ms
-}
 
@@
 
-\gls{cct} is a hyperproperty\MP{hypersafety?}\MP{what is a hp?} and, thus, not as easy to check with just monitors.\MP{chy? cit?} %Beutner et al
-To remedy this, this section defines the property \gls{scct}, a stricter variant of \gls{cct} that enforces the policy that no secret should appear on a trace.\MP{cit? did we take some inspiration}
+\gls{cct} is a hypersafety property~\cite{barthe2018sec} and, thus, difficult to check with monitors.
+This is because, intuitively, hypersafety properties can relate multiple execution traces with eachother, but monitors work on a single execution. not as easy to check with just monitors.
+To remedy this, this section defines the property \gls{scct}, a stricter variant of \gls{cct} that enforces the policy that no secret should appear on a trace (inspired by earlier work~\cite{almeida2017jasmin}).
 
@@
 The combination of \gls{ms} and \gls{scct} is the intersection of these properties, \gls{msscct}.
-It entails all program behaviours that are both memory safe and strictly cryptographic constant time.
+Since \gls{ms} and \gls{scct} are just sets of traces that, intuitively, contain all program behaviors that follow a security policy, the intersection of them contains all program behaviors that follow both security policies, i.e., it entails all program behaviours that are both \gls{ms} and \gls{scct}.
 
@@
   $
-  \msscctsafe:=\left\{\trace \left| \begin{array}{rcl}
-                                      \trace\in\mssafe & \text{ and } & \trace\in\scctsafe
-                                    \end{array}\right.\right\}
+  \msscctsafe:=\mssafe\cap\scctsafe
   $
 \end{definition}
-\MPin{
-	comment on the composition at this levle?
-}
 
@@
 
-Monitors can be used to enforce safety properties and this section presents concrete instantiations of monitors for the previousvy presented properties (\Cref{subsec:propdefs}).
-\MPin{
-	why
-}
-To lessen the burden when proving\MP{proving what}, each monitor makes use\MP{uses} of a custom trace model that contains only the relevant information related to the property the monitor checks.
-To project specification traces $\trace$ onto these monitor-level traces $\abstrace$, each property has an associated event agreement relation $\traceagree[]{\pi}{\trace}{\abstrace}$.
-This relation projects the events\MP{whaddaya mean?}, while the stutter-equivalence $\traceagree{\pi}{\trace}{\abstrace}$ remains the same for all properties $\pi$, yet is indexed by $\pi$ to emphasise the kind of event agreement.\MP{unclear sentence}
+\begin{figure}[h!]
+  \judgbox{\traceagree{\pi}{\trace}{\abstrace}}{,,Specification-level trace $\trace$ agrees with monitor-level trace $\abstrace$ with respect to property $\pi$.''}
+  %
+  \typerule{traceagree-empty}{}{
+    \traceagree{\pi}{\hole{\cdot}}{\hole{\cdot}}
+  }{traceagree-empty}
+  %
+  \typerule{traceagree-ign-L}{
+    \traceagree{\pi}{\trace}{\abstrace}
+  }{
+    \traceagree{\pi}{\emptyevent\cdot\trace}{\abstrace}
+  }{traceagree-ign-L}
+  %
+  \typerule{traceagree-ign-R}{
+    \traceagree{\pi}{\trace}{\abstrace}
+  }{
+    \traceagree{\pi}{\trace}{\absev{\emptyevent}\cdot\abstrace}
+  }{traceagree-ign-R}
+  %
+  \typerule{traceagree-cons}{
+    \traceagree[]{\pi}{\event}{\absevent} &
+    \traceagree{\pi}{\trace}{\abstrace}
+  }{
+    \traceagree{\pi}{\event\cdot\trace}{\absevent\cdot\abstrace}
+  }{traceagree-cons}
+  \caption{Trace-Agreement relation that equates specification-level traces with monitor-level traces.}\label{fig:stuttereq}
+\end{figure}
+
+Monitors enforce safety properties by accepting or rejecting traces, i.e., if it rejects a trace, the trace does not satisfy the property the monitor checks.
+Since reasoning on monitors is easier than directly on just traces, this section presents a monitor for each of the previously shown safety properties (\Cref{subsec:propdefs}).
+To lessen the burden when proving that a monitor accepts the trace of a program execution, each monitor uses a custom trace model that contains only the relevant information related to the property the monitor checks.
+To go from specification traces $\trace$ to monitor-level traces $\abstrace$, each property $\pi$ has an associated event agreement relation $\traceagree[]{\pi}{\event}{\absevent}$.
+\Cref{fig:stuttereq} shows how the event agreement is lifted to traces.
+The trace agreement is the same for all properties $\pi$ up to the use of the event agreement in \Cref{tr:traceagree-cons}.
 With agreements, this section defines monitor satisfaction for traces and then it proves that monitor satisfaction implies property satisfaction.
@@
 However, the definition can easily be adapted by, e.g., attaching a natural number serving as a counter.
-In the reflexive case of $\monitorcheck{\tmsmonitor}{\tmsmonitor}{\absevent}$, the monitor $\tmsmonitor$ must not contain allocated locations, i.e., $\tmsmonitor.allocated=\emptyset$.\MP{wut?}
-This ensures that programs do not forget to free memory.
+Contrary to other monitors in this paper, the multi-step relation of the \gls{tms} monitor is non-standard:
+\begin{center}
+  \judgbox{\monitorcheck{\tmsmonitor}{\tmsmonitor[']}{\abstrace}}{,,Monitor $\tmsmonitor$ multi-steps to $\tmsmonitor[']$ given trace $\abstrace$.''}$\;$\\
+  %
+  \typerule{tms-refl}{
+    \tmsmonitor.\text{allocated}=\emptyset
+  }{
+    \monitorcheck{\tmsmonitor}{\tmsmonitor}{\hole{\cdot}}
+  }{tms-refl}
+  %
+  \typerule{tms-ign-trans}{
+    \monitorcheck{\tmsmonitor}{\tmsmonitor[']}{\abstrace}
+  }{
+    \monitorcheck[]{\tmsmonitor}{\tmsmonitor[']}{\absev{\emptyevent}\cdot\abstrace}
+  }{tms-ign-trans}
+  %
+  \typerule{tms-trans}{
+    \monitorcheck[]{\tmsmonitor}{\tmsmonitor[']}{\absevent}
+    \monitorcheck{\tmsmonitor[']}{\tmsmonitor['']}{\abstrace}
+  }{
+    \monitorcheck[]{\tmsmonitor}{\tmsmonitor['']}{\absevent\cdot\abstrace}
+  }{tms-trans}
+\end{center}
+\Cref{tr:tms-ign-trans,tr:tms-trans} are the same for all monitors, but \Cref{tr:tms-refl} has, in this case, an additional premise that no more locations should be allocated.
+This rejects the behavior of programs that forget to free memory.
 
@@
 
-\begin{lemma}[Traces with Monitor Satisfaction are $\tmssafe$]\label{lem:mon:tmsafe}
+\begin{lemma}[Traces with Monitor Satisfaction are $\tmssafe$]\label{lem:mon:tmssafe}
   If $\montmssafe{\trace}$, then $\trace\in\tmssafe$. %\Coqed
@@
 
-\begin{example}[A program satisfying \gls{tms}]
-  Consider the following program which, when run, yields the trace $\ev{Alloc}\ \loc[_{1}]\ 17;\comp;\unlock\cdot\ev{Use}\ \loc[_{1}]\ 5;\comp;\unlock\cdot\ev{Use}\ \loc[_{1}]\ 5;\comp;\unlock\cdot\ev{Dealloc}\ \loc[_{1}];\comp;\unlock$, where $\loc[_{1}]$ is some memory address determined by the semantics of the programming language.
-  \begin{lstlisting}[language=c,basicstyle=\small\ttfamily,morekeywords={new,delete}]
-int main() {  // allocates 17 integers initialised to 2
-  int* x = new[17](2);
-
-  x[5] = 12;
-  int y = x[5];
-  delete x;
-
-  return y;
+\begin{example}[A program not satisfying \gls{tms}]
+  Consider the following \texttt{C++11} program that calls \texttt{strncpy} (\Cref{ex:strncpy}) and prints the result to the standard output stream.
+
+  \begin{lstlisting}[language=c++,basicstyle=\small\ttfamily,mathescape,commentstyle=\color{gray}]
+int main() {  // allocates 12 chars containing a greeting message
+  char* greetings = new char[12] { "Hello POPL!" }; // <- address ${\color{gray}\loc_{x}}$
+  char* to = new char[12];                          // <- address ${\color{gray}\loc_{y}}$
+
+  strncpy(12, to, greetings);
+  delete to;
+
+  printf("%cOPL\n", to[6]);
 }
   \end{lstlisting}
-  Relating the trace to abstract monitor events yields $\abstrace=\absev{Alloc}\ \loc[_{1}]\ \comp\cdot\absev{Use}\ \loc[_{1}]\ \comp\cdot\absev{Use}\ \loc[_{1}]\ \comp\cdot\absev{Dealloc}\ \loc[_{1}]\ \comp$.
-  It is easy to see that $\montmssafe{\abstrace}$.
-  From \Thmref{lem:mon:tmsafe}, it follows that the program satisfies \Thmref{def:trace:tmsdef}.
+  Up to the body of \texttt{printf}, the program execution yields the specification trace $\ev{Alloc}\ \loc[_{x}]\ 12;\comp;\unlock\cdot\ev{Alloc}\ \loc[_{y}]\ 12;\comp;\unlock\cdot\ev{Use}\ \loc_{x}\ 0;\comp;\unlock\cdot\ev{Use}\ \loc_{y}\ 0;\comp;\unlock\cdot\ev{Use}\ \loc_{x}\ 1;\comp;\unlock\cdot\ldots\cdot\ev{Use}\ \loc_{x}\ 12;\comp;\unlock\cdot\ev{Dealloc}\ \loc[_{y}];\comp;\unlock\cdot\ev{Use}\ \loc_{y}\ 6;\comp;\unlock$.
+  Relating this trace to abstract monitor events yields $\abstrace=\absev{Alloc}\ \loc[_{x}]\ \comp\cdot\absev{Alloc}\ \loc[_{y}]\ \comp\cdot\absev{Use}\ \loc[_{x}]\ \comp\cdot\absev{Use}\ \loc[_{y}]\ \comp\cdot\ldots\cdot\absev{Use}\ \loc[_{x}]\ \comp\cdot\absev{Dealloc}\ \loc[_{y}]\ \comp\cdot\absev{Use}\ \loc[_{y}]\ \comp$.
+  Remembering the definition of \texttt{strncpy} (\Cref{ex:strncpy}), it is easy to see that it does not deallocate its arguments.
+  Even though the trace contains an out-of-bounds access right before returning from \texttt{strncpy}, this is no concern for \gls{tms}, since the location $\loc[_{x}]$ is still allocated.
+  However, having returned from \texttt{strncpy}, the \texttt{main} function continues and deallocates $\loc[_{y}]$ whose subsequent use in the \texttt{printf} call is a use-after-free bug.
+  The fix would be to \lstinline[language=c++,basicstyle=\small\ttfamily]|delete greetings| instead of \texttt{to} and add a \lstinline[language=c++,basicstyle=\small\ttfamily]|delete to| after the \texttt{printf} call, which leads to the abstract monitor trace $\abstrace[']=\absev{Alloc}\ \loc[_{x}]\ \comp\cdot\absev{Alloc}\ \loc[_{y}]\ \comp\cdot\absev{Use}\ \loc[_{x}]\ \comp\cdot\absev{Use}\ \loc[_{y}]\ \comp\cdot\ldots\cdot\absev{Use}\ \loc[_{x}]\ \comp\cdot\absev{Dealloc}\ \loc[_{x}]\ \comp\cdot\absev{Use}\ \loc[_{y}]\ \comp$.
+  It is easy to see that $\montmssafe{\abstrace[']}$ and from \Thmref{lem:mon:tmssafe}, it follows that the program satisfies \Thmref{def:trace:tmsdef}, even though the program still violates \gls{sms}.
 \end{example}
-\MPin{
-	example is good, but it'd be nice to tie it to the running example?
-}
 
@@
 Dually, when the component is done, the relations take turns again and the states must be related with $\multimap_{\delta_{w_{2}}}$
-Finally, since both execution traces are similar in the sense that their projections $\msfilterLtms{\src{\trace}}$ and $\msfilterL{\trg{\trace}}$ yield specification traces that satisfy \Cref{def:monsat}, the claim follows \Cref{lem:mon:tmsafe}.
+Finally, since both execution traces are similar in the sense that their projections $\msfilterLtms{\src{\trace}}$ and $\msfilterL{\trg{\trace}}$ yield specification traces that satisfy \Cref{def:monsat}, the claim follows \Cref{lem:mon:tmssafe}.
 

### 2023-07-11 popl/main.tex
get rid of "it's easy lol"

@@
   The fix would be to \lstinline[language=c++,basicstyle=\small\ttfamily]|delete greetings| instead of \texttt{to} and add a \lstinline[language=c++,basicstyle=\small\ttfamily]|delete to| after the \texttt{printf} call, which leads to the abstract monitor trace $\abstrace[']=\absev{Alloc}\ \loc[_{x}]\ \comp\cdot\absev{Alloc}\ \loc[_{y}]\ \comp\cdot\absev{Use}\ \loc[_{x}]\ \comp\cdot\absev{Use}\ \loc[_{y}]\ \comp\cdot\ldots\cdot\absev{Use}\ \loc[_{x}]\ \comp\cdot\absev{Dealloc}\ \loc[_{x}]\ \comp\cdot\absev{Use}\ \loc[_{y}]\ \comp$.
-  It is easy to see that $\montmssafe{\abstrace[']}$ and from \Thmref{lem:mon:tmssafe}, it follows that the program satisfies \Thmref{def:trace:tmsdef}, even though the program still violates \gls{sms}.
+  It follows that $\montmssafe{\abstrace[']}$ and from \Thmref{lem:mon:tmssafe}, it follows that the program satisfies \Thmref{def:trace:tmsdef}, even though the program still violates \gls{sms}.
 \end{example}
@@
   At that stage, the induction variable \texttt{i} is equal to \texttt{2} and, unfortunately, the order of checks is such that first the cell \texttt{x[i]} is read prior to bounds checking \texttt{i < n}.
-  Because of this, there is an out of bounds\MP{hyphenate: out-of-bounds} memory access right before exiting the function.
+  Because of this, there is an out-of-bounds memory access right before exiting the function.
   This is also visible on the trace, which can be sketched as $\ldots\cdot\ev{Alloc}\ \loc_{x}\ 2;\comp;\unlock\cdot\ldots\cdot\ev{Alloc}\ \loc_{y}\ 2;\comp;\unlock\cdot\ldots\cdot\ev{Use}\ \loc_{x}\ 0;\comp;\unlock\cdot\ev{Use}\ \loc_{y}\ 0;\comp;\unlock\cdot\ev{Use}\ \loc_{x}\ 1;\comp;\unlock\cdot\ev{Use}\ \loc_{y}\ 1;\comp;\unlock\cdot\ev{Use}\ \loc_{x}\ 2;\comp;\unlock\cdot\ldots$, where $\loc_{x}$ and $\loc_{y}$ are the memory addresses associated to \texttt{x} and \texttt{y}, respectively.
@@
 
-  It is easy to see that,\MP{erase. we don't know what's easy} after the allocation events, the state of the monitor is $\left\{(\loc_{x};\comp;2),(\loc_{y};\comp;2)\right\}$.
-  Furthermore, all uses up to the last are acceptable by the monitor, but the last event does not satisfy the premise $2<2$ in \Cref{tr:sms-use}.
+  After the allocation events, the state of the monitor is $\left\{(\loc_{x};\comp;2),(\loc_{y};\comp;2)\right\}$.
+  All uses up to the last are accepted by the monitor, but the last event does not satisfy the premise $2<2$ in \Cref{tr:sms-use}.
   Therefore, the whole program (\texttt{strncpy} linked with this kind of context) is not \gls{sms}.

### 2023-07-11 popl/main.tex
get rid of "it's easy lol"

@@
 \gls{cct} is a hypersafety property~\cite{barthe2018sec} and, thus, difficult to check with monitors.
-This is because, intuitively, hypersafety properties can relate multiple execution traces with eachother, but monitors work on a single execution. not as easy to check with just monitors.
-To remedy this, this section defines the property \gls{scct}, a stricter variant of \gls{cct} that enforces the policy that no secret should appear on a trace (inspired by earlier work~\cite{almeida2017jasmin}).
+This is because, intuitively, hypersafety properties can relate multiple execution traces with eachother, but monitors work on a single execution.
+To sidestep this issue, this section defines the property \gls{scct}, a stricter variant of \gls{cct} that enforces the policy that no secret appears on a trace (inspired by earlier work~\cite{almeida2017jasmin}).
 
@@
   Relating this trace to abstract monitor events yields $\abstrace=\absev{Alloc}\ \loc[_{x}]\ \comp\cdot\absev{Alloc}\ \loc[_{y}]\ \comp\cdot\absev{Use}\ \loc[_{x}]\ \comp\cdot\absev{Use}\ \loc[_{y}]\ \comp\cdot\ldots\cdot\absev{Use}\ \loc[_{x}]\ \comp\cdot\absev{Dealloc}\ \loc[_{y}]\ \comp\cdot\absev{Use}\ \loc[_{y}]\ \comp$.
-  Remembering the definition of \texttt{strncpy} (\Cref{ex:strncpy}), it is easy to see that it does not deallocate its arguments.
+  Remembering the definition of \texttt{strncpy} (\Cref{ex:strncpy}), observe that it does not deallocate its arguments.
   Even though the trace contains an out-of-bounds access right before returning from \texttt{strncpy}, this is no concern for \gls{tms}, since the location $\loc[_{x}]$ is still allocated.
@@
 The two optimising compiler passes from $\Lms$ to $\Lms$ perform \gls{dce} and \gls{cf}, respectively.
-Both are secure with respect to \gls{ms}, the proof of which is relatively easy, because \gls{dce} and \gls{cf} do not change the way memory accesses happen.
+Both are secure with respect to \gls{ms}, the proof of which is relatively simple, because \gls{dce} and \gls{cf} do not change the way memory accesses happen.
 Moreover, since the input and output languages to these compilers are the same, attacker contexts do not have more power in the target language than in the source.

### 2023-07-11 popl/main.tex
sec 4

@@
 
-\subsubsection{Monitor for \glsfirst{tms}}
+\subsubsection{Monitor for \glsfirst{tms}}\label{subsubsec:tms}
 \begin{gather*}
@@
 
-\subsubsection{Monitor for \glsfirst{sms}}
+\subsubsection{Monitor for \glsfirst{sms}}\label{subsubsec:sms}
 \begin{gather*}
@@
 The main result is that secure compilers in the robust compilation framework~\cite{abate2019jour} compose {\em sequentially}.
-This is not intuitive in the sense that there are a number of examples in the security domain where sequential composition does not work~\cite{canetti2006univcomp,mccullough2012compo,fabian2022automatic}.%
-\MP{?canetti?} %% protocols under certain security notions are not secure when composed
-The sequential composition of compilers $\cc{\src{L}}{\trg{L}}$ and $\cc{\trg{L}}{\obj{L}}$ is defined as follows: given an $\src{L}$ program $\src{p}$, its compiled $\obj{L}$ counterpart is obtained by plugging $\src{p}$ into $\cc{\src{L}}{\trg{L}}\circ\cc{\trg{L}}{\obj{L}}$.
-Now, given \bul{$\cc{\src{L}}{\trg{L}}$ robustly preserves $\class[_{1}]$} and \rul{$\cc{\trg{L}}{\obj{L}}$ robustly preserves $\class[_{2}]$}, it follows that \oul{their sequential composition $\cc{\src{L}}{\trg{L}}\circ\cc{\trg{L}}{\obj{L}}$ robustly preserves the intersection of classes $\class[_{1}]$ and $\class[_{2}]$}.
-\MPin{
-	section 3.1 never defined the $\cap$ operator on properties, nor on classes.
-	it uses the 'and'.
-	rectify
-}
+This is not intuitive in the sense that in the security domain, composition does not work without additional generalizations~\cite{canetti2006univcomp,mccullough2012compo,fabian2022automatic}.
+The \bul{sequential composition of compilers $\cc{\src{L}}{\trg{L}}$ and $\cc{\trg{L}}{\obj{L}}$} is defined as follows:
+Given an $\src{L}$ program $\src{p}$ and compilers $\cc{\src{L}}{\trg{L}},\cc{\trg{L}}{\obj{L}}$, \rul{its compiled $\obj{L}$ counterpart is obtained by plugging $\src{p}$ into $\cc{\src{L}}{\trg{L}}\circ\cc{\trg{L}}{\obj{L}}$}.
+\begin{definition}
+  \bul{$\cc{\src{L}}{\trg{L}}\circ\cc{\trg{L}}{\obj{L}}$} $\isdef$ Given $\src{p}$, yield \rul{$\cc{\trg{L}}{\obj{L}}\left(\cc{\src{L}}{\trg{L}}(\src{p})\right)$}
+\end{definition}
+
+Consider the compilation chain for $\src{TypeScript}$.
+First, $\src{TypeScript}$ programs are translated to $\trg{JavaScript}$ which, e.g., V8~\cite{googlev8} eventually compiles in parts to $\irl{Ignition BC}$.
+The following theorem establishes what happens if all these compilation steps were robustly secure with respect to \gls{ms}:
+The resulting $\irl{Ignition BC}$ code would be \gls{ms} regardless of the context the binary runs in.
+
+Given \bul{$\cc{\src{L}}{\trg{L}}$ robustly preserves $\class[_{1}]$} and \rul{$\cc{\trg{L}}{\obj{L}}$ robustly preserves $\class[_{2}]$}, it follows that \oul{their sequential composition $\cc{\src{L}}{\trg{L}}\circ\cc{\trg{L}}{\obj{L}}$ robustly preserves the intersection of classes $\class[_{1}]$ and $\class[_{2}]$}.
 
@@
 Since the composition of secure compilers is again a secure compiler, the theorem generalises to a whole chain of $n$ secure compilers.
-Pretending that the following compilers are secure, consider the compilation chain for $\src{TypeScript}$.
-First, $\src{TypeScript}$ programs are translated to $\trg{JavaScript}$ which, e.g., V8~\cite{googlev8} eventually compiles in parts to $\irl{Ignition BC}$.
-Were all these compilation steps robustly secure with respect to \gls{ms}, the resulting $\irl{Ignition BC}$ code would be \gls{ms} regardless of the context the binary runs in.
+It is worth pointing out that \gls{ms} (\Cref{subsubsec:ms}) is an example of a practically relevant intersection of classes, namely, \gls{tms} and \gls{sms} (\Cref{subsubsec:tms,subsubsec:sms}).
+However, the properties need to be lifted~\cite{clarkson2008hyper} to sets of properties by applying the powerset operation.
 
@@
     \begin{tikzpicture}
-      \node[minimum width=3cm,minimum height=2.5cm,draw=black,very thick,align=left] (progunopt) {\begin{lstlisting}[language=ml,basicstyle=\scriptsize\ttfamily]
+      \node[minimum width=3cm,minimum height=2.5cm,draw=black,very thick,align=left] (progunopt) {\begin{lstlisting}[language=ml,basicstyle=\small\ttfamily]
 let a = true in
@@
 \end{lstlisting}};
-      \node[minimum width=3cm,minimum height=2.5cm,draw=black,very thick,align=left,right=1.0 of progunopt] (progoptcf) {\begin{lstlisting}[language=ml,basicstyle=\scriptsize\ttfamily]
+      \node[minimum width=3cm,minimum height=2.5cm,draw=black,very thick,align=left,right=1.0 of progunopt] (progoptcf) {\begin{lstlisting}[language=ml,basicstyle=\small\ttfamily]
 if true then
@@
 \end{lstlisting}};
-      \node[minimum width=3cm,minimum height=2.5cm,draw=black,very thick,align=left,right=1.0 of progoptcf] (progoptcfdce) {\begin{lstlisting}[language=ml,basicstyle=\scriptsize\ttfamily]
+      \node[minimum width=3cm,minimum height=2.5cm,draw=black,very thick,align=left,right=1.0 of progoptcf] (progoptcfdce) {\begin{lstlisting}[language=ml,basicstyle=\small\ttfamily]
 print "a"
@@
 Besides sequential composition, there are two other compositions, namely an {\em upper}, i.e., a compiler that takes multiple inputs and yields one output, and a {\em lower} composition, i.e., a compiler that takes one input and yields multiple outputs.
+\bul{Define the upper composition $\cc{\src{L}+\obj{L}}{\trg{L}}$} as follows:
+Given a program \texttt{p}, its compiled counterpart is obtained by \rul{plugging \texttt{p} into $\cc{\src{L}}{\trg{L}}$ if $\texttt{p}\in\src{L}$} or by \oul{plugging \texttt{p} into $\cc{\obj{L}}{\trg{L}}$ if $\texttt{p}\in\obj{L}$}.
+\begin{definition}[Upper Composition]
+  \bul{$\cc{\src{L}+\obj{L}}{\trg{L}}$} $\isdef$ Given \texttt{p}, yield
+  $\left\{\mbox{\begin{tabular}{c}
+    {\rul{if $\texttt{p}\in\src{L}$, then $\cc{\src{L}}{\trg{L}}(\texttt{p})$}} \\
+    \mbox{\oul{if $\texttt{p}\in\obj{L}$, then $\cc{\obj{L}}{\trg{L}}(\texttt{p})$}} \\
+  \end{tabular}}\right.$
+%
+%
+%  $\begin{cases}
+%                                                                            \rul{\cc{\src{L}}{\trg{L}}(\texttt{p}) & \text{if }\texttt{p}\in\src{L}} \\
+%                                                                            \cc{\obj{L}}{\trg{L}}(\texttt{p}) & \text{if }\texttt{p}\in\obj{L}
+%                                                                          \end{cases}$
+\end{definition}
+
 Examples of this are present in industry:
@@
 A compiler that accepts both $\src{Java}$ and $\obj{Kotlin}$ code translating to the same target language or intermediate representation performs a kind of {\em upper} composition.
-The definition of upper composition is as follows: given a program $\mathtt{p}$, its compiled $\trg{L}$ counterpart is obtained by plugging $\mathtt{p}$ into $\cc{\src{L}}{\trg{L}}$ if $\mathtt{p}\in\src{L}$ or by plugging $\mathtt{p}$ into $\cc{\obj{L}}{\trg{L}}$ if $\mathtt{p}\in\obj{L}$, written as $\cc{\src{L}+\obj{L}}{\trg{L}}$.
-\MP{
-	put into def env
-}
-Now, given \bul{$\cc{\src{L}}{\trg{L}}$ robustly preserves $\class[_{1}]$} and \rul{$\cc{\obj{L}}{\trg{L}}$ robustly preserves $\class[_{2}]$}, it follows that \oul{their upper composition $\cc{\src{L}+\obj{L}}{\trg{L}}$ robustly preserves the intersection of classes $\class[_{1}]$ and $\class[_{2}]$}.
+Now, the following theorem tells us what happens if these are secure:
+Given \bul{$\cc{\src{L}}{\trg{L}}$ robustly preserves $\class[_{1}]$} and \rul{$\cc{\obj{L}}{\trg{L}}$ robustly preserves $\class[_{2}]$}, it follows that \oul{their upper composition $\cc{\src{L}+\obj{L}}{\trg{L}}$ robustly preserves the intersection of classes $\class[_{1}]$ and $\class[_{2}]$}.
 
@@
 Dually, the {\em lower} composition is concerned about compilers that accept the same source but yield different target languages. %and the same theoretical results apply.
+\bul{Define the lower composition $\cc{\src{L}}{\trg{L}+\obj{L}}$} as follows:
+Given a program $\src{p}$, its compiled counterpart is obtained by \rul{plugging $\src{p}$ into $\cc{\src{L}}{\trg{L}}$} or by \oul{plugging \texttt{p} into $\cc{\src{L}}{\obj{L}}$}, respectively, \pul{based on the internal decision}.
+\begin{definition}[Lower Composition]
+  \bul{$\cc{\src{L}}{\trg{L}+\obj{L}}$} $\isdef$ Given $L$ and $\src{p}$, yield
+  $\left\{\mbox{\begin{tabular}{c}
+    {\pul{if $L=\trg{L}$, then} \rul{$\cc{\src{L}}{\trg{L}}(\src{p})$}} \\
+    \mbox{\pul{if $L=\obj{L}$, then} \oul{$\cc{\src{L}}{\obj{L}}(\src{p})$}} \\
+  \end{tabular}}\right.$
+%
+%
+%  $\begin{cases}
+%                                                                            \rul{\cc{\src{L}}{\trg{L}}(\texttt{p}) & \text{if }\texttt{p}\in\src{L}} \\
+%                                                                            \cc{\obj{L}}{\trg{L}}(\texttt{p}) & \text{if }\texttt{p}\in\obj{L}
+%                                                                          \end{cases}$
+\end{definition}
+
 Consider two compilers both accepting $\src{LLVM IR}$~\cite{lattner2004llvm} and one of them emits $\trg{x86\_64}$, while the other emits $\trg{ARMv8}$.
@@
 For example, the user of this compiler provides an explicit flag that instructs to emit $\trg{x86\_64}$ or the framework itself detects the target platform via heuristics, such as supported instructions.
-The %proposed
-definition of lower composition $\cc{\src{L}}{\trg{L}+\obj{L}}$ is as follows: given a program $\src{p}$, its compiled $\trg{L}$ or $\obj{L}$ program is obtained by performing an internal decision and threading $\src{p}$ into $\cc{\src{L}}{\trg{L}}$ or into $\cc{\src{L}}{\obj{L}}$ depending on that decision.
-\MP{into def}
-Now, given \bul{$\cc{\src{L}}{\trg{L}}$ robustly preserves $\class[_{1}]$} and \rul{$\cc{\src{L}}{\obj{L}}$ robustly preserves $\class[_{2}]$}, it follows that \oul{their lower composition $\cc{\src{L}}{\trg{L}+\obj{L}}$ robustly preserves the intersection of classes $\class[_{1}]$ and $\class[_{2}]$}.
+The following theorem demonstrates what happens if the involved compilres are secure:
+Given \bul{$\cc{\src{L}}{\trg{L}}$ robustly preserves $\class[_{1}]$} and \rul{$\cc{\src{L}}{\obj{L}}$ robustly preserves $\class[_{2}]$}, it follows that \oul{their lower composition $\cc{\src{L}}{\trg{L}+\obj{L}}$ robustly preserves the intersection of classes $\class[_{1}]$ and $\class[_{2}]$}.
 

### 2023-07-11 popl/main.tex
Proof

@@
 \begin{example}[A program not satisfying \gls{tms}]
-  Consider the following \texttt{C++11} program that calls \texttt{strncpy} (\Cref{ex:strncpy}) and prints the result to the standard output stream.
+  Consider the following \texttt{C++11} library that calls \texttt{strncpy} (\Cref{ex:strncpy}) and prints the result to the standard output stream.
 
   \begin{lstlisting}[language=c++,basicstyle=\small\ttfamily,mathescape,commentstyle=\color{gray}]
-int main() {  // allocates 12 chars containing a greeting message
+int greet() {  // allocates 12 chars containing a greeting message
   char* greetings = new char[12] { "Hello POPL!" }; // <- address ${\color{gray}\loc_{x}}$
@@
   Even though the trace contains an out-of-bounds access right before returning from \texttt{strncpy}, this is no concern for \gls{tms}, since the location $\loc[_{x}]$ is still allocated.
-  However, having returned from \texttt{strncpy}, the \texttt{main} function continues and deallocates $\loc[_{y}]$ whose subsequent use in the \texttt{printf} call is a use-after-free bug.
+  However, having returned from \texttt{strncpy}, the \texttt{greet} function continues and deallocates $\loc[_{y}]$ whose subsequent use in the \texttt{printf} call is a use-after-free bug.
   The fix would be to \lstinline[language=c++,basicstyle=\small\ttfamily]|delete greetings| instead of \texttt{to} and add a \lstinline[language=c++,basicstyle=\small\ttfamily]|delete to| after the \texttt{printf} call, which leads to the abstract monitor trace $\abstrace[']=\absev{Alloc}\ \loc[_{x}]\ \comp\cdot\absev{Alloc}\ \loc[_{y}]\ \comp\cdot\absev{Use}\ \loc[_{x}]\ \comp\cdot\absev{Use}\ \loc[_{y}]\ \comp\cdot\ldots\cdot\absev{Use}\ \loc[_{x}]\ \comp\cdot\absev{Dealloc}\ \loc[_{x}]\ \comp\cdot\absev{Use}\ \loc[_{y}]\ \comp$.
@@
 \end{theorem}
-\MPin{
-This you can write the proof of.
-Probably even in coq
-%
-	\MKin{ 
-	Should this appear here in the paper? It's basically identical to \Cref{thm:ccab:rtp:ms}
-
-	\MPin{yes it should}}
-}
+\begin{proof}
+  Consequence of \Thmref{thm:ccscct:rtp:scct}, \Thmref{thm:cccfccdce:rtp:ms}, and \Thmref{thm:ccab:rtp:ms} using \Thmref{thm:rtp}.
+  \MKin{%
+    \textbackslash Thmref or \textbackslash Cref?
+  }
+% Unfold the definition of $\rtp{\cc{\Ltms}{\Lscct}}{\mssafe\cap\scctsafe}$.
+% So, let $\pi\in\mssafe\cap\scctsafe$, $\src{\progvar}$ such that $\rsat{\src{\progvar}}{\pi}$ and let $\obj{\contextvar},\obj{\runtimetermvar}$ with $\progstepto{\prog[\obj]{\obj{\contextvar}}{\left(\cc{\Ltms}{\Lscct}(\src{\progvar})\right)}}{\obj{\runtimetermvar}}{\obj{\trace}}$.
+% What is left to show is $\scctfilterLscct[\emptyset]{\obj{\trace}}\in\pi$.
+% Let $\irl{\progvar}=\cc{\Ltms}{\Ltrg}\circ\cc{\Ltrg}{\Lms}\circ\cc[\gls{cf}]{\Lms}{\Lms}\circ\cc[\gls{dce}]{\Lms}{\Lms}(\src{\progvar})$ such that $\cc{\Lms}{\Lscct}(\irl{\progvar})=\cc{\Ltms}{\Lscct}(\src{\progvar})$.
+% Note that $\pi\in\scctsafe$, so apply \Thmref{thm:ccscct:rtp:scct}
+\end{proof}
 
@@
 This paper only considers pointers to arrays of simple objects, i.e., there are no pointers to pointers or structures.
-Of course, there are a lot of other approaches to memory-safety in literature, specifically as compiler instrumentations~\cite{akritidis2009baggy,younan2010paricheck,jung2021pico,shankaranarayana2023tailcheck,dhumbumroong2020boundwarden,nam2019framer,zhou2023fatptrs}, hardware-extensions~\cite{kwon2013lowfat,saileshwar2022heapcheck,chen2023flexpointer,kim2023whistle}, or programming language extensions~\cite{elliott2018checkedc,li2022formalcheckedc,jim2002cyclone,elliott2015guilt,west2005cuckoo,weis2019fyr,benoit2019uniqueness}, to name a few.
+Of course, there are a lot of other approaches to memory-safety in literature, specifically as compiler instrumentations~\cite{akritidis2009baggy,younan2010paricheck,jung2021pico,shankaranarayana2023tailcheck,dhumbumroong2020boundwarden,nam2019framer,zhou2023fatptrs}, hardware-extensions~\cite{kwon2013lowfat,saileshwar2022heapcheck,chen2023flexpointer,kim2023whistle}, or programming language extensions~\cite{elliott2018checkedc,li2022formalcheckedc,necula2005ccured,jim2002cyclone,elliott2015guilt,west2005cuckoo,weis2019fyr,benoit2019uniqueness}, to name a few.
 What differentiates this work from them is that this work uses known, compiler-based approaches to ensure memory-safety as a means to investigate secure compiler compositions.

### 2023-07-11 popl/main.tex
ex 6.2

@@
 
-The compiler $\cc{\trg{L}}{\Lms}$ only inserts bounds-checks to enforce \gls{sms} whenever reading from or writing to memory.
+The compiler $\cc{\trg{L}}{\Lms}$ only inserts bounds-checks whenever reading from or writing to memory in order to enforce \gls{sms}.
+For passing pointers, it has to pass them with their size information as well.
 To this end, the compiler introduces another, fresh identifier $\irl{x_{SIZE}}$ for each allocation that binds $\irl{x}$ to keep track of the allocation size.
-Again, linking this back to the \texttt{strncpy} example (\Cref{ex:strncpy:sms}) corresponds to the insertion of bounds check right in front of the accesses to \texttt{y[i]} and \texttt{x[i]}.
-In turn, the event $\ev{Use}\ \loc_{x}\ 2;\comp;\unlock$ would not be emitted during execution.
-\MPin{
-	show the result of the compilation as listing, and then describe the trace a bit more in detail ?
-}
+\begin{example}[Instrumented \texttt{strncpy}]
+    Consider again \texttt{strncpy}, but this time with an instrumented version of \texttt{strncpy}:
+    \begin{lstlisting}[language=c,basicstyle=\ttfamily\small, morekeywords={size_t}]
+  void strncpy(size_t n, size_t dst_size, char *dst,
+                        size_t src_size, char *src) {
+    for(size_t i = 0; i < src_size && src[i] != '\0' && i < n; ++i) {
+      if(i < src_size && i < dst_size) {
+        dst[i] = src[i];
+      }
+    }
+  }
+    \end{lstlisting}
+    When calling this in similar fashion to \Cref{ex:strncpy:sms}, the event $\ev{Use}\ \loc_{x}\ 2;\comp;\unlock$ would not be emitted during execution.
+  \MKin{
+    this is kind of problematic for several reasons. (1) we don't pass pointers (2) fns signatures should not change
+  }
+\end{example}
 
@@
 
-\Cref{thm:ccab:rtp:ms}\MP{states ... } follows from \Cref{thm:cca:rtp:tms,thm:ccb:rtp:sms} using \Cref{thm:rtp}.
+%\Cref{thm:ccab:rtp:ms}\MP{states ... } follows from \Cref{thm:cca:rtp:tms,thm:ccb:rtp:sms} using \Cref{thm:rtp}.
 
@@
 \end{theorem}
+\begin{proof}
+  Consequence of \Thmref{thm:cca:rtp:tms} and \Thmref{thm:ccb:rtp:sms} using \Thmref{thm:rtp}.
+  \MKin{%
+    \textbackslash Thmref or \textbackslash Cref?
+  }
+\end{proof}
 
@@
 
-\begin{center}
-  $$
-  \begin{array}{rcl}
+\begin{gather*}
+  \begin{align*}
     \ccdce(\irl{\lifz{true}{\expr[_{1}]}{\expr[_{2}]}}) & = & \ccdce(\irl{\expr[_{1}]}) \\
-    \ccdce(\irl{\lifz{false}{\expr[_{1}]}{\expr[_{2}]}}) & = & \ccdce(\irl{\expr[_{2}]}) \\
-  \end{array}
-  $$
-\end{center}
-
-\begin{center}
-  $$
-  \begin{array}{rcll}
-    \cccf(\irl{\llet{x}{n}{\expr}}) & = & \cccf(\irl{\expr\subst{n}{x}}) &\\
-    \cccf(\irl{\lbinop{n_{1}}{n_{2}}}) & = & \irl{n_{3}} & \text{where }\irl{n_{3}}\text{ is }\lbinop{\irl{n_{1}}}{\irl{n_{2}}}\\
-  \end{array}
-  $$
-\end{center}
-\MPin{
-	substitution are a runtime semantics thing.
-	what you want is a compiler with 2 arguments, where in one you keep track of the static substitions to perform,
-	and 2 cases on how to handle variables: if they're in the second arg, then substitute, otherwise don't
-  %
-  \MKin{
-  MK: No?
-I get how they are used as part of the runtime semantics, but a compiler can just as well make use of a substitution function that models the runtime semantics.
-I see how these two are different.
-However, I think it's easier in this case to be a bit inprecise, since adding another argument to the compiler breaks our compositionality framework (unless it's hidden with a monad...) so, IMO, we should view substitution here as a meta-level, but still language-dependent, function that the compiler has access to.
-If we don't do this, we essentially re-encode the definition of subst into the compiler
-  }
-}
+    \ccdce(\irl{\lifz{false}{\expr[_{1}]}{\expr[_{2}]}}) & = & \ccdce(\irl{\expr[_{2}]})
+  \end{align*}\\[0.25cm]
+  \begin{align*}
+    \cccf(\irl{\llet{x}{n}{\expr}}) & = \cccf(\partialeval{\irl{\expr}}{\irl{x}}{\irl{n}}) &\\
+    \cccf(\irl{\lbinop{n_{1}}{n_{2}}}) & = \irl{n_{3}} & \text{where }\irl{n_{3}}\text{ is }\lbinop{\irl{n_{1}}}{\irl{n_{2}}}
+  \end{align*}\\[0.125cm]
+  \begin{align*}
+    \partialeval{\irl{x}}{\irl{x}}{\irl{n}} & = \irl{n} &\\
+    \partialeval{\irl{y}}{\irl{x}}{\irl{n}} & = \irl{y} &\text{where }\irl{x}\not=\irl{y}\\
+    \partialeval{\left(\irl{\llet{x}{\expr[_{1}]}{\expr[_{2}]}}\right)}{\irl{x}}{\irl{n}} & = \llet[\irl]{\irl{x}}{\left(\partialeval{\irl{\expr[_{1}]}}{\irl{x}}{\irl{n}}\right)}{\irl{\expr[_{2}]}} &\\
+    \partialeval{\left(\irl{\llet{y}{\expr[_{1}]}{\expr[_{2}]}}\right)}{\irl{x}}{\irl{n}} & = \llet[\irl]{\irl{y}}{\left(\partialeval{\irl{\expr[_{1}]}}{\irl{x}}{\irl{n}}\right)}{\left(\partialeval{\irl{\expr[_{2}]}}{\irl{x}}{\irl{n}}\right)} &\text{where }\irl{x}\not=\irl{y}\\
+    \partialeval{\left(\irl{\lbinop{\expr[_{1}]}{\expr[_{2}]}}\right)}{\irl{x}}{\irl{n}} & = \lbinop[\irl]{\left(\partialeval{\irl{\expr[_{1}]}}{\irl{x}}{\irl{n}}\right)}{\left(\partialeval{\irl{\expr[_{2}]}}{\irl{x}}{\irl{n}}\right)}
+  \end{align*}
+\end{gather*}
 
 The two optimising compiler passes from $\Lms$ to $\Lms$ perform \gls{dce} and \gls{cf}, respectively.
-Both are secure with respect to \gls{ms}, the proof of which is relatively simple, because \gls{dce} and \gls{cf} do not change the way memory accesses happen.
+The \gls{dce} pass applies a naive rewrite rule on conditionals.
+For \gls{cf}, the pass uses an auxiliary function \texttt{mix} that models the substitution operation.
+Both passes are secure with respect to \gls{ms}.
+The proof for either is relatively simple, because both \gls{dce} and \gls{cf} do not change the way memory accesses happen.
 Moreover, since the input and output languages to these compilers are the same, attacker contexts do not have more power in the target language than in the source.
@@
 \begin{theorem}[Compiler $\ccdce$ is secure with respect to \gls{ms}]\label{thm:ccdce:rtp:ms}
-  $\rtp{\ccdce}{\mssafe}$ \Coqed
+  $\rtp{\ccdce}{\mssafe}$ %\Coqed
 \end{theorem}
 \begin{theorem}[Compiler $\cccf$ is secure with respect to \gls{ms}]\label{thm:cccf:rtp:ms}
-  $\rtp{\cccf}{\mssafe}$ \Coqed
+  $\rtp{\cccf}{\mssafe}$ %\Coqe
 \end{theorem}
 
-With both \Cref{thm:ccdce:rtp:ms,thm:cccf:rtp:ms} it follows from \Cref{corr:swappable} that the two passes can be interchanged arbitrarily:
+%With both \Cref{thm:ccdce:rtp:ms,thm:cccf:rtp:ms} it follows from \Cref{corr:swappable} that the two passes can be interchanged arbitrarily:
 
 \begin{theorem}[Compilers $\cccf\circ\ccdce$ and $\cccf\circ\ccdce$ are secure with respect to \gls{ms}]\label{thm:cccfccdce:rtp:ms}
-  $\rtp{\cccf\circ\ccdce}{\mssafe}$ and $\rtp{\ccdce\circ\cccf}{\mssafe}$ % \Coqed
+  $\rtp{\cccf\circ\ccdce}{\mssafe}$ and $\rtp{\ccdce\circ\cccf}{\mssafe}$. % \Coqed
 \end{theorem}
+\begin{proof}
+  Consequence of \Thmref{thm:cccf:rtp:ms} and \Thmref{thm:ccdce:rtp:ms} using \Thmref{corr:swappable}.
+  \MKin{%
+    \textbackslash Thmref or \textbackslash Cref?
+  }
+\end{proof}
 

### 2023-07-11 popl/main.tex
format

--- 2023-07-11 popl/main.tex
@@
 \begin{definition}[\glsfirst{sms}]\label{def:trace:smsdef}
-  $
+  $$
   \smssafe:=\left\{\trace \left|\begin{array}{rcl}
@@
                                 \end{array}\right.\right\}
-  $
+  $$
 \end{definition}
@@
 \begin{definition}[\glsfirst{ms}]\label{def:trace:msdef}
-  $
+  $$
   \mssafe:=\left\{\trace \left|\begin{array}{rcl}
@@
                                \end{array}\right.\right\}
-  $
+  $$
 \end{definition}
@@
 \begin{definition}[\glsfirst{scct}]\label{def:trace:scctdef}
-  $
+  $$
   \scctsafe:=\left\{\trace \left|\begin{array}{rcl}
@@
                                  \end{array}\right.\right\}
-  $
+  $$
 \end{definition}
@@
 \begin{definition}[\glsfirst{ms} and \gls{scct}]\label{def:trace:msscctdef}
-  $
+  $$
   \msscctsafe:=\mssafe\cap\scctsafe
-  $
+  $$
 \end{definition}
+\MPin{
+	why is this defined differently than MS? they should be defined in the same way
+}
 
@@
 \subsection{Monitors\pages{1.5}}\label{subsec:monitors}
-
-\begin{figure}[h!]
-  \judgbox{\traceagree{\pi}{\trace}{\abstrace}}{,,Specification-level trace $\trace$ agrees with monitor-level trace $\abstrace$ with respect to property $\pi$.''}
-  %
-  \typerule{traceagree-empty}{}{
-    \traceagree{\pi}{\hole{\cdot}}{\hole{\cdot}}
-  }{traceagree-empty}
-  %
-  \typerule{traceagree-ign-L}{
-    \traceagree{\pi}{\trace}{\abstrace}
-  }{
-    \traceagree{\pi}{\emptyevent\cdot\trace}{\abstrace}
-  }{traceagree-ign-L}
-  %
-  \typerule{traceagree-ign-R}{
-    \traceagree{\pi}{\trace}{\abstrace}
-  }{
-    \traceagree{\pi}{\trace}{\absev{\emptyevent}\cdot\abstrace}
-  }{traceagree-ign-R}
-  %
-  \typerule{traceagree-cons}{
-    \traceagree[]{\pi}{\event}{\absevent} &
-    \traceagree{\pi}{\trace}{\abstrace}
-  }{
-    \traceagree{\pi}{\event\cdot\trace}{\absevent\cdot\abstrace}
-  }{traceagree-cons}
-  \caption{Trace-Agreement relation that equates specification-level traces with monitor-level traces.}\label{fig:stuttereq}
-\end{figure}
 
@@
 \Cref{fig:stuttereq} shows how the event agreement is lifted to traces.
+\begin{figure}[!ht]
+  \judgbox{\traceagree{\pi}{\trace}{\abstrace}}{,,Specification-level trace $\trace$ agrees with monitor-level trace $\abstrace$ with respect to property $\pi$.''}
+  %
+  \typerule{traceagree-empty}{}{
+    \traceagree{\pi}{\hole{\cdot}}{\hole{\cdot}}
+  }{traceagree-empty}
+  %
+  \typerule{traceagree-ign-L}{
+    \traceagree{\pi}{\trace}{\abstrace}
+  }{
+    \traceagree{\pi}{\emptyevent\cdot\trace}{\abstrace}
+  }{traceagree-ign-L}
+  %
+  \typerule{traceagree-ign-R}{
+    \traceagree{\pi}{\trace}{\abstrace}
+  }{
+    \traceagree{\pi}{\trace}{\absev{\emptyevent}\cdot\abstrace}
+  }{traceagree-ign-R}
+  %
+  \typerule{traceagree-cons}{
+    \traceagree[]{\pi}{\event}{\absevent} &
+    \traceagree{\pi}{\trace}{\abstrace}
+  }{
+    \traceagree{\pi}{\event\cdot\trace}{\absevent\cdot\abstrace}
+  }{traceagree-cons}
+  \caption{Trace-Agreement relation that equates specification-level traces with monitor-level traces.}\label{fig:stuttereq}
+\end{figure}
 The trace agreement is the same for all properties $\pi$ up to the use of the event agreement in \Cref{tr:traceagree-cons}.
@@
 \begin{definition}[Upper Composition]
-  \bul{$\cc{\src{L}+\obj{L}}{\trg{L}}$} $\isdef$ Given \texttt{p}, yield
+  \bul{$\cc{\src{L}+\obj{L}}{\trg{L}}$} $\isdef$ 
+  % Given \texttt{p}, yield
+  $\lambda \texttt{p}\ldotp$
   $\left\{\mbox{\begin{tabular}{c}
@@
 \begin{definition}[Lower Composition]
-  \bul{$\cc{\src{L}}{\trg{L}+\obj{L}}$} $\isdef$ Given $L$ and $\src{p}$, yield
+  \bul{$\cc{\src{L}}{\trg{L}+\obj{L}}$} $\isdef$ 
+  % Given $L$ and $\src{p}$, yield
+  $\lambda \src{p}, L\ldotp$
   $\left\{\mbox{\begin{tabular}{c}
@@
 
-This section compares existing work with the one presented in this paper.
-To this end, work on robust preservation (\Cref{subsec:relw:seccomprtp}) and on other secure compiler criteria (\Cref{subsec:relw:seccompcrit}) is looked at.
-The case study of this paper (\Cref{sec:casestud:defs,sec:casestud:rtp}) implements measures for ensuring \gls{ms} as well as \gls{cct} and these are compared with previous work as well (\Cref{subsec:relw:msmechs,subsec:relw:cctmechs}).
+This section discusses work on robust preservation\MP{
+	robust compilation.
+	NOT robust preservation
+} (\Cref{subsec:relw:seccomprtp}) and on other secure compilation criteria (\Cref{subsec:relw:seccompcrit}).
+Since the case study of \Cref{sec:casestud:defs,sec:casestud:rtp} implements measures for preserving \gls{ms} and \gls{cct}, this section then presents relevant related work as well (\Cref{subsec:relw:msmechs,subsec:relw:cctmechs}).
 
@@
 
-The robust preservation of properties as a compiler-level criterion has been analyzed extensively~\cite{abate2019jour,patrignani2021rsc,abate2021extacc,patrignani2022universal,patrignani2019survey}.
-This work has made use of that previous work.
-While there exists a composition theorem~\cite{abate2019jour} already, that theorem is not concerned with composing robustly safe compilers, but rather program components and contexts.
-The work relating robust preservation with universal composability~\cite{patrignani2022universal} is closest to what this paper presents.
+The robust preservation of properties as a compiler-level criterion has been analyzed extensively~\cite{abate2019jour,patrignani2021rsc,abate2021extacc,patrignani2019survey} and thus we build on that framework.
+No existing work is concerned with composing robustly safe compilers.
+\MP{perhaps say taht we believe this scales to different trace models too and differnet kinds of props, but the ocmposition of monitors becomes tricky}
+The work relating robust preservation with universal composability~\cite{patrignani2022universal} is closest to what this paper presents.\MP{why}
 % The authors demonstrate a similar compositionality theorem to what is presented here (\Cref{sec:compcomp}) as well as in an earlier version of this work~\cite{kruse2022csc}.
@@
 The survey on formal approaches to secure compilation~\cite{patrignani2019survey} discusses a broad spectrum already, while this section presents a very high-level overview.
-The full abstraction approach~\cite{abadi1999fullabstraction} states that a compiler should preserve and reflect observational equivalence between source and target programs.
+Fully abstract compilation~\cite{abadi1999fullabstraction} states that a compiler should preserve and reflect observational equivalence between source and target programs.
 It was shown~\cite{abate2021faandrc} that fully abstract compilers robustly preserve program properties that are either trivial or meaningless.
@@
 
-Different mechanisms for memory safety exist that also consider the secure compilation domain, i.e., have an active attacker model.
+Different mechanisms for \MP{enforcing?}memory safety exist that also consider the secure compilation domain, i.e., have an active attacker model.
 For example, the ,,pointers as capabilities'' principle represents pointers as machine-level capabilities~\cite{korashy2021capableptrs}, which behave in a similar fashion to capabilities by means of linear typing~\cite{morrisett2005L3}.
@@
 That work inserts bounds-checks in front of pointer-dereferences and, for this to work, inserts meta-data information on pointer creation.
-It also works in a more advanced setting with structured fields accesses and also introduces a table-lookup for pointers that are stored in memory.
-This paper only considers pointers to arrays of simple objects, i.e., there are no pointers to pointers or structures.
-Of course, there are a lot of other approaches to memory-safety in literature, specifically as compiler instrumentations~\cite{akritidis2009baggy,younan2010paricheck,jung2021pico,shankaranarayana2023tailcheck,dhumbumroong2020boundwarden,nam2019framer,zhou2023fatptrs}, hardware-extensions~\cite{kwon2013lowfat,saileshwar2022heapcheck,chen2023flexpointer,kim2023whistle}, or programming language extensions~\cite{elliott2018checkedc,li2022formalcheckedc,jim2002cyclone,elliott2015guilt,west2005cuckoo,weis2019fyr,benoit2019uniqueness}, to name a few.
+It\MP{what} also works in a more advanced setting with structured fields accesses and also introduces a table-lookup for pointers that are stored in memory.
+This paper only considers pointers to arrays of simple objects\MP{NOT OBJECTS}, i.e., there are no pointers to pointers or structures.
+Several other approaches to memory-safety exist in literature, specifically as compiler instrumentations~\cite{akritidis2009baggy,younan2010paricheck,jung2021pico,shankaranarayana2023tailcheck,dhumbumroong2020boundwarden,nam2019framer,zhou2023fatptrs}, hardware-extensions~\cite{kwon2013lowfat,saileshwar2022heapcheck,chen2023flexpointer,kim2023whistle}, or programming language extensions~\cite{elliott2018checkedc,li2022formalcheckedc,jim2002cyclone,elliott2015guilt,west2005cuckoo,weis2019fyr,benoit2019uniqueness}.
 What differentiates this work from them is that this work uses known, compiler-based approaches to ensure memory-safety as a means to investigate secure compiler compositions.
@@
 
-The approach to cryptographic constant time in this paper is high-level, where a programming language exposes a way to switch the semantics to a data (operand) independent timing mode.
+The approach to \MP{preserving?enforcing?} cryptographic constant time in this paper is high-level, where a programming language exposes a way to switch the semantics to a data (operand) independent timing mode.
 Since identifiers in $\Lscct$ are annotated with a secrecy tag, this approach is similar to others with information flow control.
@@
 The paper exercises this framework on a case-study that consists of an optimising compilation chain that is secure with respect to \gls{msscct}.
+\MP{
+	not an appropriate recap.
+	rewrite summarising the paper
+}
+
 In future work, it would be interesting to investigate whether it is possible to provide {\em secure compiler combinators}, similarly to how parser combinators work.
@@
 This way, it is possible to obtain an executable, secure compiler.
-However, the formalisation effort for backtranslation proofs is known to be enormous, so another research avenue may be to find better ways to (semi-)automate standard secure compilation proofs.
+However, the formalisation effort for backtranslation proofs is known to be enormous\MP{cit + details}, so another research avenue may be to find better ways to (semi-)automate standard secure compilation proofs.
+\MPin{
+	not particularly informative FW. i'd rather not have it as such
+}
 
@@
 %% consistent spelling of the heading.
+
 % \begin{acks}
 %   We would like to thank the anonymous reviewers for their feedback.
+% 	This work was partially supported by a gift from
+		% the Italian Ministry of Education through funding for the Rita Levi Montalcini grant (call of 2019).
 % \end{acks}
 
+\newpage
 %%

### 2023-07-11 popl/main.tex

--- 2023-07-11 popl/main.tex
@@
 That work inserts bounds-checks in front of pointer-dereferences and, for this to work, inserts meta-data information on pointer creation.
-It also works in a more advanced setting with structured fields accesses and also introduces a table-lookup for pointers that are stored in memory.
-This paper only considers pointers to arrays of simple objects, i.e., there are no pointers to pointers or structures.
-Of course, there are a lot of other approaches to memory-safety in literature, specifically as compiler instrumentations~\cite{akritidis2009baggy,younan2010paricheck,jung2021pico,shankaranarayana2023tailcheck,dhumbumroong2020boundwarden,nam2019framer,zhou2023fatptrs}, hardware-extensions~\cite{kwon2013lowfat,saileshwar2022heapcheck,chen2023flexpointer,kim2023whistle}, or programming language extensions~\cite{elliott2018checkedc,li2022formalcheckedc,necula2005ccured,jim2002cyclone,elliott2015guilt,west2005cuckoo,weis2019fyr,benoit2019uniqueness}, to name a few.
+It\MP{what} also works in a more advanced setting with structured fields accesses and also introduces a table-lookup for pointers that are stored in memory.
+This paper only considers pointers to arrays of simple objects\MP{NOT OBJECTS}, i.e., there are no pointers to pointers or structures.
+Several other approaches to memory-safety exist in literature, specifically as compiler instrumentations~\cite{akritidis2009baggy,younan2010paricheck,jung2021pico,shankaranarayana2023tailcheck,dhumbumroong2020boundwarden,nam2019framer,zhou2023fatptrs}, hardware-extensions~\cite{kwon2013lowfat,saileshwar2022heapcheck,chen2023flexpointer,kim2023whistle}, or programming language extensions~\cite{elliott2018checkedc,li2022formalcheckedc,jim2002cyclone,elliott2015guilt,west2005cuckoo,weis2019fyr,benoit2019uniqueness}.
 What differentiates this work from them is that this work uses known, compiler-based approaches to ensure memory-safety as a means to investigate secure compiler compositions.

### 2023-07-11 popl/main.tex
Make fig thicc

--- 2023-07-11 popl/main.tex
@@
   \begin{center}
-    \begin{tikzpicture}[state/.style={minimum height=0.6cm}]
+    \begin{tikzpicture}[state/.style={minimum height=1.5cm}]
       % relative horizontal/vertical distance between states
@@
       % row of src states
-      \node[state] (srcempty) {$\src{\emptyset}$};
+      \node[state] (srcempty) {$\rtt{\src{\emptyset}}{\src{\lcall{main}{0}}}$};
       \foreach \s [remember=\s as \cur (initially empty)] in {1,w_1,p,w_2,2} {
-        \node[state,right=\hdist of src\cur] (src\s) {$\src{\Omega_{\s}}$};
+        \node[state,right=\hdist of src\cur] (src\s) {$\rtt{\src{\Omega_{\s}}}{\src{\expr[_{\s}]}}$};
       }
       % row of trg states
-      \node[state,below=\vdist of srcempty] (trgempty) {$\trg{\emptyset}$};
+      \node[state,below=\vdist of srcempty] (trgempty) {$\rtt{\trg{\emptyset}}{\trg{\lcall{main}{0}}}$};
       \foreach \s in {1,w_1,p,w_2,2} {
-        \node[state,below=\vdist of src\s] (trg\s) {$\trg{\Omega_{\s}}$};
+        \node[state,below=\vdist of src\s] (trg\s) {$\rtt{\trg{\Omega_{\s}}}{\trg{\expr[_{\s}]}}$};
       }
@@
         % backtrans wrapper 1
-        \draw[thick,loosely dotted,Peach!50,rounded corners] (src1.north east) -- (srcw\string_1.north east)
+        \draw[ultra thick,loosely dotted,Peach!50,rounded corners] (src1.north east) -- (srcw\string_1.north east)
           -- (trgw\string_1.south east) -- (trg1.south west) -- (src1.north west) -- cycle;
         % backtrans wrapper 2
-        \draw[thick,loosely dotted,Peach!50,rounded corners] (srcp.north east) -- (srcw\string_2.north east)
+        \draw[ultra thick,loosely dotted,Peach!50,rounded corners] (srcp.north east) -- (srcw\string_2.north east)
           -- (trgw\string_2.south east) -- (trgp.south west) -- (srcp.north west) -- cycle;
         % compiler correctness
-        \draw[thick,loosely dashed,Emerald!50,rounded corners] ($(srcw\string_1.north east)+(0,0.05)$) -- ($(srcp.north east)+(0,0.05)$)
+        \draw[ultra thick,loosely dashed,Emerald!50,rounded corners] ($(srcw\string_1.north east)+(0,0.05)$) -- ($(srcp.north east)+(0,0.05)$)
           -- ($(trgw\string_2.north east)+(0.05,0)$) -- ($(trgw\string_2.south east)+(0.05,-0.05)$)
@@
       % state relations
-      \path (srcempty) edge[draw=gray!25] node[pos=0.5,sloped,rotate=180,fill=white] {\scriptsize$\multimap_\emptyset$} (trgempty)
-        (src1) edge[draw=gray!25] node[pos=0.5,sloped,rotate=180,fill=white] {\scriptsize$\multimap_{\delta_1}$} (trg1)
-        (srcp) edge[draw=gray!25] node[pos=0.5,sloped,fill=white] {\scriptsize$\approx_{\delta_p}$} (trgp)
-        (src2) edge[draw=gray!25] node[pos=0.5,sloped,rotate=180,fill=white] {\scriptsize$\multimap_{\delta_2}$} (trg2)
-        (srcw\string_1) edge[draw=gray!25] node[pos=0.5,sloped,fill=white] {\scriptsize$\approx_{\delta_{w_1}}$} (trgw\string_1)
-        (srcw\string_2) edge[draw=gray!25] node[pos=0.5,sloped,rotate=180,fill=white] {\scriptsize$\multimap_{\delta_{w_2}}$} (trgw\string_2)
+      \path (srcempty) edge[very thick,draw=gray!25] node[pos=0.5,sloped,rotate=180,fill=white] {\scriptsize$\multimap_\emptyset$} (trgempty)
+        (src1) edge[very thick,draw=gray!25] node[pos=0.5,sloped,rotate=180,fill=white] {\scriptsize$\multimap_{\delta_1}$} (trg1)
+        (srcp) edge[very thick,draw=gray!25] node[pos=0.5,sloped,fill=white] {\scriptsize$\approx_{\delta_p}$} (trgp)
+        (src2) edge[very thick,draw=gray!25] node[pos=0.5,sloped,rotate=180,fill=white] {\scriptsize$\multimap_{\delta_2}$} (trg2)
+        (srcw\string_1) edge[very thick,draw=gray!25] node[pos=0.5,sloped,fill=white] {\scriptsize$\approx_{\delta_{w_1}}$} (trgw\string_1)
+        (srcw\string_2) edge[very thick,draw=gray!25] node[pos=0.5,sloped,rotate=180,fill=white] {\scriptsize$\multimap_{\delta_{w_2}}$} (trgw\string_2)
         % diagonals
-        (srcw\string_1) edge[draw=gray!25] node[pos=0.16,sloped,rotate=180,fill=white] {\scriptsize$\approx_{\delta_{w_1}}$} (trg1)
-        (src1) edge[draw=gray!25] node[pos=0.16,sloped,rotate=180,fill=white] {\scriptsize$\multimap_{\delta_1}$} (trgw\string_1)
-        (srcp) edge[draw=gray!25] node[pos=0.2,sloped,fill=white] {\scriptsize$\approx_{\delta_p}$} (trgw\string_2)
-        (trgp) edge[draw=gray!25] node[pos=0.2,sloped,fill=white] {\scriptsize$\multimap_{\delta_{w_2}}$} (srcw\string_2)
+        (srcw\string_1) edge[thick,draw=gray!25] node[pos=0.16,sloped,rotate=180,fill=white] {\scriptsize$\approx_{\delta_{w_1}}$} (trg1)
+        (src1) edge[thick,draw=gray!25] node[pos=0.16,sloped,rotate=180,fill=white] {\scriptsize$\multimap_{\delta_1}$} (trgw\string_1)
+        (srcp) edge[thick,draw=gray!25] node[pos=0.2,sloped,fill=white] {\scriptsize$\approx_{\delta_p}$} (trgw\string_2)
+        (trgp) edge[thick,draw=gray!25] node[pos=0.2,sloped,fill=white] {\scriptsize$\multimap_{\delta_{w_2}}$} (srcw\string_2)
         ;
@@
       % steps
-      \path[color=\stlccol] (srcempty) edge[draw=none] node {\ $\xrightarrow{\trace[_1]}{}{\kern-3.5pt}_{\text{ctx}}^*$} (src1)
+      \path[very thick,color=\stlccol] (srcempty) edge[draw=none] node {\ $\xrightarrow{\trace[_1]}{}{\kern-3.5pt}_{\text{ctx}}^*$} (src1)
         (src1) edge[draw=none] node {\ $\xrightarrow{\trace[_c]}{}{\kern-3.5pt}_{\text{ctx}}$} (srcw\string_1)
@@
         ;
-      \path[color=\ulccol] (trgempty) edge[draw=none] node[fill=white,inner sep=0,outer sep=0] {\ $\xrightarrow{\phantom{\trace[_1]}}{}{\kern-3.5pt}_{\text{ctx}}^*$} (trg1)
+      \path[very thick,color=\ulccol] (trgempty) edge[draw=none] node[fill=white,inner sep=0,outer sep=0] {\ $\xrightarrow{\phantom{\trace[_1]}}{}{\kern-3.5pt}_{\text{ctx}}^*$} (trg1)
         (trg1) edge[draw=none] node {\ $\xrightarrow{\phantom{\trace[_p]}}{}{\kern-3.5pt}_{\text{ctx}}^*$} (trgw\string_1)
         (trgw\string_1) edge[draw=none] node[fill=white,inner sep=0,outer sep=0] {\ $\xrightarrow{\phantom{\trace[_p]}}{}{\kern-3.5pt}_{\text{ctx}}^*$} (trgp)
-        (trgp) edge[draw=none] node {\ $\xrightarrow{\phantom{\trace[_p]}}{}{\kern-3.5pt}_{\text{ctx}}^*$} (trgw\string_2)
+        (trgp) edge[very thick,draw=none] node {\ $\xrightarrow{\phantom{\trace[_p]}}{}{\kern-3.5pt}_{\text{ctx}}^*$} (trgw\string_2)
         (trgw\string_2) edge[draw=none] node {\ $\xrightarrow{\phantom{\trace[_2]}}{}{\kern-3.5pt}_{\text{ctx}}^*$} (trg2)
@@
         };
-      \draw[thick,loosely dotted,Peach!50,rounded corners] ($(legend.north east)+(0.5,-0.4)$) -- ($(legend.north east)+(1,-0.4)$);
+      \draw[ultra thick,loosely dotted,Peach!50,rounded corners] ($(legend.north east)+(0.5,-0.4)$) -- ($(legend.north east)+(1,-0.4)$);
       \node at ($(legend.north east)+(1.5,-0.4)$) (legendwrapper) {};
       \node[right of=legendwrapper] {\tiny Backtranslation Wrapper};
-      \draw[thick,loosely dashed,Emerald!50,rounded corners] ($(legend.south east)+(0.5,0.4)$) -- ($(legend.south east)+(1,0.4)$);
+      \draw[ultra thick,loosely dashed,Emerald!50,rounded corners] ($(legend.south east)+(0.5,0.4)$) -- ($(legend.south east)+(1,0.4)$);
       \node at ($(legend.south east)+(1.5,0.4)$) (legendcorrectness) {};
@@
     \end{tikzpicture}
-    \caption{Proof diagram for \Cref{thm:cca:rtp:tms} depicting the general structure of robust preservation proofs.}\label{fig:proofdiag:rtp}
+    \caption{Proof diagram for \Cref{thm:cca:rtp:tms} depicting the general structure of robust preservation proofs. %
+      Vertical lines indicate cross language relations, while horizontal ones are execution steps. %
+      Nodes in the graph represent runtime states.
+    }\label{fig:proofdiag:rtp}
   \end{center}
 \end{figure}
-\MPin{
-	thicker lines and dots needed, not visible otherwise
-}
 

### 2023-07-11 popl/main.tex
absrtact

--- 2023-07-11 popl/main.tex
@@
 \begin{abstract}
-
-  \MKin{context}
-  Memory safety necessitates secrecy, since breaking it can lead to an arbitrarily large attack space, i.e., from simply reading a secret up to alteration of control-flow in order to execute arbitrary code.
-  However, secrecy cannot be achieved with just memory safety alone, since, e.g., private data could be leaked by differences in execution time for varying program inputs.
-  This leakage is avoidable by ensuring that programs are cryptographic constant time, i.e., different inputs do not change execution time.
-  Prior work has shown that strategies to ensure cryptographic constant time can be rendered useless, since a compiler may simply optimize them away.
-  \MKin{need}
-  This is why secure compilation is an integral ingredient to achieve security, because a secure compiler guarantees (formally) that properties that hold in the source, also hold in the target.
+% context
+To ensure that secure applications do not leak their secrets, they are required to uphold several security properties such as spatial and temporal memory safety as well as criptographic constant time.
+% need
+Existing work shows how to enforce these properties individually, in an architecture-independent way, by using secure compiler passes that each focus on an individual property.
+% task
+Unfortunately, given two secure compiler passes that each preserve a possibly-different security property, there is no way to tell what kind of security property will the composition of those secure compilers preserve.
+
+% object
+This paper is the first to study what security properties are preserved across the composition of different secure compiler passes.
+% findings
+Starting from a general theory of property composition for security-relevant properties (such as the aforementioned ones), this paper formalises a secure multi-pass compiler that preserves those security-relevant properties.
+% conclusion
+Crucially, this paper derives the security of the multi-pass compiler from the composition of the security properties preserved by its individual passes, which include security-preserving as well as optimisation passes.
+% 
+From an engineering perspective, this is the desirable approach to building secure compilers.
+
+  % \MKin{context}
+  % Memory safety necessitates secrecy, since breaking it can lead to an arbitrarily large attack space, i.e., from simply reading a secret up to alteration of control-flow in order to execute arbitrary code.
+  % However, secrecy cannot be achieved with just memory safety alone, since, e.g., private data could be leaked by differences in execution time for varying program inputs.
+  % This leakage is avoidable by ensuring that programs are cryptographic constant time, i.e., different inputs do not change execution time.
+  % Prior work has shown that strategies to ensure cryptographic constant time can be rendered useless, since a compiler may simply optimize them away.
+  % \MKin{need}
+  % This is why secure compilation is an integral ingredient to achieve security, because a secure compiler guarantees (formally) that properties that hold in the source, also hold in the target.
 %% mention attacker?
 the text describes correctness
-  \MKin{task}
-  Unfortunately, it is an open question whether secure compilers can be engineered in a modular way.
-  Ideally, compiler engineers develop two different compilation passes, one that ensures memory safety and another one that is security preserving with respect to cryptographic constant time.
+  % \MKin{task}
+  % Unfortunately, it is an open question whether secure compilers can be engineered in a modular way.
+  % Ideally, compiler engineers develop two different compilation passes, one that ensures memory safety and another one that is security preserving with respect to cryptographic constant time.
 %, because this style of engineering is well-established for its simplifying divide-and-conquer approach as well as enhanced reusability.
-  \MKin{object+findings}
-  This paper answers that question positively: The composition of compilers that are secure with respect to some security-properties yields the intersection of these properties.
-  \MKin{object}
-  This article discusses different variants of compositions of secure compilers and instantiates parts of these results in a case-study.
-  To this end, it presents an optimizing, secure compilation chain that preserves both memory safety and cryptographic constant time.
-  \MKin{conclusion}
-  The compositionality results enable modular proofs for the presented, secure compiler. % that is secure with respect to both memory safety and cryptographic constant time.
-  Some results of this work are implemented in the Coq proof assistant.
+  % \MKin{object+findings}
+  % This paper answers that question positively: The composition of compilers that are secure with respect to some security-properties yields the intersection of these properties.
+  % \MKin{object}
+  % This article discusses different variants of compositions of secure compilers and instantiates parts of these results in a case-study.
+  % To this end, it presents an optimizing, secure compilation chain that preserves both memory safety and cryptographic constant time.
+  % \MKin{conclusion}
+  % The compositionality results enable modular proofs for the presented, secure compiler. % that is secure with respect to both memory safety and cryptographic constant time.
+  % Some results of this work are implemented in the Coq proof assistant.
 

### 2023-07-11 popl/main.tex
proof

@@
 \begin{proof}
-  Consequence of \Thmref{thm:ccscct:rtp:scct}, \Thmref{thm:cccfccdce:rtp:ms}, and \Thmref{thm:ccab:rtp:ms} using \Thmref{thm:rtp}.
-  \MKin{%
-    \textbackslash Thmref or \textbackslash Cref?
-  }
+	From \Thmref{thm:ccscct:rtp:scct}, we have that any LTMS program \src{p} gets compiled into a L program \trg{p} that is also TMS.
+	Then, from \Thmref{thm:cccfccdce:rtp:ms} we have that \trg{p} gets compilet to a program ?? that is also SMS, assuming it was SMS to begin with.
+	Since TMS + SMS = MS, ?? is MS.
+	?? gets optimised by the optimisation passes and its optimisation is ???, which is still MS.
+	Finally, \Thmref{thm:ccab:rtp:ms} ensures that ??? gets compiled into ????, which is CCT assuming that the original program was CCT to begin with.
+  % \MKin{%
+  %   \textbackslash Thmref or \textbackslash Cref?
+  % }
 % Unfold the definition of $\rtp{\cc{\Ltms}{\Lscct}}{\mssafe\cap\scctsafe}$.

### 2023-07-11 popl/main.tex
relwork

--- 2023-07-11 popl/main.tex
@@
 \begin{abstract}
-% context
-To ensure that secure applications do not leak their secrets, they are required to uphold several security properties such as spatial and temporal memory safety as well as criptographic constant time.
-% need
-Existing work shows how to enforce these properties individually, in an architecture-independent way, by using secure compiler passes that each focus on an individual property.
-% task
-Unfortunately, given two secure compiler passes that each preserve a possibly-different security property, there is no way to tell what kind of security property will the composition of those secure compilers preserve.
-
-% object
-This paper is the first to study what security properties are preserved across the composition of different secure compiler passes.
-% findings
-Starting from a general theory of property composition for security-relevant properties (such as the aforementioned ones), this paper formalises a secure multi-pass compiler that preserves those security-relevant properties.
-% conclusion
-Crucially, this paper derives the security of the multi-pass compiler from the composition of the security properties preserved by its individual passes, which include security-preserving as well as optimisation passes.
-% 
-From an engineering perspective, this is the desirable approach to building secure compilers.
-
-  % \MKin{context}
-  % Memory safety necessitates secrecy, since breaking it can lead to an arbitrarily large attack space, i.e., from simply reading a secret up to alteration of control-flow in order to execute arbitrary code.
-  % However, secrecy cannot be achieved with just memory safety alone, since, e.g., private data could be leaked by differences in execution time for varying program inputs.
-  % This leakage is avoidable by ensuring that programs are cryptographic constant time, i.e., different inputs do not change execution time.
-  % Prior work has shown that strategies to ensure cryptographic constant time can be rendered useless, since a compiler may simply optimize them away.
-  % \MKin{need}
-  % This is why secure compilation is an integral ingredient to achieve security, because a secure compiler guarantees (formally) that properties that hold in the source, also hold in the target.
+
+  \MKin{context}
+  Memory safety necessitates secrecy, since breaking it can lead to an arbitrarily large attack space, i.e., from simply reading a secret up to alteration of control-flow in order to execute arbitrary code.
+  However, secrecy cannot be achieved with just memory safety alone, since, e.g., private data could be leaked by differences in execution time for varying program inputs.
+  This leakage is avoidable by ensuring that programs are cryptographic constant time, i.e., different inputs do not change execution time.
+  Prior work has shown that strategies to ensure cryptographic constant time can be rendered useless, since a compiler may simply optimize them away.
+  \MKin{need}
+  This is why secure compilation is an integral ingredient to achieve security, because a secure compiler guarantees (formally) that properties that hold in the source, also hold in the target.
 %% mention attacker?
 the text describes correctness
-  % \MKin{task}
-  % Unfortunately, it is an open question whether secure compilers can be engineered in a modular way.
-  % Ideally, compiler engineers develop two different compilation passes, one that ensures memory safety and another one that is security preserving with respect to cryptographic constant time.
+  \MKin{task}
+  Unfortunately, it is an open question whether secure compilers can be engineered in a modular way.
+  Ideally, compiler engineers develop two different compilation passes, one that ensures memory safety and another one that is security preserving with respect to cryptographic constant time.
 %, because this style of engineering is well-established for its simplifying divide-and-conquer approach as well as enhanced reusability.
-  % \MKin{object+findings}
-  % This paper answers that question positively: The composition of compilers that are secure with respect to some security-properties yields the intersection of these properties.
-  % \MKin{object}
-  % This article discusses different variants of compositions of secure compilers and instantiates parts of these results in a case-study.
-  % To this end, it presents an optimizing, secure compilation chain that preserves both memory safety and cryptographic constant time.
-  % \MKin{conclusion}
-  % The compositionality results enable modular proofs for the presented, secure compiler. % that is secure with respect to both memory safety and cryptographic constant time.
-  % Some results of this work are implemented in the Coq proof assistant.
+  \MKin{object+findings}
+  This paper answers that question positively: The composition of compilers that are secure with respect to some security-properties yields the intersection of these properties.
+  \MKin{object}
+  This article discusses different variants of compositions of secure compilers and instantiates parts of these results in a case-study.
+  To this end, it presents an optimizing, secure compilation chain that preserves both memory safety and cryptographic constant time.
+  \MKin{conclusion}
+  The compositionality results enable modular proofs for the presented, secure compiler. % that is secure with respect to both memory safety and cryptographic constant time.
+  Some results of this work are implemented in the Coq proof assistant.
 
@@
 
-This section discusses work on robust preservation\MP{
-	robust compilation.
-	NOT robust preservation
-} (\Cref{subsec:relw:seccomprtp}) and on other secure compilation criteria (\Cref{subsec:relw:seccompcrit}).
+This section discusses work on robust compilation (\Cref{subsec:relw:seccomprtp}) and on other secure compilation criteria (\Cref{subsec:relw:seccompcrit}).
 Since the case study of \Cref{sec:casestud:defs,sec:casestud:rtp} implements measures for preserving \gls{ms} and \gls{cct}, this section then presents relevant related work as well (\Cref{subsec:relw:msmechs,subsec:relw:cctmechs}).
@@
 No existing work is concerned with composing robustly safe compilers.
-\MP{perhaps say taht we believe this scales to different trace models too and differnet kinds of props, but the ocmposition of monitors becomes tricky}
-The work relating robust preservation with universal composability~\cite{patrignani2022universal} is closest to what this paper presents.\MP{why}
+\MPin{
+  perhaps say taht we believe this scales to different trace models too and differnet kinds of props, but the ocmposition of monitors becomes tricky
+  \MKin{conclusion?}
+}
+The work relating robust preservation with universal composability~\cite{patrignani2022universal} is closest to what this paper presents.
+The authors demonstrate a similar compositionalty theorem to what is presented here (\Cref{sec:compcomp}) but use it in the context of protocols.
+They do not demonstrate the scalability of the approach.
+Moreover, they are missing the upper and lower compositions.
 % The authors demonstrate a similar compositionality theorem to what is presented here (\Cref{sec:compcomp}) as well as in an earlier version of this work~\cite{kruse2022csc}.
@@
 
-Different mechanisms for \MP{enforcing?}memory safety exist that also consider the secure compilation domain, i.e., have an active attacker model.
+Different mechanisms for enforcing memory safety exist that also consider the secure compilation domain, i.e., have an active attacker model.
 For example, the ,,pointers as capabilities'' principle represents pointers as machine-level capabilities~\cite{korashy2021capableptrs}, which behave in a similar fashion to capabilities by means of linear typing~\cite{morrisett2005L3}.
@@
 Moreover, this paper considers an active attacker, while the work on $L^{3}$ only discusses whole programs and, thus, has no active attacker model.
-The instrumentation to ensure memory safety that this paper presents is inspired by seminal work in the literature~\cite{nagarakatte2009soft}.
+The instrumentation to ensure memory safety that this paper presents is inspired by Softbounds~\cite{nagarakatte2009soft}.
 That work inserts bounds-checks in front of pointer-dereferences and, for this to work, inserts meta-data information on pointer creation.
-It\MP{what} also works in a more advanced setting with structured fields accesses and also introduces a table-lookup for pointers that are stored in memory.
-This paper only considers pointers to arrays of simple objects\MP{NOT OBJECTS}, i.e., there are no pointers to pointers or structures.
+Softbounds also works in a more advanced setting with structured fields accesses and also introduces a table-lookup for pointers that are stored in memory.
+This paper only considers arrays of primitive data, i.e., there are no pointers to pointers or structures.
 Several other approaches to memory-safety exist in literature, specifically as compiler instrumentations~\cite{akritidis2009baggy,younan2010paricheck,jung2021pico,shankaranarayana2023tailcheck,dhumbumroong2020boundwarden,nam2019framer,zhou2023fatptrs}, hardware-extensions~\cite{kwon2013lowfat,saileshwar2022heapcheck,chen2023flexpointer,kim2023whistle}, or programming language extensions~\cite{elliott2018checkedc,li2022formalcheckedc,jim2002cyclone,elliott2015guilt,west2005cuckoo,weis2019fyr,benoit2019uniqueness}.
@@
 
-The approach to \MP{preserving?enforcing?} cryptographic constant time in this paper is high-level, where a programming language exposes a way to switch the semantics to a data (operand) independent timing mode.
+The approach to preserving cryptographic constant time in this paper is high-level, where a programming language exposes a way to switch the semantics to a data (operand) independent timing mode.
 Since identifiers in $\Lscct$ are annotated with a secrecy tag, this approach is similar to others with information flow control.

### 2023-07-11 popl/main.tex
fixes

--- 2023-07-11 popl/main.tex
@@
 \begin{proof}
-	From \Thmref{thm:ccscct:rtp:scct}, we have that any LTMS program \src{p} gets compiled into a L program \trg{p} that is also TMS.
-	Then, from \Thmref{thm:cccfccdce:rtp:ms} we have that \trg{p} gets compilet to a program ?? that is also SMS, assuming it was SMS to begin with.
+	From \Thmref{thm:ccscct:rtp:scct}, we have that any LTMS program $\src{p}$ gets compiled into a L program $\trg{p}$ that is also TMS.
+	Then, from \Thmref{thm:cccfccdce:rtp:ms} we have that $\trg{p}$ gets compilet to a program ?? that is also SMS, assuming it was SMS to begin with.
 	Since TMS + SMS = MS, ?? is MS.

### 2023-07-11 popl/main.tex
no proofs

@@
     When calling this in similar fashion to \Cref{ex:strncpy:sms}, the event $\ev{Use}\ \loc_{x}\ 2;\comp;\unlock$ would not be emitted during execution.
-  \MKin{
-    this is kind of problematic for several reasons. (1) we don't pass pointers (2) fns signatures should not change
-  }
 \end{example}
@@
 
-%\Cref{thm:ccab:rtp:ms}\MP{states ... } follows from \Cref{thm:cca:rtp:tms,thm:ccb:rtp:sms} using \Cref{thm:rtp}.
+\Cref{thm:ccab:rtp:ms} strates that the composition of $\cca$ and $\ccb$ is secure with respect to \gls{ms} and follows from \Cref{thm:cca:rtp:tms,thm:ccb:rtp:sms} using \Cref{thm:rtp}.
 
@@
 \end{theorem}
-\begin{proof}
-  Consequence of \Thmref{thm:cca:rtp:tms} and \Thmref{thm:ccb:rtp:sms} using \Thmref{thm:rtp}.
-  \MKin{%
-    \textbackslash Thmref or \textbackslash Cref?
-  }
-\end{proof}
 
@@
 
-%With both \Cref{thm:ccdce:rtp:ms,thm:cccf:rtp:ms} it follows from \Cref{corr:swappable} that the two passes can be interchanged arbitrarily:
+With both \Cref{thm:ccdce:rtp:ms,thm:cccf:rtp:ms} it follows from \Cref{corr:swappable} that the two passes can be interchanged arbitrarily:
 
@@
 \end{theorem}
-\begin{proof}
-  Consequence of \Thmref{thm:cccf:rtp:ms} and \Thmref{thm:ccdce:rtp:ms} using \Thmref{corr:swappable}.
-  \MKin{%
-    \textbackslash Thmref or \textbackslash Cref?
-  }
-\end{proof}
 

### 2023-07-11 popl/main.tex
minor notation change
intro

--- 2023-07-11 popl/main.tex
@@
 
-\gls{ms} is a security property that ensures that there are no out-of-bounds accesses as well as no reads or writes to uninitialised memory.
-Programs that do not attain memory safety, can easily have their private data be leaked to attackers or be corrupted~\cite{lemay2021ccc}.
-Because of this, \gls{ms} is a requirement for programs to attain an even stronger security property, namely \gls{cct}.
-Programs that attain \gls{cct} are secure with respect to timing attacks, and this is the gold standard for secure applications, including cryptographic ones, since timing attacks can expose secrets via a simple performance analysis~\cite{kocher1996timing}.
-Timing attacks have become more and more relevant recently, especially given the era of cloud-computing~\cite{aviram2010cloudtime,kumar2019cloudsecsurvey}, where programs share the same hardware but are supposed to run in a sandboxed environment that does not leak secrets.
-Unfortunately, several attacks within such a cloud environment have been demonstrated as viable strategies to extract secrets~\cite{mehmet2015getoff,flowers2022zeroday,atya2019catchme,venkatanathan2015placevul}.
-
-Even though mitigations for such attacks exist~\cite{bond2017vale,almeida2017jasmin}, some compilers may optimise these mitigations away~\cite{barthe2018sec}.
-This is of course not what programmers expect and to ensure such mitigations persist in the compiled code, one can use a {\em secure} compilers.
-More generally, secure compilers should preserve security properties even when the program runs in a hostile context such as the cloud one, i.e., where it interacts with other possibly malicious programs.
-An example of such secure compilers are those that ascribe to the theory of robust compilation~\cite{abate2019jour,abate2021extacc,patrignani2021rsc}, which takes this interaction between components and potentially malicious contexts into account.
-
-Unfortunately, it is unclear if and how robust compilation supports compositionality, and compilation passes are typically developed as a standalone module to be plugged into a compilation pipeline and composed with other passes.
-Different kinds of compilation passes exist, e.g., there are usually several compilation passes that perform optimisations~\cite{lattner2004llvm,googlev8,androidstudio,kuepper2023cryptopt,manjikian1997fusion,wegman1991ccp} or source-code instrumentations~\cite{nagarakatte2009soft,nagarakatte2010cets,akritidis2009baggy,dhumbumroong2020boundwarden,jung2021pico,nam2019framer,shankaranarayana2023tailcheck,younan2010paricheck,zhou2023fatptrs,bond2017vale,almeida2017jasmin,kuepper2023cryptopt,cauligi2019fact} that enforce that the compiled program fulfils a property of interest.
-Without a framework for composing secure compiler passes, it is not possible to enable separation of concerns, e.g., to have a secure compilation pass that ensures \gls{ms} that is developed independently of another secure pass for \gls{cct} as suggested before and as showcased in \Cref{ex:strncpy}.
+% \gls{ms} is a security property that ensures that there are no out-of-bounds accesses as well as no reads or writes to uninitialised memory.
+% Programs that do not attain memory safety, can easily have their private data be leaked to attackers or be corrupted~\cite{lemay2021ccc}.
+% Because of this, \gls{ms} is a requirement for programs to attain an even stronger security property, namely \gls{cct}.
+% Programs that attain \gls{cct} are secure with respect to timing attacks, and this is the gold standard for secure applications, including cryptographic ones, since timing attacks can expose secrets via a simple performance analysis~\cite{kocher1996timing}.
+% Timing attacks have become more and more relevant recently, especially given the era of cloud-computing~\cite{aviram2010cloudtime,kumar2019cloudsecsurvey}, where programs share the same hardware but are supposed to run in a sandboxed environment that does not leak secrets.
+% Unfortunately, several attacks within such a cloud environment have been demonstrated as viable strategies to extract secrets~\cite{mehmet2015getoff,flowers2022zeroday,atya2019catchme,venkatanathan2015placevul}.
+
+% Even though mitigations for such attacks exist~\cite{bond2017vale,almeida2017jasmin}, some compilers may optimise these mitigations away~\cite{barthe2018sec}.
+% This is of course not what programmers expect and to ensure such mitigations persist in the compiled code, one can use a {\em secure} compilers.
+% More generally, secure compilers should preserve security properties even when the program runs in a hostile context such as the cloud one, i.e., where it interacts with other possibly malicious programs.
+% An example of such secure compilers are those that ascribe to the theory of robust compilation~\cite{abate2019jour,abate2021extacc,patrignani2021rsc}, which takes this interaction between components and potentially malicious contexts into account.
+
+% Unfortunately, it is unclear if and how robust compilation supports compositionality, and compilation passes are typically developed as a standalone module to be plugged into a compilation pipeline and composed with other passes.
+% Different kinds of compilation passes exist, e.g., there are usually several compilation passes that perform optimisations~\cite{lattner2004llvm,googlev8,androidstudio,kuepper2023cryptopt,manjikian1997fusion,wegman1991ccp} or source-code instrumentations~\cite{nagarakatte2009soft,nagarakatte2010cets,akritidis2009baggy,dhumbumroong2020boundwarden,jung2021pico,nam2019framer,shankaranarayana2023tailcheck,younan2010paricheck,zhou2023fatptrs,bond2017vale,almeida2017jasmin,kuepper2023cryptopt,cauligi2019fact} that enforce that the compiled program fulfils a property of interest.
+% Without a framework for composing secure compiler passes, it is not possible to enable separation of concerns, e.g., to have a secure compilation pass that ensures \gls{ms} that is developed independently of another secure pass for \gls{cct} as suggested before and as showcased in \Cref{ex:strncpy}.
+
+\gls{ms} is a security property obtained by composing \gls{sms}, which ensures array accesses are all within bounds, and \gls{tms}, which ensures pointers are only used when they are valid~\cite{azevedo2018meaningsofms}\MP{more cits}.
+\gls{cct} is a security property that ensures sensitive data is not leaked via timing side-channels~\cite{kocher1996timing}.
+Together, \gls{sms}, \gls{tms} and \gls{cct} yield \gls{msscct}, which is the gold standard of security properties for secure applications: programs attaining \gls{msscct} do not leak sensitive data neither through erroneous memory accesses, nor through timing side-channels.
+As discussed in \Cref{ex:strncpy}, these security properties can be enforced by compiler passes~\cite{bond2017vale,almeida2017jasmin}, to ensure programmers need not be aware of the various architectural details where their code will run.
+
 \begin{example}[strncpy]\label{ex:strncpy}
-Consider the \texttt{C} program below that copies a null-terminated string \texttt{src} into \texttt{dst} up to a length of \texttt{n}:
-
+Consider the \texttt{C} function \texttt{strncpy} that copies a null-terminated string \texttt{src} into \texttt{dst} up to a length of \texttt{n}.
+% Assume that the length of the array pointed by \texttt{dst} is at least \texttt{n}, and \texttt{src} is terminated by \texttt{'\textbackslash 0'}.
+% Even under these assumptions, there is a 
+This function is subject to a subtle \gls{sms} vulnerability: The bounds check \texttt{i < n} should happen {\it before} the access to memory location \texttt{x[i]}: otherwise 
+% If \texttt{n} is at least as large as the length of the array \texttt{src}, 
+the memory location past the last element will leak.
 \begin{lstlisting}[language=c,basicstyle=\small\ttfamily,morekeywords={size_t}]
@@
 
-For sake of simplicity, assume that the length of the array \texttt{dst} points to is at least \texttt{n}, and \texttt{src} is terminated by \texttt{'\textbackslash 0'}.
-Even under these assumptions, there is a subtle \gls{ms} vulnerability in the program: The bounds check \texttt{i < n} should happen {\it before} the access to memory location \texttt{x[i]}:
-If \texttt{n} is at least as large as the length of the array \texttt{src}, the memory location past the last element will be read.
-To prevent this vulnerability, one can use a compilation pass that enforces spatial memory safety, such as Softbounds~\cite{nagarakatte2009soft} or BaggyBounds~\cite{akritidis2009baggy}.
-
-Unfortunately, fixing \gls{ms} is not enough to have a secure program, because \texttt{strncpy} is susceptible to timing attacks.
-This is because the loop can terminate early: As soon as the string-terminating character \texttt{'\textbackslash 0'} is encountered, the loop exits.
-Thus, the program execution time is proportional to the length of the array that \texttt{src} points to.
+To prevent this vulnerability, one can use a compilation pass that enforces \gls{sms}, such as Softbounds~\cite{nagarakatte2009soft} or BaggyBounds~\cite{akritidis2009baggy}.
+
+Unfortunately, fixing \gls{sms} is not enough to make \texttt{strncpy} secure, because of timing attacks.
+In fact, the loop can terminate early, as soon as the string-terminating character \texttt{'\textbackslash 0'} is encountered, making program execution time proportional to the length of the array pointed by \texttt{src}.
 % Often, programmers rewrite programs like these to have loops that run with a fixed number of iterations.
-Also in this case there exist compiler passes that can rewrite such programs~\cite{cauligi2019fact} so that they are no longer vulnerable to timing attacks.
-
-These fixes alone are not enough to make this program secure in a robust setting, i.e., a malicious context could trigger a violation of either \gls{ms} or \gls{cct}.
-Consider a program context that calls \texttt{strncpy} with an argument for \texttt{src} that points to uninitialised memory.
-This leads to a series of reads from uninitialised memory, which is an immediate \gls{ms} issue that can have devastating real-world implications~\cite{uninit-0,uninit-1,uninit-2,uninit-3,uninit-4}.
-An easy fix to sidestep this is to use non-nullable pointers as parameters instead, e.g., as presented in Cyclone~\cite{jim2002cyclone}.
-With this, a secure compiler needs to ensure that no malicious target-level context can supply unitialised memory and thereby sidestep the non-nullability guarantees of the source language.
-For compilers to target languages that provide capabilities (such as CHERI~\cite{woodruff2014CHERI}, Arm Morello~\cite{arm-morello}, or MSWasm~\cite{michael2023mswasm}), which are unforgable handles that encapsulate information about a pointer, they can leavarage the information stored in these capablities to ensure the non-nullability guarantees in the target.
+Also in this case there exist compiler passes that can rewrite such programs into \gls{cct} ones~\cite{cauligi2019fact}.
+ % so that they are no longer vulnerable to timing attacks.
+
+Alas, code is not run in isolation, so a malicious attacker could supply code that intracts with \texttt{strncpy} and trigger a violation of either \gls{ms} or \gls{cct} by calling \texttt{strncpy} with an argument for \texttt{src} that points to uninitialised memory.
+This would, in turn, triggering a series of reads from uninitialised memory, which is an immediate \gls{ms} violation with devastating real-world implications~\cite{uninit-0,uninit-1,uninit-2,uninit-3,uninit-4}.
+% An easy fix to sidestep this is to use non-nullable pointers as parameters instead, e.g., as presented in Cyclone~\cite{jim2002cyclone}.
+
+Robust compilers~\cite{abate2019jour} are a form of secure compilers that preserve security properties even in the presence of arbitrary attackers interacting with compiled code.
+Thus, robust compilers can be used to prevent these vulnerabilities, e.g., by targeting capability-based languages such as CHERI~\cite{woodruff2014CHERI}, Arm Morello~\cite{arm-morello}, or MSWasm~\cite{michael2023mswasm}, where pointers are always initialised.
 \end{example}
 
+% To ensure these compiler-preserved properties still hold even in the presence of arbitrary attackers interacting with compiled code, one can use secure compilers in the form of robust compilers~\cite{abate2019jour,abate2021extacc,patrignani2021rsc}.
+
+Unfortunately, given secure compiler passes that each preserve a possibly-different security property, there is no way to tell what kind of security property will the composition of those secure compilers preserve.
+% Different kinds of compilation passes exist, e.g., there are usually several compilation passes that perform optimisations~\cite{lattner2004llvm,googlev8,androidstudio,kuepper2023cryptopt,manjikian1997fusion,wegman1991ccp} or source-code instrumentations~\cite{nagarakatte2009soft,nagarakatte2010cets,akritidis2009baggy,dhumbumroong2020boundwarden,jung2021pico,nam2019framer,shankaranarayana2023tailcheck,younan2010paricheck,zhou2023fatptrs,bond2017vale,almeida2017jasmin,kuepper2023cryptopt,cauligi2019fact} that enforce that the compiled program fulfils a property of interest.
+Worse, without a framework for composing secure compiler passes, it is not possible to enable separation of concerns, e.g., to have a secure compilation pass that ensures \gls{ms} that is developed independently of another secure pass for \gls{cct}.
+
 This paper introduces a framework for reasoning about the composition of secure and optimising compiler passes akin to those of \Cref{ex:strncpy} and it showcases the power of this framework by instantiating it on a multi-pass compilation chain.
-To this end, this paper first discusses how to compose security properties, such as \gls{tms} and \gls{sms} into \gls{ms}, and \gls{cct}.
-Then, this paper defines compiler composition and formalises what kind of security properties does the composition of two compiler passes preserve, given what the individual passes preserve.
-The paper then defines several secure compiler passes, where each is either preserving a different security property (e.g., temporal or spatial memory safety) or performing a security-preserving optimisation, (e.g., applying \gls{cf} or \gls{dce}).
-Finally, this paper shows how to compose these secure compiler passes into a multi-pass compilation chain which is proven to provide end-to-end preservation of general memory safety and cryptographic constant time.
-Crucially, this end-to-end security preservation is obtained by having each compilation pass (i) preserve a sub-part of the overall security property individually and then (ii) compose each preserved security property as dictated by the framework of this paper.
+To this end, this paper first discusses how to compose security properties, such as \gls{tms} and \gls{sms} into \gls{ms}, and then adding \gls{cct} to the mix to obtain \gls{msscct}. 
+Then, this paper defines compiler composition and formalises that given two passes that securely preserve two (possibly-distinct) properties, their composition securely preserves the composition of those properties.
+The paper then defines several secure compiler passes, where each is either preserving a different security property (\gls{tms}, \gls{sms}, \gls{cct}) or performing a security-preserving optimisation, (e.g., applying \gls{cf} or \gls{dce}).
+Finally, this paper shows that composing these secure compiler passes into a multi-pass compilation chain results that provides end-to-end preservation of \gls{msscct}.
+Crucially, this end-to-end security preservation is obtained by composing the security property preserved by each individual pass, as dictated by the framework of this paper.
 The results showcase how the framework allows the kind of formal security reasoning that compiler writers already want (and do), obtaining precise, compositional security reasoning while providing minimal (and modular) proof effort.
@@
         %\Cref{sec:compprop} presents a formal framework to reason about compositions of security properties and demonstrate its usage with concrete examples:
-        First, this paper studies the formalisation and the composition of security properties (\Cref{sec:compprop}) by focussing on those properties that are of interest for real-world compiler writers (as identified by the plethora of work enforcing such properties individually~\cite{akritidis2009baggy,nagarakatte2009soft,nagarakatte2010cets,dhumbumroong2020boundwarden,jung2021pico,nam2019framer,shankaranarayana2023tailcheck,younan2010paricheck,zhou2023fatptrs,bond2017vale,almeida2017jasmin,kuepper2023cryptopt,cauligi2019fact}).
-        % 
-        Starting from ways to formalise and enforce those properties individually, this paper shows how to compose their formalisation.
-        % 
-        This paper then showcases the benefits of this composition by composing \gls{tms} and \gls{sms} into \gls{ms}, and ultimately adding \gls{cct} to the composition, yielding the the conjunction of \gls{ms} and \gls{cct}.
-        % 
-        The resulting security property is the golden standard of safety properties that secure cryptographic code should have~\cite{lemay2021ccc}.
-%        \MPin{
-%        	conclude that did this prop in a single pass, we show how to do it compositionally
-%        }
+        First, this paper formalises security properties (\Cref{sec:compprop}) by focussing on those properties that are of interest for real-world compiler writers (as identified by the plethora of work enforcing such properties individually~\cite{akritidis2009baggy,nagarakatte2009soft,nagarakatte2010cets,dhumbumroong2020boundwarden,jung2021pico,nam2019framer,shankaranarayana2023tailcheck,younan2010paricheck,zhou2023fatptrs,bond2017vale,almeida2017jasmin,kuepper2023cryptopt,cauligi2019fact}), namely \gls{tms}, \gls{sms} and \gls{cct}.
+        % 
+        Starting from ways to formalise and enforce those properties individually, this paper shows how to compose their formalisation as well as their enforcement.
+        % 
+        % This paper then showcases the benefits of this composition by composing \gls{tms} and \gls{sms} into \gls{ms}, and ultimately adding \gls{cct} to the composition, yielding the the conjunction of \gls{ms} and \gls{cct}.
+        % 
+        The resulting security property is \gls{msscct}, i.e., the gold standard of security properties for secure programs~\cite{lemay2021ccc}.
 
@@
         %Concretely, it investigates upper and lower compositions, as well as the sequential composition.
-        For this, this paper studies compilers with multiple inputs or multiple outputs as well as the sequential composition of compilers, as typically done in existing compilation chains.
+        For this, this paper studies sequential compiler composition as well as compilers with multiple inputs or multiple outputs, as typically done in existing compilation chains.
         %
         This paper proves that starting from two compilers that preserve two (possibly distinct) properties, their composition preserves the intersection of those properties.
-        For example, a compiler that is secure with respect to \gls{tms} can be composed with a compiler that is secure with respect to \gls{sms}, resulting in a compiler that is secure with respect to \gls{ms}.
-%       \MPin{
-%      		find way to connect to previous contrib
-%       }
-        % 
-        Finally, this paper provides an important corollary that demonstrates that the order of composition of sequential compiler passes is irrelevant for the resulting security.
-        % 
-        This corollary is crucial for reordering compiler optimisation passes as well as secure compilation passes and thus generating secure and efficient code.
+        % For example, a compiler that is secure with respect to \gls{tms} can be composed with a compiler that is secure with respect to \gls{sms}, resulting in a compiler that is secure with respect to \gls{ms}.
+        % 
+        Finally, this paper proves that the order of composition of sequential compiler passes is irrelevant for the resulting security.
+        % 
+        This is crucial for reordering optimisation passes and thus generating secure and efficient code.
 
@@
         This paper presents a case-study showcasing the conjunction of the previous contributions (\Cref{sec:casestud:defs,sec:casestud:rtp}).
-        To this end, it presents a compilation chain consisting of several passes.
-        The chain ultimately preserves a combination of \gls{ms} and \gls{cct} by means of composing the individual, secure passes concerning \gls{tms}, \gls{sms}, and a stricter variant of \gls{cct}, respectively.
-        Furthermore, the chain includes two passes for \gls{ms} that are optimising: One performs a simple \gls{dce} and the other \gls{cf}.
+        To this end, it presents a compilation chain consisting of several passes that ultimately preserves \gls{msscct} by means of composing the individual, secure passes concerning \gls{tms}, \gls{sms}, and \gls{cct}, respectively.
+        Furthermore, the chain includes two optimisation passes: One performs a simple \gls{dce} and the other \gls{cf}.
         The formalisation of this case study showcases the power of the presented framework: Namely, the divide-and-conquer approach to software engineering is a viable strategy even for the development of secure compilers.

### 2023-07-11 popl/main.tex

@@
   $$
-  \mssafe:=\left\{\trace \left|\begin{array}{rcl}
-                                 \trace\in\tmssafe & \text{ and } & \trace\in\smssafe
-                               \end{array}\right.\right\}
+  \mssafe:=\tmssafe \cap \smssafe
   $$
@@
 \end{definition}
-\MPin{
-	why is this defined differently than MS? they should be defined in the same way
-}
 

### 2023-07-11 popl/main.tex
start of 5.1

--- 2023-07-11 popl/main.tex
@@
   \begin{aligned}
-  \mi{(Expressions)}~&\expr\bnfdef\ x \mid \valueexpr \mid \lbinop{\expr[_{1}]}{\expr[_{2}]} \mid \lget{x}{\expr} \mid \lnew{\expr[_{1}]}{\expr[_{2}]} \mid \ldelete{x} \mid \lset{x}{\expr[_{1}]}{\expr[_{2}]} \\
-    & \mid \llet{x}{\expr[_{1}]}{\expr[_{2}]}  \mid \lreturn{\expr} \mid \lcall{g}{\expr} \mid \lifz{\expr}{\expr[_{1}]}{\expr[_{2}]} \mid \labort
+    \mi{(Expressions)}~&\expr\bnfdef\ x \mid \valueexpr \mid \lbinop{\expr[_{1}]}{\expr[_{2}]} \mid \lget{x}{\expr} \mid \llet{x}{\lnew{\expr[_{1}]}{\expr[_{2}]}}{\expr[_{3}]} \mid \ldelete{x} \mid \lset{x}{\expr[_{1}]}{\expr[_{2}]} \\
+                       & \mid \lpair{\expr[_{1}]}{\expr[_{2}]} \mid \lproja{\expr} \mid \lprojb{\expr} \mid \llet{x}{\expr[_{1}]}{\expr[_{2}]} \mid \lreturn{\expr} \mid \lcall{g}{\expr} \mid \lifz{\expr}{\expr[_{1}]}{\expr[_{2}]} \mid \labort
   \end{aligned}
@@
   \begin{aligned}
+  \mi{(Types)}~&\type\bnfdef\natt
+  &
   \mi{(Functions)}~&\function\bnfdef\ \lfunction{g}{x}{e}
@@
   \mi{(Programs)}~&\prog{\library_{\ctx}}{\library_{\comp}}
+  \end{aligned}
+\end{gather*}
+
+Above is the shared syntax of all the programming languages of this paper.
+Variables are referred to as $x,y,z,a,b,c,\ldots$ while functions may be referred to as multi-character words, such as $\texttt{foo}$, as well as short-forms like $f,g,h$.
+All languages share the type $\natt$ representing natural numbers.
+Functions are constrained to take one argument and can only call other functions listed in libraries, which are just lists of functions.
+A program $\prog{\library_{\ctx}}{\library_{\comp}}$ is indexed by two libraries that represent all context- and component-level functions, respectively.
+
+\begin{gather*}
+  \begin{aligned}
+    \mi{(Control~Tags)}~&\sandboxtag\bnfdef\ctx \mid \comp
+    &
+    \mi{(Communication~Tags)}~&\comm\bnfdef\ \ctxtocomp \mid\ \comptoctx \mid \nocomm
+    &
+    \mi{(Poison~Tags)}~&\poison\bnfdef \poisoned \mid \poisonless
+  \end{aligned}\\
+  \begin{aligned}
+    \mi{(Continuation~Stacks)}~&\ectxstack\bnfdef\hole{\cdot} \mid \tup{K}{g},\ectxstack
+    &
+    \mi{(Control~Flow~States)}~&\cfstate\bnfdef\tup{\library}{\ectxstack}
+  \end{aligned}\\
+  \begin{aligned}
+    \mi{(Locations)}~&\loc\in\nat
     &
     \mi{(Substitutions)}~&\substvar\bnfdef \hole{\cdot} \mid \subst{\valueexpr}{x},\gamma
+  \end{aligned}\\
+  \begin{aligned}
+    \mi{(Stores)}~&\Delta\bnfdef\hole{\cdot} \mid x\mapsto\tup{\loc}{\sandboxtag,\poison,n},\Delta
+    &
+    \mi{(Heaps)}~&H\bnfdef\hole{\cdot} | \valueexpr,H
+    \\
+    \mi{(Memory~States)}~&\memstate\bnfdef\tup{H^{\ctx}}{H^{\comp};\Delta}
+    &
+    \mi{(States)}~&\statevar\bnfdef\tup{\cfstate}{\sandboxtag,\memstate}
   \end{aligned}
 \end{gather*}
-% \MPin{
-%   this latex had many issues.
-%   the \& were at different spots
-%   it used hspace to space things instead of \&
-%   and it did not use \& to align the following things on a line
-%   and it did not use multiple aligned into the gather
-%   (what's the point of gather if you don't have several aligned inside? use a single align* instead)
-% }
-% oh... that's good to know! the tech report is also riddled with this... :/
-
-Above is a fraction of the shared syntax of all the programming languages of this paper.
-\MPin{why a fraction? what exprs are missing
-MK: Things that are missing here are: Evaluation contexts, states $\Omega$ and transitively other definitions affected by this
-}
-The syntax and semantics are mostly standard and therefore omitted.
-However, when allocating arrays with $\lnew{\expr[_{1}]}{\expr[_{2}]}$, the expression $\expr[_{1}]$ determines the initial value of the cells and $\expr[_{2}]$ the overall size of an array.
-
-\MPin{
-	below is semantics info. should go with their symbols and the semantics rules.
-	and a better description of the semantics symbols.
-  MK: this moved to appendix! should we bring it back? what exactly should be left out?
-}
-The data associated to pointers resides in the dynamic state $\Omega$ that contains sandboxed heaps, one for the context and the other for the component.
-Because of the sandboxing, locations pointing into heaps must be associated with a control-tag $\sandboxtag$ (either $\ctx$ or $\comp$) to know the heap they point into.
-Other metadata in $\Omega$ entails a list of functions with information on whether they belong to the context or the component.
-That way, when calling a function, the semantics can take care of emitting appropriate events that signal a context switch.
-For the languages considered in this paper, functions are second-class.
-So, the only values are pairs, pointers, and natural numbers.
-Another non-standard part of the languages is the concept of the poison-tag.
-The poison-tag marks an array as freed ($\poisoned$) without removing the information associated to the freed pointer from $\Omega$.
-This measure prevents the semantics from getting stuck when an already freed pointer is freed again.
-Because of this, it is possible that a program like $\llet{x}{\lnew{0}{5}}{\ldelete{x};\ldelete{x}}$ yields a trace $\ev{Alloc\ \loc\ \sandboxtag\ 5}\cdot\ev{Dealloc\ \loc\ \sandboxtag}\cdot\ev{Dealloc\ \loc\ \sandboxtag}$ when executing.
+
+States $\statevar$ are tuples containing a control flow state $\cfstate$, control tags $\sandboxtag$, and a memory state $\memstate$.
+A control flow state $\cfstate$ entails a library, which provides definitions for functions when calling, as well as a stack of continuations $\ectxstack$.
+Elements of the stack $\ectxstack$ are pairs of evaluation contexts and the name of a function associated to that evaluation context.
+Memory states $\memstate$ are tuples of two separate heaps $H$ and a store $\Delta$, which contains pointer metadata, such as the concrete memory location $\loc$, a control tag $\sandboxtag$ indicating which heap the pointer points into, a poison tag $\poison$, as well as bounds information $n$.
+The bounds information is a mere proof artifact that has no semantic significance.
+For details see the accompanying tech report~\MK{cit}.
+Memory states $\memstate$ have two separate heaps, essentially modelling a sandbox, to prevent contexts from performing pointer arithmetic and reading from or writing to the data owned by a component.
+While this prevents the effects of out of bounds accesses across the context and component boundary, the goal of the design of the languages of this paper is to be able to express security violations.
+To this end, the poison tag $\poison$ indicates whether a pointer has been freed ($\poisoned$) or is still allocated ($\poisonless$), so that uses of pointers can happen even after their deallocation without the semantics getting stuck.
 
@@
 
-Possible events that are emitted during execution are similar to the ones of the specification trace model (\Cref{subsec:propdefs}).
-The key difference is that, as standard in secure compilation work~\cite{patrignani2021rsc,abate2019jour}, the traces have a call and return event that signals context switches, which are referred to as {\em interaction events}.
+The possible events that are emitted during execution are similar to the ones of the specification trace model (\Cref{subsec:propdefs}).
+The key difference is that, as standard in secure compilation work~\cite{patrignani2021rsc,abate2019jour,korashy2021capableptrs}, the traces have a call and return event that signals context switches, which are referred to as {\em interaction events}.
 The reason for these interaction events is technical: They are a proof artifact for reconstructing a source context from a potentially malicious target context, where during that translation, the insertion of some wrapper code right before context switching may be necessary to make the proof succeed.
@@
 Contrary, $\ev{Call\ \comptoctx\ foo\ \valueexpr}$ and $\ev{Return\ \ctxtocomp\ \valueexpr}$ signal that program execution transitions from component- to context-level.
-The environmental semantics emits the tag $\nocomm$ for calls from context- to context-level or component- to component-level.
+For calls without this context switch, the environmental semantics attaches the $\nocomm$ tag.
 In the following, $\neg\ctx=\comp$ and $\neg\comp=\ctx$.
@@
 
-The environmental semantics is also mostly straightforward.
+The environmental semantics is mostly straightforward.
 In \Cref{tr:e-ret-notsame,tr:e-call-notsame} the judgement $\commlib\vdash\text{foo}:\Omega.\sandboxtag$ checks whether $\text{foo}$ is a component-level name by looking it up in the list of component-level names $\commlib$ and emits the appropriate transfer tag, i.e., either $\comptoctx$ or $\ctxtocomp$.

### 2023-07-11 popl/main.tex
5.1 first pass

@@
   &
-  \mi{(Functions)}~&\function\bnfdef\ \lfunction{g}{x}{e}
+  \mi{(Functions)}~&\function\bnfdef\ \lfunction{\textit{foo}}{x}{e}
   &
-    \mi{(Libraries)}~&\library\bnfdef\ \hole{\cdot} \mid \function,\library \\
+  \mi{(Libraries)}~&\library\bnfdef\ \hole{\cdot} \mid \function,\library
+  \end{aligned}\\
+  \begin{aligned}
+  \mi{(Component~Names)}~&\commlib\bnfdef\ \hole{\cdot} \mid \text{foo},\commlib
+  &
   \mi{(Programs)}~&\prog{\library_{\ctx}}{\library_{\comp}}
@@
 A program $\prog{\library_{\ctx}}{\library_{\comp}}$ is indexed by two libraries that represent all context- and component-level functions, respectively.
+Lists of component-level names are referred to as $\commlib$.
 
@@
     \mi{(States)}~&\statevar\bnfdef\tup{\cfstate}{\sandboxtag,\memstate}
+    &
+    \mi{(Runtime~Terms)}~&\runtimetermvar\bnfdef\rtt{\statevar}{\expr}
   \end{aligned}
@@
 
-States $\statevar$ are tuples containing a control flow state $\cfstate$, control tags $\sandboxtag$, and a memory state $\memstate$.
+States $\statevar$ are tuples\footnote{Throughout the paper, the subsitution notation is also used to update entries in states $\statevar$.} containing a control flow state $\cfstate$, control tags $\sandboxtag$, and a memory state $\memstate$.
 A control flow state $\cfstate$ entails a library, which provides definitions for functions when calling, as well as a stack of continuations $\ectxstack$.
@@
 While this prevents the effects of out of bounds accesses across the context and component boundary, the goal of the design of the languages of this paper is to be able to express security violations.
-To this end, the poison tag $\poison$ indicates whether a pointer has been freed ($\poisoned$) or is still allocated ($\poisonless$), so that uses of pointers can happen even after their deallocation without the semantics getting stuck.
+To this end, the poison tag $\poison$ indicates whether a pointer has been freed $\poisoned$ or is still allocated $\poisonless$, so that uses of pointers can happen even after their deallocation without the semantics getting stuck.
+Runtime terms are simply expressions $\expr$ paired with the operational state $\statevar$.
 
@@
   \begin{aligned}
-  \mi{(Pre\text{-}Events)}~\event[_{b}]&\bnfdef \ev{Alloc}\ \loc\ \valueexpr \mid \ev{Dealloc}\ \loc \mid \ev{Get}\ \loc\ \valueexpr \mid \ev{Set}\ \loc\ \valueexpr\ \valueexpr['] \mid \ev{Call}\ \comm\ g\ \valueexpr \mid \ev{Ret}\ \comm\ \valueexpr \mid \ev{Start} \mid \ev{End}\ \valueexpr \\
-  \mi{(Events)}~\event&\bnfdef \emptyevent \mid \lightning \mid \tup{\event[_{b}]}{\sandboxtag} \\
+  \mi{(Pre\text{-}Events)}~\event[_{b}]&\bnfdef \ev{Alloc}\ \loc\ \valueexpr \mid \ev{Dealloc}\ \loc \mid \ev{Get}\ \loc\ \valueexpr \mid \ev{Set}\ \loc\ \valueexpr\ \valueexpr['] \mid \cdots \\
+  \mi{(Events)}~\event&\bnfdef \emptyevent \mid \lightning \mid \tup{\event[_{b}]}{\sandboxtag,\securitytag} \\
   \end{aligned}
@@
 
-The possible events that are emitted during execution are similar to the ones of the specification trace model (\Cref{subsec:propdefs}).
-The key difference is that, as standard in secure compilation work~\cite{patrignani2021rsc,abate2019jour,korashy2021capableptrs}, the traces have a call and return event that signals context switches, which are referred to as {\em interaction events}.
+All languages use the same trace model.
+In this trace model, events are either the empty event \emptyevent, the program crash event $\lightning$, or a tuple consisting of a control tag $\sandboxtag$ and a security tag $\securitytag$.
+The former indicates wether the component \comp or the context \ctx is to blame for emitting this event, the latter indicates the secrecy level of values of the emitted event, i.e., either $\lock$ or $\unlock$.
+As for pre-events, the memory-related ones are allocation ($\ev{Alloc}\ \loc\ \valueexpr$), deallocation ($\ev{Dealloc}\ \loc$), reading from ($\ev{Get}\ \loc\ \valueexpr$) and writing to memory ($\ev{Set}\ \loc\ \valueexpr\ \valueexpr[']$).
+
+\begin{center}
+  \judgbox{\pstepto{\runtimetermvar}{\runtimetermvar[']}{\event}}{,,$\runtimetermvar$ does one primitive step to $\runtimetermvar[']$ emitting event $\event$.''}
+  %
+  \typerule{$e-\text{get}-\in$}{
+    \sandboxtag=\statevar.\sandboxtag &
+    \statevar.\store(x)=\tup{\loc}{\sandboxtag,\poison,m} \\
+    \loc+n\in\operatorname{dom}\statevar.\heap[^{\sandboxtag}](\loc+n)
+  }{
+    \pstepto{\rtt{\statevar}{\lget{x}{n}}}{\rtt{\statevar}{\heap[^{\sandboxtag}](\loc+n)}}{\tup{\ev{Get}\ \loc\ n}{\sandboxtag}}
+  }{e-get-in}
+  %
+  \typerule{$e-\text{set}-\notin$}{
+    \sandboxtag=\statevar.\sandboxtag &
+    \statevar.\store(x)=\tup{\loc}{\sandboxtag,\poison,m} \\
+    \loc+n\notin\operatorname{dom}\statevar.\heap[^{\sandboxtag}]
+  }{
+    \pstepto{\rtt{\statevar}{\lset{x}{n}{\valueexpr}}}{\rtt{\statevar}{\valueexpr}}{\tup{\ev{Set}\ \loc\ n\ \valueexpr}{\sandboxtag}}
+  }{e-set-notin}
+  %
+  \typerule{$e-\text{new}$}{
+    \fresh{\statevar}{z} &
+    \fresh{\statevar}{\loc} &
+    \heap[^{\sandboxtag}_1]=\growh{\statevar.\heap[^{\sandboxtag}]}{n} &
+    \store[_1]=\storeel{z}{\loc}{\statevar.\sandboxtag}{\poisonless}{n},\statevar.\store
+  }{
+    \pstepto{\rtt{\statevar}{\lnew{x}{n}{\expr}}}{\rtt{\statevar\subst{\heap[^{\sandboxtag}]}{\heap[^{\sandboxtag}_1]}\subst{\store}{\store[_1]}}{\expr\subst{z}{x}}}{\tup{\ev{Alloc}\ \loc\ n}{\sandboxtag}}
+  }{e-new}
+  %
+  \typerule{$e-\text{dealloc}$}{
+    \statevar.\store(x) = \tup{\loc}{\statevar.\sandboxtag,\poison,n} &
+    \store[_1]=\seth{\statevar.\store}{x}{\tup{\loc}{\statevar.\sandboxtag,\poisoned,n}}
+  }{
+    \pstepto{\rtt{\statevar}{\ldelete{x}}}{\rtt{\statevar\subst{\store}{\store[_1]}}{0}}{\tup{\ev{Dealloc}\ \loc}{\sandboxtag}}
+  }{e-dealloc}
+  %
+  \typerule{$e-\text{abort}$}{
+  }{
+    \pstepto{\rtt{\statevar}{\labort}}{\lightning}{\lightning}
+  }{e-abort}
+\end{center}
+
+\Cref{tr:e-dealloc,tr:e-get-in,tr:e-set-notin} demonstrate the use of the poison tag $\poison$ as metadata for pointers instead of removing them from the store $\Delta$.
+Take deallocation as an example, the premise does not care at all about the actual state of the poison tag $\poison$ and just overwrites it, marking the location as freed $\poisoned$.
+Besides that, the poison tag does not have any semantic meaning, at least in general.
+For language $\Ltms$, this tag is really just some semantic metadata that programmers have no access to.
+But, for the other languages, e.g., $\Ltrg$, the poison tag is used to check pointer validity.
+The \Cref{tr:e-new} allocates enough space on the respective heap, either $H^{\ctx}$ or $H^{\comp}$ depending on the execution context, i.e., the value of $\Omega.\sandboxtag$, and adds the appropriate metadata associated to the pointer in $\Delta$.
+Reading from and writing to memory have to cases: Either the heap is large enough or not and, depending on that, either the actual value stored at that location is read from or written to, or some garbage data is returned.
+However, note that the execution does not get stuck in such cases, performs a step, and emits an appropriate event.
+Also note that wether a pointer is poisoned or not is not reflected on the trace.
+
+\begin{align*}
+  \mi{(Pre\text{-}Events)}~\event[_{b}]&\bnfdef \cdots \mid \ev{Call}\ \comm\ g\ \valueexpr \mid \ev{Ret}\ \comm\ \valueexpr \mid \ev{Start} \mid \ev{End}\ \valueexpr \\
+\end{align*}
+
+A key difference in comparison with the specification trace model (\Cref{subsec:propdefs}) is that, as standard in secure compilation work~\cite{patrignani2021rsc,abate2019jour,korashy2021capableptrs}, the traces have a call and return event that signals context switches, which are referred to as {\em interaction events}.
 The reason for these interaction events is technical: They are a proof artifact for reconstructing a source context from a potentially malicious target context, where during that translation, the insertion of some wrapper code right before context switching may be necessary to make the proof succeed.
@@
 The environmental semantics is mostly straightforward.
-In \Cref{tr:e-ret-notsame,tr:e-call-notsame} the judgement $\commlib\vdash\text{foo}:\Omega.\sandboxtag$ checks whether $\text{foo}$ is a component-level name by looking it up in the list of component-level names $\commlib$ and emits the appropriate transfer tag, i.e., either $\comptoctx$ or $\ctxtocomp$.
+In \Cref{tr:e-ret-notsame,tr:e-call-notsame}, the judgement $\Omega.\commlib\vdash\text{foo}:\Omega.\sandboxtag$ checks whether $\text{foo}$ is a component-level name by looking it up in the list of component-level names $\commlib$ and emits the appropriate transfer tag, i.e., either $\comptoctx$ or $\ctxtocomp$.
 Additional rules that are left out ensure that, e.g., when calling the $\text{main}$ function, the event $\ev{Start}$ is emitted, which is a design choice this paper does for convenience when reasoning about call-chains.
-
+Note that the $\ev{End}\ \valueexpr$ event is not emitted if the program crashes.
+
+\MKin{%
+  should I add a rule for top level exec or keep it like this?
+}
 The top-level execution $\progstepto{\prog{\library_{\ctx}}{\library_{\comp}}}{\runtimetermvar}{\trace}$ constructs an initial state $\Omega$ by linking $\library_{\ctx}$ and $\library_{\ctx}$ and then starts execution by calling the $\text{main}$ function.

### 2023-07-11 popl/main.tex
5.2

@@
 $\src{L_{\tmssafe}}$ uses the same syntax as presented earlier (\Cref{sec:casestud:defs}) without extensions to the term level.
-But, $\src{L_{\tmssafe}}$ is statically typed, where the type system is inspired by $L^{3}$~\cite{morrisett2005L3}.\MP{fabulous paper from amal?relevant?}
-\MP{and therefore omitted?}
+But, $\src{L_{\tmssafe}}$ is statically typed, where the type system is inspired by $L^{3}$~\cite{morrisett2005L3,scherer2018fabulous}.
 The type system of $\src{L_{\tmssafe}}$ exhibits the property that every well-typed $\src{L_{\tmssafe}}$ program satisfies \gls{tms} (\Cref{thm:wt:tms}).
 The proof of this theorem relies on a projection $\msfilterLtms{\src{\event}}=\event$ from $\src{L_{\tmssafe}}$ events to specification events $\event$, because the properties defined earlier (\Cref{subsec:propdefs}) are defined in the specification trace model.
-Hereby, $\delta$ maps $\src{L_{\tmssafe}}$ locations $\src{\loc}$ to locations $\loc$ of the specification trace model whenever an event with a location is projected.
-The projection $\msfilterLtms{\src{\event}}$ for the most part does the expected, e.g., $\msfilterLtms{\src{\ev{Alloc}\ \loc\ n;\comp;\securitytag}}=\ev{Alloc}\ \delta(\loc)\ n;\comp;\securitytag$.
+Hereby, the map $\thelocmap(\src{\loc})=\loc$ maps an $\Ltms$ location to a location $\loc$ of the specification trace model.
+
+\begin{center}
+  \judgbox{\msfilterLtms{\src{\event}}=\event}{,,Project an $\Ltms$ event $\src{\event}$ to a specification event $\event$.''}
+  \typerule{$\Ltms$-filter-context}{
+    \src{\event[_b]}\not=\src{\lightning}
+  }{
+    \msfilterLtms{\src{\tup{\event[_b]}{\ctx}} = \emptyevent}
+  }{ltms-filter-context}
+  %
+  \typerule{$\Ltms$-filter-abort}{
+  }{
+    \msfilterLtms{\src{\lightning}} = \lightning
+  }{ltms-filter-abort}
+  %
+  \typerule{$\Ltms$-filter-start}{
+  }{
+    \msfilterLtms{\src{\tup{Start}{\comp}}} = \emptyevent
+  }{ltms-filter-start}
+  %
+  \typerule{$\Ltms$-filter-alloc}{
+    \locmapsto{\src{\loc}}{\loc} &
+    \src{n}=n
+  }{
+    \msfilterLtms{\src{\tup{Alloc\ \loc\ n}{\comp}}} = \tup{\ev{Alloc}\ \loc\ n}{\comp,\unlock}
+  }{ltms-filter-alloc}
+\end{center}
+
+Most rules of the projection $\msfilterLtms{\src{\event}}$ are left out since, for the most part, it does the expected, e.g., $\msfilterLtms{\src{\ev{Dealloc}\ \loc;\comp;\securitytag}}=\ev{Dealloc}\ \delta(\loc);\comp;\securitytag$.
 But, it also filters any action that a context does as well as the interaction events, since these are irrelevant for component-level \gls{tms}.
-\MPin{ok let's add details}
 

### 2023-07-11 popl/main.tex
upto 5.4

@@
   \begin{aligned}
-  \mi{(Types)}~&\type\bnfdef\natt
+  \mi{(Types)}~&\type\bnfdef\natt \mid \type[_{1}] \times \type[_{2}]
   &
@@
 
-\begin{align*}
+\begin{center}
+  \begin{aligned}
     \mi{(Pre\text{-}Events)}~\event[_{b}]&\bnfdef \cdots \mid \ev{Call}\ \comm\ g\ \valueexpr \mid \ev{Ret}\ \comm\ \valueexpr \mid \ev{Start} \mid \ev{End}\ \valueexpr \\
-\end{align*}
+  \end{aligned}
+\end{center}
 
@@
 
-$\Ltrg$ extends the syntax presented earlier (\Cref{subsec:cs:defs}) with dynamic typechecks and a way to inspect poison tags in the metadata of pointers.
-For allocated ($\trg{\poisonless}$) arrays bound to variable $\trg{x}$, the check $\trg{\lispoisoned{x}}$ yields $\trg{1}$.
+\begin{center}
+  \begin{aligned}
+    \mi{(Expressions)}~\trg{\expr}&\bnfdef \cdots \mid \trg{\lispoisoned{x}} \mid \trg{\lhast{\expr}{\type}} \\
+  \end{aligned}
+\end{center}
+
+$\Ltrg$ extends the syntax presented earlier (\Cref{subsec:cs:defs}) with dynamic typechecks $\trg{\lhast{\expr}{\type}}$ and a way to inspect poison tags $\trg{\lispoisoned{x}}$ in the metadata of pointers.
+For valid pointers ($\trg{\poisonless}$) bound to variable $\trg{x}$, the check $\trg{\lispoisoned{x}}$ yields $\trg{1}$.
 If the array bound to $\trg{x}$ was allocated, i.e., has been poisoned ($\trg{\poisoned}$), the check $\trg{\lispoisoned{x}}$ evaluates to $\trg{0}$.
+
+\begin{center}
+  \typerule{$e-\trg{\lhast{x}{\natt}}$}{
+  }{
+    \pstepto{\rtt{\trg{\statevar}}{\trg{\lhast{x}{\natt}}}}{\rtt{\trg{\statevar}}{\trg{1}}}{\emptyevent}
+  }{e-has-nat-no}
+  %
+  \typerule{$e-\trg{\lhast{n}{\natt}}$}{
+  }{
+    \pstepto{\rtt{\trg{\statevar}}{\trg{\lhast{n}{\natt}}}}{\rtt{\trg{\statevar}}{\trg{0}}}{\emptyevent}
+  }{e-has-nat-yes}
+\end{center}
+
+Typechecks $\lhast{\expr}{\type}$ match on $\trg{\expr}$ and evaluate to $\trg{0}$ if the term is of type $\trg{\type}$ and $\trg{1}$ otherwise.
+
 %
 The projection $\msfilterL{\trg{\trace}}$ is entirely similar to $\msfilterLtms{\src{\trace}}$.
-\MP{ok, details}
 
 \subsection{$\Lms$: Another Memory-Unsafe Language}\label{subsec:lms}
-To enhance readablity, this paper uses $\Lms$, despite it being exactly equal to $\Ltrg$.
+To enhance readablity, this paper uses $\Lms$, despite it being exactly equal to $\Ltrg$ (\Cref{subsec:lsms}).
 The projection $\msfilterLms{\irl{\trace}}$ is also exactly equal to $\msfilterL{\trg{\trace}}$.

### 2023-07-11 popl/main.tex
up to example 5.2

@@
 In this trace model, events are either the empty event \emptyevent, the program crash event $\lightning$, or a tuple consisting of a control tag $\sandboxtag$ and a security tag $\securitytag$.
-The former indicates wether the component \comp or the context \ctx is to blame for emitting this event, the latter indicates the secrecy level of values of the emitted event, i.e., either $\lock$ or $\unlock$.
+The former indicates whether the component \comp or the context \ctx is to blame for emitting this event, the latter indicates the secrecy level of values of the emitted event, i.e., either $\lock$ or $\unlock$.
 As for pre-events, the memory-related ones are allocation ($\ev{Alloc}\ \loc\ \valueexpr$), deallocation ($\ev{Dealloc}\ \loc$), reading from ($\ev{Get}\ \loc\ \valueexpr$) and writing to memory ($\ev{Set}\ \loc\ \valueexpr\ \valueexpr[']$).
@@
 However, note that the execution does not get stuck in such cases, performs a step, and emits an appropriate event.
-Also note that wether a pointer is poisoned or not is not reflected on the trace.
-
-\begin{center}
-  \begin{aligned}
+Also note that whether a pointer is poisoned or not is not reflected on the trace.
+
+\begin{center}
+  \begin{align*}
     \mi{(Pre\text{-}Events)}~\event[_{b}]&\bnfdef \cdots \mid \ev{Call}\ \comm\ g\ \valueexpr \mid \ev{Ret}\ \comm\ \valueexpr \mid \ev{Start} \mid \ev{End}\ \valueexpr \\
-  \end{aligned}
+  \end{align*}
 \end{center}
@@
   }{
-    \msfilterLtms{\src{\tup{\event[_b]}{\ctx}} = \emptyevent}
+    \msfilterLtms{\src{\tup{\event[_b]}{\ctx,\securitytag}} = \emptyevent}
   }{ltms-filter-context}
@@
 \begin{center}
-  \begin{aligned}
+  \begin{align*}
     \mi{(Expressions)}~\trg{\expr}&\bnfdef \cdots \mid \trg{\lispoisoned{x}} \mid \trg{\lhast{\expr}{\type}} \\
-  \end{aligned}
+  \end{align*}
 \end{center}
@@
 
-\subsection{$\Lscct$: A Language with a Data Independent Timing Mode}\label{subsec:lscct}
-
-While the previous languages (\Cref{subsec:ltms,subsec:lsms,subsec:lms}) are not concerned with \gls{cct}, $\Lscct$ extends $\Lms$ (\Cref{subsec:lms}) with a register flag for a data (operand) independent timing mode, a feature that is present in both Arm~\cite[p.~543]{arm-refman} and Intel~\cite[p.~80]{intel-refman} processors.
-For the other languages seen earlier, this mode is intuitively always-on.
-\MP{spell out the formal changes in an align* environmtn} % this necessitates that we present $\Omega$ also in the shared lang defs
-To modify the mode, $\Lscct$ extends expressions with the term $\obj{\lwrdoit{\expr}}$ that writes the value of $\obj{\expr}$ into the register and with a term $\obj{\lrddoit{x}{\expr}}$ that reads the value of the register into $\obj{x}$.
-If the register is turned on, no secrets can appear on traces.
-If the register is turned off, secrets may appear on traces as well as new events $\obj{Branch\ n}$ and $\obj{Binop\ n}$, which the semantics emits for conditionals and certain binary operations, e.g., division.
-To prevent leakage of secrets but still enable reasoning about memory safety, pre-events are extended with $\obj{\widehat{{Get}}\ \loc\ \valueexpr}$ and $\obj{\widehat{Set}\ \loc\ \valueexpr}$, which indicate reads from and writes to memory without leaking secret information involved in the access, which is taken care of by the projection to specification events.
-\Cref{ex:lscct} illustrates these changes.
-\MPin{formalisation additions in algin* env}
-
-The projection to the specification trace model is mostly straightforward.
-But for events containing the pre-events $\obj{\widehat{Get}}$ and $\obj{\widehat{Set}}$, the projection always translates the security-tag $\obj{\sigma}$ to $\unlock$, regardless of its actual value.
+\subsection{$\Lscct$: A Memory-Unsafe Language with a Data Independent Timing Mode}\label{subsec:lscct}
+
+\begin{center}
+  \begin{align*}
+    \mi{(Expressions)}~\obj{\expr}&\bnfdef \obj{n^{\securitytag}} \mid \cdots \mid \obj{\llet{x^{\securitytag}}{\expr[_{1}]}{\expr[_{2}]}} \mid \cdots \mid \obj{\lwrdoit{\expr}} \mid \obj{\lrddoit{x}{\expr}}
+  \end{align*} \\
+  \begin{align*}
+    \mi{(States)}~\obj{\statevar}&\bnfdef \obj{\tup{\cfstate}{\sandboxtag,n,\memstate}}
+  \end{align*}
+\end{center}
+
+$\Lscct$ extends $\Lms$ (\Cref{subsec:lms}) with a way to write to a model specific register that controls a data (operand) independent timing mode, a feature that is present in both Arm~\cite[p.~543]{arm-refman} and Intel~\cite[p.~80]{intel-refman} processors.
+To this end, states are extended with the value of the register, which is initially set to be not active.
+If the register is marked active, the intuition is that no secrets can appear on specification traces.
+If the register is marked inactive, secrets may appear on traces.
+For the other languages seen earlier, the mode is intuitively always-on, i.e., the mode of execution always uses data independent timing.
+The language also adds user-annotations to values and variables to know their secrecy $\obj{\securitytag}$, which is either high $\obj{\lock}$ or low $\obj{\unlock}$.
+Security tags $\obj{\securitytag}$ are on the usual secrecy lattice, where $\obj{\lock}\le\obj{\securitytag}$ and $\obj{\securitytag}\le\obj{\unlock}$.
+
+\begin{center}
+  \begin{align*}
+    \mi{(Pre\text{-}Events)}~\obj{\event[_{b}]}&\bnfdef \cdots \mid \obj{\widehat{Get}\ \loc\ \valueexpr} \mid \obj{\widehat{Set}\ \loc\ \valueexpr\ \valueexpr[']} \mid \obj{Branch\ n} \mid \obj{Binop\ n}
+  \end{align*}
+\end{center}
+
+To prevent secrets from leaking but still enable reasoning about memory safety, $\Lscct$ extends pre-events with $\obj{\widehat{Get}\ \loc\ \valueexpr}$ and $\obj{\widehat{Set}\ \loc\ \valueexpr}$.
+These indicate reads from and writes to memory without leaking secret information involved in the access.
+Moreover, the language extends pre-events with $\obj{\ev{Branch}\ n}$ and $\obj{\ev{Binop}\ n}$ that are emitted when evaluating a branch or certain binary expressions, such as division, respectively, whenever the data independent timing mode is inactive.
+
+\begin{center}
+  \typerule{$e-\obj{\oplus}-\text{noleak}$}{
+    \obj{m}\not=\obj{0} &
+    \obj{n_3}=\obj{n_1}\oplus\obj{n_2} \\
+    \obj{\securitytag['']}\le\obj{\securitytag} &
+    \obj{\securitytag['']}\le\obj{\securitytag[']}
+  }{
+    \pstepto{\rtt{\obj{\cfstate;\sandboxtag;m;\memstate}}{\obj{\lbinop{n_1^{\securitytag}}{n_2^{\securitytag[']}}}}}{\rtt{\obj{\cfstate;\sandboxtag;m;\memstate}}{\obj{n_3^{\securitytag['']}}}}{\obj{\emptyevent}}
+  }{e-binop-noleak}
+  %
+  \typerule{$e-\obj{wrdoit}$}{
+  }{
+    \pstepto{\rtt{\obj{\cfstate;\sandboxtag;m;\memstate}}{\obj{\lwrdoit{n^{\securitytag}}}}}{\rtt{\obj{\cfstate;\sandboxtag;n;\memstate}}{\obj{n^{\securitytag}}}}{\obj{\emptyevent}}
+  }{e-wrdoit}
+  %
+  \typerule{$e-\obj{ifz}-\obj{true}-\text{leak}$}{
+  }{
+    \pstepto{\rtt{\obj{\cfstate;\sandboxtag;0;\memstate}}{\obj{\lifz{0^{\securitytag}}{\expr[_1]}{\expr[_2]}}}}{\rtt{\obj{\cfstate;\sandboxtag;0;\memstate}}{\obj{\expr[_2]}}}{\obj{\ev{Branch}\ 0;\sandboxtag;\securitytag}}
+  }{e-ifz-true-leak}
+\end{center}
+
+The evaluation steps are amended to propagate the securitytags $\obj{\securitytag}$.
+When the data independent timing mode is active, pre-events $\obj{\ev{Branch}\ n}$ and $\obj{\ev{Binop}\ n}$ are emitted for conditionals and binary operations, respectively.
+To mark this mode as active or inactive, \Cref{tr:e-wrdoit} allows to access the register in $\obj{\statevar}$.
+
+\begin{center}
+  \judgbox{\scctfilterLscct{\obj{\event}}=\event}{,,Project an $\Lscct$ event $\obj{\event}$ to a specification event $\event$.''}
+  \typerule{$\Lscct$-filter-context}{
+    \obj{\event[_b]}\not=\obj{\lightning}
+  }{
+    \scctfilterLscct{\obj{\tup{\event[_b]}{\ctx,\securitytag}} = \emptyevent}
+  }{lscct-filter-context}
+  %
+  \typerule{$\Lscct$-filter-$\widehat{\text{get}}$}{
+    \locmapsto{\obj{\loc}}{\loc} &
+    \obj{n}=n
+  }{
+    \scctfilterLscct{\obj{\tup{\widehat{Get}\ \loc\ n}{\comp,\securitytag}}} = \tup{\ev{Get}\ \loc\ n}{\comp,\unlock}
+  }{lscct-filter-get-hat}
+  %
+  \typerule{$\Lscct$-filter-$\text{get}$}{
+    \locmapsto{\obj{\loc}}{\loc} &
+    \obj{n}=n
+  }{
+    \scctfilterLscct{\obj{\tup{Get\ \loc\ n}{\comp,\securitytag}}} = \tup{\ev{Get}\ \loc\ n}{\comp,\securitytag}
+  }{lscct-filter-get}
+\end{center}
+
+The projection to the specification trace model is mostly straightforward and similar to the others, e.g., \Cref{subsec:cs:tms}.
+However, for events containing the pre-events $\obj{\widehat{Get}}$ and $\obj{\widehat{Set}}$, the projection always translates the security-tag $\obj{\sigma}$ to $\unlock$, regardless of its actual value, as seen in \Cref{tr:lscct-filter-get-hat}.
 The pre-events themselves still translate to just $\ev{Get}$ and $\ev{Set}$, respectively.
-With this technical setup, the information wether\MP{spelling} a read or write happened on a secret value is not hidden by the semantics, e.g., by emitting $\obj{\emptyevent}$, but when projecting to specification events.
+With this technical setup, the information whether a read or write happened on a secret value is not hidden by the semantics, e.g., by emitting $\obj{\emptyevent}$, but when projecting to specification events.
 This allows flexibility: The trace can be checked to satisfy different properties, such as, in this case, \gls{tms}, \gls{sms}, \gls{scct}, and their combined versions.
+\Cref{ex:lscct} illustrates the differences of $\Lscct$ compared to the other languages.
 

### 2023-07-11 popl/main.tex
ex5.2

@@
 
-\begin{example}[A program not satisfying \gls{tms}]
+\begin{example}[A program not satisfying \gls{tms}]\label{ex:strncpy:cpp11}
   Consider the following \texttt{C++11} library that calls \texttt{strncpy} (\Cref{ex:strncpy}) and prints the result to the standard output stream.
@@
   \end{lstlisting}
-  Up to the body of \texttt{printf}, the program execution yields the specification trace $\ev{Alloc}\ \loc[_{x}]\ 12;\comp;\unlock\cdot\ev{Alloc}\ \loc[_{y}]\ 12;\comp;\unlock\cdot\ev{Use}\ \loc_{x}\ 0;\comp;\unlock\cdot\ev{Use}\ \loc_{y}\ 0;\comp;\unlock\cdot\ev{Use}\ \loc_{x}\ 1;\comp;\unlock\cdot\ldots\cdot\ev{Use}\ \loc_{x}\ 12;\comp;\unlock\cdot\ev{Dealloc}\ \loc[_{y}];\comp;\unlock\cdot\ev{Use}\ \loc_{y}\ 6;\comp;\unlock$.
-  Relating this trace to abstract monitor events yields $\abstrace=\absev{Alloc}\ \loc[_{x}]\ \comp\cdot\absev{Alloc}\ \loc[_{y}]\ \comp\cdot\absev{Use}\ \loc[_{x}]\ \comp\cdot\absev{Use}\ \loc[_{y}]\ \comp\cdot\ldots\cdot\absev{Use}\ \loc[_{x}]\ \comp\cdot\absev{Dealloc}\ \loc[_{y}]\ \comp\cdot\absev{Use}\ \loc[_{y}]\ \comp$.
+  Up to the body of \texttt{printf}, the program execution yields the specification trace $\ev{Alloc}\ \loc[_{x}]\ 12;\comp;\unlock\cdot\ev{Alloc}\ \loc[_{y}]\ 12;\comp;\unlock\cdot\ev{Use}\ \loc_{x}\ 0;\comp;\unlock\cdot\ev{Use}\ \loc_{x}\ 0;\comp;\unlock\cdot\ev{Use}\ \loc_{y}\ 0;\comp;\unlock\cdot\ev{Use}\ \loc_{x}\ 1;\comp;\unlock\cdot\ldots\cdot\ev{Use}\ \loc_{x}\ 12;\comp;\unlock\cdot\ev{Dealloc}\ \loc[_{y}];\comp;\unlock\cdot\ev{Use}\ \loc_{y}\ 6;\comp;\unlock$.
+  Relating this trace to abstract monitor events yields $\abstrace=\absev{Alloc}\ \loc[_{x}]\ \comp\cdot\absev{Alloc}\ \loc[_{y}]\ \comp\cdot\absev{Use}\ \loc[_{x}]\ \comp\cdot\absev{Use}\ \loc[_{x}]\ \comp\cdot\absev{Use}\ \loc[_{y}]\ \comp\cdot\ldots\cdot\absev{Use}\ \loc[_{x}]\ \comp\cdot\absev{Dealloc}\ \loc[_{y}]\ \comp\cdot\absev{Use}\ \loc[_{y}]\ \comp$.
   Remembering the definition of \texttt{strncpy} (\Cref{ex:strncpy}), observe that it does not deallocate its arguments.
@@
   However, having returned from \texttt{strncpy}, the \texttt{greet} function continues and deallocates $\loc[_{y}]$ whose subsequent use in the \texttt{printf} call is a use-after-free bug.
-  The fix would be to \lstinline[language=c++,basicstyle=\small\ttfamily]|delete greetings| instead of \texttt{to} and add a \lstinline[language=c++,basicstyle=\small\ttfamily]|delete to| after the \texttt{printf} call, which leads to the abstract monitor trace $\abstrace[']=\absev{Alloc}\ \loc[_{x}]\ \comp\cdot\absev{Alloc}\ \loc[_{y}]\ \comp\cdot\absev{Use}\ \loc[_{x}]\ \comp\cdot\absev{Use}\ \loc[_{y}]\ \comp\cdot\ldots\cdot\absev{Use}\ \loc[_{x}]\ \comp\cdot\absev{Dealloc}\ \loc[_{x}]\ \comp\cdot\absev{Use}\ \loc[_{y}]\ \comp$.
+  The fix would be to \lstinline[language=c++,basicstyle=\small\ttfamily]|delete greetings| instead of \texttt{to} and add a \lstinline[language=c++,basicstyle=\small\ttfamily]|delete to| after the \texttt{printf} call, which leads to the abstract monitor trace $\abstrace[']=\absev{Alloc}\ \loc[_{x}]\ \comp\cdot\absev{Alloc}\ \loc[_{y}]\ \comp\cdot\absev{Use}\ \loc[_{x}]\ \comp\cdot\absev{Use}\ \loc[_{x}]\ \comp\cdot\absev{Use}\ \loc[_{y}]\ \comp\cdot\ldots\cdot\absev{Use}\ \loc[_{x}]\ \comp\cdot\absev{Dealloc}\ \loc[_{x}]\ \comp\cdot\absev{Use}\ \loc[_{y}]\ \comp$.
   It follows that $\montmssafe{\abstrace[']}$ and from \Thmref{lem:mon:tmssafe}, it follows that the program satisfies \Thmref{def:trace:tmsdef}, even though the program still violates \gls{sms}.
@@
 \begin{example}[$\Lscct$ with and without data independent timing]\label{ex:lscct}
-  Consider the following $\Lscct$ program and the execution trace $\obj{\trace}$ juxtaposed with it.
-
+  Consider again the context presented in \Cref{ex:strncpy:cpp11}, where everything is marked with a security tag of high $\lock$.
+  The following table shows parts of the execution trace, read from top to bottom, in the left column with and in the right column without data independent timing.
+  The left side of the table, i.e., the two columns on the left, describe the execution trace of the program, while the right side of the table, i.e., the two columns on the right, describe the respective projections $\scctfilterLscct{\obj{\trace}}$ to the specification trace model.
+
+  $${\footnotesize
+  \begin{array}{ccccccc}
+        & \Lscct &        &   &      & Spec & \\
+    Active & \mid & Inactive & \mid & Active & \mid & Inactive \\\hline
+    \obj{\ev{Alloc}\ \loc_{x}\ 12;\comp;\unlock} & \mid & \obj{\ev{Alloc}\ \loc_{x}\ 12;\comp;\unlock} & \mid & \ev{Alloc}\ \loc_{x}\ 12;\comp;\unlock & \mid & \ev{Alloc}\ \loc_{x}\ 12;\comp;\unlock\\
+    \obj{\ev{Alloc}\ \loc_{y}\ 12;\comp;\unlock} & \mid & \obj{\ev{Alloc}\ \loc_{y}\ 12;\comp;\unlock} & \mid & \ev{Alloc}\ \loc_{y}\ 12;\comp;\unlock & \mid & \ev{Alloc}\ \loc_{x}\ 12;\comp;\unlock\\
+    \obj{\widehat{\ev{Get}}\ \loc_{x}\ 0;\comp;\lock} & \mid & \obj{\ev{Get}\ \loc_{x}\ 0;\comp;\lock} & \mid & \ev{Use}\ \loc_{x}\ 0;\comp;\unlock & \mid & \ev{Use}\ \loc_{x}\ 0;\comp;\lock\\
+    \obj{\emptyevent} & \mid & \obj{\ev{Branch}\ 0;\comp;\lock} & \mid & \emptyevent & \mid & \ev{Branch}\ 0;\comp;\lock\\
+    \obj{\widehat{\ev{Get}}\ \loc_{x}\ 0;\comp;\lock} & \mid & \obj{\ev{Get}\ \loc_{x}\ 0;\comp;\lock} & \mid & \ev{Use}\ \loc_{x}\ 0;\comp;\lock & \mid & \ev{Use}\ \loc_{x}\ 0;\comp;\lock\\
+    \obj{\widehat{\ev{Set}}\ \loc_{y}\ 0\ \mathtt{'H'};\comp;\lock} & \mid & \obj{\ev{Set}\ \loc_{y}\ 0\ \mathtt{'H'};\comp;\lock} & \mid & \ev{Use}\ \loc_{y}\ 0;\comp;\unlock & \mid & \ev{Use}\ \loc_{y}\ 0;\comp;\lock\\
+    \obj{\widehat{\ev{Get}}\ \loc_{x}\ 1;\comp;\lock} & \mid & \obj{\ev{Get}\ \loc_{x}\ 1;\comp;\lock} & \mid & \ev{Use}\ \loc_{x}\ 1;\comp;\unlock & \mid & \ev{Use}\ \loc_{x}\ 1;\comp;\lock\\
+    \obj{\emptyevent} & \mid & \obj{\ev{Branch}\ 0;\comp;\lock} & \mid & \emptyevent & \mid & \ev{Branch}\ 0;\comp;\lock\\
+    \vdots & \mid & \vdots & \mid & \vdots & \mid & \vdots \\
+    \obj{\widehat{\ev{Get}}\ \loc_{x}\ 12;\comp;\lock} & \mid & \obj{\ev{Get}\ \loc_{x}\ 12;\comp;\lock} & \mid &  \ev{Use}\ \loc_{x}\ 12;\comp;\unlock & \mid & \ev{Use}\ \loc_{x}\ 12;\comp;\lock\\
+    \obj{\emptyevent} & \mid & \obj{\ev{Branch}\ 1;\comp;\lock} & \mid &  \emptyevent & \mid & \ev{Branch}\ 1;\comp;\lock\\
+    \obj{\ev{Dealloc}\ \loc_{y};\comp;\unlock} & \mid & \obj{\ev{Dealloc}\ \loc_{y};\comp;\lock} & \mid &  \ev{Dealloc}\ \loc_{y};\comp;\unlock & \mid & \ev{Dealloc}\ \loc_{y};\comp;\unlock\\
+    \obj{\widehat{\ev{Get}}\ \loc_{y}\ 6;\comp;\lock} & \mid & \obj{\ev{Get}\ \loc_{y}\ 6;\comp;\lock} & \mid &  \ev{Use}\ \loc_{y}\ 6;\comp;\unlock & \mid & \ev{Use}\ \loc_{y}\ 6;\comp;\lock\\
+  \end{array}}
   $$
-  \begin{array}{lcc}
-  \obj{\lwrdoit{0};} & \mid & \obj{\emptyevent}\\
-  \obj{\llet{x_{\lock}}{\lnew{5}{4}}{}} & \mid & \obj{\ev{Alloc}\ \loc_{x}\ 4;\sandboxtag;\unlock} \\
-  \obj{\lset{x}{2}{\lifz{0_{\lock}}{1}{3}};} & \mid & \obj{\ev{Branch}\ 0;\sandboxtag;\lock\cdot\ev{Set}\ \loc_{x}\ 2\ 1;\sandboxtag;\lock} \\
-  \obj{\lwrdoit{1};} & \mid & \obj{\emptyevent}\\
-  \obj{\llet{y_{\lock}}{\lnew{5}{4}}{}} & \mid & \obj{\ev{Alloc}\ \loc_{y}\ 4;\sandboxtag;\unlock}\\
-  \obj{\lset{y_{\lock}}{2}{\lifz{0_{\lock}}{1}{3}};} & \mid & \obj{\emptyevent\cdot\widehat{Set}\ \loc_{y}\ 2;\sandboxtag;\lock}\\
-  \end{array}
-  $$
 
   When the data independent timing mode is off, the execution yields events in similar fashion to before (\Cref{subsec:ltms,subsec:lsms,subsec:lms}).
-  But, if it is turned on, then the branching event does not fire anymore.
-  Moreover, writing to memory now yields the pre-event $\obj{\widehat{Set}}$, whose difference to $\obj{Set}$ becomes apparent when looking at the projection to the specification trace model.
-  \MPin{shouldn't we have 2 traces, one with the bit set, one without}
-  Projecting the above trace to the specification trace model yields the following trace (emitting empty events):
-  $\ev{Alloc}\ \loc_{x}\ 4;\sandboxtag;\unlock\cdot\ev{Branch\ 0}\cdot\ev{Set}\ \loc_{x}\ 2\ 1;\sandboxtag;\lock\cdot\ev{Alloc}\ \loc_{y}\ 4;\sandboxtag;\unlock\cdot\ev{Set}\ \loc_{y}\ 2;\sandboxtag;\unlock$.
-  Clearly, secrets are hidden ($\unlock$) after this projection for the section of the program where the data independent timing mode is turned on, while still containing the appropriate events to check for \gls{ms}.
+  But, if it is turned on, then the branching event does not fire anymore and both reading and writing to memory gets ultimately translated to a specification trace with no exposed secrets.
 \end{example}
@@
 No existing work is concerned with composing robustly safe compilers.
-\MPin{
-  perhaps say taht we believe this scales to different trace models too and differnet kinds of props, but the ocmposition of monitors becomes tricky
-  \MKin{conclusion?}
-}
+These works consider languages with different trace models and our technical setup can be adapted to that as long as security properties and their monitors are still defined on the same trace model.
 The work relating robust preservation with universal composability~\cite{patrignani2022universal} is closest to what this paper presents.

### 2023-07-11 popl/main.tex
typo fix

@@
 \subsection{Robust Temporal Memory Safety Preservation}\label{subsec:cs:tms}
-\MPin{some text should introduce the compiler defs}
-
-  \begin{align*}
-%    \cca(\src{\lbinop{\expr[_{1}]}{\expr[_{2}]}}) & = & \lbinop[\trg]{\left[\cca(\src{\expr[_{1}]})\right]}{\left[\cca(\src{\expr[_{2}]})\right]} \\ [0.33cm]
-    \cca(\src{\lfunction{g}{x:\natt\to\type[_{e}]}{\expr}}) = \lfunction[\trg]{\trg{g}}{\trg{x}}{\lifz[\trg]{\trg{\lhast{x}{\natt}}}{%
+
+This subsection defines a secure compiler from $\Ltms$ to $\Ltrg$.
+To this end, the compiler needs to ensure that when execution switches from context to component, the type signatures are respected.
+It can do so by inserting dynamic typechecks prior to entering the body of a function belonging to the component.
+
+\begin{gather*}
+  \begin{aligned}
+    \cca(\src{x}) & = \trg{x} \\
+    \cca(\src{n}) & = \trg{n} \\
+    \cca(\src{\lbinop{\expr[_{1}]}{\expr[_{2}]}}) & =  \lbinop[\trg]{\left[\cca(\src{\expr[_{1}]})\right]}{\left[\cca(\src{\expr[_{2}]})\right]} \\
+    \cca(\src{\lget{x}{\expr}}) & = \lget[\trg]{\trg{x}}{\left[\cca(\src{\expr})\right]} \\
+    \cca(\src{\ldelete{x}}) & = \ldelete[\trg]{\left[\cca(\src{x})\right]} \\
+%    \cca(\src{\lset{x}{\expr[_{1}]}{\expr[_{2}]}}) & = \lset[\trg]{\left[\cca(\src{x})\right]}{\left[\cca(\src{\expr[_{1}]})\right]}{\left[\cca(\src{\expr_{2}})\right]} \\
+    \cca(\src{\lfunction{g}{x:\natt\to\type[_{e}]}{\expr}}) & = \lfunction[\trg]{\trg{g}}{\trg{x}}{\lifz[\trg]{\trg{\lhast{x}{\natt}}}{%
                                                                                                 \left[\cca(\src{\expr})\right] %
                                                                                                  }{\labort[\trg]}}
-  \end{align*}
-
-The first compiler only inserts a dynamic typecheck to respect the well-typedness of function bodies.
+  \end{aligned}
+\end{gather*}
+
 Since $\trg{L}$ has no static typechecks, it could happen that a bogus context $\trg{\library_{\ctx}}$ invokes a callable object accepting a $\src{\natt}$ with $\trg{\lpair{17}{29}}$.
@@
   \begin{array}{rcl}
-    \ccscct(\irl{\lfunction{g}{x}{\expr}}) & = & \lfunction[\obj]{\obj{g}}{\obj{x}}{\obj{\lwrdoit{1};}\ccscct(\irl{\expr})}
+    \ccscct(\irl{\lfunction{g}{x}{\expr}}) & = & \lfunction[\obj]{\obj{g}}{\obj{x}}{\obj{\lwrdoit{1};}\ccscct(\irl{\expr})} \\
+    \ccscct(\irl{\lcall{g}{\expr}}) & = & \lcall[\obj]{\obj{g}}{\ccscct(\expr)\obj{; wrdoit{1}}}
   \end{array}

### 2023-07-11 popl/main.tex
intro

--- 2023-07-11 popl/main.tex
@@
 Robust compilers~\cite{abate2019jour} are a form of secure compilers that preserve security properties even in the presence of arbitrary attackers interacting with compiled code.
-Thus, robust compilers can be used to prevent these vulnerabilities, e.g., by targeting capability-based languages such as CHERI~\cite{woodruff2014CHERI}, Arm Morello~\cite{arm-morello}, or MSWasm~\cite{michael2023mswasm}, where pointers are always initialised.
+Thus, robust compilers can be used to prevent the vulnerability resulting from uninitialised memory, e.g., by targeting capability-based languages such as CHERI~\cite{woodruff2014CHERI}, Arm Morello~\cite{arm-morello}, or MSWasm~\cite{michael2023mswasm}, where pointers are always initialised.
 \end{example}
@@
 % Different kinds of compilation passes exist, e.g., there are usually several compilation passes that perform optimisations~\cite{lattner2004llvm,googlev8,androidstudio,kuepper2023cryptopt,manjikian1997fusion,wegman1991ccp} or source-code instrumentations~\cite{nagarakatte2009soft,nagarakatte2010cets,akritidis2009baggy,dhumbumroong2020boundwarden,jung2021pico,nam2019framer,shankaranarayana2023tailcheck,younan2010paricheck,zhou2023fatptrs,bond2017vale,almeida2017jasmin,kuepper2023cryptopt,cauligi2019fact} that enforce that the compiled program fulfils a property of interest.
-Worse, without a framework for composing secure compiler passes, it is not possible to enable separation of concerns, e.g., to have a secure compilation pass that ensures \gls{ms} that is developed independently of another secure pass for \gls{cct}.
+Worse, without a framework for composing secure compiler passes, it is not possible to enable separation of concerns, e.g., to have a secure compilation pass that ensures \gls{ms} that is developed independently of another secure pass for \gls{cct}, that is developed independently of optimisation passes.
 
@@
 Finally, this paper shows that composing these secure compiler passes into a multi-pass compilation chain results that provides end-to-end preservation of \gls{msscct}.
-Crucially, this end-to-end security preservation is obtained by composing the security property preserved by each individual pass, as dictated by the framework of this paper.
-The results showcase how the framework allows the kind of formal security reasoning that compiler writers already want (and do), obtaining precise, compositional security reasoning while providing minimal (and modular) proof effort.
+Crucially, this paper derives the security of the multi-pass compiler from the composition of the security properties preserved by its individual passes.
+% Crucially, this end-to-end security preservation is obtained by composing the security property preserved by each individual pass, as dictated by the framework of this paper.
+This result showcases how the framework allows the kind of formal security reasoning that compiler writers already want (and do), obtaining precise, compositional security reasoning while providing minimal (and modular) proof effort.
 
@@
         %\Cref{sec:compprop} presents a formal framework to reason about compositions of security properties and demonstrate its usage with concrete examples:
-        First, this paper formalises security properties (\Cref{sec:compprop}) by focussing on those properties that are of interest for real-world compiler writers (as identified by the plethora of work enforcing such properties individually~\cite{akritidis2009baggy,nagarakatte2009soft,nagarakatte2010cets,dhumbumroong2020boundwarden,jung2021pico,nam2019framer,shankaranarayana2023tailcheck,younan2010paricheck,zhou2023fatptrs,bond2017vale,almeida2017jasmin,kuepper2023cryptopt,cauligi2019fact}), namely \gls{tms}, \gls{sms} and \gls{cct}.
-        % 
-        Starting from ways to formalise and enforce those properties individually, this paper shows how to compose their formalisation as well as their enforcement.
+        This paper formalises security properties (\Cref{sec:compprop}) that are of interest for real-world compiler writers (as identified by the plethora of work enforcing such properties individually~\cite{akritidis2009baggy,nagarakatte2009soft,nagarakatte2010cets,dhumbumroong2020boundwarden,jung2021pico,nam2019framer,shankaranarayana2023tailcheck,younan2010paricheck,zhou2023fatptrs,bond2017vale,almeida2017jasmin,kuepper2023cryptopt,cauligi2019fact}), namely \gls{tms}, \gls{sms} and \gls{cct}.
+        % 
+        Starting from ways to formalise those properties individually, this paper shows how to compose their formalisation.
         % 
@@
   \begin{aligned}
-    \mi{(Expressions)}~&\expr\bnfdef\ x \mid \valueexpr \mid \lbinop{\expr[_{1}]}{\expr[_{2}]} \mid \lget{x}{\expr} \mid \llet{x}{\lnew{\expr[_{1}]}{\expr[_{2}]}}{\expr[_{3}]} \mid \ldelete{x} \mid \lset{x}{\expr[_{1}]}{\expr[_{2}]} \\
-                       & \mid \lpair{\expr[_{1}]}{\expr[_{2}]} \mid \lproja{\expr} \mid \lprojb{\expr} \mid \llet{x}{\expr[_{1}]}{\expr[_{2}]} \mid \lreturn{\expr} \mid \lcall{g}{\expr} \mid \lifz{\expr}{\expr[_{1}]}{\expr[_{2}]} \mid \labort
+  \mi{(Expressions)}~&\expr\bnfdef\ x \mid \valueexpr \mid \lbinop{\expr[_{1}]}{\expr[_{2}]} \mid \lget{x}{\expr} \mid \lnew{\expr[_{1}]}{\expr[_{2}]} \mid \ldelete{x} \mid \lset{x}{\expr[_{1}]}{\expr[_{2}]} \\
+    & \mid \llet{x}{\expr[_{1}]}{\expr[_{2}]}  \mid \lreturn{\expr} \mid \lcall{g}{\expr} \mid \lifz{\expr}{\expr[_{1}]}{\expr[_{2}]} \mid \labort
   \end{aligned}
@@
   \begin{aligned}
-  \mi{(Types)}~&\type\bnfdef\natt \mid \type[_{1}] \times \type[_{2}]
+  \mi{(Functions)}~&\function\bnfdef\ \lfunction{g}{x}{e}
   &
-  \mi{(Functions)}~&\function\bnfdef\ \lfunction{\textit{foo}}{x}{e}
+  \mi{(Libraries)}~&\library\bnfdef\ \hole{\cdot} \mid \function,\library \\
+  \mi{(Programs)}~&\prog{\library_{\ctx}}{\library_{\comp}}
   &
-  \mi{(Libraries)}~&\library\bnfdef\ \hole{\cdot} \mid \function,\library
-  \end{aligned}\\
-  \begin{aligned}
-  \mi{(Component~Names)}~&\commlib\bnfdef\ \hole{\cdot} \mid \text{foo},\commlib
-  &
-  \mi{(Programs)}~&\prog{\library_{\ctx}}{\library_{\comp}}
+  \mi{(Substitutions)}~&\substvar\bnfdef \hole{\cdot} \mid \subst{\valueexpr}{x},\gamma
   \end{aligned}
 \end{gather*}
-
-Above is the shared syntax of all the programming languages of this paper.
-Variables are referred to as $x,y,z,a,b,c,\ldots$ while functions may be referred to as multi-character words, such as $\texttt{foo}$, as well as short-forms like $f,g,h$.
-All languages share the type $\natt$ representing natural numbers.
-Functions are constrained to take one argument and can only call other functions listed in libraries, which are just lists of functions.
-A program $\prog{\library_{\ctx}}{\library_{\comp}}$ is indexed by two libraries that represent all context- and component-level functions, respectively.
-Lists of component-level names are referred to as $\commlib$.
+% \MPin{
+%   this latex had many issues.
+%   the \& were at different spots
+%   it used hspace to space things instead of \&
+%   and it did not use \& to align the following things on a line
+%   and it did not use multiple aligned into the gather
+%   (what's the point of gather if you don't have several aligned inside? use a single align* instead)
+% }
+% oh... that's good to know! the tech report is also riddled with this... :/
+
+Above is a fraction of the shared syntax of all the programming languages of this paper.
+\MPin{why a fraction? what exprs are missing
+MK: Things that are missing here are: Evaluation contexts, states $\Omega$ and transitively other definitions affected by this
+}
+The syntax and semantics are mostly standard and therefore omitted.
+However, when allocating arrays with $\lnew{\expr[_{1}]}{\expr[_{2}]}$, the expression $\expr[_{1}]$ determines the initial value of the cells and $\expr[_{2}]$ the overall size of an array.
+
+\MPin{
+	below is semantics info. should go with their symbols and the semantics rules.
+	and a better description of the semantics symbols.
+  MK: this moved to appendix! should we bring it back? what exactly should be left out?
+}
+The data associated to pointers resides in the dynamic state $\Omega$ that contains sandboxed heaps, one for the context and the other for the component.
+Because of the sandboxing, locations pointing into heaps must be associated with a control-tag $\sandboxtag$ (either $\ctx$ or $\comp$) to know the heap they point into.
+Other metadata in $\Omega$ entails a list of functions with information on whether they belong to the context or the component.
+That way, when calling a function, the semantics can take care of emitting appropriate events that signal a context switch.
+For the languages considered in this paper, functions are second-class.
+So, the only values are pairs, pointers, and natural numbers.
+Another non-standard part of the languages is the concept of the poison-tag.
+The poison-tag marks an array as freed ($\poisoned$) without removing the information associated to the freed pointer from $\Omega$.
+This measure prevents the semantics from getting stuck when an already freed pointer is freed again.
+Because of this, it is possible that a program like $\llet{x}{\lnew{0}{5}}{\ldelete{x};\ldelete{x}}$ yields a trace $\ev{Alloc\ \loc\ \sandboxtag\ 5}\cdot\ev{Dealloc\ \loc\ \sandboxtag}\cdot\ev{Dealloc\ \loc\ \sandboxtag}$ when executing.
 
@@
   \begin{aligned}
-    \mi{(Control~Tags)}~&\sandboxtag\bnfdef\ctx \mid \comp
-    &
-    \mi{(Communication~Tags)}~&\comm\bnfdef\ \ctxtocomp \mid\ \comptoctx \mid \nocomm
-    &
-    \mi{(Poison~Tags)}~&\poison\bnfdef \poisoned \mid \poisonless
-  \end{aligned}\\
-  \begin{aligned}
-    \mi{(Continuation~Stacks)}~&\ectxstack\bnfdef\hole{\cdot} \mid \tup{K}{g},\ectxstack
-    &
-    \mi{(Control~Flow~States)}~&\cfstate\bnfdef\tup{\library}{\ectxstack}
-  \end{aligned}\\
-  \begin{aligned}
-    \mi{(Locations)}~&\loc\in\nat
-    &
-    \mi{(Substitutions)}~&\substvar\bnfdef \hole{\cdot} \mid \subst{\valueexpr}{x},\gamma
-  \end{aligned}\\
-  \begin{aligned}
-    \mi{(Stores)}~&\Delta\bnfdef\hole{\cdot} \mid x\mapsto\tup{\loc}{\sandboxtag,\poison,n},\Delta
-    &
-    \mi{(Heaps)}~&H\bnfdef\hole{\cdot} | \valueexpr,H
-    \\
-    \mi{(Memory~States)}~&\memstate\bnfdef\tup{H^{\ctx}}{H^{\comp};\Delta}
-    &
-    \mi{(States)}~&\statevar\bnfdef\tup{\cfstate}{\sandboxtag,\memstate}
-    &
-    \mi{(Runtime~Terms)}~&\runtimetermvar\bnfdef\rtt{\statevar}{\expr}
+  \mi{(Pre\text{-}Events)}~\event[_{b}]&\bnfdef \ev{Alloc}\ \loc\ \valueexpr \mid \ev{Dealloc}\ \loc \mid \ev{Get}\ \loc\ \valueexpr \mid \ev{Set}\ \loc\ \valueexpr\ \valueexpr['] \mid \ev{Call}\ \comm\ g\ \valueexpr \mid \ev{Ret}\ \comm\ \valueexpr \mid \ev{Start} \mid \ev{End}\ \valueexpr \\
+  \mi{(Events)}~\event&\bnfdef \emptyevent \mid \lightning \mid \tup{\event[_{b}]}{\sandboxtag} \\
   \end{aligned}
@@
 
-States $\statevar$ are tuples\footnote{Throughout the paper, the subsitution notation is also used to update entries in states $\statevar$.} containing a control flow state $\cfstate$, control tags $\sandboxtag$, and a memory state $\memstate$.
-A control flow state $\cfstate$ entails a library, which provides definitions for functions when calling, as well as a stack of continuations $\ectxstack$.
-Elements of the stack $\ectxstack$ are pairs of evaluation contexts and the name of a function associated to that evaluation context.
-Memory states $\memstate$ are tuples of two separate heaps $H$ and a store $\Delta$, which contains pointer metadata, such as the concrete memory location $\loc$, a control tag $\sandboxtag$ indicating which heap the pointer points into, a poison tag $\poison$, as well as bounds information $n$.
-The bounds information is a mere proof artifact that has no semantic significance.
-For details see the accompanying tech report~\MK{cit}.
-Memory states $\memstate$ have two separate heaps, essentially modelling a sandbox, to prevent contexts from performing pointer arithmetic and reading from or writing to the data owned by a component.
-While this prevents the effects of out of bounds accesses across the context and component boundary, the goal of the design of the languages of this paper is to be able to express security violations.
-To this end, the poison tag $\poison$ indicates whether a pointer has been freed $\poisoned$ or is still allocated $\poisonless$, so that uses of pointers can happen even after their deallocation without the semantics getting stuck.
-Runtime terms are simply expressions $\expr$ paired with the operational state $\statevar$.
-
-\begin{gather*}
-  \begin{aligned}
-  \mi{(Pre\text{-}Events)}~\event[_{b}]&\bnfdef \ev{Alloc}\ \loc\ \valueexpr \mid \ev{Dealloc}\ \loc \mid \ev{Get}\ \loc\ \valueexpr \mid \ev{Set}\ \loc\ \valueexpr\ \valueexpr['] \mid \cdots \\
-  \mi{(Events)}~\event&\bnfdef \emptyevent \mid \lightning \mid \tup{\event[_{b}]}{\sandboxtag,\securitytag} \\
-  \end{aligned}
-\end{gather*}
-
-All languages use the same trace model.
-In this trace model, events are either the empty event \emptyevent, the program crash event $\lightning$, or a tuple consisting of a control tag $\sandboxtag$ and a security tag $\securitytag$.
-The former indicates whether the component \comp or the context \ctx is to blame for emitting this event, the latter indicates the secrecy level of values of the emitted event, i.e., either $\lock$ or $\unlock$.
-As for pre-events, the memory-related ones are allocation ($\ev{Alloc}\ \loc\ \valueexpr$), deallocation ($\ev{Dealloc}\ \loc$), reading from ($\ev{Get}\ \loc\ \valueexpr$) and writing to memory ($\ev{Set}\ \loc\ \valueexpr\ \valueexpr[']$).
-
-\begin{center}
-  \judgbox{\pstepto{\runtimetermvar}{\runtimetermvar[']}{\event}}{,,$\runtimetermvar$ does one primitive step to $\runtimetermvar[']$ emitting event $\event$.''}
-  %
-  \typerule{$e-\text{get}-\in$}{
-    \sandboxtag=\statevar.\sandboxtag &
-    \statevar.\store(x)=\tup{\loc}{\sandboxtag,\poison,m} \\
-    \loc+n\in\operatorname{dom}\statevar.\heap[^{\sandboxtag}](\loc+n)
-  }{
-    \pstepto{\rtt{\statevar}{\lget{x}{n}}}{\rtt{\statevar}{\heap[^{\sandboxtag}](\loc+n)}}{\tup{\ev{Get}\ \loc\ n}{\sandboxtag}}
-  }{e-get-in}
-  %
-  \typerule{$e-\text{set}-\notin$}{
-    \sandboxtag=\statevar.\sandboxtag &
-    \statevar.\store(x)=\tup{\loc}{\sandboxtag,\poison,m} \\
-    \loc+n\notin\operatorname{dom}\statevar.\heap[^{\sandboxtag}]
-  }{
-    \pstepto{\rtt{\statevar}{\lset{x}{n}{\valueexpr}}}{\rtt{\statevar}{\valueexpr}}{\tup{\ev{Set}\ \loc\ n\ \valueexpr}{\sandboxtag}}
-  }{e-set-notin}
-  %
-  \typerule{$e-\text{new}$}{
-    \fresh{\statevar}{z} &
-    \fresh{\statevar}{\loc} &
-    \heap[^{\sandboxtag}_1]=\growh{\statevar.\heap[^{\sandboxtag}]}{n} &
-    \store[_1]=\storeel{z}{\loc}{\statevar.\sandboxtag}{\poisonless}{n},\statevar.\store
-  }{
-    \pstepto{\rtt{\statevar}{\lnew{x}{n}{\expr}}}{\rtt{\statevar\subst{\heap[^{\sandboxtag}]}{\heap[^{\sandboxtag}_1]}\subst{\store}{\store[_1]}}{\expr\subst{z}{x}}}{\tup{\ev{Alloc}\ \loc\ n}{\sandboxtag}}
-  }{e-new}
-  %
-  \typerule{$e-\text{dealloc}$}{
-    \statevar.\store(x) = \tup{\loc}{\statevar.\sandboxtag,\poison,n} &
-    \store[_1]=\seth{\statevar.\store}{x}{\tup{\loc}{\statevar.\sandboxtag,\poisoned,n}}
-  }{
-    \pstepto{\rtt{\statevar}{\ldelete{x}}}{\rtt{\statevar\subst{\store}{\store[_1]}}{0}}{\tup{\ev{Dealloc}\ \loc}{\sandboxtag}}
-  }{e-dealloc}
-  %
-  \typerule{$e-\text{abort}$}{
-  }{
-    \pstepto{\rtt{\statevar}{\labort}}{\lightning}{\lightning}
-  }{e-abort}
-\end{center}
-
-\Cref{tr:e-dealloc,tr:e-get-in,tr:e-set-notin} demonstrate the use of the poison tag $\poison$ as metadata for pointers instead of removing them from the store $\Delta$.
-Take deallocation as an example, the premise does not care at all about the actual state of the poison tag $\poison$ and just overwrites it, marking the location as freed $\poisoned$.
-Besides that, the poison tag does not have any semantic meaning, at least in general.
-For language $\Ltms$, this tag is really just some semantic metadata that programmers have no access to.
-But, for the other languages, e.g., $\Ltrg$, the poison tag is used to check pointer validity.
-The \Cref{tr:e-new} allocates enough space on the respective heap, either $H^{\ctx}$ or $H^{\comp}$ depending on the execution context, i.e., the value of $\Omega.\sandboxtag$, and adds the appropriate metadata associated to the pointer in $\Delta$.
-Reading from and writing to memory have to cases: Either the heap is large enough or not and, depending on that, either the actual value stored at that location is read from or written to, or some garbage data is returned.
-However, note that the execution does not get stuck in such cases, performs a step, and emits an appropriate event.
-Also note that whether a pointer is poisoned or not is not reflected on the trace.
-
-\begin{center}
-  \begin{align*}
-    \mi{(Pre\text{-}Events)}~\event[_{b}]&\bnfdef \cdots \mid \ev{Call}\ \comm\ g\ \valueexpr \mid \ev{Ret}\ \comm\ \valueexpr \mid \ev{Start} \mid \ev{End}\ \valueexpr \\
-  \end{align*}
-\end{center}
-
-A key difference in comparison with the specification trace model (\Cref{subsec:propdefs}) is that, as standard in secure compilation work~\cite{patrignani2021rsc,abate2019jour,korashy2021capableptrs}, the traces have a call and return event that signals context switches, which are referred to as {\em interaction events}.
+Possible events that are emitted during execution are similar to the ones of the specification trace model (\Cref{subsec:propdefs}).
+The key difference is that, as standard in secure compilation work~\cite{patrignani2021rsc,abate2019jour}, the traces have a call and return event that signals context switches, which are referred to as {\em interaction events}.
 The reason for these interaction events is technical: They are a proof artifact for reconstructing a source context from a potentially malicious target context, where during that translation, the insertion of some wrapper code right before context switching may be necessary to make the proof succeed.
@@
 Contrary, $\ev{Call\ \comptoctx\ foo\ \valueexpr}$ and $\ev{Return\ \ctxtocomp\ \valueexpr}$ signal that program execution transitions from component- to context-level.
-For calls without this context switch, the environmental semantics attaches the $\nocomm$ tag.
+The environmental semantics emits the tag $\nocomm$ for calls from context- to context-level or component- to component-level.
 In the following, $\neg\ctx=\comp$ and $\neg\comp=\ctx$.
@@
 
-The environmental semantics is mostly straightforward.
-In \Cref{tr:e-ret-notsame,tr:e-call-notsame}, the judgement $\Omega.\commlib\vdash\text{foo}:\Omega.\sandboxtag$ checks whether $\text{foo}$ is a component-level name by looking it up in the list of component-level names $\commlib$ and emits the appropriate transfer tag, i.e., either $\comptoctx$ or $\ctxtocomp$.
+The environmental semantics is also mostly straightforward.
+In \Cref{tr:e-ret-notsame,tr:e-call-notsame} the judgement $\commlib\vdash\text{foo}:\Omega.\sandboxtag$ checks whether $\text{foo}$ is a component-level name by looking it up in the list of component-level names $\commlib$ and emits the appropriate transfer tag, i.e., either $\comptoctx$ or $\ctxtocomp$.
 Additional rules that are left out ensure that, e.g., when calling the $\text{main}$ function, the event $\ev{Start}$ is emitted, which is a design choice this paper does for convenience when reasoning about call-chains.
-Note that the $\ev{End}\ \valueexpr$ event is not emitted if the program crashes.
-
-\MKin{%
-  should I add a rule for top level exec or keep it like this?
-}
+
 The top-level execution $\progstepto{\prog{\library_{\ctx}}{\library_{\comp}}}{\runtimetermvar}{\trace}$ constructs an initial state $\Omega$ by linking $\library_{\ctx}$ and $\library_{\ctx}$ and then starts execution by calling the $\text{main}$ function.
@@
 $\src{L_{\tmssafe}}$ uses the same syntax as presented earlier (\Cref{sec:casestud:defs}) without extensions to the term level.
-But, $\src{L_{\tmssafe}}$ is statically typed, where the type system is inspired by $L^{3}$~\cite{morrisett2005L3,scherer2018fabulous}.
+But, $\src{L_{\tmssafe}}$ is statically typed, where the type system is inspired by $L^{3}$~\cite{morrisett2005L3}.\MP{fabulous paper from amal?relevant?}
+\MP{and therefore omitted?}
 The type system of $\src{L_{\tmssafe}}$ exhibits the property that every well-typed $\src{L_{\tmssafe}}$ program satisfies \gls{tms} (\Cref{thm:wt:tms}).
 The proof of this theorem relies on a projection $\msfilterLtms{\src{\event}}=\event$ from $\src{L_{\tmssafe}}$ events to specification events $\event$, because the properties defined earlier (\Cref{subsec:propdefs}) are defined in the specification trace model.
-Hereby, the map $\thelocmap(\src{\loc})=\loc$ maps an $\Ltms$ location to a location $\loc$ of the specification trace model.
-
-\begin{center}
-  \judgbox{\msfilterLtms{\src{\event}}=\event}{,,Project an $\Ltms$ event $\src{\event}$ to a specification event $\event$.''}
-  \typerule{$\Ltms$-filter-context}{
-    \src{\event[_b]}\not=\src{\lightning}
-  }{
-    \msfilterLtms{\src{\tup{\event[_b]}{\ctx,\securitytag}} = \emptyevent}
-  }{ltms-filter-context}
-  %
-  \typerule{$\Ltms$-filter-abort}{
-  }{
-    \msfilterLtms{\src{\lightning}} = \lightning
-  }{ltms-filter-abort}
-  %
-  \typerule{$\Ltms$-filter-start}{
-  }{
-    \msfilterLtms{\src{\tup{Start}{\comp}}} = \emptyevent
-  }{ltms-filter-start}
-  %
-  \typerule{$\Ltms$-filter-alloc}{
-    \locmapsto{\src{\loc}}{\loc} &
-    \src{n}=n
-  }{
-    \msfilterLtms{\src{\tup{Alloc\ \loc\ n}{\comp}}} = \tup{\ev{Alloc}\ \loc\ n}{\comp,\unlock}
-  }{ltms-filter-alloc}
-\end{center}
-
-Most rules of the projection $\msfilterLtms{\src{\event}}$ are left out since, for the most part, it does the expected, e.g., $\msfilterLtms{\src{\ev{Dealloc}\ \loc;\comp;\securitytag}}=\ev{Dealloc}\ \delta(\loc);\comp;\securitytag$.
+Hereby, $\delta$ maps $\src{L_{\tmssafe}}$ locations $\src{\loc}$ to locations $\loc$ of the specification trace model whenever an event with a location is projected.
+The projection $\msfilterLtms{\src{\event}}$ for the most part does the expected, e.g., $\msfilterLtms{\src{\ev{Alloc}\ \loc\ n;\comp;\securitytag}}=\ev{Alloc}\ \delta(\loc)\ n;\comp;\securitytag$.
 But, it also filters any action that a context does as well as the interaction events, since these are irrelevant for component-level \gls{tms}.
+\MPin{ok let's add details}
 
@@
 
-\begin{center}
-  \begin{align*}
-    \mi{(Expressions)}~\trg{\expr}&\bnfdef \cdots \mid \trg{\lispoisoned{x}} \mid \trg{\lhast{\expr}{\type}} \\
-  \end{align*}
-\end{center}
-
-$\Ltrg$ extends the syntax presented earlier (\Cref{subsec:cs:defs}) with dynamic typechecks $\trg{\lhast{\expr}{\type}}$ and a way to inspect poison tags $\trg{\lispoisoned{x}}$ in the metadata of pointers.
-For valid pointers ($\trg{\poisonless}$) bound to variable $\trg{x}$, the check $\trg{\lispoisoned{x}}$ yields $\trg{1}$.
+$\Ltrg$ extends the syntax presented earlier (\Cref{subsec:cs:defs}) with dynamic typechecks and a way to inspect poison tags in the metadata of pointers.
+For allocated ($\trg{\poisonless}$) arrays bound to variable $\trg{x}$, the check $\trg{\lispoisoned{x}}$ yields $\trg{1}$.
 If the array bound to $\trg{x}$ was allocated, i.e., has been poisoned ($\trg{\poisoned}$), the check $\trg{\lispoisoned{x}}$ evaluates to $\trg{0}$.
-
-\begin{center}
-  \typerule{$e-\trg{\lhast{x}{\natt}}$}{
-  }{
-    \pstepto{\rtt{\trg{\statevar}}{\trg{\lhast{x}{\natt}}}}{\rtt{\trg{\statevar}}{\trg{1}}}{\emptyevent}
-  }{e-has-nat-no}
-  %
-  \typerule{$e-\trg{\lhast{n}{\natt}}$}{
-  }{
-    \pstepto{\rtt{\trg{\statevar}}{\trg{\lhast{n}{\natt}}}}{\rtt{\trg{\statevar}}{\trg{0}}}{\emptyevent}
-  }{e-has-nat-yes}
-\end{center}
-
-Typechecks $\lhast{\expr}{\type}$ match on $\trg{\expr}$ and evaluate to $\trg{0}$ if the term is of type $\trg{\type}$ and $\trg{1}$ otherwise.
-
 %
 The projection $\msfilterL{\trg{\trace}}$ is entirely similar to $\msfilterLtms{\src{\trace}}$.
+\MP{ok, details}
 
 \subsection{$\Lms$: Another Memory-Unsafe Language}\label{subsec:lms}
-To enhance readablity, this paper uses $\Lms$, despite it being exactly equal to $\Ltrg$ (\Cref{subsec:lsms}).
+To enhance readablity, this paper uses $\Lms$, despite it being exactly equal to $\Ltrg$.
 The projection $\msfilterLms{\irl{\trace}}$ is also exactly equal to $\msfilterL{\trg{\trace}}$.
 
-\subsection{$\Lscct$: A Memory-Unsafe Language with a Data Independent Timing Mode}\label{subsec:lscct}
-
-\begin{center}
-  \begin{align*}
-    \mi{(Expressions)}~\obj{\expr}&\bnfdef \obj{n^{\securitytag}} \mid \cdots \mid \obj{\llet{x^{\securitytag}}{\expr[_{1}]}{\expr[_{2}]}} \mid \cdots \mid \obj{\lwrdoit{\expr}} \mid \obj{\lrddoit{x}{\expr}}
-  \end{align*} \\
-  \begin{align*}
-    \mi{(States)}~\obj{\statevar}&\bnfdef \obj{\tup{\cfstate}{\sandboxtag,n,\memstate}}
-  \end{align*}
-\end{center}
-
-$\Lscct$ extends $\Lms$ (\Cref{subsec:lms}) with a way to write to a model specific register that controls a data (operand) independent timing mode, a feature that is present in both Arm~\cite[p.~543]{arm-refman} and Intel~\cite[p.~80]{intel-refman} processors.
-To this end, states are extended with the value of the register, which is initially set to be not active.
-If the register is marked active, the intuition is that no secrets can appear on specification traces.
-If the register is marked inactive, secrets may appear on traces.
-For the other languages seen earlier, the mode is intuitively always-on, i.e., the mode of execution always uses data independent timing.
-The language also adds user-annotations to values and variables to know their secrecy $\obj{\securitytag}$, which is either high $\obj{\lock}$ or low $\obj{\unlock}$.
-Security tags $\obj{\securitytag}$ are on the usual secrecy lattice, where $\obj{\lock}\le\obj{\securitytag}$ and $\obj{\securitytag}\le\obj{\unlock}$.
-
-\begin{center}
-  \begin{align*}
-    \mi{(Pre\text{-}Events)}~\obj{\event[_{b}]}&\bnfdef \cdots \mid \obj{\widehat{Get}\ \loc\ \valueexpr} \mid \obj{\widehat{Set}\ \loc\ \valueexpr\ \valueexpr[']} \mid \obj{Branch\ n} \mid \obj{Binop\ n}
-  \end{align*}
-\end{center}
-
-To prevent secrets from leaking but still enable reasoning about memory safety, $\Lscct$ extends pre-events with $\obj{\widehat{Get}\ \loc\ \valueexpr}$ and $\obj{\widehat{Set}\ \loc\ \valueexpr}$.
-These indicate reads from and writes to memory without leaking secret information involved in the access.
-Moreover, the language extends pre-events with $\obj{\ev{Branch}\ n}$ and $\obj{\ev{Binop}\ n}$ that are emitted when evaluating a branch or certain binary expressions, such as division, respectively, whenever the data independent timing mode is inactive.
-
-\begin{center}
-  \typerule{$e-\obj{\oplus}-\text{noleak}$}{
-    \obj{m}\not=\obj{0} &
-    \obj{n_3}=\obj{n_1}\oplus\obj{n_2} \\
-    \obj{\securitytag['']}\le\obj{\securitytag} &
-    \obj{\securitytag['']}\le\obj{\securitytag[']}
-  }{
-    \pstepto{\rtt{\obj{\cfstate;\sandboxtag;m;\memstate}}{\obj{\lbinop{n_1^{\securitytag}}{n_2^{\securitytag[']}}}}}{\rtt{\obj{\cfstate;\sandboxtag;m;\memstate}}{\obj{n_3^{\securitytag['']}}}}{\obj{\emptyevent}}
-  }{e-binop-noleak}
-  %
-  \typerule{$e-\obj{wrdoit}$}{
-  }{
-    \pstepto{\rtt{\obj{\cfstate;\sandboxtag;m;\memstate}}{\obj{\lwrdoit{n^{\securitytag}}}}}{\rtt{\obj{\cfstate;\sandboxtag;n;\memstate}}{\obj{n^{\securitytag}}}}{\obj{\emptyevent}}
-  }{e-wrdoit}
-  %
-  \typerule{$e-\obj{ifz}-\obj{true}-\text{leak}$}{
-  }{
-    \pstepto{\rtt{\obj{\cfstate;\sandboxtag;0;\memstate}}{\obj{\lifz{0^{\securitytag}}{\expr[_1]}{\expr[_2]}}}}{\rtt{\obj{\cfstate;\sandboxtag;0;\memstate}}{\obj{\expr[_2]}}}{\obj{\ev{Branch}\ 0;\sandboxtag;\securitytag}}
-  }{e-ifz-true-leak}
-\end{center}
-
-The evaluation steps are amended to propagate the securitytags $\obj{\securitytag}$.
-When the data independent timing mode is active, pre-events $\obj{\ev{Branch}\ n}$ and $\obj{\ev{Binop}\ n}$ are emitted for conditionals and binary operations, respectively.
-To mark this mode as active or inactive, \Cref{tr:e-wrdoit} allows to access the register in $\obj{\statevar}$.
-
-\begin{center}
-  \judgbox{\scctfilterLscct{\obj{\event}}=\event}{,,Project an $\Lscct$ event $\obj{\event}$ to a specification event $\event$.''}
-  \typerule{$\Lscct$-filter-context}{
-    \obj{\event[_b]}\not=\obj{\lightning}
-  }{
-    \scctfilterLscct{\obj{\tup{\event[_b]}{\ctx,\securitytag}} = \emptyevent}
-  }{lscct-filter-context}
-  %
-  \typerule{$\Lscct$-filter-$\widehat{\text{get}}$}{
-    \locmapsto{\obj{\loc}}{\loc} &
-    \obj{n}=n
-  }{
-    \scctfilterLscct{\obj{\tup{\widehat{Get}\ \loc\ n}{\comp,\securitytag}}} = \tup{\ev{Get}\ \loc\ n}{\comp,\unlock}
-  }{lscct-filter-get-hat}
-  %
-  \typerule{$\Lscct$-filter-$\text{get}$}{
-    \locmapsto{\obj{\loc}}{\loc} &
-    \obj{n}=n
-  }{
-    \scctfilterLscct{\obj{\tup{Get\ \loc\ n}{\comp,\securitytag}}} = \tup{\ev{Get}\ \loc\ n}{\comp,\securitytag}
-  }{lscct-filter-get}
-\end{center}
-
-The projection to the specification trace model is mostly straightforward and similar to the others, e.g., \Cref{subsec:cs:tms}.
-However, for events containing the pre-events $\obj{\widehat{Get}}$ and $\obj{\widehat{Set}}$, the projection always translates the security-tag $\obj{\sigma}$ to $\unlock$, regardless of its actual value, as seen in \Cref{tr:lscct-filter-get-hat}.
+\subsection{$\Lscct$: A Language with a Data Independent Timing Mode}\label{subsec:lscct}
+
+While the previous languages (\Cref{subsec:ltms,subsec:lsms,subsec:lms}) are not concerned with \gls{cct}, $\Lscct$ extends $\Lms$ (\Cref{subsec:lms}) with a register flag for a data (operand) independent timing mode, a feature that is present in both Arm~\cite[p.~543]{arm-refman} and Intel~\cite[p.~80]{intel-refman} processors.
+For the other languages seen earlier, this mode is intuitively always-on.
+\MP{spell out the formal changes in an align* environmtn} % this necessitates that we present $\Omega$ also in the shared lang defs
+To modify the mode, $\Lscct$ extends expressions with the term $\obj{\lwrdoit{\expr}}$ that writes the value of $\obj{\expr}$ into the register and with a term $\obj{\lrddoit{x}{\expr}}$ that reads the value of the register into $\obj{x}$.
+If the register is turned on, no secrets can appear on traces.
+If the register is turned off, secrets may appear on traces as well as new events $\obj{Branch\ n}$ and $\obj{Binop\ n}$, which the semantics emits for conditionals and certain binary operations, e.g., division.
+To prevent leakage of secrets but still enable reasoning about memory safety, pre-events are extended with $\obj{\widehat{{Get}}\ \loc\ \valueexpr}$ and $\obj{\widehat{Set}\ \loc\ \valueexpr}$, which indicate reads from and writes to memory without leaking secret information involved in the access, which is taken care of by the projection to specification events.
+\Cref{ex:lscct} illustrates these changes.
+\MPin{formalisation additions in algin* env}
+
+The projection to the specification trace model is mostly straightforward.
+But for events containing the pre-events $\obj{\widehat{Get}}$ and $\obj{\widehat{Set}}$, the projection always translates the security-tag $\obj{\sigma}$ to $\unlock$, regardless of its actual value.
 The pre-events themselves still translate to just $\ev{Get}$ and $\ev{Set}$, respectively.
-With this technical setup, the information whether a read or write happened on a secret value is not hidden by the semantics, e.g., by emitting $\obj{\emptyevent}$, but when projecting to specification events.
+With this technical setup, the information wether\MP{spelling} a read or write happened on a secret value is not hidden by the semantics, e.g., by emitting $\obj{\emptyevent}$, but when projecting to specification events.
 This allows flexibility: The trace can be checked to satisfy different properties, such as, in this case, \gls{tms}, \gls{sms}, \gls{scct}, and their combined versions.
-\Cref{ex:lscct} illustrates the differences of $\Lscct$ compared to the other languages.
 
@@
 
-This paper does a first step towards practical secure compilation chains by introducing a theoretical framework demonstrating that secure compilers compose.
-The paper exercises this framework on a case-study that consists of an optimising compilation chain that is secure with respect to \gls{msscct}.
-\MP{
-	not an appropriate recap.
-	rewrite summarising the paper
-}
-
-In future work, it would be interesting to investigate whether it is possible to provide {\em secure compiler combinators}, similarly to how parser combinators work.
-%To this end, it may be necessary to extend existing frameworks for multi-language semantics.
-It is also interesting to extend our case-study and verify all results of it in Coq.
-This way, it is possible to obtain an executable, secure compiler.
-However, the formalisation effort for backtranslation proofs is known to be enormous\MP{cit + details}, so another research avenue may be to find better ways to (semi-)automate standard secure compilation proofs.
-\MPin{
-	not particularly informative FW. i'd rather not have it as such
-}
+% This paper does a first step towards practical secure compilation chains by introducing a theoretical framework demonstrating that secure compilers compose.
+% The paper exercises this framework on a case-study that consists of an optimising compilation chain that is secure with respect to \gls{msscct}.
+% \MP{
+% 	not an appropriate recap.
+% 	rewrite summarising the paper
+% }
+
+% In future work, it would be interesting to investigate whether it is possible to provide {\em secure compiler combinators}, similarly to how parser combinators work.
+% %To this end, it may be necessary to extend existing frameworks for multi-language semantics.
+% It is also interesting to extend our case-study and verify all results of it in Coq.
+% This way, it is possible to obtain an executable, secure compiler.
+% However, the formalisation effort for backtranslation proofs is known to be enormous\MP{cit + details}, so another research avenue may be to find better ways to (semi-)automate standard secure compilation proofs.
+% \MPin{
+% 	not particularly informative FW. i'd rather not have it as such
+% }
 

### 2023-07-11 popl/main.tex
pass

@@
 \begin{abstract}
+\MPpost{
+	this abstract fails in setting up the robustness argument.
+	it'll be a good exercise for you to take inspiration from this, but then write something as concise but with a more direct focus on robust properties
+}
 % context
@@
 % Without a framework for composing secure compiler passes, it is not possible to enable separation of concerns, e.g., to have a secure compilation pass that ensures \gls{ms} that is developed independently of another secure pass for \gls{cct} as suggested before and as showcased in \Cref{ex:strncpy}.
+\MPpost{
+	this intro does a poor job at introducing the robust compilation argument.
+	perhaps it'd be best to start from robust properties, which motivates the use of robust compilers?
+	idk
+	as for the abstract, this is a good starting point for you to make a better intro
+}
 
@@
 % If \texttt{n} is at least as large as the length of the array \texttt{src}, 
-the memory location past the last element will leak.
+the memory location past the last element will be leaked to an attacker.
 \begin{lstlisting}[language=c,basicstyle=\small\ttfamily,morekeywords={size_t}]
@@
 
-Unfortunately, fixing \gls{sms} is not enough to make \texttt{strncpy} secure, because of timing attacks.
+Because of timing attacks, fixing \gls{sms} is not enough to make \texttt{strncpy} secure.
 In fact, the loop can terminate early, as soon as the string-terminating character \texttt{'\textbackslash 0'} is encountered, making program execution time proportional to the length of the array pointed by \texttt{src}.
@@
 Robust compilers~\cite{abate2019jour} are a form of secure compilers that preserve security properties even in the presence of arbitrary attackers interacting with compiled code.
-Thus, robust compilers can be used to prevent the vulnerability resulting from uninitialised memory, e.g., by targeting capability-based languages such as CHERI~\cite{woodruff2014CHERI}, Arm Morello~\cite{arm-morello}, or MSWasm~\cite{michael2023mswasm}, where pointers are always initialised.
+Thus, robust compilers can be used to prevent the vulnerability resulting from uninitialised memory, e.g., by targeting capability-based languages such as CHERI~\cite{woodruff2014CHERI}, Arm Morello~\cite{arm-morello}, or MSWasm~\cite{michael2023mswasm}, where the compiler relies on capabilities to checks that pointers are always initialised.
 \end{example}
@@
 % Different kinds of compilation passes exist, e.g., there are usually several compilation passes that perform optimisations~\cite{lattner2004llvm,googlev8,androidstudio,kuepper2023cryptopt,manjikian1997fusion,wegman1991ccp} or source-code instrumentations~\cite{nagarakatte2009soft,nagarakatte2010cets,akritidis2009baggy,dhumbumroong2020boundwarden,jung2021pico,nam2019framer,shankaranarayana2023tailcheck,younan2010paricheck,zhou2023fatptrs,bond2017vale,almeida2017jasmin,kuepper2023cryptopt,cauligi2019fact} that enforce that the compiled program fulfils a property of interest.
-Worse, without a framework for composing secure compiler passes, it is not possible to enable separation of concerns, e.g., to have a secure compilation pass that ensures \gls{ms} that is developed independently of another secure pass for \gls{cct}, that is developed independently of optimisation passes.
+Worse, without a framework for composing secure compiler passes, it is not possible to enable separation of concerns, e.g., to have a secure compilation pass that ensures \gls{ms} that is developed independently of another secure pass for \gls{cct}, that is developed independently of other passes such as optimisation ones.
 
@@
 % Crucially, this end-to-end security preservation is obtained by composing the security property preserved by each individual pass, as dictated by the framework of this paper.
-This result showcases how the framework allows the kind of formal security reasoning that compiler writers already want (and do), obtaining precise, compositional security reasoning while providing minimal (and modular) proof effort.
+This result showcases how the framework allows the kind of formal security reasoning that compiler writers already want (and already do), obtaining precise, compositional security reasoning while providing minimal (and modular) proof effort.
 
@@
         %\Cref{sec:compprop} presents a formal framework to reason about compositions of security properties and demonstrate its usage with concrete examples:
-        This paper formalises security properties (\Cref{sec:compprop}) that are of interest for real-world compiler writers (as identified by the plethora of work enforcing such properties individually~\cite{akritidis2009baggy,nagarakatte2009soft,nagarakatte2010cets,dhumbumroong2020boundwarden,jung2021pico,nam2019framer,shankaranarayana2023tailcheck,younan2010paricheck,zhou2023fatptrs,bond2017vale,almeida2017jasmin,kuepper2023cryptopt,cauligi2019fact}), namely \gls{tms}, \gls{sms} and \gls{cct}.
+        This paper formalises security properties (\Cref{sec:compprop}) that are of interest for real-world compiler writers, namely \gls{tms}, \gls{sms} and \gls{cct} (as identified by the plethora of work enforcing such properties individually~\cite{akritidis2009baggy,nagarakatte2009soft,nagarakatte2010cets,dhumbumroong2020boundwarden,jung2021pico,nam2019framer,shankaranarayana2023tailcheck,younan2010paricheck,zhou2023fatptrs,bond2017vale,almeida2017jasmin,kuepper2023cryptopt,cauligi2019fact}).
         % 
@@
         %Concretely, it investigates upper and lower compositions, as well as the sequential composition.
-        For this, this paper studies sequential compiler composition as well as compilers with multiple inputs or multiple outputs, as typically done in existing compilation chains.
+        For this, this paper studies sequential compiler composition as well as compilers with multiple input languages or multiple output ones, as used in existing compilation chains.
         %
@@
 To introduce the security argument of this paper, this section first presents the concepts of (security) properties, of their satisfaction, and of their robust satisfaction (i.e., satisfaction in the presence of an active attacker; \Cref{subsec:bg:tprop}).
-Then, borrowing from existing work~\cite{abate2019jour,abate2021extacc,patrignani2021rsc}, the section introduces secure compilers as compilers that preserve robust property satisfaction (\Cref{subsec:bg:rtp}).
+Then, borrowing from existing work~\cite{abate2019jour,abate2021extacc}, the section introduces secure compilers as compilers that preserve robust property satisfaction (\Cref{subsec:bg:rtp}).
 
@@
 This paper employs the security model where programs are written in a language whose semantics emits events $\event$.
-Events include security-relevant actions (e.g., reading from and writing to memory) and the unobservable event $\emptyevent$.
+Events include security-relevant actions (e.g., reading from and writing to memory, as detailed in \Cref{sec:compprop}) and the unobservable event $\emptyevent$.
 As programs execute, their emitted events are concatenated in traces $\trace$, which serve as the description of the behaviour of a program.%
@@
 
-Properties $\pi$ are sets of traces of admissible program behaviours, ascribing what said property considers valid~\cite{clarkson2008hyper}.
+Properties $\pi$ are sets of traces of admissible program behaviours, ascribing what said property considers valid.
 The set of all properties can be partitioned into different {\em classes} ($\class$), i.e., safety, liveness, and neither safety nor liveness~\cite{clarkson2008hyper}.
@@
 Thus, from a security perspective, this considers only a passive attacker model, where the attacker observes the execution and, e.g., retrieves secrets from that.
-To consider a stronger model similarly to what existing work does~\cite{abate2019jour,abate2021extacc,patrignani2021rsc,maffeis2008code-carrying,gordon2003authenticity,fournet2007authorization,bengtson2011refine,backes2014uniontyps}, the concept of satisfaction can be extended with {\em robustness}.
+To consider a stronger model similarly to what existing work does~\cite{abate2019jour,abate2021extacc,maffeis2008code-carrying,gordon2003authenticity,fournet2007authorization,bengtson2011refine,backes2014uniontyps}, the concept of satisfaction can be extended with {\em robustness}.
+\MPpost{
+	citations here should include mswasm, swasey's stuff and michel's and move
+}
 Robust satisfaction considers partial programs $\progvar$, i.e., components with missing imports, which cannot run until said imports are fulfilled.
@@
   The vulnerability arises due to a race condition where the context-level function \texttt{hci\_cmd\_work} was not expected to behave maliciously, since it resides in the same source-code repository where the vulnerability occurs.
-  Nevertheless, the component-level code of \texttt{hci\_dev\_do\_open} is linked with \texttt{hci\_cmd\_work} and does not atomically check whether the pointer has been freed already:
+  Nevertheless, the component-level code of \texttt{hci\_dev\_do\_open} is linked with \texttt{hci\_cmd\_work} and does not atomically check whether the pointer\MP{what pointer?} has been freed already:
   Therefore, \texttt{hci\_dev\_do\_open} does not satisfy the no-double-frees property robustly, since there is an implementation for \texttt{hci\_cmd\_work} that results the whole program to violate that property.
@@
   The pointer is valid for one $\trg{char}$-sized memory cell, as expected, but the actual read operation now returns $\trg{sizeof(int)}$ many bytes instead of just $\trg{sizeof(char)}$ many.
-  Thus, even if $\src{foo}$ may have been robust with respect to the \gls{sms}, its compiled counterpart is not and therefore the compiler fails to attain \Thmref{def:rtp}.
+  Thus, even if $\src{foo}$ may have been robust with respect to the \gls{sms}, its compiled counterpart is not and therefore the compiler fails to attain \Cref{def:rtp}.
 \end{example}
@@
 Since reasoning on monitors is easier than directly on just traces, this section presents a monitor for each of the previously shown safety properties (\Cref{subsec:propdefs}).
+\MPpost{
+	this is fine, but it'd be nice to provide more insights on why it's easier
+}
 To lessen the burden when proving that a monitor accepts the trace of a program execution, each monitor uses a custom trace model that contains only the relevant information related to the property the monitor checks.
@@
   However, having returned from \texttt{strncpy}, the \texttt{greet} function continues and deallocates $\loc[_{y}]$ whose subsequent use in the \texttt{printf} call is a use-after-free bug.
+
   The fix would be to \lstinline[language=c++,basicstyle=\small\ttfamily]|delete greetings| instead of \texttt{to} and add a \lstinline[language=c++,basicstyle=\small\ttfamily]|delete to| after the \texttt{printf} call, which leads to the abstract monitor trace $\abstrace[']=\absev{Alloc}\ \loc[_{x}]\ \comp\cdot\absev{Alloc}\ \loc[_{y}]\ \comp\cdot\absev{Use}\ \loc[_{x}]\ \comp\cdot\absev{Use}\ \loc[_{x}]\ \comp\cdot\absev{Use}\ \loc[_{y}]\ \comp\cdot\ldots\cdot\absev{Use}\ \loc[_{x}]\ \comp\cdot\absev{Dealloc}\ \loc[_{x}]\ \comp\cdot\absev{Use}\ \loc[_{y}]\ \comp$.
@@
 \end{example}
+\MPpost{
+	in general, i think embedding the traces in the text does not make them readable.
+	in the exorcising paper i think i put them in align env, or in a way to make them stand out.
+	i don't think we have the time to do this change uniformly now,
+	but at some point we should
+}
 
@@
   Consider an excerpt from a component that reads from non-confidential memory \texttt{x} and writes the result to confidential memory \texttt{y}.
+  \MPin{
+  can't we call strncpy but with src == low and dest == high?
+  }
   The trace of just this part may look like $\ev{Use}\ \loc_{x}\ 1729;\comp;\unlock\cdot\ev{Use}\ \loc_{y}\ 1729;\comp;\lock$.
@@
 Since the composition of secure compilers is again a secure compiler, the theorem generalises to a whole chain of $n$ secure compilers.
+\MPin{
+	what do the sentences below mean in this paragraph?
+}
 It is worth pointing out that \gls{ms} (\Cref{subsubsec:ms}) is an example of a practically relevant intersection of classes, namely, \gls{tms} and \gls{sms} (\Cref{subsubsec:tms,subsubsec:sms}).
@@
 
-Consider two compilers both accepting $\src{LLVM IR}$~\cite{lattner2004llvm} and one of them emits $\trg{x86\_64}$, while the other emits $\trg{ARMv8}$.
+Consider two compilers both accepting $\src{LLVM IR}$~\cite{lattner2004llvm} and one of them emits $\trg{x86\_64}$, while the other emits $\obj{ARMv8}$.
 It is intuitive that they are in some sense composed in the LLVM framework, but the decision of when to use one over the other is inherently {\em internal} to the formalisation effort of this kind of composition.
 For example, the user of this compiler provides an explicit flag that instructs to emit $\trg{x86\_64}$ or the framework itself detects the target platform via heuristics, such as supported instructions.
-The following theorem demonstrates what happens if the involved compilres are secure:
+
+The following theorem demonstrates what happens if the involved compilers are secure:
 Given \bul{$\cc{\src{L}}{\trg{L}}$ robustly preserves $\class[_{1}]$} and \rul{$\cc{\src{L}}{\obj{L}}$ robustly preserves $\class[_{2}]$}, it follows that \oul{their lower composition $\cc{\src{L}}{\trg{L}+\obj{L}}$ robustly preserves the intersection of classes $\class[_{1}]$ and $\class[_{2}]$}.
@@
 
-This paper considers \gls{cct} as an architecture-dependent property writing \gls{cct} should not be one of the programmer's concerns~\cite{cauligi2019fact}.
+This paper considers \gls{cct} as an architecture-dependent property\MPin{can we justify this}, so writing \gls{cct} should not be one of the programmer's concerns~\cite{cauligi2019fact}.
 Such consideration is backed up by architecture providing a data (operand) independent timing mode, such as processors by Arm~\cite[p.~543]{arm-refman} and Intel~\cite[p.~80]{intel-refman}.
@@
 
-States $\statevar$ are tuples\footnote{Throughout the paper, the subsitution notation is also used to update entries in states $\statevar$.} containing a control flow state $\cfstate$, control tags $\sandboxtag$, and a memory state $\memstate$.
-A control flow state $\cfstate$ entails a library, which provides definitions for functions when calling, as well as a stack of continuations $\ectxstack$.
+States $\statevar$ are tuples\footnote{Throughout the paper, the substitution notation is also used to update entries in states $\statevar$.} containing a control flow state $\cfstate$, control tags $\sandboxtag$, and a memory state $\memstate$.
+A control flow state $\cfstate$ entails a library, which provides definitions for functions calls, as well as a stack of continuations $\ectxstack$.
 Elements of the stack $\ectxstack$ are pairs of evaluation contexts and the name of a function associated to that evaluation context.
 Memory states $\memstate$ are tuples of two separate heaps $H$ and a store $\Delta$, which contains pointer metadata, such as the concrete memory location $\loc$, a control tag $\sandboxtag$ indicating which heap the pointer points into, a poison tag $\poison$, as well as bounds information $n$.
-The bounds information is a mere proof artifact that has no semantic significance.
-For details see the accompanying tech report~\MK{cit}.
-Memory states $\memstate$ have two separate heaps, essentially modelling a sandbox, to prevent contexts from performing pointer arithmetic and reading from or writing to the data owned by a component.
-While this prevents the effects of out of bounds accesses across the context and component boundary, the goal of the design of the languages of this paper is to be able to express security violations.
-To this end, the poison tag $\poison$ indicates whether a pointer has been freed $\poisoned$ or is still allocated $\poisonless$, so that uses of pointers can happen even after their deallocation without the semantics getting stuck.
+The bounds information is a mere proof artefact that has no semantic significance.
+The two separate heaps essentially model a sandbox, to prevent contexts from performing pointer arithmetic and reading from or writing to the data owned by a component.
+While this prevents the effects of out-of-bounds accesses across the context and component boundary, the goal of the design of the languages of this paper is to be able to express security violations.
+To this end, the poison tag $\poison$ indicates whether a pointer has been freed $\poisoned$ or is still allocated $\poisonless$, so that pointers can be used even after their deallocation without the semantics getting stuck.
 Runtime terms are simply expressions $\expr$ paired with the operational state $\statevar$.
@@
 
-All languages use the same trace model.
-In this trace model, events are either the empty event \emptyevent, the program crash event $\lightning$, or a tuple consisting of a control tag $\sandboxtag$ and a security tag $\securitytag$.
+All languages use the same trace model, where events are either the empty event \emptyevent, the program crash event $\lightning$, or a tuple consisting of a control tag $\sandboxtag$ and a security tag $\securitytag$.
 The former indicates whether the component \comp or the context \ctx is to blame for emitting this event, the latter indicates the secrecy level of values of the emitted event, i.e., either $\lock$ or $\unlock$.
 As for pre-events, the memory-related ones are allocation ($\ev{Alloc}\ \loc\ \valueexpr$), deallocation ($\ev{Dealloc}\ \loc$), reading from ($\ev{Get}\ \loc\ \valueexpr$) and writing to memory ($\ev{Set}\ \loc\ \valueexpr\ \valueexpr[']$).
+
+\MPpost{
+	this is a bit abrupt
+}
 
@@
 
-\Cref{tr:e-dealloc,tr:e-get-in,tr:e-set-notin} demonstrate the use of the poison tag $\poison$ as metadata for pointers instead of removing them from the store $\Delta$.
-Take deallocation as an example, the premise does not care at all about the actual state of the poison tag $\poison$ and just overwrites it, marking the location as freed $\poisoned$.
-Besides that, the poison tag does not have any semantic meaning, at least in general.
+To demonstrate the use of the poison tag $\poison$ as metadata for pointers instead of removing them from the store $\Delta$, consider \Cref{tr:e-dealloc,tr:e-get-in,tr:e-set-notin} .
+In \Cref{tr:e-dealloc}, the premise does not care at all about the actual state of the poison tag $\poison$ and just overwrites it, marking the location as freed $\poisoned$.
+Besides that, the poison tag does not have any semantic meaning.
 For language $\Ltms$, this tag is really just some semantic metadata that programmers have no access to.
 But, for the other languages, e.g., $\Ltrg$, the poison tag is used to check pointer validity.
-The \Cref{tr:e-new} allocates enough space on the respective heap, either $H^{\ctx}$ or $H^{\comp}$ depending on the execution context, i.e., the value of $\Omega.\sandboxtag$, and adds the appropriate metadata associated to the pointer in $\Delta$.
-Reading from and writing to memory have to cases: Either the heap is large enough or not and, depending on that, either the actual value stored at that location is read from or written to, or some garbage data is returned.
-However, note that the execution does not get stuck in such cases, performs a step, and emits an appropriate event.
+\Cref{tr:e-new} allocates enough space on the respective heap, either $H^{\ctx}$ or $H^{\comp}$ depending on the execution context, i.e., the value of $\Omega.\sandboxtag$, and adds the appropriate metadata associated to the pointer in $\Delta$.
+Reading from \Cref{tr:e-get-in} and writing to memory \Cref{tr:e-set-notin} have two cases: Either the heap is large enough or not and, depending on that, either the actual value stored at that location is read from or written to, or some garbage data is returned.
+However, note that the execution does not get stuck in such cases, it performs a step, and emits an appropriate event.
 Also note that whether a pointer is poisoned or not is not reflected on the trace.
@@
 For calls without this context switch, the environmental semantics attaches the $\nocomm$ tag.
-In the following, $\neg\ctx=\comp$ and $\neg\comp=\ctx$.
+In the following, $\neg\ctx=\comp$ and $\neg\comp=\ctx$.\MP{i don't think this is used?}
 
@@
 
-\MKin{%
-  should I add a rule for top level exec or keep it like this?
-}
+% \MKin{%
+%   should I add a rule for top level exec or keep it like this?
+% 	MP: like this is fine
+% }
 The top-level execution $\progstepto{\prog{\library_{\ctx}}{\library_{\comp}}}{\runtimetermvar}{\trace}$ constructs an initial state $\Omega$ by linking $\library_{\ctx}$ and $\library_{\ctx}$ and then starts execution by calling the $\text{main}$ function.
@@
 
-Typechecks $\lhast{\expr}{\type}$ match on $\trg{\expr}$ and evaluate to $\trg{0}$ if the term is of type $\trg{\type}$ and $\trg{1}$ otherwise.
+Dynamic typechecks $\trg{\lhast{\expr}{\type}}$ match on $\trg{\expr}$ and evaluate to $\trg{0}$ if the term is of type $\trg{\type}$ and $\trg{1}$ otherwise.
 
@@
 The projection $\msfilterL{\trg{\trace}}$ is entirely similar to $\msfilterLtms{\src{\trace}}$.
+\MPin{
+	entirely similar? no. equal
+}
 
@@
 
-\begin{center}
-  \begin{align*}
+\begin{gather*}
+  \begin{aligned}
     \mi{(Expressions)}~\obj{\expr}&\bnfdef \obj{n^{\securitytag}} \mid \cdots \mid \obj{\llet{x^{\securitytag}}{\expr[_{1}]}{\expr[_{2}]}} \mid \cdots \mid \obj{\lwrdoit{\expr}} \mid \obj{\lrddoit{x}{\expr}}
-  \end{align*} \\
-  \begin{align*}
+  \end{aligned} 
+  \\
+  \begin{aligned}
     \mi{(States)}~\obj{\statevar}&\bnfdef \obj{\tup{\cfstate}{\sandboxtag,n,\memstate}}
-  \end{align*}
-\end{center}
+  \end{aligned}
+\end{gather*}
 
@@
 Moreover, the language extends pre-events with $\obj{\ev{Branch}\ n}$ and $\obj{\ev{Binop}\ n}$ that are emitted when evaluating a branch or certain binary expressions, such as division, respectively, whenever the data independent timing mode is inactive.
+\MPin{rules are not explained}
 
@@
   The left side of the table, i.e., the two columns on the left, describe the execution trace of the program, while the right side of the table, i.e., the two columns on the right, describe the respective projections $\scctfilterLscct{\obj{\trace}}$ to the specification trace model.
+  \MPin{
+  very nice.
+  final bit, perhaps highlight those events taht are mentioned in the text, so they are easily recognisable?
+  like yellow BG for the events, gray BG for the locks 
+  }
 
@@
 	Finally, \Thmref{thm:ccab:rtp:ms} ensures that ??? gets compiled into ????, which is CCT assuming that the original program was CCT to begin with.
+	\MPin{please fill in the macros here}
   % \MKin{%
@@
 \section{Conclusion\pages{$\sfrac{1}{2}$}}\label{sec:concl}
+This paper tackled the problem of understanding what kind of security properties does a secure compiler preserve, when said compiler is the combination of smaller compiler passes that preserve possibly different security properties.
+% 
+For this, this paper first formalised security properties of interest and their composition.
+% 
+Then, it proved that composing secure compilers that preserve certain properties results in a secure compiler that preserves the composition of the properties.
+% 
+Finally, this paper defines a multi-pass compiler and proves that it preserves \gls{msscct}.
+Crucially, this paper derives the security of the multi-pass compiler from the composition of the security properties preserved by its individual passes, which include security-preserving as well as optimisation passes.
 

### 2023-07-12 popl/main.tex
typo fix

--- 2023-07-11 popl/main.tex
@@
 Robust compilers~\cite{abate2019jour} are a form of secure compilers that preserve security properties even in the presence of arbitrary attackers interacting with compiled code.
-Thus, robust compilers can be used to prevent the vulnerability resulting from uninitialised memory, e.g., by targeting capability-based languages such as CHERI~\cite{woodruff2014CHERI}, Arm Morello~\cite{arm-morello}, or MSWasm~\cite{michael2023mswasm}, where pointers are always initialised.
+Thus, robust compilers can be used to prevent these vulnerabilities, e.g., by targeting capability-based languages such as CHERI~\cite{woodruff2014CHERI}, Arm Morello~\cite{arm-morello}, or MSWasm~\cite{michael2023mswasm}, where pointers are always initialised.
 \end{example}
@@
 % Different kinds of compilation passes exist, e.g., there are usually several compilation passes that perform optimisations~\cite{lattner2004llvm,googlev8,androidstudio,kuepper2023cryptopt,manjikian1997fusion,wegman1991ccp} or source-code instrumentations~\cite{nagarakatte2009soft,nagarakatte2010cets,akritidis2009baggy,dhumbumroong2020boundwarden,jung2021pico,nam2019framer,shankaranarayana2023tailcheck,younan2010paricheck,zhou2023fatptrs,bond2017vale,almeida2017jasmin,kuepper2023cryptopt,cauligi2019fact} that enforce that the compiled program fulfils a property of interest.
-Worse, without a framework for composing secure compiler passes, it is not possible to enable separation of concerns, e.g., to have a secure compilation pass that ensures \gls{ms} that is developed independently of another secure pass for \gls{cct}, that is developed independently of optimisation passes.
+Worse, without a framework for composing secure compiler passes, it is not possible to enable separation of concerns, e.g., to have a secure compilation pass that ensures \gls{ms} that is developed independently of another secure pass for \gls{cct}.
 
@@
 Finally, this paper shows that composing these secure compiler passes into a multi-pass compilation chain results that provides end-to-end preservation of \gls{msscct}.
-Crucially, this paper derives the security of the multi-pass compiler from the composition of the security properties preserved by its individual passes.
-% Crucially, this end-to-end security preservation is obtained by composing the security property preserved by each individual pass, as dictated by the framework of this paper.
-This result showcases how the framework allows the kind of formal security reasoning that compiler writers already want (and do), obtaining precise, compositional security reasoning while providing minimal (and modular) proof effort.
+Crucially, this end-to-end security preservation is obtained by composing the security property preserved by each individual pass, as dictated by the framework of this paper.
+The results showcase how the framework allows the kind of formal security reasoning that compiler writers already want (and do), obtaining precise, compositional security reasoning while providing minimal (and modular) proof effort.
 
@@
         %\Cref{sec:compprop} presents a formal framework to reason about compositions of security properties and demonstrate its usage with concrete examples:
-        This paper formalises security properties (\Cref{sec:compprop}) that are of interest for real-world compiler writers (as identified by the plethora of work enforcing such properties individually~\cite{akritidis2009baggy,nagarakatte2009soft,nagarakatte2010cets,dhumbumroong2020boundwarden,jung2021pico,nam2019framer,shankaranarayana2023tailcheck,younan2010paricheck,zhou2023fatptrs,bond2017vale,almeida2017jasmin,kuepper2023cryptopt,cauligi2019fact}), namely \gls{tms}, \gls{sms} and \gls{cct}.
-        % 
-        Starting from ways to formalise those properties individually, this paper shows how to compose their formalisation.
+        First, this paper formalises security properties (\Cref{sec:compprop}) by focussing on those properties that are of interest for real-world compiler writers (as identified by the plethora of work enforcing such properties individually~\cite{akritidis2009baggy,nagarakatte2009soft,nagarakatte2010cets,dhumbumroong2020boundwarden,jung2021pico,nam2019framer,shankaranarayana2023tailcheck,younan2010paricheck,zhou2023fatptrs,bond2017vale,almeida2017jasmin,kuepper2023cryptopt,cauligi2019fact}), namely \gls{tms}, \gls{sms} and \gls{cct}.
+        % 
+        Starting from ways to formalise and enforce those properties individually, this paper shows how to compose their formalisation as well as their enforcement.
         % 
@@
 
-\Cref{thm:ccab:rtp:ms} strates that the composition of $\cca$ and $\ccb$ is secure with respect to \gls{ms} and follows from \Cref{thm:cca:rtp:tms,thm:ccb:rtp:sms} using \Cref{thm:rtp}.
+\Cref{thm:ccab:rtp:ms} states that the composition of $\cca$ and $\ccb$ is secure with respect to \gls{ms} and follows from \Cref{thm:cca:rtp:tms,thm:ccb:rtp:sms} using \Cref{thm:rtp}.
 
@@
 
-% This paper does a first step towards practical secure compilation chains by introducing a theoretical framework demonstrating that secure compilers compose.
-% The paper exercises this framework on a case-study that consists of an optimising compilation chain that is secure with respect to \gls{msscct}.
-% \MP{
-% 	not an appropriate recap.
-% 	rewrite summarising the paper
-% }
-
-% In future work, it would be interesting to investigate whether it is possible to provide {\em secure compiler combinators}, similarly to how parser combinators work.
-% %To this end, it may be necessary to extend existing frameworks for multi-language semantics.
-% It is also interesting to extend our case-study and verify all results of it in Coq.
-% This way, it is possible to obtain an executable, secure compiler.
-% However, the formalisation effort for backtranslation proofs is known to be enormous\MP{cit + details}, so another research avenue may be to find better ways to (semi-)automate standard secure compilation proofs.
-% \MPin{
-% 	not particularly informative FW. i'd rather not have it as such
-% }
+This paper does a first step towards practical secure compilation chains by introducing a theoretical framework demonstrating that secure compilers compose.
+The paper exercises this framework on a case-study that consists of an optimising compilation chain that is secure with respect to \gls{msscct}.
+\MP{
+	not an appropriate recap.
+	rewrite summarising the paper
+}
+
+In future work, it would be interesting to investigate whether it is possible to provide {\em secure compiler combinators}, similarly to how parser combinators work.
+%To this end, it may be necessary to extend existing frameworks for multi-language semantics.
+It is also interesting to extend our case-study and verify all results of it in Coq.
+This way, it is possible to obtain an executable, secure compiler.
+However, the formalisation effort for backtranslation proofs is known to be enormous\MP{cit + details}, so another research avenue may be to find better ways to (semi-)automate standard secure compilation proofs.
+\MPin{
+	not particularly informative FW. i'd rather not have it as such
+}
 

### 2023-07-12 popl/main.tex
mix

@@
   \begin{align*}
-    \cccf(\irl{\llet{x}{n}{\expr}}) & = \cccf(\partialeval{\irl{\expr}}{\irl{x}}{\irl{n}}) &\\
-    \cccf(\irl{\lbinop{n_{1}}{n_{2}}}) & = \irl{n_{3}} & \text{where }\irl{n_{3}}\text{ is }\lbinop{\irl{n_{1}}}{\irl{n_{2}}}
+    \cccf(\irl{\expr}) & = \partialeval{\irl{\expr}}{\irl{\hole{\cdot}}} &\\
   \end{align*}\\[0.125cm]
   \begin{align*}
-    \partialeval{\irl{x}}{\irl{x}}{\irl{n}} & = \irl{n} &\\
-    \partialeval{\irl{y}}{\irl{x}}{\irl{n}} & = \irl{y} &\text{where }\irl{x}\not=\irl{y}\\
-    \partialeval{\left(\irl{\llet{x}{\expr[_{1}]}{\expr[_{2}]}}\right)}{\irl{x}}{\irl{n}} & = \llet[\irl]{\irl{x}}{\left(\partialeval{\irl{\expr[_{1}]}}{\irl{x}}{\irl{n}}\right)}{\irl{\expr[_{2}]}} &\\
-    \partialeval{\left(\irl{\llet{y}{\expr[_{1}]}{\expr[_{2}]}}\right)}{\irl{x}}{\irl{n}} & = \llet[\irl]{\irl{y}}{\left(\partialeval{\irl{\expr[_{1}]}}{\irl{x}}{\irl{n}}\right)}{\left(\partialeval{\irl{\expr[_{2}]}}{\irl{x}}{\irl{n}}\right)} &\text{where }\irl{x}\not=\irl{y}\\
-    \partialeval{\left(\irl{\lbinop{\expr[_{1}]}{\expr[_{2}]}}\right)}{\irl{x}}{\irl{n}} & = \lbinop[\irl]{\left(\partialeval{\irl{\expr[_{1}]}}{\irl{x}}{\irl{n}}\right)}{\left(\partialeval{\irl{\expr[_{2}]}}{\irl{x}}{\irl{n}}\right)}
+    \partialeval{\irl{x}}{\irl{\substlist}} & = \irl{n} & \text{if } \irl{\subst{n}{x}}\in\irl{\substlist} \\
+    \partialeval{\irl{x}}{\irl{\substlist}} & = \irl{x} & \text{if } \irl{\subst{n}{x}}\notin\irl{\substlist} \\
+    \partialeval{\irl{\lbinop{n}{m}}}{\irl{\substlist}} & = \irl{k} & \text{if } \lbinop{\irl{n}}{\irl{m}}=k \\
+%    \partialeval{\irl{\lbinop{\expr[_{1}]}{\expr[_{2}]}}}{\irl{\substlist}} & = \lbinop[\irl]{\partialeval{\irl{\expr[_{1}]}}{\irl{\substlist}}}{\partialeval{\irl{\expr[_{2}]}}{\irl{\substlist}}} \\
+    \partialeval{\irl{\llet{x}{n}{\expr}}}{\irl{\substlist}} & = \partialeval{\irl{\expr}}{\irl{\subst{x}{n}},\irl{\substlist}} \\
+    \partialeval{\irl{\lget{x}{\expr}}}{\irl{\substlist}} & = \lget[\irl]{\irl{x}}{\partialeval{\irl{\expr}}{\irl{\substlist}}}
+  \end{align*}\\
+  \begin{align*}
+    \partialeval{\irl{\llet{x}{\expr[_{1}]}{\expr[_{2}]}}}{\irl{\substlist}} & = \llet[\irl]{\irl{x}}{\partialeval{\irl{\expr[_{1}]}}{\irl{\substlist}}}{\partialeval{\irl{\expr[_{2}]}}{\irl{\substlist}}} \\
+    \partialeval{\irl{\lifz{\expr[_{1}]}{\expr[_{2}]}{\expr[_{3}]}}}{\irl{\substlist}} & = \lifz[\irl]{\partialeval{\irl{\expr[_{1}]}}{\irl{\substlist}}}{\partialeval{\irl{\expr[_{2}]}}{\irl{\substlist}}}{\partialeval{\irl{\expr[_{3}]}}{\irl{\substlist}}} \\
   \end{align*}
@@
 The \gls{dce} pass applies a naive rewrite rule on conditionals.
-For \gls{cf}, the pass uses an auxiliary function \texttt{mix} that models the substitution operation.
+For \gls{cf}, the pass uses an auxiliary function \texttt{mix} that does the actual work.
+It rewrites constant binary operations, e.g., $\irl{{17}+{29}-1}$ to $\irl{46-1}$, and replaces variables that are assigned to constants with their constant, e.g., $\irl{\llet{x}{7}{x}}$ to $\irl{7}$.
 Both passes are secure with respect to \gls{ms}.
@@
     \ccscct(\irl{\lfunction{g}{x}{\expr}}) & = & \lfunction[\obj]{\obj{g}}{\obj{x}}{\obj{\lwrdoit{1};}\ccscct(\irl{\expr})} \\
-    \ccscct(\irl{\lcall{g}{\expr}}) & = & \lcall[\obj]{\obj{g}}{\ccscct(\expr)\obj{; wrdoit{1}}}
+    \ccscct(\irl{\lcall{g}{\expr}}) & = & \lcall[\obj]{\obj{g}}{\ccscct(\irl{\expr})\obj{; wrdoit{1}}}
   \end{array}

### 2023-07-12 popl/main.tex
mix
dcemix
proof sketch 6.1.1

@@
 
-The proof of \Cref{thm:cca:rtp:tms} uses standard techniques in secure compilation literature that are also used for the following theorems of this section.
-\MP{so what?}
-
-When proving robust safety property preservation for the concrete property \gls{tms}, a target-level execution of a compiled $\Ltms$ component $\src{\library_{\comp}}$ linked with an arbitrary context $\trg{\library_{\ctx}}$ is assumed, where the execution yields a trace $\trg{\trace}$.
-\MP{passive + not justified}
-The proof goal is to show that $\trg{\trace}$ satisfies \gls{tms}.
-\MP{no, the goal is to BT}
-To get this satisfaction, one way of doing so is to use the other assumption in the proof context: $\rsat{\src{\library_{\comp}}}{\tmssafe}$.
-However, this requires a source-level execution for some concrete $\src{\library_{\ctx}}$ that yields a trace $\src{\trace}$ similar to $\trg{\trace}$.
-This paper uses the well-known trace-based backtranslation technique~\cite{abate2019jour} to construct such a context $\src{\library_{\ctx}}$.
-It is shown for that obtained context $\src{\library_{\ctx}}$ that the whole program $\link{\src{\library_{\ctx}}}{\src{\library_{comp}}}$ is well-typed.\MP{passive}
-By \Cref{thm:wt:tms}, the fact that this program satisfies \gls{tms} is obtained and by typesafety, it is known that there must be some $\Ltms$ execution.\MP{passive passive passive, and imprecise}
-The proof now needs to relate the two execution with each other.
-\Cref{fig:proofdiag:rtp} visualises this situation\MP{imprec}.
 \begin{figure}[!ht]
@@
       % legend
-      \node[align=left,below right=0.3 and 0.3 of trgempty,font=\tiny] (legend) {%
-        ${\src{\trace[_c]}}\cong{\backt{\Ltms}{\Ltrg}(\trg{Call\ ?\ foo\ \valueexpr})}$\\%
-        ${\src{\trace[_r]}}\cong{\backt{\Ltms}{\Ltrg}(\trg{Ret\ !\ foo\ \valueexpr})}$
-        };
+%     \node[align=left,below right=0.3 and 0.3 of trgempty,font=\tiny] (legend) {%
+%       ${\src{\trace[_c]}}\cong{\backt{\Ltms}{\Ltrg}(\trg{Call\ ?\ foo\ \valueexpr})}$\\%
+%       ${\src{\trace[_r]}}\cong{\backt{\Ltms}{\Ltrg}(\trg{Ret\ !\ foo\ \valueexpr})}$
+%       };
+      \node[below=0.5 of trgempty,draw=none] (legend) {};
       \draw[ultra thick,loosely dotted,Peach!50,rounded corners] ($(legend.north east)+(0.5,-0.4)$) -- ($(legend.north east)+(1,-0.4)$);
@@
     \caption{Proof diagram for \Cref{thm:cca:rtp:tms} depicting the general structure of robust preservation proofs. %
+      Nodes in the graph represent runtime states. %
       Vertical lines indicate cross language relations, while horizontal ones are execution steps. %
-      Nodes in the graph represent runtime states.
+      The green dashed trapezoid encompasses the component segment, while the orange dotted rectangles entail the context switches. %
+      $\Ltrg$ traces are omitted for readability. %
+      $\Ltms$ trace segments $\src{\trace[_{c}]}$ and $\src{\trace[_{r}]}$ describe the events that happen at the boundaries, i.e., during a context switch.
+      $\src{\trace_{p}}$ is the behavior of the component and the traces $\src{\trace[_{1}]}$ and $\src{\trace[_{2}]}$ describe the context.
     }\label{fig:proofdiag:rtp}
@@
 
-At the bottom, there is the target-level execution between operational states $\trg{\emptyset},\trg{\Omega_{1}}$, and so on.
-The execution starts in the context $\trg{\library_{\ctx}}$ and is split by transitivity of $\trg{\estepto{}{}{}}$ into parts before, in, and after the component $\cc{\Ltms}{\trg{L}}(\src{p})$ executes\footnote{For sake of simplicity, this paper considers only one context-switch from context to component and back again.
-Recursive calls between context and component functions are disregarded.}.
-Initially, the empty operational states $\src{\emptyset}$ and $\trg{\emptyset}$ need to agree by means of a cross-language state relation $\multimap_{\emptyset}$.
-This relation is indexed by a function $\delta$ that maps $\trg{L}$-level locations $\trg{\loc}$ to $\Ltms$-level locations.
-The operational state of the executions up to the component $\src{p}$ need to agree via $\multimap_{\delta}$, both contexts need to behave in a similar way such that the traces also agree with each other.
-However, the relation allows $\trg{\library_{\ctx}}$ to perform internal actions that may change the operational state but are not visible on the trace, such as internal function calls, that does not $\src{\library_{\ctx}}$ need to reflect.
-Before and after the compiler wrapper code, the operational states must agree via another cross-language relation $\approx_{\delta_{w_{}}}$, which disallows these mismatches with respect to internal function calls.
-Intuitively, the behavior of the components should be tight: the source-level execution must be preserved by the compiler in the target-level execution.
-Dually, when the component is done, the relations take turns again and the states must be related with $\multimap_{\delta_{w_{2}}}$
-Finally, since both execution traces are similar in the sense that their projections $\msfilterLtms{\src{\trace}}$ and $\msfilterL{\trg{\trace}}$ yield specification traces that satisfy \Cref{def:monsat}, the claim follows \Cref{lem:mon:tmssafe}.
-
-\MPin{
-	why can we use a BT?
-	it's because the def in section 2.2 is eq to anotehr one, because the props are Safety, we use the RSC equivalent def.
-	all these intuitions are missing.
-	they are technical, so we only keep them here.
-
-  MK: but, we don't actually rewrite it to RSC when doing the proof! The compilers are not robustly preserving with respect to an arbitrary safety property, but, e.g., just for \gls{tms}!
-}
+The proof of \Cref{thm:cca:rtp:tms} serves as boilerplate for proofs of the other secure compilation theorems of this paper.
+Unfolding the statement gives the assumptions that for any $\pi\in\lceil\tmssafe\rceil$\footnote{$\lceil\cdot\rceil$ lifts the property to a hyperproperty by applying the powerset operation~\cite{clarkson2008hyper}.}, $\src{\trace}$, $\src{\runtimetermvar}$, and component $\src{\library_{comp}}$, $\rsat{\src{\library_{comp}}}{\pi}$ and $\progstepto{\prog[\trg]{\trg{\library_{\ctx}}}{\cca(\src{\library_{\comp}})}}{\trg{\runtimetermvar}}{\trg{\trace}}$, where $\trg{\library_{\ctx}}$ is arbitrary.
+The proof obligation is $\msfilterL{\trg{\trace}}\in\pi$, i.e., the specification trace associated to $\trg{\trace}$ should satisfy the property $\pi$.
+To show this, the trace $\trg{\trace}$ should be related with some $\Ltms$ trace $\src{\trace}$ and since there is a target execution associated to this trace, the task is to find an associated $\Ltms$ execution.
+The trace $\trg{\trace}$ is split into different parts, where each part contains the events that either the context or the component does, but not both.
+Because of this, all such trace segments are ,,well-bracketed'' in the sense that they start with either $\trg{\ev{Start}}$, $\trg{\ev{Call}\ \comptoctx\ foo\ \valueexpr}$, or $\trg{\ev{Ret}\ \comptoctx\ \valueexpr}$ and end with either $\trg{\ev{End}\ \valueexpr}$, $\trg{\ev{Ret}\ \ctxtocomp\ \valueexpr}$, or $\trg{\ev{Call}\ \ctxtocomp\ foo\ \valueexpr}$.
+In the following, the former is referred to as a context segment, since, for these, execution happens in $\trg{\library_{\ctx}}$, and the latter as a component segment, since, for these, execution happens in $\cca(\src{\library_{\comp}})$.
+\Cref{fig:proofdiag:rtp} visualises this split for a program execution with one call from context to component and how the target execution is related to a source execution.
+In the figure, the green dashed lines encompass the component segments while the orange boxes contain the actual context switches from context to component or vice versa.
+For the proof, it is necessary to know when two states $\src{\Omega}$ and $\trg{\Omega}$ are related.
+Two cross-language relations make this precise: (i) $\multimap_{\delta}$ relates states that are involved in a context segment, allowing the target execution to perform internal calls, and (ii) $\approx_{\delta}$ relates states that are involved in a component segment, where both states need to agree exactly, i.e., the memory and the control flow states are required to contain the same information.
+The relations are indexed with $\delta$, which is simply an injective mapping from $\Ltms$ locations $\src{\loc}$ to $\Ltrg$ locations $\trg{\loc}$.
+Note that the relations $\multimap_{\delta}$ and $\approx_{\delta}$ swap when context switching.
+
+So far, the paper explained how to relate an $\Ltms$ execution with the $\Ltrg$ execution at hand.
+However, it is not yet demonstrated how the corresponding $\Ltms$ execution is actually built.
+To do so, a trace-based backtranslation technique~\cite{abate2019jour} can be used to build a context $\src{\library_{\ctx}}$ that behaves similar to $\trg{\library_{\ctx}}$.
+Since $\Ltms$ is statically typed, the context $\src{\library_{\ctx}}$ obtained from the backtranslation needs to be well-typed when linked with $\src{\library_{\comp}}$.
+For context segments of the trace $\trg{\trace}$ it is also necessary to show that the execution behaves similarily, i.e., there is an $\Ltms$ trace segment $\src{\trace[']}$ that is related to the context segment, and that the states are related by $\multimap_{\delta}$.
+When passing control to the component, the relatedness of states and traces follows from compiler correctness.
+From all this, a source execution $\progstepto{\src{\prog{\library_{\ctx}}{\library_{\comp}}}}{\src{\runtimetermvar}}{\src{\trace}}$ can be obtained.
+
+The proof now works as follows.
+Rewrite the goal from $\msfilterL{\trg{\trace}}\in\pi$ to $\msfilterLtms{\src{\trace}}\in\pi$ after demonstrating that $\msfilterL{\trg{\trace}}=\msfilterLtms{\src{\trace}}$.
+Then, specialize the robust satisfaction assumption $\rsat{\src{\library_{\comp}}}{\pi}$ to use the context $\src{\library_{\ctx}}$, which was obtained from backtranslating the context segments of the target trace $\trg{\trace}$, and the source execution $\progstepto{\src{\prog{\library_{\ctx}}{\library_{\comp}}}}{\src{\runtimetermvar}}{\src{\trace}}$.
+This gives $\msfilterLtms{\src{\trace}}\in\pi$, which was to be demonstrated.
 
@@
 \begin{example}[Instrumented \texttt{strncpy}]
-    Consider again \texttt{strncpy}, but this time with an instrumented version of \texttt{strncpy}:
+  Consider again \texttt{strncpy}, but instrumented for \gls{sms}:
     \begin{lstlisting}[language=c,basicstyle=\ttfamily\small, morekeywords={size_t}]
@@
     \end{lstlisting}
-    When calling this in similar fashion to \Cref{ex:strncpy:sms}, the event $\ev{Use}\ \loc_{x}\ 2;\comp;\unlock$ would not be emitted during execution.
+    When calling this in similar fashion to \Cref{ex:strncpy:sms}, the event $\ev{Use}\ \loc_{x}\ 2;\comp;\unlock$ would not be emitted during execution, since the bounds check prevents the condition \texttt{src[i] != '\textbackslash 0'} from executing.
 \end{example}
@@
   \begin{align*}
-    \ccdce(\irl{\lifz{true}{\expr[_{1}]}{\expr[_{2}]}}) & = & \ccdce(\irl{\expr[_{1}]}) \\
-    \ccdce(\irl{\lifz{false}{\expr[_{1}]}{\expr[_{2}]}}) & = & \ccdce(\irl{\expr[_{2}]})
+    \ccdce(\irl{\lifz{true}{\expr[_{1}]}{\expr[_{2}]}}) & = \ccdce(\irl{\expr[_{1}]}) &\\
+    \ccdce(\irl{\lifz{false}{\expr[_{1}]}{\expr[_{2}]}}) & = \ccdce(\irl{\expr[_{2}]}) &
+  \end{align*}\\
+  \begin{align*}
+    \ccdce(\irl{\lbinop{\expr[_{1}]}{\expr[_{2}]}}) & = \lbinop[\irl]{\ccdce(\irl{\expr[_{1}]})}{\ccdce(\irl{\expr[_{2}]})} &
   \end{align*}\\[0.25cm]
   \begin{align*}
-    \cccf(\irl{\expr}) & = \partialeval{\irl{\expr}}{\irl{\hole{\cdot}}} &\\
+    \cccf(\irl{\expr}) & = \partialeval{\irl{\expr}}{\irl{\hole{\cdot}}} &
   \end{align*}\\[0.125cm]
@@
 \begin{center}
-  $$
-  \begin{array}{rcl}
-    \ccscct(\irl{\lfunction{g}{x}{\expr}}) & = & \lfunction[\obj]{\obj{g}}{\obj{x}}{\obj{\lwrdoit{1};}\ccscct(\irl{\expr})} \\
-    \ccscct(\irl{\lcall{g}{\expr}}) & = & \lcall[\obj]{\obj{g}}{\ccscct(\irl{\expr})\obj{; wrdoit{1}}}
-  \end{array}
-  $$
+  \begin{align*}
+    \ccscct(\irl{\lfunction{g}{x}{\expr}}) & = \lfunction[\obj]{\obj{g}}{\obj{x}}{\obj{\lwrdoit{1};}\ccscct(\irl{\expr})} \\
+    \ccscct(\irl{\lcall{g}{\expr}}) & = \lcall[\obj]{\obj{g}}{\ccscct(\irl{\expr})\obj{; wrdoit{1}}} \\
+    \ccscct(\irl{\lbinop{\expr[_{1}]}{\expr[_{2}]}}) & = \lbinop[\obj]{\ccscct{\irl{\expr[_{1}]}}}{\ccscct{\irl{\expr[_{2}]}}} \\
+  \end{align*}
 \end{center}

### 2023-07-12 popl/main.tex
fix

@@
 
-\gls{ms} is a security property obtained by composing \gls{sms}, which ensures array accesses are all within bounds, and \gls{tms}, which ensures pointers are only used when they are valid~\cite{azevedo2018meaningsofms}\MP{more cits}.
+\gls{ms} is a security property obtained by composing \gls{sms}, which ensures array accesses are all within bounds, and \gls{tms}, which ensures pointers are only used when they are valid~\cite{azevedo2018meaningsofms,jim2002cyclone,necula2005ccured,nagarakatte2010cets,nagarakatte2009soft,akritidis2009baggy,michael2023mswasm}.
 \gls{cct} is a security property that ensures sensitive data is not leaked via timing side-channels~\cite{kocher1996timing}.

### 2023-07-12 popl/main.tex
add star to glossaries

--- 2023-07-12 popl/main.tex
@@
 
-% \gls{ms} is a security property that ensures that there are no out-of-bounds accesses as well as no reads or writes to uninitialised memory.
+% \gls*{ms} is a security property that ensures that there are no out-of-bounds accesses as well as no reads or writes to uninitialised memory.
 % Programs that do not attain memory safety, can easily have their private data be leaked to attackers or be corrupted~\cite{lemay2021ccc}.
-% Because of this, \gls{ms} is a requirement for programs to attain an even stronger security property, namely \gls{cct}.
-% Programs that attain \gls{cct} are secure with respect to timing attacks, and this is the gold standard for secure applications, including cryptographic ones, since timing attacks can expose secrets via a simple performance analysis~\cite{kocher1996timing}.
+% Because of this, \gls*{ms} is a requirement for programs to attain an even stronger security property, namely \gls*{cct}.
+% Programs that attain \gls*{cct} are secure with respect to timing attacks, and this is the gold standard for secure applications, including cryptographic ones, since timing attacks can expose secrets via a simple performance analysis~\cite{kocher1996timing}.
 % Timing attacks have become more and more relevant recently, especially given the era of cloud-computing~\cite{aviram2010cloudtime,kumar2019cloudsecsurvey}, where programs share the same hardware but are supposed to run in a sandboxed environment that does not leak secrets.
@@
 % Different kinds of compilation passes exist, e.g., there are usually several compilation passes that perform optimisations~\cite{lattner2004llvm,googlev8,androidstudio,kuepper2023cryptopt,manjikian1997fusion,wegman1991ccp} or source-code instrumentations~\cite{nagarakatte2009soft,nagarakatte2010cets,akritidis2009baggy,dhumbumroong2020boundwarden,jung2021pico,nam2019framer,shankaranarayana2023tailcheck,younan2010paricheck,zhou2023fatptrs,bond2017vale,almeida2017jasmin,kuepper2023cryptopt,cauligi2019fact} that enforce that the compiled program fulfils a property of interest.
-% Without a framework for composing secure compiler passes, it is not possible to enable separation of concerns, e.g., to have a secure compilation pass that ensures \gls{ms} that is developed independently of another secure pass for \gls{cct} as suggested before and as showcased in \Cref{ex:strncpy}.
+% Without a framework for composing secure compiler passes, it is not possible to enable separation of concerns, e.g., to have a secure compilation pass that ensures \gls*{ms} that is developed independently of another secure pass for \gls*{cct} as suggested before and as showcased in \Cref{ex:strncpy}.
 \MPpost{
@@
 
-\gls{ms} is a security property obtained by composing \gls{sms}, which ensures array accesses are all within bounds, and \gls{tms}, which ensures pointers are only used when they are valid~\cite{azevedo2018meaningsofms}\MP{more cits}.
-\gls{cct} is a security property that ensures sensitive data is not leaked via timing side-channels~\cite{kocher1996timing}.
-Together, \gls{sms}, \gls{tms} and \gls{cct} yield \gls{msscct}, which is the gold standard of security properties for secure applications: programs attaining \gls{msscct} do not leak sensitive data neither through erroneous memory accesses, nor through timing side-channels.
+\gls*{ms} is a security property obtained by composing \gls*{sms}, which ensures array accesses are all within bounds, and \gls*{tms}, which ensures pointers are only used when they are valid~\cite{azevedo2018meaningsofms,jim2002cyclone,necula2005ccured,nagarakatte2010cets,nagarakatte2009soft,akritidis2009baggy,michael2023mswasm}.
+\gls*{cct} is a security property that ensures sensitive data is not leaked via timing side-channels~\cite{kocher1996timing}.
+Together, \gls*{sms}, \gls*{tms} and \gls*{cct} yield \gls*{msscct}, which is the gold standard of security properties for secure applications: programs attaining \gls*{msscct} do not leak sensitive data neither through erroneous memory accesses, nor through timing side-channels.
 As discussed in \Cref{ex:strncpy}, these security properties can be enforced by compiler passes~\cite{bond2017vale,almeida2017jasmin}, to ensure programmers need not be aware of the various architectural details where their code will run.
@@
 % Even under these assumptions, there is a 
-This function is subject to a subtle \gls{sms} vulnerability: The bounds check \texttt{i < n} should happen {\it before} the access to memory location \texttt{x[i]}: otherwise 
+This function is subject to a subtle \gls*{sms} vulnerability: The bounds check \texttt{i < n} should happen {\it before} the access to memory location \texttt{x[i]}: otherwise
 % If \texttt{n} is at least as large as the length of the array \texttt{src}, 
@@
 
-To prevent this vulnerability, one can use a compilation pass that enforces \gls{sms}, such as Softbounds~\cite{nagarakatte2009soft} or BaggyBounds~\cite{akritidis2009baggy}.
-
-Because of timing attacks, fixing \gls{sms} is not enough to make \texttt{strncpy} secure.
+To prevent this vulnerability, one can use a compilation pass that enforces \gls*{sms}, such as Softbounds~\cite{nagarakatte2009soft} or BaggyBounds~\cite{akritidis2009baggy}.
+
+Because of timing attacks, fixing \gls*{sms} is not enough to make \texttt{strncpy} secure.
 In fact, the loop can terminate early, as soon as the string-terminating character \texttt{'\textbackslash 0'} is encountered, making program execution time proportional to the length of the array pointed by \texttt{src}.
 % Often, programmers rewrite programs like these to have loops that run with a fixed number of iterations.
-Also in this case there exist compiler passes that can rewrite such programs into \gls{cct} ones~\cite{cauligi2019fact}.
+Also in this case there exist compiler passes that can rewrite such programs into \gls*{cct} ones~\cite{cauligi2019fact}.
  % so that they are no longer vulnerable to timing attacks.
 
-Alas, code is not run in isolation, so a malicious attacker could supply code that intracts with \texttt{strncpy} and trigger a violation of either \gls{ms} or \gls{cct} by calling \texttt{strncpy} with an argument for \texttt{src} that points to uninitialised memory.
-This would, in turn, triggering a series of reads from uninitialised memory, which is an immediate \gls{ms} violation with devastating real-world implications~\cite{uninit-0,uninit-1,uninit-2,uninit-3,uninit-4}.
+Alas, code is not run in isolation, so a malicious attacker could supply code that intracts with \texttt{strncpy} and trigger a violation of either \gls*{ms} or \gls*{cct} by calling \texttt{strncpy} with an argument for \texttt{src} that points to uninitialised memory.
+This would, in turn, triggering a series of reads from uninitialised memory, which is an immediate \gls*{ms} violation with devastating real-world implications~\cite{uninit-0,uninit-1,uninit-2,uninit-3,uninit-4}.
 % An easy fix to sidestep this is to use non-nullable pointers as parameters instead, e.g., as presented in Cyclone~\cite{jim2002cyclone}.
@@
 % Different kinds of compilation passes exist, e.g., there are usually several compilation passes that perform optimisations~\cite{lattner2004llvm,googlev8,androidstudio,kuepper2023cryptopt,manjikian1997fusion,wegman1991ccp} or source-code instrumentations~\cite{nagarakatte2009soft,nagarakatte2010cets,akritidis2009baggy,dhumbumroong2020boundwarden,jung2021pico,nam2019framer,shankaranarayana2023tailcheck,younan2010paricheck,zhou2023fatptrs,bond2017vale,almeida2017jasmin,kuepper2023cryptopt,cauligi2019fact} that enforce that the compiled program fulfils a property of interest.
-Worse, without a framework for composing secure compiler passes, it is not possible to enable separation of concerns, e.g., to have a secure compilation pass that ensures \gls{ms} that is developed independently of another secure pass for \gls{cct}, that is developed independently of other passes such as optimisation ones.
+Worse, without a framework for composing secure compiler passes, it is not possible to enable separation of concerns, e.g., to have a secure compilation pass that ensures \gls*{ms} that is developed independently of another secure pass for \gls*{cct}, that is developed independently of other passes such as optimisation ones.
 
 This paper introduces a framework for reasoning about the composition of secure and optimising compiler passes akin to those of \Cref{ex:strncpy} and it showcases the power of this framework by instantiating it on a multi-pass compilation chain.
-To this end, this paper first discusses how to compose security properties, such as \gls{tms} and \gls{sms} into \gls{ms}, and then adding \gls{cct} to the mix to obtain \gls{msscct}. 
+To this end, this paper first discusses how to compose security properties, such as \gls*{tms} and \gls*{sms} into \gls*{ms}, and then adding \gls*{cct} to the mix to obtain \gls*{msscct}.
 Then, this paper defines compiler composition and formalises that given two passes that securely preserve two (possibly-distinct) properties, their composition securely preserves the composition of those properties.
-The paper then defines several secure compiler passes, where each is either preserving a different security property (\gls{tms}, \gls{sms}, \gls{cct}) or performing a security-preserving optimisation, (e.g., applying \gls{cf} or \gls{dce}).
-Finally, this paper shows that composing these secure compiler passes into a multi-pass compilation chain results that provides end-to-end preservation of \gls{msscct}.
+The paper then defines several secure compiler passes, where each is either preserving a different security property (\gls*{tms}, \gls*{sms}, \gls*{cct}) or performing a security-preserving optimisation, (e.g., applying \gls*{cf} or \gls*{dce}).
+Finally, this paper shows that composing these secure compiler passes into a multi-pass compilation chain results that provides end-to-end preservation of \gls*{msscct}.
 Crucially, this paper derives the security of the multi-pass compiler from the composition of the security properties preserved by its individual passes.
@@
         %\Cref{sec:compprop} presents a formal framework to reason about compositions of security properties and demonstrate its usage with concrete examples:
-        This paper formalises security properties (\Cref{sec:compprop}) that are of interest for real-world compiler writers, namely \gls{tms}, \gls{sms} and \gls{cct} (as identified by the plethora of work enforcing such properties individually~\cite{akritidis2009baggy,nagarakatte2009soft,nagarakatte2010cets,dhumbumroong2020boundwarden,jung2021pico,nam2019framer,shankaranarayana2023tailcheck,younan2010paricheck,zhou2023fatptrs,bond2017vale,almeida2017jasmin,kuepper2023cryptopt,cauligi2019fact}).
+        This paper formalises security properties (\Cref{sec:compprop}) that are of interest for real-world compiler writers, namely \gls*{tms}, \gls*{sms} and \gls*{cct} (as identified by the plethora of work enforcing such properties individually~\cite{akritidis2009baggy,nagarakatte2009soft,nagarakatte2010cets,dhumbumroong2020boundwarden,jung2021pico,nam2019framer,shankaranarayana2023tailcheck,younan2010paricheck,zhou2023fatptrs,bond2017vale,almeida2017jasmin,kuepper2023cryptopt,cauligi2019fact}).
         % 
@@
         % 
-        % This paper then showcases the benefits of this composition by composing \gls{tms} and \gls{sms} into \gls{ms}, and ultimately adding \gls{cct} to the composition, yielding the the conjunction of \gls{ms} and \gls{cct}.
-        % 
-        The resulting security property is \gls{msscct}, i.e., the gold standard of security properties for secure programs~\cite{lemay2021ccc}.
+        % This paper then showcases the benefits of this composition by composing \gls*{tms} and \gls*{sms} into \gls*{ms}, and ultimately adding \gls*{cct} to the composition, yielding the the conjunction of \gls*{ms} and \gls*{cct}.
+        % 
+        The resulting security property is \gls*{msscct}, i.e., the gold standard of security properties for secure programs~\cite{lemay2021ccc}.
 
@@
         This paper proves that starting from two compilers that preserve two (possibly distinct) properties, their composition preserves the intersection of those properties.
-        % For example, a compiler that is secure with respect to \gls{tms} can be composed with a compiler that is secure with respect to \gls{sms}, resulting in a compiler that is secure with respect to \gls{ms}.
+        % For example, a compiler that is secure with respect to \gls*{tms} can be composed with a compiler that is secure with respect to \gls*{sms}, resulting in a compiler that is secure with respect to \gls*{ms}.
         % 
@@
         This paper presents a case-study showcasing the conjunction of the previous contributions (\Cref{sec:casestud:defs,sec:casestud:rtp}).
-        To this end, it presents a compilation chain consisting of several passes that ultimately preserves \gls{msscct} by means of composing the individual, secure passes concerning \gls{tms}, \gls{sms}, and \gls{cct}, respectively.
-        Furthermore, the chain includes two optimisation passes: One performs a simple \gls{dce} and the other \gls{cf}.
+        To this end, it presents a compilation chain consisting of several passes that ultimately preserves \gls*{msscct} by means of composing the individual, secure passes concerning \gls*{tms}, \gls*{sms}, and \gls*{cct}, respectively.
+        Furthermore, the chain includes two optimisation passes: One performs a simple \gls*{dce} and the other \gls*{cf}.
         The formalisation of this case study showcases the power of the presented framework: Namely, the divide-and-conquer approach to software engineering is a viable strategy even for the development of secure compilers.
@@
 This program behaviour is insecure with respect to a canonical notion of (temporal) memory safety dictating no use-after-frees of pointers~\cite{nagarakatte2010cets,azevedo2018meaningsofms}, because it reads from a memory location that was freed already.
-The previous finite trace prefix is enough to decide that the trace does not satisfy \gls{tms} and there is no way to append events to this prefix which would result in the trace being admissible.
+The previous finite trace prefix is enough to decide that the trace does not satisfy \gls*{tms} and there is no way to append events to this prefix which would result in the trace being admissible.
 In the following, the execution of a whole program $\wholeprogvar$ that terminates in state $\runtimetermvar$ according to the language semantics and produces trace $\trace$ is written as $\progstepto{\wholeprogvar}{\runtimetermvar}{\trace}$.
@@
   The pointer is valid for one $\trg{char}$-sized memory cell, as expected, but the actual read operation now returns $\trg{sizeof(int)}$ many bytes instead of just $\trg{sizeof(char)}$ many.
-  Thus, even if $\src{foo}$ may have been robust with respect to the \gls{sms}, its compiled counterpart is not and therefore the compiler fails to attain \Cref{def:rtp}.
+  Thus, even if $\src{foo}$ may have been robust with respect to the \gls*{sms}, its compiled counterpart is not and therefore the compiler fails to attain \Cref{def:rtp}.
 \end{example}
@@
 
-This section introduces a trace model and uses it to define the key properties of interest for this paper: \gls{tms}, \gls{sms}, \gls{ms}, and \gls{scct} (\Cref{subsec:propdefs}).
+This section introduces a trace model and uses it to define the key properties of interest for this paper: \gls*{tms}, \gls*{sms}, \gls*{ms}, and \gls*{scct} (\Cref{subsec:propdefs}).
 These properties are of practical importance (as mentioned in \Cref{sec:introduction}) and also of interest in the case study (\Cref{sec:casestud:defs,sec:casestud:rtp}) this paper presents later. 
@@
 The specification trace model defines events as either the empty event ($\emptyevent$), a crash ($\lightning$), or as tuples consisting of a pre-event, a control-tag, and a security-tag.
-The purpose of the model is to define key security properties of interest, such as \gls{ms} or a stricter variant of cryptographic constant time.
+The purpose of the model is to define key security properties of interest, such as \gls*{ms} or a stricter variant of cryptographic constant time.
 To this end, security-tags indicate whether an event contains sensitive information ($\lock$) or not ($\unlock$), while control-tags state whether the context ($\ctx$) or the component ($\comp$) are responsible for emitting the event.
-The latter is necessary to be able to ignore actions done by a spurious context that, e.g., immediately deallocates a memory location twice, thus violating \gls{tms}~\cite{nagarakatte2010cets}.
+The latter is necessary to be able to ignore actions done by a spurious context that, e.g., immediately deallocates a memory location twice, thus violating \gls*{tms}~\cite{nagarakatte2010cets}.
 %Even though this could be used for blame~\cite{patrignani2023blame}, this is not what this paper is concerned about.
@@
 Dually, deallocation ($\ev{Dealloc\ \loc}$) announces that the object at location $\loc$ is freed.
-These two events alone are enough to provide a partial description of \gls{tms} by requiring that, e.g., there is only one deallocation event that carries a location $\loc$.
-To be able to express \gls{sms}, there is also an event to describe reads from and writes to memory ($\ev{Use\ \loc\ n}$).
+These two events alone are enough to provide a partial description of \gls*{tms} by requiring that, e.g., there is only one deallocation event that carries a location $\loc$.
+To be able to express \gls*{sms}, there is also an event to describe reads from and writes to memory ($\ev{Use\ \loc\ n}$).
 Finally, for cryptographic code, there is a general guideline that secrets must not be visible on a trace.
@@
 
-\gls{tms} is a safety property that describes that an unallocated object must not be used in any way.
+\gls*{tms} is a safety property that describes that an unallocated object must not be used in any way.
 Moreover, the property requires that all allocated objects must be deallocated at some point.
 
-\begin{definition}[\glsfirst{tms}]\label{def:trace:tmsdef}
+\begin{definition}[\glsfirst*{tms}]\label{def:trace:tmsdef}
   $$
@@
 
-\gls{sms} prohibits out-of-bounds accesses:
-
-\begin{definition}[\glsfirst{sms}]\label{def:trace:smsdef}
+\gls*{sms} prohibits out-of-bounds accesses:
+
+\begin{definition}[\glsfirst*{sms}]\label{def:trace:smsdef}
   $$
@@
 
-Full \gls{ms} (similar to earlier work~\cite{nagarakatte2009soft,nagarakatte2010cets,jim2002cyclone,necula2005ccured,michael2023mswasm}) is then described as the conjunction of \Cref{def:trace:tmsdef,def:trace:smsdef}.
+Full \gls*{ms} (similar to earlier work~\cite{nagarakatte2009soft,nagarakatte2010cets,jim2002cyclone,necula2005ccured,michael2023mswasm}) is then described as the conjunction of \Cref{def:trace:tmsdef,def:trace:smsdef}.
 %%% Catalin's paper doesn't apply, since their definition of MS relies on noninterference to guarantee data isolation.
@@
 
-\begin{definition}[\glsfirst{ms}]\label{def:trace:msdef}
+\begin{definition}[\glsfirst*{ms}]\label{def:trace:msdef}
   $$
@@
 
-\gls{cct} is a hypersafety property~\cite{barthe2018sec} and, thus, difficult to check with monitors.
+\gls*{cct} is a hypersafety property~\cite{barthe2018sec} and, thus, difficult to check with monitors.
 This is because, intuitively, hypersafety properties can relate multiple execution traces with eachother, but monitors work on a single execution.
-To sidestep this issue, this section defines the property \gls{scct}, a stricter variant of \gls{cct} that enforces the policy that no secret appears on a trace (inspired by earlier work~\cite{almeida2017jasmin}).
-
-\begin{definition}[\glsfirst{scct}]\label{def:trace:scctdef}
+To sidestep this issue, this section defines the property \gls*{scct}, a stricter variant of \gls*{cct} that enforces the policy that no secret appears on a trace (inspired by earlier work~\cite{almeida2017jasmin}).
+
+\begin{definition}[\glsfirst*{scct}]\label{def:trace:scctdef}
   $$
@@
 
-The combination of \gls{ms} and \gls{scct} is the intersection of these properties, \gls{msscct}.
-Since \gls{ms} and \gls{scct} are just sets of traces that, intuitively, contain all program behaviors that follow a security policy, the intersection of them contains all program behaviors that follow both security policies, i.e., it entails all program behaviours that are both \gls{ms} and \gls{scct}.
-
-\begin{definition}[\glsfirst{ms} and \gls{scct}]\label{def:trace:msscctdef}
+The combination of \gls*{ms} and \gls*{scct} is the intersection of these properties, \gls*{msscct}.
+Since \gls*{ms} and \gls*{scct} are just sets of traces that, intuitively, contain all program behaviors that follow a security policy, the intersection of them contains all program behaviors that follow both security policies, i.e., it entails all program behaviours that are both \gls*{ms} and \gls*{scct}.
+
+\begin{definition}[\glsfirst*{ms} and \gls*{scct}]\label{def:trace:msscctdef}
   $$
@@
 
-\subsubsection{Monitor for \glsfirst{tms}}\label{subsubsec:tms}
+\subsubsection{Monitor for \glsfirst*{tms}}\label{subsubsec:tms}
 \begin{gather*}
@@
 
-For \gls{tms}, the state of the monitor is a record with two sets keeping track of allocated and deallocated locations.
+For \gls*{tms}, the state of the monitor is a record with two sets keeping track of allocated and deallocated locations.
 %The monitor runs on a modified trace semantics with more abstract $\absev{Alloc\ }\loc$,$\absev{Dealloc\ }\loc$,$\absev{Use\ }\loc$, and $\absterm$ events, which represent {\em allocation}, {\em deallocation}, and {\em use} of a location $\loc$ as well as abnormal program termination.
@@
 However, the definition can easily be adapted by, e.g., attaching a natural number serving as a counter.
-Contrary to other monitors in this paper, the multi-step relation of the \gls{tms} monitor is non-standard:
+Contrary to other monitors in this paper, the multi-step relation of the \gls*{tms} monitor is non-standard:
 \begin{center}
@@
 \begin{center}
-  \judgbox{\tmstraceagree[]{\event}{\absevent}}{,,Abstract event $\absevent$ is equivalent to $\event$ with respect to \gls{tms}.''}$\;$\\
+  \judgbox{\tmstraceagree[]{\event}{\absevent}}{,,Abstract event $\absevent$ is equivalent to $\event$ with respect to \gls*{tms}.''}$\;$\\
   %
@@
 
-\begin{example}[A program not satisfying \gls{tms}]\label{ex:strncpy:cpp11}
+\begin{example}[A program not satisfying \gls*{tms}]\label{ex:strncpy:cpp11}
   Consider the following \texttt{C++11} library that calls \texttt{strncpy} (\Cref{ex:strncpy}) and prints the result to the standard output stream.
@@
   Remembering the definition of \texttt{strncpy} (\Cref{ex:strncpy}), observe that it does not deallocate its arguments.
-  Even though the trace contains an out-of-bounds access right before returning from \texttt{strncpy}, this is no concern for \gls{tms}, since the location $\loc[_{x}]$ is still allocated.
+  Even though the trace contains an out-of-bounds access right before returning from \texttt{strncpy}, this is no concern for \gls*{tms}, since the location $\loc[_{x}]$ is still allocated.
   However, having returned from \texttt{strncpy}, the \texttt{greet} function continues and deallocates $\loc[_{y}]$ whose subsequent use in the \texttt{printf} call is a use-after-free bug.
@@
   The fix would be to \lstinline[language=c++,basicstyle=\small\ttfamily]|delete greetings| instead of \texttt{to} and add a \lstinline[language=c++,basicstyle=\small\ttfamily]|delete to| after the \texttt{printf} call, which leads to the abstract monitor trace $\abstrace[']=\absev{Alloc}\ \loc[_{x}]\ \comp\cdot\absev{Alloc}\ \loc[_{y}]\ \comp\cdot\absev{Use}\ \loc[_{x}]\ \comp\cdot\absev{Use}\ \loc[_{x}]\ \comp\cdot\absev{Use}\ \loc[_{y}]\ \comp\cdot\ldots\cdot\absev{Use}\ \loc[_{x}]\ \comp\cdot\absev{Dealloc}\ \loc[_{x}]\ \comp\cdot\absev{Use}\ \loc[_{y}]\ \comp$.
-  It follows that $\montmssafe{\abstrace[']}$ and from \Thmref{lem:mon:tmssafe}, it follows that the program satisfies \Thmref{def:trace:tmsdef}, even though the program still violates \gls{sms}.
+  It follows that $\montmssafe{\abstrace[']}$ and from \Thmref{lem:mon:tmssafe}, it follows that the program satisfies \Thmref{def:trace:tmsdef}, even though the program still violates \gls*{sms}.
 \end{example}
 
-\subsubsection{Monitor for \glsfirst{sms}}\label{subsubsec:sms}
+\subsubsection{Monitor for \glsfirst*{sms}}\label{subsubsec:sms}
 \begin{gather*}
@@
 \end{center}
-The state of the monitor for \gls{sms} is a set containing tuples of locations, control-tags, and the allocation size.
-In comparison to the trace model of the \gls{tms} monitor, the trace model here is extended by sizing and positional information.
+The state of the monitor for \gls*{sms} is a set containing tuples of locations, control-tags, and the allocation size.
+In comparison to the trace model of the \gls*{tms} monitor, the trace model here is extended by sizing and positional information.
 \Cref{tr:sms-use} performs a bounds check and \Cref{tr:sms-alloc} adds bounds information to the state of the monitor.
-The trace agreement is entirely straightforward and similar to the one for \gls{tms}.
+The trace agreement is entirely straightforward and similar to the one for \gls*{tms}.
 % \begin{center}
-%   \judgbox{\smstraceagree[]{\event}{\absevent}}{,,Abstract event $\absevent$ is equivalent to $\event$ with respect to \gls{sms}.''}$\;$\\
+%   \judgbox{\smstraceagree[]{\event}{\absevent}}{,,Abstract event $\absevent$ is equivalent to $\event$ with respect to \gls*{sms}.''}$\;$\\
 %   %
@@
   All uses up to the last are accepted by the monitor, but the last event does not satisfy the premise $2<2$ in \Cref{tr:sms-use}.
-  Therefore, the whole program (\texttt{strncpy} linked with this kind of context) is not \gls{sms}.
+  Therefore, the whole program (\texttt{strncpy} linked with this kind of context) is not \gls*{sms}.
 \end{example}
@@
 
-\subsubsection{Combining \glsfirst{tms} and \glsfirst{sms} Monitors to obtain \glsfirst{ms}}\label{subsubsec:ms}
+\subsubsection{Combining \glsfirst*{tms} and \glsfirst*{sms} Monitors to obtain \glsfirst*{ms}}\label{subsubsec:ms}
 
@@
 \end{center}
-The combined monitor runs the one for \gls{tms} and the one for \gls{sms} in a lockstep.
-The trace agreement similarly just relates a specification event with an abstract \gls{tms}-event $\abstrace[_{\tmssafe}]$ and with an abstract \gls{sms}-event $\abstrace[_{\smssafe}]$.
+The combined monitor runs the one for \gls*{tms} and the one for \gls*{sms} in a lockstep.
+The trace agreement similarly just relates a specification event with an abstract \gls*{tms}-event $\abstrace[_{\tmssafe}]$ and with an abstract \gls*{sms}-event $\abstrace[_{\smssafe}]$.
 % \begin{center}
-%   \judgbox{\mstraceagree{\event}{\absevent}}{,,Abstract event $\absevent$ is equivalent to $\event$ with respect to \gls{ms}.''}$\;$\\
+%   \judgbox{\mstraceagree{\event}{\absevent}}{,,Abstract event $\absevent$ is equivalent to $\event$ with respect to \gls*{ms}.''}$\;$\\
 %   %
@@
 
-\subsubsection{Monitor for \glsfirst{scct}}
+\subsubsection{Monitor for \glsfirst*{scct}}
 \begin{gather*}
@@
 \end{center}
-The monitor state for the \gls{scct} monitor is completely empty, since it does not need to keep track of information.
+The monitor state for the \gls*{scct} monitor is completely empty, since it does not need to keep track of information.
 As soon as any event is hit, the execution gets stuck, since any event is considered confidental from the perspective of this monitor.
 \begin{center}
-  \judgbox{\sccttraceagree[]{\event}{\absevent}}{,,Abstract event $\absevent$ is equivalent to $\event$ with respect to \gls{cct}.''}$\;$\\
+  \judgbox{\sccttraceagree[]{\event}{\absevent}}{,,Abstract event $\absevent$ is equivalent to $\event$ with respect to \gls*{cct}.''}$\;$\\
   %
@@
 
-\subsubsection{Combining \gls{ms} and \gls{scct} Monitors to obtain \gls{msscct}}
-
-The combination of monitors for \gls{ms} and \gls{scct} yields one for \gls{msscct}.
-The construction is entirely similar to the one for \gls{ms} (\Cref{subsubsec:ms}).
+\subsubsection{Combining \gls*{ms} and \gls*{scct} Monitors to obtain \gls*{msscct}}
+
+The combination of monitors for \gls*{ms} and \gls*{scct} yields one for \gls*{msscct}.
+The construction is entirely similar to the one for \gls*{ms} (\Cref{subsubsec:ms}).
 
@@
 First, $\src{TypeScript}$ programs are translated to $\trg{JavaScript}$ which, e.g., V8~\cite{googlev8} eventually compiles in parts to $\irl{Ignition BC}$.
-The following theorem establishes what happens if all these compilation steps were robustly secure with respect to \gls{ms}:
-The resulting $\irl{Ignition BC}$ code would be \gls{ms} regardless of the context the binary runs in.
+The following theorem establishes what happens if all these compilation steps were robustly secure with respect to \gls*{ms}:
+The resulting $\irl{Ignition BC}$ code would be \gls*{ms} regardless of the context the binary runs in.
 
@@
 }
-It is worth pointing out that \gls{ms} (\Cref{subsubsec:ms}) is an example of a practically relevant intersection of classes, namely, \gls{tms} and \gls{sms} (\Cref{subsubsec:tms,subsubsec:sms}).
+It is worth pointing out that \gls*{ms} (\Cref{subsubsec:ms}) is an example of a practically relevant intersection of classes, namely, \gls*{tms} and \gls*{sms} (\Cref{subsubsec:tms,subsubsec:sms}).
 However, the properties need to be lifted~\cite{clarkson2008hyper} to sets of properties by applying the powerset operation.
@@
 % \MP{what's the point of the previous 3 sentences?(commented)}
-Consider the following two LLVM optimisation passes: \gls{cf}, which rewrites constant expressions to the constant they evaluate to, and \gls{dce}, which removes dead code by rewriting conditional branches.
-The order in which \gls{cf} and \gls{dce} are performed influences the final result of the compilation (see \Cref{fig:cfdceex}).
+Consider the following two LLVM optimisation passes: \gls*{cf}, which rewrites constant expressions to the constant they evaluate to, and \gls*{dce}, which removes dead code by rewriting conditional branches.
+The order in which \gls*{cf} and \gls*{dce} are performed influences the final result of the compilation (see \Cref{fig:cfdceex}).
 \begin{figure}[!ht]
@@
       % edges
-      \draw[->,very thick] (progunopt) edge[loop left,left] node {\gls{dce}} (progunopt);
-      \draw[->,very thick] (progunopt) edge[sloped,above] node {\gls{cf}} (progoptcf);
-      \draw[->,very thick] (progoptcf) edge[sloped,above] node {\gls{dce}} (progoptcfdce);
+      \draw[->,very thick] (progunopt) edge[loop left,left] node {\gls*{dce}} (progunopt);
+      \draw[->,very thick] (progunopt) edge[sloped,above] node {\gls*{cf}} (progoptcf);
+      \draw[->,very thick] (progoptcf) edge[sloped,above] node {\gls*{dce}} (progoptcfdce);
     \end{tikzpicture}
   \end{center}
-  \caption{Example program where the level of optimisations differ for one pass of applying \gls{cf} and \gls{dce} in any order. %
-  Every edge is a compilation pass and the label on the edge states what the pass does, i.e., \gls{cf} or \gls{dce}. %
+  \caption{Example program where the level of optimisations differ for one pass of applying \gls*{cf} and \gls*{dce} in any order. %
+  Every edge is a compilation pass and the label on the edge states what the pass does, i.e., \gls*{cf} or \gls*{dce}. %
   The source code in the nodes is a glorified compiler intermediate representation and the code gets more optimised towards the right hand side of the figure.}\label{fig:cfdceex}
@@
 To this end, this section defines the languages $\Ltms$, $\Ltrg$, $\Lms$, and $\Lscct$ which share many common elements (presented in \Cref{subsec:cs:defs}).
-$\Ltms$ is the only statically typed language and exhibits the property that all well-typed programs are \gls{tms} (\Cref{subsec:ltms}).
-However, not all $\Ltms$ programs are \gls{sms}.
+$\Ltms$ is the only statically typed language and exhibits the property that all well-typed programs are \gls*{tms} (\Cref{subsec:ltms}).
+However, not all $\Ltms$ programs are \gls*{sms}.
 That is, there are well-typed $\Ltms$ programs that perform an out-of-bounds access.
-Language $\Ltrg$ is untyped and does not provide any guarantees with regards to \gls{ms} (\Cref{subsec:lsms}).
+Language $\Ltrg$ is untyped and does not provide any guarantees with regards to \gls*{ms} (\Cref{subsec:lsms}).
 $\Lms$ is exactly the same language as $\Ltrg$, but this paper still distinguishes the two for sake of readability (\Cref{subsec:lms}).
-All three languages --- so $\Ltms$, $\Ltrg$, and $\Lms$ --- assume \gls{cct} to hold.
-
-This paper considers \gls{cct} as an architecture-dependent property\MPin{can we justify this}, so writing \gls{cct} should not be one of the programmer's concerns~\cite{cauligi2019fact}.
+All three languages --- so $\Ltms$, $\Ltrg$, and $\Lms$ --- assume \gls*{cct} to hold.
+
+This paper considers \gls*{cct} as an architecture-dependent property\MPin{can we justify this}, so writing \gls*{cct} should not be one of the programmer's concerns~\cite{cauligi2019fact}.
 Such consideration is backed up by architecture providing a data (operand) independent timing mode, such as processors by Arm~\cite[p.~543]{arm-refman} and Intel~\cite[p.~80]{intel-refman}.
-In spirit of this, language $\Lscct$ allows violating \gls{cct} by emitting events on, e.g., branching and division, that contain secrets (\Cref{subsec:lscct}), but provides a way to read and write to a {\em model--specific register} that enables a ``\gls{cct}-mode''.
+In spirit of this, language $\Lscct$ allows violating \gls*{cct} by emitting events on, e.g., branching and division, that contain secrets (\Cref{subsec:lscct}), but provides a way to read and write to a {\em model--specific register} that enables a ``\gls*{cct}-mode''.
 
@@
 But, $\src{L_{\tmssafe}}$ is statically typed, where the type system is inspired by $L^{3}$~\cite{morrisett2005L3,scherer2018fabulous}.
-The type system of $\src{L_{\tmssafe}}$ exhibits the property that every well-typed $\src{L_{\tmssafe}}$ program satisfies \gls{tms} (\Cref{thm:wt:tms}).
+The type system of $\src{L_{\tmssafe}}$ exhibits the property that every well-typed $\src{L_{\tmssafe}}$ program satisfies \gls*{tms} (\Cref{thm:wt:tms}).
 The proof of this theorem relies on a projection $\msfilterLtms{\src{\event}}=\event$ from $\src{L_{\tmssafe}}$ events to specification events $\event$, because the properties defined earlier (\Cref{subsec:propdefs}) are defined in the specification trace model.
@@
 Most rules of the projection $\msfilterLtms{\src{\event}}$ are left out since, for the most part, it does the expected, e.g., $\msfilterLtms{\src{\ev{Dealloc}\ \loc;\comp;\securitytag}}=\ev{Dealloc}\ \delta(\loc);\comp;\securitytag$.
-But, it also filters any action that a context does as well as the interaction events, since these are irrelevant for component-level \gls{tms}.
-
-\begin{theorem}[$\src{L_{\tmssafe}}$-programs are \gls{tms}]\label{thm:wt:tms}
+But, it also filters any action that a context does as well as the interaction events, since these are irrelevant for component-level \gls*{tms}.
+
+\begin{theorem}[$\src{L_{\tmssafe}}$-programs are \gls*{tms}]\label{thm:wt:tms}
   For any $\src{\library_{\comp}}$, $\rsat{\src{\library_{\comp}}}{\tmssafe}$ \Coqed
@@
 With this technical setup, the information whether a read or write happened on a secret value is not hidden by the semantics, e.g., by emitting $\obj{\emptyevent}$, but when projecting to specification events.
-This allows flexibility: The trace can be checked to satisfy different properties, such as, in this case, \gls{tms}, \gls{sms}, \gls{scct}, and their combined versions.
+This allows flexibility: The trace can be checked to satisfy different properties, such as, in this case, \gls*{tms}, \gls*{sms}, \gls*{scct}, and their combined versions.
 \Cref{ex:lscct} illustrates the differences of $\Lscct$ compared to the other languages.
@@
 
-      \draw[->] (S) to[sloped] node[align=center,font=\scriptsize] (tmsedge) {\gls{tms}\\ \Cref{thm:cca:rtp:tms}} (T);
-      \draw[->] (T) to[sloped] node[align=center,font=\scriptsize] {\gls{sms}\\ \Cref{thm:ccb:rtp:sms}} (M);
-      \draw[->] (M) to[sloped] node[align=center,font=\scriptsize] {\gls{dce}\\ \Cref{thm:ccdce:rtp:ms}} (D0);
-      \draw[->] (M) to[sloped] node[align=center,font=\scriptsize] {\gls{cf}\\ \Cref{thm:cccf:rtp:ms}} (C0);
-      \draw[->] (D0) to[sloped] node[align=center,font=\scriptsize] {\gls{cf}\\ \Cref{thm:cccf:rtp:ms}} (E);
-      \draw[->] (C0) to[sloped] node[align=center,font=\scriptsize] {\gls{dce}\\ \Cref{thm:ccdce:rtp:ms}} (E);
-      \draw[->] (E) to[sloped,above] node[align=center,font=\scriptsize] {\gls{scct}\\ \Cref{thm:ccscct:rtp:scct}} (O);
+      \draw[->] (S) to[sloped] node[align=center,font=\scriptsize] (tmsedge) {\gls*{tms}\\ \Cref{thm:cca:rtp:tms}} (T);
+      \draw[->] (T) to[sloped] node[align=center,font=\scriptsize] {\gls*{sms}\\ \Cref{thm:ccb:rtp:sms}} (M);
+      \draw[->] (M) to[sloped] node[align=center,font=\scriptsize] {\gls*{dce}\\ \Cref{thm:ccdce:rtp:ms}} (D0);
+      \draw[->] (M) to[sloped] node[align=center,font=\scriptsize] {\gls*{cf}\\ \Cref{thm:cccf:rtp:ms}} (C0);
+      \draw[->] (D0) to[sloped] node[align=center,font=\scriptsize] {\gls*{cf}\\ \Cref{thm:cccf:rtp:ms}} (E);
+      \draw[->] (C0) to[sloped] node[align=center,font=\scriptsize] {\gls*{dce}\\ \Cref{thm:ccdce:rtp:ms}} (E);
+      \draw[->] (E) to[sloped,above] node[align=center,font=\scriptsize] {\gls*{scct}\\ \Cref{thm:ccscct:rtp:scct}} (O);
 
@@
 
-      \draw[thick,dotted,->] (S) to[bend right=33,sloped] node[align=center,font=\scriptsize] {\gls{ms}\\ \Cref{thm:ccab:rtp:ms}} (M);
-      \draw[thick,dotted,->] (M) to[bend right=0,sloped] node[align=center,font=\scriptsize] {\gls{ms}\\ \Cref{thm:cccfccdce:rtp:ms}} (E);
-      \draw[thick,dotted,->] (S) to[out=-45,in=198,sloped] node[align=center,font=\scriptsize] {\gls{ms}+\gls{scct}\\ \Cref{thm:ccall:rtp:msscct}} (O);
+      \draw[thick,dotted,->] (S) to[bend right=33,sloped] node[align=center,font=\scriptsize] {\gls*{ms}\\ \Cref{thm:ccab:rtp:ms}} (M);
+      \draw[thick,dotted,->] (M) to[bend right=0,sloped] node[align=center,font=\scriptsize] {\gls*{ms}\\ \Cref{thm:cccfccdce:rtp:ms}} (E);
+      \draw[thick,dotted,->] (S) to[out=-45,in=198,sloped] node[align=center,font=\scriptsize] {\gls*{ms}+\gls*{scct}\\ \Cref{thm:ccall:rtp:msscct}} (O);
 
@@
   \end{center}
-  \caption{Visualisation of the optimising compilation pipeline that attains a combination of \gls{ms} and \gls{cct}. %
+  \caption{Visualisation of the optimising compilation pipeline that attains a combination of \gls*{ms} and \gls*{cct}. %
     Vertices in the graph are the programming languages from earlier sections (\Cref{sec:casestud:defs}). %
@@
 \end{figure}
-The section demonstrates the power of the framework (\Cref{sec:compprop,sec:compcomp}) by composing these compilers for a secure and optimising compilation chain that robustly preserves \gls{msscct}.
-The first step in this chain is the compiler from $\src{L_{\tmssafe}}$ to $\trg{L}$ that robustly preserves just \gls{tms} (\Cref{thm:cca:rtp:tms}).
-From here, an instrumentation from $\trg{L}$ to $\irl{L_{\mssafe}}$ ensures that no out-of-bounds accesses can happen and, thus, programs at this point attain \gls{sms} (\Cref{thm:ccb:rtp:sms}).
-Since these properties compose into \gls{ms}, composing these passes yields a compiler that robustly preserves \gls{ms} (\Cref{thm:ccab:rtp:ms}).
-At this stage, the section presents two optimising translations, namely \gls{cf} and \gls{dce}, each of which robustly preserves \gls{ms} (\Cref{thm:ccdce:rtp:ms,thm:cccf:rtp:ms}).
+The section demonstrates the power of the framework (\Cref{sec:compprop,sec:compcomp}) by composing these compilers for a secure and optimising compilation chain that robustly preserves \gls*{msscct}.
+The first step in this chain is the compiler from $\src{L_{\tmssafe}}$ to $\trg{L}$ that robustly preserves just \gls*{tms} (\Cref{thm:cca:rtp:tms}).
+From here, an instrumentation from $\trg{L}$ to $\irl{L_{\mssafe}}$ ensures that no out-of-bounds accesses can happen and, thus, programs at this point attain \gls*{sms} (\Cref{thm:ccb:rtp:sms}).
+Since these properties compose into \gls*{ms}, composing these passes yields a compiler that robustly preserves \gls*{ms} (\Cref{thm:ccab:rtp:ms}).
+At this stage, the section presents two optimising translations, namely \gls*{cf} and \gls*{dce}, each of which robustly preserves \gls*{ms} (\Cref{thm:ccdce:rtp:ms,thm:cccf:rtp:ms}).
 These translations can be freely ordered in the compilation chain without compromising memory safety (\Cref{thm:cccfccdce:rtp:ms}).
-The last step of the chain ensures that code stays \gls{scct} (\Cref{thm:ccscct:rtp:scct}) when lowered from $\Lms$ to $\Lscct$.
-The final result is that the whole compilation chain robustly preserves \gls{msscct} (\Cref{thm:ccall:rtp:msscct}).
+The last step of the chain ensures that code stays \gls*{scct} (\Cref{thm:ccscct:rtp:scct}) when lowered from $\Lms$ to $\Lscct$.
+The final result is that the whole compilation chain robustly preserves \gls*{msscct} (\Cref{thm:ccall:rtp:msscct}).
 
@@
 
-\begin{theorem}[Compiler $\cca$ is secure with respect to \gls{tms}]\label{thm:cca:rtp:tms}
+\begin{theorem}[Compiler $\cca$ is secure with respect to \gls*{tms}]\label{thm:cca:rtp:tms}
   $\rtp{\cca}{\tmssafe}$ % \Coqed
@@
 
-The compiler $\cc{\trg{L}}{\Lms}$ only inserts bounds-checks whenever reading from or writing to memory in order to enforce \gls{sms}.
+The compiler $\cc{\trg{L}}{\Lms}$ only inserts bounds-checks whenever reading from or writing to memory in order to enforce \gls*{sms}.
 For passing pointers, it has to pass them with their size information as well.
@@
 \begin{example}[Instrumented \texttt{strncpy}]
-  Consider again \texttt{strncpy}, but instrumented for \gls{sms}:
+  Consider again \texttt{strncpy}, but instrumented for \gls*{sms}:
     \begin{lstlisting}[language=c,basicstyle=\ttfamily\small, morekeywords={size_t}]
@@
 
-\begin{theorem}[Compiler $\ccb$ is secure with respect to \gls{sms}]\label{thm:ccb:rtp:sms}
+\begin{theorem}[Compiler $\ccb$ is secure with respect to \gls*{sms}]\label{thm:ccb:rtp:sms}
   $\rtp{\ccb}{\smssafe}$ % \Coqed
@@
 
-\Cref{thm:ccab:rtp:ms} states that the composition of $\cca$ and $\ccb$ is secure with respect to \gls{ms} and follows from \Cref{thm:cca:rtp:tms,thm:ccb:rtp:sms} using \Cref{thm:rtp}.
-
-\begin{theorem}[Compiler $\cca\circ\ccb$ is secure with respect to \gls{ms}]\label{thm:ccab:rtp:ms}
+\Cref{thm:ccab:rtp:ms} states that the composition of $\cca$ and $\ccb$ is secure with respect to \gls*{ms} and follows from \Cref{thm:cca:rtp:tms,thm:ccb:rtp:sms} using \Cref{thm:rtp}.
+
+\begin{theorem}[Compiler $\cca\circ\ccb$ is secure with respect to \gls*{ms}]\label{thm:ccab:rtp:ms}
   $\rtp{\cca\circ\ccb}{\mssafe}$ % \Coqed
@@
 
-The two optimising compiler passes from $\Lms$ to $\Lms$ perform \gls{dce} and \gls{cf}, respectively.
-The \gls{dce} pass applies a naive rewrite rule on conditionals.
-For \gls{cf}, the pass uses an auxiliary function \texttt{mix} that does the actual work.
+The two optimising compiler passes from $\Lms$ to $\Lms$ perform \gls*{dce} and \gls*{cf}, respectively.
+The \gls*{dce} pass applies a naive rewrite rule on conditionals.
+For \gls*{cf}, the pass uses an auxiliary function \texttt{mix} that does the actual work.
 It rewrites constant binary operations, e.g., $\irl{{17}+{29}-1}$ to $\irl{46-1}$, and replaces variables that are assigned to constants with their constant, e.g., $\irl{\llet{x}{7}{x}}$ to $\irl{7}$.
-Both passes are secure with respect to \gls{ms}.
-The proof for either is relatively simple, because both \gls{dce} and \gls{cf} do not change the way memory accesses happen.
+Both passes are secure with respect to \gls*{ms}.
+The proof for either is relatively simple, because both \gls*{dce} and \gls*{cf} do not change the way memory accesses happen.
 Moreover, since the input and output languages to these compilers are the same, attacker contexts do not have more power in the target language than in the source.
 
-\begin{theorem}[Compiler $\ccdce$ is secure with respect to \gls{ms}]\label{thm:ccdce:rtp:ms}
+\begin{theorem}[Compiler $\ccdce$ is secure with respect to \gls*{ms}]\label{thm:ccdce:rtp:ms}
   $\rtp{\ccdce}{\mssafe}$ %\Coqed
 \end{theorem}
-\begin{theorem}[Compiler $\cccf$ is secure with respect to \gls{ms}]\label{thm:cccf:rtp:ms}
+\begin{theorem}[Compiler $\cccf$ is secure with respect to \gls*{ms}]\label{thm:cccf:rtp:ms}
   $\rtp{\cccf}{\mssafe}$ %\Coqe
@@
 
-\begin{theorem}[Compilers $\cccf\circ\ccdce$ and $\cccf\circ\ccdce$ are secure with respect to \gls{ms}]\label{thm:cccfccdce:rtp:ms}
+\begin{theorem}[Compilers $\cccf\circ\ccdce$ and $\cccf\circ\ccdce$ are secure with respect to \gls*{ms}]\label{thm:cccfccdce:rtp:ms}
   $\rtp{\cccf\circ\ccdce}{\mssafe}$ and $\rtp{\ccdce\circ\cccf}{\mssafe}$. % \Coqed
@@
 
-Given the fact that $\Lscct$ provides a \gls{cct}-mode that can be turned on or off, the compiler inserts wrapper code for function bodies to ensure that execution in the component always happen in this \gls{cct}-mode.
+Given the fact that $\Lscct$ provides a \gls*{cct}-mode that can be turned on or off, the compiler inserts wrapper code for function bodies to ensure that execution in the component always happen in this \gls*{cct}-mode.
 The context can overwrite the flag and exit the mode, but upon invoking a function that is part of the component, the flag would be set again.
-Because of this, the compiler is secure with respect to \gls{scct}, similarly proven as in \Cref{subsec:cs:tms}.
-
-\begin{theorem}[Compiler $\ccscct$ is secure with respect to \gls{scct}]\label{thm:ccscct:rtp:scct}
+Because of this, the compiler is secure with respect to \gls*{scct}, similarly proven as in \Cref{subsec:cs:tms}.
+
+\begin{theorem}[Compiler $\ccscct$ is secure with respect to \gls*{scct}]\label{thm:ccscct:rtp:scct}
   $\rtp{\ccscct}{\scctsafe}$ % \Coqed
@@
 
-\begin{theorem}[Compiler $\ccmsscct$ is secure with respect to \gls{scct}]\label{thm:ccall:rtp:msscct}
+\begin{theorem}[Compiler $\ccmsscct$ is secure with respect to \gls*{scct}]\label{thm:ccall:rtp:msscct}
   $\rtp{\cc{\Ltms}{\Lscct}}{\mssafe\cap\scctsafe}$ \Coqed
@@
 % What is left to show is $\scctfilterLscct[\emptyset]{\obj{\trace}}\in\pi$.
-% Let $\irl{\progvar}=\cc{\Ltms}{\Ltrg}\circ\cc{\Ltrg}{\Lms}\circ\cc[\gls{cf}]{\Lms}{\Lms}\circ\cc[\gls{dce}]{\Lms}{\Lms}(\src{\progvar})$ such that $\cc{\Lms}{\Lscct}(\irl{\progvar})=\cc{\Ltms}{\Lscct}(\src{\progvar})$.
+% Let $\irl{\progvar}=\cc{\Ltms}{\Ltrg}\circ\cc{\Ltrg}{\Lms}\circ\cc[\gls*{cf}]{\Lms}{\Lms}\circ\cc[\gls*{dce}]{\Lms}{\Lms}(\src{\progvar})$ such that $\cc{\Lms}{\Lscct}(\irl{\progvar})=\cc{\Ltms}{\Lscct}(\src{\progvar})$.
 % Note that $\pi\in\scctsafe$, so apply \Thmref{thm:ccscct:rtp:scct}
@@
 This section discusses work on robust compilation (\Cref{subsec:relw:seccomprtp}) and on other secure compilation criteria (\Cref{subsec:relw:seccompcrit}).
-Since the case study of \Cref{sec:casestud:defs,sec:casestud:rtp} implements measures for preserving \gls{ms} and \gls{cct}, this section then presents relevant related work as well (\Cref{subsec:relw:msmechs,subsec:relw:cctmechs}).
+Since the case study of \Cref{sec:casestud:defs,sec:casestud:rtp} implements measures for preserving \gls*{ms} and \gls*{cct}, this section then presents relevant related work as well (\Cref{subsec:relw:msmechs,subsec:relw:cctmechs}).
 
@@
 CT-Wasm~\cite{watt2019ctwasm} enforces constant-timeness by means of a type system.
-Different to the approach of this paper, these approaches necessitate that the programmer writes \gls{cct} code.
+Different to the approach of this paper, these approaches necessitate that the programmer writes \gls*{cct} code.
 An approach to allow programmers to write more high-level code is CryptOpt~\cite{kuepper2023cryptopt}, which generates efficient target-code by means of a randomised search.
@@
 % 
-Finally, this paper defines a multi-pass compiler and proves that it preserves \gls{msscct}.
+Finally, this paper defines a multi-pass compiler and proves that it preserves \gls*{msscct}.
 Crucially, this paper derives the security of the multi-pass compiler from the composition of the security properties preserved by its individual passes, which include security-preserving as well as optimisation passes.
@@
 % This paper does a first step towards practical secure compilation chains by introducing a theoretical framework demonstrating that secure compilers compose.
-% The paper exercises this framework on a case-study that consists of an optimising compilation chain that is secure with respect to \gls{msscct}.
+% The paper exercises this framework on a case-study that consists of an optimising compilation chain that is secure with respect to \gls*{msscct}.
 % \MP{

### 2023-07-12 popl/main.tex
cits

@@
 Thus, from a security perspective, this considers only a passive attacker model, where the attacker observes the execution and, e.g., retrieves secrets from that.
-To consider a stronger model similarly to what existing work does~\cite{abate2019jour,abate2021extacc,maffeis2008code-carrying,gordon2003authenticity,fournet2007authorization,bengtson2011refine,backes2014uniontyps}, the concept of satisfaction can be extended with {\em robustness}.
+To consider a stronger model similarly to what existing work does~\cite{abate2019jour,abate2021extacc,maffeis2008code-carrying,gordon2003authenticity,fournet2007authorization,bengtson2011refine,backes2014uniontyps,michael2023mswasm,swasey2017robust}, the concept of satisfaction can be extended with {\em robustness}.
 \MPpost{
-	citations here should include mswasm, swasey's stuff and michel's and move
+	cit michel's and move
 }

### 2023-07-12 popl/main.tex
strncpy for 3.13

@@
   The vulnerability arises due to a race condition where the context-level function \texttt{hci\_cmd\_work} was not expected to behave maliciously, since it resides in the same source-code repository where the vulnerability occurs.
-  Nevertheless, the component-level code of \texttt{hci\_dev\_do\_open} is linked with \texttt{hci\_cmd\_work} and does not atomically check whether the pointer\MP{what pointer?} has been freed already:
-  Therefore, \texttt{hci\_dev\_do\_open} does not satisfy the no-double-frees property robustly, since there is an implementation for \texttt{hci\_cmd\_work} that results the whole program to violate that property.
+  Nevertheless, the component-level code of \texttt{hci\_dev\_do\_open} is linked with \texttt{hci\_cmd\_work} and does not atomically check whether a pointer has been freed already:
+  Therefore, \texttt{hci\_dev\_do\_open} does not satisfy the no-double-frees property robustly, since there is an implementation for \texttt{hci\_cmd\_work} that leads to a violation of that property when linked with \texttt{hci\_dev\_do\_open}.
 \end{example}
@@
 \begin{example}[Data-independent timing mode]
-  Consider an excerpt from a component that reads from non-confidential memory \texttt{x} and writes the result to confidential memory \texttt{y}.
-  \MPin{
-  can't we call strncpy but with src == low and dest == high?
-  }
-  The trace of just this part may look like $\ev{Use}\ \loc_{x}\ 1729;\comp;\unlock\cdot\ev{Use}\ \loc_{y}\ 1729;\comp;\lock$.
+  Consider the call \texttt{strncpy(1,x,y)} to the \texttt{strncpy} function (\Cref{ex:strncpy}) with low ($\unlock$) security for \texttt{x} and high security ($\lock$) for \texttt{y}.
+  The trace of just the copying part inside of \texttt{strncpy} looks like $\ev{Use}\ \loc_{x}\ 0;\comp;\unlock\cdot\ev{Use}\ \loc_{y}\ 0;\comp;\lock$.
   In terms of the abstract monitor trace, this is just $\absev{Any}$.
   Running this on the monitor would result in getting stuck, since there is no matching rule to step in the presence of $\absev{Any}$ event. % pun intended
-  By means of additional features to ensure cryptographic constant time even in the presence of memory reads and loads, such as a flag to enable a data independent timing mode, which is present in both Arm~\cite[p.~543]{arm-refman} and Intel~\cite[p.~80]{intel-refman} processors, the original trace can contain $\ev{Use}\ \loc_{y}\ 1729;\comp;\unlock$ instead of $\ev{Use}\ \loc_{y}\ 1729;\comp;\lock$.
+  By means of additional features to ensure cryptographic constant time even in the presence of memory reads and loads, such as a flag to enable a data independent timing mode, which is present in both Arm~\cite[p.~543]{arm-refman} and Intel~\cite[p.~80]{intel-refman} processors, the original trace now does $\ev{Use}\ \loc_{y}\ 0;\comp;\unlock$ instead of $\ev{Use}\ \loc_{y}\ 0;\comp;\lock$.
   Because of this, the whole trace of the component equates to $\absev{\emptyevent}$ and the monitor can step without getting stuck.

### 2023-07-12 popl/main.tex
colorboxes in fig

@@
 Since the composition of secure compilers is again a secure compiler, the theorem generalises to a whole chain of $n$ secure compilers.
-\MPin{
-	what do the sentences below mean in this paragraph?
-}
-It is worth pointing out that \gls*{ms} (\Cref{subsubsec:ms}) is an example of a practically relevant intersection of classes, namely, \gls*{tms} and \gls*{sms} (\Cref{subsubsec:tms,subsubsec:sms}).
-However, the properties need to be lifted~\cite{clarkson2008hyper} to sets of properties by applying the powerset operation.
 
@@
 As for pre-events, the memory-related ones are allocation ($\ev{Alloc}\ \loc\ \valueexpr$), deallocation ($\ev{Dealloc}\ \loc$), reading from ($\ev{Get}\ \loc\ \valueexpr$) and writing to memory ($\ev{Set}\ \loc\ \valueexpr\ \valueexpr[']$).
-
-\MPpost{
-	this is a bit abrupt
-}
+The following is an excerpt of the operational semantics handling some of the memory operations.
 
@@
 For calls without this context switch, the environmental semantics attaches the $\nocomm$ tag.
-In the following, $\neg\ctx=\comp$ and $\neg\comp=\ctx$.\MP{i don't think this is used?}
+In the following, $\neg\ctx=\comp$ and $\neg\comp=\ctx$.
 
@@
 %
-The projection $\msfilterL{\trg{\trace}}$ is entirely similar to $\msfilterLtms{\src{\trace}}$.
-\MPin{
-	entirely similar? no. equal
-}
+The projection $\msfilterL{\trg{\trace}}$ is equal to $\msfilterLtms{\src{\trace}}$.
 
@@
 Moreover, the language extends pre-events with $\obj{\ev{Branch}\ n}$ and $\obj{\ev{Binop}\ n}$ that are emitted when evaluating a branch or certain binary expressions, such as division, respectively, whenever the data independent timing mode is inactive.
-\MPin{rules are not explained}
+The following rules demonstrate how this is handled semantically.
 
@@
 
-The evaluation steps are amended to propagate the securitytags $\obj{\securitytag}$.
+The evaluation steps are amended to propagate the security-tag annotations $\obj{\securitytag}$.
 When the data independent timing mode is active, pre-events $\obj{\ev{Branch}\ n}$ and $\obj{\ev{Binop}\ n}$ are emitted for conditionals and binary operations, respectively.
-To mark this mode as active or inactive, \Cref{tr:e-wrdoit} allows to access the register in $\obj{\statevar}$.
+%To mark this mode as active or inactive, \Cref{tr:e-wrdoit} allows to access the register in $\obj{\statevar}$.
 
@@
   Consider again the context presented in \Cref{ex:strncpy:cpp11}, where everything is marked with a security tag of high $\lock$.
-  The following table shows parts of the execution trace, read from top to bottom, in the left column with and in the right column without data independent timing.
-  The left side of the table, i.e., the two columns on the left, describe the execution trace of the program, while the right side of the table, i.e., the two columns on the right, describe the respective projections $\scctfilterLscct{\obj{\trace}}$ to the specification trace model.
-  \MPin{
-  very nice.
-  final bit, perhaps highlight those events taht are mentioned in the text, so they are easily recognisable?
-  like yellow BG for the events, gray BG for the locks 
-  }
+  The following table shows parts of the execution trace, read from top to bottom, in the left column (\textit{Active}) with and in the right column (\textit{Inactive}) without data independent timing.
+  The left side of the table ($\Lscct$), i.e., the two columns on the left, describes the execution trace of the program, while the right side of the table ($Spec$), i.e., the two columns on the right, describes the respective projections $\scctfilterLscct{\obj{\trace}}$ to the specification trace model.
 
@@
     \obj{\ev{Alloc}\ \loc_{y}\ 12;\comp;\unlock} & \mid & \obj{\ev{Alloc}\ \loc_{y}\ 12;\comp;\unlock} & \mid & \ev{Alloc}\ \loc_{y}\ 12;\comp;\unlock & \mid & \ev{Alloc}\ \loc_{x}\ 12;\comp;\unlock\\
-    \obj{\widehat{\ev{Get}}\ \loc_{x}\ 0;\comp;\lock} & \mid & \obj{\ev{Get}\ \loc_{x}\ 0;\comp;\lock} & \mid & \ev{Use}\ \loc_{x}\ 0;\comp;\unlock & \mid & \ev{Use}\ \loc_{x}\ 0;\comp;\lock\\
-    \obj{\emptyevent} & \mid & \obj{\ev{Branch}\ 0;\comp;\lock} & \mid & \emptyevent & \mid & \ev{Branch}\ 0;\comp;\lock\\
-    \obj{\widehat{\ev{Get}}\ \loc_{x}\ 0;\comp;\lock} & \mid & \obj{\ev{Get}\ \loc_{x}\ 0;\comp;\lock} & \mid & \ev{Use}\ \loc_{x}\ 0;\comp;\lock & \mid & \ev{Use}\ \loc_{x}\ 0;\comp;\lock\\
-    \obj{\widehat{\ev{Set}}\ \loc_{y}\ 0\ \mathtt{'H'};\comp;\lock} & \mid & \obj{\ev{Set}\ \loc_{y}\ 0\ \mathtt{'H'};\comp;\lock} & \mid & \ev{Use}\ \loc_{y}\ 0;\comp;\unlock & \mid & \ev{Use}\ \loc_{y}\ 0;\comp;\lock\\
-    \obj{\widehat{\ev{Get}}\ \loc_{x}\ 1;\comp;\lock} & \mid & \obj{\ev{Get}\ \loc_{x}\ 1;\comp;\lock} & \mid & \ev{Use}\ \loc_{x}\ 1;\comp;\unlock & \mid & \ev{Use}\ \loc_{x}\ 1;\comp;\lock\\
-    \obj{\emptyevent} & \mid & \obj{\ev{Branch}\ 0;\comp;\lock} & \mid & \emptyevent & \mid & \ev{Branch}\ 0;\comp;\lock\\
+    \obj{\widehat{\ev{Get}}\ \loc_{x}\ 0;\comp;}\text{\colorbox{lightgray}{$\obj{\lock}$}} & \mid & \obj{\ev{Get}\ \loc_{x}\ 0;\comp;\lock} & \mid & \ev{Use}\ \loc_{x}\ 0;\comp;\text{\colorbox{lightgray}{$\unlock$}} & \mid & \ev{Use}\ \loc_{x}\ 0;\comp;\lock\\
+    \text{\colorbox{yellow}{$\obj{\emptyevent}$}} & \mid & \obj{\ev{Branch}\ 0;\comp;\lock} & \mid & \emptyevent & \mid & \ev{Branch}\ 0;\comp;\lock\\
+    \obj{\widehat{\ev{Get}}\ \loc_{x}\ 0;\comp;}\text{\colorbox{lightgray}{$\obj{\lock}$}} & \mid & \obj{\ev{Get}\ \loc_{x}\ 0;\comp;\lock} & \mid & \ev{Use}\ \loc_{x}\ 0;\comp;\text{\colorbox{lightgray}{$\unlock$}} & \mid & \ev{Use}\ \loc_{x}\ 0;\comp;\lock\\
+    \obj{\widehat{\ev{Set}}\ \loc_{y}\ 0\ \mathtt{'H'};\comp;}\text{\colorbox{lightgray}{$\obj{\lock}$}} & \mid & \obj{\ev{Set}\ \loc_{y}\ 0\ \mathtt{'H'};\comp;\lock} & \mid & \ev{Use}\ \loc_{y}\ 0;\comp;\text{\colorbox{lightgray}{$\unlock$}} & \mid & \ev{Use}\ \loc_{y}\ 0;\comp;\lock\\
+    \obj{\widehat{\ev{Get}}\ \loc_{x}\ 1;\comp;}\text{\colorbox{lightgray}{$\obj{\lock}$}} & \mid & \obj{\ev{Get}\ \loc_{x}\ 1;\comp;\lock} & \mid & \ev{Use}\ \loc_{x}\ 1;\comp;\text{\colorbox{lightgray}{$\unlock$}} & \mid & \ev{Use}\ \loc_{x}\ 1;\comp;\lock\\
+    \text{\colorbox{yellow}{$\obj{\emptyevent}$}} & \mid & \obj{\ev{Branch}\ 0;\comp;\lock} & \mid & \emptyevent & \mid & \ev{Branch}\ 0;\comp;\lock\\
     \vdots & \mid & \vdots & \mid & \vdots & \mid & \vdots \\
-    \obj{\widehat{\ev{Get}}\ \loc_{x}\ 12;\comp;\lock} & \mid & \obj{\ev{Get}\ \loc_{x}\ 12;\comp;\lock} & \mid &  \ev{Use}\ \loc_{x}\ 12;\comp;\unlock & \mid & \ev{Use}\ \loc_{x}\ 12;\comp;\lock\\
-    \obj{\emptyevent} & \mid & \obj{\ev{Branch}\ 1;\comp;\lock} & \mid &  \emptyevent & \mid & \ev{Branch}\ 1;\comp;\lock\\
+    \obj{\widehat{\ev{Get}}\ \loc_{x}\ 12;\comp;}\text{\colorbox{lightgray}{$\obj{\lock}$}} & \mid & \obj{\ev{Get}\ \loc_{x}\ 12;\comp;\lock} & \mid &  \ev{Use}\ \loc_{x}\ 12;\comp;\text{\colorbox{lightgray}{$\unlock$}} & \mid & \ev{Use}\ \loc_{x}\ 12;\comp;\lock\\
+    \text{\colorbox{yellow}{$\obj{\emptyevent}$}}& \mid & \obj{\ev{Branch}\ 1;\comp;\lock} & \mid &  \emptyevent & \mid & \ev{Branch}\ 1;\comp;\lock\\
     \obj{\ev{Dealloc}\ \loc_{y};\comp;\unlock} & \mid & \obj{\ev{Dealloc}\ \loc_{y};\comp;\lock} & \mid &  \ev{Dealloc}\ \loc_{y};\comp;\unlock & \mid & \ev{Dealloc}\ \loc_{y};\comp;\unlock\\
-    \obj{\widehat{\ev{Get}}\ \loc_{y}\ 6;\comp;\lock} & \mid & \obj{\ev{Get}\ \loc_{y}\ 6;\comp;\lock} & \mid &  \ev{Use}\ \loc_{y}\ 6;\comp;\unlock & \mid & \ev{Use}\ \loc_{y}\ 6;\comp;\lock\\
+    \obj{\widehat{\ev{Get}}\ \loc_{y}\ 6;\comp;}\text{\colorbox{lightgray}{$\obj{\lock}$}} & \mid & \obj{\ev{Get}\ \loc_{y}\ 6;\comp;\lock} & \mid &  \ev{Use}\ \loc_{y}\ 6;\comp;\text{\colorbox{lightgray}{$\unlock$}} & \mid & \ev{Use}\ \loc_{y}\ 6;\comp;\lock\\
   \end{array}}
@@
   When the data independent timing mode is off, the execution yields events in similar fashion to before (\Cref{subsec:ltms,subsec:lsms,subsec:lms}).
-  But, if it is turned on, then the branching event does not fire anymore and both reading and writing to memory gets ultimately translated to a specification trace with no exposed secrets.
+  But, if it is turned on, then the branching event \colorbox{yellow}{does not fire} anymore and both reading and writing to memory gets ultimately translated to a specification trace with \colorbox{lightgray}{no exposed secrets}.
 \end{example}

### 2023-07-12 popl/main.tex
proofs

@@
 \end{theorem}
+\begin{proof}
+  From \Thmref{thm:cca:rtp:tms} it follows that for any $\Ltms$ program $\src{\progvar}$, it compiles to an $\Ltrg$ program $\trg{\progvar}$ that robustly satisfies \gls*{tms}.
+  Note that $\src{\progvar}$ robustly satisfies \gls*{tms} by the properties of the typesystem of $\Ltms$.
+  Then, \Thmref{thm:ccb:rtp:sms} demonstrates that, assuming $\trg{\progvar}$ robustly satisfies \gls*{sms}, the program $\trg{\progvar}$ compiles to an $\Lms$ program $\irl{p}$ that also robustly satisfies \gls*{sms}.
+  From \Thmref{thm:ccab:rtp:ms} it follows that $\src{p}$ compiles to $\irl{p}$ that robustly satisfies \gls*{ms}, since \gls*{ms} is the intersection of \gls*{tms} and \gls*{sms}.
+\end{proof}
 
@@
 \begin{proof}
-	From \Thmref{thm:ccscct:rtp:scct}, we have that any LTMS program $\src{p}$ gets compiled into a L program $\trg{p}$ that is also TMS.
-	Then, from \Thmref{thm:cccfccdce:rtp:ms} we have that $\trg{p}$ gets compilet to a program ?? that is also SMS, assuming it was SMS to begin with.
-	Since TMS + SMS = MS, ?? is MS.
-	?? gets optimised by the optimisation passes and its optimisation is ???, which is still MS.
-	Finally, \Thmref{thm:ccab:rtp:ms} ensures that ??? gets compiled into ????, which is CCT assuming that the original program was CCT to begin with.
-	\MPin{please fill in the macros here}
-  % \MKin{%
-  %   \textbackslash Thmref or \textbackslash Cref?
-  % }
-% Unfold the definition of $\rtp{\cc{\Ltms}{\Lscct}}{\mssafe\cap\scctsafe}$.
-% So, let $\pi\in\mssafe\cap\scctsafe$, $\src{\progvar}$ such that $\rsat{\src{\progvar}}{\pi}$ and let $\obj{\contextvar},\obj{\runtimetermvar}$ with $\progstepto{\prog[\obj]{\obj{\contextvar}}{\left(\cc{\Ltms}{\Lscct}(\src{\progvar})\right)}}{\obj{\runtimetermvar}}{\obj{\trace}}$.
-% What is left to show is $\scctfilterLscct[\emptyset]{\obj{\trace}}\in\pi$.
-% Let $\irl{\progvar}=\cc{\Ltms}{\Ltrg}\circ\cc{\Ltrg}{\Lms}\circ\cc[\gls*{cf}]{\Lms}{\Lms}\circ\cc[\gls*{dce}]{\Lms}{\Lms}(\src{\progvar})$ such that $\cc{\Lms}{\Lscct}(\irl{\progvar})=\cc{\Ltms}{\Lscct}(\src{\progvar})$.
-% Note that $\pi\in\scctsafe$, so apply \Thmref{thm:ccscct:rtp:scct}
+	From \Thmref{thm:ccab:rtp:ms}, we have that any $\Ltms$ program $\src{\progvar}$ compiles into a $\Lms$ program $\irl{\progvar}$ that robustly satisfies \gls*{ms}.
+	Then, from \Thmref{thm:cccfccdce:rtp:ms} we have that $\irl{\progvar}$ gets optimised to a program $\irl{\progvar[']}$ that is also \gls*{ms}, where the order of optimisations does not matter for $\irl{\progvar[']}$ to be \gls*{ms}.
+  Assuming $\irl{\progvar[']}$ robustly satisfies \gls*{scct}, by \Thmref{thm:ccscct:rtp:scct} it compiles to an $\Lscct$ program $\obj{\progvar}$ that robustly satisfies \gls*{scct} as well.
+  Finally, from \Thmref{thm:rtp} it follows that, given $\src{\progvar}$ robustly satisfies \gls*{scct} and \gls*{ms}, $\obj{\progvar}$ also robustly satisfies \gls*{scct} and \gls*{ms}.
 \end{proof}

### 2023-07-12 popl/main.tex
fix typo

@@
 \begin{abstract}
-\MPpost{
-	this abstract fails in setting up the robustness argument.
-	it'll be a good exercise for you to take inspiration from this, but then write something as concise but with a more direct focus on robust properties
-}
+% \MPpost{
+%   this abstract fails in setting up the robustness argument.
+%   it'll be a good exercise for you to take inspiration from this, but then write something as concise but with a more direct focus on robust properties
+% }
 % context
-To ensure that secure applications do not leak their secrets, they are required to uphold several security properties such as spatial and temporal memory safety as well as criptographic constant time.
+To ensure that secure applications do not leak their secrets, they are required to uphold several security properties such as spatial and temporal memory safety as well as cryptographic constant time.
 % need
@@
 % task
-Unfortunately, given two secure compiler passes that each preserve a possibly-different security property, there is no way to tell what kind of security property will the composition of those secure compilers preserve.
+Unfortunately, given two secure compiler passes that each preserve a possibly different security property, it is unclear what kind of security property the composition of those secure compiler passes preserve.
+%there is no way to tell what kind of security property will the composition of those secure compilers preserve.
 

### 2023-07-12 popl/main.tex
get rid of double neg

@@
 % Without a framework for composing secure compiler passes, it is not possible to enable separation of concerns, e.g., to have a secure compilation pass that ensures \gls*{ms} that is developed independently of another secure pass for \gls*{cct} as suggested before and as showcased in \Cref{ex:strncpy}.
-\MPpost{
-	this intro does a poor job at introducing the robust compilation argument.
-	perhaps it'd be best to start from robust properties, which motivates the use of robust compilers?
-	idk
-	as for the abstract, this is a good starting point for you to make a better intro
-}
+% \MPpost{
+%   this intro does a poor job at introducing the robust compilation argument.
+%   perhaps it'd be best to start from robust properties, which motivates the use of robust compilers?
+%   idk
+%   as for the abstract, this is a good starting point for you to make a better intro
+% }
 
@@
 \gls*{cct} is a security property that ensures sensitive data is not leaked via timing side-channels~\cite{kocher1996timing}.
-Together, \gls*{sms}, \gls*{tms} and \gls*{cct} yield \gls*{msscct}, which is the gold standard of security properties for secure applications: programs attaining \gls*{msscct} do not leak sensitive data neither through erroneous memory accesses, nor through timing side-channels.
+Together, \gls*{sms}, \gls*{tms} and \gls*{cct} yield \gls*{msscct}, which is the gold standard of security properties for secure applications: programs attaining \gls*{msscct} do not leak sensitive data either through erroneous memory accesses, nor through timing side-channels.
 As discussed in \Cref{ex:strncpy}, these security properties can be enforced by compiler passes~\cite{bond2017vale,almeida2017jasmin}, to ensure programmers need not be aware of the various architectural details where their code will run.

### 2023-07-12 popl/main.tex
synonym

@@
 \gls*{cct} is a security property that ensures sensitive data is not leaked via timing side-channels~\cite{kocher1996timing}.
-Together, \gls*{sms}, \gls*{tms} and \gls*{cct} yield \gls*{msscct}, which is the gold standard of security properties for secure applications: programs attaining \gls*{msscct} do not leak sensitive data either through erroneous memory accesses, nor through timing side-channels.
+Together, \gls*{sms}, \gls*{tms} and \gls*{scct}, a variant of \gls*{cct}, yield \gls*{msscct}, which is the gold standard of security properties for secure applications: programs attaining \gls*{msscct} do not leak sensitive data either through erroneous memory accesses, nor through timing side-channels.
 As discussed in \Cref{ex:strncpy}, these security properties can be enforced by compiler passes~\cite{bond2017vale,almeida2017jasmin}, to ensure programmers need not be aware of the various architectural details where their code will run.
@@
 Because of timing attacks, fixing \gls*{sms} is not enough to make \texttt{strncpy} secure.
-In fact, the loop can terminate early, as soon as the string-terminating character \texttt{'\textbackslash 0'} is encountered, making program execution time proportional to the length of the array pointed by \texttt{src}.
+In fact, the loop can terminate early, as soon as the string-terminating character \texttt{'\textbackslash 0'} is encountered, thus making program execution time proportional to the length of the array pointed by \texttt{src}.
 % Often, programmers rewrite programs like these to have loops that run with a fixed number of iterations.
@@
 Alas, code is not run in isolation, so a malicious attacker could supply code that intracts with \texttt{strncpy} and trigger a violation of either \gls*{ms} or \gls*{cct} by calling \texttt{strncpy} with an argument for \texttt{src} that points to uninitialised memory.
-This would, in turn, triggering a series of reads from uninitialised memory, which is an immediate \gls*{ms} violation with devastating real-world implications~\cite{uninit-0,uninit-1,uninit-2,uninit-3,uninit-4}.
+This would, in turn, triggering a series of reads from uninitialised memory, which is an immediate \gls*{ms} violation with devastating real-world consequences~\cite{uninit-0,uninit-1,uninit-2,uninit-3,uninit-4}.
 % An easy fix to sidestep this is to use non-nullable pointers as parameters instead, e.g., as presented in Cyclone~\cite{jim2002cyclone}.

### 2023-07-12 popl/main.tex
rewrite weird sentence

@@
 Robust compilers~\cite{abate2019jour} are a form of secure compilers that preserve security properties even in the presence of arbitrary attackers interacting with compiled code.
-Thus, robust compilers can be used to prevent the vulnerability resulting from uninitialised memory, e.g., by targeting capability-based languages such as CHERI~\cite{woodruff2014CHERI}, Arm Morello~\cite{arm-morello}, or MSWasm~\cite{michael2023mswasm}, where the compiler relies on capabilities to checks that pointers are always initialised.
+Thus, robust compilers can be used to prevent, e.g., vulnerabilities resulting from uninitialised memory, e.g., by targeting capability-based languages such as CHERI~\cite{woodruff2014CHERI}, Arm Morello~\cite{arm-morello}, or MSWasm~\cite{michael2023mswasm}, where the compiler relies on capabilities to check that pointers are always initialised.
 \end{example}

### 2023-07-12 popl/main.tex
intro

@@
 
-Unfortunately, given secure compiler passes that each preserve a possibly-different security property, there is no way to tell what kind of security property will the composition of those secure compilers preserve.
+Unfortunately, given secure compiler passes that each preserve a possibly different security property, there is no way to tell what kind of security property will the composition of those secure compilers preserve.
 % Different kinds of compilation passes exist, e.g., there are usually several compilation passes that perform optimisations~\cite{lattner2004llvm,googlev8,androidstudio,kuepper2023cryptopt,manjikian1997fusion,wegman1991ccp} or source-code instrumentations~\cite{nagarakatte2009soft,nagarakatte2010cets,akritidis2009baggy,dhumbumroong2020boundwarden,jung2021pico,nam2019framer,shankaranarayana2023tailcheck,younan2010paricheck,zhou2023fatptrs,bond2017vale,almeida2017jasmin,kuepper2023cryptopt,cauligi2019fact} that enforce that the compiled program fulfils a property of interest.
-Worse, without a framework for composing secure compiler passes, it is not possible to enable separation of concerns, e.g., to have a secure compilation pass that ensures \gls*{ms} that is developed independently of another secure pass for \gls*{cct}, that is developed independently of other passes such as optimisation ones.
+Worse, without a framework for composing secure compiler passes, it is not possible to enable separation of concerns, e.g., to have a secure compilation pass that ensures \gls*{ms} that is developed independently of another secure pass for \gls*{cct}, that is developed independently of other passes, such as optimisations.
 
 This paper introduces a framework for reasoning about the composition of secure and optimising compiler passes akin to those of \Cref{ex:strncpy} and it showcases the power of this framework by instantiating it on a multi-pass compilation chain.
-To this end, this paper first discusses how to compose security properties, such as \gls*{tms} and \gls*{sms} into \gls*{ms}, and then adding \gls*{cct} to the mix to obtain \gls*{msscct}.
-Then, this paper defines compiler composition and formalises that given two passes that securely preserve two (possibly-distinct) properties, their composition securely preserves the composition of those properties.
-The paper then defines several secure compiler passes, where each is either preserving a different security property (\gls*{tms}, \gls*{sms}, \gls*{cct}) or performing a security-preserving optimisation, (e.g., applying \gls*{cf} or \gls*{dce}).
-Finally, this paper shows that composing these secure compiler passes into a multi-pass compilation chain results that provides end-to-end preservation of \gls*{msscct}.
+To this end, this paper first discusses how to compose security properties, such as \gls*{tms} and \gls*{sms} into \gls*{ms}, and then adding \gls*{scct} to the mix to obtain \gls*{msscct}.
+Then, this paper defines compiler composition and formalises that given two passes that securely preserve two (possibly distinct) properties, their composition securely preserves the composition of those properties.
+The paper then defines several secure compiler passes, where each is either preserving a different security property (\gls*{tms}, \gls*{sms}, \gls*{scct}) or performing a security-preserving optimisation, (e.g., applying \gls*{cf} or \gls*{dce}).
+Finally, this paper shows that composing these secure compiler passes into a multi-pass compilation chain results in the end-to-end preservation of \gls*{msscct}.
 Crucially, this paper derives the security of the multi-pass compiler from the composition of the security properties preserved by its individual passes.
@@
         This paper presents a case-study showcasing the conjunction of the previous contributions (\Cref{sec:casestud:defs,sec:casestud:rtp}).
-        To this end, it presents a compilation chain consisting of several passes that ultimately preserves \gls*{msscct} by means of composing the individual, secure passes concerning \gls*{tms}, \gls*{sms}, and \gls*{cct}, respectively.
-        Furthermore, the chain includes two optimisation passes: One performs a simple \gls*{dce} and the other \gls*{cf}.
-        The formalisation of this case study showcases the power of the presented framework: Namely, the divide-and-conquer approach to software engineering is a viable strategy even for the development of secure compilers.
+        To this end, it presents a compilation chain consisting of several passes that ultimately preserves \gls*{msscct} by means of composing the individual, secure passes concerning \gls*{tms}, \gls*{sms}, and \gls*{scct}, respectively.
+        Furthermore, the chain includes two optimisation passes: One performs \gls*{dce} and the other \gls*{cf}.
+        The formalisation of this case study showcases the power of the presented framework: The divide-and-conquer approach to software engineering is a viable strategy even for the development of secure compilers.
 %        This demonstrates that the usage of the presented framework is both simpler and more expressive compared to one monolithic security proof of the whole compilation chain.

### 2023-07-12 popl/main.tex
move fig

@@
 Since reasoning on monitors is easier than directly on just traces, this section presents a monitor for each of the previously shown safety properties (\Cref{subsec:propdefs}).
-\MPpost{
-	this is fine, but it'd be nice to provide more insights on why it's easier
-}
+%\MPpost{
+%	this is fine, but it'd be nice to provide more insights on why it's easier
+%}
 To lessen the burden when proving that a monitor accepts the trace of a program execution, each monitor uses a custom trace model that contains only the relevant information related to the property the monitor checks.
@@
 \end{example}
-\MPpost{
-	in general, i think embedding the traces in the text does not make them readable.
-	in the exorcising paper i think i put them in align env, or in a way to make them stand out.
-	i don't think we have the time to do this change uniformly now,
-	but at some point we should
-}
 
@@
 
-This paper considers \gls*{cct} as an architecture-dependent property\MPin{can we justify this}, so writing \gls*{cct} should not be one of the programmer's concerns~\cite{cauligi2019fact}.
-Such consideration is backed up by architecture providing a data (operand) independent timing mode, such as processors by Arm~\cite[p.~543]{arm-refman} and Intel~\cite[p.~80]{intel-refman}.
+% This paper considers \gls*{cct} as an architecture-dependent property\MPin{can we justify this}, so
+Writing code attaining \gls*{cct} should not be of the programmer's concerns~\cite{cauligi2019fact}.
+Such consideration is also backed up by architecture providing a data (operand) independent timing mode, such as processors by Arm~\cite[p.~543]{arm-refman} and Intel~\cite[p.~80]{intel-refman}.
 In spirit of this, language $\Lscct$ allows violating \gls*{cct} by emitting events on, e.g., branching and division, that contain secrets (\Cref{subsec:lscct}), but provides a way to read and write to a {\em model--specific register} that enables a ``\gls*{cct}-mode''.
@@
 \subsubsection{Proving Robust Safety Property Preservation}
+
+The proof of \Cref{thm:cca:rtp:tms} serves as boilerplate for proofs of the other secure compilation theorems of this paper.
+Unfolding the statement gives the assumptions that for any $\pi\in\lceil\tmssafe\rceil$\footnote{$\lceil\cdot\rceil$ lifts the property to a hyperproperty by applying the powerset operation~\cite{clarkson2008hyper}.}, $\src{\trace}$, $\src{\runtimetermvar}$, and component $\src{\library_{comp}}$, $\rsat{\src{\library_{comp}}}{\pi}$ and $\progstepto{\prog[\trg]{\trg{\library_{\ctx}}}{\cca(\src{\library_{\comp}})}}{\trg{\runtimetermvar}}{\trg{\trace}}$, where $\trg{\library_{\ctx}}$ is arbitrary.
+The proof obligation is $\msfilterL{\trg{\trace}}\in\pi$, i.e., the specification trace associated to $\trg{\trace}$ should satisfy the property $\pi$.
+To show this, the trace $\trg{\trace}$ should be related with some $\Ltms$ trace $\src{\trace}$ and since there is a target execution associated to this trace, the task is to find an associated $\Ltms$ execution.
+The trace $\trg{\trace}$ is split into different parts, where each part contains the events that either the context or the component does, but not both.
+Because of this, all such trace segments are ,,well-bracketed'' in the sense that they start with either $\trg{\ev{Start}}$, $\trg{\ev{Call}\ \comptoctx\ foo\ \valueexpr}$, or $\trg{\ev{Ret}\ \comptoctx\ \valueexpr}$ and end with either $\trg{\ev{End}\ \valueexpr}$, $\trg{\ev{Ret}\ \ctxtocomp\ \valueexpr}$, or $\trg{\ev{Call}\ \ctxtocomp\ foo\ \valueexpr}$.
+In the following, the former is referred to as a context segment, since, for these, execution happens in $\trg{\library_{\ctx}}$, and the latter as a component segment, since, for these, execution happens in $\cca(\src{\library_{\comp}})$.
+\Cref{fig:proofdiag:rtp} visualises this split for a program execution with one call from context to component and how the target execution is related to a source execution.
+In the figure, the green dashed lines encompass the component segments while the orange boxes contain the actual context switches from context to component or vice versa.
+For the proof, it is necessary to know when two states $\src{\Omega}$ and $\trg{\Omega}$ are related.
+Two cross-language relations make this precise: (i) $\multimap_{\delta}$ relates states that are involved in a context segment, allowing the target execution to perform internal calls, and (ii) $\approx_{\delta}$ relates states that are involved in a component segment, where both states need to agree exactly, i.e., the memory and the control flow states are required to contain the same information.
+The relations are indexed with $\delta$, which is simply an injective mapping from $\Ltms$ locations $\src{\loc}$ to $\Ltrg$ locations $\trg{\loc}$.
+Note that the relations $\multimap_{\delta}$ and $\approx_{\delta}$ swap when context switching.
 
@@
 
-The proof of \Cref{thm:cca:rtp:tms} serves as boilerplate for proofs of the other secure compilation theorems of this paper.
-Unfolding the statement gives the assumptions that for any $\pi\in\lceil\tmssafe\rceil$\footnote{$\lceil\cdot\rceil$ lifts the property to a hyperproperty by applying the powerset operation~\cite{clarkson2008hyper}.}, $\src{\trace}$, $\src{\runtimetermvar}$, and component $\src{\library_{comp}}$, $\rsat{\src{\library_{comp}}}{\pi}$ and $\progstepto{\prog[\trg]{\trg{\library_{\ctx}}}{\cca(\src{\library_{\comp}})}}{\trg{\runtimetermvar}}{\trg{\trace}}$, where $\trg{\library_{\ctx}}$ is arbitrary.
-The proof obligation is $\msfilterL{\trg{\trace}}\in\pi$, i.e., the specification trace associated to $\trg{\trace}$ should satisfy the property $\pi$.
-To show this, the trace $\trg{\trace}$ should be related with some $\Ltms$ trace $\src{\trace}$ and since there is a target execution associated to this trace, the task is to find an associated $\Ltms$ execution.
-The trace $\trg{\trace}$ is split into different parts, where each part contains the events that either the context or the component does, but not both.
-Because of this, all such trace segments are ,,well-bracketed'' in the sense that they start with either $\trg{\ev{Start}}$, $\trg{\ev{Call}\ \comptoctx\ foo\ \valueexpr}$, or $\trg{\ev{Ret}\ \comptoctx\ \valueexpr}$ and end with either $\trg{\ev{End}\ \valueexpr}$, $\trg{\ev{Ret}\ \ctxtocomp\ \valueexpr}$, or $\trg{\ev{Call}\ \ctxtocomp\ foo\ \valueexpr}$.
-In the following, the former is referred to as a context segment, since, for these, execution happens in $\trg{\library_{\ctx}}$, and the latter as a component segment, since, for these, execution happens in $\cca(\src{\library_{\comp}})$.
-\Cref{fig:proofdiag:rtp} visualises this split for a program execution with one call from context to component and how the target execution is related to a source execution.
-In the figure, the green dashed lines encompass the component segments while the orange boxes contain the actual context switches from context to component or vice versa.
-For the proof, it is necessary to know when two states $\src{\Omega}$ and $\trg{\Omega}$ are related.
-Two cross-language relations make this precise: (i) $\multimap_{\delta}$ relates states that are involved in a context segment, allowing the target execution to perform internal calls, and (ii) $\approx_{\delta}$ relates states that are involved in a component segment, where both states need to agree exactly, i.e., the memory and the control flow states are required to contain the same information.
-The relations are indexed with $\delta$, which is simply an injective mapping from $\Ltms$ locations $\src{\loc}$ to $\Ltrg$ locations $\trg{\loc}$.
-Note that the relations $\multimap_{\delta}$ and $\approx_{\delta}$ swap when context switching.
-
 So far, the paper explained how to relate an $\Ltms$ execution with the $\Ltrg$ execution at hand.

### 2023-07-12 popl/main.tex
get rid of newpage

@@
 \section{Conclusion\pages{$\sfrac{1}{2}$}}\label{sec:concl}
-This paper tackled the problem of understanding what kind of security properties does a secure compiler preserve, when said compiler is the combination of smaller compiler passes that preserve possibly different security properties.
+This paper tackled the problem of understanding what kind of security properties does a secure compiler preserve, when said compiler is the combination of compiler passes that preserve possibly different security properties.
 % 
@@
 % 
-Then, it proved that composing secure compilers that preserve certain properties results in a secure compiler that preserves the composition of the properties.
+Then, it proved that composing secure compilers that preserve certain properties results in a secure compiler that preserves the composition of these properties.
 % 
@@
 
-\newpage
+%\newpage
 %%

### 2023-07-12 popl/main.tex
:chicken: Ltrg

@@
 Thus, from a security perspective, this considers only a passive attacker model, where the attacker observes the execution and, e.g., retrieves secrets from that.
-To consider a stronger model similarly to what existing work does~\cite{abate2019jour,abate2021extacc,maffeis2008code-carrying,gordon2003authenticity,fournet2007authorization,bengtson2011refine,backes2014uniontyps,michael2023mswasm,swasey2017robust}, the concept of satisfaction can be extended with {\em robustness}.
-\MPpost{
-	cit michel's and move
-}
+To consider a stronger model similarly to what existing work does~\cite{abate2019jour,abate2021extacc,maffeis2008code-carrying,gordon2003authenticity,fournet2007authorization,bengtson2011refine,backes2014uniontyps,michael2023mswasm,swasey2017robust,sammler2019benefits}, the concept of satisfaction can be extended with {\em robustness}.
 Robust satisfaction considers partial programs $\progvar$, i.e., components with missing imports, which cannot run until said imports are fulfilled.

### 2023-07-12 popl/main.tex
:chicken: Add Lirl Lscct

@@
 \begin{theorem}[Compiler $\ccmsscct$ is secure with respect to \gls*{scct}]\label{thm:ccall:rtp:msscct}
-  $\rtp{\cc{\Ltms}{\Lscct}}{\mssafe\cap\scctsafe}$ \Coqed
+  $\rtp{\cc{\Ltms}{\Lscct}}{\mssafe\cap\scctsafe}$ % \Coqed
 \end{theorem}

### 2023-07-12 popl/main.tex
6.1.1

--- 2023-07-12 popl/main.tex
@@
 
-The proof of \Cref{thm:cca:rtp:tms} serves as boilerplate for proofs of the other secure compilation theorems of this paper.
-Unfolding the statement gives the assumptions that for any $\pi\in\lceil\tmssafe\rceil$\footnote{$\lceil\cdot\rceil$ lifts the property to a hyperproperty by applying the powerset operation~\cite{clarkson2008hyper}.}, $\src{\trace}$, $\src{\runtimetermvar}$, and component $\src{\library_{comp}}$, $\rsat{\src{\library_{comp}}}{\pi}$ and $\progstepto{\prog[\trg]{\trg{\library_{\ctx}}}{\cca(\src{\library_{\comp}})}}{\trg{\runtimetermvar}}{\trg{\trace}}$, where $\trg{\library_{\ctx}}$ is arbitrary.
-The proof obligation is $\msfilterL{\trg{\trace}}\in\pi$, i.e., the specification trace associated to $\trg{\trace}$ should satisfy the property $\pi$.
-To show this, the trace $\trg{\trace}$ should be related with some $\Ltms$ trace $\src{\trace}$ and since there is a target execution associated to this trace, the task is to find an associated $\Ltms$ execution.
-The trace $\trg{\trace}$ is split into different parts, where each part contains the events that either the context or the component does, but not both.
+We illustrate the proof of \Cref{thm:cca:rtp:tms} since the other secure compilation proofs of this paper follow the same approach.
+Unfolding the theorem statement yields the following assumptions:  
+	for any $\pi\in\lceil\tmssafe\rceil$\footnote{$\lceil\cdot\rceil$ lifts the property to a hyperproperty by applying the powerset operation~\cite{clarkson2008hyper}.}, 
+	$\src{\trace}$, 
+	$\src{\runtimetermvar}$, 
+	and component $\src{\library_{comp}}$, 
+	we have that $\rsat{\src{\library_{comp}}}{\pi}$ and 
+	$\progstepto{\prog[\trg]{\trg{\library_{\ctx}}}{\cca(\src{\library_{\comp}})}}{\trg{\runtimetermvar}}{\trg{\trace}}$, where $\trg{\library_{\ctx}}$ is arbitrary.
+The proof obligation is $\msfilterL{\trg{\trace}}\in\pi$, i.e., the specification trace associated to $\trg{\trace}$ satisfies the property $\pi$.
+A way to show this, is relating trace $\trg{\trace}$ to some $\Ltms$ trace $\src{\trace}$ (which already satisfies the property as per the assumptions).
+The assumptions already contain a target execution associated to this trace, so the task is to find an associated $\Ltms$ execution that yields $\trg{\trace}$.
+The trace $\trg{\trace}$ is split into different parts, as commonly done in secure compilation works~\cite{}\MP{catalin's CCS and catalin+akram's paper},where each part contains the events that either the context or the component does, but not both.
 Because of this, all such trace segments are ,,well-bracketed'' in the sense that they start with either $\trg{\ev{Start}}$, $\trg{\ev{Call}\ \comptoctx\ foo\ \valueexpr}$, or $\trg{\ev{Ret}\ \comptoctx\ \valueexpr}$ and end with either $\trg{\ev{End}\ \valueexpr}$, $\trg{\ev{Ret}\ \ctxtocomp\ \valueexpr}$, or $\trg{\ev{Call}\ \ctxtocomp\ foo\ \valueexpr}$.
-In the following, the former is referred to as a context segment, since, for these, execution happens in $\trg{\library_{\ctx}}$, and the latter as a component segment, since, for these, execution happens in $\cca(\src{\library_{\comp}})$.
-\Cref{fig:proofdiag:rtp} visualises this split for a program execution with one call from context to component and how the target execution is related to a source execution.
+In the following, the former is referred to as a context segment, since these executions happen in $\trg{\library_{\ctx}}$, and the latter is referred to as a component segment, since these executions happen in $\cca(\src{\library_{\comp}})$.
+\Cref{fig:proofdiag:rtp} visualises this division for a program execution with one call from context to component and how the target execution is related to a source execution.
 In the figure, the green dashed lines encompass the component segments while the orange boxes contain the actual context switches from context to component or vice versa.
-For the proof, it is necessary to know when two states $\src{\Omega}$ and $\trg{\Omega}$ are related.
-Two cross-language relations make this precise: (i) $\multimap_{\delta}$ relates states that are involved in a context segment, allowing the target execution to perform internal calls, and (ii) $\approx_{\delta}$ relates states that are involved in a component segment, where both states need to agree exactly, i.e., the memory and the control flow states are required to contain the same information.
-The relations are indexed with $\delta$, which is simply an injective mapping from $\Ltms$ locations $\src{\loc}$ to $\Ltrg$ locations $\trg{\loc}$.
+From a technical perspective, as typically done in compilation proofs, the proof requires set up and maintain a relation between $\src{\Omega}$ and $\trg{\Omega}$ .
+Here, we use two cross-language relations make this precise: (i) $\multimap_{\delta}$ relates states that are involved in a context segment, allowing the target execution to perform internal calls, and (ii) $\approx_{\delta}$ relates states that are involved in a component segment, where both states need to agree exactly, i.e., the memory and the control flow states are required to contain the same information.
+The relations are indexed with $\delta$, which is injective mapping from $\Ltms$ locations $\src{\loc}$ to $\Ltrg$ locations $\trg{\loc}$.
 Note that the relations $\multimap_{\delta}$ and $\approx_{\delta}$ swap when context switching.
@@
 
-So far, the paper explained how to relate an $\Ltms$ execution with the $\Ltrg$ execution at hand.
-However, it is not yet demonstrated how the corresponding $\Ltms$ execution is actually built.
-To do so, a trace-based backtranslation technique~\cite{abate2019jour} can be used to build a context $\src{\library_{\ctx}}$ that behaves similar to $\trg{\library_{\ctx}}$.
-Since $\Ltms$ is statically typed, the context $\src{\library_{\ctx}}$ obtained from the backtranslation needs to be well-typed when linked with $\src{\library_{\comp}}$.
-For context segments of the trace $\trg{\trace}$ it is also necessary to show that the execution behaves similarily, i.e., there is an $\Ltms$ trace segment $\src{\trace[']}$ that is related to the context segment, and that the states are related by $\multimap_{\delta}$.
-When passing control to the component, the relatedness of states and traces follows from compiler correctness.
-From all this, a source execution $\progstepto{\src{\prog{\library_{\ctx}}{\library_{\comp}}}}{\src{\runtimetermvar}}{\src{\trace}}$ can be obtained.
+So far, the paper explained how to relate an $\Ltms$ execution with a $\Ltrg$ execution.
+The next question is therefore how to build the corresponding $\Ltms$ execution.
+This is done using a standard secure compilation proof technique called trace-based backtranslation~\cite{abate2019jour,korashy2021capableptrs,patrignani2021rsc}, which can be used to build a context $\src{\library_{\ctx}}$ that behaves similar to $\trg{\library_{\ctx}}$.
+% Since $\Ltms$ is statically typed, the context $\src{\library_{\ctx}}$ obtained from the backtranslation needs to be well-typed when linked with $\src{\library_{\comp}}$.
+For context segments of the trace $\trg{\trace}$ it is also necessary to show that the execution behaves similarily, i.e., the context obtained from the backtranslation generates trace $\src{\trace}$.
+For component segments of the trace, the relatedness of states and traces follows from a compiler correctness argument.
+These two arguments yield the following source execution $\progstepto{\src{\prog{\library_{\ctx}}{\library_{\comp}}}}{\src{\runtimetermvar}}{\src{\trace}}$.
 
 The proof now works as follows.
-Rewrite the goal from $\msfilterL{\trg{\trace}}\in\pi$ to $\msfilterLtms{\src{\trace}}\in\pi$ after demonstrating that $\msfilterL{\trg{\trace}}=\msfilterLtms{\src{\trace}}$.
-Then, specialize the robust satisfaction assumption $\rsat{\src{\library_{\comp}}}{\pi}$ to use the context $\src{\library_{\ctx}}$, which was obtained from backtranslating the context segments of the target trace $\trg{\trace}$, and the source execution $\progstepto{\src{\prog{\library_{\ctx}}{\library_{\comp}}}}{\src{\runtimetermvar}}{\src{\trace}}$.
-This gives $\msfilterLtms{\src{\trace}}\in\pi$, which was to be demonstrated.
+Given that $\msfilterL{\trg{\trace}}=\msfilterLtms{\src{\trace}}$, the proof goal changes from $\msfilterL{\trg{\trace}}\in\pi$ to $\msfilterLtms{\src{\trace}}\in\pi$.
+Then, by specializing the robust satisfaction assumption $\rsat{\src{\library_{\comp}}}{\pi}$ to use the context $\src{\library_{\ctx}}$ obtained from the backtranslation which yields
+ % the context segments of the target trace $\trg{\trace}$, and the 
+the following source execution $\progstepto{\src{\prog{\library_{\ctx}}{\library_{\comp}}}}{\src{\runtimetermvar}}{\src{\trace}}$.
+This gives $\msfilterLtms{\src{\trace}}\in\pi$, which was the proof goal.
 

### 2023-07-12 popl/main.tex
note
add cit

@@
 The assumptions already contain a target execution associated to this trace, so the task is to find an associated $\Ltms$ execution that yields $\trg{\trace}$.
-The trace $\trg{\trace}$ is split into different parts, as commonly done in secure compilation works~\cite{}\MP{catalin's CCS and catalin+akram's paper},where each part contains the events that either the context or the component does, but not both.
+The trace $\trg{\trace}$ is split into different parts, as commonly done in secure compilation works~\cite{korashy2021capableptrs,abate2018goodcomps},where each part contains the events that either the context or the component does, but not both.
 Because of this, all such trace segments are ,,well-bracketed'' in the sense that they start with either $\trg{\ev{Start}}$, $\trg{\ev{Call}\ \comptoctx\ foo\ \valueexpr}$, or $\trg{\ev{Ret}\ \comptoctx\ \valueexpr}$ and end with either $\trg{\ev{End}\ \valueexpr}$, $\trg{\ev{Ret}\ \ctxtocomp\ \valueexpr}$, or $\trg{\ev{Call}\ \ctxtocomp\ foo\ \valueexpr}$.
@@
 
+\MPpost{
+	this can be made more precise
+}
+
 \subsection{Robust (Spatial) Memory Safety Preservation}\label{subsec:cs:ms}

### 2023-07-12 popl/main.tex
nits
nit

--- 2023-07-12 popl/main.tex
@@
 \begin{center}\small\it
-	{This paper uses syntax highlighting accessible to both colourblind and black \& white readers~\citep{patrignani2020use}.
-	Specifically, it makes use of a $\src{blue}$, $\src{sans\text{-}serif}$ font for a $\src{source}$,
-	a $\trg{red}$, $\trg{bold}$ font for an $\trg{intermediate}$,
-	and a $\obj{green}$, $\obj{teletype}$ font for a $\obj{target}$ language.
+	{This paper uses syntax highlighting accessible to both colourblind and black \& white readers.
+	% ~\citep{patrignani2020use}.
+% 	% Specifically, it makes use of a $\src{blue}$, $\src{sans\text{-}serif}$ font for a $\src{source}$,
+% 	% a $\trg{red}$, $\trg{bold}$ font for an $\trg{intermediate}$,
+% 	% and a $\obj{green}$, $\obj{teletype}$ font for a $\obj{target}$ language.
 	}
@@
     \mi{(Control~Flow~States)}~&\cfstate\bnfdef\tup{\library}{\ectxstack}
-  \end{aligned}\\
-  \begin{aligned}
+  	\\
     \mi{(Locations)}~&\loc\in\nat
@@
     \mi{(Substitutions)}~&\substvar\bnfdef \hole{\cdot} \mid \subst{\valueexpr}{x},\gamma
-  \end{aligned}\\
-  \begin{aligned}
+  	\\
     \mi{(Stores)}~&\Delta\bnfdef\hole{\cdot} \mid x\mapsto\tup{\loc}{\sandboxtag,\poison,n},\Delta
@@
     \mi{(Heaps)}~&H\bnfdef\hole{\cdot} | \valueexpr,H
-    \\
+    \end{aligned}\\
+  \begin{aligned}
     \mi{(Memory~States)}~&\memstate\bnfdef\tup{H^{\ctx}}{H^{\comp};\Delta}
@@
 
+% \MPpost{
+% 	this can be made more precise
+% }
+
 \subsection{Robust (Spatial) Memory Safety Preservation}\label{subsec:cs:ms}

### 2023-07-12 popl/main.tex
nit

@@
 For \gls*{cf}, the pass uses an auxiliary function \texttt{mix} that does the actual work.
-It rewrites constant binary operations, e.g., $\irl{{17}+{29}-1}$ to $\irl{46-1}$, and replaces variables that are assigned to constants with their constant, e.g., $\irl{\llet{x}{7}{x}}$ to $\irl{7}$.
+It rewrites constant binary operations, e.g., $\irl{{17}-1}$ to $\irl{16}$, and replaces variables that are assigned to constants with their constant, e.g., $\irl{\llet{x}{7}{x}}$ to $\irl{7}$.
 Both passes are secure with respect to \gls*{ms}.

### 2023-07-12 popl/main.tex
fix a bunch of bogus sentences

@@
 The proof obligation is $\msfilterL{\trg{\trace}}\in\pi$, i.e., the specification trace associated to $\trg{\trace}$ satisfies the property $\pi$.
-A way to show this, is relating trace $\trg{\trace}$ to some $\Ltms$ trace $\src{\trace}$ (which already satisfies the property as per the assumptions).
+A way to show this is to relate trace $\trg{\trace}$ to some $\Ltms$ trace $\src{\trace}$ (which already satisfies the property as per the assumptions).
 The assumptions already contain a target execution associated to this trace, so the task is to find an associated $\Ltms$ execution that yields $\trg{\trace}$.
-The trace $\trg{\trace}$ is split into different parts, as commonly done in secure compilation works~\cite{korashy2021capableptrs,abate2018goodcomps},where each part contains the events that either the context or the component does, but not both.
+The trace $\trg{\trace}$ is split into different parts, as commonly done in secure compilation works~\cite{korashy2021capableptrs,abate2018goodcomps}, where each part contains the events that either the context or the component does, but not both.
 Because of this, all such trace segments are ,,well-bracketed'' in the sense that they start with either $\trg{\ev{Start}}$, $\trg{\ev{Call}\ \comptoctx\ foo\ \valueexpr}$, or $\trg{\ev{Ret}\ \comptoctx\ \valueexpr}$ and end with either $\trg{\ev{End}\ \valueexpr}$, $\trg{\ev{Ret}\ \ctxtocomp\ \valueexpr}$, or $\trg{\ev{Call}\ \ctxtocomp\ foo\ \valueexpr}$.
@@
 In the figure, the green dashed lines encompass the component segments while the orange boxes contain the actual context switches from context to component or vice versa.
-From a technical perspective, as typically done in compilation proofs, the proof requires set up and maintain a relation between $\src{\Omega}$ and $\trg{\Omega}$ .
-Here, we use two cross-language relations make this precise: (i) $\multimap_{\delta}$ relates states that are involved in a context segment, allowing the target execution to perform internal calls, and (ii) $\approx_{\delta}$ relates states that are involved in a component segment, where both states need to agree exactly, i.e., the memory and the control flow states are required to contain the same information.
-The relations are indexed with $\delta$, which is injective mapping from $\Ltms$ locations $\src{\loc}$ to $\Ltrg$ locations $\trg{\loc}$.
+From a technical perspective, as typically done in compilation proofs, the proof requires some setup to maintain a relation between $\src{\Omega}$ and $\trg{\Omega}$ .
+Two cross-language relations make this precise: (i) $\multimap_{\delta}$ relates states that are involved in a context segment, allowing the target execution to perform internal calls, and (ii) $\approx_{\delta}$ relates states that are involved in a component segment, where both states need to agree exactly, i.e., the memory and the control flow states are required to contain the same information.
+The relations are indexed with $\delta$, which is an injective mapping from $\Ltms$ locations $\src{\loc}$ to $\Ltrg$ locations $\trg{\loc}$.
 Note that the relations $\multimap_{\delta}$ and $\approx_{\delta}$ swap when context switching.
@@
 For component segments of the trace, the relatedness of states and traces follows from a compiler correctness argument.
-These two arguments yield the following source execution $\progstepto{\src{\prog{\library_{\ctx}}{\library_{\comp}}}}{\src{\runtimetermvar}}{\src{\trace}}$.
+These two arguments yield the source execution $\progstepto{\src{\prog{\library_{\ctx}}{\library_{\comp}}}}{\src{\runtimetermvar}}{\src{\trace}}$.
 
@@
 Given that $\msfilterL{\trg{\trace}}=\msfilterLtms{\src{\trace}}$, the proof goal changes from $\msfilterL{\trg{\trace}}\in\pi$ to $\msfilterLtms{\src{\trace}}\in\pi$.
-Then, by specializing the robust satisfaction assumption $\rsat{\src{\library_{\comp}}}{\pi}$ to use the context $\src{\library_{\ctx}}$ obtained from the backtranslation which yields
- % the context segments of the target trace $\trg{\trace}$, and the 
-the following source execution $\progstepto{\src{\prog{\library_{\ctx}}{\library_{\comp}}}}{\src{\runtimetermvar}}{\src{\trace}}$.
-This gives $\msfilterLtms{\src{\trace}}\in\pi$, which was the proof goal.
+Then, by specializing the robust satisfaction assumption $\rsat{\src{\library_{\comp}}}{\pi}$ to use the context $\src{\library_{\ctx}}$, which is obtained from the backtranslation, and to use the source execution $\progstepto{\src{\prog{\library_{\ctx}}{\library_{\comp}}}}{\src{\runtimetermvar}}{\src{\trace}}$, the proof goal $\msfilterLtms{\src{\trace}}\in\pi$ follows.
 

### 2023-07-12 popl/main.tex
reword

@@
 Given that $\msfilterL{\trg{\trace}}=\msfilterLtms{\src{\trace}}$, the proof goal changes from $\msfilterL{\trg{\trace}}\in\pi$ to $\msfilterLtms{\src{\trace}}\in\pi$.
-Then, by specializing the robust satisfaction assumption $\rsat{\src{\library_{\comp}}}{\pi}$ to use the context $\src{\library_{\ctx}}$, which is obtained from the backtranslation, and to use the source execution $\progstepto{\src{\prog{\library_{\ctx}}{\library_{\comp}}}}{\src{\runtimetermvar}}{\src{\trace}}$, the proof goal $\msfilterLtms{\src{\trace}}\in\pi$ follows.
+This follows by specializing the robust satisfaction assumption $\rsat{\src{\library_{\comp}}}{\pi}$ to use the context $\src{\library_{\ctx}}$, which is obtained from the backtranslation, and to use the source execution $\progstepto{\src{\prog{\library_{\ctx}}{\library_{\comp}}}}{\src{\runtimetermvar}}{\src{\trace}}$.
 

### 2023-07-12 popl/main.tex
nits

--- 2023-07-12 popl/main.tex
@@
 % task
-Unfortunately, given two secure compiler passes that each preserve a possibly different security property, it is unclear what kind of security property the composition of those secure compiler passes preserve.
+Unfortunately, given two secure compiler passes that each preserve a possibly different security property, it is unclear what kind of security property is preserved by the composition of those secure compiler passes.
 %there is no way to tell what kind of security property will the composition of those secure compilers preserve.
@@
 % findings
-Starting from a general theory of property composition for security-relevant properties (such as the aforementioned ones), this paper formalises a secure multi-pass compiler that preserves those security-relevant properties.
+Starting from a general theory of property composition for security-relevant properties (such as the aforementioned ones), this paper formalises a theory of composition of secure compilers.
+Then, it showcases this theory a secure multi-pass compiler that preserves the aforementioned security-relevant properties.
 % conclusion
@@
 \gls*{cct} is a security property that ensures sensitive data is not leaked via timing side-channels~\cite{kocher1996timing}.
-Together, \gls*{sms}, \gls*{tms} and \gls*{scct}, a variant of \gls*{cct}, yield \gls*{msscct}, which is the gold standard of security properties for secure applications: programs attaining \gls*{msscct} do not leak sensitive data either through erroneous memory accesses, nor through timing side-channels.
-As discussed in \Cref{ex:strncpy}, these security properties can be enforced by compiler passes~\cite{bond2017vale,almeida2017jasmin}, to ensure programmers need not be aware of the various architectural details where their code will run.
+Together, \gls*{sms}, \gls*{tms} and \gls*{scct}, an enforceable overapproximation of \gls*{cct}, yield \gls*{msscct}, which is the gold standard of security properties for secure applications.
+Programs attaining \gls*{msscct} do not leak sensitive data either through erroneous memory accesses, nor through timing side-channels.
+As discussed in \Cref{ex:strncpy}, these security properties can be enforced by compiler passes~\cite{bond2017vale,almeida2017jasmin}, to ensure programmers need not be aware of the architectural details of where their code will run.
 
@@
 Robust compilers~\cite{abate2019jour} are a form of secure compilers that preserve security properties even in the presence of arbitrary attackers interacting with compiled code.
-Thus, robust compilers can be used to prevent, e.g., vulnerabilities resulting from uninitialised memory, e.g., by targeting capability-based languages such as CHERI~\cite{woodruff2014CHERI}, Arm Morello~\cite{arm-morello}, or MSWasm~\cite{michael2023mswasm}, where the compiler relies on capabilities to check that pointers are always initialised.
+Thus, robust compilers can be used to prevent vulnerabilities resulting from uninitialised memory (as well as many other ones), e.g., by targeting capability-based languages such as CHERI~\cite{woodruff2014CHERI}, Arm Morello~\cite{arm-morello}, or MSWasm~\cite{michael2023mswasm}, where the compiler relies on capabilities to check that pointers are always initialised.
 \end{example}
@@
         The resulting security property is \gls*{msscct}, i.e., the gold standard of security properties for secure programs~\cite{lemay2021ccc}.
+
+        % \MPpast{
+        % 	some of the good things here: 
+        % 	our definitions are amenable to sec comp proofs, unlike existing definitions
+        % 	MS def is like mswasm MS def, 
+        % 	but CCT is new
+        % }
 
@@
 this paper assumes that whole programs are the result of linking partial programs referred to as {\em context} ($\ctx$) and {\em component} ($\comp$).
-The context is an arbitrary program and thus has the role of an {\em attacker} that can interact with the component by means of whatever features the programming language has.
+The context is an arbitrary program and thus has the role of an {\em attacker} that can interact with the component by means of whatever features the programming language has, and the component is what is security-relevant.
 %In this work, the semantics of the programming language is expected to differentiate between the component and the context.
@@
 This translation is considered {\em correct} if it is semantics-preserving.
-That is, for a whole program $\src{\wholeprogvar}$, the compiler should relate the $\src{L}$ semantics of $\src{\wholeprogvar}$ with the semantics of $\trg{T}$ in such a way that they are ,,compatible''.
+That is, for a whole program $\src{\wholeprogvar}$, the compiler should relate the $\src{L}$ semantics of $\src{\wholeprogvar}$ with the semantics of $\trg{T}$ of the compiled counterpart of $\src{p}$ in such a way that they are ,,compatible''.
 Unfortunately, correct compilers may be insecure compilers~\cite{patrignani2019survey,kennedy2006secure.net,abadi1999protect,ahmed2018dagstuhl} and programs translated by insecure compilers can violate security properties that the programmer assumes to hold.
@@
 \begin{example}[Types]
-  Suppose $\src{L}$ is a statically typed language similar to \texttt{C} and $\trg{T}$ is dynamically typed, where both share the same syntax up to dynamic type checks.
+  Suppose $\src{L}$ is a statically-typed language similar to \texttt{C} and $\trg{T}$ is dynamically typed, where both share the same syntax up to dynamic type checks.
   Consider the following $\src{L}$ component and its compiled version below.
@@
 \end{example}
+
+% \MPpost{
+% 	i think the trace description of traces is more uindertandable,
+% 	teh monitor one is easier to work with
+% 	this is the justification for having both that we'll have to make in the future
+% }
 

### 2023-07-12 popl/main.tex

@@
 Given that $\msfilterL{\trg{\trace}}=\msfilterLtms{\src{\trace}}$, the proof goal changes from $\msfilterL{\trg{\trace}}\in\pi$ to $\msfilterLtms{\src{\trace}}\in\pi$.
-Then, by specializing the robust satisfaction assumption $\rsat{\src{\library_{\comp}}}{\pi}$ to use the context $\src{\library_{\ctx}}$ obtained from the backtranslation which yields
- % the context segments of the target trace $\trg{\trace}$, and the 
-the following source execution $\progstepto{\src{\prog{\library_{\ctx}}{\library_{\comp}}}}{\src{\runtimetermvar}}{\src{\trace}}$.
-This gives $\msfilterLtms{\src{\trace}}\in\pi$, which was the proof goal.
+This follows by specializing the robust satisfaction assumption $\rsat{\src{\library_{\comp}}}{\pi}$ to use the context $\src{\library_{\ctx}}$, which is obtained from the backtranslation, and to use the source execution $\progstepto{\src{\prog{\library_{\ctx}}{\library_{\comp}}}}{\src{\runtimetermvar}}{\src{\trace}}$.
 

### 2023-08-01 tech/main.tex
add tech/

--- 2023-07-12 popl/main.tex
@@
-%%
-%% This is file `sample-sigconf.tex',
-%% generated with the docstrip utility.
-%%
-%% The original source files were:
-%%
-%% samples.dtx  (with options: `sigconf')
-%% 
-%% IMPORTANT NOTICE:
-%% 
-%% For the copyright see the source file.
-%% 
-%% Any modified versions of this file must be renamed
-%% with new filenames distinct from sample-sigconf.tex.
-%% 
-%% For distribution of the original source see the terms
-%% for copying and modification in the file samples.dtx.
-%% 
-%% This generated file may be distributed as long as the
-%% original source files, as listed above, are part of the
-%% same distribution. (The sources need not necessarily be
-%% in the same archive or directory.)
-%%
-%% Commands for TeXCount
-%TC:macro \cite [option:text,text]
-%TC:macro \citep [option:text,text]
-%TC:macro \citet [option:text,text]
-%TC:envir table 0 1
-%TC:envir table* 0 1
-%TC:envir tabular [ignore] word
-%TC:envir displaymath 0 word
-%TC:envir math 0 word
-%TC:envir comment 0 0
-%%
-%%
-%% The first command in your LaTeX source must be the \documentclass command.
-
 % Fixing: Too many math alphabets used in version normal.
@@
 \newcommand\bmmax{0}
-\documentclass[utf8,acmsmall,review,screen,dvipsnames,anonymous]{acmart}
+\documentclass[utf8,acmsmall,review,screen,dvipsnames]{acmart}
 
@@
 
-%% Rights management information.  This information is sent to you
-%% when you complete the rights form.  These commands have SAMPLE
-%% values in them; it is your responsibility as an author to replace
-%% the commands and values with those provided to you when you
-%% complete the rights form.
-\setcopyright{acmcopyright}
-\copyrightyear{2024}
-\acmYear{2024}
-\acmDOI{XXXXXXX.XXXXXXX}
-
-%% These commands are for a PROCEEDINGS abstract or paper.
-\acmConference[POPL '24]{51st ACM SIGPLAN Symposium on Principles of Programming Languages}{January 17-19,
-  2024}{London, UK}
-%
-%  Uncomment \acmBooktitle if th title of the proceedings is different
-%  from ``Proceedings of ...''!
-%
-%\acmBooktitle{Woodstock '18: ACM Symposium on Neural Gaze Detection,
-%  June 03-05, 2018, Woodstock, NY}
-\acmPrice{15.00}
-\acmISBN{978-1-4503-XXXX-X/18/06}
-
-%%
-%% Submission ID.
-%% Use this when submitting an article to a sponsored event. You'll
-%% receive a unique submission ID from the organizers
-%% of the event, and this ID should be used as the parameter to this command.
-%%\acmSubmissionID{123-A56-BU3}
-
-%%
-%% For managing citations, it is recommended to use bibliography
-%% files in BibTeX format.
-%%
-%% You can then either use BibTeX with the ACM-Reference-Format style,
-%% or BibLaTeX with the acmnumeric or acmauthoryear sytles, that include
-%% support for advanced citation of software artefact from the
-%% biblatex-software package, also separately available on CTAN.
-%%
-%% Look at the sample-*-biblatex.tex files for templates showcasing
-%% the biblatex styles.
-%%
-
-%%
-%% The majority of ACM publications use numbered citations and
-%% references.  The command \citestyle{authoryear} switches to the
-%% "author year" style.
-%%
-%% If you are preparing content for an event
-%% sponsored by ACM SIGGRAPH, you must use the "author year" style of
-%% citations and references.
-%% Uncommenting
-%% the next command will enable that style.
 \citestyle{acmauthoryear}
 
-%%
-%% end of the preamble, start of the body of the document source.
 \begin{document}
 
-%%
-%% The "title" command has an optional parameter,
-%% allowing the author to define a "short title" to be used in page headers.
 \title{Secure Composition of Robust and Optimising Compilers}
-% Composing Secure and Optimising Compilers
-%
-
-%%
-%% The "author" command and its associated commands are used to define
-%% the authors and their affiliations.
-%% Of note is the shared affiliation of the first two authors, and the
-%% "authornote" and "authornotemark" commands
-%% used to denote shared contribution to the research.
-\author{Matthis Kruse}
-% \authornote{Both authors contributed equally to this research.}
-\email{matthis.kruse@cispa.de}
-\orcid{0000-0003-4062-9666}
-\affiliation{%
-  \institution{CISPA Helmholtz Center for Information Security and Saarland University}
-  \streetaddress{Stuhlsatzenhaus 5}
-  \city{Saarbr{\"u}cken}
-  \state{Saarland}
-  \country{Germany}
-  \postcode{66123}
-}
-
-\author{Michael Backes}
-\email{director@cispa.de}
-%\orcid{0000-0002-7130-9211}
-\affiliation{%
-  \institution{CISPA Helmholtz Center for Information Security}
-  \streetaddress{Stuhlsatzenhaus 5}
-  \city{Saarbr{\"u}cken}
-  \state{Saarland}
-  \country{Germany}
-  \postcode{66123}
-}
-
-\author{Marco Patrignani}
-\orcid{0000-0003-3411-9678}
-\email{marco.patrignani@unitn.it}
-\affiliation{%
-  \institution{University of Trento}
-  \streetaddress{Via Sommarive, 9}
-  \city{Povo}
-  \country{Italy}
-  \postcode{38123}
-}
-
-%%
-%% By default, the full list of authors will be used in the page
-%% headers. Often, this list is too long, and will overlap
-%% other information printed in the page headers. This command allows
-%% the author to define a more concise list
-%% of authors' names for this purpose.
-\renewcommand{\shortauthors}{Kruse, Backes, and Patrignani}
-
-%%
-%% The abstract is a short summary of the work to be presented in the
-%% article.
 \begin{abstract}
-% \MPpost{
-%   this abstract fails in setting up the robustness argument.
-%   it'll be a good exercise for you to take inspiration from this, but then write something as concise but with a more direct focus on robust properties
-% }
-% context
-To ensure that secure applications do not leak their secrets, they are required to uphold several security properties such as spatial and temporal memory safety as well as cryptographic constant time.
-% need
-Existing work shows how to enforce these properties individually, in an architecture-independent way, by using secure compiler passes that each focus on an individual property.
-% task
-Unfortunately, given two secure compiler passes that each preserve a possibly different security property, it is unclear what kind of security property the composition of those secure compiler passes preserve.
-%there is no way to tell what kind of security property will the composition of those secure compilers preserve.
-
-% object
-This paper is the first to study what security properties are preserved across the composition of different secure compiler passes.
-% findings
-Starting from a general theory of property composition for security-relevant properties (such as the aforementioned ones), this paper formalises a secure multi-pass compiler that preserves those security-relevant properties.
-% conclusion
-Crucially, this paper derives the security of the multi-pass compiler from the composition of the security properties preserved by its individual passes, which include security-preserving as well as optimisation passes.
-% 
-From an engineering perspective, this is the desirable approach to building secure compilers.
-
-  % \MKin{context}
-  % Memory safety necessitates secrecy, since breaking it can lead to an arbitrarily large attack space, i.e., from simply reading a secret up to alteration of control-flow in order to execute arbitrary code.
-  % However, secrecy cannot be achieved with just memory safety alone, since, e.g., private data could be leaked by differences in execution time for varying program inputs.
-  % This leakage is avoidable by ensuring that programs are cryptographic constant time, i.e., different inputs do not change execution time.
-  % Prior work has shown that strategies to ensure cryptographic constant time can be rendered useless, since a compiler may simply optimize them away.
-  % \MKin{need}
-  % This is why secure compilation is an integral ingredient to achieve security, because a secure compiler guarantees (formally) that properties that hold in the source, also hold in the target.
-%% mention attacker?
-the text describes correctness
-  % \MKin{task}
-  % Unfortunately, it is an open question whether secure compilers can be engineered in a modular way.
-  % Ideally, compiler engineers develop two different compilation passes, one that ensures memory safety and another one that is security preserving with respect to cryptographic constant time.
-%, because this style of engineering is well-established for its simplifying divide-and-conquer approach as well as enhanced reusability.
-  % \MKin{object+findings}
-  % This paper answers that question positively: The composition of compilers that are secure with respect to some security-properties yields the intersection of these properties.
-  % \MKin{object}
-  % This article discusses different variants of compositions of secure compilers and instantiates parts of these results in a case-study.
-  % To this end, it presents an optimizing, secure compilation chain that preserves both memory safety and cryptographic constant time.
-  % \MKin{conclusion}
-  % The compositionality results enable modular proofs for the presented, secure compiler. % that is secure with respect to both memory safety and cryptographic constant time.
-  % Some results of this work are implemented in the Coq proof assistant.
-
 \begin{center}\small\it
@@
 
-%%
-%% The code below is generated by the tool at http://dl.acm.org/ccs.cfm.
-%% Please copy and paste the code instead of the example below.
-%%
-\begin{CCSXML}
-<ccs2012>
-  <concept>
-  <concept_id>10002978.10002986.10002989</concept_id>
-  <concept_desc>Security and privacy~Formal security models</concept_desc>
-  <concept_significance>500</concept_significance>
-  </concept>
-</ccs2012>
-\end{CCSXML}
-\ccsdesc[500]{Security and privacy~Formal security models}
-
-%%
-%% Keywords. The author(s) should pick words that accurately describe
-%% the work being presented. Separate the keywords with commas.
-\keywords{Memory-safety, Secure Compilation, Privacy}
-
-%%
-%% This command processes the author and affiliation and title
-%% information and builds the first part of the formatted document.
 \maketitle
 
-\section{Introduction\pages{4}}\label{sec:introduction}
-
-% \gls*{ms} is a security property that ensures that there are no out-of-bounds accesses as well as no reads or writes to uninitialised memory.
-% Programs that do not attain memory safety, can easily have their private data be leaked to attackers or be corrupted~\cite{lemay2021ccc}.
-% Because of this, \gls*{ms} is a requirement for programs to attain an even stronger security property, namely \gls*{cct}.
-% Programs that attain \gls*{cct} are secure with respect to timing attacks, and this is the gold standard for secure applications, including cryptographic ones, since timing attacks can expose secrets via a simple performance analysis~\cite{kocher1996timing}.
-% Timing attacks have become more and more relevant recently, especially given the era of cloud-computing~\cite{aviram2010cloudtime,kumar2019cloudsecsurvey}, where programs share the same hardware but are supposed to run in a sandboxed environment that does not leak secrets.
-% Unfortunately, several attacks within such a cloud environment have been demonstrated as viable strategies to extract secrets~\cite{mehmet2015getoff,flowers2022zeroday,atya2019catchme,venkatanathan2015placevul}.
-
-% Even though mitigations for such attacks exist~\cite{bond2017vale,almeida2017jasmin}, some compilers may optimise these mitigations away~\cite{barthe2018sec}.
-% This is of course not what programmers expect and to ensure such mitigations persist in the compiled code, one can use a {\em secure} compilers.
-% More generally, secure compilers should preserve security properties even when the program runs in a hostile context such as the cloud one, i.e., where it interacts with other possibly malicious programs.
-% An example of such secure compilers are those that ascribe to the theory of robust compilation~\cite{abate2019jour,abate2021extacc,patrignani2021rsc}, which takes this interaction between components and potentially malicious contexts into account.
-
-% Unfortunately, it is unclear if and how robust compilation supports compositionality, and compilation passes are typically developed as a standalone module to be plugged into a compilation pipeline and composed with other passes.
-% Different kinds of compilation passes exist, e.g., there are usually several compilation passes that perform optimisations~\cite{lattner2004llvm,googlev8,androidstudio,kuepper2023cryptopt,manjikian1997fusion,wegman1991ccp} or source-code instrumentations~\cite{nagarakatte2009soft,nagarakatte2010cets,akritidis2009baggy,dhumbumroong2020boundwarden,jung2021pico,nam2019framer,shankaranarayana2023tailcheck,younan2010paricheck,zhou2023fatptrs,bond2017vale,almeida2017jasmin,kuepper2023cryptopt,cauligi2019fact} that enforce that the compiled program fulfils a property of interest.
-% Without a framework for composing secure compiler passes, it is not possible to enable separation of concerns, e.g., to have a secure compilation pass that ensures \gls*{ms} that is developed independently of another secure pass for \gls*{cct} as suggested before and as showcased in \Cref{ex:strncpy}.
-% \MPpost{
-%   this intro does a poor job at introducing the robust compilation argument.
-%   perhaps it'd be best to start from robust properties, which motivates the use of robust compilers?
-%   idk
-%   as for the abstract, this is a good starting point for you to make a better intro
-% }
-
-\gls*{ms} is a security property obtained by composing \gls*{sms}, which ensures array accesses are all within bounds, and \gls*{tms}, which ensures pointers are only used when they are valid~\cite{azevedo2018meaningsofms,jim2002cyclone,necula2005ccured,nagarakatte2010cets,nagarakatte2009soft,akritidis2009baggy,michael2023mswasm}.
-\gls*{cct} is a security property that ensures sensitive data is not leaked via timing side-channels~\cite{kocher1996timing}.
-Together, \gls*{sms}, \gls*{tms} and \gls*{scct}, a variant of \gls*{cct}, yield \gls*{msscct}, which is the gold standard of security properties for secure applications: programs attaining \gls*{msscct} do not leak sensitive data either through erroneous memory accesses, nor through timing side-channels.
-As discussed in \Cref{ex:strncpy}, these security properties can be enforced by compiler passes~\cite{bond2017vale,almeida2017jasmin}, to ensure programmers need not be aware of the various architectural details where their code will run.
-
-\begin{example}[strncpy]\label{ex:strncpy}
-Consider the \texttt{C} function \texttt{strncpy} that copies a null-terminated string \texttt{src} into \texttt{dst} up to a length of \texttt{n}.
-% Assume that the length of the array pointed by \texttt{dst} is at least \texttt{n}, and \texttt{src} is terminated by \texttt{'\textbackslash 0'}.
-% Even under these assumptions, there is a 
-This function is subject to a subtle \gls*{sms} vulnerability: The bounds check \texttt{i < n} should happen {\it before} the access to memory location \texttt{x[i]}: otherwise
-% If \texttt{n} is at least as large as the length of the array \texttt{src}, 
-the memory location past the last element will be leaked to an attacker.
-\begin{lstlisting}[language=c,basicstyle=\small\ttfamily,morekeywords={size_t}]
-void strncpy(size_t n, char *dst, char *src) {
-  for(size_t i = 0; src[i] != '\0' && i < n; ++i) {
-    dst[i] = src[i];
-  }
-}
-\end{lstlisting}
-
-To prevent this vulnerability, one can use a compilation pass that enforces \gls*{sms}, such as Softbounds~\cite{nagarakatte2009soft} or BaggyBounds~\cite{akritidis2009baggy}.
-
-Because of timing attacks, fixing \gls*{sms} is not enough to make \texttt{strncpy} secure.
-In fact, the loop can terminate early, as soon as the string-terminating character \texttt{'\textbackslash 0'} is encountered, thus making program execution time proportional to the length of the array pointed by \texttt{src}.
-% Often, programmers rewrite programs like these to have loops that run with a fixed number of iterations.
-Also in this case there exist compiler passes that can rewrite such programs into \gls*{cct} ones~\cite{cauligi2019fact}.
- % so that they are no longer vulnerable to timing attacks.
-
-Alas, code is not run in isolation, so a malicious attacker could supply code that intracts with \texttt{strncpy} and trigger a violation of either \gls*{ms} or \gls*{cct} by calling \texttt{strncpy} with an argument for \texttt{src} that points to uninitialised memory.
-This would, in turn, triggering a series of reads from uninitialised memory, which is an immediate \gls*{ms} violation with devastating real-world consequences~\cite{uninit-0,uninit-1,uninit-2,uninit-3,uninit-4}.
-% An easy fix to sidestep this is to use non-nullable pointers as parameters instead, e.g., as presented in Cyclone~\cite{jim2002cyclone}.
-
-Robust compilers~\cite{abate2019jour} are a form of secure compilers that preserve security properties even in the presence of arbitrary attackers interacting with compiled code.
-Thus, robust compilers can be used to prevent, e.g., vulnerabilities resulting from uninitialised memory, e.g., by targeting capability-based languages such as CHERI~\cite{woodruff2014CHERI}, Arm Morello~\cite{arm-morello}, or MSWasm~\cite{michael2023mswasm}, where the compiler relies on capabilities to check that pointers are always initialised.
-\end{example}
-
-% To ensure these compiler-preserved properties still hold even in the presence of arbitrary attackers interacting with compiled code, one can use secure compilers in the form of robust compilers~\cite{abate2019jour,abate2021extacc,patrignani2021rsc}.
-
-Unfortunately, given secure compiler passes that each preserve a possibly different security property, there is no way to tell what kind of security property will the composition of those secure compilers preserve.
-% Different kinds of compilation passes exist, e.g., there are usually several compilation passes that perform optimisations~\cite{lattner2004llvm,googlev8,androidstudio,kuepper2023cryptopt,manjikian1997fusion,wegman1991ccp} or source-code instrumentations~\cite{nagarakatte2009soft,nagarakatte2010cets,akritidis2009baggy,dhumbumroong2020boundwarden,jung2021pico,nam2019framer,shankaranarayana2023tailcheck,younan2010paricheck,zhou2023fatptrs,bond2017vale,almeida2017jasmin,kuepper2023cryptopt,cauligi2019fact} that enforce that the compiled program fulfils a property of interest.
-Worse, without a framework for composing secure compiler passes, it is not possible to enable separation of concerns, e.g., to have a secure compilation pass that ensures \gls*{ms} that is developed independently of another secure pass for \gls*{cct}, that is developed independently of other passes, such as optimisations.
-
-This paper introduces a framework for reasoning about the composition of secure and optimising compiler passes akin to those of \Cref{ex:strncpy} and it showcases the power of this framework by instantiating it on a multi-pass compilation chain.
-To this end, this paper first discusses how to compose security properties, such as \gls*{tms} and \gls*{sms} into \gls*{ms}, and then adding \gls*{scct} to the mix to obtain \gls*{msscct}.
-Then, this paper defines compiler composition and formalises that given two passes that securely preserve two (possibly distinct) properties, their composition securely preserves the composition of those properties.
-The paper then defines several secure compiler passes, where each is either preserving a different security property (\gls*{tms}, \gls*{sms}, \gls*{scct}) or performing a security-preserving optimisation, (e.g., applying \gls*{cf} or \gls*{dce}).
-Finally, this paper shows that composing these secure compiler passes into a multi-pass compilation chain results in the end-to-end preservation of \gls*{msscct}.
-Crucially, this paper derives the security of the multi-pass compiler from the composition of the security properties preserved by its individual passes.
-% Crucially, this end-to-end security preservation is obtained by composing the security property preserved by each individual pass, as dictated by the framework of this paper.
-This result showcases how the framework allows the kind of formal security reasoning that compiler writers already want (and already do), obtaining precise, compositional security reasoning while providing minimal (and modular) proof effort.
-
-In summary, this paper makes the following contributions:
-\begin{itemize}
-  \item %
-        %\Cref{sec:compprop} presents a formal framework to reason about compositions of security properties and demonstrate its usage with concrete examples:
-        This paper formalises security properties (\Cref{sec:compprop}) that are of interest for real-world compiler writers, namely \gls*{tms}, \gls*{sms} and \gls*{cct} (as identified by the plethora of work enforcing such properties individually~\cite{akritidis2009baggy,nagarakatte2009soft,nagarakatte2010cets,dhumbumroong2020boundwarden,jung2021pico,nam2019framer,shankaranarayana2023tailcheck,younan2010paricheck,zhou2023fatptrs,bond2017vale,almeida2017jasmin,kuepper2023cryptopt,cauligi2019fact}).
-        % 
-        Starting from ways to formalise those properties individually, this paper shows how to compose their formalisation.
-        % 
-        % This paper then showcases the benefits of this composition by composing \gls*{tms} and \gls*{sms} into \gls*{ms}, and ultimately adding \gls*{cct} to the composition, yielding the the conjunction of \gls*{ms} and \gls*{cct}.
-        % 
-        The resulting security property is \gls*{msscct}, i.e., the gold standard of security properties for secure programs~\cite{lemay2021ccc}.
-
-  \item %
-        %\Cref{sec:compcomp} studies different forms of secure compiler compositions.
-        This paper takes the secure compilation framework of~\citep{abate2019jour} and extends it to reason about the security of all different known forms of compiler composition (\Cref{sec:compcomp}).
-        %Concretely, it investigates upper and lower compositions, as well as the sequential composition.
-        For this, this paper studies sequential compiler composition as well as compilers with multiple input languages or multiple output ones, as used in existing compilation chains.
-        %
-        This paper proves that starting from two compilers that preserve two (possibly distinct) properties, their composition preserves the intersection of those properties.
-        % For example, a compiler that is secure with respect to \gls*{tms} can be composed with a compiler that is secure with respect to \gls*{sms}, resulting in a compiler that is secure with respect to \gls*{ms}.
-        % 
-        Finally, this paper proves that the order of composition of sequential compiler passes is irrelevant for the resulting security.
-        % 
-        This is crucial for reordering optimisation passes and thus generating secure and efficient code.
-
-  \item %
-        This paper presents a case-study showcasing the conjunction of the previous contributions (\Cref{sec:casestud:defs,sec:casestud:rtp}).
-        To this end, it presents a compilation chain consisting of several passes that ultimately preserves \gls*{msscct} by means of composing the individual, secure passes concerning \gls*{tms}, \gls*{sms}, and \gls*{scct}, respectively.
-        Furthermore, the chain includes two optimisation passes: One performs \gls*{dce} and the other \gls*{cf}.
-        The formalisation of this case study showcases the power of the presented framework: The divide-and-conquer approach to software engineering is a viable strategy even for the development of secure compilers.
-%        This demonstrates that the usage of the presented framework is both simpler and more expressive compared to one monolithic security proof of the whole compilation chain.
-
-  \item The key contributions of this paper are formalised in the Coq proof assistant and the paper indicates this with \CoqSymbol.
-\end{itemize}
-
-This paper starts by introducing relevant notions of security properties and secure compilation (\Cref{sec:background}),
-and discusses related work (\Cref{sec:relwork}) before concluding (\Cref{sec:concl}).
-
-\contribution{Open Source \& Technical Report} A technical report with the omitted formal details, lemmas and proofs, as well as the Coq formalisation are available as supplementary material.
-% \MKin{cite/link}
-
-\section{Background: Security Properties and Secure Compilers\pages{1}}\label{sec:background}
-
-To introduce the security argument of this paper, this section first presents the concepts of (security) properties, of their satisfaction, and of their robust satisfaction (i.e., satisfaction in the presence of an active attacker; \Cref{subsec:bg:tprop}).
-Then, borrowing from existing work~\cite{abate2019jour,abate2021extacc}, the section introduces secure compilers as compilers that preserve robust property satisfaction (\Cref{subsec:bg:rtp}).
-
-\subsection{Properties and (Robust) Satisfaction}\label{subsec:bg:tprop}
-
-This paper employs the security model where programs are written in a language whose semantics emits events $\event$.
-Events include security-relevant actions (e.g., reading from and writing to memory, as detailed in \Cref{sec:compprop}) and the unobservable event $\emptyevent$.
-As programs execute, their emitted events are concatenated in traces $\trace$, which serve as the description of the behaviour of a program.%
-\footnote{
-Throughout the paper, sequences are indicated with an overbar (i.e., $\trace$), empty sequences with $\hole{\cdot}$, and concatenation of sequences $\trace[_{1}],\trace[_{2}]$ as $\trace[_{1}]\cdot\trace[_{2}]$.
-Prepending elements to sequences uses the same notation: $\event\cdot\trace$.
-}
-
-Properties $\pi$ are sets of traces of admissible program behaviours, ascribing what said property considers valid.
-The set of all properties can be partitioned into different {\em classes} ($\class$), i.e., safety, liveness, and neither safety nor liveness~\cite{clarkson2008hyper}.
-A class is simply a set of properties and for the class of safety properties, it is decidable whether a trace satisfies a safety property with just a finite trace prefix.
-As an example, consider a trace describing an interaction with a memory where the deallocation of an address $\loc$ precedes a read at that address in memory: $\ev{Dealloc\ \loc}\cdot\ev{Read\ \loc\ 1729}\cdot\dots$.
-This program behaviour is insecure with respect to a canonical notion of (temporal) memory safety dictating no use-after-frees of pointers~\cite{nagarakatte2010cets,azevedo2018meaningsofms}, because it reads from a memory location that was freed already.
-The previous finite trace prefix is enough to decide that the trace does not satisfy \gls*{tms} and there is no way to append events to this prefix which would result in the trace being admissible.
-In the following, the execution of a whole program $\wholeprogvar$ that terminates in state $\runtimetermvar$ according to the language semantics and produces trace $\trace$ is written as $\progstepto{\wholeprogvar}{\runtimetermvar}{\trace}$.
-With this, property satisfaction is defined as follows:
-\bul{whole programs $\wholeprogvar$ satisfy a property $\pi$} iff \iul{$\wholeprogvar$ yields a trace $\trace$} such that \oul{$\trace$ satisifies $\pi$} (\Cref{def:propsat}).
-
-\begin{definition}[Property Satisfaction]\label{def:propsat}
-  \bul{$\sat{\progvar}{\pi}$}
-  $\isdef$
-  if \iul{$\progstepto{\wholeprogvar}{\runtimetermvar}{\trace}$},
-  then \oul{$\trace\in\pi$}.
-\end{definition}
-
-Property satisfaction is defined on whole programs, i.e., programs without missing definitions.
-Thus, from a security perspective, this considers only a passive attacker model, where the attacker observes the execution and, e.g., retrieves secrets from that.
-To consider a stronger model similarly to what existing work does~\cite{abate2019jour,abate2021extacc,maffeis2008code-carrying,gordon2003authenticity,fournet2007authorization,bengtson2011refine,backes2014uniontyps,michael2023mswasm,swasey2017robust,sammler2019benefits}, the concept of satisfaction can be extended with {\em robustness}.
-Robust satisfaction considers partial programs $\progvar$, i.e., components with missing imports, which cannot run until said imports are fulfilled.
-To remedy this, {\em linking} takes two partial programs $\progvar[_{1}],\progvar[_{2}]$ and produces a whole program $\wholeprogvar$, i.e., $\link{\progvar[_{1}]}{\progvar[_{2}]}=\wholeprogvar$.
-As typically done in works that consider the execution of partial programs~\cite{abate2019jour,devriese2018parametricity,patrignani2021rsc,korashy2021capableptrs,strydonck2019lincap,devriese2017modular,bowman2015noninterference,ahmed2011equivcps,patterson2017linkingtyps},
-this paper assumes that whole programs are the result of linking partial programs referred to as {\em context} ($\ctx$) and {\em component} ($\comp$).
-The context is an arbitrary program and thus has the role of an {\em attacker} that can interact with the component by means of whatever features the programming language has.
-%In this work, the semantics of the programming language is expected to differentiate between the component and the context.
-With this, \Thmref{def:propsat} can be extended as follows: for \bul{components $\progvar$ to robustly satisfy a property $\pi$}, take an \iul{attacker context $\contextvar$ and link it with $\progvar$}, \oul{the resulting whole program must satisfy $\pi$}.
-
-\begin{definition}[Robust Satisfaction]\label{def:proprsat}
-  \bul{$\rsat{\progvar}{\pi}$}
-  $\isdef$ \iul{$\forall \contextvar$, if $\link{\contextvar}{\progvar}=\wholeprogvar$}, then \oul{$\sat{\wholeprogvar}{\pi}$}.
-\end{definition}
-
-\begin{example}[Double Free in Bluetooth Subsystem]
-  Consider \texttt{CVE-2021-3564}~\cite{doublefree-bluetooth}, one of many submissions for a double-free vulnerability.
-  The vulnerability arises due to a race condition where the context-level function \texttt{hci\_cmd\_work} was not expected to behave maliciously, since it resides in the same source-code repository where the vulnerability occurs.
-  Nevertheless, the component-level code of \texttt{hci\_dev\_do\_open} is linked with \texttt{hci\_cmd\_work} and does not atomically check whether a pointer has been freed already:
-  Therefore, \texttt{hci\_dev\_do\_open} does not satisfy the no-double-frees property robustly, since there is an implementation for \texttt{hci\_cmd\_work} that leads to a violation of that property when linked with \texttt{hci\_dev\_do\_open}.
-\end{example}
-
-\subsection{Secure Compilers}\label{subsec:bg:rtp}
-
-A {\em compiler} ($\cc{\src{L}}{\trg{L}}$) translates syntactic descriptions of programs from a {\em source} ($\src{L}$) into a {\em target} ($\trg{L}$) programming language.
-This translation is considered {\em correct} if it is semantics-preserving.
-That is, for a whole program $\src{\wholeprogvar}$, the compiler should relate the $\src{L}$ semantics of $\src{\wholeprogvar}$ with the semantics of $\trg{T}$ in such a way that they are ,,compatible''.
-Unfortunately, correct compilers may be insecure compilers~\cite{patrignani2019survey,kennedy2006secure.net,abadi1999protect,ahmed2018dagstuhl} and programs translated by insecure compilers can violate security properties that the programmer assumes to hold.
-To define when a compiler is secure, this paper uses the robust compilation framework~\cite{abate2019jour}, which the following definition summarises.
-
-For \bul{compilers $\cc{\src{L}}{\trg{L}}$ to robustly preserve a class of properties $\class$}, if for any \rul{property $\pi$ of that class $\class$ and programs $\src{p}$ written in $\src{L}$} where \iul{$\src{\progvar}$ robustly satisfies $\pi$}, then \oul{the compilation of $\src{\progvar}$, $\cc{\src{L}}{\trg{L}}\left(\src{p}\right)$, must robustly satisfy $\pi$}.
-
-\begin{definition}[Robust Compilation]\label{def:rtp}
-  % Compiler $\cc{\src{L}}{\trg{L}}$ robustly preserves $\class$, 
-  \bul{$\rtp{\cc{\src{L}}{\trg{L}}}{\class}$}
-  %, iff 
-  $\isdef$
-  \rul{$\forall \left(\pi\in\class\right) \left(\src{p}\in\Ltms\right),$} if \iul{$\rsat{\src{\progvar}}{\pi}$}, then \oul{$\rsat{\cc{\src{L}}{\trg{L}}\left(\src{p}\right)}{\pi}$}.
-\end{definition}
-
-Note that a class of properties $\class$ can represent just one property $\pi$ by lifting~\cite{clarkson2008hyper} that property to sets of properties, i.e., use the powerset of $\pi$ instead of $\pi$ itself.
-Because of this, this paper writes $\rtp{\cc{\src{L}}{\trg{L}}}{\pi}$, even though $\pi$ is a property and not a class.
-
-\begin{example}[Types]
-  Suppose $\src{L}$ is a statically typed language similar to \texttt{C} and $\trg{T}$ is dynamically typed, where both share the same syntax up to dynamic type checks.
-  Consider the following $\src{L}$ component and its compiled version below.
-  \begin{center}
-  $
-  \begin{array}{l}
-    \src{\lfunction{foo}{(char* x, int\ n)}{\lifz{valid\_ptr(x, n, sizeof(char))}{x[0]}{-1}}} \\
-    \trg{\lfunction{foo}{(\phantom{char*\ }x, \phantom{int\ }n)}{\lifz{valid\_ptr(x, n, sizeof(char))}{x[0]}{-1}}}
-  \end{array}
-  $
-  \end{center}
-  While the compiler emits code that may look correct, the generated code does not check that the provided argument is of the right type.
-  Even though the pointer $\trg{x}$ is checked for validity, the context $\trg{foo((int*)y, 1)}$ is able to provoke a read out of bounds.
-  Suppose the component transferred control to the context and passed ownership of a $\trg{char}$ pointer $\trg{y}$ sized $\trg{1}$ cells, the context can now call the component again, casting this buffer to an $\trg{int*}$ prior to that call.
-  The pointer is valid for one $\trg{char}$-sized memory cell, as expected, but the actual read operation now returns $\trg{sizeof(int)}$ many bytes instead of just $\trg{sizeof(char)}$ many.
-  Thus, even if $\src{foo}$ may have been robust with respect to the \gls*{sms}, its compiled counterpart is not and therefore the compiler fails to attain \Cref{def:rtp}.
-\end{example}
-
-\section{Security Properties: Formalisation, Enforcement and Composition\pages{1.5}}\label{sec:compprop}
-
-This section introduces a trace model and uses it to define the key properties of interest for this paper: \gls*{tms}, \gls*{sms}, \gls*{ms}, and \gls*{scct} (\Cref{subsec:propdefs}).
-These properties are of practical importance (as mentioned in \Cref{sec:introduction}) and also of interest in the case study (\Cref{sec:casestud:defs,sec:casestud:rtp}) this paper presents later. 
-Lastly, for each of the key properties, this section introduces corresponding monitors (\Cref{subsec:monitors}) that check them.
-
-\subsection{Specification Trace Model}\label{subsec:propdefs}
-
-\begin{gather*}
-  \begin{aligned}
-  \mi{(Security\ Tag)}~\sigma\bnfdef&\ \lock \mid \unlock\hspace{0.5cm}
-  \mi{(Control\ Tag)}~\sandboxtag\bnfdef \ctx\mid\comp\hspace{0.5cm}
-  \mi{(Event)}~\event\bnfdef\ \emptyevent \mid \lightning \mid \event[_{b}];\sandboxtag;\sigma \\
-  \mi{(Pre\text{-}event)}&~\event[_{b}]\bnfdef\ \ev{Alloc\ \loc\ n} \mid \ev{Dealloc\ \loc} \mid \ev{Use\ \loc\ n} \mid \ev{Branch\ n} \mid \ev{Binop\ n} \\
-  \end{aligned}
-\end{gather*}
-
-The specification trace model defines events as either the empty event ($\emptyevent$), a crash ($\lightning$), or as tuples consisting of a pre-event, a control-tag, and a security-tag.
-The purpose of the model is to define key security properties of interest, such as \gls*{ms} or a stricter variant of cryptographic constant time.
-To this end, security-tags indicate whether an event contains sensitive information ($\lock$) or not ($\unlock$), while control-tags state whether the context ($\ctx$) or the component ($\comp$) are responsible for emitting the event.
-The latter is necessary to be able to ignore actions done by a spurious context that, e.g., immediately deallocates a memory location twice, thus violating \gls*{tms}~\cite{nagarakatte2010cets}.
-%Even though this could be used for blame~\cite{patrignani2023blame}, this is not what this paper is concerned about.
-Lastly, pre-events describe the actual kind of event that happened.
-One such kind is the allocation event ($\ev{Alloc\ \loc\ n}$) that fires whenever a program claims $n$ cells of memory and stores them at address $\loc$.
-Dually, deallocation ($\ev{Dealloc\ \loc}$) announces that the object at location $\loc$ is freed.
-These two events alone are enough to provide a partial description of \gls*{tms} by requiring that, e.g., there is only one deallocation event that carries a location $\loc$.
-To be able to express \gls*{sms}, there is also an event to describe reads from and writes to memory ($\ev{Use\ \loc\ n}$).
-Finally, for cryptographic code, there is a general guideline that secrets must not be visible on a trace.
-Moreover, an instruction whose timing is data-dependent must not have a secret as an operand.
-Typical operations with data-dependent timing are branches and certain binary operations, such as division\footnote{This is highly architecture-dependent, but division is an operation that serves as a classic example for a data-dependent timing instruction, e.g., \cite[p.~755]{arm-refman}.}.
-Both operations are also modelled in the specification trace model ($\ev{Branch\ n}$ and $\ev{Binop\ n}$).
-
-\subsubsection{Temporal Memory Safety}
-
-\gls*{tms} is a safety property that describes that an unallocated object must not be used in any way.
-Moreover, the property requires that all allocated objects must be deallocated at some point.
-
-\begin{definition}[\glsfirst*{tms}]\label{def:trace:tmsdef}
-  $$
-  \tmssafe:=\left\{\trace \left| \begin{array}{rcl}
-                              \ev{Alloc\ \loc\ n;\sandboxtag;\securitytag}&\le_{\trace}&\ev{Dealloc\ \loc;\sandboxtag;\securitytag'} \\
-                              \ev{Use\ \loc\ n;\sandboxtag;\securitytag}&\le_{\trace}&\ev{Dealloc\ \loc;\sandboxtag;\securitytag'} \\
-                              \text{if }\ev{Alloc\ \loc\ n;\sandboxtag;\securitytag}\text{ in }\trace&\text{then}&\ev{Dealloc\ \loc;\sandboxtag;\securitytag'}\text{ in }\trace \\
-                              \text{at most one }\ev{Dealloc\ \loc;\sandboxtag;\securitytag}&\text{in}&\trace \\
-                              \text{at most one }\ev{Alloc\ \loc\ n;\sandboxtag;\securitytag}&\text{in}&\trace \\
-                            \end{array}\right.\right\}
-  $$
-\end{definition}
-Hereby, the notation $\event[_{1}]\le_{\trace}\event[_{2}]$ means that if $\event[_{1}]$ is in $\trace$ and if $\event[_{2}]$ is in $\trace$, then $\event[_{1}]$ appears before $\event[_{2}]$.
-
-\subsubsection{Spatial Memory Safety}
-
-\gls*{sms} prohibits out-of-bounds accesses:
-
-\begin{definition}[\glsfirst*{sms}]\label{def:trace:smsdef}
-  $$
-  \smssafe:=\left\{\trace \left|\begin{array}{rcl}
-                                  \text{If }\ev{Alloc\ \loc\ n;\sandboxtag;\securitytag}\le_{\trace}\ev{Use\ \loc\ m;\sandboxtag;\securitytag'}, \text{ then }m<n
-                                \end{array}\right.\right\}
-  $$
-\end{definition}
-
-\subsubsection{Memory Safety}
-
-Full \gls*{ms} (similar to earlier work~\cite{nagarakatte2009soft,nagarakatte2010cets,jim2002cyclone,necula2005ccured,michael2023mswasm}) is then described as the conjunction of \Cref{def:trace:tmsdef,def:trace:smsdef}.
-%%% Catalin's paper doesn't apply, since their definition of MS relies on noninterference to guarantee data isolation.
-%%% our definition does not guarantee that (on purpose, we want to compose it with cct)
-Note, however, that this definition says nothing about memory-safety issues introduced by side-channels, such as speculation.
-
-\begin{definition}[\glsfirst*{ms}]\label{def:trace:msdef}
-  $$
-  \mssafe:=\tmssafe \cap \smssafe
-  $$
-\end{definition}
-
-\subsubsection{Strict Cryptographic Constant Time}
-
-\gls*{cct} is a hypersafety property~\cite{barthe2018sec} and, thus, difficult to check with monitors.
-This is because, intuitively, hypersafety properties can relate multiple execution traces with eachother, but monitors work on a single execution.
-To sidestep this issue, this section defines the property \gls*{scct}, a stricter variant of \gls*{cct} that enforces the policy that no secret appears on a trace (inspired by earlier work~\cite{almeida2017jasmin}).
-
-\begin{definition}[\glsfirst*{scct}]\label{def:trace:scctdef}
-  $$
-  \scctsafe:=\left\{\trace \left|\begin{array}{rcl}
-                                   \trace=\hole{\cdot} & \text{ or } & \trace=\event[_{b}];\sandboxtag;\unlock\cdot\trace['] \wedge \trace[']\in\scctsafe
-                                 \end{array}\right.\right\}
-  $$
-\end{definition}
-
-\subsubsection{Memory Safe, Strict Cryptographic Constant Time}
-
-The combination of \gls*{ms} and \gls*{scct} is the intersection of these properties, \gls*{msscct}.
-Since \gls*{ms} and \gls*{scct} are just sets of traces that, intuitively, contain all program behaviors that follow a security policy, the intersection of them contains all program behaviors that follow both security policies, i.e., it entails all program behaviours that are both \gls*{ms} and \gls*{scct}.
-
-\begin{definition}[\glsfirst*{ms} and \gls*{scct}]\label{def:trace:msscctdef}
-  $$
-  \msscctsafe:=\mssafe\cap\scctsafe
-  $$
-\end{definition}
-
-%\section{Monitors and Trace Agreements\pages{1.5}}\label{subsec:monitors}
-\subsection{Monitors\pages{1.5}}\label{subsec:monitors}
-
-Monitors enforce safety properties by accepting or rejecting traces, i.e., if it rejects a trace, the trace does not satisfy the property the monitor checks.
-Since reasoning on monitors is easier than directly on just traces, this section presents a monitor for each of the previously shown safety properties (\Cref{subsec:propdefs}).
-%\MPpost{
-%	this is fine, but it'd be nice to provide more insights on why it's easier
-%}
-To lessen the burden when proving that a monitor accepts the trace of a program execution, each monitor uses a custom trace model that contains only the relevant information related to the property the monitor checks.
-To go from specification traces $\trace$ to monitor-level traces $\abstrace$, each property $\pi$ has an associated event agreement relation $\traceagree[]{\pi}{\event}{\absevent}$.
-\Cref{fig:stuttereq} shows how the event agreement is lifted to traces.
-\begin{figure}[!ht]
-  \judgbox{\traceagree{\pi}{\trace}{\abstrace}}{,,Specification-level trace $\trace$ agrees with monitor-level trace $\abstrace$ with respect to property $\pi$.''}
-  %
-  \typerule{traceagree-empty}{}{
-    \traceagree{\pi}{\hole{\cdot}}{\hole{\cdot}}
-  }{traceagree-empty}
-  %
-  \typerule{traceagree-ign-L}{
-    \traceagree{\pi}{\trace}{\abstrace}
-  }{
-    \traceagree{\pi}{\emptyevent\cdot\trace}{\abstrace}
-  }{traceagree-ign-L}
-  %
-  \typerule{traceagree-ign-R}{
-    \traceagree{\pi}{\trace}{\abstrace}
-  }{
-    \traceagree{\pi}{\trace}{\absev{\emptyevent}\cdot\abstrace}
-  }{traceagree-ign-R}
-  %
-  \typerule{traceagree-cons}{
-    \traceagree[]{\pi}{\event}{\absevent} &
-    \traceagree{\pi}{\trace}{\abstrace}
-  }{
-    \traceagree{\pi}{\event\cdot\trace}{\absevent\cdot\abstrace}
-  }{traceagree-cons}
-  \caption{Trace-Agreement relation that equates specification-level traces with monitor-level traces.}\label{fig:stuttereq}
-\end{figure}
-The trace agreement is the same for all properties $\pi$ up to the use of the event agreement in \Cref{tr:traceagree-cons}.
-With agreements, this section defines monitor satisfaction for traces and then it proves that monitor satisfaction implies property satisfaction.
-To this end, monitor satisfaction is defined as follows.
-\bul{A specification trace $\trace$ monitor-satisfies property $\pi$} iff there exists a (final) monitor state $\monitor$ and an abstract trace $\abstrace$ such that the \iul{specification trace $\trace$ agrees with abstract trace $\abstrace$} and \oul{the initial monitor\footnote{In this paper, for all monitors, the initial monitor state is denoted as $\emptyset$.} can step to the (final) monitor state $\monitor$ with abstract trace $\abstrace$}.
-
-\begin{definition}[Monitor Satisfaction]\label{def:monsat}
-  \bul{$\monsafe{\trace}{\pi}$} $\isdef$ $\exists \abstrace\ \monitor$, \iul{$\traceagree{\pi}{\trace}{\abstrace}$} and \oul{$\monitorcheck{\emptyset}{\monitor}{\abstrace}$}.
-\end{definition}
-
-\subsubsection{Monitor for \glsfirst*{tms}}\label{subsubsec:tms}
-\begin{gather*}
-  \begin{aligned}
-    \mi{(Abstract\ Store)}~\tmsmonitor\bnfdef&\ \left\{\text{allocated}:L\times\sandboxtag,\text{freed}:L\times\sandboxtag\right\} \hspace{0.33cm}%
-    \emptyset:=\ \left\{\text{allocated}:\emptyset,\text{freed}:\emptyset\right\}\\
-    \mi{(Abstract\ Events)}~\absevent\bnfdef&\ \absev{\emptyevent} \mid \absev{Alloc\ }\loc\ \sandboxtag \mid \absev{Dealloc\ }\loc\ \sandboxtag \mid \absev{Use\ }\loc\ \sandboxtag \mid \absterm\\
-  \end{aligned}
-\end{gather*}
-\begin{center}
-  \judgbox{\monitorcheck[]{\tmsmonitor}{\tmsmonitor[']}{\absevent}}{,,Monitor $\tmsmonitor$ does one step to $\tmsmonitor[']$ given event $\absevent$.''}$\;$\\
-  %
-% \typerule{tms-uninteresting}{
-% }{
-%   \monitorcheck[]{\tmsmonitor}{\tmsmonitor}{\absev{\emptyevent}}
-% }{tms-uninteresting}
-% %
-% \typerule{tms-abort}{
-% }{
-%   \monitorcheck[]{\tmsmonitor}{\tmsmonitor}{\absterm}
-% }{tms-abort}
-  %
-  \typerule{tms-use}{
-    (\loc;\sandboxtag)\in\tmsmonitor.\text{allocated} &
-    (\loc;\sandboxtag)\notin\tmsmonitor.\text{freed}
-  }{
-    \monitorcheck[]{\tmsmonitor}{\tmsmonitor}{\absev{Use\ }\loc\ \sandboxtag}
-  }{tms-use}
-  %
-  \typerule{tms-alloc}{
-    (\loc;\sandboxtag)\notin\tmsmonitor.\text{allocated} &
-    (\loc;\sandboxtag)\notin\tmsmonitor.\text{freed} \\
-    \tmsmonitor[']=\left\{\text{allocated}: \tmsmonitor.\text{allocated}\cup\left\{(\loc;\sandboxtag)\right\},\text{freed}: \tmsmonitor.\text{freed}\right\}
-  }{
-    \monitorcheck[]{\tmsmonitor}{\tmsmonitor[']}{\absev{Alloc\ }\loc\ \sandboxtag}
-  }{tms-alloc}
-  %
-  \typerule{tms-dealloc}{
-    (\loc;\sandboxtag)\in\tmsmonitor.\text{allocated} &
-    (\loc;\sandboxtag)\notin\tmsmonitor.\text{freed} \\
-    \tmsmonitor[']=\left\{\text{allocated}: \tmsmonitor.\text{allocated}\setminus\left\{(\loc;\sandboxtag)\right\},\text{freed}: \tmsmonitor.\text{freed}\cup\left\{(\loc;\sandboxtag)\right\}\right\}
-  }{
-    \monitorcheck[]{\tmsmonitor}{\tmsmonitor[']}{\absev{Dealloc\ }\loc\ \sandboxtag}
-  }{tms-dealloc}
-\end{center}
-
-For \gls*{tms}, the state of the monitor is a record with two sets keeping track of allocated and deallocated locations.
-%The monitor runs on a modified trace semantics with more abstract $\absev{Alloc\ }\loc$,$\absev{Dealloc\ }\loc$,$\absev{Use\ }\loc$, and $\absterm$ events, which represent {\em allocation}, {\em deallocation}, and {\em use} of a location $\loc$ as well as abnormal program termination.
-\Cref{tr:tms-use} simply requires that a location is (i) allocated and (ii) not freed.
-\Cref{tr:tms-alloc,tr:tms-dealloc} both require a location to not be freed already and extend the monitor state accordingly.
-This restriction effectively disallows reallocation to reassign the same location to an object.
-However, the definition can easily be adapted by, e.g., attaching a natural number serving as a counter.
-Contrary to other monitors in this paper, the multi-step relation of the \gls*{tms} monitor is non-standard:
-\begin{center}
-  \judgbox{\monitorcheck{\tmsmonitor}{\tmsmonitor[']}{\abstrace}}{,,Monitor $\tmsmonitor$ multi-steps to $\tmsmonitor[']$ given trace $\abstrace$.''}$\;$\\
-  %
-  \typerule{tms-refl}{
-    \tmsmonitor.\text{allocated}=\emptyset
-  }{
-    \monitorcheck{\tmsmonitor}{\tmsmonitor}{\hole{\cdot}}
-  }{tms-refl}
-  %
-  \typerule{tms-ign-trans}{
-    \monitorcheck{\tmsmonitor}{\tmsmonitor[']}{\abstrace}
-  }{
-    \monitorcheck[]{\tmsmonitor}{\tmsmonitor[']}{\absev{\emptyevent}\cdot\abstrace}
-  }{tms-ign-trans}
-  %
-  \typerule{tms-trans}{
-    \monitorcheck[]{\tmsmonitor}{\tmsmonitor[']}{\absevent}
-    \monitorcheck{\tmsmonitor[']}{\tmsmonitor['']}{\abstrace}
-  }{
-    \monitorcheck[]{\tmsmonitor}{\tmsmonitor['']}{\absevent\cdot\abstrace}
-  }{tms-trans}
-\end{center}
-\Cref{tr:tms-ign-trans,tr:tms-trans} are the same for all monitors, but \Cref{tr:tms-refl} has, in this case, an additional premise that no more locations should be allocated.
-This rejects the behavior of programs that forget to free memory.
-
-\begin{center}
-  \judgbox{\tmstraceagree[]{\event}{\absevent}}{,,Abstract event $\absevent$ is equivalent to $\event$ with respect to \gls*{tms}.''}$\;$\\
-  %
-  \typerule{tms-alloc-authentic}{
-  }{
-    \tmstraceagree[]{\ev{Alloc}\ \loc\ n;\sandboxtag;\securitytag}{\absev{Alloc\ }\loc\ \sandboxtag}
-  }{tms-alloc-authentic}
-  %
-% \typerule{tms-dealloc-authentic}{
-% }{
-%   \tmstraceagree[]{\ev{Dealloc}\ \loc;\sandboxtag;\securitytag}{\absev{Dealloc\ }\loc}
-% }{tms-dealloc-authentic}
-  %
-% \typerule{tms-use-authentic}{
-% }{
-%   \tmstraceagree[]{\ev{Use}\ \loc\ n;\sandboxtag;\securitytag}{\absev{Use\ }\loc}
-% }{tms-use-authentic}
-  %
-  \typerule{tms-branch-authentic}{
-  }{
-    \tmstraceagree[]{\ev{Branch\ }n}{\absev{\emptyevent}}
-  }{tms-branch-authentic}
-  %
-% \typerule{tms-binop-authentic}{
-% }{
-%   \tmstraceagree[]{\ev{Binop\ }n}{\absev{\emptyevent}}
-% }{tms-binop-authentic}
-% %
-% \typerule{tms-empty-authentic}{
-% }{
-%   \tmstraceagree[]{\emptyevent}{\absev{\emptyevent}}
-% }{tms-empty-authentic}
-% %
-  \typerule{tms-abort-authentic}{
-  }{
-    \tmstraceagree[]{\lightning}{\absterm}
-  }{tms-abort-authentic}
-\end{center}
-
-The trace agreement is entirely straightforward, so only allocation, branch, and crash are shown.
-
-\begin{lemma}[Traces with Monitor Satisfaction are $\tmssafe$]\label{lem:mon:tmssafe}
-  If $\montmssafe{\trace}$, then $\trace\in\tmssafe$. %\Coqed
-\end{lemma}
-
-\begin{example}[A program not satisfying \gls*{tms}]\label{ex:strncpy:cpp11}
-  Consider the following \texttt{C++11} library that calls \texttt{strncpy} (\Cref{ex:strncpy}) and prints the result to the standard output stream.
-
-  \begin{lstlisting}[language=c++,basicstyle=\small\ttfamily,mathescape,commentstyle=\color{gray}]
-int greet() {  // allocates 12 chars containing a greeting message
-  char* greetings = new char[12] { "Hello POPL!" }; // <- address ${\color{gray}\loc_{x}}$
-  char* to = new char[12];                          // <- address ${\color{gray}\loc_{y}}$
-
-  strncpy(12, to, greetings);
-  delete to;
-
-  printf("%cOPL\n", to[6]);
-}
-  \end{lstlisting}
-  Up to the body of \texttt{printf}, the program execution yields the specification trace $\ev{Alloc}\ \loc[_{x}]\ 12;\comp;\unlock\cdot\ev{Alloc}\ \loc[_{y}]\ 12;\comp;\unlock\cdot\ev{Use}\ \loc_{x}\ 0;\comp;\unlock\cdot\ev{Use}\ \loc_{x}\ 0;\comp;\unlock\cdot\ev{Use}\ \loc_{y}\ 0;\comp;\unlock\cdot\ev{Use}\ \loc_{x}\ 1;\comp;\unlock\cdot\ldots\cdot\ev{Use}\ \loc_{x}\ 12;\comp;\unlock\cdot\ev{Dealloc}\ \loc[_{y}];\comp;\unlock\cdot\ev{Use}\ \loc_{y}\ 6;\comp;\unlock$.
-  Relating this trace to abstract monitor events yields $\abstrace=\absev{Alloc}\ \loc[_{x}]\ \comp\cdot\absev{Alloc}\ \loc[_{y}]\ \comp\cdot\absev{Use}\ \loc[_{x}]\ \comp\cdot\absev{Use}\ \loc[_{x}]\ \comp\cdot\absev{Use}\ \loc[_{y}]\ \comp\cdot\ldots\cdot\absev{Use}\ \loc[_{x}]\ \comp\cdot\absev{Dealloc}\ \loc[_{y}]\ \comp\cdot\absev{Use}\ \loc[_{y}]\ \comp$.
-  Remembering the definition of \texttt{strncpy} (\Cref{ex:strncpy}), observe that it does not deallocate its arguments.
-  Even though the trace contains an out-of-bounds access right before returning from \texttt{strncpy}, this is no concern for \gls*{tms}, since the location $\loc[_{x}]$ is still allocated.
-  However, having returned from \texttt{strncpy}, the \texttt{greet} function continues and deallocates $\loc[_{y}]$ whose subsequent use in the \texttt{printf} call is a use-after-free bug.
-
-  The fix would be to \lstinline[language=c++,basicstyle=\small\ttfamily]|delete greetings| instead of \texttt{to} and add a \lstinline[language=c++,basicstyle=\small\ttfamily]|delete to| after the \texttt{printf} call, which leads to the abstract monitor trace $\abstrace[']=\absev{Alloc}\ \loc[_{x}]\ \comp\cdot\absev{Alloc}\ \loc[_{y}]\ \comp\cdot\absev{Use}\ \loc[_{x}]\ \comp\cdot\absev{Use}\ \loc[_{x}]\ \comp\cdot\absev{Use}\ \loc[_{y}]\ \comp\cdot\ldots\cdot\absev{Use}\ \loc[_{x}]\ \comp\cdot\absev{Dealloc}\ \loc[_{x}]\ \comp\cdot\absev{Use}\ \loc[_{y}]\ \comp$.
-  It follows that $\montmssafe{\abstrace[']}$ and from \Thmref{lem:mon:tmssafe}, it follows that the program satisfies \Thmref{def:trace:tmsdef}, even though the program still violates \gls*{sms}.
-\end{example}
-
-\subsubsection{Monitor for \glsfirst*{sms}}\label{subsubsec:sms}
-\begin{gather*}
-  \begin{aligned}
-    \mi{(Abstract\ Store)}~\smsmonitor:=\ L\times\sandboxtag\times\mb{N}\hspace{0.33cm}&
-    \mi{(Abstract\ Events)}~\absevent\bnfdef\ \absev{\varepsilon} \mid \absev{Alloc\ }\loc\ \sandboxtag\ n \mid \absev{Use\ }\loc\ \sandboxtag\ n \\
-  \end{aligned}
-\end{gather*}
-\begin{center}
-  \judgbox{\monitorcheck[]{\smsmonitor}{\smsmonitor[']}{\absevent}}{,,Monitor $\smsmonitor$ does one step to $\smsmonitor[']$ given event $\absevent$.''}$\;$\\
-  %
-% \typerule{sms-uninteresting}{
-% }{
-%   \monitorcheck[]{\smsmonitor}{\smsmonitor}{\absev{\emptyevent}}
-% }{sms-uninteresting}
-% %
-% \typerule{sms-abort}{
-% }{
-%   \monitorcheck[]{\smsmonitor}{\emptyset}{\absterm}
-% }{sms-abort}
-  %
-  \typerule{sms-use}{
-    (\loc;\sandboxtag;m)\in\smsmonitor &
-    n<m
-  }{
-    \monitorcheck[]{\smsmonitor}{\smsmonitor}{\absev{Use\ }\loc\ \sandboxtag\ n}
-  }{sms-use}
-  %
-  \typerule{sms-alloc}{
-    (\loc;\sandboxtag;m)\notin\smsmonitor
-  }{
-    \monitorcheck[]{\smsmonitor}{\smsmonitor\cup\left\{(\loc;\sandboxtag;n)\right\}}{\absev{Alloc\ }\loc\ \sandboxtag\ n}
-  }{sms-alloc}
-\end{center}
-The state of the monitor for \gls*{sms} is a set containing tuples of locations, control-tags, and the allocation size.
-In comparison to the trace model of the \gls*{tms} monitor, the trace model here is extended by sizing and positional information.
-\Cref{tr:sms-use} performs a bounds check and \Cref{tr:sms-alloc} adds bounds information to the state of the monitor.
-The trace agreement is entirely straightforward and similar to the one for \gls*{tms}.
-% \begin{center}
-%   \judgbox{\smstraceagree[]{\event}{\absevent}}{,,Abstract event $\absevent$ is equivalent to $\event$ with respect to \gls*{sms}.''}$\;$\\
-%   %
-%   \typerule{sms-alloc-authentic}{
-%   }{
-%     \smstraceagree[]{\ev{Alloc}\ \loc\ n;\sandboxtag;\securitytag}{\absev{Alloc\ }\loc\ n}
-%   }{sms-alloc-authentic}
-%   %
-%   \typerule{sms-dealloc-authentic}{
-%   }{
-%     \smstraceagree[]{\ev{Dealloc}\ \loc;\sandboxtag;\securitytag}{\absev{Dealloc\ }\loc}
-%   }{sms-dealloc-authentic}
-%   %
-%   \typerule{sms-use-authentic}{
-%   }{
-%     \smstraceagree[]{\ev{Use}\ \loc\ n;\sandboxtag;\securitytag}{\absev{Use\ }\loc\ n}
-%   }{sms-use-authentic}
-%   %
-%   \typerule{sms-branch-authentic}{
-%   }{
-%     \smstraceagree[]{\ev{Branch\ }n}{\absev{\emptyevent}}
-%   }{sms-branch-authentic}
-%   %
-%   \typerule{sms-binop-authentic}{
-%   }{
-%     \smstraceagree[]{\ev{Binop\ }n}{\absev{\emptyevent}}
-%   }{sms-binop-authentic}
-%   %
-%   \typerule{sms-empty-authentic}{
-%   }{
-%     \smstraceagree[]{\emptyevent}{\absev{\emptyevent}}
-%   }{sms-empty-authentic}
-%   %
-%   \typerule{sms-abort-authentic}{
-%   }{
-%     \smstraceagree[]{\lightning}{\absterm}
-%   }{sms-abort-authentic}
-% \end{center}
-
-\begin{lemma}[Traces with Monitor Satisfaction are $\smssafe$]\label{lem:mon:smsafe}
-  If $\monsmssafe{\trace}$, then $\trace\in\smssafe$.\Coqed
-\end{lemma}
-
-\begin{example}[Normal invocation of \texttt{strncpy}]\label{ex:strncpy:sms}
-  Consider the insecure \texttt{strncpy} function from \Cref{ex:strncpy} with a context \texttt{strncpy(2, x, y)}, where \texttt{x} and \texttt{y} are pointers to valid regions of memory with allocated space for exactly two cells and do not contain the null-terminating character \texttt{'\textbackslash 0'}.
-  For the sake of this example, the pointers have been allocated by the component and passed to the context.
-  The loop of \texttt{strncpy} will copy exactly two cells and then check the loop condition for the last time.
-  At that stage, the induction variable \texttt{i} is equal to \texttt{2} and, unfortunately, the order of checks is such that first the cell \texttt{x[i]} is read prior to bounds checking \texttt{i < n}.
-  Because of this, there is an out-of-bounds memory access right before exiting the function.
-  This is also visible on the trace, which can be sketched as $\ldots\cdot\ev{Alloc}\ \loc_{x}\ 2;\comp;\unlock\cdot\ldots\cdot\ev{Alloc}\ \loc_{y}\ 2;\comp;\unlock\cdot\ldots\cdot\ev{Use}\ \loc_{x}\ 0;\comp;\unlock\cdot\ev{Use}\ \loc_{y}\ 0;\comp;\unlock\cdot\ev{Use}\ \loc_{x}\ 1;\comp;\unlock\cdot\ev{Use}\ \loc_{y}\ 1;\comp;\unlock\cdot\ev{Use}\ \loc_{x}\ 2;\comp;\unlock\cdot\ldots$, where $\loc_{x}$ and $\loc_{y}$ are the memory addresses associated to \texttt{x} and \texttt{y}, respectively.
-  Omitting the events for all ,,$\ldots$'' for sake of brevity, the abstract monitor trace of this is $\absev{Alloc}\ \loc_{x}\ \comp\ 2\cdot\absev{Alloc}\ \loc_{y}\ \comp\ 2\cdot\absev{Use}\ \loc_{x}\ \comp\ 0\cdot\absev{Use}\ \loc_{y}\ \comp\ 0\cdot\absev{Use}\ \loc_{x}\ \comp\ 1\cdot\absev{Use}\ \loc_{y}\ \comp\ 1\cdot\absev{Use}\ \loc_{x}\ \comp\ 2$.
-
-  After the allocation events, the state of the monitor is $\left\{(\loc_{x};\comp;2),(\loc_{y};\comp;2)\right\}$.
-  All uses up to the last are accepted by the monitor, but the last event does not satisfy the premise $2<2$ in \Cref{tr:sms-use}.
-  Therefore, the whole program (\texttt{strncpy} linked with this kind of context) is not \gls*{sms}.
-\end{example}
-
-\subsubsection{Combining \glsfirst*{tms} and \glsfirst*{sms} Monitors to obtain \glsfirst*{ms}}\label{subsubsec:ms}
-
-\begin{center}
-  \judgbox{\monitorcheck[]{\msmonitor}{\msmonitor[']}{\absevent}}{,,Monitor $\msmonitor$ does one step to $\msmonitor[']$ given event $\absevent$.''}$\;$\\
-  \typerule{ms-step}{
-    \monitorcheck[]{\tmsmonitor}{\tmsmonitor[']}{\absevent_{\tmssafe}} &
-    \monitorcheck[]{\smsmonitor}{\smsmonitor[']}{\absevent_{\smssafe}} &
-  }{
-    \monitorcheck[]{(\tmsmonitor,\smsmonitor)}{(\tmsmonitor['],\smsmonitor['])}{(\absevent_{\tmssafe},\absevent_{\smssafe})}
-  }{ms-step}
-\end{center}
-The combined monitor runs the one for \gls*{tms} and the one for \gls*{sms} in a lockstep.
-The trace agreement similarly just relates a specification event with an abstract \gls*{tms}-event $\abstrace[_{\tmssafe}]$ and with an abstract \gls*{sms}-event $\abstrace[_{\smssafe}]$.
-% \begin{center}
-%   \judgbox{\mstraceagree{\event}{\absevent}}{,,Abstract event $\absevent$ is equivalent to $\event$ with respect to \gls*{ms}.''}$\;$\\
-%   %
-%   \typerule{ms-authentic}{
-%     \tmstraceagree[]{\event}{\absevent_{\tmssafe}} &
-%     \smstraceagree[]{\event}{\absevent_{\smssafe}}
-%   }{
-%     \mstraceagree[]{\event}{(\absevent_{\tmssafe},\absevent_{\smssafe})}
-%   }{ms-authentic}
-% \end{center}
-
-\begin{lemma}[Traces with Monitor Satisfaction are $\mssafe$]\label{lem:mon:msafe}
-  If $\monmssafe{\trace}$, then $\trace\in\mssafe$.\Coqed
-\end{lemma}
-
-\subsubsection{Monitor for \glsfirst*{scct}}
-\begin{gather*}
-  \begin{aligned}
-    \mi{(Abstract\ Store)}~\scctmonitor&:=\ \emptyset \hspace{0.5cm}
-    \mi{(Abstract\ Events)}~\absevent:=\ \absev{\varepsilon} \mid \absterm \mid \absev{Any} \\
-  \end{aligned}
-\end{gather*}
-\begin{center}
-  \judgbox{\monitorcheck[]{\scctmonitor}{\scctmonitor[']}{\absevent}}{,,Monitor $\scctmonitor$ does one step to $\scctmonitor[']$ given event $\absevent$.''}$\;$\\
-  %
-  \typerule{scct-none}{
-  }{
-    \monitorcheck[]{\scctmonitor}{\scctmonitor}{\absev{\emptyevent}}
-  }{scct-none}
-  %
-  \typerule{scct-abort}{
-  }{
-    \monitorcheck[]{\scctmonitor}{\scctmonitor}{\absterm}
-  }{scct-abort}
-\end{center}
-The monitor state for the \gls*{scct} monitor is completely empty, since it does not need to keep track of information.
-As soon as any event is hit, the execution gets stuck, since any event is considered confidental from the perspective of this monitor.
-\begin{center}
-  \judgbox{\sccttraceagree[]{\event}{\absevent}}{,,Abstract event $\absevent$ is equivalent to $\event$ with respect to \gls*{cct}.''}$\;$\\
-  %
-  \typerule{scct-low-authentic}{
-  }{
-    \sccttraceagree[]{\event[_b];\sandboxtag;\unlock}{\absev{\emptyevent}}
-  }{scct-low-authentic}
-  %
-  \typerule{scct-high-authentic}{
-  }{
-    \sccttraceagree[]{\event[_b];\sandboxtag;\lock}{\absev{Any}}
-  }{scct-high-authentic}
-  %
-  \typerule{scct-empty-authentic}{
-  }{
-    \sccttraceagree[]{\emptyevent}{\absev{\emptyevent}}
-  }{scct-empty-authentic}
-  %
-  \typerule{scct-abort-authentic}{
-  }{
-    \sccttraceagree[]{\lightning}{\absterm}
-  }{scct-abort-authentic}
-\end{center}
-Accordingly, the event agreement simply disregards all events that involved public data ($\unlock$) while mapping any other event that does involve private data ($\lock$) to the abstract $\absev{Any}$ event.
-
-\begin{lemma}[Monitor Traces are $\scctsafe$]\label{lem:mon:scctsafe}
-  If $\monscctsafe{\trace}$, then $\trace\in\scctsafe$. %\Coqed
-\end{lemma}
-
-\begin{example}[Data-independent timing mode]
-  Consider the call \texttt{strncpy(1,x,y)} to the \texttt{strncpy} function (\Cref{ex:strncpy}) with low ($\unlock$) security for \texttt{x} and high security ($\lock$) for \texttt{y}.
-  The trace of just the copying part inside of \texttt{strncpy} looks like $\ev{Use}\ \loc_{x}\ 0;\comp;\unlock\cdot\ev{Use}\ \loc_{y}\ 0;\comp;\lock$.
-  In terms of the abstract monitor trace, this is just $\absev{Any}$.
-  Running this on the monitor would result in getting stuck, since there is no matching rule to step in the presence of $\absev{Any}$ event. % pun intended
-  By means of additional features to ensure cryptographic constant time even in the presence of memory reads and loads, such as a flag to enable a data independent timing mode, which is present in both Arm~\cite[p.~543]{arm-refman} and Intel~\cite[p.~80]{intel-refman} processors, the original trace now does $\ev{Use}\ \loc_{y}\ 0;\comp;\unlock$ instead of $\ev{Use}\ \loc_{y}\ 0;\comp;\lock$.
-  Because of this, the whole trace of the component equates to $\absev{\emptyevent}$ and the monitor can step without getting stuck.
-\end{example}
-
-\subsubsection{Combining \gls*{ms} and \gls*{scct} Monitors to obtain \gls*{msscct}}
-
-The combination of monitors for \gls*{ms} and \gls*{scct} yields one for \gls*{msscct}.
-The construction is entirely similar to the one for \gls*{ms} (\Cref{subsubsec:ms}).
-
-\begin{lemma}[Traces with Monitor Satisfaction are $\msscctsafe$]\label{lem:mon:msscctsafe}
-  If $\monmsscctsafe{\trace}$, then $\trace\in\msscctsafe$. %\Coqed
-\end{lemma}
-
-\section{Composing Secure Compilers\pages{1}}\label{sec:compcomp}
-
-This section presents the key meta-theoretic results of this paper concerning sequential compiler composition (and of optimisation passes) (\Cref{sec:sequential}) and concerning other kinds of compiler composition (\Cref{sec:other-compos}).
-% and builds on prior work~\cite{kruse2022csc}\MK{mention this at all?
-may reveal identity even though its double-blind}.
-% MP: not needed at all, ever, workshop work does not need to be cited in the main work
-
-\subsection{Secure Sequential Composition}\label{sec:sequential}
-
-The main result is that secure compilers in the robust compilation framework~\cite{abate2019jour} compose {\em sequentially}.
-This is not intuitive in the sense that in the security domain, composition does not work without additional generalizations~\cite{canetti2006univcomp,mccullough2012compo,fabian2022automatic}.
-The \bul{sequential composition of compilers $\cc{\src{L}}{\trg{L}}$ and $\cc{\trg{L}}{\obj{L}}$} is defined as follows:
-Given an $\src{L}$ program $\src{p}$ and compilers $\cc{\src{L}}{\trg{L}},\cc{\trg{L}}{\obj{L}}$, \rul{its compiled $\obj{L}$ counterpart is obtained by plugging $\src{p}$ into $\cc{\src{L}}{\trg{L}}\circ\cc{\trg{L}}{\obj{L}}$}.
-\begin{definition}
-  \bul{$\cc{\src{L}}{\trg{L}}\circ\cc{\trg{L}}{\obj{L}}$} $\isdef$ Given $\src{p}$, yield \rul{$\cc{\trg{L}}{\obj{L}}\left(\cc{\src{L}}{\trg{L}}(\src{p})\right)$}
-\end{definition}
-
-Consider the compilation chain for $\src{TypeScript}$.
-First, $\src{TypeScript}$ programs are translated to $\trg{JavaScript}$ which, e.g., V8~\cite{googlev8} eventually compiles in parts to $\irl{Ignition BC}$.
-The following theorem establishes what happens if all these compilation steps were robustly secure with respect to \gls*{ms}:
-The resulting $\irl{Ignition BC}$ code would be \gls*{ms} regardless of the context the binary runs in.
-
-Given \bul{$\cc{\src{L}}{\trg{L}}$ robustly preserves $\class[_{1}]$} and \rul{$\cc{\trg{L}}{\obj{L}}$ robustly preserves $\class[_{2}]$}, it follows that \oul{their sequential composition $\cc{\src{L}}{\trg{L}}\circ\cc{\trg{L}}{\obj{L}}$ robustly preserves the intersection of classes $\class[_{1}]$ and $\class[_{2}]$}.
-
-\begin{theorem}[Sequential Composition of Secure Compilers]\label{thm:rtp}
-  If \bul{$\rtp{\cc{\src{L}}{\trg{L}}}{\class[_{1}]}$} and \rul{$\rtp{\cc{\trg{L}}{\obj{L}}}{\class[_{2}]}$}, then \oul{$\rtp{\cc{\src{L}}{\trg{L}}\circ\cc{\trg{L}}{\obj{L}}}{\class[_{1}]\cap\class[_{2}]}$}.
-\Coqed
-\end{theorem}
-
-Since the composition of secure compilers is again a secure compiler, the theorem generalises to a whole chain of $n$ secure compilers.
-
-\subsubsection{Securing Optimisations}
-Notably, real-world compilation chains also perform a series of (sequential) passes whose main purpose is not necessarily to translate from one language to another, but to, e.g., optimise the code or enforce a certain property.
-Both examples can be seen in practice, e.g. as in the work of~\cite{nagarakatte2009soft,nagarakatte2010cets,akritidis2009baggy,wegman1991ccp,manjikian1997fusion} and many more.
-% \MPin{survey to cite?
-this is a bit informal MK: hard to find a survey on different compiler passes...
-I can't even find one just for compiler enforced memsafety (that looks good) or for compiler translations in general}.
-% 
-% Consider the compilation passes of LLVM~\cite{lattner2004llvm}, which are handled in the \texttt{PassManager} class.
-% By manipulating the internal state of a \texttt{PassManager} object, the schedule of compiler optimisation passes can be hardcoded.
-% That way, an order of optimisations can be found by hand that yields the best results.
-% \MP{what's the point of the previous 3 sentences?(commented)}
-Consider the following two LLVM optimisation passes: \gls*{cf}, which rewrites constant expressions to the constant they evaluate to, and \gls*{dce}, which removes dead code by rewriting conditional branches.
-The order in which \gls*{cf} and \gls*{dce} are performed influences the final result of the compilation (see \Cref{fig:cfdceex}).
-\begin{figure}[!ht]
-  \begin{center}
-    \begin{tikzpicture}
-      \node[minimum width=3cm,minimum height=2.5cm,draw=black,very thick,align=left] (progunopt) {\begin{lstlisting}[language=ml,basicstyle=\small\ttfamily]
-let a = true in
-if a then
-  print "a"
-else
-  print "b"
-\end{lstlisting}};
-      \node[minimum width=3cm,minimum height=2.5cm,draw=black,very thick,align=left,right=1.0 of progunopt] (progoptcf) {\begin{lstlisting}[language=ml,basicstyle=\small\ttfamily]
-if true then
-  print "a"
-else
-  print "b"
-\end{lstlisting}};
-      \node[minimum width=3cm,minimum height=2.5cm,draw=black,very thick,align=left,right=1.0 of progoptcf] (progoptcfdce) {\begin{lstlisting}[language=ml,basicstyle=\small\ttfamily]
-print "a"
-\end{lstlisting}};% the following is a hack to get my syntax highlighting to work again \end{lstlisting}
-      % edges
-      \draw[->,very thick] (progunopt) edge[loop left,left] node {\gls*{dce}} (progunopt);
-      \draw[->,very thick] (progunopt) edge[sloped,above] node {\gls*{cf}} (progoptcf);
-      \draw[->,very thick] (progoptcf) edge[sloped,above] node {\gls*{dce}} (progoptcfdce);
-    \end{tikzpicture}
-  \end{center}
-  \caption{Example program where the level of optimisations differ for one pass of applying \gls*{cf} and \gls*{dce} in any order. %
-  Every edge is a compilation pass and the label on the edge states what the pass does, i.e., \gls*{cf} or \gls*{dce}. %
-  The source code in the nodes is a glorified compiler intermediate representation and the code gets more optimised towards the right hand side of the figure.}\label{fig:cfdceex}
-\end{figure}
-This {\em phase ordering problem} is well--known in literature and a practical solution is to simply perform a fixpoint iteration of the optimisation pipeline~\cite{click1995combining}.
-%If possible, optimisations are merged to enhance the analysis' result, since even when iterating to a fixpoint the final program may not be as optimised as it could be~\cite{click1995combining}.
-Compiler engineers typically try to find an order of optimisations that yields well-optimised programs for either code size~\cite{cooper1999geneticphases} or performance~\cite{kulkarni2006exhaustivephase}.
-\Cref{corr:swappable} justifies that any such order of compilation passes is valid with respect to security.
-So, given two compilation passes \bul{$\cc[_{1}]{\trg{L}}{\trg{L}}$, $\cc[_{2}]{\trg{L}}{\trg{L}}$, both robustly preserving class $\class[_{1}]$ or $\class[_{2}]$, respectively}, \oul{for any order of their composition the composed compiler robustly preserves the intersection of $\class[_{1}]$ and $\class[_{2}]$}.
-
-\begin{corollary}[Sequential Composition of Secure Compilers]\label{corr:swappable}
-  If \bul{$\rtp{\cc[_{1}]{\trg{L}}{\trg{L}}}{\class[_{1}]}$ and $\rtp{\cc[_{2}]{\trg{L}}{\trg{L}}}{\class[_{2}]}$}, then \oul{$\rtp{\cc[_{1}]{\trg{L}}{\trg{L}}\circ\cc[_{2}]{\trg{L}}{\trg{L}}}{\class[_{1}]\cap\class[_{2}]}$ and $\rtp{\cc[_{2}]{\trg{L}}{\trg{L}}\circ\cc[_{1}]{\trg{L}}{\trg{L}}}{\class[_{2}]\cap\class[_{1}]}$}.
-\Coqed
-\end{corollary}
-
-\subsection{Secure Upper and Lower Composition}\label{sec:other-compos}
-Besides sequential composition, there are two other compositions, namely an {\em upper}, i.e., a compiler that takes multiple inputs and yields one output, and a {\em lower} composition, i.e., a compiler that takes one input and yields multiple outputs.
-\bul{Define the upper composition $\cc{\src{L}+\obj{L}}{\trg{L}}$} as follows:
-Given a program \texttt{p}, its compiled counterpart is obtained by \rul{plugging \texttt{p} into $\cc{\src{L}}{\trg{L}}$ if $\texttt{p}\in\src{L}$} or by \oul{plugging \texttt{p} into $\cc{\obj{L}}{\trg{L}}$ if $\texttt{p}\in\obj{L}$}.
-\begin{definition}[Upper Composition]
-  \bul{$\cc{\src{L}+\obj{L}}{\trg{L}}$} $\isdef$ 
-  % Given \texttt{p}, yield
-  $\lambda \texttt{p}\ldotp$
-  $\left\{\mbox{\begin{tabular}{c}
-    {\rul{if $\texttt{p}\in\src{L}$, then $\cc{\src{L}}{\trg{L}}(\texttt{p})$}} \\
-    \mbox{\oul{if $\texttt{p}\in\obj{L}$, then $\cc{\obj{L}}{\trg{L}}(\texttt{p})$}} \\
-  \end{tabular}}\right.$
-%
-%
-%  $\begin{cases}
-%                                                                            \rul{\cc{\src{L}}{\trg{L}}(\texttt{p}) & \text{if }\texttt{p}\in\src{L}} \\
-%                                                                            \cc{\obj{L}}{\trg{L}}(\texttt{p}) & \text{if }\texttt{p}\in\obj{L}
-%                                                                          \end{cases}$
-\end{definition}
-
-Examples of this are present in industry:
-Consider the Java Virtual Machine bytecode $\trg{JVM BC}$, which is a popular target for programming language designers due to its high performance and relevance in industry.
-Compilers for several programming languages have it as their target language, some popular instances are $\src{Java}$ and $\obj{Kotlin}$.
-Technically speaking, they both compile to class files and $\obj{Kotlin}$ objects are considered to be the same as $\src{Java}$ objects at that point.
-Both languages can be used at the same time in one project~\cite{androidstudio}.
-A compiler that accepts both $\src{Java}$ and $\obj{Kotlin}$ code translating to the same target language or intermediate representation performs a kind of {\em upper} composition.
-Now, the following theorem tells us what happens if these are secure:
-Given \bul{$\cc{\src{L}}{\trg{L}}$ robustly preserves $\class[_{1}]$} and \rul{$\cc{\obj{L}}{\trg{L}}$ robustly preserves $\class[_{2}]$}, it follows that \oul{their upper composition $\cc{\src{L}+\obj{L}}{\trg{L}}$ robustly preserves the intersection of classes $\class[_{1}]$ and $\class[_{2}]$}.
-
-\begin{theorem}[Upper Composition of Secure Compilers]\label{thm:urtp}
-  If \bul{$\rtp{\cc{\src{L}}{\trg{L}}}{\class[_{1}]}$} and \rul{$\rtp{\cc{\obj{L}}{\trg{L}}}{\class[_{2}]}$}, then \oul{$\rtp{\cc{\src{L}+\obj{L}}{\trg{L}}}{\class[_{1}]\cap\class[_{2}]}$}. \Coqed
-\end{theorem}
-
-Dually, the {\em lower} composition is concerned about compilers that accept the same source but yield different target languages. %and the same theoretical results apply.
-\bul{Define the lower composition $\cc{\src{L}}{\trg{L}+\obj{L}}$} as follows:
-Given a program $\src{p}$, its compiled counterpart is obtained by \rul{plugging $\src{p}$ into $\cc{\src{L}}{\trg{L}}$} or by \oul{plugging \texttt{p} into $\cc{\src{L}}{\obj{L}}$}, respectively, \pul{based on the internal decision}.
-\begin{definition}[Lower Composition]
-  \bul{$\cc{\src{L}}{\trg{L}+\obj{L}}$} $\isdef$ 
-  % Given $L$ and $\src{p}$, yield
-  $\lambda \src{p}, L\ldotp$
-  $\left\{\mbox{\begin{tabular}{c}
-    {\pul{if $L=\trg{L}$, then} \rul{$\cc{\src{L}}{\trg{L}}(\src{p})$}} \\
-    \mbox{\pul{if $L=\obj{L}$, then} \oul{$\cc{\src{L}}{\obj{L}}(\src{p})$}} \\
-  \end{tabular}}\right.$
-%
-%
-%  $\begin{cases}
-%                                                                            \rul{\cc{\src{L}}{\trg{L}}(\texttt{p}) & \text{if }\texttt{p}\in\src{L}} \\
-%                                                                            \cc{\obj{L}}{\trg{L}}(\texttt{p}) & \text{if }\texttt{p}\in\obj{L}
-%                                                                          \end{cases}$
-\end{definition}
-
-Consider two compilers both accepting $\src{LLVM IR}$~\cite{lattner2004llvm} and one of them emits $\trg{x86\_64}$, while the other emits $\obj{ARMv8}$.
-It is intuitive that they are in some sense composed in the LLVM framework, but the decision of when to use one over the other is inherently {\em internal} to the formalisation effort of this kind of composition.
-For example, the user of this compiler provides an explicit flag that instructs to emit $\trg{x86\_64}$ or the framework itself detects the target platform via heuristics, such as supported instructions.
-
-The following theorem demonstrates what happens if the involved compilers are secure:
-Given \bul{$\cc{\src{L}}{\trg{L}}$ robustly preserves $\class[_{1}]$} and \rul{$\cc{\src{L}}{\obj{L}}$ robustly preserves $\class[_{2}]$}, it follows that \oul{their lower composition $\cc{\src{L}}{\trg{L}+\obj{L}}$ robustly preserves the intersection of classes $\class[_{1}]$ and $\class[_{2}]$}.
-
-\begin{theorem}[Lower Composition of Secure Compilers]\label{thm:lrtp}
-  If \bul{$\rtp{\cc{\src{L}}{\trg{L}}}{\class[_{1}]}$} and \rul{$\rtp{\cc{\obj{L}}{\trg{L}}}{\class[_{2}]}$}, then \oul{$\rtp{\cc{\src{L}}{\trg{L}+\obj{L}}}{\class[_{1}]\cap\class[_{2}]}$}. %\Coqed
-\end{theorem}
-
-%Either way, the theoretical results suggest that it is possible to always find a ,,most-general'', secure compiler, given two secure compilers, that robustly preserves the least-upper bound of the classes involved in their compilation process.
-
-%\section{Case Study: Memory Safe and Cryptographic Constant Time Code}\label{sec:casestud}
-\section{Case Study: Language Formalisations\pages{2}}\label{sec:casestud:defs}
-
-This section defines programming languages that the secure compilers defined in the next section will use.
-To this end, this section defines the languages $\Ltms$, $\Ltrg$, $\Lms$, and $\Lscct$ which share many common elements (presented in \Cref{subsec:cs:defs}).
-$\Ltms$ is the only statically typed language and exhibits the property that all well-typed programs are \gls*{tms} (\Cref{subsec:ltms}).
-However, not all $\Ltms$ programs are \gls*{sms}.
-That is, there are well-typed $\Ltms$ programs that perform an out-of-bounds access.
-Language $\Ltrg$ is untyped and does not provide any guarantees with regards to \gls*{ms} (\Cref{subsec:lsms}).
-$\Lms$ is exactly the same language as $\Ltrg$, but this paper still distinguishes the two for sake of readability (\Cref{subsec:lms}).
-All three languages --- so $\Ltms$, $\Ltrg$, and $\Lms$ --- assume \gls*{cct} to hold.
-
-% This paper considers \gls*{cct} as an architecture-dependent property\MPin{can we justify this}, so
-Writing code attaining \gls*{cct} should not be of the programmer's concerns~\cite{cauligi2019fact}.
-Such consideration is also backed up by architecture providing a data (operand) independent timing mode, such as processors by Arm~\cite[p.~543]{arm-refman} and Intel~\cite[p.~80]{intel-refman}.
-In spirit of this, language $\Lscct$ allows violating \gls*{cct} by emitting events on, e.g., branching and division, that contain secrets (\Cref{subsec:lscct}), but provides a way to read and write to a {\em model--specific register} that enables a ``\gls*{cct}-mode''.
-
-\subsection{Shared Language Definitions}\label{subsec:cs:defs}
-\begin{gather*}
-  \begin{aligned}
-    \mi{(Expressions)}~&\expr\bnfdef\ x \mid \valueexpr \mid \lbinop{\expr[_{1}]}{\expr[_{2}]} \mid \lget{x}{\expr} \mid \llet{x}{\lnew{\expr[_{1}]}{\expr[_{2}]}}{\expr[_{3}]} \mid \ldelete{x} \mid \lset{x}{\expr[_{1}]}{\expr[_{2}]} \\
-                       & \mid \lpair{\expr[_{1}]}{\expr[_{2}]} \mid \lproja{\expr} \mid \lprojb{\expr} \mid \llet{x}{\expr[_{1}]}{\expr[_{2}]} \mid \lreturn{\expr} \mid \lcall{g}{\expr} \mid \lifz{\expr}{\expr[_{1}]}{\expr[_{2}]} \mid \labort
-  \end{aligned}
-  \\
-  \begin{aligned}
-  \mi{(Types)}~&\type\bnfdef\natt \mid \type[_{1}] \times \type[_{2}]
-  &
-  \mi{(Functions)}~&\function\bnfdef\ \lfunction{\textit{foo}}{x}{e}
-  &
-  \mi{(Libraries)}~&\library\bnfdef\ \hole{\cdot} \mid \function,\library
-  \end{aligned}\\
-  \begin{aligned}
-  \mi{(Component~Names)}~&\commlib\bnfdef\ \hole{\cdot} \mid \text{foo},\commlib
-  &
-  \mi{(Programs)}~&\prog{\library_{\ctx}}{\library_{\comp}}
-  \end{aligned}
-\end{gather*}
-
-Above is the shared syntax of all the programming languages of this paper.
-Variables are referred to as $x,y,z,a,b,c,\ldots$ while functions may be referred to as multi-character words, such as $\texttt{foo}$, as well as short-forms like $f,g,h$.
-All languages share the type $\natt$ representing natural numbers.
-Functions are constrained to take one argument and can only call other functions listed in libraries, which are just lists of functions.
-A program $\prog{\library_{\ctx}}{\library_{\comp}}$ is indexed by two libraries that represent all context- and component-level functions, respectively.
-Lists of component-level names are referred to as $\commlib$.
-
-\begin{gather*}
-  \begin{aligned}
-    \mi{(Control~Tags)}~&\sandboxtag\bnfdef\ctx \mid \comp
-    &
-    \mi{(Communication~Tags)}~&\comm\bnfdef\ \ctxtocomp \mid\ \comptoctx \mid \nocomm
-    &
-    \mi{(Poison~Tags)}~&\poison\bnfdef \poisoned \mid \poisonless
-  \end{aligned}\\
-  \begin{aligned}
-    \mi{(Continuation~Stacks)}~&\ectxstack\bnfdef\hole{\cdot} \mid \tup{K}{g},\ectxstack
-    &
-    \mi{(Control~Flow~States)}~&\cfstate\bnfdef\tup{\library}{\ectxstack}
-  	\\
-    \mi{(Locations)}~&\loc\in\nat
-    &
-    \mi{(Substitutions)}~&\substvar\bnfdef \hole{\cdot} \mid \subst{\valueexpr}{x},\gamma
-  	\\
-    \mi{(Stores)}~&\Delta\bnfdef\hole{\cdot} \mid x\mapsto\tup{\loc}{\sandboxtag,\poison,n},\Delta
-    &
-    \mi{(Heaps)}~&H\bnfdef\hole{\cdot} | \valueexpr,H
-    \end{aligned}\\
-  \begin{aligned}
-    \mi{(Memory~States)}~&\memstate\bnfdef\tup{H^{\ctx}}{H^{\comp};\Delta}
-    &
-    \mi{(States)}~&\statevar\bnfdef\tup{\cfstate}{\sandboxtag,\memstate}
-    &
-    \mi{(Runtime~Terms)}~&\runtimetermvar\bnfdef\rtt{\statevar}{\expr}
-  \end{aligned}
-\end{gather*}
-
-States $\statevar$ are tuples\footnote{Throughout the paper, the substitution notation is also used to update entries in states $\statevar$.} containing a control flow state $\cfstate$, control tags $\sandboxtag$, and a memory state $\memstate$.
-A control flow state $\cfstate$ entails a library, which provides definitions for functions calls, as well as a stack of continuations $\ectxstack$.
-Elements of the stack $\ectxstack$ are pairs of evaluation contexts and the name of a function associated to that evaluation context.
-Memory states $\memstate$ are tuples of two separate heaps $H$ and a store $\Delta$, which contains pointer metadata, such as the concrete memory location $\loc$, a control tag $\sandboxtag$ indicating which heap the pointer points into, a poison tag $\poison$, as well as bounds information $n$.
-The bounds information is a mere proof artefact that has no semantic significance.
-The two separate heaps essentially model a sandbox, to prevent contexts from performing pointer arithmetic and reading from or writing to the data owned by a component.
-While this prevents the effects of out-of-bounds accesses across the context and component boundary, the goal of the design of the languages of this paper is to be able to express security violations.
-To this end, the poison tag $\poison$ indicates whether a pointer has been freed $\poisoned$ or is still allocated $\poisonless$, so that pointers can be used even after their deallocation without the semantics getting stuck.
-Runtime terms are simply expressions $\expr$ paired with the operational state $\statevar$.
-
-\begin{gather*}
-  \begin{aligned}
-  \mi{(Pre\text{-}Events)}~\event[_{b}]&\bnfdef \ev{Alloc}\ \loc\ \valueexpr \mid \ev{Dealloc}\ \loc \mid \ev{Get}\ \loc\ \valueexpr \mid \ev{Set}\ \loc\ \valueexpr\ \valueexpr['] \mid \cdots \\
-  \mi{(Events)}~\event&\bnfdef \emptyevent \mid \lightning \mid \tup{\event[_{b}]}{\sandboxtag,\securitytag} \\
-  \end{aligned}
-\end{gather*}
-
-All languages use the same trace model, where events are either the empty event \emptyevent, the program crash event $\lightning$, or a tuple consisting of a control tag $\sandboxtag$ and a security tag $\securitytag$.
-The former indicates whether the component \comp or the context \ctx is to blame for emitting this event, the latter indicates the secrecy level of values of the emitted event, i.e., either $\lock$ or $\unlock$.
-As for pre-events, the memory-related ones are allocation ($\ev{Alloc}\ \loc\ \valueexpr$), deallocation ($\ev{Dealloc}\ \loc$), reading from ($\ev{Get}\ \loc\ \valueexpr$) and writing to memory ($\ev{Set}\ \loc\ \valueexpr\ \valueexpr[']$).
-The following is an excerpt of the operational semantics handling some of the memory operations.
-
-\begin{center}
-  \judgbox{\pstepto{\runtimetermvar}{\runtimetermvar[']}{\event}}{,,$\runtimetermvar$ does one primitive step to $\runtimetermvar[']$ emitting event $\event$.''}
-  %
-  \typerule{$e-\text{get}-\in$}{
-    \sandboxtag=\statevar.\sandboxtag &
-    \statevar.\store(x)=\tup{\loc}{\sandboxtag,\poison,m} \\
-    \loc+n\in\operatorname{dom}\statevar.\heap[^{\sandboxtag}](\loc+n)
-  }{
-    \pstepto{\rtt{\statevar}{\lget{x}{n}}}{\rtt{\statevar}{\heap[^{\sandboxtag}](\loc+n)}}{\tup{\ev{Get}\ \loc\ n}{\sandboxtag}}
-  }{e-get-in}
-  %
-  \typerule{$e-\text{set}-\notin$}{
-    \sandboxtag=\statevar.\sandboxtag &
-    \statevar.\store(x)=\tup{\loc}{\sandboxtag,\poison,m} \\
-    \loc+n\notin\operatorname{dom}\statevar.\heap[^{\sandboxtag}]
-  }{
-    \pstepto{\rtt{\statevar}{\lset{x}{n}{\valueexpr}}}{\rtt{\statevar}{\valueexpr}}{\tup{\ev{Set}\ \loc\ n\ \valueexpr}{\sandboxtag}}
-  }{e-set-notin}
-  %
-  \typerule{$e-\text{new}$}{
-    \fresh{\statevar}{z} &
-    \fresh{\statevar}{\loc} &
-    \heap[^{\sandboxtag}_1]=\growh{\statevar.\heap[^{\sandboxtag}]}{n} &
-    \store[_1]=\storeel{z}{\loc}{\statevar.\sandboxtag}{\poisonless}{n},\statevar.\store
-  }{
-    \pstepto{\rtt{\statevar}{\lnew{x}{n}{\expr}}}{\rtt{\statevar\subst{\heap[^{\sandboxtag}]}{\heap[^{\sandboxtag}_1]}\subst{\store}{\store[_1]}}{\expr\subst{z}{x}}}{\tup{\ev{Alloc}\ \loc\ n}{\sandboxtag}}
-  }{e-new}
-  %
-  \typerule{$e-\text{dealloc}$}{
-    \statevar.\store(x) = \tup{\loc}{\statevar.\sandboxtag,\poison,n} &
-    \store[_1]=\seth{\statevar.\store}{x}{\tup{\loc}{\statevar.\sandboxtag,\poisoned,n}}
-  }{
-    \pstepto{\rtt{\statevar}{\ldelete{x}}}{\rtt{\statevar\subst{\store}{\store[_1]}}{0}}{\tup{\ev{Dealloc}\ \loc}{\sandboxtag}}
-  }{e-dealloc}
-  %
-  \typerule{$e-\text{abort}$}{
-  }{
-    \pstepto{\rtt{\statevar}{\labort}}{\lightning}{\lightning}
-  }{e-abort}
-\end{center}
-
-To demonstrate the use of the poison tag $\poison$ as metadata for pointers instead of removing them from the store $\Delta$, consider \Cref{tr:e-dealloc,tr:e-get-in,tr:e-set-notin} .
-In \Cref{tr:e-dealloc}, the premise does not care at all about the actual state of the poison tag $\poison$ and just overwrites it, marking the location as freed $\poisoned$.
-Besides that, the poison tag does not have any semantic meaning.
-For language $\Ltms$, this tag is really just some semantic metadata that programmers have no access to.
-But, for the other languages, e.g., $\Ltrg$, the poison tag is used to check pointer validity.
-\Cref{tr:e-new} allocates enough space on the respective heap, either $H^{\ctx}$ or $H^{\comp}$ depending on the execution context, i.e., the value of $\Omega.\sandboxtag$, and adds the appropriate metadata associated to the pointer in $\Delta$.
-Reading from \Cref{tr:e-get-in} and writing to memory \Cref{tr:e-set-notin} have two cases: Either the heap is large enough or not and, depending on that, either the actual value stored at that location is read from or written to, or some garbage data is returned.
-However, note that the execution does not get stuck in such cases, it performs a step, and emits an appropriate event.
-Also note that whether a pointer is poisoned or not is not reflected on the trace.
-
-\begin{center}
-  \begin{align*}
-    \mi{(Pre\text{-}Events)}~\event[_{b}]&\bnfdef \cdots \mid \ev{Call}\ \comm\ g\ \valueexpr \mid \ev{Ret}\ \comm\ \valueexpr \mid \ev{Start} \mid \ev{End}\ \valueexpr \\
-  \end{align*}
-\end{center}
-
-A key difference in comparison with the specification trace model (\Cref{subsec:propdefs}) is that, as standard in secure compilation work~\cite{patrignani2021rsc,abate2019jour,korashy2021capableptrs}, the traces have a call and return event that signals context switches, which are referred to as {\em interaction events}.
-The reason for these interaction events is technical: They are a proof artifact for reconstructing a source context from a potentially malicious target context, where during that translation, the insertion of some wrapper code right before context switching may be necessary to make the proof succeed.
-Hereby, a $\ev{Call\ \ctxtocomp\ foo\ \valueexpr}$ and $\ev{Return\ \ctxtocomp\ \valueexpr}$ signal that program execution transitions from context- to component-level.
-Contrary, $\ev{Call\ \comptoctx\ foo\ \valueexpr}$ and $\ev{Return\ \ctxtocomp\ \valueexpr}$ signal that program execution transitions from component- to context-level.
-For calls without this context switch, the environmental semantics attaches the $\nocomm$ tag.
-In the following, $\neg\ctx=\comp$ and $\neg\comp=\ctx$.
-
-\begin{center}
-  \judgbox{\estepto[]{\runtimetermvar}{\runtimetermvar}{\event}}{,,Contextual step from runtime-term $\runtimetermvar$ to $\runtimetermvar[']$ emitting event $\event$.''}
-  %
-  \typerule{$e-\text{ret}$}{
-    \statevar.\ectxstack=\tup{\ectx}{\text{foo}},\ectxstack['] &
-    \commlib\vdash\text{foo}:\Omega.\sandboxtag\dashv\comm
-  }{
-    \estepto[]{\rtt{\statevar}{\ectx[']\hole{\lreturn{\valueexpr}}}}{\rtt{\statevar\subst{\neg\statevar.\sandboxtag}{\sandboxtag}\subst{\ectxstack[']}{\ectxstack}}{\ectx\hole{\valueexpr}}}{\tup{\ev{Ret}\ \comm\ \valueexpr}{\statevar.\sandboxtag}}
-  }{e-ret-notsame}
-  %
-  \typerule{$e-\text{call}-\text{notsame}$}{
-    \lfunction{\text{foo}}{x}{e} \in \statevar.\library &
-    \commlib\vdash\text{foo}:\Omega.\sandboxtag\dashv\comm
-  }{
-    \estepto[]{\rtt{\statevar}{\ectx\hole{\lcall{\text{foo}}{\valueexpr}}}}{\rtt{\statevar\subst{\neg\statevar.\sandboxtag}{\sandboxtag}\subst{\tup{\ectx}{\text{foo}},\statevar.\ectxstack}{\ectxstack}}{\expr\subst{\valueexpr}{x}}}{\tup{\ev{Call}\ \comm\ \text{foo}\ \valueexpr}{\statevar.\sandboxtag}}
-  }{e-call-notsame}
-\end{center}
-
-The environmental semantics is mostly straightforward.
-In \Cref{tr:e-ret-notsame,tr:e-call-notsame}, the judgement $\Omega.\commlib\vdash\text{foo}:\Omega.\sandboxtag$ checks whether $\text{foo}$ is a component-level name by looking it up in the list of component-level names $\commlib$ and emits the appropriate transfer tag, i.e., either $\comptoctx$ or $\ctxtocomp$.
-Additional rules that are left out ensure that, e.g., when calling the $\text{main}$ function, the event $\ev{Start}$ is emitted, which is a design choice this paper does for convenience when reasoning about call-chains.
-Note that the $\ev{End}\ \valueexpr$ event is not emitted if the program crashes.
-
-% \MKin{%
-%   should I add a rule for top level exec or keep it like this?
-% 	MP: like this is fine
-% }
-The top-level execution $\progstepto{\prog{\library_{\ctx}}{\library_{\comp}}}{\runtimetermvar}{\trace}$ constructs an initial state $\Omega$ by linking $\library_{\ctx}$ and $\library_{\ctx}$ and then starts execution by calling the $\text{main}$ function.
-The trace $\trace$ emitted during that execution serves as abstraction of the behavior of the program enabling the use of \Cref{def:propsat,def:proprsat}.
-
-\subsection{$\src{L_{\tmssafe}}$: A Temporal but Not Spatial Memory Safe Language}\label{subsec:ltms}
-
-$\src{L_{\tmssafe}}$ uses the same syntax as presented earlier (\Cref{sec:casestud:defs}) without extensions to the term level.
-But, $\src{L_{\tmssafe}}$ is statically typed, where the type system is inspired by $L^{3}$~\cite{morrisett2005L3,scherer2018fabulous}.
-The type system of $\src{L_{\tmssafe}}$ exhibits the property that every well-typed $\src{L_{\tmssafe}}$ program satisfies \gls*{tms} (\Cref{thm:wt:tms}).
-The proof of this theorem relies on a projection $\msfilterLtms{\src{\event}}=\event$ from $\src{L_{\tmssafe}}$ events to specification events $\event$, because the properties defined earlier (\Cref{subsec:propdefs}) are defined in the specification trace model.
-Hereby, the map $\thelocmap(\src{\loc})=\loc$ maps an $\Ltms$ location to a location $\loc$ of the specification trace model.
-
-\begin{center}
-  \judgbox{\msfilterLtms{\src{\event}}=\event}{,,Project an $\Ltms$ event $\src{\event}$ to a specification event $\event$.''}
-  \typerule{$\Ltms$-filter-context}{
-    \src{\event[_b]}\not=\src{\lightning}
-  }{
-    \msfilterLtms{\src{\tup{\event[_b]}{\ctx,\securitytag}} = \emptyevent}
-  }{ltms-filter-context}
-  %
-  \typerule{$\Ltms$-filter-abort}{
-  }{
-    \msfilterLtms{\src{\lightning}} = \lightning
-  }{ltms-filter-abort}
-  %
-  \typerule{$\Ltms$-filter-start}{
-  }{
-    \msfilterLtms{\src{\tup{Start}{\comp}}} = \emptyevent
-  }{ltms-filter-start}
-  %
-  \typerule{$\Ltms$-filter-alloc}{
-    \locmapsto{\src{\loc}}{\loc} &
-    \src{n}=n
-  }{
-    \msfilterLtms{\src{\tup{Alloc\ \loc\ n}{\comp}}} = \tup{\ev{Alloc}\ \loc\ n}{\comp,\unlock}
-  }{ltms-filter-alloc}
-\end{center}
-
-Most rules of the projection $\msfilterLtms{\src{\event}}$ are left out since, for the most part, it does the expected, e.g., $\msfilterLtms{\src{\ev{Dealloc}\ \loc;\comp;\securitytag}}=\ev{Dealloc}\ \delta(\loc);\comp;\securitytag$.
-But, it also filters any action that a context does as well as the interaction events, since these are irrelevant for component-level \gls*{tms}.
-
-\begin{theorem}[$\src{L_{\tmssafe}}$-programs are \gls*{tms}]\label{thm:wt:tms}
-  For any $\src{\library_{\comp}}$, $\rsat{\src{\library_{\comp}}}{\tmssafe}$ \Coqed
-\end{theorem}
-
-\subsection{$\Ltrg$: A Memory-Unsafe Language}\label{subsec:lsms}
-
-\begin{center}
-  \begin{align*}
-    \mi{(Expressions)}~\trg{\expr}&\bnfdef \cdots \mid \trg{\lispoisoned{x}} \mid \trg{\lhast{\expr}{\type}} \\
-  \end{align*}
-\end{center}
-
-$\Ltrg$ extends the syntax presented earlier (\Cref{subsec:cs:defs}) with dynamic typechecks $\trg{\lhast{\expr}{\type}}$ and a way to inspect poison tags $\trg{\lispoisoned{x}}$ in the metadata of pointers.
-For valid pointers ($\trg{\poisonless}$) bound to variable $\trg{x}$, the check $\trg{\lispoisoned{x}}$ yields $\trg{1}$.
-If the array bound to $\trg{x}$ was allocated, i.e., has been poisoned ($\trg{\poisoned}$), the check $\trg{\lispoisoned{x}}$ evaluates to $\trg{0}$.
-
-\begin{center}
-  \typerule{$e-\trg{\lhast{x}{\natt}}$}{
-  }{
-    \pstepto{\rtt{\trg{\statevar}}{\trg{\lhast{x}{\natt}}}}{\rtt{\trg{\statevar}}{\trg{1}}}{\emptyevent}
-  }{e-has-nat-no}
-  %
-  \typerule{$e-\trg{\lhast{n}{\natt}}$}{
-  }{
-    \pstepto{\rtt{\trg{\statevar}}{\trg{\lhast{n}{\natt}}}}{\rtt{\trg{\statevar}}{\trg{0}}}{\emptyevent}
-  }{e-has-nat-yes}
-\end{center}
-
-Dynamic typechecks $\trg{\lhast{\expr}{\type}}$ match on $\trg{\expr}$ and evaluate to $\trg{0}$ if the term is of type $\trg{\type}$ and $\trg{1}$ otherwise.
-
-%
-The projection $\msfilterL{\trg{\trace}}$ is equal to $\msfilterLtms{\src{\trace}}$.
-
-\subsection{$\Lms$: Another Memory-Unsafe Language}\label{subsec:lms}
-To enhance readablity, this paper uses $\Lms$, despite it being exactly equal to $\Ltrg$ (\Cref{subsec:lsms}).
-The projection $\msfilterLms{\irl{\trace}}$ is also exactly equal to $\msfilterL{\trg{\trace}}$.
-
-\subsection{$\Lscct$: A Memory-Unsafe Language with a Data Independent Timing Mode}\label{subsec:lscct}
-
-\begin{gather*}
-  \begin{aligned}
-    \mi{(Expressions)}~\obj{\expr}&\bnfdef \obj{n^{\securitytag}} \mid \cdots \mid \obj{\llet{x^{\securitytag}}{\expr[_{1}]}{\expr[_{2}]}} \mid \cdots \mid \obj{\lwrdoit{\expr}} \mid \obj{\lrddoit{x}{\expr}}
-  \end{aligned} 
-  \\
-  \begin{aligned}
-    \mi{(States)}~\obj{\statevar}&\bnfdef \obj{\tup{\cfstate}{\sandboxtag,n,\memstate}}
-  \end{aligned}
-\end{gather*}
-
-$\Lscct$ extends $\Lms$ (\Cref{subsec:lms}) with a way to write to a model specific register that controls a data (operand) independent timing mode, a feature that is present in both Arm~\cite[p.~543]{arm-refman} and Intel~\cite[p.~80]{intel-refman} processors.
-To this end, states are extended with the value of the register, which is initially set to be not active.
-If the register is marked active, the intuition is that no secrets can appear on specification traces.
-If the register is marked inactive, secrets may appear on traces.
-For the other languages seen earlier, the mode is intuitively always-on, i.e., the mode of execution always uses data independent timing.
-The language also adds user-annotations to values and variables to know their secrecy $\obj{\securitytag}$, which is either high $\obj{\lock}$ or low $\obj{\unlock}$.
-Security tags $\obj{\securitytag}$ are on the usual secrecy lattice, where $\obj{\lock}\le\obj{\securitytag}$ and $\obj{\securitytag}\le\obj{\unlock}$.
-
-\begin{center}
-  \begin{align*}
-    \mi{(Pre\text{-}Events)}~\obj{\event[_{b}]}&\bnfdef \cdots \mid \obj{\widehat{Get}\ \loc\ \valueexpr} \mid \obj{\widehat{Set}\ \loc\ \valueexpr\ \valueexpr[']} \mid \obj{Branch\ n} \mid \obj{Binop\ n}
-  \end{align*}
-\end{center}
-
-To prevent secrets from leaking but still enable reasoning about memory safety, $\Lscct$ extends pre-events with $\obj{\widehat{Get}\ \loc\ \valueexpr}$ and $\obj{\widehat{Set}\ \loc\ \valueexpr}$.
-These indicate reads from and writes to memory without leaking secret information involved in the access.
-Moreover, the language extends pre-events with $\obj{\ev{Branch}\ n}$ and $\obj{\ev{Binop}\ n}$ that are emitted when evaluating a branch or certain binary expressions, such as division, respectively, whenever the data independent timing mode is inactive.
-The following rules demonstrate how this is handled semantically.
-
-\begin{center}
-  \typerule{$e-\obj{\oplus}-\text{noleak}$}{
-    \obj{m}\not=\obj{0} &
-    \obj{n_3}=\obj{n_1}\oplus\obj{n_2} \\
-    \obj{\securitytag['']}\le\obj{\securitytag} &
-    \obj{\securitytag['']}\le\obj{\securitytag[']}
-  }{
-    \pstepto{\rtt{\obj{\cfstate;\sandboxtag;m;\memstate}}{\obj{\lbinop{n_1^{\securitytag}}{n_2^{\securitytag[']}}}}}{\rtt{\obj{\cfstate;\sandboxtag;m;\memstate}}{\obj{n_3^{\securitytag['']}}}}{\obj{\emptyevent}}
-  }{e-binop-noleak}
-  %
-  \typerule{$e-\obj{wrdoit}$}{
-  }{
-    \pstepto{\rtt{\obj{\cfstate;\sandboxtag;m;\memstate}}{\obj{\lwrdoit{n^{\securitytag}}}}}{\rtt{\obj{\cfstate;\sandboxtag;n;\memstate}}{\obj{n^{\securitytag}}}}{\obj{\emptyevent}}
-  }{e-wrdoit}
-  %
-  \typerule{$e-\obj{ifz}-\obj{true}-\text{leak}$}{
-  }{
-    \pstepto{\rtt{\obj{\cfstate;\sandboxtag;0;\memstate}}{\obj{\lifz{0^{\securitytag}}{\expr[_1]}{\expr[_2]}}}}{\rtt{\obj{\cfstate;\sandboxtag;0;\memstate}}{\obj{\expr[_2]}}}{\obj{\ev{Branch}\ 0;\sandboxtag;\securitytag}}
-  }{e-ifz-true-leak}
-\end{center}
-
-The evaluation steps are amended to propagate the security-tag annotations $\obj{\securitytag}$.
-When the data independent timing mode is active, pre-events $\obj{\ev{Branch}\ n}$ and $\obj{\ev{Binop}\ n}$ are emitted for conditionals and binary operations, respectively.
-%To mark this mode as active or inactive, \Cref{tr:e-wrdoit} allows to access the register in $\obj{\statevar}$.
-
-\begin{center}
-  \judgbox{\scctfilterLscct{\obj{\event}}=\event}{,,Project an $\Lscct$ event $\obj{\event}$ to a specification event $\event$.''}
-  \typerule{$\Lscct$-filter-context}{
-    \obj{\event[_b]}\not=\obj{\lightning}
-  }{
-    \scctfilterLscct{\obj{\tup{\event[_b]}{\ctx,\securitytag}} = \emptyevent}
-  }{lscct-filter-context}
-  %
-  \typerule{$\Lscct$-filter-$\widehat{\text{get}}$}{
-    \locmapsto{\obj{\loc}}{\loc} &
-    \obj{n}=n
-  }{
-    \scctfilterLscct{\obj{\tup{\widehat{Get}\ \loc\ n}{\comp,\securitytag}}} = \tup{\ev{Get}\ \loc\ n}{\comp,\unlock}
-  }{lscct-filter-get-hat}
-  %
-  \typerule{$\Lscct$-filter-$\text{get}$}{
-    \locmapsto{\obj{\loc}}{\loc} &
-    \obj{n}=n
-  }{
-    \scctfilterLscct{\obj{\tup{Get\ \loc\ n}{\comp,\securitytag}}} = \tup{\ev{Get}\ \loc\ n}{\comp,\securitytag}
-  }{lscct-filter-get}
-\end{center}
-
-The projection to the specification trace model is mostly straightforward and similar to the others, e.g., \Cref{subsec:cs:tms}.
-However, for events containing the pre-events $\obj{\widehat{Get}}$ and $\obj{\widehat{Set}}$, the projection always translates the security-tag $\obj{\sigma}$ to $\unlock$, regardless of its actual value, as seen in \Cref{tr:lscct-filter-get-hat}.
-The pre-events themselves still translate to just $\ev{Get}$ and $\ev{Set}$, respectively.
-With this technical setup, the information whether a read or write happened on a secret value is not hidden by the semantics, e.g., by emitting $\obj{\emptyevent}$, but when projecting to specification events.
-This allows flexibility: The trace can be checked to satisfy different properties, such as, in this case, \gls*{tms}, \gls*{sms}, \gls*{scct}, and their combined versions.
-\Cref{ex:lscct} illustrates the differences of $\Lscct$ compared to the other languages.
-
-\begin{example}[$\Lscct$ with and without data independent timing]\label{ex:lscct}
-  Consider again the context presented in \Cref{ex:strncpy:cpp11}, where everything is marked with a security tag of high $\lock$.
-  The following table shows parts of the execution trace, read from top to bottom, in the left column (\textit{Active}) with and in the right column (\textit{Inactive}) without data independent timing.
-  The left side of the table ($\Lscct$), i.e., the two columns on the left, describes the execution trace of the program, while the right side of the table ($Spec$), i.e., the two columns on the right, describes the respective projections $\scctfilterLscct{\obj{\trace}}$ to the specification trace model.
-
-  $${\footnotesize
-  \begin{array}{ccccccc}
-        & \Lscct &        &   &      & Spec & \\
-    Active & \mid & Inactive & \mid & Active & \mid & Inactive \\\hline
-    \obj{\ev{Alloc}\ \loc_{x}\ 12;\comp;\unlock} & \mid & \obj{\ev{Alloc}\ \loc_{x}\ 12;\comp;\unlock} & \mid & \ev{Alloc}\ \loc_{x}\ 12;\comp;\unlock & \mid & \ev{Alloc}\ \loc_{x}\ 12;\comp;\unlock\\
-    \obj{\ev{Alloc}\ \loc_{y}\ 12;\comp;\unlock} & \mid & \obj{\ev{Alloc}\ \loc_{y}\ 12;\comp;\unlock} & \mid & \ev{Alloc}\ \loc_{y}\ 12;\comp;\unlock & \mid & \ev{Alloc}\ \loc_{x}\ 12;\comp;\unlock\\
-    \obj{\widehat{\ev{Get}}\ \loc_{x}\ 0;\comp;}\text{\colorbox{lightgray}{$\obj{\lock}$}} & \mid & \obj{\ev{Get}\ \loc_{x}\ 0;\comp;\lock} & \mid & \ev{Use}\ \loc_{x}\ 0;\comp;\text{\colorbox{lightgray}{$\unlock$}} & \mid & \ev{Use}\ \loc_{x}\ 0;\comp;\lock\\
-    \text{\colorbox{yellow}{$\obj{\emptyevent}$}} & \mid & \obj{\ev{Branch}\ 0;\comp;\lock} & \mid & \emptyevent & \mid & \ev{Branch}\ 0;\comp;\lock\\
-    \obj{\widehat{\ev{Get}}\ \loc_{x}\ 0;\comp;}\text{\colorbox{lightgray}{$\obj{\lock}$}} & \mid & \obj{\ev{Get}\ \loc_{x}\ 0;\comp;\lock} & \mid & \ev{Use}\ \loc_{x}\ 0;\comp;\text{\colorbox{lightgray}{$\unlock$}} & \mid & \ev{Use}\ \loc_{x}\ 0;\comp;\lock\\
-    \obj{\widehat{\ev{Set}}\ \loc_{y}\ 0\ \mathtt{'H'};\comp;}\text{\colorbox{lightgray}{$\obj{\lock}$}} & \mid & \obj{\ev{Set}\ \loc_{y}\ 0\ \mathtt{'H'};\comp;\lock} & \mid & \ev{Use}\ \loc_{y}\ 0;\comp;\text{\colorbox{lightgray}{$\unlock$}} & \mid & \ev{Use}\ \loc_{y}\ 0;\comp;\lock\\
-    \obj{\widehat{\ev{Get}}\ \loc_{x}\ 1;\comp;}\text{\colorbox{lightgray}{$\obj{\lock}$}} & \mid & \obj{\ev{Get}\ \loc_{x}\ 1;\comp;\lock} & \mid & \ev{Use}\ \loc_{x}\ 1;\comp;\text{\colorbox{lightgray}{$\unlock$}} & \mid & \ev{Use}\ \loc_{x}\ 1;\comp;\lock\\
-    \text{\colorbox{yellow}{$\obj{\emptyevent}$}} & \mid & \obj{\ev{Branch}\ 0;\comp;\lock} & \mid & \emptyevent & \mid & \ev{Branch}\ 0;\comp;\lock\\
-    \vdots & \mid & \vdots & \mid & \vdots & \mid & \vdots \\
-    \obj{\widehat{\ev{Get}}\ \loc_{x}\ 12;\comp;}\text{\colorbox{lightgray}{$\obj{\lock}$}} & \mid & \obj{\ev{Get}\ \loc_{x}\ 12;\comp;\lock} & \mid &  \ev{Use}\ \loc_{x}\ 12;\comp;\text{\colorbox{lightgray}{$\unlock$}} & \mid & \ev{Use}\ \loc_{x}\ 12;\comp;\lock\\
-    \text{\colorbox{yellow}{$\obj{\emptyevent}$}}& \mid & \obj{\ev{Branch}\ 1;\comp;\lock} & \mid &  \emptyevent & \mid & \ev{Branch}\ 1;\comp;\lock\\
-    \obj{\ev{Dealloc}\ \loc_{y};\comp;\unlock} & \mid & \obj{\ev{Dealloc}\ \loc_{y};\comp;\lock} & \mid &  \ev{Dealloc}\ \loc_{y};\comp;\unlock & \mid & \ev{Dealloc}\ \loc_{y};\comp;\unlock\\
-    \obj{\widehat{\ev{Get}}\ \loc_{y}\ 6;\comp;}\text{\colorbox{lightgray}{$\obj{\lock}$}} & \mid & \obj{\ev{Get}\ \loc_{y}\ 6;\comp;\lock} & \mid &  \ev{Use}\ \loc_{y}\ 6;\comp;\text{\colorbox{lightgray}{$\unlock$}} & \mid & \ev{Use}\ \loc_{y}\ 6;\comp;\lock\\
-  \end{array}}
-  $$
-
-  When the data independent timing mode is off, the execution yields events in similar fashion to before (\Cref{subsec:ltms,subsec:lsms,subsec:lms}).
-  But, if it is turned on, then the branching event \colorbox{yellow}{does not fire} anymore and both reading and writing to memory gets ultimately translated to a specification trace with \colorbox{lightgray}{no exposed secrets}.
-\end{example}
-
-\section{Case Study: Composing Secure Compiler Passes and Optimisations \pages{7}}\label{sec:casestud:rtp}
-This section defines several secure compilers, each of which robustly preserves a different property of interest as depicted in \Cref{fig:pipeline}.
-\begin{figure}[!h]
-  \begin{center}
-    \begin{tikzpicture}
-      \node (S) {$\src{L_{\tmssafe}}$};
-      \node[right=1.5 of S] (T) {$\trg{L}$};
-      \node[right=1.5 of T] (M) {$\irl{L_{\mssafe}}$};
-      \node[below right=0.5 and 1.0 of M] (D0) {$\irl{L_{\mssafe}}$};
-      \node[above right=0.5 and 1.0 of M] (C0) {$\irl{L_{\mssafe}}$};
-      \node[right=3.0 of M] (E) {$\irl{L_{\mssafe}}$};
-      \node[right=1.5 of E] (O) {$\obj{L_{\scctsafe}}$};
-
-      \draw[->] (S) to[sloped] node[align=center,font=\scriptsize] (tmsedge) {\gls*{tms}\\ \Cref{thm:cca:rtp:tms}} (T);
-      \draw[->] (T) to[sloped] node[align=center,font=\scriptsize] {\gls*{sms}\\ \Cref{thm:ccb:rtp:sms}} (M);
-      \draw[->] (M) to[sloped] node[align=center,font=\scriptsize] {\gls*{dce}\\ \Cref{thm:ccdce:rtp:ms}} (D0);
-      \draw[->] (M) to[sloped] node[align=center,font=\scriptsize] {\gls*{cf}\\ \Cref{thm:cccf:rtp:ms}} (C0);
-      \draw[->] (D0) to[sloped] node[align=center,font=\scriptsize] {\gls*{cf}\\ \Cref{thm:cccf:rtp:ms}} (E);
-      \draw[->] (C0) to[sloped] node[align=center,font=\scriptsize] {\gls*{dce}\\ \Cref{thm:ccdce:rtp:ms}} (E);
-      \draw[->] (E) to[sloped,above] node[align=center,font=\scriptsize] {\gls*{scct}\\ \Cref{thm:ccscct:rtp:scct}} (O);
-
-      % Sections
-      \node[above=1.0 of tmsedge] (sectms) {{\scriptsize\Cref{subsec:cs:tms}}};
-      \node[right=0.5 of sectms] (secsms) {{\scriptsize\Cref{subsec:cs:ms}}};
-      \node[right=1.75 of secsms] (secopts) {{\scriptsize\Cref{subsec:cs:opts}}};
-      \node[right=1.75 of secopts] (secscct) {{\scriptsize\Cref{subsec:cs:scct}}};
-
-      \draw[thick,dotted,->] (S) to[bend right=33,sloped] node[align=center,font=\scriptsize] {\gls*{ms}\\ \Cref{thm:ccab:rtp:ms}} (M);
-      \draw[thick,dotted,->] (M) to[bend right=0,sloped] node[align=center,font=\scriptsize] {\gls*{ms}\\ \Cref{thm:cccfccdce:rtp:ms}} (E);
-      \draw[thick,dotted,->] (S) to[out=-45,in=198,sloped] node[align=center,font=\scriptsize] {\gls*{ms}+\gls*{scct}\\ \Cref{thm:ccall:rtp:msscct}} (O);
-
-      \draw[dashed] ($(sectms)-(0.8,0)$) -- ($(sectms)-(0.8,1.25)$);
-      \draw[dashed] ($(sectms)-(-0.9,0)$) -- ($(sectms)-(-0.9,1.25)$);
-      \draw[dashed] ($(secsms)-(-1.2,0)$) -- ($(secsms)-(-1.2,1.25)$);
-      \draw[dashed] ($(secscct)-(1.2,0)$) -- ($(secscct)-(1.2,1.25)$);
-      \draw[dashed] ($(secscct)-(-1.1,0)$) -- ($(secscct)-(-1.1,1.25)$);
-    \end{tikzpicture}
-  \end{center}
-  \caption{Visualisation of the optimising compilation pipeline that attains a combination of \gls*{ms} and \gls*{cct}. %
-    Vertices in the graph are the programming languages from earlier sections (\Cref{sec:casestud:defs}). %
-    All edges are secure compilers, but dotted edges use the presented framework (\Cref{sec:compcomp}) and strikethrough edges classic proof techniques. %
-    The dashed lines partition the graph into the sections where the respective theorems are presented.
-  }\label{fig:pipeline}
-\end{figure}
-The section demonstrates the power of the framework (\Cref{sec:compprop,sec:compcomp}) by composing these compilers for a secure and optimising compilation chain that robustly preserves \gls*{msscct}.
-The first step in this chain is the compiler from $\src{L_{\tmssafe}}$ to $\trg{L}$ that robustly preserves just \gls*{tms} (\Cref{thm:cca:rtp:tms}).
-From here, an instrumentation from $\trg{L}$ to $\irl{L_{\mssafe}}$ ensures that no out-of-bounds accesses can happen and, thus, programs at this point attain \gls*{sms} (\Cref{thm:ccb:rtp:sms}).
-Since these properties compose into \gls*{ms}, composing these passes yields a compiler that robustly preserves \gls*{ms} (\Cref{thm:ccab:rtp:ms}).
-At this stage, the section presents two optimising translations, namely \gls*{cf} and \gls*{dce}, each of which robustly preserves \gls*{ms} (\Cref{thm:ccdce:rtp:ms,thm:cccf:rtp:ms}).
-These translations can be freely ordered in the compilation chain without compromising memory safety (\Cref{thm:cccfccdce:rtp:ms}).
-The last step of the chain ensures that code stays \gls*{scct} (\Cref{thm:ccscct:rtp:scct}) when lowered from $\Lms$ to $\Lscct$.
-The final result is that the whole compilation chain robustly preserves \gls*{msscct} (\Cref{thm:ccall:rtp:msscct}).
-
-\subsection{Robust Temporal Memory Safety Preservation}\label{subsec:cs:tms}
-
-This subsection defines a secure compiler from $\Ltms$ to $\Ltrg$.
-To this end, the compiler needs to ensure that when execution switches from context to component, the type signatures are respected.
-It can do so by inserting dynamic typechecks prior to entering the body of a function belonging to the component.
-
-\begin{gather*}
-  \begin{aligned}
-    \cca(\src{x}) & = \trg{x} \\
-    \cca(\src{n}) & = \trg{n} \\
-    \cca(\src{\lbinop{\expr[_{1}]}{\expr[_{2}]}}) & =  \lbinop[\trg]{\left[\cca(\src{\expr[_{1}]})\right]}{\left[\cca(\src{\expr[_{2}]})\right]} \\
-    \cca(\src{\lget{x}{\expr}}) & = \lget[\trg]{\trg{x}}{\left[\cca(\src{\expr})\right]} \\
-    \cca(\src{\ldelete{x}}) & = \ldelete[\trg]{\left[\cca(\src{x})\right]} \\
-%    \cca(\src{\lset{x}{\expr[_{1}]}{\expr[_{2}]}}) & = \lset[\trg]{\left[\cca(\src{x})\right]}{\left[\cca(\src{\expr[_{1}]})\right]}{\left[\cca(\src{\expr_{2}})\right]} \\
-    \cca(\src{\lfunction{g}{x:\natt\to\type[_{e}]}{\expr}}) & = \lfunction[\trg]{\trg{g}}{\trg{x}}{\lifz[\trg]{\trg{\lhast{x}{\natt}}}{%
-                                                                                                \left[\cca(\src{\expr})\right] %
-                                                                                                 }{\labort[\trg]}}
-  \end{aligned}
-\end{gather*}
-
-Since $\trg{L}$ has no static typechecks, it could happen that a bogus context $\trg{\library_{\ctx}}$ invokes a callable object accepting a $\src{\natt}$ with $\trg{\lpair{17}{29}}$.
-By inserting the check, the compiler ensures that execution does not proceed in such cases.
-The compiler does not insert other checks and proceeds as the identity function (which in this paper amounts to a simple re-colouring of $\src{L_{\tmssafe}}$ to $\trg{L}$ expressions).
-
-Compiling the \texttt{strncpy} function from \Cref{sec:introduction} with $\cca$, the compiler would in this case ensure that the arguments that are evaluated in the compiled \texttt{strncpy} are valid.
-
-\begin{theorem}[Compiler $\cca$ is secure with respect to \gls*{tms}]\label{thm:cca:rtp:tms}
-  $\rtp{\cca}{\tmssafe}$ % \Coqed
-\end{theorem}
-
-\subsubsection{Proving Robust Safety Property Preservation}
-
-We illustrate the proof of \Cref{thm:cca:rtp:tms} since the other secure compilation proofs of this paper follow the same approach.
-Unfolding the theorem statement yields the following assumptions:  
-	for any $\pi\in\lceil\tmssafe\rceil$\footnote{$\lceil\cdot\rceil$ lifts the property to a hyperproperty by applying the powerset operation~\cite{clarkson2008hyper}.}, 
-	$\src{\trace}$, 
-	$\src{\runtimetermvar}$, 
-	and component $\src{\library_{comp}}$, 
-	we have that $\rsat{\src{\library_{comp}}}{\pi}$ and 
-	$\progstepto{\prog[\trg]{\trg{\library_{\ctx}}}{\cca(\src{\library_{\comp}})}}{\trg{\runtimetermvar}}{\trg{\trace}}$, where $\trg{\library_{\ctx}}$ is arbitrary.
-The proof obligation is $\msfilterL{\trg{\trace}}\in\pi$, i.e., the specification trace associated to $\trg{\trace}$ satisfies the property $\pi$.
-A way to show this is to relate trace $\trg{\trace}$ to some $\Ltms$ trace $\src{\trace}$ (which already satisfies the property as per the assumptions).
-The assumptions already contain a target execution associated to this trace, so the task is to find an associated $\Ltms$ execution that yields $\trg{\trace}$.
-The trace $\trg{\trace}$ is split into different parts, as commonly done in secure compilation works~\cite{korashy2021capableptrs,abate2018goodcomps}, where each part contains the events that either the context or the component does, but not both.
-Because of this, all such trace segments are ,,well-bracketed'' in the sense that they start with either $\trg{\ev{Start}}$, $\trg{\ev{Call}\ \comptoctx\ foo\ \valueexpr}$, or $\trg{\ev{Ret}\ \comptoctx\ \valueexpr}$ and end with either $\trg{\ev{End}\ \valueexpr}$, $\trg{\ev{Ret}\ \ctxtocomp\ \valueexpr}$, or $\trg{\ev{Call}\ \ctxtocomp\ foo\ \valueexpr}$.
-In the following, the former is referred to as a context segment, since these executions happen in $\trg{\library_{\ctx}}$, and the latter is referred to as a component segment, since these executions happen in $\cca(\src{\library_{\comp}})$.
-\Cref{fig:proofdiag:rtp} visualises this division for a program execution with one call from context to component and how the target execution is related to a source execution.
-In the figure, the green dashed lines encompass the component segments while the orange boxes contain the actual context switches from context to component or vice versa.
-From a technical perspective, as typically done in compilation proofs, the proof requires some setup to maintain a relation between $\src{\Omega}$ and $\trg{\Omega}$ .
-Two cross-language relations make this precise: (i) $\multimap_{\delta}$ relates states that are involved in a context segment, allowing the target execution to perform internal calls, and (ii) $\approx_{\delta}$ relates states that are involved in a component segment, where both states need to agree exactly, i.e., the memory and the control flow states are required to contain the same information.
-The relations are indexed with $\delta$, which is an injective mapping from $\Ltms$ locations $\src{\loc}$ to $\Ltrg$ locations $\trg{\loc}$.
-Note that the relations $\multimap_{\delta}$ and $\approx_{\delta}$ swap when context switching.
-
-\begin{figure}[!ht]
-  \begin{center}
-    \begin{tikzpicture}[state/.style={minimum height=1.5cm}]
-      % relative horizontal/vertical distance between states
-      \pgfmathsetmacro{\hdist}{0.95}
-      \pgfmathsetmacro{\vdist}{1.35}
-      \pgfmathsetmacro{\halfvdist}{0.725}
-
-      % row of src states
-      \node[state] (srcempty) {$\rtt{\src{\emptyset}}{\src{\lcall{main}{0}}}$};
-      \foreach \s [remember=\s as \cur (initially empty)] in {1,w_1,p,w_2,2} {
-        \node[state,right=\hdist of src\cur] (src\s) {$\rtt{\src{\Omega_{\s}}}{\src{\expr[_{\s}]}}$};
-      }
-      % row of trg states
-      \node[state,below=\vdist of srcempty] (trgempty) {$\rtt{\trg{\emptyset}}{\trg{\lcall{main}{0}}}$};
-      \foreach \s in {1,w_1,p,w_2,2} {
-        \node[state,below=\vdist of src\s] (trg\s) {$\rtt{\trg{\Omega_{\s}}}{\trg{\expr[_{\s}]}}$};
-      }
-      %% illustrations
-        % backtrans wrapper 1
-        \draw[ultra thick,loosely dotted,Peach!50,rounded corners] (src1.north east) -- (srcw\string_1.north east)
-          -- (trgw\string_1.south east) -- (trg1.south west) -- (src1.north west) -- cycle;
-        % backtrans wrapper 2
-        \draw[ultra thick,loosely dotted,Peach!50,rounded corners] (srcp.north east) -- (srcw\string_2.north east)
-          -- (trgw\string_2.south east) -- (trgp.south west) -- (srcp.north west) -- cycle;
-        % compiler correctness
-        \draw[ultra thick,loosely dashed,Emerald!50,rounded corners] ($(srcw\string_1.north east)+(0,0.05)$) -- ($(srcp.north east)+(0,0.05)$)
-          -- ($(trgw\string_2.north east)+(0.05,0)$) -- ($(trgw\string_2.south east)+(0.05,-0.05)$)
-          -- ($(trg1.south west)+(-0.05,-0.05)$) -- ($(trg1.north west)+(-0.05,0)$) -- ($(srcw\string_1.north west)+(0,0.05)$) -- cycle;
-      % state relations
-      \path (srcempty) edge[very thick,draw=gray!25] node[pos=0.5,sloped,rotate=180,fill=white] {\scriptsize$\multimap_\emptyset$} (trgempty)
-        (src1) edge[very thick,draw=gray!25] node[pos=0.5,sloped,rotate=180,fill=white] {\scriptsize$\multimap_{\delta_1}$} (trg1)
-        (srcp) edge[very thick,draw=gray!25] node[pos=0.5,sloped,fill=white] {\scriptsize$\approx_{\delta_p}$} (trgp)
-        (src2) edge[very thick,draw=gray!25] node[pos=0.5,sloped,rotate=180,fill=white] {\scriptsize$\multimap_{\delta_2}$} (trg2)
-        (srcw\string_1) edge[very thick,draw=gray!25] node[pos=0.5,sloped,fill=white] {\scriptsize$\approx_{\delta_{w_1}}$} (trgw\string_1)
-        (srcw\string_2) edge[very thick,draw=gray!25] node[pos=0.5,sloped,rotate=180,fill=white] {\scriptsize$\multimap_{\delta_{w_2}}$} (trgw\string_2)
-        % diagonals
-        (srcw\string_1) edge[thick,draw=gray!25] node[pos=0.16,sloped,rotate=180,fill=white] {\scriptsize$\approx_{\delta_{w_1}}$} (trg1)
-        (src1) edge[thick,draw=gray!25] node[pos=0.16,sloped,rotate=180,fill=white] {\scriptsize$\multimap_{\delta_1}$} (trgw\string_1)
-        (srcp) edge[thick,draw=gray!25] node[pos=0.2,sloped,fill=white] {\scriptsize$\approx_{\delta_p}$} (trgw\string_2)
-        (trgp) edge[thick,draw=gray!25] node[pos=0.2,sloped,fill=white] {\scriptsize$\multimap_{\delta_{w_2}}$} (srcw\string_2)
-        ;
-      %\drawpolygon src1,srcw\string_1,trgw\string_1,trg1;
-      %\drawpolygon srcp,srcw\string_2,trgw\string_2,trgp;
-      %\node[font=\tiny,align=center,above=0.2 of srcw1srcp] (wrapper) {Backtranslation\\Wrapper};
-      %\path[->,draw] (wrapper) -- (srcw\string_1);
-      %\path[->,draw] (wrapper) -- (srcp);
-      % steps
-      \path[very thick,color=\stlccol] (srcempty) edge[draw=none] node {\ $\xrightarrow{\trace[_1]}{}{\kern-3.5pt}_{\text{ctx}}^*$} (src1)
-        (src1) edge[draw=none] node {\ $\xrightarrow{\trace[_c]}{}{\kern-3.5pt}_{\text{ctx}}$} (srcw\string_1)
-        (srcw\string_1) edge[draw=none] node {\ $\xrightarrow{\trace[_p]}{}{\kern-3.5pt}_{\text{ctx}}^*$} (srcp)
-        (srcp) edge[draw=none] node[fill=white,inner sep=0,outer sep=0] {\ $\xrightarrow{\trace[_r]}{}{\kern-3.5pt}_{\text{ctx}}$} (srcw\string_2)
-        (srcw\string_2) edge[draw=none] node {\ $\xrightarrow{\trace[_2]}{}{\kern-3.5pt}_{\text{ctx}}^*$} (src2)
-        ;
-      \path[very thick,color=\ulccol] (trgempty) edge[draw=none] node[fill=white,inner sep=0,outer sep=0] {\ $\xrightarrow{\phantom{\trace[_1]}}{}{\kern-3.5pt}_{\text{ctx}}^*$} (trg1)
-        (trg1) edge[draw=none] node {\ $\xrightarrow{\phantom{\trace[_p]}}{}{\kern-3.5pt}_{\text{ctx}}^*$} (trgw\string_1)
-        (trgw\string_1) edge[draw=none] node[fill=white,inner sep=0,outer sep=0] {\ $\xrightarrow{\phantom{\trace[_p]}}{}{\kern-3.5pt}_{\text{ctx}}^*$} (trgp)
-        (trgp) edge[very thick,draw=none] node {\ $\xrightarrow{\phantom{\trace[_p]}}{}{\kern-3.5pt}_{\text{ctx}}^*$} (trgw\string_2)
-        (trgw\string_2) edge[draw=none] node {\ $\xrightarrow{\phantom{\trace[_2]}}{}{\kern-3.5pt}_{\text{ctx}}^*$} (trg2)
-        ;
-      % legend
-%     \node[align=left,below right=0.3 and 0.3 of trgempty,font=\tiny] (legend) {%
-%       ${\src{\trace[_c]}}\cong{\backt{\Ltms}{\Ltrg}(\trg{Call\ ?\ foo\ \valueexpr})}$\\%
-%       ${\src{\trace[_r]}}\cong{\backt{\Ltms}{\Ltrg}(\trg{Ret\ !\ foo\ \valueexpr})}$
-%       };
-      \node[below=0.5 of trgempty,draw=none] (legend) {};
-      \draw[ultra thick,loosely dotted,Peach!50,rounded corners] ($(legend.north east)+(0.5,-0.4)$) -- ($(legend.north east)+(1,-0.4)$);
-      \node at ($(legend.north east)+(1.5,-0.4)$) (legendwrapper) {};
-      \node[right of=legendwrapper] {\tiny Backtranslation Wrapper};
-      \draw[ultra thick,loosely dashed,Emerald!50,rounded corners] ($(legend.south east)+(0.5,0.4)$) -- ($(legend.south east)+(1,0.4)$);
-      \node at ($(legend.south east)+(1.5,0.4)$) (legendcorrectness) {};
-      \node[right of=legendcorrectness] {\tiny Compiler Correctness};
-    \end{tikzpicture}
-    \caption{Proof diagram for \Cref{thm:cca:rtp:tms} depicting the general structure of robust preservation proofs. %
-      Nodes in the graph represent runtime states. %
-      Vertical lines indicate cross language relations, while horizontal ones are execution steps. %
-      The green dashed trapezoid encompasses the component segment, while the orange dotted rectangles entail the context switches. %
-      $\Ltrg$ traces are omitted for readability. %
-      $\Ltms$ trace segments $\src{\trace[_{c}]}$ and $\src{\trace[_{r}]}$ describe the events that happen at the boundaries, i.e., during a context switch.
-      $\src{\trace_{p}}$ is the behavior of the component and the traces $\src{\trace[_{1}]}$ and $\src{\trace[_{2}]}$ describe the context.
-    }\label{fig:proofdiag:rtp}
-  \end{center}
-\end{figure}
-
-So far, the paper explained how to relate an $\Ltms$ execution with a $\Ltrg$ execution.
-The next question is therefore how to build the corresponding $\Ltms$ execution.
-This is done using a standard secure compilation proof technique called trace-based backtranslation~\cite{abate2019jour,korashy2021capableptrs,patrignani2021rsc}, which can be used to build a context $\src{\library_{\ctx}}$ that behaves similar to $\trg{\library_{\ctx}}$.
-% Since $\Ltms$ is statically typed, the context $\src{\library_{\ctx}}$ obtained from the backtranslation needs to be well-typed when linked with $\src{\library_{\comp}}$.
-For context segments of the trace $\trg{\trace}$ it is also necessary to show that the execution behaves similarily, i.e., the context obtained from the backtranslation generates trace $\src{\trace}$.
-For component segments of the trace, the relatedness of states and traces follows from a compiler correctness argument.
-These two arguments yield the source execution $\progstepto{\src{\prog{\library_{\ctx}}{\library_{\comp}}}}{\src{\runtimetermvar}}{\src{\trace}}$.
-
-The proof now works as follows.
-Given that $\msfilterL{\trg{\trace}}=\msfilterLtms{\src{\trace}}$, the proof goal changes from $\msfilterL{\trg{\trace}}\in\pi$ to $\msfilterLtms{\src{\trace}}\in\pi$.
-This follows by specializing the robust satisfaction assumption $\rsat{\src{\library_{\comp}}}{\pi}$ to use the context $\src{\library_{\ctx}}$, which is obtained from the backtranslation, and to use the source execution $\progstepto{\src{\prog{\library_{\ctx}}{\library_{\comp}}}}{\src{\runtimetermvar}}{\src{\trace}}$.
-
-% \MPpost{
-% 	this can be made more precise
-% }
-
-\subsection{Robust (Spatial) Memory Safety Preservation}\label{subsec:cs:ms}
-
-\begin{center}
-  $$
-  \begin{array}{rcl}
-    \ccb(\trg{\lnew{x}{\expr[_{1}]}{\expr[_{2}]}}) & = & \llet[\irl]{\irl{x_{SIZE}}}{\ccb(\trg{\expr[_{1}]})}{\lnew[\irl]{\irl{x}}{\irl{x_{SIZE}}}{\ccb(\trg{\expr[_{2}]})}} \\
-    \ccb(\trg{\lget{x}{\expr}}) & = & \llet[\irl]{\irl{x_{n}}}{\ccb(\trg{\expr})}{\irl{\lifz{0\le x_{n}<x_{SIZE}}{\lget{x}{x_{n}}}{\labort}}} \\
-    \ccb(\trg{\lset{x}{\expr[_{1}]}{\expr[_{2}]}}) & = & \llet[\irl]{\irl{x_{n}}}{\ccb(\trg{\expr[_{1}]})}{\lifz[\irl]{\irl{0\le x_{n}<x_{SIZE}}}{\lset[\irl]{\irl{x}}{\irl{x_{n}}}{\ccb(\trg{\expr[_{2}]})}}{\irl{\labort}}} \\
-  \end{array}
-  $$
-\end{center}
-
-The compiler $\cc{\trg{L}}{\Lms}$ only inserts bounds-checks whenever reading from or writing to memory in order to enforce \gls*{sms}.
-For passing pointers, it has to pass them with their size information as well.
-To this end, the compiler introduces another, fresh identifier $\irl{x_{SIZE}}$ for each allocation that binds $\irl{x}$ to keep track of the allocation size.
-\begin{example}[Instrumented \texttt{strncpy}]
-  Consider again \texttt{strncpy}, but instrumented for \gls*{sms}:
-    \begin{lstlisting}[language=c,basicstyle=\ttfamily\small, morekeywords={size_t}]
-  void strncpy(size_t n, size_t dst_size, char *dst,
-                         size_t src_size, char *src) {
-    for(size_t i = 0; i < src_size && src[i] != '\0' && i < n; ++i) {
-      if(i < src_size && i < dst_size) {
-        dst[i] = src[i];
-      }
-    }
-  }
-    \end{lstlisting}
-    When calling this in similar fashion to \Cref{ex:strncpy:sms}, the event $\ev{Use}\ \loc_{x}\ 2;\comp;\unlock$ would not be emitted during execution, since the bounds check prevents the condition \texttt{src[i] != '\textbackslash 0'} from executing.
-\end{example}
-
-\begin{theorem}[Compiler $\ccb$ is secure with respect to \gls*{sms}]\label{thm:ccb:rtp:sms}
-  $\rtp{\ccb}{\smssafe}$ % \Coqed
-\end{theorem}
-
-\Cref{thm:ccab:rtp:ms} states that the composition of $\cca$ and $\ccb$ is secure with respect to \gls*{ms} and follows from \Cref{thm:cca:rtp:tms,thm:ccb:rtp:sms} using \Cref{thm:rtp}.
-
-\begin{theorem}[Compiler $\cca\circ\ccb$ is secure with respect to \gls*{ms}]\label{thm:ccab:rtp:ms}
-  $\rtp{\cca\circ\ccb}{\mssafe}$ % \Coqed
-\end{theorem}
-\begin{proof}
-  From \Thmref{thm:cca:rtp:tms} it follows that for any $\Ltms$ program $\src{\progvar}$, it compiles to an $\Ltrg$ program $\trg{\progvar}$ that robustly satisfies \gls*{tms}.
-  Note that $\src{\progvar}$ robustly satisfies \gls*{tms} by the properties of the typesystem of $\Ltms$.
-  Then, \Thmref{thm:ccb:rtp:sms} demonstrates that, assuming $\trg{\progvar}$ robustly satisfies \gls*{sms}, the program $\trg{\progvar}$ compiles to an $\Lms$ program $\irl{p}$ that also robustly satisfies \gls*{sms}.
-  From \Thmref{thm:ccab:rtp:ms} it follows that $\src{p}$ compiles to $\irl{p}$ that robustly satisfies \gls*{ms}, since \gls*{ms} is the intersection of \gls*{tms} and \gls*{sms}.
-\end{proof}
-
-\subsection{Optimising Compilers}\label{subsec:cs:opts}
-
-\begin{gather*}
-  \begin{align*}
-    \ccdce(\irl{\lifz{true}{\expr[_{1}]}{\expr[_{2}]}}) & = \ccdce(\irl{\expr[_{1}]}) &\\
-    \ccdce(\irl{\lifz{false}{\expr[_{1}]}{\expr[_{2}]}}) & = \ccdce(\irl{\expr[_{2}]}) &
-  \end{align*}\\
-  \begin{align*}
-    \ccdce(\irl{\lbinop{\expr[_{1}]}{\expr[_{2}]}}) & = \lbinop[\irl]{\ccdce(\irl{\expr[_{1}]})}{\ccdce(\irl{\expr[_{2}]})} &
-  \end{align*}\\[0.25cm]
-  \begin{align*}
-    \cccf(\irl{\expr}) & = \partialeval{\irl{\expr}}{\irl{\hole{\cdot}}} &
-  \end{align*}\\[0.125cm]
-  \begin{align*}
-    \partialeval{\irl{x}}{\irl{\substlist}} & = \irl{n} & \text{if } \irl{\subst{n}{x}}\in\irl{\substlist} \\
-    \partialeval{\irl{x}}{\irl{\substlist}} & = \irl{x} & \text{if } \irl{\subst{n}{x}}\notin\irl{\substlist} \\
-    \partialeval{\irl{\lbinop{n}{m}}}{\irl{\substlist}} & = \irl{k} & \text{if } \lbinop{\irl{n}}{\irl{m}}=k \\
-%    \partialeval{\irl{\lbinop{\expr[_{1}]}{\expr[_{2}]}}}{\irl{\substlist}} & = \lbinop[\irl]{\partialeval{\irl{\expr[_{1}]}}{\irl{\substlist}}}{\partialeval{\irl{\expr[_{2}]}}{\irl{\substlist}}} \\
-    \partialeval{\irl{\llet{x}{n}{\expr}}}{\irl{\substlist}} & = \partialeval{\irl{\expr}}{\irl{\subst{x}{n}},\irl{\substlist}} \\
-    \partialeval{\irl{\lget{x}{\expr}}}{\irl{\substlist}} & = \lget[\irl]{\irl{x}}{\partialeval{\irl{\expr}}{\irl{\substlist}}}
-  \end{align*}\\
-  \begin{align*}
-    \partialeval{\irl{\llet{x}{\expr[_{1}]}{\expr[_{2}]}}}{\irl{\substlist}} & = \llet[\irl]{\irl{x}}{\partialeval{\irl{\expr[_{1}]}}{\irl{\substlist}}}{\partialeval{\irl{\expr[_{2}]}}{\irl{\substlist}}} \\
-    \partialeval{\irl{\lifz{\expr[_{1}]}{\expr[_{2}]}{\expr[_{3}]}}}{\irl{\substlist}} & = \lifz[\irl]{\partialeval{\irl{\expr[_{1}]}}{\irl{\substlist}}}{\partialeval{\irl{\expr[_{2}]}}{\irl{\substlist}}}{\partialeval{\irl{\expr[_{3}]}}{\irl{\substlist}}} \\
-  \end{align*}
-\end{gather*}
-
-The two optimising compiler passes from $\Lms$ to $\Lms$ perform \gls*{dce} and \gls*{cf}, respectively.
-The \gls*{dce} pass applies a naive rewrite rule on conditionals.
-For \gls*{cf}, the pass uses an auxiliary function \texttt{mix} that does the actual work.
-It rewrites constant binary operations, e.g., $\irl{{17}-1}$ to $\irl{16}$, and replaces variables that are assigned to constants with their constant, e.g., $\irl{\llet{x}{7}{x}}$ to $\irl{7}$.
-Both passes are secure with respect to \gls*{ms}.
-The proof for either is relatively simple, because both \gls*{dce} and \gls*{cf} do not change the way memory accesses happen.
-Moreover, since the input and output languages to these compilers are the same, attacker contexts do not have more power in the target language than in the source.
-
-\begin{theorem}[Compiler $\ccdce$ is secure with respect to \gls*{ms}]\label{thm:ccdce:rtp:ms}
-  $\rtp{\ccdce}{\mssafe}$ %\Coqed
-\end{theorem}
-\begin{theorem}[Compiler $\cccf$ is secure with respect to \gls*{ms}]\label{thm:cccf:rtp:ms}
-  $\rtp{\cccf}{\mssafe}$ %\Coqe
-\end{theorem}
-
-With both \Cref{thm:ccdce:rtp:ms,thm:cccf:rtp:ms} it follows from \Cref{corr:swappable} that the two passes can be interchanged arbitrarily:
-
-\begin{theorem}[Compilers $\cccf\circ\ccdce$ and $\cccf\circ\ccdce$ are secure with respect to \gls*{ms}]\label{thm:cccfccdce:rtp:ms}
-  $\rtp{\cccf\circ\ccdce}{\mssafe}$ and $\rtp{\ccdce\circ\cccf}{\mssafe}$. % \Coqed
-\end{theorem}
-
-\subsection{Robust Strict Cryptographic Constant Time Preservation}\label{subsec:cs:scct}
-
-\begin{center}
-  \begin{align*}
-    \ccscct(\irl{\lfunction{g}{x}{\expr}}) & = \lfunction[\obj]{\obj{g}}{\obj{x}}{\obj{\lwrdoit{1};}\ccscct(\irl{\expr})} \\
-    \ccscct(\irl{\lcall{g}{\expr}}) & = \lcall[\obj]{\obj{g}}{\ccscct(\irl{\expr})\obj{; wrdoit{1}}} \\
-    \ccscct(\irl{\lbinop{\expr[_{1}]}{\expr[_{2}]}}) & = \lbinop[\obj]{\ccscct{\irl{\expr[_{1}]}}}{\ccscct{\irl{\expr[_{2}]}}} \\
-  \end{align*}
-\end{center}
-
-Given the fact that $\Lscct$ provides a \gls*{cct}-mode that can be turned on or off, the compiler inserts wrapper code for function bodies to ensure that execution in the component always happen in this \gls*{cct}-mode.
-The context can overwrite the flag and exit the mode, but upon invoking a function that is part of the component, the flag would be set again.
-Because of this, the compiler is secure with respect to \gls*{scct}, similarly proven as in \Cref{subsec:cs:tms}.
-
-\begin{theorem}[Compiler $\ccscct$ is secure with respect to \gls*{scct}]\label{thm:ccscct:rtp:scct}
-  $\rtp{\ccscct}{\scctsafe}$ % \Coqed
-\end{theorem}
-
-\subsection{Robust Preservation of Intersection of Memory Safety and Strict Cryptographic Constant Time}
-
-Let $\ccmsscct$ be the compiler that is the composition of $\cca$, $\ccb$, $\cccf$, $\ccdce$, and $\ccscct$, then the following theorem holds.
-
-\begin{theorem}[Compiler $\ccmsscct$ is secure with respect to \gls*{scct}]\label{thm:ccall:rtp:msscct}
-  $\rtp{\cc{\Ltms}{\Lscct}}{\mssafe\cap\scctsafe}$ % \Coqed
-\end{theorem}
-\begin{proof}
-	From \Thmref{thm:ccab:rtp:ms}, we have that any $\Ltms$ program $\src{\progvar}$ compiles into a $\Lms$ program $\irl{\progvar}$ that robustly satisfies \gls*{ms}.
-	Then, from \Thmref{thm:cccfccdce:rtp:ms} we have that $\irl{\progvar}$ gets optimised to a program $\irl{\progvar[']}$ that is also \gls*{ms}, where the order of optimisations does not matter for $\irl{\progvar[']}$ to be \gls*{ms}.
-  Assuming $\irl{\progvar[']}$ robustly satisfies \gls*{scct}, by \Thmref{thm:ccscct:rtp:scct} it compiles to an $\Lscct$ program $\obj{\progvar}$ that robustly satisfies \gls*{scct} as well.
-  Finally, from \Thmref{thm:rtp} it follows that, given $\src{\progvar}$ robustly satisfies \gls*{scct} and \gls*{ms}, $\obj{\progvar}$ also robustly satisfies \gls*{scct} and \gls*{ms}.
-\end{proof}
-
-\section{Related Work\pages{2}}\label{sec:relwork}
-
-This section discusses work on robust compilation (\Cref{subsec:relw:seccomprtp}) and on other secure compilation criteria (\Cref{subsec:relw:seccompcrit}).
-Since the case study of \Cref{sec:casestud:defs,sec:casestud:rtp} implements measures for preserving \gls*{ms} and \gls*{cct}, this section then presents relevant related work as well (\Cref{subsec:relw:msmechs,subsec:relw:cctmechs}).
-
-\subsection{Secure Compilation as Robust Preservation}\label{subsec:relw:seccomprtp}
-
-The robust preservation of properties as a compiler-level criterion has been analyzed extensively~\cite{abate2019jour,patrignani2021rsc,abate2021extacc,patrignani2019survey} and thus we build on that framework.
-No existing work is concerned with composing robustly safe compilers.
-These works consider languages with different trace models and our technical setup can be adapted to that as long as security properties and their monitors are still defined on the same trace model.
-The work relating robust preservation with universal composability~\cite{patrignani2022universal} is closest to what this paper presents.
-The authors demonstrate a similar compositionalty theorem to what is presented here (\Cref{sec:compcomp}) but use it in the context of protocols.
-They do not demonstrate the scalability of the approach.
-Moreover, they are missing the upper and lower compositions.
-% The authors demonstrate a similar compositionality theorem to what is presented here (\Cref{sec:compcomp}) as well as in an earlier version of this work~\cite{kruse2022csc}.
-% However, they do not demonstrate the scalability of the approach by means of a case study.
-
-\subsection{Other Secure Compilation Criteria}\label{subsec:relw:seccompcrit}
-
-While this paper focuses on the robust preservation framework~\cite{abate2019jour}, other secure compilation criteria exist.
-The survey on formal approaches to secure compilation~\cite{patrignani2019survey} discusses a broad spectrum already, while this section presents a very high-level overview.
-Fully abstract compilation~\cite{abadi1999fullabstraction} states that a compiler should preserve and reflect observational equivalence between source and target programs.
-It was shown~\cite{abate2021faandrc} that fully abstract compilers robustly preserve program properties that are either trivial or meaningless.
-As a mitigation for this, the authors presented a categorical approach based on maps of distributive laws~\cite{watanabe2002modl}, which they call many maps of distributive laws.
-Maps of distributive laws have been investigated before as a possible secure compilation criterion~\cite{tsampas2020catsc}.
-Other approaches are extensions of the compiler correctness criterion as discussed in other work~\cite{patterson2019next700} or the introduction of opaque observations~\cite{vu2021reconciling} to reconcile compiler optimisations with security.
-Note that this work also presents secure compilers that are optimising, but contrary to the other~\cite{vu2021reconciling}, provides a formal account of these in the robust preservation framework.
-% Lastly, the authors of this paper have presented ongoing work~\cite{patrignani2023blame} on a weaker robust preservation criteria based on the concept blame.
-
-\subsection{Memory Safety Mechanisms}\label{subsec:relw:msmechs}
-
-Different mechanisms for enforcing memory safety exist that also consider the secure compilation domain, i.e., have an active attacker model.
-For example, the ,,pointers as capabilities'' principle represents pointers as machine-level capabilities~\cite{korashy2021capableptrs}, which behave in a similar fashion to capabilities by means of linear typing~\cite{morrisett2005L3}.
-The approach of this paper also uses linear typing, but differs from $L^{3}$~\cite{morrisett2005L3} in the way that functions are not first-class.
-Moreover, this paper considers an active attacker, while the work on $L^{3}$ only discusses whole programs and, thus, has no active attacker model.
-The instrumentation to ensure memory safety that this paper presents is inspired by Softbounds~\cite{nagarakatte2009soft}.
-That work inserts bounds-checks in front of pointer-dereferences and, for this to work, inserts meta-data information on pointer creation.
-Softbounds also works in a more advanced setting with structured fields accesses and also introduces a table-lookup for pointers that are stored in memory.
-This paper only considers arrays of primitive data, i.e., there are no pointers to pointers or structures.
-Several other approaches to memory-safety exist in literature, specifically as compiler instrumentations~\cite{akritidis2009baggy,younan2010paricheck,jung2021pico,shankaranarayana2023tailcheck,dhumbumroong2020boundwarden,nam2019framer,zhou2023fatptrs}, hardware-extensions~\cite{kwon2013lowfat,saileshwar2022heapcheck,chen2023flexpointer,kim2023whistle}, or programming language extensions~\cite{elliott2018checkedc,li2022formalcheckedc,jim2002cyclone,elliott2015guilt,west2005cuckoo,weis2019fyr,benoit2019uniqueness}.
-What differentiates this work from them is that this work uses known, compiler-based approaches to ensure memory-safety as a means to investigate secure compiler compositions.
-This paper does not provide efficient memory-safety, but serves as a theoretical foundation for the secure compilation domain.
-
-To extend the languages in this paper with a less restricted form of pointer arithmetic, the region coloring memory safety monitor presented in earlier work~\cite{michael2023mswasm} can be used.
-The work presenting this monitor provides an approach for the robust preservation of memory safety compiling from C to WASM.
-However, they do not discuss composition of secure compilers but rather investigate an instance of a secure compiler.
-
-\subsection{Cryptographic Constant Time Mechanisms}\label{subsec:relw:cctmechs}
-
-The approach to preserving cryptographic constant time in this paper is high-level, where a programming language exposes a way to switch the semantics to a data (operand) independent timing mode.
-Since identifiers in $\Lscct$ are annotated with a secrecy tag, this approach is similar to others with information flow control.
-For example, Vale~\cite{bond2017vale} uses Dafny to ensure constant-time assembly code, while Jasmin~\cite{almeida2017jasmin} makes use of the Coq proof assistant to reject non-constant-time programs.
-CT-Wasm~\cite{watt2019ctwasm} enforces constant-timeness by means of a type system.
-Different to the approach of this paper, these approaches necessitate that the programmer writes \gls*{cct} code.
-An approach to allow programmers to write more high-level code is CryptOpt~\cite{kuepper2023cryptopt}, which generates efficient target-code by means of a randomised search.
-This paper abstracts over concrete mitigation strategies and simply assumes that there is a flag to switch to a cryptographic-constant time execution mode.
-This can be realised by employing the FaCT~\cite{cauligi2019fact} compiler, which translates common non-constant time code patterns to be constant-time, and the data (object) independent timing execution mode of modern processors.
-
-\section{Conclusion\pages{$\sfrac{1}{2}$}}\label{sec:concl}
-This paper tackled the problem of understanding what kind of security properties does a secure compiler preserve, when said compiler is the combination of compiler passes that preserve possibly different security properties.
-% 
-For this, this paper first formalised security properties of interest and their composition.
-% 
-Then, it proved that composing secure compilers that preserve certain properties results in a secure compiler that preserves the composition of these properties.
-% 
-Finally, this paper defines a multi-pass compiler and proves that it preserves \gls*{msscct}.
-Crucially, this paper derives the security of the multi-pass compiler from the composition of the security properties preserved by its individual passes, which include security-preserving as well as optimisation passes.
-
-% This paper does a first step towards practical secure compilation chains by introducing a theoretical framework demonstrating that secure compilers compose.
-% The paper exercises this framework on a case-study that consists of an optimising compilation chain that is secure with respect to \gls*{msscct}.
-% \MP{
-% 	not an appropriate recap.
-% 	rewrite summarising the paper
-% }
-
-% In future work, it would be interesting to investigate whether it is possible to provide {\em secure compiler combinators}, similarly to how parser combinators work.
-% %To this end, it may be necessary to extend existing frameworks for multi-language semantics.
-% It is also interesting to extend our case-study and verify all results of it in Coq.
-% This way, it is possible to obtain an executable, secure compiler.
-% However, the formalisation effort for backtranslation proofs is known to be enormous\MP{cit + details}, so another research avenue may be to find better ways to (semi-)automate standard secure compilation proofs.
-% \MPin{
-% 	not particularly informative FW. i'd rather not have it as such
-% }
-
-%% The acknowledgments section is defined using the "acks" environment
-%% (and NOT an unnumbered section). This ensures the proper
-%% identification of the section in the article metadata, and the
-%% consistent spelling of the heading.
-
-% \begin{acks}
-%   We would like to thank the anonymous reviewers for their feedback.
-% 	This work was partially supported by a gift from
-		% the Italian Ministry of Education through funding for the Rita Levi Montalcini grant (call of 2019).
-% \end{acks}
-
-%\newpage
-%%
-%% The next two lines define the bibliography style to be used, and
-%% the bibliography file.
 \bibliographystyle{ACM-Reference-Format}
 \bibliography{main}
-
-%%
-%% If your work has an appendix, this is the place to put it.
-\appendix
 

### 2023-08-04 tech/main.tex
tech report monitors and properties

@@
 
+\appendix
+
+\section{General Trace Model}\label{ap:def:gtm}
+
+\begin{gather*}
+  \begin{aligned}
+  \mi{(Security\ Tag)}~\sigma\bnfdef&\ \lock \mid \unlock\hspace{0.5cm}
+  \mi{(Control\ Tag)}~\sandboxtag\bnfdef \ctx\mid\comp\hspace{0.5cm}
+  \mi{(Event)}~\event\bnfdef\ \emptyevent \mid \lightning \mid \event[_{b}];\sandboxtag;\sigma \\
+  \mi{(Pre\text{-}event)}&~\event[_{b}]\bnfdef\ \ev{Alloc\ \loc\ n} \mid \ev{Dealloc\ \loc} \mid \ev{Use\ \loc\ n} \mid \ev{Branch\ n} \mid \ev{Binop\ n} \\
+  \end{aligned}
+\end{gather*}
+
+Throughout the paper, sequences are indicated with an overbar (i.e., $\trace$), empty sequences with $\hole{\cdot}$, and concatenation of sequences $\trace[_{1}],\trace[_{2}]$ as $\trace[_{1}]\cdot\trace[_{2}]$.
+Prepending elements to sequences uses the same notation: $\event\cdot\trace$.
+
+\begin{definition}\label{ap:def:property}
+  A property $\pi$ is a set of traces $\trace$.
+\end{definition}
+\begin{definition}\label{ap:def:propclass}
+  A class $\class$ is a set of properties $\pi$.
+\end{definition}
+\begin{definition}\label{ap:def:lifting:properties}
+  The lifting $\lift{\pi}$ of property $\pi$ is its powerset.
+\end{definition}
+
+\section{Properties and Monitors}\label{ap:sec:propmon}
+
+In the following, $\traceagree[]{\pi}{\event}{\absevent}$ and the structure of $\abstrace$ are left abstract.
+These are monitor-specific definitions, which are defined in \Cref{ap:subsec:tms,ap:subsec:sms,ap:subsec:ms,ap:subsec:scct,ap:subsec:msscct}.
+
+\begin{figure}[!ht]
+  \judgbox{\traceagree{\pi}{\trace}{\abstrace}}{,,Specification-level trace $\trace$ agrees with monitor-level trace $\abstrace$ with respect to property $\pi$.''}
+  %
+  \typerule{traceagree-empty}{}{
+    \traceagree{\pi}{\hole{\cdot}}{\hole{\cdot}}
+  }{traceagree-empty}
+  %
+  \typerule{traceagree-ign-L}{
+    \traceagree{\pi}{\trace}{\abstrace}
+  }{
+    \traceagree{\pi}{\emptyevent\cdot\trace}{\abstrace}
+  }{traceagree-ign-L}
+  %
+  \typerule{traceagree-ign-R}{
+    \traceagree{\pi}{\trace}{\abstrace}
+  }{
+    \traceagree{\pi}{\trace}{\absev{\emptyevent}\cdot\abstrace}
+  }{traceagree-ign-R}
+  %
+  \typerule{traceagree-cons}{
+    \traceagree[]{\pi}{\event}{\absevent} &
+    \traceagree{\pi}{\trace}{\abstrace}
+  }{
+    \traceagree{\pi}{\event\cdot\trace}{\absevent\cdot\abstrace}
+  }{traceagree-cons}
+  \caption{Trace-Agreement relation that equates specification-level traces with monitor-level traces.}\label{fig:stuttereq}
+\end{figure}
+
+\bul{A specification trace $\trace$ monitor-satisfies property $\pi$} iff there exists a (final) monitor state $\monitor$ and an abstract trace $\abstrace$ such that the \iul{specification trace $\trace$ agrees with abstract trace $\abstrace$} and \oul{the initial monitor can step to the (final) monitor state $\monitor$ with abstract trace $\abstrace$}.
+
+\begin{definition}[Monitor Satisfaction]\label{def:monsat}
+  \bul{$\monsafe{\trace}{\pi}$} $\isdef$ $\exists \abstrace\ \monitor$, \iul{$\traceagree{\pi}{\trace}{\abstrace}$} and \oul{$\monitorcheck{\emptyset}{\monitor}{\abstrace}$}.
+\end{definition}
+
+\subsection{Temporal Memory Safety}\label{ap:subsec:tms}
+
+\begin{definition}[\glsfirst*{tms}]\label{ap:def:trace:tmsdef}
+  \begin{displaymath}
+  \tmssafe:=\left\{\trace \left| \begin{array}{rcl}
+                              \ev{Alloc\ \loc\ n;\sandboxtag;\securitytag}&\le_{\trace}&\ev{Dealloc\ \loc;\sandboxtag;\securitytag'} \\
+                              \ev{Use\ \loc\ n;\sandboxtag;\securitytag}&\le_{\trace}&\ev{Dealloc\ \loc;\sandboxtag;\securitytag'} \\
+                              \text{if }\ev{Alloc\ \loc\ n;\sandboxtag;\securitytag}\text{ in }\trace&\text{then}&\ev{Dealloc\ \loc;\sandboxtag;\securitytag'}\text{ in }\trace \\
+                              \text{at most one }\ev{Dealloc\ \loc;\sandboxtag;\securitytag}&\text{in}&\trace \\
+                              \text{at most one }\ev{Alloc\ \loc\ n;\sandboxtag;\securitytag}&\text{in}&\trace \\
+                            \end{array}\right.\right\}
+  \end{displaymath}
+\end{definition}
+Hereby, the notation $\event[_{1}]\le_{\trace}\event[_{2}]$ means that if $\event[_{1}]$ is in $\trace$ and if $\event[_{2}]$ is in $\trace$, then $\event[_{1}]$ appears before $\event[_{2}]$.
+
+$$\;$$
+\hrule
+$$\;$$
+\begin{gather*}
+  \begin{aligned}
+    \mi{(Abstract\ Store)}~\tmsmonitor\bnfdef&\ \left\{\text{allocated}:L\times\sandboxtag,\text{freed}:L\times\sandboxtag\right\} \\
+    \mi{(Initial\ Abstract\ Store)}~\emptytmsmon:=&\ \left\{\text{allocated}:\emptyset,\text{freed}:\emptyset\right\}\\
+    \mi{(Abstract\ Events)}~\tmsevent\bnfdef&\ \absev{\emptyevent} \mid \absev{Alloc\ }\loc\ \sandboxtag \mid \absev{Dealloc\ }\loc\ \sandboxtag \mid \absev{Use\ }\loc\ \sandboxtag \mid \absterm\\
+  \end{aligned}
+\end{gather*}
+\begin{center}
+  \judgbox{\monitorcheck[]{\tmsmonitor}{\tmsmonitor[']}{\tmsevent}}{,,Monitor $\tmsmonitor$ does one step to $\tmsmonitor[']$ given event $\tmsevent$.''}$\;$\\
+  %
+  \typerule{tms-uninteresting}{
+  }{
+    \monitorcheck[]{\tmsmonitor}{\tmsmonitor}{\absev{\emptyevent}}
+  }{ap:tms-uninteresting}
+  %
+  \typerule{tms-abort}{
+  }{
+    \monitorcheck[]{\tmsmonitor}{\tmsmonitor}{\absterm}
+  }{ap:tms-abort}
+  %
+  \typerule{tms-use}{
+    (\loc;\sandboxtag)\in\tmsmonitor.\text{allocated} &
+    (\loc;\sandboxtag)\notin\tmsmonitor.\text{freed}
+  }{
+    \monitorcheck[]{\tmsmonitor}{\tmsmonitor}{\absev{Use\ }\loc\ \sandboxtag}
+  }{ap:tms-use}
+  %
+  \typerule{tms-alloc}{
+    (\loc;\sandboxtag)\notin\tmsmonitor.\text{allocated} &
+    (\loc;\sandboxtag)\notin\tmsmonitor.\text{freed} \\
+    \tmsmonitor[']=\left\{\text{allocated}: \tmsmonitor.\text{allocated}\cup\left\{(\loc;\sandboxtag)\right\},\text{freed}: \tmsmonitor.\text{freed}\right\}
+  }{
+    \monitorcheck[]{\tmsmonitor}{\tmsmonitor[']}{\absev{Alloc\ }\loc\ \sandboxtag}
+  }{ap:tms-alloc}
+  %
+  \typerule{tms-dealloc}{
+    (\loc;\sandboxtag)\in\tmsmonitor.\text{allocated} &
+    (\loc;\sandboxtag)\notin\tmsmonitor.\text{freed} \\
+    \tmsmonitor[']=\left\{\text{allocated}: \tmsmonitor.\text{allocated}\setminus\left\{(\loc;\sandboxtag)\right\},\text{freed}: \tmsmonitor.\text{freed}\cup\left\{(\loc;\sandboxtag)\right\}\right\}
+  }{
+    \monitorcheck[]{\tmsmonitor}{\tmsmonitor[']}{\absev{Dealloc\ }\loc\ \sandboxtag}
+  }{ap:tms-dealloc}
+\end{center}
+\begin{center}
+  \judgbox{\monitorcheck{\tmsmonitor}{\tmsmonitor[']}{\tmstrace}}{,,Monitor $\tmsmonitor$ multi-steps to $\tmsmonitor[']$ given trace $\tmstrace$.''}$\;$\\
+  %
+  \typerule{tms-refl}{
+    \tmsmonitor.\text{allocated}=\emptyset
+  }{
+    \monitorcheck{\tmsmonitor}{\tmsmonitor}{\hole{\cdot}}
+  }{ap:tms-refl}
+  %
+  \typerule{tms-ign-trans}{
+    \monitorcheck{\tmsmonitor}{\tmsmonitor[']}{\tmstrace}
+  }{
+    \monitorcheck[]{\tmsmonitor}{\tmsmonitor[']}{\absev{\emptyevent}\cdot\tmstrace}
+  }{ap:tms-ign-trans}
+  %
+  \typerule{tms-trans}{
+    \monitorcheck[]{\tmsmonitor}{\tmsmonitor[']}{\tmsevent} &
+    \monitorcheck{\tmsmonitor[']}{\tmsmonitor['']}{\tmstrace}
+  }{
+    \monitorcheck[]{\tmsmonitor}{\tmsmonitor['']}{\tmsevent\cdot\tmstrace}
+  }{ap:tms-trans}
+\end{center}
+\begin{center}
+  \judgbox{\tmstraceagree[]{\event}{\tmsevent}}{,,Abstract event $\tmsevent$ is equivalent to $\event$ with respect to \gls*{tms}.''}$\;$\\
+  %
+  \typerule{tms-alloc-authentic}{
+  }{
+    \tmstraceagree[]{\ev{Alloc}\ \loc\ n;\sandboxtag;\securitytag}{\absev{Alloc\ }\loc\ \sandboxtag}
+  }{ap:tms-alloc-authentic}
+  %
+  \typerule{tms-dealloc-authentic}{
+  }{
+    \tmstraceagree[]{\ev{Dealloc}\ \loc;\sandboxtag;\securitytag}{\absev{Dealloc\ }\loc}
+  }{ap:tms-dealloc-authentic}
+  %
+  \typerule{tms-use-authentic}{
+  }{
+    \tmstraceagree[]{\ev{Use}\ \loc\ n;\sandboxtag;\securitytag}{\absev{Use\ }\loc}
+  }{ap:tms-use-authentic}
+  %
+  \typerule{tms-branch-authentic}{
+  }{
+    \tmstraceagree[]{\ev{Branch\ }n}{\absev{\emptyevent}}
+  }{ap:tms-branch-authentic}
+  %
+  \typerule{tms-binop-authentic}{
+  }{
+    \tmstraceagree[]{\ev{Binop\ }n}{\absev{\emptyevent}}
+  }{ap:tms-binop-authentic}
+  %
+  \typerule{tms-empty-authentic}{
+  }{
+    \tmstraceagree[]{\emptyevent}{\absev{\emptyevent}}
+  }{ap:tms-empty-authentic}
+  %
+  \typerule{tms-abort-authentic}{
+  }{
+    \tmstraceagree[]{\lightning}{\absterm}
+  }{ap:tms-abort-authentic}
+\end{center}
+
+\begin{lemma}[Traces with Monitor Satisfaction are $\tmssafe$]\label{ap:lem:mon:tmssafe}
+  If $\montmssafe{\trace}$, then $\trace\in\tmssafe$. %\Coqed
+\end{lemma}
+\begin{incompleteproof}
+\end{incompleteproof}
+
+\subsection{Spatial Memory Safety}\label{ap:subsec:sms}
+
+\begin{definition}[\glsfirst*{sms}]\label{ap:def:trace:smsdef}
+  \begin{displaymath}
+  \smssafe:=\left\{\trace \left|\begin{array}{rcl}
+                                  \text{If }\ev{Alloc\ \loc\ n;\sandboxtag;\securitytag}\le_{\trace}\ev{Use\ \loc\ m;\sandboxtag;\securitytag'}, \text{ then }m<n
+                                \end{array}\right.\right\}
+  \end{displaymath}
+\end{definition}
+$$\;$$
+\hrule
+$$\;$$
+\begin{gather*}
+  \begin{aligned}
+    \mi{(Abstract\ Store)}~\smsmonitor:=&\ L\times\sandboxtag\times\mb{N}\mid\emptysmsmon\\
+    \mi{(Abstract\ Events)}~\smsevent\bnfdef&\ \absev{\varepsilon} \mid \absev{Alloc\ }\loc\ \sandboxtag\ n \mid \absev{Use\ }\loc\ \sandboxtag\ n \\
+  \end{aligned}
+\end{gather*}
+\begin{center}
+  \judgbox{\monitorcheck[]{\smsmonitor}{\smsmonitor[']}{\smsevent}}{,,Monitor $\smsmonitor$ does one step to $\smsmonitor[']$ given event $\smsevent$.''}$\;$\\
+  %
+  \typerule{sms-uninteresting}{
+  }{
+    \monitorcheck[]{\smsmonitor}{\smsmonitor}{\absev{\emptyevent}}
+  }{ap:sms-uninteresting}
+  %
+  \typerule{sms-abort}{
+  }{
+    \monitorcheck[]{\smsmonitor}{\smsmonitor}{\absterm}
+  }{ap:sms-abort}
+  %
+  \typerule{sms-use}{
+    (\loc;\sandboxtag;m)\in\smsmonitor &
+    n<m
+  }{
+    \monitorcheck[]{\smsmonitor}{\smsmonitor}{\absev{Use\ }\loc\ \sandboxtag\ n}
+  }{ap:sms-use}
+  %
+  \typerule{sms-alloc}{
+    (\loc;\sandboxtag;m)\notin\smsmonitor
+  }{
+    \monitorcheck[]{\smsmonitor}{\smsmonitor\cup\left\{(\loc;\sandboxtag;n)\right\}}{\absev{Alloc\ }\loc\ \sandboxtag\ n}
+  }{ap:sms-alloc}
+\end{center}
+\begin{center}
+  \judgbox{\monitorcheck{\smsmonitor}{\smsmonitor[']}{\smstrace}}{,,Monitor $\smsmonitor$ multi-steps to $\smsmonitor[']$ given trace $\smstrace$.''}$\;$\\
+  %
+  \typerule{sms-refl}{
+  }{
+    \monitorcheck{\smsmonitor}{\smsmonitor}{\hole{\cdot}}
+  }{ap:sms-refl}
+  %
+  \typerule{sms-ign-trans}{
+    \monitorcheck{\smsmonitor}{\smsmonitor[']}{\smstrace}
+  }{
+    \monitorcheck[]{\smsmonitor}{\smsmonitor[']}{\absev{\emptyevent}\cdot\smstrace}
+  }{ap:sms-ign-trans}
+  %
+  \typerule{sms-trans}{
+    \monitorcheck[]{\smsmonitor}{\smsmonitor[']}{\smsevent} &
+    \monitorcheck{\smsmonitor[']}{\smsmonitor['']}{\smstrace}
+  }{
+    \monitorcheck[]{\smsmonitor}{\smsmonitor['']}{\smsevent\cdot\smstrace}
+  }{ap:sms-trans}
+\end{center}
+\begin{center}
+  \judgbox{\smstraceagree[]{\event}{\smsevent}}{,,Abstract event $\smsevent$ is equivalent to $\event$ with respect to \gls*{sms}.''}$\;$\\
+  %
+  \typerule{sms-alloc-authentic}{
+  }{
+    \smstraceagree[]{\ev{Alloc}\ \loc\ n;\sandboxtag;\securitytag}{\absev{Alloc\ }\loc\ n}
+  }{ap:sms-alloc-authentic}
+  %
+  \typerule{sms-dealloc-authentic}{
+  }{
+    \smstraceagree[]{\ev{Dealloc}\ \loc;\sandboxtag;\securitytag}{\absev{Dealloc\ }\loc}
+  }{ap:sms-dealloc-authentic}
+  %
+  \typerule{sms-use-authentic}{
+  }{
+    \smstraceagree[]{\ev{Use}\ \loc\ n;\sandboxtag;\securitytag}{\absev{Use\ }\loc\ n}
+  }{ap:sms-use-authentic}
+  %
+  \typerule{sms-branch-authentic}{
+  }{
+    \smstraceagree[]{\ev{Branch\ }n}{\absev{\emptyevent}}
+  }{ap:sms-branch-authentic}
+  %
+  \typerule{sms-binop-authentic}{
+  }{
+    \smstraceagree[]{\ev{Binop\ }n}{\absev{\emptyevent}}
+  }{ap:sms-binop-authentic}
+  %
+  \typerule{sms-empty-authentic}{
+  }{
+    \smstraceagree[]{\emptyevent}{\absev{\emptyevent}}
+  }{ap:sms-empty-authentic}
+  %
+  \typerule{sms-abort-authentic}{
+  }{
+    \smstraceagree[]{\lightning}{\absterm}
+  }{ap:sms-abort-authentic}
+\end{center}
+\begin{lemma}[Traces with Monitor Satisfaction are $\smssafe$]\label{ap:lem:mon:smssafe}
+  If $\monsmssafe{\trace}$, then $\trace\in\smssafe$.\Coqed
+\end{lemma}
+\begin{proof}
+  First, generalize \Cref{ap:def:trace:smsdef} as follows:
+  \begin{displaymath}
+    \gsmssafe:=\left\{\trace \left|\begin{array}{rcl}
+                                    \text{If }\ev{Alloc\ \loc\ n;\sandboxtag;\securitytag}\le_{\trace}\ev{Use\ \loc\ m;\sandboxtag;\securitytag'}\\
+                                    \text{or }(\ev{Use\ \loc\ m;\sandboxtag;\securitytag}\text{ in }\trace\text{ and }(\loc,\sandboxtag,n)\in\smsmonitor),\\
+                                    \text{ then }m<n
+                                  \end{array}\right.\right\}
+  \end{displaymath}
+  Then, show ,,If $\mongsmssafe{\trace}$, then $\trace\in\gsmssafe$.'' by inducing on the monitor steps.
+
+  Now specialize this generalization to $\smsmonitor=\emptyset$ and the claim follows as a corollary.
+\end{proof}
+
+\subsection{Memory Safety}\label{ap:subsec:ms}
+
+\begin{definition}[\glsfirst*{ms}]\label{ap:def:trace:msdef}
+  \begin{displaymath}
+    \mssafe:=\tmssafe \cap \smssafe
+  \end{displaymath}
+\end{definition}
+$$\;$$
+\hrule
+$$\;$$
+\begin{gather*}
+  \begin{aligned}
+    \mi{(Abstract\ Store)}~\msmonitor:=&\ \tmsmonitor\times\smsmonitor\\
+    \mi{(Initial\ Abstract\ Store)}~\emptymsmon:=&\ (\emptytmsmon, \emptysmsmon) \\
+    \mi{(Abstract\ Events)}~\msevent:=&\ \tmsevent\times\smsevent \mid \absev{\emptyevent}\\
+  \end{aligned}
+\end{gather*}
+\begin{center}
+  \judgbox{\monitorcheck[]{\msmonitor}{\msmonitor[']}{\msevent}}{,,Monitor $\msmonitor$ does one step to $\msmonitor[']$ given event $\msevent$.''}$\;$\\
+  %
+  \typerule{ms-step}{
+    \monitorcheck[]{\tmsmonitor}{\tmsmonitor[']}{\tmsevent} &
+    \monitorcheck[]{\smsmonitor}{\smsmonitor[']}{\smsevent}
+  }{
+    \monitorcheck[]{(\tmsmonitor,\smsmonitor)}{(\tmsmonitor['],\smsmonitor['])}{(\tmsevent,\smsevent)}
+  }{ap:ms-step}
+\end{center}
+\begin{center}
+  \judgbox{\monitorcheck{\msmonitor}{\msmonitor[']}{\mstrace}}{,,Monitor $\msmonitor$ multi-steps to $\msmonitor[']$ given trace $\mstrace$.''}$\;$\\
+  %
+  \typerule{ms-refl}{
+    \monitorcheck{\tmsmonitor}{\tmsmonitor}{\hole{\cdot}} &
+    \monitorcheck{\smsmonitor}{\smsmonitor}{\hole{\cdot}}
+  }{
+    \monitorcheck{(\tmsmonitor,\smsmonitor)}{(\tmsmonitor,\smsmonitor)}{\hole{\cdot}}
+  }{ap:ms-refl}
+  %
+  \typerule{ms-ign-trans}{
+    \monitorcheck{\msmonitor}{\msmonitor[']}{\mstrace}
+  }{
+    \monitorcheck[]{\msmonitor}{\msmonitor[']}{\absev{\emptyevent}\cdot\mstrace}
+  }{ap:ms-ign-trans}
+  %
+  \typerule{ms-trans}{
+    \monitorcheck[]{\msmonitor}{\msmonitor[']}{\msevent} &
+    \monitorcheck{\msmonitor[']}{\msmonitor['']}{\mstrace}
+  }{
+    \monitorcheck[]{\msmonitor}{\msmonitor['']}{\msevent\cdot\mstrace}
+  }{ap:ms-trans}
+\end{center}
+\begin{center}
+  \judgbox{\mstraceagree[]{\event}{\msevent}}{,,Abstract event $\msevent$ is equivalent to $\event$ with respect to \gls*{ms}.''}$\;$\\
+  %
+  \typerule{ms-some-authentic}{
+    \tmstraceagree[]{\event}{\tmsevent} &
+    \smstraceagree[]{\event}{\smsevent}
+  }{
+    \mstraceagree[]{\event}{(\tmsevent,\smsevent)}
+  }{ap:ms-some-authentic}
+  %
+  \typerule{ms-none-authentic}{
+    \tmstraceagree[]{\event}{\absev{\emptyevent}} &
+    \smstraceagree[]{\event}{\absev{\emptyevent}}
+  }{
+    \mstraceagree[]{\event}{\absev{\emptyevent}}
+  }{ap:ms-none-authentic}
+\end{center}
+
+\begin{lemma}[Traces with Monitor Satisfaction are $\mssafe$]\label{ap:lem:mon:mssafe}
+  If $\monmssafe{\trace}$, then $\trace\in\mssafe$.\Coqed
+\end{lemma}
+\begin{proof}
+  Corollary of \Cref{ap:lem:mon:tmssafe,ap:lem:mon:smssafe}.
+\end{proof}
+
+\subsection{Strict Cryptographic Constant Time}\label{ap:subsec:scct}
+
+\begin{definition}[\glsfirst*{scct}]\label{ap:def:trace:scctdef}
+  \begin{displaymath}
+    \scctsafe:=\left\{\trace \left|\begin{array}{rcl}
+                                     \trace=\hole{\cdot} & \text{ or } & \trace=\event[_{b}];\sandboxtag;\unlock\cdot\trace['] \wedge \trace[']\in\scctsafe
+                                   \end{array}\right.\right\}
+  \end{displaymath}
+\end{definition}
+$$\;$$
+\hrule
+$$\;$$
+\begin{gather*}
+  \begin{aligned}
+    \mi{(Abstract\ Store)}~\scctmonitor:=&\ \emptyscctmon\\
+    \mi{(Abstract\ Events)}~\scctevent:=&\ \absev{\emptyevent} \mid \absev{\lightning} \mid \absev{Any} \\
+  \end{aligned}
+\end{gather*}
+\begin{center}
+  \judgbox{\monitorcheck[]{\scctmonitor}{\scctmonitor[']}{\scctevent}}{,,Monitor $\scctmonitor$ does one step to $\scctmonitor[']$ given event $\scctevent$.''}$\;$\\
+  %
+  \typerule{scct-none}{
+  }{
+    \monitorcheck[]{\scctmonitor}{\scctmonitor[']}{\absev{\emptyevent}}
+  }{ap:scct-none}
+  %
+  \typerule{scct-abort}{
+  }{
+    \monitorcheck[]{\scctmonitor}{\scctmonitor[']}{\absev{\lightning}}
+  }{ap:scct-abort}
+\end{center}
+\begin{center}
+  \judgbox{\monitorcheck{\scctmonitor}{\scctmonitor[']}{\sccttrace}}{,,Monitor $\scctmonitor$ multi-steps to $\scctmonitor[']$ given trace $\sccttrace$.''}$\;$\\
+  %
+  \typerule{scct-refl}{
+  }{
+    \monitorcheck{\scctmonitor}{\scctmonitor}{\hole{\cdot}}
+  }{ap:scct-refl}
+  %
+  \typerule{scct-ign-trans}{
+    \monitorcheck{\scctmonitor}{\scctmonitor[']}{\sccttrace}
+  }{
+    \monitorcheck[]{\scctmonitor}{\scctmonitor[']}{\absev{\emptyevent}\cdot\sccttrace}
+  }{ap:scct-ign-trans}
+  %
+  \typerule{scct-trans}{
+    \monitorcheck[]{\scctmonitor}{\scctmonitor[']}{\scctevent} &
+    \monitorcheck{\scctmonitor[']}{\scctmonitor['']}{\sccttrace}
+  }{
+    \monitorcheck[]{\scctmonitor}{\scctmonitor['']}{\scctevent\cdot\sccttrace}
+  }{ap:scct-trans}
+\end{center}
+\begin{center}
+  \judgbox{\sccttraceagree[]{\event}{\absevent}}{,,Abstract event $\absevent$ is equivalent to $\event$ with respect to \gls*{cct}.''}$\;$\\
+  %
+  \typerule{scct-low-authentic}{
+  }{
+    \sccttraceagree[]{\event[_b];\sandboxtag;\unlock}{\absev{\emptyevent}}
+  }{ap:scct-low-authentic}
+  %
+  \typerule{scct-high-authentic}{
+  }{
+    \sccttraceagree[]{\event[_b];\sandboxtag;\lock}{\absev{Any}}
+  }{ap:scct-high-authentic}
+  %
+  \typerule{scct-empty-authentic}{
+  }{
+    \sccttraceagree[]{\emptyevent}{\absev{\emptyevent}}
+  }{ap:scct-empty-authentic}
+  %
+  \typerule{scct-abort-authentic}{
+  }{
+    \sccttraceagree[]{\lightning}{\absterm}
+  }{ap:scct-abort-authentic}
+\end{center}
+\begin{lemma}[Monitor Traces are $\scctsafe$]\label{lem:mon:scctsafe}
+  If $\monscctsafe{\trace}$, then $\trace\in\scctsafe$. %\Coqed
+\end{lemma}
+\begin{incompleteproof}
+\end{incompleteproof}
+
+\subsection{Memory Safety and Strict Cryptographic Constant Time}\label{ap:subsec:msscct}
+\begin{definition}[\glsfirst*{msscct}]\label{ap:def:trace:msscctdef}
+  \begin{displaymath}
+    \msscctsafe:=\mssafe\cap\scctsafe
+  \end{displaymath}
+\end{definition}
+$$\;$$
+\hrule
+$$\;$$
+
+\begin{gather*}
+  \begin{aligned}
+    \mi{(Abstract\ Store)}~\msscctmonitor:=&\ \msmonitor\times\scctmonitor\\
+    \mi{(Initial\ Abstract\ Store)}~\emptymsscctmon:=&\ (\emptymsscctmon, \emptyscctmon) \\
+    \mi{(Abstract\ Events)}~\msscctevent:=&\ \msscctevent\times\scctevent \mid \absev{\emptyevent}\\
+  \end{aligned}
+\end{gather*}
+\begin{center}
+  \judgbox{\monitorcheck[]{\msscctmonitor}{\msscctmonitor[']}{\msscctevent}}{,,Monitor $\msscctmonitor$ does one step to $\msscctmonitor[']$ given event $\msscctevent$.''}$\;$\\
+  %
+  \typerule{msscct-step}{
+    \monitorcheck[]{\msmonitor}{\msmonitor[']}{\msevent} &
+    \monitorcheck[]{\scctmonitor}{\scctmonitor[']}{\scctevent}
+  }{
+    \monitorcheck[]{(\msmonitor,\scctmonitor)}{(\msmonitor['],\scctmonitor['])}{(\msevent,\scctevent)}
+  }{ap:msscct-step}
+\end{center}
+\begin{center}
+  \judgbox{\monitorcheck{\msscctmonitor}{\msscctmonitor[']}{\mssccttrace}}{,,Monitor $\msscctmonitor$ multi-steps to $\msscctmonitor[']$ given trace $\mssccttrace$.''}$\;$\\
+  %
+  \typerule{msscct-refl}{
+    \monitorcheck{\msmonitor}{\msmonitor}{\hole{\cdot}} &
+    \monitorcheck{\scctmonitor}{\scctmonitor}{\hole{\cdot}}
+  }{
+    \monitorcheck{(\msmonitor,\scctmonitor)}{(\msmonitor,\scctmonitor)}{\hole{\cdot}}
+  }{ap:msscct-refl}
+  %
+  \typerule{msscct-ign-trans}{
+    \monitorcheck{\msscctmonitor}{\msscctmonitor[']}{\mssccttrace}
+  }{
+    \monitorcheck[]{\msscctmonitor}{\msscctmonitor[']}{\absev{\emptyevent}\cdot\mssccttrace}
+  }{ap:msscct-ign-trans}
+  %
+  \typerule{msscct-trans}{
+    \monitorcheck[]{\msscctmonitor}{\msscctmonitor[']}{\msscctevent} &
+    \monitorcheck{\msscctmonitor[']}{\msscctmonitor['']}{\mssccttrace}
+  }{
+    \monitorcheck[]{\msscctmonitor}{\msscctmonitor['']}{\msscctevent\cdot\mssccttrace}
+  }{ap:msscct-trans}
+\end{center}
+\begin{center}
+  \judgbox{\mssccttraceagree[]{\event}{\msscctevent}}{,,Abstract event $\msscctevent$ is equivalent to $\event$ with respect to \gls*{ms}.''}$\;$\\
+  %
+  \typerule{msscct-some-authentic}{
+    \mstraceagree[]{\event}{\msevent} &
+    \sccttraceagree[]{\event}{\scctevent}
+  }{
+    \mssccttraceagree[]{\event}{(\msevent,\scctevent)}
+  }{ap:msscct-some-authentic}
+  %
+  \typerule{msscct-none-authentic}{
+    \mstraceagree[]{\event}{\absev{\emptyevent}} &
+    \sccttraceagree[]{\event}{\absev{\emptyevent}}
+  }{
+    \mssccttraceagree[]{\event}{\absev{\emptyevent}}
+  }{ap:msscct-none-authentic}
+\end{center}
+
+\begin{lemma}[Traces with Monitor Satisfaction are $\msscctsafe$]\label{ap:lem:mon:msscctsafe}
+  If $\monmsscctsafe{\trace}$, then $\trace\in\msscctsafe$.\Coqed
+\end{lemma}
+\begin{incompleteproof}
+\end{incompleteproof}
+
 \bibliographystyle{ACM-Reference-Format}

### 2023-08-05 tech/main.tex
.

@@
 
+\section{Programming Languages}\label{ap:sec:proglangs}
+
+\subsection{$\Ltms$: Statically Typed, Temporal Memory Safe, Strict Cryptographic Constant Time}
+
+$\Ltms$ trivially satisfies \gls*{scct}, because all variables/values are private ($\lock$) by default.
+
+\begin{gather*}
+  \begin{aligned}
+    \mi{(Expressions)}~&\src{\expr}\bnfdef\ \src{x} \mid \src{\valueexpr} \mid \src{\lbinop{\expr[_{1}]}{\expr[_{2}]}} \mid \src{\lget{x}{\expr}} \mid \src{\llet{x}{\lnew{\expr[_{1}]}{\expr[_{2}]}}{\expr[_{3}]}} \mid \src{\ldelete{x}} \mid \src{\lset{x}{\expr[_{1}]}{\expr[_{2}]}} \\
+                       & \mid \src{\lpair{\expr[_{1}]}{\expr[_{2}]}} \mid \src{\lproja{\expr}} \mid \src{\lprojb{\expr}} \mid \src{\llet{x}{\expr[_{1}]}{\expr[_{2}]}} \mid \src{\lreturn{\expr}} \mid \src{\lcall{g}{\expr}} \mid \src{\lifz{\expr}{\expr[_{1}]}{\expr[_{2}]}} \mid \src{\labort}
+  \end{aligned}
+  \\
+  \begin{aligned}
+  \mi{(Types)}~&\src{\type}\bnfdef\src{\natt} \mid \src{\type[_{1}]} \times \src{\type[_{2}]}
+  &
+  \mi{(Functions)}~&\src{\function}\bnfdef\ \src{\lfunction{\textit{foo}}{x}{e}}
+  &
+\mi{(Libraries)}~&\src{\library}\bnfdef\ \src{\hole{\cdot}} \mid \src{\function},\src{\library}
+  \end{aligned}\\
+  \begin{aligned}
+  \mi{(Component~Names)}~&\src{\commlib}\bnfdef\ \src{\hole{\cdot}} \mid \src{\text{foo}},\src{\commlib}
+  &
+  \mi{(Programs)}~&\src{\prog{\library_{\ctx}}{\library_{\comp}}}
+  \end{aligned}
+\end{gather*}
+
+Variables are referred to as $x,y,z,a,b,c,\ldots$ while functions may be referred to as multi-character words, such as $\texttt{foo}$, as well as short-forms like $f,g,h$.
+
+\begin{gather*}
+  \begin{aligned}
+    \mi{(Control~Tags)}~&\src{\sandboxtag}\bnfdef\src{\ctx} \mid \src{comp}
+    &
+    \mi{(Communication~Tags)}~&\src{\comm}\bnfdef\ \src{\ctxtocomp} \mid\ \src{\comptoctx} \mid \src{\nocomm}
+    &
+    \mi{(Poison~Tags)}~&\src{\poison}\bnfdef \src{\poisoned} \mid \src{\poisonless}
+  \end{aligned}\\
+  \begin{aligned}
+    \mi{(Continuation~Stacks)}~&\src{\ectxstack}\bnfdef\src{\hole{\cdot}} \mid \src{\tup{K}{g}},\src{\ectxstack}
+    &
+    \mi{(Control~Flow~States)}~&\src{\cfstate}\bnfdef\src{\tup{\library}{\ectxstack}}
+  	\\
+    \mi{(Locations)}~&\src{\loc}\in\nat
+    &
+    \mi{(Substitutions)}~&\src{\substvar}\bnfdef \src{\hole{\cdot}} \mid \src{\subst{\valueexpr}{x}},\src{\gamma}
+  	\\
+    \mi{(Stores)}~&\src{\Delta}\bnfdef\src{\hole{\cdot}} \mid \src{x\mapsto\tup{\loc}{\sandboxtag,\poison,n}},\src{\Delta}
+    &
+    \mi{(Heaps)}~&\src{H}\bnfdef\src{\hole{\cdot}} | \src{\valueexpr},\src{H}
+  \end{aligned}\\
+  \begin{aligned}
+    \mi{(Memory~States)}~&\src{\memstate}\bnfdef\src{\tup{H^{\ctx}}{H^{\comp};\Delta}}
+    &
+    \mi{(States)}~&\src{\statevar}\bnfdef\src{\tup{\cfstate}{\sandboxtag,\memstate}}
+    &
+    \mi{(Runtime~Terms)}~&\src{\runtimetermvar}\bnfdef\src{\rtt{\statevar}{\expr}} \mid \src{\lightning}
+  \end{aligned}
+\end{gather*}
+Throughout the paper, the substitution notation is also used to update entries in states $\statevar$.
+\begin{gather*}
+  \begin{aligned}
+    \mi{(Pre\text{-}Events)}~\src{\event[_{b}]}&\bnfdef \src{\ev{Alloc}\ \loc\ \valueexpr} \mid \src{\ev{Dealloc}\ \loc} \mid \src{\ev{Get}\ \loc\ \valueexpr} \mid \src{\ev{Set}\ \loc\ \valueexpr\ \valueexpr[']} \mid \src{\ev{Call}\ \comm\ g\ \valueexpr} \mid \src{\ev{Ret}\ \comm\ \valueexpr} \mid \src{\ev{Start}} \mid \src{\ev{End}\ \valueexpr}\\
+  \end{aligned}\\
+  \begin{aligned}
+    \mi{(Events)}~\src{\event}&\bnfdef \src{\emptyevent} \mid \src{\lightning} \mid \src{\tup{\event[_{b}]}{\sandboxtag,\securitytag}} \\
+  \end{aligned}
+\end{gather*}
+\begin{center}
+  \judgbox{\pstepto{\src{\runtimetermvar}}{\src{\runtimetermvar[']}}{\src{\event}}}{,,$\runtimetermvar$ does one primitive step to $\runtimetermvar[']$ emitting event $\event$.''}
+  %
+  \typerule{$\src{\Ltms}$-e-binop}{
+  }{
+    \pstepto{}{}{\emptyevent}
+  }{ap:ltms-e-binop}
+  %
+  \typerule{$e-\text{get}-\in$}{
+    \sandboxtag=\statevar.\sandboxtag &
+    \statevar.\store(x)=\tup{\loc}{\sandboxtag,\poison,m} \\
+    \loc+n\in\operatorname{dom}\statevar.\heap[^{\sandboxtag}](\loc+n)
+  }{
+    \pstepto{\rtt{\statevar}{\lget{x}{n}}}{\rtt{\statevar}{\heap[^{\sandboxtag}](\loc+n)}}{\tup{\ev{Get}\ \loc\ n}{\sandboxtag}}
+  }{e-get-in}
+  %
+  \typerule{$e-\text{set}-\notin$}{
+    \sandboxtag=\statevar.\sandboxtag &
+    \statevar.\store(x)=\tup{\loc}{\sandboxtag,\poison,m} \\
+    \loc+n\notin\operatorname{dom}\statevar.\heap[^{\sandboxtag}]
+  }{
+    \pstepto{\rtt{\statevar}{\lset{x}{n}{\valueexpr}}}{\rtt{\statevar}{\valueexpr}}{\tup{\ev{Set}\ \loc\ n\ \valueexpr}{\sandboxtag}}
+  }{e-set-notin}
+  %
+  \typerule{$e-\text{new}$}{
+    \fresh{\statevar}{z} &
+    \fresh{\statevar}{\loc} &
+    \heap[^{\sandboxtag}_1]=\growh{\statevar.\heap[^{\sandboxtag}]}{n} &
+    \store[_1]=\storeel{z}{\loc}{\statevar.\sandboxtag}{\poisonless}{n},\statevar.\store
+  }{
+    \pstepto{\rtt{\statevar}{\lnew{x}{n}{\expr}}}{\rtt{\statevar\subst{\heap[^{\sandboxtag}]}{\heap[^{\sandboxtag}_1]}\subst{\store}{\store[_1]}}{\expr\subst{z}{x}}}{\tup{\ev{Alloc}\ \loc\ n}{\sandboxtag}}
+  }{e-new}
+  %
+  \typerule{$e-\text{dealloc}$}{
+    \statevar.\store(x) = \tup{\loc}{\statevar.\sandboxtag,\poison,n} &
+    \store[_1]=\seth{\statevar.\store}{x}{\tup{\loc}{\statevar.\sandboxtag,\poisoned,n}}
+  }{
+    \pstepto{\rtt{\statevar}{\ldelete{x}}}{\rtt{\statevar\subst{\store}{\store[_1]}}{0}}{\tup{\ev{Dealloc}\ \loc}{\sandboxtag}}
+  }{e-dealloc}
+  %
+  \typerule{$e-\text{abort}$}{
+  }{
+    \pstepto{\rtt{\statevar}{\labort}}{\lightning}{\lightning}
+  }{e-abort}
+\end{center}
+
 \bibliographystyle{ACM-Reference-Format}

### 2023-08-05 tech/main.tex
add some rules
add more estep rules

--- 2023-08-04 tech/main.tex
@@
 
-\section{General Trace Model}\label{ap:def:gtm}
+\section{General Trace Model}\label{ap:sec:gtm}
 
@@
 
+\section{Programming Languages}\label{ap:sec:proglangs}
+
+\subsection{$\Ltms$: Statically Typed, Temporal Memory Safe, Strict Cryptographic Constant Time}
+
+$\Ltms$ trivially satisfies \gls*{scct}, because all variables/values are private ($\lock$) by default.
+This is ,,enforced'' when projecting $\src{\Ltms}$ events to the general trace model presented in \Cref{ap:sec:gtm}.
+
+\begin{gather*}
+  \begin{aligned}
+    \mi{(Expressions)}~&\src{\expr}\bnfdef\ \src{x} \mid \src{\valueexpr} \mid \src{\lbinop{\expr[_{1}]}{\expr[_{2}]}} \mid \src{\lget{x}{\expr}} \mid \src{\llet{x}{\lnew{\expr[_{1}]}{\expr[_{2}]}}{\expr[_{3}]}} \mid \src{\ldelete{x}} \mid \src{\lset{x}{\expr[_{1}]}{\expr[_{2}]}} \\
+                       & \mid \src{\lpair{\expr[_{1}]}{\expr[_{2}]}} \mid \src{\lproja{\expr}} \mid \src{\lprojb{\expr}} \mid \src{\llet{x}{\expr[_{1}]}{\expr[_{2}]}} \mid \src{\lreturn{\expr}} \mid \src{\lcall{g}{\expr}} \mid \src{\lifz{\expr}{\expr[_{1}]}{\expr[_{2}]}} \mid \src{\labort}
+  \end{aligned}
+  \\
+  \begin{aligned}
+  \mi{(Types)}~&\src{\type}\bnfdef\src{\natt} \mid \src{\type[_{1}]} \times \src{\type[_{2}]}
+  &
+  \mi{(Functions)}~&\src{\function}\bnfdef\ \src{\lfunction{\textit{foo}}{x}{e}}
+  &
+\mi{(Libraries)}~&\src{\library}\bnfdef\ \src{\hole{\cdot}} \mid \src{\function},\src{\library}
+  \end{aligned}\\
+  \begin{aligned}
+  \mi{(Component~Names)}~&\src{\commlib}\bnfdef\ \src{\hole{\cdot}} \mid \src{\text{foo}},\src{\commlib}
+  &
+  \mi{(Programs)}~&\src{\prog{\library_{\ctx}}{\library_{\comp}}}
+  \end{aligned}
+\end{gather*}
+
+Variables are referred to as $x,y,z,a,b,c,\ldots$ while functions may be referred to as multi-character words, such as $\texttt{foo}$, as well as short-forms like $f,g,h$.
+
+\begin{gather*}
+  \begin{aligned}
+    \mi{(Control~Tags)}~&\src{\sandboxtag}\bnfdef\src{\ctx} \mid \src{comp}
+    &
+    \mi{(Communication~Tags)}~&\src{\comm}\bnfdef\ \src{\ctxtocomp} \mid\ \src{\comptoctx} \mid \src{\nocomm}
+    &
+    \mi{(Poison~Tags)}~&\src{\poison}\bnfdef \src{\poisoned} \mid \src{\poisonless}
+  \end{aligned}\\
+  \begin{aligned}
+    \mi{(Continuation~Stacks)}~&\src{\ectxstack}\bnfdef\src{\hole{\cdot}} \mid \src{\tup{K}{g}},\src{\ectxstack}
+    &
+    \mi{(Control~Flow~States)}~&\src{\cfstate}\bnfdef\src{\tup{\library}{\ectxstack}}
+  	\\
+    \mi{(Locations)}~&\src{\loc}\in\nat
+    &
+    \mi{(Substitutions)}~&\src{\substvar}\bnfdef \src{\hole{\cdot}} \mid \src{\subst{\valueexpr}{x}},\src{\gamma}
+  	\\
+    \mi{(Stores)}~&\src{\Delta}\bnfdef\src{\hole{\cdot}} \mid \src{x\mapsto\tup{\loc}{\sandboxtag,\poison,n}},\src{\Delta}
+    &
+    \mi{(Heaps)}~&\src{H}\bnfdef\src{\hole{\cdot}} | \src{\valueexpr},\src{H}
+  \end{aligned}\\
+  \begin{aligned}
+    \mi{(Memory~States)}~&\src{\memstate}\bnfdef\src{\tup{H^{\ctx}}{H^{\comp};\Delta}}
+    &
+    \mi{(States)}~&\src{\statevar}\bnfdef\src{\tup{\cfstate}{\sandboxtag,\memstate}}
+    &
+    \mi{(Runtime~Terms)}~&\src{\runtimetermvar}\bnfdef\src{\rtt{\statevar}{\expr}} \mid \src{\lightning}
+  \end{aligned}
+\end{gather*}
+Throughout the paper, the substitution notation is also used to update entries in states $\statevar$.
+\begin{gather*}
+  \begin{aligned}
+    \mi{(Pre\text{-}Events)}~\src{\event[_{b}]}&\bnfdef \src{\ev{Alloc}\ \loc\ \valueexpr} \mid \src{\ev{Dealloc}\ \loc} \mid \src{\ev{Get}\ \loc\ \valueexpr} \mid \src{\ev{Set}\ \loc\ \valueexpr\ \valueexpr[']} \mid \src{\ev{Call}\ \comm\ g\ \valueexpr} \mid \src{\ev{Ret}\ \comm\ \valueexpr} \mid \src{\ev{Start}} \mid \src{\ev{End}\ \valueexpr}\\
+  \end{aligned}\\
+  \begin{aligned}
+    \mi{(Events)}~\src{\event}&\bnfdef \src{\emptyevent} \mid \src{\lightning} \mid \src{\tup{\event[_{b}]}{\sandboxtag}} \\
+  \end{aligned}
+\end{gather*}
+\begin{center}
+  \judgbox{\pstepto{\src{\runtimetermvar}}{\src{\runtimetermvar[']}}{\src{\event}}}{,,$\src{\runtimetermvar}$ does one primitive step to $\src{\runtimetermvar[']}$ emitting event $\src{\event}$.''}
+  %
+  \typerule{$\src{\Ltms}-e-\text{binop}$}{
+    \src{\lbinop{\expr[_1]}{\expr[_2]}}=\src{\expr[_3]}
+  }{
+    \pstepto{\src{\rtt{\statevar}{\lbinop{\expr[_1]}{\expr[_2]}}}}{\src{\rtt{\statevar}{\expr[_3]}}}{\src{\emptyevent}}
+  }{ap:ltms-e-binop}
+  %
+  \typerule{$\src{\Ltms}-e-\text{get}-\in$}{
+    \src{\sandboxtag}=\src{\statevar.\sandboxtag} &
+    \src{\statevar.\store(x)}=\src{\tup{\loc}{\sandboxtag,\poison,m}} \\
+    \src{\loc}+\src{n}\in\operatorname{dom}\src{\statevar.\heap[^{\sandboxtag}](\loc+n)}
+  }{
+    \pstepto{\src{\rtt{\statevar}{\lget{x}{n}}}}{\src{\rtt{\statevar}{\statevar.\heap[^{\sandboxtag}](\loc+n)}}}{\src{\tup{\ev{Get}\ \loc\ n}{\sandboxtag}}}
+  }{ap:ltms-e-get-in}
+  %
+  \typerule{$\src{\Ltms}-e-\text{get}-\notin$}{
+    \src{\sandboxtag}=\src{\statevar.\sandboxtag} &
+    \src{\statevar.\store(x)}=\src{\tup{\loc}{\sandboxtag,\poison,m}} \\
+    \src{\loc}+\src{n}\notin\operatorname{dom}\src{\statevar.\heap[^{\sandboxtag}](\loc+n)}
+  }{
+    \pstepto{\src{\rtt{\statevar}{\lget{x}{n}}}}{\src{\rtt{\statevar}{1729}}}{\src{\tup{\ev{Get}\ \loc\ n}{\sandboxtag}}}
+  }{ap:ltms-e-get-notin}
+  %
+  \typerule{$\src{\Ltms}-e-\text{set}-\in$}{
+    \src{\sandboxtag}=\src{\statevar.\sandboxtag} &
+    \src{\statevar.\store(x)}=\src{\tup{\loc}{\sandboxtag,\poison,m}} \\
+    \src{\loc}+\src{n}\in\operatorname{dom}\src{\statevar.\heap[^{\sandboxtag}]} &
+    \src{\heap[']}=\seth{\src{\statevar.\heap[^{\sandboxtag}]}}{\src{\loc}+\src{n}}{\src{\valueexpr}}
+  }{
+    \pstepto{\src{\rtt{\statevar}{\lset{x}{n}{\valueexpr}}}}{\src{\rtt{\statevar\subst{\heap[^{\sandboxtag}]}{\heap[']}}{\valueexpr}}}{\src{\tup{\ev{Set}\ \loc\ n\ \valueexpr}{\sandboxtag}}}
+  }{ap:ltms-e-set-in}
+  %
+  \typerule{$\src{\Ltms}-e-\text{set}-\notin$}{
+    \src{\sandboxtag}=\src{\statevar.\sandboxtag} &
+    \src{\statevar.\store(x)}=\src{\tup{\loc}{\sandboxtag,\poison,m}} \\
+    \src{\loc}+\src{n}\notin\operatorname{dom}\src{\statevar.\heap[^{\sandboxtag}]}
+  }{
+    \pstepto{\src{\rtt{\statevar}{\lset{x}{n}{\valueexpr}}}}{\src{\rtt{\statevar}{\valueexpr}}}{\src{\tup{\ev{Set}\ \loc\ n\ \valueexpr}{\sandboxtag}}}
+  }{ap:ltms-e-set-notin}
+  %
+  \typerule{$\src{\Ltms}-e-\text{new}$}{
+    \fresh{\src{\statevar}}{\src{z}} &
+    \fresh{\src{\statevar}}{\src{\loc}} &
+    \src{\heap[^{\sandboxtag}_1]}=\growh{\src{\statevar.\heap[^{\sandboxtag}]}}{\src{n}}{\src{\valueexpr}} &
+    \src{\store[_1]}=\src{\storeel{z}{\loc}{\statevar.\sandboxtag}{\poisonless}{n},\statevar.\store}
+  }{
+    \pstepto{\src{\rtt{\statevar}{\lnew{x}{\valueexpr}{n}{\expr}}}}{\src{\rtt{\statevar\subst{\heap[^{\sandboxtag}]}{\heap[^{\sandboxtag}_1]}\subst{\store}{\store[_1]}}{\expr\subst{z}{x}}}}{\src{\tup{\ev{Alloc}\ \loc\ n}{\statevar.\sandboxtag}}}
+  }{ap:ltms-e-new}
+  %
+  \typerule{$\src{\Ltms}-e-\text{dealloc}$}{
+    \src{\statevar.\store(x)} = \src{\tup{\loc}{\statevar.\sandboxtag,\poison,n}} &
+    \src{\store[_1]}=\seth{\src{\statevar.\store}}{\src{x}}{\src{\tup{\loc}{\statevar.\sandboxtag,\poisoned,n}}}
+  }{
+    \pstepto{\src{\rtt{\statevar}{\ldelete{x}}}}{\src{\rtt{\statevar\subst{\store}{\store[_1]}}{0}}}{\src{\tup{\ev{Dealloc}\ \loc}{\statevar.\sandboxtag}}}
+  }{ap:ltms-e-dealloc}
+  %
+  \typerule{$\src{\Ltms}-e-\text{let}$}{
+  }{
+    \pstepto{\src{\rtt{\statevar}{\llet{x}{\valueexpr}{\expr}}}}{\src{\rtt{\statevar}{\expr\subst{x}{\valueexpr}}}}{\src{\emptyevent}}
+  }{ap:ltms-e-let}
+  %
+  \typerule{$\src{\Ltms}-e-\text{ifz}-yes$}{
+  }{
+    \pstepto{\src{\rtt{\statevar}{\lifz{0}{\expr[_1]}{\expr[_2]}}}}{\src{\rtt{\statevar}{\expr[_1]}}}{\src{\emptyevent}}
+  }{ap:ltms-e-ifz-yes}
+  %
+  \typerule{$\src{\Ltms}-e-\text{ifz}-no$}{
+    \src{n}\not=\src{0}
+  }{
+    \pstepto{\src{\rtt{\statevar}{\lifz{n}{\expr[_1]}{\expr[_2]}}}}{\src{\rtt{\statevar}{\expr[_2]}}}{\src{\emptyevent}}
+  }{ap:ltms-e-ifz-no}
+  %
+  \typerule{$\src{\Ltms}-e-\text{proj}-a$}{
+  }{
+    \pstepto{\src{\rtt{\statevar}{\lproja{\lpair{\valueexpr[_1]}{\valueexpr[_2]}}}}}{\src{\rtt{\statevar}{\valueexpr[_1]}}}{\src{\emptyevent}}
+  }{ap:ltms-e-proj-a}
+  %
+  \typerule{$\src{\Ltms}-e-\text{proj}-b$}{
+  }{
+    \pstepto{\src{\rtt{\statevar}{\lprojb{\lpair{\valueexpr[_1]}{\valueexpr[_2]}}}}}{\src{\rtt{\statevar}{\valueexpr[_2]}}}{\src{\emptyevent}}
+  }{ap:ltms-e-proj-a}
+  %
+  \typerule{$\src{\Ltms}-e-\text{abort}$}{
+  }{
+    \pstepto{\src{\rtt{\statevar}{\labort}}}{\src{\lightning}}{\src{\lightning}}
+  }{ap:ltms-e-abort}
+\end{center}
+\begin{center}
+  \judgbox{\src{\commlib}\vdash\src{foo}:\src{\sandboxtag}\dashv\src{\comm},\src{\sandboxtag'}}{,,Determine communication tag $\src{\comm}$ and control tag $\src{\sandboxtag'}$ by checking the current control $\src{\sandboxtag}$,\\the called function $\src{foo}$, and list of component names $\src{\commlib}$.''}
+  %
+  \typerule{$\src{\Ltms}-\text{switch}-\ctx-\text{to}-\comp$}{
+    \src{foo}\in\src{\commlib}
+  }{
+    \src{\commlib}\vdash\src{foo}:\src{\ctx}\dashv\src{\ctxtocomp},\src{\comp}
+  }{ap:ltms-switch-ctx-to-comp}
+  %
+  \typerule{$\src{\Ltms}-\text{switch}-\comp-\text{to}-\ctx$}{
+    \src{foo}\notin\src{\commlib}
+  }{
+    \src{\commlib}\vdash\src{foo}:\src{\comp}\dashv\src{\comptoctx},\src{\ctx}
+  }{ap:ltms-switch-comp-to-ctx}
+  %
+  \typerule{$\src{\Ltms}-\text{noswitch}-\ctx$}{
+    \src{foo}\notin\src{\commlib}
+  }{
+    \src{\commlib}\vdash\src{foo}:\src{\ctx}\dashv\src{\nocomm},\src{\ctx}
+  }{ap:ltms-noswitch-ctx}
+  %
+  \typerule{$\src{\Ltms}-\text{noswitch}-\comp$}{
+    \src{foo}\in\src{\commlib}
+  }{
+    \src{\commlib}\vdash\src{foo}:\src{\comp}\dashv\src{\nocomm},\src{\comp}
+  }{ap:ltms-noswitch-comp}
+\end{center}
+\begin{center}
+  \judgbox{\estepto[]{\src{\runtimetermvar}}{\src{\runtimetermvar[']}}{\src{\event}}}{,,$\src{\runtimetermvar}$ does one contextual step to $\src{\runtimetermvar[']}$ emitting event $\src{\event}$.''}
+  %
+  \typerule{$\src{\Ltms}-e-\text{ectx}$}{
+    \pstepto{\src{\rtt{\statevar}{\expr}}}{\src{\rtt{\statevar[']}{\expr[']}}}{\src{\event}}
+  }{
+    \estepto[]{\src{\rtt{\statevar}{\ectx\hole{\expr}}}}{\src{\rtt{\statevar[']}{\ectx\hole{\expr[']}}}}{\src{\event}}
+  }{ap:ltms-e-ectx}
+  %
+  \typerule{$\src{\Ltms}-e-\text{stuck}$}{
+    \pstepto{\src{\rtt{\statevar}{\expr}}}{\src{\lightning}}{\src{\lightning}}
+  }{
+    \estepto[]{\src{\rtt{\statevar}{\ectx\hole{\expr}}}}{\src{\lightning}}{\src{\lightning}}
+  }{ap:ltms-e-stuck}
+  %
+  \typerule{$\src{\Ltms}-e-\text{start}$}{
+    \src{\lfunction{\text{main}}{x}{e}} \in \src{\statevar.\library} &
+    \src{\statevar.\ectxstack}=\src{\hole{\cdot}}
+  }{
+    \estepto[]{\src{\rtt{\statevar}{\lcall{\text{main}}{0}}}}{\src{\rtt{\statevar\subst{\ctx}{\sandboxtag}\subst{\tup{\hole{\cdot}}{\text{main}},\hole{\cdot}}{\ectxstack}}{\expr\subst{0}{x}}}}{\src{\tup{\ev{Start}}{\comp}}}
+  }{ap:ltms-e-start}
+  %
+  \typerule{$\src{\Ltms}-e-\text{ret}$}{
+    \src{\statevar.\ectxstack}=\src{\tup{\ectx}{\text{foo}}},\src{\ectxstack[']} &
+    \src{\commlib}\vdash\src{\text{foo}}:\src{\statevar.\sandboxtag}\dashv\src{\comm}
+  }{
+    \estepto[]{\src{\rtt{\statevar}{\ectx[']\hole{\lreturn{\valueexpr}}}}}{\src{\rtt{\statevar\subst{\neg\statevar.\sandboxtag}{\sandboxtag}\subst{\ectxstack[']}{\ectxstack}}{\ectx\hole{\valueexpr}}}}{\src{\tup{\ev{Ret}\ \comm\ \valueexpr}{\statevar.\sandboxtag}}}
+  }{ap:ltms-e-ret}
+  %
+  \typerule{$\src{\Ltms}-e-\text{call}$}{
+    \src{\lfunction{\text{foo}}{x}{e}} \in \src{\statevar.\library} &
+    \src{\commlib}\vdash\src{\text{foo}}:\src{\statevar.\sandboxtag}\dashv\src{\comm}
+  }{
+    \estepto[]{\src{\rtt{\statevar}{\ectx\hole{\lcall{\text{foo}}{\valueexpr}}}}}{\src{\rtt{\statevar\subst{\neg\statevar.\sandboxtag}{\sandboxtag}\subst{\tup{\ectx}{\text{foo}},\statevar.\ectxstack}{\ectxstack}}{\expr\subst{\valueexpr}{x}}}}{\src{\tup{\ev{Call}\ \comm\ \text{foo}\ \valueexpr}{\statevar.\sandboxtag}}}
+  }{ap:ltms-e-call}
+  %
+  \typerule{$\src{\Ltms}-e-\text{end}$}{
+    \src{\statevar.\ectxstack}=\src{(\hole{\cdot},main),\hole{\cdot}}
+  }{
+    \estepto[]{\src{\rtt{\statevar}{\ectx\hole{\lreturn{\valueexpr}}}}}{\src{\rtt{\statevar\subst{\comp}{\sandboxtag}\subst{\tup{\hole{\cdot}}{\text{main}},\hole{\cdot}}{\ectxstack}}{\valueexpr}}}{\src{\tup{\ev{End}\ \valueexpr}{\ctx}}}
+  }{ap:ltms-e-end}
+\end{center}
+
 \bibliographystyle{ACM-Reference-Format}

### 2023-08-11 tech/main.tex
Add splitting

@@
   \begin{aligned}
-    \mi{(Expressions)}~&\src{\expr}\bnfdef\ \src{x} \mid \src{\valueexpr} \mid \src{\lbinop{\expr[_{1}]}{\expr[_{2}]}} \mid \src{\lget{x}{\expr}} \mid \src{\llet{x}{\lnew{\expr[_{1}]}{\expr[_{2}]}}{\expr[_{3}]}} \mid \src{\ldelete{x}} \mid \src{\lset{x}{\expr[_{1}]}{\expr[_{2}]}} \\
+    \mi{(Expressions)}~&\src{\expr}\bnfdef\ \src{x} \mid \src{\valueexpr} \mid \src{\lbinop{\expr[_{1}]}{\expr[_{2}]}} \mid \src{\lget{x}{\expr}} \mid \src{\lnew{x}{\expr[_{1}]}{\expr[_{2}]}{\expr[_{3}]}} \mid \src{\ldelete{x}} \mid \src{\lset{x}{\expr[_{1}]}{\expr[_{2}]}} \\
                        & \mid \src{\lpair{\expr[_{1}]}{\expr[_{2}]}} \mid \src{\lproja{\expr}} \mid \src{\lprojb{\expr}} \mid \src{\llet{x}{\expr[_{1}]}{\expr[_{2}]}} \mid \src{\lreturn{\expr}} \mid \src{\lcall{g}{\expr}} \mid \src{\lifz{\expr}{\expr[_{1}]}{\expr[_{2}]}} \mid \src{\labort}
@@
   \begin{aligned}
-  \mi{(Types)}~&\src{\type}\bnfdef\src{\natt} \mid \src{\type[_{1}]} \times \src{\type[_{2}]}
+  \mi{(Types)}~&\src{\type}\bnfdef\src{\natt} \mid \src{\ptr} \mid \src{\wptr} \mid \src{\type[_{1}] \times \type[_{2}]}
   &
+  \mi{(Type~Enviroments)}~&\src{\typenv}\bnfdef\src{\hole{\cdot}} \mid \src{x:\type,\typenv} \mid \src{x:\type[_1]\to\type[_2],\typenv}
+  \end{aligned}\\
+  \begin{aligned}
   \mi{(Functions)}~&\src{\function}\bnfdef\ \src{\lfunction{\textit{foo}}{x}{e}}
@@
 Throughout the paper, the substitution notation is also used to update entries in states $\statevar$.
+
+\begin{center}
+  \judgbox{\noownedptr{\src{\type}}}{,,Type $\src{\type}$ does not entail an owned pointer.''}
+  %
+  \typerule{$\src{\Ltms}-\text{noownedptr}-\src{\natt}$}{
+  }{
+    \noownedptr{\src{\natt}}
+  }{ap:ltms-noownedptr-natt}
+  %
+  \typerule{$\src{\Ltms}-\text{noownedptr}-\src{\wptr}$}{
+  }{
+    \noownedptr{\src{\wptr}}
+  }{ap:ltms-noownedptr-wptr}
+  %
+  \typerule{$\src{\Ltms}-\text{noownedptr}-\text{pair}$}{
+    \noownedptr{\src{\type[_1]}} &
+    \noownedptr{\src{\type[_2]}}
+  }{
+    \noownedptr{\src{\type[_1]\times\type[_2]}}
+  }{ap:ltms-noownedptr-pair}
+\end{center}
+\begin{center}
+  \judgbox{\noownedptr{\src{\typenv}}}{,,$\src{\typenv}$ does not contain any owned pointers.''}
+  %
+  \typerule{$\src{\Ltms}-\text{noownedptr}-\text{empty}$}{
+  }{
+    \noownedptr{\src{\hole{\cdot}}}
+  }{ap:ltms-noownedptr-empty}
+  %
+  \typerule{$\src{\Ltms}-\text{noownedptr}-\text{cons}$}{
+    \noownedptr{\src{\type}} &
+    \noownedptr{\src{\typenv}}
+  }{
+    \noownedptr{\src{x:\type,\typenv}}
+  }{ap:ltms-noownedptr-cons}
+\end{center}
+\begin{center}
+  \judgbox{\src{\typenv}\equiv\typenvsplit{\src{\typenv[']}}{\src{\typenv['']}}}{,,Split environment $\src{\typenv}$ into $\src{\typenv[']}$ and $\src{\typenv['']}$.''}
+  %
+  \typerule{$\src{\Ltms}-$split-empty}{
+  }{
+    \src{\hole{\cdot}} \equiv \typenvsplit{\src{\hole{\cdot}}}{\src{\hole{\cdot}}}
+  }{ap:ltms-split-empty}
+  %
+  \typerule{$\src{\Ltms}-$split-emptyL}{
+  }{
+    \src{\typenv} \equiv \typenvsplit{\src{\hole{\cdot}}}{\src{\typenv}}
+  }{ap:ltms-split-emptyL}
+  %
+  \typerule{$\src{\Ltms}-$split-emptyR}{
+  }{
+    \src{\typenv} \equiv \typenvsplit{\src{\typenv}}{\src{\hole{\cdot}}}
+  }{ap:ltms-split-emptyR}
+  %
+  \typerule{$\src{\Ltms}$-split-$\natt$}{
+    \src{\typenv} \equiv \typenvsplit{\src{\typenv[_1]}}{\src{\typenv[_2]}}
+  }{
+    \src{x:\natt},\src{\typenv} \equiv \typenvsplit{\src{x:\natt},\src{\typenv[_1]}}{\src{x:\natt},\src{\typenv[_2]}}
+  }{ap:ltms-split-natt}
+  %
+  \typerule{$\src{\Ltms}$-split-wptr}{
+    \src{\typenv} \equiv \typenvsplit{\src{\typenv[_1]}}{\src{\typenv[_2]}}
+  }{
+    \src{x:\wptr},\src{\typenv} \equiv \typenvsplit{\src{x:\wptr},\src{\typenv[_1]}}{\src{x:\wptr},\src{\typenv[_2]}}
+  }{ap:ltms-split-wptr}
+  %
+  \typerule{$\src{\Ltms}$-split-pair}{
+    \noownedptr{\src{\type[_1]\times\type[_2]}} &
+    \src{\typenv} \equiv \typenvsplit{\src{\typenv[_1]}}{\src{\typenv[_2]}}
+  }{
+    \src{x:\type[_1]\times\type[_2]},\src{\typenv} \equiv \typenvsplit{\src{x:\type[_1]},\src{\typenv[_1]}}{\src{x:\type[_2]},\src{\typenv[_2]}}
+  }{ap:ltms-split-pair}
+  %
+  \typerule{$\src{\Ltms}$-split-ptrL}{
+    \src{\typenv} \equiv \typenvsplit{\src{\typenv[_1]}}{\src{\typenv[_2]}}
+  }{
+    \src{x:\ptr},\src{\typenv} \equiv \typenvsplit{\src{x:\ptr},\src{\typenv[_1]}}{\src{\typenv[_2]}}
+  }{ap:ltms-split-ptrL}
+  %
+  \typerule{$\src{\Ltms}$-split-ptrR}{
+    \src{\typenv} \equiv \typenvsplit{\src{\typenv[_1]}}{\src{\typenv[_2]}}
+  }{
+    \src{x:\ptr},\src{\typenv} \equiv \typenvsplit{\src{\typenv[_1]}}{\src{x:\ptr},\src{\typenv[_2]}}
+  }{ap:ltms-split-ptrR}
+  %
+  \typerule{$\src{\Ltms}$-split-arrow}{
+    \noownedptr{\src{\type[_1]}} &
+    \noownedptr{\src{\type[_2]}} &
+    \src{\typenv} \equiv \typenvsplit{\src{\typenv[_1]}}{\src{\typenv[_2]}}
+  }{
+  \src{foo:\type[_1]\to\type[_2]},\src{\typenv} \equiv \typenvsplit{\src{foo:\type[_1]\to\type[_2]},\src{\Gamma_1}}{\src{foo:\type[_1]\to\type[_2]},\src{\Gamma_2}}
+  }{ap:ltms-split-arrow}
+\end{center}
+
 \begin{gather*}
@@
     \pstepto{\src{\rtt{\statevar}{\lprojb{\lpair{\valueexpr[_1]}{\valueexpr[_2]}}}}}{\src{\rtt{\statevar}{\valueexpr[_2]}}}{\src{\emptyevent}}
-  }{ap:ltms-e-proj-a}
+  }{ap:ltms-e-proj-b}
   %
@@
 \end{center}
+\begin{center}
+  \judgbox{\estepton{\src{\runtimetermvar}}{\src{\runtimetermvar[']}}{\src{\trace}}}{,,$\src{\runtimetermvar}$ performs $n$ steps to $\src{\runtimetermvar[']}$ emitting trace $\src{\trace}$.''}
+  %
+  \typerule{$\src{\Ltms}-\text{estepn}-\text{refl}$}{
+  }{
+    \estepton[0]{\src{\runtimetermvar}}{\src{\runtimetermvar}}{\src{\hole{\cdot}}}
+  }{ap:ltms-estepn-refl}
+  %
+  \typerule{$\src{\Ltms}-\text{estepn}-\text{trans}-\text{important}$}{
+    \src{\event}\not=\src{\emptyevent} &
+    \estepto[]{\src{\runtimetermvar}}{\src{\runtimetermvar[']}}{\src{\event}} &
+    \estepton[n]{\src{\runtimetermvar[']}}{\src{\runtimetermvar['']}}{\src{\trace}}
+  }{
+    \estepton[n+1]{\src{\runtimetermvar}}{\src{\runtimetermvar['']}}{\src{\event\cdot\trace}}
+  }{ap:ltms-estepn-trans-important}
+  %
+  \typerule{$\src{\Ltms}-\text{estepn}-\text{trans}-\text{unimportant}$}{
+    \estepto[]{\src{\runtimetermvar}}{\src{\runtimetermvar[']}}{\src{\emptyevent}} &
+    \estepton[n]{\src{\runtimetermvar[']}}{\src{\runtimetermvar['']}}{\src{\trace}}
+  }{
+    \estepton[n+1]{\src{\runtimetermvar}}{\src{\runtimetermvar['']}}{\src{\trace}}
+  }{ap:ltms-estepn-trans-unimportant}
+\end{center}
+\begin{center}
+  \judgbox{\isfinal{\src{\runtimetermvar}}}{,,Runtime term $\src{\runtimetermvar}$ represents a final state.''}
+  %
+  \typerule{$\src{\Ltms}-\src{\lightning}-\text{is}-\text{final}$}{
+  }{
+    \isfinal{\src{\lightning}}
+  }{ap:ltms-stuck-is-final}
+  %
+  \typerule{$\src{\Ltms}-\src{\valueexpr}-\text{is}-\text{final}$}{
+  }{
+    \isfinal{\src{\rtt{\statevar}{\valueexpr}}}
+  }{ap:ltms-value-is-final}
+\end{center}
+\begin{center}
+  \judgbox{\estepto{\src{\runtimetermvar}}{\src{\runtimetermvar[']}}{\src{\trace}}}{,,$\src{\runtimetermvar}$ steps to final $\src{\runtimetermvar[']}$ emitting trace $\src{\trace}$.''}
+  %
+  \typerule{$\src{\Ltms}-\text{esteps}-\text{refl}$}{
+    \isfinal{\src{\runtimetermvar}}
+  }{
+    \estepto{\src{\runtimetermvar}}{\src{\runtimetermvar}}{\src{\hole{\cdot}}}
+  }{ap:ltms-esteps-refl}
+  %
+  \typerule{$\src{\Ltms}-\text{esteps}-\text{trans}-\text{important}$}{
+    \src{\event}\not=\src{\emptyevent} &
+    \estepto[]{\src{\runtimetermvar}}{\src{\runtimetermvar[']}}{\src{\event}} &
+    \estepto{\src{\runtimetermvar[']}}{\src{\runtimetermvar['']}}{\src{\trace}}
+  }{
+    \estepto{\src{\runtimetermvar}}{\src{\runtimetermvar['']}}{\src{\event\cdot\trace}}
+  }{ap:ltms-esteps-trans-important}
+  %
+  \typerule{$\src{\Ltms}-\text{esteps}-\text{trans}-\text{unimportant}$}{
+    \estepto[]{\src{\runtimetermvar}}{\src{\runtimetermvar[']}}{\src{\emptyevent}} &
+    \estepto{\src{\runtimetermvar[']}}{\src{\runtimetermvar['']}}{\src{\trace}}
+  }{
+    \estepto{\src{\runtimetermvar}}{\src{\runtimetermvar['']}}{\src{\trace}}
+  }{ap:ltms-esteps-trans-unimportant}
+\end{center}
 

### 2023-09-14 popl/revs/rebuttal.md
:chicken: Add comments

# Author Response for Submission 67: Secure Composition of Robust and Optimising Compilers

First of all, we would like to sincerely thank every reviewer for putting in their hard work.
We are pleased about the high-quality of the reviews and appreciate both your suggestions and questions, which we address for each reviewer below.
Prior to that, we want to explain some general concerns applying to all reviews.

- The main results of our paper are about the compositionality of secure compilers, which we have Coq'd up at the time of submitting the paper. Missing Coq proofs mainly involve the case-study, but the paper does not claim to have Coq'd the language definitions, their properties, etc.
- The compositionality scales to arbitrary properties, not just trace properties. Classes of properties are kept as general as possible and there is no restriction being made on the shape of a class without talking about a concrete one. This should also entail hyperproperties, since they are classes themselves and enjoy closedness on intersections, i.e., intersecting hyperproperties yields other hyperproperties.
- The optimisations chosen, namely Dead-Code Elimination (dce) and Constant Folding (cf), are traditional compiler optimisations. While they are simpler to define in comparison to other optimisations, we have no reason to believe that the compositionality framework breaks apart when scaling this to other passes.

## Reviewer A

We think it is important to not get lost in theory and appreciate your critical thinking about whether our theoretical framework is applicable to practical compilers.
This is a big concern for us as well.
Unfortunately, developing a compiler with formal methods is a task that has a huge workload (see CompCert).
While there is no existing "practical", formally secure compiler, we believe that our framework allows to incrementally build such a thing, because of the separation of concerns that one gets from the compositionality framework.

### General comments

> However, the approach does not study the optimizations of any real compiler and hence it is not clear if the approach would be applicable in state-of-the-art compiler development.

The paper studies dead-code elimination (dce) as well as constant folding (cf) in the setting of (formally) secure compilers.
So far, the only practical compiler that involves formal methods is CompCert, but it is not proven secure.
Nevertheless, CompCert, as well as not-formally proven compilers, such as gcc or clang, perform dce and cf.
dce and cf are foundational compiler optimisations that any optimising compiler makes use of.
Therefore, we believe our approach is as applicable as formal methods when it comes to state-of-the-art compiler development.

> Undefined behaviors play a key role in developing optimizations in the state-of-the-art compilers. I wonder if these behaviors would affect the compositionality properties. 
> It is possible that the individual components are compiled following different sequences of optimizations and linked later.

We believe this is a very interesting point that is, however, a little outside the scope of the paper.
For the framework presented in the paper, undefined behavior would be encoded in the property/-ies of interest and, therefore, be preserved in the intersection of property classes.
We are planning to do a follow-up paper where we investigate more aggressive, realistic optimising compiler passes that (1) may exploit undefined behavior, (2) reorder code - and thereby reorder events on the trace, (3) memory layout reording, and so on.

> You have chosen constant time cryptography and memory safety. It would be useful to discuss the motivation behind selecting these two security properties with their similarities and differences.
We believe that the paper addresses this in the introduction: They are key security properties for ensuring confidentiality.

### Questions

> (1) Is it possible to model the optimizations in GCC or LLVM in your approach?

Yes, the paper does this for dce and cf, which both GCC and LLVM do.
We would appreciate further clarification about particular optimising transformations that you've had on your mind.
While it is unlikely that we can extend the formalisation in the given time-frame to account for another optimisation, we do think this space warrants future work.

> (2) Is your approach applicable to or can be extended for in hyperproperty based security properties?

Yes, classes are simply sets of properties.
We verified that the proof of the sequential composition theorem also works for classes of sets of hyperproperties.
We agree that the paper should be rephrased to improve the description of classes and why the formulation is general enough, i.e., why the sequential composition theorem supports any set whatsoever (even liveness).

> (3) Do you handle dynamic languages with managed runtime? Particularly, managed runtime systems use garbage collection to deallocate memory. Does it impact your compositionality scheme?

While we haven't done an analysis of something like this explicitly, our intuition is that our compositionality framework is general enough to capture this.

### What will change

- Add to future work section that we will look at more sophisticated optimisiation passes.

## Reviewer B

While we do not agree with the reasoning for the decision to strongly reject our submission, we strongly appreciate parts of your criticism and will improve the paper as outlined below.

### General comments

> While this work sets out to study an important problem: the compositional properties of secure compilation; the technical development is straightforward and the case studies are simplistic.
> The contribution is incremental given existing work in secure compilation.

From our literature review, we were not able to identify work that discusses the composition of (formally) secure compilers.
The compositionality of secure compilers has not been discussed before, and, as you said, it is an important problem to look at.

> The properties examined is a subset of the properties that secure compilation aims to preserve.

Note that, while the universal quantification in the sequential composition theorem restricts properties to be part of a certain class (=set of properties), the class is arbitrarily fixed.
For example, in the case-study, it is fixed to the class of memory safety properties which is later composed with the class of strictly cryptographic constant timeness.

> The composition is sequential composition only.

The paper also discusses "upper" as well as "lower" composition, which fascilitates multi-language reasoning on secure compilation.
This is something we don't get from just sequential composition and allows to build more practical secure compilers incrementally.

> In terms of sequential composition, it would be interesting to show the result of composing a compilation pass that violates a certain property pi and a later pass hardens the target to ensure that property.

We strongly agree that this is an interesting point and intend to improve our submission with regards to this.
While, under the given time-frame, we cannot change the formalisation of the case-study, we will add a paragraph discussing this by two example compiler passes used in practice.
Briefly, if the second compiler generates programs that always violate the property that the first ensures, then the preserved class of properties is actually empty, since the intersection is disjoint!
So, the theorem as stated still works, i.e., it is *not* unsound, but it may trivialize, which should better be emphasized in the paper.

> Similar to the above comment, another important form of composition not considered by this work is composing/linking components that are compiled  through different compilation pass. Especially when these components have shared state.

The robust preservation criterion universally quantifies the contexts and considers a single program component, so the framework encompasses contexts that result from different compilation passes, different compilers, or from writing them by hand.
We are unsure how the presented compositionality framework seems to fail to handle this and we would need further clarification.

> Sections 5 and 6 set up toy languages and compiler passes, which made the proofs straightforward. What are the challenges of establishing the theorems? Are there interesting proof techniques used?

Since the case-study serves as motivating example for the compositionality framework, we think it is important to discuss the framework in more detail instead of the formalisations of more-or-less standard programming languages.

> Definition 3.1, "at most one Alloc/Dealloc l;t;sigma" is imprecise. This could allow two Alloc/Deallocs of the same location, with different tag. If read strictly, it does not allow memory reuse, which is commonly done in low-level languages.

We believe the concrete formalisation of memory addresses is orthogonal to the goals of the paper/the case-study and merely an implementation detail.
The main interest of study is the compositionality of secure compilers, not how to make a memory-safe programming language.

> The Rule tms-trans seems to allow the same location to be allocated twice. Is this correct?

`tms-refl`, `tms-ign-trans`, as well as `tms-trans` define the reflexive-transitive closure of the TMS-monitor step, while filtering empty events.
They do not talk about locations at all.
Maybe you meant `tms-alloc`, which has premises that ensure an address is not allocated already, i.e., it cannot happen that there is an allocation for the same location.
Due to the sandboxing (see line 645 in the paper), the work models memory addresses as number + tag indicating what heap the address lies on.
The technical development can be extended to allow re-allocation of the same memory location, but this complicates the proofs unneccessarily, since the main goal is not to build a memory-safe lanugage, but to investigate compositionality of secure compilers.

> The monitor composition simply runs both monitors, which is not optimal.

While optimality may be interesting, it is also orthogonal to the goal of the paper, which cares about doing the correct thing.
Subsequent work can optimize this, if necessary.

> Why is the monitor composition interesting? It is also unclear how the monitor connects tightly to later sections of the paper.

The composition of temporal and spatial memory safety monitors requires the monitor composition to get a monitor for memory safety (intersection of tms and sms).
We agree that this could be made more explicit in the paper.

> The rules on page 14, the return values seem arbitrary for rules of operations with ill-defined behavior. For example, when setting a memory location that doesn't exist, why return the value of the expression to be written to memory? Similarly, why deletion return 0? Why deletion does not check for existing of the location, like the set rules? It may not handle double free properly.

The purpose of the case-study is to allow memory-unsafe behavior, i.e., execution should not get stuck if there is some kind of, e.g., memory-safety violation.
While we fail to see how the `e-dealloc` rule does not check for the existence of memory locations like the `e-set-*`/`e-get-*` rules, we do want to emphasize that `e-dealloc` does not need to change or check the heap(s) in any way, since it/they is/are modeled as one big arena.
One could instead add the data the pointer points to into Δ instead of describing a two separate heaps. However, our formalisation models sandboxing more faithfully.

> The coloring of the paper is distracting.

We want to note that past iterations of POPL already have several papers that make extensive use of color.
Nevertheless, removing color is a minor change that should not influence the acceptance of the paper.

> Figure 1, allowing $\varepsilon$ events to be arbitrarily inserted does not account for timing behavior. Is this intended?

Yes, this is intended, since this precise timing is abstracted away by the data independent timing mode + techniques used in the FaCT paper (Cauligi et al., 2019. https://doi.org/10.1145/3314221.3314605).

> Definition 3.6, the monitor semantics is not introduced at this point.

The paper has several monitor semantics and Definition 3.6 generically introduces what monitor satisfaction for each of the defined monitors means.
To not re-introduce this definition several times (for tms, sms, ms, scct, and scct+ms monitors), we introduce it here and give an intuitive characterization of the definition in the underlined text, e.g., the star-squiggly-arrow corresponds to "[...] monitor can step to [...]".

### Questions

> can the theorem be applied to the case where one pass violating a property and another pass re-establishes it?

Yes, see earlier discussion. In short, one has to be careful that the intersection does not become empty, but the theorem is nonetheless applicable and sound.
It's been proven in Coq.

> What's challenging?

The compositionality framework is simple enough to appear as straightforward, which is a good thing.
The catch is that building (formally verified) secure compilers is notoriously difficult, as backed up by an extensive number of secure compilation literature.
With the presented compositionality framework, the burden of proving the composed compiler of secure compilation passes as secure is reduced dramatically.

### What will change

- Emphasize the role of classes and why the compositionality framework is general enough.
- Undermine the relevance of the combined monitor and make it clearer how, why, and where it is needed.
- Incomporate an example where a second secure compilation pass introduces violations to the property the first pass secured against.

## Reviewer C

### General Comments

### Questions

### What will change

## Reviewer D

### General Comments

### Questions

### What will change

### 2023-09-14 popl/revs/rebuttal.md
Add more text

@@
 The compositionality framework is simple enough to appear as straightforward, which is a good thing.
-The catch is that building (formally verified) secure compilers is notoriously difficult, as backed up by an extensive number of secure compilation literature.
-With the presented compositionality framework, the burden of proving the composed compiler of secure compilation passes as secure is reduced dramatically.
+The catch is that building (formally verified) secure compilers is notoriously difficult, as backed up by an extensive number of secure compilation literature, so it can be quite involved to fulfill the assumptions of the compositionality theorems, as demonstrated in the case-study.
 
@@
 
+We are thankful for the thorough investigation of our work.
+In particular, we are delighted that you've taken a look at both our Coq formalisation and the technical report.
+
 ### General Comments
+
+> - The work is very foundational, granted, but the examples are very small and are not realistic compilers, so the main contribution that remains (if the formal proof is not really finalized, and if the concrete compiler only deals with toy languages) is the formal framework to talk about these results.
+
+We agree with this.
+While the programming languages in the case-study are quite restricted, we do have - at least on paper - a full proof that the compilers are secure.
+As of now, we are not aware of a practical implementation of a secure compiler that is formalized all the way (i.e., you can *run* the formalised version of the compiler), but MSWasm (Michael et al, 2023.
+ttps://doi.org/10.1145/3554344) comes close.
+We anticipate that our compositionality framework allows for an easier, incremental development of more realistic compilers.
+
+> - 211: I think an important notion that you do not highlight is that the language of traces must be the same for the source and target
+
+This may seem like a restriction, but using standard techniques from the secure compilation literature, the theory can be lifted without much friction to account for different trace models.
+
+> - 503: are you reversing the traditional order of the composition operator?
+
+Yes. Unfortunately, there are two communities that use either the one or the other order and view it as "the traditional" order.
+We are not opposed to simply swapping this, or maybe just add a footnote detailling how this composition should be interpreted.
+(it becomes clear from the color, but having it explicit is likely better)
+
+> - 512: you have a little bit of space left, so it would've been good to give an intuition of what exactly is difficult
+> - 547: same thing, I'm left wondering whether this is a deep result at all
+
+We agree that more discussion improves both presentations dramatically.
+
+> - 623: what is the $t$ in $\mathbb N_t$?
+
+$\mathbb{N}_t$ is the "concrete" syntax of the type.
+We thought it may make sense to disambiguate the type of natural numbers from the set of natural numbers.
+However, looking at the submmission again, this appears to just add clutter to the presentation.
+
+> - 666: I think you need to make it clear that some rules are omitted (e-get-\not\in, and e-set-\in) -- it took me a while to figure out what's going on
+
+Agreed, it should be stated explicitly that only a fraction of the rules are presented and that the interested reader can find the full picture in the technical report.
+
+> - you say e-set-\not\in returns a garbage value, in this case... v? why is it garbage? why not unit? (assignments in C return the assigned value, so this seems perfectly non-garbage)
+
+You are right that the word "garbage" is not fitting and we will adapt the text accordingly.
+We use the word in an attempt to convey the intuition that no write to memory actually happened, even though it appears so (because the rule yields `v`).
+
+> - 767: I never understood how the addition of L_ms enhanced readability
+
+Fair criticism.
+We should rephrase this paragraph and write that we want to make the fact explicit that the compiled code of $L_{ms}$ is instrumented to be memory-safe.
+
+> - 790: why not change the original get and set and add an additional argument to these events?
+
+That is another option. If the presentation is better with the additional argument instead of the hat, we'll gladly change it.
+
+> - 805: do you mean INactive?
+> - 805: text is confusing, because the rule that emits Binop n is not shown -- make it clear this is omitted; another nit: you have two styles for switching over whether the CPU mode is enabled; in one you use 0 in the conclusion of the rule, in the other, you use m \neq 0 in the premise; it would be clearer to use two constant on and off
+
+Yes, it should be *in*active and we think the presentation of this benefits from using `on` and `off` instead of $m\neq 0$ and $0$.
 
@@
 
+> - Clarify novelty of monitors and the trace projections
+
+The monitors use standard techniques to identify unsafe traces.
+In most existing work, monitors are used to define the property at hand and monitor satisfaction is exactly the same as trace satisfaction, in this case.
+In our work, while the techniques for monitors are not novel, the way they are used in order to do property composition and reason about this in the context of secure compilation is novel to that area of research.
+The way the projections of one trace model to another are set up enables definitions and satisfaction proofs of properties in a modular fashion.
+Moreover, we were unable to identify existing work that proves that monitor satisfaction implies trace satisfaction.
+
+> - Clarify what was difficult in the proofs and explain where the subtleties are -- right now it seems like some of the results might not be stating anything very deep (I might be under this impression because your paper is very well written!)
+
+The proofs for the compositionality framework are easy, even in Coq.
+Proving a compiler secure (i.e. robustly preserving) is notoriously difficult, as seen by numerous works in the secure compilation area.
+So, in order to fulfill the assumptions one needs at least two formally secure compilers.
+The presented framework allows to incrementally build more and more complicated compilers by focussing on the security of just an individual compilation pass.
+While the result may not be difficult or "deep", it is most certainly of great use to scale secure compilers to more practical applications.
+
+> - Give a status report on the Coq proof; surely you must've made progress since the submission, where are you at now?
+
+The compositionality framework is formally verified in Coq.
+While it is most certainly nice to also formally verify all parts of our case-study, we don't believe this is a necessity for acceptance of our submission (the most important results, the compositionality parts, are Coq'd).
+We have paper proofs for the case-study and it merely serves a motivation to show that our compositionality framework actually works and is of use in practice.
+
+> - Discuss applicability to more involved compilers.
+
+We think the submission may benefit from adding this discussion.
+Anyhow, our case-study shows that our framework scales to compilers with many passes as well as compilers with optimising passes, whose order may be swapped.
+The concrete complexity of any compilation pass involved in a composition is orthogonal to our work.
+Nevertheless, we strongly believe that secure, more complex/realistic compilers can be developed much faster with our framework.
+Without it, one would have to prove the whole compilation chain end-to-end, while our framework allows for modular, more localized reasoning.
+
 ### What will change
+
+- Fixing the typos!
+- Add more discussion about the compositionality theorems.
+- Change $\mathbb{N}_t$ to $\mathbb{N}$
+- Everywhere were we omit rules: State this explicitly.
+- Change descriptive text of e-set-\not\in to emphasize that no write happens. `v` is not really a "garbage" value, get rid of this confusing terminology. 
+- Rephrase paragraph about $L_{ms}$: the compiled code is instrumented to be memory-safe.
+- Change data independent timing mode state to either `on` or `off`.
+- Add discussion on applicability of the framework.
+- Add hyperlinks to the Coq development of the respective proofs.
 
@@
 
+We are thankful for the thorough analysis of our work, which we will be able to improve because of your feedback.
+
 ### General Comments
+
+> This paper studies the composition of robust safety properties and of robustly safe compilers.
+
+While our examples are only looking at safety properties, the compositionality framework itself is general enough to cover arbitrary properties.
+
+> While the paper does formalize "a theory of composition of secure compilers", focusing exclusively on robust preservation of properties, it  does not include any discussion about the limitations of this approach.
+
+This is most certainly true.
+In particular, it is interesting to discuss what happens if the second compilaton pass introduces vulnerabilities that the first pass fixes.
+
+> More specifically, the framework in the paper requires that all security properties and their monitors be defined on the same trace model.
+
+We believe this has been thoroughly investigated in the secure compilation literature. (Abate et al., 2021. https://doi.org/10.1145/3554344)
+Allowing different trace models, one can apply the techniques presented in there to make our framework applicable.
+
+> in particular, the languages selected are quite close with many shared definitions (see Sec 5.1).
+
+While we agree that the languages are syntactically very similar, this is merely an artifact to keep the proofs small.
+The complexity/"uniqueness" of the concrete languages have no impact on the presented compositionality framework.
+
+> - The choice to focus on sCCT instead of CCT is troubling.
+Of course, CCT is a hypersafety property so the paper's strategy of using monitors for checking would not have worked.
+But switching to Strict CCT, which requires that no secret _ever_ appear in a trace, seems too restrictive and it feels like the choice was made just so we can use monitors for checking.
+
+sCCT is one possible overapproximation of CCT which can be made less restrictive.
+If done so, the concrete details of the proofs for the CCT preservation - of course - change.
+However, the usage of the presented composition framework would stay exactly the same.
+
+> - The paper claims upper and lower composition as a contribution (and the lack thereof as a weakness of prior work, end of Sec 7.1).
+But the paper spends very little time motivating why these forms of composition are important or challenging.
+
+We agree and will add an appropriate discussion of this.
+
+> - There are some questionable decisions about which formalism to include and which to exclude.
+For a paper about secure compilation, there is not so much detail about the compilers (and the type systems) themselves.
+This seems especially concerning for the very first pass, from $L_{TMS}$ to $L$,
+> since the type system is not standard and dynamic enforcement is not necessarily straightforward.
+> [...]
+> If a piece of formalism is indeed obvious, the authors should consider trading that space to treat the content that is _not_ straightforward.
+
+We do have the impression that the precise language definitions are not interesting enough to warrant inclusion in the paper, i.e., that it is enough to mention the high-level properties that, e.g., could in parts be established by any work on memory-safety instrumentations or CCT work.
+The most important aspect of our submission is the compositionality framework and the concrete details of the case study can be looked up in the technical report.
+
+> - 213: Why use the powerset class of $\pi$ instead of the singleton class $\{\pi\}$?
+
+In short, the equivalent hyperproperty/class of a property is its powerset.
+If the class is a singleton, the intersection will very likely be empty and the compositionality framework would be rather useless.
+For details, refer to Clarkson and Schneider, 2008. https://doi.org/10.1145/3554344
+
+> - 221: Consider showing the code exercising the bug;  describing the code in prose makes it sound more complex than it is.
+
+We show the code and explain it. Do you argue for omitting the explanation to enhance clarity?
+
+> - 261: Isn't requiring deallocation a liveness property?
+
+This is a technicality, but yes. However, since the considered programs terminate, it is possible to check this.
+Nevertheless, since leaking memory is not a memory safety violation, it may improve the presentation to remove this requirement.
+Note that we've verified that the tms-monitor satisfaction implies the satisfaction of the presented tms property, with the deallocation requirement.
+
+> - 335: Should the type of allocated and freed be $\wp(L \times t)$?
+>  The use of $\in, \notin$ on 345, 348, 353 suggest so.
+
+We admit the confusion that comes from a slight abuse of notation.
+
+> - 455: $\textbf{Any}$ is not really any event, just the private ones,
+>   so a label like $\textbf{Leak}$ might be more suggestive.
+
+We agree that changing the name from $\textbf{Any}$ to $\textbf{Leak}$ conveys the semantics in a more meaningful way.
+
+> - 492: Have the results in *Sec. 4* been proven over arbitrary trace models,
+>   or only over the model described in *Sec. 3*?
+
+The results here are arbitrary, there is no connection to the case study or Section 3, i.e., this works for any kind of property and the quantification on classes is arbitrary.
+
+> - 594: $\vdash \gamma^{L}_{L + L} : \mathbb{C}_{1} \cap \mathbb{C}_{2}$ doesn't typecheck.
+
+We suspect that the notation is confusing, because after checking we still believe that this "typechecks".
+
+> - 605: What precisely does it mean for the languages to assume CCT holds?
+
+There is no leakage at all.
+
+> - 606: I find the messaging in this paragraph to be a touch inconsistent.
+>   It starts by saying that programmers shouldn't be thinking about CCT,
+>   and they don't have to, thanks to timing-independent hardware.
+>   But the paragraph finishes by emphasizing that 
+>   $L_{scct}$ allows the programmer to *disable* CCT protections,
+>   which disagrees with the previous suggestion.
+>   I suspect that these are intended to be different types of programmers,
+>   but perhaps that distinction should be made clear in a transition.
+
+You are right that this paragraph talks about different types of programmers and we do have to rephrase it to enhance clarity.
+
+> - 688: "... have two cases": meaning, each has an *additional* case not shown in the figure, correct?
+
+Yes, the concrete details can be looked up in the technical report.
+
+> - 703: After `Call ? foo v`, should we have `Return ! v` instead of `Return ? v`?
+
+No, this is not a typo. The `?` signals the switch from context to component, while the `!` signals the switch from component to context.
+However, there is a typo in line 704, where the `Return ? v` should be a `Return ! v`.
+
+> - 759: "If the array bound to `x` was allocated ...": should this say "deallocated"?
+
+Yes
+
+> - 780: Is it really that the languages always have timing independence turned on, or is it that those languages don't even have a notion of timing independence?
+There is nothing in the semantics of the languages that mentions timing or secrecy.
+It is just one implementation (i.e., the compiler in *Sec.
+6.4*) that establishes the assumption.
+
+You are right that it is one possible implementation and in line 780 we provide an intuition for that implementation.
+
+> - 874: Is a "strikethrough edge" the same as a solid edge?
+
+Yes
+
+> - 904: This compilation strategy seems sensible for simple functions,
+> but since no details about $L_{tms}$ or its type system are given,
+> it is hard to assess whether it should hold in general.
+
+The concrete details are uninteresting for the sake of the presentation here andthe interested reader can look them up in the technical report.
+We are nevertheless going to address your questions here:
+> > Are there higher-order functions? 
+No
+> > How are capabilities (assuming the types really are inspired by L3) compiled
+No, the type system is more restrictive and this renders the proofs simpler.
+
+> > and checked dynamically?
+L_tms is checked statically, anything else dynamically.
+
+> > Or is there a restriction on function types?
+Pointers cannot be passed. Otherwise, they would need to be mershalled to allow transfer from one sandbox to another.
+
+We are aware that this is a very restrictive setting, but it nevertheless allows memory unsafe behavior that allows for an interesting case-study.
+
+> - 1113: While upper and lower compositions are claimed as a contribution,
+> their importance was not sufficiently motivated.
+
+We strongly agree that they needs more explanation and discussion.
+
+> > Why are these forms of composition important?
+They allow modular extension of secure compilers to model compiler frameworks like LLVM.
+
+> > Do you suspect that they were provable in prior work but never studied,
+> > or was there a fundamental limitation that your framework has lifted?
+We have not seen this studied in prior work, so our framework does not lift an existing limitation, it rather presents a novel, albeit simple, approach.
 
@@
 
+> What are the limitations of this approach?
+Devising secure compilers is notoriously difficult and takes a lot of time.
+Moreover, one has to be careful that the intersection of classes does not become empty, which can happen if the first compilation pass secures against property A, but the second pass introduces source-code instrumentations that violate A.
+We'll add more detailed discussion about this in the paper.
+
+> How should one approach designing the common "target" trace language?
+In similar fashion to (Abate et al., 2021. https://doi.org/10.1145/3554344)
+
+> Does one need to know ahead of time which properties will be of interest?
+Yes and no.
+Yes, because for the composition itself, you do need concrete instances of classes that apply to your setting, but this class could be arbitrarily large.
+No, because it is not a roadblock to extend the models at a later point in time.
+
+> Are there any subtleties in either the type system or compilation of $L_{tms}$?
+
+No, we believe our work here is entirely standard and, because of this, did not bother to present the details in the paper, but keep them for the technical report.
+
 ### What will change
+
+- Add emphasis that the framework works for arbitrary properties.
+- Add discussion on limitations of the compositionality framework. Especially concerning the upper and lower composition.
+- Get rid of "as you would expect", "... is straightforward", etc.
+- Add quantifiers to disambiguate definitions.
+- Add paragraph on how sCCT overapproximates CCT and what kinds of programs would not satisfy sCCT while satisfying CCT.
+- Fix the "typing" of monitor states (e.g., line 335)
+- Rewrite $\textbf{Any}$ to $\textbf{Leak}$.
+- Rephrase paragraph at 606.
+- Fix typos.
+- Change wording from "strikethrough edge" to "solid edge".

### 2023-09-15 popl/revs/rebuttal.md
notes

@@
 
-First of all, we would like to sincerely thank every reviewer for putting in their hard work.
-We are pleased about the high-quality of the reviews and appreciate both your suggestions and questions, which we address for each reviewer below.
-Prior to that, we want to explain some general concerns applying to all reviews.
+First of all, we would like to thank every reviewer for putting in their hard work.
+We really appreciate the high-quality of the reviews and your insightful suggestions and questions.
+We address the latter, for each reviewer, below.
+Prior to that, we want to address some general concerns that are shared amongst different reviews.
 
-- The main results of our paper are about the compositionality of secure compilers, which we have Coq'd up at the time of submitting the paper.
-Missing Coq proofs mainly involve the case-study, but the paper does not claim to have Coq'd the language definitions, their properties, etc.
-- The compositionality scales to arbitrary properties, not just trace properties.
-Classes of properties are kept as general as possible and there is no restriction being made on the shape of a class without talking about a concrete one.
-This should also entail hyperproperties, since they are classes themselves and enjoy closedness on intersections, i.e., intersecting hyperproperties yields other hyperproperties.
-- The optimisations chosen, namely Dead-Code Elimination (dce) and Constant Folding (cf), are traditional compiler optimisations.
+- [Mechanisation and Incrementality of the Results] [TODO: rev ? ? ?]
+The main results of our paper, which is presented in Section TODO, are about the compositionality of secure compilers.
+All these results have been Coq'd up at and they can be found in the supplementary material.
+Neither of these results do [TODO ] rely on Admits.
+While it is true that we use the theory (and the Coq formalisation) of Abate et.
+al.
+[TODO cite number], we do not believe this work can be considered incremental.
+After all, no existing work tackles the problem of identifying what security property is preserved by the composition of secure compilers (despite the secure compilation community, and the reviewers as well, identify this as an interesting problem).
+The missing Coq proofs mainly involve the case-study, but we'd like to point out that the paper does not put forth this claim.
+
+- [Limitations of the Composition Theory] [TODO: rev ? ? ?]
+The compositionality property of Theorem [TODO] is presented for arbitrary classes of properties, not just trace properties.
+Classes of properties are kept as general as possible and so the theory works for hypersafety, hyperproperties, relational hyperproperties, and so forth.
+We will make this clearer in the presentation of the main results.
+An interesting point is that since hyperproperties enjoy closedness on intersections, intersecting classes of hyperproperties yields other hyperproperties. We will clarify this point further as well.
+
+- [Real-world Application of the Results] [TODO: rev ? ? ?]
+The optimisations chosen, namely Dead-Code Elimination (dce) and Constant Folding (cf), are traditional compiler optimisations.
 While they are simpler to define in comparison to other optimisations, we have no reason to believe that the compositionality framework breaks apart when scaling this to other passes.
@@
 
+[TODO: this does not seem to address a specific point?]
 We think it is important to not get lost in theory and appreciate your critical thinking about whether our theoretical framework is applicable to practical compilers.
@@
 
-Yes, the paper does this for dce and cf, which both GCC and LLVM do.
+Yes, we believe so, and this is the reason why in the paper we chose DCE and CF, which both GCC and LLVM do.
 We would appreciate further clarification about particular optimising transformations that you've had on your mind.
 While it is unlikely that we can extend the formalisation in the given time-frame to account for another optimisation, we do think this space warrants future work.
+[nope.
+rather, point out that some optimisations may break some props.
+that is not the scope of the work.
+we will look into that in the future.
+if you have suggestions for opts to look into, they're most welcome
+]
 
@@
 
-Yes, classes are simply sets of properties.
+Yes, as mentioned in the general points above, classes are simply sets of properties.
 We verified that the proof of the sequential composition theorem also works for classes of sets of hyperproperties.
-We agree that the paper should be rephrased to improve the description of classes and why the formulation is general enough, i.e., why the sequential composition theorem supports any set whatsoever (even liveness).
+We agree that the paper should be rephrased to improve the description of classes and why the formulation is general enough, i.e., why the sequential composition theorem supports any set (even liveness).
 

### 2023-09-15 popl/revs/rebuttal.md
Start to incomporate feedback

@@
 
-- [Mechanisation and Incrementality of the Results] [TODO: rev ? ? ?]
+- **Mechanisation and Incrementality of the Results** (rev C) [TODO: rev ? ? ?]
 The main results of our paper, which is presented in Section TODO, are about the compositionality of secure compilers.
@@
 
-- [Limitations of the Composition Theory] [TODO: rev ? ? ?]
+- **Limitations of the Composition Theory** [TODO: rev ? ? ?]
 The compositionality property of Theorem [TODO] is presented for arbitrary classes of properties, not just trace properties.
@@
 
-- [Real-world Application of the Results] [TODO: rev ? ? ?]
+- **Real-world Application of the Results** [TODO: rev ? ? ?]
 The optimisations chosen, namely Dead-Code Elimination (dce) and Constant Folding (cf), are traditional compiler optimisations.
@@
 ## Reviewer A
-
-[TODO: this does not seem to address a specific point?]
-We think it is important to not get lost in theory and appreciate your critical thinking about whether our theoretical framework is applicable to practical compilers.
-This is a big concern for us as well.
-Unfortunately, developing a compiler with formal methods is a task that has a huge workload (see CompCert).
-While there is no existing "practical", formally secure compiler, we believe that our framework allows to incrementally build such a thing, because of the separation of concerns that one gets from the compositionality framework.
-
 ### General comments
@@
 > You have chosen constant time cryptography and memory safety. It would be useful to discuss the motivation behind selecting these two security properties with their similarities and differences.
+
 We believe that the paper addresses this in the introduction: They are key security properties for ensuring confidentiality.
@@
 We would appreciate further clarification about particular optimising transformations that you've had on your mind.
-While it is unlikely that we can extend the formalisation in the given time-frame to account for another optimisation, we do think this space warrants future work.
-[nope.
-rather, point out that some optimisations may break some props.
-that is not the scope of the work.
-we will look into that in the future.
-if you have suggestions for opts to look into, they're most welcome
-]
+
+Some optimisations may break security properties.
+For example, by relying on undefined behavior, a necessary security check may be removed by an optimising compiler.
+Even (a more sophisticated) DCE could lead to confidentiality violations when it removes code that would override secret memory. 
+The investigation of these pecularities is outside the scope of this work, whose main goal is to provide a tool for composing compilation passes.
 
@@
 We verified that the proof of the sequential composition theorem also works for classes of sets of hyperproperties.
-We agree that the paper should be rephrased to improve the description of classes and why the formulation is general enough, i.e., why the sequential composition theorem supports any set (even liveness).
+We agree that the paper should be rephrased to improve the description of classes and why the formulation is general enough, i.e., why the sequential composition theorem supports any kind of property (even liveness).
 
@@
 ## Reviewer B
-
-While we do not agree with the reasoning for the decision to strongly reject our submission, we strongly appreciate parts of your criticism and will improve the paper as outlined below.
 
@@
 While we fail to see how the `e-dealloc` rule does not check for the existence of memory locations like the `e-set-*`/`e-get-*` rules, we do want to emphasize that `e-dealloc` does not need to change or check the heap(s) in any way, since it/they is/are modeled as one big arena.
-One could instead add the data the pointer points to into Δ instead of describing a two separate heaps. However, our formalisation models sandboxing more faithfully.
+One could instead add the data the pointer points to into Δ instead of describing two separate heaps. However, our formalisation models sandboxing more faithfully with this separation.
 
@@
 
-We are thankful for the thorough investigation of our work.
-In particular, we are delighted that you've taken a look at both our Coq formalisation and the technical report.
-
 ### General Comments
@@
 As of now, we are not aware of a practical implementation of a secure compiler that is formalized all the way (i.e., you can *run* the formalised version of the compiler), but MSWasm (Michael et al, 2023.
-ttps://doi.org/10.1145/3554344) comes close.
+https://doi.org/10.1145/3554344) comes close.
 We anticipate that our compositionality framework allows for an easier, incremental development of more realistic compilers.
@@
 
-This may seem like a restriction, but using standard techniques from the secure compilation literature, the theory can be lifted without much friction to account for different trace models.
+This may seem like a restriction, but using standard techniques from the secure compilation literature CITETHIS, the theory can be lifted without much friction to account for different trace models.
 
@@
 
-That is another option. If the presentation is better with the additional argument instead of the hat, we'll gladly change it.
+That is another option. If the presentation is better with the additional argument instead of the hat $\widehat{\cdot}$, we'll gladly change it.
 

### 2023-09-15 popl/revs/rebuttal.md
Start to incomporate feedback

@@
 
-This may seem like a restriction, but using standard techniques from the secure compilation literature CITETHIS, the theory can be lifted without much friction to account for different trace models.
+This may seem like a restriction, but using standard techniques from the secure compilation literature (Abate et al., 2021.
+https://doi.org/10.1145/3554344), the theory can be lifted without much friction to account for different trace models.
 
@@
 The way the projections of one trace model to another are set up enables definitions and satisfaction proofs of properties in a modular fashion.
-Moreover, we were unable to identify existing work that proves that monitor satisfaction implies trace satisfaction.
+Moreover, we were unable to identify existing work that proves that monitor satisfaction implies trace satisfaction, where the property is defined with just some condition instead of the monitor.
 
@@
 
-We are thankful for the thorough analysis of our work, which we will be able to improve because of your feedback.
-
 ### General Comments
@@
 This is most certainly true.
-In particular, it is interesting to discuss what happens if the second compilaton pass introduces vulnerabilities that the first pass fixes.
+In particular, it is interesting to discuss what happens if the second compilaton pass introduces vulnerabilities that the first pass fixes, as also pointed out by reviewer B.
 
@@
 
-We believe this has been thoroughly investigated in the secure compilation literature. (Abate et al., 2021. https://doi.org/10.1145/3554344)
+We believe this has been thoroughly investigated in the secure compilation literature (Abate et al., 2021. https://doi.org/10.1145/3554344).
 Allowing different trace models, one can apply the techniques presented in there to make our framework applicable.
@@
 
-The results here are arbitrary, there is no connection to the case study or Section 3, i.e., this works for any kind of property and the quantification on classes is arbitrary.
+The results here work for arbitrary trace models, there is no connection to the case study or Section 3, i.e., this works for any kind of property and the quantification on classes is arbitrary.
 
@@
 
-We suspect that the notation is confusing, because after checking we still believe that this "typechecks".
+We do not see why this does not "typecheck".
 
@@
 
-The concrete details are uninteresting for the sake of the presentation here andthe interested reader can look them up in the technical report.
+The concrete details are uninteresting for the sake of the presentation here and the interested reader can look them up in the technical report.
 We are nevertheless going to address your questions here:
 > > Are there higher-order functions?
+
 No
+
 > > How are capabilities (assuming the types really are inspired by L3) compiled
+
 No, the type system is more restrictive and this renders the proofs simpler.
@@
 > > and checked dynamically?
-L_tms is checked statically, anything else dynamically.
+
+L_tms is checked statically, anything else has (explicit, i.e., the programmer has to insert them) dynamic typechecks.
 
 > > Or is there a restriction on function types?
-Pointers cannot be passed. Otherwise, they would need to be mershalled to allow transfer from one sandbox to another.
+
+Pointers cannot be passed. Otherwise, they would need to be mershalled to allow transfer from one sandbox to another. Moreover, the type system would indeed need an extension that tracks capabilities.
 
@@
 
-We strongly agree that they needs more explanation and discussion.
+We strongly agree that they need more explanation and discussion.
 
 > > Why are these forms of composition important?
+
 They allow modular extension of secure compilers to model compiler frameworks like LLVM.
@@
 > > or was there a fundamental limitation that your framework has lifted?
-We have not seen this studied in prior work, so our framework does not lift an existing limitation, it rather presents a novel, albeit simple, approach.
+
+We have not seen this studied in prior work, so our framework does not lift an existing limitation, it rather presents a novel approach.
 
@@
 > What are the limitations of this approach?
+
 Devising secure compilers is notoriously difficult and takes a lot of time.
@@
 > How should one approach designing the common "target" trace language?
+
 In similar fashion to (Abate et al., 2021. https://doi.org/10.1145/3554344)
@@
 > Does one need to know ahead of time which properties will be of interest?
+
 Yes and no.
@@
 
-No, we believe our work here is entirely standard and, because of this, did not bother to present the details in the paper, but keep them for the technical report.
+No, we believe our work there is entirely standard and, because of this, did not bother to present the details in the paper, but keep them for the technical report.
 
@@
 - Add emphasis that the framework works for arbitrary properties.
-- Add discussion on limitations of the compositionality framework. Especially concerning the upper and lower composition.
+- Add discussion on limitations of the compositionality framework.
+Especially concerning the upper and lower composition.
+Highlight that one has to ensure that the intersections of classes do not become empty.
 - Get rid of "as you would expect", "... is straightforward", etc.

### 2023-09-15 popl/revs/rebuttal.md
Add reviewers to top-level notes

@@
 
-- **Mechanisation and Incrementality of the Results** (rev C) [TODO: rev ? ? ?]
+- **Mechanisation and Incrementality of the Results** (reviewer B, C)
+
 The main results of our paper, which is presented in Section TODO, are about the compositionality of secure compilers.
@@
 
-- **Limitations of the Composition Theory** [TODO: rev ? ? ?]
+- **Limitations of the Composition Theory** (reviewer A, D)
+
 The compositionality property of Theorem [TODO] is presented for arbitrary classes of properties, not just trace properties.
@@
 
-- **Real-world Application of the Results** [TODO: rev ? ? ?]
+- **Real-world Application of the Results** (reviewer A, C)
+
 The optimisations chosen, namely Dead-Code Elimination (dce) and Constant Folding (cf), are traditional compiler optimisations.
 While they are simpler to define in comparison to other optimisations, we have no reason to believe that the compositionality framework breaks apart when scaling this to other passes.
+Individual secure compilation proofs for these more realistic passes will (naturally) get more involved.
 

### 2023-09-15 popl/revs/rebuttal.md
Erradicate all TODO

@@
 
-The main results of our paper, which is presented in Section TODO, are about the compositionality of secure compilers.
+The main results of our paper, which are presented in Section 4, are about the compositionality of secure compilers.
 All these results have been Coq'd up at and they can be found in the supplementary material.
-Neither of these results do [TODO ] rely on Admits.
-While it is true that we use the theory (and the Coq formalisation) of Abate et.
-al.
-[TODO cite number], we do not believe this work can be considered incremental.
-After all, no existing work tackles the problem of identifying what security property is preserved by the composition of secure compilers (despite the secure compilation community, and the reviewers as well, identify this as an interesting problem).
-The missing Coq proofs mainly involve the case-study, but we'd like to point out that the paper does not put forth this claim.
+Neither of these results rely on Admits.
+While it is true that we use the theory of (Abate et. al. 2019), we do not believe this work can be considered incremental.
+After all, no existing work tackles the problem of identifying what property is preserved by the composition of secure compilers (despite the secure compilation community, and the reviewers as well, identify this as an interesting problem).
+The missing Coq proofs mainly involve the case-study, but we'd like to point out that the paper does not put forth the claim that the case-study is fully formalised in Coq.
 
@@
 
-The compositionality property of Theorem [TODO] is presented for arbitrary classes of properties, not just trace properties.
+The compositionality property of Theorems 4.2, 4.3, 4.5, and 4.7 is presented for arbitrary classes of properties, not just trace properties.
 Classes of properties are kept as general as possible and so the theory works for hypersafety, hyperproperties, relational hyperproperties, and so forth. 
 We will make this clearer in the presentation of the main results.
-An interesting point is that since hyperproperties enjoy closedness on intersections, intersecting classes of hyperproperties yields other hyperproperties. We will clarify this point further as well.
 
@@
 Briefly, if the second compiler generates programs that always violate the property that the first ensures, then the preserved class of properties is actually empty, since the intersection is disjoint!
-So, the theorem as stated still works, i.e., it is *not* unsound, but it may trivialize, which should better be emphasized in the paper.
+So, the theorem as stated still works, i.e., it is *not* unsound, but it may trivialise, which should better be emphasised in the paper.
 
@@
 While optimality may be interesting, it is also orthogonal to the goal of the paper, which cares about doing the correct thing.
-Subsequent work can optimize this, if necessary.
+Subsequent work can optimise this, if necessary.
 
@@
 The purpose of the case-study is to allow memory-unsafe behavior, i.e., execution should not get stuck if there is some kind of, e.g., memory-safety violation.
-While we fail to see how the `e-dealloc` rule does not check for the existence of memory locations like the `e-set-*`/`e-get-*` rules, we do want to emphasize that `e-dealloc` does not need to change or check the heap(s) in any way, since it/they is/are modeled as one big arena.
+While we fail to see how the `e-dealloc` rule does not check for the existence of memory locations like the `e-set-*`/`e-get-*` rules, we do want to emphasise that `e-dealloc` does not need to change or check the heap(s) in any way, since it/they is/are modeled as one big arena.
 One could instead add the data the pointer points to into Δ instead of describing two separate heaps. However, our formalisation models sandboxing more faithfully with this separation.
@@
 
-Yes, this is intended, since this precise timing is abstracted away by the data independent timing mode + techniques used in the FaCT paper (Cauligi et al., 2019.
+Yes, this is intended, since this precise timing is abstracted away by the data independent timing mode + techniques used in the FaCT paper (Cauligi et.
+al., 2019.
 https://doi.org/10.1145/3314221.3314605).
@@
 
-- Emphasize the role of classes and why the compositionality framework is general enough.
+- Emphasise the role of classes and why the compositionality framework is general enough.
 - Undermine the relevance of the combined monitor and make it clearer how, why, and where it is needed.
@@
 
-> - The work is very foundational, granted, but the examples are very small and are not realistic compilers, so the main contribution that remains (if the formal proof is not really finalized, and if the concrete compiler only deals with toy languages) is the formal framework to talk about these results.
+> - The work is very foundational, granted, but the examples are very small and are not realistic compilers, so the main contribution that remains (if the formal proof is not really finalised, and if the concrete compiler only deals with toy languages) is the formal framework to talk about these results.
 
@@
 While the programming languages in the case-study are quite restricted, we do have - at least on paper - a full proof that the compilers are secure.
-As of now, we are not aware of a practical implementation of a secure compiler that is formalized all the way (i.e., you can *run* the formalised version of the compiler), but MSWasm (Michael et al, 2023.
+As of now, we are not aware of a practical implementation of a secure compiler that is formalised all the way (i.e., you can *run* the formalised version of the compiler), but MSWasm (Michael et.
+al, 2023.
 https://doi.org/10.1145/3554344) comes close.
@@
 
-This may seem like a restriction, but using standard techniques from the secure compilation literature (Abate et al., 2021.
+This may seem like a restriction, but using standard techniques from the secure compilation literature (Abate et.
+al., 2021.
 https://doi.org/10.1145/3554344), the theory can be lifted without much friction to account for different trace models.
@@
 Nevertheless, we strongly believe that secure, more complex/realistic compilers can be developed much faster with our framework.
-Without it, one would have to prove the whole compilation chain end-to-end, while our framework allows for modular, more localized reasoning.
+Without it, one would have to prove the whole compilation chain end-to-end, while our framework allows for modular, more localised reasoning.
 
@@
 - Everywhere were we omit rules: State this explicitly.
-- Change descriptive text of e-set-\not\in to emphasize that no write happens. `v` is not really a "garbage" value, get rid of this confusing terminology. 
+- Change descriptive text of e-set-\not\in to emphasise that no write happens. `v` is not really a "garbage" value, get rid of this confusing terminology. 
 - Rephrase paragraph about $L_{ms}$: the compiled code is instrumented to be memory-safe.
@@
 
-> While the paper does formalize "a theory of composition of secure compilers", focusing exclusively on robust preservation of properties, it  does not include any discussion about the limitations of this approach.
+> While the paper does formalise "a theory of composition of secure compilers", focusing exclusively on robust preservation of properties, it  does not include any discussion about the limitations of this approach.
 
@@
 
-We believe this has been thoroughly investigated in the secure compilation literature (Abate et al., 2021. https://doi.org/10.1145/3554344).
+We believe this has been thoroughly investigated in the secure compilation literature (Abate et. al., 2021. https://doi.org/10.1145/3554344).
 Allowing different trace models, one can apply the techniques presented in there to make our framework applicable.
@@
 
-In similar fashion to (Abate et al., 2021. https://doi.org/10.1145/3554344)
+In similar fashion to (Abate et. al., 2021. https://doi.org/10.1145/3554344)
 

### 2023-09-15 popl/revs/rebuttal.md
Wording

@@
 First of all, we would like to thank every reviewer for putting in their hard work.
-We really appreciate the high-quality of the reviews and your insightful suggestions and questions.
+We really appreciate the high quality of the reviews and your insightful suggestions and questions.
 We address the latter, for each reviewer, below.
@@
 The main results of our paper, which are presented in Section 4, are about the compositionality of secure compilers.
-All these results have been Coq'd up at and they can be found in the supplementary material.
+All these results have been Coq'd up and they can be found in the supplementary material.
 Neither of these results rely on Admits.
-While it is true that we use the theory of (Abate et. al. 2019), we do not believe this work can be considered incremental.
+While it is true that we use the theory of (Abate et. al. 2019. https://doi.org/10.48550/arXiv.2006.14969), we do not believe this work can be considered incremental.
 After all, no existing work tackles the problem of identifying what property is preserved by the composition of secure compilers (despite the secure compilation community, and the reviewers as well, identify this as an interesting problem).
@@
 
-The compositionality property of Theorems 4.2, 4.3, 4.5, and 4.7 is presented for arbitrary classes of properties, not just trace properties.
-Classes of properties are kept as general as possible and so the theory works for hypersafety, hyperproperties, relational hyperproperties, and so forth. 
+The compositionality properties stated in Theorems 4.2, 4.3, 4.5, and 4.7 are presented for arbitrary classes of properties, not just trace properties.
+Classes of properties are kept as general as possible and so the theory works for (hyper-)safety, general hyperproperties, relational hyperproperties, and so forth. 
 We will make this clearer in the presentation of the main results.
@@
 
-The optimisations chosen, namely Dead-Code Elimination (dce) and Constant Folding (cf), are traditional compiler optimisations.
+The optimisations chosen, namely Dead-Code Elimination (DCE) and Constant Folding (CF), are traditional compiler optimisations.
 While they are simpler to define in comparison to other optimisations, we have no reason to believe that the compositionality framework breaks apart when scaling this to other passes.
-Individual secure compilation proofs for these more realistic passes will (naturally) get more involved.
+Individual secure compilation proofs for these more realistic passes will (naturally) get more involved, but this localized reasoning has no influence on the compositionality framework of this work.
 
-## Reviewer A
+Reviewer A
+==========
 ### General comments
@@
 
-The paper studies dead-code elimination (dce) as well as constant folding (cf) in the setting of (formally) secure compilers.
+The paper studies dead-code elimination (DCE) as well as constant folding (CF) in the setting of (formally) secure compilers.
 So far, the only practical compiler that involves formal methods is CompCert, but it is not proven secure.
-Nevertheless, CompCert, as well as not-formally proven compilers, such as gcc or clang, perform dce and cf.
-dce and cf are foundational compiler optimisations that any optimising compiler makes use of.
-Therefore, we believe our approach is as applicable as formal methods when it comes to state-of-the-art compiler development.
+Nevertheless, CompCert, as well as not-formally proven compilers, such as gcc or clang, perform DCE and CF.
+DCE and CF are foundational compiler optimisations that any optimising compiler makes use of.
+Therefore, we believe our approach is as applicable as formal methods when it comes to state-of-the-art compiler development, given that this work presents a tool to reduce the burden when applying formal methods to compilers.
 
@@
 
-## Reviewer B
+Reviewer B
+==========
 
@@
 
-## Reviewer C
+Reviewer C
+==========
 
@@
 
-## Reviewer D
+Reviewer D
+==========
 

### 2023-09-15 popl/revs/rebuttal.md
pass

@@
 
-- **Mechanisation and Incrementality of the Results** (reviewer B, C)
+- **Mechanisation and Incrementality of the Results** (reviewers B, C)
 
@@
 While it is true that we use the theory of (Abate et. al. 2019. https://doi.org/10.48550/arXiv.2006.14969), we do not believe this work can be considered incremental.
-After all, no existing work tackles the problem of identifying what property is preserved by the composition of secure compilers (despite the secure compilation community, and the reviewers as well, identify this as an interesting problem).
+After all, no existing work tackles the problem of identifying what property is preserved by the composition of secure compilers (despite the secure compilation community, and the reviewers as well, identify this as an crucial problem).
 The missing Coq proofs mainly involve the case-study, but we'd like to point out that the paper does not put forth the claim that the case-study is fully formalised in Coq.
 
-- **Limitations of the Composition Theory** (reviewer A, D)
+- **Limitations of the Composition Theory** (reviewers A, D)
 
@@
 Classes of properties are kept as general as possible and so the theory works for (hyper-)safety, general hyperproperties, relational hyperproperties, and so forth. 
+We rely on just safety properties in the case study because of the simplicity that comes with that specific class, but the main theory of Section 4 is **not** tied to just safety.
 We will make this clearer in the presentation of the main results.
@@
 The optimisations chosen, namely Dead-Code Elimination (DCE) and Constant Folding (CF), are traditional compiler optimisations. 
+[TODO the line that they exist in rw compilers worked imho]
 While they are simpler to define in comparison to other optimisations, we have no reason to believe that the compositionality framework breaks apart when scaling this to other passes.
 Individual secure compilation proofs for these more realistic passes will (naturally) get more involved, but this localized reasoning has no influence on the compositionality framework of this work.
+[TODO: We believe studying what class of security properties various optimisations preserve, but that is beyond the scope of this work]
 
@@
 So far, the only practical compiler that involves formal methods is CompCert, but it is not proven secure.
+[TODO: nono, there's cakeml. tone this down. I'd say 'a great reference' in place of 'the only practical']
 Nevertheless, CompCert, as well as not-formally proven compilers, such as gcc or clang, perform DCE and CF.
@@
 For the framework presented in the paper, undefined behavior would be encoded in the property/-ies of interest and, therefore, be preserved in the intersection of property classes.
-We are planning to do a follow-up paper where we investigate more aggressive, realistic optimising compiler passes that (1) may exploit undefined behavior, (2) reorder code - and thereby reorder events on the trace, (3) memory layout reording, and so on.
+We are planning to do a follow-up paper where we investigate more aggressive, optimising compiler passes that (1) may exploit undefined behavior, (2) reorder code - and thereby reorder events on the trace, (3) perform memory layout reording, and so on.
+Now, it is still unclear how UB will affect the reasoning about the composition of these properties, but we believe this to be beyond the scope of the paper.
 
@@
 We believe that the paper addresses this in the introduction: They are key security properties for ensuring confidentiality.
+We will make this clearer.
 
@@
 - Add to future work section that we will look at more sophisticated optimisiation passes.
+[TODO: Rephrase intro for clarificatio ... essentially, take a good look above at the promises are re-list them here]
 
@@
 
-From our literature review, we were not able to identify work that discusses the composition of (formally) secure compilers.
+From our literature review, we were not able to identify work that discusses the composition of secure compilers.
 The compositionality of secure compilers has not been discussed before, and, as you said, it is an important problem to look at.
@@
 
-Note that, while the universal quantification in the sequential composition theorem restricts properties to be part of a certain class (=set of properties), the class is arbitrarily fixed.
-For example, in the case-study, it is fixed to the class of memory safety properties which is later composed with the class of strictly cryptographic constant timeness.
+We do not believe this statement to be correct.
+The theory of Section 4 is general in the class of property that gets considered, and this includes hypersafety, hyperproperties and so forth.
+The use cases rely on safety properties because many interesting security properties are safety, and safety properties are easier to handle in secure compilation proofs (despite these proofs being rather convoluted already).
+Also, while the universal quantification in the sequential composition theorem restricts properties to be part of a certain class (=set of properties), the class is arbitrarily fixed.
 
@@
 
+Concerning the challenges, complexity can be arbitrary here.
+Experts in secure compilation may find setting up the usual 2-3 cross-language relations (per compiler) and the backtranslation easier than non-experts.
+Fortunately we did not have to rely on novel proof techniques, and could rely on established secure compilation proof theory.
+There is quite some proof engineering work in order to maximise proof reuse though.
+
 > Definition 3.1, "at most one Alloc/Dealloc l;t;sigma" is imprecise.
@@
 While optimality may be interesting, it is also orthogonal to the goal of the paper, which cares about doing the correct thing.
+After all, we do not care about running the monitors in parallel (e.g., for efficient runtime monitornig) but just to validate that both properties hold.
 Subsequent work can optimise this, if necessary.
@@
 
-We want to note that past iterations of POPL already have several papers that make extensive use of color.
-Nevertheless, removing color is a minor change that should not influence the acceptance of the paper.
+We want to note that past iterations of POPL already have several papers that make extensive use of color to aide readers.
 
@@
 Yes, see earlier discussion. In short, one has to be careful that the intersection does not become empty, but the theorem is nonetheless applicable and sound.
-It's been proven in Coq.
+Also, please not this has been proven in Coq.
 
@@
 - Emphasise the role of classes and why the compositionality framework is general enough.
-- Undermine the relevance of the combined monitor and make it clearer how, why, and where it is needed.
+- Underline the relevance of the combined monitor and make it clearer how, why, and where it is needed.
 - Incomporate an example where a second secure compilation pass introduces violations to the property the first pass secured against.
@@
 
-We agree that more discussion improves both presentations dramatically.
+We agree that more discussion improves both presentations.
 
@@
 We thought it may make sense to disambiguate the type of natural numbers from the set of natural numbers.
-However, looking at the submmission again, this appears to just add clutter to the presentation.
+However, looking at the submmission again, this appears to just add clutter to the presentation._
 
@@
 The proofs for the compositionality framework are easy, even in Coq.
+[TODO: mmmm this may backfire.
+i'd add at least: however, we are given to understand that it is the relevance of the work that matters, not its technical complexity (which, we also provide in some of the contributions)]
 Proving a compiler secure (i.e. robustly preserving) is notoriously difficult, as seen by numerous works in the secure compilation area.
@@
 For details, refer to Clarkson and Schneider, 2008. https://doi.org/10.1145/3554344
+[TODO: this is a bit vague for a ref here. point to a specific section perhaps?]
 
@@
 
-We show the code and explain it. Do you argue for omitting the explanation to enhance clarity?
+We [TODO: will?] show the code and explain it. Do you argue for omitting the explanation to enhance clarity?
 
@@
 
-> - 594: $\vdash \gamma^{L}_{L + L} : \mathbb{C}_{1} \cap \mathbb{C}_{2}$ doesn't typecheck.
+> - 594: $\vdash \gamma^{L}_{L + L} : \mathbb{C}_{1} \cap \mathbb{C}_{2}$ doesn't typecheck._
 
-We do not see why this does not "typecheck".
+We do not see why this does not "typecheck", could you please elaborate?
 
@@
 
-Pointers cannot be passed. Otherwise, they would need to be mershalled to allow transfer from one sandbox to another. Moreover, the type system would indeed need an extension that tracks capabilities.
+Pointers cannot be passed. Otherwise, they would need to be marshalled to allow transfer from one sandbox to another. Moreover, the type system would indeed need an extension that tracks capabilities.
 
@@
 Devising secure compilers is notoriously difficult and takes a lot of time.
+[TODO: this is not a limitation]
 Moreover, one has to be careful that the intersection of classes does not become empty, which can happen if the first compilation pass secures against property A, but the second pass introduces source-code instrumentations that violate A.
 We'll add more detailed discussion about this in the paper.
+[TODO: limitations may be that we don't discuss changes in the trace model]
 
@@
 In similar fashion to (Abate et. al., 2021. https://doi.org/10.1145/3554344)
+[TODO: sammler's work dimsum too, perhsps?]
 
@@
 No, we believe our work there is entirely standard and, because of this, did not bother to present the details in the paper, but keep them for the technical report.
+[TODO: standard == needs refereces? like amals' linear paper]
 

### 2023-09-15 popl/revs/rebuttal.md
add final rebuttal

@@
 
-- **Real-world Application of the Results** (reviewer A, C)
+- **Real-world Application of the Results** (reviewers A, C)
 
-The optimisations chosen, namely Dead-Code Elimination (DCE) and Constant Folding (CF), are traditional compiler optimisations. 
-[TODO the line that they exist in rw compilers worked imho]
+The optimisations chosen, namely Dead-Code Elimination (DCE) and Constant Folding (CF), are traditional compiler optimisations that exist in real-world compilers, e.g. GCC, Clang, Rustc, and so on. 
 While they are simpler to define in comparison to other optimisations, we have no reason to believe that the compositionality framework breaks apart when scaling this to other passes.
 Individual secure compilation proofs for these more realistic passes will (naturally) get more involved, but this localized reasoning has no influence on the compositionality framework of this work.
-[TODO: We believe studying what class of security properties various optimisations preserve, but that is beyond the scope of this work]
+Eventually, we want to study what class of securityd properties various optimisations preserve, but that is beyond the scope of this work.
 
@@
 The paper studies dead-code elimination (DCE) as well as constant folding (CF) in the setting of (formally) secure compilers.
-So far, the only practical compiler that involves formal methods is CompCert, but it is not proven secure.
-[TODO: nono, there's cakeml. tone this down. I'd say 'a great reference' in place of 'the only practical']
+So far, there are few compilers based on formal methods.
+One great example of this is CompCert, a practical compiler for C that is formally proven correct, but not secure.
 Nevertheless, CompCert, as well as not-formally proven compilers, such as gcc or clang, perform DCE and CF.
@@
 Yes, as mentioned in the general points above, classes are simply sets of properties.
-We verified that the proof of the sequential composition theorem also works for classes of sets of hyperproperties.
+We verified that the proof of the compositionality theorems also works for classes of sets of hyperproperties, as well as just hyperproperties.
 We agree that the paper should be rephrased to improve the description of classes and why the formulation is general enough, i.e., why the sequential composition theorem supports any kind of property (even liveness).
@@
 - Add to future work section that we will look at more sophisticated optimisiation passes.
-[TODO: Rephrase intro for clarificatio ... essentially, take a good look above at the promises are re-list them here]
+- Rephrase introduction for clarification on why we have chosen memory safety and cryptographic constant time.
 
@@
 Yes, see earlier discussion. In short, one has to be careful that the intersection does not become empty, but the theorem is nonetheless applicable and sound.
-Also, please not this has been proven in Coq.
+Also, please note that this work has proven this in Coq.
 
@@
 We thought it may make sense to disambiguate the type of natural numbers from the set of natural numbers.
-However, looking at the submmission again, this appears to just add clutter to the presentation._
+However, looking at the submmission again, this appears to just add clutter to the presentation.
 
@@
 
-The proofs for the compositionality framework are easy, even in Coq.
-[TODO: mmmm this may backfire.
-i'd add at least: however, we are given to understand that it is the relevance of the work that matters, not its technical complexity (which, we also provide in some of the contributions)]
+The proofs for the compositionality framework are not involved.
 Proving a compiler secure (i.e. robustly preserving) is notoriously difficult, as seen by numerous works in the secure compilation area.
@@
 If the class is a singleton, the intersection will very likely be empty and the compositionality framework would be rather useless.
-For details, refer to Clarkson and Schneider, 2008. https://doi.org/10.1145/3554344
-[TODO: this is a bit vague for a ref here. point to a specific section perhaps?]
+For details, refer to page 6 in Clarkson and Schneider, 2008. https://doi.org/10.1145/3554344
 
@@
 
-We [TODO: will?] show the code and explain it. Do you argue for omitting the explanation to enhance clarity?
+In the referred paragraph, we show the code and explain how it works. 
+Do you argue for omitting the explanation to enhance clarity?
 
@@
 
-Devising secure compilers is notoriously difficult and takes a lot of time.
-[TODO: this is not a limitation]
+Devising secure compilers is notoriously difficult and takes a lot of time, but this is not a limitation of this specific work.
 Moreover, one has to be careful that the intersection of classes does not become empty, which can happen if the first compilation pass secures against property A, but the second pass introduces source-code instrumentations that violate A.
 We'll add more detailed discussion about this in the paper.
-[TODO: limitations may be that we don't discuss changes in the trace model]
+We think one limitation may be the unified trace model, for which we will add discussion.
 
@@
 In similar fashion to (Abate et. al., 2021. https://doi.org/10.1145/3554344)
-[TODO: sammler's work dimsum too, perhsps?]
 
@@
 No, we believe our work there is entirely standard and, because of this, did not bother to present the details in the paper, but keep them for the technical report.
-[TODO: standard == needs refereces? like amals' linear paper]
 
@@
 - Change wording from "strikethrough edge" to "solid edge".
+- Add discussion about the unified trace model vs. multiple different trace models

### 2023-09-15 popl/revs/rebuttal.md
Fix a bunch of typos

@@
 - Add more discussion about the compositionality theorems.
-- Change $\mathbb{N}_t$ to $\mathbb{N}$
++ Change $\mathbb{N}_t$ to $\mathbb{N}$
 - Everywhere were we omit rules: State this explicitly.
@@
 - Rephrase paragraph about $L_{ms}$: the compiled code is instrumented to be memory-safe.
-- Change data independent timing mode state to either `on` or `off`.
++ Change data independent timing mode state to either `on` or `off`.
 - Add discussion on applicability of the framework.
@@
 - Fix typos.
-- Change wording from "strikethrough edge" to "solid edge".
++ Change wording from "strikethrough edge" to "solid edge".
 - Add discussion about the unified trace model vs. multiple different trace models

### 2023-09-15 popl/main.tex
fix typo
Fix a bunch of typos

@@
 Starting from a general theory of property composition for security-relevant properties (such as the aforementioned ones), this paper formalises a theory of composition of secure compilers.
-Then, it showcases this theory a secure multi-pass compiler that preserves the aforementioned security-relevant properties.
+Then, it showcases this theory on a secure multi-pass compiler that preserves the aforementioned security-relevant properties.
 % conclusion
@@
 % Even under these assumptions, there is a 
-This function is subject to a subtle \gls*{sms} vulnerability: The bounds check \texttt{i < n} should happen {\it before} the access to memory location \texttt{x[i]}: otherwise
+This function is subject to a subtle \gls*{sms} vulnerability: the bounds check \texttt{i < n} should happen {\it before} the access to memory location \texttt{x[i]}: otherwise
 % If \texttt{n} is at least as large as the length of the array \texttt{src}, 
@@
 To prevent this vulnerability, one can use a compilation pass that enforces \gls*{sms}, such as Softbounds~\cite{nagarakatte2009soft} or BaggyBounds~\cite{akritidis2009baggy}.
+The most naive solution in this case is to insert bounds-checks in front of every access to memory.
 
@@
 
-Alas, code is not run in isolation, so a malicious attacker could supply code that intracts with \texttt{strncpy} and trigger a violation of either \gls*{ms} or \gls*{cct} by calling \texttt{strncpy} with an argument for \texttt{src} that points to uninitialised memory.
+Alas, code is not run in isolation, so a malicious attacker could supply code that interacts with \texttt{strncpy} and trigger a violation of either \gls*{ms} or \gls*{cct} by calling \texttt{strncpy} with an argument for \texttt{src} that points to uninitialised memory.
 This would, in turn, triggering a series of reads from uninitialised memory, which is an immediate \gls*{ms} violation with devastating real-world consequences~\cite{uninit-0,uninit-1,uninit-2,uninit-3,uninit-4}.
@@
   $\isdef$
-  if \iul{$\progstepto{\wholeprogvar}{\runtimetermvar}{\trace}$},
+  if \iul{$\progstepto{\progvar}{\runtimetermvar}{\trace}$},
   then \oul{$\trace\in\pi$}.
@@
 This translation is considered {\em correct} if it is semantics-preserving.
-That is, for a whole program $\src{\wholeprogvar}$, the compiler should relate the $\src{L}$ semantics of $\src{\wholeprogvar}$ with the semantics of $\trg{T}$ of the compiled counterpart of $\src{p}$ in such a way that they are ,,compatible''.
+That is, for a whole program $\src{\wholeprogvar}$, the compiler should relate the $\src{L}$ semantics of $\src{\wholeprogvar}$ with the semantics of $\trg{L}$ of the compiled counterpart of $\src{p}$ in such a way that they are ''compatible''.
 Unfortunately, correct compilers may be insecure compilers~\cite{patrignani2019survey,kennedy2006secure.net,abadi1999protect,ahmed2018dagstuhl} and programs translated by insecure compilers can violate security properties that the programmer assumes to hold.
@@
   $\isdef$
-  \rul{$\forall \left(\pi\in\class\right) \left(\src{p}\in\Ltms\right),$} if \iul{$\rsat{\src{\progvar}}{\pi}$}, then \oul{$\rsat{\cc{\src{L}}{\trg{L}}\left(\src{p}\right)}{\pi}$}.
+  \rul{$\forall \left(\pi\in\class\right) \left(\src{p}\in\src{L}\right),$} if \iul{$\rsat{\src{\progvar}}{\pi}$}, then \oul{$\rsat{\cc{\src{L}}{\trg{L}}\left(\src{p}\right)}{\pi}$}.
 \end{definition}
@@
 \gls*{cct} is a hypersafety property~\cite{barthe2018sec} and, thus, difficult to check with monitors.
-This is because, intuitively, hypersafety properties can relate multiple execution traces with eachother, but monitors work on a single execution.
+This is because, intuitively, hypersafety properties can relate multiple execution traces with each other, but monitors work on a single execution.
 To sidestep this issue, this section defines the property \gls*{scct}, a stricter variant of \gls*{cct} that enforces the policy that no secret appears on a trace (inspired by earlier work~\cite{almeida2017jasmin}).
@@
 
-Monitors enforce safety properties by accepting or rejecting traces, i.e., if it rejects a trace, the trace does not satisfy the property the monitor checks.
+Monitors enforce safety properties by accepting or rejecting traces, i.e., if they reject a trace, the trace does not satisfy the property they check.
 Since reasoning on monitors is easier than directly on just traces, this section presents a monitor for each of the previously shown safety properties (\Cref{subsec:propdefs}).
@@
   \typerule{tms-trans}{
-    \monitorcheck[]{\tmsmonitor}{\tmsmonitor[']}{\absevent}
+    \monitorcheck[]{\tmsmonitor}{\tmsmonitor[']}{\absevent} &
     \monitorcheck{\tmsmonitor[']}{\tmsmonitor['']}{\abstrace}
@@
 \end{center}
-\Cref{tr:tms-ign-trans,tr:tms-trans} are the same for all monitors, but \Cref{tr:tms-refl} has, in this case, an additional premise that no more locations should be allocated.
+\Cref{tr:tms-ign-trans,tr:tms-trans} are the same for all monitors, but \Cref{tr:tms-refl} has, in this case, an additional premise that all locations should be deallocated.
 This rejects the behavior of programs that forget to free memory.
@@
     \mi{(Abstract\ Store)}~\scctmonitor&:=\ \emptyset \hspace{0.5cm}
-    \mi{(Abstract\ Events)}~\absevent:=\ \absev{\varepsilon} \mid \absterm \mid \absev{Any} \\
+    \mi{(Abstract\ Events)}~\absevent:=\ \absev{\varepsilon} \mid \absterm \mid \absev{Leak} \\
   \end{aligned}
@@
   }{
-    \sccttraceagree[]{\event[_b];\sandboxtag;\lock}{\absev{Any}}
+    \sccttraceagree[]{\event[_b];\sandboxtag;\lock}{\absev{Leak}}
   }{scct-high-authentic}
@@
 \end{center}
-Accordingly, the event agreement simply disregards all events that involved public data ($\unlock$) while mapping any other event that does involve private data ($\lock$) to the abstract $\absev{Any}$ event.
+Accordingly, the event agreement simply disregards all events that involved public data ($\unlock$) while mapping any other event that does involve private data ($\lock$) to the abstract $\absev{Leak}$ event.
 
@@
   The trace of just the copying part inside of \texttt{strncpy} looks like $\ev{Use}\ \loc_{x}\ 0;\comp;\unlock\cdot\ev{Use}\ \loc_{y}\ 0;\comp;\lock$.
-  In terms of the abstract monitor trace, this is just $\absev{Any}$.
-  Running this on the monitor would result in getting stuck, since there is no matching rule to step in the presence of $\absev{Any}$ event. % pun intended
+  In terms of the abstract monitor trace, this is just $\absev{Leak}$.
+  Running this on the monitor would result in getting stuck, since there is no matching rule to step in the presence of $\absev{Leak}$ event. % pun intended
   By means of additional features to ensure cryptographic constant time even in the presence of memory reads and loads, such as a flag to enable a data independent timing mode, which is present in both Arm~\cite[p.~543]{arm-refman} and Intel~\cite[p.~80]{intel-refman} processors, the original trace now does $\ev{Use}\ \loc_{y}\ 0;\comp;\unlock$ instead of $\ev{Use}\ \loc_{y}\ 0;\comp;\lock$.
@@
 All languages use the same trace model, where events are either the empty event \emptyevent, the program crash event $\lightning$, or a tuple consisting of a control tag $\sandboxtag$ and a security tag $\securitytag$.
-The former indicates whether the component \comp or the context \ctx is to blame for emitting this event, the latter indicates the secrecy level of values of the emitted event, i.e., either $\lock$ or $\unlock$.
+The former indicates whether the component $\comp$ or the context $\ctx$ is to blame for emitting this event, the latter indicates the secrecy level of values of the emitted event, i.e., either $\lock$ or $\unlock$.
 As for pre-events, the memory-related ones are allocation ($\ev{Alloc}\ \loc\ \valueexpr$), deallocation ($\ev{Dealloc}\ \loc$), reading from ($\ev{Get}\ \loc\ \valueexpr$) and writing to memory ($\ev{Set}\ \loc\ \valueexpr\ \valueexpr[']$).
@@
     \statevar.\store(x)=\tup{\loc}{\sandboxtag,\poison,m} \\
-    \loc+n\in\operatorname{dom}\statevar.\heap[^{\sandboxtag}](\loc+n)
-  }{
-    \pstepto{\rtt{\statevar}{\lget{x}{n}}}{\rtt{\statevar}{\heap[^{\sandboxtag}](\loc+n)}}{\tup{\ev{Get}\ \loc\ n}{\sandboxtag}}
+    \loc+n\in\operatorname{dom}\statevar.\heap[^{\sandboxtag}]
+  }{
+    \pstepto{\rtt{\statevar}{\lget{x}{n}}}{\rtt{\statevar}{\statevar.\heap[^{\sandboxtag}](\loc+n)}}{\tup{\ev{Get}\ \loc\ n}{\sandboxtag}}
   }{e-get-in}
@@
 Hereby, a $\ev{Call\ \ctxtocomp\ foo\ \valueexpr}$ and $\ev{Return\ \ctxtocomp\ \valueexpr}$ signal that program execution transitions from context- to component-level.
-Contrary, $\ev{Call\ \comptoctx\ foo\ \valueexpr}$ and $\ev{Return\ \ctxtocomp\ \valueexpr}$ signal that program execution transitions from component- to context-level.
-For calls without this context switch, the environmental semantics attaches the $\nocomm$ tag.
+Contrary, $\ev{Call\ \comptoctx\ foo\ \valueexpr}$ and $\ev{Return\ \comptoctx\ \valueexpr}$ signal that program execution transitions from component- to context-level.
+For calls without this context switch, the contextual semantics attaches the $\nocomm$ tag.
 In the following, $\neg\ctx=\comp$ and $\neg\comp=\ctx$.
@@
 
-The environmental semantics is mostly straightforward.
-In \Cref{tr:e-ret-notsame,tr:e-call-notsame}, the judgement $\Omega.\commlib\vdash\text{foo}:\Omega.\sandboxtag$ checks whether $\text{foo}$ is a component-level name by looking it up in the list of component-level names $\commlib$ and emits the appropriate transfer tag, i.e., either $\comptoctx$ or $\ctxtocomp$.
+The contextual semantics is mostly straightforward.
+In \Cref{tr:e-ret-notsame,tr:e-call-notsame}, the judgement $\Omega.\commlib\vdash\text{foo}:\Omega.\sandboxtag\dashv \comm$ checks whether $\text{foo}$ is a component-level name by looking it up in the list of component-level names $\commlib$ and emits the appropriate transfer tag, i.e., either $\comptoctx$ or $\ctxtocomp$.
 Additional rules that are left out ensure that, e.g., when calling the $\text{main}$ function, the event $\ev{Start}$ is emitted, which is a design choice this paper does for convenience when reasoning about call-chains.
@@
   }{
-    \msfilterLtms{\src{\tup{\event[_b]}{\ctx,\securitytag}} = \emptyevent}
+    \msfilterLtms{\src{\tup{\event[_b]}{\ctx,\securitytag}}} = \emptyevent
   }{ltms-filter-context}
@@
 For valid pointers ($\trg{\poisonless}$) bound to variable $\trg{x}$, the check $\trg{\lispoisoned{x}}$ yields $\trg{1}$.
-If the array bound to $\trg{x}$ was allocated, i.e., has been poisoned ($\trg{\poisoned}$), the check $\trg{\lispoisoned{x}}$ evaluates to $\trg{0}$.
+If the array bound to $\trg{x}$ was deallocated, i.e., has been poisoned ($\trg{\poisoned}$), the check $\trg{\lispoisoned{x}}$ evaluates to $\trg{0}$.
 
@@
   \begin{aligned}
-    \mi{(Expressions)}~\obj{\expr}&\bnfdef \obj{n^{\securitytag}} \mid \cdots \mid \obj{\llet{x^{\securitytag}}{\expr[_{1}]}{\expr[_{2}]}} \mid \cdots \mid \obj{\lwrdoit{\expr}} \mid \obj{\lrddoit{x}{\expr}}
+    \mi{(Expressions)}~\obj{\expr}&\bnfdef \obj{n^{\securitytag}} \mid \cdots \mid \obj{\llet{x^{\securitytag}}{\expr[_{1}]}{\expr[_{2}]}} \mid \cdots \mid \obj{\lwrdoit{D}} \mid \obj{\lrddoit{x}{\expr}}
   \end{aligned} 
@@
   \begin{aligned}
-    \mi{(States)}~\obj{\statevar}&\bnfdef \obj{\tup{\cfstate}{\sandboxtag,n,\memstate}}
+    \mi{(States)}~\obj{\statevar}&\bnfdef \obj{\tup{\cfstate}{\sandboxtag,D,\memstate}}
+     &
+    \mi{(DIT\ Flag)}~\obj{D}&\bnfdef \obj{ON} \mid \obj{OFF}
   \end{aligned}
@@
   %
-  \typerule{$e-\obj{wrdoit}$}{
-  }{
-    \pstepto{\rtt{\obj{\cfstate;\sandboxtag;m;\memstate}}{\obj{\lwrdoit{n^{\securitytag}}}}}{\rtt{\obj{\cfstate;\sandboxtag;n;\memstate}}{\obj{n^{\securitytag}}}}{\obj{\emptyevent}}
-  }{e-wrdoit}
+  \typerule{$e-\obj{wrdoit}-\text{off}$}{
+  }{
+    \pstepto{\rtt{\obj{\cfstate;\sandboxtag;m;\memstate}}{\obj{\lwrdoit{OFF}}}}{\rtt{\obj{\cfstate;\sandboxtag;OFF;\memstate}}{\obj{0^{\securitytag}}}}{\obj{\emptyevent}}
+  }{e-wrdoit-off}
   %
@@
 The evaluation steps are amended to propagate the security-tag annotations $\obj{\securitytag}$.
-When the data independent timing mode is active, pre-events $\obj{\ev{Branch}\ n}$ and $\obj{\ev{Binop}\ n}$ are emitted for conditionals and binary operations, respectively.
+When the data independent timing mode is inactive, pre-events $\obj{\ev{Branch}\ n}$ and $\obj{\ev{Binop}\ n}$ are emitted for conditionals and binary operations, respectively.
 %To mark this mode as active or inactive, \Cref{tr:e-wrdoit} allows to access the register in $\obj{\statevar}$.
@@
     Vertices in the graph are the programming languages from earlier sections (\Cref{sec:casestud:defs}). %
-    All edges are secure compilers, but dotted edges use the presented framework (\Cref{sec:compcomp}) and strikethrough edges classic proof techniques. %
+    All edges are secure compilers, but dotted edges use the presented framework (\Cref{sec:compcomp}) and solid edges classic proof techniques. %
     The dashed lines partition the graph into the sections where the respective theorems are presented.
@@
   \begin{align*}
-    \ccscct(\irl{\lfunction{g}{x}{\expr}}) & = \lfunction[\obj]{\obj{g}}{\obj{x}}{\obj{\lwrdoit{1};}\ccscct(\irl{\expr})} \\
-    \ccscct(\irl{\lcall{g}{\expr}}) & = \lcall[\obj]{\obj{g}}{\ccscct(\irl{\expr})\obj{; wrdoit{1}}} \\
+    \ccscct(\irl{\lfunction{g}{x}{\expr}}) & = \lfunction[\obj]{\obj{g}}{\obj{x}}{\obj{\lwrdoit{ON};}\ccscct(\irl{\expr})} \\
+    \ccscct(\irl{\lcall{g}{\expr}}) & = \lcall[\obj]{\obj{g}}{\ccscct(\irl{\expr})\obj{; \lwrdoit{ON}}} \\
     \ccscct(\irl{\lbinop{\expr[_{1}]}{\expr[_{2}]}}) & = \lbinop[\obj]{\ccscct{\irl{\expr[_{1}]}}}{\ccscct{\irl{\expr[_{2}]}}} \\

### 2023-09-18 csf/main.tex
[BROKEN] try to migrate to csf template

--- 2023-09-15 popl/main.tex
@@
 \newcommand\bmmax{0}
-\documentclass[utf8,acmsmall,review,screen,dvipsnames,anonymous]{acmart}
+\documentclass[utf8,dvipsnames,conference,peerreviewca]{IEEEtran}
 
@@
 \usepackage{glossaries}
+\usepackage{orcidlink}
 \usepackage{cleveref}
@@
 
-%% NOTE that a single column version may be required for 
-%% submission and peer review. This can be done by changing
-%% the \doucmentclass[...]{acmart} in this template to 
-%% \documentclass[manuscript,screen]{acmart}
-%% 
-%% To ensure 100% compatibility, please check the white list of
-%% approved LaTeX packages to be used with the Master Article Template at
-%% https://www.acm.org/publications/taps/whitelist-of-latex-packages 
-%% before creating your document. The white list page provides 
-%% information on how to submit additional LaTeX packages for 
-%% review and adoption.
-%% Fonts used in the template cannot be substituted; margin 
-%% adjustments are not allowed.
-%%
-%%
-%% \BibTeX command to typeset BibTeX logo in the docs
 \AtBeginDocument{%
@@
     \normalfont B\kern-0.5em{\scshape i\kern-0.25em b}\kern-0.8em\TeX}}}
-
-%% Rights management information.  This information is sent to you
-%% when you complete the rights form.  These commands have SAMPLE
-%% values in them; it is your responsibility as an author to replace
-%% the commands and values with those provided to you when you
-%% complete the rights form.
-\setcopyright{acmcopyright}
-\copyrightyear{2024}
-\acmYear{2024}
-\acmDOI{XXXXXXX.XXXXXXX}
-
-%% These commands are for a PROCEEDINGS abstract or paper.
-\acmConference[POPL '24]{51st ACM SIGPLAN Symposium on Principles of Programming Languages}{January 17-19,
-  2024}{London, UK}
-%
-%  Uncomment \acmBooktitle if th title of the proceedings is different
-%  from ``Proceedings of ...''!
-%
-%\acmBooktitle{Woodstock '18: ACM Symposium on Neural Gaze Detection,
-%  June 03-05, 2018, Woodstock, NY}
-\acmPrice{15.00}
-\acmISBN{978-1-4503-XXXX-X/18/06}
-
-%%
-%% Submission ID.
-%% Use this when submitting an article to a sponsored event. You'll
-%% receive a unique submission ID from the organizers
-%% of the event, and this ID should be used as the parameter to this command.
-%%\acmSubmissionID{123-A56-BU3}
-
-%%
-%% For managing citations, it is recommended to use bibliography
-%% files in BibTeX format.
-%%
-%% You can then either use BibTeX with the ACM-Reference-Format style,
-%% or BibLaTeX with the acmnumeric or acmauthoryear sytles, that include
-%% support for advanced citation of software artefact from the
-%% biblatex-software package, also separately available on CTAN.
-%%
-%% Look at the sample-*-biblatex.tex files for templates showcasing
-%% the biblatex styles.
-%%
-
-%%
-%% The majority of ACM publications use numbered citations and
-%% references.  The command \citestyle{authoryear} switches to the
-%% "author year" style.
-%%
-%% If you are preparing content for an event
-%% sponsored by ACM SIGGRAPH, you must use the "author year" style of
-%% citations and references.
-%% Uncommenting
-%% the next command will enable that style.
-\citestyle{acmauthoryear}
-
-%%
-%% end of the preamble, start of the body of the document source.
+\makeatletter
+\newcommand{\linebreakand}{%
+  \end{@IEEEauthorhalign}
+  \hfill\mbox{}\par
+  \mbox{}\hfill\begin{@IEEEauthorhalign}
+}
+\makeatother
+
 \begin{document}
 
-%%
-%% The "title" command has an optional parameter,
-%% allowing the author to define a "short title" to be used in page headers.
 \title{Secure Composition of Robust and Optimising Compilers}
-% Composing Secure and Optimising Compilers
-%
-
-%%
-%% The "author" command and its associated commands are used to define
-%% the authors and their affiliations.
-%% Of note is the shared affiliation of the first two authors, and the
-%% "authornote" and "authornotemark" commands
-%% used to denote shared contribution to the research.
-\author{Matthis Kruse}
-% \authornote{Both authors contributed equally to this research.}
-\email{matthis.kruse@cispa.de}
-\orcid{0000-0003-4062-9666}
-\affiliation{%
-  \institution{CISPA Helmholtz Center for Information Security and Saarland University}
-  \streetaddress{Stuhlsatzenhaus 5}
-  \city{Saarbr{\"u}cken}
-  \state{Saarland}
-  \country{Germany}
-  \postcode{66123}
+
+\author{\IEEEauthorblockN{1\textsuperscript{st} Matthis Kruse\orcidlink{000-003-4062-9666}} \\
+\IEEEauthorblockA{\textit{CISPA Helmholtz Center for Information Security\\
+and Saarland Informatics Campus \\
+Saarbr\"ucken, Germany \\
+matthis.kruse@cispa.de}}%
+\and
+\IEEEauthorblockN{2\textsuperscript{nd} Michael Backes} \\
+\IEEEauthorblockA{\textit{CISPA Helmholtz Center for Information Security\\
+Saarbr\"ucken, Germany \\
+director@cispa.de}}%
+\linebreakand
+\IEEEauthorblockN{3\textsuperscript{rd} Marco Patrignani\orcidlink{000-003-3411-9678}} \\
+\IEEEauthorblockA{\textit{Trento University\\
+Trento, Italy \\
+marco.patrignani@unitn.it}}%
 }
 
-\author{Michael Backes}
-\email{director@cispa.de}
-%\orcid{0000-0002-7130-9211}
-\affiliation{%
-  \institution{CISPA Helmholtz Center for Information Security}
-  \streetaddress{Stuhlsatzenhaus 5}
-  \city{Saarbr{\"u}cken}
-  \state{Saarland}
-  \country{Germany}
-  \postcode{66123}
-}
-
-\author{Marco Patrignani}
-\orcid{0000-0003-3411-9678}
-\email{marco.patrignani@unitn.it}
-\affiliation{%
-  \institution{University of Trento}
-  \streetaddress{Via Sommarive, 9}
-  \city{Povo}
-  \country{Italy}
-  \postcode{38123}
-}
-
-%%
-%% By default, the full list of authors will be used in the page
-%% headers. Often, this list is too long, and will overlap
-%% other information printed in the page headers. This command allows
-%% the author to define a more concise list
-%% of authors' names for this purpose.
-\renewcommand{\shortauthors}{Kruse, Backes, and Patrignani}
-
-%%
-%% The abstract is a short summary of the work to be presented in the
-%% article.
+\maketitle
+\IEEEpeerreviewmaketitle
+
 \begin{abstract}
@@
 
-%%
-%% The code below is generated by the tool at http://dl.acm.org/ccs.cfm.
-%% Please copy and paste the code instead of the example below.
-%%
-\begin{CCSXML}
-<ccs2012>
-  <concept>
-  <concept_id>10002978.10002986.10002989</concept_id>
-  <concept_desc>Security and privacy~Formal security models</concept_desc>
-  <concept_significance>500</concept_significance>
-  </concept>
-</ccs2012>
-\end{CCSXML}
-\ccsdesc[500]{Security and privacy~Formal security models}
-
-%%
-%% Keywords. The author(s) should pick words that accurately describe
-%% the work being presented. Separate the keywords with commas.
-\keywords{Memory-safety, Secure Compilation, Privacy}
-
-%%
-%% This command processes the author and affiliation and title
-%% information and builds the first part of the formatted document.
-\maketitle
+\begin{IEEEkeywords}
+  memory-safety, secure compilation, privacy
+\end{IEEEkeywords}
 
@@
 the memory location past the last element will be leaked to an attacker.
-\begin{lstlisting}[language=c,basicstyle=\small\ttfamily,morekeywords={size_t}]
+\begin{lstlisting}[language=c,basicstyle=\scriptsize\ttfamily,morekeywords={size_t}]
 void strncpy(size_t n, char *dst, char *src) {

### 2023-09-19 csf/main.tex
Make CSF compile again, but remove all content for now

@@
 \newcommand\bmmax{0}
-\documentclass[utf8,dvipsnames,conference,peerreviewca]{IEEEtran}
+\documentclass[dvipsnames,conference]{IEEEtran}
 
@@
 \usepackage[inference]{semantic}
+\usepackage[switch]{lineno}
 \usepackage{fontawesome5}
@@
 
+\renewcommand\thelinenumber{\color{red}\arabic{linenumber}}
 \begin{document}
+\linenumbers
 
@@
 
-\author{\IEEEauthorblockN{1\textsuperscript{st} Matthis Kruse\orcidlink{000-003-4062-9666}} \\
-\IEEEauthorblockA{\textit{CISPA Helmholtz Center for Information Security\\
-and Saarland Informatics Campus \\
+\author{\IEEEauthorblockN{1\textsuperscript{st} Matthis Kruse\orcidlink{0000-0003-4062-9666}}
+\IEEEauthorblockA{\textit{CISPA and Saarland University} \\
 Saarbr\"ucken, Germany \\
-matthis.kruse@cispa.de}}%
+matthis.kruse@cispa.de}
 \and
-\IEEEauthorblockN{2\textsuperscript{nd} Michael Backes} \\
-\IEEEauthorblockA{\textit{CISPA Helmholtz Center for Information Security\\
+\IEEEauthorblockN{2\textsuperscript{nd} Michael Backes}
+\IEEEauthorblockA{\textit{CISPA} \\
 Saarbr\"ucken, Germany \\
-director@cispa.de}}%
-\linebreakand
-\IEEEauthorblockN{3\textsuperscript{rd} Marco Patrignani\orcidlink{000-003-3411-9678}} \\
-\IEEEauthorblockA{\textit{Trento University\\
+director@cispa.de}
+\and
+\IEEEauthorblockN{3\textsuperscript{rd} Marco Patrignani\orcidlink{0000-0003-3411-9678}}
+\IEEEauthorblockA{\textit{Trento University} \\
 Trento, Italy \\
-marco.patrignani@unitn.it}}%
+marco.patrignani@unitn.it}%
 }
@@
 \maketitle
-\IEEEpeerreviewmaketitle
 
@@
 
-% \gls*{ms} is a security property that ensures that there are no out-of-bounds accesses as well as no reads or writes to uninitialised memory.
-% Programs that do not attain memory safety, can easily have their private data be leaked to attackers or be corrupted~\cite{lemay2021ccc}.
-% Because of this, \gls*{ms} is a requirement for programs to attain an even stronger security property, namely \gls*{cct}.
-% Programs that attain \gls*{cct} are secure with respect to timing attacks, and this is the gold standard for secure applications, including cryptographic ones, since timing attacks can expose secrets via a simple performance analysis~\cite{kocher1996timing}.
-% Timing attacks have become more and more relevant recently, especially given the era of cloud-computing~\cite{aviram2010cloudtime,kumar2019cloudsecsurvey}, where programs share the same hardware but are supposed to run in a sandboxed environment that does not leak secrets.
-% Unfortunately, several attacks within such a cloud environment have been demonstrated as viable strategies to extract secrets~\cite{mehmet2015getoff,flowers2022zeroday,atya2019catchme,venkatanathan2015placevul}.
+\cite{patrignani2020use}
 
-% Even though mitigations for such attacks exist~\cite{bond2017vale,almeida2017jasmin}, some compilers may optimise these mitigations away~\cite{barthe2018sec}.
-% This is of course not what programmers expect and to ensure such mitigations persist in the compiled code, one can use a {\em secure} compilers.
-% More generally, secure compilers should preserve security properties even when the program runs in a hostile context such as the cloud one, i.e., where it interacts with other possibly malicious programs.
-% An example of such secure compilers are those that ascribe to the theory of robust compilation~\cite{abate2019jour,abate2021extacc,patrignani2021rsc}, which takes this interaction between components and potentially malicious contexts into account.
+\section*{Acknowledgements}\label{sec:acks}
 
-% Unfortunately, it is unclear if and how robust compilation supports compositionality, and compilation passes are typically developed as a standalone module to be plugged into a compilation pipeline and composed with other passes.
-% Different kinds of compilation passes exist, e.g., there are usually several compilation passes that perform optimisations~\cite{lattner2004llvm,googlev8,androidstudio,kuepper2023cryptopt,manjikian1997fusion,wegman1991ccp} or source-code instrumentations~\cite{nagarakatte2009soft,nagarakatte2010cets,akritidis2009baggy,dhumbumroong2020boundwarden,jung2021pico,nam2019framer,shankaranarayana2023tailcheck,younan2010paricheck,zhou2023fatptrs,bond2017vale,almeida2017jasmin,kuepper2023cryptopt,cauligi2019fact} that enforce that the compiled program fulfils a property of interest.
-% Without a framework for composing secure compiler passes, it is not possible to enable separation of concerns, e.g., to have a secure compilation pass that ensures \gls*{ms} that is developed independently of another secure pass for \gls*{cct} as suggested before and as showcased in \Cref{ex:strncpy}.
-% \MPpost{
-%   this intro does a poor job at introducing the robust compilation argument.
-%   perhaps it'd be best to start from robust properties, which motivates the use of robust compilers?
-%   idk
-%   as for the abstract, this is a good starting point for you to make a better intro
-% }
-
-\gls*{ms} is a security property obtained by composing \gls*{sms}, which ensures array accesses are all within bounds, and \gls*{tms}, which ensures pointers are only used when they are valid~\cite{azevedo2018meaningsofms,jim2002cyclone,necula2005ccured,nagarakatte2010cets,nagarakatte2009soft,akritidis2009baggy,michael2023mswasm}.
-\gls*{cct} is a security property that ensures sensitive data is not leaked via timing side-channels~\cite{kocher1996timing}.
-Together, \gls*{sms}, \gls*{tms} and \gls*{scct}, an enforceable overapproximation of \gls*{cct}, yield \gls*{msscct}, which is the gold standard of security properties for secure applications.
-Programs attaining \gls*{msscct} do not leak sensitive data either through erroneous memory accesses, nor through timing side-channels.
-As discussed in \Cref{ex:strncpy}, these security properties can be enforced by compiler passes~\cite{bond2017vale,almeida2017jasmin}, to ensure programmers need not be aware of the architectural details of where their code will run.
-
-\begin{example}[strncpy]\label{ex:strncpy}
-Consider the \texttt{C} function \texttt{strncpy} that copies a null-terminated string \texttt{src} into \texttt{dst} up to a length of \texttt{n}.
-% Assume that the length of the array pointed by \texttt{dst} is at least \texttt{n}, and \texttt{src} is terminated by \texttt{'\textbackslash 0'}.
-% Even under these assumptions, there is a 
-This function is subject to a subtle \gls*{sms} vulnerability: the bounds check \texttt{i < n} should happen {\it before} the access to memory location \texttt{x[i]}: otherwise
-% If \texttt{n} is at least as large as the length of the array \texttt{src}, 
-the memory location past the last element will be leaked to an attacker.
-\begin{lstlisting}[language=c,basicstyle=\scriptsize\ttfamily,morekeywords={size_t}]
-void strncpy(size_t n, char *dst, char *src) {
-  for(size_t i = 0; src[i] != '\0' && i < n; ++i) {
-    dst[i] = src[i];
-  }
-}
-\end{lstlisting}
-
-To prevent this vulnerability, one can use a compilation pass that enforces \gls*{sms}, such as Softbounds~\cite{nagarakatte2009soft} or BaggyBounds~\cite{akritidis2009baggy}.
-The most naive solution in this case is to insert bounds-checks in front of every access to memory.
-
-Because of timing attacks, fixing \gls*{sms} is not enough to make \texttt{strncpy} secure.
-In fact, the loop can terminate early, as soon as the string-terminating character \texttt{'\textbackslash 0'} is encountered, thus making program execution time proportional to the length of the array pointed by \texttt{src}.
-% Often, programmers rewrite programs like these to have loops that run with a fixed number of iterations.
-Also in this case there exist compiler passes that can rewrite such programs into \gls*{cct} ones~\cite{cauligi2019fact}.
- % so that they are no longer vulnerable to timing attacks.
-
-Alas, code is not run in isolation, so a malicious attacker could supply code that interacts with \texttt{strncpy} and trigger a violation of either \gls*{ms} or \gls*{cct} by calling \texttt{strncpy} with an argument for \texttt{src} that points to uninitialised memory.
-This would, in turn, triggering a series of reads from uninitialised memory, which is an immediate \gls*{ms} violation with devastating real-world consequences~\cite{uninit-0,uninit-1,uninit-2,uninit-3,uninit-4}.
-% An easy fix to sidestep this is to use non-nullable pointers as parameters instead, e.g., as presented in Cyclone~\cite{jim2002cyclone}.
-
-Robust compilers~\cite{abate2019jour} are a form of secure compilers that preserve security properties even in the presence of arbitrary attackers interacting with compiled code.
-Thus, robust compilers can be used to prevent vulnerabilities resulting from uninitialised memory (as well as many other ones), e.g., by targeting capability-based languages such as CHERI~\cite{woodruff2014CHERI}, Arm Morello~\cite{arm-morello}, or MSWasm~\cite{michael2023mswasm}, where the compiler relies on capabilities to check that pointers are always initialised.
-\end{example}
-
-% To ensure these compiler-preserved properties still hold even in the presence of arbitrary attackers interacting with compiled code, one can use secure compilers in the form of robust compilers~\cite{abate2019jour,abate2021extacc,patrignani2021rsc}.
-
-Unfortunately, given secure compiler passes that each preserve a possibly different security property, there is no way to tell what kind of security property will the composition of those secure compilers preserve.
-% Different kinds of compilation passes exist, e.g., there are usually several compilation passes that perform optimisations~\cite{lattner2004llvm,googlev8,androidstudio,kuepper2023cryptopt,manjikian1997fusion,wegman1991ccp} or source-code instrumentations~\cite{nagarakatte2009soft,nagarakatte2010cets,akritidis2009baggy,dhumbumroong2020boundwarden,jung2021pico,nam2019framer,shankaranarayana2023tailcheck,younan2010paricheck,zhou2023fatptrs,bond2017vale,almeida2017jasmin,kuepper2023cryptopt,cauligi2019fact} that enforce that the compiled program fulfils a property of interest.
-Worse, without a framework for composing secure compiler passes, it is not possible to enable separation of concerns, e.g., to have a secure compilation pass that ensures \gls*{ms} that is developed independently of another secure pass for \gls*{cct}, that is developed independently of other passes, such as optimisations.
-
-This paper introduces a framework for reasoning about the composition of secure and optimising compiler passes akin to those of \Cref{ex:strncpy} and it showcases the power of this framework by instantiating it on a multi-pass compilation chain.
-To this end, this paper first discusses how to compose security properties, such as \gls*{tms} and \gls*{sms} into \gls*{ms}, and then adding \gls*{scct} to the mix to obtain \gls*{msscct}.
-Then, this paper defines compiler composition and formalises that given two passes that securely preserve two (possibly distinct) properties, their composition securely preserves the composition of those properties.
-The paper then defines several secure compiler passes, where each is either preserving a different security property (\gls*{tms}, \gls*{sms}, \gls*{scct}) or performing a security-preserving optimisation, (e.g., applying \gls*{cf} or \gls*{dce}).
-Finally, this paper shows that composing these secure compiler passes into a multi-pass compilation chain results in the end-to-end preservation of \gls*{msscct}.
-Crucially, this paper derives the security of the multi-pass compiler from the composition of the security properties preserved by its individual passes.
-% Crucially, this end-to-end security preservation is obtained by composing the security property preserved by each individual pass, as dictated by the framework of this paper.
-This result showcases how the framework allows the kind of formal security reasoning that compiler writers already want (and already do), obtaining precise, compositional security reasoning while providing minimal (and modular) proof effort.
-
-In summary, this paper makes the following contributions:
-\begin{itemize}
-  \item %
-        %\Cref{sec:compprop} presents a formal framework to reason about compositions of security properties and demonstrate its usage with concrete examples:
-        This paper formalises security properties (\Cref{sec:compprop}) that are of interest for real-world compiler writers, namely \gls*{tms}, \gls*{sms} and \gls*{cct} (as identified by the plethora of work enforcing such properties individually~\cite{akritidis2009baggy,nagarakatte2009soft,nagarakatte2010cets,dhumbumroong2020boundwarden,jung2021pico,nam2019framer,shankaranarayana2023tailcheck,younan2010paricheck,zhou2023fatptrs,bond2017vale,almeida2017jasmin,kuepper2023cryptopt,cauligi2019fact}).
-        % 
-        Starting from ways to formalise those properties individually, this paper shows how to compose their formalisation.
-        % 
-        % This paper then showcases the benefits of this composition by composing \gls*{tms} and \gls*{sms} into \gls*{ms}, and ultimately adding \gls*{cct} to the composition, yielding the the conjunction of \gls*{ms} and \gls*{cct}.
-        % 
-        The resulting security property is \gls*{msscct}, i.e., the gold standard of security properties for secure programs~\cite{lemay2021ccc}.
-
-        % \MPpast{
-        % 	some of the good things here: 
-        % 	our definitions are amenable to sec comp proofs, unlike existing definitions
-        % 	MS def is like mswasm MS def, 
-        % 	but CCT is new
-        % }
-
-  \item %
-        %\Cref{sec:compcomp} studies different forms of secure compiler compositions.
-        This paper takes the secure compilation framework of~\citep{abate2019jour} and extends it to reason about the security of all different known forms of compiler composition (\Cref{sec:compcomp}).
-        %Concretely, it investigates upper and lower compositions, as well as the sequential composition.
-        For this, this paper studies sequential compiler composition as well as compilers with multiple input languages or multiple output ones, as used in existing compilation chains.
-        %
-        This paper proves that starting from two compilers that preserve two (possibly distinct) properties, their composition preserves the intersection of those properties.
-        % For example, a compiler that is secure with respect to \gls*{tms} can be composed with a compiler that is secure with respect to \gls*{sms}, resulting in a compiler that is secure with respect to \gls*{ms}.
-        % 
-        Finally, this paper proves that the order of composition of sequential compiler passes is irrelevant for the resulting security.
-        % 
-        This is crucial for reordering optimisation passes and thus generating secure and efficient code.
-
-  \item %
-        This paper presents a case-study showcasing the conjunction of the previous contributions (\Cref{sec:casestud:defs,sec:casestud:rtp}).
-        To this end, it presents a compilation chain consisting of several passes that ultimately preserves \gls*{msscct} by means of composing the individual, secure passes concerning \gls*{tms}, \gls*{sms}, and \gls*{scct}, respectively.
-        Furthermore, the chain includes two optimisation passes: One performs \gls*{dce} and the other \gls*{cf}.
-        The formalisation of this case study showcases the power of the presented framework: The divide-and-conquer approach to software engineering is a viable strategy even for the development of secure compilers.
-%        This demonstrates that the usage of the presented framework is both simpler and more expressive compared to one monolithic security proof of the whole compilation chain.
-
-  \item The key contributions of this paper are formalised in the Coq proof assistant and the paper indicates this with \CoqSymbol.
-\end{itemize}
-
-This paper starts by introducing relevant notions of security properties and secure compilation (\Cref{sec:background}),
-and discusses related work (\Cref{sec:relwork}) before concluding (\Cref{sec:concl}).
-
-\contribution{Open Source \& Technical Report} A technical report with the omitted formal details, lemmas and proofs, as well as the Coq formalisation are available as supplementary material.
-% \MKin{cite/link}
-
-\section{Background: Security Properties and Secure Compilers\pages{1}}\label{sec:background}
-
-To introduce the security argument of this paper, this section first presents the concepts of (security) properties, of their satisfaction, and of their robust satisfaction (i.e., satisfaction in the presence of an active attacker; \Cref{subsec:bg:tprop}).
-Then, borrowing from existing work~\cite{abate2019jour,abate2021extacc}, the section introduces secure compilers as compilers that preserve robust property satisfaction (\Cref{subsec:bg:rtp}).
-
-\subsection{Properties and (Robust) Satisfaction}\label{subsec:bg:tprop}
-
-This paper employs the security model where programs are written in a language whose semantics emits events $\event$.
-Events include security-relevant actions (e.g., reading from and writing to memory, as detailed in \Cref{sec:compprop}) and the unobservable event $\emptyevent$.
-As programs execute, their emitted events are concatenated in traces $\trace$, which serve as the description of the behaviour of a program.%
-\footnote{
-Throughout the paper, sequences are indicated with an overbar (i.e., $\trace$), empty sequences with $\hole{\cdot}$, and concatenation of sequences $\trace[_{1}],\trace[_{2}]$ as $\trace[_{1}]\cdot\trace[_{2}]$.
-Prepending elements to sequences uses the same notation: $\event\cdot\trace$.
-}
-
-Properties $\pi$ are sets of traces of admissible program behaviours, ascribing what said property considers valid.
-The set of all properties can be partitioned into different {\em classes} ($\class$), i.e., safety, liveness, and neither safety nor liveness~\cite{clarkson2008hyper}.
-A class is simply a set of properties and for the class of safety properties, it is decidable whether a trace satisfies a safety property with just a finite trace prefix.
-As an example, consider a trace describing an interaction with a memory where the deallocation of an address $\loc$ precedes a read at that address in memory: $\ev{Dealloc\ \loc}\cdot\ev{Read\ \loc\ 1729}\cdot\dots$.
-This program behaviour is insecure with respect to a canonical notion of (temporal) memory safety dictating no use-after-frees of pointers~\cite{nagarakatte2010cets,azevedo2018meaningsofms}, because it reads from a memory location that was freed already.
-The previous finite trace prefix is enough to decide that the trace does not satisfy \gls*{tms} and there is no way to append events to this prefix which would result in the trace being admissible.
-In the following, the execution of a whole program $\wholeprogvar$ that terminates in state $\runtimetermvar$ according to the language semantics and produces trace $\trace$ is written as $\progstepto{\wholeprogvar}{\runtimetermvar}{\trace}$.
-With this, property satisfaction is defined as follows:
-\bul{whole programs $\wholeprogvar$ satisfy a property $\pi$} iff \iul{$\wholeprogvar$ yields a trace $\trace$} such that \oul{$\trace$ satisifies $\pi$} (\Cref{def:propsat}).
-
-\begin{definition}[Property Satisfaction]\label{def:propsat}
-  \bul{$\sat{\progvar}{\pi}$}
-  $\isdef$
-  if \iul{$\progstepto{\progvar}{\runtimetermvar}{\trace}$},
-  then \oul{$\trace\in\pi$}.
-\end{definition}
-
-Property satisfaction is defined on whole programs, i.e., programs without missing definitions.
-Thus, from a security perspective, this considers only a passive attacker model, where the attacker observes the execution and, e.g., retrieves secrets from that.
-To consider a stronger model similarly to what existing work does~\cite{abate2019jour,abate2021extacc,maffeis2008code-carrying,gordon2003authenticity,fournet2007authorization,bengtson2011refine,backes2014uniontyps,michael2023mswasm,swasey2017robust,sammler2019benefits}, the concept of satisfaction can be extended with {\em robustness}.
-Robust satisfaction considers partial programs $\progvar$, i.e., components with missing imports, which cannot run until said imports are fulfilled.
-To remedy this, {\em linking} takes two partial programs $\progvar[_{1}],\progvar[_{2}]$ and produces a whole program $\wholeprogvar$, i.e., $\link{\progvar[_{1}]}{\progvar[_{2}]}=\wholeprogvar$.
-As typically done in works that consider the execution of partial programs~\cite{abate2019jour,devriese2018parametricity,patrignani2021rsc,korashy2021capableptrs,strydonck2019lincap,devriese2017modular,bowman2015noninterference,ahmed2011equivcps,patterson2017linkingtyps},
-this paper assumes that whole programs are the result of linking partial programs referred to as {\em context} ($\ctx$) and {\em component} ($\comp$).
-The context is an arbitrary program and thus has the role of an {\em attacker} that can interact with the component by means of whatever features the programming language has, and the component is what is security-relevant.
-%In this work, the semantics of the programming language is expected to differentiate between the component and the context.
-With this, \Thmref{def:propsat} can be extended as follows: for \bul{components $\progvar$ to robustly satisfy a property $\pi$}, take an \iul{attacker context $\contextvar$ and link it with $\progvar$}, \oul{the resulting whole program must satisfy $\pi$}.
-
-\begin{definition}[Robust Satisfaction]\label{def:proprsat}
-  \bul{$\rsat{\progvar}{\pi}$}
-  $\isdef$ \iul{$\forall \contextvar$, if $\link{\contextvar}{\progvar}=\wholeprogvar$}, then \oul{$\sat{\wholeprogvar}{\pi}$}.
-\end{definition}
-
-\begin{example}[Double Free in Bluetooth Subsystem]
-  Consider \texttt{CVE-2021-3564}~\cite{doublefree-bluetooth}, one of many submissions for a double-free vulnerability.
-  The vulnerability arises due to a race condition where the context-level function \texttt{hci\_cmd\_work} was not expected to behave maliciously, since it resides in the same source-code repository where the vulnerability occurs.
-  Nevertheless, the component-level code of \texttt{hci\_dev\_do\_open} is linked with \texttt{hci\_cmd\_work} and does not atomically check whether a pointer has been freed already:
-  Therefore, \texttt{hci\_dev\_do\_open} does not satisfy the no-double-frees property robustly, since there is an implementation for \texttt{hci\_cmd\_work} that leads to a violation of that property when linked with \texttt{hci\_dev\_do\_open}.
-\end{example}
-
-\subsection{Secure Compilers}\label{subsec:bg:rtp}
-
-A {\em compiler} ($\cc{\src{L}}{\trg{L}}$) translates syntactic descriptions of programs from a {\em source} ($\src{L}$) into a {\em target} ($\trg{L}$) programming language.
-This translation is considered {\em correct} if it is semantics-preserving.
-That is, for a whole program $\src{\wholeprogvar}$, the compiler should relate the $\src{L}$ semantics of $\src{\wholeprogvar}$ with the semantics of $\trg{L}$ of the compiled counterpart of $\src{p}$ in such a way that they are ''compatible''.
-Unfortunately, correct compilers may be insecure compilers~\cite{patrignani2019survey,kennedy2006secure.net,abadi1999protect,ahmed2018dagstuhl} and programs translated by insecure compilers can violate security properties that the programmer assumes to hold.
-To define when a compiler is secure, this paper uses the robust compilation framework~\cite{abate2019jour}, which the following definition summarises.
-
-For \bul{compilers $\cc{\src{L}}{\trg{L}}$ to robustly preserve a class of properties $\class$}, if for any \rul{property $\pi$ of that class $\class$ and programs $\src{p}$ written in $\src{L}$} where \iul{$\src{\progvar}$ robustly satisfies $\pi$}, then \oul{the compilation of $\src{\progvar}$, $\cc{\src{L}}{\trg{L}}\left(\src{p}\right)$, must robustly satisfy $\pi$}.
-
-\begin{definition}[Robust Compilation]\label{def:rtp}
-  % Compiler $\cc{\src{L}}{\trg{L}}$ robustly preserves $\class$, 
-  \bul{$\rtp{\cc{\src{L}}{\trg{L}}}{\class}$}
-  %, iff 
-  $\isdef$
-  \rul{$\forall \left(\pi\in\class\right) \left(\src{p}\in\src{L}\right),$} if \iul{$\rsat{\src{\progvar}}{\pi}$}, then \oul{$\rsat{\cc{\src{L}}{\trg{L}}\left(\src{p}\right)}{\pi}$}.
-\end{definition}
-
-Note that a class of properties $\class$ can represent just one property $\pi$ by lifting~\cite{clarkson2008hyper} that property to sets of properties, i.e., use the powerset of $\pi$ instead of $\pi$ itself.
-Because of this, this paper writes $\rtp{\cc{\src{L}}{\trg{L}}}{\pi}$, even though $\pi$ is a property and not a class.
-
-\begin{example}[Types]
-  Suppose $\src{L}$ is a statically-typed language similar to \texttt{C} and $\trg{T}$ is dynamically typed, where both share the same syntax up to dynamic type checks.
-  Consider the following $\src{L}$ component and its compiled version below.
-  \begin{center}
-  $
-  \begin{array}{l}
-    \src{\lfunction{foo}{(char* x, int\ n)}{\lifz{valid\_ptr(x, n, sizeof(char))}{x[0]}{-1}}} \\
-    \trg{\lfunction{foo}{(\phantom{char*\ }x, \phantom{int\ }n)}{\lifz{valid\_ptr(x, n, sizeof(char))}{x[0]}{-1}}}
-  \end{array}
-  $
-  \end{center}
-  While the compiler emits code that may look correct, the generated code does not check that the provided argument is of the right type.
-  Even though the pointer $\trg{x}$ is checked for validity, the context $\trg{foo((int*)y, 1)}$ is able to provoke a read out of bounds.
-  Suppose the component transferred control to the context and passed ownership of a $\trg{char}$ pointer $\trg{y}$ sized $\trg{1}$ cells, the context can now call the component again, casting this buffer to an $\trg{int*}$ prior to that call.
-  The pointer is valid for one $\trg{char}$-sized memory cell, as expected, but the actual read operation now returns $\trg{sizeof(int)}$ many bytes instead of just $\trg{sizeof(char)}$ many.
-  Thus, even if $\src{foo}$ may have been robust with respect to the \gls*{sms}, its compiled counterpart is not and therefore the compiler fails to attain \Cref{def:rtp}.
-\end{example}
-
-\section{Security Properties: Formalisation, Enforcement and Composition\pages{1.5}}\label{sec:compprop}
-
-This section introduces a trace model and uses it to define the key properties of interest for this paper: \gls*{tms}, \gls*{sms}, \gls*{ms}, and \gls*{scct} (\Cref{subsec:propdefs}).
-These properties are of practical importance (as mentioned in \Cref{sec:introduction}) and also of interest in the case study (\Cref{sec:casestud:defs,sec:casestud:rtp}) this paper presents later. 
-Lastly, for each of the key properties, this section introduces corresponding monitors (\Cref{subsec:monitors}) that check them.
-
-\subsection{Specification Trace Model}\label{subsec:propdefs}
-
-\begin{gather*}
-  \begin{aligned}
-  \mi{(Security\ Tag)}~\sigma\bnfdef&\ \lock \mid \unlock\hspace{0.5cm}
-  \mi{(Control\ Tag)}~\sandboxtag\bnfdef \ctx\mid\comp\hspace{0.5cm}
-  \mi{(Event)}~\event\bnfdef\ \emptyevent \mid \lightning \mid \event[_{b}];\sandboxtag;\sigma \\
-  \mi{(Pre\text{-}event)}&~\event[_{b}]\bnfdef\ \ev{Alloc\ \loc\ n} \mid \ev{Dealloc\ \loc} \mid \ev{Use\ \loc\ n} \mid \ev{Branch\ n} \mid \ev{Binop\ n} \\
-  \end{aligned}
-\end{gather*}
-
-The specification trace model defines events as either the empty event ($\emptyevent$), a crash ($\lightning$), or as tuples consisting of a pre-event, a control-tag, and a security-tag.
-The purpose of the model is to define key security properties of interest, such as \gls*{ms} or a stricter variant of cryptographic constant time.
-To this end, security-tags indicate whether an event contains sensitive information ($\lock$) or not ($\unlock$), while control-tags state whether the context ($\ctx$) or the component ($\comp$) are responsible for emitting the event.
-The latter is necessary to be able to ignore actions done by a spurious context that, e.g., immediately deallocates a memory location twice, thus violating \gls*{tms}~\cite{nagarakatte2010cets}.
-%Even though this could be used for blame~\cite{patrignani2023blame}, this is not what this paper is concerned about.
-Lastly, pre-events describe the actual kind of event that happened.
-One such kind is the allocation event ($\ev{Alloc\ \loc\ n}$) that fires whenever a program claims $n$ cells of memory and stores them at address $\loc$.
-Dually, deallocation ($\ev{Dealloc\ \loc}$) announces that the object at location $\loc$ is freed.
-These two events alone are enough to provide a partial description of \gls*{tms} by requiring that, e.g., there is only one deallocation event that carries a location $\loc$.
-To be able to express \gls*{sms}, there is also an event to describe reads from and writes to memory ($\ev{Use\ \loc\ n}$).
-Finally, for cryptographic code, there is a general guideline that secrets must not be visible on a trace.
-Moreover, an instruction whose timing is data-dependent must not have a secret as an operand.
-Typical operations with data-dependent timing are branches and certain binary operations, such as division\footnote{This is highly architecture-dependent, but division is an operation that serves as a classic example for a data-dependent timing instruction, e.g., \cite[p.~755]{arm-refman}.}.
-Both operations are also modelled in the specification trace model ($\ev{Branch\ n}$ and $\ev{Binop\ n}$).
-
-\subsubsection{Temporal Memory Safety}
-
-\gls*{tms} is a safety property that describes that an unallocated object must not be used in any way.
-Moreover, the property requires that all allocated objects must be deallocated at some point.
-
-\begin{definition}[\glsfirst*{tms}]\label{def:trace:tmsdef}
-  $$
-  \tmssafe:=\left\{\trace \left| \begin{array}{rcl}
-                              \ev{Alloc\ \loc\ n;\sandboxtag;\securitytag}&\le_{\trace}&\ev{Dealloc\ \loc;\sandboxtag;\securitytag'} \\
-                              \ev{Use\ \loc\ n;\sandboxtag;\securitytag}&\le_{\trace}&\ev{Dealloc\ \loc;\sandboxtag;\securitytag'} \\
-                              \text{if }\ev{Alloc\ \loc\ n;\sandboxtag;\securitytag}\text{ in }\trace&\text{then}&\ev{Dealloc\ \loc;\sandboxtag;\securitytag'}\text{ in }\trace \\
-                              \text{at most one }\ev{Dealloc\ \loc;\sandboxtag;\securitytag}&\text{in}&\trace \\
-                              \text{at most one }\ev{Alloc\ \loc\ n;\sandboxtag;\securitytag}&\text{in}&\trace \\
-                            \end{array}\right.\right\}
-  $$
-\end{definition}
-Hereby, the notation $\event[_{1}]\le_{\trace}\event[_{2}]$ means that if $\event[_{1}]$ is in $\trace$ and if $\event[_{2}]$ is in $\trace$, then $\event[_{1}]$ appears before $\event[_{2}]$.
-
-\subsubsection{Spatial Memory Safety}
-
-\gls*{sms} prohibits out-of-bounds accesses:
-
-\begin{definition}[\glsfirst*{sms}]\label{def:trace:smsdef}
-  $$
-  \smssafe:=\left\{\trace \left|\begin{array}{rcl}
-                                  \text{If }\ev{Alloc\ \loc\ n;\sandboxtag;\securitytag}\le_{\trace}\ev{Use\ \loc\ m;\sandboxtag;\securitytag'}, \text{ then }m<n
-                                \end{array}\right.\right\}
-  $$
-\end{definition}
-
-\subsubsection{Memory Safety}
-
-Full \gls*{ms} (similar to earlier work~\cite{nagarakatte2009soft,nagarakatte2010cets,jim2002cyclone,necula2005ccured,michael2023mswasm}) is then described as the conjunction of \Cref{def:trace:tmsdef,def:trace:smsdef}.
-%%% Catalin's paper doesn't apply, since their definition of MS relies on noninterference to guarantee data isolation.
-%%% our definition does not guarantee that (on purpose, we want to compose it with cct)
-Note, however, that this definition says nothing about memory-safety issues introduced by side-channels, such as speculation.
-
-\begin{definition}[\glsfirst*{ms}]\label{def:trace:msdef}
-  $$
-  \mssafe:=\tmssafe \cap \smssafe
-  $$
-\end{definition}
-
-\subsubsection{Strict Cryptographic Constant Time}
-
-\gls*{cct} is a hypersafety property~\cite{barthe2018sec} and, thus, difficult to check with monitors.
-This is because, intuitively, hypersafety properties can relate multiple execution traces with each other, but monitors work on a single execution.
-To sidestep this issue, this section defines the property \gls*{scct}, a stricter variant of \gls*{cct} that enforces the policy that no secret appears on a trace (inspired by earlier work~\cite{almeida2017jasmin}).
-
-\begin{definition}[\glsfirst*{scct}]\label{def:trace:scctdef}
-  $$
-  \scctsafe:=\left\{\trace \left|\begin{array}{rcl}
-                                   \trace=\hole{\cdot} & \text{ or } & \trace=\event[_{b}];\sandboxtag;\unlock\cdot\trace['] \wedge \trace[']\in\scctsafe
-                                 \end{array}\right.\right\}
-  $$
-\end{definition}
-
-\subsubsection{Memory Safe, Strict Cryptographic Constant Time}
-
-The combination of \gls*{ms} and \gls*{scct} is the intersection of these properties, \gls*{msscct}.
-Since \gls*{ms} and \gls*{scct} are just sets of traces that, intuitively, contain all program behaviors that follow a security policy, the intersection of them contains all program behaviors that follow both security policies, i.e., it entails all program behaviours that are both \gls*{ms} and \gls*{scct}.
-
-\begin{definition}[\glsfirst*{ms} and \gls*{scct}]\label{def:trace:msscctdef}
-  $$
-  \msscctsafe:=\mssafe\cap\scctsafe
-  $$
-\end{definition}
-
-%\section{Monitors and Trace Agreements\pages{1.5}}\label{subsec:monitors}
-\subsection{Monitors\pages{1.5}}\label{subsec:monitors}
-
-Monitors enforce safety properties by accepting or rejecting traces, i.e., if they reject a trace, the trace does not satisfy the property they check.
-Since reasoning on monitors is easier than directly on just traces, this section presents a monitor for each of the previously shown safety properties (\Cref{subsec:propdefs}).
-%\MPpost{
-%	this is fine, but it'd be nice to provide more insights on why it's easier
-%}
-To lessen the burden when proving that a monitor accepts the trace of a program execution, each monitor uses a custom trace model that contains only the relevant information related to the property the monitor checks.
-To go from specification traces $\trace$ to monitor-level traces $\abstrace$, each property $\pi$ has an associated event agreement relation $\traceagree[]{\pi}{\event}{\absevent}$.
-\Cref{fig:stuttereq} shows how the event agreement is lifted to traces.
-\begin{figure}[!ht]
-  \judgbox{\traceagree{\pi}{\trace}{\abstrace}}{,,Specification-level trace $\trace$ agrees with monitor-level trace $\abstrace$ with respect to property $\pi$.''}
-  %
-  \typerule{traceagree-empty}{}{
-    \traceagree{\pi}{\hole{\cdot}}{\hole{\cdot}}
-  }{traceagree-empty}
-  %
-  \typerule{traceagree-ign-L}{
-    \traceagree{\pi}{\trace}{\abstrace}
-  }{
-    \traceagree{\pi}{\emptyevent\cdot\trace}{\abstrace}
-  }{traceagree-ign-L}
-  %
-  \typerule{traceagree-ign-R}{
-    \traceagree{\pi}{\trace}{\abstrace}
-  }{
-    \traceagree{\pi}{\trace}{\absev{\emptyevent}\cdot\abstrace}
-  }{traceagree-ign-R}
-  %
-  \typerule{traceagree-cons}{
-    \traceagree[]{\pi}{\event}{\absevent} &
-    \traceagree{\pi}{\trace}{\abstrace}
-  }{
-    \traceagree{\pi}{\event\cdot\trace}{\absevent\cdot\abstrace}
-  }{traceagree-cons}
-  \caption{Trace-Agreement relation that equates specification-level traces with monitor-level traces.}\label{fig:stuttereq}
-\end{figure}
-The trace agreement is the same for all properties $\pi$ up to the use of the event agreement in \Cref{tr:traceagree-cons}.
-With agreements, this section defines monitor satisfaction for traces and then it proves that monitor satisfaction implies property satisfaction.
-To this end, monitor satisfaction is defined as follows.
-\bul{A specification trace $\trace$ monitor-satisfies property $\pi$} iff there exists a (final) monitor state $\monitor$ and an abstract trace $\abstrace$ such that the \iul{specification trace $\trace$ agrees with abstract trace $\abstrace$} and \oul{the initial monitor\footnote{In this paper, for all monitors, the initial monitor state is denoted as $\emptyset$.} can step to the (final) monitor state $\monitor$ with abstract trace $\abstrace$}.
-
-\begin{definition}[Monitor Satisfaction]\label{def:monsat}
-  \bul{$\monsafe{\trace}{\pi}$} $\isdef$ $\exists \abstrace\ \monitor$, \iul{$\traceagree{\pi}{\trace}{\abstrace}$} and \oul{$\monitorcheck{\emptyset}{\monitor}{\abstrace}$}.
-\end{definition}
-
-\subsubsection{Monitor for \glsfirst*{tms}}\label{subsubsec:tms}
-\begin{gather*}
-  \begin{aligned}
-    \mi{(Abstract\ Store)}~\tmsmonitor\bnfdef&\ \left\{\text{allocated}:L\times\sandboxtag,\text{freed}:L\times\sandboxtag\right\} \hspace{0.33cm}%
-    \emptyset:=\ \left\{\text{allocated}:\emptyset,\text{freed}:\emptyset\right\}\\
-    \mi{(Abstract\ Events)}~\absevent\bnfdef&\ \absev{\emptyevent} \mid \absev{Alloc\ }\loc\ \sandboxtag \mid \absev{Dealloc\ }\loc\ \sandboxtag \mid \absev{Use\ }\loc\ \sandboxtag \mid \absterm\\
-  \end{aligned}
-\end{gather*}
-\begin{center}
-  \judgbox{\monitorcheck[]{\tmsmonitor}{\tmsmonitor[']}{\absevent}}{,,Monitor $\tmsmonitor$ does one step to $\tmsmonitor[']$ given event $\absevent$.''}$\;$\\
-  %
-% \typerule{tms-uninteresting}{
-% }{
-%   \monitorcheck[]{\tmsmonitor}{\tmsmonitor}{\absev{\emptyevent}}
-% }{tms-uninteresting}
-% %
-% \typerule{tms-abort}{
-% }{
-%   \monitorcheck[]{\tmsmonitor}{\tmsmonitor}{\absterm}
-% }{tms-abort}
-  %
-  \typerule{tms-use}{
-    (\loc;\sandboxtag)\in\tmsmonitor.\text{allocated} &
-    (\loc;\sandboxtag)\notin\tmsmonitor.\text{freed}
-  }{
-    \monitorcheck[]{\tmsmonitor}{\tmsmonitor}{\absev{Use\ }\loc\ \sandboxtag}
-  }{tms-use}
-  %
-  \typerule{tms-alloc}{
-    (\loc;\sandboxtag)\notin\tmsmonitor.\text{allocated} &
-    (\loc;\sandboxtag)\notin\tmsmonitor.\text{freed} \\
-    \tmsmonitor[']=\left\{\text{allocated}: \tmsmonitor.\text{allocated}\cup\left\{(\loc;\sandboxtag)\right\},\text{freed}: \tmsmonitor.\text{freed}\right\}
-  }{
-    \monitorcheck[]{\tmsmonitor}{\tmsmonitor[']}{\absev{Alloc\ }\loc\ \sandboxtag}
-  }{tms-alloc}
-  %
-  \typerule{tms-dealloc}{
-    (\loc;\sandboxtag)\in\tmsmonitor.\text{allocated} &
-    (\loc;\sandboxtag)\notin\tmsmonitor.\text{freed} \\
-    \tmsmonitor[']=\left\{\text{allocated}: \tmsmonitor.\text{allocated}\setminus\left\{(\loc;\sandboxtag)\right\},\text{freed}: \tmsmonitor.\text{freed}\cup\left\{(\loc;\sandboxtag)\right\}\right\}
-  }{
-    \monitorcheck[]{\tmsmonitor}{\tmsmonitor[']}{\absev{Dealloc\ }\loc\ \sandboxtag}
-  }{tms-dealloc}
-\end{center}
-
-For \gls*{tms}, the state of the monitor is a record with two sets keeping track of allocated and deallocated locations.
-%The monitor runs on a modified trace semantics with more abstract $\absev{Alloc\ }\loc$,$\absev{Dealloc\ }\loc$,$\absev{Use\ }\loc$, and $\absterm$ events, which represent {\em allocation}, {\em deallocation}, and {\em use} of a location $\loc$ as well as abnormal program termination.
-\Cref{tr:tms-use} simply requires that a location is (i) allocated and (ii) not freed.
-\Cref{tr:tms-alloc,tr:tms-dealloc} both require a location to not be freed already and extend the monitor state accordingly.
-This restriction effectively disallows reallocation to reassign the same location to an object.
-However, the definition can easily be adapted by, e.g., attaching a natural number serving as a counter.
-Contrary to other monitors in this paper, the multi-step relation of the \gls*{tms} monitor is non-standard:
-\begin{center}
-  \judgbox{\monitorcheck{\tmsmonitor}{\tmsmonitor[']}{\abstrace}}{,,Monitor $\tmsmonitor$ multi-steps to $\tmsmonitor[']$ given trace $\abstrace$.''}$\;$\\
-  %
-  \typerule{tms-refl}{
-    \tmsmonitor.\text{allocated}=\emptyset
-  }{
-    \monitorcheck{\tmsmonitor}{\tmsmonitor}{\hole{\cdot}}
-  }{tms-refl}
-  %
-  \typerule{tms-ign-trans}{
-    \monitorcheck{\tmsmonitor}{\tmsmonitor[']}{\abstrace}
-  }{
-    \monitorcheck[]{\tmsmonitor}{\tmsmonitor[']}{\absev{\emptyevent}\cdot\abstrace}
-  }{tms-ign-trans}
-  %
-  \typerule{tms-trans}{
-    \monitorcheck[]{\tmsmonitor}{\tmsmonitor[']}{\absevent} &
-    \monitorcheck{\tmsmonitor[']}{\tmsmonitor['']}{\abstrace}
-  }{
-    \monitorcheck[]{\tmsmonitor}{\tmsmonitor['']}{\absevent\cdot\abstrace}
-  }{tms-trans}
-\end{center}
-\Cref{tr:tms-ign-trans,tr:tms-trans} are the same for all monitors, but \Cref{tr:tms-refl} has, in this case, an additional premise that all locations should be deallocated.
-This rejects the behavior of programs that forget to free memory.
-
-\begin{center}
-  \judgbox{\tmstraceagree[]{\event}{\absevent}}{,,Abstract event $\absevent$ is equivalent to $\event$ with respect to \gls*{tms}.''}$\;$\\
-  %
-  \typerule{tms-alloc-authentic}{
-  }{
-    \tmstraceagree[]{\ev{Alloc}\ \loc\ n;\sandboxtag;\securitytag}{\absev{Alloc\ }\loc\ \sandboxtag}
-  }{tms-alloc-authentic}
-  %
-% \typerule{tms-dealloc-authentic}{
-% }{
-%   \tmstraceagree[]{\ev{Dealloc}\ \loc;\sandboxtag;\securitytag}{\absev{Dealloc\ }\loc}
-% }{tms-dealloc-authentic}
-  %
-% \typerule{tms-use-authentic}{
-% }{
-%   \tmstraceagree[]{\ev{Use}\ \loc\ n;\sandboxtag;\securitytag}{\absev{Use\ }\loc}
-% }{tms-use-authentic}
-  %
-  \typerule{tms-branch-authentic}{
-  }{
-    \tmstraceagree[]{\ev{Branch\ }n}{\absev{\emptyevent}}
-  }{tms-branch-authentic}
-  %
-% \typerule{tms-binop-authentic}{
-% }{
-%   \tmstraceagree[]{\ev{Binop\ }n}{\absev{\emptyevent}}
-% }{tms-binop-authentic}
-% %
-% \typerule{tms-empty-authentic}{
-% }{
-%   \tmstraceagree[]{\emptyevent}{\absev{\emptyevent}}
-% }{tms-empty-authentic}
-% %
-  \typerule{tms-abort-authentic}{
-  }{
-    \tmstraceagree[]{\lightning}{\absterm}
-  }{tms-abort-authentic}
-\end{center}
-
-The trace agreement is entirely straightforward, so only allocation, branch, and crash are shown.
-
-\begin{lemma}[Traces with Monitor Satisfaction are $\tmssafe$]\label{lem:mon:tmssafe}
-  If $\montmssafe{\trace}$, then $\trace\in\tmssafe$. %\Coqed
-\end{lemma}
-
-\begin{example}[A program not satisfying \gls*{tms}]\label{ex:strncpy:cpp11}
-  Consider the following \texttt{C++11} library that calls \texttt{strncpy} (\Cref{ex:strncpy}) and prints the result to the standard output stream.
-
-  \begin{lstlisting}[language=c++,basicstyle=\small\ttfamily,mathescape,commentstyle=\color{gray}]
-int greet() {  // allocates 12 chars containing a greeting message
-  char* greetings = new char[12] { "Hello POPL!" }; // <- address ${\color{gray}\loc_{x}}$
-  char* to = new char[12];                          // <- address ${\color{gray}\loc_{y}}$
-
-  strncpy(12, to, greetings);
-  delete to;
-
-  printf("%cOPL\n", to[6]);
-}
-  \end{lstlisting}
-  Up to the body of \texttt{printf}, the program execution yields the specification trace $\ev{Alloc}\ \loc[_{x}]\ 12;\comp;\unlock\cdot\ev{Alloc}\ \loc[_{y}]\ 12;\comp;\unlock\cdot\ev{Use}\ \loc_{x}\ 0;\comp;\unlock\cdot\ev{Use}\ \loc_{x}\ 0;\comp;\unlock\cdot\ev{Use}\ \loc_{y}\ 0;\comp;\unlock\cdot\ev{Use}\ \loc_{x}\ 1;\comp;\unlock\cdot\ldots\cdot\ev{Use}\ \loc_{x}\ 12;\comp;\unlock\cdot\ev{Dealloc}\ \loc[_{y}];\comp;\unlock\cdot\ev{Use}\ \loc_{y}\ 6;\comp;\unlock$.
-  Relating this trace to abstract monitor events yields $\abstrace=\absev{Alloc}\ \loc[_{x}]\ \comp\cdot\absev{Alloc}\ \loc[_{y}]\ \comp\cdot\absev{Use}\ \loc[_{x}]\ \comp\cdot\absev{Use}\ \loc[_{x}]\ \comp\cdot\absev{Use}\ \loc[_{y}]\ \comp\cdot\ldots\cdot\absev{Use}\ \loc[_{x}]\ \comp\cdot\absev{Dealloc}\ \loc[_{y}]\ \comp\cdot\absev{Use}\ \loc[_{y}]\ \comp$.
-  Remembering the definition of \texttt{strncpy} (\Cref{ex:strncpy}), observe that it does not deallocate its arguments.
-  Even though the trace contains an out-of-bounds access right before returning from \texttt{strncpy}, this is no concern for \gls*{tms}, since the location $\loc[_{x}]$ is still allocated.
-  However, having returned from \texttt{strncpy}, the \texttt{greet} function continues and deallocates $\loc[_{y}]$ whose subsequent use in the \texttt{printf} call is a use-after-free bug.
-
-  The fix would be to \lstinline[language=c++,basicstyle=\small\ttfamily]|delete greetings| instead of \texttt{to} and add a \lstinline[language=c++,basicstyle=\small\ttfamily]|delete to| after the \texttt{printf} call, which leads to the abstract monitor trace $\abstrace[']=\absev{Alloc}\ \loc[_{x}]\ \comp\cdot\absev{Alloc}\ \loc[_{y}]\ \comp\cdot\absev{Use}\ \loc[_{x}]\ \comp\cdot\absev{Use}\ \loc[_{x}]\ \comp\cdot\absev{Use}\ \loc[_{y}]\ \comp\cdot\ldots\cdot\absev{Use}\ \loc[_{x}]\ \comp\cdot\absev{Dealloc}\ \loc[_{x}]\ \comp\cdot\absev{Use}\ \loc[_{y}]\ \comp$.
-  It follows that $\montmssafe{\abstrace[']}$ and from \Thmref{lem:mon:tmssafe}, it follows that the program satisfies \Thmref{def:trace:tmsdef}, even though the program still violates \gls*{sms}.
-\end{example}
-
-% \MPpost{
-% 	i think the trace description of traces is more uindertandable,
-% 	teh monitor one is easier to work with
-% 	this is the justification for having both that we'll have to make in the future
-% }
-
-\subsubsection{Monitor for \glsfirst*{sms}}\label{subsubsec:sms}
-\begin{gather*}
-  \begin{aligned}
-    \mi{(Abstract\ Store)}~\smsmonitor:=\ L\times\sandboxtag\times\mb{N}\hspace{0.33cm}&
-    \mi{(Abstract\ Events)}~\absevent\bnfdef\ \absev{\varepsilon} \mid \absev{Alloc\ }\loc\ \sandboxtag\ n \mid \absev{Use\ }\loc\ \sandboxtag\ n \\
-  \end{aligned}
-\end{gather*}
-\begin{center}
-  \judgbox{\monitorcheck[]{\smsmonitor}{\smsmonitor[']}{\absevent}}{,,Monitor $\smsmonitor$ does one step to $\smsmonitor[']$ given event $\absevent$.''}$\;$\\
-  %
-% \typerule{sms-uninteresting}{
-% }{
-%   \monitorcheck[]{\smsmonitor}{\smsmonitor}{\absev{\emptyevent}}
-% }{sms-uninteresting}
-% %
-% \typerule{sms-abort}{
-% }{
-%   \monitorcheck[]{\smsmonitor}{\emptyset}{\absterm}
-% }{sms-abort}
-  %
-  \typerule{sms-use}{
-    (\loc;\sandboxtag;m)\in\smsmonitor &
-    n<m
-  }{
-    \monitorcheck[]{\smsmonitor}{\smsmonitor}{\absev{Use\ }\loc\ \sandboxtag\ n}
-  }{sms-use}
-  %
-  \typerule{sms-alloc}{
-    (\loc;\sandboxtag;m)\notin\smsmonitor
-  }{
-    \monitorcheck[]{\smsmonitor}{\smsmonitor\cup\left\{(\loc;\sandboxtag;n)\right\}}{\absev{Alloc\ }\loc\ \sandboxtag\ n}
-  }{sms-alloc}
-\end{center}
-The state of the monitor for \gls*{sms} is a set containing tuples of locations, control-tags, and the allocation size.
-In comparison to the trace model of the \gls*{tms} monitor, the trace model here is extended by sizing and positional information.
-\Cref{tr:sms-use} performs a bounds check and \Cref{tr:sms-alloc} adds bounds information to the state of the monitor.
-The trace agreement is entirely straightforward and similar to the one for \gls*{tms}.
-% \begin{center}
-%   \judgbox{\smstraceagree[]{\event}{\absevent}}{,,Abstract event $\absevent$ is equivalent to $\event$ with respect to \gls*{sms}.''}$\;$\\
-%   %
-%   \typerule{sms-alloc-authentic}{
-%   }{
-%     \smstraceagree[]{\ev{Alloc}\ \loc\ n;\sandboxtag;\securitytag}{\absev{Alloc\ }\loc\ n}
-%   }{sms-alloc-authentic}
-%   %
-%   \typerule{sms-dealloc-authentic}{
-%   }{
-%     \smstraceagree[]{\ev{Dealloc}\ \loc;\sandboxtag;\securitytag}{\absev{Dealloc\ }\loc}
-%   }{sms-dealloc-authentic}
-%   %
-%   \typerule{sms-use-authentic}{
-%   }{
-%     \smstraceagree[]{\ev{Use}\ \loc\ n;\sandboxtag;\securitytag}{\absev{Use\ }\loc\ n}
-%   }{sms-use-authentic}
-%   %
-%   \typerule{sms-branch-authentic}{
-%   }{
-%     \smstraceagree[]{\ev{Branch\ }n}{\absev{\emptyevent}}
-%   }{sms-branch-authentic}
-%   %
-%   \typerule{sms-binop-authentic}{
-%   }{
-%     \smstraceagree[]{\ev{Binop\ }n}{\absev{\emptyevent}}
-%   }{sms-binop-authentic}
-%   %
-%   \typerule{sms-empty-authentic}{
-%   }{
-%     \smstraceagree[]{\emptyevent}{\absev{\emptyevent}}
-%   }{sms-empty-authentic}
-%   %
-%   \typerule{sms-abort-authentic}{
-%   }{
-%     \smstraceagree[]{\lightning}{\absterm}
-%   }{sms-abort-authentic}
-% \end{center}
-
-\begin{lemma}[Traces with Monitor Satisfaction are $\smssafe$]\label{lem:mon:smsafe}
-  If $\monsmssafe{\trace}$, then $\trace\in\smssafe$.\Coqed
-\end{lemma}
-
-\begin{example}[Normal invocation of \texttt{strncpy}]\label{ex:strncpy:sms}
-  Consider the insecure \texttt{strncpy} function from \Cref{ex:strncpy} with a context \texttt{strncpy(2, x, y)}, where \texttt{x} and \texttt{y} are pointers to valid regions of memory with allocated space for exactly two cells and do not contain the null-terminating character \texttt{'\textbackslash 0'}.
-  For the sake of this example, the pointers have been allocated by the component and passed to the context.
-  The loop of \texttt{strncpy} will copy exactly two cells and then check the loop condition for the last time.
-  At that stage, the induction variable \texttt{i} is equal to \texttt{2} and, unfortunately, the order of checks is such that first the cell \texttt{x[i]} is read prior to bounds checking \texttt{i < n}.
-  Because of this, there is an out-of-bounds memory access right before exiting the function.
-  This is also visible on the trace, which can be sketched as $\ldots\cdot\ev{Alloc}\ \loc_{x}\ 2;\comp;\unlock\cdot\ldots\cdot\ev{Alloc}\ \loc_{y}\ 2;\comp;\unlock\cdot\ldots\cdot\ev{Use}\ \loc_{x}\ 0;\comp;\unlock\cdot\ev{Use}\ \loc_{y}\ 0;\comp;\unlock\cdot\ev{Use}\ \loc_{x}\ 1;\comp;\unlock\cdot\ev{Use}\ \loc_{y}\ 1;\comp;\unlock\cdot\ev{Use}\ \loc_{x}\ 2;\comp;\unlock\cdot\ldots$, where $\loc_{x}$ and $\loc_{y}$ are the memory addresses associated to \texttt{x} and \texttt{y}, respectively.
-  Omitting the events for all ,,$\ldots$'' for sake of brevity, the abstract monitor trace of this is $\absev{Alloc}\ \loc_{x}\ \comp\ 2\cdot\absev{Alloc}\ \loc_{y}\ \comp\ 2\cdot\absev{Use}\ \loc_{x}\ \comp\ 0\cdot\absev{Use}\ \loc_{y}\ \comp\ 0\cdot\absev{Use}\ \loc_{x}\ \comp\ 1\cdot\absev{Use}\ \loc_{y}\ \comp\ 1\cdot\absev{Use}\ \loc_{x}\ \comp\ 2$.
-
-  After the allocation events, the state of the monitor is $\left\{(\loc_{x};\comp;2),(\loc_{y};\comp;2)\right\}$.
-  All uses up to the last are accepted by the monitor, but the last event does not satisfy the premise $2<2$ in \Cref{tr:sms-use}.
-  Therefore, the whole program (\texttt{strncpy} linked with this kind of context) is not \gls*{sms}.
-\end{example}
-
-\subsubsection{Combining \glsfirst*{tms} and \glsfirst*{sms} Monitors to obtain \glsfirst*{ms}}\label{subsubsec:ms}
-
-\begin{center}
-  \judgbox{\monitorcheck[]{\msmonitor}{\msmonitor[']}{\absevent}}{,,Monitor $\msmonitor$ does one step to $\msmonitor[']$ given event $\absevent$.''}$\;$\\
-  \typerule{ms-step}{
-    \monitorcheck[]{\tmsmonitor}{\tmsmonitor[']}{\absevent_{\tmssafe}} &
-    \monitorcheck[]{\smsmonitor}{\smsmonitor[']}{\absevent_{\smssafe}} &
-  }{
-    \monitorcheck[]{(\tmsmonitor,\smsmonitor)}{(\tmsmonitor['],\smsmonitor['])}{(\absevent_{\tmssafe},\absevent_{\smssafe})}
-  }{ms-step}
-\end{center}
-The combined monitor runs the one for \gls*{tms} and the one for \gls*{sms} in a lockstep.
-The trace agreement similarly just relates a specification event with an abstract \gls*{tms}-event $\abstrace[_{\tmssafe}]$ and with an abstract \gls*{sms}-event $\abstrace[_{\smssafe}]$.
-% \begin{center}
-%   \judgbox{\mstraceagree{\event}{\absevent}}{,,Abstract event $\absevent$ is equivalent to $\event$ with respect to \gls*{ms}.''}$\;$\\
-%   %
-%   \typerule{ms-authentic}{
-%     \tmstraceagree[]{\event}{\absevent_{\tmssafe}} &
-%     \smstraceagree[]{\event}{\absevent_{\smssafe}}
-%   }{
-%     \mstraceagree[]{\event}{(\absevent_{\tmssafe},\absevent_{\smssafe})}
-%   }{ms-authentic}
-% \end{center}
-
-\begin{lemma}[Traces with Monitor Satisfaction are $\mssafe$]\label{lem:mon:msafe}
-  If $\monmssafe{\trace}$, then $\trace\in\mssafe$.\Coqed
-\end{lemma}
-
-\subsubsection{Monitor for \glsfirst*{scct}}
-\begin{gather*}
-  \begin{aligned}
-    \mi{(Abstract\ Store)}~\scctmonitor&:=\ \emptyset \hspace{0.5cm}
-    \mi{(Abstract\ Events)}~\absevent:=\ \absev{\varepsilon} \mid \absterm \mid \absev{Leak} \\
-  \end{aligned}
-\end{gather*}
-\begin{center}
-  \judgbox{\monitorcheck[]{\scctmonitor}{\scctmonitor[']}{\absevent}}{,,Monitor $\scctmonitor$ does one step to $\scctmonitor[']$ given event $\absevent$.''}$\;$\\
-  %
-  \typerule{scct-none}{
-  }{
-    \monitorcheck[]{\scctmonitor}{\scctmonitor}{\absev{\emptyevent}}
-  }{scct-none}
-  %
-  \typerule{scct-abort}{
-  }{
-    \monitorcheck[]{\scctmonitor}{\scctmonitor}{\absterm}
-  }{scct-abort}
-\end{center}
-The monitor state for the \gls*{scct} monitor is completely empty, since it does not need to keep track of information.
-As soon as any event is hit, the execution gets stuck, since any event is considered confidental from the perspective of this monitor.
-\begin{center}
-  \judgbox{\sccttraceagree[]{\event}{\absevent}}{,,Abstract event $\absevent$ is equivalent to $\event$ with respect to \gls*{cct}.''}$\;$\\
-  %
-  \typerule{scct-low-authentic}{
-  }{
-    \sccttraceagree[]{\event[_b];\sandboxtag;\unlock}{\absev{\emptyevent}}
-  }{scct-low-authentic}
-  %
-  \typerule{scct-high-authentic}{
-  }{
-    \sccttraceagree[]{\event[_b];\sandboxtag;\lock}{\absev{Leak}}
-  }{scct-high-authentic}
-  %
-  \typerule{scct-empty-authentic}{
-  }{
-    \sccttraceagree[]{\emptyevent}{\absev{\emptyevent}}
-  }{scct-empty-authentic}
-  %
-  \typerule{scct-abort-authentic}{
-  }{
-    \sccttraceagree[]{\lightning}{\absterm}
-  }{scct-abort-authentic}
-\end{center}
-Accordingly, the event agreement simply disregards all events that involved public data ($\unlock$) while mapping any other event that does involve private data ($\lock$) to the abstract $\absev{Leak}$ event.
-
-\begin{lemma}[Monitor Traces are $\scctsafe$]\label{lem:mon:scctsafe}
-  If $\monscctsafe{\trace}$, then $\trace\in\scctsafe$. %\Coqed
-\end{lemma}
-
-\begin{example}[Data-independent timing mode]
-  Consider the call \texttt{strncpy(1,x,y)} to the \texttt{strncpy} function (\Cref{ex:strncpy}) with low ($\unlock$) security for \texttt{x} and high security ($\lock$) for \texttt{y}.
-  The trace of just the copying part inside of \texttt{strncpy} looks like $\ev{Use}\ \loc_{x}\ 0;\comp;\unlock\cdot\ev{Use}\ \loc_{y}\ 0;\comp;\lock$.
-  In terms of the abstract monitor trace, this is just $\absev{Leak}$.
-  Running this on the monitor would result in getting stuck, since there is no matching rule to step in the presence of $\absev{Leak}$ event. % pun intended
-  By means of additional features to ensure cryptographic constant time even in the presence of memory reads and loads, such as a flag to enable a data independent timing mode, which is present in both Arm~\cite[p.~543]{arm-refman} and Intel~\cite[p.~80]{intel-refman} processors, the original trace now does $\ev{Use}\ \loc_{y}\ 0;\comp;\unlock$ instead of $\ev{Use}\ \loc_{y}\ 0;\comp;\lock$.
-  Because of this, the whole trace of the component equates to $\absev{\emptyevent}$ and the monitor can step without getting stuck.
-\end{example}
-
-\subsubsection{Combining \gls*{ms} and \gls*{scct} Monitors to obtain \gls*{msscct}}
-
-The combination of monitors for \gls*{ms} and \gls*{scct} yields one for \gls*{msscct}.
-The construction is entirely similar to the one for \gls*{ms} (\Cref{subsubsec:ms}).
-
-\begin{lemma}[Traces with Monitor Satisfaction are $\msscctsafe$]\label{lem:mon:msscctsafe}
-  If $\monmsscctsafe{\trace}$, then $\trace\in\msscctsafe$. %\Coqed
-\end{lemma}
-
-\section{Composing Secure Compilers\pages{1}}\label{sec:compcomp}
-
-This section presents the key meta-theoretic results of this paper concerning sequential compiler composition (and of optimisation passes) (\Cref{sec:sequential}) and concerning other kinds of compiler composition (\Cref{sec:other-compos}).
-% and builds on prior work~\cite{kruse2022csc}\MK{mention this at all?
-may reveal identity even though its double-blind}.
-% MP: not needed at all, ever, workshop work does not need to be cited in the main work
-
-\subsection{Secure Sequential Composition}\label{sec:sequential}
-
-The main result is that secure compilers in the robust compilation framework~\cite{abate2019jour} compose {\em sequentially}.
-This is not intuitive in the sense that in the security domain, composition does not work without additional generalizations~\cite{canetti2006univcomp,mccullough2012compo,fabian2022automatic}.
-The \bul{sequential composition of compilers $\cc{\src{L}}{\trg{L}}$ and $\cc{\trg{L}}{\obj{L}}$} is defined as follows:
-Given an $\src{L}$ program $\src{p}$ and compilers $\cc{\src{L}}{\trg{L}},\cc{\trg{L}}{\obj{L}}$, \rul{its compiled $\obj{L}$ counterpart is obtained by plugging $\src{p}$ into $\cc{\src{L}}{\trg{L}}\circ\cc{\trg{L}}{\obj{L}}$}.
-\begin{definition}
-  \bul{$\cc{\src{L}}{\trg{L}}\circ\cc{\trg{L}}{\obj{L}}$} $\isdef$ Given $\src{p}$, yield \rul{$\cc{\trg{L}}{\obj{L}}\left(\cc{\src{L}}{\trg{L}}(\src{p})\right)$}
-\end{definition}
-
-Consider the compilation chain for $\src{TypeScript}$.
-First, $\src{TypeScript}$ programs are translated to $\trg{JavaScript}$ which, e.g., V8~\cite{googlev8} eventually compiles in parts to $\irl{Ignition BC}$.
-The following theorem establishes what happens if all these compilation steps were robustly secure with respect to \gls*{ms}:
-The resulting $\irl{Ignition BC}$ code would be \gls*{ms} regardless of the context the binary runs in.
-
-Given \bul{$\cc{\src{L}}{\trg{L}}$ robustly preserves $\class[_{1}]$} and \rul{$\cc{\trg{L}}{\obj{L}}$ robustly preserves $\class[_{2}]$}, it follows that \oul{their sequential composition $\cc{\src{L}}{\trg{L}}\circ\cc{\trg{L}}{\obj{L}}$ robustly preserves the intersection of classes $\class[_{1}]$ and $\class[_{2}]$}.
-
-\begin{theorem}[Sequential Composition of Secure Compilers]\label{thm:rtp}
-  If \bul{$\rtp{\cc{\src{L}}{\trg{L}}}{\class[_{1}]}$} and \rul{$\rtp{\cc{\trg{L}}{\obj{L}}}{\class[_{2}]}$}, then \oul{$\rtp{\cc{\src{L}}{\trg{L}}\circ\cc{\trg{L}}{\obj{L}}}{\class[_{1}]\cap\class[_{2}]}$}.
-\Coqed
-\end{theorem}
-
-Since the composition of secure compilers is again a secure compiler, the theorem generalises to a whole chain of $n$ secure compilers.
-
-\subsubsection{Securing Optimisations}
-Notably, real-world compilation chains also perform a series of (sequential) passes whose main purpose is not necessarily to translate from one language to another, but to, e.g., optimise the code or enforce a certain property.
-Both examples can be seen in practice, e.g. as in the work of~\cite{nagarakatte2009soft,nagarakatte2010cets,akritidis2009baggy,wegman1991ccp,manjikian1997fusion} and many more.
-% \MPin{survey to cite?
-this is a bit informal MK: hard to find a survey on different compiler passes...
-I can't even find one just for compiler enforced memsafety (that looks good) or for compiler translations in general}.
-% 
-% Consider the compilation passes of LLVM~\cite{lattner2004llvm}, which are handled in the \texttt{PassManager} class.
-% By manipulating the internal state of a \texttt{PassManager} object, the schedule of compiler optimisation passes can be hardcoded.
-% That way, an order of optimisations can be found by hand that yields the best results.
-% \MP{what's the point of the previous 3 sentences?(commented)}
-Consider the following two LLVM optimisation passes: \gls*{cf}, which rewrites constant expressions to the constant they evaluate to, and \gls*{dce}, which removes dead code by rewriting conditional branches.
-The order in which \gls*{cf} and \gls*{dce} are performed influences the final result of the compilation (see \Cref{fig:cfdceex}).
-\begin{figure}[!ht]
-  \begin{center}
-    \begin{tikzpicture}
-      \node[minimum width=3cm,minimum height=2.5cm,draw=black,very thick,align=left] (progunopt) {\begin{lstlisting}[language=ml,basicstyle=\small\ttfamily]
-let a = true in
-if a then
-  print "a"
-else
-  print "b"
-\end{lstlisting}};
-      \node[minimum width=3cm,minimum height=2.5cm,draw=black,very thick,align=left,right=1.0 of progunopt] (progoptcf) {\begin{lstlisting}[language=ml,basicstyle=\small\ttfamily]
-if true then
-  print "a"
-else
-  print "b"
-\end{lstlisting}};
-      \node[minimum width=3cm,minimum height=2.5cm,draw=black,very thick,align=left,right=1.0 of progoptcf] (progoptcfdce) {\begin{lstlisting}[language=ml,basicstyle=\small\ttfamily]
-print "a"
-\end{lstlisting}};% the following is a hack to get my syntax highlighting to work again \end{lstlisting}
-      % edges
-      \draw[->,very thick] (progunopt) edge[loop left,left] node {\gls*{dce}} (progunopt);
-      \draw[->,very thick] (progunopt) edge[sloped,above] node {\gls*{cf}} (progoptcf);
-      \draw[->,very thick] (progoptcf) edge[sloped,above] node {\gls*{dce}} (progoptcfdce);
-    \end{tikzpicture}
-  \end{center}
-  \caption{Example program where the level of optimisations differ for one pass of applying \gls*{cf} and \gls*{dce} in any order. %
-  Every edge is a compilation pass and the label on the edge states what the pass does, i.e., \gls*{cf} or \gls*{dce}. %
-  The source code in the nodes is a glorified compiler intermediate representation and the code gets more optimised towards the right hand side of the figure.}\label{fig:cfdceex}
-\end{figure}
-This {\em phase ordering problem} is well--known in literature and a practical solution is to simply perform a fixpoint iteration of the optimisation pipeline~\cite{click1995combining}.
-%If possible, optimisations are merged to enhance the analysis' result, since even when iterating to a fixpoint the final program may not be as optimised as it could be~\cite{click1995combining}.
-Compiler engineers typically try to find an order of optimisations that yields well-optimised programs for either code size~\cite{cooper1999geneticphases} or performance~\cite{kulkarni2006exhaustivephase}.
-\Cref{corr:swappable} justifies that any such order of compilation passes is valid with respect to security.
-So, given two compilation passes \bul{$\cc[_{1}]{\trg{L}}{\trg{L}}$, $\cc[_{2}]{\trg{L}}{\trg{L}}$, both robustly preserving class $\class[_{1}]$ or $\class[_{2}]$, respectively}, \oul{for any order of their composition the composed compiler robustly preserves the intersection of $\class[_{1}]$ and $\class[_{2}]$}.
-
-\begin{corollary}[Sequential Composition of Secure Compilers]\label{corr:swappable}
-  If \bul{$\rtp{\cc[_{1}]{\trg{L}}{\trg{L}}}{\class[_{1}]}$ and $\rtp{\cc[_{2}]{\trg{L}}{\trg{L}}}{\class[_{2}]}$}, then \oul{$\rtp{\cc[_{1}]{\trg{L}}{\trg{L}}\circ\cc[_{2}]{\trg{L}}{\trg{L}}}{\class[_{1}]\cap\class[_{2}]}$ and $\rtp{\cc[_{2}]{\trg{L}}{\trg{L}}\circ\cc[_{1}]{\trg{L}}{\trg{L}}}{\class[_{2}]\cap\class[_{1}]}$}.
-\Coqed
-\end{corollary}
-
-\subsection{Secure Upper and Lower Composition}\label{sec:other-compos}
-Besides sequential composition, there are two other compositions, namely an {\em upper}, i.e., a compiler that takes multiple inputs and yields one output, and a {\em lower} composition, i.e., a compiler that takes one input and yields multiple outputs.
-\bul{Define the upper composition $\cc{\src{L}+\obj{L}}{\trg{L}}$} as follows:
-Given a program \texttt{p}, its compiled counterpart is obtained by \rul{plugging \texttt{p} into $\cc{\src{L}}{\trg{L}}$ if $\texttt{p}\in\src{L}$} or by \oul{plugging \texttt{p} into $\cc{\obj{L}}{\trg{L}}$ if $\texttt{p}\in\obj{L}$}.
-\begin{definition}[Upper Composition]
-  \bul{$\cc{\src{L}+\obj{L}}{\trg{L}}$} $\isdef$ 
-  % Given \texttt{p}, yield
-  $\lambda \texttt{p}\ldotp$
-  $\left\{\mbox{\begin{tabular}{c}
-    {\rul{if $\texttt{p}\in\src{L}$, then $\cc{\src{L}}{\trg{L}}(\texttt{p})$}} \\
-    \mbox{\oul{if $\texttt{p}\in\obj{L}$, then $\cc{\obj{L}}{\trg{L}}(\texttt{p})$}} \\
-  \end{tabular}}\right.$
-%
-%
-%  $\begin{cases}
-%                                                                            \rul{\cc{\src{L}}{\trg{L}}(\texttt{p}) & \text{if }\texttt{p}\in\src{L}} \\
-%                                                                            \cc{\obj{L}}{\trg{L}}(\texttt{p}) & \text{if }\texttt{p}\in\obj{L}
-%                                                                          \end{cases}$
-\end{definition}
-
-Examples of this are present in industry:
-Consider the Java Virtual Machine bytecode $\trg{JVM BC}$, which is a popular target for programming language designers due to its high performance and relevance in industry.
-Compilers for several programming languages have it as their target language, some popular instances are $\src{Java}$ and $\obj{Kotlin}$.
-Technically speaking, they both compile to class files and $\obj{Kotlin}$ objects are considered to be the same as $\src{Java}$ objects at that point.
-Both languages can be used at the same time in one project~\cite{androidstudio}.
-A compiler that accepts both $\src{Java}$ and $\obj{Kotlin}$ code translating to the same target language or intermediate representation performs a kind of {\em upper} composition.
-Now, the following theorem tells us what happens if these are secure:
-Given \bul{$\cc{\src{L}}{\trg{L}}$ robustly preserves $\class[_{1}]$} and \rul{$\cc{\obj{L}}{\trg{L}}$ robustly preserves $\class[_{2}]$}, it follows that \oul{their upper composition $\cc{\src{L}+\obj{L}}{\trg{L}}$ robustly preserves the intersection of classes $\class[_{1}]$ and $\class[_{2}]$}.
-
-\begin{theorem}[Upper Composition of Secure Compilers]\label{thm:urtp}
-  If \bul{$\rtp{\cc{\src{L}}{\trg{L}}}{\class[_{1}]}$} and \rul{$\rtp{\cc{\obj{L}}{\trg{L}}}{\class[_{2}]}$}, then \oul{$\rtp{\cc{\src{L}+\obj{L}}{\trg{L}}}{\class[_{1}]\cap\class[_{2}]}$}. \Coqed
-\end{theorem}
-
-Dually, the {\em lower} composition is concerned about compilers that accept the same source but yield different target languages. %and the same theoretical results apply.
-\bul{Define the lower composition $\cc{\src{L}}{\trg{L}+\obj{L}}$} as follows:
-Given a program $\src{p}$, its compiled counterpart is obtained by \rul{plugging $\src{p}$ into $\cc{\src{L}}{\trg{L}}$} or by \oul{plugging \texttt{p} into $\cc{\src{L}}{\obj{L}}$}, respectively, \pul{based on the internal decision}.
-\begin{definition}[Lower Composition]
-  \bul{$\cc{\src{L}}{\trg{L}+\obj{L}}$} $\isdef$ 
-  % Given $L$ and $\src{p}$, yield
-  $\lambda \src{p}, L\ldotp$
-  $\left\{\mbox{\begin{tabular}{c}
-    {\pul{if $L=\trg{L}$, then} \rul{$\cc{\src{L}}{\trg{L}}(\src{p})$}} \\
-    \mbox{\pul{if $L=\obj{L}$, then} \oul{$\cc{\src{L}}{\obj{L}}(\src{p})$}} \\
-  \end{tabular}}\right.$
-%
-%
-%  $\begin{cases}
-%                                                                            \rul{\cc{\src{L}}{\trg{L}}(\texttt{p}) & \text{if }\texttt{p}\in\src{L}} \\
-%                                                                            \cc{\obj{L}}{\trg{L}}(\texttt{p}) & \text{if }\texttt{p}\in\obj{L}
-%                                                                          \end{cases}$
-\end{definition}
-
-Consider two compilers both accepting $\src{LLVM IR}$~\cite{lattner2004llvm} and one of them emits $\trg{x86\_64}$, while the other emits $\obj{ARMv8}$.
-It is intuitive that they are in some sense composed in the LLVM framework, but the decision of when to use one over the other is inherently {\em internal} to the formalisation effort of this kind of composition.
-For example, the user of this compiler provides an explicit flag that instructs to emit $\trg{x86\_64}$ or the framework itself detects the target platform via heuristics, such as supported instructions.
-
-The following theorem demonstrates what happens if the involved compilers are secure:
-Given \bul{$\cc{\src{L}}{\trg{L}}$ robustly preserves $\class[_{1}]$} and \rul{$\cc{\src{L}}{\obj{L}}$ robustly preserves $\class[_{2}]$}, it follows that \oul{their lower composition $\cc{\src{L}}{\trg{L}+\obj{L}}$ robustly preserves the intersection of classes $\class[_{1}]$ and $\class[_{2}]$}.
-
-\begin{theorem}[Lower Composition of Secure Compilers]\label{thm:lrtp}
-  If \bul{$\rtp{\cc{\src{L}}{\trg{L}}}{\class[_{1}]}$} and \rul{$\rtp{\cc{\obj{L}}{\trg{L}}}{\class[_{2}]}$}, then \oul{$\rtp{\cc{\src{L}}{\trg{L}+\obj{L}}}{\class[_{1}]\cap\class[_{2}]}$}. %\Coqed
-\end{theorem}
-
-%Either way, the theoretical results suggest that it is possible to always find a ,,most-general'', secure compiler, given two secure compilers, that robustly preserves the least-upper bound of the classes involved in their compilation process.
-
-%\section{Case Study: Memory Safe and Cryptographic Constant Time Code}\label{sec:casestud}
-\section{Case Study: Language Formalisations\pages{2}}\label{sec:casestud:defs}
-
-This section defines programming languages that the secure compilers defined in the next section will use.
-To this end, this section defines the languages $\Ltms$, $\Ltrg$, $\Lms$, and $\Lscct$ which share many common elements (presented in \Cref{subsec:cs:defs}).
-$\Ltms$ is the only statically typed language and exhibits the property that all well-typed programs are \gls*{tms} (\Cref{subsec:ltms}).
-However, not all $\Ltms$ programs are \gls*{sms}.
-That is, there are well-typed $\Ltms$ programs that perform an out-of-bounds access.
-Language $\Ltrg$ is untyped and does not provide any guarantees with regards to \gls*{ms} (\Cref{subsec:lsms}).
-$\Lms$ is exactly the same language as $\Ltrg$, but this paper still distinguishes the two for sake of readability (\Cref{subsec:lms}).
-All three languages --- so $\Ltms$, $\Ltrg$, and $\Lms$ --- assume \gls*{cct} to hold.
-
-% This paper considers \gls*{cct} as an architecture-dependent property\MPin{can we justify this}, so
-Writing code attaining \gls*{cct} should not be of the programmer's concerns~\cite{cauligi2019fact}.
-Such consideration is also backed up by architecture providing a data (operand) independent timing mode, such as processors by Arm~\cite[p.~543]{arm-refman} and Intel~\cite[p.~80]{intel-refman}.
-In spirit of this, language $\Lscct$ allows violating \gls*{cct} by emitting events on, e.g., branching and division, that contain secrets (\Cref{subsec:lscct}), but provides a way to read and write to a {\em model--specific register} that enables a ``\gls*{cct}-mode''.
-
-\subsection{Shared Language Definitions}\label{subsec:cs:defs}
-\begin{gather*}
-  \begin{aligned}
-    \mi{(Expressions)}~&\expr\bnfdef\ x \mid \valueexpr \mid \lbinop{\expr[_{1}]}{\expr[_{2}]} \mid \lget{x}{\expr} \mid \llet{x}{\lnew{\expr[_{1}]}{\expr[_{2}]}}{\expr[_{3}]} \mid \ldelete{x} \mid \lset{x}{\expr[_{1}]}{\expr[_{2}]} \\
-                       & \mid \lpair{\expr[_{1}]}{\expr[_{2}]} \mid \lproja{\expr} \mid \lprojb{\expr} \mid \llet{x}{\expr[_{1}]}{\expr[_{2}]} \mid \lreturn{\expr} \mid \lcall{g}{\expr} \mid \lifz{\expr}{\expr[_{1}]}{\expr[_{2}]} \mid \labort
-  \end{aligned}
-  \\
-  \begin{aligned}
-  \mi{(Types)}~&\type\bnfdef\natt \mid \type[_{1}] \times \type[_{2}]
-  &
-  \mi{(Functions)}~&\function\bnfdef\ \lfunction{\textit{foo}}{x}{e}
-  &
-  \mi{(Libraries)}~&\library\bnfdef\ \hole{\cdot} \mid \function,\library
-  \end{aligned}\\
-  \begin{aligned}
-  \mi{(Component~Names)}~&\commlib\bnfdef\ \hole{\cdot} \mid \text{foo},\commlib
-  &
-  \mi{(Programs)}~&\prog{\library_{\ctx}}{\library_{\comp}}
-  \end{aligned}
-\end{gather*}
-
-Above is the shared syntax of all the programming languages of this paper.
-Variables are referred to as $x,y,z,a,b,c,\ldots$ while functions may be referred to as multi-character words, such as $\texttt{foo}$, as well as short-forms like $f,g,h$.
-All languages share the type $\natt$ representing natural numbers.
-Functions are constrained to take one argument and can only call other functions listed in libraries, which are just lists of functions.
-A program $\prog{\library_{\ctx}}{\library_{\comp}}$ is indexed by two libraries that represent all context- and component-level functions, respectively.
-Lists of component-level names are referred to as $\commlib$.
-
-\begin{gather*}
-  \begin{aligned}
-    \mi{(Control~Tags)}~&\sandboxtag\bnfdef\ctx \mid \comp
-    &
-    \mi{(Communication~Tags)}~&\comm\bnfdef\ \ctxtocomp \mid\ \comptoctx \mid \nocomm
-    &
-    \mi{(Poison~Tags)}~&\poison\bnfdef \poisoned \mid \poisonless
-  \end{aligned}\\
-  \begin{aligned}
-    \mi{(Continuation~Stacks)}~&\ectxstack\bnfdef\hole{\cdot} \mid \tup{K}{g},\ectxstack
-    &
-    \mi{(Control~Flow~States)}~&\cfstate\bnfdef\tup{\library}{\ectxstack}
-  	\\
-    \mi{(Locations)}~&\loc\in\nat
-    &
-    \mi{(Substitutions)}~&\substvar\bnfdef \hole{\cdot} \mid \subst{\valueexpr}{x},\gamma
-  	\\
-    \mi{(Stores)}~&\Delta\bnfdef\hole{\cdot} \mid x\mapsto\tup{\loc}{\sandboxtag,\poison,n},\Delta
-    &
-    \mi{(Heaps)}~&H\bnfdef\hole{\cdot} | \valueexpr,H
-    \end{aligned}\\
-  \begin{aligned}
-    \mi{(Memory~States)}~&\memstate\bnfdef\tup{H^{\ctx}}{H^{\comp};\Delta}
-    &
-    \mi{(States)}~&\statevar\bnfdef\tup{\cfstate}{\sandboxtag,\memstate}
-    &
-    \mi{(Runtime~Terms)}~&\runtimetermvar\bnfdef\rtt{\statevar}{\expr}
-  \end{aligned}
-\end{gather*}
-
-States $\statevar$ are tuples\footnote{Throughout the paper, the substitution notation is also used to update entries in states $\statevar$.} containing a control flow state $\cfstate$, control tags $\sandboxtag$, and a memory state $\memstate$.
-A control flow state $\cfstate$ entails a library, which provides definitions for functions calls, as well as a stack of continuations $\ectxstack$.
-Elements of the stack $\ectxstack$ are pairs of evaluation contexts and the name of a function associated to that evaluation context.
-Memory states $\memstate$ are tuples of two separate heaps $H$ and a store $\Delta$, which contains pointer metadata, such as the concrete memory location $\loc$, a control tag $\sandboxtag$ indicating which heap the pointer points into, a poison tag $\poison$, as well as bounds information $n$.
-The bounds information is a mere proof artefact that has no semantic significance.
-The two separate heaps essentially model a sandbox, to prevent contexts from performing pointer arithmetic and reading from or writing to the data owned by a component.
-While this prevents the effects of out-of-bounds accesses across the context and component boundary, the goal of the design of the languages of this paper is to be able to express security violations.
-To this end, the poison tag $\poison$ indicates whether a pointer has been freed $\poisoned$ or is still allocated $\poisonless$, so that pointers can be used even after their deallocation without the semantics getting stuck.
-Runtime terms are simply expressions $\expr$ paired with the operational state $\statevar$.
-
-\begin{gather*}
-  \begin{aligned}
-  \mi{(Pre\text{-}Events)}~\event[_{b}]&\bnfdef \ev{Alloc}\ \loc\ \valueexpr \mid \ev{Dealloc}\ \loc \mid \ev{Get}\ \loc\ \valueexpr \mid \ev{Set}\ \loc\ \valueexpr\ \valueexpr['] \mid \cdots \\
-  \mi{(Events)}~\event&\bnfdef \emptyevent \mid \lightning \mid \tup{\event[_{b}]}{\sandboxtag,\securitytag} \\
-  \end{aligned}
-\end{gather*}
-
-All languages use the same trace model, where events are either the empty event \emptyevent, the program crash event $\lightning$, or a tuple consisting of a control tag $\sandboxtag$ and a security tag $\securitytag$.
-The former indicates whether the component $\comp$ or the context $\ctx$ is to blame for emitting this event, the latter indicates the secrecy level of values of the emitted event, i.e., either $\lock$ or $\unlock$.
-As for pre-events, the memory-related ones are allocation ($\ev{Alloc}\ \loc\ \valueexpr$), deallocation ($\ev{Dealloc}\ \loc$), reading from ($\ev{Get}\ \loc\ \valueexpr$) and writing to memory ($\ev{Set}\ \loc\ \valueexpr\ \valueexpr[']$).
-The following is an excerpt of the operational semantics handling some of the memory operations.
-
-\begin{center}
-  \judgbox{\pstepto{\runtimetermvar}{\runtimetermvar[']}{\event}}{,,$\runtimetermvar$ does one primitive step to $\runtimetermvar[']$ emitting event $\event$.''}
-  %
-  \typerule{$e-\text{get}-\in$}{
-    \sandboxtag=\statevar.\sandboxtag &
-    \statevar.\store(x)=\tup{\loc}{\sandboxtag,\poison,m} \\
-    \loc+n\in\operatorname{dom}\statevar.\heap[^{\sandboxtag}]
-  }{
-    \pstepto{\rtt{\statevar}{\lget{x}{n}}}{\rtt{\statevar}{\statevar.\heap[^{\sandboxtag}](\loc+n)}}{\tup{\ev{Get}\ \loc\ n}{\sandboxtag}}
-  }{e-get-in}
-  %
-  \typerule{$e-\text{set}-\notin$}{
-    \sandboxtag=\statevar.\sandboxtag &
-    \statevar.\store(x)=\tup{\loc}{\sandboxtag,\poison,m} \\
-    \loc+n\notin\operatorname{dom}\statevar.\heap[^{\sandboxtag}]
-  }{
-    \pstepto{\rtt{\statevar}{\lset{x}{n}{\valueexpr}}}{\rtt{\statevar}{\valueexpr}}{\tup{\ev{Set}\ \loc\ n\ \valueexpr}{\sandboxtag}}
-  }{e-set-notin}
-  %
-  \typerule{$e-\text{new}$}{
-    \fresh{\statevar}{z} &
-    \fresh{\statevar}{\loc} &
-    \heap[^{\sandboxtag}_1]=\growh{\statevar.\heap[^{\sandboxtag}]}{n} &
-    \store[_1]=\storeel{z}{\loc}{\statevar.\sandboxtag}{\poisonless}{n},\statevar.\store
-  }{
-    \pstepto{\rtt{\statevar}{\lnew{x}{n}{\expr}}}{\rtt{\statevar\subst{\heap[^{\sandboxtag}]}{\heap[^{\sandboxtag}_1]}\subst{\store}{\store[_1]}}{\expr\subst{z}{x}}}{\tup{\ev{Alloc}\ \loc\ n}{\sandboxtag}}
-  }{e-new}
-  %
-  \typerule{$e-\text{dealloc}$}{
-    \statevar.\store(x) = \tup{\loc}{\statevar.\sandboxtag,\poison,n} &
-    \store[_1]=\seth{\statevar.\store}{x}{\tup{\loc}{\statevar.\sandboxtag,\poisoned,n}}
-  }{
-    \pstepto{\rtt{\statevar}{\ldelete{x}}}{\rtt{\statevar\subst{\store}{\store[_1]}}{0}}{\tup{\ev{Dealloc}\ \loc}{\sandboxtag}}
-  }{e-dealloc}
-  %
-  \typerule{$e-\text{abort}$}{
-  }{
-    \pstepto{\rtt{\statevar}{\labort}}{\lightning}{\lightning}
-  }{e-abort}
-\end{center}
-
-To demonstrate the use of the poison tag $\poison$ as metadata for pointers instead of removing them from the store $\Delta$, consider \Cref{tr:e-dealloc,tr:e-get-in,tr:e-set-notin} .
-In \Cref{tr:e-dealloc}, the premise does not care at all about the actual state of the poison tag $\poison$ and just overwrites it, marking the location as freed $\poisoned$.
-Besides that, the poison tag does not have any semantic meaning.
-For language $\Ltms$, this tag is really just some semantic metadata that programmers have no access to.
-But, for the other languages, e.g., $\Ltrg$, the poison tag is used to check pointer validity.
-\Cref{tr:e-new} allocates enough space on the respective heap, either $H^{\ctx}$ or $H^{\comp}$ depending on the execution context, i.e., the value of $\Omega.\sandboxtag$, and adds the appropriate metadata associated to the pointer in $\Delta$.
-Reading from \Cref{tr:e-get-in} and writing to memory \Cref{tr:e-set-notin} have two cases: Either the heap is large enough or not and, depending on that, either the actual value stored at that location is read from or written to, or some garbage data is returned.
-However, note that the execution does not get stuck in such cases, it performs a step, and emits an appropriate event.
-Also note that whether a pointer is poisoned or not is not reflected on the trace.
-
-\begin{center}
-  \begin{align*}
-    \mi{(Pre\text{-}Events)}~\event[_{b}]&\bnfdef \cdots \mid \ev{Call}\ \comm\ g\ \valueexpr \mid \ev{Ret}\ \comm\ \valueexpr \mid \ev{Start} \mid \ev{End}\ \valueexpr \\
-  \end{align*}
-\end{center}
-
-A key difference in comparison with the specification trace model (\Cref{subsec:propdefs}) is that, as standard in secure compilation work~\cite{patrignani2021rsc,abate2019jour,korashy2021capableptrs}, the traces have a call and return event that signals context switches, which are referred to as {\em interaction events}.
-The reason for these interaction events is technical: They are a proof artifact for reconstructing a source context from a potentially malicious target context, where during that translation, the insertion of some wrapper code right before context switching may be necessary to make the proof succeed.
-Hereby, a $\ev{Call\ \ctxtocomp\ foo\ \valueexpr}$ and $\ev{Return\ \ctxtocomp\ \valueexpr}$ signal that program execution transitions from context- to component-level.
-Contrary, $\ev{Call\ \comptoctx\ foo\ \valueexpr}$ and $\ev{Return\ \comptoctx\ \valueexpr}$ signal that program execution transitions from component- to context-level.
-For calls without this context switch, the contextual semantics attaches the $\nocomm$ tag.
-In the following, $\neg\ctx=\comp$ and $\neg\comp=\ctx$.
-
-\begin{center}
-  \judgbox{\estepto[]{\runtimetermvar}{\runtimetermvar}{\event}}{,,Contextual step from runtime-term $\runtimetermvar$ to $\runtimetermvar[']$ emitting event $\event$.''}
-  %
-  \typerule{$e-\text{ret}$}{
-    \statevar.\ectxstack=\tup{\ectx}{\text{foo}},\ectxstack['] &
-    \commlib\vdash\text{foo}:\Omega.\sandboxtag\dashv\comm
-  }{
-    \estepto[]{\rtt{\statevar}{\ectx[']\hole{\lreturn{\valueexpr}}}}{\rtt{\statevar\subst{\neg\statevar.\sandboxtag}{\sandboxtag}\subst{\ectxstack[']}{\ectxstack}}{\ectx\hole{\valueexpr}}}{\tup{\ev{Ret}\ \comm\ \valueexpr}{\statevar.\sandboxtag}}
-  }{e-ret-notsame}
-  %
-  \typerule{$e-\text{call}-\text{notsame}$}{
-    \lfunction{\text{foo}}{x}{e} \in \statevar.\library &
-    \commlib\vdash\text{foo}:\Omega.\sandboxtag\dashv\comm
-  }{
-    \estepto[]{\rtt{\statevar}{\ectx\hole{\lcall{\text{foo}}{\valueexpr}}}}{\rtt{\statevar\subst{\neg\statevar.\sandboxtag}{\sandboxtag}\subst{\tup{\ectx}{\text{foo}},\statevar.\ectxstack}{\ectxstack}}{\expr\subst{\valueexpr}{x}}}{\tup{\ev{Call}\ \comm\ \text{foo}\ \valueexpr}{\statevar.\sandboxtag}}
-  }{e-call-notsame}
-\end{center}
-
-The contextual semantics is mostly straightforward.
-In \Cref{tr:e-ret-notsame,tr:e-call-notsame}, the judgement $\Omega.\commlib\vdash\text{foo}:\Omega.\sandboxtag\dashv \comm$ checks whether $\text{foo}$ is a component-level name by looking it up in the list of component-level names $\commlib$ and emits the appropriate transfer tag, i.e., either $\comptoctx$ or $\ctxtocomp$.
-Additional rules that are left out ensure that, e.g., when calling the $\text{main}$ function, the event $\ev{Start}$ is emitted, which is a design choice this paper does for convenience when reasoning about call-chains.
-Note that the $\ev{End}\ \valueexpr$ event is not emitted if the program crashes.
-
-% \MKin{%
-%   should I add a rule for top level exec or keep it like this?
-% 	MP: like this is fine
-% }
-The top-level execution $\progstepto{\prog{\library_{\ctx}}{\library_{\comp}}}{\runtimetermvar}{\trace}$ constructs an initial state $\Omega$ by linking $\library_{\ctx}$ and $\library_{\ctx}$ and then starts execution by calling the $\text{main}$ function.
-The trace $\trace$ emitted during that execution serves as abstraction of the behavior of the program enabling the use of \Cref{def:propsat,def:proprsat}.
-
-\subsection{$\src{L_{\tmssafe}}$: A Temporal but Not Spatial Memory Safe Language}\label{subsec:ltms}
-
-$\src{L_{\tmssafe}}$ uses the same syntax as presented earlier (\Cref{sec:casestud:defs}) without extensions to the term level.
-But, $\src{L_{\tmssafe}}$ is statically typed, where the type system is inspired by $L^{3}$~\cite{morrisett2005L3,scherer2018fabulous}.
-The type system of $\src{L_{\tmssafe}}$ exhibits the property that every well-typed $\src{L_{\tmssafe}}$ program satisfies \gls*{tms} (\Cref{thm:wt:tms}).
-The proof of this theorem relies on a projection $\msfilterLtms{\src{\event}}=\event$ from $\src{L_{\tmssafe}}$ events to specification events $\event$, because the properties defined earlier (\Cref{subsec:propdefs}) are defined in the specification trace model.
-Hereby, the map $\thelocmap(\src{\loc})=\loc$ maps an $\Ltms$ location to a location $\loc$ of the specification trace model.
-
-\begin{center}
-  \judgbox{\msfilterLtms{\src{\event}}=\event}{,,Project an $\Ltms$ event $\src{\event}$ to a specification event $\event$.''}
-  \typerule{$\Ltms$-filter-context}{
-    \src{\event[_b]}\not=\src{\lightning}
-  }{
-    \msfilterLtms{\src{\tup{\event[_b]}{\ctx,\securitytag}}} = \emptyevent
-  }{ltms-filter-context}
-  %
-  \typerule{$\Ltms$-filter-abort}{
-  }{
-    \msfilterLtms{\src{\lightning}} = \lightning
-  }{ltms-filter-abort}
-  %
-  \typerule{$\Ltms$-filter-start}{
-  }{
-    \msfilterLtms{\src{\tup{Start}{\comp}}} = \emptyevent
-  }{ltms-filter-start}
-  %
-  \typerule{$\Ltms$-filter-alloc}{
-    \locmapsto{\src{\loc}}{\loc} &
-    \src{n}=n
-  }{
-    \msfilterLtms{\src{\tup{Alloc\ \loc\ n}{\comp}}} = \tup{\ev{Alloc}\ \loc\ n}{\comp,\unlock}
-  }{ltms-filter-alloc}
-\end{center}
-
-Most rules of the projection $\msfilterLtms{\src{\event}}$ are left out since, for the most part, it does the expected, e.g., $\msfilterLtms{\src{\ev{Dealloc}\ \loc;\comp;\securitytag}}=\ev{Dealloc}\ \delta(\loc);\comp;\securitytag$.
-But, it also filters any action that a context does as well as the interaction events, since these are irrelevant for component-level \gls*{tms}.
-
-\begin{theorem}[$\src{L_{\tmssafe}}$-programs are \gls*{tms}]\label{thm:wt:tms}
-  For any $\src{\library_{\comp}}$, $\rsat{\src{\library_{\comp}}}{\tmssafe}$ \Coqed
-\end{theorem}
-
-\subsection{$\Ltrg$: A Memory-Unsafe Language}\label{subsec:lsms}
-
-\begin{center}
-  \begin{align*}
-    \mi{(Expressions)}~\trg{\expr}&\bnfdef \cdots \mid \trg{\lispoisoned{x}} \mid \trg{\lhast{\expr}{\type}} \\
-  \end{align*}
-\end{center}
-
-$\Ltrg$ extends the syntax presented earlier (\Cref{subsec:cs:defs}) with dynamic typechecks $\trg{\lhast{\expr}{\type}}$ and a way to inspect poison tags $\trg{\lispoisoned{x}}$ in the metadata of pointers.
-For valid pointers ($\trg{\poisonless}$) bound to variable $\trg{x}$, the check $\trg{\lispoisoned{x}}$ yields $\trg{1}$.
-If the array bound to $\trg{x}$ was deallocated, i.e., has been poisoned ($\trg{\poisoned}$), the check $\trg{\lispoisoned{x}}$ evaluates to $\trg{0}$.
-
-\begin{center}
-  \typerule{$e-\trg{\lhast{x}{\natt}}$}{
-  }{
-    \pstepto{\rtt{\trg{\statevar}}{\trg{\lhast{x}{\natt}}}}{\rtt{\trg{\statevar}}{\trg{1}}}{\emptyevent}
-  }{e-has-nat-no}
-  %
-  \typerule{$e-\trg{\lhast{n}{\natt}}$}{
-  }{
-    \pstepto{\rtt{\trg{\statevar}}{\trg{\lhast{n}{\natt}}}}{\rtt{\trg{\statevar}}{\trg{0}}}{\emptyevent}
-  }{e-has-nat-yes}
-\end{center}
-
-Dynamic typechecks $\trg{\lhast{\expr}{\type}}$ match on $\trg{\expr}$ and evaluate to $\trg{0}$ if the term is of type $\trg{\type}$ and $\trg{1}$ otherwise.
-
-%
-The projection $\msfilterL{\trg{\trace}}$ is equal to $\msfilterLtms{\src{\trace}}$.
-
-\subsection{$\Lms$: Another Memory-Unsafe Language}\label{subsec:lms}
-To enhance readablity, this paper uses $\Lms$, despite it being exactly equal to $\Ltrg$ (\Cref{subsec:lsms}).
-The projection $\msfilterLms{\irl{\trace}}$ is also exactly equal to $\msfilterL{\trg{\trace}}$.
-
-\subsection{$\Lscct$: A Memory-Unsafe Language with a Data Independent Timing Mode}\label{subsec:lscct}
-
-\begin{gather*}
-  \begin{aligned}
-    \mi{(Expressions)}~\obj{\expr}&\bnfdef \obj{n^{\securitytag}} \mid \cdots \mid \obj{\llet{x^{\securitytag}}{\expr[_{1}]}{\expr[_{2}]}} \mid \cdots \mid \obj{\lwrdoit{D}} \mid \obj{\lrddoit{x}{\expr}}
-  \end{aligned} 
-  \\
-  \begin{aligned}
-    \mi{(States)}~\obj{\statevar}&\bnfdef \obj{\tup{\cfstate}{\sandboxtag,D,\memstate}}
-     &
-    \mi{(DIT\ Flag)}~\obj{D}&\bnfdef \obj{ON} \mid \obj{OFF}
-  \end{aligned}
-\end{gather*}
-
-$\Lscct$ extends $\Lms$ (\Cref{subsec:lms}) with a way to write to a model specific register that controls a data (operand) independent timing mode, a feature that is present in both Arm~\cite[p.~543]{arm-refman} and Intel~\cite[p.~80]{intel-refman} processors.
-To this end, states are extended with the value of the register, which is initially set to be not active.
-If the register is marked active, the intuition is that no secrets can appear on specification traces.
-If the register is marked inactive, secrets may appear on traces.
-For the other languages seen earlier, the mode is intuitively always-on, i.e., the mode of execution always uses data independent timing.
-The language also adds user-annotations to values and variables to know their secrecy $\obj{\securitytag}$, which is either high $\obj{\lock}$ or low $\obj{\unlock}$.
-Security tags $\obj{\securitytag}$ are on the usual secrecy lattice, where $\obj{\lock}\le\obj{\securitytag}$ and $\obj{\securitytag}\le\obj{\unlock}$.
-
-\begin{center}
-  \begin{align*}
-    \mi{(Pre\text{-}Events)}~\obj{\event[_{b}]}&\bnfdef \cdots \mid \obj{\widehat{Get}\ \loc\ \valueexpr} \mid \obj{\widehat{Set}\ \loc\ \valueexpr\ \valueexpr[']} \mid \obj{Branch\ n} \mid \obj{Binop\ n}
-  \end{align*}
-\end{center}
-
-To prevent secrets from leaking but still enable reasoning about memory safety, $\Lscct$ extends pre-events with $\obj{\widehat{Get}\ \loc\ \valueexpr}$ and $\obj{\widehat{Set}\ \loc\ \valueexpr}$.
-These indicate reads from and writes to memory without leaking secret information involved in the access.
-Moreover, the language extends pre-events with $\obj{\ev{Branch}\ n}$ and $\obj{\ev{Binop}\ n}$ that are emitted when evaluating a branch or certain binary expressions, such as division, respectively, whenever the data independent timing mode is inactive.
-The following rules demonstrate how this is handled semantically.
-
-\begin{center}
-  \typerule{$e-\obj{\oplus}-\text{noleak}$}{
-    \obj{m}\not=\obj{0} &
-    \obj{n_3}=\obj{n_1}\oplus\obj{n_2} \\
-    \obj{\securitytag['']}\le\obj{\securitytag} &
-    \obj{\securitytag['']}\le\obj{\securitytag[']}
-  }{
-    \pstepto{\rtt{\obj{\cfstate;\sandboxtag;m;\memstate}}{\obj{\lbinop{n_1^{\securitytag}}{n_2^{\securitytag[']}}}}}{\rtt{\obj{\cfstate;\sandboxtag;m;\memstate}}{\obj{n_3^{\securitytag['']}}}}{\obj{\emptyevent}}
-  }{e-binop-noleak}
-  %
-  \typerule{$e-\obj{wrdoit}-\text{off}$}{
-  }{
-    \pstepto{\rtt{\obj{\cfstate;\sandboxtag;m;\memstate}}{\obj{\lwrdoit{OFF}}}}{\rtt{\obj{\cfstate;\sandboxtag;OFF;\memstate}}{\obj{0^{\securitytag}}}}{\obj{\emptyevent}}
-  }{e-wrdoit-off}
-  %
-  \typerule{$e-\obj{ifz}-\obj{true}-\text{leak}$}{
-  }{
-    \pstepto{\rtt{\obj{\cfstate;\sandboxtag;0;\memstate}}{\obj{\lifz{0^{\securitytag}}{\expr[_1]}{\expr[_2]}}}}{\rtt{\obj{\cfstate;\sandboxtag;0;\memstate}}{\obj{\expr[_2]}}}{\obj{\ev{Branch}\ 0;\sandboxtag;\securitytag}}
-  }{e-ifz-true-leak}
-\end{center}
-
-The evaluation steps are amended to propagate the security-tag annotations $\obj{\securitytag}$.
-When the data independent timing mode is inactive, pre-events $\obj{\ev{Branch}\ n}$ and $\obj{\ev{Binop}\ n}$ are emitted for conditionals and binary operations, respectively.
-%To mark this mode as active or inactive, \Cref{tr:e-wrdoit} allows to access the register in $\obj{\statevar}$.
-
-\begin{center}
-  \judgbox{\scctfilterLscct{\obj{\event}}=\event}{,,Project an $\Lscct$ event $\obj{\event}$ to a specification event $\event$.''}
-  \typerule{$\Lscct$-filter-context}{
-    \obj{\event[_b]}\not=\obj{\lightning}
-  }{
-    \scctfilterLscct{\obj{\tup{\event[_b]}{\ctx,\securitytag}} = \emptyevent}
-  }{lscct-filter-context}
-  %
-  \typerule{$\Lscct$-filter-$\widehat{\text{get}}$}{
-    \locmapsto{\obj{\loc}}{\loc} &
-    \obj{n}=n
-  }{
-    \scctfilterLscct{\obj{\tup{\widehat{Get}\ \loc\ n}{\comp,\securitytag}}} = \tup{\ev{Get}\ \loc\ n}{\comp,\unlock}
-  }{lscct-filter-get-hat}
-  %
-  \typerule{$\Lscct$-filter-$\text{get}$}{
-    \locmapsto{\obj{\loc}}{\loc} &
-    \obj{n}=n
-  }{
-    \scctfilterLscct{\obj{\tup{Get\ \loc\ n}{\comp,\securitytag}}} = \tup{\ev{Get}\ \loc\ n}{\comp,\securitytag}
-  }{lscct-filter-get}
-\end{center}
-
-The projection to the specification trace model is mostly straightforward and similar to the others, e.g., \Cref{subsec:cs:tms}.
-However, for events containing the pre-events $\obj{\widehat{Get}}$ and $\obj{\widehat{Set}}$, the projection always translates the security-tag $\obj{\sigma}$ to $\unlock$, regardless of its actual value, as seen in \Cref{tr:lscct-filter-get-hat}.
-The pre-events themselves still translate to just $\ev{Get}$ and $\ev{Set}$, respectively.
-With this technical setup, the information whether a read or write happened on a secret value is not hidden by the semantics, e.g., by emitting $\obj{\emptyevent}$, but when projecting to specification events.
-This allows flexibility: The trace can be checked to satisfy different properties, such as, in this case, \gls*{tms}, \gls*{sms}, \gls*{scct}, and their combined versions.
-\Cref{ex:lscct} illustrates the differences of $\Lscct$ compared to the other languages.
-
-\begin{example}[$\Lscct$ with and without data independent timing]\label{ex:lscct}
-  Consider again the context presented in \Cref{ex:strncpy:cpp11}, where everything is marked with a security tag of high $\lock$.
-  The following table shows parts of the execution trace, read from top to bottom, in the left column (\textit{Active}) with and in the right column (\textit{Inactive}) without data independent timing.
-  The left side of the table ($\Lscct$), i.e., the two columns on the left, describes the execution trace of the program, while the right side of the table ($Spec$), i.e., the two columns on the right, describes the respective projections $\scctfilterLscct{\obj{\trace}}$ to the specification trace model.
-
-  $${\footnotesize
-  \begin{array}{ccccccc}
-        & \Lscct &        &   &      & Spec & \\
-    Active & \mid & Inactive & \mid & Active & \mid & Inactive \\\hline
-    \obj{\ev{Alloc}\ \loc_{x}\ 12;\comp;\unlock} & \mid & \obj{\ev{Alloc}\ \loc_{x}\ 12;\comp;\unlock} & \mid & \ev{Alloc}\ \loc_{x}\ 12;\comp;\unlock & \mid & \ev{Alloc}\ \loc_{x}\ 12;\comp;\unlock\\
-    \obj{\ev{Alloc}\ \loc_{y}\ 12;\comp;\unlock} & \mid & \obj{\ev{Alloc}\ \loc_{y}\ 12;\comp;\unlock} & \mid & \ev{Alloc}\ \loc_{y}\ 12;\comp;\unlock & \mid & \ev{Alloc}\ \loc_{x}\ 12;\comp;\unlock\\
-    \obj{\widehat{\ev{Get}}\ \loc_{x}\ 0;\comp;}\text{\colorbox{lightgray}{$\obj{\lock}$}} & \mid & \obj{\ev{Get}\ \loc_{x}\ 0;\comp;\lock} & \mid & \ev{Use}\ \loc_{x}\ 0;\comp;\text{\colorbox{lightgray}{$\unlock$}} & \mid & \ev{Use}\ \loc_{x}\ 0;\comp;\lock\\
-    \text{\colorbox{yellow}{$\obj{\emptyevent}$}} & \mid & \obj{\ev{Branch}\ 0;\comp;\lock} & \mid & \emptyevent & \mid & \ev{Branch}\ 0;\comp;\lock\\
-    \obj{\widehat{\ev{Get}}\ \loc_{x}\ 0;\comp;}\text{\colorbox{lightgray}{$\obj{\lock}$}} & \mid & \obj{\ev{Get}\ \loc_{x}\ 0;\comp;\lock} & \mid & \ev{Use}\ \loc_{x}\ 0;\comp;\text{\colorbox{lightgray}{$\unlock$}} & \mid & \ev{Use}\ \loc_{x}\ 0;\comp;\lock\\
-    \obj{\widehat{\ev{Set}}\ \loc_{y}\ 0\ \mathtt{'H'};\comp;}\text{\colorbox{lightgray}{$\obj{\lock}$}} & \mid & \obj{\ev{Set}\ \loc_{y}\ 0\ \mathtt{'H'};\comp;\lock} & \mid & \ev{Use}\ \loc_{y}\ 0;\comp;\text{\colorbox{lightgray}{$\unlock$}} & \mid & \ev{Use}\ \loc_{y}\ 0;\comp;\lock\\
-    \obj{\widehat{\ev{Get}}\ \loc_{x}\ 1;\comp;}\text{\colorbox{lightgray}{$\obj{\lock}$}} & \mid & \obj{\ev{Get}\ \loc_{x}\ 1;\comp;\lock} & \mid & \ev{Use}\ \loc_{x}\ 1;\comp;\text{\colorbox{lightgray}{$\unlock$}} & \mid & \ev{Use}\ \loc_{x}\ 1;\comp;\lock\\
-    \text{\colorbox{yellow}{$\obj{\emptyevent}$}} & \mid & \obj{\ev{Branch}\ 0;\comp;\lock} & \mid & \emptyevent & \mid & \ev{Branch}\ 0;\comp;\lock\\
-    \vdots & \mid & \vdots & \mid & \vdots & \mid & \vdots \\
-    \obj{\widehat{\ev{Get}}\ \loc_{x}\ 12;\comp;}\text{\colorbox{lightgray}{$\obj{\lock}$}} & \mid & \obj{\ev{Get}\ \loc_{x}\ 12;\comp;\lock} & \mid &  \ev{Use}\ \loc_{x}\ 12;\comp;\text{\colorbox{lightgray}{$\unlock$}} & \mid & \ev{Use}\ \loc_{x}\ 12;\comp;\lock\\
-    \text{\colorbox{yellow}{$\obj{\emptyevent}$}}& \mid & \obj{\ev{Branch}\ 1;\comp;\lock} & \mid &  \emptyevent & \mid & \ev{Branch}\ 1;\comp;\lock\\
-    \obj{\ev{Dealloc}\ \loc_{y};\comp;\unlock} & \mid & \obj{\ev{Dealloc}\ \loc_{y};\comp;\lock} & \mid &  \ev{Dealloc}\ \loc_{y};\comp;\unlock & \mid & \ev{Dealloc}\ \loc_{y};\comp;\unlock\\
-    \obj{\widehat{\ev{Get}}\ \loc_{y}\ 6;\comp;}\text{\colorbox{lightgray}{$\obj{\lock}$}} & \mid & \obj{\ev{Get}\ \loc_{y}\ 6;\comp;\lock} & \mid &  \ev{Use}\ \loc_{y}\ 6;\comp;\text{\colorbox{lightgray}{$\unlock$}} & \mid & \ev{Use}\ \loc_{y}\ 6;\comp;\lock\\
-  \end{array}}
-  $$
-
-  When the data independent timing mode is off, the execution yields events in similar fashion to before (\Cref{subsec:ltms,subsec:lsms,subsec:lms}).
-  But, if it is turned on, then the branching event \colorbox{yellow}{does not fire} anymore and both reading and writing to memory gets ultimately translated to a specification trace with \colorbox{lightgray}{no exposed secrets}.
-\end{example}
-
-\section{Case Study: Composing Secure Compiler Passes and Optimisations \pages{7}}\label{sec:casestud:rtp}
-This section defines several secure compilers, each of which robustly preserves a different property of interest as depicted in \Cref{fig:pipeline}.
-\begin{figure}[!h]
-  \begin{center}
-    \begin{tikzpicture}
-      \node (S) {$\src{L_{\tmssafe}}$};
-      \node[right=1.5 of S] (T) {$\trg{L}$};
-      \node[right=1.5 of T] (M) {$\irl{L_{\mssafe}}$};
-      \node[below right=0.5 and 1.0 of M] (D0) {$\irl{L_{\mssafe}}$};
-      \node[above right=0.5 and 1.0 of M] (C0) {$\irl{L_{\mssafe}}$};
-      \node[right=3.0 of M] (E) {$\irl{L_{\mssafe}}$};
-      \node[right=1.5 of E] (O) {$\obj{L_{\scctsafe}}$};
-
-      \draw[->] (S) to[sloped] node[align=center,font=\scriptsize] (tmsedge) {\gls*{tms}\\ \Cref{thm:cca:rtp:tms}} (T);
-      \draw[->] (T) to[sloped] node[align=center,font=\scriptsize] {\gls*{sms}\\ \Cref{thm:ccb:rtp:sms}} (M);
-      \draw[->] (M) to[sloped] node[align=center,font=\scriptsize] {\gls*{dce}\\ \Cref{thm:ccdce:rtp:ms}} (D0);
-      \draw[->] (M) to[sloped] node[align=center,font=\scriptsize] {\gls*{cf}\\ \Cref{thm:cccf:rtp:ms}} (C0);
-      \draw[->] (D0) to[sloped] node[align=center,font=\scriptsize] {\gls*{cf}\\ \Cref{thm:cccf:rtp:ms}} (E);
-      \draw[->] (C0) to[sloped] node[align=center,font=\scriptsize] {\gls*{dce}\\ \Cref{thm:ccdce:rtp:ms}} (E);
-      \draw[->] (E) to[sloped,above] node[align=center,font=\scriptsize] {\gls*{scct}\\ \Cref{thm:ccscct:rtp:scct}} (O);
-
-      % Sections
-      \node[above=1.0 of tmsedge] (sectms) {{\scriptsize\Cref{subsec:cs:tms}}};
-      \node[right=0.5 of sectms] (secsms) {{\scriptsize\Cref{subsec:cs:ms}}};
-      \node[right=1.75 of secsms] (secopts) {{\scriptsize\Cref{subsec:cs:opts}}};
-      \node[right=1.75 of secopts] (secscct) {{\scriptsize\Cref{subsec:cs:scct}}};
-
-      \draw[thick,dotted,->] (S) to[bend right=33,sloped] node[align=center,font=\scriptsize] {\gls*{ms}\\ \Cref{thm:ccab:rtp:ms}} (M);
-      \draw[thick,dotted,->] (M) to[bend right=0,sloped] node[align=center,font=\scriptsize] {\gls*{ms}\\ \Cref{thm:cccfccdce:rtp:ms}} (E);
-      \draw[thick,dotted,->] (S) to[out=-45,in=198,sloped] node[align=center,font=\scriptsize] {\gls*{ms}+\gls*{scct}\\ \Cref{thm:ccall:rtp:msscct}} (O);
-
-      \draw[dashed] ($(sectms)-(0.8,0)$) -- ($(sectms)-(0.8,1.25)$);
-      \draw[dashed] ($(sectms)-(-0.9,0)$) -- ($(sectms)-(-0.9,1.25)$);
-      \draw[dashed] ($(secsms)-(-1.2,0)$) -- ($(secsms)-(-1.2,1.25)$);
-      \draw[dashed] ($(secscct)-(1.2,0)$) -- ($(secscct)-(1.2,1.25)$);
-      \draw[dashed] ($(secscct)-(-1.1,0)$) -- ($(secscct)-(-1.1,1.25)$);
-    \end{tikzpicture}
-  \end{center}
-  \caption{Visualisation of the optimising compilation pipeline that attains a combination of \gls*{ms} and \gls*{cct}. %
-    Vertices in the graph are the programming languages from earlier sections (\Cref{sec:casestud:defs}). %
-    All edges are secure compilers, but dotted edges use the presented framework (\Cref{sec:compcomp}) and solid edges classic proof techniques. %
-    The dashed lines partition the graph into the sections where the respective theorems are presented.
-  }\label{fig:pipeline}
-\end{figure}
-The section demonstrates the power of the framework (\Cref{sec:compprop,sec:compcomp}) by composing these compilers for a secure and optimising compilation chain that robustly preserves \gls*{msscct}.
-The first step in this chain is the compiler from $\src{L_{\tmssafe}}$ to $\trg{L}$ that robustly preserves just \gls*{tms} (\Cref{thm:cca:rtp:tms}).
-From here, an instrumentation from $\trg{L}$ to $\irl{L_{\mssafe}}$ ensures that no out-of-bounds accesses can happen and, thus, programs at this point attain \gls*{sms} (\Cref{thm:ccb:rtp:sms}).
-Since these properties compose into \gls*{ms}, composing these passes yields a compiler that robustly preserves \gls*{ms} (\Cref{thm:ccab:rtp:ms}).
-At this stage, the section presents two optimising translations, namely \gls*{cf} and \gls*{dce}, each of which robustly preserves \gls*{ms} (\Cref{thm:ccdce:rtp:ms,thm:cccf:rtp:ms}).
-These translations can be freely ordered in the compilation chain without compromising memory safety (\Cref{thm:cccfccdce:rtp:ms}).
-The last step of the chain ensures that code stays \gls*{scct} (\Cref{thm:ccscct:rtp:scct}) when lowered from $\Lms$ to $\Lscct$.
-The final result is that the whole compilation chain robustly preserves \gls*{msscct} (\Cref{thm:ccall:rtp:msscct}).
-
-\subsection{Robust Temporal Memory Safety Preservation}\label{subsec:cs:tms}
-
-This subsection defines a secure compiler from $\Ltms$ to $\Ltrg$.
-To this end, the compiler needs to ensure that when execution switches from context to component, the type signatures are respected.
-It can do so by inserting dynamic typechecks prior to entering the body of a function belonging to the component.
-
-\begin{gather*}
-  \begin{aligned}
-    \cca(\src{x}) & = \trg{x} \\
-    \cca(\src{n}) & = \trg{n} \\
-    \cca(\src{\lbinop{\expr[_{1}]}{\expr[_{2}]}}) & =  \lbinop[\trg]{\left[\cca(\src{\expr[_{1}]})\right]}{\left[\cca(\src{\expr[_{2}]})\right]} \\
-    \cca(\src{\lget{x}{\expr}}) & = \lget[\trg]{\trg{x}}{\left[\cca(\src{\expr})\right]} \\
-    \cca(\src{\ldelete{x}}) & = \ldelete[\trg]{\left[\cca(\src{x})\right]} \\
-%    \cca(\src{\lset{x}{\expr[_{1}]}{\expr[_{2}]}}) & = \lset[\trg]{\left[\cca(\src{x})\right]}{\left[\cca(\src{\expr[_{1}]})\right]}{\left[\cca(\src{\expr_{2}})\right]} \\
-    \cca(\src{\lfunction{g}{x:\natt\to\type[_{e}]}{\expr}}) & = \lfunction[\trg]{\trg{g}}{\trg{x}}{\lifz[\trg]{\trg{\lhast{x}{\natt}}}{%
-                                                                                                \left[\cca(\src{\expr})\right] %
-                                                                                                 }{\labort[\trg]}}
-  \end{aligned}
-\end{gather*}
-
-Since $\trg{L}$ has no static typechecks, it could happen that a bogus context $\trg{\library_{\ctx}}$ invokes a callable object accepting a $\src{\natt}$ with $\trg{\lpair{17}{29}}$.
-By inserting the check, the compiler ensures that execution does not proceed in such cases.
-The compiler does not insert other checks and proceeds as the identity function (which in this paper amounts to a simple re-colouring of $\src{L_{\tmssafe}}$ to $\trg{L}$ expressions).
-
-Compiling the \texttt{strncpy} function from \Cref{sec:introduction} with $\cca$, the compiler would in this case ensure that the arguments that are evaluated in the compiled \texttt{strncpy} are valid.
-
-\begin{theorem}[Compiler $\cca$ is secure with respect to \gls*{tms}]\label{thm:cca:rtp:tms}
-  $\rtp{\cca}{\tmssafe}$ % \Coqed
-\end{theorem}
-
-\subsubsection{Proving Robust Safety Property Preservation}
-
-We illustrate the proof of \Cref{thm:cca:rtp:tms} since the other secure compilation proofs of this paper follow the same approach.
-Unfolding the theorem statement yields the following assumptions:  
-	for any $\pi\in\lceil\tmssafe\rceil$\footnote{$\lceil\cdot\rceil$ lifts the property to a hyperproperty by applying the powerset operation~\cite{clarkson2008hyper}.}, 
-	$\src{\trace}$, 
-	$\src{\runtimetermvar}$, 
-	and component $\src{\library_{comp}}$, 
-	we have that $\rsat{\src{\library_{comp}}}{\pi}$ and 
-	$\progstepto{\prog[\trg]{\trg{\library_{\ctx}}}{\cca(\src{\library_{\comp}})}}{\trg{\runtimetermvar}}{\trg{\trace}}$, where $\trg{\library_{\ctx}}$ is arbitrary.
-The proof obligation is $\msfilterL{\trg{\trace}}\in\pi$, i.e., the specification trace associated to $\trg{\trace}$ satisfies the property $\pi$.
-A way to show this is to relate trace $\trg{\trace}$ to some $\Ltms$ trace $\src{\trace}$ (which already satisfies the property as per the assumptions).
-The assumptions already contain a target execution associated to this trace, so the task is to find an associated $\Ltms$ execution that yields $\trg{\trace}$.
-The trace $\trg{\trace}$ is split into different parts, as commonly done in secure compilation works~\cite{korashy2021capableptrs,abate2018goodcomps}, where each part contains the events that either the context or the component does, but not both.
-Because of this, all such trace segments are ,,well-bracketed'' in the sense that they start with either $\trg{\ev{Start}}$, $\trg{\ev{Call}\ \comptoctx\ foo\ \valueexpr}$, or $\trg{\ev{Ret}\ \comptoctx\ \valueexpr}$ and end with either $\trg{\ev{End}\ \valueexpr}$, $\trg{\ev{Ret}\ \ctxtocomp\ \valueexpr}$, or $\trg{\ev{Call}\ \ctxtocomp\ foo\ \valueexpr}$.
-In the following, the former is referred to as a context segment, since these executions happen in $\trg{\library_{\ctx}}$, and the latter is referred to as a component segment, since these executions happen in $\cca(\src{\library_{\comp}})$.
-\Cref{fig:proofdiag:rtp} visualises this division for a program execution with one call from context to component and how the target execution is related to a source execution.
-In the figure, the green dashed lines encompass the component segments while the orange boxes contain the actual context switches from context to component or vice versa.
-From a technical perspective, as typically done in compilation proofs, the proof requires some setup to maintain a relation between $\src{\Omega}$ and $\trg{\Omega}$ .
-Two cross-language relations make this precise: (i) $\multimap_{\delta}$ relates states that are involved in a context segment, allowing the target execution to perform internal calls, and (ii) $\approx_{\delta}$ relates states that are involved in a component segment, where both states need to agree exactly, i.e., the memory and the control flow states are required to contain the same information.
-The relations are indexed with $\delta$, which is an injective mapping from $\Ltms$ locations $\src{\loc}$ to $\Ltrg$ locations $\trg{\loc}$.
-Note that the relations $\multimap_{\delta}$ and $\approx_{\delta}$ swap when context switching.
-
-\begin{figure}[!ht]
-  \begin{center}
-    \begin{tikzpicture}[state/.style={minimum height=1.5cm}]
-      % relative horizontal/vertical distance between states
-      \pgfmathsetmacro{\hdist}{0.95}
-      \pgfmathsetmacro{\vdist}{1.35}
-      \pgfmathsetmacro{\halfvdist}{0.725}
-
-      % row of src states
-      \node[state] (srcempty) {$\rtt{\src{\emptyset}}{\src{\lcall{main}{0}}}$};
-      \foreach \s [remember=\s as \cur (initially empty)] in {1,w_1,p,w_2,2} {
-        \node[state,right=\hdist of src\cur] (src\s) {$\rtt{\src{\Omega_{\s}}}{\src{\expr[_{\s}]}}$};
-      }
-      % row of trg states
-      \node[state,below=\vdist of srcempty] (trgempty) {$\rtt{\trg{\emptyset}}{\trg{\lcall{main}{0}}}$};
-      \foreach \s in {1,w_1,p,w_2,2} {
-        \node[state,below=\vdist of src\s] (trg\s) {$\rtt{\trg{\Omega_{\s}}}{\trg{\expr[_{\s}]}}$};
-      }
-      %% illustrations
-        % backtrans wrapper 1
-        \draw[ultra thick,loosely dotted,Peach!50,rounded corners] (src1.north east) -- (srcw\string_1.north east)
-          -- (trgw\string_1.south east) -- (trg1.south west) -- (src1.north west) -- cycle;
-        % backtrans wrapper 2
-        \draw[ultra thick,loosely dotted,Peach!50,rounded corners] (srcp.north east) -- (srcw\string_2.north east)
-          -- (trgw\string_2.south east) -- (trgp.south west) -- (srcp.north west) -- cycle;
-        % compiler correctness
-        \draw[ultra thick,loosely dashed,Emerald!50,rounded corners] ($(srcw\string_1.north east)+(0,0.05)$) -- ($(srcp.north east)+(0,0.05)$)
-          -- ($(trgw\string_2.north east)+(0.05,0)$) -- ($(trgw\string_2.south east)+(0.05,-0.05)$)
-          -- ($(trg1.south west)+(-0.05,-0.05)$) -- ($(trg1.north west)+(-0.05,0)$) -- ($(srcw\string_1.north west)+(0,0.05)$) -- cycle;
-      % state relations
-      \path (srcempty) edge[very thick,draw=gray!25] node[pos=0.5,sloped,rotate=180,fill=white] {\scriptsize$\multimap_\emptyset$} (trgempty)
-        (src1) edge[very thick,draw=gray!25] node[pos=0.5,sloped,rotate=180,fill=white] {\scriptsize$\multimap_{\delta_1}$} (trg1)
-        (srcp) edge[very thick,draw=gray!25] node[pos=0.5,sloped,fill=white] {\scriptsize$\approx_{\delta_p}$} (trgp)
-        (src2) edge[very thick,draw=gray!25] node[pos=0.5,sloped,rotate=180,fill=white] {\scriptsize$\multimap_{\delta_2}$} (trg2)
-        (srcw\string_1) edge[very thick,draw=gray!25] node[pos=0.5,sloped,fill=white] {\scriptsize$\approx_{\delta_{w_1}}$} (trgw\string_1)
-        (srcw\string_2) edge[very thick,draw=gray!25] node[pos=0.5,sloped,rotate=180,fill=white] {\scriptsize$\multimap_{\delta_{w_2}}$} (trgw\string_2)
-        % diagonals
-        (srcw\string_1) edge[thick,draw=gray!25] node[pos=0.16,sloped,rotate=180,fill=white] {\scriptsize$\approx_{\delta_{w_1}}$} (trg1)
-        (src1) edge[thick,draw=gray!25] node[pos=0.16,sloped,rotate=180,fill=white] {\scriptsize$\multimap_{\delta_1}$} (trgw\string_1)
-        (srcp) edge[thick,draw=gray!25] node[pos=0.2,sloped,fill=white] {\scriptsize$\approx_{\delta_p}$} (trgw\string_2)
-        (trgp) edge[thick,draw=gray!25] node[pos=0.2,sloped,fill=white] {\scriptsize$\multimap_{\delta_{w_2}}$} (srcw\string_2)
-        ;
-      %\drawpolygon src1,srcw\string_1,trgw\string_1,trg1;
-      %\drawpolygon srcp,srcw\string_2,trgw\string_2,trgp;
-      %\node[font=\tiny,align=center,above=0.2 of srcw1srcp] (wrapper) {Backtranslation\\Wrapper};
-      %\path[->,draw] (wrapper) -- (srcw\string_1);
-      %\path[->,draw] (wrapper) -- (srcp);
-      % steps
-      \path[very thick,color=\stlccol] (srcempty) edge[draw=none] node {\ $\xrightarrow{\trace[_1]}{}{\kern-3.5pt}_{\text{ctx}}^*$} (src1)
-        (src1) edge[draw=none] node {\ $\xrightarrow{\trace[_c]}{}{\kern-3.5pt}_{\text{ctx}}$} (srcw\string_1)
-        (srcw\string_1) edge[draw=none] node {\ $\xrightarrow{\trace[_p]}{}{\kern-3.5pt}_{\text{ctx}}^*$} (srcp)
-        (srcp) edge[draw=none] node[fill=white,inner sep=0,outer sep=0] {\ $\xrightarrow{\trace[_r]}{}{\kern-3.5pt}_{\text{ctx}}$} (srcw\string_2)
-        (srcw\string_2) edge[draw=none] node {\ $\xrightarrow{\trace[_2]}{}{\kern-3.5pt}_{\text{ctx}}^*$} (src2)
-        ;
-      \path[very thick,color=\ulccol] (trgempty) edge[draw=none] node[fill=white,inner sep=0,outer sep=0] {\ $\xrightarrow{\phantom{\trace[_1]}}{}{\kern-3.5pt}_{\text{ctx}}^*$} (trg1)
-        (trg1) edge[draw=none] node {\ $\xrightarrow{\phantom{\trace[_p]}}{}{\kern-3.5pt}_{\text{ctx}}^*$} (trgw\string_1)
-        (trgw\string_1) edge[draw=none] node[fill=white,inner sep=0,outer sep=0] {\ $\xrightarrow{\phantom{\trace[_p]}}{}{\kern-3.5pt}_{\text{ctx}}^*$} (trgp)
-        (trgp) edge[very thick,draw=none] node {\ $\xrightarrow{\phantom{\trace[_p]}}{}{\kern-3.5pt}_{\text{ctx}}^*$} (trgw\string_2)
-        (trgw\string_2) edge[draw=none] node {\ $\xrightarrow{\phantom{\trace[_2]}}{}{\kern-3.5pt}_{\text{ctx}}^*$} (trg2)
-        ;
-      % legend
-%     \node[align=left,below right=0.3 and 0.3 of trgempty,font=\tiny] (legend) {%
-%       ${\src{\trace[_c]}}\cong{\backt{\Ltms}{\Ltrg}(\trg{Call\ ?\ foo\ \valueexpr})}$\\%
-%       ${\src{\trace[_r]}}\cong{\backt{\Ltms}{\Ltrg}(\trg{Ret\ !\ foo\ \valueexpr})}$
-%       };
-      \node[below=0.5 of trgempty,draw=none] (legend) {};
-      \draw[ultra thick,loosely dotted,Peach!50,rounded corners] ($(legend.north east)+(0.5,-0.4)$) -- ($(legend.north east)+(1,-0.4)$);
-      \node at ($(legend.north east)+(1.5,-0.4)$) (legendwrapper) {};
-      \node[right of=legendwrapper] {\tiny Backtranslation Wrapper};
-      \draw[ultra thick,loosely dashed,Emerald!50,rounded corners] ($(legend.south east)+(0.5,0.4)$) -- ($(legend.south east)+(1,0.4)$);
-      \node at ($(legend.south east)+(1.5,0.4)$) (legendcorrectness) {};
-      \node[right of=legendcorrectness] {\tiny Compiler Correctness};
-    \end{tikzpicture}
-    \caption{Proof diagram for \Cref{thm:cca:rtp:tms} depicting the general structure of robust preservation proofs. %
-      Nodes in the graph represent runtime states. %
-      Vertical lines indicate cross language relations, while horizontal ones are execution steps. %
-      The green dashed trapezoid encompasses the component segment, while the orange dotted rectangles entail the context switches. %
-      $\Ltrg$ traces are omitted for readability. %
-      $\Ltms$ trace segments $\src{\trace[_{c}]}$ and $\src{\trace[_{r}]}$ describe the events that happen at the boundaries, i.e., during a context switch.
-      $\src{\trace_{p}}$ is the behavior of the component and the traces $\src{\trace[_{1}]}$ and $\src{\trace[_{2}]}$ describe the context.
-    }\label{fig:proofdiag:rtp}
-  \end{center}
-\end{figure}
-
-So far, the paper explained how to relate an $\Ltms$ execution with a $\Ltrg$ execution.
-The next question is therefore how to build the corresponding $\Ltms$ execution.
-This is done using a standard secure compilation proof technique called trace-based backtranslation~\cite{abate2019jour,korashy2021capableptrs,patrignani2021rsc}, which can be used to build a context $\src{\library_{\ctx}}$ that behaves similar to $\trg{\library_{\ctx}}$.
-% Since $\Ltms$ is statically typed, the context $\src{\library_{\ctx}}$ obtained from the backtranslation needs to be well-typed when linked with $\src{\library_{\comp}}$.
-For context segments of the trace $\trg{\trace}$ it is also necessary to show that the execution behaves similarily, i.e., the context obtained from the backtranslation generates trace $\src{\trace}$.
-For component segments of the trace, the relatedness of states and traces follows from a compiler correctness argument.
-These two arguments yield the source execution $\progstepto{\src{\prog{\library_{\ctx}}{\library_{\comp}}}}{\src{\runtimetermvar}}{\src{\trace}}$.
-
-The proof now works as follows.
-Given that $\msfilterL{\trg{\trace}}=\msfilterLtms{\src{\trace}}$, the proof goal changes from $\msfilterL{\trg{\trace}}\in\pi$ to $\msfilterLtms{\src{\trace}}\in\pi$.
-This follows by specializing the robust satisfaction assumption $\rsat{\src{\library_{\comp}}}{\pi}$ to use the context $\src{\library_{\ctx}}$, which is obtained from the backtranslation, and to use the source execution $\progstepto{\src{\prog{\library_{\ctx}}{\library_{\comp}}}}{\src{\runtimetermvar}}{\src{\trace}}$.
-
-% \MPpost{
-% 	this can be made more precise
-% }
-
-\subsection{Robust (Spatial) Memory Safety Preservation}\label{subsec:cs:ms}
-
-\begin{center}
-  $$
-  \begin{array}{rcl}
-    \ccb(\trg{\lnew{x}{\expr[_{1}]}{\expr[_{2}]}}) & = & \llet[\irl]{\irl{x_{SIZE}}}{\ccb(\trg{\expr[_{1}]})}{\lnew[\irl]{\irl{x}}{\irl{x_{SIZE}}}{\ccb(\trg{\expr[_{2}]})}} \\
-    \ccb(\trg{\lget{x}{\expr}}) & = & \llet[\irl]{\irl{x_{n}}}{\ccb(\trg{\expr})}{\irl{\lifz{0\le x_{n}<x_{SIZE}}{\lget{x}{x_{n}}}{\labort}}} \\
-    \ccb(\trg{\lset{x}{\expr[_{1}]}{\expr[_{2}]}}) & = & \llet[\irl]{\irl{x_{n}}}{\ccb(\trg{\expr[_{1}]})}{\lifz[\irl]{\irl{0\le x_{n}<x_{SIZE}}}{\lset[\irl]{\irl{x}}{\irl{x_{n}}}{\ccb(\trg{\expr[_{2}]})}}{\irl{\labort}}} \\
-  \end{array}
-  $$
-\end{center}
-
-The compiler $\cc{\trg{L}}{\Lms}$ only inserts bounds-checks whenever reading from or writing to memory in order to enforce \gls*{sms}.
-For passing pointers, it has to pass them with their size information as well.
-To this end, the compiler introduces another, fresh identifier $\irl{x_{SIZE}}$ for each allocation that binds $\irl{x}$ to keep track of the allocation size.
-\begin{example}[Instrumented \texttt{strncpy}]
-  Consider again \texttt{strncpy}, but instrumented for \gls*{sms}:
-    \begin{lstlisting}[language=c,basicstyle=\ttfamily\small, morekeywords={size_t}]
-  void strncpy(size_t n, size_t dst_size, char *dst,
-                         size_t src_size, char *src) {
-    for(size_t i = 0; i < src_size && src[i] != '\0' && i < n; ++i) {
-      if(i < src_size && i < dst_size) {
-        dst[i] = src[i];
-      }
-    }
-  }
-    \end{lstlisting}
-    When calling this in similar fashion to \Cref{ex:strncpy:sms}, the event $\ev{Use}\ \loc_{x}\ 2;\comp;\unlock$ would not be emitted during execution, since the bounds check prevents the condition \texttt{src[i] != '\textbackslash 0'} from executing.
-\end{example}
-
-\begin{theorem}[Compiler $\ccb$ is secure with respect to \gls*{sms}]\label{thm:ccb:rtp:sms}
-  $\rtp{\ccb}{\smssafe}$ % \Coqed
-\end{theorem}
-
-\Cref{thm:ccab:rtp:ms} states that the composition of $\cca$ and $\ccb$ is secure with respect to \gls*{ms} and follows from \Cref{thm:cca:rtp:tms,thm:ccb:rtp:sms} using \Cref{thm:rtp}.
-
-\begin{theorem}[Compiler $\cca\circ\ccb$ is secure with respect to \gls*{ms}]\label{thm:ccab:rtp:ms}
-  $\rtp{\cca\circ\ccb}{\mssafe}$ % \Coqed
-\end{theorem}
-\begin{proof}
-  From \Thmref{thm:cca:rtp:tms} it follows that for any $\Ltms$ program $\src{\progvar}$, it compiles to an $\Ltrg$ program $\trg{\progvar}$ that robustly satisfies \gls*{tms}.
-  Note that $\src{\progvar}$ robustly satisfies \gls*{tms} by the properties of the typesystem of $\Ltms$.
-  Then, \Thmref{thm:ccb:rtp:sms} demonstrates that, assuming $\trg{\progvar}$ robustly satisfies \gls*{sms}, the program $\trg{\progvar}$ compiles to an $\Lms$ program $\irl{p}$ that also robustly satisfies \gls*{sms}.
-  From \Thmref{thm:ccab:rtp:ms} it follows that $\src{p}$ compiles to $\irl{p}$ that robustly satisfies \gls*{ms}, since \gls*{ms} is the intersection of \gls*{tms} and \gls*{sms}.
-\end{proof}
-
-\subsection{Optimising Compilers}\label{subsec:cs:opts}
-
-\begin{gather*}
-  \begin{align*}
-    \ccdce(\irl{\lifz{true}{\expr[_{1}]}{\expr[_{2}]}}) & = \ccdce(\irl{\expr[_{1}]}) &\\
-    \ccdce(\irl{\lifz{false}{\expr[_{1}]}{\expr[_{2}]}}) & = \ccdce(\irl{\expr[_{2}]}) &
-  \end{align*}\\
-  \begin{align*}
-    \ccdce(\irl{\lbinop{\expr[_{1}]}{\expr[_{2}]}}) & = \lbinop[\irl]{\ccdce(\irl{\expr[_{1}]})}{\ccdce(\irl{\expr[_{2}]})} &
-  \end{align*}\\[0.25cm]
-  \begin{align*}
-    \cccf(\irl{\expr}) & = \partialeval{\irl{\expr}}{\irl{\hole{\cdot}}} &
-  \end{align*}\\[0.125cm]
-  \begin{align*}
-    \partialeval{\irl{x}}{\irl{\substlist}} & = \irl{n} & \text{if } \irl{\subst{n}{x}}\in\irl{\substlist} \\
-    \partialeval{\irl{x}}{\irl{\substlist}} & = \irl{x} & \text{if } \irl{\subst{n}{x}}\notin\irl{\substlist} \\
-    \partialeval{\irl{\lbinop{n}{m}}}{\irl{\substlist}} & = \irl{k} & \text{if } \lbinop{\irl{n}}{\irl{m}}=k \\
-%    \partialeval{\irl{\lbinop{\expr[_{1}]}{\expr[_{2}]}}}{\irl{\substlist}} & = \lbinop[\irl]{\partialeval{\irl{\expr[_{1}]}}{\irl{\substlist}}}{\partialeval{\irl{\expr[_{2}]}}{\irl{\substlist}}} \\
-    \partialeval{\irl{\llet{x}{n}{\expr}}}{\irl{\substlist}} & = \partialeval{\irl{\expr}}{\irl{\subst{x}{n}},\irl{\substlist}} \\
-    \partialeval{\irl{\lget{x}{\expr}}}{\irl{\substlist}} & = \lget[\irl]{\irl{x}}{\partialeval{\irl{\expr}}{\irl{\substlist}}}
-  \end{align*}\\
-  \begin{align*}
-    \partialeval{\irl{\llet{x}{\expr[_{1}]}{\expr[_{2}]}}}{\irl{\substlist}} & = \llet[\irl]{\irl{x}}{\partialeval{\irl{\expr[_{1}]}}{\irl{\substlist}}}{\partialeval{\irl{\expr[_{2}]}}{\irl{\substlist}}} \\
-    \partialeval{\irl{\lifz{\expr[_{1}]}{\expr[_{2}]}{\expr[_{3}]}}}{\irl{\substlist}} & = \lifz[\irl]{\partialeval{\irl{\expr[_{1}]}}{\irl{\substlist}}}{\partialeval{\irl{\expr[_{2}]}}{\irl{\substlist}}}{\partialeval{\irl{\expr[_{3}]}}{\irl{\substlist}}} \\
-  \end{align*}
-\end{gather*}
-
-The two optimising compiler passes from $\Lms$ to $\Lms$ perform \gls*{dce} and \gls*{cf}, respectively.
-The \gls*{dce} pass applies a naive rewrite rule on conditionals.
-For \gls*{cf}, the pass uses an auxiliary function \texttt{mix} that does the actual work.
-It rewrites constant binary operations, e.g., $\irl{{17}-1}$ to $\irl{16}$, and replaces variables that are assigned to constants with their constant, e.g., $\irl{\llet{x}{7}{x}}$ to $\irl{7}$.
-Both passes are secure with respect to \gls*{ms}.
-The proof for either is relatively simple, because both \gls*{dce} and \gls*{cf} do not change the way memory accesses happen.
-Moreover, since the input and output languages to these compilers are the same, attacker contexts do not have more power in the target language than in the source.
-
-\begin{theorem}[Compiler $\ccdce$ is secure with respect to \gls*{ms}]\label{thm:ccdce:rtp:ms}
-  $\rtp{\ccdce}{\mssafe}$ %\Coqed
-\end{theorem}
-\begin{theorem}[Compiler $\cccf$ is secure with respect to \gls*{ms}]\label{thm:cccf:rtp:ms}
-  $\rtp{\cccf}{\mssafe}$ %\Coqe
-\end{theorem}
-
-With both \Cref{thm:ccdce:rtp:ms,thm:cccf:rtp:ms} it follows from \Cref{corr:swappable} that the two passes can be interchanged arbitrarily:
-
-\begin{theorem}[Compilers $\cccf\circ\ccdce$ and $\cccf\circ\ccdce$ are secure with respect to \gls*{ms}]\label{thm:cccfccdce:rtp:ms}
-  $\rtp{\cccf\circ\ccdce}{\mssafe}$ and $\rtp{\ccdce\circ\cccf}{\mssafe}$. % \Coqed
-\end{theorem}
-
-\subsection{Robust Strict Cryptographic Constant Time Preservation}\label{subsec:cs:scct}
-
-\begin{center}
-  \begin{align*}
-    \ccscct(\irl{\lfunction{g}{x}{\expr}}) & = \lfunction[\obj]{\obj{g}}{\obj{x}}{\obj{\lwrdoit{ON};}\ccscct(\irl{\expr})} \\
-    \ccscct(\irl{\lcall{g}{\expr}}) & = \lcall[\obj]{\obj{g}}{\ccscct(\irl{\expr})\obj{; \lwrdoit{ON}}} \\
-    \ccscct(\irl{\lbinop{\expr[_{1}]}{\expr[_{2}]}}) & = \lbinop[\obj]{\ccscct{\irl{\expr[_{1}]}}}{\ccscct{\irl{\expr[_{2}]}}} \\
-  \end{align*}
-\end{center}
-
-Given the fact that $\Lscct$ provides a \gls*{cct}-mode that can be turned on or off, the compiler inserts wrapper code for function bodies to ensure that execution in the component always happen in this \gls*{cct}-mode.
-The context can overwrite the flag and exit the mode, but upon invoking a function that is part of the component, the flag would be set again.
-Because of this, the compiler is secure with respect to \gls*{scct}, similarly proven as in \Cref{subsec:cs:tms}.
-
-\begin{theorem}[Compiler $\ccscct$ is secure with respect to \gls*{scct}]\label{thm:ccscct:rtp:scct}
-  $\rtp{\ccscct}{\scctsafe}$ % \Coqed
-\end{theorem}
-
-\subsection{Robust Preservation of Intersection of Memory Safety and Strict Cryptographic Constant Time}
-
-Let $\ccmsscct$ be the compiler that is the composition of $\cca$, $\ccb$, $\cccf$, $\ccdce$, and $\ccscct$, then the following theorem holds.
-
-\begin{theorem}[Compiler $\ccmsscct$ is secure with respect to \gls*{scct}]\label{thm:ccall:rtp:msscct}
-  $\rtp{\cc{\Ltms}{\Lscct}}{\mssafe\cap\scctsafe}$ % \Coqed
-\end{theorem}
-\begin{proof}
-	From \Thmref{thm:ccab:rtp:ms}, we have that any $\Ltms$ program $\src{\progvar}$ compiles into a $\Lms$ program $\irl{\progvar}$ that robustly satisfies \gls*{ms}.
-	Then, from \Thmref{thm:cccfccdce:rtp:ms} we have that $\irl{\progvar}$ gets optimised to a program $\irl{\progvar[']}$ that is also \gls*{ms}, where the order of optimisations does not matter for $\irl{\progvar[']}$ to be \gls*{ms}.
-  Assuming $\irl{\progvar[']}$ robustly satisfies \gls*{scct}, by \Thmref{thm:ccscct:rtp:scct} it compiles to an $\Lscct$ program $\obj{\progvar}$ that robustly satisfies \gls*{scct} as well.
-  Finally, from \Thmref{thm:rtp} it follows that, given $\src{\progvar}$ robustly satisfies \gls*{scct} and \gls*{ms}, $\obj{\progvar}$ also robustly satisfies \gls*{scct} and \gls*{ms}.
-\end{proof}
-
-\section{Related Work\pages{2}}\label{sec:relwork}
-
-This section discusses work on robust compilation (\Cref{subsec:relw:seccomprtp}) and on other secure compilation criteria (\Cref{subsec:relw:seccompcrit}).
-Since the case study of \Cref{sec:casestud:defs,sec:casestud:rtp} implements measures for preserving \gls*{ms} and \gls*{cct}, this section then presents relevant related work as well (\Cref{subsec:relw:msmechs,subsec:relw:cctmechs}).
-
-\subsection{Secure Compilation as Robust Preservation}\label{subsec:relw:seccomprtp}
-
-The robust preservation of properties as a compiler-level criterion has been analyzed extensively~\cite{abate2019jour,patrignani2021rsc,abate2021extacc,patrignani2019survey} and thus we build on that framework.
-No existing work is concerned with composing robustly safe compilers.
-These works consider languages with different trace models and our technical setup can be adapted to that as long as security properties and their monitors are still defined on the same trace model.
-The work relating robust preservation with universal composability~\cite{patrignani2022universal} is closest to what this paper presents.
-The authors demonstrate a similar compositionalty theorem to what is presented here (\Cref{sec:compcomp}) but use it in the context of protocols.
-They do not demonstrate the scalability of the approach.
-Moreover, they are missing the upper and lower compositions.
-% The authors demonstrate a similar compositionality theorem to what is presented here (\Cref{sec:compcomp}) as well as in an earlier version of this work~\cite{kruse2022csc}.
-% However, they do not demonstrate the scalability of the approach by means of a case study.
-
-\subsection{Other Secure Compilation Criteria}\label{subsec:relw:seccompcrit}
-
-While this paper focuses on the robust preservation framework~\cite{abate2019jour}, other secure compilation criteria exist.
-The survey on formal approaches to secure compilation~\cite{patrignani2019survey} discusses a broad spectrum already, while this section presents a very high-level overview.
-Fully abstract compilation~\cite{abadi1999fullabstraction} states that a compiler should preserve and reflect observational equivalence between source and target programs.
-It was shown~\cite{abate2021faandrc} that fully abstract compilers robustly preserve program properties that are either trivial or meaningless.
-As a mitigation for this, the authors presented a categorical approach based on maps of distributive laws~\cite{watanabe2002modl}, which they call many maps of distributive laws.
-Maps of distributive laws have been investigated before as a possible secure compilation criterion~\cite{tsampas2020catsc}.
-Other approaches are extensions of the compiler correctness criterion as discussed in other work~\cite{patterson2019next700} or the introduction of opaque observations~\cite{vu2021reconciling} to reconcile compiler optimisations with security.
-Note that this work also presents secure compilers that are optimising, but contrary to the other~\cite{vu2021reconciling}, provides a formal account of these in the robust preservation framework.
-% Lastly, the authors of this paper have presented ongoing work~\cite{patrignani2023blame} on a weaker robust preservation criteria based on the concept blame.
-
-\subsection{Memory Safety Mechanisms}\label{subsec:relw:msmechs}
-
-Different mechanisms for enforcing memory safety exist that also consider the secure compilation domain, i.e., have an active attacker model.
-For example, the ,,pointers as capabilities'' principle represents pointers as machine-level capabilities~\cite{korashy2021capableptrs}, which behave in a similar fashion to capabilities by means of linear typing~\cite{morrisett2005L3}.
-The approach of this paper also uses linear typing, but differs from $L^{3}$~\cite{morrisett2005L3} in the way that functions are not first-class.
-Moreover, this paper considers an active attacker, while the work on $L^{3}$ only discusses whole programs and, thus, has no active attacker model.
-The instrumentation to ensure memory safety that this paper presents is inspired by Softbounds~\cite{nagarakatte2009soft}.
-That work inserts bounds-checks in front of pointer-dereferences and, for this to work, inserts meta-data information on pointer creation.
-Softbounds also works in a more advanced setting with structured fields accesses and also introduces a table-lookup for pointers that are stored in memory.
-This paper only considers arrays of primitive data, i.e., there are no pointers to pointers or structures.
-Several other approaches to memory-safety exist in literature, specifically as compiler instrumentations~\cite{akritidis2009baggy,younan2010paricheck,jung2021pico,shankaranarayana2023tailcheck,dhumbumroong2020boundwarden,nam2019framer,zhou2023fatptrs}, hardware-extensions~\cite{kwon2013lowfat,saileshwar2022heapcheck,chen2023flexpointer,kim2023whistle}, or programming language extensions~\cite{elliott2018checkedc,li2022formalcheckedc,jim2002cyclone,elliott2015guilt,west2005cuckoo,weis2019fyr,benoit2019uniqueness}.
-What differentiates this work from them is that this work uses known, compiler-based approaches to ensure memory-safety as a means to investigate secure compiler compositions.
-This paper does not provide efficient memory-safety, but serves as a theoretical foundation for the secure compilation domain.
-
-To extend the languages in this paper with a less restricted form of pointer arithmetic, the region coloring memory safety monitor presented in earlier work~\cite{michael2023mswasm} can be used.
-The work presenting this monitor provides an approach for the robust preservation of memory safety compiling from C to WASM.
-However, they do not discuss composition of secure compilers but rather investigate an instance of a secure compiler.
-
-\subsection{Cryptographic Constant Time Mechanisms}\label{subsec:relw:cctmechs}
-
-The approach to preserving cryptographic constant time in this paper is high-level, where a programming language exposes a way to switch the semantics to a data (operand) independent timing mode.
-Since identifiers in $\Lscct$ are annotated with a secrecy tag, this approach is similar to others with information flow control.
-For example, Vale~\cite{bond2017vale} uses Dafny to ensure constant-time assembly code, while Jasmin~\cite{almeida2017jasmin} makes use of the Coq proof assistant to reject non-constant-time programs.
-CT-Wasm~\cite{watt2019ctwasm} enforces constant-timeness by means of a type system.
-Different to the approach of this paper, these approaches necessitate that the programmer writes \gls*{cct} code.
-An approach to allow programmers to write more high-level code is CryptOpt~\cite{kuepper2023cryptopt}, which generates efficient target-code by means of a randomised search.
-This paper abstracts over concrete mitigation strategies and simply assumes that there is a flag to switch to a cryptographic-constant time execution mode.
-This can be realised by employing the FaCT~\cite{cauligi2019fact} compiler, which translates common non-constant time code patterns to be constant-time, and the data (object) independent timing execution mode of modern processors.
-
-\section{Conclusion\pages{$\sfrac{1}{2}$}}\label{sec:concl}
-This paper tackled the problem of understanding what kind of security properties does a secure compiler preserve, when said compiler is the combination of compiler passes that preserve possibly different security properties.
-% 
-For this, this paper first formalised security properties of interest and their composition.
-% 
-Then, it proved that composing secure compilers that preserve certain properties results in a secure compiler that preserves the composition of these properties.
-% 
-Finally, this paper defines a multi-pass compiler and proves that it preserves \gls*{msscct}.
-Crucially, this paper derives the security of the multi-pass compiler from the composition of the security properties preserved by its individual passes, which include security-preserving as well as optimisation passes.
-
-% This paper does a first step towards practical secure compilation chains by introducing a theoretical framework demonstrating that secure compilers compose.
-% The paper exercises this framework on a case-study that consists of an optimising compilation chain that is secure with respect to \gls*{msscct}.
-% \MP{
-% 	not an appropriate recap.
-% 	rewrite summarising the paper
-% }
-
-% In future work, it would be interesting to investigate whether it is possible to provide {\em secure compiler combinators}, similarly to how parser combinators work.
-% %To this end, it may be necessary to extend existing frameworks for multi-language semantics.
-% It is also interesting to extend our case-study and verify all results of it in Coq.
-% This way, it is possible to obtain an executable, secure compiler.
-% However, the formalisation effort for backtranslation proofs is known to be enormous\MP{cit + details}, so another research avenue may be to find better ways to (semi-)automate standard secure compilation proofs.
-% \MPin{
-% 	not particularly informative FW. i'd rather not have it as such
-% }
-
-%% The acknowledgments section is defined using the "acks" environment
-%% (and NOT an unnumbered section). This ensures the proper
-%% identification of the section in the article metadata, and the
-%% consistent spelling of the heading.
-
-% \begin{acks}
-%   We would like to thank the anonymous reviewers for their feedback.
-% 	This work was partially supported by a gift from
-		% the Italian Ministry of Education through funding for the Rita Levi Montalcini grant (call of 2019).
-% \end{acks}
-
-%\newpage
-%%
-%% The next two lines define the bibliography style to be used, and
-%% the bibliography file.
-\bibliographystyle{ACM-Reference-Format}
+\bibliographystyle{IEEEtran}
 \bibliography{main}
@@
 \endinput
-%%
-%% End of file `sample-sigconf.tex'.

### 2023-09-19 csf/main.tex
finally compiles

--- 2023-09-18 csf/main.tex
@@
 \usepackage{listings}
+\usepackage{amssymb}
+\usepackage{amsmath}
+\usepackage{amsthm}
 \usepackage{xspace}
@@
 
-% https://tex.stackexchange.com/questions/648845/sans-serif-uppercase-greek-no-longer-showing-in-acmart
-\DeclareMathAlphabet{\mathsf}{OT1}{LibertinusSans-LF}{m}{n}
-\SetMathAlphabet{\mathsf}{bold}{OT1}{LibertinusSans-LF}{bx}{n}
-
-\DeclareMathAlphabet{\mathtt}{OT1}{lmtt}{m}{n}
-\SetMathAlphabet{\mathtt}{bold}{OT1}{lmtt}{bx}{n}
-
 \input{cmds.tex}
+
+\theoremstyle{definition}
+\newtheorem{exampleenv}{Example}[section]
+\newtheorem{lemma}{Lemma}[section]
+\newtheorem{theorem}{Theorem}[section]
+\newtheorem{corollary}{Corollary}[section]
+\newtheorem{definition}{Definition}[section]
 
@@
 	{This paper uses syntax highlighting accessible to both colourblind and black \& white readers.
-	% ~\citep{patrignani2020use}.
+	% ~\cite{patrignani2020use}.
 % 	% Specifically, it makes use of a $\src{blue}$, $\src{sans\text{-}serif}$ font for a $\src{source}$,
@@
 
-\begin{example}[strncpy]\label{ex:strncpy}
+\begin{exampleenv}[strncpy]\label{ex:strncpy}
 Consider the \texttt{C} function \texttt{strncpy} that copies a null-terminated string \texttt{src} into \texttt{dst} up to a length of \texttt{n}.
@@
 Thus, robust compilers can be used to prevent vulnerabilities resulting from uninitialised memory (as well as many other ones), e.g., by targeting capability-based languages such as CHERI~\cite{woodruff2014CHERI}, Arm Morello~\cite{arm-morello}, or MSWasm~\cite{michael2023mswasm}, where the compiler relies on capabilities to check that pointers are always initialised.
-\end{example}
+\end{exampleenv}
 
@@
         %\Cref{sec:compcomp} studies different forms of secure compiler compositions.
-        This paper takes the secure compilation framework of~\citep{abate2019jour} and extends it to reason about the security of all different known forms of compiler composition (\Cref{sec:compcomp}).
+        This paper takes the secure compilation framework of~\cite{abate2019jour} and extends it to reason about the security of all different known forms of compiler composition (\Cref{sec:compcomp}).
         %Concretely, it investigates upper and lower compositions, as well as the sequential composition.
@@
 
-\begin{example}[Double Free in Bluetooth Subsystem]
+\begin{exampleenv}[Double Free in Bluetooth Subsystem]
   Consider \texttt{CVE-2021-3564}~\cite{doublefree-bluetooth}, one of many submissions for a double-free vulnerability.
@@
   Therefore, \texttt{hci\_dev\_do\_open} does not satisfy the no-double-frees property robustly, since there is an implementation for \texttt{hci\_cmd\_work} that leads to a violation of that property when linked with \texttt{hci\_dev\_do\_open}.
-\end{example}
+\end{exampleenv}
 
@@
 
-\begin{example}[Types]
-  Suppose $\src{L}$ is a statically-typed language similar to \texttt{C} and $\trg{T}$ is dynamically typed, where both share the same syntax up to dynamic type checks.
-  Consider the following $\src{L}$ component and its compiled version below.
-  \begin{center}
-  $
-  \begin{array}{l}
-    \src{\lfunction{foo}{(char* x, int\ n)}{\lifz{valid\_ptr(x, n, sizeof(char))}{x[0]}{-1}}} \\
-    \trg{\lfunction{foo}{(\phantom{char*\ }x, \phantom{int\ }n)}{\lifz{valid\_ptr(x, n, sizeof(char))}{x[0]}{-1}}}
-  \end{array}
-  $
-  \end{center}
-  While the compiler emits code that may look correct, the generated code does not check that the provided argument is of the right type.
-  Even though the pointer $\trg{x}$ is checked for validity, the context $\trg{foo((int*)y, 1)}$ is able to provoke a read out of bounds.
-  Suppose the component transferred control to the context and passed ownership of a $\trg{char}$ pointer $\trg{y}$ sized $\trg{1}$ cells, the context can now call the component again, casting this buffer to an $\trg{int*}$ prior to that call.
-  The pointer is valid for one $\trg{char}$-sized memory cell, as expected, but the actual read operation now returns $\trg{sizeof(int)}$ many bytes instead of just $\trg{sizeof(char)}$ many.
-  Thus, even if $\src{foo}$ may have been robust with respect to the \gls*{sms}, its compiled counterpart is not and therefore the compiler fails to attain \Cref{def:rtp}.
-\end{example}
+% \begin{exampleenv}[Types]
+%   Suppose $\src{L}$ is a statically-typed language similar to \texttt{C} and $\trg{T}$ is dynamically typed, where both share the same syntax up to dynamic type checks.
+%   Consider the following $\src{L}$ component and its compiled version below.
+%   \begin{center}
+%   $
+%   \begin{array}{l}
+%     \src{\lfunction{foo}{(char* x, int\ n)}{\lifz{valid\_ptr(x, n, sizeof(char))}{x[0]}{-1}}} \\
+%     \trg{\lfunction{foo}{(\phantom{char*\ }x, \phantom{int\ }n)}{\lifz{valid\_ptr(x, n, sizeof(char))}{x[0]}{-1}}}
+%   \end{array}
+%   $
+%   \end{center}
+%   While the compiler emits code that may look correct, the generated code does not check that the provided argument is of the right type.
+%   Even though the pointer $\trg{x}$ is checked for validity, the context $\trg{foo((int*)y, 1)}$ is able to provoke a read out of bounds.
+%   Suppose the component transferred control to the context and passed ownership of a $\trg{char}$ pointer $\trg{y}$ sized $\trg{1}$ cells, the context can now call the component again, casting this buffer to an $\trg{int*}$ prior to that call.
+%   The pointer is valid for one $\trg{char}$-sized memory cell, as expected, but the actual read operation now returns $\trg{sizeof(int)}$ many bytes instead of just $\trg{sizeof(char)}$ many.
+%   Thus, even if $\src{foo}$ may have been robust with respect to the \gls*{sms}, its compiled counterpart is not and therefore the compiler fails to attain \Cref{def:rtp}.
+% \end{exampleenv}
 
@@
   \mi{(Security\ Tag)}~\sigma\bnfdef&\ \lock \mid \unlock\hspace{0.5cm}
-  \mi{(Control\ Tag)}~\sandboxtag\bnfdef \ctx\mid\comp\hspace{0.5cm}
-  \mi{(Event)}~\event\bnfdef\ \emptyevent \mid \lightning \mid \event[_{b}];\sandboxtag;\sigma \\
-  \mi{(Pre\text{-}event)}&~\event[_{b}]\bnfdef\ \ev{Alloc\ \loc\ n} \mid \ev{Dealloc\ \loc} \mid \ev{Use\ \loc\ n} \mid \ev{Branch\ n} \mid \ev{Binop\ n} \\
+  \mi{(Control\ Tag)}~\sandboxtag\bnfdef \ctx\mid\comp\\
+  \mi{(Event)}~\event\bnfdef&\ \emptyevent \mid \lightning \mid \event[_{b}];\sandboxtag;\sigma \\
+  \mi{(Pre\text{-}event)}~\event[_{b}]\bnfdef&\ \ev{Alloc\ \loc\ n} \mid \ev{Dealloc\ \loc} \\
+                                             & \mid \ev{Use\ \loc\ n} \mid \ev{Branch\ n} \mid \ev{Binop\ n} \\
   \end{aligned}
@@
 \begin{definition}[\glsfirst*{tms}]\label{def:trace:tmsdef}
-  $$
+  \scriptsize$$
   \tmssafe:=\left\{\trace \left| \begin{array}{rcl}
@@
 \begin{definition}[\glsfirst*{ms}]\label{def:trace:msdef}
-  $$
+  $
   \mssafe:=\tmssafe \cap \smssafe
-  $$
+  $
 \end{definition}
@@
 \begin{definition}[\glsfirst*{ms} and \gls*{scct}]\label{def:trace:msscctdef}
-  $$
+  $
   \msscctsafe:=\mssafe\cap\scctsafe
-  $$
+  $
 \end{definition}
@@
 \begin{figure}[!ht]
-  \judgbox{\traceagree{\pi}{\trace}{\abstrace}}{,,Specification-level trace $\trace$ agrees with monitor-level trace $\abstrace$ with respect to property $\pi$.''}
+  \judgbox{\traceagree{\pi}{\trace}{\abstrace}}{,,Specification-level trace $\trace$ agrees with monitor-level\\trace $\abstrace$ with respect to property $\pi$.''}
   %
@@
   \begin{aligned}
-    \mi{(Abstract\ Store)}~\tmsmonitor\bnfdef&\ \left\{\text{allocated}:L\times\sandboxtag,\text{freed}:L\times\sandboxtag\right\} \hspace{0.33cm}%
-    \emptyset:=\ \left\{\text{allocated}:\emptyset,\text{freed}:\emptyset\right\}\\
-    \mi{(Abstract\ Events)}~\absevent\bnfdef&\ \absev{\emptyevent} \mid \absev{Alloc\ }\loc\ \sandboxtag \mid \absev{Dealloc\ }\loc\ \sandboxtag \mid \absev{Use\ }\loc\ \sandboxtag \mid \absterm\\
+    \mi{(Abstract\ Store)}~\tmsmonitor\bnfdef&\ \left\{\text{A}:L\times\sandboxtag,\text{F}:L\times\sandboxtag\right\} \\%
+    \emptyset:=&\ \left\{\text{A}:\emptyset,\text{F}:\emptyset\right\}\\
+    \mi{(Abstract\ Events)}~\absevent\bnfdef&\ \absev{\emptyevent} \mid \absev{Alloc\ }\loc\ \sandboxtag \mid \absev{Dealloc\ }\loc\ \sandboxtag \\
+                                            & \mid \absev{Use\ }\loc\ \sandboxtag \mid \absterm\\
   \end{aligned}
@@
 \begin{center}
-  \judgbox{\monitorcheck[]{\tmsmonitor}{\tmsmonitor[']}{\absevent}}{,,Monitor $\tmsmonitor$ does one step to $\tmsmonitor[']$ given event $\absevent$.''}$\;$\\
+  \judgbox{\monitorcheck[]{\tmsmonitor}{\tmsmonitor[']}{\absevent}}{,,Monitor $\tmsmonitor$ does one step to $\tmsmonitor[']$\\given event $\absevent$.''}$\;$\\
   %
@@
   \typerule{tms-use}{
-    (\loc;\sandboxtag)\in\tmsmonitor.\text{allocated} &
-    (\loc;\sandboxtag)\notin\tmsmonitor.\text{freed}
+    (\loc;\sandboxtag)\in\tmsmonitor.\text{A} &
+    (\loc;\sandboxtag)\notin\tmsmonitor.\text{F}
   }{
@@
   \typerule{tms-alloc}{
-    (\loc;\sandboxtag)\notin\tmsmonitor.\text{allocated} &
-    (\loc;\sandboxtag)\notin\tmsmonitor.\text{freed} \\
-    \tmsmonitor[']=\left\{\text{allocated}: \tmsmonitor.\text{allocated}\cup\left\{(\loc;\sandboxtag)\right\},\text{freed}: \tmsmonitor.\text{freed}\right\}
+    (\loc;\sandboxtag)\notin\tmsmonitor.\text{A} &
+    (\loc;\sandboxtag)\notin\tmsmonitor.\text{F} \\
+    \tmsmonitor[']=\left\{\text{A}: \tmsmonitor.\text{A}\cup\left\{(\loc;\sandboxtag)\right\},\text{F}: \tmsmonitor.\text{F}\right\}
   }{
@@
   \typerule{tms-dealloc}{
-    (\loc;\sandboxtag)\in\tmsmonitor.\text{allocated} &
-    (\loc;\sandboxtag)\notin\tmsmonitor.\text{freed} \\
-    \tmsmonitor[']=\left\{\text{allocated}: \tmsmonitor.\text{allocated}\setminus\left\{(\loc;\sandboxtag)\right\},\text{freed}: \tmsmonitor.\text{freed}\cup\left\{(\loc;\sandboxtag)\right\}\right\}
+    (\loc;\sandboxtag)\in\tmsmonitor.\text{A} &
+    (\loc;\sandboxtag)\notin\tmsmonitor.\text{F} \\
+    \tmsmonitor[']=\left\{\text{A}: \tmsmonitor.\text{A}\setminus\left\{(\loc;\sandboxtag)\right\}\right.,\\\left.\text{F}: \tmsmonitor.\text{F}\cup\left\{(\loc;\sandboxtag)\right\}\right\}
   }{
@@
 \begin{center}
-  \judgbox{\monitorcheck{\tmsmonitor}{\tmsmonitor[']}{\abstrace}}{,,Monitor $\tmsmonitor$ multi-steps to $\tmsmonitor[']$ given trace $\abstrace$.''}$\;$\\
+  \judgbox{\monitorcheck{\tmsmonitor}{\tmsmonitor[']}{\abstrace}}{,,Monitor $\tmsmonitor$ multi-steps to $\tmsmonitor[']$\\given trace $\abstrace$.''}$\;$\\
   %
@@
 \begin{center}
-  \judgbox{\tmstraceagree[]{\event}{\absevent}}{,,Abstract event $\absevent$ is equivalent to $\event$ with respect to \gls*{tms}.''}$\;$\\
+  \judgbox{\tmstraceagree[]{\event}{\absevent}}{,,Abstract event $\absevent$ is equivalent to $\event$\\with respect to \gls*{tms}.''}$\;$\\
   %
@@
 
-\begin{example}[A program not satisfying \gls*{tms}]\label{ex:strncpy:cpp11}
+\begin{exampleenv}[A program not satisfying \gls*{tms}]\label{ex:strncpy:cpp11}
   Consider the following \texttt{C++11} library that calls \texttt{strncpy} (\Cref{ex:strncpy}) and prints the result to the standard output stream.
@@
   \begin{lstlisting}[language=c++,basicstyle=\small\ttfamily,mathescape,commentstyle=\color{gray}]
-int greet() {  // allocates 12 chars containing a greeting message
-  char* greetings = new char[12] { "Hello POPL!" }; // <- address ${\color{gray}\loc_{x}}$
-  char* to = new char[12];                          // <- address ${\color{gray}\loc_{y}}$
-
-  strncpy(12, to, greetings);
+int hello() {
+  char* greet = new char[11] { "Hello CSF!" };
+  // greet address ${\color{gray}\loc_{x}}$
+  char* to = new char[11];
+  // to address ${\color{gray}\loc_{y}}$
+  strncpy(11, to, greetings);
   delete to;
-
-  printf("%cOPL\n", to[6]);
+  printf("%cSF\n", to[6]);
 }
@@
   It follows that $\montmssafe{\abstrace[']}$ and from \Thmref{lem:mon:tmssafe}, it follows that the program satisfies \Thmref{def:trace:tmsdef}, even though the program still violates \gls*{sms}.
-\end{example}
+\end{exampleenv}
 
@@
   \begin{aligned}
-    \mi{(Abstract\ Store)}~\smsmonitor:=\ L\times\sandboxtag\times\mb{N}\hspace{0.33cm}&
-    \mi{(Abstract\ Events)}~\absevent\bnfdef\ \absev{\varepsilon} \mid \absev{Alloc\ }\loc\ \sandboxtag\ n \mid \absev{Use\ }\loc\ \sandboxtag\ n \\
+    \mi{(Abstract\ Store)}~\smsmonitor:=&\ L\times\sandboxtag\times\mb{N}\\
+    \mi{(Abstract\ Events)}~\absevent\bnfdef&\ \absev{\varepsilon} \mid \absev{Alloc\ }\loc\ \sandboxtag\ n \mid \absev{Use\ }\loc\ \sandboxtag\ n \\
   \end{aligned}
@@
 \begin{center}
-  \judgbox{\monitorcheck[]{\smsmonitor}{\smsmonitor[']}{\absevent}}{,,Monitor $\smsmonitor$ does one step to $\smsmonitor[']$ given event $\absevent$.''}$\;$\\
+  \judgbox{\monitorcheck[]{\smsmonitor}{\smsmonitor[']}{\absevent}}{,,Monitor $\smsmonitor$ does one step to $\smsmonitor[']$\\given event $\absevent$.''}$\;$\\
   %
@@
 
-\begin{example}[Normal invocation of \texttt{strncpy}]\label{ex:strncpy:sms}
+\begin{exampleenv}[Normal invocation of \texttt{strncpy}]\label{ex:strncpy:sms}
   Consider the insecure \texttt{strncpy} function from \Cref{ex:strncpy} with a context \texttt{strncpy(2, x, y)}, where \texttt{x} and \texttt{y} are pointers to valid regions of memory with allocated space for exactly two cells and do not contain the null-terminating character \texttt{'\textbackslash 0'}.
@@
   Therefore, the whole program (\texttt{strncpy} linked with this kind of context) is not \gls*{sms}.
-\end{example}
+\end{exampleenv}
 
@@
 \begin{center}
-  \judgbox{\monitorcheck[]{\msmonitor}{\msmonitor[']}{\absevent}}{,,Monitor $\msmonitor$ does one step to $\msmonitor[']$ given event $\absevent$.''}$\;$\\
+  \judgbox{\monitorcheck[]{\msmonitor}{\msmonitor[']}{\absevent}}{,,Monitor $\msmonitor$ does one step to $\msmonitor[']$\\given event $\absevent$.''}$\;$\\
   \typerule{ms-step}{
@@
   \begin{aligned}
-    \mi{(Abstract\ Store)}~\scctmonitor&:=\ \emptyset \hspace{0.5cm}
-    \mi{(Abstract\ Events)}~\absevent:=\ \absev{\varepsilon} \mid \absterm \mid \absev{Leak} \\
+    \mi{(Abstract\ Store)}~\scctmonitor:=&\ \emptyset \\
+    \mi{(Abstract\ Events)}~\absevent:=&\ \absev{\varepsilon} \mid \absterm \mid \absev{Leak} \\
   \end{aligned}
@@
 \begin{center}
-  \judgbox{\monitorcheck[]{\scctmonitor}{\scctmonitor[']}{\absevent}}{,,Monitor $\scctmonitor$ does one step to $\scctmonitor[']$ given event $\absevent$.''}$\;$\\
+  \judgbox{\monitorcheck[]{\scctmonitor}{\scctmonitor[']}{\absevent}}{,,Monitor $\scctmonitor$ does one step to $\scctmonitor[']$\\given event $\absevent$.''}$\;$\\
   %
@@
 \begin{center}
-  \judgbox{\sccttraceagree[]{\event}{\absevent}}{,,Abstract event $\absevent$ is equivalent to $\event$ with respect to \gls*{cct}.''}$\;$\\
+  \judgbox{\sccttraceagree[]{\event}{\absevent}}{,,Abstract event $\absevent$ is equivalent to $\event$\\with respect to \gls*{cct}.''}$\;$\\
   %
@@
 
-\begin{example}[Data-independent timing mode]
+\begin{exampleenv}[Data-independent timing mode]
   Consider the call \texttt{strncpy(1,x,y)} to the \texttt{strncpy} function (\Cref{ex:strncpy}) with low ($\unlock$) security for \texttt{x} and high security ($\lock$) for \texttt{y}.
@@
   Because of this, the whole trace of the component equates to $\absev{\emptyevent}$ and the monitor can step without getting stuck.
-\end{example}
+\end{exampleenv}
 
@@
   \begin{center}
-    \begin{tikzpicture}
+    \begin{tikzpicture}[scale=0.6, every node/.style={transform shape}]
       \node[minimum width=3cm,minimum height=2.5cm,draw=black,very thick,align=left] (progunopt) {\begin{lstlisting}[language=ml,basicstyle=\small\ttfamily]
@@
   % Given \texttt{p}, yield
-  $\lambda \texttt{p}\ldotp$
-  $\left\{\mbox{\begin{tabular}{c}
+  $$\lambda \texttt{p}\ldotp
+  \left\{\mbox{\begin{tabular}{c}
     {\rul{if $\texttt{p}\in\src{L}$, then $\cc{\src{L}}{\trg{L}}(\texttt{p})$}} \\
     \mbox{\oul{if $\texttt{p}\in\obj{L}$, then $\cc{\obj{L}}{\trg{L}}(\texttt{p})$}} \\
-  \end{tabular}}\right.$
+  \end{tabular}}\right.$$
 %
@@
   % Given $L$ and $\src{p}$, yield
-  $\lambda \src{p}, L\ldotp$
-  $\left\{\mbox{\begin{tabular}{c}
+  $$\lambda \src{p}, L\ldotp
+  \left\{\mbox{\begin{tabular}{c}
     {\pul{if $L=\trg{L}$, then} \rul{$\cc{\src{L}}{\trg{L}}(\src{p})$}} \\
     \mbox{\pul{if $L=\obj{L}$, then} \oul{$\cc{\src{L}}{\obj{L}}(\src{p})$}} \\
-  \end{tabular}}\right.$
+  \end{tabular}}\right.$$
 %
@@
 
-%\section{Case Study: Memory Safe and Cryptographic Constant Time Code}\label{sec:casestud}
 \section{Case Study: Language Formalisations\pages{2}}\label{sec:casestud:defs}
@@
 
-\begin{example}[$\Lscct$ with and without data independent timing]\label{ex:lscct}
+\begin{exampleenv}[$\Lscct$ with and without data independent timing]\label{ex:lscct}
   Consider again the context presented in \Cref{ex:strncpy:cpp11}, where everything is marked with a security tag of high $\lock$.
@@
   But, if it is turned on, then the branching event \colorbox{yellow}{does not fire} anymore and both reading and writing to memory gets ultimately translated to a specification trace with \colorbox{lightgray}{no exposed secrets}.
-\end{example}
+\end{exampleenv}
 
@@
 
-\begin{figure}[!ht]
+\begin{figure*}[!ht]
   \begin{center}
@@
   \end{center}
-\end{figure}
+\end{figure*}
 
@@
 To this end, the compiler introduces another, fresh identifier $\irl{x_{SIZE}}$ for each allocation that binds $\irl{x}$ to keep track of the allocation size.
-\begin{example}[Instrumented \texttt{strncpy}]
+\begin{exampleenv}[Instrumented \texttt{strncpy}]
   Consider again \texttt{strncpy}, but instrumented for \gls*{sms}:
@@
     When calling this in similar fashion to \Cref{ex:strncpy:sms}, the event $\ev{Use}\ \loc_{x}\ 2;\comp;\unlock$ would not be emitted during execution, since the bounds check prevents the condition \texttt{src[i] != '\textbackslash 0'} from executing.
-\end{example}
+\end{exampleenv}
 
@@
 
-%\newpage
-%%
-%% The next two lines define the bibliography style to be used, and
-%% the bibliography file.
-\bibliographystyle{ACM-Reference-Format}
+\section*{Acknowledgements}\label{sec:acks}
+
+\bibliographystyle{IEEEtran}
 \bibliography{main}

### 2023-09-19 csf/main.tex
Fix grave layouting issues

@@
   \begin{aligned}
-    \mi{(Expressions)}~&\expr\bnfdef\ x \mid \valueexpr \mid \lbinop{\expr[_{1}]}{\expr[_{2}]} \mid \lget{x}{\expr} \mid \llet{x}{\lnew{\expr[_{1}]}{\expr[_{2}]}}{\expr[_{3}]} \mid \ldelete{x} \mid \lset{x}{\expr[_{1}]}{\expr[_{2}]} \\
-                       & \mid \lpair{\expr[_{1}]}{\expr[_{2}]} \mid \lproja{\expr} \mid \lprojb{\expr} \mid \llet{x}{\expr[_{1}]}{\expr[_{2}]} \mid \lreturn{\expr} \mid \lcall{g}{\expr} \mid \lifz{\expr}{\expr[_{1}]}{\expr[_{2}]} \mid \labort
+    \mi{(Expressions)}~&\expr\bnfdef\ x \mid \valueexpr \mid \lbinop{\expr[_{1}]}{\expr[_{2}]} \mid \lget{x}{\expr}\\
+                       &\mid \llet{x}{\lnew{\expr[_{1}]}{\expr[_{2}]}}{\expr[_{3}]} \mid \ldelete{x}\\
+                       &\mid \lset{x}{\expr[_{1}]}{\expr[_{2}]} \mid \lpair{\expr[_{1}]}{\expr[_{2}]} \mid \lproja{\expr} \mid \lprojb{\expr}\\
+                       &\mid \llet{x}{\expr[_{1}]}{\expr[_{2}]} \mid \lreturn{\expr} \mid \lcall{g}{\expr}\\
+                       &\mid \lifz{\expr}{\expr[_{1}]}{\expr[_{2}]} \mid \labort
   \end{aligned}
@@
   \begin{aligned}
-  \mi{(Types)}~&\type\bnfdef\natt \mid \type[_{1}] \times \type[_{2}]
-  &
-  \mi{(Functions)}~&\function\bnfdef\ \lfunction{\textit{foo}}{x}{e}
-  &
-  \mi{(Libraries)}~&\library\bnfdef\ \hole{\cdot} \mid \function,\library
+  \mi{(Functions)}~\function\bnfdef&\ \lfunction{\textit{foo}}{x}{e}
+  \end{aligned}
+  \\
+  \begin{aligned}
+  \mi{(Types)}~\type\bnfdef\natt \mid \type[_{1}] \times \type[_{2}]\hspace{0.5cm}
+  \mi{(Libraries)}~\library\bnfdef\ \hole{\cdot} \mid \function,\library
   \end{aligned}\\
   \begin{aligned}
-  \mi{(Component~Names)}~&\commlib\bnfdef\ \hole{\cdot} \mid \text{foo},\commlib
-  &
-  \mi{(Programs)}~&\prog{\library_{\ctx}}{\library_{\comp}}
+  \mi{(Component~Names)}~\commlib\bnfdef\ \hole{\cdot} \mid \text{foo},\commlib
+  \end{aligned}
+  \\
+  \begin{aligned}
+  \mi{(Programs)}~\prog{\library_{\ctx}}{\library_{\comp}}
   \end{aligned}
@@
   \begin{aligned}
-    \mi{(Control~Tags)}~&\sandboxtag\bnfdef\ctx \mid \comp
-    &
-    \mi{(Communication~Tags)}~&\comm\bnfdef\ \ctxtocomp \mid\ \comptoctx \mid \nocomm
-    &
-    \mi{(Poison~Tags)}~&\poison\bnfdef \poisoned \mid \poisonless
-  \end{aligned}\\
-  \begin{aligned}
-    \mi{(Continuation~Stacks)}~&\ectxstack\bnfdef\hole{\cdot} \mid \tup{K}{g},\ectxstack
-    &
+    \mi{(Control~Tags)}~&\sandboxtag\bnfdef\ctx \mid \comp \\
+    \mi{(Communication~Tags)}~&\comm\bnfdef\ \ctxtocomp \mid\ \comptoctx \mid \nocomm \\
+    \mi{(Poison~Tags)}~&\poison\bnfdef \poisoned \mid \poisonless \\
+    \mi{(Continuation~Stacks)}~&\ectxstack\bnfdef\hole{\cdot} \mid \tup{K}{g},\ectxstack \\
     \mi{(Control~Flow~States)}~&\cfstate\bnfdef\tup{\library}{\ectxstack}
-  	\\
-    \mi{(Locations)}~&\loc\in\nat
-    &
+    \hspace{0.5cm}
+    \mi{(Locations)}~\loc\in\nat \\
     \mi{(Substitutions)}~&\substvar\bnfdef \hole{\cdot} \mid \subst{\valueexpr}{x},\gamma
   	\\
-    \mi{(Stores)}~&\Delta\bnfdef\hole{\cdot} \mid x\mapsto\tup{\loc}{\sandboxtag,\poison,n},\Delta
-    &
-    \mi{(Heaps)}~&H\bnfdef\hole{\cdot} | \valueexpr,H
-    \end{aligned}\\
-  \begin{aligned}
-    \mi{(Memory~States)}~&\memstate\bnfdef\tup{H^{\ctx}}{H^{\comp};\Delta}
-    &
-    \mi{(States)}~&\statevar\bnfdef\tup{\cfstate}{\sandboxtag,\memstate}
-    &
+    \mi{(Stores)}~&\Delta\bnfdef\hole{\cdot} \mid x\mapsto\tup{\loc}{\sandboxtag,\poison,n},\Delta \\
+    \mi{(Heaps)}~&H\bnfdef\hole{\cdot} | \valueexpr,H \\
+    \mi{(Memory~States)}~&\memstate\bnfdef\tup{H^{\ctx}}{H^{\comp};\Delta} \\
+    \mi{(States)}~&\statevar\bnfdef\tup{\cfstate}{\sandboxtag,\memstate} \\
     \mi{(Runtime~Terms)}~&\runtimetermvar\bnfdef\rtt{\statevar}{\expr}
@@
   \begin{aligned}
-  \mi{(Pre\text{-}Events)}~\event[_{b}]&\bnfdef \ev{Alloc}\ \loc\ \valueexpr \mid \ev{Dealloc}\ \loc \mid \ev{Get}\ \loc\ \valueexpr \mid \ev{Set}\ \loc\ \valueexpr\ \valueexpr['] \mid \cdots \\
+  \mi{(Pre\text{-}Events)}~\event[_{b}]\bnfdef&\ \ev{Alloc}\ \loc\ \valueexpr \mid \ev{Dealloc}\ \loc\\
+                                       &\mid \ev{Get}\ \loc\ \valueexpr \mid \ev{Set}\ \loc\ \valueexpr\ \valueexpr['] \mid \cdots \\
   \mi{(Events)}~\event&\bnfdef \emptyevent \mid \lightning \mid \tup{\event[_{b}]}{\sandboxtag,\securitytag} \\
@@
     \fresh{\statevar}{\loc} &
-    \heap[^{\sandboxtag}_1]=\growh{\statevar.\heap[^{\sandboxtag}]}{n} &
-    \store[_1]=\storeel{z}{\loc}{\statevar.\sandboxtag}{\poisonless}{n},\statevar.\store
-  }{
-    \pstepto{\rtt{\statevar}{\lnew{x}{n}{\expr}}}{\rtt{\statevar\subst{\heap[^{\sandboxtag}]}{\heap[^{\sandboxtag}_1]}\subst{\store}{\store[_1]}}{\expr\subst{z}{x}}}{\tup{\ev{Alloc}\ \loc\ n}{\sandboxtag}}
+    \heap[^{\sandboxtag}_1]=\growh{\statevar.\heap[^{\sandboxtag}]}{n} \\
+    \store[_1]=\storeel{z}{\loc}{\statevar.\sandboxtag}{\poisonless}{n},\statevar.\store \\
+    \statevar['] = \statevar\subst{\heap[^{\sandboxtag}]}{\heap[^{\sandboxtag}_1]}\subst{\store}{\store[_1]}
+  }{
+    \pstepto{\rtt{\statevar}{\llet{x}{\lnew{\valueexpr}{n}}{e}}}{\rtt{\statevar[']}{\expr\subst{z}{x}}}{\tup{\ev{Alloc}\ \loc\ n}{\sandboxtag}}
   }{e-new}
@@
 
-\begin{center}
-  \judgbox{\estepto[]{\runtimetermvar}{\runtimetermvar}{\event}}{,,Contextual step from runtime-term $\runtimetermvar$ to $\runtimetermvar[']$ emitting event $\event$.''}
-  %
-  \typerule{$e-\text{ret}$}{
-    \statevar.\ectxstack=\tup{\ectx}{\text{foo}},\ectxstack['] &
-    \commlib\vdash\text{foo}:\Omega.\sandboxtag\dashv\comm
-  }{
-    \estepto[]{\rtt{\statevar}{\ectx[']\hole{\lreturn{\valueexpr}}}}{\rtt{\statevar\subst{\neg\statevar.\sandboxtag}{\sandboxtag}\subst{\ectxstack[']}{\ectxstack}}{\ectx\hole{\valueexpr}}}{\tup{\ev{Ret}\ \comm\ \valueexpr}{\statevar.\sandboxtag}}
-  }{e-ret-notsame}
-  %
-  \typerule{$e-\text{call}-\text{notsame}$}{
-    \lfunction{\text{foo}}{x}{e} \in \statevar.\library &
-    \commlib\vdash\text{foo}:\Omega.\sandboxtag\dashv\comm
-  }{
-    \estepto[]{\rtt{\statevar}{\ectx\hole{\lcall{\text{foo}}{\valueexpr}}}}{\rtt{\statevar\subst{\neg\statevar.\sandboxtag}{\sandboxtag}\subst{\tup{\ectx}{\text{foo}},\statevar.\ectxstack}{\ectxstack}}{\expr\subst{\valueexpr}{x}}}{\tup{\ev{Call}\ \comm\ \text{foo}\ \valueexpr}{\statevar.\sandboxtag}}
-  }{e-call-notsame}
-\end{center}
+% \begin{center}
+%   \judgbox{\estepto[]{\runtimetermvar}{\runtimetermvar}{\event}}{,,Contextual step from runtime-term $\runtimetermvar$ to $\runtimetermvar[']$ emitting event $\event$.''}
+%   %
+%   \typerule{$e-\text{ret}$}{
+%     \statevar.\ectxstack=\tup{\ectx}{\text{foo}},\ectxstack['] &
+%     \commlib\vdash\text{foo}:\Omega.\sandboxtag\dashv\comm
+%   }{
+%     \estepto[]{\rtt{\statevar}{\ectx[']\hole{\lreturn{\valueexpr}}}}{\rtt{\statevar\subst{\neg\statevar.\sandboxtag}{\sandboxtag}\subst{\ectxstack[']}{\ectxstack}}{\ectx\hole{\valueexpr}}}{\tup{\ev{Ret}\ \comm\ \valueexpr}{\statevar.\sandboxtag}}
+%   }{e-ret-notsame}
+%   %
+%   \typerule{$e-\text{call}-\text{notsame}$}{
+%     \lfunction{\text{foo}}{x}{e} \in \statevar.\library &
+%     \commlib\vdash\text{foo}:\Omega.\sandboxtag\dashv\comm
+%   }{
+%     \estepto[]{\rtt{\statevar}{\ectx\hole{\lcall{\text{foo}}{\valueexpr}}}}{\rtt{\statevar\subst{\neg\statevar.\sandboxtag}{\sandboxtag}\subst{\tup{\ectx}{\text{foo}},\statevar.\ectxstack}{\ectxstack}}{\expr\subst{\valueexpr}{x}}}{\tup{\ev{Call}\ \comm\ \text{foo}\ \valueexpr}{\statevar.\sandboxtag}}
+%   }{e-call-notsame}
+% \end{center}
 
@@
 \begin{center}
-  \judgbox{\msfilterLtms{\src{\event}}=\event}{,,Project an $\Ltms$ event $\src{\event}$ to a specification event $\event$.''}
+  \judgbox{\msfilterLtms{\src{\event}}=\event}{,,Project an $\Ltms$ event $\src{\event}$ to\\a specification event $\event$.''}
   \typerule{$\Ltms$-filter-context}{
@@
   \begin{align*}
-    \mi{(Expressions)}~\trg{\expr}&\bnfdef \cdots \mid \trg{\lispoisoned{x}} \mid \trg{\lhast{\expr}{\type}} \\
+    \mi{(Expressions)}~\trg{\expr}\bnfdef& \cdots \mid \trg{\lispoisoned{x}} \mid \trg{\lhast{\expr}{\type}} \\
   \end{align*}
@@
   \begin{aligned}
-    \mi{(Expressions)}~\obj{\expr}&\bnfdef \obj{n^{\securitytag}} \mid \cdots \mid \obj{\llet{x^{\securitytag}}{\expr[_{1}]}{\expr[_{2}]}} \mid \cdots \mid \obj{\lwrdoit{D}} \mid \obj{\lrddoit{x}{\expr}}
+    \mi{(Expressions)}~\obj{\expr}\bnfdef&\ \obj{n^{\securitytag}} \mid \cdots \mid \obj{\llet{x^{\securitytag}}{\expr[_{1}]}{\expr[_{2}]}} \mid \cdots\\&\mid \obj{\lwrdoit{D}} \mid \obj{\lrddoit{x}{\expr}}
   \end{aligned} 
@@
   \begin{align*}
-    \mi{(Pre\text{-}Events)}~\obj{\event[_{b}]}&\bnfdef \cdots \mid \obj{\widehat{Get}\ \loc\ \valueexpr} \mid \obj{\widehat{Set}\ \loc\ \valueexpr\ \valueexpr[']} \mid \obj{Branch\ n} \mid \obj{Binop\ n}
+    \mi{(Pre\text{-}Events)}~\obj{\event[_{b}]}\bnfdef&\ \cdots \mid \obj{\widehat{Get}\ \loc\ \valueexpr} \mid \obj{\widehat{Set}\ \loc\ \valueexpr\ \valueexpr[']}\\
+                                                      &\mid \obj{Branch\ n} \mid \obj{Binop\ n}
   \end{align*}
@@
 \begin{center}
-  \judgbox{\scctfilterLscct{\obj{\event}}=\event}{,,Project an $\Lscct$ event $\obj{\event}$ to a specification event $\event$.''}
+  \judgbox{\scctfilterLscct{\obj{\event}}=\event}{,,Project an $\Lscct$ event $\obj{\event}$\\to a specification event $\event$.''}
   \typerule{$\Lscct$-filter-context}{
@@
 
+\begin{figure*}[!h]
 \begin{exampleenv}[$\Lscct$ with and without data independent timing]\label{ex:lscct}
@@
 \end{exampleenv}
+\end{figure*}
 
@@
 This section defines several secure compilers, each of which robustly preserves a different property of interest as depicted in \Cref{fig:pipeline}.
-\begin{figure}[!h]
+\begin{figure*}[!h]
   \begin{center}
@@
   }\label{fig:pipeline}
-\end{figure}
+\end{figure*}
 The section demonstrates the power of the framework (\Cref{sec:compprop,sec:compcomp}) by composing these compilers for a secure and optimising compilation chain that robustly preserves \gls*{msscct}.
@@
   \begin{aligned}
-    \cca(\src{x}) & = \trg{x} \\
-    \cca(\src{n}) & = \trg{n} \\
-    \cca(\src{\lbinop{\expr[_{1}]}{\expr[_{2}]}}) & =  \lbinop[\trg]{\left[\cca(\src{\expr[_{1}]})\right]}{\left[\cca(\src{\expr[_{2}]})\right]} \\
-    \cca(\src{\lget{x}{\expr}}) & = \lget[\trg]{\trg{x}}{\left[\cca(\src{\expr})\right]} \\
-    \cca(\src{\ldelete{x}}) & = \ldelete[\trg]{\left[\cca(\src{x})\right]} \\
+    \cca(\src{x}) = & \trg{x} \\
+    \cca(\src{n}) = & \trg{n} \\
+    \cca(\src{\lbinop{\expr[_{1}]}{\expr[_{2}]}}) = & \lbinop[\trg]{\left[\cca(\src{\expr[_{1}]})\right]}{\left[\cca(\src{\expr[_{2}]})\right]} \\
+    \cca(\src{\lget{x}{\expr}}) = & \lget[\trg]{\trg{x}}{\left[\cca(\src{\expr})\right]} \\
+    \cca(\src{\ldelete{x}}) = & \ldelete[\trg]{\left[\cca(\src{x})\right]} \\
 %    \cca(\src{\lset{x}{\expr[_{1}]}{\expr[_{2}]}}) & = \lset[\trg]{\left[\cca(\src{x})\right]}{\left[\cca(\src{\expr[_{1}]})\right]}{\left[\cca(\src{\expr_{2}})\right]} \\
-    \cca(\src{\lfunction{g}{x:\natt\to\type[_{e}]}{\expr}}) & = \lfunction[\trg]{\trg{g}}{\trg{x}}{\lifz[\trg]{\trg{\lhast{x}{\natt}}}{%
-                                                                                                \left[\cca(\src{\expr})\right] %
+\cca(\src{\lfunction{g}{x:\natt\to\type[_{e}]}{\expr}})  = & \lfunction[\trg]{\trg{g}}{\trg{x}}{\lifz[\trg]{\trg{\lhast{x}{\natt}}}{\\%
+                                                            &\left[\cca(\src{\expr})\right] %
                                                                                                  }{\labort[\trg]}}
@@
   \begin{array}{rcl}
-    \ccb(\trg{\lnew{x}{\expr[_{1}]}{\expr[_{2}]}}) & = & \llet[\irl]{\irl{x_{SIZE}}}{\ccb(\trg{\expr[_{1}]})}{\lnew[\irl]{\irl{x}}{\irl{x_{SIZE}}}{\ccb(\trg{\expr[_{2}]})}} \\
-    \ccb(\trg{\lget{x}{\expr}}) & = & \llet[\irl]{\irl{x_{n}}}{\ccb(\trg{\expr})}{\irl{\lifz{0\le x_{n}<x_{SIZE}}{\lget{x}{x_{n}}}{\labort}}} \\
-    \ccb(\trg{\lset{x}{\expr[_{1}]}{\expr[_{2}]}}) & = & \llet[\irl]{\irl{x_{n}}}{\ccb(\trg{\expr[_{1}]})}{\lifz[\irl]{\irl{0\le x_{n}<x_{SIZE}}}{\lset[\irl]{\irl{x}}{\irl{x_{n}}}{\ccb(\trg{\expr[_{2}]})}}{\irl{\labort}}} \\
+    \ccb(\trg{\lnew{x}{\expr[_{1}]}{\expr[_{2}]}}) & = & \llet[\irl]{\irl{x_{SIZE}}}{\ccb(\trg{\expr[_{1}]})}{\\&&\lnew[\irl]{\irl{x}}{\irl{x_{SIZE}}}{\ccb(\trg{\expr[_{2}]})}} \\
+  \ccb(\trg{\lget{x}{\expr}}) & = & \llet[\irl]{\irl{x_{n}}}{\ccb(\trg{\expr})}{\\&&\lifz[\irl]{\irl{0\le x_{n}<x_{SIZE}}}{\irl{\lget{x}{x_{n}}}}{\\&&\irl{\labort}}} \\
+  \ccb(\trg{\lset{x}{\expr[_{1}]}{\expr[_{2}]}}) & = & \llet[\irl]{\irl{x_{n}}}{\ccb(\trg{\expr[_{1}]})}{\\&&\lifz[\irl]{\irl{0\le x_{n}<x_{SIZE}}}{\lset[\irl]{\irl{x}}{\irl{x_{n}}}{\\&&\ccb(\trg{\expr[_{2}]})}}{\irl{\labort}}} \\
   \end{array}
@@
   Consider again \texttt{strncpy}, but instrumented for \gls*{sms}:
-    \begin{lstlisting}[language=c,basicstyle=\ttfamily\small, morekeywords={size_t}]
+    \begin{lstlisting}[language=c,basicstyle=\ttfamily\scriptsize, morekeywords={size_t}]
   void strncpy(size_t n, size_t dst_size, char *dst,
                          size_t src_size, char *src) {
-    for(size_t i = 0; i < src_size && src[i] != '\0' && i < n; ++i) {
+    for(size_t i = 0; i < src_size && src[i] != '\0'
+                   && i < n; ++i) {
       if(i < src_size && i < dst_size) {
@@
   \begin{align*}
-    \partialeval{\irl{\llet{x}{\expr[_{1}]}{\expr[_{2}]}}}{\irl{\substlist}} & = \llet[\irl]{\irl{x}}{\partialeval{\irl{\expr[_{1}]}}{\irl{\substlist}}}{\partialeval{\irl{\expr[_{2}]}}{\irl{\substlist}}} \\
-    \partialeval{\irl{\lifz{\expr[_{1}]}{\expr[_{2}]}{\expr[_{3}]}}}{\irl{\substlist}} & = \lifz[\irl]{\partialeval{\irl{\expr[_{1}]}}{\irl{\substlist}}}{\partialeval{\irl{\expr[_{2}]}}{\irl{\substlist}}}{\partialeval{\irl{\expr[_{3}]}}{\irl{\substlist}}} \\
+    \partialeval{\irl{\llet{x}{\expr[_{1}]}{\expr[_{2}]}}}{\irl{\substlist}} & = \llet[\irl]{\irl{x}}{\partialeval{\irl{\expr[_{1}]}}{\irl{\substlist}}}{\\&\partialeval{\irl{\expr[_{2}]}}{\irl{\substlist}}} \\
+    \partialeval{\irl{\lifz{\expr[_{1}]}{\expr[_{2}]}{\expr[_{3}]}}}{\irl{\substlist}} & = \lifz[\irl]{\partialeval{\irl{\expr[_{1}]}}{\irl{\substlist}}}{\\&\partialeval{\irl{\expr[_{2}]}}{\irl{\substlist}}}{\partialeval{\irl{\expr[_{3}]}}{\irl{\substlist}}} \\
   \end{align*}

### 2023-09-21 csf/main.tex
fix cleverref
Remove larger chunks of text and fix some figures.

@@
 \input{cmds.tex}
+
+% allow page breaks inside multiline alignment displays
+\allowdisplaybreaks
+
+\Crefname{exampleenv}{Example}{Examples}
 
@@
 A class is simply a set of properties and for the class of safety properties, it is decidable whether a trace satisfies a safety property with just a finite trace prefix.
+The compositionality framework (\Cref{sec:compcomp}) presented in this paper considers arbitrary classes, while the case-study (\Cref{sec:casestud:rtp}) fixes them to concrete instances.
 As an example, consider a trace describing an interaction with a memory where the deallocation of an address $\loc$ precedes a read at that address in memory: $\ev{Dealloc\ \loc}\cdot\ev{Read\ \loc\ 1729}\cdot\dots$.
@@
   $\isdef$
-  if \iul{$\progstepto{\progvar}{\runtimetermvar}{\trace}$},
+  if \iul{$\forall\runtimetermvar\ \trace,\progstepto{\progvar}{\runtimetermvar}{\trace}$},
   then \oul{$\trace\in\pi$}.
@@
   \bul{$\rsat{\progvar}{\pi}$}
-  $\isdef$ \iul{$\forall \contextvar$, if $\link{\contextvar}{\progvar}=\wholeprogvar$}, then \oul{$\sat{\wholeprogvar}{\pi}$}.
+  $\isdef$ \iul{$\forall \contextvar\ \wholeprogvar$, if $\link{\contextvar}{\progvar}=\wholeprogvar$}, then \oul{$\sat{\wholeprogvar}{\pi}$}.
 \end{definition}
 
-\begin{exampleenv}[Double Free in Bluetooth Subsystem]
-  Consider \texttt{CVE-2021-3564}~\cite{doublefree-bluetooth}, one of many submissions for a double-free vulnerability.
-  The vulnerability arises due to a race condition where the context-level function \texttt{hci\_cmd\_work} was not expected to behave maliciously, since it resides in the same source-code repository where the vulnerability occurs.
-  Nevertheless, the component-level code of \texttt{hci\_dev\_do\_open} is linked with \texttt{hci\_cmd\_work} and does not atomically check whether a pointer has been freed already:
-  Therefore, \texttt{hci\_dev\_do\_open} does not satisfy the no-double-frees property robustly, since there is an implementation for \texttt{hci\_cmd\_work} that leads to a violation of that property when linked with \texttt{hci\_dev\_do\_open}.
-\end{exampleenv}
+% \begin{exampleenv}[Double Free in Bluetooth Subsystem]
+%   Consider \texttt{CVE-2021-3564}~\cite{doublefree-bluetooth}, one of many submissions for a double-free vulnerability.
+%   The vulnerability arises due to a race condition where the context-level function \texttt{hci\_cmd\_work} was not expected to behave maliciously, since it resides in the same source-code repository where the vulnerability occurs.
+%   Nevertheless, the component-level code of \texttt{hci\_dev\_do\_open} is linked with \texttt{hci\_cmd\_work} and does not atomically check whether a pointer has been freed already:
+%   Therefore, \texttt{hci\_dev\_do\_open} does not satisfy the no-double-frees property robustly, since there is an implementation for \texttt{hci\_cmd\_work} that leads to a violation of that property when linked with \texttt{hci\_dev\_do\_open}.
+% \end{exampleenv}
 
@@
 \begin{definition}[\glsfirst*{tms}]\label{def:trace:tmsdef}
-  \scriptsize$$
+  \footnotesize$$
   \tmssafe:=\left\{\trace \left| \begin{array}{rcl}
@@
 To go from specification traces $\trace$ to monitor-level traces $\abstrace$, each property $\pi$ has an associated event agreement relation $\traceagree[]{\pi}{\event}{\absevent}$.
-\Cref{fig:stuttereq} shows how the event agreement is lifted to traces.
-\begin{figure}[!ht]
-  \judgbox{\traceagree{\pi}{\trace}{\abstrace}}{,,Specification-level trace $\trace$ agrees with monitor-level\\trace $\abstrace$ with respect to property $\pi$.''}
-  %
-  \typerule{traceagree-empty}{}{
-    \traceagree{\pi}{\hole{\cdot}}{\hole{\cdot}}
-  }{traceagree-empty}
-  %
-  \typerule{traceagree-ign-L}{
-    \traceagree{\pi}{\trace}{\abstrace}
-  }{
-    \traceagree{\pi}{\emptyevent\cdot\trace}{\abstrace}
-  }{traceagree-ign-L}
-  %
-  \typerule{traceagree-ign-R}{
-    \traceagree{\pi}{\trace}{\abstrace}
-  }{
-    \traceagree{\pi}{\trace}{\absev{\emptyevent}\cdot\abstrace}
-  }{traceagree-ign-R}
-  %
-  \typerule{traceagree-cons}{
-    \traceagree[]{\pi}{\event}{\absevent} &
-    \traceagree{\pi}{\trace}{\abstrace}
-  }{
-    \traceagree{\pi}{\event\cdot\trace}{\absevent\cdot\abstrace}
-  }{traceagree-cons}
-  \caption{Trace-Agreement relation that equates specification-level traces with monitor-level traces.}\label{fig:stuttereq}
-\end{figure}
-The trace agreement is the same for all properties $\pi$ up to the use of the event agreement in \Cref{tr:traceagree-cons}.
+Lifting the event agreement to traces works as expected, i.e., the empty traces are related, both are allowed to stutter with the empty event, and otherwise the events $\event$ and $\absevent$ must fulfill $\traceagree[]{\pi}{\event}{\absevent}$.
+The trace agreement is the same for all properties $\pi$. 
 With agreements, this section defines monitor satisfaction for traces and then it proves that monitor satisfaction implies property satisfaction.
@@
 
-\begin{exampleenv}[A program not satisfying \gls*{tms}]\label{ex:strncpy:cpp11}
-  Consider the following \texttt{C++11} library that calls \texttt{strncpy} (\Cref{ex:strncpy}) and prints the result to the standard output stream.
-
-  \begin{lstlisting}[language=c++,basicstyle=\small\ttfamily,mathescape,commentstyle=\color{gray}]
-int hello() {
-  char* greet = new char[11] { "Hello CSF!" };
-  // greet address ${\color{gray}\loc_{x}}$
-  char* to = new char[11];
-  // to address ${\color{gray}\loc_{y}}$
-  strncpy(11, to, greetings);
-  delete to;
-  printf("%cSF\n", to[6]);
-}
-  \end{lstlisting}
-  Up to the body of \texttt{printf}, the program execution yields the specification trace $\ev{Alloc}\ \loc[_{x}]\ 12;\comp;\unlock\cdot\ev{Alloc}\ \loc[_{y}]\ 12;\comp;\unlock\cdot\ev{Use}\ \loc_{x}\ 0;\comp;\unlock\cdot\ev{Use}\ \loc_{x}\ 0;\comp;\unlock\cdot\ev{Use}\ \loc_{y}\ 0;\comp;\unlock\cdot\ev{Use}\ \loc_{x}\ 1;\comp;\unlock\cdot\ldots\cdot\ev{Use}\ \loc_{x}\ 12;\comp;\unlock\cdot\ev{Dealloc}\ \loc[_{y}];\comp;\unlock\cdot\ev{Use}\ \loc_{y}\ 6;\comp;\unlock$.
-  Relating this trace to abstract monitor events yields $\abstrace=\absev{Alloc}\ \loc[_{x}]\ \comp\cdot\absev{Alloc}\ \loc[_{y}]\ \comp\cdot\absev{Use}\ \loc[_{x}]\ \comp\cdot\absev{Use}\ \loc[_{x}]\ \comp\cdot\absev{Use}\ \loc[_{y}]\ \comp\cdot\ldots\cdot\absev{Use}\ \loc[_{x}]\ \comp\cdot\absev{Dealloc}\ \loc[_{y}]\ \comp\cdot\absev{Use}\ \loc[_{y}]\ \comp$.
-  Remembering the definition of \texttt{strncpy} (\Cref{ex:strncpy}), observe that it does not deallocate its arguments.
-  Even though the trace contains an out-of-bounds access right before returning from \texttt{strncpy}, this is no concern for \gls*{tms}, since the location $\loc[_{x}]$ is still allocated.
-  However, having returned from \texttt{strncpy}, the \texttt{greet} function continues and deallocates $\loc[_{y}]$ whose subsequent use in the \texttt{printf} call is a use-after-free bug.
-
-  The fix would be to \lstinline[language=c++,basicstyle=\small\ttfamily]|delete greetings| instead of \texttt{to} and add a \lstinline[language=c++,basicstyle=\small\ttfamily]|delete to| after the \texttt{printf} call, which leads to the abstract monitor trace $\abstrace[']=\absev{Alloc}\ \loc[_{x}]\ \comp\cdot\absev{Alloc}\ \loc[_{y}]\ \comp\cdot\absev{Use}\ \loc[_{x}]\ \comp\cdot\absev{Use}\ \loc[_{x}]\ \comp\cdot\absev{Use}\ \loc[_{y}]\ \comp\cdot\ldots\cdot\absev{Use}\ \loc[_{x}]\ \comp\cdot\absev{Dealloc}\ \loc[_{x}]\ \comp\cdot\absev{Use}\ \loc[_{y}]\ \comp$.
-  It follows that $\montmssafe{\abstrace[']}$ and from \Thmref{lem:mon:tmssafe}, it follows that the program satisfies \Thmref{def:trace:tmsdef}, even though the program still violates \gls*{sms}.
-\end{exampleenv}
+% \begin{exampleenv}[A program not satisfying \gls*{tms}]\label{ex:strncpy:cpp11}
+%   Consider the following \texttt{C++11} library that calls \texttt{strncpy} (\Cref{ex:strncpy}) and prints the result to the standard output stream.
+
+%   \begin{lstlisting}[language=c++,basicstyle=\small\ttfamily,mathescape,commentstyle=\color{gray}]
+% int hello() {
+%   char* greet = new char[11] { "Hello CSF!" };
+%   // greet address ${\color{gray}\loc_{x}}$
+%   char* to = new char[11];
+%   // to address ${\color{gray}\loc_{y}}$
+%   strncpy(11, to, greetings);
+%   delete to;
+%   printf("%cSF\n", to[6]);
+% }
+%   \end{lstlisting}
+%   Up to the body of \texttt{printf}, the program execution yields the specification trace $\ev{Alloc}\ \loc[_{x}]\ 12;\comp;\unlock\cdot\ev{Alloc}\ \loc[_{y}]\ 12;\comp;\unlock\cdot\ev{Use}\ \loc_{x}\ 0;\comp;\unlock\cdot\ev{Use}\ \loc_{x}\ 0;\comp;\unlock\cdot\ev{Use}\ \loc_{y}\ 0;\comp;\unlock\cdot\ev{Use}\ \loc_{x}\ 1;\comp;\unlock\cdot\ldots\cdot\ev{Use}\ \loc_{x}\ 12;\comp;\unlock\cdot\ev{Dealloc}\ \loc[_{y}];\comp;\unlock\cdot\ev{Use}\ \loc_{y}\ 6;\comp;\unlock$.
+%   Relating this trace to abstract monitor events yields $\abstrace=\absev{Alloc}\ \loc[_{x}]\ \comp\cdot\absev{Alloc}\ \loc[_{y}]\ \comp\cdot\absev{Use}\ \loc[_{x}]\ \comp\cdot\absev{Use}\ \loc[_{x}]\ \comp\cdot\absev{Use}\ \loc[_{y}]\ \comp\cdot\ldots\cdot\absev{Use}\ \loc[_{x}]\ \comp\cdot\absev{Dealloc}\ \loc[_{y}]\ \comp\cdot\absev{Use}\ \loc[_{y}]\ \comp$.
+%   Remembering the definition of \texttt{strncpy} (\Cref{ex:strncpy}), observe that it does not deallocate its arguments.
+%   Even though the trace contains an out-of-bounds access right before returning from \texttt{strncpy}, this is no concern for \gls*{tms}, since the location $\loc[_{x}]$ is still allocated.
+%   However, having returned from \texttt{strncpy}, the \texttt{greet} function continues and deallocates $\loc[_{y}]$ whose subsequent use in the \texttt{printf} call is a use-after-free bug.
+
+%   The fix would be to \lstinline[language=c++,basicstyle=\small\ttfamily]|delete greetings| instead of \texttt{to} and add a \lstinline[language=c++,basicstyle=\small\ttfamily]|delete to| after the \texttt{printf} call, which leads to the abstract monitor trace $\abstrace[']=\absev{Alloc}\ \loc[_{x}]\ \comp\cdot\absev{Alloc}\ \loc[_{y}]\ \comp\cdot\absev{Use}\ \loc[_{x}]\ \comp\cdot\absev{Use}\ \loc[_{x}]\ \comp\cdot\absev{Use}\ \loc[_{y}]\ \comp\cdot\ldots\cdot\absev{Use}\ \loc[_{x}]\ \comp\cdot\absev{Dealloc}\ \loc[_{x}]\ \comp\cdot\absev{Use}\ \loc[_{y}]\ \comp$.
+%   It follows that $\montmssafe{\abstrace[']}$ and from \Thmref{lem:mon:tmssafe}, it follows that the program satisfies \Thmref{def:trace:tmsdef}, even though the program still violates \gls*{sms}.
+% \end{exampleenv}
 
@@
 
-\begin{exampleenv}[Normal invocation of \texttt{strncpy}]\label{ex:strncpy:sms}
-  Consider the insecure \texttt{strncpy} function from \Cref{ex:strncpy} with a context \texttt{strncpy(2, x, y)}, where \texttt{x} and \texttt{y} are pointers to valid regions of memory with allocated space for exactly two cells and do not contain the null-terminating character \texttt{'\textbackslash 0'}.
-  For the sake of this example, the pointers have been allocated by the component and passed to the context.
-  The loop of \texttt{strncpy} will copy exactly two cells and then check the loop condition for the last time.
-  At that stage, the induction variable \texttt{i} is equal to \texttt{2} and, unfortunately, the order of checks is such that first the cell \texttt{x[i]} is read prior to bounds checking \texttt{i < n}.
-  Because of this, there is an out-of-bounds memory access right before exiting the function.
-  This is also visible on the trace, which can be sketched as $\ldots\cdot\ev{Alloc}\ \loc_{x}\ 2;\comp;\unlock\cdot\ldots\cdot\ev{Alloc}\ \loc_{y}\ 2;\comp;\unlock\cdot\ldots\cdot\ev{Use}\ \loc_{x}\ 0;\comp;\unlock\cdot\ev{Use}\ \loc_{y}\ 0;\comp;\unlock\cdot\ev{Use}\ \loc_{x}\ 1;\comp;\unlock\cdot\ev{Use}\ \loc_{y}\ 1;\comp;\unlock\cdot\ev{Use}\ \loc_{x}\ 2;\comp;\unlock\cdot\ldots$, where $\loc_{x}$ and $\loc_{y}$ are the memory addresses associated to \texttt{x} and \texttt{y}, respectively.
-  Omitting the events for all ,,$\ldots$'' for sake of brevity, the abstract monitor trace of this is $\absev{Alloc}\ \loc_{x}\ \comp\ 2\cdot\absev{Alloc}\ \loc_{y}\ \comp\ 2\cdot\absev{Use}\ \loc_{x}\ \comp\ 0\cdot\absev{Use}\ \loc_{y}\ \comp\ 0\cdot\absev{Use}\ \loc_{x}\ \comp\ 1\cdot\absev{Use}\ \loc_{y}\ \comp\ 1\cdot\absev{Use}\ \loc_{x}\ \comp\ 2$.
-
-  After the allocation events, the state of the monitor is $\left\{(\loc_{x};\comp;2),(\loc_{y};\comp;2)\right\}$.
-  All uses up to the last are accepted by the monitor, but the last event does not satisfy the premise $2<2$ in \Cref{tr:sms-use}.
-  Therefore, the whole program (\texttt{strncpy} linked with this kind of context) is not \gls*{sms}.
-\end{exampleenv}
+% \begin{exampleenv}[Normal invocation of \texttt{strncpy}]\label{ex:strncpy:sms}
+%   Consider the insecure \texttt{strncpy} function from \Cref{ex:strncpy} with a context \texttt{strncpy(2, x, y)}, where \texttt{x} and \texttt{y} are pointers to valid regions of memory with allocated space for exactly two cells and do not contain the null-terminating character \texttt{'\textbackslash 0'}.
+%   For the sake of this example, the pointers have been allocated by the component and passed to the context.
+%   The loop of \texttt{strncpy} will copy exactly two cells and then check the loop condition for the last time.
+%   At that stage, the induction variable \texttt{i} is equal to \texttt{2} and, unfortunately, the order of checks is such that first the cell \texttt{x[i]} is read prior to bounds checking \texttt{i < n}.
+%   Because of this, there is an out-of-bounds memory access right before exiting the function.
+%   This is also visible on the trace, which can be sketched as $\ldots\cdot\ev{Alloc}\ \loc_{x}\ 2;\comp;\unlock\cdot\ldots\cdot\ev{Alloc}\ \loc_{y}\ 2;\comp;\unlock\cdot\ldots\cdot\ev{Use}\ \loc_{x}\ 0;\comp;\unlock\cdot\ev{Use}\ \loc_{y}\ 0;\comp;\unlock\cdot\ev{Use}\ \loc_{x}\ 1;\comp;\unlock\cdot\ev{Use}\ \loc_{y}\ 1;\comp;\unlock\cdot\ev{Use}\ \loc_{x}\ 2;\comp;\unlock\cdot\ldots$, where $\loc_{x}$ and $\loc_{y}$ are the memory addresses associated to \texttt{x} and \texttt{y}, respectively.
+%   Omitting the events for all ,,$\ldots$'' for sake of brevity, the abstract monitor trace of this is $\absev{Alloc}\ \loc_{x}\ \comp\ 2\cdot\absev{Alloc}\ \loc_{y}\ \comp\ 2\cdot\absev{Use}\ \loc_{x}\ \comp\ 0\cdot\absev{Use}\ \loc_{y}\ \comp\ 0\cdot\absev{Use}\ \loc_{x}\ \comp\ 1\cdot\absev{Use}\ \loc_{y}\ \comp\ 1\cdot\absev{Use}\ \loc_{x}\ \comp\ 2$.
+
+%   After the allocation events, the state of the monitor is $\left\{(\loc_{x};\comp;2),(\loc_{y};\comp;2)\right\}$.
+%   All uses up to the last are accepted by the monitor, but the last event does not satisfy the premise $2<2$ in \Cref{tr:sms-use}.
+%   Therefore, the whole program (\texttt{strncpy} linked with this kind of context) is not \gls*{sms}.
+% \end{exampleenv}
 
@@
   \begin{center}
-    \begin{tikzpicture}[scale=0.6, every node/.style={transform shape}]
+    \begin{tikzpicture}[scale=0.62, every node/.style={transform shape}]
       \node[minimum width=3cm,minimum height=2.5cm,draw=black,very thick,align=left] (progunopt) {\begin{lstlisting}[language=ml,basicstyle=\small\ttfamily]
@@
 \end{lstlisting}};
-      \node[minimum width=3cm,minimum height=2.5cm,draw=black,very thick,align=left,right=1.0 of progunopt] (progoptcf) {\begin{lstlisting}[language=ml,basicstyle=\small\ttfamily]
+      \node[minimum width=3cm,minimum height=2.5cm,draw=black,very thick,align=left,right=1.0 of progunopt] (progoptcf) {\begin{lstlisting}[language=ml,basicstyle=\large\ttfamily]
 if true then
@@
 \end{lstlisting}};
-      \node[minimum width=3cm,minimum height=2.5cm,draw=black,very thick,align=left,right=1.0 of progoptcf] (progoptcfdce) {\begin{lstlisting}[language=ml,basicstyle=\small\ttfamily]
+      \node[minimum width=3cm,minimum height=2.5cm,draw=black,very thick,align=left,right=1.0 of progoptcf] (progoptcfdce) {\begin{lstlisting}[language=ml,basicstyle=\Large\ttfamily]
 print "a"
@@
 \begin{definition}[Upper Composition]
-  \bul{$\cc{\src{L}+\obj{L}}{\trg{L}}$} $\isdef$ 
+  $$\text{\bul{$\cc{\src{L}+\obj{L}}{\trg{L}}$}}\isdef
   % Given \texttt{p}, yield
-  $$\lambda \texttt{p}\ldotp
+  \lambda \texttt{p}\ldotp
   \left\{\mbox{\begin{tabular}{c}
@@
 \begin{definition}[Lower Composition]
-  \bul{$\cc{\src{L}}{\trg{L}+\obj{L}}$} $\isdef$ 
+  $$\text{\bul{$\cc{\src{L}}{\trg{L}+\obj{L}}$}}\isdef
   % Given $L$ and $\src{p}$, yield
-  $$\lambda \src{p}, L\ldotp
+  \lambda \src{p}, L\ldotp
   \left\{\mbox{\begin{tabular}{c}
@@
 Such consideration is also backed up by architecture providing a data (operand) independent timing mode, such as processors by Arm~\cite[p.~543]{arm-refman} and Intel~\cite[p.~80]{intel-refman}.
-In spirit of this, language $\Lscct$ allows violating \gls*{cct} by emitting events on, e.g., branching and division, that contain secrets (\Cref{subsec:lscct}), but provides a way to read and write to a {\em model--specific register} that enables a ``\gls*{cct}-mode''.
+In spirit of this, all languages besides $\Lscct$ are trivially \gls*{scct}. 
+$\Lscct$ (\Cref{subsec:lscct}) programs have access to a ``\gls*{cct}-mode'' and can change the leakage of emitted events according to the value of this mode (either $\obj{ON}$ or $\obj{OFF}$).
 
@@
   \begin{aligned}
-  \mi{(Types)}~\type\bnfdef\natt \mid \type[_{1}] \times \type[_{2}]\hspace{0.5cm}
-  \mi{(Libraries)}~\library\bnfdef\ \hole{\cdot} \mid \function,\library
-  \end{aligned}\\
-  \begin{aligned}
-  \mi{(Component~Names)}~\commlib\bnfdef\ \hole{\cdot} \mid \text{foo},\commlib
-  \end{aligned}
-  \\
-  \begin{aligned}
+  \mi{(Libraries)}~\library\bnfdef&\ \hole{\cdot} \mid \function,\library\hspace{0.5cm}
   \mi{(Programs)}~\prog{\library_{\ctx}}{\library_{\comp}}
@@
 \end{gather*}
-
-Above is the shared syntax of all the programming languages of this paper.
-Variables are referred to as $x,y,z,a,b,c,\ldots$ while functions may be referred to as multi-character words, such as $\texttt{foo}$, as well as short-forms like $f,g,h$.
-All languages share the type $\natt$ representing natural numbers.
-Functions are constrained to take one argument and can only call other functions listed in libraries, which are just lists of functions.
-A program $\prog{\library_{\ctx}}{\library_{\comp}}$ is indexed by two libraries that represent all context- and component-level functions, respectively.
-Lists of component-level names are referred to as $\commlib$.
-
-\begin{gather*}
-  \begin{aligned}
-    \mi{(Control~Tags)}~&\sandboxtag\bnfdef\ctx \mid \comp \\
-    \mi{(Communication~Tags)}~&\comm\bnfdef\ \ctxtocomp \mid\ \comptoctx \mid \nocomm \\
-    \mi{(Poison~Tags)}~&\poison\bnfdef \poisoned \mid \poisonless \\
-    \mi{(Continuation~Stacks)}~&\ectxstack\bnfdef\hole{\cdot} \mid \tup{K}{g},\ectxstack \\
-    \mi{(Control~Flow~States)}~&\cfstate\bnfdef\tup{\library}{\ectxstack}
-    \hspace{0.5cm}
-    \mi{(Locations)}~\loc\in\nat \\
-    \mi{(Substitutions)}~&\substvar\bnfdef \hole{\cdot} \mid \subst{\valueexpr}{x},\gamma
-  	\\
-    \mi{(Stores)}~&\Delta\bnfdef\hole{\cdot} \mid x\mapsto\tup{\loc}{\sandboxtag,\poison,n},\Delta \\
-    \mi{(Heaps)}~&H\bnfdef\hole{\cdot} | \valueexpr,H \\
-    \mi{(Memory~States)}~&\memstate\bnfdef\tup{H^{\ctx}}{H^{\comp};\Delta} \\
-    \mi{(States)}~&\statevar\bnfdef\tup{\cfstate}{\sandboxtag,\memstate} \\
-    \mi{(Runtime~Terms)}~&\runtimetermvar\bnfdef\rtt{\statevar}{\expr}
-  \end{aligned}
-\end{gather*}
-
-States $\statevar$ are tuples\footnote{Throughout the paper, the substitution notation is also used to update entries in states $\statevar$.} containing a control flow state $\cfstate$, control tags $\sandboxtag$, and a memory state $\memstate$.
-A control flow state $\cfstate$ entails a library, which provides definitions for functions calls, as well as a stack of continuations $\ectxstack$.
-Elements of the stack $\ectxstack$ are pairs of evaluation contexts and the name of a function associated to that evaluation context.
-Memory states $\memstate$ are tuples of two separate heaps $H$ and a store $\Delta$, which contains pointer metadata, such as the concrete memory location $\loc$, a control tag $\sandboxtag$ indicating which heap the pointer points into, a poison tag $\poison$, as well as bounds information $n$.
-The bounds information is a mere proof artefact that has no semantic significance.
-The two separate heaps essentially model a sandbox, to prevent contexts from performing pointer arithmetic and reading from or writing to the data owned by a component.
-While this prevents the effects of out-of-bounds accesses across the context and component boundary, the goal of the design of the languages of this paper is to be able to express security violations.
-To this end, the poison tag $\poison$ indicates whether a pointer has been freed $\poisoned$ or is still allocated $\poisonless$, so that pointers can be used even after their deallocation without the semantics getting stuck.
-Runtime terms are simply expressions $\expr$ paired with the operational state $\statevar$.
-
-\begin{gather*}
-  \begin{aligned}
-  \mi{(Pre\text{-}Events)}~\event[_{b}]\bnfdef&\ \ev{Alloc}\ \loc\ \valueexpr \mid \ev{Dealloc}\ \loc\\
-                                       &\mid \ev{Get}\ \loc\ \valueexpr \mid \ev{Set}\ \loc\ \valueexpr\ \valueexpr['] \mid \cdots \\
-  \mi{(Events)}~\event&\bnfdef \emptyevent \mid \lightning \mid \tup{\event[_{b}]}{\sandboxtag,\securitytag} \\
-  \end{aligned}
-\end{gather*}
-
-All languages use the same trace model, where events are either the empty event \emptyevent, the program crash event $\lightning$, or a tuple consisting of a control tag $\sandboxtag$ and a security tag $\securitytag$.
-The former indicates whether the component $\comp$ or the context $\ctx$ is to blame for emitting this event, the latter indicates the secrecy level of values of the emitted event, i.e., either $\lock$ or $\unlock$.
-As for pre-events, the memory-related ones are allocation ($\ev{Alloc}\ \loc\ \valueexpr$), deallocation ($\ev{Dealloc}\ \loc$), reading from ($\ev{Get}\ \loc\ \valueexpr$) and writing to memory ($\ev{Set}\ \loc\ \valueexpr\ \valueexpr[']$).
-The following is an excerpt of the operational semantics handling some of the memory operations.
-
-\begin{center}
-  \judgbox{\pstepto{\runtimetermvar}{\runtimetermvar[']}{\event}}{,,$\runtimetermvar$ does one primitive step to $\runtimetermvar[']$ emitting event $\event$.''}
-  %
-  \typerule{$e-\text{get}-\in$}{
-    \sandboxtag=\statevar.\sandboxtag &
-    \statevar.\store(x)=\tup{\loc}{\sandboxtag,\poison,m} \\
-    \loc+n\in\operatorname{dom}\statevar.\heap[^{\sandboxtag}]
-  }{
-    \pstepto{\rtt{\statevar}{\lget{x}{n}}}{\rtt{\statevar}{\statevar.\heap[^{\sandboxtag}](\loc+n)}}{\tup{\ev{Get}\ \loc\ n}{\sandboxtag}}
-  }{e-get-in}
-  %
-  \typerule{$e-\text{set}-\notin$}{
-    \sandboxtag=\statevar.\sandboxtag &
-    \statevar.\store(x)=\tup{\loc}{\sandboxtag,\poison,m} \\
-    \loc+n\notin\operatorname{dom}\statevar.\heap[^{\sandboxtag}]
-  }{
-    \pstepto{\rtt{\statevar}{\lset{x}{n}{\valueexpr}}}{\rtt{\statevar}{\valueexpr}}{\tup{\ev{Set}\ \loc\ n\ \valueexpr}{\sandboxtag}}
-  }{e-set-notin}
-  %
-  \typerule{$e-\text{new}$}{
-    \fresh{\statevar}{z} &
-    \fresh{\statevar}{\loc} &
-    \heap[^{\sandboxtag}_1]=\growh{\statevar.\heap[^{\sandboxtag}]}{n} \\
-    \store[_1]=\storeel{z}{\loc}{\statevar.\sandboxtag}{\poisonless}{n},\statevar.\store \\
-    \statevar['] = \statevar\subst{\heap[^{\sandboxtag}]}{\heap[^{\sandboxtag}_1]}\subst{\store}{\store[_1]}
-  }{
-    \pstepto{\rtt{\statevar}{\llet{x}{\lnew{\valueexpr}{n}}{e}}}{\rtt{\statevar[']}{\expr\subst{z}{x}}}{\tup{\ev{Alloc}\ \loc\ n}{\sandboxtag}}
-  }{e-new}
-  %
-  \typerule{$e-\text{dealloc}$}{
-    \statevar.\store(x) = \tup{\loc}{\statevar.\sandboxtag,\poison,n} &
-    \store[_1]=\seth{\statevar.\store}{x}{\tup{\loc}{\statevar.\sandboxtag,\poisoned,n}}
-  }{
-    \pstepto{\rtt{\statevar}{\ldelete{x}}}{\rtt{\statevar\subst{\store}{\store[_1]}}{0}}{\tup{\ev{Dealloc}\ \loc}{\sandboxtag}}
-  }{e-dealloc}
-  %
-  \typerule{$e-\text{abort}$}{
-  }{
-    \pstepto{\rtt{\statevar}{\labort}}{\lightning}{\lightning}
-  }{e-abort}
-\end{center}
-
-To demonstrate the use of the poison tag $\poison$ as metadata for pointers instead of removing them from the store $\Delta$, consider \Cref{tr:e-dealloc,tr:e-get-in,tr:e-set-notin} .
-In \Cref{tr:e-dealloc}, the premise does not care at all about the actual state of the poison tag $\poison$ and just overwrites it, marking the location as freed $\poisoned$.
-Besides that, the poison tag does not have any semantic meaning.
-For language $\Ltms$, this tag is really just some semantic metadata that programmers have no access to.
-But, for the other languages, e.g., $\Ltrg$, the poison tag is used to check pointer validity.
-\Cref{tr:e-new} allocates enough space on the respective heap, either $H^{\ctx}$ or $H^{\comp}$ depending on the execution context, i.e., the value of $\Omega.\sandboxtag$, and adds the appropriate metadata associated to the pointer in $\Delta$.
-Reading from \Cref{tr:e-get-in} and writing to memory \Cref{tr:e-set-notin} have two cases: Either the heap is large enough or not and, depending on that, either the actual value stored at that location is read from or written to, or some garbage data is returned.
-However, note that the execution does not get stuck in such cases, it performs a step, and emits an appropriate event.
-Also note that whether a pointer is poisoned or not is not reflected on the trace.
-
-\begin{center}
-  \begin{align*}
-    \mi{(Pre\text{-}Events)}~\event[_{b}]&\bnfdef \cdots \mid \ev{Call}\ \comm\ g\ \valueexpr \mid \ev{Ret}\ \comm\ \valueexpr \mid \ev{Start} \mid \ev{End}\ \valueexpr \\
-  \end{align*}
-\end{center}
-
-A key difference in comparison with the specification trace model (\Cref{subsec:propdefs}) is that, as standard in secure compilation work~\cite{patrignani2021rsc,abate2019jour,korashy2021capableptrs}, the traces have a call and return event that signals context switches, which are referred to as {\em interaction events}.
-The reason for these interaction events is technical: They are a proof artifact for reconstructing a source context from a potentially malicious target context, where during that translation, the insertion of some wrapper code right before context switching may be necessary to make the proof succeed.
-Hereby, a $\ev{Call\ \ctxtocomp\ foo\ \valueexpr}$ and $\ev{Return\ \ctxtocomp\ \valueexpr}$ signal that program execution transitions from context- to component-level.
-Contrary, $\ev{Call\ \comptoctx\ foo\ \valueexpr}$ and $\ev{Return\ \comptoctx\ \valueexpr}$ signal that program execution transitions from component- to context-level.
-For calls without this context switch, the contextual semantics attaches the $\nocomm$ tag.
-In the following, $\neg\ctx=\comp$ and $\neg\comp=\ctx$.
-
-% \begin{center}
-%   \judgbox{\estepto[]{\runtimetermvar}{\runtimetermvar}{\event}}{,,Contextual step from runtime-term $\runtimetermvar$ to $\runtimetermvar[']$ emitting event $\event$.''}
-%   %
-%   \typerule{$e-\text{ret}$}{
-%     \statevar.\ectxstack=\tup{\ectx}{\text{foo}},\ectxstack['] &
-%     \commlib\vdash\text{foo}:\Omega.\sandboxtag\dashv\comm
-%   }{
-%     \estepto[]{\rtt{\statevar}{\ectx[']\hole{\lreturn{\valueexpr}}}}{\rtt{\statevar\subst{\neg\statevar.\sandboxtag}{\sandboxtag}\subst{\ectxstack[']}{\ectxstack}}{\ectx\hole{\valueexpr}}}{\tup{\ev{Ret}\ \comm\ \valueexpr}{\statevar.\sandboxtag}}
-%   }{e-ret-notsame}
-%   %
-%   \typerule{$e-\text{call}-\text{notsame}$}{
-%     \lfunction{\text{foo}}{x}{e} \in \statevar.\library &
-%     \commlib\vdash\text{foo}:\Omega.\sandboxtag\dashv\comm
-%   }{
-%     \estepto[]{\rtt{\statevar}{\ectx\hole{\lcall{\text{foo}}{\valueexpr}}}}{\rtt{\statevar\subst{\neg\statevar.\sandboxtag}{\sandboxtag}\subst{\tup{\ectx}{\text{foo}},\statevar.\ectxstack}{\ectxstack}}{\expr\subst{\valueexpr}{x}}}{\tup{\ev{Call}\ \comm\ \text{foo}\ \valueexpr}{\statevar.\sandboxtag}}
-%   }{e-call-notsame}
-% \end{center}
-
-The contextual semantics is mostly straightforward.
-In \Cref{tr:e-ret-notsame,tr:e-call-notsame}, the judgement $\Omega.\commlib\vdash\text{foo}:\Omega.\sandboxtag\dashv \comm$ checks whether $\text{foo}$ is a component-level name by looking it up in the list of component-level names $\commlib$ and emits the appropriate transfer tag, i.e., either $\comptoctx$ or $\ctxtocomp$.
-Additional rules that are left out ensure that, e.g., when calling the $\text{main}$ function, the event $\ev{Start}$ is emitted, which is a design choice this paper does for convenience when reasoning about call-chains.
-Note that the $\ev{End}\ \valueexpr$ event is not emitted if the program crashes.
-
-% \MKin{%
-%   should I add a rule for top level exec or keep it like this?
-% 	MP: like this is fine
-% }
-The top-level execution $\progstepto{\prog{\library_{\ctx}}{\library_{\comp}}}{\runtimetermvar}{\trace}$ constructs an initial state $\Omega$ by linking $\library_{\ctx}$ and $\library_{\ctx}$ and then starts execution by calling the $\text{main}$ function.
-The trace $\trace$ emitted during that execution serves as abstraction of the behavior of the program enabling the use of \Cref{def:propsat,def:proprsat}.
 
@@
 
-\begin{center}
-  \typerule{$e-\trg{\lhast{x}{\natt}}$}{
-  }{
-    \pstepto{\rtt{\trg{\statevar}}{\trg{\lhast{x}{\natt}}}}{\rtt{\trg{\statevar}}{\trg{1}}}{\emptyevent}
-  }{e-has-nat-no}
-  %
-  \typerule{$e-\trg{\lhast{n}{\natt}}$}{
-  }{
-    \pstepto{\rtt{\trg{\statevar}}{\trg{\lhast{n}{\natt}}}}{\rtt{\trg{\statevar}}{\trg{0}}}{\emptyevent}
-  }{e-has-nat-yes}
-\end{center}
-
 Dynamic typechecks $\trg{\lhast{\expr}{\type}}$ match on $\trg{\expr}$ and evaluate to $\trg{0}$ if the term is of type $\trg{\type}$ and $\trg{1}$ otherwise.
@@
 \subsection{$\Lms$: Another Memory-Unsafe Language}\label{subsec:lms}
-To enhance readablity, this paper uses $\Lms$, despite it being exactly equal to $\Ltrg$ (\Cref{subsec:lsms}).
-The projection $\msfilterLms{\irl{\trace}}$ is also exactly equal to $\msfilterL{\trg{\trace}}$.
+$\Lms$ is exactly equal to $\Ltrg$ (\Cref{subsec:lsms}), but used to signify that the code considered is after applying $\cc{\Ltrg}{\Lms}$.
 
@@
 The following rules demonstrate how this is handled semantically.
-
-\begin{center}
-  \typerule{$e-\obj{\oplus}-\text{noleak}$}{
-    \obj{m}\not=\obj{0} &
-    \obj{n_3}=\obj{n_1}\oplus\obj{n_2} \\
-    \obj{\securitytag['']}\le\obj{\securitytag} &
-    \obj{\securitytag['']}\le\obj{\securitytag[']}
-  }{
-    \pstepto{\rtt{\obj{\cfstate;\sandboxtag;m;\memstate}}{\obj{\lbinop{n_1^{\securitytag}}{n_2^{\securitytag[']}}}}}{\rtt{\obj{\cfstate;\sandboxtag;m;\memstate}}{\obj{n_3^{\securitytag['']}}}}{\obj{\emptyevent}}
-  }{e-binop-noleak}
-  %
-  \typerule{$e-\obj{wrdoit}-\text{off}$}{
-  }{
-    \pstepto{\rtt{\obj{\cfstate;\sandboxtag;m;\memstate}}{\obj{\lwrdoit{OFF}}}}{\rtt{\obj{\cfstate;\sandboxtag;OFF;\memstate}}{\obj{0^{\securitytag}}}}{\obj{\emptyevent}}
-  }{e-wrdoit-off}
-  %
-  \typerule{$e-\obj{ifz}-\obj{true}-\text{leak}$}{
-  }{
-    \pstepto{\rtt{\obj{\cfstate;\sandboxtag;0;\memstate}}{\obj{\lifz{0^{\securitytag}}{\expr[_1]}{\expr[_2]}}}}{\rtt{\obj{\cfstate;\sandboxtag;0;\memstate}}{\obj{\expr[_2]}}}{\obj{\ev{Branch}\ 0;\sandboxtag;\securitytag}}
-  }{e-ifz-true-leak}
-\end{center}
 
@@
 \begin{figure*}[!h]
-  \begin{center}
+  \centering
   \begin{tikzpicture}
@@
     \node[right=0.5 of sectms] (secsms) {{\scriptsize\Cref{subsec:cs:ms}}};
-      \node[right=1.75 of secsms] (secopts) {{\scriptsize\Cref{subsec:cs:opts}}};
-      \node[right=1.75 of secopts] (secscct) {{\scriptsize\Cref{subsec:cs:scct}}};
+    \node[right=1.5 of secsms] (secopts) {{\scriptsize\Cref{subsec:cs:opts}}};
+    \node[right=1.25 of secopts] (secscct) {{\scriptsize\Cref{subsec:cs:scct}}};
 
@@
 
-      \draw[dashed] ($(sectms)-(0.8,0)$) -- ($(sectms)-(0.8,1.25)$);
-      \draw[dashed] ($(sectms)-(-0.9,0)$) -- ($(sectms)-(-0.9,1.25)$);
+    \draw[dashed] ($(sectms)-(1.2,0)$) -- ($(sectms)-(1.2,1.25)$);
+    \draw[dashed] ($(sectms)-(-1.0,0)$) -- ($(sectms)-(-1.0,1.25)$);
     \draw[dashed] ($(secsms)-(-1.2,0)$) -- ($(secsms)-(-1.2,1.25)$);
-      \draw[dashed] ($(secscct)-(1.2,0)$) -- ($(secscct)-(1.2,1.25)$);
-      \draw[dashed] ($(secscct)-(-1.1,0)$) -- ($(secscct)-(-1.1,1.25)$);
+    \draw[dashed] ($(secscct)-(1.1,0)$) -- ($(secscct)-(1.1,1.25)$);
+    \draw[dashed] ($(secscct)-(-1.4,0)$) -- ($(secscct)-(-1.4,1.25)$);
   \end{tikzpicture}
-  \end{center}
+  \vspace{-2.5em}
   \caption{Visualisation of the optimising compilation pipeline that attains a combination of \gls*{ms} and \gls*{cct}. %
@@
   \end{align*}\\[0.125cm]
-  \begin{align*}
-    \partialeval{\irl{x}}{\irl{\substlist}} & = \irl{n} & \text{if } \irl{\subst{n}{x}}\in\irl{\substlist} \\
-    \partialeval{\irl{x}}{\irl{\substlist}} & = \irl{x} & \text{if } \irl{\subst{n}{x}}\notin\irl{\substlist} \\
-    \partialeval{\irl{\lbinop{n}{m}}}{\irl{\substlist}} & = \irl{k} & \text{if } \lbinop{\irl{n}}{\irl{m}}=k \\
+% \begin{align*}
+%   \partialeval{\irl{x}}{\irl{\substlist}} & = \irl{n} & \text{if } \irl{\subst{n}{x}}\in\irl{\substlist} \\
+%   \partialeval{\irl{x}}{\irl{\substlist}} & = \irl{x} & \text{if } \irl{\subst{n}{x}}\notin\irl{\substlist} \\
+%   \partialeval{\irl{\lbinop{n}{m}}}{\irl{\substlist}} & = \irl{k} & \text{if } \lbinop{\irl{n}}{\irl{m}}=k \\
 %    \partialeval{\irl{\lbinop{\expr[_{1}]}{\expr[_{2}]}}}{\irl{\substlist}} & = \lbinop[\irl]{\partialeval{\irl{\expr[_{1}]}}{\irl{\substlist}}}{\partialeval{\irl{\expr[_{2}]}}{\irl{\substlist}}} \\
-    \partialeval{\irl{\llet{x}{n}{\expr}}}{\irl{\substlist}} & = \partialeval{\irl{\expr}}{\irl{\subst{x}{n}},\irl{\substlist}} \\
-    \partialeval{\irl{\lget{x}{\expr}}}{\irl{\substlist}} & = \lget[\irl]{\irl{x}}{\partialeval{\irl{\expr}}{\irl{\substlist}}}
-  \end{align*}\\
-  \begin{align*}
-    \partialeval{\irl{\llet{x}{\expr[_{1}]}{\expr[_{2}]}}}{\irl{\substlist}} & = \llet[\irl]{\irl{x}}{\partialeval{\irl{\expr[_{1}]}}{\irl{\substlist}}}{\\&\partialeval{\irl{\expr[_{2}]}}{\irl{\substlist}}} \\
-    \partialeval{\irl{\lifz{\expr[_{1}]}{\expr[_{2}]}{\expr[_{3}]}}}{\irl{\substlist}} & = \lifz[\irl]{\partialeval{\irl{\expr[_{1}]}}{\irl{\substlist}}}{\\&\partialeval{\irl{\expr[_{2}]}}{\irl{\substlist}}}{\partialeval{\irl{\expr[_{3}]}}{\irl{\substlist}}} \\
-  \end{align*}
+%   \partialeval{\irl{\llet{x}{n}{\expr}}}{\irl{\substlist}} & = \partialeval{\irl{\expr}}{\irl{\subst{x}{n}},\irl{\substlist}} \\
+%   \partialeval{\irl{\lget{x}{\expr}}}{\irl{\substlist}} & = \lget[\irl]{\irl{x}}{\partialeval{\irl{\expr}}{\irl{\substlist}}}
+% \end{align*}\\
+% \begin{align*}
+%   \partialeval{\irl{\llet{x}{\expr[_{1}]}{\expr[_{2}]}}}{\irl{\substlist}} & = \llet[\irl]{\irl{x}}{\partialeval{\irl{\expr[_{1}]}}{\irl{\substlist}}}{\\&\partialeval{\irl{\expr[_{2}]}}{\irl{\substlist}}} \\
+%   \partialeval{\irl{\lifz{\expr[_{1}]}{\expr[_{2}]}{\expr[_{3}]}}}{\irl{\substlist}} & = \lifz[\irl]{\partialeval{\irl{\expr[_{1}]}}{\irl{\substlist}}}{\\&\partialeval{\irl{\expr[_{2}]}}{\irl{\substlist}}}{\partialeval{\irl{\expr[_{3}]}}{\irl{\substlist}}} \\
+% \end{align*}
 \end{gather*}
@@
 The \gls*{dce} pass applies a naive rewrite rule on conditionals.
-For \gls*{cf}, the pass uses an auxiliary function \texttt{mix} that does the actual work.
+For \gls*{cf}, the pass uses an auxiliary function \texttt{mix} that does the actual work:
 It rewrites constant binary operations, e.g., $\irl{{17}-1}$ to $\irl{16}$, and replaces variables that are assigned to constants with their constant, e.g., $\irl{\llet{x}{7}{x}}$ to $\irl{7}$.
@@
 \subsection{Robust Strict Cryptographic Constant Time Preservation}\label{subsec:cs:scct}
-
+%
 \begin{center}
@@
 \end{center}
-
+%
 Given the fact that $\Lscct$ provides a \gls*{cct}-mode that can be turned on or off, the compiler inserts wrapper code for function bodies to ensure that execution in the component always happen in this \gls*{cct}-mode.

### 2023-09-22 csf/main.tex
Text

@@
 
-This section presents the key meta-theoretic results of this paper concerning sequential compiler composition (and of optimisation passes) (\Cref{sec:sequential}) and concerning other kinds of compiler composition (\Cref{sec:other-compos}).
+This section presents the key meta-theoretic results of this paper concerning secure compiler compositions (\Cref{sec:sequential,sec:other-compos}).
+Note that the presented framework works with arbitrary classes.
 % and builds on prior work~\cite{kruse2022csc}\MK{mention this at all?
@@
 Since the composition of secure compilers is again a secure compiler, the theorem generalises to a whole chain of $n$ secure compilers.
+
+\textbf{Triviality.} Note that the theorem may give trivial results, i.e., the intersection of classes $\class[_1]$ and $\class[_2]$ may be empty. 
+For example, consider $\rtp{\cc{\src{L}}{\trg{L}}}{\tmssafe}$ and $\rtp{\cc{\trg{L}}{\obj{L}}}{\smssafe}$, where $\cc{\trg{L}}{\obj{L}}$ introduces code that allocates a table that stores bounds information of pointers, but fails to insert code that deallocates this table.
+By \Cref{def:trace:tmsdef}, this violates $\tmssafe$.
+Therefore, no behavior resulting from programs produced by the $\cc{\trg{L}}{\obj{L}}$ compiler is secure with respect to $\tmssafe$, but only for $\smssafe$.
+In turn, % TODO TODO TODO FIXME
 

### 2023-09-25 csf/main.tex
Add note

@@
 
-\textbf{Triviality.} Note that the theorem may give trivial results, i.e., the intersection of classes $\class[_1]$ and $\class[_2]$ may be empty. 
+Note that the theorem may not be applicable in certain cases.
 For example, consider $\rtp{\cc{\src{L}}{\trg{L}}}{\tmssafe}$ and $\rtp{\cc{\trg{L}}{\obj{L}}}{\smssafe}$, where $\cc{\trg{L}}{\obj{L}}$ introduces code that allocates a table that stores bounds information of pointers, but fails to insert code that deallocates this table.
@@
 Therefore, no behavior resulting from programs produced by the $\cc{\trg{L}}{\obj{L}}$ compiler is secure with respect to $\tmssafe$, but only for $\smssafe$.
-In turn, % TODO TODO TODO FIXME
+This, however, is a requirement of the preservation statement (the property of interest must be in $\class[_1]\cap\class[_2]$).
 

### 2023-09-25 csf/main.tex
Do TODOs

@@
 the memory location past the last element will be leaked to an attacker.
-\begin{lstlisting}[language=c,basicstyle=\scriptsize\ttfamily,morekeywords={size_t}]
+\begin{lstlisting}[language=c,basicstyle=\footnotesize\ttfamily,morekeywords={size_t}]
 void strncpy(size_t n, char *dst, char *src) {
@@
 \begin{definition}[Property Satisfaction]\label{def:propsat}
+  $\;$\\
+  \vspace{-1em}
+  \begin{nscenter}
     \bul{$\sat{\progvar}{\pi}$}
@@
     then \oul{$\trace\in\pi$}.
+  \end{nscenter}
 \end{definition}
@@
 \begin{definition}[Robust Satisfaction]\label{def:proprsat}
+  $\;$\\
+  \vspace{-1em}
+  \begin{nscenter}
   \bul{$\rsat{\progvar}{\pi}$}
   $\isdef$ \iul{$\forall \contextvar\ \wholeprogvar$, if $\link{\contextvar}{\progvar}=\wholeprogvar$}, then \oul{$\sat{\wholeprogvar}{\pi}$}.
+  \end{nscenter}
 \end{definition}
@@
 \begin{definition}[Robust Compilation]\label{def:rtp}
+  $\;$\\
+  \vspace{-1em}
+  \begin{nscenter}
   % Compiler $\cc{\src{L}}{\trg{L}}$ robustly preserves $\class$, 
@@
   $\isdef$
-  \rul{$\forall \left(\pi\in\class\right) \left(\src{p}\in\src{L}\right),$} if \iul{$\rsat{\src{\progvar}}{\pi}$}, then \oul{$\rsat{\cc{\src{L}}{\trg{L}}\left(\src{p}\right)}{\pi}$}.
+  \rul{$\forall \pi\in\class, \src{p}\in\src{L},$} if \iul{$\rsat{\src{\progvar}}{\pi}$}, then \oul{$\rsat{\cc{\src{L}}{\trg{L}}\left(\src{p}\right)}{\pi}$}.
+  \end{nscenter}
 \end{definition}
@@
 
+\vspace{-2em}
 \begin{gather*}
@@
 
+While \gls*{scct} may seem overly strict, since it may seem that secrets must not occur on a trace (since $\securitytag$ is forced to be $\unlock$). 
+However, programs that operate on secrets and have a ''data independent timing mode'' enabled do not leak these secrets, despite performing computation on them (see \Cref{ex:ditm}).
+
 \subsubsection{Memory Safe, Strict Cryptographic Constant Time}
@@
 The combination of \gls*{ms} and \gls*{scct} is the intersection of these properties, \gls*{msscct}.
-Since \gls*{ms} and \gls*{scct} are just sets of traces that, intuitively, contain all program behaviors that follow a security policy, the intersection of them contains all program behaviors that follow both security policies, i.e., it entails all program behaviours that are both \gls*{ms} and \gls*{scct}.
+%Since \gls*{ms} and \gls*{scct} are just sets of traces that, intuitively, contain all program behaviors that follow a security policy, the intersection of them contains all program behaviors that follow both security policies, i.e., it entails all program behaviours that are both \gls*{ms} and \gls*{scct}.
 
@@
 \begin{definition}[Monitor Satisfaction]\label{def:monsat}
+  $\;$ 
+  \begin{nscenter}
   \bul{$\monsafe{\trace}{\pi}$} $\isdef$ $\exists \abstrace\ \monitor$, \iul{$\traceagree{\pi}{\trace}{\abstrace}$} and \oul{$\monitorcheck{\emptyset}{\monitor}{\abstrace}$}.
+  \end{nscenter}
 \end{definition}
@@
 \begin{lemma}[Traces with Monitor Satisfaction are $\tmssafe$]\label{lem:mon:tmssafe}
-  If $\montmssafe{\trace}$, then $\trace\in\tmssafe$. %\Coqed
+  $\;\;$ If $\montmssafe{\trace}$, then $\trace\in\tmssafe$. %\Coqed
 \end{lemma}
@@
   \typerule{sms-alloc}{
-    (\loc;\sandboxtag;m)\notin\smsmonitor
+    (\loc;\sandboxtag;m)\notin\smsmonitor %\\
+    % unfortunately, this does not render sms-use and sms-alloc side by side
+    % my only other idea is \scalebox...ugh
+    %\smsmonitor[']=\smsmonitor\cup\left\{(\loc;\sandboxtag;n)\right\}
   }{
@@
 \begin{lemma}[Traces with Monitor Satisfaction are $\smssafe$]\label{lem:mon:smsafe}
-  If $\monsmssafe{\trace}$, then $\trace\in\smssafe$.\Coqed
+  $\;\;$
+  If $\monsmssafe{\trace}$, then $\trace\in\smssafe$. %\Coqed
 \end{lemma}
@@
 
+Any technique to combine existing monitors $T_A$, $T_B$ (that check for properties $A$ and $B$) into one $T_{AB}$ (that checks for both $A$ and $B$) is sufficient for the sake of the analysis of this paper.
+Note that monitors only apply to trace-properties, but not, e.g., relational hyperproperties~\cite{abate2019jour}.
+
 \subsubsection{Monitor for \glsfirst*{scct}}
@@
 
-\begin{exampleenv}[Data-independent timing mode]
+\begin{exampleenv}[Data-independent timing mode]\label{ex:ditm}
   Consider the call \texttt{strncpy(1,x,y)} to the \texttt{strncpy} function (\Cref{ex:strncpy}) with low ($\unlock$) security for \texttt{x} and high security ($\lock$) for \texttt{y}.
@@
 \begin{theorem}[Lower Composition of Secure Compilers]\label{thm:lrtp}
-  If \bul{$\rtp{\cc{\src{L}}{\trg{L}}}{\class[_{1}]}$} and \rul{$\rtp{\cc{\obj{L}}{\trg{L}}}{\class[_{2}]}$}, then \oul{$\rtp{\cc{\src{L}}{\trg{L}+\obj{L}}}{\class[_{1}]\cap\class[_{2}]}$}. %\Coqed
+  If \bul{$\rtp{\cc{\src{L}}{\trg{L}}}{\class[_{1}]}$} and \rul{$\rtp{\cc{\obj{L}}{\trg{L}}}{\class[_{2}]}$}, then \oul{$\rtp{\cc{\src{L}}{\trg{L}+\obj{L}}}{\class[_{1}]\cap\class[_{2}]}$}. \Coqed
 \end{theorem}
@@
 %Either way, the theoretical results suggest that it is possible to always find a ,,most-general'', secure compiler, given two secure compilers, that robustly preserves the least-upper bound of the classes involved in their compilation process.
+
+\subsection{Applicability of the Compositionality Framework}
+
+It is worth emphasizing that the framework is general: It supports both ordinary trace properties as well as hyperproperties or even relational hyperproperties~\cite{abate2019jour}.
+Moreover, while this paper assumes a unified trace-model, it is merely a technical exercise to allow for different trace-models between programming languages~\cite{abate2021extacc}.
 
@@
 
+\vspace{-2em}
 \begin{center}
@@
 \end{center}
+\vspace{-2em}
 
@@
 
+\vspace{-1em}
 \begin{gather*}
@@
 \begin{exampleenv}[$\Lscct$ with and without data independent timing]\label{ex:lscct}
-  Consider again the context presented in \Cref{ex:strncpy:cpp11}, where everything is marked with a security tag of high $\lock$.
+  Consider again \Cref{ex:strncpy}, with a context copying the string \texttt{Hello World}, where everything is marked with a security tag of high $\lock$.
   The following table shows parts of the execution trace, read from top to bottom, in the left column (\textit{Active}) with and in the right column (\textit{Inactive}) without data independent timing.
@@
 
+\vspace{-2em}
   $${\footnotesize
@@
 
+\vspace{-1em}
 \begin{center}
@@
   Consider again \texttt{strncpy}, but instrumented for \gls*{sms}:
-    \begin{lstlisting}[language=c,basicstyle=\ttfamily\scriptsize, morekeywords={size_t}]
+    \begin{lstlisting}[language=c,basicstyle=\ttfamily\footnotesize, morekeywords={size_t}]
 void strncpy(size_t n, size_t dst_size, char *dst,
@@
     \end{lstlisting}
-    When calling this in similar fashion to \Cref{ex:strncpy:sms}, the event $\ev{Use}\ \loc_{x}\ 2;\comp;\unlock$ would not be emitted during execution, since the bounds check prevents the condition \texttt{src[i] != '\textbackslash 0'} from executing.
+    When calling this with \texttt{strncpy(2, x, y)}, where \texttt{x} and \texttt{y} are pointers to valid regions of memory with allocated space for exactly two cells and do not contain the null-terminating character \texttt{'\textbackslash 0'}, the event $\ev{Use}\ \loc_{x}\ 2;\comp;\unlock$ would not be emitted during execution, since the bounds check prevents the condition \texttt{src[i] != '\textbackslash 0'} from executing.
 \end{exampleenv}
@@
 \end{gather*}
+\vspace{-3em}
 
@@
 The proof for either is relatively simple, because both \gls*{dce} and \gls*{cf} do not change the way memory accesses happen.
-Moreover, since the input and output languages to these compilers are the same, attacker contexts do not have more power in the target language than in the source.
+Moreover, since the input and output languages to these compilers are the same, attacker contexts do not have more power in the target language than in the source, since both languages are equal.
 
@@
 \subsection{Robust Strict Cryptographic Constant Time Preservation}\label{subsec:cs:scct}
-%
+
+\vspace{-2em}
 \begin{center}
@@
 \end{center}
+\vspace{-2em}
 %
@@
 Then, it proved that composing secure compilers that preserve certain properties results in a secure compiler that preserves the composition of these properties.
+While the presented security property composition relied on monitors that check only trace-properties, the composition of secure compilers does not make any restriction towards the kind of property involved in the composition.
+It is subject to future work to develop techniques to verify relational hyperproperties~\cite{abate2019jour}, while the composition of hyperproperties could be very similar to the composition of ordinary properties as presented in this paper, since hyperproperties can be checked with an automata-based model-checker~\cite{beutner23hyperltl}.
 % 
@@
 Crucially, this paper derives the security of the multi-pass compiler from the composition of the security properties preserved by its individual passes, which include security-preserving as well as optimisation passes.
+For future work, it is interesting to look at more sophisticated optimisation passes that, e.g., reorder events that appear on traces.
 

### 2023-09-25 csf/main.tex
Uncomment examples

@@
 
-% \begin{exampleenv}[Double Free in Bluetooth Subsystem]
-%   Consider \texttt{CVE-2021-3564}~\cite{doublefree-bluetooth}, one of many submissions for a double-free vulnerability.
-%   The vulnerability arises due to a race condition where the context-level function \texttt{hci\_cmd\_work} was not expected to behave maliciously, since it resides in the same source-code repository where the vulnerability occurs.
-%   Nevertheless, the component-level code of \texttt{hci\_dev\_do\_open} is linked with \texttt{hci\_cmd\_work} and does not atomically check whether a pointer has been freed already:
-%   Therefore, \texttt{hci\_dev\_do\_open} does not satisfy the no-double-frees property robustly, since there is an implementation for \texttt{hci\_cmd\_work} that leads to a violation of that property when linked with \texttt{hci\_dev\_do\_open}.
-% \end{exampleenv}
+\begin{exampleenv}[Double Free in Bluetooth Subsystem]
+  Consider \texttt{CVE-2021-3564}~\cite{doublefree-bluetooth}, one of many submissions for a double-free vulnerability.
+  The vulnerability arises due to a race condition where the context-level function \texttt{hci\_cmd\_work} was not expected to behave maliciously, since it resides in the same source-code repository where the vulnerability occurs.
+  Nevertheless, the component-level code of \texttt{hci\_dev\_do\_open} is linked with \texttt{hci\_cmd\_work} and does not atomically check whether a pointer has been freed already:
+  Therefore, \texttt{hci\_dev\_do\_open} does not satisfy the no-double-frees property robustly, since there is an implementation for \texttt{hci\_cmd\_work} that leads to a violation of that property when linked with \texttt{hci\_dev\_do\_open}.
+\end{exampleenv}
 
@@
 
-% \begin{exampleenv}[A program not satisfying \gls*{tms}]\label{ex:strncpy:cpp11}
-%   Consider the following \texttt{C++11} library that calls \texttt{strncpy} (\Cref{ex:strncpy}) and prints the result to the standard output stream.
-
-%   \begin{lstlisting}[language=c++,basicstyle=\small\ttfamily,mathescape,commentstyle=\color{gray}]
-% int hello() {
-%   char* greet = new char[11] { "Hello CSF!" };
-%   // greet address ${\color{gray}\loc_{x}}$
-%   char* to = new char[11];
-%   // to address ${\color{gray}\loc_{y}}$
-%   strncpy(11, to, greetings);
-%   delete to;
-%   printf("%cSF\n", to[6]);
-% }
-%   \end{lstlisting}
-%   Up to the body of \texttt{printf}, the program execution yields the specification trace $\ev{Alloc}\ \loc[_{x}]\ 12;\comp;\unlock\cdot\ev{Alloc}\ \loc[_{y}]\ 12;\comp;\unlock\cdot\ev{Use}\ \loc_{x}\ 0;\comp;\unlock\cdot\ev{Use}\ \loc_{x}\ 0;\comp;\unlock\cdot\ev{Use}\ \loc_{y}\ 0;\comp;\unlock\cdot\ev{Use}\ \loc_{x}\ 1;\comp;\unlock\cdot\ldots\cdot\ev{Use}\ \loc_{x}\ 12;\comp;\unlock\cdot\ev{Dealloc}\ \loc[_{y}];\comp;\unlock\cdot\ev{Use}\ \loc_{y}\ 6;\comp;\unlock$.
-%   Relating this trace to abstract monitor events yields $\abstrace=\absev{Alloc}\ \loc[_{x}]\ \comp\cdot\absev{Alloc}\ \loc[_{y}]\ \comp\cdot\absev{Use}\ \loc[_{x}]\ \comp\cdot\absev{Use}\ \loc[_{x}]\ \comp\cdot\absev{Use}\ \loc[_{y}]\ \comp\cdot\ldots\cdot\absev{Use}\ \loc[_{x}]\ \comp\cdot\absev{Dealloc}\ \loc[_{y}]\ \comp\cdot\absev{Use}\ \loc[_{y}]\ \comp$.
-%   Remembering the definition of \texttt{strncpy} (\Cref{ex:strncpy}), observe that it does not deallocate its arguments.
-%   Even though the trace contains an out-of-bounds access right before returning from \texttt{strncpy}, this is no concern for \gls*{tms}, since the location $\loc[_{x}]$ is still allocated.
-%   However, having returned from \texttt{strncpy}, the \texttt{greet} function continues and deallocates $\loc[_{y}]$ whose subsequent use in the \texttt{printf} call is a use-after-free bug.
-
-%   The fix would be to \lstinline[language=c++,basicstyle=\small\ttfamily]|delete greetings| instead of \texttt{to} and add a \lstinline[language=c++,basicstyle=\small\ttfamily]|delete to| after the \texttt{printf} call, which leads to the abstract monitor trace $\abstrace[']=\absev{Alloc}\ \loc[_{x}]\ \comp\cdot\absev{Alloc}\ \loc[_{y}]\ \comp\cdot\absev{Use}\ \loc[_{x}]\ \comp\cdot\absev{Use}\ \loc[_{x}]\ \comp\cdot\absev{Use}\ \loc[_{y}]\ \comp\cdot\ldots\cdot\absev{Use}\ \loc[_{x}]\ \comp\cdot\absev{Dealloc}\ \loc[_{x}]\ \comp\cdot\absev{Use}\ \loc[_{y}]\ \comp$.
-%   It follows that $\montmssafe{\abstrace[']}$ and from \Thmref{lem:mon:tmssafe}, it follows that the program satisfies \Thmref{def:trace:tmsdef}, even though the program still violates \gls*{sms}.
-% \end{exampleenv}
+\begin{exampleenv}[A program not satisfying \gls*{tms}]\label{ex:strncpy:cpp11}
+  Consider the following \texttt{C++11} library that calls \texttt{strncpy} (\Cref{ex:strncpy}) and prints the result to the standard output stream.
+
+  \begin{lstlisting}[language=c++,basicstyle=\small\ttfamily,mathescape,commentstyle=\color{gray}]
+int hello() {
+  char* greet = new char[12] { "Hello CSF!!" };
+  // greet address ${\color{gray}\loc_{x}}$
+  char* to = new char[12];
+  // to address ${\color{gray}\loc_{y}}$
+  strncpy(11, to, greetings);
+  delete to;
+  printf("%cSF\n", to[6]);
+}
+  \end{lstlisting}
+  Up to the body of \texttt{printf}, the program execution yields the specification trace $\ev{Alloc}\ \loc[_{x}]\ 12;\comp;\unlock\cdot\ev{Alloc}\ \loc[_{y}]\ 12;\comp;\unlock\cdot\ev{Use}\ \loc_{x}\ 0;\comp;\unlock\cdot\ev{Use}\ \loc_{x}\ 0;\comp;\unlock\cdot\ev{Use}\ \loc_{y}\ 0;\comp;\unlock\cdot\ev{Use}\ \loc_{x}\ 1;\comp;\unlock\cdot\ldots\cdot\ev{Use}\ \loc_{x}\ 12;\comp;\unlock\cdot\ev{Dealloc}\ \loc[_{y}];\comp;\unlock\cdot\ev{Use}\ \loc_{y}\ 6;\comp;\unlock$.
+  Relating this trace to abstract monitor events yields $\abstrace=\absev{Alloc}\ \loc[_{x}]\ \comp\cdot\absev{Alloc}\ \loc[_{y}]\ \comp\cdot\absev{Use}\ \loc[_{x}]\ \comp\cdot\absev{Use}\ \loc[_{x}]\ \comp\cdot\absev{Use}\ \loc[_{y}]\ \comp\cdot\ldots\cdot\absev{Use}\ \loc[_{x}]\ \comp\cdot\absev{Dealloc}\ \loc[_{y}]\ \comp\cdot\absev{Use}\ \loc[_{y}]\ \comp$.
+  Remembering the definition of \texttt{strncpy} (\Cref{ex:strncpy}), observe that it does not deallocate its arguments.
+  Even though the trace contains an out-of-bounds access right before returning from \texttt{strncpy}, this is no concern for \gls*{tms}, since the location $\loc[_{x}]$ is still allocated.
+  However, having returned from \texttt{strncpy}, the \texttt{greet} function continues and deallocates $\loc[_{y}]$ whose subsequent use in the \texttt{printf} call is a use-after-free bug.
+
+  The fix would be to \lstinline[language=c++,basicstyle=\small\ttfamily]|delete greetings| instead of \texttt{to} and add a \lstinline[language=c++,basicstyle=\small\ttfamily]|delete to| after the \texttt{printf} call, which leads to the abstract monitor trace $\abstrace[']=\absev{Alloc}\ \loc[_{x}]\ \comp\cdot\absev{Alloc}\ \loc[_{y}]\ \comp\cdot\absev{Use}\ \loc[_{x}]\ \comp\cdot\absev{Use}\ \loc[_{x}]\ \comp\cdot\absev{Use}\ \loc[_{y}]\ \comp\cdot\ldots\cdot\absev{Use}\ \loc[_{x}]\ \comp\cdot\absev{Dealloc}\ \loc[_{x}]\ \comp\cdot\absev{Use}\ \loc[_{y}]\ \comp$.
+  It follows that $\montmssafe{\abstrace[']}$ and from \Thmref{lem:mon:tmssafe}, it follows that the program satisfies \Thmref{def:trace:tmsdef}, even though the program still violates \gls*{sms}.
+\end{exampleenv}
 
@@
 
-% \begin{exampleenv}[Normal invocation of \texttt{strncpy}]\label{ex:strncpy:sms}
-%   Consider the insecure \texttt{strncpy} function from \Cref{ex:strncpy} with a context \texttt{strncpy(2, x, y)}, where \texttt{x} and \texttt{y} are pointers to valid regions of memory with allocated space for exactly two cells and do not contain the null-terminating character \texttt{'\textbackslash 0'}.
-%   For the sake of this example, the pointers have been allocated by the component and passed to the context.
-%   The loop of \texttt{strncpy} will copy exactly two cells and then check the loop condition for the last time.
-%   At that stage, the induction variable \texttt{i} is equal to \texttt{2} and, unfortunately, the order of checks is such that first the cell \texttt{x[i]} is read prior to bounds checking \texttt{i < n}.
-%   Because of this, there is an out-of-bounds memory access right before exiting the function.
-%   This is also visible on the trace, which can be sketched as $\ldots\cdot\ev{Alloc}\ \loc_{x}\ 2;\comp;\unlock\cdot\ldots\cdot\ev{Alloc}\ \loc_{y}\ 2;\comp;\unlock\cdot\ldots\cdot\ev{Use}\ \loc_{x}\ 0;\comp;\unlock\cdot\ev{Use}\ \loc_{y}\ 0;\comp;\unlock\cdot\ev{Use}\ \loc_{x}\ 1;\comp;\unlock\cdot\ev{Use}\ \loc_{y}\ 1;\comp;\unlock\cdot\ev{Use}\ \loc_{x}\ 2;\comp;\unlock\cdot\ldots$, where $\loc_{x}$ and $\loc_{y}$ are the memory addresses associated to \texttt{x} and \texttt{y}, respectively.
-%   Omitting the events for all ,,$\ldots$'' for sake of brevity, the abstract monitor trace of this is $\absev{Alloc}\ \loc_{x}\ \comp\ 2\cdot\absev{Alloc}\ \loc_{y}\ \comp\ 2\cdot\absev{Use}\ \loc_{x}\ \comp\ 0\cdot\absev{Use}\ \loc_{y}\ \comp\ 0\cdot\absev{Use}\ \loc_{x}\ \comp\ 1\cdot\absev{Use}\ \loc_{y}\ \comp\ 1\cdot\absev{Use}\ \loc_{x}\ \comp\ 2$.
-
-%   After the allocation events, the state of the monitor is $\left\{(\loc_{x};\comp;2),(\loc_{y};\comp;2)\right\}$.
-%   All uses up to the last are accepted by the monitor, but the last event does not satisfy the premise $2<2$ in \Cref{tr:sms-use}.
-%   Therefore, the whole program (\texttt{strncpy} linked with this kind of context) is not \gls*{sms}.
-% \end{exampleenv}
+\begin{exampleenv}[Normal invocation of \texttt{strncpy}]\label{ex:strncpy:sms}
+  Consider the insecure \texttt{strncpy} function from \Cref{ex:strncpy} with a context \texttt{strncpy(2, x, y)}, where \texttt{x} and \texttt{y} are pointers to valid regions of memory with allocated space for exactly two cells and do not contain the null-terminating character \texttt{'\textbackslash 0'}.
+  For the sake of this example, the pointers have been allocated by the component and passed to the context.
+  The loop of \texttt{strncpy} will copy exactly two cells and then check the loop condition for the last time.
+  At that stage, the induction variable \texttt{i} is equal to \texttt{2} and, unfortunately, the order of checks is such that first the cell \texttt{x[i]} is read prior to bounds checking \texttt{i < n}.
+  Because of this, there is an out-of-bounds memory access right before exiting the function.
+  This is also visible on the trace, which can be sketched as $\ldots\cdot\ev{Alloc}\ \loc_{x}\ 2;\comp;\unlock\cdot\ldots\cdot\ev{Alloc}\ \loc_{y}\ 2;\comp;\unlock\cdot\ldots\cdot\ev{Use}\ \loc_{x}\ 0;\comp;\unlock\cdot\ev{Use}\ \loc_{y}\ 0;\comp;\unlock\cdot\ev{Use}\ \loc_{x}\ 1;\comp;\unlock\cdot\ev{Use}\ \loc_{y}\ 1;\comp;\unlock\cdot\ev{Use}\ \loc_{x}\ 2;\comp;\unlock\cdot\ldots$, where $\loc_{x}$ and $\loc_{y}$ are the memory addresses associated to \texttt{x} and \texttt{y}, respectively.
+  Omitting the events for all ,,$\ldots$'' for sake of brevity, the abstract monitor trace of this is $\absev{Alloc}\ \loc_{x}\ \comp\ 2\cdot\absev{Alloc}\ \loc_{y}\ \comp\ 2\cdot\absev{Use}\ \loc_{x}\ \comp\ 0\cdot\absev{Use}\ \loc_{y}\ \comp\ 0\cdot\absev{Use}\ \loc_{x}\ \comp\ 1\cdot\absev{Use}\ \loc_{y}\ \comp\ 1\cdot\absev{Use}\ \loc_{x}\ \comp\ 2$.
+
+  After the allocation events, the state of the monitor is $\left\{(\loc_{x};\comp;2),(\loc_{y};\comp;2)\right\}$.
+  All uses up to the last are accepted by the monitor, but the last event does not satisfy the premise $2<2$ in \Cref{tr:sms-use}.
+  Therefore, the whole program (\texttt{strncpy} linked with this kind of context) is not \gls*{sms}.
+\end{exampleenv}
 
@@
 
-\subsubsection{Proving Robust Safety Property Preservation}
-
-We illustrate the proof of \Cref{thm:cca:rtp:tms} since the other secure compilation proofs of this paper follow the same approach.
-Unfolding the theorem statement yields the following assumptions:  
-	for any $\pi\in\lceil\tmssafe\rceil$\footnote{$\lceil\cdot\rceil$ lifts the property to a hyperproperty by applying the powerset operation~\cite{clarkson2008hyper}.}, 
-	$\src{\trace}$, 
-	$\src{\runtimetermvar}$, 
-	and component $\src{\library_{comp}}$, 
-	we have that $\rsat{\src{\library_{comp}}}{\pi}$ and 
-	$\progstepto{\prog[\trg]{\trg{\library_{\ctx}}}{\cca(\src{\library_{\comp}})}}{\trg{\runtimetermvar}}{\trg{\trace}}$, where $\trg{\library_{\ctx}}$ is arbitrary.
-The proof obligation is $\msfilterL{\trg{\trace}}\in\pi$, i.e., the specification trace associated to $\trg{\trace}$ satisfies the property $\pi$.
-A way to show this is to relate trace $\trg{\trace}$ to some $\Ltms$ trace $\src{\trace}$ (which already satisfies the property as per the assumptions).
-The assumptions already contain a target execution associated to this trace, so the task is to find an associated $\Ltms$ execution that yields $\trg{\trace}$.
-The trace $\trg{\trace}$ is split into different parts, as commonly done in secure compilation works~\cite{korashy2021capableptrs,abate2018goodcomps}, where each part contains the events that either the context or the component does, but not both.
-Because of this, all such trace segments are ,,well-bracketed'' in the sense that they start with either $\trg{\ev{Start}}$, $\trg{\ev{Call}\ \comptoctx\ foo\ \valueexpr}$, or $\trg{\ev{Ret}\ \comptoctx\ \valueexpr}$ and end with either $\trg{\ev{End}\ \valueexpr}$, $\trg{\ev{Ret}\ \ctxtocomp\ \valueexpr}$, or $\trg{\ev{Call}\ \ctxtocomp\ foo\ \valueexpr}$.
-In the following, the former is referred to as a context segment, since these executions happen in $\trg{\library_{\ctx}}$, and the latter is referred to as a component segment, since these executions happen in $\cca(\src{\library_{\comp}})$.
-\Cref{fig:proofdiag:rtp} visualises this division for a program execution with one call from context to component and how the target execution is related to a source execution.
-In the figure, the green dashed lines encompass the component segments while the orange boxes contain the actual context switches from context to component or vice versa.
-From a technical perspective, as typically done in compilation proofs, the proof requires some setup to maintain a relation between $\src{\Omega}$ and $\trg{\Omega}$ .
-Two cross-language relations make this precise: (i) $\multimap_{\delta}$ relates states that are involved in a context segment, allowing the target execution to perform internal calls, and (ii) $\approx_{\delta}$ relates states that are involved in a component segment, where both states need to agree exactly, i.e., the memory and the control flow states are required to contain the same information.
-The relations are indexed with $\delta$, which is an injective mapping from $\Ltms$ locations $\src{\loc}$ to $\Ltrg$ locations $\trg{\loc}$.
-Note that the relations $\multimap_{\delta}$ and $\approx_{\delta}$ swap when context switching.
-
-\begin{figure*}[!ht]
-  \begin{center}
-    \begin{tikzpicture}[state/.style={minimum height=1.5cm}]
-      % relative horizontal/vertical distance between states
-      \pgfmathsetmacro{\hdist}{0.95}
-      \pgfmathsetmacro{\vdist}{1.35}
-      \pgfmathsetmacro{\halfvdist}{0.725}
-
-      % row of src states
-      \node[state] (srcempty) {$\rtt{\src{\emptyset}}{\src{\lcall{main}{0}}}$};
-      \foreach \s [remember=\s as \cur (initially empty)] in {1,w_1,p,w_2,2} {
-        \node[state,right=\hdist of src\cur] (src\s) {$\rtt{\src{\Omega_{\s}}}{\src{\expr[_{\s}]}}$};
-      }
-      % row of trg states
-      \node[state,below=\vdist of srcempty] (trgempty) {$\rtt{\trg{\emptyset}}{\trg{\lcall{main}{0}}}$};
-      \foreach \s in {1,w_1,p,w_2,2} {
-        \node[state,below=\vdist of src\s] (trg\s) {$\rtt{\trg{\Omega_{\s}}}{\trg{\expr[_{\s}]}}$};
-      }
-      %% illustrations
-        % backtrans wrapper 1
-        \draw[ultra thick,loosely dotted,Peach!50,rounded corners] (src1.north east) -- (srcw\string_1.north east)
-          -- (trgw\string_1.south east) -- (trg1.south west) -- (src1.north west) -- cycle;
-        % backtrans wrapper 2
-        \draw[ultra thick,loosely dotted,Peach!50,rounded corners] (srcp.north east) -- (srcw\string_2.north east)
-          -- (trgw\string_2.south east) -- (trgp.south west) -- (srcp.north west) -- cycle;
-        % compiler correctness
-        \draw[ultra thick,loosely dashed,Emerald!50,rounded corners] ($(srcw\string_1.north east)+(0,0.05)$) -- ($(srcp.north east)+(0,0.05)$)
-          -- ($(trgw\string_2.north east)+(0.05,0)$) -- ($(trgw\string_2.south east)+(0.05,-0.05)$)
-          -- ($(trg1.south west)+(-0.05,-0.05)$) -- ($(trg1.north west)+(-0.05,0)$) -- ($(srcw\string_1.north west)+(0,0.05)$) -- cycle;
-      % state relations
-      \path (srcempty) edge[very thick,draw=gray!25] node[pos=0.5,sloped,rotate=180,fill=white] {\scriptsize$\multimap_\emptyset$} (trgempty)
-        (src1) edge[very thick,draw=gray!25] node[pos=0.5,sloped,rotate=180,fill=white] {\scriptsize$\multimap_{\delta_1}$} (trg1)
-        (srcp) edge[very thick,draw=gray!25] node[pos=0.5,sloped,fill=white] {\scriptsize$\approx_{\delta_p}$} (trgp)
-        (src2) edge[very thick,draw=gray!25] node[pos=0.5,sloped,rotate=180,fill=white] {\scriptsize$\multimap_{\delta_2}$} (trg2)
-        (srcw\string_1) edge[very thick,draw=gray!25] node[pos=0.5,sloped,fill=white] {\scriptsize$\approx_{\delta_{w_1}}$} (trgw\string_1)
-        (srcw\string_2) edge[very thick,draw=gray!25] node[pos=0.5,sloped,rotate=180,fill=white] {\scriptsize$\multimap_{\delta_{w_2}}$} (trgw\string_2)
-        % diagonals
-        (srcw\string_1) edge[thick,draw=gray!25] node[pos=0.16,sloped,rotate=180,fill=white] {\scriptsize$\approx_{\delta_{w_1}}$} (trg1)
-        (src1) edge[thick,draw=gray!25] node[pos=0.16,sloped,rotate=180,fill=white] {\scriptsize$\multimap_{\delta_1}$} (trgw\string_1)
-        (srcp) edge[thick,draw=gray!25] node[pos=0.2,sloped,fill=white] {\scriptsize$\approx_{\delta_p}$} (trgw\string_2)
-        (trgp) edge[thick,draw=gray!25] node[pos=0.2,sloped,fill=white] {\scriptsize$\multimap_{\delta_{w_2}}$} (srcw\string_2)
-        ;
-      %\drawpolygon src1,srcw\string_1,trgw\string_1,trg1;
-      %\drawpolygon srcp,srcw\string_2,trgw\string_2,trgp;
-      %\node[font=\tiny,align=center,above=0.2 of srcw1srcp] (wrapper) {Backtranslation\\Wrapper};
-      %\path[->,draw] (wrapper) -- (srcw\string_1);
-      %\path[->,draw] (wrapper) -- (srcp);
-      % steps
-      \path[very thick,color=\stlccol] (srcempty) edge[draw=none] node {\ $\xrightarrow{\trace[_1]}{}{\kern-3.5pt}_{\text{ctx}}^*$} (src1)
-        (src1) edge[draw=none] node {\ $\xrightarrow{\trace[_c]}{}{\kern-3.5pt}_{\text{ctx}}$} (srcw\string_1)
-        (srcw\string_1) edge[draw=none] node {\ $\xrightarrow{\trace[_p]}{}{\kern-3.5pt}_{\text{ctx}}^*$} (srcp)
-        (srcp) edge[draw=none] node[fill=white,inner sep=0,outer sep=0] {\ $\xrightarrow{\trace[_r]}{}{\kern-3.5pt}_{\text{ctx}}$} (srcw\string_2)
-        (srcw\string_2) edge[draw=none] node {\ $\xrightarrow{\trace[_2]}{}{\kern-3.5pt}_{\text{ctx}}^*$} (src2)
-        ;
-      \path[very thick,color=\ulccol] (trgempty) edge[draw=none] node[fill=white,inner sep=0,outer sep=0] {\ $\xrightarrow{\phantom{\trace[_1]}}{}{\kern-3.5pt}_{\text{ctx}}^*$} (trg1)
-        (trg1) edge[draw=none] node {\ $\xrightarrow{\phantom{\trace[_p]}}{}{\kern-3.5pt}_{\text{ctx}}^*$} (trgw\string_1)
-        (trgw\string_1) edge[draw=none] node[fill=white,inner sep=0,outer sep=0] {\ $\xrightarrow{\phantom{\trace[_p]}}{}{\kern-3.5pt}_{\text{ctx}}^*$} (trgp)
-        (trgp) edge[very thick,draw=none] node {\ $\xrightarrow{\phantom{\trace[_p]}}{}{\kern-3.5pt}_{\text{ctx}}^*$} (trgw\string_2)
-        (trgw\string_2) edge[draw=none] node {\ $\xrightarrow{\phantom{\trace[_2]}}{}{\kern-3.5pt}_{\text{ctx}}^*$} (trg2)
-        ;
-      % legend
-%     \node[align=left,below right=0.3 and 0.3 of trgempty,font=\tiny] (legend) {%
-%       ${\src{\trace[_c]}}\cong{\backt{\Ltms}{\Ltrg}(\trg{Call\ ?\ foo\ \valueexpr})}$\\%
-%       ${\src{\trace[_r]}}\cong{\backt{\Ltms}{\Ltrg}(\trg{Ret\ !\ foo\ \valueexpr})}$
-%       };
-      \node[below=0.5 of trgempty,draw=none] (legend) {};
-      \draw[ultra thick,loosely dotted,Peach!50,rounded corners] ($(legend.north east)+(0.5,-0.4)$) -- ($(legend.north east)+(1,-0.4)$);
-      \node at ($(legend.north east)+(1.5,-0.4)$) (legendwrapper) {};
-      \node[right of=legendwrapper] {\tiny Backtranslation Wrapper};
-      \draw[ultra thick,loosely dashed,Emerald!50,rounded corners] ($(legend.south east)+(0.5,0.4)$) -- ($(legend.south east)+(1,0.4)$);
-      \node at ($(legend.south east)+(1.5,0.4)$) (legendcorrectness) {};
-      \node[right of=legendcorrectness] {\tiny Compiler Correctness};
-    \end{tikzpicture}
-    \caption{Proof diagram for \Cref{thm:cca:rtp:tms} depicting the general structure of robust preservation proofs. %
-      Nodes in the graph represent runtime states. %
-      Vertical lines indicate cross language relations, while horizontal ones are execution steps. %
-      The green dashed trapezoid encompasses the component segment, while the orange dotted rectangles entail the context switches. %
-      $\Ltrg$ traces are omitted for readability. %
-      $\Ltms$ trace segments $\src{\trace[_{c}]}$ and $\src{\trace[_{r}]}$ describe the events that happen at the boundaries, i.e., during a context switch.
-      $\src{\trace_{p}}$ is the behavior of the component and the traces $\src{\trace[_{1}]}$ and $\src{\trace[_{2}]}$ describe the context.
-    }\label{fig:proofdiag:rtp}
-  \end{center}
-\end{figure*}
-
-So far, the paper explained how to relate an $\Ltms$ execution with a $\Ltrg$ execution.
-The next question is therefore how to build the corresponding $\Ltms$ execution.
-This is done using a standard secure compilation proof technique called trace-based backtranslation~\cite{abate2019jour,korashy2021capableptrs,patrignani2021rsc}, which can be used to build a context $\src{\library_{\ctx}}$ that behaves similar to $\trg{\library_{\ctx}}$.
-% Since $\Ltms$ is statically typed, the context $\src{\library_{\ctx}}$ obtained from the backtranslation needs to be well-typed when linked with $\src{\library_{\comp}}$.
-For context segments of the trace $\trg{\trace}$ it is also necessary to show that the execution behaves similarily, i.e., the context obtained from the backtranslation generates trace $\src{\trace}$.
-For component segments of the trace, the relatedness of states and traces follows from a compiler correctness argument.
-These two arguments yield the source execution $\progstepto{\src{\prog{\library_{\ctx}}{\library_{\comp}}}}{\src{\runtimetermvar}}{\src{\trace}}$.
-
-The proof now works as follows.
-Given that $\msfilterL{\trg{\trace}}=\msfilterLtms{\src{\trace}}$, the proof goal changes from $\msfilterL{\trg{\trace}}\in\pi$ to $\msfilterLtms{\src{\trace}}\in\pi$.
-This follows by specializing the robust satisfaction assumption $\rsat{\src{\library_{\comp}}}{\pi}$ to use the context $\src{\library_{\ctx}}$, which is obtained from the backtranslation, and to use the source execution $\progstepto{\src{\prog{\library_{\ctx}}{\library_{\comp}}}}{\src{\runtimetermvar}}{\src{\trace}}$.
+% \subsubsection{Proving Robust Safety Property Preservation}
+
+% We illustrate the proof of \Cref{thm:cca:rtp:tms} since the other secure compilation proofs of this paper follow the same approach.
+% Unfolding the theorem statement yields the following assumptions:  
+%   for any $\pi\in\lceil\tmssafe\rceil$\footnote{$\lceil\cdot\rceil$ lifts the property to a hyperproperty by applying the powerset operation~\cite{clarkson2008hyper}.}, 
+%   $\src{\trace}$, 
+%   $\src{\runtimetermvar}$, 
+%   and component $\src{\library_{comp}}$, 
+%   we have that $\rsat{\src{\library_{comp}}}{\pi}$ and 
+%   $\progstepto{\prog[\trg]{\trg{\library_{\ctx}}}{\cca(\src{\library_{\comp}})}}{\trg{\runtimetermvar}}{\trg{\trace}}$, where $\trg{\library_{\ctx}}$ is arbitrary.
+% The proof obligation is $\msfilterL{\trg{\trace}}\in\pi$, i.e., the specification trace associated to $\trg{\trace}$ satisfies the property $\pi$.
+% A way to show this is to relate trace $\trg{\trace}$ to some $\Ltms$ trace $\src{\trace}$ (which already satisfies the property as per the assumptions).
+% The assumptions already contain a target execution associated to this trace, so the task is to find an associated $\Ltms$ execution that yields $\trg{\trace}$.
+% The trace $\trg{\trace}$ is split into different parts, as commonly done in secure compilation works~\cite{korashy2021capableptrs,abate2018goodcomps}, where each part contains the events that either the context or the component does, but not both.
+% Because of this, all such trace segments are ,,well-bracketed'' in the sense that they start with either $\trg{\ev{Start}}$, $\trg{\ev{Call}\ \comptoctx\ foo\ \valueexpr}$, or $\trg{\ev{Ret}\ \comptoctx\ \valueexpr}$ and end with either $\trg{\ev{End}\ \valueexpr}$, $\trg{\ev{Ret}\ \ctxtocomp\ \valueexpr}$, or $\trg{\ev{Call}\ \ctxtocomp\ foo\ \valueexpr}$.
+% In the following, the former is referred to as a context segment, since these executions happen in $\trg{\library_{\ctx}}$, and the latter is referred to as a component segment, since these executions happen in $\cca(\src{\library_{\comp}})$.
+% \Cref{fig:proofdiag:rtp} visualises this division for a program execution with one call from context to component and how the target execution is related to a source execution.
+% In the figure, the green dashed lines encompass the component segments while the orange boxes contain the actual context switches from context to component or vice versa.
+% From a technical perspective, as typically done in compilation proofs, the proof requires some setup to maintain a relation between $\src{\Omega}$ and $\trg{\Omega}$ .
+% Two cross-language relations make this precise: (i) $\multimap_{\delta}$ relates states that are involved in a context segment, allowing the target execution to perform internal calls, and (ii) $\approx_{\delta}$ relates states that are involved in a component segment, where both states need to agree exactly, i.e., the memory and the control flow states are required to contain the same information.
+% The relations are indexed with $\delta$, which is an injective mapping from $\Ltms$ locations $\src{\loc}$ to $\Ltrg$ locations $\trg{\loc}$.
+% %Note that the relations $\multimap_{\delta}$ and $\approx_{\delta}$ swap when context switching.
+
+% So far, the paper explained how to relate an $\Ltms$ execution with a $\Ltrg$ execution.
+% The next question is therefore how to build the corresponding $\Ltms$ execution.
+% This is done using a standard secure compilation proof technique called trace-based backtranslation~\cite{abate2019jour,korashy2021capableptrs,patrignani2021rsc}, which can be used to build a context $\src{\library_{\ctx}}$ that behaves similar to $\trg{\library_{\ctx}}$.
+% % Since $\Ltms$ is statically typed, the context $\src{\library_{\ctx}}$ obtained from the backtranslation needs to be well-typed when linked with $\src{\library_{\comp}}$.
+% For context segments of the trace $\trg{\trace}$ it is also necessary to show that the execution behaves similarily, i.e., the context obtained from the backtranslation generates trace $\src{\trace}$.
+% For component segments of the trace, the relatedness of states and traces follows from a compiler correctness argument.
+% These two arguments yield the source execution $\progstepto{\src{\prog{\library_{\ctx}}{\library_{\comp}}}}{\src{\runtimetermvar}}{\src{\trace}}$.
+
+% The proof now works as follows.
+% Given that $\msfilterL{\trg{\trace}}=\msfilterLtms{\src{\trace}}$, the proof goal changes from $\msfilterL{\trg{\trace}}\in\pi$ to $\msfilterLtms{\src{\trace}}\in\pi$.
+% This follows by specializing the robust satisfaction assumption $\rsat{\src{\library_{\comp}}}{\pi}$ to use the context $\src{\library_{\ctx}}$, which is obtained from the backtranslation, and to use the source execution $\progstepto{\src{\prog{\library_{\ctx}}{\library_{\comp}}}}{\src{\runtimetermvar}}{\src{\trace}}$.
+
+% \begin{figure*}[!ht]
+%   \begin{center}
+%     \begin{tikzpicture}[state/.style={minimum height=1.5cm}]
+%       % relative horizontal/vertical distance between states
+%       \pgfmathsetmacro{\hdist}{0.95}
+%       \pgfmathsetmacro{\vdist}{1.35}
+%       \pgfmathsetmacro{\halfvdist}{0.725}
+
+%       % row of src states
+%       \node[state] (srcempty) {$\rtt{\src{\emptyset}}{\src{\lcall{main}{0}}}$};
+%       \foreach \s [remember=\s as \cur (initially empty)] in {1,w_1,p,w_2,2} {
+%         \node[state,right=\hdist of src\cur] (src\s) {$\rtt{\src{\Omega_{\s}}}{\src{\expr[_{\s}]}}$};
+%       }
+%       % row of trg states
+%       \node[state,below=\vdist of srcempty] (trgempty) {$\rtt{\trg{\emptyset}}{\trg{\lcall{main}{0}}}$};
+%       \foreach \s in {1,w_1,p,w_2,2} {
+%         \node[state,below=\vdist of src\s] (trg\s) {$\rtt{\trg{\Omega_{\s}}}{\trg{\expr[_{\s}]}}$};
+%       }
+%       %% illustrations
+%         % backtrans wrapper 1
+%         \draw[ultra thick,loosely dotted,Peach!50,rounded corners] (src1.north east) -- (srcw\string_1.north east)
+%           -- (trgw\string_1.south east) -- (trg1.south west) -- (src1.north west) -- cycle;
+%         % backtrans wrapper 2
+%         \draw[ultra thick,loosely dotted,Peach!50,rounded corners] (srcp.north east) -- (srcw\string_2.north east)
+%           -- (trgw\string_2.south east) -- (trgp.south west) -- (srcp.north west) -- cycle;
+%         % compiler correctness
+%         \draw[ultra thick,loosely dashed,Emerald!50,rounded corners] ($(srcw\string_1.north east)+(0,0.05)$) -- ($(srcp.north east)+(0,0.05)$)
+%           -- ($(trgw\string_2.north east)+(0.05,0)$) -- ($(trgw\string_2.south east)+(0.05,-0.05)$)
+%           -- ($(trg1.south west)+(-0.05,-0.05)$) -- ($(trg1.north west)+(-0.05,0)$) -- ($(srcw\string_1.north west)+(0,0.05)$) -- cycle;
+%       % state relations
+%       \path (srcempty) edge[very thick,draw=gray!25] node[pos=0.5,sloped,rotate=180,fill=white] {\scriptsize$\multimap_\emptyset$} (trgempty)
+%         (src1) edge[very thick,draw=gray!25] node[pos=0.5,sloped,rotate=180,fill=white] {\scriptsize$\multimap_{\delta_1}$} (trg1)
+%         (srcp) edge[very thick,draw=gray!25] node[pos=0.5,sloped,fill=white] {\scriptsize$\approx_{\delta_p}$} (trgp)
+%         (src2) edge[very thick,draw=gray!25] node[pos=0.5,sloped,rotate=180,fill=white] {\scriptsize$\multimap_{\delta_2}$} (trg2)
+%         (srcw\string_1) edge[very thick,draw=gray!25] node[pos=0.5,sloped,fill=white] {\scriptsize$\approx_{\delta_{w_1}}$} (trgw\string_1)
+%         (srcw\string_2) edge[very thick,draw=gray!25] node[pos=0.5,sloped,rotate=180,fill=white] {\scriptsize$\multimap_{\delta_{w_2}}$} (trgw\string_2)
+%         % diagonals
+%         (srcw\string_1) edge[thick,draw=gray!25] node[pos=0.16,sloped,rotate=180,fill=white] {\scriptsize$\approx_{\delta_{w_1}}$} (trg1)
+%         (src1) edge[thick,draw=gray!25] node[pos=0.16,sloped,rotate=180,fill=white] {\scriptsize$\multimap_{\delta_1}$} (trgw\string_1)
+%         (srcp) edge[thick,draw=gray!25] node[pos=0.2,sloped,fill=white] {\scriptsize$\approx_{\delta_p}$} (trgw\string_2)
+%         (trgp) edge[thick,draw=gray!25] node[pos=0.2,sloped,fill=white] {\scriptsize$\multimap_{\delta_{w_2}}$} (srcw\string_2)
+%         ;
+%       %\drawpolygon src1,srcw\string_1,trgw\string_1,trg1;
+%       %\drawpolygon srcp,srcw\string_2,trgw\string_2,trgp;
+%       %\node[font=\tiny,align=center,above=0.2 of srcw1srcp] (wrapper) {Backtranslation\\Wrapper};
+%       %\path[->,draw] (wrapper) -- (srcw\string_1);
+%       %\path[->,draw] (wrapper) -- (srcp);
+%       % steps
+%       \path[very thick,color=\stlccol] (srcempty) edge[draw=none] node {\ $\xrightarrow{\trace[_1]}{}{\kern-3.5pt}_{\text{ctx}}^*$} (src1)
+%         (src1) edge[draw=none] node {\ $\xrightarrow{\trace[_c]}{}{\kern-3.5pt}_{\text{ctx}}$} (srcw\string_1)
+%         (srcw\string_1) edge[draw=none] node {\ $\xrightarrow{\trace[_p]}{}{\kern-3.5pt}_{\text{ctx}}^*$} (srcp)
+%         (srcp) edge[draw=none] node[fill=white,inner sep=0,outer sep=0] {\ $\xrightarrow{\trace[_r]}{}{\kern-3.5pt}_{\text{ctx}}$} (srcw\string_2)
+%         (srcw\string_2) edge[draw=none] node {\ $\xrightarrow{\trace[_2]}{}{\kern-3.5pt}_{\text{ctx}}^*$} (src2)
+%         ;
+%       \path[very thick,color=\ulccol] (trgempty) edge[draw=none] node[fill=white,inner sep=0,outer sep=0] {\ $\xrightarrow{\phantom{\trace[_1]}}{}{\kern-3.5pt}_{\text{ctx}}^*$} (trg1)
+%         (trg1) edge[draw=none] node {\ $\xrightarrow{\phantom{\trace[_p]}}{}{\kern-3.5pt}_{\text{ctx}}^*$} (trgw\string_1)
+%         (trgw\string_1) edge[draw=none] node[fill=white,inner sep=0,outer sep=0] {\ $\xrightarrow{\phantom{\trace[_p]}}{}{\kern-3.5pt}_{\text{ctx}}^*$} (trgp)
+%         (trgp) edge[very thick,draw=none] node {\ $\xrightarrow{\phantom{\trace[_p]}}{}{\kern-3.5pt}_{\text{ctx}}^*$} (trgw\string_2)
+%         (trgw\string_2) edge[draw=none] node {\ $\xrightarrow{\phantom{\trace[_2]}}{}{\kern-3.5pt}_{\text{ctx}}^*$} (trg2)
+%         ;
+%       % legend
+% %     \node[align=left,below right=0.3 and 0.3 of trgempty,font=\tiny] (legend) {%
+% %       ${\src{\trace[_c]}}\cong{\backt{\Ltms}{\Ltrg}(\trg{Call\ ?\ foo\ \valueexpr})}$\\%
+% %       ${\src{\trace[_r]}}\cong{\backt{\Ltms}{\Ltrg}(\trg{Ret\ !\ foo\ \valueexpr})}$
+% %       };
+%       \node[below=0.5 of trgempty,draw=none] (legend) {};
+%       \draw[ultra thick,loosely dotted,Peach!50,rounded corners] ($(legend.north east)+(0.5,-0.4)$) -- ($(legend.north east)+(1,-0.4)$);
+%       \node at ($(legend.north east)+(1.5,-0.4)$) (legendwrapper) {};
+%       \node[right of=legendwrapper] {\tiny Backtranslation Wrapper};
+%       \draw[ultra thick,loosely dashed,Emerald!50,rounded corners] ($(legend.south east)+(0.5,0.4)$) -- ($(legend.south east)+(1,0.4)$);
+%       \node at ($(legend.south east)+(1.5,0.4)$) (legendcorrectness) {};
+%       \node[right of=legendcorrectness] {\tiny Compiler Correctness};
+%     \end{tikzpicture}
+%     \caption{Proof diagram for \Cref{thm:cca:rtp:tms} depicting the general structure of robust preservation proofs. %
+%       Nodes in the graph represent runtime states. %
+%       Vertical lines indicate cross language relations, while horizontal ones are execution steps. %
+%       The green dashed trapezoid encompasses the component segment, while the orange dotted rectangles entail the context switches. %
+%       $\Ltrg$ traces are omitted for readability. %
+%       $\Ltms$ trace segments $\src{\trace[_{c}]}$ and $\src{\trace[_{r}]}$ describe the events that happen at the boundaries, i.e., during a context switch.
+%       $\src{\trace_{p}}$ is the behavior of the component and the traces $\src{\trace[_{1}]}$ and $\src{\trace[_{2}]}$ describe the context.
+%     }\label{fig:proofdiag:rtp}
+%   \end{center}
+% \end{figure*}
 
@@
     \end{lstlisting}
-    When calling this with \texttt{strncpy(2, x, y)}, where \texttt{x} and \texttt{y} are pointers to valid regions of memory with allocated space for exactly two cells and do not contain the null-terminating character \texttt{'\textbackslash 0'}, the event $\ev{Use}\ \loc_{x}\ 2;\comp;\unlock$ would not be emitted during execution, since the bounds check prevents the condition \texttt{src[i] != '\textbackslash 0'} from executing.
+    When calling this with a context like in \Cref{ex:strncpy:sms}, the event $\ev{Use}\ \loc_{x}\ 2;\comp;\unlock$ would not be emitted during execution, since the bounds check prevents the condition \texttt{src[i] != '\textbackslash 0'} from executing.
 \end{exampleenv}

### 2023-09-25 csf/main.tex
comments up to end of 5

@@
 
+\hypersetup{
+    colorlinks,
+    linkcolor={red!50!black},
+    citecolor={blue!50!black},
+    urlcolor={blue!80!black}
+}
+
 \setul{0.95ex}{0.3ex}
@@
 
-\title{Secure Composition of Robust and Optimising Compilers}
-
-\author{\IEEEauthorblockN{1\textsuperscript{st} Matthis Kruse\orcidlink{0000-0003-4062-9666}}
-\IEEEauthorblockA{\textit{CISPA and Saarland University} \\
-Saarbr\"ucken, Germany \\
-matthis.kruse@cispa.de}
-\and
-\IEEEauthorblockN{2\textsuperscript{nd} Michael Backes}
-\IEEEauthorblockA{\textit{CISPA} \\
-Saarbr\"ucken, Germany \\
-director@cispa.de}
-\and
-\IEEEauthorblockN{3\textsuperscript{rd} Marco Patrignani\orcidlink{0000-0003-3411-9678}}
-\IEEEauthorblockA{\textit{Trento University} \\
-Trento, Italy \\
-marco.patrignani@unitn.it}%
+\title{Secure Composition \\ of Robust and Optimising Compilers}
+
+% \author{
+% \IEEEauthorblockN{1\textsuperscript{st} Matthis Kruse\orcidlink{0000-0003-4062-9666}}
+% \IEEEauthorblockA{\textit{CISPA and Saarland University} \\
+% Saarbr\"ucken, Germany \\
+% matthis.kruse@cispa.de}
+% \and
+% \IEEEauthorblockN{2\textsuperscript{nd} Michael Backes}
+% \IEEEauthorblockA{\textit{CISPA} \\
+% Saarbr\"ucken, Germany \\
+% director@cispa.de}
+% \and
+% \IEEEauthorblockN{3\textsuperscript{rd} Marco Patrignani\orcidlink{0000-0003-3411-9678}}
+% \IEEEauthorblockA{\textit{Trento University} \\
+% Trento, Italy \\
+% marco.patrignani@unitn.it}%
+% }
+\author{
+
 }
@@
 \begin{IEEEkeywords}
-  memory-safety, secure compilation, privacy
+  memory-safety, cryptographic constant-time, secure compilation, robustness
 \end{IEEEkeywords}
@@
 % Even under these assumptions, there is a 
-This function is subject to a subtle \gls*{sms} vulnerability: the bounds check \texttt{i < n} should happen {\it before} the access to memory location \texttt{x[i]}: otherwise
+This function is subject to a subtle \gls*{sms} vulnerability: the bounds check \texttt{i < n} should happen {\it before} the access to memory location \texttt{src[i]}: otherwise
 % If \texttt{n} is at least as large as the length of the array \texttt{src}, 
@@
 
-Because of timing attacks, fixing \gls*{sms} is not enough to make \texttt{strncpy} secure.
-In fact, the loop can terminate early, as soon as the string-terminating character \texttt{'\textbackslash 0'} is encountered, thus making program execution time proportional to the length of the array pointed by \texttt{src}.
+Because of timing attacks, addressing \gls*{sms} is not enough to make \texttt{strncpy} secure.
+In fact, the loop can terminate early, as soon as the string-terminating character \texttt{'\textbackslash 0'} is encountered, thus making program execution time proportional to the length of the array pointed by \texttt{src}, and violating the \gls*{cct} property.
 % Often, programmers rewrite programs like these to have loops that run with a fixed number of iterations.
@@
         %\Cref{sec:compcomp} studies different forms of secure compiler compositions.
-        This paper takes the secure compilation framework of~\cite{abate2019jour} and extends it to reason about the security of all different known forms of compiler composition (\Cref{sec:compcomp}).
+        This paper takes the secure compilation framework of Abate et al.~\cite{abate2019jour} and extends it to reason about the security of all different known forms of compiler composition (\Cref{sec:compcomp}).
         %Concretely, it investigates upper and lower compositions, as well as the sequential composition.
@@
         This is crucial for reordering optimisation passes and thus generating secure and efficient code.
+      	\MP{
+      		doublecheck.
+      		also, the corollary .. ?
+      	}
 
@@
 
-\begin{exampleenv}[Double Free in Bluetooth Subsystem]
-  Consider \texttt{CVE-2021-3564}~\cite{doublefree-bluetooth}, one of many submissions for a double-free vulnerability.
-  The vulnerability arises due to a race condition where the context-level function \texttt{hci\_cmd\_work} was not expected to behave maliciously, since it resides in the same source-code repository where the vulnerability occurs.
-  Nevertheless, the component-level code of \texttt{hci\_dev\_do\_open} is linked with \texttt{hci\_cmd\_work} and does not atomically check whether a pointer has been freed already:
-  Therefore, \texttt{hci\_dev\_do\_open} does not satisfy the no-double-frees property robustly, since there is an implementation for \texttt{hci\_cmd\_work} that leads to a violation of that property when linked with \texttt{hci\_dev\_do\_open}.
-\end{exampleenv}
+% \begin{exampleenv}[Double Free in Bluetooth Subsystem]
+%   Consider \texttt{CVE-2021-3564}~\cite{doublefree-bluetooth}, one of many submissions for a double-free vulnerability.
+%   The vulnerability arises due to a race condition where the context-level function \texttt{hci\_cmd\_work} was not expected to behave maliciously, since it resides in the same source-code repository where the vulnerability occurs.
+%   Nevertheless, the component-level code of \texttt{hci\_dev\_do\_open} is linked with \texttt{hci\_cmd\_work} and does not atomically check whether a pointer has been freed already:
+%   Therefore, \texttt{hci\_dev\_do\_open} does not satisfy the no-double-frees property robustly, since there is an implementation for \texttt{hci\_cmd\_work} that leads to a violation of that property when linked with \texttt{hci\_dev\_do\_open}.
+% \end{exampleenv}
 
@@
 Because of this, this paper writes $\rtp{\cc{\src{L}}{\trg{L}}}{\pi}$, even though $\pi$ is a property and not a class.
+\MP{
+	check that we say classes are hp
+}
 
@@
 \vspace{-2em}
-\begin{gather*}
+\begin{gather*}\small
   \begin{aligned}
-  \mi{(Security\ Tag)}~\sigma\bnfdef&\ \lock \mid \unlock\hspace{0.5cm}
+  \mi{(Security\ Tag)}~\sigma\bnfdef&\ \lock \mid \unlock
+  \qquad
   \mi{(Control\ Tag)}~\sandboxtag\bnfdef \ctx\mid\comp\\
   \mi{(Event)}~\event\bnfdef&\ \emptyevent \mid \lightning \mid \event[_{b}];\sandboxtag;\sigma \\
-  \mi{(Pre\text{-}event)}~\event[_{b}]\bnfdef&\ \ev{Alloc\ \loc\ n} \mid \ev{Dealloc\ \loc} \\
-                                             & \mid \ev{Use\ \loc\ n} \mid \ev{Branch\ n} \mid \ev{Binop\ n} \\
+  \mi{(Pre\text{-}event)}~\event[_{b}]\bnfdef&\ \ev{Alloc\ \loc\ n} \mid \ev{Dealloc\ \loc} 
+  \\
+                                             & 
+                                             \mid \ev{Use\ \loc\ n} \mid \ev{Branch\ n} \mid \ev{Binop\ n} \\
   \end{aligned}
 \end{gather*}
-
 The specification trace model defines events as either the empty event ($\emptyevent$), a crash ($\lightning$), or as tuples consisting of a pre-event, a control-tag, and a security-tag.
-The purpose of the model is to define key security properties of interest, such as \gls*{ms} or a stricter variant of cryptographic constant time.
-To this end, security-tags indicate whether an event contains sensitive information ($\lock$) or not ($\unlock$), while control-tags state whether the context ($\ctx$) or the component ($\comp$) are responsible for emitting the event.
+Since the purpose of the model is to define key security properties of interest, such as \gls*{ms} or a stricter variant of cryptographic constant time, security-tags indicate whether an event contains sensitive information ($\lock$) or not ($\unlock$), while control-tags state whether the context ($\ctx$) or the component ($\comp$) are responsible for emitting the event.
 The latter is necessary to be able to ignore actions done by a spurious context that, e.g., immediately deallocates a memory location twice, thus violating \gls*{tms}~\cite{nagarakatte2010cets}.
@@
 Moreover, an instruction whose timing is data-dependent must not have a secret as an operand.
-Typical operations with data-dependent timing are branches and certain binary operations, such as division\footnote{This is highly architecture-dependent, but division is an operation that serves as a classic example for a data-dependent timing instruction, e.g., \cite[p.~755]{arm-refman}.}.
+Typical operations with data-dependent timing are branches and certain binary operations, such as division.%
+\footnote{
+	This is architecture-dependent, but division is an operation that serves as a classic example for a data-dependent timing instruction~\cite[p.~755]{arm-refman}.
+}
 Both operations are also modelled in the specification trace model ($\ev{Branch\ n}$ and $\ev{Binop\ n}$).
@@
 
-\gls*{sms} prohibits out-of-bounds accesses:
+\gls*{sms} prohibits out-of-bounds accesses.
 
 \begin{definition}[\glsfirst*{sms}]\label{def:trace:smsdef}
+\small
   $$
@@
 %%% our definition does not guarantee that (on purpose, we want to compose it with cct)
-Note, however, that this definition says nothing about memory-safety issues introduced by side-channels, such as speculation.
+Note that this definition says nothing about memory-safety issues introduced by side-channels.
 
@@
 
-While \gls*{scct} may seem overly strict, since it may seem that secrets must not occur on a trace (since $\securitytag$ is forced to be $\unlock$). 
-However, programs that operate on secrets and have a ''data independent timing mode'' enabled do not leak these secrets, despite performing computation on them (see \Cref{ex:ditm}).
+\gls*{scct} may seem overly strict, since it may seem that secrets must not occur on a trace (since $\securitytag$ is forced to be $\unlock$). 
+However, programs that operate on secrets and have a ``data independent timing mode'' enabled do not leak these secrets, despite performing computation on them (see \Cref{ex:ditm}).
 
@@
 \begin{definition}[\glsfirst*{ms} and \gls*{scct}]\label{def:trace:msscctdef}
-  $
+  $$
   \msscctsafe:=\mssafe\cap\scctsafe
-  $
+  $$
 \end{definition}
@@
 Monitors enforce safety properties by accepting or rejecting traces, i.e., if they reject a trace, the trace does not satisfy the property they check.
-Since reasoning on monitors is easier than directly on just traces, this section presents a monitor for each of the previously shown safety properties (\Cref{subsec:propdefs}).
+Since reasoning on monitors is easier than on just traces, this section presents a monitor for each of the previously shown safety properties (\Cref{subsec:propdefs}).
 %\MPpost{
@@
 The trace agreement is entirely straightforward, so only allocation, branch, and crash are shown.
+\MP{
+	we should say something .. ?
+}
 
@@
   char* greet = new char[12] { "Hello CSF!!" };
-  // greet address ${\color{gray}\loc_{x}}$
   char* to = new char[12];
-  // to address ${\color{gray}\loc_{y}}$
   strncpy(11, to, greetings);
@@
   \end{lstlisting}
-  Up to the body of \texttt{printf}, the program execution yields the specification trace $\ev{Alloc}\ \loc[_{x}]\ 12;\comp;\unlock\cdot\ev{Alloc}\ \loc[_{y}]\ 12;\comp;\unlock\cdot\ev{Use}\ \loc_{x}\ 0;\comp;\unlock\cdot\ev{Use}\ \loc_{x}\ 0;\comp;\unlock\cdot\ev{Use}\ \loc_{y}\ 0;\comp;\unlock\cdot\ev{Use}\ \loc_{x}\ 1;\comp;\unlock\cdot\ldots\cdot\ev{Use}\ \loc_{x}\ 12;\comp;\unlock\cdot\ev{Dealloc}\ \loc[_{y}];\comp;\unlock\cdot\ev{Use}\ \loc_{y}\ 6;\comp;\unlock$.
-  Relating this trace to abstract monitor events yields $\abstrace=\absev{Alloc}\ \loc[_{x}]\ \comp\cdot\absev{Alloc}\ \loc[_{y}]\ \comp\cdot\absev{Use}\ \loc[_{x}]\ \comp\cdot\absev{Use}\ \loc[_{x}]\ \comp\cdot\absev{Use}\ \loc[_{y}]\ \comp\cdot\ldots\cdot\absev{Use}\ \loc[_{x}]\ \comp\cdot\absev{Dealloc}\ \loc[_{y}]\ \comp\cdot\absev{Use}\ \loc[_{y}]\ \comp$.
+    % // greet address ${\color{gray}\loc_{x}}$
+  % // to address ${\color{gray}\loc_{y}}$
+% 
+  Up to the body of \texttt{printf}, the program execution yields the specification trace 
+  	% $\ev{Alloc}\ \loc[_{x}]\ 12;\comp;\unlock\cdot\ev{Alloc}\ \loc[_{y}]\ 12;\comp;\unlock\cdot\ev{Use}\ \loc_{x}\ 0;\comp;\unlock\cdot\ev{Use}\ \loc_{x}\ 0;\comp;\unlock\cdot\ev{Use}\ \loc_{y}\ 0;\comp;\unlock\cdot\ev{Use}\ \loc_{x}\ 1;\comp;\unlock\cdot\ldots\cdot\ev{Use}\ \loc_{x}\ 12;\comp;\unlock\cdot\ev{Dealloc}\ \loc[_{y}];\comp;\unlock\cdot\ev{Use}\ \loc_{y}\ 6;\comp;\unlock$
+
+  	\vspace{-1em}
+  	{\small
+  	\begin{align*}
+  		&
+  		\ev{Alloc}\ \loc[_{x}]\ 12;\comp;\unlock
+  		\cdot
+  		\ev{Alloc}\ \loc[_{y}]\ 12;\comp;\unlock
+  		\cdot
+  		\ev{Use}\ \loc_{x}\ 0;\comp;\unlock
+  		\\
+  		&
+  		\cdot
+  		\ev{Use}\ \loc_{x}\ 0;\comp;\unlock
+  		\cdot
+  		\ev{Use}\ \loc_{y}\ 0;\comp;\unlock
+  		\cdot
+  		\ev{Use}\ \loc_{x}\ 1;\comp;\unlock
+  		\cdot
+  		\ldots
+  		\\
+  		&
+  		\cdot
+  		\ev{Use}\ \loc_{x}\ 12;\comp;\unlock
+  		\cdot
+  		\ev{Dealloc}\ \loc[_{y}];\comp;\unlock
+  		\cdot
+  		\ev{Use}\ \loc_{y}\ 6;\comp;\unlock
+  	\end{align*}
+  	}
+Relating this trace to abstract monitor events yields trace
+
+  \vspace{-1em}
+  {\small
+  \begin{align*}
+   	% \abstrace=
+   	&
+   	\absev{Alloc}\ \loc[_{x}]\ \comp
+   	\cdot
+   	\absev{Alloc}\ \loc[_{y}]\ \comp
+   	\cdot
+   	\absev{Use}\ \loc[_{x}]\ \comp
+   	\\
+   	&
+   	\cdot
+   	\absev{Use}\ \loc[_{x}]\ \comp
+   	\cdot
+   	\absev{Use}\ \loc[_{y}]\ \comp
+   	\cdot
+   	\ldots
+   	\\
+   	&
+   	\cdot
+   	\absev{Use}\ \loc[_{x}]\ \comp
+   	\cdot
+   	\absev{Dealloc}\ \loc[_{y}]\ \comp
+   	\cdot
+   	\absev{Use}\ \loc[_{y}]\ \comp
+   \end{align*} 
+  	}
   Remembering the definition of \texttt{strncpy} (\Cref{ex:strncpy}), observe that it does not deallocate its arguments.
-  Even though the trace contains an out-of-bounds access right before returning from \texttt{strncpy}, this is no concern for \gls*{tms}, since the location $\loc[_{x}]$ is still allocated.
-  However, having returned from \texttt{strncpy}, the \texttt{greet} function continues and deallocates $\loc[_{y}]$ whose subsequent use in the \texttt{printf} call is a use-after-free bug.
-
-  The fix would be to \lstinline[language=c++,basicstyle=\small\ttfamily]|delete greetings| instead of \texttt{to} and add a \lstinline[language=c++,basicstyle=\small\ttfamily]|delete to| after the \texttt{printf} call, which leads to the abstract monitor trace $\abstrace[']=\absev{Alloc}\ \loc[_{x}]\ \comp\cdot\absev{Alloc}\ \loc[_{y}]\ \comp\cdot\absev{Use}\ \loc[_{x}]\ \comp\cdot\absev{Use}\ \loc[_{x}]\ \comp\cdot\absev{Use}\ \loc[_{y}]\ \comp\cdot\ldots\cdot\absev{Use}\ \loc[_{x}]\ \comp\cdot\absev{Dealloc}\ \loc[_{x}]\ \comp\cdot\absev{Use}\ \loc[_{y}]\ \comp$.
+  Even though the trace contains an out-of-bounds access right before returning from \texttt{strncpy}, this is no concern for \gls*{tms}, since location $\loc[_{x}]$ is still allocated.
+  However, having returned from \texttt{strncpy}, the \texttt{greet} function continues and deallocates $\loc[_{y}]$ whose subsequent use in \texttt{printf} is a use-after-free bug.
+
+  The fix would be to \lstinline[language=c++,basicstyle=\small\ttfamily]|delete greetings| instead of \texttt{to} and add a \lstinline[language=c++,basicstyle=\small\ttfamily]|delete to| after the \texttt{printf} call, which leads to the abstract monitor trace
+
+\vspace{-1em}
+  {\small
+  \begin{align*}
+  	\abstrace[']=
+  	&
+  	\absev{Alloc}\ \loc[_{x}]\ \comp
+  	\cdot
+  	\absev{Alloc}\ \loc[_{y}]\ \comp
+  	\cdot
+  	\absev{Use}\ \loc[_{x}]\ \comp
+  	\\
+  	&
+  	\cdot
+  	\absev{Use}\ \loc[_{x}]\ \comp
+  	\cdot
+  	\absev{Use}\ \loc[_{y}]\ \comp
+  	\cdot
+  	\ldots
+  	\cdot
+  	\\
+  	&
+  	\absev{Use}\ \loc[_{x}]\ \comp
+  	\cdot
+  	\absev{Dealloc}\ \loc[_{x}]\ \comp
+  	\cdot
+  	\absev{Use}\ \loc[_{y}]\ \comp	
+  \end{align*}
+  }
   It follows that $\montmssafe{\abstrace[']}$ and from \Thmref{lem:mon:tmssafe}, it follows that the program satisfies \Thmref{def:trace:tmsdef}, even though the program still violates \gls*{sms}.
@@
   The loop of \texttt{strncpy} will copy exactly two cells and then check the loop condition for the last time.
-  At that stage, the induction variable \texttt{i} is equal to \texttt{2} and, unfortunately, the order of checks is such that first the cell \texttt{x[i]} is read prior to bounds checking \texttt{i < n}.
-  Because of this, there is an out-of-bounds memory access right before exiting the function.
-  This is also visible on the trace, which can be sketched as $\ldots\cdot\ev{Alloc}\ \loc_{x}\ 2;\comp;\unlock\cdot\ldots\cdot\ev{Alloc}\ \loc_{y}\ 2;\comp;\unlock\cdot\ldots\cdot\ev{Use}\ \loc_{x}\ 0;\comp;\unlock\cdot\ev{Use}\ \loc_{y}\ 0;\comp;\unlock\cdot\ev{Use}\ \loc_{x}\ 1;\comp;\unlock\cdot\ev{Use}\ \loc_{y}\ 1;\comp;\unlock\cdot\ev{Use}\ \loc_{x}\ 2;\comp;\unlock\cdot\ldots$, where $\loc_{x}$ and $\loc_{y}$ are the memory addresses associated to \texttt{x} and \texttt{y}, respectively.
-  Omitting the events for all ,,$\ldots$'' for sake of brevity, the abstract monitor trace of this is $\absev{Alloc}\ \loc_{x}\ \comp\ 2\cdot\absev{Alloc}\ \loc_{y}\ \comp\ 2\cdot\absev{Use}\ \loc_{x}\ \comp\ 0\cdot\absev{Use}\ \loc_{y}\ \comp\ 0\cdot\absev{Use}\ \loc_{x}\ \comp\ 1\cdot\absev{Use}\ \loc_{y}\ \comp\ 1\cdot\absev{Use}\ \loc_{x}\ \comp\ 2$.
+  At that stage, variable \texttt{i} is equal to \texttt{2} and, unfortunately, the order of checks is such that first the cell \texttt{x[i]} is read prior to bounds checking \texttt{i < n}, so there is an out-of-bounds memory access right before exiting the function.
+  This is also visible on the trace, which can be sketched as
+
+  	{\small
+  	\begin{align*}
+  	&
+  		\ldots
+  		\cdot
+  		\ev{Alloc}\ \loc_{x}\ 2;\comp;\unlock
+  		\cdot
+  		\ldots
+  		\cdot
+  		\ev{Alloc}\ \loc_{y}\ 2;\comp;\unlock
+  		\cdot
+  		\ldots
+  		\cdot
+  	\\
+  	&
+  		\ev{Use}\ \loc_{x}\ 0;\comp;\unlock
+  		\cdot
+  		\ev{Use}\ \loc_{y}\ 0;\comp;\unlock
+  		\cdot
+  		\ev{Use}\ \loc_{x}\ 1;\comp;\unlock
+  	\\
+  	&
+  		\cdot
+  		\ev{Use}\ \loc_{y}\ 1;\comp;\unlock
+  		\cdot
+  		\ev{Use}\ \loc_{x}\ 2;\comp;\unlock
+  		\cdot
+  		\ldots	
+  	\end{align*}
+  	}
+
+\noindent
+where $\loc_{x}$ and $\loc_{y}$ are the memory addresses associated to \texttt{x} and \texttt{y}, respectively.
+  Omitting the events for all ,,$\ldots$'' for sake of brevity, the abstract monitor trace of this is
+
+  	{\small
+  	\begin{align*}
+  		&
+  		\absev{Alloc}\ \loc_{x}\ \comp\ 2
+  		\cdot
+  		\absev{Alloc}\ \loc_{y}\ \comp\ 2
+  		\cdot
+  		\absev{Use}\ \loc_{x}\ \comp\ 0
+  		\\
+  		&
+  		\cdot
+  		\absev{Use}\ \loc_{y}\ \comp\ 0
+  		\cdot
+  		\absev{Use}\ \loc_{x}\ \comp\ 1
+  		\cdot
+  		\absev{Use}\ \loc_{y}\ \comp\ 1
+  		\cdot
+  		\absev{Use}\ \loc_{x}\ \comp\ 2
+  	\end{align*}
+  	}
 
@@
 
-\subsubsection{Combining \glsfirst*{tms} and \glsfirst*{sms} Monitors to obtain \glsfirst*{ms}}\label{subsubsec:ms}
+\subsubsection{Combining \glsfirst*{tms} and \glsfirst*{sms} Monitors into \glsfirst*{ms}}\label{subsubsec:ms}
 
@@
 \begin{gather*}
+% \small
   \begin{aligned}
-    \mi{(Abstract\ Store)}~\scctmonitor:=&\ \emptyset \\
-    \mi{(Abstract\ Events)}~\absevent:=&\ \absev{\varepsilon} \mid \absterm \mid \absev{Leak} \\
+    \mi{(Abstract\ Store)}~\scctmonitor:=&\ \emptyset 
+    % &
+    \\
+    \mi{(Abstract\ Events)}~\absevent:=&\ \absev{\varepsilon} \mid \absterm \mid \absev{Leak} 
+    % \\
   \end{aligned}
@@
   Consider the call \texttt{strncpy(1,x,y)} to the \texttt{strncpy} function (\Cref{ex:strncpy}) with low ($\unlock$) security for \texttt{x} and high security ($\lock$) for \texttt{y}.
-  The trace of just the copying part inside of \texttt{strncpy} looks like $\ev{Use}\ \loc_{x}\ 0;\comp;\unlock\cdot\ev{Use}\ \loc_{y}\ 0;\comp;\lock$.
+  The trace for the copying part of \texttt{strncpy} is:
+
+\vspace{-1em}
+  {\small
+  \begin{align*}
+  	\ev{Use}\ \loc_{x}\ 0;\comp;\unlock\cdot\ev{Use}\ \loc_{y}\ 0;\comp;\lock	
+  \end{align*}
+  }
   In terms of the abstract monitor trace, this is just $\absev{Leak}$.
@@
 Note that the presented framework works with arbitrary classes.
+\MP{check, we ened more than this}
 % and builds on prior work~\cite{kruse2022csc}\MK{mention this at all?
@@
 This is not intuitive in the sense that in the security domain, composition does not work without additional generalizations~\cite{canetti2006univcomp,mccullough2012compo,fabian2022automatic}.
+\MP{what deos this mean?}
 The \bul{sequential composition of compilers $\cc{\src{L}}{\trg{L}}$ and $\cc{\trg{L}}{\obj{L}}$} is defined as follows:
@@
 
+\MP{move this}
 Note that the theorem may not be applicable in certain cases.
@@
 So, given two compilation passes \bul{$\cc[_{1}]{\trg{L}}{\trg{L}}$, $\cc[_{2}]{\trg{L}}{\trg{L}}$, both robustly preserving class $\class[_{1}]$ or $\class[_{2}]$, respectively}, \oul{for any order of their composition the composed compiler robustly preserves the intersection of $\class[_{1}]$ and $\class[_{2}]$}.
+% 
+\MP{
+	later on, say that optimisation passes remove elements from traces, so they should be fine with respect to the result with $\tau$.
+	also, yes, if $\tau$ adds we have a problem, if it removes we don't
+}
 
@@
 It is worth emphasizing that the framework is general: It supports both ordinary trace properties as well as hyperproperties or even relational hyperproperties~\cite{abate2019jour}.
+\MP{pull this out to earlier}
 Moreover, while this paper assumes a unified trace-model, it is merely a technical exercise to allow for different trace-models between programming languages~\cite{abate2021extacc}.

### 2023-09-25 csf/main.tex
pass

@@
 
-\begin{IEEEkeywords}
-  memory-safety, cryptographic constant-time, secure compilation, robustness
-\end{IEEEkeywords}
+% \begin{IEEEkeywords}
+%   memory-safety, cryptographic constant-time, secure compilation, robustness
+% \end{IEEEkeywords}
 
@@
 Note that a class of properties $\class$ can represent just one property $\pi$ by lifting~\cite{clarkson2008hyper} that property to sets of properties, i.e., use the powerset of $\pi$ instead of $\pi$ itself.
-Because of this, this paper writes $\rtp{\cc{\src{L}}{\trg{L}}}{\pi}$, even though $\pi$ is a property and not a class.
-\MP{
-	check that we say classes are hp
-}
+Thus we write $\rtp{\cc{\src{L}}{\trg{L}}}{\pi}$, even though $\pi$ is a property and not a class.
+% \MP{
+% 	check that we say classes are hp
+% }
 
@@
   \begin{aligned}
-    \mi{(Abstract\ Store)}~\tmsmonitor\bnfdef&\ \left\{\text{A}:L\times\sandboxtag,\text{F}:L\times\sandboxtag\right\} \\%
-    \emptyset:=&\ \left\{\text{A}:\emptyset,\text{F}:\emptyset\right\}\\
+    \mi{(Abstract\ Store)}~\tmsmonitor\bnfdef&\ \left\{\text{A}:L\times\sandboxtag,\text{F}:L\times\sandboxtag\right\} 
+    \\%
+    % \emptyset:=&\ \left\{\text{A}:\emptyset,\text{F}:\emptyset\right\}
+    % \\
     \mi{(Abstract\ Events)}~\absevent\bnfdef&\ \absev{\emptyevent} \mid \absev{Alloc\ }\loc\ \sandboxtag \mid \absev{Dealloc\ }\loc\ \sandboxtag \\
@@
 \end{gather*}
-\begin{center}
-  \judgbox{\monitorcheck[]{\tmsmonitor}{\tmsmonitor[']}{\absevent}}{,,Monitor $\tmsmonitor$ does one step to $\tmsmonitor[']$\\given event $\absevent$.''}$\;$\\
+For \gls*{tms}, the state of the monitor is a record with two sets keeping track of allocated and deallocated locations, and the initial state $\emptyset$ is defined as $\left\{\text{A}:\emptyset,\text{F}:\emptyset\right\}$.
+% 
+Monitor events include {\em allocation}, {\em deallocation}, and {\em use} of a location $\loc$ as well as abnormal program termination.
+
+\begin{center}\small
+  \judgbox{\monitorcheck[]{\tmsmonitor}{\tmsmonitor[']}{\absevent}}{,,Monitor $\tmsmonitor$ steps to $\tmsmonitor[']$\\given event $\absevent$.''}$\;$\\
   %
@@
   \typerule{tms-use}{
-    (\loc;\sandboxtag)\in\tmsmonitor.\text{A} &
+    (\loc;\sandboxtag)\in\tmsmonitor.\text{A} 
+    &
+    % \\
     (\loc;\sandboxtag)\notin\tmsmonitor.\text{F}
@@
   \typerule{tms-dealloc}{
-    (\loc;\sandboxtag)\in\tmsmonitor.\text{A} &
-    (\loc;\sandboxtag)\notin\tmsmonitor.\text{F} \\
-    \tmsmonitor[']=\left\{\text{A}: \tmsmonitor.\text{A}\setminus\left\{(\loc;\sandboxtag)\right\}\right.,\\\left.\text{F}: \tmsmonitor.\text{F}\cup\left\{(\loc;\sandboxtag)\right\}\right\}
+    (\loc;\sandboxtag)\in\tmsmonitor.\text{A} 
+    &
+    (\loc;\sandboxtag)\notin\tmsmonitor.\text{F} 
+    \\
+    \tmsmonitor[']=\left\{\text{A}: \tmsmonitor.\text{A}\setminus\left\{(\loc;\sandboxtag)\right\} \text{F}: \tmsmonitor.\text{F}\cup\left\{(\loc;\sandboxtag)\right\}\right\}
   }{
@@
   }{tms-dealloc}
+  %
 \end{center}
 
-For \gls*{tms}, the state of the monitor is a record with two sets keeping track of allocated and deallocated locations.
-%The monitor runs on a modified trace semantics with more abstract $\absev{Alloc\ }\loc$,$\absev{Dealloc\ }\loc$,$\absev{Use\ }\loc$, and $\absterm$ events, which represent {\em allocation}, {\em deallocation}, and {\em use} of a location $\loc$ as well as abnormal program termination.
 \Cref{tr:tms-use} simply requires that a location is (i) allocated and (ii) not freed.
@@
 
-\begin{center}
-  \judgbox{\tmstraceagree[]{\event}{\absevent}}{,,Abstract event $\absevent$ is equivalent to $\event$\\with respect to \gls*{tms}.''}$\;$\\
+\begin{center}\small
+  \judgbox{\tmstraceagree[]{\event}{\absevent}}{,,Abstract event $\absevent$ is equivalent to $\event$ w.r.t. \gls*{tms}.''}$\;$\\
   %
@@
 % %
-  \typerule{tms-abort-authentic}{
-  }{
-    \tmstraceagree[]{\lightning}{\absterm}
-  }{tms-abort-authentic}
+  % \typerule{tms-abort-authentic}{
+  % }{
+  %   \tmstraceagree[]{\lightning}{\absterm}
+  % }{tms-abort-authentic}
 \end{center}
 
-The trace agreement is entirely straightforward, so only allocation, branch, and crash are shown.
-\MP{
-	we should say something .. ?
-}
+The trace agreement maps actions to their respective (e.g., \Cref{tr:tms-alloc-authentic}), so we elide most cases.
+% 
+Since branches are not relevant for \gls*{tms}, they are mapped to $\absev{\emptyevent}$ (\Cref{tr:tms-branch-authentic}).
+% \MP{
+% 	we should say something .. ?
+% }
 
@@
 % 
-  Up to the body of \texttt{printf}, the program execution yields the specification trace 
+  Up to the body of \texttt{printf}, the program execution yields the specification trace: 
   	% $\ev{Alloc}\ \loc[_{x}]\ 12;\comp;\unlock\cdot\ev{Alloc}\ \loc[_{y}]\ 12;\comp;\unlock\cdot\ev{Use}\ \loc_{x}\ 0;\comp;\unlock\cdot\ev{Use}\ \loc_{x}\ 0;\comp;\unlock\cdot\ev{Use}\ \loc_{y}\ 0;\comp;\unlock\cdot\ev{Use}\ \loc_{x}\ 1;\comp;\unlock\cdot\ldots\cdot\ev{Use}\ \loc_{x}\ 12;\comp;\unlock\cdot\ev{Dealloc}\ \loc[_{y}];\comp;\unlock\cdot\ev{Use}\ \loc_{y}\ 6;\comp;\unlock$
@@
   	}
-Relating this trace to abstract monitor events yields trace
+Relating this trace to abstract monitor events yields trace:
 
@@
 
-  The fix would be to \lstinline[language=c++,basicstyle=\small\ttfamily]|delete greetings| instead of \texttt{to} and add a \lstinline[language=c++,basicstyle=\small\ttfamily]|delete to| after the \texttt{printf} call, which leads to the abstract monitor trace
+  The fix would be to \lstinline[language=c++,basicstyle=\small\ttfamily]|delete greetings| instead of \texttt{to} and add a \lstinline[language=c++,basicstyle=\small\ttfamily]|delete to| after the \texttt{printf} call, which leads to trace:
 
@@
 \begin{center}
-  \judgbox{\monitorcheck[]{\smsmonitor}{\smsmonitor[']}{\absevent}}{,,Monitor $\smsmonitor$ does one step to $\smsmonitor[']$\\given event $\absevent$.''}$\;$\\
+  \judgbox{\monitorcheck[]{\smsmonitor}{\smsmonitor[']}{\absevent}}{,,Monitor $\smsmonitor$ steps to $\smsmonitor[']$\\given event $\absevent$.''}$\;$\\
   %
@@
   At that stage, variable \texttt{i} is equal to \texttt{2} and, unfortunately, the order of checks is such that first the cell \texttt{x[i]} is read prior to bounds checking \texttt{i < n}, so there is an out-of-bounds memory access right before exiting the function.
-  This is also visible on the trace, which can be sketched as
+  This is also visible on the trace:
 
@@
 where $\loc_{x}$ and $\loc_{y}$ are the memory addresses associated to \texttt{x} and \texttt{y}, respectively.
-  Omitting the events for all ,,$\ldots$'' for sake of brevity, the abstract monitor trace of this is
+  Omitting the events for all ,,$\ldots$'' for sake of brevity, the abstract monitor trace of this is:
 
@@
 \begin{center}
-  \judgbox{\monitorcheck[]{\msmonitor}{\msmonitor[']}{\absevent}}{,,Monitor $\msmonitor$ does one step to $\msmonitor[']$\\given event $\absevent$.''}$\;$\\
+  \judgbox{\monitorcheck[]{\msmonitor}{\msmonitor[']}{\absevent}}{,,Monitor $\msmonitor$ steps to $\msmonitor[']$ given event $\absevent$.''}$\;$\\
   \typerule{ms-step}{
@@
 Note that monitors only apply to trace-properties, but not, e.g., relational hyperproperties~\cite{abate2019jour}.
+\MP{what}
 
@@
 \begin{center}
-  \judgbox{\monitorcheck[]{\scctmonitor}{\scctmonitor[']}{\absevent}}{,,Monitor $\scctmonitor$ does one step to $\scctmonitor[']$\\given event $\absevent$.''}$\;$\\
+  \judgbox{\monitorcheck[]{\scctmonitor}{\scctmonitor[']}{\absevent}}{,,Monitor $\scctmonitor$ steps to $\scctmonitor[']$\\given event $\absevent$.''}$\;$\\
   %
@@
 \begin{center}
-  \judgbox{\sccttraceagree[]{\event}{\absevent}}{,,Abstract event $\absevent$ is equivalent to $\event$\\with respect to \gls*{cct}.''}$\;$\\
+  \judgbox{\sccttraceagree[]{\event}{\absevent}}{,,Abstract event $\absevent$ is equivalent to $\event$ w.r.t. \gls*{cct}.''}$\;$\\
   %
@@
   %
-  \typerule{scct-empty-authentic}{
-  }{
-    \sccttraceagree[]{\emptyevent}{\absev{\emptyevent}}
-  }{scct-empty-authentic}
-  %
-  \typerule{scct-abort-authentic}{
-  }{
-    \sccttraceagree[]{\lightning}{\absterm}
-  }{scct-abort-authentic}
+  % \typerule{scct-empty-authentic}{
+  % }{
+  %   \sccttraceagree[]{\emptyevent}{\absev{\emptyevent}}
+  % }{scct-empty-authentic}
+  % %
+  % \typerule{scct-abort-authentic}{
+  % }{
+  %   \sccttraceagree[]{\lightning}{\absterm}
+  % }{scct-abort-authentic}
 \end{center}
@@
 This section presents the key meta-theoretic results of this paper concerning secure compiler compositions (\Cref{sec:sequential,sec:other-compos}).
-Note that the presented framework works with arbitrary classes.
-\MP{check, we ened more than this}
+% Note that the presented framework works with arbitrary classes.
+% \MP{check, we ened more than this}
 % and builds on prior work~\cite{kruse2022csc}\MK{mention this at all?
@@
 % MP: not needed at all, ever, workshop work does not need to be cited in the main work
+Note that since it deals with the preservation of \emph{classes}, which include both trace properties as well as hyperproperties, the presented framework is general: It describes robust compilers preserving ordinary trace properties, hyperproperties or even relational hyperproperties.
+% \MP{pull this out to earlier}
 
@@
 
-The main result is that secure compilers in the robust compilation framework~\cite{abate2019jour} compose {\em sequentially}.
-This is not intuitive in the sense that in the security domain, composition does not work without additional generalizations~\cite{canetti2006univcomp,mccullough2012compo,fabian2022automatic}.
-\MP{what deos this mean?}
+Our main result is that secure compilers in the robust compilation framework~\cite{abate2019jour} compose {\em sequentially}.
+% This is not intuitive in the sense that in the security domain, composition does not work without additional generalizations~\cite{canetti2006univcomp,mccullough2012compo,fabian2022automatic}.
+% \MP{what deos this mean?}
 The \bul{sequential composition of compilers $\cc{\src{L}}{\trg{L}}$ and $\cc{\trg{L}}{\obj{L}}$} is defined as follows:
@@
 Since the composition of secure compilers is again a secure compiler, the theorem generalises to a whole chain of $n$ secure compilers.
-
-\MP{move this}
-Note that the theorem may not be applicable in certain cases.
-For example, consider $\rtp{\cc{\src{L}}{\trg{L}}}{\tmssafe}$ and $\rtp{\cc{\trg{L}}{\obj{L}}}{\smssafe}$, where $\cc{\trg{L}}{\obj{L}}$ introduces code that allocates a table that stores bounds information of pointers, but fails to insert code that deallocates this table.
-By \Cref{def:trace:tmsdef}, this violates $\tmssafe$.
-Therefore, no behavior resulting from programs produced by the $\cc{\trg{L}}{\obj{L}}$ compiler is secure with respect to $\tmssafe$, but only for $\smssafe$.
-This, however, is a requirement of the preservation statement (the property of interest must be in $\class[_1]\cap\class[_2]$).
 
@@
 
-It is worth emphasizing that the framework is general: It supports both ordinary trace properties as well as hyperproperties or even relational hyperproperties~\cite{abate2019jour}.
-\MP{pull this out to earlier}
 Moreover, while this paper assumes a unified trace-model, it is merely a technical exercise to allow for different trace-models between programming languages~\cite{abate2021extacc}.
+
+\MP{moved this}
+Note that the theorem may not be applicable in certain cases.
+For example, consider $\rtp{\cc{\src{L}}{\trg{L}}}{\tmssafe}$ and $\rtp{\cc{\trg{L}}{\obj{L}}}{\smssafe}$, where $\cc{\trg{L}}{\obj{L}}$ introduces code that allocates a table that stores bounds information of pointers, but fails to insert code that deallocates this table.
+By \Cref{def:trace:tmsdef}, this violates $\tmssafe$.
+Therefore, no behavior resulting from programs produced by the $\cc{\trg{L}}{\obj{L}}$ compiler is secure with respect to $\tmssafe$, but only for $\smssafe$.
+This, however, is a requirement of the preservation statement (the property of interest must be in $\class[_1]\cap\class[_2]$).
 
@@
 \subsection{Shared Language Definitions}\label{subsec:cs:defs}
+\MPin{
+	note to fix as agreed upon
+	+ red syntax
+	+ green syntax
+}
 \begin{gather*}
@@
 $\src{L_{\tmssafe}}$ uses the same syntax as presented earlier (\Cref{sec:casestud:defs}) without extensions to the term level.
-But, $\src{L_{\tmssafe}}$ is statically typed, where the type system is inspired by $L^{3}$~\cite{morrisett2005L3,scherer2018fabulous}.
-The type system of $\src{L_{\tmssafe}}$ exhibits the property that every well-typed $\src{L_{\tmssafe}}$ program satisfies \gls*{tms} (\Cref{thm:wt:tms}).
+$\src{L_{\tmssafe}}$ is statically typed, where the type system is inspired by $L^{3}$~\cite{morrisett2005L3,scherer2018fabulous}.
+The type system of $\src{L_{\tmssafe}}$ enforces that every well-typed $\src{L_{\tmssafe}}$ program satisfies \gls*{tms} (\Cref{thm:wt:tms}).
 The proof of this theorem relies on a projection $\msfilterLtms{\src{\event}}=\event$ from $\src{L_{\tmssafe}}$ events to specification events $\event$, because the properties defined earlier (\Cref{subsec:propdefs}) are defined in the specification trace model.
@@
 
-\begin{center}
+\begin{center}\small
   \judgbox{\msfilterLtms{\src{\event}}=\event}{,,Project an $\Ltms$ event $\src{\event}$ to\\a specification event $\event$.''}
@@
 
-Most rules of the projection $\msfilterLtms{\src{\event}}$ are left out since, for the most part, it does the expected, e.g., $\msfilterLtms{\src{\ev{Dealloc}\ \loc;\comp;\securitytag}}=\ev{Dealloc}\ \delta(\loc);\comp;\securitytag$.
-But, it also filters any action that a context does as well as the interaction events, since these are irrelevant for component-level \gls*{tms}.
+Projection $\msfilterLtms{\src{\event}}$ translates an action into its correspondent (e.g., \Cref{tr:ltms-filter-alloc}); since these rules are straightforward they are mostly omitted.
+Crucially, the projection filters irrelevant actions such as context-made ones (\Cref{tr:ltms-filter-context}) and interaction events (\Cref{tr:ltms-filter-start}).
+
+% Most rules of the projection $\msfilterLtms{\src{\event}}$ are left out 
+% since, for the most part, it does the expected, e.g., $\msfilterLtms{\src{\ev{Dealloc}\ \loc;\comp;\securitytag}}=\ev{Dealloc}\ \delta(\loc);\comp;\securitytag$.
+% But, it also filters any action that a context does as well as the interaction events, since these are irrelevant for component-level \gls*{tms}.
+% \MP{rewrite these lines}
 
@@
 
-\vspace{-2em}
-\begin{center}
+\vspace{-1em}
+% \begin{center}
+{\small
   \begin{align*}
-    \mi{(Pre\text{-}Events)}~\obj{\event[_{b}]}\bnfdef&\ \cdots \mid \obj{\widehat{Get}\ \loc\ \valueexpr} \mid \obj{\widehat{Set}\ \loc\ \valueexpr\ \valueexpr[']}\\
-                                                      &\mid \obj{Branch\ n} \mid \obj{Binop\ n}
+    \mi{(Pre\text{-}Events)}~\obj{\event[_{b}]}\bnfdef&\ \cdots \mid \obj{\widehat{Get}\ \loc\ \valueexpr} \mid \obj{\widehat{Set}\ \loc\ \valueexpr\ \valueexpr[']}
+    % \\
+                                                      % &
+                                                      \mid \obj{Branch\ n} \mid \obj{Binop\ n}
   \end{align*}
-\end{center}
+% \end{center}
+}
 
@@
 This allows flexibility: The trace can be checked to satisfy different properties, such as, in this case, \gls*{tms}, \gls*{sms}, \gls*{scct}, and their combined versions.
-\Cref{ex:lscct} illustrates the differences of $\Lscct$ compared to the other languages.
+
+\Cref{ex:lscct} illustrates the differences between $\Lscct$ and other languages.
+
+\begin{exampleenv}[$\Lscct$ with and without data independent timing]\label{ex:lscct}
+  Consider again \Cref{ex:strncpy}, with a context copying the string \texttt{Hello World}, where everything is marked with a high security tag: $\lock$.
+  The left half of \Cref{fig:ex-cct} ($\Lscct$, the two columns on the left), describes the execution trace of the program, while the right side of the table ($Spec$, the two columns on the right), describes the respective projections $\scctfilterLscct{\obj{\trace}}$ to the specification trace model.
+  Read from top to bottom, the figure shows parts of the execution trace. 
+  In each half, the left column (\textit{Active}) has data-independent timing $\obj{ON}$ and the right one (\textit{Inactive}) has it $\obj{OFF}$.
+
+  When the data independent timing mode is off, the execution yields events in similar fashion to before (\Cref{subsec:ltms,subsec:lsms,subsec:lms}).
+  But, if it is turned on, then the branching event \colorbox{yellow}{does not fire} anymore and both reading and writing to memory gets ultimately translated to a specification trace with \colorbox{lightgray}{no exposed secrets}.
+\end{exampleenv}
 
 \begin{figure*}[!h]
-\begin{exampleenv}[$\Lscct$ with and without data independent timing]\label{ex:lscct}
-  Consider again \Cref{ex:strncpy}, with a context copying the string \texttt{Hello World}, where everything is marked with a security tag of high $\lock$.
-  The following table shows parts of the execution trace, read from top to bottom, in the left column (\textit{Active}) with and in the right column (\textit{Inactive}) without data independent timing.
-  The left side of the table ($\Lscct$), i.e., the two columns on the left, describes the execution trace of the program, while the right side of the table ($Spec$), i.e., the two columns on the right, describes the respective projections $\scctfilterLscct{\obj{\trace}}$ to the specification trace model.
-
-\vspace{-2em}
-  $${\footnotesize
+% \vspace{-2em}
+  $$
+  % {
+  \small
   \begin{array}{ccccccc}
@@
     \obj{\widehat{\ev{Get}}\ \loc_{y}\ 6;\comp;}\text{\colorbox{lightgray}{$\obj{\lock}$}} & \mid & \obj{\ev{Get}\ \loc_{y}\ 6;\comp;\lock} & \mid &  \ev{Use}\ \loc_{y}\ 6;\comp;\text{\colorbox{lightgray}{$\unlock$}} & \mid & \ev{Use}\ \loc_{y}\ 6;\comp;\lock\\
-  \end{array}}
+  \end{array}
+  % }
   $$
-
-  When the data independent timing mode is off, the execution yields events in similar fashion to before (\Cref{subsec:ltms,subsec:lsms,subsec:lms}).
-  But, if it is turned on, then the branching event \colorbox{yellow}{does not fire} anymore and both reading and writing to memory gets ultimately translated to a specification trace with \colorbox{lightgray}{no exposed secrets}.
-\end{exampleenv}
+  \caption{Traces for \Cref{ex:lscct}.}
+  \label{fig:ex-cct}
 \end{figure*}
@@
   \centering
+  \begin{minipage}[c]{0.64\textwidth}
   \begin{tikzpicture}
@@
   \end{tikzpicture}
-  \vspace{-2.5em}
+  \end{minipage}
+  \begin{minipage}{0.33\textwidth}
+  % \vspace{-2.5em}
   \caption{Visualisation of the optimising compilation pipeline that attains a combination of \gls*{ms} and \gls*{cct}. %
@@
   }\label{fig:pipeline}
+  \end{minipage}
 \end{figure*}
@@
 Since these properties compose into \gls*{ms}, composing these passes yields a compiler that robustly preserves \gls*{ms} (\Cref{thm:ccab:rtp:ms}).
-At this stage, the section presents two optimising translations, namely \gls*{cf} and \gls*{dce}, each of which robustly preserves \gls*{ms} (\Cref{thm:ccdce:rtp:ms,thm:cccf:rtp:ms}).
+Then, the section presents two optimising translations, namely \gls*{cf} and \gls*{dce}, each of which robustly preserves \gls*{ms} (\Cref{thm:ccdce:rtp:ms,thm:cccf:rtp:ms}).
 These translations can be freely ordered in the compilation chain without compromising memory safety (\Cref{thm:cccfccdce:rtp:ms}).
@@
 \vspace{-1em}
-\begin{gather*}
+\begin{gather*}\small
   \begin{aligned}
-    \cca(\src{x}) = & \trg{x} \\
-    \cca(\src{n}) = & \trg{n} \\
-    \cca(\src{\lbinop{\expr[_{1}]}{\expr[_{2}]}}) = & \lbinop[\trg]{\left[\cca(\src{\expr[_{1}]})\right]}{\left[\cca(\src{\expr[_{2}]})\right]} \\
-    \cca(\src{\lget{x}{\expr}}) = & \lget[\trg]{\trg{x}}{\left[\cca(\src{\expr})\right]} \\
-    \cca(\src{\ldelete{x}}) = & \ldelete[\trg]{\left[\cca(\src{x})\right]} \\
+    \cca(\src{x}) = &\ \trg{x} 
+    % \\
+  	&
+    \cca(\src{\lbinop{\expr[_{1}]}{\expr[_{2}]}}) = &\ \lbinop[\trg]{\left[\cca(\src{\expr[_{1}]})\right]}{\left[\cca(\src{\expr[_{2}]})\right]} \\
+    \cca(\src{n}) = &\ \trg{n} 
+    % \\
+    &
+    \cca(\src{\lget{x}{\expr}}) = &\ \lget[\trg]{\trg{x}}{\left[\cca(\src{\expr})\right]}
+  \end{aligned}
+    \\
+  \begin{aligned}
+    \cca(\src{\ldelete{x}}) = &\ \ldelete[\trg]{\left[\cca(\src{x})\right]} \\
 %    \cca(\src{\lset{x}{\expr[_{1}]}{\expr[_{2}]}}) & = \lset[\trg]{\left[\cca(\src{x})\right]}{\left[\cca(\src{\expr[_{1}]})\right]}{\left[\cca(\src{\expr_{2}})\right]} \\
-\cca(\src{\lfunction{g}{x:\natt\to\type[_{e}]}{\expr}})  = & \lfunction[\trg]{\trg{g}}{\trg{x}}{\lifz[\trg]{\trg{\lhast{x}{\natt}}}{\\%
+\cca(\src{\lfunction{g}{x:\natt\to\type[_{e}]}{\expr}})  = &\ \lfunction[\trg]{\trg{g}}{\trg{x}}{\lifz[\trg]{\trg{\lhast{x}{\natt}}}{\\%
                                                             &\left[\cca(\src{\expr})\right] %
@@
 Compiling the \texttt{strncpy} function from \Cref{sec:introduction} with $\cca$, the compiler would in this case ensure that the arguments that are evaluated in the compiled \texttt{strncpy} are valid.
+% \MP{put the proof section in an appendix}
 
@@
 \vspace{-1em}
-\begin{center}
+\begin{center}\small
   $$
@@
 \end{theorem}
-\begin{proof}
-  From \Thmref{thm:cca:rtp:tms} it follows that for any $\Ltms$ program $\src{\progvar}$, it compiles to an $\Ltrg$ program $\trg{\progvar}$ that robustly satisfies \gls*{tms}.
-  Note that $\src{\progvar}$ robustly satisfies \gls*{tms} by the properties of the typesystem of $\Ltms$.
-  Then, \Thmref{thm:ccb:rtp:sms} demonstrates that, assuming $\trg{\progvar}$ robustly satisfies \gls*{sms}, the program $\trg{\progvar}$ compiles to an $\Lms$ program $\irl{p}$ that also robustly satisfies \gls*{sms}.
-  From \Thmref{thm:ccab:rtp:ms} it follows that $\src{p}$ compiles to $\irl{p}$ that robustly satisfies \gls*{ms}, since \gls*{ms} is the intersection of \gls*{tms} and \gls*{sms}.
-\end{proof}
+% \begin{proof}\MP{keep?}
+%   From \Thmref{thm:cca:rtp:tms} it follows that for any $\Ltms$ program $\src{\progvar}$, it compiles to an $\Ltrg$ program $\trg{\progvar}$ that robustly satisfies \gls*{tms}.
+%   Note that $\src{\progvar}$ robustly satisfies \gls*{tms} by the properties of the typesystem of $\Ltms$.
+%   Then, \Thmref{thm:ccb:rtp:sms} demonstrates that, assuming $\trg{\progvar}$ robustly satisfies \gls*{sms}, the program $\trg{\progvar}$ compiles to an $\Lms$ program $\irl{p}$ that also robustly satisfies \gls*{sms}.
+%   From \Thmref{thm:ccab:rtp:ms} it follows that $\src{p}$ compiles to $\irl{p}$ that robustly satisfies \gls*{ms}, since \gls*{ms} is the intersection of \gls*{tms} and \gls*{sms}.
+% \end{proof}
 
@@
     \ccdce(\irl{\lifz{false}{\expr[_{1}]}{\expr[_{2}]}}) & = \ccdce(\irl{\expr[_{2}]}) &
-  \end{align*}\\
-  \begin{align*}
-    \ccdce(\irl{\lbinop{\expr[_{1}]}{\expr[_{2}]}}) & = \lbinop[\irl]{\ccdce(\irl{\expr[_{1}]})}{\ccdce(\irl{\expr[_{2}]})} &
-  \end{align*}\\[0.25cm]
+  \end{align*}
+  \\
+  % \begin{align*}
+  %   \ccdce(\irl{\lbinop{\expr[_{1}]}{\expr[_{2}]}}) & = \lbinop[\irl]{\ccdce(\irl{\expr[_{1}]})}{\ccdce(\irl{\expr[_{2}]})} &
+  % \end{align*}
+  % \\[0.25cm]
   \begin{align*}
     \cccf(\irl{\expr}) & = \partialeval{\irl{\expr}}{\irl{\hole{\cdot}}} &
-  \end{align*}\\[0.125cm]
+  % \end{align*}
+  \\
+  % [0.125cm]
 % \begin{align*}
-%   \partialeval{\irl{x}}{\irl{\substlist}} & = \irl{n} & \text{if } \irl{\subst{n}{x}}\in\irl{\substlist} \\
-%   \partialeval{\irl{x}}{\irl{\substlist}} & = \irl{x} & \text{if } \irl{\subst{n}{x}}\notin\irl{\substlist} \\
-%   \partialeval{\irl{\lbinop{n}{m}}}{\irl{\substlist}} & = \irl{k} & \text{if } \lbinop{\irl{n}}{\irl{m}}=k \\
+  \partialeval{\irl{x}}{\irl{\substlist}} & = \irl{n} 
+  	\qquad\qquad \text{if } \irl{\subst{n}{x}}\in\irl{\substlist} \\
+  \partialeval{\irl{x}}{\irl{\substlist}} & = \irl{x} 
+  \qquad\qquad \text{if } \irl{\subst{n}{x}}\notin\irl{\substlist} \\
+  \partialeval{\irl{\lbinop{n}{m}}}{\irl{\substlist}} & = \irl{k} 
+  \qquad\qquad \text{if } \lbinop{\irl{n}}{\irl{m}}=k \\
    % \partialeval{\irl{\lbinop{\expr[_{1}]}{\expr[_{2}]}}}{\irl{\substlist}} & = \lbinop[\irl]{\partialeval{\irl{\expr[_{1}]}}{\irl{\substlist}}}{\partialeval{\irl{\expr[_{2}]}}{\irl{\substlist}}} \\
-%   \partialeval{\irl{\llet{x}{n}{\expr}}}{\irl{\substlist}} & = \partialeval{\irl{\expr}}{\irl{\subst{x}{n}},\irl{\substlist}} \\
+  \partialeval{\irl{\llet{x}{n}{\expr}}}{\irl{\substlist}} & = \partialeval{\irl{\expr}}{\irl{\subst{x}{n}\cdot\substlist}} 
+  % \\
   % \partialeval{\irl{\lget{x}{\expr}}}{\irl{\substlist}} & = \lget[\irl]{\irl{x}}{\partialeval{\irl{\expr}}{\irl{\substlist}}}
-% \end{align*}\\
+\end{align*}
+% \\
 % \begin{align*}
@@
 \end{gather*}
-\vspace{-3em}
-
+% \vspace{-3em}
+
+% \MP{mix ??}
 The two optimising compiler passes from $\Lms$ to $\Lms$ perform \gls*{dce} and \gls*{cf}, respectively.
 The \gls*{dce} pass applies a naive rewrite rule on conditionals.
-For \gls*{cf}, the pass uses an auxiliary function \texttt{mix} that does the actual work:
-It rewrites constant binary operations, e.g., $\irl{{17}-1}$ to $\irl{16}$, and replaces variables that are assigned to constants with their constant, e.g., $\irl{\llet{x}{7}{x}}$ to $\irl{7}$.
+The \gls*{cf} pass relies on an auxiliary function \texttt{mix} that uses a substitutions accumulator $\irl{\substlist}$ in order to rewrite constant binary operations, e.g., $\irl{{17}-1}$ to $\irl{16}$, and replace variables that are assigned to constants with their constant, e.g., $\irl{\llet{x}{7}{x}}$ to $\irl{7}$.
 Both passes are secure with respect to \gls*{ms}.
@@
 
-\vspace{-2em}
-\begin{center}
+\vspace{-1em}
+\begin{center}\small
   \begin{align*}
     \ccscct(\irl{\lfunction{g}{x}{\expr}}) & = \lfunction[\obj]{\obj{g}}{\obj{x}}{\obj{\lwrdoit{ON};}\ccscct(\irl{\expr})} \\
-    \ccscct(\irl{\lcall{g}{\expr}}) & = \lcall[\obj]{\obj{g}}{\ccscct(\irl{\expr})\obj{; \lwrdoit{ON}}} \\
-    \ccscct(\irl{\lbinop{\expr[_{1}]}{\expr[_{2}]}}) & = \lbinop[\obj]{\ccscct{\irl{\expr[_{1}]}}}{\ccscct{\irl{\expr[_{2}]}}} \\
+    \ccscct(\irl{\lcall{g}{\expr}}) & = \lcall[\obj]{\obj{g}}{\ccscct(\irl{\expr})\obj{; \lwrdoit{ON}}} 
+    % \\
+    % \ccscct(\irl{\lbinop{\expr[_{1}]}{\expr[_{2}]}}) & = \lbinop[\obj]{\ccscct{\irl{\expr[_{1}]}}}{\ccscct{\irl{\expr[_{2}]}}} 
+    % \\
   \end{align*}
 \end{center}
-\vspace{-2em}
+% \vspace{-2em}
 %
@@
 \end{theorem}
-\begin{proof}
-	From \Thmref{thm:ccab:rtp:ms}, we have that any $\Ltms$ program $\src{\progvar}$ compiles into a $\Lms$ program $\irl{\progvar}$ that robustly satisfies \gls*{ms}.
-	Then, from \Thmref{thm:cccfccdce:rtp:ms} we have that $\irl{\progvar}$ gets optimised to a program $\irl{\progvar[']}$ that is also \gls*{ms}, where the order of optimisations does not matter for $\irl{\progvar[']}$ to be \gls*{ms}.
-  Assuming $\irl{\progvar[']}$ robustly satisfies \gls*{scct}, by \Thmref{thm:ccscct:rtp:scct} it compiles to an $\Lscct$ program $\obj{\progvar}$ that robustly satisfies \gls*{scct} as well.
-  Finally, from \Thmref{thm:rtp} it follows that, given $\src{\progvar}$ robustly satisfies \gls*{scct} and \gls*{ms}, $\obj{\progvar}$ also robustly satisfies \gls*{scct} and \gls*{ms}.
-\end{proof}
+% \begin{proof}\MP{keep?}
+% 	From \Thmref{thm:ccab:rtp:ms}, we have that any $\Ltms$ program $\src{\progvar}$ compiles into a $\Lms$ program $\irl{\progvar}$ that robustly satisfies \gls*{ms}.
+% 	Then, from \Thmref{thm:cccfccdce:rtp:ms} we have that $\irl{\progvar}$ gets optimised to a program $\irl{\progvar[']}$ that is also \gls*{ms}, where the order of optimisations does not matter for $\irl{\progvar[']}$ to be \gls*{ms}.
+%   Assuming $\irl{\progvar[']}$ robustly satisfies \gls*{scct}, by \Thmref{thm:ccscct:rtp:scct} it compiles to an $\Lscct$ program $\obj{\progvar}$ that robustly satisfies \gls*{scct} as well.
+%   Finally, from \Thmref{thm:rtp} it follows that, given $\src{\progvar}$ robustly satisfies \gls*{scct} and \gls*{ms}, $\obj{\progvar}$ also robustly satisfies \gls*{scct} and \gls*{ms}.
+% \end{proof}
 
@@
 Then, it proved that composing secure compilers that preserve certain properties results in a secure compiler that preserves the composition of these properties.
-While the presented security property composition relied on monitors that check only trace-properties, the composition of secure compilers does not make any restriction towards the kind of property involved in the composition.
-It is subject to future work to develop techniques to verify relational hyperproperties~\cite{abate2019jour}, while the composition of hyperproperties could be very similar to the composition of ordinary properties as presented in this paper, since hyperproperties can be checked with an automata-based model-checker~\cite{beutner23hyperltl}.
+% While the presented security property composition relied on monitors that check only trace-properties, the composition of secure compilers does not make any restriction towards the kind of property involved in the composition.
+% It is subject to future work to develop techniques to verify relational hyperproperties~\cite{abate2019jour}, while the composition of hyperproperties could be very similar to the composition of ordinary properties as presented in this paper, since hyperproperties can be checked with an automata-based model-checker~\cite{beutner23hyperltl}.
 % 
@@
 Crucially, this paper derives the security of the multi-pass compiler from the composition of the security properties preserved by its individual passes, which include security-preserving as well as optimisation passes.
-For future work, it is interesting to look at more sophisticated optimisation passes that, e.g., reorder events that appear on traces.
+% For future work, it is interesting to look at more sophisticated optimisation passes that, e.g., reorder events that appear on traces.
 
@@
 
-\section*{Acknowledgements}\label{sec:acks}
+\newpage
+% \section*{Acknowledgements}\label{sec:acks}
+
+\appendices
+
+\section*{Appendix}
+
+The appendix contains insights on the structure of proofs for the theorems of \Cref{sec:casestud:rtp} (\Cref{sec:proof-sketch}).
+Then it contains the proof of \Thmref{thm:ccab:rtp:ms} (\Cref{sec:proof1}) and of \Thmref{thm:ccall:rtp:msscct} (\Cref{sec:proof2}).
+
+\section{Proving Robust Safety Property Preservation}\label{sec:proof-sketch}
+
+We illustrate the proof of \Cref{thm:cca:rtp:tms} since the other secure compilation proofs of this paper follow the same approach.
+Unfolding the theorem statement yields the following assumptions:  
+  for any $\pi\in\lceil\tmssafe\rceil$\footnote{$\lceil\cdot\rceil$ lifts the property to a hyperproperty by applying the powerset operation~\cite{clarkson2008hyper}.}, 
+  $\src{\trace}$, 
+  $\src{\runtimetermvar}$, 
+  and component $\src{\library_{comp}}$, 
+  we have that $\rsat{\src{\library_{comp}}}{\pi}$ and 
+  $\progstepto{\prog[\trg]{\trg{\library_{\ctx}}}{\cca(\src{\library_{\comp}})}}{\trg{\runtimetermvar}}{\trg{\trace}}$, where $\trg{\library_{\ctx}}$ is arbitrary.
+The proof obligation is $\msfilterL{\trg{\trace}}\in\pi$, i.e., the specification trace associated to $\trg{\trace}$ satisfies the property $\pi$.
+A way to show this is to relate trace $\trg{\trace}$ to some $\Ltms$ trace $\src{\trace}$ (which already satisfies the property as per the assumptions).
+The assumptions already contain a target execution associated to this trace, so the task is to find an associated $\Ltms$ execution that yields $\trg{\trace}$.
+The trace $\trg{\trace}$ is split into different parts, as commonly done in secure compilation works~\cite{korashy2021capableptrs,abate2018goodcomps}, where each part contains the events that either the context or the component does, but not both.
+Because of this, all such trace segments are ,,well-bracketed'' in the sense that they start with either $\trg{\ev{Start}}$, $\trg{\ev{Call}\ \comptoctx\ foo\ \valueexpr}$, or $\trg{\ev{Ret}\ \comptoctx\ \valueexpr}$ and end with either $\trg{\ev{End}\ \valueexpr}$, $\trg{\ev{Ret}\ \ctxtocomp\ \valueexpr}$, or $\trg{\ev{Call}\ \ctxtocomp\ foo\ \valueexpr}$.
+In the following, the former is referred to as a context segment, since these executions happen in $\trg{\library_{\ctx}}$, and the latter is referred to as a component segment, since these executions happen in $\cca(\src{\library_{\comp}})$.
+\Cref{fig:proofdiag:rtp} visualises this division for a program execution with one call from context to component and how the target execution is related to a source execution.
+In the figure, the green dashed lines encompass the component segments while the orange boxes contain the actual context switches from context to component or vice versa.
+From a technical perspective, as typically done in compilation proofs, the proof requires some setup to maintain a relation between $\src{\Omega}$ and $\trg{\Omega}$ .
+Two cross-language relations make this precise: (i) $\multimap_{\delta}$ relates states that are involved in a context segment, allowing the target execution to perform internal calls, and (ii) $\approx_{\delta}$ relates states that are involved in a component segment, where both states need to agree exactly, i.e., the memory and the control flow states are required to contain the same information.
+The relations are indexed with $\delta$, which is an injective mapping from $\Ltms$ locations $\src{\loc}$ to $\Ltrg$ locations $\trg{\loc}$.
+%Note that the relations $\multimap_{\delta}$ and $\approx_{\delta}$ swap when context switching.
+
+So far, the paper explained how to relate an $\Ltms$ execution with a $\Ltrg$ execution.
+The next question is therefore how to build the corresponding $\Ltms$ execution.
+This is done using a standard secure compilation proof technique called trace-based backtranslation~\cite{abate2019jour,korashy2021capableptrs,patrignani2021rsc}, which can be used to build a context $\src{\library_{\ctx}}$ that behaves similar to $\trg{\library_{\ctx}}$.
+% Since $\Ltms$ is statically typed, the context $\src{\library_{\ctx}}$ obtained from the backtranslation needs to be well-typed when linked with $\src{\library_{\comp}}$.
+For context segments of the trace $\trg{\trace}$ it is also necessary to show that the execution behaves similarily, i.e., the context obtained from the backtranslation generates trace $\src{\trace}$.
+For component segments of the trace, the relatedness of states and traces follows from a compiler correctness argument.
+These two arguments yield the source execution $\progstepto{\src{\prog{\library_{\ctx}}{\library_{\comp}}}}{\src{\runtimetermvar}}{\src{\trace}}$.
+
+The proof now works as follows.
+Given that $\msfilterL{\trg{\trace}}=\msfilterLtms{\src{\trace}}$, the proof goal changes from $\msfilterL{\trg{\trace}}\in\pi$ to $\msfilterLtms{\src{\trace}}\in\pi$.
+This follows by specializing the robust satisfaction assumption $\rsat{\src{\library_{\comp}}}{\pi}$ to use the context $\src{\library_{\ctx}}$, which is obtained from the backtranslation, and to use the source execution $\progstepto{\src{\prog{\library_{\ctx}}{\library_{\comp}}}}{\src{\runtimetermvar}}{\src{\trace}}$.
+
+\begin{figure*}[!ht]
+  \begin{center}
+    \begin{tikzpicture}[state/.style={minimum height=1.5cm}]
+      % relative horizontal/vertical distance between states
+      \pgfmathsetmacro{\hdist}{0.95}
+      \pgfmathsetmacro{\vdist}{1.35}
+      \pgfmathsetmacro{\halfvdist}{0.725}
+
+      % row of src states
+      \node[state] (srcempty) {$\rtt{\src{\emptyset}}{\src{\lcall{main}{0}}}$};
+      \foreach \s [remember=\s as \cur (initially empty)] in {1,w_1,p,w_2,2} {
+        \node[state,right=\hdist of src\cur] (src\s) {$\rtt{\src{\Omega_{\s}}}{\src{\expr[_{\s}]}}$};
+      }
+      % row of trg states
+      \node[state,below=\vdist of srcempty] (trgempty) {$\rtt{\trg{\emptyset}}{\trg{\lcall{main}{0}}}$};
+      \foreach \s in {1,w_1,p,w_2,2} {
+        \node[state,below=\vdist of src\s] (trg\s) {$\rtt{\trg{\Omega_{\s}}}{\trg{\expr[_{\s}]}}$};
+      }
+      %% illustrations
+        % backtrans wrapper 1
+        \draw[ultra thick,loosely dotted,Peach!50,rounded corners] (src1.north east) -- (srcw\string_1.north east)
+          -- (trgw\string_1.south east) -- (trg1.south west) -- (src1.north west) -- cycle;
+        % backtrans wrapper 2
+        \draw[ultra thick,loosely dotted,Peach!50,rounded corners] (srcp.north east) -- (srcw\string_2.north east)
+          -- (trgw\string_2.south east) -- (trgp.south west) -- (srcp.north west) -- cycle;
+        % compiler correctness
+        \draw[ultra thick,loosely dashed,Emerald!50,rounded corners] ($(srcw\string_1.north east)+(0,0.05)$) -- ($(srcp.north east)+(0,0.05)$)
+          -- ($(trgw\string_2.north east)+(0.05,0)$) -- ($(trgw\string_2.south east)+(0.05,-0.05)$)
+          -- ($(trg1.south west)+(-0.05,-0.05)$) -- ($(trg1.north west)+(-0.05,0)$) -- ($(srcw\string_1.north west)+(0,0.05)$) -- cycle;
+      % state relations
+      \path (srcempty) edge[very thick,draw=gray!25] node[pos=0.5,sloped,rotate=180,fill=white] {\scriptsize$\multimap_\emptyset$} (trgempty)
+        (src1) edge[very thick,draw=gray!25] node[pos=0.5,sloped,rotate=180,fill=white] {\scriptsize$\multimap_{\delta_1}$} (trg1)
+        (srcp) edge[very thick,draw=gray!25] node[pos=0.5,sloped,fill=white] {\scriptsize$\approx_{\delta_p}$} (trgp)
+        (src2) edge[very thick,draw=gray!25] node[pos=0.5,sloped,rotate=180,fill=white] {\scriptsize$\multimap_{\delta_2}$} (trg2)
+        (srcw\string_1) edge[very thick,draw=gray!25] node[pos=0.5,sloped,fill=white] {\scriptsize$\approx_{\delta_{w_1}}$} (trgw\string_1)
+        (srcw\string_2) edge[very thick,draw=gray!25] node[pos=0.5,sloped,rotate=180,fill=white] {\scriptsize$\multimap_{\delta_{w_2}}$} (trgw\string_2)
+        % diagonals
+        (srcw\string_1) edge[thick,draw=gray!25] node[pos=0.16,sloped,rotate=180,fill=white] {\scriptsize$\approx_{\delta_{w_1}}$} (trg1)
+        (src1) edge[thick,draw=gray!25] node[pos=0.16,sloped,rotate=180,fill=white] {\scriptsize$\multimap_{\delta_1}$} (trgw\string_1)
+        (srcp) edge[thick,draw=gray!25] node[pos=0.2,sloped,fill=white] {\scriptsize$\approx_{\delta_p}$} (trgw\string_2)
+        (trgp) edge[thick,draw=gray!25] node[pos=0.2,sloped,fill=white] {\scriptsize$\multimap_{\delta_{w_2}}$} (srcw\string_2)
+        ;
+      %\drawpolygon src1,srcw\string_1,trgw\string_1,trg1;
+      %\drawpolygon srcp,srcw\string_2,trgw\string_2,trgp;
+      %\node[font=\tiny,align=center,above=0.2 of srcw1srcp] (wrapper) {Backtranslation\\Wrapper};
+      %\path[->,draw] (wrapper) -- (srcw\string_1);
+      %\path[->,draw] (wrapper) -- (srcp);
+      % steps
+      \path[very thick,color=\stlccol] (srcempty) edge[draw=none] node {\ $\xrightarrow{\trace[_1]}{}{\kern-3.5pt}_{\text{ctx}}^*$} (src1)
+        (src1) edge[draw=none] node {\ $\xrightarrow{\trace[_c]}{}{\kern-3.5pt}_{\text{ctx}}$} (srcw\string_1)
+        (srcw\string_1) edge[draw=none] node {\ $\xrightarrow{\trace[_p]}{}{\kern-3.5pt}_{\text{ctx}}^*$} (srcp)
+        (srcp) edge[draw=none] node[fill=white,inner sep=0,outer sep=0] {\ $\xrightarrow{\trace[_r]}{}{\kern-3.5pt}_{\text{ctx}}$} (srcw\string_2)
+        (srcw\string_2) edge[draw=none] node {\ $\xrightarrow{\trace[_2]}{}{\kern-3.5pt}_{\text{ctx}}^*$} (src2)
+        ;
+      \path[very thick,color=\ulccol] (trgempty) edge[draw=none] node[fill=white,inner sep=0,outer sep=0] {\ $\xrightarrow{\phantom{\trace[_1]}}{}{\kern-3.5pt}_{\text{ctx}}^*$} (trg1)
+        (trg1) edge[draw=none] node {\ $\xrightarrow{\phantom{\trace[_p]}}{}{\kern-3.5pt}_{\text{ctx}}^*$} (trgw\string_1)
+        (trgw\string_1) edge[draw=none] node[fill=white,inner sep=0,outer sep=0] {\ $\xrightarrow{\phantom{\trace[_p]}}{}{\kern-3.5pt}_{\text{ctx}}^*$} (trgp)
+        (trgp) edge[very thick,draw=none] node {\ $\xrightarrow{\phantom{\trace[_p]}}{}{\kern-3.5pt}_{\text{ctx}}^*$} (trgw\string_2)
+        (trgw\string_2) edge[draw=none] node {\ $\xrightarrow{\phantom{\trace[_2]}}{}{\kern-3.5pt}_{\text{ctx}}^*$} (trg2)
+        ;
+      % legend
+%     \node[align=left,below right=0.3 and 0.3 of trgempty,font=\tiny] (legend) {%
+%       ${\src{\trace[_c]}}\cong{\backt{\Ltms}{\Ltrg}(\trg{Call\ ?\ foo\ \valueexpr})}$\\%
+%       ${\src{\trace[_r]}}\cong{\backt{\Ltms}{\Ltrg}(\trg{Ret\ !\ foo\ \valueexpr})}$
+%       };
+      \node[below=0.5 of trgempty,draw=none] (legend) {};
+      \draw[ultra thick,loosely dotted,Peach!50,rounded corners] ($(legend.north east)+(0.5,-0.4)$) -- ($(legend.north east)+(1,-0.4)$);
+      \node at ($(legend.north east)+(1.5,-0.4)$) (legendwrapper) {};
+      \node[right of=legendwrapper] {\tiny Backtranslation Wrapper};
+      \draw[ultra thick,loosely dashed,Emerald!50,rounded corners] ($(legend.south east)+(0.5,0.4)$) -- ($(legend.south east)+(1,0.4)$);
+      \node at ($(legend.south east)+(1.5,0.4)$) (legendcorrectness) {};
+      \node[right of=legendcorrectness] {\tiny Compiler Correctness};
+    \end{tikzpicture}
+    \caption{Proof diagram for \Cref{thm:cca:rtp:tms} depicting the general structure of robust preservation proofs. %
+      Nodes in the graph represent runtime states. %
+      Vertical lines indicate cross language relations, while horizontal ones are execution steps. %
+      The green dashed trapezoid encompasses the component segment, while the orange dotted rectangles entail the context switches. %
+      $\Ltrg$ traces are omitted for readability. %
+      $\Ltms$ trace segments $\src{\trace[_{c}]}$ and $\src{\trace[_{r}]}$ describe the events that happen at the boundaries, i.e., during a context switch.
+      $\src{\trace_{p}}$ is the behavior of the component and the traces $\src{\trace[_{1}]}$ and $\src{\trace[_{2}]}$ describe the context.
+    }\label{fig:proofdiag:rtp}
+  \end{center}
+\end{figure*}
+
+\section{Proof of \Cref{thm:ccab:rtp:ms}} \label{sec:proof1}
+  From \Thmref{thm:cca:rtp:tms} it follows that for any $\Ltms$ program $\src{\progvar}$, it compiles to an $\Ltrg$ program $\trg{\progvar}$ that robustly satisfies \gls*{tms}.
+  Note that $\src{\progvar}$ robustly satisfies \gls*{tms} by the properties of the typesystem of $\Ltms$.
+  Then, \Thmref{thm:ccb:rtp:sms} demonstrates that, assuming $\trg{\progvar}$ robustly satisfies \gls*{sms}, the program $\trg{\progvar}$ compiles to an $\Lms$ program $\irl{p}$ that also robustly satisfies \gls*{sms}.
+  From \Thmref{thm:ccab:rtp:ms} it follows that $\src{p}$ compiles to $\irl{p}$ that robustly satisfies \gls*{ms}, since \gls*{ms} is the intersection of \gls*{tms} and \gls*{sms}.
+
+\section{Proof of \Cref{thm:ccall:rtp:msscct}} \label{sec:proof2}
+From \Thmref{thm:ccab:rtp:ms}, we have that any $\Ltms$ program $\src{\progvar}$ compiles into a $\Lms$ program $\irl{\progvar}$ that robustly satisfies \gls*{ms}.
+Then, from \Thmref{thm:cccfccdce:rtp:ms} we have that $\irl{\progvar}$ gets optimised to a program $\irl{\progvar[']}$ that is also \gls*{ms}, where the order of optimisations does not matter for $\irl{\progvar[']}$ to be \gls*{ms}.
+Assuming $\irl{\progvar[']}$ robustly satisfies \gls*{scct}, by \Thmref{thm:ccscct:rtp:scct} it compiles to an $\Lscct$ program $\obj{\progvar}$ that robustly satisfies \gls*{scct} as well.
+Finally, from \Thmref{thm:rtp} it follows that, given $\src{\progvar}$ robustly satisfies \gls*{scct} and \gls*{ms}, $\obj{\progvar}$ also robustly satisfies \gls*{scct} and \gls*{ms}.
+
+\newpage
 
@@
 %% If your work has an appendix, this is the place to put it.
-\appendix
+% \appendix
 

### 2023-09-27 csf/main.tex
Change text
Additions to seqcompo

@@
 Any technique to combine existing monitors $T_A$, $T_B$ (that check for properties $A$ and $B$) into one $T_{AB}$ (that checks for both $A$ and $B$) is sufficient for the sake of the analysis of this paper.
-Note that monitors only apply to trace-properties, but not, e.g., relational hyperproperties~\cite{abate2019jour}.
-\MP{what}
 
@@
 Given an $\src{L}$ program $\src{p}$ and compilers $\cc{\src{L}}{\trg{L}},\cc{\trg{L}}{\obj{L}}$, \rul{its compiled $\obj{L}$ counterpart is obtained by plugging $\src{p}$ into $\cc{\src{L}}{\trg{L}}\circ\cc{\trg{L}}{\obj{L}}$}.
-\begin{definition}
+\begin{definition}[Sequential Composition]
   \bul{$\cc{\src{L}}{\trg{L}}\circ\cc{\trg{L}}{\obj{L}}$} $\isdef$ Given $\src{p}$, yield \rul{$\cc{\trg{L}}{\obj{L}}\left(\cc{\src{L}}{\trg{L}}(\src{p})\right)$}
@@
 
-Moreover, while this paper assumes a unified trace-model, it is merely a technical exercise to allow for different trace-models between programming languages~\cite{abate2021extacc}.
-
-\MP{moved this}
-Note that the theorem may not be applicable in certain cases.
+So far, this paper assumes a unified trace-model and secure compilers that do not change the behavior.
+It is worth noting that in this setting, the theorem may not be applicable.
 For example, consider $\rtp{\cc{\src{L}}{\trg{L}}}{\tmssafe}$ and $\rtp{\cc{\trg{L}}{\obj{L}}}{\smssafe}$, where $\cc{\trg{L}}{\obj{L}}$ introduces code that allocates a table that stores bounds information of pointers, but fails to insert code that deallocates this table.
-By \Cref{def:trace:tmsdef}, this violates $\tmssafe$.
+By \Thmref{def:trace:tmsdef}, this violates $\tmssafe$.
 Therefore, no behavior resulting from programs produced by the $\cc{\trg{L}}{\obj{L}}$ compiler is secure with respect to $\tmssafe$, but only for $\smssafe$.
-This, however, is a requirement of the preservation statement (the property of interest must be in $\class[_1]\cap\class[_2]$).
+$\cc{\trg{L}}{\obj{L}}$ modifies the behavior and because of this, the theorem is inapplicable.
+
+This section re-introduces secure compiler composition without this limitation. 
+To this end, it utilises existing work on secure compilation~\cite{abate2021extacc} that defines Galois connections for several real-world compiler scenarios, such as a compilation pass that exploits undefined behavior.
+This section is concerned about source-level properties, and henceforth uses the existential image of a source-level property $\src{\pi}$ to map $\src{\pi}$ to a target-level property:
+
+\begin{definition}[Existential Image of Properties]\label{def:eximg:prop}
+  Given a cross-language trace relation $\sim$,
+  $$
+  \tau_\sim(\src{\pi}) \isdef \left\{
+    \trg{\trace} \mid \exists\src{\trace}.\src{\trace}\sim\trg{\trace} \text{ and }
+    \src{\trace}\in\src{\pi}
+  \right\}
+  $$
+\end{definition}
+
+\Thmref{def:eximg:prop} can be lifted to hyperproperties/classes as follows:
+\begin{definition}[Existential Image of Hyperproperties]\label{def:eximg:hprop}
+  Given a cross-language trace relation $\sim$,
+  $$
+  \tilde{\tau}_\sim(\src{\Pi}) \isdef \left\{
+    \trg{\pi} \mid \exists\src{\pi}.\src{\pi}\in\src{\Pi} \text{ and }
+    \trg{\pi}=\tau_\sim\left(\src{\pi}\right)
+  \right\}
+  $$
+\end{definition}
+
+\bul{A cross-language relation $\sim$ is well-formed with respect to $\src{\Pi}$} iff \rul{the existential images preserve membership}.
+\begin{definition}[Well-formedness of Cross-Language Trace Relations for a Concrete Property $\src{\Pi}$]\label{def:wf:tracerel}
+  $$
+  \text{\bul{$\wf{\sim}{\src{\Pi}}$}} \isdef \text{\rul{$\forall \src{\pi}\in\src{\Pi}, \tau_\sim(\src{\pi})\in\tilde{\tau}_\sim(\src{\Pi})$}}
+  $$
+\end{definition}
+\begin{definition}[Well-formedness of Cross-Language Trace Relations for a Class $\src{\class}$]\label{def:wfc:tracerel}
+  $$
+  \wfc{\sim}{\src{\class}} \isdef \forall \src{\Pi}\subseteq\src{\class}, \tau_\sim(\src{\Pi})\subseteq\tilde{\tau}_\sim(\src{\class})
+  $$
+\end{definition}
+\begin{lemma}[Well-Formedness Equivalence for Classes]\label{lem:wf:equiv:wfc}
+  $\;\;$
+  $ \wf{\sim}{\src{\class}} \Leftrightarrow\ \wfc{\sim}{\src{\class}} $\Coqed
+\end{lemma}
+
+\Thmref{def:rtp} needs to be adapted to account for different target-level trace-models:
+
+\begin{definition}[Robust Compilation with $\sim$]\label{def:rtpsim}
+  $$\rtpsim{\cc{\src{L}}{\trg{L}}}{\src{\class}}{\sim} \isdef
+  \forall \src{\pi} \src{p}.
+    \text{if }\rsat{\src{\progvar}}{\src{\pi}},
+    \text{then }\rsat{\cc{\src{L}}{\trg{L}}\left(\src{p}\right)}{\tilde{\tau}_\sim\left(\src{\pi}\right)}.
+  $$
+\end{definition}
+
+To state a similar theorem as \Thmref{thm:rtp} but for \Cref{def:rtpsim}, it is necessary to define composition of cross-language trace relations:
+\begin{definition}[Composing Cross-Language Trace Relations]
+  Given $\sim_1 : \src{\trace}\to\irl{\trace}\to\text{Prop},\sim_2 : \irl{\trace}\to\trg{\trace}\to\text{Prop}$, then 
+  $$  
+  \sim_1\bullet\sim_2\ \isdef \lambda \src{\trace}\ \trg{\trace}. \exists \irl{\trace}. \src{\trace}\sim_1\irl{\trace} \text{ and } \irl{\trace}\sim_2\trg{\trace}
+  $$
+\end{definition}
+
+\begin{theorem}[Sequential Composition of Secure Compilers]\label{thm:rtpsim}
+  If $\rtpsim{\cc{\src{L}}{\trg{L}}}{\class[_{1}]}{\sim_1}$, $\rtpsim{\cc{\trg{L}}{\obj{L}}}{\tilde{\tau}_{\sim_1}\left(\class[_{2}]\right)}{\sim_2}$, and $\wf{\sim_1}{\class[_2]}$, then $\rtpsim{\cc{\src{L}}{\trg{L}}\circ\cc{\trg{L}}{\obj{L}}}{\class[_{1}]\cap\class[_{2}]}{\sim_1\bullet\sim_2}$.
+\Coqed
+\end{theorem}
+
+Reconsidering the example from the beginning of the section, suppose $\sim_{{\smssafe}}$ is a suitable trace relation such that $\trg{\trace}\sim_{{\smssafe}}\irl{\trace}$ establishes the fact that $\irl{\trace}$ contains an allocation of a table containing bounds information which is never deallocated.
+It is impossible that $\wf{\sim_{{\smssafe}}}{{\tmssafe}}$, since every behavior of compiled $\trg{L}$ programs never contains a deallocation of the bounds-information table.
+Still, supposing $\rtpsim{\cc{\src{L}}{\trg{L}}}{\tmssafe}{{\sim_{\tmssafe}}}$, $\rtpsim{\cc{\trg{L}}{\irl{L}}}{\smssafe}{{\sim_{\smssafe}}}$,
+and $\wf{{\sim_{\tmssafe}}}{{\smssafe}}$ allows to apply \Thmref{thm:rtpsim}.
+However, the resulting class $\tilde{\tau}_{{{\sim_{\tmssafe}}\bullet{\sim_{\smssafe}}}}\left(\tmssafe\cap\smssafe\right)=\emptyset$, which the following lemma shows to hold in general.
+
+\begin{lemma}[Composition is Useless]\label{lem:useless}
+  If $\wf{\sim_1}{\class}$, $\neg\left(\wf{\sim_2}{\tilde{\tau}_{\sim_1}\class}\right)$, then $\tilde{\tau}_{\sim_1\bullet\sim_2}\left(\class\right)=\emptyset$. \Coqed
+\end{lemma}
+
+With this, the previous claim follows as a simple corollary:
+\begin{corollary}
+  $\tilde{\tau}_{{{\sim_{\tmssafe}}\bullet{\sim_{\smssafe}}}}\left(\tmssafe\cap\smssafe\right)=\emptyset$.
+\end{corollary}
+\begin{proof}
+  Note that $\neg\left(\wf{\sim_{{\smssafe}}}{{\tmssafe}}\right)$, since the bounds-table is never deallocated and therefore, for any $\trg{\trace}\in\tmssafe$, $\tau_{\sim_{\tmssafe}}\left(\trg{\trace}\right)\notin\tilde{\tau}_{\sim_{\tmssafe}}\left({\tmssafe}\right)$.
+  However, by assumption, $\wf{\sim_{{\tmssafe}}}{{\tmssafe}}$.
+  Using \Thmref{lem:useless} it follows that $\tilde{\tau}_{{{\sim_{\tmssafe}}\bullet{\sim_{\smssafe}}}}\left(\tmssafe\right)=\emptyset$.
+\end{proof}
 

### 2023-09-27 csf/main.tex
crunch TODOs

@@
         % 
-        Finally, this paper proves that the order of composition of sequential compiler passes is irrelevant for the resulting security.
-        % 
-        This is crucial for reordering optimisation passes and thus generating secure and efficient code.
-      	\MP{
-      		doublecheck.
-      		also, the corollary .. ?
-      	}
+        From this, the paper proves that the order of composition of sequential compiler passes is irrelevant for the resulting security.
+        % 
+        This is crucial for reordering individual optimisation passes and thus generating secure and efficient code.
+        %
+        Finally, this paper extends parts of the compositionality framework to support different trace-models between compilation passes.
 
@@
 % 
-\MP{
+\MPin{
 	later on, say that optimisation passes remove elements from traces, so they should be fine with respect to the result with $\tau$.
 	also, yes, if $\tau$ adds we have a problem, if it removes we don't
+\MKin{ Optimisations that remove elements from traces are not necessarily sound,
+ e.g., remove all deallocations to "optimise" memory management by offloading
+ the work to the OS.
+}
 }
@@
 \subsection{Shared Language Definitions}\label{subsec:cs:defs}
-\MPin{
-	note to fix as agreed upon
-	+ red syntax
-	+ green syntax
-}
-\begin{gather*}
-  \begin{aligned}
-    \mi{(Expressions)}~&\expr\bnfdef\ x \mid \valueexpr \mid \lbinop{\expr[_{1}]}{\expr[_{2}]} \mid \lget{x}{\expr}\\
-                       &\mid \llet{x}{\lnew{\expr[_{1}]}{\expr[_{2}]}}{\expr[_{3}]} \mid \ldelete{x}\\
-                       &\mid \lset{x}{\expr[_{1}]}{\expr[_{2}]} \mid \lpair{\expr[_{1}]}{\expr[_{2}]} \mid \lproja{\expr} \mid \lprojb{\expr}\\
-                       &\mid \llet{x}{\expr[_{1}]}{\expr[_{2}]} \mid \lreturn{\expr} \mid \lcall{g}{\expr}\\
-                       &\mid \lifz{\expr}{\expr[_{1}]}{\expr[_{2}]} \mid \labort
-  \end{aligned}
-  \\
-  \begin{aligned}
-  \mi{(Functions)}~\function\bnfdef&\ \lfunction{\textit{foo}}{x}{e}
-  \end{aligned}
-  \\
-  \begin{aligned}
-  \mi{(Libraries)}~\library\bnfdef&\ \hole{\cdot} \mid \function,\library\hspace{0.5cm}
-  \mi{(Programs)}~\prog{\library_{\ctx}}{\library_{\comp}}
-  \end{aligned}
-\end{gather*}
+
+All presented programming languages share a common fragment. 
+Programs ($\prog{\library_{\ctx}}{\library_{\comp}}$) consist of libraries ($\library$), which are lists of functions ($\function$).
+Functions ($\lfunction{\textit{foo}}{x}{\expr}$) have a single argument $x$ and an expression ($\expr$) as a body which may have $x$ as a free variable.
+Expressions can be variables ($x$), values ($\valueexpr$), binary operations ($\lbinop{\expr[_1]}{\expr[_2]}$), reading from ($\lget{x}{\expr}$) and writing to ($\lset{x}{\expr[_1]}{\expr[_2]}$) memory, memory allocation ($\llet{x}{\lnew{\expr[_1]}{\expr[_2]}}{\expr[_3]}$) as well as deallocation ($\ldelete{x}$), let expressions ($\llet{x}{\expr[_1]}{\expr[_2]}$), pairs ($\lpair{\expr[_1]}{\expr[_2]}$), their first ($\lproja{\expr}$) and second projection ($\lprojb{\expr}$), conditionals ($\lifz{\expr[_1]}{\expr[_2]}{\expr[_3]}$), function calls ($\lcall{\texttt{foo}}{\expr}$), returns ($\lreturn{\expr}$), and an aborting expression ($\labort$).
 
@@
 \begin{theorem}[$\src{L_{\tmssafe}}$-programs are \gls*{tms}]\label{thm:wt:tms}
+  $\;$
+  \begin{nscenter}
   For any $\src{\library_{\comp}}$, $\rsat{\src{\library_{\comp}}}{\tmssafe}$ \Coqed
+  \end{nscenter}
 \end{theorem}
@@
 
-\vspace{-2em}
-\begin{center}
-  \begin{align*}
-    \mi{(Expressions)}~\trg{\expr}\bnfdef& \cdots \mid \trg{\lispoisoned{x}} \mid \trg{\lhast{\expr}{\type}} \\
-  \end{align*}
-\end{center}
-\vspace{-2em}
-
-$\Ltrg$ extends the syntax presented earlier (\Cref{subsec:cs:defs}) with dynamic typechecks $\trg{\lhast{\expr}{\type}}$ and a way to inspect poison tags $\trg{\lispoisoned{x}}$ in the metadata of pointers.
+$\Ltrg$ extends the syntax presented earlier (\Cref{subsec:cs:defs}) with dynamic typechecks ($\trg{\lhast{\expr}{\type}}$) and a way to inspect poison tags ($\trg{\lispoisoned{x}}$) in the metadata of pointers.
 For valid pointers ($\trg{\poisonless}$) bound to variable $\trg{x}$, the check $\trg{\lispoisoned{x}}$ yields $\trg{1}$.
@@
 
-\vspace{-1em}
-\begin{gather*}
-  \begin{aligned}
-    \mi{(Expressions)}~\obj{\expr}\bnfdef&\ \obj{n^{\securitytag}} \mid \cdots \mid \obj{\llet{x^{\securitytag}}{\expr[_{1}]}{\expr[_{2}]}} \mid \cdots\\&\mid \obj{\lwrdoit{D}} \mid \obj{\lrddoit{x}{\expr}}
-  \end{aligned} 
-  \\
-  \begin{aligned}
-    \mi{(States)}~\obj{\statevar}&\bnfdef \obj{\tup{\cfstate}{\sandboxtag,D,\memstate}}
-     &
-    \mi{(DIT\ Flag)}~\obj{D}&\bnfdef \obj{ON} \mid \obj{OFF}
-  \end{aligned}
-\end{gather*}
-
-$\Lscct$ extends $\Lms$ (\Cref{subsec:lms}) with a way to write to a model specific register that controls a data (operand) independent timing mode, a feature that is present in both Arm~\cite[p.~543]{arm-refman} and Intel~\cite[p.~80]{intel-refman} processors.
-To this end, states are extended with the value of the register, which is initially set to be not active.
-If the register is marked active, the intuition is that no secrets can appear on specification traces.
+$\Lscct$ extends $\Lms$ (\Cref{subsec:lms}) with a way to read from ($\obj{\lrddoit{x}{\expr}}$) and to write to ($\obj{\lwrdoit{D}}$) a model specific register that controls a data (operand) independent timing mode, a feature that is present in both Arm~\cite[p.~543]{arm-refman} and Intel~\cite[p.~80]{intel-refman} processors.
+To this end, states are extended with the value of the register ($\obj{D}$), which is initially set to be not active ($\obj{OFF}$).
+If the register is marked active ($\obj{ON}$), the intuition is that no secrets can appear on specification traces.
 If the register is marked inactive, secrets may appear on traces.
@@
 Security tags $\obj{\securitytag}$ are on the usual secrecy lattice, where $\obj{\lock}\le\obj{\securitytag}$ and $\obj{\securitytag}\le\obj{\unlock}$.
-
-\vspace{-1em}
-% \begin{center}
-{\small
-  \begin{align*}
-    \mi{(Pre\text{-}Events)}~\obj{\event[_{b}]}\bnfdef&\ \cdots \mid \obj{\widehat{Get}\ \loc\ \valueexpr} \mid \obj{\widehat{Set}\ \loc\ \valueexpr\ \valueexpr[']}
-    % \\
-                                                      % &
-                                                      \mid \obj{Branch\ n} \mid \obj{Binop\ n}
-  \end{align*}
-% \end{center}
-}
-
 To prevent secrets from leaking but still enable reasoning about memory safety, $\Lscct$ extends pre-events with $\obj{\widehat{Get}\ \loc\ \valueexpr}$ and $\obj{\widehat{Set}\ \loc\ \valueexpr}$.

### 2023-09-28 csf/main.tex
todos

@@
 
-\vspace{-2em}
+\vspace{-1em}
 \begin{gather*}\small
@@
 % 
-\MPin{
-	later on, say that optimisation passes remove elements from traces, so they should be fine with respect to the result with $\tau$.
-	also, yes, if $\tau$ adds we have a problem, if it removes we don't
-\MKin{ Optimisations that remove elements from traces are not necessarily sound,
- e.g., remove all deallocations to "optimise" memory management by offloading
- the work to the OS.
-}
-}
+% \MPin{
+% 	later on, say that optimisation passes remove elements from traces, so they should be fine with respect to the result with $\tau$.
+% 	also, yes, if $\tau$ adds we have a problem, if it removes we don't
+% \MKin{ Optimisations that remove elements from traces are not necessarily sound,
+%  e.g., remove all deallocations to "optimise" memory management by offloading
+%  the work to the OS.
+% }
+% }
 
@@
 $\cc{\trg{L}}{\obj{L}}$ modifies the behavior and because of this, the theorem is inapplicable.
+\MPin{
+	this is still applicable though
+	many use it, and we showcase it too.
+}
 
@@
 To this end, it utilises existing work on secure compilation~\cite{abate2021extacc} that defines Galois connections for several real-world compiler scenarios, such as a compilation pass that exploits undefined behavior.
-This section is concerned about source-level properties, and henceforth uses the existential image of a source-level property $\src{\pi}$ to map $\src{\pi}$ to a target-level property:
+This section is concerned about source-level properties, and henceforth 
+\MPin{
+	different source and target trace models
+	.
+	traces are related
+	.
+	there is a function tau that lefts the relation from individual traces to props/hprops
+}
+\MPin{
+	drop next 3 defs
+}
+
+uses the existential image of a source-level property $\src{\pi}$ to map $\src{\pi}$ to a target-level property:
 
@@
 \end{definition}
+\MPin{
+	explain with text and underline
+}
 \begin{lemma}[Well-Formedness Equivalence for Classes]\label{lem:wf:equiv:wfc}
@@
 To state a similar theorem as \Thmref{thm:rtp} but for \Cref{def:rtpsim}, it is necessary to define composition of cross-language trace relations:
+\MPin{
+	do in text
+}
 \begin{definition}[Composing Cross-Language Trace Relations]
@@
 
+\MPin{
+	explain
+}
 \begin{theorem}[Sequential Composition of Secure Compilers]\label{thm:rtpsim}
@@
 
+\MPin{
+	drop
+}
 With this, the previous claim follows as a simple corollary:
@@
 \end{proof}
+
+\MPin{
+	what about sim2 and C1?
+	-> no relation 
+}
+\MPin{
+	For the sake of simplicity we go back to the singel trace model and go back to the framework of single trace model
+}
 

### 2023-09-28 csf/main.tex
crunch TODOs

@@
 
-\subsection{Applicability of the Compositionality Framework}
+\subsection{Applicability of the Compositionality Framework}\label{subsec:app}
 
 So far, this paper assumes a unified trace-model and secure compilers that do not change the behavior.
-It is worth noting that in this setting, the theorem may not be applicable.
+It is worth noting that, as demonstrated in our case study, quite a few compilation passes do not change the behavior in such a way that the projections to the specification trace model become unequal to each other, i.e., source and target language of the compiler essentially have a different trace model.
+However, there are instances of passes where this is not the case.
 For example, consider $\rtp{\cc{\src{L}}{\trg{L}}}{\tmssafe}$ and $\rtp{\cc{\trg{L}}{\obj{L}}}{\smssafe}$, where $\cc{\trg{L}}{\obj{L}}$ introduces code that allocates a table that stores bounds information of pointers, but fails to insert code that deallocates this table.
@@
 Therefore, no behavior resulting from programs produced by the $\cc{\trg{L}}{\obj{L}}$ compiler is secure with respect to $\tmssafe$, but only for $\smssafe$.
-$\cc{\trg{L}}{\obj{L}}$ modifies the behavior and because of this, the theorem is inapplicable.
-\MPin{
-	this is still applicable though
-	many use it, and we showcase it too.
-}
-
-This section re-introduces secure compiler composition without this limitation. 
-To this end, it utilises existing work on secure compilation~\cite{abate2021extacc} that defines Galois connections for several real-world compiler scenarios, such as a compilation pass that exploits undefined behavior.
-This section is concerned about source-level properties, and henceforth 
-\MPin{
-	different source and target trace models
-	.
-	traces are related
-	.
-	there is a function tau that lefts the relation from individual traces to props/hprops
-}
-\MPin{
-	drop next 3 defs
-}
-
-uses the existential image of a source-level property $\src{\pi}$ to map $\src{\pi}$ to a target-level property:
-
-\begin{definition}[Existential Image of Properties]\label{def:eximg:prop}
-  Given a cross-language trace relation $\sim$,
+$\cc{\trg{L}}{\obj{L}}$ modifies the behavior and this section re-introduces the compositionality framework with support for these situations.
+
+To this end, the section utilises existing work on secure compilation~\cite{abate2021extacc} that defines Galois connections for several real-world compiler scenarios, such as a compilation pass that exploits undefined behavior.
+This section is concerned about a compilation setting with different source and target trace models. 
+To this end, there needs to be a way to say that two traces $\src{\trace}$ and $\trg{\trace}$ are related ($\sim$) and a way to translate a source-level property $\src{\pi}$ into a target-level property ($\tau_\sim\left( \src{\pi} \right)$~\cite{abate2021extacc}) using this trace relation.
+Moreover, the function to map source-level properties to target-level properties can be lifted to translate source-level classes of properties $\src{\class}$ to target-level classes ($\tilde{\tau}_\sim\left(\src{\class}\right)$).
+The section refers to these functions as existential images.
+
+With this, \bul{a cross-language relation $\sim$ is well-formed with respect to a class of source-level properties $\src{\class}$} iff \rul{the existential images preserve set inclusion}.
+\begin{definition}[Well-formedness of Cross-Language Trace Relations for a Class $\src{\class}$]\label{def:wfc:tracerel}
   $$
-  \tau_\sim(\src{\pi}) \isdef \left\{
-    \trg{\trace} \mid \exists\src{\trace}.\src{\trace}\sim\trg{\trace} \text{ and }
-    \src{\trace}\in\src{\pi}
-  \right\}
+  \text{\bul{$\wfc{\sim}{\src{\class}}$}} \isdef \text{\rul{$\forall \src{\Pi}\subseteq\src{\class}, \tau_\sim(\src{\Pi})\subseteq\tilde{\tau}_\sim(\src{\class})$}}
   $$
@@
 
-\Thmref{def:eximg:prop} can be lifted to hyperproperties/classes as follows:
-\begin{definition}[Existential Image of Hyperproperties]\label{def:eximg:hprop}
-  Given a cross-language trace relation $\sim$,
-  $$
-  \tilde{\tau}_\sim(\src{\Pi}) \isdef \left\{
-    \trg{\pi} \mid \exists\src{\pi}.\src{\pi}\in\src{\Pi} \text{ and }
-    \trg{\pi}=\tau_\sim\left(\src{\pi}\right)
-  \right\}
-  $$
-\end{definition}
-
-\bul{A cross-language relation $\sim$ is well-formed with respect to $\src{\Pi}$} iff \rul{the existential images preserve membership}.
-\begin{definition}[Well-formedness of Cross-Language Trace Relations for a Concrete Property $\src{\Pi}$]\label{def:wf:tracerel}
-  $$
-  \text{\bul{$\wf{\sim}{\src{\Pi}}$}} \isdef \text{\rul{$\forall \src{\pi}\in\src{\Pi}, \tau_\sim(\src{\pi})\in\tilde{\tau}_\sim(\src{\Pi})$}}
-  $$
-\end{definition}
-\begin{definition}[Well-formedness of Cross-Language Trace Relations for a Class $\src{\class}$]\label{def:wfc:tracerel}
-  $$
-  \wfc{\sim}{\src{\class}} \isdef \forall \src{\Pi}\subseteq\src{\class}, \tau_\sim(\src{\Pi})\subseteq\tilde{\tau}_\sim(\src{\class})
-  $$
-\end{definition}
-\MPin{
-	explain with text and underline
-}
-\begin{lemma}[Well-Formedness Equivalence for Classes]\label{lem:wf:equiv:wfc}
-  $\;\;$
-  $ \wf{\sim}{\src{\class}} \Leftrightarrow\ \wfc{\sim}{\src{\class}} $\Coqed
-\end{lemma}
-
-\Thmref{def:rtp} needs to be adapted to account for different target-level trace-models:
+\Thmref{def:rtp} needs to be adapted to account for different target-level trace-models~\cite{abate2021extacc}:
 
@@
 
-To state a similar theorem as \Thmref{thm:rtp} but for \Cref{def:rtpsim}, it is necessary to define composition of cross-language trace relations:
-\MPin{
-	do in text
-}
-\begin{definition}[Composing Cross-Language Trace Relations]
-  Given $\sim_1 : \src{\trace}\to\irl{\trace}\to\text{Prop},\sim_2 : \irl{\trace}\to\trg{\trace}\to\text{Prop}$, then 
-  $$  
-  \sim_1\bullet\sim_2\ \isdef \lambda \src{\trace}\ \trg{\trace}. \exists \irl{\trace}. \src{\trace}\sim_1\irl{\trace} \text{ and } \irl{\trace}\sim_2\trg{\trace}
-  $$
-\end{definition}
-
-\MPin{
-	explain
-}
+To state a similar theorem as \Thmref{thm:rtp} but for \Cref{def:rtpsim}, it is necessary to define composition of cross-language trace relations.
+Suppose there are traces $\src{\trace}$ and $\trg{\trace}$ as well as two trace relations $\sim_1$ and $\sim_2$. 
+The traces are related $\src{\trace}\sim_1\bullet\sim_2\trg{\trace}$ if there is another trace $\irl{\trace}$ such that $\src{\trace}\sim_1\irl{\trace}$ and $\irl{\trace}\sim_2\trg{\trace}$.
+
+For the adapted sequential composition theorem, let \bul{$\cc{\src{L}}{\trg{L}}$ robustly preserve the class $\src{\class[_1]}$ under $\sim_1$} and let \rul{$\cc{\trg{L}}{\obj{L}}$ robustly preserve the class $\tilde{\tau}_{\sim_1}\left(\src{\class[_2]}\right)$ under $\sim_2$}.
+Then, \iul{when the cross-language relation $\sim_1$ is well-formed with respect to class $\src{\class[_2]}$}, it follows that \oul{the composed compiler $\cc{\src{L}}{\trg{L}}\circ\cc{\trg{L}}{\obj{L}}$ robustly preserves the intersection of classes $\src{\class[_1]}\cap\src{\class[_2]}$ under $\sim_1\bullet\sim_2$}.
 \begin{theorem}[Sequential Composition of Secure Compilers]\label{thm:rtpsim}
-  If $\rtpsim{\cc{\src{L}}{\trg{L}}}{\class[_{1}]}{\sim_1}$, $\rtpsim{\cc{\trg{L}}{\obj{L}}}{\tilde{\tau}_{\sim_1}\left(\class[_{2}]\right)}{\sim_2}$, and $\wf{\sim_1}{\class[_2]}$, then $\rtpsim{\cc{\src{L}}{\trg{L}}\circ\cc{\trg{L}}{\obj{L}}}{\class[_{1}]\cap\class[_{2}]}{\sim_1\bullet\sim_2}$.
+  If \bul{$\rtpsim{\cc{\src{L}}{\trg{L}}}{\class[_{1}]}{\sim_1}$}, \rul{$\rtpsim{\cc{\trg{L}}{\obj{L}}}{\tilde{\tau}_{\sim_1}\left(\class[_{2}]\right)}{\sim_2}$}, and \iul{$\wf{\sim_1}{\class[_2]}$}, then \oul{$\rtpsim{\cc{\src{L}}{\trg{L}}\circ\cc{\trg{L}}{\obj{L}}}{\class[_{1}]\cap\class[_{2}]}{\sim_1\bullet\sim_2}$}.
 \Coqed
@@
 
-Reconsidering the example from the beginning of the section, suppose $\sim_{{\smssafe}}$ is a suitable trace relation such that $\trg{\trace}\sim_{{\smssafe}}\irl{\trace}$ establishes the fact that $\irl{\trace}$ contains an allocation of a table containing bounds information which is never deallocated.
+For space reasons, \Thmref{thm:urtp} and \Thmref{thm:lrtp} are not presented here, but can be found in the technical appendix ($\CoqSymbol$).
+Note that it is not necessary to require $\sim_2$ to be well-formed with respect to $\class[_1]$.
+
+However, reconsidering the example from the beginning of the section, suppose $\sim_{{\smssafe}}$ is a suitable trace relation such that $\trg{\trace}\sim_{{\smssafe}}\irl{\trace}$ establishes the fact that $\irl{\trace}$ contains an allocation of a table containing bounds information which is never deallocated.
 It is impossible that $\wf{\sim_{{\smssafe}}}{{\tmssafe}}$, since every behavior of compiled $\trg{L}$ programs never contains a deallocation of the bounds-information table.
@@
 
-\MPin{
-	drop
-}
-With this, the previous claim follows as a simple corollary:
-\begin{corollary}
-  $\tilde{\tau}_{{{\sim_{\tmssafe}}\bullet{\sim_{\smssafe}}}}\left(\tmssafe\cap\smssafe\right)=\emptyset$.
-\end{corollary}
-\begin{proof}
-  Note that $\neg\left(\wf{\sim_{{\smssafe}}}{{\tmssafe}}\right)$, since the bounds-table is never deallocated and therefore, for any $\trg{\trace}\in\tmssafe$, $\tau_{\sim_{\tmssafe}}\left(\trg{\trace}\right)\notin\tilde{\tau}_{\sim_{\tmssafe}}\left({\tmssafe}\right)$.
-  However, by assumption, $\wf{\sim_{{\tmssafe}}}{{\tmssafe}}$.
-  Using \Thmref{lem:useless} it follows that $\tilde{\tau}_{{{\sim_{\tmssafe}}\bullet{\sim_{\smssafe}}}}\left(\tmssafe\right)=\emptyset$.
-\end{proof}
-
-\MPin{
-	what about sim2 and C1?
-	-> no relation 
-}
-\MPin{
-	For the sake of simplicity we go back to the singel trace model and go back to the framework of single trace model
-}
+For sake of simplicity, the case-study presented in the following sections does not consider different trace-models: The traces of all programs of the presented languages are projected to a unified trace model.
+The case-study demonstrates that, while the compositionality framework without the consideration for different trace-models may seem overly simplistic, it is useful and simpler in practice, if applicable.
 
@@
 
-\vspace{-1em}
 \begin{gather*}\small
@@
 
+\section{Auxiliary definitions for \Cref{subsec:app}}
+\begin{definition}[Existential Image of Properties]\label{def:eximg:prop}
+  Given a cross-language trace relation $\sim$,
+  $$
+  \tau_\sim(\src{\pi}) \isdef \left\{
+    \trg{\trace} \mid \exists\src{\trace}.\src{\trace}\sim\trg{\trace} \text{ and }
+    \src{\trace}\in\src{\pi}
+  \right\}
+  $$
+\end{definition}
+
+\Thmref{def:eximg:prop} can be lifted to hyperproperties/classes as follows:
+\begin{definition}[Existential Image of Hyperproperties]\label{def:eximg:hprop}
+  Given a cross-language trace relation $\sim$,
+  $$
+  \tilde{\tau}_\sim(\src{\Pi}) \isdef \left\{
+    \trg{\pi} \mid \exists\src{\pi}.\src{\pi}\in\src{\Pi} \text{ and }
+    \trg{\pi}=\tau_\sim\left(\src{\pi}\right)
+  \right\}
+  $$
+\end{definition}
+\bul{A cross-language relation $\sim$ is well-formed with respect to a source-level (hyper-)property $\src{\Pi}$} iff \rul{the existential images preserve membership}.
+\begin{definition}[Well-formedness of Cross-Language Trace Relations for a Properties $\src{\Pi}$]\label{def:wf:tracerel}
+  $$
+  \text{\bul{$\wfc{\sim}{\src{\Pi}}$}} \isdef \text{\rul{$\forall \src{\pi}\in\src{\Pi}, \tau_\sim(\src{\pi})\in\tilde{\tau}_\sim(\src{\Pi})$}}
+  $$
+\end{definition}
+\begin{lemma}[Well-Formedness Equivalence for Classes]\label{lem:wf:equiv:wfc}
+  $\;\;$
+  $ \wf{\sim}{\src{\class}} \Leftrightarrow\ \wfc{\sim}{\src{\class}} $\Coqed
+\end{lemma}
+
 \newpage

### 2023-09-28 csf/main.tex
notation and typo fix

@@
 \begin{theorem}[Sequential Composition of Secure Compilers]\label{thm:rtpsim}
-  If \bul{$\rtpsim{\cc{\src{L}}{\trg{L}}}{\class[_{1}]}{\sim_1}$}, \rul{$\rtpsim{\cc{\trg{L}}{\obj{L}}}{\tilde{\tau}_{\sim_1}\left(\class[_{2}]\right)}{\sim_2}$}, and \iul{$\wf{\sim_1}{\class[_2]}$}, then \oul{$\rtpsim{\cc{\src{L}}{\trg{L}}\circ\cc{\trg{L}}{\obj{L}}}{\class[_{1}]\cap\class[_{2}]}{\sim_1\bullet\sim_2}$}.
+  If \bul{$\rtpsim{\cc{\src{L}}{\trg{L}}}{\src{\class[_{1}]}}{\sim_1}$}, \rul{$\rtpsim{\cc{\trg{L}}{\obj{L}}}{\tilde{\tau}_{\sim_1}\left(\src{\class[_{2}]}\right)}{\sim_2}$}, and \iul{$\wf{\sim_1}{\src{\class[_2]}}$}, then \oul{$\rtpsim{\cc{\src{L}}{\trg{L}}\circ\cc{\trg{L}}{\obj{L}}}{\src{\class[_{1}]}\cap\src{\class[_{2}]}}{\sim_1\bullet\sim_2}$}.
 \Coqed
@@
 For space reasons, \Thmref{thm:urtp} and \Thmref{thm:lrtp} are not presented here, but can be found in the technical appendix ($\CoqSymbol$).
-Note that it is not necessary to require $\sim_2$ to be well-formed with respect to $\class[_1]$.
+Note that it is not necessary to require $\sim_2$ to be well-formed with respect to $\src{\class[_1]}$.
 

### 2023-09-29 csf/main.tex
push

--- 2023-09-28 csf/main.tex
@@
 
-So far, this paper assumes a unified trace-model and secure compilers that do not change the behavior.
-It is worth noting that, as demonstrated in our case study, quite a few compilation passes do not change the behavior in such a way that the projections to the specification trace model become unequal to each other, i.e., source and target language of the compiler essentially have a different trace model.
+So far, this paper assumes a unified trace-model between source and target languages.
+Moreover secure compilers (in the sense of \Cref{def:rtp}) cannot change the behavior of compiled programs.
+It is worth noting that, as our case study in \Cref{sec:casestud:defs} demonstrates, several compilation passes do not change the behavior of compiled programs
+% in such a way that the projections to the specification trace model become unequal to each other, 
+% i.e., source and target language of the compiler essentially have a different trace model. 
 However, there are instances of passes where this is not the case.
@@
 
-To this end, the section utilises existing work on secure compilation~\cite{abate2021extacc} that defines Galois connections for several real-world compiler scenarios, such as a compilation pass that exploits undefined behavior.
-This section is concerned about a compilation setting with different source and target trace models. 
-To this end, there needs to be a way to say that two traces $\src{\trace}$ and $\trg{\trace}$ are related ($\sim$) and a way to translate a source-level property $\src{\pi}$ into a target-level property ($\tau_\sim\left( \src{\pi} \right)$~\cite{abate2021extacc}) using this trace relation.
-Moreover, the function to map source-level properties to target-level properties can be lifted to translate source-level classes of properties $\src{\class}$ to target-level classes ($\tilde{\tau}_\sim\left(\src{\class}\right)$).
+To reason about these kinds of compilers, we rely on existing work~\cite{abate2021extacc} that lifts the presented secure compilation theory to compilers whose source and target trace models are different.
+% These compilers require a Galois connections for several real-world compiler scenarios, such as a compilation pass that exploits undefined behavior.
+% This section is concerned about a compilation setting with different source and target trace models. 
+These compilers require a way to say that two traces $\src{\trace}$ and $\trg{\trace}$ are related ($\sim$) and a function to translate a source-level property $\src{\pi}$ into a target-level property ($\tau_\sim\left( \src{\pi} \right)$~\cite{abate2021extacc}) using this trace relation.
+The function to map source-level properties to target-level properties can be lifted to translate source-level classes of properties $\src{\class}$ to target-level classes ($\tilde{\tau}_\sim\left(\src{\class}\right)$).
 The section refers to these functions as existential images.
@@
 
-\Thmref{def:rtp} needs to be adapted to account for different target-level trace-models~\cite{abate2021extacc}:
+\Cref{def:rtpsim} shows how to adapt \Thmref{def:rtp} to account for different target-level trace-models~\cite{abate2021extacc}:
 
@@
 For the adapted sequential composition theorem, let \bul{$\cc{\src{L}}{\trg{L}}$ robustly preserve the class $\src{\class[_1]}$ under $\sim_1$} and let \rul{$\cc{\trg{L}}{\obj{L}}$ robustly preserve the class $\tilde{\tau}_{\sim_1}\left(\src{\class[_2]}\right)$ under $\sim_2$}.
-Then, \iul{when the cross-language relation $\sim_1$ is well-formed with respect to class $\src{\class[_2]}$}, it follows that \oul{the composed compiler $\cc{\src{L}}{\trg{L}}\circ\cc{\trg{L}}{\obj{L}}$ robustly preserves the intersection of classes $\src{\class[_1]}\cap\src{\class[_2]}$ under $\sim_1\bullet\sim_2$}.
+Then, \iul{when the cross-language relation $\sim_1$ is well-formed with respect to class $\src{\class[_2]}$}, and \pul{relation $\sim_2$ is well-formed with respect to $\src{\class[_1]}$}, it follows that \oul{the composed compiler $\cc{\src{L}}{\trg{L}}\circ\cc{\trg{L}}{\obj{L}}$ robustly preserves the intersection of classes $\src{\class[_1]}\cap\src{\class[_2]}$ under $\sim_1\bullet\sim_2$}.
+% 
+For space reasons, \Thmref{thm:urtp} and \Thmref{thm:lrtp} are not presented here, but can be found in the technical appendix ($\CoqSymbol$).
+% Note that it is not necessary to require $\sim_2$ to be well-formed with respect to $\src{\class[_1]}$, since the 
+% \MPin{why}
 \begin{theorem}[Sequential Composition of Secure Compilers]\label{thm:rtpsim}
-  If \bul{$\rtpsim{\cc{\src{L}}{\trg{L}}}{\class[_{1}]}{\sim_1}$}, \rul{$\rtpsim{\cc{\trg{L}}{\obj{L}}}{\tilde{\tau}_{\sim_1}\left(\class[_{2}]\right)}{\sim_2}$}, and \iul{$\wf{\sim_1}{\class[_2]}$}, then \oul{$\rtpsim{\cc{\src{L}}{\trg{L}}\circ\cc{\trg{L}}{\obj{L}}}{\class[_{1}]\cap\class[_{2}]}{\sim_1\bullet\sim_2}$}.
+  If \bul{$\rtpsim{\cc{\src{L}}{\trg{L}}}{\src{\class[_{1}]}}{\sim_1}$}, \rul{$\rtpsim{\cc{\trg{L}}{\obj{L}}}{\tilde{\tau}_{\sim_1}\left(\src{\class[_{2}]}\right)}{\sim_2}$}, and \iul{$\wf{\sim_1}{\src{\class[_2]}}$}, and \pul{$\wf{\sim_2}{\src{\class[_1]}}$}, then \oul{$\rtpsim{\cc{\src{L}}{\trg{L}}\circ\cc{\trg{L}}{\obj{L}}}{\src{\class[_{1}]}\cap\src{\class[_{2}]}}{\sim_1\bullet\sim_2}$}.
 \Coqed
@@
 
-For space reasons, \Thmref{thm:urtp} and \Thmref{thm:lrtp} are not presented here, but can be found in the technical appendix ($\CoqSymbol$).
-Note that it is not necessary to require $\sim_2$ to be well-formed with respect to $\class[_1]$.
-
-However, reconsidering the example from the beginning of the section, suppose $\sim_{{\smssafe}}$ is a suitable trace relation such that $\trg{\trace}\sim_{{\smssafe}}\irl{\trace}$ establishes the fact that $\irl{\trace}$ contains an allocation of a table containing bounds information which is never deallocated.
+The reason we need assumptions about both relations is motivated by \Cref{lem:useless}.
+% 
+In fact, if \bul{that a relation $\sim_1$ is well-formed for a property $\class$} but \rul{another relation $\sim_2$ is not well-formed for the property mapped with the first relation $\tilde{\tau}_{\sim_1}\class$} then \iul{their composition is empty}.
+% 
+An empty intersection means that there is no correct behaviour to preserve for the compiler, which nullifies its security relevance.
+\begin{lemma}[Composition is Useless]\label{lem:useless}
+  If \bul{$\wf{\sim_1}{\class}$}, \rul{$\neg\left(\wf{\sim_2}{\tilde{\tau}_{\sim_1}\class}\right)$}, then \iul{$\tilde{\tau}_{\sim_1\bullet\sim_2}\left(\class\right)=\emptyset$}. \Coqed
+\end{lemma}
+
+Let us now reconsider the example from the beginning of the section, suppose $\sim_{{\smssafe}}$ is a suitable trace relation such that $\trg{\trace}\sim_{{\smssafe}}\irl{\trace}$ establishes the fact that $\irl{\trace}$ contains an allocation of a table containing bounds information which is never deallocated.
 It is impossible that $\wf{\sim_{{\smssafe}}}{{\tmssafe}}$, since every behavior of compiled $\trg{L}$ programs never contains a deallocation of the bounds-information table.
@@
 and $\wf{{\sim_{\tmssafe}}}{{\smssafe}}$ allows to apply \Thmref{thm:rtpsim}.
-However, the resulting class $\tilde{\tau}_{{{\sim_{\tmssafe}}\bullet{\sim_{\smssafe}}}}\left(\tmssafe\cap\smssafe\right)=\emptyset$, which the following lemma shows to hold in general.
-
-\begin{lemma}[Composition is Useless]\label{lem:useless}
-  If $\wf{\sim_1}{\class}$, $\neg\left(\wf{\sim_2}{\tilde{\tau}_{\sim_1}\class}\right)$, then $\tilde{\tau}_{\sim_1\bullet\sim_2}\left(\class\right)=\emptyset$. \Coqed
-\end{lemma}
+However, the resulting class $\tilde{\tau}_{{{\sim_{\tmssafe}}\bullet{\sim_{\smssafe}}}}\left(\tmssafe\cap\smssafe\right)=\emptyset$.
+
+\smallskip
 
 For sake of simplicity, the case-study presented in the following sections does not consider different trace-models: The traces of all programs of the presented languages are projected to a unified trace model.
-The case-study demonstrates that, while the compositionality framework without the consideration for different trace-models may seem overly simplistic, it is useful and simpler in practice, if applicable.
+The case-study demonstrates that, the compositionality framework without the different trace-models is already applicable to a variety of security-relevant passes as well as existing optimisations.
+% may seem overly simplistic, it is useful and simpler in practice, if applicable.
 

### 2023-09-29 csf/main.tex
Minor typo fix

@@
 \begin{theorem}[Upper Composition of Secure Compilers]\label{thm:urtp}
-  If \bul{$\rtp{\cc{\src{L}}{\trg{L}}}{\class[_{1}]}$} and \rul{$\rtp{\cc{\obj{L}}{\trg{L}}}{\class[_{2}]}$}, then \oul{$\rtp{\cc{\src{L}+\obj{L}}{\trg{L}}}{\class[_{1}]\cap\class[_{2}]}$}. \Coqed
+  If \bul{$\rtp{\cc{\src{L}}{\trg{L}}}{\class[_{1}]}$} and \rul{$\rtp{\cc{\obj{L}}{\trg{L}}}{\class[_{2}]}$}, then \oul{$\rtp{\cc{\src{L}+\obj{L}}{\trg{L}}}{\class[_{1}]\cap\class[_{2}]}$}. %\Coqed
 \end{theorem}
@@
 \begin{theorem}[Lower Composition of Secure Compilers]\label{thm:lrtp}
-  If \bul{$\rtp{\cc{\src{L}}{\trg{L}}}{\class[_{1}]}$} and \rul{$\rtp{\cc{\obj{L}}{\trg{L}}}{\class[_{2}]}$}, then \oul{$\rtp{\cc{\src{L}}{\trg{L}+\obj{L}}}{\class[_{1}]\cap\class[_{2}]}$}. \Coqed
+  If \bul{$\rtp{\cc{\src{L}}{\trg{L}}}{\class[_{1}]}$} and \rul{$\rtp{\cc{\obj{L}}{\trg{L}}}{\class[_{2}]}$}, then \oul{$\rtp{\cc{\src{L}}{\trg{L}+\obj{L}}}{\class[_{1}]\cap\class[_{2}]}$}. % \Coqed
 \end{theorem}
@@
 Moreover secure compilers (in the sense of \Cref{def:rtp}) cannot change the behavior of compiled programs.
-It is worth noting that, as our case study in \Cref{sec:casestud:defs} demonstrates, several compilation passes do not change the behavior of compiled programs
+It is worth noting that, as our case study in \Cref{sec:casestud:defs} demonstrates, several compilation passes do not change the behavior of compiled programs.
 % in such a way that the projections to the specification trace model become unequal to each other, 
@@
 These compilers require a way to say that two traces $\src{\trace}$ and $\trg{\trace}$ are related ($\sim$) and a function to translate a source-level property $\src{\pi}$ into a target-level property ($\tau_\sim\left( \src{\pi} \right)$~\cite{abate2021extacc}) using this trace relation.
-The function to map source-level properties to target-level properties can be lifted to translate source-level classes of properties $\src{\class}$ to target-level classes ($\tilde{\tau}_\sim\left(\src{\class}\right)$).
+The function to map source-level properties to target-level properties can be lifted to translate source-level classes of properties ($\src{\class}$) to target-level classes ($\tilde{\tau}_\sim\left(\src{\class}\right)$).
 The section refers to these functions as existential images.
@@
 % 
-For space reasons, \Thmref{thm:urtp} and \Thmref{thm:lrtp} are not presented here, but can be found in the technical appendix ($\CoqSymbol$).
+%For space reasons, \Thmref{thm:urtp} and \Thmref{thm:lrtp} are not presented here, but can be found in the technical appendix ($\CoqSymbol$).
 % Note that it is not necessary to require $\sim_2$ to be well-formed with respect to $\src{\class[_1]}$, since the 

### 2023-10-09 esop/main.tex
esop

--- 2023-09-29 csf/main.tex
@@
 \newcommand\bmmax{0}
-\documentclass[dvipsnames,conference]{IEEEtran}
+\documentclass[dvipsnames]{llncs}
+
+\usepackage[T1]{fontenc}
 
@@
 \usepackage[inference]{semantic}
-\usepackage[switch]{lineno}
+\usepackage{lineno}
 \usepackage{fontawesome5}
 \usepackage{listofitems}
+\usepackage{breakcites}
 \usepackage{glossaries}
-\usepackage{orcidlink}
+%\usepackage{hyperref}
 \usepackage{cleveref}
@@
 \usepackage{amsmath}
-\usepackage{amsthm}
+\usepackage{nameref}
+%\usepackage{amsthm}
 \usepackage{xspace}
@@
 
-\hypersetup{
-    colorlinks,
-    linkcolor={red!50!black},
-    citecolor={blue!50!black},
-    urlcolor={blue!80!black}
-}
+%\renewcommand\UrlFont{\color{blue}\rmfamily}
+\newcommand{\url}[1]{\lstinline{#1}}
 
@@
 
-% allow page breaks inside multiline alignment displays
-\allowdisplaybreaks
-
 \Crefname{exampleenv}{Example}{Examples}
 
-\theoremstyle{definition}
 \newtheorem{exampleenv}{Example}[section]
-\newtheorem{lemma}{Lemma}[section]
-\newtheorem{theorem}{Theorem}[section]
-\newtheorem{corollary}{Corollary}[section]
-\newtheorem{definition}{Definition}[section]
+%\newtheorem{lemma}{Lemma}[section]
+%\newtheorem{theorem}{Theorem}[section]
+%\newtheorem{corollary}{Corollary}[section]
+%\newtheorem{definition}{Definition}[section]
 
@@
 \makeglossaries
-
-\AtBeginDocument{%
-  \providecommand\BibTeX{{%
-    \normalfont B\kern-0.5em{\scshape i\kern-0.25em b}\kern-0.8em\TeX}}}
-\makeatletter
-\newcommand{\linebreakand}{%
-  \end{@IEEEauthorhalign}
-  \hfill\mbox{}\par
-  \mbox{}\hfill\begin{@IEEEauthorhalign}
-}
-\makeatother
 
@@
 % \author{
-% \IEEEauthorblockN{1\textsuperscript{st} Matthis Kruse\orcidlink{0000-0003-4062-9666}}
-% \IEEEauthorblockA{\textit{CISPA and Saarland University} \\
-% Saarbr\"ucken, Germany \\
-% matthis.kruse@cispa.de}
+% Matthis Kruse\orcidID{0000-0003-4062-9666}
 % \and
-% \IEEEauthorblockN{2\textsuperscript{nd} Michael Backes}
-% \IEEEauthorblockA{\textit{CISPA} \\
-% Saarbr\"ucken, Germany \\
-% director@cispa.de}
+% Michael Backes
 % \and
-% \IEEEauthorblockN{3\textsuperscript{rd} Marco Patrignani\orcidlink{0000-0003-3411-9678}}
-% \IEEEauthorblockA{\textit{Trento University} \\
-% Trento, Italy \\
-% marco.patrignani@unitn.it}%
+% Marco Patrignani
 % }
-\author{
-
-}
+% \authorrunning{M. Kruse et al.}
+
+% \institute{
+% CISPA Helmholtz Center for Information Security and Saarland University
+% \and
+% CISPA Helmholtz Center for Information Security
+% \and
+% Trento University
+% }
+
+% \email{matthis.kruse@cispa.de \and director@cispa.de \and marco.patrignani@unitn.it}
 
@@
 \end{center}
+\keywords{memory-safety, secure compilation, formal methods}
 \end{abstract}
-
-% \begin{IEEEkeywords}
-%   memory-safety, cryptographic constant-time, secure compilation, robustness
-% \end{IEEEkeywords}
 
@@
  % so that they are no longer vulnerable to timing attacks.
-
+%
 Alas, code is not run in isolation, so a malicious attacker could supply code that interacts with \texttt{strncpy} and trigger a violation of either \gls*{ms} or \gls*{cct} by calling \texttt{strncpy} with an argument for \texttt{src} that points to uninitialised memory.
@@
 This result showcases how the framework allows the kind of formal security reasoning that compiler writers already want (and already do), obtaining precise, compositional security reasoning while providing minimal (and modular) proof effort.
-
+%
 In summary, this paper makes the following contributions:
@@
 \begin{definition}[Property Satisfaction]\label{def:propsat}
-  $\;$\\
-  \vspace{-1em}
-  \begin{nscenter}
     \bul{$\sat{\progvar}{\pi}$}
@@
     then \oul{$\trace\in\pi$}.
-  \end{nscenter}
 \end{definition}
@@
 \begin{definition}[Robust Satisfaction]\label{def:proprsat}
-  $\;$\\
-  \vspace{-1em}
   \begin{nscenter}
@@
 
-\vspace{-1em}
-\begin{gather*}\small
-  \begin{aligned}
-  \mi{(Security\ Tag)}~\sigma\bnfdef&\ \lock \mid \unlock
-  \qquad
-  \mi{(Control\ Tag)}~\sandboxtag\bnfdef \ctx\mid\comp\\
-  \mi{(Event)}~\event\bnfdef&\ \emptyevent \mid \lightning \mid \event[_{b}];\sandboxtag;\sigma \\
-  \mi{(Pre\text{-}event)}~\event[_{b}]\bnfdef&\ \ev{Alloc\ \loc\ n} \mid \ev{Dealloc\ \loc} 
-  \\
-                                             & 
-                                             \mid \ev{Use\ \loc\ n} \mid \ev{Branch\ n} \mid \ev{Binop\ n} \\
-  \end{aligned}
-\end{gather*}
+% \begin{gather*}\small
+%   \begin{aligned}
+%   \mi{(Security\ Tag)}~\sigma\bnfdef&\ \lock \mid \unlock
+%   \qquad
+%   \mi{(Control\ Tag)}~\sandboxtag\bnfdef \ctx\mid\comp\\
+%   \mi{(Event)}~\event\bnfdef&\ \emptyevent \mid \lightning \mid \event[_{b}];\sandboxtag;\sigma \\
+%   \mi{(Pre\text{-}event)}~\event[_{b}]\bnfdef&\ \ev{Alloc\ \loc\ n} \mid \ev{Dealloc\ \loc} 
+%   \\
+%                                              & 
+%                                              \mid \ev{Use\ \loc\ n} \mid \ev{Branch\ n} \mid \ev{Binop\ n} \\
+%   \end{aligned}
+% \end{gather*}
 The specification trace model defines events as either the empty event ($\emptyevent$), a crash ($\lightning$), or as tuples consisting of a pre-event, a control-tag, and a security-tag.
@@
 \small
-  $$
+  $
   \smssafe:=\left\{\trace \left|\begin{array}{rcl}
@@
                                 \end{array}\right.\right\}
-  $$
+  $
 \end{definition}
@@
 \begin{definition}[\glsfirst*{scct}]\label{def:trace:scctdef}
-  $$
+  $
   \scctsafe:=\left\{\trace \left|\begin{array}{rcl}
@@
                                  \end{array}\right.\right\}
-  $$
+  $
 \end{definition}
@@
 \begin{definition}[Monitor Satisfaction]\label{def:monsat}
-  $\;$ 
-  \begin{nscenter}
   \bul{$\monsafe{\trace}{\pi}$} $\isdef$ $\exists \abstrace\ \monitor$, \iul{$\traceagree{\pi}{\trace}{\abstrace}$} and \oul{$\monitorcheck{\emptyset}{\monitor}{\abstrace}$}.
-  \end{nscenter}
 \end{definition}
@@
 \begin{center}\small
-  \judgbox{\monitorcheck[]{\tmsmonitor}{\tmsmonitor[']}{\absevent}}{,,Monitor $\tmsmonitor$ steps to $\tmsmonitor[']$\\given event $\absevent$.''}$\;$\\
+  \judgbox{\monitorcheck[]{\tmsmonitor}{\tmsmonitor[']}{\absevent}}{,,Monitor $\tmsmonitor$ steps to $\tmsmonitor[']$ given event $\absevent$.''}$\;$\\
   %
@@
 \begin{center}
-  \judgbox{\monitorcheck{\tmsmonitor}{\tmsmonitor[']}{\abstrace}}{,,Monitor $\tmsmonitor$ multi-steps to $\tmsmonitor[']$\\given trace $\abstrace$.''}$\;$\\
+  \judgbox{\monitorcheck{\tmsmonitor}{\tmsmonitor[']}{\abstrace}}{,,Monitor $\tmsmonitor$ multi-steps to $\tmsmonitor[']$ given trace $\abstrace$.''}$\;$\\
   %
@@
 
-\begin{lemma}[Traces with Monitor Satisfaction are $\tmssafe$]\label{lem:mon:tmssafe}
+\begin{lemma}[Monitor Satisfaction is $\tmssafe$]\label{lem:mon:tmssafe}
   $\;\;$ If $\montmssafe{\trace}$, then $\trace\in\tmssafe$. %\Coqed
@@
   Even though the trace contains an out-of-bounds access right before returning from \texttt{strncpy}, this is no concern for \gls*{tms}, since location $\loc[_{x}]$ is still allocated.
-  However, having returned from \texttt{strncpy}, the \texttt{greet} function continues and deallocates $\loc[_{y}]$ whose subsequent use in \texttt{printf} is a use-after-free bug.
+  Having returned from \texttt{strncpy}, the \texttt{greet} function continues and deallocates $\loc[_{y}]$ whose subsequent use in \texttt{printf} is a use-after-free bug.
 
@@
 \begin{center}
-  \judgbox{\monitorcheck[]{\smsmonitor}{\smsmonitor[']}{\absevent}}{,,Monitor $\smsmonitor$ steps to $\smsmonitor[']$\\given event $\absevent$.''}$\;$\\
+  \judgbox{\monitorcheck[]{\smsmonitor}{\smsmonitor[']}{\absevent}}{,,Monitor $\smsmonitor$ steps to $\smsmonitor[']$ given event $\absevent$.''}$\;$\\
   %
@@
 
-\begin{lemma}[Traces with Monitor Satisfaction are $\smssafe$]\label{lem:mon:smsafe}
+\begin{lemma}[Monitor Satisfaction is $\smssafe$]\label{lem:mon:smsafe}
   $\;\;$
@@
 
-\begin{lemma}[Traces with Monitor Satisfaction are $\mssafe$]\label{lem:mon:msafe}
+\begin{lemma}[Monitor Satisfaction is $\mssafe$]\label{lem:mon:msafe}
   If $\monmssafe{\trace}$, then $\trace\in\mssafe$.\Coqed
@@
     % &
-    \\
+    &
     \mi{(Abstract\ Events)}~\absevent:=&\ \absev{\varepsilon} \mid \absterm \mid \absev{Leak} 
@@
 
-\begin{lemma}[Monitor Traces are $\scctsafe$]\label{lem:mon:scctsafe}
+\begin{lemma}[Monitor Satisfaction is $\scctsafe$]\label{lem:mon:scctsafe}
   If $\monscctsafe{\trace}$, then $\trace\in\scctsafe$. %\Coqed
@@
 \begin{exampleenv}[Data-independent timing mode]\label{ex:ditm}
-  Consider the call \texttt{strncpy(1,x,y)} to the \texttt{strncpy} function (\Cref{ex:strncpy}) with low ($\unlock$) security for \texttt{x} and high security ($\lock$) for \texttt{y}.
+  Consider the function call \texttt{strncpy(1,x,y)} to the \texttt{strncpy} function (\Cref{ex:strncpy}) with low ($\unlock$) security for \texttt{x} and high security ($\lock$) for \texttt{y}.
   The trace for the copying part of \texttt{strncpy} is:
@@
 
-\begin{lemma}[Traces with Monitor Satisfaction are $\msscctsafe$]\label{lem:mon:msscctsafe}
+\begin{lemma}[Monitor Satisfaction is $\msscctsafe$]\label{lem:mon:msscctsafe}
+  $\;$\\\vspace{-1em}
+  \begin{nscenter}
   If $\monmsscctsafe{\trace}$, then $\trace\in\msscctsafe$. %\Coqed
+  \end{nscenter}
 \end{lemma}
@@
 With this, \bul{a cross-language relation $\sim$ is well-formed with respect to a class of source-level properties $\src{\class}$} iff \rul{the existential images preserve set inclusion}.
-\begin{definition}[Well-formedness of Cross-Language Trace Relations for a Class $\src{\class}$]\label{def:wfc:tracerel}
+\begin{definition}[Well-formedness of Trace Relations for a Class $\src{\class}$]\label{def:wfc:tracerel}
   $$
@@
 \begin{theorem}[$\src{L_{\tmssafe}}$-programs are \gls*{tms}]\label{thm:wt:tms}
-  $\;$
-  \begin{nscenter}
   For any $\src{\library_{\comp}}$, $\rsat{\src{\library_{\comp}}}{\tmssafe}$ \Coqed
-  \end{nscenter}
 \end{theorem}
@@
 If the array bound to $\trg{x}$ was deallocated, i.e., has been poisoned ($\trg{\poisoned}$), the check $\trg{\lispoisoned{x}}$ evaluates to $\trg{0}$.
-
 Dynamic typechecks $\trg{\lhast{\expr}{\type}}$ match on $\trg{\expr}$ and evaluate to $\trg{0}$ if the term is of type $\trg{\type}$ and $\trg{1}$ otherwise.
-
-%
 The projection $\msfilterL{\trg{\trace}}$ is equal to $\msfilterLtms{\src{\trace}}$.
@@
 
-\begin{figure*}[!h]
-% \vspace{-2em}
+\begin{figure}[!h]
+ \vspace{-2em}
   $$
+  \scriptsize
   % {
-  \small
   \begin{array}{ccccccc}
@@
   \label{fig:ex-cct}
-\end{figure*}
+\end{figure}
 
@@
 This section defines several secure compilers, each of which robustly preserves a different property of interest as depicted in \Cref{fig:pipeline}.
-\begin{figure*}[!h]
+\begin{figure}[!h]
   \centering
-  \begin{minipage}[c]{0.64\textwidth}
   \begin{tikzpicture}
@@
   \end{tikzpicture}
-  \end{minipage}
-  \begin{minipage}{0.33\textwidth}
   % \vspace{-2.5em}
@@
   }\label{fig:pipeline}
-  \end{minipage}
-\end{figure*}
+\end{figure}
 The section demonstrates the power of the framework (\Cref{sec:compprop,sec:compcomp}) by composing these compilers for a secure and optimising compilation chain that robustly preserves \gls*{msscct}.
@@
 
-\begin{theorem}[Compiler $\cca$ is secure with respect to \gls*{tms}]\label{thm:cca:rtp:tms}
+\begin{theorem}[$\cca$ is secure with respect to \gls*{tms}]\label{thm:cca:rtp:tms}
   $\rtp{\cca}{\tmssafe}$ % \Coqed
@@
 
-\begin{theorem}[Compiler $\ccb$ is secure with respect to \gls*{sms}]\label{thm:ccb:rtp:sms}
+\begin{theorem}[$\ccb$ is secure with respect to \gls*{sms}]\label{thm:ccb:rtp:sms}
   $\rtp{\ccb}{\smssafe}$ % \Coqed
@@
 
-\begin{theorem}[Compiler $\cca\circ\ccb$ is secure with respect to \gls*{ms}]\label{thm:ccab:rtp:ms}
+\begin{theorem}[$\cca\circ\ccb$ is secure with respect to \gls*{ms}]\label{thm:ccab:rtp:ms}
   $\rtp{\cca\circ\ccb}{\mssafe}$ % \Coqed
@@
 % \begin{align*}
-  \partialeval{\irl{x}}{\irl{\substlist}} & = \irl{n} 
-  	\qquad\qquad \text{if } \irl{\subst{n}{x}}\in\irl{\substlist} \\
-  \partialeval{\irl{x}}{\irl{\substlist}} & = \irl{x} 
-  \qquad\qquad \text{if } \irl{\subst{n}{x}}\notin\irl{\substlist} \\
-  \partialeval{\irl{\lbinop{n}{m}}}{\irl{\substlist}} & = \irl{k} 
-  \qquad\qquad \text{if } \lbinop{\irl{n}}{\irl{m}}=k \\
+%  \partialeval{\irl{x}}{\irl{\substlist}} & = \irl{n} 
+%  	\qquad\qquad \text{if } \irl{\subst{n}{x}}\in\irl{\substlist} \\
+%  \partialeval{\irl{x}}{\irl{\substlist}} & = \irl{x} 
+%  \qquad\qquad \text{if } \irl{\subst{n}{x}}\notin\irl{\substlist} \\
+%  \partialeval{\irl{\lbinop{n}{m}}}{\irl{\substlist}} & = \irl{k} 
+%  \qquad\qquad \text{if } \lbinop{\irl{n}}{\irl{m}}=k \\
    % \partialeval{\irl{\lbinop{\expr[_{1}]}{\expr[_{2}]}}}{\irl{\substlist}} & = \lbinop[\irl]{\partialeval{\irl{\expr[_{1}]}}{\irl{\substlist}}}{\partialeval{\irl{\expr[_{2}]}}{\irl{\substlist}}} \\
-  \partialeval{\irl{\llet{x}{n}{\expr}}}{\irl{\substlist}} & = \partialeval{\irl{\expr}}{\irl{\subst{x}{n}\cdot\substlist}} 
+%  \partialeval{\irl{\llet{x}{\valueexpr}{\expr}}}{\irl{\substlist}} & = \partialeval{\irl{\expr}}{\irl{\subst{x}{\valueexpr}\cdot\substlist}} 
   % \\
@@
 
-\begin{theorem}[Compiler $\ccdce$ is secure with respect to \gls*{ms}]\label{thm:ccdce:rtp:ms}
+\begin{theorem}[$\ccdce$ is secure with respect to \gls*{ms}]\label{thm:ccdce:rtp:ms}
   $\rtp{\ccdce}{\mssafe}$ %\Coqed
 \end{theorem}
-\begin{theorem}[Compiler $\cccf$ is secure with respect to \gls*{ms}]\label{thm:cccf:rtp:ms}
+\begin{theorem}[$\cccf$ is secure with respect to \gls*{ms}]\label{thm:cccf:rtp:ms}
   $\rtp{\cccf}{\mssafe}$ %\Coqe
@@
 
-\begin{theorem}[Compilers $\cccf\circ\ccdce$ and $\cccf\circ\ccdce$ are secure with respect to \gls*{ms}]\label{thm:cccfccdce:rtp:ms}
+\begin{theorem}[$\cccf\circ\ccdce$ and $\cccf\circ\ccdce$ are secure with respect to \gls*{ms}]\label{thm:cccfccdce:rtp:ms}
   $\rtp{\cccf\circ\ccdce}{\mssafe}$ and $\rtp{\ccdce\circ\cccf}{\mssafe}$. % \Coqed
@@
 
-\begin{theorem}[Compiler $\ccscct$ is secure with respect to \gls*{scct}]\label{thm:ccscct:rtp:scct}
+\begin{theorem}[$\ccscct$ is secure with respect to \gls*{scct}]\label{thm:ccscct:rtp:scct}
   $\rtp{\ccscct}{\scctsafe}$ % \Coqed
@@
 
-\begin{theorem}[Compiler $\ccmsscct$ is secure with respect to \gls*{scct}]\label{thm:ccall:rtp:msscct}
+\begin{theorem}[$\ccmsscct$ is secure with respect to \gls*{scct}]\label{thm:ccall:rtp:msscct}
   $\rtp{\cc{\Ltms}{\Lscct}}{\mssafe\cap\scctsafe}$ % \Coqed
@@
 
-\newpage
 % \section*{Acknowledgements}\label{sec:acks}
 
-\appendices
-
-\section*{Appendix}
-
-The appendix contains insights on the structure of proofs for the theorems of \Cref{sec:casestud:rtp} (\Cref{sec:proof-sketch}).
-Then it contains the proof of \Thmref{thm:ccab:rtp:ms} (\Cref{sec:proof1}) and of \Thmref{thm:ccall:rtp:msscct} (\Cref{sec:proof2}).
-
-\section{Proving Robust Safety Property Preservation}\label{sec:proof-sketch}
-
-We illustrate the proof of \Cref{thm:cca:rtp:tms} since the other secure compilation proofs of this paper follow the same approach.
-Unfolding the theorem statement yields the following assumptions:  
-  for any $\pi\in\lceil\tmssafe\rceil$\footnote{$\lceil\cdot\rceil$ lifts the property to a hyperproperty by applying the powerset operation~\cite{clarkson2008hyper}.}, 
-  $\src{\trace}$, 
-  $\src{\runtimetermvar}$, 
-  and component $\src{\library_{comp}}$, 
-  we have that $\rsat{\src{\library_{comp}}}{\pi}$ and 
-  $\progstepto{\prog[\trg]{\trg{\library_{\ctx}}}{\cca(\src{\library_{\comp}})}}{\trg{\runtimetermvar}}{\trg{\trace}}$, where $\trg{\library_{\ctx}}$ is arbitrary.
-The proof obligation is $\msfilterL{\trg{\trace}}\in\pi$, i.e., the specification trace associated to $\trg{\trace}$ satisfies the property $\pi$.
-A way to show this is to relate trace $\trg{\trace}$ to some $\Ltms$ trace $\src{\trace}$ (which already satisfies the property as per the assumptions).
-The assumptions already contain a target execution associated to this trace, so the task is to find an associated $\Ltms$ execution that yields $\trg{\trace}$.
-The trace $\trg{\trace}$ is split into different parts, as commonly done in secure compilation works~\cite{korashy2021capableptrs,abate2018goodcomps}, where each part contains the events that either the context or the component does, but not both.
-Because of this, all such trace segments are ,,well-bracketed'' in the sense that they start with either $\trg{\ev{Start}}$, $\trg{\ev{Call}\ \comptoctx\ foo\ \valueexpr}$, or $\trg{\ev{Ret}\ \comptoctx\ \valueexpr}$ and end with either $\trg{\ev{End}\ \valueexpr}$, $\trg{\ev{Ret}\ \ctxtocomp\ \valueexpr}$, or $\trg{\ev{Call}\ \ctxtocomp\ foo\ \valueexpr}$.
-In the following, the former is referred to as a context segment, since these executions happen in $\trg{\library_{\ctx}}$, and the latter is referred to as a component segment, since these executions happen in $\cca(\src{\library_{\comp}})$.
-\Cref{fig:proofdiag:rtp} visualises this division for a program execution with one call from context to component and how the target execution is related to a source execution.
-In the figure, the green dashed lines encompass the component segments while the orange boxes contain the actual context switches from context to component or vice versa.
-From a technical perspective, as typically done in compilation proofs, the proof requires some setup to maintain a relation between $\src{\Omega}$ and $\trg{\Omega}$ .
-Two cross-language relations make this precise: (i) $\multimap_{\delta}$ relates states that are involved in a context segment, allowing the target execution to perform internal calls, and (ii) $\approx_{\delta}$ relates states that are involved in a component segment, where both states need to agree exactly, i.e., the memory and the control flow states are required to contain the same information.
-The relations are indexed with $\delta$, which is an injective mapping from $\Ltms$ locations $\src{\loc}$ to $\Ltrg$ locations $\trg{\loc}$.
-%Note that the relations $\multimap_{\delta}$ and $\approx_{\delta}$ swap when context switching.
-
-So far, the paper explained how to relate an $\Ltms$ execution with a $\Ltrg$ execution.
-The next question is therefore how to build the corresponding $\Ltms$ execution.
-This is done using a standard secure compilation proof technique called trace-based backtranslation~\cite{abate2019jour,korashy2021capableptrs,patrignani2021rsc}, which can be used to build a context $\src{\library_{\ctx}}$ that behaves similar to $\trg{\library_{\ctx}}$.
-% Since $\Ltms$ is statically typed, the context $\src{\library_{\ctx}}$ obtained from the backtranslation needs to be well-typed when linked with $\src{\library_{\comp}}$.
-For context segments of the trace $\trg{\trace}$ it is also necessary to show that the execution behaves similarily, i.e., the context obtained from the backtranslation generates trace $\src{\trace}$.
-For component segments of the trace, the relatedness of states and traces follows from a compiler correctness argument.
-These two arguments yield the source execution $\progstepto{\src{\prog{\library_{\ctx}}{\library_{\comp}}}}{\src{\runtimetermvar}}{\src{\trace}}$.
-
-The proof now works as follows.
-Given that $\msfilterL{\trg{\trace}}=\msfilterLtms{\src{\trace}}$, the proof goal changes from $\msfilterL{\trg{\trace}}\in\pi$ to $\msfilterLtms{\src{\trace}}\in\pi$.
-This follows by specializing the robust satisfaction assumption $\rsat{\src{\library_{\comp}}}{\pi}$ to use the context $\src{\library_{\ctx}}$, which is obtained from the backtranslation, and to use the source execution $\progstepto{\src{\prog{\library_{\ctx}}{\library_{\comp}}}}{\src{\runtimetermvar}}{\src{\trace}}$.
-
-\begin{figure*}[!ht]
-  \begin{center}
-    \begin{tikzpicture}[state/.style={minimum height=1.5cm}]
-      % relative horizontal/vertical distance between states
-      \pgfmathsetmacro{\hdist}{0.95}
-      \pgfmathsetmacro{\vdist}{1.35}
-      \pgfmathsetmacro{\halfvdist}{0.725}
-
-      % row of src states
-      \node[state] (srcempty) {$\rtt{\src{\emptyset}}{\src{\lcall{main}{0}}}$};
-      \foreach \s [remember=\s as \cur (initially empty)] in {1,w_1,p,w_2,2} {
-        \node[state,right=\hdist of src\cur] (src\s) {$\rtt{\src{\Omega_{\s}}}{\src{\expr[_{\s}]}}$};
-      }
-      % row of trg states
-      \node[state,below=\vdist of srcempty] (trgempty) {$\rtt{\trg{\emptyset}}{\trg{\lcall{main}{0}}}$};
-      \foreach \s in {1,w_1,p,w_2,2} {
-        \node[state,below=\vdist of src\s] (trg\s) {$\rtt{\trg{\Omega_{\s}}}{\trg{\expr[_{\s}]}}$};
-      }
-      %% illustrations
-        % backtrans wrapper 1
-        \draw[ultra thick,loosely dotted,Peach!50,rounded corners] (src1.north east) -- (srcw\string_1.north east)
-          -- (trgw\string_1.south east) -- (trg1.south west) -- (src1.north west) -- cycle;
-        % backtrans wrapper 2
-        \draw[ultra thick,loosely dotted,Peach!50,rounded corners] (srcp.north east) -- (srcw\string_2.north east)
-          -- (trgw\string_2.south east) -- (trgp.south west) -- (srcp.north west) -- cycle;
-        % compiler correctness
-        \draw[ultra thick,loosely dashed,Emerald!50,rounded corners] ($(srcw\string_1.north east)+(0,0.05)$) -- ($(srcp.north east)+(0,0.05)$)
-          -- ($(trgw\string_2.north east)+(0.05,0)$) -- ($(trgw\string_2.south east)+(0.05,-0.05)$)
-          -- ($(trg1.south west)+(-0.05,-0.05)$) -- ($(trg1.north west)+(-0.05,0)$) -- ($(srcw\string_1.north west)+(0,0.05)$) -- cycle;
-      % state relations
-      \path (srcempty) edge[very thick,draw=gray!25] node[pos=0.5,sloped,rotate=180,fill=white] {\scriptsize$\multimap_\emptyset$} (trgempty)
-        (src1) edge[very thick,draw=gray!25] node[pos=0.5,sloped,rotate=180,fill=white] {\scriptsize$\multimap_{\delta_1}$} (trg1)
-        (srcp) edge[very thick,draw=gray!25] node[pos=0.5,sloped,fill=white] {\scriptsize$\approx_{\delta_p}$} (trgp)
-        (src2) edge[very thick,draw=gray!25] node[pos=0.5,sloped,rotate=180,fill=white] {\scriptsize$\multimap_{\delta_2}$} (trg2)
-        (srcw\string_1) edge[very thick,draw=gray!25] node[pos=0.5,sloped,fill=white] {\scriptsize$\approx_{\delta_{w_1}}$} (trgw\string_1)
-        (srcw\string_2) edge[very thick,draw=gray!25] node[pos=0.5,sloped,rotate=180,fill=white] {\scriptsize$\multimap_{\delta_{w_2}}$} (trgw\string_2)
-        % diagonals
-        (srcw\string_1) edge[thick,draw=gray!25] node[pos=0.16,sloped,rotate=180,fill=white] {\scriptsize$\approx_{\delta_{w_1}}$} (trg1)
-        (src1) edge[thick,draw=gray!25] node[pos=0.16,sloped,rotate=180,fill=white] {\scriptsize$\multimap_{\delta_1}$} (trgw\string_1)
-        (srcp) edge[thick,draw=gray!25] node[pos=0.2,sloped,fill=white] {\scriptsize$\approx_{\delta_p}$} (trgw\string_2)
-        (trgp) edge[thick,draw=gray!25] node[pos=0.2,sloped,fill=white] {\scriptsize$\multimap_{\delta_{w_2}}$} (srcw\string_2)
-        ;
-      %\drawpolygon src1,srcw\string_1,trgw\string_1,trg1;
-      %\drawpolygon srcp,srcw\string_2,trgw\string_2,trgp;
-      %\node[font=\tiny,align=center,above=0.2 of srcw1srcp] (wrapper) {Backtranslation\\Wrapper};
-      %\path[->,draw] (wrapper) -- (srcw\string_1);
-      %\path[->,draw] (wrapper) -- (srcp);
-      % steps
-      \path[very thick,color=\stlccol] (srcempty) edge[draw=none] node {\ $\xrightarrow{\trace[_1]}{}{\kern-3.5pt}_{\text{ctx}}^*$} (src1)
-        (src1) edge[draw=none] node {\ $\xrightarrow{\trace[_c]}{}{\kern-3.5pt}_{\text{ctx}}$} (srcw\string_1)
-        (srcw\string_1) edge[draw=none] node {\ $\xrightarrow{\trace[_p]}{}{\kern-3.5pt}_{\text{ctx}}^*$} (srcp)
-        (srcp) edge[draw=none] node[fill=white,inner sep=0,outer sep=0] {\ $\xrightarrow{\trace[_r]}{}{\kern-3.5pt}_{\text{ctx}}$} (srcw\string_2)
-        (srcw\string_2) edge[draw=none] node {\ $\xrightarrow{\trace[_2]}{}{\kern-3.5pt}_{\text{ctx}}^*$} (src2)
-        ;
-      \path[very thick,color=\ulccol] (trgempty) edge[draw=none] node[fill=white,inner sep=0,outer sep=0] {\ $\xrightarrow{\phantom{\trace[_1]}}{}{\kern-3.5pt}_{\text{ctx}}^*$} (trg1)
-        (trg1) edge[draw=none] node {\ $\xrightarrow{\phantom{\trace[_p]}}{}{\kern-3.5pt}_{\text{ctx}}^*$} (trgw\string_1)
-        (trgw\string_1) edge[draw=none] node[fill=white,inner sep=0,outer sep=0] {\ $\xrightarrow{\phantom{\trace[_p]}}{}{\kern-3.5pt}_{\text{ctx}}^*$} (trgp)
-        (trgp) edge[very thick,draw=none] node {\ $\xrightarrow{\phantom{\trace[_p]}}{}{\kern-3.5pt}_{\text{ctx}}^*$} (trgw\string_2)
-        (trgw\string_2) edge[draw=none] node {\ $\xrightarrow{\phantom{\trace[_2]}}{}{\kern-3.5pt}_{\text{ctx}}^*$} (trg2)
-        ;
-      % legend
-%     \node[align=left,below right=0.3 and 0.3 of trgempty,font=\tiny] (legend) {%
-%       ${\src{\trace[_c]}}\cong{\backt{\Ltms}{\Ltrg}(\trg{Call\ ?\ foo\ \valueexpr})}$\\%
-%       ${\src{\trace[_r]}}\cong{\backt{\Ltms}{\Ltrg}(\trg{Ret\ !\ foo\ \valueexpr})}$
-%       };
-      \node[below=0.5 of trgempty,draw=none] (legend) {};
-      \draw[ultra thick,loosely dotted,Peach!50,rounded corners] ($(legend.north east)+(0.5,-0.4)$) -- ($(legend.north east)+(1,-0.4)$);
-      \node at ($(legend.north east)+(1.5,-0.4)$) (legendwrapper) {};
-      \node[right of=legendwrapper] {\tiny Backtranslation Wrapper};
-      \draw[ultra thick,loosely dashed,Emerald!50,rounded corners] ($(legend.south east)+(0.5,0.4)$) -- ($(legend.south east)+(1,0.4)$);
-      \node at ($(legend.south east)+(1.5,0.4)$) (legendcorrectness) {};
-      \node[right of=legendcorrectness] {\tiny Compiler Correctness};
-    \end{tikzpicture}
-    \caption{Proof diagram for \Cref{thm:cca:rtp:tms} depicting the general structure of robust preservation proofs. %
-      Nodes in the graph represent runtime states. %
-      Vertical lines indicate cross language relations, while horizontal ones are execution steps. %
-      The green dashed trapezoid encompasses the component segment, while the orange dotted rectangles entail the context switches. %
-      $\Ltrg$ traces are omitted for readability. %
-      $\Ltms$ trace segments $\src{\trace[_{c}]}$ and $\src{\trace[_{r}]}$ describe the events that happen at the boundaries, i.e., during a context switch.
-      $\src{\trace_{p}}$ is the behavior of the component and the traces $\src{\trace[_{1}]}$ and $\src{\trace[_{2}]}$ describe the context.
-    }\label{fig:proofdiag:rtp}
-  \end{center}
-\end{figure*}
-
-\section{Proof of \Cref{thm:ccab:rtp:ms}} \label{sec:proof1}
-  From \Thmref{thm:cca:rtp:tms} it follows that for any $\Ltms$ program $\src{\progvar}$, it compiles to an $\Ltrg$ program $\trg{\progvar}$ that robustly satisfies \gls*{tms}.
-  Note that $\src{\progvar}$ robustly satisfies \gls*{tms} by the properties of the typesystem of $\Ltms$.
-  Then, \Thmref{thm:ccb:rtp:sms} demonstrates that, assuming $\trg{\progvar}$ robustly satisfies \gls*{sms}, the program $\trg{\progvar}$ compiles to an $\Lms$ program $\irl{p}$ that also robustly satisfies \gls*{sms}.
-  From \Thmref{thm:ccab:rtp:ms} it follows that $\src{p}$ compiles to $\irl{p}$ that robustly satisfies \gls*{ms}, since \gls*{ms} is the intersection of \gls*{tms} and \gls*{sms}.
-
-\section{Proof of \Cref{thm:ccall:rtp:msscct}} \label{sec:proof2}
-From \Thmref{thm:ccab:rtp:ms}, we have that any $\Ltms$ program $\src{\progvar}$ compiles into a $\Lms$ program $\irl{\progvar}$ that robustly satisfies \gls*{ms}.
-Then, from \Thmref{thm:cccfccdce:rtp:ms} we have that $\irl{\progvar}$ gets optimised to a program $\irl{\progvar[']}$ that is also \gls*{ms}, where the order of optimisations does not matter for $\irl{\progvar[']}$ to be \gls*{ms}.
-Assuming $\irl{\progvar[']}$ robustly satisfies \gls*{scct}, by \Thmref{thm:ccscct:rtp:scct} it compiles to an $\Lscct$ program $\obj{\progvar}$ that robustly satisfies \gls*{scct} as well.
-Finally, from \Thmref{thm:rtp} it follows that, given $\src{\progvar}$ robustly satisfies \gls*{scct} and \gls*{ms}, $\obj{\progvar}$ also robustly satisfies \gls*{scct} and \gls*{ms}.
-
-\section{Auxiliary definitions for \Cref{subsec:app}}
-\begin{definition}[Existential Image of Properties]\label{def:eximg:prop}
-  Given a cross-language trace relation $\sim$,
-  $$
-  \tau_\sim(\src{\pi}) \isdef \left\{
-    \trg{\trace} \mid \exists\src{\trace}.\src{\trace}\sim\trg{\trace} \text{ and }
-    \src{\trace}\in\src{\pi}
-  \right\}
-  $$
-\end{definition}
-
-\Thmref{def:eximg:prop} can be lifted to hyperproperties/classes as follows:
-\begin{definition}[Existential Image of Hyperproperties]\label{def:eximg:hprop}
-  Given a cross-language trace relation $\sim$,
-  $$
-  \tilde{\tau}_\sim(\src{\Pi}) \isdef \left\{
-    \trg{\pi} \mid \exists\src{\pi}.\src{\pi}\in\src{\Pi} \text{ and }
-    \trg{\pi}=\tau_\sim\left(\src{\pi}\right)
-  \right\}
-  $$
-\end{definition}
-\bul{A cross-language relation $\sim$ is well-formed with respect to a source-level (hyper-)property $\src{\Pi}$} iff \rul{the existential images preserve membership}.
-\begin{definition}[Well-formedness of Cross-Language Trace Relations for a Properties $\src{\Pi}$]\label{def:wf:tracerel}
-  $$
-  \text{\bul{$\wfc{\sim}{\src{\Pi}}$}} \isdef \text{\rul{$\forall \src{\pi}\in\src{\Pi}, \tau_\sim(\src{\pi})\in\tilde{\tau}_\sim(\src{\Pi})$}}
-  $$
-\end{definition}
-\begin{lemma}[Well-Formedness Equivalence for Classes]\label{lem:wf:equiv:wfc}
-  $\;\;$
-  $ \wf{\sim}{\src{\class}} \Leftrightarrow\ \wfc{\sim}{\src{\class}} $\Coqed
-\end{lemma}
-
-\newpage
-
-\bibliographystyle{IEEEtran}
+% \section*{Appendix}
+
+% The appendix contains insights on the structure of proofs for the theorems of \Cref{sec:casestud:rtp} (\Cref{sec:proof-sketch}).
+% Then it contains the proof of \Thmref{thm:ccab:rtp:ms} (\Cref{sec:proof1}) and of \Thmref{thm:ccall:rtp:msscct} (\Cref{sec:proof2}).
+
+% \section{Proving Robust Safety Property Preservation}\label{sec:proof-sketch}
+
+% We illustrate the proof of \Cref{thm:cca:rtp:tms} since the other secure compilation proofs of this paper follow the same approach.
+% Unfolding the theorem statement yields the following assumptions:  
+%   for any $\pi\in\lceil\tmssafe\rceil$\footnote{$\lceil\cdot\rceil$ lifts the property to a hyperproperty by applying the powerset operation~\cite{clarkson2008hyper}.}, 
+%   $\src{\trace}$, 
+%   $\src{\runtimetermvar}$, 
+%   and component $\src{\library_{comp}}$, 
+%   we have that $\rsat{\src{\library_{comp}}}{\pi}$ and 
+%   $\progstepto{\prog[\trg]{\trg{\library_{\ctx}}}{\cca(\src{\library_{\comp}})}}{\trg{\runtimetermvar}}{\trg{\trace}}$, where $\trg{\library_{\ctx}}$ is arbitrary.
+% The proof obligation is $\msfilterL{\trg{\trace}}\in\pi$, i.e., the specification trace associated to $\trg{\trace}$ satisfies the property $\pi$.
+% A way to show this is to relate trace $\trg{\trace}$ to some $\Ltms$ trace $\src{\trace}$ (which already satisfies the property as per the assumptions).
+% The assumptions already contain a target execution associated to this trace, so the task is to find an associated $\Ltms$ execution that yields $\trg{\trace}$.
+% The trace $\trg{\trace}$ is split into different parts, as commonly done in secure compilation works~\cite{korashy2021capableptrs,abate2018goodcomps}, where each part contains the events that either the context or the component does, but not both.
+% Because of this, all such trace segments are ,,well-bracketed'' in the sense that they start with either $\trg{\ev{Start}}$, $\trg{\ev{Call}\ \comptoctx\ foo\ \valueexpr}$, or $\trg{\ev{Ret}\ \comptoctx\ \valueexpr}$ and end with either $\trg{\ev{End}\ \valueexpr}$, $\trg{\ev{Ret}\ \ctxtocomp\ \valueexpr}$, or $\trg{\ev{Call}\ \ctxtocomp\ foo\ \valueexpr}$.
+% In the following, the former is referred to as a context segment, since these executions happen in $\trg{\library_{\ctx}}$, and the latter is referred to as a component segment, since these executions happen in $\cca(\src{\library_{\comp}})$.
+% \Cref{fig:proofdiag:rtp} visualises this division for a program execution with one call from context to component and how the target execution is related to a source execution.
+% In the figure, the green dashed lines encompass the component segments while the orange boxes contain the actual context switches from context to component or vice versa.
+% From a technical perspective, as typically done in compilation proofs, the proof requires some setup to maintain a relation between $\src{\Omega}$ and $\trg{\Omega}$ .
+% Two cross-language relations make this precise: (i) $\multimap_{\delta}$ relates states that are involved in a context segment, allowing the target execution to perform internal calls, and (ii) $\approx_{\delta}$ relates states that are involved in a component segment, where both states need to agree exactly, i.e., the memory and the control flow states are required to contain the same information.
+% The relations are indexed with $\delta$, which is an injective mapping from $\Ltms$ locations $\src{\loc}$ to $\Ltrg$ locations $\trg{\loc}$.
+% %Note that the relations $\multimap_{\delta}$ and $\approx_{\delta}$ swap when context switching.
+
+% So far, the paper explained how to relate an $\Ltms$ execution with a $\Ltrg$ execution.
+% The next question is therefore how to build the corresponding $\Ltms$ execution.
+% This is done using a standard secure compilation proof technique called trace-based backtranslation~\cite{abate2019jour,korashy2021capableptrs,patrignani2021rsc}, which can be used to build a context $\src{\library_{\ctx}}$ that behaves similar to $\trg{\library_{\ctx}}$.
+% % Since $\Ltms$ is statically typed, the context $\src{\library_{\ctx}}$ obtained from the backtranslation needs to be well-typed when linked with $\src{\library_{\comp}}$.
+% For context segments of the trace $\trg{\trace}$ it is also necessary to show that the execution behaves similarily, i.e., the context obtained from the backtranslation generates trace $\src{\trace}$.
+% For component segments of the trace, the relatedness of states and traces follows from a compiler correctness argument.
+% These two arguments yield the source execution $\progstepto{\src{\prog{\library_{\ctx}}{\library_{\comp}}}}{\src{\runtimetermvar}}{\src{\trace}}$.
+
+% The proof now works as follows.
+% Given that $\msfilterL{\trg{\trace}}=\msfilterLtms{\src{\trace}}$, the proof goal changes from $\msfilterL{\trg{\trace}}\in\pi$ to $\msfilterLtms{\src{\trace}}\in\pi$.
+% This follows by specializing the robust satisfaction assumption $\rsat{\src{\library_{\comp}}}{\pi}$ to use the context $\src{\library_{\ctx}}$, which is obtained from the backtranslation, and to use the source execution $\progstepto{\src{\prog{\library_{\ctx}}{\library_{\comp}}}}{\src{\runtimetermvar}}{\src{\trace}}$.
+
+% \begin{figure*}[!ht]
+%   \begin{center}
+%     \begin{tikzpicture}[state/.style={minimum height=1.5cm}]
+%       % relative horizontal/vertical distance between states
+%       \pgfmathsetmacro{\hdist}{0.95}
+%       \pgfmathsetmacro{\vdist}{1.35}
+%       \pgfmathsetmacro{\halfvdist}{0.725}
+
+%       % row of src states
+%       \node[state] (srcempty) {$\rtt{\src{\emptyset}}{\src{\lcall{main}{0}}}$};
+%       \foreach \s [remember=\s as \cur (initially empty)] in {1,w_1,p,w_2,2} {
+%         \node[state,right=\hdist of src\cur] (src\s) {$\rtt{\src{\Omega_{\s}}}{\src{\expr[_{\s}]}}$};
+%       }
+%       % row of trg states
+%       \node[state,below=\vdist of srcempty] (trgempty) {$\rtt{\trg{\emptyset}}{\trg{\lcall{main}{0}}}$};
+%       \foreach \s in {1,w_1,p,w_2,2} {
+%         \node[state,below=\vdist of src\s] (trg\s) {$\rtt{\trg{\Omega_{\s}}}{\trg{\expr[_{\s}]}}$};
+%       }
+%       %% illustrations
+%         % backtrans wrapper 1
+%         \draw[ultra thick,loosely dotted,Peach!50,rounded corners] (src1.north east) -- (srcw\string_1.north east)
+%           -- (trgw\string_1.south east) -- (trg1.south west) -- (src1.north west) -- cycle;
+%         % backtrans wrapper 2
+%         \draw[ultra thick,loosely dotted,Peach!50,rounded corners] (srcp.north east) -- (srcw\string_2.north east)
+%           -- (trgw\string_2.south east) -- (trgp.south west) -- (srcp.north west) -- cycle;
+%         % compiler correctness
+%         \draw[ultra thick,loosely dashed,Emerald!50,rounded corners] ($(srcw\string_1.north east)+(0,0.05)$) -- ($(srcp.north east)+(0,0.05)$)
+%           -- ($(trgw\string_2.north east)+(0.05,0)$) -- ($(trgw\string_2.south east)+(0.05,-0.05)$)
+%           -- ($(trg1.south west)+(-0.05,-0.05)$) -- ($(trg1.north west)+(-0.05,0)$) -- ($(srcw\string_1.north west)+(0,0.05)$) -- cycle;
+%       % state relations
+%       \path (srcempty) edge[very thick,draw=gray!25] node[pos=0.5,sloped,rotate=180,fill=white] {\scriptsize$\multimap_\emptyset$} (trgempty)
+%         (src1) edge[very thick,draw=gray!25] node[pos=0.5,sloped,rotate=180,fill=white] {\scriptsize$\multimap_{\delta_1}$} (trg1)
+%         (srcp) edge[very thick,draw=gray!25] node[pos=0.5,sloped,fill=white] {\scriptsize$\approx_{\delta_p}$} (trgp)
+%         (src2) edge[very thick,draw=gray!25] node[pos=0.5,sloped,rotate=180,fill=white] {\scriptsize$\multimap_{\delta_2}$} (trg2)
+%         (srcw\string_1) edge[very thick,draw=gray!25] node[pos=0.5,sloped,fill=white] {\scriptsize$\approx_{\delta_{w_1}}$} (trgw\string_1)
+%         (srcw\string_2) edge[very thick,draw=gray!25] node[pos=0.5,sloped,rotate=180,fill=white] {\scriptsize$\multimap_{\delta_{w_2}}$} (trgw\string_2)
+%         % diagonals
+%         (srcw\string_1) edge[thick,draw=gray!25] node[pos=0.16,sloped,rotate=180,fill=white] {\scriptsize$\approx_{\delta_{w_1}}$} (trg1)
+%         (src1) edge[thick,draw=gray!25] node[pos=0.16,sloped,rotate=180,fill=white] {\scriptsize$\multimap_{\delta_1}$} (trgw\string_1)
+%         (srcp) edge[thick,draw=gray!25] node[pos=0.2,sloped,fill=white] {\scriptsize$\approx_{\delta_p}$} (trgw\string_2)
+%         (trgp) edge[thick,draw=gray!25] node[pos=0.2,sloped,fill=white] {\scriptsize$\multimap_{\delta_{w_2}}$} (srcw\string_2)
+%         ;
+%       %\drawpolygon src1,srcw\string_1,trgw\string_1,trg1;
+%       %\drawpolygon srcp,srcw\string_2,trgw\string_2,trgp;
+%       %\node[font=\tiny,align=center,above=0.2 of srcw1srcp] (wrapper) {Backtranslation\\Wrapper};
+%       %\path[->,draw] (wrapper) -- (srcw\string_1);
+%       %\path[->,draw] (wrapper) -- (srcp);
+%       % steps
+%       \path[very thick,color=\stlccol] (srcempty) edge[draw=none] node {\ $\xrightarrow{\trace[_1]}{}{\kern-3.5pt}_{\text{ctx}}^*$} (src1)
+%         (src1) edge[draw=none] node {\ $\xrightarrow{\trace[_c]}{}{\kern-3.5pt}_{\text{ctx}}$} (srcw\string_1)
+%         (srcw\string_1) edge[draw=none] node {\ $\xrightarrow{\trace[_p]}{}{\kern-3.5pt}_{\text{ctx}}^*$} (srcp)
+%         (srcp) edge[draw=none] node[fill=white,inner sep=0,outer sep=0] {\ $\xrightarrow{\trace[_r]}{}{\kern-3.5pt}_{\text{ctx}}$} (srcw\string_2)
+%         (srcw\string_2) edge[draw=none] node {\ $\xrightarrow{\trace[_2]}{}{\kern-3.5pt}_{\text{ctx}}^*$} (src2)
+%         ;
+%       \path[very thick,color=\ulccol] (trgempty) edge[draw=none] node[fill=white,inner sep=0,outer sep=0] {\ $\xrightarrow{\phantom{\trace[_1]}}{}{\kern-3.5pt}_{\text{ctx}}^*$} (trg1)
+%         (trg1) edge[draw=none] node {\ $\xrightarrow{\phantom{\trace[_p]}}{}{\kern-3.5pt}_{\text{ctx}}^*$} (trgw\string_1)
+%         (trgw\string_1) edge[draw=none] node[fill=white,inner sep=0,outer sep=0] {\ $\xrightarrow{\phantom{\trace[_p]}}{}{\kern-3.5pt}_{\text{ctx}}^*$} (trgp)
+%         (trgp) edge[very thick,draw=none] node {\ $\xrightarrow{\phantom{\trace[_p]}}{}{\kern-3.5pt}_{\text{ctx}}^*$} (trgw\string_2)
+%         (trgw\string_2) edge[draw=none] node {\ $\xrightarrow{\phantom{\trace[_2]}}{}{\kern-3.5pt}_{\text{ctx}}^*$} (trg2)
+%         ;
+%       % legend
+% %     \node[align=left,below right=0.3 and 0.3 of trgempty,font=\tiny] (legend) {%
+% %       ${\src{\trace[_c]}}\cong{\backt{\Ltms}{\Ltrg}(\trg{Call\ ?\ foo\ \valueexpr})}$\\%
+% %       ${\src{\trace[_r]}}\cong{\backt{\Ltms}{\Ltrg}(\trg{Ret\ !\ foo\ \valueexpr})}$
+% %       };
+%       \node[below=0.5 of trgempty,draw=none] (legend) {};
+%       \draw[ultra thick,loosely dotted,Peach!50,rounded corners] ($(legend.north east)+(0.5,-0.4)$) -- ($(legend.north east)+(1,-0.4)$);
+%       \node at ($(legend.north east)+(1.5,-0.4)$) (legendwrapper) {};
+%       \node[right of=legendwrapper] {\tiny Backtranslation Wrapper};
+%       \draw[ultra thick,loosely dashed,Emerald!50,rounded corners] ($(legend.south east)+(0.5,0.4)$) -- ($(legend.south east)+(1,0.4)$);
+%       \node at ($(legend.south east)+(1.5,0.4)$) (legendcorrectness) {};
+%       \node[right of=legendcorrectness] {\tiny Compiler Correctness};
+%     \end{tikzpicture}
+%     \caption{Proof diagram for \Cref{thm:cca:rtp:tms} depicting the general structure of robust preservation proofs. %
+%       Nodes in the graph represent runtime states. %
+%       Vertical lines indicate cross language relations, while horizontal ones are execution steps. %
+%       The green dashed trapezoid encompasses the component segment, while the orange dotted rectangles entail the context switches. %
+%       $\Ltrg$ traces are omitted for readability. %
+%       $\Ltms$ trace segments $\src{\trace[_{c}]}$ and $\src{\trace[_{r}]}$ describe the events that happen at the boundaries, i.e., during a context switch.
+%       $\src{\trace_{p}}$ is the behavior of the component and the traces $\src{\trace[_{1}]}$ and $\src{\trace[_{2}]}$ describe the context.
+%     }\label{fig:proofdiag:rtp}
+%   \end{center}
+% \end{figure*}
+
+% \section{Proof of \Cref{thm:ccab:rtp:ms}} \label{sec:proof1}
+%   From \Thmref{thm:cca:rtp:tms} it follows that for any $\Ltms$ program $\src{\progvar}$, it compiles to an $\Ltrg$ program $\trg{\progvar}$ that robustly satisfies \gls*{tms}.
+%   Note that $\src{\progvar}$ robustly satisfies \gls*{tms} by the properties of the typesystem of $\Ltms$.
+%   Then, \Thmref{thm:ccb:rtp:sms} demonstrates that, assuming $\trg{\progvar}$ robustly satisfies \gls*{sms}, the program $\trg{\progvar}$ compiles to an $\Lms$ program $\irl{p}$ that also robustly satisfies \gls*{sms}.
+%   From \Thmref{thm:ccab:rtp:ms} it follows that $\src{p}$ compiles to $\irl{p}$ that robustly satisfies \gls*{ms}, since \gls*{ms} is the intersection of \gls*{tms} and \gls*{sms}.
+
+% \section{Proof of \Cref{thm:ccall:rtp:msscct}} \label{sec:proof2}
+% From \Thmref{thm:ccab:rtp:ms}, we have that any $\Ltms$ program $\src{\progvar}$ compiles into a $\Lms$ program $\irl{\progvar}$ that robustly satisfies \gls*{ms}.
+% Then, from \Thmref{thm:cccfccdce:rtp:ms} we have that $\irl{\progvar}$ gets optimised to a program $\irl{\progvar[']}$ that is also \gls*{ms}, where the order of optimisations does not matter for $\irl{\progvar[']}$ to be \gls*{ms}.
+% Assuming $\irl{\progvar[']}$ robustly satisfies \gls*{scct}, by \Thmref{thm:ccscct:rtp:scct} it compiles to an $\Lscct$ program $\obj{\progvar}$ that robustly satisfies \gls*{scct} as well.
+% Finally, from \Thmref{thm:rtp} it follows that, given $\src{\progvar}$ robustly satisfies \gls*{scct} and \gls*{ms}, $\obj{\progvar}$ also robustly satisfies \gls*{scct} and \gls*{ms}.
+
+% \section{Auxiliary definitions for \Cref{subsec:app}}
+% \begin{definition}[Existential Image of Properties]\label{def:eximg:prop}
+%   Given a cross-language trace relation $\sim$,
+%   $$
+%   \tau_\sim(\src{\pi}) \isdef \left\{
+%     \trg{\trace} \mid \exists\src{\trace}.\src{\trace}\sim\trg{\trace} \text{ and }
+%     \src{\trace}\in\src{\pi}
+%   \right\}
+%   $$
+% \end{definition}
+
+% \Thmref{def:eximg:prop} can be lifted to hyperproperties/classes as follows:
+% \begin{definition}[Existential Image of Hyperproperties]\label{def:eximg:hprop}
+%   Given a cross-language trace relation $\sim$,
+%   $$
+%   \tilde{\tau}_\sim(\src{\Pi}) \isdef \left\{
+%     \trg{\pi} \mid \exists\src{\pi}.\src{\pi}\in\src{\Pi} \text{ and }
+%     \trg{\pi}=\tau_\sim\left(\src{\pi}\right)
+%   \right\}
+%   $$
+% \end{definition}
+% \bul{A cross-language relation $\sim$ is well-formed with respect to a source-level (hyper-)property $\src{\Pi}$} iff \rul{the existential images preserve membership}.
+% \begin{definition}[Well-formedness of Cross-Language Trace Relations for a Properties $\src{\Pi}$]\label{def:wf:tracerel}
+%   $$
+%   \text{\bul{$\wfc{\sim}{\src{\Pi}}$}} \isdef \text{\rul{$\forall \src{\pi}\in\src{\Pi}, \tau_\sim(\src{\pi})\in\tilde{\tau}_\sim(\src{\Pi})$}}
+%   $$
+% \end{definition}
+% \begin{lemma}[Well-Formedness Equivalence for Classes]\label{lem:wf:equiv:wfc}
+%   $\;\;$
+%   $ \wf{\sim}{\src{\class}} \Leftrightarrow\ \wfc{\sim}{\src{\class}} $\Coqed
+% \end{lemma}
+
+% \newpage
+
+\bibliographystyle{splncs04}
 \bibliography{main}

### 2023-10-09 esop/main.tex
change csf to esop in code sample

@@
 int hello() {
-  char* greet = new char[12] { "Hello CSF!!" };
+  char* greet = new char[12] { "Hello ESOP!" };
   char* to = new char[12];
@@
   delete to;
-  printf("%cSF\n", to[6]);
+  printf("%cSOP\n", to[6]);
 }

### 2023-10-09 esop/main.tex
space + font

@@
 \usepackage[T1]{fontenc}
+\usepackage[scaled=.83]{beramono}
 
@@
 
-% \author{
+\author{}
 % Matthis Kruse\orcidID{0000-0003-4062-9666}
@@
 
-% \institute{
+\institute{}
 % CISPA Helmholtz Center for Information Security and Saarland University
@@
 \begin{definition}[Robust Satisfaction]\label{def:proprsat}
-  \begin{nscenter}
+  % \begin{nscenter}
   \bul{$\rsat{\progvar}{\pi}$}
   $\isdef$ \iul{$\forall \contextvar\ \wholeprogvar$, if $\link{\contextvar}{\progvar}=\wholeprogvar$}, then \oul{$\sat{\wholeprogvar}{\pi}$}.
-  \end{nscenter}
+  % \end{nscenter}
 \end{definition}
@@
 \begin{definition}[Robust Compilation]\label{def:rtp}
-  $\;$\\
-  \vspace{-1em}
-  \begin{nscenter}
+  % $\;$\\
+  % \vspace{-1em}
+  % \begin{nscenter}
   % Compiler $\cc{\src{L}}{\trg{L}}$ robustly preserves $\class$, 
@@
   \rul{$\forall \pi\in\class, \src{p}\in\src{L},$} if \iul{$\rsat{\src{\progvar}}{\pi}$}, then \oul{$\rsat{\cc{\src{L}}{\trg{L}}\left(\src{p}\right)}{\pi}$}.
-  \end{nscenter}
+  % \end{nscenter}
 \end{definition}
-
+% 
 Note that a class of properties $\class$ can represent just one property $\pi$ by lifting~\cite{clarkson2008hyper} that property to sets of properties, i.e., use the powerset of $\pi$ instead of $\pi$ itself.
@@
 \begin{definition}[\glsfirst*{sms}]\label{def:trace:smsdef}
-\small
+% \small
+  % \begin{nscenter}
+
+  \noindent
   $
@@
                                 $
+  % \end{nscenter}
 \end{definition}
@@
 \begin{definition}[\glsfirst*{scct}]\label{def:trace:scctdef}
-  $
+% \begin{nscenter}
+
+  \noindent$
   \scctsafe:=\left\{\trace \left|\begin{array}{rcl}
@@
   $
+% \end{nscenter}
 \end{definition}
@@
     % \\
-    \mi{(Abstract\ Events)}~\absevent\bnfdef&\ \absev{\emptyevent} \mid \absev{Alloc\ }\loc\ \sandboxtag \mid \absev{Dealloc\ }\loc\ \sandboxtag \\
-                                            & \mid \absev{Use\ }\loc\ \sandboxtag \mid \absterm\\
+    \mi{(Abstract\ Events)}~\absevent\bnfdef&\ \absev{\emptyevent} \mid \absev{Alloc\ }\loc\ \sandboxtag \mid \absev{Dealloc\ }\loc\ \sandboxtag 
+    % \\
+                                            % & 
+                                            \mid \absev{Use\ }\loc\ \sandboxtag \mid \absterm\\
   \end{aligned}
@@
     (\loc;\sandboxtag)\notin\tmsmonitor.\text{A} &
-    (\loc;\sandboxtag)\notin\tmsmonitor.\text{F} \\
+    (\loc;\sandboxtag)\notin\tmsmonitor.\text{F} 
+    % \\
+    &
     \tmsmonitor[']=\left\{\text{A}: \tmsmonitor.\text{A}\cup\left\{(\loc;\sandboxtag)\right\},\text{F}: \tmsmonitor.\text{F}\right\}
@@
    	\absev{Use}\ \loc[_{x}]\ \comp
-   	\\
-   	&
+   	% \\
+   	% &
    	\cdot
@@
   	\absev{Use}\ \loc[_{x}]\ \comp
-  	\\
-  	&
+  	% \\
+  	% &
   	\cdot
@@
   		\cdot
+  	% \\
+  	% &
+  		\ev{Use}\ \loc_{x}\ 0;\comp;\unlock
+  		\cdot
   		\\
   	&
-  		\ev{Use}\ \loc_{x}\ 0;\comp;\unlock
-  		\cdot
   		\ev{Use}\ \loc_{y}\ 0;\comp;\unlock
@@
   		\ev{Use}\ \loc_{x}\ 1;\comp;\unlock
-  	\\
-  	&
+  	% \\
+  	% &
   		\cdot
@@
   Omitting the events for all ,,$\ldots$'' for sake of brevity, the abstract monitor trace of this is:
-
+  	% 
   	{\small
@@
   By means of additional features to ensure cryptographic constant time even in the presence of memory reads and loads, such as a flag to enable a data independent timing mode, which is present in both Arm~\cite[p.~543]{arm-refman} and Intel~\cite[p.~80]{intel-refman} processors, the original trace now does $\ev{Use}\ \loc_{y}\ 0;\comp;\unlock$ instead of $\ev{Use}\ \loc_{y}\ 0;\comp;\lock$.
-  Because of this, the whole trace of the component equates to $\absev{\emptyevent}$ and the monitor can step without getting stuck.
+  Because of this, the whole trace of the component is $\absev{\emptyevent}$ and the monitor can step without getting stuck.
 \end{exampleenv}
@@
 \begin{lemma}[Monitor Satisfaction is $\msscctsafe$]\label{lem:mon:msscctsafe}
-  $\;$\\\vspace{-1em}
-  \begin{nscenter}
+  % $\;$\\\vspace{-1em}
+  % \begin{nscenter}
+  % \noindent
   If $\monmsscctsafe{\trace}$, then $\trace\in\msscctsafe$. %\Coqed
-  \end{nscenter}
+  % \end{nscenter}
 \end{lemma}
@@
 \begin{definition}[Well-formedness of Trace Relations for a Class $\src{\class}$]\label{def:wfc:tracerel}
-  $$
+  % $$
+
+  \noindent
   \text{\bul{$\wfc{\sim}{\src{\class}}$}} \isdef \text{\rul{$\forall \src{\Pi}\subseteq\src{\class}, \tau_\sim(\src{\Pi})\subseteq\tilde{\tau}_\sim(\src{\class})$}}
-  $$
+  % $$
 \end{definition}
@@
 \begin{definition}[Robust Compilation with $\sim$]\label{def:rtpsim}
-  $$\rtpsim{\cc{\src{L}}{\trg{L}}}{\src{\class}}{\sim} \isdef
+  $
+  \rtpsim{\cc{\src{L}}{\trg{L}}}{\src{\class}}{\sim} \isdef
   \forall \src{\pi} \src{p}.
@@
     \text{then }\rsat{\cc{\src{L}}{\trg{L}}\left(\src{p}\right)}{\tilde{\tau}_\sim\left(\src{\pi}\right)}.
-  $$
+  $
 \end{definition}
@@
   \begin{array}{rcl}
-    \ccb(\trg{\lnew{x}{\expr[_{1}]}{\expr[_{2}]}}) & = & \llet[\irl]{\irl{x_{SIZE}}}{\ccb(\trg{\expr[_{1}]})}{\\&&\lnew[\irl]{\irl{x}}{\irl{x_{SIZE}}}{\ccb(\trg{\expr[_{2}]})}} \\
-  \ccb(\trg{\lget{x}{\expr}}) & = & \llet[\irl]{\irl{x_{n}}}{\ccb(\trg{\expr})}{\\&&\lifz[\irl]{\irl{0\le x_{n}<x_{SIZE}}}{\irl{\lget{x}{x_{n}}}}{\\&&\irl{\labort}}} \\
-  \ccb(\trg{\lset{x}{\expr[_{1}]}{\expr[_{2}]}}) & = & \llet[\irl]{\irl{x_{n}}}{\ccb(\trg{\expr[_{1}]})}{\\&&\lifz[\irl]{\irl{0\le x_{n}<x_{SIZE}}}{\lset[\irl]{\irl{x}}{\irl{x_{n}}}{\\&&\ccb(\trg{\expr[_{2}]})}}{\irl{\labort}}} \\
+    \ccb(\trg{\lnew{x}{\expr[_{1}]}{\expr[_{2}]}}) & = 
+    	& \llet[\irl]{\irl{x_{SIZE}}}{\ccb(\trg{\expr[_{1}]})}{
+    		% \\&&
+    		\lnew[\irl]{\irl{x}}{\irl{x_{SIZE}}}{\ccb(\trg{\expr[_{2}]})}
+    		} \\
+  \ccb(\trg{\lget{x}{\expr}}) & = 
+  	& \llet[\irl]{\irl{x_{n}}}{\ccb(\trg{\expr})}{
+  	% \\&&
+  	\lifz[\irl]{\irl{0\le x_{n}<x_{SIZE}}}{\irl{\lget{x}{x_{n}}}}{
+  		% \\&&
+  		\irl{\labort}
+  		}
+  		} \\
+  \ccb(\trg{\lset{x}{\expr[_{1}]}{\expr[_{2}]}}) & = 
+  	& \llet[\irl]{\irl{x_{n}}}{\ccb(\trg{\expr[_{1}]})}{
+  		\\&&
+  		\lifz[\irl]{\irl{0\le x_{n}<x_{SIZE}}}{\lset[\irl]{\irl{x}}{\irl{x_{n}}}{
+  		% \\&&
+  		\ccb(\trg{\expr[_{2}]})
+  		}}{\irl{\labort}}
+  		} 
+  		% \\
   \end{array}
@@
     \begin{lstlisting}[language=c,basicstyle=\ttfamily\footnotesize, morekeywords={size_t}]
-void strncpy(size_t n, size_t dst_size, char *dst,
-                       size_t src_size, char *src) {
-  for(size_t i = 0; i < src_size && src[i] != '\0'
-                 && i < n; ++i) {
+void strncpy(size_t n, size_t dst_size, char *dst, size_t src_size, char *src) {
+  for(size_t i = 0; i < src_size && src[i] != '\0' && i < n; ++i) {
     if(i < src_size && i < dst_size) {

### 2023-10-11 esop/main.tex
nit

@@
 \end{itemize}
-
+% 
 This paper starts by introducing relevant notions of security properties and secure compilation (\Cref{sec:background}),
@@
 \begin{definition}[\glsfirst*{tms}]\label{def:trace:tmsdef}
-  \footnotesize$$
+  % \footnotesize
+  $$
   \tmssafe:=\left\{\trace \left| \begin{array}{rcl}
@@
 \end{center}
-
+% 
 \Cref{tr:tms-use} simply requires that a location is (i) allocated and (ii) not freed.
@@
     &
-    \mi{(Abstract\ Events)}~\absevent:=&\ \absev{\varepsilon} \mid \absterm \mid \absev{Leak} 
+    \qquad\mi{(Abstract\ Events)}~\absevent:=&\ \absev{\varepsilon} \mid \absterm \mid \absev{Leak} 
     % \\
@@
 The monitor state for the \gls*{scct} monitor is completely empty, since it does not need to keep track of information.
-As soon as any event is hit, the execution gets stuck, since any event is considered confidental from the perspective of this monitor.
+% As soon as 
+If
+any event is hit, the execution gets stuck, since any event is considered confidental from the perspective of this monitor.
 \begin{center}
@@
 First, $\src{TypeScript}$ programs are translated to $\trg{JavaScript}$ which, e.g., V8~\cite{googlev8} eventually compiles in parts to $\irl{Ignition BC}$.
-The following theorem establishes what happens if all these compilation steps were robustly secure with respect to \gls*{ms}:
+The following theorem establishes what happens if all these compilation steps were robustly secure with respect to e.g., \gls*{ms}:
 The resulting $\irl{Ignition BC}$ code would be \gls*{ms} regardless of the context the binary runs in.
@@
 Therefore, no behavior resulting from programs produced by the $\cc{\trg{L}}{\obj{L}}$ compiler is secure with respect to $\tmssafe$, but only for $\smssafe$.
-$\cc{\trg{L}}{\obj{L}}$ modifies the behavior and this section re-introduces the compositionality framework with support for these situations.
+$\cc{\trg{L}}{\obj{L}}$ modifies the behavior of compiled code and this section introduces support for these situations in the compositionality framework.
 
@@
 % 
-In fact, if \bul{that a relation $\sim_1$ is well-formed for a property $\class$} but \rul{another relation $\sim_2$ is not well-formed for the property mapped with the first relation $\tilde{\tau}_{\sim_1}\class$} then \iul{their composition is empty}.
+In fact, if \bul{a relation $\sim_1$ is well-formed for a property $\class$} but \rul{another relation $\sim_2$ is not well-formed for the property mapped with the first relation $\tilde{\tau}_{\sim_1}\class$} then \iul{their composition is empty}.
 % 
@@
 All presented programming languages share a common fragment. 
-Programs ($\prog{\library_{\ctx}}{\library_{\comp}}$) consist of libraries ($\library$), which are lists of functions ($\function$).
+Programs ($\prog{\library_{\ctx}}{\library_{\comp}}$) consist of libraries ($\library$) that belong to a context ($\ctx$) or to a component ($\comp$), and that are lists of functions ($\function$).
 Functions ($\lfunction{\textit{foo}}{x}{\expr}$) have a single argument $x$ and an expression ($\expr$) as a body which may have $x$ as a free variable.
@@
 The projection $\msfilterL{\trg{\trace}}$ is equal to $\msfilterLtms{\src{\trace}}$.
+\MPin{
+	what is poison. do we need the symbol for valid pointers. 
+}
 
@@
 
-\begin{gather*}\small
+\begin{gather*}
+% \small
   \begin{aligned}
@@
   	&
+  	\qquad
     \cca(\src{\lbinop{\expr[_{1}]}{\expr[_{2}]}}) = &\ \lbinop[\trg]{\left[\cca(\src{\expr[_{1}]})\right]}{\left[\cca(\src{\expr[_{2}]})\right]} \\
@@
     &
+    \qquad
     \cca(\src{\lget{x}{\expr}}) = &\ \lget[\trg]{\trg{x}}{\left[\cca(\src{\expr})\right]}
@@
 %    \cca(\src{\lset{x}{\expr[_{1}]}{\expr[_{2}]}}) & = \lset[\trg]{\left[\cca(\src{x})\right]}{\left[\cca(\src{\expr[_{1}]})\right]}{\left[\cca(\src{\expr_{2}})\right]} \\
-\cca(\src{\lfunction{g}{x:\natt\to\type[_{e}]}{\expr}})  = &\ \lfunction[\trg]{\trg{g}}{\trg{x}}{\lifz[\trg]{\trg{\lhast{x}{\natt}}}{\\%
-                                                            &\left[\cca(\src{\expr})\right] %
+\cca(\src{\lfunction{g}{x:\natt\to\type[_{e}]}{\expr}})  = &\ \lfunction[\trg]{\trg{g}}{\trg{x}}{\lifz[\trg]{\trg{\lhast{x}{\natt}}}{
+\\%
+                                                            &
+                                                            \left[\cca(\src{\expr})\right] %
                                                                                                  }{\labort[\trg]}}
@@
 \vspace{-1em}
-\begin{center}\small
+\begin{center}
+% \small
   $$

### 2023-10-11 esop/main.tex
Make sure text defines poison tags

--- 2023-10-09 esop/main.tex
@@
 
-$\Ltrg$ extends the syntax presented earlier (\Cref{subsec:cs:defs}) with dynamic typechecks ($\trg{\lhast{\expr}{\type}}$) and a way to inspect poison tags ($\trg{\lispoisoned{x}}$) in the metadata of pointers.
-For valid pointers ($\trg{\poisonless}$) bound to variable $\trg{x}$, the check $\trg{\lispoisoned{x}}$ yields $\trg{1}$.
-If the array bound to $\trg{x}$ was deallocated, i.e., has been poisoned ($\trg{\poisoned}$), the check $\trg{\lispoisoned{x}}$ evaluates to $\trg{0}$.
+$\Ltrg$ extends the syntax presented earlier (\Cref{subsec:cs:defs}) with dynamic typechecks ($\trg{\lhast{\expr}{\type}}$).
+The pointers carry poison tags, which signify a pointer being allocated ($\trg{\poisoned}$) or freed ($\trg{\poisonless}$). 
+The syntax of $\Ltrg$ is extended further with a way to inspect whether a pointer is invalid ($\trg{\lispoisoned{x}}$).
 Dynamic typechecks $\trg{\lhast{\expr}{\type}}$ match on $\trg{\expr}$ and evaluate to $\trg{0}$ if the term is of type $\trg{\type}$ and $\trg{1}$ otherwise.
-The projection $\msfilterL{\trg{\trace}}$ is equal to $\msfilterLtms{\src{\trace}}$.
+The projection $\msfilterL{\trg{\trace}}$ is structurally equal to $\msfilterLtms{\src{\trace}}$.
 

### 2023-10-11 esop/main.tex

--- 2023-10-11 esop/main.tex
@@
 
-$\Ltrg$ extends the syntax presented earlier (\Cref{subsec:cs:defs}) with dynamic typechecks ($\trg{\lhast{\expr}{\type}}$) and a way to inspect poison tags ($\trg{\lispoisoned{x}}$) in the metadata of pointers.
-For valid pointers ($\trg{\poisonless}$) bound to variable $\trg{x}$, the check $\trg{\lispoisoned{x}}$ yields $\trg{1}$.
-If the array bound to $\trg{x}$ was deallocated, i.e., has been poisoned ($\trg{\poisoned}$), the check $\trg{\lispoisoned{x}}$ evaluates to $\trg{0}$.
-Dynamic typechecks $\trg{\lhast{\expr}{\type}}$ match on $\trg{\expr}$ and evaluate to $\trg{0}$ if the term is of type $\trg{\type}$ and $\trg{1}$ otherwise.
-The projection $\msfilterL{\trg{\trace}}$ is equal to $\msfilterLtms{\src{\trace}}$.
-\MPin{
-	what is poison. do we need the symbol for valid pointers. 
-}
+$\Ltrg$ extends the syntax presented earlier (\Cref{subsec:cs:defs}) with dynamic typechecks ($\trg{\lhast{\expr}{\type}}$), evaluating to $\trg{0}$ if the type matches with the shape of $\trg{\expr}$ and $\trg{1}$ otherwise.
+%The pointers carry poison tags, which signify a pointer being allocated ($\trg{\poisoned}$) or freed ($\trg{\poisonless}$). 
+Furthermore, the syntax of $\Ltrg$ is extended with a way to inspect whether a pointer is freed ($\trg{\lispoisoned{x}}$), evaluating to $\trg{0}$ if it is freed and to $\trg{1}$ otherwise.
+The projection $\msfilterL{\trg{\trace}}$ is structurally equal to $\msfilterLtms{\src{\trace}}$.
 

### 2023-10-26 prisc-spectre/prisc.tex
Add prisc-spectre

%% For double-blind review submission, w/o CCS and ACM Reference (max submission space)
%\documentclass[sigplan,review,anonymous]{acmart}\settopmatter{printfolios=true,printccs=false,printacmref=false}
%% For double-blind review submission, w/ CCS and ACM Reference
%\documentclass[sigplan,review,anonymous]{acmart}\settopmatter{printfolios=true}
%% For single-blind review submission, w/o CCS and ACM Reference (max submission space)
%\documentclass[sigplan,review]{acmart}\settopmatter{printfolios=true,printccs=false,printacmref=false}
%% For single-blind review submission, w/ CCS and ACM Reference
%\documentclass[sigplan,review]{acmart}\settopmatter{printfolios=true}
%% For final camera-ready submission, w/ required CCS and ACM Reference
\documentclass[sigplan,dvipsnames]{acmart}\settopmatter{}
%% "not anonymous"
\settopmatter{printacmref=false} % Removes citation information below abstract
\renewcommand\footnotetextcopyrightpermission[1]{}

%% Conference information
%% Supplied to authors by publisher for camera-ready submission;
%% use defaults for review submission.
\acmConference[PriSC'23]{ACM SIGPLAN Workshop on Principles of Secure Compilation}{January 20nd, 2023}{London, United Kingdom}
\acmYear{2023}
%\acmISBN{} % \acmISBN{978-x-xxxx-xxxx-x/YY/MM}
%\acmDOI{} % \acmDOI{10.1145/nnnnnnn.nnnnnnn}
%\startPage{1}

%% Copyright information
%% Supplied to authors (based on authors' rights management selection;
%% see authors.acm.org) by publisher for camera-ready submission;
%% use 'none' for review submission.
\setcopyright{none}
%\setcopyright{acmcopyright}
%\setcopyright{acmlicensed}
%\setcopyright{rightsretained}
%\copyrightyear{2022}           %% If different from \acmYear

%% Bibliography style
\bibliographystyle{ACM-Reference-Format}
%% Citation style
%\citestyle{acmauthoryear}  %% For author/year citations
%\citestyle{acmnumeric}     %% For numeric citations
%\setcitestyle{nosort}      %% With 'acmnumeric', to disable automatic
                            %% sorting of references within a single citation;
                            %% e.g., \cite{Smith99,Carpenter05,Baker12}
                            %% rendered as [14,5,2] rather than [2,5,14].
%\setcitesyle{nocompress}   %% With 'acmnumeric', to disable automatic
                            %% compression of sequential references within a
                            %% single citation;
                            %% e.g., \cite{Baker12,Baker14,Baker16}
                            %% rendered as [2,3,4] rather than [2-4].

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%% Note: Authors migrating a paper from traditional SIGPLAN
%% proceedings format to PACMPL format must update the
%% '\documentclass' and topmatter commands above; see
%% 'acmart-pacmpl-template.tex'.
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\newcommand\hmmax{0}
\newcommand\bmmax{0}

%% Some recommended packages.
\usepackage[colorinlistoftodos]{todonotes}
\newcommand{\MK}[1]{\todo[color=orange!30]{TODO: #1}}
\newcommand{\MKin}[1]{\todo[inline,color=orange!30]{TODO: #1}}
\newcommand{\MP}[1]{\todo[color=blue!30]{TODO: #1}}
\newcommand{\MPin}[1]{\todo[inline,color=blue!30]{TODO: #1}}

\usepackage{booktabs}   %% For formal tables:
                        %% http://ctan.org/pkg/booktabs
\usepackage{subcaption} %% For complex figures with subfigures/subcaptions
                        %% http://ctan.org/pkg/subcaption
\usepackage{stmaryrd}
\usepackage{mathrsfs}
\usepackage{mathtools}
\usepackage{xcolor}
\usepackage{amsmath}
\usepackage{amsthm}
\usepackage{glossaries}
\usepackage{tikz}
\usepackage{cleveref}
\usepackage{listings}
\usepackage{mmmacros}
\usepackage[switch]{lineno}
%\modulolinenumbers[2]
\renewcommand{\linenumberfont}{\normalfont\bfseries\small\color{red}}

\usetikzlibrary{decorations,positioning,arrows}

\newacronym{rc}{RC}{Robust Compilation}
\newacronym{ltlp}{LTL-Past}{Linear Temporal Logic with Past}
\newacronym{imd}{IMD}{Index-masking Defense}
\newacronym{lfence}{LFENCE}{Insertion of \texttt{lfence}s}
\newacronym{slh}{SLH}{Speculative Load Hardening}
\newacronym{retpoline}{Retpoline}{Return Trampoline}
\newacronym{speccfi}{SpecCFI}{Speculative Control Flow Integrity}
\newacronym{msr}{MSR}{Set Model Specific Register Flags}
\newacronym{pi}{PI}{Process Isolation}
\newacronym{ss}{SS}{Speculative Safety}

\begin{document}

%% Title information
\title{Secure Composition of SPECTRE Variants}         %% [Short Title] is optional;
                                        %% when present, will be used in
                                        %% header instead of Full Title.
%\titlenote{with title note}             %% \titlenote is optional;
                                        %% can be repeated if necessary;
                                        %% contents suppressed with 'anonymous'
%\subtitle{Subtitle}                     %% \subtitle is optional
%\subtitlenote{with subtitle note}       %% \subtitlenote is optional;
                                        %% can be repeated if necessary;
                                        %% contents suppressed with 'anonymous'

%% Author information
%% Contents and number of authors suppressed with 'anonymous'.
%% Each author should be introduced by \author, followed by
%% \authornote (optional), \orcid (optional), \affiliation, and
%% \email.
%% An author may have multiple affiliations and/or emails; repeat the
%% appropriate command.
%% Many elements are not rendered, but should be provided for metadata
%% extraction tools.

%% Author with single affiliation.
\author{Matthis Kruse}
%\authornote{with author1 note}          %% \authornote is optional;
                                        %% can be repeated if necessary
\orcid{0000-0003-4062-9666}             %% \orcid is optional
\affiliation{
%  \position{Position1}
%  \department{Department1}              %% \department is recommended
  \institution{CISPA Helmholtz Center for Information Security}            %% \institution is required
%  \streetaddress{Street1 Address1}
%  \city{City1}
%  \state{State1}
%  \postcode{Post-Code1}
  \country{Germany}                    %% \country is recommended
}
\email{matthis.kruse@cispa.de}          %% \email is recommended

%% Author with two affiliations and emails.
\author{Michael Backes}
%\authornote{with author2 note}          %% \authornote is optional;
                                        %% can be repeated if necessary
%\orcid{0000-0003-3411-9678}             %% \orcid is optional
\affiliation{
%  \position{Position2a}
%  \department{Department2a}             %% \department is recommended
  \institution{CISPA Helmholtz Center for Information Security}            %% \institution is required
%  \streetaddress{Street2a Address2a}
%  \city{City2a}
%  \state{State2a}
%  \postcode{Post-Code2a}
  \country{Germany}                   %% \country is recommended
}
\email{director@cispa.de}         %% \email is recommended

%% Abstract
%% Note: \begin{abstract}...\end{abstract} environment must come
%% before \maketitle command
%\begin{abstract}
%Text of abstract \ldots.
%\end{abstract}

%% 2012 ACM Computing Classification System (CSS) concepts
%% Generate at 'http://dl.acm.org/ccs/ccs.cfm'.
%\begin{CCSXML}
%<ccs2012>
%<concept>
%<concept_id>10011007.10011006.10011008</concept_id>
%<concept_desc>Software and its engineering~General programming languages</concept_desc>
%<concept_significance>500</concept_significance>
%</concept>
%<concept>
%<concept_id>10003456.10003457.10003521.10003525</concept_id>
%<concept_desc>Social and professional topics~History of programming languages</concept_desc>
%<concept_significance>300</concept_significance>
%</concept>
%</ccs2012>
%\end{CCSXML}

%\ccsdesc[500]{Software and its engineering~General programming languages}
%\ccsdesc[300]{Social and professional topics~History of programming languages}
%% End of generated code

%% Keywords
%% comma separated list
%\keywords{compilers, security}  %% \keywords are mandatory in final camera-ready submission

%% \maketitle
%% Note: \maketitle command must come after title commands, author
%% commands, abstract environment, Computing Classification System
%% environment and commands, and keywords command.
\maketitle

\linenumbers

\section{Introduction}

Some programming languages of today enjoy strong security guarantees that programmers want to have, such as the memory-safety guarantees given by the semantic analysis pass of a compiler for the Rust programming lanugage.
While programmers may expect that these guarantees hold even after translating their program to the target programming language, it has been shown that this expectation is false in the general case~\MK{cit constant time examples}.
Even correct compilers do not provide satisfactory guarantees and one has to resort to secure compilers~\cite{abate19,abate20,patrignani21} which preserve property satisfaction even when the compiled program is linked with arbitrary target-level code.

Compilers that give these strong security guarantees usually instrument the code with mitigation mechanisms, such as bounds checks for (spatial) memory safety.
While such source-code instrumentations can be done in a local manner, i.e., there is one compilation pass that does just this, a compiler typically consists of several compilation passes, e.g., translating to continuation-passing-style or performing optimisations, like loop unrolling or vectorisation. 
From a software engineering perspective, these passes are tackled in an individual manner and there have been efforts to make compiler correctness proofs modular as well. 
A recent line of work\MK{cit} demonstrates what is necessary for secure compilers to enjoy this modularity, effectively enabling a divide-and-conquer approach to the engineering of secure compilers.

To prove individual compilation passes secure and compose them into one compiler, the authors, inspired by earlier work\MK{catalin}, give a semantic interpretation of the modifications done by a compilation pass. 
To instantiate their framework, one has to define a cross-language trace relation that witnesses the behavioral changes of a program when compiled.
It is worth noting that some compiler compositions are not useful in practice, such as one where the last compilation pass invalidates the security guarantees given by other passes, like a pass that inserts accesses to some arbitrary memory address without any checks, thereby violating memory-safety.
Whether a composition is useful or not can be captured by compatibility of the corresponding cross-language trace relation, thus bypassing the need to reason about the concrete, syntactic changes a compilation pass does.

\begin{definition}[Compatibility]
tbd
\MKin{todo}
\end{definition}

%Compatibility states that any source-level trace $\src{\trace}$ must have a corresponding target-level trace $\trg{\trace}$ such that the traces are related $\src{\trace}\sim\trg{\trace}$ and $\trg{\trace}$ satisfies a target-level property of interest.

Another prominent example for \textit{in}compatiblity is the \gls{imd}\MK{cit} that prevents SPECTREv1 attacks that exploit speculative execution of loads happening after a branch. 
\gls{imd} works by masking the index of the load with the size of the corresponding array.
Unfortunately, while this mitigation prevents SPECTREv1 attacks, it can introduce SPECTREv4 vulnerabilities, where loads/stores are run out-of-order and, thus, it can happen that the index is not masked in a subsequent, speculative load.
This demonstrates that a secure compilation pipeline that aims at instrumenting the source-code to be hardened against SPECTREv1 and SPECTREv4 has two options: (1) do not use \gls{imd} for SPECTREv1 mitigation and adopt another technique that is compatible with SPECTREv4 or (2) \textit{first} do \gls{imd}, then a source-code instrumentation to prevent SPECTREv4 vulnerabilities that does not assume that the original source-code satisfied absence of SPECTREv4 vulnerabilities\footnote{Of course, in both options, the mitigation for SPECTREv4 must not be incompatible with the one used for SPECTREv1.}.
For the latter, this paper refers to these kinds of compilation passes as (robust) \textit{enforcements}, because they are much stronger than (robust) preservations:

\begin{figure}[h]
  \small
\[
  \begin{array}{l|l}
    \textbf{Robust Preservation} & \textbf{Robust Enforcement} \\\hline
    \text{if }\src{p}\text{ robustly satisfies }\tilde{\sigma}(\trg{\pi}), & \\
    \text{then }\gamma_{\text{pres}}(\src{p})\text{ robustly satisfies }\trg{\pi} & \gamma_{\text{enf}}(\src{p})\text{ robustly satisfies }\trg{\pi}
  \end{array}
\]
\caption{Comparison of formal statements of robust preservation and robust enforcement. $\gamma_{\text{pres}}$ is a robustly preserving compiler and $\gamma_{\text{enf}}$ a robust enforcing compiler. $\trg{\pi}$ is an arbitrary target-level property. $\tilde{\sigma}$ is the universal image.}
\end{figure}
.\MK{cite ,,an extended account'' in caption}

Within the framework of recent work\MK{cit kruse et al (arxiv)}, this extended abstract aims to formalize the semantic effect of mitigation mechanisms for SPECTRE variants and investigate their compatibility with each other.

This work aims to investigate the compatibility, within the framework of recent work\MK{kruse et al}, of source-code instrumentation passes that insert SPECTRE mitigations. 
Moreover, the paper investigates not only security-preserving compilers, but also security-\textit{enforcing} compilers. 
Enforcements make no assumptions to the code of a source-level component\footnote{Although the source-level language may impose restrictions by means of a type-system that nonetheless provide security guarantees.}.
%% MK:
%% motivation lacking for why we do it for just SPECTRE
%% literally no explanation for spectre whatsoever

\section{SPECTREv\texttt{X} Mitigations}

In the following, it is assumed that the source-level traces $\src{\trace}$ are trivially absent of speculative leaks. 
This is an acceptable assumption to make, since higher-level programming languages typically do not have a semantics where speculation is considered. 
Consequently, if no speculation happens, no speculative leak can occur.
For the target-level traces $\trg{\trace}$, it is assumed that all vulnerabilities are in some sense expressible in the trace model of the target language. 
At the very least, this paper assumes the existence of some kind of allocation, memory load/store and branch events as well as a marker event for the beginning of a speculative execution, a barrie, and a rollback event. \MK{cite spectector, list concrete names in parentheses}
The following bulletlist describes the semantic effect of popular SPECTRE mitigation techniques and borrows the succinct notation from \gls{ltlp}\MK{cit}\footnote{$\square \equiv$ ,,always'', $\lozenge \equiv$ ,,eventually'', $\triangleright \equiv$ ,,next'', and $\triangleleft \equiv$ ,,previous''}.
Note that this paper only considers security tags $\trg{\sigma}$ in the target language, the source language is implicitly secure with respect to speculation, since it does not model any in its dynamic semantics.

\begin{itemize}
  \item \acrfull{imd} (\texttt{v1})
$$
\begin{array}{rcl}
  \src{\trace} \sim^{v1}_{\texttt{\tiny IMD}} \trg{\trace} & \equiv & \text{if }\src{Alloc(\loc;n)},\src{Get(\loc;idx;v)}\in\src{\trace} \\
                                                           && \text{then } \exists \trg{\loc\ n\ idx\ v}, \trg{Alloc(\loc;n)}\in\trg{\trace}\\
                                                           && \text{s.t. } \trg{Get(\loc;idx;v)}\in\trg{\trace} \\
                                                           && \text{and } \exists m, \trg{n} = 2^m \text{ and } \trg{idx} \le 2^m \\
                                                           && \text{and } \src{\loc}\approx\trg{\loc} \text{ and }\src{v}\approx\trg{v}
\end{array}
$$
  \gls{imd} changes memory allocation to be powers-of-2 and masks all indices by the bounds of the array.

  \item \gls{lfence} (\texttt{v1}, \texttt{v4})
$$
\begin{array}{rcl}
  \src{\trace} \sim^{v1,v4}_{\texttt{\tiny LFENCE}} \trg{\trace} & \equiv & \trg{\trace}\vDash\square(\trg{Branch(\_)}\implies\triangleright\trg{Barrier})\\
                                                                 && \text{ and }\trg{\trace}\vDash\square(\trg{Get(\_)}\implies\triangleleft\trg{Barrier})
\end{array}
$$
  \gls{lfence} simply puts a barrier after any branch or before any load instruction. 
  While the literature provides (partial) solutions that do not insert the barrier \textit{everywhere}, due to the significant performance penalty, this pass is simple, stupid and puts the barrier ,,everywhere'', i.e., in front of loads and after branches.

%   \item \gls{slh} (\texttt{v1})
% $$
% \src{\trace} \sim^{v1}_{\texttt{\tiny SLH}} \trg{\trace} \equiv ?
% $$

  \item \gls{retpoline} (\texttt{v2})
$$
\begin{array}{rcl}
  \src{\trace} \sim^{v2}_{\texttt{\tiny Retpoline}} \trg{\trace} & \equiv & \forall i, \src{\trace}[i]=\src{IndirectBranch(to)} \implies \\
                                                                 && \text{\color{red}todo} \\
                                                                 && \text{ and }\trg{\trace}\vDash\square(\trg{Get(\_)}\implies\triangleleft\trg{Barrier})
\end{array}
$$
The \gls{retpoline} applies for every indirect branch on the source-level trace. 
Each indirect branch at source-level must have an associated retpoline at target-level. 
That is, the address of the indirect call must be pushed onto the stack to be used in the return instruction and speculation busy waits.
\MKin{This is tricky to express.
There needs to be an abstract way to say "this source-level indirect branch corresponds to this sequence of events in the target trace"}

%   \item \gls{speccfi} (\texttt{v2}, \texttt{v5})
% $$
% \src{\trace} \sim^{v2}_{\texttt{\tiny SpecCFI}} \trg{\trace} \equiv ?
% $$

  \item \gls{msr} (\texttt{v2, v4, v5})
$$
\src{\trace} \sim^{v2,v4,v5}_{\texttt{\tiny MSR}} \trg{\trace} \equiv \forall\trg{\event^\sigma}, \trg{\sigma}=\trg{S}
$$
\MK{fix def to match intuition in text}
Modern processors have flags to turn off speculation features, resulting in complete absence of speculation (for these variants) and all taints being secure $\trg{S}$.

%   \item Hunter (v1, v4)
% $$
% \src{\trace} \sim^{all}_{\texttt{\tiny Hunter}} \trg{\trace} \equiv ?
% $$

%   \item Serberus (all versions)
% $$
% \src{\trace} \sim^{all}_{\texttt{\tiny SERBERUS}} \trg{\trace} \equiv ?
% $$
\end{itemize}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\section{Compatibility of SPECTREv\texttt{X} Mitigations}

The target-level property of interest (that is to be robustly preserved/enforced) would be a variant of \gls{ss}\MK{cit exorcising spectres}.
Contrary to the original definition of SS, \Cref{def:ss} does not require tags to be always of the shape $\trg{S}$, i.e., the ,,secure'' tag.

\begin{definition}[\gls{ss} for Variant $x$]\label{def:ss}
  \[
    \trg{\pi}_{\trg{vX}} = \left\{ \trg{\trace} \mid \forall \trg{\event^\sigma}\in\trg{\trace}, \trg{\sigma} \not= \trg{vX} \right\}
  \]
\end{definition}
Here, it is assumed that the trace model of the target language has taint tags $\trg{\sigma}$.
Instead, tags should be unequal to the kind of variant that one is interested in.
Due to this, it is possible to recover SS, hereby simply referred to as $\trg{\pi}_{ss}$:
$$\trg{\pi}_{ss} = \displaystyle\bigcap_{\trg{vX}}\trg{\pi}_{\trg{vX}}$$

\begin{figure}[h]
  \fbox{%
  \begin{tikzpicture}
    \node (imd) {\gls{imd}};
    \node[right=1.0 of imd] (lfence) {\gls{lfence}};

    \draw[->] (imd.east) to[bend left=20] (lfence.west);
  \end{tikzpicture}}
  \caption{Compatibility of source-code instrumentations to prevent attacks of individual or multiple SPECTRE-variants.
  Nodes are compilers that perform the respective source-code instrumentation. 
  Edges are directed and represent compatibility of the composition. 
  Tha origin of an edge is the compiler that should be run first, the target of an edge is the compiler that should be run afterwards.}
\end{figure}

\section{Conclusion}

This short paper sketched a real-world application of the compositionality framework for secure compilers as presented in earlier work\MK{cit}. 
While few source-code transformations have been formally proven secure, it is nonetheless interesting to investigate their compatiblity relationships with other transformations on a semantic level, pretending the passes have been proven secure.
Future work considers more source-code instrumentations, not only on just speculation. 
Most interestingly, the authors would like to investigate the compatiblity of common compiler optimisations.

\newpage

%% Bibliography
\bibliography{library}
% \bibliography{./../library}

%% Appendix
%\appendix
%\section{Appendix}

%Text of appendix \ldots

\end{document}

### 2023-10-29 prisc-spectre/prisc.tex
Minor changes to prisc

@@
 
-Some programming languages of today enjoy strong security guarantees that programmers want to have, such as the memory-safety guarantees given by the semantic analysis pass of a compiler for the Rust programming lanugage.
+Some programming languages of today enjoy strong security guarantees, such as the memory-safety guarantees given by the semantic analysis pass of a compiler for the Rust programming lanugage.
 While programmers may expect that these guarantees hold even after translating their program to the target programming language, it has been shown that this expectation is false in the general case~\MK{cit constant time examples}.
@@
 \begin{definition}[Compatibility]
-tbd
-\MKin{todo}
+  Given languages $\src{S}$ and $\trg{T}$, a cross-language trace relation $\sim$ between traces of $\src{S}$ and $\trg{T}$, and a $\trg{T}-$level class of properties $\trg{\mathbb{C}}$.
+  $\sim$ is compatible with $\trg{\mathbb{C}}$ iff for any $\trg{\Pi}\subseteq\trg{\mathbb{C}}$ it holds that $\tilde{\sigma}(\trg{\Pi})\subseteq\tilde{\sigma}(\trg{\mathbb{C}})$.
+
+%Definition gwfσ {S I : Language}
+%  (rel : Trace (Event S) -> Trace (Event I) -> Prop)
+%  (C : Class (Event I)) :=
+%    forall (Π : Hyperproperty (Event I)),
+%      belongs_to (Event I) Π C ->
+%      belongs_to (Event S) (σ~ rel Π) (σ~ rel C)
 \end{definition}
@@
 Within the framework of recent work\MK{cit kruse et al (arxiv)}, this extended abstract aims to formalize the semantic effect of mitigation mechanisms for SPECTRE variants and investigate their compatibility with each other.
-
-This work aims to investigate the compatibility, within the framework of recent work\MK{kruse et al}, of source-code instrumentation passes that insert SPECTRE mitigations. 
+%
+%
+%This work aims to investigate the compatibility, within the framework of recent work\MK{kruse et al}, of source-code instrumentation passes that insert SPECTRE mitigations. 
 Moreover, the paper investigates not only security-preserving compilers, but also security-\textit{enforcing} compilers. 

### 2023-10-31 prisc-spectre/prisc.tex
polish and fill out fig

@@
                         %% http://ctan.org/pkg/subcaption
+\usepackage{semantic}
 \usepackage{stmaryrd}
@@
 \newacronym{rc}{RC}{Robust Compilation}
-\newacronym{ltlp}{LTL-Past}{Linear Temporal Logic with Past}
+\newacronym{ltlp}{LTL+Past}{Linear Temporal Logic with Past}
 \newacronym{imd}{IMD}{Index-masking Defense}
@@
 From a software engineering perspective, these passes are tackled in an individual manner and there have been efforts to make compiler correctness proofs modular as well. 
-A recent line of work\MK{cit} demonstrates what is necessary for secure compilers to enjoy this modularity, effectively enabling a divide-and-conquer approach to the engineering of secure compilers.
+A recent line of work~\cite{kruse22,kruse23,abate21} demonstrates what is necessary for secure compilers to enjoy this modularity, effectively enabling a divide-and-conquer approach to the engineering of secure compilers.
 
-To prove individual compilation passes secure and compose them into one compiler, the authors, inspired by earlier work\MK{catalin}, give a semantic interpretation of the modifications done by a compilation pass.
+To prove individual compilation passes secure and compose them into one compiler, there is a semantic interpretation of the modifications done by a compilation pass~\cite{kruse23,abate21}. 
 To instantiate their framework, one has to define a cross-language trace relation that witnesses the behavioral changes of a program when compiled.
@@
 \begin{definition}[Compatibility]
-  Given languages $\src{S}$ and $\trg{T}$, a cross-language trace relation $\sim$ between traces of $\src{S}$ and $\trg{T}$, and a $\trg{T}-$level class of properties $\trg{\mathbb{C}}$.
-  $\sim$ is compatible with $\trg{\mathbb{C}}$ iff for any $\trg{\Pi}\subseteq\trg{\mathbb{C}}$ it holds that $\tilde{\sigma}(\trg{\Pi})\subseteq\tilde{\sigma}(\trg{\mathbb{C}})$.
+  Given languages $\src{S}$ and $\trg{T}$, a cross-language trace relation $\sim$ between traces of $\src{S}$ and $\trg{T}$, and a $\trg{T}-$level class of properties $\trg{\mathbb{C}}$, then $\sim$ is compatible with $\trg{\mathbb{C}}$ iff for any $\trg{\Pi}\subseteq\trg{\mathbb{C}}$ it holds that $\tilde{\sigma}(\trg{\Pi})\subseteq\tilde{\sigma}(\trg{\mathbb{C}})$.
 
@@
 
-Another prominent example for \textit{in}compatiblity is the \gls{imd}\MK{cit} that prevents SPECTREv1 attacks that exploit speculative execution of loads happening after a branch. 
-\gls{imd} works by masking the index of the load with the size of the corresponding array.
-Unfortunately, while this mitigation prevents SPECTREv1 attacks, it can introduce SPECTREv4 vulnerabilities, where loads/stores are run out-of-order and, thus, it can happen that the index is not masked in a subsequent, speculative load.
+Another prominent example for \textit{in}compatiblity is the \gls{imd}~\cite{pizlo18} that prevents SPECTREv1 attacks that exploit speculative execution of loads happening after a branch. 
+\gls{imd} works by masking the index of a load with the size of the corresponding memory region.
+Unfortunately, while this mitigation prevents SPECTREv1 attacks, it can introduce SPECTREv4 vulnerabilities, where loads/stores are run speculatively and, thus, may bypass the masking of the index.
 This demonstrates that a secure compilation pipeline that aims at instrumenting the source-code to be hardened against SPECTREv1 and SPECTREv4 has two options: (1) do not use \gls{imd} for SPECTREv1 mitigation and adopt another technique that is compatible with SPECTREv4 or (2) \textit{first} do \gls{imd}, then a source-code instrumentation to prevent SPECTREv4 vulnerabilities that does not assume that the original source-code satisfied absence of SPECTREv4 vulnerabilities\footnote{Of course, in both options, the mitigation for SPECTREv4 must not be incompatible with the one used for SPECTREv1.}.
-For the latter, this paper refers to these kinds of compilation passes as (robust) \textit{enforcements}, because they are much stronger than (robust) preservations:
+For the latter, this extended abstract refers to these kinds of compilation passes as (robust) \textit{enforcements}, because they are much stronger than (robust) preservations:
 
@@
 $\trg{\pi}$ is an arbitrary target-level property.
-$\tilde{\sigma}$ is the universal image.}
+$\tilde{\sigma}$ is taken from recent work~\cite{abate21}.}
 \end{figure}
-.\MK{cite ,,an extended account'' in caption}
 
-Within the framework of recent work\MK{cit kruse et al (arxiv)}, this extended abstract aims to formalize the semantic effect of mitigation mechanisms for SPECTRE variants and investigate their compatibility with each other.
+Within the framework of recent work~\cite{kruse22,kruse23}, this extended abstract aims to formalize the semantic effect of mitigation mechanisms for SPECTRE variants and investigate their compatibility with each other.
 %
@@
 Moreover, the paper investigates not only security-preserving compilers, but also security-\textit{enforcing} compilers. 
-Enforcements make no assumptions to the code of a source-level component\footnote{Although the source-level language may impose restrictions by means of a type-system that nonetheless provide security guarantees.}.
+Enforcements make no assumptions to the code of a source-level component\footnote{Although the source-level language may impose restrictions by means of a type-system that nonetheless provide security guarantees, since ill-typed programs are not compiled in practice.}.
 %% MK:
@@
 For the target-level traces $\trg{\trace}$, it is assumed that all vulnerabilities are in some sense expressible in the trace model of the target language. 
-At the very least, this paper assumes the existence of some kind of allocation, memory load/store and branch events as well as a marker event for the beginning of a speculative execution, a barrie, and a rollback event.
-\MK{cite spectector, list concrete names in parentheses}
-The following bulletlist describes the semantic effect of popular SPECTRE mitigation techniques and borrows the succinct notation from \gls{ltlp}\MK{cit}\footnote{$\square \equiv$ ,,always'', $\lozenge \equiv$ ,,eventually'', $\triangleright \equiv$ ,,next'', and $\triangleleft \equiv$ ,,previous''}.
+At the very least, this paper assumes the existence of some kind of allocation ($\trg{Alloc(\loc;n)}$), memory load/store ($\trg{Get(\loc;idx;n)}$ and $\trg{Set(\loc;idx;n;v)}$), branch ($\trg{If(b)}$), and indirect branch events ($\trg{Ibranch(v)}$) as well as a marker event for a barrier ($\trg{\times}$), and a rollback event ($\trg{Rlb}$)~\cite{patrignani21b}.
+The following sketches the semantic effect of popular SPECTRE mitigation techniques and borrows the succinct notation from \gls{ltlp}~\cite{markey03}\footnote{$\square \equiv$ ,,always'', $\lozenge \equiv$ ,,eventually'', $\triangleright \equiv$ ,,next'', and $\triangleleft \equiv$ ,,previous''}.
 Note that this paper only considers security tags $\trg{\sigma}$ in the target language, the source language is implicitly secure with respect to speculation, since it does not model any in its dynamic semantics.
@@
   \src{\trace} \sim^{v1}_{\texttt{\tiny IMD}} \trg{\trace} & \equiv & \text{if }\src{Alloc(\loc;n)},\src{Get(\loc;idx;v)}\in\src{\trace} \\
-                                                           && \text{then } \exists \trg{\loc\ n\ idx\ v}, \trg{Alloc(\loc;n)}\in\trg{\trace}\\
-                                                           && \text{s.t. } \trg{Get(\loc;idx;v)}\in\trg{\trace} \\
+                                                           && \text{then } \exists \trg{\loc\ n\ idx\ v}, \trg{Alloc(\loc;n)^\sigma}\in\trg{\trace}\\
+                                                           && \text{s.t. } \trg{Get(\loc;idx;v)^{\sigma'}}\in\trg{\trace} \\
                                                            && \text{and } \exists m, \trg{n} = 2^m \text{ and } \trg{idx} \le 2^m \\
@@
 $$
-  \gls{imd} changes memory allocation to be powers-of-2 and masks all indices by the bounds of the array.
+  \gls{imd} changes memory allocation to be powers-of-2 and masks all indices with the bounds of the array.
 
@@
 \begin{array}{rcl}
-  \src{\trace} \sim^{v1,v4}_{\texttt{\tiny LFENCE}} \trg{\trace} & \equiv & \trg{\trace}\vDash\square(\trg{Branch(\_)}\implies\triangleright\trg{Barrier})\\
-                                                                 && \text{ and }\trg{\trace}\vDash\square(\trg{Get(\_)}\implies\triangleleft\trg{Barrier})
+  \src{\trace} \sim^{v1,v4}_{\texttt{\tiny LFENCE}} \trg{\trace} & \equiv & \trg{\trace}\vDash\square(\trg{If(\_)^\sigma}\implies\triangleright\trg{\times^{\sigma'}})\\
+                                                                 && \text{ and }\trg{\trace}\vDash\square(\trg{Get(\_)^{\sigma''}}\implies\triangleleft\trg{\times^{\sigma'''}})
 \end{array}
@@
   While the literature provides (partial) solutions that do not insert the barrier \textit{everywhere}, due to the significant performance penalty, this pass is simple, stupid and puts the barrier ,,everywhere'', i.e., in front of loads and after branches.
+  Since the $\src{S}$-level trace is completely irrelevant, this is an example for an enforcement.
 
@@
   \item \gls{retpoline} (\texttt{v2})
+
+\begin{center}
+  \typerule{retpol-ibranch}{\src{v}\approx\trg{v} \\
+    \trg{\trace}=\trg{Set(\loc;sp;v)^\sigma}\cdot\trg{Ret^{\sigma'}}\cdot\trg{\overline{Jmp}^{\overline{\sigma''}}}\cdot\trg{Rlb}\cdot\trg{\trace'}
+  }{\src{Ibranch(v)}\cdot\src{\trace}\sim^{v2}_{\texttt{\tiny Retpoline}}\trg{\trace}}{retpol-ibranch}
+\end{center}
 $$
 \begin{array}{rcl}
-  \src{\trace} \sim^{v2}_{\texttt{\tiny Retpoline}} \trg{\trace} & \equiv & \forall i, \src{\trace}[i]=\src{IndirectBranch(to)} \implies \\
+  \src{\trace} \sim^{v2}_{\texttt{\tiny Retpoline}} \trg{\trace} & \equiv & \forall i, \src{\trace}[i]=\src{Ibranch(v)} \implies \\
                                                                  && \text{\color{red}todo} \\
-                                                                 && \text{ and }\trg{\trace}\vDash\square(\trg{Get(\_)}\implies\triangleleft\trg{Barrier})
 \end{array}
@@
 The \gls{retpoline} applies for every indirect branch on the source-level trace. 
-Each indirect branch at source-level must have an associated retpoline at target-level. 
-That is, the address of the indirect call must be pushed onto the stack to be used in the return instruction and speculation busy waits.
-\MKin{This is tricky to express.
-There needs to be an abstract way to say "this source-level indirect branch corresponds to this sequence of events in the target trace"}
+Each indirect branch at source-level must have an associated retpoline at target-level, as sketched with the \cref{tr:retpol-ibranch}. 
+That is, the address of the indirect call must be pushed onto the stack to be used in the return instruction and speculation busy waits until the rollback happens.
 
@@
 $$
-\MK{fix def to match intuition in text}
 Modern processors have flags to turn off speculation features, resulting in complete absence of speculation (for these variants) and all taints being secure $\trg{S}$.
+This is another example of an enforcement.
 
@@
 
-The target-level property of interest (that is to be robustly preserved/enforced) would be a variant of \gls{ss}\MK{cit exorcising spectres}.
+The target-level property of interest (that is to be robustly preserved/enforced) would be a variant of \gls{ss}~\cite{patrignani21b}.
 Contrary to the original definition of SS, \Cref{def:ss} does not require tags to be always of the shape $\trg{S}$, i.e., the ,,secure'' tag.
+Instead, tags should be unequal to the kind of variant that one is interested in:
 
-\begin{definition}[\gls{ss} for Variant $x$]\label{def:ss}
-  \[
+\begin{definition}[\gls{ss} for variant $\trg{vX}$]\label{def:ss}
+  \small$
     \trg{\pi}_{\trg{vX}} = \left\{ \trg{\trace} \mid \forall \trg{\event^\sigma}\in\trg{\trace}, \trg{\sigma} \not= \trg{vX} \right\}
-  \]
+  $
 \end{definition}
-Here, it is assumed that the trace model of the target language has taint tags $\trg{\sigma}$.
-Instead, tags should be unequal to the kind of variant that one is interested in.
-Due to this, it is possible to recover SS, hereby simply referred to as $\trg{\pi}_{ss}$:
+Due to this, it is possible to recover SS~\cite{patrignani21b}, hereby simply referred to as $\trg{\pi}_{ss}$:
 $$\trg{\pi}_{ss} = \displaystyle\bigcap_{\trg{vX}}\trg{\pi}_{\trg{vX}}$$
+
+\Cref{fig:rels} sketches the compatibility of SPECTRE mitigations given the semantic interpretation as presented in the previous section.
 
@@
     \node[right=1.0 of imd] (lfence) {\gls{lfence}};
+    \node[below=1.0 of imd] (retpoline) {\gls{retpoline}};
+    \node[right=1.0 of retpoline] (msr) {\gls{msr}};
 
-    \draw[->] (imd.east) to[bend left=20] (lfence.west);
+    \draw[->,very thick] (lfence) edge[loop right,right] node {$\trg{\pi_{v1,v4}}$} (lfence);
+    \draw[->,very thick] (msr) edge[loop right,right] node {$\trg{\pi_{v2,v4,v5}}$} (msr);
+    \draw[->,very thick] (imd) edge[loop left,left] node {$\trg{\pi_{v1}}$} (imd);
+    \draw[->,very thick] (retpoline) edge[loop left,left] node {$\trg{\pi_{v2}}$} (retpoline);
+
+    \draw[->,very thick] (imd.east) to[bend left=20,above] node {$\trg{\pi_{v1,v4}}$} (lfence.west);
+    \draw[->,very thick] (retpoline.east) to[bend right=20,below] node {$\trg{\pi_{v2,v4,v5}}$} (msr.west);
+    \draw[<->,very thick] (msr.north) to[bend right=20,right] node {$\trg{\pi}_{ss}$} (lfence.south);
+    \draw[<->,very thick] (retpoline.north) to[bend left=20,below right] node {$\trg{\pi_{v1,v2,v4}}$} (lfence.south);
+    \draw[<->,very thick] (msr.north) to[bend left=20,left] node {$\trg{\pi}_{ss}$} (imd.south);
+    \draw[<->,very thick] (retpoline.north) to[bend left=20,left] node {$\trg{\pi_{v1,v2}}$} (imd.south);
   \end{tikzpicture}}
@@
   Edges are directed and represent compatibility of the composition. 
-  Tha origin of an edge is the compiler that should be run first, the target of an edge is the compiler that should be run afterwards.}
+  The origin of an edge is the compiler that should be run first, the target of an edge is the compiler that should be run afterwards.
+  Edge labels indicate the \gls{ss} variants.}
+  \label{fig:rels}
 \end{figure}
@@
 
-This short paper sketched a real-world application of the compositionality framework for secure compilers as presented in earlier work\MK{cit}. 
+This short paper sketched a real-world application of the compositionality framework for secure compilers as presented in earlier work~\cite{abate21,kruse22,kruse23}. 
 While few source-code transformations have been formally proven secure, it is nonetheless interesting to investigate their compatiblity relationships with other transformations on a semantic level, pretending the passes have been proven secure.

### 2023-11-01 prisc-spectre/prisc.tex
nit

@@
 
-Some programming languages of today enjoy strong security guarantees, such as the memory-safety guarantees given by the semantic analysis pass of a compiler for the Rust programming lanugage.
+Some modern programming languages enjoy strong security guarantees, for example the Rust programming languages has memory safety guarantees given by its compiler performing a semantic analysis pass.
 While programmers may expect that these guarantees hold even after translating their program to the target programming language, it has been shown that this expectation is false in the general case~\MK{cit constant time examples}.
-Even correct compilers do not provide satisfactory guarantees and one has to resort to secure compilers~\cite{abate19,abate20,patrignani21} which preserve property satisfaction even when the compiled program is linked with arbitrary target-level code.
+Even correct compilers do not provide satisfactory guarantees and thus one has to resort to secure compilers~\cite{abate19,abate20,patrignani21} which preserve property satisfaction even when the compiled program is linked with arbitrary target-level code.
 
-Compilers that give these strong security guarantees usually instrument the code with mitigation mechanisms, such as bounds checks for (spatial) memory safety.
-While such source-code instrumentations can be done in a local manner, i.e., there is one compilation pass that does just this, a compiler typically consists of several compilation passes, e.g., translating to continuation-passing-style or performing optimisations, like loop unrolling or vectorisation.
-From a software engineering perspective, these passes are tackled in an individual manner and there have been efforts to make compiler correctness proofs modular as well. 
-A recent line of work~\cite{kruse22,kruse23,abate21} demonstrates what is necessary for secure compilers to enjoy this modularity, effectively enabling a divide-and-conquer approach to the engineering of secure compilers.
+Compilers that give these strong security guarantees usually instrument the code with mitigation mechanisms, for example they insert bounds checks for (spatial) memory safety.
+While such source-code instrumentations can be done in a local manner, i.e., there is one compilation pass that does just this, a compiler typically consists of several compilation passes, e.g., translating to continuation-passing-style or performing optimisations like loop unrolling.
+From a software engineering perspective, these passes are tackled in an individual manner, so there have been efforts to make compiler correctness proofs modular. 
+Recent work~\cite{kruse22,kruse23,abate21} demonstrates what is necessary for secure compilers to enjoy this modularity too, effectively enabling a divide-and-conquer approach to the engineering of secure compilers.
 
@@
 
-\section{Conclusion}
+% \section{Conclusion}
 
-This short paper sketched a real-world application of the compositionality framework for secure compilers as presented in earlier work~\cite{abate21,kruse22,kruse23}. 
-While few source-code transformations have been formally proven secure, it is nonetheless interesting to investigate their compatiblity relationships with other transformations on a semantic level, pretending the passes have been proven secure.
-Future work considers more source-code instrumentations, not only on just speculation. 
-Most interestingly, the authors would like to investigate the compatiblity of common compiler optimisations.
+% This short paper sketched a real-world application of the compositionality framework for secure compilers as presented in earlier work~\cite{abate21,kruse22,kruse23}. 
+% While few source-code transformations have been formally proven secure, it is nonetheless interesting to investigate their compatiblity relationships with other transformations on a semantic level, pretending the passes have been proven secure.
+% Future work considers more source-code instrumentations, not only on just speculation. 
+% Most interestingly, the authors would like to investigate the compatiblity of common compiler optimisations.
 

### 2023-11-02 prisc-spectre/prisc.tex
Merge
Incorporate some Feedback

@@
 %% Title information
-\title{Secure Composition of SPECTRE Variants}         %% [Short Title] is optional;
+\title{Secure Composition of SPECTRE Mitigations}         %% [Short Title] is optional;
                                         %% when present, will be used in
@@
 Some modern programming languages enjoy strong security guarantees, for example the Rust programming languages has memory safety guarantees given by its compiler performing a semantic analysis pass.
-While programmers may expect that these guarantees hold even after translating their program to the target programming language, it has been shown that this expectation is false in the general case~\MK{cit constant time examples}.
+While programmers may expect that these guarantees hold even after translating their program to the target programming language, it has been shown that this expectation is false in the general case~\cite{barthe18}.
 Even correct compilers do not provide satisfactory guarantees and thus one has to resort to secure compilers~\cite{abate19,abate20,patrignani21} which preserve property satisfaction even when the compiled program is linked with arbitrary target-level code.
@@
 
-To prove individual compilation passes secure and compose them into one compiler, there is a semantic interpretation of the modifications done by a compilation pass~\cite{kruse23,abate21}. 
-To instantiate their framework, one has to define a cross-language trace relation that witnesses the behavioral changes of a program when compiled.
-It is worth noting that some compiler compositions are not useful in practice, such as one where the last compilation pass invalidates the security guarantees given by other passes, like a pass that inserts accesses to some arbitrary memory address without any checks, thereby violating memory-safety.
-Whether a composition is useful or not can be captured by compatibility of the corresponding cross-language trace relation, thus bypassing the need to reason about the concrete, syntactic changes a compilation pass does.
+To prove individual compilation passes secure and compose them into one compiler, previous work introduces a framework for a semantic interpretation of the modifications done by a compilation pass~\cite{kruse23,abate21}.
+To instantiate this framework, one has to define a cross-language trace relation that witnesses the behavioral changes of a program when compiled.
+It is worth noting that some compiler compositions may not give the wanted security properties, such as one where the last compilation pass invalidates the security guarantees given by other passes, like a pass that inserts accesses to some arbitrary memory address without any checks, thereby violating memory-safety.
+Compilation passes that do not violate the security properties of earlier ones fulfill a notion of compatibility of a cross-language trace relation that describes the effect of the compiler. 
+Because compatiblity is defined on cross-language trace relations, there is no need to reason about the concrete, syntactic changes a compilation pass does.
 
@@
 
-Within the framework of recent work~\cite{kruse22,kruse23}, this extended abstract aims to formalize the semantic effect of mitigation mechanisms for SPECTRE variants and investigate their compatibility with each other.
+Within the framework of recent work~\cite{kruse22,kruse23}, this extended abstract aims to give cross-language trace relations that describe the semantic effect of mitigation mechanisms for SPECTRE variants and investigate their compatibility with each other.
 %
@@
 
-In the following, it is assumed that the source-level traces $\src{\trace}$ are trivially absent of speculative leaks. 
+The following assumes that the source-level traces $\src{\trace}$ are trivially absent of speculative leaks. 
 This is an acceptable assumption to make, since higher-level programming languages typically do not have a semantics where speculation is considered. 
@@
 For the target-level traces $\trg{\trace}$, it is assumed that all vulnerabilities are in some sense expressible in the trace model of the target language. 
-At the very least, this paper assumes the existence of some kind of allocation ($\trg{Alloc(\loc;n)}$), memory load/store ($\trg{Get(\loc;idx;n)}$ and $\trg{Set(\loc;idx;n;v)}$), branch ($\trg{If(b)}$), and indirect branch events ($\trg{Ibranch(v)}$) as well as a marker event for a barrier ($\trg{\times}$), and a rollback event ($\trg{Rlb}$)~\cite{patrignani21b}.
-The following sketches the semantic effect of popular SPECTRE mitigation techniques and borrows the succinct notation from \gls{ltlp}~\cite{markey03}\footnote{$\square \equiv$ ,,always'', $\lozenge \equiv$ ,,eventually'', $\triangleright \equiv$ ,,next'', and $\triangleleft \equiv$ ,,previous''}.
+This section also assumes the existence of some kind of allocation ($\trg{Alloc(\loc;n)}$), memory load/store ($\trg{Get(\loc;idx;n)}$ and $\trg{Set(\loc;idx;n;v)}$), branch ($\trg{If(b)}$), and indirect branch events ($\trg{Ibranch(v)}$) as well as a marker event for a barrier ($\trg{\times}$), and a rollback event ($\trg{Rlb}$)~\cite{patrignani21b}.
+The following sketches the cross-language trace relations describing the semantic effect of popular SPECTRE mitigation techniques.
 Note that this paper only considers security tags $\trg{\sigma}$ in the target language, the source language is implicitly secure with respect to speculation, since it does not model any in its dynamic semantics.
@@
 \begin{array}{rcl}
-  \src{\trace} \sim^{v1,v4}_{\texttt{\tiny LFENCE}} \trg{\trace} & \equiv & \trg{\trace}\vDash\square(\trg{If(\_)^\sigma}\implies\triangleright\trg{\times^{\sigma'}})\\
-                                                                 && \text{ and }\trg{\trace}\vDash\square(\trg{Get(\_)^{\sigma''}}\implies\triangleleft\trg{\times^{\sigma'''}})
+  \src{\trace} \sim^{v1,v4}_{\texttt{\tiny LFENCE}} \trg{\trace} & \equiv & (\text{if }\forall i,\trg{\trace}[i]=\trg{If(\_)^\sigma}\text{ then }\trg{\trace}[i+1]=\trg{\times^{\sigma'}})\\
+                                                                 && \text{and }(\text{if }\forall i>0,\trg{\trace}[i]=\trg{Get(\_)^{\sigma''}}\\
+                                                                 && \ \ \ \ \ \ \ \text{ then }\trg{\trace}[i-1]=\trg{\times^{\sigma'''}})
 \end{array}

### 2023-11-02 prisc-spectre/prisc.tex
Trampoline goes away
Feedback

@@
-%% For double-blind review submission, w/o CCS and ACM Reference (max submission space)
+%% For double-blind review submission, w/o CCS and ACM Reference (max submission spaceomp
 %\documentclass[sigplan,review,anonymous]{acmart}\settopmatter{printfolios=true,printccs=false,printacmref=false}
@@
 \newacronym{ltlp}{LTL+Past}{Linear Temporal Logic with Past}
-\newacronym{imd}{IMD}{Index-masking Defense}
+\newacronym{imd}{IMD}{Index-Masking Defense}
 \newacronym{lfence}{LFENCE}{Insertion of \texttt{lfence}s}
@@
 While programmers may expect that these guarantees hold even after translating their program to the target programming language, it has been shown that this expectation is false in the general case~\cite{barthe18}.
-Even correct compilers do not provide satisfactory guarantees and thus one has to resort to secure compilers~\cite{abate19,abate20,patrignani21} which preserve property satisfaction even when the compiled program is linked with arbitrary target-level code.
+Even correct compilers do not provide satisfactory guarantees and thus one has to resort to secure compilers~\cite{abate19,abate20,patrignani21} which preserve property satisfaction even when the compiled program is linked with arbitrary target-level code, i.e., the compiled program is robust.
 
@@
 From a software engineering perspective, these passes are tackled in an individual manner, so there have been efforts to make compiler correctness proofs modular. 
-Recent work~\cite{kruse22,kruse23,abate21} demonstrates what is necessary for secure compilers to enjoy this modularity too, effectively enabling a divide-and-conquer approach to the engineering of secure compilers.
+Recent work~\cite{kruse23} demonstrates what is necessary for secure compilers to enjoy this modularity too, effectively enabling a divide-and-conquer approach to the engineering of secure compilers.
 
@@
 
-\begin{definition}[Compatibility]
-  Given languages $\src{S}$ and $\trg{T}$, a cross-language trace relation $\sim$ between traces of $\src{S}$ and $\trg{T}$, and a $\trg{T}-$level class of properties $\trg{\mathbb{C}}$, then $\sim$ is compatible with $\trg{\mathbb{C}}$ iff for any $\trg{\Pi}\subseteq\trg{\mathbb{C}}$ it holds that $\tilde{\sigma}(\trg{\Pi})\subseteq\tilde{\sigma}(\trg{\mathbb{C}})$.
+\begin{definition}[Compatibility]\label{def:compat}
+  Given languages $\src{S}$ and $\trg{T}$, a cross-language trace relation $\sim$ between traces of $\src{S}$ and $\trg{T}$, and a $\trg{T}-$level collection of properties $\trg{\mathbb{C}}$, then $\sim$ is compatible with $\trg{\mathbb{C}}$ iff for any $\trg{\pi}\in\trg{\mathbb{C}}$ it holds that $\sigma_\sim(\trg{\pi})\in\sigma_\sim(\trg{\mathbb{C}})$.
 
@@
     \textbf{Robust Preservation} & \textbf{Robust Enforcement} \\\hline
-    \text{if }\src{p}\text{ robustly satisfies }\tilde{\sigma}(\trg{\pi}), & \\
+    \text{if }\src{p}\text{ robustly satisfies }\sigma_\sim(\trg{\pi}), & \\
     \text{then }\gamma_{\text{pres}}(\src{p})\text{ robustly satisfies }\trg{\pi} & \gamma_{\text{enf}}(\src{p})\text{ robustly satisfies }\trg{\pi}
@@
 $\trg{\pi}$ is an arbitrary target-level property.
-$\tilde{\sigma}$ is taken from recent work~\cite{abate21}.}
+$\sigma_\sim$ is taken from recent work~\cite{abate21}.}\label{fig:pres}
 \end{figure}
 
-Within the framework of recent work~\cite{kruse22,kruse23}, this extended abstract aims to give cross-language trace relations that describe the semantic effect of mitigation mechanisms for SPECTRE variants and investigate their compatibility with each other.
+Within the framework of recent work~\cite{kruse22,kruse23}, this extended abstract aims to give cross-language trace relations that describe the semantic effect of mitigation mechanisms for SPECTRE variants and sketch their compatibility with each other.
 %
@@
 Enforcements make no assumptions to the code of a source-level component\footnote{Although the source-level language may impose restrictions by means of a type-system that nonetheless provide security guarantees, since ill-typed programs are not compiled in practice.}.
-%% MK:
-%% motivation lacking for why we do it for just SPECTRE
-%% literally no explanation for spectre whatsoever
 
-\section{SPECTREv\texttt{X} Mitigations}
+\section{Composing Secure Compilers}
 
-The following assumes that the source-level traces $\src{\trace}$ are trivially absent of speculative leaks. 
-This is an acceptable assumption to make, since higher-level programming languages typically do not have a semantics where speculation is considered. 
-Consequently, if no speculation happens, no speculative leak can occur.
-For the target-level traces $\trg{\trace}$, it is assumed that all vulnerabilities are in some sense expressible in the trace model of the target language. 
-This section also assumes the existence of some kind of allocation ($\trg{Alloc(\loc;n)}$), memory load/store ($\trg{Get(\loc;idx;n)}$ and $\trg{Set(\loc;idx;n;v)}$), branch ($\trg{If(b)}$), and indirect branch events ($\trg{Ibranch(v)}$) as well as a marker event for a barrier ($\trg{\times}$), and a rollback event ($\trg{Rlb}$)~\cite{patrignani21b}.
-The following sketches the cross-language trace relations describing the semantic effect of popular SPECTRE mitigation techniques.
-Note that this paper only considers security tags $\trg{\sigma}$ in the target language, the source language is implicitly secure with respect to speculation, since it does not model any in its dynamic semantics.
+The composition of secure compilers requires two theorems to be proven:
+(1) either of the statements in \Cref{fig:pres} and (2) \Thmref{def:compat}.
+The rest of the paper assumes that, for the presented SPECTRE mitigations, the statements in \Cref{fig:pres} have been shown for the respective compilation pass. 
+The property that all mitigations aim to robustly preserve/enforce is a variant of \gls{ss}~\cite{patrignani21b}, which relies on a taint-tracking mechanism and taints ($\sigma$) that tag events as safe ($S$) or unsafe.
+Contrary to the original definition of \gls{ss}, this paper states \gls{ss} such that tags should be unequal to the tag of the kind of variant ($\text{vX}$) that one is interested in:
+
+\begin{definition}[\gls{ss} for variant $\trg{vX}$]\label{def:ss}
+  \small$
+    \trg{\pi}_{\trg{vX}} = \left\{ \trg{\trace} \mid \forall \trg{\event^\sigma}\in\trg{\trace}, \trg{\sigma} \not= \trg{vX} \right\}
+  $
+\end{definition}
+\noindent The original \gls{ss}~\cite{patrignani21b}, hereby named $\trg{\pi}_{ss}$, can be recovered:
+$$\trg{\pi}_{ss} = \displaystyle\bigcap_{\trg{vX}}\trg{\pi}_{\trg{vX}}$$
+
+Both robust preservation/enforcement (\Cref{fig:pres}) and compatiblity (\Cref{def:compat}) require a cross-language trace relation that describes the corresponding compilers semantically.
+Therefore, for the rest of the extended abstract, it is assumed that there are source ($\src{S}$) and target ($\trg{T}$) languages, which share a large portion of their trace model. 
+The trace models must have some kind of allocation ($\text{Alloc}{(\loc;n)}$), memory load/store ($\text{Get}{(\loc;idx;n)}$ and $\text{Set}{(\loc;idx;n;v)}$), branch ($\text{If}(b)$), and indirect branch events ($\text{Ibranch(v)}$), as well as a marker event for a barrier ($\times$), and a rollback event ($\text{Rlb}$)~\cite{patrignani21b}.
+Trace events are annotated with taint tags. 
+For sake of readability, trace events tagged with the secure tag (S) are written without the tag.
+This is enough setup to sketch the cross-language trace relations describing the semantic changes each considered mitigation does:
 
 \begin{itemize}
-  \item \acrfull{imd} (\texttt{v1})
+  \item \acrfull{imd}~\cite{pizlo18} (\texttt{v1})
 $$
@@
 
-  \item \gls{lfence} (\texttt{v1}, \texttt{v4})
+  \item \gls{lfence}~\cite{taram22} (\texttt{v1}, \texttt{v4})
 $$
@@
   \src{\trace} \sim^{v1,v4}_{\texttt{\tiny LFENCE}} \trg{\trace} & \equiv & (\text{if }\forall i,\trg{\trace}[i]=\trg{If(\_)^\sigma}\text{ then }\trg{\trace}[i+1]=\trg{\times^{\sigma'}})\\
-                                                                 && \text{and }(\text{if }\forall i>0,\trg{\trace}[i]=\trg{Get(\_)^{\sigma''}}\\
-                                                                 && \ \ \ \ \ \ \ \text{ then }\trg{\trace}[i-1]=\trg{\times^{\sigma'''}})
+                                                                 & \text{and}&(\text{if }\forall i>0,\trg{\trace}[i]=\trg{Get(\_)^{\sigma''}}\\
+                                                                 && \text{ then }\trg{\trace}[i-1]=\trg{\times^{\sigma'''}})
 \end{array}
@@
 
-  \item \gls{retpoline} (\texttt{v2})
+  \item \gls{retpoline}~\cite{turnerret} (\texttt{v2})
 
@@
 \end{center}
-$$
-\begin{array}{rcl}
-  \src{\trace} \sim^{v2}_{\texttt{\tiny Retpoline}} \trg{\trace} & \equiv & \forall i, \src{\trace}[i]=\src{Ibranch(v)} \implies \\
-                                                                 && \text{\color{red}todo} \\
-\end{array}
-$$
 The \gls{retpoline} applies for every indirect branch on the source-level trace. 
@@
 
-  \item \gls{msr} (\texttt{v2, v4, v5})
+  \item \gls{msr}~\cite{redhatmsr} (\texttt{\scriptsize v2, v4, v5})
 $$
@@
 
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-
-\section{Compatibility of SPECTREv\texttt{X} Mitigations}
-
-The target-level property of interest (that is to be robustly preserved/enforced) would be a variant of \gls{ss}~\cite{patrignani21b}.
-Contrary to the original definition of SS, \Cref{def:ss} does not require tags to be always of the shape $\trg{S}$, i.e., the ,,secure'' tag.
-Instead, tags should be unequal to the kind of variant that one is interested in:
-
-\begin{definition}[\gls{ss} for variant $\trg{vX}$]\label{def:ss}
-  \small$
-    \trg{\pi}_{\trg{vX}} = \left\{ \trg{\trace} \mid \forall \trg{\event^\sigma}\in\trg{\trace}, \trg{\sigma} \not= \trg{vX} \right\}
-  $
-\end{definition}
-Due to this, it is possible to recover SS~\cite{patrignani21b}, hereby simply referred to as $\trg{\pi}_{ss}$:
-$$\trg{\pi}_{ss} = \displaystyle\bigcap_{\trg{vX}}\trg{\pi}_{\trg{vX}}$$
-
-\Cref{fig:rels} sketches the compatibility of SPECTRE mitigations given the semantic interpretation as presented in the previous section.
+Having shown either theorem stated in \Cref{fig:pres} for each of the mitigations in the list above, it remains to show \Cref{def:compat}.
+Without a proof of \Cref{def:compat}, the composition of mitigations may not provide the security guarantees of interest.
+For space reasons, this extended abstract does not provide formal proof for all possible compositions of above mitigations, but at least sketches the anticipated proofs of compatibility theorems in \Cref{fig:rels}.
 
 \begin{figure}[h]
-  \fbox{%
   \begin{tikzpicture}
-    \node (imd) {\gls{imd}};
-    \node[right=1.0 of imd] (lfence) {\gls{lfence}};
-    \node[below=1.0 of imd] (retpoline) {\gls{retpoline}};
-    \node[right=1.0 of retpoline] (msr) {\gls{msr}};
+    \node (imd) [inner sep = 5pt]{\acrshort{imd}};
+    \node[right=1.5 of imd, inner sep = 5pt] (lfence) {\acrshort{lfence}};
+    \node[below=1.5 of imd, inner sep = 5pt] (retpoline) {\acrshort{retpoline}};
+    \node[right=1.5 of retpoline, inner sep = 5pt] (msr) {\acrshort{msr}};
 
-    \draw[->,very thick] (lfence) edge[loop right,right] node {$\trg{\pi_{v1,v4}}$} (lfence);
-    \draw[->,very thick] (msr) edge[loop right,right] node {$\trg{\pi_{v2,v4,v5}}$} (msr);
-    \draw[->,very thick] (imd) edge[loop left,left] node {$\trg{\pi_{v1}}$} (imd);
-    \draw[->,very thick] (retpoline) edge[loop left,left] node {$\trg{\pi_{v2}}$} (retpoline);
+    \draw[->,thick] (lfence) edge[loop right,right] node {$\trg{\pi_{v1,v4}}$} (lfence);
+    \draw[->,thick] (msr) edge[loop right,right] node {$\trg{\pi_{v2,v4,v5}}$} (msr);
+    \draw[->,thick] (imd) edge[loop left,left] node {$\trg{\pi_{v1}}$} (imd);
+    \draw[->,thick] (retpoline) edge[loop left,left] node {$\trg{\pi_{v2}}$} (retpoline);
 
-    \draw[->,very thick] (imd.east) to[bend left=20,above] node {$\trg{\pi_{v1,v4}}$} (lfence.west);
-    \draw[->,very thick] (retpoline.east) to[bend right=20,below] node {$\trg{\pi_{v2,v4,v5}}$} (msr.west);
-    \draw[<->,very thick] (msr.north) to[bend right=20,right] node {$\trg{\pi}_{ss}$} (lfence.south);
-    \draw[<->,very thick] (retpoline.north) to[bend left=20,below right] node {$\trg{\pi_{v1,v2,v4}}$} (lfence.south);
-    \draw[<->,very thick] (msr.north) to[bend left=20,left] node {$\trg{\pi}_{ss}$} (imd.south);
-    \draw[<->,very thick] (retpoline.north) to[bend left=20,left] node {$\trg{\pi_{v1,v2}}$} (imd.south);
-  \end{tikzpicture}}
+    \draw[->,thick] (imd) to[bend left=20,above] node[above] {$\trg{\pi_{v1,v4}}$} (lfence);
+    \draw[->,thick] (retpoline) to[bend right=20,below] node[below] {$\trg{\pi_{v2,v4,v5}}$} (msr);
+    \draw[<->,thick] (msr) to[bend right=20,right] node[right] {$\trg{\pi}_{ss}$} (lfence);
+    \draw[<->,thick] (retpoline) to[bend left=20,below right] node[below = 5.5pt, pos = .9, ] {$\trg{\pi_{v1,v2,v4}}$} (lfence);
+    \draw[<->,thick] (msr) to[bend left=20,left] node[above, pos=.15] {$\trg{\pi}_{ss}$} (imd);
+    \draw[<->,thick] (retpoline) to[bend left=20,left] node[left] {$\trg{\pi_{v1,v2}}$} (imd);
+  \end{tikzpicture}
   \caption{Compatibility of source-code instrumentations to prevent attacks of individual or multiple SPECTRE-variants.
-  Nodes are compilers that perform the respective source-code instrumentation. 
+  Nodes are mitigations that perform the respective source-code instrumentation. 
   Edges are directed and represent compatibility of the composition. 

### 2023-11-03 prisc-spectre/prisc.tex
Make another pass over PriSC submission

@@
 While programmers may expect that these guarantees hold even after translating their program to the target programming language, it has been shown that this expectation is false in the general case~\cite{barthe18}.
-Even correct compilers do not provide satisfactory guarantees and thus one has to resort to secure compilers~\cite{abate19,abate20,patrignani21} which preserve property satisfaction even when the compiled program is linked with arbitrary target-level code, i.e., the compiled program is robust.
 
-Compilers that give these strong security guarantees usually instrument the code with mitigation mechanisms, for example they insert bounds checks for (spatial) memory safety.
-While such source-code instrumentations can be done in a local manner, i.e., there is one compilation pass that does just this, a compiler typically consists of several compilation passes, e.g., translating to continuation-passing-style or performing optimisations like loop unrolling.
-From a software engineering perspective, these passes are tackled in an individual manner, so there have been efforts to make compiler correctness proofs modular. 
-Recent work~\cite{kruse23} demonstrates what is necessary for secure compilers to enjoy this modularity too, effectively enabling a divide-and-conquer approach to the engineering of secure compilers.
+Correct compilers do not necessarily provide satisfactory guarantees~\cite{barthe18} and thus one has to resort to secure compilers~\cite{abate19,abate20,patrignani21} which preserve property satisfaction even when the compiled program is linked with arbitrary target-level code, i.e., the compiled program is robust.
+A recent framework~\cite{kruse23} describes how to compose secure compilers, thus allowing a divide-and-conquer approach to the engineering of secure compilers. 
+This framework primarily focuses on compilers that do not change the traces of the original program.
+However, real-world compilers perform source-code transformations that may change the trace, such as applying source-code instrumentations that enhance security by, e.g., inserting bounds-checks. 
+Other work~\cite{abate21} showed that there are essentially two approaches to this, where the robust preservation~\cite{abate19,abate20,patrignani21} statement is changed to lift the restriction of a unified trace-model as follows:
+{
+  \footnotesize
+\[
+  \begin{array}{l|l}
+    \textbf{Top-down Approach} & \textbf{Bottom-up Approach} \\\hline
+     \text{if }\src{p}\text{ robustly satisfies }\src{\pi}, & \text{if }\src{p}\text{ robustly satisfies }\sigma_\sim(\trg{\pi}), \\
+     \text{then }\gamma(\src{p})\text{ robustly satisfies }\tau_\sim(\src{\pi}) & \text{then }\gamma(\src{p})\text{ robustly satisfies }\trg{\pi}
+  \end{array}
+\]
+}
+Hereby, $\gamma$ is a compiler, $\sim$ a cross-language trace relation from $\src{S}$-level traces to $\trg{T}$-level traces describing the semantic effect of $\gamma$, and $\tau_\sim$/$\sigma_\sim$ project a source/target property to a target/source property.
+While the compositionality framework~\cite{kruse23} does consider the top-down approach, it does not entail composition of the bottom-up one.
+However, the bottom-up approach is crucial for security properties that can only be described in the target-level, such as absence of SPECTRE vulnerabilities~\cite{kocher19}, since higher-level languages do not model speculation in their semantics.
 
-To prove individual compilation passes secure and compose them into one compiler, previous work introduces a framework for a semantic interpretation of the modifications done by a compilation pass~\cite{kruse23,abate21}.
-To instantiate this framework, one has to define a cross-language trace relation that witnesses the behavioral changes of a program when compiled.
-It is worth noting that some compiler compositions may not give the wanted security properties, such as one where the last compilation pass invalidates the security guarantees given by other passes, like a pass that inserts accesses to some arbitrary memory address without any checks, thereby violating memory-safety.
+It is worth noting that some compiler compositions may not give the wanted security properties, such as when composing with \gls{imd}~\cite{pizlo18} that prevents SPECTREv1 attacks that exploit speculative execution of loads happening after a branch.
+While \gls{imd} prevents SPECTREv1 attacks, it can introduce SPECTREv4 vulnerabilities and, thus, can render a SPECTREv4 mitigation run prior to \gls{imd} useless.
+
 Compilation passes that do not violate the security properties of earlier ones fulfill a notion of compatibility of a cross-language trace relation that describes the effect of the compiler. 
@@
 
-Another prominent example for \textit{in}compatiblity is the \gls{imd}~\cite{pizlo18} that prevents SPECTREv1 attacks that exploit speculative execution of loads happening after a branch. 
-\gls{imd} works by masking the index of a load with the size of the corresponding memory region.
-Unfortunately, while this mitigation prevents SPECTREv1 attacks, it can introduce SPECTREv4 vulnerabilities, where loads/stores are run speculatively and, thus, may bypass the masking of the index.
-This demonstrates that a secure compilation pipeline that aims at instrumenting the source-code to be hardened against SPECTREv1 and SPECTREv4 has two options: (1) do not use \gls{imd} for SPECTREv1 mitigation and adopt another technique that is compatible with SPECTREv4 or (2) \textit{first} do \gls{imd}, then a source-code instrumentation to prevent SPECTREv4 vulnerabilities that does not assume that the original source-code satisfied absence of SPECTREv4 vulnerabilities\footnote{Of course, in both options, the mitigation for SPECTREv4 must not be incompatible with the one used for SPECTREv1.}.
-For the latter, this extended abstract refers to these kinds of compilation passes as (robust) \textit{enforcements}, because they are much stronger than (robust) preservations:
-
-\begin{figure}[h]
-  \small
-\[
-  \begin{array}{l|l}
-    \textbf{Robust Preservation} & \textbf{Robust Enforcement} \\\hline
-    \text{if }\src{p}\text{ robustly satisfies }\sigma_\sim(\trg{\pi}), & \\
-    \text{then }\gamma_{\text{pres}}(\src{p})\text{ robustly satisfies }\trg{\pi} & \gamma_{\text{enf}}(\src{p})\text{ robustly satisfies }\trg{\pi}
-  \end{array}
-\]
-\caption{Comparison of formal statements of robust preservation and robust enforcement.
-$\gamma_{\text{pres}}$ is a robustly preserving compiler and $\gamma_{\text{enf}}$ a robust enforcing compiler.
-$\trg{\pi}$ is an arbitrary target-level property.
-$\sigma_\sim$ is taken from recent work~\cite{abate21}.}\label{fig:pres}
-\end{figure}
-
-Within the framework of recent work~\cite{kruse22,kruse23}, this extended abstract aims to give cross-language trace relations that describe the semantic effect of mitigation mechanisms for SPECTRE variants and sketch their compatibility with each other.
+This extended abstract extends prior work~\cite{kruse23} to consider bottom-up secure compiler composition and aims to apply that theory to a selection of mitigations for SPECTRE variants. 
+It is demonstrated that it suffices to setup a cross-language trace relation that describes the semantic effect of a secure compiler and prove compatibility with properties of interest in order to compose the secure compiler without giving up on security guarantees.
 %
@@
 %This work aims to investigate the compatibility, within the framework of recent work\MK{kruse et al}, of source-code instrumentation passes that insert SPECTRE mitigations.
-Moreover, the paper investigates not only security-preserving compilers, but also security-\textit{enforcing} compilers. 
-Enforcements make no assumptions to the code of a source-level component\footnote{Although the source-level language may impose restrictions by means of a type-system that nonetheless provide security guarantees, since ill-typed programs are not compiled in practice.}.
 
@@
 The composition of secure compilers requires two theorems to be proven:
-(1) either of the statements in \Cref{fig:pres} and (2) \Thmref{def:compat}.
-The rest of the paper assumes that, for the presented SPECTRE mitigations, the statements in \Cref{fig:pres} have been shown for the respective compilation pass. 
-The property that all mitigations aim to robustly preserve/enforce is a variant of \gls{ss}~\cite{patrignani21b}, which relies on a taint-tracking mechanism and taints ($\sigma$) that tag events as safe ($S$) or unsafe.
+(1) robust preservation, either with a unified trace-model~\cite{abate19}, top-down, or bottom-up, and (2) \Thmref{def:compat}.
+The rest of the paper assumes that the presented SPECTRE mitigations have been proven secure as in (1).
+The property that all mitigations aim to robustly preserve is a variant of \gls{ss}~\cite{patrignani21b}, which relies on a taint-tracking mechanism and taints ($\sigma$) that tag events as safe ($S$) or unsafe.
 Contrary to the original definition of \gls{ss}, this paper states \gls{ss} such that tags should be unequal to the tag of the kind of variant ($\text{vX}$) that one is interested in:
@@
 
-Both robust preservation/enforcement (\Cref{fig:pres}) and compatiblity (\Cref{def:compat}) require a cross-language trace relation that describes the corresponding compilers semantically.
+Robust preservation (only for top-down or bottom-up) and compatiblity (\Cref{def:compat}) require a cross-language trace relation that describes the effect of a corresponding compiler semantically.
 Therefore, for the rest of the extended abstract, it is assumed that there are source ($\src{S}$) and target ($\trg{T}$) languages, which share a large portion of their trace model. 
-The trace models must have some kind of allocation ($\text{Alloc}{(\loc;n)}$), memory load/store ($\text{Get}{(\loc;idx;n)}$ and $\text{Set}{(\loc;idx;n;v)}$), branch ($\text{If}(b)$), and indirect branch events ($\text{Ibranch(v)}$), as well as a marker event for a barrier ($\times$), and a rollback event ($\text{Rlb}$)~\cite{patrignani21b}.
-Trace events are annotated with taint tags. 
-For sake of readability, trace events tagged with the secure tag (S) are written without the tag.
+The trace models must have some kind of allocation ($\text{Alloc}{(\loc;n)}$), memory load/store ($\text{Get}{(\loc;idx;n)}$ and $\text{Set}{(\loc;idx;n;v)}$), branch ($\text{If}(b)$), and indirect branch events ($\text{Ibranch}(v)$), jumps ($\text{Jmp}(v)$), as well as a marker event for a barrier ($\times$), and a rollback event ($\text{Rlb}$)~\cite{patrignani21b}.
+Trace events are annotated with taint tags and for sake of readability, trace events tagged with the secure tag (S) are written without the tag.
 This is enough setup to sketch the cross-language trace relations describing the semantic changes each considered mitigation does:
@@
   \gls{lfence} simply puts a barrier after any branch or before any load instruction. 
-  While the literature provides (partial) solutions that do not insert the barrier \textit{everywhere}, due to the significant performance penalty, this pass is simple, stupid and puts the barrier ,,everywhere'', i.e., in front of loads and after branches.
+  While the literature provides (partial) solutions that do not insert the barrier \textit{everywhere}, due to the significant performance penalty, the considered pass is simple and puts the barrier ,,everywhere'', i.e., in front of loads and after branches.
   Since the $\src{S}$-level trace is completely irrelevant, this is an example for an enforcement.
@@
 \begin{center}
-  \typerule{retpol-ibranch}{\src{v}\approx\trg{v} \\
+  \typerule{retpol-ibranch}{\src{v}\approx\trg{v} &
     \trg{\trace}=\trg{Set(\loc;sp;v)^\sigma}\cdot\trg{Ret^{\sigma'}}\cdot\trg{\overline{Jmp}^{\overline{\sigma''}}}\cdot\trg{Rlb}\cdot\trg{\trace'}
@@
 $$
-\src{\trace} \sim^{v2,v4,v5}_{\texttt{\tiny MSR}} \trg{\trace} \equiv \forall\trg{\event^\sigma}, \trg{\sigma}=\trg{S}
+\src{\trace} \sim^{v2,v4,v5}_{\texttt{\tiny MSR}} \trg{\trace} \equiv \forall\trg{\event^\sigma}, \trg{\sigma}\not\in\{\trg{v2},\trg{v4},\trg{v5}\}
 $$
-Modern processors have flags to turn off speculation features, resulting in complete absence of speculation (for these variants) and all taints being secure $\trg{S}$.
+Modern processors have flags to turn off speculation features, resulting in complete absence of speculation (for these variants). 
 This is another example of an enforcement.
@@
 
-Having shown either theorem stated in \Cref{fig:pres} for each of the mitigations in the list above, it remains to show \Cref{def:compat}.
-Without a proof of \Cref{def:compat}, the composition of mitigations may not provide the security guarantees of interest.
+It remains to show \Thmref{def:compat}.
+Without a proof of \Thmref{def:compat}, the composition of mitigations may not provide the security guarantees of interest, since one could intuitively ,,undo'' what another one did.
 For space reasons, this extended abstract does not provide formal proof for all possible compositions of above mitigations, but at least sketches the anticipated proofs of compatibility theorems in \Cref{fig:rels}.

### 2023-11-03 prisc-spectre/prisc.tex
.

@@
 Without a proof of \Thmref{def:compat}, the composition of mitigations may not provide the security guarantees of interest, since one could intuitively ,,undo'' what another one did.
-For space reasons, this extended abstract does not provide formal proof for all possible compositions of above mitigations, but at least sketches the anticipated proofs of compatibility theorems in \Cref{fig:rels}.
+This extended abstract does not provide formal proof for all possible compositions of above mitigations, but sketches the anticipated proofs of compatibility theorems in \Cref{fig:rels}.
 

### 2023-12-06 esop/revs/rebuttal.md
Add draft for ESOP rebuttal

# Author Response for submission

We thank the reviewers for their comments and questions.
We address the latter below, first by giving some general comments that apply to all reviews and then one by one for each individual review.

Reviewer 1
==========

>  Don't you mean "*a* whole program"?   And do you really mean *a* trace there?  I'd expect all the traces of the program to have to satisfy π.

We are thankful for identifying this typo.

> Many languages have some kind of nondeterminism, e.g. to loosely specify the evaluation order.  That too has to be taken into account in the notion of correctness, and a priori it's quite unclear how one could specify that a compiler for such a language prevents information flow leakage via the resolution of those nondeterministic choices. Again I see no mention of this in the paper.

Section 4.3 presents extensions to the framework that allow to take undefinedness, non-determinism, resource exhaustion, and similar settings into account.
Here, it is crucial to find a suitable cross-language trace relation `~`.
Example instantiations of `~` are in section 1 in the paper `An Extended Account of Trace-relating Compiler Correctness and Secure Compilation` by Abate et al., 2021.

> 226 Is Use l n supposed to be an access to the first n bytes from l?  That's not realistic - many accesses will be to non-prefixes of allocations.

It's a random access of a single value at (intuitively) location `l + n`, where `l` is an abstract memory address.

> - why insist that there is a Dealloc?  A program that runs indefinitely using a persistent allocation does not intuitively have a temporal memory safety violation.

We agree that whether TMS covers existence of a Dealloc event is debatable.
However, this debate is orthogonal to the main goal of the work, since this enforcement of Dealloc events does not affect the usage of the compositionality framework at large.
However, it does complicate the case-study in the sense that the static semantics of `L_{tms}` enforces this requirement.

> - are locations l supposed to be concrete addresses, or unique allocation IDs, or combinations of the two?

Given the sandboxed setting, a location only makes sense together with an appropriate tag that witnesses the kind of heap this location points to.
Then, the location represents a concrete location in the respective heap.

> If they're concrete addresses, it's again weird, as any real implementation will reallocate at the same address.

Monotonic buffer allocators do not reallocate at the same address and are arguably the most simple allocator, yet very useful in practice, e.g., for hot loops that allocate each iteration.
We would like to point out that the concrete allocation strategy is orthogonal to the main line of work of this paper. 
The case-study is simple enough to keep the workload for the proofs not relevant to the grand-scheme of the compositionality framework, but relevant to the intricate technical setup to make use of it, small.
While an extension to support a more involved allocator - or even polymorphic memory allocators - does add a great amount of possibly interesting technical detail, the key theorems in the paper would not change at all from an observers perspective. (of course, the proof does change given the cross-language relations need to be patched accordingly)

> In all of this section, it's unclear where the security tag information is supposed to come from.  Are we supposed to imagine an annotated operational semantics that propagates sensitive tags?   If so, are control-flow choices on a sensitive value supposed to infect all later computation?
> [...]
> The idea that any access in compiled code can be uniquely associated to either the component or context also seems naive: in reality there'll be a nontrivial calling convention, with both caller-save and callee-save registers; how are those accesses supposed to be labelled?  Are we presuming that there's no inlining or other optimisation that crosses those boundaries?

The case-study of our work needs a bit of setup to fulfill the assumptions of the framework, e.g., that the compilers used in the case-study are secure.
This setup is proven with standard techniques and, thus, uninteresting for the sake of presentation of this work.

> 252 "sCCT may seem overly strict" - yes, it does.

We would like to point out - as done in the paper as well - that modern processors have a data-independent timing mode that, when turned on, aims to provide a guarantee that a certain set of instructions runs in constant-time.

> Do the monitors exactly characterize the properties?   You state only one direction of implication.  If not, why are they useful?

Focussing on the case-study, the secure compilation statements used here talk about traces satisfying a property by being an element of that property.
The established imlications allow to lift this to the monitor-level, where inductive reasoning is much more useful than on just traces.

Reviewer 2
==========

> The main weakness of this work is that the form of composition is limited. It requires the compilers (or compiler passes) to use the same trace model and are robustly safe, which limit the applicability of the results.

We kindly want to point out that section 4.3 discusses this and shows how to extend the framework to support variying trace-models between languages.

> The framework set up here may not be suitable for studying composition of other forms of properties (e.g., hyper-properties) or not robustly safe compiler composition.

The framework supports hyperproperties but the paper does not demonstrate this in the case-study since the proofs would be more involved.
Because of this, we overapproximate a well-known hyperproperty, i.e., cryptographic constant-time, using an ordinary trace property, i.e., strict cryptographic constant-time.
This is also a common technique for the verification of hyperproperties.

> Finally, the Coq formalization is not complete. It is unclear if there are fundamental difficulties in completing the proofs, or it is due to the limitation of time.

The main bulk of missing Coq proofs is with regards to the security of the compilers, each of which can easily take 20klocs per proof.
Because the development of these proofs is mostly standard, but unfortunately technically involved, we left them out in the interest of time: Proving these results in Coq does not give any additional insight to the key message of the paper, i.e., the compositionality framework.

Reviewer 3
==========

> I was particularly puzzled by the composition property.  Naively, it seemed straightforward to me that if a pass preserves A and B, then it preserves A; and if two passes preserve A, then their composition preserves A; given this, it seems straightforward that composition preserves (at least) the intersection of properties.  I'm sure there is something more to the results in the paper, but I wasn't entirely sure what; a clarification would be very helpful.

Your intuition is entirely correct and we agree that the results seems intuitive to expert readers.
Despite their intuitiveness, as the other reviewers acknowledge as well, the results are important.
The importance can be demonstrated in a concrete setting, where the framework is applied and the paper contains a case-study that does this.
Unfortunately, instantiations of the compositionality framework are non-trivial in the sense that secure compilation proofs need quite a lot of technical setup, as demonstrated by the case-study from the paper, which takes the vast majority of space in our supplementary material.

> I would in particular like to understand this better from the perspective of theorems 12 and 13.  What is really going on in these two, and can you sketch the proof?  It seems that the theorem here would say that /any/ function that is MS would, when composed with the lowering to Lscct, also be Lscct; is that the case?

The theorems do not talk about any function/compiler γ, but about concrete instances that have been set-up in the case-study.
But, most of their definitions are left out given their straightforwardness, e.g., the beginning of section 6.4 shows what the compiler has to do to enforce scct, but leaves out all the "recolorings" from yellow to red.

The proof for theorem 12 is a standard secure compilation proof that uses a trace-based backtranslation technique to get a source-level context that behaves similar to a given target-level adverserial.
The proof of theorem 12 is entirely similar in structure to theorems 6, 7, 9, and 10. 
As for theorem 13, this uses previous results from the compositionality framework and is, thus, more interesting from the perspective of your question.
With hopes to make it more readable, we decompose `γ^{L_{tms}}_{L_{scct}}` as follows:
 - `γA` is `γ^{L_{tms}}_{L}` 
 - `γB` is `γA` composed with `γ^{L}_{L_{ms}}` 
 - `γC` is `γB` composed with `γ_{CF}` 
 - `γD` is `γC` composed with `γ_{DCE}` 
 - `γE` is `γD` composed with `γ^{L_{ms}}_{L_{scct}}` 
The proof now works as follows:

 - Unfold the definition and simplify:
     - Let `π` be an element of `ms ∩ scct` and `p_{tms}` an `L_{tms}` component.
     - (H1) Assume that `p_{tms}` robustly satisifies `ms ∩ scct`.
     - Goal is `γE (p_{tms})` robustly satisfies `ms ∩ scct`.
- Apply Theorem 12 on the goal, changing it to `γD (p_{tms})` robustly satisfies `ms ∩ scct`.
- Apply Theorem 9 on the goal, changing it to `γC (p_{tms})` robustly satisfies `ms ∩ scct`.
- Apply Theorem 10 on the goal, changing it to `γB (p_{tms})` robustly satisfies `ms ∩ scct`.
- Apply Theorem 8 on the goal, changing it to `p_{tms}` robustly satisfies `ms ∩ scct`.
  -> solved by assumption (H1)

> The definition of strncpy is not the one I'm used to; the one I know pads `dst` with 0s.

We admit that the paper presents a more naive version of strncpy for sake of presentation.

### 2023-12-07 esop/revs/rebuttal.md
notes

@@
 We address the latter below, first by giving some general comments that apply to all reviews and then one by one for each individual review.
+
+==========
+TODO: where are the general comments ? 
+==========
+
+========================================================
+General Points
+========================================================
 
@@
 We are thankful for identifying this typo.
+To be clear -- besides the typo -- notion we are using is the canonical definition for program behaviour, as taken from Compcert and other similar work.
 
@@
 Example instantiations of `~` are in section 1 in the paper `An Extended Account of Trace-relating Compiler Correctness and Secure Compilation` by Abate et al., 2021.
+Thus, our framework supports reasoning about that kind of behaviour.
+We do not take that into account in our evaluation study because we focus on security properties, whose preservation does not rely on exploiting undefined behaviour.
 
@@
 We agree that whether TMS covers existence of a Dealloc event is debatable.
+========
+-TODO: is there existing work that uses this? i think our MSWasm paper perhaps
+========
 However, this debate is orthogonal to the main goal of the work, since this enforcement of Dealloc events does not affect the usage of the compositionality framework at large.
@@
 We would like to point out that the concrete allocation strategy is orthogonal to the main line of work of this paper. 
-The case-study is simple enough to keep the workload for the proofs not relevant to the grand-scheme of the compositionality framework, but relevant to the intricate technical setup to make use of it, small.
+The case-study is simple enough to keep the workload for the proofs not relevant to the grand-scheme of the compositionality framework, but relevant to the intricate technical setup to make use of it.
 While an extension to support a more involved allocator - or even polymorphic memory allocators - does add a great amount of possibly interesting technical detail, the key theorems in the paper would not change at all from an observers perspective.
-(of course, the proof does change given the cross-language relations need to be patched accordingly)
+(of course, the proof does change given the cross-language relations need to be patched accordingly, but existing work has already shown how to do this kind of complex proofs)
 
@@
 
-We would like to point out - as done in the paper as well - that modern processors have a data-independent timing mode that, when turned on, aims to provide a guarantee that a certain set of instructions runs in constant-time.
+We would like to point out - as done in the paper as well - that modern processors have a data-independent timing mode that, when turned on, aims to provide a guarantee that a certain set of instructions runs in constant-time, just like sCCT.
 
@@
 Focussing on the case-study, the secure compilation statements used here talk about traces satisfying a property by being an element of that property.
-The established imlications allow to lift this to the monitor-level, where inductive reasoning is much more useful than on just traces.
+The established implications allow to lift this to the monitor-level, where inductive reasoning is much more useful than on just traces, and thus the other direction of the implication is not necessary.
 
@@
 
-We kindly want to point out that section 4.3 discusses this and shows how to extend the framework to support variying trace-models between languages.
-
 > The framework set up here may not be suitable for studying composition of other forms of properties (e.g., hyper-properties) or not robustly safe compiler composition.
 
-The framework supports hyperproperties but the paper does not demonstrate this in the case-study since the proofs would be more involved.
-Because of this, we overapproximate a well-known hyperproperty, i.e., cryptographic constant-time, using an ordinary trace property, i.e., strict cryptographic constant-time.
+Wrt the first point, we'd like to to point out that section 4.3 discusses this and shows how to extend the framework to support variying trace-models between languages.
+
+The framework supports reasoning about the composition of compilers preserving hyperproperties but the paper does not demonstrate this in the case-study since the proofs would be more involved.
+Because of this, we overapproximate a well-known hyperproperty, i.e., cryptographic constant-time, using an ordinary trace property, i.e., strict cryptographic constant-time, which yields more manageable proofs.
 This is also a common technique for the verification of hyperproperties.
@@
 
-The main bulk of missing Coq proofs is with regards to the security of the compilers, each of which can easily take 20klocs per proof.
-Because the development of these proofs is mostly standard, but unfortunately technically involved, we left them out in the interest of time: Proving these results in Coq does not give any additional insight to the key message of the paper, i.e., the compositionality framework.
+==========
+TODO -> focus on the pros!
+==========
+The Coq proofs concern the main technical results (TODO Sections), and these results are all completely covered by the Coq development.
+What is missing in terms of Coq proofs is the secure compilation proofs, each of which can easily take 20klocs per proof, as demonstrated by the work of El-Korashi et al, CSF'22.
+We plan to study the modularisation of these proofs in future work, but since they follow a known pattern (albeit very complex to mechanise), we do not believe the lack of this mechanisation to be a significant drawback.
 
@@
 
-The proof for theorem 12 is a standard secure compilation proof that uses a trace-based backtranslation technique to get a source-level context that behaves similar to a given target-level adverserial.
+The proof for theorem 12 is a standard secure compilation proof that uses a trace-based backtranslation technique to get a source-level context that behaves similar to a given target-level adversary.
 The proof of theorem 12 is entirely similar in structure to theorems 6, 7, 9, and 10. 
@@
 We admit that the paper presents a more naive version of strncpy for sake of presentation.
+We will add a note for this.

### 2023-12-07 esop/revs/rebuttal.md
notes

@@
 We thank the reviewers for their comments and questions.
-We address the latter below, first by giving some general comments that apply to all reviews and then one by one for each individual review.
+Below, we address each individual review, addressing the questions for authors first, and then other questions raised in the reviews.
 
-==========
-TODO: where are the general comments ? 
-==========
+==========================================================================================
+Reviewer 1
+==========================================================================================
+---------------------------------------------------------------------------------
+---------------------------------------------------------------------------------
+> However, the basic setup seems to be highly idealized, ... 
+---------------------------------------------------------------------------------
+Reply:
+------
 
-========================================================
-General Points
-========================================================
+The setup of this paper is the same that many other work in correct and secure compilation use, namely a language with a notion of components, and with a trace semantics that is used to both express the security properties of interest as well as to aide the proofs.
+Some simplifications regard e.g., the memory model, but these are orthogonal to the applicability of the framework and just aim to simplify the (complex) secure compilation proofs.
+Since much of the language formalisation is not presented, we can add a clarification in the language section that the model is rather 'standard'.
 
-Reviewer 1
-==========
+---------------------------------------------------------------------------------
+---------------------------------------------------------------------------------
+> Instead, we get rather grand claims, e.g.
+> - "This paper is the first to study what security properties are preserved across the composition of different secure compiler passes."
+> - "From an engineering perspective, this is the desirable approach to building secure compilers."
+> which are not really well supported.
+---------------------------------------------------------------------------------
+Reply:
+------
 
+The claims we make are factually correct, it is the first to study composition of secure compilers.
+Moreover, given how compilers are built (i.e., sequences of passes), the theory we devise works with the current compiler construction approach.
+We use (some) real-world compiler passes to highlight the real-world applicability of our approach, but we are happy to add explanatory text as
+
+---------------------------------------------------------------------------------
+---------------------------------------------------------------------------------
 >  Don't you mean "*a* whole program"?   And do you really mean *a* trace there?  I'd expect all the traces of the program to have to satisfy π.
+---------------------------------------------------------------------------------
+Reply:
+------
 
@@
 
+---------------------------------------------------------------------------------
+---------------------------------------------------------------------------------
 > Many languages have some kind of nondeterminism, e.g.
@@
 Again I see no mention of this in the paper.
+---------------------------------------------------------------------------------
+Reply:
+------
 
@@
 
+---------------------------------------------------------------------------------
+---------------------------------------------------------------------------------
 > 226 Is Use l n supposed to be an access to the first n bytes from l?  That's not realistic - many accesses will be to non-prefixes of allocations.
+---------------------------------------------------------------------------------
+Reply:
+------
 
@@
 
+---------------------------------------------------------------------------------
+---------------------------------------------------------------------------------
 > - why insist that there is a Dealloc?  A program that runs indefinitely using a persistent allocation does not intuitively have a temporal memory safety violation.
+---------------------------------------------------------------------------------
+Reply:
+------
 
@@
 
+---------------------------------------------------------------------------------
+---------------------------------------------------------------------------------
 > - are locations l supposed to be concrete addresses, or unique allocation IDs, or combinations of the two?
+---------------------------------------------------------------------------------
+Reply:
+------
 
@@
 
+---------------------------------------------------------------------------------
+---------------------------------------------------------------------------------
 > If they're concrete addresses, it's again weird, as any real implementation will reallocate at the same address.
+---------------------------------------------------------------------------------
+Reply:
+------
 
@@
 
+---------------------------------------------------------------------------------
+---------------------------------------------------------------------------------
 > In all of this section, it's unclear where the security tag information is supposed to come from.
@@
 Are we presuming that there's no inlining or other optimisation that crosses those boundaries?
+---------------------------------------------------------------------------------
+Reply:
+------
 
@@
 
+---------------------------------------------------------------------------------
+---------------------------------------------------------------------------------
 > 252 "sCCT may seem overly strict" - yes, it does.
+---------------------------------------------------------------------------------
+Reply:
+------
 
@@
 
+---------------------------------------------------------------------------------
+---------------------------------------------------------------------------------
 > Do the monitors exactly characterize the properties?   You state only one direction of implication.  If not, why are they useful?
+---------------------------------------------------------------------------------
+Reply:
+------
 
@@
 
+==========================================================================================
 Reviewer 2
-==========
+==========================================================================================
 
+---------------------------------------------------------------------------------
+---------------------------------------------------------------------------------
 > The main weakness of this work is that the form of composition is limited.
 It requires the compilers (or compiler passes) to use the same trace model and are robustly safe, which limit the applicability of the results.
-
 > The framework set up here may not be suitable for studying composition of other forms of properties (e.g., hyper-properties) or not robustly safe compiler composition.
+---------------------------------------------------------------------------------
+Reply:
+------
 
 Wrt the first point, we'd like to to point out that section 4.3 discusses this and shows how to extend the framework to support variying trace-models between languages.
+Given the security relevance and applicability of the framework without the extension (as witnessed by the case study), we chose to investigate this extension from a case study perspective in future work.
 
-The framework supports reasoning about the composition of compilers preserving hyperproperties but the paper does not demonstrate this in the case-study since the proofs would be more involved.
+Nothing in the general compositionality theory we develop prevents considering compilers that preserve hyperproperties though.
+In fact, the framework supports reasoning about the composition of compilers preserving hyperproperties but the paper does not demonstrate this in the case-study since the proofs would be more involved.
 Because of this, we overapproximate a well-known hyperproperty, i.e., cryptographic constant-time, using an ordinary trace property, i.e., strict cryptographic constant-time, which yields more manageable proofs.
@@
 
+---------------------------------------------------------------------------------
+---------------------------------------------------------------------------------
 > Finally, the Coq formalization is not complete. It is unclear if there are fundamental difficulties in completing the proofs, or it is due to the limitation of time.
+---------------------------------------------------------------------------------
+Reply:
+------
 
-==========
-TODO -> focus on the pros!
-==========
 The Coq proofs concern the main technical results (TODO Sections), and these results are all completely covered by the Coq development.
@@
 
+==========================================================================================
 Reviewer 3
-==========
+==========================================================================================
 
+---------------------------------------------------------------------------------
+---------------------------------------------------------------------------------
 > I was particularly puzzled by the composition property.
@@
 I'm sure there is something more to the results in the paper, but I wasn't entirely sure what; a clarification would be very helpful.
+---------------------------------------------------------------------------------
+Reply:
+------
 
-Your intuition is entirely correct and we agree that the results seems intuitive to expert readers.
-Despite their intuitiveness, as the other reviewers acknowledge as well, the results are important.
-The importance can be demonstrated in a concrete setting, where the framework is applied and the paper contains a case-study that does this.
-Unfortunately, instantiations of the compositionality framework are non-trivial in the sense that secure compilation proofs need quite a lot of technical setup, as demonstrated by the case-study from the paper, which takes the vast majority of space in our supplementary material.
+Your intuition is entirely correct and we agree that the results can seem intuitive to expert readers.
+Despite their intuitiveness, as the other reviewers acknowledge as well, the results are important, and their importance can be demonstrated in a concrete setting, where the framework is applied and the paper contains a case-study that does this.
+Also, an intuition that may be harder to convey is: what does it mean to perform the intersection of 2 properties *in practice*?
+A reason we have to provide such a rich case study is also to give a better intuition of this fact.
+We can clarify this point in the related section.
 
+TODO: drop next sentence?
+From a technical standpoint, instantiations of the compositionality framework are non-trivial in the sense that secure compilation proofs need quite a lot of technical setup, as demonstrated by the case-study from the paper, which takes the vast majority of space in our supplementary material.
+
+---------------------------------------------------------------------------------
+---------------------------------------------------------------------------------
 > I would in particular like to understand this better from the perspective of theorems 12 and 13.
@@
 It seems that the theorem here would say that /any/ function that is MS would, when composed with the lowering to Lscct, also be Lscct; is that the case?
+---------------------------------------------------------------------------------
+Reply:
+------
 
@@
 
+---------------------------------------------------------------------------------
+---------------------------------------------------------------------------------
 > The definition of strncpy is not the one I'm used to; the one I know pads `dst` with 0s.
+---------------------------------------------------------------------------------
+Reply:
+------
 

### 2023-12-07 esop/revs/rebuttal.md
some annots

--- 2023-12-07 esop/revs/rebuttal.md
@@
 TODO: where are the general comments ? 
+  just get rid of general points?
 ==========
@@
 -TODO: is there existing work that uses this? i think our MSWasm paper perhaps
+               no, we may as well get rid of the requirement. But alas, we are able to show something more interesting...
 ========
@@
 TODO -> focus on the pros!
+   already done?
 ==========
-The Coq proofs concern the main technical results (TODO Sections), and these results are all completely covered by the Coq development.
-What is missing in terms of Coq proofs is the secure compilation proofs, each of which can easily take 20klocs per proof, as demonstrated by the work of El-Korashi et al, CSF'22.
+The Coq proofs concern the main technical results (section 4, parts of sections 5 and 6 as well as 3), and these results are all completely covered by the Coq development.
+What is missing in terms of Coq proofs is the secure compilation proofs, each of which can easily take 20klocs per proof, as demonstrated by the work of El-Korashy et al, CSF'22.
 We plan to study the modularisation of these proofs in future work, but since they follow a known pattern (albeit very complex to mechanise), we do not believe the lack of this mechanisation to be a significant drawback.

### 2023-12-07 esop/revs/rebuttal.md

--- 2023-12-07 esop/revs/rebuttal.md
@@
 We thank the reviewers for their comments and questions.
+
+==========
+TODO: where are the general comments ? 
+  just get rid of general points?
+==========
+
+========================================================
+General Points
+========================================================
+
 Below, we address each individual review, addressing the questions for authors first, and then other questions raised in the reviews.
@@
 -TODO: is there existing work that uses this? i think our MSWasm paper perhaps
+               no, the monitor in mswasm does not enforce this.
+               we may as well get rid of the requirement. But alas, we are able to show something more interesting...
 ========
@@
 
-The Coq proofs concern the main technical results (TODO Sections), and these results are all completely covered by the Coq development.
-What is missing in terms of Coq proofs is the secure compilation proofs, each of which can easily take 20klocs per proof, as demonstrated by the work of El-Korashi et al, CSF'22.
+TODO -> focus on the pros!
+   already done?
+The Coq proofs concern the main technical results (section 4, parts of sections 5 and 6 as well as 3), and these results are all completely covered by the Coq development.
+What is missing in terms of Coq proofs is the secure compilation proofs, each of which can easily take 20klocs per proof, as demonstrated by the work of El-Korashy et al, CSF'22.
 We plan to study the modularisation of these proofs in future work, but since they follow a known pattern (albeit very complex to mechanise), we do not believe the lack of this mechanisation to be a significant drawback.

### 2023-12-07 esop/revs/rebuttal.md
final ?

@@
 We thank the reviewers for their comments and questions.
-
-==========
-TODO: where are the general comments ? 
-  just get rid of general points?
-==========
-
-========================================================
-General Points
-========================================================
-
 Below, we address each individual review, addressing the questions for authors first, and then other questions raised in the reviews.
@@
 
-We agree that whether TMS covers existence of a Dealloc event is debatable.
-========
--TODO: is there existing work that uses this? i think our MSWasm paper perhaps
-               no, the monitor in mswasm does not enforce this.
-               we may as well get rid of the requirement. But alas, we are able to show something more interesting...
-========
-However, this debate is orthogonal to the main goal of the work, since this enforcement of Dealloc events does not affect the usage of the compositionality framework at large.
-However, it does complicate the case-study in the sense that the static semantics of `L_{tms}` enforces this requirement.
+Temporal memory safety is concerned with use-after-free and double-frees, therefore not having a free (Dealloc) event renders the TMS argument void.
+We choose to consider the Dealloc event in our trace model -- as other work has done before, e.g., the work of Michael et al in POPL23 -- because we want a meaningful version of TMS.
+Also, note that the Dealloc argument is orthogonal to the main goal of the work, since the enforcement of Dealloc events does not affect the usage of the compositionality framework at large.
 
@@
 
-TODO -> focus on the pros!
-   already done?
 The Coq proofs concern the main technical results (section 4, parts of sections 5 and 6 as well as 3), and these results are all completely covered by the Coq development.
@@
 We can clarify this point in the related section.
-
-TODO: drop next sentence?
-From a technical standpoint, instantiations of the compositionality framework are non-trivial in the sense that secure compilation proofs need quite a lot of technical setup, as demonstrated by the case-study from the paper, which takes the vast majority of space in our supplementary material.
 

### 2023-12-07 esop/revs/rebuttal.md
nits

@@
 Moreover, given how compilers are built (i.e., sequences of passes), the theory we devise works with the current compiler construction approach.
-We use (some) real-world compiler passes to highlight the real-world applicability of our approach, but we are happy to add explanatory text as
+We use (some) real-world compiler passes to highlight the real-world applicability of our approach, but we are happy to add explanatory text to support our claims.
 
@@
 Thus, our framework supports reasoning about that kind of behaviour.
-We do not take that into account in our evaluation study because we focus on security properties, whose preservation does not rely on exploiting undefined behaviour.
+We do not take that into account in our evaluation study because we focus on security properties whose preservation does not rely on exploiting undefined behaviour.
 

### 2023-12-08 esop/revs/rebuttal.md
some annots

@@
 We are thankful for identifying this typo.
-To be clear -- besides the typo -- notion we are using is the canonical definition for program behaviour, as taken from Compcert and other similar work.
+To be clear -- besides the typo -- the notion we are using is the canonical definition for program behaviour, as taken from Compcert and other similar work.
 

### 2024-01-04 csf/main.tex
notes

@@
 
-\section{Background: Security Properties and Secure Compilers\pages{1}}\label{sec:background}
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+
+\section{Background: Security Properties \\and Secure Compilers\pages{1}}\label{sec:background}
 
@@
 % \end{exampleenv}
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+\MPin{
+  moved to get a feeling of what's what
+}
+
+\section{Composing Secure Compilers\pages{1}}\label{sec:compcomp}
+
+This section presents the key meta-theoretic results of this paper concerning secure compiler compositions (\Cref{sec:sequential,sec:other-compos}).
+% Note that the presented framework works with arbitrary classes.
+% \MP{check, we ened more than this}
+% and builds on prior work~\cite{kruse2022csc}\MK{mention this at all?
+may reveal identity even though its double-blind}.
+% MP: not needed at all, ever, workshop work does not need to be cited in the main work
+Note that since it deals with the preservation of \emph{classes}, which include both trace properties as well as hyperproperties, the presented framework is general: It describes robust compilers preserving ordinary trace properties, hyperproperties or even relational hyperproperties.
+% \MP{pull this out to earlier}
+
+\subsection{Secure Sequential Composition}\label{sec:sequential}
+
+Our main result is that secure compilers in the robust compilation framework~\cite{abate2019jour} compose {\em sequentially}.
+% This is not intuitive in the sense that in the security domain, composition does not work without additional generalizations~\cite{canetti2006univcomp,mccullough2012compo,fabian2022automatic}.
+% \MP{what deos this mean?}
+The \bul{sequential composition of compilers $\cc{\src{L}}{\trg{L}}$ and $\cc{\trg{L}}{\obj{L}}$} is defined as follows:
+Given an $\src{L}$ program $\src{p}$ and compilers $\cc{\src{L}}{\trg{L}},\cc{\trg{L}}{\obj{L}}$, \rul{its compiled $\obj{L}$ counterpart is obtained by plugging $\src{p}$ into $\cc{\src{L}}{\trg{L}}\circ\cc{\trg{L}}{\obj{L}}$}.
+\begin{definition}[Sequential Composition]
+  \bul{$\cc{\src{L}}{\trg{L}}\circ\cc{\trg{L}}{\obj{L}}$} $\isdef$ Given $\src{p}$, yield \rul{$\cc{\trg{L}}{\obj{L}}\left(\cc{\src{L}}{\trg{L}}(\src{p})\right)$}
+\end{definition}
+
+Consider the compilation chain for $\src{TypeScript}$.
+First, $\src{TypeScript}$ programs are translated to $\trg{JavaScript}$ which, e.g., V8~\cite{googlev8} eventually compiles in parts to $\irl{Ignition BC}$.
+The following theorem establishes what happens if all these compilation steps were robustly secure with respect to \gls*{ms}:
+The resulting $\irl{Ignition BC}$ code would be \gls*{ms} regardless of the context the binary runs in.
+
+Given \bul{$\cc{\src{L}}{\trg{L}}$ robustly preserves $\class[_{1}]$} and \rul{$\cc{\trg{L}}{\obj{L}}$ robustly preserves $\class[_{2}]$}, it follows that \oul{their sequential composition $\cc{\src{L}}{\trg{L}}\circ\cc{\trg{L}}{\obj{L}}$ robustly preserves the intersection of classes $\class[_{1}]$ and $\class[_{2}]$}.
+
+\begin{theorem}[Sequential Composition of Secure Compilers]\label{thm:rtp}
+  If \bul{$\rtp{\cc{\src{L}}{\trg{L}}}{\class[_{1}]}$} and \rul{$\rtp{\cc{\trg{L}}{\obj{L}}}{\class[_{2}]}$}, then \oul{$\rtp{\cc{\src{L}}{\trg{L}}\circ\cc{\trg{L}}{\obj{L}}}{\class[_{1}]\cap\class[_{2}]}$}.
+\Coqed
+\end{theorem}
+
+Since the composition of secure compilers is again a secure compiler, the theorem generalises to a whole chain of $n$ secure compilers.
+
+\subsubsection{Securing Optimisations}
+Notably, real-world compilation chains also perform a series of (sequential) passes whose main purpose is not necessarily to translate from one language to another, but to, e.g., optimise the code or enforce a certain property.
+Both examples can be seen in practice, e.g. as in the work of~\cite{nagarakatte2009soft,nagarakatte2010cets,akritidis2009baggy,wegman1991ccp,manjikian1997fusion} and many more.
+% \MPin{survey to cite?
+this is a bit informal MK: hard to find a survey on different compiler passes...
+I can't even find one just for compiler enforced memsafety (that looks good) or for compiler translations in general}.
+% 
+% Consider the compilation passes of LLVM~\cite{lattner2004llvm}, which are handled in the \texttt{PassManager} class.
+% By manipulating the internal state of a \texttt{PassManager} object, the schedule of compiler optimisation passes can be hardcoded.
+% That way, an order of optimisations can be found by hand that yields the best results.
+% \MP{what's the point of the previous 3 sentences?(commented)}
+Consider the following two LLVM optimisation passes: \gls*{cf}, which rewrites constant expressions to the constant they evaluate to, and \gls*{dce}, which removes dead code by rewriting conditional branches.
+The order in which \gls*{cf} and \gls*{dce} are performed influences the final result of the compilation (see \Cref{fig:cfdceex}).
+\begin{figure}[!ht]
+  \begin{center}
+    \begin{tikzpicture}[scale=0.62, every node/.style={transform shape}]
+      \node[minimum width=3cm,minimum height=2.5cm,draw=black,very thick,align=left] (progunopt) {\begin{lstlisting}[language=ml,basicstyle=\small\ttfamily]
+let a = true in
+if a then
+  print "a"
+else
+  print "b"
+\end{lstlisting}};
+      \node[minimum width=3cm,minimum height=2.5cm,draw=black,very thick,align=left,right=1.0 of progunopt] (progoptcf) {\begin{lstlisting}[language=ml,basicstyle=\large\ttfamily]
+if true then
+  print "a"
+else
+  print "b"
+\end{lstlisting}};
+      \node[minimum width=3cm,minimum height=2.5cm,draw=black,very thick,align=left,right=1.0 of progoptcf] (progoptcfdce) {\begin{lstlisting}[language=ml,basicstyle=\Large\ttfamily]
+print "a"
+\end{lstlisting}};% the following is a hack to get my syntax highlighting to work again \end{lstlisting}
+      % edges
+      \draw[->,very thick] (progunopt) edge[loop left,left] node {\gls*{dce}} (progunopt);
+      \draw[->,very thick] (progunopt) edge[sloped,above] node {\gls*{cf}} (progoptcf);
+      \draw[->,very thick] (progoptcf) edge[sloped,above] node {\gls*{dce}} (progoptcfdce);
+    \end{tikzpicture}
+  \end{center}
+  \caption{Example program where the level of optimisations differ for one pass of applying \gls*{cf} and \gls*{dce} in any order. %
+  Every edge is a compilation pass and the label on the edge states what the pass does, i.e., \gls*{cf} or \gls*{dce}. %
+  The source code in the nodes is a glorified compiler intermediate representation and the code gets more optimised towards the right hand side of the figure.}\label{fig:cfdceex}
+\end{figure}
+This {\em phase ordering problem} is well--known in literature and a practical solution is to simply perform a fixpoint iteration of the optimisation pipeline~\cite{click1995combining}.
+%If possible, optimisations are merged to enhance the analysis' result, since even when iterating to a fixpoint the final program may not be as optimised as it could be~\cite{click1995combining}.
+Compiler engineers typically try to find an order of optimisations that yields well-optimised programs for either code size~\cite{cooper1999geneticphases} or performance~\cite{kulkarni2006exhaustivephase}.
+\Cref{corr:swappable} justifies that any such order of compilation passes is valid with respect to security.
+So, given two compilation passes \bul{$\cc[_{1}]{\trg{L}}{\trg{L}}$, $\cc[_{2}]{\trg{L}}{\trg{L}}$, both robustly preserving class $\class[_{1}]$ or $\class[_{2}]$, respectively}, \oul{for any order of their composition the composed compiler robustly preserves the intersection of $\class[_{1}]$ and $\class[_{2}]$}.
+% 
+% \MPin{
+%   later on, say that optimisation passes remove elements from traces, so they should be fine with respect to the result with $\tau$.
+%   also, yes, if $\tau$ adds we have a problem, if it removes we don't
+% \MKin{ Optimisations that remove elements from traces are not necessarily sound,
+%  e.g., remove all deallocations to "optimise" memory management by offloading
+%  the work to the OS.
+% }
+% }
+
+\begin{corollary}[Sequential Composition of Secure Compilers]\label{corr:swappable}
+  If \bul{$\rtp{\cc[_{1}]{\trg{L}}{\trg{L}}}{\class[_{1}]}$ and $\rtp{\cc[_{2}]{\trg{L}}{\trg{L}}}{\class[_{2}]}$}, then \oul{$\rtp{\cc[_{1}]{\trg{L}}{\trg{L}}\circ\cc[_{2}]{\trg{L}}{\trg{L}}}{\class[_{1}]\cap\class[_{2}]}$ and $\rtp{\cc[_{2}]{\trg{L}}{\trg{L}}\circ\cc[_{1}]{\trg{L}}{\trg{L}}}{\class[_{2}]\cap\class[_{1}]}$}.
+\Coqed
+\end{corollary}
+
+\subsection{Secure Upper and Lower Composition}\label{sec:other-compos}
+Besides sequential composition, there are two other compositions, namely an {\em upper}, i.e., a compiler that takes multiple inputs and yields one output, and a {\em lower} composition, i.e., a compiler that takes one input and yields multiple outputs.
+\bul{Define the upper composition $\cc{\src{L}+\obj{L}}{\trg{L}}$} as follows:
+Given a program \texttt{p}, its compiled counterpart is obtained by \rul{plugging \texttt{p} into $\cc{\src{L}}{\trg{L}}$ if $\texttt{p}\in\src{L}$} or by \oul{plugging \texttt{p} into $\cc{\obj{L}}{\trg{L}}$ if $\texttt{p}\in\obj{L}$}.
+\begin{definition}[Upper Composition]
+  $$\text{\bul{$\cc{\src{L}+\obj{L}}{\trg{L}}$}}\isdef
+  % Given \texttt{p}, yield
+  \lambda \texttt{p}\ldotp
+  \left\{\mbox{\begin{tabular}{c}
+    {\rul{if $\texttt{p}\in\src{L}$, then $\cc{\src{L}}{\trg{L}}(\texttt{p})$}} \\
+    \mbox{\oul{if $\texttt{p}\in\obj{L}$, then $\cc{\obj{L}}{\trg{L}}(\texttt{p})$}} \\
+  \end{tabular}}\right.$$
+%
+%
+%  $\begin{cases}
+%                                                                            \rul{\cc{\src{L}}{\trg{L}}(\texttt{p}) & \text{if }\texttt{p}\in\src{L}} \\
+%                                                                            \cc{\obj{L}}{\trg{L}}(\texttt{p}) & \text{if }\texttt{p}\in\obj{L}
+%                                                                          \end{cases}$
+\end{definition}
+
+Examples of this are present in industry:
+Consider the Java Virtual Machine bytecode $\trg{JVM BC}$, which is a popular target for programming language designers due to its high performance and relevance in industry.
+Compilers for several programming languages have it as their target language, some popular instances are $\src{Java}$ and $\obj{Kotlin}$.
+Technically speaking, they both compile to class files and $\obj{Kotlin}$ objects are considered to be the same as $\src{Java}$ objects at that point.
+Both languages can be used at the same time in one project~\cite{androidstudio}.
+A compiler that accepts both $\src{Java}$ and $\obj{Kotlin}$ code translating to the same target language or intermediate representation performs a kind of {\em upper} composition.
+Now, the following theorem tells us what happens if these are secure:
+Given \bul{$\cc{\src{L}}{\trg{L}}$ robustly preserves $\class[_{1}]$} and \rul{$\cc{\obj{L}}{\trg{L}}$ robustly preserves $\class[_{2}]$}, it follows that \oul{their upper composition $\cc{\src{L}+\obj{L}}{\trg{L}}$ robustly preserves the intersection of classes $\class[_{1}]$ and $\class[_{2}]$}.
+
+\begin{theorem}[Upper Composition of Secure Compilers]\label{thm:urtp}
+  If \bul{$\rtp{\cc{\src{L}}{\trg{L}}}{\class[_{1}]}$} and \rul{$\rtp{\cc{\obj{L}}{\trg{L}}}{\class[_{2}]}$}, then \oul{$\rtp{\cc{\src{L}+\obj{L}}{\trg{L}}}{\class[_{1}]\cap\class[_{2}]}$}. %\Coqed
+\end{theorem}
+
+Dually, the {\em lower} composition is concerned about compilers that accept the same source but yield different target languages. %and the same theoretical results apply.
+\bul{Define the lower composition $\cc{\src{L}}{\trg{L}+\obj{L}}$} as follows:
+Given a program $\src{p}$, its compiled counterpart is obtained by \rul{plugging $\src{p}$ into $\cc{\src{L}}{\trg{L}}$} or by \oul{plugging \texttt{p} into $\cc{\src{L}}{\obj{L}}$}, respectively, \pul{based on the internal decision}.
+\begin{definition}[Lower Composition]
+  $$\text{\bul{$\cc{\src{L}}{\trg{L}+\obj{L}}$}}\isdef
+  % Given $L$ and $\src{p}$, yield
+  \lambda \src{p}, L\ldotp
+  \left\{\mbox{\begin{tabular}{c}
+    {\pul{if $L=\trg{L}$, then} \rul{$\cc{\src{L}}{\trg{L}}(\src{p})$}} \\
+    \mbox{\pul{if $L=\obj{L}$, then} \oul{$\cc{\src{L}}{\obj{L}}(\src{p})$}} \\
+  \end{tabular}}\right.$$
+%
+%
+%  $\begin{cases}
+%                                                                            \rul{\cc{\src{L}}{\trg{L}}(\texttt{p}) & \text{if }\texttt{p}\in\src{L}} \\
+%                                                                            \cc{\obj{L}}{\trg{L}}(\texttt{p}) & \text{if }\texttt{p}\in\obj{L}
+%                                                                          \end{cases}$
+\end{definition}
+
+Consider two compilers both accepting $\src{LLVM IR}$~\cite{lattner2004llvm} and one of them emits $\trg{x86\_64}$, while the other emits $\obj{ARMv8}$.
+It is intuitive that they are in some sense composed in the LLVM framework, but the decision of when to use one over the other is inherently {\em internal} to the formalisation effort of this kind of composition.
+For example, the user of this compiler provides an explicit flag that instructs to emit $\trg{x86\_64}$ or the framework itself detects the target platform via heuristics, such as supported instructions.
+
+The following theorem demonstrates what happens if the involved compilers are secure:
+Given \bul{$\cc{\src{L}}{\trg{L}}$ robustly preserves $\class[_{1}]$} and \rul{$\cc{\src{L}}{\obj{L}}$ robustly preserves $\class[_{2}]$}, it follows that \oul{their lower composition $\cc{\src{L}}{\trg{L}+\obj{L}}$ robustly preserves the intersection of classes $\class[_{1}]$ and $\class[_{2}]$}.
+
+\begin{theorem}[Lower Composition of Secure Compilers]\label{thm:lrtp}
+  If \bul{$\rtp{\cc{\src{L}}{\trg{L}}}{\class[_{1}]}$} and \rul{$\rtp{\cc{\obj{L}}{\trg{L}}}{\class[_{2}]}$}, then \oul{$\rtp{\cc{\src{L}}{\trg{L}+\obj{L}}}{\class[_{1}]\cap\class[_{2}]}$}. % \Coqed
+\end{theorem}
+
+%Either way, the theoretical results suggest that it is possible to always find a ,,most-general'', secure compiler, given two secure compilers, that robustly preserves the least-upper bound of the classes involved in their compilation process.
+
+\subsection{Applicability of the Compositionality Framework}\label{subsec:app}
+
+So far, this paper assumes a unified trace-model between source and target languages.
+Moreover secure compilers (in the sense of \Cref{def:rtp}) cannot change the behavior of compiled programs.
+It is worth noting that, as our case study in \Cref{sec:casestud:defs} demonstrates, several compilation passes do not change the behavior of compiled programs.
+% in such a way that the projections to the specification trace model become unequal to each other, 
+% i.e., source and target language of the compiler essentially have a different trace model. 
+However, there are instances of passes where this is not the case.
+For example, consider $\rtp{\cc{\src{L}}{\trg{L}}}{\tmssafe}$ and $\rtp{\cc{\trg{L}}{\obj{L}}}{\smssafe}$, where $\cc{\trg{L}}{\obj{L}}$ introduces code that allocates a table that stores bounds information of pointers, but fails to insert code that deallocates this table.
+By \Thmref{def:trace:tmsdef}, this violates $\tmssafe$.
+Therefore, no behavior resulting from programs produced by the $\cc{\trg{L}}{\obj{L}}$ compiler is secure with respect to $\tmssafe$, but only for $\smssafe$.
+$\cc{\trg{L}}{\obj{L}}$ modifies the behavior and this section re-introduces the compositionality framework with support for these situations.
+
+To reason about these kinds of compilers, we rely on existing work~\cite{abate2021extacc} that lifts the presented secure compilation theory to compilers whose source and target trace models are different.
+% These compilers require a Galois connections for several real-world compiler scenarios, such as a compilation pass that exploits undefined behavior.
+% This section is concerned about a compilation setting with different source and target trace models. 
+These compilers require a way to say that two traces $\src{\trace}$ and $\trg{\trace}$ are related ($\sim$) and a function to translate a source-level property $\src{\pi}$ into a target-level property ($\tau_\sim\left( \src{\pi} \right)$~\cite{abate2021extacc}) using this trace relation.
+The function to map source-level properties to target-level properties can be lifted to translate source-level classes of properties ($\src{\class}$) to target-level classes ($\tilde{\tau}_\sim\left(\src{\class}\right)$).
+The section refers to these functions as existential images.
+
+With this, \bul{a cross-language relation $\sim$ is well-formed with respect to a class of source-level properties $\src{\class}$} iff \rul{the existential images preserve set inclusion}.
+\begin{definition}[Well-formedness of Cross-Language Trace Relations for a Class $\src{\class}$]\label{def:wfc:tracerel}
+  $$
+  \text{\bul{$\wfc{\sim}{\src{\class}}$}} \isdef \text{\rul{$\forall \src{\Pi}\subseteq\src{\class}, \tau_\sim(\src{\Pi})\subseteq\tilde{\tau}_\sim(\src{\class})$}}
+  $$
+\end{definition}
+
+\Cref{def:rtpsim} shows how to adapt \Thmref{def:rtp} to account for different target-level trace-models~\cite{abate2021extacc}:
+
+\begin{definition}[Robust Compilation with $\sim$]\label{def:rtpsim}
+  $$\rtpsim{\cc{\src{L}}{\trg{L}}}{\src{\class}}{\sim} \isdef
+  \forall \src{\pi} \src{p}.
+    \text{if }\rsat{\src{\progvar}}{\src{\pi}},
+    \text{then }\rsat{\cc{\src{L}}{\trg{L}}\left(\src{p}\right)}{\tilde{\tau}_\sim\left(\src{\pi}\right)}.
+  $$
+\end{definition}
+
+To state a similar theorem as \Thmref{thm:rtp} but for \Cref{def:rtpsim}, it is necessary to define composition of cross-language trace relations.
+Suppose there are traces $\src{\trace}$ and $\trg{\trace}$ as well as two trace relations $\sim_1$ and $\sim_2$. 
+The traces are related $\src{\trace}\sim_1\bullet\sim_2\trg{\trace}$ if there is another trace $\irl{\trace}$ such that $\src{\trace}\sim_1\irl{\trace}$ and $\irl{\trace}\sim_2\trg{\trace}$.
+
+For the adapted sequential composition theorem, let \bul{$\cc{\src{L}}{\trg{L}}$ robustly preserve the class $\src{\class[_1]}$ under $\sim_1$} and let \rul{$\cc{\trg{L}}{\obj{L}}$ robustly preserve the class $\tilde{\tau}_{\sim_1}\left(\src{\class[_2]}\right)$ under $\sim_2$}.
+Then, \iul{when the cross-language relation $\sim_1$ is well-formed with respect to class $\src{\class[_2]}$}, and \pul{relation $\sim_2$ is well-formed with respect to $\src{\class[_1]}$}, it follows that \oul{the composed compiler $\cc{\src{L}}{\trg{L}}\circ\cc{\trg{L}}{\obj{L}}$ robustly preserves the intersection of classes $\src{\class[_1]}\cap\src{\class[_2]}$ under $\sim_1\bullet\sim_2$}.
+% 
+%For space reasons, \Thmref{thm:urtp} and \Thmref{thm:lrtp} are not presented here, but can be found in the technical appendix ($\CoqSymbol$).
+% Note that it is not necessary to require $\sim_2$ to be well-formed with respect to $\src{\class[_1]}$, since the 
+% \MPin{why}
+\begin{theorem}[Sequential Composition of Secure Compilers]\label{thm:rtpsim}
+  If \bul{$\rtpsim{\cc{\src{L}}{\trg{L}}}{\src{\class[_{1}]}}{\sim_1}$}, \rul{$\rtpsim{\cc{\trg{L}}{\obj{L}}}{\tilde{\tau}_{\sim_1}\left(\src{\class[_{2}]}\right)}{\sim_2}$}, and \iul{$\wf{\sim_1}{\src{\class[_2]}}$}, and \pul{$\wf{\sim_2}{\src{\class[_1]}}$}, then \oul{$\rtpsim{\cc{\src{L}}{\trg{L}}\circ\cc{\trg{L}}{\obj{L}}}{\src{\class[_{1}]}\cap\src{\class[_{2}]}}{\sim_1\bullet\sim_2}$}.
+\Coqed
+\end{theorem}
+
+The reason we need assumptions about both relations is motivated by \Cref{lem:useless}.
+% 
+In fact, if \bul{that a relation $\sim_1$ is well-formed for a property $\class$} but \rul{another relation $\sim_2$ is not well-formed for the property mapped with the first relation $\tilde{\tau}_{\sim_1}\class$} then \iul{their composition is empty}.
+% 
+An empty intersection means that there is no correct behaviour to preserve for the compiler, which nullifies its security relevance.
+\begin{lemma}[Composition is Useless]\label{lem:useless}
+  If \bul{$\wf{\sim_1}{\class}$}, \rul{$\neg\left(\wf{\sim_2}{\tilde{\tau}_{\sim_1}\class}\right)$}, then \iul{$\tilde{\tau}_{\sim_1\bullet\sim_2}\left(\class\right)=\emptyset$}. \Coqed
+\end{lemma}
+
+Let us now reconsider the example from the beginning of the section, suppose $\sim_{{\smssafe}}$ is a suitable trace relation such that $\trg{\trace}\sim_{{\smssafe}}\irl{\trace}$ establishes the fact that $\irl{\trace}$ contains an allocation of a table containing bounds information which is never deallocated.
+It is impossible that $\wf{\sim_{{\smssafe}}}{{\tmssafe}}$, since every behavior of compiled $\trg{L}$ programs never contains a deallocation of the bounds-information table.
+Still, supposing $\rtpsim{\cc{\src{L}}{\trg{L}}}{\tmssafe}{{\sim_{\tmssafe}}}$, $\rtpsim{\cc{\trg{L}}{\irl{L}}}{\smssafe}{{\sim_{\smssafe}}}$,
+and $\wf{{\sim_{\tmssafe}}}{{\smssafe}}$ allows to apply \Thmref{thm:rtpsim}.
+However, the resulting class $\tilde{\tau}_{{{\sim_{\tmssafe}}\bullet{\sim_{\smssafe}}}}\left(\tmssafe\cap\smssafe\right)=\emptyset$.
+
+\smallskip
+
+For sake of simplicity, the case-study presented in the following sections does not consider different trace-models: The traces of all programs of the presented languages are projected to a unified trace model.
+The case-study demonstrates that, the compositionality framework without the different trace-models is already applicable to a variety of security-relevant passes as well as existing optimisations.
+% may seem overly simplistic, it is useful and simpler in practice, if applicable.
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
 
@@
 \subsection{Monitors\pages{1.5}}\label{subsec:monitors}
+
+\MPin{
+  if we cut all of this, we gain 2 pages.
+  we still need this subsection where we say that:
+  - proofs using the traces definitions above are tricky
+  - for simplicity, we use an equivalent encoding as monitors, which is doable for safety properties, and we use safety properties
+  - for each property we have a monitor (do we need to say how we indicate that monitor? or will later code not use it)
+  - for each property, we show its monitor correctly captures it
+  - we need to add another trace model for teh monitor and a relation between the property and the monitor's traces but that is standard
+  - details are in the appendices
+}
+\MPin{
+  likely, we can get away with the aforementioned stuff in 1 column
+}
+\MPin{
+  with the extra space:
+  - define SS as taken from exorcising (and probably some other work, Xaver should know), and its trace model extension
+  - reintroduce the language definition using a align* environment in V.A
+  - add V.F, defining L-spect, a speculatively-unsafe language, with an example showing the canonical Spectre V1 gadget (it normally uses heap-allocated arrays, which we don't have in our language, but the memory is N-indexed, so we should be able to do this ?)
+  - add VI.F defining the fence-inserting compiler, defining its trace relation and the extra lemmas that make it interesting
+}
+\MPin{
+  extra things to note:
+  - start talking about SS in the intro, 
+  - check the paper (like, conclusions) for ms+scct and replace with the new acronym (ms+scct+ss == ? we should find a single name, like SMC: speculative memory safety and CT)
+  - drop the last paragraph of III and replace with 'another example is SS, which we do in ... ...'
+  - change the title of III.C ? if we do a pass with the cross-trace relation, then 
+  - how is III.A.1 affected by III.C ? do we get another/a different corollary for TAU optimising passes? is it worthwile moving III.A.1 at the end, as III.D, with a more general discussion?
+}
 
@@
 
-\section{Composing Secure Compilers\pages{1}}\label{sec:compcomp}
-
-This section presents the key meta-theoretic results of this paper concerning secure compiler compositions (\Cref{sec:sequential,sec:other-compos}).
-% Note that the presented framework works with arbitrary classes.
-% \MP{check, we ened more than this}
-% and builds on prior work~\cite{kruse2022csc}\MK{mention this at all?
-may reveal identity even though its double-blind}.
-% MP: not needed at all, ever, workshop work does not need to be cited in the main work
-Note that since it deals with the preservation of \emph{classes}, which include both trace properties as well as hyperproperties, the presented framework is general: It describes robust compilers preserving ordinary trace properties, hyperproperties or even relational hyperproperties.
-% \MP{pull this out to earlier}
-
-\subsection{Secure Sequential Composition}\label{sec:sequential}
-
-Our main result is that secure compilers in the robust compilation framework~\cite{abate2019jour} compose {\em sequentially}.
-% This is not intuitive in the sense that in the security domain, composition does not work without additional generalizations~\cite{canetti2006univcomp,mccullough2012compo,fabian2022automatic}.
-% \MP{what deos this mean?}
-The \bul{sequential composition of compilers $\cc{\src{L}}{\trg{L}}$ and $\cc{\trg{L}}{\obj{L}}$} is defined as follows:
-Given an $\src{L}$ program $\src{p}$ and compilers $\cc{\src{L}}{\trg{L}},\cc{\trg{L}}{\obj{L}}$, \rul{its compiled $\obj{L}$ counterpart is obtained by plugging $\src{p}$ into $\cc{\src{L}}{\trg{L}}\circ\cc{\trg{L}}{\obj{L}}$}.
-\begin{definition}[Sequential Composition]
-  \bul{$\cc{\src{L}}{\trg{L}}\circ\cc{\trg{L}}{\obj{L}}$} $\isdef$ Given $\src{p}$, yield \rul{$\cc{\trg{L}}{\obj{L}}\left(\cc{\src{L}}{\trg{L}}(\src{p})\right)$}
-\end{definition}
-
-Consider the compilation chain for $\src{TypeScript}$.
-First, $\src{TypeScript}$ programs are translated to $\trg{JavaScript}$ which, e.g., V8~\cite{googlev8} eventually compiles in parts to $\irl{Ignition BC}$.
-The following theorem establishes what happens if all these compilation steps were robustly secure with respect to \gls*{ms}:
-The resulting $\irl{Ignition BC}$ code would be \gls*{ms} regardless of the context the binary runs in.
-
-Given \bul{$\cc{\src{L}}{\trg{L}}$ robustly preserves $\class[_{1}]$} and \rul{$\cc{\trg{L}}{\obj{L}}$ robustly preserves $\class[_{2}]$}, it follows that \oul{their sequential composition $\cc{\src{L}}{\trg{L}}\circ\cc{\trg{L}}{\obj{L}}$ robustly preserves the intersection of classes $\class[_{1}]$ and $\class[_{2}]$}.
-
-\begin{theorem}[Sequential Composition of Secure Compilers]\label{thm:rtp}
-  If \bul{$\rtp{\cc{\src{L}}{\trg{L}}}{\class[_{1}]}$} and \rul{$\rtp{\cc{\trg{L}}{\obj{L}}}{\class[_{2}]}$}, then \oul{$\rtp{\cc{\src{L}}{\trg{L}}\circ\cc{\trg{L}}{\obj{L}}}{\class[_{1}]\cap\class[_{2}]}$}.
-\Coqed
-\end{theorem}
-
-Since the composition of secure compilers is again a secure compiler, the theorem generalises to a whole chain of $n$ secure compilers.
-
-\subsubsection{Securing Optimisations}
-Notably, real-world compilation chains also perform a series of (sequential) passes whose main purpose is not necessarily to translate from one language to another, but to, e.g., optimise the code or enforce a certain property.
-Both examples can be seen in practice, e.g. as in the work of~\cite{nagarakatte2009soft,nagarakatte2010cets,akritidis2009baggy,wegman1991ccp,manjikian1997fusion} and many more.
-% \MPin{survey to cite?
-this is a bit informal MK: hard to find a survey on different compiler passes...
-I can't even find one just for compiler enforced memsafety (that looks good) or for compiler translations in general}.
-% 
-% Consider the compilation passes of LLVM~\cite{lattner2004llvm}, which are handled in the \texttt{PassManager} class.
-% By manipulating the internal state of a \texttt{PassManager} object, the schedule of compiler optimisation passes can be hardcoded.
-% That way, an order of optimisations can be found by hand that yields the best results.
-% \MP{what's the point of the previous 3 sentences?(commented)}
-Consider the following two LLVM optimisation passes: \gls*{cf}, which rewrites constant expressions to the constant they evaluate to, and \gls*{dce}, which removes dead code by rewriting conditional branches.
-The order in which \gls*{cf} and \gls*{dce} are performed influences the final result of the compilation (see \Cref{fig:cfdceex}).
-\begin{figure}[!ht]
-  \begin{center}
-    \begin{tikzpicture}[scale=0.62, every node/.style={transform shape}]
-      \node[minimum width=3cm,minimum height=2.5cm,draw=black,very thick,align=left] (progunopt) {\begin{lstlisting}[language=ml,basicstyle=\small\ttfamily]
-let a = true in
-if a then
-  print "a"
-else
-  print "b"
-\end{lstlisting}};
-      \node[minimum width=3cm,minimum height=2.5cm,draw=black,very thick,align=left,right=1.0 of progunopt] (progoptcf) {\begin{lstlisting}[language=ml,basicstyle=\large\ttfamily]
-if true then
-  print "a"
-else
-  print "b"
-\end{lstlisting}};
-      \node[minimum width=3cm,minimum height=2.5cm,draw=black,very thick,align=left,right=1.0 of progoptcf] (progoptcfdce) {\begin{lstlisting}[language=ml,basicstyle=\Large\ttfamily]
-print "a"
-\end{lstlisting}};% the following is a hack to get my syntax highlighting to work again \end{lstlisting}
-      % edges
-      \draw[->,very thick] (progunopt) edge[loop left,left] node {\gls*{dce}} (progunopt);
-      \draw[->,very thick] (progunopt) edge[sloped,above] node {\gls*{cf}} (progoptcf);
-      \draw[->,very thick] (progoptcf) edge[sloped,above] node {\gls*{dce}} (progoptcfdce);
-    \end{tikzpicture}
-  \end{center}
-  \caption{Example program where the level of optimisations differ for one pass of applying \gls*{cf} and \gls*{dce} in any order. %
-  Every edge is a compilation pass and the label on the edge states what the pass does, i.e., \gls*{cf} or \gls*{dce}. %
-  The source code in the nodes is a glorified compiler intermediate representation and the code gets more optimised towards the right hand side of the figure.}\label{fig:cfdceex}
-\end{figure}
-This {\em phase ordering problem} is well--known in literature and a practical solution is to simply perform a fixpoint iteration of the optimisation pipeline~\cite{click1995combining}.
-%If possible, optimisations are merged to enhance the analysis' result, since even when iterating to a fixpoint the final program may not be as optimised as it could be~\cite{click1995combining}.
-Compiler engineers typically try to find an order of optimisations that yields well-optimised programs for either code size~\cite{cooper1999geneticphases} or performance~\cite{kulkarni2006exhaustivephase}.
-\Cref{corr:swappable} justifies that any such order of compilation passes is valid with respect to security.
-So, given two compilation passes \bul{$\cc[_{1}]{\trg{L}}{\trg{L}}$, $\cc[_{2}]{\trg{L}}{\trg{L}}$, both robustly preserving class $\class[_{1}]$ or $\class[_{2}]$, respectively}, \oul{for any order of their composition the composed compiler robustly preserves the intersection of $\class[_{1}]$ and $\class[_{2}]$}.
-% 
-% \MPin{
-% 	later on, say that optimisation passes remove elements from traces, so they should be fine with respect to the result with $\tau$.
-% 	also, yes, if $\tau$ adds we have a problem, if it removes we don't
-% \MKin{ Optimisations that remove elements from traces are not necessarily sound,
-%  e.g., remove all deallocations to "optimise" memory management by offloading
-%  the work to the OS.
-% }
-% }
-
-\begin{corollary}[Sequential Composition of Secure Compilers]\label{corr:swappable}
-  If \bul{$\rtp{\cc[_{1}]{\trg{L}}{\trg{L}}}{\class[_{1}]}$ and $\rtp{\cc[_{2}]{\trg{L}}{\trg{L}}}{\class[_{2}]}$}, then \oul{$\rtp{\cc[_{1}]{\trg{L}}{\trg{L}}\circ\cc[_{2}]{\trg{L}}{\trg{L}}}{\class[_{1}]\cap\class[_{2}]}$ and $\rtp{\cc[_{2}]{\trg{L}}{\trg{L}}\circ\cc[_{1}]{\trg{L}}{\trg{L}}}{\class[_{2}]\cap\class[_{1}]}$}.
-\Coqed
-\end{corollary}
-
-\subsection{Secure Upper and Lower Composition}\label{sec:other-compos}
-Besides sequential composition, there are two other compositions, namely an {\em upper}, i.e., a compiler that takes multiple inputs and yields one output, and a {\em lower} composition, i.e., a compiler that takes one input and yields multiple outputs.
-\bul{Define the upper composition $\cc{\src{L}+\obj{L}}{\trg{L}}$} as follows:
-Given a program \texttt{p}, its compiled counterpart is obtained by \rul{plugging \texttt{p} into $\cc{\src{L}}{\trg{L}}$ if $\texttt{p}\in\src{L}$} or by \oul{plugging \texttt{p} into $\cc{\obj{L}}{\trg{L}}$ if $\texttt{p}\in\obj{L}$}.
-\begin{definition}[Upper Composition]
-  $$\text{\bul{$\cc{\src{L}+\obj{L}}{\trg{L}}$}}\isdef
-  % Given \texttt{p}, yield
-  \lambda \texttt{p}\ldotp
-  \left\{\mbox{\begin{tabular}{c}
-    {\rul{if $\texttt{p}\in\src{L}$, then $\cc{\src{L}}{\trg{L}}(\texttt{p})$}} \\
-    \mbox{\oul{if $\texttt{p}\in\obj{L}$, then $\cc{\obj{L}}{\trg{L}}(\texttt{p})$}} \\
-  \end{tabular}}\right.$$
-%
-%
-%  $\begin{cases}
-%                                                                            \rul{\cc{\src{L}}{\trg{L}}(\texttt{p}) & \text{if }\texttt{p}\in\src{L}} \\
-%                                                                            \cc{\obj{L}}{\trg{L}}(\texttt{p}) & \text{if }\texttt{p}\in\obj{L}
-%                                                                          \end{cases}$
-\end{definition}
-
-Examples of this are present in industry:
-Consider the Java Virtual Machine bytecode $\trg{JVM BC}$, which is a popular target for programming language designers due to its high performance and relevance in industry.
-Compilers for several programming languages have it as their target language, some popular instances are $\src{Java}$ and $\obj{Kotlin}$.
-Technically speaking, they both compile to class files and $\obj{Kotlin}$ objects are considered to be the same as $\src{Java}$ objects at that point.
-Both languages can be used at the same time in one project~\cite{androidstudio}.
-A compiler that accepts both $\src{Java}$ and $\obj{Kotlin}$ code translating to the same target language or intermediate representation performs a kind of {\em upper} composition.
-Now, the following theorem tells us what happens if these are secure:
-Given \bul{$\cc{\src{L}}{\trg{L}}$ robustly preserves $\class[_{1}]$} and \rul{$\cc{\obj{L}}{\trg{L}}$ robustly preserves $\class[_{2}]$}, it follows that \oul{their upper composition $\cc{\src{L}+\obj{L}}{\trg{L}}$ robustly preserves the intersection of classes $\class[_{1}]$ and $\class[_{2}]$}.
-
-\begin{theorem}[Upper Composition of Secure Compilers]\label{thm:urtp}
-  If \bul{$\rtp{\cc{\src{L}}{\trg{L}}}{\class[_{1}]}$} and \rul{$\rtp{\cc{\obj{L}}{\trg{L}}}{\class[_{2}]}$}, then \oul{$\rtp{\cc{\src{L}+\obj{L}}{\trg{L}}}{\class[_{1}]\cap\class[_{2}]}$}. %\Coqed
-\end{theorem}
-
-Dually, the {\em lower} composition is concerned about compilers that accept the same source but yield different target languages. %and the same theoretical results apply.
-\bul{Define the lower composition $\cc{\src{L}}{\trg{L}+\obj{L}}$} as follows:
-Given a program $\src{p}$, its compiled counterpart is obtained by \rul{plugging $\src{p}$ into $\cc{\src{L}}{\trg{L}}$} or by \oul{plugging \texttt{p} into $\cc{\src{L}}{\obj{L}}$}, respectively, \pul{based on the internal decision}.
-\begin{definition}[Lower Composition]
-  $$\text{\bul{$\cc{\src{L}}{\trg{L}+\obj{L}}$}}\isdef
-  % Given $L$ and $\src{p}$, yield
-  \lambda \src{p}, L\ldotp
-  \left\{\mbox{\begin{tabular}{c}
-    {\pul{if $L=\trg{L}$, then} \rul{$\cc{\src{L}}{\trg{L}}(\src{p})$}} \\
-    \mbox{\pul{if $L=\obj{L}$, then} \oul{$\cc{\src{L}}{\obj{L}}(\src{p})$}} \\
-  \end{tabular}}\right.$$
-%
-%
-%  $\begin{cases}
-%                                                                            \rul{\cc{\src{L}}{\trg{L}}(\texttt{p}) & \text{if }\texttt{p}\in\src{L}} \\
-%                                                                            \cc{\obj{L}}{\trg{L}}(\texttt{p}) & \text{if }\texttt{p}\in\obj{L}
-%                                                                          \end{cases}$
-\end{definition}
-
-Consider two compilers both accepting $\src{LLVM IR}$~\cite{lattner2004llvm} and one of them emits $\trg{x86\_64}$, while the other emits $\obj{ARMv8}$.
-It is intuitive that they are in some sense composed in the LLVM framework, but the decision of when to use one over the other is inherently {\em internal} to the formalisation effort of this kind of composition.
-For example, the user of this compiler provides an explicit flag that instructs to emit $\trg{x86\_64}$ or the framework itself detects the target platform via heuristics, such as supported instructions.
-
-The following theorem demonstrates what happens if the involved compilers are secure:
-Given \bul{$\cc{\src{L}}{\trg{L}}$ robustly preserves $\class[_{1}]$} and \rul{$\cc{\src{L}}{\obj{L}}$ robustly preserves $\class[_{2}]$}, it follows that \oul{their lower composition $\cc{\src{L}}{\trg{L}+\obj{L}}$ robustly preserves the intersection of classes $\class[_{1}]$ and $\class[_{2}]$}.
-
-\begin{theorem}[Lower Composition of Secure Compilers]\label{thm:lrtp}
-  If \bul{$\rtp{\cc{\src{L}}{\trg{L}}}{\class[_{1}]}$} and \rul{$\rtp{\cc{\obj{L}}{\trg{L}}}{\class[_{2}]}$}, then \oul{$\rtp{\cc{\src{L}}{\trg{L}+\obj{L}}}{\class[_{1}]\cap\class[_{2}]}$}. % \Coqed
-\end{theorem}
-
-%Either way, the theoretical results suggest that it is possible to always find a ,,most-general'', secure compiler, given two secure compilers, that robustly preserves the least-upper bound of the classes involved in their compilation process.
-
-\subsection{Applicability of the Compositionality Framework}\label{subsec:app}
-
-So far, this paper assumes a unified trace-model between source and target languages.
-Moreover secure compilers (in the sense of \Cref{def:rtp}) cannot change the behavior of compiled programs.
-It is worth noting that, as our case study in \Cref{sec:casestud:defs} demonstrates, several compilation passes do not change the behavior of compiled programs.
-% in such a way that the projections to the specification trace model become unequal to each other, 
-% i.e., source and target language of the compiler essentially have a different trace model. 
-However, there are instances of passes where this is not the case.
-For example, consider $\rtp{\cc{\src{L}}{\trg{L}}}{\tmssafe}$ and $\rtp{\cc{\trg{L}}{\obj{L}}}{\smssafe}$, where $\cc{\trg{L}}{\obj{L}}$ introduces code that allocates a table that stores bounds information of pointers, but fails to insert code that deallocates this table.
-By \Thmref{def:trace:tmsdef}, this violates $\tmssafe$.
-Therefore, no behavior resulting from programs produced by the $\cc{\trg{L}}{\obj{L}}$ compiler is secure with respect to $\tmssafe$, but only for $\smssafe$.
-$\cc{\trg{L}}{\obj{L}}$ modifies the behavior and this section re-introduces the compositionality framework with support for these situations.
-
-To reason about these kinds of compilers, we rely on existing work~\cite{abate2021extacc} that lifts the presented secure compilation theory to compilers whose source and target trace models are different.
-% These compilers require a Galois connections for several real-world compiler scenarios, such as a compilation pass that exploits undefined behavior.
-% This section is concerned about a compilation setting with different source and target trace models. 
-These compilers require a way to say that two traces $\src{\trace}$ and $\trg{\trace}$ are related ($\sim$) and a function to translate a source-level property $\src{\pi}$ into a target-level property ($\tau_\sim\left( \src{\pi} \right)$~\cite{abate2021extacc}) using this trace relation.
-The function to map source-level properties to target-level properties can be lifted to translate source-level classes of properties ($\src{\class}$) to target-level classes ($\tilde{\tau}_\sim\left(\src{\class}\right)$).
-The section refers to these functions as existential images.
-
-With this, \bul{a cross-language relation $\sim$ is well-formed with respect to a class of source-level properties $\src{\class}$} iff \rul{the existential images preserve set inclusion}.
-\begin{definition}[Well-formedness of Cross-Language Trace Relations for a Class $\src{\class}$]\label{def:wfc:tracerel}
-  $$
-  \text{\bul{$\wfc{\sim}{\src{\class}}$}} \isdef \text{\rul{$\forall \src{\Pi}\subseteq\src{\class}, \tau_\sim(\src{\Pi})\subseteq\tilde{\tau}_\sim(\src{\class})$}}
-  $$
-\end{definition}
-
-\Cref{def:rtpsim} shows how to adapt \Thmref{def:rtp} to account for different target-level trace-models~\cite{abate2021extacc}:
-
-\begin{definition}[Robust Compilation with $\sim$]\label{def:rtpsim}
-  $$\rtpsim{\cc{\src{L}}{\trg{L}}}{\src{\class}}{\sim} \isdef
-  \forall \src{\pi} \src{p}.
-    \text{if }\rsat{\src{\progvar}}{\src{\pi}},
-    \text{then }\rsat{\cc{\src{L}}{\trg{L}}\left(\src{p}\right)}{\tilde{\tau}_\sim\left(\src{\pi}\right)}.
-  $$
-\end{definition}
-
-To state a similar theorem as \Thmref{thm:rtp} but for \Cref{def:rtpsim}, it is necessary to define composition of cross-language trace relations.
-Suppose there are traces $\src{\trace}$ and $\trg{\trace}$ as well as two trace relations $\sim_1$ and $\sim_2$. 
-The traces are related $\src{\trace}\sim_1\bullet\sim_2\trg{\trace}$ if there is another trace $\irl{\trace}$ such that $\src{\trace}\sim_1\irl{\trace}$ and $\irl{\trace}\sim_2\trg{\trace}$.
-
-For the adapted sequential composition theorem, let \bul{$\cc{\src{L}}{\trg{L}}$ robustly preserve the class $\src{\class[_1]}$ under $\sim_1$} and let \rul{$\cc{\trg{L}}{\obj{L}}$ robustly preserve the class $\tilde{\tau}_{\sim_1}\left(\src{\class[_2]}\right)$ under $\sim_2$}.
-Then, \iul{when the cross-language relation $\sim_1$ is well-formed with respect to class $\src{\class[_2]}$}, and \pul{relation $\sim_2$ is well-formed with respect to $\src{\class[_1]}$}, it follows that \oul{the composed compiler $\cc{\src{L}}{\trg{L}}\circ\cc{\trg{L}}{\obj{L}}$ robustly preserves the intersection of classes $\src{\class[_1]}\cap\src{\class[_2]}$ under $\sim_1\bullet\sim_2$}.
-% 
-%For space reasons, \Thmref{thm:urtp} and \Thmref{thm:lrtp} are not presented here, but can be found in the technical appendix ($\CoqSymbol$).
-% Note that it is not necessary to require $\sim_2$ to be well-formed with respect to $\src{\class[_1]}$, since the 
-% \MPin{why}
-\begin{theorem}[Sequential Composition of Secure Compilers]\label{thm:rtpsim}
-  If \bul{$\rtpsim{\cc{\src{L}}{\trg{L}}}{\src{\class[_{1}]}}{\sim_1}$}, \rul{$\rtpsim{\cc{\trg{L}}{\obj{L}}}{\tilde{\tau}_{\sim_1}\left(\src{\class[_{2}]}\right)}{\sim_2}$}, and \iul{$\wf{\sim_1}{\src{\class[_2]}}$}, and \pul{$\wf{\sim_2}{\src{\class[_1]}}$}, then \oul{$\rtpsim{\cc{\src{L}}{\trg{L}}\circ\cc{\trg{L}}{\obj{L}}}{\src{\class[_{1}]}\cap\src{\class[_{2}]}}{\sim_1\bullet\sim_2}$}.
-\Coqed
-\end{theorem}
-
-The reason we need assumptions about both relations is motivated by \Cref{lem:useless}.
-% 
-In fact, if \bul{that a relation $\sim_1$ is well-formed for a property $\class$} but \rul{another relation $\sim_2$ is not well-formed for the property mapped with the first relation $\tilde{\tau}_{\sim_1}\class$} then \iul{their composition is empty}.
-% 
-An empty intersection means that there is no correct behaviour to preserve for the compiler, which nullifies its security relevance.
-\begin{lemma}[Composition is Useless]\label{lem:useless}
-  If \bul{$\wf{\sim_1}{\class}$}, \rul{$\neg\left(\wf{\sim_2}{\tilde{\tau}_{\sim_1}\class}\right)$}, then \iul{$\tilde{\tau}_{\sim_1\bullet\sim_2}\left(\class\right)=\emptyset$}. \Coqed
-\end{lemma}
-
-Let us now reconsider the example from the beginning of the section, suppose $\sim_{{\smssafe}}$ is a suitable trace relation such that $\trg{\trace}\sim_{{\smssafe}}\irl{\trace}$ establishes the fact that $\irl{\trace}$ contains an allocation of a table containing bounds information which is never deallocated.
-It is impossible that $\wf{\sim_{{\smssafe}}}{{\tmssafe}}$, since every behavior of compiled $\trg{L}$ programs never contains a deallocation of the bounds-information table.
-Still, supposing $\rtpsim{\cc{\src{L}}{\trg{L}}}{\tmssafe}{{\sim_{\tmssafe}}}$, $\rtpsim{\cc{\trg{L}}{\irl{L}}}{\smssafe}{{\sim_{\smssafe}}}$,
-and $\wf{{\sim_{\tmssafe}}}{{\smssafe}}$ allows to apply \Thmref{thm:rtpsim}.
-However, the resulting class $\tilde{\tau}_{{{\sim_{\tmssafe}}\bullet{\sim_{\smssafe}}}}\left(\tmssafe\cap\smssafe\right)=\emptyset$.
-
-\smallskip
-
-For sake of simplicity, the case-study presented in the following sections does not consider different trace-models: The traces of all programs of the presented languages are projected to a unified trace model.
-The case-study demonstrates that, the compositionality framework without the different trace-models is already applicable to a variety of security-relevant passes as well as existing optimisations.
-% may seem overly simplistic, it is useful and simpler in practice, if applicable.
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
 
@@
 Moreover, the language extends pre-events with $\obj{\ev{Branch}\ n}$ and $\obj{\ev{Binop}\ n}$ that are emitted when evaluating a branch or certain binary expressions, such as division, respectively, whenever the data independent timing mode is inactive.
-The following rules demonstrate how this is handled semantically.
+The following rules demonstrate how this is handled semantically. \MP{which ones ? refs?}
 
@@
 
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+
 \section{Case Study: Composing Secure Compiler Passes and Optimisations \pages{7}}\label{sec:casestud:rtp}
@@
 %    \cca(\src{\lset{x}{\expr[_{1}]}{\expr[_{2}]}}) & = \lset[\trg]{\left[\cca(\src{x})\right]}{\left[\cca(\src{\expr[_{1}]})\right]}{\left[\cca(\src{\expr_{2}})\right]} \\
-\cca(\src{\lfunction{g}{x:\natt\to\type[_{e}]}{\expr}})  = &\ \lfunction[\trg]{\trg{g}}{\trg{x}}{\lifz[\trg]{\trg{\lhast{x}{\natt}}}{\\%
+\cca(\src{\lfunction{g}{x:\natt\to\type[_{e}]}{\expr}})  = &\ \lfunction[\trg]{\trg{g}}{\trg{x}}{\lifzt[\trg]{\trg{\lhast{x}{\natt}}}{\\%
                                                             &\left[\cca(\src{\expr})\right] %
@@
 
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+
 \section{Related Work\pages{2}}\label{sec:relwork}
@@
 This can be realised by employing the FaCT~\cite{cauligi2019fact} compiler, which translates common non-constant time code patterns to be constant-time, and the data (object) independent timing execution mode of modern processors.
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
 

### 2024-05-10 csf2/acronyms.tex
add csf2

\newacronym{rtp}{RTP}{Robust Trace-Property Preservation}
\newacronym{tms}{TMS}{Temporal Memory Safety}
\newacronym{sms}{SMS}{Spatial Memory Safety}
\newacronym{ms}{MS}{Memory Safety}
\newacronym{cct}{CCT}{Cryptographic Constant Time}
\newacronym{scct}{sCCT}{Strict Cryptographic Constant Time}
\newacronym{msscct}{MS+sCCT}{Memory Safety and Strict Cryptographic Constant Time}
\newacronym{dce}{DCE}{Dead Code Elimination}
\newacronym{cf}{CF}{Constant Folding}

### 2024-05-10 csf2/main.tex
add csf2
figure fitting

--- 2023-10-11 esop/main.tex
@@
-% Fixing: Too many math alphabets used in version normal.
+% Fixing: Too many math alphabets used in version normal
 \newcommand\hmmax{0}
 \newcommand\bmmax{0}
-\documentclass[dvipsnames]{llncs}
+\documentclass[dvipsnames,conference]{IEEEtran}
 
@@
 \usepackage[inference]{semantic}
-\usepackage{lineno}
+\usepackage[switch]{lineno}
 \usepackage{fontawesome5}
@@
 \usepackage{glossaries}
-%\usepackage{hyperref}
+\usepackage{hyperref}
 \usepackage{cleveref}
@@
 \usepackage{nameref}
-%\usepackage{amsthm}
+\usepackage{amsthm}
 \usepackage{xspace}
@@
 %\renewcommand\UrlFont{\color{blue}\rmfamily}
-\newcommand{\url}[1]{\lstinline{#1}}
+%\newcommand{\url}[1]{\lstinline{#1}}
 
@@
 
+% allow page breaks inside multiline alignment displays
+\allowdisplaybreaks
+
 \Crefname{exampleenv}{Example}{Examples}
 
+\theoremstyle{definition}
 \newtheorem{exampleenv}{Example}[section]
-%\newtheorem{lemma}{Lemma}[section]
-%\newtheorem{theorem}{Theorem}[section]
-%\newtheorem{corollary}{Corollary}[section]
-%\newtheorem{definition}{Definition}[section]
+\newtheorem{lemma}{Lemma}[section]
+\newtheorem{theorem}{Theorem}[section]
+\newtheorem{corollary}{Corollary}[section]
+\newtheorem{definition}{Definition}[section]
 
@@
 \makeglossaries
+
+\AtBeginDocument{%
+  \providecommand\BibTeX{{%
+    \normalfont B\kern-0.5em{\scshape i\kern-0.25em b}\kern-0.8em\TeX}}}
+\makeatletter
+\newcommand{\linebreakand}{%
+  \end{@IEEEauthorhalign}
+  \hfill\mbox{}\par
+  \mbox{}\hfill\begin{@IEEEauthorhalign}
+}
+\makeatother
 
@@
 % \authorrunning{M. Kruse et al.}
-
-\institute{}
-% CISPA Helmholtz Center for Information Security and Saarland University
-% \and
-% CISPA Helmholtz Center for Information Security
-% \and
-% Trento University
-% }
-
-% \email{matthis.kruse@cispa.de \and director@cispa.de \and marco.patrignani@unitn.it}
 
@@
 \end{center}
-\keywords{memory-safety, secure compilation, formal methods}
 \end{abstract}
+
+\begin{IEEEkeywords}
+  memory-safety, cryptographic constant-time, secure compilation, robustness
+\end{IEEEkeywords}
 
@@
 These properties are of practical importance (as mentioned in \Cref{sec:introduction}) and also of interest in the case study (\Cref{sec:casestud:defs,sec:casestud:rtp}) this paper presents later.
-Lastly, for each of the key properties, this section introduces corresponding monitors (\Cref{subsec:monitors}) that check them.
 
@@
   \noindent
-  $
+  $$
   \smssafe:=\left\{\trace \left|\begin{array}{rcl}
@@
                                 \end{array}\right.\right\}
-                                $
+                                $$
   % \end{nscenter}
@@
 
-  \noindent$
+  \noindent$$
   \scctsafe:=\left\{\trace \left|\begin{array}{rcl}
@@
                                  \end{array}\right.\right\}
-  $
+  $$
 % \end{nscenter}
@@
 %\section{Monitors and Trace Agreements\pages{1.5}}\label{subsec:monitors}
-\subsection{Monitors\pages{1.5}}\label{subsec:monitors}
-
-Monitors enforce safety properties by accepting or rejecting traces, i.e., if they reject a trace, the trace does not satisfy the property they check.
-Since reasoning on monitors is easier than on just traces, this section presents a monitor for each of the previously shown safety properties (\Cref{subsec:propdefs}).
-%\MPpost{
-%	this is fine, but it'd be nice to provide more insights on why it's easier
-%}
-To lessen the burden when proving that a monitor accepts the trace of a program execution, each monitor uses a custom trace model that contains only the relevant information related to the property the monitor checks.
-To go from specification traces $\trace$ to monitor-level traces $\abstrace$, each property $\pi$ has an associated event agreement relation $\traceagree[]{\pi}{\event}{\absevent}$.
-Lifting the event agreement to traces works as expected, i.e., the empty traces are related, both are allowed to stutter with the empty event, and otherwise the events $\event$ and $\absevent$ must fulfill $\traceagree[]{\pi}{\event}{\absevent}$.
-The trace agreement is the same for all properties $\pi$. 
-With agreements, this section defines monitor satisfaction for traces and then it proves that monitor satisfaction implies property satisfaction.
-To this end, monitor satisfaction is defined as follows.
-\bul{A specification trace $\trace$ monitor-satisfies property $\pi$} iff there exists a (final) monitor state $\monitor$ and an abstract trace $\abstrace$ such that the \iul{specification trace $\trace$ agrees with abstract trace $\abstrace$} and \oul{the initial monitor\footnote{In this paper, for all monitors, the initial monitor state is denoted as $\emptyset$.} can step to the (final) monitor state $\monitor$ with abstract trace $\abstrace$}.
-
-\begin{definition}[Monitor Satisfaction]\label{def:monsat}
-  \bul{$\monsafe{\trace}{\pi}$} $\isdef$ $\exists \abstrace\ \monitor$, \iul{$\traceagree{\pi}{\trace}{\abstrace}$} and \oul{$\monitorcheck{\emptyset}{\monitor}{\abstrace}$}.
-\end{definition}
-
-\subsubsection{Monitor for \glsfirst*{tms}}\label{subsubsec:tms}
-\begin{gather*}
-  \begin{aligned}
-    \mi{(Abstract\ Store)}~\tmsmonitor\bnfdef&\ \left\{\text{A}:L\times\sandboxtag,\text{F}:L\times\sandboxtag\right\} 
-    \\%
-    % \emptyset:=&\ \left\{\text{A}:\emptyset,\text{F}:\emptyset\right\}
+% \subsection{Monitors\pages{1.5}}\label{subsec:monitors}
+
+% Monitors enforce safety properties by accepting or rejecting traces, i.e., if they reject a trace, the trace does not satisfy the property they check.
+% Since reasoning on monitors is easier than on just traces, this section presents a monitor for each of the previously shown safety properties (\Cref{subsec:propdefs}).
+% %\MPpost{
+% %	this is fine, but it'd be nice to provide more insights on why it's easier
+% %}
+% To lessen the burden when proving that a monitor accepts the trace of a program execution, each monitor uses a custom trace model that contains only the relevant information related to the property the monitor checks.
+% To go from specification traces $\trace$ to monitor-level traces $\abstrace$, each property $\pi$ has an associated event agreement relation $\traceagree[]{\pi}{\event}{\absevent}$.
+% Lifting the event agreement to traces works as expected, i.e., the empty traces are related, both are allowed to stutter with the empty event, and otherwise the events $\event$ and $\absevent$ must fulfill $\traceagree[]{\pi}{\event}{\absevent}$.
+% The trace agreement is the same for all properties $\pi$. 
+% With agreements, this section defines monitor satisfaction for traces and then it proves that monitor satisfaction implies property satisfaction.
+% To this end, monitor satisfaction is defined as follows.
+% \bul{A specification trace $\trace$ monitor-satisfies property $\pi$} iff there exists a (final) monitor state $\monitor$ and an abstract trace $\abstrace$ such that the \iul{specification trace $\trace$ agrees with abstract trace $\abstrace$} and \oul{the initial monitor\footnote{In this paper, for all monitors, the initial monitor state is denoted as $\emptyset$.} can step to the (final) monitor state $\monitor$ with abstract trace $\abstrace$}.
+
+% \begin{definition}[Monitor Satisfaction]\label{def:monsat}
+%   \bul{$\monsafe{\trace}{\pi}$} $\isdef$ $\exists \abstrace\ \monitor$, \iul{$\traceagree{\pi}{\trace}{\abstrace}$} and \oul{$\monitorcheck{\emptyset}{\monitor}{\abstrace}$}.
+% \end{definition}
+
+% \subsubsection{Monitor for \glsfirst*{tms}}\label{subsubsec:tms}
+% \begin{gather*}
+%   \begin{aligned}
+%     \mi{(Abstract\ Store)}~\tmsmonitor\bnfdef&\ \left\{\text{A}:L\times\sandboxtag,\text{F}:L\times\sandboxtag\right\} 
+%     \\%
+%     % \emptyset:=&\ \left\{\text{A}:\emptyset,\text{F}:\emptyset\right\}
+%     % \\
+%     \mi{(Abstract\ Events)}~\absevent\bnfdef&\ \absev{\emptyevent} \mid \absev{Alloc\ }\loc\ \sandboxtag \mid \absev{Dealloc\ }\loc\ \sandboxtag 
+%     % \\
+%                                             % & 
+%                                             \mid \absev{Use\ }\loc\ \sandboxtag \mid \absterm\\
+%   \end{aligned}
+% \end{gather*}
+% For \gls*{tms}, the state of the monitor is a record with two sets keeping track of allocated and deallocated locations, and the initial state $\emptyset$ is defined as $\left\{\text{A}:\emptyset,\text{F}:\emptyset\right\}$.
+% % 
+% Monitor events include {\em allocation}, {\em deallocation}, and {\em use} of a location $\loc$ as well as abnormal program termination.
+
+% \begin{center}\small
+%   \judgbox{\monitorcheck[]{\tmsmonitor}{\tmsmonitor[']}{\absevent}}{,,Monitor $\tmsmonitor$ steps to $\tmsmonitor[']$ given event $\absevent$.''}$\;$\\
+%   %
+% % \typerule{tms-uninteresting}{
+% % }{
+% %   \monitorcheck[]{\tmsmonitor}{\tmsmonitor}{\absev{\emptyevent}}
+% % }{tms-uninteresting}
+% % %
+% % \typerule{tms-abort}{
+% % }{
+% %   \monitorcheck[]{\tmsmonitor}{\tmsmonitor}{\absterm}
+% % }{tms-abort}
+%   % 
+%   \typerule{tms-use}{
+%     (\loc;\sandboxtag)\in\tmsmonitor.\text{A} 
+%     &
+%     % \\
+%     (\loc;\sandboxtag)\notin\tmsmonitor.\text{F}
+%   }{
+%     \monitorcheck[]{\tmsmonitor}{\tmsmonitor}{\absev{Use\ }\loc\ \sandboxtag}
+%   }{tms-use}
+%   %
+%   \typerule{tms-alloc}{
+%     (\loc;\sandboxtag)\notin\tmsmonitor.\text{A} &
+%     (\loc;\sandboxtag)\notin\tmsmonitor.\text{F} 
+%     % \\
+%     &
+%     \tmsmonitor[']=\left\{\text{A}: \tmsmonitor.\text{A}\cup\left\{(\loc;\sandboxtag)\right\},\text{F}: \tmsmonitor.\text{F}\right\}
+%   }{
+%     \monitorcheck[]{\tmsmonitor}{\tmsmonitor[']}{\absev{Alloc\ }\loc\ \sandboxtag}
+%   }{tms-alloc}
+%   %
+%   \typerule{tms-dealloc}{
+%     (\loc;\sandboxtag)\in\tmsmonitor.\text{A} 
+%     &
+%     (\loc;\sandboxtag)\notin\tmsmonitor.\text{F} 
 %     \\
-    \mi{(Abstract\ Events)}~\absevent\bnfdef&\ \absev{\emptyevent} \mid \absev{Alloc\ }\loc\ \sandboxtag \mid \absev{Dealloc\ }\loc\ \sandboxtag 
+%     \tmsmonitor[']=\left\{\text{A}: \tmsmonitor.\text{A}\setminus\left\{(\loc;\sandboxtag)\right\} \text{F}: \tmsmonitor.\text{F}\cup\left\{(\loc;\sandboxtag)\right\}\right\}
+%   }{
+%     \monitorcheck[]{\tmsmonitor}{\tmsmonitor[']}{\absev{Dealloc\ }\loc\ \sandboxtag}
+%   }{tms-dealloc}
+%   %
+% \end{center}
+% % 
+% \Cref{tr:tms-use} simply requires that a location is (i) allocated and (ii) not freed.
+% \Cref{tr:tms-alloc,tr:tms-dealloc} both require a location to not be freed already and extend the monitor state accordingly.
+% This restriction effectively disallows reallocation to reassign the same location to an object.
+% However, the definition can easily be adapted by, e.g., attaching a natural number serving as a counter.
+% Contrary to other monitors in this paper, the multi-step relation of the \gls*{tms} monitor is non-standard:
+% \begin{center}
+%   \judgbox{\monitorcheck{\tmsmonitor}{\tmsmonitor[']}{\abstrace}}{,,Monitor $\tmsmonitor$ multi-steps to $\tmsmonitor[']$ given trace $\abstrace$.''}$\;$\\
+%   %
+%   \typerule{tms-refl}{
+%     \tmsmonitor.\text{allocated}=\emptyset
+%   }{
+%     \monitorcheck{\tmsmonitor}{\tmsmonitor}{\hole{\cdot}}
+%   }{tms-refl}
+%   %
+%   \typerule{tms-ign-trans}{
+%     \monitorcheck{\tmsmonitor}{\tmsmonitor[']}{\abstrace}
+%   }{
+%     \monitorcheck[]{\tmsmonitor}{\tmsmonitor[']}{\absev{\emptyevent}\cdot\abstrace}
+%   }{tms-ign-trans}
+%   %
+%   \typerule{tms-trans}{
+%     \monitorcheck[]{\tmsmonitor}{\tmsmonitor[']}{\absevent} &
+%     \monitorcheck{\tmsmonitor[']}{\tmsmonitor['']}{\abstrace}
+%   }{
+%     \monitorcheck[]{\tmsmonitor}{\tmsmonitor['']}{\absevent\cdot\abstrace}
+%   }{tms-trans}
+% \end{center}
+% \Cref{tr:tms-ign-trans,tr:tms-trans} are the same for all monitors, but \Cref{tr:tms-refl} has, in this case, an additional premise that all locations should be deallocated.
+% This rejects the behavior of programs that forget to free memory.
+
+% \begin{center}\small
+%   \judgbox{\tmstraceagree[]{\event}{\absevent}}{,,Abstract event $\absevent$ is equivalent to $\event$ w.r.t. \gls*{tms}.''}$\;$\\
+%   %
+%   \typerule{tms-alloc-authentic}{
+%   }{
+%     \tmstraceagree[]{\ev{Alloc}\ \loc\ n;\sandboxtag;\securitytag}{\absev{Alloc\ }\loc\ \sandboxtag}
+%   }{tms-alloc-authentic}
+%   %
+% % \typerule{tms-dealloc-authentic}{
+% % }{
+% %   \tmstraceagree[]{\ev{Dealloc}\ \loc;\sandboxtag;\securitytag}{\absev{Dealloc\ }\loc}
+% % }{tms-dealloc-authentic}
+%   %
+% % \typerule{tms-use-authentic}{
+% % }{
+% %   \tmstraceagree[]{\ev{Use}\ \loc\ n;\sandboxtag;\securitytag}{\absev{Use\ }\loc}
+% % }{tms-use-authentic}
+%   %
+%   \typerule{tms-branch-authentic}{
+%   }{
+%     \tmstraceagree[]{\ev{Branch\ }n}{\absev{\emptyevent}}
+%   }{tms-branch-authentic}
+%   %
+% % \typerule{tms-binop-authentic}{
+% % }{
+% %   \tmstraceagree[]{\ev{Binop\ }n}{\absev{\emptyevent}}
+% % }{tms-binop-authentic}
+% % %
+% % \typerule{tms-empty-authentic}{
+% % }{
+% %   \tmstraceagree[]{\emptyevent}{\absev{\emptyevent}}
+% % }{tms-empty-authentic}
+% % %
+%   % \typerule{tms-abort-authentic}{
+%   % }{
+%   %   \tmstraceagree[]{\lightning}{\absterm}
+%   % }{tms-abort-authentic}
+% \end{center}
+
+% The trace agreement maps actions to their respective (e.g., \Cref{tr:tms-alloc-authentic}), so we elide most cases.
+% % 
+% Since branches are not relevant for \gls*{tms}, they are mapped to $\absev{\emptyevent}$ (\Cref{tr:tms-branch-authentic}).
+% % \MP{
+% % 	we should say something .. ?
+% % }
+
+% \begin{lemma}[Monitor Satisfaction is $\tmssafe$]\label{lem:mon:tmssafe}
+%   $\;\;$ If $\montmssafe{\trace}$, then $\trace\in\tmssafe$. %\Coqed
+% \end{lemma}
+
+% \begin{exampleenv}[A program not satisfying \gls*{tms}]\label{ex:strncpy:cpp11}
+%   Consider the following \texttt{C++11} library that calls \texttt{strncpy} (\Cref{ex:strncpy}) and prints the result to the standard output stream.
+
+%   \begin{lstlisting}[language=c++,basicstyle=\small\ttfamily,mathescape,commentstyle=\color{gray}]
+% int hello() {
+%   char* greet = new char[12] { "Hello ESOP!" };
+%   char* to = new char[12];
+%   strncpy(11, to, greetings);
+%   delete to;
+%   printf("%cSOP\n", to[6]);
+% }
+%   \end{lstlisting}
+%     % // greet address ${\color{gray}\loc_{x}}$
+%   % // to address ${\color{gray}\loc_{y}}$
+% % 
+%   Up to the body of \texttt{printf}, the program execution yields the specification trace: 
+%     % $\ev{Alloc}\ \loc[_{x}]\ 12;\comp;\unlock\cdot\ev{Alloc}\ \loc[_{y}]\ 12;\comp;\unlock\cdot\ev{Use}\ \loc_{x}\ 0;\comp;\unlock\cdot\ev{Use}\ \loc_{x}\ 0;\comp;\unlock\cdot\ev{Use}\ \loc_{y}\ 0;\comp;\unlock\cdot\ev{Use}\ \loc_{x}\ 1;\comp;\unlock\cdot\ldots\cdot\ev{Use}\ \loc_{x}\ 12;\comp;\unlock\cdot\ev{Dealloc}\ \loc[_{y}];\comp;\unlock\cdot\ev{Use}\ \loc_{y}\ 6;\comp;\unlock$
+
+%     \vspace{-1em}
+%     {\small
+%     \begin{align*}
+%       &
+%       \ev{Alloc}\ \loc[_{x}]\ 12;\comp;\unlock
+%       \cdot
+%       \ev{Alloc}\ \loc[_{y}]\ 12;\comp;\unlock
+%       \cdot
+%       \ev{Use}\ \loc_{x}\ 0;\comp;\unlock
 %       \\
 %       &
-                                            \mid \absev{Use\ }\loc\ \sandboxtag \mid \absterm\\
-  \end{aligned}
-\end{gather*}
-For \gls*{tms}, the state of the monitor is a record with two sets keeping track of allocated and deallocated locations, and the initial state $\emptyset$ is defined as $\left\{\text{A}:\emptyset,\text{F}:\emptyset\right\}$.
-% 
-Monitor events include {\em allocation}, {\em deallocation}, and {\em use} of a location $\loc$ as well as abnormal program termination.
-
-\begin{center}\small
-  \judgbox{\monitorcheck[]{\tmsmonitor}{\tmsmonitor[']}{\absevent}}{,,Monitor $\tmsmonitor$ steps to $\tmsmonitor[']$ given event $\absevent$.''}$\;$\\
-  %
-% \typerule{tms-uninteresting}{
-% }{
-%   \monitorcheck[]{\tmsmonitor}{\tmsmonitor}{\absev{\emptyevent}}
-% }{tms-uninteresting}
-% %
-% \typerule{tms-abort}{
-% }{
-%   \monitorcheck[]{\tmsmonitor}{\tmsmonitor}{\absterm}
-% }{tms-abort}
-  % 
-  \typerule{tms-use}{
-    (\loc;\sandboxtag)\in\tmsmonitor.\text{A} 
-    &
-    % \\
-    (\loc;\sandboxtag)\notin\tmsmonitor.\text{F}
-  }{
-    \monitorcheck[]{\tmsmonitor}{\tmsmonitor}{\absev{Use\ }\loc\ \sandboxtag}
-  }{tms-use}
-  %
-  \typerule{tms-alloc}{
-    (\loc;\sandboxtag)\notin\tmsmonitor.\text{A} &
-    (\loc;\sandboxtag)\notin\tmsmonitor.\text{F} 
-    % \\
-    &
-    \tmsmonitor[']=\left\{\text{A}: \tmsmonitor.\text{A}\cup\left\{(\loc;\sandboxtag)\right\},\text{F}: \tmsmonitor.\text{F}\right\}
-  }{
-    \monitorcheck[]{\tmsmonitor}{\tmsmonitor[']}{\absev{Alloc\ }\loc\ \sandboxtag}
-  }{tms-alloc}
-  %
-  \typerule{tms-dealloc}{
-    (\loc;\sandboxtag)\in\tmsmonitor.\text{A} 
-    &
-    (\loc;\sandboxtag)\notin\tmsmonitor.\text{F} 
-    \\
-    \tmsmonitor[']=\left\{\text{A}: \tmsmonitor.\text{A}\setminus\left\{(\loc;\sandboxtag)\right\} \text{F}: \tmsmonitor.\text{F}\cup\left\{(\loc;\sandboxtag)\right\}\right\}
-  }{
-    \monitorcheck[]{\tmsmonitor}{\tmsmonitor[']}{\absev{Dealloc\ }\loc\ \sandboxtag}
-  }{tms-dealloc}
-  %
-\end{center}
-% 
-\Cref{tr:tms-use} simply requires that a location is (i) allocated and (ii) not freed.
-\Cref{tr:tms-alloc,tr:tms-dealloc} both require a location to not be freed already and extend the monitor state accordingly.
-This restriction effectively disallows reallocation to reassign the same location to an object.
-However, the definition can easily be adapted by, e.g., attaching a natural number serving as a counter.
-Contrary to other monitors in this paper, the multi-step relation of the \gls*{tms} monitor is non-standard:
-\begin{center}
-  \judgbox{\monitorcheck{\tmsmonitor}{\tmsmonitor[']}{\abstrace}}{,,Monitor $\tmsmonitor$ multi-steps to $\tmsmonitor[']$ given trace $\abstrace$.''}$\;$\\
-  %
-  \typerule{tms-refl}{
-    \tmsmonitor.\text{allocated}=\emptyset
-  }{
-    \monitorcheck{\tmsmonitor}{\tmsmonitor}{\hole{\cdot}}
-  }{tms-refl}
-  %
-  \typerule{tms-ign-trans}{
-    \monitorcheck{\tmsmonitor}{\tmsmonitor[']}{\abstrace}
-  }{
-    \monitorcheck[]{\tmsmonitor}{\tmsmonitor[']}{\absev{\emptyevent}\cdot\abstrace}
-  }{tms-ign-trans}
-  %
-  \typerule{tms-trans}{
-    \monitorcheck[]{\tmsmonitor}{\tmsmonitor[']}{\absevent} &
-    \monitorcheck{\tmsmonitor[']}{\tmsmonitor['']}{\abstrace}
-  }{
-    \monitorcheck[]{\tmsmonitor}{\tmsmonitor['']}{\absevent\cdot\abstrace}
-  }{tms-trans}
-\end{center}
-\Cref{tr:tms-ign-trans,tr:tms-trans} are the same for all monitors, but \Cref{tr:tms-refl} has, in this case, an additional premise that all locations should be deallocated.
-This rejects the behavior of programs that forget to free memory.
-
-\begin{center}\small
-  \judgbox{\tmstraceagree[]{\event}{\absevent}}{,,Abstract event $\absevent$ is equivalent to $\event$ w.r.t. \gls*{tms}.''}$\;$\\
-  %
-  \typerule{tms-alloc-authentic}{
-  }{
-    \tmstraceagree[]{\ev{Alloc}\ \loc\ n;\sandboxtag;\securitytag}{\absev{Alloc\ }\loc\ \sandboxtag}
-  }{tms-alloc-authentic}
-  %
-% \typerule{tms-dealloc-authentic}{
-% }{
-%   \tmstraceagree[]{\ev{Dealloc}\ \loc;\sandboxtag;\securitytag}{\absev{Dealloc\ }\loc}
-% }{tms-dealloc-authentic}
-  %
-% \typerule{tms-use-authentic}{
-% }{
-%   \tmstraceagree[]{\ev{Use}\ \loc\ n;\sandboxtag;\securitytag}{\absev{Use\ }\loc}
-% }{tms-use-authentic}
-  %
-  \typerule{tms-branch-authentic}{
-  }{
-    \tmstraceagree[]{\ev{Branch\ }n}{\absev{\emptyevent}}
-  }{tms-branch-authentic}
-  %
-% \typerule{tms-binop-authentic}{
-% }{
-%   \tmstraceagree[]{\ev{Binop\ }n}{\absev{\emptyevent}}
-% }{tms-binop-authentic}
-% %
-% \typerule{tms-empty-authentic}{
-% }{
-%   \tmstraceagree[]{\emptyevent}{\absev{\emptyevent}}
-% }{tms-empty-authentic}
-% %
-  % \typerule{tms-abort-authentic}{
-  % }{
-  %   \tmstraceagree[]{\lightning}{\absterm}
-  % }{tms-abort-authentic}
-\end{center}
-
-The trace agreement maps actions to their respective (e.g., \Cref{tr:tms-alloc-authentic}), so we elide most cases.
-% 
-Since branches are not relevant for \gls*{tms}, they are mapped to $\absev{\emptyevent}$ (\Cref{tr:tms-branch-authentic}).
-% \MP{
-% 	we should say something .. ?
-% }
-
-\begin{lemma}[Monitor Satisfaction is $\tmssafe$]\label{lem:mon:tmssafe}
-  $\;\;$ If $\montmssafe{\trace}$, then $\trace\in\tmssafe$. %\Coqed
-\end{lemma}
-
-\begin{exampleenv}[A program not satisfying \gls*{tms}]\label{ex:strncpy:cpp11}
-  Consider the following \texttt{C++11} library that calls \texttt{strncpy} (\Cref{ex:strncpy}) and prints the result to the standard output stream.
-
-  \begin{lstlisting}[language=c++,basicstyle=\small\ttfamily,mathescape,commentstyle=\color{gray}]
-int hello() {
-  char* greet = new char[12] { "Hello ESOP!" };
-  char* to = new char[12];
-  strncpy(11, to, greetings);
-  delete to;
-  printf("%cSOP\n", to[6]);
-}
-  \end{lstlisting}
-    % // greet address ${\color{gray}\loc_{x}}$
-  % // to address ${\color{gray}\loc_{y}}$
-% 
-  Up to the body of \texttt{printf}, the program execution yields the specification trace: 
-  	% $\ev{Alloc}\ \loc[_{x}]\ 12;\comp;\unlock\cdot\ev{Alloc}\ \loc[_{y}]\ 12;\comp;\unlock\cdot\ev{Use}\ \loc_{x}\ 0;\comp;\unlock\cdot\ev{Use}\ \loc_{x}\ 0;\comp;\unlock\cdot\ev{Use}\ \loc_{y}\ 0;\comp;\unlock\cdot\ev{Use}\ \loc_{x}\ 1;\comp;\unlock\cdot\ldots\cdot\ev{Use}\ \loc_{x}\ 12;\comp;\unlock\cdot\ev{Dealloc}\ \loc[_{y}];\comp;\unlock\cdot\ev{Use}\ \loc_{y}\ 6;\comp;\unlock$
-
-  	\vspace{-1em}
-  	{\small
-  	\begin{align*}
-  		&
-  		\ev{Alloc}\ \loc[_{x}]\ 12;\comp;\unlock
-  		\cdot
-  		\ev{Alloc}\ \loc[_{y}]\ 12;\comp;\unlock
-  		\cdot
-  		\ev{Use}\ \loc_{x}\ 0;\comp;\unlock
-  		\\
-  		&
-  		\cdot
-  		\ev{Use}\ \loc_{x}\ 0;\comp;\unlock
-  		\cdot
-  		\ev{Use}\ \loc_{y}\ 0;\comp;\unlock
-  		\cdot
-  		\ev{Use}\ \loc_{x}\ 1;\comp;\unlock
-  		\cdot
-  		\ldots
-  		\\
-  		&
-  		\cdot
-  		\ev{Use}\ \loc_{x}\ 12;\comp;\unlock
-  		\cdot
-  		\ev{Dealloc}\ \loc[_{y}];\comp;\unlock
-  		\cdot
-  		\ev{Use}\ \loc_{y}\ 6;\comp;\unlock
-  	\end{align*}
-  	}
-Relating this trace to abstract monitor events yields trace:
-
-  \vspace{-1em}
-  {\small
-  \begin{align*}
-   	% \abstrace=
-   	&
-   	\absev{Alloc}\ \loc[_{x}]\ \comp
-   	\cdot
-   	\absev{Alloc}\ \loc[_{y}]\ \comp
-   	\cdot
-   	\absev{Use}\ \loc[_{x}]\ \comp
+%       \cdot
+%       \ev{Use}\ \loc_{x}\ 0;\comp;\unlock
+%       \cdot
+%       \ev{Use}\ \loc_{y}\ 0;\comp;\unlock
+%       \cdot
+%       \ev{Use}\ \loc_{x}\ 1;\comp;\unlock
+%       \cdot
+%       \ldots
 %       \\
 %       &
-   	\cdot
-   	\absev{Use}\ \loc[_{x}]\ \comp
-   	\cdot
-   	\absev{Use}\ \loc[_{y}]\ \comp
-   	\cdot
-   	\ldots
-   	\\
-   	&
-   	\cdot
-   	\absev{Use}\ \loc[_{x}]\ \comp
-   	\cdot
-   	\absev{Dealloc}\ \loc[_{y}]\ \comp
-   	\cdot
-   	\absev{Use}\ \loc[_{y}]\ \comp
-   \end{align*} 
-  	}
-  Remembering the definition of \texttt{strncpy} (\Cref{ex:strncpy}), observe that it does not deallocate its arguments.
-  Even though the trace contains an out-of-bounds access right before returning from \texttt{strncpy}, this is no concern for \gls*{tms}, since location $\loc[_{x}]$ is still allocated.
-  Having returned from \texttt{strncpy}, the \texttt{greet} function continues and deallocates $\loc[_{y}]$ whose subsequent use in \texttt{printf} is a use-after-free bug.
-
-  The fix would be to \lstinline[language=c++,basicstyle=\small\ttfamily]|delete greetings| instead of \texttt{to} and add a \lstinline[language=c++,basicstyle=\small\ttfamily]|delete to| after the \texttt{printf} call, which leads to trace:
-
-\vspace{-1em}
-  {\small
-  \begin{align*}
-  	\abstrace[']=
-  	&
-  	\absev{Alloc}\ \loc[_{x}]\ \comp
-  	\cdot
-  	\absev{Alloc}\ \loc[_{y}]\ \comp
-  	\cdot
-  	\absev{Use}\ \loc[_{x}]\ \comp
+%       \cdot
+%       \ev{Use}\ \loc_{x}\ 12;\comp;\unlock
+%       \cdot
+%       \ev{Dealloc}\ \loc[_{y}];\comp;\unlock
+%       \cdot
+%       \ev{Use}\ \loc_{y}\ 6;\comp;\unlock
+%     \end{align*}
+%     }
+% Relating this trace to abstract monitor events yields trace:
+%   
+%   \vspace{-1em}
+%   {\small
+%   \begin{align*}
+%     % \abstrace=
+%     &
+%     \absev{Alloc}\ \loc[_{x}]\ \comp
+%     \cdot
+%     \absev{Alloc}\ \loc[_{y}]\ \comp
+%     \cdot
+%     \absev{Use}\ \loc[_{x}]\ \comp
+%     % \\
+%     % &
+%     \cdot
+%     \absev{Use}\ \loc[_{x}]\ \comp
+%     \cdot
+%     \absev{Use}\ \loc[_{y}]\ \comp
+%     \cdot
+%     \ldots
 %     \\
 %     &
-  	\cdot
-  	\absev{Use}\ \loc[_{x}]\ \comp
-  	\cdot
-  	\absev{Use}\ \loc[_{y}]\ \comp
-  	\cdot
-  	\ldots
-  	\cdot
-  	\\
-  	&
-  	\absev{Use}\ \loc[_{x}]\ \comp
-  	\cdot
-  	\absev{Dealloc}\ \loc[_{x}]\ \comp
-  	\cdot
-  	\absev{Use}\ \loc[_{y}]\ \comp	
-  \end{align*}
-  }
-  It follows that $\montmssafe{\abstrace[']}$ and from \Thmref{lem:mon:tmssafe}, it follows that the program satisfies \Thmref{def:trace:tmsdef}, even though the program still violates \gls*{sms}.
-\end{exampleenv}
-
-% \MPpost{
-% 	i think the trace description of traces is more uindertandable,
-% 	teh monitor one is easier to work with
-% 	this is the justification for having both that we'll have to make in the future
-% }
-
-\subsubsection{Monitor for \glsfirst*{sms}}\label{subsubsec:sms}
-\begin{gather*}
-  \begin{aligned}
-    \mi{(Abstract\ Store)}~\smsmonitor:=&\ L\times\sandboxtag\times\mb{N}\\
-    \mi{(Abstract\ Events)}~\absevent\bnfdef&\ \absev{\varepsilon} \mid \absev{Alloc\ }\loc\ \sandboxtag\ n \mid \absev{Use\ }\loc\ \sandboxtag\ n \\
-  \end{aligned}
-\end{gather*}
-\begin{center}
-  \judgbox{\monitorcheck[]{\smsmonitor}{\smsmonitor[']}{\absevent}}{,,Monitor $\smsmonitor$ steps to $\smsmonitor[']$ given event $\absevent$.''}$\;$\\
-  %
-% \typerule{sms-uninteresting}{
-% }{
-%   \monitorcheck[]{\smsmonitor}{\smsmonitor}{\absev{\emptyevent}}
-% }{sms-uninteresting}
-% %
-% \typerule{sms-abort}{
-% }{
-%   \monitorcheck[]{\smsmonitor}{\emptyset}{\absterm}
-% }{sms-abort}
-  %
-  \typerule{sms-use}{
-    (\loc;\sandboxtag;m)\in\smsmonitor &
-    n<m
-  }{
-    \monitorcheck[]{\smsmonitor}{\smsmonitor}{\absev{Use\ }\loc\ \sandboxtag\ n}
-  }{sms-use}
-  %
-  \typerule{sms-alloc}{
-    (\loc;\sandboxtag;m)\notin\smsmonitor %\\
-    % unfortunately, this does not render sms-use and sms-alloc side by side
-    % my only other idea is \scalebox...ugh
-    %\smsmonitor[']=\smsmonitor\cup\left\{(\loc;\sandboxtag;n)\right\}
-  }{
-    \monitorcheck[]{\smsmonitor}{\smsmonitor\cup\left\{(\loc;\sandboxtag;n)\right\}}{\absev{Alloc\ }\loc\ \sandboxtag\ n}
-  }{sms-alloc}
-\end{center}
-The state of the monitor for \gls*{sms} is a set containing tuples of locations, control-tags, and the allocation size.
-In comparison to the trace model of the \gls*{tms} monitor, the trace model here is extended by sizing and positional information.
-\Cref{tr:sms-use} performs a bounds check and \Cref{tr:sms-alloc} adds bounds information to the state of the monitor.
-The trace agreement is entirely straightforward and similar to the one for \gls*{tms}.
-% \begin{center}
-%   \judgbox{\smstraceagree[]{\event}{\absevent}}{,,Abstract event $\absevent$ is equivalent to $\event$ with respect to \gls*{sms}.''}$\;$\\
-%   %
-%   \typerule{sms-alloc-authentic}{
-%   }{
-%     \smstraceagree[]{\ev{Alloc}\ \loc\ n;\sandboxtag;\securitytag}{\absev{Alloc\ }\loc\ n}
-%   }{sms-alloc-authentic}
-%   %
-%   \typerule{sms-dealloc-authentic}{
-%   }{
-%     \smstraceagree[]{\ev{Dealloc}\ \loc;\sandboxtag;\securitytag}{\absev{Dealloc\ }\loc}
-%   }{sms-dealloc-authentic}
-%   %
-%   \typerule{sms-use-authentic}{
-%   }{
-%     \smstraceagree[]{\ev{Use}\ \loc\ n;\sandboxtag;\securitytag}{\absev{Use\ }\loc\ n}
-%   }{sms-use-authentic}
-%   %
-%   \typerule{sms-branch-authentic}{
-%   }{
-%     \smstraceagree[]{\ev{Branch\ }n}{\absev{\emptyevent}}
-%   }{sms-branch-authentic}
-%   %
-%   \typerule{sms-binop-authentic}{
-%   }{
-%     \smstraceagree[]{\ev{Binop\ }n}{\absev{\emptyevent}}
-%   }{sms-binop-authentic}
-%   %
-%   \typerule{sms-empty-authentic}{
-%   }{
-%     \smstraceagree[]{\emptyevent}{\absev{\emptyevent}}
-%   }{sms-empty-authentic}
-%   %
-%   \typerule{sms-abort-authentic}{
-%   }{
-%     \smstraceagree[]{\lightning}{\absterm}
-%   }{sms-abort-authentic}
-% \end{center}
-
-\begin{lemma}[Monitor Satisfaction is $\smssafe$]\label{lem:mon:smsafe}
-  $\;\;$
-  If $\monsmssafe{\trace}$, then $\trace\in\smssafe$. %\Coqed
-\end{lemma}
-
-\begin{exampleenv}[Normal invocation of \texttt{strncpy}]\label{ex:strncpy:sms}
-  Consider the insecure \texttt{strncpy} function from \Cref{ex:strncpy} with a context \texttt{strncpy(2, x, y)}, where \texttt{x} and \texttt{y} are pointers to valid regions of memory with allocated space for exactly two cells and do not contain the null-terminating character \texttt{'\textbackslash 0'}.
-  For the sake of this example, the pointers have been allocated by the component and passed to the context.
-  The loop of \texttt{strncpy} will copy exactly two cells and then check the loop condition for the last time.
-  At that stage, variable \texttt{i} is equal to \texttt{2} and, unfortunately, the order of checks is such that first the cell \texttt{x[i]} is read prior to bounds checking \texttt{i < n}, so there is an out-of-bounds memory access right before exiting the function.
-  This is also visible on the trace:
-
-  	{\small
-  	\begin{align*}
-  	&
-  		\ldots
-  		\cdot
-  		\ev{Alloc}\ \loc_{x}\ 2;\comp;\unlock
-  		\cdot
-  		\ldots
-  		\cdot
-  		\ev{Alloc}\ \loc_{y}\ 2;\comp;\unlock
-  		\cdot
-  		\ldots
-  		\cdot
+%     \cdot
+%     \absev{Use}\ \loc[_{x}]\ \comp
+%     \cdot
+%     \absev{Dealloc}\ \loc[_{y}]\ \comp
+%     \cdot
+%     \absev{Use}\ \loc[_{y}]\ \comp
+%    \end{align*} 
+%     }
+%   Remembering the definition of \texttt{strncpy} (\Cref{ex:strncpy}), observe that it does not deallocate its arguments.
+%   Even though the trace contains an out-of-bounds access right before returning from \texttt{strncpy}, this is no concern for \gls*{tms}, since location $\loc[_{x}]$ is still allocated.
+%   Having returned from \texttt{strncpy}, the \texttt{greet} function continues and deallocates $\loc[_{y}]$ whose subsequent use in \texttt{printf} is a use-after-free bug.
+
+%   The fix would be to \lstinline[language=c++,basicstyle=\small\ttfamily]|delete greetings| instead of \texttt{to} and add a \lstinline[language=c++,basicstyle=\small\ttfamily]|delete to| after the \texttt{printf} call, which leads to trace:
+%   
+% \vspace{-1em}
+%   {\small
+%   \begin{align*}
+%     \abstrace[']=
+%     &
+%     \absev{Alloc}\ \loc[_{x}]\ \comp
+%     \cdot
+%     \absev{Alloc}\ \loc[_{y}]\ \comp
+%     \cdot
+%     \absev{Use}\ \loc[_{x}]\ \comp
+%     % \\
+%     % &
+%     \cdot
+%     \absev{Use}\ \loc[_{x}]\ \comp
+%     \cdot
+%     \absev{Use}\ \loc[_{y}]\ \comp
+%     \cdot
+%     \ldots
+%     \cdot
 %     \\
 %     &
-  		\ev{Use}\ \loc_{x}\ 0;\comp;\unlock
-  		\cdot
-  		\\
-  	&
-  		\ev{Use}\ \loc_{y}\ 0;\comp;\unlock
-  		\cdot
-  		\ev{Use}\ \loc_{x}\ 1;\comp;\unlock
+%     \absev{Use}\ \loc[_{x}]\ \comp
+%     \cdot
+%     \absev{Dealloc}\ \loc[_{x}]\ \comp
+%     \cdot
+%     \absev{Use}\ \loc[_{y}]\ \comp	
+%   \end{align*}
+%   }
+%   It follows that $\montmssafe{\abstrace[']}$ and from \Thmref{lem:mon:tmssafe}, it follows that the program satisfies \Thmref{def:trace:tmsdef}, even though the program still violates \gls*{sms}.
+% \end{exampleenv}
+
+% % \MPpost{
+% % 	i think the trace description of traces is more uindertandable,
+% % 	teh monitor one is easier to work with
+% % 	this is the justification for having both that we'll have to make in the future
+% % }
+
+% \subsubsection{Monitor for \glsfirst*{sms}}\label{subsubsec:sms}
+% \begin{gather*}
+%   \begin{aligned}
+%     \mi{(Abstract\ Store)}~\smsmonitor:=&\ L\times\sandboxtag\times\mb{N}\\
+%     \mi{(Abstract\ Events)}~\absevent\bnfdef&\ \absev{\varepsilon} \mid \absev{Alloc\ }\loc\ \sandboxtag\ n \mid \absev{Use\ }\loc\ \sandboxtag\ n \\
+%   \end{aligned}
+% \end{gather*}
+% \begin{center}
+%   \judgbox{\monitorcheck[]{\smsmonitor}{\smsmonitor[']}{\absevent}}{,,Monitor $\smsmonitor$ steps to $\smsmonitor[']$ given event $\absevent$.''}$\;$\\
+%   %
+% % \typerule{sms-uninteresting}{
+% % }{
+% %   \monitorcheck[]{\smsmonitor}{\smsmonitor}{\absev{\emptyevent}}
+% % }{sms-uninteresting}
+% % %
+% % \typerule{sms-abort}{
+% % }{
+% %   \monitorcheck[]{\smsmonitor}{\emptyset}{\absterm}
+% % }{sms-abort}
+%   %
+%   \typerule{sms-use}{
+%     (\loc;\sandboxtag;m)\in\smsmonitor &
+%     n<m
+%   }{
+%     \monitorcheck[]{\smsmonitor}{\smsmonitor}{\absev{Use\ }\loc\ \sandboxtag\ n}
+%   }{sms-use}
+%   %
+%   \typerule{sms-alloc}{
+%     (\loc;\sandboxtag;m)\notin\smsmonitor %\\
+%     % unfortunately, this does not render sms-use and sms-alloc side by side
+%     % my only other idea is \scalebox...ugh
+%     %\smsmonitor[']=\smsmonitor\cup\left\{(\loc;\sandboxtag;n)\right\}
+%   }{
+%     \monitorcheck[]{\smsmonitor}{\smsmonitor\cup\left\{(\loc;\sandboxtag;n)\right\}}{\absev{Alloc\ }\loc\ \sandboxtag\ n}
+%   }{sms-alloc}
+% \end{center}
+% The state of the monitor for \gls*{sms} is a set containing tuples of locations, control-tags, and the allocation size.
+% In comparison to the trace model of the \gls*{tms} monitor, the trace model here is extended by sizing and positional information.
+% \Cref{tr:sms-use} performs a bounds check and \Cref{tr:sms-alloc} adds bounds information to the state of the monitor.
+% The trace agreement is entirely straightforward and similar to the one for \gls*{tms}.
+% % \begin{center}
+% %   \judgbox{\smstraceagree[]{\event}{\absevent}}{,,Abstract event $\absevent$ is equivalent to $\event$ with respect to \gls*{sms}.''}$\;$\\
+% %   %
+% %   \typerule{sms-alloc-authentic}{
+% %   }{
+% %     \smstraceagree[]{\ev{Alloc}\ \loc\ n;\sandboxtag;\securitytag}{\absev{Alloc\ }\loc\ n}
+% %   }{sms-alloc-authentic}
+% %   %
+% %   \typerule{sms-dealloc-authentic}{
+% %   }{
+% %     \smstraceagree[]{\ev{Dealloc}\ \loc;\sandboxtag;\securitytag}{\absev{Dealloc\ }\loc}
+% %   }{sms-dealloc-authentic}
+% %   %
+% %   \typerule{sms-use-authentic}{
+% %   }{
+% %     \smstraceagree[]{\ev{Use}\ \loc\ n;\sandboxtag;\securitytag}{\absev{Use\ }\loc\ n}
+% %   }{sms-use-authentic}
+% %   %
+% %   \typerule{sms-branch-authentic}{
+% %   }{
+% %     \smstraceagree[]{\ev{Branch\ }n}{\absev{\emptyevent}}
+% %   }{sms-branch-authentic}
+% %   %
+% %   \typerule{sms-binop-authentic}{
+% %   }{
+% %     \smstraceagree[]{\ev{Binop\ }n}{\absev{\emptyevent}}
+% %   }{sms-binop-authentic}
+% %   %
+% %   \typerule{sms-empty-authentic}{
+% %   }{
+% %     \smstraceagree[]{\emptyevent}{\absev{\emptyevent}}
+% %   }{sms-empty-authentic}
+% %   %
+% %   \typerule{sms-abort-authentic}{
+% %   }{
+% %     \smstraceagree[]{\lightning}{\absterm}
+% %   }{sms-abort-authentic}
+% % \end{center}
+
+% \begin{lemma}[Monitor Satisfaction is $\smssafe$]\label{lem:mon:smsafe}
+%   $\;\;$
+%   If $\monsmssafe{\trace}$, then $\trace\in\smssafe$. %\Coqed
+% \end{lemma}
+
+% \begin{exampleenv}[Normal invocation of \texttt{strncpy}]\label{ex:strncpy:sms}
+%   Consider the insecure \texttt{strncpy} function from \Cref{ex:strncpy} with a context \texttt{strncpy(2, x, y)}, where \texttt{x} and \texttt{y} are pointers to valid regions of memory with allocated space for exactly two cells and do not contain the null-terminating character \texttt{'\textbackslash 0'}.
+%   For the sake of this example, the pointers have been allocated by the component and passed to the context.
+%   The loop of \texttt{strncpy} will copy exactly two cells and then check the loop condition for the last time.
+%   At that stage, variable \texttt{i} is equal to \texttt{2} and, unfortunately, the order of checks is such that first the cell \texttt{x[i]} is read prior to bounds checking \texttt{i < n}, so there is an out-of-bounds memory access right before exiting the function.
+%   This is also visible on the trace:
+%     
+%     {\small
+%     \begin{align*}
+%     &
+%       \ldots
+%       \cdot
+%       \ev{Alloc}\ \loc_{x}\ 2;\comp;\unlock
+%       \cdot
+%       \ldots
+%       \cdot
+%       \ev{Alloc}\ \loc_{y}\ 2;\comp;\unlock
+%       \cdot
+%       \ldots
+%       \cdot
+%     % \\
+%     % &
+%       \ev{Use}\ \loc_{x}\ 0;\comp;\unlock
+%       \cdot
 %       \\
 %     &
-  		\cdot
-  		\ev{Use}\ \loc_{y}\ 1;\comp;\unlock
-  		\cdot
-  		\ev{Use}\ \loc_{x}\ 2;\comp;\unlock
-  		\cdot
-  		\ldots	
-  	\end{align*}
-  	}
-
-\noindent
-where $\loc_{x}$ and $\loc_{y}$ are the memory addresses associated to \texttt{x} and \texttt{y}, respectively.
-  Omitting the events for all ,,$\ldots$'' for sake of brevity, the abstract monitor trace of this is:
-  	% 
-  	{\small
-  	\begin{align*}
-  		&
-  		\absev{Alloc}\ \loc_{x}\ \comp\ 2
-  		\cdot
-  		\absev{Alloc}\ \loc_{y}\ \comp\ 2
-  		\cdot
-  		\absev{Use}\ \loc_{x}\ \comp\ 0
-  		\\
-  		&
-  		\cdot
-  		\absev{Use}\ \loc_{y}\ \comp\ 0
-  		\cdot
-  		\absev{Use}\ \loc_{x}\ \comp\ 1
-  		\cdot
-  		\absev{Use}\ \loc_{y}\ \comp\ 1
-  		\cdot
-  		\absev{Use}\ \loc_{x}\ \comp\ 2
-  	\end{align*}
-  	}
-
-  After the allocation events, the state of the monitor is $\left\{(\loc_{x};\comp;2),(\loc_{y};\comp;2)\right\}$.
-  All uses up to the last are accepted by the monitor, but the last event does not satisfy the premise $2<2$ in \Cref{tr:sms-use}.
-  Therefore, the whole program (\texttt{strncpy} linked with this kind of context) is not \gls*{sms}.
-\end{exampleenv}
-
-\subsubsection{Combining \glsfirst*{tms} and \glsfirst*{sms} Monitors into \glsfirst*{ms}}\label{subsubsec:ms}
-
-\begin{center}
-  \judgbox{\monitorcheck[]{\msmonitor}{\msmonitor[']}{\absevent}}{,,Monitor $\msmonitor$ steps to $\msmonitor[']$ given event $\absevent$.''}$\;$\\
-  \typerule{ms-step}{
-    \monitorcheck[]{\tmsmonitor}{\tmsmonitor[']}{\absevent_{\tmssafe}} &
-    \monitorcheck[]{\smsmonitor}{\smsmonitor[']}{\absevent_{\smssafe}} &
-  }{
-    \monitorcheck[]{(\tmsmonitor,\smsmonitor)}{(\tmsmonitor['],\smsmonitor['])}{(\absevent_{\tmssafe},\absevent_{\smssafe})}
-  }{ms-step}
-\end{center}
-The combined monitor runs the one for \gls*{tms} and the one for \gls*{sms} in a lockstep.
+%       \ev{Use}\ \loc_{y}\ 0;\comp;\unlock
+%       \cdot
+%       \ev{Use}\ \loc_{x}\ 1;\comp;\unlock
+%     % \\
+%     % &
+%       \cdot
+%       \ev{Use}\ \loc_{y}\ 1;\comp;\unlock
+%       \cdot
+%       \ev{Use}\ \loc_{x}\ 2;\comp;\unlock
+%       \cdot
+%       \ldots	
+%     \end{align*}
+%     }
+
+% \noindent
+% where $\loc_{x}$ and $\loc_{y}$ are the memory addresses associated to \texttt{x} and \texttt{y}, respectively.
+%   Omitting the events for all ,,$\ldots$'' for sake of brevity, the abstract monitor trace of this is:
+%     % 
+%     {\small
+%     \begin{align*}
+%       &
+%       \absev{Alloc}\ \loc_{x}\ \comp\ 2
+%       \cdot
+%       \absev{Alloc}\ \loc_{y}\ \comp\ 2
+%       \cdot
+%       \absev{Use}\ \loc_{x}\ \comp\ 0
+%       \\
+%       &
+%       \cdot
+%       \absev{Use}\ \loc_{y}\ \comp\ 0
+%       \cdot
+%       \absev{Use}\ \loc_{x}\ \comp\ 1
+%       \cdot
+%       \absev{Use}\ \loc_{y}\ \comp\ 1
+%       \cdot
+%       \absev{Use}\ \loc_{x}\ \comp\ 2
+%     \end{align*}
+%     }
+
+%   After the allocation events, the state of the monitor is $\left\{(\loc_{x};\comp;2),(\loc_{y};\comp;2)\right\}$.
+%   All uses up to the last are accepted by the monitor, but the last event does not satisfy the premise $2<2$ in \Cref{tr:sms-use}.
+%   Therefore, the whole program (\texttt{strncpy} linked with this kind of context) is not \gls*{sms}.
+% \end{exampleenv}
+
+% \subsubsection{Combining \glsfirst*{tms} and \glsfirst*{sms} Monitors into \glsfirst*{ms}}\label{subsubsec:ms}
+
+% \begin{center}
+%   \judgbox{\monitorcheck[]{\msmonitor}{\msmonitor[']}{\absevent}}{,,Monitor $\msmonitor$ steps to $\msmonitor[']$ given event $\absevent$.''}$\;$\\
+%   \typerule{ms-step}{
+%     \monitorcheck[]{\tmsmonitor}{\tmsmonitor[']}{\absevent_{\tmssafe}} &
+%     \monitorcheck[]{\smsmonitor}{\smsmonitor[']}{\absevent_{\smssafe}} &
+%   }{
+%     \monitorcheck[]{(\tmsmonitor,\smsmonitor)}{(\tmsmonitor['],\smsmonitor['])}{(\absevent_{\tmssafe},\absevent_{\smssafe})}
+%   }{ms-step}
+% \end{center}
+% The combined monitor runs the one for \gls*{tms} and the one for \gls*{sms} in a lockstep.
 The trace agreement similarly just relates a specification event with an abstract \gls*{tms}-event $\abstrace[_{\tmssafe}]$ and with an abstract \gls*{sms}-event $\abstrace[_{\smssafe}]$.
+% % \begin{center}
+% %   \judgbox{\mstraceagree{\event}{\absevent}}{,,Abstract event $\absevent$ is equivalent to $\event$ with respect to \gls*{ms}.''}$\;$\\
+% %   %
+% %   \typerule{ms-authentic}{
+% %     \tmstraceagree[]{\event}{\absevent_{\tmssafe}} &
+% %     \smstraceagree[]{\event}{\absevent_{\smssafe}}
+% %   }{
+% %     \mstraceagree[]{\event}{(\absevent_{\tmssafe},\absevent_{\smssafe})}
+% %   }{ms-authentic}
+% % \end{center}
+
+% \begin{lemma}[Monitor Satisfaction is $\mssafe$]\label{lem:mon:msafe}
+%   If $\monmssafe{\trace}$, then $\trace\in\mssafe$.\Coqed
+% \end{lemma}
+
+% Any technique to combine existing monitors $T_A$, $T_B$ (that check for properties $A$ and $B$) into one $T_{AB}$ (that checks for both $A$ and $B$) is sufficient for the sake of the analysis of this paper.
+
+% \subsubsection{Monitor for \glsfirst*{scct}}
+% \begin{gather*}
+% % \small
+%   \begin{aligned}
+%     \mi{(Abstract\ Store)}~\scctmonitor:=&\ \emptyset 
+%     % &
+%     &
+%     \qquad\mi{(Abstract\ Events)}~\absevent:=&\ \absev{\varepsilon} \mid \absterm \mid \absev{Leak} 
+%     % \\
+%   \end{aligned}
+% \end{gather*}
 % \begin{center}
-%   \judgbox{\mstraceagree{\event}{\absevent}}{,,Abstract event $\absevent$ is equivalent to $\event$ with respect to \gls*{ms}.''}$\;$\\
-%   %
-%   \typerule{ms-authentic}{
-%     \tmstraceagree[]{\event}{\absevent_{\tmssafe}} &
-%     \smstraceagree[]{\event}{\absevent_{\smssafe}}
+%   \judgbox{\monitorcheck[]{\scctmonitor}{\scctmonitor[']}{\absevent}}{,,Monitor $\scctmonitor$ steps to $\scctmonitor[']$\\given event $\absevent$.''}$\;$\\
+%   %
+%   \typerule{scct-none}{
 %   }{
-%     \mstraceagree[]{\event}{(\absevent_{\tmssafe},\absevent_{\smssafe})}
-%   }{ms-authentic}
+%     \monitorcheck[]{\scctmonitor}{\scctmonitor}{\absev{\emptyevent}}
+%   }{scct-none}
+%   %
+%   \typerule{scct-abort}{
+%   }{
+%     \monitorcheck[]{\scctmonitor}{\scctmonitor}{\absterm}
+%   }{scct-abort}
 % \end{center}
-
-\begin{lemma}[Monitor Satisfaction is $\mssafe$]\label{lem:mon:msafe}
-  If $\monmssafe{\trace}$, then $\trace\in\mssafe$.\Coqed
-\end{lemma}
-
-Any technique to combine existing monitors $T_A$, $T_B$ (that check for properties $A$ and $B$) into one $T_{AB}$ (that checks for both $A$ and $B$) is sufficient for the sake of the analysis of this paper.
-
-\subsubsection{Monitor for \glsfirst*{scct}}
-\begin{gather*}
-% \small
-  \begin{aligned}
-    \mi{(Abstract\ Store)}~\scctmonitor:=&\ \emptyset 
-    % &
-    &
-    \qquad\mi{(Abstract\ Events)}~\absevent:=&\ \absev{\varepsilon} \mid \absterm \mid \absev{Leak} 
-    % \\
-  \end{aligned}
-\end{gather*}
-\begin{center}
-  \judgbox{\monitorcheck[]{\scctmonitor}{\scctmonitor[']}{\absevent}}{,,Monitor $\scctmonitor$ steps to $\scctmonitor[']$\\given event $\absevent$.''}$\;$\\
-  %
-  \typerule{scct-none}{
-  }{
-    \monitorcheck[]{\scctmonitor}{\scctmonitor}{\absev{\emptyevent}}
-  }{scct-none}
-  %
-  \typerule{scct-abort}{
-  }{
-    \monitorcheck[]{\scctmonitor}{\scctmonitor}{\absterm}
-  }{scct-abort}
-\end{center}
-The monitor state for the \gls*{scct} monitor is completely empty, since it does not need to keep track of information.
-% As soon as 
-If
-any event is hit, the execution gets stuck, since any event is considered confidental from the perspective of this monitor.
-\begin{center}
-  \judgbox{\sccttraceagree[]{\event}{\absevent}}{,,Abstract event $\absevent$ is equivalent to $\event$ w.r.t. \gls*{cct}.''}$\;$\\
-  %
-  \typerule{scct-low-authentic}{
-  }{
-    \sccttraceagree[]{\event[_b];\sandboxtag;\unlock}{\absev{\emptyevent}}
-  }{scct-low-authentic}
-  %
-  \typerule{scct-high-authentic}{
-  }{
-    \sccttraceagree[]{\event[_b];\sandboxtag;\lock}{\absev{Leak}}
-  }{scct-high-authentic}
-  %
-  % \typerule{scct-empty-authentic}{
+% The monitor state for the \gls*{scct} monitor is completely empty, since it does not need to keep track of information.
+% % As soon as 
+% If
+% any event is hit, the execution gets stuck, since any event is considered confidental from the perspective of this monitor.
+% \begin{center}
+%   \judgbox{\sccttraceagree[]{\event}{\absevent}}{,,Abstract event $\absevent$ is equivalent to $\event$ w.r.t. \gls*{cct}.''}$\;$\\
+%   %
+%   \typerule{scct-low-authentic}{
 %   }{
-  %   \sccttraceagree[]{\emptyevent}{\absev{\emptyevent}}
-  % }{scct-empty-authentic}
-  % %
-  % \typerule{scct-abort-authentic}{
+%     \sccttraceagree[]{\event[_b];\sandboxtag;\unlock}{\absev{\emptyevent}}
+%   }{scct-low-authentic}
+%   %
+%   \typerule{scct-high-authentic}{
 %   }{
-  %   \sccttraceagree[]{\lightning}{\absterm}
-  % }{scct-abort-authentic}
-\end{center}
-Accordingly, the event agreement simply disregards all events that involved public data ($\unlock$) while mapping any other event that does involve private data ($\lock$) to the abstract $\absev{Leak}$ event.
-
-\begin{lemma}[Monitor Satisfaction is $\scctsafe$]\label{lem:mon:scctsafe}
-  If $\monscctsafe{\trace}$, then $\trace\in\scctsafe$. %\Coqed
-\end{lemma}
-
-\begin{exampleenv}[Data-independent timing mode]\label{ex:ditm}
-  Consider the function call \texttt{strncpy(1,x,y)} to the \texttt{strncpy} function (\Cref{ex:strncpy}) with low ($\unlock$) security for \texttt{x} and high security ($\lock$) for \texttt{y}.
-  The trace for the copying part of \texttt{strncpy} is:
-
-\vspace{-1em}
-  {\small
-  \begin{align*}
-  	\ev{Use}\ \loc_{x}\ 0;\comp;\unlock\cdot\ev{Use}\ \loc_{y}\ 0;\comp;\lock	
-  \end{align*}
-  }
-  In terms of the abstract monitor trace, this is just $\absev{Leak}$.
-  Running this on the monitor would result in getting stuck, since there is no matching rule to step in the presence of $\absev{Leak}$ event. % pun intended
-  By means of additional features to ensure cryptographic constant time even in the presence of memory reads and loads, such as a flag to enable a data independent timing mode, which is present in both Arm~\cite[p.~543]{arm-refman} and Intel~\cite[p.~80]{intel-refman} processors, the original trace now does $\ev{Use}\ \loc_{y}\ 0;\comp;\unlock$ instead of $\ev{Use}\ \loc_{y}\ 0;\comp;\lock$.
-  Because of this, the whole trace of the component is $\absev{\emptyevent}$ and the monitor can step without getting stuck.
-\end{exampleenv}
-
-\subsubsection{Combining \gls*{ms} and \gls*{scct} Monitors to obtain \gls*{msscct}}
-
-The combination of monitors for \gls*{ms} and \gls*{scct} yields one for \gls*{msscct}.
-The construction is entirely similar to the one for \gls*{ms} (\Cref{subsubsec:ms}).
-
-\begin{lemma}[Monitor Satisfaction is $\msscctsafe$]\label{lem:mon:msscctsafe}
-  % $\;$\\\vspace{-1em}
-  % \begin{nscenter}
-  % \noindent
-  If $\monmsscctsafe{\trace}$, then $\trace\in\msscctsafe$. %\Coqed
-  % \end{nscenter}
-\end{lemma}
-
-\section{Composing Secure Compilers\pages{1}}\label{sec:compcomp}
-
-This section presents the key meta-theoretic results of this paper concerning secure compiler compositions (\Cref{sec:sequential,sec:other-compos}).
-% Note that the presented framework works with arbitrary classes.
-% \MP{check, we ened more than this}
-% and builds on prior work~\cite{kruse2022csc}\MK{mention this at all?
+%     \sccttraceagree[]{\event[_b];\sandboxtag;\lock}{\absev{Leak}}
+%   }{scct-high-authentic}
+%   %
+%   % \typerule{scct-empty-authentic}{
+%   % }{
+%   %   \sccttraceagree[]{\emptyevent}{\absev{\emptyevent}}
+%   % }{scct-empty-authentic}
+%   % %
+%   % \typerule{scct-abort-authentic}{
+%   % }{
+%   %   \sccttraceagree[]{\lightning}{\absterm}
+%   % }{scct-abort-authentic}
+% \end{center}
+% Accordingly, the event agreement simply disregards all events that involved public data ($\unlock$) while mapping any other event that does involve private data ($\lock$) to the abstract $\absev{Leak}$ event.
+
+% \begin{lemma}[Monitor Satisfaction is $\scctsafe$]\label{lem:mon:scctsafe}
+%   If $\monscctsafe{\trace}$, then $\trace\in\scctsafe$. %\Coqed
+% \end{lemma}
+
+% \begin{exampleenv}[Data-independent timing mode]\label{ex:ditm}
+%   Consider the function call \texttt{strncpy(1,x,y)} to the \texttt{strncpy} function (\Cref{ex:strncpy}) with low ($\unlock$) security for \texttt{x} and high security ($\lock$) for \texttt{y}.
+%   The trace for the copying part of \texttt{strncpy} is:
+
+% \vspace{-1em}
+%   {\small
+%   \begin{align*}
+%     \ev{Use}\ \loc_{x}\ 0;\comp;\unlock\cdot\ev{Use}\ \loc_{y}\ 0;\comp;\lock	
+%   \end{align*}
+%   }
+%   In terms of the abstract monitor trace, this is just $\absev{Leak}$.
+%   Running this on the monitor would result in getting stuck, since there is no matching rule to step in the presence of $\absev{Leak}$ event. % pun intended
+%   By means of additional features to ensure cryptographic constant time even in the presence of memory reads and loads, such as a flag to enable a data independent timing mode, which is present in both Arm~\cite[p.~543]{arm-refman} and Intel~\cite[p.~80]{intel-refman} processors, the original trace now does $\ev{Use}\ \loc_{y}\ 0;\comp;\unlock$ instead of $\ev{Use}\ \loc_{y}\ 0;\comp;\lock$.
+%   Because of this, the whole trace of the component is $\absev{\emptyevent}$ and the monitor can step without getting stuck.
+% \end{exampleenv}
+
+% \subsubsection{Combining \gls*{ms} and \gls*{scct} Monitors to obtain \gls*{msscct}}
+
+% The combination of monitors for \gls*{ms} and \gls*{scct} yields one for \gls*{msscct}.
+% The construction is entirely similar to the one for \gls*{ms} (\Cref{subsubsec:ms}).
+
+% \begin{lemma}[Monitor Satisfaction is $\msscctsafe$]\label{lem:mon:msscctsafe}
+%   % $\;$\\\vspace{-1em}
+%   % \begin{nscenter}
+%   % \noindent
+%   If $\monmsscctsafe{\trace}$, then $\trace\in\msscctsafe$. %\Coqed
+%   % \end{nscenter}
+% \end{lemma}
+
+% \section{Composing Secure Compilers\pages{1}}\label{sec:compcomp}
+
+% This section presents the key meta-theoretic results of this paper concerning secure compiler compositions (\Cref{sec:sequential,sec:other-compos}).
+% % Note that the presented framework works with arbitrary classes.
+% % \MP{check, we ened more than this}
+% % and builds on prior work~\cite{kruse2022csc}\MK{mention this at all?
 may reveal identity even though its double-blind}.
 % MP: not needed at all, ever, workshop work does not need to be cited in the main work
-Note that since it deals with the preservation of \emph{classes}, which include both trace properties as well as hyperproperties, the presented framework is general: It describes robust compilers preserving ordinary trace properties, hyperproperties or even relational hyperproperties.
-% \MP{pull this out to earlier}
+% Note that since it deals with the preservation of \emph{classes}, which include both trace properties as well as hyperproperties, the presented framework is general: It describes robust compilers preserving ordinary trace properties, hyperproperties or even relational hyperproperties.
+% % \MP{pull this out to earlier}
 
@@
 
-\begin{figure}[!h]
+\begin{figure*}[!h]
  \vspace{-2em}
@@
   \label{fig:ex-cct}
-\end{figure}
+\end{figure*}
 
@@
 This section defines several secure compilers, each of which robustly preserves a different property of interest as depicted in \Cref{fig:pipeline}.
-\begin{figure}[!h]
+\begin{figure*}[!h]
   \centering
@@
     \ccb(\trg{\lnew{x}{\expr[_{1}]}{\expr[_{2}]}}) & = 
-    	& \llet[\irl]{\irl{x_{SIZE}}}{\ccb(\trg{\expr[_{1}]})}{
-    		% \\&&
+                                                   & \llet[\irl]{\irl{x_{SIZE}}}{\ccb(\trg{\expr[_{1}]})}{}
+    		\\&&
     		\lnew[\irl]{\irl{x}}{\irl{x_{SIZE}}}{\ccb(\trg{\expr[_{2}]})}
-    		} \\
+    		 \\
   \ccb(\trg{\lget{x}{\expr}}) & = 
-  	& \llet[\irl]{\irl{x_{n}}}{\ccb(\trg{\expr})}{
-  	% \\&&
-  	\lifz[\irl]{\irl{0\le x_{n}<x_{SIZE}}}{\irl{\lget{x}{x_{n}}}}{
-  		% \\&&
+                              & \llet[\irl]{\irl{x_{n}}}{\ccb(\trg{\expr})}{}
+  	\\&&
+  \lifz[\irl]{\irl{0\le x_{n}<x_{SIZE}}}{\\&&\irl{\lget{x}{x_{n}}}}{}
   		\irl{\labort}
-  		}
-  		} \\
+  	  \\
   \ccb(\trg{\lset{x}{\expr[_{1}]}{\expr[_{2}]}}) & = 
-  	& \llet[\irl]{\irl{x_{n}}}{\ccb(\trg{\expr[_{1}]})}{
+                                                 & \llet[\irl]{\irl{x_{n}}}{\ccb(\trg{\expr[_{1}]})}{}
   		\\&&
-  		\lifz[\irl]{\irl{0\le x_{n}<x_{SIZE}}}{\lset[\irl]{\irl{x}}{\irl{x_{n}}}{
-  		% \\&&
+  \lifz[\irl]{\irl{0\le x_{n}<x_{SIZE}}}{\\&&\lset[\irl]{\irl{x}}{\irl{x_{n}}}{}
   		\ccb(\trg{\expr[_{2}]})
-  		}}{\irl{\labort}}
-  		} 
+  		}{\irl{\labort}} 
   		% \\
@@
     \begin{lstlisting}[language=c,basicstyle=\ttfamily\footnotesize, morekeywords={size_t}]
-void strncpy(size_t n, size_t dst_size, char *dst, size_t src_size, char *src) {
+void strncpy(size_t n, size_t dst_size, char *dst,
+             size_t src_size, char *src) {
   for(size_t i = 0; i < src_size && src[i] != '\0' && i < n; ++i) {
@@
     \cccf(\irl{\expr}) & = \partialeval{\irl{\expr}}{\irl{\hole{\cdot}}} &
-  % \end{align*}
-  \\
-  % [0.125cm]
-% \begin{align*}
-%  \partialeval{\irl{x}}{\irl{\substlist}} & = \irl{n} 
-%  	\qquad\qquad \text{if } \irl{\subst{n}{x}}\in\irl{\substlist} \\
-%  \partialeval{\irl{x}}{\irl{\substlist}} & = \irl{x} 
-%  \qquad\qquad \text{if } \irl{\subst{n}{x}}\notin\irl{\substlist} \\
-%  \partialeval{\irl{\lbinop{n}{m}}}{\irl{\substlist}} & = \irl{k} 
-%  \qquad\qquad \text{if } \lbinop{\irl{n}}{\irl{m}}=k \\
-   % \partialeval{\irl{\lbinop{\expr[_{1}]}{\expr[_{2}]}}}{\irl{\substlist}} & = \lbinop[\irl]{\partialeval{\irl{\expr[_{1}]}}{\irl{\substlist}}}{\partialeval{\irl{\expr[_{2}]}}{\irl{\substlist}}} \\
-%  \partialeval{\irl{\llet{x}{\valueexpr}{\expr}}}{\irl{\substlist}} & = \partialeval{\irl{\expr}}{\irl{\subst{x}{\valueexpr}\cdot\substlist}} 
+  \end{align*}
+  \\[0.125cm]
+  \begin{align*}
+   \partialeval{\irl{x}}{\irl{\substlist}} & = \irl{n} 
+   	\qquad\qquad \text{if } \irl{\subst{n}{x}}\in\irl{\substlist} \\
+   \partialeval{\irl{x}}{\irl{\substlist}} & = \irl{x} 
+   \qquad\qquad \text{if } \irl{\subst{n}{x}}\notin\irl{\substlist} \\
+   \partialeval{\irl{\lbinop{n}{m}}}{\irl{\substlist}} & = \irl{k} 
+   \qquad\qquad \text{if } \lbinop{\irl{n}}{\irl{m}}=k \\
+   %\partialeval{\irl{\lbinop{\expr[_{1}]}{\expr[_{2}]}}}{\irl{\substlist}} & = \lbinop[\irl]{\partialeval{\irl{\expr[_{1}]}}{\irl{\substlist}}}{\partialeval{\irl{\expr[_{2}]}}{\irl{\substlist}}} \\
+   \partialeval{\irl{\llet{x}{\valueexpr}{\expr}}}{\irl{\substlist}} & = \partialeval{\irl{\expr}}{\irl{\subst{x}{\valueexpr}\cdot\substlist}} 
   % \\
   % \partialeval{\irl{\lget{x}{\expr}}}{\irl{\substlist}} & = \lget[\irl]{\irl{x}}{\partialeval{\irl{\expr}}{\irl{\substlist}}}
+% \end{align*}
 \end{align*}
@@
 
-% \MP{mix ??}
 The two optimising compiler passes from $\Lms$ to $\Lms$ perform \gls*{dce} and \gls*{cf}, respectively.
@@
 
-\bibliographystyle{splncs04}
+\bibliographystyle{IEEEtran}
 \bibliography{main}

### 2024-05-14 csf2/main.tex
remove projection figures

@@
 
-\begin{center}\small
-  \judgbox{\msfilterLtms{\src{\event}}=\event}{,,Project an $\Ltms$ event $\src{\event}$ to\\a specification event $\event$.''}
-  \typerule{$\Ltms$-filter-context}{
-    \src{\event[_b]}\not=\src{\lightning}
-  }{
-    \msfilterLtms{\src{\tup{\event[_b]}{\ctx,\securitytag}}} = \emptyevent
-  }{ltms-filter-context}
-  %
-  \typerule{$\Ltms$-filter-abort}{
-  }{
-    \msfilterLtms{\src{\lightning}} = \lightning
-  }{ltms-filter-abort}
-  %
-  \typerule{$\Ltms$-filter-start}{
-  }{
-    \msfilterLtms{\src{\tup{Start}{\comp}}} = \emptyevent
-  }{ltms-filter-start}
-  %
-  \typerule{$\Ltms$-filter-alloc}{
-    \locmapsto{\src{\loc}}{\loc} &
-    \src{n}=n
-  }{
-    \msfilterLtms{\src{\tup{Alloc\ \loc\ n}{\comp}}} = \tup{\ev{Alloc}\ \loc\ n}{\comp,\unlock}
-  }{ltms-filter-alloc}
-\end{center}
-
 Projection $\msfilterLtms{\src{\event}}$ translates an action into its correspondent (e.g., \Cref{tr:ltms-filter-alloc}); since these rules are straightforward they are mostly omitted.
@@
 %To mark this mode as active or inactive, \Cref{tr:e-wrdoit} allows to access the register in $\obj{\statevar}$.
-
-\begin{center}
-  \judgbox{\scctfilterLscct{\obj{\event}}=\event}{,,Project an $\Lscct$ event $\obj{\event}$\\to a specification event $\event$.''}
-  \typerule{$\Lscct$-filter-context}{
-    \obj{\event[_b]}\not=\obj{\lightning}
-  }{
-    \scctfilterLscct{\obj{\tup{\event[_b]}{\ctx,\securitytag}} = \emptyevent}
-  }{lscct-filter-context}
-  %
-  \typerule{$\Lscct$-filter-$\widehat{\text{get}}$}{
-    \locmapsto{\obj{\loc}}{\loc} &
-    \obj{n}=n
-  }{
-    \scctfilterLscct{\obj{\tup{\widehat{Get}\ \loc\ n}{\comp,\securitytag}}} = \tup{\ev{Get}\ \loc\ n}{\comp,\unlock}
-  }{lscct-filter-get-hat}
-  %
-  \typerule{$\Lscct$-filter-$\text{get}$}{
-    \locmapsto{\obj{\loc}}{\loc} &
-    \obj{n}=n
-  }{
-    \scctfilterLscct{\obj{\tup{Get\ \loc\ n}{\comp,\securitytag}}} = \tup{\ev{Get}\ \loc\ n}{\comp,\securitytag}
-  }{lscct-filter-get}
-\end{center}
-
-The projection to the specification trace model is mostly straightforward and similar to the others, e.g., \Cref{subsec:cs:tms}.
-However, for events containing the pre-events $\obj{\widehat{Get}}$ and $\obj{\widehat{Set}}$, the projection always translates the security-tag $\obj{\sigma}$ to $\unlock$, regardless of its actual value, as seen in \Cref{tr:lscct-filter-get-hat}.
-The pre-events themselves still translate to just $\ev{Get}$ and $\ev{Set}$, respectively.
-With this technical setup, the information whether a read or write happened on a secret value is not hidden by the semantics, e.g., by emitting $\obj{\emptyevent}$, but when projecting to specification events.
-This allows flexibility: The trace can be checked to satisfy different properties, such as, in this case, \gls*{tms}, \gls*{sms}, \gls*{scct}, and their combined versions.
 

### 2024-05-14 csf2/main.tex
minor tweaks

@@
 
-Our main result is that secure compilers in the robust compilation framework~\cite{abate2019jour} compose {\em sequentially}.
+The main result is that secure compilers in the robust compilation framework~\cite{abate2019jour} compose {\em sequentially}.
+This is true even in the context of compilers that map from one trace model to another~\cite{abate2021extacc}.
 % This is not intuitive in the sense that in the security domain, composition does not work without additional generalizations~\cite{canetti2006univcomp,mccullough2012compo,fabian2022automatic}.

### 2024-05-14 csf2/main.tex
move around a bunch of defs

--- 2024-05-14 csf2/main.tex
@@
 Unfortunately, correct compilers may be insecure compilers~\cite{patrignani2019survey,kennedy2006secure.net,abadi1999protect,ahmed2018dagstuhl} and programs translated by insecure compilers can violate security properties that the programmer assumes to hold.
-To define when a compiler is secure, this paper uses the robust compilation framework~\cite{abate2019jour}, which the following definition summarises.
-
-For \bul{compilers $\cc{\src{L}}{\trg{L}}$ to robustly preserve a class of properties $\class$}, if for any \rul{property $\pi$ of that class $\class$ and programs $\src{p}$ written in $\src{L}$} where \iul{$\src{\progvar}$ robustly satisfies $\pi$}, then \oul{the compilation of $\src{\progvar}$, $\cc{\src{L}}{\trg{L}}\left(\src{p}\right)$, must robustly satisfy $\pi$}.
+This is why {\em robust preservation} is a good candidate as a compiler-level security property~\cite{abate2019jour}.
+
+This paper uses a more general notion~\cite{abate2021extacc} that applies to compilers translating from one trace model into another.
+To this end, considering a source trace $\src{\trace}$ and a target trace $\trg{\trace}$, there is a relation ($\sim$) describing the effect of a corresponding compiler. 
+\MKin{reference example from case study}
+Using this relation, there are two projection functions: (1) the existential image $\tau_\sim\left(\src{\pi}\right)$ and (2) the universal image $\sigma_\sim\left(\trg{\pi}\right)$.
+These projections map source-level (resp. target-level) properties to target-level (resp. source-level) properties and can be lifted to work with classes, too, by taking the powerset.
+
+For a \bul{compiler $\cc{\src{L}}{\trg{L}}$ to robustly preserve a class of source properties $\src{\class}$}, if for any \rul{property $\src{\pi}$ of that class $\src{\class}$ and programs $\src{p}$ written in $\src{L}$} where \iul{$\src{\progvar}$ robustly satisfies $\src{\pi}$}, then \oul{the compilation of $\src{\progvar}$, $\cc{\src{L}}{\trg{L}}\left(\src{p}\right)$, must robustly satisfy $\tau_\sim\left(\src{\pi}\right)$}.
+
+\begin{definition}[Robust Compilation with $\tau_\sim$]\label{def:rtp:tau}
+  % $\;$\\
+  % \vspace{-1em}
+  % \begin{nscenter}
+  % Compiler $\cc{\src{L}}{\trg{L}}$ robustly preserves $\class$, 
+  \bul{$\rtptau{\cc{\src{L}}{\trg{L}}}{\src{\class}}{\sim}$}
+  %, iff 
+  $\isdef$
+  \rul{$\forall \src{\pi}\in\src{\class}, \src{p}\in\src{L},$} if \iul{$\rsat{\src{\progvar}}{\src{\pi}}$}, then \oul{$\rsat{\cc{\src{L}}{\trg{L}}\left(\src{p}\right)}{\tau_\sim{\left(\src{\pi}\right)}}$}.
+  % \end{nscenter}
+\end{definition}
+% 
+Note that a class of properties $\class$ can represent just one property $\pi$ by lifting~\cite{clarkson2008hyper} that property to sets of properties, i.e., use the powerset of $\pi$ instead of $\pi$ itself.
+Because of this, this paper may write $\rtptau{\cc{\src{L}}{\trg{L}}}{\src{\pi}}{\sim}$, even though $\src{\pi}$ is a property and not a class.
+
+Dually to \Cref{def:rtp:tau}, for a \bul{compiler $\cc{\src{L}}{\trg{L}}$ to robustly preserve a class of target properties $\trg{\class}$}, if for any \rul{property $\trg{\pi}$ of that class $\trg{\class}$ and programs $\src{p}$ written in $\src{L}$} where \iul{$\src{\progvar}$ robustly satisfies $\sigma_\sim\left(\trg{\pi}\right)$}, then \oul{the compilation of $\src{\progvar}$, $\cc{\src{L}}{\trg{L}}\left(\src{p}\right)$, must robustly satisfy $\trg{\pi}$}.
+
+\begin{definition}[Robust Compilation with $\sigma_\sim$]\label{def:rtp:sigma}
+  % $\;$\\
+  % \vspace{-1em}
+  % \begin{nscenter}
+  % Compiler $\cc{\src{L}}{\trg{L}}$ robustly preserves $\class$, 
+  \bul{$\rtpsig{\cc{\src{L}}{\trg{L}}}{\trg{\class}}{\sim}$}
+  %, iff 
+  $\isdef$
+  \rul{$\forall \trg{\pi}\in\trg{\class}, \src{p}\in\src{L},$} if \iul{$\rsat{\src{\progvar}}{\sigma_\sim\left(\trg{\pi}\right)}$}, then \oul{$\rsat{\cc{\src{L}}{\trg{L}}\left(\src{p}\right)}{\trg{\pi}}$}.
+  % \end{nscenter}
+\end{definition}
+
+A degenerate case where $\sim$ coincides with equality, i.e., the trace model is the same for both source and target programs, is the classic definition of robust preservation~\cite{abate2019jour}:
 
@@
 \end{definition}
-% 
-Note that a class of properties $\class$ can represent just one property $\pi$ by lifting~\cite{clarkson2008hyper} that property to sets of properties, i.e., use the powerset of $\pi$ instead of $\pi$ itself.
-Thus we write $\rtp{\cc{\src{L}}{\trg{L}}}{\pi}$, even though $\pi$ is a property and not a class.
+
 % \MP{
@@
 % \end{exampleenv}
+
+\subsection{Secure Sequential Composition}\label{sec:sequential}
+
+Notably, real-world compilation chains also perform a series of (sequential) passes whose main purpose is not necessarily to translate from one language to another, but to, e.g., optimise the code or enforce a certain property.
+Both examples can be seen in practice, e.g. as in the work of~\cite{nagarakatte2009soft,nagarakatte2010cets,akritidis2009baggy,wegman1991ccp,manjikian1997fusion} and many more.
+% \MPin{survey to cite?
+this is a bit informal MK: hard to find a survey on different compiler passes...
+I can't even find one just for compiler enforced memsafety (that looks good) or for compiler translations in general}.
+% 
+% Consider the compilation passes of LLVM~\cite{lattner2004llvm}, which are handled in the \texttt{PassManager} class.
+% By manipulating the internal state of a \texttt{PassManager} object, the schedule of compiler optimisation passes can be hardcoded.
+% That way, an order of optimisations can be found by hand that yields the best results.
+% \MP{what's the point of the previous 3 sentences?(commented)}
+Consider the following two LLVM optimisation passes: \gls*{cf}, which rewrites constant expressions to the constant they evaluate to, and \gls*{dce}, which removes dead code by rewriting conditional branches.
+The order in which \gls*{cf} and \gls*{dce} are performed influences the final result of the compilation (see \Cref{fig:cfdceex}).
+\begin{figure}[!ht]
+  \begin{center}
+    \begin{tikzpicture}[scale=0.62, every node/.style={transform shape}]
+      \node[minimum width=3cm,minimum height=2.5cm,draw=black,very thick,align=left] (progunopt) {\begin{lstlisting}[language=ml,basicstyle=\small\ttfamily]
+let a = true in
+if a then
+  print "a"
+else
+  print "b"
+\end{lstlisting}};
+      \node[minimum width=3cm,minimum height=2.5cm,draw=black,very thick,align=left,right=1.0 of progunopt] (progoptcf) {\begin{lstlisting}[language=ml,basicstyle=\large\ttfamily]
+if true then
+  print "a"
+else
+  print "b"
+\end{lstlisting}};
+      \node[minimum width=3cm,minimum height=2.5cm,draw=black,very thick,align=left,right=1.0 of progoptcf] (progoptcfdce) {\begin{lstlisting}[language=ml,basicstyle=\Large\ttfamily]
+print "a"
+\end{lstlisting}};% the following is a hack to get my syntax highlighting to work again \end{lstlisting}
+      % edges
+      \draw[->,very thick] (progunopt) edge[loop left,left] node {\gls*{dce}} (progunopt);
+      \draw[->,very thick] (progunopt) edge[sloped,above] node {\gls*{cf}} (progoptcf);
+      \draw[->,very thick] (progoptcf) edge[sloped,above] node {\gls*{dce}} (progoptcfdce);
+    \end{tikzpicture}
+  \end{center}
+  \caption{Example program where the level of optimisations differ for one pass of applying \gls*{cf} and \gls*{dce} in any order. %
+  Every edge is a compilation pass and the label on the edge states what the pass does, i.e., \gls*{cf} or \gls*{dce}. %
+  The source code in the nodes is a glorified compiler intermediate representation and the code gets more optimised towards the right hand side of the figure.}\label{fig:cfdceex}
+\end{figure}
+This {\em phase ordering problem} is well--known in literature and a practical solution is to simply perform a fixpoint iteration of the optimisation pipeline~\cite{click1995combining}.
+
+Suppose there are traces $\src{\trace}$ and $\trg{\trace}$ as well as two trace relations $\sim_1$ and $\sim_2$. 
+The traces are related $\src{\trace}\sim_1\bullet\sim_2\trg{\trace}$ if there is another trace $\irl{\trace}$ such that $\src{\trace}\sim_1\irl{\trace}$ and $\irl{\trace}\sim_2\trg{\trace}$.
+
+\bul{A cross-language relation $\sim$ is well-formed with respect to a class of source-level properties $\src{\class}$} iff \rul{the existential image preserves set membership}.
+\begin{definition}[Well-formedness of Trace Relations for a Class $\src{\class}$]\label{def:wfc:tau:tracerel}
+  % $$
+
+  \noindent
+  \text{\bul{$\wfctau{\sim}{\src{\class}}$}} \isdef \text{\rul{$\forall \src{\Pi}\in\src{\class}, \tau_\sim(\src{\Pi})\in\tilde{\tau}_\sim(\src{\class})$}}
+  % $$
+\end{definition}
+
+\bul{A cross-language relation $\sim$ is well-formed with respect to a class of target-level properties $\trg{\class}$} iff \rul{the universal image preserves set membership}.
+\begin{definition}[Well-formedness of Trace Relations for a Class $\trg{\class}$]\label{def:wfc:sig:tracerel}
+  % $$
+
+  \noindent
+  \text{\bul{$\wfcsig{\sim}{\trg{\class}}$}} \isdef \text{\rul{$\forall \trg{\Pi}\in\trg{\class}, \sigma_\sim(\trg{\Pi})\in\sigma_\sim(\trg{\class})$}}
+  % $$
+\end{definition}
+
+With these notions of well-formedness, it is now possible to prove the main result that secure compilers in the robust compilation framework~\cite{abate2021extacc} compose {\em sequentially}.
+
+Let \bul{$\cc{\src{L}}{\trg{L}}$ robustly preserve the class $\src{\class[_1]}$ under $\sim_1$} and let \rul{$\cc{\trg{L}}{\obj{L}}$ robustly preserve the class $\tilde{\tau}_{\sim_1}\left(\src{\class[_2]}\right)$ under $\sim_2$}.
+Then, \iul{when the cross-language relation $\sim_1$ is well-formed with respect to class $\src{\class[_2]}$}, it follows that \oul{the composed compiler $\cc{\src{L}}{\trg{L}}\circ\cc{\trg{L}}{\obj{L}}$ robustly preserves the intersection of classes $\src{\class[_1]}\cap\src{\class[_2]}$ under $\sim_1\bullet\sim_2$}.
+% 
+%For space reasons, \Thmref{thm:urtp} and \Thmref{thm:lrtp} are not presented here, but can be found in the technical appendix ($\CoqSymbol$).
+% Note that it is not necessary to require $\sim_2$ to be well-formed with respect to $\src{\class[_1]}$, since the 
+% \MPin{why}
+\begin{theorem}[Sequential Composition of Secure Compilers with respect to $\tau$]\label{thm:rtpsim:tau}
+  If \bul{$\rtptau{\cc{\src{L}}{\trg{L}}}{\src{\class[_{1}]}}{\sim_1}$}, \rul{$\rtptau{\cc{\trg{L}}{\obj{L}}}{\tilde{\tau}_{\sim_1}\left(\src{\class[_{2}]}\right)}{\sim_2}$}, and \iul{$\wfctau{\sim_1}{\src{\class[_2]}}$}, then \oul{$\rtptau{\cc{\src{L}}{\trg{L}}\circ\cc{\trg{L}}{\obj{L}}}{\src{\class[_{1}]}\cap\src{\class[_{2}]}}{\sim_1\bullet\sim_2}$}.
+\Coqed
+\end{theorem}
+
+Let \bul{$\cc{\src{L}}{\trg{L}}$ robustly preserve the class $\sigma_{\sim_2}\left(\trg{\class[_1]}\right)$ under $\sim_1$} and let \rul{$\cc{\trg{L}}{\obj{L}}$ robustly preserve the class $\trg{\class[_2]}$ under $\sim_2$}.
+Then, \iul{when the cross-language relation $\sim_2$ is well-formed with respect to class $\trg{\class[_1]}$}, it follows that \oul{the composed compiler $\cc{\src{L}}{\trg{L}}\circ\cc{\trg{L}}{\obj{L}}$ robustly preserves the intersection of classes $\trg{\class[_1]}\cap\trg{\class[_2]}$ under $\sim_1\bullet\sim_2$}.
+% 
+%For space reasons, \Thmref{thm:urtp} and \Thmref{thm:lrtp} are not presented here, but can be found in the technical appendix ($\CoqSymbol$).
+% Note that it is not necessary to require $\sim_2$ to be well-formed with respect to $\src{\class[_1]}$, since the 
+% \MPin{why}
+\begin{theorem}[Sequential Composition of Secure Compilers with respect to $\sigma$]\label{thm:rtpsim:sig}
+  If \bul{$\rtpsig{\cc{\src{L}}{\trg{L}}}{\tilde{\sigma}_{\sim_2}\left(\trg{\class[_{1}]}\right)}{\sim_1}$}, \rul{$\rtpsig{\cc{\trg{L}}{\obj{L}}}{\trg{\class[_2]}}{\sim_2}$}, and \iul{$\wfcsig{\sim_1}{\trg{\class[_2]}}$}, then \oul{$\rtpsig{\cc{\src{L}}{\trg{L}}\circ\cc{\trg{L}}{\obj{L}}}{\trg{\class[_{1}]}\cap\trg{\class[_{2}]}}{\sim_1\bullet\sim_2}$}.
+\Coqed
+\end{theorem}
+
+The degenerate case where the cross-language trace relation is equality has a simpler compositionality result:
+Given \bul{$\cc{\src{L}}{\trg{L}}$ robustly preserves $\class[_{1}]$} and \rul{$\cc{\trg{L}}{\obj{L}}$ robustly preserves $\class[_{2}]$}, it follows that \oul{their sequential composition $\cc{\src{L}}{\trg{L}}\circ\cc{\trg{L}}{\obj{L}}$ robustly preserves the intersection of classes $\class[_{1}]$ and $\class[_{2}]$}.
+
+\begin{theorem}[Sequential Composition of Secure Compilers]\label{thm:rtp}
+  If \bul{$\rtp{\cc{\src{L}}{\trg{L}}}{\class[_{1}]}$} and \rul{$\rtp{\cc{\trg{L}}{\obj{L}}}{\class[_{2}]}$}, then \oul{$\rtp{\cc{\src{L}}{\trg{L}}\circ\cc{\trg{L}}{\obj{L}}}{\class[_{1}]\cap\class[_{2}]}$}.
+\Coqed
+\end{theorem}
+
+Since the composition of secure compilers is again a secure compiler, the theorem generalises to a whole chain of $n$ secure compilers.
+
+%If possible, optimisations are merged to enhance the analysis' result, since even when iterating to a fixpoint the final program may not be as optimised as it could be~\cite{click1995combining}.
+Compiler engineers typically try to find an order of optimisations that yields well-optimised programs for either code size~\cite{cooper1999geneticphases} or performance~\cite{kulkarni2006exhaustivephase}.
+\Cref{corr:swappable} justifies that any such order of compilation passes is valid with respect to security.
+So, given two compilation passes \bul{$\cc[_{1}]{\trg{L}}{\trg{L}}$, $\cc[_{2}]{\trg{L}}{\trg{L}}$, both robustly preserving class $\class[_{1}]$ or $\class[_{2}]$, respectively}, \oul{for any order of their composition the composed compiler robustly preserves the intersection of $\class[_{1}]$ and $\class[_{2}]$}.
+% 
+% \MPin{
+% 	later on, say that optimisation passes remove elements from traces, so they should be fine with respect to the result with $\tau$.
+% 	also, yes, if $\tau$ adds we have a problem, if it removes we don't
+% \MKin{ Optimisations that remove elements from traces are not necessarily sound,
+%  e.g., remove all deallocations to "optimise" memory management by offloading
+%  the work to the OS.
+% }
+% }
+
+\begin{corollary}[Sequential Composition of Secure Compilers]\label{corr:swappable}
+  If \bul{$\rtp{\cc[_{1}]{\trg{L}}{\trg{L}}}{\class[_{1}]}$ and $\rtp{\cc[_{2}]{\trg{L}}{\trg{L}}}{\class[_{2}]}$}, then \oul{$\rtp{\cc[_{1}]{\trg{L}}{\trg{L}}\circ\cc[_{2}]{\trg{L}}{\trg{L}}}{\class[_{1}]\cap\class[_{2}]}$ and $\rtp{\cc[_{2}]{\trg{L}}{\trg{L}}\circ\cc[_{1}]{\trg{L}}{\trg{L}}}{\class[_{2}]\cap\class[_{1}]}$}.
+\Coqed
+\end{corollary}
+
+\subsection{Secure Upper and Lower Composition}\label{sec:other-compos}
+Besides sequential composition, there are two other compositions, namely an {\em upper}, i.e., a compiler that takes multiple inputs and yields one output, and a {\em lower} composition, i.e., a compiler that takes one input and yields multiple outputs.
+\bul{Define the upper composition $\cc{\src{L}+\obj{L}}{\trg{L}}$} as follows:
+Given a program \texttt{p}, its compiled counterpart is obtained by \rul{plugging \texttt{p} into $\cc{\src{L}}{\trg{L}}$ if $\texttt{p}\in\src{L}$} or by \oul{plugging \texttt{p} into $\cc{\obj{L}}{\trg{L}}$ if $\texttt{p}\in\obj{L}$}.
+\begin{definition}[Upper Composition]
+  $$\text{\bul{$\cc{\src{L}+\obj{L}}{\trg{L}}$}}\isdef
+  % Given \texttt{p}, yield
+  \lambda \texttt{p}\ldotp
+  \left\{\mbox{\begin{tabular}{c}
+    {\rul{if $\texttt{p}\in\src{L}$, then $\cc{\src{L}}{\trg{L}}(\texttt{p})$}} \\
+    \mbox{\oul{if $\texttt{p}\in\obj{L}$, then $\cc{\obj{L}}{\trg{L}}(\texttt{p})$}} \\
+  \end{tabular}}\right.$$
+%
+%
+%  $\begin{cases}
+%                                                                            \rul{\cc{\src{L}}{\trg{L}}(\texttt{p}) & \text{if }\texttt{p}\in\src{L}} \\
+%                                                                            \cc{\obj{L}}{\trg{L}}(\texttt{p}) & \text{if }\texttt{p}\in\obj{L}
+%                                                                          \end{cases}$
+\end{definition}
+
+Examples of this are present in industry:
+Consider the Java Virtual Machine bytecode $\trg{JVM BC}$, which is a popular target for programming language designers due to its high performance and relevance in industry.
+Compilers for several programming languages have it as their target language, some popular instances are $\src{Java}$ and $\obj{Kotlin}$.
+Technically speaking, they both compile to class files and $\obj{Kotlin}$ objects are considered to be the same as $\src{Java}$ objects at that point.
+Both languages can be used at the same time in one project~\cite{androidstudio}.
+A compiler that accepts both $\src{Java}$ and $\obj{Kotlin}$ code translating to the same target language or intermediate representation performs a kind of {\em upper} composition.
+Now, the following theorem tells us what happens if these are secure:
+Given \bul{$\cc{\src{L}}{\trg{L}}$ robustly preserves $\class[_{1}]$} and \rul{$\cc{\obj{L}}{\trg{L}}$ robustly preserves $\class[_{2}]$}, it follows that \oul{their upper composition $\cc{\src{L}+\obj{L}}{\trg{L}}$ robustly preserves the intersection of classes $\class[_{1}]$ and $\class[_{2}]$}.
+
+\begin{theorem}[Upper Composition of Secure Compilers]\label{thm:urtp}
+  If \bul{$\rtp{\cc{\src{L}}{\trg{L}}}{\class[_{1}]}$} and \rul{$\rtp{\cc{\obj{L}}{\trg{L}}}{\class[_{2}]}$}, then \oul{$\rtp{\cc{\src{L}+\obj{L}}{\trg{L}}}{\class[_{1}]\cap\class[_{2}]}$}. %\Coqed
+\end{theorem}
+
+Dually, the {\em lower} composition is concerned about compilers that accept the same source but yield different target languages. %and the same theoretical results apply.
+\bul{Define the lower composition $\cc{\src{L}}{\trg{L}+\obj{L}}$} as follows:
+Given a program $\src{p}$, its compiled counterpart is obtained by \rul{plugging $\src{p}$ into $\cc{\src{L}}{\trg{L}}$} or by \oul{plugging \texttt{p} into $\cc{\src{L}}{\obj{L}}$}, respectively, \pul{based on the internal decision}.
+\begin{definition}[Lower Composition]
+  $$\text{\bul{$\cc{\src{L}}{\trg{L}+\obj{L}}$}}\isdef
+  % Given $L$ and $\src{p}$, yield
+  \lambda \src{p}, L\ldotp
+  \left\{\mbox{\begin{tabular}{c}
+    {\pul{if $L=\trg{L}$, then} \rul{$\cc{\src{L}}{\trg{L}}(\src{p})$}} \\
+    \mbox{\pul{if $L=\obj{L}$, then} \oul{$\cc{\src{L}}{\obj{L}}(\src{p})$}} \\
+  \end{tabular}}\right.$$
+%
+%
+%  $\begin{cases}
+%                                                                            \rul{\cc{\src{L}}{\trg{L}}(\texttt{p}) & \text{if }\texttt{p}\in\src{L}} \\
+%                                                                            \cc{\obj{L}}{\trg{L}}(\texttt{p}) & \text{if }\texttt{p}\in\obj{L}
+%                                                                          \end{cases}$
+\end{definition}
+
+Consider two compilers both accepting $\src{LLVM IR}$~\cite{lattner2004llvm} and one of them emits $\trg{x86\_64}$, while the other emits $\obj{ARMv8}$.
+It is intuitive that they are in some sense composed in the LLVM framework, but the decision of when to use one over the other is inherently {\em internal} to the formalisation effort of this kind of composition.
+For example, the user of this compiler provides an explicit flag that instructs to emit $\trg{x86\_64}$ or the framework itself detects the target platform via heuristics, such as supported instructions.
+
+The following theorem demonstrates what happens if the involved compilers are secure:
+Given \bul{$\cc{\src{L}}{\trg{L}}$ robustly preserves $\class[_{1}]$} and \rul{$\cc{\src{L}}{\obj{L}}$ robustly preserves $\class[_{2}]$}, it follows that \oul{their lower composition $\cc{\src{L}}{\trg{L}+\obj{L}}$ robustly preserves the intersection of classes $\class[_{1}]$ and $\class[_{2}]$}.
+
+\begin{theorem}[Lower Composition of Secure Compilers]\label{thm:lrtp}
+  If \bul{$\rtp{\cc{\src{L}}{\trg{L}}}{\class[_{1}]}$} and \rul{$\rtp{\cc{\obj{L}}{\trg{L}}}{\class[_{2}]}$}, then \oul{$\rtp{\cc{\src{L}}{\trg{L}+\obj{L}}}{\class[_{1}]\cap\class[_{2}]}$}. % \Coqed
+\end{theorem}
+
+%Either way, the theoretical results suggest that it is possible to always find a ,,most-general'', secure compiler, given two secure compilers, that robustly preserves the least-upper bound of the classes involved in their compilation process.
+
+\subsection{Applicability of the Compositionality Framework}\label{subsec:app}
+
+So far, this paper assumes a unified trace-model between source and target languages.
+Moreover secure compilers (in the sense of \Cref{def:rtp}) cannot change the behavior of compiled programs.
+It is worth noting that, as our case study in \Cref{sec:casestud:defs} demonstrates, several compilation passes do not change the behavior of compiled programs.
+% in such a way that the projections to the specification trace model become unequal to each other, 
+% i.e., source and target language of the compiler essentially have a different trace model. 
+However, there are instances of passes where this is not the case.
+For example, consider $\rtp{\cc{\src{L}}{\trg{L}}}{\tmssafe}$ and $\rtp{\cc{\trg{L}}{\obj{L}}}{\smssafe}$, where $\cc{\trg{L}}{\obj{L}}$ introduces code that allocates a table that stores bounds information of pointers, but fails to insert code that deallocates this table.
+By \Thmref{def:trace:tmsdef}, this violates $\tmssafe$.
+Therefore, no behavior resulting from programs produced by the $\cc{\trg{L}}{\obj{L}}$ compiler is secure with respect to $\tmssafe$, but only for $\smssafe$.
+$\cc{\trg{L}}{\obj{L}}$ modifies the behavior of compiled code and this section introduces support for these situations in the compositionality framework.
+
+To reason about these kinds of compilers, we rely on existing work~\cite{abate2021extacc} that lifts the presented secure compilation theory to compilers whose source and target trace models are different.
+% These compilers require a Galois connections for several real-world compiler scenarios, such as a compilation pass that exploits undefined behavior.
+% This section is concerned about a compilation setting with different source and target trace models. 
+These compilers require a way to say that two traces $\src{\trace}$ and $\trg{\trace}$ are related ($\sim$) and a function to translate a source-level property $\src{\pi}$ into a target-level property ($\tau_\sim\left( \src{\pi} \right)$~\cite{abate2021extacc}) using this trace relation.
+The function to map source-level properties to target-level properties can be lifted to translate source-level classes of properties ($\src{\class}$) to target-level classes ($\tilde{\tau}_\sim\left(\src{\class}\right)$).
+The section refers to these functions as existential images.
+
+With this, \bul{a cross-language relation $\sim$ is well-formed with respect to a class of source-level properties $\src{\class}$} iff \rul{the existential images preserve set inclusion}.
+\begin{definition}[Well-formedness of Trace Relations for a Class $\src{\class}$]\label{def:wfc:tracerel}
+  % $$
+
+  \noindent
+  \text{\bul{$\wfc{\sim}{\src{\class}}$}} \isdef \text{\rul{$\forall \src{\Pi}\subseteq\src{\class}, \tau_\sim(\src{\Pi})\subseteq\tilde{\tau}_\sim(\src{\class})$}}
+  % $$
+\end{definition}
+
+\Cref{def:rtpsim} shows how to adapt \Thmref{def:rtp} to account for different target-level trace-models~\cite{abate2021extacc}:
+
+\begin{definition}[Robust Compilation with $\sim$]\label{def:rtpsim}
+  $
+  \rtpsim{\cc{\src{L}}{\trg{L}}}{\src{\class}}{\sim} \isdef
+  \forall \src{\pi} \src{p}.
+    \text{if }\rsat{\src{\progvar}}{\src{\pi}},
+    \text{then }\rsat{\cc{\src{L}}{\trg{L}}\left(\src{p}\right)}{\tilde{\tau}_\sim\left(\src{\pi}\right)}.
+  $
+\end{definition}
+
+To state a similar theorem as \Thmref{thm:rtp} but for \Cref{def:rtpsim}, it is necessary to define composition of cross-language trace relations.
+
+The reason we need assumptions about both relations is motivated by \Cref{lem:useless}.
+% 
+In fact, if \bul{a relation $\sim_1$ is well-formed for a property $\class$} but \rul{another relation $\sim_2$ is not well-formed for the property mapped with the first relation $\tilde{\tau}_{\sim_1}\class$} then \iul{their composition is empty}.
+% 
+An empty intersection means that there is no correct behaviour to preserve for the compiler, which nullifies its security relevance.
+\begin{lemma}[Composition is Useless]\label{lem:useless}
+  If \bul{$\wf{\sim_1}{\class}$}, \rul{$\neg\left(\wf{\sim_2}{\tilde{\tau}_{\sim_1}\class}\right)$}, then \iul{$\tilde{\tau}_{\sim_1\bullet\sim_2}\left(\class\right)=\emptyset$}. \Coqed
+\end{lemma}
+
+Let us now reconsider the example from the beginning of the section, suppose $\sim_{{\smssafe}}$ is a suitable trace relation such that $\trg{\trace}\sim_{{\smssafe}}\irl{\trace}$ establishes the fact that $\irl{\trace}$ contains an allocation of a table containing bounds information which is never deallocated.
+It is impossible that $\wf{\sim_{{\smssafe}}}{{\tmssafe}}$, since every behavior of compiled $\trg{L}$ programs never contains a deallocation of the bounds-information table.
+Still, supposing $\rtpsim{\cc{\src{L}}{\trg{L}}}{\tmssafe}{{\sim_{\tmssafe}}}$, $\rtpsim{\cc{\trg{L}}{\irl{L}}}{\smssafe}{{\sim_{\smssafe}}}$,
+and $\wf{{\sim_{\tmssafe}}}{{\smssafe}}$ allows to apply \Thmref{thm:rtpsim}.
+However, the resulting class $\tilde{\tau}_{{{\sim_{\tmssafe}}\bullet{\sim_{\smssafe}}}}\left(\tmssafe\cap\smssafe\right)=\emptyset$.
+
+\smallskip
+
+For sake of simplicity, the case-study presented in the following sections does not consider different trace-models: The traces of all programs of the presented languages are projected to a unified trace model.
+The case-study demonstrates that, the compositionality framework without the different trace-models is already applicable to a variety of security-relevant passes as well as existing optimisations.
+% may seem overly simplistic, it is useful and simpler in practice, if applicable.
 
@@
   \tmssafe:=\left\{\trace \left| \begin{array}{rcl}
-                              \ev{Alloc\ \loc\ n;\sandboxtag;\securitytag}&\le_{\trace}&\ev{Dealloc\ \loc;\sandboxtag;\securitytag'} \\
-                              \ev{Use\ \loc\ n;\sandboxtag;\securitytag}&\le_{\trace}&\ev{Dealloc\ \loc;\sandboxtag;\securitytag'} \\
-                              \text{if }\ev{Alloc\ \loc\ n;\sandboxtag;\securitytag}\text{ in }\trace&\text{then}&\ev{Dealloc\ \loc;\sandboxtag;\securitytag'}\text{ in }\trace \\
-                              \text{at most one }\ev{Dealloc\ \loc;\sandboxtag;\securitytag}&\text{in}&\trace \\
-                              \text{at most one }\ev{Alloc\ \loc\ n;\sandboxtag;\securitytag}&\text{in}&\trace \\
+                              \ev{Alloc\ \loc\ n;\sandboxtag}&\le_{\trace}&\ev{Dealloc\ \loc;\sandboxtag} \\
+                              \ev{Use\ \loc\ n;\sandboxtag}&\le_{\trace}&\ev{Dealloc\ \loc;\sandboxtag} \\
+                              \text{if }\ev{Alloc\ \loc\ n;\sandboxtag}\text{ in }\trace&\text{then}&\ev{Dealloc\ \loc;\sandboxtag}\text{ in }\trace \\
+                              \text{at most one }\ev{Dealloc\ \loc;\sandboxtag}&\text{in}&\trace \\
+                              \text{at most one }\ev{Alloc\ \loc\ n;\sandboxtag}&\text{in}&\trace \\
                             \end{array}\right.\right\}
@@
   \smssafe:=\left\{\trace \left|\begin{array}{rcl}
-                                  \text{If }\ev{Alloc\ \loc\ n;\sandboxtag;\securitytag}\le_{\trace}\ev{Use\ \loc\ m;\sandboxtag;\securitytag'}, \text{ then }m<n
+                                  \text{If }\ev{Alloc\ \loc\ n;\sandboxtag}\le_{\trace}\ev{Use\ \loc\ m;\sandboxtag}, \text{ then }m<n
                                 \end{array}\right.\right\}
@@
 \end{definition}
+
+\subsubsection{Speculative Safety}
+
+\MKin{fill in the blank based off of spectetor}
 
@@
 % % \MP{pull this out to earlier}
-
-\subsection{Secure Sequential Composition}\label{sec:sequential}
-
-Our main result is that secure compilers in the robust compilation framework~\cite{abate2019jour} compose {\em sequentially}.
-% This is not intuitive in the sense that in the security domain, composition does not work without additional generalizations~\cite{canetti2006univcomp,mccullough2012compo,fabian2022automatic}.
-% \MP{what deos this mean?}
-The \bul{sequential composition of compilers $\cc{\src{L}}{\trg{L}}$ and $\cc{\trg{L}}{\obj{L}}$} is defined as follows:
-Given an $\src{L}$ program $\src{p}$ and compilers $\cc{\src{L}}{\trg{L}},\cc{\trg{L}}{\obj{L}}$, \rul{its compiled $\obj{L}$ counterpart is obtained by plugging $\src{p}$ into $\cc{\src{L}}{\trg{L}}\circ\cc{\trg{L}}{\obj{L}}$}.
-\begin{definition}[Sequential Composition]
-  \bul{$\cc{\src{L}}{\trg{L}}\circ\cc{\trg{L}}{\obj{L}}$} $\isdef$ Given $\src{p}$, yield \rul{$\cc{\trg{L}}{\obj{L}}\left(\cc{\src{L}}{\trg{L}}(\src{p})\right)$}
-\end{definition}
-
-Consider the compilation chain for $\src{TypeScript}$.
-First, $\src{TypeScript}$ programs are translated to $\trg{JavaScript}$ which, e.g., V8~\cite{googlev8} eventually compiles in parts to $\irl{Ignition BC}$.
-The following theorem establishes what happens if all these compilation steps were robustly secure with respect to e.g., \gls*{ms}:
-The resulting $\irl{Ignition BC}$ code would be \gls*{ms} regardless of the context the binary runs in.
-
-Given \bul{$\cc{\src{L}}{\trg{L}}$ robustly preserves $\class[_{1}]$} and \rul{$\cc{\trg{L}}{\obj{L}}$ robustly preserves $\class[_{2}]$}, it follows that \oul{their sequential composition $\cc{\src{L}}{\trg{L}}\circ\cc{\trg{L}}{\obj{L}}$ robustly preserves the intersection of classes $\class[_{1}]$ and $\class[_{2}]$}.
-
-\begin{theorem}[Sequential Composition of Secure Compilers]\label{thm:rtp}
-  If \bul{$\rtp{\cc{\src{L}}{\trg{L}}}{\class[_{1}]}$} and \rul{$\rtp{\cc{\trg{L}}{\obj{L}}}{\class[_{2}]}$}, then \oul{$\rtp{\cc{\src{L}}{\trg{L}}\circ\cc{\trg{L}}{\obj{L}}}{\class[_{1}]\cap\class[_{2}]}$}.
-\Coqed
-\end{theorem}
-
-Since the composition of secure compilers is again a secure compiler, the theorem generalises to a whole chain of $n$ secure compilers.
-
-\subsubsection{Securing Optimisations}
-Notably, real-world compilation chains also perform a series of (sequential) passes whose main purpose is not necessarily to translate from one language to another, but to, e.g., optimise the code or enforce a certain property.
-Both examples can be seen in practice, e.g. as in the work of~\cite{nagarakatte2009soft,nagarakatte2010cets,akritidis2009baggy,wegman1991ccp,manjikian1997fusion} and many more.
-% \MPin{survey to cite?
-this is a bit informal MK: hard to find a survey on different compiler passes...
-I can't even find one just for compiler enforced memsafety (that looks good) or for compiler translations in general}.
-% 
-% Consider the compilation passes of LLVM~\cite{lattner2004llvm}, which are handled in the \texttt{PassManager} class.
-% By manipulating the internal state of a \texttt{PassManager} object, the schedule of compiler optimisation passes can be hardcoded.
-% That way, an order of optimisations can be found by hand that yields the best results.
-% \MP{what's the point of the previous 3 sentences?(commented)}
-Consider the following two LLVM optimisation passes: \gls*{cf}, which rewrites constant expressions to the constant they evaluate to, and \gls*{dce}, which removes dead code by rewriting conditional branches.
-The order in which \gls*{cf} and \gls*{dce} are performed influences the final result of the compilation (see \Cref{fig:cfdceex}).
-\begin{figure}[!ht]
-  \begin{center}
-    \begin{tikzpicture}[scale=0.62, every node/.style={transform shape}]
-      \node[minimum width=3cm,minimum height=2.5cm,draw=black,very thick,align=left] (progunopt) {\begin{lstlisting}[language=ml,basicstyle=\small\ttfamily]
-let a = true in
-if a then
-  print "a"
-else
-  print "b"
-\end{lstlisting}};
-      \node[minimum width=3cm,minimum height=2.5cm,draw=black,very thick,align=left,right=1.0 of progunopt] (progoptcf) {\begin{lstlisting}[language=ml,basicstyle=\large\ttfamily]
-if true then
-  print "a"
-else
-  print "b"
-\end{lstlisting}};
-      \node[minimum width=3cm,minimum height=2.5cm,draw=black,very thick,align=left,right=1.0 of progoptcf] (progoptcfdce) {\begin{lstlisting}[language=ml,basicstyle=\Large\ttfamily]
-print "a"
-\end{lstlisting}};% the following is a hack to get my syntax highlighting to work again \end{lstlisting}
-      % edges
-      \draw[->,very thick] (progunopt) edge[loop left,left] node {\gls*{dce}} (progunopt);
-      \draw[->,very thick] (progunopt) edge[sloped,above] node {\gls*{cf}} (progoptcf);
-      \draw[->,very thick] (progoptcf) edge[sloped,above] node {\gls*{dce}} (progoptcfdce);
-    \end{tikzpicture}
-  \end{center}
-  \caption{Example program where the level of optimisations differ for one pass of applying \gls*{cf} and \gls*{dce} in any order. %
-  Every edge is a compilation pass and the label on the edge states what the pass does, i.e., \gls*{cf} or \gls*{dce}. %
-  The source code in the nodes is a glorified compiler intermediate representation and the code gets more optimised towards the right hand side of the figure.}\label{fig:cfdceex}
-\end{figure}
-This {\em phase ordering problem} is well--known in literature and a practical solution is to simply perform a fixpoint iteration of the optimisation pipeline~\cite{click1995combining}.
-%If possible, optimisations are merged to enhance the analysis' result, since even when iterating to a fixpoint the final program may not be as optimised as it could be~\cite{click1995combining}.
-Compiler engineers typically try to find an order of optimisations that yields well-optimised programs for either code size~\cite{cooper1999geneticphases} or performance~\cite{kulkarni2006exhaustivephase}.
-\Cref{corr:swappable} justifies that any such order of compilation passes is valid with respect to security.
-So, given two compilation passes \bul{$\cc[_{1}]{\trg{L}}{\trg{L}}$, $\cc[_{2}]{\trg{L}}{\trg{L}}$, both robustly preserving class $\class[_{1}]$ or $\class[_{2}]$, respectively}, \oul{for any order of their composition the composed compiler robustly preserves the intersection of $\class[_{1}]$ and $\class[_{2}]$}.
-% 
-% \MPin{
-% 	later on, say that optimisation passes remove elements from traces, so they should be fine with respect to the result with $\tau$.
-% 	also, yes, if $\tau$ adds we have a problem, if it removes we don't
-% \MKin{ Optimisations that remove elements from traces are not necessarily sound,
-%  e.g., remove all deallocations to "optimise" memory management by offloading
-%  the work to the OS.
-% }
-% }
-
-\begin{corollary}[Sequential Composition of Secure Compilers]\label{corr:swappable}
-  If \bul{$\rtp{\cc[_{1}]{\trg{L}}{\trg{L}}}{\class[_{1}]}$ and $\rtp{\cc[_{2}]{\trg{L}}{\trg{L}}}{\class[_{2}]}$}, then \oul{$\rtp{\cc[_{1}]{\trg{L}}{\trg{L}}\circ\cc[_{2}]{\trg{L}}{\trg{L}}}{\class[_{1}]\cap\class[_{2}]}$ and $\rtp{\cc[_{2}]{\trg{L}}{\trg{L}}\circ\cc[_{1}]{\trg{L}}{\trg{L}}}{\class[_{2}]\cap\class[_{1}]}$}.
-\Coqed
-\end{corollary}
-
-\subsection{Secure Upper and Lower Composition}\label{sec:other-compos}
-Besides sequential composition, there are two other compositions, namely an {\em upper}, i.e., a compiler that takes multiple inputs and yields one output, and a {\em lower} composition, i.e., a compiler that takes one input and yields multiple outputs.
-\bul{Define the upper composition $\cc{\src{L}+\obj{L}}{\trg{L}}$} as follows:
-Given a program \texttt{p}, its compiled counterpart is obtained by \rul{plugging \texttt{p} into $\cc{\src{L}}{\trg{L}}$ if $\texttt{p}\in\src{L}$} or by \oul{plugging \texttt{p} into $\cc{\obj{L}}{\trg{L}}$ if $\texttt{p}\in\obj{L}$}.
-\begin{definition}[Upper Composition]
-  $$\text{\bul{$\cc{\src{L}+\obj{L}}{\trg{L}}$}}\isdef
-  % Given \texttt{p}, yield
-  \lambda \texttt{p}\ldotp
-  \left\{\mbox{\begin{tabular}{c}
-    {\rul{if $\texttt{p}\in\src{L}$, then $\cc{\src{L}}{\trg{L}}(\texttt{p})$}} \\
-    \mbox{\oul{if $\texttt{p}\in\obj{L}$, then $\cc{\obj{L}}{\trg{L}}(\texttt{p})$}} \\
-  \end{tabular}}\right.$$
-%
-%
-%  $\begin{cases}
-%                                                                            \rul{\cc{\src{L}}{\trg{L}}(\texttt{p}) & \text{if }\texttt{p}\in\src{L}} \\
-%                                                                            \cc{\obj{L}}{\trg{L}}(\texttt{p}) & \text{if }\texttt{p}\in\obj{L}
-%                                                                          \end{cases}$
-\end{definition}
-
-Examples of this are present in industry:
-Consider the Java Virtual Machine bytecode $\trg{JVM BC}$, which is a popular target for programming language designers due to its high performance and relevance in industry.
-Compilers for several programming languages have it as their target language, some popular instances are $\src{Java}$ and $\obj{Kotlin}$.
-Technically speaking, they both compile to class files and $\obj{Kotlin}$ objects are considered to be the same as $\src{Java}$ objects at that point.
-Both languages can be used at the same time in one project~\cite{androidstudio}.
-A compiler that accepts both $\src{Java}$ and $\obj{Kotlin}$ code translating to the same target language or intermediate representation performs a kind of {\em upper} composition.
-Now, the following theorem tells us what happens if these are secure:
-Given \bul{$\cc{\src{L}}{\trg{L}}$ robustly preserves $\class[_{1}]$} and \rul{$\cc{\obj{L}}{\trg{L}}$ robustly preserves $\class[_{2}]$}, it follows that \oul{their upper composition $\cc{\src{L}+\obj{L}}{\trg{L}}$ robustly preserves the intersection of classes $\class[_{1}]$ and $\class[_{2}]$}.
-
-\begin{theorem}[Upper Composition of Secure Compilers]\label{thm:urtp}
-  If \bul{$\rtp{\cc{\src{L}}{\trg{L}}}{\class[_{1}]}$} and \rul{$\rtp{\cc{\obj{L}}{\trg{L}}}{\class[_{2}]}$}, then \oul{$\rtp{\cc{\src{L}+\obj{L}}{\trg{L}}}{\class[_{1}]\cap\class[_{2}]}$}. %\Coqed
-\end{theorem}
-
-Dually, the {\em lower} composition is concerned about compilers that accept the same source but yield different target languages. %and the same theoretical results apply.
-\bul{Define the lower composition $\cc{\src{L}}{\trg{L}+\obj{L}}$} as follows:
-Given a program $\src{p}$, its compiled counterpart is obtained by \rul{plugging $\src{p}$ into $\cc{\src{L}}{\trg{L}}$} or by \oul{plugging \texttt{p} into $\cc{\src{L}}{\obj{L}}$}, respectively, \pul{based on the internal decision}.
-\begin{definition}[Lower Composition]
-  $$\text{\bul{$\cc{\src{L}}{\trg{L}+\obj{L}}$}}\isdef
-  % Given $L$ and $\src{p}$, yield
-  \lambda \src{p}, L\ldotp
-  \left\{\mbox{\begin{tabular}{c}
-    {\pul{if $L=\trg{L}$, then} \rul{$\cc{\src{L}}{\trg{L}}(\src{p})$}} \\
-    \mbox{\pul{if $L=\obj{L}$, then} \oul{$\cc{\src{L}}{\obj{L}}(\src{p})$}} \\
-  \end{tabular}}\right.$$
-%
-%
-%  $\begin{cases}
-%                                                                            \rul{\cc{\src{L}}{\trg{L}}(\texttt{p}) & \text{if }\texttt{p}\in\src{L}} \\
-%                                                                            \cc{\obj{L}}{\trg{L}}(\texttt{p}) & \text{if }\texttt{p}\in\obj{L}
-%                                                                          \end{cases}$
-\end{definition}
-
-Consider two compilers both accepting $\src{LLVM IR}$~\cite{lattner2004llvm} and one of them emits $\trg{x86\_64}$, while the other emits $\obj{ARMv8}$.
-It is intuitive that they are in some sense composed in the LLVM framework, but the decision of when to use one over the other is inherently {\em internal} to the formalisation effort of this kind of composition.
-For example, the user of this compiler provides an explicit flag that instructs to emit $\trg{x86\_64}$ or the framework itself detects the target platform via heuristics, such as supported instructions.
-
-The following theorem demonstrates what happens if the involved compilers are secure:
-Given \bul{$\cc{\src{L}}{\trg{L}}$ robustly preserves $\class[_{1}]$} and \rul{$\cc{\src{L}}{\obj{L}}$ robustly preserves $\class[_{2}]$}, it follows that \oul{their lower composition $\cc{\src{L}}{\trg{L}+\obj{L}}$ robustly preserves the intersection of classes $\class[_{1}]$ and $\class[_{2}]$}.
-
-\begin{theorem}[Lower Composition of Secure Compilers]\label{thm:lrtp}
-  If \bul{$\rtp{\cc{\src{L}}{\trg{L}}}{\class[_{1}]}$} and \rul{$\rtp{\cc{\obj{L}}{\trg{L}}}{\class[_{2}]}$}, then \oul{$\rtp{\cc{\src{L}}{\trg{L}+\obj{L}}}{\class[_{1}]\cap\class[_{2}]}$}. % \Coqed
-\end{theorem}
-
-%Either way, the theoretical results suggest that it is possible to always find a ,,most-general'', secure compiler, given two secure compilers, that robustly preserves the least-upper bound of the classes involved in their compilation process.
-
-\subsection{Applicability of the Compositionality Framework}\label{subsec:app}
-
-So far, this paper assumes a unified trace-model between source and target languages.
-Moreover secure compilers (in the sense of \Cref{def:rtp}) cannot change the behavior of compiled programs.
-It is worth noting that, as our case study in \Cref{sec:casestud:defs} demonstrates, several compilation passes do not change the behavior of compiled programs.
-% in such a way that the projections to the specification trace model become unequal to each other, 
-% i.e., source and target language of the compiler essentially have a different trace model. 
-However, there are instances of passes where this is not the case.
-For example, consider $\rtp{\cc{\src{L}}{\trg{L}}}{\tmssafe}$ and $\rtp{\cc{\trg{L}}{\obj{L}}}{\smssafe}$, where $\cc{\trg{L}}{\obj{L}}$ introduces code that allocates a table that stores bounds information of pointers, but fails to insert code that deallocates this table.
-By \Thmref{def:trace:tmsdef}, this violates $\tmssafe$.
-Therefore, no behavior resulting from programs produced by the $\cc{\trg{L}}{\obj{L}}$ compiler is secure with respect to $\tmssafe$, but only for $\smssafe$.
-$\cc{\trg{L}}{\obj{L}}$ modifies the behavior of compiled code and this section introduces support for these situations in the compositionality framework.
-
-To reason about these kinds of compilers, we rely on existing work~\cite{abate2021extacc} that lifts the presented secure compilation theory to compilers whose source and target trace models are different.
-% These compilers require a Galois connections for several real-world compiler scenarios, such as a compilation pass that exploits undefined behavior.
-% This section is concerned about a compilation setting with different source and target trace models. 
-These compilers require a way to say that two traces $\src{\trace}$ and $\trg{\trace}$ are related ($\sim$) and a function to translate a source-level property $\src{\pi}$ into a target-level property ($\tau_\sim\left( \src{\pi} \right)$~\cite{abate2021extacc}) using this trace relation.
-The function to map source-level properties to target-level properties can be lifted to translate source-level classes of properties ($\src{\class}$) to target-level classes ($\tilde{\tau}_\sim\left(\src{\class}\right)$).
-The section refers to these functions as existential images.
-
-With this, \bul{a cross-language relation $\sim$ is well-formed with respect to a class of source-level properties $\src{\class}$} iff \rul{the existential images preserve set inclusion}.
-\begin{definition}[Well-formedness of Trace Relations for a Class $\src{\class}$]\label{def:wfc:tracerel}
-  % $$
-
-  \noindent
-  \text{\bul{$\wfc{\sim}{\src{\class}}$}} \isdef \text{\rul{$\forall \src{\Pi}\subseteq\src{\class}, \tau_\sim(\src{\Pi})\subseteq\tilde{\tau}_\sim(\src{\class})$}}
-  % $$
-\end{definition}
-
-\Cref{def:rtpsim} shows how to adapt \Thmref{def:rtp} to account for different target-level trace-models~\cite{abate2021extacc}:
-
-\begin{definition}[Robust Compilation with $\sim$]\label{def:rtpsim}
-  $
-  \rtpsim{\cc{\src{L}}{\trg{L}}}{\src{\class}}{\sim} \isdef
-  \forall \src{\pi} \src{p}.
-    \text{if }\rsat{\src{\progvar}}{\src{\pi}},
-    \text{then }\rsat{\cc{\src{L}}{\trg{L}}\left(\src{p}\right)}{\tilde{\tau}_\sim\left(\src{\pi}\right)}.
-  $
-\end{definition}
-
-To state a similar theorem as \Thmref{thm:rtp} but for \Cref{def:rtpsim}, it is necessary to define composition of cross-language trace relations.
-Suppose there are traces $\src{\trace}$ and $\trg{\trace}$ as well as two trace relations $\sim_1$ and $\sim_2$. 
-The traces are related $\src{\trace}\sim_1\bullet\sim_2\trg{\trace}$ if there is another trace $\irl{\trace}$ such that $\src{\trace}\sim_1\irl{\trace}$ and $\irl{\trace}\sim_2\trg{\trace}$.
-
-For the adapted sequential composition theorem, let \bul{$\cc{\src{L}}{\trg{L}}$ robustly preserve the class $\src{\class[_1]}$ under $\sim_1$} and let \rul{$\cc{\trg{L}}{\obj{L}}$ robustly preserve the class $\tilde{\tau}_{\sim_1}\left(\src{\class[_2]}\right)$ under $\sim_2$}.
-Then, \iul{when the cross-language relation $\sim_1$ is well-formed with respect to class $\src{\class[_2]}$}, and \pul{relation $\sim_2$ is well-formed with respect to $\src{\class[_1]}$}, it follows that \oul{the composed compiler $\cc{\src{L}}{\trg{L}}\circ\cc{\trg{L}}{\obj{L}}$ robustly preserves the intersection of classes $\src{\class[_1]}\cap\src{\class[_2]}$ under $\sim_1\bullet\sim_2$}.
-% 
-%For space reasons, \Thmref{thm:urtp} and \Thmref{thm:lrtp} are not presented here, but can be found in the technical appendix ($\CoqSymbol$).
-% Note that it is not necessary to require $\sim_2$ to be well-formed with respect to $\src{\class[_1]}$, since the 
-% \MPin{why}
-\begin{theorem}[Sequential Composition of Secure Compilers]\label{thm:rtpsim}
-  If \bul{$\rtpsim{\cc{\src{L}}{\trg{L}}}{\src{\class[_{1}]}}{\sim_1}$}, \rul{$\rtpsim{\cc{\trg{L}}{\obj{L}}}{\tilde{\tau}_{\sim_1}\left(\src{\class[_{2}]}\right)}{\sim_2}$}, and \iul{$\wf{\sim_1}{\src{\class[_2]}}$}, and \pul{$\wf{\sim_2}{\src{\class[_1]}}$}, then \oul{$\rtpsim{\cc{\src{L}}{\trg{L}}\circ\cc{\trg{L}}{\obj{L}}}{\src{\class[_{1}]}\cap\src{\class[_{2}]}}{\sim_1\bullet\sim_2}$}.
-\Coqed
-\end{theorem}
-
-The reason we need assumptions about both relations is motivated by \Cref{lem:useless}.
-% 
-In fact, if \bul{a relation $\sim_1$ is well-formed for a property $\class$} but \rul{another relation $\sim_2$ is not well-formed for the property mapped with the first relation $\tilde{\tau}_{\sim_1}\class$} then \iul{their composition is empty}.
-% 
-An empty intersection means that there is no correct behaviour to preserve for the compiler, which nullifies its security relevance.
-\begin{lemma}[Composition is Useless]\label{lem:useless}
-  If \bul{$\wf{\sim_1}{\class}$}, \rul{$\neg\left(\wf{\sim_2}{\tilde{\tau}_{\sim_1}\class}\right)$}, then \iul{$\tilde{\tau}_{\sim_1\bullet\sim_2}\left(\class\right)=\emptyset$}. \Coqed
-\end{lemma}
-
-Let us now reconsider the example from the beginning of the section, suppose $\sim_{{\smssafe}}$ is a suitable trace relation such that $\trg{\trace}\sim_{{\smssafe}}\irl{\trace}$ establishes the fact that $\irl{\trace}$ contains an allocation of a table containing bounds information which is never deallocated.
-It is impossible that $\wf{\sim_{{\smssafe}}}{{\tmssafe}}$, since every behavior of compiled $\trg{L}$ programs never contains a deallocation of the bounds-information table.
-Still, supposing $\rtpsim{\cc{\src{L}}{\trg{L}}}{\tmssafe}{{\sim_{\tmssafe}}}$, $\rtpsim{\cc{\trg{L}}{\irl{L}}}{\smssafe}{{\sim_{\smssafe}}}$,
-and $\wf{{\sim_{\tmssafe}}}{{\smssafe}}$ allows to apply \Thmref{thm:rtpsim}.
-However, the resulting class $\tilde{\tau}_{{{\sim_{\tmssafe}}\bullet{\sim_{\smssafe}}}}\left(\tmssafe\cap\smssafe\right)=\emptyset$.
-
-\smallskip
-
-For sake of simplicity, the case-study presented in the following sections does not consider different trace-models: The traces of all programs of the presented languages are projected to a unified trace model.
-The case-study demonstrates that, the compositionality framework without the different trace-models is already applicable to a variety of security-relevant passes as well as existing optimisations.
-% may seem overly simplistic, it is useful and simpler in practice, if applicable.
 
@@
 The work relating robust preservation with universal composability~\cite{patrignani2022universal} is closest to what this paper presents.
-The authors demonstrate a similar compositionalty theorem to what is presented here (\Cref{sec:compcomp}) but use it in the context of protocols.
+The authors demonstrate a similar compositionality theorem to what is presented here (\Cref{sec:compcomp}) but use it in the context of protocols.
 They do not demonstrate the scalability of the approach.

### 2024-05-15 csf2/main.tex
add text

@@
 
-This paper uses a more general notion~\cite{abate2021extacc} that applies to compilers translating from one trace model into another.
+This paper uses a more general notion of robust preservation~\cite{abate2021extacc} that applies to compilers that use languages with different trace models. 
 To this end, considering a source trace $\src{\trace}$ and a target trace $\trg{\trace}$, there is a relation ($\sim$) describing the effect of a corresponding compiler. 
@@
 Using this relation, there are two projection functions: (1) the existential image $\tau_\sim\left(\src{\pi}\right)$ and (2) the universal image $\sigma_\sim\left(\trg{\pi}\right)$.
-These projections map source-level (resp. target-level) properties to target-level (resp. source-level) properties and can be lifted to work with classes, too, by taking the powerset.
-
+These projections map source-level (resp. target-level) properties to target-level (resp. source-level) properties. 
+The case study of this paper exclusively uses the universal image, since some considered properties, such as \gls*{ss}, are not definable in a higher-level language that, e.g., does not model speculation.
+
+Using these projection functions, the more general version of robust preservation is defined as follows~\cite{abate2021extacc}:
 For a \bul{compiler $\cc{\src{L}}{\trg{L}}$ to robustly preserve a class of source properties $\src{\class}$}, if for any \rul{property $\src{\pi}$ of that class $\src{\class}$ and programs $\src{p}$ written in $\src{L}$} where \iul{$\src{\progvar}$ robustly satisfies $\src{\pi}$}, then \oul{the compilation of $\src{\progvar}$, $\cc{\src{L}}{\trg{L}}\left(\src{p}\right)$, must robustly satisfy $\tau_\sim\left(\src{\pi}\right)$}.
 
-\begin{definition}[Robust Compilation with $\tau_\sim$]\label{def:rtp:tau}
+\begin{definition}[Robust Preservation with $\tau_\sim$]\label{def:rtp:tau}
   % $\;$\\
@@
 Because of this, this paper may write $\rtptau{\cc{\src{L}}{\trg{L}}}{\src{\pi}}{\sim}$, even though $\src{\pi}$ is a property and not a class.
+A similar construction can be applied to the projection functions.
 
@@
 
-\begin{definition}[Robust Compilation with $\sigma_\sim$]\label{def:rtp:sigma}
+\begin{definition}[Robust Preservation with $\sigma_\sim$]\label{def:rtp:sigma}
   % $\;$\\
@@
 
-\begin{definition}[Robust Compilation]\label{def:rtp}
+\begin{definition}[Robust Preservation]\label{def:rtp}
   % $\;$\\
@@
 \end{definition}
-
-% \MP{
-% 	check that we say classes are hp
-% }
-
-% \begin{exampleenv}[Types]
-%   Suppose $\src{L}$ is a statically-typed language similar to \texttt{C} and $\trg{T}$ is dynamically typed, where both share the same syntax up to dynamic type checks.
-%   Consider the following $\src{L}$ component and its compiled version below.
-%   \begin{center}
-%   $
-%   \begin{array}{l}
-%     \src{\lfunction{foo}{(char* x, int\ n)}{\lifz{valid\_ptr(x, n, sizeof(char))}{x[0]}{-1}}} \\
-%     \trg{\lfunction{foo}{(\phantom{char*\ }x, \phantom{int\ }n)}{\lifz{valid\_ptr(x, n, sizeof(char))}{x[0]}{-1}}}
-%   \end{array}
-%   $
-%   \end{center}
-%   While the compiler emits code that may look correct, the generated code does not check that the provided argument is of the right type.
-%   Even though the pointer $\trg{x}$ is checked for validity, the context $\trg{foo((int*)y, 1)}$ is able to provoke a read out of bounds.
-%   Suppose the component transferred control to the context and passed ownership of a $\trg{char}$ pointer $\trg{y}$ sized $\trg{1}$ cells, the context can now call the component again, casting this buffer to an $\trg{int*}$ prior to that call.
-%   The pointer is valid for one $\trg{char}$-sized memory cell, as expected, but the actual read operation now returns $\trg{sizeof(int)}$ many bytes instead of just $\trg{sizeof(char)}$ many.
-%   Thus, even if $\src{foo}$ may have been robust with respect to the \gls*{sms}, its compiled counterpart is not and therefore the compiler fails to attain \Cref{def:rtp}.
-% \end{exampleenv}
 
@@
 This {\em phase ordering problem} is well--known in literature and a practical solution is to simply perform a fixpoint iteration of the optimisation pipeline~\cite{click1995combining}.
-
-Suppose there are traces $\src{\trace}$ and $\trg{\trace}$ as well as two trace relations $\sim_1$ and $\sim_2$. 
+Hence, it is of interest to analyze the security of compilation passes and their interation within a compilation pipeline.
+
+For starters, suppose there are traces $\src{\trace}$ and $\trg{\trace}$ as well as two trace relations $\sim_1$ and $\sim_2$. 
 The traces are related $\src{\trace}\sim_1\bullet\sim_2\trg{\trace}$ if there is another trace $\irl{\trace}$ such that $\src{\trace}\sim_1\irl{\trace}$ and $\irl{\trace}\sim_2\trg{\trace}$.
@@
 
-Let \bul{$\cc{\src{L}}{\trg{L}}$ robustly preserve the class $\sigma_{\sim_2}\left(\trg{\class[_1]}\right)$ under $\sim_1$} and let \rul{$\cc{\trg{L}}{\obj{L}}$ robustly preserve the class $\trg{\class[_2]}$ under $\sim_2$}.
+The dual theorem also holds: let \bul{$\cc{\src{L}}{\trg{L}}$ robustly preserve the class $\sigma_{\sim_2}\left(\trg{\class[_1]}\right)$ under $\sim_1$} and let \rul{$\cc{\trg{L}}{\obj{L}}$ robustly preserve the class $\trg{\class[_2]}$ under $\sim_2$}.
 Then, \iul{when the cross-language relation $\sim_2$ is well-formed with respect to class $\trg{\class[_1]}$}, it follows that \oul{the composed compiler $\cc{\src{L}}{\trg{L}}\circ\cc{\trg{L}}{\obj{L}}$ robustly preserves the intersection of classes $\trg{\class[_1]}\cap\trg{\class[_2]}$ under $\sim_1\bullet\sim_2$}.
@@
 
-The degenerate case where the cross-language trace relation is equality has a simpler compositionality result:
+When the cross-language trace relation is an equality, \Cref{thm:rtpsim:tau,thm:rtpsim:sig} collapse:
 Given \bul{$\cc{\src{L}}{\trg{L}}$ robustly preserves $\class[_{1}]$} and \rul{$\cc{\trg{L}}{\obj{L}}$ robustly preserves $\class[_{2}]$}, it follows that \oul{their sequential composition $\cc{\src{L}}{\trg{L}}\circ\cc{\trg{L}}{\obj{L}}$ robustly preserves the intersection of classes $\class[_{1}]$ and $\class[_{2}]$}.
@@
 
-Since the composition of secure compilers is again a secure compiler, the theorem generalises to a whole chain of $n$ secure compilers.
+Since the composition of secure compilers is again a secure compiler, the theorems generalise to a whole chain of $n$ secure compilers.
 
@@
 \end{corollary}
+
+The technical appendix describes a more general version of \Cref{corr:swappable} that is applicable when trace models differ.
 
@@
 
-Examples of this are present in industry:
-Consider the Java Virtual Machine bytecode $\trg{JVM BC}$, which is a popular target for programming language designers due to its high performance and relevance in industry.
-Compilers for several programming languages have it as their target language, some popular instances are $\src{Java}$ and $\obj{Kotlin}$.
-Technically speaking, they both compile to class files and $\obj{Kotlin}$ objects are considered to be the same as $\src{Java}$ objects at that point.
-Both languages can be used at the same time in one project~\cite{androidstudio}.
-A compiler that accepts both $\src{Java}$ and $\obj{Kotlin}$ code translating to the same target language or intermediate representation performs a kind of {\em upper} composition.
+% Examples of this are present in industry:
+% Consider the Java Virtual Machine bytecode $\trg{JVM BC}$, which is a popular target for programming language designers due to its high performance and relevance in industry.
+% Compilers for several programming languages have it as their target language, some popular instances are $\src{Java}$ and $\obj{Kotlin}$.
+% Technically speaking, they both compile to class files and $\obj{Kotlin}$ objects are considered to be the same as $\src{Java}$ objects at that point.
+% Both languages can be used at the same time in one project~\cite{androidstudio}.
+% A compiler that accepts both $\src{Java}$ and $\obj{Kotlin}$ code translating to the same target language or intermediate representation performs a kind of {\em upper} composition.
 Now, the following theorem tells us what happens if these are secure:
@@
 
-Consider two compilers both accepting $\src{LLVM IR}$~\cite{lattner2004llvm} and one of them emits $\trg{x86\_64}$, while the other emits $\obj{ARMv8}$.
-It is intuitive that they are in some sense composed in the LLVM framework, but the decision of when to use one over the other is inherently {\em internal} to the formalisation effort of this kind of composition.
-For example, the user of this compiler provides an explicit flag that instructs to emit $\trg{x86\_64}$ or the framework itself detects the target platform via heuristics, such as supported instructions.
-
-The following theorem demonstrates what happens if the involved compilers are secure:
+% Consider two compilers both accepting $\src{LLVM IR}$~\cite{lattner2004llvm} and one of them emits $\trg{x86\_64}$, while the other emits $\obj{ARMv8}$.
+% It is intuitive that they are in some sense composed in the LLVM framework, but the decision of when to use one over the other is inherently {\em internal} to the formalisation effort of this kind of composition.
+% For example, the user of this compiler provides an explicit flag that instructs to emit $\trg{x86\_64}$ or the framework itself detects the target platform via heuristics, such as supported instructions.
+
+% The following theorem demonstrates what happens if the involved compilers are secure:
 Given \bul{$\cc{\src{L}}{\trg{L}}$ robustly preserves $\class[_{1}]$} and \rul{$\cc{\src{L}}{\obj{L}}$ robustly preserves $\class[_{2}]$}, it follows that \oul{their lower composition $\cc{\src{L}}{\trg{L}+\obj{L}}$ robustly preserves the intersection of classes $\class[_{1}]$ and $\class[_{2}]$}.
@@
 
-\subsection{Applicability of the Compositionality Framework}\label{subsec:app}
-
-So far, this paper assumes a unified trace-model between source and target languages.
-Moreover secure compilers (in the sense of \Cref{def:rtp}) cannot change the behavior of compiled programs.
-It is worth noting that, as our case study in \Cref{sec:casestud:defs} demonstrates, several compilation passes do not change the behavior of compiled programs.
-% in such a way that the projections to the specification trace model become unequal to each other, 
-% i.e., source and target language of the compiler essentially have a different trace model. 
-However, there are instances of passes where this is not the case.
-For example, consider $\rtp{\cc{\src{L}}{\trg{L}}}{\tmssafe}$ and $\rtp{\cc{\trg{L}}{\obj{L}}}{\smssafe}$, where $\cc{\trg{L}}{\obj{L}}$ introduces code that allocates a table that stores bounds information of pointers, but fails to insert code that deallocates this table.
-By \Thmref{def:trace:tmsdef}, this violates $\tmssafe$.
-Therefore, no behavior resulting from programs produced by the $\cc{\trg{L}}{\obj{L}}$ compiler is secure with respect to $\tmssafe$, but only for $\smssafe$.
-$\cc{\trg{L}}{\obj{L}}$ modifies the behavior of compiled code and this section introduces support for these situations in the compositionality framework.
-
-To reason about these kinds of compilers, we rely on existing work~\cite{abate2021extacc} that lifts the presented secure compilation theory to compilers whose source and target trace models are different.
-% These compilers require a Galois connections for several real-world compiler scenarios, such as a compilation pass that exploits undefined behavior.
-% This section is concerned about a compilation setting with different source and target trace models. 
-These compilers require a way to say that two traces $\src{\trace}$ and $\trg{\trace}$ are related ($\sim$) and a function to translate a source-level property $\src{\pi}$ into a target-level property ($\tau_\sim\left( \src{\pi} \right)$~\cite{abate2021extacc}) using this trace relation.
-The function to map source-level properties to target-level properties can be lifted to translate source-level classes of properties ($\src{\class}$) to target-level classes ($\tilde{\tau}_\sim\left(\src{\class}\right)$).
-The section refers to these functions as existential images.
-
-With this, \bul{a cross-language relation $\sim$ is well-formed with respect to a class of source-level properties $\src{\class}$} iff \rul{the existential images preserve set inclusion}.
-\begin{definition}[Well-formedness of Trace Relations for a Class $\src{\class}$]\label{def:wfc:tracerel}
-  % $$
-
-  \noindent
-  \text{\bul{$\wfc{\sim}{\src{\class}}$}} \isdef \text{\rul{$\forall \src{\Pi}\subseteq\src{\class}, \tau_\sim(\src{\Pi})\subseteq\tilde{\tau}_\sim(\src{\class})$}}
-  % $$
-\end{definition}
-
-\Cref{def:rtpsim} shows how to adapt \Thmref{def:rtp} to account for different target-level trace-models~\cite{abate2021extacc}:
-
-\begin{definition}[Robust Compilation with $\sim$]\label{def:rtpsim}
-  $
-  \rtpsim{\cc{\src{L}}{\trg{L}}}{\src{\class}}{\sim} \isdef
-  \forall \src{\pi} \src{p}.
-    \text{if }\rsat{\src{\progvar}}{\src{\pi}},
-    \text{then }\rsat{\cc{\src{L}}{\trg{L}}\left(\src{p}\right)}{\tilde{\tau}_\sim\left(\src{\pi}\right)}.
-  $
-\end{definition}
-
-To state a similar theorem as \Thmref{thm:rtp} but for \Cref{def:rtpsim}, it is necessary to define composition of cross-language trace relations.
-
-The reason we need assumptions about both relations is motivated by \Cref{lem:useless}.
-% 
-In fact, if \bul{a relation $\sim_1$ is well-formed for a property $\class$} but \rul{another relation $\sim_2$ is not well-formed for the property mapped with the first relation $\tilde{\tau}_{\sim_1}\class$} then \iul{their composition is empty}.
-% 
-An empty intersection means that there is no correct behaviour to preserve for the compiler, which nullifies its security relevance.
-\begin{lemma}[Composition is Useless]\label{lem:useless}
-  If \bul{$\wf{\sim_1}{\class}$}, \rul{$\neg\left(\wf{\sim_2}{\tilde{\tau}_{\sim_1}\class}\right)$}, then \iul{$\tilde{\tau}_{\sim_1\bullet\sim_2}\left(\class\right)=\emptyset$}. \Coqed
-\end{lemma}
-
-Let us now reconsider the example from the beginning of the section, suppose $\sim_{{\smssafe}}$ is a suitable trace relation such that $\trg{\trace}\sim_{{\smssafe}}\irl{\trace}$ establishes the fact that $\irl{\trace}$ contains an allocation of a table containing bounds information which is never deallocated.
-It is impossible that $\wf{\sim_{{\smssafe}}}{{\tmssafe}}$, since every behavior of compiled $\trg{L}$ programs never contains a deallocation of the bounds-information table.
-Still, supposing $\rtpsim{\cc{\src{L}}{\trg{L}}}{\tmssafe}{{\sim_{\tmssafe}}}$, $\rtpsim{\cc{\trg{L}}{\irl{L}}}{\smssafe}{{\sim_{\smssafe}}}$,
-and $\wf{{\sim_{\tmssafe}}}{{\smssafe}}$ allows to apply \Thmref{thm:rtpsim}.
-However, the resulting class $\tilde{\tau}_{{{\sim_{\tmssafe}}\bullet{\sim_{\smssafe}}}}\left(\tmssafe\cap\smssafe\right)=\emptyset$.
-
-\smallskip
-
-For sake of simplicity, the case-study presented in the following sections does not consider different trace-models: The traces of all programs of the presented languages are projected to a unified trace model.
-The case-study demonstrates that, the compositionality framework without the different trace-models is already applicable to a variety of security-relevant passes as well as existing optimisations.
-% may seem overly simplistic, it is useful and simpler in practice, if applicable.
-
 \section{Security Properties: Formalisation, Enforcement and Composition\pages{1.5}}\label{sec:compprop}
 
-This section introduces a trace model and uses it to define the key properties of interest for this paper: \gls*{tms}, \gls*{sms}, \gls*{ms}, and \gls*{scct} (\Cref{subsec:propdefs}).
-These properties are of practical importance (as mentioned in \Cref{sec:introduction}) and also of interest in the case study (\Cref{sec:casestud:defs,sec:casestud:rtp}) this paper presents later.
-
-\subsection{Specification Trace Model}\label{subsec:propdefs}
-
-% \begin{gather*}\small
-%   \begin{aligned}
-%   \mi{(Security\ Tag)}~\sigma\bnfdef&\ \lock \mid \unlock
-%   \qquad
-%   \mi{(Control\ Tag)}~\sandboxtag\bnfdef \ctx\mid\comp\\
-%   \mi{(Event)}~\event\bnfdef&\ \emptyevent \mid \lightning \mid \event[_{b}];\sandboxtag;\sigma \\
-%   \mi{(Pre\text{-}event)}~\event[_{b}]\bnfdef&\ \ev{Alloc\ \loc\ n} \mid \ev{Dealloc\ \loc} 
-%   \\
-%                                              & 
-%                                              \mid \ev{Use\ \loc\ n} \mid \ev{Branch\ n} \mid \ev{Binop\ n} \\
-%   \end{aligned}
-% \end{gather*}
-The specification trace model defines events as either the empty event ($\emptyevent$), a crash ($\lightning$), or as tuples consisting of a pre-event, a control-tag, and a security-tag.
-Since the purpose of the model is to define key security properties of interest, such as \gls*{ms} or a stricter variant of cryptographic constant time, security-tags indicate whether an event contains sensitive information ($\lock$) or not ($\unlock$), while control-tags state whether the context ($\ctx$) or the component ($\comp$) are responsible for emitting the event.
-The latter is necessary to be able to ignore actions done by a spurious context that, e.g., immediately deallocates a memory location twice, thus violating \gls*{tms}~\cite{nagarakatte2010cets}.
+This section introduces trace properties of interest for this paper: \gls*{tms}, \gls*{sms}, \gls*{ms}, \gls*{scct}, and \gls*{ss} (\Cref{subsec:propdefs}).
+These properties are of practical importance (as mentioned in \Cref{sec:introduction}) and also of interest in the case study (\Cref{sec:casestud:defs,sec:casestud:rtp}) presented later.
+
+\subsection{Basic Memory Safety Trace Model}\label{subsec:propdefs}\label{subsec:basic:memsafety:tracemodel}
+
+For simple memory safety composed of temporal and spatial memory safety, the trace model defines events as either the empty event ($\emptyevent$), a crash ($\lightning$), or as tuples consisting of a pre-event and a control-tag.
+Control-tags state whether the context ($\ctx$) or the component ($\comp$) are responsible for emitting the event, a necessary technical detail to know who is to blame for, e.g., an allocation.
+With these control-tags it is possible to ignore effects performed by a spurious context that, e.g., immediately deallocates a memory location twice, thus violating \gls*{tms}~\cite{nagarakatte2010cets}.
 %Even though this could be used for blame~\cite{patrignani2023blame}, this is not what this paper is concerned about.
-Lastly, pre-events describe the actual kind of event that happened.
-One such kind is the allocation event ($\ev{Alloc\ \loc\ n}$) that fires whenever a program claims $n$ cells of memory and stores them at address $\loc$.
-Dually, deallocation ($\ev{Dealloc\ \loc}$) announces that the object at location $\loc$ is freed.
-These two events alone are enough to provide a partial description of \gls*{tms} by requiring that, e.g., there is only one deallocation event that carries a location $\loc$.
-To be able to express \gls*{sms}, there is also an event to describe reads from and writes to memory ($\ev{Use\ \loc\ n}$).
-Finally, for cryptographic code, there is a general guideline that secrets must not be visible on a trace.
-Moreover, an instruction whose timing is data-dependent must not have a secret as an operand.
-Typical operations with data-dependent timing are branches and certain binary operations, such as division.%
-\footnote{
-	This is architecture-dependent, but division is an operation that serves as a classic example for a data-dependent timing instruction~\cite[p.~755]{arm-refman}.
-}
-Both operations are also modelled in the specification trace model ($\ev{Branch\ n}$ and $\ev{Binop\ n}$).
+Pre-events describe the actual kind of event that happened.
+For the basic memory-safety properties, these are three variants:
+First, the allocation event ($\ev{Alloc\ \loc\ n}$) that fires whenever a program claims $n$ cells of memory and stores them at address $\loc$, where addresses are assumed to be unique.
+Second, deallocation ($\ev{Dealloc\ \loc}$) announces that the object at location $\loc$ is freed.
+Third, an event to describe reads from and writes to memory ($\ev{Use\ \loc\ n}$).
+%Typical operations with data-dependent timing are branches and certain binary operations, such as division.%
+%\footnote{
+%	This is architecture-dependent, but division is an operation that serves as a classic example for a data-dependent timing instruction~\cite[p.~755]{arm-refman}.
+%}
+%Both operations are also modelled in the specification trace model ($\ev{Branch\ n}$ and $\ev{Binop\ n}$).
 
@@
 \gls*{tms} is a safety property that describes that an unallocated object must not be used in any way.
-Moreover, the property requires that all allocated objects must be deallocated at some point.
+%Moreover, the property requires that all allocated objects must be deallocated at some point.
 
@@
                               \ev{Use\ \loc\ n;\sandboxtag}&\le_{\trace}&\ev{Dealloc\ \loc;\sandboxtag} \\
-                              \text{if }\ev{Alloc\ \loc\ n;\sandboxtag}\text{ in }\trace&\text{then}&\ev{Dealloc\ \loc;\sandboxtag}\text{ in }\trace \\
+                            %  \text{if }\ev{Alloc\ \loc\ n;\sandboxtag}\text{ in }\trace&\text{then}&\ev{Dealloc\ \loc;\sandboxtag}\text{ in }\trace \\
                               \text{at most one }\ev{Dealloc\ \loc;\sandboxtag}&\text{in}&\trace \\
@@
 %%% our definition does not guarantee that (on purpose, we want to compose it with cct)
-Note that this definition says nothing about memory-safety issues introduced by side-channels.
 
@@
 \end{definition}
+
+Note that \Cref{def:trace:msdef} says nothing about memory-safety issues introduced by side-channels.
+
+\subsection{Memory Safety Trace Model}
+
+This section starts off by extending the basic memory safety trace model (\Cref{subsec:basic:memsafety:tracemodel}) with a {\em security tag} ($\securitytag$) that indicates whether the involved event contains sensitive information ($\lock$) or not ($\unlock$).
+For cryptographic code, there is a general guideline that secrets must not be visible on a trace, i.e., marked as $\unlock$.
+Moreover, an instruction whose timing is data-dependent must not have a secret as an operand.
+Typical operations with data-dependent timing are branches and certain binary operations, such as division. 
+\footnote{
+	This is architecture-dependent, but division is an operation that serves as a classic example for a data-dependent timing instruction~\cite[p.~755]{arm-refman}.
+}
+Both operations are also modelled in the trace model by extending the set of pre-events with branches on some value $n$ ($\ev{Branch\ n}$) and binary operations with result $n$ ($\ev{Binop\ n}$).
 
@@
 
-\gls*{scct} may seem overly strict, since it may seem that secrets must not occur on a trace (since $\securitytag$ is forced to be $\unlock$). 
+\gls*{scct} may seem overly strict, since it seems that secrets must not occur on a trace (since $\securitytag$ is forced to be $\unlock$). 
 However, programs that operate on secrets and have a ``data independent timing mode'' enabled do not leak these secrets, despite performing computation on them (see \Cref{ex:ditm}).
@@
 
-The combination of \gls*{ms} and \gls*{scct} is the intersection of these properties, \gls*{msscct}.
+The combination of \gls*{ms} and \gls*{scct} is, intuitively, the intersection of these properties, \gls*{msscct}.
+However, \gls*{ms} uses a simpler trace model than \gls*{scct}, so intersecting them would trivially yield the empty set. 
+To remedy this issue, let $\sim_{msscct}$ be a cross-language trace relation that relates the basic memory safety trace model with the memory safety trace model by ignoring security tags and replacing the new pre-events with the empty event:
+\[
+  \typerulenolabel{mscctrel:drop:tag}{}{\event[_b];\sandboxtag\sim_{msscct}\event[_b];\sandboxtag;\securitytag}
+\]
+\[
+  \typerulenolabel{mscctrel:drop:branch}{}{\emptyevent\sim_{msscct}\ev{Branch n};\sandboxtag;\securitytag}
+  \typerulenolabel{mscctrel:drop:binop}{}{\emptyevent\sim_{msscct}\ev{Binop n};\sandboxtag;\securitytag}
+\]
+
+It is now possible to define \gls*{msscct} using the universal image:
+
 %Since \gls*{ms} and \gls*{scct} are just sets of traces that, intuitively, contain all program behaviors that follow a security policy, the intersection of them contains all program behaviors that follow both security policies, i.e., it entails all program behaviours that are both \gls*{ms} and \gls*{scct}.
@@
 \begin{definition}[\glsfirst*{ms} and \gls*{scct}]\label{def:trace:msscctdef}
-  $
-  \msscctsafe:=\mssafe\cap\scctsafe
-  $
+  \[
+  \msscctsafe:=\sigma_{\sim_{msscct}}\left(\mssafe\right)\cap\scctsafe
+  \]
 \end{definition}
 
+\subsection{Speculation Trace Model}
+
+So far, the considered trace models do not take speculative execution into account. 
+For this, extend the memory safety trace model so that the security tags ($\securitytag$) carry additional information about the kind of private data leakage, i.e., the type of speculative leak.
+Moreover, add pre-events signaling the beginning of a speculative execution ($\ev{Spec}$) as well as a rollback event ($\ev{Rlb}$).
+
 \subsubsection{Speculative Safety}
 
-\MKin{fill in the blank based off of spectetor}
-
-%\section{Monitors and Trace Agreements\pages{1.5}}\label{subsec:monitors}
-% \subsection{Monitors\pages{1.5}}\label{subsec:monitors}
-
-% Monitors enforce safety properties by accepting or rejecting traces, i.e., if they reject a trace, the trace does not satisfy the property they check.
-% Since reasoning on monitors is easier than on just traces, this section presents a monitor for each of the previously shown safety properties (\Cref{subsec:propdefs}).
-% %\MPpost{
-% %	this is fine, but it'd be nice to provide more insights on why it's easier
-% %}
-% To lessen the burden when proving that a monitor accepts the trace of a program execution, each monitor uses a custom trace model that contains only the relevant information related to the property the monitor checks.
-% To go from specification traces $\trace$ to monitor-level traces $\abstrace$, each property $\pi$ has an associated event agreement relation $\traceagree[]{\pi}{\event}{\absevent}$.
-% Lifting the event agreement to traces works as expected, i.e., the empty traces are related, both are allowed to stutter with the empty event, and otherwise the events $\event$ and $\absevent$ must fulfill $\traceagree[]{\pi}{\event}{\absevent}$.
-% The trace agreement is the same for all properties $\pi$. 
-% With agreements, this section defines monitor satisfaction for traces and then it proves that monitor satisfaction implies property satisfaction.
-% To this end, monitor satisfaction is defined as follows.
-% \bul{A specification trace $\trace$ monitor-satisfies property $\pi$} iff there exists a (final) monitor state $\monitor$ and an abstract trace $\abstrace$ such that the \iul{specification trace $\trace$ agrees with abstract trace $\abstrace$} and \oul{the initial monitor\footnote{In this paper, for all monitors, the initial monitor state is denoted as $\emptyset$.} can step to the (final) monitor state $\monitor$ with abstract trace $\abstrace$}.
-
-% \begin{definition}[Monitor Satisfaction]\label{def:monsat}
-%   \bul{$\monsafe{\trace}{\pi}$} $\isdef$ $\exists \abstrace\ \monitor$, \iul{$\traceagree{\pi}{\trace}{\abstrace}$} and \oul{$\monitorcheck{\emptyset}{\monitor}{\abstrace}$}.
-% \end{definition}
-
-% \subsubsection{Monitor for \glsfirst*{tms}}\label{subsubsec:tms}
-% \begin{gather*}
-%   \begin{aligned}
-%     \mi{(Abstract\ Store)}~\tmsmonitor\bnfdef&\ \left\{\text{A}:L\times\sandboxtag,\text{F}:L\times\sandboxtag\right\} 
-%     \\%
-%     % \emptyset:=&\ \left\{\text{A}:\emptyset,\text{F}:\emptyset\right\}
-%     % \\
-%     \mi{(Abstract\ Events)}~\absevent\bnfdef&\ \absev{\emptyevent} \mid \absev{Alloc\ }\loc\ \sandboxtag \mid \absev{Dealloc\ }\loc\ \sandboxtag 
-%     % \\
-%                                             % & 
-%                                             \mid \absev{Use\ }\loc\ \sandboxtag \mid \absterm\\
-%   \end{aligned}
-% \end{gather*}
-% For \gls*{tms}, the state of the monitor is a record with two sets keeping track of allocated and deallocated locations, and the initial state $\emptyset$ is defined as $\left\{\text{A}:\emptyset,\text{F}:\emptyset\right\}$.
-% % 
-% Monitor events include {\em allocation}, {\em deallocation}, and {\em use} of a location $\loc$ as well as abnormal program termination.
-
-% \begin{center}\small
-%   \judgbox{\monitorcheck[]{\tmsmonitor}{\tmsmonitor[']}{\absevent}}{,,Monitor $\tmsmonitor$ steps to $\tmsmonitor[']$ given event $\absevent$.''}$\;$\\
-%   %
-% % \typerule{tms-uninteresting}{
-% % }{
-% %   \monitorcheck[]{\tmsmonitor}{\tmsmonitor}{\absev{\emptyevent}}
-% % }{tms-uninteresting}
-% % %
-% % \typerule{tms-abort}{
-% % }{
-% %   \monitorcheck[]{\tmsmonitor}{\tmsmonitor}{\absterm}
-% % }{tms-abort}
-%   % 
-%   \typerule{tms-use}{
-%     (\loc;\sandboxtag)\in\tmsmonitor.\text{A} 
-%     &
-%     % \\
-%     (\loc;\sandboxtag)\notin\tmsmonitor.\text{F}
-%   }{
-%     \monitorcheck[]{\tmsmonitor}{\tmsmonitor}{\absev{Use\ }\loc\ \sandboxtag}
-%   }{tms-use}
-%   %
-%   \typerule{tms-alloc}{
-%     (\loc;\sandboxtag)\notin\tmsmonitor.\text{A} &
-%     (\loc;\sandboxtag)\notin\tmsmonitor.\text{F} 
-%     % \\
-%     &
-%     \tmsmonitor[']=\left\{\text{A}: \tmsmonitor.\text{A}\cup\left\{(\loc;\sandboxtag)\right\},\text{F}: \tmsmonitor.\text{F}\right\}
-%   }{
-%     \monitorcheck[]{\tmsmonitor}{\tmsmonitor[']}{\absev{Alloc\ }\loc\ \sandboxtag}
-%   }{tms-alloc}
-%   %
-%   \typerule{tms-dealloc}{
-%     (\loc;\sandboxtag)\in\tmsmonitor.\text{A} 
-%     &
-%     (\loc;\sandboxtag)\notin\tmsmonitor.\text{F} 
-%     \\
-%     \tmsmonitor[']=\left\{\text{A}: \tmsmonitor.\text{A}\setminus\left\{(\loc;\sandboxtag)\right\} \text{F}: \tmsmonitor.\text{F}\cup\left\{(\loc;\sandboxtag)\right\}\right\}
-%   }{
-%     \monitorcheck[]{\tmsmonitor}{\tmsmonitor[']}{\absev{Dealloc\ }\loc\ \sandboxtag}
-%   }{tms-dealloc}
-%   %
-% \end{center}
-% % 
-% \Cref{tr:tms-use} simply requires that a location is (i) allocated and (ii) not freed.
-% \Cref{tr:tms-alloc,tr:tms-dealloc} both require a location to not be freed already and extend the monitor state accordingly.
-% This restriction effectively disallows reallocation to reassign the same location to an object.
-% However, the definition can easily be adapted by, e.g., attaching a natural number serving as a counter.
-% Contrary to other monitors in this paper, the multi-step relation of the \gls*{tms} monitor is non-standard:
-% \begin{center}
-%   \judgbox{\monitorcheck{\tmsmonitor}{\tmsmonitor[']}{\abstrace}}{,,Monitor $\tmsmonitor$ multi-steps to $\tmsmonitor[']$ given trace $\abstrace$.''}$\;$\\
-%   %
+\gls*{ss}~\cite{patrignani2021exorcising}, similar to \gls*{scct}, is an overapproximation to a variant of noninterference.
+The key difference is just the underlying trace model, while the definition of the property itself turns out to be syntactically equal to \gls*{scct}.
+
+\begin{definition}[\glsfirst*{ss}]\label{def:trace:ss}
+  \noindent\[
+  \sssafe:=\left\{\trace \left|\begin{array}{rcl}
+                                   \trace=\hole{\cdot} & \text{ or } & \trace=\event[_{b}];\sandboxtag;\unlock\cdot\trace['] \wedge \trace[']\in\sssafe
+                                 \end{array}\right.\right\}
+  \]
+\end{definition}
+
+\subsubsection{Speculation Memory Safety}
+
+Similar to \gls*{msscct}, there is a need to relate the different trace models with each other, so that the memory safety property without speculation can be lifted to speculation. 
+To this end, let $\sim_{spec}$ be a cross-language trace relation that relates the non-speculative memory safety trace model with the one that has speculation by simply ignoring any events related to speculation and ignoring the additional variant in the security-tag:
+\[
+  \typerulenolabel{specrel:drop:spec}{}{\emptyevent\sim_{spec}\ev{Spec};\sandboxtag;\securitytag}
+  \typerulenolabel{specrel:drop:rlb}{}{\emptyevent\sim_{spec}\ev{Rlb};\sandboxtag;\securitytag}
+\]
+\[
+  \typerulenolabel{specrel:drop:tag}{}{\event[_b];\sandboxtag;\lock\sim_{spec}\event[_b];\sandboxtag;\lock_{vX}}
+\]
+
+With this, define \gls*{specms} as follows:
+\begin{definition}[\glsfirst*{specms}]\label{def:trace:specmsdef}
+  \[
+  \specmssafe=\sigma_{\sim_{spec}}\left(\msscctsafe\right)\cap\sssafe
+  \]
+\end{definition}
+
 %   \typerule{tms-refl}{

### 2024-05-16 csf2/main.tex
add text

@@
 
+\subsection{$\Lspec$: A Memory-Unsafe Language with Speculation}\label{subsec:lspec}
+
+$\Lspec$ extends $\Lscct$ (\Cref{subsec:lscct}) with speculative dynamic semantics. 
+After branching or accessing memory, speculation can start and run subsequent code in a predetermined window until a rollback of operational state is performed. 
+Within such windows, data is leaked, marked as high $\ird{\lock}$ and with an annotation that indicates the respective speculative execution variant.
+In terms of language features, $\Lspec$ supports a barrier operation $\ird{\lbarrier}$ that blocks speculative execution. 
+The trace model is extended with $\ird{\ev{Spec}}$, $\ird{\ev{Rlb}}$, and $\ird{\ev{Barrier}}$ events that signal the respective operational action. 
+A barrier prevents any events whatsoever besides $\ird{\ev{Rlb}}$ when run in speculative execution mode and does nothing when run in normal mode.
+
 \begin{figure*}[!h]
@@
 
+\subsection{Robust Speculative Safety Preservation}\label{subsec:cs:ss}
+
+\vspace{-1em}
+\begin{center}\small
+  \begin{align*}
+    \ccspec(\obj{\lifz{\expr[_0]}{\expr[_1]}{\expr[_2]}}) & = \lifz[\ird]{\ccspec{\left(\obj{\expr[_0]}\right)}}{\\&\ird{\lbarrier;}\ccspec{\left(\obj{\expr[_1]}\right)}\\&}{\ird{\lbarrier;}\ccspec{\left(\ird{\expr[_2]}\right)}} \\
+  \end{align*}
+\end{center}
+% \vspace{-2em}
+%
+\begin{theorem}[$\ccspec$ is secure with respect to \gls*{ss}]\label{thm:ccspec:rtp:spec}
+  $\rtp{\ccspec}{\specmssafe}$ % \Coqed
+\end{theorem}
+
 \section{Related Work\pages{2}}\label{sec:relwork}

### 2024-05-19 csf2/main.tex
some bnfs

@@
 
-\bul{A cross-language relation $\sim$ is well-formed with respect to a class of source-level properties $\src{\class}$} iff \rul{the existential image preserves set membership}.
-\begin{definition}[Well-formedness of Trace Relations for a Class $\src{\class}$]\label{def:wfc:tau:tracerel}
-  % $$
-
-  \noindent
-  \text{\bul{$\wfctau{\sim}{\src{\class}}$}} \isdef \text{\rul{$\forall \src{\Pi}\in\src{\class}, \tau_\sim(\src{\Pi})\in\tilde{\tau}_\sim(\src{\class})$}}
-  % $$
-\end{definition}
-
 \bul{A cross-language relation $\sim$ is well-formed with respect to a class of target-level properties $\trg{\class}$} iff \rul{the universal image preserves set membership}.
@@
 
-Let \bul{$\cc{\src{L}}{\trg{L}}$ robustly preserve the class $\src{\class[_1]}$ under $\sim_1$} and let \rul{$\cc{\trg{L}}{\obj{L}}$ robustly preserve the class $\tilde{\tau}_{\sim_1}\left(\src{\class[_2]}\right)$ under $\sim_2$}.
-Then, \iul{when the cross-language relation $\sim_1$ is well-formed with respect to class $\src{\class[_2]}$}, it follows that \oul{the composed compiler $\cc{\src{L}}{\trg{L}}\circ\cc{\trg{L}}{\obj{L}}$ robustly preserves the intersection of classes $\src{\class[_1]}\cap\src{\class[_2]}$ under $\sim_1\bullet\sim_2$}.
-% 
-%For space reasons, \Thmref{thm:urtp} and \Thmref{thm:lrtp} are not presented here, but can be found in the technical appendix ($\CoqSymbol$).
-% Note that it is not necessary to require $\sim_2$ to be well-formed with respect to $\src{\class[_1]}$, since the 
-% \MPin{why}
-\begin{theorem}[Sequential Composition of Secure Compilers with respect to $\tau$]\label{thm:rtpsim:tau}
-  If \bul{$\rtptau{\cc{\src{L}}{\trg{L}}}{\src{\class[_{1}]}}{\sim_1}$}, \rul{$\rtptau{\cc{\trg{L}}{\obj{L}}}{\tilde{\tau}_{\sim_1}\left(\src{\class[_{2}]}\right)}{\sim_2}$}, and \iul{$\wfctau{\sim_1}{\src{\class[_2]}}$}, then \oul{$\rtptau{\cc{\src{L}}{\trg{L}}\circ\cc{\trg{L}}{\obj{L}}}{\src{\class[_{1}]}\cap\src{\class[_{2}]}}{\sim_1\bullet\sim_2}$}.
-\Coqed
-\end{theorem}
-
-The dual theorem also holds: let \bul{$\cc{\src{L}}{\trg{L}}$ robustly preserve the class $\sigma_{\sim_2}\left(\trg{\class[_1]}\right)$ under $\sim_1$} and let \rul{$\cc{\trg{L}}{\obj{L}}$ robustly preserve the class $\trg{\class[_2]}$ under $\sim_2$}.
+Let \bul{$\cc{\src{L}}{\trg{L}}$ robustly preserve the class $\sigma_{\sim_2}\left(\trg{\class[_1]}\right)$ under $\sim_1$} and let \rul{$\cc{\trg{L}}{\obj{L}}$ robustly preserve the class $\trg{\class[_2]}$ under $\sim_2$}.
 Then, \iul{when the cross-language relation $\sim_2$ is well-formed with respect to class $\trg{\class[_1]}$}, it follows that \oul{the composed compiler $\cc{\src{L}}{\trg{L}}\circ\cc{\trg{L}}{\obj{L}}$ robustly preserves the intersection of classes $\trg{\class[_1]}\cap\trg{\class[_2]}$ under $\sim_1\bullet\sim_2$}.
@@
 
-When the cross-language trace relation is an equality, \Cref{thm:rtpsim:tau,thm:rtpsim:sig} collapse:
+When the cross-language trace relation is an equality, \Cref{thm:rtpsim:sig} collapse:
 Given \bul{$\cc{\src{L}}{\trg{L}}$ robustly preserves $\class[_{1}]$} and \rul{$\cc{\trg{L}}{\obj{L}}$ robustly preserves $\class[_{2}]$}, it follows that \oul{their sequential composition $\cc{\src{L}}{\trg{L}}\circ\cc{\trg{L}}{\obj{L}}$ robustly preserves the intersection of classes $\class[_{1}]$ and $\class[_{2}]$}.
@@
 
-\begin{corollary}[Sequential Composition of Secure Compilers]\label{corr:swappable}
+\begin{corollary}[Swapping Secure Compiler Passes]\label{corr:swappable}
   If \bul{$\rtp{\cc[_{1}]{\trg{L}}{\trg{L}}}{\class[_{1}]}$ and $\rtp{\cc[_{2}]{\trg{L}}{\trg{L}}}{\class[_{2}]}$}, then \oul{$\rtp{\cc[_{1}]{\trg{L}}{\trg{L}}\circ\cc[_{2}]{\trg{L}}{\trg{L}}}{\class[_{1}]\cap\class[_{2}]}$ and $\rtp{\cc[_{2}]{\trg{L}}{\trg{L}}\circ\cc[_{1}]{\trg{L}}{\trg{L}}}{\class[_{2}]\cap\class[_{1}]}$}.
@@
 These properties are of practical importance (as mentioned in \Cref{sec:introduction}) and also of interest in the case study (\Cref{sec:casestud:defs,sec:casestud:rtp}) presented later. 
-
-\subsection{Basic Memory Safety Trace Model}\label{subsec:propdefs}\label{subsec:basic:memsafety:tracemodel}
+For each of these properties, the technical appendix accompanying the paper defines corresponding monitors, which refine a given property.
+
+\subsection{Memory Safety Trace Model}\label{subsec:propdefs}\label{subsec:basic:memsafety:tracemodel}
+
+\vspace{-1.5em}
+
+\[
+  \begin{array}{rcll}
+    \event[_{ms}] &:=& \ev{Alloc \loc\ n} \mid \ev{Dealloc \loc} \mid \ev{Use \loc\ n} \mid \emptyevent \mid \lightning \\ 
+  \end{array}
+\]
 
@@
   \tmssafe:=\left\{\trace \left| \begin{array}{rcl}
-                              \ev{Alloc\ \loc\ n;\sandboxtag}&\le_{\trace}&\ev{Dealloc\ \loc;\sandboxtag} \\
-                              \ev{Use\ \loc\ n;\sandboxtag}&\le_{\trace}&\ev{Dealloc\ \loc;\sandboxtag} \\
+                              \ev{Alloc\ \loc\ n}&\le_{\trace}&\ev{Dealloc\ \loc} \\
+                              \ev{Use\ \loc\ n}&\le_{\trace}&\ev{Dealloc\ \loc} \\
                             %  \text{if }\ev{Alloc\ \loc\ n;\sandboxtag}\text{ in }\trace&\text{then}&\ev{Dealloc\ \loc;\sandboxtag}\text{ in }\trace \\
-                              \text{at most one }\ev{Dealloc\ \loc;\sandboxtag}&\text{in}&\trace \\
-                              \text{at most one }\ev{Alloc\ \loc\ n;\sandboxtag}&\text{in}&\trace \\
+                              \text{at most one }\ev{Dealloc\ \loc}&\text{in}&\trace \\
+                              \text{at most one }\ev{Alloc\ \loc\ n}&\text{in}&\trace \\
                             \end{array}\right.\right\}
@@
   \smssafe:=\left\{\trace \left|\begin{array}{rcl}
-                                  \text{If }\ev{Alloc\ \loc\ n;\sandboxtag}\le_{\trace}\ev{Use\ \loc\ m;\sandboxtag}, \text{ then }m<n
+                                  \text{If }\ev{Alloc\ \loc\ n}\le_{\trace}\ev{Use\ \loc\ m}, \text{ then }m<n
                                 \end{array}\right.\right\}
@@
 
-\subsection{Memory Safety Trace Model}
+\subsection{Memory Safety Trace Model with Constant Timeness}
+
+\[
+  \begin{array}{rcll}
+    \event[^b_{ct}] &:=& \\
+    \event[_{ct}] &:=& \ev{Alloc \loc\ n} \mid \ev{Dealloc \loc} \mid \ev{Use \loc\ n} \mid \emptyevent \mid \lightning \\ 
+  \end{array}
+\]
 
@@
   \scctsafe:=\left\{\trace \left|\begin{array}{rcl}
-                                   \trace=\hole{\cdot} & \text{ or } & \trace=\event[_{b}];\sandboxtag;\unlock\cdot\trace['] \wedge \trace[']\in\scctsafe
+                                   \trace=\hole{\cdot} & \text{ or } & \trace=\event[_{b}];\unlock\cdot\trace['] \wedge \trace[']\in\scctsafe
                                  \end{array}\right.\right\}
@@
 \[
-  \typerulenolabel{mscctrel:drop:tag}{}{\event[_b];\sandboxtag\sim_{msscct}\event[_b];\sandboxtag;\securitytag}
+  \typerulenolabel{mscctrel:drop:tag}{}{\event[_b];\sim_{msscct}\event[_b];\securitytag}
 \]
 \[
-  \typerulenolabel{mscctrel:drop:branch}{}{\emptyevent\sim_{msscct}\ev{Branch n};\sandboxtag;\securitytag}
-  \typerulenolabel{mscctrel:drop:binop}{}{\emptyevent\sim_{msscct}\ev{Binop n};\sandboxtag;\securitytag}
+  \typerulenolabel{mscctrel:drop:branch}{}{\emptyevent\sim_{msscct}\ev{Branch n};\securitytag}
+  \typerulenolabel{mscctrel:drop:binop}{}{\emptyevent\sim_{msscct}\ev{Binop n};\securitytag}
 \]
@@
 
-\subsection{Speculation Trace Model}
+\subsection{Memory-Safe, Constant-Time Trace Model with Speculation}
 
@@
   \sssafe:=\left\{\trace \left|\begin{array}{rcl}
-                                   \trace=\hole{\cdot} & \text{ or } & \trace=\event[_{b}];\sandboxtag;\unlock\cdot\trace['] \wedge \trace[']\in\sssafe
+                                   \trace=\hole{\cdot} & \text{ or } & \trace=\event[_{b}];\unlock\cdot\trace['] \wedge \trace[']\in\sssafe
                                  \end{array}\right.\right\}
@@
 \[
-  \typerulenolabel{specrel:drop:spec}{}{\emptyevent\sim_{spec}\ev{Spec};\sandboxtag;\securitytag}
-  \typerulenolabel{specrel:drop:rlb}{}{\emptyevent\sim_{spec}\ev{Rlb};\sandboxtag;\securitytag}
+  \typerulenolabel{specrel:drop:spec}{}{\emptyevent\sim_{spec}\ev{Spec};\securitytag}
+  \typerulenolabel{specrel:drop:rlb}{}{\emptyevent\sim_{spec}\ev{Rlb};\securitytag}
 \]
 \[
-  \typerulenolabel{specrel:drop:tag}{}{\event[_b];\sandboxtag;\lock\sim_{spec}\event[_b];\sandboxtag;\lock_{vX}}
+  \typerulenolabel{specrel:drop:tag}{}{\event[_b];\lock\sim_{spec}\event[_b];\lock_{vX}}
 \]
@@
 \end{definition}
-
-%   \typerule{tms-refl}{
-%     \tmsmonitor.\text{allocated}=\emptyset
-%   }{
-%     \monitorcheck{\tmsmonitor}{\tmsmonitor}{\hole{\cdot}}
-%   }{tms-refl}
-%   %
-%   \typerule{tms-ign-trans}{
-%     \monitorcheck{\tmsmonitor}{\tmsmonitor[']}{\abstrace}
-%   }{
-%     \monitorcheck[]{\tmsmonitor}{\tmsmonitor[']}{\absev{\emptyevent}\cdot\abstrace}
-%   }{tms-ign-trans}
-%   %
-%   \typerule{tms-trans}{
-%     \monitorcheck[]{\tmsmonitor}{\tmsmonitor[']}{\absevent} &
-%     \monitorcheck{\tmsmonitor[']}{\tmsmonitor['']}{\abstrace}
-%   }{
-%     \monitorcheck[]{\tmsmonitor}{\tmsmonitor['']}{\absevent\cdot\abstrace}
-%   }{tms-trans}
-% \end{center}
-% \Cref{tr:tms-ign-trans,tr:tms-trans} are the same for all monitors, but \Cref{tr:tms-refl} has, in this case, an additional premise that all locations should be deallocated.
-% This rejects the behavior of programs that forget to free memory.
-
-% \begin{center}\small
-%   \judgbox{\tmstraceagree[]{\event}{\absevent}}{,,Abstract event $\absevent$ is equivalent to $\event$ w.r.t. \gls*{tms}.''}$\;$\\
-%   %
-%   \typerule{tms-alloc-authentic}{
-%   }{
-%     \tmstraceagree[]{\ev{Alloc}\ \loc\ n;\sandboxtag;\securitytag}{\absev{Alloc\ }\loc\ \sandboxtag}
-%   }{tms-alloc-authentic}
-%   %
-% % \typerule{tms-dealloc-authentic}{
-% % }{
-% %   \tmstraceagree[]{\ev{Dealloc}\ \loc;\sandboxtag;\securitytag}{\absev{Dealloc\ }\loc}
-% % }{tms-dealloc-authentic}
-%   %
-% % \typerule{tms-use-authentic}{
-% % }{
-% %   \tmstraceagree[]{\ev{Use}\ \loc\ n;\sandboxtag;\securitytag}{\absev{Use\ }\loc}
-% % }{tms-use-authentic}
-%   %
-%   \typerule{tms-branch-authentic}{
-%   }{
-%     \tmstraceagree[]{\ev{Branch\ }n}{\absev{\emptyevent}}
-%   }{tms-branch-authentic}
-%   %
-% % \typerule{tms-binop-authentic}{
-% % }{
-% %   \tmstraceagree[]{\ev{Binop\ }n}{\absev{\emptyevent}}
-% % }{tms-binop-authentic}
-% % %
-% % \typerule{tms-empty-authentic}{
-% % }{
-% %   \tmstraceagree[]{\emptyevent}{\absev{\emptyevent}}
-% % }{tms-empty-authentic}
-% % %
-%   % \typerule{tms-abort-authentic}{
-%   % }{
-%   %   \tmstraceagree[]{\lightning}{\absterm}
-%   % }{tms-abort-authentic}
-% \end{center}
-
-% The trace agreement maps actions to their respective (e.g., \Cref{tr:tms-alloc-authentic}), so we elide most cases.
-% % 
-% Since branches are not relevant for \gls*{tms}, they are mapped to $\absev{\emptyevent}$ (\Cref{tr:tms-branch-authentic}).
-% % \MP{
-% % 	we should say something .. ?
-% % }
-
-% \begin{lemma}[Monitor Satisfaction is $\tmssafe$]\label{lem:mon:tmssafe}
-%   $\;\;$ If $\montmssafe{\trace}$, then $\trace\in\tmssafe$. %\Coqed
-% \end{lemma}
-
-% \begin{exampleenv}[A program not satisfying \gls*{tms}]\label{ex:strncpy:cpp11}
-%   Consider the following \texttt{C++11} library that calls \texttt{strncpy} (\Cref{ex:strncpy}) and prints the result to the standard output stream.
-
-%   \begin{lstlisting}[language=c++,basicstyle=\small\ttfamily,mathescape,commentstyle=\color{gray}]
-% int hello() {
-%   char* greet = new char[12] { "Hello ESOP!" };
-%   char* to = new char[12];
-%   strncpy(11, to, greetings);
-%   delete to;
-%   printf("%cSOP\n", to[6]);
-% }
-%   \end{lstlisting}
-%     % // greet address ${\color{gray}\loc_{x}}$
-%   % // to address ${\color{gray}\loc_{y}}$
-% % 
-%   Up to the body of \texttt{printf}, the program execution yields the specification trace: 
-%     % $\ev{Alloc}\ \loc[_{x}]\ 12;\comp;\unlock\cdot\ev{Alloc}\ \loc[_{y}]\ 12;\comp;\unlock\cdot\ev{Use}\ \loc_{x}\ 0;\comp;\unlock\cdot\ev{Use}\ \loc_{x}\ 0;\comp;\unlock\cdot\ev{Use}\ \loc_{y}\ 0;\comp;\unlock\cdot\ev{Use}\ \loc_{x}\ 1;\comp;\unlock\cdot\ldots\cdot\ev{Use}\ \loc_{x}\ 12;\comp;\unlock\cdot\ev{Dealloc}\ \loc[_{y}];\comp;\unlock\cdot\ev{Use}\ \loc_{y}\ 6;\comp;\unlock$
-
-%     \vspace{-1em}
-%     {\small
-%     \begin{align*}
-%       &
-%       \ev{Alloc}\ \loc[_{x}]\ 12;\comp;\unlock
-%       \cdot
-%       \ev{Alloc}\ \loc[_{y}]\ 12;\comp;\unlock
-%       \cdot
-%       \ev{Use}\ \loc_{x}\ 0;\comp;\unlock
-%       \\
-%       &
-%       \cdot
-%       \ev{Use}\ \loc_{x}\ 0;\comp;\unlock
-%       \cdot
-%       \ev{Use}\ \loc_{y}\ 0;\comp;\unlock
-%       \cdot
-%       \ev{Use}\ \loc_{x}\ 1;\comp;\unlock
-%       \cdot
-%       \ldots
-%       \\
-%       &
-%       \cdot
-%       \ev{Use}\ \loc_{x}\ 12;\comp;\unlock
-%       \cdot
-%       \ev{Dealloc}\ \loc[_{y}];\comp;\unlock
-%       \cdot
-%       \ev{Use}\ \loc_{y}\ 6;\comp;\unlock
-%     \end{align*}
-%     }
-% Relating this trace to abstract monitor events yields trace:
-%   
-%   \vspace{-1em}
-%   {\small
-%   \begin{align*}
-%     % \abstrace=
-%     &
-%     \absev{Alloc}\ \loc[_{x}]\ \comp
-%     \cdot
-%     \absev{Alloc}\ \loc[_{y}]\ \comp
-%     \cdot
-%     \absev{Use}\ \loc[_{x}]\ \comp
-%     % \\
-%     % &
-%     \cdot
-%     \absev{Use}\ \loc[_{x}]\ \comp
-%     \cdot
-%     \absev{Use}\ \loc[_{y}]\ \comp
-%     \cdot
-%     \ldots
-%     \\
-%     &
-%     \cdot
-%     \absev{Use}\ \loc[_{x}]\ \comp
-%     \cdot
-%     \absev{Dealloc}\ \loc[_{y}]\ \comp
-%     \cdot
-%     \absev{Use}\ \loc[_{y}]\ \comp
-%    \end{align*} 
-%     }
-%   Remembering the definition of \texttt{strncpy} (\Cref{ex:strncpy}), observe that it does not deallocate its arguments.
-%   Even though the trace contains an out-of-bounds access right before returning from \texttt{strncpy}, this is no concern for \gls*{tms}, since location $\loc[_{x}]$ is still allocated.
-%   Having returned from \texttt{strncpy}, the \texttt{greet} function continues and deallocates $\loc[_{y}]$ whose subsequent use in \texttt{printf} is a use-after-free bug.
-
-%   The fix would be to \lstinline[language=c++,basicstyle=\small\ttfamily]|delete greetings| instead of \texttt{to} and add a \lstinline[language=c++,basicstyle=\small\ttfamily]|delete to| after the \texttt{printf} call, which leads to trace:
-%   
-% \vspace{-1em}
-%   {\small
-%   \begin{align*}
-%     \abstrace[']=
-%     &
-%     \absev{Alloc}\ \loc[_{x}]\ \comp
-%     \cdot
-%     \absev{Alloc}\ \loc[_{y}]\ \comp
-%     \cdot
-%     \absev{Use}\ \loc[_{x}]\ \comp
-%     % \\
-%     % &
-%     \cdot
-%     \absev{Use}\ \loc[_{x}]\ \comp
-%     \cdot
-%     \absev{Use}\ \loc[_{y}]\ \comp
-%     \cdot
-%     \ldots
-%     \cdot
-%     \\
-%     &
-%     \absev{Use}\ \loc[_{x}]\ \comp
-%     \cdot
-%     \absev{Dealloc}\ \loc[_{x}]\ \comp
-%     \cdot
-%     \absev{Use}\ \loc[_{y}]\ \comp	
-%   \end{align*}
-%   }
-%   It follows that $\montmssafe{\abstrace[']}$ and from \Thmref{lem:mon:tmssafe}, it follows that the program satisfies \Thmref{def:trace:tmsdef}, even though the program still violates \gls*{sms}.
-% \end{exampleenv}
-
-% % \MPpost{
-% % 	i think the trace description of traces is more uindertandable,
-% % 	teh monitor one is easier to work with
-% % 	this is the justification for having both that we'll have to make in the future
-% % }
-
-% \subsubsection{Monitor for \glsfirst*{sms}}\label{subsubsec:sms}
-% \begin{gather*}
-%   \begin{aligned}
-%     \mi{(Abstract\ Store)}~\smsmonitor:=&\ L\times\sandboxtag\times\mb{N}\\
-%     \mi{(Abstract\ Events)}~\absevent\bnfdef&\ \absev{\varepsilon} \mid \absev{Alloc\ }\loc\ \sandboxtag\ n \mid \absev{Use\ }\loc\ \sandboxtag\ n \\
-%   \end{aligned}
-% \end{gather*}
-% \begin{center}
-%   \judgbox{\monitorcheck[]{\smsmonitor}{\smsmonitor[']}{\absevent}}{,,Monitor $\smsmonitor$ steps to $\smsmonitor[']$ given event $\absevent$.''}$\;$\\
-%   %
-% % \typerule{sms-uninteresting}{
-% % }{
-% %   \monitorcheck[]{\smsmonitor}{\smsmonitor}{\absev{\emptyevent}}
-% % }{sms-uninteresting}
-% % %
-% % \typerule{sms-abort}{
-% % }{
-% %   \monitorcheck[]{\smsmonitor}{\emptyset}{\absterm}
-% % }{sms-abort}
-%   %
-%   \typerule{sms-use}{
-%     (\loc;\sandboxtag;m)\in\smsmonitor &
-%     n<m
-%   }{
-%     \monitorcheck[]{\smsmonitor}{\smsmonitor}{\absev{Use\ }\loc\ \sandboxtag\ n}
-%   }{sms-use}
-%   %
-%   \typerule{sms-alloc}{
-%     (\loc;\sandboxtag;m)\notin\smsmonitor %\\
-%     % unfortunately, this does not render sms-use and sms-alloc side by side
-%     % my only other idea is \scalebox...ugh
-%     %\smsmonitor[']=\smsmonitor\cup\left\{(\loc;\sandboxtag;n)\right\}
-%   }{
-%     \monitorcheck[]{\smsmonitor}{\smsmonitor\cup\left\{(\loc;\sandboxtag;n)\right\}}{\absev{Alloc\ }\loc\ \sandboxtag\ n}
-%   }{sms-alloc}
-% \end{center}
-% The state of the monitor for \gls*{sms} is a set containing tuples of locations, control-tags, and the allocation size.
-% In comparison to the trace model of the \gls*{tms} monitor, the trace model here is extended by sizing and positional information.
-% \Cref{tr:sms-use} performs a bounds check and \Cref{tr:sms-alloc} adds bounds information to the state of the monitor.
-% The trace agreement is entirely straightforward and similar to the one for \gls*{tms}.
-% % \begin{center}
-% %   \judgbox{\smstraceagree[]{\event}{\absevent}}{,,Abstract event $\absevent$ is equivalent to $\event$ with respect to \gls*{sms}.''}$\;$\\
-% %   %
-% %   \typerule{sms-alloc-authentic}{
-% %   }{
-% %     \smstraceagree[]{\ev{Alloc}\ \loc\ n;\sandboxtag;\securitytag}{\absev{Alloc\ }\loc\ n}
-% %   }{sms-alloc-authentic}
-% %   %
-% %   \typerule{sms-dealloc-authentic}{
-% %   }{
-% %     \smstraceagree[]{\ev{Dealloc}\ \loc;\sandboxtag;\securitytag}{\absev{Dealloc\ }\loc}
-% %   }{sms-dealloc-authentic}
-% %   %
-% %   \typerule{sms-use-authentic}{
-% %   }{
-% %     \smstraceagree[]{\ev{Use}\ \loc\ n;\sandboxtag;\securitytag}{\absev{Use\ }\loc\ n}
-% %   }{sms-use-authentic}
-% %   %
-% %   \typerule{sms-branch-authentic}{
-% %   }{
-% %     \smstraceagree[]{\ev{Branch\ }n}{\absev{\emptyevent}}
-% %   }{sms-branch-authentic}
-% %   %
-% %   \typerule{sms-binop-authentic}{
-% %   }{
-% %     \smstraceagree[]{\ev{Binop\ }n}{\absev{\emptyevent}}
-% %   }{sms-binop-authentic}
-% %   %
-% %   \typerule{sms-empty-authentic}{
-% %   }{
-% %     \smstraceagree[]{\emptyevent}{\absev{\emptyevent}}
-% %   }{sms-empty-authentic}
-% %   %
-% %   \typerule{sms-abort-authentic}{
-% %   }{
-% %     \smstraceagree[]{\lightning}{\absterm}
-% %   }{sms-abort-authentic}
-% % \end{center}
-
-% \begin{lemma}[Monitor Satisfaction is $\smssafe$]\label{lem:mon:smsafe}
-%   $\;\;$
-%   If $\monsmssafe{\trace}$, then $\trace\in\smssafe$. %\Coqed
-% \end{lemma}
-
-% \begin{exampleenv}[Normal invocation of \texttt{strncpy}]\label{ex:strncpy:sms}
-%   Consider the insecure \texttt{strncpy} function from \Cref{ex:strncpy} with a context \texttt{strncpy(2, x, y)}, where \texttt{x} and \texttt{y} are pointers to valid regions of memory with allocated space for exactly two cells and do not contain the null-terminating character \texttt{'\textbackslash 0'}.
-%   For the sake of this example, the pointers have been allocated by the component and passed to the context.
-%   The loop of \texttt{strncpy} will copy exactly two cells and then check the loop condition for the last time.
-%   At that stage, variable \texttt{i} is equal to \texttt{2} and, unfortunately, the order of checks is such that first the cell \texttt{x[i]} is read prior to bounds checking \texttt{i < n}, so there is an out-of-bounds memory access right before exiting the function.
-%   This is also visible on the trace:
-%     
-%     {\small
-%     \begin{align*}
-%     &
-%       \ldots
-%       \cdot
-%       \ev{Alloc}\ \loc_{x}\ 2;\comp;\unlock
-%       \cdot
-%       \ldots
-%       \cdot
-%       \ev{Alloc}\ \loc_{y}\ 2;\comp;\unlock
-%       \cdot
-%       \ldots
-%       \cdot
-%     % \\
-%     % &
-%       \ev{Use}\ \loc_{x}\ 0;\comp;\unlock
-%       \cdot
-%       \\
-%     &
-%       \ev{Use}\ \loc_{y}\ 0;\comp;\unlock
-%       \cdot
-%       \ev{Use}\ \loc_{x}\ 1;\comp;\unlock
-%     % \\
-%     % &
-%       \cdot
-%       \ev{Use}\ \loc_{y}\ 1;\comp;\unlock
-%       \cdot
-%       \ev{Use}\ \loc_{x}\ 2;\comp;\unlock
-%       \cdot
-%       \ldots	
-%     \end{align*}
-%     }
-
-% \noindent
-% where $\loc_{x}$ and $\loc_{y}$ are the memory addresses associated to \texttt{x} and \texttt{y}, respectively.
-%   Omitting the events for all ,,$\ldots$'' for sake of brevity, the abstract monitor trace of this is:
-%     % 
-%     {\small
-%     \begin{align*}
-%       &
-%       \absev{Alloc}\ \loc_{x}\ \comp\ 2
-%       \cdot
-%       \absev{Alloc}\ \loc_{y}\ \comp\ 2
-%       \cdot
-%       \absev{Use}\ \loc_{x}\ \comp\ 0
-%       \\
-%       &
-%       \cdot
-%       \absev{Use}\ \loc_{y}\ \comp\ 0
-%       \cdot
-%       \absev{Use}\ \loc_{x}\ \comp\ 1
-%       \cdot
-%       \absev{Use}\ \loc_{y}\ \comp\ 1
-%       \cdot
-%       \absev{Use}\ \loc_{x}\ \comp\ 2
-%     \end{align*}
-%     }
-
-%   After the allocation events, the state of the monitor is $\left\{(\loc_{x};\comp;2),(\loc_{y};\comp;2)\right\}$.
-%   All uses up to the last are accepted by the monitor, but the last event does not satisfy the premise $2<2$ in \Cref{tr:sms-use}.
-%   Therefore, the whole program (\texttt{strncpy} linked with this kind of context) is not \gls*{sms}.
-% \end{exampleenv}
-
-% \subsubsection{Combining \glsfirst*{tms} and \glsfirst*{sms} Monitors into \glsfirst*{ms}}\label{subsubsec:ms}
-
-% \begin{center}
-%   \judgbox{\monitorcheck[]{\msmonitor}{\msmonitor[']}{\absevent}}{,,Monitor $\msmonitor$ steps to $\msmonitor[']$ given event $\absevent$.''}$\;$\\
-%   \typerule{ms-step}{
-%     \monitorcheck[]{\tmsmonitor}{\tmsmonitor[']}{\absevent_{\tmssafe}} &
-%     \monitorcheck[]{\smsmonitor}{\smsmonitor[']}{\absevent_{\smssafe}} &
-%   }{
-%     \monitorcheck[]{(\tmsmonitor,\smsmonitor)}{(\tmsmonitor['],\smsmonitor['])}{(\absevent_{\tmssafe},\absevent_{\smssafe})}
-%   }{ms-step}
-% \end{center}
-% The combined monitor runs the one for \gls*{tms} and the one for \gls*{sms} in a lockstep.
-The trace agreement similarly just relates a specification event with an abstract \gls*{tms}-event $\abstrace[_{\tmssafe}]$ and with an abstract \gls*{sms}-event $\abstrace[_{\smssafe}]$.
-% % \begin{center}
-% %   \judgbox{\mstraceagree{\event}{\absevent}}{,,Abstract event $\absevent$ is equivalent to $\event$ with respect to \gls*{ms}.''}$\;$\\
-% %   %
-% %   \typerule{ms-authentic}{
-% %     \tmstraceagree[]{\event}{\absevent_{\tmssafe}} &
-% %     \smstraceagree[]{\event}{\absevent_{\smssafe}}
-% %   }{
-% %     \mstraceagree[]{\event}{(\absevent_{\tmssafe},\absevent_{\smssafe})}
-% %   }{ms-authentic}
-% % \end{center}
-
-% \begin{lemma}[Monitor Satisfaction is $\mssafe$]\label{lem:mon:msafe}
-%   If $\monmssafe{\trace}$, then $\trace\in\mssafe$.\Coqed
-% \end{lemma}
-
-% Any technique to combine existing monitors $T_A$, $T_B$ (that check for properties $A$ and $B$) into one $T_{AB}$ (that checks for both $A$ and $B$) is sufficient for the sake of the analysis of this paper.
-
-% \subsubsection{Monitor for \glsfirst*{scct}}
-% \begin{gather*}
-% % \small
-%   \begin{aligned}
-%     \mi{(Abstract\ Store)}~\scctmonitor:=&\ \emptyset 
-%     % &
-%     &
-%     \qquad\mi{(Abstract\ Events)}~\absevent:=&\ \absev{\varepsilon} \mid \absterm \mid \absev{Leak} 
-%     % \\
-%   \end{aligned}
-% \end{gather*}
-% \begin{center}
-%   \judgbox{\monitorcheck[]{\scctmonitor}{\scctmonitor[']}{\absevent}}{,,Monitor $\scctmonitor$ steps to $\scctmonitor[']$\\given event $\absevent$.''}$\;$\\
-%   %
-%   \typerule{scct-none}{
-%   }{
-%     \monitorcheck[]{\scctmonitor}{\scctmonitor}{\absev{\emptyevent}}
-%   }{scct-none}
-%   %
-%   \typerule{scct-abort}{
-%   }{
-%     \monitorcheck[]{\scctmonitor}{\scctmonitor}{\absterm}
-%   }{scct-abort}
-% \end{center}
-% The monitor state for the \gls*{scct} monitor is completely empty, since it does not need to keep track of information.
-% % As soon as 
-% If
-% any event is hit, the execution gets stuck, since any event is considered confidental from the perspective of this monitor.
-% \begin{center}
-%   \judgbox{\sccttraceagree[]{\event}{\absevent}}{,,Abstract event $\absevent$ is equivalent to $\event$ w.r.t. \gls*{cct}.''}$\;$\\
-%   %
-%   \typerule{scct-low-authentic}{
-%   }{
-%     \sccttraceagree[]{\event[_b];\sandboxtag;\unlock}{\absev{\emptyevent}}
-%   }{scct-low-authentic}
-%   %
-%   \typerule{scct-high-authentic}{
-%   }{
-%     \sccttraceagree[]{\event[_b];\sandboxtag;\lock}{\absev{Leak}}
-%   }{scct-high-authentic}
-%   %
-%   % \typerule{scct-empty-authentic}{
-%   % }{
-%   %   \sccttraceagree[]{\emptyevent}{\absev{\emptyevent}}
-%   % }{scct-empty-authentic}
-%   % %
-%   % \typerule{scct-abort-authentic}{
-%   % }{
-%   %   \sccttraceagree[]{\lightning}{\absterm}
-%   % }{scct-abort-authentic}
-% \end{center}
-% Accordingly, the event agreement simply disregards all events that involved public data ($\unlock$) while mapping any other event that does involve private data ($\lock$) to the abstract $\absev{Leak}$ event.
-
-% \begin{lemma}[Monitor Satisfaction is $\scctsafe$]\label{lem:mon:scctsafe}
-%   If $\monscctsafe{\trace}$, then $\trace\in\scctsafe$. %\Coqed
-% \end{lemma}
-
-% \begin{exampleenv}[Data-independent timing mode]\label{ex:ditm}
-%   Consider the function call \texttt{strncpy(1,x,y)} to the \texttt{strncpy} function (\Cref{ex:strncpy}) with low ($\unlock$) security for \texttt{x} and high security ($\lock$) for \texttt{y}.
-%   The trace for the copying part of \texttt{strncpy} is:
-
-% \vspace{-1em}
-%   {\small
-%   \begin{align*}
-%     \ev{Use}\ \loc_{x}\ 0;\comp;\unlock\cdot\ev{Use}\ \loc_{y}\ 0;\comp;\lock	
-%   \end{align*}
-%   }
-%   In terms of the abstract monitor trace, this is just $\absev{Leak}$.
-%   Running this on the monitor would result in getting stuck, since there is no matching rule to step in the presence of $\absev{Leak}$ event. % pun intended
-%   By means of additional features to ensure cryptographic constant time even in the presence of memory reads and loads, such as a flag to enable a data independent timing mode, which is present in both Arm~\cite[p.~543]{arm-refman} and Intel~\cite[p.~80]{intel-refman} processors, the original trace now does $\ev{Use}\ \loc_{y}\ 0;\comp;\unlock$ instead of $\ev{Use}\ \loc_{y}\ 0;\comp;\lock$.
-%   Because of this, the whole trace of the component is $\absev{\emptyevent}$ and the monitor can step without getting stuck.
-% \end{exampleenv}
-
-% \subsubsection{Combining \gls*{ms} and \gls*{scct} Monitors to obtain \gls*{msscct}}
-
-% The combination of monitors for \gls*{ms} and \gls*{scct} yields one for \gls*{msscct}.
-% The construction is entirely similar to the one for \gls*{ms} (\Cref{subsubsec:ms}).
-
-% \begin{lemma}[Monitor Satisfaction is $\msscctsafe$]\label{lem:mon:msscctsafe}
-%   % $\;$\\\vspace{-1em}
-%   % \begin{nscenter}
-%   % \noindent
-%   If $\monmsscctsafe{\trace}$, then $\trace\in\msscctsafe$. %\Coqed
-%   % \end{nscenter}
-% \end{lemma}
-
-% \section{Composing Secure Compilers\pages{1}}\label{sec:compcomp}
-
-% This section presents the key meta-theoretic results of this paper concerning secure compiler compositions (\Cref{sec:sequential,sec:other-compos}).
-% % Note that the presented framework works with arbitrary classes.
-% % \MP{check, we ened more than this}
-% % and builds on prior work~\cite{kruse2022csc}\MK{mention this at all?
-may reveal identity even though its double-blind}.
-% MP: not needed at all, ever, workshop work does not need to be cited in the main work
-% Note that since it deals with the preservation of \emph{classes}, which include both trace properties as well as hyperproperties, the presented framework is general: It describes robust compilers preserving ordinary trace properties, hyperproperties or even relational hyperproperties.
-% % \MP{pull this out to earlier}
 

### 2024-05-20 csf2/main.tex
implement some todos

@@
 \usepackage[switch]{lineno}
+\usepackage{halloweenmath}
 \usepackage{fontawesome5}
@@
 \[
-  \begin{array}{rcll}
-    \event[^b_{ct}] &:=& \\
-    \event[_{ct}] &:=& \ev{Alloc \loc\ n} \mid \ev{Dealloc \loc} \mid \ev{Use \loc\ n} \mid \emptyevent \mid \lightning \\ 
+  \begin{array}{rrcl}
+    (\text{Pre-Event}) & \event[^b_{ct}] &:=& \ev{Alloc \loc\ n} \mid \ev{Dealloc \loc} \mid \ev{Use \loc\ n}\\
+                       &&\mid& \ev{Branch\ }n \mid \ev{Binop\ }n\\
+    (\text{Security Tag}) & \securitytag &:=& \lock \mid \unlock\\ 
+    (\text{Event}) & \event[_{ct}] &:=& \event[^b_{ct}];\securitytag \mid \emptyevent \mid \lightning \\ 
   \end{array}
@@
 \[
-  \typerulenolabel{mscctrel:drop:tag}{}{\event[_b];\sim_{msscct}\event[_b];\securitytag}
+  \typerulenolabel{mscctrel:drop:tag}{}{\event[_{ms}]\sim_{msscct}\event[^b_{ct}];\securitytag}
 \]
@@
 \subsection{Memory-Safe, Constant-Time Trace Model with Speculation}
+
+\[
+  \begin{array}{rrcl}
+    (\text{Pre-Event}) & \event[^b_{\mathghost}] &:=& \ev{Alloc \loc\ n} \mid \ev{Dealloc \loc} \mid \ev{Use \loc\ n}\\
+                       &&\mid& \ev{Branch\ }n \mid \ev{Binop\ }n \mid \ev{Spec} \mid \ev{Rlb}\\
+    (\text{Spectre Variant}) & vX &:=& \operatorname{PHT} \\
+    (\text{Security Tag}) & \securitytag &:=& \lock_{vX} \mid \unlock\\ 
+    (\text{Event}) & \event[_{\mathghost}] &:=& \event[^b_{\mathghost}];\securitytag \mid \emptyevent \mid \lightning \\ 
+  \end{array}
+\]
 
@@
 This section defines programming languages that the secure compilers defined in the next section will use.
-To this end, this section defines the languages $\Ltms$, $\Ltrg$, $\Lms$, and $\Lscct$ which share many common elements (presented in \Cref{subsec:cs:defs}).
+To this end, this section defines the languages $\Ltms$, $\Ltrg$, $\Lms$, $\Lscct$, and $\Lspec$ which share many common elements (presented in \Cref{subsec:cs:defs}).
 $\Ltms$ is the only statically typed language and exhibits the property that all well-typed programs are \gls*{tms} (\Cref{subsec:ltms}).
@@
 $\Lscct$ (\Cref{subsec:lscct}) programs have access to a ``\gls*{cct}-mode'' and can change the leakage of emitted events according to the value of this mode (either $\obj{ON}$ or $\obj{OFF}$).
+
+Lastly, modern processors also employ speculative execution to achieve speedups, but fail to rollback properly, so that data may be left over at a cache, so that an attacker can exploit the difference in runtime for cache-hits versus cache-misses.
+$\Lspec$ extends $\Lscct$ to model this situation, while all previous languages trivially satisfy \gls*{ss}, since they do not support speculative execution at all.
 

### 2024-05-21 csf2/main.tex
implement some todos

@@
 \gls*{cct} is a security property that ensures sensitive data is not leaked via timing side-channels~\cite{kocher1996timing}.
-Together, \gls*{sms}, \gls*{tms} and \gls*{scct}, an enforceable overapproximation of \gls*{cct}, yield \gls*{msscct}, which is the gold standard of security properties for secure applications.
-Programs attaining \gls*{msscct} do not leak sensitive data either through erroneous memory accesses, nor through timing side-channels.
+Together, \gls*{sms}, \gls*{tms}, \gls*{scct}, an enforceable overapproximation of \gls*{cct}, and \gls*{ss}, yield \gls*{mssscct}, which is the gold standard of security properties for secure applications.
+Programs attaining \gls*{mssscct} do not leak sensitive data either through erroneous memory accesses, nor through timing side-channels, even under speculative execution.
 As discussed in \Cref{ex:strncpy}, these security properties can be enforced by compiler passes~\cite{bond2017vale,almeida2017jasmin}, to ensure programmers need not be aware of the architectural details of where their code will run.
@@
 Then, this paper defines compiler composition and formalises that given two passes that securely preserve two (possibly distinct) properties, their composition securely preserves the composition of those properties.
-The paper then defines several secure compiler passes, where each is either preserving a different security property (\gls*{tms}, \gls*{sms}, \gls*{scct}) or performing a security-preserving optimisation, (e.g., applying \gls*{cf} or \gls*{dce}).
-Finally, this paper shows that composing these secure compiler passes into a multi-pass compilation chain results in the end-to-end preservation of \gls*{msscct}.
+The paper then defines several secure compiler passes, where each is either preserving a different security property (\gls*{tms}, \gls*{sms}, \gls*{scct}, \gls*{ss}) or performing a security-preserving optimisation, (e.g., applying \gls*{cf} or \gls*{dce}).
+Finally, this paper shows that composing these secure compiler passes into a multi-pass compilation chain results in the end-to-end preservation of \gls*{mssscct}.
 Crucially, this paper derives the security of the multi-pass compiler from the composition of the security properties preserved by its individual passes.
@@
   \item %
-        %\Cref{sec:compprop} presents a formal framework to reason about compositions of security properties and demonstrate its usage with concrete examples:
-        This paper formalises security properties (\Cref{sec:compprop}) that are of interest for real-world compiler writers, namely \gls*{tms}, \gls*{sms} and \gls*{cct} (as identified by the plethora of work enforcing such properties individually~\cite{akritidis2009baggy,nagarakatte2009soft,nagarakatte2010cets,dhumbumroong2020boundwarden,jung2021pico,nam2019framer,shankaranarayana2023tailcheck,younan2010paricheck,zhou2023fatptrs,bond2017vale,almeida2017jasmin,kuepper2023cryptopt,cauligi2019fact}).
-        % 
-        Starting from ways to formalise those properties individually, this paper shows how to compose their formalisation.
-        % 
-        % This paper then showcases the benefits of this composition by composing \gls*{tms} and \gls*{sms} into \gls*{ms}, and ultimately adding \gls*{cct} to the composition, yielding the the conjunction of \gls*{ms} and \gls*{cct}.
-        % 
-        The resulting security property is \gls*{msscct}, i.e., the gold standard of security properties for secure programs~\cite{lemay2021ccc}.
-
-        % \MPpast{
-        % 	some of the good things here: 
-        % 	our definitions are amenable to sec comp proofs, unlike existing definitions
-        % 	MS def is like mswasm MS def, 
-        % 	but CCT is new
-        % }
+        %\Cref{sec:compcomp} studies different forms of secure compiler compositions.
+        This paper takes the secure compilation framework of Abate et al.~\cite{abate2021extacc} and extends it for compositionality. (\Cref{sec:sequential}).
+        %Concretely, it investigates upper and lower compositions, as well as the sequential composition.
+        %For this, this paper studies compiler composition as well as compilers with multiple input languages or multiple output ones, as used in existing compilation chains.
+        %
+        This paper proves that starting from two compilers that preserve two (possibly distinct) properties, their composition preserves roughly the intersection of those properties.
+        % For example, a compiler that is secure with respect to \gls*{tms} can be composed with a compiler that is secure with respect to \gls*{sms}, resulting in a compiler that is secure with respect to \gls*{ms}.
+        % 
+        From this, the paper proves that the order of composition of sequential compiler passes is irrelevant for security.
+        % 
+        This is crucial for reordering individual optimisation passes and thus generating secure and efficient code.
+        %
+        %Finally, this paper extends parts of the compositionality framework to support different trace-models between compilation passes.
 
   \item %
-        %\Cref{sec:compcomp} studies different forms of secure compiler compositions.
-        This paper takes the secure compilation framework of Abate et al.~\cite{abate2019jour} and extends it to reason about the security of all different known forms of compiler composition (\Cref{sec:compcomp}).
-        %Concretely, it investigates upper and lower compositions, as well as the sequential composition.
-        For this, this paper studies sequential compiler composition as well as compilers with multiple input languages or multiple output ones, as used in existing compilation chains.
-        %
-        This paper proves that starting from two compilers that preserve two (possibly distinct) properties, their composition preserves the intersection of those properties.
-        % For example, a compiler that is secure with respect to \gls*{tms} can be composed with a compiler that is secure with respect to \gls*{sms}, resulting in a compiler that is secure with respect to \gls*{ms}.
-        % 
-        From this, the paper proves that the order of composition of sequential compiler passes is irrelevant for the resulting security.
-        % 
-        This is crucial for reordering individual optimisation passes and thus generating secure and efficient code.
-        %
-        Finally, this paper extends parts of the compositionality framework to support different trace-models between compilation passes.
-
-  \item %
-        This paper presents a case-study showcasing the conjunction of the previous contributions (\Cref{sec:casestud:defs,sec:casestud:rtp}).
-        To this end, it presents a compilation chain consisting of several passes that ultimately preserves \gls*{msscct} by means of composing the individual, secure passes concerning \gls*{tms}, \gls*{sms}, and \gls*{scct}, respectively.
-        Furthermore, the chain includes two optimisation passes: One performs \gls*{dce} and the other \gls*{cf}.
+        This paper presents a case-study with five programming languages showcasing the previous contribution (\Cref{sec:casestud:defs,sec:casestud:rtp}).
+        To this end, it presents a compilation chain consisting of six passes, two of which are well-known optimisations \gls*{dce} and \gls*{cf}, that ultimately preserve \gls*{mssscct} by means of composing the individual, secure passes concerning \gls*{tms}, \gls*{sms}, \gls*{scct}, and \gls*{ss}, respectively.
+        %Furthermore, the chain includes two optimisation passes: One performs \gls*{dce} and the other \gls*{cf}.
         The formalisation of this case study showcases the power of the presented framework: The divide-and-conquer approach to software engineering is a viable strategy even for the development of secure compilers.
 %        This demonstrates that the usage of the presented framework is both simpler and more expressive compared to one monolithic security proof of the whole compilation chain.
+
+  \item %
+        The considered compilation passes are analyzed in terms of their compatibility with each other (\Cref{sec:formalities}).
+        In doing so, the paper provides formal insights into some technicalities of the concrete proofs necessary as technical setup to reap the benefits of the presented compositionality framework.
 
@@
 
-\subsection{Secure Sequential Composition}\label{sec:sequential}
+\section{Secure Composition}\label{sec:sequential}
 
@@
 
+Since the composition of secure compilers is again a secure compiler, the theorems generalise to a whole chain of $n$ secure compilers. 
+Moreover, the previously defined theory dually holds for the existential image $\tau_\sim\left(\src{\pi}\right)$, but these results have been moved to the technical report in the interest of space and interest.
+
+\subsection{Secure Compiler Composition with Same Trace Models}
 When the cross-language trace relation is an equality, \Cref{thm:rtpsim:sig} collapse:
@@
 
-\begin{theorem}[Sequential Composition of Secure Compilers]\label{thm:rtp}
+\begin{corollary}[Sequential Composition of Secure Compilers]\label{corr:rtp}
   If \bul{$\rtp{\cc{\src{L}}{\trg{L}}}{\class[_{1}]}$} and \rul{$\rtp{\cc{\trg{L}}{\obj{L}}}{\class[_{2}]}$}, then \oul{$\rtp{\cc{\src{L}}{\trg{L}}\circ\cc{\trg{L}}{\obj{L}}}{\class[_{1}]\cap\class[_{2}]}$}.
 \Coqed
-\end{theorem}
-
-Since the composition of secure compilers is again a secure compiler, the theorems generalise to a whole chain of $n$ secure compilers.
+\end{corollary}
+
+\Cref{corr:rtp} provides an easy to use theorem to compose secure compilers without any caveats. 
+However, while \Cref{thm:rtpsim:sig} explicitly requires that $\sim_1$ is well-formed with respect to $\trg{\class[_2]}$, if $\sim_2$ is not well-formed with respect to $\trg{\class[_1]}$, care must be taken.
+Worst-case is that the resulting class $\sigma_{\sim_1\bullet\sim_2}\left(\trg{\class[_1]}\cap\trg{\class[_2]}\right)$ becomes empty, even though $\trg{\class[_1]}\cap\trg{\class[_2]}$ is not necessarily empty.
+\Cref{sec:formalities:compoex} provides an example of this phenomenon.
 
@@
 
-This section introduces trace properties of interest for this paper: \gls*{tms}, \gls*{sms}, \gls*{ms}, \gls*{scct}, and \gls*{ss} (\Cref{subsec:propdefs}).
+This section introduces trace properties of interest for this paper: \gls*{tms}, \gls*{sms}, \gls*{ms}, \gls*{scct}, and \gls*{ss}.
 These properties are of practical importance (as mentioned in \Cref{sec:introduction}) and also of interest in the case study (\Cref{sec:casestud:defs,sec:casestud:rtp}) presented later. 
-For each of these properties, the technical appendix accompanying the paper defines corresponding monitors, which refine a given property.
-
-\subsection{Memory Safety Trace Model}\label{subsec:propdefs}\label{subsec:basic:memsafety:tracemodel}
+While up to \gls*{scct}, none of the properties is novel, for self-containment of the paper, their definition is necessary for the case study.
+For each of the presented properties, the technical appendix accompanying the paper defines corresponding monitors, which refine a given property.
+
+\subsection{Memory Safety Trace Model}\label{subsec:basic:memsafety:tracemodel}
 
@@
 
-All presented programming languages share a common fragment. 
-Programs ($\prog{\library_{\ctx}}{\library_{\comp}}$) consist of libraries ($\library$) that belong to a context ($\ctx$) or to a component ($\comp$), and that are lists of functions ($\function$).
-Functions ($\lfunction{\textit{foo}}{x}{\expr}$) have a single argument $x$ and an expression ($\expr$) as a body which may have $x$ as a free variable.
-Expressions can be variables ($x$), values ($\valueexpr$), binary operations ($\lbinop{\expr[_1]}{\expr[_2]}$), reading from ($\lget{x}{\expr}$) and writing to ($\lset{x}{\expr[_1]}{\expr[_2]}$) memory, memory allocation ($\llet{x}{\lnew{\expr[_1]}{\expr[_2]}}{\expr[_3]}$) as well as deallocation ($\ldelete{x}$), let expressions ($\llet{x}{\expr[_1]}{\expr[_2]}$), pairs ($\lpair{\expr[_1]}{\expr[_2]}$), their first ($\lproja{\expr}$) and second projection ($\lprojb{\expr}$), conditionals ($\lifz{\expr[_1]}{\expr[_2]}{\expr[_3]}$), function calls ($\lcall{\texttt{foo}}{\expr}$), returns ($\lreturn{\expr}$), and an aborting expression ($\labort$).
+All presented programming languages share a common fragment which is partially presented here and in full detail in the technical report. 
+The trace models of all languages are similar to those presented earlier (\Cref{sec:compprop}).
+One technical detail is the addition of a sandboxtag, indicating who is to blame for an emitted action: context ($\ctx$) or component ($\comp$).
+
+% Programs ($\prog{\library_{\ctx}}{\library_{\comp}}$) consist of libraries ($\library$) that belong to a context ($\ctx$) or to a component ($\comp$), and that are lists of functions ($\function$).
+% Functions ($\lfunction{\textit{foo}}{x}{\expr}$) have a single argument $x$ and an expression ($\expr$) as a body which may have $x$ as a free variable.
+% Expressions can be variables ($x$), values ($\valueexpr$), binary operations ($\lbinop{\expr[_1]}{\expr[_2]}$), reading from ($\lget{x}{\expr}$) and writing to ($\lset{x}{\expr[_1]}{\expr[_2]}$) memory, memory allocation ($\llet{x}{\lnew{\expr[_1]}{\expr[_2]}}{\expr[_3]}$) as well as deallocation ($\ldelete{x}$), let expressions ($\llet{x}{\expr[_1]}{\expr[_2]}$), pairs ($\lpair{\expr[_1]}{\expr[_2]}$), their first ($\lproja{\expr}$) and second projection ($\lprojb{\expr}$), conditionals ($\lifz{\expr[_1]}{\expr[_2]}{\expr[_3]}$), function calls ($\lcall{\texttt{foo}}{\expr}$), returns ($\lreturn{\expr}$), and an aborting expression ($\labort$).
 
@@
 The type system of $\src{L_{\tmssafe}}$ enforces that every well-typed $\src{L_{\tmssafe}}$ program satisfies \gls*{tms} (\Cref{thm:wt:tms}).
-The proof of this theorem relies on a projection $\msfilterLtms{\src{\event}}=\event$ from $\src{L_{\tmssafe}}$ events to specification events $\event$, because the properties defined earlier (\Cref{subsec:propdefs}) are defined in the specification trace model.
 Hereby, the map $\thelocmap(\src{\loc})=\loc$ maps an $\Ltms$ location to a location $\loc$ of the specification trace model.
-
-Projection $\msfilterLtms{\src{\event}}$ translates an action into its correspondent (e.g., \Cref{tr:ltms-filter-alloc}); since these rules are straightforward they are mostly omitted.
-Crucially, the projection filters irrelevant actions such as context-made ones (\Cref{tr:ltms-filter-context}) and interaction events (\Cref{tr:ltms-filter-start}).
 
@@
 Furthermore, the syntax of $\Ltrg$ is extended with a way to inspect whether a pointer is freed ($\trg{\lispoisoned{x}}$), evaluating to $\trg{0}$ if it is freed and to $\trg{1}$ otherwise.
-The projection $\msfilterL{\trg{\trace}}$ is structurally equal to $\msfilterLtms{\src{\trace}}$.
 
@@
 Security tags $\obj{\securitytag}$ are on the usual secrecy lattice, where $\obj{\lock}\le\obj{\securitytag}$ and $\obj{\securitytag}\le\obj{\unlock}$.
-To prevent secrets from leaking but still enable reasoning about memory safety, $\Lscct$ extends pre-events with $\obj{\widehat{Get}\ \loc\ \valueexpr}$ and $\obj{\widehat{Set}\ \loc\ \valueexpr}$.
+To prevent secrets from leaking but still enable reasoning about memory safety, $\Lscct$ extends pre-events with $\obj{\ev{iGet}\ \loc\ \valueexpr}$ and $\obj{\ev{iSet}\ \loc\ \valueexpr}$.
 These indicate reads from and writes to memory without leaking secret information involved in the access.
@@
 A barrier prevents any events whatsoever besides $\ird{\ev{Rlb}}$ when run in speculative execution mode and does nothing when run in normal mode.
+The speculative semantics are taken from existing work~\cite{fabian:mispredict}.
 
@@
     \obj{\ev{Alloc}\ \loc_{y}\ 12;\comp;\unlock} & \mid & \obj{\ev{Alloc}\ \loc_{y}\ 12;\comp;\unlock} & \mid & \ev{Alloc}\ \loc_{y}\ 12;\comp;\unlock & \mid & \ev{Alloc}\ \loc_{x}\ 12;\comp;\unlock\\
-    \obj{\widehat{\ev{Get}}\ \loc_{x}\ 0;\comp;}\text{\colorbox{lightgray}{$\obj{\lock}$}} & \mid & \obj{\ev{Get}\ \loc_{x}\ 0;\comp;\lock} & \mid & \ev{Use}\ \loc_{x}\ 0;\comp;\text{\colorbox{lightgray}{$\unlock$}} & \mid & \ev{Use}\ \loc_{x}\ 0;\comp;\lock\\
+    \obj{{\ev{iGet}}\ \loc_{x}\ 0;\comp;}\text{\colorbox{lightgray}{$\obj{\lock}$}} & \mid & \obj{\ev{Get}\ \loc_{x}\ 0;\comp;\lock} & \mid & \ev{Use}\ \loc_{x}\ 0;\comp;\text{\colorbox{lightgray}{$\unlock$}} & \mid & \ev{Use}\ \loc_{x}\ 0;\comp;\lock\\
     \text{\colorbox{yellow}{$\obj{\emptyevent}$}} & \mid & \obj{\ev{Branch}\ 0;\comp;\lock} & \mid & \emptyevent & \mid & \ev{Branch}\ 0;\comp;\lock\\
-    \obj{\widehat{\ev{Get}}\ \loc_{x}\ 0;\comp;}\text{\colorbox{lightgray}{$\obj{\lock}$}} & \mid & \obj{\ev{Get}\ \loc_{x}\ 0;\comp;\lock} & \mid & \ev{Use}\ \loc_{x}\ 0;\comp;\text{\colorbox{lightgray}{$\unlock$}} & \mid & \ev{Use}\ \loc_{x}\ 0;\comp;\lock\\
-    \obj{\widehat{\ev{Set}}\ \loc_{y}\ 0\ \mathtt{'H'};\comp;}\text{\colorbox{lightgray}{$\obj{\lock}$}} & \mid & \obj{\ev{Set}\ \loc_{y}\ 0\ \mathtt{'H'};\comp;\lock} & \mid & \ev{Use}\ \loc_{y}\ 0;\comp;\text{\colorbox{lightgray}{$\unlock$}} & \mid & \ev{Use}\ \loc_{y}\ 0;\comp;\lock\\
-    \obj{\widehat{\ev{Get}}\ \loc_{x}\ 1;\comp;}\text{\colorbox{lightgray}{$\obj{\lock}$}} & \mid & \obj{\ev{Get}\ \loc_{x}\ 1;\comp;\lock} & \mid & \ev{Use}\ \loc_{x}\ 1;\comp;\text{\colorbox{lightgray}{$\unlock$}} & \mid & \ev{Use}\ \loc_{x}\ 1;\comp;\lock\\
+    \obj{{\ev{iGet}}\ \loc_{x}\ 0;\comp;}\text{\colorbox{lightgray}{$\obj{\lock}$}} & \mid & \obj{\ev{Get}\ \loc_{x}\ 0;\comp;\lock} & \mid & \ev{Use}\ \loc_{x}\ 0;\comp;\text{\colorbox{lightgray}{$\unlock$}} & \mid & \ev{Use}\ \loc_{x}\ 0;\comp;\lock\\
+    \obj{{\ev{iSet}}\ \loc_{y}\ 0\ \mathtt{'H'};\comp;}\text{\colorbox{lightgray}{$\obj{\lock}$}} & \mid & \obj{\ev{Set}\ \loc_{y}\ 0\ \mathtt{'H'};\comp;\lock} & \mid & \ev{Use}\ \loc_{y}\ 0;\comp;\text{\colorbox{lightgray}{$\unlock$}} & \mid & \ev{Use}\ \loc_{y}\ 0;\comp;\lock\\
+    \obj{{\ev{iGet}}\ \loc_{x}\ 1;\comp;}\text{\colorbox{lightgray}{$\obj{\lock}$}} & \mid & \obj{\ev{Get}\ \loc_{x}\ 1;\comp;\lock} & \mid & \ev{Use}\ \loc_{x}\ 1;\comp;\text{\colorbox{lightgray}{$\unlock$}} & \mid & \ev{Use}\ \loc_{x}\ 1;\comp;\lock\\
     \text{\colorbox{yellow}{$\obj{\emptyevent}$}} & \mid & \obj{\ev{Branch}\ 0;\comp;\lock} & \mid & \emptyevent & \mid & \ev{Branch}\ 0;\comp;\lock\\
     \vdots & \mid & \vdots & \mid & \vdots & \mid & \vdots \\
-    \obj{\widehat{\ev{Get}}\ \loc_{x}\ 12;\comp;}\text{\colorbox{lightgray}{$\obj{\lock}$}} & \mid & \obj{\ev{Get}\ \loc_{x}\ 12;\comp;\lock} & \mid &  \ev{Use}\ \loc_{x}\ 12;\comp;\text{\colorbox{lightgray}{$\unlock$}} & \mid & \ev{Use}\ \loc_{x}\ 12;\comp;\lock\\
+    \obj{{\ev{iGet}}\ \loc_{x}\ 12;\comp;}\text{\colorbox{lightgray}{$\obj{\lock}$}} & \mid & \obj{\ev{Get}\ \loc_{x}\ 12;\comp;\lock} & \mid &  \ev{Use}\ \loc_{x}\ 12;\comp;\text{\colorbox{lightgray}{$\unlock$}} & \mid & \ev{Use}\ \loc_{x}\ 12;\comp;\lock\\
     \text{\colorbox{yellow}{$\obj{\emptyevent}$}}& \mid & \obj{\ev{Branch}\ 1;\comp;\lock} & \mid &  \emptyevent & \mid & \ev{Branch}\ 1;\comp;\lock\\
     \obj{\ev{Dealloc}\ \loc_{y};\comp;\unlock} & \mid & \obj{\ev{Dealloc}\ \loc_{y};\comp;\lock} & \mid &  \ev{Dealloc}\ \loc_{y};\comp;\unlock & \mid & \ev{Dealloc}\ \loc_{y};\comp;\unlock\\
-    \obj{\widehat{\ev{Get}}\ \loc_{y}\ 6;\comp;}\text{\colorbox{lightgray}{$\obj{\lock}$}} & \mid & \obj{\ev{Get}\ \loc_{y}\ 6;\comp;\lock} & \mid &  \ev{Use}\ \loc_{y}\ 6;\comp;\text{\colorbox{lightgray}{$\unlock$}} & \mid & \ev{Use}\ \loc_{y}\ 6;\comp;\lock\\
+    \obj{{\ev{iGet}}\ \loc_{y}\ 6;\comp;}\text{\colorbox{lightgray}{$\obj{\lock}$}} & \mid & \obj{\ev{Get}\ \loc_{y}\ 6;\comp;\lock} & \mid &  \ev{Use}\ \loc_{y}\ 6;\comp;\text{\colorbox{lightgray}{$\unlock$}} & \mid & \ev{Use}\ \loc_{y}\ 6;\comp;\lock\\
   \end{array}

### 2024-05-21 csf2/main.tex
Add bnfs

@@
 
-\subsection{Memory-Safe, Constant-Time Trace Model with Speculation}
+\subsection{Memory-Safe, Constant-Time Trace Model with Speculation}\label{subsec:msctss:tracemodel}
 
@@
 One technical detail is the addition of a sandboxtag, indicating who is to blame for an emitted action: context ($\ctx$) or component ($\comp$).
+This is a standard technique in secure compilation in order to rule out irrelevant context-level events. 
+For example, a context immediately deallocating an allocated memory region twice trivially violates memory safety, but the main interest in secure compilation is at component-level events and whether those signal, e.g., a memory safety issue.
 
@@
 
-$\src{L_{\tmssafe}}$ uses the same syntax as presented earlier (\Cref{sec:casestud:defs}) without extensions to the term level.
+\vspace{-1.5em}
+\[
+  \begin{array}{rcl}
+    \src{\event[^b]} &:=& \src{\ev{Alloc \loc\ n}} \mid \src{\ev{Dealloc \loc}} \mid \src{\ev{Get \loc\ n}} \\
+                     && \mid \src{\ev{Set \loc\ n}} \mid \src{\ev{Branch\ n}} \mid \src{\ev{Binop\ n}}\\
+   && \src{\sandboxtag} := \src{\ctx} \mid \src{\comp} \hspace{3em}
+   \src{\event} := \src{\event[^b];\sandboxtag} \mid \src{\emptyevent} \mid \src{\lightning} \\ 
+    \src{\expr} &:=& \src{x} \mid \src{n} \mid \src{\lbinop{\expr[_1]}{\expr[_2]}} \mid \src{\lifz{\expr[_1]}{\expr[_2]}{\expr[_3]}} \mid \src{\llet{x}{\expr[_1]}{\expr[_2]}} \\
+                && \src{\library_{\ctx}} : \src{Vars} \to \src{\expr} \hspace{4em}
+    \src{\library_{\comp}} : \src{Vars} \to \src{\expr} \\
+  \end{array}
+\]
+
 $\src{L_{\tmssafe}}$ is statically typed, where the type system is inspired by $L^{3}$~\cite{morrisett2005L3,scherer2018fabulous}.
 The type system of $\src{L_{\tmssafe}}$ enforces that every well-typed $\src{L_{\tmssafe}}$ program satisfies \gls*{tms} (\Cref{thm:wt:tms}).
-Hereby, the map $\thelocmap(\src{\loc})=\loc$ maps an $\Ltms$ location to a location $\loc$ of the specification trace model.
+%Hereby, the map $\thelocmap(\src{\loc})=\loc$ maps an $\Ltms$ location to a location $\loc$ of the specification trace model.
 
@@
 
+The trace model is largely similar to the memory safety trace model (\Cref{subsec:basic:memsafety:tracemodel}).
+$\Ltms$ dynamic semantics do not get stuck if a memory-safety violation is encountered, such as a write out of bounds.
+
 \subsection{$\Ltrg$: A Memory-Unsafe Language}\label{subsec:lsms}
+
+\vspace{-1.5em}
+\[
+  \begin{array}{rcl}
+    \trg{\expr} &:=& \dots \mid \trg{\lhast{\expr}{\tau}} \mid \trg{\lispoisoned{x}}
+  \end{array}
+\]
 
@@
 Furthermore, the syntax of $\Ltrg$ is extended with a way to inspect whether a pointer is freed ($\trg{\lispoisoned{x}}$), evaluating to $\trg{0}$ if it is freed and to $\trg{1}$ otherwise.
+No changes are done to the trace model, but $\Ltrg$ has no static typing, making double-free patterns possible.
 
 \subsection{$\Lms$: Another Memory-Unsafe Language}\label{subsec:lms}
-$\Lms$ is exactly equal to $\Ltrg$ (\Cref{subsec:lsms}), but used to signify that the code considered is after applying $\cc{\Ltrg}{\Lms}$.
+$\Lms$ is exactly equal to $\Ltrg$ (\Cref{subsec:lsms}), but used to emphasize that this is code after applying $\cc{\Ltrg}{\Lms}$ (\Cref{subsec:cs:ms}).
 
 \subsection{$\Lscct$: A Memory-Unsafe Language with a Data Independent Timing Mode}\label{subsec:lscct}
+
+\vspace{-1.5em}
+\[
+  \begin{array}{rcl}
+    \obj{D} &:=& \obj{ON} \mid \obj{OFF} \hspace{2em}
+    \obj{\securitytag} := \obj{\lock} \mid \obj{\unlock} \\
+    \obj{\event[^b]} &:=& \dots \mid \obj{iGet\ \loc\ v} \mid \obj{iSet\ \loc\ v} \mid \obj{Binop\ n} \mid \obj{Branch\ n} \\
+    \obj{\expr} &:=& \dots \mid \obj{\lrddoit{x}{\expr}} \mid \obj{\lwrdoit{D}} \mid \obj{\llet{x^{\securitytag}}{\expr[_1]}{\expr[_2]}}\\
+    \obj{\event} &:=& \obj{\event[^b];\sandboxtag;\securitytag} \mid \dots
+  \end{array}
+\]
 
@@
 For the other languages seen earlier, the mode is intuitively always-on, i.e., the mode of execution always uses data independent timing.
-The language also adds user-annotations to values and variables to know their secrecy $\obj{\securitytag}$, which is either high $\obj{\lock}$ or low $\obj{\unlock}$.
+The language also adds user-annotations for secrecy $\obj{\securitytag}$, which is either high $\obj{\lock}$ or low $\obj{\unlock}$.
 Security tags $\obj{\securitytag}$ are on the usual secrecy lattice, where $\obj{\lock}\le\obj{\securitytag}$ and $\obj{\securitytag}\le\obj{\unlock}$.
 To prevent secrets from leaking but still enable reasoning about memory safety, $\Lscct$ extends pre-events with $\obj{\ev{iGet}\ \loc\ \valueexpr}$ and $\obj{\ev{iSet}\ \loc\ \valueexpr}$.
-These indicate reads from and writes to memory without leaking secret information involved in the access.
+This way, memory accesses to secret data are still exposed as memory accesses, despite being protected by data-independent timing mode.
 Moreover, the language extends pre-events with $\obj{\ev{Branch}\ n}$ and $\obj{\ev{Binop}\ n}$ that are emitted when evaluating a branch or certain binary expressions, such as division, respectively, whenever the data independent timing mode is inactive.
-The following rules demonstrate how this is handled semantically.
+Events are extended with security-tag to signal the secrecy of the involved data.
 
@@
 When the data independent timing mode is inactive, pre-events $\obj{\ev{Branch}\ n}$ and $\obj{\ev{Binop}\ n}$ are emitted for conditionals and binary operations, respectively.
+Otherwise, just like in the semantics of the earlier languages, $\obj{\emptyevent}$ is emitted for binary and branching operations.
 %To mark this mode as active or inactive, \Cref{tr:e-wrdoit} allows to access the register in $\obj{\statevar}$.
@@
   Consider again \Cref{ex:strncpy}, with a context copying the string \texttt{Hello World}, where everything is marked with a high security tag: $\lock$.
-  The left half of \Cref{fig:ex-cct} ($\Lscct$, the two columns on the left), describes the execution trace of the program, while the right side of the table ($Spec$, the two columns on the right), describes the respective projections $\scctfilterLscct{\obj{\trace}}$ to the specification trace model.
+  The left half of \Cref{fig:ex-cct} ($\Lscct$, the two columns on the left), describes the execution trace of the program, while the right side of the table ($Spec$, the two columns on the right), describes the related specification trace (\Cref{subsec:msctss:tracemodel}).
   Read from top to bottom, the figure shows parts of the execution trace. 
@@
 \subsection{$\Lspec$: A Memory-Unsafe Language with Speculation}\label{subsec:lspec}
+
+\vspace{-1.5em}
+\[
+  \begin{array}{rcl}
+    \ird{\event[^b]} &:=& \dots \mid \ird{Spec} \mid \ird{Rlb} \mid \ird{Barrier}\\
+    \ird{\expr[^b]} &:=& \dots \mid \ird{\lbarrier}
+  \end{array}
+\]
 
@@
 The trace model is extended with $\ird{\ev{Spec}}$, $\ird{\ev{Rlb}}$, and $\ird{\ev{Barrier}}$ events that signal the respective operational action. 
-A barrier prevents any events whatsoever besides $\ird{\ev{Rlb}}$ when run in speculative execution mode and does nothing when run in normal mode.
+A $\ird{\lbarrier}$ prevents any execution whatsoever besides $\ird{\ev{Rlb}}$ when run in speculative execution mode and does nothing when run in normal mode.
 The speculative semantics are taken from existing work~\cite{fabian:mispredict}.
@@
     Active & \mid & Inactive & \mid & Active & \mid & Inactive \\\hline
-    \obj{\ev{Alloc}\ \loc_{x}\ 12;\comp;\unlock} & \mid & \obj{\ev{Alloc}\ \loc_{x}\ 12;\comp;\unlock} & \mid & \ev{Alloc}\ \loc_{x}\ 12;\comp;\unlock & \mid & \ev{Alloc}\ \loc_{x}\ 12;\comp;\unlock\\
-    \obj{\ev{Alloc}\ \loc_{y}\ 12;\comp;\unlock} & \mid & \obj{\ev{Alloc}\ \loc_{y}\ 12;\comp;\unlock} & \mid & \ev{Alloc}\ \loc_{y}\ 12;\comp;\unlock & \mid & \ev{Alloc}\ \loc_{x}\ 12;\comp;\unlock\\
-    \obj{{\ev{iGet}}\ \loc_{x}\ 0;\comp;}\text{\colorbox{lightgray}{$\obj{\lock}$}} & \mid & \obj{\ev{Get}\ \loc_{x}\ 0;\comp;\lock} & \mid & \ev{Use}\ \loc_{x}\ 0;\comp;\text{\colorbox{lightgray}{$\unlock$}} & \mid & \ev{Use}\ \loc_{x}\ 0;\comp;\lock\\
-    \text{\colorbox{yellow}{$\obj{\emptyevent}$}} & \mid & \obj{\ev{Branch}\ 0;\comp;\lock} & \mid & \emptyevent & \mid & \ev{Branch}\ 0;\comp;\lock\\
-    \obj{{\ev{iGet}}\ \loc_{x}\ 0;\comp;}\text{\colorbox{lightgray}{$\obj{\lock}$}} & \mid & \obj{\ev{Get}\ \loc_{x}\ 0;\comp;\lock} & \mid & \ev{Use}\ \loc_{x}\ 0;\comp;\text{\colorbox{lightgray}{$\unlock$}} & \mid & \ev{Use}\ \loc_{x}\ 0;\comp;\lock\\
-    \obj{{\ev{iSet}}\ \loc_{y}\ 0\ \mathtt{'H'};\comp;}\text{\colorbox{lightgray}{$\obj{\lock}$}} & \mid & \obj{\ev{Set}\ \loc_{y}\ 0\ \mathtt{'H'};\comp;\lock} & \mid & \ev{Use}\ \loc_{y}\ 0;\comp;\text{\colorbox{lightgray}{$\unlock$}} & \mid & \ev{Use}\ \loc_{y}\ 0;\comp;\lock\\
-    \obj{{\ev{iGet}}\ \loc_{x}\ 1;\comp;}\text{\colorbox{lightgray}{$\obj{\lock}$}} & \mid & \obj{\ev{Get}\ \loc_{x}\ 1;\comp;\lock} & \mid & \ev{Use}\ \loc_{x}\ 1;\comp;\text{\colorbox{lightgray}{$\unlock$}} & \mid & \ev{Use}\ \loc_{x}\ 1;\comp;\lock\\
-    \text{\colorbox{yellow}{$\obj{\emptyevent}$}} & \mid & \obj{\ev{Branch}\ 0;\comp;\lock} & \mid & \emptyevent & \mid & \ev{Branch}\ 0;\comp;\lock\\
+    \obj{\ev{Alloc}\ \loc_{x}\ 12;\comp;\unlock} & \mid & \obj{\ev{Alloc}\ \loc_{x}\ 12;\comp;\unlock} & \mid & \ev{Alloc}\ \loc_{x}\ 12;\unlock & \mid & \ev{Alloc}\ \loc_{x}\ 12;\unlock\\
+    \obj{\ev{Alloc}\ \loc_{y}\ 12;\comp;\unlock} & \mid & \obj{\ev{Alloc}\ \loc_{y}\ 12;\comp;\unlock} & \mid & \ev{Alloc}\ \loc_{y}\ 12;\unlock & \mid & \ev{Alloc}\ \loc_{x}\ 12;\unlock\\
+    \obj{{\ev{iGet}}\ \loc_{x}\ 0;\comp;}\text{\colorbox{lightgray}{$\obj{\lock}$}} & \mid & \obj{\ev{Get}\ \loc_{x}\ 0;\comp;\lock} & \mid & \ev{Use}\ \loc_{x}\ 0;\text{\colorbox{lightgray}{$\unlock$}} & \mid & \ev{Use}\ \loc_{x}\ 0;\lock\\
+    \text{\colorbox{yellow}{$\obj{\emptyevent}$}} & \mid & \obj{\ev{Branch}\ 0;\comp;\lock} & \mid & \emptyevent & \mid & \ev{Branch}\ 0;\lock\\
+    \obj{{\ev{iGet}}\ \loc_{x}\ 0;\comp;}\text{\colorbox{lightgray}{$\obj{\lock}$}} & \mid & \obj{\ev{Get}\ \loc_{x}\ 0;\comp;\lock} & \mid & \ev{Use}\ \loc_{x}\ 0;\text{\colorbox{lightgray}{$\unlock$}} & \mid & \ev{Use}\ \loc_{x}\ 0;\lock\\
+    \obj{{\ev{iSet}}\ \loc_{y}\ 0\ \mathtt{'H'};\comp;}\text{\colorbox{lightgray}{$\obj{\lock}$}} & \mid & \obj{\ev{Set}\ \loc_{y}\ 0\ \mathtt{'H'};\comp;\lock} & \mid & \ev{Use}\ \loc_{y}\ 0;\text{\colorbox{lightgray}{$\unlock$}} & \mid & \ev{Use}\ \loc_{y}\ 0;\lock\\
+    \obj{{\ev{iGet}}\ \loc_{x}\ 1;\comp;}\text{\colorbox{lightgray}{$\obj{\lock}$}} & \mid & \obj{\ev{Get}\ \loc_{x}\ 1;\comp;\lock} & \mid & \ev{Use}\ \loc_{x}\ 1;\text{\colorbox{lightgray}{$\unlock$}} & \mid & \ev{Use}\ \loc_{x}\ 1;\lock\\
+    \text{\colorbox{yellow}{$\obj{\emptyevent}$}} & \mid & \obj{\ev{Branch}\ 0;\comp;\lock} & \mid & \emptyevent & \mid & \ev{Branch}\ 0;\lock\\
     \vdots & \mid & \vdots & \mid & \vdots & \mid & \vdots \\
-    \obj{{\ev{iGet}}\ \loc_{x}\ 12;\comp;}\text{\colorbox{lightgray}{$\obj{\lock}$}} & \mid & \obj{\ev{Get}\ \loc_{x}\ 12;\comp;\lock} & \mid &  \ev{Use}\ \loc_{x}\ 12;\comp;\text{\colorbox{lightgray}{$\unlock$}} & \mid & \ev{Use}\ \loc_{x}\ 12;\comp;\lock\\
-    \text{\colorbox{yellow}{$\obj{\emptyevent}$}}& \mid & \obj{\ev{Branch}\ 1;\comp;\lock} & \mid &  \emptyevent & \mid & \ev{Branch}\ 1;\comp;\lock\\
-    \obj{\ev{Dealloc}\ \loc_{y};\comp;\unlock} & \mid & \obj{\ev{Dealloc}\ \loc_{y};\comp;\lock} & \mid &  \ev{Dealloc}\ \loc_{y};\comp;\unlock & \mid & \ev{Dealloc}\ \loc_{y};\comp;\unlock\\
-    \obj{{\ev{iGet}}\ \loc_{y}\ 6;\comp;}\text{\colorbox{lightgray}{$\obj{\lock}$}} & \mid & \obj{\ev{Get}\ \loc_{y}\ 6;\comp;\lock} & \mid &  \ev{Use}\ \loc_{y}\ 6;\comp;\text{\colorbox{lightgray}{$\unlock$}} & \mid & \ev{Use}\ \loc_{y}\ 6;\comp;\lock\\
+    \obj{{\ev{iGet}}\ \loc_{x}\ 12;\comp;}\text{\colorbox{lightgray}{$\obj{\lock}$}} & \mid & \obj{\ev{Get}\ \loc_{x}\ 12;\comp;\lock} & \mid &  \ev{Use}\ \loc_{x}\ 12;\text{\colorbox{lightgray}{$\unlock$}} & \mid & \ev{Use}\ \loc_{x}\ 12;\lock\\
+    \text{\colorbox{yellow}{$\obj{\emptyevent}$}}& \mid & \obj{\ev{Branch}\ 1;\comp;\lock} & \mid &  \emptyevent & \mid & \ev{Branch}\ 1;\lock\\
+    \obj{\ev{Dealloc}\ \loc_{y};\comp;\unlock} & \mid & \obj{\ev{Dealloc}\ \loc_{y};\comp;\lock} & \mid &  \ev{Dealloc}\ \loc_{y};\unlock & \mid & \ev{Dealloc}\ \loc_{y};\unlock\\
+    \obj{{\ev{iGet}}\ \loc_{y}\ 6;\comp;}\text{\colorbox{lightgray}{$\obj{\lock}$}} & \mid & \obj{\ev{Get}\ \loc_{y}\ 6;\comp;\lock} & \mid &  \ev{Use}\ \loc_{y}\ 6;\text{\colorbox{lightgray}{$\unlock$}} & \mid & \ev{Use}\ \loc_{y}\ 6;\lock\\
   \end{array}
@@
 
-\subsection{Robust (Spatial) Memory Safety Preservation}\label{subsec:cs:ms}
+\subsection{Robust Spatial Memory Safety Preservation}\label{subsec:cs:ms}
 

### 2024-05-21 csf2/main.tex
figs

@@
 A $\ird{\lbarrier}$ prevents any execution whatsoever besides $\ird{\ev{Rlb}}$ when run in speculative execution mode and does nothing when run in normal mode.
-The speculative semantics are taken from existing work~\cite{fabian:mispredict}.
+The speculative semantics are inspired by existing work~\cite{guarnieri2018spectector}.
 
@@
     \node[right=1.5 of E] (O) {$\obj{L_{\scctsafe}}$};
+    \node[right=1.5 of O] (Z) {$\ird{L_{\mathghost}}$};
 
@@
     \draw[->] (E) to[sloped,above] node[align=center,font=\scriptsize] {\gls*{scct}\\ \Cref{thm:ccscct:rtp:scct}} (O);
+    \draw[->] (O) to[sloped,above] node[align=center,font=\scriptsize] {\gls*{ss}\\ \Cref{thm:ccspec:rtp:spec}} (Z);
 
@@
     \node[right=1.25 of secopts] (secscct) {{\scriptsize\Cref{subsec:cs:scct}}};
+    \node[right=1.00 of secscct] (secss) {{\scriptsize\Cref{subsec:cs:ss}}};
 
@@
     \draw[thick,dotted,->] (S) to[out=-45,in=198,sloped] node[align=center,font=\scriptsize] {\gls*{ms}+\gls*{scct}\\ \Cref{thm:ccall:rtp:msscct}} (O);
+    \draw[thick,dotted,->] (S) to[out=-50,in=210,sloped] node[align=center,font=\scriptsize] {\gls*{mssscct}\\ \Cref{thm:ccall:rtp:mssscct}} (Z);
 
@@
     \draw[dashed] ($(secscct)-(-1.4,0)$) -- ($(secscct)-(-1.4,1.25)$);
+    \draw[dashed] ($(secss)-(-1.0,0)$) -- ($(secss)-(-1.0,1.25)$);
   \end{tikzpicture}

### 2024-05-22 csf2/main.tex
add theorem 6.9

@@
 
-\subsection{Robust Preservation of Intersection of Memory Safety and Strict Cryptographic Constant Time}
-
-Let $\ccmsscct$ be the compiler that is the composition of $\cca$, $\ccb$, $\cccf$, $\ccdce$, and $\ccscct$, then the following theorem holds.
-
-\begin{theorem}[$\ccmsscct$ is secure with respect to \gls*{scct}]\label{thm:ccall:rtp:msscct}
-  $\rtp{\cc{\Ltms}{\Lscct}}{\mssafe\cap\scctsafe}$ % \Coqed
-\end{theorem}
-% \begin{proof}\MP{keep?}
-% 	From \Thmref{thm:ccab:rtp:ms}, we have that any $\Ltms$ program $\src{\progvar}$ compiles into a $\Lms$ program $\irl{\progvar}$ that robustly satisfies \gls*{ms}.
-% 	Then, from \Thmref{thm:cccfccdce:rtp:ms} we have that $\irl{\progvar}$ gets optimised to a program $\irl{\progvar[']}$ that is also \gls*{ms}, where the order of optimisations does not matter for $\irl{\progvar[']}$ to be \gls*{ms}.
-%   Assuming $\irl{\progvar[']}$ robustly satisfies \gls*{scct}, by \Thmref{thm:ccscct:rtp:scct} it compiles to an $\Lscct$ program $\obj{\progvar}$ that robustly satisfies \gls*{scct} as well.
-%   Finally, from \Thmref{thm:rtp} it follows that, given $\src{\progvar}$ robustly satisfies \gls*{scct} and \gls*{ms}, $\obj{\progvar}$ also robustly satisfies \gls*{scct} and \gls*{ms}.
-% \end{proof}
-
 \subsection{Robust Speculative Safety Preservation}\label{subsec:cs:ss}
@@
   $\rtp{\ccspec}{\specmssafe}$ % \Coqed
+\end{theorem}
+
+\subsection{Robust Preservation of Intersection of Memory Safety, Strict Cryptographic Constant Time, and Speculative Safety}
+
+Let $\ccmssscct$ be the compiler that is the composition of $\cca$, $\ccb$, $\cccf$, $\ccdce$, $\ccscct$, and $\ccspec$, then the following theorem holds.
+
+\begin{theorem}[$\ccmssscct$ is secure with respect to \gls*{mssscct}]\label{thm:ccall:rtp:mssscct}
+  $\rtp{\cc{\Ltms}{\Lspec}}{\mssafe\cap\scctsafe\cap\sssafe}$ % \Coqed
 \end{theorem}

### 2024-05-22 csf2/main.tex
related work and such

@@
 % context
-To ensure that secure applications do not leak their secrets, they are required to uphold several security properties such as spatial and temporal memory safety as well as cryptographic constant time.
+To ensure that secure applications do not leak their secrets, they are required to uphold several security properties such as spatial and temporal memory safety, cryptographic constant time, as well as speculative safety.
 % need
@@
 \gls*{ms} is a security property obtained by composing \gls*{sms}, which ensures array accesses are all within bounds, and \gls*{tms}, which ensures pointers are only used when they are valid~\cite{azevedo2018meaningsofms,jim2002cyclone,necula2005ccured,nagarakatte2010cets,nagarakatte2009soft,akritidis2009baggy,michael2023mswasm}.
-\gls*{cct} is a security property that ensures sensitive data is not leaked via timing side-channels~\cite{kocher1996timing}.
+\gls*{cct} is a security property that ensures sensitive data is not leaked via timing side-channels~\cite{kocher1996timing} and \gls*{ss} is a security property that enforces the same but under a speculative semantics~\cite{guarnieri2018spectector}.
 Together, \gls*{sms}, \gls*{tms}, \gls*{scct}, an enforceable overapproximation of \gls*{cct}, and \gls*{ss}, yield \gls*{mssscct}, which is the gold standard of security properties for secure applications.
@@
 Alas, code is not run in isolation, so a malicious attacker could supply code that interacts with \texttt{strncpy} and trigger a violation of either \gls*{ms} or \gls*{cct} by calling \texttt{strncpy} with an argument for \texttt{src} that points to uninitialised memory.
-This would, in turn, triggering a series of reads from uninitialised memory, which is an immediate \gls*{ms} violation with devastating real-world consequences~\cite{uninit-0,uninit-1,uninit-2,uninit-3,uninit-4}.
+This would, in turn, trigger a series of reads from uninitialised memory, which is an immediate \gls*{ms} violation with devastating real-world consequences~\cite{uninit-0,uninit-1,uninit-2,uninit-3,uninit-4}.
 % An easy fix to sidestep this is to use non-nullable pointers as parameters instead, e.g., as presented in Cyclone~\cite{jim2002cyclone}.
@@
 This paper introduces a framework for reasoning about the composition of secure and optimising compiler passes akin to those of \Cref{ex:strncpy} and it showcases the power of this framework by instantiating it on a multi-pass compilation chain.
-To this end, this paper first discusses how to compose security properties, such as \gls*{tms} and \gls*{sms} into \gls*{ms}, and then adding \gls*{scct} to the mix to obtain \gls*{msscct}.
-Then, this paper defines compiler composition and formalises that given two passes that securely preserve two (possibly distinct) properties, their composition securely preserves the composition of those properties.
+To this end, this paper first discusses how to compose security properties, such as \gls*{tms} and \gls*{sms} into \gls*{ms}, and then adds \gls*{scct} as well as \gls*{ss} into the mix to obtain \gls*{mssscct}.
+%Then, this paper defines compiler composition and formalises that given two passes that securely preserve two (possibly distinct) properties, their composition securely preserves the composition of those properties.
 The paper then defines several secure compiler passes, where each is either preserving a different security property (\gls*{tms}, \gls*{sms}, \gls*{scct}, \gls*{ss}) or performing a security-preserving optimisation, (e.g., applying \gls*{cf} or \gls*{dce}).
-Finally, this paper shows that composing these secure compiler passes into a multi-pass compilation chain results in the end-to-end preservation of \gls*{mssscct}.
+Finally, this paper shows that composing these secure compiler passes into a multi-pass compilation chain results in the end-to-end (robust) preservation of \gls*{mssscct}.
 Crucially, this paper derives the security of the multi-pass compiler from the composition of the security properties preserved by its individual passes.
@@
 
+\section{Formal Insights}\label{sec:formalities}
+
+This section provides some formal insights into the case study.
+
+\subsection{Preconditions for Compositionality}
+
+The composition theorems can only be used on compilers which have been proven secure in the sense that they robustly preserve a certain class of properties. 
+Traditional proof techniques apply, the usual strategy is to establish a backwards simulation between source components and their compiled target counterpart. 
+For the universally quantified target context, a technique known as backtranslation is well-established, which constructs a source-level context either by backtranslating the target-level context, amounting to a ,,decompilation'', or by constructing it using the target-level trace, so that the source-context emits parts of the whole execution trace that are related to the parts emitted by the target-level context.
+Up to backtranslations for $\ccdce$ and $\cccf$, all backtranslations in the case-study are trace based, while for $\ccdce$ and $\cccf$, they are context-based. 
+Since these optimisations have no significant semantic effect and source- and target-language are equal, the backtranslation amounts to an identity translation.
+
+\subsection{Compatibility of Secure Compiler Passes}
+
+\Cref{thm:rtpsim:sig} is always applicable, up to well-formedness of $\sim_1$ with respect to $\trg{\class[_2]}$.
+However, consider $\ccb$ and $\ccspec$.
+
 \section{Related Work\pages{2}}\label{sec:relwork}
@@
 This section discusses work on robust compilation (\Cref{subsec:relw:seccomprtp}) and on other secure compilation criteria (\Cref{subsec:relw:seccompcrit}).
-Since the case study of \Cref{sec:casestud:defs,sec:casestud:rtp} implements measures for preserving \gls*{ms} and \gls*{cct}, this section then presents relevant related work as well (\Cref{subsec:relw:msmechs,subsec:relw:cctmechs}).
+Since the case study of \Cref{sec:casestud:defs,sec:casestud:rtp} implements measures for preserving \gls*{ms}, \gls*{cct}, and \gls*{ss}, this section then presents relevant related work as well (\Cref{subsec:relw:msmechs,subsec:relw:cctmechs,subsec:relw:ssmechs}).
 
@@
 No existing work is concerned with composing robustly safe compilers.
-These works consider languages with different trace models and our technical setup can be adapted to that as long as security properties and their monitors are still defined on the same trace model.
+%Parts of these works consider languages with different trace models and our technical setup can handle this. 
 The work relating robust preservation with universal composability~\cite{patrignani2022universal} is closest to what this paper presents.
-The authors demonstrate a similar compositionality theorem to what is presented here (\Cref{sec:compcomp}) but use it in the context of protocols.
-They do not demonstrate the scalability of the approach.
-Moreover, they are missing the upper and lower compositions.
+The authors demonstrate a similar compositionality theorem to what is presented here (\Cref{sec:sequential}) but use it in the context of protocols.
+The work is missing the upper and lower compositions as well as the generality to support different trace models.
 % The authors demonstrate a similar compositionality theorem to what is presented here (\Cref{sec:compcomp}) as well as in an earlier version of this work~\cite{kruse2022csc}.
@@
 
+\subsection{Speculation Safety Mechanisms}\label{subsec:relw:ssmechs}
+
+With the advent of speculative execution in processors, it is possible to exploit leftover data in caches to identify secrets.
+SPECTECTOR~\cite{guarnieri2018spectector} models speculative execution formally and the present work has taken inspiration in the sematics of $\Lspec$ and in \Cref{def:trace:ss} from SPECTECTOR.
+A subsequent line of work~\cite{fabian2022automatic} shows how to modularly define and compose formal semantics that models different speculative execution variants. 
+Similar in spirit, recent work~\cite{fabian2024lift} demonstrates how to lift existing secure compilation proofs for security properties related to speculative execution to more powerful properties, i.e., properties defined for languages with a trace model that enables more vulnerabilities, such as combinations of different speculative execution exploits.
+To this end, they define a lifting framework and demonstrate how a secure compilation proof for a weaker property can be lifted to a stronger one. 
+Contrary to the compositionality framework presented in this paper, these works are only concerned with composing semantics of the respective programming languages, while in this paper, the secure compilers themselves are composed.
+
 \section{Conclusion\pages{$\sfrac{1}{2}$}}\label{sec:concl}
-This paper tackled the problem of understanding what kind of security properties does a secure compiler preserve, when said compiler is the combination of compiler passes that preserve possibly different security properties.
-% 
-For this, this paper first formalised security properties of interest and their composition.
-% 
-Then, it proved that composing secure compilers that preserve certain properties results in a secure compiler that preserves the composition of these properties.
+This paper tackles the problem of understanding what kind of security properties a secure compiler preserves, when said compiler is the combination of compiler passes that preserve possibly different security properties.
+% 
+%For this, this paper first formalised security properties of interest and their composition.
+% 
+The paper proves that composing secure compilers that preserve certain properties results in a secure compiler that preserves the composition of these properties.
 % While the presented security property composition relied on monitors that check only trace-properties, the composition of secure compilers does not make any restriction towards the kind of property involved in the composition.
@@
 % 
-Finally, this paper defines a multi-pass compiler and proves that it preserves \gls*{msscct}.
+Finally, this paper defines a multi-pass compiler and proves that it preserves \gls*{mssscct}.
 Crucially, this paper derives the security of the multi-pass compiler from the composition of the security properties preserved by its individual passes, which include security-preserving as well as optimisation passes.
 % For future work, it is interesting to look at more sophisticated optimisation passes that, e.g., reorder events that appear on traces.
-
-% This paper does a first step towards practical secure compilation chains by introducing a theoretical framework demonstrating that secure compilers compose.
-% The paper exercises this framework on a case-study that consists of an optimising compilation chain that is secure with respect to \gls*{msscct}.
-% \MP{
-% 	not an appropriate recap.
-% 	rewrite summarising the paper
-% }
-
-% In future work, it would be interesting to investigate whether it is possible to provide {\em secure compiler combinators}, similarly to how parser combinators work.
-% %To this end, it may be necessary to extend existing frameworks for multi-language semantics.
-% It is also interesting to extend our case-study and verify all results of it in Coq.
-% This way, it is possible to obtain an executable, secure compiler.
-% However, the formalisation effort for backtranslation proofs is known to be enormous\MP{cit + details}, so another research avenue may be to find better ways to (semi-)automate standard secure compilation proofs.
-% \MPin{
-% 	not particularly informative FW. i'd rather not have it as such
-% }
-
-%% The acknowledgments section is defined using the "acks" environment
-%% (and NOT an unnumbered section). This ensures the proper
-%% identification of the section in the article metadata, and the
-%% consistent spelling of the heading.
-
-% \begin{acks}
-%   We would like to thank the anonymous reviewers for their feedback.
-% 	This work was partially supported by a gift from
-		% the Italian Ministry of Education through funding for the Rita Levi Montalcini grant (call of 2019).
-% \end{acks}
-
-% \section*{Acknowledgements}\label{sec:acks}
-
-% \section*{Appendix}
-
-% The appendix contains insights on the structure of proofs for the theorems of \Cref{sec:casestud:rtp} (\Cref{sec:proof-sketch}).
-% Then it contains the proof of \Thmref{thm:ccab:rtp:ms} (\Cref{sec:proof1}) and of \Thmref{thm:ccall:rtp:msscct} (\Cref{sec:proof2}).
-
-% \section{Proving Robust Safety Property Preservation}\label{sec:proof-sketch}
-
-% We illustrate the proof of \Cref{thm:cca:rtp:tms} since the other secure compilation proofs of this paper follow the same approach.
-% Unfolding the theorem statement yields the following assumptions:  
-%   for any $\pi\in\lceil\tmssafe\rceil$\footnote{$\lceil\cdot\rceil$ lifts the property to a hyperproperty by applying the powerset operation~\cite{clarkson2008hyper}.}, 
-%   $\src{\trace}$, 
-%   $\src{\runtimetermvar}$, 
-%   and component $\src{\library_{comp}}$, 
-%   we have that $\rsat{\src{\library_{comp}}}{\pi}$ and 
-%   $\progstepto{\prog[\trg]{\trg{\library_{\ctx}}}{\cca(\src{\library_{\comp}})}}{\trg{\runtimetermvar}}{\trg{\trace}}$, where $\trg{\library_{\ctx}}$ is arbitrary.
-% The proof obligation is $\msfilterL{\trg{\trace}}\in\pi$, i.e., the specification trace associated to $\trg{\trace}$ satisfies the property $\pi$.
-% A way to show this is to relate trace $\trg{\trace}$ to some $\Ltms$ trace $\src{\trace}$ (which already satisfies the property as per the assumptions).
-% The assumptions already contain a target execution associated to this trace, so the task is to find an associated $\Ltms$ execution that yields $\trg{\trace}$.
-% The trace $\trg{\trace}$ is split into different parts, as commonly done in secure compilation works~\cite{korashy2021capableptrs,abate2018goodcomps}, where each part contains the events that either the context or the component does, but not both.
-% Because of this, all such trace segments are ,,well-bracketed'' in the sense that they start with either $\trg{\ev{Start}}$, $\trg{\ev{Call}\ \comptoctx\ foo\ \valueexpr}$, or $\trg{\ev{Ret}\ \comptoctx\ \valueexpr}$ and end with either $\trg{\ev{End}\ \valueexpr}$, $\trg{\ev{Ret}\ \ctxtocomp\ \valueexpr}$, or $\trg{\ev{Call}\ \ctxtocomp\ foo\ \valueexpr}$.
-% In the following, the former is referred to as a context segment, since these executions happen in $\trg{\library_{\ctx}}$, and the latter is referred to as a component segment, since these executions happen in $\cca(\src{\library_{\comp}})$.
-% \Cref{fig:proofdiag:rtp} visualises this division for a program execution with one call from context to component and how the target execution is related to a source execution.
-% In the figure, the green dashed lines encompass the component segments while the orange boxes contain the actual context switches from context to component or vice versa.
-% From a technical perspective, as typically done in compilation proofs, the proof requires some setup to maintain a relation between $\src{\Omega}$ and $\trg{\Omega}$ .
-% Two cross-language relations make this precise: (i) $\multimap_{\delta}$ relates states that are involved in a context segment, allowing the target execution to perform internal calls, and (ii) $\approx_{\delta}$ relates states that are involved in a component segment, where both states need to agree exactly, i.e., the memory and the control flow states are required to contain the same information.
-% The relations are indexed with $\delta$, which is an injective mapping from $\Ltms$ locations $\src{\loc}$ to $\Ltrg$ locations $\trg{\loc}$.
-% %Note that the relations $\multimap_{\delta}$ and $\approx_{\delta}$ swap when context switching.
-
-% So far, the paper explained how to relate an $\Ltms$ execution with a $\Ltrg$ execution.
-% The next question is therefore how to build the corresponding $\Ltms$ execution.
-% This is done using a standard secure compilation proof technique called trace-based backtranslation~\cite{abate2019jour,korashy2021capableptrs,patrignani2021rsc}, which can be used to build a context $\src{\library_{\ctx}}$ that behaves similar to $\trg{\library_{\ctx}}$.
-% % Since $\Ltms$ is statically typed, the context $\src{\library_{\ctx}}$ obtained from the backtranslation needs to be well-typed when linked with $\src{\library_{\comp}}$.
-% For context segments of the trace $\trg{\trace}$ it is also necessary to show that the execution behaves similarily, i.e., the context obtained from the backtranslation generates trace $\src{\trace}$.
-% For component segments of the trace, the relatedness of states and traces follows from a compiler correctness argument.
-% These two arguments yield the source execution $\progstepto{\src{\prog{\library_{\ctx}}{\library_{\comp}}}}{\src{\runtimetermvar}}{\src{\trace}}$.
-
-% The proof now works as follows.
-% Given that $\msfilterL{\trg{\trace}}=\msfilterLtms{\src{\trace}}$, the proof goal changes from $\msfilterL{\trg{\trace}}\in\pi$ to $\msfilterLtms{\src{\trace}}\in\pi$.
-% This follows by specializing the robust satisfaction assumption $\rsat{\src{\library_{\comp}}}{\pi}$ to use the context $\src{\library_{\ctx}}$, which is obtained from the backtranslation, and to use the source execution $\progstepto{\src{\prog{\library_{\ctx}}{\library_{\comp}}}}{\src{\runtimetermvar}}{\src{\trace}}$.
-
-% \begin{figure*}[!ht]
-%   \begin{center}
-%     \begin{tikzpicture}[state/.style={minimum height=1.5cm}]
-%       % relative horizontal/vertical distance between states
-%       \pgfmathsetmacro{\hdist}{0.95}
-%       \pgfmathsetmacro{\vdist}{1.35}
-%       \pgfmathsetmacro{\halfvdist}{0.725}
-
-%       % row of src states
-%       \node[state] (srcempty) {$\rtt{\src{\emptyset}}{\src{\lcall{main}{0}}}$};
-%       \foreach \s [remember=\s as \cur (initially empty)] in {1,w_1,p,w_2,2} {
-%         \node[state,right=\hdist of src\cur] (src\s) {$\rtt{\src{\Omega_{\s}}}{\src{\expr[_{\s}]}}$};
-%       }
-%       % row of trg states
-%       \node[state,below=\vdist of srcempty] (trgempty) {$\rtt{\trg{\emptyset}}{\trg{\lcall{main}{0}}}$};
-%       \foreach \s in {1,w_1,p,w_2,2} {
-%         \node[state,below=\vdist of src\s] (trg\s) {$\rtt{\trg{\Omega_{\s}}}{\trg{\expr[_{\s}]}}$};
-%       }
-%       %% illustrations
-%         % backtrans wrapper 1
-%         \draw[ultra thick,loosely dotted,Peach!50,rounded corners] (src1.north east) -- (srcw\string_1.north east)
-%           -- (trgw\string_1.south east) -- (trg1.south west) -- (src1.north west) -- cycle;
-%         % backtrans wrapper 2
-%         \draw[ultra thick,loosely dotted,Peach!50,rounded corners] (srcp.north east) -- (srcw\string_2.north east)
-%           -- (trgw\string_2.south east) -- (trgp.south west) -- (srcp.north west) -- cycle;
-%         % compiler correctness
-%         \draw[ultra thick,loosely dashed,Emerald!50,rounded corners] ($(srcw\string_1.north east)+(0,0.05)$) -- ($(srcp.north east)+(0,0.05)$)
-%           -- ($(trgw\string_2.north east)+(0.05,0)$) -- ($(trgw\string_2.south east)+(0.05,-0.05)$)
-%           -- ($(trg1.south west)+(-0.05,-0.05)$) -- ($(trg1.north west)+(-0.05,0)$) -- ($(srcw\string_1.north west)+(0,0.05)$) -- cycle;
-%       % state relations
-%       \path (srcempty) edge[very thick,draw=gray!25] node[pos=0.5,sloped,rotate=180,fill=white] {\scriptsize$\multimap_\emptyset$} (trgempty)
-%         (src1) edge[very thick,draw=gray!25] node[pos=0.5,sloped,rotate=180,fill=white] {\scriptsize$\multimap_{\delta_1}$} (trg1)
-%         (srcp) edge[very thick,draw=gray!25] node[pos=0.5,sloped,fill=white] {\scriptsize$\approx_{\delta_p}$} (trgp)
-%         (src2) edge[very thick,draw=gray!25] node[pos=0.5,sloped,rotate=180,fill=white] {\scriptsize$\multimap_{\delta_2}$} (trg2)
-%         (srcw\string_1) edge[very thick,draw=gray!25] node[pos=0.5,sloped,fill=white] {\scriptsize$\approx_{\delta_{w_1}}$} (trgw\string_1)
-%         (srcw\string_2) edge[very thick,draw=gray!25] node[pos=0.5,sloped,rotate=180,fill=white] {\scriptsize$\multimap_{\delta_{w_2}}$} (trgw\string_2)
-%         % diagonals
-%         (srcw\string_1) edge[thick,draw=gray!25] node[pos=0.16,sloped,rotate=180,fill=white] {\scriptsize$\approx_{\delta_{w_1}}$} (trg1)
-%         (src1) edge[thick,draw=gray!25] node[pos=0.16,sloped,rotate=180,fill=white] {\scriptsize$\multimap_{\delta_1}$} (trgw\string_1)
-%         (srcp) edge[thick,draw=gray!25] node[pos=0.2,sloped,fill=white] {\scriptsize$\approx_{\delta_p}$} (trgw\string_2)
-%         (trgp) edge[thick,draw=gray!25] node[pos=0.2,sloped,fill=white] {\scriptsize$\multimap_{\delta_{w_2}}$} (srcw\string_2)
-%         ;
-%       %\drawpolygon src1,srcw\string_1,trgw\string_1,trg1;
-%       %\drawpolygon srcp,srcw\string_2,trgw\string_2,trgp;
-%       %\node[font=\tiny,align=center,above=0.2 of srcw1srcp] (wrapper) {Backtranslation\\Wrapper};
-%       %\path[->,draw] (wrapper) -- (srcw\string_1);
-%       %\path[->,draw] (wrapper) -- (srcp);
-%       % steps
-%       \path[very thick,color=\stlccol] (srcempty) edge[draw=none] node {\ $\xrightarrow{\trace[_1]}{}{\kern-3.5pt}_{\text{ctx}}^*$} (src1)
-%         (src1) edge[draw=none] node {\ $\xrightarrow{\trace[_c]}{}{\kern-3.5pt}_{\text{ctx}}$} (srcw\string_1)
-%         (srcw\string_1) edge[draw=none] node {\ $\xrightarrow{\trace[_p]}{}{\kern-3.5pt}_{\text{ctx}}^*$} (srcp)
-%         (srcp) edge[draw=none] node[fill=white,inner sep=0,outer sep=0] {\ $\xrightarrow{\trace[_r]}{}{\kern-3.5pt}_{\text{ctx}}$} (srcw\string_2)
-%         (srcw\string_2) edge[draw=none] node {\ $\xrightarrow{\trace[_2]}{}{\kern-3.5pt}_{\text{ctx}}^*$} (src2)
-%         ;
-%       \path[very thick,color=\ulccol] (trgempty) edge[draw=none] node[fill=white,inner sep=0,outer sep=0] {\ $\xrightarrow{\phantom{\trace[_1]}}{}{\kern-3.5pt}_{\text{ctx}}^*$} (trg1)
-%         (trg1) edge[draw=none] node {\ $\xrightarrow{\phantom{\trace[_p]}}{}{\kern-3.5pt}_{\text{ctx}}^*$} (trgw\string_1)
-%         (trgw\string_1) edge[draw=none] node[fill=white,inner sep=0,outer sep=0] {\ $\xrightarrow{\phantom{\trace[_p]}}{}{\kern-3.5pt}_{\text{ctx}}^*$} (trgp)
-%         (trgp) edge[very thick,draw=none] node {\ $\xrightarrow{\phantom{\trace[_p]}}{}{\kern-3.5pt}_{\text{ctx}}^*$} (trgw\string_2)
-%         (trgw\string_2) edge[draw=none] node {\ $\xrightarrow{\phantom{\trace[_2]}}{}{\kern-3.5pt}_{\text{ctx}}^*$} (trg2)
-%         ;
-%       % legend
-% %     \node[align=left,below right=0.3 and 0.3 of trgempty,font=\tiny] (legend) {%
-% %       ${\src{\trace[_c]}}\cong{\backt{\Ltms}{\Ltrg}(\trg{Call\ ?\ foo\ \valueexpr})}$\\%
-% %       ${\src{\trace[_r]}}\cong{\backt{\Ltms}{\Ltrg}(\trg{Ret\ !\ foo\ \valueexpr})}$
-% %       };
-%       \node[below=0.5 of trgempty,draw=none] (legend) {};
-%       \draw[ultra thick,loosely dotted,Peach!50,rounded corners] ($(legend.north east)+(0.5,-0.4)$) -- ($(legend.north east)+(1,-0.4)$);
-%       \node at ($(legend.north east)+(1.5,-0.4)$) (legendwrapper) {};
-%       \node[right of=legendwrapper] {\tiny Backtranslation Wrapper};
-%       \draw[ultra thick,loosely dashed,Emerald!50,rounded corners] ($(legend.south east)+(0.5,0.4)$) -- ($(legend.south east)+(1,0.4)$);
-%       \node at ($(legend.south east)+(1.5,0.4)$) (legendcorrectness) {};
-%       \node[right of=legendcorrectness] {\tiny Compiler Correctness};
-%     \end{tikzpicture}
-%     \caption{Proof diagram for \Cref{thm:cca:rtp:tms} depicting the general structure of robust preservation proofs. %
-%       Nodes in the graph represent runtime states. %
-%       Vertical lines indicate cross language relations, while horizontal ones are execution steps. %
-%       The green dashed trapezoid encompasses the component segment, while the orange dotted rectangles entail the context switches. %
-%       $\Ltrg$ traces are omitted for readability. %
-%       $\Ltms$ trace segments $\src{\trace[_{c}]}$ and $\src{\trace[_{r}]}$ describe the events that happen at the boundaries, i.e., during a context switch.
-%       $\src{\trace_{p}}$ is the behavior of the component and the traces $\src{\trace[_{1}]}$ and $\src{\trace[_{2}]}$ describe the context.
-%     }\label{fig:proofdiag:rtp}
-%   \end{center}
-% \end{figure*}
-
-% \section{Proof of \Cref{thm:ccab:rtp:ms}} \label{sec:proof1}
-%   From \Thmref{thm:cca:rtp:tms} it follows that for any $\Ltms$ program $\src{\progvar}$, it compiles to an $\Ltrg$ program $\trg{\progvar}$ that robustly satisfies \gls*{tms}.
-%   Note that $\src{\progvar}$ robustly satisfies \gls*{tms} by the properties of the typesystem of $\Ltms$.
-%   Then, \Thmref{thm:ccb:rtp:sms} demonstrates that, assuming $\trg{\progvar}$ robustly satisfies \gls*{sms}, the program $\trg{\progvar}$ compiles to an $\Lms$ program $\irl{p}$ that also robustly satisfies \gls*{sms}.
-%   From \Thmref{thm:ccab:rtp:ms} it follows that $\src{p}$ compiles to $\irl{p}$ that robustly satisfies \gls*{ms}, since \gls*{ms} is the intersection of \gls*{tms} and \gls*{sms}.
-
-% \section{Proof of \Cref{thm:ccall:rtp:msscct}} \label{sec:proof2}
-% From \Thmref{thm:ccab:rtp:ms}, we have that any $\Ltms$ program $\src{\progvar}$ compiles into a $\Lms$ program $\irl{\progvar}$ that robustly satisfies \gls*{ms}.
-% Then, from \Thmref{thm:cccfccdce:rtp:ms} we have that $\irl{\progvar}$ gets optimised to a program $\irl{\progvar[']}$ that is also \gls*{ms}, where the order of optimisations does not matter for $\irl{\progvar[']}$ to be \gls*{ms}.
-% Assuming $\irl{\progvar[']}$ robustly satisfies \gls*{scct}, by \Thmref{thm:ccscct:rtp:scct} it compiles to an $\Lscct$ program $\obj{\progvar}$ that robustly satisfies \gls*{scct} as well.
-% Finally, from \Thmref{thm:rtp} it follows that, given $\src{\progvar}$ robustly satisfies \gls*{scct} and \gls*{ms}, $\obj{\progvar}$ also robustly satisfies \gls*{scct} and \gls*{ms}.
-
-% \section{Auxiliary definitions for \Cref{subsec:app}}
-% \begin{definition}[Existential Image of Properties]\label{def:eximg:prop}
-%   Given a cross-language trace relation $\sim$,
-%   $$
-%   \tau_\sim(\src{\pi}) \isdef \left\{
-%     \trg{\trace} \mid \exists\src{\trace}.\src{\trace}\sim\trg{\trace} \text{ and }
-%     \src{\trace}\in\src{\pi}
-%   \right\}
-%   $$
-% \end{definition}
-
-% \Thmref{def:eximg:prop} can be lifted to hyperproperties/classes as follows:
-% \begin{definition}[Existential Image of Hyperproperties]\label{def:eximg:hprop}
-%   Given a cross-language trace relation $\sim$,
-%   $$
-%   \tilde{\tau}_\sim(\src{\Pi}) \isdef \left\{
-%     \trg{\pi} \mid \exists\src{\pi}.\src{\pi}\in\src{\Pi} \text{ and }
-%     \trg{\pi}=\tau_\sim\left(\src{\pi}\right)
-%   \right\}
-%   $$
-% \end{definition}
-% \bul{A cross-language relation $\sim$ is well-formed with respect to a source-level (hyper-)property $\src{\Pi}$} iff \rul{the existential images preserve membership}.
-% \begin{definition}[Well-formedness of Cross-Language Trace Relations for a Properties $\src{\Pi}$]\label{def:wf:tracerel}
-%   $$
-%   \text{\bul{$\wfc{\sim}{\src{\Pi}}$}} \isdef \text{\rul{$\forall \src{\pi}\in\src{\Pi}, \tau_\sim(\src{\pi})\in\tilde{\tau}_\sim(\src{\Pi})$}}
-%   $$
-% \end{definition}
-% \begin{lemma}[Well-Formedness Equivalence for Classes]\label{lem:wf:equiv:wfc}
-%   $\;\;$
-%   $ \wf{\sim}{\src{\class}} \Leftrightarrow\ \wfc{\sim}{\src{\class}} $\Coqed
-% \end{lemma}
 

### 2024-05-23 csf2/main.tex
add some rules and text

@@
 \]
+\begin{center}
+\footnotesize
+  \typerule{$e-\obj{wrdoit}-\text{off}$}{
+  }{
+    \pstepto{\rtt{\obj{\cfstate;\sandboxtag;m;\memstate}}{\obj{\lwrdoit{OFF}}}}{\rtt{\obj{\cfstate;\sandboxtag;OFF;\memstate}}{\obj{0^{\securitytag}}}}{\obj{\emptyevent}}
+  }{e-wrdoit-off}
+\newcommand{\expreval}[5]{{#1}\triangleright\xspace {#2}\xrightarrow{#5}\ {#3}\triangleright\xspace {#4}\xspace}
+\newcommand{\exprevalo}[5]{\expreval{\obj{#1}}{\obj{#2}}{\obj{#3}}{\obj{#4}}{\obj{#5}}}
+  \typerule{$e-\obj{get}-\in-$noleak}{
+    \obj{\memstate}=\obj{H^{\ctx};H^{\comp};\Delta_1,x\mapsto(\loc;\sandboxtag;\poison;m),\Delta_2} &
+    \obj{\loc}+\obj{n}\in\text{dom }\obj{H^{\sandboxtag}} \\
+    \obj{\sigma''}=\obj{\sigma\sqcap\sigma'}
+  }{
+    \exprevalo{\cfstate;\sandboxtag';m+1;\memstate}{x^\sigma[n^{\sigma'}]}{\cfstate;\sandboxtag';m+1;\memstate}{(H^{\sandboxtag}(\loc+n))^{\sigma''}}{{iGet}\ \loc\ n;\sandboxtag;\sigma''}
+  }{to-e-get-in-noleak}
+\end{center}
 
@@
 \]
+
+\begin{center}
+  \footnotesize
+\newcommand{\expreval}[5]{{#1}\triangleright\xspace {#2}\xrightarrow{#5}\ {#3}\triangleright\xspace {#4}\xspace}
+\newcommand{\exprevald}[5]{\expreval{\ird{#1}}{\ird{#2}}{\ird{#3}}{\ird{#4}}{\ird{#5}}}
+
+  %
+  \typerule{$e-\ird{ifz}-\ird{true}-\text{spec}$}{
+    \dots \text{todo}
+  }{
+    \pstepto{\rtt{\ird{\cfstate;\sandboxtag;0;\memstate}}{\ird{\lifz{0^{\securitytag}}{\expr[_1]}{\expr[_2]}}}}{\begin{array}{c}\rtt{\ird{\cfstate;\sandboxtag;0;\memstate}}{\ird{\expr[_1]}}\\\rtt{\ird{\cfstate;\sandboxtag;0;\memstate}}{\ird{\expr[_2]}}\end{array}}{\ird{\ev{Branch}\ 0;\sandboxtag;\securitytag}\cdot\ird{\ev{Spec};\sandboxtag;\securitytag}}
+  }{e-ifz-true-leak}
+\end{center}
 
@@
 However, consider $\ccb$ and $\ccspec$.
+Even though swapping these compilers in the compilation chain is syntactically impossible, let us pretend that it is possible by just extending $\ccb$ for the constructs present in $\Lscct$ that are not present in $\Ltrg$.
+Clearly, as evidenced in the case study earlier, $\ccb\cdot\ccspec$ is an acceptable composition that yields an interesting class of security properties, i.e., $\sigma_{\sim_{\irl{ms}}\bullet\sim_{\ird{\mathghost}}}\left(\mssafe\cap\scctsafe\cap\sssafe\right)$.
+However, swapping them $\ccspec\cdot\ccb$ means that $\ccb$ inserts new branches into the code that are not necessarily protected by a speculative execution barrier! 
+It is crucial to ensure that the resulting class is meaningful, much similar to ensuring that a definition models what the formal methods expert intends to model. 
+For the $\ccspec\cdot\ccb$ case, the class is $\sigma_{\sim_{\ird{\mathghost}}\bullet\sim_{\irl{ms}}}\left(\mssafe\cap\scctsafe\cap\sssafe\right)$
+Since $\ev{Use\ \loc\ n}\sim_{\irl{ms}}\ev{Branch n}\cdot\ev{Spec}\cdots$, where $\cdots$ does not contain a $\ird{\ev{Barrier}}$ event, the resulting class of properties is only a subset of the intended one.
+Because of this, it is crucial to analyze the precise shape of the class of properties after mapping a target-level property to a source-level property\footnote{Dito for vice-versa.}.
 

### 2024-05-23 csf2/main.tex
nit
change fig3 and hyperref

@@
 
+\hypersetup{
+    colorlinks,
+    linkcolor={red!50!black},
+    citecolor={blue!50!black},
+    urlcolor={blue!80!black}
+}
 %\renewcommand\UrlFont{\color{blue}\rmfamily}
@@
   % Unfortunately, it is an open question whether secure compilers can be engineered in a modular way.
-  % Ideally, compiler engineers develop two different compilation passes, one that ensures memory safety and another one that is security preserving with respect to cryptographic constant time.
+  % Ideally, compiler engineers develop two different compilation passes, one that ensures memory safety and another one that is security preserving w.r.t.
+cryptographic constant time.
 %, because this style of engineering is well-established for its simplifying divide-and-conquer approach as well as enhanced reusability.
   % \MKin{object+findings}
-  % This paper answers that question positively: The composition of compilers that are secure with respect to some security-properties yields the intersection of these properties.
+  % This paper answers that question positively: The composition of compilers that are secure w.r.t. some security-properties yields the intersection of these properties.
   % \MKin{object}
@@
   % \MKin{conclusion}
-  % The compositionality results enable modular proofs for the presented, secure compiler. % that is secure with respect to both memory safety and cryptographic constant time.
+  % The compositionality results enable modular proofs for the presented, secure compiler. % that is secure w.r.t. both memory safety and cryptographic constant time.
   % Some results of this work are implemented in the Coq proof assistant.
@@
 % Because of this, \gls*{ms} is a requirement for programs to attain an even stronger security property, namely \gls*{cct}.
-% Programs that attain \gls*{cct} are secure with respect to timing attacks, and this is the gold standard for secure applications, including cryptographic ones, since timing attacks can expose secrets via a simple performance analysis~\cite{kocher1996timing}.
+% Programs that attain \gls*{cct} are secure w.r.t.
+timing attacks, and this is the gold standard for secure applications, including cryptographic ones, since timing attacks can expose secrets via a simple performance analysis~\cite{kocher1996timing}.
 % Timing attacks have become more and more relevant recently, especially given the era of cloud-computing~\cite{aviram2010cloudtime,kumar2019cloudsecsurvey}, where programs share the same hardware but are supposed to run in a sandboxed environment that does not leak secrets.
@@
         This paper proves that starting from two compilers that preserve two (possibly distinct) properties, their composition preserves roughly the intersection of those properties.
-        % For example, a compiler that is secure with respect to \gls*{tms} can be composed with a compiler that is secure with respect to \gls*{sms}, resulting in a compiler that is secure with respect to \gls*{ms}.
+        % For example, a compiler that is secure w.r.t. \gls*{tms} can be composed with a compiler that is secure w.r.t. \gls*{sms}, resulting in a compiler that is secure w.r.t. \gls*{ms}.
         % 
@@
 As an example, consider a trace describing an interaction with a memory where the deallocation of an address $\loc$ precedes a read at that address in memory: $\ev{Dealloc\ \loc}\cdot\ev{Read\ \loc\ 1729}\cdot\dots$.
-This program behaviour is insecure with respect to a canonical notion of (temporal) memory safety dictating no use-after-frees of pointers~\cite{nagarakatte2010cets,azevedo2018meaningsofms}, because it reads from a memory location that was freed already.
+This program behaviour is insecure w.r.t.
+a canonical notion of (temporal) memory safety dictating no use-after-frees of pointers~\cite{nagarakatte2010cets,azevedo2018meaningsofms}, because it reads from a memory location that was freed already.
 The previous finite trace prefix is enough to decide that the trace does not satisfy \gls*{tms} and there is no way to append events to this prefix which would result in the trace being admissible.
@@
 
-\bul{A cross-language relation $\sim$ is well-formed with respect to a class of target-level properties $\trg{\class}$} iff \rul{the universal image preserves set membership}.
+\bul{A cross-language relation $\sim$ is well-formed w.r.t. a class of target-level properties $\trg{\class}$} iff \rul{the universal image preserves set membership}.
 \begin{definition}[Well-formedness of Trace Relations for a Class $\trg{\class}$]\label{def:wfc:sig:tracerel}
@@
 Let \bul{$\cc{\src{L}}{\trg{L}}$ robustly preserve the class $\sigma_{\sim_2}\left(\trg{\class[_1]}\right)$ under $\sim_1$} and let \rul{$\cc{\trg{L}}{\obj{L}}$ robustly preserve the class $\trg{\class[_2]}$ under $\sim_2$}.
-Then, \iul{when the cross-language relation $\sim_2$ is well-formed with respect to class $\trg{\class[_1]}$}, it follows that \oul{the composed compiler $\cc{\src{L}}{\trg{L}}\circ\cc{\trg{L}}{\obj{L}}$ robustly preserves the intersection of classes $\trg{\class[_1]}\cap\trg{\class[_2]}$ under $\sim_1\bullet\sim_2$}.
+Then, \iul{when the cross-language relation $\sim_2$ is well-formed w.r.t.
+class $\trg{\class[_1]}$}, it follows that \oul{the composed compiler $\cc{\src{L}}{\trg{L}}\circ\cc{\trg{L}}{\obj{L}}$ robustly preserves the intersection of classes $\trg{\class[_1]}\cap\trg{\class[_2]}$ under $\sim_1\bullet\sim_2$}.
 % 
 %For space reasons, \Thmref{thm:urtp} and \Thmref{thm:lrtp} are not presented here, but can be found in the technical appendix ($\CoqSymbol$).
-% Note that it is not necessary to require $\sim_2$ to be well-formed with respect to $\src{\class[_1]}$, since the 
+% Note that it is not necessary to require $\sim_2$ to be well-formed w.r.t. $\src{\class[_1]}$, since the 
 % \MPin{why}
-\begin{theorem}[Sequential Composition of Secure Compilers with respect to $\sigma$]\label{thm:rtpsim:sig}
+\begin{theorem}[Sequential Composition of Secure Compilers w.r.t. $\sigma$]\label{thm:rtpsim:sig}
   If \bul{$\rtpsig{\cc{\src{L}}{\trg{L}}}{\tilde{\sigma}_{\sim_2}\left(\trg{\class[_{1}]}\right)}{\sim_1}$}, \rul{$\rtpsig{\cc{\trg{L}}{\obj{L}}}{\trg{\class[_2]}}{\sim_2}$}, and \iul{$\wfcsig{\sim_1}{\trg{\class[_2]}}$}, then \oul{$\rtpsig{\cc{\src{L}}{\trg{L}}\circ\cc{\trg{L}}{\obj{L}}}{\trg{\class[_{1}]}\cap\trg{\class[_{2}]}}{\sim_1\bullet\sim_2}$}.
@@
 \Cref{corr:rtp} provides an easy to use theorem to compose secure compilers without any caveats. 
-However, while \Cref{thm:rtpsim:sig} explicitly requires that $\sim_1$ is well-formed with respect to $\trg{\class[_2]}$, if $\sim_2$ is not well-formed with respect to $\trg{\class[_1]}$, care must be taken.
+However, while \Cref{thm:rtpsim:sig} explicitly requires that $\sim_1$ is well-formed w.r.t. $\trg{\class[_2]}$, if $\sim_2$ is not well-formed w.r.t. $\trg{\class[_1]}$, care must be taken. 
 Worst-case is that the resulting class $\sigma_{\sim_1\bullet\sim_2}\left(\trg{\class[_1]}\cap\trg{\class[_2]}\right)$ becomes empty, even though $\trg{\class[_1]}\cap\trg{\class[_2]}$ is not necessarily empty.
@@
 Compiler engineers typically try to find an order of optimisations that yields well-optimised programs for either code size~\cite{cooper1999geneticphases} or performance~\cite{kulkarni2006exhaustivephase}.
-\Cref{corr:swappable} justifies that any such order of compilation passes is valid with respect to security.
+\Cref{corr:swappable} justifies that any such order of compilation passes is valid w.r.t. security.
 So, given two compilation passes \bul{$\cc[_{1}]{\trg{L}}{\trg{L}}$, $\cc[_{2}]{\trg{L}}{\trg{L}}$, both robustly preserving class $\class[_{1}]$ or $\class[_{2}]$, respectively}, \oul{for any order of their composition the composed compiler robustly preserves the intersection of $\class[_{1}]$ and $\class[_{2}]$}.
@@
 % \MPin{
-% 	later on, say that optimisation passes remove elements from traces, so they should be fine with respect to the result with $\tau$.
+% 	later on, say that optimisation passes remove elements from traces, so they should be fine w.r.t. the result with $\tau$.
 % 	also, yes, if $\tau$ adds we have a problem, if it removes we don't
@@
 Besides sequential composition, there are two other compositions, namely an {\em upper}, i.e., a compiler that takes multiple inputs and yields one output, and a {\em lower} composition, i.e., a compiler that takes one input and yields multiple outputs.
-\bul{Define the upper composition $\cc{\src{L}+\obj{L}}{\trg{L}}$} as follows:
-Given a program \texttt{p}, its compiled counterpart is obtained by \rul{plugging \texttt{p} into $\cc{\src{L}}{\trg{L}}$ if $\texttt{p}\in\src{L}$} or by \oul{plugging \texttt{p} into $\cc{\obj{L}}{\trg{L}}$ if $\texttt{p}\in\obj{L}$}.
+{Define the upper composition $\cc{\src{L}+\obj{L}}{\trg{L}}$} as follows:
+Given a program \texttt{p}, its compiled counterpart is obtained by {plugging \texttt{p} into $\cc{\src{L}}{\trg{L}}$ if $\texttt{p}\in\src{L}$} or by {plugging \texttt{p} into $\cc{\obj{L}}{\trg{L}}$ if $\texttt{p}\in\obj{L}$}.
 \begin{definition}[Upper Composition]
-  $$\text{\bul{$\cc{\src{L}+\obj{L}}{\trg{L}}$}}\isdef
+  $$\text{{$\cc{\src{L}+\obj{L}}{\trg{L}}$}}\isdef
   % Given \texttt{p}, yield
@@
   \left\{\mbox{\begin{tabular}{c}
-    {\rul{if $\texttt{p}\in\src{L}$, then $\cc{\src{L}}{\trg{L}}(\texttt{p})$}} \\
-    \mbox{\oul{if $\texttt{p}\in\obj{L}$, then $\cc{\obj{L}}{\trg{L}}(\texttt{p})$}} \\
+    {{if $\texttt{p}\in\src{L}$, then $\cc{\src{L}}{\trg{L}}(\texttt{p})$}} \\
+    \mbox{{if $\texttt{p}\in\obj{L}$, then $\cc{\obj{L}}{\trg{L}}(\texttt{p})$}} \\
   \end{tabular}}\right.$$
@@
 Now, the following theorem tells us what happens if these are secure:
-Given \bul{$\cc{\src{L}}{\trg{L}}$ robustly preserves $\class[_{1}]$} and \rul{$\cc{\obj{L}}{\trg{L}}$ robustly preserves $\class[_{2}]$}, it follows that \oul{their upper composition $\cc{\src{L}+\obj{L}}{\trg{L}}$ robustly preserves the intersection of classes $\class[_{1}]$ and $\class[_{2}]$}.
+Given {$\cc{\src{L}}{\trg{L}}$ robustly preserves $\class[_{1}]$} and {$\cc{\obj{L}}{\trg{L}}$ robustly preserves $\class[_{2}]$}, it follows that {their upper composition $\cc{\src{L}+\obj{L}}{\trg{L}}$ robustly preserves the intersection of classes $\class[_{1}]$ and $\class[_{2}]$}.
 
 \begin{theorem}[Upper Composition of Secure Compilers]\label{thm:urtp}
-  If \bul{$\rtp{\cc{\src{L}}{\trg{L}}}{\class[_{1}]}$} and \rul{$\rtp{\cc{\obj{L}}{\trg{L}}}{\class[_{2}]}$}, then \oul{$\rtp{\cc{\src{L}+\obj{L}}{\trg{L}}}{\class[_{1}]\cap\class[_{2}]}$}. %\Coqed
+  If {$\rtp{\cc{\src{L}}{\trg{L}}}{\class[_{1}]}$} and {$\rtp{\cc{\obj{L}}{\trg{L}}}{\class[_{2}]}$}, then {$\rtp{\cc{\src{L}+\obj{L}}{\trg{L}}}{\class[_{1}]\cap\class[_{2}]}$}. %\Coqed
 \end{theorem}
@@
 Dually, the {\em lower} composition is concerned about compilers that accept the same source but yield different target languages. %and the same theoretical results apply.
-\bul{Define the lower composition $\cc{\src{L}}{\trg{L}+\obj{L}}$} as follows:
-Given a program $\src{p}$, its compiled counterpart is obtained by \rul{plugging $\src{p}$ into $\cc{\src{L}}{\trg{L}}$} or by \oul{plugging \texttt{p} into $\cc{\src{L}}{\obj{L}}$}, respectively, \pul{based on the internal decision}.
+{Define the lower composition $\cc{\src{L}}{\trg{L}+\obj{L}}$} as follows:
+Given a program $\src{p}$, its compiled counterpart is obtained by {plugging $\src{p}$ into $\cc{\src{L}}{\trg{L}}$} or by {plugging \texttt{p} into $\cc{\src{L}}{\obj{L}}$}, respectively, {based on the internal decision}.
 \begin{definition}[Lower Composition]
-  $$\text{\bul{$\cc{\src{L}}{\trg{L}+\obj{L}}$}}\isdef
+  $$\text{{$\cc{\src{L}}{\trg{L}+\obj{L}}$}}\isdef
   % Given $L$ and $\src{p}$, yield
@@
   \left\{\mbox{\begin{tabular}{c}
-    {\pul{if $L=\trg{L}$, then} \rul{$\cc{\src{L}}{\trg{L}}(\src{p})$}} \\
-    \mbox{\pul{if $L=\obj{L}$, then} \oul{$\cc{\src{L}}{\obj{L}}(\src{p})$}} \\
+    {{if $L=\trg{L}$, then} {$\cc{\src{L}}{\trg{L}}(\src{p})$}} \\
+    \mbox{{if $L=\obj{L}$, then} {$\cc{\src{L}}{\obj{L}}(\src{p})$}} \\
   \end{tabular}}\right.$$
@@
 % The following theorem demonstrates what happens if the involved compilers are secure:
-Given \bul{$\cc{\src{L}}{\trg{L}}$ robustly preserves $\class[_{1}]$} and \rul{$\cc{\src{L}}{\obj{L}}$ robustly preserves $\class[_{2}]$}, it follows that \oul{their lower composition $\cc{\src{L}}{\trg{L}+\obj{L}}$ robustly preserves the intersection of classes $\class[_{1}]$ and $\class[_{2}]$}.
+Given {$\cc{\src{L}}{\trg{L}}$ robustly preserves $\class[_{1}]$} and {$\cc{\src{L}}{\obj{L}}$ robustly preserves $\class[_{2}]$}, it follows that {their lower composition $\cc{\src{L}}{\trg{L}+\obj{L}}$ robustly preserves the intersection of classes $\class[_{1}]$ and $\class[_{2}]$}.
 
 \begin{theorem}[Lower Composition of Secure Compilers]\label{thm:lrtp}
-  If \bul{$\rtp{\cc{\src{L}}{\trg{L}}}{\class[_{1}]}$} and \rul{$\rtp{\cc{\obj{L}}{\trg{L}}}{\class[_{2}]}$}, then \oul{$\rtp{\cc{\src{L}}{\trg{L}+\obj{L}}}{\class[_{1}]\cap\class[_{2}]}$}. % \Coqed
+  If {$\rtp{\cc{\src{L}}{\trg{L}}}{\class[_{1}]}$} and {$\rtp{\cc{\obj{L}}{\trg{L}}}{\class[_{2}]}$}, then {$\rtp{\cc{\src{L}}{\trg{L}+\obj{L}}}{\class[_{1}]\cap\class[_{2}]}$}. % \Coqed
 \end{theorem}
@@
 
-\begin{figure*}[!h]
+\begin{figure}[!htb]
  \vspace{-2em}
@@
     \node (S) {$\src{L_{\tmssafe}}$};
-    \node[right=1.5 of S] (T) {$\trg{L}$};
-    \node[right=1.5 of T] (M) {$\irl{L_{\mssafe}}$};
-    \node[below right=0.5 and 1.0 of M] (D0) {$\irl{L_{\mssafe}}$};
-    \node[above right=0.5 and 1.0 of M] (C0) {$\irl{L_{\mssafe}}$};
+    \node[right=2.0 of S] (T) {$\trg{L}$};
+    \node[right=2.0 of T] (M) {$\irl{L_{\mssafe}}$};
+    \node[below right=1.5 and 1.0 of M] (D0) {$\irl{L_{\mssafe}}$};
+    \node[above right=1.5 and 1.0 of M] (C0) {$\irl{L_{\mssafe}}$};
     \node[right=3.0 of M] (E) {$\irl{L_{\mssafe}}$};
-    \node[right=1.5 of E] (O) {$\obj{L_{\scctsafe}}$};
-    \node[right=1.5 of O] (Z) {$\ird{L_{\mathghost}}$};
-
-    \draw[->] (S) to[sloped] node[align=center,font=\scriptsize] (tmsedge) {\gls*{tms}\\ \Cref{thm:cca:rtp:tms}} (T);
-    \draw[->] (T) to[sloped] node[align=center,font=\scriptsize] {\gls*{sms}\\ \Cref{thm:ccb:rtp:sms}} (M);
-    \draw[->] (M) to[sloped] node[align=center,font=\scriptsize] {\gls*{dce}\\ \Cref{thm:ccdce:rtp:ms}} (D0);
-    \draw[->] (M) to[sloped] node[align=center,font=\scriptsize] {\gls*{cf}\\ \Cref{thm:cccf:rtp:ms}} (C0);
-    \draw[->] (D0) to[sloped] node[align=center,font=\scriptsize] {\gls*{cf}\\ \Cref{thm:cccf:rtp:ms}} (E);
-    \draw[->] (C0) to[sloped] node[align=center,font=\scriptsize] {\gls*{dce}\\ \Cref{thm:ccdce:rtp:ms}} (E);
-    \draw[->] (E) to[sloped,above] node[align=center,font=\scriptsize] {\gls*{scct}\\ \Cref{thm:ccscct:rtp:scct}} (O);
-    \draw[->] (O) to[sloped,above] node[align=center,font=\scriptsize] {\gls*{ss}\\ \Cref{thm:ccspec:rtp:spec}} (Z);
+    \node[right=2.0 of E] (O) {$\obj{L_{\scctsafe}}$};
+    \node[right=2.0 of O] (Z) {$\ird{L_{\mathghost}}$};
+
+    \draw[->] (S) to[sloped] node[align=center] (tmsedge) {\gls*{tms}\\ \Cref{thm:cca:rtp:tms}} (T);
+    \draw[->] (T) to[sloped] node[align=center] {\gls*{sms}\\ \Cref{thm:ccb:rtp:sms}} (M);
+    \draw[->] (M) to[sloped] node[align=center] {\gls*{dce}\\ \Cref{thm:ccdce:rtp:ms}} (D0);
+    \draw[->] (M) to[sloped] node[align=center] {\gls*{cf}\\ \Cref{thm:cccf:rtp:ms}} (C0);
+    \draw[->] (D0) to[sloped] node[align=center] {\gls*{cf}\\ \Cref{thm:cccf:rtp:ms}} (E);
+    \draw[->] (C0) to[sloped] node[align=center] {\gls*{dce}\\ \Cref{thm:ccdce:rtp:ms}} (E);
+    \draw[->] (E) to[sloped,above] node[align=center] {\gls*{scct}\\ \Cref{thm:ccscct:rtp:scct}} (O);
+    \draw[->] (O) to[sloped,above] node[align=center] {\gls*{ss}\\ \Cref{thm:ccspec:rtp:spec}} (Z);
 
     % Sections
-    \node[above=1.0 of tmsedge] (sectms) {{\scriptsize\Cref{subsec:cs:tms}}};
-    \node[right=0.5 of sectms] (secsms) {{\scriptsize\Cref{subsec:cs:ms}}};
-    \node[right=1.5 of secsms] (secopts) {{\scriptsize\Cref{subsec:cs:opts}}};
-    \node[right=1.25 of secopts] (secscct) {{\scriptsize\Cref{subsec:cs:scct}}};
-    \node[right=1.00 of secscct] (secss) {{\scriptsize\Cref{subsec:cs:ss}}};
-
-    \draw[thick,dotted,->] (S) to[bend right=33,sloped] node[align=center,font=\scriptsize] {\gls*{ms}\\ \Cref{thm:ccab:rtp:ms}} (M);
-    \draw[thick,dotted,->] (M) to[bend right=0,sloped] node[align=center,font=\scriptsize] {\gls*{ms}\\ \Cref{thm:cccfccdce:rtp:ms}} (E);
-    \draw[thick,dotted,->] (S) to[out=-45,in=198,sloped] node[align=center,font=\scriptsize] {\gls*{ms}+\gls*{scct}\\ \Cref{thm:ccall:rtp:msscct}} (O);
-    \draw[thick,dotted,->] (S) to[out=-50,in=210,sloped] node[align=center,font=\scriptsize] {\gls*{mssscct}\\ \Cref{thm:ccall:rtp:mssscct}} (Z);
-
-    \draw[dashed] ($(sectms)-(1.2,0)$) -- ($(sectms)-(1.2,1.25)$);
-    \draw[dashed] ($(sectms)-(-1.0,0)$) -- ($(sectms)-(-1.0,1.25)$);
-    \draw[dashed] ($(secsms)-(-1.2,0)$) -- ($(secsms)-(-1.2,1.25)$);
-    \draw[dashed] ($(secscct)-(1.1,0)$) -- ($(secscct)-(1.1,1.25)$);
-    \draw[dashed] ($(secscct)-(-1.4,0)$) -- ($(secscct)-(-1.4,1.25)$);
-    \draw[dashed] ($(secss)-(-1.0,0)$) -- ($(secss)-(-1.0,1.25)$);
+    \node (sectms) at ($(S)+(1.25,3)$) {{\Cref{subsec:cs:tms}}};
+    \node (secsms) at ($(T)+(1.25,3)$) {{\Cref{subsec:cs:ms}}};
+    \node (secopts) at ($(M)+(2.0,3)$) {{\Cref{subsec:cs:opts}}};
+    \node (secscct) at ($(E)+(1.5,3)$) {{\Cref{subsec:cs:scct}}};
+    \node (secss) at ($(O)+(1.5,3)$) {{\Cref{subsec:cs:ss}}};
+
+    \draw[thick,dotted,->] (S) to[bend right=33,sloped] node[align=center] {\gls*{ms}\\ \Cref{thm:ccab:rtp:ms}} (M);
+    \draw[thick,dotted,->] (M) to[bend right=0,sloped] node[align=center] {\gls*{ms}\\ \Cref{thm:cccfccdce:rtp:ms}} (E);
+%    \draw[thick,dotted,->] (S) to[out=-45,in=198,sloped] node[align=center] {\gls*{ms}+\gls*{scct}\\ \Cref{thm:ccall:rtp:msscct}} (O);
+    \draw[thick,dotted,->] (S) to[out=-40,in=210,sloped] node[align=center] {\gls*{mssscct}\\ \Cref{thm:ccall:rtp:mssscct}} (Z);
+
+    \draw[dashed] ($(S)-(0.0,-0.5)$) -- ($(S)-(0.0,-3)$);
+    \draw[dashed] ($(T)-(0.0,-0.5)$) -- ($(T)-(0.0,-3)$);
+    \draw[dashed] ($(M)-(0.0,-0.5)$) -- ($(M)-(0.0,-3)$);
+    \draw[dashed] ($(E)-(0.0,-0.5)$) -- ($(E)-(0.0,-3)$);
+    \draw[dashed] ($(O)-(0.0,-0.5)$) -- ($(O)-(0.0,-3)$);
+    \draw[dashed] ($(Z)-(0.0,-0.5)$) -- ($(Z)-(0.0,-3)$);
   \end{tikzpicture}
@@
 
-\begin{theorem}[$\cca$ is secure with respect to \gls*{tms}]\label{thm:cca:rtp:tms}
+\begin{theorem}[$\cca$ is secure w.r.t. \gls*{tms}]\label{thm:cca:rtp:tms}
   $\rtp{\cca}{\tmssafe}$ % \Coqed
@@
 
-\begin{theorem}[$\ccb$ is secure with respect to \gls*{sms}]\label{thm:ccb:rtp:sms}
+\begin{theorem}[$\ccb$ is secure w.r.t. \gls*{sms}]\label{thm:ccb:rtp:sms}
   $\rtp{\ccb}{\smssafe}$ % \Coqed
@@
 
-\Cref{thm:ccab:rtp:ms} states that the composition of $\cca$ and $\ccb$ is secure with respect to \gls*{ms} and follows from \Cref{thm:cca:rtp:tms,thm:ccb:rtp:sms} using \Cref{thm:rtp}.
-
-\begin{theorem}[$\cca\circ\ccb$ is secure with respect to \gls*{ms}]\label{thm:ccab:rtp:ms}
+\Cref{thm:ccab:rtp:ms} states that the composition of $\cca$ and $\ccb$ is secure w.r.t. \gls*{ms} and follows from \Cref{thm:cca:rtp:tms,thm:ccb:rtp:sms} using \Cref{thm:rtp}.
+
+\begin{theorem}[$\cca\circ\ccb$ is secure w.r.t. \gls*{ms}]\label{thm:ccab:rtp:ms}
   $\rtp{\cca\circ\ccb}{\mssafe}$ % \Coqed
@@
 The \gls*{cf} pass relies on an auxiliary function \texttt{mix} that uses a substitutions accumulator $\irl{\substlist}$ in order to rewrite constant binary operations, e.g., $\irl{{17}-1}$ to $\irl{16}$, and replace variables that are assigned to constants with their constant, e.g., $\irl{\llet{x}{7}{x}}$ to $\irl{7}$.
-Both passes are secure with respect to \gls*{ms}.
+Both passes are secure w.r.t. \gls*{ms}.
 The proof for either is relatively simple, because both \gls*{dce} and \gls*{cf} do not change the way memory accesses happen.
@@
 
-\begin{theorem}[$\ccdce$ is secure with respect to \gls*{ms}]\label{thm:ccdce:rtp:ms}
+\begin{theorem}[$\ccdce$ is secure w.r.t. \gls*{ms}]\label{thm:ccdce:rtp:ms}
   $\rtp{\ccdce}{\mssafe}$ %\Coqed
 \end{theorem}
-\begin{theorem}[$\cccf$ is secure with respect to \gls*{ms}]\label{thm:cccf:rtp:ms}
+\begin{theorem}[$\cccf$ is secure w.r.t. \gls*{ms}]\label{thm:cccf:rtp:ms}
   $\rtp{\cccf}{\mssafe}$ %\Coqe
@@
 
-\begin{theorem}[$\cccf\circ\ccdce$ and $\cccf\circ\ccdce$ are secure with respect to \gls*{ms}]\label{thm:cccfccdce:rtp:ms}
+\begin{theorem}[$\cccf\circ\ccdce$ and $\cccf\circ\ccdce$ are secure w.r.t. \gls*{ms}]\label{thm:cccfccdce:rtp:ms}
   $\rtp{\cccf\circ\ccdce}{\mssafe}$ and $\rtp{\ccdce\circ\cccf}{\mssafe}$. % \Coqed
@@
 The context can overwrite the flag and exit the mode, but upon invoking a function that is part of the component, the flag would be set again.
-Because of this, the compiler is secure with respect to \gls*{scct}, similarly proven as in \Cref{subsec:cs:tms}.
-
-\begin{theorem}[$\ccscct$ is secure with respect to \gls*{scct}]\label{thm:ccscct:rtp:scct}
+Because of this, the compiler is secure w.r.t. \gls*{scct}, similarly proven as in \Cref{subsec:cs:tms}.
+
+\begin{theorem}[$\ccscct$ is secure w.r.t. \gls*{scct}]\label{thm:ccscct:rtp:scct}
   $\rtp{\ccscct}{\scctsafe}$ % \Coqed
@@
 %
-\begin{theorem}[$\ccspec$ is secure with respect to \gls*{ss}]\label{thm:ccspec:rtp:spec}
+\begin{theorem}[$\ccspec$ is secure w.r.t. \gls*{ss}]\label{thm:ccspec:rtp:spec}
   $\rtp{\ccspec}{\specmssafe}$ % \Coqed
@@
 
-\begin{theorem}[$\ccmssscct$ is secure with respect to \gls*{mssscct}]\label{thm:ccall:rtp:mssscct}
+\begin{theorem}[$\ccmssscct$ is secure w.r.t. \gls*{mssscct}]\label{thm:ccall:rtp:mssscct}
   $\rtp{\cc{\Ltms}{\Lspec}}{\mssafe\cap\scctsafe\cap\sssafe}$ % \Coqed
@@
 
-\subsection{Preconditions for Compositionality}
-
-The composition theorems can only be used on compilers which have been proven secure in the sense that they robustly preserve a certain class of properties. 
-Traditional proof techniques apply, the usual strategy is to establish a backwards simulation between source components and their compiled target counterpart. 
-For the universally quantified target context, a technique known as backtranslation is well-established, which constructs a source-level context either by backtranslating the target-level context, amounting to a ,,decompilation'', or by constructing it using the target-level trace, so that the source-context emits parts of the whole execution trace that are related to the parts emitted by the target-level context.
-Up to backtranslations for $\ccdce$ and $\cccf$, all backtranslations in the case-study are trace based, while for $\ccdce$ and $\cccf$, they are context-based. 
-Since these optimisations have no significant semantic effect and source- and target-language are equal, the backtranslation amounts to an identity translation.
+\subsection{Well-formedness of Cross Language Trace Relations}
 
@@
 
-\Cref{thm:rtpsim:sig} is always applicable, up to well-formedness of $\sim_1$ with respect to $\trg{\class[_2]}$.
+\Cref{thm:rtpsim:sig} is always applicable, up to well-formedness of $\sim_1$ w.r.t. $\trg{\class[_2]}$.
 However, consider $\ccb$ and $\ccspec$.
@@
 Since $\ev{Use\ \loc\ n}\sim_{\irl{ms}}\ev{Branch n}\cdot\ev{Spec}\cdots$, where $\cdots$ does not contain a $\ird{\ev{Barrier}}$ event, the resulting class of properties is only a subset of the intended one.
-Because of this, it is crucial to analyze the precise shape of the class of properties after mapping a target-level property to a source-level property\footnote{Dito for vice-versa.}.
+Because of this, it is crucial to analyze the precise shape of the class of properties after mapping a target-level property to a source-level property\footnote{Dito for mapping a source-level property to a target-level property.}.
+
+\subsection{Secure Compilation Proofs}
+
+The composition theorems can only be used on compilers which have been proven secure in the sense that they robustly preserve a certain class of properties. 
+Traditional proof techniques apply, the usual strategy is to establish a backwards simulation between source components and their compiled target counterpart. 
+For the universally quantified target context, a technique known as backtranslation is well-established, which constructs a source-level context either by backtranslating the target-level context, amounting to a ,,decompilation'', or by constructing it using the target-level trace, so that the source-context emits parts of the whole execution trace that are related to the parts emitted by the target-level context.
+Up to backtranslations for $\ccdce$ and $\cccf$, all backtranslations in the case-study are trace based, while for $\ccdce$ and $\cccf$, they are context-based. 
+Since these optimisations have no significant semantic effect and source- and target-language are equal, the backtranslation amounts to an identity translation.
 

### 2024-05-23 csf2/main.tex
move table columns

@@
 \begin{figure}[!htb]
- \vspace{-2em}
   $$
-  \scriptsize
   % {
-  \begin{array}{ccccccc}
-    & \Lscct &        &   &      & Spec & \\
-    Active & \mid & Inactive & \mid & Active & \mid & Inactive \\\hline
-    \obj{\ev{Alloc}\ \loc_{x}\ 12;\comp;\unlock} & \mid & \obj{\ev{Alloc}\ \loc_{x}\ 12;\comp;\unlock} & \mid & \ev{Alloc}\ \loc_{x}\ 12;\unlock & \mid & \ev{Alloc}\ \loc_{x}\ 12;\unlock\\
-    \obj{\ev{Alloc}\ \loc_{y}\ 12;\comp;\unlock} & \mid & \obj{\ev{Alloc}\ \loc_{y}\ 12;\comp;\unlock} & \mid & \ev{Alloc}\ \loc_{y}\ 12;\unlock & \mid & \ev{Alloc}\ \loc_{x}\ 12;\unlock\\
-    \obj{{\ev{iGet}}\ \loc_{x}\ 0;\comp;}\text{\colorbox{lightgray}{$\obj{\lock}$}} & \mid & \obj{\ev{Get}\ \loc_{x}\ 0;\comp;\lock} & \mid & \ev{Use}\ \loc_{x}\ 0;\text{\colorbox{lightgray}{$\unlock$}} & \mid & \ev{Use}\ \loc_{x}\ 0;\lock\\
-    \text{\colorbox{yellow}{$\obj{\emptyevent}$}} & \mid & \obj{\ev{Branch}\ 0;\comp;\lock} & \mid & \emptyevent & \mid & \ev{Branch}\ 0;\lock\\
-    \obj{{\ev{iGet}}\ \loc_{x}\ 0;\comp;}\text{\colorbox{lightgray}{$\obj{\lock}$}} & \mid & \obj{\ev{Get}\ \loc_{x}\ 0;\comp;\lock} & \mid & \ev{Use}\ \loc_{x}\ 0;\text{\colorbox{lightgray}{$\unlock$}} & \mid & \ev{Use}\ \loc_{x}\ 0;\lock\\
-    \obj{{\ev{iSet}}\ \loc_{y}\ 0\ \mathtt{'H'};\comp;}\text{\colorbox{lightgray}{$\obj{\lock}$}} & \mid & \obj{\ev{Set}\ \loc_{y}\ 0\ \mathtt{'H'};\comp;\lock} & \mid & \ev{Use}\ \loc_{y}\ 0;\text{\colorbox{lightgray}{$\unlock$}} & \mid & \ev{Use}\ \loc_{y}\ 0;\lock\\
-    \obj{{\ev{iGet}}\ \loc_{x}\ 1;\comp;}\text{\colorbox{lightgray}{$\obj{\lock}$}} & \mid & \obj{\ev{Get}\ \loc_{x}\ 1;\comp;\lock} & \mid & \ev{Use}\ \loc_{x}\ 1;\text{\colorbox{lightgray}{$\unlock$}} & \mid & \ev{Use}\ \loc_{x}\ 1;\lock\\
-    \text{\colorbox{yellow}{$\obj{\emptyevent}$}} & \mid & \obj{\ev{Branch}\ 0;\comp;\lock} & \mid & \emptyevent & \mid & \ev{Branch}\ 0;\lock\\
-    \vdots & \mid & \vdots & \mid & \vdots & \mid & \vdots \\
-    \obj{{\ev{iGet}}\ \loc_{x}\ 12;\comp;}\text{\colorbox{lightgray}{$\obj{\lock}$}} & \mid & \obj{\ev{Get}\ \loc_{x}\ 12;\comp;\lock} & \mid &  \ev{Use}\ \loc_{x}\ 12;\text{\colorbox{lightgray}{$\unlock$}} & \mid & \ev{Use}\ \loc_{x}\ 12;\lock\\
-    \text{\colorbox{yellow}{$\obj{\emptyevent}$}}& \mid & \obj{\ev{Branch}\ 1;\comp;\lock} & \mid &  \emptyevent & \mid & \ev{Branch}\ 1;\lock\\
-    \obj{\ev{Dealloc}\ \loc_{y};\comp;\unlock} & \mid & \obj{\ev{Dealloc}\ \loc_{y};\comp;\lock} & \mid &  \ev{Dealloc}\ \loc_{y};\unlock & \mid & \ev{Dealloc}\ \loc_{y};\unlock\\
-    \obj{{\ev{iGet}}\ \loc_{y}\ 6;\comp;}\text{\colorbox{lightgray}{$\obj{\lock}$}} & \mid & \obj{\ev{Get}\ \loc_{y}\ 6;\comp;\lock} & \mid &  \ev{Use}\ \loc_{y}\ 6;\text{\colorbox{lightgray}{$\unlock$}} & \mid & \ev{Use}\ \loc_{y}\ 6;\lock\\
+  \begin{array}{ccc}
+    & \Lscct &        \\
+    Active & \mid & Inactive \\\hline
+    \obj{\ev{Alloc}\ \loc_{x}\ 12;\comp;\unlock} & \mid & \obj{\ev{Alloc}\ \loc_{x}\ 12;\comp;\unlock} \\
+    \obj{\ev{Alloc}\ \loc_{y}\ 12;\comp;\unlock} & \mid & \obj{\ev{Alloc}\ \loc_{y}\ 12;\comp;\unlock} \\
+    \obj{{\ev{iGet}}\ \loc_{x}\ 0;\comp;}\text{\colorbox{lightgray}{$\obj{\lock}$}} & \mid & \obj{\ev{Get}\ \loc_{x}\ 0;\comp;\lock} \\
+    \text{\colorbox{yellow}{$\obj{\emptyevent}$}} & \mid & \obj{\ev{Branch}\ 0;\comp;\lock} \\
+    \obj{{\ev{iGet}}\ \loc_{x}\ 0;\comp;}\text{\colorbox{lightgray}{$\obj{\lock}$}} & \mid & \obj{\ev{Get}\ \loc_{x}\ 0;\comp;\lock} \\
+    \obj{{\ev{iSet}}\ \loc_{y}\ 0\ \mathtt{'H'};\comp;}\text{\colorbox{lightgray}{$\obj{\lock}$}} & \mid & \obj{\ev{Set}\ \loc_{y}\ 0\ \mathtt{'H'};\comp;\lock} \\
+    \obj{{\ev{iGet}}\ \loc_{x}\ 1;\comp;}\text{\colorbox{lightgray}{$\obj{\lock}$}} & \mid & \obj{\ev{Get}\ \loc_{x}\ 1;\comp;\lock} \\
+    \text{\colorbox{yellow}{$\obj{\emptyevent}$}} & \mid & \obj{\ev{Branch}\ 0;\comp;\lock} \\
+    \vdots & \mid & \vdots \\
+    \obj{{\ev{iGet}}\ \loc_{x}\ 12;\comp;}\text{\colorbox{lightgray}{$\obj{\lock}$}} & \mid & \obj{\ev{Get}\ \loc_{x}\ 12;\comp;\lock} \\
+    \text{\colorbox{yellow}{$\obj{\emptyevent}$}}& \mid & \obj{\ev{Branch}\ 1;\comp;\lock} \\
+    \obj{\ev{Dealloc}\ \loc_{y};\comp;\unlock} & \mid & \obj{\ev{Dealloc}\ \loc_{y};\comp;\lock} \\
+    \obj{{\ev{iGet}}\ \loc_{y}\ 6;\comp;}\text{\colorbox{lightgray}{$\obj{\lock}$}} & \mid & \obj{\ev{Get}\ \loc_{y}\ 6;\comp;\lock} \\
+  \end{array}
+  % }
+  $$
+  $$
+  % {
+  \begin{array}{ccc}
+          & Spec & \\
+     Active & \mid & Inactive \\\hline
+     \ev{Alloc}\ \loc_{x}\ 12;\unlock & \mid & \ev{Alloc}\ \loc_{x}\ 12;\unlock\\
+     \ev{Alloc}\ \loc_{y}\ 12;\unlock & \mid & \ev{Alloc}\ \loc_{x}\ 12;\unlock\\
+     \ev{Use}\ \loc_{x}\ 0;\text{\colorbox{lightgray}{$\unlock$}} & \mid & \ev{Use}\ \loc_{x}\ 0;\lock\\
+     \emptyevent & \mid & \ev{Branch}\ 0;\lock\\
+     \ev{Use}\ \loc_{x}\ 0;\text{\colorbox{lightgray}{$\unlock$}} & \mid & \ev{Use}\ \loc_{x}\ 0;\lock\\
+     \ev{Use}\ \loc_{y}\ 0;\text{\colorbox{lightgray}{$\unlock$}} & \mid & \ev{Use}\ \loc_{y}\ 0;\lock\\
+     \ev{Use}\ \loc_{x}\ 1;\text{\colorbox{lightgray}{$\unlock$}} & \mid & \ev{Use}\ \loc_{x}\ 1;\lock\\
+     \emptyevent & \mid & \ev{Branch}\ 0;\lock\\
+     \vdots & \mid & \vdots \\
+      \ev{Use}\ \loc_{x}\ 12;\text{\colorbox{lightgray}{$\unlock$}} & \mid & \ev{Use}\ \loc_{x}\ 12;\lock\\
+      \emptyevent & \mid & \ev{Branch}\ 1;\lock\\
+      \ev{Dealloc}\ \loc_{y};\unlock & \mid & \ev{Dealloc}\ \loc_{y};\unlock\\
+      \ev{Use}\ \loc_{y}\ 6;\text{\colorbox{lightgray}{$\unlock$}} & \mid & \ev{Use}\ \loc_{y}\ 6;\lock\\
   \end{array}

### 2024-05-23 csf2/main.tex
Describe some syntax

@@
 
+$\src{L_{\tmssafe}}$ is a relatively naive language with second class functionns, the only values being numbers. 
+It has simple binary operations, case analyses, C-like memory operations on arrays, as well as C-like function calls.
+Programs have sets of pre-determined functions called libraries and it is assumed that there are functions part of some component or context.
+
 \vspace{-1.5em}
 \[
-  \begin{array}{rcl}
-    \src{\event[^b]} &:=& \src{\ev{Alloc \loc\ n}} \mid \src{\ev{Dealloc \loc}} \mid \src{\ev{Get \loc\ n}} \\
-                     && \mid \src{\ev{Set \loc\ n}} \mid \src{\ev{Branch\ n}} \mid \src{\ev{Binop\ n}}\\
-   && \src{\sandboxtag} := \src{\ctx} \mid \src{\comp} \hspace{3em}
-   \src{\event} := \src{\event[^b];\sandboxtag} \mid \src{\emptyevent} \mid \src{\lightning} \\ 
-    \src{\expr} &:=& \src{x} \mid \src{n} \mid \src{\lbinop{\expr[_1]}{\expr[_2]}} \mid \src{\lifz{\expr[_1]}{\expr[_2]}{\expr[_3]}} \mid \src{\llet{x}{\expr[_1]}{\expr[_2]}} \\
-                && \src{\library_{\ctx}} : \src{Vars} \to \src{\expr} \hspace{4em}
+  \begin{array}{rrcl}
+    \text{(Pre-Events)} & \src{\preevent} &:=& \src{\ev{Alloc \loc\ n}} \mid \src{\ev{Dealloc \loc}} \mid \src{\ev{Get \loc\ n}} \mid \src{\ev{Set \loc\ n}} \\
+    \text{(Sandboxtag)} & \src{\sandboxtag} &:=& \src{\ctx} \mid \src{\comp} \\
+    \text{(Events)} & \src{\event} &:=& \src{\preevent;\sandboxtag} \mid \src{\emptyevent} \mid \src{\lightning} \\ 
+    \text{(Expressions)} & \src{\expr} &:=& \src{x} \mid \src{n} \mid \src{\lbinop{\expr[_1]}{\expr[_2]}} \mid \src{\lifz{\expr[_1]}{\expr[_2]}{\expr[_3]}} \\ 
+                         &&&\mid \src{\llet{x}{\expr[_1]}{\expr[_2]}} \mid \src{\llet{x}{new\ \expr[_1]}{\expr[_2]}} \\
+                         &&&\mid \src{\ldelete{x}} \mid \src{\lget{x}{\expr}} \mid \src{\lset{x}{\expr[_1]}{\expr[]}} \mid \src{stuck}\\
+                         &&&\mid \src{\lcall{f}{\expr}} \mid \src{\lreturn{\expr}} \\
+    \text{(Libraries)} &&& \src{\library_{\ctx}} : \src{Vars} \to \src{\expr} \hspace{2em}
     \src{\library_{\comp}} : \src{Vars} \to \src{\expr} \\
@@
 
-$\src{L_{\tmssafe}}$ is statically typed, where the type system is inspired by $L^{3}$~\cite{morrisett2005L3,scherer2018fabulous}.
+Crucially, $\src{L_{\tmssafe}}$ is statically typed, where the type system is inspired by $L^{3}$~\cite{morrisett2005L3,scherer2018fabulous}.
 The type system of $\src{L_{\tmssafe}}$ enforces that every well-typed $\src{L_{\tmssafe}}$ program satisfies \gls*{tms} (\Cref{thm:wt:tms}).
@@
 \[
-  \begin{array}{rcl}
-    \trg{\expr} &:=& \dots \mid \trg{\lhast{\expr}{\tau}} \mid \trg{\lispoisoned{x}}
+  \begin{array}{rrcl}
+    \text{(Expressions)} & \trg{\expr} &:=& \dots \mid \trg{\lhast{\expr}{\tau}} \mid \trg{\lispoisoned{x}}
   \end{array}
@@
 \[
-  \begin{array}{rcl}
-    \obj{D} &:=& \obj{ON} \mid \obj{OFF} \hspace{2em}
-    \obj{\securitytag} := \obj{\lock} \mid \obj{\unlock} \\
-    \obj{\event[^b]} &:=& \dots \mid \obj{iGet\ \loc\ v} \mid \obj{iSet\ \loc\ v} \mid \obj{Binop\ n} \mid \obj{Branch\ n} \\
-    \obj{\expr} &:=& \dots \mid \obj{\lrddoit{x}{\expr}} \mid \obj{\lwrdoit{D}} \mid \obj{\llet{x^{\securitytag}}{\expr[_1]}{\expr[_2]}}\\
-    \obj{\event} &:=& \obj{\event[^b];\sandboxtag;\securitytag} \mid \dots
+  \begin{array}{rrcl}
+    \text{(Mode Values)} & \obj{D} &:=& \obj{ON} \mid \obj{OFF} \\
+    \text{(Security Tags)} & \obj{\securitytag} &:=& \obj{\lock} \mid \obj{\unlock} \\
+    \text{(Pre-Event)} & \obj{\preevent} &:=& \dots \mid \obj{iGet\ \loc\ v} \mid \obj{iSet\ \loc\ v} \\
+                       &&& \mid\ \obj{Binop\ n} \mid \obj{Branch\ n} \\
+    \text{(Event)} & \obj{\event} &:=& \obj{\event[^b];\sandboxtag;\securitytag} \mid \obj{\emptyevent} \mid \obj{\lightning} \\
+    \text{(Expressions)} & \obj{\expr} &:=& \dots \mid \obj{\lrddoit{x}{\expr}} \mid \obj{\lwrdoit{D}} \\
+                         &&&\mid\ \obj{\llet{x^{\securitytag}}{\expr[_1]}{\expr[_2]}}\\
   \end{array}
@@
 \[
-  \begin{array}{rcl}
-    \ird{\event[^b]} &:=& \dots \mid \ird{Spec} \mid \ird{Rlb} \mid \ird{Barrier}\\
-    \ird{\expr[^b]} &:=& \dots \mid \ird{\lbarrier}
+  \begin{array}{rrcl}
+    \text{(Pre-Events)} & \ird{\preevent} &:=& \dots \mid \ird{Spec} \mid \ird{Rlb} \mid \ird{Barrier}\\
+    \text{(Expressions)} & \ird{\expr} &:=& \dots \mid \ird{\lbarrier}
   \end{array}

### 2024-05-23 csf2/main.tex
pass over lang section

@@
 
-$\src{L_{\tmssafe}}$ is a relatively naive language with second class functionns, the only values being numbers. 
-It has simple binary operations, case analyses, C-like memory operations on arrays, as well as C-like function calls.
+$\src{L_{\tmssafe}}$ is a relatively naive language with second class functions, the only values being numbers. 
+It has simple binary operations, case analyses, C-like function calls, as well as C-like memory operations. 
+Pointers are left second-class and can only be referenced by an identifier, so addresses cannot be guessed by contexts.
 Programs have sets of pre-determined functions called libraries and it is assumed that there are functions part of some component or context.
@@
 \begin{theorem}[$\src{L_{\tmssafe}}$-programs are \gls*{tms}]\label{thm:wt:tms}
-  For any $\src{\library_{\comp}}$, $\rsat{\src{\library_{\comp}}}{\tmssafe}$ \Coqed
+  For any component $\src{\library_{\comp}}$, $\rsat{\src{\library_{\comp}}}{\tmssafe}$ \Coqed
 \end{theorem}
@@
 \subsection{$\Ltrg$: A Memory-Unsafe Language}\label{subsec:lsms}
+
+$\Ltrg$ extends the syntax presented earlier (\Cref{subsec:cs:defs}) with dynamic typechecks ($\trg{\lhast{\expr}{\type}}$), evaluating to $\trg{0}$ if the type matches with the shape of $\trg{\expr}$ and $\trg{1}$ otherwise.
+%The pointers carry poison tags, which signify a pointer being allocated ($\trg{\poisoned}$) or freed ($\trg{\poisonless}$). 
+Furthermore, the syntax of $\Ltrg$ is extended with a way to inspect whether a pointer is freed ($\trg{\lispoisoned{x}}$), evaluating to $\trg{0}$ if it is freed and to $\trg{1}$ otherwise.
 
@@
 
-$\Ltrg$ extends the syntax presented earlier (\Cref{subsec:cs:defs}) with dynamic typechecks ($\trg{\lhast{\expr}{\type}}$), evaluating to $\trg{0}$ if the type matches with the shape of $\trg{\expr}$ and $\trg{1}$ otherwise.
-%The pointers carry poison tags, which signify a pointer being allocated ($\trg{\poisoned}$) or freed ($\trg{\poisonless}$). 
-Furthermore, the syntax of $\Ltrg$ is extended with a way to inspect whether a pointer is freed ($\trg{\lispoisoned{x}}$), evaluating to $\trg{0}$ if it is freed and to $\trg{1}$ otherwise.
 No changes are done to the trace model, but $\Ltrg$ has no static typing, making double-free patterns possible.
@@
 \subsection{$\Lscct$: A Memory-Unsafe Language with a Data Independent Timing Mode}\label{subsec:lscct}
+
+$\Lscct$ extends $\Lms$ (\Cref{subsec:lms}) with a way to read from ($\obj{\lrddoit{x}{\expr}}$) and to write to ($\obj{\lwrdoit{D}}$) a model specific register that controls a data (operand) independent timing mode, a feature that is present in both Arm~\cite[p.~543]{arm-refman} and Intel~\cite[p.~80]{intel-refman} processors.
+To this end, states are extended with the value of the register ($\obj{D}$), which is initially set to be not active ($\obj{OFF}$).
+If the register is marked active ($\obj{ON}$), the intuition is that no secrets can appear on specification traces.
+If the register is marked inactive, secrets may appear on traces.
+The state of the register can be read and written to, e.g., \Cref{tr:e-wrdoit-off}.
+For the other languages seen earlier, the mode is intuitively always-on, i.e., the mode of execution always uses data independent timing.
 
@@
 
-$\Lscct$ extends $\Lms$ (\Cref{subsec:lms}) with a way to read from ($\obj{\lrddoit{x}{\expr}}$) and to write to ($\obj{\lwrdoit{D}}$) a model specific register that controls a data (operand) independent timing mode, a feature that is present in both Arm~\cite[p.~543]{arm-refman} and Intel~\cite[p.~80]{intel-refman} processors.
-To this end, states are extended with the value of the register ($\obj{D}$), which is initially set to be not active ($\obj{OFF}$).
-If the register is marked active ($\obj{ON}$), the intuition is that no secrets can appear on specification traces.
-If the register is marked inactive, secrets may appear on traces.
-For the other languages seen earlier, the mode is intuitively always-on, i.e., the mode of execution always uses data independent timing.
 The language also adds user-annotations for secrecy $\obj{\securitytag}$, which is either high $\obj{\lock}$ or low $\obj{\unlock}$.
@@
 To prevent secrets from leaking but still enable reasoning about memory safety, $\Lscct$ extends pre-events with $\obj{\ev{iGet}\ \loc\ \valueexpr}$ and $\obj{\ev{iSet}\ \loc\ \valueexpr}$.
-This way, memory accesses to secret data are still exposed as memory accesses, despite being protected by data-independent timing mode.
+This way, memory accesses to secret data are still exposed as memory accesses, despite being protected by data-independent timing mode, e.g, \Cref{tr:to-e-get-in-noleak}.
 Moreover, the language extends pre-events with $\obj{\ev{Branch}\ n}$ and $\obj{\ev{Binop}\ n}$ that are emitted when evaluating a branch or certain binary expressions, such as division, respectively, whenever the data independent timing mode is inactive.
@@
   Consider again \Cref{ex:strncpy}, with a context copying the string \texttt{Hello World}, where everything is marked with a high security tag: $\lock$.
-  The left half of \Cref{fig:ex-cct} ($\Lscct$, the two columns on the left), describes the execution trace of the program, while the right side of the table ($Spec$, the two columns on the right), describes the related specification trace (\Cref{subsec:msctss:tracemodel}).
-  Read from top to bottom, the figure shows parts of the execution trace. 
+  The top half of \Cref{fig:ex-cct} (titled $\Lscct$), describes the execution trace of the program, while the bottom side of the table (titled $Spec$), describes the related specification trace (\Cref{subsec:msctss:tracemodel}).
+  Read in parallel from top to bottom, the figure shows parts of the execution trace. 
   In each half, the left column (\textit{Active}) has data-independent timing $\obj{ON}$ and the right one (\textit{Inactive}) has it $\obj{OFF}$.
@@
   When the data independent timing mode is off, the execution yields events in similar fashion to before (\Cref{subsec:ltms,subsec:lsms,subsec:lms}).
-  But, if it is turned on, then the branching event \colorbox{yellow}{does not fire} anymore and both reading and writing to memory gets ultimately translated to a specification trace with \colorbox{lightgray}{no exposed secrets}.
+  But, if it is turned on, then the branching event \colorbox{yellow}{does not fire} anymore and both reading and writing to memory is related to a specification trace with \colorbox{lightgray}{no exposed secrets}.
 \end{exampleenv}
@@
 
-\vspace{-1.5em}
+$\Lspec$ extends $\Lscct$ (\Cref{subsec:lscct}) with speculative dynamic semantics. 
+After branching or accessing memory, speculation can start and run subsequent code in a predetermined window until a rollback of operational state is performed. 
+Within such windows, data is leaked, marked as high $\ird{\lock}$ and with an annotation that indicates the respective speculative execution variant, in this case just SPECTRE-PHT~\cite{kocher2019spectre}, whose semantics is sketched with \Cref{tr:e-ifz-true-spec}, where $\omega$ is the size of the speculation window.
+The semantics is kept general enough to allow for future extension to support different variants.
+In terms of language features, $\Lspec$ supports a barrier operation $\ird{\lbarrier}$ that blocks speculative execution. 
+To facilitate speculative execution in a non-assembly-like language, a stack of operational state is used and speculation is active if that stack has a size greater than one, indicated by writing the operational states vertically.
+This is exploited in \Cref{tr:e-spec-eat} to ,,eat away'' any leftover speculation.
+
+\vspace{-1em}
 \[
@@
   \typerule{$e-\ird{ifz}-\ird{true}-\text{spec}$}{
-    \dots \text{todo}
+    \ird{\cfstate[']}.\ird{\operatorname{win}} = \omega
   }{
-    \pstepto{\rtt{\ird{\cfstate;\sandboxtag;0;\memstate}}{\ird{\lifz{0^{\securitytag}}{\expr[_1]}{\expr[_2]}}}}{\begin{array}{c}\rtt{\ird{\cfstate;\sandboxtag;0;\memstate}}{\ird{\expr[_1]}}\\\rtt{\ird{\cfstate;\sandboxtag;0;\memstate}}{\ird{\expr[_2]}}\end{array}}{\ird{\ev{Branch}\ 0;\sandboxtag;\securitytag}\cdot\ird{\ev{Spec};\sandboxtag;\securitytag}}
-  }{e-ifz-true-leak}
+    \pstepto{\rtt{\ird{\cfstate;\sandboxtag;0;\memstate}}{\ird{\lifz{0^{\securitytag}}{\expr[_1]}{\expr[_2]}}}}{\begin{array}{c}\rtt{\ird{\cfstate['];\sandboxtag;0;\memstate}}{\ird{\expr[_1]}}\\\rtt{\ird{\cfstate;\sandboxtag;0;\memstate}}{\ird{\expr[_2]}}\end{array}}{\ird{\ev{Branch}\ 0;\sandboxtag;\securitytag}\cdot\ird{\ev{Spec};\sandboxtag;\securitytag}}
+  }{e-ifz-true-spec}
+  %
+  \typerule{$e-\ird{spec}-\text{eat}$}{
+    \ird{\cfstate}.\ird{\operatorname{win}} > 0 &
+    \pstepto{\rtt{\ird{\cfstate;\sandboxtag;0;\memstate}}{\ird{\expr}}}{\rtt{\ird{\cfstate['];\sandboxtag;0;\memstate}}{\ird{\expr[']}}}{\ird{\event}} &
+    \ird{\cfstate['']}.\ird{\operatorname{win}} = \ird{\cfstate[']}.\ird{\operatorname{win}}-1
+  }{
+    \pstepto{\begin{array}{c}\rtt{\ird{\cfstate;\sandboxtag;0;\memstate}}{\ird{\expr}}\\\ird{r}\end{array}}{\begin{array}{c}\rtt{\ird{\cfstate[''];\sandboxtag;0;\memstate}}{\ird{\expr[']}}\\\ird{r}\end{array}}{\ird{\event}}
+  }{e-spec-eat}
+  % 
+  \typerule{$e-\ird{spec}-\text{eaten}$}{
+    \ird{r}.\ird{\cfstate}.\ird{\operatorname{win}} = 0
+  }{
+    \pstepto{\begin{array}{c}\ird{r}\\\ird{r'}\end{array}}{\ird{r'}}{\ird{Rlb}}
+  }{e-spec-eaten}
+  %
+  \typerule{$e-\ird{spec}-\text{barrier}$}{
+    \ird{r'}.\ird{\cfstate} = \ird{r}.\ird{\cfstate}[.\ird{\operatorname{win}} = 0]
+  }{
+    \pstepto{\begin{array}{c}\ird{r}\\\ird{r''}\end{array}}{\begin{array}{c}\ird{r'}\\\ird{r''}\end{array}}{\ird{Barrier}}
+  }{e-spec-barrier}
 \end{center}
 
-$\Lspec$ extends $\Lscct$ (\Cref{subsec:lscct}) with speculative dynamic semantics. 
-After branching or accessing memory, speculation can start and run subsequent code in a predetermined window until a rollback of operational state is performed. 
-Within such windows, data is leaked, marked as high $\ird{\lock}$ and with an annotation that indicates the respective speculative execution variant.
-In terms of language features, $\Lspec$ supports a barrier operation $\ird{\lbarrier}$ that blocks speculative execution. 
 The trace model is extended with $\ird{\ev{Spec}}$, $\ird{\ev{Rlb}}$, and $\ird{\ev{Barrier}}$ events that signal the respective operational action. 
-A $\ird{\lbarrier}$ prevents any execution whatsoever besides $\ird{\ev{Rlb}}$ when run in speculative execution mode and does nothing when run in normal mode.
+A $\ird{\lbarrier}$ prevents any execution whatsoever besides $\ird{\ev{Rlb}}$ when run in speculative execution mode (see \Cref{tr:e-spec-barrier}) and does nothing when run in normal mode.
 The speculative semantics are inspired by existing work~\cite{guarnieri2018spectector}.

### 2024-05-24 csf2/main.tex
use ghostarrow in spec

@@
   }{
-    \pstepto{\begin{array}{c}\rtt{\ird{\cfstate;\sandboxtag;0;\memstate}}{\ird{\expr}}\\\ird{r}\end{array}}{\begin{array}{c}\rtt{\ird{\cfstate[''];\sandboxtag;0;\memstate}}{\ird{\expr[']}}\\\ird{r}\end{array}}{\ird{\event}}
+    \ghoststepto{\begin{array}{c}\rtt{\ird{\cfstate;\sandboxtag;0;\memstate}}{\ird{\expr}}\\\ird{r}\end{array}}{\begin{array}{c}\rtt{\ird{\cfstate[''];\sandboxtag;0;\memstate}}{\ird{\expr[']}}\\\ird{r}\end{array}}{\ird{\event}}
   }{e-spec-eat}
@@
   }{
-    \pstepto{\begin{array}{c}\ird{r}\\\ird{r'}\end{array}}{\ird{r'}}{\ird{Rlb}}
+    \ghoststepto{\begin{array}{c}\ird{r}\\\ird{r'}\end{array}}{\ird{r'}}{\ird{Rlb}}
   }{e-spec-eaten}
@@
   }{
-    \pstepto{\begin{array}{c}\ird{r}\\\ird{r''}\end{array}}{\begin{array}{c}\ird{r'}\\\ird{r''}\end{array}}{\ird{Barrier}}
+    \ghoststepto{\begin{array}{c}\ird{r}\\\ird{r''}\end{array}}{\begin{array}{c}\ird{r'}\\\ird{r''}\end{array}}{\ird{Barrier}}
   }{e-spec-barrier}

### 2024-05-24 csf2/main.tex
fixing typos and some space opts

@@
 \begin{definition}[Property Satisfaction]\label{def:propsat}
+$\;$
+
+  \begin{nscenter}
     \bul{$\sat{\progvar}{\pi}$}
@@
 \begin{definition}[Robust Satisfaction]\label{def:proprsat}
-  % \begin{nscenter}
+  $\;$
+
+  \begin{nscenter}
   \bul{$\rsat{\progvar}{\pi}$}
   $\isdef$ \iul{$\forall \contextvar\ \wholeprogvar$, if $\link{\contextvar}{\progvar}=\wholeprogvar$}, then \oul{$\sat{\wholeprogvar}{\pi}$}.
-  % \end{nscenter}
+  \end{nscenter}
 \end{definition}
@@
 The case study of this paper exclusively uses the universal image, since some considered properties, such as \gls*{ss}, are not definable in a higher-level language that, e.g., does not model speculation.
+For sake of self-containment, the definition of the universal image is reiterated~\cite{abate2021extacc}:
+
+\begin{definition}[Universal Image]\label{def:universal:img}\label{def:sigma}
+  \[ 
+    \sigma_\sim\left(\trg{\pi}\right) := \left\{ \src{\trace} \mid \forall \trg{\trace}, \text{if }\src{\trace}\sim\trg{\trace}, \text{ then } \trg{\trace}\in\trg{\pi} \right\}
+  \]
+\end{definition}
 
@@
 
-Dually to \Cref{def:rtp:tau}, for a \bul{compiler $\cc{\src{L}}{\trg{L}}$ to robustly preserve a class of target properties $\trg{\class}$}, if for any \rul{property $\trg{\pi}$ of that class $\trg{\class}$ and programs $\src{p}$ written in $\src{L}$} where \iul{$\src{\progvar}$ robustly satisfies $\sigma_\sim\left(\trg{\pi}\right)$}, then \oul{the compilation of $\src{\progvar}$, $\cc{\src{L}}{\trg{L}}\left(\src{p}\right)$, must robustly satisfy $\trg{\pi}$}.
+Dually to \Cref{def:rtp:tau}, a \bul{compiler $\cc{\src{L}}{\trg{L}}$ to robustly preserves a class of target properties $\trg{\class}$}, if for any \rul{property $\trg{\pi}$ of class $\trg{\class}$ and programs $\src{p}$}, where \iul{$\src{\progvar}$ robustly satisfies $\sigma_\sim\left(\trg{\pi}\right)$}, \oul{the compilation of $\src{\progvar}$, $\cc{\src{L}}{\trg{L}}\left(\src{p}\right)$, must robustly satisfy $\trg{\pi}$}.
 
@@
 
-A degenerate case where $\sim$ coincides with equality, i.e., the trace model is the same for both source and target programs, is the classic definition of robust preservation~\cite{abate2019jour}:
+A degenerate case where $\sim$ coincides with equality, i.e., the trace model is the same for both source and target programs and the compiler does not have an effect on the resulting traces, is the classic definition of robust preservation~\cite{abate2019jour}:
 
@@
   % Compiler $\cc{\src{L}}{\trg{L}}$ robustly preserves $\class$, 
+  $\;$
+
   \bul{$\rtp{\cc{\src{L}}{\trg{L}}}{\class}$}
@@
 
-\bul{A cross-language relation $\sim$ is well-formed w.r.t. a class of target-level properties $\trg{\class}$} iff \rul{the universal image preserves set membership}.
-\begin{definition}[Well-formedness of Trace Relations for a Class $\trg{\class}$]\label{def:wfc:sig:tracerel}
+\bul{A relation $\sim$ is well-formed w.r.t. a class of target-level properties $\trg{\class}$} iff \rul{the universal image preserves set membership}.
+\begin{definition}[Well-formedness of $\sim$ for a Class $\trg{\class}$]\label{def:wfc:sig:tracerel}
   % $$
 
+  \begin{nscenter}
   \noindent
   \text{\bul{$\wfcsig{\sim}{\trg{\class}}$}} \isdef \text{\rul{$\forall \trg{\Pi}\in\trg{\class}, \sigma_\sim(\trg{\Pi})\in\sigma_\sim(\trg{\class})$}}
+  \end{nscenter}
   % $$
@@
 % \MPin{why}
-\begin{theorem}[Sequential Composition of Secure Compilers w.r.t. $\sigma$]\label{thm:rtpsim:sig}
-  If \bul{$\rtpsig{\cc{\src{L}}{\trg{L}}}{\tilde{\sigma}_{\sim_2}\left(\trg{\class[_{1}]}\right)}{\sim_1}$}, \rul{$\rtpsig{\cc{\trg{L}}{\obj{L}}}{\trg{\class[_2]}}{\sim_2}$}, and \iul{$\wfcsig{\sim_1}{\trg{\class[_2]}}$}, then \oul{$\rtpsig{\cc{\src{L}}{\trg{L}}\circ\cc{\trg{L}}{\obj{L}}}{\trg{\class[_{1}]}\cap\trg{\class[_{2}]}}{\sim_1\bullet\sim_2}$}.
+\begin{theorem}[Composition of Secure Compilers w.r.t. $\sigma$]\label{thm:rtpsim:sig}
+  $\;$
+
+  If \bul{$\rtpsig{\cc{\src{L}}{\trg{L}}}{\tilde{\sigma}_{\sim_2}\left(\trg{\class[_{1}]}\right)}{\sim_1}$}, \rul{$\rtpsig{\cc{\trg{L}}{\obj{L}}}{\trg{\class[_2]}}{\sim_2}$}, and \iul{$\wfcsig{\sim_1}{\trg{\class[_2]}}$}, \\ then \oul{$\rtpsig{\cc{\src{L}}{\trg{L}}\circ\cc{\trg{L}}{\obj{L}}}{\trg{\class[_{1}]}\cap\trg{\class[_{2}]}}{\sim_1\bullet\sim_2}$}.
 \Coqed
@@
 
-\begin{corollary}[Sequential Composition of Secure Compilers]\label{corr:rtp}
+\begin{corollary}[Composition of Secure Compilers]\label{corr:rtp}
+  $\;$
+
   If \bul{$\rtp{\cc{\src{L}}{\trg{L}}}{\class[_{1}]}$} and \rul{$\rtp{\cc{\trg{L}}{\obj{L}}}{\class[_{2}]}$}, then \oul{$\rtp{\cc{\src{L}}{\trg{L}}\circ\cc{\trg{L}}{\obj{L}}}{\class[_{1}]\cap\class[_{2}]}$}.
@@
 However, while \Cref{thm:rtpsim:sig} explicitly requires that $\sim_1$ is well-formed w.r.t. $\trg{\class[_2]}$, if $\sim_2$ is not well-formed w.r.t. $\trg{\class[_1]}$, care must be taken. 
-Worst-case is that the resulting class $\sigma_{\sim_1\bullet\sim_2}\left(\trg{\class[_1]}\cap\trg{\class[_2]}\right)$ becomes empty, even though $\trg{\class[_1]}\cap\trg{\class[_2]}$ is not necessarily empty.
-\Cref{sec:formalities:compoex} provides an example of this phenomenon.
+%Worst-case is that the resulting class $\sigma_{\sim_1\bullet\sim_2}\left(\trg{\class[_1]}\cap\trg{\class[_2]}\right)$ becomes empty, even though $\trg{\class[_1]}\cap\trg{\class[_2]}$ is not necessarily empty.
+\Cref{subsec:compatsecpasses} provides an example of this phenomenon.
 
@@
 \begin{corollary}[Swapping Secure Compiler Passes]\label{corr:swappable}
+  $\;$
+
   If \bul{$\rtp{\cc[_{1}]{\trg{L}}{\trg{L}}}{\class[_{1}]}$ and $\rtp{\cc[_{2}]{\trg{L}}{\trg{L}}}{\class[_{2}]}$}, then \oul{$\rtp{\cc[_{1}]{\trg{L}}{\trg{L}}\circ\cc[_{2}]{\trg{L}}{\trg{L}}}{\class[_{1}]\cap\class[_{2}]}$ and $\rtp{\cc[_{2}]{\trg{L}}{\trg{L}}\circ\cc[_{1}]{\trg{L}}{\trg{L}}}{\class[_{2}]\cap\class[_{1}]}$}.
@@
 \begin{theorem}[Upper Composition of Secure Compilers]\label{thm:urtp}
+  $\;$
+
   If {$\rtp{\cc{\src{L}}{\trg{L}}}{\class[_{1}]}$} and {$\rtp{\cc{\obj{L}}{\trg{L}}}{\class[_{2}]}$}, then {$\rtp{\cc{\src{L}+\obj{L}}{\trg{L}}}{\class[_{1}]\cap\class[_{2}]}$}. %\Coqed
@@
 \begin{theorem}[Lower Composition of Secure Compilers]\label{thm:lrtp}
+  $\;$
+
   If {$\rtp{\cc{\src{L}}{\trg{L}}}{\class[_{1}]}$} and {$\rtp{\cc{\obj{L}}{\trg{L}}}{\class[_{2}]}$}, then {$\rtp{\cc{\src{L}}{\trg{L}+\obj{L}}}{\class[_{1}]\cap\class[_{2}]}$}. % \Coqed
@@
 While up to \gls*{scct}, none of the properties is novel, for self-containment of the paper, their definition is necessary for the case study.
-For each of the presented properties, the technical appendix accompanying the paper defines corresponding monitors, which refine a given property.
+For each of the presented properties, the technical appendix accompanying the paper defines corresponding monitors, which refine a given property and have a major role in the proofs of this paper.
 
@@
 
-\vspace{-1.5em}
-
+For simple memory safety composed of temporal and spatial memory safety, the trace model defines events as either the empty event ($\emptyevent$), a crash ($\lightning$), or a base-event ($\preevent[_{\mssafe}]$).
+
+\vspace{-1.0em}
 \[
   \begin{array}{rcll}
-    \event[_{ms}] &:=& \ev{Alloc \loc\ n} \mid \ev{Dealloc \loc} \mid \ev{Use \loc\ n} \mid \emptyevent \mid \lightning \\ 
+    \text{(Base-Events)}&\preevent[_{\mssafe}] &:=& \ev{Alloc \loc\ n} \mid \ev{Dealloc \loc} \mid \ev{Use \loc\ n} \\
+    \text{(Events)}&\event[_{\mssafe}] &:=& \preevent[_{\mssafe}] \mid \emptyevent \mid \lightning \\ 
   \end{array}
@@
 
-For simple memory safety composed of temporal and spatial memory safety, the trace model defines events as either the empty event ($\emptyevent$), a crash ($\lightning$), or as tuples consisting of a pre-event and a control-tag.
-Control-tags state whether the context ($\ctx$) or the component ($\comp$) are responsible for emitting the event, a necessary technical detail to know who is to blame for, e.g., an allocation.
-With these control-tags it is possible to ignore effects performed by a spurious context that, e.g., immediately deallocates a memory location twice, thus violating \gls*{tms}~\cite{nagarakatte2010cets}.
+%Control-tags state whether the context ($\ctx$) or the component ($\comp$) are responsible for emitting the event, a necessary technical detail to know who is to blame for, e.g., an allocation.
+%AAAAAAAAAAAAAWith these control-tags it is possible to ignore effects performed by a spurious context that, e.g., immediately deallocates a memory location twice, thus violating \gls*{tms}~\cite{nagarakatte2010cets}.
 %Even though this could be used for blame~\cite{patrignani2023blame}, this is not what this paper is concerned about.
-Pre-events describe the actual kind of event that happened.
+Base-events describe the actual kind of event that happened.
 For the basic memory-safety properties, these are three variants:
@@
 Third, an event to describe reads from and writes to memory ($\ev{Use\ \loc\ n}$).
-%Typical operations with data-dependent timing are branches and certain binary operations, such as division.%
-%\footnote{
-%	This is architecture-dependent, but division is an operation that serves as a classic example for a data-dependent timing instruction~\cite[p.~755]{arm-refman}.
-%}
-%Both operations are also modelled in the specification trace model ($\ev{Branch\ n}$ and $\ev{Binop\ n}$).
 
@@
 
-\gls*{tms} is a safety property that describes that an unallocated object must not be used in any way.
-%Moreover, the property requires that all allocated objects must be deallocated at some point.
+\gls*{tms}~\cite{nagarakatte2010cets} is a safety property that describes that an unallocated object must not be (re-)used.
 
@@
   $$
-  \tmssafe:=\left\{\trace \left| \begin{array}{rcl}
-                              \ev{Alloc\ \loc\ n}&\le_{\trace}&\ev{Dealloc\ \loc} \\
-                              \ev{Use\ \loc\ n}&\le_{\trace}&\ev{Dealloc\ \loc} \\
-                            %  \text{if }\ev{Alloc\ \loc\ n;\sandboxtag}\text{ in }\trace&\text{then}&\ev{Dealloc\ \loc;\sandboxtag}\text{ in }\trace \\
-                              \text{at most one }\ev{Dealloc\ \loc}&\text{in}&\trace \\
-                              \text{at most one }\ev{Alloc\ \loc\ n}&\text{in}&\trace \\
+  \tmssafe:=\left\{\trace[_{\mssafe}] \left| \begin{array}{rcl}
+    \ev{Alloc\ \loc\ n}&\le_{\trace[_{\mssafe}]}&\ev{Dealloc\ \loc} \\
+    \ev{Use\ \loc\ n}&\le_{\trace[_{\mssafe}]}&\ev{Dealloc\ \loc} \\
+    \text{at most one }\ev{Dealloc\ \loc}&\text{in}&\trace[_{\mssafe}] \\
+    \text{at most one }\ev{Alloc\ \loc\ n}&\text{in}&\trace[_{\mssafe}] \\
   \end{array}\right.\right\}
@@
 
-\gls*{sms} prohibits out-of-bounds accesses.
+\gls*{sms}~\cite{nagarakatte2009soft} is a safety property that prohibits out-of-bounds accesses.
 
@@
   \noindent
-  $$
-  \smssafe:=\left\{\trace \left|\begin{array}{rcl}
-                                  \text{If }\ev{Alloc\ \loc\ n}\le_{\trace}\ev{Use\ \loc\ m}, \text{ then }m<n
+  \[
+  \smssafe:=\left\{\trace[_{\mssafe}] \left|\begin{array}{rcl}
+      \text{If }\ev{Alloc\ \loc\ n}\le_{\trace_{\mssafe}}\ev{Use\ \loc\ m}, \text{ then }m<n
   \end{array}\right.\right\}
-                                $$
+  \]
   % \end{nscenter}
@@
 
-Full \gls*{ms} (similar to earlier work~\cite{nagarakatte2009soft,nagarakatte2010cets,jim2002cyclone,necula2005ccured,michael2023mswasm}) is then described as the conjunction of \Cref{def:trace:tmsdef,def:trace:smsdef}.
-%%% Catalin's paper doesn't apply, since their definition of MS relies on noninterference to guarantee data isolation.
-%%% our definition does not guarantee that (on purpose, we want to compose it with cct)
+In spirit of earlier work~\cite{nagarakatte2009soft,nagarakatte2010cets,jim2002cyclone,necula2005ccured,michael2023mswasm}, full \gls*{ms} is the intersection of \Cref{def:trace:tmsdef,def:trace:smsdef}.
 
@@
 
-Note that \Cref{def:trace:msdef} says nothing about memory-safety issues introduced by side-channels.
-
-\subsection{Memory Safety Trace Model with Constant Timeness}
-
+Note that \Cref{def:trace:msdef} ignores data isolation, there may still be memory-safety issues introduced by side-channels.
+
+\subsection{Memory Safety Trace Model with Constant Timeness}\label{subsec:scct:tracemodel}
+
+This section starts off by extending the memory safety trace model (\Cref{subsec:basic:memsafety:tracemodel}) with a {\em security tag} ($\securitytag$) that indicates whether the involved event contains sensitive information ($\lock$) or not ($\unlock$).
+
+\vspace{-1em}
 \[
   \begin{array}{rrcl}
-    (\text{Pre-Event}) & \event[^b_{ct}] &:=& \ev{Alloc \loc\ n} \mid \ev{Dealloc \loc} \mid \ev{Use \loc\ n}\\
-                       &&\mid& \ev{Branch\ }n \mid \ev{Binop\ }n\\
-    (\text{Security Tag}) & \securitytag &:=& \lock \mid \unlock\\ 
-    (\text{Event}) & \event[_{ct}] &:=& \event[^b_{ct}];\securitytag \mid \emptyevent \mid \lightning \\ 
+    (\text{Base-Events}) & \preevent[_{\ctsafe}] &:=& \preevent[_{\mssafe}] \mid \ev{Branch\ }n \mid \ev{Binop\ }n\\
+    (\text{Security Tags}) & \securitytag &:=& \lock \mid \unlock\\ 
+    (\text{Events}) & \event[_{\ctsafe}] &:=& \preevent[_{\ctsafe}];\securitytag \mid \emptyevent \mid \lightning \\ 
   \end{array}
@@
 
-This section starts off by extending the basic memory safety trace model (\Cref{subsec:basic:memsafety:tracemodel}) with a {\em security tag} ($\securitytag$) that indicates whether the involved event contains sensitive information ($\lock$) or not ($\unlock$).
 For cryptographic code, there is a general guideline that secrets must not be visible on a trace, i.e., marked as $\unlock$.
@@
 }
-Both operations are also modelled in the trace model by extending the set of pre-events with branches on some value $n$ ($\ev{Branch\ n}$) and binary operations with result $n$ ($\ev{Binop\ n}$).
+Both operations are also modelled in the trace model by extending the set of base-events with branches on some value $n$ ($\ev{Branch\ n}$) and binary operations with result $n$ ($\ev{Binop\ n}$).
 
@@
 
-  \noindent$$
-  \scctsafe:=\left\{\trace \left|\begin{array}{rcl}
-                                   \trace=\hole{\cdot} & \text{ or } & \trace=\event[_{b}];\unlock\cdot\trace['] \wedge \trace[']\in\scctsafe
+  \noindent\[
+  \scctsafe:=\left\{\trace[_{\ctsafe}] \left|\begin{array}{l}
+      \trace[_{\ctsafe}]=\hole{\cdot} \text{ or } \\\exists\trace[_{\ctsafe}'],\trace[_{\ctsafe}]=\preevent[_{\ctsafe}];\unlock\cdot\trace[_{\ctsafe}'] \wedge \trace[_{\ctsafe}']\in\scctsafe
     \end{array}\right.\right\}
-  $$
+  \]
 % \end{nscenter}
@@
 
-\gls*{scct} may seem overly strict, since it seems that secrets must not occur on a trace (since $\securitytag$ is forced to be $\unlock$). 
-However, programs that operate on secrets and have a ``data independent timing mode'' enabled do not leak these secrets, despite performing computation on them (see \Cref{ex:ditm}).
+\gls*{scct} may appear overly strict, since it seems that secrets must not occur on a trace (since $\securitytag$ is forced to be $\unlock$). 
+However, this is considered standard practice in terms of coding guidelines~\cite{ctguidelines}.
+Moreover, programs that operate on secrets and have a ``data independent timing mode'' enabled do not leak these secrets, despite performing computation on them (see \Cref{ex:lscct}).
 
@@
 
-The combination of \gls*{ms} and \gls*{scct} is, intuitively, the intersection of these properties, \gls*{msscct}.
-However, \gls*{ms} uses a simpler trace model than \gls*{scct}, so intersecting them would trivially yield the empty set. 
-To remedy this issue, let $\sim_{msscct}$ be a cross-language trace relation that relates the basic memory safety trace model with the memory safety trace model by ignoring security tags and replacing the new pre-events with the empty event:
+The combination of \gls*{ms} and \gls*{scct} is the intersection of these properties, \gls*{msscct}.
+However, \gls*{ms} uses a different trace model than \gls*{scct}, so intersecting them would trivially yield the empty set. 
+To remedy this issue, let $\sim_{\ctsafe}$ be a cross-language trace relation whose definition can be sketched as follows:
 \[
-  \typerulenolabel{mscctrel:drop:tag}{}{\event[_{ms}]\sim_{msscct}\event[^b_{ct}];\securitytag}
+  \typerulenolabel{mscctrel:drop:tag}{}{\preevent[_{\mssafe}]\sim_{\ctsafe}\preevent[_{\mssafe}];\securitytag}
+  \typerulenolabel{mscctrel:drop:crash}{}{\lightning\sim_{\ctsafe}\lightning}
 \]
 \[
-  \typerulenolabel{mscctrel:drop:branch}{}{\emptyevent\sim_{msscct}\ev{Branch n};\securitytag}
-  \typerulenolabel{mscctrel:drop:binop}{}{\emptyevent\sim_{msscct}\ev{Binop n};\securitytag}
+  \typerulenolabel{mscctrel:drop:branch}{}{\emptyevent\sim_{\ctsafe}\ev{Branch n};\securitytag}
+  \typerulenolabel{mscctrel:drop:binop}{}{\emptyevent\sim_{\ctsafe}\ev{Binop n};\securitytag}
 \]
 
-It is now possible to define \gls*{msscct} using the universal image:
-
-%Since \gls*{ms} and \gls*{scct} are just sets of traces that, intuitively, contain all program behaviors that follow a security policy, the intersection of them contains all program behaviors that follow both security policies, i.e., it entails all program behaviours that are both \gls*{ms} and \gls*{scct}.
+Essentially, the only work $\sim_{\ctsafe}$ has to do, is to ignore both the new $\ev{Branch\ n}$ and $\ev{Binop\ n}$ base-events as well as the security tag, since they are irrelevant to \gls*{ms}.
+The relation is extended from just events to traces in a standard way, skipping the empty event $\emptyevent$ on either side, and it is now possible to define \gls*{msscct} using the universal image:
 
 \begin{definition}[\glsfirst*{ms} and \gls*{scct}]\label{def:trace:msscctdef}
-  \[
+  $
   \msscctsafe:=\sigma_{\sim_{msscct}}\left(\mssafe\right)\cap\scctsafe
-  \]
+  $
 \end{definition}
 
-\subsection{Memory-Safe, Constant-Time Trace Model with Speculation}\label{subsec:msctss:tracemodel}
-
+\subsection{Memory-Safety, Constant-Time Trace Model with Speculation}\label{subsec:msctss:tracemodel}
+
+So far, the considered trace models do not take speculative execution into account. 
+For this, extend the earlier trace model (see \Cref{subsec:scct:tracemodel}) so that the security tags ($\securitytag$) carry additional information about the kind of private data leakage, i.e., the type of speculative leak.
+Moreover, add pre-events signaling the beginning of a speculative execution ($\ev{Spec}$), a barrier ($\ev{Barrier}$) that signals that any speculative execution may not go past it, as well as a rollback event ($\ev{Rlb}$), which signals that execution resumes to where speculation started.
+
+\vspace{-1em}
 \[
   \begin{array}{rrcl}
-    (\text{Pre-Event}) & \event[^b_{\mathghost}] &:=& \ev{Alloc \loc\ n} \mid \ev{Dealloc \loc} \mid \ev{Use \loc\ n}\\
-                       &&\mid& \ev{Branch\ }n \mid \ev{Binop\ }n \mid \ev{Spec} \mid \ev{Rlb}\\
-    (\text{Spectre Variant}) & vX &:=& \operatorname{PHT} \\
-    (\text{Security Tag}) & \securitytag &:=& \lock_{vX} \mid \unlock\\ 
-    (\text{Event}) & \event[_{\mathghost}] &:=& \event[^b_{\mathghost}];\securitytag \mid \emptyevent \mid \lightning \\ 
+    (\text{Base-Events}) & \preevent[_{\mathghost}] &:=& \preevent[_{\ctsafe}] \\
+    (\text{Spectre Variants}) & vX &:=& \operatorname{NONE} \mid \operatorname{PHT} \\
+    (\text{Security Tags}) & \securitytag &:=& \lock_{vX} \mid \unlock\\ 
+    (\text{Events}) & \event[_{\mathghost}] &:=& \preevent[_{\mathghost}];\securitytag \mid \emptyevent \mid \lightning \mid \ev{Spec} \mid \ev{Rlb} \mid \ev{Barrier}\\ 
   \end{array}
@@
 
-So far, the considered trace models do not take speculative execution into account. 
-For this, extend the memory safety trace model so that the security tags ($\securitytag$) carry additional information about the kind of private data leakage, i.e., the type of speculative leak.
-Moreover, add pre-events signaling the beginning of a speculative execution ($\ev{Spec}$) as well as a rollback event ($\ev{Rlb}$).
+Even though the considered Spectre variants are just SPECTRE-PHT~\cite{kocher2019spectre}, NONE just describes secret data as in \gls*{scct} (see \Cref{subsec:scct:tracemodel}), the trace model is general enough to allow for potential future extension with different variants.
+%For sake of readability, this paper just uses the notation $\lock$ in place of $\lock_{\text{PHT}}$.
 
@@
 \gls*{ss}~\cite{patrignani2021exorcising}, similar to \gls*{scct}, is an overapproximation to a variant of noninterference.
-The key difference is just the underlying trace model, while the definition of the property itself turns out to be syntactically equal to \gls*{scct}.
+The definition of the property itself turns out to be syntactically equal to \gls*{scct}, modulo the underlying trace model.
 
@@
   \noindent\[
-  \sssafe:=\left\{\trace \left|\begin{array}{rcl}
-                                   \trace=\hole{\cdot} & \text{ or } & \trace=\event[_{b}];\unlock\cdot\trace['] \wedge \trace[']\in\sssafe
+    \sssafe:=\left\{\trace[_{\mathghost}] \left|\begin{array}{rcl}
+        \trace[_{\mathghost}]=\hole{\cdot} & \text{ or } & \trace[_{\mathghost}]=\preevent[_{\mathghost}];\unlock\cdot\trace[_{\mathghost}'] \wedge \trace[_{\mathghost}']\in\sssafe
                                  \end{array}\right.\right\}
@@
 Similar to \gls*{msscct}, there is a need to relate the different trace models with each other, so that the memory safety property without speculation can be lifted to speculation. 
-To this end, let $\sim_{spec}$ be a cross-language trace relation that relates the non-speculative memory safety trace model with the one that has speculation by simply ignoring any events related to speculation and ignoring the additional variant in the security-tag:
+To this end, let $\sim_{\mathghost}$ be a cross-language trace relation whose definition can be sketched as follows:
+
 \[
-  \typerulenolabel{specrel:drop:spec}{}{\emptyevent\sim_{spec}\ev{Spec};\securitytag}
-  \typerulenolabel{specrel:drop:rlb}{}{\emptyevent\sim_{spec}\ev{Rlb};\securitytag}
+  \typerulenolabel{specrel:drop:tag}{}{\preevent[_{\ctsafe}];\lock\sim_{\mathghost}\preevent[_{\ctsafe}];\lock_{\text{NONE}}}
+  \typerulenolabel{specrel:drop:spectag}{}{\emptyevent\sim_{\mathghost}\preevent[_{\mathghost}];\lock_{\text{PHT}}}
 \]
 \[
-  \typerulenolabel{specrel:drop:tag}{}{\event[_b];\lock\sim_{spec}\event[_b];\lock_{vX}}
+  \typerulenolabel{specrel:drop:pub}{}{\preevent[_{\ctsafe}];\unlock\sim_{\mathghost}\preevent[_{\ctsafe}];\unlock}
+  \typerulenolabel{specrel:drop:crash}{}{\lightning\sim_{\mathghost}\lightning}
 \]
-
-With this, define \gls*{specms} as follows:
+\[
+  \typerulenolabel{specrel:drop:spec}{}{\emptyevent\sim_{\mathghost}\ev{Spec}}
+  \typerulenolabel{specrel:drop:rlb}{}{\emptyevent\sim_{\mathghost}\ev{Rlb}}
+  \typerulenolabel{specrel:drop:barrier}{}{\emptyevent\sim_{\mathghost}\ev{Barrier}}
+\]
+
+The intuition is that speculation is ignored, which amounts to dropping events $\ev{Spec}$, $\ev{Rlb}$, or $\ev{Barrier}$, as well as all base events tagged with $\lock_{\text{PHT}}$. 
+Now, define \gls*{specms} as follows:
 \begin{definition}[\glsfirst*{specms}]\label{def:trace:specmsdef}
-  \[
-  \specmssafe=\sigma_{\sim_{spec}}\left(\msscctsafe\right)\cap\sssafe
-  \]
+  $
+  \specmssafe=\sigma_{\sim_{\mathghost}}\left(\msscctsafe\right)\cap\sssafe
+  $
 \end{definition}
@@
 
-\subsection{Compatibility of Secure Compiler Passes}
+\subsection{Compatibility of Secure Compiler Passes}\label{subsec:compatsecpasses}
 

### 2024-05-24 csf2/main.tex
do more todos

@@
   \item %
-        %\Cref{sec:compcomp} studies different forms of secure compiler compositions.
+        %\Cref{sec:sequential} studies different forms of secure compiler compositions.
         This paper takes the secure compilation framework of Abate et al.~\cite{abate2021extacc} and extends it for compositionality. (\Cref{sec:sequential}).
@@
 A class is simply a set of properties and for the class of safety properties, it is decidable whether a trace satisfies a safety property with just a finite trace prefix.
-The compositionality framework (\Cref{sec:compcomp}) presented in this paper considers arbitrary classes, while the case-study (\Cref{sec:casestud:rtp}) fixes them to concrete instances.
+The compositionality framework (\Cref{sec:sequential}) presented in this paper considers arbitrary classes, while the case-study (\Cref{sec:casestud:rtp}) fixes them to concrete instances.
 As an example, consider a trace describing an interaction with a memory where the deallocation of an address $\loc$ precedes a read at that address in memory: $\ev{Dealloc\ \loc}\cdot\ev{Read\ \loc\ 1729}\cdot\dots$.
@@
 
-This section defines programming languages that the secure compilers defined in the next section will use.
-To this end, this section defines the languages $\Ltms$, $\Ltrg$, $\Lms$, $\Lscct$, and $\Lspec$ which share many common elements (presented in \Cref{subsec:cs:defs}).
+This section defines programming languages $\Ltms$, $\Ltrg$, $\Lms$, $\Lscct$, and $\Lspec$, all of which share common elements (presented in \Cref{subsec:cs:defs}).
 $\Ltms$ is the only statically typed language and exhibits the property that all well-typed programs are \gls*{tms} (\Cref{subsec:ltms}).
@@
 $\Lms$ is exactly the same language as $\Ltrg$, but this paper still distinguishes the two for sake of readability (\Cref{subsec:lms}).
-All three languages --- so $\Ltms$, $\Ltrg$, and $\Lms$ --- assume \gls*{cct} to hold.
+All three languages --- so $\Ltms$, $\Ltrg$, and $\Lms$ --- assume \gls*{cct} to hold, since this is -- in an ideal world -- what the programmer would expect, too: It is the job of the compiler to preserve and (potentially) enforce security~\cite{cauligi2019fact}.
 
 % This paper considers \gls*{cct} as an architecture-dependent property\MPin{can we justify this}, so
-Writing code attaining \gls*{cct} should not be of the programmer's concerns~\cite{cauligi2019fact}.
 Such consideration is also backed up by architecture providing a data (operand) independent timing mode, such as processors by Arm~\cite[p.~543]{arm-refman} and Intel~\cite[p.~80]{intel-refman}.
-In spirit of this, all languages besides $\Lscct$ are trivially \gls*{scct}. 
+%In spirit of this, all languages besides $\Lscct$ are trivially \gls*{scct}. 
 $\Lscct$ (\Cref{subsec:lscct}) programs have access to a ``\gls*{cct}-mode'' and can change the leakage of emitted events according to the value of this mode (either $\obj{ON}$ or $\obj{OFF}$).
 
-Lastly, modern processors also employ speculative execution to achieve speedups, but fail to rollback properly, so that data may be left over at a cache, so that an attacker can exploit the difference in runtime for cache-hits versus cache-misses.
-$\Lspec$ extends $\Lscct$ to model this situation, while all previous languages trivially satisfy \gls*{ss}, since they do not support speculative execution at all.
+Lastly, modern processors also employ speculative execution to achieve speedups, but fail to roll back properly, so that an attacker can exploit side-channels~\cite{kocher2019spectre}. 
+$\Lspec$ (\Cref{subsec:lspec}) extends $\Lscct$ to model this situation, while all previous languages trivially satisfy \gls*{ss}, since they do not support speculative execution at all.
+All languages share large parts of syntactic elements, which are presented in the next subsection (\Cref{subsec:cs:defs}).
+Subsequent subsections extend on this accordingly.
 
@@
 This is a standard technique in secure compilation in order to rule out irrelevant context-level events. 
-For example, a context immediately deallocating an allocated memory region twice trivially violates memory safety, but the main interest in secure compilation is at component-level events and whether those signal, e.g., a memory safety issue.
+For example, a context immediately deallocating an allocated memory region twice trivially violates memory safety, but the main interest in secure compilation is at component-level events and if those signal, e.g., a memory safety issue.
 
@@
 Pointers are left second-class and can only be referenced by an identifier, so addresses cannot be guessed by contexts.
-Programs have sets of pre-determined functions called libraries and it is assumed that there are functions part of some component or context.
+Programs have sets of pre-determined functions called libraries and it is assumed that functions are marked as being part of some component or context.
+For sake of simplicity, contexts may not pass pointers to the component, but only integers.
+%Extending this is orthogonal to this work.
 
@@
                          &&&\mid \src{\lcall{f}{\expr}} \mid \src{\lreturn{\expr}} \\
-    \text{(Libraries)} &&& \src{\library_{\ctx}} : \src{Vars} \to \src{\expr} \hspace{2em}
-    \src{\library_{\comp}} : \src{Vars} \to \src{\expr} \\
+    \text{(Functions)} & \src{F} &:=& (\src{x};\src{\natt\to\natt};\src{\expr}) \\
+    \text{(Libraries)} &&& \src{\library_{\ctx}} : \src{Vars} \to \src{F} \hspace{2em}
+    \src{\library_{\comp}} : \src{Vars} \to \src{F} \\
   \end{array}
@@
 The type system of $\src{L_{\tmssafe}}$ enforces that every well-typed $\src{L_{\tmssafe}}$ program satisfies \gls*{tms} (\Cref{thm:wt:tms}).
-%Hereby, the map $\thelocmap(\src{\loc})=\loc$ maps an $\Ltms$ location to a location $\loc$ of the specification trace model.
-
-% Most rules of the projection $\msfilterLtms{\src{\event}}$ are left out 
-% since, for the most part, it does the expected, e.g., $\msfilterLtms{\src{\ev{Dealloc}\ \loc;\comp;\securitytag}}=\ev{Dealloc}\ \delta(\loc);\comp;\securitytag$.
-% But, it also filters any action that a context does as well as the interaction events, since these are irrelevant for component-level \gls*{tms}.
-% \MP{rewrite these lines}
 
 \begin{theorem}[$\src{L_{\tmssafe}}$-programs are \gls*{tms}]\label{thm:wt:tms}
+  $\;$
+
+  \begin{nscenter}
   For any component $\src{\library_{\comp}}$, $\rsat{\src{\library_{\comp}}}{\tmssafe}$ \Coqed
+  \end{nscenter}
 \end{theorem}
 
-The trace model is largely similar to the memory safety trace model (\Cref{subsec:basic:memsafety:tracemodel}).
 $\Ltms$ dynamic semantics do not get stuck if a memory-safety violation is encountered, such as a write out of bounds.
+The trace model is largely similar to the memory safety trace model (\Cref{subsec:basic:memsafety:tracemodel}), where the only real differences are separate events for reading and writing memory, instead of just using it, and the sandboxtags.
+With these tags, it is possible to ignore effects performed by a spurious context that, e.g., immediately deallocates a memory location twice, thus violating \gls*{tms}.
+Even though this could be used for blame preservation~\cite{patrignani2023blame}, this is not what this paper is concerned about.
 
@@
 Furthermore, the syntax of $\Ltrg$ is extended with a way to inspect whether a pointer is freed ($\trg{\lispoisoned{x}}$), evaluating to $\trg{0}$ if it is freed and to $\trg{1}$ otherwise.
+Functions may receive arguments that are not $\trg{\natt}$.
 
@@
 
-No changes are done to the trace model, but $\Ltrg$ has no static typing, making double-free patterns possible.
+No changes are done to the trace model, but $\Ltrg$ has no static typing, making double-free code patterns possible.
 
@@
     \text{(Security Tags)} & \obj{\securitytag} &:=& \obj{\lock} \mid \obj{\unlock} \\
-    \text{(Pre-Event)} & \obj{\preevent} &:=& \dots \mid \obj{iGet\ \loc\ v} \mid \obj{iSet\ \loc\ v} \\
+    \text{(Base-Event)} & \obj{\preevent} &:=& \dots \mid \obj{iGet\ \loc\ v} \mid \obj{iSet\ \loc\ v} \\
                        &&& \mid\ \obj{Binop\ n} \mid \obj{Branch\ n} \\
-    \text{(Event)} & \obj{\event} &:=& \obj{\event[^b];\sandboxtag;\securitytag} \mid \obj{\emptyevent} \mid \obj{\lightning} \\
+    \text{(Event)} & \obj{\event} &:=& \obj{\preevent;\sandboxtag;\securitytag} \mid \obj{\emptyevent} \mid \obj{\lightning} \\
     \text{(Expressions)} & \obj{\expr} &:=& \dots \mid \obj{\lrddoit{x}{\expr}} \mid \obj{\lwrdoit{D}} \\
@@
 Security tags $\obj{\securitytag}$ are on the usual secrecy lattice, where $\obj{\lock}\le\obj{\securitytag}$ and $\obj{\securitytag}\le\obj{\unlock}$.
-To prevent secrets from leaking but still enable reasoning about memory safety, $\Lscct$ extends pre-events with $\obj{\ev{iGet}\ \loc\ \valueexpr}$ and $\obj{\ev{iSet}\ \loc\ \valueexpr}$.
+To prevent secrets from leaking but still enable reasoning about memory safety, $\Lscct$ extends base-events with $\obj{\ev{iGet}\ \loc\ \valueexpr}$ and $\obj{\ev{iSet}\ \loc\ \valueexpr}$.
 This way, memory accesses to secret data are still exposed as memory accesses, despite being protected by data-independent timing mode, e.g, \Cref{tr:to-e-get-in-noleak}.
 Moreover, the language extends pre-events with $\obj{\ev{Branch}\ n}$ and $\obj{\ev{Binop}\ n}$ that are emitted when evaluating a branch or certain binary expressions, such as division, respectively, whenever the data independent timing mode is inactive.
-Events are extended with security-tag to signal the secrecy of the involved data.
+Events are extended with a security-tag to signal the secrecy of the involved data.
 
@@
 Within such windows, data is leaked, marked as high $\ird{\lock}$ and with an annotation that indicates the respective speculative execution variant, in this case just SPECTRE-PHT~\cite{kocher2019spectre}, whose semantics is sketched with \Cref{tr:e-ifz-true-spec}, where $\omega$ is the size of the speculation window.
+Additionally, $\ird{\lock}$ may also carry a $\ird{NONE}$ annotation, to signal that this is a leak irrespective of speculation, i.e., as defined earlier (\Cref{subsec:lscct}).
 The semantics is kept general enough to allow for future extension to support different variants.
+
 In terms of language features, $\Lspec$ supports a barrier operation $\ird{\lbarrier}$ that blocks speculative execution. 
@@
 
-    \draw[thick,dotted,->] (S) to[bend right=33,sloped] node[align=center] {\gls*{ms}\\ \Cref{thm:ccab:rtp:ms}} (M);
+    \draw[thick,dotted,->] (S) to[bend right=33,sloped] node[align=center] {\gls*{ms}\\ \Cref{corr:ccab:rtp:ms}} (M);
     \draw[thick,dotted,->] (M) to[bend right=0,sloped] node[align=center] {\gls*{ms}\\ \Cref{thm:cccfccdce:rtp:ms}} (E);
@@
     Vertices in the graph are the programming languages from earlier sections (\Cref{sec:casestud:defs}). %
-    All edges are secure compilers, but dotted edges use the presented framework (\Cref{sec:compcomp}) and solid edges classic proof techniques. %
+    All edges are secure compilers, but dotted edges use the presented framework (\Cref{sec:sequential}) and solid edges classic proof techniques. %
     The dashed lines partition the graph into the sections where the respective theorems are presented.
@@
 \end{figure*}
-The section demonstrates the power of the framework (\Cref{sec:compprop,sec:compcomp}) by composing these compilers for a secure and optimising compilation chain that robustly preserves \gls*{msscct}.
+The section demonstrates the power of the framework (\Cref{sec:sequential}) by composing these compilers for a secure and optimising compilation chain that robustly preserves \gls*{msscct}.
 The first step in this chain is the compiler from $\src{L_{\tmssafe}}$ to $\trg{L}$ that robustly preserves just \gls*{tms} (\Cref{thm:cca:rtp:tms}).
 From here, an instrumentation from $\trg{L}$ to $\irl{L_{\mssafe}}$ ensures that no out-of-bounds accesses can happen and, thus, programs at this point attain \gls*{sms} (\Cref{thm:ccb:rtp:sms}).
-Since these properties compose into \gls*{ms}, composing these passes yields a compiler that robustly preserves \gls*{ms} (\Cref{thm:ccab:rtp:ms}).
+Since these properties compose into \gls*{ms}, composing these passes yields a compiler that robustly preserves \gls*{ms} (\Cref{corr:ccab:rtp:ms}).
 Then, the section presents two optimising translations, namely \gls*{cf} and \gls*{dce}, each of which robustly preserves \gls*{ms} (\Cref{thm:ccdce:rtp:ms,thm:cccf:rtp:ms}).
 These translations can be freely ordered in the compilation chain without compromising memory safety (\Cref{thm:cccfccdce:rtp:ms}).
-The last step of the chain ensures that code stays \gls*{scct} (\Cref{thm:ccscct:rtp:scct}) when lowered from $\Lms$ to $\Lscct$.
-The final result is that the whole compilation chain robustly preserves \gls*{msscct} (\Cref{thm:ccall:rtp:msscct}).
+The next step in the chain ensures that code stays \gls*{scct} (\Cref{thm:ccscct:rtp:scct}) when lowered from $\Lms$ to $\Lscct$, which is done by switching on a data-independent timing mode for the computation.
+Lastly, by introducing barriers immediately after branches, speculative leaks via SPECTRE-PHT are prevented when compiling $\Lscct$ to $\Lspec$.
+The final result is that the whole compilation chain robustly preserves \gls*{mssscct} (\Cref{thm:ccall:rtp:mssscct}).
 
@@
 To this end, the compiler needs to ensure that when execution switches from context to component, the type signatures are respected.
-It can do so by inserting dynamic typechecks prior to entering the body of a function belonging to the component.
-
-\begin{gather*}
+It can do so by inserting dynamic typechecks prior to entering the body of a function belonging to the component:
+
 % \small
-  \begin{aligned}
-    \cca(\src{x}) = &\ \trg{x} 
+\[
+  \begin{array}{rcl}
+    %\cca(\src{x}) = &\ \trg{x} 
     % \\
-  	&
-  	\qquad
-    \cca(\src{\lbinop{\expr[_{1}]}{\expr[_{2}]}}) = &\ \lbinop[\trg]{\left[\cca(\src{\expr[_{1}]})\right]}{\left[\cca(\src{\expr[_{2}]})\right]} \\
-    \cca(\src{n}) = &\ \trg{n} 
+  	%&
+  	%\qquad
+    %\cca(\src{\lbinop{\expr[_{1}]}{\expr[_{2}]}}) = &\ \lbinop[\trg]{\left[\cca(\src{\expr[_{1}]})\right]}{\left[\cca(\src{\expr[_{2}]})\right]} \\
+    %\cca(\src{n}) = &\ \trg{n} 
     % \\
-    &
-    \qquad
-    \cca(\src{\lget{x}{\expr}}) = &\ \lget[\trg]{\trg{x}}{\left[\cca(\src{\expr})\right]}
-  \end{aligned}
-    \\
-  \begin{aligned}
-    \cca(\src{\ldelete{x}}) = &\ \ldelete[\trg]{\left[\cca(\src{x})\right]} \\
+    %&
+    \cca(\src{\lget{x}{\expr}}) &= &\ \lget[\trg]{\trg{x}}{\left[\cca(\src{\expr})\right]} \\
+    \cca(\src{\ldelete{x}}) &= &\ \ldelete[\trg]{\left[\cca(\src{x})\right]} \\
 %    \cca(\src{\lset{x}{\expr[_{1}]}{\expr[_{2}]}}) & = \lset[\trg]{\left[\cca(\src{x})\right]}{\left[\cca(\src{\expr[_{1}]})\right]}{\left[\cca(\src{\expr_{2}})\right]} \\
-\cca(\src{\lfunction{g}{x:\natt\to\type[_{e}]}{\expr}})  = &\ \lfunction[\trg]{\trg{g}}{\trg{x}}{\lifz[\trg]{\trg{\lhast{x}{\natt}}}{
+\cca(\src{\lfunction{g}{x:\natt\to\type[_{e}]}{\expr}})  &= &\ \lfunction[\trg]{\trg{g}}{\trg{x}}{\lifz[\trg]{\trg{\lhast{x}{\natt}}}{
 \\%
-                                                            &
+                                                         &  &
                                                             \left[\cca(\src{\expr})\right] %
                                                                                                  }{\labort[\trg]}}
-  \end{aligned}
-\end{gather*}
-
-Since $\trg{L}$ has no static typechecks, it could happen that a bogus context $\trg{\library_{\ctx}}$ invokes a callable object accepting a $\src{\natt}$ with $\trg{\lpair{17}{29}}$.
-By inserting the check, the compiler ensures that execution does not proceed in such cases.
-The compiler does not insert other checks and proceeds as the identity function (which in this paper amounts to a simple re-colouring of $\src{L_{\tmssafe}}$ to $\trg{L}$ expressions).
+  \end{array}
+\]
+
+The above sketches interesting parts, anything left out is a trivial ,,re-colouring'' from $\src{blue}$ to $\trg{red}$.
+
+Since $\trg{L}$ has no static typing, a bogus context $\trg{\library_{\ctx}}$ can invoke a callable object accepting a $\src{\natt}$ with $\trg{\lpair{17}{29}}$.
+By inserting a check, the compiler ensures that execution does not proceed in such cases.
+%The compiler does not insert other checks and proceeds as the identity function (which in this paper amounts to a simple re-colouring of $\src{L_{\tmssafe}}$ to $\trg{L}$ expressions).
 
 Compiling the \texttt{strncpy} function from \Cref{sec:introduction} with $\cca$, the compiler would in this case ensure that the arguments that are evaluated in the compiled \texttt{strncpy} are valid.
-% \MP{put the proof section in an appendix}
-
-\begin{theorem}[$\cca$ is secure w.r.t. \gls*{tms}]\label{thm:cca:rtp:tms}
+This effectively has no influence on the resulting trace, so a corresponding cross-language trace relation $\sim^{\Ltms}_{\Ltrg}$ is, up to one trace being $\src{blue}$ and the other $\trg{red}$, defined as equality.
+There is only one catch: $\sim^{\Ltms}_{\Ltrg}$ ignores irrelevant context actions, since there is no interest in contexts who are to blame for a security violation {\em inside the context itself}!
+Subsequent cross-language trace relations for the other compilers do it similarly, the explicit details can be found in the technical report.
+
+$\cca$ is robustly preserving (\Cref{def:rtp}) \gls*{tms}:
+\begin{theorem}[$\cca$ secure w.r.t. \gls*{tms}]\label{thm:cca:rtp:tms}
     $\rtp{\cca}{\tmssafe}$ % \Coqed
@@
 
-% \subsubsection{Proving Robust Safety Property Preservation}
-
-% We illustrate the proof of \Cref{thm:cca:rtp:tms} since the other secure compilation proofs of this paper follow the same approach.
-% Unfolding the theorem statement yields the following assumptions:  
-%   for any $\pi\in\lceil\tmssafe\rceil$\footnote{$\lceil\cdot\rceil$ lifts the property to a hyperproperty by applying the powerset operation~\cite{clarkson2008hyper}.}, 
-%   $\src{\trace}$, 
-%   $\src{\runtimetermvar}$, 
-%   and component $\src{\library_{comp}}$, 
-%   we have that $\rsat{\src{\library_{comp}}}{\pi}$ and 
-%   $\progstepto{\prog[\trg]{\trg{\library_{\ctx}}}{\cca(\src{\library_{\comp}})}}{\trg{\runtimetermvar}}{\trg{\trace}}$, where $\trg{\library_{\ctx}}$ is arbitrary.
-% The proof obligation is $\msfilterL{\trg{\trace}}\in\pi$, i.e., the specification trace associated to $\trg{\trace}$ satisfies the property $\pi$.
-% A way to show this is to relate trace $\trg{\trace}$ to some $\Ltms$ trace $\src{\trace}$ (which already satisfies the property as per the assumptions).
-% The assumptions already contain a target execution associated to this trace, so the task is to find an associated $\Ltms$ execution that yields $\trg{\trace}$.
-% The trace $\trg{\trace}$ is split into different parts, as commonly done in secure compilation works~\cite{korashy2021capableptrs,abate2018goodcomps}, where each part contains the events that either the context or the component does, but not both.
-% Because of this, all such trace segments are ,,well-bracketed'' in the sense that they start with either $\trg{\ev{Start}}$, $\trg{\ev{Call}\ \comptoctx\ foo\ \valueexpr}$, or $\trg{\ev{Ret}\ \comptoctx\ \valueexpr}$ and end with either $\trg{\ev{End}\ \valueexpr}$, $\trg{\ev{Ret}\ \ctxtocomp\ \valueexpr}$, or $\trg{\ev{Call}\ \ctxtocomp\ foo\ \valueexpr}$.
-% In the following, the former is referred to as a context segment, since these executions happen in $\trg{\library_{\ctx}}$, and the latter is referred to as a component segment, since these executions happen in $\cca(\src{\library_{\comp}})$.
-% \Cref{fig:proofdiag:rtp} visualises this division for a program execution with one call from context to component and how the target execution is related to a source execution.
-% In the figure, the green dashed lines encompass the component segments while the orange boxes contain the actual context switches from context to component or vice versa.
-% From a technical perspective, as typically done in compilation proofs, the proof requires some setup to maintain a relation between $\src{\Omega}$ and $\trg{\Omega}$ .
-% Two cross-language relations make this precise: (i) $\multimap_{\delta}$ relates states that are involved in a context segment, allowing the target execution to perform internal calls, and (ii) $\approx_{\delta}$ relates states that are involved in a component segment, where both states need to agree exactly, i.e., the memory and the control flow states are required to contain the same information.
-% The relations are indexed with $\delta$, which is an injective mapping from $\Ltms$ locations $\src{\loc}$ to $\Ltrg$ locations $\trg{\loc}$.
-% %Note that the relations $\multimap_{\delta}$ and $\approx_{\delta}$ swap when context switching.
-
-% So far, the paper explained how to relate an $\Ltms$ execution with a $\Ltrg$ execution.
-% The next question is therefore how to build the corresponding $\Ltms$ execution.
-% This is done using a standard secure compilation proof technique called trace-based backtranslation~\cite{abate2019jour,korashy2021capableptrs,patrignani2021rsc}, which can be used to build a context $\src{\library_{\ctx}}$ that behaves similar to $\trg{\library_{\ctx}}$.
-% % Since $\Ltms$ is statically typed, the context $\src{\library_{\ctx}}$ obtained from the backtranslation needs to be well-typed when linked with $\src{\library_{\comp}}$.
-% For context segments of the trace $\trg{\trace}$ it is also necessary to show that the execution behaves similarily, i.e., the context obtained from the backtranslation generates trace $\src{\trace}$.
-% For component segments of the trace, the relatedness of states and traces follows from a compiler correctness argument.
-% These two arguments yield the source execution $\progstepto{\src{\prog{\library_{\ctx}}{\library_{\comp}}}}{\src{\runtimetermvar}}{\src{\trace}}$.
-
-% The proof now works as follows.
-% Given that $\msfilterL{\trg{\trace}}=\msfilterLtms{\src{\trace}}$, the proof goal changes from $\msfilterL{\trg{\trace}}\in\pi$ to $\msfilterLtms{\src{\trace}}\in\pi$.
-% This follows by specializing the robust satisfaction assumption $\rsat{\src{\library_{\comp}}}{\pi}$ to use the context $\src{\library_{\ctx}}$, which is obtained from the backtranslation, and to use the source execution $\progstepto{\src{\prog{\library_{\ctx}}{\library_{\comp}}}}{\src{\runtimetermvar}}{\src{\trace}}$.
-
-% \begin{figure*}[!ht]
-%   \begin{center}
-%     \begin{tikzpicture}[state/.style={minimum height=1.5cm}]
-%       % relative horizontal/vertical distance between states
-%       \pgfmathsetmacro{\hdist}{0.95}
-%       \pgfmathsetmacro{\vdist}{1.35}
-%       \pgfmathsetmacro{\halfvdist}{0.725}
-
-%       % row of src states
-%       \node[state] (srcempty) {$\rtt{\src{\emptyset}}{\src{\lcall{main}{0}}}$};
-%       \foreach \s [remember=\s as \cur (initially empty)] in {1,w_1,p,w_2,2} {
-%         \node[state,right=\hdist of src\cur] (src\s) {$\rtt{\src{\Omega_{\s}}}{\src{\expr[_{\s}]}}$};
-%       }
-%       % row of trg states
-%       \node[state,below=\vdist of srcempty] (trgempty) {$\rtt{\trg{\emptyset}}{\trg{\lcall{main}{0}}}$};
-%       \foreach \s in {1,w_1,p,w_2,2} {
-%         \node[state,below=\vdist of src\s] (trg\s) {$\rtt{\trg{\Omega_{\s}}}{\trg{\expr[_{\s}]}}$};
-%       }
-%       %% illustrations
-%         % backtrans wrapper 1
-%         \draw[ultra thick,loosely dotted,Peach!50,rounded corners] (src1.north east) -- (srcw\string_1.north east)
-%           -- (trgw\string_1.south east) -- (trg1.south west) -- (src1.north west) -- cycle;
-%         % backtrans wrapper 2
-%         \draw[ultra thick,loosely dotted,Peach!50,rounded corners] (srcp.north east) -- (srcw\string_2.north east)
-%           -- (trgw\string_2.south east) -- (trgp.south west) -- (srcp.north west) -- cycle;
-%         % compiler correctness
-%         \draw[ultra thick,loosely dashed,Emerald!50,rounded corners] ($(srcw\string_1.north east)+(0,0.05)$) -- ($(srcp.north east)+(0,0.05)$)
-%           -- ($(trgw\string_2.north east)+(0.05,0)$) -- ($(trgw\string_2.south east)+(0.05,-0.05)$)
-%           -- ($(trg1.south west)+(-0.05,-0.05)$) -- ($(trg1.north west)+(-0.05,0)$) -- ($(srcw\string_1.north west)+(0,0.05)$) -- cycle;
-%       % state relations
-%       \path (srcempty) edge[very thick,draw=gray!25] node[pos=0.5,sloped,rotate=180,fill=white] {\scriptsize$\multimap_\emptyset$} (trgempty)
-%         (src1) edge[very thick,draw=gray!25] node[pos=0.5,sloped,rotate=180,fill=white] {\scriptsize$\multimap_{\delta_1}$} (trg1)
-%         (srcp) edge[very thick,draw=gray!25] node[pos=0.5,sloped,fill=white] {\scriptsize$\approx_{\delta_p}$} (trgp)
-%         (src2) edge[very thick,draw=gray!25] node[pos=0.5,sloped,rotate=180,fill=white] {\scriptsize$\multimap_{\delta_2}$} (trg2)
-%         (srcw\string_1) edge[very thick,draw=gray!25] node[pos=0.5,sloped,fill=white] {\scriptsize$\approx_{\delta_{w_1}}$} (trgw\string_1)
-%         (srcw\string_2) edge[very thick,draw=gray!25] node[pos=0.5,sloped,rotate=180,fill=white] {\scriptsize$\multimap_{\delta_{w_2}}$} (trgw\string_2)
-%         % diagonals
-%         (srcw\string_1) edge[thick,draw=gray!25] node[pos=0.16,sloped,rotate=180,fill=white] {\scriptsize$\approx_{\delta_{w_1}}$} (trg1)
-%         (src1) edge[thick,draw=gray!25] node[pos=0.16,sloped,rotate=180,fill=white] {\scriptsize$\multimap_{\delta_1}$} (trgw\string_1)
-%         (srcp) edge[thick,draw=gray!25] node[pos=0.2,sloped,fill=white] {\scriptsize$\approx_{\delta_p}$} (trgw\string_2)
-%         (trgp) edge[thick,draw=gray!25] node[pos=0.2,sloped,fill=white] {\scriptsize$\multimap_{\delta_{w_2}}$} (srcw\string_2)
-%         ;
-%       %\drawpolygon src1,srcw\string_1,trgw\string_1,trg1;
-%       %\drawpolygon srcp,srcw\string_2,trgw\string_2,trgp;
-%       %\node[font=\tiny,align=center,above=0.2 of srcw1srcp] (wrapper) {Backtranslation\\Wrapper};
-%       %\path[->,draw] (wrapper) -- (srcw\string_1);
-%       %\path[->,draw] (wrapper) -- (srcp);
-%       % steps
-%       \path[very thick,color=\stlccol] (srcempty) edge[draw=none] node {\ $\xrightarrow{\trace[_1]}{}{\kern-3.5pt}_{\text{ctx}}^*$} (src1)
-%         (src1) edge[draw=none] node {\ $\xrightarrow{\trace[_c]}{}{\kern-3.5pt}_{\text{ctx}}$} (srcw\string_1)
-%         (srcw\string_1) edge[draw=none] node {\ $\xrightarrow{\trace[_p]}{}{\kern-3.5pt}_{\text{ctx}}^*$} (srcp)
-%         (srcp) edge[draw=none] node[fill=white,inner sep=0,outer sep=0] {\ $\xrightarrow{\trace[_r]}{}{\kern-3.5pt}_{\text{ctx}}$} (srcw\string_2)
-%         (srcw\string_2) edge[draw=none] node {\ $\xrightarrow{\trace[_2]}{}{\kern-3.5pt}_{\text{ctx}}^*$} (src2)
-%         ;
-%       \path[very thick,color=\ulccol] (trgempty) edge[draw=none] node[fill=white,inner sep=0,outer sep=0] {\ $\xrightarrow{\phantom{\trace[_1]}}{}{\kern-3.5pt}_{\text{ctx}}^*$} (trg1)
-%         (trg1) edge[draw=none] node {\ $\xrightarrow{\phantom{\trace[_p]}}{}{\kern-3.5pt}_{\text{ctx}}^*$} (trgw\string_1)
-%         (trgw\string_1) edge[draw=none] node[fill=white,inner sep=0,outer sep=0] {\ $\xrightarrow{\phantom{\trace[_p]}}{}{\kern-3.5pt}_{\text{ctx}}^*$} (trgp)
-%         (trgp) edge[very thick,draw=none] node {\ $\xrightarrow{\phantom{\trace[_p]}}{}{\kern-3.5pt}_{\text{ctx}}^*$} (trgw\string_2)
-%         (trgw\string_2) edge[draw=none] node {\ $\xrightarrow{\phantom{\trace[_2]}}{}{\kern-3.5pt}_{\text{ctx}}^*$} (trg2)
-%         ;
-%       % legend
-% %     \node[align=left,below right=0.3 and 0.3 of trgempty,font=\tiny] (legend) {%
-% %       ${\src{\trace[_c]}}\cong{\backt{\Ltms}{\Ltrg}(\trg{Call\ ?\ foo\ \valueexpr})}$\\%
-% %       ${\src{\trace[_r]}}\cong{\backt{\Ltms}{\Ltrg}(\trg{Ret\ !\ foo\ \valueexpr})}$
-% %       };
-%       \node[below=0.5 of trgempty,draw=none] (legend) {};
-%       \draw[ultra thick,loosely dotted,Peach!50,rounded corners] ($(legend.north east)+(0.5,-0.4)$) -- ($(legend.north east)+(1,-0.4)$);
-%       \node at ($(legend.north east)+(1.5,-0.4)$) (legendwrapper) {};
-%       \node[right of=legendwrapper] {\tiny Backtranslation Wrapper};
-%       \draw[ultra thick,loosely dashed,Emerald!50,rounded corners] ($(legend.south east)+(0.5,0.4)$) -- ($(legend.south east)+(1,0.4)$);
-%       \node at ($(legend.south east)+(1.5,0.4)$) (legendcorrectness) {};
-%       \node[right of=legendcorrectness] {\tiny Compiler Correctness};
-%     \end{tikzpicture}
-%     \caption{Proof diagram for \Cref{thm:cca:rtp:tms} depicting the general structure of robust preservation proofs. %
-%       Nodes in the graph represent runtime states. %
-%       Vertical lines indicate cross language relations, while horizontal ones are execution steps. %
-%       The green dashed trapezoid encompasses the component segment, while the orange dotted rectangles entail the context switches. %
-%       $\Ltrg$ traces are omitted for readability. %
-%       $\Ltms$ trace segments $\src{\trace[_{c}]}$ and $\src{\trace[_{r}]}$ describe the events that happen at the boundaries, i.e., during a context switch.
-%       $\src{\trace_{p}}$ is the behavior of the component and the traces $\src{\trace[_{1}]}$ and $\src{\trace[_{2}]}$ describe the context.
-%     }\label{fig:proofdiag:rtp}
-%   \end{center}
-% \end{figure*}
-
-% \MPpost{
-% 	this can be made more precise
-% }
-
 \subsection{Robust Spatial Memory Safety Preservation}\label{subsec:cs:ms}
 
-\vspace{-1em}
-\begin{center}
+This subsection defines spatial memory safety preserving compiler from $\Ltrg$ to $\Lms$, $\ccb$.
+The compiler $\cc{\trg{L}}{\Lms}$ only inserts bounds-checks whenever reading from or writing to memory in order to enforce \gls*{sms}.
+These bounds checks need the bounds information, which $\ccb$ keeps around by introducing a fresh identifier $\irl{x_{SIZE}}$ for each allocation that binds $\irl{x}$.
+Then, it is simply a matter of referring to that variable and ensuring that memory accesses are in the interval $[\irl{0},\irl{x_{SIZE}})$.
+For the case where it does not, the compiler inserts a panic.
+
+\begin{nscenter}
 % \small
@@
   $$
-\end{center}
-
-The compiler $\cc{\trg{L}}{\Lms}$ only inserts bounds-checks whenever reading from or writing to memory in order to enforce \gls*{sms}.
-For passing pointers, it has to pass them with their size information as well.
-To this end, the compiler introduces another, fresh identifier $\irl{x_{SIZE}}$ for each allocation that binds $\irl{x}$ to keep track of the allocation size.
+\end{nscenter}
+
 \begin{exampleenv}[Instrumented \texttt{strncpy}]
@@
 void strncpy(size_t n, size_t dst_size, char *dst,
-             size_t src_size, char *src) {
-  for(size_t i = 0; i < src_size && src[i] != '\0' && i < n; ++i) {
-    if(i < src_size && i < dst_size) {
+             size_t src_size, char *src)
+{
+  for(size_t i = 0; i < src_size && src[i] != '\0' && i < n; ++i)
+  {
+    if(i < src_size && i < dst_size)
+    {
       dst[i] = src[i];
@@
     \end{lstlisting}
-    When calling this with a context like in \Cref{ex:strncpy:sms}, the event $\ev{Use}\ \loc_{x}\ 2;\comp;\unlock$ would not be emitted during execution, since the bounds check prevents the condition \texttt{src[i] != '\textbackslash 0'} from executing.
+    Suppose context \texttt{strncpy(2, x, y)}, where \texttt{x} and \texttt{y} are pointers to valid regions of memory with allocated space for exactly two cells and do not contain the null-terminating character \texttt{'\textbackslash 0'}.
+    When calling \texttt{strncpy} like this, the event $\ev{Use}\ \loc_{x}\ 2;\comp;\unlock$ would not be emitted during execution, since the bounds check prevents the condition \texttt{src[i] != '\textbackslash 0'} from executing.
 \end{exampleenv}
 
-\begin{theorem}[$\ccb$ is secure w.r.t. \gls*{sms}]\label{thm:ccb:rtp:sms}
-  $\rtp{\ccb}{\smssafe}$ % \Coqed
+Contrary to $\cca$, $\ccb$ may in fact change the trace of the original $\Ltrg$ program: if there is a memory access, it needs to be protected with a bounds check.
+The corresponding relation $\sim^{\Ltrg}_{\Lms}$ that describes this semantic effect of the compiler may be sketched as follows:
+
+\[
+  \small
+  \typerulenolabel{xrel:sms:read}{n\text{ in bounds}}{\trg{Get\ \loc\ n;\comp}\sim^{\Ltrg}_{\Lms}\irl{Get\ \loc\ n;\comp}}
+  \typerulenolabel{xrel:sms:panic}{n\text{ out of bounds}}{\trg{Get\ \loc\ n;\comp}\sim^{\Ltrg}_{\Lms}\irl{\lightning}}
+\]
+
+And so on for the other, $\trg{Set}$ is analogous and any other event is just a trivial re-colouring.
+Note that neither $\Ltrg$ nor $\Lms$ expose branches on the trace.
+The bounds check is left abstract above, since the precise definition needs some technicalities that only inhibit readability without any technical insights.
+
+It is possible to prove that $\ccb$ is robustly preserving \gls*{sms}.
+\begin{theorem}[$\ccb$ secure w.r.t. \gls*{sms}]\label{thm:ccb:rtp:sms}
+  $\rtpsim{\ccb}{\smssafe}{\sim^{\Ltrg}_{\Lms}}$ % \Coqed
 \end{theorem}
 
-\Cref{thm:ccab:rtp:ms} states that the composition of $\cca$ and $\ccb$ is secure w.r.t. \gls*{ms} and follows from \Cref{thm:cca:rtp:tms,thm:ccb:rtp:sms} using \Cref{thm:rtp}.
-
-\begin{theorem}[$\cca\circ\ccb$ is secure w.r.t. \gls*{ms}]\label{thm:ccab:rtp:ms}
-  $\rtp{\cca\circ\ccb}{\mssafe}$ % \Coqed
-\end{theorem}
-% \begin{proof}\MP{keep?}
-%   From \Thmref{thm:cca:rtp:tms} it follows that for any $\Ltms$ program $\src{\progvar}$, it compiles to an $\Ltrg$ program $\trg{\progvar}$ that robustly satisfies \gls*{tms}.
-%   Note that $\src{\progvar}$ robustly satisfies \gls*{tms} by the properties of the typesystem of $\Ltms$.
-%   Then, \Thmref{thm:ccb:rtp:sms} demonstrates that, assuming $\trg{\progvar}$ robustly satisfies \gls*{sms}, the program $\trg{\progvar}$ compiles to an $\Lms$ program $\irl{p}$ that also robustly satisfies \gls*{sms}.
-%   From \Thmref{thm:ccab:rtp:ms} it follows that $\src{p}$ compiles to $\irl{p}$ that robustly satisfies \gls*{ms}, since \gls*{ms} is the intersection of \gls*{tms} and \gls*{sms}.
-% \end{proof}
+Let $\src{\trace}\sim^{\Ltms}_{\Lms}\irl{\trace}:=\src{\trace}\sim^{\Ltms}_{\Ltrg}\bullet\sim^{\Ltrg}_{\Lms}\irl{\trace}$.
+
+\Cref{corr:ccab:rtp:ms} states that the composition of $\cca$ and $\ccb$ is secure w.r.t. \gls*{ms} and follows from \Cref{thm:cca:rtp:tms,thm:ccb:rtp:sms} using \Cref{thm:rtpsim:sig}.
+
+\begin{corollary}[$\cca\circ\ccb$ secure w.r.t. \gls*{ms}]\label{corr:ccab:rtp:ms}
+  $\;$
+
+  \begin{nscenter}
+    $\rtpsim{\cca\circ\ccb}{\mssafe}{\sim^{\Ltms}_{\Lms}}$ % \Coqed
+  \end{nscenter}
+\end{corollary}
+
+\Cref{thm:rtpsim:sig} has another precondition besides \Cref{thm:ccb:rtp:sms,thm:cca:rtp:tms}: $\sim^{\Ltms}_{\Ltrg}$ needs to be well-formed with respect to $\smssafe$.
+This fact is easily established, since $\sim^{\Ltms}_{\Ltrg}$ is an equality.
+
+\begin{lemma}[$\sim^{\Ltms}_{\Ltrg}$ well-formed w.r.t. $\smssafe$]\label{lem:wf:ltmsltrg}
+  $\wfcsig{\sim^{\Ltms}_{\Ltrg}}{\smssafe}$
+\end{lemma}
 
 \subsection{Optimising Compilers}\label{subsec:cs:opts}
+
+This section defines two optimising compiler passes from $\Lms$ to $\Lms$ which perform \gls*{dce} and \gls*{cf}, respectively.
+The \gls*{dce} pass applies a naive rewrite rule on conditionals.
+The \gls*{cf} pass relies on an auxiliary function \texttt{mix} that uses a substitutions accumulator $\irl{\substlist}$ in order to rewrite constant binary operations, e.g., $\irl{{17}-1}$ to $\irl{16}$, and replace variables that are assigned to constants with their constant, e.g., $\irl{\llet{x}{7}{x}}$ to $\irl{7}$.
 
@@
 
-The two optimising compiler passes from $\Lms$ to $\Lms$ perform \gls*{dce} and \gls*{cf}, respectively.
-The \gls*{dce} pass applies a naive rewrite rule on conditionals.
-The \gls*{cf} pass relies on an auxiliary function \texttt{mix} that uses a substitutions accumulator $\irl{\substlist}$ in order to rewrite constant binary operations, e.g., $\irl{{17}-1}$ to $\irl{16}$, and replace variables that are assigned to constants with their constant, e.g., $\irl{\llet{x}{7}{x}}$ to $\irl{7}$.
-Both passes are secure w.r.t. \gls*{ms}.
-The proof for either is relatively simple, because both \gls*{dce} and \gls*{cf} do not change the way memory accesses happen.
-Moreover, since the input and output languages to these compilers are the same, attacker contexts do not have more power in the target language than in the source, since both languages are equal.
-
-\begin{theorem}[$\ccdce$ is secure w.r.t. \gls*{ms}]\label{thm:ccdce:rtp:ms}
+Note that both passes have no effect on the resulting trace of a program, up to $\emptyevent$-steps. 
+Because of this, both passes have equalit as corresponding cross language tracce relation. 
+Moreover, it is straightforward to prove both passes as secure (\Cref{def:rtp}) w.r.t. \gls*{ms}.
+
+\begin{theorem}[$\ccdce$ secure w.r.t. \gls*{ms}]\label{thm:ccdce:rtp:ms}
   $\rtp{\ccdce}{\mssafe}$ %\Coqed
 \end{theorem}
-\begin{theorem}[$\cccf$ is secure w.r.t. \gls*{ms}]\label{thm:cccf:rtp:ms}
+\begin{theorem}[$\cccf$ secure w.r.t. \gls*{ms}]\label{thm:cccf:rtp:ms}
   $\rtp{\cccf}{\mssafe}$ %\Coqe
@@
 
+This section defines a compiler $\ccscct$ from $\Lms$ to $\Lscct$ that robustly preserves \gls*{scct}. 
+Given the fact that $\Lscct$ provides a \gls*{cct}-mode that can be turned on or off, the compiler inserts wrapper code for function calls and function bodies to ensure that execution in the component always happen in \gls*{cct}-mode.
+
 \vspace{-1em}
-\begin{center}\small
-  \begin{align*}
-    \ccscct(\irl{\lfunction{g}{x}{\expr}}) & = \lfunction[\obj]{\obj{g}}{\obj{x}}{\obj{\lwrdoit{ON};}\ccscct(\irl{\expr})} \\
-    \ccscct(\irl{\lcall{g}{\expr}}) & = \lcall[\obj]{\obj{g}}{\ccscct(\irl{\expr})\obj{; \lwrdoit{ON}}} 
+\[
+\begin{array}{rcl}
+  \ccscct(\irl{\lfunction{g}{x}{\expr}}) & = & \lfunction[\obj]{\obj{g}}{\obj{x}}{\obj{\lwrdoit{ON};}\ccscct(\irl{\expr})} \\
+  \ccscct(\irl{\lcall{g}{\expr}}) & = & \lcall[\obj]{\obj{g}}{\ccscct(\irl{\expr})\obj{; \lwrdoit{ON}}} 
     % \\
@@
     % \\
-  \end{align*}
+\end{array}
+\]
+% \vspace{-2em}
+%
+The context can overwrite the flag and exit the mode, but upon invoking a function that is part of the component, the flag would be set again.
+Because of this, the semantic effect of the compiler, i.e., the corresponding cross-language trace relation $\sim^{\Lms}_{\Lscct}$, only relates events without secrets\footnote{For context actions, it can be whatever.}:
+
+\begin{center} 
+  \typerulenolabel{xrel:scct:noleak}{}{\irl{\preevent;\comp}\sim^{\Lms}_{\Lscct}\obj{\preevent;\comp;\unlock}}
 \end{center}
+
+The compiler is secure w.r.t. \gls*{scct}: %, similarly proven as in \Cref{subsec:cs:tms}.
+
+\begin{theorem}[$\ccscct$ secure w.r.t. \gls*{scct}]\label{thm:ccscct:rtp:scct}
+  \small
+  $\rtpsim{\ccscct}{\scctsafe}{\sim^{\Lms}_{\Lscct}}$ % \Coqed
+\end{theorem}
+
+\subsection{Robust Speculative Safety Preservation}\label{subsec:cs:ss}
+
+This section defines the final compilation pass $\ccspec$, which ensures that $\Lscct$ programs, which are assumed to be \gls*{ss}, stay \gls*{ss} at $\Lspec$-level. 
+To do so, $\ccspec$ inserts a speculation barrier after branches, which is sufficient to harden the program against speculative leaks, since SPECTRE-PHT~\cite{kocher2019spectre} is the only speculative leak modeled in the semantics of $\Lspec$.
+
+\vspace{-1em}
+\[
+\begin{array}{rcl}
+  \ccspec(\obj{\lifz{\expr[_0]}{\expr[_1]}{\expr[_2]}}) & = & \lifz[\ird]{\ccspec{\left(\obj{\expr[_0]}\right)}}{\\&&\ \ \ \ \ird{\lbarrier;}\ccspec{\left(\obj{\expr[_1]}\right)}\\&&}{\\&&\ \ \ \ \ird{\lbarrier;}\ccspec{\left(\obj{\expr[_2]}\right)}} \\
+\end{array}
+\]
 % \vspace{-2em}
 %
-Given the fact that $\Lscct$ provides a \gls*{cct}-mode that can be turned on or off, the compiler inserts wrapper code for function bodies to ensure that execution in the component always happen in this \gls*{cct}-mode.
-The context can overwrite the flag and exit the mode, but upon invoking a function that is part of the component, the flag would be set again.
-Because of this, the compiler is secure w.r.t. \gls*{scct}, similarly proven as in \Cref{subsec:cs:tms}.
-
-\begin{theorem}[$\ccscct$ is secure w.r.t. \gls*{scct}]\label{thm:ccscct:rtp:scct}
-  $\rtp{\ccscct}{\scctsafe}$ % \Coqed
+
+Clearly, the corresponding cross-language trace relation $\sim^{\Lscct}_{\Lspec}$ has only one non-trivial case: For branches, only relate them where speculation is blocked by a barrier:
+
+\begin{center}
+  \typerulenolabel{xrel:spec:if}{}{\obj{Branch\ n}\sim^{\Lscct}_{\Lspec}\ird{Branch\ n}\cdot\ird{Spec}\cdot\ird{Barrier}\cdot\ird{Rlb}}
+\end{center}
+
+In the above, because of space constraints, pretend that each base-event has an additional $\obj{\comp;\securitytag}$, $\ird{\comp;\securitytag}$ annotation, so that full events are related.
+
+$\ccspec$ is secure with respect to \gls*{ss}.
+\begin{theorem}[$\ccspec$ secure w.r.t. \gls*{ss}]\label{thm:ccspec:rtp:spec}
+  \small$\rtpsim{\ccspec}{\sssafe}{\sim^{\Lscct}_{\Lspec}}$ % \Coqed
 \end{theorem}
 
-\subsection{Robust Speculative Safety Preservation}\label{subsec:cs:ss}
-
-\vspace{-1em}
-\begin{center}\small
-  \begin{align*}
-    \ccspec(\obj{\lifz{\expr[_0]}{\expr[_1]}{\expr[_2]}}) & = \lifz[\ird]{\ccspec{\left(\obj{\expr[_0]}\right)}}{\\&\ird{\lbarrier;}\ccspec{\left(\obj{\expr[_1]}\right)}\\&}{\ird{\lbarrier;}\ccspec{\left(\ird{\expr[_2]}\right)}} \\
-  \end{align*}
-\end{center}
-% \vspace{-2em}
-%
-\begin{theorem}[$\ccspec$ is secure w.r.t. \gls*{ss}]\label{thm:ccspec:rtp:spec}
-  $\rtp{\ccspec}{\specmssafe}$ % \Coqed
-\end{theorem}
-
 \subsection{Robust Preservation of Intersection of Memory Safety, Strict Cryptographic Constant Time, and Speculative Safety}
 
-Let $\ccmssscct$ be the compiler that is the composition of $\cca$, $\ccb$, $\cccf$, $\ccdce$, $\ccscct$, and $\ccspec$, then the following theorem holds.
-
-\begin{theorem}[$\ccmssscct$ is secure w.r.t. \gls*{mssscct}]\label{thm:ccall:rtp:mssscct}
-  $\rtp{\cc{\Ltms}{\Lspec}}{\mssafe\cap\scctsafe\cap\sssafe}$ % \Coqed
-\end{theorem}
+Finally, this subsection combines all previous results into one compilation chain to get that it preserves full \gls*{mssscct}.
+Let $\ccmssscct$ be the compiler that is the composition of $\cca$, $\ccb$, $\cccf$, $\ccdce$, $\ccscct$, and $\ccspec$. 
+Let $\sim^{\Ltms}_{\Lspec}$ be the composition of $\sim^{\Ltms}_{\Lms}$, $\sim^{\Lms}_{\Lscct}$, and $\sim^{\Lscct}_{\Lspec}$.
+Then, the following corollary holds.
+
+\begin{corollary}[$\ccmssscct$ secure w.r.t. \gls*{mssscct}]\label{thm:ccall:rtp:mssscct}
+  $\;$
+
+  \begin{nscenter}
+    $\rtpsim{\cc{\Ltms}{\Lspec}}{\mssafe\cap\scctsafe\cap\sssafe}{\sim^{\Ltms}_{\Lspec}}$ % \Coqed
+  \end{nscenter}
+\end{corollary}
+
+As with \Cref{corr:ccab:rtp:ms}, it is important to ensure that the respective cross language trace relations are well-formed (\Cref{def:wfc:sig:tracerel}).
+It is already known that $\sim^{\Ltms}_{\Lms}$ is well-formed with respect to $\mssafe$ (\Cref{lem:wf:ltmsltrg}).
+Next in the chain is $\sim^{\Lms}_{\Lscct}$, which has to be well-formed w.r.t. $\scctsafe$.
+This lemma holds, since a trace that was $\scctsafe$ is $\scctsafe$ even after applying $\sim^{\Lms}_{\Lscct}$: The relation enforces that $\Lscct$ traces related to $\Lms$ traces have no leaks of secrets whatsoever.
+
+\begin{lemma}[$\sim^{\Lms}_{\Lscct}$ well-formed w.r.t. $\scctsafe$]\label{lem:wf:lsmslscct}
+  $\wfcsig{\sim^{\Lms}_{\Lscct}}{\scctsafe}$
+\end{lemma}
+
+The last relation is $\sim^{\Lscct}_{\Lspec}$ which needs to be well-formed w.r.t. $\sssafe$.
+Similarly to the previous relation, this holds, since $\sim^{\Lscct}_{\Lspec}$ only relates $\Lspec$ traces, which do not have speculative leaks, with $\Lscct$ traces.
+
+\begin{lemma}[$\sim^{\Lscct}_{\Lspec}$ well-formed w.r.t. $\sssafe$]\label{lem:wf:lscctlspec}
+  $\wfcsig{\sim^{\Lscct}_{\Lspec}}{\sssafe}$
+\end{lemma}
 
@@
 
-This section provides some formal insights into the case study.
-
-\subsection{Well-formedness of Cross Language Trace Relations}
+This section provides some additional formal insights into the case study. 
+\Cref{subsec:compatsecpasses} discusses a caveat with \Cref{thm:rtpsim:sig}, while \Cref{subsec:seccompproofs} gives some additional background on the secure compilation proofs that do not use the compositionality framework (\Cref{sec:sequential}).
 
@@
 
-\subsection{Secure Compilation Proofs}
+\subsection{Secure Compilation Proofs}\label{subsec:seccompproofs}
 
@@
 The work is missing the upper and lower compositions as well as the generality to support different trace models.
-% The authors demonstrate a similar compositionality theorem to what is presented here (\Cref{sec:compcomp}) as well as in an earlier version of this work~\cite{kruse2022csc}.
+% The authors demonstrate a similar compositionality theorem to what is presented here (\Cref{sec:sequential}) as well as in an earlier version of this work~\cite{kruse2022csc}.
 % However, they do not demonstrate the scalability of the approach by means of a case study.

### 2024-05-24 csf2/main.tex
first draft done

@@
 
-  % \MKin{context}
-  % Memory safety necessitates secrecy, since breaking it can lead to an arbitrarily large attack space, i.e., from simply reading a secret up to alteration of control-flow in order to execute arbitrary code.
-  % However, secrecy cannot be achieved with just memory safety alone, since, e.g., private data could be leaked by differences in execution time for varying program inputs.
-  % This leakage is avoidable by ensuring that programs are cryptographic constant time, i.e., different inputs do not change execution time.
-  % Prior work has shown that strategies to ensure cryptographic constant time can be rendered useless, since a compiler may simply optimize them away.
-  % \MKin{need}
-  % This is why secure compilation is an integral ingredient to achieve security, because a secure compiler guarantees (formally) that properties that hold in the source, also hold in the target.
-%% mention attacker?
-the text describes correctness
-  % \MKin{task}
-  % Unfortunately, it is an open question whether secure compilers can be engineered in a modular way.
-  % Ideally, compiler engineers develop two different compilation passes, one that ensures memory safety and another one that is security preserving w.r.t.
-cryptographic constant time.
-%, because this style of engineering is well-established for its simplifying divide-and-conquer approach as well as enhanced reusability.
-  % \MKin{object+findings}
-  % This paper answers that question positively: The composition of compilers that are secure w.r.t. some security-properties yields the intersection of these properties.
-  % \MKin{object}
-  % This article discusses different variants of compositions of secure compilers and instantiates parts of these results in a case-study.
-  % To this end, it presents an optimizing, secure compilation chain that preserves both memory safety and cryptographic constant time.
-  % \MKin{conclusion}
-  % The compositionality results enable modular proofs for the presented, secure compiler. % that is secure w.r.t. both memory safety and cryptographic constant time.
-  % Some results of this work are implemented in the Coq proof assistant.
-
 \begin{center}\small\it
@@
 This paper uses a more general notion of robust preservation~\cite{abate2021extacc} that applies to compilers that use languages with different trace models. 
-To this end, considering a source trace $\src{\trace}$ and a target trace $\trg{\trace}$, there is a relation ($\sim$) describing the effect of a corresponding compiler. 
-\MKin{reference example from case study}
+To this end, considering a source trace $\src{\trace}$ and a target trace $\trg{\trace}$, there is a relation ($\sim$) describing the effect of a corresponding compiler (see \Cref{sec:casestud:rtp}). 
 Using this relation, there are two projection functions: (1) the existential image $\tau_\sim\left(\src{\pi}\right)$ and (2) the universal image $\sigma_\sim\left(\trg{\pi}\right)$.
@@
 
-%Control-tags state whether the context ($\ctx$) or the component ($\comp$) are responsible for emitting the event, a necessary technical detail to know who is to blame for, e.g., an allocation.
-%AAAAAAAAAAAAAWith these control-tags it is possible to ignore effects performed by a spurious context that, e.g., immediately deallocates a memory location twice, thus violating \gls*{tms}~\cite{nagarakatte2010cets}.
-%Even though this could be used for blame~\cite{patrignani2023blame}, this is not what this paper is concerned about.
 Base-events describe the actual kind of event that happened.
@@
 For example, a context immediately deallocating an allocated memory region twice trivially violates memory safety, but the main interest in secure compilation is at component-level events and if those signal, e.g., a memory safety issue.
-
-% Programs ($\prog{\library_{\ctx}}{\library_{\comp}}$) consist of libraries ($\library$) that belong to a context ($\ctx$) or to a component ($\comp$), and that are lists of functions ($\function$).
-% Functions ($\lfunction{\textit{foo}}{x}{\expr}$) have a single argument $x$ and an expression ($\expr$) as a body which may have $x$ as a free variable.
-% Expressions can be variables ($x$), values ($\valueexpr$), binary operations ($\lbinop{\expr[_1]}{\expr[_2]}$), reading from ($\lget{x}{\expr}$) and writing to ($\lset{x}{\expr[_1]}{\expr[_2]}$) memory, memory allocation ($\llet{x}{\lnew{\expr[_1]}{\expr[_2]}}{\expr[_3]}$) as well as deallocation ($\ldelete{x}$), let expressions ($\llet{x}{\expr[_1]}{\expr[_2]}$), pairs ($\lpair{\expr[_1]}{\expr[_2]}$), their first ($\lproja{\expr}$) and second projection ($\lprojb{\expr}$), conditionals ($\lifz{\expr[_1]}{\expr[_2]}{\expr[_3]}$), function calls ($\lcall{\texttt{foo}}{\expr}$), returns ($\lreturn{\expr}$), and an aborting expression ($\labort$).
 
@@
 This section provides some additional formal insights into the case study. 
+\Cref{subsec:formalities:props} demonstrates that the resulting projected security property using the universal image (\Cref{def:universal:img}) is faithful to the original property. 
 \Cref{subsec:compatsecpasses} discusses a caveat with \Cref{thm:rtpsim:sig}, while \Cref{subsec:seccompproofs} gives some additional background on the secure compilation proofs that do not use the compositionality framework (\Cref{sec:sequential}).
+Lastly, \Cref{subsec:formalities:maps} discusses a minor technicality related to the ,,coloring'' of the investigated properties.
+
+\subsection{Security Properties}\label{subsec:formalities:props}
+
+When modeling some thing using formal methods, the intent is to do so faithfully. 
+That is why it is important to ensure that the properties considered in \Cref{sec:casestud:rtp} follow this ideal.
+It is relatively straightforward to see that, e.g., $\mssafe$ models a basic memory safety property. 
+But, what about $\sigma_{\sim^{\Ltms}_{\Lms}}\left(\mssafe\right)$?
+It is not immediately obvious that the resulting property after projecting it from $\Lms$-level to $\Ltms$-level is similar in spirit to the original $\mssafe$ property (\Cref{def:trace:msdef}) and this is what needs to be discussed when using the presented framework.
+Let us investigate the involved projections from the case study.
+
+\paragraph{$\sigma_{\sim^{\Ltms}_{\Lms}}\left(\mssafe\cap\scctsafe\cap\sssafe\right)$}
+Consider a trace $\irl{\trace}\in\mssafe$ and remember that $\sim^{\Ltms}_{\Lms}$ is the composition of $\sim^{\Ltms}_{\Ltrg}$ and $\sim^{\Ltrg}_{\Lms}$.
+{\em All} traces $\trg{\trace}$ with $\trg{\trace}\sim^{\Ltrg}_{\Lms}\irl{\trace}$ are almost identical to $\irl{\trace}$. 
+They differ in memory accesses and crashes, where $\irl{\lightning}$ can be related to a $\trg{\lightning}$ or a memory event, e.g., $\trg{Get\ \loc\ 2}$.
+In both cases, the memory access cannot be out of bounds.
+
+Now, consider all $\src{\trace}$ with $\src{\trace}\sim^{\Ltms}_{\Ltrg}\trg{\trace}$, which is like equality\footnote{Both $\sim^{\Ltms}_{\Ltrg}$ and $\sim^{\Ltrg}_{\Lms}$ filter context actions.}.
+
+Therefore, the resulting projected property should faithfully model $\mssafe$, which translates to both $\scctsafe$ and $\sssafe$, since these properties trivially hold for $\Ltms$, $\Ltrg$, and $\Lms$ traces.
+
+\paragraph{$\sigma_{\sim^{\Lms}_{\Lscct}}\left(\mssafe\cap\scctsafe\cap\sssafe\right)$}
+Consider a trace $\obj{\trace}\in\scctsafe$ and all $\irl{\trace}$ such that $\irl{\trace}\sim^{\Lms}_{\Lscct}\obj{\trace}$. 
+In turn, $\obj{\trace}$ cannot contain any leaks of secret values, so it cannot violate $\scctsafe$.
+Toggling data-independent timing mode has no impact on $\mssafe$ or $\sssafe$.
+
+\paragraph{$\sigma_{\sim^{\Lscct}_{\Lspec}}\left(\mssafe\cap\scctsafe\cap\sssafe\right)$}
+
+The argument here is analogous to the previous one, introducing $\ird{barrier}$ never violates $\sssafe$ and has no effect on $\mssafe$ or $\scctsafe$.
 
@@
 Even though swapping these compilers in the compilation chain is syntactically impossible, let us pretend that it is possible by just extending $\ccb$ for the constructs present in $\Lscct$ that are not present in $\Ltrg$.
-Clearly, as evidenced in the case study earlier, $\ccb\cdot\ccspec$ is an acceptable composition that yields an interesting class of security properties, i.e., $\sigma_{\sim_{\irl{ms}}\bullet\sim_{\ird{\mathghost}}}\left(\mssafe\cap\scctsafe\cap\sssafe\right)$.
+Clearly, as evidenced earlier (\Cref{subsec:formalities:props}), $\ccb\cdot\ccspec$ is an acceptable composition that yields an interesting class of security properties, i.e., $\sigma_{\sim_{\irl{ms}}\bullet\sim_{\ird{\mathghost}}}\left(\mssafe\cap\scctsafe\cap\sssafe\right)$.
 However, swapping them $\ccspec\cdot\ccb$ means that $\ccb$ inserts new branches into the code that are not necessarily protected by a speculative execution barrier! 
@@
 Since these optimisations have no significant semantic effect and source- and target-language are equal, the backtranslation amounts to an identity translation.
+
+\subsection{Mapping Language-Specific to Property-Specific Traces}\label{subsec:formalities:maps}
+
+Previous sections disregarded a glaring but minor technical issue: Take $\tmssafe$ (\Cref{def:trace:tmsdef}) as an example.
+It contains traces as specified in \Cref{subsec:basic:memsafety:tracemodel}.
+However, these are different from $\Ltms$ traces (\Cref{subsec:ltms}) which means that \Cref{thm:wt:tms} is ,,ill-typed''.
+Similarly for the other theorems.
+In the interest of reducing visual baggage, the notation left out yet another cross-language trace relation, translating language-specific traces into these non-language specific traces. 
+In doing so, context-level actions are dropped and, e.g., both $\src{Get\ \loc\ n}$ and $\src{Set\ \loc\ n}$ are translated into $\ev{Use\ \loc\ n}$.
 
@@
 
-With the advent of speculative execution in processors, it is possible to exploit leftover data in caches to identify secrets.
+With the advent of speculative execution in processors, it is possible to exploit this unusual paradigm identify secrets.
 SPECTECTOR~\cite{guarnieri2018spectector} models speculative execution formally and the present work has taken inspiration in the sematics of $\Lspec$ and in \Cref{def:trace:ss} from SPECTECTOR.

### 2024-05-27 csf2/main.tex
Beautify

@@
 
-\subsubsection{Memory Safe, Strict Cryptographic Constant Time}
+\subsubsection{\gls*{ms}, Strict Cryptographic Constant Time}
 
@@
 
-\subsection{Memory-Safety, Constant-Time Trace Model with Speculation}\label{subsec:msctss:tracemodel}
+\subsection{\gls*{ms}, \gls*{scct} Trace Model with Speculation}\label{subsec:msctss:tracemodel}
 
@@
 \vspace{-1em}
+{
+\small
 \[
@@
 \]
+}
 
@@
 \vspace{-1.5em}
+{\small
 \[
@@
   \begin{nscenter}
-  For any component $\src{\library_{\comp}}$, $\rsat{\src{\library_{\comp}}}{\tmssafe}$ \Coqed
+    \hfill For any component $\src{\library_{\comp}}$, $\rsat{\src{\library_{\comp}}}{\tmssafe}$ \Coqed
   \end{nscenter}
@@
   \typerule{$e-\ird{ifz}-\ird{true}-\text{spec}$}{
-    \ird{\cfstate[']}.\ird{\operatorname{win}} = \omega
+    \ird{\cfstate[']}.\ird{\operatorname{win}} = \omega &
+    \ird{\trace} = \ird{\ev{Branch}\ 0;\sandboxtag;\securitytag}\cdot\ird{\ev{Spec};\sandboxtag;\securitytag}
   }{
-    \pstepto{\rtt{\ird{\cfstate;\sandboxtag;0;\memstate}}{\ird{\lifz{0^{\securitytag}}{\expr[_1]}{\expr[_2]}}}}{\begin{array}{c}\rtt{\ird{\cfstate['];\sandboxtag;0;\memstate}}{\ird{\expr[_1]}}\\\rtt{\ird{\cfstate;\sandboxtag;0;\memstate}}{\ird{\expr[_2]}}\end{array}}{\ird{\ev{Branch}\ 0;\sandboxtag;\securitytag}\cdot\ird{\ev{Spec};\sandboxtag;\securitytag}}
+    \pstepto{\rtt{\ird{\cfstate;\sandboxtag;0;\memstate}}{\ird{\lifz{0^{\securitytag}}{\expr[_1]}{\expr[_2]}}}}{\begin{array}{c}\rtt{\ird{\cfstate['];\sandboxtag;0;\memstate}}{\ird{\expr[_1]}}\\\rtt{\ird{\cfstate;\sandboxtag;0;\memstate}}{\ird{\expr[_2]}}\end{array}}{\ird{\trace}}
   }{e-ifz-true-spec}
@@
 
-% \small
+\vspace{-1em}
+{\small
 \[
@@
 void strncpy(size_t n, size_t dst_size, char *dst,
-             size_t src_size, char *src)
-{
-  for(size_t i = 0; i < src_size && src[i] != '\0' && i < n; ++i)
-  {
-    if(i < src_size && i < dst_size)
-    {
+             size_t src_size, char *src) {
+  for(size_t i = 0; i < src_size
+      && src[i] != '\0' && i < n; ++i) {
+    if(i < src_size && i < dst_size) {
       dst[i] = src[i];
@@
 
-\[
-  \small
+{
+  \footnotesize
+\[
   \typerulenolabel{xrel:sms:read}{n\text{ in bounds}}{\trg{Get\ \loc\ n;\comp}\sim^{\Ltrg}_{\Lms}\irl{Get\ \loc\ n;\comp}}

### 2024-05-27 csf2/main.tex
Space opts

@@
 The previous finite trace prefix is enough to decide that the trace does not satisfy \gls*{tms} and there is no way to append events to this prefix which would result in the trace being admissible.
+
 In the following, the execution of a whole program $\wholeprogvar$ that terminates in state $\runtimetermvar$ according to the language semantics and produces trace $\trace$ is written as $\progstepto{\wholeprogvar}{\runtimetermvar}{\trace}$.
@@
   \begin{nscenter}
-    \bul{$\sat{\progvar}{\pi}$}
+    \bul{$\sat{\wholeprogvar}{\pi}$}
     $\isdef$
-    if \iul{$\forall\runtimetermvar\ \trace,\progstepto{\progvar}{\runtimetermvar}{\trace}$},
+    if \iul{$\forall\runtimetermvar\ \trace,\progstepto{\wholeprogvar}{\runtimetermvar}{\trace}$},
     then \oul{$\trace\in\pi$}.
@@
 The case study of this paper exclusively uses the universal image, since some considered properties, such as \gls*{ss}, are not definable in a higher-level language that, e.g., does not model speculation.
-For sake of self-containment, the definition of the universal image is reiterated~\cite{abate2021extacc}:
+For self-containment, the universal image is defined as follows~\cite{abate2021extacc}:
 
@@
 Using these projection functions, the more general version of robust preservation is defined as follows~\cite{abate2021extacc}:
-For a \bul{compiler $\cc{\src{L}}{\trg{L}}$ to robustly preserve a class of source properties $\src{\class}$}, if for any \rul{property $\src{\pi}$ of that class $\src{\class}$ and programs $\src{p}$ written in $\src{L}$} where \iul{$\src{\progvar}$ robustly satisfies $\src{\pi}$}, then \oul{the compilation of $\src{\progvar}$, $\cc{\src{L}}{\trg{L}}\left(\src{p}\right)$, must robustly satisfy $\tau_\sim\left(\src{\pi}\right)$}.
-
-\begin{definition}[Robust Preservation with $\tau_\sim$]\label{def:rtp:tau}
-  % $\;$\\
-  % \vspace{-1em}
-  % \begin{nscenter}
-  % Compiler $\cc{\src{L}}{\trg{L}}$ robustly preserves $\class$, 
-  \bul{$\rtptau{\cc{\src{L}}{\trg{L}}}{\src{\class}}{\sim}$}
-  %, iff 
-  $\isdef$
-  \rul{$\forall \src{\pi}\in\src{\class}, \src{p}\in\src{L},$} if \iul{$\rsat{\src{\progvar}}{\src{\pi}}$}, then \oul{$\rsat{\cc{\src{L}}{\trg{L}}\left(\src{p}\right)}{\tau_\sim{\left(\src{\pi}\right)}}$}.
-  % \end{nscenter}
-\end{definition}
-% 
-Note that a class of properties $\class$ can represent just one property $\pi$ by lifting~\cite{clarkson2008hyper} that property to sets of properties, i.e., use the powerset of $\pi$ instead of $\pi$ itself.
-Because of this, this paper may write $\rtptau{\cc{\src{L}}{\trg{L}}}{\src{\pi}}{\sim}$, even though $\src{\pi}$ is a property and not a class.
-A similar construction can be applied to the projection functions.
-
-Dually to \Cref{def:rtp:tau}, a \bul{compiler $\cc{\src{L}}{\trg{L}}$ to robustly preserves a class of target properties $\trg{\class}$}, if for any \rul{property $\trg{\pi}$ of class $\trg{\class}$ and programs $\src{p}$}, where \iul{$\src{\progvar}$ robustly satisfies $\sigma_\sim\left(\trg{\pi}\right)$}, \oul{the compilation of $\src{\progvar}$, $\cc{\src{L}}{\trg{L}}\left(\src{p}\right)$, must robustly satisfy $\trg{\pi}$}.
+% For a \bul{compiler $\cc{\src{L}}{\trg{L}}$ to robustly preserve a class of source properties $\src{\class}$}, if for any \rul{property $\src{\pi}$ of that class $\src{\class}$ and programs $\src{p}$ written in $\src{L}$} where \iul{$\src{\progvar}$ robustly satisfies $\src{\pi}$}, then \oul{the compilation of $\src{\progvar}$, $\cc{\src{L}}{\trg{L}}\left(\src{p}\right)$, must robustly satisfy $\tau_\sim\left(\src{\pi}\right)$}.
+
+% \begin{definition}[Robust Preservation with $\tau_\sim$]\label{def:rtp:tau}
+%   % $\;$\\
+%   % \vspace{-1em}
+%   % \begin{nscenter}
+%   % Compiler $\cc{\src{L}}{\trg{L}}$ robustly preserves $\class$, 
+%   \bul{$\rtptau{\cc{\src{L}}{\trg{L}}}{\src{\class}}{\sim}$}
+%   %, iff 
+%   $\isdef$
+%   \rul{$\forall \src{\pi}\in\src{\class}, \src{p}\in\src{L},$} if \iul{$\rsat{\src{\progvar}}{\src{\pi}}$}, then \oul{$\rsat{\cc{\src{L}}{\trg{L}}\left(\src{p}\right)}{\tau_\sim{\left(\src{\pi}\right)}}$}.
+%   % \end{nscenter}
+% \end{definition}
+% 
+A \bul{compiler $\cc{\src{L}}{\trg{L}}$ to robustly preserves a class of target properties $\trg{\class}$}, if for any \rul{property $\trg{\pi}$ of class $\trg{\class}$ and programs $\src{p}$}, where \iul{$\src{\progvar}$ robustly satisfies $\sigma_\sim\left(\trg{\pi}\right)$}, \oul{the compilation of $\src{\progvar}$, $\cc{\src{L}}{\trg{L}}\left(\src{p}\right)$, must robustly satisfy $\trg{\pi}$}.
 
@@
 \end{definition}
+Note that a class of properties $\class$ can represent just one property $\pi$ by lifting~\cite{clarkson2008hyper} that property to sets of properties, i.e., use the powerset of $\pi$ instead of $\pi$ itself.
+Because of this, this paper may write $\rtptau{\cc{\src{L}}{\trg{L}}}{\trg{\pi}}{\sim}$, even though $\trg{\pi}$ is a property and not a class.
+A similar construction can be applied to the projection functions.
 
@@
 This {\em phase ordering problem} is well--known in literature and a practical solution is to simply perform a fixpoint iteration of the optimisation pipeline~\cite{click1995combining}.
-Hence, it is of interest to analyze the security of compilation passes and their interation within a compilation pipeline.
+Hence, it is of interest to analyze the security of compilation passes and their interaction within a compilation pipeline.
 
@@
   \noindent
-  \text{\bul{$\wfcsig{\sim}{\trg{\class}}$}} \isdef \text{\rul{$\forall \trg{\Pi}\in\trg{\class}, \sigma_\sim(\trg{\Pi})\in\sigma_\sim(\trg{\class})$}}
+  \text{\bul{$\wfcsig{\sim}{\trg{\class}}$}} \isdef \text{\rul{$\forall \trg{\pi}\in\trg{\class}, \sigma_\sim(\trg{\pi})\in\sigma_\sim(\trg{\class})$}}
   \end{nscenter}
@@
 Since the composition of secure compilers is again a secure compiler, the theorems generalise to a whole chain of $n$ secure compilers. 
-Moreover, the previously defined theory dually holds for the existential image $\tau_\sim\left(\src{\pi}\right)$, but these results have been moved to the technical report in the interest of space and interest.
+Moreover, the previously defined theory dually holds for the existential image $\tau_\sim\left(\src{\pi}\right)$, but these results have been moved to the technical report in the interest of space.
 
 \subsection{Secure Compiler Composition with Same Trace Models}
-When the cross-language trace relation is an equality, \Cref{thm:rtpsim:sig} collapse:
+When the cross-language trace relation is an equality, \Cref{thm:rtpsim:sig} collapses:
 Given \bul{$\cc{\src{L}}{\trg{L}}$ robustly preserves $\class[_{1}]$} and \rul{$\cc{\trg{L}}{\obj{L}}$ robustly preserves $\class[_{2}]$}, it follows that \oul{their sequential composition $\cc{\src{L}}{\trg{L}}\circ\cc{\trg{L}}{\obj{L}}$ robustly preserves the intersection of classes $\class[_{1}]$ and $\class[_{2}]$}.
@@
 
-\Cref{corr:rtp} provides an easy to use theorem to compose secure compilers without any caveats. 
+\Cref{corr:rtp} provides an easy to use way to compose secure compilers without any caveats. 
 However, while \Cref{thm:rtpsim:sig} explicitly requires that $\sim_1$ is well-formed w.r.t. $\trg{\class[_2]}$, if $\sim_2$ is not well-formed w.r.t. $\trg{\class[_1]}$, care must be taken. 
@@
 \Cref{corr:swappable} justifies that any such order of compilation passes is valid w.r.t. security.
-So, given two compilation passes \bul{$\cc[_{1}]{\trg{L}}{\trg{L}}$, $\cc[_{2}]{\trg{L}}{\trg{L}}$, both robustly preserving class $\class[_{1}]$ or $\class[_{2}]$, respectively}, \oul{for any order of their composition the composed compiler robustly preserves the intersection of $\class[_{1}]$ and $\class[_{2}]$}.
+So, given two compilation passes \bul{$\cc[_{1}]{\trg{L}}{\trg{L}}$, $\cc[_{2}]{\trg{L}}{\trg{L}}$, both robustly preserving class $\class[_{1}]$ or $\class[_{2}]$, respectively}, \oul{for any order of their composition, the composed compiler robustly preserves the intersection of $\class[_{1}]$ and $\class[_{2}]$}.
 % 
@@
 
+Future work intends to use upper and lower composition as means to build a secure compiler incrementally.
+
 \section{Security Properties: Formalisation, Enforcement and Composition\pages{1.5}}\label{sec:compprop}
@@
 While up to \gls*{scct}, none of the properties is novel, for self-containment of the paper, their definition is necessary for the case study.
-For each of the presented properties, the technical appendix accompanying the paper defines corresponding monitors, which refine a given property and have a major role in the proofs of this paper.
+For each of the presented properties, the technical appendix accompanying the paper defines corresponding monitors, which refine a given property and have a key role in the proofs of this paper.
 
@@
 
-For cryptographic code, there is a general guideline that secrets must not be visible on a trace, i.e., marked as $\unlock$.
-Moreover, an instruction whose timing is data-dependent must not have a secret as an operand.
+For cryptographic code, there is a general guideline that secrets must not be visible on a trace~\cite{ctguidelines}, i.e., marked as $\unlock$.
+In turn, an instruction whose timing is data-dependent must not have a secret as an operand.
 Typical operations with data-dependent timing are branches and certain binary operations, such as division. 
@@
 }
-Both operations are also modelled in the trace model by extending the set of base-events with branches on some value $n$ ($\ev{Branch\ n}$) and binary operations with result $n$ ($\ev{Binop\ n}$).
+Both operations are also modelled in the trace model by extending the set of base-events with branches ($\ev{Branch\ n}$) and binary operations ($\ev{Binop\ n}$).
 
@@
 However, this is considered standard practice in terms of coding guidelines~\cite{ctguidelines}.
-Moreover, programs that operate on secrets and have a ``data independent timing mode'' enabled do not leak these secrets, despite performing computation on them (see \Cref{ex:lscct}).
+Moreover, programs that operate on secrets and have a ``data independent timing mode'' enabled do not leak these secrets, despite performing computation on them~\cite{arm-refman,intel-refman} (see \Cref{ex:lscct}).
 
@@
 Essentially, the only work $\sim_{\ctsafe}$ has to do, is to ignore both the new $\ev{Branch\ n}$ and $\ev{Binop\ n}$ base-events as well as the security tag, since they are irrelevant to \gls*{ms}.
-The relation is extended from just events to traces in a standard way, skipping the empty event $\emptyevent$ on either side, and it is now possible to define \gls*{msscct} using the universal image:
+Put differently, traces of \gls*{ms} trivially satisfy \gls*{scct}.
+
+The above relation is extended to traces in a standard way, skipping the empty event $\emptyevent$ on either side, and it is now possible to define \gls*{msscct} using the universal image:
 
@@
 For this, extend the earlier trace model (see \Cref{subsec:scct:tracemodel}) so that the security tags ($\securitytag$) carry additional information about the kind of private data leakage, i.e., the type of speculative leak.
-Moreover, add pre-events signaling the beginning of a speculative execution ($\ev{Spec}$), a barrier ($\ev{Barrier}$) that signals that any speculative execution may not go past it, as well as a rollback event ($\ev{Rlb}$), which signals that execution resumes to where speculation started.
+Moreover, add base-events signaling the beginning of a speculative execution ($\ev{Spec}$), a barrier ($\ev{Barrier}$) that signals that any speculative execution may not go past it, as well as a rollback event ($\ev{Rlb}$), which signals that execution resumes to where speculation started.
 
@@
 
-\gls*{ss}~\cite{patrignani2021exorcising}, similar to \gls*{scct}, is an overapproximation to a variant of noninterference.
+\gls*{ss}~\cite{patrignani2021exorcising}, similar to \gls*{scct}, is an overapproximation of a variant of noninterference.
 The definition of the property itself turns out to be syntactically equal to \gls*{scct}, modulo the underlying trace model.
@@
 \begin{definition}[\glsfirst*{ss}]\label{def:trace:ss}
-  \noindent\[
+  \noindent
+
+  \begin{nscenter}
+  $
     \sssafe:=\left\{\trace[_{\mathghost}] \left|\begin{array}{rcl}
@@
                                  \end{array}\right.\right\}
-  \]
+  $ 
+  \end{nscenter}
 \end{definition}
@@
 
-The intuition is that speculation is ignored, which amounts to dropping events $\ev{Spec}$, $\ev{Rlb}$, or $\ev{Barrier}$, as well as all base events tagged with $\lock_{\text{PHT}}$. 
-Now, define \gls*{specms} as follows:
+The intuition is that \gls*{ss} is trivially satisfied in \gls*{scct}, since speculation is inexpressible there, which amounts to dropping events $\ev{Spec}$, $\ev{Rlb}$, or $\ev{Barrier}$, as well as all base events tagged with $\lock_{\text{PHT}}$.
+
+Now, define \gls*{specms} as: 
 \begin{definition}[\glsfirst*{specms}]\label{def:trace:specmsdef}
@@
   \begin{array}{rrcl}
-    \text{(Pre-Events)} & \src{\preevent} &:=& \src{\ev{Alloc \loc\ n}} \mid \src{\ev{Dealloc \loc}} \mid \src{\ev{Get \loc\ n}} \mid \src{\ev{Set \loc\ n}} \\
+    \text{(Base-Events)} & \src{\preevent} &:=& \src{\ev{Alloc \loc\ n}} \mid \src{\ev{Dealloc \loc}} \mid \src{\ev{Get \loc\ n}} \mid \src{\ev{Set \loc\ n}} \\
     \text{(Sandboxtag)} & \src{\sandboxtag} &:=& \src{\ctx} \mid \src{\comp} \\
@@
 This way, memory accesses to secret data are still exposed as memory accesses, despite being protected by data-independent timing mode, e.g, \Cref{tr:to-e-get-in-noleak}.
-Moreover, the language extends pre-events with $\obj{\ev{Branch}\ n}$ and $\obj{\ev{Binop}\ n}$ that are emitted when evaluating a branch or certain binary expressions, such as division, respectively, whenever the data independent timing mode is inactive.
+Moreover, the language extends base-events with $\obj{\ev{Branch}\ n}$ and $\obj{\ev{Binop}\ n}$ that are emitted when evaluating a branch or certain binary expressions, such as division, respectively, whenever the data independent timing mode is inactive.
 Events are extended with a security-tag to signal the secrecy of the involved data.
@@
 The evaluation steps are amended to propagate the security-tag annotations $\obj{\securitytag}$.
-When the data independent timing mode is inactive, pre-events $\obj{\ev{Branch}\ n}$ and $\obj{\ev{Binop}\ n}$ are emitted for conditionals and binary operations, respectively.
+When the data independent timing mode is inactive, base-events $\obj{\ev{Branch}\ n}$ and $\obj{\ev{Binop}\ n}$ are emitted for conditionals and binary operations, respectively.
 Otherwise, just like in the semantics of the earlier languages, $\obj{\emptyevent}$ is emitted for binary and branching operations.
@@
   \begin{array}{rrcl}
-    \text{(Pre-Events)} & \ird{\preevent} &:=& \dots \mid \ird{Spec} \mid \ird{Rlb} \mid \ird{Barrier}\\
+    \text{(Base-Events)} & \ird{\preevent} &:=& \dots \mid \ird{Spec} \mid \ird{Rlb} \mid \ird{Barrier}\\
     \text{(Expressions)} & \ird{\expr} &:=& \dots \mid \ird{\lbarrier}

### 2024-05-27 csf2/main.tex
typos

@@
 This section defines programming languages $\Ltms$, $\Ltrg$, $\Lms$, $\Lscct$, and $\Lspec$, all of which share common elements (presented in \Cref{subsec:cs:defs}).
-$\Ltms$ is the only statically typed language and exhibits the property that all well-typed programs are \gls*{tms} (\Cref{subsec:ltms}).
+In the case study of this paper, $\Ltms$ is the only statically typed language and exhibits the property that all well-typed programs are \gls*{tms} (\Cref{subsec:ltms}).
 However, not all $\Ltms$ programs are \gls*{sms}.
@@
 $\Lms$ is exactly the same language as $\Ltrg$, but this paper still distinguishes the two for sake of readability (\Cref{subsec:lms}).
-All three languages --- so $\Ltms$, $\Ltrg$, and $\Lms$ --- assume \gls*{cct} to hold, since this is -- in an ideal world -- what the programmer would expect, too: It is the job of the compiler to preserve and (potentially) enforce security~\cite{cauligi2019fact}.
+All three languages --- so $\Ltms$, $\Ltrg$, and $\Lms$ --- assume \gls*{cct} to hold, since this is -- in an ideal world -- what the programmer would expect, too: It is the job of the compiler to preserve and (potentially) enforce security~\cite{cauligi2019fact,nagarakatte2010cets,nagarakatte2009soft,akritidis2009baggy}.
 
@@
 $\src{L_{\tmssafe}}$ is a relatively naive language with second class functions, the only values being numbers. 
-It has simple binary operations, case analyses, C-like function calls, as well as C-like memory operations. 
-Pointers are left second-class and can only be referenced by an identifier, so addresses cannot be guessed by contexts.
-Programs have sets of pre-determined functions called libraries and it is assumed that functions are marked as being part of some component or context.
-For sake of simplicity, contexts may not pass pointers to the component, but only integers.
+It has simple binary operations ($\src{\lbinop{\expr[_1]}{\expr[_2]}}$), branches ($\src{\lifz{\expr[_1]}{\expr[_2]}{\expr[_3]}}$), C-like function calls ($\src{\lcall{f}{\expr}}$) and returns ($\src{\lreturn{\expr}}$), as well as C-like memory operations.
+%Pointers are left second-class and can only be referenced by an identifier, so addresses cannot be guessed by contexts.
+Programs have sets of pre-determined functions called libraries and it is assumed that functions are marked as being part of some component ($\src{\library_{\comp}}$) or context ($\src{\library_{\ctx}}$).
+Functions ($(\src{x};\src{\natt\to\natt};\src{\expr})$) are modelled as tuples and can only accept integers, which also means that contexts may not pass pointers to the component.
 %Extending this is orthogonal to this work.
@@
 Furthermore, the syntax of $\Ltrg$ is extended with a way to inspect whether a pointer is freed ($\trg{\lispoisoned{x}}$), evaluating to $\trg{0}$ if it is freed and to $\trg{1}$ otherwise.
-Functions may receive arguments that are not $\trg{\natt}$.
+Functions may receive arguments that are not $\trg{\natt}$ and values are extended with tuples.
 
@@
 Moreover, the language extends base-events with $\obj{\ev{Branch}\ n}$ and $\obj{\ev{Binop}\ n}$ that are emitted when evaluating a branch or certain binary expressions, such as division, respectively, whenever the data independent timing mode is inactive.
-Events are extended with a security-tag to signal the secrecy of the involved data.
+Events are extended with a security-tag ($\obj{\securitytag}$) to signal the secrecy of the involved data.
 

### 2024-05-27 csf2/main.tex
typos

@@
 \end{figure*}
-The section demonstrates the power of the framework (\Cref{sec:sequential}) by composing these compilers for a secure and optimising compilation chain that robustly preserves \gls*{msscct}.
+The section demonstrates the power of the framework (\Cref{sec:sequential}) by composing these compilers for a secure and optimising compilation chain that robustly preserves \gls*{mssscct}.
 The first step in this chain is the compiler from $\src{L_{\tmssafe}}$ to $\trg{L}$ that robustly preserves just \gls*{tms} (\Cref{thm:cca:rtp:tms}).
@@
 This effectively has no influence on the resulting trace, so a corresponding cross-language trace relation $\sim^{\Ltms}_{\Ltrg}$ is, up to one trace being $\src{blue}$ and the other $\trg{red}$, defined as equality.
-There is only one catch: $\sim^{\Ltms}_{\Ltrg}$ ignores irrelevant context actions, since there is no interest in contexts who are to blame for a security violation {\em inside the context itself}!
-Subsequent cross-language trace relations for the other compilers do it similarly, the explicit details can be found in the technical report.
 
@@
 {
-  \footnotesize
-\[
-  \typerulenolabel{xrel:sms:read}{n\text{ in bounds}}{\trg{Get\ \loc\ n;\comp}\sim^{\Ltrg}_{\Lms}\irl{Get\ \loc\ n;\comp}}
-  \typerulenolabel{xrel:sms:panic}{n\text{ out of bounds}}{\trg{Get\ \loc\ n;\comp}\sim^{\Ltrg}_{\Lms}\irl{\lightning}}
+\[
+  \typerulenolabel{xrel:sms:read}{\trg{n}\text{ in bounds}}{\trg{Get\ \loc\ n;\comp}\sim^{\Ltrg}_{\Lms}\irl{Get\ \loc\ n;\comp}}
+\]
+\[
+  \typerulenolabel{xrel:sms:panic}{\trg{n}\text{ out of bounds}}{\trg{Get\ \loc\ n;\comp}\sim^{\Ltrg}_{\Lms}\irl{\lightning}}
 \]
@@
 The \gls*{dce} pass applies a naive rewrite rule on conditionals.
-The \gls*{cf} pass relies on an auxiliary function \texttt{mix} that uses a substitutions accumulator $\irl{\substlist}$ in order to rewrite constant binary operations, e.g., $\irl{{17}-1}$ to $\irl{16}$, and replace variables that are assigned to constants with their constant, e.g., $\irl{\llet{x}{7}{x}}$ to $\irl{7}$.
+The \gls*{cf} pass relies on an auxiliary function \texttt{mix} that uses a substitutions accumulator $\irl{\substlist}$ in order to rewrite constant binary operations, e.g., $\irl{{17}-1}$ to $\irl{16}$, and replace variables that are assigned to constants, e.g., $\irl{\llet{x}{7}{x}}$ to $\irl{7}$.
 
@@
 Note that both passes have no effect on the resulting trace of a program, up to $\emptyevent$-steps. 
-Because of this, both passes have equalit as corresponding cross language tracce relation. 
+Because of this, both passes have equality as corresponding cross language trace relation. 
 Moreover, it is straightforward to prove both passes as secure (\Cref{def:rtp}) w.r.t. \gls*{ms}.
@@
 \begin{array}{rcl}
-  \ccspec(\obj{\lifz{\expr[_0]}{\expr[_1]}{\expr[_2]}}) & = & \lifz[\ird]{\ccspec{\left(\obj{\expr[_0]}\right)}}{\\&&\ \ \ \ \ird{\lbarrier;}\ccspec{\left(\obj{\expr[_1]}\right)}\\&&}{\\&&\ \ \ \ \ird{\lbarrier;}\ccspec{\left(\obj{\expr[_2]}\right)}} \\
+  \ccspec(\obj{\lifz{\expr[_0]}{\expr[_1]}{\expr[_2]}}) & = & \lifz[\ird]{\ccspec{\left(\obj{\expr[_0]}\right)}}{\\&&\ \ \ \ird{\lbarrier;}\ccspec{\left(\obj{\expr[_1]}\right)}\\&&}{\\&&\ \ \ \ird{\lbarrier;}\ccspec{\left(\obj{\expr[_2]}\right)}} \\
 \end{array}
@@
 
-$\ccspec$ is secure with respect to \gls*{ss}.
+With this, $\ccspec$ is secure with respect to \gls*{ss}.
 \begin{theorem}[$\ccspec$ secure w.r.t. \gls*{ss}]\label{thm:ccspec:rtp:spec}

### 2024-05-27 csf2/main.tex
nits of spacing + intro

@@
 
-\author{}
+% \author{}
 % Matthis Kruse\orcidID{0000-0003-4062-9666}
@@
 From an engineering perspective, this is the desirable approach to building secure compilers.
-
 \begin{center}\small\it
@@
 % 	% and a $\obj{green}$, $\obj{teletype}$ font for a $\obj{target}$ language.	
+	Please print or view in colours.
 	}
@@
 \gls*{ms} is a security property obtained by composing \gls*{sms}, which ensures array accesses are all within bounds, and \gls*{tms}, which ensures pointers are only used when they are valid~\cite{azevedo2018meaningsofms,jim2002cyclone,necula2005ccured,nagarakatte2010cets,nagarakatte2009soft,akritidis2009baggy,michael2023mswasm}.
-\gls*{cct} is a security property that ensures sensitive data is not leaked via timing side-channels~\cite{kocher1996timing} and \gls*{ss} is a security property that enforces the same but under a speculative semantics~\cite{guarnieri2018spectector}.
-Together, \gls*{sms}, \gls*{tms}, \gls*{scct}, an enforceable overapproximation of \gls*{cct}, and \gls*{ss}, yield \gls*{mssscct}, which is the gold standard of security properties for secure applications.
+\gls*{cct} is a security property that ensures sensitive data is not leaked via timing side-channels~\cite{kocher1996timing} and \gls*{ss} is a security property that enforces the same but under a speculative semantics~\cite{guarnieri2018spectector}\MP{cit xaver ccs} that captures speculative execution attacks such as spectre\MP{citations}.
+Together, \gls*{sms}, \gls*{tms}, \gls*{scct}---an enforceable overapproximation of \gls*{cct}---and \gls*{ss}, yield \gls*{mssscct}, which is the gold standard of security properties for secure applications.
 Programs attaining \gls*{mssscct} do not leak sensitive data either through erroneous memory accesses, nor through timing side-channels, even under speculative execution.
-As discussed in \Cref{ex:strncpy}, these security properties can be enforced by compiler passes~\cite{bond2017vale,almeida2017jasmin}, to ensure programmers need not be aware of the architectural details of where their code will run.
+\Cref{ex:strncpy} below discusses how these security properties can be enforced by compiler passes~\cite{bond2017vale,almeida2017jasmin}, to ensure programmers need not be aware of the architectural details of where their code will run.
 
@@
 the memory location past the last element will be leaked to an attacker.
-\begin{lstlisting}[language=c,basicstyle=\footnotesize\ttfamily,morekeywords={size_t}]
+\begin{lstlisting}[language=c,basicstyle=\small\ttfamily,morekeywords={size_t}]
 void strncpy(size_t n, char *dst, char *src) {
@@
 In fact, the loop can terminate early, as soon as the string-terminating character \texttt{'\textbackslash 0'} is encountered, thus making program execution time proportional to the length of the array pointed by \texttt{src}, and violating the \gls*{cct} property.
+\MP{can we use scct here, instead of cct?}
 % Often, programmers rewrite programs like these to have loops that run with a fixed number of iterations.
@@
 
+\MP{flow}
 Robust compilers~\cite{abate2019jour} are a form of secure compilers that preserve security properties even in the presence of arbitrary attackers interacting with compiled code.
@@
 % Different kinds of compilation passes exist, e.g., there are usually several compilation passes that perform optimisations~\cite{lattner2004llvm,googlev8,androidstudio,kuepper2023cryptopt,manjikian1997fusion,wegman1991ccp} or source-code instrumentations~\cite{nagarakatte2009soft,nagarakatte2010cets,akritidis2009baggy,dhumbumroong2020boundwarden,jung2021pico,nam2019framer,shankaranarayana2023tailcheck,younan2010paricheck,zhou2023fatptrs,bond2017vale,almeida2017jasmin,kuepper2023cryptopt,cauligi2019fact} that enforce that the compiled program fulfils a property of interest.
-Worse, without a framework for composing secure compiler passes, it is not possible to enable separation of concerns, e.g., to have a secure compilation pass that ensures \gls*{ms} that is developed independently of another secure pass for \gls*{cct}, that is developed independently of other passes, such as optimisations.
-
-This paper introduces a framework for reasoning about the composition of secure and optimising compiler passes akin to those of \Cref{ex:strncpy} and it showcases the power of this framework by instantiating it on a multi-pass compilation chain.
+Worse, without a framework for composing secure compiler passes, it is not possible to enable separation of concerns, e.g., to have a secure compilation pass that ensures \gls*{ms} that is developed independently of another secure pass for \gls*{cct}, that is developed independently of other passes, such as optimisation passes.
+
+This paper introduces a framework for reasoning about the composition of secure and optimising compiler passes akin to those of \Cref{ex:strncpy}\MP{there aren't really passes there} and it showcases the power of this framework by instantiating it on a multi-pass compilation chain.
 To this end, this paper first discusses how to compose security properties, such as \gls*{tms} and \gls*{sms} into \gls*{ms}, and then adds \gls*{scct} as well as \gls*{ss} into the mix to obtain \gls*{mssscct}.
@@
         %Finally, this paper extends parts of the compositionality framework to support different trace-models between compilation passes.
+        \MP{this isn't really true, is it?. the order matters for the relations' lemmas}
 
@@
         This paper presents a case-study with five programming languages showcasing the previous contribution (\Cref{sec:casestud:defs,sec:casestud:rtp}).
-        To this end, it presents a compilation chain consisting of six passes, two of which are well-known optimisations \gls*{dce} and \gls*{cf}, that ultimately preserve \gls*{mssscct} by means of composing the individual, secure passes concerning \gls*{tms}, \gls*{sms}, \gls*{scct}, and \gls*{ss}, respectively.
+        To this end, it presents a compilation chain consisting of six passes that ultimately preserve \gls*{mssscct} by means of composing four secure passes that \emph{individually} preserve \gls*{tms}, \gls*{sms}, \gls*{scct}, and \gls*{ss}, and two passes which are well-known optimisations \gls*{dce} and \gls*{cf},.
         %Furthermore, the chain includes two optimisation passes: One performs \gls*{dce} and the other \gls*{cf}.
@@
   \item %
-        The considered compilation passes are analyzed in terms of their compatibility with each other (\Cref{sec:formalities}).
-        In doing so, the paper provides formal insights into some technicalities of the concrete proofs necessary as technical setup to reap the benefits of the presented compositionality framework.
+  	The paper provides formal proofs of the security of each presented pass, and it analyzes each pass in terms of their compatibility with each other (\Cref{sec:formalities}).
+        % In doing so, the paper provides formal insights into some technicalities of the concrete proofs necessary as technical setup to reap the benefits of the presented compositionality framework.
 
@@
 {
-\small
 \[
@@
 
-\vspace{-1.5em}
-{\small
+\vspace{-.5em}
+{
+% \small
 \[
@@
 
-\vspace{-1.5em}
+\vspace{-.5em}
 \[
@@
 \begin{center}
-\footnotesize
+\newcommand{\expreval}[5]{{#1}\triangleright\xspace {#2}\xrightarrow{#5}\ {#3}\triangleright\xspace {#4}\xspace}
+\newcommand{\exprevalo}[5]{\expreval{\obj{#1}}{\obj{#2}}{\obj{#3}}{\obj{#4}}{\obj{#5}}}
   \typerule{$e-\obj{wrdoit}-\text{off}$}{
@@
   }{e-wrdoit-off}
-\newcommand{\expreval}[5]{{#1}\triangleright\xspace {#2}\xrightarrow{#5}\ {#3}\triangleright\xspace {#4}\xspace}
-\newcommand{\exprevalo}[5]{\expreval{\obj{#1}}{\obj{#2}}{\obj{#3}}{\obj{#4}}{\obj{#5}}}
+
   \typerule{$e-\obj{get}-\in-$noleak}{
-    \obj{\memstate}=\obj{H^{\ctx};H^{\comp};\Delta_1,x\mapsto(\loc;\sandboxtag;\poison;m),\Delta_2} &
-    \obj{\loc}+\obj{n}\in\text{dom }\obj{H^{\sandboxtag}} \\
+    \obj{\memstate}=\obj{H^{\ctx};H^{\comp};\Delta_1,x\mapsto(\loc;\sandboxtag;\poison;m),\Delta_2} \\
+    \obj{\loc}+\obj{n}\in\text{dom }\obj{H^{\sandboxtag}} &
     \obj{\sigma''}=\obj{\sigma\sqcap\sigma'}
   }{
-    \exprevalo{\cfstate;\sandboxtag';m+1;\memstate}{x^\sigma[n^{\sigma'}]}{\cfstate;\sandboxtag';m+1;\memstate}{(H^{\sandboxtag}(\loc+n))^{\sigma''}}{{iGet}\ \loc\ n;\sandboxtag;\sigma''}
+    \exprevalo{
+    	\cfstate;\sandboxtag';m+1;\memstate
+    }{
+    	x^\sigma[n^{\sigma'}]
+    }{
+    	\cfstate;\sandboxtag';m+1;\memstate
+    }{
+    	(H^{\sandboxtag}(\loc+n))^{\sigma''}
+    }{
+    	{iGet}\ \loc\ n;\sandboxtag;\sigma''
+    }
   }{to-e-get-in-noleak}
@@
 
-\vspace{-1em}
+\vspace{-.5em}
 \[
@@
 \begin{center}
-  \footnotesize
 \newcommand{\expreval}[5]{{#1}\triangleright\xspace {#2}\xrightarrow{#5}\ {#3}\triangleright\xspace {#4}\xspace}
@@
     \ird{\cfstate}.\ird{\operatorname{win}} > 0 &
-    \pstepto{\rtt{\ird{\cfstate;\sandboxtag;0;\memstate}}{\ird{\expr}}}{\rtt{\ird{\cfstate['];\sandboxtag;0;\memstate}}{\ird{\expr[']}}}{\ird{\event}} &
+    \pstepto{\rtt{\ird{\cfstate;\sandboxtag;0;\memstate}}{\ird{\expr}}}{\rtt{\ird{\cfstate['];\sandboxtag;0;\memstate}}{\ird{\expr[']}}}{\ird{\event}}
+    \\
     \ird{\cfstate['']}.\ird{\operatorname{win}} = \ird{\cfstate[']}.\ird{\operatorname{win}}-1
@@
     \draw[->] (C0) to[sloped] node[align=center] {\gls*{dce}\\ \Cref{thm:ccdce:rtp:ms}} (E);
-    \draw[->] (E) to[sloped,above] node[align=center] {\gls*{scct}\\ \Cref{thm:ccscct:rtp:scct}} (O);
-    \draw[->] (O) to[sloped,above] node[align=center] {\gls*{ss}\\ \Cref{thm:ccspec:rtp:spec}} (Z);
+    \draw[->] (E) to[sloped] node[align=center] {\gls*{scct}\\ \Cref{thm:ccscct:rtp:scct}} (O);
+    \draw[->] (O) to[sloped] node[align=center] {\gls*{ss}\\ \Cref{thm:ccspec:rtp:spec}} (Z);
 
     % Sections
-    \node (sectms) at ($(S)+(1.25,3)$) {{\Cref{subsec:cs:tms}}};
-    \node (secsms) at ($(T)+(1.25,3)$) {{\Cref{subsec:cs:ms}}};
-    \node (secopts) at ($(M)+(2.0,3)$) {{\Cref{subsec:cs:opts}}};
-    \node (secscct) at ($(E)+(1.5,3)$) {{\Cref{subsec:cs:scct}}};
-    \node (secss) at ($(O)+(1.5,3)$) {{\Cref{subsec:cs:ss}}};
-
-    \draw[thick,dotted,->] (S) to[bend right=33,sloped] node[align=center] {\gls*{ms}\\ \Cref{corr:ccab:rtp:ms}} (M);
+    \node (sectms) at ($(S)+(1.25,2.5)$) {{\Cref{subsec:cs:tms}}};
+    \node (secsms) at ($(T)+(1.25,2.5)$) {{\Cref{subsec:cs:ms}}};
+    \node (secopts) at ($(M)+(2.0,2.5)$) {{\Cref{subsec:cs:opts}}};
+    \node (secscct) at ($(E)+(1.5,2.5)$) {{\Cref{subsec:cs:scct}}};
+    \node (secss) at ($(O)+(1.5,2.5)$) {{\Cref{subsec:cs:ss}}};
+
+    \draw[thick,dotted,->] ($(S) + (0.2,0.0)$) to ($(S) - (-0.2,1)$) to node[align=center] {\gls*{ms}\\ \Cref{corr:ccab:rtp:ms}} ($(M) - (0.0,1)$) to (M);
+
+    \draw[thick,dotted,->] (S) to ($(S) - (0.0,2.5)$) to node[align=center,pos=.65] {\gls*{mssscct}\\ \Cref{thm:ccall:rtp:mssscct}} ($(Z) - (0.0,2.5)$) to (Z);
+
+    % \draw[thick,dotted,->] (S) to[bend right=33,sloped] node[align=center] {\gls*{ms}\\ \Cref{corr:ccab:rtp:ms}} (M);
     \draw[thick,dotted,->] (M) to[bend right=0,sloped] node[align=center] {\gls*{ms}\\ \Cref{thm:cccfccdce:rtp:ms}} (E);
 %    \draw[thick,dotted,->] (S) to[out=-45,in=198,sloped] node[align=center] {\gls*{ms}+\gls*{scct}\\ \Cref{thm:ccall:rtp:msscct}} (O);
-    \draw[thick,dotted,->] (S) to[out=-40,in=210,sloped] node[align=center] {\gls*{mssscct}\\ \Cref{thm:ccall:rtp:mssscct}} (Z);
-
-    \draw[dashed] ($(S)-(0.0,-0.5)$) -- ($(S)-(0.0,-3)$);
-    \draw[dashed] ($(T)-(0.0,-0.5)$) -- ($(T)-(0.0,-3)$);
-    \draw[dashed] ($(M)-(0.0,-0.5)$) -- ($(M)-(0.0,-3)$);
-    \draw[dashed] ($(E)-(0.0,-0.5)$) -- ($(E)-(0.0,-3)$);
-    \draw[dashed] ($(O)-(0.0,-0.5)$) -- ($(O)-(0.0,-3)$);
-    \draw[dashed] ($(Z)-(0.0,-0.5)$) -- ($(Z)-(0.0,-3)$);
+    % \draw[thick,dotted,->] (S) to[out=-40,in=210,sloped] node[align=center,pos=.8] {\gls*{mssscct}\\ \Cref{thm:ccall:rtp:mssscct}} (Z);
+
+    \draw[dashed] ($(S)-(0.0,-0.5)$) -- ($(S)-(0.0,-2.5)$);
+    \draw[dashed] ($(T)-(0.0,-0.5)$) -- ($(T)-(0.0,-2.5)$);
+    \draw[dashed] ($(M)-(0.0,-0.5)$) -- ($(M)-(0.0,-2.5)$);
+    \draw[dashed] ($(E)-(0.0,-0.5)$) -- ($(E)-(0.0,-2.5)$);
+    \draw[dashed] ($(O)-(0.0,-0.5)$) -- ($(O)-(0.0,-2.5)$);
+    \draw[dashed] ($(Z)-(0.0,-0.5)$) -- ($(Z)-(0.0,-2.5)$);
   \end{tikzpicture}
-  % \vspace{-2.5em}
+  \vspace{-1em}
   \caption{Visualisation of the optimising compilation pipeline that attains a combination of \gls*{ms} and \gls*{cct}. %
@@
 \vspace{-1em}
-{\small
+{
 \[
@@
   Consider again \texttt{strncpy}, but instrumented for \gls*{sms}:
-    \begin{lstlisting}[language=c,basicstyle=\ttfamily\footnotesize, morekeywords={size_t}]
+    \begin{lstlisting}[language=c,basicstyle=\ttfamily, morekeywords={size_t}]
 void strncpy(size_t n, size_t dst_size, char *dst,
@@
 \[
-\begin{array}{rcl}
-  \ccspec(\obj{\lifz{\expr[_0]}{\expr[_1]}{\expr[_2]}}) & = & \lifz[\ird]{\ccspec{\left(\obj{\expr[_0]}\right)}}{\\&&\ \ \ \ird{\lbarrier;}\ccspec{\left(\obj{\expr[_1]}\right)}\\&&}{\\&&\ \ \ \ird{\lbarrier;}\ccspec{\left(\obj{\expr[_2]}\right)}} \\
+\begin{array}{cl}
+  &\ccspec{(\obj{\lifz{\expr[_0]}{\expr[_1]}{\expr[_2]}})} = 
+  \\
+  &\qquad\qquad \lifz[\ird]{\ccspec{\left(\obj{\expr[_0]}\right)}}{\ird{\lbarrier;}\ccspec{\left(\obj{\expr[_1]}\right)} \\&\qquad\qquad}{ \ird{\lbarrier;}\ccspec{\left(\obj{\expr[_2]}\right)}} 
 \end{array}

### 2024-05-27 csf2/main.tex
contribs nits

@@
         % 
-        From this, the paper proves that the order of composition of sequential compiler passes is irrelevant for security.
-        % 
-        This is crucial for reordering individual optimisation passes and thus generating secure and efficient code.
+        % From this, the paper proves that the order of composition of sequential compiler passes is irrelevant for security.
+        % % 
+        % This is crucial for reordering individual optimisation passes and thus generating secure and efficient code.
         %
         %Finally, this paper extends parts of the compositionality framework to support different trace-models between compilation passes.
-        \MP{this isn't really true, is it?. the order matters for the relations' lemmas}
+        % \MP{this isn't really true, is it?. the order matters for the relations' lemmas}
+        Then, this paper identifies which conditions make the composition of secure compilers meaningful from a security perspective, and which other conditions allow passes to be swapped without losing security meaning.
 
@@
   \item %
-  	The paper provides formal proofs of the security of each presented pass, and it analyzes each pass in terms of their compatibility with each other (\Cref{sec:formalities}).
+  		This paper provides formal proofs of the security properties preserved by each of the presented pass. 
+  		% 
+  		Additionally this paper analyzes each pass in terms of their compatibility with each other (\Cref{sec:formalities}), proving how does each pass fulfill the conditions that make their composition meaningful from a security standpoint.
         % In doing so, the paper provides formal insights into some technicalities of the concrete proofs necessary as technical setup to reap the benefits of the presented compositionality framework.

### 2024-05-27 csf2/main.tex
sizing
typos

@@
 \gls*{ms} is a security property obtained by composing \gls*{sms}, which ensures array accesses are all within bounds, and \gls*{tms}, which ensures pointers are only used when they are valid~\cite{azevedo2018meaningsofms,jim2002cyclone,necula2005ccured,nagarakatte2010cets,nagarakatte2009soft,akritidis2009baggy,michael2023mswasm}.
-\gls*{cct} is a security property that ensures sensitive data is not leaked via timing side-channels~\cite{kocher1996timing} and \gls*{ss} is a security property that enforces the same but under a speculative semantics~\cite{guarnieri2018spectector}\MP{cit xaver ccs} that captures speculative execution attacks such as spectre\MP{citations}.
-Together, \gls*{sms}, \gls*{tms}, \gls*{scct}---an enforceable overapproximation of \gls*{cct}---and \gls*{ss}, yield \gls*{mssscct}, which is the gold standard of security properties for secure applications.
+\gls*{cct} is a security property that ensures sensitive data is not leaked via timing side-channels~\cite{kocher1996timing} and \gls*{ss} is a security property that enforces the same but under a speculative semantics~\cite{guarnieri2018spectector,fabian2022automatic} that captures speculative execution attacks such as spectre~\cite{kocher2019spectre}.
+Together, \gls*{sms}, \gls*{tms}, \gls*{cct} and \gls*{ss}, yield \gls*{mssscct}, which is the gold standard of security properties for secure applications.
 Programs attaining \gls*{mssscct} do not leak sensitive data either through erroneous memory accesses, nor through timing side-channels, even under speculative execution.
@@
 In fact, the loop can terminate early, as soon as the string-terminating character \texttt{'\textbackslash 0'} is encountered, thus making program execution time proportional to the length of the array pointed by \texttt{src}, and violating the \gls*{cct} property.
-\MP{can we use scct here, instead of cct?}
 % Often, programmers rewrite programs like these to have loops that run with a fixed number of iterations.
@@
 
-This paper introduces a framework for reasoning about the composition of secure and optimising compiler passes akin to those of \Cref{ex:strncpy}\MP{there aren't really passes there} and it showcases the power of this framework by instantiating it on a multi-pass compilation chain.
-To this end, this paper first discusses how to compose security properties, such as \gls*{tms} and \gls*{sms} into \gls*{ms}, and then adds \gls*{scct} as well as \gls*{ss} into the mix to obtain \gls*{mssscct}.
+This paper introduces a framework for reasoning about the composition of secure and optimising compiler passes %akin to those of \Cref{ex:strncpy}\MP{there aren't really passes there} 
+and it showcases the power of this framework by instantiating it on a multi-pass compilation chain.
+To this end, this paper first discusses how to compose security properties, such as \gls*{tms} and \gls*{sms} into \gls*{ms}, and then adds \gls*{cct} as well as \gls*{ss} into the mix to obtain \gls*{mssscct}.
 %Then, this paper defines compiler composition and formalises that given two passes that securely preserve two (possibly distinct) properties, their composition securely preserves the composition of those properties.
@@
   \item %
-  		This paper provides formal proofs of the security properties preserved by each of the presented pass. 
+  		This paper provides formal proofs of the security properties preserved by each of the presented passes. 
   		% 
@@
   }{
-    \pstepto{\rtt{\obj{\cfstate;\sandboxtag;m;\memstate}}{\obj{\lwrdoit{OFF}}}}{\rtt{\obj{\cfstate;\sandboxtag;OFF;\memstate}}{\obj{0^{\securitytag}}}}{\obj{\emptyevent}}
+    \pstepto{\rtt{\obj{\cfstate;\sandboxtag;D;\memstate}}{\obj{\lwrdoit{OFF}}}}{\rtt{\obj{\cfstate;\sandboxtag;OFF;\memstate}}{\obj{0^{\securitytag}}}}{\obj{\emptyevent}}
   }{e-wrdoit-off}
@@
   \typerule{$e-\obj{get}-\in-$noleak}{
-    \obj{\memstate}=\obj{H^{\ctx};H^{\comp};\Delta_1,x\mapsto(\loc;\sandboxtag;\poison;m),\Delta_2} \\
-    \obj{\loc}+\obj{n}\in\text{dom }\obj{H^{\sandboxtag}} &
-    \obj{\sigma''}=\obj{\sigma\sqcap\sigma'}
+    \obj{\memstate}=\obj{H^{\ctx};H^{\comp};\Delta_1,x\mapsto(\loc;\sandboxtag;\poison;D),\Delta_2} &
+    \obj{\loc}+\obj{n}\in\text{dom }\obj{H^{\sandboxtag}} \\
+    \obj{\sigma''}=\obj{\sigma\sqcap\sigma'} &
+    \obj{\event} = \obj{{iGet}\ \loc\ n;\sandboxtag;\sigma''}
   }{
     \exprevalo{
-    	\cfstate;\sandboxtag';m+1;\memstate
+    	\cfstate;\sandboxtag';ON;\memstate
     }{
@@
     }{
-    	\cfstate;\sandboxtag';m+1;\memstate
+    	\cfstate;\sandboxtag';ON;\memstate
     }{
@@
     }{
-    	{iGet}\ \loc\ n;\sandboxtag;\sigma''
+      \event
     }

### 2024-05-27 csf2/main.tex
address todos from discussion

@@
     \text{(Base-Events)} & \src{\preevent} &:=& \src{\ev{Alloc \loc\ n}} \mid \src{\ev{Dealloc \loc}} \mid \src{\ev{Get \loc\ n}} \mid \src{\ev{Set \loc\ n}} \\
-    \text{(Sandboxtag)} & \src{\sandboxtag} &:=& \src{\ctx} \mid \src{\comp} \\
+    \text{(Control Tags)} & \src{\sandboxtag} &:=& \src{\ctx} \mid \src{\comp} \\
     \text{(Events)} & \src{\event} &:=& \src{\preevent;\sandboxtag} \mid \src{\emptyevent} \mid \src{\lightning} \\ 
@@
     \src{\library_{\comp}} : \src{Vars} \to \src{F} \\
+    \text{(Heaps)} & \src{H} &:=& \src{\hole{\cdot}} \mid \src{n},\src{H}\hspace{0.5cm}    \text{(Pointer Maps)}\ \ \ \src{\Delta} \\
+    \text{(Memory States)} & \src{\memstate} &:=& \left(\src{H^{\ctx}};\src{H^{\comp}};\src{\Delta}\right)\\
+  \text{(Control States)} & \src{\cfstate} & & \text{(States)}\ \ \ \src{\Omega}\ :=\ \left(\src{\cfstate};\src{\sandboxtag};\src{\memstate}\right)\\
   \end{array}
@@
 
-$\Ltms$ dynamic semantics do not get stuck if a memory-safety violation is encountered, such as a write out of bounds.
+Operational state ($\src{\Omega}$) is a triple consisting of control-flow state ($\src{\cfstate}$), a control tag ($\src{\sandboxtag}$), and a memory state ($\memstate$), which carries information about the memory.
+Because pointers are kept ,,alive'' in pointer maps ($\src{\Delta}$), $\Ltms$ dynamic semantics do not get stuck if a memory-safety violation is encountered, such as a double--free.
+The memory state carries two heaps to model sandboxing between a context and a component.
+
 The trace model is largely similar to the memory safety trace model (\Cref{subsec:basic:memsafety:tracemodel}), where the only real differences are separate events for reading and writing memory, instead of just using it, and the sandboxtags.
@@
 $\Lscct$ extends $\Lms$ (\Cref{subsec:lms}) with a way to read from ($\obj{\lrddoit{x}{\expr}}$) and to write to ($\obj{\lwrdoit{D}}$) a model specific register that controls a data (operand) independent timing mode, a feature that is present in both Arm~\cite[p.~543]{arm-refman} and Intel~\cite[p.~80]{intel-refman} processors.
-To this end, states are extended with the value of the register ($\obj{D}$), which is initially set to be not active ($\obj{OFF}$).
+To this end, states ($\src{\Omega}$) are extended with the value of the register ($\obj{D}$), which is initially set to be not active ($\obj{OFF}$).
 If the register is marked active ($\obj{ON}$), the intuition is that no secrets can appear on specification traces.
@@
     \text{(Security Tags)} & \obj{\securitytag} &:=& \obj{\lock} \mid \obj{\unlock} \\
-    \text{(Base-Event)} & \obj{\preevent} &:=& \dots \mid \obj{iGet\ \loc\ v} \mid \obj{iSet\ \loc\ v} \\
+    \text{(Base-Events)} & \obj{\preevent} &:=& \dots \mid \obj{iGet\ \loc\ v} \mid \obj{iSet\ \loc\ v} \\
                        &&& \mid\ \obj{Binop\ n} \mid \obj{Branch\ n} \\
-    \text{(Event)} & \obj{\event} &:=& \obj{\preevent;\sandboxtag;\securitytag} \mid \obj{\emptyevent} \mid \obj{\lightning} \\
+    \text{(Events)} & \obj{\event} &:=& \obj{\preevent;\sandboxtag;\securitytag} \mid \obj{\emptyevent} \mid \obj{\lightning} \\
     \text{(Expressions)} & \obj{\expr} &:=& \dots \mid \obj{\lrddoit{x}{\expr}} \mid \obj{\lwrdoit{D}} \\
                          &&&\mid\ \obj{\llet{x^{\securitytag}}{\expr[_1]}{\expr[_2]}}\\
+    \text{(States)} & \obj{\Omega} & := & \left(\obj{\cfstate};\obj{\sandboxtag};\obj{D};\obj{\memstate}\right)\\
   \end{array}
@@
   \typerule{$e-\obj{get}-\in-$noleak}{
-    \obj{\memstate}=\obj{H^{\ctx};H^{\comp};\Delta_1,x\mapsto(\loc;\sandboxtag;\poison;D),\Delta_2} &
+    \obj{\memstate}=\obj{H^{\ctx};H^{\comp};\Delta_1,x\mapsto(\loc;\sandboxtag;\poison),\Delta_2} &
     \obj{\loc}+\obj{n}\in\text{dom }\obj{H^{\sandboxtag}} \\
@@
 This way, memory accesses to secret data are still exposed as memory accesses, despite being protected by data-independent timing mode, e.g, \Cref{tr:to-e-get-in-noleak}.
-Moreover, the language extends base-events with $\obj{\ev{Branch}\ n}$ and $\obj{\ev{Binop}\ n}$ that are emitted when evaluating a branch or certain binary expressions, such as division, respectively, whenever the data independent timing mode is inactive.
+Due to the technical setup, the rule needs to check if the access is in bounds ($\obj{\loc}+\obj{n}\in\obj{H^t}$) and update the secrecy tag ($\obj{\securitytag''}=\obj{\securitytag\sqcap\securitytag'}$) according to the lattice described earlier.
+The precise information carried by the event, e.g., location ($\obj{\loc}$) is taken from the pointer map, which carries some more information irrelevant to this rule ($\obj{\poison}$).
+
+The language extends base-events with $\obj{\ev{Branch}\ n}$ and $\obj{\ev{Binop}\ n}$ that are emitted when evaluating a branch or certain binary expressions, such as division, respectively, whenever the data independent timing mode is inactive.
 Events are extended with a security-tag ($\obj{\securitytag}$) to signal the secrecy of the involved data.

### 2024-05-27 csf2/main.tex
intro

--- 2024-05-27 csf2/main.tex
@@
 
-% \author{}
+\author{\vspace*{-2em}\phantom{a}}
 % Matthis Kruse\orcidID{0000-0003-4062-9666}
@@
 \gls*{cct} is a security property that ensures sensitive data is not leaked via timing side-channels~\cite{kocher1996timing} and \gls*{ss} is a security property that enforces the same but under a speculative semantics~\cite{guarnieri2018spectector}\MP{cit xaver ccs} that captures speculative execution attacks such as spectre\MP{citations}.
-Together, \gls*{sms}, \gls*{tms}, \gls*{scct}---an enforceable overapproximation of \gls*{cct}---and \gls*{ss}, yield \gls*{mssscct}, which is the gold standard of security properties for secure applications.
+Together, \gls*{sms}, \gls*{tms}, \gls*{cct} and \gls*{ss}, yield \gls*{mssscct}, which is the gold standard of security properties for secure applications.
 Programs attaining \gls*{mssscct} do not leak sensitive data either through erroneous memory accesses, nor through timing side-channels, even under speculative execution.
@@
 This paper introduces a framework for reasoning about the composition of secure and optimising compiler passes akin to those of \Cref{ex:strncpy}\MP{there aren't really passes there} and it showcases the power of this framework by instantiating it on a multi-pass compilation chain.
-To this end, this paper first discusses how to compose security properties, such as \gls*{tms} and \gls*{sms} into \gls*{ms}, and then adds \gls*{scct} as well as \gls*{ss} into the mix to obtain \gls*{mssscct}.
+To this end, this paper first discusses how to compose security properties, such as \gls*{tms} and \gls*{sms} into \gls*{ms}, and then adds \gls*{scct}---an enforceable overapproximation of \gls*{cct}---as well as \gls*{ss} into the mix to obtain \gls*{mssscct}.
 %Then, this paper defines compiler composition and formalises that given two passes that securely preserve two (possibly distinct) properties, their composition securely preserves the composition of those properties.

### 2024-05-27 csf2/main.tex
flow?

--- 2024-05-27 csf2/main.tex
@@
 % An easy fix to sidestep this is to use non-nullable pointers as parameters instead, e.g., as presented in Cyclone~\cite{jim2002cyclone}.
+\end{exampleenv}
 
 \MP{flow}
+\MKin{
+  % last sentence is before the example, which talks about compiler passes enforcing props
+Whether or not compiler passes enforce certain security properties, it is important for a security-aware compiler to consider partial programs that interact with potentially malicious library code, since that code may lead to, e.g., memory--safety issues in the considered partial program.
+}
+
 Robust compilers~\cite{abate2019jour} are a form of secure compilers that preserve security properties even in the presence of arbitrary attackers interacting with compiled code.
 Thus, robust compilers can be used to prevent vulnerabilities resulting from uninitialised memory (as well as many other ones), e.g., by targeting capability-based languages such as CHERI~\cite{woodruff2014CHERI}, Arm Morello~\cite{arm-morello}, or MSWasm~\cite{michael2023mswasm}, where the compiler relies on capabilities to check that pointers are always initialised.
-\end{exampleenv}
 

### 2024-05-27 csf2/main.tex
bg

--- 2024-05-27 csf2/main.tex
@@
 
-To introduce the security argument of this paper, this section first presents the concepts of (security) properties, of their satisfaction, and of their robust satisfaction (i.e., satisfaction in the presence of an active attacker; \Cref{subsec:bg:tprop}).
-Then, borrowing from existing work~\cite{abate2019jour,abate2021extacc}, the section introduces secure compilers as compilers that preserve robust property satisfaction (\Cref{subsec:bg:rtp}).
+To introduce the security argument of this paper, this section defines (security) properties, their satisfaction, and their robust satisfaction (i.e., satisfaction in the presence of an active attacker; \Cref{subsec:bg:tprop}).
+Then, borrowing from existing work~\cite{abate2019jour,abate2021extacc}, this section introduces secure compilers as compilers that preserve robust property satisfaction (\Cref{subsec:bg:rtp}).
 
@@
 Throughout the paper, sequences are indicated with an overbar (i.e., $\trace$), empty sequences with $\hole{\cdot}$, and concatenation of sequences $\trace[_{1}],\trace[_{2}]$ as $\trace[_{1}]\cdot\trace[_{2}]$.
-Prepending elements to sequences uses the same notation: $\event\cdot\trace$.
+Prepending element $\event$ to a sequence $\trace$ uses the same notation: $\event\cdot\trace$.
 }
@@
 Properties $\pi$ are sets of traces of admissible program behaviours, ascribing what said property considers valid.
-The set of all properties can be partitioned into different {\em classes} ($\class$), i.e., safety, liveness, and neither safety nor liveness~\cite{clarkson2008hyper}.
-A class is simply a set of properties and for the class of safety properties, it is decidable whether a trace satisfies a safety property with just a finite trace prefix.
-The compositionality framework (\Cref{sec:sequential}) presented in this paper considers arbitrary classes, while the case-study (\Cref{sec:casestud:rtp}) fixes them to concrete instances.
-As an example, consider a trace describing an interaction with a memory where the deallocation of an address $\loc$ precedes a read at that address in memory: $\ev{Dealloc\ \loc}\cdot\ev{Read\ \loc\ 1729}\cdot\dots$.
+% 
+The set of all properties can be partitioned into different {\em classes} ($\class$), i.e., safety, liveness, and neither safety nor liveness~\cite{clarkson2008hyper}, so a class is simply a set of properties.
+% 
+The compositionality framework (\Cref{sec:sequential}) presented in this paper considers arbitrary classes, while the case-study (\Cref{sec:casestud:rtp}) fixes them to concrete instances of safety properties, since it is decidable whether a trace satisfies a safety property with just a finite trace (i.e., a \emph{prefix}).
+As an example, consider the trace:
+$$\ev{Dealloc\ \loc}\cdot\ev{Read\ \loc\ 1729}\cdot\dots$$
+which describes the interaction with a memory where the deallocation of an address $\loc$ precedes a read (of some value $1729$) at that address in memory.
+% 
 This program behaviour is insecure w.r.t.
@@
 In the following, the execution of a whole program $\wholeprogvar$ that terminates in state $\runtimetermvar$ according to the language semantics and produces trace $\trace$ is written as $\progstepto{\wholeprogvar}{\runtimetermvar}{\trace}$.
-With this, property satisfaction is defined as follows:
-\bul{whole programs $\wholeprogvar$ satisfy a property $\pi$} iff \iul{$\wholeprogvar$ yields a trace $\trace$} such that \oul{$\trace$ satisifies $\pi$} (\Cref{def:propsat}).
+With this, we defined property satisfaction as follows:
+\bul{a whole program $\wholeprogvar$ satisfies a property $\pi$} iff \iul{$\wholeprogvar$ yields a trace $\trace$} such that \oul{$\trace$ satisfies $\pi$}.
 
@@
 Property satisfaction is defined on whole programs, i.e., programs without missing definitions.
-Thus, from a security perspective, this considers only a passive attacker model, where the attacker observes the execution and, e.g., retrieves secrets from that.
-To consider a stronger model similarly to what existing work does~\cite{abate2019jour,abate2021extacc,maffeis2008code-carrying,gordon2003authenticity,fournet2007authorization,bengtson2011refine,backes2014uniontyps,michael2023mswasm,swasey2017robust,sammler2019benefits}, the concept of satisfaction can be extended with {\em robustness}.
+Thus, from a security perspective, \Cref{def:propsat} considers only a passive attacker model, where the attacker observes the execution and, e.g., retrieves secrets from that.
+To consider a stronger model similarly to what existing work does~\cite{abate2019jour,abate2021extacc,maffeis2008code-carrying,gordon2003authenticity,fournet2007authorization,bengtson2011refine,backes2014uniontyps,michael2023mswasm,swasey2017robust,sammler2019benefits}, the we extend the concept of satisfaction with {\em robustness}.
 Robust satisfaction considers partial programs $\progvar$, i.e., components with missing imports, which cannot run until said imports are fulfilled.
@@
 this paper assumes that whole programs are the result of linking partial programs referred to as {\em context} ($\ctx$) and {\em component} ($\comp$).
-The context is an arbitrary program and thus has the role of an {\em attacker} that can interact with the component by means of whatever features the programming language has, and the component is what is security-relevant.
+The context is an arbitrary program and thus has the role of an {\em attacker} that can interact with the component by means of any features the programming language has, and the component is what is security-relevant.
 %In this work, the semantics of the programming language is expected to differentiate between the component and the context.
-With this, \Thmref{def:propsat} can be extended as follows: for \bul{components $\progvar$ to robustly satisfy a property $\pi$}, take an \iul{attacker context $\contextvar$ and link it with $\progvar$}, \oul{the resulting whole program must satisfy $\pi$}.
+With this, \Thmref{def:propsat} can be extended as follows: for \bul{a component $\progvar$ to robustly satisfy a property $\pi$}, take an \iul{attacker context $\contextvar$ and link it with $\progvar$}, \oul{the resulting whole program must satisfy $\pi$}.
 
@@
 This translation is considered {\em correct} if it is semantics-preserving.
-That is, for a whole program $\src{\wholeprogvar}$, the compiler should relate the $\src{L}$ semantics of $\src{\wholeprogvar}$ with the semantics of $\trg{L}$ of the compiled counterpart of $\src{p}$ in such a way that they are ''compatible''.
+That is, for a whole program $\src{\wholeprogvar}$, the compiler should relate the $\src{L}$ semantics of $\src{\wholeprogvar}$ with the semantics of $\trg{L}$ of the compiled counterpart of $\src{p}$ in such a way that they are ``compatible''.
 Unfortunately, correct compilers may be insecure compilers~\cite{patrignani2019survey,kennedy2006secure.net,abadi1999protect,ahmed2018dagstuhl} and programs translated by insecure compilers can violate security properties that the programmer assumes to hold.
@@
 
-This paper uses a more general notion of robust preservation~\cite{abate2021extacc} that applies to compilers that use languages with different trace models. 
+This paper uses a general notion of robust preservation~\cite{abate2021extacc} that considers compilers that use languages with different trace models. 
 To this end, considering a source trace $\src{\trace}$ and a target trace $\trg{\trace}$, there is a relation ($\sim$) describing the effect of a corresponding compiler (see \Cref{sec:casestud:rtp}). 
-Using this relation, there are two projection functions: (1) the existential image $\tau_\sim\left(\src{\pi}\right)$ and (2) the universal image $\sigma_\sim\left(\trg{\pi}\right)$.
-These projections map source-level (resp. target-level) properties to target-level (resp. source-level) properties. 
+This relation induces the following two projection functions: (1) the existential image $\tau_\sim\left(\src{\pi}\right)$ and (2) the universal image $\sigma_\sim\left(\trg{\pi}\right)$.
+These projections map source-level (resp. target-level) properties to target-level (resp. source-level) properties in a way that identifies the ``same'' property across languages. 
 The case study of this paper exclusively uses the universal image, since some considered properties, such as \gls*{ss}, are not definable in a higher-level language that, e.g., does not model speculation.
@@
 
-Using these projection functions, the more general version of robust preservation is defined as follows~\cite{abate2021extacc}:
+Using these projection functions, the more general version of robust preservation is defined as follows~\cite{abate2021extacc}.
 % For a \bul{compiler $\cc{\src{L}}{\trg{L}}$ to robustly preserve a class of source properties $\src{\class}$}, if for any \rul{property $\src{\pi}$ of that class $\src{\class}$ and programs $\src{p}$ written in $\src{L}$} where \iul{$\src{\progvar}$ robustly satisfies $\src{\pi}$}, then \oul{the compilation of $\src{\progvar}$, $\cc{\src{L}}{\trg{L}}\left(\src{p}\right)$, must robustly satisfy $\tau_\sim\left(\src{\pi}\right)$}.
@@
 % 
-A \bul{compiler $\cc{\src{L}}{\trg{L}}$ to robustly preserves a class of target properties $\trg{\class}$}, if for any \rul{property $\trg{\pi}$ of class $\trg{\class}$ and programs $\src{p}$}, where \iul{$\src{\progvar}$ robustly satisfies $\sigma_\sim\left(\trg{\pi}\right)$}, \oul{the compilation of $\src{\progvar}$, $\cc{\src{L}}{\trg{L}}\left(\src{p}\right)$, must robustly satisfy $\trg{\pi}$}.
+A \bul{compiler $\cc{\src{L}}{\trg{L}}$ robustly preserves a class of target properties $\trg{\class}$}, if for any \rul{property $\trg{\pi}$ of class $\trg{\class}$ and programs $\src{p}$}, where \iul{$\src{\progvar}$ robustly satisfies $\sigma_\sim\left(\trg{\pi}\right)$}, \oul{the compilation of $\src{\progvar}$, $\cc{\src{L}}{\trg{L}}\left(\src{p}\right)$, must robustly satisfy $\trg{\pi}$}.
 
@@
 Because of this, this paper may write $\rtptau{\cc{\src{L}}{\trg{L}}}{\trg{\pi}}{\sim}$, even though $\trg{\pi}$ is a property and not a class.
-A similar construction can be applied to the projection functions.
-
-A degenerate case where $\sim$ coincides with equality, i.e., the trace model is the same for both source and target programs and the compiler does not have an effect on the resulting traces, is the classic definition of robust preservation~\cite{abate2019jour}:
-
+A similar construction can be applied to the projection functions.\MP{wut?}
+
+In case the trace model is the same for both source and target programs (and thus $\sim$ is equality), we have the classic definition of robust preservation~\cite{abate2019jour}.
 \begin{definition}[Robust Preservation]\label{def:rtp}

### 2024-05-27 csf2/main.tex
flow nit

@@
 
-\MP{flow}
-\MKin{
   % last sentence is before the example, which talks about compiler passes enforcing props
-Whether or not compiler passes enforce certain security properties, it is important for a security-aware compiler to consider partial programs that interact with potentially malicious library code, since that code may lead to, e.g., memory--safety issues in the considered partial program.
-}
+Whether or not compiler passes enforce certain security properties, it is important for a secure compiler to consider partial programs that interact with potentially malicious code, since the latter may lead to, e.g., memory-safety issues in the considered partial program.
 

### 2024-05-27 csf2/main.tex
spec

--- 2024-05-27 csf2/main.tex
@@
 
-To introduce the security argument of this paper, this section defines (security) properties, their satisfaction, and their robust satisfaction (i.e., satisfaction in the presence of an active attacker; \Cref{subsec:bg:tprop}).
-Then, borrowing from existing work~\cite{abate2019jour,abate2021extacc}, this section introduces secure compilers as compilers that preserve robust property satisfaction (\Cref{subsec:bg:rtp}).
+To introduce the security argument of this paper, this section first presents the concepts of (security) properties, of their satisfaction, and of their robust satisfaction (i.e., satisfaction in the presence of an active attacker; \Cref{subsec:bg:tprop}).
+Then, borrowing from existing work~\cite{abate2019jour,abate2021extacc}, the section introduces secure compilers as compilers that preserve robust property satisfaction (\Cref{subsec:bg:rtp}).
 
@@
 Throughout the paper, sequences are indicated with an overbar (i.e., $\trace$), empty sequences with $\hole{\cdot}$, and concatenation of sequences $\trace[_{1}],\trace[_{2}]$ as $\trace[_{1}]\cdot\trace[_{2}]$.
-Prepending element $\event$ to a sequence $\trace$ uses the same notation: $\event\cdot\trace$.
+Prepending elements to sequences uses the same notation: $\event\cdot\trace$.
 }
@@
 Properties $\pi$ are sets of traces of admissible program behaviours, ascribing what said property considers valid.
-% 
-The set of all properties can be partitioned into different {\em classes} ($\class$), i.e., safety, liveness, and neither safety nor liveness~\cite{clarkson2008hyper}, so a class is simply a set of properties.
-% 
-The compositionality framework (\Cref{sec:sequential}) presented in this paper considers arbitrary classes, while the case-study (\Cref{sec:casestud:rtp}) fixes them to concrete instances of safety properties, since it is decidable whether a trace satisfies a safety property with just a finite trace (i.e., a \emph{prefix}).
-As an example, consider the trace:
-$$\ev{Dealloc\ \loc}\cdot\ev{Read\ \loc\ 1729}\cdot\dots$$
-which describes the interaction with a memory where the deallocation of an address $\loc$ precedes a read (of some value $1729$) at that address in memory.
-% 
+The set of all properties can be partitioned into different {\em classes} ($\class$), i.e., safety, liveness, and neither safety nor liveness~\cite{clarkson2008hyper}.
+A class is simply a set of properties and for the class of safety properties, it is decidable whether a trace satisfies a safety property with just a finite trace prefix.
+The compositionality framework (\Cref{sec:sequential}) presented in this paper considers arbitrary classes, while the case-study (\Cref{sec:casestud:rtp}) fixes them to concrete instances.
+As an example, consider a trace describing an interaction with a memory where the deallocation of an address $\loc$ precedes a read at that address in memory: $\ev{Dealloc\ \loc}\cdot\ev{Read\ \loc\ 1729}\cdot\dots$.
 This program behaviour is insecure w.r.t.
@@
 In the following, the execution of a whole program $\wholeprogvar$ that terminates in state $\runtimetermvar$ according to the language semantics and produces trace $\trace$ is written as $\progstepto{\wholeprogvar}{\runtimetermvar}{\trace}$.
-With this, we defined property satisfaction as follows:
-\bul{a whole program $\wholeprogvar$ satisfies a property $\pi$} iff \iul{$\wholeprogvar$ yields a trace $\trace$} such that \oul{$\trace$ satisfies $\pi$}.
+With this, property satisfaction is defined as follows:
+\bul{whole programs $\wholeprogvar$ satisfy a property $\pi$} iff \iul{$\wholeprogvar$ yields a trace $\trace$} such that \oul{$\trace$ satisifies $\pi$} (\Cref{def:propsat}).
 
@@
 Property satisfaction is defined on whole programs, i.e., programs without missing definitions.
-Thus, from a security perspective, \Cref{def:propsat} considers only a passive attacker model, where the attacker observes the execution and, e.g., retrieves secrets from that.
-To consider a stronger model similarly to what existing work does~\cite{abate2019jour,abate2021extacc,maffeis2008code-carrying,gordon2003authenticity,fournet2007authorization,bengtson2011refine,backes2014uniontyps,michael2023mswasm,swasey2017robust,sammler2019benefits}, the we extend the concept of satisfaction with {\em robustness}.
+Thus, from a security perspective, this considers only a passive attacker model, where the attacker observes the execution and, e.g., retrieves secrets from that.
+To consider a stronger model similarly to what existing work does~\cite{abate2019jour,abate2021extacc,maffeis2008code-carrying,gordon2003authenticity,fournet2007authorization,bengtson2011refine,backes2014uniontyps,michael2023mswasm,swasey2017robust,sammler2019benefits}, the concept of satisfaction can be extended with {\em robustness}.
 Robust satisfaction considers partial programs $\progvar$, i.e., components with missing imports, which cannot run until said imports are fulfilled.
@@
 this paper assumes that whole programs are the result of linking partial programs referred to as {\em context} ($\ctx$) and {\em component} ($\comp$).
-The context is an arbitrary program and thus has the role of an {\em attacker} that can interact with the component by means of any features the programming language has, and the component is what is security-relevant.
+The context is an arbitrary program and thus has the role of an {\em attacker} that can interact with the component by means of whatever features the programming language has, and the component is what is security-relevant.
 %In this work, the semantics of the programming language is expected to differentiate between the component and the context.
-With this, \Thmref{def:propsat} can be extended as follows: for \bul{a component $\progvar$ to robustly satisfy a property $\pi$}, take an \iul{attacker context $\contextvar$ and link it with $\progvar$}, \oul{the resulting whole program must satisfy $\pi$}.
+With this, \Thmref{def:propsat} can be extended as follows: for \bul{components $\progvar$ to robustly satisfy a property $\pi$}, take an \iul{attacker context $\contextvar$ and link it with $\progvar$}, \oul{the resulting whole program must satisfy $\pi$}.
 
@@
 This translation is considered {\em correct} if it is semantics-preserving.
-That is, for a whole program $\src{\wholeprogvar}$, the compiler should relate the $\src{L}$ semantics of $\src{\wholeprogvar}$ with the semantics of $\trg{L}$ of the compiled counterpart of $\src{p}$ in such a way that they are ``compatible''.
+That is, for a whole program $\src{\wholeprogvar}$, the compiler should relate the $\src{L}$ semantics of $\src{\wholeprogvar}$ with the semantics of $\trg{L}$ of the compiled counterpart of $\src{p}$ in such a way that they are ''compatible''.
 Unfortunately, correct compilers may be insecure compilers~\cite{patrignani2019survey,kennedy2006secure.net,abadi1999protect,ahmed2018dagstuhl} and programs translated by insecure compilers can violate security properties that the programmer assumes to hold.
@@
 
-This paper uses a general notion of robust preservation~\cite{abate2021extacc} that considers compilers that use languages with different trace models. 
+This paper uses a more general notion of robust preservation~\cite{abate2021extacc} that applies to compilers that use languages with different trace models. 
 To this end, considering a source trace $\src{\trace}$ and a target trace $\trg{\trace}$, there is a relation ($\sim$) describing the effect of a corresponding compiler (see \Cref{sec:casestud:rtp}). 
-This relation induces the following two projection functions: (1) the existential image $\tau_\sim\left(\src{\pi}\right)$ and (2) the universal image $\sigma_\sim\left(\trg{\pi}\right)$.
-These projections map source-level (resp. target-level) properties to target-level (resp. source-level) properties in a way that identifies the ``same'' property across languages. 
+Using this relation, there are two projection functions: (1) the existential image $\tau_\sim\left(\src{\pi}\right)$ and (2) the universal image $\sigma_\sim\left(\trg{\pi}\right)$.
+These projections map source-level (resp. target-level) properties to target-level (resp. source-level) properties. 
 The case study of this paper exclusively uses the universal image, since some considered properties, such as \gls*{ss}, are not definable in a higher-level language that, e.g., does not model speculation.
@@
 
-Using these projection functions, the more general version of robust preservation is defined as follows~\cite{abate2021extacc}.
+Using these projection functions, the more general version of robust preservation is defined as follows~\cite{abate2021extacc}:
 % For a \bul{compiler $\cc{\src{L}}{\trg{L}}$ to robustly preserve a class of source properties $\src{\class}$}, if for any \rul{property $\src{\pi}$ of that class $\src{\class}$ and programs $\src{p}$ written in $\src{L}$} where \iul{$\src{\progvar}$ robustly satisfies $\src{\pi}$}, then \oul{the compilation of $\src{\progvar}$, $\cc{\src{L}}{\trg{L}}\left(\src{p}\right)$, must robustly satisfy $\tau_\sim\left(\src{\pi}\right)$}.
@@
 % 
-A \bul{compiler $\cc{\src{L}}{\trg{L}}$ robustly preserves a class of target properties $\trg{\class}$}, if for any \rul{property $\trg{\pi}$ of class $\trg{\class}$ and programs $\src{p}$}, where \iul{$\src{\progvar}$ robustly satisfies $\sigma_\sim\left(\trg{\pi}\right)$}, \oul{the compilation of $\src{\progvar}$, $\cc{\src{L}}{\trg{L}}\left(\src{p}\right)$, must robustly satisfy $\trg{\pi}$}.
+A \bul{compiler $\cc{\src{L}}{\trg{L}}$ to robustly preserves a class of target properties $\trg{\class}$}, if for any \rul{property $\trg{\pi}$ of class $\trg{\class}$ and programs $\src{p}$}, where \iul{$\src{\progvar}$ robustly satisfies $\sigma_\sim\left(\trg{\pi}\right)$}, \oul{the compilation of $\src{\progvar}$, $\cc{\src{L}}{\trg{L}}\left(\src{p}\right)$, must robustly satisfy $\trg{\pi}$}.
 
@@
 Because of this, this paper may write $\rtptau{\cc{\src{L}}{\trg{L}}}{\trg{\pi}}{\sim}$, even though $\trg{\pi}$ is a property and not a class.
-A similar construction can be applied to the projection functions.\MP{wut?}
-
-In case the trace model is the same for both source and target programs (and thus $\sim$ is equality), we have the classic definition of robust preservation~\cite{abate2019jour}.
+A similar construction can be applied to the projection functions.
+
+A degenerate case where $\sim$ coincides with equality, i.e., the trace model is the same for both source and target programs and the compiler does not have an effect on the resulting traces, is the classic definition of robust preservation~\cite{abate2019jour}:
+
 \begin{definition}[Robust Preservation]\label{def:rtp}
@@
 $\Lspec$ extends $\Lscct$ (\Cref{subsec:lscct}) with speculative dynamic semantics. 
-After branching or accessing memory, speculation can start and run subsequent code in a predetermined window until a rollback of operational state is performed. 
+After branching or accessing memory, speculation can start, pushes the current configuration on the speculation register ($\ird{S}$), and run subsequent code in a predetermined window until a rollback of operational state is performed.
+The window is set inside the control-flow state ($\ird{\cfstate}$), whose details are in the technical report.
 Within such windows, data is leaked, marked as high $\ird{\lock}$ and with an annotation that indicates the respective speculative execution variant, in this case just SPECTRE-PHT~\cite{kocher2019spectre}, whose semantics is sketched with \Cref{tr:e-ifz-true-spec}, where $\omega$ is the size of the speculation window.
@@
     \text{(Base-Events)} & \ird{\preevent} &:=& \dots \mid \ird{Spec} \mid \ird{Rlb} \mid \ird{Barrier}\\
-    \text{(Expressions)} & \ird{\expr} &:=& \dots \mid \ird{\lbarrier}
+    \text{(Expressions)} & \ird{\expr} &:=& \dots \mid \ird{\lbarrier} \\
+    \text{(Speculation Register)} & \ird{S} &:=& \ird{\hole{\cdot}} \mid \left(\ird{\Omega};\ird{\expr}\right)
   \end{array}
@@
     \ird{\cfstate[']}.\ird{\operatorname{win}} = \omega &
-    \ird{\trace} = \ird{\ev{Branch}\ 0;\sandboxtag;\securitytag}\cdot\ird{\ev{Spec};\sandboxtag;\securitytag}
+    \ird{\trace} = \ird{\ev{Branch}\ 0;\sandboxtag;\securitytag}\cdot\ird{\ev{Spec};\sandboxtag;\securitytag} \\
+    \ird{S} = ({\ird{\cfstate;\sandboxtag;D;\memstate}};{\ird{\expr[_2]}}),\ird{\hole{\cdot}}
   }{
-    \pstepto{\rtt{\ird{\cfstate;\sandboxtag;0;\memstate}}{\ird{\lifz{0^{\securitytag}}{\expr[_1]}{\expr[_2]}}}}{\begin{array}{c}\rtt{\ird{\cfstate['];\sandboxtag;0;\memstate}}{\ird{\expr[_1]}}\\\rtt{\ird{\cfstate;\sandboxtag;0;\memstate}}{\ird{\expr[_2]}}\end{array}}{\ird{\trace}}
+    \pstepto{\rtt{(\ird{\cfstate;\sandboxtag;D;\memstate});\ird{\hole{\cdot}}}{\ird{\lifz{0^{\securitytag}}{\expr[_1]}{\expr[_2]}}}}{\rtt{(\ird{\cfstate['];\sandboxtag;D;\memstate});\ird{S}}{\ird{\expr[_1]}}}{\ird{\trace}}
   }{e-ifz-true-spec}
@@
     \ird{\cfstate}.\ird{\operatorname{win}} > 0 &
-    \pstepto{\rtt{\ird{\cfstate;\sandboxtag;0;\memstate}}{\ird{\expr}}}{\rtt{\ird{\cfstate['];\sandboxtag;0;\memstate}}{\ird{\expr[']}}}{\ird{\event}}
+    \pstepto{\rtt{(\ird{\cfstate;\sandboxtag;D;\memstate};\ird{\hole{\cdot}})}{\ird{\expr}}}{\rtt{(\ird{\cfstate['];\sandboxtag;D;\memstate};\ird{\hole{\cdot}})}{\ird{\expr[']}}}{\ird{\event}}
     \\
-    \ird{\cfstate['']}.\ird{\operatorname{win}} = \ird{\cfstate[']}.\ird{\operatorname{win}}-1
+    \ird{\cfstate['']}.\ird{\operatorname{win}} = \ird{\cfstate[']}.\ird{\operatorname{win}}-1 &
+    \ird{S} \not=\ird{\hole{\cdot}}
   }{
-    \ghoststepto{\begin{array}{c}\rtt{\ird{\cfstate;\sandboxtag;0;\memstate}}{\ird{\expr}}\\\ird{r}\end{array}}{\begin{array}{c}\rtt{\ird{\cfstate[''];\sandboxtag;0;\memstate}}{\ird{\expr[']}}\\\ird{r}\end{array}}{\ird{\event}}
+    \ghoststepto{\rtt{(\ird{\cfstate;\sandboxtag;D;\memstate};\ird{S})}{\ird{\expr}}}{\rtt{(\ird{\cfstate[''];\sandboxtag;D;\memstate};\ird{S})}{\ird{\expr[']}}}{\ird{\event}}
   }{e-spec-eat}
@@
   \typerule{$e-\ird{spec}-\text{eaten}$}{
-    \ird{r}.\ird{\cfstate}.\ird{\operatorname{win}} = 0
+    \ird{\cfstate}.\ird{\operatorname{win}} = 0 &
+    \ird{S} = \left(\ird{\Omega};\ird{\expr[']}\right)
   }{
-    \ghoststepto{\begin{array}{c}\ird{r}\\\ird{r'}\end{array}}{\ird{r'}}{\ird{Rlb}}
+    \ghoststepto{\rtt{(\ird{\cfstate;\sandboxtag;D;\memstate};\ird{S})}{\ird{\expr}}}{\rtt{(\ird{\Omega};\ird{\hole{\cdot}})}{\ird{\expr[']}}}{\ird{\event}}
   }{e-spec-eaten}
@@
   \typerule{$e-\ird{spec}-\text{barrier}$}{
-    \ird{r'}.\ird{\cfstate} = \ird{r}.\ird{\cfstate}[.\ird{\operatorname{win}} = 0]
+    \ird{\cfstate'} = \ird{\cfstate}[.\ird{\operatorname{win}} = 0] &
+    \ird{S} \not=\ird{\hole{\cdot}}
   }{
-    \ghoststepto{\begin{array}{c}\ird{r}\\\ird{r''}\end{array}}{\begin{array}{c}\ird{r'}\\\ird{r''}\end{array}}{\ird{Barrier}}
+    \ghoststepto{\rtt{(\ird{\cfstate;\sandboxtag;D;\memstate};\ird{S})}{\ird{\lbarrier}}}{\rtt{(\ird{\cfstate';\sandboxtag;D;\memstate};\ird{\hole{\cdot}})}{\ird{\lbarrier}}}{\ird{\event}}
   }{e-spec-barrier}
@@
 The trace model is extended with $\ird{\ev{Spec}}$, $\ird{\ev{Rlb}}$, and $\ird{\ev{Barrier}}$ events that signal the respective operational action. 
-A $\ird{\lbarrier}$ prevents any execution whatsoever besides $\ird{\ev{Rlb}}$ when run in speculative execution mode (see \Cref{tr:e-spec-barrier}) and does nothing when run in normal mode.
+A $\ird{\lbarrier}$ prevents any execution whatsoever besides $\ird{\ev{Rlb}}$ when run in speculative execution mode and does nothing when run in normal mode, e.g., \Cref{tr:e-spec-barrier}.
+\footnote{In the rule, the notation $\cfstate' = \cfstate[.\text{win} = 0]$ means that $\cfstate'$ is a copy of $\cfstate$ up to field $\text{win}$, which should be set to $0$.}
+
 The speculative semantics are inspired by existing work~\cite{guarnieri2018spectector}.

### 2024-05-27 csf2/main.tex
remove inline todo
sec 3

--- 2024-05-27 csf2/main.tex
@@
 
+\smallskip%MP: the underline compacts the sections too much
+
 \section{Secure Composition}\label{sec:sequential}
@@
 Notably, real-world compilation chains also perform a series of (sequential) passes whose main purpose is not necessarily to translate from one language to another, but to, e.g., optimise the code or enforce a certain property.
-Both examples can be seen in practice, e.g. as in the work of~\cite{nagarakatte2009soft,nagarakatte2010cets,akritidis2009baggy,wegman1991ccp,manjikian1997fusion} and many more.
+Both examples can be seen in practice, e.g.,~\cite{nagarakatte2009soft,nagarakatte2010cets,akritidis2009baggy,wegman1991ccp,manjikian1997fusion} and many more.
 % \MPin{survey to cite?
@@
 
-For starters, suppose there are traces $\src{\trace}$ and $\trg{\trace}$ as well as two trace relations $\sim_1$ and $\sim_2$. 
+\smallskip
+
+The key notions for our main results are the definition of a trace relation, and the definition of when is a trace relation well-formed with respect to a class.
+%
+Consider traces $\src{\trace}$ and $\trg{\trace}$ as well as two trace relations $\sim_1$ and $\sim_2$. 
+% 
 The traces are related $\src{\trace}\sim_1\bullet\sim_2\trg{\trace}$ if there is another trace $\irl{\trace}$ such that $\src{\trace}\sim_1\irl{\trace}$ and $\irl{\trace}\sim_2\trg{\trace}$.
-
+% 
 \bul{A relation $\sim$ is well-formed w.r.t. a class of target-level properties $\trg{\class}$} iff \rul{the universal image preserves set membership}.
@@
 
-With these notions of well-formedness, it is now possible to prove the main result that secure compilers in the robust compilation framework~\cite{abate2021extacc} compose {\em sequentially}.
+We can now state our main result: secure compilers in the robust compilation framework~\cite{abate2021extacc} compose {\em sequentially}. 
+%
 
@@
 Since the composition of secure compilers is again a secure compiler, the theorems generalise to a whole chain of $n$ secure compilers. 
-Moreover, the previously defined theory dually holds for the existential image $\tau_\sim\left(\src{\pi}\right)$, but these results have been moved to the technical report in the interest of space.
+% 
+Note that the previous results also hold for the existential image $\tau_\sim\left(\src{\pi}\right)$, but these results have been moved to the technical report in the interest of space.
 
@@
 
-\Cref{corr:rtp} provides an easy to use way to compose secure compilers without any caveats. 
+\Cref{corr:rtp} provides an easy way to compose secure compilers without well-formedness of trace relations. 
+% 
 However, while \Cref{thm:rtpsim:sig} explicitly requires that $\sim_1$ is well-formed w.r.t. $\trg{\class[_2]}$, if $\sim_2$ is not well-formed w.r.t. $\trg{\class[_1]}$, care must be taken. 
 %Worst-case is that the resulting class $\sigma_{\sim_1\bullet\sim_2}\left(\trg{\class[_1]}\cap\trg{\class[_2]}\right)$ becomes empty, even though $\trg{\class[_1]}\cap\trg{\class[_2]}$ is not necessarily empty.
-\Cref{subsec:compatsecpasses} provides an example of this phenomenon.
+This is further discussed in \Cref{subsec:compatsecpasses}.
 
@@
 Besides sequential composition, there are two other compositions, namely an {\em upper}, i.e., a compiler that takes multiple inputs and yields one output, and a {\em lower} composition, i.e., a compiler that takes one input and yields multiple outputs.
-{Define the upper composition $\cc{\src{L}+\obj{L}}{\trg{L}}$} as follows:
+We {fefine the upper composition $\cc{\src{L}+\obj{L}}{\trg{L}}$} as follows:
 Given a program \texttt{p}, its compiled counterpart is obtained by {plugging \texttt{p} into $\cc{\src{L}}{\trg{L}}$ if $\texttt{p}\in\src{L}$} or by {plugging \texttt{p} into $\cc{\obj{L}}{\trg{L}}$ if $\texttt{p}\in\obj{L}$}.
@@
 Future work intends to use upper and lower composition as means to build a secure compiler incrementally.
+\MPin{
+	we should add briefly which real-world compilers (taken from the elided examples) are upper and lower compos
+}
+\MP{no coq here ?}
 
@@
 This section introduces trace properties of interest for this paper: \gls*{tms}, \gls*{sms}, \gls*{ms}, \gls*{scct}, and \gls*{ss}.
-These properties are of practical importance (as mentioned in \Cref{sec:introduction}) and also of interest in the case study (\Cref{sec:casestud:defs,sec:casestud:rtp}) presented later. 
+These properties are of practical importance (as mentioned in \Cref{sec:introduction}) and also of interest in the case study presented later (\Cref{sec:casestud:defs,sec:casestud:rtp}). 
 While up to \gls*{scct}, none of the properties is novel, for self-containment of the paper, their definition is necessary for the case study.
@@
 
-Note that \Cref{def:trace:msdef} ignores data isolation, there may still be memory-safety issues introduced by side-channels.
-
-\subsection{Memory Safety Trace Model with Constant Timeness}\label{subsec:scct:tracemodel}
-
-This section starts off by extending the memory safety trace model (\Cref{subsec:basic:memsafety:tracemodel}) with a {\em security tag} ($\securitytag$) that indicates whether the involved event contains sensitive information ($\lock$) or not ($\unlock$).
-
-\vspace{-1em}
+Note that \Cref{def:trace:msdef} ignores data isolation, so there may still be memory-safety issues introduced by side-channels.
+
+\subsection{Memory Safety Trace Model with Constant Time}\label{subsec:scct:tracemodel}
+
+This section starts by extending the memory safety trace model (\Cref{subsec:basic:memsafety:tracemodel}) with a {\em security tag} ($\securitytag$) that indicates whether the involved event contains sensitive information ($\lock$) or not ($\unlock$).
+
+\vspace{-.5em}
 \[
@@
 In turn, an instruction whose timing is data-dependent must not have a secret as an operand.
-Typical operations with data-dependent timing are branches and certain binary operations, such as division. 
+Typical operations with data-dependent timing are branches and certain binary operations, such as division.%
 \footnote{
@@
 }
-Both operations are also modelled in the trace model by extending the set of base-events with branches ($\ev{Branch\ n}$) and binary operations ($\ev{Binop\ n}$).
+Both operations are represented in the trace model by extending the set of base-events with branches ($\ev{Branch\ n}$) and binary operations ($\ev{Binop\ n}$).
 

### 2024-05-27 csf2/main.tex
more on four

@@
 However, \gls*{ms} uses a different trace model than \gls*{scct}, so intersecting them would trivially yield the empty set. 
-To remedy this issue, let $\sim_{\ctsafe}$ be a cross-language trace relation whose definition can be sketched as follows:
+To remedy this issue, let $\sim_{\ctsafe}$ be a cross-language trace relation whose definition can be sketched as follows:\MP{why just sketched?}
 \[
@@
 
-\begin{definition}[\glsfirst*{ms} and \gls*{scct}]\label{def:trace:msscctdef}
+\begin{definition}[\gls*{ms} and \gls*{scct}]\label{def:trace:msscctdef}
   $
-  \msscctsafe:=\sigma_{\sim_{msscct}}\left(\mssafe\right)\cap\scctsafe
+  \msscctsafe:=\sigma_{\sim_{mct}}\left(\mssafe\right)\cap\scctsafe
   $
@@
 
-So far, the considered trace models do not take speculative execution into account. 
-For this, extend the earlier trace model (see \Cref{subsec:scct:tracemodel}) so that the security tags ($\securitytag$) carry additional information about the kind of private data leakage, i.e., the type of speculative leak.
+So far, the considered trace models do not let us express speculative execution attacks such as Spectre~\cite{}\MP{cits}. 
+For this, we extend the earlier trace model (see \Cref{subsec:scct:tracemodel}) so that the security tags ($\securitytag$) carry additional information about the kind of private data leakage, i.e., the type of speculative leak.
 Moreover, add base-events signaling the beginning of a speculative execution ($\ev{Spec}$), a barrier ($\ev{Barrier}$) that signals that any speculative execution may not go past it, as well as a rollback event ($\ev{Rlb}$), which signals that execution resumes to where speculation started.

### 2024-05-27 csf2/main.tex

@@
 Whether or not compiler passes enforce certain security properties, it is important for a secure compiler to consider partial programs that interact with potentially malicious code, since the latter may lead to, e.g., memory-safety issues in the considered partial program.
-
 Robust compilers~\cite{abate2019jour} are a form of secure compilers that preserve security properties even in the presence of arbitrary attackers interacting with compiled code.

### 2024-05-27 csf2/main.tex
section 4 done

@@
 For this, we extend the earlier trace model (see \Cref{subsec:scct:tracemodel}) so that the security tags ($\securitytag$) carry additional information about the kind of private data leakage, i.e., the type of speculative leak.
-Moreover, add base-events signaling the beginning of a speculative execution ($\ev{Spec}$), a barrier ($\ev{Barrier}$) that signals that any speculative execution may not go past it, as well as a rollback event ($\ev{Rlb}$), which signals that execution resumes to where speculation started.
+Moreover, we add base-events signalling the beginning of a speculative execution ($\ev{Spec}$), a barrier ($\ev{Barrier}$) that signals that any speculative execution may not go past it, as well as a rollback event ($\ev{Rlb}$), which signals that execution resumes to where speculation started.
 
@@
 
-Even though the considered Spectre variants are just SPECTRE-PHT~\cite{kocher2019spectre}, NONE just describes secret data as in \gls*{scct} (see \Cref{subsec:scct:tracemodel}), the trace model is general enough to allow for potential future extension with different variants.
+Even though the considered Spectre variants are just SPECTRE-PHT~\cite{kocher2019spectre}, NONE just describes secret data as in \gls*{scct} (see \Cref{subsec:scct:tracemodel}), the trace model is general enough to allow for potential future extension with different variants.\MP{cits}
 %For sake of readability, this paper just uses the notation $\lock$ in place of $\lock_{\text{PHT}}$.
@@
 The definition of the property itself turns out to be syntactically equal to \gls*{scct}, modulo the underlying trace model.
+\MPin{recap its meaning for security}
 
@@
 \end{definition}
+\MPin{why isn't it ok to have locks with NONE in the trace?}
 
@@
 Similar to \gls*{msscct}, there is a need to relate the different trace models with each other, so that the memory safety property without speculation can be lifted to speculation. 
-To this end, let $\sim_{\mathghost}$ be a cross-language trace relation whose definition can be sketched as follows:
+To this end, let $\sim_{\mathghost}$ be a cross-language trace relation whose definition can be sketched as follows:\MP{sketch?}
 
@@
 
-Now, define \gls*{specms} as: 
-\begin{definition}[\glsfirst*{specms}]\label{def:trace:specmsdef}
+We conclude this section by defining the ultimate property of interest for secure compilers: \gls*{specms}.
+\begin{definition}[\gls*{specms}]\label{def:trace:specmsdef}
   $

### 2024-05-27 csf2/main.tex
lang section todos

@@
 
-%Either way, the theoretical results suggest that it is possible to always find a ,,most-general'', secure compiler, given two secure compilers, that robustly preserves the least-upper bound of the classes involved in their compilation process.
+%Either way, the theoretical results suggest that it is possible to always find a ``most-general'', secure compiler, given two secure compilers, that robustly preserves the least-upper bound of the classes involved in their compilation process.
 
@@
 This section defines programming languages $\Ltms$, $\Ltrg$, $\Lms$, $\Lscct$, and $\Lspec$, all of which share common elements (presented in \Cref{subsec:cs:defs}).
-In the case study of this paper, $\Ltms$ is the only statically typed language and exhibits the property that all well-typed programs are \gls*{tms} (\Cref{subsec:ltms}).
+$\Ltms$ is the only statically typed language and exhibits the property that all well-typed programs are \gls*{tms} (\Cref{subsec:ltms}).
 However, not all $\Ltms$ programs are \gls*{sms}.
@@
 All three languages --- so $\Ltms$, $\Ltrg$, and $\Lms$ --- assume \gls*{cct} to hold, since this is -- in an ideal world -- what the programmer would expect, too: It is the job of the compiler to preserve and (potentially) enforce security~\cite{cauligi2019fact,nagarakatte2010cets,nagarakatte2009soft,akritidis2009baggy}.
-
 % This paper considers \gls*{cct} as an architecture-dependent property\MPin{can we justify this}, so
+
 Such consideration is also backed up by architecture providing a data (operand) independent timing mode, such as processors by Arm~\cite[p.~543]{arm-refman} and Intel~\cite[p.~80]{intel-refman}.
 %In spirit of this, all languages besides $\Lscct$ are trivially \gls*{scct}. 
-$\Lscct$ (\Cref{subsec:lscct}) programs have access to a ``\gls*{cct}-mode'' and can change the leakage of emitted events according to the value of this mode (either $\obj{ON}$ or $\obj{OFF}$).
-
-Lastly, modern processors also employ speculative execution to achieve speedups, but fail to roll back properly, so that an attacker can exploit side-channels~\cite{kocher2019spectre}. 
-$\Lspec$ (\Cref{subsec:lspec}) extends $\Lscct$ to model this situation, while all previous languages trivially satisfy \gls*{ss}, since they do not support speculative execution at all.
-All languages share large parts of syntactic elements, which are presented in the next subsection (\Cref{subsec:cs:defs}).
-Subsequent subsections extend on this accordingly.
+This kind of processor feature is modelled in language $\Lscct$ (\Cref{subsec:lscct}), where programs have access to a ``\gls*{cct}-mode'' and can change the leakage of emitted events according to the value of this mode (either $\obj{ON}$ or $\obj{OFF}$).
+
+Finally, modern processors also employ speculative execution to achieve speedups---and unfortunately generate Spectre attacks~\cite{kocher2019spectre}---and this is the extension of $\Lspec$ (\Cref{subsec:lspec}).
+ % extends $\Lscct$ to model speculative execution.
+%
+Thus, all previous languages trivially satisfy \gls*{ss}, since they do not support speculative execution at all.
+% All languages share large parts of syntactic elements, which are presented in the next subsection (\Cref{subsec:cs:defs}).
+% Subsequent subsections extend on this accordingly.
 
@@
 For example, a context immediately deallocating an allocated memory region twice trivially violates memory safety, but the main interest in secure compilation is at component-level events and if those signal, e.g., a memory safety issue.
+
+\MPin{this section does not define anything, not even the sandbox tag metavariable}
+\MPin{
+	move b/t/a/e/F/CHI/H/Delta/Phi/Psi to here.
+	their explanation too.
+	make black.
+}
+\MPin{
+	what are pointer maps and Control states?
+}
 
@@
 
-Operational state ($\src{\Omega}$) is a triple consisting of control-flow state ($\src{\cfstate}$), a control tag ($\src{\sandboxtag}$), and a memory state ($\memstate$), which carries information about the memory.
-Because pointers are kept ,,alive'' in pointer maps ($\src{\Delta}$), $\Ltms$ dynamic semantics do not get stuck if a memory-safety violation is encountered, such as a double--free.
+Operational state ($\src{\Omega}$) is a triple consisting of control-flow state ($\src{\cfstate}$), a control tag ($\src{\sandboxtag}$), and a memory state ($\src{\memstate}$), which carries information about the memory.
+Because pointers are kept ``alive'' in pointer maps ($\src{\Delta}$), $\Ltms$ dynamic semantics does not get stuck if it encounters a memory-safety violation such as a double--free.
 The memory state carries two heaps to model sandboxing between a context and a component.
+\MP{notation}
 
@@
 With these tags, it is possible to ignore effects performed by a spurious context that, e.g., immediately deallocates a memory location twice, thus violating \gls*{tms}.
-Even though this could be used for blame preservation~\cite{patrignani2023blame}, this is not what this paper is concerned about.
+Even though this could be used for blame preservation~\cite{patrignani2023blame}, this beyond the focus of this paper.
 
@@
 Furthermore, the syntax of $\Ltrg$ is extended with a way to inspect whether a pointer is freed ($\trg{\lispoisoned{x}}$), evaluating to $\trg{0}$ if it is freed and to $\trg{1}$ otherwise.
-Functions may receive arguments that are not $\trg{\natt}$ and values are extended with tuples.
+Functions may receive arguments that are not $\trg{\natt}$ and values are extended with tuples\MP{add this?}\MP{no values in previous grammar. add?}.
 
@@
 $\Lscct$ extends $\Lms$ (\Cref{subsec:lms}) with a way to read from ($\obj{\lrddoit{x}{\expr}}$) and to write to ($\obj{\lwrdoit{D}}$) a model specific register that controls a data (operand) independent timing mode, a feature that is present in both Arm~\cite[p.~543]{arm-refman} and Intel~\cite[p.~80]{intel-refman} processors.
-To this end, states ($\src{\Omega}$) are extended with the value of the register ($\obj{D}$), which is initially set to be not active ($\obj{OFF}$).
+To this end, states ($\obj{\Omega}$) are extended with the value of the register ($\obj{D}$), which is initially set to be not active ($\obj{OFF}$).
 If the register is marked active ($\obj{ON}$), the intuition is that no secrets can appear on specification traces.
@@
 
-The language also adds user-annotations for secrecy $\obj{\securitytag}$, which is either high $\obj{\lock}$ or low $\obj{\unlock}$.
-Security tags $\obj{\securitytag}$ are on the usual secrecy lattice, where $\obj{\lock}\le\obj{\securitytag}$ and $\obj{\securitytag}\le\obj{\unlock}$.
+The language also adds user annotations $\obj{\securitytag}$ for the secrecy of variables, which can be either private (high secrecy) $\obj{\lock}$ or public (low secrecy) $\obj{\unlock}$.
+Security tags $\obj{\securitytag}$ are arranged in the usual secrecy lattice\MP{NI cits}, where $\obj{\lock}\le\obj{\securitytag}$ and $\obj{\securitytag}\le\obj{\unlock}$.
 To prevent secrets from leaking but still enable reasoning about memory safety, $\Lscct$ extends base-events with $\obj{\ev{iGet}\ \loc\ \valueexpr}$ and $\obj{\ev{iSet}\ \loc\ \valueexpr}$.
+\MPin{what's the i for? explain. intuition missing}
 This way, memory accesses to secret data are still exposed as memory accesses, despite being protected by data-independent timing mode, e.g, \Cref{tr:to-e-get-in-noleak}.
@@
 \end{exampleenv}
-
-\subsection{$\Lspec$: A Memory-Unsafe Language with Speculation}\label{subsec:lspec}
-
-$\Lspec$ extends $\Lscct$ (\Cref{subsec:lscct}) with speculative dynamic semantics. 
-After branching or accessing memory, speculation can start, pushes the current configuration on the speculation register ($\ird{S}$), and run subsequent code in a predetermined window until a rollback of operational state is performed.
-The window is set inside the control-flow state ($\ird{\cfstate}$), whose details are in the technical report.
-Within such windows, data is leaked, marked as high $\ird{\lock}$ and with an annotation that indicates the respective speculative execution variant, in this case just SPECTRE-PHT~\cite{kocher2019spectre}, whose semantics is sketched with \Cref{tr:e-ifz-true-spec}, where $\omega$ is the size of the speculation window.
-Additionally, $\ird{\lock}$ may also carry a $\ird{NONE}$ annotation, to signal that this is a leak irrespective of speculation, i.e., as defined earlier (\Cref{subsec:lscct}).
-The semantics is kept general enough to allow for future extension to support different variants.
-
-In terms of language features, $\Lspec$ supports a barrier operation $\ird{\lbarrier}$ that blocks speculative execution. 
-To facilitate speculative execution in a non-assembly-like language, a stack of operational state is used and speculation is active if that stack has a size greater than one, indicated by writing the operational states vertically.
-This is exploited in \Cref{tr:e-spec-eat} to ,,eat away'' any leftover speculation.
-
-\vspace{-.5em}
-\[
-  \begin{array}{rrcl}
-    \text{(Base-Events)} & \ird{\preevent} &:=& \dots \mid \ird{Spec} \mid \ird{Rlb} \mid \ird{Barrier}\\
-    \text{(Expressions)} & \ird{\expr} &:=& \dots \mid \ird{\lbarrier} \\
-    \text{(Speculation Register)} & \ird{S} &:=& \ird{\hole{\cdot}} \mid \left(\ird{\Omega};\ird{\expr}\right)
-  \end{array}
-\]
-
-\begin{center}
-\newcommand{\expreval}[5]{{#1}\triangleright\xspace {#2}\xrightarrow{#5}\ {#3}\triangleright\xspace {#4}\xspace}
-\newcommand{\exprevald}[5]{\expreval{\ird{#1}}{\ird{#2}}{\ird{#3}}{\ird{#4}}{\ird{#5}}}
-
-  %
-  \typerule{$e-\ird{ifz}-\ird{true}-\text{spec}$}{
-    \ird{\cfstate[']}.\ird{\operatorname{win}} = \omega &
-    \ird{\trace} = \ird{\ev{Branch}\ 0;\sandboxtag;\securitytag}\cdot\ird{\ev{Spec};\sandboxtag;\securitytag} \\
-    \ird{S} = ({\ird{\cfstate;\sandboxtag;D;\memstate}};{\ird{\expr[_2]}}),\ird{\hole{\cdot}}
-  }{
-    \pstepto{\rtt{(\ird{\cfstate;\sandboxtag;D;\memstate});\ird{\hole{\cdot}}}{\ird{\lifz{0^{\securitytag}}{\expr[_1]}{\expr[_2]}}}}{\rtt{(\ird{\cfstate['];\sandboxtag;D;\memstate});\ird{S}}{\ird{\expr[_1]}}}{\ird{\trace}}
-  }{e-ifz-true-spec}
-  %
-  \typerule{$e-\ird{spec}-\text{eat}$}{
-    \ird{\cfstate}.\ird{\operatorname{win}} > 0 &
-    \pstepto{\rtt{(\ird{\cfstate;\sandboxtag;D;\memstate};\ird{\hole{\cdot}})}{\ird{\expr}}}{\rtt{(\ird{\cfstate['];\sandboxtag;D;\memstate};\ird{\hole{\cdot}})}{\ird{\expr[']}}}{\ird{\event}}
-    \\
-    \ird{\cfstate['']}.\ird{\operatorname{win}} = \ird{\cfstate[']}.\ird{\operatorname{win}}-1 &
-    \ird{S} \not=\ird{\hole{\cdot}}
-  }{
-    \ghoststepto{\rtt{(\ird{\cfstate;\sandboxtag;D;\memstate};\ird{S})}{\ird{\expr}}}{\rtt{(\ird{\cfstate[''];\sandboxtag;D;\memstate};\ird{S})}{\ird{\expr[']}}}{\ird{\event}}
-  }{e-spec-eat}
-  % 
-  \typerule{$e-\ird{spec}-\text{eaten}$}{
-    \ird{\cfstate}.\ird{\operatorname{win}} = 0 &
-    \ird{S} = \left(\ird{\Omega};\ird{\expr[']}\right)
-  }{
-    \ghoststepto{\rtt{(\ird{\cfstate;\sandboxtag;D;\memstate};\ird{S})}{\ird{\expr}}}{\rtt{(\ird{\Omega};\ird{\hole{\cdot}})}{\ird{\expr[']}}}{\ird{\event}}
-  }{e-spec-eaten}
-  %
-  \typerule{$e-\ird{spec}-\text{barrier}$}{
-    \ird{\cfstate'} = \ird{\cfstate}[.\ird{\operatorname{win}} = 0] &
-    \ird{S} \not=\ird{\hole{\cdot}}
-  }{
-    \ghoststepto{\rtt{(\ird{\cfstate;\sandboxtag;D;\memstate};\ird{S})}{\ird{\lbarrier}}}{\rtt{(\ird{\cfstate';\sandboxtag;D;\memstate};\ird{\hole{\cdot}})}{\ird{\lbarrier}}}{\ird{\event}}
-  }{e-spec-barrier}
-\end{center}
-
-The trace model is extended with $\ird{\ev{Spec}}$, $\ird{\ev{Rlb}}$, and $\ird{\ev{Barrier}}$ events that signal the respective operational action. 
-A $\ird{\lbarrier}$ prevents any execution whatsoever besides $\ird{\ev{Rlb}}$ when run in speculative execution mode and does nothing when run in normal mode, e.g., \Cref{tr:e-spec-barrier}.
-\footnote{In the rule, the notation $\cfstate' = \cfstate[.\text{win} = 0]$ means that $\cfstate'$ is a copy of $\cfstate$ up to field $\text{win}$, which should be set to $0$.}
-
-The speculative semantics are inspired by existing work~\cite{guarnieri2018spectector}.
-
 \begin{figure}[!htb]
@@
 
+\subsection{$\Lspec$: A Memory-Unsafe Language with Speculation}\label{subsec:lspec}
+
+$\Lspec$ extends $\Lscct$ (\Cref{subsec:lscct}) with speculative dynamic semantics that is inspired by existing work~\cite{guarnieri2018spectector}\MP{cit xaver's}.
+After branching or accessing memory\MP{?}, speculation starts by pushing the current configuration on the speculation register ($\ird{S}$), and run subsequent code in a predetermined window until a rollback of operational state is performed.
+The window is set inside the control-flow state ($\ird{\cfstate}$), whose details are in the technical report.
+Within such windows, data is leaked, marked as high $\ird{\lock}$ and with an annotation that indicates the respective speculative execution variant, in this case just SPECTRE-PHT~\cite{kocher2019spectre}, whose semantics is sketched with \Cref{tr:e-ifz-true-spec}, where $\omega$ is the size of the speculation window\MP{role of omega?}.
+Additionally, $\ird{\lock}$ may also carry a $\ird{NONE}$ annotation, to signal that this is a leak irrespective of speculation, i.e., as defined earlier (\Cref{subsec:lscct}).
+The semantics is kept general enough to allow for future extension to support different variants.
+
+In terms of language features, $\Lspec$ supports a barrier operation $\ird{\lbarrier}$ that blocks speculative execution. 
+To facilitate speculative execution in a non-assembly-like language, a stack of operational state is used and speculation is active if that stack has a size greater than one, indicated by writing the operational states vertically.
+This is exploited in \Cref{tr:e-spec-eat} to consume any leftover speculation.
+
+\vspace{-.5em}
+\[
+  \begin{array}{rrcl}
+    \text{(Base-Events)} & \ird{\preevent} &:=& \dots \mid \ird{Spec} \mid \ird{Rlb} \mid \ird{Barrier}\\
+    \text{(Expressions)} & \ird{\expr} &:=& \dots \mid \ird{\lbarrier} \\
+    \text{(Speculation Register)} & \ird{S} &:=& \ird{\hole{\cdot}} \mid \left(\ird{\Omega};\ird{\expr}\right)
+  \end{array}
+\]
+
+\begin{center}
+\newcommand{\expreval}[5]{{#1}\triangleright\xspace {#2}\xrightarrow{#5}\ {#3}\triangleright\xspace {#4}\xspace}
+\newcommand{\exprevald}[5]{\expreval{\ird{#1}}{\ird{#2}}{\ird{#3}}{\ird{#4}}{\ird{#5}}}
+
+  %
+  \typerule{$e-\ird{ifz}-\ird{true}-\text{spec}$}{
+    \ird{\cfstate[']}.\ird{\operatorname{win}} = \omega &
+    \ird{\trace} = \ird{\ev{Branch}\ 0;\sandboxtag;\securitytag}\cdot\ird{\ev{Spec};\sandboxtag;\securitytag} \\
+    \ird{S} = ({\ird{\cfstate;\sandboxtag;D;\memstate}};{\ird{\expr[_2]}}),\ird{\hole{\cdot}}
+  }{
+    \pstepto{\rtt{(\ird{\cfstate;\sandboxtag;D;\memstate});\ird{\hole{\cdot}}}{\ird{\lifz{0^{\securitytag}}{\expr[_1]}{\expr[_2]}}}}{\rtt{(\ird{\cfstate['];\sandboxtag;D;\memstate});\ird{S}}{\ird{\expr[_1]}}}{\ird{\trace}}
+  }{e-ifz-true-spec}
+  %
+  \typerule{$e-\ird{spec}-\text{eat}$}{
+    \ird{\cfstate}.\ird{\operatorname{win}} > 0 &
+    \pstepto{\rtt{(\ird{\cfstate;\sandboxtag;D;\memstate};\ird{\hole{\cdot}})}{\ird{\expr}}}{\rtt{(\ird{\cfstate['];\sandboxtag;D;\memstate};\ird{\hole{\cdot}})}{\ird{\expr[']}}}{\ird{\event}}
+    \\
+    \ird{\cfstate['']}.\ird{\operatorname{win}} = \ird{\cfstate[']}.\ird{\operatorname{win}}-1 &
+    \ird{S} \not=\ird{\hole{\cdot}}
+  }{
+    \ghoststepto{\rtt{(\ird{\cfstate;\sandboxtag;D;\memstate};\ird{S})}{\ird{\expr}}}{\rtt{(\ird{\cfstate[''];\sandboxtag;D;\memstate};\ird{S})}{\ird{\expr[']}}}{\ird{\event}}
+  }{e-spec-eat}
+  % 
+  \typerule{$e-\ird{spec}-\text{eaten}$}{
+    \ird{\cfstate}.\ird{\operatorname{win}} = 0 &
+    \ird{S} = \left(\ird{\Omega};\ird{\expr[']}\right)
+  }{
+    \ghoststepto{\rtt{(\ird{\cfstate;\sandboxtag;D;\memstate};\ird{S})}{\ird{\expr}}}{\rtt{(\ird{\Omega};\ird{\hole{\cdot}})}{\ird{\expr[']}}}{\ird{\event}}
+  }{e-spec-eaten}
+  %
+  \typerule{$e-\ird{spec}-\text{barrier}$}{
+    \ird{\cfstate'} = \ird{\cfstate}[.\ird{\operatorname{win}} = 0] &
+    \ird{S} \not=\ird{\hole{\cdot}}
+  }{
+    \ghoststepto{\rtt{(\ird{\cfstate;\sandboxtag;D;\memstate};\ird{S})}{\ird{\lbarrier}}}{\rtt{(\ird{\cfstate';\sandboxtag;D;\memstate};\ird{\hole{\cdot}})}{\ird{\lbarrier}}}{\ird{\event}}
+  }{e-spec-barrier}
+\end{center}
+
+The trace model is extended with $\ird{\ev{Spec}}$, $\ird{\ev{Rlb}}$, and $\ird{\ev{Barrier}}$ events that signal the respective operational action. 
+A $\ird{\lbarrier}$ prevents any execution whatsoever besides $\ird{\ev{Rlb}}$ when run in speculative execution mode and does nothing when run in normal mode, e.g., \Cref{tr:e-spec-barrier}.%
+\footnote{In the rule, the notation $\cfstate' = \cfstate[.\text{win} = 0]$ means that $\cfstate'$ is a copy of $\cfstate$ up to field $\text{win}$, which should be set to $0$.}
+
 \section{Case Study: Composing Secure Compiler Passes and Optimisations \pages{7}}\label{sec:casestud:rtp}
@@
 
-The above sketches interesting parts, anything left out is a trivial ,,re-colouring'' from $\src{blue}$ to $\trg{red}$.
+The above sketches interesting parts, anything left out is a trivial ``re-colouring'' from $\src{blue}$ to $\trg{red}$.
 
@@
 \Cref{subsec:compatsecpasses} discusses a caveat with \Cref{thm:rtpsim:sig}, while \Cref{subsec:seccompproofs} gives some additional background on the secure compilation proofs that do not use the compositionality framework (\Cref{sec:sequential}).
-Lastly, \Cref{subsec:formalities:maps} discusses a minor technicality related to the ,,coloring'' of the investigated properties.
+Lastly, \Cref{subsec:formalities:maps} discusses a minor technicality related to the ``coloring'' of the investigated properties.
 
@@
 Traditional proof techniques apply, the usual strategy is to establish a backwards simulation between source components and their compiled target counterpart. 
-For the universally quantified target context, a technique known as backtranslation is well-established, which constructs a source-level context either by backtranslating the target-level context, amounting to a ,,decompilation'', or by constructing it using the target-level trace, so that the source-context emits parts of the whole execution trace that are related to the parts emitted by the target-level context.
+For the universally quantified target context, a technique known as backtranslation is well-established, which constructs a source-level context either by backtranslating the target-level context, amounting to a ``decompilation'', or by constructing it using the target-level trace, so that the source-context emits parts of the whole execution trace that are related to the parts emitted by the target-level context.
 Up to backtranslations for $\ccdce$ and $\cccf$, all backtranslations in the case-study are trace based, while for $\ccdce$ and $\cccf$, they are context-based. 
@@
 It contains traces as specified in \Cref{subsec:basic:memsafety:tracemodel}.
-However, these are different from $\Ltms$ traces (\Cref{subsec:ltms}) which means that \Cref{thm:wt:tms} is ,,ill-typed''.
+However, these are different from $\Ltms$ traces (\Cref{subsec:ltms}) which means that \Cref{thm:wt:tms} is ``ill-typed''.
 Similarly for the other theorems.
@@
 Different mechanisms for enforcing memory safety exist that also consider the secure compilation domain, i.e., have an active attacker model.
-For example, the ,,pointers as capabilities'' principle represents pointers as machine-level capabilities~\cite{korashy2021capableptrs}, which behave in a similar fashion to capabilities by means of linear typing~\cite{morrisett2005L3}.
+For example, the ``pointers as capabilities'' principle represents pointers as machine-level capabilities~\cite{korashy2021capableptrs}, which behave in a similar fashion to capabilities by means of linear typing~\cite{morrisett2005L3}.
 The approach of this paper also uses linear typing, but differs from $L^{3}$~\cite{morrisett2005L3} in the way that functions are not first-class.

### 2024-05-27 csf2/main.tex
comments on 6

@@
   \vspace{-1em}
-  \caption{Visualisation of the optimising compilation pipeline that attains a combination of \gls*{ms} and \gls*{cct}. %
+  \caption{Visualisation of the optimising compilation pipeline that preserves \gls*{mssscct}. %
     Vertices in the graph are the programming languages from earlier sections (\Cref{sec:casestud:defs}). %
-    All edges are secure compilers, but dotted edges use the presented framework (\Cref{sec:sequential}) and solid edges classic proof techniques. %
+    Full edges are secure compilers passes.
+    Dotted edges are composition of passes and use the presented framework (\Cref{sec:sequential}) to indicate the property they preserve. %
     The dashed lines partition the graph into the sections where the respective theorems are presented.
@@
 
-This subsection defines a secure compiler from $\Ltms$ to $\Ltrg$.
-To this end, the compiler needs to ensure that when execution switches from context to component, the type signatures are respected.
-It can do so by inserting dynamic typechecks prior to entering the body of a function belonging to the component:
+The secure compiler from $\Ltms$ to $\Ltrg$ needs to ensure that when execution switches from context to component, the type signatures are respected.
+Thus, the compiler inserts the following dynamic typechecks before entering the body of a component-defined function (anything elided is a trivial identity function from source to target):
 
@@
 
-The above sketches interesting parts, anything left out is a trivial ``re-colouring'' from $\src{blue}$ to $\trg{red}$.
-
-Since $\trg{L}$ has no static typing, a bogus context $\trg{\library_{\ctx}}$ can invoke a callable object accepting a $\src{\natt}$ with $\trg{\lpair{17}{29}}$.
-By inserting a check, the compiler ensures that execution does not proceed in such cases.
+\noindent Since $\trg{L}$ has no static typing, an attacker $\trg{\library_{\ctx}}$ can invoke a component function accepting a $\src{\natt}$ with $\trg{\lpair{17}{29}}$.
+With the dynamic check, the compiler ensures that execution aborts in such cases.
 %The compiler does not insert other checks and proceeds as the identity function (which in this paper amounts to a simple re-colouring of $\src{L_{\tmssafe}}$ to $\trg{L}$ expressions).
@@
 Compiling the \texttt{strncpy} function from \Cref{sec:introduction} with $\cca$, the compiler would in this case ensure that the arguments that are evaluated in the compiled \texttt{strncpy} are valid.
-This effectively has no influence on the resulting trace, so a corresponding cross-language trace relation $\sim^{\Ltms}_{\Ltrg}$ is, up to one trace being $\src{blue}$ and the other $\trg{red}$, defined as equality.
+This effectively has no influence on the resulting trace, so a corresponding cross-language trace relation $\sim^{\Ltms}_{\Ltrg}$ is just equality.
+\MPin{what's the point of this relation? where is it used? why define it here?}
 
@@
 
-This subsection defines spatial memory safety preserving compiler from $\Ltrg$ to $\Lms$, $\ccb$.
-The compiler $\cc{\trg{L}}{\Lms}$ only inserts bounds-checks whenever reading from or writing to memory in order to enforce \gls*{sms}.
-These bounds checks need the bounds information, which $\ccb$ keeps around by introducing a fresh identifier $\irl{x_{SIZE}}$ for each allocation that binds $\irl{x}$.
+The spatial memory safety preserving compiler from $\Ltrg$ to $\Lms$ only inserts bounds-checks whenever reading from or writing to memory in order to enforce \gls*{sms}.
+These bounds checks need the bounds information, which the compiler keeps around by introducing a fresh identifier $\irl{x_{SIZE}}$ for each allocation that binds $\irl{x}$.
 Then, it is simply a matter of referring to that variable and ensuring that memory accesses are in the interval $[\irl{0},\irl{x_{SIZE}})$.
-For the case where it does not, the compiler inserts a panic.
+When the check fails, the code aborts.
 
@@
     \end{lstlisting}
-    Suppose context \texttt{strncpy(2, x, y)}, where \texttt{x} and \texttt{y} are pointers to valid regions of memory with allocated space for exactly two cells and do not contain the null-terminating character \texttt{'\textbackslash 0'}.
+Consider context \texttt{strncpy(2, x, y)}, where \texttt{x} and \texttt{y} are pointers to valid regions of memory with allocated space for exactly two cells and do not contain the null-terminating character \texttt{'\textbackslash 0'}.
+% 
 When calling \texttt{strncpy} like this, the event $\ev{Use}\ \loc_{x}\ 2;\comp;\unlock$ would not be emitted during execution, since the bounds check prevents the condition \texttt{src[i] != '\textbackslash 0'} from executing.
+\MP{what?}
 \end{exampleenv}
 
-Contrary to $\cca$, $\ccb$ may in fact change the trace of the original $\Ltrg$ program: if there is a memory access, it needs to be protected with a bounds check.
-The corresponding relation $\sim^{\Ltrg}_{\Lms}$ that describes this semantic effect of the compiler may be sketched as follows:
+Contrary to the previous compiler, $\ccb$ may change the trace of the original $\Ltrg$ program: \MP{make a uniformity check: always uppercase or always lowercase after :}if there is a memory access, it needs to be protected with a bounds check.
+The corresponding relation $\sim^{\Ltrg}_{\Lms}$ that describes this semantic effect of the compiler may be sketched as follows.
+\MPin{no sketching. rephrase / say that omitted things are omitted for a reason}
 
@@
 
-And so on for the other, $\trg{Set}$ is analogous and any other event is just a trivial re-colouring.
-Note that neither $\Ltrg$ nor $\Lms$ expose branches on the trace.
-The bounds check is left abstract above, since the precise definition needs some technicalities that only inhibit readability without any technical insights.
-
-It is possible to prove that $\ccb$ is robustly preserving \gls*{sms}.
+Action $\trg{Set}$ is treated analogously; any other event is just a trivial re-colouring.\MP{no explicit mentioning of colours. identity functions}
+Note that neither $\Ltrg$ nor $\Lms$ expose branches on the trace.\MP{so what?}
+The bounds check is left abstract above, since the precise definition needs some technicalities that only inhibit readability without any technical insights.\MP{?? unclear}
+
+Compiler $\ccb$ is robustly preserving \gls*{sms}.
 \begin{theorem}[$\ccb$ secure w.r.t. \gls*{sms}]\label{thm:ccb:rtp:sms}
@@
 
-\Cref{corr:ccab:rtp:ms} states that the composition of $\cca$ and $\ccb$ is secure w.r.t. \gls*{ms} and follows from \Cref{thm:cca:rtp:tms,thm:ccb:rtp:sms} using \Cref{thm:rtpsim:sig}.
+\Cref{corr:ccab:rtp:ms} states that the composition of $\cca$ and $\ccb$ is secure w.r.t.
+\gls*{ms} and \MP{it.
+many missing such subjects}follows from \Cref{thm:cca:rtp:tms,thm:ccb:rtp:sms} using \Cref{thm:rtpsim:sig}.
 
@@
 The context can overwrite the flag and exit the mode, but upon invoking a function that is part of the component, the flag would be set again.
-Because of this, the semantic effect of the compiler, i.e., the corresponding cross-language trace relation $\sim^{\Lms}_{\Lscct}$, only relates events without secrets\footnote{For context actions, it can be whatever.}:
+Because of this, the corresponding cross-language trace relation $\sim^{\Lms}_{\Lscct}$, only relates events without secrets:%
+\footnote{For context actions, the relation is arbitrary.
+\MPin{is it really?}
+}
 
@@
 
-In the above, because of space constraints, pretend that each base-event has an additional $\obj{\comp;\securitytag}$, $\ird{\comp;\securitytag}$ annotation, so that full events are related.
+The base-event relation above scales to full events by ensuring the missing annotations ($\obj{\comp;\securitytag}$ and $\ird{\comp;\securitytag}$) are the same.
 
@@
 
-\subsection{Robust Preservation of Intersection of Memory Safety, Strict Cryptographic Constant Time, and Speculative Safety}
+\subsection{Robust Preservation of Memory Safety, Strict Cryptographic Constant Time, and Speculative Safety}
 

### 2024-05-27 csf2/main.tex
nit

@@
 {
-\[
-  \begin{array}{rcl}
+\begin{gather}
+  \begin{align*}%{rcl}
     %\cca(\src{x}) = &\ \trg{x} 
@@
     %&
-    \cca(\src{\lget{x}{\expr}}) &= &\ \lget[\trg]{\trg{x}}{\left[\cca(\src{\expr})\right]} \\
-    \cca(\src{\ldelete{x}}) &= &\ \ldelete[\trg]{\left[\cca(\src{x})\right]} \\
+    \cca(\src{\lget{x}{\expr}}) = &\ \lget[\trg]{\trg{x}}{\left[\cca(\src{\expr})\right]} \\
+    \cca(\src{\ldelete{x}}) = &\ \ldelete[\trg]{\left[\cca(\src{x})\right]} 
+    \\
+    \cca(\src{\lfunction{g}{x:\natt\to\type[_{e}]}{\expr}})  =&
+  \end{align*}
+  \\
+  \begin{align*}
 %    \cca(\src{\lset{x}{\expr[_{1}]}{\expr[_{2}]}}) & = \lset[\trg]{\left[\cca(\src{x})\right]}{\left[\cca(\src{\expr[_{1}]})\right]}{\left[\cca(\src{\expr_{2}})\right]} \\
-\cca(\src{\lfunction{g}{x:\natt\to\type[_{e}]}{\expr}})  &= &\ \lfunction[\trg]{\trg{g}}{\trg{x}}{\lifz[\trg]{\trg{\lhast{x}{\natt}}}{
-\\%
-                                                         &  &
+% \\ 
+\lfunction[\trg]{\trg{g}}{\trg{x}}{\lifz[\trg]{\trg{\lhast{x}{\natt}}}{
                                                             \left[\cca(\src{\expr})\right] %
                                                                                                  }{\labort[\trg]}}
-  \end{array}
-\]
+  \end{align*}
+\end{gather}
 }
@@
 
-Action $\trg{Set}$ is treated analogously; any other event is just a trivial re-colouring.\MP{no explicit mentioning of colours. identity functions}
-Note that neither $\Ltrg$ nor $\Lms$ expose branches on the trace.\MP{so what?}
-The bounds check is left abstract above, since the precise definition needs some technicalities that only inhibit readability without any technical insights.\MP{?? unclear}
+Action $\trg{Set}$ is treated analogously; any other event is just a trivial re-colouring.\MPin{no explicit mentioning of colours. identity functions}
+% Note that neither $\Ltrg$ nor $\Lms$ expose branches on the trace.\MP{so what?}
+For simplicity, we elide the environment that this relation carries around in order to bind each location to its metadata (such as its size), and resolve the ``$\trg{n}$ in/out of bounds'' premise.
+% The bounds check mentioned in the premises of the rules above is left abstract for the sake of simplicity, the interested reader can find
+% since the precise definition needs some technicalities that only inhibit readability without any technical insights.
 
@@
 
+\MPin{
+	can this be made formal?
+	what's the point?
+}
+
 \subsection{Compatibility of Secure Compiler Passes}\label{subsec:compatsecpasses}

### 2024-05-27 csf2/main.tex
rw

@@
 \begin{figure}[!htb]
+	\vspace{-1em}
   $$
@@
 
-This section discusses work on robust compilation (\Cref{subsec:relw:seccomprtp}) and on other secure compilation criteria (\Cref{subsec:relw:seccompcrit}).
-Since the case study of \Cref{sec:casestud:defs,sec:casestud:rtp} implements measures for preserving \gls*{ms}, \gls*{cct}, and \gls*{ss}, this section then presents relevant related work as well (\Cref{subsec:relw:msmechs,subsec:relw:cctmechs,subsec:relw:ssmechs}).
-
-\subsection{Secure Compilation as Robust Preservation}\label{subsec:relw:seccomprtp}
+This section discusses robust compilation, other secure compilation criteria and work related to the properties preserved in the case study.
+% (\Cref{subsec:relw:seccomprtp}) 
+% (\Cref{subsec:relw:seccompcrit}).
+% Since the case study of \Cref{sec:casestud:defs,sec:casestud:rtp} implements measures for preserving \gls*{ms}, \gls*{cct}, and \gls*{ss}, 
+% Then, this section discusses work related 
+% relevant related work as well 
+% (\Cref{subsec:relw:msmechs,subsec:relw:cctmechs,subsec:relw:ssmechs}).
+
+\paragraph*{Secure Compilation as Robust Preservation}\label{subsec:relw:seccomprtp}
 
@@
 The authors demonstrate a similar compositionality theorem to what is presented here (\Cref{sec:sequential}) but use it in the context of protocols.
-The work is missing the upper and lower compositions as well as the generality to support different trace models.
+The work does not consider the generality to support different trace models.
 % The authors demonstrate a similar compositionality theorem to what is presented here (\Cref{sec:sequential}) as well as in an earlier version of this work~\cite{kruse2022csc}.
@@
 
-\subsection{Other Secure Compilation Criteria}\label{subsec:relw:seccompcrit}
+\paragraph*{Other Secure Compilation Criteria}\label{subsec:relw:seccompcrit}
 
@@
 Fully abstract compilation~\cite{abadi1999fullabstraction} states that a compiler should preserve and reflect observational equivalence between source and target programs.
-It was shown~\cite{abate2021faandrc} that fully abstract compilers robustly preserve program properties that are either trivial or meaningless.
+Abate \emph{et al.}~\cite{abate2021faandrc} showed that fully abstract compilers robustly preserve program properties that are either trivial or meaningless.
 As a mitigation for this, the authors presented a categorical approach based on maps of distributive laws~\cite{watanabe2002modl}, which they call many maps of distributive laws.
@@
 
-\subsection{Memory Safety Mechanisms}\label{subsec:relw:msmechs}
+\paragraph*{Memory Safety Mechanisms}\label{subsec:relw:msmechs}
 
@@
 
-\subsection{Cryptographic Constant Time Mechanisms}\label{subsec:relw:cctmechs}
+\paragraph*{Cryptographic Constant Time Mechanisms}\label{subsec:relw:cctmechs}
 
@@
 
-\subsection{Speculation Safety Mechanisms}\label{subsec:relw:ssmechs}
-
-With the advent of speculative execution in processors, it is possible to exploit this unusual paradigm identify secrets.
-SPECTECTOR~\cite{guarnieri2018spectector} models speculative execution formally and the present work has taken inspiration in the sematics of $\Lspec$ and in \Cref{def:trace:ss} from SPECTECTOR.
-A subsequent line of work~\cite{fabian2022automatic} shows how to modularly define and compose formal semantics that models different speculative execution variants. 
-Similar in spirit, recent work~\cite{fabian2024lift} demonstrates how to lift existing secure compilation proofs for security properties related to speculative execution to more powerful properties, i.e., properties defined for languages with a trace model that enables more vulnerabilities, such as combinations of different speculative execution exploits.
-To this end, they define a lifting framework and demonstrate how a secure compilation proof for a weaker property can be lifted to a stronger one. 
-Contrary to the compositionality framework presented in this paper, these works are only concerned with composing semantics of the respective programming languages, while in this paper, the secure compilers themselves are composed.
+\paragraph*{Speculation Safety Mechanisms}\label{subsec:relw:ssmechs}
+
+With the advent of speculative execution in processors, it is possible to exploit this unusual paradigm leak secrets.
+This work has taken inspiration in the sematics of $\Lspec$ and in \Cref{def:trace:ss} from recent formalisations~\cite{guarnieri2018spectector,fabian2022automatic}.
+\MPin{
+  we take SS from exorcising.
+  we do not take SNI from spectector because it is a hypersafety.
+  other work defined similar props.
+}
+% A subsequent line of work~\cite{fabian2022automatic} shows how to modularly define and compose formal semantics that models different speculative execution variants. 
+% Recent work~\cite{fabian2024lift} demonstrates how to lift existing secure compilation proofs for security properties related to speculative execution to more powerful properties, i.e., properties defined for languages with a trace model that enables more vulnerabilities, such as combinations of different speculative execution exploits.
+% To this end, they define a lifting framework and demonstrate how a secure compilation proof for a weaker property can be lifted to a stronger one. 
+% Contrary to the compositionality framework presented in this paper, these works are only concerned with composing semantics of the respective programming languages, while in this paper, the secure compilers themselves are composed.
 

### 2024-05-27 csf2/main.tex
7a

@@
 
-So far, the considered trace models do not let us express speculative execution attacks such as Spectre~\cite{}\MP{cits}. 
+So far, the considered trace models do not let us express speculative execution attacks such as Spectre~\cite{kocher2019spectre}. 
 For this, we extend the earlier trace model (see \Cref{subsec:scct:tracemodel}) so that the security tags ($\securitytag$) carry additional information about the kind of private data leakage, i.e., the type of speculative leak.
@@
   $
-  \specmssafe=\sigma_{\sim_{\mathghost}}\left(\msscctsafe\right)\cap\sssafe
+  \specmssafe := \sigma_{\sim_{\mathghost}}\left(\msscctsafe\right)\cap\sssafe
   $
@@
 
-When modeling some thing using formal methods, the intent is to do so faithfully. 
-That is why it is important to ensure that the properties considered in \Cref{sec:casestud:rtp} follow this ideal.
-It is relatively straightforward to see that, e.g., $\mssafe$ models a basic memory safety property. 
-But, what about $\sigma_{\sim^{\Ltms}_{\Lms}}\left(\mssafe\right)$?
-It is not immediately obvious that the resulting property after projecting it from $\Lms$-level to $\Ltms$-level is similar in spirit to the original $\mssafe$ property (\Cref{def:trace:msdef}) and this is what needs to be discussed when using the presented framework.
-Let us investigate the involved projections from the case study.
+Compiler designers need to be careful that their implementation preserves the original source language semantics. 
+Similar to this, it is important to consider the cross-language relation, which gives a semantic description of the compiler in terms of traces, and check that the original target (resp.
+source) property is properly translated using the universal (resp.
+existential) image.
+The following discusses this for relations presented in \Cref{sec:casestud:rtp}.
 

### 2024-05-27 csf2/main.tex
oob

@@
 {\em All} traces $\trg{\trace}$ with $\trg{\trace}\sim^{\Ltrg}_{\Lms}\irl{\trace}$ are almost identical to $\irl{\trace}$. 
-They differ in memory accesses and crashes, where $\irl{\lightning}$ can be related to a $\trg{\lightning}$ or a memory event, e.g., $\trg{Get\ \loc\ 2}$.
+They differ in memory accesses and crashes, where $\irl{\lightning}$ can be related to a $\trg{\lightning}$ or a memory event, e.g., $\trg{Get\ \loc\ 2}$, for an out-of-bounds access.
 In both cases, the memory access cannot be out of bounds.

### 2024-05-27 csf2/main.tex
not oob
allow nesting of speculation

@@
         To this end, it presents a compilation chain consisting of six passes that ultimately preserve \gls*{mssscct} by means of composing four secure passes that \emph{individually} preserve \gls*{tms}, \gls*{sms}, \gls*{scct}, and \gls*{ss}, and two passes which are well-known optimisations \gls*{dce} and \gls*{cf},.
-        %Furthermore, the chain includes two optimisation passes: One performs \gls*{dce} and the other \gls*{cf}.
-        The formalisation of this case study showcases the power of the presented framework: The divide-and-conquer approach to software engineering is a viable strategy even for the development of secure compilers.
+        %Furthermore, the chain includes two optimisation passes: one performs \gls*{dce} and the other \gls*{cf}.
+        The formalisation of this case study showcases the power of the presented framework: the divide-and-conquer approach to software engineering is a viable strategy even for the development of secure compilers.
 %        This demonstrates that the usage of the presented framework is both simpler and more expressive compared to one monolithic security proof of the whole compilation chain.
@@
 However, \gls*{ms} uses a different trace model than \gls*{scct}, so intersecting them would trivially yield the empty set. 
-To remedy this issue, let $\sim_{\ctsafe}$ be a cross-language trace relation whose definition can be sketched as follows:\MP{why just sketched?}
+To remedy this issue, let $\sim_{\ctsafe}$ be a cross-language trace relation whose definition is partially follows:
 \[
@@
 Similar to \gls*{msscct}, there is a need to relate the different trace models with each other, so that the memory safety property without speculation can be lifted to speculation. 
-To this end, let $\sim_{\mathghost}$ be a cross-language trace relation whose definition can be sketched as follows:\MP{sketch?}
+To this end, let $\sim_{\mathghost}$ be a cross-language trace relation whose definition is partially as follows:
 
@@
 $\Lms$ is exactly the same language as $\Ltrg$, but this paper still distinguishes the two for sake of readability (\Cref{subsec:lms}).
-All three languages --- so $\Ltms$, $\Ltrg$, and $\Lms$ --- assume \gls*{cct} to hold, since this is -- in an ideal world -- what the programmer would expect, too: It is the job of the compiler to preserve and (potentially) enforce security~\cite{cauligi2019fact,nagarakatte2010cets,nagarakatte2009soft,akritidis2009baggy}.
+All three languages --- so $\Ltms$, $\Ltrg$, and $\Lms$ --- assume \gls*{cct} to hold, since this is -- in an ideal world -- what the programmer would expect, too: it is the job of the compiler to preserve and (potentially) enforce security~\cite{cauligi2019fact,nagarakatte2010cets,nagarakatte2009soft,akritidis2009baggy}.
 % This paper considers \gls*{cct} as an architecture-dependent property\MPin{can we justify this}, so
@@
     \text{(Events)} & \src{\event} &:=& \src{\preevent;\sandboxtag} \mid \src{\emptyevent} \mid \src{\lightning} \\ 
+    \text{(Values)} & \src{v} &:=& \src{n} \\
     \text{(Expressions)} & \src{\expr} &:=& \src{x} \mid \src{n} \mid \src{\lbinop{\expr[_1]}{\expr[_2]}} \mid \src{\lifz{\expr[_1]}{\expr[_2]}{\expr[_3]}} \\ 
@@
 Because pointers are kept ``alive'' in pointer maps ($\src{\Delta}$), $\Ltms$ dynamic semantics does not get stuck if it encounters a memory-safety violation such as a double--free.
-The memory state carries two heaps to model sandboxing between a context and a component.
-\MP{notation}
+The memory state carries two heaps to model sandboxing between a context ($\src{H^\ctx}$) and a component ($\src{H^\comp}$).
 
@@
 Furthermore, the syntax of $\Ltrg$ is extended with a way to inspect whether a pointer is freed ($\trg{\lispoisoned{x}}$), evaluating to $\trg{0}$ if it is freed and to $\trg{1}$ otherwise.
-Functions may receive arguments that are not $\trg{\natt}$ and values are extended with tuples\MP{add this?}\MP{no values in previous grammar. add?}.
+Functions may receive arguments that are not $\trg{\natt}$ and values are extended with tuples.
 
@@
   \begin{array}{rrcl}
+    \text{(Values)} & \trg{v} &:=& \dots \mid \trg{\lpair{v_1}{v_2}} \\
     \text{(Expressions)} & \trg{\expr} &:=& \dots \mid \trg{\lhast{\expr}{\tau}} \mid \trg{\lispoisoned{x}}
@@
   Consider again \Cref{ex:strncpy}, with a context copying the string \texttt{Hello World}, where everything is marked with a high security tag: $\lock$.
-  The top half of \Cref{fig:ex-cct} (titled $\Lscct$), describes the execution trace of the program, while the bottom side of the table (titled $Spec$), describes the related specification trace (\Cref{subsec:msctss:tracemodel}).
+  The top half of \Cref{fig:ex-cct} (titled $\Lscct$), describes the execution trace of the program, while the bottom side of the table (titled ,,Specification''), describes the related specification trace (\Cref{subsec:msctss:tracemodel}).
   Read in parallel from top to bottom, the figure shows parts of the execution trace. 
@@
     & \Lscct &        \\
-    Active & \mid & Inactive \\\hline
+    \text{Active} & \mid & \text{Inactive} \\\hline
     \obj{\ev{Alloc}\ \loc_{x}\ 12;\comp;\unlock} & \mid & \obj{\ev{Alloc}\ \loc_{x}\ 12;\comp;\unlock} \\
@@
   \begin{array}{ccc}
-          & Spec & \\
-     Active & \mid & Inactive \\\hline
+          & \text{Specification} & \\
+  \text{Active} & \mid & \text{Inactive} \\\hline
      \ev{Alloc}\ \loc_{x}\ 12;\unlock & \mid & \ev{Alloc}\ \loc_{x}\ 12;\unlock\\
@@
 
-$\Lspec$ extends $\Lscct$ (\Cref{subsec:lscct}) with speculative dynamic semantics that is inspired by existing work~\cite{guarnieri2018spectector}\MP{cit xaver's}.
-After branching or accessing memory\MP{?}, speculation starts by pushing the current configuration on the speculation register ($\ird{S}$), and run subsequent code in a predetermined window until a rollback of operational state is performed.
-The window is set inside the control-flow state ($\ird{\cfstate}$), whose details are in the technical report.
-Within such windows, data is leaked, marked as high $\ird{\lock}$ and with an annotation that indicates the respective speculative execution variant, in this case just SPECTRE-PHT~\cite{kocher2019spectre}, whose semantics is sketched with \Cref{tr:e-ifz-true-spec}, where $\omega$ is the size of the speculation window\MP{role of omega?}.
+$\Lspec$ extends $\Lscct$ (\Cref{subsec:lscct}) with speculative dynamic semantics that is inspired by existing work~\cite{guarnieri2018spectector,fabian2022automatic}.
+After branching, speculation starts by pushing the current configuration into the speculation state ($\ird{S}$), and run subsequent code in a predetermined window (whose size is $\omega$) until a rollback of operational state is performed.
+The window is set inside the speculation state.
+Within such windows, data may leak, marked as high $\ird{\lock}$ and with an annotation that indicates the respective speculative execution variant, in this case just SPECTRE-PHT~\cite{kocher2019spectre}, whose starting-point is \Cref{tr:e-ifz-true-spec}.
+% where $\omega$ is the size of the speculation window, i.e., a number of reduction steps to be taken under speculative semantics.
 Additionally, $\ird{\lock}$ may also carry a $\ird{NONE}$ annotation, to signal that this is a leak irrespective of speculation, i.e., as defined earlier (\Cref{subsec:lscct}).
@@
 In terms of language features, $\Lspec$ supports a barrier operation $\ird{\lbarrier}$ that blocks speculative execution. 
-To facilitate speculative execution in a non-assembly-like language, a stack of operational state is used and speculation is active if that stack has a size greater than one, indicated by writing the operational states vertically.
+To facilitate speculative execution in a non-assembly-like language, a stack of operational state is used and speculation is active if that stack has a size greater than one.
 This is exploited in \Cref{tr:e-spec-eat} to consume any leftover speculation.
@@
     \text{(Expressions)} & \ird{\expr} &:=& \dots \mid \ird{\lbarrier} \\
-    \text{(Speculation Register)} & \ird{S} &:=& \ird{\hole{\cdot}} \mid \left(\ird{\Omega};\ird{\expr}\right)
+    \text{(Speculation State)} & \ird{S} &:=& \ird{\Omega\triangleright\expr} \mid \ird{S},\left(\ird{\Omega};\ird{n};\ird{\expr}\right)
   \end{array}
@@
   \typerule{$e-\ird{ifz}-\ird{true}-\text{spec}$}{
-    \ird{\cfstate[']}.\ird{\operatorname{win}} = \omega &
-    \ird{\trace} = \ird{\ev{Branch}\ 0;\sandboxtag;\securitytag}\cdot\ird{\ev{Spec};\sandboxtag;\securitytag} \\
-    \ird{S} = ({\ird{\cfstate;\sandboxtag;D;\memstate}};{\ird{\expr[_2]}}),\ird{\hole{\cdot}}
+    \ird{S} = \rtt{(\ird{\cfstate['];\sandboxtag;D;\memstate});\ird{S}}{\ird{\expr[_1]}},({\ird{\cfstate;\sandboxtag;D;\memstate}};\omega;{\ird{\expr[_2]}})
   }{
-    \pstepto{\rtt{(\ird{\cfstate;\sandboxtag;D;\memstate});\ird{\hole{\cdot}}}{\ird{\lifz{0^{\securitytag}}{\expr[_1]}{\expr[_2]}}}}{\rtt{(\ird{\cfstate['];\sandboxtag;D;\memstate});\ird{S}}{\ird{\expr[_1]}}}{\ird{\trace}}
+    \pstepto{\rtt{\ird{\cfstate;\sandboxtag;D;\memstate}}{\ird{\lifz{0^{\securitytag}}{\expr[_1]}{\expr[_2]}}}}{\ird{S}}{\ird{\ev{Branch}\ 0;\sandboxtag;\securitytag}\cdot\ird{\ev{Spec};\sandboxtag;\securitytag}}
   }{e-ifz-true-spec}
@@
   \typerule{$e-\ird{spec}-\text{eat}$}{
-    \ird{\cfstate}.\ird{\operatorname{win}} > 0 &
-    \pstepto{\rtt{(\ird{\cfstate;\sandboxtag;D;\memstate};\ird{\hole{\cdot}})}{\ird{\expr}}}{\rtt{(\ird{\cfstate['];\sandboxtag;D;\memstate};\ird{\hole{\cdot}})}{\ird{\expr[']}}}{\ird{\event}}
-    \\
-    \ird{\cfstate['']}.\ird{\operatorname{win}} = \ird{\cfstate[']}.\ird{\operatorname{win}}-1 &
-    \ird{S} \not=\ird{\hole{\cdot}}
+    \ird{n} > 0 &
+    \pstepto{\rtt{\ird{\Omega}}{\ird{\expr}}}{\rtt{\ird{\Omega'}}{\ird{\expr[']}}}{\ird{\event}}
   }{
-    \ghoststepto{\rtt{(\ird{\cfstate;\sandboxtag;D;\memstate};\ird{S})}{\ird{\expr}}}{\rtt{(\ird{\cfstate[''];\sandboxtag;D;\memstate};\ird{S})}{\ird{\expr[']}}}{\ird{\event}}
+    \ghoststepto{\ird{S},(\ird{\Omega};\ird{n};\ird{\expr})}{\ird{S},(\ird{\Omega'};\ird{n}-1;\ird{\expr'})}{\ird{\event}}
   }{e-spec-eat}
@@
   \typerule{$e-\ird{spec}-\text{eaten}$}{
-    \ird{\cfstate}.\ird{\operatorname{win}} = 0 &
-    \ird{S} = \left(\ird{\Omega};\ird{\expr[']}\right)
   }{
-    \ghoststepto{\rtt{(\ird{\cfstate;\sandboxtag;D;\memstate};\ird{S})}{\ird{\expr}}}{\rtt{(\ird{\Omega};\ird{\hole{\cdot}})}{\ird{\expr[']}}}{\ird{\event}}
+    \ghoststepto{\ird{S},(\ird{\Omega};\ird{0};\ird{\expr})}{\ird{S}}{\ird{Rlb;\Omega.\sandboxtag;\unlock}}
   }{e-spec-eaten}
@@
   \typerule{$e-\ird{spec}-\text{barrier}$}{
-    \ird{\cfstate'} = \ird{\cfstate}[.\ird{\operatorname{win}} = 0] &
-    \ird{S} \not=\ird{\hole{\cdot}}
   }{
-    \ghoststepto{\rtt{(\ird{\cfstate;\sandboxtag;D;\memstate};\ird{S})}{\ird{\lbarrier}}}{\rtt{(\ird{\cfstate';\sandboxtag;D;\memstate};\ird{\hole{\cdot}})}{\ird{\lbarrier}}}{\ird{\event}}
+    \ghoststepto{\ird{S},(\ird{\Omega};\ird{n};\ird{\lbarrier})}{\ird{S},(\ird{\Omega};\ird{0};\ird{0})}{\ird{Barrier;\Omega.\sandboxtag;\unlock}}
   }{e-spec-barrier}
@@
 \footnote{In the rule, the notation $\cfstate' = \cfstate[.\text{win} = 0]$ means that $\cfstate'$ is a copy of $\cfstate$ up to field $\text{win}$, which should be set to $0$.}
+Lastly, $\ird{Rlb}$ is emitted when the speculation window is zero and the operationl state is rolled back (see \Cref{tr:e-spec-eaten}).
 
@@
 Compiling the \texttt{strncpy} function from \Cref{sec:introduction} with $\cca$, the compiler would in this case ensure that the arguments that are evaluated in the compiled \texttt{strncpy} are valid.
-This effectively has no influence on the resulting trace, so a corresponding cross-language trace relation $\sim^{\Ltms}_{\Ltrg}$ is just equality.
-\MPin{what's the point of this relation? where is it used? why define it here?}
+%This effectively has no influence on the resulting trace, so a corresponding cross-language trace relation $\sim^{\Ltms}_{\Ltrg}$ is just equality.
 
@@
 % 
-When calling \texttt{strncpy} like this, the event $\ev{Use}\ \loc_{x}\ 2;\comp;\unlock$ would not be emitted during execution, since the bounds check prevents the condition \texttt{src[i] != '\textbackslash 0'} from executing.
-\MP{what?}
+Without \gls*{sms} instrumentation, the event $\ev{Use}\ \loc_{x}\ 2;\comp;\unlock$ would appear on the trace, which is an out-of-bounds access. 
+But, with \gls*{sms} mitigation in place, the use event would not appear during execution, since the bounds check prevents the condition \texttt{src[i] != '\textbackslash 0'} from executing.
 \end{exampleenv}
 
-Contrary to the previous compiler, $\ccb$ may change the trace of the original $\Ltrg$ program: \MP{make a uniformity check: always uppercase or always lowercase after :}if there is a memory access, it needs to be protected with a bounds check.
-The corresponding relation $\sim^{\Ltrg}_{\Lms}$ that describes this semantic effect of the compiler may be sketched as follows.
-\MPin{no sketching. rephrase / say that omitted things are omitted for a reason}
+Contrary to the previous compiler, $\ccb$ may change the trace of the original $\Ltrg$ program: if there is a memory access, it needs to be protected with a bounds check.
+The corresponding relation $\sim^{\Ltrg}_{\Lms}$ that describes this semantic effect of the compiler is partially as follows\footnote{In the interest of space, the full details are in the technical report.}:
 
@@
 
-Action $\trg{Set}$ is treated analogously; any other event is just a trivial re-colouring.\MPin{no explicit mentioning of colours. identity functions}
+Action $\trg{Set}$ is treated analogously; any other event is ,,the same''\footnote{Up to colours.}.
 % Note that neither $\Ltrg$ nor $\Lms$ expose branches on the trace.\MP{so what?}
@@
 
-\Cref{corr:ccab:rtp:ms} states that the composition of $\cca$ and $\ccb$ is secure w.r.t.
-\gls*{ms} and \MP{it.
-many missing such subjects}follows from \Cref{thm:cca:rtp:tms,thm:ccb:rtp:sms} using \Cref{thm:rtpsim:sig}.
+\Cref{corr:ccab:rtp:ms} states that the composition of $\cca$ and $\ccb$ is secure w.r.t. \gls*{ms} and it follows from \Cref{thm:cca:rtp:tms,thm:ccb:rtp:sms} using \Cref{thm:rtpsim:sig}.
 
@@
 Because of this, the corresponding cross-language trace relation $\sim^{\Lms}_{\Lscct}$, only relates events without secrets:%
-\footnote{For context actions, the relation is arbitrary.
-\MPin{is it really?}
-}
 
@@
 
-Clearly, the corresponding cross-language trace relation $\sim^{\Lscct}_{\Lspec}$ has only one non-trivial case: For branches, only relate them where speculation is blocked by a barrier:
+Clearly, the corresponding cross-language trace relation $\sim^{\Lscct}_{\Lspec}$ has only one non-trivial case: for branches, only relate them where speculation is blocked by a barrier:
 
@@
 Next in the chain is $\sim^{\Lms}_{\Lscct}$, which has to be well-formed w.r.t. $\scctsafe$.
-This lemma holds, since a trace that was $\scctsafe$ is $\scctsafe$ even after applying $\sim^{\Lms}_{\Lscct}$: The relation enforces that $\Lscct$ traces related to $\Lms$ traces have no leaks of secrets whatsoever.
+This lemma holds, since a trace that was $\scctsafe$ is $\scctsafe$ even after applying $\sim^{\Lms}_{\Lscct}$: the relation enforces that $\Lscct$ traces related to $\Lms$ traces have no leaks of secrets whatsoever.
 
@@
 They differ in memory accesses and crashes, where $\irl{\lightning}$ can be related to a $\trg{\lightning}$ or a memory event, e.g., $\trg{Get\ \loc\ 2}$, for an out-of-bounds access.
-In both cases, the memory access cannot be out of bounds.
+In both cases, the translated event does not represent an out-of-bounds memory access.
 
@@
 
-Previous sections disregarded a glaring but minor technical issue: Take $\tmssafe$ (\Cref{def:trace:tmsdef}) as an example.
+Previous sections disregarded a glaring but minor technical issue: take $\tmssafe$ (\Cref{def:trace:tmsdef}) as an example.
 It contains traces as specified in \Cref{subsec:basic:memsafety:tracemodel}.

### 2024-05-27 csf2/main.tex
some reformulations

@@
 Because of this, this paper may write $\rtptau{\cc{\src{L}}{\trg{L}}}{\trg{\pi}}{\sim}$, even though $\trg{\pi}$ is a property and not a class.
-A similar construction can be applied to the projection functions.\MP{wut?}
+A similar construction can be used to the projection functions (see \Cref{def:universal:img}) by applying them to the lifted version of $\trg{\pi}$ instead of just $\trg{\pi}$.
 
@@
 Security tags $\obj{\securitytag}$ are arranged in the usual secrecy lattice\MP{NI cits}, where $\obj{\lock}\le\obj{\securitytag}$ and $\obj{\securitytag}\le\obj{\unlock}$.
-To prevent secrets from leaking but still enable reasoning about memory safety, $\Lscct$ extends base-events with $\obj{\ev{iGet}\ \loc\ \valueexpr}$ and $\obj{\ev{iSet}\ \loc\ \valueexpr}$.
-\MPin{what's the i for? explain. intuition missing}
-This way, memory accesses to secret data are still exposed as memory accesses, despite being protected by data-independent timing mode, e.g, \Cref{tr:to-e-get-in-noleak}.
+
+Memory accesses to secret data need to be present to reason about memory safety, even when execution is in data-independent timing mode, e.g, \Cref{tr:to-e-get-in-noleak}.
+$\Lscct$ extends base-events with $\obj{\ev{iGet}\ \loc\ \valueexpr}$ and $\obj{\ev{iSet}\ \loc\ \valueexpr}$ to prevent secrets from leaking but still enable reasoning about memory safety. 
 Due to the technical setup, the rule needs to check if the access is in bounds ($\obj{\loc}+\obj{n}\in\obj{H^t}$) and update the secrecy tag ($\obj{\securitytag''}=\obj{\securitytag\sqcap\securitytag'}$) according to the lattice described earlier.

### 2024-05-28 csf2/main.tex
address some todos

@@
 
-Future work intends to use upper and lower composition as means to build a secure compiler incrementally.
-\MPin{
-	we should add briefly which real-world compilers (taken from the elided examples) are upper and lower compos
-}
-\MP{no coq here ?}
+Future work intends to use upper and lower composition as means to build a secure compiler incrementally
+This follows current non-formal-methods based compiler engineering, such as LLVM~\cite{lattner2004llvm}, where languages such as \texttt{C++} and \texttt{Fortran} translate into LLVM IR and languages such as \texttt{x86} or \texttt{ARMv8} assembly variants translate out of LLVM IR.
+
+% no coq, because above definitions are non-formal. 
+%   How would you precisely define the lower compo compiler, given two compilers?
 
@@
 
-\gls*{ss}~\cite{patrignani2021exorcising}, similar to \gls*{scct}, is an overapproximation of a variant of noninterference.
-The definition of the property itself turns out to be syntactically equal to \gls*{scct}, modulo the underlying trace model.
-\MPin{recap its meaning for security}
+\gls*{ss}~\cite{patrignani2021exorcising}, similar to \gls*{scct}, is a sound overapproximation of a variant of noninterference.
 
@@
   $
-    \sssafe:=\left\{\trace[_{\mathghost}] \left|\begin{array}{rcl}
-        \trace[_{\mathghost}]=\hole{\cdot} & \text{ or } & \trace[_{\mathghost}]=\preevent[_{\mathghost}];\unlock\cdot\trace[_{\mathghost}'] \wedge \trace[_{\mathghost}']\in\sssafe
+    \sssafe:=\left\{\trace[_{\mathghost}] \left|\begin{array}{l}
+      \trace[_{\mathghost}]=\hole{\cdot} \text{ or } \exists\trace[_{\mathghost}'].\\
+      \left(\trace[_{\mathghost}]=\preevent[_{\mathghost}];\unlock\cdot\trace[_{\mathghost}'] \text{or }\trace[_{\mathghost}]=\preevent[_{\mathghost}];\lock_{\text{NONE}}\cdot\trace[_{\mathghost}']\right)\\
+      \text{and }\ \trace[_{\mathghost}']\in\sssafe
                                  \end{array}\right.\right\}
@@
 \end{definition}
-\MPin{why isn't it ok to have locks with NONE in the trace?}
+The technical setup so far leads to the above definition, where only locks annotated with $\text{SPECTRE-PHT}$ are disallowed to occur on the trace.
+That way, programs attaining \gls*{ss} do not necessarily attain \gls*{scct}.
 
@@
 All presented programming languages share a common fragment which is partially presented here and in full detail in the technical report.
-The trace models of all languages are similar to those presented earlier (\Cref{sec:compprop}).
-One technical detail is the addition of a sandboxtag, indicating who is to blame for an emitted action: context ($\ctx$) or component ($\comp$).
-This is a standard technique in secure compilation in order to rule out irrelevant context-level events. 
-For example, a context immediately deallocating an allocated memory region twice trivially violates memory safety, but the main interest in secure compilation is at component-level events and if those signal, e.g., a memory safety issue.
-
-\MPin{this section does not define anything, not even the sandbox tag metavariable}
-\MPin{
-	move b/t/a/e/F/CHI/H/Delta/Phi/Psi to here.
-	their explanation too.
-	make black.
-}
-\MPin{
-	what are pointer maps and Control states?
-}
-
-\subsection{$\src{L_{\tmssafe}}$: A Temporal but Not Spatial Memory Safe Language}\label{subsec:ltms}
-
-$\src{L_{\tmssafe}}$ is a relatively naive language with second class functions, the only values being numbers. 
-It has simple binary operations ($\src{\lbinop{\expr[_1]}{\expr[_2]}}$), branches ($\src{\lifz{\expr[_1]}{\expr[_2]}{\expr[_3]}}$), C-like function calls ($\src{\lcall{f}{\expr}}$) and returns ($\src{\lreturn{\expr}}$), as well as C-like memory operations.
-%Pointers are left second-class and can only be referenced by an identifier, so addresses cannot be guessed by contexts.
-Programs have sets of pre-determined functions called libraries and it is assumed that functions are marked as being part of some component ($\src{\library_{\comp}}$) or context ($\src{\library_{\ctx}}$).
-Functions ($(\src{x};\src{\natt\to\natt};\src{\expr})$) are modelled as tuples and can only accept integers, which also means that contexts may not pass pointers to the component.
-%Extending this is orthogonal to this work.
 
@@
 {
+  \renewcommand{\src}[1]{\mathsf{#1}}
 % \small
@@
                          &&&\mid \src{\lcall{f}{\expr}} \mid \src{\lreturn{\expr}} \\
-    \text{(Functions)} & \src{F} &:=& (\src{x};\src{\natt\to\natt};\src{\expr}) \\
+    \text{(Functions)} & \src{F} &:=& (\src{x};\src{\expr}) \\
     \text{(Libraries)} &&& \src{\library_{\ctx}} : \src{Vars} \to \src{F} \hspace{2em}
@@
 }
-
-Crucially, $\src{L_{\tmssafe}}$ is statically typed, where the type system is inspired by $L^{3}$~\cite{morrisett2005L3,scherer2018fabulous}.
-The type system of $\src{L_{\tmssafe}}$ enforces that every well-typed $\src{L_{\tmssafe}}$ program satisfies \gls*{tms} (\Cref{thm:wt:tms}).
+{
+  \renewcommand{\src}[1]{\mathsf{#1}}
+The trace models of all languages are similar to those presented earlier (\Cref{sec:compprop}).
+One technical detail is the addition of a control tag, indicating who is to blame for an emitted action: context ($\ctx$) or component ($\comp$).
+This is a standard technique in secure compilation in order to rule out irrelevant context-level events. 
+For example, a context immediately deallocating an allocated memory region twice trivially violates memory safety, but the main interest in secure compilation is at component-level events and if those signal, e.g., a memory safety issue.
+Even though this tagging could be used for blame preservation~\cite{patrignani2023blame}, this beyond the focus of this paper.
+Another key difference is that memory accesses are now explicitly modeled as reads ($\src{\ev{Read\ \loc\ n}}$) and writes ($\src{\ev{Write\ \loc\ n}}$) instead of just uses.
+
+All languages have at least numbers as values ($\src{v}$) and second class functions ($\src{F}$).
+Functions are modeled as pairs containing the name of one argument and the body of the function.
+Bodies are just ordinary expressions, which can be simple binary operations ($\src{\lbinop{\expr[_1]}{\expr[_2]}}$), conditionals ($\src{\lifz{\expr[_1]}{\expr[_2]}{\expr[_3]}}$), C-like function calls ($\src{\lcall{f}{e}}$) and returns ($\src{\lreturn{\expr}}$), as well as C-like memory operations.
+Programs have sets of pre-determined functions called libraries and it is assumed that functions are marked as being part of some component ($\src{\library_{\comp}}$) or context ($\src{\library_{\ctx}}$).
+Moreover, operational state ($\src{\Omega}$) is a triple consisting of a control-flow state ($\src{\cfstate}$), a control tag ($\src{\sandboxtag}$), and a memory state ($\src{\memstate}$). 
+The latter carrying information about pointers that are kept ,,alive'' in pointer maps ($\src{\Delta}$), so that dynamic semantics do not get stuck when encountering, e.g., a double-free.
+The memory state also carries two heaps to model sandboxing between a context ($\src{H^{\ctx}}$) and a component ($\src{H^{\comp}}$).
+
+}
+
+\subsection{$\src{L_{\tmssafe}}$: A Temporal but Not Spatial Memory Safe Language}\label{subsec:ltms}
+
+$\src{L_{\tmssafe}}$ is the only statically typed language in this case study and restricts functions ($\src{F}$) to the typing signature $\src{\natt\to\natt}$. 
+The type system of $\src{L_{\tmssafe}}$ is inspired by $L^{3}$~\cite{morrisett2005L3,scherer2018fabulous} and enforces that every well-typed $\src{L_{\tmssafe}}$ program satisfies \gls*{tms} (\Cref{thm:wt:tms}).
 
@@
 \end{theorem}
-
-Operational state ($\src{\Omega}$) is a triple consisting of control-flow state ($\src{\cfstate}$), a control tag ($\src{\sandboxtag}$), and a memory state ($\src{\memstate}$), which carries information about the memory.
-Because pointers are kept ``alive'' in pointer maps ($\src{\Delta}$), $\Ltms$ dynamic semantics does not get stuck if it encounters a memory-safety violation such as a double--free.
-The memory state carries two heaps to model sandboxing between a context ($\src{H^\ctx}$) and a component ($\src{H^\comp}$).
-
-The trace model is largely similar to the memory safety trace model (\Cref{subsec:basic:memsafety:tracemodel}), where the only real differences are separate events for reading and writing memory, instead of just using it, and the sandboxtags.
-With these tags, it is possible to ignore effects performed by a spurious context that, e.g., immediately deallocates a memory location twice, thus violating \gls*{tms}.
-Even though this could be used for blame preservation~\cite{patrignani2023blame}, this beyond the focus of this paper.
 
@@
 
-\MPin{
-	can this be made formal?
-	what's the point?
-}
+%\MPin{
+%	can this be made formal?
+%	what's the point?
+%}
 
@@
 This work has taken inspiration in the sematics of $\Lspec$ and in \Cref{def:trace:ss} from recent formalisations~\cite{guarnieri2018spectector,fabian2022automatic}.
-\MPin{
-  we take SS from exorcising.
-  we do not take SNI from spectector because it is a hypersafety.
-  other work defined similar props.
-}
+% \MPin{
+%  we take SS from exorcising.
+%  we do not take SNI from spectector because it is a hypersafety.
+%  other work defined similar props. 
+% }
 % A subsequent line of work~\cite{fabian2022automatic} shows how to modularly define and compose formal semantics that models different speculative execution variants. 

### 2024-05-28 csf2/main.tex
niut

--- 2024-05-27 csf2/main.tex
@@
         To this end, it presents a compilation chain consisting of six passes that ultimately preserve \gls*{mssscct} by means of composing four secure passes that \emph{individually} preserve \gls*{tms}, \gls*{sms}, \gls*{scct}, and \gls*{ss}, and two passes which are well-known optimisations \gls*{dce} and \gls*{cf},.
-        %Furthermore, the chain includes two optimisation passes: one performs \gls*{dce} and the other \gls*{cf}.
-        The formalisation of this case study showcases the power of the presented framework: the divide-and-conquer approach to software engineering is a viable strategy even for the development of secure compilers.
+        %Furthermore, the chain includes two optimisation passes: One performs \gls*{dce} and the other \gls*{cf}.
+        The formalisation of this case study showcases the power of the presented framework: The divide-and-conquer approach to software engineering is a viable strategy even for the development of secure compilers.
 %        This demonstrates that the usage of the presented framework is both simpler and more expressive compared to one monolithic security proof of the whole compilation chain.
@@
 However, \gls*{ms} uses a different trace model than \gls*{scct}, so intersecting them would trivially yield the empty set. 
-To remedy this issue, let $\sim_{\ctsafe}$ be a cross-language trace relation whose definition is partially follows:
+To remedy this issue, let $\sim_{\ctsafe}$ be a cross-language trace relation whose definition can be sketched as follows:\MP{why just sketched?}
 \[
@@
 
-So far, the considered trace models do not let us express speculative execution attacks such as Spectre~\cite{kocher2019spectre}. 
+So far, the considered trace models do not let us express speculative execution attacks such as Spectre~\cite{}\MP{cits}. 
 For this, we extend the earlier trace model (see \Cref{subsec:scct:tracemodel}) so that the security tags ($\securitytag$) carry additional information about the kind of private data leakage, i.e., the type of speculative leak.
@@
 Similar to \gls*{msscct}, there is a need to relate the different trace models with each other, so that the memory safety property without speculation can be lifted to speculation. 
-To this end, let $\sim_{\mathghost}$ be a cross-language trace relation whose definition is partially as follows:
+To this end, let $\sim_{\mathghost}$ be a cross-language trace relation whose definition can be sketched as follows:\MP{sketch?}
 
@@
   $
-  \specmssafe := \sigma_{\sim_{\mathghost}}\left(\msscctsafe\right)\cap\sssafe
+  \specmssafe=\sigma_{\sim_{\mathghost}}\left(\msscctsafe\right)\cap\sssafe
   $
@@
 $\Lms$ is exactly the same language as $\Ltrg$, but this paper still distinguishes the two for sake of readability (\Cref{subsec:lms}).
-All three languages --- so $\Ltms$, $\Ltrg$, and $\Lms$ --- assume \gls*{cct} to hold, since this is -- in an ideal world -- what the programmer would expect, too: it is the job of the compiler to preserve and (potentially) enforce security~\cite{cauligi2019fact,nagarakatte2010cets,nagarakatte2009soft,akritidis2009baggy}.
+All three languages --- so $\Ltms$, $\Ltrg$, and $\Lms$ --- assume \gls*{cct} to hold, since this is -- in an ideal world -- what the programmer would expect, too: It is the job of the compiler to preserve and (potentially) enforce security~\cite{cauligi2019fact,nagarakatte2010cets,nagarakatte2009soft,akritidis2009baggy}.
 % This paper considers \gls*{cct} as an architecture-dependent property\MPin{can we justify this}, so
@@
     \text{(Events)} & \src{\event} &:=& \src{\preevent;\sandboxtag} \mid \src{\emptyevent} \mid \src{\lightning} \\ 
-    \text{(Values)} & \src{v} &:=& \src{n} \\
     \text{(Expressions)} & \src{\expr} &:=& \src{x} \mid \src{n} \mid \src{\lbinop{\expr[_1]}{\expr[_2]}} \mid \src{\lifz{\expr[_1]}{\expr[_2]}{\expr[_3]}} \\ 
@@
 Because pointers are kept ``alive'' in pointer maps ($\src{\Delta}$), $\Ltms$ dynamic semantics does not get stuck if it encounters a memory-safety violation such as a double--free.
-The memory state carries two heaps to model sandboxing between a context ($\src{H^\ctx}$) and a component ($\src{H^\comp}$).
+The memory state carries two heaps to model sandboxing between a context and a component.
+\MP{notation}
 
@@
   \begin{array}{rrcl}
-    \text{(Values)} & \trg{v} &:=& \dots \mid \trg{\lpair{v_1}{v_2}} \\
     \text{(Expressions)} & \trg{\expr} &:=& \dots \mid \trg{\lhast{\expr}{\tau}} \mid \trg{\lispoisoned{x}}
@@
   Consider again \Cref{ex:strncpy}, with a context copying the string \texttt{Hello World}, where everything is marked with a high security tag: $\lock$.
-  The top half of \Cref{fig:ex-cct} (titled $\Lscct$), describes the execution trace of the program, while the bottom side of the table (titled ,,Specification''), describes the related specification trace (\Cref{subsec:msctss:tracemodel}).
+  The top half of \Cref{fig:ex-cct} (titled $\Lscct$), describes the execution trace of the program, while the bottom side of the table (titled $Spec$), describes the related specification trace (\Cref{subsec:msctss:tracemodel}).
   Read in parallel from top to bottom, the figure shows parts of the execution trace. 
@@
     & \Lscct &        \\
-    \text{Active} & \mid & \text{Inactive} \\\hline
+    Active & \mid & Inactive \\\hline
     \obj{\ev{Alloc}\ \loc_{x}\ 12;\comp;\unlock} & \mid & \obj{\ev{Alloc}\ \loc_{x}\ 12;\comp;\unlock} \\
@@
   \begin{array}{ccc}
-          & \text{Specification} & \\
-  \text{Active} & \mid & \text{Inactive} \\\hline
+          & Spec & \\
+     Active & \mid & Inactive \\\hline
      \ev{Alloc}\ \loc_{x}\ 12;\unlock & \mid & \ev{Alloc}\ \loc_{x}\ 12;\unlock\\
@@
 
-$\Lspec$ extends $\Lscct$ (\Cref{subsec:lscct}) with speculative dynamic semantics that is inspired by existing work~\cite{guarnieri2018spectector,fabian2022automatic}.
-After branching, speculation starts by pushing the current configuration into the speculation state ($\ird{S}$), and run subsequent code in a predetermined window (whose size is $\omega$) until a rollback of operational state is performed.
-The window is set inside the speculation state.
-Within such windows, data may leak, marked as high $\ird{\lock}$ and with an annotation that indicates the respective speculative execution variant, in this case just SPECTRE-PHT~\cite{kocher2019spectre}, whose starting-point is \Cref{tr:e-ifz-true-spec}.
-% where $\omega$ is the size of the speculation window, i.e., a number of reduction steps to be taken under speculative semantics.
+$\Lspec$ extends $\Lscct$ (\Cref{subsec:lscct}) with speculative dynamic semantics that is inspired by existing work~\cite{guarnieri2018spectector}\MP{cit xaver's}.
+After branching or accessing memory\MP{?}, speculation starts by pushing the current configuration on the speculation register ($\ird{S}$), and run subsequent code in a predetermined window until a rollback of operational state is performed.
+The window is set inside the control-flow state ($\ird{\cfstate}$), whose details are in the technical report.
+Within such windows, data is leaked, marked as high $\ird{\lock}$ and with an annotation that indicates the respective speculative execution variant, in this case just SPECTRE-PHT~\cite{kocher2019spectre}, whose semantics is sketched with \Cref{tr:e-ifz-true-spec}, where $\omega$ is the size of the speculation window\MP{role of omega?}.
 Additionally, $\ird{\lock}$ may also carry a $\ird{NONE}$ annotation, to signal that this is a leak irrespective of speculation, i.e., as defined earlier (\Cref{subsec:lscct}).
@@
 In terms of language features, $\Lspec$ supports a barrier operation $\ird{\lbarrier}$ that blocks speculative execution. 
-To facilitate speculative execution in a non-assembly-like language, a stack of operational state is used and speculation is active if that stack has a size greater than one.
+To facilitate speculative execution in a non-assembly-like language, a stack of operational state is used and speculation is active if that stack has a size greater than one, indicated by writing the operational states vertically.
 This is exploited in \Cref{tr:e-spec-eat} to consume any leftover speculation.
@@
     \text{(Expressions)} & \ird{\expr} &:=& \dots \mid \ird{\lbarrier} \\
-    \text{(Speculation State)} & \ird{S} &:=& \ird{\Omega\triangleright\expr} \mid \ird{S},\left(\ird{\Omega};\ird{n};\ird{\expr}\right)
+    \text{(Speculation Register)} & \ird{S} &:=& \ird{\hole{\cdot}} \mid \left(\ird{\Omega};\ird{\expr}\right)
   \end{array}
@@
   \typerule{$e-\ird{ifz}-\ird{true}-\text{spec}$}{
-    \ird{S} = \rtt{(\ird{\cfstate['];\sandboxtag;D;\memstate});\ird{S}}{\ird{\expr[_1]}},({\ird{\cfstate;\sandboxtag;D;\memstate}};\omega;{\ird{\expr[_2]}})
+    \ird{\cfstate[']}.\ird{\operatorname{win}} = \omega &
+    \ird{\trace} = \ird{\ev{Branch}\ 0;\sandboxtag;\securitytag}\cdot\ird{\ev{Spec};\sandboxtag;\securitytag} \\
+    \ird{S} = ({\ird{\cfstate;\sandboxtag;D;\memstate}};{\ird{\expr[_2]}}),\ird{\hole{\cdot}}
   }{
-    \pstepto{\rtt{\ird{\cfstate;\sandboxtag;D;\memstate}}{\ird{\lifz{0^{\securitytag}}{\expr[_1]}{\expr[_2]}}}}{\ird{S}}{\ird{\ev{Branch}\ 0;\sandboxtag;\securitytag}\cdot\ird{\ev{Spec};\sandboxtag;\securitytag}}
+    \pstepto{\rtt{(\ird{\cfstate;\sandboxtag;D;\memstate});\ird{\hole{\cdot}}}{\ird{\lifz{0^{\securitytag}}{\expr[_1]}{\expr[_2]}}}}{\rtt{(\ird{\cfstate['];\sandboxtag;D;\memstate});\ird{S}}{\ird{\expr[_1]}}}{\ird{\trace}}
   }{e-ifz-true-spec}
@@
   \typerule{$e-\ird{spec}-\text{eat}$}{
-    \ird{n} > 0 &
-    \pstepto{\rtt{\ird{\Omega}}{\ird{\expr}}}{\rtt{\ird{\Omega'}}{\ird{\expr[']}}}{\ird{\event}}
+    \ird{\cfstate}.\ird{\operatorname{win}} > 0 &
+    \pstepto{\rtt{(\ird{\cfstate;\sandboxtag;D;\memstate};\ird{\hole{\cdot}})}{\ird{\expr}}}{\rtt{(\ird{\cfstate['];\sandboxtag;D;\memstate};\ird{\hole{\cdot}})}{\ird{\expr[']}}}{\ird{\event}}
+    \\
+    \ird{\cfstate['']}.\ird{\operatorname{win}} = \ird{\cfstate[']}.\ird{\operatorname{win}}-1 &
+    \ird{S} \not=\ird{\hole{\cdot}}
   }{
-    \ghoststepto{\ird{S},(\ird{\Omega};\ird{n};\ird{\expr})}{\ird{S},(\ird{\Omega'};\ird{n}-1;\ird{\expr'})}{\ird{\event}}
+    \ghoststepto{\rtt{(\ird{\cfstate;\sandboxtag;D;\memstate};\ird{S})}{\ird{\expr}}}{\rtt{(\ird{\cfstate[''];\sandboxtag;D;\memstate};\ird{S})}{\ird{\expr[']}}}{\ird{\event}}
   }{e-spec-eat}
@@
   \typerule{$e-\ird{spec}-\text{eaten}$}{
+    \ird{\cfstate}.\ird{\operatorname{win}} = 0 &
+    \ird{S} = \left(\ird{\Omega};\ird{\expr[']}\right)
   }{
-    \ghoststepto{\ird{S},(\ird{\Omega};\ird{0};\ird{\expr})}{\ird{S}}{\ird{Rlb;\Omega.\sandboxtag;\unlock}}
+    \ghoststepto{\rtt{(\ird{\cfstate;\sandboxtag;D;\memstate};\ird{S})}{\ird{\expr}}}{\rtt{(\ird{\Omega};\ird{\hole{\cdot}})}{\ird{\expr[']}}}{\ird{\event}}
   }{e-spec-eaten}
@@
   \typerule{$e-\ird{spec}-\text{barrier}$}{
+    \ird{\cfstate'} = \ird{\cfstate}[.\ird{\operatorname{win}} = 0] &
+    \ird{S} \not=\ird{\hole{\cdot}}
   }{
-    \ghoststepto{\ird{S},(\ird{\Omega};\ird{n};\ird{\lbarrier})}{\ird{S},(\ird{\Omega};\ird{0};\ird{0})}{\ird{Barrier;\Omega.\sandboxtag;\unlock}}
+    \ghoststepto{\rtt{(\ird{\cfstate;\sandboxtag;D;\memstate};\ird{S})}{\ird{\lbarrier}}}{\rtt{(\ird{\cfstate';\sandboxtag;D;\memstate};\ird{\hole{\cdot}})}{\ird{\lbarrier}}}{\ird{\event}}
   }{e-spec-barrier}
@@
 \footnote{In the rule, the notation $\cfstate' = \cfstate[.\text{win} = 0]$ means that $\cfstate'$ is a copy of $\cfstate$ up to field $\text{win}$, which should be set to $0$.}
-Lastly, $\ird{Rlb}$ is emitted when the speculation window is zero and the operationl state is rolled back (see \Cref{tr:e-spec-eaten}).
 
@@
 Compiling the \texttt{strncpy} function from \Cref{sec:introduction} with $\cca$, the compiler would in this case ensure that the arguments that are evaluated in the compiled \texttt{strncpy} are valid.
-%This effectively has no influence on the resulting trace, so a corresponding cross-language trace relation $\sim^{\Ltms}_{\Ltrg}$ is just equality.
+This effectively has no influence on the resulting trace, so a corresponding cross-language trace relation $\sim^{\Ltms}_{\Ltrg}$ is just equality.
+\MPin{what's the point of this relation? where is it used? why define it here?}
 
@@
 % 
-Without \gls*{sms} instrumentation, the event $\ev{Use}\ \loc_{x}\ 2;\comp;\unlock$ would appear on the trace, which is an out-of-bounds access. 
-But, with \gls*{sms} mitigation in place, the use event would not appear during execution, since the bounds check prevents the condition \texttt{src[i] != '\textbackslash 0'} from executing.
+When calling \texttt{strncpy} like this, the event $\ev{Use}\ \loc_{x}\ 2;\comp;\unlock$ would not be emitted during execution, since the bounds check prevents the condition \texttt{src[i] != '\textbackslash 0'} from executing.
+\MP{what?}
 \end{exampleenv}
 
-Contrary to the previous compiler, $\ccb$ may change the trace of the original $\Ltrg$ program: if there is a memory access, it needs to be protected with a bounds check.
-The corresponding relation $\sim^{\Ltrg}_{\Lms}$ that describes this semantic effect of the compiler is partially as follows\footnote{In the interest of space, the full details are in the technical report.}:
+Contrary to the previous compiler, $\ccb$ may change the trace of the original $\Ltrg$ program: \MP{make a uniformity check: always uppercase or always lowercase after :}if there is a memory access, it needs to be protected with a bounds check.
+The corresponding relation $\sim^{\Ltrg}_{\Lms}$ that describes this semantic effect of the compiler may be sketched as follows.
+\MPin{no sketching. rephrase / say that omitted things are omitted for a reason}
 
@@
 
-Action $\trg{Set}$ is treated analogously; any other event is ,,the same''\footnote{Up to colours.}.
+Action $\trg{Set}$ is treated analogously; any other event is just a trivial re-colouring.\MPin{no explicit mentioning of colours. identity functions}
 % Note that neither $\Ltrg$ nor $\Lms$ expose branches on the trace.\MP{so what?}
@@
 
-\Cref{corr:ccab:rtp:ms} states that the composition of $\cca$ and $\ccb$ is secure w.r.t. \gls*{ms} and it follows from \Cref{thm:cca:rtp:tms,thm:ccb:rtp:sms} using \Cref{thm:rtpsim:sig}.
+\Cref{corr:ccab:rtp:ms} states that the composition of $\cca$ and $\ccb$ is secure w.r.t.
+\gls*{ms} and \MP{it.
+many missing such subjects}follows from \Cref{thm:cca:rtp:tms,thm:ccb:rtp:sms} using \Cref{thm:rtpsim:sig}.
 
@@
 Because of this, the corresponding cross-language trace relation $\sim^{\Lms}_{\Lscct}$, only relates events without secrets:%
+\footnote{For context actions, the relation is arbitrary.
+\MPin{is it really?}
+}
 
@@
 
-Clearly, the corresponding cross-language trace relation $\sim^{\Lscct}_{\Lspec}$ has only one non-trivial case: for branches, only relate them where speculation is blocked by a barrier:
+Clearly, the corresponding cross-language trace relation $\sim^{\Lscct}_{\Lspec}$ has only one non-trivial case: For branches, only relate them where speculation is blocked by a barrier:
 
@@
 Next in the chain is $\sim^{\Lms}_{\Lscct}$, which has to be well-formed w.r.t. $\scctsafe$.
-This lemma holds, since a trace that was $\scctsafe$ is $\scctsafe$ even after applying $\sim^{\Lms}_{\Lscct}$: the relation enforces that $\Lscct$ traces related to $\Lms$ traces have no leaks of secrets whatsoever.
+This lemma holds, since a trace that was $\scctsafe$ is $\scctsafe$ even after applying $\sim^{\Lms}_{\Lscct}$: The relation enforces that $\Lscct$ traces related to $\Lms$ traces have no leaks of secrets whatsoever.
 
@@
 
-Compiler designers need to be careful that their implementation preserves the original source language semantics. 
-Similar to this, it is important to consider the cross-language relation, which gives a semantic description of the compiler in terms of traces, and check that the original target (resp.
-source) property is properly translated using the universal (resp.
-existential) image.
-The following discusses this for relations presented in \Cref{sec:casestud:rtp}.
+When modeling some thing using formal methods, the intent is to do so faithfully. 
+That is why it is important to ensure that the properties considered in \Cref{sec:casestud:rtp} follow this ideal.
+It is relatively straightforward to see that, e.g., $\mssafe$ models a basic memory safety property. 
+But, what about $\sigma_{\sim^{\Ltms}_{\Lms}}\left(\mssafe\right)$?
+It is not immediately obvious that the resulting property after projecting it from $\Lms$-level to $\Ltms$-level is similar in spirit to the original $\mssafe$ property (\Cref{def:trace:msdef}) and this is what needs to be discussed when using the presented framework.
+Let us investigate the involved projections from the case study.
 
@@
 {\em All} traces $\trg{\trace}$ with $\trg{\trace}\sim^{\Ltrg}_{\Lms}\irl{\trace}$ are almost identical to $\irl{\trace}$. 
-They differ in memory accesses and crashes, where $\irl{\lightning}$ can be related to a $\trg{\lightning}$ or a memory event, e.g., $\trg{Get\ \loc\ 2}$, for an out-of-bounds access.
-In both cases, the translated event does not represent an out-of-bounds memory access.
+They differ in memory accesses and crashes, where $\irl{\lightning}$ can be related to a $\trg{\lightning}$ or a memory event, e.g., $\trg{Get\ \loc\ 2}$.
+In both cases, the memory access cannot be out of bounds.
 
@@
 
-Previous sections disregarded a glaring but minor technical issue: take $\tmssafe$ (\Cref{def:trace:tmsdef}) as an example.
+Previous sections disregarded a glaring but minor technical issue: Take $\tmssafe$ (\Cref{def:trace:tmsdef}) as an example.
 It contains traces as specified in \Cref{subsec:basic:memsafety:tracemodel}.
@@
   we do not take SNI from spectector because it is a hypersafety.
-  other work defined similar props.
+  other work defined similar props.d
 }

### 2024-05-28 csf2/main.tex

--- 2024-05-28 csf2/main.tex
@@
 
-Even though the considered Spectre variants are just SPECTRE-PHT~\cite{kocher2019spectre}, NONE just describes secret data as in \gls*{scct} (see \Cref{subsec:scct:tracemodel}), the trace model is general enough to allow for potential future extension with different variants.\MP{cits}
+Even though the considered Spectre variants are just SPECTRE-PHT~\cite{kocher2019spectre}, NONE just describes secret data as in \gls*{scct} (see \Cref{subsec:scct:tracemodel}), the trace model is general enough to allow for potential future extension with different variants~\cite{kocher2019spectre,maisuradze2018ret2spec,horn2019zero}.
 %For sake of readability, this paper just uses the notation $\lock$ in place of $\lock_{\text{PHT}}$.
@@
 The language also adds user annotations $\obj{\securitytag}$ for the secrecy of variables, which can be either private (high secrecy) $\obj{\lock}$ or public (low secrecy) $\obj{\unlock}$.
-Security tags $\obj{\securitytag}$ are arranged in the usual secrecy lattice\MP{NI cits}, where $\obj{\lock}\le\obj{\securitytag}$ and $\obj{\securitytag}\le\obj{\unlock}$.
+Security tags $\obj{\securitytag}$ are arranged in the usual secrecy lattice~\MP{NI cits}, where $\obj{\lock}\le\obj{\securitytag}$ and $\obj{\securitytag}\le\obj{\unlock}$.
 
@@
 Additionally, $\ird{\lock}$ may also carry a $\ird{NONE}$ annotation, to signal that this is a leak irrespective of speculation, i.e., as defined earlier (\Cref{subsec:lscct}).
-The semantics is kept general enough to allow for future extension to support different variants.
+The semantics is kept general enough to allow for future extension to support different variants~\cite{kocher2019spectre,maisuradze2018ret2spec,horn2019zero}.
 

### 2024-05-28 csf2/main.tex
spacing, sec1

@@
 \usepackage{bm}
+\usepackage{paralist}
 
@@
 
-\title{Secure Composition \\ of Robust and Optimising Compilers}
-
-\author{\vspace*{-2em}\phantom{a}}
+\title{
+  Secure Composition 
+  \\ 
+  of Robust and Optimising Compilers
+  \vspace{-1.5em}
+}
+
+% \author{
 % Matthis Kruse\orcidID{0000-0003-4062-9666}
@@
 % 	% and a $\obj{green}$, $\obj{teletype}$ font for a $\obj{target}$ language.	
-	Please print or view in colours.
+	For a better experience, please print or view in colours.
 	}
@@
 \begin{exampleenv}[strncpy]\label{ex:strncpy}
-Consider the \texttt{C} function \texttt{strncpy} that copies a null-terminated string \texttt{src} into \texttt{dst} up to a length of \texttt{n}.
-% Assume that the length of the array pointed by \texttt{dst} is at least \texttt{n}, and \texttt{src} is terminated by \texttt{'\textbackslash 0'}.
-% Even under these assumptions, there is a 
+Consider the following \texttt{C} function \texttt{strncpy} that copies a null-terminated string \texttt{src} into \texttt{dst} up to a length of \texttt{n} characters.
 This function is subject to a subtle \gls*{sms} vulnerability: the bounds check \texttt{i < n} should happen {\it before} the access to memory location \texttt{src[i]}: otherwise
 % If \texttt{n} is at least as large as the length of the array \texttt{src}, 
-the memory location past the last element will be leaked to an attacker.
+the memory location past the last element is leaked to an attacker.
 \begin{lstlisting}[language=c,basicstyle=\small\ttfamily,morekeywords={size_t}]
@@
 \end{lstlisting}
-
+% Assume that the length of the array pointed by \texttt{dst} is at least \texttt{n}, and \texttt{src} is terminated by \texttt{'\textbackslash 0'}.
+% Even under these assumptions, there is a 
 To prevent this vulnerability, one can use a compilation pass that enforces \gls*{sms}, such as Softbounds~\cite{nagarakatte2009soft} or BaggyBounds~\cite{akritidis2009baggy}.
-The most naive solution in this case is to insert bounds-checks in front of every access to memory.
+The most na\"ive solution in this case is to insert bounds-checks in front of every access to memory.
 
@@
 %
-Alas, code is not run in isolation, so a malicious attacker could supply code that interacts with \texttt{strncpy} and trigger a violation of either \gls*{ms} or \gls*{cct} by calling \texttt{strncpy} with an argument for \texttt{src} that points to uninitialised memory.
+Finally, even with these precautions, code is not run in isolation, so a malicious attacker could supply code that interacts with \texttt{strncpy} and trigger a violation of either \gls*{ms} or \gls*{cct} by calling \texttt{strncpy} with an argument for \texttt{src} that points to uninitialised memory.
 This would, in turn, trigger a series of reads from uninitialised memory, which is an immediate \gls*{ms} violation with devastating real-world consequences~\cite{uninit-0,uninit-1,uninit-2,uninit-3,uninit-4}.
@@
 % Different kinds of compilation passes exist, e.g., there are usually several compilation passes that perform optimisations~\cite{lattner2004llvm,googlev8,androidstudio,kuepper2023cryptopt,manjikian1997fusion,wegman1991ccp} or source-code instrumentations~\cite{nagarakatte2009soft,nagarakatte2010cets,akritidis2009baggy,dhumbumroong2020boundwarden,jung2021pico,nam2019framer,shankaranarayana2023tailcheck,younan2010paricheck,zhou2023fatptrs,bond2017vale,almeida2017jasmin,kuepper2023cryptopt,cauligi2019fact} that enforce that the compiled program fulfils a property of interest.
-Worse, without a framework for composing secure compiler passes, it is not possible to enable separation of concerns, e.g., to have a secure compilation pass that ensures \gls*{ms} that is developed independently of another secure pass for \gls*{cct}, that is developed independently of other passes, such as optimisation passes.
+Worse, without a framework for composing secure compiler passes, it is not possible to enable separation of concerns, e.g., to have a secure compilation pass that ensures \gls*{ms} that is developed independently of another secure pass for \gls*{cct}, that is developed independently of other passes, such as optimisation ones.
 
@@
 %Then, this paper defines compiler composition and formalises that given two passes that securely preserve two (possibly distinct) properties, their composition securely preserves the composition of those properties.
-The paper then defines several secure compiler passes, where each is either preserving a different security property (\gls*{tms}, \gls*{sms}, \gls*{scct}, \gls*{ss}) or performing a security-preserving optimisation, (e.g., applying \gls*{cf} or \gls*{dce}).
+The paper then defines several secure compiler passes, where each is either preserving a different security property from the list above
+% (\gls*{tms}, \gls*{sms}, \gls*{scct}, \gls*{ss}) 
+or performing a security-preserving optimisation.
+% , (e.g., applying \gls*{cf} or \gls*{dce}).
 Finally, this paper shows that composing these secure compiler passes into a multi-pass compilation chain results in the end-to-end (robust) preservation of \gls*{mssscct}.
@@
 In summary, this paper makes the following contributions:
-\begin{itemize}
+% \begin{itemize}[leftmargin=0em,nosep,label=$\blacktriangleright$]
+\begin{asparaitem}[$\blacktriangleright$]
   \item %
@@
   \item The key contributions of this paper are formalised in the Coq proof assistant and the paper indicates this with \CoqSymbol.
-\end{itemize}
+\end{asparaitem}
 % 
@@
 
-\contribution{Open Source \& Technical Report} A technical report with the omitted formal details, lemmas and proofs, as well as the Coq formalisation are available as supplementary material.
+% \contribution{Technical Report} 
+The omitted formal details, lemmas and proofs, as well as the Coq formalisation are available as supplementary material.
 % \MKin{cite/link}
 
-\section{Background: Security Properties and Secure Compilers\pages{1}}\label{sec:background}
+\section{Background: Properties and Secure Compilers\pages{1}}\label{sec:background}
 
@@
 
-\section{Security Properties: Formalisation, Enforcement and Composition\pages{1.5}}\label{sec:compprop}
+\section{Security Properties Formalisation \& Composition\pages{1.5}}\label{sec:compprop}
 
@@
 This section defines two optimising compiler passes from $\Lms$ to $\Lms$ which perform \gls*{dce} and \gls*{cf}, respectively.
-The \gls*{dce} pass applies a naive rewrite rule on conditionals.
+The \gls*{dce} pass applies a na\"ive rewrite rule on conditionals.
 The \gls*{cf} pass relies on an auxiliary function \texttt{mix} that uses a substitutions accumulator $\irl{\substlist}$ in order to rewrite constant binary operations, e.g., $\irl{{17}-1}$ to $\irl{16}$, and replace variables that are assigned to constants, e.g., $\irl{\llet{x}{7}{x}}$ to $\irl{7}$.

### 2024-05-28 csf2/main.tex
fix placement sigma
replace securitytag sigma with s
four

@@
         This paper presents a case-study with five programming languages showcasing the previous contribution (\Cref{sec:casestud:defs,sec:casestud:rtp}).
-        To this end, it presents a compilation chain consisting of six passes that ultimately preserve \gls*{mssscct} by means of composing four secure passes that \emph{individually} preserve \gls*{tms}, \gls*{sms}, \gls*{scct}, and \gls*{ss}, and two passes which are well-known optimisations \gls*{dce} and \gls*{cf},.
+        To this end, it presents a compilation chain consisting of six passes that ultimately preserve \gls*{mssscct} by means of composing four secure passes that \emph{individually} preserve \gls*{tms}, \gls*{sms}, \gls*{scct}---an enforceable version of \gls*{cct}---, and \gls*{ss}, and two passes which are well-known optimisations \gls*{dce} and \gls*{cf},.
         %Furthermore, the chain includes two optimisation passes: one performs \gls*{dce} and the other \gls*{cf}.
@@
 To this end, considering a source trace $\src{\trace}$ and a target trace $\trg{\trace}$, there is a relation ($\sim$) describing the effect of a corresponding compiler (see \Cref{sec:casestud:rtp}). 
-This relation induces the following two projection functions: (1) the existential image $\tau_\sim\left(\src{\pi}\right)$ and (2) the universal image $\sigma_\sim\left(\trg{\pi}\right)$.
+This relation induces the following two projection functions: (1) the \emph{existential image} $\tau_\sim\left(\src{\pi}\right)$ and (2) the \emph{universal image} $\sigma_\sim\left(\trg{\pi}\right)$.
 These projections map source-level (resp. target-level) properties to target-level (resp. source-level) properties in a way that identifies the ``same'' property across languages. 
@@
 This {\em phase ordering problem} is well--known in literature and a practical solution is to simply perform a fixpoint iteration of the optimisation pipeline~\cite{click1995combining}.
-Hence, it is of interest to analyze the security of compilation passes and their interaction within a compilation pipeline.
 
@@
 
-The key notions for our main results are the definition of a trace relation, and the definition of when is a trace relation well-formed with respect to a class.
+To analyze the security of compilation passes and their interaction within a compilation pipeline, we rely on a few key notions: the definition of a trace relation, and the definition of when is a trace relation well-formed with respect to a class.
 %
@@
 Besides sequential composition, there are two other compositions, namely an {\em upper}, i.e., a compiler that takes multiple inputs and yields one output, and a {\em lower} composition, i.e., a compiler that takes one input and yields multiple outputs.
-We {fefine the upper composition $\cc{\src{L}+\obj{L}}{\trg{L}}$} as follows:
+We {define the upper composition $\cc{\src{L}+\obj{L}}{\trg{L}}$} as follows:
 Given a program \texttt{p}, its compiled counterpart is obtained by {plugging \texttt{p} into $\cc{\src{L}}{\trg{L}}$ if $\texttt{p}\in\src{L}$} or by {plugging \texttt{p} into $\cc{\obj{L}}{\trg{L}}$ if $\texttt{p}\in\obj{L}$}.
@@
 These properties are of practical importance (as mentioned in \Cref{sec:introduction}) and also of interest in the case study presented later (\Cref{sec:casestud:defs,sec:casestud:rtp}). 
-While up to \gls*{scct}, none of the properties is novel, for self-containment of the paper, their definition is necessary for the case study.
+For clarity, this section presents all of them, despite the fact that only property \gls*{scct} is novel, while the other ones are inspired by existing work.
+% , none of the properties is novel, for self-containment of the paper, their definition is necessary for the case study.
 For each of the presented properties, the technical appendix accompanying the paper defines corresponding monitors, which refine a given property and have a key role in the proofs of this paper.
@@
 
-For simple memory safety composed of temporal and spatial memory safety, the trace model defines events as either the empty event ($\emptyevent$), a crash ($\lightning$), or a base-event ($\preevent[_{\mssafe}]$).
+For simple memory safety composed of temporal and spatial memory safety, the trace model defines events ($\event[_{\mssafe}]$) as either the empty event ($\emptyevent$), a crash ($\lightning$), or a base-event ($\preevent[_{\mssafe}]$).
 
@@
 
-This section starts by extending the memory safety trace model (\Cref{subsec:basic:memsafety:tracemodel}) with a {\em security tag} ($\securitytag$) that indicates whether the involved event contains sensitive information ($\lock$) or not ($\unlock$).
+To express Constant Time, we extend the memory safety trace model (\Cref{subsec:basic:memsafety:tracemodel}) with a {\em security tag} ($\securitytag$) that indicates whether the involved event contains sensitive information ($\lock$) or not ($\unlock$).
 
@@
 However, \gls*{ms} uses a different trace model than \gls*{scct}, so intersecting them would trivially yield the empty set. 
-To remedy this issue, let $\sim_{\ctsafe}$ be a cross-language trace relation whose definition is partially follows:
+To remedy this issue, we introduce $\sim_{\ctsafe}: \preevent[_{\ctsafe}] \times \preevent[_{\mssafe}] $, a cross-language trace relation (whose key cases are presented below), that we use to intuitively unify the trace model in which the two properties are expressed:
 \[
@@
 
-The above relation is extended to traces in a standard way, skipping the empty event $\emptyevent$ on either side, and it is now possible to define \gls*{msscct} using the universal image:
+The above relation is extended point-wise to traces, skipping the empty event $\emptyevent$ on either side, and it is now possible to define \gls*{msscct} using the universal image:
 
@@
 
-Similar to \gls*{msscct}, there is a need to relate the different trace models with each other, so that the memory safety property without speculation can be lifted to speculation. 
-To this end, let $\sim_{\mathghost}$ be a cross-language trace relation whose definition is partially as follows:
+As before, we need to relate the different trace models with each other, so that the memory safety property without speculation can be lifted to speculation. 
+To this end, let $\sim_{\mathghost}: \preevent[_{\mathghost}] \times \preevent[_{\ctsafe}]$ be a cross-language trace relation whose key cases are below.
+The intuition is that \gls*{ss} is trivially satisfied in \gls*{scct}, since speculation is inexpressible there, which amounts to dropping events $\ev{Spec}$, $\ev{Rlb}$, or $\ev{Barrier}$, as well as all base events tagged with $\lock_{\text{PHT}}$.
 
@@
 
-The intuition is that \gls*{ss} is trivially satisfied in \gls*{scct}, since speculation is inexpressible there, which amounts to dropping events $\ev{Spec}$, $\ev{Rlb}$, or $\ev{Barrier}$, as well as all base events tagged with $\lock_{\text{PHT}}$.
-
-We conclude this section by defining the ultimate property of interest for secure compilers: \gls*{specms}.
+We conclude by defining the ultimate property of interest for secure compilers: \gls*{specms}.
 \begin{definition}[\gls*{specms}]\label{def:trace:specmsdef}

### 2024-05-28 csf2/main.tex
comperror

@@
 
-To express Constant Time, we extend the memory safety trace model (\Cref{subsec:basic:memsafety:tracemodel}) with a {\em security tag} ($\securitytag$) that indicates whether the involved event contains sensitive information ($\lock$) or not ($\unlock$).
+To express Constant Time, we extend the memory safety trace model (\Cref{subsec:basic:memsafety:tracemodel}) with a {\em security tag} ($\securitytag{}$) that indicates whether the involved event contains sensitive information ($\lock$) or not ($\unlock$).
 
@@
     (\text{Base-Events}) & \preevent[_{\ctsafe}] &:=& \preevent[_{\mssafe}] \mid \ev{Branch\ }n \mid \ev{Binop\ }n\\
-    (\text{Security Tags}) & \securitytag &:=& \lock \mid \unlock\\ 
-    (\text{Events}) & \event[_{\ctsafe}] &:=& \preevent[_{\ctsafe}];\securitytag \mid \emptyevent \mid \lightning \\ 
+    (\text{Security Tags}) & \securitytag{} &:=& \lock \mid \unlock\\ 
+    (\text{Events}) & \event[_{\ctsafe}] &:=& \preevent[_{\ctsafe}];\securitytag{} \mid \emptyevent \mid \lightning \\ 
   \end{array}
@@
 
-\gls*{scct} may appear overly strict, since it seems that secrets must not occur on a trace (since $\securitytag$ is forced to be $\unlock$). 
+\gls*{scct} may appear overly strict, since it seems that secrets must not occur on a trace (since $\securitytag{}$ is forced to be $\unlock$). 
 However, this is considered standard practice in terms of coding guidelines~\cite{ctguidelines}.
@@
 \[
-  \typerulenolabel{mscctrel:drop:tag}{}{\preevent[_{\mssafe}]\sim_{\ctsafe}\preevent[_{\mssafe}];\securitytag}
+  \typerulenolabel{mscctrel:drop:tag}{}{\preevent[_{\mssafe}]\sim_{\ctsafe}\preevent[_{\mssafe}];\securitytag{}}
   \typerulenolabel{mscctrel:drop:crash}{}{\lightning\sim_{\ctsafe}\lightning}
@@
 \[
-  \typerulenolabel{mscctrel:drop:branch}{}{\emptyevent\sim_{\ctsafe}\ev{Branch n};\securitytag}
-  \typerulenolabel{mscctrel:drop:binop}{}{\emptyevent\sim_{\ctsafe}\ev{Binop n};\securitytag}
+  \typerulenolabel{mscctrel:drop:branch}{}{\emptyevent\sim_{\ctsafe}\ev{Branch n};\securitytag{}}
+  \typerulenolabel{mscctrel:drop:binop}{}{\emptyevent\sim_{\ctsafe}\ev{Binop n};\securitytag{}}
 \]
@@
   $
-  \msscctsafe:=\sigma_{\sim_{mct}}\left(\mssafe\right)\cap\scctsafe
+  \msscctsafe:=\mssafe\cap\sigma_{\sim_{\ctsafe}}\left(\mssafe\right)
   $
@@
 So far, the considered trace models do not let us express speculative execution attacks such as Spectre~\cite{kocher2019spectre}. 
-For this, we extend the earlier trace model (see \Cref{subsec:scct:tracemodel}) so that the security tags ($\securitytag$) carry additional information about the kind of private data leakage, i.e., the type of speculative leak.
+For this, we extend the earlier trace model (see \Cref{subsec:scct:tracemodel}) so that the security tags ($\securitytag{}$) carry additional information about the kind of private data leakage, i.e., the type of speculative leak.
 Moreover, we add base-events signalling the beginning of a speculative execution ($\ev{Spec}$), a barrier ($\ev{Barrier}$) that signals that any speculative execution may not go past it, as well as a rollback event ($\ev{Rlb}$), which signals that execution resumes to where speculation started.
@@
     (\text{Spectre Variants}) & vX &:=& \operatorname{NONE} \mid \operatorname{PHT} \\
-    (\text{Security Tags}) & \securitytag &:=& \lock_{vX} \mid \unlock\\ 
-    (\text{Events}) & \event[_{\mathghost}] &:=& \preevent[_{\mathghost}];\securitytag \mid \emptyevent \mid \lightning \mid \ev{Spec} \mid \ev{Rlb} \mid \ev{Barrier}\\ 
+    (\text{Security Tags}) & \securitytag{} &:=& \lock_{vX} \mid \unlock\\ 
+    (\text{Events}) & \event[_{\mathghost}] &:=& \preevent[_{\mathghost}];\securitytag{} \mid \emptyevent \mid \lightning \mid \ev{Spec} \mid \ev{Rlb} \mid \ev{Barrier}\\ 
   \end{array}
@@
   $
-  \specmssafe := \sigma_{\sim_{\mathghost}}\left(\msscctsafe\right)\cap\sssafe
+  \specmssafe := \msscctsafe\cap\sigma_{\sim_{\mathghost}}\left(\sssafe\right)
   $
@@
     \text{(Base-Events)} & \src{\preevent} &:=& \src{\ev{Alloc \loc\ n}} \mid \src{\ev{Dealloc \loc}} \mid \src{\ev{Get \loc\ n}} \mid \src{\ev{Set \loc\ n}} \\
-    \text{(Control Tags)} & \src{\sandboxtag} &:=& \src{\ctx} \mid \src{\comp} \\
-    \text{(Events)} & \src{\event} &:=& \src{\preevent;\sandboxtag} \mid \src{\emptyevent} \mid \src{\lightning} \\ 
+    \text{(Control Tags)} & \src{\sandboxtag{}} &:=& \src{\ctx} \mid \src{\comp} \\
+    \text{(Events)} & \src{\event} &:=& \src{\preevent;\sandboxtag{}} \mid \src{\emptyevent} \mid \src{\lightning} \\ 
     \text{(Values)} & \src{v} &:=& \src{n} \\
@@
     \text{(Memory States)} & \src{\memstate} &:=& \left(\src{H^{\ctx}};\src{H^{\comp}};\src{\Delta}\right)\\
-  \text{(Control States)} & \src{\cfstate} & & \text{(States)}\ \ \ \src{\Omega}\ :=\ \left(\src{\cfstate};\src{\sandboxtag};\src{\memstate}\right)\\
+  \text{(Control States)} & \src{\cfstate} & & \text{(States)}\ \ \ \src{\Omega}\ :=\ \left(\src{\cfstate};\src{\sandboxtag{}};\src{\memstate}\right)\\
   \end{array}
@@
 Programs have sets of pre-determined functions called libraries and it is assumed that functions are marked as being part of some component ($\src{\library_{\comp}}$) or context ($\src{\library_{\ctx}}$).
-Moreover, operational state ($\src{\Omega}$) is a triple consisting of a control-flow state ($\src{\cfstate}$), a control tag ($\src{\sandboxtag}$), and a memory state ($\src{\memstate}$). 
+Moreover, operational state ($\src{\Omega}$) is a triple consisting of a control-flow state ($\src{\cfstate}$), a control tag ($\src{\sandboxtag{}}$), and a memory state ($\src{\memstate}$). 
 The latter carrying information about pointers that are kept ,,alive'' in pointer maps ($\src{\Delta}$), so that dynamic semantics do not get stuck when encountering, e.g., a double-free.
@@
     \text{(Mode Values)} & \obj{D} &:=& \obj{ON} \mid \obj{OFF} \\
-    \text{(Security Tags)} & \obj{\securitytag} &:=& \obj{\lock} \mid \obj{\unlock} \\
+    \text{(Security Tags)} & \obj{\securitytag{}} &:=& \obj{\lock} \mid \obj{\unlock} \\
     \text{(Base-Events)} & \obj{\preevent} &:=& \dots \mid \obj{iGet\ \loc\ v} \mid \obj{iSet\ \loc\ v} \\
                        &&& \mid\ \obj{Binop\ n} \mid \obj{Branch\ n} \\
-    \text{(Events)} & \obj{\event} &:=& \obj{\preevent;\sandboxtag;\securitytag} \mid \obj{\emptyevent} \mid \obj{\lightning} \\
+    \text{(Events)} & \obj{\event} &:=& \obj{\preevent;\sandboxtag{};\securitytag{}} \mid \obj{\emptyevent} \mid \obj{\lightning} \\
     \text{(Expressions)} & \obj{\expr} &:=& \dots \mid \obj{\lrddoit{x}{\expr}} \mid \obj{\lwrdoit{D}} \\
-                         &&&\mid\ \obj{\llet{x^{\securitytag}}{\expr[_1]}{\expr[_2]}}\\
-    \text{(States)} & \obj{\Omega} & := & \left(\obj{\cfstate};\obj{\sandboxtag};\obj{D};\obj{\memstate}\right)\\
+                         &&&\mid\ \obj{\llet{x^{\securitytag{}}}{\expr[_1]}{\expr[_2]}}\\
+    \text{(States)} & \obj{\Omega} & := & \left(\obj{\cfstate};\obj{\sandboxtag{}};\obj{D};\obj{\memstate}\right)\\
   \end{array}
@@
   }{
-    \pstepto{\rtt{\obj{\cfstate;\sandboxtag;D;\memstate}}{\obj{\lwrdoit{OFF}}}}{\rtt{\obj{\cfstate;\sandboxtag;OFF;\memstate}}{\obj{0^{\securitytag}}}}{\obj{\emptyevent}}
+    \pstepto{\rtt{\obj{\cfstate;\sandboxtag{};D;\memstate}}{\obj{\lwrdoit{OFF}}}}{\rtt{\obj{\cfstate;\sandboxtag{};OFF;\memstate}}{\obj{0^{\securitytag{}}}}}{\obj{\emptyevent}}
   }{e-wrdoit-off}
@@
   \typerule{$e-\obj{get}-\in-$noleak}{
-    \obj{\memstate}=\obj{H^{\ctx};H^{\comp};\Delta_1,x\mapsto(\loc;\sandboxtag;\poison),\Delta_2} &
-    \obj{\loc}+\obj{n}\in\text{dom }\obj{H^{\sandboxtag}} \\
-    \obj{\sigma''}=\obj{\sigma\sqcap\sigma'} &
-    \obj{\event} = \obj{{iGet}\ \loc\ n;\sandboxtag;\sigma''}
+    \obj{\memstate}=\obj{H^{\ctx};H^{\comp};\Delta_1,x\mapsto(\loc;\sandboxtag{};\poison),\Delta_2} &
+    \obj{\loc}+\obj{n}\in\text{dom }\obj{H^{\sandboxtag{}}} \\
+    \obj{\securitytag{''}}=\obj{\securitytag{}\sqcap\securitytag{'}} &
+    \obj{\event} = \obj{{iGet}\ \loc\ n;\sandboxtag{};\securitytag{''}}
   }{
     \exprevalo{
-    	\cfstate;\sandboxtag';ON;\memstate
+    	\cfstate;\sandboxtag{'};ON;\memstate
     }{
-    	x^\sigma[n^{\sigma'}]
+      	x^{\securitytag[n^{\securitytag{'}}] }
     }{
-    	\cfstate;\sandboxtag';ON;\memstate
+    	\cfstate;\sandboxtag{'};ON;\memstate
     }{
-    	(H^{\sandboxtag}(\loc+n))^{\sigma''}
+    	(H^{\sandboxtag{}}(\loc+n))^{\securitytag{''}}
     }{
@@
 
-The language also adds user annotations $\obj{\securitytag}$ for the secrecy of variables, which can be either private (high secrecy) $\obj{\lock}$ or public (low secrecy) $\obj{\unlock}$.
-Security tags $\obj{\securitytag}$ are arranged in the usual secrecy lattice~\MP{NI cits}, where $\obj{\lock}\le\obj{\securitytag}$ and $\obj{\securitytag}\le\obj{\unlock}$.
+The language also adds user annotations $\obj{\securitytag{}}$ for the secrecy of variables, which can be either private (high secrecy) $\obj{\lock}$ or public (low secrecy) $\obj{\unlock}$.
+Security tags $\obj{\securitytag{}}$ are arranged in the usual secrecy lattice~\MP{NI cits}, where $\obj{\lock}\le\obj{\securitytag{}}$ and $\obj{\securitytag{}}\le\obj{\unlock}$.
 
@@
 $\Lscct$ extends base-events with $\obj{\ev{iGet}\ \loc\ \valueexpr}$ and $\obj{\ev{iSet}\ \loc\ \valueexpr}$ to prevent secrets from leaking but still enable reasoning about memory safety. 
-Due to the technical setup, the rule needs to check if the access is in bounds ($\obj{\loc}+\obj{n}\in\obj{H^t}$) and update the secrecy tag ($\obj{\securitytag''}=\obj{\securitytag\sqcap\securitytag'}$) according to the lattice described earlier.
+Due to the technical setup, the rule needs to check if the access is in bounds ($\obj{\loc}+\obj{n}\in\obj{H^t}$) and update the secrecy tag ($\obj{\securitytag{''}}=\obj{\securitytag{}\sqcap\securitytag{'}}$) according to the lattice described earlier.
 The precise information carried by the event, e.g., location ($\obj{\loc}$) is taken from the pointer map, which carries some more information irrelevant to this rule ($\obj{\poison}$).
@@
 The language extends base-events with $\obj{\ev{Branch}\ n}$ and $\obj{\ev{Binop}\ n}$ that are emitted when evaluating a branch or certain binary expressions, such as division, respectively, whenever the data independent timing mode is inactive.
-Events are extended with a security-tag ($\obj{\securitytag}$) to signal the secrecy of the involved data.
-
-The evaluation steps are amended to propagate the security-tag annotations $\obj{\securitytag}$.
+Events are extended with a security-tag ($\obj{\securitytag{}}$) to signal the secrecy of the involved data.
+
+The evaluation steps are amended to propagate the security-tag annotations $\obj{\securitytag{}}$.
 When the data independent timing mode is inactive, base-events $\obj{\ev{Branch}\ n}$ and $\obj{\ev{Binop}\ n}$ are emitted for conditionals and binary operations, respectively.
@@
   \typerule{$e-\ird{ifz}-\ird{true}-\text{spec}$}{
-    \ird{S} = \rtt{(\ird{\cfstate['];\sandboxtag;D;\memstate});\ird{S}}{\ird{\expr[_1]}},({\ird{\cfstate;\sandboxtag;D;\memstate}};\omega;{\ird{\expr[_2]}})
+    \ird{S} = \rtt{(\ird{\cfstate['];\sandboxtag{};D;\memstate});\ird{S}}{\ird{\expr[_1]}},({\ird{\cfstate;\sandboxtag{};D;\memstate}};\omega;{\ird{\expr[_2]}})
   }{
-    \pstepto{\rtt{\ird{\cfstate;\sandboxtag;D;\memstate}}{\ird{\lifz{0^{\securitytag}}{\expr[_1]}{\expr[_2]}}}}{\ird{S}}{\ird{\ev{Branch}\ 0;\sandboxtag;\securitytag}\cdot\ird{\ev{Spec};\sandboxtag;\securitytag}}
+    \pstepto{\rtt{\ird{\cfstate;\sandboxtag{};D;\memstate}}{\ird{\lifz{0^{\securitytag{}}}{\expr[_1]}{\expr[_2]}}}}{\ird{S}}{\ird{\ev{Branch}\ 0;\sandboxtag{};\securitytag{}}\cdot\ird{\ev{Spec};\sandboxtag{};\securitytag{}}}
   }{e-ifz-true-spec}
@@
   }{
-    \ghoststepto{\ird{S},(\ird{\Omega};\ird{0};\ird{\expr})}{\ird{S}}{\ird{Rlb;\Omega.\sandboxtag;\unlock}}
+    \ghoststepto{\ird{S},(\ird{\Omega};\ird{0};\ird{\expr})}{\ird{S}}{\ird{Rlb;\Omega.\sandboxtag{};\unlock}}
   }{e-spec-eaten}
@@
   }{
-    \ghoststepto{\ird{S},(\ird{\Omega};\ird{n};\ird{\lbarrier})}{\ird{S},(\ird{\Omega};\ird{0};\ird{0})}{\ird{Barrier;\Omega.\sandboxtag;\unlock}}
+    \ghoststepto{\ird{S},(\ird{\Omega};\ird{n};\ird{\lbarrier})}{\ird{S},(\ird{\Omega};\ird{0};\ird{0})}{\ird{Barrier;\Omega.\sandboxtag{};\unlock}}
   }{e-spec-barrier}
@@
 
-The base-event relation above scales to full events by ensuring the missing annotations ($\obj{\comp;\securitytag}$ and $\ird{\comp;\securitytag}$) are the same.
+The base-event relation above scales to full events by ensuring the missing annotations ($\obj{\comp;\securitytag{}}$ and $\ird{\comp;\securitytag{}}$) are the same.
 

### 2024-05-28 csf2/main.tex
5

@@
 This section defines programming languages $\Ltms$, $\Ltrg$, $\Lms$, $\Lscct$, and $\Lspec$, all of which share common elements (presented in \Cref{subsec:cs:defs}).
-$\Ltms$ is the only statically typed language and exhibits the property that all well-typed programs are \gls*{tms} (\Cref{subsec:ltms}).
+$\Ltms$ is the only statically-typed language, and it exhibits the property that all well-typed programs are \gls*{tms} (\Cref{subsec:ltms}).
 However, not all $\Ltms$ programs are \gls*{sms}.
@@
 $\Lms$ is exactly the same language as $\Ltrg$, but this paper still distinguishes the two for sake of readability (\Cref{subsec:lms}).
-All three languages --- so $\Ltms$, $\Ltrg$, and $\Lms$ --- assume \gls*{cct} to hold, since this is -- in an ideal world -- what the programmer would expect, too: it is the job of the compiler to preserve and (potentially) enforce security~\cite{cauligi2019fact,nagarakatte2010cets,nagarakatte2009soft,akritidis2009baggy}.
+All three languages --- so $\Ltms$, $\Ltrg$, and $\Lms$ --- assume \gls*{scct} to hold, since this is -- in an ideal world -- what the programmer would expect, too: it is the job of the compiler to preserve and (potentially) enforce \gls*{scct} security~\cite{cauligi2019fact,nagarakatte2010cets,nagarakatte2009soft,akritidis2009baggy}.
 % This paper considers \gls*{cct} as an architecture-dependent property\MPin{can we justify this}, so
@@
 This is a standard technique in secure compilation in order to rule out irrelevant context-level events. 
-For example, a context immediately deallocating an allocated memory region twice trivially violates memory safety, but the main interest in secure compilation is at component-level events and if those signal, e.g., a memory safety issue.
+For example, a context immediately deallocating an allocated memory region twice trivially violates memory safety, but the main interest in secure compilation is to preserve component-level security, and thus component events.
 Even though this tagging could be used for blame preservation~\cite{patrignani2023blame}, this beyond the focus of this paper.
-Another key difference is that memory accesses are now explicitly modeled as reads ($\src{\ev{Read\ \loc\ n}}$) and writes ($\src{\ev{Write\ \loc\ n}}$) instead of just uses.
+Another key difference is that memory accesses are now explicitly modelled as reads ($\src{\ev{Read\ \loc\ n}}$) and writes ($\src{\ev{Write\ \loc\ n}}$) instead of just uses.
 
 All languages have at least numbers as values ($\src{v}$) and second class functions ($\src{F}$).
-Functions are modeled as pairs containing the name of one argument and the body of the function.
-Bodies are just ordinary expressions, which can be simple binary operations ($\src{\lbinop{\expr[_1]}{\expr[_2]}}$), conditionals ($\src{\lifz{\expr[_1]}{\expr[_2]}{\expr[_3]}}$), C-like function calls ($\src{\lcall{f}{e}}$) and returns ($\src{\lreturn{\expr}}$), as well as C-like memory operations.
-Programs have sets of pre-determined functions called libraries and it is assumed that functions are marked as being part of some component ($\src{\library_{\comp}}$) or context ($\src{\library_{\ctx}}$).
-Moreover, operational state ($\src{\Omega}$) is a triple consisting of a control-flow state ($\src{\cfstate}$), a control tag ($\src{\sandboxtag{}}$), and a memory state ($\src{\memstate}$). 
-The latter carrying information about pointers that are kept ,,alive'' in pointer maps ($\src{\Delta}$), so that dynamic semantics do not get stuck when encountering, e.g., a double-free.
+Functions are modelled as pairs containing the name of one argument and the body of the function.
+Bodies are just ordinary expressions, which can be simple binary operations ($\src{\lbinop{\expr[_1]}{\expr[_2]}}$), conditionals ($\src{\lifz{\expr[_1]}{\expr[_2]}{\expr[_3]}}$), function calls ($\src{\lcall{f}{e}}$) and returns ($\src{\lreturn{\expr}}$), as well as C-like memory operations.
+Programs have sets of pre-determined functions called libraries and they are marked as being part of some component ($\src{\library_{\comp}}$) or context ($\src{\library_{\ctx}}$).
+
+For the operational semantics, the runtime state ($\src{\Omega}$) is a triple consisting of a control-flow state ($\src{\cfstate}$), a control tag ($\src{\sandboxtag{}}$), and a memory state ($\src{\memstate}$).
+The latter carries information about pointers that are kept ``alive'' in pointer maps ($\src{\Delta}$), so that the semantics does not get stuck when encountering, e.g., a double-free.
 The memory state also carries two heaps to model sandboxing between a context ($\src{H^{\ctx}}$) and a component ($\src{H^{\comp}}$).
@@
 
-$\src{L_{\tmssafe}}$ is the only statically typed language in this case study and restricts functions ($\src{F}$) to the typing signature $\src{\natt\to\natt}$. 
-The type system of $\src{L_{\tmssafe}}$ is inspired by $L^{3}$~\cite{morrisett2005L3,scherer2018fabulous} and enforces that every well-typed $\src{L_{\tmssafe}}$ program satisfies \gls*{tms} (\Cref{thm:wt:tms}).
+$\src{L_{\tmssafe}}$ is the only statically-typed language in this case study and restricts functions ($\src{F}$) to the typing signature $\src{\natt\to\natt}$. 
+The type system of $\src{L_{\tmssafe}}$ is inspired by $L^{3}$~\cite{morrisett2005L3,scherer2018fabulous} and enforces that every well-typed $\src{L_{\tmssafe}}$ program satisfies \gls*{tms}.
+ % (\Cref{thm:wt:tms}).
 
 \begin{theorem}[$\src{L_{\tmssafe}}$-programs are \gls*{tms}]\label{thm:wt:tms}
-  $\;$
-
-  \begin{nscenter}
-    \hfill For any component $\src{\library_{\comp}}$, $\rsat{\src{\library_{\comp}}}{\tmssafe}$ \Coqed
-  \end{nscenter}
+$\rsat{\src{\library_{\comp}}}{\tmssafe}$ \Coqed
+  % $\;$ 
+% 
+  % \begin{nscenter}
+  %   \hfill For any component $\src{\library_{\comp}}$, $\rsat{\src{\library_{\comp}}}{\tmssafe}$ \Coqed
+  % \end{nscenter}
 \end{theorem}
@@
 
-\vspace{-1.5em}
+\vspace{-1em}
 \[
@@
 
-$\Lscct$ extends $\Lms$ (\Cref{subsec:lms}) with a way to read from ($\obj{\lrddoit{x}{\expr}}$) and to write to ($\obj{\lwrdoit{D}}$) a model specific register that controls a data (operand) independent timing mode, a feature that is present in both Arm~\cite[p.~543]{arm-refman} and Intel~\cite[p.~80]{intel-refman} processors.
+$\Lscct$ extends $\Lms$ (\Cref{subsec:lms}) with a way to read from ($\obj{\lrddoit{x}{\expr}}$) and to write to ($\obj{\lwrdoit{D}}$) a model-specific register that controls a data (operand) independent timing mode, a feature that is present in both Arm~\cite[p.~543]{arm-refman} and Intel~\cite[p.~80]{intel-refman} processors.
 To this end, states ($\obj{\Omega}$) are extended with the value of the register ($\obj{D}$), which is initially set to be not active ($\obj{OFF}$).
@@
 The language also adds user annotations $\obj{\securitytag{}}$ for the secrecy of variables, which can be either private (high secrecy) $\obj{\lock}$ or public (low secrecy) $\obj{\unlock}$.
-Security tags $\obj{\securitytag{}}$ are arranged in the usual secrecy lattice~\MP{NI cits}, where $\obj{\lock}\le\obj{\securitytag{}}$ and $\obj{\securitytag{}}\le\obj{\unlock}$.
+Security tags $\obj{\securitytag{}}$ are arranged in the usual secrecy lattice~\MP{NI cits}, where $\obj{\lock}\sqsubseteq\obj{\securitytag{}}$ and $\obj{\securitytag{}}\sqsubseteq\obj{\unlock}$.
 
 Memory accesses to secret data need to be present to reason about memory safety, even when execution is in data-independent timing mode, e.g, \Cref{tr:to-e-get-in-noleak}.
-$\Lscct$ extends base-events with $\obj{\ev{iGet}\ \loc\ \valueexpr}$ and $\obj{\ev{iSet}\ \loc\ \valueexpr}$ to prevent secrets from leaking but still enable reasoning about memory safety. 
-Due to the technical setup, the rule needs to check if the access is in bounds ($\obj{\loc}+\obj{n}\in\obj{H^t}$) and update the secrecy tag ($\obj{\securitytag{''}}=\obj{\securitytag{}\sqcap\securitytag{'}}$) according to the lattice described earlier.
-The precise information carried by the event, e.g., location ($\obj{\loc}$) is taken from the pointer map, which carries some more information irrelevant to this rule ($\obj{\poison}$).
-
-The language extends base-events with $\obj{\ev{Branch}\ n}$ and $\obj{\ev{Binop}\ n}$ that are emitted when evaluating a branch or certain binary expressions, such as division, respectively, whenever the data independent timing mode is inactive.
+$\Lscct$ extends base-events with $\obj{\ev{iGet}\ \loc\ \valueexpr}$ and $\obj{\ev{iSet}\ \loc\ \valueexpr}$ (for data $\obj{i}$ndependent get and set) to prevent secrets from leaking but still enable reasoning about memory safety.
+Due to the technical setup, the rule needs to check if the access is in bounds ($\obj{\loc}+\obj{n}\in\obj{H^t}$) and update the secrecy tag ($\obj{\securitytag{''}}=\obj{\securitytag{}\sqcap\securitytag{'}}$) with the least upper bound of the tags, according to the aforementioned lattice.
+The precise information carried by the event, e.g., location ($\obj{\loc}$) is taken from the pointer map, which carries information irrelevant to this rule ($\obj{\poison}$).
+
+Base-events include $\obj{\ev{Branch}\ n}$ and $\obj{\ev{Binop}\ n}$ that are emitted when evaluating a branch or certain binary expressions, such as division, respectively, whenever the data independent timing mode is inactive.
 Events are extended with a security-tag ($\obj{\securitytag{}}$) to signal the secrecy of the involved data.
@@
   Consider again \Cref{ex:strncpy}, with a context copying the string \texttt{Hello World}, where everything is marked with a high security tag: $\lock$.
-  The top half of \Cref{fig:ex-cct} (titled $\Lscct$), describes the execution trace of the program, while the bottom side of the table (titled ,,Specification''), describes the related specification trace (\Cref{subsec:msctss:tracemodel}).
+  The top half of \Cref{fig:ex-cct} (titled $\Lscct$), describes the execution trace of the program, while the bottom side of the table (titled ``Specification''), describes the related specification trace (\Cref{subsec:msctss:tracemodel}).
   Read in parallel from top to bottom, the figure shows parts of the execution trace. 
@@
 The window is set inside the speculation state.
-Within such windows, data may leak, marked as high $\ird{\lock}$ and with an annotation that indicates the respective speculative execution variant, in this case just SPECTRE-PHT~\cite{kocher2019spectre}, whose starting-point is \Cref{tr:e-ifz-true-spec}.
+Within such windows, data may leak, this is marked as high $\ird{\lock}$ and with an annotation that indicates the respective speculative execution variant. 
+In this paper, we just consider SPECTRE-PHT~\cite{kocher2019spectre}, whose starting point is \Cref{tr:e-ifz-true-spec}.
 % where $\omega$ is the size of the speculation window, i.e., a number of reduction steps to be taken under speculative semantics.
@@
 
-In terms of language features, $\Lspec$ supports a barrier operation $\ird{\lbarrier}$ that blocks speculative execution. 
+In terms of language features, $\Lspec$ includes a new barrier operation $\ird{\lbarrier}$ that blocks speculative execution. 
 To facilitate speculative execution in a non-assembly-like language, a stack of operational state is used and speculation is active if that stack has a size greater than one.
@@
 \end{figure*}
-The section demonstrates the power of the framework (\Cref{sec:sequential}) by composing these compilers for a secure and optimising compilation chain that robustly preserves \gls*{mssscct}.
+The section demonstrates the power of the framework (\Cref{sec:sequential}) by composing these compilers for a secure and optimising compilation chain that robustly preserves \gls*{mssscct}. %specms
 The first step in this chain is the compiler from $\src{L_{\tmssafe}}$ to $\trg{L}$ that robustly preserves just \gls*{tms} (\Cref{thm:cca:rtp:tms}).
@@
 
-Action $\trg{Set}$ is treated analogously; any other event is ,,the same''\footnote{Up to colours.}.
+Action $\trg{Set}$ is treated analogously; any other event is ``the same''\footnote{Up to colours.}.
 % Note that neither $\Ltrg$ nor $\Lms$ expose branches on the trace.\MP{so what?}

### 2024-05-28 csf2/main.tex
6

@@
 A $\ird{\lbarrier}$ prevents any execution whatsoever besides $\ird{\ev{Rlb}}$ when run in speculative execution mode and does nothing when run in normal mode, e.g., \Cref{tr:e-spec-barrier}.%
-\footnote{In the rule, the notation $\cfstate' = \cfstate[.\text{win} = 0]$ means that $\cfstate'$ is a copy of $\cfstate$ up to field $\text{win}$, which should be set to $0$.}
+\footnote{In the rule, the notation $\cfstate' = \cfstate[.\text{win} = 0]$ means that $\cfstate'$ is a copy of $\cfstate$ up to field $\text{win}$, which is set to $0$.}
 Lastly, $\ird{Rlb}$ is emitted when the speculation window is zero and the operationl state is rolled back (see \Cref{tr:e-spec-eaten}).
@@
 \end{figure*}
-The section demonstrates the power of the framework (\Cref{sec:sequential}) by composing these compilers for a secure and optimising compilation chain that robustly preserves \gls*{mssscct}. %specms
+The section demonstrates the power of the framework (\Cref{sec:sequential}) by composing these compilers for a secure and optimising compilation chain that robustly preserves \gls*{mssscct}.\MP{ replace with specms?}
 The first step in this chain is the compiler from $\src{L_{\tmssafe}}$ to $\trg{L}$ that robustly preserves just \gls*{tms} (\Cref{thm:cca:rtp:tms}).
-From here, an instrumentation from $\trg{L}$ to $\irl{L_{\mssafe}}$ ensures that no out-of-bounds accesses can happen and, thus, programs at this point attain \gls*{sms} (\Cref{thm:ccb:rtp:sms}).
+From here, another pass from $\trg{L}$ to $\irl{L_{\mssafe}}$ ensures that no out-of-bounds accesses can happen and, thus, programs at this point attain \gls*{sms} (\Cref{thm:ccb:rtp:sms}).
 Since these properties compose into \gls*{ms}, composing these passes yields a compiler that robustly preserves \gls*{ms} (\Cref{corr:ccab:rtp:ms}).
-Then, the section presents two optimising translations, namely \gls*{cf} and \gls*{dce}, each of which robustly preserves \gls*{ms} (\Cref{thm:ccdce:rtp:ms,thm:cccf:rtp:ms}).
-These translations can be freely ordered in the compilation chain without compromising memory safety (\Cref{thm:cccfccdce:rtp:ms}).
-The next step in the chain ensures that code stays \gls*{scct} (\Cref{thm:ccscct:rtp:scct}) when lowered from $\Lms$ to $\Lscct$, which is done by switching on a data-independent timing mode for the computation.
+Then, the section presents two optimisation passes, namely \gls*{cf} and \gls*{dce}, each of which robustly preserves \gls*{ms} (\Cref{thm:ccdce:rtp:ms,thm:cccf:rtp:ms}).
+These passes can be freely ordered in the compilation chain without compromising memory safety (\Cref{thm:cccfccdce:rtp:ms}).
+The next step in the chain ensures that code stays \gls*{scct} (\Cref{thm:ccscct:rtp:scct}) when compiled from $\Lms$ to $\Lscct$, which is done by switching on a data-independent timing mode for the computation.
 Lastly, by introducing barriers immediately after branches, speculative leaks via SPECTRE-PHT are prevented when compiling $\Lscct$ to $\Lspec$.
@@
 % 
-Without \gls*{sms} instrumentation, the event $\ev{Use}\ \loc_{x}\ 2;\comp;\unlock$ would appear on the trace, which is an out-of-bounds access. 
-But, with \gls*{sms} mitigation in place, the use event would not appear during execution, since the bounds check prevents the condition \texttt{src[i] != '\textbackslash 0'} from executing.
+Without the \gls*{sms} pass, the event $\ev{Use}\ \loc_{x}\ 2;\comp;\unlock$ would appear on the trace, but that indicates an out-of-bounds access! 
+Fortunately, with \gls*{sms} mitigation in place, that event does not appear during execution, since the bounds check prevents the condition \texttt{src[i] != '\textbackslash 0'} from executing.
 \end{exampleenv}
@@
 Contrary to the previous compiler, $\ccb$ may change the trace of the original $\Ltrg$ program: if there is a memory access, it needs to be protected with a bounds check.
-The corresponding relation $\sim^{\Ltrg}_{\Lms}$ that describes this semantic effect of the compiler is partially as follows\footnote{In the interest of space, the full details are in the technical report.}:
+% 
+The corresponding relation $\sim^{\Ltrg}_{\Lms}$ that describes this semantic effect of the compiler is defined partially below.
+% 
+We omit action $\trg{Set}$, which is treated analogously, and any other event, which is related to its cross-language equivalent.
+% \footnote{Up to colours.}
+% \footnote{In the interest of space, the full details are in the technical report.}
 
@@
 
-Action $\trg{Set}$ is treated analogously; any other event is ``the same''\footnote{Up to colours.}.
 % Note that neither $\Ltrg$ nor $\Lms$ expose branches on the trace.\MP{so what?}
@@
 % since the precise definition needs some technicalities that only inhibit readability without any technical insights.
-
-Compiler $\ccb$ is robustly preserving \gls*{sms}.
+% 
+We can now prove that compiler $\ccb$ robustly preserves \gls*{sms}.
 \begin{theorem}[$\ccb$ secure w.r.t. \gls*{sms}]\label{thm:ccb:rtp:sms}
@@
 
+\MPin{
+  trace rel eq is not defined.
+  jump.
+  why arewe defining this ?
+}
+\MPin{
+  when defining the sim, state its type (as i did before). here and later too.
+}
 Let $\src{\trace}\sim^{\Ltms}_{\Lms}\irl{\trace}:=\src{\trace}\sim^{\Ltms}_{\Ltrg}\bullet\sim^{\Ltrg}_{\Lms}\irl{\trace}$.
@@
 \begin{lemma}[$\sim^{\Ltms}_{\Ltrg}$ well-formed w.r.t. $\smssafe$]\label{lem:wf:ltmsltrg}
+$\;$
+
+  \begin{nscenter}
   $\wfcsig{\sim^{\Ltms}_{\Ltrg}}{\smssafe}$
+  \end{nscenter}
 \end{lemma}
@@
 \begin{theorem}[$\cccf\circ\ccdce$ and $\cccf\circ\ccdce$ are secure w.r.t. \gls*{ms}]\label{thm:cccfccdce:rtp:ms}
-  $\rtp{\cccf\circ\ccdce}{\mssafe}$ and $\rtp{\ccdce\circ\cccf}{\mssafe}$. % \Coqed
+$\;$
+
+  \begin{nscenter}
+  \phantom{and~}\!\!$\rtp{\cccf\circ\ccdce}{\mssafe}$
+
+  and~$\rtp{\ccdce\circ\cccf}{\mssafe}$ % \Coqed
+  \end{nscenter}
 \end{theorem}
@@
 %
-The context can overwrite the flag and exit the mode, but upon invoking a function that is part of the component, the flag would be set again.
+The context can overwrite the flag and exit the mode, but upon invoking a function that is part of the component, the flag is set again.
 Because of this, the corresponding cross-language trace relation $\sim^{\Lms}_{\Lscct}$, only relates events without secrets:%
@@
 The base-event relation above scales to full events by ensuring the missing annotations ($\obj{\comp;\securitytag{}}$ and $\ird{\comp;\securitytag{}}$) are the same.
-
-With this, $\ccspec$ is secure with respect to \gls*{ss}.
+% 
+With this relation, we prove that $\ccspec$ is secure with respect to \gls*{ss}.
 \begin{theorem}[$\ccspec$ secure w.r.t. \gls*{ss}]\label{thm:ccspec:rtp:spec}
@@
 \begin{lemma}[$\sim^{\Lms}_{\Lscct}$ well-formed w.r.t. $\scctsafe$]\label{lem:wf:lsmslscct}
+$\;$
+
+  \begin{nscenter}
   $\wfcsig{\sim^{\Lms}_{\Lscct}}{\scctsafe}$
+  \end{nscenter}
 \end{lemma}

### 2024-05-28 csf2/main.tex
add phd thesis

--- 2024-05-28 csf2/main.tex
@@
         This paper presents a case-study with five programming languages showcasing the previous contribution (\Cref{sec:casestud:defs,sec:casestud:rtp}).
-        To this end, it presents a compilation chain consisting of six passes that ultimately preserve \gls*{mssscct} by means of composing four secure passes that \emph{individually} preserve \gls*{tms}, \gls*{sms}, \gls*{scct}---an enforceable version of \gls*{cct}---, and \gls*{ss}, and two passes which are well-known optimisations \gls*{dce} and \gls*{cf},.
+        To this end, it presents a compilation chain consisting of six passes that ultimately preserve \gls*{mssscct} by means of composing four secure passes that \emph{individually} preserve \gls*{tms}, \gls*{sms}, \gls*{scct}, and \gls*{ss}, and two passes which are well-known optimisations \gls*{dce} and \gls*{cf},.
         %Furthermore, the chain includes two optimisation passes: one performs \gls*{dce} and the other \gls*{cf}.
@@
 To this end, considering a source trace $\src{\trace}$ and a target trace $\trg{\trace}$, there is a relation ($\sim$) describing the effect of a corresponding compiler (see \Cref{sec:casestud:rtp}). 
-This relation induces the following two projection functions: (1) the \emph{existential image} $\tau_\sim\left(\src{\pi}\right)$ and (2) the \emph{universal image} $\sigma_\sim\left(\trg{\pi}\right)$.
+This relation induces the following two projection functions: (1) the existential image $\tau_\sim\left(\src{\pi}\right)$ and (2) the universal image $\sigma_\sim\left(\trg{\pi}\right)$.
 These projections map source-level (resp. target-level) properties to target-level (resp. source-level) properties in a way that identifies the ``same'' property across languages. 
@@
 This {\em phase ordering problem} is well--known in literature and a practical solution is to simply perform a fixpoint iteration of the optimisation pipeline~\cite{click1995combining}.
+Hence, it is of interest to analyze the security of compilation passes and their interaction within a compilation pipeline.
 
@@
 
-To analyze the security of compilation passes and their interaction within a compilation pipeline, we rely on a few key notions: the definition of a trace relation, and the definition of when is a trace relation well-formed with respect to a class.
+The key notions for our main results are the definition of a trace relation, and the definition of when is a trace relation well-formed with respect to a class.
 %
@@
 Besides sequential composition, there are two other compositions, namely an {\em upper}, i.e., a compiler that takes multiple inputs and yields one output, and a {\em lower} composition, i.e., a compiler that takes one input and yields multiple outputs.
-We {define the upper composition $\cc{\src{L}+\obj{L}}{\trg{L}}$} as follows:
+We {fefine the upper composition $\cc{\src{L}+\obj{L}}{\trg{L}}$} as follows:
 Given a program \texttt{p}, its compiled counterpart is obtained by {plugging \texttt{p} into $\cc{\src{L}}{\trg{L}}$ if $\texttt{p}\in\src{L}$} or by {plugging \texttt{p} into $\cc{\obj{L}}{\trg{L}}$ if $\texttt{p}\in\obj{L}$}.
@@
 These properties are of practical importance (as mentioned in \Cref{sec:introduction}) and also of interest in the case study presented later (\Cref{sec:casestud:defs,sec:casestud:rtp}). 
-For clarity, this section presents all of them, despite the fact that only property \gls*{scct} is novel, while the other ones are inspired by existing work.
-% , none of the properties is novel, for self-containment of the paper, their definition is necessary for the case study.
+While up to \gls*{scct}, none of the properties is novel, for self-containment of the paper, their definition is necessary for the case study.
 For each of the presented properties, the technical appendix accompanying the paper defines corresponding monitors, which refine a given property and have a key role in the proofs of this paper.
@@
 
-For simple memory safety composed of temporal and spatial memory safety, the trace model defines events ($\event[_{\mssafe}]$) as either the empty event ($\emptyevent$), a crash ($\lightning$), or a base-event ($\preevent[_{\mssafe}]$).
+For simple memory safety composed of temporal and spatial memory safety, the trace model defines events as either the empty event ($\emptyevent$), a crash ($\lightning$), or a base-event ($\preevent[_{\mssafe}]$).
 
@@
 
-To express Constant Time, we extend the memory safety trace model (\Cref{subsec:basic:memsafety:tracemodel}) with a {\em security tag} ($\securitytag{}$) that indicates whether the involved event contains sensitive information ($\lock$) or not ($\unlock$).
+This section starts by extending the memory safety trace model (\Cref{subsec:basic:memsafety:tracemodel}) with a {\em security tag} ($\securitytag$) that indicates whether the involved event contains sensitive information ($\lock$) or not ($\unlock$).
 
@@
     (\text{Base-Events}) & \preevent[_{\ctsafe}] &:=& \preevent[_{\mssafe}] \mid \ev{Branch\ }n \mid \ev{Binop\ }n\\
-    (\text{Security Tags}) & \securitytag{} &:=& \lock \mid \unlock\\ 
-    (\text{Events}) & \event[_{\ctsafe}] &:=& \preevent[_{\ctsafe}];\securitytag{} \mid \emptyevent \mid \lightning \\ 
+    (\text{Security Tags}) & \securitytag &:=& \lock \mid \unlock\\ 
+    (\text{Events}) & \event[_{\ctsafe}] &:=& \preevent[_{\ctsafe}];\securitytag \mid \emptyevent \mid \lightning \\ 
   \end{array}
@@
 
-\gls*{scct} may appear overly strict, since it seems that secrets must not occur on a trace (since $\securitytag{}$ is forced to be $\unlock$). 
+\gls*{scct} may appear overly strict, since it seems that secrets must not occur on a trace (since $\securitytag$ is forced to be $\unlock$). 
 However, this is considered standard practice in terms of coding guidelines~\cite{ctguidelines}.
@@
 However, \gls*{ms} uses a different trace model than \gls*{scct}, so intersecting them would trivially yield the empty set. 
-To remedy this issue, we introduce $\sim_{\ctsafe}: \preevent[_{\ctsafe}] \times \preevent[_{\mssafe}] $, a cross-language trace relation (whose key cases are presented below), that we use to intuitively unify the trace model in which the two properties are expressed:
-\[
-  \typerulenolabel{mscctrel:drop:tag}{}{\preevent[_{\mssafe}]\sim_{\ctsafe}\preevent[_{\mssafe}];\securitytag{}}
+To remedy this issue, let $\sim_{\ctsafe}$ be a cross-language trace relation whose definition is partially follows:
+\[
+  \typerulenolabel{mscctrel:drop:tag}{}{\preevent[_{\mssafe}]\sim_{\ctsafe}\preevent[_{\mssafe}];\securitytag}
   \typerulenolabel{mscctrel:drop:crash}{}{\lightning\sim_{\ctsafe}\lightning}
@@
 \[
-  \typerulenolabel{mscctrel:drop:branch}{}{\emptyevent\sim_{\ctsafe}\ev{Branch n};\securitytag{}}
-  \typerulenolabel{mscctrel:drop:binop}{}{\emptyevent\sim_{\ctsafe}\ev{Binop n};\securitytag{}}
+  \typerulenolabel{mscctrel:drop:branch}{}{\emptyevent\sim_{\ctsafe}\ev{Branch n};\securitytag}
+  \typerulenolabel{mscctrel:drop:binop}{}{\emptyevent\sim_{\ctsafe}\ev{Binop n};\securitytag}
 \]
@@
 
-The above relation is extended point-wise to traces, skipping the empty event $\emptyevent$ on either side, and it is now possible to define \gls*{msscct} using the universal image:
+The above relation is extended to traces in a standard way, skipping the empty event $\emptyevent$ on either side, and it is now possible to define \gls*{msscct} using the universal image:
 
@@
 So far, the considered trace models do not let us express speculative execution attacks such as Spectre~\cite{kocher2019spectre}. 
-For this, we extend the earlier trace model (see \Cref{subsec:scct:tracemodel}) so that the security tags ($\securitytag{}$) carry additional information about the kind of private data leakage, i.e., the type of speculative leak.
+For this, we extend the earlier trace model (see \Cref{subsec:scct:tracemodel}) so that the security tags ($\securitytag$) carry additional information about the kind of private data leakage, i.e., the type of speculative leak.
 Moreover, we add base-events signalling the beginning of a speculative execution ($\ev{Spec}$), a barrier ($\ev{Barrier}$) that signals that any speculative execution may not go past it, as well as a rollback event ($\ev{Rlb}$), which signals that execution resumes to where speculation started.
@@
     (\text{Spectre Variants}) & vX &:=& \operatorname{NONE} \mid \operatorname{PHT} \\
-    (\text{Security Tags}) & \securitytag{} &:=& \lock_{vX} \mid \unlock\\ 
-    (\text{Events}) & \event[_{\mathghost}] &:=& \preevent[_{\mathghost}];\securitytag{} \mid \emptyevent \mid \lightning \mid \ev{Spec} \mid \ev{Rlb} \mid \ev{Barrier}\\ 
+    (\text{Security Tags}) & \securitytag &:=& \lock_{vX} \mid \unlock\\ 
+    (\text{Events}) & \event[_{\mathghost}] &:=& \preevent[_{\mathghost}];\securitytag \mid \emptyevent \mid \lightning \mid \ev{Spec} \mid \ev{Rlb} \mid \ev{Barrier}\\ 
   \end{array}
@@
 
-As before, we need to relate the different trace models with each other, so that the memory safety property without speculation can be lifted to speculation. 
-To this end, let $\sim_{\mathghost}: \preevent[_{\mathghost}] \times \preevent[_{\ctsafe}]$ be a cross-language trace relation whose key cases are below.
-The intuition is that \gls*{ss} is trivially satisfied in \gls*{scct}, since speculation is inexpressible there, which amounts to dropping events $\ev{Spec}$, $\ev{Rlb}$, or $\ev{Barrier}$, as well as all base events tagged with $\lock_{\text{PHT}}$.
+Similar to \gls*{msscct}, there is a need to relate the different trace models with each other, so that the memory safety property without speculation can be lifted to speculation. 
+To this end, let $\sim_{\mathghost}$ be a cross-language trace relation whose definition is partially as follows:
 
@@
 
-We conclude by defining the ultimate property of interest for secure compilers: \gls*{specms}.
+The intuition is that \gls*{ss} is trivially satisfied in \gls*{scct}, since speculation is inexpressible there, which amounts to dropping events $\ev{Spec}$, $\ev{Rlb}$, or $\ev{Barrier}$, as well as all base events tagged with $\lock_{\text{PHT}}$.
+
+We conclude this section by defining the ultimate property of interest for secure compilers: \gls*{specms}.
 \begin{definition}[\gls*{specms}]\label{def:trace:specmsdef}
@@
     \text{(Base-Events)} & \src{\preevent} &:=& \src{\ev{Alloc \loc\ n}} \mid \src{\ev{Dealloc \loc}} \mid \src{\ev{Get \loc\ n}} \mid \src{\ev{Set \loc\ n}} \\
-    \text{(Control Tags)} & \src{\sandboxtag{}} &:=& \src{\ctx} \mid \src{\comp} \\
-    \text{(Events)} & \src{\event} &:=& \src{\preevent;\sandboxtag{}} \mid \src{\emptyevent} \mid \src{\lightning} \\ 
+    \text{(Control Tags)} & \src{\sandboxtag} &:=& \src{\ctx} \mid \src{\comp} \\
+    \text{(Events)} & \src{\event} &:=& \src{\preevent;\sandboxtag} \mid \src{\emptyevent} \mid \src{\lightning} \\ 
     \text{(Values)} & \src{v} &:=& \src{n} \\
@@
     \text{(Memory States)} & \src{\memstate} &:=& \left(\src{H^{\ctx}};\src{H^{\comp}};\src{\Delta}\right)\\
-  \text{(Control States)} & \src{\cfstate} & & \text{(States)}\ \ \ \src{\Omega}\ :=\ \left(\src{\cfstate};\src{\sandboxtag{}};\src{\memstate}\right)\\
+  \text{(Control States)} & \src{\cfstate} & & \text{(States)}\ \ \ \src{\Omega}\ :=\ \left(\src{\cfstate};\src{\sandboxtag};\src{\memstate}\right)\\
   \end{array}
@@
 Programs have sets of pre-determined functions called libraries and it is assumed that functions are marked as being part of some component ($\src{\library_{\comp}}$) or context ($\src{\library_{\ctx}}$).
-Moreover, operational state ($\src{\Omega}$) is a triple consisting of a control-flow state ($\src{\cfstate}$), a control tag ($\src{\sandboxtag{}}$), and a memory state ($\src{\memstate}$). 
+Moreover, operational state ($\src{\Omega}$) is a triple consisting of a control-flow state ($\src{\cfstate}$), a control tag ($\src{\sandboxtag}$), and a memory state ($\src{\memstate}$). 
 The latter carrying information about pointers that are kept ,,alive'' in pointer maps ($\src{\Delta}$), so that dynamic semantics do not get stuck when encountering, e.g., a double-free.
@@
     \text{(Mode Values)} & \obj{D} &:=& \obj{ON} \mid \obj{OFF} \\
-    \text{(Security Tags)} & \obj{\securitytag{}} &:=& \obj{\lock} \mid \obj{\unlock} \\
+    \text{(Security Tags)} & \obj{\securitytag} &:=& \obj{\lock} \mid \obj{\unlock} \\
     \text{(Base-Events)} & \obj{\preevent} &:=& \dots \mid \obj{iGet\ \loc\ v} \mid \obj{iSet\ \loc\ v} \\
                        &&& \mid\ \obj{Binop\ n} \mid \obj{Branch\ n} \\
-    \text{(Events)} & \obj{\event} &:=& \obj{\preevent;\sandboxtag{};\securitytag{}} \mid \obj{\emptyevent} \mid \obj{\lightning} \\
+    \text{(Events)} & \obj{\event} &:=& \obj{\preevent;\sandboxtag;\securitytag} \mid \obj{\emptyevent} \mid \obj{\lightning} \\
     \text{(Expressions)} & \obj{\expr} &:=& \dots \mid \obj{\lrddoit{x}{\expr}} \mid \obj{\lwrdoit{D}} \\
-                         &&&\mid\ \obj{\llet{x^{\securitytag{}}}{\expr[_1]}{\expr[_2]}}\\
-    \text{(States)} & \obj{\Omega} & := & \left(\obj{\cfstate};\obj{\sandboxtag{}};\obj{D};\obj{\memstate}\right)\\
+                         &&&\mid\ \obj{\llet{x^{\securitytag}}{\expr[_1]}{\expr[_2]}}\\
+    \text{(States)} & \obj{\Omega} & := & \left(\obj{\cfstate};\obj{\sandboxtag};\obj{D};\obj{\memstate}\right)\\
   \end{array}
@@
   }{
-    \pstepto{\rtt{\obj{\cfstate;\sandboxtag{};D;\memstate}}{\obj{\lwrdoit{OFF}}}}{\rtt{\obj{\cfstate;\sandboxtag{};OFF;\memstate}}{\obj{0^{\securitytag{}}}}}{\obj{\emptyevent}}
+    \pstepto{\rtt{\obj{\cfstate;\sandboxtag;D;\memstate}}{\obj{\lwrdoit{OFF}}}}{\rtt{\obj{\cfstate;\sandboxtag;OFF;\memstate}}{\obj{0^{\securitytag}}}}{\obj{\emptyevent}}
   }{e-wrdoit-off}
@@
   \typerule{$e-\obj{get}-\in-$noleak}{
-    \obj{\memstate}=\obj{H^{\ctx};H^{\comp};\Delta_1,x\mapsto(\loc;\sandboxtag{};\poison),\Delta_2} &
-    \obj{\loc}+\obj{n}\in\text{dom }\obj{H^{\sandboxtag{}}} \\
-    \obj{\securitytag{''}}=\obj{\securitytag{}\sqcap\securitytag{'}} &
-    \obj{\event} = \obj{{iGet}\ \loc\ n;\sandboxtag{};\securitytag{''}}
+    \obj{\memstate}=\obj{H^{\ctx};H^{\comp};\Delta_1,x\mapsto(\loc;\sandboxtag;\poison),\Delta_2} &
+    \obj{\loc}+\obj{n}\in\text{dom }\obj{H^{\sandboxtag}} \\
+    \obj{\securitytag''}=\obj{\securitytag\sqcap\securitytag'} &
+    \obj{\event} = \obj{{iGet}\ \loc\ n;\sandboxtag;\securitytag''}
   }{
     \exprevalo{
-    	\cfstate;\sandboxtag{'};ON;\memstate
+    	\cfstate;\sandboxtag';ON;\memstate
     }{
-      	x^{\securitytag[n^{\securitytag{'}}] }
+    	x^\securitytag[n^{\securitytag'}]
     }{
-    	\cfstate;\sandboxtag{'};ON;\memstate
+    	\cfstate;\sandboxtag';ON;\memstate
     }{
-    	(H^{\sandboxtag{}}(\loc+n))^{\securitytag{''}}
+    	(H^{\sandboxtag}(\loc+n))^{\securitytag''}
     }{
@@
 
-The language also adds user annotations $\obj{\securitytag{}}$ for the secrecy of variables, which can be either private (high secrecy) $\obj{\lock}$ or public (low secrecy) $\obj{\unlock}$.
-Security tags $\obj{\securitytag{}}$ are arranged in the usual secrecy lattice~\MP{NI cits}, where $\obj{\lock}\le\obj{\securitytag{}}$ and $\obj{\securitytag{}}\le\obj{\unlock}$.
+The language also adds user annotations $\obj{\securitytag}$ for the secrecy of variables, which can be either private (high secrecy) $\obj{\lock}$ or public (low secrecy) $\obj{\unlock}$.
+Security tags $\obj{\securitytag}$ are arranged in the usual secrecy lattice~\cite{zdancewicphd}, where $\obj{\lock}\le\obj{\securitytag}$ and $\obj{\securitytag}\le\obj{\unlock}$.
 
@@
 $\Lscct$ extends base-events with $\obj{\ev{iGet}\ \loc\ \valueexpr}$ and $\obj{\ev{iSet}\ \loc\ \valueexpr}$ to prevent secrets from leaking but still enable reasoning about memory safety. 
-Due to the technical setup, the rule needs to check if the access is in bounds ($\obj{\loc}+\obj{n}\in\obj{H^t}$) and update the secrecy tag ($\obj{\securitytag{''}}=\obj{\securitytag{}\sqcap\securitytag{'}}$) according to the lattice described earlier.
+Due to the technical setup, the rule needs to check if the access is in bounds ($\obj{\loc}+\obj{n}\in\obj{H^t}$) and update the secrecy tag ($\obj{\securitytag''}=\obj{\securitytag\sqcap\securitytag'}$) according to the lattice described earlier.
 The precise information carried by the event, e.g., location ($\obj{\loc}$) is taken from the pointer map, which carries some more information irrelevant to this rule ($\obj{\poison}$).
@@
 The language extends base-events with $\obj{\ev{Branch}\ n}$ and $\obj{\ev{Binop}\ n}$ that are emitted when evaluating a branch or certain binary expressions, such as division, respectively, whenever the data independent timing mode is inactive.
-Events are extended with a security-tag ($\obj{\securitytag{}}$) to signal the secrecy of the involved data.
-
-The evaluation steps are amended to propagate the security-tag annotations $\obj{\securitytag{}}$.
+Events are extended with a security-tag ($\obj{\securitytag}$) to signal the secrecy of the involved data.
+
+The evaluation steps are amended to propagate the security-tag annotations $\obj{\securitytag}$.
 When the data independent timing mode is inactive, base-events $\obj{\ev{Branch}\ n}$ and $\obj{\ev{Binop}\ n}$ are emitted for conditionals and binary operations, respectively.
@@
   \typerule{$e-\ird{ifz}-\ird{true}-\text{spec}$}{
-    \ird{S} = \rtt{(\ird{\cfstate['];\sandboxtag{};D;\memstate});\ird{S}}{\ird{\expr[_1]}},({\ird{\cfstate;\sandboxtag{};D;\memstate}};\omega;{\ird{\expr[_2]}})
+    \ird{S} = \rtt{(\ird{\cfstate['];\sandboxtag;D;\memstate});\ird{S}}{\ird{\expr[_1]}},({\ird{\cfstate;\sandboxtag;D;\memstate}};\omega;{\ird{\expr[_2]}})
   }{
-    \pstepto{\rtt{\ird{\cfstate;\sandboxtag{};D;\memstate}}{\ird{\lifz{0^{\securitytag{}}}{\expr[_1]}{\expr[_2]}}}}{\ird{S}}{\ird{\ev{Branch}\ 0;\sandboxtag{};\securitytag{}}\cdot\ird{\ev{Spec};\sandboxtag{};\securitytag{}}}
+    \pstepto{\rtt{\ird{\cfstate;\sandboxtag;D;\memstate}}{\ird{\lifz{0^{\securitytag}}{\expr[_1]}{\expr[_2]}}}}{\ird{S}}{\ird{\ev{Branch}\ 0;\sandboxtag;\securitytag}\cdot\ird{\ev{Spec};\sandboxtag;\securitytag}}
   }{e-ifz-true-spec}
@@
   }{
-    \ghoststepto{\ird{S},(\ird{\Omega};\ird{0};\ird{\expr})}{\ird{S}}{\ird{Rlb;\Omega.\sandboxtag{};\unlock}}
+    \ghoststepto{\ird{S},(\ird{\Omega};\ird{0};\ird{\expr})}{\ird{S}}{\ird{Rlb;\Omega.\sandboxtag;\unlock}}
   }{e-spec-eaten}
@@
   }{
-    \ghoststepto{\ird{S},(\ird{\Omega};\ird{n};\ird{\lbarrier})}{\ird{S},(\ird{\Omega};\ird{0};\ird{0})}{\ird{Barrier;\Omega.\sandboxtag{};\unlock}}
+    \ghoststepto{\ird{S},(\ird{\Omega};\ird{n};\ird{\lbarrier})}{\ird{S},(\ird{\Omega};\ird{0};\ird{0})}{\ird{Barrier;\Omega.\sandboxtag;\unlock}}
   }{e-spec-barrier}
@@
 
-The base-event relation above scales to full events by ensuring the missing annotations ($\obj{\comp;\securitytag{}}$ and $\ird{\comp;\securitytag{}}$) are the same.
+The base-event relation above scales to full events by ensuring the missing annotations ($\obj{\comp;\securitytag}$ and $\ird{\comp;\securitytag}$) are the same.
 

### 2024-05-28 csf2/main.tex

--- 2024-05-28 csf2/main.tex
@@
 The language also adds user annotations $\obj{\securitytag{}}$ for the secrecy of variables, which can be either private (high secrecy) $\obj{\lock}$ or public (low secrecy) $\obj{\unlock}$.
-Security tags $\obj{\securitytag{}}$ are arranged in the usual secrecy lattice~\MP{NI cits}, where $\obj{\lock}\sqsubseteq\obj{\securitytag{}}$ and $\obj{\securitytag{}}\sqsubseteq\obj{\unlock}$.
+Security tags $\obj{\securitytag{}}$ are arranged in the usual secrecy lattice~\cite{zdancewicphd}, where $\obj{\lock}\sqsubseteq\obj{\securitytag{}}$ and $\obj{\securitytag{}}\sqsubseteq\obj{\unlock}$.
 

### 2024-05-28 csf2/main.tex
replace mssscct with specms

@@
 \gls*{cct} is a security property that ensures sensitive data is not leaked via timing side-channels~\cite{kocher1996timing} and \gls*{ss} is a security property that enforces the same but under a speculative semantics~\cite{guarnieri2018spectector,fabian2022automatic} that captures speculative execution attacks such as spectre~\cite{kocher2019spectre}.
-Together, \gls*{sms}, \gls*{tms}, \gls*{cct} and \gls*{ss}, yield \gls*{mssscct}, which is the gold standard of security properties for secure applications.
-Programs attaining \gls*{mssscct} do not leak sensitive data either through erroneous memory accesses, nor through timing side-channels, even under speculative execution.
+Together, \gls*{sms}, \gls*{tms}, \gls*{cct} and \gls*{ss}, yield \gls*{specms}, which is the gold standard of security properties for secure applications.
+Programs attaining \gls*{specms} do not leak sensitive data either through erroneous memory accesses, nor through timing side-channels, even under speculative execution.
 \Cref{ex:strncpy} below discusses how these security properties can be enforced by compiler passes~\cite{bond2017vale,almeida2017jasmin}, to ensure programmers need not be aware of the architectural details of where their code will run.
@@
 and it showcases the power of this framework by instantiating it on a multi-pass compilation chain.
-To this end, this paper first discusses how to compose security properties, such as \gls*{tms} and \gls*{sms} into \gls*{ms}, and then adds \gls*{cct} as well as \gls*{ss} into the mix to obtain \gls*{mssscct}.
+To this end, this paper first discusses how to compose security properties, such as \gls*{tms} and \gls*{sms} into \gls*{ms}, and then adds \gls*{cct} as well as \gls*{ss} into the mix to obtain \gls*{specms}.
 %Then, this paper defines compiler composition and formalises that given two passes that securely preserve two (possibly distinct) properties, their composition securely preserves the composition of those properties.
@@
 % , (e.g., applying \gls*{cf} or \gls*{dce}).
-Finally, this paper shows that composing these secure compiler passes into a multi-pass compilation chain results in the end-to-end (robust) preservation of \gls*{mssscct}.
+Finally, this paper shows that composing these secure compiler passes into a multi-pass compilation chain results in the end-to-end (robust) preservation of \gls*{specms}.
 Crucially, this paper derives the security of the multi-pass compiler from the composition of the security properties preserved by its individual passes.
@@
         This paper presents a case-study with five programming languages showcasing the previous contribution (\Cref{sec:casestud:defs,sec:casestud:rtp}).
-        To this end, it presents a compilation chain consisting of six passes that ultimately preserve \gls*{mssscct} by means of composing four secure passes that \emph{individually} preserve \gls*{tms}, \gls*{sms}, \gls*{scct}---an enforceable version of \gls*{cct}---, and \gls*{ss}, and two passes which are well-known optimisations \gls*{dce} and \gls*{cf},.
+        To this end, it presents a compilation chain consisting of six passes that ultimately preserve \gls*{specms} by means of composing four secure passes that \emph{individually} preserve \gls*{tms}, \gls*{sms}, \gls*{scct}---an enforceable version of \gls*{cct}---, and \gls*{ss}, and two passes which are well-known optimisations \gls*{dce} and \gls*{cf},.
         %Furthermore, the chain includes two optimisation passes: one performs \gls*{dce} and the other \gls*{cf}.
@@
 
-    \draw[thick,dotted,->] (S) to ($(S) - (0.0,2.5)$) to node[align=center,pos=.65] {\gls*{mssscct}\\ \Cref{thm:ccall:rtp:mssscct}} ($(Z) - (0.0,2.5)$) to (Z);
+    \draw[thick,dotted,->] (S) to ($(S) - (0.0,2.5)$) to node[align=center,pos=.65] {\gls*{specms}\\ \Cref{thm:ccall:rtp:specms}} ($(Z) - (0.0,2.5)$) to (Z);
 
@@
 %    \draw[thick,dotted,->] (S) to[out=-45,in=198,sloped] node[align=center] {\gls*{ms}+\gls*{scct}\\ \Cref{thm:ccall:rtp:msscct}} (O);
-    % \draw[thick,dotted,->] (S) to[out=-40,in=210,sloped] node[align=center,pos=.8] {\gls*{mssscct}\\ \Cref{thm:ccall:rtp:mssscct}} (Z);
+    % \draw[thick,dotted,->] (S) to[out=-40,in=210,sloped] node[align=center,pos=.8] {\gls*{specms}\\ \Cref{thm:ccall:rtp:specms}} (Z);
 
@@
   \vspace{-1em}
-  \caption{Visualisation of the optimising compilation pipeline that preserves \gls*{mssscct}. %
+  \caption{Visualisation of the optimising compilation pipeline that preserves \gls*{specms}. %
     Vertices in the graph are the programming languages from earlier sections (\Cref{sec:casestud:defs}). %
@@
 \end{figure*}
-The section demonstrates the power of the framework (\Cref{sec:sequential}) by composing these compilers for a secure and optimising compilation chain that robustly preserves \gls*{mssscct}.\MP{ replace with specms?}
+The section demonstrates the power of the framework (\Cref{sec:sequential}) by composing these compilers for a secure and optimising compilation chain that robustly preserves \gls*{specms}.
 The first step in this chain is the compiler from $\src{L_{\tmssafe}}$ to $\trg{L}$ that robustly preserves just \gls*{tms} (\Cref{thm:cca:rtp:tms}).
@@
 Lastly, by introducing barriers immediately after branches, speculative leaks via SPECTRE-PHT are prevented when compiling $\Lscct$ to $\Lspec$.
-The final result is that the whole compilation chain robustly preserves \gls*{mssscct} (\Cref{thm:ccall:rtp:mssscct}).
+The final result is that the whole compilation chain robustly preserves \gls*{specms} (\Cref{thm:ccall:rtp:specms}).
 
@@
 
-Finally, this subsection combines all previous results into one compilation chain to get that it preserves full \gls*{mssscct}.
-Let $\ccmssscct$ be the compiler that is the composition of $\cca$, $\ccb$, $\cccf$, $\ccdce$, $\ccscct$, and $\ccspec$. 
+Finally, this subsection combines all previous results into one compilation chain to get that it preserves full \gls*{specms}.
+Let $\ccspecms$ be the compiler that is the composition of $\cca$, $\ccb$, $\cccf$, $\ccdce$, $\ccscct$, and $\ccspec$. 
 Let $\sim^{\Ltms}_{\Lspec}$ be the composition of $\sim^{\Ltms}_{\Lms}$, $\sim^{\Lms}_{\Lscct}$, and $\sim^{\Lscct}_{\Lspec}$.
@@
 
-\begin{corollary}[$\ccmssscct$ secure w.r.t. \gls*{mssscct}]\label{thm:ccall:rtp:mssscct}
+\begin{corollary}[$\ccspecms$ secure w.r.t. \gls*{specms}]\label{thm:ccall:rtp:specms}
   $\;$
@@
 % 
-Finally, this paper defines a multi-pass compiler and proves that it preserves \gls*{mssscct}.
+Finally, this paper defines a multi-pass compiler and proves that it preserves \gls*{specms}.
 Crucially, this paper derives the security of the multi-pass compiler from the composition of the security properties preserved by its individual passes, which include security-preserving as well as optimisation passes.

### 2024-05-28 csf2/main.tex
compilation + intro 7

@@
 
-This section provides some additional formal insights into the case study. 
-\Cref{subsec:formalities:props} demonstrates that the resulting projected security property using the universal image (\Cref{def:universal:img}) is faithful to the original property. 
-\Cref{subsec:compatsecpasses} discusses a caveat with \Cref{thm:rtpsim:sig}, while \Cref{subsec:seccompproofs} gives some additional background on the secure compilation proofs that do not use the compositionality framework (\Cref{sec:sequential}).
-Lastly, \Cref{subsec:formalities:maps} discusses a minor technicality related to the ``coloring'' of the investigated properties.
+This section provides additional insights for the formal development of the paper.
+First, this section demonstrates that the security property resulting from the universal image projection is faithful to the original property (\Cref{subsec:formalities:props}). 
+Then, it discusses why the order of compiler passes matters, and how does our framework help with identifying insecure compositions (\Cref{subsec:compatsecpasses}).
+Finally, this section then gives additional technical insights on the secure compilation proofs (\Cref{subsec:seccompproofs}) and it discusses how to connect each language-specific security property to the general security properties presented in \Cref{sec:compprop} (\Cref{subsec:formalities:maps}).
 

### 2024-05-28 csf2/main.tex
new theorem

@@
 % 
-The corresponding relation $\sim^{\Ltrg}_{\Lms}$ that describes this semantic effect of the compiler is defined partially below.
+The corresponding relation $\sim^{\Ltrg}_{\Lms} : \trg{\trace}\times\irl{\trace}$ that describes this semantic effect of the compiler is defined partially below.
 % 
@@
 
-\MPin{
-  trace rel eq is not defined.
-  jump.
-  why arewe defining this ?
-}
-\MPin{
-  when defining the sim, state its type (as i did before). here and later too.
-}
-Let $\src{\trace}\sim^{\Ltms}_{\Lms}\irl{\trace}:=\src{\trace}\sim^{\Ltms}_{\Ltrg}\bullet\sim^{\Ltrg}_{\Lms}\irl{\trace}$.
-
-\Cref{corr:ccab:rtp:ms} states that the composition of $\cca$ and $\ccb$ is secure w.r.t. \gls*{ms} and it follows from \Cref{thm:cca:rtp:tms,thm:ccb:rtp:sms} using \Cref{thm:rtpsim:sig}.
-
+At this point we can compose $\ccb$ with the previous compiler ($\cca$), but in order to do so, we need a trace relation from $\Ltms$ to $\Lms$.
+% 
+We can obtain this relation by composing the trace relation we just defined ($\sim^{\Ltrg}_{\Lms}$) with the one used by the previous compiler: $\sim^{\Ltms}_{\Ltrg} : \src{\trace}\times\trg{\trace}$.
+% 
+The latter has not been previously defined (nor has it been used in the related theorem) because that is just an equality relation, since the trace models of $\Ltms$ and of $\Ltrg$ are the same.
+% 
+Thus, we formally define $\sim^{\Ltms}_{\Lms}: \src{\trace} \times \irl{\trace}$ as the following composition: $\sim^{\Ltms}_{\Ltrg}\bullet\sim^{\Ltrg}_{\Lms}$.
+% 
+With this relation, \Cref{corr:ccab:rtp:ms} states that the composition of $\cca$ and $\ccb$ is secure w.r.t.
+\gls*{ms} and it follows from \Cref{thm:cca:rtp:tms,thm:ccb:rtp:sms} using \Cref{thm:rtpsim:sig}.
+% 
 \begin{corollary}[$\cca\circ\ccb$ secure w.r.t. \gls*{ms}]\label{corr:ccab:rtp:ms}
@@
 \end{corollary}
-
-\Cref{thm:rtpsim:sig} has another precondition besides \Cref{thm:ccb:rtp:sms,thm:cca:rtp:tms}: $\sim^{\Ltms}_{\Ltrg}$ needs to be well-formed with respect to $\smssafe$.
-This fact is easily established, since $\sim^{\Ltms}_{\Ltrg}$ is an equality.
-
+% 
+This proof requires another precondition besides \Cref{thm:ccb:rtp:sms,thm:cca:rtp:tms}: $\sim^{\Ltms}_{\Ltrg}$ needs to be well-formed with respect to $\smssafe$.
+This follows trivially since $\sim^{\Ltms}_{\Ltrg}$ is an equality. 
+% 
 \begin{lemma}[$\sim^{\Ltms}_{\Ltrg}$ well-formed w.r.t. $\smssafe$]\label{lem:wf:ltmsltrg}
@@
 This section provides additional insights for the formal development of the paper.
-First, this section demonstrates that the security property resulting from the universal image projection is faithful to the original property (\Cref{subsec:formalities:props}). 
-Then, it discusses why the order of compiler passes matters, and how does our framework help with identifying insecure compositions (\Cref{subsec:compatsecpasses}).
-Finally, this section then gives additional technical insights on the secure compilation proofs (\Cref{subsec:seccompproofs}) and it discusses how to connect each language-specific security property to the general security properties presented in \Cref{sec:compprop} (\Cref{subsec:formalities:maps}).
-
-\subsection{Security Properties}\label{subsec:formalities:props}
-
-Compiler designers need to be careful that their implementation preserves the original source language semantics. 
-Similar to this, it is important to consider the cross-language relation, which gives a semantic description of the compiler in terms of traces, and check that the original target (resp.
-source) property is properly translated using the universal (resp.
-existential) image.
-The following discusses this for relations presented in \Cref{sec:casestud:rtp}.
+First, this section discusses how to connect each language-specific security property to the general security properties presented in \Cref{sec:compprop} (\Cref{subsec:formalities:maps}), then it demonstrates that the security property resulting from the universal image projection is faithful to the original property (\Cref{subsec:formalities:props}).
+Finally, this section discusses why the order of compiler passes matters, and how does our framework help with identifying insecure compositions (\Cref{subsec:compatsecpasses}), and it gives additional technical insights on the secure compilation proofs (\Cref{subsec:seccompproofs}).
+
+\subsection{From Language Traces to General Ones}\label{subsec:formalities:maps}
+
+The previous theorems talk about preserving properties expressed in the trace model of the languages of each compiler.
+However, these trace models are not the same trace model we used to specify the properties of \Cref{sec:compprop} (which we denote with \ev{L}), and which serves as the ``ground truth'' of the meaning of a certain property.
+% 
+To bridge this gap, the formal development requires specifying additional trace relations, from each of the language trace models to the \ev{L} trace model, that, for example, relate $\src{Get\ \loc\ n}$ and $\src{Set\ \loc\ n}$ to $\ev{Use\ \loc\ n}$.
+% 
+One key insight of these relations is that they all omit context-made actions for two reasons: (1) contexts (which are universally quantified) can trivially invalidate any property and (2) we are interested in the component upholding the properties.
+%
+
+ % disregarded a glaring but minor technical issue: take $\tmssafe$ (\Cref{def:trace:tmsdef}) as an example.
+% It contains traces as specified in \Cref{subsec:basic:memsafety:tracemodel}.
+% However, these are different from $\Ltms$ traces (\Cref{subsec:ltms}) which means that \Cref{thm:wt:tms} is ``ill-typed''.
+% Similarly for the other theorems.
+% In the interest of reducing visual baggage, the notation left out yet another cross-language trace relation, translating language-specific traces into these non-language specific traces. 
+% In doing so, context-level actions are dropped and, e.g., both $\src{Get\ \loc\ n}$ and $\src{Set\ \loc\ n}$ are translated into $\ev{Use\ \loc\ n}$.
+
+\subsection{Security Properties and Their Meaning}\label{subsec:formalities:props}
+Each of the presented compilers use a cross-language trace relation, which is also used to translate the property from one language to the other one (via the universal image).
+% 
+Whenever such translations occur, it is important to check whether there has been a change of meaning.
+% 
+That is, we could be using a trace relation that translates a property in a language to a different one in another language.
+% 
+To ensure such a bad translation of properties does not occur, \Cref{thm:prop-rel-corr} states (in a general fashion) that that each security property is properly translated using the universal image according to the cross-language trace relation used by the compiler.
+
+\begin{theorem}[Properties Relation Correctness]\label{thm:prop-rel-corr}
+  \begin{align*}
+    \forall& \pi \in \{\tmssafe,\smssafe,\scctsafe,\sssafe\}, 
+    \\
+    \text{ for each }& \text{pair of languages } \src{L} \text{ and } \trg{L} \text{ used by the compilers},
+    \\
+    \text{ if }& \sigma_{\sim^{\trg{L}}_{\ev{L}}}(\pi) \sim^{\trg{L}}_{\ev{L}} \pi
+    \text{ and } \sigma_{\sim^{\src{L}}_{\trg{L}}}(\trg{\pi}) \sim^{\src{L}}_{\trg{L}} \sigma_{\sim^{\trg{L}}_{\ev{L}}}(\pi)
+    \\
+    \text{ then }& \sigma_{\sim^{\src{L}}_{\trg{L}}}(\trg{\pi}) \sim^{\src{L}}_{\ev{L}} \pi
+  \end{align*}
+\end{theorem}
+The complexity of this theorem is that the relation in the conclusion cannot be obtained by composing the two relations in the premises.
+% 
+What happens in practice is that programmers are given the relations from each language to \ev{L} and they need to build a compiler that induces a relation $\sim^{\src{L}}_{\trg{L}}$ that respects this theorem.
+% 
+% We now informally argue why this theorem holds in all of its specialised cases.
 
@@
 
-Now, consider all $\src{\trace}$ with $\src{\trace}\sim^{\Ltms}_{\Ltrg}\trg{\trace}$, which is like equality\footnote{Both $\sim^{\Ltms}_{\Ltrg}$ and $\sim^{\Ltrg}_{\Lms}$ filter context actions.}.
-
-Therefore, the resulting projected property should faithfully model $\mssafe$, which translates to both $\scctsafe$ and $\sssafe$, since these properties trivially hold for $\Ltms$, $\Ltrg$, and $\Lms$ traces.
+Recall that $\sim^{\Ltms}_{\Ltrg}$ is an equality, so if $\src{\trace}\sim^{\Ltms}_{\Ltrg}\trg{\trace}$, we know that $\src{\trace}$ and $\trg{\trace}$ are identical.
+% \footnote{Both $\sim^{\Ltms}_{\Ltrg}$ and $\sim^{\Ltrg}_{\Lms}$ filter context actions.}.
+
+Now, starting from $\mssafe$ expressed in $\Lms$, we check that its translation in $\Ltms$ is still $\mssafe$.
+% 
+We can see that the property resulting from the projection of $\Lms$ traces back to $\Ltms$ faithfully models $\mssafe$, which translates to both $\scctsafe$ and $\sssafe$, 
+\MPin{ equiv  +  other rel only adds traces that fail }
+since these properties trivially hold for $\Ltms$, $\Ltrg$, and $\Lms$ traces.
 
 \paragraph{$\sigma_{\sim^{\Lms}_{\Lscct}}\left(\mssafe\cap\scctsafe\cap\sssafe\right)$}
+
 Consider a trace $\obj{\trace}\in\scctsafe$ and all $\irl{\trace}$ such that $\irl{\trace}\sim^{\Lms}_{\Lscct}\obj{\trace}$. 
@@
 Since these optimisations have no significant semantic effect and source- and target-language are equal, the backtranslation amounts to an identity translation.
-
-\subsection{Mapping Language-Specific to Property-Specific Traces}\label{subsec:formalities:maps}
-
-Previous sections disregarded a glaring but minor technical issue: take $\tmssafe$ (\Cref{def:trace:tmsdef}) as an example.
-It contains traces as specified in \Cref{subsec:basic:memsafety:tracemodel}.
-However, these are different from $\Ltms$ traces (\Cref{subsec:ltms}) which means that \Cref{thm:wt:tms} is ``ill-typed''.
-Similarly for the other theorems.
-In the interest of reducing visual baggage, the notation left out yet another cross-language trace relation, translating language-specific traces into these non-language specific traces. 
-In doing so, context-level actions are dropped and, e.g., both $\src{Get\ \loc\ n}$ and $\src{Set\ \loc\ n}$ are translated into $\ev{Use\ \loc\ n}$.
 
@@
 
-% \newpage
+\newpage
 

### 2024-05-28 csf2/main.tex
7

@@
 To this end, considering a source trace $\src{\trace}$ and a target trace $\trg{\trace}$, there is a relation ($\sim$) describing the effect of a corresponding compiler (see \Cref{sec:casestud:rtp}). 
-This relation induces the following two projection functions: (1) the \emph{existential image} $\tau_\sim\left(\src{\pi}\right)$ and (2) the \emph{universal image} $\sigma_\sim\left(\trg{\pi}\right)$.
+This relation induces the following two projection functions~\cite{abate2021extacc}: (1) the \emph{existential image} $\tau_\sim\left(\src{\pi}\right)$ and (2) the \emph{universal image} $\sigma_\sim\left(\trg{\pi}\right)$.
 These projections map source-level (resp. target-level) properties to target-level (resp. source-level) properties in a way that identifies the ``same'' property across languages. 
-The case study of this paper exclusively uses the universal image, since some considered properties, such as \gls*{ss}, are not definable in a higher-level language that, e.g., does not model speculation.
-For self-containment, the universal image is defined as follows~\cite{abate2021extacc}:
-
-\begin{definition}[Universal Image]\label{def:universal:img}\label{def:sigma}
+The case study of this paper uses the universal image, since some considered properties, such as \gls*{ss}, are not definable in a higher-level language that, e.g., does not model speculation.
+% For self-containment, the universal image is defined as follows~\cite{abate2021extacc}:
+
+\begin{definition}[Universal Image]
+% [Existential and Universal Image]
+\label{def:universal:img}\label{def:sigma}
+  % \[
+  %   \tau_\sim\left(\src{\pi}\right) := 
+  %     \left\{ 
+  %       \trg{\trace} \mid \exists \src{\trace}\ldotp \src{\trace}\sim\trg{\trace}, \text{ and } \src{\trace}\in\src{\pi} 
+  %     \right\}
+  % \]
   \[ 
-    \sigma_\sim\left(\trg{\pi}\right) := \left\{ \src{\trace} \mid \forall \trg{\trace}, \text{if }\src{\trace}\sim\trg{\trace}, \text{ then } \trg{\trace}\in\trg{\pi} \right\}
+    \sigma_\sim\left(\trg{\pi}\right) := 
+      \left\{ 
+        \src{\trace} \mid \forall \trg{\trace}\ldotp \text{if }\src{\trace}\sim\trg{\trace}, \text{ then } \trg{\trace}\in\trg{\pi} 
+      \right\}
   \]
@@
 
-\subsubsection{\gls*{ms}, Strict Cryptographic Constant Time}
+\subsubsection{\gls*{ms}, Strict Cryptographic Constant Time}\label{sec:msscct-rel}
 
@@
 \[
-  \typerulenolabel{mscctrel:drop:tag}{}{\preevent[_{\mssafe}]\sim_{\ctsafe}\preevent[_{\mssafe}];\securitytag{}}
+  \typerulenolabel{mscctrel:drop:tag}{}{\preevent[_{\mssafe}]\sim_{\ctsafe}\preevent[_{\mssafe}];\unlock{}}
   \typerulenolabel{mscctrel:drop:crash}{}{\lightning\sim_{\ctsafe}\lightning}
@@
 
-Essentially, the only work $\sim_{\ctsafe}$ has to do, is to ignore both the new $\ev{Branch\ n}$ and $\ev{Binop\ n}$ base-events as well as the security tag, since they are irrelevant to \gls*{ms}.
-Put differently, traces of \gls*{ms} trivially satisfy \gls*{scct}.
+Essentially, $\sim_{\ctsafe}$ ignores both the new $\ev{Branch\ n}$ and $\ev{Binop\ n}$ base-events as it relates security-insensitive actions ($\unlock$) to their equivalent counterparts.
+Thus, traces of \gls*{ms} trivially satisfy \gls*{scct}.
 
@@
 
-\subsubsection{Speculation Memory Safety}
+\subsubsection{Speculation Memory Safety}\label{sec:spec-ms-rel}
 
@@
 \]
-\[
-  \typerulenolabel{xrel:sms:panic}{\trg{n}\text{ out of bounds}}{\trg{Get\ \loc\ n;\comp}\sim^{\Ltrg}_{\Lms}\irl{\lightning}}
-\]
+% \[
+%   \typerulenolabel{xrel:sms:panic}{\trg{n}\text{ out of bounds}}{\trg{Get\ \loc\ n;\comp}\sim^{\Ltrg}_{\Lms}\irl{\lightning}}
+% \]
 }
@@
 
-This section provides additional insights for the formal development of the paper.
-First, this section discusses how to connect each language-specific security property to the general security properties presented in \Cref{sec:compprop} (\Cref{subsec:formalities:maps}), then it demonstrates that the security property resulting from the universal image projection is faithful to the original property (\Cref{subsec:formalities:props}).
+This section first discusses how to connect each language-specific security property to the general security properties presented in \Cref{sec:compprop} (\Cref{subsec:formalities:maps}), then it demonstrates that the security property resulting from the universal image projection is faithful to the original property (\Cref{subsec:formalities:props}).
 Finally, this section discusses why the order of compiler passes matters, and how does our framework help with identifying insecure compositions (\Cref{subsec:compatsecpasses}), and it gives additional technical insights on the secure compilation proofs (\Cref{subsec:seccompproofs}).
@@
 The previous theorems talk about preserving properties expressed in the trace model of the languages of each compiler.
-However, these trace models are not the same trace model we used to specify the properties of \Cref{sec:compprop} (which we denote with \ev{L}), and which serves as the ``ground truth'' of the meaning of a certain property.
-% 
-To bridge this gap, the formal development requires specifying additional trace relations, from each of the language trace models to the \ev{L} trace model, that, for example, relate $\src{Get\ \loc\ n}$ and $\src{Set\ \loc\ n}$ to $\ev{Use\ \loc\ n}$.
+However, these trace models are not the same trace model we used to specify the properties of \Cref{sec:compprop} (indicated with \ev{L}), which serves as the ``ground truth'' for the meaning of our properties.
+% 
+To bridge this gap, the formal development requires specifying additional trace relations, from each of the language trace models to the \ev{L} one, that, for example, relate $\src{Get\ \loc\ n}$ and $\src{Set\ \loc\ n}$ to $\ev{Use\ \loc\ n}$ (and that induce a related universal image that we use in \Cref{subsec:formalities:props}).
 % 
@@
     \\
-    \text{ if }& \sigma_{\sim^{\trg{L}}_{\ev{L}}}(\pi) \sim^{\trg{L}}_{\ev{L}} \pi
-    \text{ and } \sigma_{\sim^{\src{L}}_{\trg{L}}}(\trg{\pi}) \sim^{\src{L}}_{\trg{L}} \sigma_{\sim^{\trg{L}}_{\ev{L}}}(\pi)
+    \text{ if }& 
+    % \tau_{\sim^{\trg{L}}_{\ev{L}}}(\pi) \sim^{\trg{L}}_{\ev{L}} \pi
+    \sigma_{\sim^{\trg{L}}_{\ev{L}}}(\pi) \sim^{\trg{L}}_{\ev{L}} \pi
+    \text{ and } 
+    % \sigma_{\sim^{\src{L}}_{\trg{L}}}(\tau_{\sim^{\trg{L}}_{\ev{L}}}(\pi)) \sim^{\src{L}}_{\trg{L}} \tau_{\sim^{\trg{L}}_{\ev{L}}}(\pi)
+    \sigma_{\sim^{\src{L}}_{\trg{L}}}(\sigma_{\sim^{\trg{L}}_{\ev{L}}}(\pi)) \sim^{\src{L}}_{\trg{L}} \sigma_{\sim^{\trg{L}}_{\ev{L}}}(\pi)
     \\
-    \text{ then }& \sigma_{\sim^{\src{L}}_{\trg{L}}}(\trg{\pi}) \sim^{\src{L}}_{\ev{L}} \pi
+    \text{ then }& 
+    % \tau_{\sim^{\src{L}}_{\ev{L}}}(\pi) = \sigma_{\sim^{\src{L}}_{\trg{L}}}(\tau_{\sim^{\trg{L}}_{\ev{L}}}(\pi)) 
+    \sigma_{\sim^{\src{L}}_{\trg{L}}}(\sigma_{\sim^{\trg{L}}_{\ev{L}}}(\pi)) \sim^{\src{L}}_{\ev{L}} \pi
   \end{align*}
@@
 
-\paragraph{$\sigma_{\sim^{\Ltms}_{\Lms}}\left(\mssafe\cap\scctsafe\cap\sssafe\right)$}
-Consider a trace $\irl{\trace}\in\mssafe$ and remember that $\sim^{\Ltms}_{\Lms}$ is the composition of $\sim^{\Ltms}_{\Ltrg}$ and $\sim^{\Ltrg}_{\Lms}$.
-{\em All} traces $\trg{\trace}$ with $\trg{\trace}\sim^{\Ltrg}_{\Lms}\irl{\trace}$ are almost identical to $\irl{\trace}$. 
-They differ in memory accesses and crashes, where $\irl{\lightning}$ can be related to a $\trg{\lightning}$ or a memory event, e.g., $\trg{Get\ \loc\ 2}$, for an out-of-bounds access.
-In both cases, the translated event does not represent an out-of-bounds memory access.
-
-Recall that $\sim^{\Ltms}_{\Ltrg}$ is an equality, so if $\src{\trace}\sim^{\Ltms}_{\Ltrg}\trg{\trace}$, we know that $\src{\trace}$ and $\trg{\trace}$ are identical.
+\paragraph{$\sigma_{\sim^{\Ltms}_{\Lms}}\left(\tmssafe\cap\smssafe\cap\scctsafe\cap\sssafe\right)$}
+
+For the four properties considered here, the trace models of $\Ltms$, $\Ltrg$, and $\Lms$ do not consider actions related to $\scctsafe$ and $\sssafe$, so these two properties are trivially translated correctly.
+% 
+We now discuss the remaining $\tmssafe$ and $\smssafe$ in the form of their intersection $\mssafe$.
+
+Recall that $\sim^{\Ltms}_{\Lms}$ is the composition of $\sim^{\Ltms}_{\Ltrg}$ and $\sim^{\Ltms}_{\Lms}$.
+% 
+Since $\sim^{\Ltms}_{\Ltrg}$ is an equality, this relation trivially preserves the meaning of translated properties: related traces are identical!
+
+Finally, let us consider a trace $\irl{\trace}\in\mssafe$ and understand what is that is related to via $\sim^{\Ltrg}_{\Lms}$.
+% 
+{\em All} traces $\trg{\trace}$ with $\trg{\trace}\sim^{\Ltrg}_{\Lms}\irl{\trace}$ are identical to $\irl{\trace}$ except for get and set actions, which require for in-bound accesses: this clearly respect $\mssafe$.
+%
+
+% In both cases, the translated event does not represent an out-of-bounds memory access.
+
 % \footnote{Both $\sim^{\Ltms}_{\Ltrg}$ and $\sim^{\Ltrg}_{\Lms}$ filter context actions.}.
 
-Now, starting from $\mssafe$ expressed in $\Lms$, we check that its translation in $\Ltms$ is still $\mssafe$.
-% 
-We can see that the property resulting from the projection of $\Lms$ traces back to $\Ltms$ faithfully models $\mssafe$, which translates to both $\scctsafe$ and $\sssafe$, 
-\MPin{ equiv  +  other rel only adds traces that fail }
-since these properties trivially hold for $\Ltms$, $\Ltrg$, and $\Lms$ traces.
-
-\paragraph{$\sigma_{\sim^{\Lms}_{\Lscct}}\left(\mssafe\cap\scctsafe\cap\sssafe\right)$}
-
-Consider a trace $\obj{\trace}\in\scctsafe$ and all $\irl{\trace}$ such that $\irl{\trace}\sim^{\Lms}_{\Lscct}\obj{\trace}$. 
-In turn, $\obj{\trace}$ cannot contain any leaks of secret values, so it cannot violate $\scctsafe$.
-Toggling data-independent timing mode has no impact on $\mssafe$ or $\sssafe$.
-
-\paragraph{$\sigma_{\sim^{\Lscct}_{\Lspec}}\left(\mssafe\cap\scctsafe\cap\sssafe\right)$}
-
-The argument here is analogous to the previous one, introducing $\ird{barrier}$ never violates $\sssafe$ and has no effect on $\mssafe$ or $\scctsafe$.
+% Now, starting from $\mssafe$ expressed in $\Lms$, we check that its translation in $\Ltms$ is still $\mssafe$.
+% % 
+% We can see that the property resulting from the projection of $\Lms$ traces back to $\Ltms$ faithfully models $\mssafe$, which translates to both $\scctsafe$ and $\sssafe$, 
+% \MPin{ equiv  +  other rel only adds traces that fail }
+% since these properties trivially hold for $\Ltms$, $\Ltrg$, and $\Lms$ traces.
+
+\paragraph{$\sigma_{\sim^{\Lms}_{\Lscct}}\left(\tmssafe\cap\smssafe\cap\scctsafe\cap\sssafe\right)$}
+
+As before, the trace models of $\Lscct$ and $\Lms$ cannot express $\sssafe$, so that is trivially translated correctly.
+% 
+Also, the trace model of $\Lscct$ extends the one of $\Lms$ with respect to $\tmssafe$ and $\smssafe$ events, so the translation argument regarding those two properties is the same as before.
+% 
+Thus, we need to reason about whether $\scctsafe$ is translated correctly.
+
+By definition, $\sim^{\Lms}_{\Lscct}$ only relates $\obj{\unlock}$ events, which is also the same relation induced by $\sim_{\ctsafe}$ in \Cref{sec:msscct-rel}.
+% 
+This ensures that composing the relations only relates $\unlock$ events, and thus the property is translated correctly.
+
+% Consider a trace $\obj{\trace}\in\scctsafe$ and all $\irl{\trace}$ such that $\irl{\trace}\sim^{\Lms}_{\Lscct}\obj{\trace}$. 
+% In turn, $\obj{\trace}$ cannot contain any leaks of secret values, so it cannot violate $\scctsafe$.
+% Toggling data-independent timing mode has no impact on $\mssafe$ or $\sssafe$.
+
+\paragraph{$\sigma_{\sim^{\Lscct}_{\Lspec}}\left(\tmssafe\cap\smssafe\cap\scctsafe\cap\sssafe\right)$}
+The trace model of $\Lspec$ extends the one of $\Lscct$ with respect to $\tmssafe$, $\smssafe$, and $\scctsafe$, so the translation argument for those three properties is the same as before.
+% 
+Concerning $\sssafe$, $\sim^{\Lscct}_{\Lspec}$ relates $\Lspec$ speculation traces with $\Lscct$ branches with the $\lock_{\text{NONE}}$ tag, so the property is translated correctly, according to the relation defined in \Cref{sec:spec-ms-rel}.
+ % introducing $\ird{barrier}$ never violates $\sssafe$ and has no effect on $\mssafe$ or $\scctsafe$.
 
@@
 
-\Cref{thm:rtpsim:sig} is always applicable, up to well-formedness of $\sim_1$ w.r.t. $\trg{\class[_2]}$.
-However, consider $\ccb$ and $\ccspec$.
-Even though swapping these compilers in the compilation chain is syntactically impossible, let us pretend that it is possible by just extending $\ccb$ for the constructs present in $\Lscct$ that are not present in $\Ltrg$.
-Clearly, as evidenced earlier (\Cref{subsec:formalities:props}), $\ccb\cdot\ccspec$ is an acceptable composition that yields an interesting class of security properties, i.e., $\sigma_{\sim_{\irl{ms}}\bullet\sim_{\ird{\mathghost}}}\left(\mssafe\cap\scctsafe\cap\sssafe\right)$.
-However, swapping them $\ccspec\cdot\ccb$ means that $\ccb$ inserts new branches into the code that are not necessarily protected by a speculative execution barrier! 
-It is crucial to ensure that the resulting class is meaningful, much similar to ensuring that a definition models what the formal methods expert intends to model. 
-For the $\ccspec\cdot\ccb$ case, the class is $\sigma_{\sim_{\ird{\mathghost}}\bullet\sim_{\irl{ms}}}\left(\mssafe\cap\scctsafe\cap\sssafe\right)$
-Since $\ev{Use\ \loc\ n}\sim_{\irl{ms}}\ev{Branch n}\cdot\ev{Spec}\cdots$, where $\cdots$ does not contain a $\ird{\ev{Barrier}}$ event, the resulting class of properties is only a subset of the intended one.
-Because of this, it is crucial to analyze the precise shape of the class of properties after mapping a target-level property to a source-level property\footnote{Dito for mapping a source-level property to a target-level property.}.
+% \Cref{thm:rtpsim:sig} is applicable, up to the well-formedness of $\sim_1$ w.r.t. $\trg{\class[_2]}$.
+Consider applying $\ccspec$ first and then $\ccb$ (albeit currently syntactically impossible).
+% , let us pretend that it is possible by just extending $\ccb$ for the constructs present in $\Lscct$ that are not present in $\Ltrg$.
+% Clearly, as evidenced earlier (\Cref{subsec:formalities:props}), $\ccb\cdot\ccspec$ is an acceptable composition that yields an interesting class of security properties, i.e., $\sigma_{\sim_{\irl{ms}}\bullet\sim_{\ird{\mathghost}}}\left(\mssafe\cap\scctsafe\cap\sssafe\right)$.
+In this case, $\ccb$ would insert new branches into the code that are not protected by a speculation barrier! 
+This concern is reflected in the proofs that establish that the security class resulting of the composition of the trace relations is meaningful.
+% , much similar to ensuring that a definition models what the formal methods expert intends to model. 
+For the $\ccspec\cdot\ccb$ case, the class is $\sigma_{\sim_{\ird{\mathghost}}\bullet\sim_{\irl{ms}}}\left(\mssafe\cap\scctsafe\cap\sssafe\right)$.
+Since $\ev{Use\ \loc\ n}\sim_{\irl{ms}}\irl{Branch\ n}\cdot\irl{Spec}\cdots$, where $\cdots$ does not contain a $\ird{\ev{Barrier}}$ event, the resulting class is \emph{not} the original $\sssafe$ that is intended and it would break the corresponding \Cref{thm:prop-rel-corr}.
+% Because of this, it is crucial to analyze the precise shape of the class of properties after mapping a target-level property to a source-level property\footnote{Dito for mapping a source-level property to a target-level property.}.
 
@@
 
-The composition theorems can only be used on compilers which have been proven secure in the sense that they robustly preserve a certain class of properties. 
-Traditional proof techniques apply, the usual strategy is to establish a backwards simulation between source components and their compiled target counterpart. 
-For the universally quantified target context, a technique known as backtranslation is well-established, which constructs a source-level context either by backtranslating the target-level context, amounting to a ``decompilation'', or by constructing it using the target-level trace, so that the source-context emits parts of the whole execution trace that are related to the parts emitted by the target-level context.
-Up to backtranslations for $\ccdce$ and $\cccf$, all backtranslations in the case-study are trace based, while for $\ccdce$ and $\cccf$, they are context-based. 
-Since these optimisations have no significant semantic effect and source- and target-language are equal, the backtranslation amounts to an identity translation.
+Our secure compilation proofs rely on established proof techniques called backtranslation~\cite{abate2019jour,patrignani2021rsc}, which let one construct a source context starting from either target traces (aka trace-based backtranslations) or target contexts (aka context-based backtranslations).
+These backtranslations also require setting up cross-language relations between various language elements such as expressions and program states, so we leave these details for the technical reports.
+% , the usual strategy is to establish a backwards simulation between source components and their compiled target counterpart. 
+% For the universally quantified target context, a technique known as backtranslation is well-established, which constructs a source-level context either by backtranslating the target-level context, amounting to a ``decompilation'', or by constructing it using the target-level trace, so that the source-context emits parts of the whole execution trace that are related to the parts emitted by the target-level context.
+All backtranslations in the case-study are trace based except for those required by $\ccdce$ and $\cccf$, which are context-based (and they are essentially an identity function). 
+% Since these optimisations have no significant semantic effect and source- and target-language are equal, the backtranslation amounts to an identity translation.
 

### 2024-05-28 csf2/main.tex
suibmittable

@@
 % For self-containment, the universal image is defined as follows~\cite{abate2021extacc}:
-
+% 
 \begin{definition}[Universal Image]
@@
 \end{definition}
-
-Using these projection functions, the more general version of robust preservation is defined as follows~\cite{abate2021extacc}.
+\noindent
+With this projection function, we define a more general version of robust preservation as follows~\cite{abate2021extacc}.
 % For a \bul{compiler $\cc{\src{L}}{\trg{L}}$ to robustly preserve a class of source properties $\src{\class}$}, if for any \rul{property $\src{\pi}$ of that class $\src{\class}$ and programs $\src{p}$ written in $\src{L}$} where \iul{$\src{\progvar}$ robustly satisfies $\src{\pi}$}, then \oul{the compilation of $\src{\progvar}$, $\cc{\src{L}}{\trg{L}}\left(\src{p}\right)$, must robustly satisfy $\tau_\sim\left(\src{\pi}\right)$}.
-
+% 
 % \begin{definition}[Robust Preservation with $\tau_\sim$]\label{def:rtp:tau}
@@
 % 
-A \bul{compiler $\cc{\src{L}}{\trg{L}}$ robustly preserves a class of target properties $\trg{\class}$}, if for any \rul{property $\trg{\pi}$ of class $\trg{\class}$ and programs $\src{p}$}, where \iul{$\src{\progvar}$ robustly satisfies $\sigma_\sim\left(\trg{\pi}\right)$}, \oul{the compilation of $\src{\progvar}$, $\cc{\src{L}}{\trg{L}}\left(\src{p}\right)$, must robustly satisfy $\trg{\pi}$}.
+A \bul{compiler $\cc{\src{L}}{\trg{L}}$ robustly preserves a class of target properties $\trg{\class}$}, if for any \rul{property $\trg{\pi}$ of class $\trg{\class}$ and programs $\src{p}$}, where \iul{$\src{\progvar}$ robustly satisfies $\sigma_\sim\left(\trg{\pi}\right)$}, \oul{the compilation of $\src{\progvar}$, $\cc{\src{L}}{\trg{L}}\left(\src{p}\right)$, robustly satisfies $\trg{\pi}$}.
 
@@
 
-\subsection{Memory Safety Trace Model}\label{subsec:basic:memsafety:tracemodel}
+\subsection{A Trace Model for Memory Safety}\label{subsec:basic:memsafety:tracemodel}
 
@@
 
-\subsection{Memory Safety Trace Model with Constant Time}\label{subsec:scct:tracemodel}
-
-To express Constant Time, we extend the memory safety trace model (\Cref{subsec:basic:memsafety:tracemodel}) with a {\em security tag} ($\securitytag{}$) that indicates whether the involved event contains sensitive information ($\lock$) or not ($\unlock$).
+\subsection{A Trace Model for Memory Safety with Constant Time}\label{subsec:scct:tracemodel}
+
+To express Constant Time, we extend the memory safety trace model with a {\em security tag} ($\securitytag{}$) that indicates whether events contains sensitive information ($\lock$) or not ($\unlock$).
 
@@
 Essentially, $\sim_{\ctsafe}$ ignores both the new $\ev{Branch\ n}$ and $\ev{Binop\ n}$ base-events as it relates security-insensitive actions ($\unlock$) to their equivalent counterparts.
-Thus, traces of \gls*{ms} trivially satisfy \gls*{scct}.
-
+Thus, \gls*{ms} traces trivially satisfy \gls*{scct}.
+% 
 The above relation is extended point-wise to traces, skipping the empty event $\emptyevent$ on either side, and it is now possible to define \gls*{msscct} using the universal image:
@@
   $
-  \msscctsafe:=\mssafe\cap\sigma_{\sim_{\ctsafe}}\left(\mssafe\right)
+  \msscctsafe:=\mssafe\cap\sigma_{\sim_{\ctsafe}}\left(\scctsafe\right)
   $
@@
 
-\subsection{\gls*{ms}, \gls*{scct} Trace Model with Speculation}\label{subsec:msctss:tracemodel}
+\subsection{Extending the Trace Model with Speculation}\label{subsec:msctss:tracemodel}
 

### 2024-05-29 csf2/main.tex
nits

@@
 % 
-A \bul{compiler $\cc{\src{L}}{\trg{L}}$ robustly preserves a class of target properties $\trg{\class}$}, if for any \rul{property $\trg{\pi}$ of class $\trg{\class}$ and programs $\src{p}$}, where \iul{$\src{\progvar}$ robustly satisfies $\sigma_\sim\left(\trg{\pi}\right)$}, \oul{the compilation of $\src{\progvar}$, $\cc{\src{L}}{\trg{L}}\left(\src{p}\right)$, robustly satisfies $\trg{\pi}$}.
+A \bul{compiler $\cc{\src{L}}{\trg{L}}$ robustly preserves a class of target properties $\trg{\class}$}, if for any \rul{property $\trg{\pi}$ of class $\trg{\class}$ and programs $\src{p}$}, where \iul{$\src{\progvar}$ robustly satisfies $\sigma_\sim\left(\trg{\pi}\right)$}, \oul{the compilation of $\src{\progvar}$, we have that $\cc{\src{L}}{\trg{L}}\left(\src{p}\right)$ robustly satisfies $\trg{\pi}$}.
 
@@
 The language also adds user annotations $\obj{\securitytag{}}$ for the secrecy of variables, which can be either private (high secrecy) $\obj{\lock}$ or public (low secrecy) $\obj{\unlock}$.
-Security tags $\obj{\securitytag{}}$ are arranged in the usual secrecy lattice~\cite{zdancewicphd}, where $\obj{\lock}\sqsubseteq\obj{\securitytag{}}$ and $\obj{\securitytag{}}\sqsubseteq\obj{\unlock}$.
+Security tags $\obj{\securitytag{}}$ are arranged in the usual secrecy lattice~\cite{zdancewicphd}, where $\obj{\unlock}\sqsubseteq\obj{\lock}$.
+% $\obj{\lock}\sqsubseteq\obj{\securitytag{}}$ and $\obj{\securitytag{}}\sqsubseteq\obj{\unlock}$.
 
@@
 % 
-{\em All} traces $\trg{\trace}$ with $\trg{\trace}\sim^{\Ltrg}_{\Lms}\irl{\trace}$ are identical to $\irl{\trace}$ except for get and set actions, which require for in-bound accesses: this clearly respect $\mssafe$.
+{\em All} traces $\trg{\trace}$ with $\trg{\trace}\sim^{\Ltrg}_{\Lms}\irl{\trace}$ are identical to $\irl{\trace}$ except for get and set actions, which require for in-bound accesses (as stated in \Cref{subsec:cs:ms}): this clearly respect $\mssafe$.
 %

### 2024-05-29 csf2/main.tex
nit

@@
 These properties are of practical importance (as mentioned in \Cref{sec:introduction}) and also of interest in the case study presented later (\Cref{sec:casestud:defs,sec:casestud:rtp}). 
-For clarity, this section presents all of them, despite the fact that only property \gls*{scct} is novel, while the other ones are inspired by existing work.
+This section presents all of them, despite the fact that only property \gls*{scct} is novel and the others are inspired by existing work, in order to showcase all that is required for a formal proof of security for a realistic compilation toolchain.
 % , none of the properties is novel, for self-containment of the paper, their definition is necessary for the case study.
-For each of the presented properties, the technical appendix accompanying the paper defines corresponding monitors, which refine a given property and have a key role in the proofs of this paper.
+For each of the presented properties, the technical appendix defines corresponding monitors, which refine a given property and have a key role in the proofs of this paper.
 

### 2024-05-29 csf2/main.tex
more nits
last languege is italics now

@@
 Second, deallocation ($\ev{Dealloc\ \loc}$) announces that the object at location $\loc$ is freed.
-Third, an event to describe reads from and writes to memory ($\ev{Use\ \loc\ n}$).
+Third, an event to describe reads from and writes to the $n$-th memory cell from address $\loc$ ($\ev{Use\ \loc\ n}$).
 
@@
     \text{(Functions)} & \src{F} &:=& (\src{x};\src{\expr}) \\
-    \text{(Libraries)} &&& \src{\library_{\ctx}} : \src{Vars} \to \src{F} \hspace{2em}
-    \src{\library_{\comp}} : \src{Vars} \to \src{F} \\
-    \text{(Heaps)} & \src{H} &:=& \src{\hole{\cdot}} \mid \src{n},\src{H}\hspace{0.5cm}    \text{(Pointer Maps)}\ \ \ \src{\Delta} \\
+    \text{(Libraries)} &&&\hspace{-3.8em} \src{\library_{\ctx}},\src{\library_{\comp}} : \src{Vars} \to \src{F} 
+    % \hspace{2em}
+    %  : \src{Vars} \to \src{F} 
+     \\
+    \text{(Heaps)} & \src{H} &:=& \src{\hole{\cdot}} \mid \src{n},\src{H}
+    % \hspace{0.5cm}    
+    \\
+    \text{(Pointer Maps)} & \src{\Delta} && \text{ omitted for simplicity}
+    \\
     \text{(Memory States)} & \src{\memstate} &:=& \left(\src{H^{\ctx}};\src{H^{\comp}};\src{\Delta}\right)\\
-  \text{(Control States)} & \src{\cfstate} & & \text{(States)}\ \ \ \src{\Omega}\ :=\ \left(\src{\cfstate};\src{\sandboxtag{}};\src{\memstate}\right)\\
+    \text{(Control States)} & \src{\cfstate} & & \text{ omitted for simplicity} 
+    \\
+    \text{(States)} &\src{\Omega} &:=& \left(\src{\cfstate};\src{\sandboxtag{}};\src{\memstate}\right)\\
   \end{array}
@@
 \begin{figure}[!htb]
-	\vspace{-1em}
+	\vspace{-1em}\small
   $$
@@
   }{
-    \pstepto{\rtt{\ird{\cfstate;\sandboxtag{};D;\memstate}}{\ird{\lifz{0^{\securitytag{}}}{\expr[_1]}{\expr[_2]}}}}{\ird{S}}{\ird{\ev{Branch}\ 0;\sandboxtag{};\securitytag{}}\cdot\ird{\ev{Spec};\sandboxtag{};\securitytag{}}}
+    \pstepto{\rtt{\ird{\cfstate;\sandboxtag{};D;\memstate}}{\ird{\lifz{0^{\securitytag{}}}{\expr[_1]}{\expr[_2]}}}}{\ird{S}}{\ird{{Branch}\ 0;\sandboxtag{};\securitytag{}}\cdot\ird{{Spec};\sandboxtag{};\securitytag{}}}
   }{e-ifz-true-spec}
@@
 
-The trace model is extended with $\ird{\ev{Spec}}$, $\ird{\ev{Rlb}}$, and $\ird{\ev{Barrier}}$ events that signal the respective operational action. 
-A $\ird{\lbarrier}$ prevents any execution whatsoever besides $\ird{\ev{Rlb}}$ when run in speculative execution mode and does nothing when run in normal mode, e.g., \Cref{tr:e-spec-barrier}.%
+The trace model is extended with $\ird{{Spec}}$, $\ird{{Rlb}}$, and $\ird{{Barrier}}$ events that signal the respective operational action. 
+A $\ird{\lbarrier}$ prevents any execution whatsoever besides $\ird{{Rlb}}$ when run in speculative execution mode and does nothing when run in normal mode, e.g., \Cref{tr:e-spec-barrier}.%
 \footnote{In the rule, the notation $\cfstate' = \cfstate[.\text{win} = 0]$ means that $\cfstate'$ is a copy of $\cfstate$ up to field $\text{win}$, which is set to $0$.}
@@
 For the $\ccspec\cdot\ccb$ case, the class is $\sigma_{\sim_{\ird{\mathghost}}\bullet\sim_{\irl{ms}}}\left(\mssafe\cap\scctsafe\cap\sssafe\right)$.
-Since $\ev{Use\ \loc\ n}\sim_{\irl{ms}}\irl{Branch\ n}\cdot\irl{Spec}\cdots$, where $\cdots$ does not contain a $\ird{\ev{Barrier}}$ event, the resulting class is \emph{not} the original $\sssafe$ that is intended and it would break the corresponding \Cref{thm:prop-rel-corr}.
+Since $\ev{Use\ \loc\ n}\sim_{\irl{ms}}\irl{Branch\ n}\cdot\irl{Spec}\cdots$, where $\cdots$ does not contain a $\ev{{Barrier}}$ event, the resulting class is \emph{not} the original $\sssafe$ that is intended and it would break the corresponding \Cref{thm:prop-rel-corr}.
 % Because of this, it is crucial to analyze the precise shape of the class of properties after mapping a target-level property to a source-level property\footnote{Dito for mapping a source-level property to a target-level property.}.

### 2024-05-29 csf2/main.tex
nit

@@
   		% 
-  		Additionally this paper analyzes each pass in terms of their compatibility with each other (\Cref{sec:formalities}), proving how does each pass fulfill the conditions that make their composition meaningful from a security standpoint.
+  		Additionally this paper analyses each pass in terms of their compatibility with each other (\Cref{sec:formalities}), proving how does each pass fulfill the conditions that make their composition meaningful from a security standpoint.
         % In doing so, the paper provides formal insights into some technicalities of the concrete proofs necessary as technical setup to reap the benefits of the presented compositionality framework.
@@
 
-To analyze the security of compilation passes and their interaction within a compilation pipeline, we rely on a few key notions: the definition of a trace relation, and the definition of when is a trace relation well-formed with respect to a class.
+To analyse the security of compilation passes and their interaction within a compilation pipeline, we rely on a few key notions: the definition of a trace relation, and the definition of when is a trace relation well-formed with respect to a class.
 %
@@
 \footnote{In the rule, the notation $\cfstate' = \cfstate[.\text{win} = 0]$ means that $\cfstate'$ is a copy of $\cfstate$ up to field $\text{win}$, which is set to $0$.}
-Lastly, $\ird{Rlb}$ is emitted when the speculation window is zero and the operationl state is rolled back (see \Cref{tr:e-spec-eaten}).
+Lastly, $\ird{Rlb}$ is emitted when the speculation window is zero and the operational state is rolled back (see \Cref{tr:e-spec-eaten}).
 
@@
 Since $\ev{Use\ \loc\ n}\sim_{\irl{ms}}\irl{Branch\ n}\cdot\irl{Spec}\cdots$, where $\cdots$ does not contain a $\ev{{Barrier}}$ event, the resulting class is \emph{not} the original $\sssafe$ that is intended and it would break the corresponding \Cref{thm:prop-rel-corr}.
-% Because of this, it is crucial to analyze the precise shape of the class of properties after mapping a target-level property to a source-level property\footnote{Dito for mapping a source-level property to a target-level property.}.
+% Because of this, it is crucial to analyse the precise shape of the class of properties after mapping a target-level property to a source-level property\footnote{Dito for mapping a source-level property to a target-level property.}.
 
@@
 
-The robust preservation of properties as a compiler-level criterion has been analyzed extensively~\cite{abate2019jour,patrignani2021rsc,abate2021extacc,patrignani2019survey} and thus we build on that framework.
+The robust preservation of properties as a compiler-level criterion has been analysed extensively~\cite{abate2019jour,patrignani2021rsc,abate2021extacc,patrignani2019survey} and thus we build on that framework.
 No existing work is concerned with composing robustly safe compilers.
@@
 
-To extend the languages in this paper with a less restricted form of pointer arithmetic, the region coloring memory safety monitor presented in earlier work~\cite{michael2023mswasm} can be used.
+To extend the languages in this paper with a less restricted form of pointer arithmetic, the region colouring memory safety monitor presented in earlier work~\cite{michael2023mswasm} can be used.
 The work presenting this monitor provides an approach for the robust preservation of memory safety compiling from C to WASM.

### 2024-05-29 csf2/main.tex
fix typo in def of mix
remove last paragraph

@@
    %\partialeval{\irl{\lbinop{\expr[_{1}]}{\expr[_{2}]}}}{\irl{\substlist}} & = \lbinop[\irl]{\partialeval{\irl{\expr[_{1}]}}{\irl{\substlist}}}{\partialeval{\irl{\expr[_{2}]}}{\irl{\substlist}}} \\
-   \partialeval{\irl{\llet{x}{\valueexpr}{\expr}}}{\irl{\substlist}} & = \partialeval{\irl{\expr}}{\irl{\subst{x}{\valueexpr}\cdot\substlist}} 
+   \partialeval{\irl{\llet{x}{n}{\expr}}}{\irl{\substlist}} & = \partialeval{\irl{\expr}}{\irl{\subst{n}{x}\cdot\substlist}} 
   % \\
@@
 
-\paragraph*{Speculation Safety Mechanisms}\label{subsec:relw:ssmechs}
-
-With the advent of speculative execution in processors, it is possible to exploit this unusual paradigm leak secrets.
-This work has taken inspiration in the sematics of $\Lspec$ and in \Cref{def:trace:ss} from recent formalisations~\cite{guarnieri2018spectector,fabian2022automatic}.
+%\paragraph*{Speculation Safety Mechanisms}\label{subsec:relw:ssmechs}
+
+%With the advent of speculative execution in processors, it is possible to exploit this unusual paradigm leak secrets.
+%This work has taken inspiration in the sematics of $\Lspec$ and in \Cref{def:trace:ss} from recent formalisations~\cite{guarnieri2018spectector,fabian2022automatic}.
 % \MPin{
@@
 
-\bibliographystyle{IEEEtran}
+\bibliographystyle{IEEEtranS}
 \bibliography{main}

### 2024-06-13 csf2/main.tex
small spec section

@@
 
-%\paragraph*{Speculation Safety Mechanisms}\label{subsec:relw:ssmechs}
-
-%With the advent of speculative execution in processors, it is possible to exploit this unusual paradigm leak secrets.
+\paragraph*{Speculation Safety Mechanisms}\label{subsec:relw:ssmechs}
+
+This paper uses a taint-tracking mechanism inspired by existing work~\cite{guarnieri2018spectector,fabian2022automatic}.
+These taints are used to express absence of any speculative leaks in \gls*{ss}~\cite{guarnieri2018spectector}. 
+The semantics of $\Lspec$ hardcodes the kind of speculative leaks to just SPECTRE-PHT~\cite{kocher2019spectre}, but future work could use semantics composition~\cite{fabian2022automatic} to support more variants.
+Note that our framework composes compilers and not semantics.
+
 %This work has taken inspiration in the sematics of $\Lspec$ and in \Cref{def:trace:ss} from recent formalisations~\cite{guarnieri2018spectector,fabian2022automatic}.

### 2024-09-05 tech/main.tex
Add typechecks
.

--- 2023-08-05 tech/main.tex
@@
 \newcommand\bmmax{0}
-\documentclass[utf8,acmsmall,review,screen,dvipsnames]{acmart}
-
-\usepackage[colorinlistoftodos]{todonotes}
-\usepackage[inference]{semantic}
-\usepackage{fontawesome5}
-\usepackage{listofitems}
-\usepackage{glossaries}
-\usepackage{cleveref}
-\usepackage{stmaryrd}
-\usepackage{marvosym}
-\usepackage{listings}
-\usepackage{xspace}
-\usepackage{xfrac}
-\usepackage{tikz}
-\usepackage{soul}
-\usepackage{bm}
-
-\setul{0.95ex}{0.3ex}
+\documentclass[a4paper,12pt]{article}
+\usepackage[left=1.5cm,top=1.5cm,right=1.5cm,bottom=1.5cm]{geometry}
+\usepackage{parskip}
 
@@
 \SetMathAlphabet{\mathsf}{bold}{OT1}{LibertinusSans-LF}{bx}{n}
-
 \DeclareMathAlphabet{\mathtt}{OT1}{lmtt}{m}{n}
@@
 
-%% NOTE that a single column version may be required for 
-%% submission and peer review. This can be done by changing
-%% the \doucmentclass[...]{acmart} in this template to 
-%% \documentclass[manuscript,screen]{acmart}
-%% 
-%% To ensure 100% compatibility, please check the white list of
-%% approved LaTeX packages to be used with the Master Article Template at
-%% https://www.acm.org/publications/taps/whitelist-of-latex-packages 
-%% before creating your document. The white list page provides 
-%% information on how to submit additional LaTeX packages for 
-%% review and adoption.
-%% Fonts used in the template cannot be substituted; margin 
-%% adjustments are not allowed.
-%%
-%%
-%% \BibTeX command to typeset BibTeX logo in the docs
-\AtBeginDocument{%
-  \providecommand\BibTeX{{%
-    \normalfont B\kern-0.5em{\scshape i\kern-0.25em b}\kern-0.8em\TeX}}}
-
-\citestyle{acmauthoryear}
-
+\title{Secure Composition of Robust and Optimising Compilers}
 \begin{document}
 
-\title{Secure Composition of Robust and Optimising Compilers}
+\maketitle
+
 \begin{abstract}
+% context
+To ensure that secure applications do not leak their secrets, they are required to uphold several security properties such as spatial and temporal memory safety, cryptographic constant time, as well as speculative safety.
+% need
+Existing work shows how to enforce these properties individually, in an architecture-independent way, by using secure compiler passes that each focus on an individual property.
+% task
+Unfortunately, given two secure compiler passes that each preserve a possibly different security property, it is unclear what kind of security property is preserved by the composition of those secure compiler passes.
+%there is no way to tell what kind of security property will the composition of those secure compilers preserve.
+
+% object
+This paper is the first to study what security properties are preserved across the composition of different secure compiler passes.
+% findings
+Starting from a general theory of property composition for security-relevant properties (such as the aforementioned ones), this paper formalises a theory of composition of secure compilers.
+Then, it showcases this theory on a secure multi-pass compiler that preserves the aforementioned security-relevant properties.
+% conclusion
+Crucially, this paper derives the security of the multi-pass compiler from the composition of the security properties preserved by its individual passes, which include security-preserving as well as optimisation passes.
+% 
+From an engineering perspective, this is the desirable approach to building secure compilers.
 \begin{center}\small\it
-	{This paper uses syntax highlighting accessible to both colourblind and black \& white readers.
-	% ~\citep{patrignani2020use}.
-% 	% Specifically, it makes use of a $\src{blue}$, $\src{sans\text{-}serif}$ font for a $\src{source}$,
-% 	% a $\trg{red}$, $\trg{bold}$ font for an $\trg{intermediate}$,
-% 	% and a $\obj{green}$, $\obj{teletype}$ font for a $\obj{target}$ language.
-	}
+	{This paper uses syntax highlighting accessible to both colourblind and black \& white readers.}
 \end{center}
@@
 
-\maketitle
-
 \appendix
 
-\section{General Trace Model}\label{ap:def:gtm}
+\section{Background}
 
-\begin{gather*}
-  \begin{aligned}
-  \mi{(Security\ Tag)}~\sigma\bnfdef&\ \lock \mid \unlock\hspace{0.5cm}
-  \mi{(Control\ Tag)}~\sandboxtag\bnfdef \ctx\mid\comp\hspace{0.5cm}
-  \mi{(Event)}~\event\bnfdef\ \emptyevent \mid \lightning \mid \event[_{b}];\sandboxtag;\sigma \\
-  \mi{(Pre\text{-}event)}&~\event[_{b}]\bnfdef\ \ev{Alloc\ \loc\ n} \mid \ev{Dealloc\ \loc} \mid \ev{Use\ \loc\ n} \mid \ev{Branch\ n} \mid \ev{Binop\ n} \\
-  \end{aligned}
-\end{gather*}
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+\subsection{Satisfaction}
+\bul{A whole program $\varWholeProg$ satisfies a property $\varProperty$} iff \iul{$\varWholeProg$ yields a trace $\varTrace$} such that \oul{$\varTrace$ satisfies $\varProperty$}.
 
-Throughout the paper, sequences are indicated with an overbar (i.e., $\trace$), empty sequences with $\hole{\cdot}$, and concatenation of sequences $\trace[_{1}],\trace[_{2}]$ as $\trace[_{1}]\cdot\trace[_{2}]$.
-Prepending elements to sequences uses the same notation: $\event\cdot\trace$.
-
-\begin{definition}\label{ap:def:property}
-  A property $\pi$ is a set of traces $\trace$.
-\end{definition}
-\begin{definition}\label{ap:def:propclass}
-  A class $\class$ is a set of properties $\pi$.
-\end{definition}
-\begin{definition}\label{ap:def:lifting:properties}
-  The lifting $\lift{\pi}$ of property $\pi$ is its powerset.
+\begin{definition}{\definitionlabel[Property Satisfaction]{prop-sat}}
+    \bul{$\sat{\varWholeProg}{\varProperty}$}
+    $\isdef$
+    if \iul{$\forall\varRuntimeTerm\ \varTrace, %
+      \progstepto{\varWholeProg}{\varRuntimeTerm}{\varTrace}$
+    },
+    then \oul{$\varTrace\in\varProperty$}.
 \end{definition}
 
-\section{Properties and Monitors}\label{ap:sec:propmon}
+With this, \definitionref{prop-sat} can be extended as follows: for \bul{a component $\varComponent$ to robustly satisfy a property $\varProperty$}, take an \iul{attacker context $\varContext$ and link it with $\varComponent$}, \oul{the resulting whole program must satisfy $\varProperty$}.
 
-In the following, $\traceagree[]{\pi}{\event}{\absevent}$ and the structure of $\abstrace$ are left abstract.
-These are monitor-specific definitions, which are defined in \Cref{ap:subsec:tms,ap:subsec:sms,ap:subsec:ms,ap:subsec:scct,ap:subsec:msscct}.
-
-\begin{figure}[!ht]
-  \judgbox{\traceagree{\pi}{\trace}{\abstrace}}{,,Specification-level trace $\trace$ agrees with monitor-level trace $\abstrace$ with respect to property $\pi$.''}
-  %
-  \typerule{traceagree-empty}{}{
-    \traceagree{\pi}{\hole{\cdot}}{\hole{\cdot}}
-  }{traceagree-empty}
-  %
-  \typerule{traceagree-ign-L}{
-    \traceagree{\pi}{\trace}{\abstrace}
-  }{
-    \traceagree{\pi}{\emptyevent\cdot\trace}{\abstrace}
-  }{traceagree-ign-L}
-  %
-  \typerule{traceagree-ign-R}{
-    \traceagree{\pi}{\trace}{\abstrace}
-  }{
-    \traceagree{\pi}{\trace}{\absev{\emptyevent}\cdot\abstrace}
-  }{traceagree-ign-R}
-  %
-  \typerule{traceagree-cons}{
-    \traceagree[]{\pi}{\event}{\absevent} &
-    \traceagree{\pi}{\trace}{\abstrace}
-  }{
-    \traceagree{\pi}{\event\cdot\trace}{\absevent\cdot\abstrace}
-  }{traceagree-cons}
-  \caption{Trace-Agreement relation that equates specification-level traces with monitor-level traces.}\label{fig:stuttereq}
-\end{figure}
-
-\bul{A specification trace $\trace$ monitor-satisfies property $\pi$} iff there exists a (final) monitor state $\monitor$ and an abstract trace $\abstrace$ such that the \iul{specification trace $\trace$ agrees with abstract trace $\abstrace$} and \oul{the initial monitor can step to the (final) monitor state $\monitor$ with abstract trace $\abstrace$}.
-
-\begin{definition}[Monitor Satisfaction]\label{def:monsat}
-  \bul{$\monsafe{\trace}{\pi}$} $\isdef$ $\exists \abstrace\ \monitor$, \iul{$\traceagree{\pi}{\trace}{\abstrace}$} and \oul{$\monitorcheck{\emptyset}{\monitor}{\abstrace}$}.
+\begin{definition}{\definitionlabel[Robust Property Satisfaction]{prop-rsat}}
+  \bul{$\rsat{\varComponent}{\varProperty}$} %
+  $\isdef$ %
+  \iul{$\forall \varContext\ \varWholeProg$, %
+    if $\bopLink{\varContext}{\varComponent} = \varWholeProg$}, %
+    then \oul{$\sat{\varWholeProg}{\varProperty}$%
+  }.
 \end{definition}
 
-\subsection{Temporal Memory Safety}\label{ap:subsec:tms}
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+\subsection{Property Mappings}
 
-\begin{definition}[\glsfirst*{tms}]\label{ap:def:trace:tmsdef}
-  \begin{displaymath}
-  \tmssafe:=\left\{\trace \left| \begin{array}{rcl}
-                              \ev{Alloc\ \loc\ n;\sandboxtag;\securitytag}&\le_{\trace}&\ev{Dealloc\ \loc;\sandboxtag;\securitytag'} \\
-                              \ev{Use\ \loc\ n;\sandboxtag;\securitytag}&\le_{\trace}&\ev{Dealloc\ \loc;\sandboxtag;\securitytag'} \\
-                              \text{if }\ev{Alloc\ \loc\ n;\sandboxtag;\securitytag}\text{ in }\trace&\text{then}&\ev{Dealloc\ \loc;\sandboxtag;\securitytag'}\text{ in }\trace \\
-                              \text{at most one }\ev{Dealloc\ \loc;\sandboxtag;\securitytag}&\text{in}&\trace \\
-                              \text{at most one }\ev{Alloc\ \loc\ n;\sandboxtag;\securitytag}&\text{in}&\trace \\
-                            \end{array}\right.\right\}
-  \end{displaymath}
+\begin{definition}{\definitionlabel[Universal Image]{universal-image}}
+  $ 
+    \mapUniversal{\sim}{\trgProperty} := 
+      \lrbraces{\srcTrace \mid %
+        \forall \trgTrace\ldotp \text{if }\xrelTraces{\srcTrace}{\trgTrace}, \text{ then }%
+        \trgTrace\in\trgProperty
+      }
+  $
 \end{definition}
-Hereby, the notation $\event[_{1}]\le_{\trace}\event[_{2}]$ means that if $\event[_{1}]$ is in $\trace$ and if $\event[_{2}]$ is in $\trace$, then $\event[_{1}]$ appears before $\event[_{2}]$.
+\begin{definition}{\definitionlabel[Existential Image]{existential-image}}
+  $ 
+    \mapExistential{\sim}{\srcProperty} := 
+      \lrbraces{\trgTrace \mid %
+        \exists \srcTrace\ldotp \xrelTraces{\srcTrace}{\trgTrace} \text{ and }%
+        \srcTrace\in\srcProperty
+      }
+  $
+\end{definition}
 
-$$\;$$
-\hrule
-$$\;$$
-\begin{gather*}
-  \begin{aligned}
-    \mi{(Abstract\ Store)}~\tmsmonitor\bnfdef&\ \left\{\text{allocated}:L\times\sandboxtag,\text{freed}:L\times\sandboxtag\right\} \\
-    \mi{(Initial\ Abstract\ Store)}~\emptytmsmon:=&\ \left\{\text{allocated}:\emptyset,\text{freed}:\emptyset\right\}\\
-    \mi{(Abstract\ Events)}~\tmsevent\bnfdef&\ \absev{\emptyevent} \mid \absev{Alloc\ }\loc\ \sandboxtag \mid \absev{Dealloc\ }\loc\ \sandboxtag \mid \absev{Use\ }\loc\ \sandboxtag \mid \absterm\\
-  \end{aligned}
-\end{gather*}
+The lifting of \definitionref{existential-image} and \definitionref{universal-image} to classes of properties is standard, i.e., simply element-wise.
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+\subsection{Robust Preservation}
+
+\subsubsection{Differing Trace Models}
+
+A \bul{compiler $\ccST$ for languages $\src{S}$ and $\trg{T}$ robustly preserves a $\trg{T}$-level class $\trgClass$}, if for any \rul{property $\trgProperty$ of class $\trgClass$ and programs $\srcComponent$}, where \iul{$\srcComponent$ robustly satisfies $\mapUniversal{\sim}{\trgProperty}$}, \oul{the compilation of $\srcComponent$, we have that $\ccST\lrpars{\srcComponent}$ robustly satisfies $\trgProperty$}.
+
+\begin{definition}{\definitionlabel[Robust Preservation with $\mapUniversal{\sim}{\cdot}$]{rtp-universal}}
   \begin{center}
-  \judgbox{\monitorcheck[]{\tmsmonitor}{\tmsmonitor[']}{\tmsevent}}{,,Monitor $\tmsmonitor$ does one step to $\tmsmonitor[']$ given event $\tmsevent$.''}$\;$\\
-  %
-  \typerule{tms-uninteresting}{
-  }{
-    \monitorcheck[]{\tmsmonitor}{\tmsmonitor}{\absev{\emptyevent}}
-  }{ap:tms-uninteresting}
-  %
-  \typerule{tms-abort}{
-  }{
-    \monitorcheck[]{\tmsmonitor}{\tmsmonitor}{\absterm}
-  }{ap:tms-abort}
-  %
-  \typerule{tms-use}{
-    (\loc;\sandboxtag)\in\tmsmonitor.\text{allocated} &
-    (\loc;\sandboxtag)\notin\tmsmonitor.\text{freed}
-  }{
-    \monitorcheck[]{\tmsmonitor}{\tmsmonitor}{\absev{Use\ }\loc\ \sandboxtag}
-  }{ap:tms-use}
-  %
-  \typerule{tms-alloc}{
-    (\loc;\sandboxtag)\notin\tmsmonitor.\text{allocated} &
-    (\loc;\sandboxtag)\notin\tmsmonitor.\text{freed} \\
-    \tmsmonitor[']=\left\{\text{allocated}: \tmsmonitor.\text{allocated}\cup\left\{(\loc;\sandboxtag)\right\},\text{freed}: \tmsmonitor.\text{freed}\right\}
-  }{
-    \monitorcheck[]{\tmsmonitor}{\tmsmonitor[']}{\absev{Alloc\ }\loc\ \sandboxtag}
-  }{ap:tms-alloc}
-  %
-  \typerule{tms-dealloc}{
-    (\loc;\sandboxtag)\in\tmsmonitor.\text{allocated} &
-    (\loc;\sandboxtag)\notin\tmsmonitor.\text{freed} \\
-    \tmsmonitor[']=\left\{\text{allocated}: \tmsmonitor.\text{allocated}\setminus\left\{(\loc;\sandboxtag)\right\},\text{freed}: \tmsmonitor.\text{freed}\cup\left\{(\loc;\sandboxtag)\right\}\right\}
-  }{
-    \monitorcheck[]{\tmsmonitor}{\tmsmonitor[']}{\absev{Dealloc\ }\loc\ \sandboxtag}
-  }{ap:tms-dealloc}
+    \bul{$\rtpUniversal{\ccST}{\trgClass}{\sim}$}
+    %, iff 
+    $\isdef$
+    \rul{$\forall \trgProperty\in\trgClass, \srcComponent\in\src{S},$} %
+    if \iul{$\rsat{\srcComponent}{\mapUniversal{\sim}{\trgProperty}}$},
+    then \oul{$\rsat{\ccST\lrpars{\srcComponent}}{\trgProperty}$}.
   \end{center}
+\end{definition}
+A \bul{compiler $\ccST$ for languages $\src{S}$ and $\trg{T}$ robustly preserves a $\src{S}$-level class $\srcClass$}, if for any \rul{property $\srcProperty$ of class $\srcClass$ and programs $\srcComponent$}, where \iul{$\srcComponent$ robustly satisfies $\srcProperty$}, \oul{the compilation of $\srcComponent$, we have that $\ccST\lrpars{\srcComponent}$ robustly satisfies $\mapExistential{\sim}{\srcProperty}$}.
+
+\begin{definition}{\definitionlabel[Robust Preservation with $\mapExistential{\sim}{\cdot}$]{rtp-existential}}
   \begin{center}
-  \judgbox{\monitorcheck{\tmsmonitor}{\tmsmonitor[']}{\tmstrace}}{,,Monitor $\tmsmonitor$ multi-steps to $\tmsmonitor[']$ given trace $\tmstrace$.''}$\;$\\
-  %
-  \typerule{tms-refl}{
-    \tmsmonitor.\text{allocated}=\emptyset
-  }{
-    \monitorcheck{\tmsmonitor}{\tmsmonitor}{\hole{\cdot}}
-  }{ap:tms-refl}
-  %
-  \typerule{tms-ign-trans}{
-    \monitorcheck{\tmsmonitor}{\tmsmonitor[']}{\tmstrace}
-  }{
-    \monitorcheck[]{\tmsmonitor}{\tmsmonitor[']}{\absev{\emptyevent}\cdot\tmstrace}
-  }{ap:tms-ign-trans}
-  %
-  \typerule{tms-trans}{
-    \monitorcheck[]{\tmsmonitor}{\tmsmonitor[']}{\tmsevent} &
-    \monitorcheck{\tmsmonitor[']}{\tmsmonitor['']}{\tmstrace}
-  }{
-    \monitorcheck[]{\tmsmonitor}{\tmsmonitor['']}{\tmsevent\cdot\tmstrace}
-  }{ap:tms-trans}
+    \bul{$\rtpExistential{\ccST}{\srcClass}{\sim}$}
+    %, iff 
+    $\isdef$
+    \rul{$\forall \srcProperty\in\srcClass, \srcComponent\in\src{S},$} %
+    if \iul{$\rsat{\srcComponent}{\srcProperty}$},
+    then \oul{$\rsat{\ccST\lrpars{\srcComponent}}{\mapExistential{\sim}{\srcProperty}}$}.
   \end{center}
+\end{definition}
+
+\subsubsection{Same Trace Models}
+
+\begin{definition}{\definitionlabel[Robust Preservation]{rtp}}
   \begin{center}
-  \judgbox{\tmstraceagree[]{\event}{\tmsevent}}{,,Abstract event $\tmsevent$ is equivalent to $\event$ with respect to \gls*{tms}.''}$\;$\\
-  %
-  \typerule{tms-alloc-authentic}{
-  }{
-    \tmstraceagree[]{\ev{Alloc}\ \loc\ n;\sandboxtag;\securitytag}{\absev{Alloc\ }\loc\ \sandboxtag}
-  }{ap:tms-alloc-authentic}
-  %
-  \typerule{tms-dealloc-authentic}{
-  }{
-    \tmstraceagree[]{\ev{Dealloc}\ \loc;\sandboxtag;\securitytag}{\absev{Dealloc\ }\loc}
-  }{ap:tms-dealloc-authentic}
-  %
-  \typerule{tms-use-authentic}{
-  }{
-    \tmstraceagree[]{\ev{Use}\ \loc\ n;\sandboxtag;\securitytag}{\absev{Use\ }\loc}
-  }{ap:tms-use-authentic}
-  %
-  \typerule{tms-branch-authentic}{
-  }{
-    \tmstraceagree[]{\ev{Branch\ }n}{\absev{\emptyevent}}
-  }{ap:tms-branch-authentic}
-  %
-  \typerule{tms-binop-authentic}{
-  }{
-    \tmstraceagree[]{\ev{Binop\ }n}{\absev{\emptyevent}}
-  }{ap:tms-binop-authentic}
-  %
-  \typerule{tms-empty-authentic}{
-  }{
-    \tmstraceagree[]{\emptyevent}{\absev{\emptyevent}}
-  }{ap:tms-empty-authentic}
-  %
-  \typerule{tms-abort-authentic}{
-  }{
-    \tmstraceagree[]{\lightning}{\absterm}
-  }{ap:tms-abort-authentic}
+    $\rtp{\ccST}{\varClass}$
+    %, iff 
+    $\isdef$
+    $\forall \varProperty\in\varClass, \srcComponent\in\src{S},$ %
+    if $\rsat{\srcComponent}{\varProperty}$,
+    then $\rsat{\ccST\lrpars{\srcComponent}}{\varProperty}$.
   \end{center}
-
-\begin{lemma}[Traces with Monitor Satisfaction are $\tmssafe$]\label{ap:lem:mon:tmssafe}
-  If $\montmssafe{\trace}$, then $\trace\in\tmssafe$. %\Coqed
-\end{lemma}
-\begin{incompleteproof}
-\end{incompleteproof}
-
-\subsection{Spatial Memory Safety}\label{ap:subsec:sms}
-
-\begin{definition}[\glsfirst*{sms}]\label{ap:def:trace:smsdef}
-  \begin{displaymath}
-  \smssafe:=\left\{\trace \left|\begin{array}{rcl}
-                                  \text{If }\ev{Alloc\ \loc\ n;\sandboxtag;\securitytag}\le_{\trace}\ev{Use\ \loc\ m;\sandboxtag;\securitytag'}, \text{ then }m<n
-                                \end{array}\right.\right\}
-  \end{displaymath}
 \end{definition}
-$$\;$$
-\hrule
-$$\;$$
-\begin{gather*}
-  \begin{aligned}
-    \mi{(Abstract\ Store)}~\smsmonitor:=&\ L\times\sandboxtag\times\mb{N}\mid\emptysmsmon\\
-    \mi{(Abstract\ Events)}~\smsevent\bnfdef&\ \absev{\varepsilon} \mid \absev{Alloc\ }\loc\ \sandboxtag\ n \mid \absev{Use\ }\loc\ \sandboxtag\ n \\
-  \end{aligned}
-\end{gather*}
-\begin{center}
-  \judgbox{\monitorcheck[]{\smsmonitor}{\smsmonitor[']}{\smsevent}}{,,Monitor $\smsmonitor$ does one step to $\smsmonitor[']$ given event $\smsevent$.''}$\;$\\
-  %
-  \typerule{sms-uninteresting}{
-  }{
-    \monitorcheck[]{\smsmonitor}{\smsmonitor}{\absev{\emptyevent}}
-  }{ap:sms-uninteresting}
-  %
-  \typerule{sms-abort}{
-  }{
-    \monitorcheck[]{\smsmonitor}{\smsmonitor}{\absterm}
-  }{ap:sms-abort}
-  %
-  \typerule{sms-use}{
-    (\loc;\sandboxtag;m)\in\smsmonitor &
-    n<m
-  }{
-    \monitorcheck[]{\smsmonitor}{\smsmonitor}{\absev{Use\ }\loc\ \sandboxtag\ n}
-  }{ap:sms-use}
-  %
-  \typerule{sms-alloc}{
-    (\loc;\sandboxtag;m)\notin\smsmonitor
-  }{
-    \monitorcheck[]{\smsmonitor}{\smsmonitor\cup\left\{(\loc;\sandboxtag;n)\right\}}{\absev{Alloc\ }\loc\ \sandboxtag\ n}
-  }{ap:sms-alloc}
-\end{center}
-\begin{center}
-  \judgbox{\monitorcheck{\smsmonitor}{\smsmonitor[']}{\smstrace}}{,,Monitor $\smsmonitor$ multi-steps to $\smsmonitor[']$ given trace $\smstrace$.''}$\;$\\
-  %
-  \typerule{sms-refl}{
-  }{
-    \monitorcheck{\smsmonitor}{\smsmonitor}{\hole{\cdot}}
-  }{ap:sms-refl}
-  %
-  \typerule{sms-ign-trans}{
-    \monitorcheck{\smsmonitor}{\smsmonitor[']}{\smstrace}
-  }{
-    \monitorcheck[]{\smsmonitor}{\smsmonitor[']}{\absev{\emptyevent}\cdot\smstrace}
-  }{ap:sms-ign-trans}
-  %
-  \typerule{sms-trans}{
-    \monitorcheck[]{\smsmonitor}{\smsmonitor[']}{\smsevent} &
-    \monitorcheck{\smsmonitor[']}{\smsmonitor['']}{\smstrace}
-  }{
-    \monitorcheck[]{\smsmonitor}{\smsmonitor['']}{\smsevent\cdot\smstrace}
-  }{ap:sms-trans}
-\end{center}
-\begin{center}
-  \judgbox{\smstraceagree[]{\event}{\smsevent}}{,,Abstract event $\smsevent$ is equivalent to $\event$ with respect to \gls*{sms}.''}$\;$\\
-  %
-  \typerule{sms-alloc-authentic}{
-  }{
-    \smstraceagree[]{\ev{Alloc}\ \loc\ n;\sandboxtag;\securitytag}{\absev{Alloc\ }\loc\ n}
-  }{ap:sms-alloc-authentic}
-  %
-  \typerule{sms-dealloc-authentic}{
-  }{
-    \smstraceagree[]{\ev{Dealloc}\ \loc;\sandboxtag;\securitytag}{\absev{Dealloc\ }\loc}
-  }{ap:sms-dealloc-authentic}
-  %
-  \typerule{sms-use-authentic}{
-  }{
-    \smstraceagree[]{\ev{Use}\ \loc\ n;\sandboxtag;\securitytag}{\absev{Use\ }\loc\ n}
-  }{ap:sms-use-authentic}
-  %
-  \typerule{sms-branch-authentic}{
-  }{
-    \smstraceagree[]{\ev{Branch\ }n}{\absev{\emptyevent}}
-  }{ap:sms-branch-authentic}
-  %
-  \typerule{sms-binop-authentic}{
-  }{
-    \smstraceagree[]{\ev{Binop\ }n}{\absev{\emptyevent}}
-  }{ap:sms-binop-authentic}
-  %
-  \typerule{sms-empty-authentic}{
-  }{
-    \smstraceagree[]{\emptyevent}{\absev{\emptyevent}}
-  }{ap:sms-empty-authentic}
-  %
-  \typerule{sms-abort-authentic}{
-  }{
-    \smstraceagree[]{\lightning}{\absterm}
-  }{ap:sms-abort-authentic}
-\end{center}
-\begin{lemma}[Traces with Monitor Satisfaction are $\smssafe$]\label{ap:lem:mon:smssafe}
-  If $\monsmssafe{\trace}$, then $\trace\in\smssafe$.\Coqed
-\end{lemma}
-\begin{proof}
-  First, generalize \Cref{ap:def:trace:smsdef} as follows:
-  \begin{displaymath}
-    \gsmssafe:=\left\{\trace \left|\begin{array}{rcl}
-                                    \text{If }\ev{Alloc\ \loc\ n;\sandboxtag;\securitytag}\le_{\trace}\ev{Use\ \loc\ m;\sandboxtag;\securitytag'}\\
-                                    \text{or }(\ev{Use\ \loc\ m;\sandboxtag;\securitytag}\text{ in }\trace\text{ and }(\loc,\sandboxtag,n)\in\smsmonitor),\\
-                                    \text{ then }m<n
-                                  \end{array}\right.\right\}
-  \end{displaymath}
-  Then, show ,,If $\mongsmssafe{\trace}$, then $\trace\in\gsmssafe$.'' by inducing on the monitor steps.
-
-  Now specialize this generalization to $\smsmonitor=\emptyset$ and the claim follows as a corollary.
-\end{proof}
-
-\subsection{Memory Safety}\label{ap:subsec:ms}
-
-\begin{definition}[\glsfirst*{ms}]\label{ap:def:trace:msdef}
-  \begin{displaymath}
-    \mssafe:=\tmssafe \cap \smssafe
-  \end{displaymath}
-\end{definition}
-$$\;$$
-\hrule
-$$\;$$
-\begin{gather*}
-  \begin{aligned}
-    \mi{(Abstract\ Store)}~\msmonitor:=&\ \tmsmonitor\times\smsmonitor\\
-    \mi{(Initial\ Abstract\ Store)}~\emptymsmon:=&\ (\emptytmsmon, \emptysmsmon) \\
-    \mi{(Abstract\ Events)}~\msevent:=&\ \tmsevent\times\smsevent \mid \absev{\emptyevent}\\
-  \end{aligned}
-\end{gather*}
-\begin{center}
-  \judgbox{\monitorcheck[]{\msmonitor}{\msmonitor[']}{\msevent}}{,,Monitor $\msmonitor$ does one step to $\msmonitor[']$ given event $\msevent$.''}$\;$\\
-  %
-  \typerule{ms-step}{
-    \monitorcheck[]{\tmsmonitor}{\tmsmonitor[']}{\tmsevent} &
-    \monitorcheck[]{\smsmonitor}{\smsmonitor[']}{\smsevent}
-  }{
-    \monitorcheck[]{(\tmsmonitor,\smsmonitor)}{(\tmsmonitor['],\smsmonitor['])}{(\tmsevent,\smsevent)}
-  }{ap:ms-step}
-\end{center}
-\begin{center}
-  \judgbox{\monitorcheck{\msmonitor}{\msmonitor[']}{\mstrace}}{,,Monitor $\msmonitor$ multi-steps to $\msmonitor[']$ given trace $\mstrace$.''}$\;$\\
-  %
-  \typerule{ms-refl}{
-    \monitorcheck{\tmsmonitor}{\tmsmonitor}{\hole{\cdot}} &
-    \monitorcheck{\smsmonitor}{\smsmonitor}{\hole{\cdot}}
-  }{
-    \monitorcheck{(\tmsmonitor,\smsmonitor)}{(\tmsmonitor,\smsmonitor)}{\hole{\cdot}}
-  }{ap:ms-refl}
-  %
-  \typerule{ms-ign-trans}{
-    \monitorcheck{\msmonitor}{\msmonitor[']}{\mstrace}
-  }{
-    \monitorcheck[]{\msmonitor}{\msmonitor[']}{\absev{\emptyevent}\cdot\mstrace}
-  }{ap:ms-ign-trans}
-  %
-  \typerule{ms-trans}{
-    \monitorcheck[]{\msmonitor}{\msmonitor[']}{\msevent} &
-    \monitorcheck{\msmonitor[']}{\msmonitor['']}{\mstrace}
-  }{
-    \monitorcheck[]{\msmonitor}{\msmonitor['']}{\msevent\cdot\mstrace}
-  }{ap:ms-trans}
-\end{center}
-\begin{center}
-  \judgbox{\mstraceagree[]{\event}{\msevent}}{,,Abstract event $\msevent$ is equivalent to $\event$ with respect to \gls*{ms}.''}$\;$\\
-  %
-  \typerule{ms-some-authentic}{
-    \tmstraceagree[]{\event}{\tmsevent} &
-    \smstraceagree[]{\event}{\smsevent}
-  }{
-    \mstraceagree[]{\event}{(\tmsevent,\smsevent)}
-  }{ap:ms-some-authentic}
-  %
-  \typerule{ms-none-authentic}{
-    \tmstraceagree[]{\event}{\absev{\emptyevent}} &
-    \smstraceagree[]{\event}{\absev{\emptyevent}}
-  }{
-    \mstraceagree[]{\event}{\absev{\emptyevent}}
-  }{ap:ms-none-authentic}
-\end{center}
-
-\begin{lemma}[Traces with Monitor Satisfaction are $\mssafe$]\label{ap:lem:mon:mssafe}
-  If $\monmssafe{\trace}$, then $\trace\in\mssafe$.\Coqed
-\end{lemma}
-\begin{proof}
-  Corollary of \Cref{ap:lem:mon:tmssafe,ap:lem:mon:smssafe}.
-\end{proof}
-
-\subsection{Strict Cryptographic Constant Time}\label{ap:subsec:scct}
-
-\begin{definition}[\glsfirst*{scct}]\label{ap:def:trace:scctdef}
-  \begin{displaymath}
-    \scctsafe:=\left\{\trace \left|\begin{array}{rcl}
-                                     \trace=\hole{\cdot} & \text{ or } & \trace=\event[_{b}];\sandboxtag;\unlock\cdot\trace['] \wedge \trace[']\in\scctsafe
-                                   \end{array}\right.\right\}
-  \end{displaymath}
-\end{definition}
-$$\;$$
-\hrule
-$$\;$$
-\begin{gather*}
-  \begin{aligned}
-    \mi{(Abstract\ Store)}~\scctmonitor:=&\ \emptyscctmon\\
-    \mi{(Abstract\ Events)}~\scctevent:=&\ \absev{\emptyevent} \mid \absev{\lightning} \mid \absev{Any} \\
-  \end{aligned}
-\end{gather*}
-\begin{center}
-  \judgbox{\monitorcheck[]{\scctmonitor}{\scctmonitor[']}{\scctevent}}{,,Monitor $\scctmonitor$ does one step to $\scctmonitor[']$ given event $\scctevent$.''}$\;$\\
-  %
-  \typerule{scct-none}{
-  }{
-    \monitorcheck[]{\scctmonitor}{\scctmonitor[']}{\absev{\emptyevent}}
-  }{ap:scct-none}
-  %
-  \typerule{scct-abort}{
-  }{
-    \monitorcheck[]{\scctmonitor}{\scctmonitor[']}{\absev{\lightning}}
-  }{ap:scct-abort}
-\end{center}
-\begin{center}
-  \judgbox{\monitorcheck{\scctmonitor}{\scctmonitor[']}{\sccttrace}}{,,Monitor $\scctmonitor$ multi-steps to $\scctmonitor[']$ given trace $\sccttrace$.''}$\;$\\
-  %
-  \typerule{scct-refl}{
-  }{
-    \monitorcheck{\scctmonitor}{\scctmonitor}{\hole{\cdot}}
-  }{ap:scct-refl}
-  %
-  \typerule{scct-ign-trans}{
-    \monitorcheck{\scctmonitor}{\scctmonitor[']}{\sccttrace}
-  }{
-    \monitorcheck[]{\scctmonitor}{\scctmonitor[']}{\absev{\emptyevent}\cdot\sccttrace}
-  }{ap:scct-ign-trans}
-  %
-  \typerule{scct-trans}{
-    \monitorcheck[]{\scctmonitor}{\scctmonitor[']}{\scctevent} &
-    \monitorcheck{\scctmonitor[']}{\scctmonitor['']}{\sccttrace}
-  }{
-    \monitorcheck[]{\scctmonitor}{\scctmonitor['']}{\scctevent\cdot\sccttrace}
-  }{ap:scct-trans}
-\end{center}
-\begin{center}
-  \judgbox{\sccttraceagree[]{\event}{\absevent}}{,,Abstract event $\absevent$ is equivalent to $\event$ with respect to \gls*{cct}.''}$\;$\\
-  %
-  \typerule{scct-low-authentic}{
-  }{
-    \sccttraceagree[]{\event[_b];\sandboxtag;\unlock}{\absev{\emptyevent}}
-  }{ap:scct-low-authentic}
-  %
-  \typerule{scct-high-authentic}{
-  }{
-    \sccttraceagree[]{\event[_b];\sandboxtag;\lock}{\absev{Any}}
-  }{ap:scct-high-authentic}
-  %
-  \typerule{scct-empty-authentic}{
-  }{
-    \sccttraceagree[]{\emptyevent}{\absev{\emptyevent}}
-  }{ap:scct-empty-authentic}
-  %
-  \typerule{scct-abort-authentic}{
-  }{
-    \sccttraceagree[]{\lightning}{\absterm}
-  }{ap:scct-abort-authentic}
-\end{center}
-\begin{lemma}[Monitor Traces are $\scctsafe$]\label{lem:mon:scctsafe}
-  If $\monscctsafe{\trace}$, then $\trace\in\scctsafe$. %\Coqed
-\end{lemma}
-\begin{incompleteproof}
-\end{incompleteproof}
-
-\subsection{Memory Safety and Strict Cryptographic Constant Time}\label{ap:subsec:msscct}
-\begin{definition}[\glsfirst*{msscct}]\label{ap:def:trace:msscctdef}
-  \begin{displaymath}
-    \msscctsafe:=\mssafe\cap\scctsafe
-  \end{displaymath}
-\end{definition}
-$$\;$$
-\hrule
-$$\;$$
-
-\begin{gather*}
-  \begin{aligned}
-    \mi{(Abstract\ Store)}~\msscctmonitor:=&\ \msmonitor\times\scctmonitor\\
-    \mi{(Initial\ Abstract\ Store)}~\emptymsscctmon:=&\ (\emptymsscctmon, \emptyscctmon) \\
-    \mi{(Abstract\ Events)}~\msscctevent:=&\ \msscctevent\times\scctevent \mid \absev{\emptyevent}\\
-  \end{aligned}
-\end{gather*}
-\begin{center}
-  \judgbox{\monitorcheck[]{\msscctmonitor}{\msscctmonitor[']}{\msscctevent}}{,,Monitor $\msscctmonitor$ does one step to $\msscctmonitor[']$ given event $\msscctevent$.''}$\;$\\
-  %
-  \typerule{msscct-step}{
-    \monitorcheck[]{\msmonitor}{\msmonitor[']}{\msevent} &
-    \monitorcheck[]{\scctmonitor}{\scctmonitor[']}{\scctevent}
-  }{
-    \monitorcheck[]{(\msmonitor,\scctmonitor)}{(\msmonitor['],\scctmonitor['])}{(\msevent,\scctevent)}
-  }{ap:msscct-step}
-\end{center}
-\begin{center}
-  \judgbox{\monitorcheck{\msscctmonitor}{\msscctmonitor[']}{\mssccttrace}}{,,Monitor $\msscctmonitor$ multi-steps to $\msscctmonitor[']$ given trace $\mssccttrace$.''}$\;$\\
-  %
-  \typerule{msscct-refl}{
-    \monitorcheck{\msmonitor}{\msmonitor}{\hole{\cdot}} &
-    \monitorcheck{\scctmonitor}{\scctmonitor}{\hole{\cdot}}
-  }{
-    \monitorcheck{(\msmonitor,\scctmonitor)}{(\msmonitor,\scctmonitor)}{\hole{\cdot}}
-  }{ap:msscct-refl}
-  %
-  \typerule{msscct-ign-trans}{
-    \monitorcheck{\msscctmonitor}{\msscctmonitor[']}{\mssccttrace}
-  }{
-    \monitorcheck[]{\msscctmonitor}{\msscctmonitor[']}{\absev{\emptyevent}\cdot\mssccttrace}
-  }{ap:msscct-ign-trans}
-  %
-  \typerule{msscct-trans}{
-    \monitorcheck[]{\msscctmonitor}{\msscctmonitor[']}{\msscctevent} &
-    \monitorcheck{\msscctmonitor[']}{\msscctmonitor['']}{\mssccttrace}
-  }{
-    \monitorcheck[]{\msscctmonitor}{\msscctmonitor['']}{\msscctevent\cdot\mssccttrace}
-  }{ap:msscct-trans}
-\end{center}
-\begin{center}
-  \judgbox{\mssccttraceagree[]{\event}{\msscctevent}}{,,Abstract event $\msscctevent$ is equivalent to $\event$ with respect to \gls*{ms}.''}$\;$\\
-  %
-  \typerule{msscct-some-authentic}{
-    \mstraceagree[]{\event}{\msevent} &
-    \sccttraceagree[]{\event}{\scctevent}
-  }{
-    \mssccttraceagree[]{\event}{(\msevent,\scctevent)}
-  }{ap:msscct-some-authentic}
-  %
-  \typerule{msscct-none-authentic}{
-    \mstraceagree[]{\event}{\absev{\emptyevent}} &
-    \sccttraceagree[]{\event}{\absev{\emptyevent}}
-  }{
-    \mssccttraceagree[]{\event}{\absev{\emptyevent}}
-  }{ap:msscct-none-authentic}
-\end{center}
-
-\begin{lemma}[Traces with Monitor Satisfaction are $\msscctsafe$]\label{ap:lem:mon:msscctsafe}
-  If $\monmsscctsafe{\trace}$, then $\trace\in\msscctsafe$.\Coqed
-\end{lemma}
-\begin{incompleteproof}
-\end{incompleteproof}
-
-\section{Programming Languages}\label{ap:sec:proglangs}
-
-\subsection{$\Ltms$: Statically Typed, Temporal Memory Safe, Strict Cryptographic Constant Time}
-
-$\Ltms$ trivially satisfies \gls*{scct}, because all variables/values are private ($\lock$) by default.
-
-\begin{gather*}
-  \begin{aligned}
-    \mi{(Expressions)}~&\src{\expr}\bnfdef\ \src{x} \mid \src{\valueexpr} \mid \src{\lbinop{\expr[_{1}]}{\expr[_{2}]}} \mid \src{\lget{x}{\expr}} \mid \src{\llet{x}{\lnew{\expr[_{1}]}{\expr[_{2}]}}{\expr[_{3}]}} \mid \src{\ldelete{x}} \mid \src{\lset{x}{\expr[_{1}]}{\expr[_{2}]}} \\
-                       & \mid \src{\lpair{\expr[_{1}]}{\expr[_{2}]}} \mid \src{\lproja{\expr}} \mid \src{\lprojb{\expr}} \mid \src{\llet{x}{\expr[_{1}]}{\expr[_{2}]}} \mid \src{\lreturn{\expr}} \mid \src{\lcall{g}{\expr}} \mid \src{\lifz{\expr}{\expr[_{1}]}{\expr[_{2}]}} \mid \src{\labort}
-  \end{aligned}
-  \\
-  \begin{aligned}
-  \mi{(Types)}~&\src{\type}\bnfdef\src{\natt} \mid \src{\type[_{1}]} \times \src{\type[_{2}]}
-  &
-  \mi{(Functions)}~&\src{\function}\bnfdef\ \src{\lfunction{\textit{foo}}{x}{e}}
-  &
-\mi{(Libraries)}~&\src{\library}\bnfdef\ \src{\hole{\cdot}} \mid \src{\function},\src{\library}
-  \end{aligned}\\
-  \begin{aligned}
-  \mi{(Component~Names)}~&\src{\commlib}\bnfdef\ \src{\hole{\cdot}} \mid \src{\text{foo}},\src{\commlib}
-  &
-  \mi{(Programs)}~&\src{\prog{\library_{\ctx}}{\library_{\comp}}}
-  \end{aligned}
-\end{gather*}
-
-Variables are referred to as $x,y,z,a,b,c,\ldots$ while functions may be referred to as multi-character words, such as $\texttt{foo}$, as well as short-forms like $f,g,h$.
-
-\begin{gather*}
-  \begin{aligned}
-    \mi{(Control~Tags)}~&\src{\sandboxtag}\bnfdef\src{\ctx} \mid \src{comp}
-    &
-    \mi{(Communication~Tags)}~&\src{\comm}\bnfdef\ \src{\ctxtocomp} \mid\ \src{\comptoctx} \mid \src{\nocomm}
-    &
-    \mi{(Poison~Tags)}~&\src{\poison}\bnfdef \src{\poisoned} \mid \src{\poisonless}
-  \end{aligned}\\
-  \begin{aligned}
-    \mi{(Continuation~Stacks)}~&\src{\ectxstack}\bnfdef\src{\hole{\cdot}} \mid \src{\tup{K}{g}},\src{\ectxstack}
-    &
-    \mi{(Control~Flow~States)}~&\src{\cfstate}\bnfdef\src{\tup{\library}{\ectxstack}}
-  	\\
-    \mi{(Locations)}~&\src{\loc}\in\nat
-    &
-    \mi{(Substitutions)}~&\src{\substvar}\bnfdef \src{\hole{\cdot}} \mid \src{\subst{\valueexpr}{x}},\src{\gamma}
-  	\\
-    \mi{(Stores)}~&\src{\Delta}\bnfdef\src{\hole{\cdot}} \mid \src{x\mapsto\tup{\loc}{\sandboxtag,\poison,n}},\src{\Delta}
-    &
-    \mi{(Heaps)}~&\src{H}\bnfdef\src{\hole{\cdot}} | \src{\valueexpr},\src{H}
-  \end{aligned}\\
-  \begin{aligned}
-    \mi{(Memory~States)}~&\src{\memstate}\bnfdef\src{\tup{H^{\ctx}}{H^{\comp};\Delta}}
-    &
-    \mi{(States)}~&\src{\statevar}\bnfdef\src{\tup{\cfstate}{\sandboxtag,\memstate}}
-    &
-    \mi{(Runtime~Terms)}~&\src{\runtimetermvar}\bnfdef\src{\rtt{\statevar}{\expr}} \mid \src{\lightning}
-  \end{aligned}
-\end{gather*}
-Throughout the paper, the substitution notation is also used to update entries in states $\statevar$.
-\begin{gather*}
-  \begin{aligned}
-    \mi{(Pre\text{-}Events)}~\src{\event[_{b}]}&\bnfdef \src{\ev{Alloc}\ \loc\ \valueexpr} \mid \src{\ev{Dealloc}\ \loc} \mid \src{\ev{Get}\ \loc\ \valueexpr} \mid \src{\ev{Set}\ \loc\ \valueexpr\ \valueexpr[']} \mid \src{\ev{Call}\ \comm\ g\ \valueexpr} \mid \src{\ev{Ret}\ \comm\ \valueexpr} \mid \src{\ev{Start}} \mid \src{\ev{End}\ \valueexpr}\\
-  \end{aligned}\\
-  \begin{aligned}
-    \mi{(Events)}~\src{\event}&\bnfdef \src{\emptyevent} \mid \src{\lightning} \mid \src{\tup{\event[_{b}]}{\sandboxtag,\securitytag}} \\
-  \end{aligned}
-\end{gather*}
-\begin{center}
-  \judgbox{\pstepto{\src{\runtimetermvar}}{\src{\runtimetermvar[']}}{\src{\event}}}{,,$\runtimetermvar$ does one primitive step to $\runtimetermvar[']$ emitting event $\event$.''}
-  %
-  \typerule{$\src{\Ltms}$-e-binop}{
-  }{
-    \pstepto{}{}{\emptyevent}
-  }{ap:ltms-e-binop}
-  %
-  \typerule{$e-\text{get}-\in$}{
-    \sandboxtag=\statevar.\sandboxtag &
-    \statevar.\store(x)=\tup{\loc}{\sandboxtag,\poison,m} \\
-    \loc+n\in\operatorname{dom}\statevar.\heap[^{\sandboxtag}](\loc+n)
-  }{
-    \pstepto{\rtt{\statevar}{\lget{x}{n}}}{\rtt{\statevar}{\heap[^{\sandboxtag}](\loc+n)}}{\tup{\ev{Get}\ \loc\ n}{\sandboxtag}}
-  }{e-get-in}
-  %
-  \typerule{$e-\text{set}-\notin$}{
-    \sandboxtag=\statevar.\sandboxtag &
-    \statevar.\store(x)=\tup{\loc}{\sandboxtag,\poison,m} \\
-    \loc+n\notin\operatorname{dom}\statevar.\heap[^{\sandboxtag}]
-  }{
-    \pstepto{\rtt{\statevar}{\lset{x}{n}{\valueexpr}}}{\rtt{\statevar}{\valueexpr}}{\tup{\ev{Set}\ \loc\ n\ \valueexpr}{\sandboxtag}}
-  }{e-set-notin}
-  %
-  \typerule{$e-\text{new}$}{
-    \fresh{\statevar}{z} &
-    \fresh{\statevar}{\loc} &
-    \heap[^{\sandboxtag}_1]=\growh{\statevar.\heap[^{\sandboxtag}]}{n} &
-    \store[_1]=\storeel{z}{\loc}{\statevar.\sandboxtag}{\poisonless}{n},\statevar.\store
-  }{
-    \pstepto{\rtt{\statevar}{\lnew{x}{n}{\expr}}}{\rtt{\statevar\subst{\heap[^{\sandboxtag}]}{\heap[^{\sandboxtag}_1]}\subst{\store}{\store[_1]}}{\expr\subst{z}{x}}}{\tup{\ev{Alloc}\ \loc\ n}{\sandboxtag}}
-  }{e-new}
-  %
-  \typerule{$e-\text{dealloc}$}{
-    \statevar.\store(x) = \tup{\loc}{\statevar.\sandboxtag,\poison,n} &
-    \store[_1]=\seth{\statevar.\store}{x}{\tup{\loc}{\statevar.\sandboxtag,\poisoned,n}}
-  }{
-    \pstepto{\rtt{\statevar}{\ldelete{x}}}{\rtt{\statevar\subst{\store}{\store[_1]}}{0}}{\tup{\ev{Dealloc}\ \loc}{\sandboxtag}}
-  }{e-dealloc}
-  %
-  \typerule{$e-\text{abort}$}{
-  }{
-    \pstepto{\rtt{\statevar}{\labort}}{\lightning}{\lightning}
-  }{e-abort}
-\end{center}
 
@@
 
+\printglossary
+
 \end{document}
 \endinput
-%%
-%% End of file `sample-sigconf.tex'.

### 2024-09-07 tech/main.tex
Some techreport work

@@
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+\subsection{Preliminaries}
+
+\begin{definition}{\definitionlabel[Function Composition]{function-composition}}
+  $\fncompo{f}{g}\isdef \lambda x.f(g\;x)$
+\end{definition}
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
 \subsection{Satisfaction}
@@
 
+This work relies on property mappings introduced in related work~\cite{abate2021extacc}.
+
 \begin{definition}{\definitionlabel[Universal Image]{universal-image}}
@@
 \subsection{Robust Preservation}
+
+The following definitions for secure compilers are standard~\cite{abate2019jour,abate2021extacc}.
 
@@
 
+\section{Composition of Secure Compilers}\label{sec:rtpc}
+
+This section presents the composition framework.
+
+\subsection{Same Trace Models}\label{subsec:rtpc-same-trace-models}
+
+In the setting of same trace models, composition of secure compilers (as in \defref{rtp}) is simple:
+
+\begin{lemma}{\lemmalabel[Composition of RTP]{rtp-compo}}
+  If 
+  \begin{assumptions}
+    \asm{rtp-compo-comp1}{\rtp{\ccSI}{\varClass[_1]}}
+    \asm{rtp-compo-comp2}{\rtp{\ccIT}{\varClass[_2]}}
+  \end{assumptions}
+  then
+  \begin{goals}
+    \goal{rtp-compo}{\rtp{\fncompo{\ccSI}{\ccIT}}{\varClass[_1]\cap\varClass[_2]}}
+  \end{goals}
+\end{lemma}
+\begin{proof}
+  \incompleteProof
+\end{proof}
+The following corollary is a nice consequence of \lemref{rtp-compo}.
+\begin{corollary}{\corollarylabel[Commutativity of RTP]{rtp-swappable}}
+  If 
+  \begin{assumptions}
+    \asm{rtp-swappable-comp1}{\rtp{\ccII}{\varClass[_1]}}
+    \asm{rtp-swappable-comp2}{\rtp{\ccII[\delta]}{\varClass[_2]}}
+  \end{assumptions}
+  then
+  \begin{goals}
+    \goal{rtp-swappable1}{\rtp{\fncompo{\ccII}{\ccII[\delta]}}{\varClass[_1]\cap\varClass[_2]}}
+    \goal{rtp-swappable2}{\rtp{\fncompo{\ccII[\delta]}{\ccII}}{\varClass[_1]\cap\varClass[_2]}}
+  \end{goals}
+\end{corollary}
+\begin{proof}
+  \goalref{rtp-swappable1} and \goalref{rtp-swappable2} are an immediate consequence based on \lemmaref{rtp-compo} with \asmref{rtp-swappable-comp1} and \asmref{rtp-swappable-comp2}, noting the commutativity of set intersection.
+\end{proof}
+
+\subsection{Different Trace Models}\label{subsec:rtpc-different-trace-models}
+
+The situation in \Cref{subsec:rtpc-same-trace-models} is idealistic in the sense that a somewhat realistic robustly preserving compiler will likely have to insert checks for when control switches from context to component.
+For example, compiling from a statically typed language into a dynamically typed one, where the static language enforces functions to only accept numbers, the compiler has to insert appropriate checks to make sure that well-typedness is robustly preserved.
+Without such checks, a context could simply supply, e.g., a pair, violating the contract that a function can only receive numbers.
+
+These checks yield a mismatch between the source-level trace and the target-level trace, rendering \definitionref{rtp} not provable.
+Instead, one would have to use \definitionref{rtp-universal} or \definitionref{rtp-existential} and use an appropriate, cross-language trace relation that precisely describes the semantic effect on traces when compiling.
+For the type-checking function interfaces example, parts of this cross-language trace relation may be sketched as:
+
+\begin{center}
+  \typerule{ex-rel-interfaces-agree}{
+    \trg{v}\text{ is a number} \rulesep
+    \src{v}\approx\trg{v}
+  }{
+    \xrelTraces{\src{Call\;\ctxtocomp\;v}}{\trg{Call\;\ctxtocomp\;v}}
+  }{ex-rel-interfaces-agree}
+  \typerule{ex-rel-interfaces-disagree}{
+    \trg{v}\text{ not a number}
+  }{
+    \xrelTraces{\src{Abort}}{\trg{Call\;\ctxtocomp\;v}\cdot\trg{Abort}}
+  }{ex-rel-interfaces-disagree}
+\end{center}
+\trref{ex-rel-interfaces-agree} ensures that both in the source and target-level execution, when the context calls into the component, the values agree and are numerical.
+
+\trref{ex-rel-interfaces-disagree} handles the case where the target-level context supplied a $\trg{v}$ that is not a number.
+Here, the backtranslation cannot construct a context that gives a well-typed program {\em and} produces the same trace as the target, since the behavior is ill-defined to begin with.
+So, the only option for the backtranslation is to immediately crash.
+This, however, is a mismatch with the target-level trace, which only aborts in the wrapper that the compiler inserted.
+
+\begin{lemma}{\lemmalabel[Composition of RTP-Universal]{rtp-universal-compo}}
+  If 
+  \begin{assumptions}
+    \asm{rtp-universal-compo-comp1}{\rtpUniversal{\ccSI}{\mapUniversal{\sim_2}{\trgClass[_1]}}{\sim_1}}
+    \asm{rtp-universal-compo-comp2}{\rtpUniversal{\ccIT}{\trgClass[_2]}{\sim_2}}
+  \end{assumptions}
+  then
+  \begin{goals}
+    \goal{rtp-universal-compo}{\rtpUniversal{\fncompo{\ccSI}{\ccIT}}{\trgClass[_1]\cap\trgClass[_2]}{\fncompo{\sim_1}{\sim_2}}}
+  \end{goals}
+\end{lemma}
+\begin{proof}
+  \incompleteProof
+\end{proof}
+
+\begin{lemma}{\lemmalabel[Composition of RTP-Existential]{rtp-existential-compo}}
+  If 
+  \begin{assumptions}
+    \asm{rtp-existential-compo-comp1}{\rtpExistential{\ccSI}{\srcClass[_1]}{\sim_1}}
+    \asm{rtp-existential-compo-comp2}{\rtpExistential{\ccIT}{\mapExistential{\sim_1}{\srcClass[_2]}}{\sim_2}}
+  \end{assumptions}
+  then
+  \begin{goals}
+    \goal{rtp-universal-compo}{\rtpExistential{\fncompo{\ccSI}{\ccIT}}{\srcClass[_1]\cap\srcClass[_2]}{\fncompo{\sim_1}{\sim_2}}}
+  \end{goals}
+\end{lemma}
+\begin{proof}
+  \incompleteProof
+\end{proof}
+
+While compiler passes can be incompatible with each other in the sense that the resulting, mapped property is useless/not what is desired, it is enough to reason about compatibility of compilers at the level of traces.
+This is what this work refers to as ,,semantic effect of a compiler''.
+Instead of having to do syntactic arguments for the interaction of several compilation passes, one only has to do them for the semantic notion, i.e., the cross-language trace relations, which are much simpler proofs by an order of magnitude.
+\MKin{
+  cite section backing up this claim 
+}
+
+\clearpage
 \bibliographystyle{ACM-Reference-Format}

### 2024-09-08 tech/main.tex

@@
 \begin{definition}{\definitionlabel[Function Composition]{function-composition}}
-  $\fncompo{f}{g}\isdef \lambda x.f(g\;x)$
+  $\fncompo{f}{g}\isdef \lambda x.g(f\;x)$
 \end{definition}
@@
 The lifting of \definitionref{existential-image} and \definitionref{universal-image} to classes of properties is standard, i.e., simply element-wise.
+
+\begin{lemma}{\lemmalabel[Trace Relation distributes on Universal Image]{xtrace-distribute-universal}}
+  Let $X$ be either a trace-property or a class. Then:
+  \begin{goals}
+    \goal{xtrace-distribute-universal}{\mapUniversal{\fncompo{\sim_1}{\sim_2}}{X} = \mapUniversal{\sim_2}{\mapUniversal{\sim_1}{X}}}
+  \end{goals}
+\end{lemma}
+\begin{proof}
+  Straightforward unfolding of \definitionref{universal-image} and \definitionref{function-composition} in \goalref{xtrace-distribute-universal}.
+\end{proof}
+\begin{lemma}{\lemmalabel[Trace Relation distributes on Existential Image]{xtrace-distribute-existential}}
+  Let $X$ be either a trace-property or a class. Then:
+  \begin{goals}
+    \goal{xtrace-distribute-existential}{\mapExistential{\fncompo{\sim_1}{\sim_2}}{X} = \mapExistential{\sim_2}{\mapExistential{\sim_1}{X}}}
+  \end{goals}
+\end{lemma}
+\begin{proof}
+  Straightforward unfolding of \definitionref{existential-image} and \definitionref{function-composition} in \goalref{xtrace-distribute-existential}.
+\end{proof}
 

### 2024-09-08 tech/main.tex
Minor changes

@@
 
-\begin{lemma}{\lemmalabel[Trace Relation distributes on Universal Image]{xtrace-distribute-universal}}
+\begin{lemma}[\Coqed]{\lemmalabel[Trace Relation distributes on Universal Image]{xtrace-distribute-universal}}
   Let $X$ be either a trace-property or a class. Then:
@@
 \end{proof}
-\begin{lemma}{\lemmalabel[Trace Relation distributes on Existential Image]{xtrace-distribute-existential}}
+\begin{lemma}[\Coqed]{\lemmalabel[Trace Relation distributes on Existential Image]{xtrace-distribute-existential}}
   Let $X$ be either a trace-property or a class. Then:
@@
 
-\begin{lemma}{\lemmalabel[Composition of RTP]{rtp-compo}}
+\begin{lemma}[\Coqed]{\lemmalabel[Composition of RTP]{rtp-compo}}
   If 
@@
 The following corollary is a nice consequence of \lemref{rtp-compo}.
-\begin{corollary}{\corollarylabel[Commutativity of RTP]{rtp-swappable}}
+\begin{corollary}[\Coqed]{\corollarylabel[Commutativity of RTP]{rtp-swappable}}
   If 
@@
 
-\begin{lemma}{\lemmalabel[Composition of RTP-Universal]{rtp-universal-compo}}
+\begin{lemma}[\Coqed]{\lemmalabel[Composition of RTP-Universal]{rtp-universal-compo}}
   If 
@@
 
-\begin{lemma}{\lemmalabel[Composition of RTP-Existential]{rtp-existential-compo}}
+\begin{lemma}[\Coqed]{\lemmalabel[Composition of RTP-Existential]{rtp-existential-compo}}
   If 
@@
   \begin{goals}
-    \goal{rtp-universal-compo}{\rtpExistential{\fncompo{\ccSI}{\ccIT}}{\srcClass[_1]\cap\srcClass[_2]}{\fncompo{\sim_1}{\sim_2}}}
+    \goal{rtp-existential-compo}{\rtpExistential{\fncompo{\ccSI}{\ccIT}}{\srcClass[_1]\cap\srcClass[_2]}{\fncompo{\sim_1}{\sim_2}}}
   \end{goals}
@@
 
+%%%%%%%%%%%%%%%%%%%%%%%
+%%%%%%%%%%%%%%%%%%%%%%%
+%%%%%%%%%%%%%%%%%%%%%%%
+\section{Case Study}
+
 \clearpage

### 2024-09-12 csf2/main.tex
Paper changes

@@
 
-In case the trace model is the same for both source and target programs (and thus $\sim$ is equality), we have the classic definition of robust preservation~\cite{abate2019jour}.
+In case the trace model is the same for both source and target programs (and thus $\sim$ is equality), we obtain~\cite{abate2019jour}:
 \begin{definition}[Robust Preservation]\label{def:rtp}
@@
 % 
-Note that the previous results also hold for the existential image $\tau_\sim\left(\src{\pi}\right)$, but these results have been moved to the technical report in the interest of space.
+Note that the previous results also hold for the existential image $\tau_\sim\left(\src{\pi}\right)$ as well as for classes of hyperproperties, but these results have been moved to the technical report in the interest of space.
 
@@
 
-\subsection{Secure Upper and Lower Composition}\label{sec:other-compos}
-Besides sequential composition, there are two other compositions, namely an {\em upper}, i.e., a compiler that takes multiple inputs and yields one output, and a {\em lower} composition, i.e., a compiler that takes one input and yields multiple outputs.
-We {define the upper composition $\cc{\src{L}+\obj{L}}{\trg{L}}$} as follows:
-Given a program \texttt{p}, its compiled counterpart is obtained by {plugging \texttt{p} into $\cc{\src{L}}{\trg{L}}$ if $\texttt{p}\in\src{L}$} or by {plugging \texttt{p} into $\cc{\obj{L}}{\trg{L}}$ if $\texttt{p}\in\obj{L}$}.
-\begin{definition}[Upper Composition]
-  $$\text{{$\cc{\src{L}+\obj{L}}{\trg{L}}$}}\isdef
-  % Given \texttt{p}, yield
-  \lambda \texttt{p}\ldotp
-  \left\{\mbox{\begin{tabular}{c}
-    {{if $\texttt{p}\in\src{L}$, then $\cc{\src{L}}{\trg{L}}(\texttt{p})$}} \\
-    \mbox{{if $\texttt{p}\in\obj{L}$, then $\cc{\obj{L}}{\trg{L}}(\texttt{p})$}} \\
-  \end{tabular}}\right.$$
+%\subsection{Secure Upper and Lower Composition}\label{sec:other-compos}
+%Besides sequential composition, there are two other compositions, namely an {\em upper}, i.e., a compiler that takes multiple inputs and yields one output, and a {\em lower} composition, i.e., a compiler that takes one input and yields multiple outputs.
+%We {define the upper composition $\cc{\src{L}+\obj{L}}{\trg{L}}$} as follows:
+%Given a program \texttt{p}, its compiled counterpart is obtained by {plugging \texttt{p} into $\cc{\src{L}}{\trg{L}}$ if $\texttt{p}\in\src{L}$} or by {plugging \texttt{p} into $\cc{\obj{L}}{\trg{L}}$ if $\texttt{p}\in\obj{L}$}.
+%\begin{definition}[Upper Composition]
+%  $$\text{{$\cc{\src{L}+\obj{L}}{\trg{L}}$}}\isdef
+%  % Given \texttt{p}, yield
+%  \lambda \texttt{p}\ldotp
+%  \left\{\mbox{\begin{tabular}{c}
+%    {{if $\texttt{p}\in\src{L}$, then $\cc{\src{L}}{\trg{L}}(\texttt{p})$}} \\
+%    \mbox{{if $\texttt{p}\in\obj{L}$, then $\cc{\obj{L}}{\trg{L}}(\texttt{p})$}} \\
+%  \end{tabular}}\right.$$
 %
@@
 %                                                                          \end{cases}$
-\end{definition}
+%\end{definition}
 
@@
 % A compiler that accepts both $\src{Java}$ and $\obj{Kotlin}$ code translating to the same target language or intermediate representation performs a kind of {\em upper} composition.
-Now, the following theorem tells us what happens if these are secure:
-Given {$\cc{\src{L}}{\trg{L}}$ robustly preserves $\class[_{1}]$} and {$\cc{\obj{L}}{\trg{L}}$ robustly preserves $\class[_{2}]$}, it follows that {their upper composition $\cc{\src{L}+\obj{L}}{\trg{L}}$ robustly preserves the intersection of classes $\class[_{1}]$ and $\class[_{2}]$}.
-
-\begin{theorem}[Upper Composition of Secure Compilers]\label{thm:urtp}
-  $\;$
-
-  If {$\rtp{\cc{\src{L}}{\trg{L}}}{\class[_{1}]}$} and {$\rtp{\cc{\obj{L}}{\trg{L}}}{\class[_{2}]}$}, then {$\rtp{\cc{\src{L}+\obj{L}}{\trg{L}}}{\class[_{1}]\cap\class[_{2}]}$}. %\Coqed
-\end{theorem}
-
-Dually, the {\em lower} composition is concerned about compilers that accept the same source but yield different target languages. %and the same theoretical results apply.
-{Define the lower composition $\cc{\src{L}}{\trg{L}+\obj{L}}$} as follows:
-Given a program $\src{p}$, its compiled counterpart is obtained by {plugging $\src{p}$ into $\cc{\src{L}}{\trg{L}}$} or by {plugging \texttt{p} into $\cc{\src{L}}{\obj{L}}$}, respectively, {based on the internal decision}.
-\begin{definition}[Lower Composition]
-  $$\text{{$\cc{\src{L}}{\trg{L}+\obj{L}}$}}\isdef
-  % Given $L$ and $\src{p}$, yield
-  \lambda \src{p}, L\ldotp
-  \left\{\mbox{\begin{tabular}{c}
-    {{if $L=\trg{L}$, then} {$\cc{\src{L}}{\trg{L}}(\src{p})$}} \\
-    \mbox{{if $L=\obj{L}$, then} {$\cc{\src{L}}{\obj{L}}(\src{p})$}} \\
-  \end{tabular}}\right.$$
+%Now, the following theorem tells us what happens if these are secure:
+%Given {$\cc{\src{L}}{\trg{L}}$ robustly preserves $\class[_{1}]$} and {$\cc{\obj{L}}{\trg{L}}$ robustly preserves $\class[_{2}]$}, it follows that {their upper composition $\cc{\src{L}+\obj{L}}{\trg{L}}$ robustly preserves the intersection of classes $\class[_{1}]$ and $\class[_{2}]$}.
+
+%\begin{theorem}[Upper Composition of Secure Compilers]\label{thm:urtp}
+%  $\;$
+%
+%  If {$\rtp{\cc{\src{L}}{\trg{L}}}{\class[_{1}]}$} and {$\rtp{\cc{\obj{L}}{\trg{L}}}{\class[_{2}]}$}, then {$\rtp{\cc{\src{L}+\obj{L}}{\trg{L}}}{\class[_{1}]\cap\class[_{2}]}$}. %\Coqed
+%\end{theorem}
+
+%Dually, the {\em lower} composition is concerned about compilers that accept the same source but yield different target languages. %and the same theoretical results apply.
+%{Define the lower composition $\cc{\src{L}}{\trg{L}+\obj{L}}$} as follows:
+%Given a program $\src{p}$, its compiled counterpart is obtained by {plugging $\src{p}$ into $\cc{\src{L}}{\trg{L}}$} or by {plugging \texttt{p} into $\cc{\src{L}}{\obj{L}}$}, respectively, {based on the internal decision}.
+%\begin{definition}[Lower Composition]
+%  $$\text{{$\cc{\src{L}}{\trg{L}+\obj{L}}$}}\isdef
+%  % Given $L$ and $\src{p}$, yield
+%  \lambda \src{p}, L\ldotp
+%  \left\{\mbox{\begin{tabular}{c}
+%    {{if $L=\trg{L}$, then} {$\cc{\src{L}}{\trg{L}}(\src{p})$}} \\
+%    \mbox{{if $L=\obj{L}$, then} {$\cc{\src{L}}{\obj{L}}(\src{p})$}} \\
+%  \end{tabular}}\right.$$
 %
@@
 % The following theorem demonstrates what happens if the involved compilers are secure:
-Given {$\cc{\src{L}}{\trg{L}}$ robustly preserves $\class[_{1}]$} and {$\cc{\src{L}}{\obj{L}}$ robustly preserves $\class[_{2}]$}, it follows that {their lower composition $\cc{\src{L}}{\trg{L}+\obj{L}}$ robustly preserves the intersection of classes $\class[_{1}]$ and $\class[_{2}]$}.
-
-\begin{theorem}[Lower Composition of Secure Compilers]\label{thm:lrtp}
-  $\;$
-
-  If {$\rtp{\cc{\src{L}}{\trg{L}}}{\class[_{1}]}$} and {$\rtp{\cc{\obj{L}}{\trg{L}}}{\class[_{2}]}$}, then {$\rtp{\cc{\src{L}}{\trg{L}+\obj{L}}}{\class[_{1}]\cap\class[_{2}]}$}. % \Coqed
-\end{theorem}
+%Given {$\cc{\src{L}}{\trg{L}}$ robustly preserves $\class[_{1}]$} and {$\cc{\src{L}}{\obj{L}}$ robustly preserves $\class[_{2}]$}, it follows that {their lower composition $\cc{\src{L}}{\trg{L}+\obj{L}}$ robustly preserves the intersection of classes $\class[_{1}]$ and $\class[_{2}]$}.
+
+%\begin{theorem}[Lower Composition of Secure Compilers]\label{thm:lrtp}
+%  $\;$ 
+%
+%  If {$\rtp{\cc{\src{L}}{\trg{L}}}{\class[_{1}]}$} and {$\rtp{\cc{\obj{L}}{\trg{L}}}{\class[_{2}]}$}, then {$\rtp{\cc{\src{L}}{\trg{L}+\obj{L}}}{\class[_{1}]\cap\class[_{2}]}$}. % \Coqed
+%\end{theorem}
 
@@
 
-Future work intends to use upper and lower composition as means to build a secure compiler incrementally
-This follows current non-formal-methods based compiler engineering, such as LLVM~\cite{lattner2004llvm}, where languages such as \texttt{C++} and \texttt{Fortran} translate into LLVM IR and languages such as \texttt{x86} or \texttt{ARMv8} assembly variants translate out of LLVM IR.
+%Future work intends to use upper and lower composition as means to build a secure compiler incrementally
+%This follows current non-formal-methods based compiler engineering, such as LLVM~\cite{lattner2004llvm}, where languages such as \texttt{C++} and \texttt{Fortran} translate into LLVM IR and languages such as \texttt{x86} or \texttt{ARMv8} assembly variants translate out of LLVM IR.
 
@@
 However, this is considered standard practice in terms of coding guidelines~\cite{ctguidelines}.
-Moreover, programs that operate on secrets and have a ``data independent timing mode'' enabled do not leak these secrets, despite performing computation on them~\cite{arm-refman,intel-refman} (see \Cref{ex:lscct}).
+Moreover, programs that operate on secrets, have been compiled with FaCT~\cite{cauligi2019fact}, and a ``data independent timing mode'' enabled, do not leak these secrets, despite performing computation on them~\cite{arm-refman,intel-refman} (see \Cref{ex:lscct}).
 
@@
 
-\subsection{$\Lscct$: A Memory-Unsafe Language with a Data Independent Timing Mode}\label{subsec:lscct}
-
-$\Lscct$ extends $\Lms$ (\Cref{subsec:lms}) with a way to read from ($\obj{\lrddoit{x}{\expr}}$) and to write to ($\obj{\lwrdoit{D}}$) a model-specific register that controls a data (operand) independent timing mode, a feature that is present in both Arm~\cite[p.~543]{arm-refman} and Intel~\cite[p.~80]{intel-refman} processors.
+\subsection{$\Lscct$: A Memory-Unsafe Language with a Constant Time Mode}\label{subsec:lscct}
+
+$\Lscct$ extends $\Lms$ (\Cref{subsec:lms}) with a way to read from ($\obj{\lrddoit{x}{\expr}}$) and to write to ($\obj{\lwrdoit{D}}$) a model-specific register that controls a constant-time mode. 
+This mode is inspired by a combination of the effect of the FaCT compiler~\cite{cauligi2019fact} as well as a data-independent timing mode, a feature that is present in both Arm~\cite[p.~543]{arm-refman} and Intel~\cite[p.~80]{intel-refman} processors.
 To this end, states ($\obj{\Omega}$) are extended with the value of the register ($\obj{D}$), which is initially set to be not active ($\obj{OFF}$).
@@
 The state of the register can be read and written to, e.g., \Cref{tr:e-wrdoit-off}.
-For the other languages seen earlier, the mode is intuitively always-on, i.e., the mode of execution always uses data independent timing.
+For the other languages seen earlier, the mode is intuitively always-on, i.e., the mode of execution always uses constant-time mode.
 
@@
 
-Memory accesses to secret data need to be present to reason about memory safety, even when execution is in data-independent timing mode, e.g, \Cref{tr:to-e-get-in-noleak}.
+Memory accesses to secret data need to be present to reason about memory safety, even when execution is in constant-time mode, e.g, \Cref{tr:to-e-get-in-noleak}.
 $\Lscct$ extends base-events with $\obj{\ev{iGet}\ \loc\ \valueexpr}$ and $\obj{\ev{iSet}\ \loc\ \valueexpr}$ (for data $\obj{i}$ndependent get and set) to prevent secrets from leaking but still enable reasoning about memory safety.
@@
 
-Base-events include $\obj{\ev{Branch}\ n}$ and $\obj{\ev{Binop}\ n}$ that are emitted when evaluating a branch or certain binary expressions, such as division, respectively, whenever the data independent timing mode is inactive.
+Base-events include $\obj{\ev{Branch}\ n}$ and $\obj{\ev{Binop}\ n}$ that are emitted when evaluating a branch or certain binary expressions, such as division, respectively, whenever the constant-time mode is inactive.
 Events are extended with a security-tag ($\obj{\securitytag{}}$) to signal the secrecy of the involved data.
@@
 The evaluation steps are amended to propagate the security-tag annotations $\obj{\securitytag{}}$.
-When the data independent timing mode is inactive, base-events $\obj{\ev{Branch}\ n}$ and $\obj{\ev{Binop}\ n}$ are emitted for conditionals and binary operations, respectively.
+When the constant-time mode is inactive, base-events $\obj{\ev{Branch}\ n}$ and $\obj{\ev{Binop}\ n}$ are emitted for conditionals and binary operations, respectively.
 Otherwise, just like in the semantics of the earlier languages, $\obj{\emptyevent}$ is emitted for binary and branching operations.
@@
 
-\begin{exampleenv}[$\Lscct$ with and without data independent timing]\label{ex:lscct}
+\begin{exampleenv}[$\Lscct$ with and without constant-time mode]\label{ex:lscct}
   Consider again \Cref{ex:strncpy}, with a context copying the string \texttt{Hello World}, where everything is marked with a high security tag: $\lock$.
@@
   Read in parallel from top to bottom, the figure shows parts of the execution trace. 
-  In each half, the left column (\textit{Active}) has data-independent timing $\obj{ON}$ and the right one (\textit{Inactive}) has it $\obj{OFF}$.
-
-  When the data independent timing mode is off, the execution yields events in similar fashion to before (\Cref{subsec:ltms,subsec:lsms,subsec:lms}).
+  In each half, the left column (\textit{Active}) has constant-time mode $\obj{ON}$ and the right one (\textit{Inactive}) has it $\obj{OFF}$.
+
+  When the constant-time mode is off, the execution yields events in similar fashion to before (\Cref{subsec:ltms,subsec:lsms,subsec:lms}).
   But, if it is turned on, then the branching event \colorbox{yellow}{does not fire} anymore and both reading and writing to memory is related to a specification trace with \colorbox{lightgray}{no exposed secrets}.
@@
 These passes can be freely ordered in the compilation chain without compromising memory safety (\Cref{thm:cccfccdce:rtp:ms}).
-The next step in the chain ensures that code stays \gls*{scct} (\Cref{thm:ccscct:rtp:scct}) when compiled from $\Lms$ to $\Lscct$, which is done by switching on a data-independent timing mode for the computation.
+The next step in the chain ensures that code stays \gls*{scct} (\Cref{thm:ccscct:rtp:scct}) when compiled from $\Lms$ to $\Lscct$, which is done by switching on a constant-tame mode for the computation.
 Lastly, by introducing barriers immediately after branches, speculative leaks via SPECTRE-PHT are prevented when compiling $\Lscct$ to $\Lspec$.
@@
 Given the fact that $\Lscct$ provides a \gls*{cct}-mode that can be turned on or off, the compiler inserts wrapper code for function calls and function bodies to ensure that execution in the component always happen in \gls*{cct}-mode.
+This simple flag combines the effect of FaCT~\cite{cauligi2019fact}
 
@@
 
-Our secure compilation proofs rely on established proof techniques called backtranslation~\cite{abate2019jour,patrignani2021rsc}, which let one construct a source context starting from either target traces (aka trace-based backtranslations) or target contexts (aka context-based backtranslations).
+Our secure compilation proofs rely on backtranslations~\cite{abate2019jour,patrignani2021rsc}, which let one construct a source context starting from either target traces (aka trace-based backtranslations) or target contexts (aka context-based backtranslations).
 These backtranslations also require setting up cross-language relations between various language elements such as expressions and program states, so we leave these details for the technical reports.
@@
 % For the universally quantified target context, a technique known as backtranslation is well-established, which constructs a source-level context either by backtranslating the target-level context, amounting to a ``decompilation'', or by constructing it using the target-level trace, so that the source-context emits parts of the whole execution trace that are related to the parts emitted by the target-level context.
-All backtranslations in the case-study are trace based except for those required by $\ccdce$ and $\cccf$, which are context-based (and they are essentially an identity function). 
+All backtranslations in the case-study are trace based except for those required by $\ccdce$ and $\cccf$, which are context-based (and they are an identity function). 
 % Since these optimisations have no significant semantic effect and source- and target-language are equal, the backtranslation amounts to an identity translation.
@@
 % However, they do not demonstrate the scalability of the approach by means of a case study.
+
+%\paragraph*{Composition of Secure Compilers}\label{subsec:relw:seccompcomp}
+
+There is work on lifting exploits for single compilers to the whole chain~\cite{paykin2019weird}.
+Our framework does not consider exploits.
+%To the best of our knowledge, there does not exist published work discussing the compositionality of the robust preservation criterion, the gold-standard of secure compilation.
 

### 2024-09-12 tech/main.tex
more daitl

@@
 
+\begin{definition}{\definitionlabel[Closedness w.r.t. Universal Image]{universal-map-wf}}
+  \[
+    \mapUniversalWF{\sim}{\trgClass} \isdef \forall \trgProperty\in\trgClass, \mapUniversal{\sim}{\trgProperty}\in\mapUniversal{\sim}{\trgClass}
+  \]
+\end{definition}
+\begin{definition}{\definitionlabel[Closedness w.r.t. Existential Image]{existential-map-wf}}
+  \[
+    \mapExistentialWF{\sim}{\srcClass} \isdef \forall \srcProperty\in\srcClass, \mapExistential{\sim}{\srcProperty}\in\mapExistential{\sim}{\srcClass}
+  \]
+\end{definition}
+
 \begin{lemma}[\Coqed]{\lemmalabel[Composition of RTP-Universal]{rtp-universal-compo}}
@@
   \begin{assumptions}
+    \asm{rtp-universal-compo-wfsig}{\mapUniversalWF{\sim_2}{\trgClass[_1]}}
     \asm{rtp-universal-compo-comp1}{\rtpUniversal{\ccSI}{\mapUniversal{\sim_2}{\trgClass[_1]}}{\sim_1}}
@@
 
+To conclude a similar colloray as \corollaryref{rtp-swappable}, it is of great use to show that a given cross-language trace relation has no effect on the class of security properties.
+Then, it is possible to derive the following consequences:
+
+\begin{corollary}[\Coqed]{\lemmalabel[Commutativity of RTP-Universal]{rtp-universal-swappable}}
+  If 
+  \begin{assumptions}
+    \asm{rtp-universal-compo-wfsig1}{\mapUniversalWF{\sim_1}{\irlClass[_2]}}
+    \asm{rtp-universal-compo-wfsig2}{\mapUniversalWF{\sim_2}{\irlClass[_1]}}
+    \asm{rtp-universal-compo-eq2}{\mapUniversal{\sim_2}{\irlClass[_1]}=\irlClass[_1]}
+    \asm{rtp-universal-compo-eq1}{\mapUniversal{\sim_1}{\irlClass[_2]}=\irlClass[_2]}
+    \asm{rtp-universal-swappable-comp1}{\rtpUniversal{\ccII}{\irlClass[_1]}{\sim_1}}
+    \asm{rtp-universal-swappable-comp2}{\rtpUniversal{\ccII[\delta]}{\irlClass[_2]}{\sim_2}}
+  \end{assumptions}
+  then
+  \begin{goals}
+    \goal{rtp-universal-swappable1}{\rtpUniversal{\fncompo{\ccII}{\ccII[\delta]}}{\irlClass[_1]\cap\irlClass[_2]}{\fncompo{\sim_1}{\sim_2}}}
+    \goal{rtp-universal-swappable2}{\rtpUniversal{\fncompo{\ccII[\delta]}{\ccII}}{\irlClass[_1]\cap\irlClass[_2]}{\fncompo{\sim_2}{\sim_1}}}
+  \end{goals}
+\end{corollary}
+\begin{proof}
+  todo
+\end{proof}
+
+\begin{corollary}[\Coqed]{\lemmalabel[Commutativity of RTP-existential]{rtp-existential-swappable}}
+  If 
+  \begin{assumptions}
+    \asm{rtp-existential-compo-wfsig1}{\mapExistentialWF{\sim_1}{\irlClass[_2]}}
+    \asm{rtp-existential-compo-wfsig2}{\mapExistentialWF{\sim_2}{\irlClass[_1]}}
+    \asm{rtp-existential-compo-eq2}{\mapExistential{\sim_2}{\irlClass[_1]}=\irlClass[_1]}
+    \asm{rtp-existential-compo-eq1}{\mapExistential{\sim_1}{\irlClass[_2]}=\irlClass[_2]}
+    \asm{rtp-existential-swappable-comp1}{\rtpExistential{\ccII}{\irlClass[_1]}{\sim_1}}
+    \asm{rtp-existential-swappable-comp2}{\rtpExistential{\ccII[\delta]}{\irlClass[_2]}{\sim_2}}
+  \end{assumptions}
+  then
+  \begin{goals}
+    \goal{rtp-existential-swappable1}{\rtpExistential{\fncompo{\ccII}{\ccII[\delta]}}{\irlClass[_1]\cap\irlClass[_2]}{\fncompo{\sim_1}{\sim_2}}}
+    \goal{rtp-existential-swappable2}{\rtpExistential{\fncompo{\ccII[\delta]}{\ccII}}{\irlClass[_1]\cap\irlClass[_2]}{\fncompo{\sim_2}{\sim_1}}}
+  \end{goals}
+\end{corollary}
+\begin{proof}
+  todo
+\end{proof}
+
+if $\sim_1,\sim_2:\irl{I}\to\irl{I}$ are galois, then simpler commutativity
+
 %%%%%%%%%%%%%%%%%%%%%%%

### 2024-09-13 csf2/main.tex
Paper additions

@@
 
-  \bul{$\rtp{\cc{\src{L}}{\trg{L}}}{\class}$}
+  {$\rtp{\cc{\src{L}}{\trg{L}}}{\class}$}
   %, iff 
   $\isdef$
-  \rul{$\forall \pi\in\class, \src{p}\in\src{L},$} if \iul{$\rsat{\src{\progvar}}{\pi}$}, then \oul{$\rsat{\cc{\src{L}}{\trg{L}}\left(\src{p}\right)}{\pi}$}.
+  {$\forall \pi\in\class, \src{p}\in\src{L},$} if {$\rsat{\src{\progvar}}{\pi}$}, then {$\rsat{\cc{\src{L}}{\trg{L}}\left(\src{p}\right)}{\pi}$}.
   % \end{nscenter}
@@
 
-\smallskip%MP: the underline compacts the sections too much
+Consider an examples of compilers fulfilling \Cref{def:rtp}.
+The first example is a mechanized compiler that preserves all safety properties for a language with pointers compiled to an assembly-like language~\cite{korashy2022secureptrs}.
+The mechanization has well over 20k lines in total, which demonstrates that proving a compiler secure with respect to \Cref{def:rtp} is a significant effort.
+
+% i don't see how adding a def of cct is beneficial
+The second example is a compiler ensuring the \gls*{cct} property. %, whose precise definition can be found in \Cref{sec:cct}.
+While we were not able to find compilers in literature that provably fulfill \Cref{def:rtp}, we anticipate that FaCT~\cite{cauligi2019fact} would satisfy this hyperproperty.
+As common, FaCT proves constant-timeness by means of overapproximation, i.e., instead of directly proving the hyperproperty, they fall back to a property that overapproximates the corresponding hyperproperty.
+
+%\smallskip%MP: the underline compacts the sections too much
 
@@
 
-\subsection{Secure Compiler Composition with Same Trace Models}
-When the cross-language trace relation is an equality, \Cref{thm:rtpsim:sig} collapses:
-Given \bul{$\cc{\src{L}}{\trg{L}}$ robustly preserves $\class[_{1}]$} and \rul{$\cc{\trg{L}}{\obj{L}}$ robustly preserves $\class[_{2}]$}, it follows that \oul{their sequential composition $\cc{\src{L}}{\trg{L}}\circ\cc{\trg{L}}{\obj{L}}$ robustly preserves the intersection of classes $\class[_{1}]$ and $\class[_{2}]$}.
-
-\begin{corollary}[Composition of Secure Compilers]\label{corr:rtp}
-  $\;$
-
-  If \bul{$\rtp{\cc{\src{L}}{\trg{L}}}{\class[_{1}]}$} and \rul{$\rtp{\cc{\trg{L}}{\obj{L}}}{\class[_{2}]}$}, then \oul{$\rtp{\cc{\src{L}}{\trg{L}}\circ\cc{\trg{L}}{\obj{L}}}{\class[_{1}]\cap\class[_{2}]}$}.
-\Coqed
-\end{corollary}
-
-\Cref{corr:rtp} provides an easy way to compose secure compilers without well-formedness of trace relations. 
-% 
-However, while \Cref{thm:rtpsim:sig} explicitly requires that $\sim_1$ is well-formed w.r.t. $\trg{\class[_2]}$, if $\sim_2$ is not well-formed w.r.t. $\trg{\class[_1]}$, care must be taken. 
-%Worst-case is that the resulting class $\sigma_{\sim_1\bullet\sim_2}\left(\trg{\class[_1]}\cap\trg{\class[_2]}\right)$ becomes empty, even though $\trg{\class[_1]}\cap\trg{\class[_2]}$ is not necessarily empty.
-This is further discussed in \Cref{subsec:compatsecpasses}.
-
-%If possible, optimisations are merged to enhance the analysis' result, since even when iterating to a fixpoint the final program may not be as optimised as it could be~\cite{click1995combining}.
+It is furthermore possible to conflate the notion of classes and hyperproperties by intersecting all elements of a class of hyperproperties, to obtain one combined hyperproperty.
+This idea has been discussed in other work~\cite{clarkson2008hyper}.
+
+Plugging the examples given in \Cref{subsec:bg:rtp} into \Cref{thm:rtpsim:sig}, we get a compiler that robustly preserves safety properties and the \gls*{cct} hyperproperty. 
+The intersection simply uses the lifted version of the respective safety properties.
+
 Compiler engineers typically try to find an order of optimisations that yields well-optimised programs for either code size~\cite{cooper1999geneticphases} or performance~\cite{kulkarni2006exhaustivephase}.
-\Cref{corr:swappable} justifies that any such order of compilation passes is valid w.r.t. security.
-So, given two compilation passes \bul{$\cc[_{1}]{\trg{L}}{\trg{L}}$, $\cc[_{2}]{\trg{L}}{\trg{L}}$, both robustly preserving class $\class[_{1}]$ or $\class[_{2}]$, respectively}, \oul{for any order of their composition, the composed compiler robustly preserves the intersection of $\class[_{1}]$ and $\class[_{2}]$}.
+\Cref{corr:swappable} justifies that any such order of compilation passes is valid w.r.t. security, as long as the trace-relations have no effect on the respective classes.
+
+So, given two compilation passes \bul{$\cc[_{1}]{\trg{L}}{\trg{L}}$, $\cc[_{2}]{\trg{L}}{\trg{L}}$, both robustly preserving class $\trg{\class[_{1}]}$ or $\trg{\class[_{2}]}$, respectively}, \rul{their corresponding well-formed trace-relations}, \iul{and indifference of the classes with respect to these trace relations}, \oul{for any order of their composition, the composed compiler robustly preserves the intersection of $\trg{\class[_{1}]}$ and $\trg{\class[_{2}]}$}.
 % 
@@
 
-  If \bul{$\rtp{\cc[_{1}]{\trg{L}}{\trg{L}}}{\class[_{1}]}$ and $\rtp{\cc[_{2}]{\trg{L}}{\trg{L}}}{\class[_{2}]}$}, then \oul{$\rtp{\cc[_{1}]{\trg{L}}{\trg{L}}\circ\cc[_{2}]{\trg{L}}{\trg{L}}}{\class[_{1}]\cap\class[_{2}]}$ and $\rtp{\cc[_{2}]{\trg{L}}{\trg{L}}\circ\cc[_{1}]{\trg{L}}{\trg{L}}}{\class[_{2}]\cap\class[_{1}]}$}.
+  If \bul{$\rtpsig{\cc[_{1}]{\trg{L}}{\trg{L}}}{\trg{\class[_{1}]}}{\sim_1}$ and $\rtpsig{\cc[_{2}]{\trg{L}}{\trg{L}}}{\trg{\class[_{2}]}}{\sim_2}$}, %
+  \rul{$\wfcsig{\sim_1}{\trg{\class[_2]}}$ and $\wfcsig{\sim_2}{\trg{\class[_1]}}$}, %
+  and \iul{$\tilde{\sigma}_{\sim_2}\left(\trg{\class[_1]}\right)=\trg{\class[_1]}$ as well as $\tilde{\sigma}_{\sim_1}\left(\trg{\class[_2]}\right)=\trg{\class[_2]}$},
+  then \oul{$\rtpsig{\cc[_{1}]{\trg{L}}{\trg{L}}\circ\cc[_{2}]{\trg{L}}{\trg{L}}}{\trg{\class[_{1}]}\cap\trg{\class[_{2}]}}{\sim_1\circ\sim_2}$ and $\rtpsig{\cc[_{2}]{\trg{L}}{\trg{L}}\circ\cc[_{1}]{\trg{L}}{\trg{L}}}{\trg{\class[_{2}]}\cap\trg{\class[_{1}]}}{\sim_2\circ\sim_1}$}.
 \Coqed
@@
 
-The technical appendix describes a more general version of \Cref{corr:swappable} that is applicable when trace models differ.
-
-%\subsection{Secure Upper and Lower Composition}\label{sec:other-compos}
-%Besides sequential composition, there are two other compositions, namely an {\em upper}, i.e., a compiler that takes multiple inputs and yields one output, and a {\em lower} composition, i.e., a compiler that takes one input and yields multiple outputs.
-%We {define the upper composition $\cc{\src{L}+\obj{L}}{\trg{L}}$} as follows:
-%Given a program \texttt{p}, its compiled counterpart is obtained by {plugging \texttt{p} into $\cc{\src{L}}{\trg{L}}$ if $\texttt{p}\in\src{L}$} or by {plugging \texttt{p} into $\cc{\obj{L}}{\trg{L}}$ if $\texttt{p}\in\obj{L}$}.
-%\begin{definition}[Upper Composition]
-%  $$\text{{$\cc{\src{L}+\obj{L}}{\trg{L}}$}}\isdef
-%  % Given \texttt{p}, yield
-%  \lambda \texttt{p}\ldotp
-%  \left\{\mbox{\begin{tabular}{c}
-%    {{if $\texttt{p}\in\src{L}$, then $\cc{\src{L}}{\trg{L}}(\texttt{p})$}} \\
-%    \mbox{{if $\texttt{p}\in\obj{L}$, then $\cc{\obj{L}}{\trg{L}}(\texttt{p})$}} \\
-%  \end{tabular}}\right.$$
-%
-%
-%  $\begin{cases}
-%                                                                            \rul{\cc{\src{L}}{\trg{L}}(\texttt{p}) & \text{if }\texttt{p}\in\src{L}} \\
-%                                                                            \cc{\obj{L}}{\trg{L}}(\texttt{p}) & \text{if }\texttt{p}\in\obj{L}
-%                                                                          \end{cases}$
-%\end{definition}
-
-% Examples of this are present in industry:
-% Consider the Java Virtual Machine bytecode $\trg{JVM BC}$, which is a popular target for programming language designers due to its high performance and relevance in industry.
-% Compilers for several programming languages have it as their target language, some popular instances are $\src{Java}$ and $\obj{Kotlin}$.
-% Technically speaking, they both compile to class files and $\obj{Kotlin}$ objects are considered to be the same as $\src{Java}$ objects at that point.
-% Both languages can be used at the same time in one project~\cite{androidstudio}.
-% A compiler that accepts both $\src{Java}$ and $\obj{Kotlin}$ code translating to the same target language or intermediate representation performs a kind of {\em upper} composition.
-%Now, the following theorem tells us what happens if these are secure:
-%Given {$\cc{\src{L}}{\trg{L}}$ robustly preserves $\class[_{1}]$} and {$\cc{\obj{L}}{\trg{L}}$ robustly preserves $\class[_{2}]$}, it follows that {their upper composition $\cc{\src{L}+\obj{L}}{\trg{L}}$ robustly preserves the intersection of classes $\class[_{1}]$ and $\class[_{2}]$}.
-
-%\begin{theorem}[Upper Composition of Secure Compilers]\label{thm:urtp}
-%  $\;$
-%
-%  If {$\rtp{\cc{\src{L}}{\trg{L}}}{\class[_{1}]}$} and {$\rtp{\cc{\obj{L}}{\trg{L}}}{\class[_{2}]}$}, then {$\rtp{\cc{\src{L}+\obj{L}}{\trg{L}}}{\class[_{1}]\cap\class[_{2}]}$}. %\Coqed
-%\end{theorem}
-
-%Dually, the {\em lower} composition is concerned about compilers that accept the same source but yield different target languages. %and the same theoretical results apply.
-%{Define the lower composition $\cc{\src{L}}{\trg{L}+\obj{L}}$} as follows:
-%Given a program $\src{p}$, its compiled counterpart is obtained by {plugging $\src{p}$ into $\cc{\src{L}}{\trg{L}}$} or by {plugging \texttt{p} into $\cc{\src{L}}{\obj{L}}$}, respectively, {based on the internal decision}.
-%\begin{definition}[Lower Composition]
-%  $$\text{{$\cc{\src{L}}{\trg{L}+\obj{L}}$}}\isdef
-%  % Given $L$ and $\src{p}$, yield
-%  \lambda \src{p}, L\ldotp
-%  \left\{\mbox{\begin{tabular}{c}
-%    {{if $L=\trg{L}$, then} {$\cc{\src{L}}{\trg{L}}(\src{p})$}} \\
-%    \mbox{{if $L=\obj{L}$, then} {$\cc{\src{L}}{\obj{L}}(\src{p})$}} \\
-%  \end{tabular}}\right.$$
-%
-%
-%  $\begin{cases}
-%                                                                            \rul{\cc{\src{L}}{\trg{L}}(\texttt{p}) & \text{if }\texttt{p}\in\src{L}} \\
-%                                                                            \cc{\obj{L}}{\trg{L}}(\texttt{p}) & \text{if }\texttt{p}\in\obj{L}
-%                                                                          \end{cases}$
-%\end{definition}
-
-% Consider two compilers both accepting $\src{LLVM IR}$~\cite{lattner2004llvm} and one of them emits $\trg{x86\_64}$, while the other emits $\obj{ARMv8}$.
-% It is intuitive that they are in some sense composed in the LLVM framework, but the decision of when to use one over the other is inherently {\em internal} to the formalisation effort of this kind of composition.
-% For example, the user of this compiler provides an explicit flag that instructs to emit $\trg{x86\_64}$ or the framework itself detects the target platform via heuristics, such as supported instructions.
-
-% The following theorem demonstrates what happens if the involved compilers are secure:
-%Given {$\cc{\src{L}}{\trg{L}}$ robustly preserves $\class[_{1}]$} and {$\cc{\src{L}}{\obj{L}}$ robustly preserves $\class[_{2}]$}, it follows that {their lower composition $\cc{\src{L}}{\trg{L}+\obj{L}}$ robustly preserves the intersection of classes $\class[_{1}]$ and $\class[_{2}]$}.
-
-%\begin{theorem}[Lower Composition of Secure Compilers]\label{thm:lrtp}
-%  $\;$ 
-%
-%  If {$\rtp{\cc{\src{L}}{\trg{L}}}{\class[_{1}]}$} and {$\rtp{\cc{\obj{L}}{\trg{L}}}{\class[_{2}]}$}, then {$\rtp{\cc{\src{L}}{\trg{L}+\obj{L}}}{\class[_{1}]\cap\class[_{2}]}$}. % \Coqed
-%\end{theorem}
-
-%Either way, the theoretical results suggest that it is possible to always find a ``most-general'', secure compiler, given two secure compilers, that robustly preserves the least-upper bound of the classes involved in their compilation process.
-
-%Future work intends to use upper and lower composition as means to build a secure compiler incrementally
-%This follows current non-formal-methods based compiler engineering, such as LLVM~\cite{lattner2004llvm}, where languages such as \texttt{C++} and \texttt{Fortran} translate into LLVM IR and languages such as \texttt{x86} or \texttt{ARMv8} assembly variants translate out of LLVM IR.
-
-% no coq, because above definitions are non-formal. 
-%   How would you precisely define the lower compo compiler, given two compilers?
+\subsection{Secure Compiler Composition with Same Trace Models}
+When the cross-language trace relation is an equality, \Cref{thm:rtpsim:sig} collapses:
+Given \bul{$\cc{\src{L}}{\trg{L}}$ robustly preserves $\class[_{1}]$} and \rul{$\cc{\trg{L}}{\obj{L}}$ robustly preserves $\class[_{2}]$}, it follows that \oul{their sequential composition $\cc{\src{L}}{\trg{L}}\circ\cc{\trg{L}}{\obj{L}}$ robustly preserves the intersection of classes $\class[_{1}]$ and $\class[_{2}]$}.
+
+\begin{corollary}[Composition of Secure Compilers]\label{corr:rtp}
+  $\;$
+
+  If \bul{$\rtp{\cc{\src{L}}{\trg{L}}}{\class[_{1}]}$} and \rul{$\rtp{\cc{\trg{L}}{\obj{L}}}{\class[_{2}]}$}, then \oul{$\rtp{\cc{\src{L}}{\trg{L}}\circ\cc{\trg{L}}{\obj{L}}}{\class[_{1}]\cap\class[_{2}]}$}.
+\Coqed
+\end{corollary}
+
+\Cref{corr:rtp} provides an easy way to compose secure compilers without well-formedness of trace relations. 
+% 
+However, while \Cref{thm:rtpsim:sig} explicitly requires that $\sim_1$ is well-formed w.r.t. $\trg{\class[_2]}$, if $\sim_2$ is not well-formed w.r.t. $\trg{\class[_1]}$, care must be taken. 
+%Worst-case is that the resulting class $\sigma_{\sim_1\bullet\sim_2}\left(\trg{\class[_1]}\cap\trg{\class[_2]}\right)$ becomes empty, even though $\trg{\class[_1]}\cap\trg{\class[_2]}$ is not necessarily empty.
+This is further discussed in \Cref{subsec:compatsecpasses}.
+
+%If possible, optimisations are merged to enhance the analysis' result, since even when iterating to a fixpoint the final program may not be as optimised as it could be~\cite{click1995combining}.
+We can also obtain a specialised version of \Cref{corr:swappable}:
+%So, given two compilation passes \bul{$\cc[_{1}]{\trg{L}}{\trg{L}}$, $\cc[_{2}]{\trg{L}}{\trg{L}}$, both robustly preserving class $\class[_{1}]$ or $\class[_{2}]$, respectively}, \oul{for any order of their composition, the composed compiler robustly preserves the intersection of $\class[_{1}]$ and $\class[_{2}]$}.
+% 
+% \MPin{
+% 	later on, say that optimisation passes remove elements from traces, so they should be fine w.r.t. the result with $\tau$.
+% 	also, yes, if $\tau$ adds we have a problem, if it removes we don't
+% \MKin{ Optimisations that remove elements from traces are not necessarily sound,
+%  e.g., remove all deallocations to "optimise" memory management by offloading
+%  the work to the OS.
+% }
+% }
+
+\begin{corollary}[Swapping Secure Compiler Passes]\label{corr:swappable:one}
+  $\;$
+
+  If {$\rtp{\cc[_{1}]{\trg{L}}{\trg{L}}}{\class[_{1}]}$ and $\rtp{\cc[_{2}]{\trg{L}}{\trg{L}}}{\class[_{2}]}$}, then {$\rtp{\cc[_{1}]{\trg{L}}{\trg{L}}\circ\cc[_{2}]{\trg{L}}{\trg{L}}}{\class[_{1}]\cap\class[_{2}]}$ and $\rtp{\cc[_{2}]{\trg{L}}{\trg{L}}\circ\cc[_{1}]{\trg{L}}{\trg{L}}}{\class[_{2}]\cap\class[_{1}]}$}.
+\Coqed
+\end{corollary}
 
@@
 %% If your work has an appendix, this is the place to put it.
-% \appendix
+\appendix
+\subsection{Secure Upper and Lower Composition}\label{sec:other-compos}
+Besides sequential composition, there are two other compositions, namely an {\em upper}, i.e., a compiler that takes multiple inputs and yields one output, and a {\em lower} composition, i.e., a compiler that takes one input and yields multiple outputs.
+We {define the upper composition $\cc{\src{L}+\obj{L}}{\trg{L}}$} as follows:
+Given a program \texttt{p}, its compiled counterpart is obtained by {plugging \texttt{p} into $\cc{\src{L}}{\trg{L}}$ if $\texttt{p}\in\src{L}$} or by {plugging \texttt{p} into $\cc{\obj{L}}{\trg{L}}$ if $\texttt{p}\in\obj{L}$}.
+\begin{definition}[Upper Composition]
+  \[
+    \text{{$\cc{\src{L}+\obj{L}}{\trg{L}}$}}\isdef
+  % Given \texttt{p}, yield
+  \lambda \texttt{p}\ldotp
+  \left\{\mbox{\begin{tabular}{c}
+    {{if $\texttt{p}\in\src{L}$, then $\cc{\src{L}}{\trg{L}}(\texttt{p})$}} \\
+    \mbox{{if $\texttt{p}\in\obj{L}$, then $\cc{\obj{L}}{\trg{L}}(\texttt{p})$}} \\
+  \end{tabular}}\right. 
+  \]
+\end{definition}
+
+Examples of this are present in industry:
+Consider the Java Virtual Machine bytecode $\trg{JVM BC}$, which is a popular target for programming language designers due to its high performance and relevance in industry.
+Compilers for several programming languages have it as their target language, some popular instances are $\src{Java}$ and $\obj{Kotlin}$.
+Technically speaking, they both compile to class files and $\obj{Kotlin}$ objects are considered to be the same as $\src{Java}$ objects at that point.
+Both languages can be used at the same time in one project~\cite{androidstudio}.
+A compiler that accepts both $\src{Java}$ and $\obj{Kotlin}$ code translating to the same target language or intermediate representation performs a kind of {\em upper} composition.
+Now, the following theorem tells us what happens if these are secure:
+Given {$\cc{\src{L}}{\trg{L}}$ robustly preserves $\class[_{1}]$} and {$\cc{\obj{L}}{\trg{L}}$ robustly preserves $\class[_{2}]$}, it follows that {their upper composition $\cc{\src{L}+\obj{L}}{\trg{L}}$ robustly preserves the intersection of classes $\class[_{1}]$ and $\class[_{2}]$}.
+
+\begin{theorem}[Upper Composition of Secure Compilers]\label{thm:urtp}
+  $\;$
+
+  If {$\rtp{\cc{\src{L}}{\trg{L}}}{\class[_{1}]}$} and {$\rtp{\cc{\obj{L}}{\trg{L}}}{\class[_{2}]}$}, then {$\rtp{\cc{\src{L}+\obj{L}}{\trg{L}}}{\class[_{1}]\cap\class[_{2}]}$}. %\Coqed
+\end{theorem}
+
+Dually, the {\em lower} composition is concerned about compilers that accept the same source but yield different target languages. %and the same theoretical results apply.
+{Define the lower composition $\cc{\src{L}}{\trg{L}+\obj{L}}$} as follows:
+Given a program $\src{p}$, its compiled counterpart is obtained by {plugging $\src{p}$ into $\cc{\src{L}}{\trg{L}}$} or by {plugging \texttt{p} into $\cc{\src{L}}{\obj{L}}$}, respectively, {based on the internal decision}.
+\begin{definition}[Lower Composition]
+  \[
+    \text{{$\cc{\src{L}}{\trg{L}+\obj{L}}$}}\isdef
+  % Given $L$ and $\src{p}$, yield
+  \lambda \src{p}, L\ldotp
+  \left\{\mbox{\begin{tabular}{c}
+    {{if $L=\trg{L}$, then} {$\cc{\src{L}}{\trg{L}}(\src{p})$}} \\
+    \mbox{{if $L=\obj{L}$, then} {$\cc{\src{L}}{\obj{L}}(\src{p})$}} \\
+  \end{tabular}}\right.\]
+\end{definition}
+
+ Consider two compilers both accepting $\src{LLVM IR}$~\cite{lattner2004llvm} and one of them emits $\trg{x86\_64}$, while the other emits $\obj{ARMv8}$.
+ It is intuitive that they are in some sense composed in the LLVM framework, but the decision of when to use one over the other is inherently {\em internal} to the formalisation effort of this kind of composition.
+ For example, the user of this compiler provides an explicit flag that instructs to emit $\trg{x86\_64}$ or the framework itself detects the target platform via heuristics, such as supported instructions.
+
+ The following theorem demonstrates what happens if the involved compilers are secure:
+Given {$\cc{\src{L}}{\trg{L}}$ robustly preserves $\class[_{1}]$} and {$\cc{\src{L}}{\obj{L}}$ robustly preserves $\class[_{2}]$}, it follows that {their lower composition $\cc{\src{L}}{\trg{L}+\obj{L}}$ robustly preserves the intersection of classes $\class[_{1}]$ and $\class[_{2}]$}.
+
+\begin{theorem}[Lower Composition of Secure Compilers]\label{thm:lrtp}
+  $\;$
+
+  If {$\rtp{\cc{\src{L}}{\trg{L}}}{\class[_{1}]}$} and {$\rtp{\cc{\obj{L}}{\trg{L}}}{\class[_{2}]}$}, then {$\rtp{\cc{\src{L}}{\trg{L}+\obj{L}}}{\class[_{1}]\cap\class[_{2}]}$}. % \Coqed
+\end{theorem}
+
+Either way, the theoretical results suggest that it is possible to always find a ``most-general'', secure compiler, given two secure compilers, that robustly preserves the least-upper bound of the classes involved in their compilation process.
+
+%Future work intends to use upper and lower composition as means to build a secure compiler incrementally
+%This follows current non-formal-methods based compiler engineering, such as LLVM~\cite{lattner2004llvm}, where languages such as \texttt{C++} and \texttt{Fortran} translate into LLVM IR and languages such as \texttt{x86} or \texttt{ARMv8} assembly variants translate out of LLVM IR.
+
+% no coq, because above definitions are non-formal. 
+%   How would you precisely define the lower compo compiler, given two compilers?
+
+%\section{Cryptographic Constant Time}\label{sec:cct}
 

### 2024-09-16 csf2/main.tex
add text

@@
   \renewcommand{\src}[1]{\mathsf{#1}}
-% \small
+ \small
 \[
@@
 \subsection{Security Properties and Their Meaning}\label{subsec:formalities:props}
-Each of the presented compilers use a cross-language trace relation, which is also used to translate the property from one language to the other one (via the universal image).
-% 
-Whenever such translations occur, it is important to check whether there has been a change of meaning.
-% 
-That is, we could be using a trace relation that translates a property in a language to a different one in another language.
-% 
-To ensure such a bad translation of properties does not occur, \Cref{thm:prop-rel-corr} states (in a general fashion) that that each security property is properly translated using the universal image according to the cross-language trace relation used by the compiler.
+Each of the presented compilers use a cross-language trace relation, which is also used to translate the property from one language to the other one (via the existential or universal images).
+% 
+%Whenever such translations occur, it is important to check whether there has been a change of meaning.
+%Secure compiler engineers need to ensure that such translations of properties do not change meaning in a fundamental wrong way. 
+% lifting or lowering reasoning from target- to source-level or vice-versa,
+While the meaning of projected properties does change with a translation, the change should not allow for a flawed compilation pipeline.
+% 
+For example, we could be using a trace relation that translates a property in a language to a totally different one in another language.
+% 
+To raise the trust into the translation of properties, \Cref{thm:prop-rel-corr} states (in a general fashion) that each security property is faithfully translated using the universal image according to the cross-language trace relation induced by the compiler.
 
@@
   \begin{align*}
-    \forall& \pi \in \{\tmssafe,\smssafe,\scctsafe,\sssafe\}, 
+    \forall& \pi \in 2^{\{\tmssafe,\smssafe,\scctsafe,\sssafe\}}\setminus\{\emptyset\}, 
     \\
@@
 % 
-What happens in practice is that programmers are given the relations from each language to \ev{L} and they need to build a compiler that induces a relation $\sim^{\src{L}}_{\trg{L}}$ that respects this theorem.
-% 
-% We now informally argue why this theorem holds in all of its specialised cases.
+%% MK: the following does not say anything about the theorem, only about the syntactic compiler
+%What happens in practice is that compiler engineers are given the relations from each language to \ev{L} and they need to build a compiler that induces a relation $\sim^{\src{L}}_{\trg{L}}$ that respects this theorem.
+%
+
+We now informally argue why this theorem holds for the composition of all considered properties from \Cref{sec:compprop}.
 
@@
 
+However, the composition is still technically possible and it is the job of the compiler engineers to ensure that the secure compilation pipeline happens in an order that ensures that the mapped security property is the intended one.
+
 \subsection{Secure Compilation Proofs}\label{subsec:seccompproofs}
@@
 The robust preservation of properties as a compiler-level criterion has been analysed extensively~\cite{abate2019jour,patrignani2021rsc,abate2021extacc,patrignani2019survey} and thus we build on that framework.
-No existing work is concerned with composing robustly safe compilers.
+No existing work is concerned with composing secure compilers, however, existing work~\cite{abate2021extacc} sketches composition of trace-relating compiler correctness in a similar way to what has been presented here.
 %Parts of these works consider languages with different trace models and our technical setup can handle this. 
@@
 The authors demonstrate a similar compositionality theorem to what is presented here (\Cref{sec:sequential}) but use it in the context of protocols.
-The work does not consider the generality to support different trace models.
+The work does not consider the generality to support different trace models or composition of compilers which robustly preserve different classes.
 % The authors demonstrate a similar compositionality theorem to what is presented here (\Cref{sec:sequential}) as well as in an earlier version of this work~\cite{kruse2022csc}.

### 2024-09-18 csf2/main.tex
quite some todos

@@
   of Robust and Optimising Compilers
-  \vspace{-1.5em}
+  % \vspace{-1.5em}
 }
@@
         This paper presents a case-study with five programming languages showcasing the previous contribution (\Cref{sec:casestud:defs,sec:casestud:rtp}).
-        To this end, it presents a compilation chain consisting of six passes that ultimately preserve \gls*{specms} by means of composing four secure passes that \emph{individually} preserve \gls*{tms}, \gls*{sms}, \gls*{scct}---an enforceable version of \gls*{cct}---, and \gls*{ss}, and two passes which are well-known optimisations \gls*{dce} and \gls*{cf},.
+        To this end, it presents a compilation chain consisting of six passes that ultimately preserve \gls*{specms} by means of composing four secure passes that \emph{individually} preserve \gls*{tms}, \gls*{sms}, \gls*{scct}---an enforceable version of \gls*{cct}---, and \gls*{ss}, and two passes which are well-known optimisations \gls*{dce} and \gls*{cf}.
         %Furthermore, the chain includes two optimisation passes: one performs \gls*{dce} and the other \gls*{cf}.
@@
 
-Consider an examples of compilers fulfilling \Cref{def:rtp}.
-The first example is a mechanized compiler that preserves all safety properties for a language with pointers compiled to an assembly-like language~\cite{korashy2022secureptrs}.
-The mechanization has well over 20k lines in total, which demonstrates that proving a compiler secure with respect to \Cref{def:rtp} is a significant effort.
-
+Examples of compilers fulfilling \Cref{def:rtp} exist in the literature~\cite{..}\MP{add}.
+For example, \MP{name?} preserves all safety properties for a language with pointers compiled to an assembly-like language~\cite{korashy2022secureptrs},
+% The mechanization has well over 20k lines in total, which demonstrates that proving a compiler secure with respect to \Cref{def:rtp} is a significant effort. \MP{move this line to later ?}
 % i don't see how adding a def of cct is beneficial
-The second example is a compiler ensuring the \gls*{cct} property. %, whose precise definition can be found in \Cref{sec:cct}.
-While we were not able to find compilers in literature that provably fulfill \Cref{def:rtp}, we anticipate that FaCT~\cite{cauligi2019fact} would satisfy this hyperproperty.
-As common, FaCT proves constant-timeness by means of overapproximation, i.e., instead of directly proving the hyperproperty, they fall back to a property that overapproximates the corresponding hyperproperty.
+and FaCT~\cite{cauligi2019fact} preserves the \gls*{cct} property.\MP{for what lang?} %, whose precise definition can be found in \Cref{sec:cct}.
+% While we were not able to find compilers in literature that provably fulfill \Cref{def:rtp}, we anticipate that 
+ % would satisfy this hyperproperty.
+% FaCT proves constant-timeness by means of overapproximation, i.e., instead of directly proving the hyperproperty, they fall back to a property that overapproximates the corresponding hyperproperty.
 
@@
 % 
-Note that the previous results also hold for the existential image $\tau_\sim\left(\src{\pi}\right)$ as well as for classes of hyperproperties, but these results have been moved to the technical report in the interest of space.
-
-It is furthermore possible to conflate the notion of classes and hyperproperties by intersecting all elements of a class of hyperproperties, to obtain one combined hyperproperty.
-This idea has been discussed in other work~\cite{clarkson2008hyper}.
-
-Plugging the examples given in \Cref{subsec:bg:rtp} into \Cref{thm:rtpsim:sig}, we get a compiler that robustly preserves safety properties and the \gls*{cct} hyperproperty. 
-The intersection simply uses the lifted version of the respective safety properties.
+\Cref{thm:rtpsim:sig} can also be stated for the existential image $\tau_\sim\left(\src{\pi}\right)$, but in the interest of space that result has been moved to the appendix\MP{move}.
+% 
+Crucially, \Cref{thm:rtpsim:sig} also holds for \emph{classes of hyperproperties}, and thus, compilers that robustly preserve hyperproperties can be composed with each other as well as with compilers that robustly preserve properties.
+% 
+Unfortunately, proofs of compilers that robustly preserve hyperproperties (including hypersafety) are scarce in the literature due to their complexity.
+\MP{ so what ? }
+%  we don't discuss this further?
+%
+
+% It is furthermore possible to conflate the notion of classes and hyperproperties by intersecting all elements of a class of hyperproperties, to obtain one combined hyperproperty.
+% This idea has been discussed in other work~\cite{clarkson2008hyper}.
+%  MP: i think i stated the point of these 2 lines above
+
+If we take \MP{name} and FaCT from \Cref{subsec:bg:rtp} and compose them according to \Cref{thm:rtpsim:sig}, we obtain a compiler that robustly preserves the intersection of safety properties and the \gls*{cct} hyperproperty.
+% The intersection simply uses the lifted version of the respective safety properties.
+\MP{meaning?}
 
@@
 \Cref{corr:swappable} justifies that any such order of compilation passes is valid w.r.t. security, as long as the trace-relations have no effect on the respective classes.
+Coming back to the example composing \MP{name} with FaCT, \MP{does the order matter?}
 
@@
 This section presents all of them, despite the fact that only property \gls*{scct} is novel and the others are inspired by existing work, in order to showcase all that is required for a formal proof of security for a realistic compilation toolchain.
+\MP{ say something about this in the rebuttal: only the defs are borrowed, the universal images and all that work is new}
 % , none of the properties is novel, for self-containment of the paper, their definition is necessary for the case study.
-For each of the presented properties, the technical appendix defines corresponding monitors, which refine a given property and have a key role in the proofs of this paper.
+For each of the presented properties, the technical appendix\MP{ or report ?} defines corresponding monitors, which refine\MP{enforce?
+and enforce?} a given property and have a key role in the proofs of this paper.
 
@@
 
-For cryptographic code, there is a general guideline that secrets must not be visible on a trace~\cite{ctguidelines}, i.e., marked as $\unlock$.
+For cryptographic code, there is a general guideline that secrets must not be visible on a trace~\cite{ctguidelines}, i.e., secrets should not be marked as $\unlock$.
 In turn, an instruction whose timing is data-dependent must not have a secret as an operand.
@@
 This is because, intuitively, hypersafety properties can relate multiple execution traces with each other, but monitors work on a single execution.
-To sidestep this issue, this section defines the property \gls*{scct}, a stricter variant of \gls*{cct} that enforces the policy that no secret appears on a trace (inspired by earlier work~\cite{almeida2017jasmin}).
+To sidestep this issue, this section defines the property \gls*{scct}, a stricter variant of \gls*{cct} (inspired by earlier work~\cite{almeida2017jasmin}) that enforces the policy that no secret appears on a trace.
 
@@
 All presented programming languages share a common fragment which is partially presented here and in full detail in the technical report. 
-
-\vspace{-.5em}
+\MP{put omitted stuff in appendix + ref there (and add textual explanation)}
+
+% \vspace{-.5em}
 {
@@
 Functions may receive arguments that are not $\trg{\natt}$ and values are extended with tuples.
-
-\vspace{-1em}
+\MP{ only vals? no proj expressions ? }
+
+% \vspace{-1em}
 \[
@@
 $\Lscct$ extends $\Lms$ (\Cref{subsec:lms}) with a way to read from ($\obj{\lrddoit{x}{\expr}}$) and to write to ($\obj{\lwrdoit{D}}$) a model-specific register that controls a constant-time mode. 
-This mode is inspired by a combination of the effect of the FaCT compiler~\cite{cauligi2019fact} as well as a data-independent timing mode, a feature that is present in both Arm~\cite[p.~543]{arm-refman} and Intel~\cite[p.~80]{intel-refman} processors.
+Moreover, this language \MP{ assumes CT operations? -- how do we manifest the fact pass in the language}
+This language is a combination of the effect of the FaCT compiler~\cite{cauligi2019fact} as well as a data-independent timing mode, a feature that is present in both Arm~\cite[p.~543]{arm-refman} and Intel~\cite[p.~80]{intel-refman} processors.
+\MP{ can we be more specific about the fact pass? }
 To this end, states ($\obj{\Omega}$) are extended with the value of the register ($\obj{D}$), which is initially set to be not active ($\obj{OFF}$).
@@
 
-\vspace{-.5em}
+% \vspace{-.5em}
 \[
@@
 To raise the trust into the translation of properties, \Cref{thm:prop-rel-corr} states (in a general fashion) that each security property is faithfully translated using the universal image according to the cross-language trace relation induced by the compiler.
+\MP{aren't tms,sms etc all sets ?}
+\MP{ can we list all these thms in the appendix?}
 
@@
 Our framework does not consider exploits.
+\MP{ a bit dry and short}
 %To the best of our knowledge, there does not exist published work discussing the compositionality of the robust preservation criterion, the gold-standard of secure compilation.

### 2024-09-19 csf2/main.tex
don't make it dry

@@
 There is work on lifting exploits for single compilers to the whole chain~\cite{paykin2019weird}.
-Our framework does not consider exploits.
-\MP{ a bit dry and short}
+%Our framework does not consider exploits.
+%\MP{ a bit dry and short}
+While that work considers {\em in}secure compilation and composition thereof in terms of exploits, the composition they are interested in allows to lift an exploit for one compiler pass to the whole compilation chain.
+Our framework takes the opposite direction and provides compositionality results for secure compilers.
+%
+%
 %To the best of our knowledge, there does not exist published work discussing the compositionality of the robust preservation criterion, the gold-standard of secure compilation.

### 2024-09-19 csf2/main.tex
FaCT

@@
 
-Examples of compilers fulfilling \Cref{def:rtp} exist in the literature~\cite{..}\MP{add}.
-For example, \MP{name?} preserves all safety properties for a language with pointers compiled to an assembly-like language~\cite{korashy2022secureptrs},
+Examples of compilers fulfilling \Cref{def:rtp} exist in the literature~\cite{korashy2022secureptrs,korashy2021capableptrs,abate2021extacc,abate2019jour,patrignani2021rsc}.
+For example, SecurePtrs~\cite{korashy2022secureptrs} gives a compiler that robustly preserves all safety properties for a C-like language to an assembly-like language. 
 % The mechanization has well over 20k lines in total, which demonstrates that proving a compiler secure with respect to \Cref{def:rtp} is a significant effort. \MP{move this line to later ?}
 % i don't see how adding a def of cct is beneficial
-and FaCT~\cite{cauligi2019fact} preserves the \gls*{cct} property.\MP{for what lang?} %, whose precise definition can be found in \Cref{sec:cct}.
+As another example, even though it is not strictly satisfying \Cref{def:rtp}, the FaCT~\cite{cauligi2019fact} compiler preserves the \gls*{cct} property for a C-like language with constant-time primitives, e.g., \texttt{ctselect} for branching.
+%, whose precise definition can be found in \Cref{sec:cct}.
+Throughout this work, it is assumed that FaCT satisfies \Cref{def:rtp}.
 % While we were not able to find compilers in literature that provably fulfill \Cref{def:rtp}, we anticipate that 

### 2024-09-19 csf2/main.tex
FaCT

@@
 % 
-Unfortunately, proofs of compilers that robustly preserve hyperproperties (including hypersafety) are scarce in the literature due to their complexity.
-\MP{ so what ? }
+%Unfortunately, proofs of compilers that robustly preserve hyperproperties (including hypersafety) are scarce in the literature due to their complexity.
+%\MP{ so what ? }
 %  we don't discuss this further?
@@
 
-If we take \MP{name} and FaCT from \Cref{subsec:bg:rtp} and compose them according to \Cref{thm:rtpsim:sig}, we obtain a compiler that robustly preserves the intersection of safety properties and the \gls*{cct} hyperproperty.
+If we take SecurePtrs and FaCT from \Cref{subsec:bg:rtp} and compose them according to \Cref{thm:rtpsim:sig}, we obtain a compiler that robustly preserves the intersection of safety properties and the \gls*{cct} hyperproperty.
+That is, for a source component that robustly satisfies any set of safety properties {\em and} \gls*{cct}, the compiled target component also robustly satisfies the same set of safety properties and \gls*{cct}.
 % The intersection simply uses the lifted version of the respective safety properties.
-\MP{meaning?}
+%\MP{meaning?}
 
@@
 \Cref{corr:swappable} justifies that any such order of compilation passes is valid w.r.t. security, as long as the trace-relations have no effect on the respective classes.
-Coming back to the example composing \MP{name} with FaCT, \MP{does the order matter?}
+Coming back to the example composing SecurePtrs with FaCT, \MP{does the order matter?}
 

### 2024-09-19 csf2/main.tex
add comment
add unrendered comments
projections

@@
 \MP{ say something about this in the rebuttal: only the defs are borrowed, the universal images and all that work is new}
+\MK{ universal images are not new. }
 % , none of the properties is novel, for self-containment of the paper, their definition is necessary for the case study.
-For each of the presented properties, the technical appendix\MP{ or report ?} defines corresponding monitors, which refine\MP{enforce?
-and enforce?} a given property and have a key role in the proofs of this paper.
+For each of the presented properties, the technical report defines corresponding monitors, which refine\MP{enforce? and enforce?} a given property and have a key role in the proofs of this paper. 
+% managed to prove 
+%  if tms-monitor --[trace]-->* ∅  
+%  then trace ∈ tms
+% but not
+%  if trace ∈ tms
+%  then tms-monitor --[trace]-->* ∅
 
@@
 Furthermore, the syntax of $\Ltrg$ is extended with a way to inspect whether a pointer is freed ($\trg{\lispoisoned{x}}$), evaluating to $\trg{0}$ if it is freed and to $\trg{1}$ otherwise.
-Functions may receive arguments that are not $\trg{\natt}$ and values are extended with tuples.
-\MP{ only vals? no proj expressions ? }
+Functions may receive arguments that are not $\trg{\natt}$, values are extended with tuples, and expressions are also extended with pair projections.
 

### 2024-09-19 csf2/main.tex
some todo crunching

@@
 
-  If \bul{$\rtpsig{\cc{\src{L}}{\trg{L}}}{\tilde{\sigma}_{\sim_2}\left(\trg{\class[_{1}]}\right)}{\sim_1}$}, \rul{$\rtpsig{\cc{\trg{L}}{\obj{L}}}{\trg{\class[_2]}}{\sim_2}$}, and \iul{$\wfcsig{\sim_1}{\trg{\class[_2]}}$}, \\ then \oul{$\rtpsig{\cc{\src{L}}{\trg{L}}\circ\cc{\trg{L}}{\obj{L}}}{\trg{\class[_{1}]}\cap\trg{\class[_{2}]}}{\sim_1\bullet\sim_2}$}.
+  If \bul{$\rtpsig{\cc{\src{L}}{\trg{L}}}{\tilde{\sigma}_{\sim_2}\left(\trg{\class[_{1}]}\right)}{\sim_1}$}, \rul{$\rtpsig{\cc{\trg{L}}{\obj{L}}}{\trg{\class[_2]}}{\sim_2}$}, and \iul{$\wfcsig{\sim_2}{\trg{\class[_1]}}$}, \\ then \oul{$\rtpsig{\cc{\src{L}}{\trg{L}}\circ\cc{\trg{L}}{\obj{L}}}{\trg{\class[_{1}]}\cap\trg{\class[_{2}]}}{\sim_1\bullet\sim_2}$}.
 \Coqed
@@
 % 
-\Cref{thm:rtpsim:sig} can also be stated for the existential image $\tau_\sim\left(\src{\pi}\right)$, but in the interest of space that result has been moved to the appendix\MP{move}.
+\Cref{thm:rtpsim:sig} can also be stated for the existential image $\tau_\sim\left(\src{\pi}\right)$, but in the interest of space that result has been moved to the appendix (\Cref{subsec:extimg}).
 % 
@@
 \Cref{corr:swappable} justifies that any such order of compilation passes is valid w.r.t. security, as long as the trace-relations have no effect on the respective classes.
-Coming back to the example composing SecurePtrs with FaCT, \MP{does the order matter?}
 
@@
 
+Coming back to the example composing SecurePtrs with FaCT, it is likely the case that \Cref{corr:swappable} is not applicable.
+While first running SecurePtrs and then FaCT should be fine, the other direction has potential security hazards, since the SecurePtrs compiler does not account for cryptographic-constant time primitives, such as \texttt{ctselect}.
+
 \subsection{Secure Compiler Composition with Same Trace Models}
@@
 This section presents all of them, despite the fact that only property \gls*{scct} is novel and the others are inspired by existing work, in order to showcase all that is required for a formal proof of security for a realistic compilation toolchain.
-\MP{ say something about this in the rebuttal: only the defs are borrowed, the universal images and all that work is new}
-\MK{ universal images are not new. }
+%\MP{ say something about this in the rebuttal: only the defs are borrowed, the universal images and all that work is new}
+%\MK{ universal images are not new. }
 % , none of the properties is novel, for self-containment of the paper, their definition is necessary for the case study.
-For each of the presented properties, the technical report defines corresponding monitors, which refine\MP{enforce? and enforce?} a given property and have a key role in the proofs of this paper. 
+For each of the presented properties, the technical report defines corresponding monitors, which refine a given property and have a key role in the proofs of this paper. 
 % managed to prove 
@@
 All presented programming languages share a common fragment which is partially presented here and in full detail in the technical report. 
-\MP{put omitted stuff in appendix + ref there (and add textual explanation)}
+%\MP{put omitted stuff in appendix + ref there (and add textual explanation)}
 
@@
 
-    \draw[thick,dotted,->] ($(S) + (0.2,0.0)$) to ($(S) - (-0.2,1)$) to node[align=center] {\gls*{ms}\\ \Cref{corr:ccab:rtp:ms}} ($(M) - (0.0,1)$) to (M);
+    \draw[thick,dotted,->] ($(S) + (0.2,-0.27)$) to ($(S) - (-0.2,1)$) to node[align=center] {\gls*{ms}\\ \Cref{corr:ccab:rtp:ms}} ($(M) - (0.0,1)$) to (M);
 
@@
 To raise the trust into the translation of properties, \Cref{thm:prop-rel-corr} states (in a general fashion) that each security property is faithfully translated using the universal image according to the cross-language trace relation induced by the compiler.
-\MP{aren't tms,sms etc all sets ?}
-\MP{ can we list all these thms in the appendix?}
 
@@
   \begin{align*}
-    \forall& \pi \in 2^{\{\tmssafe,\smssafe,\scctsafe,\sssafe\}}\setminus\{\emptyset\}, 
+    \forall& \pi \in \{\tmssafe,\smssafe,\scctsafe,\sssafe\}, 
     \\
@@
 \appendix
+\subsection{Existential Image}\label{subsec:extimg}
+
+\begin{definition}[Existential Image]
+% [Existential and Universal Image]
+\label{def:existential:img}\label{def:tau}
+  \[
+    \tau_\sim\left(\src{\pi}\right) := 
+      \left\{ 
+        \trg{\trace} \mid \exists \src{\trace}\ldotp \src{\trace}\sim\trg{\trace}, \text{ and } \src{\trace}\in\src{\pi} 
+      \right\}
+  \]
+  %\[ 
+  %  \sigma_\sim\left(\trg{\pi}\right) := 
+  %    \left\{ 
+  %      \src{\trace} \mid \forall \trg{\trace}\ldotp \text{if }\src{\trace}\sim\trg{\trace}, \text{ then } \trg{\trace}\in\trg{\pi} 
+  %    \right\}
+  %\]
+\end{definition}
+
+\begin{definition}[Robust Preservation with $\tau_\sim$]\label{def:rtp:tau}
+  % $\;$\\
+  % \vspace{-1em}
+  % \begin{nscenter}
+  % Compiler $\cc{\src{L}}{\trg{L}}$ robustly preserves $\class$, 
+  {$\rtptau{\cc{\src{L}}{\trg{L}}}{\src{\class}}{\sim}$}
+  %, iff 
+  $\isdef$
+    {$\forall \src{\pi}\in\src{\class}, \src{p}\in\src{L},$} if {$\rsat{\src{\progvar}}{\src{\pi}}$}, then {$\rsat{\cc{\src{L}}{\trg{L}}\left(\src{p}\right)}{\tau_\sim\left(\src{\pi}\right)}$}.
+  % \end{nscenter}
+\end{definition}
+
+\begin{theorem}[Composition of Secure Compilers w.r.t. $\tau$]\label{thm:rtpsim:tau}
+  $\;$
+
+  If {$\rtptau{\cc{\src{L}}{\trg{L}}}{\src{\class[_{1}]}}{\sim_1}$}, {$\rtptau{\cc{\trg{L}}{\obj{L}}}{\tilde{\tau}_{\sim_1}\left(\src{\class[_2]}\right)}{\sim_2}$}, and {$\wfctau{\sim_1}{\src{\class[_2]}}$}, \\ then {$\rtptau{\cc{\src{L}}{\trg{L}}\circ\cc{\trg{L}}{\obj{L}}}{\src{\class[_{1}]}\cap\src{\class[_{2}]}}{\sim_1\bullet\sim_2}$}.
+\Coqed
+\end{theorem}
+
+\begin{corollary}[Swapping Secure Compiler Passes]\label{corr:swappable:tau}
+  $\;$
+
+  If {$\rtptau{\cc[_{1}]{\trg{L}}{\trg{L}}}{\src{\class[_{1}]}}{\sim_1}$ and $\rtptau{\cc[_{2}]{\trg{L}}{\trg{L}}}{\src{\class[_{2}]}}{\sim_2}$}, %
+  {$\wfctau{\sim_1}{\src{\class[_2]}}$ and $\wfctau{\sim_2}{\src{\class[_1]}}$}, %
+  and {$\tilde{\tau}_{\sim_1}\left(\src{\class[_2]}\right)=\src{\class[_2]}$ as well as $\tilde{\tau}_{\sim_2}\left(\src{\class[_1]}\right)=\src{\class[_1]}$},
+  then {$\rtptau{\cc[_{1}]{\trg{L}}{\trg{L}}\circ\cc[_{2}]{\trg{L}}{\trg{L}}}{\src{\class[_{1}]}\cap\src{\class[_{2}]}}{\sim_1\circ\sim_2}$ and $\rtptau{\cc[_{2}]{\trg{L}}{\trg{L}}\circ\cc[_{1}]{\trg{L}}{\trg{L}}}{\src{\class[_{2}]}\cap\src{\class[_{1}]}}{\sim_2\circ\sim_1}$}.
+\Coqed
+\end{corollary}
+
 \subsection{Secure Upper and Lower Composition}\label{sec:other-compos}

### 2024-09-19 csf2/main.tex
Lscct

@@
 
-$\Lscct$ extends $\Lms$ (\Cref{subsec:lms}) with a way to read from ($\obj{\lrddoit{x}{\expr}}$) and to write to ($\obj{\lwrdoit{D}}$) a model-specific register that controls a constant-time mode. 
-Moreover, this language \MP{ assumes CT operations? -- how do we manifest the fact pass in the language}
-This language is a combination of the effect of the FaCT compiler~\cite{cauligi2019fact} as well as a data-independent timing mode, a feature that is present in both Arm~\cite[p.~543]{arm-refman} and Intel~\cite[p.~80]{intel-refman} processors.
-\MP{ can we be more specific about the fact pass? }
-To this end, states ($\obj{\Omega}$) are extended with the value of the register ($\obj{D}$), which is initially set to be not active ($\obj{OFF}$).
-If the register is marked active ($\obj{ON}$), the intuition is that no secrets can appear on specification traces.
-If the register is marked inactive, secrets may appear on traces.
-The state of the register can be read and written to, e.g., \Cref{tr:e-wrdoit-off}.
-For the other languages seen earlier, the mode is intuitively always-on, i.e., the mode of execution always uses constant-time mode.
+$\Lscct$ extends $\Lms$ (\Cref{subsec:lms}) with a ,,constant-time mode''. 
+The activation of the mode can be checked ($\obj{\lrddoit{x}{\expr}}$), changed ($\oobj{\lwrdoit{D}}$), and is stored in program states ($\obj{\Omega}$).
+At the beginning of program execution of $\Lscct$ programs, the mode is turned off ($\obj{OFF}$).
+If the mode is enabled (i.e., set to $\obj{ON}$), the intuition is that no secrets are leaked. 
+This models the real-world\footnote{As present in Intel~\cite[p.80]{intel-refman} and ARM~\cite[p.~543]{arm-refman} processors.} data-independent timing mode as well as the result of compiling a program with FaCT~\cite{cauligi2019fact}.
+For example, FaCT rewrite code that branches to use a constant-time selection primitive.
+To not obfuscate our formalisation unnecessarily by duplicating syntax, we simply added the ,,constant-time mode'' to $\Lscct$.
+
+% $\Lscct$ extends $\Lms$ (\Cref{subsec:lms}) with a way to read from ($\obj{\lrddoit{x}{\expr}}$) and to write to ($\obj{\lwrdoit{D}}$) a model-specific register that controls a constant-time mode. 
+% This language is a combination of the effect of the FaCT compiler~\cite{cauligi2019fact} as well as a data-independent timing mode, a feature that is present in both Arm~\cite[p.~543]{arm-refman} and Intel~\cite[p.~80]{intel-refman} processors.
+% \MP{ can we be more specific about the fact pass? }
+% To this end, states ($\obj{\Omega}$) are extended with the value of the register ($\obj{D}$), which is initially set to be not active ($\obj{OFF}$).
+% If the register is marked active ($\obj{ON}$), the intuition is that no secrets can appear on specification traces.
+% If the register is marked inactive, secrets may appear on traces.
+% The state of the register can be read and written to, e.g., \Cref{tr:e-wrdoit-off}.
+% For the other languages seen earlier, the mode is intuitively always-on, i.e., the mode of execution always uses constant-time mode.
 

### 2024-09-19 csf2/main.tex
typo fix

@@
 $\Lscct$ extends $\Lms$ (\Cref{subsec:lms}) with a ,,constant-time mode''. 
-The activation of the mode can be checked ($\obj{\lrddoit{x}{\expr}}$), changed ($\oobj{\lwrdoit{D}}$), and is stored in program states ($\obj{\Omega}$).
+The activation of the mode can be checked ($\obj{\lrddoit{x}{\expr}}$), changed ($\obj{\lwrdoit{D}}$), and is stored in program states ($\obj{\Omega}$).
 At the beginning of program execution of $\Lscct$ programs, the mode is turned off ($\obj{OFF}$).

### 2024-09-23 csf2/main.tex
incorporate some feedback and do rebuttal

@@
 Thus, from a security perspective, \Cref{def:propsat} considers only a passive attacker model, where the attacker observes the execution and, e.g., retrieves secrets from that.
-To consider a stronger model similarly to what existing work does~\cite{abate2019jour,abate2021extacc,maffeis2008code-carrying,gordon2003authenticity,fournet2007authorization,bengtson2011refine,backes2014uniontyps,michael2023mswasm,swasey2017robust,sammler2019benefits}, the we extend the concept of satisfaction with {\em robustness}.
+To consider a stronger model similarly to what existing work does~\cite{abate2019jour,abate2021extacc,maffeis2008code-carrying,gordon2003authenticity,fournet2007authorization,bengtson2011refine,backes2014uniontyps,michael2023mswasm,swasey2017robust,sammler2019benefits}, we extend the concept of satisfaction with {\em robustness}.
 Robust satisfaction considers partial programs $\progvar$, i.e., components with missing imports, which cannot run until said imports are fulfilled.
@@
 This is because, intuitively, hypersafety properties can relate multiple execution traces with each other, but monitors work on a single execution.
-To sidestep this issue, this section defines the property \gls*{scct}, a stricter variant of \gls*{cct} (inspired by earlier work~\cite{almeida2017jasmin}) that enforces the policy that no secret appears on a trace.
+It is a common trick to sidestep this issue by means of overapproximation: this section defines the property \gls*{scct}, a stricter variant of \gls*{cct} (inspired by earlier work~\cite{almeida2017jasmin}) that enforces the policy that no secret appears on a trace.
+Programs that satisfy \gls*{scct} also satisfy \gls*{cct}, but programs that satisfy \gls*{cct} may not satisfy \gls*{scct}.
 
@@
 However, this is considered standard practice in terms of coding guidelines~\cite{ctguidelines}.
-Moreover, programs that operate on secrets, have been compiled with FaCT~\cite{cauligi2019fact}, and a ``data independent timing mode'' enabled, do not leak these secrets, despite performing computation on them~\cite{arm-refman,intel-refman} (see \Cref{ex:lscct}).
+Moreover, programs that have been compiled with FaCT~\cite{cauligi2019fact} and run with a ``data independent timing mode''~\cite{arm-refman,intel-refman} enabled do not leak secrets (see \Cref{ex:lscct}).
 
@@
 For example, a context immediately deallocating an allocated memory region twice trivially violates memory safety, but the main interest in secure compilation is to preserve component-level security, and thus component events.
-Even though this tagging could be used for blame preservation~\cite{patrignani2023blame}, this beyond the focus of this paper.
+Even though this tagging could be used for blame preservation~\cite{patrignani2023blame}, this is beyond the focus of this paper.
 Another key difference is that memory accesses are now explicitly modelled as reads ($\src{\ev{Read\ \loc\ n}}$) and writes ($\src{\ev{Write\ \loc\ n}}$) instead of just uses.

### 2024-09-24 csf2/main.tex
drop novelty claim

@@
 These properties are of practical importance (as mentioned in \Cref{sec:introduction}) and also of interest in the case study presented later (\Cref{sec:casestud:defs,sec:casestud:rtp}). 
-This section presents all of them, despite the fact that only property \gls*{scct} is novel and the others are inspired by existing work, in order to showcase all that is required for a formal proof of security for a realistic compilation toolchain.
+This section presents all of them, despite the fact that they are inspired by existing work, in order to showcase all that is required for a formal proof of security for a realistic compilation toolchain.
 %\MP{ say something about this in the rebuttal: only the defs are borrowed, the universal images and all that work is new}

### 2024-09-26 csf2/main.tex
pagelimit

@@
   of Robust and Optimising Compilers
-  % \vspace{-1.5em}
+  \vspace{-1.5em}
 }
@@
 From an engineering perspective, this is the desirable approach to building secure compilers.
-\begin{center}\small\it
-	{This paper uses syntax highlighting accessible to both colourblind and black \& white readers.
-	% ~\cite{patrignani2020use}.
-% 	% Specifically, it makes use of a $\src{blue}$, $\src{sans\text{-}serif}$ font for a $\src{source}$,
-% 	% a $\trg{red}$, $\trg{bold}$ font for an $\trg{intermediate}$,
-% 	% and a $\obj{green}$, $\obj{teletype}$ font for a $\obj{target}$ language.	
-	For a better experience, please print or view in colours.
-	}
-\end{center}
+% \begin{center}\small\it
+% 	{This paper uses syntax highlighting accessible to both colourblind and black \& white readers.
+% 	% ~\cite{patrignani2020use}.
+% % 	% Specifically, it makes use of a $\src{blue}$, $\src{sans\text{-}serif}$ font for a $\src{source}$,
+% % 	% a $\trg{red}$, $\trg{bold}$ font for an $\trg{intermediate}$,
+% % 	% and a $\obj{green}$, $\obj{teletype}$ font for a $\obj{target}$ language.	
+% 	For a better experience, please print or view in colours.
+% 	}
+% \end{center}
 \end{abstract}
 
-\begin{IEEEkeywords}
-  memory-safety, cryptographic constant-time, secure compilation, robustness
-\end{IEEEkeywords}
+% \begin{IEEEkeywords}
+%   memory-safety, cryptographic constant-time, secure compilation, robustness
+% \end{IEEEkeywords}
 
@@
 % 
-\Cref{thm:rtpsim:sig} can also be stated for the existential image $\tau_\sim\left(\src{\pi}\right)$, but in the interest of space that result has been moved to the appendix (\Cref{subsec:extimg}).
+\Cref{thm:rtpsim:sig} can also be stated for the existential image $\tau_\sim\left(\src{\pi}\right)$, but in the interest of space that result has been moved to the appendix.
+ % (\Cref{subsec:extimg}).
 % 
@@
 % , none of the properties is novel, for self-containment of the paper, their definition is necessary for the case study.
-For each of the presented properties, the technical report defines corresponding monitors, which refine a given property and have a key role in the proofs of this paper. 
+% For each of the presented properties, the technical report defines corresponding monitors, which refine a given property and have a key role in the proofs of this paper. 
+The technical report defines monitors for each of the presented properties.
+% 
+Monitors refine each property and have a key tole in the proofs of this paper.
 % managed to prove 
@@
 
-This section first discusses how to connect each language-specific security property to the general security properties presented in \Cref{sec:compprop} (\Cref{subsec:formalities:maps}), then it demonstrates that the security property resulting from the universal image projection is faithful to the original property (\Cref{subsec:formalities:props}).
-Finally, this section discusses why the order of compiler passes matters, and how does our framework help with identifying insecure compositions (\Cref{subsec:compatsecpasses}), and it gives additional technical insights on the secure compilation proofs (\Cref{subsec:seccompproofs}).
+This section discusses how to connect each language-specific security property to the general security properties of \Cref{sec:compprop} (\Cref{subsec:formalities:maps}), and it demonstrates that the security property resulting from the universal image projection is faithful to the original property (\Cref{subsec:formalities:props}).
+% 
+Then, this section discusses why the order of compiler passes matters, and how does our framework help with identifying insecure compositions (\Cref{subsec:compatsecpasses}), and it gives additional technical insights on the secure compilation proofs (\Cref{subsec:seccompproofs}).
 
@@
 The work relating robust preservation with universal composability~\cite{patrignani2022universal} is closest to what this paper presents.
-The authors demonstrate a similar compositionality theorem to what is presented here (\Cref{sec:sequential}) but use it in the context of protocols.
+% 
+The authors demonstrate a similar compositionality theorem to ours (\Cref{sec:sequential}) but use it in the context of protocols.
+% 
 The work does not consider the generality to support different trace models or composition of compilers which robustly preserve different classes.

### 2024-10-10 tech/main.tex
some proofs

@@
 \begin{proof}
-  \incompleteProof
+  Unfold \goalref{rtp-compo} and introduce the assumptions. 
+  \begin{passumptions}
+    \asm{rtp-compo-prop}{\varProperty\in\varClass[_1]\cap\varClass[_2]}
+    \asm{rtp-compo-trg-prog}{\srcComponent\in\src{S}}
+    \asm{rtp-compo-rsat}{\rsat{\srcComponent}{\varProperty}}
+  \end{passumptions}
+  What is left to show is:
+  \begin{goals}
+    \goal{rtp-compo-rsat}{\rsat{\ccIT\lrpars{\ccSI\lrpars{\srcComponent}}}%
+                               {\varProperty}}
+  \end{goals}
+  Apply \asmref{rtp-compo-comp2} on \goalref{rtp-compo-rsat} and note that due to \asmref{rtp-compo-prop}, $\varProperty\in\varClass[_2]$ also holds.
+  \begin{goals}
+    \goal{rtp-compo-rsatc}{\rsat{\ccSI\lrpars{\srcComponent}}{\varProperty}}
+  \end{goals}
+  Similarly, apply \asmref{rtp-compo-comp1} on \goalref{rtp-compo-rsatc} and note that due to \asmref{rtp-compo-prop}, $\varProperty\in\varClass[_1]$ also holds.
+  \begin{goals}
+    \goal{rtp-compo-rsatp}{\rsat{\srcComponent}{\varProperty}}
+  \end{goals}
+  \goalref{rtp-compo-rsatp} is resolved by \asmref{rtp-compo-rsat}.
 \end{proof}
+%%%%%%%%%%%%%%%%%%%%%%%%%%
 The following corollary is a nice consequence of \lemref{rtp-compo}.
@@
 \begin{proof}
-  \incompleteProof
+  Unfold \goalref{rtp-universal-compo} and introduce the assumptions. 
+  \begin{passumptions}
+    \asm{rtp-universal-compo-prop}{\trgProperty\in\trgClass[_1]\cap\trgClass[_2]}
+    \asm{rtp-universal-compo-trg-prog}{\srcComponent\in\src{S}}
+    \asm{rtp-universal-compo-rsat}{\rsat{\srcComponent}{\mapUniversal{\fncompo{\sim_1}{\sim_2}}{\trgProperty}}}
+  \end{passumptions}
+  What is left to show is:
+  \begin{goals}
+    \goal{rtp-universal-compo-rsat}{\rsat{\ccIT\lrpars{\ccSI\lrpars{\srcComponent}}}%
+                                         {\trgProperty}}
+  \end{goals}
+  Apply \asmref{rtp-universal-compo-comp2} on \goalref{rtp-universal-compo-rsat} and note that due to \asmref{rtp-universal-compo-prop}, $\trgProperty\in\trgClass[_2]$ also holds.
+  \begin{goals}
+    \goal{rtp-universal-compo-rsatc}{\rsat{\ccSI\lrpars{\srcComponent}}{\mapUniversal{\sim_2}{\trgProperty}}}
+  \end{goals}
+  Similarly, apply \asmref{rtp-universal-compo-comp1} on \goalref{rtp-universal-compo-rsatc} and note that due to \asmref{rtp-universal-compo-prop} and \asmref{rtp-universal-compo-wfsig}, $\trgProperty\in\mapUniversal{\sim_2}{\trgClass[_1]}$ also holds.
+  \begin{goals}
+    \goal{rtp-universal-compo-rsatp}{\rsat{\srcComponent}{\mapUniversal{\sim_2}{\mapUniversal{\sim_1}{\trgProperty}}}}
+  \end{goals}
+  Note that $\mapUniversal{\sim_2}{\mapUniversal{\sim_1}{\trgProperty}}=\mapUniversal{\fncompo{\sim_1}{\sim_2}}{\trgProperty}$, and thus \goalref{rtp-universal-compo-rsatp} is resolved by \asmref{rtp-universal-compo-rsat}.
 \end{proof}
@@
 \begin{proof}
-  \incompleteProof
+  Similar to \lemmaref{rtp-universal-compo}.
 \end{proof}
@@
 \begin{proof}
-  todo
+  \goalref{rtp-universal-swappable1} and \goalref{rtp-universal-swappable2} are an immediate consequence based on \lemmaref{rtp-universal-compo} with \asmref{rtp-universal-swappable-comp1} and \asmref{rtp-universal-swappable-comp2}, noting the commutativity of set intersection.
 \end{proof}
@@
 \begin{proof}
-  todo
+  \goalref{rtp-existential-swappable1} and \goalref{rtp-existential-swappable2} are an immediate consequence based on \lemmaref{rtp-existential-compo} with \asmref{rtp-existential-swappable-comp1} and \asmref{rtp-existential-swappable-comp2}, noting the commutativity of set intersection.
 \end{proof}
-
-if $\sim_1,\sim_2:\irl{I}\to\irl{I}$ are galois, then simpler commutativity
 

### 2024-10-10 tech/main.tex
tech

@@
 \documentclass[a4paper,12pt]{article}
-\usepackage[left=1.5cm,top=1.5cm,right=1.5cm,bottom=1.5cm]{geometry}
+\usepackage[left=1.0cm,top=1.5cm,right=1.0cm,bottom=1.5cm]{geometry}
 \usepackage{parskip}
@@
 %%%%%%%%%%%%%%%%%%%%%%%
-\section{Case Study}
+\section{Case Study}\label{sec-cs}
+This part instantiates the previously presented theory. 
+\Cref{subsec-cs-properties} presents all properties.
+
+\subsection{Properties}\label{subsec-cs-properties}
+
+The properties considered in this case-study are a mixture of trace-properties as welll as hyperproperties.
+However, to carry out proofs, all properties are overapproximated by a monitor.
+
+\subsubsection{Formal Definition}\label{subsubsec-cs-properties-formaldef}
+
+\paragraph{Temporal Memory Safety}
+
+\gls*{tms}~\cite{nagarakatte2010cets} is a safety property that describes that an unallocated object must not be (re-)used.
+
+\begin{definition}{\definitionlabel[\glsfirst{tms}]{def-trace-tmsdef}}
+  \[
+  \tmssafe:=\left\{\varTrace[_{\mssafe}]
+    \left| 
+      \begin{array}{rcl}
+        \ev{Alloc\ \varLoc\ n}&\le_{\varTrace[_{\mssafe}]}&\ev{Dealloc\ \varLoc} \\
+        \ev{Use\ \varLoc\ n}&\le_{\varTrace[_{\mssafe}]}&\ev{Dealloc\ \varLoc} \\
+        \text{at most one }\ev{Dealloc\ \varLoc}&\text{in}&\varTrace[_{\mssafe}] \\
+        \text{at most one }\ev{Alloc\ \varLoc\ n}&\text{in}&\varTrace[_{\mssafe}] \\
+      \end{array}
+    \right.
+  \right\}
+  \]
+\end{definition}
+Hereby, the notation $\varEvent[_{1}]\le_{\varTrace}\varEvent[_{2}]$ means that if $\varEvent[_{1}]$ is in $\varTrace$ and if $\varEvent[_{2}]$ is in $\varTrace$, then $\varEvent[_{1}]$ appears before $\varEvent[_{2}]$.
+
+\paragraph{Spatial Memory Safety}
+\gls*{sms}~\cite{nagarakatte2009soft} is a safety property that prohibits out-of-bounds accesses.
+
+\begin{definition}{\definitionlabel[\glsfirst{sms}]{def-trace-smsdef}}
+  \[
+  \smssafe:=\left\{\varTrace[_{\mssafe}] \left|\begin{array}{rcl}
+      \text{If }\ev{Alloc\ \varLoc\ n}\le_{\varTrace_{\mssafe}}\ev{Use\ \varLoc\ m}, \text{ then }m<n
+  \end{array}\right.\right\}
+  \]
+\end{definition}
+
+\paragraph{Memory Safety}
+
+In spirit of earlier work~\cite{nagarakatte2009soft,nagarakatte2010cets,jim2002cyclone,necula2005ccured,michael2023mswasm}, full \gls*{ms} is the intersection of \Cref{def:trace:tmsdef,def:trace:smsdef}.
+
+\begin{definition}{\definitionlabel[\glsfirst{ms}]{def-trace-msdef}}
+  \[
+  \mssafe:=\tmssafe \cap \smssafe
+  \]
+\end{definition}
+
+Note that \Cref{def:trace:msdef} ignores data isolation, so there may still be memory-safety issues introduced by side-channels.
+
+\subsubsection{Strict Cryptographic Constant Time}
+
+\gls*{cct} is a hypersafety property~\cite{barthe2018sec} and, thus, difficult to check with monitors.
+This is because, intuitively, hypersafety properties can relate multiple execution traces with each other, but monitors work on a single execution.
+It is a common trick to sidestep this issue by means of overapproximation: this section defines the property \gls*{scct}, a stricter variant of \gls*{cct} (inspired by earlier work~\cite{almeida2017jasmin}) that enforces the policy that no secret appears on a trace.
+Programs that satisfy \gls*{scct} also satisfy \gls*{cct}, but programs that satisfy \gls*{cct} may not satisfy \gls*{scct}.
+
+\begin{definition}{\definitionlabel[\glsfirst{scct}]{def-trace-scctdef}}
+  \noindent\[
+  \scctsafe:=\left\{\varTrace[_{\ctsafe}] 
+      \left|
+        \begin{array}{l}
+          \varTrace[_{\ctsafe}]=\hole{\cdot} \text{ or } \\
+          \exists\varTrace[_{\ctsafe}'],\varTrace[_{\ctsafe}]=\varEvent[_{\ctsafe}];\unlock\cdot\varTrace[_{\ctsafe}'] \wedge \varTrace[_{\ctsafe}']\in\scctsafe
+        \end{array}
+      \right.
+    \right\}
+  \]
+\end{definition}
+
+\gls*{scct} may appear overly strict, since it seems that secrets must not occur on a trace (since $\varSecuritytag$ is forced to be $\unlock$). 
+However, this is considered standard practice in terms of coding guidelines~\cite{ctguidelines}.
+Moreover, programs that have been compiled with FaCT~\cite{cauligi2019fact} and run with a ``data independent timing mode''~\cite{arm-refman,intel-refman} enabled do not leak secrets.
+
+\paragraph{\gls*{ms}, Strict Cryptographic Constant Time}
+
+The combination of \gls*{ms} and \gls*{scct} is the intersection of these properties, \gls*{msscct}.
+However, \gls*{ms} uses a different trace model than \gls*{scct}, so intersecting them would trivially yield the empty set. 
+To remedy this issue, we introduce $\sim_{\ctsafe}: \preevent[_{\ctsafe}] \times \preevent[_{\mssafe}] $, a cross-language trace relation (whose key cases are presented below), that we use to intuitively unify the trace model in which the two properties are expressed:
+\[
+  \typerulenolabel{}{\varEvent[_{\mssafe}]\sim_{\ctsafe}\varEvent[_{\mssafe}];\unlock{}}
+  \hspace{2ex}
+  \typerulenolabel{}{\lightning\sim_{\ctsafe}\lightning}
+  \hspace{2ex}
+  \typerulenolabel{}{\emptyevent\sim_{\ctsafe}\ev{Branch\;n};\varSecuritytag{}}
+  \hspace{2ex}
+  \typerulenolabel{}{\emptyevent\sim_{\ctsafe}\ev{Binop\;n};\varSecuritytag{}}
+\]
+
+Essentially, $\sim_{\ctsafe}$ ignores both the new $\ev{Branch\ n}$ and $\ev{Binop\ n}$ base-events as it relates security-insensitive actions ($\unlock$) to their equivalent counterparts.
+Thus, \gls*{ms} traces trivially satisfy \gls*{scct}.
+% 
+The above relation is extended point-wise to traces, skipping the empty event $\emptyevent$ on either side, and it is now possible to define \gls*{msscct} using the universal image:
+
+\begin{definition}{\definitionlabel[\glsfirst{ms} and \glsfirst{scct}]{def-trace-msscctdef}}
+  \[
+  \msscctsafe:=\mssafe\cap\sigma_{\sim_{\ctsafe}}\left(\scctsafe\right)
+  \]
+\end{definition}
+
+\paragraph{Extending the Trace Model with Speculation}
+
+So far, the considered trace models do not let us express speculative execution attacks such as Spectre~\cite{kocher2019spectre}. 
+For this, we extend the earlier trace model so that the security tags ($\varSecuritytag{}$) carry additional information about the kind of private data leakage, i.e., the type of speculative leak.
+Moreover, we add base-events signalling the beginning of a speculative execution ($\ev{Spec}$), a barrier ($\ev{Barrier}$) that signals that any speculative execution may not go past it, as well as a rollback event ($\ev{Rlb}$), which signals that execution resumes to where speculation started.
+
+\vspace{-1em}
+{
+\[
+  \begin{array}{rrcl}
+    (\text{Base-Events}) & \varEvent[_{\mathghost}] &:=& \varEvent[_{\ctsafe}] \\
+    (\text{Spectre Variants}) & vX &:=& \operatorname{NONE} \mid \operatorname{PHT} \\
+    (\text{Security Tags}) & \varSecuritytag{} &:=& \lock_{vX} \mid \unlock\\ 
+    (\text{Events}) & \varEvent[_{\mathghost}] &:=& \varEvent[_{\mathghost}];\varSecuritytag{} \mid \emptyevent \mid \lightning \mid \ev{Spec} \mid \ev{Rlb} \mid \ev{Barrier}\\ 
+  \end{array}
+\]
+}
+
+Even though the considered Spectre variants are just SPECTRE-PHT~\cite{kocher2019spectre}, NONE just describes secret data as in \gls*{scct} (see \Cref{subsec:scct:tracemodel}), the trace model is general enough to allow for potential future extension with different variants~\cite{kocher2019spectre,maisuradze2018ret2spec,horn2019zero}.
+%For sake of readability, this paper just uses the notation $\lock$ in place of $\lock_{\text{PHT}}$.
+
+\subsubsection{Speculative Safety}
+
+\gls*{ss}~\cite{patrignani2021exorcising}, similar to \gls*{scct}, is a sound overapproximation of a variant of noninterference.
+
+\begin{definition}{\definitionlabel[\glsfirst{ss}]{def-trace-ssdef}}
+  \noindent
+
+  \begin{nscenter}
+    \[
+      \sssafe := \left\{
+        \varTrace[_{\mathghost}] 
+          \left|
+            \begin{array}{l}
+              \varTrace[_{\mathghost}]=\hole{\cdot} 
+                \text{ or } 
+              \exists\varTrace[_{\mathghost}'].\\
+              %
+              \left(\varTrace[_{\mathghost}]=\varEvent[_{\mathghost}];\unlock\cdot\varTrace[_{\mathghost}'] 
+                \text{or }
+              \varTrace[_{\mathghost}]=\varEvent[_{\mathghost}];\lock_{\text{NONE}}\cdot\varTrace[_{\mathghost}']\right)\\
+              %
+              \text{and }%
+              \ \varTrace[_{\mathghost}']\in\sssafe
+            \end{array}
+          \right.
+        \right\}
+    \] 
+  \end{nscenter}
+\end{definition}
+The technical setup so far leads to the above definition, where only locks annotated with $\text{SPECTRE-PHT}$ are disallowed to occur on the trace.
+That way, programs attaining \gls*{ss} do not necessarily attain \gls*{scct}.
+
+\subsubsection{Speculation Memory Safety}\label{sec:spec-ms-rel}
+
+As before, we need to relate the different trace models with each other, so that the memory safety property without speculation can be lifted to speculation. 
+To this end, let $\sim_{\mathghost}: \varEvent[_{\mathghost}] \times \varEvent[_{\ctsafe}]$ be a cross-language trace relation whose key cases are below.
+The intuition is that \gls*{ss} is trivially satisfied in \gls*{scct}, since speculation is inexpressible there, which amounts to dropping events $\ev{Spec}$, $\ev{Rlb}$, or $\ev{Barrier}$, as well as all base events tagged with $\lock_{\text{PHT}}$.
+
+\[
+  \typerulenolabel{}{\varEvent[_{\ctsafe}];\lock\sim_{\mathghost}\varEvent[_{\ctsafe}];\lock_{\text{NONE}}}
+  \hspace{2ex}
+  \typerulenolabel{}{\emptyevent\sim_{\mathghost}\varEvent[_{\mathghost}];\lock_{\text{PHT}}}
+  \hspace{2ex}
+  \typerulenolabel{}{\varEvent[_{\ctsafe}];\unlock\sim_{\mathghost}\varEvent[_{\ctsafe}];\unlock}
+  \hspace{2ex}
+  \typerulenolabel{}{\lightning\sim_{\mathghost}\lightning}
+  \hspace{2ex}
+  \typerulenolabel{}{\emptyevent\sim_{\mathghost}\ev{Spec}}
+  \hspace{2ex}
+  \typerulenolabel{}{\emptyevent\sim_{\mathghost}\ev{Rlb}}
+  \hspace{2ex}
+  \typerulenolabel{}{\emptyevent\sim_{\mathghost}\ev{Barrier}}
+\]
+
+We conclude by defining the ultimate property of interest for secure compilers: \gls*{specms}.
+\begin{definition}{\definitionlabel[\glsfirst{specms}]{def-trace-specmsdef}}
+  \[
+  \specmssafe := \msscctsafe\cap\sigma_{\sim_{\mathghost}}\left(\sssafe\right)
+  \]
+\end{definition}
+
+\subsubsection{Monitors}\label{subsubsec-cs-properties-monitors}
 

### 2024-10-11 tech/main.tex
tech

@@
 
+For simple memory safety composed of temporal and spatial memory safety, the trace model defines events ($\varEvent[_{\mssafe}]$) as either the empty event ($\emptyevent$), a crash ($\lightning$), or a base-event ($\varEvent[_{\mssafe}]$).
+
+\begin{definition}{\definitionlabel[\glsfirst{ms} Trace Model]{mstraces}}
+\[
+  \begin{array}{rcll}
+    \text{(Base-Events)}&\varEvent[_{\mssafe}] &:=& \msAlloc \mid \msDealloc \mid \msUse \\
+    \text{(Events)}&\varEvent[_{\mssafe}] &:=& \varEvent[_{\mssafe}] \mid \emptyevent \mid \lightning \\ 
+  \end{array}
+\]
+\end{definition}
+
+Base-events describe the actual kind of event that happened.
+For the basic memory-safety properties, these are three variants:
+First, the allocation event ($\msAlloc$) that fires whenever a program claims $n$ cells of memory and stores them at address $\varLoc$, where addresses are assumed to be unique.
+Second, deallocation ($\msDealloc$) announces that the object at location $\varLoc$ is freed.
+Third, an event to describe reads from and writes to the $n$-th memory cell from address $\varLoc$ ($\msUse$).
+
 \paragraph{Temporal Memory Safety}
@@
 \gls*{tms}~\cite{nagarakatte2010cets} is a safety property that describes that an unallocated object must not be (re-)used.
+Since it does not care about bounds of an object in memory, we emphasize this fact by deviating from the unified, memory-safety trace model and instead introduce traces for temporal memory safety:
+\begin{definition}{\definitionlabel[\glsfirst{tms} Traces]{tmstraces}}
+\[
+  \begin{array}{rcll}
+    \text{(Base-Events)}&\tmsEvent &:=& \tmsAlloc \mid \tmsDealloc \mid \tmsUse \\
+    \text{(Events)}&\tmsEvent &:=& \tmsEvent \mid \emptyevent \mid \lightning \\ 
+  \end{array}
+\]
+\end{definition}
 
-\begin{definition}{\definitionlabel[\glsfirst{tms}]{def-trace-tmsdef}}
+With this, \gls*{tms} is defined as follows.
+\begin{definition}{\definitionlabel[\glsfirst{tms}]{tmsdef}}
   \[
-  \tmssafe:=\left\{\varTrace[_{\mssafe}]
+  \tmssafe:=\left\{\tmsTrace
     \left| 
       \begin{array}{rcl}
-        \ev{Alloc\ \varLoc\ n}&\le_{\varTrace[_{\mssafe}]}&\ev{Dealloc\ \varLoc} \\
-        \ev{Use\ \varLoc\ n}&\le_{\varTrace[_{\mssafe}]}&\ev{Dealloc\ \varLoc} \\
-        \text{at most one }\ev{Dealloc\ \varLoc}&\text{in}&\varTrace[_{\mssafe}] \\
-        \text{at most one }\ev{Alloc\ \varLoc\ n}&\text{in}&\varTrace[_{\mssafe}] \\
+        \tmsAlloc&\le_{\tmsTrace}&\tmsDealloc \\
+        \tmsUse&\le_{\tmsTrace}&\tmsDealloc \\
       \end{array}
@@
 \gls*{sms}~\cite{nagarakatte2009soft} is a safety property that prohibits out-of-bounds accesses.
+For emphasis, we also use a custom trace-model for spatial memory safety.
+\begin{definition}{\definitionlabel[\glsfirst{sms} Traces]{smstraces}}
+\[
+  \begin{array}{rcll}
+    \text{(Base-Events)}&\smsEvent &:=& \smsAlloc \mid \smsDealloc \mid \smsUse \\
+    \text{(Events)}&\smsEvent &:=& \smsEvent \mid \emptyevent \mid \lightning \\ 
+  \end{array}
+\]
+\end{definition}
 
-\begin{definition}{\definitionlabel[\glsfirst{sms}]{def-trace-smsdef}}
+With this, \gls*{sms} is defined as follows.
+\begin{definition}{\definitionlabel[\glsfirst{sms}]{smsdef}}
   \[
-  \smssafe:=\left\{\varTrace[_{\mssafe}] \left|\begin{array}{rcl}
-      \text{If }\ev{Alloc\ \varLoc\ n}\le_{\varTrace_{\mssafe}}\ev{Use\ \varLoc\ m}, \text{ then }m<n
+  \smssafe:=\left\{\smsTrace \left|\begin{array}{rcl}
+      \text{If }\smsAlloc\le_{\smsTrace}\smsUse, \text{ then }m<n
   \end{array}\right.\right\}
@@
 
-In spirit of earlier work~\cite{nagarakatte2009soft,nagarakatte2010cets,jim2002cyclone,necula2005ccured,michael2023mswasm}, full \gls*{ms} is the intersection of \Cref{def:trace:tmsdef,def:trace:smsdef}.
+In spirit of earlier work~\cite{nagarakatte2009soft,nagarakatte2010cets,jim2002cyclone,necula2005ccured,michael2023mswasm}, full \gls*{ms} is the intersection of \definitionref{tmsdef} and \definitionref{smsdef}.
+However, both properties use a different trace model, so intersecting them alone would trivially yield the empty set.
+To remedy this, we define cross-language trace relations that lift both \definitionref{tmstraces} and \definitionref{smstraces} to \definitionref{mstraces}.
+\begin{definition}{\definitionlabel[Mapping \glsfirst{ms} to \glsfirst{tms}]{mapping-tms-ms}}
+  \newcommand{\mmap}[2]{{#1}\;{\xlangrel{\mssafe}{\tmssafe}}\;{#2}}
+  \begin{center}
+    \judgbox{\mmap{\msEvent}{\tmsEvent}}{,,Translate \gls*{ms}-events to \gls*{tms}-events.''}
 
-\begin{definition}{\definitionlabel[\glsfirst{ms}]{def-trace-msdef}}
+    \typerule{mapping-ms-tms-empty}{}{
+      \mmap{\msev{\emptyevent}}{\tmsev{\emptyevent}}
+    }{mapping-ms-tms-empty}
+    %
+    \typerule{mapping-ms-tms-crash}{}{
+      \mmap{\msev{\lightning}}{\tmsev{\lightning}}
+    }{mapping-ms-tms-crash}
+    %
+    \typerule{mapping-ms-tms-alloc}{}{
+      \mmap{\msAlloc}{\tmsAlloc}
+    }{mapping-ms-tms-alloc}
+    %
+    \typerule{mapping-ms-tms-dealloc}{}{
+      \mmap{\msDealloc}{\tmsDealloc}
+    }{mapping-ms-tms-dealloc}
+    %
+    \typerule{mapping-ms-tms-use}{}{
+      \mmap{\msUse}{\tmsUse}
+    }{mapping-ms-tms-use}
+  \end{center}
+  \begin{center}
+    \judgbox{\mmap{\msTrace}{\tmsTrace}}{,,Translate \gls*{ms}-traces to \gls*{tms}-traces.''}
+
+    \typerule{mapping-ms-tms-empty}{}{
+      \mmap{\emptyTrace}{\emptyTrace}
+    }{mapping-ms-tms-empty}
+    %
+    \typerule{mapping-ms-tms-cons}{}{
+      \mmap{\consTrace{\msEvent}{\msTrace}}{\consTrace{\tmsEvent}{\tmsTrace}}
+    }{mapping-ms-tms-cons}
+  \end{center}
+\end{definition}
+\begin{definition}{\definitionlabel[Mapping \glsfirst{ms} to \glsfirst{sms}]{mapping-sms-ms}}
+  \newcommand{\mmap}[2]{{#1}\;{\xlangrel{\mssafe}{\smssafe}}\;{#2}}
+  \begin{center}
+    \judgbox{\mmap{\msEvent}{\smsEvent}}{,,Translate \gls*{ms}-events to \gls*{sms}-events.''}
+
+    \typerule{mapping-ms-sms-empty}{}{
+      \mmap{\msev{\emptyevent}}{\smsev{\emptyevent}}
+    }{mapping-ms-sms-empty}
+    %
+    \typerule{mapping-ms-sms-crash}{}{
+      \mmap{\msev{\lightning}}{\smsev{\lightning}}
+    }{mapping-ms-sms-crash}
+    %
+    \typerule{mapping-ms-sms-alloc}{}{
+      \mmap{\msAlloc}{\smsAlloc}
+    }{mapping-ms-sms-alloc}
+    %
+    \typerule{mapping-ms-sms-dealloc}{}{
+      \mmap{\msDealloc}{\smsDealloc}
+    }{mapping-ms-sms-dealloc}
+    %
+    \typerule{mapping-ms-sms-use}{}{
+      \mmap{\msUse}{\smsUse}
+    }{mapping-ms-sms-use}
+  \end{center}
+  \begin{center}
+    \judgbox{\mmap{\msTrace}{\smsTrace}}{,,Translate \gls*{ms}-traces to \gls*{sms}-traces.''}
+
+    \typerule{mapping-ms-sms-empty}{}{
+      \mmap{\emptyTrace}{\emptyTrace}
+    }{mapping-ms-sms-empty}
+    %
+    \typerule{mapping-ms-sms-cons}{}{
+      \mmap{\consTrace{\msEvent}{\msTrace}}{\consTrace{\smsEvent}{\smsTrace}}
+    }{mapping-ms-sms-cons}
+  \end{center}
+\end{definition}
+
+\begin{definition}{\definitionlabel[\glsfirst{ms}]{msdef}}
   \[
-  \mssafe:=\tmssafe \cap \smssafe
+    \mssafe:=\mapUniversal{\xlangrel{\mssafe}{\tmssafe}}{\tmssafe} \cap \mapUniversal{\xlangrel{\mssafe}{\smssafe}}{\smssafe}
   \]
@@
 
-Note that \Cref{def:trace:msdef} ignores data isolation, so there may still be memory-safety issues introduced by side-channels.
+Note that \definitionref{msdef} ignores data isolation, so there may still be memory-safety issues introduced by side-channels.
 
@@
 However, \gls*{ms} uses a different trace model than \gls*{scct}, so intersecting them would trivially yield the empty set. 
-To remedy this issue, we introduce $\sim_{\ctsafe}: \preevent[_{\ctsafe}] \times \preevent[_{\mssafe}] $, a cross-language trace relation (whose key cases are presented below), that we use to intuitively unify the trace model in which the two properties are expressed:
+To remedy this issue, we introduce $\sim_{\ctsafe}: \varEvent[_{\ctsafe}] \times \varEvent[_{\mssafe}] $, a cross-language trace relation (whose key cases are presented below), that we use to intuitively unify the trace model in which the two properties are expressed:
 \[
@@
 \begin{definition}{\definitionlabel[\glsfirst{ss}]{def-trace-ssdef}}
-  \noindent
-
-  \begin{nscenter}
   \[
@@
   \] 
-  \end{nscenter}
 \end{definition}

### 2024-10-14 tech/main.tex
tech

@@
 
+\begin{definition}{\definitionlabel[\glsfirst{scct} Traces]{sccttraces}}
+\[
+  \begin{array}{rcll}
+    \text{(Secrecy Tags)}&\varSecuritytag &:=& \unlock \mid \lock \\
+    \text{(Events)}&\tmsEvent &:=& \scctAny \mid \emptyevent \mid \lightning \\ 
+  \end{array}
+\]
+\end{definition}
+
 \gls*{cct} is a hypersafety property~\cite{barthe2018sec} and, thus, difficult to check with monitors.
@@
 
-\begin{definition}{\definitionlabel[\glsfirst{scct}]{def-trace-scctdef}}
+\begin{definition}{\definitionlabel[\glsfirst{scct}]{scctdef}}
   \noindent\[
-  \scctsafe:=\left\{\varTrace[_{\ctsafe}] 
+  \scctsafe:=\left\{\scctTrace 
       \left|
         \begin{array}{l}
-          \varTrace[_{\ctsafe}]=\hole{\cdot} \text{ or } \\
-          \exists\varTrace[_{\ctsafe}'],\varTrace[_{\ctsafe}]=\varEvent[_{\ctsafe}];\unlock\cdot\varTrace[_{\ctsafe}'] \wedge \varTrace[_{\ctsafe}']\in\scctsafe
+          \scctTrace=\hole{\cdot} \text{ or } \scctTrace=\lightning \\
+          \exists\scctTrace['],\left(\scctTrace=\scctAny[\unlock]\cdot\varTrace[_{\ctsafe}'] \text{ or } \right.\\
+          \hspace{2.8em}\left.\scctTrace=\emptyevent\cdot\scctTrace[']
+          \wedge \varTrace[_{\ctsafe}']\in\scctsafe \right)
         \end{array}
@@
 Moreover, programs that have been compiled with FaCT~\cite{cauligi2019fact} and run with a ``data independent timing mode''~\cite{arm-refman,intel-refman} enabled do not leak secrets. 
+We'd like to emphasize that our goal is to motivate the core theory presented in \Cref{sec:rtpc} and not develop a new programming language.
 
@@
 However, \gls*{ms} uses a different trace model than \gls*{scct}, so intersecting them would trivially yield the empty set. 
-To remedy this issue, we introduce $\sim_{\ctsafe}: \varEvent[_{\ctsafe}] \times \varEvent[_{\mssafe}] $, a cross-language trace relation (whose key cases are presented below), that we use to intuitively unify the trace model in which the two properties are expressed:
+To remedy this issue, we introduce $\xlangrel{\scctsafe}{\mssafe} : \scctEvent \to \msEvent \to \mathbb{P}$, a cross-language trace relation (whose key cases are presented below), that we use to intuitively unify the trace model in which the two properties are expressed:
+
+\begin{definition}{\definitionlabel[Mapping \glsfirst{ms} to \glsfirst{scct}]{mapping-ms-scct}}
+  \newcommand{\mmap}[2]{{#1}\;{\xlangrel{\mssafe}{\scctsafe}}\;{#2}}
+  \begin{center}
+    \judgbox{\mmap{\msEvent}{\scctEvent}}{,,Translate \gls*{ms}-events to \gls*{scct}-events.''}
+
+    \typerule{mapping-ms-scct-empty}{}{
+      \mmap{\msev{\emptyevent}}{\scctev{\emptyevent}}
+    }{mapping-ms-scct-empty}
+    %
+    \typerule{mapping-ms-scct-crash}{}{
+      \mmap{\msev{\lightning}}{\scctev{\lightning}}
+    }{mapping-ms-scct-crash}
+    %
+    \typerule{mapping-ms-scct-alloc}{}{
+      \mmap{\msAlloc}{\scctAny[\unlock]}
+    }{mapping-ms-scct-alloc}
+    %
+    \typerule{mapping-ms-scct-dealloc}{}{
+      \mmap{\msDealloc}{\scctAny[\unlock]}
+    }{mapping-ms-scct-dealloc}
+    %
+    \typerule{mapping-ms-scct-use}{}{
+      \mmap{\msUse}{\scctAny[\unlock]}
+    }{mapping-ms-scct-use}
+  \end{center}
+  \begin{center}
+    \judgbox{\mmap{\msTrace}{\scctTrace}}{,,Translate \gls*{ms}-traces to \gls*{scct}-traces.''}
+
+    \typerule{mapping-ms-scct-empty}{}{
+      \mmap{\emptyTrace}{\emptyTrace}
+    }{mapping-ms-scct-empty}
+    %
+    \typerule{mapping-ms-scct-cons}{}{
+      \mmap{\consTrace{\msEvent}{\msTrace}}{\consTrace{\scctEvent}{\scctTrace}}
+    }{mapping-ms-scct-cons}
+  \end{center}
+\end{definition}
+
+%Essentially, $\sim_{\ctsafe}$ ignores both the new $\ev{Branch\ n}$ and $\ev{Binop\ n}$ base-events as it relates security-insensitive actions ($\unlock$) to their equivalent counterparts.
+\gls*{ms} traces trivially satisfy \gls*{scct}, since it is assumed here that they are the result of a cryptographic constant time computation.
+This may be ensured by a compiler, such as FaCT, or the semantics of the language itself.
+We also don't have any other information in \gls*{scct} traces besides the secrecy tags, since the semantics of a language can simply tag relevant events with the respective vulnerability, such as a leak of secret data.
+% 
+It is now possible to define \gls*{msscct} using the universal image:
+
+\begin{definition}{\definitionlabel[\glsfirst{ms} and \glsfirst{scct}]{msscctdef}}
   \[
-  \typerulenolabel{}{\varEvent[_{\mssafe}]\sim_{\ctsafe}\varEvent[_{\mssafe}];\unlock{}}
-  \hspace{2ex}
-  \typerulenolabel{}{\lightning\sim_{\ctsafe}\lightning}
-  \hspace{2ex}
-  \typerulenolabel{}{\emptyevent\sim_{\ctsafe}\ev{Branch\;n};\varSecuritytag{}}
-  \hspace{2ex}
-  \typerulenolabel{}{\emptyevent\sim_{\ctsafe}\ev{Binop\;n};\varSecuritytag{}}
-\]
-
-Essentially, $\sim_{\ctsafe}$ ignores both the new $\ev{Branch\ n}$ and $\ev{Binop\ n}$ base-events as it relates security-insensitive actions ($\unlock$) to their equivalent counterparts.
-Thus, \gls*{ms} traces trivially satisfy \gls*{scct}.
-% 
-The above relation is extended point-wise to traces, skipping the empty event $\emptyevent$ on either side, and it is now possible to define \gls*{msscct} using the universal image:
-
-\begin{definition}{\definitionlabel[\glsfirst{ms} and \glsfirst{scct}]{def-trace-msscctdef}}
-  \[
-  \msscctsafe:=\mssafe\cap\sigma_{\sim_{\ctsafe}}\left(\scctsafe\right)
+    \msscctsafe:=\mssafe\cap\mapUniversal{\xlangrel{\mssafe}{\scctsafe}}{\scctsafe}
   \]
@@
 So far, the considered trace models do not let us express speculative execution attacks such as Spectre~\cite{kocher2019spectre}. 
-For this, we extend the earlier trace model so that the security tags ($\varSecuritytag{}$) carry additional information about the kind of private data leakage, i.e., the type of speculative leak.
-Moreover, we add base-events signalling the beginning of a speculative execution ($\ev{Spec}$), a barrier ($\ev{Barrier}$) that signals that any speculative execution may not go past it, as well as a rollback event ($\ev{Rlb}$), which signals that execution resumes to where speculation started.
+For this, we extend the earlier trace model so that the secrecy tags ($\varSecuritytag{}$) carry additional information about the kind of private data leakage, i.e., the type of speculative leak.
+This tag may be emitted by the semantics of a concrete programming language, as seen in earlier work~\cite{fabian2022automatic}. 
+%Moreover, we add base-events signalling the beginning of a speculative execution ($\ev{Spec}$), a barrier ($\ev{Barrier}$) that signals that any speculative execution may not go past it, as well as a rollback event ($\ev{Rlb}$), which signals that execution resumes to where speculation started.
 
-\vspace{-1em}
+\begin{definition}{\definitionlabel[\glsfirst{ss} traces]{sstraces}}
 {
@@
   \begin{array}{rrcl}
-    (\text{Base-Events}) & \varEvent[_{\mathghost}] &:=& \varEvent[_{\ctsafe}] \\
-    (\text{Spectre Variants}) & vX &:=& \operatorname{NONE} \mid \operatorname{PHT} \\
-    (\text{Security Tags}) & \varSecuritytag{} &:=& \lock_{vX} \mid \unlock\\ 
-    (\text{Events}) & \varEvent[_{\mathghost}] &:=& \varEvent[_{\mathghost}];\varSecuritytag{} \mid \emptyevent \mid \lightning \mid \ev{Spec} \mid \ev{Rlb} \mid \ev{Barrier}\\ 
+    (\text{Spectre Variants}) & vX &:=& \operatorname{NONE} \mid \operatorname{PHT} \mid \dots \\
+    (\text{Secrecy Tags}) & \varSecuritytag{} &:=& \lock_{vX} \mid \unlock\\ 
+    (\text{Events}) & \specEvent &:=& \specEvent \mid \emptyevent \mid \lightning \\ 
   \end{array}
@@
 }
+\end{definition}
 
-Even though the considered Spectre variants are just SPECTRE-PHT~\cite{kocher2019spectre}, NONE just describes secret data as in \gls*{scct} (see \Cref{subsec:scct:tracemodel}), the trace model is general enough to allow for potential future extension with different variants~\cite{kocher2019spectre,maisuradze2018ret2spec,horn2019zero}.
+%Even though the considered Spectre variants are just SPECTRE-PHT~\cite{kocher2019spectre}, NONE just describes secret data as in \gls*{scct} (see \Cref{subsec:scct:tracemodel}), the trace model is general enough to allow for potential future extension with different variants~\cite{kocher2019spectre,maisuradze2018ret2spec,horn2019zero}.
+NONE describes secret data as in \gls*{scct} (see \Cref{subsec:scct:tracemodel}).
+Any other secrecy tag, such as PHT, describes the concrete reason for the respective leak.
+The trace model in \defref{sstraces} is general enough to allow for potential future extension with different variants~\cite{kocher2019spectre,maisuradze2018ret2spec,horn2019zero}.
 %For sake of readability, this paper just uses the notation $\lock$ in place of $\lock_{\text{PHT}}$.
 
-\subsubsection{Speculative Safety}
+\subsubsection{Speculation Safety}
 
@@
 
-\begin{definition}{\definitionlabel[\glsfirst{ss}]{def-trace-ssdef}}
+\begin{definition}{\definitionlabel[\glsfirst{ss}]{ssdef}}
   \[
     \sssafe := \left\{
-      \varTrace[_{\mathghost}] 
+      \specTrace
         \left|
           \begin{array}{l}
-            \varTrace[_{\mathghost}]=\hole{\cdot} 
+            \specTrace=\hole{\cdot} 
               \text{ or } 
-            \exists\varTrace[_{\mathghost}'].\\
+            \specTrace=\lightning
+              \text{ or }
+            \exists\specTrace['].\\
             %
-            \left(\varTrace[_{\mathghost}]=\varEvent[_{\mathghost}];\unlock\cdot\varTrace[_{\mathghost}'] 
+            \left(\specTrace=\specAny[\unlock]\cdot\specTrace['] 
               \text{or }
-            \varTrace[_{\mathghost}]=\varEvent[_{\mathghost}];\lock_{\text{NONE}}\cdot\varTrace[_{\mathghost}']\right)\\
+            \varTrace[_{\mathghost}]=\specAny[\lock_{\text{NONE}}]\cdot\specTrace[']\right.\\
+            \hspace{2em}\text{or }
+            \left.\varTrace[_{\mathghost}]=\emptyevent\cdot\specTrace[']\right)\\
             %
@@
 As before, we need to relate the different trace models with each other, so that the memory safety property without speculation can be lifted to speculation. 
-To this end, let $\sim_{\mathghost}: \varEvent[_{\mathghost}] \times \varEvent[_{\ctsafe}]$ be a cross-language trace relation whose key cases are below.
-The intuition is that \gls*{ss} is trivially satisfied in \gls*{scct}, since speculation is inexpressible there, which amounts to dropping events $\ev{Spec}$, $\ev{Rlb}$, or $\ev{Barrier}$, as well as all base events tagged with $\lock_{\text{PHT}}$.
+To this end, let $\xlangrel{ms}{ss}: \msEvent \to \specEvent\to\mathbb{P}$ be a cross-language trace relation whose key cases are below.
+The intuition is that \gls*{ss} is trivially satisfied in \gls*{ms}, since speculation is inexpressible there and a language attaining \gls*{ms} by means of semantics can enforce \gls*{ss} simply by not having a speculative dynamic semantics.
+In turn, all base events tagged with $\lock_{\text{PHT}}$ are simply dropped.
 
+\begin{definition}{\definitionlabel[Mapping \glsfirst{ms} to \glsfirst{ss}]{mapping-ms-spec}}
+  \newcommand{\mmap}[2]{{#1}\;{\xlangrel{\mssafe}{\sssafe}}\;{#2}}
+  \begin{center}
+    \judgbox{\mmap{\msEvent}{\specEvent}}{,,Translate \gls*{ms}-events to \gls*{ss}-events.''}
+
+    \typerule{mapping-ms-spec-empty}{}{
+      \mmap{\msev{\emptyevent}}{\specev{\emptyevent}}
+    }{mapping-ms-spec-empty}
+    %
+    \typerule{mapping-ms-spec-crash}{}{
+      \mmap{\msev{\lightning}}{\specev{\lightning}}
+    }{mapping-ms-spec-crash}
+    %
+    \typerule{mapping-ms-spec-alloc}{}{
+      \mmap{\msAlloc}{\specAny[\unlock]}
+    }{mapping-ms-spec-alloc}
+    %
+    \typerule{mapping-ms-spec-dealloc}{}{
+      \mmap{\msDealloc}{\specAny[\unlock]}
+    }{mapping-ms-spec-dealloc}
+    %
+    \typerule{mapping-ms-spec-use}{}{
+      \mmap{\msUse}{\specAny[\unlock]}
+    }{mapping-ms-spec-use}
+  \end{center}
+  \begin{center}
+    \judgbox{\mmap{\msTrace}{\specTrace}}{,,Translate \gls*{ms}-traces to \gls*{spec}-traces.''}
+
+    \typerule{mapping-ms-spec-empty}{}{
+      \mmap{\emptyTrace}{\emptyTrace}
+    }{mapping-ms-spec-empty}
+    %
+    \typerule{mapping-ms-spec-cons}{}{
+      \mmap{\consTrace{\msEvent}{\msTrace}}{\consTrace{\specEvent}{\specTrace}}
+    }{mapping-ms-spec-cons}
+  \end{center}
+\end{definition}
+
+We conclude with the ultimate property of interest for our secure compiler: \gls*{specms}.
+\begin{definition}{\definitionlabel[\glsfirst{specms}]{specmsdef}}
   \[
-  \typerulenolabel{}{\varEvent[_{\ctsafe}];\lock\sim_{\mathghost}\varEvent[_{\ctsafe}];\lock_{\text{NONE}}}
-  \hspace{2ex}
-  \typerulenolabel{}{\emptyevent\sim_{\mathghost}\varEvent[_{\mathghost}];\lock_{\text{PHT}}}
-  \hspace{2ex}
-  \typerulenolabel{}{\varEvent[_{\ctsafe}];\unlock\sim_{\mathghost}\varEvent[_{\ctsafe}];\unlock}
-  \hspace{2ex}
-  \typerulenolabel{}{\lightning\sim_{\mathghost}\lightning}
-  \hspace{2ex}
-  \typerulenolabel{}{\emptyevent\sim_{\mathghost}\ev{Spec}}
-  \hspace{2ex}
-  \typerulenolabel{}{\emptyevent\sim_{\mathghost}\ev{Rlb}}
-  \hspace{2ex}
-  \typerulenolabel{}{\emptyevent\sim_{\mathghost}\ev{Barrier}}
-\]
-
-We conclude by defining the ultimate property of interest for secure compilers: \gls*{specms}.
-\begin{definition}{\definitionlabel[\glsfirst{specms}]{def-trace-specmsdef}}
-  \[
-  \specmssafe := \msscctsafe\cap\sigma_{\sim_{\mathghost}}\left(\sssafe\right)
+    \specmssafe := \msscctsafe\cap\mapUniversal{\xlangrel{\mssafe}{\sssafe}}{\sssafe}
   \]
@@
 
-\subsubsection{Monitors}\label{subsubsec-cs-properties-monitors}
+\subsection{Monitors}\label{subsec-cs-monitors}
+
+This part presents monitors for the properties presented in \Cref{subsec-cs-properties}.
+Since all properties are safety properties, it is enough to consider trace-prefixes, i.e., finite lists of events.
+
+\subsubsection{\gls*{tms} Monitor}\label{subsubsec-mon-tms}
+In order to check \gls*{tms} on traces mechanically, we introduce a monitor that works on more abstract monitor-actions, without any other events besides those relevant to temporal memory safety.
+
+\begin{definition}{\definitionlabel[\gls*{tms} Monitor]{monitor-tms}}
+  \begin{gather*}
+    \begin{aligned}
+      \text{(Abstract Store)}~\tmsMonitor &\isdef\{A : \mi{Locs}, F : \mi{Locs}\} \\
+      \tmsMonitor\subseteq_F \tmsMonitor['] &\text{ iff } \tmsMonitor[.F] \subseteq \tmsMonitor['.F]\\
+      \varLoc\in \tmsMonitor &\text{ iff } \varLoc \in {\tmsMonitor[.A]} \wedge \varLoc \notin {\tmsMonitor[.F]}\\
+      \varLoc\notin \tmsMonitor &\text{ iff } \varLoc \notin {\tmsMonitor[.A]} \wedge \varLoc \notin {\tmsMonitor[.F]}\\
+      \{\varLoc\}\cup \tmsMonitor &= \{A : \{\varLoc\}\cup \tmsMonitor[.A], F : \tmsMonitor[.F]\}\\
+      \tmsMonitor\setminus\{\varLoc\} &= \{A : \tmsMonitor[.A]\setminus\{\varLoc\}, F : \tmsMonitor[.F]\cup\left\{\varLoc\right\}\}\\
+      \tmsMonitor\cup \tmsMonitor' &= \{A : \tmsMonitor[.A]\cup \tmsMonitor['.A], F : \tmsMonitor[.F]\cup \tmsMonitor['.F]\}\\
+    \end{aligned}
+  \end{gather*}
+\end{definition}
+
+The reductions are defined as follows:
+\begin{definition}{\definitionlabel[\gls*{tms} Monitor Reductions]{monitor-tms-steps}}
+  \begin{center}
+    \judgbox{\monitorcheck[]{\tmsMonitor}{\tmsMonitor[']}{\tmsEvent}}{,,Given abstract state $\tmsMonitor$, step to abstract state $\tmsMonitor'$\\ emitting an abstract event $\tmsEvent$.''}
+
+    \typerule{tmsmon-empty}{}{
+      \monitorcheck[]{\tmsMonitor}{\tmsMonitor}{\emptyevent}
+    }{tmsmon-empty}
+    \typerule{tmsmon-alloc}{
+      \varLoc\not\in\tmsMonitor
+    }{
+      \monitorcheck[]{\tmsMonitor}{\{\varLoc\}\cup\tmsMonitor}{\tmsAlloc}
+    }{tmsmon-alloc}
+    \typerule{tmsmon-dealloc}{
+      \varLoc\in\tmsMonitor
+    }{
+      \monitorcheck[]{\tmsMonitor}{\tmsMonitor\setminus\{\varLoc\}}{\tmsDealloc}
+    }{tmsmon-dealloc}
+    \typerule{tmsmon-use}{
+      \varLoc\in\tmsMonitor
+    }{
+      \monitorcheck[]{\tmsMonitor}{\tmsMonitor}{\tmsUse}
+    }{tmsmon-use}
+    \typerule{tmsmon-crash}{}{
+      \monitorcheck[]{\tmsMonitor}{\tmsMonitor}{\lightning}
+    }{tmsmon-crash}
+  \end{center}
+  \begin{center}
+    \judgbox{\monitorcheck{\tmsMonitor}{\tmsMonitor[']}{\tmsEvent}}{,,Reflexive-Transitive closure of above judgement.\\$\tmsTrace$ is a list of events.''}
+
+    \typerule{tmsmon-refl}{
+      \tmsMonitor[.A]=\emptyset
+    }{
+      \monitorcheck{\tmsMonitor}{\tmsMonitor}{\hole{\cdot}}
+    }{tmsmon-refl}
+    %
+    \typerule{tmsmon-trans}{
+      \monitorcheck[]{\tmsMonitor}{\tmsMonitor[']}{\tmsEvent}&
+      \monitorcheck{\tmsMonitor[']}{\tmsMonitor['']}{\tmsTrace}&
+    }{
+      \monitorcheck{\tmsMonitor}{\tmsMonitor['']}{\tmsEvent\cdot\tmsTrace}
+    }{tms-trans}
+  \end{center}
+\end{definition}
+
+\subsubsection{\gls*{sms} Monitor}\label{subsubsec-mon-sms}
+placeholder
+\subsubsection{\gls*{ms} Monitor}\label{subsubsec-mon-ms}
+placeholder
+\subsubsection{\gls*{scct} Monitor}\label{subsubsec-mon-scct}
+placeholder
+\subsubsection{\gls*{msscct} Monitor}\label{subsubsec-mon-msscct}
+placeholder
+\subsubsection{\gls*{ss} Monitor}\label{subsubsec-mon-spec}
+placeholder
+\subsubsection{\gls*{specms} Monitor}\label{subsubsec-mon-specms}
+placeholder
+
+%%%%%%%%%%%%%
+
+\subsection{$\src{RusT}$- A Temporal Memory Safe Language}\label{subsec-rust}
+placeholder
+\subsubsection{Syntax}\label{subsec-rust-syntax}
+placeholder
+\subsubsection{Static Semantics}\label{subsec-rust-static}
+placeholder
+\subsubsection{Dynamic Semantics}\label{subsec-rust-dynamic}
+placeholder
+\subsubsection{Properties}\label{subsec-rust-properties}
+placeholder
+
+%%%%%%%%%%%%%
+
+\subsection{$\irl{CePeEs}$- An Intermediate Representation}\label{subsec-cepees}
+placeholder
+\subsubsection{Syntax}\label{subsec-cepees-syntax}
+placeholder
+\subsubsection{Static Semantics}\label{subsec-cepees-static}
+placeholder
+\subsubsection{Dynamic Semantics}\label{subsec-cepees-dynamic}
+placeholder
+\subsubsection{Properties}\label{subsec-cepees-properties}
+placeholder
+
+%%%%%%%%%%%%%
+
+\subsection{Compiler from $\src{RusT}$ to $\irl{CePeEs}$}\label{subsec-rust-to-cepees}
+placeholder
+\subsubsection{Secure Compiler}\label{subsec-rust-to-cepees-seccomp}
+placeholder
+
+\subsection{Bounds-Check Instrumentation in $\irl{CePeEs}$}\label{subsec-softbounds-in-cepees}
+placeholder
+\subsubsection{Secure Compiler}\label{subsec-softbounds-cepees-seccomp}
+placeholder
+
+\subsection{Improved Compiler from $\src{RusT}$ to $\irl{CePeEs}$}\label{subsec-improved-rust-to-cepees}
+placeholder
+\subsubsection{Secure Compiler}\label{subsec-improved-rust-to-cepees-seccomp}
+placeholder
+
+\subsection{Dead Code Elimination in $\irl{CePeEs}$}\label{subsec-dce-in-cepees}
+placeholder
+\subsubsection{Secure Compiler}\label{subsec-dce-cepees-seccomp}
+placeholder
+
+\subsection{Constant Folding in $\irl{CePeEs}$}\label{subsec-cf-in-cepees}
+placeholder
+\subsubsection{Secure Compiler}\label{subsec-cf-cepees-seccomp}
+placeholder
+
+\subsection{Optimising Compiler from $\src{RusT}$ to $\irl{CePeEs}$}\label{subsec-optimising-rust-to-cepees}
+placeholder
+\subsubsection{Secure Compiler}\label{subsec-optimising-rust-to-cepees-seccomp}
+placeholder
+
+%%%%%%%%%%%%%
+
+\subsection{$\obj{A}$ - An Assembly-Like Target}\label{subsec-a}
+placeholder
+\subsubsection{Syntax}\label{subsec-a-syntax}
+placeholder
+\subsubsection{Static Semantics}\label{subsec-a-static}
+placeholder
+\subsubsection{Dynamic Semantics}\label{subsec-a-dynamic}
+placeholder
+\subsubsection{Properties}\label{subsec-a-properties}
+placeholder
+
+%%%%%%%%%%%%
+
+\subsection{Compiler from $\irl{CePeEs}$ to $\obj{A}$}\label{subsec-cepees-to-a}
+placeholder
+\subsubsection{Secure Compiler}\label{subsec-cepees-to-a-seccomp}
+placeholder
+
+\subsection{Compiler from $\src{RusT}$ to $\obj{A}$}\label{subsec-rust-to-a}
+placeholder
+\subsubsection{Secure Compiler}\label{subsec-rust-to-a-seccomp}
+placeholder
+
+%%%%%%%%%%%%
+
+\subsection{$\ird{SpA}$ - An Assembly-Like Target with Speculation}\label{subsec-spa}
+placeholder
+\subsubsection{Syntax}\label{subsec-spa-syntax}
+placeholder
+\subsubsection{Static Semantics}\label{subsec-spa-static}
+placeholder
+\subsubsection{Dynamic Semantics}\label{subsec-spa-dynamic}
+placeholder
+\subsubsection{Properties}\label{subsec-spa-properties}
+placeholder
+
+%%%%%%%%%%%%
+
+\subsection{Compiler from $\obj{A}$ to $\ird{SpA}$}\label{subsec-a-to-spa}
+placeholder
+\subsubsection{Secure Compiler}\label{subsec-a-to-spa-seccomp}
+placeholder
+
+\subsection{Compiler from $\src{RusT}$ to $\ird{SpA}$}\label{subsec-rust-to-spa}
+placeholder
+\subsubsection{Secure Compiler}\label{subsec-rust-to-spa-seccomp}
+placeholder
 

### 2024-10-14 tech/main.tex
tech

@@
   \begin{array}{rcll}
-    \text{(Base-Events)}&\smsEvent &:=& \smsAlloc \mid \smsDealloc \mid \smsUse \\
+    \text{(Base-Events)}&\smsEvent &:=& \smsAlloc \mid \smsUse \\
     \text{(Events)}&\smsEvent &:=& \smsEvent \mid \emptyevent \mid \lightning \\ 
@@
     \typerule{mapping-ms-sms-dealloc}{}{
-      \mmap{\msDealloc}{\smsDealloc}
+      \mmap{\msDealloc}{\emptyevent}
     }{mapping-ms-sms-dealloc}
     %
-    \typerule{mapping-ms-sms-use}{}{
+    \typerule{mapping-ms-sms-use}{
+      n=m
+    }{
       \mmap{\msUse}{\smsUse}
@@
   \begin{center}
-    \judgbox{\mmap{\msTrace}{\specTrace}}{,,Translate \gls*{ms}-traces to \gls*{spec}-traces.''}
+    \judgbox{\mmap{\msTrace}{\specTrace}}{,,Translate \gls*{ms}-traces to \gls*{ss}-traces.''}
 
@@
 \subsubsection{\gls*{tms} Monitor}\label{subsubsec-mon-tms}
-In order to check \gls*{tms} on traces mechanically, we introduce a monitor that works on more abstract monitor-actions, without any other events besides those relevant to temporal memory safety.
 
@@
     \typerule{tmsmon-trans}{
-      \monitorcheck[]{\tmsMonitor}{\tmsMonitor[']}{\tmsEvent}&
-      \monitorcheck{\tmsMonitor[']}{\tmsMonitor['']}{\tmsTrace}&
+      \monitorcheck[]{\tmsMonitor}{\tmsMonitor[']}{\tmsEvent} \rulesep
+      \monitorcheck{\tmsMonitor[']}{\tmsMonitor['']}{\tmsTrace}
     }{
@@
 
+\begin{definition}{\definitionlabel[Trace is \gls*{tms} via Monitor]{trace-is-tms-via-mon}}
+  \begin{align*}
+    \tmssafe\vdash\tmsTrace &\isdef \exists\tmsMonitor,\monitorcheck{\emptyset}{\tmsMonitor}{\tmsTrace}
+  \end{align*}
+\end{definition}
+
+\begin{lemma}{\lemmalabel[Monitor \gls*{tms} is $\tmssafe$]{monitor-tms-is-tms}}
+  If
+  \begin{assumptions}
+    \asm{monitor-tms-is-tms}{\tmssafe\vdash\tmsTrace}
+  \end{assumptions}
+  then
+  \begin{goals}
+    \goal{monitor-tms-is-tms}{\tmsTrace\in\tmssafe}
+  \end{goals}
+\end{lemma}
+\begin{proof}
+  \incompleteProof
+\end{proof}
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+
 \subsubsection{\gls*{sms} Monitor}\label{subsubsec-mon-sms}
-placeholder
+\begin{definition}{\definitionlabel[\gls*{sms} Monitor]{monitor-sms}}
+  \begin{gather*}
+    \begin{aligned}
+      \text{(Abstract Store)}~\smsMonitor &\isdef \mi{Locs}\times\mathbb{N} \\
+      \operatorname{dom}\lrpars{\smsMonitor}&=\{\varLoc \mid \forall n, \lrpars{\varLoc;n}\in\smsMonitor\}
+    \end{aligned}
+  \end{gather*}
+\end{definition}
+
+The reductions are defined as follows:
+\begin{definition}{\definitionlabel[\gls*{sms} Monitor Reductions]{monitor-sms-steps}}
+  \begin{center}
+    \judgbox{\monitorcheck[]{\smsMonitor}{\smsMonitor[']}{\smsEvent}}{,,Given abstract state $\smsMonitor$, step to abstract state $\smsMonitor'$\\ emitting an abstract event $\smsEvent$.''}
+
+    \typerule{smsmon-empty}{}{
+      \monitorcheck[]{\smsMonitor}{\smsMonitor}{\emptyevent}
+    }{smsmon-empty}
+    \typerule{smsmon-alloc}{
+      \varLoc\not\in\operatorname{dom}\lrpars{\smsMonitor}
+    }{
+      \monitorcheck[]{\smsMonitor}{\{\varLoc\}\cup\smsMonitor}{\smsAlloc}
+    }{smsmon-alloc}
+    \typerule{smsmon-use}{
+      \lrpars{\varLoc;n}\in\smsMonitor \rulesep
+      m<n
+    }{
+      \monitorcheck[]{\smsMonitor}{\smsMonitor}{\smsUse}
+    }{smsmon-use}
+    \typerule{smsmon-crash}{}{
+      \monitorcheck[]{\smsMonitor}{\smsMonitor}{\lightning}
+    }{smsmon-crash}
+  \end{center}
+  \begin{center}
+    \judgbox{\monitorcheck{\smsMonitor}{\smsMonitor[']}{\smsEvent}}{,,Reflexive-Transitive closure of above judgement.\\$\smsTrace$ is a list of events.''}
+
+    \typerule{smsmon-refl}{
+      \smsMonitor=\emptyset
+    }{
+      \monitorcheck{\smsMonitor}{\smsMonitor}{\hole{\cdot}}
+    }{smsmon-refl}
+    %
+    \typerule{smsmon-trans}{
+      \monitorcheck[]{\smsMonitor}{\smsMonitor[']}{\smsEvent} \rulesep
+      \monitorcheck{\smsMonitor[']}{\smsMonitor['']}{\smsTrace}
+    }{
+      \monitorcheck{\smsMonitor}{\smsMonitor['']}{\smsEvent\cdot\smsTrace}
+    }{smsmon-trans}
+  \end{center}
+\end{definition}
+
+\begin{definition}{\definitionlabel[Trace is \gls*{sms} via Monitor]{trace-is-sms-via-mon}}
+  \begin{align*}
+    \smssafe\vdash\smsTrace &\isdef \exists\smsMonitor,\monitorcheck{\emptyset}{\smsMonitor}{\smsTrace}
+  \end{align*}
+\end{definition}
+
+\begin{lemma}{\lemmalabel[Monitor \gls*{sms} is $\smssafe$]{monitor-sms-is-sms}}
+  If
+  \begin{assumptions}
+    \asm{monitor-sms-is-sms}{\smssafe\vdash\smsTrace}
+  \end{assumptions}
+  then
+  \begin{goals}
+    \goal{monitor-sms-is-sms}{\smsTrace\in\smssafe}
+  \end{goals}
+\end{lemma}
+\begin{proof}
+  \incompleteProof
+\end{proof}
+
+\subsubsection{Dual-Monitor}\label{subsubsec-mon-dual}
+
+\begingroup
+  \newcommand{\aM}[1][]{\varMonitor[A]#1}
+  \newcommand{\bM}[1][]{\varMonitor[B]#1}
+  \newcommand{\aE}[1][]{\varEvent[A]#1}
+  \newcommand{\bE}[1][]{\varEvent[B]#1}
+
+For sake of this section, let $A,B$ be monitors with respective monitor states $\aM$ and $\bM$.
+Then, both monitor states can be combined as follows:
+
+\begin{definition}{\definitionlabel[Dual-Monitor State]{monitor-dual}}
+  \begin{gather*}
+    \begin{aligned}
+      \text{(Abstract State)}~\dualMonitor &\isdef \aM \times \bM\\
+    \end{aligned}
+  \end{gather*}
+\end{definition}
+
+The reductions are defined as follows:
+\begin{definition}{\definitionlabel[Dual-Monitor Reductions]{monitor-dual-steps}}
+  \begin{center}
+    \typerule{dualmon-refl}{}{
+      \monitorcheck{\lrpars{\emptyset_A;\emptyset_B}}{\lrpars{\emptyset_A;\emptyset_B}}{\hole{\cdot}}
+    }{dualmon-refl}
+    %
+    \typerule{dualmon-trans-emptyA}{
+      \dualMonitor=\lrpars{\aM;\bM} \rulesep
+      \monitorcheck[]{\bM}{\bM}{\bE} \rulesep
+      \monitorcheck{\dualMonitor}{\dualMonitor[']}{\varTrace[_{A;B}]}
+    }{
+      \monitorcheck{\dualMonitor}{\dualMonitor[']}{\lrpars{\emptyevent;\bE}\cdot\varTrace[_{A;B}]}
+    }{dualmon-trans-emptyA}
+    %
+    \typerule{dualmon-trans-emptyB}{
+      \dualMonitor=\lrpars{\aM;\bM} \rulesep
+      \aE\not=\emptyevent \rulesep
+      \monitorcheck[]{\aM}{\aM}{\aE} \rulesep
+      \monitorcheck{\dualMonitor}{\dualMonitor[']}{\varTrace[_{A;B}]}
+    }{
+      \monitorcheck{\dualMonitor}{\dualMonitor}{\lrpars{\aE;\emptyevent}\cdot\varTrace[_{A;B}]}
+    }{dualmon-trans-emptyB}
+    %
+    \typerule{dualmon-trans}{
+      \lrpars{\aE;\bE}\not=\lrpars{\emptyevent;\emptyevent} \rulesep
+      \monitorcheck[]{\aM}{\aM[']}{\aE} \rulesep
+      \monitorcheck[]{\bM}{\bM[']}{\bE} \rulesep
+      \monitorcheck{\dualMonitor}{\dualMonitor[']}{\varTrace[_{A;B}]}
+    }{
+      \monitorcheck{\dualMonitor}{\dualMonitor}{\lrpars{\aE;\bE}\cdot\varTrace[_{A;B}]}
+    }{dualmon-trans}
+  \end{center}
+\end{definition}
+
+\begin{definition}{\definitionlabel[Dual-Trace Satisfaction via Monitor]{trace-is-ab-via-mon}}
+  Let $A,B$ monitors and $\varProperty$ a property,
+  \begin{align*}
+    \varProperty\vdash\varTrace[A;B] &\isdef \exists\aM\;\bM,\monitorcheck{\emptyset}{\lrpars{\aM;\bM}}{\varTrace[A;B]}
+  \end{align*}
+\end{definition}
+
+% \begin{lemma}{\lemmalabel[Monitor \gls*{sms} is $\smssafe$]{monitor-sms-is-sms}}
+%   If
+%   \begin{assumptions}
+%     \asm{monitor-sms-is-sms}{\smssafe\vdash\smsTrace}
+%   \end{assumptions}
+%   then
+%   \begin{goals}
+%     \goal{monitor-sms-is-sms}{\smsTrace\in\smssafe}
+%   \end{goals}
+% \end{lemma}
+% \begin{proof}
+%   \incompleteProof
+% \end{proof}
+\endgroup
+
 \subsubsection{\gls*{ms} Monitor}\label{subsubsec-mon-ms}
-placeholder
+Instantiate \Cref{subsubsec-mon-dual} with \Cref{subsubsec-mon-tms,subsubsec-mon-sms}. 
+What is left to give is a translation from traces $\msTrace$ to traces $\varTrace[\tmssafe;\smssafe]$.
+This translation is done as follows:
+\begin{definition}{\definitionlabel[\gls*{ms} to \gls*{tms} and \gls*{sms}]{traces-ms-to-tmssms}}
+  ...
+\end{definition}
+
 \subsubsection{\gls*{scct} Monitor}\label{subsubsec-mon-scct}

### 2024-10-15 tech/main.tex
tech

--- 2024-10-11 tech/main.tex
@@
 
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\subsection{Preliminaries}
+\input{tex/preliminaries}
+\input{tex/background/sat}
+\input{tex/background/propmaps}
+\input{tex/background/rp}
 
-\begin{definition}{\definitionlabel[Function Composition]{function-composition}}
-  $\fncompo{f}{g}\isdef \lambda x.g(f\;x)$
-\end{definition}
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\subsection{Satisfaction}
-\bul{A whole program $\varWholeProg$ satisfies a property $\varProperty$} iff \iul{$\varWholeProg$ yields a trace $\varTrace$} such that \oul{$\varTrace$ satisfies $\varProperty$}.
-
-\begin{definition}{\definitionlabel[Property Satisfaction]{prop-sat}}
-    \bul{$\sat{\varWholeProg}{\varProperty}$}
-    $\isdef$
-    if \iul{$\forall\varRuntimeTerm\ \varTrace, %
-      \progstepto{\varWholeProg}{\varRuntimeTerm}{\varTrace}$
-    },
-    then \oul{$\varTrace\in\varProperty$}.
-\end{definition}
-
-With this, \definitionref{prop-sat} can be extended as follows: for \bul{a component $\varComponent$ to robustly satisfy a property $\varProperty$}, take an \iul{attacker context $\varContext$ and link it with $\varComponent$}, \oul{the resulting whole program must satisfy $\varProperty$}.
-
-\begin{definition}{\definitionlabel[Robust Property Satisfaction]{prop-rsat}}
-  \bul{$\rsat{\varComponent}{\varProperty}$} %
-  $\isdef$ %
-  \iul{$\forall \varContext\ \varWholeProg$, %
-    if $\bopLink{\varContext}{\varComponent} = \varWholeProg$}, %
-    then \oul{$\sat{\varWholeProg}{\varProperty}$%
-  }.
-\end{definition}
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\subsection{Property Mappings}
-
-This work relies on property mappings introduced in related work~\cite{abate2021extacc}.
-
-\begin{definition}{\definitionlabel[Universal Image]{universal-image}}
-  $ 
-    \mapUniversal{\sim}{\trgProperty} := 
-      \lrbraces{\srcTrace \mid %
-        \forall \trgTrace\ldotp \text{if }\xrelTraces{\srcTrace}{\trgTrace}, \text{ then }%
-        \trgTrace\in\trgProperty
-      }
-  $
-\end{definition}
-\begin{definition}{\definitionlabel[Existential Image]{existential-image}}
-  $ 
-    \mapExistential{\sim}{\srcProperty} := 
-      \lrbraces{\trgTrace \mid %
-        \exists \srcTrace\ldotp \xrelTraces{\srcTrace}{\trgTrace} \text{ and }%
-        \srcTrace\in\srcProperty
-      }
-  $
-\end{definition}
-
-The lifting of \definitionref{existential-image} and \definitionref{universal-image} to classes of properties is standard, i.e., simply element-wise.
-
-\begin{lemma}[\Coqed]{\lemmalabel[Trace Relation distributes on Universal Image]{xtrace-distribute-universal}}
-  Let $X$ be either a trace-property or a class. Then:
-  \begin{goals}
-    \goal{xtrace-distribute-universal}{\mapUniversal{\fncompo{\sim_1}{\sim_2}}{X} = \mapUniversal{\sim_2}{\mapUniversal{\sim_1}{X}}}
-  \end{goals}
-\end{lemma}
-\begin{proof}
-  Straightforward unfolding of \definitionref{universal-image} and \definitionref{function-composition} in \goalref{xtrace-distribute-universal}.
-\end{proof}
-\begin{lemma}[\Coqed]{\lemmalabel[Trace Relation distributes on Existential Image]{xtrace-distribute-existential}}
-  Let $X$ be either a trace-property or a class. Then:
-  \begin{goals}
-    \goal{xtrace-distribute-existential}{\mapExistential{\fncompo{\sim_1}{\sim_2}}{X} = \mapExistential{\sim_2}{\mapExistential{\sim_1}{X}}}
-  \end{goals}
-\end{lemma}
-\begin{proof}
-  Straightforward unfolding of \definitionref{existential-image} and \definitionref{function-composition} in \goalref{xtrace-distribute-existential}.
-\end{proof}
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\subsection{Robust Preservation}
-
-The following definitions for secure compilers are standard~\cite{abate2019jour,abate2021extacc}.
-
-\subsubsection{Differing Trace Models}
-
-A \bul{compiler $\ccST$ for languages $\src{S}$ and $\trg{T}$ robustly preserves a $\trg{T}$-level class $\trgClass$}, if for any \rul{property $\trgProperty$ of class $\trgClass$ and programs $\srcComponent$}, where \iul{$\srcComponent$ robustly satisfies $\mapUniversal{\sim}{\trgProperty}$}, \oul{the compilation of $\srcComponent$, we have that $\ccST\lrpars{\srcComponent}$ robustly satisfies $\trgProperty$}.
-
-\begin{definition}{\definitionlabel[Robust Preservation with $\mapUniversal{\sim}{\cdot}$]{rtp-universal}}
-  \begin{center}
-    \bul{$\rtpUniversal{\ccST}{\trgClass}{\sim}$}
-    %, iff 
-    $\isdef$
-    \rul{$\forall \trgProperty\in\trgClass, \srcComponent\in\src{S},$} %
-    if \iul{$\rsat{\srcComponent}{\mapUniversal{\sim}{\trgProperty}}$},
-    then \oul{$\rsat{\ccST\lrpars{\srcComponent}}{\trgProperty}$}.
-  \end{center}
-\end{definition}
-A \bul{compiler $\ccST$ for languages $\src{S}$ and $\trg{T}$ robustly preserves a $\src{S}$-level class $\srcClass$}, if for any \rul{property $\srcProperty$ of class $\srcClass$ and programs $\srcComponent$}, where \iul{$\srcComponent$ robustly satisfies $\srcProperty$}, \oul{the compilation of $\srcComponent$, we have that $\ccST\lrpars{\srcComponent}$ robustly satisfies $\mapExistential{\sim}{\srcProperty}$}.
-
-\begin{definition}{\definitionlabel[Robust Preservation with $\mapExistential{\sim}{\cdot}$]{rtp-existential}}
-  \begin{center}
-    \bul{$\rtpExistential{\ccST}{\srcClass}{\sim}$}
-    %, iff 
-    $\isdef$
-    \rul{$\forall \srcProperty\in\srcClass, \srcComponent\in\src{S},$} %
-    if \iul{$\rsat{\srcComponent}{\srcProperty}$},
-    then \oul{$\rsat{\ccST\lrpars{\srcComponent}}{\mapExistential{\sim}{\srcProperty}}$}.
-  \end{center}
-\end{definition}
-
-\subsubsection{Same Trace Models}
-
-\begin{definition}{\definitionlabel[Robust Preservation]{rtp}}
-  \begin{center}
-    $\rtp{\ccST}{\varClass}$
-    %, iff 
-    $\isdef$
-    $\forall \varProperty\in\varClass, \srcComponent\in\src{S},$ %
-    if $\rsat{\srcComponent}{\varProperty}$,
-    then $\rsat{\ccST\lrpars{\srcComponent}}{\varProperty}$.
-  \end{center}
-\end{definition}
-
-\section{Composition of Secure Compilers}\label{sec:rtpc}
-
-This section presents the composition framework.
-
-\subsection{Same Trace Models}\label{subsec:rtpc-same-trace-models}
-
-In the setting of same trace models, composition of secure compilers (as in \defref{rtp}) is simple:
-
-\begin{lemma}[\Coqed]{\lemmalabel[Composition of RTP]{rtp-compo}}
-  If 
-  \begin{assumptions}
-    \asm{rtp-compo-comp1}{\rtp{\ccSI}{\varClass[_1]}}
-    \asm{rtp-compo-comp2}{\rtp{\ccIT}{\varClass[_2]}}
-  \end{assumptions}
-  then
-  \begin{goals}
-    \goal{rtp-compo}{\rtp{\fncompo{\ccSI}{\ccIT}}{\varClass[_1]\cap\varClass[_2]}}
-  \end{goals}
-\end{lemma}
-\begin{proof}
-  Unfold \goalref{rtp-compo} and introduce the assumptions. 
-  \begin{passumptions}
-    \asm{rtp-compo-prop}{\varProperty\in\varClass[_1]\cap\varClass[_2]}
-    \asm{rtp-compo-trg-prog}{\srcComponent\in\src{S}}
-    \asm{rtp-compo-rsat}{\rsat{\srcComponent}{\varProperty}}
-  \end{passumptions}
-  What is left to show is:
-  \begin{goals}
-    \goal{rtp-compo-rsat}{\rsat{\ccIT\lrpars{\ccSI\lrpars{\srcComponent}}}%
-                               {\varProperty}}
-  \end{goals}
-  Apply \asmref{rtp-compo-comp2} on \goalref{rtp-compo-rsat} and note that due to \asmref{rtp-compo-prop}, $\varProperty\in\varClass[_2]$ also holds.
-  \begin{goals}
-    \goal{rtp-compo-rsatc}{\rsat{\ccSI\lrpars{\srcComponent}}{\varProperty}}
-  \end{goals}
-  Similarly, apply \asmref{rtp-compo-comp1} on \goalref{rtp-compo-rsatc} and note that due to \asmref{rtp-compo-prop}, $\varProperty\in\varClass[_1]$ also holds.
-  \begin{goals}
-    \goal{rtp-compo-rsatp}{\rsat{\srcComponent}{\varProperty}}
-  \end{goals}
-  \goalref{rtp-compo-rsatp} is resolved by \asmref{rtp-compo-rsat}.
-\end{proof}
-%%%%%%%%%%%%%%%%%%%%%%%%%%
-The following corollary is a nice consequence of \lemref{rtp-compo}.
-\begin{corollary}[\Coqed]{\corollarylabel[Commutativity of RTP]{rtp-swappable}}
-  If 
-  \begin{assumptions}
-    \asm{rtp-swappable-comp1}{\rtp{\ccII}{\varClass[_1]}}
-    \asm{rtp-swappable-comp2}{\rtp{\ccII[\delta]}{\varClass[_2]}}
-  \end{assumptions}
-  then
-  \begin{goals}
-    \goal{rtp-swappable1}{\rtp{\fncompo{\ccII}{\ccII[\delta]}}{\varClass[_1]\cap\varClass[_2]}}
-    \goal{rtp-swappable2}{\rtp{\fncompo{\ccII[\delta]}{\ccII}}{\varClass[_1]\cap\varClass[_2]}}
-  \end{goals}
-\end{corollary}
-\begin{proof}
-  \goalref{rtp-swappable1} and \goalref{rtp-swappable2} are an immediate consequence based on \lemmaref{rtp-compo} with \asmref{rtp-swappable-comp1} and \asmref{rtp-swappable-comp2}, noting the commutativity of set intersection.
-\end{proof}
-
-\subsection{Different Trace Models}\label{subsec:rtpc-different-trace-models}
-
-The situation in \Cref{subsec:rtpc-same-trace-models} is idealistic in the sense that a somewhat realistic robustly preserving compiler will likely have to insert checks for when control switches from context to component.
-For example, compiling from a statically typed language into a dynamically typed one, where the static language enforces functions to only accept numbers, the compiler has to insert appropriate checks to make sure that well-typedness is robustly preserved.
-Without such checks, a context could simply supply, e.g., a pair, violating the contract that a function can only receive numbers.
-
-These checks yield a mismatch between the source-level trace and the target-level trace, rendering \definitionref{rtp} not provable.
-Instead, one would have to use \definitionref{rtp-universal} or \definitionref{rtp-existential} and use an appropriate, cross-language trace relation that precisely describes the semantic effect on traces when compiling.
-For the type-checking function interfaces example, parts of this cross-language trace relation may be sketched as:
-
-\begin{center}
-  \typerule{ex-rel-interfaces-agree}{
-    \trg{v}\text{ is a number} \rulesep
-    \src{v}\approx\trg{v}
-  }{
-    \xrelTraces{\src{Call\;\ctxtocomp\;v}}{\trg{Call\;\ctxtocomp\;v}}
-  }{ex-rel-interfaces-agree}
-  \typerule{ex-rel-interfaces-disagree}{
-    \trg{v}\text{ not a number}
-  }{
-    \xrelTraces{\src{Abort}}{\trg{Call\;\ctxtocomp\;v}\cdot\trg{Abort}}
-  }{ex-rel-interfaces-disagree}
-\end{center}
-\trref{ex-rel-interfaces-agree} ensures that both in the source and target-level execution, when the context calls into the component, the values agree and are numerical.
-
-\trref{ex-rel-interfaces-disagree} handles the case where the target-level context supplied a $\trg{v}$ that is not a number.
-Here, the backtranslation cannot construct a context that gives a well-typed program {\em and} produces the same trace as the target, since the behavior is ill-defined to begin with.
-So, the only option for the backtranslation is to immediately crash.
-This, however, is a mismatch with the target-level trace, which only aborts in the wrapper that the compiler inserted.
-
-\begin{definition}{\definitionlabel[Closedness w.r.t. Universal Image]{universal-map-wf}}
-  \[
-    \mapUniversalWF{\sim}{\trgClass} \isdef \forall \trgProperty\in\trgClass, \mapUniversal{\sim}{\trgProperty}\in\mapUniversal{\sim}{\trgClass}
-  \]
-\end{definition}
-\begin{definition}{\definitionlabel[Closedness w.r.t. Existential Image]{existential-map-wf}}
-  \[
-    \mapExistentialWF{\sim}{\srcClass} \isdef \forall \srcProperty\in\srcClass, \mapExistential{\sim}{\srcProperty}\in\mapExistential{\sim}{\srcClass}
-  \]
-\end{definition}
-
-\begin{lemma}[\Coqed]{\lemmalabel[Composition of RTP-Universal]{rtp-universal-compo}}
-  If 
-  \begin{assumptions}
-    \asm{rtp-universal-compo-wfsig}{\mapUniversalWF{\sim_2}{\trgClass[_1]}}
-    \asm{rtp-universal-compo-comp1}{\rtpUniversal{\ccSI}{\mapUniversal{\sim_2}{\trgClass[_1]}}{\sim_1}}
-    \asm{rtp-universal-compo-comp2}{\rtpUniversal{\ccIT}{\trgClass[_2]}{\sim_2}}
-  \end{assumptions}
-  then
-  \begin{goals}
-    \goal{rtp-universal-compo}{\rtpUniversal{\fncompo{\ccSI}{\ccIT}}{\trgClass[_1]\cap\trgClass[_2]}{\fncompo{\sim_1}{\sim_2}}}
-  \end{goals}
-\end{lemma}
-\begin{proof}
-  Unfold \goalref{rtp-universal-compo} and introduce the assumptions. 
-  \begin{passumptions}
-    \asm{rtp-universal-compo-prop}{\trgProperty\in\trgClass[_1]\cap\trgClass[_2]}
-    \asm{rtp-universal-compo-trg-prog}{\srcComponent\in\src{S}}
-    \asm{rtp-universal-compo-rsat}{\rsat{\srcComponent}{\mapUniversal{\fncompo{\sim_1}{\sim_2}}{\trgProperty}}}
-  \end{passumptions}
-  What is left to show is:
-  \begin{goals}
-    \goal{rtp-universal-compo-rsat}{\rsat{\ccIT\lrpars{\ccSI\lrpars{\srcComponent}}}%
-                                         {\trgProperty}}
-  \end{goals}
-  Apply \asmref{rtp-universal-compo-comp2} on \goalref{rtp-universal-compo-rsat} and note that due to \asmref{rtp-universal-compo-prop}, $\trgProperty\in\trgClass[_2]$ also holds.
-  \begin{goals}
-    \goal{rtp-universal-compo-rsatc}{\rsat{\ccSI\lrpars{\srcComponent}}{\mapUniversal{\sim_2}{\trgProperty}}}
-  \end{goals}
-  Similarly, apply \asmref{rtp-universal-compo-comp1} on \goalref{rtp-universal-compo-rsatc} and note that due to \asmref{rtp-universal-compo-prop} and \asmref{rtp-universal-compo-wfsig}, $\trgProperty\in\mapUniversal{\sim_2}{\trgClass[_1]}$ also holds.
-  \begin{goals}
-    \goal{rtp-universal-compo-rsatp}{\rsat{\srcComponent}{\mapUniversal{\sim_2}{\mapUniversal{\sim_1}{\trgProperty}}}}
-  \end{goals}
-  Note that $\mapUniversal{\sim_2}{\mapUniversal{\sim_1}{\trgProperty}}=\mapUniversal{\fncompo{\sim_1}{\sim_2}}{\trgProperty}$, and thus \goalref{rtp-universal-compo-rsatp} is resolved by \asmref{rtp-universal-compo-rsat}.
-\end{proof}
-
-\begin{lemma}[\Coqed]{\lemmalabel[Composition of RTP-Existential]{rtp-existential-compo}}
-  If 
-  \begin{assumptions}
-    \asm{rtp-existential-compo-comp1}{\rtpExistential{\ccSI}{\srcClass[_1]}{\sim_1}}
-    \asm{rtp-existential-compo-comp2}{\rtpExistential{\ccIT}{\mapExistential{\sim_1}{\srcClass[_2]}}{\sim_2}}
-  \end{assumptions}
-  then
-  \begin{goals}
-    \goal{rtp-existential-compo}{\rtpExistential{\fncompo{\ccSI}{\ccIT}}{\srcClass[_1]\cap\srcClass[_2]}{\fncompo{\sim_1}{\sim_2}}}
-  \end{goals}
-\end{lemma}
-\begin{proof}
-  Similar to \lemmaref{rtp-universal-compo}.
-\end{proof}
-
-While compiler passes can be incompatible with each other in the sense that the resulting, mapped property is useless/not what is desired, it is enough to reason about compatibility of compilers at the level of traces.
-This is what this work refers to as ,,semantic effect of a compiler''.
-Instead of having to do syntactic arguments for the interaction of several compilation passes, one only has to do them for the semantic notion, i.e., the cross-language trace relations, which are much simpler proofs by an order of magnitude.
-\MKin{
-  cite section backing up this claim
-}
-
-To conclude a similar colloray as \corollaryref{rtp-swappable}, it is of great use to show that a given cross-language trace relation has no effect on the class of security properties.
-Then, it is possible to derive the following consequences:
-
-\begin{corollary}[\Coqed]{\lemmalabel[Commutativity of RTP-Universal]{rtp-universal-swappable}}
-  If 
-  \begin{assumptions}
-    \asm{rtp-universal-compo-wfsig1}{\mapUniversalWF{\sim_1}{\irlClass[_2]}}
-    \asm{rtp-universal-compo-wfsig2}{\mapUniversalWF{\sim_2}{\irlClass[_1]}}
-    \asm{rtp-universal-compo-eq2}{\mapUniversal{\sim_2}{\irlClass[_1]}=\irlClass[_1]}
-    \asm{rtp-universal-compo-eq1}{\mapUniversal{\sim_1}{\irlClass[_2]}=\irlClass[_2]}
-    \asm{rtp-universal-swappable-comp1}{\rtpUniversal{\ccII}{\irlClass[_1]}{\sim_1}}
-    \asm{rtp-universal-swappable-comp2}{\rtpUniversal{\ccII[\delta]}{\irlClass[_2]}{\sim_2}}
-  \end{assumptions}
-  then
-  \begin{goals}
-    \goal{rtp-universal-swappable1}{\rtpUniversal{\fncompo{\ccII}{\ccII[\delta]}}{\irlClass[_1]\cap\irlClass[_2]}{\fncompo{\sim_1}{\sim_2}}}
-    \goal{rtp-universal-swappable2}{\rtpUniversal{\fncompo{\ccII[\delta]}{\ccII}}{\irlClass[_1]\cap\irlClass[_2]}{\fncompo{\sim_2}{\sim_1}}}
-  \end{goals}
-\end{corollary}
-\begin{proof}
-  \goalref{rtp-universal-swappable1} and \goalref{rtp-universal-swappable2} are an immediate consequence based on \lemmaref{rtp-universal-compo} with \asmref{rtp-universal-swappable-comp1} and \asmref{rtp-universal-swappable-comp2}, noting the commutativity of set intersection.
-\end{proof}
-
-\begin{corollary}[\Coqed]{\lemmalabel[Commutativity of RTP-existential]{rtp-existential-swappable}}
-  If 
-  \begin{assumptions}
-    \asm{rtp-existential-compo-wfsig1}{\mapExistentialWF{\sim_1}{\irlClass[_2]}}
-    \asm{rtp-existential-compo-wfsig2}{\mapExistentialWF{\sim_2}{\irlClass[_1]}}
-    \asm{rtp-existential-compo-eq2}{\mapExistential{\sim_2}{\irlClass[_1]}=\irlClass[_1]}
-    \asm{rtp-existential-compo-eq1}{\mapExistential{\sim_1}{\irlClass[_2]}=\irlClass[_2]}
-    \asm{rtp-existential-swappable-comp1}{\rtpExistential{\ccII}{\irlClass[_1]}{\sim_1}}
-    \asm{rtp-existential-swappable-comp2}{\rtpExistential{\ccII[\delta]}{\irlClass[_2]}{\sim_2}}
-  \end{assumptions}
-  then
-  \begin{goals}
-    \goal{rtp-existential-swappable1}{\rtpExistential{\fncompo{\ccII}{\ccII[\delta]}}{\irlClass[_1]\cap\irlClass[_2]}{\fncompo{\sim_1}{\sim_2}}}
-    \goal{rtp-existential-swappable2}{\rtpExistential{\fncompo{\ccII[\delta]}{\ccII}}{\irlClass[_1]\cap\irlClass[_2]}{\fncompo{\sim_2}{\sim_1}}}
-  \end{goals}
-\end{corollary}
-\begin{proof}
-  \goalref{rtp-existential-swappable1} and \goalref{rtp-existential-swappable2} are an immediate consequence based on \lemmaref{rtp-existential-compo} with \asmref{rtp-existential-swappable-comp1} and \asmref{rtp-existential-swappable-comp2}, noting the commutativity of set intersection.
-\end{proof}
+\input{tex/compo}
 
@@
 
-\subsection{Properties}\label{subsec-cs-properties}
+\input{tex/properties}
+\input{tex/monitors}
 
-The properties considered in this case-study are a mixture of trace-properties as welll as hyperproperties.
-However, to carry out proofs, all properties are overapproximated by a monitor.
+\input{tex/pls/rust}
+\input{tex/pls/cepees}
 
-\subsubsection{Formal Definition}\label{subsubsec-cs-properties-formaldef}
+%%%%%%%%%%%%%
 
-For simple memory safety composed of temporal and spatial memory safety, the trace model defines events ($\varEvent[_{\mssafe}]$) as either the empty event ($\emptyevent$), a crash ($\lightning$), or a base-event ($\varEvent[_{\mssafe}]$).
+\subsection{Compiler from $\src{RusT}$ to $\irl{CePeEs}$}\label{subsec-rust-to-cepees}
+placeholder
+\subsubsection{Secure Compiler}\label{subsec-rust-to-cepees-seccomp}
+placeholder
 
-\begin{definition}{\definitionlabel[\glsfirst{ms} Trace Model]{mstraces}}
-\[
-  \begin{array}{rcll}
-    \text{(Base-Events)}&\varEvent[_{\mssafe}] &:=& \msAlloc \mid \msDealloc \mid \msUse \\
-    \text{(Events)}&\varEvent[_{\mssafe}] &:=& \varEvent[_{\mssafe}] \mid \emptyevent \mid \lightning \\ 
-  \end{array}
-\]
-\end{definition}
+\subsection{Bounds-Check Instrumentation in $\irl{CePeEs}$}\label{subsec-softbounds-in-cepees}
+placeholder
+\subsubsection{Secure Compiler}\label{subsec-softbounds-cepees-seccomp}
+placeholder
 
-Base-events describe the actual kind of event that happened.
-For the basic memory-safety properties, these are three variants:
-First, the allocation event ($\msAlloc$) that fires whenever a program claims $n$ cells of memory and stores them at address $\varLoc$, where addresses are assumed to be unique.
-Second, deallocation ($\msDealloc$) announces that the object at location $\varLoc$ is freed.
-Third, an event to describe reads from and writes to the $n$-th memory cell from address $\varLoc$ ($\msUse$).
+\subsection{Improved Compiler from $\src{RusT}$ to $\irl{CePeEs}$}\label{subsec-improved-rust-to-cepees}
+placeholder
+\subsubsection{Secure Compiler}\label{subsec-improved-rust-to-cepees-seccomp}
+placeholder
 
-\paragraph{Temporal Memory Safety}
+\subsection{Dead Code Elimination in $\irl{CePeEs}$}\label{subsec-dce-in-cepees}
+placeholder
+\subsubsection{Secure Compiler}\label{subsec-dce-cepees-seccomp}
+placeholder
 
-\gls*{tms}~\cite{nagarakatte2010cets} is a safety property that describes that an unallocated object must not be (re-)used.
-Since it does not care about bounds of an object in memory, we emphasize this fact by deviating from the unified, memory-safety trace model and instead introduce traces for temporal memory safety:
-\begin{definition}{\definitionlabel[\glsfirst{tms} Traces]{tmstraces}}
-\[
-  \begin{array}{rcll}
-    \text{(Base-Events)}&\tmsEvent &:=& \tmsAlloc \mid \tmsDealloc \mid \tmsUse \\
-    \text{(Events)}&\tmsEvent &:=& \tmsEvent \mid \emptyevent \mid \lightning \\ 
-  \end{array}
-\]
-\end{definition}
+\subsection{Constant Folding in $\irl{CePeEs}$}\label{subsec-cf-in-cepees}
+placeholder
+\subsubsection{Secure Compiler}\label{subsec-cf-cepees-seccomp}
+placeholder
 
-With this, \gls*{tms} is defined as follows.
-\begin{definition}{\definitionlabel[\glsfirst{tms}]{tmsdef}}
-  \[
-  \tmssafe:=\left\{\tmsTrace
-    \left| 
-      \begin{array}{rcl}
-        \tmsAlloc&\le_{\tmsTrace}&\tmsDealloc \\
-        \tmsUse&\le_{\tmsTrace}&\tmsDealloc \\
-      \end{array}
-    \right.
-  \right\}
-  \]
-\end{definition}
-Hereby, the notation $\varEvent[_{1}]\le_{\varTrace}\varEvent[_{2}]$ means that if $\varEvent[_{1}]$ is in $\varTrace$ and if $\varEvent[_{2}]$ is in $\varTrace$, then $\varEvent[_{1}]$ appears before $\varEvent[_{2}]$.
+\subsection{Optimising Compiler from $\src{RusT}$ to $\irl{CePeEs}$}\label{subsec-optimising-rust-to-cepees}
+placeholder
+\subsubsection{Secure Compiler}\label{subsec-optimising-rust-to-cepees-seccomp}
+placeholder
 
-\paragraph{Spatial Memory Safety}
-\gls*{sms}~\cite{nagarakatte2009soft} is a safety property that prohibits out-of-bounds accesses.
-For emphasis, we also use a custom trace-model for spatial memory safety.
-\begin{definition}{\definitionlabel[\glsfirst{sms} Traces]{smstraces}}
-\[
-  \begin{array}{rcll}
-    \text{(Base-Events)}&\smsEvent &:=& \smsAlloc \mid \smsDealloc \mid \smsUse \\
-    \text{(Events)}&\smsEvent &:=& \smsEvent \mid \emptyevent \mid \lightning \\ 
-  \end{array}
-\]
-\end{definition}
+\input{tex/pls/a}
 
-With this, \gls*{sms} is defined as follows.
-\begin{definition}{\definitionlabel[\glsfirst{sms}]{smsdef}}
-  \[
-  \smssafe:=\left\{\smsTrace \left|\begin{array}{rcl}
-      \text{If }\smsAlloc\le_{\smsTrace}\smsUse, \text{ then }m<n
-  \end{array}\right.\right\}
-  \]
-\end{definition}
+\subsection{Compiler from $\irl{CePeEs}$ to $\obj{A}$}\label{subsec-cepees-to-a}
+placeholder
+\subsubsection{Secure Compiler}\label{subsec-cepees-to-a-seccomp}
+placeholder
 
-\paragraph{Memory Safety}
+\subsection{Compiler from $\src{RusT}$ to $\obj{A}$}\label{subsec-rust-to-a}
+placeholder
+\subsubsection{Secure Compiler}\label{subsec-rust-to-a-seccomp}
+placeholder
 
-In spirit of earlier work~\cite{nagarakatte2009soft,nagarakatte2010cets,jim2002cyclone,necula2005ccured,michael2023mswasm}, full \gls*{ms} is the intersection of \definitionref{tmsdef} and \definitionref{smsdef}.
-However, both properties use a different trace model, so intersecting them alone would trivially yield the empty set.
-To remedy this, we define cross-language trace relations that lift both \definitionref{tmstraces} and \definitionref{smstraces} to \definitionref{mstraces}.
-\begin{definition}{\definitionlabel[Mapping \glsfirst{ms} to \glsfirst{tms}]{mapping-tms-ms}}
-  \newcommand{\mmap}[2]{{#1}\;{\xlangrel{\mssafe}{\tmssafe}}\;{#2}}
-  \begin{center}
-    \judgbox{\mmap{\msEvent}{\tmsEvent}}{,,Translate \gls*{ms}-events to \gls*{tms}-events.''}
+\input{tex/pls/spa}
 
-    \typerule{mapping-ms-tms-empty}{}{
-      \mmap{\msev{\emptyevent}}{\tmsev{\emptyevent}}
-    }{mapping-ms-tms-empty}
-    %
-    \typerule{mapping-ms-tms-crash}{}{
-      \mmap{\msev{\lightning}}{\tmsev{\lightning}}
-    }{mapping-ms-tms-crash}
-    %
-    \typerule{mapping-ms-tms-alloc}{}{
-      \mmap{\msAlloc}{\tmsAlloc}
-    }{mapping-ms-tms-alloc}
-    %
-    \typerule{mapping-ms-tms-dealloc}{}{
-      \mmap{\msDealloc}{\tmsDealloc}
-    }{mapping-ms-tms-dealloc}
-    %
-    \typerule{mapping-ms-tms-use}{}{
-      \mmap{\msUse}{\tmsUse}
-    }{mapping-ms-tms-use}
-  \end{center}
-  \begin{center}
-    \judgbox{\mmap{\msTrace}{\tmsTrace}}{,,Translate \gls*{ms}-traces to \gls*{tms}-traces.''}
+%%%%%%%%%%%%
 
-    \typerule{mapping-ms-tms-empty}{}{
-      \mmap{\emptyTrace}{\emptyTrace}
-    }{mapping-ms-tms-empty}
-    %
-    \typerule{mapping-ms-tms-cons}{}{
-      \mmap{\consTrace{\msEvent}{\msTrace}}{\consTrace{\tmsEvent}{\tmsTrace}}
-    }{mapping-ms-tms-cons}
-  \end{center}
-\end{definition}
-\begin{definition}{\definitionlabel[Mapping \glsfirst{ms} to \glsfirst{sms}]{mapping-sms-ms}}
-  \newcommand{\mmap}[2]{{#1}\;{\xlangrel{\mssafe}{\smssafe}}\;{#2}}
-  \begin{center}
-    \judgbox{\mmap{\msEvent}{\smsEvent}}{,,Translate \gls*{ms}-events to \gls*{sms}-events.''}
+\subsection{Compiler from $\obj{A}$ to $\ird{SpA}$}\label{subsec-a-to-spa}
+placeholder
+\subsubsection{Secure Compiler}\label{subsec-a-to-spa-seccomp}
+placeholder
 
-    \typerule{mapping-ms-sms-empty}{}{
-      \mmap{\msev{\emptyevent}}{\smsev{\emptyevent}}
-    }{mapping-ms-sms-empty}
-    %
-    \typerule{mapping-ms-sms-crash}{}{
-      \mmap{\msev{\lightning}}{\smsev{\lightning}}
-    }{mapping-ms-sms-crash}
-    %
-    \typerule{mapping-ms-sms-alloc}{}{
-      \mmap{\msAlloc}{\smsAlloc}
-    }{mapping-ms-sms-alloc}
-    %
-    \typerule{mapping-ms-sms-dealloc}{}{
-      \mmap{\msDealloc}{\smsDealloc}
-    }{mapping-ms-sms-dealloc}
-    %
-    \typerule{mapping-ms-sms-use}{}{
-      \mmap{\msUse}{\smsUse}
-    }{mapping-ms-sms-use}
-  \end{center}
-  \begin{center}
-    \judgbox{\mmap{\msTrace}{\smsTrace}}{,,Translate \gls*{ms}-traces to \gls*{sms}-traces.''}
-
-    \typerule{mapping-ms-sms-empty}{}{
-      \mmap{\emptyTrace}{\emptyTrace}
-    }{mapping-ms-sms-empty}
-    %
-    \typerule{mapping-ms-sms-cons}{}{
-      \mmap{\consTrace{\msEvent}{\msTrace}}{\consTrace{\smsEvent}{\smsTrace}}
-    }{mapping-ms-sms-cons}
-  \end{center}
-\end{definition}
-
-\begin{definition}{\definitionlabel[\glsfirst{ms}]{msdef}}
-  \[
-    \mssafe:=\mapUniversal{\xlangrel{\mssafe}{\tmssafe}}{\tmssafe} \cap \mapUniversal{\xlangrel{\mssafe}{\smssafe}}{\smssafe}
-  \]
-\end{definition}
-
-Note that \definitionref{msdef} ignores data isolation, so there may still be memory-safety issues introduced by side-channels.
-
-\subsubsection{Strict Cryptographic Constant Time}
-
-\gls*{cct} is a hypersafety property~\cite{barthe2018sec} and, thus, difficult to check with monitors.
-This is because, intuitively, hypersafety properties can relate multiple execution traces with each other, but monitors work on a single execution.
-It is a common trick to sidestep this issue by means of overapproximation: this section defines the property \gls*{scct}, a stricter variant of \gls*{cct} (inspired by earlier work~\cite{almeida2017jasmin}) that enforces the policy that no secret appears on a trace.
-Programs that satisfy \gls*{scct} also satisfy \gls*{cct}, but programs that satisfy \gls*{cct} may not satisfy \gls*{scct}.
-
-\begin{definition}{\definitionlabel[\glsfirst{scct}]{def-trace-scctdef}}
-  \noindent\[
-  \scctsafe:=\left\{\varTrace[_{\ctsafe}] 
-      \left|
-        \begin{array}{l}
-          \varTrace[_{\ctsafe}]=\hole{\cdot} \text{ or } \\
-          \exists\varTrace[_{\ctsafe}'],\varTrace[_{\ctsafe}]=\varEvent[_{\ctsafe}];\unlock\cdot\varTrace[_{\ctsafe}'] \wedge \varTrace[_{\ctsafe}']\in\scctsafe
-        \end{array}
-      \right.
-    \right\}
-  \]
-\end{definition}
-
-\gls*{scct} may appear overly strict, since it seems that secrets must not occur on a trace (since $\varSecuritytag$ is forced to be $\unlock$). 
-However, this is considered standard practice in terms of coding guidelines~\cite{ctguidelines}.
-Moreover, programs that have been compiled with FaCT~\cite{cauligi2019fact} and run with a ``data independent timing mode''~\cite{arm-refman,intel-refman} enabled do not leak secrets.
-
-\paragraph{\gls*{ms}, Strict Cryptographic Constant Time}
-
-The combination of \gls*{ms} and \gls*{scct} is the intersection of these properties, \gls*{msscct}.
-However, \gls*{ms} uses a different trace model than \gls*{scct}, so intersecting them would trivially yield the empty set. 
-To remedy this issue, we introduce $\sim_{\ctsafe}: \varEvent[_{\ctsafe}] \times \varEvent[_{\mssafe}] $, a cross-language trace relation (whose key cases are presented below), that we use to intuitively unify the trace model in which the two properties are expressed:
-\[
-  \typerulenolabel{}{\varEvent[_{\mssafe}]\sim_{\ctsafe}\varEvent[_{\mssafe}];\unlock{}}
-  \hspace{2ex}
-  \typerulenolabel{}{\lightning\sim_{\ctsafe}\lightning}
-  \hspace{2ex}
-  \typerulenolabel{}{\emptyevent\sim_{\ctsafe}\ev{Branch\;n};\varSecuritytag{}}
-  \hspace{2ex}
-  \typerulenolabel{}{\emptyevent\sim_{\ctsafe}\ev{Binop\;n};\varSecuritytag{}}
-\]
-
-Essentially, $\sim_{\ctsafe}$ ignores both the new $\ev{Branch\ n}$ and $\ev{Binop\ n}$ base-events as it relates security-insensitive actions ($\unlock$) to their equivalent counterparts.
-Thus, \gls*{ms} traces trivially satisfy \gls*{scct}.
-% 
-The above relation is extended point-wise to traces, skipping the empty event $\emptyevent$ on either side, and it is now possible to define \gls*{msscct} using the universal image:
-
-\begin{definition}{\definitionlabel[\glsfirst{ms} and \glsfirst{scct}]{def-trace-msscctdef}}
-  \[
-  \msscctsafe:=\mssafe\cap\sigma_{\sim_{\ctsafe}}\left(\scctsafe\right)
-  \]
-\end{definition}
-
-\paragraph{Extending the Trace Model with Speculation}
-
-So far, the considered trace models do not let us express speculative execution attacks such as Spectre~\cite{kocher2019spectre}. 
-For this, we extend the earlier trace model so that the security tags ($\varSecuritytag{}$) carry additional information about the kind of private data leakage, i.e., the type of speculative leak.
-Moreover, we add base-events signalling the beginning of a speculative execution ($\ev{Spec}$), a barrier ($\ev{Barrier}$) that signals that any speculative execution may not go past it, as well as a rollback event ($\ev{Rlb}$), which signals that execution resumes to where speculation started.
-
-\vspace{-1em}
-{
-\[
-  \begin{array}{rrcl}
-    (\text{Base-Events}) & \varEvent[_{\mathghost}] &:=& \varEvent[_{\ctsafe}] \\
-    (\text{Spectre Variants}) & vX &:=& \operatorname{NONE} \mid \operatorname{PHT} \\
-    (\text{Security Tags}) & \varSecuritytag{} &:=& \lock_{vX} \mid \unlock\\ 
-    (\text{Events}) & \varEvent[_{\mathghost}] &:=& \varEvent[_{\mathghost}];\varSecuritytag{} \mid \emptyevent \mid \lightning \mid \ev{Spec} \mid \ev{Rlb} \mid \ev{Barrier}\\ 
-  \end{array}
-\]
-}
-
-Even though the considered Spectre variants are just SPECTRE-PHT~\cite{kocher2019spectre}, NONE just describes secret data as in \gls*{scct} (see \Cref{subsec:scct:tracemodel}), the trace model is general enough to allow for potential future extension with different variants~\cite{kocher2019spectre,maisuradze2018ret2spec,horn2019zero}.
-%For sake of readability, this paper just uses the notation $\lock$ in place of $\lock_{\text{PHT}}$.
-
-\subsubsection{Speculative Safety}
-
-\gls*{ss}~\cite{patrignani2021exorcising}, similar to \gls*{scct}, is a sound overapproximation of a variant of noninterference.
-
-\begin{definition}{\definitionlabel[\glsfirst{ss}]{def-trace-ssdef}}
-  \[
-    \sssafe := \left\{
-      \varTrace[_{\mathghost}] 
-        \left|
-          \begin{array}{l}
-            \varTrace[_{\mathghost}]=\hole{\cdot} 
-              \text{ or } 
-            \exists\varTrace[_{\mathghost}'].\\
-            %
-            \left(\varTrace[_{\mathghost}]=\varEvent[_{\mathghost}];\unlock\cdot\varTrace[_{\mathghost}'] 
-              \text{or }
-            \varTrace[_{\mathghost}]=\varEvent[_{\mathghost}];\lock_{\text{NONE}}\cdot\varTrace[_{\mathghost}']\right)\\
-            %
-            \text{and }%
-            \ \varTrace[_{\mathghost}']\in\sssafe
-          \end{array}
-        \right.
-      \right\}
-  \] 
-\end{definition}
-The technical setup so far leads to the above definition, where only locks annotated with $\text{SPECTRE-PHT}$ are disallowed to occur on the trace.
-That way, programs attaining \gls*{ss} do not necessarily attain \gls*{scct}.
-
-\subsubsection{Speculation Memory Safety}\label{sec:spec-ms-rel}
-
-As before, we need to relate the different trace models with each other, so that the memory safety property without speculation can be lifted to speculation. 
-To this end, let $\sim_{\mathghost}: \varEvent[_{\mathghost}] \times \varEvent[_{\ctsafe}]$ be a cross-language trace relation whose key cases are below.
-The intuition is that \gls*{ss} is trivially satisfied in \gls*{scct}, since speculation is inexpressible there, which amounts to dropping events $\ev{Spec}$, $\ev{Rlb}$, or $\ev{Barrier}$, as well as all base events tagged with $\lock_{\text{PHT}}$.
-
-\[
-  \typerulenolabel{}{\varEvent[_{\ctsafe}];\lock\sim_{\mathghost}\varEvent[_{\ctsafe}];\lock_{\text{NONE}}}
-  \hspace{2ex}
-  \typerulenolabel{}{\emptyevent\sim_{\mathghost}\varEvent[_{\mathghost}];\lock_{\text{PHT}}}
-  \hspace{2ex}
-  \typerulenolabel{}{\varEvent[_{\ctsafe}];\unlock\sim_{\mathghost}\varEvent[_{\ctsafe}];\unlock}
-  \hspace{2ex}
-  \typerulenolabel{}{\lightning\sim_{\mathghost}\lightning}
-  \hspace{2ex}
-  \typerulenolabel{}{\emptyevent\sim_{\mathghost}\ev{Spec}}
-  \hspace{2ex}
-  \typerulenolabel{}{\emptyevent\sim_{\mathghost}\ev{Rlb}}
-  \hspace{2ex}
-  \typerulenolabel{}{\emptyevent\sim_{\mathghost}\ev{Barrier}}
-\]
-
-We conclude by defining the ultimate property of interest for secure compilers: \gls*{specms}.
-\begin{definition}{\definitionlabel[\glsfirst{specms}]{def-trace-specmsdef}}
-  \[
-  \specmssafe := \msscctsafe\cap\sigma_{\sim_{\mathghost}}\left(\sssafe\right)
-  \]
-\end{definition}
-
-\subsubsection{Monitors}\label{subsubsec-cs-properties-monitors}
+\subsection{Compiler from $\src{RusT}$ to $\ird{SpA}$}\label{subsec-rust-to-spa}
+placeholder
+\subsubsection{Secure Compiler}\label{subsec-rust-to-spa-seccomp}
+placeholder
 

### 2024-10-15 tech/tex/background/propmaps.tex
tech

\subsection{Property Mappings}

This work relies on property mappings introduced in related work~\cite{abate2021extacc}.

\begin{definition}{\definitionlabel[Universal Image]{universal-image}}
  $ 
    \mapUniversal{\sim}{\trgProperty} := 
      \lrbraces{\srcTrace \mid %
        \forall \trgTrace\ldotp \text{if }\xrelTraces{\srcTrace}{\trgTrace}, \text{ then }%
        \trgTrace\in\trgProperty
      }
  $
\end{definition}
\begin{definition}{\definitionlabel[Existential Image]{existential-image}}
  $ 
    \mapExistential{\sim}{\srcProperty} := 
      \lrbraces{\trgTrace \mid %
        \exists \srcTrace\ldotp \xrelTraces{\srcTrace}{\trgTrace} \text{ and }%
        \srcTrace\in\srcProperty
      }
  $
\end{definition}

The lifting of \definitionref{existential-image} and \definitionref{universal-image} to classes of properties is standard, i.e., simply element-wise.

\begin{lemma}[\Coqed]{\lemmalabel[Trace Relation distributes on Universal Image]{xtrace-distribute-universal}}
  Let $X$ be either a trace-property or a class. Then:
  \begin{goals}
    \goal{xtrace-distribute-universal}{\mapUniversal{\fncompo{\sim_1}{\sim_2}}{X} = \mapUniversal{\sim_2}{\mapUniversal{\sim_1}{X}}}
  \end{goals}
\end{lemma}
\begin{proof}
  Straightforward unfolding of \definitionref{universal-image} and \definitionref{function-composition} in \goalref{xtrace-distribute-universal}.
\end{proof}
\begin{lemma}[\Coqed]{\lemmalabel[Trace Relation distributes on Existential Image]{xtrace-distribute-existential}}
  Let $X$ be either a trace-property or a class. Then:
  \begin{goals}
    \goal{xtrace-distribute-existential}{\mapExistential{\fncompo{\sim_1}{\sim_2}}{X} = \mapExistential{\sim_2}{\mapExistential{\sim_1}{X}}}
  \end{goals}
\end{lemma}
\begin{proof}
  Straightforward unfolding of \definitionref{existential-image} and \definitionref{function-composition} in \goalref{xtrace-distribute-existential}.
\end{proof}

### 2024-10-15 tech/tex/background/rp.tex
tech

\subsection{Robust Preservation}

The following definitions for secure compilers are standard~\cite{abate2019jour,abate2021extacc}.

\subsubsection{Differing Trace Models}

A \bul{compiler $\ccST$ for languages $\src{S}$ and $\trg{T}$ robustly preserves a $\trg{T}$-level class $\trgClass$}, if for any \rul{property $\trgProperty$ of class $\trgClass$ and programs $\srcComponent$}, where \iul{$\srcComponent$ robustly satisfies $\mapUniversal{\sim}{\trgProperty}$}, \oul{the compilation of $\srcComponent$, we have that $\ccST\lrpars{\srcComponent}$ robustly satisfies $\trgProperty$}.

\begin{definition}{\definitionlabel[Robust Preservation with $\mapUniversal{\sim}{\cdot}$]{rtp-universal}}
  \begin{center}
    \bul{$\rtpUniversal{\ccST}{\trgClass}{\sim}$}
    %, iff 
    $\isdef$
    \rul{$\forall \trgProperty\in\trgClass, \srcComponent\in\src{S},$} %
    if \iul{$\rsat{\srcComponent}{\mapUniversal{\sim}{\trgProperty}}$},
    then \oul{$\rsat{\ccST\lrpars{\srcComponent}}{\trgProperty}$}.
  \end{center}
\end{definition}
A \bul{compiler $\ccST$ for languages $\src{S}$ and $\trg{T}$ robustly preserves a $\src{S}$-level class $\srcClass$}, if for any \rul{property $\srcProperty$ of class $\srcClass$ and programs $\srcComponent$}, where \iul{$\srcComponent$ robustly satisfies $\srcProperty$}, \oul{the compilation of $\srcComponent$, we have that $\ccST\lrpars{\srcComponent}$ robustly satisfies $\mapExistential{\sim}{\srcProperty}$}.

\begin{definition}{\definitionlabel[Robust Preservation with $\mapExistential{\sim}{\cdot}$]{rtp-existential}}
  \begin{center}
    \bul{$\rtpExistential{\ccST}{\srcClass}{\sim}$}
    %, iff 
    $\isdef$
    \rul{$\forall \srcProperty\in\srcClass, \srcComponent\in\src{S},$} %
    if \iul{$\rsat{\srcComponent}{\srcProperty}$},
    then \oul{$\rsat{\ccST\lrpars{\srcComponent}}{\mapExistential{\sim}{\srcProperty}}$}.
  \end{center}
\end{definition}

\subsubsection{Same Trace Models}

\begin{definition}{\definitionlabel[Robust Preservation]{rtp}}
  \begin{center}
    $\rtp{\ccST}{\varClass}$
    %, iff 
    $\isdef$
    $\forall \varProperty\in\varClass, \srcComponent\in\src{S},$ %
    if $\rsat{\srcComponent}{\varProperty}$,
    then $\rsat{\ccST\lrpars{\srcComponent}}{\varProperty}$.
  \end{center}
\end{definition}

### 2024-10-15 tech/tex/background/sat.tex
tech

\subsection{Satisfaction}
\bul{A whole program $\varWholeProg$ satisfies a property $\varProperty$} iff \iul{$\varWholeProg$ yields a trace $\varTrace$} such that \oul{$\varTrace$ satisfies $\varProperty$}.

\begin{definition}{\definitionlabel[Property Satisfaction]{prop-sat}}
    \bul{$\sat{\varWholeProg}{\varProperty}$}
    $\isdef$
    if \iul{$\forall\varRuntimeTerm\ \varTrace, %
      \progstepto{\varWholeProg}{\varRuntimeTerm}{\varTrace}$
    },
    then \oul{$\varTrace\in\varProperty$}.
\end{definition}

With this, \definitionref{prop-sat} can be extended as follows: for \bul{a component $\varComponent$ to robustly satisfy a property $\varProperty$}, take an \iul{attacker context $\varContext$ and link it with $\varComponent$}, \oul{the resulting whole program must satisfy $\varProperty$}.

\begin{definition}{\definitionlabel[Robust Property Satisfaction]{prop-rsat}}
  \bul{$\rsat{\varComponent}{\varProperty}$} %
  $\isdef$ %
  \iul{$\forall \varContext\ \varWholeProg$, %
    if $\bopLink{\varContext}{\varComponent} = \varWholeProg$}, %
    then \oul{$\sat{\varWholeProg}{\varProperty}$%
  }.
\end{definition}

### 2024-10-15 tech/tex/compo.tex
tech

\section{Composition of Secure Compilers}\label{sec:rtpc}

This section presents the composition framework.

\subsection{Same Trace Models}\label{subsec:rtpc-same-trace-models}

In the setting of same trace models, composition of secure compilers (as in \defref{rtp}) is simple:

\begin{lemma}[\Coqed]{\lemmalabel[Composition of RTP]{rtp-compo}}
  If 
  \begin{assumptions}
    \asm{rtp-compo-comp1}{\rtp{\ccSI}{\varClass[_1]}}
    \asm{rtp-compo-comp2}{\rtp{\ccIT}{\varClass[_2]}}
  \end{assumptions}
  then
  \begin{goals}
    \goal{rtp-compo}{\rtp{\fncompo{\ccSI}{\ccIT}}{\varClass[_1]\cap\varClass[_2]}}
  \end{goals}
\end{lemma}
\begin{proof}
  Unfold \goalref{rtp-compo} and introduce the assumptions. 
  \begin{passumptions}
    \asm{rtp-compo-prop}{\varProperty\in\varClass[_1]\cap\varClass[_2]}
    \asm{rtp-compo-trg-prog}{\srcComponent\in\src{S}}
    \asm{rtp-compo-rsat}{\rsat{\srcComponent}{\varProperty}}
  \end{passumptions}
  What is left to show is:
  \begin{goals}
    \goal{rtp-compo-rsat}{\rsat{\ccIT\lrpars{\ccSI\lrpars{\srcComponent}}}%
                               {\varProperty}}
  \end{goals}
  Apply \asmref{rtp-compo-comp2} on \goalref{rtp-compo-rsat} and note that due to \asmref{rtp-compo-prop}, $\varProperty\in\varClass[_2]$ also holds.
  \begin{goals}
    \goal{rtp-compo-rsatc}{\rsat{\ccSI\lrpars{\srcComponent}}{\varProperty}}
  \end{goals}
  Similarly, apply \asmref{rtp-compo-comp1} on \goalref{rtp-compo-rsatc} and note that due to \asmref{rtp-compo-prop}, $\varProperty\in\varClass[_1]$ also holds.
  \begin{goals}
    \goal{rtp-compo-rsatp}{\rsat{\srcComponent}{\varProperty}}
  \end{goals}
  \goalref{rtp-compo-rsatp} is resolved by \asmref{rtp-compo-rsat}.
\end{proof}
%%%%%%%%%%%%%%%%%%%%%%%%%%
The following corollary is a nice consequence of \lemref{rtp-compo}.
\begin{corollary}[\Coqed]{\corollarylabel[Commutativity of RTP]{rtp-swappable}}
  If 
  \begin{assumptions}
    \asm{rtp-swappable-comp1}{\rtp{\ccII}{\varClass[_1]}}
    \asm{rtp-swappable-comp2}{\rtp{\ccII[\delta]}{\varClass[_2]}}
  \end{assumptions}
  then
  \begin{goals}
    \goal{rtp-swappable1}{\rtp{\fncompo{\ccII}{\ccII[\delta]}}{\varClass[_1]\cap\varClass[_2]}}
    \goal{rtp-swappable2}{\rtp{\fncompo{\ccII[\delta]}{\ccII}}{\varClass[_1]\cap\varClass[_2]}}
  \end{goals}
\end{corollary}
\begin{proof}
  \goalref{rtp-swappable1} and \goalref{rtp-swappable2} are an immediate consequence based on \lemmaref{rtp-compo} with \asmref{rtp-swappable-comp1} and \asmref{rtp-swappable-comp2}, noting the commutativity of set intersection.
\end{proof}

\subsection{Different Trace Models}\label{subsec:rtpc-different-trace-models}

The situation in \Cref{subsec:rtpc-same-trace-models} is idealistic in the sense that a somewhat realistic robustly preserving compiler will likely have to insert checks for when control switches from context to component.
For example, compiling from a statically typed language into a dynamically typed one, where the static language enforces functions to only accept numbers, the compiler has to insert appropriate checks to make sure that well-typedness is robustly preserved.
Without such checks, a context could simply supply, e.g., a pair, violating the contract that a function can only receive numbers.

These checks yield a mismatch between the source-level trace and the target-level trace, rendering \definitionref{rtp} not provable.
Instead, one would have to use \definitionref{rtp-universal} or \definitionref{rtp-existential} and use an appropriate, cross-language trace relation that precisely describes the semantic effect on traces when compiling.
For the type-checking function interfaces example, parts of this cross-language trace relation may be sketched as:

\begin{center}
  \typerule{ex-rel-interfaces-agree}{
    \trg{v}\text{ is a number} \rulesep
    \src{v}\approx\trg{v}
  }{
    \xrelTraces{\src{Call\;\ctxtocomp\;v}}{\trg{Call\;\ctxtocomp\;v}}
  }{ex-rel-interfaces-agree}
  \typerule{ex-rel-interfaces-disagree}{
    \trg{v}\text{ not a number}
  }{
    \xrelTraces{\src{Abort}}{\trg{Call\;\ctxtocomp\;v}\cdot\trg{Abort}}
  }{ex-rel-interfaces-disagree}
\end{center}
\trref{ex-rel-interfaces-agree} ensures that both in the source and target-level execution, when the context calls into the component, the values agree and are numerical.

\trref{ex-rel-interfaces-disagree} handles the case where the target-level context supplied a $\trg{v}$ that is not a number.
Here, the backtranslation cannot construct a context that gives a well-typed program {\em and} produces the same trace as the target, since the behavior is ill-defined to begin with.
So, the only option for the backtranslation is to immediately crash.
This, however, is a mismatch with the target-level trace, which only aborts in the wrapper that the compiler inserted.

\begin{definition}{\definitionlabel[Closedness w.r.t. Universal Image]{universal-map-wf}}
  \[
    \mapUniversalWF{\sim}{\trgClass} \isdef \forall \trgProperty\in\trgClass, \mapUniversal{\sim}{\trgProperty}\in\mapUniversal{\sim}{\trgClass}
  \]
\end{definition}
\begin{definition}{\definitionlabel[Closedness w.r.t. Existential Image]{existential-map-wf}}
  \[
    \mapExistentialWF{\sim}{\srcClass} \isdef \forall \srcProperty\in\srcClass, \mapExistential{\sim}{\srcProperty}\in\mapExistential{\sim}{\srcClass}
  \]
\end{definition}

\begin{lemma}[\Coqed]{\lemmalabel[Composition of RTP-Universal]{rtp-universal-compo}}
  If 
  \begin{assumptions}
    \asm{rtp-universal-compo-wfsig}{\mapUniversalWF{\sim_2}{\trgClass[_1]}}
    \asm{rtp-universal-compo-comp1}{\rtpUniversal{\ccSI}{\mapUniversal{\sim_2}{\trgClass[_1]}}{\sim_1}}
    \asm{rtp-universal-compo-comp2}{\rtpUniversal{\ccIT}{\trgClass[_2]}{\sim_2}}
  \end{assumptions}
  then
  \begin{goals}
    \goal{rtp-universal-compo}{\rtpUniversal{\fncompo{\ccSI}{\ccIT}}{\trgClass[_1]\cap\trgClass[_2]}{\fncompo{\sim_1}{\sim_2}}}
  \end{goals}
\end{lemma}
\begin{proof}
  Unfold \goalref{rtp-universal-compo} and introduce the assumptions. 
  \begin{passumptions}
    \asm{rtp-universal-compo-prop}{\trgProperty\in\trgClass[_1]\cap\trgClass[_2]}
    \asm{rtp-universal-compo-trg-prog}{\srcComponent\in\src{S}}
    \asm{rtp-universal-compo-rsat}{\rsat{\srcComponent}{\mapUniversal{\fncompo{\sim_1}{\sim_2}}{\trgProperty}}}
  \end{passumptions}
  What is left to show is:
  \begin{goals}
    \goal{rtp-universal-compo-rsat}{\rsat{\ccIT\lrpars{\ccSI\lrpars{\srcComponent}}}%
                                         {\trgProperty}}
  \end{goals}
  Apply \asmref{rtp-universal-compo-comp2} on \goalref{rtp-universal-compo-rsat} and note that due to \asmref{rtp-universal-compo-prop}, $\trgProperty\in\trgClass[_2]$ also holds.
  \begin{goals}
    \goal{rtp-universal-compo-rsatc}{\rsat{\ccSI\lrpars{\srcComponent}}{\mapUniversal{\sim_2}{\trgProperty}}}
  \end{goals}
  Similarly, apply \asmref{rtp-universal-compo-comp1} on \goalref{rtp-universal-compo-rsatc} and note that due to \asmref{rtp-universal-compo-prop} and \asmref{rtp-universal-compo-wfsig}, $\trgProperty\in\mapUniversal{\sim_2}{\trgClass[_1]}$ also holds.
  \begin{goals}
    \goal{rtp-universal-compo-rsatp}{\rsat{\srcComponent}{\mapUniversal{\sim_2}{\mapUniversal{\sim_1}{\trgProperty}}}}
  \end{goals}
  Note that $\mapUniversal{\sim_2}{\mapUniversal{\sim_1}{\trgProperty}}=\mapUniversal{\fncompo{\sim_1}{\sim_2}}{\trgProperty}$, and thus \goalref{rtp-universal-compo-rsatp} is resolved by \asmref{rtp-universal-compo-rsat}.
\end{proof}

\begin{lemma}[\Coqed]{\lemmalabel[Composition of RTP-Existential]{rtp-existential-compo}}
  If 
  \begin{assumptions}
    \asm{rtp-existential-compo-comp1}{\rtpExistential{\ccSI}{\srcClass[_1]}{\sim_1}}
    \asm{rtp-existential-compo-comp2}{\rtpExistential{\ccIT}{\mapExistential{\sim_1}{\srcClass[_2]}}{\sim_2}}
  \end{assumptions}
  then
  \begin{goals}
    \goal{rtp-existential-compo}{\rtpExistential{\fncompo{\ccSI}{\ccIT}}{\srcClass[_1]\cap\srcClass[_2]}{\fncompo{\sim_1}{\sim_2}}}
  \end{goals}
\end{lemma}
\begin{proof}
  Similar to \lemmaref{rtp-universal-compo}.
\end{proof}

While compiler passes can be incompatible with each other in the sense that the resulting, mapped property is useless/not what is desired, it is enough to reason about compatibility of compilers at the level of traces.
This is what this work refers to as ,,semantic effect of a compiler''.
Instead of having to do syntactic arguments for the interaction of several compilation passes, one only has to do them for the semantic notion, i.e., the cross-language trace relations, which are much simpler proofs by an order of magnitude. 
\MKin{
  cite section backing up this claim
}

To conclude a similar colloray as \corollaryref{rtp-swappable}, it is of great use to show that a given cross-language trace relation has no effect on the class of security properties.
Then, it is possible to derive the following consequences:

\begin{corollary}[\Coqed]{\lemmalabel[Commutativity of RTP-Universal]{rtp-universal-swappable}}
  If 
  \begin{assumptions}
    \asm{rtp-universal-compo-wfsig1}{\mapUniversalWF{\sim_1}{\irlClass[_2]}}
    \asm{rtp-universal-compo-wfsig2}{\mapUniversalWF{\sim_2}{\irlClass[_1]}}
    \asm{rtp-universal-compo-eq2}{\mapUniversal{\sim_2}{\irlClass[_1]}=\irlClass[_1]}
    \asm{rtp-universal-compo-eq1}{\mapUniversal{\sim_1}{\irlClass[_2]}=\irlClass[_2]}
    \asm{rtp-universal-swappable-comp1}{\rtpUniversal{\ccII}{\irlClass[_1]}{\sim_1}}
    \asm{rtp-universal-swappable-comp2}{\rtpUniversal{\ccII[\delta]}{\irlClass[_2]}{\sim_2}}
  \end{assumptions}
  then
  \begin{goals}
    \goal{rtp-universal-swappable1}{\rtpUniversal{\fncompo{\ccII}{\ccII[\delta]}}{\irlClass[_1]\cap\irlClass[_2]}{\fncompo{\sim_1}{\sim_2}}}
    \goal{rtp-universal-swappable2}{\rtpUniversal{\fncompo{\ccII[\delta]}{\ccII}}{\irlClass[_1]\cap\irlClass[_2]}{\fncompo{\sim_2}{\sim_1}}}
  \end{goals}
\end{corollary}
\begin{proof}
  \goalref{rtp-universal-swappable1} and \goalref{rtp-universal-swappable2} are an immediate consequence based on \lemmaref{rtp-universal-compo} with \asmref{rtp-universal-swappable-comp1} and \asmref{rtp-universal-swappable-comp2}, noting the commutativity of set intersection.
\end{proof}

\begin{corollary}[\Coqed]{\lemmalabel[Commutativity of RTP-existential]{rtp-existential-swappable}}
  If 
  \begin{assumptions}
    \asm{rtp-existential-compo-wfsig1}{\mapExistentialWF{\sim_1}{\irlClass[_2]}}
    \asm{rtp-existential-compo-wfsig2}{\mapExistentialWF{\sim_2}{\irlClass[_1]}}
    \asm{rtp-existential-compo-eq2}{\mapExistential{\sim_2}{\irlClass[_1]}=\irlClass[_1]}
    \asm{rtp-existential-compo-eq1}{\mapExistential{\sim_1}{\irlClass[_2]}=\irlClass[_2]}
    \asm{rtp-existential-swappable-comp1}{\rtpExistential{\ccII}{\irlClass[_1]}{\sim_1}}
    \asm{rtp-existential-swappable-comp2}{\rtpExistential{\ccII[\delta]}{\irlClass[_2]}{\sim_2}}
  \end{assumptions}
  then
  \begin{goals}
    \goal{rtp-existential-swappable1}{\rtpExistential{\fncompo{\ccII}{\ccII[\delta]}}{\irlClass[_1]\cap\irlClass[_2]}{\fncompo{\sim_1}{\sim_2}}}
    \goal{rtp-existential-swappable2}{\rtpExistential{\fncompo{\ccII[\delta]}{\ccII}}{\irlClass[_1]\cap\irlClass[_2]}{\fncompo{\sim_2}{\sim_1}}}
  \end{goals}
\end{corollary}
\begin{proof}
  \goalref{rtp-existential-swappable1} and \goalref{rtp-existential-swappable2} are an immediate consequence based on \lemmaref{rtp-existential-compo} with \asmref{rtp-existential-swappable-comp1} and \asmref{rtp-existential-swappable-comp2}, noting the commutativity of set intersection.
\end{proof}

### 2024-10-15 tech/tex/monitors.tex
tech

\subsection{Monitors}\label{subsec-cs-monitors}

This part presents monitors for the properties presented in \Cref{subsec-cs-properties}.
Since all properties are safety properties, it is enough to consider trace-prefixes, i.e., finite lists of events.

In this section, we use the syntax $\exists\noncrashMonitor{\varMonitor}, \phi$ as shorthand for $\exists\varMonitor, \varMonitor\not=\lightning \text{ and } \phi$, where $\phi$ is some proposition that may depend on $\varMonitor$.
Essentially, whenever $\noncrashMonitor{\varMonitor}$ occurs, there is a constraint that $\varMonitor$ must not be $\lightning$.

\input{tex/monitors/tmsmon}
\input{tex/monitors/smsmon}
\input{tex/monitors/dual}
\input{tex/monitors/msmon}
\input{tex/monitors/scctmon}
\input{tex/monitors/msscctmon}
\input{tex/monitors/ssmon}
\input{tex/monitors/specmsmon}

### 2024-10-15 tech/tex/monitors/msmon.tex
tech

\subsubsection{\gls*{ms} Monitor}\label{subsubsec-mon-ms}
Instantiate \Cref{subsubsec-mon-dual} with \Cref{subsubsec-mon-tms,subsubsec-mon-sms}. 
We refer to the resulting monitor state as $\msMonitor$.
What is left to give is a translation from traces $\msTrace$ to traces $\varTrace[\tmssafe;\smssafe]$.
This translation is done as follows:
\begin{definition}{\definitionlabel[\gls*{ms} to \gls*{tms} and \gls*{sms}]{traces-ms-to-tmssms}}
  \begin{center}
    \judgbox{\msTrace\xlangrel{\mssafe}{\lrpars{\tmssafe;\smssafe}}\varTrace[\tmssafe;\smssafe]}{,,Project \gls*{ms}-traces to dual-trace \gls*{tms}-\gls*{sms}.''}

    \typerule{ms-to-tmssms-empty}{
    }{
      \hole{\cdot}\xlangrel{\mssafe}{\lrpars{\tmssafe;\smssafe}}\hole{\cdot}
    }{ms-to-tmssms-empty}
    %
    \typerule{ms-to-tmssms-cons}{
      \msEvent\xlangrel{\mssafe}{\tmssafe}\tmsEvent \rulesep
      \msEvent\xlangrel{\mssafe}{\smssafe}\smsEvent \rulesep
      \msTrace\xlangrel{\mssafe}{\lrpars{\tmssafe;\smssafe}}\varTrace[\tmssafe;\smssafe]
    }{
      \msEvent\cdot\msTrace\xlangrel{\mssafe}{\lrpars{\tmssafe;\smssafe}}\lrpars{\tmsEvent;\smsEvent}\cdot\varTrace[\tmssafe;\smssafe]
    }{ms-to-tmssms-cons}
  \end{center}
\end{definition}

\begin{definition}{\definitionlabel[\gls*{ms}-Satisfaction via Monitor]{trace-is-ms-via-mon}}
  \begin{align*}
    \mssafe\vdash\msTrace &\isdef \exists\varTrace[\tmssafe;\smssafe]\;\noncrashMonitor{\msMonitor},\monitorcheck{\emptyset}{\msMonitor}{\varTrace[\tmssafe;\smssafe]}\text{ and }
    \msTrace\xlangrel{\mssafe}{\lrpars{\tmssafe;\smssafe}}\varTrace[\tmssafe;\smssafe]
  \end{align*}
\end{definition}

### 2024-10-15 tech/tex/monitors/msscctmon.tex
tech

\subsubsection{\gls*{msscct} Monitor}\label{subsubsec-mon-msscct}

Instantiate \Cref{subsubsec-mon-dual} with \Cref{subsubsec-mon-ms,subsubsec-mon-scct}. 
We refer to the resulting monitor state as $\msscctMonitor$.
What is left to give is a translation from traces $\msTrace$ to traces $\varTrace[\mssafe;\scctsafe]$.
This translation is done as follows:
\begin{definition}{\definitionlabel[\gls*{ms} to \gls*{ms} and \gls*{scct}]{traces-ms-to-msscct}}
  \begin{center}
    \judgbox{\msTrace\xlangrel{\mssafe}{\lrpars{\mssafe;\scctsafe}}\varTrace[\mssafe;\scctsafe]}{,,Project \gls*{ms}-traces to dual-trace \gls*{ms}-\gls*{scct}.''}

    \typerule{ms-to-msscct-empty}{
    }{
      \hole{\cdot}\xlangrel{\mssafe}{\lrpars{\mssafe;\scctsafe}}\hole{\cdot}
    }{ms-to-mscct-empty}
    %
    \typerule{ms-to-msscct-cons}{
      \msEvent\xlangrel{\mssafe}{\scctsafe}\scctEvent \rulesep
      \msTrace\xlangrel{\mssafe}{\lrpars{\mssafe;\scctsafe}}\varTrace[\mssafe;\scctsafe]
    }{
      \msEvent\cdot\msTrace\xlangrel{\mssafe}{\lrpars{\mssafe;\scctsafe}}\lrpars{\msEvent;\scctEvent}\cdot\varTrace[\mssafe;\scctsafe]
    }{ms-to-msscct-cons}
  \end{center}
\end{definition}

\begin{definition}{\definitionlabel[\gls*{msscct}-Satisfaction via Monitor]{trace-is-msscct-via-mon}}
  \begin{align*}
    \msscctsafe\vdash\msTrace &\isdef \exists\varTrace[\mssafe;\scctsafe]\;\noncrashMonitor{\msscctMonitor},\monitorcheck{\emptyset}{\msscctMonitor}{\varTrace[\mssafe;\scctsafe]}\text{ and }
    \msTrace\xlangrel{\mssafe}{\lrpars{\mssafe;\scctsafe}}\varTrace[\mssafe;\scctsafe]
  \end{align*}
\end{definition}

### 2024-10-15 tech/tex/monitors/scctmon.tex
tech

\subsubsection{\gls*{scct} Monitor}\label{subsubsec-mon-scct}

\begin{definition}{\definitionlabel[\gls*{scct} Monitor State]{monitor-scct}}
  \begin{gather*}
    \begin{aligned}
      \text{(Abstract State)}~\scctMonitor &\isdef \emptyset \mid \lightning\\
    \end{aligned}
  \end{gather*}
\end{definition}

The reductions are defined as follows:
\begin{definition}{\definitionlabel[\gls*{scct} Monitor Reductions]{monitor-scct-steps}}
  \begin{center}
    \judgbox{\monitorcheck[]{\scctMonitor}{\scctMonitor[']}{\scctEvent}}{,,Given abstract state $\scctMonitor$, step to abstract state $\scctMonitor'$\\ emitting an abstract event $\scctEvent$.''}

    \typerule{scctmon-empty}{}{
      \monitorcheck[]{\scctMonitor}{\scctMonitor}{\emptyevent}
    }{scctmon-empty}
    \typerule{scctmon-any}{
    }{
      \monitorcheck[]{\scctMonitor}{\scctMonitor}{\scctAny[\unlock]}
    }{scctmon-any}
    \typerule{scctmon-crash}{}{
      \monitorcheck[]{\scctMonitor}{\lightning}{\lightning}
    }{scctmon-crash}
  \end{center}
  \begin{center}
    \judgbox{\monitorcheck{\scctMonitor}{\scctMonitor[']}{\scctEvent}}{,,Reflexive-Transitive closure of above judgement.\\$\scctTrace$ is a list of events.''}

    \typerule{scctmon-refl}{
    }{
      \monitorcheck{\scctMonitor}{\scctMonitor}{\hole{\cdot}}
    }{scctmon-refl}
    %
    \typerule{scctmon-trans}{
      \monitorcheck[]{\scctMonitor}{\scctMonitor[']}{\scctEvent} \rulesep
      \monitorcheck{\scctMonitor[']}{\scctMonitor['']}{\scctTrace}
    }{
      \monitorcheck{\scctMonitor}{\scctMonitor['']}{\scctEvent\cdot\scctTrace}
    }{scctmon-trans}
  \end{center}
\end{definition}

\begin{definition}{\definitionlabel[\gls*{scct}-Satisfaction via Monitor]{trace-is-scct-via-mon}}
  \begin{align*}
    \scctsafe\vdash\scctTrace &\isdef \exists\noncrashMonitor{\scctMonitor},\monitorcheck{\emptyset}{\scctMonitor}{\scctTrace}
  \end{align*}
\end{definition}

### 2024-10-15 tech/tex/monitors/specmsmon.tex
tech

\subsubsection{\gls*{specms} Monitor}\label{subsubsec-mon-specms}
Instantiate \Cref{subsubsec-mon-dual} with \Cref{subsubsec-mon-msscct,subsubsec-mon-spec}. 
We refer to the resulting monitor state as $\specmsMonitor$.

\begin{definition}{\definitionlabel[\gls*{msscct}-Satisfaction via Monitor]{trace-is-msscct-via-mon}}
  \begin{align*}
    \msscctsafe\vdash\varTrace[\mssafe;\sssafe] &\isdef \exists\noncrashMonitor{\msscctMonitor},\monitorcheck{\emptyset}{\msscctMonitor}{\varTrace[\mssafe;\sssafe]}
  \end{align*}
\end{definition}

### 2024-10-15 tech/tex/monitors/ssmon.tex
tech

\subsubsection{\gls*{ss} Monitor}\label{subsubsec-mon-spec}

\begin{definition}{\definitionlabel[\gls*{ss} Monitor State]{monitor-spec}}
  \begin{gather*}
    \begin{aligned}
      \text{(Abstract State)}~\specMonitor &\isdef \emptyset \mid \lightning\\
    \end{aligned}
  \end{gather*}
\end{definition}

The reductions are defined as follows:
\begin{definition}{\definitionlabel[\gls*{ss} Monitor Reductions]{monitor-spec-steps}}
  \begin{center}
    \judgbox{\monitorcheck[]{\specMonitor}{\specMonitor[']}{\specEvent}}{,,Given abstract state $\specMonitor$, step to abstract state $\specMonitor'$\\ emitting an abstract event $\specEvent$.''}

    \typerule{specmon-empty}{}{
      \monitorcheck[]{\specMonitor}{\specMonitor}{\emptyevent}
    }{specmon-empty}
    \typerule{specmon-any-public}{
    }{
      \monitorcheck[]{\specMonitor}{\specMonitor}{\specAny[\unlock]}
    }{specmon-any-public}
    \typerule{specmon-any-private}{
    }{
      \monitorcheck[]{\specMonitor}{\specMonitor}{\specAny[\lock_{\text{NONE}}]}
    }{specmon-any-private}
    \typerule{specmon-crash}{}{
      \monitorcheck[]{\specMonitor}{\lightning}{\lightning}
    }{specmon-crash}
  \end{center}
  \begin{center}
    \judgbox{\monitorcheck{\specMonitor}{\specMonitor[']}{\specEvent}}{,,Reflexive-Transitive closure of above judgement.\\$\specTrace$ is a list of events.''}

    \typerule{specmon-refl}{
    }{
      \monitorcheck{\specMonitor}{\specMonitor}{\hole{\cdot}}
    }{specmon-refl}
    %
    \typerule{specmon-trans}{
      \monitorcheck[]{\specMonitor}{\specMonitor[']}{\specEvent} \rulesep
      \monitorcheck{\specMonitor[']}{\specMonitor['']}{\specTrace}
    }{
      \monitorcheck{\specMonitor}{\specMonitor['']}{\specEvent\cdot\specTrace}
    }{specmon-trans}
  \end{center}
\end{definition}
Note that \trref{specmon-any-private} allows secrets to be exposed on the trace if the leak is not because of speculation.

\begin{definition}{\definitionlabel[\gls*{ss}-Satisfaction via Monitor]{trace-is-spec-via-mon}}
  \begin{align*}
    \sssafe\vdash\specTrace &\isdef \exists\noncrashMonitor{\specMonitor},\monitorcheck{\emptyset}{\specMonitor}{\specTrace}
  \end{align*}
\end{definition}

### 2024-10-15 tech/tex/properties.tex
tech

\subsection{Properties}\label{subsec-cs-properties}

The properties considered in this case-study are a mixture of trace-properties as welll as hyperproperties.
However, to carry out proofs, all properties are overapproximated by a monitor.

\subsubsection{Formal Definition}\label{subsubsec-cs-properties-formaldef}

For simple memory safety composed of temporal and spatial memory safety, the trace model defines events ($\varEvent[_{\mssafe}]$) as either the empty event ($\emptyevent$), a crash ($\lightning$), or a base-event ($\varEvent[_{\mssafe}]$).

\begin{definition}{\definitionlabel[\glsfirst{ms} Trace Model]{mstraces}}
\[
  \begin{array}{rcll}
    \text{(Base-Events)}&\varEvent[_{\mssafe}] &:=& \msAlloc \mid \msDealloc \mid \msUse \\
    \text{(Events)}&\varEvent[_{\mssafe}] &:=& \varEvent[_{\mssafe}] \mid \emptyevent \mid \lightning \\ 
  \end{array}
\]
\end{definition}

Base-events describe the actual kind of event that happened.
For the basic memory-safety properties, these are three variants:
First, the allocation event ($\msAlloc$) that fires whenever a program claims $n$ cells of memory and stores them at address $\varLoc$, where addresses are assumed to be unique.
Second, deallocation ($\msDealloc$) announces that the object at location $\varLoc$ is freed.
Third, an event to describe reads from and writes to the $n$-th memory cell from address $\varLoc$ ($\msUse$).

\paragraph{Temporal Memory Safety}

\gls*{tms}~\cite{nagarakatte2010cets} is a safety property that describes that an unallocated object must not be (re-)used.
Since it does not care about bounds of an object in memory, we emphasize this fact by deviating from the unified, memory-safety trace model and instead introduce traces for temporal memory safety:
\begin{definition}{\definitionlabel[\glsfirst{tms} Traces]{tmstraces}}
\[
  \begin{array}{rcll}
    \text{(Base-Events)}&\tmsEvent &:=& \tmsAlloc \mid \tmsDealloc \mid \tmsUse \\
    \text{(Events)}&\tmsEvent &:=& \tmsEvent \mid \emptyevent \mid \lightning \\ 
  \end{array}
\]
\end{definition}

With this, \gls*{tms} is defined as follows.
\begin{definition}{\definitionlabel[\glsfirst{tms}]{tmsdef}}
  \[
  \tmssafe:=\left\{\tmsTrace
    \left| 
      \begin{array}{rcl}
        \tmsAlloc&\le_{\tmsTrace}&\tmsDealloc \\
        \tmsUse&\le_{\tmsTrace}&\tmsDealloc \\
      \end{array}
    \right.
  \right\}
  \]
\end{definition}
Hereby, the notation $\varEvent[_{1}]\le_{\varTrace}\varEvent[_{2}]$ means that if $\varEvent[_{1}]$ is in $\varTrace$ and if $\varEvent[_{2}]$ is in $\varTrace$, then $\varEvent[_{1}]$ appears before $\varEvent[_{2}]$.

\paragraph{Spatial Memory Safety}
\gls*{sms}~\cite{nagarakatte2009soft} is a safety property that prohibits out-of-bounds accesses.
For emphasis, we also use a custom trace-model for spatial memory safety.
\begin{definition}{\definitionlabel[\glsfirst{sms} Traces]{smstraces}}
\[
  \begin{array}{rcll}
    \text{(Base-Events)}&\smsEvent &:=& \smsAlloc \mid \smsUse \\
    \text{(Events)}&\smsEvent &:=& \smsEvent \mid \emptyevent \mid \lightning \\ 
  \end{array}
\]
\end{definition}

With this, \gls*{sms} is defined as follows.
\begin{definition}{\definitionlabel[\glsfirst{sms}]{smsdef}}
  \[
  \smssafe:=\left\{\smsTrace \left|\begin{array}{rcl}
      \text{If }\smsAlloc\le_{\smsTrace}\smsUse, \text{ then }m<n
  \end{array}\right.\right\}
  \]
\end{definition}

\paragraph{Memory Safety}

In spirit of earlier work~\cite{nagarakatte2009soft,nagarakatte2010cets,jim2002cyclone,necula2005ccured,michael2023mswasm}, full \gls*{ms} is the intersection of \definitionref{tmsdef} and \definitionref{smsdef}.
However, both properties use a different trace model, so intersecting them alone would trivially yield the empty set.
To remedy this, we define cross-language trace relations that lift both \definitionref{tmstraces} and \definitionref{smstraces} to \definitionref{mstraces}.
\begin{definition}{\definitionlabel[Mapping \glsfirst{ms} to \glsfirst{tms}]{mapping-tms-ms}}
  \newcommand{\mmap}[2]{{#1}\;{\xlangrel{\mssafe}{\tmssafe}}\;{#2}}
  \begin{center}
    \judgbox{\mmap{\msEvent}{\tmsEvent}}{,,Translate \gls*{ms}-events to \gls*{tms}-events.''}

    \typerule{mapping-ms-tms-empty}{}{
      \mmap{\msev{\emptyevent}}{\tmsev{\emptyevent}}
    }{mapping-ms-tms-empty}
    %
    \typerule{mapping-ms-tms-crash}{}{
      \mmap{\msev{\lightning}}{\tmsev{\lightning}}
    }{mapping-ms-tms-crash}
    %
    \typerule{mapping-ms-tms-alloc}{}{
      \mmap{\msAlloc}{\tmsAlloc}
    }{mapping-ms-tms-alloc}
    %
    \typerule{mapping-ms-tms-dealloc}{}{
      \mmap{\msDealloc}{\tmsDealloc}
    }{mapping-ms-tms-dealloc}
    %
    \typerule{mapping-ms-tms-use}{}{
      \mmap{\msUse}{\tmsUse}
    }{mapping-ms-tms-use}
  \end{center}
  \begin{center}
    \judgbox{\mmap{\msTrace}{\tmsTrace}}{,,Translate \gls*{ms}-traces to \gls*{tms}-traces.''}

    \typerule{mapping-ms-tms-refl}{}{
      \mmap{\emptyTrace}{\emptyTrace}
    }{mapping-ms-tms-refl}
    %
    \typerule{mapping-ms-tms-cons}{}{
      \mmap{\consTrace{\msEvent}{\msTrace}}{\consTrace{\tmsEvent}{\tmsTrace}}
    }{mapping-ms-tms-cons}
  \end{center}
\end{definition}
\begin{definition}{\definitionlabel[Mapping \glsfirst{ms} to \glsfirst{sms}]{mapping-sms-ms}}
  \newcommand{\mmap}[2]{{#1}\;{\xlangrel{\mssafe}{\smssafe}}\;{#2}}
  \begin{center}
    \judgbox{\mmap{\msEvent}{\smsEvent}}{,,Translate \gls*{ms}-events to \gls*{sms}-events.''}

    \typerule{mapping-ms-sms-empty}{}{
      \mmap{\msev{\emptyevent}}{\smsev{\emptyevent}}
    }{mapping-ms-sms-empty}
    %
    \typerule{mapping-ms-sms-crash}{}{
      \mmap{\msev{\lightning}}{\smsev{\lightning}}
    }{mapping-ms-sms-crash}
    %
    \typerule{mapping-ms-sms-alloc}{}{
      \mmap{\msAlloc}{\smsAlloc}
    }{mapping-ms-sms-alloc}
    %
    \typerule{mapping-ms-sms-dealloc}{}{
      \mmap{\msDealloc}{\emptyevent}
    }{mapping-ms-sms-dealloc}
    %
    \typerule{mapping-ms-sms-use}{
      n=m
    }{
      \mmap{\msUse}{\smsUse}
    }{mapping-ms-sms-use}
  \end{center}
  \begin{center}
    \judgbox{\mmap{\msTrace}{\smsTrace}}{,,Translate \gls*{ms}-traces to \gls*{sms}-traces.''}

    \typerule{mapping-ms-sms-refl}{}{
      \mmap{\emptyTrace}{\emptyTrace}
    }{mapping-ms-sms-refl}
    %
    \typerule{mapping-ms-sms-cons}{}{
      \mmap{\consTrace{\msEvent}{\msTrace}}{\consTrace{\smsEvent}{\smsTrace}}
    }{mapping-ms-sms-cons}
  \end{center}
\end{definition}

\begin{definition}{\definitionlabel[\glsfirst{ms}]{msdef}}
  \[
    \mssafe:=\mapUniversal{\xlangrel{\mssafe}{\tmssafe}}{\tmssafe} \cap \mapUniversal{\xlangrel{\mssafe}{\smssafe}}{\smssafe}
  \]
\end{definition}

Note that \definitionref{msdef} ignores data isolation, so there may still be memory-safety issues introduced by side-channels.

\subsubsection{Strict Cryptographic Constant Time}

\begin{definition}{\definitionlabel[\glsfirst{scct} Traces]{sccttraces}}
\[
  \begin{array}{rcll}
    \text{(Secrecy Tags)}&\varSecuritytag &:=& \unlock \mid \lock \\
    \text{(Events)}&\tmsEvent &:=& \scctAny \mid \emptyevent \mid \lightning \\ 
  \end{array}
\]
\end{definition}

\gls*{cct} is a hypersafety property~\cite{barthe2018sec} and, thus, difficult to check with monitors.
This is because, intuitively, hypersafety properties can relate multiple execution traces with each other, but monitors work on a single execution.
It is a common trick to sidestep this issue by means of overapproximation: this section defines the property \gls*{scct}, a stricter variant of \gls*{cct} (inspired by earlier work~\cite{almeida2017jasmin}) that enforces the policy that no secret appears on a trace.
Programs that satisfy \gls*{scct} also satisfy \gls*{cct}, but programs that satisfy \gls*{cct} may not satisfy \gls*{scct}.

\begin{definition}{\definitionlabel[\glsfirst{scct}]{scctdef}}
  \noindent\[
  \scctsafe:=\left\{\scctTrace 
      \left|
        \begin{array}{l}
          \scctTrace=\hole{\cdot} \text{ or } \scctTrace=\lightning \\
          \exists\scctTrace['],\left(\scctTrace=\scctAny[\unlock]\cdot\varTrace[_{\ctsafe}'] \text{ or } \right.\\
          \hspace{2.8em}\left.\scctTrace=\emptyevent\cdot\scctTrace[']
          \wedge \varTrace[_{\ctsafe}']\in\scctsafe \right)
        \end{array}
      \right.
    \right\}
  \]
\end{definition}

\gls*{scct} may appear overly strict, since it seems that secrets must not occur on a trace (since $\varSecuritytag$ is forced to be $\unlock$). 
However, this is considered standard practice in terms of coding guidelines~\cite{ctguidelines}.
Moreover, programs that have been compiled with FaCT~\cite{cauligi2019fact} and run with a ``data independent timing mode''~\cite{arm-refman,intel-refman} enabled do not leak secrets. 
We'd like to emphasize that our goal is to motivate the core theory presented in \Cref{sec:rtpc} and not develop a new programming language.

\paragraph{\gls*{ms}, Strict Cryptographic Constant Time}

The combination of \gls*{ms} and \gls*{scct} is the intersection of these properties, \gls*{msscct}.
However, \gls*{ms} uses a different trace model than \gls*{scct}, so intersecting them would trivially yield the empty set. 
To remedy this issue, we introduce $\xlangrel{\scctsafe}{\mssafe} : \scctEvent \to \msEvent \to \mathbb{P}$, a cross-language trace relation (whose key cases are presented below), that we use to intuitively unify the trace model in which the two properties are expressed:

\begin{definition}{\definitionlabel[Mapping \glsfirst{ms} to \glsfirst{scct}]{mapping-ms-scct}}
  \newcommand{\mmap}[2]{{#1}\;{\xlangrel{\mssafe}{\scctsafe}}\;{#2}}
  \begin{center}
    \judgbox{\mmap{\msEvent}{\scctEvent}}{,,Translate \gls*{ms}-events to \gls*{scct}-events.''}

    \typerule{mapping-ms-scct-empty}{}{
      \mmap{\msev{\emptyevent}}{\scctev{\emptyevent}}
    }{mapping-ms-scct-empty}
    %
    \typerule{mapping-ms-scct-crash}{}{
      \mmap{\msev{\lightning}}{\scctev{\lightning}}
    }{mapping-ms-scct-crash}
    %
    \typerule{mapping-ms-scct-alloc}{}{
      \mmap{\msAlloc}{\scctAny[\unlock]}
    }{mapping-ms-scct-alloc}
    %
    \typerule{mapping-ms-scct-dealloc}{}{
      \mmap{\msDealloc}{\scctAny[\unlock]}
    }{mapping-ms-scct-dealloc}
    %
    \typerule{mapping-ms-scct-use}{}{
      \mmap{\msUse}{\scctAny[\unlock]}
    }{mapping-ms-scct-use}
  \end{center}
  \begin{center}
    \judgbox{\mmap{\msTrace}{\scctTrace}}{,,Translate \gls*{ms}-traces to \gls*{scct}-traces.''}

    \typerule{mapping-ms-scct-refl}{}{
      \mmap{\emptyTrace}{\emptyTrace}
    }{mapping-ms-scct-refl}
    %
    \typerule{mapping-ms-scct-cons}{}{
      \mmap{\consTrace{\msEvent}{\msTrace}}{\consTrace{\scctEvent}{\scctTrace}}
    }{mapping-ms-scct-cons}
  \end{center}
\end{definition}

%Essentially, $\sim_{\ctsafe}$ ignores both the new $\ev{Branch\ n}$ and $\ev{Binop\ n}$ base-events as it relates security-insensitive actions ($\unlock$) to their equivalent counterparts.
\gls*{ms} traces trivially satisfy \gls*{scct}, since it is assumed here that they are the result of a cryptographic constant time computation.
This may be ensured by a compiler, such as FaCT, or the semantics of the language itself.
We also don't have any other information in \gls*{scct} traces besides the secrecy tags, since the semantics of a language can simply tag relevant events with the respective vulnerability, such as a leak of secret data.
% 
It is now possible to define \gls*{msscct} using the universal image:

\begin{definition}{\definitionlabel[\glsfirst{ms} and \glsfirst{scct}]{msscctdef}}
  \[
    \msscctsafe:=\mssafe\cap\mapUniversal{\xlangrel{\mssafe}{\scctsafe}}{\scctsafe}
  \]
\end{definition}

\paragraph{Extending the Trace Model with Speculation}

So far, the considered trace models do not let us express speculative execution attacks such as Spectre~\cite{kocher2019spectre}. 
For this, we extend the earlier trace model so that the secrecy tags ($\varSecuritytag{}$) carry additional information about the kind of private data leakage, i.e., the type of speculative leak.
This tag may be emitted by the semantics of a concrete programming language, as seen in earlier work~\cite{fabian2022automatic}. 
%Moreover, we add base-events signalling the beginning of a speculative execution ($\ev{Spec}$), a barrier ($\ev{Barrier}$) that signals that any speculative execution may not go past it, as well as a rollback event ($\ev{Rlb}$), which signals that execution resumes to where speculation started.

\begin{definition}{\definitionlabel[\glsfirst{ss} traces]{sstraces}}
{
\[
  \begin{array}{rrcl}
    (\text{Spectre Variants}) & vX &:=& \operatorname{NONE} \mid \operatorname{PHT} \mid \dots \\
    (\text{Secrecy Tags}) & \varSecuritytag{} &:=& \lock_{vX} \mid \unlock\\ 
    (\text{Events}) & \specEvent &:=& \specEvent \mid \emptyevent \mid \lightning \\ 
  \end{array}
\]
}
\end{definition}

%Even though the considered Spectre variants are just SPECTRE-PHT~\cite{kocher2019spectre}, NONE just describes secret data as in \gls*{scct} (see \Cref{subsec:scct:tracemodel}), the trace model is general enough to allow for potential future extension with different variants~\cite{kocher2019spectre,maisuradze2018ret2spec,horn2019zero}.
NONE describes secret data as in \gls*{scct} (see \Cref{subsec:scct:tracemodel}).
Any other secrecy tag, such as PHT, describes the concrete reason for the respective leak.
The trace model in \defref{sstraces} is general enough to allow for potential future extension with different variants~\cite{kocher2019spectre,maisuradze2018ret2spec,horn2019zero}.
%For sake of readability, this paper just uses the notation $\lock$ in place of $\lock_{\text{PHT}}$.

\subsubsection{Speculation Safety}

\gls*{ss}~\cite{patrignani2021exorcising}, similar to \gls*{scct}, is a sound overapproximation of a variant of noninterference.

\begin{definition}{\definitionlabel[\glsfirst{ss}]{ssdef}}
  \[
    \sssafe := \left\{
      \specTrace
        \left|
          \begin{array}{l}
            \specTrace=\hole{\cdot} 
              \text{ or } 
            \specTrace=\lightning
              \text{ or }
            \exists\specTrace['].\\
            %
            \left(\specTrace=\specAny[\unlock]\cdot\specTrace['] 
              \text{or }
            \varTrace[_{\mathghost}]=\specAny[\lock_{\text{NONE}}]\cdot\specTrace[']\right.\\
            \hspace{2em}\text{or }
            \left.\varTrace[_{\mathghost}]=\emptyevent\cdot\specTrace[']\right)\\
            %
            \text{and }%
            \ \varTrace[_{\mathghost}']\in\sssafe
          \end{array}
        \right.
      \right\}
  \] 
\end{definition}
The technical setup so far leads to the above definition, where only locks annotated with $\text{SPECTRE-PHT}$ are disallowed to occur on the trace.
That way, programs attaining \gls*{ss} do not necessarily attain \gls*{scct}.

\subsubsection{Speculation Memory Safety}\label{sec:spec-ms-rel}

As before, we need to relate the different trace models with each other, so that the memory safety property without speculation can be lifted to speculation. 
To this end, let $\xlangrel{ms}{ss}: \msEvent \to \specEvent\to\mathbb{P}$ be a cross-language trace relation whose key cases are below.
The intuition is that \gls*{ss} is trivially satisfied in \gls*{ms}, since speculation is inexpressible there and a language attaining \gls*{ms} by means of semantics can enforce \gls*{ss} simply by not having a speculative dynamic semantics. 
In turn, all base events tagged with $\lock_{\text{PHT}}$ are simply dropped.

\begin{definition}{\definitionlabel[Mapping \glsfirst{ms} to \glsfirst{ss}]{mapping-ms-spec}}
  \newcommand{\mmap}[2]{{#1}\;{\xlangrel{\mssafe}{\sssafe}}\;{#2}}
  \begin{center}
    \judgbox{\mmap{\msEvent}{\specEvent}}{,,Translate \gls*{ms}-events to \gls*{ss}-events.''}

    \typerule{mapping-ms-spec-empty}{}{
      \mmap{\msev{\emptyevent}}{\specev{\emptyevent}}
    }{mapping-ms-spec-empty}
    %
    \typerule{mapping-ms-spec-crash}{}{
      \mmap{\msev{\lightning}}{\specev{\lightning}}
    }{mapping-ms-spec-crash}
    %
    \typerule{mapping-ms-spec-alloc}{}{
      \mmap{\msAlloc}{\specAny[\unlock]}
    }{mapping-ms-spec-alloc}
    %
    \typerule{mapping-ms-spec-dealloc}{}{
      \mmap{\msDealloc}{\specAny[\unlock]}
    }{mapping-ms-spec-dealloc}
    %
    \typerule{mapping-ms-spec-use}{}{
      \mmap{\msUse}{\specAny[\unlock]}
    }{mapping-ms-spec-use}
  \end{center}
  \begin{center}
    \judgbox{\mmap{\msTrace}{\specTrace}}{,,Translate \gls*{ms}-traces to \gls*{ss}-traces.''}

    \typerule{mapping-ms-spec-refl}{}{
      \mmap{\emptyTrace}{\emptyTrace}
    }{mapping-ms-spec-refl}
    %
    \typerule{mapping-ms-spec-cons}{}{
      \mmap{\consTrace{\msEvent}{\msTrace}}{\consTrace{\specEvent}{\specTrace}}
    }{mapping-ms-spec-cons}
  \end{center}
\end{definition}

We conclude with the ultimate property of interest for our secure compiler: \gls*{specms}.
\begin{definition}{\definitionlabel[\glsfirst{specms}]{specmsdef}}
  \[
    \specmssafe := \msscctsafe\cap\mapUniversal{\xlangrel{\mssafe}{\sssafe}}{\sssafe}
  \]
\end{definition}

### 2024-10-22 spectre/acronyms.tex
fill parts of spectre table

\newacronym{rtp}{RTP}{Robust Trace-Property Preservation}
\newacronym{tms}{TMS}{Temporal Memory Safety}
\newacronym{sms}{SMS}{Spatial Memory Safety}
\newacronym{ms}{MS}{Memory Safety}
\newacronym{cct}{CCT}{Cryptographic Constant Time}
\newacronym{scct}{sCT}{Strict Cryptographic Constant Time}
\newacronym{msscct}{MS+sCCT}{Memory Safety and Strict Cryptographic Constant Time}
\newacronym{dce}{DCE}{Dead Code Elimination}
\newacronym{cf}{CF}{Constant Folding}
\newacronym{ss}{SS}{Speculation Safety}
\newacronym{specms}{SpecMS}{Speculation Memory Safety}

\newacronym{s}{S}{Software}
\newacronym{h}{H}{Hardware}
\newacronym{hs}{HS}{Hardware-Simulation}

\newacronym{seq}{SEQ}{Sequential, in-order execution of instructions.}
\newacronym{spec}{SPEC}{Speculative, in-order execution of instructions.}
\newacronym{seq-ooo}{SEQOOO}{Sequential, out-of-order execution of instructions.}
\newacronym{spec-ooo}{SPECOOO}{Speculative, out-of-order execution of instructions.}

\newacronym{memobs}{MEMobs}{Sees loads and store memory addresses.}
\newacronym{ctobs}{CTobs}{Sees loads, stores, and program counter memory addresses.}
\newacronym{pathctobs}{PATH-CTobs}{Constant-Time observer. Sees loads, stores, and program counter memory addresses, as well as all possible (speculative) paths.}
\newacronym{ct+pcobs}{CT+PCobs}{\acrshort{ctobs}-obsverver, but in speculative-mode only sees program counter.}
\newacronym{archobs}{ARCHobs}{Architectural observer. Sees any microarchitectural covert channel.}

\newacronym{ccseqct}{$\ccseqct$}{\acrshort{seq}-execution; \acrshort{ctobs}-observations.}
\newacronym{ccspecmem}{$\ccspecmem$}{\acrshort{spec}-execution; \acrshort{memobs}-observations.}
\newacronym{ccspecct}{$\ccspecct$}{\acrshort{spec}-execution; \acrshort{ctobs}-observations.}
\newacronym{ccspecpathct}{$\ccspecpathct$}{\acrshort{spec}-execution; \acrshort{pathctobs}-observations}
\newacronym{ccseqarch}{$\ccseqarch$}{\acrshort{seq}-execution; \acrshort{archobs}-observations}
\newacronym{ccspecarch}{$\ccspecarch$}{\acrshort{spec}-execution; \acrshort{archobs}-observations}
\newacronym{ccseqspecctpc}{$\ccseqspecctpc$}{\acrshort{seq}- and \acrshort{spec}-execution; \acrshort{ct+pcobs} observer}

\newacronym{fi}{FI}{Fence Insertion}
\newacronym{cmr}{CMR}{Cancelable Memory Requests}
\newacronym{dtt}{DTT}{Dynamic Taint Tracking}
\newacronym{msr}{MSR}{Model-Specific Register}
\newacronym{iso}{ISO}{Process Isolation}
\newacronym{rz}{RZ}{Register Zeroing}
\newacronym{sst}{SSt}{Shadow Stack}
\newacronym{ri}{RI}{Register Interlocking}
\newacronym{slh}{SLH}{Speculative Load Hardening}

### 2024-12-10 icfp/main.tex
Add icfp folder

--- 2024-09-26 csf2/main.tex
@@
 \newcommand\bmmax{0}
-\documentclass[dvipsnames,conference]{IEEEtran}
+\documentclass[acmsmall]{acmart}
+%%
+%% \BibTeX command to typeset BibTeX logo in the docs
+\AtBeginDocument{%
+  \providecommand\BibTeX{{%
+    Bib\TeX}}}
+
+%% Rights management information.  This information is sent to you
+%% when you complete the rights form.  These commands have SAMPLE
+%% values in them; it is your responsibility as an author to replace
+%% the commands and values with those provided to you when you
+%% complete the rights form.
+\setcopyright{acmlicensed}
+\copyrightyear{2018}
+\acmYear{2018}
+\acmDOI{XXXXXXX.XXXXXXX}
+
+%%
+%% These commands are for a JOURNAL article.
+\acmJournal{JACM}
+\acmVolume{37}
+\acmNumber{4}
+\acmArticle{111}
+\acmMonth{8}
+
+%%
+%% Submission ID.
+%% Use this when submitting an article to a sponsored event. You'll
+%% receive a unique submission ID from the organizers
+%% of the event, and this ID should be used as the parameter to this command.
+%%\acmSubmissionID{123-A56-BU3}
+
+%%
+%% For managing citations, it is recommended to use bibliography
+%% files in BibTeX format.
+%%
+%% You can then either use BibTeX with the ACM-Reference-Format style,
+%% or BibLaTeX with the acmnumeric or acmauthoryear sytles, that include
+%% support for advanced citation of software artefact from the
+%% biblatex-software package, also separately available on CTAN.
+%%
+%% Look at the sample-*-biblatex.tex files for templates showcasing
+%% the biblatex styles.
+%%
+
+%%
+%% The majority of ACM publications use numbered citations and
+%% references.  The command \citestyle{authoryear} switches to the
+%% "author year" style.
+%%
+%% If you are preparing content for an event
+%% sponsored by ACM SIGGRAPH, you must use the "author year" style of
+%% citations and references.
+%% Uncommenting
+%% the next command will enable that style.
+%%\citestyle{acmauthoryear}
 
@@
 \usepackage[inference]{semantic}
+\usepackage[dvipsnames]{xcolor}
 \usepackage[switch]{lineno}
@@
 \usepackage{listings}
-\usepackage{amssymb}
 \usepackage{amsmath}
@@
 \makeglossaries
-
-\AtBeginDocument{%
-  \providecommand\BibTeX{{%
-    \normalfont B\kern-0.5em{\scshape i\kern-0.25em b}\kern-0.8em\TeX}}}
-\makeatletter
-\newcommand{\linebreakand}{%
-  \end{@IEEEauthorhalign}
-  \hfill\mbox{}\par
-  \mbox{}\hfill\begin{@IEEEauthorhalign}
-}
-\makeatother
-
 \renewcommand\thelinenumber{\color{red}\arabic{linenumber}}
@@
 \linenumbers
-
 \title{
@@
   of Robust and Optimising Compilers
-  \vspace{-1.5em}
 }
 
-% \author{
-% Matthis Kruse\orcidID{0000-0003-4062-9666}
-% \and
-% Michael Backes
-% \and
-% Marco Patrignani
-% }
-% \authorrunning{M. Kruse et al.}
-
-\maketitle
+\author{Matthis Kruse}
+\affiliation{%
+   \institution{Saarland University and CISPA}
+   \city{Saarbr\"ucken}
+   \state{Saarland}
+   \country{Germany}}
+\email{matthis.kruse@cispa.de}
+\author{Marco Patrignani}
+\affiliation{%
+   \institution{Trento University}
+   \city{Trento}
+   \state{South Tyrol}
+   \country{Italy}}
+\email{marco.patrignani@unitn.it}
+%%
+%% By default, the full list of authors will be used in the page
+%% headers. Often, this list is too long, and will overlap
+%% other information printed in the page headers. This command allows
+%% the author to define a more concise list
+%% of authors' names for this purpose.
+\renewcommand{\shortauthors}{Kruse and Patrignani}
 
@@
 
-% \begin{IEEEkeywords}
-%   memory-safety, cryptographic constant-time, secure compilation, robustness
-% \end{IEEEkeywords}
+%%
+%% The code below is generated by the tool at http://dl.acm.org/ccs.cfm.
+%% Please copy and paste the code instead of the example below.
+%%
+\begin{CCSXML}
+<ccs2012>
+ <concept>
+  <concept_id>00000000.0000000.0000000</concept_id>
+  <concept_desc>Do Not Use This Code, Generate the Correct Terms for Your Paper</concept_desc>
+  <concept_significance>500</concept_significance>
+ </concept>
+ <concept>
+  <concept_id>00000000.00000000.00000000</concept_id>
+  <concept_desc>Do Not Use This Code, Generate the Correct Terms for Your Paper</concept_desc>
+  <concept_significance>300</concept_significance>
+ </concept>
+ <concept>
+  <concept_id>00000000.00000000.00000000</concept_id>
+  <concept_desc>Do Not Use This Code, Generate the Correct Terms for Your Paper</concept_desc>
+  <concept_significance>100</concept_significance>
+ </concept>
+ <concept>
+  <concept_id>00000000.00000000.00000000</concept_id>
+  <concept_desc>Do Not Use This Code, Generate the Correct Terms for Your Paper</concept_desc>
+  <concept_significance>100</concept_significance>
+ </concept>
+</ccs2012>
+\end{CCSXML}
+
+\ccsdesc[500]{Do Not Use This Code~Generate the Correct Terms for Your Paper}
+\ccsdesc[300]{Do Not Use This Code~Generate the Correct Terms for Your Paper}
+\ccsdesc{Do Not Use This Code~Generate the Correct Terms for Your Paper}
+\ccsdesc[100]{Do Not Use This Code~Generate the Correct Terms for Your Paper}
+
+%%
+%% Keywords. The author(s) should pick words that accurately describe
+%% the work being presented. Separate the keywords with commas.
+\keywords{Do, Not, Us, This, Code, Put, the, Correct, Terms, for,
+  Your, Paper}
+
+\received{20 February 2007}
+\received[revised]{12 March 2009}
+\received[accepted]{5 June 2009}
+
+\maketitle
 
@@
   		% 
-  		Additionally this paper analyses each pass in terms of their compatibility with each other (\Cref{sec:formalities}), proving how does each pass fulfill the conditions that make their composition meaningful from a security standpoint.
+  		Additionally this paper analyses each pass in terms of their compatibility with each other (\Cref{sec:formalities}), proving how each pass fulfills the conditions that make their composition meaningful from a security standpoint.
         % In doing so, the paper provides formal insights into some technicalities of the concrete proofs necessary as technical setup to reap the benefits of the presented compositionality framework.
@@
 % 
-The set of all properties can be partitioned into different {\em classes} ($\class$), i.e., safety, liveness, and neither safety nor liveness~\cite{clarkson2008hyper}, so a class is simply a set of properties.
+The set of all properties can be split-up into different {\em classes} ($\class$), i.e., safety, liveness, and neither safety nor liveness~\cite{clarkson2008hyper}, so a class is simply a set of properties.
 % 
@@
 % 
-A \bul{compiler $\cc{\src{L}}{\trg{L}}$ robustly preserves a class of target properties $\trg{\class}$}, if for any \rul{property $\trg{\pi}$ of class $\trg{\class}$ and programs $\src{p}$}, where \iul{$\src{\progvar}$ robustly satisfies $\sigma_\sim\left(\trg{\pi}\right)$}, \oul{the compilation of $\src{\progvar}$, we have that $\cc{\src{L}}{\trg{L}}\left(\src{p}\right)$ robustly satisfies $\trg{\pi}$}.
+A \bul{compiler $\cc{\src{L}}{\trg{L}}$ robustly preserves a class of target properties $\trg{\class}$}, if for any \rul{property $\trg{\pi}$ of class $\trg{\class}$ and programs $\src{p}$}, where \iul{$\src{\progvar}$ robustly satisfies $\sigma_\sim\left(\trg{\pi}\right)$}, \oul{for the compilation of $\src{\progvar}$, we have that $\cc{\src{L}}{\trg{L}}\left(\src{p}\right)$ robustly satisfies $\trg{\pi}$}.
 
@@
 
-Let \bul{$\cc{\src{L}}{\trg{L}}$ robustly preserve the class $\sigma_{\sim_2}\left(\trg{\class[_1]}\right)$ under $\sim_1$} and let \rul{$\cc{\trg{L}}{\obj{L}}$ robustly preserve the class $\trg{\class[_2]}$ under $\sim_2$}.
+Let \bul{$\cc{\src{L}}{\trg{L}}$ robustly preserve the class $\sigma_{\sim_2}\left(\obj{\class[_1]}\right)$ under $\sim_1$} and let \rul{$\cc{\trg{L}}{\obj{L}}$ robustly preserve the class $\obj{\class[_2]}$ under $\sim_2$}.
 Then, \iul{when the cross-language relation $\sim_2$ is well-formed w.r.t.
-class $\trg{\class[_1]}$}, it follows that \oul{the composed compiler $\cc{\src{L}}{\trg{L}}\circ\cc{\trg{L}}{\obj{L}}$ robustly preserves the intersection of classes $\trg{\class[_1]}\cap\trg{\class[_2]}$ under $\sim_1\bullet\sim_2$}.
+class $\obj{\class[_1]}$}, it follows that \oul{the composed compiler\footnote{Here, $\cc{\src{L}}{\trg{L}}\circ\cc{\trg{L}}{\obj{L}}$ means first running $\cc{\src{L}}{\trg{L}}$ and then $\cc{\trg{L}}{\obj{L}}$.} $\cc{\src{L}}{\trg{L}}\circ\cc{\trg{L}}{\obj{L}}$ robustly preserves the intersection of classes $\obj{\class[_1]}\cap\obj{\class[_2]}$ under $\sim_1\bullet\sim_2$}.
 % 
@@
 
-  If \bul{$\rtpsig{\cc{\src{L}}{\trg{L}}}{\tilde{\sigma}_{\sim_2}\left(\trg{\class[_{1}]}\right)}{\sim_1}$}, \rul{$\rtpsig{\cc{\trg{L}}{\obj{L}}}{\trg{\class[_2]}}{\sim_2}$}, and \iul{$\wfcsig{\sim_2}{\trg{\class[_1]}}$}, \\ then \oul{$\rtpsig{\cc{\src{L}}{\trg{L}}\circ\cc{\trg{L}}{\obj{L}}}{\trg{\class[_{1}]}\cap\trg{\class[_{2}]}}{\sim_1\bullet\sim_2}$}.
+  If \bul{$\rtpsig{\cc{\src{L}}{\trg{L}}}{\tilde{\sigma}_{\sim_2}\left(\obj{\class[_{1}]}\right)}{\sim_1}$}, \rul{$\rtpsig{\cc{\trg{L}}{\obj{L}}}{\obj{\class[_2]}}{\sim_2}$}, and \iul{$\wfcsig{\sim_2}{\obj{\class[_1]}}$}, \\ then \oul{$\rtpsig{\cc{\src{L}}{\trg{L}}\circ\cc{\trg{L}}{\obj{L}}}{\obj{\class[_{1}]}\cap\obj{\class[_{2}]}}{\sim_1\bullet\sim_2}$}.
 \Coqed
@@
 % 
-Monitors refine each property and have a key tole in the proofs of this paper.
+Monitors refine each property and have a key role in the proofs of this paper.
 % managed to prove 
@@
 
-To express Constant Time, we extend the memory safety trace model with a {\em security tag} ($\securitytag{}$) that indicates whether events contains sensitive information ($\lock$) or not ($\unlock$).
+To express Constant Time, we extend the memory safety trace model with a {\em security tag} ($\securitytag{}$) that indicates whether events contain sensitive information ($\lock$) or not ($\unlock$).
 
@@
 \[
-  \typerulenolabel{specrel:drop:tag}{}{\preevent[_{\ctsafe}];\lock\sim_{\mathghost}\preevent[_{\ctsafe}];\lock_{\text{NONE}}}
+  \typerulenolabel{specrel:drop:tag}{}{\preevent[_{\ctsafe}];\lock\sim_{\mathghost}\preevent[_{\mathghost}];\lock_{\text{NONE}}}
   \typerulenolabel{specrel:drop:spectag}{}{\emptyevent\sim_{\mathghost}\preevent[_{\mathghost}];\lock_{\text{PHT}}}
@@
 \[
-  \typerulenolabel{specrel:drop:pub}{}{\preevent[_{\ctsafe}];\unlock\sim_{\mathghost}\preevent[_{\ctsafe}];\unlock}
+  \typerulenolabel{specrel:drop:pub}{}{\preevent[_{\ctsafe}];\unlock\sim_{\mathghost}\preevent[_{\mathghost}];\unlock}
   \typerulenolabel{specrel:drop:crash}{}{\lightning\sim_{\mathghost}\lightning}
@@
     \text{(Values)} & \trg{v} &:=& \dots \mid \trg{\lpair{v_1}{v_2}} \\
-    \text{(Expressions)} & \trg{\expr} &:=& \dots \mid \trg{\lhast{\expr}{\tau}} \mid \trg{\lispoisoned{x}}
+    \text{(Expressions)} & \trg{\expr} &:=& \dots \mid \trg{\lhast{\expr}{\tau}} \mid \trg{\lispoisoned{x}} \mid \trg{e.1} \mid \trg{e.2}
   \end{array}
@@
 
-  When the constant-time mode is off, the execution yields events in similar fashion to before (\Cref{subsec:ltms,subsec:lsms,subsec:lms}).
+  When the constant-time mode is off, the execution yields events without constant-time leaks (\Cref{subsec:ltms,subsec:lsms,subsec:lms}).
   But, if it is turned on, then the branching event \colorbox{yellow}{does not fire} anymore and both reading and writing to memory is related to a specification trace with \colorbox{lightgray}{no exposed secrets}.
@@
 These passes can be freely ordered in the compilation chain without compromising memory safety (\Cref{thm:cccfccdce:rtp:ms}).
-The next step in the chain ensures that code stays \gls*{scct} (\Cref{thm:ccscct:rtp:scct}) when compiled from $\Lms$ to $\Lscct$, which is done by switching on a constant-tame mode for the computation.
+The next step in the chain ensures that code stays \gls*{scct} (\Cref{thm:ccscct:rtp:scct}) when compiled from $\Lms$ to $\Lscct$, which is done by switching on a constant-time mode for the computation.
 Lastly, by introducing barriers immediately after branches, speculative leaks via SPECTRE-PHT are prevented when compiling $\Lscct$ to $\Lspec$.
@@
 \subsection{Security Properties and Their Meaning}\label{subsec:formalities:props}
-Each of the presented compilers use a cross-language trace relation, which is also used to translate the property from one language to the other one (via the existential or universal images).
+Each of the presented compilers uses a cross-language trace relation, which is also used to translate the property from one language to the other one (via the existential or universal images).
 % 
@@
 
-Finally, let us consider a trace $\irl{\trace}\in\mssafe$ and understand what is that is related to via $\sim^{\Ltrg}_{\Lms}$.
+Finally, let us consider a trace $\irl{\trace}\in\mssafe$ and understand what is related to via $\sim^{\Ltrg}_{\Lms}$.
 % 

### 2025-01-29 icfp/main.tex
hyper

@@
 %\smallskip%MP: the underline compacts the sections too much
+
+\subsection{Hyperproperties and (Robust) Satisfaction}
+
+Hyperproperties~\cite{clarkson2008hyper} are strictly more expressive than ordinary properties. 
+While a property is a set of traces, a hyperproperty is a set of properties. 
+We can lift properties to hyperproperties by applying the powerset.
+Hyperproperty satisfaction is defined as follows.
+
+It turns out that classes of hyperproperties, that is, sets of hyperproperties, are equally expressive than simply hyperproperties themselves~\cite{clarkson2008hyper}.
+
+% and so on
 

### 2025-02-03 icfp/main.tex
sketch of intro

@@
 \title{
-  Secure Composition 
-  \\ 
-  of Robust and Optimising Compilers
+  Secure Composition of Robust and Optimising Compilers
 }
@@
 
-\section{Introduction\pages{4}}\label{sec:introduction}
-
-% \gls*{ms} is a security property that ensures that there are no out-of-bounds accesses as well as no reads or writes to uninitialised memory.
-% Programs that do not attain memory safety, can easily have their private data be leaked to attackers or be corrupted~\cite{lemay2021ccc}.
-% Because of this, \gls*{ms} is a requirement for programs to attain an even stronger security property, namely \gls*{cct}.
-% Programs that attain \gls*{cct} are secure w.r.t.
-timing attacks, and this is the gold standard for secure applications, including cryptographic ones, since timing attacks can expose secrets via a simple performance analysis~\cite{kocher1996timing}.
-% Timing attacks have become more and more relevant recently, especially given the era of cloud-computing~\cite{aviram2010cloudtime,kumar2019cloudsecsurvey}, where programs share the same hardware but are supposed to run in a sandboxed environment that does not leak secrets.
-% Unfortunately, several attacks within such a cloud environment have been demonstrated as viable strategies to extract secrets~\cite{mehmet2015getoff,flowers2022zeroday,atya2019catchme,venkatanathan2015placevul}.
-
-% Even though mitigations for such attacks exist~\cite{bond2017vale,almeida2017jasmin}, some compilers may optimise these mitigations away~\cite{barthe2018sec}.
-% This is of course not what programmers expect and to ensure such mitigations persist in the compiled code, one can use a {\em secure} compilers.
-% More generally, secure compilers should preserve security properties even when the program runs in a hostile context such as the cloud one, i.e., where it interacts with other possibly malicious programs.
-% An example of such secure compilers are those that ascribe to the theory of robust compilation~\cite{abate2019jour,abate2021extacc,patrignani2021rsc}, which takes this interaction between components and potentially malicious contexts into account.
-
-% Unfortunately, it is unclear if and how robust compilation supports compositionality, and compilation passes are typically developed as a standalone module to be plugged into a compilation pipeline and composed with other passes.
-% Different kinds of compilation passes exist, e.g., there are usually several compilation passes that perform optimisations~\cite{lattner2004llvm,googlev8,androidstudio,kuepper2023cryptopt,manjikian1997fusion,wegman1991ccp} or source-code instrumentations~\cite{nagarakatte2009soft,nagarakatte2010cets,akritidis2009baggy,dhumbumroong2020boundwarden,jung2021pico,nam2019framer,shankaranarayana2023tailcheck,younan2010paricheck,zhou2023fatptrs,bond2017vale,almeida2017jasmin,kuepper2023cryptopt,cauligi2019fact} that enforce that the compiled program fulfils a property of interest.
-% Without a framework for composing secure compiler passes, it is not possible to enable separation of concerns, e.g., to have a secure compilation pass that ensures \gls*{ms} that is developed independently of another secure pass for \gls*{cct} as suggested before and as showcased in \Cref{ex:strncpy}.
-% \MPpost{
-%   this intro does a poor job at introducing the robust compilation argument.
-%   perhaps it'd be best to start from robust properties, which motivates the use of robust compilers?
-%   idk
-%   as for the abstract, this is a good starting point for you to make a better intro
-% }
-
-\gls*{ms} is a security property obtained by composing \gls*{sms}, which ensures array accesses are all within bounds, and \gls*{tms}, which ensures pointers are only used when they are valid~\cite{azevedo2018meaningsofms,jim2002cyclone,necula2005ccured,nagarakatte2010cets,nagarakatte2009soft,akritidis2009baggy,michael2023mswasm}.
-\gls*{cct} is a security property that ensures sensitive data is not leaked via timing side-channels~\cite{kocher1996timing} and \gls*{ss} is a security property that enforces the same but under a speculative semantics~\cite{guarnieri2018spectector,fabian2022automatic} that captures speculative execution attacks such as spectre~\cite{kocher2019spectre}.
-Together, \gls*{sms}, \gls*{tms}, \gls*{cct} and \gls*{ss}, yield \gls*{specms}, which is the gold standard of security properties for secure applications.
-Programs attaining \gls*{specms} do not leak sensitive data either through erroneous memory accesses, nor through timing side-channels, even under speculative execution.
-\Cref{ex:strncpy} below discusses how these security properties can be enforced by compiler passes~\cite{bond2017vale,almeida2017jasmin}, to ensure programmers need not be aware of the architectural details of where their code will run.
-
-\begin{exampleenv}[strncpy]\label{ex:strncpy}
-Consider the following \texttt{C} function \texttt{strncpy} that copies a null-terminated string \texttt{src} into \texttt{dst} up to a length of \texttt{n} characters.
-This function is subject to a subtle \gls*{sms} vulnerability: the bounds check \texttt{i < n} should happen {\it before} the access to memory location \texttt{src[i]}: otherwise
-% If \texttt{n} is at least as large as the length of the array \texttt{src}, 
-the memory location past the last element is leaked to an attacker.
-\begin{lstlisting}[language=c,basicstyle=\small\ttfamily,morekeywords={size_t}]
-void strncpy(size_t n, char *dst, char *src) {
-  for(size_t i = 0; src[i] != '\0' && i < n; ++i) {
-    dst[i] = src[i];
-  }
-}
-\end{lstlisting}
-% Assume that the length of the array pointed by \texttt{dst} is at least \texttt{n}, and \texttt{src} is terminated by \texttt{'\textbackslash 0'}.
-% Even under these assumptions, there is a 
-To prevent this vulnerability, one can use a compilation pass that enforces \gls*{sms}, such as Softbounds~\cite{nagarakatte2009soft} or BaggyBounds~\cite{akritidis2009baggy}.
-The most na\"ive solution in this case is to insert bounds-checks in front of every access to memory.
-
-Because of timing attacks, addressing \gls*{sms} is not enough to make \texttt{strncpy} secure.
-In fact, the loop can terminate early, as soon as the string-terminating character \texttt{'\textbackslash 0'} is encountered, thus making program execution time proportional to the length of the array pointed by \texttt{src}, and violating the \gls*{cct} property.
-% Often, programmers rewrite programs like these to have loops that run with a fixed number of iterations.
-Also in this case there exist compiler passes that can rewrite such programs into \gls*{cct} ones~\cite{cauligi2019fact}.
- % so that they are no longer vulnerable to timing attacks.
-%
-Finally, even with these precautions, code is not run in isolation, so a malicious attacker could supply code that interacts with \texttt{strncpy} and trigger a violation of either \gls*{ms} or \gls*{cct} by calling \texttt{strncpy} with an argument for \texttt{src} that points to uninitialised memory.
-This would, in turn, trigger a series of reads from uninitialised memory, which is an immediate \gls*{ms} violation with devastating real-world consequences~\cite{uninit-0,uninit-1,uninit-2,uninit-3,uninit-4}.
-% An easy fix to sidestep this is to use non-nullable pointers as parameters instead, e.g., as presented in Cyclone~\cite{jim2002cyclone}.
-\end{exampleenv}
-
-  % last sentence is before the example, which talks about compiler passes enforcing props
-Whether or not compiler passes enforce certain security properties, it is important for a secure compiler to consider partial programs that interact with potentially malicious code, since the latter may lead to, e.g., memory-safety issues in the considered partial program.
-Robust compilers~\cite{abate2019jour} are a form of secure compilers that preserve security properties even in the presence of arbitrary attackers interacting with compiled code.
-Thus, robust compilers can be used to prevent vulnerabilities resulting from uninitialised memory (as well as many other ones), e.g., by targeting capability-based languages such as CHERI~\cite{woodruff2014CHERI}, Arm Morello~\cite{arm-morello}, or MSWasm~\cite{michael2023mswasm}, where the compiler relies on capabilities to check that pointers are always initialised.
-
-% To ensure these compiler-preserved properties still hold even in the presence of arbitrary attackers interacting with compiled code, one can use secure compilers in the form of robust compilers~\cite{abate2019jour,abate2021extacc,patrignani2021rsc}.
-
-Unfortunately, given secure compiler passes that each preserve a possibly different security property, there is no way to tell what kind of security property will the composition of those secure compilers preserve.
-% Different kinds of compilation passes exist, e.g., there are usually several compilation passes that perform optimisations~\cite{lattner2004llvm,googlev8,androidstudio,kuepper2023cryptopt,manjikian1997fusion,wegman1991ccp} or source-code instrumentations~\cite{nagarakatte2009soft,nagarakatte2010cets,akritidis2009baggy,dhumbumroong2020boundwarden,jung2021pico,nam2019framer,shankaranarayana2023tailcheck,younan2010paricheck,zhou2023fatptrs,bond2017vale,almeida2017jasmin,kuepper2023cryptopt,cauligi2019fact} that enforce that the compiled program fulfils a property of interest.
-Worse, without a framework for composing secure compiler passes, it is not possible to enable separation of concerns, e.g., to have a secure compilation pass that ensures \gls*{ms} that is developed independently of another secure pass for \gls*{cct}, that is developed independently of other passes, such as optimisation ones.
-
-This paper introduces a framework for reasoning about the composition of secure and optimising compiler passes %akin to those of \Cref{ex:strncpy}\MP{there aren't really passes there} 
-and it showcases the power of this framework by instantiating it on a multi-pass compilation chain.
-To this end, this paper first discusses how to compose security properties, such as \gls*{tms} and \gls*{sms} into \gls*{ms}, and then adds \gls*{cct} as well as \gls*{ss} into the mix to obtain \gls*{specms}.
-%Then, this paper defines compiler composition and formalises that given two passes that securely preserve two (possibly distinct) properties, their composition securely preserves the composition of those properties.
-The paper then defines several secure compiler passes, where each is either preserving a different security property from the list above
-% (\gls*{tms}, \gls*{sms}, \gls*{scct}, \gls*{ss}) 
-or performing a security-preserving optimisation.
-% , (e.g., applying \gls*{cf} or \gls*{dce}).
-Finally, this paper shows that composing these secure compiler passes into a multi-pass compilation chain results in the end-to-end (robust) preservation of \gls*{specms}.
-Crucially, this paper derives the security of the multi-pass compiler from the composition of the security properties preserved by its individual passes.
-% Crucially, this end-to-end security preservation is obtained by composing the security property preserved by each individual pass, as dictated by the framework of this paper.
-This result showcases how the framework allows the kind of formal security reasoning that compiler writers already want (and already do), obtaining precise, compositional security reasoning while providing minimal (and modular) proof effort.
-%
-In summary, this paper makes the following contributions:
-% \begin{itemize}[leftmargin=0em,nosep,label=$\blacktriangleright$]
-\begin{asparaitem}[$\blacktriangleright$]
-  \item %
-        %\Cref{sec:sequential} studies different forms of secure compiler compositions.
-        This paper takes the secure compilation framework of Abate et al.~\cite{abate2021extacc} and extends it for compositionality. (\Cref{sec:sequential}).
-        %Concretely, it investigates upper and lower compositions, as well as the sequential composition.
-        %For this, this paper studies compiler composition as well as compilers with multiple input languages or multiple output ones, as used in existing compilation chains.
-        %
-        This paper proves that starting from two compilers that preserve two (possibly distinct) properties, their composition preserves roughly the intersection of those properties.
-        % For example, a compiler that is secure w.r.t. \gls*{tms} can be composed with a compiler that is secure w.r.t. \gls*{sms}, resulting in a compiler that is secure w.r.t. \gls*{ms}.
-        % 
-        % From this, the paper proves that the order of composition of sequential compiler passes is irrelevant for security.
-        % % 
-        % This is crucial for reordering individual optimisation passes and thus generating secure and efficient code.
-        %
-        %Finally, this paper extends parts of the compositionality framework to support different trace-models between compilation passes.
-        % \MP{this isn't really true, is it?. the order matters for the relations' lemmas}
-        Then, this paper identifies which conditions make the composition of secure compilers meaningful from a security perspective, and which other conditions allow passes to be swapped without losing security meaning.
-
-  \item %
-        This paper presents a case-study with five programming languages showcasing the previous contribution (\Cref{sec:casestud:defs,sec:casestud:rtp}).
-        To this end, it presents a compilation chain consisting of six passes that ultimately preserve \gls*{specms} by means of composing four secure passes that \emph{individually} preserve \gls*{tms}, \gls*{sms}, \gls*{scct}---an enforceable version of \gls*{cct}---, and \gls*{ss}, and two passes which are well-known optimisations \gls*{dce} and \gls*{cf}.
-        %Furthermore, the chain includes two optimisation passes: one performs \gls*{dce} and the other \gls*{cf}.
-        The formalisation of this case study showcases the power of the presented framework: the divide-and-conquer approach to software engineering is a viable strategy even for the development of secure compilers.
-%        This demonstrates that the usage of the presented framework is both simpler and more expressive compared to one monolithic security proof of the whole compilation chain.
-
-  \item %
-  		This paper provides formal proofs of the security properties preserved by each of the presented passes. 
-  		% 
-  		Additionally this paper analyses each pass in terms of their compatibility with each other (\Cref{sec:formalities}), proving how each pass fulfills the conditions that make their composition meaningful from a security standpoint.
-        % In doing so, the paper provides formal insights into some technicalities of the concrete proofs necessary as technical setup to reap the benefits of the presented compositionality framework.
-
-  \item The key contributions of this paper are formalised in the Coq proof assistant and the paper indicates this with \CoqSymbol.
-\end{asparaitem}
-% 
-This paper starts by introducing relevant notions of security properties and secure compilation (\Cref{sec:background}),
-and discusses related work (\Cref{sec:relwork}) before concluding (\Cref{sec:concl}).
-
-% \contribution{Technical Report} 
-The omitted formal details, lemmas and proofs, as well as the Coq formalisation are available as supplementary material.
-% \MKin{cite/link}
+\input{sections/intro.tex}
 

### 2025-02-03 icfp/sections/intro.tex
sketch of intro

\section{Introduction}\label{sec:introduction}

Programming languages provide security enforcements, e.g., Rust memory safety by means of ownership semantics.

Compilers may fail to preserve these enforcements, even when proven correct.

This is why secure compilers have become an interesting avenue for research, but they are difficult to prove.

Compositionality to the rescue, proofs should compose to make it easier.

% a secure compiler, by definition, is supports source-independent linking
% vertical composition is a consequence of existence of backtranslations

- Demonstrate compositionality of secure compilation.

- Mechanization of core theory.

- Case Study demonstrating composition.

### 2025-02-04 icfp/sections/intro.tex
sketch of intro

@@
 
-Programming languages provide security enforcements, e.g., Rust memory safety by means of ownership semantics.
+As of today, there exist a number of programming languages that enforce security-relevant properties.
+One of the most used programming languages in that space is Rust~\cite{}, which guarantees basic memory safety.
+That is, there is a guarantee that all memory accesses are {\em temporally} as well as {\em spatially} safe. 
+Temporal memory safety is the absence of use-after-free and use-after-reallocation bugs and necessitates tracking of pointer provenance\footnote{Due to potential use-after-reallocation bugs.}.
+%TODO: example?
+Provenance of pointers is also crucial information, together with bounds, for spatial memory safety, which ensures that all accesses are within bounds of an allocated object.
 
-Compilers may fail to preserve these enforcements, even when proven correct.
+Compilers may fail to preserve these enforcements, even when proven correct or thoroughly tested~\cite{}.
+While properties can be enforced at source-level by means of static analyses, after compiling to languages without such abstractions, the properties may break.
+For example, RISC-V, a possible target-language of the most commonly used Rust compiler, is untyped and provides neither memory safety nor type safety guarantees.
+Intuitively, this is why we want compilers to be {\em correct}, so that they provably preserve the meaning of the source-language.
+Unfortunately, software is seldolmy developed in isolation and programs may be linked with target-level libraries after compilation.
+These libraries can be compiled with different compilers or may even be hand-written, so the correctness-guarantees provided by one compiler may not hold.
 

### 2025-02-05 icfp/sections/intro.tex
Rewrite intro

@@
 While properties can be enforced at source-level by means of static analyses, after compiling to languages without such abstractions, the properties may break.
-For example, RISC-V, a possible target-language of the most commonly used Rust compiler, is untyped and provides neither memory safety nor type safety guarantees.
+For example, RISC-V, a possible target-language of the most commonly used Rust compiler,  provides neither memory nor type safety guarantees.
 Intuitively, this is why we want compilers to be {\em correct}, so that they provably preserve the meaning of the source-language.
+With preservation of meaning follows preservation of safety guarantees.
 Unfortunately, software is seldolmy developed in isolation and programs may be linked with target-level libraries after compilation.
@@
 
-This is why secure compilers have become an interesting avenue for research, but they are difficult to prove.
+In previous work, it has been argued that correct compilation is a spectrum~\cite{} with different restrictions applying to how the compiled partial programs can interact with target-level libraries.
+At the extreme end of the spectrum with basically zero restrictions\footnote{The minimal restriction that has to be applied is interface compatibility.} lies robust preservation of meaning.
+Here, partial program components can interact with arbitrary contexts, i.e., libraries, and the compiler has to ensure that crossing the boundary between components and contexts does not break any properties whatsoever.
+For example, in order to robustly ensure memory safety in Rust when compiling to RISC-V, the compiler could compartmentalize all calls and returns via the foreign function interface.
+This compartmentalization can be done in different ways, one of which is via CHERI capabilities~\cite{}.
 
-Compositionality to the rescue, proofs should compose to make it easier.
+For the context of this paper, we are interested in formally verified compilers.
+That is, they need to be proven to be robustly preserving, which is a labor-intensive and in parts difficult task~\cite{}. %lots of cits
+Worse, compilers are not just simple functions translating from one domain into another. 
+A practical compiler consists of several different compilation passes, which itself can be seen as compilers that translate from one intermediate representation into another.
+The CompCert project~\cite{}, while having more restrictions on target-level contexts, attests the herculian effort necessary to prove correctness for realistic compilers with several different passes.
+Moreover, for optimising passes, it can be beneficial to swap the order or iterate them until a fixpoint is reached, but this necessitates modifications to a top-level meaning-preservation proof of a compiler.
+Therefore, it is helpful to have local meaning-preservation proofs for each pass and be able to compose them to obtain a whole-proof for the complete pipeline.
 
-% a secure compiler, by definition, is supports source-independent linking
-% vertical composition is a consequence of existence of backtranslations
+Compositionality of meaning-preservation proofs for compilers has been investigated in the context of different flavors of compiler correctness. 
+But, to the best of our knowledge, compositionality for robustly preserving compilers is still unclear.
+Moreover, there are two key compositionality properties~\cite{} that researchers care about: (1) vertical compositionality and (2) source-independent linking. 
+Vertical compositionality asserts that it should be possible to verify a compiler pass in isolation.
+For source-independent linking, note that when verifying a compiler, source and target terms need to be related somehow.
+However, this is only necessary for the component, i.e., the partial program that is being compiled, and so source-independent linking asserts that the target context need not be related to any source-context.
 
-- Demonstrate compositionality of secure compilation.
+Our aim with this paper is to lay the groundwork for secure compiler composition:
+\begin{itemize}
+  \item 
+    We present a set of theorems (\Cref{sec:sequential}) that can be used for practical secure compiler verification.
+    Intuitively, we prove that two compilers that robustly preserve two (possibly distinct) properties compose into a compiler that robustly preserves the intersection of the properties.
+    We show special-cases, such as swapping secure optimisation passes or composition of compilers robustly preserving different properties. 
+    Moreover, we discuss how robust preservation relates to vertical compositionality and source-independent linking.
+  \item
+    There is a mechanization of our core-theory in the Coq\footnote{To be renamed into Rocq.} proof assistant and we mark these instances with (\CoqSymbol) throughout the paper.
+  \item 
+    We provide an extensive case-study of different security properties (\Cref{sec:compprop}), a set of programming languages (\Cref{sec:casestud:defs}), and compilers between them (\Cref{sec:casestud:rtp}).
+    The case-study consists of different pen-and-paper verified compiler passes comprised of robust preservation or enforcement mechanisms which individually ensure relevant security properties, such as memory safety in a speculative execution model.
+\end{itemize}
+The paper also provides background (\Cref{sec:background}), some formal insights (\Cref{sec:formalities}) and a discussion of related work (\Cref{sec:relwork}) before concluding (\Cref{sec:concl}).
 
-- Mechanization of core theory.
-
-- Case Study demonstrating composition.
+Pen-and-paper proofs as well as the mechanization are available as supplementary material.

### 2025-02-05 icfp/main.tex
Move Sections around

@@
 
-\section{Background: Properties and Secure Compilers\pages{1}}\label{sec:background}
-
-To introduce the security argument of this paper, this section defines (security) properties, their satisfaction, and their robust satisfaction (i.e., satisfaction in the presence of an active attacker; \Cref{subsec:bg:tprop}).
-Then, borrowing from existing work~\cite{abate2019jour,abate2021extacc}, this section introduces secure compilers as compilers that preserve robust property satisfaction (\Cref{subsec:bg:rtp}).
-
-\subsection{Properties and (Robust) Satisfaction}\label{subsec:bg:tprop}
-
-This paper employs the security model where programs are written in a language whose semantics emits events $\event$.
-Events include security-relevant actions (e.g., reading from and writing to memory, as detailed in \Cref{sec:compprop}) and the unobservable event $\emptyevent$.
-As programs execute, their emitted events are concatenated in traces $\trace$, which serve as the description of the behaviour of a program.%
-\footnote{
-Throughout the paper, sequences are indicated with an overbar (i.e., $\trace$), empty sequences with $\hole{\cdot}$, and concatenation of sequences $\trace[_{1}],\trace[_{2}]$ as $\trace[_{1}]\cdot\trace[_{2}]$.
-Prepending element $\event$ to a sequence $\trace$ uses the same notation: $\event\cdot\trace$.
-}
-
-Properties $\pi$ are sets of traces of admissible program behaviours, ascribing what said property considers valid.
-% 
-The set of all properties can be split-up into different {\em classes} ($\class$), i.e., safety, liveness, and neither safety nor liveness~\cite{clarkson2008hyper}, so a class is simply a set of properties.
-% 
-The compositionality framework (\Cref{sec:sequential}) presented in this paper considers arbitrary classes, while the case-study (\Cref{sec:casestud:rtp}) fixes them to concrete instances of safety properties, since it is decidable whether a trace satisfies a safety property with just a finite trace (i.e., a \emph{prefix}).
-As an example, consider the trace:
-$$\ev{Dealloc\ \loc}\cdot\ev{Read\ \loc\ 1729}\cdot\dots$$
-which describes the interaction with a memory where the deallocation of an address $\loc$ precedes a read (of some value $1729$) at that address in memory.
-% 
-This program behaviour is insecure w.r.t.
-a canonical notion of (temporal) memory safety dictating no use-after-frees of pointers~\cite{nagarakatte2010cets,azevedo2018meaningsofms}, because it reads from a memory location that was freed already.
-The previous finite trace prefix is enough to decide that the trace does not satisfy \gls*{tms} and there is no way to append events to this prefix which would result in the trace being admissible.
-
-In the following, the execution of a whole program $\wholeprogvar$ that terminates in state $\runtimetermvar$ according to the language semantics and produces trace $\trace$ is written as $\progstepto{\wholeprogvar}{\runtimetermvar}{\trace}$.
-With this, we defined property satisfaction as follows:
-\bul{a whole program $\wholeprogvar$ satisfies a property $\pi$} iff \iul{$\wholeprogvar$ yields a trace $\trace$} such that \oul{$\trace$ satisfies $\pi$}.
-
-\begin{definition}[Property Satisfaction]\label{def:propsat}
-$\;$
-
-  \begin{nscenter}
-    \bul{$\sat{\wholeprogvar}{\pi}$}
-    $\isdef$
-    if \iul{$\forall\runtimetermvar\ \trace,\progstepto{\wholeprogvar}{\runtimetermvar}{\trace}$},
-    then \oul{$\trace\in\pi$}.
-  \end{nscenter}
-\end{definition}
-
-Property satisfaction is defined on whole programs, i.e., programs without missing definitions.
-Thus, from a security perspective, \Cref{def:propsat} considers only a passive attacker model, where the attacker observes the execution and, e.g., retrieves secrets from that.
-To consider a stronger model similarly to what existing work does~\cite{abate2019jour,abate2021extacc,maffeis2008code-carrying,gordon2003authenticity,fournet2007authorization,bengtson2011refine,backes2014uniontyps,michael2023mswasm,swasey2017robust,sammler2019benefits}, we extend the concept of satisfaction with {\em robustness}.
-Robust satisfaction considers partial programs $\progvar$, i.e., components with missing imports, which cannot run until said imports are fulfilled.
-To remedy this, {\em linking} takes two partial programs $\progvar[_{1}],\progvar[_{2}]$ and produces a whole program $\wholeprogvar$, i.e., $\link{\progvar[_{1}]}{\progvar[_{2}]}=\wholeprogvar$.
-As typically done in works that consider the execution of partial programs~\cite{abate2019jour,devriese2018parametricity,patrignani2021rsc,korashy2021capableptrs,strydonck2019lincap,devriese2017modular,bowman2015noninterference,ahmed2011equivcps,patterson2017linkingtyps},
-this paper assumes that whole programs are the result of linking partial programs referred to as {\em context} ($\ctx$) and {\em component} ($\comp$).
-The context is an arbitrary program and thus has the role of an {\em attacker} that can interact with the component by means of any features the programming language has, and the component is what is security-relevant.
-%In this work, the semantics of the programming language is expected to differentiate between the component and the context.
-With this, \Thmref{def:propsat} can be extended as follows: for \bul{a component $\progvar$ to robustly satisfy a property $\pi$}, take an \iul{attacker context $\contextvar$ and link it with $\progvar$}, \oul{the resulting whole program must satisfy $\pi$}.
-
-\begin{definition}[Robust Satisfaction]\label{def:proprsat}
-  $\;$
-
-  \begin{nscenter}
-  \bul{$\rsat{\progvar}{\pi}$}
-  $\isdef$ \iul{$\forall \contextvar\ \wholeprogvar$, if $\link{\contextvar}{\progvar}=\wholeprogvar$}, then \oul{$\sat{\wholeprogvar}{\pi}$}.
-  \end{nscenter}
-\end{definition}
-
-% \begin{exampleenv}[Double Free in Bluetooth Subsystem]
-%   Consider \texttt{CVE-2021-3564}~\cite{doublefree-bluetooth}, one of many submissions for a double-free vulnerability.
-%   The vulnerability arises due to a race condition where the context-level function \texttt{hci\_cmd\_work} was not expected to behave maliciously, since it resides in the same source-code repository where the vulnerability occurs.
-%   Nevertheless, the component-level code of \texttt{hci\_dev\_do\_open} is linked with \texttt{hci\_cmd\_work} and does not atomically check whether a pointer has been freed already:
-%   Therefore, \texttt{hci\_dev\_do\_open} does not satisfy the no-double-frees property robustly, since there is an implementation for \texttt{hci\_cmd\_work} that leads to a violation of that property when linked with \texttt{hci\_dev\_do\_open}.
-% \end{exampleenv}
-
-\subsection{Secure Compilers}\label{subsec:bg:rtp}
-
-A {\em compiler} ($\cc{\src{L}}{\trg{L}}$) translates syntactic descriptions of programs from a {\em source} ($\src{L}$) into a {\em target} ($\trg{L}$) programming language.
-This translation is considered {\em correct} if it is semantics-preserving.
-That is, for a whole program $\src{\wholeprogvar}$, the compiler should relate the $\src{L}$ semantics of $\src{\wholeprogvar}$ with the semantics of $\trg{L}$ of the compiled counterpart of $\src{p}$ in such a way that they are ``compatible''.
-Unfortunately, correct compilers may be insecure compilers~\cite{patrignani2019survey,kennedy2006secure.net,abadi1999protect,ahmed2018dagstuhl} and programs translated by insecure compilers can violate security properties that the programmer assumes to hold.
-This is why {\em robust preservation} is a good candidate as a compiler-level security property~\cite{abate2019jour}.
-
-This paper uses a general notion of robust preservation~\cite{abate2021extacc} that considers compilers that use languages with different trace models. 
-To this end, considering a source trace $\src{\trace}$ and a target trace $\trg{\trace}$, there is a relation ($\sim$) describing the effect of a corresponding compiler (see \Cref{sec:casestud:rtp}). 
-This relation induces the following two projection functions~\cite{abate2021extacc}: (1) the \emph{existential image} $\tau_\sim\left(\src{\pi}\right)$ and (2) the \emph{universal image} $\sigma_\sim\left(\trg{\pi}\right)$.
-These projections map source-level (resp. target-level) properties to target-level (resp. source-level) properties in a way that identifies the ``same'' property across languages. 
-The case study of this paper uses the universal image, since some considered properties, such as \gls*{ss}, are not definable in a higher-level language that, e.g., does not model speculation.
-% For self-containment, the universal image is defined as follows~\cite{abate2021extacc}:
-% 
-\begin{definition}[Universal Image]
-% [Existential and Universal Image]
-\label{def:universal:img}\label{def:sigma}
-  % \[
-  %   \tau_\sim\left(\src{\pi}\right) := 
-  %     \left\{ 
-  %       \trg{\trace} \mid \exists \src{\trace}\ldotp \src{\trace}\sim\trg{\trace}, \text{ and } \src{\trace}\in\src{\pi} 
-  %     \right\}
-  % \]
-  \[ 
-    \sigma_\sim\left(\trg{\pi}\right) := 
-      \left\{ 
-        \src{\trace} \mid \forall \trg{\trace}\ldotp \text{if }\src{\trace}\sim\trg{\trace}, \text{ then } \trg{\trace}\in\trg{\pi} 
-      \right\}
-  \]
-\end{definition}
-\noindent
-With this projection function, we define a more general version of robust preservation as follows~\cite{abate2021extacc}.
-% For a \bul{compiler $\cc{\src{L}}{\trg{L}}$ to robustly preserve a class of source properties $\src{\class}$}, if for any \rul{property $\src{\pi}$ of that class $\src{\class}$ and programs $\src{p}$ written in $\src{L}$} where \iul{$\src{\progvar}$ robustly satisfies $\src{\pi}$}, then \oul{the compilation of $\src{\progvar}$, $\cc{\src{L}}{\trg{L}}\left(\src{p}\right)$, must robustly satisfy $\tau_\sim\left(\src{\pi}\right)$}.
-% 
-% \begin{definition}[Robust Preservation with $\tau_\sim$]\label{def:rtp:tau}
-%   % $\;$\\
-%   % \vspace{-1em}
-%   % \begin{nscenter}
-%   % Compiler $\cc{\src{L}}{\trg{L}}$ robustly preserves $\class$, 
-%   \bul{$\rtptau{\cc{\src{L}}{\trg{L}}}{\src{\class}}{\sim}$}
-%   %, iff 
-%   $\isdef$
-%   \rul{$\forall \src{\pi}\in\src{\class}, \src{p}\in\src{L},$} if \iul{$\rsat{\src{\progvar}}{\src{\pi}}$}, then \oul{$\rsat{\cc{\src{L}}{\trg{L}}\left(\src{p}\right)}{\tau_\sim{\left(\src{\pi}\right)}}$}.
-%   % \end{nscenter}
-% \end{definition}
-% 
-A \bul{compiler $\cc{\src{L}}{\trg{L}}$ robustly preserves a class of target properties $\trg{\class}$}, if for any \rul{property $\trg{\pi}$ of class $\trg{\class}$ and programs $\src{p}$}, where \iul{$\src{\progvar}$ robustly satisfies $\sigma_\sim\left(\trg{\pi}\right)$}, \oul{for the compilation of $\src{\progvar}$, we have that $\cc{\src{L}}{\trg{L}}\left(\src{p}\right)$ robustly satisfies $\trg{\pi}$}.
-
-\begin{definition}[Robust Preservation with $\sigma_\sim$]\label{def:rtp:sigma}
-  % $\;$\\
-  % \vspace{-1em}
-  % \begin{nscenter}
-  % Compiler $\cc{\src{L}}{\trg{L}}$ robustly preserves $\class$, 
-  \bul{$\rtpsig{\cc{\src{L}}{\trg{L}}}{\trg{\class}}{\sim}$}
-  %, iff 
-  $\isdef$
-  \rul{$\forall \trg{\pi}\in\trg{\class}, \src{p}\in\src{L},$} if \iul{$\rsat{\src{\progvar}}{\sigma_\sim\left(\trg{\pi}\right)}$}, then \oul{$\rsat{\cc{\src{L}}{\trg{L}}\left(\src{p}\right)}{\trg{\pi}}$}.
-  % \end{nscenter}
-\end{definition}
-Note that a class of properties $\class$ can represent just one property $\pi$ by lifting~\cite{clarkson2008hyper} that property to sets of properties, i.e., use the powerset of $\pi$ instead of $\pi$ itself.
-Because of this, this paper may write $\rtptau{\cc{\src{L}}{\trg{L}}}{\trg{\pi}}{\sim}$, even though $\trg{\pi}$ is a property and not a class.
-A similar construction can be used to the projection functions (see \Cref{def:universal:img}) by applying them to the lifted version of $\trg{\pi}$ instead of just $\trg{\pi}$.
-
-In case the trace model is the same for both source and target programs (and thus $\sim$ is equality), we obtain~\cite{abate2019jour}:
-\begin{definition}[Robust Preservation]\label{def:rtp}
-  % $\;$\\
-  % \vspace{-1em}
-  % \begin{nscenter}
-  % Compiler $\cc{\src{L}}{\trg{L}}$ robustly preserves $\class$, 
-  $\;$
-
-  {$\rtp{\cc{\src{L}}{\trg{L}}}{\class}$}
-  %, iff 
-  $\isdef$
-  {$\forall \pi\in\class, \src{p}\in\src{L},$} if {$\rsat{\src{\progvar}}{\pi}$}, then {$\rsat{\cc{\src{L}}{\trg{L}}\left(\src{p}\right)}{\pi}$}.
-  % \end{nscenter}
-\end{definition}
-
-Examples of compilers fulfilling \Cref{def:rtp} exist in the literature~\cite{korashy2022secureptrs,korashy2021capableptrs,abate2021extacc,abate2019jour,patrignani2021rsc}.
-For example, SecurePtrs~\cite{korashy2022secureptrs} gives a compiler that robustly preserves all safety properties for a C-like language to an assembly-like language. 
-% The mechanization has well over 20k lines in total, which demonstrates that proving a compiler secure with respect to \Cref{def:rtp} is a significant effort. \MP{move this line to later ?}
-% i don't see how adding a def of cct is beneficial
-As another example, even though it is not strictly satisfying \Cref{def:rtp}, the FaCT~\cite{cauligi2019fact} compiler preserves the \gls*{cct} property for a C-like language with constant-time primitives, e.g., \texttt{ctselect} for branching.
-%, whose precise definition can be found in \Cref{sec:cct}.
-Throughout this work, it is assumed that FaCT satisfies \Cref{def:rtp}.
-% While we were not able to find compilers in literature that provably fulfill \Cref{def:rtp}, we anticipate that 
- % would satisfy this hyperproperty.
-% FaCT proves constant-timeness by means of overapproximation, i.e., instead of directly proving the hyperproperty, they fall back to a property that overapproximates the corresponding hyperproperty.
-
-%\smallskip%MP: the underline compacts the sections too much
-
-\subsection{Hyperproperties and (Robust) Satisfaction}
-
-Hyperproperties~\cite{clarkson2008hyper} are strictly more expressive than ordinary properties. 
-While a property is a set of traces, a hyperproperty is a set of properties. 
-We can lift properties to hyperproperties by applying the powerset.
-Hyperproperty satisfaction is defined as follows.
-
-It turns out that classes of hyperproperties, that is, sets of hyperproperties, are equally expressive than simply hyperproperties themselves~\cite{clarkson2008hyper}.
-
-% and so on
+\input{sections/background.tex}
 
@@
 
-\section{Conclusion\pages{$\sfrac{1}{2}$}}\label{sec:concl}
-This paper tackles the problem of understanding what kind of security properties a secure compiler preserves, when said compiler is the combination of compiler passes that preserve possibly different security properties.
-% 
-%For this, this paper first formalised security properties of interest and their composition.
-% 
-The paper proves that composing secure compilers that preserve certain properties results in a secure compiler that preserves the composition of these properties.
-% While the presented security property composition relied on monitors that check only trace-properties, the composition of secure compilers does not make any restriction towards the kind of property involved in the composition.
-% It is subject to future work to develop techniques to verify relational hyperproperties~\cite{abate2019jour}, while the composition of hyperproperties could be very similar to the composition of ordinary properties as presented in this paper, since hyperproperties can be checked with an automata-based model-checker~\cite{beutner23hyperltl}.
-% 
-Finally, this paper defines a multi-pass compiler and proves that it preserves \gls*{specms}.
-Crucially, this paper derives the security of the multi-pass compiler from the composition of the security properties preserved by its individual passes, which include security-preserving as well as optimisation passes.
-% For future work, it is interesting to look at more sophisticated optimisation passes that, e.g., reorder events that appear on traces.
+\input{sections/conclusion.tex}
 

### 2025-02-05 icfp/sections/background.tex
Move Sections around

--- 2024-09-24 csf2/main.tex
@@
-% Fixing: Too many math alphabets used in version normal
-\newcommand\hmmax{0}
-\newcommand\bmmax{0}
-\documentclass[dvipsnames,conference]{IEEEtran}
-
-\usepackage[T1]{fontenc}
-\usepackage[scaled=.83]{beramono}
-
-\usepackage[colorinlistoftodos]{todonotes}
-\usepackage[inference]{semantic}
-\usepackage[switch]{lineno}
-\usepackage{halloweenmath}
-\usepackage{fontawesome5}
-\usepackage{listofitems}
-\usepackage{breakcites}
-\usepackage{glossaries}
-\usepackage{hyperref}
-\usepackage{cleveref}
-\usepackage{stmaryrd}
-\usepackage{marvosym}
-\usepackage{listings}
-\usepackage{amssymb}
-\usepackage{amsmath}
-\usepackage{nameref}
-\usepackage{amsthm}
-\usepackage{xspace}
-\usepackage{xfrac}
-\usepackage{tikz}
-\usepackage{soul}
-\usepackage{bm}
-\usepackage{paralist}
-
-\hypersetup{
-    colorlinks,
-    linkcolor={red!50!black},
-    citecolor={blue!50!black},
-    urlcolor={blue!80!black}
-}
-%\renewcommand\UrlFont{\color{blue}\rmfamily}
-%\newcommand{\url}[1]{\lstinline{#1}}
-
-\setul{0.95ex}{0.3ex}
-
-\input{cmds.tex}
-
-% allow page breaks inside multiline alignment displays
-\allowdisplaybreaks
-
-\Crefname{exampleenv}{Example}{Examples}
-
-\theoremstyle{definition}
-\newtheorem{exampleenv}{Example}[section]
-\newtheorem{lemma}{Lemma}[section]
-\newtheorem{theorem}{Theorem}[section]
-\newtheorem{corollary}{Corollary}[section]
-\newtheorem{definition}{Definition}[section]
-
-\loadglsentries{acronyms}
-\makeglossaries
-
-\AtBeginDocument{%
-  \providecommand\BibTeX{{%
-    \normalfont B\kern-0.5em{\scshape i\kern-0.25em b}\kern-0.8em\TeX}}}
-\makeatletter
-\newcommand{\linebreakand}{%
-  \end{@IEEEauthorhalign}
-  \hfill\mbox{}\par
-  \mbox{}\hfill\begin{@IEEEauthorhalign}
-}
-\makeatother
-
-\renewcommand\thelinenumber{\color{red}\arabic{linenumber}}
-\begin{document}
-\linenumbers
-
-\title{
-  Secure Composition 
-  \\ 
-  of Robust and Optimising Compilers
-  % \vspace{-1.5em}
-}
-
-% \author{
-% Matthis Kruse\orcidID{0000-0003-4062-9666}
-% \and
-% Michael Backes
-% \and
-% Marco Patrignani
-% }
-% \authorrunning{M. Kruse et al.}
-
-\maketitle
-
-\begin{abstract}
-% \MPpost{
-%   this abstract fails in setting up the robustness argument.
-%   it'll be a good exercise for you to take inspiration from this, but then write something as concise but with a more direct focus on robust properties
-% }
-% context
-To ensure that secure applications do not leak their secrets, they are required to uphold several security properties such as spatial and temporal memory safety, cryptographic constant time, as well as speculative safety.
-% need
-Existing work shows how to enforce these properties individually, in an architecture-independent way, by using secure compiler passes that each focus on an individual property.
-% task
-Unfortunately, given two secure compiler passes that each preserve a possibly different security property, it is unclear what kind of security property is preserved by the composition of those secure compiler passes.
-%there is no way to tell what kind of security property will the composition of those secure compilers preserve.
-
-% object
-This paper is the first to study what security properties are preserved across the composition of different secure compiler passes.
-% findings
-Starting from a general theory of property composition for security-relevant properties (such as the aforementioned ones), this paper formalises a theory of composition of secure compilers.
-Then, it showcases this theory on a secure multi-pass compiler that preserves the aforementioned security-relevant properties.
-% conclusion
-Crucially, this paper derives the security of the multi-pass compiler from the composition of the security properties preserved by its individual passes, which include security-preserving as well as optimisation passes.
-% 
-From an engineering perspective, this is the desirable approach to building secure compilers.
-\begin{center}\small\it
-	{This paper uses syntax highlighting accessible to both colourblind and black \& white readers.
-	% ~\cite{patrignani2020use}.
-% 	% Specifically, it makes use of a $\src{blue}$, $\src{sans\text{-}serif}$ font for a $\src{source}$,
-% 	% a $\trg{red}$, $\trg{bold}$ font for an $\trg{intermediate}$,
-% 	% and a $\obj{green}$, $\obj{teletype}$ font for a $\obj{target}$ language.	
-	For a better experience, please print or view in colours.
-	}
-\end{center}
-\end{abstract}
-
-\begin{IEEEkeywords}
-  memory-safety, cryptographic constant-time, secure compilation, robustness
-\end{IEEEkeywords}
-
-\section{Introduction\pages{4}}\label{sec:introduction}
-
-% \gls*{ms} is a security property that ensures that there are no out-of-bounds accesses as well as no reads or writes to uninitialised memory.
-% Programs that do not attain memory safety, can easily have their private data be leaked to attackers or be corrupted~\cite{lemay2021ccc}.
-% Because of this, \gls*{ms} is a requirement for programs to attain an even stronger security property, namely \gls*{cct}.
-% Programs that attain \gls*{cct} are secure w.r.t.
-timing attacks, and this is the gold standard for secure applications, including cryptographic ones, since timing attacks can expose secrets via a simple performance analysis~\cite{kocher1996timing}.
-% Timing attacks have become more and more relevant recently, especially given the era of cloud-computing~\cite{aviram2010cloudtime,kumar2019cloudsecsurvey}, where programs share the same hardware but are supposed to run in a sandboxed environment that does not leak secrets.
-% Unfortunately, several attacks within such a cloud environment have been demonstrated as viable strategies to extract secrets~\cite{mehmet2015getoff,flowers2022zeroday,atya2019catchme,venkatanathan2015placevul}.
-
-% Even though mitigations for such attacks exist~\cite{bond2017vale,almeida2017jasmin}, some compilers may optimise these mitigations away~\cite{barthe2018sec}.
-% This is of course not what programmers expect and to ensure such mitigations persist in the compiled code, one can use a {\em secure} compilers.
-% More generally, secure compilers should preserve security properties even when the program runs in a hostile context such as the cloud one, i.e., where it interacts with other possibly malicious programs.
-% An example of such secure compilers are those that ascribe to the theory of robust compilation~\cite{abate2019jour,abate2021extacc,patrignani2021rsc}, which takes this interaction between components and potentially malicious contexts into account.
-
-% Unfortunately, it is unclear if and how robust compilation supports compositionality, and compilation passes are typically developed as a standalone module to be plugged into a compilation pipeline and composed with other passes.
-% Different kinds of compilation passes exist, e.g., there are usually several compilation passes that perform optimisations~\cite{lattner2004llvm,googlev8,androidstudio,kuepper2023cryptopt,manjikian1997fusion,wegman1991ccp} or source-code instrumentations~\cite{nagarakatte2009soft,nagarakatte2010cets,akritidis2009baggy,dhumbumroong2020boundwarden,jung2021pico,nam2019framer,shankaranarayana2023tailcheck,younan2010paricheck,zhou2023fatptrs,bond2017vale,almeida2017jasmin,kuepper2023cryptopt,cauligi2019fact} that enforce that the compiled program fulfils a property of interest.
-% Without a framework for composing secure compiler passes, it is not possible to enable separation of concerns, e.g., to have a secure compilation pass that ensures \gls*{ms} that is developed independently of another secure pass for \gls*{cct} as suggested before and as showcased in \Cref{ex:strncpy}.
-% \MPpost{
-%   this intro does a poor job at introducing the robust compilation argument.
-%   perhaps it'd be best to start from robust properties, which motivates the use of robust compilers?
-%   idk
-%   as for the abstract, this is a good starting point for you to make a better intro
-% }
-
-\gls*{ms} is a security property obtained by composing \gls*{sms}, which ensures array accesses are all within bounds, and \gls*{tms}, which ensures pointers are only used when they are valid~\cite{azevedo2018meaningsofms,jim2002cyclone,necula2005ccured,nagarakatte2010cets,nagarakatte2009soft,akritidis2009baggy,michael2023mswasm}.
-\gls*{cct} is a security property that ensures sensitive data is not leaked via timing side-channels~\cite{kocher1996timing} and \gls*{ss} is a security property that enforces the same but under a speculative semantics~\cite{guarnieri2018spectector,fabian2022automatic} that captures speculative execution attacks such as spectre~\cite{kocher2019spectre}.
-Together, \gls*{sms}, \gls*{tms}, \gls*{cct} and \gls*{ss}, yield \gls*{specms}, which is the gold standard of security properties for secure applications.
-Programs attaining \gls*{specms} do not leak sensitive data either through erroneous memory accesses, nor through timing side-channels, even under speculative execution.
-\Cref{ex:strncpy} below discusses how these security properties can be enforced by compiler passes~\cite{bond2017vale,almeida2017jasmin}, to ensure programmers need not be aware of the architectural details of where their code will run.
-
-\begin{exampleenv}[strncpy]\label{ex:strncpy}
-Consider the following \texttt{C} function \texttt{strncpy} that copies a null-terminated string \texttt{src} into \texttt{dst} up to a length of \texttt{n} characters.
-This function is subject to a subtle \gls*{sms} vulnerability: the bounds check \texttt{i < n} should happen {\it before} the access to memory location \texttt{src[i]}: otherwise
-% If \texttt{n} is at least as large as the length of the array \texttt{src}, 
-the memory location past the last element is leaked to an attacker.
-\begin{lstlisting}[language=c,basicstyle=\small\ttfamily,morekeywords={size_t}]
-void strncpy(size_t n, char *dst, char *src) {
-  for(size_t i = 0; src[i] != '\0' && i < n; ++i) {
-    dst[i] = src[i];
-  }
-}
-\end{lstlisting}
-% Assume that the length of the array pointed by \texttt{dst} is at least \texttt{n}, and \texttt{src} is terminated by \texttt{'\textbackslash 0'}.
-% Even under these assumptions, there is a 
-To prevent this vulnerability, one can use a compilation pass that enforces \gls*{sms}, such as Softbounds~\cite{nagarakatte2009soft} or BaggyBounds~\cite{akritidis2009baggy}.
-The most na\"ive solution in this case is to insert bounds-checks in front of every access to memory.
-
-Because of timing attacks, addressing \gls*{sms} is not enough to make \texttt{strncpy} secure.
-In fact, the loop can terminate early, as soon as the string-terminating character \texttt{'\textbackslash 0'} is encountered, thus making program execution time proportional to the length of the array pointed by \texttt{src}, and violating the \gls*{cct} property.
-% Often, programmers rewrite programs like these to have loops that run with a fixed number of iterations.
-Also in this case there exist compiler passes that can rewrite such programs into \gls*{cct} ones~\cite{cauligi2019fact}.
- % so that they are no longer vulnerable to timing attacks.
-%
-Finally, even with these precautions, code is not run in isolation, so a malicious attacker could supply code that interacts with \texttt{strncpy} and trigger a violation of either \gls*{ms} or \gls*{cct} by calling \texttt{strncpy} with an argument for \texttt{src} that points to uninitialised memory.
-This would, in turn, trigger a series of reads from uninitialised memory, which is an immediate \gls*{ms} violation with devastating real-world consequences~\cite{uninit-0,uninit-1,uninit-2,uninit-3,uninit-4}.
-% An easy fix to sidestep this is to use non-nullable pointers as parameters instead, e.g., as presented in Cyclone~\cite{jim2002cyclone}.
-\end{exampleenv}
-
-  % last sentence is before the example, which talks about compiler passes enforcing props
-Whether or not compiler passes enforce certain security properties, it is important for a secure compiler to consider partial programs that interact with potentially malicious code, since the latter may lead to, e.g., memory-safety issues in the considered partial program.
-Robust compilers~\cite{abate2019jour} are a form of secure compilers that preserve security properties even in the presence of arbitrary attackers interacting with compiled code.
-Thus, robust compilers can be used to prevent vulnerabilities resulting from uninitialised memory (as well as many other ones), e.g., by targeting capability-based languages such as CHERI~\cite{woodruff2014CHERI}, Arm Morello~\cite{arm-morello}, or MSWasm~\cite{michael2023mswasm}, where the compiler relies on capabilities to check that pointers are always initialised.
-
-% To ensure these compiler-preserved properties still hold even in the presence of arbitrary attackers interacting with compiled code, one can use secure compilers in the form of robust compilers~\cite{abate2019jour,abate2021extacc,patrignani2021rsc}.
-
-Unfortunately, given secure compiler passes that each preserve a possibly different security property, there is no way to tell what kind of security property will the composition of those secure compilers preserve.
-% Different kinds of compilation passes exist, e.g., there are usually several compilation passes that perform optimisations~\cite{lattner2004llvm,googlev8,androidstudio,kuepper2023cryptopt,manjikian1997fusion,wegman1991ccp} or source-code instrumentations~\cite{nagarakatte2009soft,nagarakatte2010cets,akritidis2009baggy,dhumbumroong2020boundwarden,jung2021pico,nam2019framer,shankaranarayana2023tailcheck,younan2010paricheck,zhou2023fatptrs,bond2017vale,almeida2017jasmin,kuepper2023cryptopt,cauligi2019fact} that enforce that the compiled program fulfils a property of interest.
-Worse, without a framework for composing secure compiler passes, it is not possible to enable separation of concerns, e.g., to have a secure compilation pass that ensures \gls*{ms} that is developed independently of another secure pass for \gls*{cct}, that is developed independently of other passes, such as optimisation ones.
-
-This paper introduces a framework for reasoning about the composition of secure and optimising compiler passes %akin to those of \Cref{ex:strncpy}\MP{there aren't really passes there} 
-and it showcases the power of this framework by instantiating it on a multi-pass compilation chain.
-To this end, this paper first discusses how to compose security properties, such as \gls*{tms} and \gls*{sms} into \gls*{ms}, and then adds \gls*{cct} as well as \gls*{ss} into the mix to obtain \gls*{specms}.
-%Then, this paper defines compiler composition and formalises that given two passes that securely preserve two (possibly distinct) properties, their composition securely preserves the composition of those properties.
-The paper then defines several secure compiler passes, where each is either preserving a different security property from the list above
-% (\gls*{tms}, \gls*{sms}, \gls*{scct}, \gls*{ss}) 
-or performing a security-preserving optimisation.
-% , (e.g., applying \gls*{cf} or \gls*{dce}).
-Finally, this paper shows that composing these secure compiler passes into a multi-pass compilation chain results in the end-to-end (robust) preservation of \gls*{specms}.
-Crucially, this paper derives the security of the multi-pass compiler from the composition of the security properties preserved by its individual passes.
-% Crucially, this end-to-end security preservation is obtained by composing the security property preserved by each individual pass, as dictated by the framework of this paper.
-This result showcases how the framework allows the kind of formal security reasoning that compiler writers already want (and already do), obtaining precise, compositional security reasoning while providing minimal (and modular) proof effort.
-%
-In summary, this paper makes the following contributions:
-% \begin{itemize}[leftmargin=0em,nosep,label=$\blacktriangleright$]
-\begin{asparaitem}[$\blacktriangleright$]
-  \item %
-        %\Cref{sec:sequential} studies different forms of secure compiler compositions.
-        This paper takes the secure compilation framework of Abate et al.~\cite{abate2021extacc} and extends it for compositionality. (\Cref{sec:sequential}).
-        %Concretely, it investigates upper and lower compositions, as well as the sequential composition.
-        %For this, this paper studies compiler composition as well as compilers with multiple input languages or multiple output ones, as used in existing compilation chains.
-        %
-        This paper proves that starting from two compilers that preserve two (possibly distinct) properties, their composition preserves roughly the intersection of those properties.
-        % For example, a compiler that is secure w.r.t. \gls*{tms} can be composed with a compiler that is secure w.r.t. \gls*{sms}, resulting in a compiler that is secure w.r.t. \gls*{ms}.
-        % 
-        % From this, the paper proves that the order of composition of sequential compiler passes is irrelevant for security.
-        % % 
-        % This is crucial for reordering individual optimisation passes and thus generating secure and efficient code.
-        %
-        %Finally, this paper extends parts of the compositionality framework to support different trace-models between compilation passes.
-        % \MP{this isn't really true, is it?. the order matters for the relations' lemmas}
-        Then, this paper identifies which conditions make the composition of secure compilers meaningful from a security perspective, and which other conditions allow passes to be swapped without losing security meaning.
-
-  \item %
-        This paper presents a case-study with five programming languages showcasing the previous contribution (\Cref{sec:casestud:defs,sec:casestud:rtp}).
-        To this end, it presents a compilation chain consisting of six passes that ultimately preserve \gls*{specms} by means of composing four secure passes that \emph{individually} preserve \gls*{tms}, \gls*{sms}, \gls*{scct}---an enforceable version of \gls*{cct}---, and \gls*{ss}, and two passes which are well-known optimisations \gls*{dce} and \gls*{cf}.
-        %Furthermore, the chain includes two optimisation passes: one performs \gls*{dce} and the other \gls*{cf}.
-        The formalisation of this case study showcases the power of the presented framework: the divide-and-conquer approach to software engineering is a viable strategy even for the development of secure compilers.
-%        This demonstrates that the usage of the presented framework is both simpler and more expressive compared to one monolithic security proof of the whole compilation chain.
-
-  \item %
-  		This paper provides formal proofs of the security properties preserved by each of the presented passes. 
-  		% 
-  		Additionally this paper analyses each pass in terms of their compatibility with each other (\Cref{sec:formalities}), proving how does each pass fulfill the conditions that make their composition meaningful from a security standpoint.
-        % In doing so, the paper provides formal insights into some technicalities of the concrete proofs necessary as technical setup to reap the benefits of the presented compositionality framework.
-
-  \item The key contributions of this paper are formalised in the Coq proof assistant and the paper indicates this with \CoqSymbol.
-\end{asparaitem}
-% 
-This paper starts by introducing relevant notions of security properties and secure compilation (\Cref{sec:background}),
-and discusses related work (\Cref{sec:relwork}) before concluding (\Cref{sec:concl}).
-
-% \contribution{Technical Report} 
-The omitted formal details, lemmas and proofs, as well as the Coq formalisation are available as supplementary material.
-% \MKin{cite/link}
-
 \section{Background: Properties and Secure Compilers\pages{1}}\label{sec:background}
@@
 
-\subsection{Properties and (Robust) Satisfaction}\label{subsec:bg:tprop}
-
-This paper employs the security model where programs are written in a language whose semantics emits events $\event$.
-Events include security-relevant actions (e.g., reading from and writing to memory, as detailed in \Cref{sec:compprop}) and the unobservable event $\emptyevent$.
-As programs execute, their emitted events are concatenated in traces $\trace$, which serve as the description of the behaviour of a program.%
-\footnote{
-Throughout the paper, sequences are indicated with an overbar (i.e., $\trace$), empty sequences with $\hole{\cdot}$, and concatenation of sequences $\trace[_{1}],\trace[_{2}]$ as $\trace[_{1}]\cdot\trace[_{2}]$.
-Prepending element $\event$ to a sequence $\trace$ uses the same notation: $\event\cdot\trace$.
-}
-
-Properties $\pi$ are sets of traces of admissible program behaviours, ascribing what said property considers valid.
-% 
-The set of all properties can be partitioned into different {\em classes} ($\class$), i.e., safety, liveness, and neither safety nor liveness~\cite{clarkson2008hyper}, so a class is simply a set of properties.
-% 
-The compositionality framework (\Cref{sec:sequential}) presented in this paper considers arbitrary classes, while the case-study (\Cref{sec:casestud:rtp}) fixes them to concrete instances of safety properties, since it is decidable whether a trace satisfies a safety property with just a finite trace (i.e., a \emph{prefix}).
-As an example, consider the trace:
-$$\ev{Dealloc\ \loc}\cdot\ev{Read\ \loc\ 1729}\cdot\dots$$
-which describes the interaction with a memory where the deallocation of an address $\loc$ precedes a read (of some value $1729$) at that address in memory.
-% 
-This program behaviour is insecure w.r.t.
-a canonical notion of (temporal) memory safety dictating no use-after-frees of pointers~\cite{nagarakatte2010cets,azevedo2018meaningsofms}, because it reads from a memory location that was freed already.
-The previous finite trace prefix is enough to decide that the trace does not satisfy \gls*{tms} and there is no way to append events to this prefix which would result in the trace being admissible.
-
-In the following, the execution of a whole program $\wholeprogvar$ that terminates in state $\runtimetermvar$ according to the language semantics and produces trace $\trace$ is written as $\progstepto{\wholeprogvar}{\runtimetermvar}{\trace}$.
-With this, we defined property satisfaction as follows:
-\bul{a whole program $\wholeprogvar$ satisfies a property $\pi$} iff \iul{$\wholeprogvar$ yields a trace $\trace$} such that \oul{$\trace$ satisfies $\pi$}.
-
-\begin{definition}[Property Satisfaction]\label{def:propsat}
-$\;$
-
-  \begin{nscenter}
-    \bul{$\sat{\wholeprogvar}{\pi}$}
-    $\isdef$
-    if \iul{$\forall\runtimetermvar\ \trace,\progstepto{\wholeprogvar}{\runtimetermvar}{\trace}$},
-    then \oul{$\trace\in\pi$}.
-  \end{nscenter}
-\end{definition}
-
-Property satisfaction is defined on whole programs, i.e., programs without missing definitions.
-Thus, from a security perspective, \Cref{def:propsat} considers only a passive attacker model, where the attacker observes the execution and, e.g., retrieves secrets from that.
-To consider a stronger model similarly to what existing work does~\cite{abate2019jour,abate2021extacc,maffeis2008code-carrying,gordon2003authenticity,fournet2007authorization,bengtson2011refine,backes2014uniontyps,michael2023mswasm,swasey2017robust,sammler2019benefits}, we extend the concept of satisfaction with {\em robustness}.
-Robust satisfaction considers partial programs $\progvar$, i.e., components with missing imports, which cannot run until said imports are fulfilled.
-To remedy this, {\em linking} takes two partial programs $\progvar[_{1}],\progvar[_{2}]$ and produces a whole program $\wholeprogvar$, i.e., $\link{\progvar[_{1}]}{\progvar[_{2}]}=\wholeprogvar$.
-As typically done in works that consider the execution of partial programs~\cite{abate2019jour,devriese2018parametricity,patrignani2021rsc,korashy2021capableptrs,strydonck2019lincap,devriese2017modular,bowman2015noninterference,ahmed2011equivcps,patterson2017linkingtyps},
-this paper assumes that whole programs are the result of linking partial programs referred to as {\em context} ($\ctx$) and {\em component} ($\comp$).
-The context is an arbitrary program and thus has the role of an {\em attacker} that can interact with the component by means of any features the programming language has, and the component is what is security-relevant.
-%In this work, the semantics of the programming language is expected to differentiate between the component and the context.
-With this, \Thmref{def:propsat} can be extended as follows: for \bul{a component $\progvar$ to robustly satisfy a property $\pi$}, take an \iul{attacker context $\contextvar$ and link it with $\progvar$}, \oul{the resulting whole program must satisfy $\pi$}.
-
-\begin{definition}[Robust Satisfaction]\label{def:proprsat}
-  $\;$
-
-  \begin{nscenter}
-  \bul{$\rsat{\progvar}{\pi}$}
-  $\isdef$ \iul{$\forall \contextvar\ \wholeprogvar$, if $\link{\contextvar}{\progvar}=\wholeprogvar$}, then \oul{$\sat{\wholeprogvar}{\pi}$}.
-  \end{nscenter}
-\end{definition}
-
-% \begin{exampleenv}[Double Free in Bluetooth Subsystem]
-%   Consider \texttt{CVE-2021-3564}~\cite{doublefree-bluetooth}, one of many submissions for a double-free vulnerability.
-%   The vulnerability arises due to a race condition where the context-level function \texttt{hci\_cmd\_work} was not expected to behave maliciously, since it resides in the same source-code repository where the vulnerability occurs.
-%   Nevertheless, the component-level code of \texttt{hci\_dev\_do\_open} is linked with \texttt{hci\_cmd\_work} and does not atomically check whether a pointer has been freed already:
-%   Therefore, \texttt{hci\_dev\_do\_open} does not satisfy the no-double-frees property robustly, since there is an implementation for \texttt{hci\_cmd\_work} that leads to a violation of that property when linked with \texttt{hci\_dev\_do\_open}.
-% \end{exampleenv}
-
-\subsection{Secure Compilers}\label{subsec:bg:rtp}
-
-A {\em compiler} ($\cc{\src{L}}{\trg{L}}$) translates syntactic descriptions of programs from a {\em source} ($\src{L}$) into a {\em target} ($\trg{L}$) programming language.
-This translation is considered {\em correct} if it is semantics-preserving.
-That is, for a whole program $\src{\wholeprogvar}$, the compiler should relate the $\src{L}$ semantics of $\src{\wholeprogvar}$ with the semantics of $\trg{L}$ of the compiled counterpart of $\src{p}$ in such a way that they are ``compatible''.
-Unfortunately, correct compilers may be insecure compilers~\cite{patrignani2019survey,kennedy2006secure.net,abadi1999protect,ahmed2018dagstuhl} and programs translated by insecure compilers can violate security properties that the programmer assumes to hold.
-This is why {\em robust preservation} is a good candidate as a compiler-level security property~\cite{abate2019jour}.
-
-This paper uses a general notion of robust preservation~\cite{abate2021extacc} that considers compilers that use languages with different trace models. 
-To this end, considering a source trace $\src{\trace}$ and a target trace $\trg{\trace}$, there is a relation ($\sim$) describing the effect of a corresponding compiler (see \Cref{sec:casestud:rtp}). 
-This relation induces the following two projection functions~\cite{abate2021extacc}: (1) the \emph{existential image} $\tau_\sim\left(\src{\pi}\right)$ and (2) the \emph{universal image} $\sigma_\sim\left(\trg{\pi}\right)$.
-These projections map source-level (resp. target-level) properties to target-level (resp. source-level) properties in a way that identifies the ``same'' property across languages. 
-The case study of this paper uses the universal image, since some considered properties, such as \gls*{ss}, are not definable in a higher-level language that, e.g., does not model speculation.
-% For self-containment, the universal image is defined as follows~\cite{abate2021extacc}:
-% 
-\begin{definition}[Universal Image]
-% [Existential and Universal Image]
-\label{def:universal:img}\label{def:sigma}
-  % \[
-  %   \tau_\sim\left(\src{\pi}\right) := 
-  %     \left\{ 
-  %       \trg{\trace} \mid \exists \src{\trace}\ldotp \src{\trace}\sim\trg{\trace}, \text{ and } \src{\trace}\in\src{\pi} 
-  %     \right\}
-  % \]
-  \[ 
-    \sigma_\sim\left(\trg{\pi}\right) := 
-      \left\{ 
-        \src{\trace} \mid \forall \trg{\trace}\ldotp \text{if }\src{\trace}\sim\trg{\trace}, \text{ then } \trg{\trace}\in\trg{\pi} 
-      \right\}
-  \]
-\end{definition}
-\noindent
-With this projection function, we define a more general version of robust preservation as follows~\cite{abate2021extacc}.
-% For a \bul{compiler $\cc{\src{L}}{\trg{L}}$ to robustly preserve a class of source properties $\src{\class}$}, if for any \rul{property $\src{\pi}$ of that class $\src{\class}$ and programs $\src{p}$ written in $\src{L}$} where \iul{$\src{\progvar}$ robustly satisfies $\src{\pi}$}, then \oul{the compilation of $\src{\progvar}$, $\cc{\src{L}}{\trg{L}}\left(\src{p}\right)$, must robustly satisfy $\tau_\sim\left(\src{\pi}\right)$}.
-% 
-% \begin{definition}[Robust Preservation with $\tau_\sim$]\label{def:rtp:tau}
-%   % $\;$\\
-%   % \vspace{-1em}
-%   % \begin{nscenter}
-%   % Compiler $\cc{\src{L}}{\trg{L}}$ robustly preserves $\class$, 
-%   \bul{$\rtptau{\cc{\src{L}}{\trg{L}}}{\src{\class}}{\sim}$}
-%   %, iff 
-%   $\isdef$
-%   \rul{$\forall \src{\pi}\in\src{\class}, \src{p}\in\src{L},$} if \iul{$\rsat{\src{\progvar}}{\src{\pi}}$}, then \oul{$\rsat{\cc{\src{L}}{\trg{L}}\left(\src{p}\right)}{\tau_\sim{\left(\src{\pi}\right)}}$}.
-%   % \end{nscenter}
-% \end{definition}
-% 
-A \bul{compiler $\cc{\src{L}}{\trg{L}}$ robustly preserves a class of target properties $\trg{\class}$}, if for any \rul{property $\trg{\pi}$ of class $\trg{\class}$ and programs $\src{p}$}, where \iul{$\src{\progvar}$ robustly satisfies $\sigma_\sim\left(\trg{\pi}\right)$}, \oul{the compilation of $\src{\progvar}$, we have that $\cc{\src{L}}{\trg{L}}\left(\src{p}\right)$ robustly satisfies $\trg{\pi}$}.
-
-\begin{definition}[Robust Preservation with $\sigma_\sim$]\label{def:rtp:sigma}
-  % $\;$\\
-  % \vspace{-1em}
-  % \begin{nscenter}
-  % Compiler $\cc{\src{L}}{\trg{L}}$ robustly preserves $\class$, 
-  \bul{$\rtpsig{\cc{\src{L}}{\trg{L}}}{\trg{\class}}{\sim}$}
-  %, iff 
-  $\isdef$
-  \rul{$\forall \trg{\pi}\in\trg{\class}, \src{p}\in\src{L},$} if \iul{$\rsat{\src{\progvar}}{\sigma_\sim\left(\trg{\pi}\right)}$}, then \oul{$\rsat{\cc{\src{L}}{\trg{L}}\left(\src{p}\right)}{\trg{\pi}}$}.
-  % \end{nscenter}
-\end{definition}
-Note that a class of properties $\class$ can represent just one property $\pi$ by lifting~\cite{clarkson2008hyper} that property to sets of properties, i.e., use the powerset of $\pi$ instead of $\pi$ itself.
-Because of this, this paper may write $\rtptau{\cc{\src{L}}{\trg{L}}}{\trg{\pi}}{\sim}$, even though $\trg{\pi}$ is a property and not a class.
-A similar construction can be used to the projection functions (see \Cref{def:universal:img}) by applying them to the lifted version of $\trg{\pi}$ instead of just $\trg{\pi}$.
-
-In case the trace model is the same for both source and target programs (and thus $\sim$ is equality), we obtain~\cite{abate2019jour}:
-\begin{definition}[Robust Preservation]\label{def:rtp}
-  % $\;$\\
-  % \vspace{-1em}
-  % \begin{nscenter}
-  % Compiler $\cc{\src{L}}{\trg{L}}$ robustly preserves $\class$, 
-  $\;$
-
-  {$\rtp{\cc{\src{L}}{\trg{L}}}{\class}$}
-  %, iff 
-  $\isdef$
-  {$\forall \pi\in\class, \src{p}\in\src{L},$} if {$\rsat{\src{\progvar}}{\pi}$}, then {$\rsat{\cc{\src{L}}{\trg{L}}\left(\src{p}\right)}{\pi}$}.
-  % \end{nscenter}
-\end{definition}
-
-Examples of compilers fulfilling \Cref{def:rtp} exist in the literature~\cite{korashy2022secureptrs,korashy2021capableptrs,abate2021extacc,abate2019jour,patrignani2021rsc}.
-For example, SecurePtrs~\cite{korashy2022secureptrs} gives a compiler that robustly preserves all safety properties for a C-like language to an assembly-like language. 
-% The mechanization has well over 20k lines in total, which demonstrates that proving a compiler secure with respect to \Cref{def:rtp} is a significant effort. \MP{move this line to later ?}
-% i don't see how adding a def of cct is beneficial
-As another example, even though it is not strictly satisfying \Cref{def:rtp}, the FaCT~\cite{cauligi2019fact} compiler preserves the \gls*{cct} property for a C-like language with constant-time primitives, e.g., \texttt{ctselect} for branching.
-%, whose precise definition can be found in \Cref{sec:cct}.
-Throughout this work, it is assumed that FaCT satisfies \Cref{def:rtp}.
-% While we were not able to find compilers in literature that provably fulfill \Cref{def:rtp}, we anticipate that 
- % would satisfy this hyperproperty.
-% FaCT proves constant-timeness by means of overapproximation, i.e., instead of directly proving the hyperproperty, they fall back to a property that overapproximates the corresponding hyperproperty.
-
-%\smallskip%MP: the underline compacts the sections too much
-
-\section{Secure Composition}\label{sec:sequential}
-
-Notably, real-world compilation chains also perform a series of (sequential) passes whose main purpose is not necessarily to translate from one language to another, but to, e.g., optimise the code or enforce a certain property.
-Both examples can be seen in practice, e.g.,~\cite{nagarakatte2009soft,nagarakatte2010cets,akritidis2009baggy,wegman1991ccp,manjikian1997fusion} and many more.
-% \MPin{survey to cite?
-this is a bit informal MK: hard to find a survey on different compiler passes...
-I can't even find one just for compiler enforced memsafety (that looks good) or for compiler translations in general}.
-% 
-% Consider the compilation passes of LLVM~\cite{lattner2004llvm}, which are handled in the \texttt{PassManager} class.
-% By manipulating the internal state of a \texttt{PassManager} object, the schedule of compiler optimisation passes can be hardcoded.
-% That way, an order of optimisations can be found by hand that yields the best results.
-% \MP{what's the point of the previous 3 sentences?(commented)}
-Consider the following two LLVM optimisation passes: \gls*{cf}, which rewrites constant expressions to the constant they evaluate to, and \gls*{dce}, which removes dead code by rewriting conditional branches.
-The order in which \gls*{cf} and \gls*{dce} are performed influences the final result of the compilation (see \Cref{fig:cfdceex}).
-\begin{figure}[!ht]
-  \begin{center}
-    \begin{tikzpicture}[scale=0.62, every node/.style={transform shape}]
-      \node[minimum width=3cm,minimum height=2.5cm,draw=black,very thick,align=left] (progunopt) {\begin{lstlisting}[language=ml,basicstyle=\small\ttfamily]
-let a = true in
-if a then
-  print "a"
-else
-  print "b"
-\end{lstlisting}};
-      \node[minimum width=3cm,minimum height=2.5cm,draw=black,very thick,align=left,right=1.0 of progunopt] (progoptcf) {\begin{lstlisting}[language=ml,basicstyle=\large\ttfamily]
-if true then
-  print "a"
-else
-  print "b"
-\end{lstlisting}};
-      \node[minimum width=3cm,minimum height=2.5cm,draw=black,very thick,align=left,right=1.0 of progoptcf] (progoptcfdce) {\begin{lstlisting}[language=ml,basicstyle=\Large\ttfamily]
-print "a"
-\end{lstlisting}};% the following is a hack to get my syntax highlighting to work again \end{lstlisting}
-      % edges
-      \draw[->,very thick] (progunopt) edge[loop left,left] node {\gls*{dce}} (progunopt);
-      \draw[->,very thick] (progunopt) edge[sloped,above] node {\gls*{cf}} (progoptcf);
-      \draw[->,very thick] (progoptcf) edge[sloped,above] node {\gls*{dce}} (progoptcfdce);
-    \end{tikzpicture}
-  \end{center}
-  \caption{Example program where the level of optimisations differ for one pass of applying \gls*{cf} and \gls*{dce} in any order. %
-  Every edge is a compilation pass and the label on the edge states what the pass does, i.e., \gls*{cf} or \gls*{dce}. %
-  The source code in the nodes is a glorified compiler intermediate representation and the code gets more optimised towards the right hand side of the figure.}\label{fig:cfdceex}
-\end{figure}
-This {\em phase ordering problem} is well--known in literature and a practical solution is to simply perform a fixpoint iteration of the optimisation pipeline~\cite{click1995combining}.
-
-\smallskip
-
-To analyse the security of compilation passes and their interaction within a compilation pipeline, we rely on a few key notions: the definition of a trace relation, and the definition of when is a trace relation well-formed with respect to a class.
-%
-Consider traces $\src{\trace}$ and $\trg{\trace}$ as well as two trace relations $\sim_1$ and $\sim_2$. 
-% 
-The traces are related $\src{\trace}\sim_1\bullet\sim_2\trg{\trace}$ if there is another trace $\irl{\trace}$ such that $\src{\trace}\sim_1\irl{\trace}$ and $\irl{\trace}\sim_2\trg{\trace}$.
-% 
-\bul{A relation $\sim$ is well-formed w.r.t. a class of target-level properties $\trg{\class}$} iff \rul{the universal image preserves set membership}.
-\begin{definition}[Well-formedness of $\sim$ for a Class $\trg{\class}$]\label{def:wfc:sig:tracerel}
-  % $$
-
-  \begin{nscenter}
-  \noindent
-  \text{\bul{$\wfcsig{\sim}{\trg{\class}}$}} \isdef \text{\rul{$\forall \trg{\pi}\in\trg{\class}, \sigma_\sim(\trg{\pi})\in\sigma_\sim(\trg{\class})$}}
-  \end{nscenter}
-  % $$
-\end{definition}
-
-We can now state our main result: secure compilers in the robust compilation framework~\cite{abate2021extacc} compose {\em sequentially}. 
-%
-
-Let \bul{$\cc{\src{L}}{\trg{L}}$ robustly preserve the class $\sigma_{\sim_2}\left(\trg{\class[_1]}\right)$ under $\sim_1$} and let \rul{$\cc{\trg{L}}{\obj{L}}$ robustly preserve the class $\trg{\class[_2]}$ under $\sim_2$}.
-Then, \iul{when the cross-language relation $\sim_2$ is well-formed w.r.t.
-class $\trg{\class[_1]}$}, it follows that \oul{the composed compiler $\cc{\src{L}}{\trg{L}}\circ\cc{\trg{L}}{\obj{L}}$ robustly preserves the intersection of classes $\trg{\class[_1]}\cap\trg{\class[_2]}$ under $\sim_1\bullet\sim_2$}.
-% 
-%For space reasons, \Thmref{thm:urtp} and \Thmref{thm:lrtp} are not presented here, but can be found in the technical appendix ($\CoqSymbol$).
-% Note that it is not necessary to require $\sim_2$ to be well-formed w.r.t. $\src{\class[_1]}$, since the 
-% \MPin{why}
-\begin{theorem}[Composition of Secure Compilers w.r.t. $\sigma$]\label{thm:rtpsim:sig}
-  $\;$
-
-  If \bul{$\rtpsig{\cc{\src{L}}{\trg{L}}}{\tilde{\sigma}_{\sim_2}\left(\trg{\class[_{1}]}\right)}{\sim_1}$}, \rul{$\rtpsig{\cc{\trg{L}}{\obj{L}}}{\trg{\class[_2]}}{\sim_2}$}, and \iul{$\wfcsig{\sim_2}{\trg{\class[_1]}}$}, \\ then \oul{$\rtpsig{\cc{\src{L}}{\trg{L}}\circ\cc{\trg{L}}{\obj{L}}}{\trg{\class[_{1}]}\cap\trg{\class[_{2}]}}{\sim_1\bullet\sim_2}$}.
-\Coqed
-\end{theorem}
-
-Since the composition of secure compilers is again a secure compiler, the theorems generalise to a whole chain of $n$ secure compilers. 
-% 
-\Cref{thm:rtpsim:sig} can also be stated for the existential image $\tau_\sim\left(\src{\pi}\right)$, but in the interest of space that result has been moved to the appendix (\Cref{subsec:extimg}).
-% 
-Crucially, \Cref{thm:rtpsim:sig} also holds for \emph{classes of hyperproperties}, and thus, compilers that robustly preserve hyperproperties can be composed with each other as well as with compilers that robustly preserve properties.
-% 
-%Unfortunately, proofs of compilers that robustly preserve hyperproperties (including hypersafety) are scarce in the literature due to their complexity.
-%\MP{ so what ? }
-%  we don't discuss this further?
-%
-
-% It is furthermore possible to conflate the notion of classes and hyperproperties by intersecting all elements of a class of hyperproperties, to obtain one combined hyperproperty.
-% This idea has been discussed in other work~\cite{clarkson2008hyper}.
-%  MP: i think i stated the point of these 2 lines above
-
-If we take SecurePtrs and FaCT from \Cref{subsec:bg:rtp} and compose them according to \Cref{thm:rtpsim:sig}, we obtain a compiler that robustly preserves the intersection of safety properties and the \gls*{cct} hyperproperty.
-That is, for a source component that robustly satisfies any set of safety properties {\em and} \gls*{cct}, the compiled target component also robustly satisfies the same set of safety properties and \gls*{cct}.
-% The intersection simply uses the lifted version of the respective safety properties.
-%\MP{meaning?}
-
-Compiler engineers typically try to find an order of optimisations that yields well-optimised programs for either code size~\cite{cooper1999geneticphases} or performance~\cite{kulkarni2006exhaustivephase}.
-\Cref{corr:swappable} justifies that any such order of compilation passes is valid w.r.t. security, as long as the trace-relations have no effect on the respective classes.
-
-So, given two compilation passes \bul{$\cc[_{1}]{\trg{L}}{\trg{L}}$, $\cc[_{2}]{\trg{L}}{\trg{L}}$, both robustly preserving class $\trg{\class[_{1}]}$ or $\trg{\class[_{2}]}$, respectively}, \rul{their corresponding well-formed trace-relations}, \iul{and indifference of the classes with respect to these trace relations}, \oul{for any order of their composition, the composed compiler robustly preserves the intersection of $\trg{\class[_{1}]}$ and $\trg{\class[_{2}]}$}.
-% 
-% \MPin{
-% 	later on, say that optimisation passes remove elements from traces, so they should be fine w.r.t. the result with $\tau$.
-% 	also, yes, if $\tau$ adds we have a problem, if it removes we don't
-% \MKin{ Optimisations that remove elements from traces are not necessarily sound,
-%  e.g., remove all deallocations to "optimise" memory management by offloading
-%  the work to the OS.
-% }
-% }
-
-\begin{corollary}[Swapping Secure Compiler Passes]\label{corr:swappable}
-  $\;$
-
-  If \bul{$\rtpsig{\cc[_{1}]{\trg{L}}{\trg{L}}}{\trg{\class[_{1}]}}{\sim_1}$ and $\rtpsig{\cc[_{2}]{\trg{L}}{\trg{L}}}{\trg{\class[_{2}]}}{\sim_2}$}, %
-  \rul{$\wfcsig{\sim_1}{\trg{\class[_2]}}$ and $\wfcsig{\sim_2}{\trg{\class[_1]}}$}, %
-  and \iul{$\tilde{\sigma}_{\sim_2}\left(\trg{\class[_1]}\right)=\trg{\class[_1]}$ as well as $\tilde{\sigma}_{\sim_1}\left(\trg{\class[_2]}\right)=\trg{\class[_2]}$},
-  then \oul{$\rtpsig{\cc[_{1}]{\trg{L}}{\trg{L}}\circ\cc[_{2}]{\trg{L}}{\trg{L}}}{\trg{\class[_{1}]}\cap\trg{\class[_{2}]}}{\sim_1\circ\sim_2}$ and $\rtpsig{\cc[_{2}]{\trg{L}}{\trg{L}}\circ\cc[_{1}]{\trg{L}}{\trg{L}}}{\trg{\class[_{2}]}\cap\trg{\class[_{1}]}}{\sim_2\circ\sim_1}$}.
-\Coqed
-\end{corollary}
-
-Coming back to the example composing SecurePtrs with FaCT, it is likely the case that \Cref{corr:swappable} is not applicable.
-While first running SecurePtrs and then FaCT should be fine, the other direction has potential security hazards, since the SecurePtrs compiler does not account for cryptographic-constant time primitives, such as \texttt{ctselect}.
-
-\subsection{Secure Compiler Composition with Same Trace Models}
-When the cross-language trace relation is an equality, \Cref{thm:rtpsim:sig} collapses:
-Given \bul{$\cc{\src{L}}{\trg{L}}$ robustly preserves $\class[_{1}]$} and \rul{$\cc{\trg{L}}{\obj{L}}$ robustly preserves $\class[_{2}]$}, it follows that \oul{their sequential composition $\cc{\src{L}}{\trg{L}}\circ\cc{\trg{L}}{\obj{L}}$ robustly preserves the intersection of classes $\class[_{1}]$ and $\class[_{2}]$}.
-
-\begin{corollary}[Composition of Secure Compilers]\label{corr:rtp}
-  $\;$
-
-  If \bul{$\rtp{\cc{\src{L}}{\trg{L}}}{\class[_{1}]}$} and \rul{$\rtp{\cc{\trg{L}}{\obj{L}}}{\class[_{2}]}$}, then \oul{$\rtp{\cc{\src{L}}{\trg{L}}\circ\cc{\trg{L}}{\obj{L}}}{\class[_{1}]\cap\class[_{2}]}$}.
-\Coqed
-\end{corollary}
-
-\Cref{corr:rtp} provides an easy way to compose secure compilers without well-formedness of trace relations. 
-% 
-However, while \Cref{thm:rtpsim:sig} explicitly requires that $\sim_1$ is well-formed w.r.t. $\trg{\class[_2]}$, if $\sim_2$ is not well-formed w.r.t. $\trg{\class[_1]}$, care must be taken. 
-%Worst-case is that the resulting class $\sigma_{\sim_1\bullet\sim_2}\left(\trg{\class[_1]}\cap\trg{\class[_2]}\right)$ becomes empty, even though $\trg{\class[_1]}\cap\trg{\class[_2]}$ is not necessarily empty.
-This is further discussed in \Cref{subsec:compatsecpasses}.
-
-%If possible, optimisations are merged to enhance the analysis' result, since even when iterating to a fixpoint the final program may not be as optimised as it could be~\cite{click1995combining}.
-We can also obtain a specialised version of \Cref{corr:swappable}:
-%So, given two compilation passes \bul{$\cc[_{1}]{\trg{L}}{\trg{L}}$, $\cc[_{2}]{\trg{L}}{\trg{L}}$, both robustly preserving class $\class[_{1}]$ or $\class[_{2}]$, respectively}, \oul{for any order of their composition, the composed compiler robustly preserves the intersection of $\class[_{1}]$ and $\class[_{2}]$}.
-% 
-% \MPin{
-% 	later on, say that optimisation passes remove elements from traces, so they should be fine w.r.t. the result with $\tau$.
-% 	also, yes, if $\tau$ adds we have a problem, if it removes we don't
-% \MKin{ Optimisations that remove elements from traces are not necessarily sound,
-%  e.g., remove all deallocations to "optimise" memory management by offloading
-%  the work to the OS.
-% }
-% }
-
-\begin{corollary}[Swapping Secure Compiler Passes]\label{corr:swappable:one}
-  $\;$
-
-  If {$\rtp{\cc[_{1}]{\trg{L}}{\trg{L}}}{\class[_{1}]}$ and $\rtp{\cc[_{2}]{\trg{L}}{\trg{L}}}{\class[_{2}]}$}, then {$\rtp{\cc[_{1}]{\trg{L}}{\trg{L}}\circ\cc[_{2}]{\trg{L}}{\trg{L}}}{\class[_{1}]\cap\class[_{2}]}$ and $\rtp{\cc[_{2}]{\trg{L}}{\trg{L}}\circ\cc[_{1}]{\trg{L}}{\trg{L}}}{\class[_{2}]\cap\class[_{1}]}$}.
-\Coqed
-\end{corollary}
-
-\section{Security Properties Formalisation \& Composition\pages{1.5}}\label{sec:compprop}
-
-This section introduces trace properties of interest for this paper: \gls*{tms}, \gls*{sms}, \gls*{ms}, \gls*{scct}, and \gls*{ss}.
-These properties are of practical importance (as mentioned in \Cref{sec:introduction}) and also of interest in the case study presented later (\Cref{sec:casestud:defs,sec:casestud:rtp}). 
-This section presents all of them, despite the fact that they are inspired by existing work, in order to showcase all that is required for a formal proof of security for a realistic compilation toolchain.
-%\MP{ say something about this in the rebuttal: only the defs are borrowed, the universal images and all that work is new}
-%\MK{ universal images are not new. }
-% , none of the properties is novel, for self-containment of the paper, their definition is necessary for the case study.
-For each of the presented properties, the technical report defines corresponding monitors, which refine a given property and have a key role in the proofs of this paper. 
-% managed to prove 
-%  if tms-monitor --[trace]-->* ∅  
-%  then trace ∈ tms
-% but not
-%  if trace ∈ tms
-%  then tms-monitor --[trace]-->* ∅
-
-\subsection{A Trace Model for Memory Safety}\label{subsec:basic:memsafety:tracemodel}
-
-For simple memory safety composed of temporal and spatial memory safety, the trace model defines events ($\event[_{\mssafe}]$) as either the empty event ($\emptyevent$), a crash ($\lightning$), or a base-event ($\preevent[_{\mssafe}]$).
-
-\vspace{-1.0em}
-\[
-  \begin{array}{rcll}
-    \text{(Base-Events)}&\preevent[_{\mssafe}] &:=& \ev{Alloc \loc\ n} \mid \ev{Dealloc \loc} \mid \ev{Use \loc\ n} \\
-    \text{(Events)}&\event[_{\mssafe}] &:=& \preevent[_{\mssafe}] \mid \emptyevent \mid \lightning \\ 
-  \end{array}
-\]
-
-Base-events describe the actual kind of event that happened.
-For the basic memory-safety properties, these are three variants:
-First, the allocation event ($\ev{Alloc\ \loc\ n}$) that fires whenever a program claims $n$ cells of memory and stores them at address $\loc$, where addresses are assumed to be unique.
-Second, deallocation ($\ev{Dealloc\ \loc}$) announces that the object at location $\loc$ is freed.
-Third, an event to describe reads from and writes to the $n$-th memory cell from address $\loc$ ($\ev{Use\ \loc\ n}$).
-
-\subsubsection{Temporal Memory Safety}
-
-\gls*{tms}~\cite{nagarakatte2010cets} is a safety property that describes that an unallocated object must not be (re-)used.
-
-\begin{definition}[\glsfirst*{tms}]\label{def:trace:tmsdef}
-  % \footnotesize
-  $$
-  \tmssafe:=\left\{\trace[_{\mssafe}] \left| \begin{array}{rcl}
-    \ev{Alloc\ \loc\ n}&\le_{\trace[_{\mssafe}]}&\ev{Dealloc\ \loc} \\
-    \ev{Use\ \loc\ n}&\le_{\trace[_{\mssafe}]}&\ev{Dealloc\ \loc} \\
-    \text{at most one }\ev{Dealloc\ \loc}&\text{in}&\trace[_{\mssafe}] \\
-    \text{at most one }\ev{Alloc\ \loc\ n}&\text{in}&\trace[_{\mssafe}] \\
-  \end{array}\right.\right\}
-  $$
-\end{definition}
-Hereby, the notation $\event[_{1}]\le_{\trace}\event[_{2}]$ means that if $\event[_{1}]$ is in $\trace$ and if $\event[_{2}]$ is in $\trace$, then $\event[_{1}]$ appears before $\event[_{2}]$.
-
-\subsubsection{Spatial Memory Safety}
-
-\gls*{sms}~\cite{nagarakatte2009soft} is a safety property that prohibits out-of-bounds accesses.
-
-\begin{definition}[\glsfirst*{sms}]\label{def:trace:smsdef}
-% \small
-  % \begin{nscenter}
-
-  \noindent
-  \[
-  \smssafe:=\left\{\trace[_{\mssafe}] \left|\begin{array}{rcl}
-      \text{If }\ev{Alloc\ \loc\ n}\le_{\trace_{\mssafe}}\ev{Use\ \loc\ m}, \text{ then }m<n
-  \end{array}\right.\right\}
-  \]
-  % \end{nscenter}
-\end{definition}
-
-\subsubsection{Memory Safety}
-
-In spirit of earlier work~\cite{nagarakatte2009soft,nagarakatte2010cets,jim2002cyclone,necula2005ccured,michael2023mswasm}, full \gls*{ms} is the intersection of \Cref{def:trace:tmsdef,def:trace:smsdef}.
-
-\begin{definition}[\glsfirst*{ms}]\label{def:trace:msdef}
-  $
-  \mssafe:=\tmssafe \cap \smssafe
-  $
-\end{definition}
-
-Note that \Cref{def:trace:msdef} ignores data isolation, so there may still be memory-safety issues introduced by side-channels.
-
-\subsection{A Trace Model for Memory Safety with Constant Time}\label{subsec:scct:tracemodel}
-
-To express Constant Time, we extend the memory safety trace model with a {\em security tag} ($\securitytag{}$) that indicates whether events contains sensitive information ($\lock$) or not ($\unlock$).
-
-\vspace{-.5em}
-\[
-  \begin{array}{rrcl}
-    (\text{Base-Events}) & \preevent[_{\ctsafe}] &:=& \preevent[_{\mssafe}] \mid \ev{Branch\ }n \mid \ev{Binop\ }n\\
-    (\text{Security Tags}) & \securitytag{} &:=& \lock \mid \unlock\\ 
-    (\text{Events}) & \event[_{\ctsafe}] &:=& \preevent[_{\ctsafe}];\securitytag{} \mid \emptyevent \mid \lightning \\ 
-  \end{array}
-\]
-
-For cryptographic code, there is a general guideline that secrets must not be visible on a trace~\cite{ctguidelines}, i.e., secrets should not be marked as $\unlock$.
-In turn, an instruction whose timing is data-dependent must not have a secret as an operand.
-Typical operations with data-dependent timing are branches and certain binary operations, such as division.%
-\footnote{
-	This is architecture-dependent, but division is an operation that serves as a classic example for a data-dependent timing instruction~\cite[p.~755]{arm-refman}.
-}
-Both operations are represented in the trace model by extending the set of base-events with branches ($\ev{Branch\ n}$) and binary operations ($\ev{Binop\ n}$).
-
-\subsubsection{Strict Cryptographic Constant Time}
-
-\gls*{cct} is a hypersafety property~\cite{barthe2018sec} and, thus, difficult to check with monitors.
-This is because, intuitively, hypersafety properties can relate multiple execution traces with each other, but monitors work on a single execution.
-It is a common trick to sidestep this issue by means of overapproximation: this section defines the property \gls*{scct}, a stricter variant of \gls*{cct} (inspired by earlier work~\cite{almeida2017jasmin}) that enforces the policy that no secret appears on a trace.
-Programs that satisfy \gls*{scct} also satisfy \gls*{cct}, but programs that satisfy \gls*{cct} may not satisfy \gls*{scct}.
-
-\begin{definition}[\glsfirst*{scct}]\label{def:trace:scctdef}
-% \begin{nscenter}
-
-  \noindent\[
-  \scctsafe:=\left\{\trace[_{\ctsafe}] \left|\begin{array}{l}
-      \trace[_{\ctsafe}]=\hole{\cdot} \text{ or } \\\exists\trace[_{\ctsafe}'],\trace[_{\ctsafe}]=\preevent[_{\ctsafe}];\unlock\cdot\trace[_{\ctsafe}'] \wedge \trace[_{\ctsafe}']\in\scctsafe
-    \end{array}\right.\right\}
-  \]
-% \end{nscenter}
-\end{definition}
-
-\gls*{scct} may appear overly strict, since it seems that secrets must not occur on a trace (since $\securitytag{}$ is forced to be $\unlock$). 
-However, this is considered standard practice in terms of coding guidelines~\cite{ctguidelines}.
-Moreover, programs that have been compiled with FaCT~\cite{cauligi2019fact} and run with a ``data independent timing mode''~\cite{arm-refman,intel-refman} enabled do not leak secrets (see \Cref{ex:lscct}).
-
-\subsubsection{\gls*{ms}, Strict Cryptographic Constant Time}\label{sec:msscct-rel}
-
-The combination of \gls*{ms} and \gls*{scct} is the intersection of these properties, \gls*{msscct}.
-However, \gls*{ms} uses a different trace model than \gls*{scct}, so intersecting them would trivially yield the empty set. 
-To remedy this issue, we introduce $\sim_{\ctsafe}: \preevent[_{\ctsafe}] \times \preevent[_{\mssafe}] $, a cross-language trace relation (whose key cases are presented below), that we use to intuitively unify the trace model in which the two properties are expressed:
-\[
-  \typerulenolabel{mscctrel:drop:tag}{}{\preevent[_{\mssafe}]\sim_{\ctsafe}\preevent[_{\mssafe}];\unlock{}}
-  \typerulenolabel{mscctrel:drop:crash}{}{\lightning\sim_{\ctsafe}\lightning}
-\]
-\[
-  \typerulenolabel{mscctrel:drop:branch}{}{\emptyevent\sim_{\ctsafe}\ev{Branch n};\securitytag{}}
-  \typerulenolabel{mscctrel:drop:binop}{}{\emptyevent\sim_{\ctsafe}\ev{Binop n};\securitytag{}}
-\]
-
-Essentially, $\sim_{\ctsafe}$ ignores both the new $\ev{Branch\ n}$ and $\ev{Binop\ n}$ base-events as it relates security-insensitive actions ($\unlock$) to their equivalent counterparts.
-Thus, \gls*{ms} traces trivially satisfy \gls*{scct}.
-% 
-The above relation is extended point-wise to traces, skipping the empty event $\emptyevent$ on either side, and it is now possible to define \gls*{msscct} using the universal image:
-
-\begin{definition}[\gls*{ms} and \gls*{scct}]\label{def:trace:msscctdef}
-  $
-  \msscctsafe:=\mssafe\cap\sigma_{\sim_{\ctsafe}}\left(\scctsafe\right)
-  $
-\end{definition}
-
-\subsection{Extending the Trace Model with Speculation}\label{subsec:msctss:tracemodel}
-
-So far, the considered trace models do not let us express speculative execution attacks such as Spectre~\cite{kocher2019spectre}. 
-For this, we extend the earlier trace model (see \Cref{subsec:scct:tracemodel}) so that the security tags ($\securitytag{}$) carry additional information about the kind of private data leakage, i.e., the type of speculative leak.
-Moreover, we add base-events signalling the beginning of a speculative execution ($\ev{Spec}$), a barrier ($\ev{Barrier}$) that signals that any speculative execution may not go past it, as well as a rollback event ($\ev{Rlb}$), which signals that execution resumes to where speculation started.
-
-\vspace{-1em}
-{
-\[
-  \begin{array}{rrcl}
-    (\text{Base-Events}) & \preevent[_{\mathghost}] &:=& \preevent[_{\ctsafe}] \\
-    (\text{Spectre Variants}) & vX &:=& \operatorname{NONE} \mid \operatorname{PHT} \\
-    (\text{Security Tags}) & \securitytag{} &:=& \lock_{vX} \mid \unlock\\ 
-    (\text{Events}) & \event[_{\mathghost}] &:=& \preevent[_{\mathghost}];\securitytag{} \mid \emptyevent \mid \lightning \mid \ev{Spec} \mid \ev{Rlb} \mid \ev{Barrier}\\ 
-  \end{array}
-\]
-}
-
-Even though the considered Spectre variants are just SPECTRE-PHT~\cite{kocher2019spectre}, NONE just describes secret data as in \gls*{scct} (see \Cref{subsec:scct:tracemodel}), the trace model is general enough to allow for potential future extension with different variants~\cite{kocher2019spectre,maisuradze2018ret2spec,horn2019zero}.
-%For sake of readability, this paper just uses the notation $\lock$ in place of $\lock_{\text{PHT}}$.
-
-\subsubsection{Speculative Safety}
-
-\gls*{ss}~\cite{patrignani2021exorcising}, similar to \gls*{scct}, is a sound overapproximation of a variant of noninterference.
-
-\begin{definition}[\glsfirst*{ss}]\label{def:trace:ss}
-  \noindent
-
-  \begin{nscenter}
-  $
-    \sssafe:=\left\{\trace[_{\mathghost}] \left|\begin{array}{l}
-      \trace[_{\mathghost}]=\hole{\cdot} \text{ or } \exists\trace[_{\mathghost}'].\\
-      \left(\trace[_{\mathghost}]=\preevent[_{\mathghost}];\unlock\cdot\trace[_{\mathghost}'] \text{or }\trace[_{\mathghost}]=\preevent[_{\mathghost}];\lock_{\text{NONE}}\cdot\trace[_{\mathghost}']\right)\\
-      \text{and }\ \trace[_{\mathghost}']\in\sssafe
-                                 \end{array}\right.\right\}
-  $ 
-  \end{nscenter}
-\end{definition}
-The technical setup so far leads to the above definition, where only locks annotated with $\text{SPECTRE-PHT}$ are disallowed to occur on the trace.
-That way, programs attaining \gls*{ss} do not necessarily attain \gls*{scct}.
-
-\subsubsection{Speculation Memory Safety}\label{sec:spec-ms-rel}
-
-As before, we need to relate the different trace models with each other, so that the memory safety property without speculation can be lifted to speculation. 
-To this end, let $\sim_{\mathghost}: \preevent[_{\mathghost}] \times \preevent[_{\ctsafe}]$ be a cross-language trace relation whose key cases are below.
-The intuition is that \gls*{ss} is trivially satisfied in \gls*{scct}, since speculation is inexpressible there, which amounts to dropping events $\ev{Spec}$, $\ev{Rlb}$, or $\ev{Barrier}$, as well as all base events tagged with $\lock_{\text{PHT}}$.
-
-\[
-  \typerulenolabel{specrel:drop:tag}{}{\preevent[_{\ctsafe}];\lock\sim_{\mathghost}\preevent[_{\ctsafe}];\lock_{\text{NONE}}}
-  \typerulenolabel{specrel:drop:spectag}{}{\emptyevent\sim_{\mathghost}\preevent[_{\mathghost}];\lock_{\text{PHT}}}
-\]
-\[
-  \typerulenolabel{specrel:drop:pub}{}{\preevent[_{\ctsafe}];\unlock\sim_{\mathghost}\preevent[_{\ctsafe}];\unlock}
-  \typerulenolabel{specrel:drop:crash}{}{\lightning\sim_{\mathghost}\lightning}
-\]
-\[
-  \typerulenolabel{specrel:drop:spec}{}{\emptyevent\sim_{\mathghost}\ev{Spec}}
-  \typerulenolabel{specrel:drop:rlb}{}{\emptyevent\sim_{\mathghost}\ev{Rlb}}
-  \typerulenolabel{specrel:drop:barrier}{}{\emptyevent\sim_{\mathghost}\ev{Barrier}}
-\]
-
-We conclude by defining the ultimate property of interest for secure compilers: \gls*{specms}.
-\begin{definition}[\gls*{specms}]\label{def:trace:specmsdef}
-  $
-  \specmssafe := \msscctsafe\cap\sigma_{\sim_{\mathghost}}\left(\sssafe\right)
-  $
-\end{definition}
-
-\section{Case Study: Language Formalisations\pages{2}}\label{sec:casestud:defs}
-
-This section defines programming languages $\Ltms$, $\Ltrg$, $\Lms$, $\Lscct$, and $\Lspec$, all of which share common elements (presented in \Cref{subsec:cs:defs}).
-$\Ltms$ is the only statically-typed language, and it exhibits the property that all well-typed programs are \gls*{tms} (\Cref{subsec:ltms}).
-However, not all $\Ltms$ programs are \gls*{sms}.
-That is, there are well-typed $\Ltms$ programs that perform an out-of-bounds access.
-Language $\Ltrg$ is untyped and does not provide any guarantees with regards to \gls*{ms} (\Cref{subsec:lsms}).
-$\Lms$ is exactly the same language as $\Ltrg$, but this paper still distinguishes the two for sake of readability (\Cref{subsec:lms}).
-All three languages --- so $\Ltms$, $\Ltrg$, and $\Lms$ --- assume \gls*{scct} to hold, since this is -- in an ideal world -- what the programmer would expect, too: it is the job of the compiler to preserve and (potentially) enforce \gls*{scct} security~\cite{cauligi2019fact,nagarakatte2010cets,nagarakatte2009soft,akritidis2009baggy}.
-% This paper considers \gls*{cct} as an architecture-dependent property\MPin{can we justify this}, so
-
-Such consideration is also backed up by architecture providing a data (operand) independent timing mode, such as processors by Arm~\cite[p.~543]{arm-refman} and Intel~\cite[p.~80]{intel-refman}.
-%In spirit of this, all languages besides $\Lscct$ are trivially \gls*{scct}. 
-This kind of processor feature is modelled in language $\Lscct$ (\Cref{subsec:lscct}), where programs have access to a ``\gls*{cct}-mode'' and can change the leakage of emitted events according to the value of this mode (either $\obj{ON}$ or $\obj{OFF}$).
-
-Finally, modern processors also employ speculative execution to achieve speedups---and unfortunately generate Spectre attacks~\cite{kocher2019spectre}---and this is the extension of $\Lspec$ (\Cref{subsec:lspec}).
- % extends $\Lscct$ to model speculative execution.
-%
-Thus, all previous languages trivially satisfy \gls*{ss}, since they do not support speculative execution at all.
-% All languages share large parts of syntactic elements, which are presented in the next subsection (\Cref{subsec:cs:defs}).
-% Subsequent subsections extend on this accordingly.
-
-\subsection{Shared Language Definitions}\label{subsec:cs:defs}
-
-All presented programming languages share a common fragment which is partially presented here and in full detail in the technical report. 
-%\MP{put omitted stuff in appendix + ref there (and add textual explanation)}
-
-% \vspace{-.5em}
-{
-  \renewcommand{\src}[1]{\mathsf{#1}}
- \small
-\[
-  \begin{array}{rrcl}
-    \text{(Base-Events)} & \src{\preevent} &:=& \src{\ev{Alloc \loc\ n}} \mid \src{\ev{Dealloc \loc}} \mid \src{\ev{Get \loc\ n}} \mid \src{\ev{Set \loc\ n}} \\
-    \text{(Control Tags)} & \src{\sandboxtag{}} &:=& \src{\ctx} \mid \src{\comp} \\
-    \text{(Events)} & \src{\event} &:=& \src{\preevent;\sandboxtag{}} \mid \src{\emptyevent} \mid \src{\lightning} \\ 
-    \text{(Values)} & \src{v} &:=& \src{n} \\
-    \text{(Expressions)} & \src{\expr} &:=& \src{x} \mid \src{n} \mid \src{\lbinop{\expr[_1]}{\expr[_2]}} \mid \src{\lifz{\expr[_1]}{\expr[_2]}{\expr[_3]}} \\ 
-                         &&&\mid \src{\llet{x}{\expr[_1]}{\expr[_2]}} \mid \src{\llet{x}{new\ \expr[_1]}{\expr[_2]}} \\
-                         &&&\mid \src{\ldelete{x}} \mid \src{\lget{x}{\expr}} \mid \src{\lset{x}{\expr[_1]}{\expr[]}} \mid \src{stuck}\\
-                         &&&\mid \src{\lcall{f}{\expr}} \mid \src{\lreturn{\expr}} \\
-    \text{(Functions)} & \src{F} &:=& (\src{x};\src{\expr}) \\
-    \text{(Libraries)} &&&\hspace{-3.8em} \src{\library_{\ctx}},\src{\library_{\comp}} : \src{Vars} \to \src{F} 
-    % \hspace{2em}
-    %  : \src{Vars} \to \src{F} 
-     \\
-    \text{(Heaps)} & \src{H} &:=& \src{\hole{\cdot}} \mid \src{n},\src{H}
-    % \hspace{0.5cm}    
-    \\
-    \text{(Pointer Maps)} & \src{\Delta} && \text{ omitted for simplicity}
-    \\
-    \text{(Memory States)} & \src{\memstate} &:=& \left(\src{H^{\ctx}};\src{H^{\comp}};\src{\Delta}\right)\\
-    \text{(Control States)} & \src{\cfstate} & & \text{ omitted for simplicity} 
-    \\
-    \text{(States)} &\src{\Omega} &:=& \left(\src{\cfstate};\src{\sandboxtag{}};\src{\memstate}\right)\\
-  \end{array}
-\]
-}
-{
-  \renewcommand{\src}[1]{\mathsf{#1}}
-The trace models of all languages are similar to those presented earlier (\Cref{sec:compprop}).
-One technical detail is the addition of a control tag, indicating who is to blame for an emitted action: context ($\ctx$) or component ($\comp$).
-This is a standard technique in secure compilation in order to rule out irrelevant context-level events. 
-For example, a context immediately deallocating an allocated memory region twice trivially violates memory safety, but the main interest in secure compilation is to preserve component-level security, and thus component events.
-Even though this tagging could be used for blame preservation~\cite{patrignani2023blame}, this is beyond the focus of this paper.
-Another key difference is that memory accesses are now explicitly modelled as reads ($\src{\ev{Read\ \loc\ n}}$) and writes ($\src{\ev{Write\ \loc\ n}}$) instead of just uses.
-
-All languages have at least numbers as values ($\src{v}$) and second class functions ($\src{F}$).
-Functions are modelled as pairs containing the name of one argument and the body of the function.
-Bodies are just ordinary expressions, which can be simple binary operations ($\src{\lbinop{\expr[_1]}{\expr[_2]}}$), conditionals ($\src{\lifz{\expr[_1]}{\expr[_2]}{\expr[_3]}}$), function calls ($\src{\lcall{f}{e}}$) and returns ($\src{\lreturn{\expr}}$), as well as C-like memory operations.
-Programs have sets of pre-determined functions called libraries and they are marked as being part of some component ($\src{\library_{\comp}}$) or context ($\src{\library_{\ctx}}$).
-
-For the operational semantics, the runtime state ($\src{\Omega}$) is a triple consisting of a control-flow state ($\src{\cfstate}$), a control tag ($\src{\sandboxtag{}}$), and a memory state ($\src{\memstate}$).
-The latter carries information about pointers that are kept ``alive'' in pointer maps ($\src{\Delta}$), so that the semantics does not get stuck when encountering, e.g., a double-free.
-The memory state also carries two heaps to model sandboxing between a context ($\src{H^{\ctx}}$) and a component ($\src{H^{\comp}}$).
-
-}
-
-\subsection{$\src{L_{\tmssafe}}$: A Temporal but Not Spatial Memory Safe Language}\label{subsec:ltms}
-
-$\src{L_{\tmssafe}}$ is the only statically-typed language in this case study and restricts functions ($\src{F}$) to the typing signature $\src{\natt\to\natt}$. 
-The type system of $\src{L_{\tmssafe}}$ is inspired by $L^{3}$~\cite{morrisett2005L3,scherer2018fabulous} and enforces that every well-typed $\src{L_{\tmssafe}}$ program satisfies \gls*{tms}.
- % (\Cref{thm:wt:tms}).
-
-\begin{theorem}[$\src{L_{\tmssafe}}$-programs are \gls*{tms}]\label{thm:wt:tms}
-$\rsat{\src{\library_{\comp}}}{\tmssafe}$ \Coqed
-  % $\;$ 
-% 
-  % \begin{nscenter}
-  %   \hfill For any component $\src{\library_{\comp}}$, $\rsat{\src{\library_{\comp}}}{\tmssafe}$ \Coqed
-  % \end{nscenter}
-\end{theorem}
-
-\subsection{$\Ltrg$: A Memory-Unsafe Language}\label{subsec:lsms}
-
-$\Ltrg$ extends the syntax presented earlier (\Cref{subsec:cs:defs}) with dynamic typechecks ($\trg{\lhast{\expr}{\type}}$), evaluating to $\trg{0}$ if the type matches with the shape of $\trg{\expr}$ and $\trg{1}$ otherwise.
-%The pointers carry poison tags, which signify a pointer being allocated ($\trg{\poisoned}$) or freed ($\trg{\poisonless}$). 
-Furthermore, the syntax of $\Ltrg$ is extended with a way to inspect whether a pointer is freed ($\trg{\lispoisoned{x}}$), evaluating to $\trg{0}$ if it is freed and to $\trg{1}$ otherwise.
-Functions may receive arguments that are not $\trg{\natt}$, values are extended with tuples, and expressions are also extended with pair projections.
-
-% \vspace{-1em}
-\[
-  \begin{array}{rrcl}
-    \text{(Values)} & \trg{v} &:=& \dots \mid \trg{\lpair{v_1}{v_2}} \\
-    \text{(Expressions)} & \trg{\expr} &:=& \dots \mid \trg{\lhast{\expr}{\tau}} \mid \trg{\lispoisoned{x}}
-  \end{array}
-\]
-
-No changes are done to the trace model, but $\Ltrg$ has no static typing, making double-free code patterns possible.
-
-\subsection{$\Lms$: Another Memory-Unsafe Language}\label{subsec:lms}
-$\Lms$ is exactly equal to $\Ltrg$ (\Cref{subsec:lsms}), but used to emphasize that this is code after applying $\cc{\Ltrg}{\Lms}$ (\Cref{subsec:cs:ms}).
-
-\subsection{$\Lscct$: A Memory-Unsafe Language with a Constant Time Mode}\label{subsec:lscct}
-
-$\Lscct$ extends $\Lms$ (\Cref{subsec:lms}) with a ,,constant-time mode''. 
-The activation of the mode can be checked ($\obj{\lrddoit{x}{\expr}}$), changed ($\obj{\lwrdoit{D}}$), and is stored in program states ($\obj{\Omega}$).
-At the beginning of program execution of $\Lscct$ programs, the mode is turned off ($\obj{OFF}$).
-If the mode is enabled (i.e., set to $\obj{ON}$), the intuition is that no secrets are leaked. 
-This models the real-world\footnote{As present in Intel~\cite[p.80]{intel-refman} and ARM~\cite[p.~543]{arm-refman} processors.} data-independent timing mode as well as the result of compiling a program with FaCT~\cite{cauligi2019fact}.
-For example, FaCT rewrite code that branches to use a constant-time selection primitive.
-To not obfuscate our formalisation unnecessarily by duplicating syntax, we simply added the ,,constant-time mode'' to $\Lscct$.
-
-% $\Lscct$ extends $\Lms$ (\Cref{subsec:lms}) with a way to read from ($\obj{\lrddoit{x}{\expr}}$) and to write to ($\obj{\lwrdoit{D}}$) a model-specific register that controls a constant-time mode. 
-% This language is a combination of the effect of the FaCT compiler~\cite{cauligi2019fact} as well as a data-independent timing mode, a feature that is present in both Arm~\cite[p.~543]{arm-refman} and Intel~\cite[p.~80]{intel-refman} processors.
-% \MP{ can we be more specific about the fact pass? }
-% To this end, states ($\obj{\Omega}$) are extended with the value of the register ($\obj{D}$), which is initially set to be not active ($\obj{OFF}$).
-% If the register is marked active ($\obj{ON}$), the intuition is that no secrets can appear on specification traces.
-% If the register is marked inactive, secrets may appear on traces.
-% The state of the register can be read and written to, e.g., \Cref{tr:e-wrdoit-off}.
-% For the other languages seen earlier, the mode is intuitively always-on, i.e., the mode of execution always uses constant-time mode.
-
-% \vspace{-.5em}
-\[
-  \begin{array}{rrcl}
-    \text{(Mode Values)} & \obj{D} &:=& \obj{ON} \mid \obj{OFF} \\
-    \text{(Security Tags)} & \obj{\securitytag{}} &:=& \obj{\lock} \mid \obj{\unlock} \\
-    \text{(Base-Events)} & \obj{\preevent} &:=& \dots \mid \obj{iGet\ \loc\ v} \mid \obj{iSet\ \loc\ v} \\
-                       &&& \mid\ \obj{Binop\ n} \mid \obj{Branch\ n} \\
-    \text{(Events)} & \obj{\event} &:=& \obj{\preevent;\sandboxtag{};\securitytag{}} \mid \obj{\emptyevent} \mid \obj{\lightning} \\
-    \text{(Expressions)} & \obj{\expr} &:=& \dots \mid \obj{\lrddoit{x}{\expr}} \mid \obj{\lwrdoit{D}} \\
-                         &&&\mid\ \obj{\llet{x^{\securitytag{}}}{\expr[_1]}{\expr[_2]}}\\
-    \text{(States)} & \obj{\Omega} & := & \left(\obj{\cfstate};\obj{\sandboxtag{}};\obj{D};\obj{\memstate}\right)\\
-  \end{array}
-\]
-\begin{center}
-\newcommand{\expreval}[5]{{#1}\triangleright\xspace {#2}\xrightarrow{#5}\ {#3}\triangleright\xspace {#4}\xspace}
-\newcommand{\exprevalo}[5]{\expreval{\obj{#1}}{\obj{#2}}{\obj{#3}}{\obj{#4}}{\obj{#5}}}
-  \typerule{$e-\obj{wrdoit}-\text{off}$}{
-  }{
-    \pstepto{\rtt{\obj{\cfstate;\sandboxtag{};D;\memstate}}{\obj{\lwrdoit{OFF}}}}{\rtt{\obj{\cfstate;\sandboxtag{};OFF;\memstate}}{\obj{0^{\securitytag{}}}}}{\obj{\emptyevent}}
-  }{e-wrdoit-off}
-
-  \typerule{$e-\obj{get}-\in-$noleak}{
-    \obj{\memstate}=\obj{H^{\ctx};H^{\comp};\Delta_1,x\mapsto(\loc;\sandboxtag{};\poison),\Delta_2} &
-    \obj{\loc}+\obj{n}\in\text{dom }\obj{H^{\sandboxtag{}}} \\
-    \obj{\securitytag{''}}=\obj{\securitytag{}\sqcap\securitytag{'}} &
-    \obj{\event} = \obj{{iGet}\ \loc\ n;\sandboxtag{};\securitytag{''}}
-  }{
-    \exprevalo{
-    	\cfstate;\sandboxtag{'};ON;\memstate
-    }{
-      	x^{\securitytag[n^{\securitytag{'}}] }
-    }{
-    	\cfstate;\sandboxtag{'};ON;\memstate
-    }{
-    	(H^{\sandboxtag{}}(\loc+n))^{\securitytag{''}}
-    }{
-      \event
-    }
-  }{to-e-get-in-noleak}
-\end{center}
-
-The language also adds user annotations $\obj{\securitytag{}}$ for the secrecy of variables, which can be either private (high secrecy) $\obj{\lock}$ or public (low secrecy) $\obj{\unlock}$.
-Security tags $\obj{\securitytag{}}$ are arranged in the usual secrecy lattice~\cite{zdancewicphd}, where $\obj{\unlock}\sqsubseteq\obj{\lock}$.
-% $\obj{\lock}\sqsubseteq\obj{\securitytag{}}$ and $\obj{\securitytag{}}\sqsubseteq\obj{\unlock}$.
-
-Memory accesses to secret data need to be present to reason about memory safety, even when execution is in constant-time mode, e.g, \Cref{tr:to-e-get-in-noleak}.
-$\Lscct$ extends base-events with $\obj{\ev{iGet}\ \loc\ \valueexpr}$ and $\obj{\ev{iSet}\ \loc\ \valueexpr}$ (for data $\obj{i}$ndependent get and set) to prevent secrets from leaking but still enable reasoning about memory safety.
-Due to the technical setup, the rule needs to check if the access is in bounds ($\obj{\loc}+\obj{n}\in\obj{H^t}$) and update the secrecy tag ($\obj{\securitytag{''}}=\obj{\securitytag{}\sqcap\securitytag{'}}$) with the least upper bound of the tags, according to the aforementioned lattice.
-The precise information carried by the event, e.g., location ($\obj{\loc}$) is taken from the pointer map, which carries information irrelevant to this rule ($\obj{\poison}$).
-
-Base-events include $\obj{\ev{Branch}\ n}$ and $\obj{\ev{Binop}\ n}$ that are emitted when evaluating a branch or certain binary expressions, such as division, respectively, whenever the constant-time mode is inactive.
-Events are extended with a security-tag ($\obj{\securitytag{}}$) to signal the secrecy of the involved data.
-
-The evaluation steps are amended to propagate the security-tag annotations $\obj{\securitytag{}}$.
-When the constant-time mode is inactive, base-events $\obj{\ev{Branch}\ n}$ and $\obj{\ev{Binop}\ n}$ are emitted for conditionals and binary operations, respectively.
-Otherwise, just like in the semantics of the earlier languages, $\obj{\emptyevent}$ is emitted for binary and branching operations.
-%To mark this mode as active or inactive, \Cref{tr:e-wrdoit} allows to access the register in $\obj{\statevar}$.
-
-\Cref{ex:lscct} illustrates the differences between $\Lscct$ and other languages.
-
-\begin{exampleenv}[$\Lscct$ with and without constant-time mode]\label{ex:lscct}
-  Consider again \Cref{ex:strncpy}, with a context copying the string \texttt{Hello World}, where everything is marked with a high security tag: $\lock$.
-  The top half of \Cref{fig:ex-cct} (titled $\Lscct$), describes the execution trace of the program, while the bottom side of the table (titled ``Specification''), describes the related specification trace (\Cref{subsec:msctss:tracemodel}).
-  Read in parallel from top to bottom, the figure shows parts of the execution trace. 
-  In each half, the left column (\textit{Active}) has constant-time mode $\obj{ON}$ and the right one (\textit{Inactive}) has it $\obj{OFF}$.
-
-  When the constant-time mode is off, the execution yields events in similar fashion to before (\Cref{subsec:ltms,subsec:lsms,subsec:lms}).
-  But, if it is turned on, then the branching event \colorbox{yellow}{does not fire} anymore and both reading and writing to memory is related to a specification trace with \colorbox{lightgray}{no exposed secrets}.
-\end{exampleenv}
-\begin{figure}[!htb]
-	\vspace{-1em}\small
-  $$
-  % {
-  \begin{array}{ccc}
-    & \Lscct &        \\
-    \text{Active} & \mid & \text{Inactive} \\\hline
-    \obj{\ev{Alloc}\ \loc_{x}\ 12;\comp;\unlock} & \mid & \obj{\ev{Alloc}\ \loc_{x}\ 12;\comp;\unlock} \\
-    \obj{\ev{Alloc}\ \loc_{y}\ 12;\comp;\unlock} & \mid & \obj{\ev{Alloc}\ \loc_{y}\ 12;\comp;\unlock} \\
-    \obj{{\ev{iGet}}\ \loc_{x}\ 0;\comp;}\text{\colorbox{lightgray}{$\obj{\lock}$}} & \mid & \obj{\ev{Get}\ \loc_{x}\ 0;\comp;\lock} \\
-    \text{\colorbox{yellow}{$\obj{\emptyevent}$}} & \mid & \obj{\ev{Branch}\ 0;\comp;\lock} \\
-    \obj{{\ev{iGet}}\ \loc_{x}\ 0;\comp;}\text{\colorbox{lightgray}{$\obj{\lock}$}} & \mid & \obj{\ev{Get}\ \loc_{x}\ 0;\comp;\lock} \\
-    \obj{{\ev{iSet}}\ \loc_{y}\ 0\ \mathtt{'H'};\comp;}\text{\colorbox{lightgray}{$\obj{\lock}$}} & \mid & \obj{\ev{Set}\ \loc_{y}\ 0\ \mathtt{'H'};\comp;\lock} \\
-    \obj{{\ev{iGet}}\ \loc_{x}\ 1;\comp;}\text{\colorbox{lightgray}{$\obj{\lock}$}} & \mid & \obj{\ev{Get}\ \loc_{x}\ 1;\comp;\lock} \\
-    \text{\colorbox{yellow}{$\obj{\emptyevent}$}} & \mid & \obj{\ev{Branch}\ 0;\comp;\lock} \\
-    \vdots & \mid & \vdots \\
-    \obj{{\ev{iGet}}\ \loc_{x}\ 12;\comp;}\text{\colorbox{lightgray}{$\obj{\lock}$}} & \mid & \obj{\ev{Get}\ \loc_{x}\ 12;\comp;\lock} \\
-    \text{\colorbox{yellow}{$\obj{\emptyevent}$}}& \mid & \obj{\ev{Branch}\ 1;\comp;\lock} \\
-    \obj{\ev{Dealloc}\ \loc_{y};\comp;\unlock} & \mid & \obj{\ev{Dealloc}\ \loc_{y};\comp;\lock} \\
-    \obj{{\ev{iGet}}\ \loc_{y}\ 6;\comp;}\text{\colorbox{lightgray}{$\obj{\lock}$}} & \mid & \obj{\ev{Get}\ \loc_{y}\ 6;\comp;\lock} \\
-  \end{array}
-  % }
-  $$
-  $$
-  % {
-  \begin{array}{ccc}
-          & \text{Specification} & \\
-  \text{Active} & \mid & \text{Inactive} \\\hline
-     \ev{Alloc}\ \loc_{x}\ 12;\unlock & \mid & \ev{Alloc}\ \loc_{x}\ 12;\unlock\\
-     \ev{Alloc}\ \loc_{y}\ 12;\unlock & \mid & \ev{Alloc}\ \loc_{x}\ 12;\unlock\\
-     \ev{Use}\ \loc_{x}\ 0;\text{\colorbox{lightgray}{$\unlock$}} & \mid & \ev{Use}\ \loc_{x}\ 0;\lock\\
-     \emptyevent & \mid & \ev{Branch}\ 0;\lock\\
-     \ev{Use}\ \loc_{x}\ 0;\text{\colorbox{lightgray}{$\unlock$}} & \mid & \ev{Use}\ \loc_{x}\ 0;\lock\\
-     \ev{Use}\ \loc_{y}\ 0;\text{\colorbox{lightgray}{$\unlock$}} & \mid & \ev{Use}\ \loc_{y}\ 0;\lock\\
-     \ev{Use}\ \loc_{x}\ 1;\text{\colorbox{lightgray}{$\unlock$}} & \mid & \ev{Use}\ \loc_{x}\ 1;\lock\\
-     \emptyevent & \mid & \ev{Branch}\ 0;\lock\\
-     \vdots & \mid & \vdots \\
-      \ev{Use}\ \loc_{x}\ 12;\text{\colorbox{lightgray}{$\unlock$}} & \mid & \ev{Use}\ \loc_{x}\ 12;\lock\\
-      \emptyevent & \mid & \ev{Branch}\ 1;\lock\\
-      \ev{Dealloc}\ \loc_{y};\unlock & \mid & \ev{Dealloc}\ \loc_{y};\unlock\\
-      \ev{Use}\ \loc_{y}\ 6;\text{\colorbox{lightgray}{$\unlock$}} & \mid & \ev{Use}\ \loc_{y}\ 6;\lock\\
-  \end{array}
-  % }
-  $$
-  \caption{Traces for \Cref{ex:lscct}.}
-  \label{fig:ex-cct}
-\end{figure}
-
-\subsection{$\Lspec$: A Memory-Unsafe Language with Speculation}\label{subsec:lspec}
-
-$\Lspec$ extends $\Lscct$ (\Cref{subsec:lscct}) with speculative dynamic semantics that is inspired by existing work~\cite{guarnieri2018spectector,fabian2022automatic}.
-After branching, speculation starts by pushing the current configuration into the speculation state ($\ird{S}$), and run subsequent code in a predetermined window (whose size is $\omega$) until a rollback of operational state is performed.
-The window is set inside the speculation state.
-Within such windows, data may leak, this is marked as high $\ird{\lock}$ and with an annotation that indicates the respective speculative execution variant. 
-In this paper, we just consider SPECTRE-PHT~\cite{kocher2019spectre}, whose starting point is \Cref{tr:e-ifz-true-spec}.
-% where $\omega$ is the size of the speculation window, i.e., a number of reduction steps to be taken under speculative semantics.
-Additionally, $\ird{\lock}$ may also carry a $\ird{NONE}$ annotation, to signal that this is a leak irrespective of speculation, i.e., as defined earlier (\Cref{subsec:lscct}).
-The semantics is kept general enough to allow for future extension to support different variants~\cite{kocher2019spectre,maisuradze2018ret2spec,horn2019zero}.
-
-In terms of language features, $\Lspec$ includes a new barrier operation $\ird{\lbarrier}$ that blocks speculative execution. 
-To facilitate speculative execution in a non-assembly-like language, a stack of operational state is used and speculation is active if that stack has a size greater than one.
-This is exploited in \Cref{tr:e-spec-eat} to consume any leftover speculation.
-
-\vspace{-.5em}
-\[
-  \begin{array}{rrcl}
-    \text{(Base-Events)} & \ird{\preevent} &:=& \dots \mid \ird{Spec} \mid \ird{Rlb} \mid \ird{Barrier}\\
-    \text{(Expressions)} & \ird{\expr} &:=& \dots \mid \ird{\lbarrier} \\
-    \text{(Speculation State)} & \ird{S} &:=& \ird{\Omega\triangleright\expr} \mid \ird{S},\left(\ird{\Omega};\ird{n};\ird{\expr}\right)
-  \end{array}
-\]
-
-\begin{center}
-\newcommand{\expreval}[5]{{#1}\triangleright\xspace {#2}\xrightarrow{#5}\ {#3}\triangleright\xspace {#4}\xspace}
-\newcommand{\exprevald}[5]{\expreval{\ird{#1}}{\ird{#2}}{\ird{#3}}{\ird{#4}}{\ird{#5}}}
-
-  %
-  \typerule{$e-\ird{ifz}-\ird{true}-\text{spec}$}{
-    \ird{S} = \rtt{(\ird{\cfstate['];\sandboxtag{};D;\memstate});\ird{S}}{\ird{\expr[_1]}},({\ird{\cfstate;\sandboxtag{};D;\memstate}};\omega;{\ird{\expr[_2]}})
-  }{
-    \pstepto{\rtt{\ird{\cfstate;\sandboxtag{};D;\memstate}}{\ird{\lifz{0^{\securitytag{}}}{\expr[_1]}{\expr[_2]}}}}{\ird{S}}{\ird{{Branch}\ 0;\sandboxtag{};\securitytag{}}\cdot\ird{{Spec};\sandboxtag{};\securitytag{}}}
-  }{e-ifz-true-spec}
-  %
-  \typerule{$e-\ird{spec}-\text{eat}$}{
-    \ird{n} > 0 &
-    \pstepto{\rtt{\ird{\Omega}}{\ird{\expr}}}{\rtt{\ird{\Omega'}}{\ird{\expr[']}}}{\ird{\event}}
-  }{
-    \ghoststepto{\ird{S},(\ird{\Omega};\ird{n};\ird{\expr})}{\ird{S},(\ird{\Omega'};\ird{n}-1;\ird{\expr'})}{\ird{\event}}
-  }{e-spec-eat}
-  % 
-  \typerule{$e-\ird{spec}-\text{eaten}$}{
-  }{
-    \ghoststepto{\ird{S},(\ird{\Omega};\ird{0};\ird{\expr})}{\ird{S}}{\ird{Rlb;\Omega.\sandboxtag{};\unlock}}
-  }{e-spec-eaten}
-  %
-  \typerule{$e-\ird{spec}-\text{barrier}$}{
-  }{
-    \ghoststepto{\ird{S},(\ird{\Omega};\ird{n};\ird{\lbarrier})}{\ird{S},(\ird{\Omega};\ird{0};\ird{0})}{\ird{Barrier;\Omega.\sandboxtag{};\unlock}}
-  }{e-spec-barrier}
-\end{center}
-
-The trace model is extended with $\ird{{Spec}}$, $\ird{{Rlb}}$, and $\ird{{Barrier}}$ events that signal the respective operational action. 
-A $\ird{\lbarrier}$ prevents any execution whatsoever besides $\ird{{Rlb}}$ when run in speculative execution mode and does nothing when run in normal mode, e.g., \Cref{tr:e-spec-barrier}.%
-\footnote{In the rule, the notation $\cfstate' = \cfstate[.\text{win} = 0]$ means that $\cfstate'$ is a copy of $\cfstate$ up to field $\text{win}$, which is set to $0$.}
-Lastly, $\ird{Rlb}$ is emitted when the speculation window is zero and the operational state is rolled back (see \Cref{tr:e-spec-eaten}).
-
-\section{Case Study: Composing Secure Compiler Passes and Optimisations \pages{7}}\label{sec:casestud:rtp}
-This section defines several secure compilers, each of which robustly preserves a different property of interest as depicted in \Cref{fig:pipeline}.
-\begin{figure*}[!h]
-  \centering
-  \begin{tikzpicture}
-    \node (S) {$\src{L_{\tmssafe}}$};
-    \node[right=2.0 of S] (T) {$\trg{L}$};
-    \node[right=2.0 of T] (M) {$\irl{L_{\mssafe}}$};
-    \node[below right=1.5 and 1.0 of M] (D0) {$\irl{L_{\mssafe}}$};
-    \node[above right=1.5 and 1.0 of M] (C0) {$\irl{L_{\mssafe}}$};
-    \node[right=3.0 of M] (E) {$\irl{L_{\mssafe}}$};
-    \node[right=2.0 of E] (O) {$\obj{L_{\scctsafe}}$};
-    \node[right=2.0 of O] (Z) {$\ird{L_{\mathghost}}$};
-
-    \draw[->] (S) to[sloped] node[align=center] (tmsedge) {\gls*{tms}\\ \Cref{thm:cca:rtp:tms}} (T);
-    \draw[->] (T) to[sloped] node[align=center] {\gls*{sms}\\ \Cref{thm:ccb:rtp:sms}} (M);
-    \draw[->] (M) to[sloped] node[align=center] {\gls*{dce}\\ \Cref{thm:ccdce:rtp:ms}} (D0);
-    \draw[->] (M) to[sloped] node[align=center] {\gls*{cf}\\ \Cref{thm:cccf:rtp:ms}} (C0);
-    \draw[->] (D0) to[sloped] node[align=center] {\gls*{cf}\\ \Cref{thm:cccf:rtp:ms}} (E);
-    \draw[->] (C0) to[sloped] node[align=center] {\gls*{dce}\\ \Cref{thm:ccdce:rtp:ms}} (E);
-    \draw[->] (E) to[sloped] node[align=center] {\gls*{scct}\\ \Cref{thm:ccscct:rtp:scct}} (O);
-    \draw[->] (O) to[sloped] node[align=center] {\gls*{ss}\\ \Cref{thm:ccspec:rtp:spec}} (Z);
-
-    % Sections
-    \node (sectms) at ($(S)+(1.25,2.5)$) {{\Cref{subsec:cs:tms}}};
-    \node (secsms) at ($(T)+(1.25,2.5)$) {{\Cref{subsec:cs:ms}}};
-    \node (secopts) at ($(M)+(2.0,2.5)$) {{\Cref{subsec:cs:opts}}};
-    \node (secscct) at ($(E)+(1.5,2.5)$) {{\Cref{subsec:cs:scct}}};
-    \node (secss) at ($(O)+(1.5,2.5)$) {{\Cref{subsec:cs:ss}}};
-
-    \draw[thick,dotted,->] ($(S) + (0.2,-0.27)$) to ($(S) - (-0.2,1)$) to node[align=center] {\gls*{ms}\\ \Cref{corr:ccab:rtp:ms}} ($(M) - (0.0,1)$) to (M);
-
-    \draw[thick,dotted,->] (S) to ($(S) - (0.0,2.5)$) to node[align=center,pos=.65] {\gls*{specms}\\ \Cref{thm:ccall:rtp:specms}} ($(Z) - (0.0,2.5)$) to (Z);
-
-    % \draw[thick,dotted,->] (S) to[bend right=33,sloped] node[align=center] {\gls*{ms}\\ \Cref{corr:ccab:rtp:ms}} (M);
-    \draw[thick,dotted,->] (M) to[bend right=0,sloped] node[align=center] {\gls*{ms}\\ \Cref{thm:cccfccdce:rtp:ms}} (E);
-%    \draw[thick,dotted,->] (S) to[out=-45,in=198,sloped] node[align=center] {\gls*{ms}+\gls*{scct}\\ \Cref{thm:ccall:rtp:msscct}} (O);
-    % \draw[thick,dotted,->] (S) to[out=-40,in=210,sloped] node[align=center,pos=.8] {\gls*{specms}\\ \Cref{thm:ccall:rtp:specms}} (Z);
-
-    \draw[dashed] ($(S)-(0.0,-0.5)$) -- ($(S)-(0.0,-2.5)$);
-    \draw[dashed] ($(T)-(0.0,-0.5)$) -- ($(T)-(0.0,-2.5)$);
-    \draw[dashed] ($(M)-(0.0,-0.5)$) -- ($(M)-(0.0,-2.5)$);
-    \draw[dashed] ($(E)-(0.0,-0.5)$) -- ($(E)-(0.0,-2.5)$);
-    \draw[dashed] ($(O)-(0.0,-0.5)$) -- ($(O)-(0.0,-2.5)$);
-    \draw[dashed] ($(Z)-(0.0,-0.5)$) -- ($(Z)-(0.0,-2.5)$);
-  \end{tikzpicture}
-  \vspace{-1em}
-  \caption{Visualisation of the optimising compilation pipeline that preserves \gls*{specms}. %
-    Vertices in the graph are the programming languages from earlier sections (\Cref{sec:casestud:defs}). %
-    Full edges are secure compilers passes.
-    Dotted edges are composition of passes and use the presented framework (\Cref{sec:sequential}) to indicate the property they preserve. %
-    The dashed lines partition the graph into the sections where the respective theorems are presented.
-  }\label{fig:pipeline}
-\end{figure*}
-The section demonstrates the power of the framework (\Cref{sec:sequential}) by composing these compilers for a secure and optimising compilation chain that robustly preserves \gls*{specms}.
-The first step in this chain is the compiler from $\src{L_{\tmssafe}}$ to $\trg{L}$ that robustly preserves just \gls*{tms} (\Cref{thm:cca:rtp:tms}).
-From here, another pass from $\trg{L}$ to $\irl{L_{\mssafe}}$ ensures that no out-of-bounds accesses can happen and, thus, programs at this point attain \gls*{sms} (\Cref{thm:ccb:rtp:sms}).
-Since these properties compose into \gls*{ms}, composing these passes yields a compiler that robustly preserves \gls*{ms} (\Cref{corr:ccab:rtp:ms}).
-Then, the section presents two optimisation passes, namely \gls*{cf} and \gls*{dce}, each of which robustly preserves \gls*{ms} (\Cref{thm:ccdce:rtp:ms,thm:cccf:rtp:ms}).
-These passes can be freely ordered in the compilation chain without compromising memory safety (\Cref{thm:cccfccdce:rtp:ms}).
-The next step in the chain ensures that code stays \gls*{scct} (\Cref{thm:ccscct:rtp:scct}) when compiled from $\Lms$ to $\Lscct$, which is done by switching on a constant-tame mode for the computation.
-Lastly, by introducing barriers immediately after branches, speculative leaks via SPECTRE-PHT are prevented when compiling $\Lscct$ to $\Lspec$.
-The final result is that the whole compilation chain robustly preserves \gls*{specms} (\Cref{thm:ccall:rtp:specms}).
-
-\subsection{Robust Temporal Memory Safety Preservation}\label{subsec:cs:tms}
-
-The secure compiler from $\Ltms$ to $\Ltrg$ needs to ensure that when execution switches from context to component, the type signatures are respected.
-Thus, the compiler inserts the following dynamic typechecks before entering the body of a component-defined function (anything elided is a trivial identity function from source to target):
-
-\vspace{-1em}
-{
-\begin{gather}
-  \begin{align*}%{rcl}
-    %\cca(\src{x}) = &\ \trg{x} 
-    % \\
-  	%&
-  	%\qquad
-    %\cca(\src{\lbinop{\expr[_{1}]}{\expr[_{2}]}}) = &\ \lbinop[\trg]{\left[\cca(\src{\expr[_{1}]})\right]}{\left[\cca(\src{\expr[_{2}]})\right]} \\
-    %\cca(\src{n}) = &\ \trg{n} 
-    % \\
-    %&
-    \cca(\src{\lget{x}{\expr}}) = &\ \lget[\trg]{\trg{x}}{\left[\cca(\src{\expr})\right]} \\
-    \cca(\src{\ldelete{x}}) = &\ \ldelete[\trg]{\left[\cca(\src{x})\right]} 
-    \\
-    \cca(\src{\lfunction{g}{x:\natt\to\type[_{e}]}{\expr}})  =&
-  \end{align*}
-  \\
-  \begin{align*}
-%    \cca(\src{\lset{x}{\expr[_{1}]}{\expr[_{2}]}}) & = \lset[\trg]{\left[\cca(\src{x})\right]}{\left[\cca(\src{\expr[_{1}]})\right]}{\left[\cca(\src{\expr_{2}})\right]} \\
-% \\ 
-\lfunction[\trg]{\trg{g}}{\trg{x}}{\lifz[\trg]{\trg{\lhast{x}{\natt}}}{
-                                                            \left[\cca(\src{\expr})\right] %
-                                                                                                 }{\labort[\trg]}}
-  \end{align*}
-\end{gather}
-}
-
-\noindent Since $\trg{L}$ has no static typing, an attacker $\trg{\library_{\ctx}}$ can invoke a component function accepting a $\src{\natt}$ with $\trg{\lpair{17}{29}}$.
-With the dynamic check, the compiler ensures that execution aborts in such cases.
-%The compiler does not insert other checks and proceeds as the identity function (which in this paper amounts to a simple re-colouring of $\src{L_{\tmssafe}}$ to $\trg{L}$ expressions).
-
-Compiling the \texttt{strncpy} function from \Cref{sec:introduction} with $\cca$, the compiler would in this case ensure that the arguments that are evaluated in the compiled \texttt{strncpy} are valid.
-%This effectively has no influence on the resulting trace, so a corresponding cross-language trace relation $\sim^{\Ltms}_{\Ltrg}$ is just equality.
-
-$\cca$ is robustly preserving (\Cref{def:rtp}) \gls*{tms}:
-\begin{theorem}[$\cca$ secure w.r.t. \gls*{tms}]\label{thm:cca:rtp:tms}
-    $\rtp{\cca}{\tmssafe}$ % \Coqed
-\end{theorem}
-
-\subsection{Robust Spatial Memory Safety Preservation}\label{subsec:cs:ms}
-
-The spatial memory safety preserving compiler from $\Ltrg$ to $\Lms$ only inserts bounds-checks whenever reading from or writing to memory in order to enforce \gls*{sms}.
-These bounds checks need the bounds information, which the compiler keeps around by introducing a fresh identifier $\irl{x_{SIZE}}$ for each allocation that binds $\irl{x}$.
-Then, it is simply a matter of referring to that variable and ensuring that memory accesses are in the interval $[\irl{0},\irl{x_{SIZE}})$.
-When the check fails, the code aborts.
-
-\begin{nscenter}
-% \small
-  $$
-  \begin{array}{rcl}
-    \ccb(\trg{\lnew{x}{\expr[_{1}]}{\expr[_{2}]}}) & = 
-                                                   & \llet[\irl]{\irl{x_{SIZE}}}{\ccb(\trg{\expr[_{1}]})}{}
-    		\\&&
-    		\lnew[\irl]{\irl{x}}{\irl{x_{SIZE}}}{\ccb(\trg{\expr[_{2}]})}
-    		 \\
-  \ccb(\trg{\lget{x}{\expr}}) & = 
-                              & \llet[\irl]{\irl{x_{n}}}{\ccb(\trg{\expr})}{}
-  	\\&&
-  \lifz[\irl]{\irl{0\le x_{n}<x_{SIZE}}}{\\&&\irl{\lget{x}{x_{n}}}}{}
-  		\irl{\labort}
-  	  \\
-  \ccb(\trg{\lset{x}{\expr[_{1}]}{\expr[_{2}]}}) & = 
-                                                 & \llet[\irl]{\irl{x_{n}}}{\ccb(\trg{\expr[_{1}]})}{}
-  		\\&&
-  \lifz[\irl]{\irl{0\le x_{n}<x_{SIZE}}}{\\&&\lset[\irl]{\irl{x}}{\irl{x_{n}}}{}
-  		\ccb(\trg{\expr[_{2}]})
-  		}{\irl{\labort}} 
-  		% \\
-  \end{array}
-  $$
-\end{nscenter}
-
-\begin{exampleenv}[Instrumented \texttt{strncpy}]
-  Consider again \texttt{strncpy}, but instrumented for \gls*{sms}:
-    \begin{lstlisting}[language=c,basicstyle=\ttfamily, morekeywords={size_t}]
-void strncpy(size_t n, size_t dst_size, char *dst,
-             size_t src_size, char *src) {
-  for(size_t i = 0; i < src_size
-      && src[i] != '\0' && i < n; ++i) {
-    if(i < src_size && i < dst_size) {
-      dst[i] = src[i];
-    }
-  }
-}
-    \end{lstlisting}
-Consider context \texttt{strncpy(2, x, y)}, where \texttt{x} and \texttt{y} are pointers to valid regions of memory with allocated space for exactly two cells and do not contain the null-terminating character \texttt{'\textbackslash 0'}.
-% 
-Without the \gls*{sms} pass, the event $\ev{Use}\ \loc_{x}\ 2;\comp;\unlock$ would appear on the trace, but that indicates an out-of-bounds access! 
-Fortunately, with \gls*{sms} mitigation in place, that event does not appear during execution, since the bounds check prevents the condition \texttt{src[i] != '\textbackslash 0'} from executing.
-\end{exampleenv}
-
-Contrary to the previous compiler, $\ccb$ may change the trace of the original $\Ltrg$ program: if there is a memory access, it needs to be protected with a bounds check.
-% 
-The corresponding relation $\sim^{\Ltrg}_{\Lms} : \trg{\trace}\times\irl{\trace}$ that describes this semantic effect of the compiler is defined partially below.
-% 
-We omit action $\trg{Set}$, which is treated analogously, and any other event, which is related to its cross-language equivalent.
-% \footnote{Up to colours.}
-% \footnote{In the interest of space, the full details are in the technical report.}
-
-{
-\[
-  \typerulenolabel{xrel:sms:read}{\trg{n}\text{ in bounds}}{\trg{Get\ \loc\ n;\comp}\sim^{\Ltrg}_{\Lms}\irl{Get\ \loc\ n;\comp}}
-\]
-% \[
-%   \typerulenolabel{xrel:sms:panic}{\trg{n}\text{ out of bounds}}{\trg{Get\ \loc\ n;\comp}\sim^{\Ltrg}_{\Lms}\irl{\lightning}}
-% \]
-}
-
-% Note that neither $\Ltrg$ nor $\Lms$ expose branches on the trace.\MP{so what?}
-For simplicity, we elide the environment that this relation carries around in order to bind each location to its metadata (such as its size), and resolve the ``$\trg{n}$ in/out of bounds'' premise.
-% The bounds check mentioned in the premises of the rules above is left abstract for the sake of simplicity, the interested reader can find
-% since the precise definition needs some technicalities that only inhibit readability without any technical insights.
-% 
-We can now prove that compiler $\ccb$ robustly preserves \gls*{sms}.
-\begin{theorem}[$\ccb$ secure w.r.t. \gls*{sms}]\label{thm:ccb:rtp:sms}
-  $\rtpsim{\ccb}{\smssafe}{\sim^{\Ltrg}_{\Lms}}$ % \Coqed
-\end{theorem}
-
-At this point we can compose $\ccb$ with the previous compiler ($\cca$), but in order to do so, we need a trace relation from $\Ltms$ to $\Lms$.
-% 
-We can obtain this relation by composing the trace relation we just defined ($\sim^{\Ltrg}_{\Lms}$) with the one used by the previous compiler: $\sim^{\Ltms}_{\Ltrg} : \src{\trace}\times\trg{\trace}$.
-% 
-The latter has not been previously defined (nor has it been used in the related theorem) because that is just an equality relation, since the trace models of $\Ltms$ and of $\Ltrg$ are the same.
-% 
-Thus, we formally define $\sim^{\Ltms}_{\Lms}: \src{\trace} \times \irl{\trace}$ as the following composition: $\sim^{\Ltms}_{\Ltrg}\bullet\sim^{\Ltrg}_{\Lms}$.
-% 
-With this relation, \Cref{corr:ccab:rtp:ms} states that the composition of $\cca$ and $\ccb$ is secure w.r.t.
-\gls*{ms} and it follows from \Cref{thm:cca:rtp:tms,thm:ccb:rtp:sms} using \Cref{thm:rtpsim:sig}.
-% 
-\begin{corollary}[$\cca\circ\ccb$ secure w.r.t. \gls*{ms}]\label{corr:ccab:rtp:ms}
-  $\;$
-
-  \begin{nscenter}
-    $\rtpsim{\cca\circ\ccb}{\mssafe}{\sim^{\Ltms}_{\Lms}}$ % \Coqed
-  \end{nscenter}
-\end{corollary}
-% 
-This proof requires another precondition besides \Cref{thm:ccb:rtp:sms,thm:cca:rtp:tms}: $\sim^{\Ltms}_{\Ltrg}$ needs to be well-formed with respect to $\smssafe$.
-This follows trivially since $\sim^{\Ltms}_{\Ltrg}$ is an equality. 
-% 
-\begin{lemma}[$\sim^{\Ltms}_{\Ltrg}$ well-formed w.r.t. $\smssafe$]\label{lem:wf:ltmsltrg}
-$\;$
-
-  \begin{nscenter}
-  $\wfcsig{\sim^{\Ltms}_{\Ltrg}}{\smssafe}$
-  \end{nscenter}
-\end{lemma}
-
-\subsection{Optimising Compilers}\label{subsec:cs:opts}
-
-This section defines two optimising compiler passes from $\Lms$ to $\Lms$ which perform \gls*{dce} and \gls*{cf}, respectively.
-The \gls*{dce} pass applies a na\"ive rewrite rule on conditionals.
-The \gls*{cf} pass relies on an auxiliary function \texttt{mix} that uses a substitutions accumulator $\irl{\substlist}$ in order to rewrite constant binary operations, e.g., $\irl{{17}-1}$ to $\irl{16}$, and replace variables that are assigned to constants, e.g., $\irl{\llet{x}{7}{x}}$ to $\irl{7}$.
-
-\vspace{-1em}
-\begin{gather*}
-  \begin{align*}
-    \ccdce(\irl{\lifz{true}{\expr[_{1}]}{\expr[_{2}]}}) & = \ccdce(\irl{\expr[_{1}]}) &\\
-    \ccdce(\irl{\lifz{false}{\expr[_{1}]}{\expr[_{2}]}}) & = \ccdce(\irl{\expr[_{2}]}) &
-  \end{align*}
-  \\
-  % \begin{align*}
-  %   \ccdce(\irl{\lbinop{\expr[_{1}]}{\expr[_{2}]}}) & = \lbinop[\irl]{\ccdce(\irl{\expr[_{1}]})}{\ccdce(\irl{\expr[_{2}]})} &
-  % \end{align*}
-  % \\[0.25cm]
-  \begin{align*}
-    \cccf(\irl{\expr}) & = \partialeval{\irl{\expr}}{\irl{\hole{\cdot}}} &
-  \end{align*}
-  \\[0.125cm]
-  \begin{align*}
-   \partialeval{\irl{x}}{\irl{\substlist}} & = \irl{n} 
-   	\qquad\qquad \text{if } \irl{\subst{n}{x}}\in\irl{\substlist} \\
-   \partialeval{\irl{x}}{\irl{\substlist}} & = \irl{x} 
-   \qquad\qquad \text{if } \irl{\subst{n}{x}}\notin\irl{\substlist} \\
-   \partialeval{\irl{\lbinop{n}{m}}}{\irl{\substlist}} & = \irl{k} 
-   \qquad\qquad \text{if } \lbinop{\irl{n}}{\irl{m}}=k \\
-   %\partialeval{\irl{\lbinop{\expr[_{1}]}{\expr[_{2}]}}}{\irl{\substlist}} & = \lbinop[\irl]{\partialeval{\irl{\expr[_{1}]}}{\irl{\substlist}}}{\partialeval{\irl{\expr[_{2}]}}{\irl{\substlist}}} \\
-   \partialeval{\irl{\llet{x}{n}{\expr}}}{\irl{\substlist}} & = \partialeval{\irl{\expr}}{\irl{\subst{n}{x}\cdot\substlist}} 
-  % \\
-  % \partialeval{\irl{\lget{x}{\expr}}}{\irl{\substlist}} & = \lget[\irl]{\irl{x}}{\partialeval{\irl{\expr}}{\irl{\substlist}}}
-% \end{align*}
-\end{align*}
-% \\
-% \begin{align*}
-%   \partialeval{\irl{\llet{x}{\expr[_{1}]}{\expr[_{2}]}}}{\irl{\substlist}} & = \llet[\irl]{\irl{x}}{\partialeval{\irl{\expr[_{1}]}}{\irl{\substlist}}}{\\&\partialeval{\irl{\expr[_{2}]}}{\irl{\substlist}}} \\
-%   \partialeval{\irl{\lifz{\expr[_{1}]}{\expr[_{2}]}{\expr[_{3}]}}}{\irl{\substlist}} & = \lifz[\irl]{\partialeval{\irl{\expr[_{1}]}}{\irl{\substlist}}}{\\&\partialeval{\irl{\expr[_{2}]}}{\irl{\substlist}}}{\partialeval{\irl{\expr[_{3}]}}{\irl{\substlist}}} \\
-% \end{align*}
-\end{gather*}
-% \vspace{-3em}
-
-Note that both passes have no effect on the resulting trace of a program, up to $\emptyevent$-steps. 
-Because of this, both passes have equality as corresponding cross language trace relation. 
-Moreover, it is straightforward to prove both passes as secure (\Cref{def:rtp}) w.r.t. \gls*{ms}.
-
-\begin{theorem}[$\ccdce$ secure w.r.t. \gls*{ms}]\label{thm:ccdce:rtp:ms}
-  $\rtp{\ccdce}{\mssafe}$ %\Coqed
-\end{theorem}
-\begin{theorem}[$\cccf$ secure w.r.t. \gls*{ms}]\label{thm:cccf:rtp:ms}
-  $\rtp{\cccf}{\mssafe}$ %\Coqe
-\end{theorem}
-
-With both \Cref{thm:ccdce:rtp:ms,thm:cccf:rtp:ms} it follows from \Cref{corr:swappable} that the two passes can be interchanged arbitrarily:
-
-\begin{theorem}[$\cccf\circ\ccdce$ and $\cccf\circ\ccdce$ are secure w.r.t. \gls*{ms}]\label{thm:cccfccdce:rtp:ms}
-$\;$
-
-  \begin{nscenter}
-  \phantom{and~}\!\!$\rtp{\cccf\circ\ccdce}{\mssafe}$
-
-  and~$\rtp{\ccdce\circ\cccf}{\mssafe}$ % \Coqed
-  \end{nscenter}
-\end{theorem}
-
-\subsection{Robust Strict Cryptographic Constant Time Preservation}\label{subsec:cs:scct}
-
-This section defines a compiler $\ccscct$ from $\Lms$ to $\Lscct$ that robustly preserves \gls*{scct}. 
-Given the fact that $\Lscct$ provides a \gls*{cct}-mode that can be turned on or off, the compiler inserts wrapper code for function calls and function bodies to ensure that execution in the component always happen in \gls*{cct}-mode.
-This simple flag combines the effect of FaCT~\cite{cauligi2019fact}
-
-\vspace{-1em}
-\[
-\begin{array}{rcl}
-  \ccscct(\irl{\lfunction{g}{x}{\expr}}) & = & \lfunction[\obj]{\obj{g}}{\obj{x}}{\obj{\lwrdoit{ON};}\ccscct(\irl{\expr})} \\
-  \ccscct(\irl{\lcall{g}{\expr}}) & = & \lcall[\obj]{\obj{g}}{\ccscct(\irl{\expr})\obj{; \lwrdoit{ON}}} 
-    % \\
-    % \ccscct(\irl{\lbinop{\expr[_{1}]}{\expr[_{2}]}}) & = \lbinop[\obj]{\ccscct{\irl{\expr[_{1}]}}}{\ccscct{\irl{\expr[_{2}]}}} 
-    % \\
-\end{array}
-\]
-% \vspace{-2em}
-%
-The context can overwrite the flag and exit the mode, but upon invoking a function that is part of the component, the flag is set again.
-Because of this, the corresponding cross-language trace relation $\sim^{\Lms}_{\Lscct}$, only relates events without secrets:%
-
-\begin{center} 
-  \typerulenolabel{xrel:scct:noleak}{}{\irl{\preevent;\comp}\sim^{\Lms}_{\Lscct}\obj{\preevent;\comp;\unlock}}
-\end{center}
-
-The compiler is secure w.r.t. \gls*{scct}: %, similarly proven as in \Cref{subsec:cs:tms}.
-
-\begin{theorem}[$\ccscct$ secure w.r.t. \gls*{scct}]\label{thm:ccscct:rtp:scct}
-  \small
-  $\rtpsim{\ccscct}{\scctsafe}{\sim^{\Lms}_{\Lscct}}$ % \Coqed
-\end{theorem}
-
-\subsection{Robust Speculative Safety Preservation}\label{subsec:cs:ss}
-
-This section defines the final compilation pass $\ccspec$, which ensures that $\Lscct$ programs, which are assumed to be \gls*{ss}, stay \gls*{ss} at $\Lspec$-level. 
-To do so, $\ccspec$ inserts a speculation barrier after branches, which is sufficient to harden the program against speculative leaks, since SPECTRE-PHT~\cite{kocher2019spectre} is the only speculative leak modeled in the semantics of $\Lspec$.
-
-\vspace{-1em}
-\[
-\begin{array}{cl}
-  &\ccspec{(\obj{\lifz{\expr[_0]}{\expr[_1]}{\expr[_2]}})} = 
-  \\
-  &\qquad\qquad \lifz[\ird]{\ccspec{\left(\obj{\expr[_0]}\right)}}{\ird{\lbarrier;}\ccspec{\left(\obj{\expr[_1]}\right)} \\&\qquad\qquad}{ \ird{\lbarrier;}\ccspec{\left(\obj{\expr[_2]}\right)}} 
-\end{array}
-\]
-% \vspace{-2em}
-%
-
-Clearly, the corresponding cross-language trace relation $\sim^{\Lscct}_{\Lspec}$ has only one non-trivial case: for branches, only relate them where speculation is blocked by a barrier:
-
-\begin{center}
-  \typerulenolabel{xrel:spec:if}{}{\obj{Branch\ n}\sim^{\Lscct}_{\Lspec}\ird{Branch\ n}\cdot\ird{Spec}\cdot\ird{Barrier}\cdot\ird{Rlb}}
-\end{center}
-
-The base-event relation above scales to full events by ensuring the missing annotations ($\obj{\comp;\securitytag{}}$ and $\ird{\comp;\securitytag{}}$) are the same.
-% 
-With this relation, we prove that $\ccspec$ is secure with respect to \gls*{ss}.
-\begin{theorem}[$\ccspec$ secure w.r.t. \gls*{ss}]\label{thm:ccspec:rtp:spec}
-  \small$\rtpsim{\ccspec}{\sssafe}{\sim^{\Lscct}_{\Lspec}}$ % \Coqed
-\end{theorem}
-
-\subsection{Robust Preservation of Memory Safety, Strict Cryptographic Constant Time, and Speculative Safety}
-
-Finally, this subsection combines all previous results into one compilation chain to get that it preserves full \gls*{specms}.
-Let $\ccspecms$ be the compiler that is the composition of $\cca$, $\ccb$, $\cccf$, $\ccdce$, $\ccscct$, and $\ccspec$. 
-Let $\sim^{\Ltms}_{\Lspec}$ be the composition of $\sim^{\Ltms}_{\Lms}$, $\sim^{\Lms}_{\Lscct}$, and $\sim^{\Lscct}_{\Lspec}$.
-Then, the following corollary holds.
-
-\begin{corollary}[$\ccspecms$ secure w.r.t. \gls*{specms}]\label{thm:ccall:rtp:specms}
-  $\;$
-
-  \begin{nscenter}
-    $\rtpsim{\cc{\Ltms}{\Lspec}}{\mssafe\cap\scctsafe\cap\sssafe}{\sim^{\Ltms}_{\Lspec}}$ % \Coqed
-  \end{nscenter}
-\end{corollary}
-
-As with \Cref{corr:ccab:rtp:ms}, it is important to ensure that the respective cross language trace relations are well-formed (\Cref{def:wfc:sig:tracerel}).
-It is already known that $\sim^{\Ltms}_{\Lms}$ is well-formed with respect to $\mssafe$ (\Cref{lem:wf:ltmsltrg}).
-Next in the chain is $\sim^{\Lms}_{\Lscct}$, which has to be well-formed w.r.t. $\scctsafe$.
-This lemma holds, since a trace that was $\scctsafe$ is $\scctsafe$ even after applying $\sim^{\Lms}_{\Lscct}$: the relation enforces that $\Lscct$ traces related to $\Lms$ traces have no leaks of secrets whatsoever.
-
-\begin{lemma}[$\sim^{\Lms}_{\Lscct}$ well-formed w.r.t. $\scctsafe$]\label{lem:wf:lsmslscct}
-$\;$
-
-  \begin{nscenter}
-  $\wfcsig{\sim^{\Lms}_{\Lscct}}{\scctsafe}$
-  \end{nscenter}
-\end{lemma}
-
-The last relation is $\sim^{\Lscct}_{\Lspec}$ which needs to be well-formed w.r.t. $\sssafe$.
-Similarly to the previous relation, this holds, since $\sim^{\Lscct}_{\Lspec}$ only relates $\Lspec$ traces, which do not have speculative leaks, with $\Lscct$ traces.
-
-\begin{lemma}[$\sim^{\Lscct}_{\Lspec}$ well-formed w.r.t. $\sssafe$]\label{lem:wf:lscctlspec}
-  $\wfcsig{\sim^{\Lscct}_{\Lspec}}{\sssafe}$
-\end{lemma}
-
-\section{Formal Insights}\label{sec:formalities}
-
-This section first discusses how to connect each language-specific security property to the general security properties presented in \Cref{sec:compprop} (\Cref{subsec:formalities:maps}), then it demonstrates that the security property resulting from the universal image projection is faithful to the original property (\Cref{subsec:formalities:props}).
-Finally, this section discusses why the order of compiler passes matters, and how does our framework help with identifying insecure compositions (\Cref{subsec:compatsecpasses}), and it gives additional technical insights on the secure compilation proofs (\Cref{subsec:seccompproofs}).
-
-\subsection{From Language Traces to General Ones}\label{subsec:formalities:maps}
-
-The previous theorems talk about preserving properties expressed in the trace model of the languages of each compiler.
-However, these trace models are not the same trace model we used to specify the properties of \Cref{sec:compprop} (indicated with \ev{L}), which serves as the ``ground truth'' for the meaning of our properties.
-% 
-To bridge this gap, the formal development requires specifying additional trace relations, from each of the language trace models to the \ev{L} one, that, for example, relate $\src{Get\ \loc\ n}$ and $\src{Set\ \loc\ n}$ to $\ev{Use\ \loc\ n}$ (and that induce a related universal image that we use in \Cref{subsec:formalities:props}).
-% 
-One key insight of these relations is that they all omit context-made actions for two reasons: (1) contexts (which are universally quantified) can trivially invalidate any property and (2) we are interested in the component upholding the properties.
-%
-
- % disregarded a glaring but minor technical issue: take $\tmssafe$ (\Cref{def:trace:tmsdef}) as an example.
-% It contains traces as specified in \Cref{subsec:basic:memsafety:tracemodel}.
-% However, these are different from $\Ltms$ traces (\Cref{subsec:ltms}) which means that \Cref{thm:wt:tms} is ``ill-typed''.
-% Similarly for the other theorems.
-% In the interest of reducing visual baggage, the notation left out yet another cross-language trace relation, translating language-specific traces into these non-language specific traces. 
-% In doing so, context-level actions are dropped and, e.g., both $\src{Get\ \loc\ n}$ and $\src{Set\ \loc\ n}$ are translated into $\ev{Use\ \loc\ n}$.
-
-\subsection{Security Properties and Their Meaning}\label{subsec:formalities:props}
-Each of the presented compilers use a cross-language trace relation, which is also used to translate the property from one language to the other one (via the existential or universal images).
-% 
-%Whenever such translations occur, it is important to check whether there has been a change of meaning.
-%Secure compiler engineers need to ensure that such translations of properties do not change meaning in a fundamental wrong way. 
-% lifting or lowering reasoning from target- to source-level or vice-versa,
-While the meaning of projected properties does change with a translation, the change should not allow for a flawed compilation pipeline.
-% 
-For example, we could be using a trace relation that translates a property in a language to a totally different one in another language.
-% 
-To raise the trust into the translation of properties, \Cref{thm:prop-rel-corr} states (in a general fashion) that each security property is faithfully translated using the universal image according to the cross-language trace relation induced by the compiler.
-
-\begin{theorem}[Properties Relation Correctness]\label{thm:prop-rel-corr}
-  \begin{align*}
-    \forall& \pi \in \{\tmssafe,\smssafe,\scctsafe,\sssafe\}, 
-    \\
-    \text{ for each }& \text{pair of languages } \src{L} \text{ and } \trg{L} \text{ used by the compilers},
-    \\
-    \text{ if }& 
-    % \tau_{\sim^{\trg{L}}_{\ev{L}}}(\pi) \sim^{\trg{L}}_{\ev{L}} \pi
-    \sigma_{\sim^{\trg{L}}_{\ev{L}}}(\pi) \sim^{\trg{L}}_{\ev{L}} \pi
-    \text{ and } 
-    % \sigma_{\sim^{\src{L}}_{\trg{L}}}(\tau_{\sim^{\trg{L}}_{\ev{L}}}(\pi)) \sim^{\src{L}}_{\trg{L}} \tau_{\sim^{\trg{L}}_{\ev{L}}}(\pi)
-    \sigma_{\sim^{\src{L}}_{\trg{L}}}(\sigma_{\sim^{\trg{L}}_{\ev{L}}}(\pi)) \sim^{\src{L}}_{\trg{L}} \sigma_{\sim^{\trg{L}}_{\ev{L}}}(\pi)
-    \\
-    \text{ then }& 
-    % \tau_{\sim^{\src{L}}_{\ev{L}}}(\pi) = \sigma_{\sim^{\src{L}}_{\trg{L}}}(\tau_{\sim^{\trg{L}}_{\ev{L}}}(\pi)) 
-    \sigma_{\sim^{\src{L}}_{\trg{L}}}(\sigma_{\sim^{\trg{L}}_{\ev{L}}}(\pi)) \sim^{\src{L}}_{\ev{L}} \pi
-  \end{align*}
-\end{theorem}
-The complexity of this theorem is that the relation in the conclusion cannot be obtained by composing the two relations in the premises.
-% 
-%% MK: the following does not say anything about the theorem, only about the syntactic compiler
-%What happens in practice is that compiler engineers are given the relations from each language to \ev{L} and they need to build a compiler that induces a relation $\sim^{\src{L}}_{\trg{L}}$ that respects this theorem.
-%
-
-We now informally argue why this theorem holds for the composition of all considered properties from \Cref{sec:compprop}.
-
-\paragraph{$\sigma_{\sim^{\Ltms}_{\Lms}}\left(\tmssafe\cap\smssafe\cap\scctsafe\cap\sssafe\right)$}
-
-For the four properties considered here, the trace models of $\Ltms$, $\Ltrg$, and $\Lms$ do not consider actions related to $\scctsafe$ and $\sssafe$, so these two properties are trivially translated correctly.
-% 
-We now discuss the remaining $\tmssafe$ and $\smssafe$ in the form of their intersection $\mssafe$.
-
-Recall that $\sim^{\Ltms}_{\Lms}$ is the composition of $\sim^{\Ltms}_{\Ltrg}$ and $\sim^{\Ltms}_{\Lms}$.
-% 
-Since $\sim^{\Ltms}_{\Ltrg}$ is an equality, this relation trivially preserves the meaning of translated properties: related traces are identical!
-
-Finally, let us consider a trace $\irl{\trace}\in\mssafe$ and understand what is that is related to via $\sim^{\Ltrg}_{\Lms}$.
-% 
-{\em All} traces $\trg{\trace}$ with $\trg{\trace}\sim^{\Ltrg}_{\Lms}\irl{\trace}$ are identical to $\irl{\trace}$ except for get and set actions, which require for in-bound accesses (as stated in \Cref{subsec:cs:ms}): this clearly respect $\mssafe$.
-%
-
-% In both cases, the translated event does not represent an out-of-bounds memory access.
-
-% \footnote{Both $\sim^{\Ltms}_{\Ltrg}$ and $\sim^{\Ltrg}_{\Lms}$ filter context actions.}.
-
-% Now, starting from $\mssafe$ expressed in $\Lms$, we check that its translation in $\Ltms$ is still $\mssafe$.
-% % 
-% We can see that the property resulting from the projection of $\Lms$ traces back to $\Ltms$ faithfully models $\mssafe$, which translates to both $\scctsafe$ and $\sssafe$, 
-% \MPin{ equiv  +  other rel only adds traces that fail }
-% since these properties trivially hold for $\Ltms$, $\Ltrg$, and $\Lms$ traces.
-
-\paragraph{$\sigma_{\sim^{\Lms}_{\Lscct}}\left(\tmssafe\cap\smssafe\cap\scctsafe\cap\sssafe\right)$}
-
-As before, the trace models of $\Lscct$ and $\Lms$ cannot express $\sssafe$, so that is trivially translated correctly.
-% 
-Also, the trace model of $\Lscct$ extends the one of $\Lms$ with respect to $\tmssafe$ and $\smssafe$ events, so the translation argument regarding those two properties is the same as before.
-% 
-Thus, we need to reason about whether $\scctsafe$ is translated correctly.
-
-By definition, $\sim^{\Lms}_{\Lscct}$ only relates $\obj{\unlock}$ events, which is also the same relation induced by $\sim_{\ctsafe}$ in \Cref{sec:msscct-rel}.
-% 
-This ensures that composing the relations only relates $\unlock$ events, and thus the property is translated correctly.
-
-% Consider a trace $\obj{\trace}\in\scctsafe$ and all $\irl{\trace}$ such that $\irl{\trace}\sim^{\Lms}_{\Lscct}\obj{\trace}$. 
-% In turn, $\obj{\trace}$ cannot contain any leaks of secret values, so it cannot violate $\scctsafe$.
-% Toggling data-independent timing mode has no impact on $\mssafe$ or $\sssafe$.
-
-\paragraph{$\sigma_{\sim^{\Lscct}_{\Lspec}}\left(\tmssafe\cap\smssafe\cap\scctsafe\cap\sssafe\right)$}
-The trace model of $\Lspec$ extends the one of $\Lscct$ with respect to $\tmssafe$, $\smssafe$, and $\scctsafe$, so the translation argument for those three properties is the same as before.
-% 
-Concerning $\sssafe$, $\sim^{\Lscct}_{\Lspec}$ relates $\Lspec$ speculation traces with $\Lscct$ branches with the $\lock_{\text{NONE}}$ tag, so the property is translated correctly, according to the relation defined in \Cref{sec:spec-ms-rel}.
- % introducing $\ird{barrier}$ never violates $\sssafe$ and has no effect on $\mssafe$ or $\scctsafe$.
-
-%\MPin{
-%	can this be made formal?
-%	what's the point?
-%}
-
-\subsection{Compatibility of Secure Compiler Passes}\label{subsec:compatsecpasses}
-
-% \Cref{thm:rtpsim:sig} is applicable, up to the well-formedness of $\sim_1$ w.r.t. $\trg{\class[_2]}$.
-Consider applying $\ccspec$ first and then $\ccb$ (albeit currently syntactically impossible).
-% , let us pretend that it is possible by just extending $\ccb$ for the constructs present in $\Lscct$ that are not present in $\Ltrg$.
-% Clearly, as evidenced earlier (\Cref{subsec:formalities:props}), $\ccb\cdot\ccspec$ is an acceptable composition that yields an interesting class of security properties, i.e., $\sigma_{\sim_{\irl{ms}}\bullet\sim_{\ird{\mathghost}}}\left(\mssafe\cap\scctsafe\cap\sssafe\right)$.
-In this case, $\ccb$ would insert new branches into the code that are not protected by a speculation barrier! 
-This concern is reflected in the proofs that establish that the security class resulting of the composition of the trace relations is meaningful.
-% , much similar to ensuring that a definition models what the formal methods expert intends to model. 
-For the $\ccspec\cdot\ccb$ case, the class is $\sigma_{\sim_{\ird{\mathghost}}\bullet\sim_{\irl{ms}}}\left(\mssafe\cap\scctsafe\cap\sssafe\right)$.
-Since $\ev{Use\ \loc\ n}\sim_{\irl{ms}}\irl{Branch\ n}\cdot\irl{Spec}\cdots$, where $\cdots$ does not contain a $\ev{{Barrier}}$ event, the resulting class is \emph{not} the original $\sssafe$ that is intended and it would break the corresponding \Cref{thm:prop-rel-corr}.
-% Because of this, it is crucial to analyse the precise shape of the class of properties after mapping a target-level property to a source-level property\footnote{Dito for mapping a source-level property to a target-level property.}.
-
-However, the composition is still technically possible and it is the job of the compiler engineers to ensure that the secure compilation pipeline happens in an order that ensures that the mapped security property is the intended one.
-
-\subsection{Secure Compilation Proofs}\label{subsec:seccompproofs}
-
-Our secure compilation proofs rely on backtranslations~\cite{abate2019jour,patrignani2021rsc}, which let one construct a source context starting from either target traces (aka trace-based backtranslations) or target contexts (aka context-based backtranslations).
-These backtranslations also require setting up cross-language relations between various language elements such as expressions and program states, so we leave these details for the technical reports.
-% , the usual strategy is to establish a backwards simulation between source components and their compiled target counterpart. 
-% For the universally quantified target context, a technique known as backtranslation is well-established, which constructs a source-level context either by backtranslating the target-level context, amounting to a ``decompilation'', or by constructing it using the target-level trace, so that the source-context emits parts of the whole execution trace that are related to the parts emitted by the target-level context.
-All backtranslations in the case-study are trace based except for those required by $\ccdce$ and $\cccf$, which are context-based (and they are an identity function). 
-% Since these optimisations have no significant semantic effect and source- and target-language are equal, the backtranslation amounts to an identity translation.
-
-\section{Related Work\pages{2}}\label{sec:relwork}
-
-This section discusses robust compilation, other secure compilation criteria and work related to the properties preserved in the case study.
-% (\Cref{subsec:relw:seccomprtp}) 
-% (\Cref{subsec:relw:seccompcrit}).
-% Since the case study of \Cref{sec:casestud:defs,sec:casestud:rtp} implements measures for preserving \gls*{ms}, \gls*{cct}, and \gls*{ss}, 
-% Then, this section discusses work related 
-% relevant related work as well 
-% (\Cref{subsec:relw:msmechs,subsec:relw:cctmechs,subsec:relw:ssmechs}).
-
-\paragraph*{Secure Compilation as Robust Preservation}\label{subsec:relw:seccomprtp}
-
-The robust preservation of properties as a compiler-level criterion has been analysed extensively~\cite{abate2019jour,patrignani2021rsc,abate2021extacc,patrignani2019survey} and thus we build on that framework.
-No existing work is concerned with composing secure compilers, however, existing work~\cite{abate2021extacc} sketches composition of trace-relating compiler correctness in a similar way to what has been presented here.
-%Parts of these works consider languages with different trace models and our technical setup can handle this. 
-The work relating robust preservation with universal composability~\cite{patrignani2022universal} is closest to what this paper presents.
-The authors demonstrate a similar compositionality theorem to what is presented here (\Cref{sec:sequential}) but use it in the context of protocols.
-The work does not consider the generality to support different trace models or composition of compilers which robustly preserve different classes.
-% The authors demonstrate a similar compositionality theorem to what is presented here (\Cref{sec:sequential}) as well as in an earlier version of this work~\cite{kruse2022csc}.
-% However, they do not demonstrate the scalability of the approach by means of a case study.
-
-%\paragraph*{Composition of Secure Compilers}\label{subsec:relw:seccompcomp}
-
-There is work on lifting exploits for single compilers to the whole chain~\cite{paykin2019weird}.
-%Our framework does not consider exploits.
-%\MP{ a bit dry and short}
-While that work considers {\em in}secure compilation and composition thereof in terms of exploits, the composition they are interested in allows to lift an exploit for one compiler pass to the whole compilation chain.
-Our framework takes the opposite direction and provides compositionality results for secure compilers.
-%
-%
-%To the best of our knowledge, there does not exist published work discussing the compositionality of the robust preservation criterion, the gold-standard of secure compilation.
-
-\paragraph*{Other Secure Compilation Criteria}\label{subsec:relw:seccompcrit}
-
-While this paper focuses on the robust preservation framework~\cite{abate2019jour}, other secure compilation criteria exist.
-The survey on formal approaches to secure compilation~\cite{patrignani2019survey} discusses a broad spectrum already, while this section presents a very high-level overview.
-Fully abstract compilation~\cite{abadi1999fullabstraction} states that a compiler should preserve and reflect observational equivalence between source and target programs.
-Abate \emph{et al.}~\cite{abate2021faandrc} showed that fully abstract compilers robustly preserve program properties that are either trivial or meaningless.
-As a mitigation for this, the authors presented a categorical approach based on maps of distributive laws~\cite{watanabe2002modl}, which they call many maps of distributive laws.
-Maps of distributive laws have been investigated before as a possible secure compilation criterion~\cite{tsampas2020catsc}.
-Other approaches are extensions of the compiler correctness criterion as discussed in other work~\cite{patterson2019next700} or the introduction of opaque observations~\cite{vu2021reconciling} to reconcile compiler optimisations with security.
-Note that this work also presents secure compilers that are optimising, but contrary to the other~\cite{vu2021reconciling}, provides a formal account of these in the robust preservation framework.
-% Lastly, the authors of this paper have presented ongoing work~\cite{patrignani2023blame} on a weaker robust preservation criteria based on the concept blame.
-
-\paragraph*{Memory Safety Mechanisms}\label{subsec:relw:msmechs}
-
-Different mechanisms for enforcing memory safety exist that also consider the secure compilation domain, i.e., have an active attacker model.
-For example, the ``pointers as capabilities'' principle represents pointers as machine-level capabilities~\cite{korashy2021capableptrs}, which behave in a similar fashion to capabilities by means of linear typing~\cite{morrisett2005L3}.
-The approach of this paper also uses linear typing, but differs from $L^{3}$~\cite{morrisett2005L3} in the way that functions are not first-class.
-Moreover, this paper considers an active attacker, while the work on $L^{3}$ only discusses whole programs and, thus, has no active attacker model.
-The instrumentation to ensure memory safety that this paper presents is inspired by Softbounds~\cite{nagarakatte2009soft}.
-That work inserts bounds-checks in front of pointer-dereferences and, for this to work, inserts meta-data information on pointer creation.
-Softbounds also works in a more advanced setting with structured fields accesses and also introduces a table-lookup for pointers that are stored in memory.
-This paper only considers arrays of primitive data, i.e., there are no pointers to pointers or structures.
-Several other approaches to memory-safety exist in literature, specifically as compiler instrumentations~\cite{akritidis2009baggy,younan2010paricheck,jung2021pico,shankaranarayana2023tailcheck,dhumbumroong2020boundwarden,nam2019framer,zhou2023fatptrs}, hardware-extensions~\cite{kwon2013lowfat,saileshwar2022heapcheck,chen2023flexpointer,kim2023whistle}, or programming language extensions~\cite{elliott2018checkedc,li2022formalcheckedc,jim2002cyclone,elliott2015guilt,west2005cuckoo,weis2019fyr,benoit2019uniqueness}.
-What differentiates this work from them is that this work uses known, compiler-based approaches to ensure memory-safety as a means to investigate secure compiler compositions.
-This paper does not provide efficient memory-safety, but serves as a theoretical foundation for the secure compilation domain.
-
-To extend the languages in this paper with a less restricted form of pointer arithmetic, the region colouring memory safety monitor presented in earlier work~\cite{michael2023mswasm} can be used.
-The work presenting this monitor provides an approach for the robust preservation of memory safety compiling from C to WASM.
-However, they do not discuss composition of secure compilers but rather investigate an instance of a secure compiler.
-
-\paragraph*{Cryptographic Constant Time Mechanisms}\label{subsec:relw:cctmechs}
-
-The approach to preserving cryptographic constant time in this paper is high-level, where a programming language exposes a way to switch the semantics to a data (operand) independent timing mode.
-Since identifiers in $\Lscct$ are annotated with a secrecy tag, this approach is similar to others with information flow control.
-For example, Vale~\cite{bond2017vale} uses Dafny to ensure constant-time assembly code, while Jasmin~\cite{almeida2017jasmin} makes use of the Coq proof assistant to reject non-constant-time programs.
-CT-Wasm~\cite{watt2019ctwasm} enforces constant-timeness by means of a type system.
-Different to the approach of this paper, these approaches necessitate that the programmer writes \gls*{cct} code.
-An approach to allow programmers to write more high-level code is CryptOpt~\cite{kuepper2023cryptopt}, which generates efficient target-code by means of a randomised search.
-This paper abstracts over concrete mitigation strategies and simply assumes that there is a flag to switch to a cryptographic-constant time execution mode.
-This can be realised by employing the FaCT~\cite{cauligi2019fact} compiler, which translates common non-constant time code patterns to be constant-time, and the data (object) independent timing execution mode of modern processors.
-
-\paragraph*{Speculation Safety Mechanisms}\label{subsec:relw:ssmechs}
-
-This paper uses a taint-tracking mechanism inspired by existing work~\cite{guarnieri2018spectector,fabian2022automatic}.
-These taints are used to express absence of any speculative leaks in \gls*{ss}~\cite{guarnieri2018spectector}. 
-The semantics of $\Lspec$ hardcodes the kind of speculative leaks to just SPECTRE-PHT~\cite{kocher2019spectre}, but future work could use semantics composition~\cite{fabian2022automatic} to support more variants.
-Note that our framework composes compilers and not semantics.
-
-%This work has taken inspiration in the sematics of $\Lspec$ and in \Cref{def:trace:ss} from recent formalisations~\cite{guarnieri2018spectector,fabian2022automatic}.
-% \MPin{
-%  we take SS from exorcising.
-%  we do not take SNI from spectector because it is a hypersafety.
-%  other work defined similar props. 
-% }
-% A subsequent line of work~\cite{fabian2022automatic} shows how to modularly define and compose formal semantics that models different speculative execution variants. 
-% Recent work~\cite{fabian2024lift} demonstrates how to lift existing secure compilation proofs for security properties related to speculative execution to more powerful properties, i.e., properties defined for languages with a trace model that enables more vulnerabilities, such as combinations of different speculative execution exploits.
-% To this end, they define a lifting framework and demonstrate how a secure compilation proof for a weaker property can be lifted to a stronger one. 
-% Contrary to the compositionality framework presented in this paper, these works are only concerned with composing semantics of the respective programming languages, while in this paper, the secure compilers themselves are composed.
-
-\section{Conclusion\pages{$\sfrac{1}{2}$}}\label{sec:concl}
-This paper tackles the problem of understanding what kind of security properties a secure compiler preserves, when said compiler is the combination of compiler passes that preserve possibly different security properties.
-% 
-%For this, this paper first formalised security properties of interest and their composition.
-% 
-The paper proves that composing secure compilers that preserve certain properties results in a secure compiler that preserves the composition of these properties.
-% While the presented security property composition relied on monitors that check only trace-properties, the composition of secure compilers does not make any restriction towards the kind of property involved in the composition.
-% It is subject to future work to develop techniques to verify relational hyperproperties~\cite{abate2019jour}, while the composition of hyperproperties could be very similar to the composition of ordinary properties as presented in this paper, since hyperproperties can be checked with an automata-based model-checker~\cite{beutner23hyperltl}.
-% 
-Finally, this paper defines a multi-pass compiler and proves that it preserves \gls*{specms}.
-Crucially, this paper derives the security of the multi-pass compiler from the composition of the security properties preserved by its individual passes, which include security-preserving as well as optimisation passes.
-% For future work, it is interesting to look at more sophisticated optimisation passes that, e.g., reorder events that appear on traces.
-
-\newpage
-
-\bibliographystyle{IEEEtranS}
-\bibliography{main}
-
-%%
-%% If your work has an appendix, this is the place to put it.
-\appendix
-\subsection{Existential Image}\label{subsec:extimg}
-
-\begin{definition}[Existential Image]
-% [Existential and Universal Image]
-\label{def:existential:img}\label{def:tau}
-  \[
-    \tau_\sim\left(\src{\pi}\right) := 
-      \left\{ 
-        \trg{\trace} \mid \exists \src{\trace}\ldotp \src{\trace}\sim\trg{\trace}, \text{ and } \src{\trace}\in\src{\pi} 
-      \right\}
-  \]
-  %\[ 
-  %  \sigma_\sim\left(\trg{\pi}\right) := 
-  %    \left\{ 
-  %      \src{\trace} \mid \forall \trg{\trace}\ldotp \text{if }\src{\trace}\sim\trg{\trace}, \text{ then } \trg{\trace}\in\trg{\pi} 
-  %    \right\}
-  %\]
-\end{definition}
-
-\begin{definition}[Robust Preservation with $\tau_\sim$]\label{def:rtp:tau}
-  % $\;$\\
-  % \vspace{-1em}
-  % \begin{nscenter}
-  % Compiler $\cc{\src{L}}{\trg{L}}$ robustly preserves $\class$, 
-  {$\rtptau{\cc{\src{L}}{\trg{L}}}{\src{\class}}{\sim}$}
-  %, iff 
-  $\isdef$
-    {$\forall \src{\pi}\in\src{\class}, \src{p}\in\src{L},$} if {$\rsat{\src{\progvar}}{\src{\pi}}$}, then {$\rsat{\cc{\src{L}}{\trg{L}}\left(\src{p}\right)}{\tau_\sim\left(\src{\pi}\right)}$}.
-  % \end{nscenter}
-\end{definition}
-
-\begin{theorem}[Composition of Secure Compilers w.r.t. $\tau$]\label{thm:rtpsim:tau}
-  $\;$
-
-  If {$\rtptau{\cc{\src{L}}{\trg{L}}}{\src{\class[_{1}]}}{\sim_1}$}, {$\rtptau{\cc{\trg{L}}{\obj{L}}}{\tilde{\tau}_{\sim_1}\left(\src{\class[_2]}\right)}{\sim_2}$}, and {$\wfctau{\sim_1}{\src{\class[_2]}}$}, \\ then {$\rtptau{\cc{\src{L}}{\trg{L}}\circ\cc{\trg{L}}{\obj{L}}}{\src{\class[_{1}]}\cap\src{\class[_{2}]}}{\sim_1\bullet\sim_2}$}.
-\Coqed
-\end{theorem}
-
-\begin{corollary}[Swapping Secure Compiler Passes]\label{corr:swappable:tau}
-  $\;$
-
-  If {$\rtptau{\cc[_{1}]{\trg{L}}{\trg{L}}}{\src{\class[_{1}]}}{\sim_1}$ and $\rtptau{\cc[_{2}]{\trg{L}}{\trg{L}}}{\src{\class[_{2}]}}{\sim_2}$}, %
-  {$\wfctau{\sim_1}{\src{\class[_2]}}$ and $\wfctau{\sim_2}{\src{\class[_1]}}$}, %
-  and {$\tilde{\tau}_{\sim_1}\left(\src{\class[_2]}\right)=\src{\class[_2]}$ as well as $\tilde{\tau}_{\sim_2}\left(\src{\class[_1]}\right)=\src{\class[_1]}$},
-  then {$\rtptau{\cc[_{1}]{\trg{L}}{\trg{L}}\circ\cc[_{2}]{\trg{L}}{\trg{L}}}{\src{\class[_{1}]}\cap\src{\class[_{2}]}}{\sim_1\circ\sim_2}$ and $\rtptau{\cc[_{2}]{\trg{L}}{\trg{L}}\circ\cc[_{1}]{\trg{L}}{\trg{L}}}{\src{\class[_{2}]}\cap\src{\class[_{1}]}}{\sim_2\circ\sim_1}$}.
-\Coqed
-\end{corollary}
-
-\subsection{Secure Upper and Lower Composition}\label{sec:other-compos}
-Besides sequential composition, there are two other compositions, namely an {\em upper}, i.e., a compiler that takes multiple inputs and yields one output, and a {\em lower} composition, i.e., a compiler that takes one input and yields multiple outputs.
-We {define the upper composition $\cc{\src{L}+\obj{L}}{\trg{L}}$} as follows:
-Given a program \texttt{p}, its compiled counterpart is obtained by {plugging \texttt{p} into $\cc{\src{L}}{\trg{L}}$ if $\texttt{p}\in\src{L}$} or by {plugging \texttt{p} into $\cc{\obj{L}}{\trg{L}}$ if $\texttt{p}\in\obj{L}$}.
-\begin{definition}[Upper Composition]
-  \[
-    \text{{$\cc{\src{L}+\obj{L}}{\trg{L}}$}}\isdef
-  % Given \texttt{p}, yield
-  \lambda \texttt{p}\ldotp
-  \left\{\mbox{\begin{tabular}{c}
-    {{if $\texttt{p}\in\src{L}$, then $\cc{\src{L}}{\trg{L}}(\texttt{p})$}} \\
-    \mbox{{if $\texttt{p}\in\obj{L}$, then $\cc{\obj{L}}{\trg{L}}(\texttt{p})$}} \\
-  \end{tabular}}\right. 
-  \]
-\end{definition}
-
-Examples of this are present in industry:
-Consider the Java Virtual Machine bytecode $\trg{JVM BC}$, which is a popular target for programming language designers due to its high performance and relevance in industry.
-Compilers for several programming languages have it as their target language, some popular instances are $\src{Java}$ and $\obj{Kotlin}$.
-Technically speaking, they both compile to class files and $\obj{Kotlin}$ objects are considered to be the same as $\src{Java}$ objects at that point.
-Both languages can be used at the same time in one project~\cite{androidstudio}.
-A compiler that accepts both $\src{Java}$ and $\obj{Kotlin}$ code translating to the same target language or intermediate representation performs a kind of {\em upper} composition.
-Now, the following theorem tells us what happens if these are secure:
-Given {$\cc{\src{L}}{\trg{L}}$ robustly preserves $\class[_{1}]$} and {$\cc{\obj{L}}{\trg{L}}$ robustly preserves $\class[_{2}]$}, it follows that {their upper composition $\cc{\src{L}+\obj{L}}{\trg{L}}$ robustly preserves the intersection of classes $\class[_{1}]$ and $\class[_{2}]$}.
-
-\begin{theorem}[Upper Composition of Secure Compilers]\label{thm:urtp}
-  $\;$
-
-  If {$\rtp{\cc{\src{L}}{\trg{L}}}{\class[_{1}]}$} and {$\rtp{\cc{\obj{L}}{\trg{L}}}{\class[_{2}]}$}, then {$\rtp{\cc{\src{L}+\obj{L}}{\trg{L}}}{\class[_{1}]\cap\class[_{2}]}$}. %\Coqed
-\end{theorem}
-
-Dually, the {\em lower} composition is concerned about compilers that accept the same source but yield different target languages. %and the same theoretical results apply.
-{Define the lower composition $\cc{\src{L}}{\trg{L}+\obj{L}}$} as follows:
-Given a program $\src{p}$, its compiled counterpart is obtained by {plugging $\src{p}$ into $\cc{\src{L}}{\trg{L}}$} or by {plugging \texttt{p} into $\cc{\src{L}}{\obj{L}}$}, respectively, {based on the internal decision}.
-\begin{definition}[Lower Composition]
-  \[
-    \text{{$\cc{\src{L}}{\trg{L}+\obj{L}}$}}\isdef
-  % Given $L$ and $\src{p}$, yield
-  \lambda \src{p}, L\ldotp
-  \left\{\mbox{\begin{tabular}{c}
-    {{if $L=\trg{L}$, then} {$\cc{\src{L}}{\trg{L}}(\src{p})$}} \\
-    \mbox{{if $L=\obj{L}$, then} {$\cc{\src{L}}{\obj{L}}(\src{p})$}} \\
-  \end{tabular}}\right.\]
-\end{definition}
-
- Consider two compilers both accepting $\src{LLVM IR}$~\cite{lattner2004llvm} and one of them emits $\trg{x86\_64}$, while the other emits $\obj{ARMv8}$.
- It is intuitive that they are in some sense composed in the LLVM framework, but the decision of when to use one over the other is inherently {\em internal} to the formalisation effort of this kind of composition.
- For example, the user of this compiler provides an explicit flag that instructs to emit $\trg{x86\_64}$ or the framework itself detects the target platform via heuristics, such as supported instructions.
-
- The following theorem demonstrates what happens if the involved compilers are secure:
-Given {$\cc{\src{L}}{\trg{L}}$ robustly preserves $\class[_{1}]$} and {$\cc{\src{L}}{\obj{L}}$ robustly preserves $\class[_{2}]$}, it follows that {their lower composition $\cc{\src{L}}{\trg{L}+\obj{L}}$ robustly preserves the intersection of classes $\class[_{1}]$ and $\class[_{2}]$}.
-
-\begin{theorem}[Lower Composition of Secure Compilers]\label{thm:lrtp}
-  $\;$
-
-  If {$\rtp{\cc{\src{L}}{\trg{L}}}{\class[_{1}]}$} and {$\rtp{\cc{\obj{L}}{\trg{L}}}{\class[_{2}]}$}, then {$\rtp{\cc{\src{L}}{\trg{L}+\obj{L}}}{\class[_{1}]\cap\class[_{2}]}$}. % \Coqed
-\end{theorem}
-
-Either way, the theoretical results suggest that it is possible to always find a ``most-general'', secure compiler, given two secure compilers, that robustly preserves the least-upper bound of the classes involved in their compilation process.
-
-%Future work intends to use upper and lower composition as means to build a secure compiler incrementally
-%This follows current non-formal-methods based compiler engineering, such as LLVM~\cite{lattner2004llvm}, where languages such as \texttt{C++} and \texttt{Fortran} translate into LLVM IR and languages such as \texttt{x86} or \texttt{ARMv8} assembly variants translate out of LLVM IR.
-
-% no coq, because above definitions are non-formal. 
-%   How would you precisely define the lower compo compiler, given two compilers?
-
-%\section{Cryptographic Constant Time}\label{sec:cct}
-
-\end{document}
-\endinput
+\input{sections/bg/sat.tex}
+\input{sections/bg/comp.tex}
+\input{sections/bg/hprop.tex}

### 2025-02-05 icfp/sections/bg/comp.tex
Move Sections around

--- 2024-09-24 csf2/main.tex
@@
-% Fixing: Too many math alphabets used in version normal
-\newcommand\hmmax{0}
-\newcommand\bmmax{0}
-\documentclass[dvipsnames,conference]{IEEEtran}
-
-\usepackage[T1]{fontenc}
-\usepackage[scaled=.83]{beramono}
-
-\usepackage[colorinlistoftodos]{todonotes}
-\usepackage[inference]{semantic}
-\usepackage[switch]{lineno}
-\usepackage{halloweenmath}
-\usepackage{fontawesome5}
-\usepackage{listofitems}
-\usepackage{breakcites}
-\usepackage{glossaries}
-\usepackage{hyperref}
-\usepackage{cleveref}
-\usepackage{stmaryrd}
-\usepackage{marvosym}
-\usepackage{listings}
-\usepackage{amssymb}
-\usepackage{amsmath}
-\usepackage{nameref}
-\usepackage{amsthm}
-\usepackage{xspace}
-\usepackage{xfrac}
-\usepackage{tikz}
-\usepackage{soul}
-\usepackage{bm}
-\usepackage{paralist}
-
-\hypersetup{
-    colorlinks,
-    linkcolor={red!50!black},
-    citecolor={blue!50!black},
-    urlcolor={blue!80!black}
-}
-%\renewcommand\UrlFont{\color{blue}\rmfamily}
-%\newcommand{\url}[1]{\lstinline{#1}}
-
-\setul{0.95ex}{0.3ex}
-
-\input{cmds.tex}
-
-% allow page breaks inside multiline alignment displays
-\allowdisplaybreaks
-
-\Crefname{exampleenv}{Example}{Examples}
-
-\theoremstyle{definition}
-\newtheorem{exampleenv}{Example}[section]
-\newtheorem{lemma}{Lemma}[section]
-\newtheorem{theorem}{Theorem}[section]
-\newtheorem{corollary}{Corollary}[section]
-\newtheorem{definition}{Definition}[section]
-
-\loadglsentries{acronyms}
-\makeglossaries
-
-\AtBeginDocument{%
-  \providecommand\BibTeX{{%
-    \normalfont B\kern-0.5em{\scshape i\kern-0.25em b}\kern-0.8em\TeX}}}
-\makeatletter
-\newcommand{\linebreakand}{%
-  \end{@IEEEauthorhalign}
-  \hfill\mbox{}\par
-  \mbox{}\hfill\begin{@IEEEauthorhalign}
-}
-\makeatother
-
-\renewcommand\thelinenumber{\color{red}\arabic{linenumber}}
-\begin{document}
-\linenumbers
-
-\title{
-  Secure Composition 
-  \\ 
-  of Robust and Optimising Compilers
-  % \vspace{-1.5em}
-}
-
-% \author{
-% Matthis Kruse\orcidID{0000-0003-4062-9666}
-% \and
-% Michael Backes
-% \and
-% Marco Patrignani
-% }
-% \authorrunning{M. Kruse et al.}
-
-\maketitle
-
-\begin{abstract}
-% \MPpost{
-%   this abstract fails in setting up the robustness argument.
-%   it'll be a good exercise for you to take inspiration from this, but then write something as concise but with a more direct focus on robust properties
-% }
-% context
-To ensure that secure applications do not leak their secrets, they are required to uphold several security properties such as spatial and temporal memory safety, cryptographic constant time, as well as speculative safety.
-% need
-Existing work shows how to enforce these properties individually, in an architecture-independent way, by using secure compiler passes that each focus on an individual property.
-% task
-Unfortunately, given two secure compiler passes that each preserve a possibly different security property, it is unclear what kind of security property is preserved by the composition of those secure compiler passes.
-%there is no way to tell what kind of security property will the composition of those secure compilers preserve.
-
-% object
-This paper is the first to study what security properties are preserved across the composition of different secure compiler passes.
-% findings
-Starting from a general theory of property composition for security-relevant properties (such as the aforementioned ones), this paper formalises a theory of composition of secure compilers.
-Then, it showcases this theory on a secure multi-pass compiler that preserves the aforementioned security-relevant properties.
-% conclusion
-Crucially, this paper derives the security of the multi-pass compiler from the composition of the security properties preserved by its individual passes, which include security-preserving as well as optimisation passes.
-% 
-From an engineering perspective, this is the desirable approach to building secure compilers.
-\begin{center}\small\it
-	{This paper uses syntax highlighting accessible to both colourblind and black \& white readers.
-	% ~\cite{patrignani2020use}.
-% 	% Specifically, it makes use of a $\src{blue}$, $\src{sans\text{-}serif}$ font for a $\src{source}$,
-% 	% a $\trg{red}$, $\trg{bold}$ font for an $\trg{intermediate}$,
-% 	% and a $\obj{green}$, $\obj{teletype}$ font for a $\obj{target}$ language.	
-	For a better experience, please print or view in colours.
-	}
-\end{center}
-\end{abstract}
-
-\begin{IEEEkeywords}
-  memory-safety, cryptographic constant-time, secure compilation, robustness
-\end{IEEEkeywords}
-
-\section{Introduction\pages{4}}\label{sec:introduction}
-
-% \gls*{ms} is a security property that ensures that there are no out-of-bounds accesses as well as no reads or writes to uninitialised memory.
-% Programs that do not attain memory safety, can easily have their private data be leaked to attackers or be corrupted~\cite{lemay2021ccc}.
-% Because of this, \gls*{ms} is a requirement for programs to attain an even stronger security property, namely \gls*{cct}.
-% Programs that attain \gls*{cct} are secure w.r.t.
-timing attacks, and this is the gold standard for secure applications, including cryptographic ones, since timing attacks can expose secrets via a simple performance analysis~\cite{kocher1996timing}.
-% Timing attacks have become more and more relevant recently, especially given the era of cloud-computing~\cite{aviram2010cloudtime,kumar2019cloudsecsurvey}, where programs share the same hardware but are supposed to run in a sandboxed environment that does not leak secrets.
-% Unfortunately, several attacks within such a cloud environment have been demonstrated as viable strategies to extract secrets~\cite{mehmet2015getoff,flowers2022zeroday,atya2019catchme,venkatanathan2015placevul}.
-
-% Even though mitigations for such attacks exist~\cite{bond2017vale,almeida2017jasmin}, some compilers may optimise these mitigations away~\cite{barthe2018sec}.
-% This is of course not what programmers expect and to ensure such mitigations persist in the compiled code, one can use a {\em secure} compilers.
-% More generally, secure compilers should preserve security properties even when the program runs in a hostile context such as the cloud one, i.e., where it interacts with other possibly malicious programs.
-% An example of such secure compilers are those that ascribe to the theory of robust compilation~\cite{abate2019jour,abate2021extacc,patrignani2021rsc}, which takes this interaction between components and potentially malicious contexts into account.
-
-% Unfortunately, it is unclear if and how robust compilation supports compositionality, and compilation passes are typically developed as a standalone module to be plugged into a compilation pipeline and composed with other passes.
-% Different kinds of compilation passes exist, e.g., there are usually several compilation passes that perform optimisations~\cite{lattner2004llvm,googlev8,androidstudio,kuepper2023cryptopt,manjikian1997fusion,wegman1991ccp} or source-code instrumentations~\cite{nagarakatte2009soft,nagarakatte2010cets,akritidis2009baggy,dhumbumroong2020boundwarden,jung2021pico,nam2019framer,shankaranarayana2023tailcheck,younan2010paricheck,zhou2023fatptrs,bond2017vale,almeida2017jasmin,kuepper2023cryptopt,cauligi2019fact} that enforce that the compiled program fulfils a property of interest.
-% Without a framework for composing secure compiler passes, it is not possible to enable separation of concerns, e.g., to have a secure compilation pass that ensures \gls*{ms} that is developed independently of another secure pass for \gls*{cct} as suggested before and as showcased in \Cref{ex:strncpy}.
-% \MPpost{
-%   this intro does a poor job at introducing the robust compilation argument.
-%   perhaps it'd be best to start from robust properties, which motivates the use of robust compilers?
-%   idk
-%   as for the abstract, this is a good starting point for you to make a better intro
-% }
-
-\gls*{ms} is a security property obtained by composing \gls*{sms}, which ensures array accesses are all within bounds, and \gls*{tms}, which ensures pointers are only used when they are valid~\cite{azevedo2018meaningsofms,jim2002cyclone,necula2005ccured,nagarakatte2010cets,nagarakatte2009soft,akritidis2009baggy,michael2023mswasm}.
-\gls*{cct} is a security property that ensures sensitive data is not leaked via timing side-channels~\cite{kocher1996timing} and \gls*{ss} is a security property that enforces the same but under a speculative semantics~\cite{guarnieri2018spectector,fabian2022automatic} that captures speculative execution attacks such as spectre~\cite{kocher2019spectre}.
-Together, \gls*{sms}, \gls*{tms}, \gls*{cct} and \gls*{ss}, yield \gls*{specms}, which is the gold standard of security properties for secure applications.
-Programs attaining \gls*{specms} do not leak sensitive data either through erroneous memory accesses, nor through timing side-channels, even under speculative execution.
-\Cref{ex:strncpy} below discusses how these security properties can be enforced by compiler passes~\cite{bond2017vale,almeida2017jasmin}, to ensure programmers need not be aware of the architectural details of where their code will run.
-
-\begin{exampleenv}[strncpy]\label{ex:strncpy}
-Consider the following \texttt{C} function \texttt{strncpy} that copies a null-terminated string \texttt{src} into \texttt{dst} up to a length of \texttt{n} characters.
-This function is subject to a subtle \gls*{sms} vulnerability: the bounds check \texttt{i < n} should happen {\it before} the access to memory location \texttt{src[i]}: otherwise
-% If \texttt{n} is at least as large as the length of the array \texttt{src}, 
-the memory location past the last element is leaked to an attacker.
-\begin{lstlisting}[language=c,basicstyle=\small\ttfamily,morekeywords={size_t}]
-void strncpy(size_t n, char *dst, char *src) {
-  for(size_t i = 0; src[i] != '\0' && i < n; ++i) {
-    dst[i] = src[i];
-  }
-}
-\end{lstlisting}
-% Assume that the length of the array pointed by \texttt{dst} is at least \texttt{n}, and \texttt{src} is terminated by \texttt{'\textbackslash 0'}.
-% Even under these assumptions, there is a 
-To prevent this vulnerability, one can use a compilation pass that enforces \gls*{sms}, such as Softbounds~\cite{nagarakatte2009soft} or BaggyBounds~\cite{akritidis2009baggy}.
-The most na\"ive solution in this case is to insert bounds-checks in front of every access to memory.
-
-Because of timing attacks, addressing \gls*{sms} is not enough to make \texttt{strncpy} secure.
-In fact, the loop can terminate early, as soon as the string-terminating character \texttt{'\textbackslash 0'} is encountered, thus making program execution time proportional to the length of the array pointed by \texttt{src}, and violating the \gls*{cct} property.
-% Often, programmers rewrite programs like these to have loops that run with a fixed number of iterations.
-Also in this case there exist compiler passes that can rewrite such programs into \gls*{cct} ones~\cite{cauligi2019fact}.
- % so that they are no longer vulnerable to timing attacks.
-%
-Finally, even with these precautions, code is not run in isolation, so a malicious attacker could supply code that interacts with \texttt{strncpy} and trigger a violation of either \gls*{ms} or \gls*{cct} by calling \texttt{strncpy} with an argument for \texttt{src} that points to uninitialised memory.
-This would, in turn, trigger a series of reads from uninitialised memory, which is an immediate \gls*{ms} violation with devastating real-world consequences~\cite{uninit-0,uninit-1,uninit-2,uninit-3,uninit-4}.
-% An easy fix to sidestep this is to use non-nullable pointers as parameters instead, e.g., as presented in Cyclone~\cite{jim2002cyclone}.
-\end{exampleenv}
-
-  % last sentence is before the example, which talks about compiler passes enforcing props
-Whether or not compiler passes enforce certain security properties, it is important for a secure compiler to consider partial programs that interact with potentially malicious code, since the latter may lead to, e.g., memory-safety issues in the considered partial program.
-Robust compilers~\cite{abate2019jour} are a form of secure compilers that preserve security properties even in the presence of arbitrary attackers interacting with compiled code.
-Thus, robust compilers can be used to prevent vulnerabilities resulting from uninitialised memory (as well as many other ones), e.g., by targeting capability-based languages such as CHERI~\cite{woodruff2014CHERI}, Arm Morello~\cite{arm-morello}, or MSWasm~\cite{michael2023mswasm}, where the compiler relies on capabilities to check that pointers are always initialised.
-
-% To ensure these compiler-preserved properties still hold even in the presence of arbitrary attackers interacting with compiled code, one can use secure compilers in the form of robust compilers~\cite{abate2019jour,abate2021extacc,patrignani2021rsc}.
-
-Unfortunately, given secure compiler passes that each preserve a possibly different security property, there is no way to tell what kind of security property will the composition of those secure compilers preserve.
-% Different kinds of compilation passes exist, e.g., there are usually several compilation passes that perform optimisations~\cite{lattner2004llvm,googlev8,androidstudio,kuepper2023cryptopt,manjikian1997fusion,wegman1991ccp} or source-code instrumentations~\cite{nagarakatte2009soft,nagarakatte2010cets,akritidis2009baggy,dhumbumroong2020boundwarden,jung2021pico,nam2019framer,shankaranarayana2023tailcheck,younan2010paricheck,zhou2023fatptrs,bond2017vale,almeida2017jasmin,kuepper2023cryptopt,cauligi2019fact} that enforce that the compiled program fulfils a property of interest.
-Worse, without a framework for composing secure compiler passes, it is not possible to enable separation of concerns, e.g., to have a secure compilation pass that ensures \gls*{ms} that is developed independently of another secure pass for \gls*{cct}, that is developed independently of other passes, such as optimisation ones.
-
-This paper introduces a framework for reasoning about the composition of secure and optimising compiler passes %akin to those of \Cref{ex:strncpy}\MP{there aren't really passes there} 
-and it showcases the power of this framework by instantiating it on a multi-pass compilation chain.
-To this end, this paper first discusses how to compose security properties, such as \gls*{tms} and \gls*{sms} into \gls*{ms}, and then adds \gls*{cct} as well as \gls*{ss} into the mix to obtain \gls*{specms}.
-%Then, this paper defines compiler composition and formalises that given two passes that securely preserve two (possibly distinct) properties, their composition securely preserves the composition of those properties.
-The paper then defines several secure compiler passes, where each is either preserving a different security property from the list above
-% (\gls*{tms}, \gls*{sms}, \gls*{scct}, \gls*{ss}) 
-or performing a security-preserving optimisation.
-% , (e.g., applying \gls*{cf} or \gls*{dce}).
-Finally, this paper shows that composing these secure compiler passes into a multi-pass compilation chain results in the end-to-end (robust) preservation of \gls*{specms}.
-Crucially, this paper derives the security of the multi-pass compiler from the composition of the security properties preserved by its individual passes.
-% Crucially, this end-to-end security preservation is obtained by composing the security property preserved by each individual pass, as dictated by the framework of this paper.
-This result showcases how the framework allows the kind of formal security reasoning that compiler writers already want (and already do), obtaining precise, compositional security reasoning while providing minimal (and modular) proof effort.
-%
-In summary, this paper makes the following contributions:
-% \begin{itemize}[leftmargin=0em,nosep,label=$\blacktriangleright$]
-\begin{asparaitem}[$\blacktriangleright$]
-  \item %
-        %\Cref{sec:sequential} studies different forms of secure compiler compositions.
-        This paper takes the secure compilation framework of Abate et al.~\cite{abate2021extacc} and extends it for compositionality. (\Cref{sec:sequential}).
-        %Concretely, it investigates upper and lower compositions, as well as the sequential composition.
-        %For this, this paper studies compiler composition as well as compilers with multiple input languages or multiple output ones, as used in existing compilation chains.
-        %
-        This paper proves that starting from two compilers that preserve two (possibly distinct) properties, their composition preserves roughly the intersection of those properties.
-        % For example, a compiler that is secure w.r.t. \gls*{tms} can be composed with a compiler that is secure w.r.t. \gls*{sms}, resulting in a compiler that is secure w.r.t. \gls*{ms}.
-        % 
-        % From this, the paper proves that the order of composition of sequential compiler passes is irrelevant for security.
-        % % 
-        % This is crucial for reordering individual optimisation passes and thus generating secure and efficient code.
-        %
-        %Finally, this paper extends parts of the compositionality framework to support different trace-models between compilation passes.
-        % \MP{this isn't really true, is it?. the order matters for the relations' lemmas}
-        Then, this paper identifies which conditions make the composition of secure compilers meaningful from a security perspective, and which other conditions allow passes to be swapped without losing security meaning.
-
-  \item %
-        This paper presents a case-study with five programming languages showcasing the previous contribution (\Cref{sec:casestud:defs,sec:casestud:rtp}).
-        To this end, it presents a compilation chain consisting of six passes that ultimately preserve \gls*{specms} by means of composing four secure passes that \emph{individually} preserve \gls*{tms}, \gls*{sms}, \gls*{scct}---an enforceable version of \gls*{cct}---, and \gls*{ss}, and two passes which are well-known optimisations \gls*{dce} and \gls*{cf}.
-        %Furthermore, the chain includes two optimisation passes: one performs \gls*{dce} and the other \gls*{cf}.
-        The formalisation of this case study showcases the power of the presented framework: the divide-and-conquer approach to software engineering is a viable strategy even for the development of secure compilers.
-%        This demonstrates that the usage of the presented framework is both simpler and more expressive compared to one monolithic security proof of the whole compilation chain.
-
-  \item %
-  		This paper provides formal proofs of the security properties preserved by each of the presented passes. 
-  		% 
-  		Additionally this paper analyses each pass in terms of their compatibility with each other (\Cref{sec:formalities}), proving how does each pass fulfill the conditions that make their composition meaningful from a security standpoint.
-        % In doing so, the paper provides formal insights into some technicalities of the concrete proofs necessary as technical setup to reap the benefits of the presented compositionality framework.
-
-  \item The key contributions of this paper are formalised in the Coq proof assistant and the paper indicates this with \CoqSymbol.
-\end{asparaitem}
-% 
-This paper starts by introducing relevant notions of security properties and secure compilation (\Cref{sec:background}),
-and discusses related work (\Cref{sec:relwork}) before concluding (\Cref{sec:concl}).
-
-% \contribution{Technical Report} 
-The omitted formal details, lemmas and proofs, as well as the Coq formalisation are available as supplementary material.
-% \MKin{cite/link}
-
-\section{Background: Properties and Secure Compilers\pages{1}}\label{sec:background}
-
-To introduce the security argument of this paper, this section defines (security) properties, their satisfaction, and their robust satisfaction (i.e., satisfaction in the presence of an active attacker; \Cref{subsec:bg:tprop}).
-Then, borrowing from existing work~\cite{abate2019jour,abate2021extacc}, this section introduces secure compilers as compilers that preserve robust property satisfaction (\Cref{subsec:bg:rtp}).
-
-\subsection{Properties and (Robust) Satisfaction}\label{subsec:bg:tprop}
-
-This paper employs the security model where programs are written in a language whose semantics emits events $\event$.
-Events include security-relevant actions (e.g., reading from and writing to memory, as detailed in \Cref{sec:compprop}) and the unobservable event $\emptyevent$.
-As programs execute, their emitted events are concatenated in traces $\trace$, which serve as the description of the behaviour of a program.%
-\footnote{
-Throughout the paper, sequences are indicated with an overbar (i.e., $\trace$), empty sequences with $\hole{\cdot}$, and concatenation of sequences $\trace[_{1}],\trace[_{2}]$ as $\trace[_{1}]\cdot\trace[_{2}]$.
-Prepending element $\event$ to a sequence $\trace$ uses the same notation: $\event\cdot\trace$.
-}
-
-Properties $\pi$ are sets of traces of admissible program behaviours, ascribing what said property considers valid.
-% 
-The set of all properties can be partitioned into different {\em classes} ($\class$), i.e., safety, liveness, and neither safety nor liveness~\cite{clarkson2008hyper}, so a class is simply a set of properties.
-% 
-The compositionality framework (\Cref{sec:sequential}) presented in this paper considers arbitrary classes, while the case-study (\Cref{sec:casestud:rtp}) fixes them to concrete instances of safety properties, since it is decidable whether a trace satisfies a safety property with just a finite trace (i.e., a \emph{prefix}).
-As an example, consider the trace:
-$$\ev{Dealloc\ \loc}\cdot\ev{Read\ \loc\ 1729}\cdot\dots$$
-which describes the interaction with a memory where the deallocation of an address $\loc$ precedes a read (of some value $1729$) at that address in memory.
-% 
-This program behaviour is insecure w.r.t.
-a canonical notion of (temporal) memory safety dictating no use-after-frees of pointers~\cite{nagarakatte2010cets,azevedo2018meaningsofms}, because it reads from a memory location that was freed already.
-The previous finite trace prefix is enough to decide that the trace does not satisfy \gls*{tms} and there is no way to append events to this prefix which would result in the trace being admissible.
-
-In the following, the execution of a whole program $\wholeprogvar$ that terminates in state $\runtimetermvar$ according to the language semantics and produces trace $\trace$ is written as $\progstepto{\wholeprogvar}{\runtimetermvar}{\trace}$.
-With this, we defined property satisfaction as follows:
-\bul{a whole program $\wholeprogvar$ satisfies a property $\pi$} iff \iul{$\wholeprogvar$ yields a trace $\trace$} such that \oul{$\trace$ satisfies $\pi$}.
-
-\begin{definition}[Property Satisfaction]\label{def:propsat}
-$\;$
-
-  \begin{nscenter}
-    \bul{$\sat{\wholeprogvar}{\pi}$}
-    $\isdef$
-    if \iul{$\forall\runtimetermvar\ \trace,\progstepto{\wholeprogvar}{\runtimetermvar}{\trace}$},
-    then \oul{$\trace\in\pi$}.
-  \end{nscenter}
-\end{definition}
-
-Property satisfaction is defined on whole programs, i.e., programs without missing definitions.
-Thus, from a security perspective, \Cref{def:propsat} considers only a passive attacker model, where the attacker observes the execution and, e.g., retrieves secrets from that.
-To consider a stronger model similarly to what existing work does~\cite{abate2019jour,abate2021extacc,maffeis2008code-carrying,gordon2003authenticity,fournet2007authorization,bengtson2011refine,backes2014uniontyps,michael2023mswasm,swasey2017robust,sammler2019benefits}, we extend the concept of satisfaction with {\em robustness}.
-Robust satisfaction considers partial programs $\progvar$, i.e., components with missing imports, which cannot run until said imports are fulfilled.
-To remedy this, {\em linking} takes two partial programs $\progvar[_{1}],\progvar[_{2}]$ and produces a whole program $\wholeprogvar$, i.e., $\link{\progvar[_{1}]}{\progvar[_{2}]}=\wholeprogvar$.
-As typically done in works that consider the execution of partial programs~\cite{abate2019jour,devriese2018parametricity,patrignani2021rsc,korashy2021capableptrs,strydonck2019lincap,devriese2017modular,bowman2015noninterference,ahmed2011equivcps,patterson2017linkingtyps},
-this paper assumes that whole programs are the result of linking partial programs referred to as {\em context} ($\ctx$) and {\em component} ($\comp$).
-The context is an arbitrary program and thus has the role of an {\em attacker} that can interact with the component by means of any features the programming language has, and the component is what is security-relevant.
-%In this work, the semantics of the programming language is expected to differentiate between the component and the context.
-With this, \Thmref{def:propsat} can be extended as follows: for \bul{a component $\progvar$ to robustly satisfy a property $\pi$}, take an \iul{attacker context $\contextvar$ and link it with $\progvar$}, \oul{the resulting whole program must satisfy $\pi$}.
-
-\begin{definition}[Robust Satisfaction]\label{def:proprsat}
-  $\;$
-
-  \begin{nscenter}
-  \bul{$\rsat{\progvar}{\pi}$}
-  $\isdef$ \iul{$\forall \contextvar\ \wholeprogvar$, if $\link{\contextvar}{\progvar}=\wholeprogvar$}, then \oul{$\sat{\wholeprogvar}{\pi}$}.
-  \end{nscenter}
-\end{definition}
-
-% \begin{exampleenv}[Double Free in Bluetooth Subsystem]
-%   Consider \texttt{CVE-2021-3564}~\cite{doublefree-bluetooth}, one of many submissions for a double-free vulnerability.
-%   The vulnerability arises due to a race condition where the context-level function \texttt{hci\_cmd\_work} was not expected to behave maliciously, since it resides in the same source-code repository where the vulnerability occurs.
-%   Nevertheless, the component-level code of \texttt{hci\_dev\_do\_open} is linked with \texttt{hci\_cmd\_work} and does not atomically check whether a pointer has been freed already:
-%   Therefore, \texttt{hci\_dev\_do\_open} does not satisfy the no-double-frees property robustly, since there is an implementation for \texttt{hci\_cmd\_work} that leads to a violation of that property when linked with \texttt{hci\_dev\_do\_open}.
-% \end{exampleenv}
-
 \subsection{Secure Compilers}\label{subsec:bg:rtp}
@@
 %\smallskip%MP: the underline compacts the sections too much
-
-\section{Secure Composition}\label{sec:sequential}
-
-Notably, real-world compilation chains also perform a series of (sequential) passes whose main purpose is not necessarily to translate from one language to another, but to, e.g., optimise the code or enforce a certain property.
-Both examples can be seen in practice, e.g.,~\cite{nagarakatte2009soft,nagarakatte2010cets,akritidis2009baggy,wegman1991ccp,manjikian1997fusion} and many more.
-% \MPin{survey to cite?
-this is a bit informal MK: hard to find a survey on different compiler passes...
-I can't even find one just for compiler enforced memsafety (that looks good) or for compiler translations in general}.
-% 
-% Consider the compilation passes of LLVM~\cite{lattner2004llvm}, which are handled in the \texttt{PassManager} class.
-% By manipulating the internal state of a \texttt{PassManager} object, the schedule of compiler optimisation passes can be hardcoded.
-% That way, an order of optimisations can be found by hand that yields the best results.
-% \MP{what's the point of the previous 3 sentences?(commented)}
-Consider the following two LLVM optimisation passes: \gls*{cf}, which rewrites constant expressions to the constant they evaluate to, and \gls*{dce}, which removes dead code by rewriting conditional branches.
-The order in which \gls*{cf} and \gls*{dce} are performed influences the final result of the compilation (see \Cref{fig:cfdceex}).
-\begin{figure}[!ht]
-  \begin{center}
-    \begin{tikzpicture}[scale=0.62, every node/.style={transform shape}]
-      \node[minimum width=3cm,minimum height=2.5cm,draw=black,very thick,align=left] (progunopt) {\begin{lstlisting}[language=ml,basicstyle=\small\ttfamily]
-let a = true in
-if a then
-  print "a"
-else
-  print "b"
-\end{lstlisting}};
-      \node[minimum width=3cm,minimum height=2.5cm,draw=black,very thick,align=left,right=1.0 of progunopt] (progoptcf) {\begin{lstlisting}[language=ml,basicstyle=\large\ttfamily]
-if true then
-  print "a"
-else
-  print "b"
-\end{lstlisting}};
-      \node[minimum width=3cm,minimum height=2.5cm,draw=black,very thick,align=left,right=1.0 of progoptcf] (progoptcfdce) {\begin{lstlisting}[language=ml,basicstyle=\Large\ttfamily]
-print "a"
-\end{lstlisting}};% the following is a hack to get my syntax highlighting to work again \end{lstlisting}
-      % edges
-      \draw[->,very thick] (progunopt) edge[loop left,left] node {\gls*{dce}} (progunopt);
-      \draw[->,very thick] (progunopt) edge[sloped,above] node {\gls*{cf}} (progoptcf);
-      \draw[->,very thick] (progoptcf) edge[sloped,above] node {\gls*{dce}} (progoptcfdce);
-    \end{tikzpicture}
-  \end{center}
-  \caption{Example program where the level of optimisations differ for one pass of applying \gls*{cf} and \gls*{dce} in any order. %
-  Every edge is a compilation pass and the label on the edge states what the pass does, i.e., \gls*{cf} or \gls*{dce}. %
-  The source code in the nodes is a glorified compiler intermediate representation and the code gets more optimised towards the right hand side of the figure.}\label{fig:cfdceex}
-\end{figure}
-This {\em phase ordering problem} is well--known in literature and a practical solution is to simply perform a fixpoint iteration of the optimisation pipeline~\cite{click1995combining}.
-
-\smallskip
-
-To analyse the security of compilation passes and their interaction within a compilation pipeline, we rely on a few key notions: the definition of a trace relation, and the definition of when is a trace relation well-formed with respect to a class.
-%
-Consider traces $\src{\trace}$ and $\trg{\trace}$ as well as two trace relations $\sim_1$ and $\sim_2$. 
-% 
-The traces are related $\src{\trace}\sim_1\bullet\sim_2\trg{\trace}$ if there is another trace $\irl{\trace}$ such that $\src{\trace}\sim_1\irl{\trace}$ and $\irl{\trace}\sim_2\trg{\trace}$.
-% 
-\bul{A relation $\sim$ is well-formed w.r.t. a class of target-level properties $\trg{\class}$} iff \rul{the universal image preserves set membership}.
-\begin{definition}[Well-formedness of $\sim$ for a Class $\trg{\class}$]\label{def:wfc:sig:tracerel}
-  % $$
-
-  \begin{nscenter}
-  \noindent
-  \text{\bul{$\wfcsig{\sim}{\trg{\class}}$}} \isdef \text{\rul{$\forall \trg{\pi}\in\trg{\class}, \sigma_\sim(\trg{\pi})\in\sigma_\sim(\trg{\class})$}}
-  \end{nscenter}
-  % $$
-\end{definition}
-
-We can now state our main result: secure compilers in the robust compilation framework~\cite{abate2021extacc} compose {\em sequentially}. 
-%
-
-Let \bul{$\cc{\src{L}}{\trg{L}}$ robustly preserve the class $\sigma_{\sim_2}\left(\trg{\class[_1]}\right)$ under $\sim_1$} and let \rul{$\cc{\trg{L}}{\obj{L}}$ robustly preserve the class $\trg{\class[_2]}$ under $\sim_2$}.
-Then, \iul{when the cross-language relation $\sim_2$ is well-formed w.r.t.
-class $\trg{\class[_1]}$}, it follows that \oul{the composed compiler $\cc{\src{L}}{\trg{L}}\circ\cc{\trg{L}}{\obj{L}}$ robustly preserves the intersection of classes $\trg{\class[_1]}\cap\trg{\class[_2]}$ under $\sim_1\bullet\sim_2$}.
-% 
-%For space reasons, \Thmref{thm:urtp} and \Thmref{thm:lrtp} are not presented here, but can be found in the technical appendix ($\CoqSymbol$).
-% Note that it is not necessary to require $\sim_2$ to be well-formed w.r.t. $\src{\class[_1]}$, since the 
-% \MPin{why}
-\begin{theorem}[Composition of Secure Compilers w.r.t. $\sigma$]\label{thm:rtpsim:sig}
-  $\;$
-
-  If \bul{$\rtpsig{\cc{\src{L}}{\trg{L}}}{\tilde{\sigma}_{\sim_2}\left(\trg{\class[_{1}]}\right)}{\sim_1}$}, \rul{$\rtpsig{\cc{\trg{L}}{\obj{L}}}{\trg{\class[_2]}}{\sim_2}$}, and \iul{$\wfcsig{\sim_2}{\trg{\class[_1]}}$}, \\ then \oul{$\rtpsig{\cc{\src{L}}{\trg{L}}\circ\cc{\trg{L}}{\obj{L}}}{\trg{\class[_{1}]}\cap\trg{\class[_{2}]}}{\sim_1\bullet\sim_2}$}.
-\Coqed
-\end{theorem}
-
-Since the composition of secure compilers is again a secure compiler, the theorems generalise to a whole chain of $n$ secure compilers. 
-% 
-\Cref{thm:rtpsim:sig} can also be stated for the existential image $\tau_\sim\left(\src{\pi}\right)$, but in the interest of space that result has been moved to the appendix (\Cref{subsec:extimg}).
-% 
-Crucially, \Cref{thm:rtpsim:sig} also holds for \emph{classes of hyperproperties}, and thus, compilers that robustly preserve hyperproperties can be composed with each other as well as with compilers that robustly preserve properties.
-% 
-%Unfortunately, proofs of compilers that robustly preserve hyperproperties (including hypersafety) are scarce in the literature due to their complexity.
-%\MP{ so what ? }
-%  we don't discuss this further?
-%
-
-% It is furthermore possible to conflate the notion of classes and hyperproperties by intersecting all elements of a class of hyperproperties, to obtain one combined hyperproperty.
-% This idea has been discussed in other work~\cite{clarkson2008hyper}.
-%  MP: i think i stated the point of these 2 lines above
-
-If we take SecurePtrs and FaCT from \Cref{subsec:bg:rtp} and compose them according to \Cref{thm:rtpsim:sig}, we obtain a compiler that robustly preserves the intersection of safety properties and the \gls*{cct} hyperproperty.
-That is, for a source component that robustly satisfies any set of safety properties {\em and} \gls*{cct}, the compiled target component also robustly satisfies the same set of safety properties and \gls*{cct}.
-% The intersection simply uses the lifted version of the respective safety properties.
-%\MP{meaning?}
-
-Compiler engineers typically try to find an order of optimisations that yields well-optimised programs for either code size~\cite{cooper1999geneticphases} or performance~\cite{kulkarni2006exhaustivephase}.
-\Cref{corr:swappable} justifies that any such order of compilation passes is valid w.r.t. security, as long as the trace-relations have no effect on the respective classes.
-
-So, given two compilation passes \bul{$\cc[_{1}]{\trg{L}}{\trg{L}}$, $\cc[_{2}]{\trg{L}}{\trg{L}}$, both robustly preserving class $\trg{\class[_{1}]}$ or $\trg{\class[_{2}]}$, respectively}, \rul{their corresponding well-formed trace-relations}, \iul{and indifference of the classes with respect to these trace relations}, \oul{for any order of their composition, the composed compiler robustly preserves the intersection of $\trg{\class[_{1}]}$ and $\trg{\class[_{2}]}$}.
-% 
-% \MPin{
-% 	later on, say that optimisation passes remove elements from traces, so they should be fine w.r.t. the result with $\tau$.
-% 	also, yes, if $\tau$ adds we have a problem, if it removes we don't
-% \MKin{ Optimisations that remove elements from traces are not necessarily sound,
-%  e.g., remove all deallocations to "optimise" memory management by offloading
-%  the work to the OS.
-% }
-% }
-
-\begin{corollary}[Swapping Secure Compiler Passes]\label{corr:swappable}
-  $\;$
-
-  If \bul{$\rtpsig{\cc[_{1}]{\trg{L}}{\trg{L}}}{\trg{\class[_{1}]}}{\sim_1}$ and $\rtpsig{\cc[_{2}]{\trg{L}}{\trg{L}}}{\trg{\class[_{2}]}}{\sim_2}$}, %
-  \rul{$\wfcsig{\sim_1}{\trg{\class[_2]}}$ and $\wfcsig{\sim_2}{\trg{\class[_1]}}$}, %
-  and \iul{$\tilde{\sigma}_{\sim_2}\left(\trg{\class[_1]}\right)=\trg{\class[_1]}$ as well as $\tilde{\sigma}_{\sim_1}\left(\trg{\class[_2]}\right)=\trg{\class[_2]}$},
-  then \oul{$\rtpsig{\cc[_{1}]{\trg{L}}{\trg{L}}\circ\cc[_{2}]{\trg{L}}{\trg{L}}}{\trg{\class[_{1}]}\cap\trg{\class[_{2}]}}{\sim_1\circ\sim_2}$ and $\rtpsig{\cc[_{2}]{\trg{L}}{\trg{L}}\circ\cc[_{1}]{\trg{L}}{\trg{L}}}{\trg{\class[_{2}]}\cap\trg{\class[_{1}]}}{\sim_2\circ\sim_1}$}.
-\Coqed
-\end{corollary}
-
-Coming back to the example composing SecurePtrs with FaCT, it is likely the case that \Cref{corr:swappable} is not applicable.
-While first running SecurePtrs and then FaCT should be fine, the other direction has potential security hazards, since the SecurePtrs compiler does not account for cryptographic-constant time primitives, such as \texttt{ctselect}.
-
-\subsection{Secure Compiler Composition with Same Trace Models}
-When the cross-language trace relation is an equality, \Cref{thm:rtpsim:sig} collapses:
-Given \bul{$\cc{\src{L}}{\trg{L}}$ robustly preserves $\class[_{1}]$} and \rul{$\cc{\trg{L}}{\obj{L}}$ robustly preserves $\class[_{2}]$}, it follows that \oul{their sequential composition $\cc{\src{L}}{\trg{L}}\circ\cc{\trg{L}}{\obj{L}}$ robustly preserves the intersection of classes $\class[_{1}]$ and $\class[_{2}]$}.
-
-\begin{corollary}[Composition of Secure Compilers]\label{corr:rtp}
-  $\;$
-
-  If \bul{$\rtp{\cc{\src{L}}{\trg{L}}}{\class[_{1}]}$} and \rul{$\rtp{\cc{\trg{L}}{\obj{L}}}{\class[_{2}]}$}, then \oul{$\rtp{\cc{\src{L}}{\trg{L}}\circ\cc{\trg{L}}{\obj{L}}}{\class[_{1}]\cap\class[_{2}]}$}.
-\Coqed
-\end{corollary}
-
-\Cref{corr:rtp} provides an easy way to compose secure compilers without well-formedness of trace relations. 
-% 
-However, while \Cref{thm:rtpsim:sig} explicitly requires that $\sim_1$ is well-formed w.r.t. $\trg{\class[_2]}$, if $\sim_2$ is not well-formed w.r.t. $\trg{\class[_1]}$, care must be taken. 
-%Worst-case is that the resulting class $\sigma_{\sim_1\bullet\sim_2}\left(\trg{\class[_1]}\cap\trg{\class[_2]}\right)$ becomes empty, even though $\trg{\class[_1]}\cap\trg{\class[_2]}$ is not necessarily empty.
-This is further discussed in \Cref{subsec:compatsecpasses}.
-
-%If possible, optimisations are merged to enhance the analysis' result, since even when iterating to a fixpoint the final program may not be as optimised as it could be~\cite{click1995combining}.
-We can also obtain a specialised version of \Cref{corr:swappable}:
-%So, given two compilation passes \bul{$\cc[_{1}]{\trg{L}}{\trg{L}}$, $\cc[_{2}]{\trg{L}}{\trg{L}}$, both robustly preserving class $\class[_{1}]$ or $\class[_{2}]$, respectively}, \oul{for any order of their composition, the composed compiler robustly preserves the intersection of $\class[_{1}]$ and $\class[_{2}]$}.
-% 
-% \MPin{
-% 	later on, say that optimisation passes remove elements from traces, so they should be fine w.r.t. the result with $\tau$.
-% 	also, yes, if $\tau$ adds we have a problem, if it removes we don't
-% \MKin{ Optimisations that remove elements from traces are not necessarily sound,
-%  e.g., remove all deallocations to "optimise" memory management by offloading
-%  the work to the OS.
-% }
-% }
-
-\begin{corollary}[Swapping Secure Compiler Passes]\label{corr:swappable:one}
-  $\;$
-
-  If {$\rtp{\cc[_{1}]{\trg{L}}{\trg{L}}}{\class[_{1}]}$ and $\rtp{\cc[_{2}]{\trg{L}}{\trg{L}}}{\class[_{2}]}$}, then {$\rtp{\cc[_{1}]{\trg{L}}{\trg{L}}\circ\cc[_{2}]{\trg{L}}{\trg{L}}}{\class[_{1}]\cap\class[_{2}]}$ and $\rtp{\cc[_{2}]{\trg{L}}{\trg{L}}\circ\cc[_{1}]{\trg{L}}{\trg{L}}}{\class[_{2}]\cap\class[_{1}]}$}.
-\Coqed
-\end{corollary}
-
-\section{Security Properties Formalisation \& Composition\pages{1.5}}\label{sec:compprop}
-
-This section introduces trace properties of interest for this paper: \gls*{tms}, \gls*{sms}, \gls*{ms}, \gls*{scct}, and \gls*{ss}.
-These properties are of practical importance (as mentioned in \Cref{sec:introduction}) and also of interest in the case study presented later (\Cref{sec:casestud:defs,sec:casestud:rtp}). 
-This section presents all of them, despite the fact that they are inspired by existing work, in order to showcase all that is required for a formal proof of security for a realistic compilation toolchain.
-%\MP{ say something about this in the rebuttal: only the defs are borrowed, the universal images and all that work is new}
-%\MK{ universal images are not new. }
-% , none of the properties is novel, for self-containment of the paper, their definition is necessary for the case study.
-For each of the presented properties, the technical report defines corresponding monitors, which refine a given property and have a key role in the proofs of this paper. 
-% managed to prove 
-%  if tms-monitor --[trace]-->* ∅  
-%  then trace ∈ tms
-% but not
-%  if trace ∈ tms
-%  then tms-monitor --[trace]-->* ∅
-
-\subsection{A Trace Model for Memory Safety}\label{subsec:basic:memsafety:tracemodel}
-
-For simple memory safety composed of temporal and spatial memory safety, the trace model defines events ($\event[_{\mssafe}]$) as either the empty event ($\emptyevent$), a crash ($\lightning$), or a base-event ($\preevent[_{\mssafe}]$).
-
-\vspace{-1.0em}
-\[
-  \begin{array}{rcll}
-    \text{(Base-Events)}&\preevent[_{\mssafe}] &:=& \ev{Alloc \loc\ n} \mid \ev{Dealloc \loc} \mid \ev{Use \loc\ n} \\
-    \text{(Events)}&\event[_{\mssafe}] &:=& \preevent[_{\mssafe}] \mid \emptyevent \mid \lightning \\ 
-  \end{array}
-\]
-
-Base-events describe the actual kind of event that happened.
-For the basic memory-safety properties, these are three variants:
-First, the allocation event ($\ev{Alloc\ \loc\ n}$) that fires whenever a program claims $n$ cells of memory and stores them at address $\loc$, where addresses are assumed to be unique.
-Second, deallocation ($\ev{Dealloc\ \loc}$) announces that the object at location $\loc$ is freed.
-Third, an event to describe reads from and writes to the $n$-th memory cell from address $\loc$ ($\ev{Use\ \loc\ n}$).
-
-\subsubsection{Temporal Memory Safety}
-
-\gls*{tms}~\cite{nagarakatte2010cets} is a safety property that describes that an unallocated object must not be (re-)used.
-
-\begin{definition}[\glsfirst*{tms}]\label{def:trace:tmsdef}
-  % \footnotesize
-  $$
-  \tmssafe:=\left\{\trace[_{\mssafe}] \left| \begin{array}{rcl}
-    \ev{Alloc\ \loc\ n}&\le_{\trace[_{\mssafe}]}&\ev{Dealloc\ \loc} \\
-    \ev{Use\ \loc\ n}&\le_{\trace[_{\mssafe}]}&\ev{Dealloc\ \loc} \\
-    \text{at most one }\ev{Dealloc\ \loc}&\text{in}&\trace[_{\mssafe}] \\
-    \text{at most one }\ev{Alloc\ \loc\ n}&\text{in}&\trace[_{\mssafe}] \\
-  \end{array}\right.\right\}
-  $$
-\end{definition}
-Hereby, the notation $\event[_{1}]\le_{\trace}\event[_{2}]$ means that if $\event[_{1}]$ is in $\trace$ and if $\event[_{2}]$ is in $\trace$, then $\event[_{1}]$ appears before $\event[_{2}]$.
-
-\subsubsection{Spatial Memory Safety}
-
-\gls*{sms}~\cite{nagarakatte2009soft} is a safety property that prohibits out-of-bounds accesses.
-
-\begin{definition}[\glsfirst*{sms}]\label{def:trace:smsdef}
-% \small
-  % \begin{nscenter}
-
-  \noindent
-  \[
-  \smssafe:=\left\{\trace[_{\mssafe}] \left|\begin{array}{rcl}
-      \text{If }\ev{Alloc\ \loc\ n}\le_{\trace_{\mssafe}}\ev{Use\ \loc\ m}, \text{ then }m<n
-  \end{array}\right.\right\}
-  \]
-  % \end{nscenter}
-\end{definition}
-
-\subsubsection{Memory Safety}
-
-In spirit of earlier work~\cite{nagarakatte2009soft,nagarakatte2010cets,jim2002cyclone,necula2005ccured,michael2023mswasm}, full \gls*{ms} is the intersection of \Cref{def:trace:tmsdef,def:trace:smsdef}.
-
-\begin{definition}[\glsfirst*{ms}]\label{def:trace:msdef}
-  $
-  \mssafe:=\tmssafe \cap \smssafe
-  $
-\end{definition}
-
-Note that \Cref{def:trace:msdef} ignores data isolation, so there may still be memory-safety issues introduced by side-channels.
-
-\subsection{A Trace Model for Memory Safety with Constant Time}\label{subsec:scct:tracemodel}
-
-To express Constant Time, we extend the memory safety trace model with a {\em security tag} ($\securitytag{}$) that indicates whether events contains sensitive information ($\lock$) or not ($\unlock$).
-
-\vspace{-.5em}
-\[
-  \begin{array}{rrcl}
-    (\text{Base-Events}) & \preevent[_{\ctsafe}] &:=& \preevent[_{\mssafe}] \mid \ev{Branch\ }n \mid \ev{Binop\ }n\\
-    (\text{Security Tags}) & \securitytag{} &:=& \lock \mid \unlock\\ 
-    (\text{Events}) & \event[_{\ctsafe}] &:=& \preevent[_{\ctsafe}];\securitytag{} \mid \emptyevent \mid \lightning \\ 
-  \end{array}
-\]
-
-For cryptographic code, there is a general guideline that secrets must not be visible on a trace~\cite{ctguidelines}, i.e., secrets should not be marked as $\unlock$.
-In turn, an instruction whose timing is data-dependent must not have a secret as an operand.
-Typical operations with data-dependent timing are branches and certain binary operations, such as division.%
-\footnote{
-	This is architecture-dependent, but division is an operation that serves as a classic example for a data-dependent timing instruction~\cite[p.~755]{arm-refman}.
-}
-Both operations are represented in the trace model by extending the set of base-events with branches ($\ev{Branch\ n}$) and binary operations ($\ev{Binop\ n}$).
-
-\subsubsection{Strict Cryptographic Constant Time}
-
-\gls*{cct} is a hypersafety property~\cite{barthe2018sec} and, thus, difficult to check with monitors.
-This is because, intuitively, hypersafety properties can relate multiple execution traces with each other, but monitors work on a single execution.
-It is a common trick to sidestep this issue by means of overapproximation: this section defines the property \gls*{scct}, a stricter variant of \gls*{cct} (inspired by earlier work~\cite{almeida2017jasmin}) that enforces the policy that no secret appears on a trace.
-Programs that satisfy \gls*{scct} also satisfy \gls*{cct}, but programs that satisfy \gls*{cct} may not satisfy \gls*{scct}.
-
-\begin{definition}[\glsfirst*{scct}]\label{def:trace:scctdef}
-% \begin{nscenter}
-
-  \noindent\[
-  \scctsafe:=\left\{\trace[_{\ctsafe}] \left|\begin{array}{l}
-      \trace[_{\ctsafe}]=\hole{\cdot} \text{ or } \\\exists\trace[_{\ctsafe}'],\trace[_{\ctsafe}]=\preevent[_{\ctsafe}];\unlock\cdot\trace[_{\ctsafe}'] \wedge \trace[_{\ctsafe}']\in\scctsafe
-    \end{array}\right.\right\}
-  \]
-% \end{nscenter}
-\end{definition}
-
-\gls*{scct} may appear overly strict, since it seems that secrets must not occur on a trace (since $\securitytag{}$ is forced to be $\unlock$). 
-However, this is considered standard practice in terms of coding guidelines~\cite{ctguidelines}.
-Moreover, programs that have been compiled with FaCT~\cite{cauligi2019fact} and run with a ``data independent timing mode''~\cite{arm-refman,intel-refman} enabled do not leak secrets (see \Cref{ex:lscct}).
-
-\subsubsection{\gls*{ms}, Strict Cryptographic Constant Time}\label{sec:msscct-rel}
-
-The combination of \gls*{ms} and \gls*{scct} is the intersection of these properties, \gls*{msscct}.
-However, \gls*{ms} uses a different trace model than \gls*{scct}, so intersecting them would trivially yield the empty set. 
-To remedy this issue, we introduce $\sim_{\ctsafe}: \preevent[_{\ctsafe}] \times \preevent[_{\mssafe}] $, a cross-language trace relation (whose key cases are presented below), that we use to intuitively unify the trace model in which the two properties are expressed:
-\[
-  \typerulenolabel{mscctrel:drop:tag}{}{\preevent[_{\mssafe}]\sim_{\ctsafe}\preevent[_{\mssafe}];\unlock{}}
-  \typerulenolabel{mscctrel:drop:crash}{}{\lightning\sim_{\ctsafe}\lightning}
-\]
-\[
-  \typerulenolabel{mscctrel:drop:branch}{}{\emptyevent\sim_{\ctsafe}\ev{Branch n};\securitytag{}}
-  \typerulenolabel{mscctrel:drop:binop}{}{\emptyevent\sim_{\ctsafe}\ev{Binop n};\securitytag{}}
-\]
-
-Essentially, $\sim_{\ctsafe}$ ignores both the new $\ev{Branch\ n}$ and $\ev{Binop\ n}$ base-events as it relates security-insensitive actions ($\unlock$) to their equivalent counterparts.
-Thus, \gls*{ms} traces trivially satisfy \gls*{scct}.
-% 
-The above relation is extended point-wise to traces, skipping the empty event $\emptyevent$ on either side, and it is now possible to define \gls*{msscct} using the universal image:
-
-\begin{definition}[\gls*{ms} and \gls*{scct}]\label{def:trace:msscctdef}
-  $
-  \msscctsafe:=\mssafe\cap\sigma_{\sim_{\ctsafe}}\left(\scctsafe\right)
-  $
-\end{definition}
-
-\subsection{Extending the Trace Model with Speculation}\label{subsec:msctss:tracemodel}
-
-So far, the considered trace models do not let us express speculative execution attacks such as Spectre~\cite{kocher2019spectre}. 
-For this, we extend the earlier trace model (see \Cref{subsec:scct:tracemodel}) so that the security tags ($\securitytag{}$) carry additional information about the kind of private data leakage, i.e., the type of speculative leak.
-Moreover, we add base-events signalling the beginning of a speculative execution ($\ev{Spec}$), a barrier ($\ev{Barrier}$) that signals that any speculative execution may not go past it, as well as a rollback event ($\ev{Rlb}$), which signals that execution resumes to where speculation started.
-
-\vspace{-1em}
-{
-\[
-  \begin{array}{rrcl}
-    (\text{Base-Events}) & \preevent[_{\mathghost}] &:=& \preevent[_{\ctsafe}] \\
-    (\text{Spectre Variants}) & vX &:=& \operatorname{NONE} \mid \operatorname{PHT} \\
-    (\text{Security Tags}) & \securitytag{} &:=& \lock_{vX} \mid \unlock\\ 
-    (\text{Events}) & \event[_{\mathghost}] &:=& \preevent[_{\mathghost}];\securitytag{} \mid \emptyevent \mid \lightning \mid \ev{Spec} \mid \ev{Rlb} \mid \ev{Barrier}\\ 
-  \end{array}
-\]
-}
-
-Even though the considered Spectre variants are just SPECTRE-PHT~\cite{kocher2019spectre}, NONE just describes secret data as in \gls*{scct} (see \Cref{subsec:scct:tracemodel}), the trace model is general enough to allow for potential future extension with different variants~\cite{kocher2019spectre,maisuradze2018ret2spec,horn2019zero}.
-%For sake of readability, this paper just uses the notation $\lock$ in place of $\lock_{\text{PHT}}$.
-
-\subsubsection{Speculative Safety}
-
-\gls*{ss}~\cite{patrignani2021exorcising}, similar to \gls*{scct}, is a sound overapproximation of a variant of noninterference.
-
-\begin{definition}[\glsfirst*{ss}]\label{def:trace:ss}
-  \noindent
-
-  \begin{nscenter}
-  $
-    \sssafe:=\left\{\trace[_{\mathghost}] \left|\begin{array}{l}
-      \trace[_{\mathghost}]=\hole{\cdot} \text{ or } \exists\trace[_{\mathghost}'].\\
-      \left(\trace[_{\mathghost}]=\preevent[_{\mathghost}];\unlock\cdot\trace[_{\mathghost}'] \text{or }\trace[_{\mathghost}]=\preevent[_{\mathghost}];\lock_{\text{NONE}}\cdot\trace[_{\mathghost}']\right)\\
-      \text{and }\ \trace[_{\mathghost}']\in\sssafe
-                                 \end{array}\right.\right\}
-  $ 
-  \end{nscenter}
-\end{definition}
-The technical setup so far leads to the above definition, where only locks annotated with $\text{SPECTRE-PHT}$ are disallowed to occur on the trace.
-That way, programs attaining \gls*{ss} do not necessarily attain \gls*{scct}.
-
-\subsubsection{Speculation Memory Safety}\label{sec:spec-ms-rel}
-
-As before, we need to relate the different trace models with each other, so that the memory safety property without speculation can be lifted to speculation. 
-To this end, let $\sim_{\mathghost}: \preevent[_{\mathghost}] \times \preevent[_{\ctsafe}]$ be a cross-language trace relation whose key cases are below.
-The intuition is that \gls*{ss} is trivially satisfied in \gls*{scct}, since speculation is inexpressible there, which amounts to dropping events $\ev{Spec}$, $\ev{Rlb}$, or $\ev{Barrier}$, as well as all base events tagged with $\lock_{\text{PHT}}$.
-
-\[
-  \typerulenolabel{specrel:drop:tag}{}{\preevent[_{\ctsafe}];\lock\sim_{\mathghost}\preevent[_{\ctsafe}];\lock_{\text{NONE}}}
-  \typerulenolabel{specrel:drop:spectag}{}{\emptyevent\sim_{\mathghost}\preevent[_{\mathghost}];\lock_{\text{PHT}}}
-\]
-\[
-  \typerulenolabel{specrel:drop:pub}{}{\preevent[_{\ctsafe}];\unlock\sim_{\mathghost}\preevent[_{\ctsafe}];\unlock}
-  \typerulenolabel{specrel:drop:crash}{}{\lightning\sim_{\mathghost}\lightning}
-\]
-\[
-  \typerulenolabel{specrel:drop:spec}{}{\emptyevent\sim_{\mathghost}\ev{Spec}}
-  \typerulenolabel{specrel:drop:rlb}{}{\emptyevent\sim_{\mathghost}\ev{Rlb}}
-  \typerulenolabel{specrel:drop:barrier}{}{\emptyevent\sim_{\mathghost}\ev{Barrier}}
-\]
-
-We conclude by defining the ultimate property of interest for secure compilers: \gls*{specms}.
-\begin{definition}[\gls*{specms}]\label{def:trace:specmsdef}
-  $
-  \specmssafe := \msscctsafe\cap\sigma_{\sim_{\mathghost}}\left(\sssafe\right)
-  $
-\end{definition}
-
-\section{Case Study: Language Formalisations\pages{2}}\label{sec:casestud:defs}
-
-This section defines programming languages $\Ltms$, $\Ltrg$, $\Lms$, $\Lscct$, and $\Lspec$, all of which share common elements (presented in \Cref{subsec:cs:defs}).
-$\Ltms$ is the only statically-typed language, and it exhibits the property that all well-typed programs are \gls*{tms} (\Cref{subsec:ltms}).
-However, not all $\Ltms$ programs are \gls*{sms}.
-That is, there are well-typed $\Ltms$ programs that perform an out-of-bounds access.
-Language $\Ltrg$ is untyped and does not provide any guarantees with regards to \gls*{ms} (\Cref{subsec:lsms}).
-$\Lms$ is exactly the same language as $\Ltrg$, but this paper still distinguishes the two for sake of readability (\Cref{subsec:lms}).
-All three languages --- so $\Ltms$, $\Ltrg$, and $\Lms$ --- assume \gls*{scct} to hold, since this is -- in an ideal world -- what the programmer would expect, too: it is the job of the compiler to preserve and (potentially) enforce \gls*{scct} security~\cite{cauligi2019fact,nagarakatte2010cets,nagarakatte2009soft,akritidis2009baggy}.
-% This paper considers \gls*{cct} as an architecture-dependent property\MPin{can we justify this}, so
-
-Such consideration is also backed up by architecture providing a data (operand) independent timing mode, such as processors by Arm~\cite[p.~543]{arm-refman} and Intel~\cite[p.~80]{intel-refman}.
-%In spirit of this, all languages besides $\Lscct$ are trivially \gls*{scct}. 
-This kind of processor feature is modelled in language $\Lscct$ (\Cref{subsec:lscct}), where programs have access to a ``\gls*{cct}-mode'' and can change the leakage of emitted events according to the value of this mode (either $\obj{ON}$ or $\obj{OFF}$).
-
-Finally, modern processors also employ speculative execution to achieve speedups---and unfortunately generate Spectre attacks~\cite{kocher2019spectre}---and this is the extension of $\Lspec$ (\Cref{subsec:lspec}).
- % extends $\Lscct$ to model speculative execution.
-%
-Thus, all previous languages trivially satisfy \gls*{ss}, since they do not support speculative execution at all.
-% All languages share large parts of syntactic elements, which are presented in the next subsection (\Cref{subsec:cs:defs}).
-% Subsequent subsections extend on this accordingly.
-
-\subsection{Shared Language Definitions}\label{subsec:cs:defs}
-
-All presented programming languages share a common fragment which is partially presented here and in full detail in the technical report. 
-%\MP{put omitted stuff in appendix + ref there (and add textual explanation)}
-
-% \vspace{-.5em}
-{
-  \renewcommand{\src}[1]{\mathsf{#1}}
- \small
-\[
-  \begin{array}{rrcl}
-    \text{(Base-Events)} & \src{\preevent} &:=& \src{\ev{Alloc \loc\ n}} \mid \src{\ev{Dealloc \loc}} \mid \src{\ev{Get \loc\ n}} \mid \src{\ev{Set \loc\ n}} \\
-    \text{(Control Tags)} & \src{\sandboxtag{}} &:=& \src{\ctx} \mid \src{\comp} \\
-    \text{(Events)} & \src{\event} &:=& \src{\preevent;\sandboxtag{}} \mid \src{\emptyevent} \mid \src{\lightning} \\ 
-    \text{(Values)} & \src{v} &:=& \src{n} \\
-    \text{(Expressions)} & \src{\expr} &:=& \src{x} \mid \src{n} \mid \src{\lbinop{\expr[_1]}{\expr[_2]}} \mid \src{\lifz{\expr[_1]}{\expr[_2]}{\expr[_3]}} \\ 
-                         &&&\mid \src{\llet{x}{\expr[_1]}{\expr[_2]}} \mid \src{\llet{x}{new\ \expr[_1]}{\expr[_2]}} \\
-                         &&&\mid \src{\ldelete{x}} \mid \src{\lget{x}{\expr}} \mid \src{\lset{x}{\expr[_1]}{\expr[]}} \mid \src{stuck}\\
-                         &&&\mid \src{\lcall{f}{\expr}} \mid \src{\lreturn{\expr}} \\
-    \text{(Functions)} & \src{F} &:=& (\src{x};\src{\expr}) \\
-    \text{(Libraries)} &&&\hspace{-3.8em} \src{\library_{\ctx}},\src{\library_{\comp}} : \src{Vars} \to \src{F} 
-    % \hspace{2em}
-    %  : \src{Vars} \to \src{F} 
-     \\
-    \text{(Heaps)} & \src{H} &:=& \src{\hole{\cdot}} \mid \src{n},\src{H}
-    % \hspace{0.5cm}    
-    \\
-    \text{(Pointer Maps)} & \src{\Delta} && \text{ omitted for simplicity}
-    \\
-    \text{(Memory States)} & \src{\memstate} &:=& \left(\src{H^{\ctx}};\src{H^{\comp}};\src{\Delta}\right)\\
-    \text{(Control States)} & \src{\cfstate} & & \text{ omitted for simplicity} 
-    \\
-    \text{(States)} &\src{\Omega} &:=& \left(\src{\cfstate};\src{\sandboxtag{}};\src{\memstate}\right)\\
-  \end{array}
-\]
-}
-{
-  \renewcommand{\src}[1]{\mathsf{#1}}
-The trace models of all languages are similar to those presented earlier (\Cref{sec:compprop}).
-One technical detail is the addition of a control tag, indicating who is to blame for an emitted action: context ($\ctx$) or component ($\comp$).
-This is a standard technique in secure compilation in order to rule out irrelevant context-level events. 
-For example, a context immediately deallocating an allocated memory region twice trivially violates memory safety, but the main interest in secure compilation is to preserve component-level security, and thus component events.
-Even though this tagging could be used for blame preservation~\cite{patrignani2023blame}, this is beyond the focus of this paper.
-Another key difference is that memory accesses are now explicitly modelled as reads ($\src{\ev{Read\ \loc\ n}}$) and writes ($\src{\ev{Write\ \loc\ n}}$) instead of just uses.
-
-All languages have at least numbers as values ($\src{v}$) and second class functions ($\src{F}$).
-Functions are modelled as pairs containing the name of one argument and the body of the function.
-Bodies are just ordinary expressions, which can be simple binary operations ($\src{\lbinop{\expr[_1]}{\expr[_2]}}$), conditionals ($\src{\lifz{\expr[_1]}{\expr[_2]}{\expr[_3]}}$), function calls ($\src{\lcall{f}{e}}$) and returns ($\src{\lreturn{\expr}}$), as well as C-like memory operations.
-Programs have sets of pre-determined functions called libraries and they are marked as being part of some component ($\src{\library_{\comp}}$) or context ($\src{\library_{\ctx}}$).
-
-For the operational semantics, the runtime state ($\src{\Omega}$) is a triple consisting of a control-flow state ($\src{\cfstate}$), a control tag ($\src{\sandboxtag{}}$), and a memory state ($\src{\memstate}$).
-The latter carries information about pointers that are kept ``alive'' in pointer maps ($\src{\Delta}$), so that the semantics does not get stuck when encountering, e.g., a double-free.
-The memory state also carries two heaps to model sandboxing between a context ($\src{H^{\ctx}}$) and a component ($\src{H^{\comp}}$).
-
-}
-
-\subsection{$\src{L_{\tmssafe}}$: A Temporal but Not Spatial Memory Safe Language}\label{subsec:ltms}
-
-$\src{L_{\tmssafe}}$ is the only statically-typed language in this case study and restricts functions ($\src{F}$) to the typing signature $\src{\natt\to\natt}$. 
-The type system of $\src{L_{\tmssafe}}$ is inspired by $L^{3}$~\cite{morrisett2005L3,scherer2018fabulous} and enforces that every well-typed $\src{L_{\tmssafe}}$ program satisfies \gls*{tms}.
- % (\Cref{thm:wt:tms}).
-
-\begin{theorem}[$\src{L_{\tmssafe}}$-programs are \gls*{tms}]\label{thm:wt:tms}
-$\rsat{\src{\library_{\comp}}}{\tmssafe}$ \Coqed
-  % $\;$ 
-% 
-  % \begin{nscenter}
-  %   \hfill For any component $\src{\library_{\comp}}$, $\rsat{\src{\library_{\comp}}}{\tmssafe}$ \Coqed
-  % \end{nscenter}
-\end{theorem}
-
-\subsection{$\Ltrg$: A Memory-Unsafe Language}\label{subsec:lsms}
-
-$\Ltrg$ extends the syntax presented earlier (\Cref{subsec:cs:defs}) with dynamic typechecks ($\trg{\lhast{\expr}{\type}}$), evaluating to $\trg{0}$ if the type matches with the shape of $\trg{\expr}$ and $\trg{1}$ otherwise.
-%The pointers carry poison tags, which signify a pointer being allocated ($\trg{\poisoned}$) or freed ($\trg{\poisonless}$). 
-Furthermore, the syntax of $\Ltrg$ is extended with a way to inspect whether a pointer is freed ($\trg{\lispoisoned{x}}$), evaluating to $\trg{0}$ if it is freed and to $\trg{1}$ otherwise.
-Functions may receive arguments that are not $\trg{\natt}$, values are extended with tuples, and expressions are also extended with pair projections.
-
-% \vspace{-1em}
-\[
-  \begin{array}{rrcl}
-    \text{(Values)} & \trg{v} &:=& \dots \mid \trg{\lpair{v_1}{v_2}} \\
-    \text{(Expressions)} & \trg{\expr} &:=& \dots \mid \trg{\lhast{\expr}{\tau}} \mid \trg{\lispoisoned{x}}
-  \end{array}
-\]
-
-No changes are done to the trace model, but $\Ltrg$ has no static typing, making double-free code patterns possible.
-
-\subsection{$\Lms$: Another Memory-Unsafe Language}\label{subsec:lms}
-$\Lms$ is exactly equal to $\Ltrg$ (\Cref{subsec:lsms}), but used to emphasize that this is code after applying $\cc{\Ltrg}{\Lms}$ (\Cref{subsec:cs:ms}).
-
-\subsection{$\Lscct$: A Memory-Unsafe Language with a Constant Time Mode}\label{subsec:lscct}
-
-$\Lscct$ extends $\Lms$ (\Cref{subsec:lms}) with a ,,constant-time mode''. 
-The activation of the mode can be checked ($\obj{\lrddoit{x}{\expr}}$), changed ($\obj{\lwrdoit{D}}$), and is stored in program states ($\obj{\Omega}$).
-At the beginning of program execution of $\Lscct$ programs, the mode is turned off ($\obj{OFF}$).
-If the mode is enabled (i.e., set to $\obj{ON}$), the intuition is that no secrets are leaked. 
-This models the real-world\footnote{As present in Intel~\cite[p.80]{intel-refman} and ARM~\cite[p.~543]{arm-refman} processors.} data-independent timing mode as well as the result of compiling a program with FaCT~\cite{cauligi2019fact}.
-For example, FaCT rewrite code that branches to use a constant-time selection primitive.
-To not obfuscate our formalisation unnecessarily by duplicating syntax, we simply added the ,,constant-time mode'' to $\Lscct$.
-
-% $\Lscct$ extends $\Lms$ (\Cref{subsec:lms}) with a way to read from ($\obj{\lrddoit{x}{\expr}}$) and to write to ($\obj{\lwrdoit{D}}$) a model-specific register that controls a constant-time mode. 
-% This language is a combination of the effect of the FaCT compiler~\cite{cauligi2019fact} as well as a data-independent timing mode, a feature that is present in both Arm~\cite[p.~543]{arm-refman} and Intel~\cite[p.~80]{intel-refman} processors.
-% \MP{ can we be more specific about the fact pass? }
-% To this end, states ($\obj{\Omega}$) are extended with the value of the register ($\obj{D}$), which is initially set to be not active ($\obj{OFF}$).
-% If the register is marked active ($\obj{ON}$), the intuition is that no secrets can appear on specification traces.
-% If the register is marked inactive, secrets may appear on traces.
-% The state of the register can be read and written to, e.g., \Cref{tr:e-wrdoit-off}.
-% For the other languages seen earlier, the mode is intuitively always-on, i.e., the mode of execution always uses constant-time mode.
-
-% \vspace{-.5em}
-\[
-  \begin{array}{rrcl}
-    \text{(Mode Values)} & \obj{D} &:=& \obj{ON} \mid \obj{OFF} \\
-    \text{(Security Tags)} & \obj{\securitytag{}} &:=& \obj{\lock} \mid \obj{\unlock} \\
-    \text{(Base-Events)} & \obj{\preevent} &:=& \dots \mid \obj{iGet\ \loc\ v} \mid \obj{iSet\ \loc\ v} \\
-                       &&& \mid\ \obj{Binop\ n} \mid \obj{Branch\ n} \\
-    \text{(Events)} & \obj{\event} &:=& \obj{\preevent;\sandboxtag{};\securitytag{}} \mid \obj{\emptyevent} \mid \obj{\lightning} \\
-    \text{(Expressions)} & \obj{\expr} &:=& \dots \mid \obj{\lrddoit{x}{\expr}} \mid \obj{\lwrdoit{D}} \\
-                         &&&\mid\ \obj{\llet{x^{\securitytag{}}}{\expr[_1]}{\expr[_2]}}\\
-    \text{(States)} & \obj{\Omega} & := & \left(\obj{\cfstate};\obj{\sandboxtag{}};\obj{D};\obj{\memstate}\right)\\
-  \end{array}
-\]
-\begin{center}
-\newcommand{\expreval}[5]{{#1}\triangleright\xspace {#2}\xrightarrow{#5}\ {#3}\triangleright\xspace {#4}\xspace}
-\newcommand{\exprevalo}[5]{\expreval{\obj{#1}}{\obj{#2}}{\obj{#3}}{\obj{#4}}{\obj{#5}}}
-  \typerule{$e-\obj{wrdoit}-\text{off}$}{
-  }{
-    \pstepto{\rtt{\obj{\cfstate;\sandboxtag{};D;\memstate}}{\obj{\lwrdoit{OFF}}}}{\rtt{\obj{\cfstate;\sandboxtag{};OFF;\memstate}}{\obj{0^{\securitytag{}}}}}{\obj{\emptyevent}}
-  }{e-wrdoit-off}
-
-  \typerule{$e-\obj{get}-\in-$noleak}{
-    \obj{\memstate}=\obj{H^{\ctx};H^{\comp};\Delta_1,x\mapsto(\loc;\sandboxtag{};\poison),\Delta_2} &
-    \obj{\loc}+\obj{n}\in\text{dom }\obj{H^{\sandboxtag{}}} \\
-    \obj{\securitytag{''}}=\obj{\securitytag{}\sqcap\securitytag{'}} &
-    \obj{\event} = \obj{{iGet}\ \loc\ n;\sandboxtag{};\securitytag{''}}
-  }{
-    \exprevalo{
-    	\cfstate;\sandboxtag{'};ON;\memstate
-    }{
-      	x^{\securitytag[n^{\securitytag{'}}] }
-    }{
-    	\cfstate;\sandboxtag{'};ON;\memstate
-    }{
-    	(H^{\sandboxtag{}}(\loc+n))^{\securitytag{''}}
-    }{
-      \event
-    }
-  }{to-e-get-in-noleak}
-\end{center}
-
-The language also adds user annotations $\obj{\securitytag{}}$ for the secrecy of variables, which can be either private (high secrecy) $\obj{\lock}$ or public (low secrecy) $\obj{\unlock}$.
-Security tags $\obj{\securitytag{}}$ are arranged in the usual secrecy lattice~\cite{zdancewicphd}, where $\obj{\unlock}\sqsubseteq\obj{\lock}$.
-% $\obj{\lock}\sqsubseteq\obj{\securitytag{}}$ and $\obj{\securitytag{}}\sqsubseteq\obj{\unlock}$.
-
-Memory accesses to secret data need to be present to reason about memory safety, even when execution is in constant-time mode, e.g, \Cref{tr:to-e-get-in-noleak}.
-$\Lscct$ extends base-events with $\obj{\ev{iGet}\ \loc\ \valueexpr}$ and $\obj{\ev{iSet}\ \loc\ \valueexpr}$ (for data $\obj{i}$ndependent get and set) to prevent secrets from leaking but still enable reasoning about memory safety.
-Due to the technical setup, the rule needs to check if the access is in bounds ($\obj{\loc}+\obj{n}\in\obj{H^t}$) and update the secrecy tag ($\obj{\securitytag{''}}=\obj{\securitytag{}\sqcap\securitytag{'}}$) with the least upper bound of the tags, according to the aforementioned lattice.
-The precise information carried by the event, e.g., location ($\obj{\loc}$) is taken from the pointer map, which carries information irrelevant to this rule ($\obj{\poison}$).
-
-Base-events include $\obj{\ev{Branch}\ n}$ and $\obj{\ev{Binop}\ n}$ that are emitted when evaluating a branch or certain binary expressions, such as division, respectively, whenever the constant-time mode is inactive.
-Events are extended with a security-tag ($\obj{\securitytag{}}$) to signal the secrecy of the involved data.
-
-The evaluation steps are amended to propagate the security-tag annotations $\obj{\securitytag{}}$.
-When the constant-time mode is inactive, base-events $\obj{\ev{Branch}\ n}$ and $\obj{\ev{Binop}\ n}$ are emitted for conditionals and binary operations, respectively.
-Otherwise, just like in the semantics of the earlier languages, $\obj{\emptyevent}$ is emitted for binary and branching operations.
-%To mark this mode as active or inactive, \Cref{tr:e-wrdoit} allows to access the register in $\obj{\statevar}$.
-
-\Cref{ex:lscct} illustrates the differences between $\Lscct$ and other languages.
-
-\begin{exampleenv}[$\Lscct$ with and without constant-time mode]\label{ex:lscct}
-  Consider again \Cref{ex:strncpy}, with a context copying the string \texttt{Hello World}, where everything is marked with a high security tag: $\lock$.
-  The top half of \Cref{fig:ex-cct} (titled $\Lscct$), describes the execution trace of the program, while the bottom side of the table (titled ``Specification''), describes the related specification trace (\Cref{subsec:msctss:tracemodel}).
-  Read in parallel from top to bottom, the figure shows parts of the execution trace. 
-  In each half, the left column (\textit{Active}) has constant-time mode $\obj{ON}$ and the right one (\textit{Inactive}) has it $\obj{OFF}$.
-
-  When the constant-time mode is off, the execution yields events in similar fashion to before (\Cref{subsec:ltms,subsec:lsms,subsec:lms}).
-  But, if it is turned on, then the branching event \colorbox{yellow}{does not fire} anymore and both reading and writing to memory is related to a specification trace with \colorbox{lightgray}{no exposed secrets}.
-\end{exampleenv}
-\begin{figure}[!htb]
-	\vspace{-1em}\small
-  $$
-  % {
-  \begin{array}{ccc}
-    & \Lscct &        \\
-    \text{Active} & \mid & \text{Inactive} \\\hline
-    \obj{\ev{Alloc}\ \loc_{x}\ 12;\comp;\unlock} & \mid & \obj{\ev{Alloc}\ \loc_{x}\ 12;\comp;\unlock} \\
-    \obj{\ev{Alloc}\ \loc_{y}\ 12;\comp;\unlock} & \mid & \obj{\ev{Alloc}\ \loc_{y}\ 12;\comp;\unlock} \\
-    \obj{{\ev{iGet}}\ \loc_{x}\ 0;\comp;}\text{\colorbox{lightgray}{$\obj{\lock}$}} & \mid & \obj{\ev{Get}\ \loc_{x}\ 0;\comp;\lock} \\
-    \text{\colorbox{yellow}{$\obj{\emptyevent}$}} & \mid & \obj{\ev{Branch}\ 0;\comp;\lock} \\
-    \obj{{\ev{iGet}}\ \loc_{x}\ 0;\comp;}\text{\colorbox{lightgray}{$\obj{\lock}$}} & \mid & \obj{\ev{Get}\ \loc_{x}\ 0;\comp;\lock} \\
-    \obj{{\ev{iSet}}\ \loc_{y}\ 0\ \mathtt{'H'};\comp;}\text{\colorbox{lightgray}{$\obj{\lock}$}} & \mid & \obj{\ev{Set}\ \loc_{y}\ 0\ \mathtt{'H'};\comp;\lock} \\
-    \obj{{\ev{iGet}}\ \loc_{x}\ 1;\comp;}\text{\colorbox{lightgray}{$\obj{\lock}$}} & \mid & \obj{\ev{Get}\ \loc_{x}\ 1;\comp;\lock} \\
-    \text{\colorbox{yellow}{$\obj{\emptyevent}$}} & \mid & \obj{\ev{Branch}\ 0;\comp;\lock} \\
-    \vdots & \mid & \vdots \\
-    \obj{{\ev{iGet}}\ \loc_{x}\ 12;\comp;}\text{\colorbox{lightgray}{$\obj{\lock}$}} & \mid & \obj{\ev{Get}\ \loc_{x}\ 12;\comp;\lock} \\
-    \text{\colorbox{yellow}{$\obj{\emptyevent}$}}& \mid & \obj{\ev{Branch}\ 1;\comp;\lock} \\
-    \obj{\ev{Dealloc}\ \loc_{y};\comp;\unlock} & \mid & \obj{\ev{Dealloc}\ \loc_{y};\comp;\lock} \\
-    \obj{{\ev{iGet}}\ \loc_{y}\ 6;\comp;}\text{\colorbox{lightgray}{$\obj{\lock}$}} & \mid & \obj{\ev{Get}\ \loc_{y}\ 6;\comp;\lock} \\
-  \end{array}
-  % }
-  $$
-  $$
-  % {
-  \begin{array}{ccc}
-          & \text{Specification} & \\
-  \text{Active} & \mid & \text{Inactive} \\\hline
-     \ev{Alloc}\ \loc_{x}\ 12;\unlock & \mid & \ev{Alloc}\ \loc_{x}\ 12;\unlock\\
-     \ev{Alloc}\ \loc_{y}\ 12;\unlock & \mid & \ev{Alloc}\ \loc_{x}\ 12;\unlock\\
-     \ev{Use}\ \loc_{x}\ 0;\text{\colorbox{lightgray}{$\unlock$}} & \mid & \ev{Use}\ \loc_{x}\ 0;\lock\\
-     \emptyevent & \mid & \ev{Branch}\ 0;\lock\\
-     \ev{Use}\ \loc_{x}\ 0;\text{\colorbox{lightgray}{$\unlock$}} & \mid & \ev{Use}\ \loc_{x}\ 0;\lock\\
-     \ev{Use}\ \loc_{y}\ 0;\text{\colorbox{lightgray}{$\unlock$}} & \mid & \ev{Use}\ \loc_{y}\ 0;\lock\\
-     \ev{Use}\ \loc_{x}\ 1;\text{\colorbox{lightgray}{$\unlock$}} & \mid & \ev{Use}\ \loc_{x}\ 1;\lock\\
-     \emptyevent & \mid & \ev{Branch}\ 0;\lock\\
-     \vdots & \mid & \vdots \\
-      \ev{Use}\ \loc_{x}\ 12;\text{\colorbox{lightgray}{$\unlock$}} & \mid & \ev{Use}\ \loc_{x}\ 12;\lock\\
-      \emptyevent & \mid & \ev{Branch}\ 1;\lock\\
-      \ev{Dealloc}\ \loc_{y};\unlock & \mid & \ev{Dealloc}\ \loc_{y};\unlock\\
-      \ev{Use}\ \loc_{y}\ 6;\text{\colorbox{lightgray}{$\unlock$}} & \mid & \ev{Use}\ \loc_{y}\ 6;\lock\\
-  \end{array}
-  % }
-  $$
-  \caption{Traces for \Cref{ex:lscct}.}
-  \label{fig:ex-cct}
-\end{figure}
-
-\subsection{$\Lspec$: A Memory-Unsafe Language with Speculation}\label{subsec:lspec}
-
-$\Lspec$ extends $\Lscct$ (\Cref{subsec:lscct}) with speculative dynamic semantics that is inspired by existing work~\cite{guarnieri2018spectector,fabian2022automatic}.
-After branching, speculation starts by pushing the current configuration into the speculation state ($\ird{S}$), and run subsequent code in a predetermined window (whose size is $\omega$) until a rollback of operational state is performed.
-The window is set inside the speculation state.
-Within such windows, data may leak, this is marked as high $\ird{\lock}$ and with an annotation that indicates the respective speculative execution variant. 
-In this paper, we just consider SPECTRE-PHT~\cite{kocher2019spectre}, whose starting point is \Cref{tr:e-ifz-true-spec}.
-% where $\omega$ is the size of the speculation window, i.e., a number of reduction steps to be taken under speculative semantics.
-Additionally, $\ird{\lock}$ may also carry a $\ird{NONE}$ annotation, to signal that this is a leak irrespective of speculation, i.e., as defined earlier (\Cref{subsec:lscct}).
-The semantics is kept general enough to allow for future extension to support different variants~\cite{kocher2019spectre,maisuradze2018ret2spec,horn2019zero}.
-
-In terms of language features, $\Lspec$ includes a new barrier operation $\ird{\lbarrier}$ that blocks speculative execution. 
-To facilitate speculative execution in a non-assembly-like language, a stack of operational state is used and speculation is active if that stack has a size greater than one.
-This is exploited in \Cref{tr:e-spec-eat} to consume any leftover speculation.
-
-\vspace{-.5em}
-\[
-  \begin{array}{rrcl}
-    \text{(Base-Events)} & \ird{\preevent} &:=& \dots \mid \ird{Spec} \mid \ird{Rlb} \mid \ird{Barrier}\\
-    \text{(Expressions)} & \ird{\expr} &:=& \dots \mid \ird{\lbarrier} \\
-    \text{(Speculation State)} & \ird{S} &:=& \ird{\Omega\triangleright\expr} \mid \ird{S},\left(\ird{\Omega};\ird{n};\ird{\expr}\right)
-  \end{array}
-\]
-
-\begin{center}
-\newcommand{\expreval}[5]{{#1}\triangleright\xspace {#2}\xrightarrow{#5}\ {#3}\triangleright\xspace {#4}\xspace}
-\newcommand{\exprevald}[5]{\expreval{\ird{#1}}{\ird{#2}}{\ird{#3}}{\ird{#4}}{\ird{#5}}}
-
-  %
-  \typerule{$e-\ird{ifz}-\ird{true}-\text{spec}$}{
-    \ird{S} = \rtt{(\ird{\cfstate['];\sandboxtag{};D;\memstate});\ird{S}}{\ird{\expr[_1]}},({\ird{\cfstate;\sandboxtag{};D;\memstate}};\omega;{\ird{\expr[_2]}})
-  }{
-    \pstepto{\rtt{\ird{\cfstate;\sandboxtag{};D;\memstate}}{\ird{\lifz{0^{\securitytag{}}}{\expr[_1]}{\expr[_2]}}}}{\ird{S}}{\ird{{Branch}\ 0;\sandboxtag{};\securitytag{}}\cdot\ird{{Spec};\sandboxtag{};\securitytag{}}}
-  }{e-ifz-true-spec}
-  %
-  \typerule{$e-\ird{spec}-\text{eat}$}{
-    \ird{n} > 0 &
-    \pstepto{\rtt{\ird{\Omega}}{\ird{\expr}}}{\rtt{\ird{\Omega'}}{\ird{\expr[']}}}{\ird{\event}}
-  }{
-    \ghoststepto{\ird{S},(\ird{\Omega};\ird{n};\ird{\expr})}{\ird{S},(\ird{\Omega'};\ird{n}-1;\ird{\expr'})}{\ird{\event}}
-  }{e-spec-eat}
-  % 
-  \typerule{$e-\ird{spec}-\text{eaten}$}{
-  }{
-    \ghoststepto{\ird{S},(\ird{\Omega};\ird{0};\ird{\expr})}{\ird{S}}{\ird{Rlb;\Omega.\sandboxtag{};\unlock}}
-  }{e-spec-eaten}
-  %
-  \typerule{$e-\ird{spec}-\text{barrier}$}{
-  }{
-    \ghoststepto{\ird{S},(\ird{\Omega};\ird{n};\ird{\lbarrier})}{\ird{S},(\ird{\Omega};\ird{0};\ird{0})}{\ird{Barrier;\Omega.\sandboxtag{};\unlock}}
-  }{e-spec-barrier}
-\end{center}
-
-The trace model is extended with $\ird{{Spec}}$, $\ird{{Rlb}}$, and $\ird{{Barrier}}$ events that signal the respective operational action. 
-A $\ird{\lbarrier}$ prevents any execution whatsoever besides $\ird{{Rlb}}$ when run in speculative execution mode and does nothing when run in normal mode, e.g., \Cref{tr:e-spec-barrier}.%
-\footnote{In the rule, the notation $\cfstate' = \cfstate[.\text{win} = 0]$ means that $\cfstate'$ is a copy of $\cfstate$ up to field $\text{win}$, which is set to $0$.}
-Lastly, $\ird{Rlb}$ is emitted when the speculation window is zero and the operational state is rolled back (see \Cref{tr:e-spec-eaten}).
-
-\section{Case Study: Composing Secure Compiler Passes and Optimisations \pages{7}}\label{sec:casestud:rtp}
-This section defines several secure compilers, each of which robustly preserves a different property of interest as depicted in \Cref{fig:pipeline}.
-\begin{figure*}[!h]
-  \centering
-  \begin{tikzpicture}
-    \node (S) {$\src{L_{\tmssafe}}$};
-    \node[right=2.0 of S] (T) {$\trg{L}$};
-    \node[right=2.0 of T] (M) {$\irl{L_{\mssafe}}$};
-    \node[below right=1.5 and 1.0 of M] (D0) {$\irl{L_{\mssafe}}$};
-    \node[above right=1.5 and 1.0 of M] (C0) {$\irl{L_{\mssafe}}$};
-    \node[right=3.0 of M] (E) {$\irl{L_{\mssafe}}$};
-    \node[right=2.0 of E] (O) {$\obj{L_{\scctsafe}}$};
-    \node[right=2.0 of O] (Z) {$\ird{L_{\mathghost}}$};
-
-    \draw[->] (S) to[sloped] node[align=center] (tmsedge) {\gls*{tms}\\ \Cref{thm:cca:rtp:tms}} (T);
-    \draw[->] (T) to[sloped] node[align=center] {\gls*{sms}\\ \Cref{thm:ccb:rtp:sms}} (M);
-    \draw[->] (M) to[sloped] node[align=center] {\gls*{dce}\\ \Cref{thm:ccdce:rtp:ms}} (D0);
-    \draw[->] (M) to[sloped] node[align=center] {\gls*{cf}\\ \Cref{thm:cccf:rtp:ms}} (C0);
-    \draw[->] (D0) to[sloped] node[align=center] {\gls*{cf}\\ \Cref{thm:cccf:rtp:ms}} (E);
-    \draw[->] (C0) to[sloped] node[align=center] {\gls*{dce}\\ \Cref{thm:ccdce:rtp:ms}} (E);
-    \draw[->] (E) to[sloped] node[align=center] {\gls*{scct}\\ \Cref{thm:ccscct:rtp:scct}} (O);
-    \draw[->] (O) to[sloped] node[align=center] {\gls*{ss}\\ \Cref{thm:ccspec:rtp:spec}} (Z);
-
-    % Sections
-    \node (sectms) at ($(S)+(1.25,2.5)$) {{\Cref{subsec:cs:tms}}};
-    \node (secsms) at ($(T)+(1.25,2.5)$) {{\Cref{subsec:cs:ms}}};
-    \node (secopts) at ($(M)+(2.0,2.5)$) {{\Cref{subsec:cs:opts}}};
-    \node (secscct) at ($(E)+(1.5,2.5)$) {{\Cref{subsec:cs:scct}}};
-    \node (secss) at ($(O)+(1.5,2.5)$) {{\Cref{subsec:cs:ss}}};
-
-    \draw[thick,dotted,->] ($(S) + (0.2,-0.27)$) to ($(S) - (-0.2,1)$) to node[align=center] {\gls*{ms}\\ \Cref{corr:ccab:rtp:ms}} ($(M) - (0.0,1)$) to (M);
-
-    \draw[thick,dotted,->] (S) to ($(S) - (0.0,2.5)$) to node[align=center,pos=.65] {\gls*{specms}\\ \Cref{thm:ccall:rtp:specms}} ($(Z) - (0.0,2.5)$) to (Z);
-
-    % \draw[thick,dotted,->] (S) to[bend right=33,sloped] node[align=center] {\gls*{ms}\\ \Cref{corr:ccab:rtp:ms}} (M);
-    \draw[thick,dotted,->] (M) to[bend right=0,sloped] node[align=center] {\gls*{ms}\\ \Cref{thm:cccfccdce:rtp:ms}} (E);
-%    \draw[thick,dotted,->] (S) to[out=-45,in=198,sloped] node[align=center] {\gls*{ms}+\gls*{scct}\\ \Cref{thm:ccall:rtp:msscct}} (O);
-    % \draw[thick,dotted,->] (S) to[out=-40,in=210,sloped] node[align=center,pos=.8] {\gls*{specms}\\ \Cref{thm:ccall:rtp:specms}} (Z);
-
-    \draw[dashed] ($(S)-(0.0,-0.5)$) -- ($(S)-(0.0,-2.5)$);
-    \draw[dashed] ($(T)-(0.0,-0.5)$) -- ($(T)-(0.0,-2.5)$);
-    \draw[dashed] ($(M)-(0.0,-0.5)$) -- ($(M)-(0.0,-2.5)$);
-    \draw[dashed] ($(E)-(0.0,-0.5)$) -- ($(E)-(0.0,-2.5)$);
-    \draw[dashed] ($(O)-(0.0,-0.5)$) -- ($(O)-(0.0,-2.5)$);
-    \draw[dashed] ($(Z)-(0.0,-0.5)$) -- ($(Z)-(0.0,-2.5)$);
-  \end{tikzpicture}
-  \vspace{-1em}
-  \caption{Visualisation of the optimising compilation pipeline that preserves \gls*{specms}. %
-    Vertices in the graph are the programming languages from earlier sections (\Cref{sec:casestud:defs}). %
-    Full edges are secure compilers passes.
-    Dotted edges are composition of passes and use the presented framework (\Cref{sec:sequential}) to indicate the property they preserve. %
-    The dashed lines partition the graph into the sections where the respective theorems are presented.
-  }\label{fig:pipeline}
-\end{figure*}
-The section demonstrates the power of the framework (\Cref{sec:sequential}) by composing these compilers for a secure and optimising compilation chain that robustly preserves \gls*{specms}.
-The first step in this chain is the compiler from $\src{L_{\tmssafe}}$ to $\trg{L}$ that robustly preserves just \gls*{tms} (\Cref{thm:cca:rtp:tms}).
-From here, another pass from $\trg{L}$ to $\irl{L_{\mssafe}}$ ensures that no out-of-bounds accesses can happen and, thus, programs at this point attain \gls*{sms} (\Cref{thm:ccb:rtp:sms}).
-Since these properties compose into \gls*{ms}, composing these passes yields a compiler that robustly preserves \gls*{ms} (\Cref{corr:ccab:rtp:ms}).
-Then, the section presents two optimisation passes, namely \gls*{cf} and \gls*{dce}, each of which robustly preserves \gls*{ms} (\Cref{thm:ccdce:rtp:ms,thm:cccf:rtp:ms}).
-These passes can be freely ordered in the compilation chain without compromising memory safety (\Cref{thm:cccfccdce:rtp:ms}).
-The next step in the chain ensures that code stays \gls*{scct} (\Cref{thm:ccscct:rtp:scct}) when compiled from $\Lms$ to $\Lscct$, which is done by switching on a constant-tame mode for the computation.
-Lastly, by introducing barriers immediately after branches, speculative leaks via SPECTRE-PHT are prevented when compiling $\Lscct$ to $\Lspec$.
-The final result is that the whole compilation chain robustly preserves \gls*{specms} (\Cref{thm:ccall:rtp:specms}).
-
-\subsection{Robust Temporal Memory Safety Preservation}\label{subsec:cs:tms}
-
-The secure compiler from $\Ltms$ to $\Ltrg$ needs to ensure that when execution switches from context to component, the type signatures are respected.
-Thus, the compiler inserts the following dynamic typechecks before entering the body of a component-defined function (anything elided is a trivial identity function from source to target):
-
-\vspace{-1em}
-{
-\begin{gather}
-  \begin{align*}%{rcl}
-    %\cca(\src{x}) = &\ \trg{x} 
-    % \\
-  	%&
-  	%\qquad
-    %\cca(\src{\lbinop{\expr[_{1}]}{\expr[_{2}]}}) = &\ \lbinop[\trg]{\left[\cca(\src{\expr[_{1}]})\right]}{\left[\cca(\src{\expr[_{2}]})\right]} \\
-    %\cca(\src{n}) = &\ \trg{n} 
-    % \\
-    %&
-    \cca(\src{\lget{x}{\expr}}) = &\ \lget[\trg]{\trg{x}}{\left[\cca(\src{\expr})\right]} \\
-    \cca(\src{\ldelete{x}}) = &\ \ldelete[\trg]{\left[\cca(\src{x})\right]} 
-    \\
-    \cca(\src{\lfunction{g}{x:\natt\to\type[_{e}]}{\expr}})  =&
-  \end{align*}
-  \\
-  \begin{align*}
-%    \cca(\src{\lset{x}{\expr[_{1}]}{\expr[_{2}]}}) & = \lset[\trg]{\left[\cca(\src{x})\right]}{\left[\cca(\src{\expr[_{1}]})\right]}{\left[\cca(\src{\expr_{2}})\right]} \\
-% \\ 
-\lfunction[\trg]{\trg{g}}{\trg{x}}{\lifz[\trg]{\trg{\lhast{x}{\natt}}}{
-                                                            \left[\cca(\src{\expr})\right] %
-                                                                                                 }{\labort[\trg]}}
-  \end{align*}
-\end{gather}
-}
-
-\noindent Since $\trg{L}$ has no static typing, an attacker $\trg{\library_{\ctx}}$ can invoke a component function accepting a $\src{\natt}$ with $\trg{\lpair{17}{29}}$.
-With the dynamic check, the compiler ensures that execution aborts in such cases.
-%The compiler does not insert other checks and proceeds as the identity function (which in this paper amounts to a simple re-colouring of $\src{L_{\tmssafe}}$ to $\trg{L}$ expressions).
-
-Compiling the \texttt{strncpy} function from \Cref{sec:introduction} with $\cca$, the compiler would in this case ensure that the arguments that are evaluated in the compiled \texttt{strncpy} are valid.
-%This effectively has no influence on the resulting trace, so a corresponding cross-language trace relation $\sim^{\Ltms}_{\Ltrg}$ is just equality.
-
-$\cca$ is robustly preserving (\Cref{def:rtp}) \gls*{tms}:
-\begin{theorem}[$\cca$ secure w.r.t. \gls*{tms}]\label{thm:cca:rtp:tms}
-    $\rtp{\cca}{\tmssafe}$ % \Coqed
-\end{theorem}
-
-\subsection{Robust Spatial Memory Safety Preservation}\label{subsec:cs:ms}
-
-The spatial memory safety preserving compiler from $\Ltrg$ to $\Lms$ only inserts bounds-checks whenever reading from or writing to memory in order to enforce \gls*{sms}.
-These bounds checks need the bounds information, which the compiler keeps around by introducing a fresh identifier $\irl{x_{SIZE}}$ for each allocation that binds $\irl{x}$.
-Then, it is simply a matter of referring to that variable and ensuring that memory accesses are in the interval $[\irl{0},\irl{x_{SIZE}})$.
-When the check fails, the code aborts.
-
-\begin{nscenter}
-% \small
-  $$
-  \begin{array}{rcl}
-    \ccb(\trg{\lnew{x}{\expr[_{1}]}{\expr[_{2}]}}) & = 
-                                                   & \llet[\irl]{\irl{x_{SIZE}}}{\ccb(\trg{\expr[_{1}]})}{}
-    		\\&&
-    		\lnew[\irl]{\irl{x}}{\irl{x_{SIZE}}}{\ccb(\trg{\expr[_{2}]})}
-    		 \\
-  \ccb(\trg{\lget{x}{\expr}}) & = 
-                              & \llet[\irl]{\irl{x_{n}}}{\ccb(\trg{\expr})}{}
-  	\\&&
-  \lifz[\irl]{\irl{0\le x_{n}<x_{SIZE}}}{\\&&\irl{\lget{x}{x_{n}}}}{}
-  		\irl{\labort}
-  	  \\
-  \ccb(\trg{\lset{x}{\expr[_{1}]}{\expr[_{2}]}}) & = 
-                                                 & \llet[\irl]{\irl{x_{n}}}{\ccb(\trg{\expr[_{1}]})}{}
-  		\\&&
-  \lifz[\irl]{\irl{0\le x_{n}<x_{SIZE}}}{\\&&\lset[\irl]{\irl{x}}{\irl{x_{n}}}{}
-  		\ccb(\trg{\expr[_{2}]})
-  		}{\irl{\labort}} 
-  		% \\
-  \end{array}
-  $$
-\end{nscenter}
-
-\begin{exampleenv}[Instrumented \texttt{strncpy}]
-  Consider again \texttt{strncpy}, but instrumented for \gls*{sms}:
-    \begin{lstlisting}[language=c,basicstyle=\ttfamily, morekeywords={size_t}]
-void strncpy(size_t n, size_t dst_size, char *dst,
-             size_t src_size, char *src) {
-  for(size_t i = 0; i < src_size
-      && src[i] != '\0' && i < n; ++i) {
-    if(i < src_size && i < dst_size) {
-      dst[i] = src[i];
-    }
-  }
-}
-    \end{lstlisting}
-Consider context \texttt{strncpy(2, x, y)}, where \texttt{x} and \texttt{y} are pointers to valid regions of memory with allocated space for exactly two cells and do not contain the null-terminating character \texttt{'\textbackslash 0'}.
-% 
-Without the \gls*{sms} pass, the event $\ev{Use}\ \loc_{x}\ 2;\comp;\unlock$ would appear on the trace, but that indicates an out-of-bounds access! 
-Fortunately, with \gls*{sms} mitigation in place, that event does not appear during execution, since the bounds check prevents the condition \texttt{src[i] != '\textbackslash 0'} from executing.
-\end{exampleenv}
-
-Contrary to the previous compiler, $\ccb$ may change the trace of the original $\Ltrg$ program: if there is a memory access, it needs to be protected with a bounds check.
-% 
-The corresponding relation $\sim^{\Ltrg}_{\Lms} : \trg{\trace}\times\irl{\trace}$ that describes this semantic effect of the compiler is defined partially below.
-% 
-We omit action $\trg{Set}$, which is treated analogously, and any other event, which is related to its cross-language equivalent.
-% \footnote{Up to colours.}
-% \footnote{In the interest of space, the full details are in the technical report.}
-
-{
-\[
-  \typerulenolabel{xrel:sms:read}{\trg{n}\text{ in bounds}}{\trg{Get\ \loc\ n;\comp}\sim^{\Ltrg}_{\Lms}\irl{Get\ \loc\ n;\comp}}
-\]
-% \[
-%   \typerulenolabel{xrel:sms:panic}{\trg{n}\text{ out of bounds}}{\trg{Get\ \loc\ n;\comp}\sim^{\Ltrg}_{\Lms}\irl{\lightning}}
-% \]
-}
-
-% Note that neither $\Ltrg$ nor $\Lms$ expose branches on the trace.\MP{so what?}
-For simplicity, we elide the environment that this relation carries around in order to bind each location to its metadata (such as its size), and resolve the ``$\trg{n}$ in/out of bounds'' premise.
-% The bounds check mentioned in the premises of the rules above is left abstract for the sake of simplicity, the interested reader can find
-% since the precise definition needs some technicalities that only inhibit readability without any technical insights.
-% 
-We can now prove that compiler $\ccb$ robustly preserves \gls*{sms}.
-\begin{theorem}[$\ccb$ secure w.r.t. \gls*{sms}]\label{thm:ccb:rtp:sms}
-  $\rtpsim{\ccb}{\smssafe}{\sim^{\Ltrg}_{\Lms}}$ % \Coqed
-\end{theorem}
-
-At this point we can compose $\ccb$ with the previous compiler ($\cca$), but in order to do so, we need a trace relation from $\Ltms$ to $\Lms$.
-% 
-We can obtain this relation by composing the trace relation we just defined ($\sim^{\Ltrg}_{\Lms}$) with the one used by the previous compiler: $\sim^{\Ltms}_{\Ltrg} : \src{\trace}\times\trg{\trace}$.
-% 
-The latter has not been previously defined (nor has it been used in the related theorem) because that is just an equality relation, since the trace models of $\Ltms$ and of $\Ltrg$ are the same.
-% 
-Thus, we formally define $\sim^{\Ltms}_{\Lms}: \src{\trace} \times \irl{\trace}$ as the following composition: $\sim^{\Ltms}_{\Ltrg}\bullet\sim^{\Ltrg}_{\Lms}$.
-% 
-With this relation, \Cref{corr:ccab:rtp:ms} states that the composition of $\cca$ and $\ccb$ is secure w.r.t.
-\gls*{ms} and it follows from \Cref{thm:cca:rtp:tms,thm:ccb:rtp:sms} using \Cref{thm:rtpsim:sig}.
-% 
-\begin{corollary}[$\cca\circ\ccb$ secure w.r.t. \gls*{ms}]\label{corr:ccab:rtp:ms}
-  $\;$
-
-  \begin{nscenter}
-    $\rtpsim{\cca\circ\ccb}{\mssafe}{\sim^{\Ltms}_{\Lms}}$ % \Coqed
-  \end{nscenter}
-\end{corollary}
-% 
-This proof requires another precondition besides \Cref{thm:ccb:rtp:sms,thm:cca:rtp:tms}: $\sim^{\Ltms}_{\Ltrg}$ needs to be well-formed with respect to $\smssafe$.
-This follows trivially since $\sim^{\Ltms}_{\Ltrg}$ is an equality. 
-% 
-\begin{lemma}[$\sim^{\Ltms}_{\Ltrg}$ well-formed w.r.t. $\smssafe$]\label{lem:wf:ltmsltrg}
-$\;$
-
-  \begin{nscenter}
-  $\wfcsig{\sim^{\Ltms}_{\Ltrg}}{\smssafe}$
-  \end{nscenter}
-\end{lemma}
-
-\subsection{Optimising Compilers}\label{subsec:cs:opts}
-
-This section defines two optimising compiler passes from $\Lms$ to $\Lms$ which perform \gls*{dce} and \gls*{cf}, respectively.
-The \gls*{dce} pass applies a na\"ive rewrite rule on conditionals.
-The \gls*{cf} pass relies on an auxiliary function \texttt{mix} that uses a substitutions accumulator $\irl{\substlist}$ in order to rewrite constant binary operations, e.g., $\irl{{17}-1}$ to $\irl{16}$, and replace variables that are assigned to constants, e.g., $\irl{\llet{x}{7}{x}}$ to $\irl{7}$.
-
-\vspace{-1em}
-\begin{gather*}
-  \begin{align*}
-    \ccdce(\irl{\lifz{true}{\expr[_{1}]}{\expr[_{2}]}}) & = \ccdce(\irl{\expr[_{1}]}) &\\
-    \ccdce(\irl{\lifz{false}{\expr[_{1}]}{\expr[_{2}]}}) & = \ccdce(\irl{\expr[_{2}]}) &
-  \end{align*}
-  \\
-  % \begin{align*}
-  %   \ccdce(\irl{\lbinop{\expr[_{1}]}{\expr[_{2}]}}) & = \lbinop[\irl]{\ccdce(\irl{\expr[_{1}]})}{\ccdce(\irl{\expr[_{2}]})} &
-  % \end{align*}
-  % \\[0.25cm]
-  \begin{align*}
-    \cccf(\irl{\expr}) & = \partialeval{\irl{\expr}}{\irl{\hole{\cdot}}} &
-  \end{align*}
-  \\[0.125cm]
-  \begin{align*}
-   \partialeval{\irl{x}}{\irl{\substlist}} & = \irl{n} 
-   	\qquad\qquad \text{if } \irl{\subst{n}{x}}\in\irl{\substlist} \\
-   \partialeval{\irl{x}}{\irl{\substlist}} & = \irl{x} 
-   \qquad\qquad \text{if } \irl{\subst{n}{x}}\notin\irl{\substlist} \\
-   \partialeval{\irl{\lbinop{n}{m}}}{\irl{\substlist}} & = \irl{k} 
-   \qquad\qquad \text{if } \lbinop{\irl{n}}{\irl{m}}=k \\
-   %\partialeval{\irl{\lbinop{\expr[_{1}]}{\expr[_{2}]}}}{\irl{\substlist}} & = \lbinop[\irl]{\partialeval{\irl{\expr[_{1}]}}{\irl{\substlist}}}{\partialeval{\irl{\expr[_{2}]}}{\irl{\substlist}}} \\
-   \partialeval{\irl{\llet{x}{n}{\expr}}}{\irl{\substlist}} & = \partialeval{\irl{\expr}}{\irl{\subst{n}{x}\cdot\substlist}} 
-  % \\
-  % \partialeval{\irl{\lget{x}{\expr}}}{\irl{\substlist}} & = \lget[\irl]{\irl{x}}{\partialeval{\irl{\expr}}{\irl{\substlist}}}
-% \end{align*}
-\end{align*}
-% \\
-% \begin{align*}
-%   \partialeval{\irl{\llet{x}{\expr[_{1}]}{\expr[_{2}]}}}{\irl{\substlist}} & = \llet[\irl]{\irl{x}}{\partialeval{\irl{\expr[_{1}]}}{\irl{\substlist}}}{\\&\partialeval{\irl{\expr[_{2}]}}{\irl{\substlist}}} \\
-%   \partialeval{\irl{\lifz{\expr[_{1}]}{\expr[_{2}]}{\expr[_{3}]}}}{\irl{\substlist}} & = \lifz[\irl]{\partialeval{\irl{\expr[_{1}]}}{\irl{\substlist}}}{\\&\partialeval{\irl{\expr[_{2}]}}{\irl{\substlist}}}{\partialeval{\irl{\expr[_{3}]}}{\irl{\substlist}}} \\
-% \end{align*}
-\end{gather*}
-% \vspace{-3em}
-
-Note that both passes have no effect on the resulting trace of a program, up to $\emptyevent$-steps. 
-Because of this, both passes have equality as corresponding cross language trace relation. 
-Moreover, it is straightforward to prove both passes as secure (\Cref{def:rtp}) w.r.t. \gls*{ms}.
-
-\begin{theorem}[$\ccdce$ secure w.r.t. \gls*{ms}]\label{thm:ccdce:rtp:ms}
-  $\rtp{\ccdce}{\mssafe}$ %\Coqed
-\end{theorem}
-\begin{theorem}[$\cccf$ secure w.r.t. \gls*{ms}]\label{thm:cccf:rtp:ms}
-  $\rtp{\cccf}{\mssafe}$ %\Coqe
-\end{theorem}
-
-With both \Cref{thm:ccdce:rtp:ms,thm:cccf:rtp:ms} it follows from \Cref{corr:swappable} that the two passes can be interchanged arbitrarily:
-
-\begin{theorem}[$\cccf\circ\ccdce$ and $\cccf\circ\ccdce$ are secure w.r.t. \gls*{ms}]\label{thm:cccfccdce:rtp:ms}
-$\;$
-
-  \begin{nscenter}
-  \phantom{and~}\!\!$\rtp{\cccf\circ\ccdce}{\mssafe}$
-
-  and~$\rtp{\ccdce\circ\cccf}{\mssafe}$ % \Coqed
-  \end{nscenter}
-\end{theorem}
-
-\subsection{Robust Strict Cryptographic Constant Time Preservation}\label{subsec:cs:scct}
-
-This section defines a compiler $\ccscct$ from $\Lms$ to $\Lscct$ that robustly preserves \gls*{scct}. 
-Given the fact that $\Lscct$ provides a \gls*{cct}-mode that can be turned on or off, the compiler inserts wrapper code for function calls and function bodies to ensure that execution in the component always happen in \gls*{cct}-mode.
-This simple flag combines the effect of FaCT~\cite{cauligi2019fact}
-
-\vspace{-1em}
-\[
-\begin{array}{rcl}
-  \ccscct(\irl{\lfunction{g}{x}{\expr}}) & = & \lfunction[\obj]{\obj{g}}{\obj{x}}{\obj{\lwrdoit{ON};}\ccscct(\irl{\expr})} \\
-  \ccscct(\irl{\lcall{g}{\expr}}) & = & \lcall[\obj]{\obj{g}}{\ccscct(\irl{\expr})\obj{; \lwrdoit{ON}}} 
-    % \\
-    % \ccscct(\irl{\lbinop{\expr[_{1}]}{\expr[_{2}]}}) & = \lbinop[\obj]{\ccscct{\irl{\expr[_{1}]}}}{\ccscct{\irl{\expr[_{2}]}}} 
-    % \\
-\end{array}
-\]
-% \vspace{-2em}
-%
-The context can overwrite the flag and exit the mode, but upon invoking a function that is part of the component, the flag is set again.
-Because of this, the corresponding cross-language trace relation $\sim^{\Lms}_{\Lscct}$, only relates events without secrets:%
-
-\begin{center} 
-  \typerulenolabel{xrel:scct:noleak}{}{\irl{\preevent;\comp}\sim^{\Lms}_{\Lscct}\obj{\preevent;\comp;\unlock}}
-\end{center}
-
-The compiler is secure w.r.t. \gls*{scct}: %, similarly proven as in \Cref{subsec:cs:tms}.
-
-\begin{theorem}[$\ccscct$ secure w.r.t. \gls*{scct}]\label{thm:ccscct:rtp:scct}
-  \small
-  $\rtpsim{\ccscct}{\scctsafe}{\sim^{\Lms}_{\Lscct}}$ % \Coqed
-\end{theorem}
-
-\subsection{Robust Speculative Safety Preservation}\label{subsec:cs:ss}
-
-This section defines the final compilation pass $\ccspec$, which ensures that $\Lscct$ programs, which are assumed to be \gls*{ss}, stay \gls*{ss} at $\Lspec$-level. 
-To do so, $\ccspec$ inserts a speculation barrier after branches, which is sufficient to harden the program against speculative leaks, since SPECTRE-PHT~\cite{kocher2019spectre} is the only speculative leak modeled in the semantics of $\Lspec$.
-
-\vspace{-1em}
-\[
-\begin{array}{cl}
-  &\ccspec{(\obj{\lifz{\expr[_0]}{\expr[_1]}{\expr[_2]}})} = 
-  \\
-  &\qquad\qquad \lifz[\ird]{\ccspec{\left(\obj{\expr[_0]}\right)}}{\ird{\lbarrier;}\ccspec{\left(\obj{\expr[_1]}\right)} \\&\qquad\qquad}{ \ird{\lbarrier;}\ccspec{\left(\obj{\expr[_2]}\right)}} 
-\end{array}
-\]
-% \vspace{-2em}
-%
-
-Clearly, the corresponding cross-language trace relation $\sim^{\Lscct}_{\Lspec}$ has only one non-trivial case: for branches, only relate them where speculation is blocked by a barrier:
-
-\begin{center}
-  \typerulenolabel{xrel:spec:if}{}{\obj{Branch\ n}\sim^{\Lscct}_{\Lspec}\ird{Branch\ n}\cdot\ird{Spec}\cdot\ird{Barrier}\cdot\ird{Rlb}}
-\end{center}
-
-The base-event relation above scales to full events by ensuring the missing annotations ($\obj{\comp;\securitytag{}}$ and $\ird{\comp;\securitytag{}}$) are the same.
-% 
-With this relation, we prove that $\ccspec$ is secure with respect to \gls*{ss}.
-\begin{theorem}[$\ccspec$ secure w.r.t. \gls*{ss}]\label{thm:ccspec:rtp:spec}
-  \small$\rtpsim{\ccspec}{\sssafe}{\sim^{\Lscct}_{\Lspec}}$ % \Coqed
-\end{theorem}
-
-\subsection{Robust Preservation of Memory Safety, Strict Cryptographic Constant Time, and Speculative Safety}
-
-Finally, this subsection combines all previous results into one compilation chain to get that it preserves full \gls*{specms}.
-Let $\ccspecms$ be the compiler that is the composition of $\cca$, $\ccb$, $\cccf$, $\ccdce$, $\ccscct$, and $\ccspec$. 
-Let $\sim^{\Ltms}_{\Lspec}$ be the composition of $\sim^{\Ltms}_{\Lms}$, $\sim^{\Lms}_{\Lscct}$, and $\sim^{\Lscct}_{\Lspec}$.
-Then, the following corollary holds.
-
-\begin{corollary}[$\ccspecms$ secure w.r.t. \gls*{specms}]\label{thm:ccall:rtp:specms}
-  $\;$
-
-  \begin{nscenter}
-    $\rtpsim{\cc{\Ltms}{\Lspec}}{\mssafe\cap\scctsafe\cap\sssafe}{\sim^{\Ltms}_{\Lspec}}$ % \Coqed
-  \end{nscenter}
-\end{corollary}
-
-As with \Cref{corr:ccab:rtp:ms}, it is important to ensure that the respective cross language trace relations are well-formed (\Cref{def:wfc:sig:tracerel}).
-It is already known that $\sim^{\Ltms}_{\Lms}$ is well-formed with respect to $\mssafe$ (\Cref{lem:wf:ltmsltrg}).
-Next in the chain is $\sim^{\Lms}_{\Lscct}$, which has to be well-formed w.r.t. $\scctsafe$.
-This lemma holds, since a trace that was $\scctsafe$ is $\scctsafe$ even after applying $\sim^{\Lms}_{\Lscct}$: the relation enforces that $\Lscct$ traces related to $\Lms$ traces have no leaks of secrets whatsoever.
-
-\begin{lemma}[$\sim^{\Lms}_{\Lscct}$ well-formed w.r.t. $\scctsafe$]\label{lem:wf:lsmslscct}
-$\;$
-
-  \begin{nscenter}
-  $\wfcsig{\sim^{\Lms}_{\Lscct}}{\scctsafe}$
-  \end{nscenter}
-\end{lemma}
-
-The last relation is $\sim^{\Lscct}_{\Lspec}$ which needs to be well-formed w.r.t. $\sssafe$.
-Similarly to the previous relation, this holds, since $\sim^{\Lscct}_{\Lspec}$ only relates $\Lspec$ traces, which do not have speculative leaks, with $\Lscct$ traces.
-
-\begin{lemma}[$\sim^{\Lscct}_{\Lspec}$ well-formed w.r.t. $\sssafe$]\label{lem:wf:lscctlspec}
-  $\wfcsig{\sim^{\Lscct}_{\Lspec}}{\sssafe}$
-\end{lemma}
-
-\section{Formal Insights}\label{sec:formalities}
-
-This section first discusses how to connect each language-specific security property to the general security properties presented in \Cref{sec:compprop} (\Cref{subsec:formalities:maps}), then it demonstrates that the security property resulting from the universal image projection is faithful to the original property (\Cref{subsec:formalities:props}).
-Finally, this section discusses why the order of compiler passes matters, and how does our framework help with identifying insecure compositions (\Cref{subsec:compatsecpasses}), and it gives additional technical insights on the secure compilation proofs (\Cref{subsec:seccompproofs}).
-
-\subsection{From Language Traces to General Ones}\label{subsec:formalities:maps}
-
-The previous theorems talk about preserving properties expressed in the trace model of the languages of each compiler.
-However, these trace models are not the same trace model we used to specify the properties of \Cref{sec:compprop} (indicated with \ev{L}), which serves as the ``ground truth'' for the meaning of our properties.
-% 
-To bridge this gap, the formal development requires specifying additional trace relations, from each of the language trace models to the \ev{L} one, that, for example, relate $\src{Get\ \loc\ n}$ and $\src{Set\ \loc\ n}$ to $\ev{Use\ \loc\ n}$ (and that induce a related universal image that we use in \Cref{subsec:formalities:props}).
-% 
-One key insight of these relations is that they all omit context-made actions for two reasons: (1) contexts (which are universally quantified) can trivially invalidate any property and (2) we are interested in the component upholding the properties.
-%
-
- % disregarded a glaring but minor technical issue: take $\tmssafe$ (\Cref{def:trace:tmsdef}) as an example.
-% It contains traces as specified in \Cref{subsec:basic:memsafety:tracemodel}.
-% However, these are different from $\Ltms$ traces (\Cref{subsec:ltms}) which means that \Cref{thm:wt:tms} is ``ill-typed''.
-% Similarly for the other theorems.
-% In the interest of reducing visual baggage, the notation left out yet another cross-language trace relation, translating language-specific traces into these non-language specific traces. 
-% In doing so, context-level actions are dropped and, e.g., both $\src{Get\ \loc\ n}$ and $\src{Set\ \loc\ n}$ are translated into $\ev{Use\ \loc\ n}$.
-
-\subsection{Security Properties and Their Meaning}\label{subsec:formalities:props}
-Each of the presented compilers use a cross-language trace relation, which is also used to translate the property from one language to the other one (via the existential or universal images).
-% 
-%Whenever such translations occur, it is important to check whether there has been a change of meaning.
-%Secure compiler engineers need to ensure that such translations of properties do not change meaning in a fundamental wrong way. 
-% lifting or lowering reasoning from target- to source-level or vice-versa,
-While the meaning of projected properties does change with a translation, the change should not allow for a flawed compilation pipeline.
-% 
-For example, we could be using a trace relation that translates a property in a language to a totally different one in another language.
-% 
-To raise the trust into the translation of properties, \Cref{thm:prop-rel-corr} states (in a general fashion) that each security property is faithfully translated using the universal image according to the cross-language trace relation induced by the compiler.
-
-\begin{theorem}[Properties Relation Correctness]\label{thm:prop-rel-corr}
-  \begin{align*}
-    \forall& \pi \in \{\tmssafe,\smssafe,\scctsafe,\sssafe\}, 
-    \\
-    \text{ for each }& \text{pair of languages } \src{L} \text{ and } \trg{L} \text{ used by the compilers},
-    \\
-    \text{ if }& 
-    % \tau_{\sim^{\trg{L}}_{\ev{L}}}(\pi) \sim^{\trg{L}}_{\ev{L}} \pi
-    \sigma_{\sim^{\trg{L}}_{\ev{L}}}(\pi) \sim^{\trg{L}}_{\ev{L}} \pi
-    \text{ and } 
-    % \sigma_{\sim^{\src{L}}_{\trg{L}}}(\tau_{\sim^{\trg{L}}_{\ev{L}}}(\pi)) \sim^{\src{L}}_{\trg{L}} \tau_{\sim^{\trg{L}}_{\ev{L}}}(\pi)
-    \sigma_{\sim^{\src{L}}_{\trg{L}}}(\sigma_{\sim^{\trg{L}}_{\ev{L}}}(\pi)) \sim^{\src{L}}_{\trg{L}} \sigma_{\sim^{\trg{L}}_{\ev{L}}}(\pi)
-    \\
-    \text{ then }& 
-    % \tau_{\sim^{\src{L}}_{\ev{L}}}(\pi) = \sigma_{\sim^{\src{L}}_{\trg{L}}}(\tau_{\sim^{\trg{L}}_{\ev{L}}}(\pi)) 
-    \sigma_{\sim^{\src{L}}_{\trg{L}}}(\sigma_{\sim^{\trg{L}}_{\ev{L}}}(\pi)) \sim^{\src{L}}_{\ev{L}} \pi
-  \end{align*}
-\end{theorem}
-The complexity of this theorem is that the relation in the conclusion cannot be obtained by composing the two relations in the premises.
-% 
-%% MK: the following does not say anything about the theorem, only about the syntactic compiler
-%What happens in practice is that compiler engineers are given the relations from each language to \ev{L} and they need to build a compiler that induces a relation $\sim^{\src{L}}_{\trg{L}}$ that respects this theorem.
-%
-
-We now informally argue why this theorem holds for the composition of all considered properties from \Cref{sec:compprop}.
-
-\paragraph{$\sigma_{\sim^{\Ltms}_{\Lms}}\left(\tmssafe\cap\smssafe\cap\scctsafe\cap\sssafe\right)$}
-
-For the four properties considered here, the trace models of $\Ltms$, $\Ltrg$, and $\Lms$ do not consider actions related to $\scctsafe$ and $\sssafe$, so these two properties are trivially translated correctly.
-% 
-We now discuss the remaining $\tmssafe$ and $\smssafe$ in the form of their intersection $\mssafe$.
-
-Recall that $\sim^{\Ltms}_{\Lms}$ is the composition of $\sim^{\Ltms}_{\Ltrg}$ and $\sim^{\Ltms}_{\Lms}$.
-% 
-Since $\sim^{\Ltms}_{\Ltrg}$ is an equality, this relation trivially preserves the meaning of translated properties: related traces are identical!
-
-Finally, let us consider a trace $\irl{\trace}\in\mssafe$ and understand what is that is related to via $\sim^{\Ltrg}_{\Lms}$.
-% 
-{\em All} traces $\trg{\trace}$ with $\trg{\trace}\sim^{\Ltrg}_{\Lms}\irl{\trace}$ are identical to $\irl{\trace}$ except for get and set actions, which require for in-bound accesses (as stated in \Cref{subsec:cs:ms}): this clearly respect $\mssafe$.
-%
-
-% In both cases, the translated event does not represent an out-of-bounds memory access.
-
-% \footnote{Both $\sim^{\Ltms}_{\Ltrg}$ and $\sim^{\Ltrg}_{\Lms}$ filter context actions.}.
-
-% Now, starting from $\mssafe$ expressed in $\Lms$, we check that its translation in $\Ltms$ is still $\mssafe$.
-% % 
-% We can see that the property resulting from the projection of $\Lms$ traces back to $\Ltms$ faithfully models $\mssafe$, which translates to both $\scctsafe$ and $\sssafe$, 
-% \MPin{ equiv  +  other rel only adds traces that fail }
-% since these properties trivially hold for $\Ltms$, $\Ltrg$, and $\Lms$ traces.
-
-\paragraph{$\sigma_{\sim^{\Lms}_{\Lscct}}\left(\tmssafe\cap\smssafe\cap\scctsafe\cap\sssafe\right)$}
-
-As before, the trace models of $\Lscct$ and $\Lms$ cannot express $\sssafe$, so that is trivially translated correctly.
-% 
-Also, the trace model of $\Lscct$ extends the one of $\Lms$ with respect to $\tmssafe$ and $\smssafe$ events, so the translation argument regarding those two properties is the same as before.
-% 
-Thus, we need to reason about whether $\scctsafe$ is translated correctly.
-
-By definition, $\sim^{\Lms}_{\Lscct}$ only relates $\obj{\unlock}$ events, which is also the same relation induced by $\sim_{\ctsafe}$ in \Cref{sec:msscct-rel}.
-% 
-This ensures that composing the relations only relates $\unlock$ events, and thus the property is translated correctly.
-
-% Consider a trace $\obj{\trace}\in\scctsafe$ and all $\irl{\trace}$ such that $\irl{\trace}\sim^{\Lms}_{\Lscct}\obj{\trace}$. 
-% In turn, $\obj{\trace}$ cannot contain any leaks of secret values, so it cannot violate $\scctsafe$.
-% Toggling data-independent timing mode has no impact on $\mssafe$ or $\sssafe$.
-
-\paragraph{$\sigma_{\sim^{\Lscct}_{\Lspec}}\left(\tmssafe\cap\smssafe\cap\scctsafe\cap\sssafe\right)$}
-The trace model of $\Lspec$ extends the one of $\Lscct$ with respect to $\tmssafe$, $\smssafe$, and $\scctsafe$, so the translation argument for those three properties is the same as before.
-% 
-Concerning $\sssafe$, $\sim^{\Lscct}_{\Lspec}$ relates $\Lspec$ speculation traces with $\Lscct$ branches with the $\lock_{\text{NONE}}$ tag, so the property is translated correctly, according to the relation defined in \Cref{sec:spec-ms-rel}.
- % introducing $\ird{barrier}$ never violates $\sssafe$ and has no effect on $\mssafe$ or $\scctsafe$.
-
-%\MPin{
-%	can this be made formal?
-%	what's the point?
-%}
-
-\subsection{Compatibility of Secure Compiler Passes}\label{subsec:compatsecpasses}
-
-% \Cref{thm:rtpsim:sig} is applicable, up to the well-formedness of $\sim_1$ w.r.t. $\trg{\class[_2]}$.
-Consider applying $\ccspec$ first and then $\ccb$ (albeit currently syntactically impossible).
-% , let us pretend that it is possible by just extending $\ccb$ for the constructs present in $\Lscct$ that are not present in $\Ltrg$.
-% Clearly, as evidenced earlier (\Cref{subsec:formalities:props}), $\ccb\cdot\ccspec$ is an acceptable composition that yields an interesting class of security properties, i.e., $\sigma_{\sim_{\irl{ms}}\bullet\sim_{\ird{\mathghost}}}\left(\mssafe\cap\scctsafe\cap\sssafe\right)$.
-In this case, $\ccb$ would insert new branches into the code that are not protected by a speculation barrier! 
-This concern is reflected in the proofs that establish that the security class resulting of the composition of the trace relations is meaningful.
-% , much similar to ensuring that a definition models what the formal methods expert intends to model. 
-For the $\ccspec\cdot\ccb$ case, the class is $\sigma_{\sim_{\ird{\mathghost}}\bullet\sim_{\irl{ms}}}\left(\mssafe\cap\scctsafe\cap\sssafe\right)$.
-Since $\ev{Use\ \loc\ n}\sim_{\irl{ms}}\irl{Branch\ n}\cdot\irl{Spec}\cdots$, where $\cdots$ does not contain a $\ev{{Barrier}}$ event, the resulting class is \emph{not} the original $\sssafe$ that is intended and it would break the corresponding \Cref{thm:prop-rel-corr}.
-% Because of this, it is crucial to analyse the precise shape of the class of properties after mapping a target-level property to a source-level property\footnote{Dito for mapping a source-level property to a target-level property.}.
-
-However, the composition is still technically possible and it is the job of the compiler engineers to ensure that the secure compilation pipeline happens in an order that ensures that the mapped security property is the intended one.
-
-\subsection{Secure Compilation Proofs}\label{subsec:seccompproofs}
-
-Our secure compilation proofs rely on backtranslations~\cite{abate2019jour,patrignani2021rsc}, which let one construct a source context starting from either target traces (aka trace-based backtranslations) or target contexts (aka context-based backtranslations).
-These backtranslations also require setting up cross-language relations between various language elements such as expressions and program states, so we leave these details for the technical reports.
-% , the usual strategy is to establish a backwards simulation between source components and their compiled target counterpart. 
-% For the universally quantified target context, a technique known as backtranslation is well-established, which constructs a source-level context either by backtranslating the target-level context, amounting to a ``decompilation'', or by constructing it using the target-level trace, so that the source-context emits parts of the whole execution trace that are related to the parts emitted by the target-level context.
-All backtranslations in the case-study are trace based except for those required by $\ccdce$ and $\cccf$, which are context-based (and they are an identity function). 
-% Since these optimisations have no significant semantic effect and source- and target-language are equal, the backtranslation amounts to an identity translation.
-
-\section{Related Work\pages{2}}\label{sec:relwork}
-
-This section discusses robust compilation, other secure compilation criteria and work related to the properties preserved in the case study.
-% (\Cref{subsec:relw:seccomprtp}) 
-% (\Cref{subsec:relw:seccompcrit}).
-% Since the case study of \Cref{sec:casestud:defs,sec:casestud:rtp} implements measures for preserving \gls*{ms}, \gls*{cct}, and \gls*{ss}, 
-% Then, this section discusses work related 
-% relevant related work as well 
-% (\Cref{subsec:relw:msmechs,subsec:relw:cctmechs,subsec:relw:ssmechs}).
-
-\paragraph*{Secure Compilation as Robust Preservation}\label{subsec:relw:seccomprtp}
-
-The robust preservation of properties as a compiler-level criterion has been analysed extensively~\cite{abate2019jour,patrignani2021rsc,abate2021extacc,patrignani2019survey} and thus we build on that framework.
-No existing work is concerned with composing secure compilers, however, existing work~\cite{abate2021extacc} sketches composition of trace-relating compiler correctness in a similar way to what has been presented here.
-%Parts of these works consider languages with different trace models and our technical setup can handle this. 
-The work relating robust preservation with universal composability~\cite{patrignani2022universal} is closest to what this paper presents.
-The authors demonstrate a similar compositionality theorem to what is presented here (\Cref{sec:sequential}) but use it in the context of protocols.
-The work does not consider the generality to support different trace models or composition of compilers which robustly preserve different classes.
-% The authors demonstrate a similar compositionality theorem to what is presented here (\Cref{sec:sequential}) as well as in an earlier version of this work~\cite{kruse2022csc}.
-% However, they do not demonstrate the scalability of the approach by means of a case study.
-
-%\paragraph*{Composition of Secure Compilers}\label{subsec:relw:seccompcomp}
-
-There is work on lifting exploits for single compilers to the whole chain~\cite{paykin2019weird}.
-%Our framework does not consider exploits.
-%\MP{ a bit dry and short}
-While that work considers {\em in}secure compilation and composition thereof in terms of exploits, the composition they are interested in allows to lift an exploit for one compiler pass to the whole compilation chain.
-Our framework takes the opposite direction and provides compositionality results for secure compilers.
-%
-%
-%To the best of our knowledge, there does not exist published work discussing the compositionality of the robust preservation criterion, the gold-standard of secure compilation.
-
-\paragraph*{Other Secure Compilation Criteria}\label{subsec:relw:seccompcrit}
-
-While this paper focuses on the robust preservation framework~\cite{abate2019jour}, other secure compilation criteria exist.
-The survey on formal approaches to secure compilation~\cite{patrignani2019survey} discusses a broad spectrum already, while this section presents a very high-level overview.
-Fully abstract compilation~\cite{abadi1999fullabstraction} states that a compiler should preserve and reflect observational equivalence between source and target programs.
-Abate \emph{et al.}~\cite{abate2021faandrc} showed that fully abstract compilers robustly preserve program properties that are either trivial or meaningless.
-As a mitigation for this, the authors presented a categorical approach based on maps of distributive laws~\cite{watanabe2002modl}, which they call many maps of distributive laws.
-Maps of distributive laws have been investigated before as a possible secure compilation criterion~\cite{tsampas2020catsc}.
-Other approaches are extensions of the compiler correctness criterion as discussed in other work~\cite{patterson2019next700} or the introduction of opaque observations~\cite{vu2021reconciling} to reconcile compiler optimisations with security.
-Note that this work also presents secure compilers that are optimising, but contrary to the other~\cite{vu2021reconciling}, provides a formal account of these in the robust preservation framework.
-% Lastly, the authors of this paper have presented ongoing work~\cite{patrignani2023blame} on a weaker robust preservation criteria based on the concept blame.
-
-\paragraph*{Memory Safety Mechanisms}\label{subsec:relw:msmechs}
-
-Different mechanisms for enforcing memory safety exist that also consider the secure compilation domain, i.e., have an active attacker model.
-For example, the ``pointers as capabilities'' principle represents pointers as machine-level capabilities~\cite{korashy2021capableptrs}, which behave in a similar fashion to capabilities by means of linear typing~\cite{morrisett2005L3}.
-The approach of this paper also uses linear typing, but differs from $L^{3}$~\cite{morrisett2005L3} in the way that functions are not first-class.
-Moreover, this paper considers an active attacker, while the work on $L^{3}$ only discusses whole programs and, thus, has no active attacker model.
-The instrumentation to ensure memory safety that this paper presents is inspired by Softbounds~\cite{nagarakatte2009soft}.
-That work inserts bounds-checks in front of pointer-dereferences and, for this to work, inserts meta-data information on pointer creation.
-Softbounds also works in a more advanced setting with structured fields accesses and also introduces a table-lookup for pointers that are stored in memory.
-This paper only considers arrays of primitive data, i.e., there are no pointers to pointers or structures.
-Several other approaches to memory-safety exist in literature, specifically as compiler instrumentations~\cite{akritidis2009baggy,younan2010paricheck,jung2021pico,shankaranarayana2023tailcheck,dhumbumroong2020boundwarden,nam2019framer,zhou2023fatptrs}, hardware-extensions~\cite{kwon2013lowfat,saileshwar2022heapcheck,chen2023flexpointer,kim2023whistle}, or programming language extensions~\cite{elliott2018checkedc,li2022formalcheckedc,jim2002cyclone,elliott2015guilt,west2005cuckoo,weis2019fyr,benoit2019uniqueness}.
-What differentiates this work from them is that this work uses known, compiler-based approaches to ensure memory-safety as a means to investigate secure compiler compositions.
-This paper does not provide efficient memory-safety, but serves as a theoretical foundation for the secure compilation domain.
-
-To extend the languages in this paper with a less restricted form of pointer arithmetic, the region colouring memory safety monitor presented in earlier work~\cite{michael2023mswasm} can be used.
-The work presenting this monitor provides an approach for the robust preservation of memory safety compiling from C to WASM.
-However, they do not discuss composition of secure compilers but rather investigate an instance of a secure compiler.
-
-\paragraph*{Cryptographic Constant Time Mechanisms}\label{subsec:relw:cctmechs}
-
-The approach to preserving cryptographic constant time in this paper is high-level, where a programming language exposes a way to switch the semantics to a data (operand) independent timing mode.
-Since identifiers in $\Lscct$ are annotated with a secrecy tag, this approach is similar to others with information flow control.
-For example, Vale~\cite{bond2017vale} uses Dafny to ensure constant-time assembly code, while Jasmin~\cite{almeida2017jasmin} makes use of the Coq proof assistant to reject non-constant-time programs.
-CT-Wasm~\cite{watt2019ctwasm} enforces constant-timeness by means of a type system.
-Different to the approach of this paper, these approaches necessitate that the programmer writes \gls*{cct} code.
-An approach to allow programmers to write more high-level code is CryptOpt~\cite{kuepper2023cryptopt}, which generates efficient target-code by means of a randomised search.
-This paper abstracts over concrete mitigation strategies and simply assumes that there is a flag to switch to a cryptographic-constant time execution mode.
-This can be realised by employing the FaCT~\cite{cauligi2019fact} compiler, which translates common non-constant time code patterns to be constant-time, and the data (object) independent timing execution mode of modern processors.
-
-\paragraph*{Speculation Safety Mechanisms}\label{subsec:relw:ssmechs}
-
-This paper uses a taint-tracking mechanism inspired by existing work~\cite{guarnieri2018spectector,fabian2022automatic}.
-These taints are used to express absence of any speculative leaks in \gls*{ss}~\cite{guarnieri2018spectector}. 
-The semantics of $\Lspec$ hardcodes the kind of speculative leaks to just SPECTRE-PHT~\cite{kocher2019spectre}, but future work could use semantics composition~\cite{fabian2022automatic} to support more variants.
-Note that our framework composes compilers and not semantics.
-
-%This work has taken inspiration in the sematics of $\Lspec$ and in \Cref{def:trace:ss} from recent formalisations~\cite{guarnieri2018spectector,fabian2022automatic}.
-% \MPin{
-%  we take SS from exorcising.
-%  we do not take SNI from spectector because it is a hypersafety.
-%  other work defined similar props. 
-% }
-% A subsequent line of work~\cite{fabian2022automatic} shows how to modularly define and compose formal semantics that models different speculative execution variants. 
-% Recent work~\cite{fabian2024lift} demonstrates how to lift existing secure compilation proofs for security properties related to speculative execution to more powerful properties, i.e., properties defined for languages with a trace model that enables more vulnerabilities, such as combinations of different speculative execution exploits.
-% To this end, they define a lifting framework and demonstrate how a secure compilation proof for a weaker property can be lifted to a stronger one. 
-% Contrary to the compositionality framework presented in this paper, these works are only concerned with composing semantics of the respective programming languages, while in this paper, the secure compilers themselves are composed.
-
-\section{Conclusion\pages{$\sfrac{1}{2}$}}\label{sec:concl}
-This paper tackles the problem of understanding what kind of security properties a secure compiler preserves, when said compiler is the combination of compiler passes that preserve possibly different security properties.
-% 
-%For this, this paper first formalised security properties of interest and their composition.
-% 
-The paper proves that composing secure compilers that preserve certain properties results in a secure compiler that preserves the composition of these properties.
-% While the presented security property composition relied on monitors that check only trace-properties, the composition of secure compilers does not make any restriction towards the kind of property involved in the composition.
-% It is subject to future work to develop techniques to verify relational hyperproperties~\cite{abate2019jour}, while the composition of hyperproperties could be very similar to the composition of ordinary properties as presented in this paper, since hyperproperties can be checked with an automata-based model-checker~\cite{beutner23hyperltl}.
-% 
-Finally, this paper defines a multi-pass compiler and proves that it preserves \gls*{specms}.
-Crucially, this paper derives the security of the multi-pass compiler from the composition of the security properties preserved by its individual passes, which include security-preserving as well as optimisation passes.
-% For future work, it is interesting to look at more sophisticated optimisation passes that, e.g., reorder events that appear on traces.
-
-\newpage
-
-\bibliographystyle{IEEEtranS}
-\bibliography{main}
-
-%%
-%% If your work has an appendix, this is the place to put it.
-\appendix
-\subsection{Existential Image}\label{subsec:extimg}
-
-\begin{definition}[Existential Image]
-% [Existential and Universal Image]
-\label{def:existential:img}\label{def:tau}
-  \[
-    \tau_\sim\left(\src{\pi}\right) := 
-      \left\{ 
-        \trg{\trace} \mid \exists \src{\trace}\ldotp \src{\trace}\sim\trg{\trace}, \text{ and } \src{\trace}\in\src{\pi} 
-      \right\}
-  \]
-  %\[ 
-  %  \sigma_\sim\left(\trg{\pi}\right) := 
-  %    \left\{ 
-  %      \src{\trace} \mid \forall \trg{\trace}\ldotp \text{if }\src{\trace}\sim\trg{\trace}, \text{ then } \trg{\trace}\in\trg{\pi} 
-  %    \right\}
-  %\]
-\end{definition}
-
-\begin{definition}[Robust Preservation with $\tau_\sim$]\label{def:rtp:tau}
-  % $\;$\\
-  % \vspace{-1em}
-  % \begin{nscenter}
-  % Compiler $\cc{\src{L}}{\trg{L}}$ robustly preserves $\class$, 
-  {$\rtptau{\cc{\src{L}}{\trg{L}}}{\src{\class}}{\sim}$}
-  %, iff 
-  $\isdef$
-    {$\forall \src{\pi}\in\src{\class}, \src{p}\in\src{L},$} if {$\rsat{\src{\progvar}}{\src{\pi}}$}, then {$\rsat{\cc{\src{L}}{\trg{L}}\left(\src{p}\right)}{\tau_\sim\left(\src{\pi}\right)}$}.
-  % \end{nscenter}
-\end{definition}
-
-\begin{theorem}[Composition of Secure Compilers w.r.t. $\tau$]\label{thm:rtpsim:tau}
-  $\;$
-
-  If {$\rtptau{\cc{\src{L}}{\trg{L}}}{\src{\class[_{1}]}}{\sim_1}$}, {$\rtptau{\cc{\trg{L}}{\obj{L}}}{\tilde{\tau}_{\sim_1}\left(\src{\class[_2]}\right)}{\sim_2}$}, and {$\wfctau{\sim_1}{\src{\class[_2]}}$}, \\ then {$\rtptau{\cc{\src{L}}{\trg{L}}\circ\cc{\trg{L}}{\obj{L}}}{\src{\class[_{1}]}\cap\src{\class[_{2}]}}{\sim_1\bullet\sim_2}$}.
-\Coqed
-\end{theorem}
-
-\begin{corollary}[Swapping Secure Compiler Passes]\label{corr:swappable:tau}
-  $\;$
-
-  If {$\rtptau{\cc[_{1}]{\trg{L}}{\trg{L}}}{\src{\class[_{1}]}}{\sim_1}$ and $\rtptau{\cc[_{2}]{\trg{L}}{\trg{L}}}{\src{\class[_{2}]}}{\sim_2}$}, %
-  {$\wfctau{\sim_1}{\src{\class[_2]}}$ and $\wfctau{\sim_2}{\src{\class[_1]}}$}, %
-  and {$\tilde{\tau}_{\sim_1}\left(\src{\class[_2]}\right)=\src{\class[_2]}$ as well as $\tilde{\tau}_{\sim_2}\left(\src{\class[_1]}\right)=\src{\class[_1]}$},
-  then {$\rtptau{\cc[_{1}]{\trg{L}}{\trg{L}}\circ\cc[_{2}]{\trg{L}}{\trg{L}}}{\src{\class[_{1}]}\cap\src{\class[_{2}]}}{\sim_1\circ\sim_2}$ and $\rtptau{\cc[_{2}]{\trg{L}}{\trg{L}}\circ\cc[_{1}]{\trg{L}}{\trg{L}}}{\src{\class[_{2}]}\cap\src{\class[_{1}]}}{\sim_2\circ\sim_1}$}.
-\Coqed
-\end{corollary}
-
-\subsection{Secure Upper and Lower Composition}\label{sec:other-compos}
-Besides sequential composition, there are two other compositions, namely an {\em upper}, i.e., a compiler that takes multiple inputs and yields one output, and a {\em lower} composition, i.e., a compiler that takes one input and yields multiple outputs.
-We {define the upper composition $\cc{\src{L}+\obj{L}}{\trg{L}}$} as follows:
-Given a program \texttt{p}, its compiled counterpart is obtained by {plugging \texttt{p} into $\cc{\src{L}}{\trg{L}}$ if $\texttt{p}\in\src{L}$} or by {plugging \texttt{p} into $\cc{\obj{L}}{\trg{L}}$ if $\texttt{p}\in\obj{L}$}.
-\begin{definition}[Upper Composition]
-  \[
-    \text{{$\cc{\src{L}+\obj{L}}{\trg{L}}$}}\isdef
-  % Given \texttt{p}, yield
-  \lambda \texttt{p}\ldotp
-  \left\{\mbox{\begin{tabular}{c}
-    {{if $\texttt{p}\in\src{L}$, then $\cc{\src{L}}{\trg{L}}(\texttt{p})$}} \\
-    \mbox{{if $\texttt{p}\in\obj{L}$, then $\cc{\obj{L}}{\trg{L}}(\texttt{p})$}} \\
-  \end{tabular}}\right. 
-  \]
-\end{definition}
-
-Examples of this are present in industry:
-Consider the Java Virtual Machine bytecode $\trg{JVM BC}$, which is a popular target for programming language designers due to its high performance and relevance in industry.
-Compilers for several programming languages have it as their target language, some popular instances are $\src{Java}$ and $\obj{Kotlin}$.
-Technically speaking, they both compile to class files and $\obj{Kotlin}$ objects are considered to be the same as $\src{Java}$ objects at that point.
-Both languages can be used at the same time in one project~\cite{androidstudio}.
-A compiler that accepts both $\src{Java}$ and $\obj{Kotlin}$ code translating to the same target language or intermediate representation performs a kind of {\em upper} composition.
-Now, the following theorem tells us what happens if these are secure:
-Given {$\cc{\src{L}}{\trg{L}}$ robustly preserves $\class[_{1}]$} and {$\cc{\obj{L}}{\trg{L}}$ robustly preserves $\class[_{2}]$}, it follows that {their upper composition $\cc{\src{L}+\obj{L}}{\trg{L}}$ robustly preserves the intersection of classes $\class[_{1}]$ and $\class[_{2}]$}.
-
-\begin{theorem}[Upper Composition of Secure Compilers]\label{thm:urtp}
-  $\;$
-
-  If {$\rtp{\cc{\src{L}}{\trg{L}}}{\class[_{1}]}$} and {$\rtp{\cc{\obj{L}}{\trg{L}}}{\class[_{2}]}$}, then {$\rtp{\cc{\src{L}+\obj{L}}{\trg{L}}}{\class[_{1}]\cap\class[_{2}]}$}. %\Coqed
-\end{theorem}
-
-Dually, the {\em lower} composition is concerned about compilers that accept the same source but yield different target languages. %and the same theoretical results apply.
-{Define the lower composition $\cc{\src{L}}{\trg{L}+\obj{L}}$} as follows:
-Given a program $\src{p}$, its compiled counterpart is obtained by {plugging $\src{p}$ into $\cc{\src{L}}{\trg{L}}$} or by {plugging \texttt{p} into $\cc{\src{L}}{\obj{L}}$}, respectively, {based on the internal decision}.
-\begin{definition}[Lower Composition]
-  \[
-    \text{{$\cc{\src{L}}{\trg{L}+\obj{L}}$}}\isdef
-  % Given $L$ and $\src{p}$, yield
-  \lambda \src{p}, L\ldotp
-  \left\{\mbox{\begin{tabular}{c}
-    {{if $L=\trg{L}$, then} {$\cc{\src{L}}{\trg{L}}(\src{p})$}} \\
-    \mbox{{if $L=\obj{L}$, then} {$\cc{\src{L}}{\obj{L}}(\src{p})$}} \\
-  \end{tabular}}\right.\]
-\end{definition}
-
- Consider two compilers both accepting $\src{LLVM IR}$~\cite{lattner2004llvm} and one of them emits $\trg{x86\_64}$, while the other emits $\obj{ARMv8}$.
- It is intuitive that they are in some sense composed in the LLVM framework, but the decision of when to use one over the other is inherently {\em internal} to the formalisation effort of this kind of composition.
- For example, the user of this compiler provides an explicit flag that instructs to emit $\trg{x86\_64}$ or the framework itself detects the target platform via heuristics, such as supported instructions.
-
- The following theorem demonstrates what happens if the involved compilers are secure:
-Given {$\cc{\src{L}}{\trg{L}}$ robustly preserves $\class[_{1}]$} and {$\cc{\src{L}}{\obj{L}}$ robustly preserves $\class[_{2}]$}, it follows that {their lower composition $\cc{\src{L}}{\trg{L}+\obj{L}}$ robustly preserves the intersection of classes $\class[_{1}]$ and $\class[_{2}]$}.
-
-\begin{theorem}[Lower Composition of Secure Compilers]\label{thm:lrtp}
-  $\;$
-
-  If {$\rtp{\cc{\src{L}}{\trg{L}}}{\class[_{1}]}$} and {$\rtp{\cc{\obj{L}}{\trg{L}}}{\class[_{2}]}$}, then {$\rtp{\cc{\src{L}}{\trg{L}+\obj{L}}}{\class[_{1}]\cap\class[_{2}]}$}. % \Coqed
-\end{theorem}
-
-Either way, the theoretical results suggest that it is possible to always find a ``most-general'', secure compiler, given two secure compilers, that robustly preserves the least-upper bound of the classes involved in their compilation process.
-
-%Future work intends to use upper and lower composition as means to build a secure compiler incrementally
-%This follows current non-formal-methods based compiler engineering, such as LLVM~\cite{lattner2004llvm}, where languages such as \texttt{C++} and \texttt{Fortran} translate into LLVM IR and languages such as \texttt{x86} or \texttt{ARMv8} assembly variants translate out of LLVM IR.
-
-% no coq, because above definitions are non-formal. 
-%   How would you precisely define the lower compo compiler, given two compilers?
-
-%\section{Cryptographic Constant Time}\label{sec:cct}
-
-\end{document}
-\endinput

### 2025-02-05 icfp/sections/bg/hprop.tex
Move Sections around

\subsection{Hyperproperties and (Robust) Satisfaction}

Hyperproperties~\cite{clarkson2008hyper} are strictly more expressive than ordinary properties. 
While a property is a set of traces, a hyperproperty is a set of properties. 
We can lift properties to hyperproperties by applying the powerset.
Hyperproperty satisfaction is defined as follows.

It turns out that classes of hyperproperties, that is, sets of hyperproperties, are equally expressive than simply hyperproperties themselves~\cite{clarkson2008hyper}.

% and so on

### 2025-02-05 icfp/sections/bg/sat.tex
Move Sections around

--- 2024-09-24 csf2/main.tex
@@
-% Fixing: Too many math alphabets used in version normal
-\newcommand\hmmax{0}
-\newcommand\bmmax{0}
-\documentclass[dvipsnames,conference]{IEEEtran}
-
-\usepackage[T1]{fontenc}
-\usepackage[scaled=.83]{beramono}
-
-\usepackage[colorinlistoftodos]{todonotes}
-\usepackage[inference]{semantic}
-\usepackage[switch]{lineno}
-\usepackage{halloweenmath}
-\usepackage{fontawesome5}
-\usepackage{listofitems}
-\usepackage{breakcites}
-\usepackage{glossaries}
-\usepackage{hyperref}
-\usepackage{cleveref}
-\usepackage{stmaryrd}
-\usepackage{marvosym}
-\usepackage{listings}
-\usepackage{amssymb}
-\usepackage{amsmath}
-\usepackage{nameref}
-\usepackage{amsthm}
-\usepackage{xspace}
-\usepackage{xfrac}
-\usepackage{tikz}
-\usepackage{soul}
-\usepackage{bm}
-\usepackage{paralist}
-
-\hypersetup{
-    colorlinks,
-    linkcolor={red!50!black},
-    citecolor={blue!50!black},
-    urlcolor={blue!80!black}
-}
-%\renewcommand\UrlFont{\color{blue}\rmfamily}
-%\newcommand{\url}[1]{\lstinline{#1}}
-
-\setul{0.95ex}{0.3ex}
-
-\input{cmds.tex}
-
-% allow page breaks inside multiline alignment displays
-\allowdisplaybreaks
-
-\Crefname{exampleenv}{Example}{Examples}
-
-\theoremstyle{definition}
-\newtheorem{exampleenv}{Example}[section]
-\newtheorem{lemma}{Lemma}[section]
-\newtheorem{theorem}{Theorem}[section]
-\newtheorem{corollary}{Corollary}[section]
-\newtheorem{definition}{Definition}[section]
-
-\loadglsentries{acronyms}
-\makeglossaries
-
-\AtBeginDocument{%
-  \providecommand\BibTeX{{%
-    \normalfont B\kern-0.5em{\scshape i\kern-0.25em b}\kern-0.8em\TeX}}}
-\makeatletter
-\newcommand{\linebreakand}{%
-  \end{@IEEEauthorhalign}
-  \hfill\mbox{}\par
-  \mbox{}\hfill\begin{@IEEEauthorhalign}
-}
-\makeatother
-
-\renewcommand\thelinenumber{\color{red}\arabic{linenumber}}
-\begin{document}
-\linenumbers
-
-\title{
-  Secure Composition 
-  \\ 
-  of Robust and Optimising Compilers
-  % \vspace{-1.5em}
-}
-
-% \author{
-% Matthis Kruse\orcidID{0000-0003-4062-9666}
-% \and
-% Michael Backes
-% \and
-% Marco Patrignani
-% }
-% \authorrunning{M. Kruse et al.}
-
-\maketitle
-
-\begin{abstract}
-% \MPpost{
-%   this abstract fails in setting up the robustness argument.
-%   it'll be a good exercise for you to take inspiration from this, but then write something as concise but with a more direct focus on robust properties
-% }
-% context
-To ensure that secure applications do not leak their secrets, they are required to uphold several security properties such as spatial and temporal memory safety, cryptographic constant time, as well as speculative safety.
-% need
-Existing work shows how to enforce these properties individually, in an architecture-independent way, by using secure compiler passes that each focus on an individual property.
-% task
-Unfortunately, given two secure compiler passes that each preserve a possibly different security property, it is unclear what kind of security property is preserved by the composition of those secure compiler passes.
-%there is no way to tell what kind of security property will the composition of those secure compilers preserve.
-
-% object
-This paper is the first to study what security properties are preserved across the composition of different secure compiler passes.
-% findings
-Starting from a general theory of property composition for security-relevant properties (such as the aforementioned ones), this paper formalises a theory of composition of secure compilers.
-Then, it showcases this theory on a secure multi-pass compiler that preserves the aforementioned security-relevant properties.
-% conclusion
-Crucially, this paper derives the security of the multi-pass compiler from the composition of the security properties preserved by its individual passes, which include security-preserving as well as optimisation passes.
-% 
-From an engineering perspective, this is the desirable approach to building secure compilers.
-\begin{center}\small\it
-	{This paper uses syntax highlighting accessible to both colourblind and black \& white readers.
-	% ~\cite{patrignani2020use}.
-% 	% Specifically, it makes use of a $\src{blue}$, $\src{sans\text{-}serif}$ font for a $\src{source}$,
-% 	% a $\trg{red}$, $\trg{bold}$ font for an $\trg{intermediate}$,
-% 	% and a $\obj{green}$, $\obj{teletype}$ font for a $\obj{target}$ language.	
-	For a better experience, please print or view in colours.
-	}
-\end{center}
-\end{abstract}
-
-\begin{IEEEkeywords}
-  memory-safety, cryptographic constant-time, secure compilation, robustness
-\end{IEEEkeywords}
-
-\section{Introduction\pages{4}}\label{sec:introduction}
-
-% \gls*{ms} is a security property that ensures that there are no out-of-bounds accesses as well as no reads or writes to uninitialised memory.
-% Programs that do not attain memory safety, can easily have their private data be leaked to attackers or be corrupted~\cite{lemay2021ccc}.
-% Because of this, \gls*{ms} is a requirement for programs to attain an even stronger security property, namely \gls*{cct}.
-% Programs that attain \gls*{cct} are secure w.r.t.
-timing attacks, and this is the gold standard for secure applications, including cryptographic ones, since timing attacks can expose secrets via a simple performance analysis~\cite{kocher1996timing}.
-% Timing attacks have become more and more relevant recently, especially given the era of cloud-computing~\cite{aviram2010cloudtime,kumar2019cloudsecsurvey}, where programs share the same hardware but are supposed to run in a sandboxed environment that does not leak secrets.
-% Unfortunately, several attacks within such a cloud environment have been demonstrated as viable strategies to extract secrets~\cite{mehmet2015getoff,flowers2022zeroday,atya2019catchme,venkatanathan2015placevul}.
-
-% Even though mitigations for such attacks exist~\cite{bond2017vale,almeida2017jasmin}, some compilers may optimise these mitigations away~\cite{barthe2018sec}.
-% This is of course not what programmers expect and to ensure such mitigations persist in the compiled code, one can use a {\em secure} compilers.
-% More generally, secure compilers should preserve security properties even when the program runs in a hostile context such as the cloud one, i.e., where it interacts with other possibly malicious programs.
-% An example of such secure compilers are those that ascribe to the theory of robust compilation~\cite{abate2019jour,abate2021extacc,patrignani2021rsc}, which takes this interaction between components and potentially malicious contexts into account.
-
-% Unfortunately, it is unclear if and how robust compilation supports compositionality, and compilation passes are typically developed as a standalone module to be plugged into a compilation pipeline and composed with other passes.
-% Different kinds of compilation passes exist, e.g., there are usually several compilation passes that perform optimisations~\cite{lattner2004llvm,googlev8,androidstudio,kuepper2023cryptopt,manjikian1997fusion,wegman1991ccp} or source-code instrumentations~\cite{nagarakatte2009soft,nagarakatte2010cets,akritidis2009baggy,dhumbumroong2020boundwarden,jung2021pico,nam2019framer,shankaranarayana2023tailcheck,younan2010paricheck,zhou2023fatptrs,bond2017vale,almeida2017jasmin,kuepper2023cryptopt,cauligi2019fact} that enforce that the compiled program fulfils a property of interest.
-% Without a framework for composing secure compiler passes, it is not possible to enable separation of concerns, e.g., to have a secure compilation pass that ensures \gls*{ms} that is developed independently of another secure pass for \gls*{cct} as suggested before and as showcased in \Cref{ex:strncpy}.
-% \MPpost{
-%   this intro does a poor job at introducing the robust compilation argument.
-%   perhaps it'd be best to start from robust properties, which motivates the use of robust compilers?
-%   idk
-%   as for the abstract, this is a good starting point for you to make a better intro
-% }
-
-\gls*{ms} is a security property obtained by composing \gls*{sms}, which ensures array accesses are all within bounds, and \gls*{tms}, which ensures pointers are only used when they are valid~\cite{azevedo2018meaningsofms,jim2002cyclone,necula2005ccured,nagarakatte2010cets,nagarakatte2009soft,akritidis2009baggy,michael2023mswasm}.
-\gls*{cct} is a security property that ensures sensitive data is not leaked via timing side-channels~\cite{kocher1996timing} and \gls*{ss} is a security property that enforces the same but under a speculative semantics~\cite{guarnieri2018spectector,fabian2022automatic} that captures speculative execution attacks such as spectre~\cite{kocher2019spectre}.
-Together, \gls*{sms}, \gls*{tms}, \gls*{cct} and \gls*{ss}, yield \gls*{specms}, which is the gold standard of security properties for secure applications.
-Programs attaining \gls*{specms} do not leak sensitive data either through erroneous memory accesses, nor through timing side-channels, even under speculative execution.
-\Cref{ex:strncpy} below discusses how these security properties can be enforced by compiler passes~\cite{bond2017vale,almeida2017jasmin}, to ensure programmers need not be aware of the architectural details of where their code will run.
-
-\begin{exampleenv}[strncpy]\label{ex:strncpy}
-Consider the following \texttt{C} function \texttt{strncpy} that copies a null-terminated string \texttt{src} into \texttt{dst} up to a length of \texttt{n} characters.
-This function is subject to a subtle \gls*{sms} vulnerability: the bounds check \texttt{i < n} should happen {\it before} the access to memory location \texttt{src[i]}: otherwise
-% If \texttt{n} is at least as large as the length of the array \texttt{src}, 
-the memory location past the last element is leaked to an attacker.
-\begin{lstlisting}[language=c,basicstyle=\small\ttfamily,morekeywords={size_t}]
-void strncpy(size_t n, char *dst, char *src) {
-  for(size_t i = 0; src[i] != '\0' && i < n; ++i) {
-    dst[i] = src[i];
-  }
-}
-\end{lstlisting}
-% Assume that the length of the array pointed by \texttt{dst} is at least \texttt{n}, and \texttt{src} is terminated by \texttt{'\textbackslash 0'}.
-% Even under these assumptions, there is a 
-To prevent this vulnerability, one can use a compilation pass that enforces \gls*{sms}, such as Softbounds~\cite{nagarakatte2009soft} or BaggyBounds~\cite{akritidis2009baggy}.
-The most na\"ive solution in this case is to insert bounds-checks in front of every access to memory.
-
-Because of timing attacks, addressing \gls*{sms} is not enough to make \texttt{strncpy} secure.
-In fact, the loop can terminate early, as soon as the string-terminating character \texttt{'\textbackslash 0'} is encountered, thus making program execution time proportional to the length of the array pointed by \texttt{src}, and violating the \gls*{cct} property.
-% Often, programmers rewrite programs like these to have loops that run with a fixed number of iterations.
-Also in this case there exist compiler passes that can rewrite such programs into \gls*{cct} ones~\cite{cauligi2019fact}.
- % so that they are no longer vulnerable to timing attacks.
-%
-Finally, even with these precautions, code is not run in isolation, so a malicious attacker could supply code that interacts with \texttt{strncpy} and trigger a violation of either \gls*{ms} or \gls*{cct} by calling \texttt{strncpy} with an argument for \texttt{src} that points to uninitialised memory.
-This would, in turn, trigger a series of reads from uninitialised memory, which is an immediate \gls*{ms} violation with devastating real-world consequences~\cite{uninit-0,uninit-1,uninit-2,uninit-3,uninit-4}.
-% An easy fix to sidestep this is to use non-nullable pointers as parameters instead, e.g., as presented in Cyclone~\cite{jim2002cyclone}.
-\end{exampleenv}
-
-  % last sentence is before the example, which talks about compiler passes enforcing props
-Whether or not compiler passes enforce certain security properties, it is important for a secure compiler to consider partial programs that interact with potentially malicious code, since the latter may lead to, e.g., memory-safety issues in the considered partial program.
-Robust compilers~\cite{abate2019jour} are a form of secure compilers that preserve security properties even in the presence of arbitrary attackers interacting with compiled code.
-Thus, robust compilers can be used to prevent vulnerabilities resulting from uninitialised memory (as well as many other ones), e.g., by targeting capability-based languages such as CHERI~\cite{woodruff2014CHERI}, Arm Morello~\cite{arm-morello}, or MSWasm~\cite{michael2023mswasm}, where the compiler relies on capabilities to check that pointers are always initialised.
-
-% To ensure these compiler-preserved properties still hold even in the presence of arbitrary attackers interacting with compiled code, one can use secure compilers in the form of robust compilers~\cite{abate2019jour,abate2021extacc,patrignani2021rsc}.
-
-Unfortunately, given secure compiler passes that each preserve a possibly different security property, there is no way to tell what kind of security property will the composition of those secure compilers preserve.
-% Different kinds of compilation passes exist, e.g., there are usually several compilation passes that perform optimisations~\cite{lattner2004llvm,googlev8,androidstudio,kuepper2023cryptopt,manjikian1997fusion,wegman1991ccp} or source-code instrumentations~\cite{nagarakatte2009soft,nagarakatte2010cets,akritidis2009baggy,dhumbumroong2020boundwarden,jung2021pico,nam2019framer,shankaranarayana2023tailcheck,younan2010paricheck,zhou2023fatptrs,bond2017vale,almeida2017jasmin,kuepper2023cryptopt,cauligi2019fact} that enforce that the compiled program fulfils a property of interest.
-Worse, without a framework for composing secure compiler passes, it is not possible to enable separation of concerns, e.g., to have a secure compilation pass that ensures \gls*{ms} that is developed independently of another secure pass for \gls*{cct}, that is developed independently of other passes, such as optimisation ones.
-
-This paper introduces a framework for reasoning about the composition of secure and optimising compiler passes %akin to those of \Cref{ex:strncpy}\MP{there aren't really passes there} 
-and it showcases the power of this framework by instantiating it on a multi-pass compilation chain.
-To this end, this paper first discusses how to compose security properties, such as \gls*{tms} and \gls*{sms} into \gls*{ms}, and then adds \gls*{cct} as well as \gls*{ss} into the mix to obtain \gls*{specms}.
-%Then, this paper defines compiler composition and formalises that given two passes that securely preserve two (possibly distinct) properties, their composition securely preserves the composition of those properties.
-The paper then defines several secure compiler passes, where each is either preserving a different security property from the list above
-% (\gls*{tms}, \gls*{sms}, \gls*{scct}, \gls*{ss}) 
-or performing a security-preserving optimisation.
-% , (e.g., applying \gls*{cf} or \gls*{dce}).
-Finally, this paper shows that composing these secure compiler passes into a multi-pass compilation chain results in the end-to-end (robust) preservation of \gls*{specms}.
-Crucially, this paper derives the security of the multi-pass compiler from the composition of the security properties preserved by its individual passes.
-% Crucially, this end-to-end security preservation is obtained by composing the security property preserved by each individual pass, as dictated by the framework of this paper.
-This result showcases how the framework allows the kind of formal security reasoning that compiler writers already want (and already do), obtaining precise, compositional security reasoning while providing minimal (and modular) proof effort.
-%
-In summary, this paper makes the following contributions:
-% \begin{itemize}[leftmargin=0em,nosep,label=$\blacktriangleright$]
-\begin{asparaitem}[$\blacktriangleright$]
-  \item %
-        %\Cref{sec:sequential} studies different forms of secure compiler compositions.
-        This paper takes the secure compilation framework of Abate et al.~\cite{abate2021extacc} and extends it for compositionality. (\Cref{sec:sequential}).
-        %Concretely, it investigates upper and lower compositions, as well as the sequential composition.
-        %For this, this paper studies compiler composition as well as compilers with multiple input languages or multiple output ones, as used in existing compilation chains.
-        %
-        This paper proves that starting from two compilers that preserve two (possibly distinct) properties, their composition preserves roughly the intersection of those properties.
-        % For example, a compiler that is secure w.r.t. \gls*{tms} can be composed with a compiler that is secure w.r.t. \gls*{sms}, resulting in a compiler that is secure w.r.t. \gls*{ms}.
-        % 
-        % From this, the paper proves that the order of composition of sequential compiler passes is irrelevant for security.
-        % % 
-        % This is crucial for reordering individual optimisation passes and thus generating secure and efficient code.
-        %
-        %Finally, this paper extends parts of the compositionality framework to support different trace-models between compilation passes.
-        % \MP{this isn't really true, is it?. the order matters for the relations' lemmas}
-        Then, this paper identifies which conditions make the composition of secure compilers meaningful from a security perspective, and which other conditions allow passes to be swapped without losing security meaning.
-
-  \item %
-        This paper presents a case-study with five programming languages showcasing the previous contribution (\Cref{sec:casestud:defs,sec:casestud:rtp}).
-        To this end, it presents a compilation chain consisting of six passes that ultimately preserve \gls*{specms} by means of composing four secure passes that \emph{individually} preserve \gls*{tms}, \gls*{sms}, \gls*{scct}---an enforceable version of \gls*{cct}---, and \gls*{ss}, and two passes which are well-known optimisations \gls*{dce} and \gls*{cf}.
-        %Furthermore, the chain includes two optimisation passes: one performs \gls*{dce} and the other \gls*{cf}.
-        The formalisation of this case study showcases the power of the presented framework: the divide-and-conquer approach to software engineering is a viable strategy even for the development of secure compilers.
-%        This demonstrates that the usage of the presented framework is both simpler and more expressive compared to one monolithic security proof of the whole compilation chain.
-
-  \item %
-  		This paper provides formal proofs of the security properties preserved by each of the presented passes. 
-  		% 
-  		Additionally this paper analyses each pass in terms of their compatibility with each other (\Cref{sec:formalities}), proving how does each pass fulfill the conditions that make their composition meaningful from a security standpoint.
-        % In doing so, the paper provides formal insights into some technicalities of the concrete proofs necessary as technical setup to reap the benefits of the presented compositionality framework.
-
-  \item The key contributions of this paper are formalised in the Coq proof assistant and the paper indicates this with \CoqSymbol.
-\end{asparaitem}
-% 
-This paper starts by introducing relevant notions of security properties and secure compilation (\Cref{sec:background}),
-and discusses related work (\Cref{sec:relwork}) before concluding (\Cref{sec:concl}).
-
-% \contribution{Technical Report} 
-The omitted formal details, lemmas and proofs, as well as the Coq formalisation are available as supplementary material.
-% \MKin{cite/link}
-
-\section{Background: Properties and Secure Compilers\pages{1}}\label{sec:background}
-
-To introduce the security argument of this paper, this section defines (security) properties, their satisfaction, and their robust satisfaction (i.e., satisfaction in the presence of an active attacker; \Cref{subsec:bg:tprop}).
-Then, borrowing from existing work~\cite{abate2019jour,abate2021extacc}, this section introduces secure compilers as compilers that preserve robust property satisfaction (\Cref{subsec:bg:rtp}).
-
 \subsection{Properties and (Robust) Satisfaction}\label{subsec:bg:tprop}
@@
 % \end{exampleenv}
-
-\subsection{Secure Compilers}\label{subsec:bg:rtp}
-
-A {\em compiler} ($\cc{\src{L}}{\trg{L}}$) translates syntactic descriptions of programs from a {\em source} ($\src{L}$) into a {\em target} ($\trg{L}$) programming language.
-This translation is considered {\em correct} if it is semantics-preserving.
-That is, for a whole program $\src{\wholeprogvar}$, the compiler should relate the $\src{L}$ semantics of $\src{\wholeprogvar}$ with the semantics of $\trg{L}$ of the compiled counterpart of $\src{p}$ in such a way that they are ``compatible''.
-Unfortunately, correct compilers may be insecure compilers~\cite{patrignani2019survey,kennedy2006secure.net,abadi1999protect,ahmed2018dagstuhl} and programs translated by insecure compilers can violate security properties that the programmer assumes to hold.
-This is why {\em robust preservation} is a good candidate as a compiler-level security property~\cite{abate2019jour}.
-
-This paper uses a general notion of robust preservation~\cite{abate2021extacc} that considers compilers that use languages with different trace models. 
-To this end, considering a source trace $\src{\trace}$ and a target trace $\trg{\trace}$, there is a relation ($\sim$) describing the effect of a corresponding compiler (see \Cref{sec:casestud:rtp}). 
-This relation induces the following two projection functions~\cite{abate2021extacc}: (1) the \emph{existential image} $\tau_\sim\left(\src{\pi}\right)$ and (2) the \emph{universal image} $\sigma_\sim\left(\trg{\pi}\right)$.
-These projections map source-level (resp. target-level) properties to target-level (resp. source-level) properties in a way that identifies the ``same'' property across languages. 
-The case study of this paper uses the universal image, since some considered properties, such as \gls*{ss}, are not definable in a higher-level language that, e.g., does not model speculation.
-% For self-containment, the universal image is defined as follows~\cite{abate2021extacc}:
-% 
-\begin{definition}[Universal Image]
-% [Existential and Universal Image]
-\label{def:universal:img}\label{def:sigma}
-  % \[
-  %   \tau_\sim\left(\src{\pi}\right) := 
-  %     \left\{ 
-  %       \trg{\trace} \mid \exists \src{\trace}\ldotp \src{\trace}\sim\trg{\trace}, \text{ and } \src{\trace}\in\src{\pi} 
-  %     \right\}
-  % \]
-  \[ 
-    \sigma_\sim\left(\trg{\pi}\right) := 
-      \left\{ 
-        \src{\trace} \mid \forall \trg{\trace}\ldotp \text{if }\src{\trace}\sim\trg{\trace}, \text{ then } \trg{\trace}\in\trg{\pi} 
-      \right\}
-  \]
-\end{definition}
-\noindent
-With this projection function, we define a more general version of robust preservation as follows~\cite{abate2021extacc}.
-% For a \bul{compiler $\cc{\src{L}}{\trg{L}}$ to robustly preserve a class of source properties $\src{\class}$}, if for any \rul{property $\src{\pi}$ of that class $\src{\class}$ and programs $\src{p}$ written in $\src{L}$} where \iul{$\src{\progvar}$ robustly satisfies $\src{\pi}$}, then \oul{the compilation of $\src{\progvar}$, $\cc{\src{L}}{\trg{L}}\left(\src{p}\right)$, must robustly satisfy $\tau_\sim\left(\src{\pi}\right)$}.
-% 
-% \begin{definition}[Robust Preservation with $\tau_\sim$]\label{def:rtp:tau}
-%   % $\;$\\
-%   % \vspace{-1em}
-%   % \begin{nscenter}
-%   % Compiler $\cc{\src{L}}{\trg{L}}$ robustly preserves $\class$, 
-%   \bul{$\rtptau{\cc{\src{L}}{\trg{L}}}{\src{\class}}{\sim}$}
-%   %, iff 
-%   $\isdef$
-%   \rul{$\forall \src{\pi}\in\src{\class}, \src{p}\in\src{L},$} if \iul{$\rsat{\src{\progvar}}{\src{\pi}}$}, then \oul{$\rsat{\cc{\src{L}}{\trg{L}}\left(\src{p}\right)}{\tau_\sim{\left(\src{\pi}\right)}}$}.
-%   % \end{nscenter}
-% \end{definition}
-% 
-A \bul{compiler $\cc{\src{L}}{\trg{L}}$ robustly preserves a class of target properties $\trg{\class}$}, if for any \rul{property $\trg{\pi}$ of class $\trg{\class}$ and programs $\src{p}$}, where \iul{$\src{\progvar}$ robustly satisfies $\sigma_\sim\left(\trg{\pi}\right)$}, \oul{the compilation of $\src{\progvar}$, we have that $\cc{\src{L}}{\trg{L}}\left(\src{p}\right)$ robustly satisfies $\trg{\pi}$}.
-
-\begin{definition}[Robust Preservation with $\sigma_\sim$]\label{def:rtp:sigma}
-  % $\;$\\
-  % \vspace{-1em}
-  % \begin{nscenter}
-  % Compiler $\cc{\src{L}}{\trg{L}}$ robustly preserves $\class$, 
-  \bul{$\rtpsig{\cc{\src{L}}{\trg{L}}}{\trg{\class}}{\sim}$}
-  %, iff 
-  $\isdef$
-  \rul{$\forall \trg{\pi}\in\trg{\class}, \src{p}\in\src{L},$} if \iul{$\rsat{\src{\progvar}}{\sigma_\sim\left(\trg{\pi}\right)}$}, then \oul{$\rsat{\cc{\src{L}}{\trg{L}}\left(\src{p}\right)}{\trg{\pi}}$}.
-  % \end{nscenter}
-\end{definition}
-Note that a class of properties $\class$ can represent just one property $\pi$ by lifting~\cite{clarkson2008hyper} that property to sets of properties, i.e., use the powerset of $\pi$ instead of $\pi$ itself.
-Because of this, this paper may write $\rtptau{\cc{\src{L}}{\trg{L}}}{\trg{\pi}}{\sim}$, even though $\trg{\pi}$ is a property and not a class.
-A similar construction can be used to the projection functions (see \Cref{def:universal:img}) by applying them to the lifted version of $\trg{\pi}$ instead of just $\trg{\pi}$.
-
-In case the trace model is the same for both source and target programs (and thus $\sim$ is equality), we obtain~\cite{abate2019jour}:
-\begin{definition}[Robust Preservation]\label{def:rtp}
-  % $\;$\\
-  % \vspace{-1em}
-  % \begin{nscenter}
-  % Compiler $\cc{\src{L}}{\trg{L}}$ robustly preserves $\class$, 
-  $\;$
-
-  {$\rtp{\cc{\src{L}}{\trg{L}}}{\class}$}
-  %, iff 
-  $\isdef$
-  {$\forall \pi\in\class, \src{p}\in\src{L},$} if {$\rsat{\src{\progvar}}{\pi}$}, then {$\rsat{\cc{\src{L}}{\trg{L}}\left(\src{p}\right)}{\pi}$}.
-  % \end{nscenter}
-\end{definition}
-
-Examples of compilers fulfilling \Cref{def:rtp} exist in the literature~\cite{korashy2022secureptrs,korashy2021capableptrs,abate2021extacc,abate2019jour,patrignani2021rsc}.
-For example, SecurePtrs~\cite{korashy2022secureptrs} gives a compiler that robustly preserves all safety properties for a C-like language to an assembly-like language. 
-% The mechanization has well over 20k lines in total, which demonstrates that proving a compiler secure with respect to \Cref{def:rtp} is a significant effort. \MP{move this line to later ?}
-% i don't see how adding a def of cct is beneficial
-As another example, even though it is not strictly satisfying \Cref{def:rtp}, the FaCT~\cite{cauligi2019fact} compiler preserves the \gls*{cct} property for a C-like language with constant-time primitives, e.g., \texttt{ctselect} for branching.
-%, whose precise definition can be found in \Cref{sec:cct}.
-Throughout this work, it is assumed that FaCT satisfies \Cref{def:rtp}.
-% While we were not able to find compilers in literature that provably fulfill \Cref{def:rtp}, we anticipate that 
- % would satisfy this hyperproperty.
-% FaCT proves constant-timeness by means of overapproximation, i.e., instead of directly proving the hyperproperty, they fall back to a property that overapproximates the corresponding hyperproperty.
-
-%\smallskip%MP: the underline compacts the sections too much
-
-\section{Secure Composition}\label{sec:sequential}
-
-Notably, real-world compilation chains also perform a series of (sequential) passes whose main purpose is not necessarily to translate from one language to another, but to, e.g., optimise the code or enforce a certain property.
-Both examples can be seen in practice, e.g.,~\cite{nagarakatte2009soft,nagarakatte2010cets,akritidis2009baggy,wegman1991ccp,manjikian1997fusion} and many more.
-% \MPin{survey to cite?
-this is a bit informal MK: hard to find a survey on different compiler passes...
-I can't even find one just for compiler enforced memsafety (that looks good) or for compiler translations in general}.
-% 
-% Consider the compilation passes of LLVM~\cite{lattner2004llvm}, which are handled in the \texttt{PassManager} class.
-% By manipulating the internal state of a \texttt{PassManager} object, the schedule of compiler optimisation passes can be hardcoded.
-% That way, an order of optimisations can be found by hand that yields the best results.
-% \MP{what's the point of the previous 3 sentences?(commented)}
-Consider the following two LLVM optimisation passes: \gls*{cf}, which rewrites constant expressions to the constant they evaluate to, and \gls*{dce}, which removes dead code by rewriting conditional branches.
-The order in which \gls*{cf} and \gls*{dce} are performed influences the final result of the compilation (see \Cref{fig:cfdceex}).
-\begin{figure}[!ht]
-  \begin{center}
-    \begin{tikzpicture}[scale=0.62, every node/.style={transform shape}]
-      \node[minimum width=3cm,minimum height=2.5cm,draw=black,very thick,align=left] (progunopt) {\begin{lstlisting}[language=ml,basicstyle=\small\ttfamily]
-let a = true in
-if a then
-  print "a"
-else
-  print "b"
-\end{lstlisting}};
-      \node[minimum width=3cm,minimum height=2.5cm,draw=black,very thick,align=left,right=1.0 of progunopt] (progoptcf) {\begin{lstlisting}[language=ml,basicstyle=\large\ttfamily]
-if true then
-  print "a"
-else
-  print "b"
-\end{lstlisting}};
-      \node[minimum width=3cm,minimum height=2.5cm,draw=black,very thick,align=left,right=1.0 of progoptcf] (progoptcfdce) {\begin{lstlisting}[language=ml,basicstyle=\Large\ttfamily]
-print "a"
-\end{lstlisting}};% the following is a hack to get my syntax highlighting to work again \end{lstlisting}
-      % edges
-      \draw[->,very thick] (progunopt) edge[loop left,left] node {\gls*{dce}} (progunopt);
-      \draw[->,very thick] (progunopt) edge[sloped,above] node {\gls*{cf}} (progoptcf);
-      \draw[->,very thick] (progoptcf) edge[sloped,above] node {\gls*{dce}} (progoptcfdce);
-    \end{tikzpicture}
-  \end{center}
-  \caption{Example program where the level of optimisations differ for one pass of applying \gls*{cf} and \gls*{dce} in any order. %
-  Every edge is a compilation pass and the label on the edge states what the pass does, i.e., \gls*{cf} or \gls*{dce}. %
-  The source code in the nodes is a glorified compiler intermediate representation and the code gets more optimised towards the right hand side of the figure.}\label{fig:cfdceex}
-\end{figure}
-This {\em phase ordering problem} is well--known in literature and a practical solution is to simply perform a fixpoint iteration of the optimisation pipeline~\cite{click1995combining}.
-
-\smallskip
-
-To analyse the security of compilation passes and their interaction within a compilation pipeline, we rely on a few key notions: the definition of a trace relation, and the definition of when is a trace relation well-formed with respect to a class.
-%
-Consider traces $\src{\trace}$ and $\trg{\trace}$ as well as two trace relations $\sim_1$ and $\sim_2$. 
-% 
-The traces are related $\src{\trace}\sim_1\bullet\sim_2\trg{\trace}$ if there is another trace $\irl{\trace}$ such that $\src{\trace}\sim_1\irl{\trace}$ and $\irl{\trace}\sim_2\trg{\trace}$.
-% 
-\bul{A relation $\sim$ is well-formed w.r.t. a class of target-level properties $\trg{\class}$} iff \rul{the universal image preserves set membership}.
-\begin{definition}[Well-formedness of $\sim$ for a Class $\trg{\class}$]\label{def:wfc:sig:tracerel}
-  % $$
-
-  \begin{nscenter}
-  \noindent
-  \text{\bul{$\wfcsig{\sim}{\trg{\class}}$}} \isdef \text{\rul{$\forall \trg{\pi}\in\trg{\class}, \sigma_\sim(\trg{\pi})\in\sigma_\sim(\trg{\class})$}}
-  \end{nscenter}
-  % $$
-\end{definition}
-
-We can now state our main result: secure compilers in the robust compilation framework~\cite{abate2021extacc} compose {\em sequentially}. 
-%
-
-Let \bul{$\cc{\src{L}}{\trg{L}}$ robustly preserve the class $\sigma_{\sim_2}\left(\trg{\class[_1]}\right)$ under $\sim_1$} and let \rul{$\cc{\trg{L}}{\obj{L}}$ robustly preserve the class $\trg{\class[_2]}$ under $\sim_2$}.
-Then, \iul{when the cross-language relation $\sim_2$ is well-formed w.r.t.
-class $\trg{\class[_1]}$}, it follows that \oul{the composed compiler $\cc{\src{L}}{\trg{L}}\circ\cc{\trg{L}}{\obj{L}}$ robustly preserves the intersection of classes $\trg{\class[_1]}\cap\trg{\class[_2]}$ under $\sim_1\bullet\sim_2$}.
-% 
-%For space reasons, \Thmref{thm:urtp} and \Thmref{thm:lrtp} are not presented here, but can be found in the technical appendix ($\CoqSymbol$).
-% Note that it is not necessary to require $\sim_2$ to be well-formed w.r.t. $\src{\class[_1]}$, since the 
-% \MPin{why}
-\begin{theorem}[Composition of Secure Compilers w.r.t. $\sigma$]\label{thm:rtpsim:sig}
-  $\;$
-
-  If \bul{$\rtpsig{\cc{\src{L}}{\trg{L}}}{\tilde{\sigma}_{\sim_2}\left(\trg{\class[_{1}]}\right)}{\sim_1}$}, \rul{$\rtpsig{\cc{\trg{L}}{\obj{L}}}{\trg{\class[_2]}}{\sim_2}$}, and \iul{$\wfcsig{\sim_2}{\trg{\class[_1]}}$}, \\ then \oul{$\rtpsig{\cc{\src{L}}{\trg{L}}\circ\cc{\trg{L}}{\obj{L}}}{\trg{\class[_{1}]}\cap\trg{\class[_{2}]}}{\sim_1\bullet\sim_2}$}.
-\Coqed
-\end{theorem}
-
-Since the composition of secure compilers is again a secure compiler, the theorems generalise to a whole chain of $n$ secure compilers. 
-% 
-\Cref{thm:rtpsim:sig} can also be stated for the existential image $\tau_\sim\left(\src{\pi}\right)$, but in the interest of space that result has been moved to the appendix (\Cref{subsec:extimg}).
-% 
-Crucially, \Cref{thm:rtpsim:sig} also holds for \emph{classes of hyperproperties}, and thus, compilers that robustly preserve hyperproperties can be composed with each other as well as with compilers that robustly preserve properties.
-% 
-%Unfortunately, proofs of compilers that robustly preserve hyperproperties (including hypersafety) are scarce in the literature due to their complexity.
-%\MP{ so what ? }
-%  we don't discuss this further?
-%
-
-% It is furthermore possible to conflate the notion of classes and hyperproperties by intersecting all elements of a class of hyperproperties, to obtain one combined hyperproperty.
-% This idea has been discussed in other work~\cite{clarkson2008hyper}.
-%  MP: i think i stated the point of these 2 lines above
-
-If we take SecurePtrs and FaCT from \Cref{subsec:bg:rtp} and compose them according to \Cref{thm:rtpsim:sig}, we obtain a compiler that robustly preserves the intersection of safety properties and the \gls*{cct} hyperproperty.
-That is, for a source component that robustly satisfies any set of safety properties {\em and} \gls*{cct}, the compiled target component also robustly satisfies the same set of safety properties and \gls*{cct}.
-% The intersection simply uses the lifted version of the respective safety properties.
-%\MP{meaning?}
-
-Compiler engineers typically try to find an order of optimisations that yields well-optimised programs for either code size~\cite{cooper1999geneticphases} or performance~\cite{kulkarni2006exhaustivephase}.
-\Cref{corr:swappable} justifies that any such order of compilation passes is valid w.r.t. security, as long as the trace-relations have no effect on the respective classes.
-
-So, given two compilation passes \bul{$\cc[_{1}]{\trg{L}}{\trg{L}}$, $\cc[_{2}]{\trg{L}}{\trg{L}}$, both robustly preserving class $\trg{\class[_{1}]}$ or $\trg{\class[_{2}]}$, respectively}, \rul{their corresponding well-formed trace-relations}, \iul{and indifference of the classes with respect to these trace relations}, \oul{for any order of their composition, the composed compiler robustly preserves the intersection of $\trg{\class[_{1}]}$ and $\trg{\class[_{2}]}$}.
-% 
-% \MPin{
-% 	later on, say that optimisation passes remove elements from traces, so they should be fine w.r.t. the result with $\tau$.
-% 	also, yes, if $\tau$ adds we have a problem, if it removes we don't
-% \MKin{ Optimisations that remove elements from traces are not necessarily sound,
-%  e.g., remove all deallocations to "optimise" memory management by offloading
-%  the work to the OS.
-% }
-% }
-
-\begin{corollary}[Swapping Secure Compiler Passes]\label{corr:swappable}
-  $\;$
-
-  If \bul{$\rtpsig{\cc[_{1}]{\trg{L}}{\trg{L}}}{\trg{\class[_{1}]}}{\sim_1}$ and $\rtpsig{\cc[_{2}]{\trg{L}}{\trg{L}}}{\trg{\class[_{2}]}}{\sim_2}$}, %
-  \rul{$\wfcsig{\sim_1}{\trg{\class[_2]}}$ and $\wfcsig{\sim_2}{\trg{\class[_1]}}$}, %
-  and \iul{$\tilde{\sigma}_{\sim_2}\left(\trg{\class[_1]}\right)=\trg{\class[_1]}$ as well as $\tilde{\sigma}_{\sim_1}\left(\trg{\class[_2]}\right)=\trg{\class[_2]}$},
-  then \oul{$\rtpsig{\cc[_{1}]{\trg{L}}{\trg{L}}\circ\cc[_{2}]{\trg{L}}{\trg{L}}}{\trg{\class[_{1}]}\cap\trg{\class[_{2}]}}{\sim_1\circ\sim_2}$ and $\rtpsig{\cc[_{2}]{\trg{L}}{\trg{L}}\circ\cc[_{1}]{\trg{L}}{\trg{L}}}{\trg{\class[_{2}]}\cap\trg{\class[_{1}]}}{\sim_2\circ\sim_1}$}.
-\Coqed
-\end{corollary}
-
-Coming back to the example composing SecurePtrs with FaCT, it is likely the case that \Cref{corr:swappable} is not applicable.
-While first running SecurePtrs and then FaCT should be fine, the other direction has potential security hazards, since the SecurePtrs compiler does not account for cryptographic-constant time primitives, such as \texttt{ctselect}.
-
-\subsection{Secure Compiler Composition with Same Trace Models}
-When the cross-language trace relation is an equality, \Cref{thm:rtpsim:sig} collapses:
-Given \bul{$\cc{\src{L}}{\trg{L}}$ robustly preserves $\class[_{1}]$} and \rul{$\cc{\trg{L}}{\obj{L}}$ robustly preserves $\class[_{2}]$}, it follows that \oul{their sequential composition $\cc{\src{L}}{\trg{L}}\circ\cc{\trg{L}}{\obj{L}}$ robustly preserves the intersection of classes $\class[_{1}]$ and $\class[_{2}]$}.
-
-\begin{corollary}[Composition of Secure Compilers]\label{corr:rtp}
-  $\;$
-
-  If \bul{$\rtp{\cc{\src{L}}{\trg{L}}}{\class[_{1}]}$} and \rul{$\rtp{\cc{\trg{L}}{\obj{L}}}{\class[_{2}]}$}, then \oul{$\rtp{\cc{\src{L}}{\trg{L}}\circ\cc{\trg{L}}{\obj{L}}}{\class[_{1}]\cap\class[_{2}]}$}.
-\Coqed
-\end{corollary}
-
-\Cref{corr:rtp} provides an easy way to compose secure compilers without well-formedness of trace relations. 
-% 
-However, while \Cref{thm:rtpsim:sig} explicitly requires that $\sim_1$ is well-formed w.r.t. $\trg{\class[_2]}$, if $\sim_2$ is not well-formed w.r.t. $\trg{\class[_1]}$, care must be taken. 
-%Worst-case is that the resulting class $\sigma_{\sim_1\bullet\sim_2}\left(\trg{\class[_1]}\cap\trg{\class[_2]}\right)$ becomes empty, even though $\trg{\class[_1]}\cap\trg{\class[_2]}$ is not necessarily empty.
-This is further discussed in \Cref{subsec:compatsecpasses}.
-
-%If possible, optimisations are merged to enhance the analysis' result, since even when iterating to a fixpoint the final program may not be as optimised as it could be~\cite{click1995combining}.
-We can also obtain a specialised version of \Cref{corr:swappable}:
-%So, given two compilation passes \bul{$\cc[_{1}]{\trg{L}}{\trg{L}}$, $\cc[_{2}]{\trg{L}}{\trg{L}}$, both robustly preserving class $\class[_{1}]$ or $\class[_{2}]$, respectively}, \oul{for any order of their composition, the composed compiler robustly preserves the intersection of $\class[_{1}]$ and $\class[_{2}]$}.
-% 
-% \MPin{
-% 	later on, say that optimisation passes remove elements from traces, so they should be fine w.r.t. the result with $\tau$.
-% 	also, yes, if $\tau$ adds we have a problem, if it removes we don't
-% \MKin{ Optimisations that remove elements from traces are not necessarily sound,
-%  e.g., remove all deallocations to "optimise" memory management by offloading
-%  the work to the OS.
-% }
-% }
-
-\begin{corollary}[Swapping Secure Compiler Passes]\label{corr:swappable:one}
-  $\;$
-
-  If {$\rtp{\cc[_{1}]{\trg{L}}{\trg{L}}}{\class[_{1}]}$ and $\rtp{\cc[_{2}]{\trg{L}}{\trg{L}}}{\class[_{2}]}$}, then {$\rtp{\cc[_{1}]{\trg{L}}{\trg{L}}\circ\cc[_{2}]{\trg{L}}{\trg{L}}}{\class[_{1}]\cap\class[_{2}]}$ and $\rtp{\cc[_{2}]{\trg{L}}{\trg{L}}\circ\cc[_{1}]{\trg{L}}{\trg{L}}}{\class[_{2}]\cap\class[_{1}]}$}.
-\Coqed
-\end{corollary}
-
-\section{Security Properties Formalisation \& Composition\pages{1.5}}\label{sec:compprop}
-
-This section introduces trace properties of interest for this paper: \gls*{tms}, \gls*{sms}, \gls*{ms}, \gls*{scct}, and \gls*{ss}.
-These properties are of practical importance (as mentioned in \Cref{sec:introduction}) and also of interest in the case study presented later (\Cref{sec:casestud:defs,sec:casestud:rtp}). 
-This section presents all of them, despite the fact that they are inspired by existing work, in order to showcase all that is required for a formal proof of security for a realistic compilation toolchain.
-%\MP{ say something about this in the rebuttal: only the defs are borrowed, the universal images and all that work is new}
-%\MK{ universal images are not new. }
-% , none of the properties is novel, for self-containment of the paper, their definition is necessary for the case study.
-For each of the presented properties, the technical report defines corresponding monitors, which refine a given property and have a key role in the proofs of this paper. 
-% managed to prove 
-%  if tms-monitor --[trace]-->* ∅  
-%  then trace ∈ tms
-% but not
-%  if trace ∈ tms
-%  then tms-monitor --[trace]-->* ∅
-
-\subsection{A Trace Model for Memory Safety}\label{subsec:basic:memsafety:tracemodel}
-
-For simple memory safety composed of temporal and spatial memory safety, the trace model defines events ($\event[_{\mssafe}]$) as either the empty event ($\emptyevent$), a crash ($\lightning$), or a base-event ($\preevent[_{\mssafe}]$).
-
-\vspace{-1.0em}
-\[
-  \begin{array}{rcll}
-    \text{(Base-Events)}&\preevent[_{\mssafe}] &:=& \ev{Alloc \loc\ n} \mid \ev{Dealloc \loc} \mid \ev{Use \loc\ n} \\
-    \text{(Events)}&\event[_{\mssafe}] &:=& \preevent[_{\mssafe}] \mid \emptyevent \mid \lightning \\ 
-  \end{array}
-\]
-
-Base-events describe the actual kind of event that happened.
-For the basic memory-safety properties, these are three variants:
-First, the allocation event ($\ev{Alloc\ \loc\ n}$) that fires whenever a program claims $n$ cells of memory and stores them at address $\loc$, where addresses are assumed to be unique.
-Second, deallocation ($\ev{Dealloc\ \loc}$) announces that the object at location $\loc$ is freed.
-Third, an event to describe reads from and writes to the $n$-th memory cell from address $\loc$ ($\ev{Use\ \loc\ n}$).
-
-\subsubsection{Temporal Memory Safety}
-
-\gls*{tms}~\cite{nagarakatte2010cets} is a safety property that describes that an unallocated object must not be (re-)used.
-
-\begin{definition}[\glsfirst*{tms}]\label{def:trace:tmsdef}
-  % \footnotesize
-  $$
-  \tmssafe:=\left\{\trace[_{\mssafe}] \left| \begin{array}{rcl}
-    \ev{Alloc\ \loc\ n}&\le_{\trace[_{\mssafe}]}&\ev{Dealloc\ \loc} \\
-    \ev{Use\ \loc\ n}&\le_{\trace[_{\mssafe}]}&\ev{Dealloc\ \loc} \\
-    \text{at most one }\ev{Dealloc\ \loc}&\text{in}&\trace[_{\mssafe}] \\
-    \text{at most one }\ev{Alloc\ \loc\ n}&\text{in}&\trace[_{\mssafe}] \\
-  \end{array}\right.\right\}
-  $$
-\end{definition}
-Hereby, the notation $\event[_{1}]\le_{\trace}\event[_{2}]$ means that if $\event[_{1}]$ is in $\trace$ and if $\event[_{2}]$ is in $\trace$, then $\event[_{1}]$ appears before $\event[_{2}]$.
-
-\subsubsection{Spatial Memory Safety}
-
-\gls*{sms}~\cite{nagarakatte2009soft} is a safety property that prohibits out-of-bounds accesses.
-
-\begin{definition}[\glsfirst*{sms}]\label{def:trace:smsdef}
-% \small
-  % \begin{nscenter}
-
-  \noindent
-  \[
-  \smssafe:=\left\{\trace[_{\mssafe}] \left|\begin{array}{rcl}
-      \text{If }\ev{Alloc\ \loc\ n}\le_{\trace_{\mssafe}}\ev{Use\ \loc\ m}, \text{ then }m<n
-  \end{array}\right.\right\}
-  \]
-  % \end{nscenter}
-\end{definition}
-
-\subsubsection{Memory Safety}
-
-In spirit of earlier work~\cite{nagarakatte2009soft,nagarakatte2010cets,jim2002cyclone,necula2005ccured,michael2023mswasm}, full \gls*{ms} is the intersection of \Cref{def:trace:tmsdef,def:trace:smsdef}.
-
-\begin{definition}[\glsfirst*{ms}]\label{def:trace:msdef}
-  $
-  \mssafe:=\tmssafe \cap \smssafe
-  $
-\end{definition}
-
-Note that \Cref{def:trace:msdef} ignores data isolation, so there may still be memory-safety issues introduced by side-channels.
-
-\subsection{A Trace Model for Memory Safety with Constant Time}\label{subsec:scct:tracemodel}
-
-To express Constant Time, we extend the memory safety trace model with a {\em security tag} ($\securitytag{}$) that indicates whether events contains sensitive information ($\lock$) or not ($\unlock$).
-
-\vspace{-.5em}
-\[
-  \begin{array}{rrcl}
-    (\text{Base-Events}) & \preevent[_{\ctsafe}] &:=& \preevent[_{\mssafe}] \mid \ev{Branch\ }n \mid \ev{Binop\ }n\\
-    (\text{Security Tags}) & \securitytag{} &:=& \lock \mid \unlock\\ 
-    (\text{Events}) & \event[_{\ctsafe}] &:=& \preevent[_{\ctsafe}];\securitytag{} \mid \emptyevent \mid \lightning \\ 
-  \end{array}
-\]
-
-For cryptographic code, there is a general guideline that secrets must not be visible on a trace~\cite{ctguidelines}, i.e., secrets should not be marked as $\unlock$.
-In turn, an instruction whose timing is data-dependent must not have a secret as an operand.
-Typical operations with data-dependent timing are branches and certain binary operations, such as division.%
-\footnote{
-	This is architecture-dependent, but division is an operation that serves as a classic example for a data-dependent timing instruction~\cite[p.~755]{arm-refman}.
-}
-Both operations are represented in the trace model by extending the set of base-events with branches ($\ev{Branch\ n}$) and binary operations ($\ev{Binop\ n}$).
-
-\subsubsection{Strict Cryptographic Constant Time}
-
-\gls*{cct} is a hypersafety property~\cite{barthe2018sec} and, thus, difficult to check with monitors.
-This is because, intuitively, hypersafety properties can relate multiple execution traces with each other, but monitors work on a single execution.
-It is a common trick to sidestep this issue by means of overapproximation: this section defines the property \gls*{scct}, a stricter variant of \gls*{cct} (inspired by earlier work~\cite{almeida2017jasmin}) that enforces the policy that no secret appears on a trace.
-Programs that satisfy \gls*{scct} also satisfy \gls*{cct}, but programs that satisfy \gls*{cct} may not satisfy \gls*{scct}.
-
-\begin{definition}[\glsfirst*{scct}]\label{def:trace:scctdef}
-% \begin{nscenter}
-
-  \noindent\[
-  \scctsafe:=\left\{\trace[_{\ctsafe}] \left|\begin{array}{l}
-      \trace[_{\ctsafe}]=\hole{\cdot} \text{ or } \\\exists\trace[_{\ctsafe}'],\trace[_{\ctsafe}]=\preevent[_{\ctsafe}];\unlock\cdot\trace[_{\ctsafe}'] \wedge \trace[_{\ctsafe}']\in\scctsafe
-    \end{array}\right.\right\}
-  \]
-% \end{nscenter}
-\end{definition}
-
-\gls*{scct} may appear overly strict, since it seems that secrets must not occur on a trace (since $\securitytag{}$ is forced to be $\unlock$). 
-However, this is considered standard practice in terms of coding guidelines~\cite{ctguidelines}.
-Moreover, programs that have been compiled with FaCT~\cite{cauligi2019fact} and run with a ``data independent timing mode''~\cite{arm-refman,intel-refman} enabled do not leak secrets (see \Cref{ex:lscct}).
-
-\subsubsection{\gls*{ms}, Strict Cryptographic Constant Time}\label{sec:msscct-rel}
-
-The combination of \gls*{ms} and \gls*{scct} is the intersection of these properties, \gls*{msscct}.
-However, \gls*{ms} uses a different trace model than \gls*{scct}, so intersecting them would trivially yield the empty set. 
-To remedy this issue, we introduce $\sim_{\ctsafe}: \preevent[_{\ctsafe}] \times \preevent[_{\mssafe}] $, a cross-language trace relation (whose key cases are presented below), that we use to intuitively unify the trace model in which the two properties are expressed:
-\[
-  \typerulenolabel{mscctrel:drop:tag}{}{\preevent[_{\mssafe}]\sim_{\ctsafe}\preevent[_{\mssafe}];\unlock{}}
-  \typerulenolabel{mscctrel:drop:crash}{}{\lightning\sim_{\ctsafe}\lightning}
-\]
-\[
-  \typerulenolabel{mscctrel:drop:branch}{}{\emptyevent\sim_{\ctsafe}\ev{Branch n};\securitytag{}}
-  \typerulenolabel{mscctrel:drop:binop}{}{\emptyevent\sim_{\ctsafe}\ev{Binop n};\securitytag{}}
-\]
-
-Essentially, $\sim_{\ctsafe}$ ignores both the new $\ev{Branch\ n}$ and $\ev{Binop\ n}$ base-events as it relates security-insensitive actions ($\unlock$) to their equivalent counterparts.
-Thus, \gls*{ms} traces trivially satisfy \gls*{scct}.
-% 
-The above relation is extended point-wise to traces, skipping the empty event $\emptyevent$ on either side, and it is now possible to define \gls*{msscct} using the universal image:
-
-\begin{definition}[\gls*{ms} and \gls*{scct}]\label{def:trace:msscctdef}
-  $
-  \msscctsafe:=\mssafe\cap\sigma_{\sim_{\ctsafe}}\left(\scctsafe\right)
-  $
-\end{definition}
-
-\subsection{Extending the Trace Model with Speculation}\label{subsec:msctss:tracemodel}
-
-So far, the considered trace models do not let us express speculative execution attacks such as Spectre~\cite{kocher2019spectre}. 
-For this, we extend the earlier trace model (see \Cref{subsec:scct:tracemodel}) so that the security tags ($\securitytag{}$) carry additional information about the kind of private data leakage, i.e., the type of speculative leak.
-Moreover, we add base-events signalling the beginning of a speculative execution ($\ev{Spec}$), a barrier ($\ev{Barrier}$) that signals that any speculative execution may not go past it, as well as a rollback event ($\ev{Rlb}$), which signals that execution resumes to where speculation started.
-
-\vspace{-1em}
-{
-\[
-  \begin{array}{rrcl}
-    (\text{Base-Events}) & \preevent[_{\mathghost}] &:=& \preevent[_{\ctsafe}] \\
-    (\text{Spectre Variants}) & vX &:=& \operatorname{NONE} \mid \operatorname{PHT} \\
-    (\text{Security Tags}) & \securitytag{} &:=& \lock_{vX} \mid \unlock\\ 
-    (\text{Events}) & \event[_{\mathghost}] &:=& \preevent[_{\mathghost}];\securitytag{} \mid \emptyevent \mid \lightning \mid \ev{Spec} \mid \ev{Rlb} \mid \ev{Barrier}\\ 
-  \end{array}
-\]
-}
-
-Even though the considered Spectre variants are just SPECTRE-PHT~\cite{kocher2019spectre}, NONE just describes secret data as in \gls*{scct} (see \Cref{subsec:scct:tracemodel}), the trace model is general enough to allow for potential future extension with different variants~\cite{kocher2019spectre,maisuradze2018ret2spec,horn2019zero}.
-%For sake of readability, this paper just uses the notation $\lock$ in place of $\lock_{\text{PHT}}$.
-
-\subsubsection{Speculative Safety}
-
-\gls*{ss}~\cite{patrignani2021exorcising}, similar to \gls*{scct}, is a sound overapproximation of a variant of noninterference.
-
-\begin{definition}[\glsfirst*{ss}]\label{def:trace:ss}
-  \noindent
-
-  \begin{nscenter}
-  $
-    \sssafe:=\left\{\trace[_{\mathghost}] \left|\begin{array}{l}
-      \trace[_{\mathghost}]=\hole{\cdot} \text{ or } \exists\trace[_{\mathghost}'].\\
-      \left(\trace[_{\mathghost}]=\preevent[_{\mathghost}];\unlock\cdot\trace[_{\mathghost}'] \text{or }\trace[_{\mathghost}]=\preevent[_{\mathghost}];\lock_{\text{NONE}}\cdot\trace[_{\mathghost}']\right)\\
-      \text{and }\ \trace[_{\mathghost}']\in\sssafe
-                                 \end{array}\right.\right\}
-  $ 
-  \end{nscenter}
-\end{definition}
-The technical setup so far leads to the above definition, where only locks annotated with $\text{SPECTRE-PHT}$ are disallowed to occur on the trace.
-That way, programs attaining \gls*{ss} do not necessarily attain \gls*{scct}.
-
-\subsubsection{Speculation Memory Safety}\label{sec:spec-ms-rel}
-
-As before, we need to relate the different trace models with each other, so that the memory safety property without speculation can be lifted to speculation. 
-To this end, let $\sim_{\mathghost}: \preevent[_{\mathghost}] \times \preevent[_{\ctsafe}]$ be a cross-language trace relation whose key cases are below.
-The intuition is that \gls*{ss} is trivially satisfied in \gls*{scct}, since speculation is inexpressible there, which amounts to dropping events $\ev{Spec}$, $\ev{Rlb}$, or $\ev{Barrier}$, as well as all base events tagged with $\lock_{\text{PHT}}$.
-
-\[
-  \typerulenolabel{specrel:drop:tag}{}{\preevent[_{\ctsafe}];\lock\sim_{\mathghost}\preevent[_{\ctsafe}];\lock_{\text{NONE}}}
-  \typerulenolabel{specrel:drop:spectag}{}{\emptyevent\sim_{\mathghost}\preevent[_{\mathghost}];\lock_{\text{PHT}}}
-\]
-\[
-  \typerulenolabel{specrel:drop:pub}{}{\preevent[_{\ctsafe}];\unlock\sim_{\mathghost}\preevent[_{\ctsafe}];\unlock}
-  \typerulenolabel{specrel:drop:crash}{}{\lightning\sim_{\mathghost}\lightning}
-\]
-\[
-  \typerulenolabel{specrel:drop:spec}{}{\emptyevent\sim_{\mathghost}\ev{Spec}}
-  \typerulenolabel{specrel:drop:rlb}{}{\emptyevent\sim_{\mathghost}\ev{Rlb}}
-  \typerulenolabel{specrel:drop:barrier}{}{\emptyevent\sim_{\mathghost}\ev{Barrier}}
-\]
-
-We conclude by defining the ultimate property of interest for secure compilers: \gls*{specms}.
-\begin{definition}[\gls*{specms}]\label{def:trace:specmsdef}
-  $
-  \specmssafe := \msscctsafe\cap\sigma_{\sim_{\mathghost}}\left(\sssafe\right)
-  $
-\end{definition}
-
-\section{Case Study: Language Formalisations\pages{2}}\label{sec:casestud:defs}
-
-This section defines programming languages $\Ltms$, $\Ltrg$, $\Lms$, $\Lscct$, and $\Lspec$, all of which share common elements (presented in \Cref{subsec:cs:defs}).
-$\Ltms$ is the only statically-typed language, and it exhibits the property that all well-typed programs are \gls*{tms} (\Cref{subsec:ltms}).
-However, not all $\Ltms$ programs are \gls*{sms}.
-That is, there are well-typed $\Ltms$ programs that perform an out-of-bounds access.
-Language $\Ltrg$ is untyped and does not provide any guarantees with regards to \gls*{ms} (\Cref{subsec:lsms}).
-$\Lms$ is exactly the same language as $\Ltrg$, but this paper still distinguishes the two for sake of readability (\Cref{subsec:lms}).
-All three languages --- so $\Ltms$, $\Ltrg$, and $\Lms$ --- assume \gls*{scct} to hold, since this is -- in an ideal world -- what the programmer would expect, too: it is the job of the compiler to preserve and (potentially) enforce \gls*{scct} security~\cite{cauligi2019fact,nagarakatte2010cets,nagarakatte2009soft,akritidis2009baggy}.
-% This paper considers \gls*{cct} as an architecture-dependent property\MPin{can we justify this}, so
-
-Such consideration is also backed up by architecture providing a data (operand) independent timing mode, such as processors by Arm~\cite[p.~543]{arm-refman} and Intel~\cite[p.~80]{intel-refman}.
-%In spirit of this, all languages besides $\Lscct$ are trivially \gls*{scct}. 
-This kind of processor feature is modelled in language $\Lscct$ (\Cref{subsec:lscct}), where programs have access to a ``\gls*{cct}-mode'' and can change the leakage of emitted events according to the value of this mode (either $\obj{ON}$ or $\obj{OFF}$).
-
-Finally, modern processors also employ speculative execution to achieve speedups---and unfortunately generate Spectre attacks~\cite{kocher2019spectre}---and this is the extension of $\Lspec$ (\Cref{subsec:lspec}).
- % extends $\Lscct$ to model speculative execution.
-%
-Thus, all previous languages trivially satisfy \gls*{ss}, since they do not support speculative execution at all.
-% All languages share large parts of syntactic elements, which are presented in the next subsection (\Cref{subsec:cs:defs}).
-% Subsequent subsections extend on this accordingly.
-
-\subsection{Shared Language Definitions}\label{subsec:cs:defs}
-
-All presented programming languages share a common fragment which is partially presented here and in full detail in the technical report. 
-%\MP{put omitted stuff in appendix + ref there (and add textual explanation)}
-
-% \vspace{-.5em}
-{
-  \renewcommand{\src}[1]{\mathsf{#1}}
- \small
-\[
-  \begin{array}{rrcl}
-    \text{(Base-Events)} & \src{\preevent} &:=& \src{\ev{Alloc \loc\ n}} \mid \src{\ev{Dealloc \loc}} \mid \src{\ev{Get \loc\ n}} \mid \src{\ev{Set \loc\ n}} \\
-    \text{(Control Tags)} & \src{\sandboxtag{}} &:=& \src{\ctx} \mid \src{\comp} \\
-    \text{(Events)} & \src{\event} &:=& \src{\preevent;\sandboxtag{}} \mid \src{\emptyevent} \mid \src{\lightning} \\ 
-    \text{(Values)} & \src{v} &:=& \src{n} \\
-    \text{(Expressions)} & \src{\expr} &:=& \src{x} \mid \src{n} \mid \src{\lbinop{\expr[_1]}{\expr[_2]}} \mid \src{\lifz{\expr[_1]}{\expr[_2]}{\expr[_3]}} \\ 
-                         &&&\mid \src{\llet{x}{\expr[_1]}{\expr[_2]}} \mid \src{\llet{x}{new\ \expr[_1]}{\expr[_2]}} \\
-                         &&&\mid \src{\ldelete{x}} \mid \src{\lget{x}{\expr}} \mid \src{\lset{x}{\expr[_1]}{\expr[]}} \mid \src{stuck}\\
-                         &&&\mid \src{\lcall{f}{\expr}} \mid \src{\lreturn{\expr}} \\
-    \text{(Functions)} & \src{F} &:=& (\src{x};\src{\expr}) \\
-    \text{(Libraries)} &&&\hspace{-3.8em} \src{\library_{\ctx}},\src{\library_{\comp}} : \src{Vars} \to \src{F} 
-    % \hspace{2em}
-    %  : \src{Vars} \to \src{F} 
-     \\
-    \text{(Heaps)} & \src{H} &:=& \src{\hole{\cdot}} \mid \src{n},\src{H}
-    % \hspace{0.5cm}    
-    \\
-    \text{(Pointer Maps)} & \src{\Delta} && \text{ omitted for simplicity}
-    \\
-    \text{(Memory States)} & \src{\memstate} &:=& \left(\src{H^{\ctx}};\src{H^{\comp}};\src{\Delta}\right)\\
-    \text{(Control States)} & \src{\cfstate} & & \text{ omitted for simplicity} 
-    \\
-    \text{(States)} &\src{\Omega} &:=& \left(\src{\cfstate};\src{\sandboxtag{}};\src{\memstate}\right)\\
-  \end{array}
-\]
-}
-{
-  \renewcommand{\src}[1]{\mathsf{#1}}
-The trace models of all languages are similar to those presented earlier (\Cref{sec:compprop}).
-One technical detail is the addition of a control tag, indicating who is to blame for an emitted action: context ($\ctx$) or component ($\comp$).
-This is a standard technique in secure compilation in order to rule out irrelevant context-level events. 
-For example, a context immediately deallocating an allocated memory region twice trivially violates memory safety, but the main interest in secure compilation is to preserve component-level security, and thus component events.
-Even though this tagging could be used for blame preservation~\cite{patrignani2023blame}, this is beyond the focus of this paper.
-Another key difference is that memory accesses are now explicitly modelled as reads ($\src{\ev{Read\ \loc\ n}}$) and writes ($\src{\ev{Write\ \loc\ n}}$) instead of just uses.
-
-All languages have at least numbers as values ($\src{v}$) and second class functions ($\src{F}$).
-Functions are modelled as pairs containing the name of one argument and the body of the function.
-Bodies are just ordinary expressions, which can be simple binary operations ($\src{\lbinop{\expr[_1]}{\expr[_2]}}$), conditionals ($\src{\lifz{\expr[_1]}{\expr[_2]}{\expr[_3]}}$), function calls ($\src{\lcall{f}{e}}$) and returns ($\src{\lreturn{\expr}}$), as well as C-like memory operations.
-Programs have sets of pre-determined functions called libraries and they are marked as being part of some component ($\src{\library_{\comp}}$) or context ($\src{\library_{\ctx}}$).
-
-For the operational semantics, the runtime state ($\src{\Omega}$) is a triple consisting of a control-flow state ($\src{\cfstate}$), a control tag ($\src{\sandboxtag{}}$), and a memory state ($\src{\memstate}$).
-The latter carries information about pointers that are kept ``alive'' in pointer maps ($\src{\Delta}$), so that the semantics does not get stuck when encountering, e.g., a double-free.
-The memory state also carries two heaps to model sandboxing between a context ($\src{H^{\ctx}}$) and a component ($\src{H^{\comp}}$).
-
-}
-
-\subsection{$\src{L_{\tmssafe}}$: A Temporal but Not Spatial Memory Safe Language}\label{subsec:ltms}
-
-$\src{L_{\tmssafe}}$ is the only statically-typed language in this case study and restricts functions ($\src{F}$) to the typing signature $\src{\natt\to\natt}$. 
-The type system of $\src{L_{\tmssafe}}$ is inspired by $L^{3}$~\cite{morrisett2005L3,scherer2018fabulous} and enforces that every well-typed $\src{L_{\tmssafe}}$ program satisfies \gls*{tms}.
- % (\Cref{thm:wt:tms}).
-
-\begin{theorem}[$\src{L_{\tmssafe}}$-programs are \gls*{tms}]\label{thm:wt:tms}
-$\rsat{\src{\library_{\comp}}}{\tmssafe}$ \Coqed
-  % $\;$ 
-% 
-  % \begin{nscenter}
-  %   \hfill For any component $\src{\library_{\comp}}$, $\rsat{\src{\library_{\comp}}}{\tmssafe}$ \Coqed
-  % \end{nscenter}
-\end{theorem}
-
-\subsection{$\Ltrg$: A Memory-Unsafe Language}\label{subsec:lsms}
-
-$\Ltrg$ extends the syntax presented earlier (\Cref{subsec:cs:defs}) with dynamic typechecks ($\trg{\lhast{\expr}{\type}}$), evaluating to $\trg{0}$ if the type matches with the shape of $\trg{\expr}$ and $\trg{1}$ otherwise.
-%The pointers carry poison tags, which signify a pointer being allocated ($\trg{\poisoned}$) or freed ($\trg{\poisonless}$). 
-Furthermore, the syntax of $\Ltrg$ is extended with a way to inspect whether a pointer is freed ($\trg{\lispoisoned{x}}$), evaluating to $\trg{0}$ if it is freed and to $\trg{1}$ otherwise.
-Functions may receive arguments that are not $\trg{\natt}$, values are extended with tuples, and expressions are also extended with pair projections.
-
-% \vspace{-1em}
-\[
-  \begin{array}{rrcl}
-    \text{(Values)} & \trg{v} &:=& \dots \mid \trg{\lpair{v_1}{v_2}} \\
-    \text{(Expressions)} & \trg{\expr} &:=& \dots \mid \trg{\lhast{\expr}{\tau}} \mid \trg{\lispoisoned{x}}
-  \end{array}
-\]
-
-No changes are done to the trace model, but $\Ltrg$ has no static typing, making double-free code patterns possible.
-
-\subsection{$\Lms$: Another Memory-Unsafe Language}\label{subsec:lms}
-$\Lms$ is exactly equal to $\Ltrg$ (\Cref{subsec:lsms}), but used to emphasize that this is code after applying $\cc{\Ltrg}{\Lms}$ (\Cref{subsec:cs:ms}).
-
-\subsection{$\Lscct$: A Memory-Unsafe Language with a Constant Time Mode}\label{subsec:lscct}
-
-$\Lscct$ extends $\Lms$ (\Cref{subsec:lms}) with a ,,constant-time mode''. 
-The activation of the mode can be checked ($\obj{\lrddoit{x}{\expr}}$), changed ($\obj{\lwrdoit{D}}$), and is stored in program states ($\obj{\Omega}$).
-At the beginning of program execution of $\Lscct$ programs, the mode is turned off ($\obj{OFF}$).
-If the mode is enabled (i.e., set to $\obj{ON}$), the intuition is that no secrets are leaked. 
-This models the real-world\footnote{As present in Intel~\cite[p.80]{intel-refman} and ARM~\cite[p.~543]{arm-refman} processors.} data-independent timing mode as well as the result of compiling a program with FaCT~\cite{cauligi2019fact}.
-For example, FaCT rewrite code that branches to use a constant-time selection primitive.
-To not obfuscate our formalisation unnecessarily by duplicating syntax, we simply added the ,,constant-time mode'' to $\Lscct$.
-
-% $\Lscct$ extends $\Lms$ (\Cref{subsec:lms}) with a way to read from ($\obj{\lrddoit{x}{\expr}}$) and to write to ($\obj{\lwrdoit{D}}$) a model-specific register that controls a constant-time mode. 
-% This language is a combination of the effect of the FaCT compiler~\cite{cauligi2019fact} as well as a data-independent timing mode, a feature that is present in both Arm~\cite[p.~543]{arm-refman} and Intel~\cite[p.~80]{intel-refman} processors.
-% \MP{ can we be more specific about the fact pass? }
-% To this end, states ($\obj{\Omega}$) are extended with the value of the register ($\obj{D}$), which is initially set to be not active ($\obj{OFF}$).
-% If the register is marked active ($\obj{ON}$), the intuition is that no secrets can appear on specification traces.
-% If the register is marked inactive, secrets may appear on traces.
-% The state of the register can be read and written to, e.g., \Cref{tr:e-wrdoit-off}.
-% For the other languages seen earlier, the mode is intuitively always-on, i.e., the mode of execution always uses constant-time mode.
-
-% \vspace{-.5em}
-\[
-  \begin{array}{rrcl}
-    \text{(Mode Values)} & \obj{D} &:=& \obj{ON} \mid \obj{OFF} \\
-    \text{(Security Tags)} & \obj{\securitytag{}} &:=& \obj{\lock} \mid \obj{\unlock} \\
-    \text{(Base-Events)} & \obj{\preevent} &:=& \dots \mid \obj{iGet\ \loc\ v} \mid \obj{iSet\ \loc\ v} \\
-                       &&& \mid\ \obj{Binop\ n} \mid \obj{Branch\ n} \\
-    \text{(Events)} & \obj{\event} &:=& \obj{\preevent;\sandboxtag{};\securitytag{}} \mid \obj{\emptyevent} \mid \obj{\lightning} \\
-    \text{(Expressions)} & \obj{\expr} &:=& \dots \mid \obj{\lrddoit{x}{\expr}} \mid \obj{\lwrdoit{D}} \\
-                         &&&\mid\ \obj{\llet{x^{\securitytag{}}}{\expr[_1]}{\expr[_2]}}\\
-    \text{(States)} & \obj{\Omega} & := & \left(\obj{\cfstate};\obj{\sandboxtag{}};\obj{D};\obj{\memstate}\right)\\
-  \end{array}
-\]
-\begin{center}
-\newcommand{\expreval}[5]{{#1}\triangleright\xspace {#2}\xrightarrow{#5}\ {#3}\triangleright\xspace {#4}\xspace}
-\newcommand{\exprevalo}[5]{\expreval{\obj{#1}}{\obj{#2}}{\obj{#3}}{\obj{#4}}{\obj{#5}}}
-  \typerule{$e-\obj{wrdoit}-\text{off}$}{
-  }{
-    \pstepto{\rtt{\obj{\cfstate;\sandboxtag{};D;\memstate}}{\obj{\lwrdoit{OFF}}}}{\rtt{\obj{\cfstate;\sandboxtag{};OFF;\memstate}}{\obj{0^{\securitytag{}}}}}{\obj{\emptyevent}}
-  }{e-wrdoit-off}
-
-  \typerule{$e-\obj{get}-\in-$noleak}{
-    \obj{\memstate}=\obj{H^{\ctx};H^{\comp};\Delta_1,x\mapsto(\loc;\sandboxtag{};\poison),\Delta_2} &
-    \obj{\loc}+\obj{n}\in\text{dom }\obj{H^{\sandboxtag{}}} \\
-    \obj{\securitytag{''}}=\obj{\securitytag{}\sqcap\securitytag{'}} &
-    \obj{\event} = \obj{{iGet}\ \loc\ n;\sandboxtag{};\securitytag{''}}
-  }{
-    \exprevalo{
-    	\cfstate;\sandboxtag{'};ON;\memstate
-    }{
-      	x^{\securitytag[n^{\securitytag{'}}] }
-    }{
-    	\cfstate;\sandboxtag{'};ON;\memstate
-    }{
-    	(H^{\sandboxtag{}}(\loc+n))^{\securitytag{''}}
-    }{
-      \event
-    }
-  }{to-e-get-in-noleak}
-\end{center}
-
-The language also adds user annotations $\obj{\securitytag{}}$ for the secrecy of variables, which can be either private (high secrecy) $\obj{\lock}$ or public (low secrecy) $\obj{\unlock}$.
-Security tags $\obj{\securitytag{}}$ are arranged in the usual secrecy lattice~\cite{zdancewicphd}, where $\obj{\unlock}\sqsubseteq\obj{\lock}$.
-% $\obj{\lock}\sqsubseteq\obj{\securitytag{}}$ and $\obj{\securitytag{}}\sqsubseteq\obj{\unlock}$.
-
-Memory accesses to secret data need to be present to reason about memory safety, even when execution is in constant-time mode, e.g, \Cref{tr:to-e-get-in-noleak}.
-$\Lscct$ extends base-events with $\obj{\ev{iGet}\ \loc\ \valueexpr}$ and $\obj{\ev{iSet}\ \loc\ \valueexpr}$ (for data $\obj{i}$ndependent get and set) to prevent secrets from leaking but still enable reasoning about memory safety.
-Due to the technical setup, the rule needs to check if the access is in bounds ($\obj{\loc}+\obj{n}\in\obj{H^t}$) and update the secrecy tag ($\obj{\securitytag{''}}=\obj{\securitytag{}\sqcap\securitytag{'}}$) with the least upper bound of the tags, according to the aforementioned lattice.
-The precise information carried by the event, e.g., location ($\obj{\loc}$) is taken from the pointer map, which carries information irrelevant to this rule ($\obj{\poison}$).
-
-Base-events include $\obj{\ev{Branch}\ n}$ and $\obj{\ev{Binop}\ n}$ that are emitted when evaluating a branch or certain binary expressions, such as division, respectively, whenever the constant-time mode is inactive.
-Events are extended with a security-tag ($\obj{\securitytag{}}$) to signal the secrecy of the involved data.
-
-The evaluation steps are amended to propagate the security-tag annotations $\obj{\securitytag{}}$.
-When the constant-time mode is inactive, base-events $\obj{\ev{Branch}\ n}$ and $\obj{\ev{Binop}\ n}$ are emitted for conditionals and binary operations, respectively.
-Otherwise, just like in the semantics of the earlier languages, $\obj{\emptyevent}$ is emitted for binary and branching operations.
-%To mark this mode as active or inactive, \Cref{tr:e-wrdoit} allows to access the register in $\obj{\statevar}$.
-
-\Cref{ex:lscct} illustrates the differences between $\Lscct$ and other languages.
-
-\begin{exampleenv}[$\Lscct$ with and without constant-time mode]\label{ex:lscct}
-  Consider again \Cref{ex:strncpy}, with a context copying the string \texttt{Hello World}, where everything is marked with a high security tag: $\lock$.
-  The top half of \Cref{fig:ex-cct} (titled $\Lscct$), describes the execution trace of the program, while the bottom side of the table (titled ``Specification''), describes the related specification trace (\Cref{subsec:msctss:tracemodel}).
-  Read in parallel from top to bottom, the figure shows parts of the execution trace. 
-  In each half, the left column (\textit{Active}) has constant-time mode $\obj{ON}$ and the right one (\textit{Inactive}) has it $\obj{OFF}$.
-
-  When the constant-time mode is off, the execution yields events in similar fashion to before (\Cref{subsec:ltms,subsec:lsms,subsec:lms}).
-  But, if it is turned on, then the branching event \colorbox{yellow}{does not fire} anymore and both reading and writing to memory is related to a specification trace with \colorbox{lightgray}{no exposed secrets}.
-\end{exampleenv}
-\begin{figure}[!htb]
-	\vspace{-1em}\small
-  $$
-  % {
-  \begin{array}{ccc}
-    & \Lscct &        \\
-    \text{Active} & \mid & \text{Inactive} \\\hline
-    \obj{\ev{Alloc}\ \loc_{x}\ 12;\comp;\unlock} & \mid & \obj{\ev{Alloc}\ \loc_{x}\ 12;\comp;\unlock} \\
-    \obj{\ev{Alloc}\ \loc_{y}\ 12;\comp;\unlock} & \mid & \obj{\ev{Alloc}\ \loc_{y}\ 12;\comp;\unlock} \\
-    \obj{{\ev{iGet}}\ \loc_{x}\ 0;\comp;}\text{\colorbox{lightgray}{$\obj{\lock}$}} & \mid & \obj{\ev{Get}\ \loc_{x}\ 0;\comp;\lock} \\
-    \text{\colorbox{yellow}{$\obj{\emptyevent}$}} & \mid & \obj{\ev{Branch}\ 0;\comp;\lock} \\
-    \obj{{\ev{iGet}}\ \loc_{x}\ 0;\comp;}\text{\colorbox{lightgray}{$\obj{\lock}$}} & \mid & \obj{\ev{Get}\ \loc_{x}\ 0;\comp;\lock} \\
-    \obj{{\ev{iSet}}\ \loc_{y}\ 0\ \mathtt{'H'};\comp;}\text{\colorbox{lightgray}{$\obj{\lock}$}} & \mid & \obj{\ev{Set}\ \loc_{y}\ 0\ \mathtt{'H'};\comp;\lock} \\
-    \obj{{\ev{iGet}}\ \loc_{x}\ 1;\comp;}\text{\colorbox{lightgray}{$\obj{\lock}$}} & \mid & \obj{\ev{Get}\ \loc_{x}\ 1;\comp;\lock} \\
-    \text{\colorbox{yellow}{$\obj{\emptyevent}$}} & \mid & \obj{\ev{Branch}\ 0;\comp;\lock} \\
-    \vdots & \mid & \vdots \\
-    \obj{{\ev{iGet}}\ \loc_{x}\ 12;\comp;}\text{\colorbox{lightgray}{$\obj{\lock}$}} & \mid & \obj{\ev{Get}\ \loc_{x}\ 12;\comp;\lock} \\
-    \text{\colorbox{yellow}{$\obj{\emptyevent}$}}& \mid & \obj{\ev{Branch}\ 1;\comp;\lock} \\
-    \obj{\ev{Dealloc}\ \loc_{y};\comp;\unlock} & \mid & \obj{\ev{Dealloc}\ \loc_{y};\comp;\lock} \\
-    \obj{{\ev{iGet}}\ \loc_{y}\ 6;\comp;}\text{\colorbox{lightgray}{$\obj{\lock}$}} & \mid & \obj{\ev{Get}\ \loc_{y}\ 6;\comp;\lock} \\
-  \end{array}
-  % }
-  $$
-  $$
-  % {
-  \begin{array}{ccc}
-          & \text{Specification} & \\
-  \text{Active} & \mid & \text{Inactive} \\\hline
-     \ev{Alloc}\ \loc_{x}\ 12;\unlock & \mid & \ev{Alloc}\ \loc_{x}\ 12;\unlock\\
-     \ev{Alloc}\ \loc_{y}\ 12;\unlock & \mid & \ev{Alloc}\ \loc_{x}\ 12;\unlock\\
-     \ev{Use}\ \loc_{x}\ 0;\text{\colorbox{lightgray}{$\unlock$}} & \mid & \ev{Use}\ \loc_{x}\ 0;\lock\\
-     \emptyevent & \mid & \ev{Branch}\ 0;\lock\\
-     \ev{Use}\ \loc_{x}\ 0;\text{\colorbox{lightgray}{$\unlock$}} & \mid & \ev{Use}\ \loc_{x}\ 0;\lock\\
-     \ev{Use}\ \loc_{y}\ 0;\text{\colorbox{lightgray}{$\unlock$}} & \mid & \ev{Use}\ \loc_{y}\ 0;\lock\\
-     \ev{Use}\ \loc_{x}\ 1;\text{\colorbox{lightgray}{$\unlock$}} & \mid & \ev{Use}\ \loc_{x}\ 1;\lock\\
-     \emptyevent & \mid & \ev{Branch}\ 0;\lock\\
-     \vdots & \mid & \vdots \\
-      \ev{Use}\ \loc_{x}\ 12;\text{\colorbox{lightgray}{$\unlock$}} & \mid & \ev{Use}\ \loc_{x}\ 12;\lock\\
-      \emptyevent & \mid & \ev{Branch}\ 1;\lock\\
-      \ev{Dealloc}\ \loc_{y};\unlock & \mid & \ev{Dealloc}\ \loc_{y};\unlock\\
-      \ev{Use}\ \loc_{y}\ 6;\text{\colorbox{lightgray}{$\unlock$}} & \mid & \ev{Use}\ \loc_{y}\ 6;\lock\\
-  \end{array}
-  % }
-  $$
-  \caption{Traces for \Cref{ex:lscct}.}
-  \label{fig:ex-cct}
-\end{figure}
-
-\subsection{$\Lspec$: A Memory-Unsafe Language with Speculation}\label{subsec:lspec}
-
-$\Lspec$ extends $\Lscct$ (\Cref{subsec:lscct}) with speculative dynamic semantics that is inspired by existing work~\cite{guarnieri2018spectector,fabian2022automatic}.
-After branching, speculation starts by pushing the current configuration into the speculation state ($\ird{S}$), and run subsequent code in a predetermined window (whose size is $\omega$) until a rollback of operational state is performed.
-The window is set inside the speculation state.
-Within such windows, data may leak, this is marked as high $\ird{\lock}$ and with an annotation that indicates the respective speculative execution variant. 
-In this paper, we just consider SPECTRE-PHT~\cite{kocher2019spectre}, whose starting point is \Cref{tr:e-ifz-true-spec}.
-% where $\omega$ is the size of the speculation window, i.e., a number of reduction steps to be taken under speculative semantics.
-Additionally, $\ird{\lock}$ may also carry a $\ird{NONE}$ annotation, to signal that this is a leak irrespective of speculation, i.e., as defined earlier (\Cref{subsec:lscct}).
-The semantics is kept general enough to allow for future extension to support different variants~\cite{kocher2019spectre,maisuradze2018ret2spec,horn2019zero}.
-
-In terms of language features, $\Lspec$ includes a new barrier operation $\ird{\lbarrier}$ that blocks speculative execution. 
-To facilitate speculative execution in a non-assembly-like language, a stack of operational state is used and speculation is active if that stack has a size greater than one.
-This is exploited in \Cref{tr:e-spec-eat} to consume any leftover speculation.
-
-\vspace{-.5em}
-\[
-  \begin{array}{rrcl}
-    \text{(Base-Events)} & \ird{\preevent} &:=& \dots \mid \ird{Spec} \mid \ird{Rlb} \mid \ird{Barrier}\\
-    \text{(Expressions)} & \ird{\expr} &:=& \dots \mid \ird{\lbarrier} \\
-    \text{(Speculation State)} & \ird{S} &:=& \ird{\Omega\triangleright\expr} \mid \ird{S},\left(\ird{\Omega};\ird{n};\ird{\expr}\right)
-  \end{array}
-\]
-
-\begin{center}
-\newcommand{\expreval}[5]{{#1}\triangleright\xspace {#2}\xrightarrow{#5}\ {#3}\triangleright\xspace {#4}\xspace}
-\newcommand{\exprevald}[5]{\expreval{\ird{#1}}{\ird{#2}}{\ird{#3}}{\ird{#4}}{\ird{#5}}}
-
-  %
-  \typerule{$e-\ird{ifz}-\ird{true}-\text{spec}$}{
-    \ird{S} = \rtt{(\ird{\cfstate['];\sandboxtag{};D;\memstate});\ird{S}}{\ird{\expr[_1]}},({\ird{\cfstate;\sandboxtag{};D;\memstate}};\omega;{\ird{\expr[_2]}})
-  }{
-    \pstepto{\rtt{\ird{\cfstate;\sandboxtag{};D;\memstate}}{\ird{\lifz{0^{\securitytag{}}}{\expr[_1]}{\expr[_2]}}}}{\ird{S}}{\ird{{Branch}\ 0;\sandboxtag{};\securitytag{}}\cdot\ird{{Spec};\sandboxtag{};\securitytag{}}}
-  }{e-ifz-true-spec}
-  %
-  \typerule{$e-\ird{spec}-\text{eat}$}{
-    \ird{n} > 0 &
-    \pstepto{\rtt{\ird{\Omega}}{\ird{\expr}}}{\rtt{\ird{\Omega'}}{\ird{\expr[']}}}{\ird{\event}}
-  }{
-    \ghoststepto{\ird{S},(\ird{\Omega};\ird{n};\ird{\expr})}{\ird{S},(\ird{\Omega'};\ird{n}-1;\ird{\expr'})}{\ird{\event}}
-  }{e-spec-eat}
-  % 
-  \typerule{$e-\ird{spec}-\text{eaten}$}{
-  }{
-    \ghoststepto{\ird{S},(\ird{\Omega};\ird{0};\ird{\expr})}{\ird{S}}{\ird{Rlb;\Omega.\sandboxtag{};\unlock}}
-  }{e-spec-eaten}
-  %
-  \typerule{$e-\ird{spec}-\text{barrier}$}{
-  }{
-    \ghoststepto{\ird{S},(\ird{\Omega};\ird{n};\ird{\lbarrier})}{\ird{S},(\ird{\Omega};\ird{0};\ird{0})}{\ird{Barrier;\Omega.\sandboxtag{};\unlock}}
-  }{e-spec-barrier}
-\end{center}
-
-The trace model is extended with $\ird{{Spec}}$, $\ird{{Rlb}}$, and $\ird{{Barrier}}$ events that signal the respective operational action. 
-A $\ird{\lbarrier}$ prevents any execution whatsoever besides $\ird{{Rlb}}$ when run in speculative execution mode and does nothing when run in normal mode, e.g., \Cref{tr:e-spec-barrier}.%
-\footnote{In the rule, the notation $\cfstate' = \cfstate[.\text{win} = 0]$ means that $\cfstate'$ is a copy of $\cfstate$ up to field $\text{win}$, which is set to $0$.}
-Lastly, $\ird{Rlb}$ is emitted when the speculation window is zero and the operational state is rolled back (see \Cref{tr:e-spec-eaten}).
-
-\section{Case Study: Composing Secure Compiler Passes and Optimisations \pages{7}}\label{sec:casestud:rtp}
-This section defines several secure compilers, each of which robustly preserves a different property of interest as depicted in \Cref{fig:pipeline}.
-\begin{figure*}[!h]
-  \centering
-  \begin{tikzpicture}
-    \node (S) {$\src{L_{\tmssafe}}$};
-    \node[right=2.0 of S] (T) {$\trg{L}$};
-    \node[right=2.0 of T] (M) {$\irl{L_{\mssafe}}$};
-    \node[below right=1.5 and 1.0 of M] (D0) {$\irl{L_{\mssafe}}$};
-    \node[above right=1.5 and 1.0 of M] (C0) {$\irl{L_{\mssafe}}$};
-    \node[right=3.0 of M] (E) {$\irl{L_{\mssafe}}$};
-    \node[right=2.0 of E] (O) {$\obj{L_{\scctsafe}}$};
-    \node[right=2.0 of O] (Z) {$\ird{L_{\mathghost}}$};
-
-    \draw[->] (S) to[sloped] node[align=center] (tmsedge) {\gls*{tms}\\ \Cref{thm:cca:rtp:tms}} (T);
-    \draw[->] (T) to[sloped] node[align=center] {\gls*{sms}\\ \Cref{thm:ccb:rtp:sms}} (M);
-    \draw[->] (M) to[sloped] node[align=center] {\gls*{dce}\\ \Cref{thm:ccdce:rtp:ms}} (D0);
-    \draw[->] (M) to[sloped] node[align=center] {\gls*{cf}\\ \Cref{thm:cccf:rtp:ms}} (C0);
-    \draw[->] (D0) to[sloped] node[align=center] {\gls*{cf}\\ \Cref{thm:cccf:rtp:ms}} (E);
-    \draw[->] (C0) to[sloped] node[align=center] {\gls*{dce}\\ \Cref{thm:ccdce:rtp:ms}} (E);
-    \draw[->] (E) to[sloped] node[align=center] {\gls*{scct}\\ \Cref{thm:ccscct:rtp:scct}} (O);
-    \draw[->] (O) to[sloped] node[align=center] {\gls*{ss}\\ \Cref{thm:ccspec:rtp:spec}} (Z);
-
-    % Sections
-    \node (sectms) at ($(S)+(1.25,2.5)$) {{\Cref{subsec:cs:tms}}};
-    \node (secsms) at ($(T)+(1.25,2.5)$) {{\Cref{subsec:cs:ms}}};
-    \node (secopts) at ($(M)+(2.0,2.5)$) {{\Cref{subsec:cs:opts}}};
-    \node (secscct) at ($(E)+(1.5,2.5)$) {{\Cref{subsec:cs:scct}}};
-    \node (secss) at ($(O)+(1.5,2.5)$) {{\Cref{subsec:cs:ss}}};
-
-    \draw[thick,dotted,->] ($(S) + (0.2,-0.27)$) to ($(S) - (-0.2,1)$) to node[align=center] {\gls*{ms}\\ \Cref{corr:ccab:rtp:ms}} ($(M) - (0.0,1)$) to (M);
-
-    \draw[thick,dotted,->] (S) to ($(S) - (0.0,2.5)$) to node[align=center,pos=.65] {\gls*{specms}\\ \Cref{thm:ccall:rtp:specms}} ($(Z) - (0.0,2.5)$) to (Z);
-
-    % \draw[thick,dotted,->] (S) to[bend right=33,sloped] node[align=center] {\gls*{ms}\\ \Cref{corr:ccab:rtp:ms}} (M);
-    \draw[thick,dotted,->] (M) to[bend right=0,sloped] node[align=center] {\gls*{ms}\\ \Cref{thm:cccfccdce:rtp:ms}} (E);
-%    \draw[thick,dotted,->] (S) to[out=-45,in=198,sloped] node[align=center] {\gls*{ms}+\gls*{scct}\\ \Cref{thm:ccall:rtp:msscct}} (O);
-    % \draw[thick,dotted,->] (S) to[out=-40,in=210,sloped] node[align=center,pos=.8] {\gls*{specms}\\ \Cref{thm:ccall:rtp:specms}} (Z);
-
-    \draw[dashed] ($(S)-(0.0,-0.5)$) -- ($(S)-(0.0,-2.5)$);
-    \draw[dashed] ($(T)-(0.0,-0.5)$) -- ($(T)-(0.0,-2.5)$);
-    \draw[dashed] ($(M)-(0.0,-0.5)$) -- ($(M)-(0.0,-2.5)$);
-    \draw[dashed] ($(E)-(0.0,-0.5)$) -- ($(E)-(0.0,-2.5)$);
-    \draw[dashed] ($(O)-(0.0,-0.5)$) -- ($(O)-(0.0,-2.5)$);
-    \draw[dashed] ($(Z)-(0.0,-0.5)$) -- ($(Z)-(0.0,-2.5)$);
-  \end{tikzpicture}
-  \vspace{-1em}
-  \caption{Visualisation of the optimising compilation pipeline that preserves \gls*{specms}. %
-    Vertices in the graph are the programming languages from earlier sections (\Cref{sec:casestud:defs}). %
-    Full edges are secure compilers passes.
-    Dotted edges are composition of passes and use the presented framework (\Cref{sec:sequential}) to indicate the property they preserve. %
-    The dashed lines partition the graph into the sections where the respective theorems are presented.
-  }\label{fig:pipeline}
-\end{figure*}
-The section demonstrates the power of the framework (\Cref{sec:sequential}) by composing these compilers for a secure and optimising compilation chain that robustly preserves \gls*{specms}.
-The first step in this chain is the compiler from $\src{L_{\tmssafe}}$ to $\trg{L}$ that robustly preserves just \gls*{tms} (\Cref{thm:cca:rtp:tms}).
-From here, another pass from $\trg{L}$ to $\irl{L_{\mssafe}}$ ensures that no out-of-bounds accesses can happen and, thus, programs at this point attain \gls*{sms} (\Cref{thm:ccb:rtp:sms}).
-Since these properties compose into \gls*{ms}, composing these passes yields a compiler that robustly preserves \gls*{ms} (\Cref{corr:ccab:rtp:ms}).
-Then, the section presents two optimisation passes, namely \gls*{cf} and \gls*{dce}, each of which robustly preserves \gls*{ms} (\Cref{thm:ccdce:rtp:ms,thm:cccf:rtp:ms}).
-These passes can be freely ordered in the compilation chain without compromising memory safety (\Cref{thm:cccfccdce:rtp:ms}).
-The next step in the chain ensures that code stays \gls*{scct} (\Cref{thm:ccscct:rtp:scct}) when compiled from $\Lms$ to $\Lscct$, which is done by switching on a constant-tame mode for the computation.
-Lastly, by introducing barriers immediately after branches, speculative leaks via SPECTRE-PHT are prevented when compiling $\Lscct$ to $\Lspec$.
-The final result is that the whole compilation chain robustly preserves \gls*{specms} (\Cref{thm:ccall:rtp:specms}).
-
-\subsection{Robust Temporal Memory Safety Preservation}\label{subsec:cs:tms}
-
-The secure compiler from $\Ltms$ to $\Ltrg$ needs to ensure that when execution switches from context to component, the type signatures are respected.
-Thus, the compiler inserts the following dynamic typechecks before entering the body of a component-defined function (anything elided is a trivial identity function from source to target):
-
-\vspace{-1em}
-{
-\begin{gather}
-  \begin{align*}%{rcl}
-    %\cca(\src{x}) = &\ \trg{x} 
-    % \\
-  	%&
-  	%\qquad
-    %\cca(\src{\lbinop{\expr[_{1}]}{\expr[_{2}]}}) = &\ \lbinop[\trg]{\left[\cca(\src{\expr[_{1}]})\right]}{\left[\cca(\src{\expr[_{2}]})\right]} \\
-    %\cca(\src{n}) = &\ \trg{n} 
-    % \\
-    %&
-    \cca(\src{\lget{x}{\expr}}) = &\ \lget[\trg]{\trg{x}}{\left[\cca(\src{\expr})\right]} \\
-    \cca(\src{\ldelete{x}}) = &\ \ldelete[\trg]{\left[\cca(\src{x})\right]} 
-    \\
-    \cca(\src{\lfunction{g}{x:\natt\to\type[_{e}]}{\expr}})  =&
-  \end{align*}
-  \\
-  \begin{align*}
-%    \cca(\src{\lset{x}{\expr[_{1}]}{\expr[_{2}]}}) & = \lset[\trg]{\left[\cca(\src{x})\right]}{\left[\cca(\src{\expr[_{1}]})\right]}{\left[\cca(\src{\expr_{2}})\right]} \\
-% \\ 
-\lfunction[\trg]{\trg{g}}{\trg{x}}{\lifz[\trg]{\trg{\lhast{x}{\natt}}}{
-                                                            \left[\cca(\src{\expr})\right] %
-                                                                                                 }{\labort[\trg]}}
-  \end{align*}
-\end{gather}
-}
-
-\noindent Since $\trg{L}$ has no static typing, an attacker $\trg{\library_{\ctx}}$ can invoke a component function accepting a $\src{\natt}$ with $\trg{\lpair{17}{29}}$.
-With the dynamic check, the compiler ensures that execution aborts in such cases.
-%The compiler does not insert other checks and proceeds as the identity function (which in this paper amounts to a simple re-colouring of $\src{L_{\tmssafe}}$ to $\trg{L}$ expressions).
-
-Compiling the \texttt{strncpy} function from \Cref{sec:introduction} with $\cca$, the compiler would in this case ensure that the arguments that are evaluated in the compiled \texttt{strncpy} are valid.
-%This effectively has no influence on the resulting trace, so a corresponding cross-language trace relation $\sim^{\Ltms}_{\Ltrg}$ is just equality.
-
-$\cca$ is robustly preserving (\Cref{def:rtp}) \gls*{tms}:
-\begin{theorem}[$\cca$ secure w.r.t. \gls*{tms}]\label{thm:cca:rtp:tms}
-    $\rtp{\cca}{\tmssafe}$ % \Coqed
-\end{theorem}
-
-\subsection{Robust Spatial Memory Safety Preservation}\label{subsec:cs:ms}
-
-The spatial memory safety preserving compiler from $\Ltrg$ to $\Lms$ only inserts bounds-checks whenever reading from or writing to memory in order to enforce \gls*{sms}.
-These bounds checks need the bounds information, which the compiler keeps around by introducing a fresh identifier $\irl{x_{SIZE}}$ for each allocation that binds $\irl{x}$.
-Then, it is simply a matter of referring to that variable and ensuring that memory accesses are in the interval $[\irl{0},\irl{x_{SIZE}})$.
-When the check fails, the code aborts.
-
-\begin{nscenter}
-% \small
-  $$
-  \begin{array}{rcl}
-    \ccb(\trg{\lnew{x}{\expr[_{1}]}{\expr[_{2}]}}) & = 
-                                                   & \llet[\irl]{\irl{x_{SIZE}}}{\ccb(\trg{\expr[_{1}]})}{}
-    		\\&&
-    		\lnew[\irl]{\irl{x}}{\irl{x_{SIZE}}}{\ccb(\trg{\expr[_{2}]})}
-    		 \\
-  \ccb(\trg{\lget{x}{\expr}}) & = 
-                              & \llet[\irl]{\irl{x_{n}}}{\ccb(\trg{\expr})}{}
-  	\\&&
-  \lifz[\irl]{\irl{0\le x_{n}<x_{SIZE}}}{\\&&\irl{\lget{x}{x_{n}}}}{}
-  		\irl{\labort}
-  	  \\
-  \ccb(\trg{\lset{x}{\expr[_{1}]}{\expr[_{2}]}}) & = 
-                                                 & \llet[\irl]{\irl{x_{n}}}{\ccb(\trg{\expr[_{1}]})}{}
-  		\\&&
-  \lifz[\irl]{\irl{0\le x_{n}<x_{SIZE}}}{\\&&\lset[\irl]{\irl{x}}{\irl{x_{n}}}{}
-  		\ccb(\trg{\expr[_{2}]})
-  		}{\irl{\labort}} 
-  		% \\
-  \end{array}
-  $$
-\end{nscenter}
-
-\begin{exampleenv}[Instrumented \texttt{strncpy}]
-  Consider again \texttt{strncpy}, but instrumented for \gls*{sms}:
-    \begin{lstlisting}[language=c,basicstyle=\ttfamily, morekeywords={size_t}]
-void strncpy(size_t n, size_t dst_size, char *dst,
-             size_t src_size, char *src) {
-  for(size_t i = 0; i < src_size
-      && src[i] != '\0' && i < n; ++i) {
-    if(i < src_size && i < dst_size) {
-      dst[i] = src[i];
-    }
-  }
-}
-    \end{lstlisting}
-Consider context \texttt{strncpy(2, x, y)}, where \texttt{x} and \texttt{y} are pointers to valid regions of memory with allocated space for exactly two cells and do not contain the null-terminating character \texttt{'\textbackslash 0'}.
-% 
-Without the \gls*{sms} pass, the event $\ev{Use}\ \loc_{x}\ 2;\comp;\unlock$ would appear on the trace, but that indicates an out-of-bounds access! 
-Fortunately, with \gls*{sms} mitigation in place, that event does not appear during execution, since the bounds check prevents the condition \texttt{src[i] != '\textbackslash 0'} from executing.
-\end{exampleenv}
-
-Contrary to the previous compiler, $\ccb$ may change the trace of the original $\Ltrg$ program: if there is a memory access, it needs to be protected with a bounds check.
-% 
-The corresponding relation $\sim^{\Ltrg}_{\Lms} : \trg{\trace}\times\irl{\trace}$ that describes this semantic effect of the compiler is defined partially below.
-% 
-We omit action $\trg{Set}$, which is treated analogously, and any other event, which is related to its cross-language equivalent.
-% \footnote{Up to colours.}
-% \footnote{In the interest of space, the full details are in the technical report.}
-
-{
-\[
-  \typerulenolabel{xrel:sms:read}{\trg{n}\text{ in bounds}}{\trg{Get\ \loc\ n;\comp}\sim^{\Ltrg}_{\Lms}\irl{Get\ \loc\ n;\comp}}
-\]
-% \[
-%   \typerulenolabel{xrel:sms:panic}{\trg{n}\text{ out of bounds}}{\trg{Get\ \loc\ n;\comp}\sim^{\Ltrg}_{\Lms}\irl{\lightning}}
-% \]
-}
-
-% Note that neither $\Ltrg$ nor $\Lms$ expose branches on the trace.\MP{so what?}
-For simplicity, we elide the environment that this relation carries around in order to bind each location to its metadata (such as its size), and resolve the ``$\trg{n}$ in/out of bounds'' premise.
-% The bounds check mentioned in the premises of the rules above is left abstract for the sake of simplicity, the interested reader can find
-% since the precise definition needs some technicalities that only inhibit readability without any technical insights.
-% 
-We can now prove that compiler $\ccb$ robustly preserves \gls*{sms}.
-\begin{theorem}[$\ccb$ secure w.r.t. \gls*{sms}]\label{thm:ccb:rtp:sms}
-  $\rtpsim{\ccb}{\smssafe}{\sim^{\Ltrg}_{\Lms}}$ % \Coqed
-\end{theorem}
-
-At this point we can compose $\ccb$ with the previous compiler ($\cca$), but in order to do so, we need a trace relation from $\Ltms$ to $\Lms$.
-% 
-We can obtain this relation by composing the trace relation we just defined ($\sim^{\Ltrg}_{\Lms}$) with the one used by the previous compiler: $\sim^{\Ltms}_{\Ltrg} : \src{\trace}\times\trg{\trace}$.
-% 
-The latter has not been previously defined (nor has it been used in the related theorem) because that is just an equality relation, since the trace models of $\Ltms$ and of $\Ltrg$ are the same.
-% 
-Thus, we formally define $\sim^{\Ltms}_{\Lms}: \src{\trace} \times \irl{\trace}$ as the following composition: $\sim^{\Ltms}_{\Ltrg}\bullet\sim^{\Ltrg}_{\Lms}$.
-% 
-With this relation, \Cref{corr:ccab:rtp:ms} states that the composition of $\cca$ and $\ccb$ is secure w.r.t.
-\gls*{ms} and it follows from \Cref{thm:cca:rtp:tms,thm:ccb:rtp:sms} using \Cref{thm:rtpsim:sig}.
-% 
-\begin{corollary}[$\cca\circ\ccb$ secure w.r.t. \gls*{ms}]\label{corr:ccab:rtp:ms}
-  $\;$
-
-  \begin{nscenter}
-    $\rtpsim{\cca\circ\ccb}{\mssafe}{\sim^{\Ltms}_{\Lms}}$ % \Coqed
-  \end{nscenter}
-\end{corollary}
-% 
-This proof requires another precondition besides \Cref{thm:ccb:rtp:sms,thm:cca:rtp:tms}: $\sim^{\Ltms}_{\Ltrg}$ needs to be well-formed with respect to $\smssafe$.
-This follows trivially since $\sim^{\Ltms}_{\Ltrg}$ is an equality. 
-% 
-\begin{lemma}[$\sim^{\Ltms}_{\Ltrg}$ well-formed w.r.t. $\smssafe$]\label{lem:wf:ltmsltrg}
-$\;$
-
-  \begin{nscenter}
-  $\wfcsig{\sim^{\Ltms}_{\Ltrg}}{\smssafe}$
-  \end{nscenter}
-\end{lemma}
-
-\subsection{Optimising Compilers}\label{subsec:cs:opts}
-
-This section defines two optimising compiler passes from $\Lms$ to $\Lms$ which perform \gls*{dce} and \gls*{cf}, respectively.
-The \gls*{dce} pass applies a na\"ive rewrite rule on conditionals.
-The \gls*{cf} pass relies on an auxiliary function \texttt{mix} that uses a substitutions accumulator $\irl{\substlist}$ in order to rewrite constant binary operations, e.g., $\irl{{17}-1}$ to $\irl{16}$, and replace variables that are assigned to constants, e.g., $\irl{\llet{x}{7}{x}}$ to $\irl{7}$.
-
-\vspace{-1em}
-\begin{gather*}
-  \begin{align*}
-    \ccdce(\irl{\lifz{true}{\expr[_{1}]}{\expr[_{2}]}}) & = \ccdce(\irl{\expr[_{1}]}) &\\
-    \ccdce(\irl{\lifz{false}{\expr[_{1}]}{\expr[_{2}]}}) & = \ccdce(\irl{\expr[_{2}]}) &
-  \end{align*}
-  \\
-  % \begin{align*}
-  %   \ccdce(\irl{\lbinop{\expr[_{1}]}{\expr[_{2}]}}) & = \lbinop[\irl]{\ccdce(\irl{\expr[_{1}]})}{\ccdce(\irl{\expr[_{2}]})} &
-  % \end{align*}
-  % \\[0.25cm]
-  \begin{align*}
-    \cccf(\irl{\expr}) & = \partialeval{\irl{\expr}}{\irl{\hole{\cdot}}} &
-  \end{align*}
-  \\[0.125cm]
-  \begin{align*}
-   \partialeval{\irl{x}}{\irl{\substlist}} & = \irl{n} 
-   	\qquad\qquad \text{if } \irl{\subst{n}{x}}\in\irl{\substlist} \\
-   \partialeval{\irl{x}}{\irl{\substlist}} & = \irl{x} 
-   \qquad\qquad \text{if } \irl{\subst{n}{x}}\notin\irl{\substlist} \\
-   \partialeval{\irl{\lbinop{n}{m}}}{\irl{\substlist}} & = \irl{k} 
-   \qquad\qquad \text{if } \lbinop{\irl{n}}{\irl{m}}=k \\
-   %\partialeval{\irl{\lbinop{\expr[_{1}]}{\expr[_{2}]}}}{\irl{\substlist}} & = \lbinop[\irl]{\partialeval{\irl{\expr[_{1}]}}{\irl{\substlist}}}{\partialeval{\irl{\expr[_{2}]}}{\irl{\substlist}}} \\
-   \partialeval{\irl{\llet{x}{n}{\expr}}}{\irl{\substlist}} & = \partialeval{\irl{\expr}}{\irl{\subst{n}{x}\cdot\substlist}} 
-  % \\
-  % \partialeval{\irl{\lget{x}{\expr}}}{\irl{\substlist}} & = \lget[\irl]{\irl{x}}{\partialeval{\irl{\expr}}{\irl{\substlist}}}
-% \end{align*}
-\end{align*}
-% \\
-% \begin{align*}
-%   \partialeval{\irl{\llet{x}{\expr[_{1}]}{\expr[_{2}]}}}{\irl{\substlist}} & = \llet[\irl]{\irl{x}}{\partialeval{\irl{\expr[_{1}]}}{\irl{\substlist}}}{\\&\partialeval{\irl{\expr[_{2}]}}{\irl{\substlist}}} \\
-%   \partialeval{\irl{\lifz{\expr[_{1}]}{\expr[_{2}]}{\expr[_{3}]}}}{\irl{\substlist}} & = \lifz[\irl]{\partialeval{\irl{\expr[_{1}]}}{\irl{\substlist}}}{\\&\partialeval{\irl{\expr[_{2}]}}{\irl{\substlist}}}{\partialeval{\irl{\expr[_{3}]}}{\irl{\substlist}}} \\
-% \end{align*}
-\end{gather*}
-% \vspace{-3em}
-
-Note that both passes have no effect on the resulting trace of a program, up to $\emptyevent$-steps. 
-Because of this, both passes have equality as corresponding cross language trace relation. 
-Moreover, it is straightforward to prove both passes as secure (\Cref{def:rtp}) w.r.t. \gls*{ms}.
-
-\begin{theorem}[$\ccdce$ secure w.r.t. \gls*{ms}]\label{thm:ccdce:rtp:ms}
-  $\rtp{\ccdce}{\mssafe}$ %\Coqed
-\end{theorem}
-\begin{theorem}[$\cccf$ secure w.r.t. \gls*{ms}]\label{thm:cccf:rtp:ms}
-  $\rtp{\cccf}{\mssafe}$ %\Coqe
-\end{theorem}
-
-With both \Cref{thm:ccdce:rtp:ms,thm:cccf:rtp:ms} it follows from \Cref{corr:swappable} that the two passes can be interchanged arbitrarily:
-
-\begin{theorem}[$\cccf\circ\ccdce$ and $\cccf\circ\ccdce$ are secure w.r.t. \gls*{ms}]\label{thm:cccfccdce:rtp:ms}
-$\;$
-
-  \begin{nscenter}
-  \phantom{and~}\!\!$\rtp{\cccf\circ\ccdce}{\mssafe}$
-
-  and~$\rtp{\ccdce\circ\cccf}{\mssafe}$ % \Coqed
-  \end{nscenter}
-\end{theorem}
-
-\subsection{Robust Strict Cryptographic Constant Time Preservation}\label{subsec:cs:scct}
-
-This section defines a compiler $\ccscct$ from $\Lms$ to $\Lscct$ that robustly preserves \gls*{scct}. 
-Given the fact that $\Lscct$ provides a \gls*{cct}-mode that can be turned on or off, the compiler inserts wrapper code for function calls and function bodies to ensure that execution in the component always happen in \gls*{cct}-mode.
-This simple flag combines the effect of FaCT~\cite{cauligi2019fact}
-
-\vspace{-1em}
-\[
-\begin{array}{rcl}
-  \ccscct(\irl{\lfunction{g}{x}{\expr}}) & = & \lfunction[\obj]{\obj{g}}{\obj{x}}{\obj{\lwrdoit{ON};}\ccscct(\irl{\expr})} \\
-  \ccscct(\irl{\lcall{g}{\expr}}) & = & \lcall[\obj]{\obj{g}}{\ccscct(\irl{\expr})\obj{; \lwrdoit{ON}}} 
-    % \\
-    % \ccscct(\irl{\lbinop{\expr[_{1}]}{\expr[_{2}]}}) & = \lbinop[\obj]{\ccscct{\irl{\expr[_{1}]}}}{\ccscct{\irl{\expr[_{2}]}}} 
-    % \\
-\end{array}
-\]
-% \vspace{-2em}
-%
-The context can overwrite the flag and exit the mode, but upon invoking a function that is part of the component, the flag is set again.
-Because of this, the corresponding cross-language trace relation $\sim^{\Lms}_{\Lscct}$, only relates events without secrets:%
-
-\begin{center} 
-  \typerulenolabel{xrel:scct:noleak}{}{\irl{\preevent;\comp}\sim^{\Lms}_{\Lscct}\obj{\preevent;\comp;\unlock}}
-\end{center}
-
-The compiler is secure w.r.t. \gls*{scct}: %, similarly proven as in \Cref{subsec:cs:tms}.
-
-\begin{theorem}[$\ccscct$ secure w.r.t. \gls*{scct}]\label{thm:ccscct:rtp:scct}
-  \small
-  $\rtpsim{\ccscct}{\scctsafe}{\sim^{\Lms}_{\Lscct}}$ % \Coqed
-\end{theorem}
-
-\subsection{Robust Speculative Safety Preservation}\label{subsec:cs:ss}
-
-This section defines the final compilation pass $\ccspec$, which ensures that $\Lscct$ programs, which are assumed to be \gls*{ss}, stay \gls*{ss} at $\Lspec$-level. 
-To do so, $\ccspec$ inserts a speculation barrier after branches, which is sufficient to harden the program against speculative leaks, since SPECTRE-PHT~\cite{kocher2019spectre} is the only speculative leak modeled in the semantics of $\Lspec$.
-
-\vspace{-1em}
-\[
-\begin{array}{cl}
-  &\ccspec{(\obj{\lifz{\expr[_0]}{\expr[_1]}{\expr[_2]}})} = 
-  \\
-  &\qquad\qquad \lifz[\ird]{\ccspec{\left(\obj{\expr[_0]}\right)}}{\ird{\lbarrier;}\ccspec{\left(\obj{\expr[_1]}\right)} \\&\qquad\qquad}{ \ird{\lbarrier;}\ccspec{\left(\obj{\expr[_2]}\right)}} 
-\end{array}
-\]
-% \vspace{-2em}
-%
-
-Clearly, the corresponding cross-language trace relation $\sim^{\Lscct}_{\Lspec}$ has only one non-trivial case: for branches, only relate them where speculation is blocked by a barrier:
-
-\begin{center}
-  \typerulenolabel{xrel:spec:if}{}{\obj{Branch\ n}\sim^{\Lscct}_{\Lspec}\ird{Branch\ n}\cdot\ird{Spec}\cdot\ird{Barrier}\cdot\ird{Rlb}}
-\end{center}
-
-The base-event relation above scales to full events by ensuring the missing annotations ($\obj{\comp;\securitytag{}}$ and $\ird{\comp;\securitytag{}}$) are the same.
-% 
-With this relation, we prove that $\ccspec$ is secure with respect to \gls*{ss}.
-\begin{theorem}[$\ccspec$ secure w.r.t. \gls*{ss}]\label{thm:ccspec:rtp:spec}
-  \small$\rtpsim{\ccspec}{\sssafe}{\sim^{\Lscct}_{\Lspec}}$ % \Coqed
-\end{theorem}
-
-\subsection{Robust Preservation of Memory Safety, Strict Cryptographic Constant Time, and Speculative Safety}
-
-Finally, this subsection combines all previous results into one compilation chain to get that it preserves full \gls*{specms}.
-Let $\ccspecms$ be the compiler that is the composition of $\cca$, $\ccb$, $\cccf$, $\ccdce$, $\ccscct$, and $\ccspec$. 
-Let $\sim^{\Ltms}_{\Lspec}$ be the composition of $\sim^{\Ltms}_{\Lms}$, $\sim^{\Lms}_{\Lscct}$, and $\sim^{\Lscct}_{\Lspec}$.
-Then, the following corollary holds.
-
-\begin{corollary}[$\ccspecms$ secure w.r.t. \gls*{specms}]\label{thm:ccall:rtp:specms}
-  $\;$
-
-  \begin{nscenter}
-    $\rtpsim{\cc{\Ltms}{\Lspec}}{\mssafe\cap\scctsafe\cap\sssafe}{\sim^{\Ltms}_{\Lspec}}$ % \Coqed
-  \end{nscenter}
-\end{corollary}
-
-As with \Cref{corr:ccab:rtp:ms}, it is important to ensure that the respective cross language trace relations are well-formed (\Cref{def:wfc:sig:tracerel}).
-It is already known that $\sim^{\Ltms}_{\Lms}$ is well-formed with respect to $\mssafe$ (\Cref{lem:wf:ltmsltrg}).
-Next in the chain is $\sim^{\Lms}_{\Lscct}$, which has to be well-formed w.r.t. $\scctsafe$.
-This lemma holds, since a trace that was $\scctsafe$ is $\scctsafe$ even after applying $\sim^{\Lms}_{\Lscct}$: the relation enforces that $\Lscct$ traces related to $\Lms$ traces have no leaks of secrets whatsoever.
-
-\begin{lemma}[$\sim^{\Lms}_{\Lscct}$ well-formed w.r.t. $\scctsafe$]\label{lem:wf:lsmslscct}
-$\;$
-
-  \begin{nscenter}
-  $\wfcsig{\sim^{\Lms}_{\Lscct}}{\scctsafe}$
-  \end{nscenter}
-\end{lemma}
-
-The last relation is $\sim^{\Lscct}_{\Lspec}$ which needs to be well-formed w.r.t. $\sssafe$.
-Similarly to the previous relation, this holds, since $\sim^{\Lscct}_{\Lspec}$ only relates $\Lspec$ traces, which do not have speculative leaks, with $\Lscct$ traces.
-
-\begin{lemma}[$\sim^{\Lscct}_{\Lspec}$ well-formed w.r.t. $\sssafe$]\label{lem:wf:lscctlspec}
-  $\wfcsig{\sim^{\Lscct}_{\Lspec}}{\sssafe}$
-\end{lemma}
-
-\section{Formal Insights}\label{sec:formalities}
-
-This section first discusses how to connect each language-specific security property to the general security properties presented in \Cref{sec:compprop} (\Cref{subsec:formalities:maps}), then it demonstrates that the security property resulting from the universal image projection is faithful to the original property (\Cref{subsec:formalities:props}).
-Finally, this section discusses why the order of compiler passes matters, and how does our framework help with identifying insecure compositions (\Cref{subsec:compatsecpasses}), and it gives additional technical insights on the secure compilation proofs (\Cref{subsec:seccompproofs}).
-
-\subsection{From Language Traces to General Ones}\label{subsec:formalities:maps}
-
-The previous theorems talk about preserving properties expressed in the trace model of the languages of each compiler.
-However, these trace models are not the same trace model we used to specify the properties of \Cref{sec:compprop} (indicated with \ev{L}), which serves as the ``ground truth'' for the meaning of our properties.
-% 
-To bridge this gap, the formal development requires specifying additional trace relations, from each of the language trace models to the \ev{L} one, that, for example, relate $\src{Get\ \loc\ n}$ and $\src{Set\ \loc\ n}$ to $\ev{Use\ \loc\ n}$ (and that induce a related universal image that we use in \Cref{subsec:formalities:props}).
-% 
-One key insight of these relations is that they all omit context-made actions for two reasons: (1) contexts (which are universally quantified) can trivially invalidate any property and (2) we are interested in the component upholding the properties.
-%
-
- % disregarded a glaring but minor technical issue: take $\tmssafe$ (\Cref{def:trace:tmsdef}) as an example.
-% It contains traces as specified in \Cref{subsec:basic:memsafety:tracemodel}.
-% However, these are different from $\Ltms$ traces (\Cref{subsec:ltms}) which means that \Cref{thm:wt:tms} is ``ill-typed''.
-% Similarly for the other theorems.
-% In the interest of reducing visual baggage, the notation left out yet another cross-language trace relation, translating language-specific traces into these non-language specific traces. 
-% In doing so, context-level actions are dropped and, e.g., both $\src{Get\ \loc\ n}$ and $\src{Set\ \loc\ n}$ are translated into $\ev{Use\ \loc\ n}$.
-
-\subsection{Security Properties and Their Meaning}\label{subsec:formalities:props}
-Each of the presented compilers use a cross-language trace relation, which is also used to translate the property from one language to the other one (via the existential or universal images).
-% 
-%Whenever such translations occur, it is important to check whether there has been a change of meaning.
-%Secure compiler engineers need to ensure that such translations of properties do not change meaning in a fundamental wrong way. 
-% lifting or lowering reasoning from target- to source-level or vice-versa,
-While the meaning of projected properties does change with a translation, the change should not allow for a flawed compilation pipeline.
-% 
-For example, we could be using a trace relation that translates a property in a language to a totally different one in another language.
-% 
-To raise the trust into the translation of properties, \Cref{thm:prop-rel-corr} states (in a general fashion) that each security property is faithfully translated using the universal image according to the cross-language trace relation induced by the compiler.
-
-\begin{theorem}[Properties Relation Correctness]\label{thm:prop-rel-corr}
-  \begin{align*}
-    \forall& \pi \in \{\tmssafe,\smssafe,\scctsafe,\sssafe\}, 
-    \\
-    \text{ for each }& \text{pair of languages } \src{L} \text{ and } \trg{L} \text{ used by the compilers},
-    \\
-    \text{ if }& 
-    % \tau_{\sim^{\trg{L}}_{\ev{L}}}(\pi) \sim^{\trg{L}}_{\ev{L}} \pi
-    \sigma_{\sim^{\trg{L}}_{\ev{L}}}(\pi) \sim^{\trg{L}}_{\ev{L}} \pi
-    \text{ and } 
-    % \sigma_{\sim^{\src{L}}_{\trg{L}}}(\tau_{\sim^{\trg{L}}_{\ev{L}}}(\pi)) \sim^{\src{L}}_{\trg{L}} \tau_{\sim^{\trg{L}}_{\ev{L}}}(\pi)
-    \sigma_{\sim^{\src{L}}_{\trg{L}}}(\sigma_{\sim^{\trg{L}}_{\ev{L}}}(\pi)) \sim^{\src{L}}_{\trg{L}} \sigma_{\sim^{\trg{L}}_{\ev{L}}}(\pi)
-    \\
-    \text{ then }& 
-    % \tau_{\sim^{\src{L}}_{\ev{L}}}(\pi) = \sigma_{\sim^{\src{L}}_{\trg{L}}}(\tau_{\sim^{\trg{L}}_{\ev{L}}}(\pi)) 
-    \sigma_{\sim^{\src{L}}_{\trg{L}}}(\sigma_{\sim^{\trg{L}}_{\ev{L}}}(\pi)) \sim^{\src{L}}_{\ev{L}} \pi
-  \end{align*}
-\end{theorem}
-The complexity of this theorem is that the relation in the conclusion cannot be obtained by composing the two relations in the premises.
-% 
-%% MK: the following does not say anything about the theorem, only about the syntactic compiler
-%What happens in practice is that compiler engineers are given the relations from each language to \ev{L} and they need to build a compiler that induces a relation $\sim^{\src{L}}_{\trg{L}}$ that respects this theorem.
-%
-
-We now informally argue why this theorem holds for the composition of all considered properties from \Cref{sec:compprop}.
-
-\paragraph{$\sigma_{\sim^{\Ltms}_{\Lms}}\left(\tmssafe\cap\smssafe\cap\scctsafe\cap\sssafe\right)$}
-
-For the four properties considered here, the trace models of $\Ltms$, $\Ltrg$, and $\Lms$ do not consider actions related to $\scctsafe$ and $\sssafe$, so these two properties are trivially translated correctly.
-% 
-We now discuss the remaining $\tmssafe$ and $\smssafe$ in the form of their intersection $\mssafe$.
-
-Recall that $\sim^{\Ltms}_{\Lms}$ is the composition of $\sim^{\Ltms}_{\Ltrg}$ and $\sim^{\Ltms}_{\Lms}$.
-% 
-Since $\sim^{\Ltms}_{\Ltrg}$ is an equality, this relation trivially preserves the meaning of translated properties: related traces are identical!
-
-Finally, let us consider a trace $\irl{\trace}\in\mssafe$ and understand what is that is related to via $\sim^{\Ltrg}_{\Lms}$.
-% 
-{\em All} traces $\trg{\trace}$ with $\trg{\trace}\sim^{\Ltrg}_{\Lms}\irl{\trace}$ are identical to $\irl{\trace}$ except for get and set actions, which require for in-bound accesses (as stated in \Cref{subsec:cs:ms}): this clearly respect $\mssafe$.
-%
-
-% In both cases, the translated event does not represent an out-of-bounds memory access.
-
-% \footnote{Both $\sim^{\Ltms}_{\Ltrg}$ and $\sim^{\Ltrg}_{\Lms}$ filter context actions.}.
-
-% Now, starting from $\mssafe$ expressed in $\Lms$, we check that its translation in $\Ltms$ is still $\mssafe$.
-% % 
-% We can see that the property resulting from the projection of $\Lms$ traces back to $\Ltms$ faithfully models $\mssafe$, which translates to both $\scctsafe$ and $\sssafe$, 
-% \MPin{ equiv  +  other rel only adds traces that fail }
-% since these properties trivially hold for $\Ltms$, $\Ltrg$, and $\Lms$ traces.
-
-\paragraph{$\sigma_{\sim^{\Lms}_{\Lscct}}\left(\tmssafe\cap\smssafe\cap\scctsafe\cap\sssafe\right)$}
-
-As before, the trace models of $\Lscct$ and $\Lms$ cannot express $\sssafe$, so that is trivially translated correctly.
-% 
-Also, the trace model of $\Lscct$ extends the one of $\Lms$ with respect to $\tmssafe$ and $\smssafe$ events, so the translation argument regarding those two properties is the same as before.
-% 
-Thus, we need to reason about whether $\scctsafe$ is translated correctly.
-
-By definition, $\sim^{\Lms}_{\Lscct}$ only relates $\obj{\unlock}$ events, which is also the same relation induced by $\sim_{\ctsafe}$ in \Cref{sec:msscct-rel}.
-% 
-This ensures that composing the relations only relates $\unlock$ events, and thus the property is translated correctly.
-
-% Consider a trace $\obj{\trace}\in\scctsafe$ and all $\irl{\trace}$ such that $\irl{\trace}\sim^{\Lms}_{\Lscct}\obj{\trace}$. 
-% In turn, $\obj{\trace}$ cannot contain any leaks of secret values, so it cannot violate $\scctsafe$.
-% Toggling data-independent timing mode has no impact on $\mssafe$ or $\sssafe$.
-
-\paragraph{$\sigma_{\sim^{\Lscct}_{\Lspec}}\left(\tmssafe\cap\smssafe\cap\scctsafe\cap\sssafe\right)$}
-The trace model of $\Lspec$ extends the one of $\Lscct$ with respect to $\tmssafe$, $\smssafe$, and $\scctsafe$, so the translation argument for those three properties is the same as before.
-% 
-Concerning $\sssafe$, $\sim^{\Lscct}_{\Lspec}$ relates $\Lspec$ speculation traces with $\Lscct$ branches with the $\lock_{\text{NONE}}$ tag, so the property is translated correctly, according to the relation defined in \Cref{sec:spec-ms-rel}.
- % introducing $\ird{barrier}$ never violates $\sssafe$ and has no effect on $\mssafe$ or $\scctsafe$.
-
-%\MPin{
-%	can this be made formal?
-%	what's the point?
-%}
-
-\subsection{Compatibility of Secure Compiler Passes}\label{subsec:compatsecpasses}
-
-% \Cref{thm:rtpsim:sig} is applicable, up to the well-formedness of $\sim_1$ w.r.t. $\trg{\class[_2]}$.
-Consider applying $\ccspec$ first and then $\ccb$ (albeit currently syntactically impossible).
-% , let us pretend that it is possible by just extending $\ccb$ for the constructs present in $\Lscct$ that are not present in $\Ltrg$.
-% Clearly, as evidenced earlier (\Cref{subsec:formalities:props}), $\ccb\cdot\ccspec$ is an acceptable composition that yields an interesting class of security properties, i.e., $\sigma_{\sim_{\irl{ms}}\bullet\sim_{\ird{\mathghost}}}\left(\mssafe\cap\scctsafe\cap\sssafe\right)$.
-In this case, $\ccb$ would insert new branches into the code that are not protected by a speculation barrier! 
-This concern is reflected in the proofs that establish that the security class resulting of the composition of the trace relations is meaningful.
-% , much similar to ensuring that a definition models what the formal methods expert intends to model. 
-For the $\ccspec\cdot\ccb$ case, the class is $\sigma_{\sim_{\ird{\mathghost}}\bullet\sim_{\irl{ms}}}\left(\mssafe\cap\scctsafe\cap\sssafe\right)$.
-Since $\ev{Use\ \loc\ n}\sim_{\irl{ms}}\irl{Branch\ n}\cdot\irl{Spec}\cdots$, where $\cdots$ does not contain a $\ev{{Barrier}}$ event, the resulting class is \emph{not} the original $\sssafe$ that is intended and it would break the corresponding \Cref{thm:prop-rel-corr}.
-% Because of this, it is crucial to analyse the precise shape of the class of properties after mapping a target-level property to a source-level property\footnote{Dito for mapping a source-level property to a target-level property.}.
-
-However, the composition is still technically possible and it is the job of the compiler engineers to ensure that the secure compilation pipeline happens in an order that ensures that the mapped security property is the intended one.
-
-\subsection{Secure Compilation Proofs}\label{subsec:seccompproofs}
-
-Our secure compilation proofs rely on backtranslations~\cite{abate2019jour,patrignani2021rsc}, which let one construct a source context starting from either target traces (aka trace-based backtranslations) or target contexts (aka context-based backtranslations).
-These backtranslations also require setting up cross-language relations between various language elements such as expressions and program states, so we leave these details for the technical reports.
-% , the usual strategy is to establish a backwards simulation between source components and their compiled target counterpart. 
-% For the universally quantified target context, a technique known as backtranslation is well-established, which constructs a source-level context either by backtranslating the target-level context, amounting to a ``decompilation'', or by constructing it using the target-level trace, so that the source-context emits parts of the whole execution trace that are related to the parts emitted by the target-level context.
-All backtranslations in the case-study are trace based except for those required by $\ccdce$ and $\cccf$, which are context-based (and they are an identity function). 
-% Since these optimisations have no significant semantic effect and source- and target-language are equal, the backtranslation amounts to an identity translation.
-
-\section{Related Work\pages{2}}\label{sec:relwork}
-
-This section discusses robust compilation, other secure compilation criteria and work related to the properties preserved in the case study.
-% (\Cref{subsec:relw:seccomprtp}) 
-% (\Cref{subsec:relw:seccompcrit}).
-% Since the case study of \Cref{sec:casestud:defs,sec:casestud:rtp} implements measures for preserving \gls*{ms}, \gls*{cct}, and \gls*{ss}, 
-% Then, this section discusses work related 
-% relevant related work as well 
-% (\Cref{subsec:relw:msmechs,subsec:relw:cctmechs,subsec:relw:ssmechs}).
-
-\paragraph*{Secure Compilation as Robust Preservation}\label{subsec:relw:seccomprtp}
-
-The robust preservation of properties as a compiler-level criterion has been analysed extensively~\cite{abate2019jour,patrignani2021rsc,abate2021extacc,patrignani2019survey} and thus we build on that framework.
-No existing work is concerned with composing secure compilers, however, existing work~\cite{abate2021extacc} sketches composition of trace-relating compiler correctness in a similar way to what has been presented here.
-%Parts of these works consider languages with different trace models and our technical setup can handle this. 
-The work relating robust preservation with universal composability~\cite{patrignani2022universal} is closest to what this paper presents.
-The authors demonstrate a similar compositionality theorem to what is presented here (\Cref{sec:sequential}) but use it in the context of protocols.
-The work does not consider the generality to support different trace models or composition of compilers which robustly preserve different classes.
-% The authors demonstrate a similar compositionality theorem to what is presented here (\Cref{sec:sequential}) as well as in an earlier version of this work~\cite{kruse2022csc}.
-% However, they do not demonstrate the scalability of the approach by means of a case study.
-
-%\paragraph*{Composition of Secure Compilers}\label{subsec:relw:seccompcomp}
-
-There is work on lifting exploits for single compilers to the whole chain~\cite{paykin2019weird}.
-%Our framework does not consider exploits.
-%\MP{ a bit dry and short}
-While that work considers {\em in}secure compilation and composition thereof in terms of exploits, the composition they are interested in allows to lift an exploit for one compiler pass to the whole compilation chain.
-Our framework takes the opposite direction and provides compositionality results for secure compilers.
-%
-%
-%To the best of our knowledge, there does not exist published work discussing the compositionality of the robust preservation criterion, the gold-standard of secure compilation.
-
-\paragraph*{Other Secure Compilation Criteria}\label{subsec:relw:seccompcrit}
-
-While this paper focuses on the robust preservation framework~\cite{abate2019jour}, other secure compilation criteria exist.
-The survey on formal approaches to secure compilation~\cite{patrignani2019survey} discusses a broad spectrum already, while this section presents a very high-level overview.
-Fully abstract compilation~\cite{abadi1999fullabstraction} states that a compiler should preserve and reflect observational equivalence between source and target programs.
-Abate \emph{et al.}~\cite{abate2021faandrc} showed that fully abstract compilers robustly preserve program properties that are either trivial or meaningless.
-As a mitigation for this, the authors presented a categorical approach based on maps of distributive laws~\cite{watanabe2002modl}, which they call many maps of distributive laws.
-Maps of distributive laws have been investigated before as a possible secure compilation criterion~\cite{tsampas2020catsc}.
-Other approaches are extensions of the compiler correctness criterion as discussed in other work~\cite{patterson2019next700} or the introduction of opaque observations~\cite{vu2021reconciling} to reconcile compiler optimisations with security.
-Note that this work also presents secure compilers that are optimising, but contrary to the other~\cite{vu2021reconciling}, provides a formal account of these in the robust preservation framework.
-% Lastly, the authors of this paper have presented ongoing work~\cite{patrignani2023blame} on a weaker robust preservation criteria based on the concept blame.
-
-\paragraph*{Memory Safety Mechanisms}\label{subsec:relw:msmechs}
-
-Different mechanisms for enforcing memory safety exist that also consider the secure compilation domain, i.e., have an active attacker model.
-For example, the ``pointers as capabilities'' principle represents pointers as machine-level capabilities~\cite{korashy2021capableptrs}, which behave in a similar fashion to capabilities by means of linear typing~\cite{morrisett2005L3}.
-The approach of this paper also uses linear typing, but differs from $L^{3}$~\cite{morrisett2005L3} in the way that functions are not first-class.
-Moreover, this paper considers an active attacker, while the work on $L^{3}$ only discusses whole programs and, thus, has no active attacker model.
-The instrumentation to ensure memory safety that this paper presents is inspired by Softbounds~\cite{nagarakatte2009soft}.
-That work inserts bounds-checks in front of pointer-dereferences and, for this to work, inserts meta-data information on pointer creation.
-Softbounds also works in a more advanced setting with structured fields accesses and also introduces a table-lookup for pointers that are stored in memory.
-This paper only considers arrays of primitive data, i.e., there are no pointers to pointers or structures.
-Several other approaches to memory-safety exist in literature, specifically as compiler instrumentations~\cite{akritidis2009baggy,younan2010paricheck,jung2021pico,shankaranarayana2023tailcheck,dhumbumroong2020boundwarden,nam2019framer,zhou2023fatptrs}, hardware-extensions~\cite{kwon2013lowfat,saileshwar2022heapcheck,chen2023flexpointer,kim2023whistle}, or programming language extensions~\cite{elliott2018checkedc,li2022formalcheckedc,jim2002cyclone,elliott2015guilt,west2005cuckoo,weis2019fyr,benoit2019uniqueness}.
-What differentiates this work from them is that this work uses known, compiler-based approaches to ensure memory-safety as a means to investigate secure compiler compositions.
-This paper does not provide efficient memory-safety, but serves as a theoretical foundation for the secure compilation domain.
-
-To extend the languages in this paper with a less restricted form of pointer arithmetic, the region colouring memory safety monitor presented in earlier work~\cite{michael2023mswasm} can be used.
-The work presenting this monitor provides an approach for the robust preservation of memory safety compiling from C to WASM.
-However, they do not discuss composition of secure compilers but rather investigate an instance of a secure compiler.
-
-\paragraph*{Cryptographic Constant Time Mechanisms}\label{subsec:relw:cctmechs}
-
-The approach to preserving cryptographic constant time in this paper is high-level, where a programming language exposes a way to switch the semantics to a data (operand) independent timing mode.
-Since identifiers in $\Lscct$ are annotated with a secrecy tag, this approach is similar to others with information flow control.
-For example, Vale~\cite{bond2017vale} uses Dafny to ensure constant-time assembly code, while Jasmin~\cite{almeida2017jasmin} makes use of the Coq proof assistant to reject non-constant-time programs.
-CT-Wasm~\cite{watt2019ctwasm} enforces constant-timeness by means of a type system.
-Different to the approach of this paper, these approaches necessitate that the programmer writes \gls*{cct} code.
-An approach to allow programmers to write more high-level code is CryptOpt~\cite{kuepper2023cryptopt}, which generates efficient target-code by means of a randomised search.
-This paper abstracts over concrete mitigation strategies and simply assumes that there is a flag to switch to a cryptographic-constant time execution mode.
-This can be realised by employing the FaCT~\cite{cauligi2019fact} compiler, which translates common non-constant time code patterns to be constant-time, and the data (object) independent timing execution mode of modern processors.
-
-\paragraph*{Speculation Safety Mechanisms}\label{subsec:relw:ssmechs}
-
-This paper uses a taint-tracking mechanism inspired by existing work~\cite{guarnieri2018spectector,fabian2022automatic}.
-These taints are used to express absence of any speculative leaks in \gls*{ss}~\cite{guarnieri2018spectector}. 
-The semantics of $\Lspec$ hardcodes the kind of speculative leaks to just SPECTRE-PHT~\cite{kocher2019spectre}, but future work could use semantics composition~\cite{fabian2022automatic} to support more variants.
-Note that our framework composes compilers and not semantics.
-
-%This work has taken inspiration in the sematics of $\Lspec$ and in \Cref{def:trace:ss} from recent formalisations~\cite{guarnieri2018spectector,fabian2022automatic}.
-% \MPin{
-%  we take SS from exorcising.
-%  we do not take SNI from spectector because it is a hypersafety.
-%  other work defined similar props. 
-% }
-% A subsequent line of work~\cite{fabian2022automatic} shows how to modularly define and compose formal semantics that models different speculative execution variants. 
-% Recent work~\cite{fabian2024lift} demonstrates how to lift existing secure compilation proofs for security properties related to speculative execution to more powerful properties, i.e., properties defined for languages with a trace model that enables more vulnerabilities, such as combinations of different speculative execution exploits.
-% To this end, they define a lifting framework and demonstrate how a secure compilation proof for a weaker property can be lifted to a stronger one. 
-% Contrary to the compositionality framework presented in this paper, these works are only concerned with composing semantics of the respective programming languages, while in this paper, the secure compilers themselves are composed.
-
-\section{Conclusion\pages{$\sfrac{1}{2}$}}\label{sec:concl}
-This paper tackles the problem of understanding what kind of security properties a secure compiler preserves, when said compiler is the combination of compiler passes that preserve possibly different security properties.
-% 
-%For this, this paper first formalised security properties of interest and their composition.
-% 
-The paper proves that composing secure compilers that preserve certain properties results in a secure compiler that preserves the composition of these properties.
-% While the presented security property composition relied on monitors that check only trace-properties, the composition of secure compilers does not make any restriction towards the kind of property involved in the composition.
-% It is subject to future work to develop techniques to verify relational hyperproperties~\cite{abate2019jour}, while the composition of hyperproperties could be very similar to the composition of ordinary properties as presented in this paper, since hyperproperties can be checked with an automata-based model-checker~\cite{beutner23hyperltl}.
-% 
-Finally, this paper defines a multi-pass compiler and proves that it preserves \gls*{specms}.
-Crucially, this paper derives the security of the multi-pass compiler from the composition of the security properties preserved by its individual passes, which include security-preserving as well as optimisation passes.
-% For future work, it is interesting to look at more sophisticated optimisation passes that, e.g., reorder events that appear on traces.
-
-\newpage
-
-\bibliographystyle{IEEEtranS}
-\bibliography{main}
-
-%%
-%% If your work has an appendix, this is the place to put it.
-\appendix
-\subsection{Existential Image}\label{subsec:extimg}
-
-\begin{definition}[Existential Image]
-% [Existential and Universal Image]
-\label{def:existential:img}\label{def:tau}
-  \[
-    \tau_\sim\left(\src{\pi}\right) := 
-      \left\{ 
-        \trg{\trace} \mid \exists \src{\trace}\ldotp \src{\trace}\sim\trg{\trace}, \text{ and } \src{\trace}\in\src{\pi} 
-      \right\}
-  \]
-  %\[ 
-  %  \sigma_\sim\left(\trg{\pi}\right) := 
-  %    \left\{ 
-  %      \src{\trace} \mid \forall \trg{\trace}\ldotp \text{if }\src{\trace}\sim\trg{\trace}, \text{ then } \trg{\trace}\in\trg{\pi} 
-  %    \right\}
-  %\]
-\end{definition}
-
-\begin{definition}[Robust Preservation with $\tau_\sim$]\label{def:rtp:tau}
-  % $\;$\\
-  % \vspace{-1em}
-  % \begin{nscenter}
-  % Compiler $\cc{\src{L}}{\trg{L}}$ robustly preserves $\class$, 
-  {$\rtptau{\cc{\src{L}}{\trg{L}}}{\src{\class}}{\sim}$}
-  %, iff 
-  $\isdef$
-    {$\forall \src{\pi}\in\src{\class}, \src{p}\in\src{L},$} if {$\rsat{\src{\progvar}}{\src{\pi}}$}, then {$\rsat{\cc{\src{L}}{\trg{L}}\left(\src{p}\right)}{\tau_\sim\left(\src{\pi}\right)}$}.
-  % \end{nscenter}
-\end{definition}
-
-\begin{theorem}[Composition of Secure Compilers w.r.t. $\tau$]\label{thm:rtpsim:tau}
-  $\;$
-
-  If {$\rtptau{\cc{\src{L}}{\trg{L}}}{\src{\class[_{1}]}}{\sim_1}$}, {$\rtptau{\cc{\trg{L}}{\obj{L}}}{\tilde{\tau}_{\sim_1}\left(\src{\class[_2]}\right)}{\sim_2}$}, and {$\wfctau{\sim_1}{\src{\class[_2]}}$}, \\ then {$\rtptau{\cc{\src{L}}{\trg{L}}\circ\cc{\trg{L}}{\obj{L}}}{\src{\class[_{1}]}\cap\src{\class[_{2}]}}{\sim_1\bullet\sim_2}$}.
-\Coqed
-\end{theorem}
-
-\begin{corollary}[Swapping Secure Compiler Passes]\label{corr:swappable:tau}
-  $\;$
-
-  If {$\rtptau{\cc[_{1}]{\trg{L}}{\trg{L}}}{\src{\class[_{1}]}}{\sim_1}$ and $\rtptau{\cc[_{2}]{\trg{L}}{\trg{L}}}{\src{\class[_{2}]}}{\sim_2}$}, %
-  {$\wfctau{\sim_1}{\src{\class[_2]}}$ and $\wfctau{\sim_2}{\src{\class[_1]}}$}, %
-  and {$\tilde{\tau}_{\sim_1}\left(\src{\class[_2]}\right)=\src{\class[_2]}$ as well as $\tilde{\tau}_{\sim_2}\left(\src{\class[_1]}\right)=\src{\class[_1]}$},
-  then {$\rtptau{\cc[_{1}]{\trg{L}}{\trg{L}}\circ\cc[_{2}]{\trg{L}}{\trg{L}}}{\src{\class[_{1}]}\cap\src{\class[_{2}]}}{\sim_1\circ\sim_2}$ and $\rtptau{\cc[_{2}]{\trg{L}}{\trg{L}}\circ\cc[_{1}]{\trg{L}}{\trg{L}}}{\src{\class[_{2}]}\cap\src{\class[_{1}]}}{\sim_2\circ\sim_1}$}.
-\Coqed
-\end{corollary}
-
-\subsection{Secure Upper and Lower Composition}\label{sec:other-compos}
-Besides sequential composition, there are two other compositions, namely an {\em upper}, i.e., a compiler that takes multiple inputs and yields one output, and a {\em lower} composition, i.e., a compiler that takes one input and yields multiple outputs.
-We {define the upper composition $\cc{\src{L}+\obj{L}}{\trg{L}}$} as follows:
-Given a program \texttt{p}, its compiled counterpart is obtained by {plugging \texttt{p} into $\cc{\src{L}}{\trg{L}}$ if $\texttt{p}\in\src{L}$} or by {plugging \texttt{p} into $\cc{\obj{L}}{\trg{L}}$ if $\texttt{p}\in\obj{L}$}.
-\begin{definition}[Upper Composition]
-  \[
-    \text{{$\cc{\src{L}+\obj{L}}{\trg{L}}$}}\isdef
-  % Given \texttt{p}, yield
-  \lambda \texttt{p}\ldotp
-  \left\{\mbox{\begin{tabular}{c}
-    {{if $\texttt{p}\in\src{L}$, then $\cc{\src{L}}{\trg{L}}(\texttt{p})$}} \\
-    \mbox{{if $\texttt{p}\in\obj{L}$, then $\cc{\obj{L}}{\trg{L}}(\texttt{p})$}} \\
-  \end{tabular}}\right. 
-  \]
-\end{definition}
-
-Examples of this are present in industry:
-Consider the Java Virtual Machine bytecode $\trg{JVM BC}$, which is a popular target for programming language designers due to its high performance and relevance in industry.
-Compilers for several programming languages have it as their target language, some popular instances are $\src{Java}$ and $\obj{Kotlin}$.
-Technically speaking, they both compile to class files and $\obj{Kotlin}$ objects are considered to be the same as $\src{Java}$ objects at that point.
-Both languages can be used at the same time in one project~\cite{androidstudio}.
-A compiler that accepts both $\src{Java}$ and $\obj{Kotlin}$ code translating to the same target language or intermediate representation performs a kind of {\em upper} composition.
-Now, the following theorem tells us what happens if these are secure:
-Given {$\cc{\src{L}}{\trg{L}}$ robustly preserves $\class[_{1}]$} and {$\cc{\obj{L}}{\trg{L}}$ robustly preserves $\class[_{2}]$}, it follows that {their upper composition $\cc{\src{L}+\obj{L}}{\trg{L}}$ robustly preserves the intersection of classes $\class[_{1}]$ and $\class[_{2}]$}.
-
-\begin{theorem}[Upper Composition of Secure Compilers]\label{thm:urtp}
-  $\;$
-
-  If {$\rtp{\cc{\src{L}}{\trg{L}}}{\class[_{1}]}$} and {$\rtp{\cc{\obj{L}}{\trg{L}}}{\class[_{2}]}$}, then {$\rtp{\cc{\src{L}+\obj{L}}{\trg{L}}}{\class[_{1}]\cap\class[_{2}]}$}. %\Coqed
-\end{theorem}
-
-Dually, the {\em lower} composition is concerned about compilers that accept the same source but yield different target languages. %and the same theoretical results apply.
-{Define the lower composition $\cc{\src{L}}{\trg{L}+\obj{L}}$} as follows:
-Given a program $\src{p}$, its compiled counterpart is obtained by {plugging $\src{p}$ into $\cc{\src{L}}{\trg{L}}$} or by {plugging \texttt{p} into $\cc{\src{L}}{\obj{L}}$}, respectively, {based on the internal decision}.
-\begin{definition}[Lower Composition]
-  \[
-    \text{{$\cc{\src{L}}{\trg{L}+\obj{L}}$}}\isdef
-  % Given $L$ and $\src{p}$, yield
-  \lambda \src{p}, L\ldotp
-  \left\{\mbox{\begin{tabular}{c}
-    {{if $L=\trg{L}$, then} {$\cc{\src{L}}{\trg{L}}(\src{p})$}} \\
-    \mbox{{if $L=\obj{L}$, then} {$\cc{\src{L}}{\obj{L}}(\src{p})$}} \\
-  \end{tabular}}\right.\]
-\end{definition}
-
- Consider two compilers both accepting $\src{LLVM IR}$~\cite{lattner2004llvm} and one of them emits $\trg{x86\_64}$, while the other emits $\obj{ARMv8}$.
- It is intuitive that they are in some sense composed in the LLVM framework, but the decision of when to use one over the other is inherently {\em internal} to the formalisation effort of this kind of composition.
- For example, the user of this compiler provides an explicit flag that instructs to emit $\trg{x86\_64}$ or the framework itself detects the target platform via heuristics, such as supported instructions.
-
- The following theorem demonstrates what happens if the involved compilers are secure:
-Given {$\cc{\src{L}}{\trg{L}}$ robustly preserves $\class[_{1}]$} and {$\cc{\src{L}}{\obj{L}}$ robustly preserves $\class[_{2}]$}, it follows that {their lower composition $\cc{\src{L}}{\trg{L}+\obj{L}}$ robustly preserves the intersection of classes $\class[_{1}]$ and $\class[_{2}]$}.
-
-\begin{theorem}[Lower Composition of Secure Compilers]\label{thm:lrtp}
-  $\;$
-
-  If {$\rtp{\cc{\src{L}}{\trg{L}}}{\class[_{1}]}$} and {$\rtp{\cc{\obj{L}}{\trg{L}}}{\class[_{2}]}$}, then {$\rtp{\cc{\src{L}}{\trg{L}+\obj{L}}}{\class[_{1}]\cap\class[_{2}]}$}. % \Coqed
-\end{theorem}
-
-Either way, the theoretical results suggest that it is possible to always find a ``most-general'', secure compiler, given two secure compilers, that robustly preserves the least-upper bound of the classes involved in their compilation process.
-
-%Future work intends to use upper and lower composition as means to build a secure compiler incrementally
-%This follows current non-formal-methods based compiler engineering, such as LLVM~\cite{lattner2004llvm}, where languages such as \texttt{C++} and \texttt{Fortran} translate into LLVM IR and languages such as \texttt{x86} or \texttt{ARMv8} assembly variants translate out of LLVM IR.
-
-% no coq, because above definitions are non-formal. 
-%   How would you precisely define the lower compo compiler, given two compilers?
-
-%\section{Cryptographic Constant Time}\label{sec:cct}
-
-\end{document}
-\endinput

### 2025-02-05 icfp/sections/conclusion.tex
Move Sections around

--- 2024-09-24 csf2/main.tex
@@
-% Fixing: Too many math alphabets used in version normal
-\newcommand\hmmax{0}
-\newcommand\bmmax{0}
-\documentclass[dvipsnames,conference]{IEEEtran}
-
-\usepackage[T1]{fontenc}
-\usepackage[scaled=.83]{beramono}
-
-\usepackage[colorinlistoftodos]{todonotes}
-\usepackage[inference]{semantic}
-\usepackage[switch]{lineno}
-\usepackage{halloweenmath}
-\usepackage{fontawesome5}
-\usepackage{listofitems}
-\usepackage{breakcites}
-\usepackage{glossaries}
-\usepackage{hyperref}
-\usepackage{cleveref}
-\usepackage{stmaryrd}
-\usepackage{marvosym}
-\usepackage{listings}
-\usepackage{amssymb}
-\usepackage{amsmath}
-\usepackage{nameref}
-\usepackage{amsthm}
-\usepackage{xspace}
-\usepackage{xfrac}
-\usepackage{tikz}
-\usepackage{soul}
-\usepackage{bm}
-\usepackage{paralist}
-
-\hypersetup{
-    colorlinks,
-    linkcolor={red!50!black},
-    citecolor={blue!50!black},
-    urlcolor={blue!80!black}
-}
-%\renewcommand\UrlFont{\color{blue}\rmfamily}
-%\newcommand{\url}[1]{\lstinline{#1}}
-
-\setul{0.95ex}{0.3ex}
-
-\input{cmds.tex}
-
-% allow page breaks inside multiline alignment displays
-\allowdisplaybreaks
-
-\Crefname{exampleenv}{Example}{Examples}
-
-\theoremstyle{definition}
-\newtheorem{exampleenv}{Example}[section]
-\newtheorem{lemma}{Lemma}[section]
-\newtheorem{theorem}{Theorem}[section]
-\newtheorem{corollary}{Corollary}[section]
-\newtheorem{definition}{Definition}[section]
-
-\loadglsentries{acronyms}
-\makeglossaries
-
-\AtBeginDocument{%
-  \providecommand\BibTeX{{%
-    \normalfont B\kern-0.5em{\scshape i\kern-0.25em b}\kern-0.8em\TeX}}}
-\makeatletter
-\newcommand{\linebreakand}{%
-  \end{@IEEEauthorhalign}
-  \hfill\mbox{}\par
-  \mbox{}\hfill\begin{@IEEEauthorhalign}
-}
-\makeatother
-
-\renewcommand\thelinenumber{\color{red}\arabic{linenumber}}
-\begin{document}
-\linenumbers
-
-\title{
-  Secure Composition 
-  \\ 
-  of Robust and Optimising Compilers
-  % \vspace{-1.5em}
-}
-
-% \author{
-% Matthis Kruse\orcidID{0000-0003-4062-9666}
-% \and
-% Michael Backes
-% \and
-% Marco Patrignani
-% }
-% \authorrunning{M. Kruse et al.}
-
-\maketitle
-
-\begin{abstract}
-% \MPpost{
-%   this abstract fails in setting up the robustness argument.
-%   it'll be a good exercise for you to take inspiration from this, but then write something as concise but with a more direct focus on robust properties
-% }
-% context
-To ensure that secure applications do not leak their secrets, they are required to uphold several security properties such as spatial and temporal memory safety, cryptographic constant time, as well as speculative safety.
-% need
-Existing work shows how to enforce these properties individually, in an architecture-independent way, by using secure compiler passes that each focus on an individual property.
-% task
-Unfortunately, given two secure compiler passes that each preserve a possibly different security property, it is unclear what kind of security property is preserved by the composition of those secure compiler passes.
-%there is no way to tell what kind of security property will the composition of those secure compilers preserve.
-
-% object
-This paper is the first to study what security properties are preserved across the composition of different secure compiler passes.
-% findings
-Starting from a general theory of property composition for security-relevant properties (such as the aforementioned ones), this paper formalises a theory of composition of secure compilers.
-Then, it showcases this theory on a secure multi-pass compiler that preserves the aforementioned security-relevant properties.
-% conclusion
-Crucially, this paper derives the security of the multi-pass compiler from the composition of the security properties preserved by its individual passes, which include security-preserving as well as optimisation passes.
-% 
-From an engineering perspective, this is the desirable approach to building secure compilers.
-\begin{center}\small\it
-	{This paper uses syntax highlighting accessible to both colourblind and black \& white readers.
-	% ~\cite{patrignani2020use}.
-% 	% Specifically, it makes use of a $\src{blue}$, $\src{sans\text{-}serif}$ font for a $\src{source}$,
-% 	% a $\trg{red}$, $\trg{bold}$ font for an $\trg{intermediate}$,
-% 	% and a $\obj{green}$, $\obj{teletype}$ font for a $\obj{target}$ language.	
-	For a better experience, please print or view in colours.
-	}
-\end{center}
-\end{abstract}
-
-\begin{IEEEkeywords}
-  memory-safety, cryptographic constant-time, secure compilation, robustness
-\end{IEEEkeywords}
-
-\section{Introduction\pages{4}}\label{sec:introduction}
-
-% \gls*{ms} is a security property that ensures that there are no out-of-bounds accesses as well as no reads or writes to uninitialised memory.
-% Programs that do not attain memory safety, can easily have their private data be leaked to attackers or be corrupted~\cite{lemay2021ccc}.
-% Because of this, \gls*{ms} is a requirement for programs to attain an even stronger security property, namely \gls*{cct}.
-% Programs that attain \gls*{cct} are secure w.r.t.
-timing attacks, and this is the gold standard for secure applications, including cryptographic ones, since timing attacks can expose secrets via a simple performance analysis~\cite{kocher1996timing}.
-% Timing attacks have become more and more relevant recently, especially given the era of cloud-computing~\cite{aviram2010cloudtime,kumar2019cloudsecsurvey}, where programs share the same hardware but are supposed to run in a sandboxed environment that does not leak secrets.
-% Unfortunately, several attacks within such a cloud environment have been demonstrated as viable strategies to extract secrets~\cite{mehmet2015getoff,flowers2022zeroday,atya2019catchme,venkatanathan2015placevul}.
-
-% Even though mitigations for such attacks exist~\cite{bond2017vale,almeida2017jasmin}, some compilers may optimise these mitigations away~\cite{barthe2018sec}.
-% This is of course not what programmers expect and to ensure such mitigations persist in the compiled code, one can use a {\em secure} compilers.
-% More generally, secure compilers should preserve security properties even when the program runs in a hostile context such as the cloud one, i.e., where it interacts with other possibly malicious programs.
-% An example of such secure compilers are those that ascribe to the theory of robust compilation~\cite{abate2019jour,abate2021extacc,patrignani2021rsc}, which takes this interaction between components and potentially malicious contexts into account.
-
-% Unfortunately, it is unclear if and how robust compilation supports compositionality, and compilation passes are typically developed as a standalone module to be plugged into a compilation pipeline and composed with other passes.
-% Different kinds of compilation passes exist, e.g., there are usually several compilation passes that perform optimisations~\cite{lattner2004llvm,googlev8,androidstudio,kuepper2023cryptopt,manjikian1997fusion,wegman1991ccp} or source-code instrumentations~\cite{nagarakatte2009soft,nagarakatte2010cets,akritidis2009baggy,dhumbumroong2020boundwarden,jung2021pico,nam2019framer,shankaranarayana2023tailcheck,younan2010paricheck,zhou2023fatptrs,bond2017vale,almeida2017jasmin,kuepper2023cryptopt,cauligi2019fact} that enforce that the compiled program fulfils a property of interest.
-% Without a framework for composing secure compiler passes, it is not possible to enable separation of concerns, e.g., to have a secure compilation pass that ensures \gls*{ms} that is developed independently of another secure pass for \gls*{cct} as suggested before and as showcased in \Cref{ex:strncpy}.
-% \MPpost{
-%   this intro does a poor job at introducing the robust compilation argument.
-%   perhaps it'd be best to start from robust properties, which motivates the use of robust compilers?
-%   idk
-%   as for the abstract, this is a good starting point for you to make a better intro
-% }
-
-\gls*{ms} is a security property obtained by composing \gls*{sms}, which ensures array accesses are all within bounds, and \gls*{tms}, which ensures pointers are only used when they are valid~\cite{azevedo2018meaningsofms,jim2002cyclone,necula2005ccured,nagarakatte2010cets,nagarakatte2009soft,akritidis2009baggy,michael2023mswasm}.
-\gls*{cct} is a security property that ensures sensitive data is not leaked via timing side-channels~\cite{kocher1996timing} and \gls*{ss} is a security property that enforces the same but under a speculative semantics~\cite{guarnieri2018spectector,fabian2022automatic} that captures speculative execution attacks such as spectre~\cite{kocher2019spectre}.
-Together, \gls*{sms}, \gls*{tms}, \gls*{cct} and \gls*{ss}, yield \gls*{specms}, which is the gold standard of security properties for secure applications.
-Programs attaining \gls*{specms} do not leak sensitive data either through erroneous memory accesses, nor through timing side-channels, even under speculative execution.
-\Cref{ex:strncpy} below discusses how these security properties can be enforced by compiler passes~\cite{bond2017vale,almeida2017jasmin}, to ensure programmers need not be aware of the architectural details of where their code will run.
-
-\begin{exampleenv}[strncpy]\label{ex:strncpy}
-Consider the following \texttt{C} function \texttt{strncpy} that copies a null-terminated string \texttt{src} into \texttt{dst} up to a length of \texttt{n} characters.
-This function is subject to a subtle \gls*{sms} vulnerability: the bounds check \texttt{i < n} should happen {\it before} the access to memory location \texttt{src[i]}: otherwise
-% If \texttt{n} is at least as large as the length of the array \texttt{src}, 
-the memory location past the last element is leaked to an attacker.
-\begin{lstlisting}[language=c,basicstyle=\small\ttfamily,morekeywords={size_t}]
-void strncpy(size_t n, char *dst, char *src) {
-  for(size_t i = 0; src[i] != '\0' && i < n; ++i) {
-    dst[i] = src[i];
-  }
-}
-\end{lstlisting}
-% Assume that the length of the array pointed by \texttt{dst} is at least \texttt{n}, and \texttt{src} is terminated by \texttt{'\textbackslash 0'}.
-% Even under these assumptions, there is a 
-To prevent this vulnerability, one can use a compilation pass that enforces \gls*{sms}, such as Softbounds~\cite{nagarakatte2009soft} or BaggyBounds~\cite{akritidis2009baggy}.
-The most na\"ive solution in this case is to insert bounds-checks in front of every access to memory.
-
-Because of timing attacks, addressing \gls*{sms} is not enough to make \texttt{strncpy} secure.
-In fact, the loop can terminate early, as soon as the string-terminating character \texttt{'\textbackslash 0'} is encountered, thus making program execution time proportional to the length of the array pointed by \texttt{src}, and violating the \gls*{cct} property.
-% Often, programmers rewrite programs like these to have loops that run with a fixed number of iterations.
-Also in this case there exist compiler passes that can rewrite such programs into \gls*{cct} ones~\cite{cauligi2019fact}.
- % so that they are no longer vulnerable to timing attacks.
-%
-Finally, even with these precautions, code is not run in isolation, so a malicious attacker could supply code that interacts with \texttt{strncpy} and trigger a violation of either \gls*{ms} or \gls*{cct} by calling \texttt{strncpy} with an argument for \texttt{src} that points to uninitialised memory.
-This would, in turn, trigger a series of reads from uninitialised memory, which is an immediate \gls*{ms} violation with devastating real-world consequences~\cite{uninit-0,uninit-1,uninit-2,uninit-3,uninit-4}.
-% An easy fix to sidestep this is to use non-nullable pointers as parameters instead, e.g., as presented in Cyclone~\cite{jim2002cyclone}.
-\end{exampleenv}
-
-  % last sentence is before the example, which talks about compiler passes enforcing props
-Whether or not compiler passes enforce certain security properties, it is important for a secure compiler to consider partial programs that interact with potentially malicious code, since the latter may lead to, e.g., memory-safety issues in the considered partial program.
-Robust compilers~\cite{abate2019jour} are a form of secure compilers that preserve security properties even in the presence of arbitrary attackers interacting with compiled code.
-Thus, robust compilers can be used to prevent vulnerabilities resulting from uninitialised memory (as well as many other ones), e.g., by targeting capability-based languages such as CHERI~\cite{woodruff2014CHERI}, Arm Morello~\cite{arm-morello}, or MSWasm~\cite{michael2023mswasm}, where the compiler relies on capabilities to check that pointers are always initialised.
-
-% To ensure these compiler-preserved properties still hold even in the presence of arbitrary attackers interacting with compiled code, one can use secure compilers in the form of robust compilers~\cite{abate2019jour,abate2021extacc,patrignani2021rsc}.
-
-Unfortunately, given secure compiler passes that each preserve a possibly different security property, there is no way to tell what kind of security property will the composition of those secure compilers preserve.
-% Different kinds of compilation passes exist, e.g., there are usually several compilation passes that perform optimisations~\cite{lattner2004llvm,googlev8,androidstudio,kuepper2023cryptopt,manjikian1997fusion,wegman1991ccp} or source-code instrumentations~\cite{nagarakatte2009soft,nagarakatte2010cets,akritidis2009baggy,dhumbumroong2020boundwarden,jung2021pico,nam2019framer,shankaranarayana2023tailcheck,younan2010paricheck,zhou2023fatptrs,bond2017vale,almeida2017jasmin,kuepper2023cryptopt,cauligi2019fact} that enforce that the compiled program fulfils a property of interest.
-Worse, without a framework for composing secure compiler passes, it is not possible to enable separation of concerns, e.g., to have a secure compilation pass that ensures \gls*{ms} that is developed independently of another secure pass for \gls*{cct}, that is developed independently of other passes, such as optimisation ones.
-
-This paper introduces a framework for reasoning about the composition of secure and optimising compiler passes %akin to those of \Cref{ex:strncpy}\MP{there aren't really passes there} 
-and it showcases the power of this framework by instantiating it on a multi-pass compilation chain.
-To this end, this paper first discusses how to compose security properties, such as \gls*{tms} and \gls*{sms} into \gls*{ms}, and then adds \gls*{cct} as well as \gls*{ss} into the mix to obtain \gls*{specms}.
-%Then, this paper defines compiler composition and formalises that given two passes that securely preserve two (possibly distinct) properties, their composition securely preserves the composition of those properties.
-The paper then defines several secure compiler passes, where each is either preserving a different security property from the list above
-% (\gls*{tms}, \gls*{sms}, \gls*{scct}, \gls*{ss}) 
-or performing a security-preserving optimisation.
-% , (e.g., applying \gls*{cf} or \gls*{dce}).
-Finally, this paper shows that composing these secure compiler passes into a multi-pass compilation chain results in the end-to-end (robust) preservation of \gls*{specms}.
-Crucially, this paper derives the security of the multi-pass compiler from the composition of the security properties preserved by its individual passes.
-% Crucially, this end-to-end security preservation is obtained by composing the security property preserved by each individual pass, as dictated by the framework of this paper.
-This result showcases how the framework allows the kind of formal security reasoning that compiler writers already want (and already do), obtaining precise, compositional security reasoning while providing minimal (and modular) proof effort.
-%
-In summary, this paper makes the following contributions:
-% \begin{itemize}[leftmargin=0em,nosep,label=$\blacktriangleright$]
-\begin{asparaitem}[$\blacktriangleright$]
-  \item %
-        %\Cref{sec:sequential} studies different forms of secure compiler compositions.
-        This paper takes the secure compilation framework of Abate et al.~\cite{abate2021extacc} and extends it for compositionality. (\Cref{sec:sequential}).
-        %Concretely, it investigates upper and lower compositions, as well as the sequential composition.
-        %For this, this paper studies compiler composition as well as compilers with multiple input languages or multiple output ones, as used in existing compilation chains.
-        %
-        This paper proves that starting from two compilers that preserve two (possibly distinct) properties, their composition preserves roughly the intersection of those properties.
-        % For example, a compiler that is secure w.r.t. \gls*{tms} can be composed with a compiler that is secure w.r.t. \gls*{sms}, resulting in a compiler that is secure w.r.t. \gls*{ms}.
-        % 
-        % From this, the paper proves that the order of composition of sequential compiler passes is irrelevant for security.
-        % % 
-        % This is crucial for reordering individual optimisation passes and thus generating secure and efficient code.
-        %
-        %Finally, this paper extends parts of the compositionality framework to support different trace-models between compilation passes.
-        % \MP{this isn't really true, is it?. the order matters for the relations' lemmas}
-        Then, this paper identifies which conditions make the composition of secure compilers meaningful from a security perspective, and which other conditions allow passes to be swapped without losing security meaning.
-
-  \item %
-        This paper presents a case-study with five programming languages showcasing the previous contribution (\Cref{sec:casestud:defs,sec:casestud:rtp}).
-        To this end, it presents a compilation chain consisting of six passes that ultimately preserve \gls*{specms} by means of composing four secure passes that \emph{individually} preserve \gls*{tms}, \gls*{sms}, \gls*{scct}---an enforceable version of \gls*{cct}---, and \gls*{ss}, and two passes which are well-known optimisations \gls*{dce} and \gls*{cf}.
-        %Furthermore, the chain includes two optimisation passes: one performs \gls*{dce} and the other \gls*{cf}.
-        The formalisation of this case study showcases the power of the presented framework: the divide-and-conquer approach to software engineering is a viable strategy even for the development of secure compilers.
-%        This demonstrates that the usage of the presented framework is both simpler and more expressive compared to one monolithic security proof of the whole compilation chain.
-
-  \item %
-  		This paper provides formal proofs of the security properties preserved by each of the presented passes. 
-  		% 
-  		Additionally this paper analyses each pass in terms of their compatibility with each other (\Cref{sec:formalities}), proving how does each pass fulfill the conditions that make their composition meaningful from a security standpoint.
-        % In doing so, the paper provides formal insights into some technicalities of the concrete proofs necessary as technical setup to reap the benefits of the presented compositionality framework.
-
-  \item The key contributions of this paper are formalised in the Coq proof assistant and the paper indicates this with \CoqSymbol.
-\end{asparaitem}
-% 
-This paper starts by introducing relevant notions of security properties and secure compilation (\Cref{sec:background}),
-and discusses related work (\Cref{sec:relwork}) before concluding (\Cref{sec:concl}).
-
-% \contribution{Technical Report} 
-The omitted formal details, lemmas and proofs, as well as the Coq formalisation are available as supplementary material.
-% \MKin{cite/link}
-
-\section{Background: Properties and Secure Compilers\pages{1}}\label{sec:background}
-
-To introduce the security argument of this paper, this section defines (security) properties, their satisfaction, and their robust satisfaction (i.e., satisfaction in the presence of an active attacker; \Cref{subsec:bg:tprop}).
-Then, borrowing from existing work~\cite{abate2019jour,abate2021extacc}, this section introduces secure compilers as compilers that preserve robust property satisfaction (\Cref{subsec:bg:rtp}).
-
-\subsection{Properties and (Robust) Satisfaction}\label{subsec:bg:tprop}
-
-This paper employs the security model where programs are written in a language whose semantics emits events $\event$.
-Events include security-relevant actions (e.g., reading from and writing to memory, as detailed in \Cref{sec:compprop}) and the unobservable event $\emptyevent$.
-As programs execute, their emitted events are concatenated in traces $\trace$, which serve as the description of the behaviour of a program.%
-\footnote{
-Throughout the paper, sequences are indicated with an overbar (i.e., $\trace$), empty sequences with $\hole{\cdot}$, and concatenation of sequences $\trace[_{1}],\trace[_{2}]$ as $\trace[_{1}]\cdot\trace[_{2}]$.
-Prepending element $\event$ to a sequence $\trace$ uses the same notation: $\event\cdot\trace$.
-}
-
-Properties $\pi$ are sets of traces of admissible program behaviours, ascribing what said property considers valid.
-% 
-The set of all properties can be partitioned into different {\em classes} ($\class$), i.e., safety, liveness, and neither safety nor liveness~\cite{clarkson2008hyper}, so a class is simply a set of properties.
-% 
-The compositionality framework (\Cref{sec:sequential}) presented in this paper considers arbitrary classes, while the case-study (\Cref{sec:casestud:rtp}) fixes them to concrete instances of safety properties, since it is decidable whether a trace satisfies a safety property with just a finite trace (i.e., a \emph{prefix}).
-As an example, consider the trace:
-$$\ev{Dealloc\ \loc}\cdot\ev{Read\ \loc\ 1729}\cdot\dots$$
-which describes the interaction with a memory where the deallocation of an address $\loc$ precedes a read (of some value $1729$) at that address in memory.
-% 
-This program behaviour is insecure w.r.t.
-a canonical notion of (temporal) memory safety dictating no use-after-frees of pointers~\cite{nagarakatte2010cets,azevedo2018meaningsofms}, because it reads from a memory location that was freed already.
-The previous finite trace prefix is enough to decide that the trace does not satisfy \gls*{tms} and there is no way to append events to this prefix which would result in the trace being admissible.
-
-In the following, the execution of a whole program $\wholeprogvar$ that terminates in state $\runtimetermvar$ according to the language semantics and produces trace $\trace$ is written as $\progstepto{\wholeprogvar}{\runtimetermvar}{\trace}$.
-With this, we defined property satisfaction as follows:
-\bul{a whole program $\wholeprogvar$ satisfies a property $\pi$} iff \iul{$\wholeprogvar$ yields a trace $\trace$} such that \oul{$\trace$ satisfies $\pi$}.
-
-\begin{definition}[Property Satisfaction]\label{def:propsat}
-$\;$
-
-  \begin{nscenter}
-    \bul{$\sat{\wholeprogvar}{\pi}$}
-    $\isdef$
-    if \iul{$\forall\runtimetermvar\ \trace,\progstepto{\wholeprogvar}{\runtimetermvar}{\trace}$},
-    then \oul{$\trace\in\pi$}.
-  \end{nscenter}
-\end{definition}
-
-Property satisfaction is defined on whole programs, i.e., programs without missing definitions.
-Thus, from a security perspective, \Cref{def:propsat} considers only a passive attacker model, where the attacker observes the execution and, e.g., retrieves secrets from that.
-To consider a stronger model similarly to what existing work does~\cite{abate2019jour,abate2021extacc,maffeis2008code-carrying,gordon2003authenticity,fournet2007authorization,bengtson2011refine,backes2014uniontyps,michael2023mswasm,swasey2017robust,sammler2019benefits}, we extend the concept of satisfaction with {\em robustness}.
-Robust satisfaction considers partial programs $\progvar$, i.e., components with missing imports, which cannot run until said imports are fulfilled.
-To remedy this, {\em linking} takes two partial programs $\progvar[_{1}],\progvar[_{2}]$ and produces a whole program $\wholeprogvar$, i.e., $\link{\progvar[_{1}]}{\progvar[_{2}]}=\wholeprogvar$.
-As typically done in works that consider the execution of partial programs~\cite{abate2019jour,devriese2018parametricity,patrignani2021rsc,korashy2021capableptrs,strydonck2019lincap,devriese2017modular,bowman2015noninterference,ahmed2011equivcps,patterson2017linkingtyps},
-this paper assumes that whole programs are the result of linking partial programs referred to as {\em context} ($\ctx$) and {\em component} ($\comp$).
-The context is an arbitrary program and thus has the role of an {\em attacker} that can interact with the component by means of any features the programming language has, and the component is what is security-relevant.
-%In this work, the semantics of the programming language is expected to differentiate between the component and the context.
-With this, \Thmref{def:propsat} can be extended as follows: for \bul{a component $\progvar$ to robustly satisfy a property $\pi$}, take an \iul{attacker context $\contextvar$ and link it with $\progvar$}, \oul{the resulting whole program must satisfy $\pi$}.
-
-\begin{definition}[Robust Satisfaction]\label{def:proprsat}
-  $\;$
-
-  \begin{nscenter}
-  \bul{$\rsat{\progvar}{\pi}$}
-  $\isdef$ \iul{$\forall \contextvar\ \wholeprogvar$, if $\link{\contextvar}{\progvar}=\wholeprogvar$}, then \oul{$\sat{\wholeprogvar}{\pi}$}.
-  \end{nscenter}
-\end{definition}
-
-% \begin{exampleenv}[Double Free in Bluetooth Subsystem]
-%   Consider \texttt{CVE-2021-3564}~\cite{doublefree-bluetooth}, one of many submissions for a double-free vulnerability.
-%   The vulnerability arises due to a race condition where the context-level function \texttt{hci\_cmd\_work} was not expected to behave maliciously, since it resides in the same source-code repository where the vulnerability occurs.
-%   Nevertheless, the component-level code of \texttt{hci\_dev\_do\_open} is linked with \texttt{hci\_cmd\_work} and does not atomically check whether a pointer has been freed already:
-%   Therefore, \texttt{hci\_dev\_do\_open} does not satisfy the no-double-frees property robustly, since there is an implementation for \texttt{hci\_cmd\_work} that leads to a violation of that property when linked with \texttt{hci\_dev\_do\_open}.
-% \end{exampleenv}
-
-\subsection{Secure Compilers}\label{subsec:bg:rtp}
-
-A {\em compiler} ($\cc{\src{L}}{\trg{L}}$) translates syntactic descriptions of programs from a {\em source} ($\src{L}$) into a {\em target} ($\trg{L}$) programming language.
-This translation is considered {\em correct} if it is semantics-preserving.
-That is, for a whole program $\src{\wholeprogvar}$, the compiler should relate the $\src{L}$ semantics of $\src{\wholeprogvar}$ with the semantics of $\trg{L}$ of the compiled counterpart of $\src{p}$ in such a way that they are ``compatible''.
-Unfortunately, correct compilers may be insecure compilers~\cite{patrignani2019survey,kennedy2006secure.net,abadi1999protect,ahmed2018dagstuhl} and programs translated by insecure compilers can violate security properties that the programmer assumes to hold.
-This is why {\em robust preservation} is a good candidate as a compiler-level security property~\cite{abate2019jour}.
-
-This paper uses a general notion of robust preservation~\cite{abate2021extacc} that considers compilers that use languages with different trace models. 
-To this end, considering a source trace $\src{\trace}$ and a target trace $\trg{\trace}$, there is a relation ($\sim$) describing the effect of a corresponding compiler (see \Cref{sec:casestud:rtp}). 
-This relation induces the following two projection functions~\cite{abate2021extacc}: (1) the \emph{existential image} $\tau_\sim\left(\src{\pi}\right)$ and (2) the \emph{universal image} $\sigma_\sim\left(\trg{\pi}\right)$.
-These projections map source-level (resp. target-level) properties to target-level (resp. source-level) properties in a way that identifies the ``same'' property across languages. 
-The case study of this paper uses the universal image, since some considered properties, such as \gls*{ss}, are not definable in a higher-level language that, e.g., does not model speculation.
-% For self-containment, the universal image is defined as follows~\cite{abate2021extacc}:
-% 
-\begin{definition}[Universal Image]
-% [Existential and Universal Image]
-\label{def:universal:img}\label{def:sigma}
-  % \[
-  %   \tau_\sim\left(\src{\pi}\right) := 
-  %     \left\{ 
-  %       \trg{\trace} \mid \exists \src{\trace}\ldotp \src{\trace}\sim\trg{\trace}, \text{ and } \src{\trace}\in\src{\pi} 
-  %     \right\}
-  % \]
-  \[ 
-    \sigma_\sim\left(\trg{\pi}\right) := 
-      \left\{ 
-        \src{\trace} \mid \forall \trg{\trace}\ldotp \text{if }\src{\trace}\sim\trg{\trace}, \text{ then } \trg{\trace}\in\trg{\pi} 
-      \right\}
-  \]
-\end{definition}
-\noindent
-With this projection function, we define a more general version of robust preservation as follows~\cite{abate2021extacc}.
-% For a \bul{compiler $\cc{\src{L}}{\trg{L}}$ to robustly preserve a class of source properties $\src{\class}$}, if for any \rul{property $\src{\pi}$ of that class $\src{\class}$ and programs $\src{p}$ written in $\src{L}$} where \iul{$\src{\progvar}$ robustly satisfies $\src{\pi}$}, then \oul{the compilation of $\src{\progvar}$, $\cc{\src{L}}{\trg{L}}\left(\src{p}\right)$, must robustly satisfy $\tau_\sim\left(\src{\pi}\right)$}.
-% 
-% \begin{definition}[Robust Preservation with $\tau_\sim$]\label{def:rtp:tau}
-%   % $\;$\\
-%   % \vspace{-1em}
-%   % \begin{nscenter}
-%   % Compiler $\cc{\src{L}}{\trg{L}}$ robustly preserves $\class$, 
-%   \bul{$\rtptau{\cc{\src{L}}{\trg{L}}}{\src{\class}}{\sim}$}
-%   %, iff 
-%   $\isdef$
-%   \rul{$\forall \src{\pi}\in\src{\class}, \src{p}\in\src{L},$} if \iul{$\rsat{\src{\progvar}}{\src{\pi}}$}, then \oul{$\rsat{\cc{\src{L}}{\trg{L}}\left(\src{p}\right)}{\tau_\sim{\left(\src{\pi}\right)}}$}.
-%   % \end{nscenter}
-% \end{definition}
-% 
-A \bul{compiler $\cc{\src{L}}{\trg{L}}$ robustly preserves a class of target properties $\trg{\class}$}, if for any \rul{property $\trg{\pi}$ of class $\trg{\class}$ and programs $\src{p}$}, where \iul{$\src{\progvar}$ robustly satisfies $\sigma_\sim\left(\trg{\pi}\right)$}, \oul{the compilation of $\src{\progvar}$, we have that $\cc{\src{L}}{\trg{L}}\left(\src{p}\right)$ robustly satisfies $\trg{\pi}$}.
-
-\begin{definition}[Robust Preservation with $\sigma_\sim$]\label{def:rtp:sigma}
-  % $\;$\\
-  % \vspace{-1em}
-  % \begin{nscenter}
-  % Compiler $\cc{\src{L}}{\trg{L}}$ robustly preserves $\class$, 
-  \bul{$\rtpsig{\cc{\src{L}}{\trg{L}}}{\trg{\class}}{\sim}$}
-  %, iff 
-  $\isdef$
-  \rul{$\forall \trg{\pi}\in\trg{\class}, \src{p}\in\src{L},$} if \iul{$\rsat{\src{\progvar}}{\sigma_\sim\left(\trg{\pi}\right)}$}, then \oul{$\rsat{\cc{\src{L}}{\trg{L}}\left(\src{p}\right)}{\trg{\pi}}$}.
-  % \end{nscenter}
-\end{definition}
-Note that a class of properties $\class$ can represent just one property $\pi$ by lifting~\cite{clarkson2008hyper} that property to sets of properties, i.e., use the powerset of $\pi$ instead of $\pi$ itself.
-Because of this, this paper may write $\rtptau{\cc{\src{L}}{\trg{L}}}{\trg{\pi}}{\sim}$, even though $\trg{\pi}$ is a property and not a class.
-A similar construction can be used to the projection functions (see \Cref{def:universal:img}) by applying them to the lifted version of $\trg{\pi}$ instead of just $\trg{\pi}$.
-
-In case the trace model is the same for both source and target programs (and thus $\sim$ is equality), we obtain~\cite{abate2019jour}:
-\begin{definition}[Robust Preservation]\label{def:rtp}
-  % $\;$\\
-  % \vspace{-1em}
-  % \begin{nscenter}
-  % Compiler $\cc{\src{L}}{\trg{L}}$ robustly preserves $\class$, 
-  $\;$
-
-  {$\rtp{\cc{\src{L}}{\trg{L}}}{\class}$}
-  %, iff 
-  $\isdef$
-  {$\forall \pi\in\class, \src{p}\in\src{L},$} if {$\rsat{\src{\progvar}}{\pi}$}, then {$\rsat{\cc{\src{L}}{\trg{L}}\left(\src{p}\right)}{\pi}$}.
-  % \end{nscenter}
-\end{definition}
-
-Examples of compilers fulfilling \Cref{def:rtp} exist in the literature~\cite{korashy2022secureptrs,korashy2021capableptrs,abate2021extacc,abate2019jour,patrignani2021rsc}.
-For example, SecurePtrs~\cite{korashy2022secureptrs} gives a compiler that robustly preserves all safety properties for a C-like language to an assembly-like language. 
-% The mechanization has well over 20k lines in total, which demonstrates that proving a compiler secure with respect to \Cref{def:rtp} is a significant effort. \MP{move this line to later ?}
-% i don't see how adding a def of cct is beneficial
-As another example, even though it is not strictly satisfying \Cref{def:rtp}, the FaCT~\cite{cauligi2019fact} compiler preserves the \gls*{cct} property for a C-like language with constant-time primitives, e.g., \texttt{ctselect} for branching.
-%, whose precise definition can be found in \Cref{sec:cct}.
-Throughout this work, it is assumed that FaCT satisfies \Cref{def:rtp}.
-% While we were not able to find compilers in literature that provably fulfill \Cref{def:rtp}, we anticipate that 
- % would satisfy this hyperproperty.
-% FaCT proves constant-timeness by means of overapproximation, i.e., instead of directly proving the hyperproperty, they fall back to a property that overapproximates the corresponding hyperproperty.
-
-%\smallskip%MP: the underline compacts the sections too much
-
-\section{Secure Composition}\label{sec:sequential}
-
-Notably, real-world compilation chains also perform a series of (sequential) passes whose main purpose is not necessarily to translate from one language to another, but to, e.g., optimise the code or enforce a certain property.
-Both examples can be seen in practice, e.g.,~\cite{nagarakatte2009soft,nagarakatte2010cets,akritidis2009baggy,wegman1991ccp,manjikian1997fusion} and many more.
-% \MPin{survey to cite?
-this is a bit informal MK: hard to find a survey on different compiler passes...
-I can't even find one just for compiler enforced memsafety (that looks good) or for compiler translations in general}.
-% 
-% Consider the compilation passes of LLVM~\cite{lattner2004llvm}, which are handled in the \texttt{PassManager} class.
-% By manipulating the internal state of a \texttt{PassManager} object, the schedule of compiler optimisation passes can be hardcoded.
-% That way, an order of optimisations can be found by hand that yields the best results.
-% \MP{what's the point of the previous 3 sentences?(commented)}
-Consider the following two LLVM optimisation passes: \gls*{cf}, which rewrites constant expressions to the constant they evaluate to, and \gls*{dce}, which removes dead code by rewriting conditional branches.
-The order in which \gls*{cf} and \gls*{dce} are performed influences the final result of the compilation (see \Cref{fig:cfdceex}).
-\begin{figure}[!ht]
-  \begin{center}
-    \begin{tikzpicture}[scale=0.62, every node/.style={transform shape}]
-      \node[minimum width=3cm,minimum height=2.5cm,draw=black,very thick,align=left] (progunopt) {\begin{lstlisting}[language=ml,basicstyle=\small\ttfamily]
-let a = true in
-if a then
-  print "a"
-else
-  print "b"
-\end{lstlisting}};
-      \node[minimum width=3cm,minimum height=2.5cm,draw=black,very thick,align=left,right=1.0 of progunopt] (progoptcf) {\begin{lstlisting}[language=ml,basicstyle=\large\ttfamily]
-if true then
-  print "a"
-else
-  print "b"
-\end{lstlisting}};
-      \node[minimum width=3cm,minimum height=2.5cm,draw=black,very thick,align=left,right=1.0 of progoptcf] (progoptcfdce) {\begin{lstlisting}[language=ml,basicstyle=\Large\ttfamily]
-print "a"
-\end{lstlisting}};% the following is a hack to get my syntax highlighting to work again \end{lstlisting}
-      % edges
-      \draw[->,very thick] (progunopt) edge[loop left,left] node {\gls*{dce}} (progunopt);
-      \draw[->,very thick] (progunopt) edge[sloped,above] node {\gls*{cf}} (progoptcf);
-      \draw[->,very thick] (progoptcf) edge[sloped,above] node {\gls*{dce}} (progoptcfdce);
-    \end{tikzpicture}
-  \end{center}
-  \caption{Example program where the level of optimisations differ for one pass of applying \gls*{cf} and \gls*{dce} in any order. %
-  Every edge is a compilation pass and the label on the edge states what the pass does, i.e., \gls*{cf} or \gls*{dce}. %
-  The source code in the nodes is a glorified compiler intermediate representation and the code gets more optimised towards the right hand side of the figure.}\label{fig:cfdceex}
-\end{figure}
-This {\em phase ordering problem} is well--known in literature and a practical solution is to simply perform a fixpoint iteration of the optimisation pipeline~\cite{click1995combining}.
-
-\smallskip
-
-To analyse the security of compilation passes and their interaction within a compilation pipeline, we rely on a few key notions: the definition of a trace relation, and the definition of when is a trace relation well-formed with respect to a class.
-%
-Consider traces $\src{\trace}$ and $\trg{\trace}$ as well as two trace relations $\sim_1$ and $\sim_2$. 
-% 
-The traces are related $\src{\trace}\sim_1\bullet\sim_2\trg{\trace}$ if there is another trace $\irl{\trace}$ such that $\src{\trace}\sim_1\irl{\trace}$ and $\irl{\trace}\sim_2\trg{\trace}$.
-% 
-\bul{A relation $\sim$ is well-formed w.r.t. a class of target-level properties $\trg{\class}$} iff \rul{the universal image preserves set membership}.
-\begin{definition}[Well-formedness of $\sim$ for a Class $\trg{\class}$]\label{def:wfc:sig:tracerel}
-  % $$
-
-  \begin{nscenter}
-  \noindent
-  \text{\bul{$\wfcsig{\sim}{\trg{\class}}$}} \isdef \text{\rul{$\forall \trg{\pi}\in\trg{\class}, \sigma_\sim(\trg{\pi})\in\sigma_\sim(\trg{\class})$}}
-  \end{nscenter}
-  % $$
-\end{definition}
-
-We can now state our main result: secure compilers in the robust compilation framework~\cite{abate2021extacc} compose {\em sequentially}. 
-%
-
-Let \bul{$\cc{\src{L}}{\trg{L}}$ robustly preserve the class $\sigma_{\sim_2}\left(\trg{\class[_1]}\right)$ under $\sim_1$} and let \rul{$\cc{\trg{L}}{\obj{L}}$ robustly preserve the class $\trg{\class[_2]}$ under $\sim_2$}.
-Then, \iul{when the cross-language relation $\sim_2$ is well-formed w.r.t.
-class $\trg{\class[_1]}$}, it follows that \oul{the composed compiler $\cc{\src{L}}{\trg{L}}\circ\cc{\trg{L}}{\obj{L}}$ robustly preserves the intersection of classes $\trg{\class[_1]}\cap\trg{\class[_2]}$ under $\sim_1\bullet\sim_2$}.
-% 
-%For space reasons, \Thmref{thm:urtp} and \Thmref{thm:lrtp} are not presented here, but can be found in the technical appendix ($\CoqSymbol$).
-% Note that it is not necessary to require $\sim_2$ to be well-formed w.r.t. $\src{\class[_1]}$, since the 
-% \MPin{why}
-\begin{theorem}[Composition of Secure Compilers w.r.t. $\sigma$]\label{thm:rtpsim:sig}
-  $\;$
-
-  If \bul{$\rtpsig{\cc{\src{L}}{\trg{L}}}{\tilde{\sigma}_{\sim_2}\left(\trg{\class[_{1}]}\right)}{\sim_1}$}, \rul{$\rtpsig{\cc{\trg{L}}{\obj{L}}}{\trg{\class[_2]}}{\sim_2}$}, and \iul{$\wfcsig{\sim_2}{\trg{\class[_1]}}$}, \\ then \oul{$\rtpsig{\cc{\src{L}}{\trg{L}}\circ\cc{\trg{L}}{\obj{L}}}{\trg{\class[_{1}]}\cap\trg{\class[_{2}]}}{\sim_1\bullet\sim_2}$}.
-\Coqed
-\end{theorem}
-
-Since the composition of secure compilers is again a secure compiler, the theorems generalise to a whole chain of $n$ secure compilers. 
-% 
-\Cref{thm:rtpsim:sig} can also be stated for the existential image $\tau_\sim\left(\src{\pi}\right)$, but in the interest of space that result has been moved to the appendix (\Cref{subsec:extimg}).
-% 
-Crucially, \Cref{thm:rtpsim:sig} also holds for \emph{classes of hyperproperties}, and thus, compilers that robustly preserve hyperproperties can be composed with each other as well as with compilers that robustly preserve properties.
-% 
-%Unfortunately, proofs of compilers that robustly preserve hyperproperties (including hypersafety) are scarce in the literature due to their complexity.
-%\MP{ so what ? }
-%  we don't discuss this further?
-%
-
-% It is furthermore possible to conflate the notion of classes and hyperproperties by intersecting all elements of a class of hyperproperties, to obtain one combined hyperproperty.
-% This idea has been discussed in other work~\cite{clarkson2008hyper}.
-%  MP: i think i stated the point of these 2 lines above
-
-If we take SecurePtrs and FaCT from \Cref{subsec:bg:rtp} and compose them according to \Cref{thm:rtpsim:sig}, we obtain a compiler that robustly preserves the intersection of safety properties and the \gls*{cct} hyperproperty.
-That is, for a source component that robustly satisfies any set of safety properties {\em and} \gls*{cct}, the compiled target component also robustly satisfies the same set of safety properties and \gls*{cct}.
-% The intersection simply uses the lifted version of the respective safety properties.
-%\MP{meaning?}
-
-Compiler engineers typically try to find an order of optimisations that yields well-optimised programs for either code size~\cite{cooper1999geneticphases} or performance~\cite{kulkarni2006exhaustivephase}.
-\Cref{corr:swappable} justifies that any such order of compilation passes is valid w.r.t. security, as long as the trace-relations have no effect on the respective classes.
-
-So, given two compilation passes \bul{$\cc[_{1}]{\trg{L}}{\trg{L}}$, $\cc[_{2}]{\trg{L}}{\trg{L}}$, both robustly preserving class $\trg{\class[_{1}]}$ or $\trg{\class[_{2}]}$, respectively}, \rul{their corresponding well-formed trace-relations}, \iul{and indifference of the classes with respect to these trace relations}, \oul{for any order of their composition, the composed compiler robustly preserves the intersection of $\trg{\class[_{1}]}$ and $\trg{\class[_{2}]}$}.
-% 
-% \MPin{
-% 	later on, say that optimisation passes remove elements from traces, so they should be fine w.r.t. the result with $\tau$.
-% 	also, yes, if $\tau$ adds we have a problem, if it removes we don't
-% \MKin{ Optimisations that remove elements from traces are not necessarily sound,
-%  e.g., remove all deallocations to "optimise" memory management by offloading
-%  the work to the OS.
-% }
-% }
-
-\begin{corollary}[Swapping Secure Compiler Passes]\label{corr:swappable}
-  $\;$
-
-  If \bul{$\rtpsig{\cc[_{1}]{\trg{L}}{\trg{L}}}{\trg{\class[_{1}]}}{\sim_1}$ and $\rtpsig{\cc[_{2}]{\trg{L}}{\trg{L}}}{\trg{\class[_{2}]}}{\sim_2}$}, %
-  \rul{$\wfcsig{\sim_1}{\trg{\class[_2]}}$ and $\wfcsig{\sim_2}{\trg{\class[_1]}}$}, %
-  and \iul{$\tilde{\sigma}_{\sim_2}\left(\trg{\class[_1]}\right)=\trg{\class[_1]}$ as well as $\tilde{\sigma}_{\sim_1}\left(\trg{\class[_2]}\right)=\trg{\class[_2]}$},
-  then \oul{$\rtpsig{\cc[_{1}]{\trg{L}}{\trg{L}}\circ\cc[_{2}]{\trg{L}}{\trg{L}}}{\trg{\class[_{1}]}\cap\trg{\class[_{2}]}}{\sim_1\circ\sim_2}$ and $\rtpsig{\cc[_{2}]{\trg{L}}{\trg{L}}\circ\cc[_{1}]{\trg{L}}{\trg{L}}}{\trg{\class[_{2}]}\cap\trg{\class[_{1}]}}{\sim_2\circ\sim_1}$}.
-\Coqed
-\end{corollary}
-
-Coming back to the example composing SecurePtrs with FaCT, it is likely the case that \Cref{corr:swappable} is not applicable.
-While first running SecurePtrs and then FaCT should be fine, the other direction has potential security hazards, since the SecurePtrs compiler does not account for cryptographic-constant time primitives, such as \texttt{ctselect}.
-
-\subsection{Secure Compiler Composition with Same Trace Models}
-When the cross-language trace relation is an equality, \Cref{thm:rtpsim:sig} collapses:
-Given \bul{$\cc{\src{L}}{\trg{L}}$ robustly preserves $\class[_{1}]$} and \rul{$\cc{\trg{L}}{\obj{L}}$ robustly preserves $\class[_{2}]$}, it follows that \oul{their sequential composition $\cc{\src{L}}{\trg{L}}\circ\cc{\trg{L}}{\obj{L}}$ robustly preserves the intersection of classes $\class[_{1}]$ and $\class[_{2}]$}.
-
-\begin{corollary}[Composition of Secure Compilers]\label{corr:rtp}
-  $\;$
-
-  If \bul{$\rtp{\cc{\src{L}}{\trg{L}}}{\class[_{1}]}$} and \rul{$\rtp{\cc{\trg{L}}{\obj{L}}}{\class[_{2}]}$}, then \oul{$\rtp{\cc{\src{L}}{\trg{L}}\circ\cc{\trg{L}}{\obj{L}}}{\class[_{1}]\cap\class[_{2}]}$}.
-\Coqed
-\end{corollary}
-
-\Cref{corr:rtp} provides an easy way to compose secure compilers without well-formedness of trace relations. 
-% 
-However, while \Cref{thm:rtpsim:sig} explicitly requires that $\sim_1$ is well-formed w.r.t. $\trg{\class[_2]}$, if $\sim_2$ is not well-formed w.r.t. $\trg{\class[_1]}$, care must be taken. 
-%Worst-case is that the resulting class $\sigma_{\sim_1\bullet\sim_2}\left(\trg{\class[_1]}\cap\trg{\class[_2]}\right)$ becomes empty, even though $\trg{\class[_1]}\cap\trg{\class[_2]}$ is not necessarily empty.
-This is further discussed in \Cref{subsec:compatsecpasses}.
-
-%If possible, optimisations are merged to enhance the analysis' result, since even when iterating to a fixpoint the final program may not be as optimised as it could be~\cite{click1995combining}.
-We can also obtain a specialised version of \Cref{corr:swappable}:
-%So, given two compilation passes \bul{$\cc[_{1}]{\trg{L}}{\trg{L}}$, $\cc[_{2}]{\trg{L}}{\trg{L}}$, both robustly preserving class $\class[_{1}]$ or $\class[_{2}]$, respectively}, \oul{for any order of their composition, the composed compiler robustly preserves the intersection of $\class[_{1}]$ and $\class[_{2}]$}.
-% 
-% \MPin{
-% 	later on, say that optimisation passes remove elements from traces, so they should be fine w.r.t. the result with $\tau$.
-% 	also, yes, if $\tau$ adds we have a problem, if it removes we don't
-% \MKin{ Optimisations that remove elements from traces are not necessarily sound,
-%  e.g., remove all deallocations to "optimise" memory management by offloading
-%  the work to the OS.
-% }
-% }
-
-\begin{corollary}[Swapping Secure Compiler Passes]\label{corr:swappable:one}
-  $\;$
-
-  If {$\rtp{\cc[_{1}]{\trg{L}}{\trg{L}}}{\class[_{1}]}$ and $\rtp{\cc[_{2}]{\trg{L}}{\trg{L}}}{\class[_{2}]}$}, then {$\rtp{\cc[_{1}]{\trg{L}}{\trg{L}}\circ\cc[_{2}]{\trg{L}}{\trg{L}}}{\class[_{1}]\cap\class[_{2}]}$ and $\rtp{\cc[_{2}]{\trg{L}}{\trg{L}}\circ\cc[_{1}]{\trg{L}}{\trg{L}}}{\class[_{2}]\cap\class[_{1}]}$}.
-\Coqed
-\end{corollary}
-
-\section{Security Properties Formalisation \& Composition\pages{1.5}}\label{sec:compprop}
-
-This section introduces trace properties of interest for this paper: \gls*{tms}, \gls*{sms}, \gls*{ms}, \gls*{scct}, and \gls*{ss}.
-These properties are of practical importance (as mentioned in \Cref{sec:introduction}) and also of interest in the case study presented later (\Cref{sec:casestud:defs,sec:casestud:rtp}). 
-This section presents all of them, despite the fact that they are inspired by existing work, in order to showcase all that is required for a formal proof of security for a realistic compilation toolchain.
-%\MP{ say something about this in the rebuttal: only the defs are borrowed, the universal images and all that work is new}
-%\MK{ universal images are not new. }
-% , none of the properties is novel, for self-containment of the paper, their definition is necessary for the case study.
-For each of the presented properties, the technical report defines corresponding monitors, which refine a given property and have a key role in the proofs of this paper. 
-% managed to prove 
-%  if tms-monitor --[trace]-->* ∅  
-%  then trace ∈ tms
-% but not
-%  if trace ∈ tms
-%  then tms-monitor --[trace]-->* ∅
-
-\subsection{A Trace Model for Memory Safety}\label{subsec:basic:memsafety:tracemodel}
-
-For simple memory safety composed of temporal and spatial memory safety, the trace model defines events ($\event[_{\mssafe}]$) as either the empty event ($\emptyevent$), a crash ($\lightning$), or a base-event ($\preevent[_{\mssafe}]$).
-
-\vspace{-1.0em}
-\[
-  \begin{array}{rcll}
-    \text{(Base-Events)}&\preevent[_{\mssafe}] &:=& \ev{Alloc \loc\ n} \mid \ev{Dealloc \loc} \mid \ev{Use \loc\ n} \\
-    \text{(Events)}&\event[_{\mssafe}] &:=& \preevent[_{\mssafe}] \mid \emptyevent \mid \lightning \\ 
-  \end{array}
-\]
-
-Base-events describe the actual kind of event that happened.
-For the basic memory-safety properties, these are three variants:
-First, the allocation event ($\ev{Alloc\ \loc\ n}$) that fires whenever a program claims $n$ cells of memory and stores them at address $\loc$, where addresses are assumed to be unique.
-Second, deallocation ($\ev{Dealloc\ \loc}$) announces that the object at location $\loc$ is freed.
-Third, an event to describe reads from and writes to the $n$-th memory cell from address $\loc$ ($\ev{Use\ \loc\ n}$).
-
-\subsubsection{Temporal Memory Safety}
-
-\gls*{tms}~\cite{nagarakatte2010cets} is a safety property that describes that an unallocated object must not be (re-)used.
-
-\begin{definition}[\glsfirst*{tms}]\label{def:trace:tmsdef}
-  % \footnotesize
-  $$
-  \tmssafe:=\left\{\trace[_{\mssafe}] \left| \begin{array}{rcl}
-    \ev{Alloc\ \loc\ n}&\le_{\trace[_{\mssafe}]}&\ev{Dealloc\ \loc} \\
-    \ev{Use\ \loc\ n}&\le_{\trace[_{\mssafe}]}&\ev{Dealloc\ \loc} \\
-    \text{at most one }\ev{Dealloc\ \loc}&\text{in}&\trace[_{\mssafe}] \\
-    \text{at most one }\ev{Alloc\ \loc\ n}&\text{in}&\trace[_{\mssafe}] \\
-  \end{array}\right.\right\}
-  $$
-\end{definition}
-Hereby, the notation $\event[_{1}]\le_{\trace}\event[_{2}]$ means that if $\event[_{1}]$ is in $\trace$ and if $\event[_{2}]$ is in $\trace$, then $\event[_{1}]$ appears before $\event[_{2}]$.
-
-\subsubsection{Spatial Memory Safety}
-
-\gls*{sms}~\cite{nagarakatte2009soft} is a safety property that prohibits out-of-bounds accesses.
-
-\begin{definition}[\glsfirst*{sms}]\label{def:trace:smsdef}
-% \small
-  % \begin{nscenter}
-
-  \noindent
-  \[
-  \smssafe:=\left\{\trace[_{\mssafe}] \left|\begin{array}{rcl}
-      \text{If }\ev{Alloc\ \loc\ n}\le_{\trace_{\mssafe}}\ev{Use\ \loc\ m}, \text{ then }m<n
-  \end{array}\right.\right\}
-  \]
-  % \end{nscenter}
-\end{definition}
-
-\subsubsection{Memory Safety}
-
-In spirit of earlier work~\cite{nagarakatte2009soft,nagarakatte2010cets,jim2002cyclone,necula2005ccured,michael2023mswasm}, full \gls*{ms} is the intersection of \Cref{def:trace:tmsdef,def:trace:smsdef}.
-
-\begin{definition}[\glsfirst*{ms}]\label{def:trace:msdef}
-  $
-  \mssafe:=\tmssafe \cap \smssafe
-  $
-\end{definition}
-
-Note that \Cref{def:trace:msdef} ignores data isolation, so there may still be memory-safety issues introduced by side-channels.
-
-\subsection{A Trace Model for Memory Safety with Constant Time}\label{subsec:scct:tracemodel}
-
-To express Constant Time, we extend the memory safety trace model with a {\em security tag} ($\securitytag{}$) that indicates whether events contains sensitive information ($\lock$) or not ($\unlock$).
-
-\vspace{-.5em}
-\[
-  \begin{array}{rrcl}
-    (\text{Base-Events}) & \preevent[_{\ctsafe}] &:=& \preevent[_{\mssafe}] \mid \ev{Branch\ }n \mid \ev{Binop\ }n\\
-    (\text{Security Tags}) & \securitytag{} &:=& \lock \mid \unlock\\ 
-    (\text{Events}) & \event[_{\ctsafe}] &:=& \preevent[_{\ctsafe}];\securitytag{} \mid \emptyevent \mid \lightning \\ 
-  \end{array}
-\]
-
-For cryptographic code, there is a general guideline that secrets must not be visible on a trace~\cite{ctguidelines}, i.e., secrets should not be marked as $\unlock$.
-In turn, an instruction whose timing is data-dependent must not have a secret as an operand.
-Typical operations with data-dependent timing are branches and certain binary operations, such as division.%
-\footnote{
-	This is architecture-dependent, but division is an operation that serves as a classic example for a data-dependent timing instruction~\cite[p.~755]{arm-refman}.
-}
-Both operations are represented in the trace model by extending the set of base-events with branches ($\ev{Branch\ n}$) and binary operations ($\ev{Binop\ n}$).
-
-\subsubsection{Strict Cryptographic Constant Time}
-
-\gls*{cct} is a hypersafety property~\cite{barthe2018sec} and, thus, difficult to check with monitors.
-This is because, intuitively, hypersafety properties can relate multiple execution traces with each other, but monitors work on a single execution.
-It is a common trick to sidestep this issue by means of overapproximation: this section defines the property \gls*{scct}, a stricter variant of \gls*{cct} (inspired by earlier work~\cite{almeida2017jasmin}) that enforces the policy that no secret appears on a trace.
-Programs that satisfy \gls*{scct} also satisfy \gls*{cct}, but programs that satisfy \gls*{cct} may not satisfy \gls*{scct}.
-
-\begin{definition}[\glsfirst*{scct}]\label{def:trace:scctdef}
-% \begin{nscenter}
-
-  \noindent\[
-  \scctsafe:=\left\{\trace[_{\ctsafe}] \left|\begin{array}{l}
-      \trace[_{\ctsafe}]=\hole{\cdot} \text{ or } \\\exists\trace[_{\ctsafe}'],\trace[_{\ctsafe}]=\preevent[_{\ctsafe}];\unlock\cdot\trace[_{\ctsafe}'] \wedge \trace[_{\ctsafe}']\in\scctsafe
-    \end{array}\right.\right\}
-  \]
-% \end{nscenter}
-\end{definition}
-
-\gls*{scct} may appear overly strict, since it seems that secrets must not occur on a trace (since $\securitytag{}$ is forced to be $\unlock$). 
-However, this is considered standard practice in terms of coding guidelines~\cite{ctguidelines}.
-Moreover, programs that have been compiled with FaCT~\cite{cauligi2019fact} and run with a ``data independent timing mode''~\cite{arm-refman,intel-refman} enabled do not leak secrets (see \Cref{ex:lscct}).
-
-\subsubsection{\gls*{ms}, Strict Cryptographic Constant Time}\label{sec:msscct-rel}
-
-The combination of \gls*{ms} and \gls*{scct} is the intersection of these properties, \gls*{msscct}.
-However, \gls*{ms} uses a different trace model than \gls*{scct}, so intersecting them would trivially yield the empty set. 
-To remedy this issue, we introduce $\sim_{\ctsafe}: \preevent[_{\ctsafe}] \times \preevent[_{\mssafe}] $, a cross-language trace relation (whose key cases are presented below), that we use to intuitively unify the trace model in which the two properties are expressed:
-\[
-  \typerulenolabel{mscctrel:drop:tag}{}{\preevent[_{\mssafe}]\sim_{\ctsafe}\preevent[_{\mssafe}];\unlock{}}
-  \typerulenolabel{mscctrel:drop:crash}{}{\lightning\sim_{\ctsafe}\lightning}
-\]
-\[
-  \typerulenolabel{mscctrel:drop:branch}{}{\emptyevent\sim_{\ctsafe}\ev{Branch n};\securitytag{}}
-  \typerulenolabel{mscctrel:drop:binop}{}{\emptyevent\sim_{\ctsafe}\ev{Binop n};\securitytag{}}
-\]
-
-Essentially, $\sim_{\ctsafe}$ ignores both the new $\ev{Branch\ n}$ and $\ev{Binop\ n}$ base-events as it relates security-insensitive actions ($\unlock$) to their equivalent counterparts.
-Thus, \gls*{ms} traces trivially satisfy \gls*{scct}.
-% 
-The above relation is extended point-wise to traces, skipping the empty event $\emptyevent$ on either side, and it is now possible to define \gls*{msscct} using the universal image:
-
-\begin{definition}[\gls*{ms} and \gls*{scct}]\label{def:trace:msscctdef}
-  $
-  \msscctsafe:=\mssafe\cap\sigma_{\sim_{\ctsafe}}\left(\scctsafe\right)
-  $
-\end{definition}
-
-\subsection{Extending the Trace Model with Speculation}\label{subsec:msctss:tracemodel}
-
-So far, the considered trace models do not let us express speculative execution attacks such as Spectre~\cite{kocher2019spectre}. 
-For this, we extend the earlier trace model (see \Cref{subsec:scct:tracemodel}) so that the security tags ($\securitytag{}$) carry additional information about the kind of private data leakage, i.e., the type of speculative leak.
-Moreover, we add base-events signalling the beginning of a speculative execution ($\ev{Spec}$), a barrier ($\ev{Barrier}$) that signals that any speculative execution may not go past it, as well as a rollback event ($\ev{Rlb}$), which signals that execution resumes to where speculation started.
-
-\vspace{-1em}
-{
-\[
-  \begin{array}{rrcl}
-    (\text{Base-Events}) & \preevent[_{\mathghost}] &:=& \preevent[_{\ctsafe}] \\
-    (\text{Spectre Variants}) & vX &:=& \operatorname{NONE} \mid \operatorname{PHT} \\
-    (\text{Security Tags}) & \securitytag{} &:=& \lock_{vX} \mid \unlock\\ 
-    (\text{Events}) & \event[_{\mathghost}] &:=& \preevent[_{\mathghost}];\securitytag{} \mid \emptyevent \mid \lightning \mid \ev{Spec} \mid \ev{Rlb} \mid \ev{Barrier}\\ 
-  \end{array}
-\]
-}
-
-Even though the considered Spectre variants are just SPECTRE-PHT~\cite{kocher2019spectre}, NONE just describes secret data as in \gls*{scct} (see \Cref{subsec:scct:tracemodel}), the trace model is general enough to allow for potential future extension with different variants~\cite{kocher2019spectre,maisuradze2018ret2spec,horn2019zero}.
-%For sake of readability, this paper just uses the notation $\lock$ in place of $\lock_{\text{PHT}}$.
-
-\subsubsection{Speculative Safety}
-
-\gls*{ss}~\cite{patrignani2021exorcising}, similar to \gls*{scct}, is a sound overapproximation of a variant of noninterference.
-
-\begin{definition}[\glsfirst*{ss}]\label{def:trace:ss}
-  \noindent
-
-  \begin{nscenter}
-  $
-    \sssafe:=\left\{\trace[_{\mathghost}] \left|\begin{array}{l}
-      \trace[_{\mathghost}]=\hole{\cdot} \text{ or } \exists\trace[_{\mathghost}'].\\
-      \left(\trace[_{\mathghost}]=\preevent[_{\mathghost}];\unlock\cdot\trace[_{\mathghost}'] \text{or }\trace[_{\mathghost}]=\preevent[_{\mathghost}];\lock_{\text{NONE}}\cdot\trace[_{\mathghost}']\right)\\
-      \text{and }\ \trace[_{\mathghost}']\in\sssafe
-                                 \end{array}\right.\right\}
-  $ 
-  \end{nscenter}
-\end{definition}
-The technical setup so far leads to the above definition, where only locks annotated with $\text{SPECTRE-PHT}$ are disallowed to occur on the trace.
-That way, programs attaining \gls*{ss} do not necessarily attain \gls*{scct}.
-
-\subsubsection{Speculation Memory Safety}\label{sec:spec-ms-rel}
-
-As before, we need to relate the different trace models with each other, so that the memory safety property without speculation can be lifted to speculation. 
-To this end, let $\sim_{\mathghost}: \preevent[_{\mathghost}] \times \preevent[_{\ctsafe}]$ be a cross-language trace relation whose key cases are below.
-The intuition is that \gls*{ss} is trivially satisfied in \gls*{scct}, since speculation is inexpressible there, which amounts to dropping events $\ev{Spec}$, $\ev{Rlb}$, or $\ev{Barrier}$, as well as all base events tagged with $\lock_{\text{PHT}}$.
-
-\[
-  \typerulenolabel{specrel:drop:tag}{}{\preevent[_{\ctsafe}];\lock\sim_{\mathghost}\preevent[_{\ctsafe}];\lock_{\text{NONE}}}
-  \typerulenolabel{specrel:drop:spectag}{}{\emptyevent\sim_{\mathghost}\preevent[_{\mathghost}];\lock_{\text{PHT}}}
-\]
-\[
-  \typerulenolabel{specrel:drop:pub}{}{\preevent[_{\ctsafe}];\unlock\sim_{\mathghost}\preevent[_{\ctsafe}];\unlock}
-  \typerulenolabel{specrel:drop:crash}{}{\lightning\sim_{\mathghost}\lightning}
-\]
-\[
-  \typerulenolabel{specrel:drop:spec}{}{\emptyevent\sim_{\mathghost}\ev{Spec}}
-  \typerulenolabel{specrel:drop:rlb}{}{\emptyevent\sim_{\mathghost}\ev{Rlb}}
-  \typerulenolabel{specrel:drop:barrier}{}{\emptyevent\sim_{\mathghost}\ev{Barrier}}
-\]
-
-We conclude by defining the ultimate property of interest for secure compilers: \gls*{specms}.
-\begin{definition}[\gls*{specms}]\label{def:trace:specmsdef}
-  $
-  \specmssafe := \msscctsafe\cap\sigma_{\sim_{\mathghost}}\left(\sssafe\right)
-  $
-\end{definition}
-
-\section{Case Study: Language Formalisations\pages{2}}\label{sec:casestud:defs}
-
-This section defines programming languages $\Ltms$, $\Ltrg$, $\Lms$, $\Lscct$, and $\Lspec$, all of which share common elements (presented in \Cref{subsec:cs:defs}).
-$\Ltms$ is the only statically-typed language, and it exhibits the property that all well-typed programs are \gls*{tms} (\Cref{subsec:ltms}).
-However, not all $\Ltms$ programs are \gls*{sms}.
-That is, there are well-typed $\Ltms$ programs that perform an out-of-bounds access.
-Language $\Ltrg$ is untyped and does not provide any guarantees with regards to \gls*{ms} (\Cref{subsec:lsms}).
-$\Lms$ is exactly the same language as $\Ltrg$, but this paper still distinguishes the two for sake of readability (\Cref{subsec:lms}).
-All three languages --- so $\Ltms$, $\Ltrg$, and $\Lms$ --- assume \gls*{scct} to hold, since this is -- in an ideal world -- what the programmer would expect, too: it is the job of the compiler to preserve and (potentially) enforce \gls*{scct} security~\cite{cauligi2019fact,nagarakatte2010cets,nagarakatte2009soft,akritidis2009baggy}.
-% This paper considers \gls*{cct} as an architecture-dependent property\MPin{can we justify this}, so
-
-Such consideration is also backed up by architecture providing a data (operand) independent timing mode, such as processors by Arm~\cite[p.~543]{arm-refman} and Intel~\cite[p.~80]{intel-refman}.
-%In spirit of this, all languages besides $\Lscct$ are trivially \gls*{scct}. 
-This kind of processor feature is modelled in language $\Lscct$ (\Cref{subsec:lscct}), where programs have access to a ``\gls*{cct}-mode'' and can change the leakage of emitted events according to the value of this mode (either $\obj{ON}$ or $\obj{OFF}$).
-
-Finally, modern processors also employ speculative execution to achieve speedups---and unfortunately generate Spectre attacks~\cite{kocher2019spectre}---and this is the extension of $\Lspec$ (\Cref{subsec:lspec}).
- % extends $\Lscct$ to model speculative execution.
-%
-Thus, all previous languages trivially satisfy \gls*{ss}, since they do not support speculative execution at all.
-% All languages share large parts of syntactic elements, which are presented in the next subsection (\Cref{subsec:cs:defs}).
-% Subsequent subsections extend on this accordingly.
-
-\subsection{Shared Language Definitions}\label{subsec:cs:defs}
-
-All presented programming languages share a common fragment which is partially presented here and in full detail in the technical report. 
-%\MP{put omitted stuff in appendix + ref there (and add textual explanation)}
-
-% \vspace{-.5em}
-{
-  \renewcommand{\src}[1]{\mathsf{#1}}
- \small
-\[
-  \begin{array}{rrcl}
-    \text{(Base-Events)} & \src{\preevent} &:=& \src{\ev{Alloc \loc\ n}} \mid \src{\ev{Dealloc \loc}} \mid \src{\ev{Get \loc\ n}} \mid \src{\ev{Set \loc\ n}} \\
-    \text{(Control Tags)} & \src{\sandboxtag{}} &:=& \src{\ctx} \mid \src{\comp} \\
-    \text{(Events)} & \src{\event} &:=& \src{\preevent;\sandboxtag{}} \mid \src{\emptyevent} \mid \src{\lightning} \\ 
-    \text{(Values)} & \src{v} &:=& \src{n} \\
-    \text{(Expressions)} & \src{\expr} &:=& \src{x} \mid \src{n} \mid \src{\lbinop{\expr[_1]}{\expr[_2]}} \mid \src{\lifz{\expr[_1]}{\expr[_2]}{\expr[_3]}} \\ 
-                         &&&\mid \src{\llet{x}{\expr[_1]}{\expr[_2]}} \mid \src{\llet{x}{new\ \expr[_1]}{\expr[_2]}} \\
-                         &&&\mid \src{\ldelete{x}} \mid \src{\lget{x}{\expr}} \mid \src{\lset{x}{\expr[_1]}{\expr[]}} \mid \src{stuck}\\
-                         &&&\mid \src{\lcall{f}{\expr}} \mid \src{\lreturn{\expr}} \\
-    \text{(Functions)} & \src{F} &:=& (\src{x};\src{\expr}) \\
-    \text{(Libraries)} &&&\hspace{-3.8em} \src{\library_{\ctx}},\src{\library_{\comp}} : \src{Vars} \to \src{F} 
-    % \hspace{2em}
-    %  : \src{Vars} \to \src{F} 
-     \\
-    \text{(Heaps)} & \src{H} &:=& \src{\hole{\cdot}} \mid \src{n},\src{H}
-    % \hspace{0.5cm}    
-    \\
-    \text{(Pointer Maps)} & \src{\Delta} && \text{ omitted for simplicity}
-    \\
-    \text{(Memory States)} & \src{\memstate} &:=& \left(\src{H^{\ctx}};\src{H^{\comp}};\src{\Delta}\right)\\
-    \text{(Control States)} & \src{\cfstate} & & \text{ omitted for simplicity} 
-    \\
-    \text{(States)} &\src{\Omega} &:=& \left(\src{\cfstate};\src{\sandboxtag{}};\src{\memstate}\right)\\
-  \end{array}
-\]
-}
-{
-  \renewcommand{\src}[1]{\mathsf{#1}}
-The trace models of all languages are similar to those presented earlier (\Cref{sec:compprop}).
-One technical detail is the addition of a control tag, indicating who is to blame for an emitted action: context ($\ctx$) or component ($\comp$).
-This is a standard technique in secure compilation in order to rule out irrelevant context-level events. 
-For example, a context immediately deallocating an allocated memory region twice trivially violates memory safety, but the main interest in secure compilation is to preserve component-level security, and thus component events.
-Even though this tagging could be used for blame preservation~\cite{patrignani2023blame}, this is beyond the focus of this paper.
-Another key difference is that memory accesses are now explicitly modelled as reads ($\src{\ev{Read\ \loc\ n}}$) and writes ($\src{\ev{Write\ \loc\ n}}$) instead of just uses.
-
-All languages have at least numbers as values ($\src{v}$) and second class functions ($\src{F}$).
-Functions are modelled as pairs containing the name of one argument and the body of the function.
-Bodies are just ordinary expressions, which can be simple binary operations ($\src{\lbinop{\expr[_1]}{\expr[_2]}}$), conditionals ($\src{\lifz{\expr[_1]}{\expr[_2]}{\expr[_3]}}$), function calls ($\src{\lcall{f}{e}}$) and returns ($\src{\lreturn{\expr}}$), as well as C-like memory operations.
-Programs have sets of pre-determined functions called libraries and they are marked as being part of some component ($\src{\library_{\comp}}$) or context ($\src{\library_{\ctx}}$).
-
-For the operational semantics, the runtime state ($\src{\Omega}$) is a triple consisting of a control-flow state ($\src{\cfstate}$), a control tag ($\src{\sandboxtag{}}$), and a memory state ($\src{\memstate}$).
-The latter carries information about pointers that are kept ``alive'' in pointer maps ($\src{\Delta}$), so that the semantics does not get stuck when encountering, e.g., a double-free.
-The memory state also carries two heaps to model sandboxing between a context ($\src{H^{\ctx}}$) and a component ($\src{H^{\comp}}$).
-
-}
-
-\subsection{$\src{L_{\tmssafe}}$: A Temporal but Not Spatial Memory Safe Language}\label{subsec:ltms}
-
-$\src{L_{\tmssafe}}$ is the only statically-typed language in this case study and restricts functions ($\src{F}$) to the typing signature $\src{\natt\to\natt}$. 
-The type system of $\src{L_{\tmssafe}}$ is inspired by $L^{3}$~\cite{morrisett2005L3,scherer2018fabulous} and enforces that every well-typed $\src{L_{\tmssafe}}$ program satisfies \gls*{tms}.
- % (\Cref{thm:wt:tms}).
-
-\begin{theorem}[$\src{L_{\tmssafe}}$-programs are \gls*{tms}]\label{thm:wt:tms}
-$\rsat{\src{\library_{\comp}}}{\tmssafe}$ \Coqed
-  % $\;$ 
-% 
-  % \begin{nscenter}
-  %   \hfill For any component $\src{\library_{\comp}}$, $\rsat{\src{\library_{\comp}}}{\tmssafe}$ \Coqed
-  % \end{nscenter}
-\end{theorem}
-
-\subsection{$\Ltrg$: A Memory-Unsafe Language}\label{subsec:lsms}
-
-$\Ltrg$ extends the syntax presented earlier (\Cref{subsec:cs:defs}) with dynamic typechecks ($\trg{\lhast{\expr}{\type}}$), evaluating to $\trg{0}$ if the type matches with the shape of $\trg{\expr}$ and $\trg{1}$ otherwise.
-%The pointers carry poison tags, which signify a pointer being allocated ($\trg{\poisoned}$) or freed ($\trg{\poisonless}$). 
-Furthermore, the syntax of $\Ltrg$ is extended with a way to inspect whether a pointer is freed ($\trg{\lispoisoned{x}}$), evaluating to $\trg{0}$ if it is freed and to $\trg{1}$ otherwise.
-Functions may receive arguments that are not $\trg{\natt}$, values are extended with tuples, and expressions are also extended with pair projections.
-
-% \vspace{-1em}
-\[
-  \begin{array}{rrcl}
-    \text{(Values)} & \trg{v} &:=& \dots \mid \trg{\lpair{v_1}{v_2}} \\
-    \text{(Expressions)} & \trg{\expr} &:=& \dots \mid \trg{\lhast{\expr}{\tau}} \mid \trg{\lispoisoned{x}}
-  \end{array}
-\]
-
-No changes are done to the trace model, but $\Ltrg$ has no static typing, making double-free code patterns possible.
-
-\subsection{$\Lms$: Another Memory-Unsafe Language}\label{subsec:lms}
-$\Lms$ is exactly equal to $\Ltrg$ (\Cref{subsec:lsms}), but used to emphasize that this is code after applying $\cc{\Ltrg}{\Lms}$ (\Cref{subsec:cs:ms}).
-
-\subsection{$\Lscct$: A Memory-Unsafe Language with a Constant Time Mode}\label{subsec:lscct}
-
-$\Lscct$ extends $\Lms$ (\Cref{subsec:lms}) with a ,,constant-time mode''. 
-The activation of the mode can be checked ($\obj{\lrddoit{x}{\expr}}$), changed ($\obj{\lwrdoit{D}}$), and is stored in program states ($\obj{\Omega}$).
-At the beginning of program execution of $\Lscct$ programs, the mode is turned off ($\obj{OFF}$).
-If the mode is enabled (i.e., set to $\obj{ON}$), the intuition is that no secrets are leaked. 
-This models the real-world\footnote{As present in Intel~\cite[p.80]{intel-refman} and ARM~\cite[p.~543]{arm-refman} processors.} data-independent timing mode as well as the result of compiling a program with FaCT~\cite{cauligi2019fact}.
-For example, FaCT rewrite code that branches to use a constant-time selection primitive.
-To not obfuscate our formalisation unnecessarily by duplicating syntax, we simply added the ,,constant-time mode'' to $\Lscct$.
-
-% $\Lscct$ extends $\Lms$ (\Cref{subsec:lms}) with a way to read from ($\obj{\lrddoit{x}{\expr}}$) and to write to ($\obj{\lwrdoit{D}}$) a model-specific register that controls a constant-time mode. 
-% This language is a combination of the effect of the FaCT compiler~\cite{cauligi2019fact} as well as a data-independent timing mode, a feature that is present in both Arm~\cite[p.~543]{arm-refman} and Intel~\cite[p.~80]{intel-refman} processors.
-% \MP{ can we be more specific about the fact pass? }
-% To this end, states ($\obj{\Omega}$) are extended with the value of the register ($\obj{D}$), which is initially set to be not active ($\obj{OFF}$).
-% If the register is marked active ($\obj{ON}$), the intuition is that no secrets can appear on specification traces.
-% If the register is marked inactive, secrets may appear on traces.
-% The state of the register can be read and written to, e.g., \Cref{tr:e-wrdoit-off}.
-% For the other languages seen earlier, the mode is intuitively always-on, i.e., the mode of execution always uses constant-time mode.
-
-% \vspace{-.5em}
-\[
-  \begin{array}{rrcl}
-    \text{(Mode Values)} & \obj{D} &:=& \obj{ON} \mid \obj{OFF} \\
-    \text{(Security Tags)} & \obj{\securitytag{}} &:=& \obj{\lock} \mid \obj{\unlock} \\
-    \text{(Base-Events)} & \obj{\preevent} &:=& \dots \mid \obj{iGet\ \loc\ v} \mid \obj{iSet\ \loc\ v} \\
-                       &&& \mid\ \obj{Binop\ n} \mid \obj{Branch\ n} \\
-    \text{(Events)} & \obj{\event} &:=& \obj{\preevent;\sandboxtag{};\securitytag{}} \mid \obj{\emptyevent} \mid \obj{\lightning} \\
-    \text{(Expressions)} & \obj{\expr} &:=& \dots \mid \obj{\lrddoit{x}{\expr}} \mid \obj{\lwrdoit{D}} \\
-                         &&&\mid\ \obj{\llet{x^{\securitytag{}}}{\expr[_1]}{\expr[_2]}}\\
-    \text{(States)} & \obj{\Omega} & := & \left(\obj{\cfstate};\obj{\sandboxtag{}};\obj{D};\obj{\memstate}\right)\\
-  \end{array}
-\]
-\begin{center}
-\newcommand{\expreval}[5]{{#1}\triangleright\xspace {#2}\xrightarrow{#5}\ {#3}\triangleright\xspace {#4}\xspace}
-\newcommand{\exprevalo}[5]{\expreval{\obj{#1}}{\obj{#2}}{\obj{#3}}{\obj{#4}}{\obj{#5}}}
-  \typerule{$e-\obj{wrdoit}-\text{off}$}{
-  }{
-    \pstepto{\rtt{\obj{\cfstate;\sandboxtag{};D;\memstate}}{\obj{\lwrdoit{OFF}}}}{\rtt{\obj{\cfstate;\sandboxtag{};OFF;\memstate}}{\obj{0^{\securitytag{}}}}}{\obj{\emptyevent}}
-  }{e-wrdoit-off}
-
-  \typerule{$e-\obj{get}-\in-$noleak}{
-    \obj{\memstate}=\obj{H^{\ctx};H^{\comp};\Delta_1,x\mapsto(\loc;\sandboxtag{};\poison),\Delta_2} &
-    \obj{\loc}+\obj{n}\in\text{dom }\obj{H^{\sandboxtag{}}} \\
-    \obj{\securitytag{''}}=\obj{\securitytag{}\sqcap\securitytag{'}} &
-    \obj{\event} = \obj{{iGet}\ \loc\ n;\sandboxtag{};\securitytag{''}}
-  }{
-    \exprevalo{
-    	\cfstate;\sandboxtag{'};ON;\memstate
-    }{
-      	x^{\securitytag[n^{\securitytag{'}}] }
-    }{
-    	\cfstate;\sandboxtag{'};ON;\memstate
-    }{
-    	(H^{\sandboxtag{}}(\loc+n))^{\securitytag{''}}
-    }{
-      \event
-    }
-  }{to-e-get-in-noleak}
-\end{center}
-
-The language also adds user annotations $\obj{\securitytag{}}$ for the secrecy of variables, which can be either private (high secrecy) $\obj{\lock}$ or public (low secrecy) $\obj{\unlock}$.
-Security tags $\obj{\securitytag{}}$ are arranged in the usual secrecy lattice~\cite{zdancewicphd}, where $\obj{\unlock}\sqsubseteq\obj{\lock}$.
-% $\obj{\lock}\sqsubseteq\obj{\securitytag{}}$ and $\obj{\securitytag{}}\sqsubseteq\obj{\unlock}$.
-
-Memory accesses to secret data need to be present to reason about memory safety, even when execution is in constant-time mode, e.g, \Cref{tr:to-e-get-in-noleak}.
-$\Lscct$ extends base-events with $\obj{\ev{iGet}\ \loc\ \valueexpr}$ and $\obj{\ev{iSet}\ \loc\ \valueexpr}$ (for data $\obj{i}$ndependent get and set) to prevent secrets from leaking but still enable reasoning about memory safety.
-Due to the technical setup, the rule needs to check if the access is in bounds ($\obj{\loc}+\obj{n}\in\obj{H^t}$) and update the secrecy tag ($\obj{\securitytag{''}}=\obj{\securitytag{}\sqcap\securitytag{'}}$) with the least upper bound of the tags, according to the aforementioned lattice.
-The precise information carried by the event, e.g., location ($\obj{\loc}$) is taken from the pointer map, which carries information irrelevant to this rule ($\obj{\poison}$).
-
-Base-events include $\obj{\ev{Branch}\ n}$ and $\obj{\ev{Binop}\ n}$ that are emitted when evaluating a branch or certain binary expressions, such as division, respectively, whenever the constant-time mode is inactive.
-Events are extended with a security-tag ($\obj{\securitytag{}}$) to signal the secrecy of the involved data.
-
-The evaluation steps are amended to propagate the security-tag annotations $\obj{\securitytag{}}$.
-When the constant-time mode is inactive, base-events $\obj{\ev{Branch}\ n}$ and $\obj{\ev{Binop}\ n}$ are emitted for conditionals and binary operations, respectively.
-Otherwise, just like in the semantics of the earlier languages, $\obj{\emptyevent}$ is emitted for binary and branching operations.
-%To mark this mode as active or inactive, \Cref{tr:e-wrdoit} allows to access the register in $\obj{\statevar}$.
-
-\Cref{ex:lscct} illustrates the differences between $\Lscct$ and other languages.
-
-\begin{exampleenv}[$\Lscct$ with and without constant-time mode]\label{ex:lscct}
-  Consider again \Cref{ex:strncpy}, with a context copying the string \texttt{Hello World}, where everything is marked with a high security tag: $\lock$.
-  The top half of \Cref{fig:ex-cct} (titled $\Lscct$), describes the execution trace of the program, while the bottom side of the table (titled ``Specification''), describes the related specification trace (\Cref{subsec:msctss:tracemodel}).
-  Read in parallel from top to bottom, the figure shows parts of the execution trace. 
-  In each half, the left column (\textit{Active}) has constant-time mode $\obj{ON}$ and the right one (\textit{Inactive}) has it $\obj{OFF}$.
-
-  When the constant-time mode is off, the execution yields events in similar fashion to before (\Cref{subsec:ltms,subsec:lsms,subsec:lms}).
-  But, if it is turned on, then the branching event \colorbox{yellow}{does not fire} anymore and both reading and writing to memory is related to a specification trace with \colorbox{lightgray}{no exposed secrets}.
-\end{exampleenv}
-\begin{figure}[!htb]
-	\vspace{-1em}\small
-  $$
-  % {
-  \begin{array}{ccc}
-    & \Lscct &        \\
-    \text{Active} & \mid & \text{Inactive} \\\hline
-    \obj{\ev{Alloc}\ \loc_{x}\ 12;\comp;\unlock} & \mid & \obj{\ev{Alloc}\ \loc_{x}\ 12;\comp;\unlock} \\
-    \obj{\ev{Alloc}\ \loc_{y}\ 12;\comp;\unlock} & \mid & \obj{\ev{Alloc}\ \loc_{y}\ 12;\comp;\unlock} \\
-    \obj{{\ev{iGet}}\ \loc_{x}\ 0;\comp;}\text{\colorbox{lightgray}{$\obj{\lock}$}} & \mid & \obj{\ev{Get}\ \loc_{x}\ 0;\comp;\lock} \\
-    \text{\colorbox{yellow}{$\obj{\emptyevent}$}} & \mid & \obj{\ev{Branch}\ 0;\comp;\lock} \\
-    \obj{{\ev{iGet}}\ \loc_{x}\ 0;\comp;}\text{\colorbox{lightgray}{$\obj{\lock}$}} & \mid & \obj{\ev{Get}\ \loc_{x}\ 0;\comp;\lock} \\
-    \obj{{\ev{iSet}}\ \loc_{y}\ 0\ \mathtt{'H'};\comp;}\text{\colorbox{lightgray}{$\obj{\lock}$}} & \mid & \obj{\ev{Set}\ \loc_{y}\ 0\ \mathtt{'H'};\comp;\lock} \\
-    \obj{{\ev{iGet}}\ \loc_{x}\ 1;\comp;}\text{\colorbox{lightgray}{$\obj{\lock}$}} & \mid & \obj{\ev{Get}\ \loc_{x}\ 1;\comp;\lock} \\
-    \text{\colorbox{yellow}{$\obj{\emptyevent}$}} & \mid & \obj{\ev{Branch}\ 0;\comp;\lock} \\
-    \vdots & \mid & \vdots \\
-    \obj{{\ev{iGet}}\ \loc_{x}\ 12;\comp;}\text{\colorbox{lightgray}{$\obj{\lock}$}} & \mid & \obj{\ev{Get}\ \loc_{x}\ 12;\comp;\lock} \\
-    \text{\colorbox{yellow}{$\obj{\emptyevent}$}}& \mid & \obj{\ev{Branch}\ 1;\comp;\lock} \\
-    \obj{\ev{Dealloc}\ \loc_{y};\comp;\unlock} & \mid & \obj{\ev{Dealloc}\ \loc_{y};\comp;\lock} \\
-    \obj{{\ev{iGet}}\ \loc_{y}\ 6;\comp;}\text{\colorbox{lightgray}{$\obj{\lock}$}} & \mid & \obj{\ev{Get}\ \loc_{y}\ 6;\comp;\lock} \\
-  \end{array}
-  % }
-  $$
-  $$
-  % {
-  \begin{array}{ccc}
-          & \text{Specification} & \\
-  \text{Active} & \mid & \text{Inactive} \\\hline
-     \ev{Alloc}\ \loc_{x}\ 12;\unlock & \mid & \ev{Alloc}\ \loc_{x}\ 12;\unlock\\
-     \ev{Alloc}\ \loc_{y}\ 12;\unlock & \mid & \ev{Alloc}\ \loc_{x}\ 12;\unlock\\
-     \ev{Use}\ \loc_{x}\ 0;\text{\colorbox{lightgray}{$\unlock$}} & \mid & \ev{Use}\ \loc_{x}\ 0;\lock\\
-     \emptyevent & \mid & \ev{Branch}\ 0;\lock\\
-     \ev{Use}\ \loc_{x}\ 0;\text{\colorbox{lightgray}{$\unlock$}} & \mid & \ev{Use}\ \loc_{x}\ 0;\lock\\
-     \ev{Use}\ \loc_{y}\ 0;\text{\colorbox{lightgray}{$\unlock$}} & \mid & \ev{Use}\ \loc_{y}\ 0;\lock\\
-     \ev{Use}\ \loc_{x}\ 1;\text{\colorbox{lightgray}{$\unlock$}} & \mid & \ev{Use}\ \loc_{x}\ 1;\lock\\
-     \emptyevent & \mid & \ev{Branch}\ 0;\lock\\
-     \vdots & \mid & \vdots \\
-      \ev{Use}\ \loc_{x}\ 12;\text{\colorbox{lightgray}{$\unlock$}} & \mid & \ev{Use}\ \loc_{x}\ 12;\lock\\
-      \emptyevent & \mid & \ev{Branch}\ 1;\lock\\
-      \ev{Dealloc}\ \loc_{y};\unlock & \mid & \ev{Dealloc}\ \loc_{y};\unlock\\
-      \ev{Use}\ \loc_{y}\ 6;\text{\colorbox{lightgray}{$\unlock$}} & \mid & \ev{Use}\ \loc_{y}\ 6;\lock\\
-  \end{array}
-  % }
-  $$
-  \caption{Traces for \Cref{ex:lscct}.}
-  \label{fig:ex-cct}
-\end{figure}
-
-\subsection{$\Lspec$: A Memory-Unsafe Language with Speculation}\label{subsec:lspec}
-
-$\Lspec$ extends $\Lscct$ (\Cref{subsec:lscct}) with speculative dynamic semantics that is inspired by existing work~\cite{guarnieri2018spectector,fabian2022automatic}.
-After branching, speculation starts by pushing the current configuration into the speculation state ($\ird{S}$), and run subsequent code in a predetermined window (whose size is $\omega$) until a rollback of operational state is performed.
-The window is set inside the speculation state.
-Within such windows, data may leak, this is marked as high $\ird{\lock}$ and with an annotation that indicates the respective speculative execution variant. 
-In this paper, we just consider SPECTRE-PHT~\cite{kocher2019spectre}, whose starting point is \Cref{tr:e-ifz-true-spec}.
-% where $\omega$ is the size of the speculation window, i.e., a number of reduction steps to be taken under speculative semantics.
-Additionally, $\ird{\lock}$ may also carry a $\ird{NONE}$ annotation, to signal that this is a leak irrespective of speculation, i.e., as defined earlier (\Cref{subsec:lscct}).
-The semantics is kept general enough to allow for future extension to support different variants~\cite{kocher2019spectre,maisuradze2018ret2spec,horn2019zero}.
-
-In terms of language features, $\Lspec$ includes a new barrier operation $\ird{\lbarrier}$ that blocks speculative execution. 
-To facilitate speculative execution in a non-assembly-like language, a stack of operational state is used and speculation is active if that stack has a size greater than one.
-This is exploited in \Cref{tr:e-spec-eat} to consume any leftover speculation.
-
-\vspace{-.5em}
-\[
-  \begin{array}{rrcl}
-    \text{(Base-Events)} & \ird{\preevent} &:=& \dots \mid \ird{Spec} \mid \ird{Rlb} \mid \ird{Barrier}\\
-    \text{(Expressions)} & \ird{\expr} &:=& \dots \mid \ird{\lbarrier} \\
-    \text{(Speculation State)} & \ird{S} &:=& \ird{\Omega\triangleright\expr} \mid \ird{S},\left(\ird{\Omega};\ird{n};\ird{\expr}\right)
-  \end{array}
-\]
-
-\begin{center}
-\newcommand{\expreval}[5]{{#1}\triangleright\xspace {#2}\xrightarrow{#5}\ {#3}\triangleright\xspace {#4}\xspace}
-\newcommand{\exprevald}[5]{\expreval{\ird{#1}}{\ird{#2}}{\ird{#3}}{\ird{#4}}{\ird{#5}}}
-
-  %
-  \typerule{$e-\ird{ifz}-\ird{true}-\text{spec}$}{
-    \ird{S} = \rtt{(\ird{\cfstate['];\sandboxtag{};D;\memstate});\ird{S}}{\ird{\expr[_1]}},({\ird{\cfstate;\sandboxtag{};D;\memstate}};\omega;{\ird{\expr[_2]}})
-  }{
-    \pstepto{\rtt{\ird{\cfstate;\sandboxtag{};D;\memstate}}{\ird{\lifz{0^{\securitytag{}}}{\expr[_1]}{\expr[_2]}}}}{\ird{S}}{\ird{{Branch}\ 0;\sandboxtag{};\securitytag{}}\cdot\ird{{Spec};\sandboxtag{};\securitytag{}}}
-  }{e-ifz-true-spec}
-  %
-  \typerule{$e-\ird{spec}-\text{eat}$}{
-    \ird{n} > 0 &
-    \pstepto{\rtt{\ird{\Omega}}{\ird{\expr}}}{\rtt{\ird{\Omega'}}{\ird{\expr[']}}}{\ird{\event}}
-  }{
-    \ghoststepto{\ird{S},(\ird{\Omega};\ird{n};\ird{\expr})}{\ird{S},(\ird{\Omega'};\ird{n}-1;\ird{\expr'})}{\ird{\event}}
-  }{e-spec-eat}
-  % 
-  \typerule{$e-\ird{spec}-\text{eaten}$}{
-  }{
-    \ghoststepto{\ird{S},(\ird{\Omega};\ird{0};\ird{\expr})}{\ird{S}}{\ird{Rlb;\Omega.\sandboxtag{};\unlock}}
-  }{e-spec-eaten}
-  %
-  \typerule{$e-\ird{spec}-\text{barrier}$}{
-  }{
-    \ghoststepto{\ird{S},(\ird{\Omega};\ird{n};\ird{\lbarrier})}{\ird{S},(\ird{\Omega};\ird{0};\ird{0})}{\ird{Barrier;\Omega.\sandboxtag{};\unlock}}
-  }{e-spec-barrier}
-\end{center}
-
-The trace model is extended with $\ird{{Spec}}$, $\ird{{Rlb}}$, and $\ird{{Barrier}}$ events that signal the respective operational action. 
-A $\ird{\lbarrier}$ prevents any execution whatsoever besides $\ird{{Rlb}}$ when run in speculative execution mode and does nothing when run in normal mode, e.g., \Cref{tr:e-spec-barrier}.%
-\footnote{In the rule, the notation $\cfstate' = \cfstate[.\text{win} = 0]$ means that $\cfstate'$ is a copy of $\cfstate$ up to field $\text{win}$, which is set to $0$.}
-Lastly, $\ird{Rlb}$ is emitted when the speculation window is zero and the operational state is rolled back (see \Cref{tr:e-spec-eaten}).
-
-\section{Case Study: Composing Secure Compiler Passes and Optimisations \pages{7}}\label{sec:casestud:rtp}
-This section defines several secure compilers, each of which robustly preserves a different property of interest as depicted in \Cref{fig:pipeline}.
-\begin{figure*}[!h]
-  \centering
-  \begin{tikzpicture}
-    \node (S) {$\src{L_{\tmssafe}}$};
-    \node[right=2.0 of S] (T) {$\trg{L}$};
-    \node[right=2.0 of T] (M) {$\irl{L_{\mssafe}}$};
-    \node[below right=1.5 and 1.0 of M] (D0) {$\irl{L_{\mssafe}}$};
-    \node[above right=1.5 and 1.0 of M] (C0) {$\irl{L_{\mssafe}}$};
-    \node[right=3.0 of M] (E) {$\irl{L_{\mssafe}}$};
-    \node[right=2.0 of E] (O) {$\obj{L_{\scctsafe}}$};
-    \node[right=2.0 of O] (Z) {$\ird{L_{\mathghost}}$};
-
-    \draw[->] (S) to[sloped] node[align=center] (tmsedge) {\gls*{tms}\\ \Cref{thm:cca:rtp:tms}} (T);
-    \draw[->] (T) to[sloped] node[align=center] {\gls*{sms}\\ \Cref{thm:ccb:rtp:sms}} (M);
-    \draw[->] (M) to[sloped] node[align=center] {\gls*{dce}\\ \Cref{thm:ccdce:rtp:ms}} (D0);
-    \draw[->] (M) to[sloped] node[align=center] {\gls*{cf}\\ \Cref{thm:cccf:rtp:ms}} (C0);
-    \draw[->] (D0) to[sloped] node[align=center] {\gls*{cf}\\ \Cref{thm:cccf:rtp:ms}} (E);
-    \draw[->] (C0) to[sloped] node[align=center] {\gls*{dce}\\ \Cref{thm:ccdce:rtp:ms}} (E);
-    \draw[->] (E) to[sloped] node[align=center] {\gls*{scct}\\ \Cref{thm:ccscct:rtp:scct}} (O);
-    \draw[->] (O) to[sloped] node[align=center] {\gls*{ss}\\ \Cref{thm:ccspec:rtp:spec}} (Z);
-
-    % Sections
-    \node (sectms) at ($(S)+(1.25,2.5)$) {{\Cref{subsec:cs:tms}}};
-    \node (secsms) at ($(T)+(1.25,2.5)$) {{\Cref{subsec:cs:ms}}};
-    \node (secopts) at ($(M)+(2.0,2.5)$) {{\Cref{subsec:cs:opts}}};
-    \node (secscct) at ($(E)+(1.5,2.5)$) {{\Cref{subsec:cs:scct}}};
-    \node (secss) at ($(O)+(1.5,2.5)$) {{\Cref{subsec:cs:ss}}};
-
-    \draw[thick,dotted,->] ($(S) + (0.2,-0.27)$) to ($(S) - (-0.2,1)$) to node[align=center] {\gls*{ms}\\ \Cref{corr:ccab:rtp:ms}} ($(M) - (0.0,1)$) to (M);
-
-    \draw[thick,dotted,->] (S) to ($(S) - (0.0,2.5)$) to node[align=center,pos=.65] {\gls*{specms}\\ \Cref{thm:ccall:rtp:specms}} ($(Z) - (0.0,2.5)$) to (Z);
-
-    % \draw[thick,dotted,->] (S) to[bend right=33,sloped] node[align=center] {\gls*{ms}\\ \Cref{corr:ccab:rtp:ms}} (M);
-    \draw[thick,dotted,->] (M) to[bend right=0,sloped] node[align=center] {\gls*{ms}\\ \Cref{thm:cccfccdce:rtp:ms}} (E);
-%    \draw[thick,dotted,->] (S) to[out=-45,in=198,sloped] node[align=center] {\gls*{ms}+\gls*{scct}\\ \Cref{thm:ccall:rtp:msscct}} (O);
-    % \draw[thick,dotted,->] (S) to[out=-40,in=210,sloped] node[align=center,pos=.8] {\gls*{specms}\\ \Cref{thm:ccall:rtp:specms}} (Z);
-
-    \draw[dashed] ($(S)-(0.0,-0.5)$) -- ($(S)-(0.0,-2.5)$);
-    \draw[dashed] ($(T)-(0.0,-0.5)$) -- ($(T)-(0.0,-2.5)$);
-    \draw[dashed] ($(M)-(0.0,-0.5)$) -- ($(M)-(0.0,-2.5)$);
-    \draw[dashed] ($(E)-(0.0,-0.5)$) -- ($(E)-(0.0,-2.5)$);
-    \draw[dashed] ($(O)-(0.0,-0.5)$) -- ($(O)-(0.0,-2.5)$);
-    \draw[dashed] ($(Z)-(0.0,-0.5)$) -- ($(Z)-(0.0,-2.5)$);
-  \end{tikzpicture}
-  \vspace{-1em}
-  \caption{Visualisation of the optimising compilation pipeline that preserves \gls*{specms}. %
-    Vertices in the graph are the programming languages from earlier sections (\Cref{sec:casestud:defs}). %
-    Full edges are secure compilers passes.
-    Dotted edges are composition of passes and use the presented framework (\Cref{sec:sequential}) to indicate the property they preserve. %
-    The dashed lines partition the graph into the sections where the respective theorems are presented.
-  }\label{fig:pipeline}
-\end{figure*}
-The section demonstrates the power of the framework (\Cref{sec:sequential}) by composing these compilers for a secure and optimising compilation chain that robustly preserves \gls*{specms}.
-The first step in this chain is the compiler from $\src{L_{\tmssafe}}$ to $\trg{L}$ that robustly preserves just \gls*{tms} (\Cref{thm:cca:rtp:tms}).
-From here, another pass from $\trg{L}$ to $\irl{L_{\mssafe}}$ ensures that no out-of-bounds accesses can happen and, thus, programs at this point attain \gls*{sms} (\Cref{thm:ccb:rtp:sms}).
-Since these properties compose into \gls*{ms}, composing these passes yields a compiler that robustly preserves \gls*{ms} (\Cref{corr:ccab:rtp:ms}).
-Then, the section presents two optimisation passes, namely \gls*{cf} and \gls*{dce}, each of which robustly preserves \gls*{ms} (\Cref{thm:ccdce:rtp:ms,thm:cccf:rtp:ms}).
-These passes can be freely ordered in the compilation chain without compromising memory safety (\Cref{thm:cccfccdce:rtp:ms}).
-The next step in the chain ensures that code stays \gls*{scct} (\Cref{thm:ccscct:rtp:scct}) when compiled from $\Lms$ to $\Lscct$, which is done by switching on a constant-tame mode for the computation.
-Lastly, by introducing barriers immediately after branches, speculative leaks via SPECTRE-PHT are prevented when compiling $\Lscct$ to $\Lspec$.
-The final result is that the whole compilation chain robustly preserves \gls*{specms} (\Cref{thm:ccall:rtp:specms}).
-
-\subsection{Robust Temporal Memory Safety Preservation}\label{subsec:cs:tms}
-
-The secure compiler from $\Ltms$ to $\Ltrg$ needs to ensure that when execution switches from context to component, the type signatures are respected.
-Thus, the compiler inserts the following dynamic typechecks before entering the body of a component-defined function (anything elided is a trivial identity function from source to target):
-
-\vspace{-1em}
-{
-\begin{gather}
-  \begin{align*}%{rcl}
-    %\cca(\src{x}) = &\ \trg{x} 
-    % \\
-  	%&
-  	%\qquad
-    %\cca(\src{\lbinop{\expr[_{1}]}{\expr[_{2}]}}) = &\ \lbinop[\trg]{\left[\cca(\src{\expr[_{1}]})\right]}{\left[\cca(\src{\expr[_{2}]})\right]} \\
-    %\cca(\src{n}) = &\ \trg{n} 
-    % \\
-    %&
-    \cca(\src{\lget{x}{\expr}}) = &\ \lget[\trg]{\trg{x}}{\left[\cca(\src{\expr})\right]} \\
-    \cca(\src{\ldelete{x}}) = &\ \ldelete[\trg]{\left[\cca(\src{x})\right]} 
-    \\
-    \cca(\src{\lfunction{g}{x:\natt\to\type[_{e}]}{\expr}})  =&
-  \end{align*}
-  \\
-  \begin{align*}
-%    \cca(\src{\lset{x}{\expr[_{1}]}{\expr[_{2}]}}) & = \lset[\trg]{\left[\cca(\src{x})\right]}{\left[\cca(\src{\expr[_{1}]})\right]}{\left[\cca(\src{\expr_{2}})\right]} \\
-% \\ 
-\lfunction[\trg]{\trg{g}}{\trg{x}}{\lifz[\trg]{\trg{\lhast{x}{\natt}}}{
-                                                            \left[\cca(\src{\expr})\right] %
-                                                                                                 }{\labort[\trg]}}
-  \end{align*}
-\end{gather}
-}
-
-\noindent Since $\trg{L}$ has no static typing, an attacker $\trg{\library_{\ctx}}$ can invoke a component function accepting a $\src{\natt}$ with $\trg{\lpair{17}{29}}$.
-With the dynamic check, the compiler ensures that execution aborts in such cases.
-%The compiler does not insert other checks and proceeds as the identity function (which in this paper amounts to a simple re-colouring of $\src{L_{\tmssafe}}$ to $\trg{L}$ expressions).
-
-Compiling the \texttt{strncpy} function from \Cref{sec:introduction} with $\cca$, the compiler would in this case ensure that the arguments that are evaluated in the compiled \texttt{strncpy} are valid.
-%This effectively has no influence on the resulting trace, so a corresponding cross-language trace relation $\sim^{\Ltms}_{\Ltrg}$ is just equality.
-
-$\cca$ is robustly preserving (\Cref{def:rtp}) \gls*{tms}:
-\begin{theorem}[$\cca$ secure w.r.t. \gls*{tms}]\label{thm:cca:rtp:tms}
-    $\rtp{\cca}{\tmssafe}$ % \Coqed
-\end{theorem}
-
-\subsection{Robust Spatial Memory Safety Preservation}\label{subsec:cs:ms}
-
-The spatial memory safety preserving compiler from $\Ltrg$ to $\Lms$ only inserts bounds-checks whenever reading from or writing to memory in order to enforce \gls*{sms}.
-These bounds checks need the bounds information, which the compiler keeps around by introducing a fresh identifier $\irl{x_{SIZE}}$ for each allocation that binds $\irl{x}$.
-Then, it is simply a matter of referring to that variable and ensuring that memory accesses are in the interval $[\irl{0},\irl{x_{SIZE}})$.
-When the check fails, the code aborts.
-
-\begin{nscenter}
-% \small
-  $$
-  \begin{array}{rcl}
-    \ccb(\trg{\lnew{x}{\expr[_{1}]}{\expr[_{2}]}}) & = 
-                                                   & \llet[\irl]{\irl{x_{SIZE}}}{\ccb(\trg{\expr[_{1}]})}{}
-    		\\&&
-    		\lnew[\irl]{\irl{x}}{\irl{x_{SIZE}}}{\ccb(\trg{\expr[_{2}]})}
-    		 \\
-  \ccb(\trg{\lget{x}{\expr}}) & = 
-                              & \llet[\irl]{\irl{x_{n}}}{\ccb(\trg{\expr})}{}
-  	\\&&
-  \lifz[\irl]{\irl{0\le x_{n}<x_{SIZE}}}{\\&&\irl{\lget{x}{x_{n}}}}{}
-  		\irl{\labort}
-  	  \\
-  \ccb(\trg{\lset{x}{\expr[_{1}]}{\expr[_{2}]}}) & = 
-                                                 & \llet[\irl]{\irl{x_{n}}}{\ccb(\trg{\expr[_{1}]})}{}
-  		\\&&
-  \lifz[\irl]{\irl{0\le x_{n}<x_{SIZE}}}{\\&&\lset[\irl]{\irl{x}}{\irl{x_{n}}}{}
-  		\ccb(\trg{\expr[_{2}]})
-  		}{\irl{\labort}} 
-  		% \\
-  \end{array}
-  $$
-\end{nscenter}
-
-\begin{exampleenv}[Instrumented \texttt{strncpy}]
-  Consider again \texttt{strncpy}, but instrumented for \gls*{sms}:
-    \begin{lstlisting}[language=c,basicstyle=\ttfamily, morekeywords={size_t}]
-void strncpy(size_t n, size_t dst_size, char *dst,
-             size_t src_size, char *src) {
-  for(size_t i = 0; i < src_size
-      && src[i] != '\0' && i < n; ++i) {
-    if(i < src_size && i < dst_size) {
-      dst[i] = src[i];
-    }
-  }
-}
-    \end{lstlisting}
-Consider context \texttt{strncpy(2, x, y)}, where \texttt{x} and \texttt{y} are pointers to valid regions of memory with allocated space for exactly two cells and do not contain the null-terminating character \texttt{'\textbackslash 0'}.
-% 
-Without the \gls*{sms} pass, the event $\ev{Use}\ \loc_{x}\ 2;\comp;\unlock$ would appear on the trace, but that indicates an out-of-bounds access! 
-Fortunately, with \gls*{sms} mitigation in place, that event does not appear during execution, since the bounds check prevents the condition \texttt{src[i] != '\textbackslash 0'} from executing.
-\end{exampleenv}
-
-Contrary to the previous compiler, $\ccb$ may change the trace of the original $\Ltrg$ program: if there is a memory access, it needs to be protected with a bounds check.
-% 
-The corresponding relation $\sim^{\Ltrg}_{\Lms} : \trg{\trace}\times\irl{\trace}$ that describes this semantic effect of the compiler is defined partially below.
-% 
-We omit action $\trg{Set}$, which is treated analogously, and any other event, which is related to its cross-language equivalent.
-% \footnote{Up to colours.}
-% \footnote{In the interest of space, the full details are in the technical report.}
-
-{
-\[
-  \typerulenolabel{xrel:sms:read}{\trg{n}\text{ in bounds}}{\trg{Get\ \loc\ n;\comp}\sim^{\Ltrg}_{\Lms}\irl{Get\ \loc\ n;\comp}}
-\]
-% \[
-%   \typerulenolabel{xrel:sms:panic}{\trg{n}\text{ out of bounds}}{\trg{Get\ \loc\ n;\comp}\sim^{\Ltrg}_{\Lms}\irl{\lightning}}
-% \]
-}
-
-% Note that neither $\Ltrg$ nor $\Lms$ expose branches on the trace.\MP{so what?}
-For simplicity, we elide the environment that this relation carries around in order to bind each location to its metadata (such as its size), and resolve the ``$\trg{n}$ in/out of bounds'' premise.
-% The bounds check mentioned in the premises of the rules above is left abstract for the sake of simplicity, the interested reader can find
-% since the precise definition needs some technicalities that only inhibit readability without any technical insights.
-% 
-We can now prove that compiler $\ccb$ robustly preserves \gls*{sms}.
-\begin{theorem}[$\ccb$ secure w.r.t. \gls*{sms}]\label{thm:ccb:rtp:sms}
-  $\rtpsim{\ccb}{\smssafe}{\sim^{\Ltrg}_{\Lms}}$ % \Coqed
-\end{theorem}
-
-At this point we can compose $\ccb$ with the previous compiler ($\cca$), but in order to do so, we need a trace relation from $\Ltms$ to $\Lms$.
-% 
-We can obtain this relation by composing the trace relation we just defined ($\sim^{\Ltrg}_{\Lms}$) with the one used by the previous compiler: $\sim^{\Ltms}_{\Ltrg} : \src{\trace}\times\trg{\trace}$.
-% 
-The latter has not been previously defined (nor has it been used in the related theorem) because that is just an equality relation, since the trace models of $\Ltms$ and of $\Ltrg$ are the same.
-% 
-Thus, we formally define $\sim^{\Ltms}_{\Lms}: \src{\trace} \times \irl{\trace}$ as the following composition: $\sim^{\Ltms}_{\Ltrg}\bullet\sim^{\Ltrg}_{\Lms}$.
-% 
-With this relation, \Cref{corr:ccab:rtp:ms} states that the composition of $\cca$ and $\ccb$ is secure w.r.t.
-\gls*{ms} and it follows from \Cref{thm:cca:rtp:tms,thm:ccb:rtp:sms} using \Cref{thm:rtpsim:sig}.
-% 
-\begin{corollary}[$\cca\circ\ccb$ secure w.r.t. \gls*{ms}]\label{corr:ccab:rtp:ms}
-  $\;$
-
-  \begin{nscenter}
-    $\rtpsim{\cca\circ\ccb}{\mssafe}{\sim^{\Ltms}_{\Lms}}$ % \Coqed
-  \end{nscenter}
-\end{corollary}
-% 
-This proof requires another precondition besides \Cref{thm:ccb:rtp:sms,thm:cca:rtp:tms}: $\sim^{\Ltms}_{\Ltrg}$ needs to be well-formed with respect to $\smssafe$.
-This follows trivially since $\sim^{\Ltms}_{\Ltrg}$ is an equality. 
-% 
-\begin{lemma}[$\sim^{\Ltms}_{\Ltrg}$ well-formed w.r.t. $\smssafe$]\label{lem:wf:ltmsltrg}
-$\;$
-
-  \begin{nscenter}
-  $\wfcsig{\sim^{\Ltms}_{\Ltrg}}{\smssafe}$
-  \end{nscenter}
-\end{lemma}
-
-\subsection{Optimising Compilers}\label{subsec:cs:opts}
-
-This section defines two optimising compiler passes from $\Lms$ to $\Lms$ which perform \gls*{dce} and \gls*{cf}, respectively.
-The \gls*{dce} pass applies a na\"ive rewrite rule on conditionals.
-The \gls*{cf} pass relies on an auxiliary function \texttt{mix} that uses a substitutions accumulator $\irl{\substlist}$ in order to rewrite constant binary operations, e.g., $\irl{{17}-1}$ to $\irl{16}$, and replace variables that are assigned to constants, e.g., $\irl{\llet{x}{7}{x}}$ to $\irl{7}$.
-
-\vspace{-1em}
-\begin{gather*}
-  \begin{align*}
-    \ccdce(\irl{\lifz{true}{\expr[_{1}]}{\expr[_{2}]}}) & = \ccdce(\irl{\expr[_{1}]}) &\\
-    \ccdce(\irl{\lifz{false}{\expr[_{1}]}{\expr[_{2}]}}) & = \ccdce(\irl{\expr[_{2}]}) &
-  \end{align*}
-  \\
-  % \begin{align*}
-  %   \ccdce(\irl{\lbinop{\expr[_{1}]}{\expr[_{2}]}}) & = \lbinop[\irl]{\ccdce(\irl{\expr[_{1}]})}{\ccdce(\irl{\expr[_{2}]})} &
-  % \end{align*}
-  % \\[0.25cm]
-  \begin{align*}
-    \cccf(\irl{\expr}) & = \partialeval{\irl{\expr}}{\irl{\hole{\cdot}}} &
-  \end{align*}
-  \\[0.125cm]
-  \begin{align*}
-   \partialeval{\irl{x}}{\irl{\substlist}} & = \irl{n} 
-   	\qquad\qquad \text{if } \irl{\subst{n}{x}}\in\irl{\substlist} \\
-   \partialeval{\irl{x}}{\irl{\substlist}} & = \irl{x} 
-   \qquad\qquad \text{if } \irl{\subst{n}{x}}\notin\irl{\substlist} \\
-   \partialeval{\irl{\lbinop{n}{m}}}{\irl{\substlist}} & = \irl{k} 
-   \qquad\qquad \text{if } \lbinop{\irl{n}}{\irl{m}}=k \\
-   %\partialeval{\irl{\lbinop{\expr[_{1}]}{\expr[_{2}]}}}{\irl{\substlist}} & = \lbinop[\irl]{\partialeval{\irl{\expr[_{1}]}}{\irl{\substlist}}}{\partialeval{\irl{\expr[_{2}]}}{\irl{\substlist}}} \\
-   \partialeval{\irl{\llet{x}{n}{\expr}}}{\irl{\substlist}} & = \partialeval{\irl{\expr}}{\irl{\subst{n}{x}\cdot\substlist}} 
-  % \\
-  % \partialeval{\irl{\lget{x}{\expr}}}{\irl{\substlist}} & = \lget[\irl]{\irl{x}}{\partialeval{\irl{\expr}}{\irl{\substlist}}}
-% \end{align*}
-\end{align*}
-% \\
-% \begin{align*}
-%   \partialeval{\irl{\llet{x}{\expr[_{1}]}{\expr[_{2}]}}}{\irl{\substlist}} & = \llet[\irl]{\irl{x}}{\partialeval{\irl{\expr[_{1}]}}{\irl{\substlist}}}{\\&\partialeval{\irl{\expr[_{2}]}}{\irl{\substlist}}} \\
-%   \partialeval{\irl{\lifz{\expr[_{1}]}{\expr[_{2}]}{\expr[_{3}]}}}{\irl{\substlist}} & = \lifz[\irl]{\partialeval{\irl{\expr[_{1}]}}{\irl{\substlist}}}{\\&\partialeval{\irl{\expr[_{2}]}}{\irl{\substlist}}}{\partialeval{\irl{\expr[_{3}]}}{\irl{\substlist}}} \\
-% \end{align*}
-\end{gather*}
-% \vspace{-3em}
-
-Note that both passes have no effect on the resulting trace of a program, up to $\emptyevent$-steps. 
-Because of this, both passes have equality as corresponding cross language trace relation. 
-Moreover, it is straightforward to prove both passes as secure (\Cref{def:rtp}) w.r.t. \gls*{ms}.
-
-\begin{theorem}[$\ccdce$ secure w.r.t. \gls*{ms}]\label{thm:ccdce:rtp:ms}
-  $\rtp{\ccdce}{\mssafe}$ %\Coqed
-\end{theorem}
-\begin{theorem}[$\cccf$ secure w.r.t. \gls*{ms}]\label{thm:cccf:rtp:ms}
-  $\rtp{\cccf}{\mssafe}$ %\Coqe
-\end{theorem}
-
-With both \Cref{thm:ccdce:rtp:ms,thm:cccf:rtp:ms} it follows from \Cref{corr:swappable} that the two passes can be interchanged arbitrarily:
-
-\begin{theorem}[$\cccf\circ\ccdce$ and $\cccf\circ\ccdce$ are secure w.r.t. \gls*{ms}]\label{thm:cccfccdce:rtp:ms}
-$\;$
-
-  \begin{nscenter}
-  \phantom{and~}\!\!$\rtp{\cccf\circ\ccdce}{\mssafe}$
-
-  and~$\rtp{\ccdce\circ\cccf}{\mssafe}$ % \Coqed
-  \end{nscenter}
-\end{theorem}
-
-\subsection{Robust Strict Cryptographic Constant Time Preservation}\label{subsec:cs:scct}
-
-This section defines a compiler $\ccscct$ from $\Lms$ to $\Lscct$ that robustly preserves \gls*{scct}. 
-Given the fact that $\Lscct$ provides a \gls*{cct}-mode that can be turned on or off, the compiler inserts wrapper code for function calls and function bodies to ensure that execution in the component always happen in \gls*{cct}-mode.
-This simple flag combines the effect of FaCT~\cite{cauligi2019fact}
-
-\vspace{-1em}
-\[
-\begin{array}{rcl}
-  \ccscct(\irl{\lfunction{g}{x}{\expr}}) & = & \lfunction[\obj]{\obj{g}}{\obj{x}}{\obj{\lwrdoit{ON};}\ccscct(\irl{\expr})} \\
-  \ccscct(\irl{\lcall{g}{\expr}}) & = & \lcall[\obj]{\obj{g}}{\ccscct(\irl{\expr})\obj{; \lwrdoit{ON}}} 
-    % \\
-    % \ccscct(\irl{\lbinop{\expr[_{1}]}{\expr[_{2}]}}) & = \lbinop[\obj]{\ccscct{\irl{\expr[_{1}]}}}{\ccscct{\irl{\expr[_{2}]}}} 
-    % \\
-\end{array}
-\]
-% \vspace{-2em}
-%
-The context can overwrite the flag and exit the mode, but upon invoking a function that is part of the component, the flag is set again.
-Because of this, the corresponding cross-language trace relation $\sim^{\Lms}_{\Lscct}$, only relates events without secrets:%
-
-\begin{center} 
-  \typerulenolabel{xrel:scct:noleak}{}{\irl{\preevent;\comp}\sim^{\Lms}_{\Lscct}\obj{\preevent;\comp;\unlock}}
-\end{center}
-
-The compiler is secure w.r.t. \gls*{scct}: %, similarly proven as in \Cref{subsec:cs:tms}.
-
-\begin{theorem}[$\ccscct$ secure w.r.t. \gls*{scct}]\label{thm:ccscct:rtp:scct}
-  \small
-  $\rtpsim{\ccscct}{\scctsafe}{\sim^{\Lms}_{\Lscct}}$ % \Coqed
-\end{theorem}
-
-\subsection{Robust Speculative Safety Preservation}\label{subsec:cs:ss}
-
-This section defines the final compilation pass $\ccspec$, which ensures that $\Lscct$ programs, which are assumed to be \gls*{ss}, stay \gls*{ss} at $\Lspec$-level. 
-To do so, $\ccspec$ inserts a speculation barrier after branches, which is sufficient to harden the program against speculative leaks, since SPECTRE-PHT~\cite{kocher2019spectre} is the only speculative leak modeled in the semantics of $\Lspec$.
-
-\vspace{-1em}
-\[
-\begin{array}{cl}
-  &\ccspec{(\obj{\lifz{\expr[_0]}{\expr[_1]}{\expr[_2]}})} = 
-  \\
-  &\qquad\qquad \lifz[\ird]{\ccspec{\left(\obj{\expr[_0]}\right)}}{\ird{\lbarrier;}\ccspec{\left(\obj{\expr[_1]}\right)} \\&\qquad\qquad}{ \ird{\lbarrier;}\ccspec{\left(\obj{\expr[_2]}\right)}} 
-\end{array}
-\]
-% \vspace{-2em}
-%
-
-Clearly, the corresponding cross-language trace relation $\sim^{\Lscct}_{\Lspec}$ has only one non-trivial case: for branches, only relate them where speculation is blocked by a barrier:
-
-\begin{center}
-  \typerulenolabel{xrel:spec:if}{}{\obj{Branch\ n}\sim^{\Lscct}_{\Lspec}\ird{Branch\ n}\cdot\ird{Spec}\cdot\ird{Barrier}\cdot\ird{Rlb}}
-\end{center}
-
-The base-event relation above scales to full events by ensuring the missing annotations ($\obj{\comp;\securitytag{}}$ and $\ird{\comp;\securitytag{}}$) are the same.
-% 
-With this relation, we prove that $\ccspec$ is secure with respect to \gls*{ss}.
-\begin{theorem}[$\ccspec$ secure w.r.t. \gls*{ss}]\label{thm:ccspec:rtp:spec}
-  \small$\rtpsim{\ccspec}{\sssafe}{\sim^{\Lscct}_{\Lspec}}$ % \Coqed
-\end{theorem}
-
-\subsection{Robust Preservation of Memory Safety, Strict Cryptographic Constant Time, and Speculative Safety}
-
-Finally, this subsection combines all previous results into one compilation chain to get that it preserves full \gls*{specms}.
-Let $\ccspecms$ be the compiler that is the composition of $\cca$, $\ccb$, $\cccf$, $\ccdce$, $\ccscct$, and $\ccspec$. 
-Let $\sim^{\Ltms}_{\Lspec}$ be the composition of $\sim^{\Ltms}_{\Lms}$, $\sim^{\Lms}_{\Lscct}$, and $\sim^{\Lscct}_{\Lspec}$.
-Then, the following corollary holds.
-
-\begin{corollary}[$\ccspecms$ secure w.r.t. \gls*{specms}]\label{thm:ccall:rtp:specms}
-  $\;$
-
-  \begin{nscenter}
-    $\rtpsim{\cc{\Ltms}{\Lspec}}{\mssafe\cap\scctsafe\cap\sssafe}{\sim^{\Ltms}_{\Lspec}}$ % \Coqed
-  \end{nscenter}
-\end{corollary}
-
-As with \Cref{corr:ccab:rtp:ms}, it is important to ensure that the respective cross language trace relations are well-formed (\Cref{def:wfc:sig:tracerel}).
-It is already known that $\sim^{\Ltms}_{\Lms}$ is well-formed with respect to $\mssafe$ (\Cref{lem:wf:ltmsltrg}).
-Next in the chain is $\sim^{\Lms}_{\Lscct}$, which has to be well-formed w.r.t. $\scctsafe$.
-This lemma holds, since a trace that was $\scctsafe$ is $\scctsafe$ even after applying $\sim^{\Lms}_{\Lscct}$: the relation enforces that $\Lscct$ traces related to $\Lms$ traces have no leaks of secrets whatsoever.
-
-\begin{lemma}[$\sim^{\Lms}_{\Lscct}$ well-formed w.r.t. $\scctsafe$]\label{lem:wf:lsmslscct}
-$\;$
-
-  \begin{nscenter}
-  $\wfcsig{\sim^{\Lms}_{\Lscct}}{\scctsafe}$
-  \end{nscenter}
-\end{lemma}
-
-The last relation is $\sim^{\Lscct}_{\Lspec}$ which needs to be well-formed w.r.t. $\sssafe$.
-Similarly to the previous relation, this holds, since $\sim^{\Lscct}_{\Lspec}$ only relates $\Lspec$ traces, which do not have speculative leaks, with $\Lscct$ traces.
-
-\begin{lemma}[$\sim^{\Lscct}_{\Lspec}$ well-formed w.r.t. $\sssafe$]\label{lem:wf:lscctlspec}
-  $\wfcsig{\sim^{\Lscct}_{\Lspec}}{\sssafe}$
-\end{lemma}
-
-\section{Formal Insights}\label{sec:formalities}
-
-This section first discusses how to connect each language-specific security property to the general security properties presented in \Cref{sec:compprop} (\Cref{subsec:formalities:maps}), then it demonstrates that the security property resulting from the universal image projection is faithful to the original property (\Cref{subsec:formalities:props}).
-Finally, this section discusses why the order of compiler passes matters, and how does our framework help with identifying insecure compositions (\Cref{subsec:compatsecpasses}), and it gives additional technical insights on the secure compilation proofs (\Cref{subsec:seccompproofs}).
-
-\subsection{From Language Traces to General Ones}\label{subsec:formalities:maps}
-
-The previous theorems talk about preserving properties expressed in the trace model of the languages of each compiler.
-However, these trace models are not the same trace model we used to specify the properties of \Cref{sec:compprop} (indicated with \ev{L}), which serves as the ``ground truth'' for the meaning of our properties.
-% 
-To bridge this gap, the formal development requires specifying additional trace relations, from each of the language trace models to the \ev{L} one, that, for example, relate $\src{Get\ \loc\ n}$ and $\src{Set\ \loc\ n}$ to $\ev{Use\ \loc\ n}$ (and that induce a related universal image that we use in \Cref{subsec:formalities:props}).
-% 
-One key insight of these relations is that they all omit context-made actions for two reasons: (1) contexts (which are universally quantified) can trivially invalidate any property and (2) we are interested in the component upholding the properties.
-%
-
- % disregarded a glaring but minor technical issue: take $\tmssafe$ (\Cref{def:trace:tmsdef}) as an example.
-% It contains traces as specified in \Cref{subsec:basic:memsafety:tracemodel}.
-% However, these are different from $\Ltms$ traces (\Cref{subsec:ltms}) which means that \Cref{thm:wt:tms} is ``ill-typed''.
-% Similarly for the other theorems.
-% In the interest of reducing visual baggage, the notation left out yet another cross-language trace relation, translating language-specific traces into these non-language specific traces. 
-% In doing so, context-level actions are dropped and, e.g., both $\src{Get\ \loc\ n}$ and $\src{Set\ \loc\ n}$ are translated into $\ev{Use\ \loc\ n}$.
-
-\subsection{Security Properties and Their Meaning}\label{subsec:formalities:props}
-Each of the presented compilers use a cross-language trace relation, which is also used to translate the property from one language to the other one (via the existential or universal images).
-% 
-%Whenever such translations occur, it is important to check whether there has been a change of meaning.
-%Secure compiler engineers need to ensure that such translations of properties do not change meaning in a fundamental wrong way. 
-% lifting or lowering reasoning from target- to source-level or vice-versa,
-While the meaning of projected properties does change with a translation, the change should not allow for a flawed compilation pipeline.
-% 
-For example, we could be using a trace relation that translates a property in a language to a totally different one in another language.
-% 
-To raise the trust into the translation of properties, \Cref{thm:prop-rel-corr} states (in a general fashion) that each security property is faithfully translated using the universal image according to the cross-language trace relation induced by the compiler.
-
-\begin{theorem}[Properties Relation Correctness]\label{thm:prop-rel-corr}
-  \begin{align*}
-    \forall& \pi \in \{\tmssafe,\smssafe,\scctsafe,\sssafe\}, 
-    \\
-    \text{ for each }& \text{pair of languages } \src{L} \text{ and } \trg{L} \text{ used by the compilers},
-    \\
-    \text{ if }& 
-    % \tau_{\sim^{\trg{L}}_{\ev{L}}}(\pi) \sim^{\trg{L}}_{\ev{L}} \pi
-    \sigma_{\sim^{\trg{L}}_{\ev{L}}}(\pi) \sim^{\trg{L}}_{\ev{L}} \pi
-    \text{ and } 
-    % \sigma_{\sim^{\src{L}}_{\trg{L}}}(\tau_{\sim^{\trg{L}}_{\ev{L}}}(\pi)) \sim^{\src{L}}_{\trg{L}} \tau_{\sim^{\trg{L}}_{\ev{L}}}(\pi)
-    \sigma_{\sim^{\src{L}}_{\trg{L}}}(\sigma_{\sim^{\trg{L}}_{\ev{L}}}(\pi)) \sim^{\src{L}}_{\trg{L}} \sigma_{\sim^{\trg{L}}_{\ev{L}}}(\pi)
-    \\
-    \text{ then }& 
-    % \tau_{\sim^{\src{L}}_{\ev{L}}}(\pi) = \sigma_{\sim^{\src{L}}_{\trg{L}}}(\tau_{\sim^{\trg{L}}_{\ev{L}}}(\pi)) 
-    \sigma_{\sim^{\src{L}}_{\trg{L}}}(\sigma_{\sim^{\trg{L}}_{\ev{L}}}(\pi)) \sim^{\src{L}}_{\ev{L}} \pi
-  \end{align*}
-\end{theorem}
-The complexity of this theorem is that the relation in the conclusion cannot be obtained by composing the two relations in the premises.
-% 
-%% MK: the following does not say anything about the theorem, only about the syntactic compiler
-%What happens in practice is that compiler engineers are given the relations from each language to \ev{L} and they need to build a compiler that induces a relation $\sim^{\src{L}}_{\trg{L}}$ that respects this theorem.
-%
-
-We now informally argue why this theorem holds for the composition of all considered properties from \Cref{sec:compprop}.
-
-\paragraph{$\sigma_{\sim^{\Ltms}_{\Lms}}\left(\tmssafe\cap\smssafe\cap\scctsafe\cap\sssafe\right)$}
-
-For the four properties considered here, the trace models of $\Ltms$, $\Ltrg$, and $\Lms$ do not consider actions related to $\scctsafe$ and $\sssafe$, so these two properties are trivially translated correctly.
-% 
-We now discuss the remaining $\tmssafe$ and $\smssafe$ in the form of their intersection $\mssafe$.
-
-Recall that $\sim^{\Ltms}_{\Lms}$ is the composition of $\sim^{\Ltms}_{\Ltrg}$ and $\sim^{\Ltms}_{\Lms}$.
-% 
-Since $\sim^{\Ltms}_{\Ltrg}$ is an equality, this relation trivially preserves the meaning of translated properties: related traces are identical!
-
-Finally, let us consider a trace $\irl{\trace}\in\mssafe$ and understand what is that is related to via $\sim^{\Ltrg}_{\Lms}$.
-% 
-{\em All} traces $\trg{\trace}$ with $\trg{\trace}\sim^{\Ltrg}_{\Lms}\irl{\trace}$ are identical to $\irl{\trace}$ except for get and set actions, which require for in-bound accesses (as stated in \Cref{subsec:cs:ms}): this clearly respect $\mssafe$.
-%
-
-% In both cases, the translated event does not represent an out-of-bounds memory access.
-
-% \footnote{Both $\sim^{\Ltms}_{\Ltrg}$ and $\sim^{\Ltrg}_{\Lms}$ filter context actions.}.
-
-% Now, starting from $\mssafe$ expressed in $\Lms$, we check that its translation in $\Ltms$ is still $\mssafe$.
-% % 
-% We can see that the property resulting from the projection of $\Lms$ traces back to $\Ltms$ faithfully models $\mssafe$, which translates to both $\scctsafe$ and $\sssafe$, 
-% \MPin{ equiv  +  other rel only adds traces that fail }
-% since these properties trivially hold for $\Ltms$, $\Ltrg$, and $\Lms$ traces.
-
-\paragraph{$\sigma_{\sim^{\Lms}_{\Lscct}}\left(\tmssafe\cap\smssafe\cap\scctsafe\cap\sssafe\right)$}
-
-As before, the trace models of $\Lscct$ and $\Lms$ cannot express $\sssafe$, so that is trivially translated correctly.
-% 
-Also, the trace model of $\Lscct$ extends the one of $\Lms$ with respect to $\tmssafe$ and $\smssafe$ events, so the translation argument regarding those two properties is the same as before.
-% 
-Thus, we need to reason about whether $\scctsafe$ is translated correctly.
-
-By definition, $\sim^{\Lms}_{\Lscct}$ only relates $\obj{\unlock}$ events, which is also the same relation induced by $\sim_{\ctsafe}$ in \Cref{sec:msscct-rel}.
-% 
-This ensures that composing the relations only relates $\unlock$ events, and thus the property is translated correctly.
-
-% Consider a trace $\obj{\trace}\in\scctsafe$ and all $\irl{\trace}$ such that $\irl{\trace}\sim^{\Lms}_{\Lscct}\obj{\trace}$. 
-% In turn, $\obj{\trace}$ cannot contain any leaks of secret values, so it cannot violate $\scctsafe$.
-% Toggling data-independent timing mode has no impact on $\mssafe$ or $\sssafe$.
-
-\paragraph{$\sigma_{\sim^{\Lscct}_{\Lspec}}\left(\tmssafe\cap\smssafe\cap\scctsafe\cap\sssafe\right)$}
-The trace model of $\Lspec$ extends the one of $\Lscct$ with respect to $\tmssafe$, $\smssafe$, and $\scctsafe$, so the translation argument for those three properties is the same as before.
-% 
-Concerning $\sssafe$, $\sim^{\Lscct}_{\Lspec}$ relates $\Lspec$ speculation traces with $\Lscct$ branches with the $\lock_{\text{NONE}}$ tag, so the property is translated correctly, according to the relation defined in \Cref{sec:spec-ms-rel}.
- % introducing $\ird{barrier}$ never violates $\sssafe$ and has no effect on $\mssafe$ or $\scctsafe$.
-
-%\MPin{
-%	can this be made formal?
-%	what's the point?
-%}
-
-\subsection{Compatibility of Secure Compiler Passes}\label{subsec:compatsecpasses}
-
-% \Cref{thm:rtpsim:sig} is applicable, up to the well-formedness of $\sim_1$ w.r.t. $\trg{\class[_2]}$.
-Consider applying $\ccspec$ first and then $\ccb$ (albeit currently syntactically impossible).
-% , let us pretend that it is possible by just extending $\ccb$ for the constructs present in $\Lscct$ that are not present in $\Ltrg$.
-% Clearly, as evidenced earlier (\Cref{subsec:formalities:props}), $\ccb\cdot\ccspec$ is an acceptable composition that yields an interesting class of security properties, i.e., $\sigma_{\sim_{\irl{ms}}\bullet\sim_{\ird{\mathghost}}}\left(\mssafe\cap\scctsafe\cap\sssafe\right)$.
-In this case, $\ccb$ would insert new branches into the code that are not protected by a speculation barrier! 
-This concern is reflected in the proofs that establish that the security class resulting of the composition of the trace relations is meaningful.
-% , much similar to ensuring that a definition models what the formal methods expert intends to model. 
-For the $\ccspec\cdot\ccb$ case, the class is $\sigma_{\sim_{\ird{\mathghost}}\bullet\sim_{\irl{ms}}}\left(\mssafe\cap\scctsafe\cap\sssafe\right)$.
-Since $\ev{Use\ \loc\ n}\sim_{\irl{ms}}\irl{Branch\ n}\cdot\irl{Spec}\cdots$, where $\cdots$ does not contain a $\ev{{Barrier}}$ event, the resulting class is \emph{not} the original $\sssafe$ that is intended and it would break the corresponding \Cref{thm:prop-rel-corr}.
-% Because of this, it is crucial to analyse the precise shape of the class of properties after mapping a target-level property to a source-level property\footnote{Dito for mapping a source-level property to a target-level property.}.
-
-However, the composition is still technically possible and it is the job of the compiler engineers to ensure that the secure compilation pipeline happens in an order that ensures that the mapped security property is the intended one.
-
-\subsection{Secure Compilation Proofs}\label{subsec:seccompproofs}
-
-Our secure compilation proofs rely on backtranslations~\cite{abate2019jour,patrignani2021rsc}, which let one construct a source context starting from either target traces (aka trace-based backtranslations) or target contexts (aka context-based backtranslations).
-These backtranslations also require setting up cross-language relations between various language elements such as expressions and program states, so we leave these details for the technical reports.
-% , the usual strategy is to establish a backwards simulation between source components and their compiled target counterpart. 
-% For the universally quantified target context, a technique known as backtranslation is well-established, which constructs a source-level context either by backtranslating the target-level context, amounting to a ``decompilation'', or by constructing it using the target-level trace, so that the source-context emits parts of the whole execution trace that are related to the parts emitted by the target-level context.
-All backtranslations in the case-study are trace based except for those required by $\ccdce$ and $\cccf$, which are context-based (and they are an identity function). 
-% Since these optimisations have no significant semantic effect and source- and target-language are equal, the backtranslation amounts to an identity translation.
-
-\section{Related Work\pages{2}}\label{sec:relwork}
-
-This section discusses robust compilation, other secure compilation criteria and work related to the properties preserved in the case study.
-% (\Cref{subsec:relw:seccomprtp}) 
-% (\Cref{subsec:relw:seccompcrit}).
-% Since the case study of \Cref{sec:casestud:defs,sec:casestud:rtp} implements measures for preserving \gls*{ms}, \gls*{cct}, and \gls*{ss}, 
-% Then, this section discusses work related 
-% relevant related work as well 
-% (\Cref{subsec:relw:msmechs,subsec:relw:cctmechs,subsec:relw:ssmechs}).
-
-\paragraph*{Secure Compilation as Robust Preservation}\label{subsec:relw:seccomprtp}
-
-The robust preservation of properties as a compiler-level criterion has been analysed extensively~\cite{abate2019jour,patrignani2021rsc,abate2021extacc,patrignani2019survey} and thus we build on that framework.
-No existing work is concerned with composing secure compilers, however, existing work~\cite{abate2021extacc} sketches composition of trace-relating compiler correctness in a similar way to what has been presented here.
-%Parts of these works consider languages with different trace models and our technical setup can handle this. 
-The work relating robust preservation with universal composability~\cite{patrignani2022universal} is closest to what this paper presents.
-The authors demonstrate a similar compositionality theorem to what is presented here (\Cref{sec:sequential}) but use it in the context of protocols.
-The work does not consider the generality to support different trace models or composition of compilers which robustly preserve different classes.
-% The authors demonstrate a similar compositionality theorem to what is presented here (\Cref{sec:sequential}) as well as in an earlier version of this work~\cite{kruse2022csc}.
-% However, they do not demonstrate the scalability of the approach by means of a case study.
-
-%\paragraph*{Composition of Secure Compilers}\label{subsec:relw:seccompcomp}
-
-There is work on lifting exploits for single compilers to the whole chain~\cite{paykin2019weird}.
-%Our framework does not consider exploits.
-%\MP{ a bit dry and short}
-While that work considers {\em in}secure compilation and composition thereof in terms of exploits, the composition they are interested in allows to lift an exploit for one compiler pass to the whole compilation chain.
-Our framework takes the opposite direction and provides compositionality results for secure compilers.
-%
-%
-%To the best of our knowledge, there does not exist published work discussing the compositionality of the robust preservation criterion, the gold-standard of secure compilation.
-
-\paragraph*{Other Secure Compilation Criteria}\label{subsec:relw:seccompcrit}
-
-While this paper focuses on the robust preservation framework~\cite{abate2019jour}, other secure compilation criteria exist.
-The survey on formal approaches to secure compilation~\cite{patrignani2019survey} discusses a broad spectrum already, while this section presents a very high-level overview.
-Fully abstract compilation~\cite{abadi1999fullabstraction} states that a compiler should preserve and reflect observational equivalence between source and target programs.
-Abate \emph{et al.}~\cite{abate2021faandrc} showed that fully abstract compilers robustly preserve program properties that are either trivial or meaningless.
-As a mitigation for this, the authors presented a categorical approach based on maps of distributive laws~\cite{watanabe2002modl}, which they call many maps of distributive laws.
-Maps of distributive laws have been investigated before as a possible secure compilation criterion~\cite{tsampas2020catsc}.
-Other approaches are extensions of the compiler correctness criterion as discussed in other work~\cite{patterson2019next700} or the introduction of opaque observations~\cite{vu2021reconciling} to reconcile compiler optimisations with security.
-Note that this work also presents secure compilers that are optimising, but contrary to the other~\cite{vu2021reconciling}, provides a formal account of these in the robust preservation framework.
-% Lastly, the authors of this paper have presented ongoing work~\cite{patrignani2023blame} on a weaker robust preservation criteria based on the concept blame.
-
-\paragraph*{Memory Safety Mechanisms}\label{subsec:relw:msmechs}
-
-Different mechanisms for enforcing memory safety exist that also consider the secure compilation domain, i.e., have an active attacker model.
-For example, the ``pointers as capabilities'' principle represents pointers as machine-level capabilities~\cite{korashy2021capableptrs}, which behave in a similar fashion to capabilities by means of linear typing~\cite{morrisett2005L3}.
-The approach of this paper also uses linear typing, but differs from $L^{3}$~\cite{morrisett2005L3} in the way that functions are not first-class.
-Moreover, this paper considers an active attacker, while the work on $L^{3}$ only discusses whole programs and, thus, has no active attacker model.
-The instrumentation to ensure memory safety that this paper presents is inspired by Softbounds~\cite{nagarakatte2009soft}.
-That work inserts bounds-checks in front of pointer-dereferences and, for this to work, inserts meta-data information on pointer creation.
-Softbounds also works in a more advanced setting with structured fields accesses and also introduces a table-lookup for pointers that are stored in memory.
-This paper only considers arrays of primitive data, i.e., there are no pointers to pointers or structures.
-Several other approaches to memory-safety exist in literature, specifically as compiler instrumentations~\cite{akritidis2009baggy,younan2010paricheck,jung2021pico,shankaranarayana2023tailcheck,dhumbumroong2020boundwarden,nam2019framer,zhou2023fatptrs}, hardware-extensions~\cite{kwon2013lowfat,saileshwar2022heapcheck,chen2023flexpointer,kim2023whistle}, or programming language extensions~\cite{elliott2018checkedc,li2022formalcheckedc,jim2002cyclone,elliott2015guilt,west2005cuckoo,weis2019fyr,benoit2019uniqueness}.
-What differentiates this work from them is that this work uses known, compiler-based approaches to ensure memory-safety as a means to investigate secure compiler compositions.
-This paper does not provide efficient memory-safety, but serves as a theoretical foundation for the secure compilation domain.
-
-To extend the languages in this paper with a less restricted form of pointer arithmetic, the region colouring memory safety monitor presented in earlier work~\cite{michael2023mswasm} can be used.
-The work presenting this monitor provides an approach for the robust preservation of memory safety compiling from C to WASM.
-However, they do not discuss composition of secure compilers but rather investigate an instance of a secure compiler.
-
-\paragraph*{Cryptographic Constant Time Mechanisms}\label{subsec:relw:cctmechs}
-
-The approach to preserving cryptographic constant time in this paper is high-level, where a programming language exposes a way to switch the semantics to a data (operand) independent timing mode.
-Since identifiers in $\Lscct$ are annotated with a secrecy tag, this approach is similar to others with information flow control.
-For example, Vale~\cite{bond2017vale} uses Dafny to ensure constant-time assembly code, while Jasmin~\cite{almeida2017jasmin} makes use of the Coq proof assistant to reject non-constant-time programs.
-CT-Wasm~\cite{watt2019ctwasm} enforces constant-timeness by means of a type system.
-Different to the approach of this paper, these approaches necessitate that the programmer writes \gls*{cct} code.
-An approach to allow programmers to write more high-level code is CryptOpt~\cite{kuepper2023cryptopt}, which generates efficient target-code by means of a randomised search.
-This paper abstracts over concrete mitigation strategies and simply assumes that there is a flag to switch to a cryptographic-constant time execution mode.
-This can be realised by employing the FaCT~\cite{cauligi2019fact} compiler, which translates common non-constant time code patterns to be constant-time, and the data (object) independent timing execution mode of modern processors.
-
-\paragraph*{Speculation Safety Mechanisms}\label{subsec:relw:ssmechs}
-
-This paper uses a taint-tracking mechanism inspired by existing work~\cite{guarnieri2018spectector,fabian2022automatic}.
-These taints are used to express absence of any speculative leaks in \gls*{ss}~\cite{guarnieri2018spectector}. 
-The semantics of $\Lspec$ hardcodes the kind of speculative leaks to just SPECTRE-PHT~\cite{kocher2019spectre}, but future work could use semantics composition~\cite{fabian2022automatic} to support more variants.
-Note that our framework composes compilers and not semantics.
-
-%This work has taken inspiration in the sematics of $\Lspec$ and in \Cref{def:trace:ss} from recent formalisations~\cite{guarnieri2018spectector,fabian2022automatic}.
-% \MPin{
-%  we take SS from exorcising.
-%  we do not take SNI from spectector because it is a hypersafety.
-%  other work defined similar props. 
-% }
-% A subsequent line of work~\cite{fabian2022automatic} shows how to modularly define and compose formal semantics that models different speculative execution variants. 
-% Recent work~\cite{fabian2024lift} demonstrates how to lift existing secure compilation proofs for security properties related to speculative execution to more powerful properties, i.e., properties defined for languages with a trace model that enables more vulnerabilities, such as combinations of different speculative execution exploits.
-% To this end, they define a lifting framework and demonstrate how a secure compilation proof for a weaker property can be lifted to a stronger one. 
-% Contrary to the compositionality framework presented in this paper, these works are only concerned with composing semantics of the respective programming languages, while in this paper, the secure compilers themselves are composed.
-
 \section{Conclusion\pages{$\sfrac{1}{2}$}}\label{sec:concl}
@@
 % For future work, it is interesting to look at more sophisticated optimisation passes that, e.g., reorder events that appear on traces.
-
-\newpage
-
-\bibliographystyle{IEEEtranS}
-\bibliography{main}
-
-%%
-%% If your work has an appendix, this is the place to put it.
-\appendix
-\subsection{Existential Image}\label{subsec:extimg}
-
-\begin{definition}[Existential Image]
-% [Existential and Universal Image]
-\label{def:existential:img}\label{def:tau}
-  \[
-    \tau_\sim\left(\src{\pi}\right) := 
-      \left\{ 
-        \trg{\trace} \mid \exists \src{\trace}\ldotp \src{\trace}\sim\trg{\trace}, \text{ and } \src{\trace}\in\src{\pi} 
-      \right\}
-  \]
-  %\[ 
-  %  \sigma_\sim\left(\trg{\pi}\right) := 
-  %    \left\{ 
-  %      \src{\trace} \mid \forall \trg{\trace}\ldotp \text{if }\src{\trace}\sim\trg{\trace}, \text{ then } \trg{\trace}\in\trg{\pi} 
-  %    \right\}
-  %\]
-\end{definition}
-
-\begin{definition}[Robust Preservation with $\tau_\sim$]\label{def:rtp:tau}
-  % $\;$\\
-  % \vspace{-1em}
-  % \begin{nscenter}
-  % Compiler $\cc{\src{L}}{\trg{L}}$ robustly preserves $\class$, 
-  {$\rtptau{\cc{\src{L}}{\trg{L}}}{\src{\class}}{\sim}$}
-  %, iff 
-  $\isdef$
-    {$\forall \src{\pi}\in\src{\class}, \src{p}\in\src{L},$} if {$\rsat{\src{\progvar}}{\src{\pi}}$}, then {$\rsat{\cc{\src{L}}{\trg{L}}\left(\src{p}\right)}{\tau_\sim\left(\src{\pi}\right)}$}.
-  % \end{nscenter}
-\end{definition}
-
-\begin{theorem}[Composition of Secure Compilers w.r.t. $\tau$]\label{thm:rtpsim:tau}
-  $\;$
-
-  If {$\rtptau{\cc{\src{L}}{\trg{L}}}{\src{\class[_{1}]}}{\sim_1}$}, {$\rtptau{\cc{\trg{L}}{\obj{L}}}{\tilde{\tau}_{\sim_1}\left(\src{\class[_2]}\right)}{\sim_2}$}, and {$\wfctau{\sim_1}{\src{\class[_2]}}$}, \\ then {$\rtptau{\cc{\src{L}}{\trg{L}}\circ\cc{\trg{L}}{\obj{L}}}{\src{\class[_{1}]}\cap\src{\class[_{2}]}}{\sim_1\bullet\sim_2}$}.
-\Coqed
-\end{theorem}
-
-\begin{corollary}[Swapping Secure Compiler Passes]\label{corr:swappable:tau}
-  $\;$
-
-  If {$\rtptau{\cc[_{1}]{\trg{L}}{\trg{L}}}{\src{\class[_{1}]}}{\sim_1}$ and $\rtptau{\cc[_{2}]{\trg{L}}{\trg{L}}}{\src{\class[_{2}]}}{\sim_2}$}, %
-  {$\wfctau{\sim_1}{\src{\class[_2]}}$ and $\wfctau{\sim_2}{\src{\class[_1]}}$}, %
-  and {$\tilde{\tau}_{\sim_1}\left(\src{\class[_2]}\right)=\src{\class[_2]}$ as well as $\tilde{\tau}_{\sim_2}\left(\src{\class[_1]}\right)=\src{\class[_1]}$},
-  then {$\rtptau{\cc[_{1}]{\trg{L}}{\trg{L}}\circ\cc[_{2}]{\trg{L}}{\trg{L}}}{\src{\class[_{1}]}\cap\src{\class[_{2}]}}{\sim_1\circ\sim_2}$ and $\rtptau{\cc[_{2}]{\trg{L}}{\trg{L}}\circ\cc[_{1}]{\trg{L}}{\trg{L}}}{\src{\class[_{2}]}\cap\src{\class[_{1}]}}{\sim_2\circ\sim_1}$}.
-\Coqed
-\end{corollary}
-
-\subsection{Secure Upper and Lower Composition}\label{sec:other-compos}
-Besides sequential composition, there are two other compositions, namely an {\em upper}, i.e., a compiler that takes multiple inputs and yields one output, and a {\em lower} composition, i.e., a compiler that takes one input and yields multiple outputs.
-We {define the upper composition $\cc{\src{L}+\obj{L}}{\trg{L}}$} as follows:
-Given a program \texttt{p}, its compiled counterpart is obtained by {plugging \texttt{p} into $\cc{\src{L}}{\trg{L}}$ if $\texttt{p}\in\src{L}$} or by {plugging \texttt{p} into $\cc{\obj{L}}{\trg{L}}$ if $\texttt{p}\in\obj{L}$}.
-\begin{definition}[Upper Composition]
-  \[
-    \text{{$\cc{\src{L}+\obj{L}}{\trg{L}}$}}\isdef
-  % Given \texttt{p}, yield
-  \lambda \texttt{p}\ldotp
-  \left\{\mbox{\begin{tabular}{c}
-    {{if $\texttt{p}\in\src{L}$, then $\cc{\src{L}}{\trg{L}}(\texttt{p})$}} \\
-    \mbox{{if $\texttt{p}\in\obj{L}$, then $\cc{\obj{L}}{\trg{L}}(\texttt{p})$}} \\
-  \end{tabular}}\right. 
-  \]
-\end{definition}
-
-Examples of this are present in industry:
-Consider the Java Virtual Machine bytecode $\trg{JVM BC}$, which is a popular target for programming language designers due to its high performance and relevance in industry.
-Compilers for several programming languages have it as their target language, some popular instances are $\src{Java}$ and $\obj{Kotlin}$.
-Technically speaking, they both compile to class files and $\obj{Kotlin}$ objects are considered to be the same as $\src{Java}$ objects at that point.
-Both languages can be used at the same time in one project~\cite{androidstudio}.
-A compiler that accepts both $\src{Java}$ and $\obj{Kotlin}$ code translating to the same target language or intermediate representation performs a kind of {\em upper} composition.
-Now, the following theorem tells us what happens if these are secure:
-Given {$\cc{\src{L}}{\trg{L}}$ robustly preserves $\class[_{1}]$} and {$\cc{\obj{L}}{\trg{L}}$ robustly preserves $\class[_{2}]$}, it follows that {their upper composition $\cc{\src{L}+\obj{L}}{\trg{L}}$ robustly preserves the intersection of classes $\class[_{1}]$ and $\class[_{2}]$}.
-
-\begin{theorem}[Upper Composition of Secure Compilers]\label{thm:urtp}
-  $\;$
-
-  If {$\rtp{\cc{\src{L}}{\trg{L}}}{\class[_{1}]}$} and {$\rtp{\cc{\obj{L}}{\trg{L}}}{\class[_{2}]}$}, then {$\rtp{\cc{\src{L}+\obj{L}}{\trg{L}}}{\class[_{1}]\cap\class[_{2}]}$}. %\Coqed
-\end{theorem}
-
-Dually, the {\em lower} composition is concerned about compilers that accept the same source but yield different target languages. %and the same theoretical results apply.
-{Define the lower composition $\cc{\src{L}}{\trg{L}+\obj{L}}$} as follows:
-Given a program $\src{p}$, its compiled counterpart is obtained by {plugging $\src{p}$ into $\cc{\src{L}}{\trg{L}}$} or by {plugging \texttt{p} into $\cc{\src{L}}{\obj{L}}$}, respectively, {based on the internal decision}.
-\begin{definition}[Lower Composition]
-  \[
-    \text{{$\cc{\src{L}}{\trg{L}+\obj{L}}$}}\isdef
-  % Given $L$ and $\src{p}$, yield
-  \lambda \src{p}, L\ldotp
-  \left\{\mbox{\begin{tabular}{c}
-    {{if $L=\trg{L}$, then} {$\cc{\src{L}}{\trg{L}}(\src{p})$}} \\
-    \mbox{{if $L=\obj{L}$, then} {$\cc{\src{L}}{\obj{L}}(\src{p})$}} \\
-  \end{tabular}}\right.\]
-\end{definition}
-
- Consider two compilers both accepting $\src{LLVM IR}$~\cite{lattner2004llvm} and one of them emits $\trg{x86\_64}$, while the other emits $\obj{ARMv8}$.
- It is intuitive that they are in some sense composed in the LLVM framework, but the decision of when to use one over the other is inherently {\em internal} to the formalisation effort of this kind of composition.
- For example, the user of this compiler provides an explicit flag that instructs to emit $\trg{x86\_64}$ or the framework itself detects the target platform via heuristics, such as supported instructions.
-
- The following theorem demonstrates what happens if the involved compilers are secure:
-Given {$\cc{\src{L}}{\trg{L}}$ robustly preserves $\class[_{1}]$} and {$\cc{\src{L}}{\obj{L}}$ robustly preserves $\class[_{2}]$}, it follows that {their lower composition $\cc{\src{L}}{\trg{L}+\obj{L}}$ robustly preserves the intersection of classes $\class[_{1}]$ and $\class[_{2}]$}.
-
-\begin{theorem}[Lower Composition of Secure Compilers]\label{thm:lrtp}
-  $\;$
-
-  If {$\rtp{\cc{\src{L}}{\trg{L}}}{\class[_{1}]}$} and {$\rtp{\cc{\obj{L}}{\trg{L}}}{\class[_{2}]}$}, then {$\rtp{\cc{\src{L}}{\trg{L}+\obj{L}}}{\class[_{1}]\cap\class[_{2}]}$}. % \Coqed
-\end{theorem}
-
-Either way, the theoretical results suggest that it is possible to always find a ``most-general'', secure compiler, given two secure compilers, that robustly preserves the least-upper bound of the classes involved in their compilation process.
-
-%Future work intends to use upper and lower composition as means to build a secure compiler incrementally
-%This follows current non-formal-methods based compiler engineering, such as LLVM~\cite{lattner2004llvm}, where languages such as \texttt{C++} and \texttt{Fortran} translate into LLVM IR and languages such as \texttt{x86} or \texttt{ARMv8} assembly variants translate out of LLVM IR.
-
-% no coq, because above definitions are non-formal. 
-%   How would you precisely define the lower compo compiler, given two compilers?
-
-%\section{Cryptographic Constant Time}\label{sec:cct}
-
-\end{document}
-\endinput

### 2025-02-06 icfp/sections/intro.tex
Intro

@@
 Provenance of pointers is also crucial information, together with bounds, for spatial memory safety, which ensures that all accesses are within bounds of an allocated object.
+There is a large body of prior work modifying existing memory-unsafe languages~\cite{} or providing compiler-level enforcements~\cite{} for both temporal and spatial memory safety.
+Moreover, modern processors can speculatively load memory which has become the source of bugs not fixable by temporal or spatial memory safety.
+Here, other mechanisms~\cite{} are necessary in order to enforce privacy of data.
 

### 2025-02-10 icfp/sections/intro.tex
Intro

@@
 
+  % Context
   As of today, there exist a number of programming languages that enforce security-relevant properties.
@@
   Temporal memory safety is the absence of use-after-free and use-after-reallocation bugs and necessitates tracking of pointer provenance\footnote{Due to potential use-after-reallocation bugs.}.
-%TODO: example?
-Provenance of pointers is also crucial information, together with bounds, for spatial memory safety, which ensures that all accesses are within bounds of an allocated object.
+  Provenance of pointers is also crucial information, together with bounds, for spatial memory safety, which ensures that all accesses are within an allocated object.
   There is a large body of prior work modifying existing memory-unsafe languages~\cite{} or providing compiler-level enforcements~\cite{} for both temporal and spatial memory safety.
-Moreover, modern processors can speculatively load memory which has become the source of bugs not fixable by temporal or spatial memory safety.
+  However, there is more to memory-safety than just temporal and spatial.
+  For example, processors of the current age can speculatively load memory and this can be exploited with timing attacks to obtain secrets from ,,memory-safe'' applications~\cite{}.
   Here, other mechanisms~\cite{} are necessary in order to enforce privacy of data.
@@
   Compilers may fail to preserve these enforcements, even when proven correct or thoroughly tested~\cite{}.
-While properties can be enforced at source-level by means of static analyses, after compiling to languages without such abstractions, the properties may break.
+  While properties can be enforced at source-level by means of, e.g., static analyses, after compiling to languages without such abstractions, the properties may break.
   For example, RISC-V, a possible target-language of the most commonly used Rust compiler, provides neither memory nor type safety guarantees.
@@
   Unfortunately, software is seldolmy developed in isolation and programs may be linked with target-level libraries after compilation.
-These libraries can be compiled with different compilers or may even be hand-written, so the correctness-guarantees provided by one compiler may not hold.
+  These libraries can be compiled with different compilers or may even be hand-written.
+  In turn, the correctness-guarantees provided by one compiler do not extend to this setting.
 
@@
 
+  % Gap
   For the context of this paper, we are interested in formally verified compilers.
@@
   A practical compiler consists of several different compilation passes, which itself can be seen as compilers that translate from one intermediate representation into another.
-The CompCert project~\cite{}, while having more restrictions on target-level contexts, attests the herculian effort necessary to prove correctness for realistic compilers with several different passes.
+  %%MK: the following is not necessary, i think
+  %The CompCert project~\cite{}, while having more restrictions on target-level contexts, attests the herculian effort necessary to prove correctness for realistic compilers with several different passes.
   Moreover, for optimising passes, it can be beneficial to swap the order or iterate them until a fixpoint is reached, but this necessitates modifications to a top-level meaning-preservation proof of a compiler.
@@
 
-Our aim with this paper is to lay the groundwork for secure compiler composition:
+  % Innovation
+  Our aim with this paper is to lay the groundwork for secure compiler composition.
+  To this end, we make the following contributions.
   \begin{itemize}
@@
       Intuitively, we prove that two compilers that robustly preserve two (possibly distinct) properties compose into a compiler that robustly preserves the intersection of the properties.
-    We show special-cases, such as swapping secure optimisation passes or composition of compilers robustly preserving different properties. 
-    Moreover, we discuss how robust preservation relates to vertical compositionality and source-independent linking.
+      We show special-cases, such as swapping secure optimisation passes or composition of compilers robustly preserving different properties, and we discuss how robust preservation relates to vertical compositionality and source-independent linking.
     \item
-    There is a mechanization of our core-theory in the Coq\footnote{To be renamed into Rocq.} proof assistant and we mark these instances with (\CoqSymbol) throughout the paper.
+      We mechanized our core-theory in the Coq\footnote{To be renamed into Rocq.} proof assistant and we mark these instances with (\CoqSymbol) throughout the paper.
     \item 
@@
 
-Pen-and-paper proofs as well as the mechanization are available as supplementary material.
+  Pen-and-paper proofs as well as the mechanization are available as supplementary material~\cite{}.

### 2025-02-12 icfp/sections/bg/sat.tex
Rewrite first part of 2.1

@@
 
-This paper employs the security model where programs are written in a language whose semantics emits events $\event$.
-Events include security-relevant actions (e.g., reading from and writing to memory, as detailed in \Cref{sec:compprop}) and the unobservable event $\emptyevent$.
-As programs execute, their emitted events are concatenated in traces $\trace$, which serve as the description of the behaviour of a program.%
-\footnote{
+Running a (whole) program yields a trace ($\trace$), i.e., a potentially infinite sequence\footnote{
 Throughout the paper, sequences are indicated with an overbar (i.e., $\trace$), empty sequences with $\hole{\cdot}$, and concatenation of sequences $\trace[_{1}],\trace[_{2}]$ as $\trace[_{1}]\cdot\trace[_{2}]$.
 Prepending element $\event$ to a sequence $\trace$ uses the same notation: $\event\cdot\trace$.
-}
-
-Properties $\pi$ are sets of traces of admissible program behaviours, ascribing what said property considers valid.
+} of events ($\event$).
+Events include security-relevant actions (e.g., reading from and writing to memory, as detailed in \Cref{sec:compprop}), the unobservable event ($\emptyevent$), and a crash ($\lightning$).
+For a given program, we refer to the set of traces as its behaviour. 
+Moreover, sets of traces are called properties ($\pi$) which describe admissible program behaviours. % ascribing what said property considers valid.
 % 
@@
 % 
-The compositionality framework (\Cref{sec:sequential}) presented in this paper considers arbitrary classes, while the case-study (\Cref{sec:casestud:rtp}) fixes them to concrete instances of safety properties, since it is decidable whether a trace satisfies a safety property with just a finite trace (i.e., a \emph{prefix}).
-As an example, consider the trace:
-$$\ev{Dealloc\ \loc}\cdot\ev{Read\ \loc\ 1729}\cdot\dots$$
-which describes the interaction with a memory where the deallocation of an address $\loc$ precedes a read (of some value $1729$) at that address in memory.
+%The compositionality framework (\Cref{sec:sequential}) presented in this paper considers arbitrary classes, while the case-study (\Cref{sec:casestud:rtp}) fixes them to concrete instances of safety properties, since it is decidable whether a trace satisfies a safety property with just a finite trace (i.e., a \emph{prefix}).
+
+As an example, consider the trace $\ev{Dealloc\ \loc}\cdot\ev{Read\ \loc\ 1729}\cdot\dots$ which describes the interaction with memory where the deallocation of an address ($\loc$) precedes a read (of some value $1729$) at that address in memory.
 % 
-This program behaviour is insecure w.r.t.
-a canonical notion of (temporal) memory safety dictating no use-after-frees of pointers~\cite{nagarakatte2010cets,azevedo2018meaningsofms}, because it reads from a memory location that was freed already.
-The previous finite trace prefix is enough to decide that the trace does not satisfy \gls*{tms} and there is no way to append events to this prefix which would result in the trace being admissible.
+This behaviour is insecure w.r.t.
+the canonical notion of \gls*{tms} dictating no use-after-frees~\cite{nagarakatte2010cets,azevedo2018meaningsofms}, because it reads from a memory location that was freed already.
+The first two events in the shown trace are enough to decide that the trace does not satisfy \gls*{tms}, since there is no way to append events to this prefix which would result in the trace being admissible for \gls*{tms}.
 
-In the following, the execution of a whole program $\wholeprogvar$ that terminates in state $\runtimetermvar$ according to the language semantics and produces trace $\trace$ is written as $\progstepto{\wholeprogvar}{\runtimetermvar}{\trace}$.
-With this, we defined property satisfaction as follows:
-\bul{a whole program $\wholeprogvar$ satisfies a property $\pi$} iff \iul{$\wholeprogvar$ yields a trace $\trace$} such that \oul{$\trace$ satisfies $\pi$}.
+%In the following, the execution of a whole program $\wholeprogvar$ that terminates in state $\runtimetermvar$ according to the language semantics and produces trace $\trace$ is written as $\progstepto{\wholeprogvar}{\runtimetermvar}{\trace}$.
+%With this, we defined property satisfaction as follows:
+\bul{A whole program $\wholeprogvar$ satisfies a property $\pi$} iff \iul{$\wholeprogvar$ yields a trace $\trace$} such that \oul{$\trace$ satisfies $\pi$}.
 
 \begin{definition}[Property Satisfaction]\label{def:propsat}
-$\;$
-
-  \begin{nscenter}
   \bul{$\sat{\wholeprogvar}{\pi}$}
   $\isdef$
-    if \iul{$\forall\runtimetermvar\ \trace,\progstepto{\wholeprogvar}{\runtimetermvar}{\trace}$},
+  if \iul{$\forall\trace,\tracegen{\wholeprogvar}{\trace}$},
   then \oul{$\trace\in\pi$}.
-  \end{nscenter}
 \end{definition}

### 2025-02-12 icfp/sections/bg/hsat.tex
Sketch some hprop stuff

\subsection{Hyperproperties and (Robust) Hyper-Satisfaction}\label{subsec:bg:hprop}

The formal language of trace-based properties is incapable of expressing certain security-relevant restrictions on program behaviours.
For example, consider {\em noninterference}~\cite{}, which demands that high-security inputs have no influence on low-security outputs.
To describe this property, it is necessary to refer to two possibly distinct traces.
It is possible to describe such properties as so-called hyperproperties~\cite{clarkson2008hyper}.

A hyperproperty ($\varHProp$) is a set of program behaviours, i.e., a set of sets of traces, which have been referred to as {\em class} (see \Cref{subsec:bg:tprop}).
Much like trace-based properties, hyperproperties can also be classified into different categories, e.g., hypersafety or hyperliveness~\cite{}.

We extend \Cref{def:propsat,def:proprsat} as follows:

%TODO: this is wrong. rework to explain it via program behaviours instead
\begin{definition}[Hyper-Satisfaction]\label{def:prophsat}
  $\sat{\wholeprogvar}{\varHProp}$
  $\isdef$
  $\forall\varProp\in\varHProp,\sat{\wholeprogvar}{\varProp}$.
\end{definition}
\begin{definition}[Robust Hyper-Satisfaction]\label{def:proprhsat}
  $\rsat{\progvar}{\varHProp}$
  $\isdef$ 
  $\forall\varProp\in\varHProp,\rsat{\progvar}{\varProp}$.
\end{definition}

The definitions in \Cref{subsec:bg:rtp} are trivially extended to the hyperproperty case, so we omit it.
However, it is worth noting that program refinement does not commute with hyperproperty refinement, while it does so for traces.
% TODO: write more
Consider 
\[
\begin{array}{rcl}
  \operatorname{behav}\left(\src{p}\right) &=& \left\{ \src{\event_0} \right\} \\
  \operatorname{behav}\left(\src{p'}\right) &=& \left\{ \src{\event_0}, \src{\event_1} \right\} \\[.33cm]
  \varHProp &=& \left\{ \left\{ \src{\event_0} \right\}, \left\{ \src{\event_1} \right\} \right\} \\
  \varHProp' &=& \left\{ \left\{ \src{\event_1} \right\} \right\} \\
\end{array}
\]
Clearly, $\operatorname{behav}(\src{p})\subseteq\operatorname{behav}(\src{p'})$, so $\src{p}$ {\em refines} $\src{p'}$.
Similarly, $\varHProp'\subseteq\varHProp$.
Note that $\sat{\src{p}}{\varHProp}$, but $\not\sat{\src{p}}{\varHProp'}$.
This does not break for ordinary trace-based properties, since refinement coincides with satisfaction.

Based off of \Cref{def:prophsat,def:propsat}, it is well known that an ordinary trace-based property can be lifted to an equivalent hyperproperty by means of simply taking its powerset~\cite{clarkson2008hyper}.

### 2025-02-12 icfp/sections/bg/sat.tex
Sketch some hprop stuff

@@
-\subsection{Properties and (Robust) Satisfaction}\label{subsec:bg:tprop}
+\subsection{Trace Properties and (Robust) Satisfaction}\label{subsec:bg:tprop}
 
@@
 For a given program, we refer to the set of traces as its behaviour. 
-Moreover, sets of traces are called properties ($\pi$) which describe admissible program behaviours. % ascribing what said property considers valid.
+Moreover, sets of traces are called properties ($\varProp$) which describe admissible program behaviours. % ascribing what said property considers valid.
 % 
@@
 %With this, we defined property satisfaction as follows:
-\bul{A whole program $\wholeprogvar$ satisfies a property $\pi$} iff \iul{$\wholeprogvar$ yields a trace $\trace$} such that \oul{$\trace$ satisfies $\pi$}.
+\bul{A whole program $\wholeprogvar$ satisfies a property $\varProp$} iff \iul{$\wholeprogvar$ yields a trace $\trace$} such that \oul{$\trace$ satisfies $\varProp$}.
 
 \begin{definition}[Property Satisfaction]\label{def:propsat}
-  \bul{$\sat{\wholeprogvar}{\pi}$}
+  \bul{$\sat{\wholeprogvar}{\varProp}$}
   $\isdef$
   if \iul{$\forall\trace,\tracegen{\wholeprogvar}{\trace}$},
-  then \oul{$\trace\in\pi$}.
+  then \oul{$\trace\in\varProp$}.
 \end{definition}
@@
 Robust satisfaction considers partial programs $\progvar$, i.e., components with missing imports, which cannot run until said imports are fulfilled.
-To remedy this, {\em linking} takes two partial programs $\progvar[_{1}],\progvar[_{2}]$ and produces a whole program $\wholeprogvar$, i.e., $\link{\progvar[_{1}]}{\progvar[_{2}]}=\wholeprogvar$.
+To fill those gaps, {\em linking} takes two partial programs $\progvar[_{1}],\progvar[_{2}]$ and produces a whole program $\wholeprogvar$, i.e., $\link{\progvar[_{1}]}{\progvar[_{2}]}=\wholeprogvar$.
+
 As typically done in works that consider the execution of partial programs~\cite{abate2019jour,devriese2018parametricity,patrignani2021rsc,korashy2021capableptrs,strydonck2019lincap,devriese2017modular,bowman2015noninterference,ahmed2011equivcps,patterson2017linkingtyps},
 this paper assumes that whole programs are the result of linking partial programs referred to as {\em context} ($\ctx$) and {\em component} ($\comp$).
-The context is an arbitrary program and thus has the role of an {\em attacker} that can interact with the component by means of any features the programming language has, and the component is what is security-relevant.
+The context is an arbitrary program and thus has the role of an {\em attacker} that can interact with the component by means of any features the programming language has, while the component is what is security-relevant.
 %In this work, the semantics of the programming language is expected to differentiate between the component and the context.
-With this, \Thmref{def:propsat} can be extended as follows: for \bul{a component $\progvar$ to robustly satisfy a property $\pi$}, take an \iul{attacker context $\contextvar$ and link it with $\progvar$}, \oul{the resulting whole program must satisfy $\pi$}.
+With this, \Thmref{def:propsat} can be extended as follows: for \bul{a component $\progvar$ to robustly satisfy a property $\varProp$}, take an \iul{attacker context $\contextvar$ and link it with $\progvar$} such that \oul{the resulting whole program must satisfy $\varProp$}.
 
 \begin{definition}[Robust Satisfaction]\label{def:proprsat}
-  $\;$
-
-  \begin{nscenter}
-  \bul{$\rsat{\progvar}{\pi}$}
-  $\isdef$ \iul{$\forall \contextvar\ \wholeprogvar$, if $\link{\contextvar}{\progvar}=\wholeprogvar$}, then \oul{$\sat{\wholeprogvar}{\pi}$}.
-  \end{nscenter}
+  \bul{$\rsat{\progvar}{\varProp}$}
+  $\isdef$ \iul{$\forall \contextvar\ \wholeprogvar$, if $\link{\contextvar}{\progvar}=\wholeprogvar$}, then \oul{$\sat{\wholeprogvar}{\varProp}$}.
 \end{definition}

### 2025-02-13 icfp/sections/intro.tex
comments

@@
+% !TEX root = ./../main.tex
+
 \section{Introduction}\label{sec:introduction}
@@
   That is, there is a guarantee that all memory accesses are {\em temporally} as well as {\em spatially} safe. 
-  Temporal memory safety is the absence of use-after-free and use-after-reallocation bugs and necessitates tracking of pointer provenance\footnote{Due to potential use-after-reallocation bugs.}.
-  Provenance of pointers is also crucial information, together with bounds, for spatial memory safety, which ensures that all accesses are within an allocated object.
+  Temporal memory safety is the absence of use-after-free and use-after-reallocation bugs and necessitates tracking of pointer provenance.\footnote{Due to potential use-after-reallocation bugs.}
+  Provenance of pointers is also crucial information, together with bounds, for spatial memory safety, which ensures that all pointer accesses are within the bounds of an allocated object.
   There is a large body of prior work modifying existing memory-unsafe languages~\cite{} or providing compiler-level enforcements~\cite{} for both temporal and spatial memory safety.
@@
   Intuitively, this is why we want compilers to be {\em correct}, so that they provably preserve the meaning of the source-language.
+  \MP{not flowing, logically}
   With preservation of meaning follows preservation of safety guarantees.
-  Unfortunately, software is seldolmy developed in isolation and programs may be linked with target-level libraries after compilation.
+  Unfortunately, software is seldom developed in isolation and programs may be linked with target-level libraries after compilation.
   These libraries can be compiled with different compilers or may even be hand-written.
   In turn, the correctness-guarantees provided by one compiler do not extend to this setting.
+  \MP{the correctness thing breaks the flow of the argument}
 
-  In previous work, it has been argued that correct compilation is a spectrum~\cite{} with different restrictions applying to how the compiled partial programs can interact with target-level libraries.
+\MP{abate et al, not spectrum thing}
+  Previous work argued that correct compilation is a spectrum~\cite{} with different restrictions applying to how the compiled partial programs can interact with target-level libraries.
   At the extreme end of the spectrum with basically zero restrictions\footnote{The minimal restriction that has to be applied is interface compatibility.} lies robust preservation of meaning.
@@
   For example, in order to robustly ensure memory safety in Rust when compiling to RISC-V, the compiler could compartmentalize all calls and returns via the foreign function interface.
-  This compartmentalization can be done in different ways, one of which is via CHERI capabilities~\cite{}.
+  This compartmentalization can be done in different ways, one of which is via CHERI capabilities~\cite{}. \MP{ don't need}
 
   % Gap
-  For the context of this paper, we are interested in formally verified compilers.
-  That is, they need to be proven to be robustly preserving, which is a labor-intensive and in parts difficult task~\cite{}. %lots of cits
+  For the context of this paper, we are interested in formally verified compilers. \MP{don't need}
+  That is, they need to be proven to be robustly preserving, which is a labor-intensive and in parts difficult task~\cite{}. \MP{don't need}
+  %lots of cits 
   Worse, compilers are not just simple functions translating from one domain into another. 
@@
   Therefore, it is helpful to have local meaning-preservation proofs for each pass and be able to compose them to obtain a whole-proof for the complete pipeline.
+  \MP{ can likely  }
 
@@
   However, this is only necessary for the component, i.e., the partial program that is being compiled, and so source-independent linking asserts that the target context need not be related to any source-context.
+  \MP{wut?}
 

### 2025-02-13 icfp/sections/bg/hsat.tex
Sketch out hsat

@@
 A hyperproperty ($\varHProp$) is a set of program behaviours, i.e., a set of sets of traces, which have been referred to as {\em class} (see \Cref{subsec:bg:tprop}).
-Much like trace-based properties, hyperproperties can also be classified into different categories, e.g., hypersafety or hyperliveness~\cite{}.
+Much like trace-based properties, hyperproperties can also be classified into different categories, e.g., hypersafety or hyperliveness~\cite{clarkson2008hyper}.
+Based off of \Cref{def:prophsat,def:propsat}, it is well known that an ordinary trace-based property can be lifted to an equivalent hyperproperty by means of simply taking its powerset~\cite{clarkson2008hyper}.
 
@@
 
-%TODO: this is wrong. rework to explain it via program behaviours instead
 \begin{definition}[Hyper-Satisfaction]\label{def:prophsat}
@@
   $\isdef$
-  $\forall\varProp\in\varHProp,\sat{\wholeprogvar}{\varProp}$.
+  $\behav{\wholeprogvar}\in\varHProp$.
 \end{definition}
@@
   $\isdef$ 
-  $\forall\varProp\in\varHProp,\rsat{\progvar}{\varProp}$.
+  $\forall\contextvar\ \wholeprogvar,\text{if}\ \link{\contextvar}{\progvar}=\wholeprogvar, \text{then}\sat{\wholeprogvar}{\varHProp}$.
 \end{definition}
 
-The definitions in \Cref{subsec:bg:rtp} are trivially extended to the hyperproperty case, so we omit it.
-However, it is worth noting that program refinement does not commute with hyperproperty refinement, while it does so for traces.
-% TODO: write more
-Consider 
-\[
-\begin{array}{rcl}
-  \operatorname{behav}\left(\src{p}\right) &=& \left\{ \src{\event_0} \right\} \\
-  \operatorname{behav}\left(\src{p'}\right) &=& \left\{ \src{\event_0}, \src{\event_1} \right\} \\[.33cm]
-  \varHProp &=& \left\{ \left\{ \src{\event_0} \right\}, \left\{ \src{\event_1} \right\} \right\} \\
-  \varHProp' &=& \left\{ \left\{ \src{\event_1} \right\} \right\} \\
-\end{array}
-\]
-Clearly, $\operatorname{behav}(\src{p})\subseteq\operatorname{behav}(\src{p'})$, so $\src{p}$ {\em refines} $\src{p'}$.
-Similarly, $\varHProp'\subseteq\varHProp$.
-Note that $\sat{\src{p}}{\varHProp}$, but $\not\sat{\src{p}}{\varHProp'}$.
-This does not break for ordinary trace-based properties, since refinement coincides with satisfaction.
+The definitions in \Cref{subsec:bg:rtp} are trivially extended to the hyperproperty case, so we omit them.
+However, it is worth noting that program refinement does not commute with hyper-satisfaction, while it does commute for ordinary satisfaction.
+To see this, consider the following example:
+ %\[
+ %\begin{array}{c}
+ %  \operatorname{behav}\left(\src{\wholeprogvar}\right) = \left\{ \src{\trace[_0]} \right\} \quad
+ %  \operatorname{behav}\left(\src{\wholeprogvar'}\right) = \left\{ \src{\trace[_0]}, \src{\trace[_1]} \right\} \qquad
+ %  \varHProp = \left\{ \left\{ \src{\trace[_0]} \right\}, \left\{ \src{\trace[_1]} \right\} \right\} \quad
+ %  \varHProp' = \left\{ \left\{ \src{\trace[_1]} \right\} \right\} \\
+ %\end{array}
+ %\]
+\begin{center}
+\begin{tikzpicture}
+  \node (a) {$\behav{\src{\wholeprogvar}}=\left\{\src{\trace[_0]}\right\}$};
+  \node[right=2cm of a] (b) {$\behav{\src{\wholeprogvar[']}}=\left\{\src{\trace[_0]},\src{\trace[_1]}\right\}$};
+  \node[below=1cm of b] (d) {$\varHProp=\left\{\left\{\src{\trace[_0]},\src{\trace[_1]}\right\},\left\{\src{\trace[_1]}\right\}\right\}$};
 
-Based off of \Cref{def:prophsat,def:propsat}, it is well known that an ordinary trace-based property can be lifted to an equivalent hyperproperty by means of simply taking its powerset~\cite{clarkson2008hyper}.
+  \node[right=.8cm of a] (ab) {$\subseteq$};
+  \node[below=.5cm of b,rotate=270] (bd) {$\in$};
+  \node[below=.3cm of ab,rotate=-33] (ad) {$\not\in$};
+\end{tikzpicture}
+\end{center}
+
+Clearly, $\operatorname{behav}(\src{\wholeprogvar})\subseteq\operatorname{behav}(\src{\wholeprogvar'})$, so $\src{\wholeprogvar}$ {\em refines} $\src{\wholeprogvar'}$.
+However, while $\src{\wholeprogvar[']}$ satisfies $\varHProp$ ($\sat{\src{\wholeprogvar[']}}{\varHProp}$), we have that $\not\sat{\src{\wholeprogvar}}{\varHProp}$.
+This does not break for ordinary trace-based properties, since refinement coincides with satisfaction, i.e., taking subsets is a transitive operation.

### 2025-02-13 icfp/sections/bg/sat.tex
Sketch out hsat

@@
 
-Running a (whole) program yields a trace ($\trace$), i.e., a potentially infinite sequence\footnote{
+Running a (whole) program ($\wholeprogvar$) yields a trace ($\trace$), i.e., a potentially infinite sequence\footnote{
 Throughout the paper, sequences are indicated with an overbar (i.e., $\trace$), empty sequences with $\hole{\cdot}$, and concatenation of sequences $\trace[_{1}],\trace[_{2}]$ as $\trace[_{1}]\cdot\trace[_{2}]$.
@@
 Events include security-relevant actions (e.g., reading from and writing to memory, as detailed in \Cref{sec:compprop}), the unobservable event ($\emptyevent$), and a crash ($\lightning$).
-For a given program, we refer to the set of traces as its behaviour. 
+For a given program, we refer to the set of traces as its behaviour $\behav{\wholeprogvar}$. 
 Moreover, sets of traces are called properties ($\varProp$) which describe admissible program behaviours. % ascribing what said property considers valid.
@@
 
-As an example, consider the trace $\ev{Dealloc\ \loc}\cdot\ev{Read\ \loc\ 1729}\cdot\dots$ which describes the interaction with memory where the deallocation of an address ($\loc$) precedes a read (of some value $1729$) at that address in memory.
+As an example for a safety property, consider the trace $\ev{Dealloc\ \loc}\cdot\ev{Read\ \loc\ 1729}\cdot\dots$ which describes the interaction with memory where the deallocation of an address ($\loc$) precedes a read (of some value $1729$) at that address in memory.
 % 
@@
   if \iul{$\forall\trace,\tracegen{\wholeprogvar}{\trace}$},
-  then \oul{$\trace\in\varProp$}.
+  then \oul{$\trace\in\varProp$} $\equiv\behav{\wholeprogvar}\subseteq\varProp$.
 \end{definition}

### 2025-02-13 icfp/main.tex
comments
move

@@
+\PassOptionsToPackage{dvipsnames}{xcolor} % Added for new colors.
 % Fixing: Too many math alphabets used in version normal
@@
 \input{sections/intro.tex}
-
 \input{sections/background.tex}
-
-\section{Secure Composition}\label{sec:sequential}
-
-Notably, real-world compilation chains also perform a series of (sequential) passes whose main purpose is not necessarily to translate from one language to another, but to, e.g., optimise the code or enforce a certain property.
-Both examples can be seen in practice, e.g.,~\cite{nagarakatte2009soft,nagarakatte2010cets,akritidis2009baggy,wegman1991ccp,manjikian1997fusion} and many more.
-% \MPin{survey to cite?
-this is a bit informal MK: hard to find a survey on different compiler passes...
-I can't even find one just for compiler enforced memsafety (that looks good) or for compiler translations in general}.
-% 
-% Consider the compilation passes of LLVM~\cite{lattner2004llvm}, which are handled in the \texttt{PassManager} class.
-% By manipulating the internal state of a \texttt{PassManager} object, the schedule of compiler optimisation passes can be hardcoded.
-% That way, an order of optimisations can be found by hand that yields the best results.
-% \MP{what's the point of the previous 3 sentences?(commented)}
-Consider the following two LLVM optimisation passes: \gls*{cf}, which rewrites constant expressions to the constant they evaluate to, and \gls*{dce}, which removes dead code by rewriting conditional branches.
-The order in which \gls*{cf} and \gls*{dce} are performed influences the final result of the compilation (see \Cref{fig:cfdceex}).
-\begin{figure}[!ht]
-  \begin{center}
-    \begin{tikzpicture}[scale=0.62, every node/.style={transform shape}]
-      \node[minimum width=3cm,minimum height=2.5cm,draw=black,very thick,align=left] (progunopt) {\begin{lstlisting}[language=ml,basicstyle=\small\ttfamily]
-let a = true in
-if a then
-  print "a"
-else
-  print "b"
-\end{lstlisting}};
-      \node[minimum width=3cm,minimum height=2.5cm,draw=black,very thick,align=left,right=1.0 of progunopt] (progoptcf) {\begin{lstlisting}[language=ml,basicstyle=\large\ttfamily]
-if true then
-  print "a"
-else
-  print "b"
-\end{lstlisting}};
-      \node[minimum width=3cm,minimum height=2.5cm,draw=black,very thick,align=left,right=1.0 of progoptcf] (progoptcfdce) {\begin{lstlisting}[language=ml,basicstyle=\Large\ttfamily]
-print "a"
-\end{lstlisting}};% the following is a hack to get my syntax highlighting to work again \end{lstlisting}
-      % edges
-      \draw[->,very thick] (progunopt) edge[loop left,left] node {\gls*{dce}} (progunopt);
-      \draw[->,very thick] (progunopt) edge[sloped,above] node {\gls*{cf}} (progoptcf);
-      \draw[->,very thick] (progoptcf) edge[sloped,above] node {\gls*{dce}} (progoptcfdce);
-    \end{tikzpicture}
-  \end{center}
-  \caption{Example program where the level of optimisations differ for one pass of applying \gls*{cf} and \gls*{dce} in any order. %
-  Every edge is a compilation pass and the label on the edge states what the pass does, i.e., \gls*{cf} or \gls*{dce}. %
-  The source code in the nodes is a glorified compiler intermediate representation and the code gets more optimised towards the right hand side of the figure.}\label{fig:cfdceex}
-\end{figure}
-This {\em phase ordering problem} is well--known in literature and a practical solution is to simply perform a fixpoint iteration of the optimisation pipeline~\cite{click1995combining}.
-
-\smallskip
-
-To analyse the security of compilation passes and their interaction within a compilation pipeline, we rely on a few key notions: the definition of a trace relation, and the definition of when is a trace relation well-formed with respect to a class.
-%
-Consider traces $\src{\trace}$ and $\trg{\trace}$ as well as two trace relations $\sim_1$ and $\sim_2$. 
-% 
-The traces are related $\src{\trace}\sim_1\bullet\sim_2\trg{\trace}$ if there is another trace $\irl{\trace}$ such that $\src{\trace}\sim_1\irl{\trace}$ and $\irl{\trace}\sim_2\trg{\trace}$.
-% 
-\bul{A relation $\sim$ is well-formed w.r.t. a class of target-level properties $\trg{\class}$} iff \rul{the universal image preserves set membership}.
-\begin{definition}[Well-formedness of $\sim$ for a Class $\trg{\class}$]\label{def:wfc:sig:tracerel}
-  % $$
-
-  \begin{nscenter}
-  \noindent
-  \text{\bul{$\wfcsig{\sim}{\trg{\class}}$}} \isdef \text{\rul{$\forall \trg{\pi}\in\trg{\class}, \sigma_\sim(\trg{\pi})\in\sigma_\sim(\trg{\class})$}}
-  \end{nscenter}
-  % $$
-\end{definition}
-
-We can now state our main result: secure compilers in the robust compilation framework~\cite{abate2021extacc} compose {\em sequentially}. 
-%
-
-Let \bul{$\cc{\src{L}}{\trg{L}}$ robustly preserve the class $\sigma_{\sim_2}\left(\obj{\class[_1]}\right)$ under $\sim_1$} and let \rul{$\cc{\trg{L}}{\obj{L}}$ robustly preserve the class $\obj{\class[_2]}$ under $\sim_2$}.
-Then, \iul{when the cross-language relation $\sim_2$ is well-formed w.r.t.
-class $\obj{\class[_1]}$}, it follows that \oul{the composed compiler\footnote{Here, $\cc{\src{L}}{\trg{L}}\circ\cc{\trg{L}}{\obj{L}}$ means first running $\cc{\src{L}}{\trg{L}}$ and then $\cc{\trg{L}}{\obj{L}}$.} $\cc{\src{L}}{\trg{L}}\circ\cc{\trg{L}}{\obj{L}}$ robustly preserves the intersection of classes $\obj{\class[_1]}\cap\obj{\class[_2]}$ under $\sim_1\bullet\sim_2$}.
-% 
-%For space reasons, \Thmref{thm:urtp} and \Thmref{thm:lrtp} are not presented here, but can be found in the technical appendix ($\CoqSymbol$).
-% Note that it is not necessary to require $\sim_2$ to be well-formed w.r.t. $\src{\class[_1]}$, since the 
-% \MPin{why}
-\begin{theorem}[Composition of Secure Compilers w.r.t. $\sigma$]\label{thm:rtpsim:sig}
-  $\;$
-
-  If \bul{$\rtpsig{\cc{\src{L}}{\trg{L}}}{\tilde{\sigma}_{\sim_2}\left(\obj{\class[_{1}]}\right)}{\sim_1}$}, \rul{$\rtpsig{\cc{\trg{L}}{\obj{L}}}{\obj{\class[_2]}}{\sim_2}$}, and \iul{$\wfcsig{\sim_2}{\obj{\class[_1]}}$}, \\ then \oul{$\rtpsig{\cc{\src{L}}{\trg{L}}\circ\cc{\trg{L}}{\obj{L}}}{\obj{\class[_{1}]}\cap\obj{\class[_{2}]}}{\sim_1\bullet\sim_2}$}.
-\Coqed
-\end{theorem}
-
-Since the composition of secure compilers is again a secure compiler, the theorems generalise to a whole chain of $n$ secure compilers. 
-% 
-\Cref{thm:rtpsim:sig} can also be stated for the existential image $\tau_\sim\left(\src{\pi}\right)$, but in the interest of space that result has been moved to the appendix.
- % (\Cref{subsec:extimg}).
-% 
-Crucially, \Cref{thm:rtpsim:sig} also holds for \emph{classes of hyperproperties}, and thus, compilers that robustly preserve hyperproperties can be composed with each other as well as with compilers that robustly preserve properties.
-% 
-%Unfortunately, proofs of compilers that robustly preserve hyperproperties (including hypersafety) are scarce in the literature due to their complexity.
-%\MP{ so what ? }
-%  we don't discuss this further?
-%
-
-% It is furthermore possible to conflate the notion of classes and hyperproperties by intersecting all elements of a class of hyperproperties, to obtain one combined hyperproperty.
-% This idea has been discussed in other work~\cite{clarkson2008hyper}.
-%  MP: i think i stated the point of these 2 lines above
-
-If we take SecurePtrs and FaCT from \Cref{subsec:bg:rtp} and compose them according to \Cref{thm:rtpsim:sig}, we obtain a compiler that robustly preserves the intersection of safety properties and the \gls*{cct} hyperproperty.
-That is, for a source component that robustly satisfies any set of safety properties {\em and} \gls*{cct}, the compiled target component also robustly satisfies the same set of safety properties and \gls*{cct}.
-% The intersection simply uses the lifted version of the respective safety properties.
-%\MP{meaning?}
-
-Compiler engineers typically try to find an order of optimisations that yields well-optimised programs for either code size~\cite{cooper1999geneticphases} or performance~\cite{kulkarni2006exhaustivephase}.
-\Cref{corr:swappable} justifies that any such order of compilation passes is valid w.r.t. security, as long as the trace-relations have no effect on the respective classes.
-
-So, given two compilation passes \bul{$\cc[_{1}]{\trg{L}}{\trg{L}}$, $\cc[_{2}]{\trg{L}}{\trg{L}}$, both robustly preserving class $\trg{\class[_{1}]}$ or $\trg{\class[_{2}]}$, respectively}, \rul{their corresponding well-formed trace-relations}, \iul{and indifference of the classes with respect to these trace relations}, \oul{for any order of their composition, the composed compiler robustly preserves the intersection of $\trg{\class[_{1}]}$ and $\trg{\class[_{2}]}$}.
-% 
-% \MPin{
-% 	later on, say that optimisation passes remove elements from traces, so they should be fine w.r.t. the result with $\tau$.
-% 	also, yes, if $\tau$ adds we have a problem, if it removes we don't
-% \MKin{ Optimisations that remove elements from traces are not necessarily sound,
-%  e.g., remove all deallocations to "optimise" memory management by offloading
-%  the work to the OS.
-% }
-% }
-
-\begin{corollary}[Swapping Secure Compiler Passes]\label{corr:swappable}
-  $\;$
-
-  If \bul{$\rtpsig{\cc[_{1}]{\trg{L}}{\trg{L}}}{\trg{\class[_{1}]}}{\sim_1}$ and $\rtpsig{\cc[_{2}]{\trg{L}}{\trg{L}}}{\trg{\class[_{2}]}}{\sim_2}$}, %
-  \rul{$\wfcsig{\sim_1}{\trg{\class[_2]}}$ and $\wfcsig{\sim_2}{\trg{\class[_1]}}$}, %
-  and \iul{$\tilde{\sigma}_{\sim_2}\left(\trg{\class[_1]}\right)=\trg{\class[_1]}$ as well as $\tilde{\sigma}_{\sim_1}\left(\trg{\class[_2]}\right)=\trg{\class[_2]}$},
-  then \oul{$\rtpsig{\cc[_{1}]{\trg{L}}{\trg{L}}\circ\cc[_{2}]{\trg{L}}{\trg{L}}}{\trg{\class[_{1}]}\cap\trg{\class[_{2}]}}{\sim_1\circ\sim_2}$ and $\rtpsig{\cc[_{2}]{\trg{L}}{\trg{L}}\circ\cc[_{1}]{\trg{L}}{\trg{L}}}{\trg{\class[_{2}]}\cap\trg{\class[_{1}]}}{\sim_2\circ\sim_1}$}.
-\Coqed
-\end{corollary}
-
-Coming back to the example composing SecurePtrs with FaCT, it is likely the case that \Cref{corr:swappable} is not applicable.
-While first running SecurePtrs and then FaCT should be fine, the other direction has potential security hazards, since the SecurePtrs compiler does not account for cryptographic-constant time primitives, such as \texttt{ctselect}.
-
-\subsection{Secure Compiler Composition with Same Trace Models}
-When the cross-language trace relation is an equality, \Cref{thm:rtpsim:sig} collapses:
-Given \bul{$\cc{\src{L}}{\trg{L}}$ robustly preserves $\class[_{1}]$} and \rul{$\cc{\trg{L}}{\obj{L}}$ robustly preserves $\class[_{2}]$}, it follows that \oul{their sequential composition $\cc{\src{L}}{\trg{L}}\circ\cc{\trg{L}}{\obj{L}}$ robustly preserves the intersection of classes $\class[_{1}]$ and $\class[_{2}]$}.
-
-\begin{corollary}[Composition of Secure Compilers]\label{corr:rtp}
-  $\;$
-
-  If \bul{$\rtp{\cc{\src{L}}{\trg{L}}}{\class[_{1}]}$} and \rul{$\rtp{\cc{\trg{L}}{\obj{L}}}{\class[_{2}]}$}, then \oul{$\rtp{\cc{\src{L}}{\trg{L}}\circ\cc{\trg{L}}{\obj{L}}}{\class[_{1}]\cap\class[_{2}]}$}.
-\Coqed
-\end{corollary}
-
-\Cref{corr:rtp} provides an easy way to compose secure compilers without well-formedness of trace relations. 
-% 
-However, while \Cref{thm:rtpsim:sig} explicitly requires that $\sim_1$ is well-formed w.r.t. $\trg{\class[_2]}$, if $\sim_2$ is not well-formed w.r.t. $\trg{\class[_1]}$, care must be taken. 
-%Worst-case is that the resulting class $\sigma_{\sim_1\bullet\sim_2}\left(\trg{\class[_1]}\cap\trg{\class[_2]}\right)$ becomes empty, even though $\trg{\class[_1]}\cap\trg{\class[_2]}$ is not necessarily empty.
-This is further discussed in \Cref{subsec:compatsecpasses}.
-
-%If possible, optimisations are merged to enhance the analysis' result, since even when iterating to a fixpoint the final program may not be as optimised as it could be~\cite{click1995combining}.
-We can also obtain a specialised version of \Cref{corr:swappable}:
-%So, given two compilation passes \bul{$\cc[_{1}]{\trg{L}}{\trg{L}}$, $\cc[_{2}]{\trg{L}}{\trg{L}}$, both robustly preserving class $\class[_{1}]$ or $\class[_{2}]$, respectively}, \oul{for any order of their composition, the composed compiler robustly preserves the intersection of $\class[_{1}]$ and $\class[_{2}]$}.
-% 
-% \MPin{
-% 	later on, say that optimisation passes remove elements from traces, so they should be fine w.r.t. the result with $\tau$.
-% 	also, yes, if $\tau$ adds we have a problem, if it removes we don't
-% \MKin{ Optimisations that remove elements from traces are not necessarily sound,
-%  e.g., remove all deallocations to "optimise" memory management by offloading
-%  the work to the OS.
-% }
-% }
-
-\begin{corollary}[Swapping Secure Compiler Passes]\label{corr:swappable:one}
-  $\;$
-
-  If {$\rtp{\cc[_{1}]{\trg{L}}{\trg{L}}}{\class[_{1}]}$ and $\rtp{\cc[_{2}]{\trg{L}}{\trg{L}}}{\class[_{2}]}$}, then {$\rtp{\cc[_{1}]{\trg{L}}{\trg{L}}\circ\cc[_{2}]{\trg{L}}{\trg{L}}}{\class[_{1}]\cap\class[_{2}]}$ and $\rtp{\cc[_{2}]{\trg{L}}{\trg{L}}\circ\cc[_{1}]{\trg{L}}{\trg{L}}}{\class[_{2}]\cap\class[_{1}]}$}.
-\Coqed
-\end{corollary}
-
-\section{Security Properties Formalisation \& Composition\pages{1.5}}\label{sec:compprop}
-
-This section introduces trace properties of interest for this paper: \gls*{tms}, \gls*{sms}, \gls*{ms}, \gls*{scct}, and \gls*{ss}.
-These properties are of practical importance (as mentioned in \Cref{sec:introduction}) and also of interest in the case study presented later (\Cref{sec:casestud:defs,sec:casestud:rtp}). 
-This section presents all of them, despite the fact that they are inspired by existing work, in order to showcase all that is required for a formal proof of security for a realistic compilation toolchain.
-%\MP{ say something about this in the rebuttal: only the defs are borrowed, the universal images and all that work is new}
-%\MK{ universal images are not new. }
-% , none of the properties is novel, for self-containment of the paper, their definition is necessary for the case study.
-% For each of the presented properties, the technical report defines corresponding monitors, which refine a given property and have a key role in the proofs of this paper. 
-The technical report defines monitors for each of the presented properties.
-% 
-Monitors refine each property and have a key role in the proofs of this paper.
-% managed to prove 
-%  if tms-monitor --[trace]-->* ∅  
-%  then trace ∈ tms
-% but not
-%  if trace ∈ tms
-%  then tms-monitor --[trace]-->* ∅
-
-\subsection{A Trace Model for Memory Safety}\label{subsec:basic:memsafety:tracemodel}
-
-For simple memory safety composed of temporal and spatial memory safety, the trace model defines events ($\event[_{\mssafe}]$) as either the empty event ($\emptyevent$), a crash ($\lightning$), or a base-event ($\preevent[_{\mssafe}]$).
-
-\vspace{-1.0em}
-\[
-  \begin{array}{rcll}
-    \text{(Base-Events)}&\preevent[_{\mssafe}] &:=& \ev{Alloc \loc\ n} \mid \ev{Dealloc \loc} \mid \ev{Use \loc\ n} \\
-    \text{(Events)}&\event[_{\mssafe}] &:=& \preevent[_{\mssafe}] \mid \emptyevent \mid \lightning \\ 
-  \end{array}
-\]
-
-Base-events describe the actual kind of event that happened.
-For the basic memory-safety properties, these are three variants:
-First, the allocation event ($\ev{Alloc\ \loc\ n}$) that fires whenever a program claims $n$ cells of memory and stores them at address $\loc$, where addresses are assumed to be unique.
-Second, deallocation ($\ev{Dealloc\ \loc}$) announces that the object at location $\loc$ is freed.
-Third, an event to describe reads from and writes to the $n$-th memory cell from address $\loc$ ($\ev{Use\ \loc\ n}$).
-
-\subsubsection{Temporal Memory Safety}
-
-\gls*{tms}~\cite{nagarakatte2010cets} is a safety property that describes that an unallocated object must not be (re-)used.
-
-\begin{definition}[\glsfirst*{tms}]\label{def:trace:tmsdef}
-  % \footnotesize
-  $$
-  \tmssafe:=\left\{\trace[_{\mssafe}] \left| \begin{array}{rcl}
-    \ev{Alloc\ \loc\ n}&\le_{\trace[_{\mssafe}]}&\ev{Dealloc\ \loc} \\
-    \ev{Use\ \loc\ n}&\le_{\trace[_{\mssafe}]}&\ev{Dealloc\ \loc} \\
-    \text{at most one }\ev{Dealloc\ \loc}&\text{in}&\trace[_{\mssafe}] \\
-    \text{at most one }\ev{Alloc\ \loc\ n}&\text{in}&\trace[_{\mssafe}] \\
-  \end{array}\right.\right\}
-  $$
-\end{definition}
-Hereby, the notation $\event[_{1}]\le_{\trace}\event[_{2}]$ means that if $\event[_{1}]$ is in $\trace$ and if $\event[_{2}]$ is in $\trace$, then $\event[_{1}]$ appears before $\event[_{2}]$.
-
-\subsubsection{Spatial Memory Safety}
-
-\gls*{sms}~\cite{nagarakatte2009soft} is a safety property that prohibits out-of-bounds accesses.
-
-\begin{definition}[\glsfirst*{sms}]\label{def:trace:smsdef}
-% \small
-  % \begin{nscenter}
-
-  \noindent
-  \[
-  \smssafe:=\left\{\trace[_{\mssafe}] \left|\begin{array}{rcl}
-      \text{If }\ev{Alloc\ \loc\ n}\le_{\trace_{\mssafe}}\ev{Use\ \loc\ m}, \text{ then }m<n
-  \end{array}\right.\right\}
-  \]
-  % \end{nscenter}
-\end{definition}
-
-\subsubsection{Memory Safety}
-
-In spirit of earlier work~\cite{nagarakatte2009soft,nagarakatte2010cets,jim2002cyclone,necula2005ccured,michael2023mswasm}, full \gls*{ms} is the intersection of \Cref{def:trace:tmsdef,def:trace:smsdef}.
-
-\begin{definition}[\glsfirst*{ms}]\label{def:trace:msdef}
-  $
-  \mssafe:=\tmssafe \cap \smssafe
-  $
-\end{definition}
-
-Note that \Cref{def:trace:msdef} ignores data isolation, so there may still be memory-safety issues introduced by side-channels.
-
-\subsection{A Trace Model for Memory Safety with Constant Time}\label{subsec:scct:tracemodel}
-
-To express Constant Time, we extend the memory safety trace model with a {\em security tag} ($\securitytag{}$) that indicates whether events contain sensitive information ($\lock$) or not ($\unlock$).
-
-\vspace{-.5em}
-\[
-  \begin{array}{rrcl}
-    (\text{Base-Events}) & \preevent[_{\ctsafe}] &:=& \preevent[_{\mssafe}] \mid \ev{Branch\ }n \mid \ev{Binop\ }n\\
-    (\text{Security Tags}) & \securitytag{} &:=& \lock \mid \unlock\\ 
-    (\text{Events}) & \event[_{\ctsafe}] &:=& \preevent[_{\ctsafe}];\securitytag{} \mid \emptyevent \mid \lightning \\ 
-  \end{array}
-\]
-
-For cryptographic code, there is a general guideline that secrets must not be visible on a trace~\cite{ctguidelines}, i.e., secrets should not be marked as $\unlock$.
-In turn, an instruction whose timing is data-dependent must not have a secret as an operand.
-Typical operations with data-dependent timing are branches and certain binary operations, such as division.%
-\footnote{
-	This is architecture-dependent, but division is an operation that serves as a classic example for a data-dependent timing instruction~\cite[p.~755]{arm-refman}.
-}
-Both operations are represented in the trace model by extending the set of base-events with branches ($\ev{Branch\ n}$) and binary operations ($\ev{Binop\ n}$).
-
-\subsubsection{Strict Cryptographic Constant Time}
-
-\gls*{cct} is a hypersafety property~\cite{barthe2018sec} and, thus, difficult to check with monitors.
-This is because, intuitively, hypersafety properties can relate multiple execution traces with each other, but monitors work on a single execution.
-It is a common trick to sidestep this issue by means of overapproximation: this section defines the property \gls*{scct}, a stricter variant of \gls*{cct} (inspired by earlier work~\cite{almeida2017jasmin}) that enforces the policy that no secret appears on a trace.
-Programs that satisfy \gls*{scct} also satisfy \gls*{cct}, but programs that satisfy \gls*{cct} may not satisfy \gls*{scct}.
-
-\begin{definition}[\glsfirst*{scct}]\label{def:trace:scctdef}
-% \begin{nscenter}
-
-  \noindent\[
-  \scctsafe:=\left\{\trace[_{\ctsafe}] \left|\begin{array}{l}
-      \trace[_{\ctsafe}]=\hole{\cdot} \text{ or } \\\exists\trace[_{\ctsafe}'],\trace[_{\ctsafe}]=\preevent[_{\ctsafe}];\unlock\cdot\trace[_{\ctsafe}'] \wedge \trace[_{\ctsafe}']\in\scctsafe
-    \end{array}\right.\right\}
-  \]
-% \end{nscenter}
-\end{definition}
-
-\gls*{scct} may appear overly strict, since it seems that secrets must not occur on a trace (since $\securitytag{}$ is forced to be $\unlock$). 
-However, this is considered standard practice in terms of coding guidelines~\cite{ctguidelines}.
-Moreover, programs that have been compiled with FaCT~\cite{cauligi2019fact} and run with a ``data independent timing mode''~\cite{arm-refman,intel-refman} enabled do not leak secrets (see \Cref{ex:lscct}).
-
-\subsubsection{\gls*{ms}, Strict Cryptographic Constant Time}\label{sec:msscct-rel}
-
-The combination of \gls*{ms} and \gls*{scct} is the intersection of these properties, \gls*{msscct}.
-However, \gls*{ms} uses a different trace model than \gls*{scct}, so intersecting them would trivially yield the empty set. 
-To remedy this issue, we introduce $\sim_{\ctsafe}: \preevent[_{\ctsafe}] \times \preevent[_{\mssafe}] $, a cross-language trace relation (whose key cases are presented below), that we use to intuitively unify the trace model in which the two properties are expressed:
-\[
-  \typerulenolabel{mscctrel:drop:tag}{}{\preevent[_{\mssafe}]\sim_{\ctsafe}\preevent[_{\mssafe}];\unlock{}}
-  \typerulenolabel{mscctrel:drop:crash}{}{\lightning\sim_{\ctsafe}\lightning}
-\]
-\[
-  \typerulenolabel{mscctrel:drop:branch}{}{\emptyevent\sim_{\ctsafe}\ev{Branch n};\securitytag{}}
-  \typerulenolabel{mscctrel:drop:binop}{}{\emptyevent\sim_{\ctsafe}\ev{Binop n};\securitytag{}}
-\]
-
-Essentially, $\sim_{\ctsafe}$ ignores both the new $\ev{Branch\ n}$ and $\ev{Binop\ n}$ base-events as it relates security-insensitive actions ($\unlock$) to their equivalent counterparts.
-Thus, \gls*{ms} traces trivially satisfy \gls*{scct}.
-% 
-The above relation is extended point-wise to traces, skipping the empty event $\emptyevent$ on either side, and it is now possible to define \gls*{msscct} using the universal image:
-
-\begin{definition}[\gls*{ms} and \gls*{scct}]\label{def:trace:msscctdef}
-  $
-  \msscctsafe:=\mssafe\cap\sigma_{\sim_{\ctsafe}}\left(\scctsafe\right)
-  $
-\end{definition}
-
-\subsection{Extending the Trace Model with Speculation}\label{subsec:msctss:tracemodel}
-
-So far, the considered trace models do not let us express speculative execution attacks such as Spectre~\cite{kocher2019spectre}. 
-For this, we extend the earlier trace model (see \Cref{subsec:scct:tracemodel}) so that the security tags ($\securitytag{}$) carry additional information about the kind of private data leakage, i.e., the type of speculative leak.
-Moreover, we add base-events signalling the beginning of a speculative execution ($\ev{Spec}$), a barrier ($\ev{Barrier}$) that signals that any speculative execution may not go past it, as well as a rollback event ($\ev{Rlb}$), which signals that execution resumes to where speculation started.
-
-\vspace{-1em}
-{
-\[
-  \begin{array}{rrcl}
-    (\text{Base-Events}) & \preevent[_{\mathghost}] &:=& \preevent[_{\ctsafe}] \\
-    (\text{Spectre Variants}) & vX &:=& \operatorname{NONE} \mid \operatorname{PHT} \\
-    (\text{Security Tags}) & \securitytag{} &:=& \lock_{vX} \mid \unlock\\ 
-    (\text{Events}) & \event[_{\mathghost}] &:=& \preevent[_{\mathghost}];\securitytag{} \mid \emptyevent \mid \lightning \mid \ev{Spec} \mid \ev{Rlb} \mid \ev{Barrier}\\ 
-  \end{array}
-\]
-}
-
-Even though the considered Spectre variants are just SPECTRE-PHT~\cite{kocher2019spectre}, NONE just describes secret data as in \gls*{scct} (see \Cref{subsec:scct:tracemodel}), the trace model is general enough to allow for potential future extension with different variants~\cite{kocher2019spectre,maisuradze2018ret2spec,horn2019zero}.
-%For sake of readability, this paper just uses the notation $\lock$ in place of $\lock_{\text{PHT}}$.
-
-\subsubsection{Speculative Safety}
-
-\gls*{ss}~\cite{patrignani2021exorcising}, similar to \gls*{scct}, is a sound overapproximation of a variant of noninterference.
-
-\begin{definition}[\glsfirst*{ss}]\label{def:trace:ss}
-  \noindent
-
-  \begin{nscenter}
-  $
-    \sssafe:=\left\{\trace[_{\mathghost}] \left|\begin{array}{l}
-      \trace[_{\mathghost}]=\hole{\cdot} \text{ or } \exists\trace[_{\mathghost}'].\\
-      \left(\trace[_{\mathghost}]=\preevent[_{\mathghost}];\unlock\cdot\trace[_{\mathghost}'] \text{or }\trace[_{\mathghost}]=\preevent[_{\mathghost}];\lock_{\text{NONE}}\cdot\trace[_{\mathghost}']\right)\\
-      \text{and }\ \trace[_{\mathghost}']\in\sssafe
-                                 \end{array}\right.\right\}
-  $ 
-  \end{nscenter}
-\end{definition}
-The technical setup so far leads to the above definition, where only locks annotated with $\text{SPECTRE-PHT}$ are disallowed to occur on the trace.
-That way, programs attaining \gls*{ss} do not necessarily attain \gls*{scct}.
-
-\subsubsection{Speculation Memory Safety}\label{sec:spec-ms-rel}
-
-As before, we need to relate the different trace models with each other, so that the memory safety property without speculation can be lifted to speculation. 
-To this end, let $\sim_{\mathghost}: \preevent[_{\mathghost}] \times \preevent[_{\ctsafe}]$ be a cross-language trace relation whose key cases are below.
-The intuition is that \gls*{ss} is trivially satisfied in \gls*{scct}, since speculation is inexpressible there, which amounts to dropping events $\ev{Spec}$, $\ev{Rlb}$, or $\ev{Barrier}$, as well as all base events tagged with $\lock_{\text{PHT}}$.
-
-\[
-  \typerulenolabel{specrel:drop:tag}{}{\preevent[_{\ctsafe}];\lock\sim_{\mathghost}\preevent[_{\mathghost}];\lock_{\text{NONE}}}
-  \typerulenolabel{specrel:drop:spectag}{}{\emptyevent\sim_{\mathghost}\preevent[_{\mathghost}];\lock_{\text{PHT}}}
-\]
-\[
-  \typerulenolabel{specrel:drop:pub}{}{\preevent[_{\ctsafe}];\unlock\sim_{\mathghost}\preevent[_{\mathghost}];\unlock}
-  \typerulenolabel{specrel:drop:crash}{}{\lightning\sim_{\mathghost}\lightning}
-\]
-\[
-  \typerulenolabel{specrel:drop:spec}{}{\emptyevent\sim_{\mathghost}\ev{Spec}}
-  \typerulenolabel{specrel:drop:rlb}{}{\emptyevent\sim_{\mathghost}\ev{Rlb}}
-  \typerulenolabel{specrel:drop:barrier}{}{\emptyevent\sim_{\mathghost}\ev{Barrier}}
-\]
-
-We conclude by defining the ultimate property of interest for secure compilers: \gls*{specms}.
-\begin{definition}[\gls*{specms}]\label{def:trace:specmsdef}
-  $
-  \specmssafe := \msscctsafe\cap\sigma_{\sim_{\mathghost}}\left(\sssafe\right)
-  $
-\end{definition}
-
-\section{Case Study: Language Formalisations\pages{2}}\label{sec:casestud:defs}
-
-This section defines programming languages $\Ltms$, $\Ltrg$, $\Lms$, $\Lscct$, and $\Lspec$, all of which share common elements (presented in \Cref{subsec:cs:defs}).
-$\Ltms$ is the only statically-typed language, and it exhibits the property that all well-typed programs are \gls*{tms} (\Cref{subsec:ltms}).
-However, not all $\Ltms$ programs are \gls*{sms}.
-That is, there are well-typed $\Ltms$ programs that perform an out-of-bounds access.
-Language $\Ltrg$ is untyped and does not provide any guarantees with regards to \gls*{ms} (\Cref{subsec:lsms}).
-$\Lms$ is exactly the same language as $\Ltrg$, but this paper still distinguishes the two for sake of readability (\Cref{subsec:lms}).
-All three languages --- so $\Ltms$, $\Ltrg$, and $\Lms$ --- assume \gls*{scct} to hold, since this is -- in an ideal world -- what the programmer would expect, too: it is the job of the compiler to preserve and (potentially) enforce \gls*{scct} security~\cite{cauligi2019fact,nagarakatte2010cets,nagarakatte2009soft,akritidis2009baggy}.
-% This paper considers \gls*{cct} as an architecture-dependent property\MPin{can we justify this}, so
-
-Such consideration is also backed up by architecture providing a data (operand) independent timing mode, such as processors by Arm~\cite[p.~543]{arm-refman} and Intel~\cite[p.~80]{intel-refman}.
-%In spirit of this, all languages besides $\Lscct$ are trivially \gls*{scct}. 
-This kind of processor feature is modelled in language $\Lscct$ (\Cref{subsec:lscct}), where programs have access to a ``\gls*{cct}-mode'' and can change the leakage of emitted events according to the value of this mode (either $\obj{ON}$ or $\obj{OFF}$).
-
-Finally, modern processors also employ speculative execution to achieve speedups---and unfortunately generate Spectre attacks~\cite{kocher2019spectre}---and this is the extension of $\Lspec$ (\Cref{subsec:lspec}).
- % extends $\Lscct$ to model speculative execution.
-%
-Thus, all previous languages trivially satisfy \gls*{ss}, since they do not support speculative execution at all.
-% All languages share large parts of syntactic elements, which are presented in the next subsection (\Cref{subsec:cs:defs}).
-% Subsequent subsections extend on this accordingly.
-
-\subsection{Shared Language Definitions}\label{subsec:cs:defs}
-
-All presented programming languages share a common fragment which is partially presented here and in full detail in the technical report. 
-%\MP{put omitted stuff in appendix + ref there (and add textual explanation)}
-
-% \vspace{-.5em}
-{
-  \renewcommand{\src}[1]{\mathsf{#1}}
- \small
-\[
-  \begin{array}{rrcl}
-    \text{(Base-Events)} & \src{\preevent} &:=& \src{\ev{Alloc \loc\ n}} \mid \src{\ev{Dealloc \loc}} \mid \src{\ev{Get \loc\ n}} \mid \src{\ev{Set \loc\ n}} \\
-    \text{(Control Tags)} & \src{\sandboxtag{}} &:=& \src{\ctx} \mid \src{\comp} \\
-    \text{(Events)} & \src{\event} &:=& \src{\preevent;\sandboxtag{}} \mid \src{\emptyevent} \mid \src{\lightning} \\ 
-    \text{(Values)} & \src{v} &:=& \src{n} \\
-    \text{(Expressions)} & \src{\expr} &:=& \src{x} \mid \src{n} \mid \src{\lbinop{\expr[_1]}{\expr[_2]}} \mid \src{\lifz{\expr[_1]}{\expr[_2]}{\expr[_3]}} \\ 
-                         &&&\mid \src{\llet{x}{\expr[_1]}{\expr[_2]}} \mid \src{\llet{x}{new\ \expr[_1]}{\expr[_2]}} \\
-                         &&&\mid \src{\ldelete{x}} \mid \src{\lget{x}{\expr}} \mid \src{\lset{x}{\expr[_1]}{\expr[]}} \mid \src{stuck}\\
-                         &&&\mid \src{\lcall{f}{\expr}} \mid \src{\lreturn{\expr}} \\
-    \text{(Functions)} & \src{F} &:=& (\src{x};\src{\expr}) \\
-    \text{(Libraries)} &&&\hspace{-3.8em} \src{\library_{\ctx}},\src{\library_{\comp}} : \src{Vars} \to \src{F} 
-    % \hspace{2em}
-    %  : \src{Vars} \to \src{F} 
-     \\
-    \text{(Heaps)} & \src{H} &:=& \src{\hole{\cdot}} \mid \src{n},\src{H}
-    % \hspace{0.5cm}    
-    \\
-    \text{(Pointer Maps)} & \src{\Delta} && \text{ omitted for simplicity}
-    \\
-    \text{(Memory States)} & \src{\memstate} &:=& \left(\src{H^{\ctx}};\src{H^{\comp}};\src{\Delta}\right)\\
-    \text{(Control States)} & \src{\cfstate} & & \text{ omitted for simplicity} 
-    \\
-    \text{(States)} &\src{\Omega} &:=& \left(\src{\cfstate};\src{\sandboxtag{}};\src{\memstate}\right)\\
-  \end{array}
-\]
-}
-{
-  \renewcommand{\src}[1]{\mathsf{#1}}
-The trace models of all languages are similar to those presented earlier (\Cref{sec:compprop}).
-One technical detail is the addition of a control tag, indicating who is to blame for an emitted action: context ($\ctx$) or component ($\comp$).
-This is a standard technique in secure compilation in order to rule out irrelevant context-level events. 
-For example, a context immediately deallocating an allocated memory region twice trivially violates memory safety, but the main interest in secure compilation is to preserve component-level security, and thus component events.
-Even though this tagging could be used for blame preservation~\cite{patrignani2023blame}, this is beyond the focus of this paper.
-Another key difference is that memory accesses are now explicitly modelled as reads ($\src{\ev{Read\ \loc\ n}}$) and writes ($\src{\ev{Write\ \loc\ n}}$) instead of just uses.
-
-All languages have at least numbers as values ($\src{v}$) and second class functions ($\src{F}$).
-Functions are modelled as pairs containing the name of one argument and the body of the function.
-Bodies are just ordinary expressions, which can be simple binary operations ($\src{\lbinop{\expr[_1]}{\expr[_2]}}$), conditionals ($\src{\lifz{\expr[_1]}{\expr[_2]}{\expr[_3]}}$), function calls ($\src{\lcall{f}{e}}$) and returns ($\src{\lreturn{\expr}}$), as well as C-like memory operations.
-Programs have sets of pre-determined functions called libraries and they are marked as being part of some component ($\src{\library_{\comp}}$) or context ($\src{\library_{\ctx}}$).
-
-For the operational semantics, the runtime state ($\src{\Omega}$) is a triple consisting of a control-flow state ($\src{\cfstate}$), a control tag ($\src{\sandboxtag{}}$), and a memory state ($\src{\memstate}$).
-The latter carries information about pointers that are kept ``alive'' in pointer maps ($\src{\Delta}$), so that the semantics does not get stuck when encountering, e.g., a double-free.
-The memory state also carries two heaps to model sandboxing between a context ($\src{H^{\ctx}}$) and a component ($\src{H^{\comp}}$).
-
-}
-
-\subsection{$\src{L_{\tmssafe}}$: A Temporal but Not Spatial Memory Safe Language}\label{subsec:ltms}
-
-$\src{L_{\tmssafe}}$ is the only statically-typed language in this case study and restricts functions ($\src{F}$) to the typing signature $\src{\natt\to\natt}$. 
-The type system of $\src{L_{\tmssafe}}$ is inspired by $L^{3}$~\cite{morrisett2005L3,scherer2018fabulous} and enforces that every well-typed $\src{L_{\tmssafe}}$ program satisfies \gls*{tms}.
- % (\Cref{thm:wt:tms}).
-
-\begin{theorem}[$\src{L_{\tmssafe}}$-programs are \gls*{tms}]\label{thm:wt:tms}
-$\rsat{\src{\library_{\comp}}}{\tmssafe}$ \Coqed
-  % $\;$ 
-% 
-  % \begin{nscenter}
-  %   \hfill For any component $\src{\library_{\comp}}$, $\rsat{\src{\library_{\comp}}}{\tmssafe}$ \Coqed
-  % \end{nscenter}
-\end{theorem}
-
-\subsection{$\Ltrg$: A Memory-Unsafe Language}\label{subsec:lsms}
-
-$\Ltrg$ extends the syntax presented earlier (\Cref{subsec:cs:defs}) with dynamic typechecks ($\trg{\lhast{\expr}{\type}}$), evaluating to $\trg{0}$ if the type matches with the shape of $\trg{\expr}$ and $\trg{1}$ otherwise.
-%The pointers carry poison tags, which signify a pointer being allocated ($\trg{\poisoned}$) or freed ($\trg{\poisonless}$). 
-Furthermore, the syntax of $\Ltrg$ is extended with a way to inspect whether a pointer is freed ($\trg{\lispoisoned{x}}$), evaluating to $\trg{0}$ if it is freed and to $\trg{1}$ otherwise.
-Functions may receive arguments that are not $\trg{\natt}$, values are extended with tuples, and expressions are also extended with pair projections.
-
-% \vspace{-1em}
-\[
-  \begin{array}{rrcl}
-    \text{(Values)} & \trg{v} &:=& \dots \mid \trg{\lpair{v_1}{v_2}} \\
-    \text{(Expressions)} & \trg{\expr} &:=& \dots \mid \trg{\lhast{\expr}{\tau}} \mid \trg{\lispoisoned{x}} \mid \trg{e.1} \mid \trg{e.2}
-  \end{array}
-\]
-
-No changes are done to the trace model, but $\Ltrg$ has no static typing, making double-free code patterns possible.
-
-\subsection{$\Lms$: Another Memory-Unsafe Language}\label{subsec:lms}
-$\Lms$ is exactly equal to $\Ltrg$ (\Cref{subsec:lsms}), but used to emphasize that this is code after applying $\cc{\Ltrg}{\Lms}$ (\Cref{subsec:cs:ms}).
-
-\subsection{$\Lscct$: A Memory-Unsafe Language with a Constant Time Mode}\label{subsec:lscct}
-
-$\Lscct$ extends $\Lms$ (\Cref{subsec:lms}) with a ,,constant-time mode''. 
-The activation of the mode can be checked ($\obj{\lrddoit{x}{\expr}}$), changed ($\obj{\lwrdoit{D}}$), and is stored in program states ($\obj{\Omega}$).
-At the beginning of program execution of $\Lscct$ programs, the mode is turned off ($\obj{OFF}$).
-If the mode is enabled (i.e., set to $\obj{ON}$), the intuition is that no secrets are leaked. 
-This models the real-world\footnote{As present in Intel~\cite[p.80]{intel-refman} and ARM~\cite[p.~543]{arm-refman} processors.} data-independent timing mode as well as the result of compiling a program with FaCT~\cite{cauligi2019fact}.
-For example, FaCT rewrite code that branches to use a constant-time selection primitive.
-To not obfuscate our formalisation unnecessarily by duplicating syntax, we simply added the ,,constant-time mode'' to $\Lscct$.
-
-% $\Lscct$ extends $\Lms$ (\Cref{subsec:lms}) with a way to read from ($\obj{\lrddoit{x}{\expr}}$) and to write to ($\obj{\lwrdoit{D}}$) a model-specific register that controls a constant-time mode. 
-% This language is a combination of the effect of the FaCT compiler~\cite{cauligi2019fact} as well as a data-independent timing mode, a feature that is present in both Arm~\cite[p.~543]{arm-refman} and Intel~\cite[p.~80]{intel-refman} processors.
-% \MP{ can we be more specific about the fact pass? }
-% To this end, states ($\obj{\Omega}$) are extended with the value of the register ($\obj{D}$), which is initially set to be not active ($\obj{OFF}$).
-% If the register is marked active ($\obj{ON}$), the intuition is that no secrets can appear on specification traces.
-% If the register is marked inactive, secrets may appear on traces.
-% The state of the register can be read and written to, e.g., \Cref{tr:e-wrdoit-off}.
-% For the other languages seen earlier, the mode is intuitively always-on, i.e., the mode of execution always uses constant-time mode.
-
-% \vspace{-.5em}
-\[
-  \begin{array}{rrcl}
-    \text{(Mode Values)} & \obj{D} &:=& \obj{ON} \mid \obj{OFF} \\
-    \text{(Security Tags)} & \obj{\securitytag{}} &:=& \obj{\lock} \mid \obj{\unlock} \\
-    \text{(Base-Events)} & \obj{\preevent} &:=& \dots \mid \obj{iGet\ \loc\ v} \mid \obj{iSet\ \loc\ v} \\
-                       &&& \mid\ \obj{Binop\ n} \mid \obj{Branch\ n} \\
-    \text{(Events)} & \obj{\event} &:=& \obj{\preevent;\sandboxtag{};\securitytag{}} \mid \obj{\emptyevent} \mid \obj{\lightning} \\
-    \text{(Expressions)} & \obj{\expr} &:=& \dots \mid \obj{\lrddoit{x}{\expr}} \mid \obj{\lwrdoit{D}} \\
-                         &&&\mid\ \obj{\llet{x^{\securitytag{}}}{\expr[_1]}{\expr[_2]}}\\
-    \text{(States)} & \obj{\Omega} & := & \left(\obj{\cfstate};\obj{\sandboxtag{}};\obj{D};\obj{\memstate}\right)\\
-  \end{array}
-\]
-\begin{center}
-\newcommand{\expreval}[5]{{#1}\triangleright\xspace {#2}\xrightarrow{#5}\ {#3}\triangleright\xspace {#4}\xspace}
-\newcommand{\exprevalo}[5]{\expreval{\obj{#1}}{\obj{#2}}{\obj{#3}}{\obj{#4}}{\obj{#5}}}
-  \typerule{$e-\obj{wrdoit}-\text{off}$}{
-  }{
-    \pstepto{\rtt{\obj{\cfstate;\sandboxtag{};D;\memstate}}{\obj{\lwrdoit{OFF}}}}{\rtt{\obj{\cfstate;\sandboxtag{};OFF;\memstate}}{\obj{0^{\securitytag{}}}}}{\obj{\emptyevent}}
-  }{e-wrdoit-off}
-
-  \typerule{$e-\obj{get}-\in-$noleak}{
-    \obj{\memstate}=\obj{H^{\ctx};H^{\comp};\Delta_1,x\mapsto(\loc;\sandboxtag{};\poison),\Delta_2} &
-    \obj{\loc}+\obj{n}\in\text{dom }\obj{H^{\sandboxtag{}}} \\
-    \obj{\securitytag{''}}=\obj{\securitytag{}\sqcap\securitytag{'}} &
-    \obj{\event} = \obj{{iGet}\ \loc\ n;\sandboxtag{};\securitytag{''}}
-  }{
-    \exprevalo{
-    	\cfstate;\sandboxtag{'};ON;\memstate
-    }{
-      	x^{\securitytag[n^{\securitytag{'}}] }
-    }{
-    	\cfstate;\sandboxtag{'};ON;\memstate
-    }{
-    	(H^{\sandboxtag{}}(\loc+n))^{\securitytag{''}}
-    }{
-      \event
-    }
-  }{to-e-get-in-noleak}
-\end{center}
-
-The language also adds user annotations $\obj{\securitytag{}}$ for the secrecy of variables, which can be either private (high secrecy) $\obj{\lock}$ or public (low secrecy) $\obj{\unlock}$.
-Security tags $\obj{\securitytag{}}$ are arranged in the usual secrecy lattice~\cite{zdancewicphd}, where $\obj{\unlock}\sqsubseteq\obj{\lock}$.
-% $\obj{\lock}\sqsubseteq\obj{\securitytag{}}$ and $\obj{\securitytag{}}\sqsubseteq\obj{\unlock}$.
-
-Memory accesses to secret data need to be present to reason about memory safety, even when execution is in constant-time mode, e.g, \Cref{tr:to-e-get-in-noleak}.
-$\Lscct$ extends base-events with $\obj{\ev{iGet}\ \loc\ \valueexpr}$ and $\obj{\ev{iSet}\ \loc\ \valueexpr}$ (for data $\obj{i}$ndependent get and set) to prevent secrets from leaking but still enable reasoning about memory safety.
-Due to the technical setup, the rule needs to check if the access is in bounds ($\obj{\loc}+\obj{n}\in\obj{H^t}$) and update the secrecy tag ($\obj{\securitytag{''}}=\obj{\securitytag{}\sqcap\securitytag{'}}$) with the least upper bound of the tags, according to the aforementioned lattice.
-The precise information carried by the event, e.g., location ($\obj{\loc}$) is taken from the pointer map, which carries information irrelevant to this rule ($\obj{\poison}$).
-
-Base-events include $\obj{\ev{Branch}\ n}$ and $\obj{\ev{Binop}\ n}$ that are emitted when evaluating a branch or certain binary expressions, such as division, respectively, whenever the constant-time mode is inactive.
-Events are extended with a security-tag ($\obj{\securitytag{}}$) to signal the secrecy of the involved data.
-
-The evaluation steps are amended to propagate the security-tag annotations $\obj{\securitytag{}}$.
-When the constant-time mode is inactive, base-events $\obj{\ev{Branch}\ n}$ and $\obj{\ev{Binop}\ n}$ are emitted for conditionals and binary operations, respectively.
-Otherwise, just like in the semantics of the earlier languages, $\obj{\emptyevent}$ is emitted for binary and branching operations.
-%To mark this mode as active or inactive, \Cref{tr:e-wrdoit} allows to access the register in $\obj{\statevar}$.
-
-\Cref{ex:lscct} illustrates the differences between $\Lscct$ and other languages.
-
-\begin{exampleenv}[$\Lscct$ with and without constant-time mode]\label{ex:lscct}
-  Consider again \Cref{ex:strncpy}, with a context copying the string \texttt{Hello World}, where everything is marked with a high security tag: $\lock$.
-  The top half of \Cref{fig:ex-cct} (titled $\Lscct$), describes the execution trace of the program, while the bottom side of the table (titled ``Specification''), describes the related specification trace (\Cref{subsec:msctss:tracemodel}).
-  Read in parallel from top to bottom, the figure shows parts of the execution trace. 
-  In each half, the left column (\textit{Active}) has constant-time mode $\obj{ON}$ and the right one (\textit{Inactive}) has it $\obj{OFF}$.
-
-  When the constant-time mode is off, the execution yields events without constant-time leaks (\Cref{subsec:ltms,subsec:lsms,subsec:lms}).
-  But, if it is turned on, then the branching event \colorbox{yellow}{does not fire} anymore and both reading and writing to memory is related to a specification trace with \colorbox{lightgray}{no exposed secrets}.
-\end{exampleenv}
-\begin{figure}[!htb]
-	\vspace{-1em}\small
-  $$
-  % {
-  \begin{array}{ccc}
-    & \Lscct &        \\
-    \text{Active} & \mid & \text{Inactive} \\\hline
-    \obj{\ev{Alloc}\ \loc_{x}\ 12;\comp;\unlock} & \mid & \obj{\ev{Alloc}\ \loc_{x}\ 12;\comp;\unlock} \\
-    \obj{\ev{Alloc}\ \loc_{y}\ 12;\comp;\unlock} & \mid & \obj{\ev{Alloc}\ \loc_{y}\ 12;\comp;\unlock} \\
-    \obj{{\ev{iGet}}\ \loc_{x}\ 0;\comp;}\text{\colorbox{lightgray}{$\obj{\lock}$}} & \mid & \obj{\ev{Get}\ \loc_{x}\ 0;\comp;\lock} \\
-    \text{\colorbox{yellow}{$\obj{\emptyevent}$}} & \mid & \obj{\ev{Branch}\ 0;\comp;\lock} \\
-    \obj{{\ev{iGet}}\ \loc_{x}\ 0;\comp;}\text{\colorbox{lightgray}{$\obj{\lock}$}} & \mid & \obj{\ev{Get}\ \loc_{x}\ 0;\comp;\lock} \\
-    \obj{{\ev{iSet}}\ \loc_{y}\ 0\ \mathtt{'H'};\comp;}\text{\colorbox{lightgray}{$\obj{\lock}$}} & \mid & \obj{\ev{Set}\ \loc_{y}\ 0\ \mathtt{'H'};\comp;\lock} \\
-    \obj{{\ev{iGet}}\ \loc_{x}\ 1;\comp;}\text{\colorbox{lightgray}{$\obj{\lock}$}} & \mid & \obj{\ev{Get}\ \loc_{x}\ 1;\comp;\lock} \\
-    \text{\colorbox{yellow}{$\obj{\emptyevent}$}} & \mid & \obj{\ev{Branch}\ 0;\comp;\lock} \\
-    \vdots & \mid & \vdots \\
-    \obj{{\ev{iGet}}\ \loc_{x}\ 12;\comp;}\text{\colorbox{lightgray}{$\obj{\lock}$}} & \mid & \obj{\ev{Get}\ \loc_{x}\ 12;\comp;\lock} \\
-    \text{\colorbox{yellow}{$\obj{\emptyevent}$}}& \mid & \obj{\ev{Branch}\ 1;\comp;\lock} \\
-    \obj{\ev{Dealloc}\ \loc_{y};\comp;\unlock} & \mid & \obj{\ev{Dealloc}\ \loc_{y};\comp;\lock} \\
-    \obj{{\ev{iGet}}\ \loc_{y}\ 6;\comp;}\text{\colorbox{lightgray}{$\obj{\lock}$}} & \mid & \obj{\ev{Get}\ \loc_{y}\ 6;\comp;\lock} \\
-  \end{array}
-  % }
-  $$
-  $$
-  % {
-  \begin{array}{ccc}
-          & \text{Specification} & \\
-  \text{Active} & \mid & \text{Inactive} \\\hline
-     \ev{Alloc}\ \loc_{x}\ 12;\unlock & \mid & \ev{Alloc}\ \loc_{x}\ 12;\unlock\\
-     \ev{Alloc}\ \loc_{y}\ 12;\unlock & \mid & \ev{Alloc}\ \loc_{x}\ 12;\unlock\\
-     \ev{Use}\ \loc_{x}\ 0;\text{\colorbox{lightgray}{$\unlock$}} & \mid & \ev{Use}\ \loc_{x}\ 0;\lock\\
-     \emptyevent & \mid & \ev{Branch}\ 0;\lock\\
-     \ev{Use}\ \loc_{x}\ 0;\text{\colorbox{lightgray}{$\unlock$}} & \mid & \ev{Use}\ \loc_{x}\ 0;\lock\\
-     \ev{Use}\ \loc_{y}\ 0;\text{\colorbox{lightgray}{$\unlock$}} & \mid & \ev{Use}\ \loc_{y}\ 0;\lock\\
-     \ev{Use}\ \loc_{x}\ 1;\text{\colorbox{lightgray}{$\unlock$}} & \mid & \ev{Use}\ \loc_{x}\ 1;\lock\\
-     \emptyevent & \mid & \ev{Branch}\ 0;\lock\\
-     \vdots & \mid & \vdots \\
-      \ev{Use}\ \loc_{x}\ 12;\text{\colorbox{lightgray}{$\unlock$}} & \mid & \ev{Use}\ \loc_{x}\ 12;\lock\\
-      \emptyevent & \mid & \ev{Branch}\ 1;\lock\\
-      \ev{Dealloc}\ \loc_{y};\unlock & \mid & \ev{Dealloc}\ \loc_{y};\unlock\\
-      \ev{Use}\ \loc_{y}\ 6;\text{\colorbox{lightgray}{$\unlock$}} & \mid & \ev{Use}\ \loc_{y}\ 6;\lock\\
-  \end{array}
-  % }
-  $$
-  \caption{Traces for \Cref{ex:lscct}.}
-  \label{fig:ex-cct}
-\end{figure}
-
-\subsection{$\Lspec$: A Memory-Unsafe Language with Speculation}\label{subsec:lspec}
-
-$\Lspec$ extends $\Lscct$ (\Cref{subsec:lscct}) with speculative dynamic semantics that is inspired by existing work~\cite{guarnieri2018spectector,fabian2022automatic}.
-After branching, speculation starts by pushing the current configuration into the speculation state ($\ird{S}$), and run subsequent code in a predetermined window (whose size is $\omega$) until a rollback of operational state is performed.
-The window is set inside the speculation state.
-Within such windows, data may leak, this is marked as high $\ird{\lock}$ and with an annotation that indicates the respective speculative execution variant. 
-In this paper, we just consider SPECTRE-PHT~\cite{kocher2019spectre}, whose starting point is \Cref{tr:e-ifz-true-spec}.
-% where $\omega$ is the size of the speculation window, i.e., a number of reduction steps to be taken under speculative semantics.
-Additionally, $\ird{\lock}$ may also carry a $\ird{NONE}$ annotation, to signal that this is a leak irrespective of speculation, i.e., as defined earlier (\Cref{subsec:lscct}).
-The semantics is kept general enough to allow for future extension to support different variants~\cite{kocher2019spectre,maisuradze2018ret2spec,horn2019zero}.
-
-In terms of language features, $\Lspec$ includes a new barrier operation $\ird{\lbarrier}$ that blocks speculative execution. 
-To facilitate speculative execution in a non-assembly-like language, a stack of operational state is used and speculation is active if that stack has a size greater than one.
-This is exploited in \Cref{tr:e-spec-eat} to consume any leftover speculation.
-
-\vspace{-.5em}
-\[
-  \begin{array}{rrcl}
-    \text{(Base-Events)} & \ird{\preevent} &:=& \dots \mid \ird{Spec} \mid \ird{Rlb} \mid \ird{Barrier}\\
-    \text{(Expressions)} & \ird{\expr} &:=& \dots \mid \ird{\lbarrier} \\
-    \text{(Speculation State)} & \ird{S} &:=& \ird{\Omega\triangleright\expr} \mid \ird{S},\left(\ird{\Omega};\ird{n};\ird{\expr}\right)
-  \end{array}
-\]
-
-\begin{center}
-\newcommand{\expreval}[5]{{#1}\triangleright\xspace {#2}\xrightarrow{#5}\ {#3}\triangleright\xspace {#4}\xspace}
-\newcommand{\exprevald}[5]{\expreval{\ird{#1}}{\ird{#2}}{\ird{#3}}{\ird{#4}}{\ird{#5}}}
-
-  %
-  \typerule{$e-\ird{ifz}-\ird{true}-\text{spec}$}{
-    \ird{S} = \rtt{(\ird{\cfstate['];\sandboxtag{};D;\memstate});\ird{S}}{\ird{\expr[_1]}},({\ird{\cfstate;\sandboxtag{};D;\memstate}};\omega;{\ird{\expr[_2]}})
-  }{
-    \pstepto{\rtt{\ird{\cfstate;\sandboxtag{};D;\memstate}}{\ird{\lifz{0^{\securitytag{}}}{\expr[_1]}{\expr[_2]}}}}{\ird{S}}{\ird{{Branch}\ 0;\sandboxtag{};\securitytag{}}\cdot\ird{{Spec};\sandboxtag{};\securitytag{}}}
-  }{e-ifz-true-spec}
-  %
-  \typerule{$e-\ird{spec}-\text{eat}$}{
-    \ird{n} > 0 &
-    \pstepto{\rtt{\ird{\Omega}}{\ird{\expr}}}{\rtt{\ird{\Omega'}}{\ird{\expr[']}}}{\ird{\event}}
-  }{
-    \ghoststepto{\ird{S},(\ird{\Omega};\ird{n};\ird{\expr})}{\ird{S},(\ird{\Omega'};\ird{n}-1;\ird{\expr'})}{\ird{\event}}
-  }{e-spec-eat}
-  % 
-  \typerule{$e-\ird{spec}-\text{eaten}$}{
-  }{
-    \ghoststepto{\ird{S},(\ird{\Omega};\ird{0};\ird{\expr})}{\ird{S}}{\ird{Rlb;\Omega.\sandboxtag{};\unlock}}
-  }{e-spec-eaten}
-  %
-  \typerule{$e-\ird{spec}-\text{barrier}$}{
-  }{
-    \ghoststepto{\ird{S},(\ird{\Omega};\ird{n};\ird{\lbarrier})}{\ird{S},(\ird{\Omega};\ird{0};\ird{0})}{\ird{Barrier;\Omega.\sandboxtag{};\unlock}}
-  }{e-spec-barrier}
-\end{center}
-
-The trace model is extended with $\ird{{Spec}}$, $\ird{{Rlb}}$, and $\ird{{Barrier}}$ events that signal the respective operational action. 
-A $\ird{\lbarrier}$ prevents any execution whatsoever besides $\ird{{Rlb}}$ when run in speculative execution mode and does nothing when run in normal mode, e.g., \Cref{tr:e-spec-barrier}.%
-\footnote{In the rule, the notation $\cfstate' = \cfstate[.\text{win} = 0]$ means that $\cfstate'$ is a copy of $\cfstate$ up to field $\text{win}$, which is set to $0$.}
-Lastly, $\ird{Rlb}$ is emitted when the speculation window is zero and the operational state is rolled back (see \Cref{tr:e-spec-eaten}).
-
-\section{Case Study: Composing Secure Compiler Passes and Optimisations \pages{7}}\label{sec:casestud:rtp}
-This section defines several secure compilers, each of which robustly preserves a different property of interest as depicted in \Cref{fig:pipeline}.
-\begin{figure*}[!h]
-  \centering
-  \begin{tikzpicture}
-    \node (S) {$\src{L_{\tmssafe}}$};
-    \node[right=2.0 of S] (T) {$\trg{L}$};
-    \node[right=2.0 of T] (M) {$\irl{L_{\mssafe}}$};
-    \node[below right=1.5 and 1.0 of M] (D0) {$\irl{L_{\mssafe}}$};
-    \node[above right=1.5 and 1.0 of M] (C0) {$\irl{L_{\mssafe}}$};
-    \node[right=3.0 of M] (E) {$\irl{L_{\mssafe}}$};
-    \node[right=2.0 of E] (O) {$\obj{L_{\scctsafe}}$};
-    \node[right=2.0 of O] (Z) {$\ird{L_{\mathghost}}$};
-
-    \draw[->] (S) to[sloped] node[align=center] (tmsedge) {\gls*{tms}\\ \Cref{thm:cca:rtp:tms}} (T);
-    \draw[->] (T) to[sloped] node[align=center] {\gls*{sms}\\ \Cref{thm:ccb:rtp:sms}} (M);
-    \draw[->] (M) to[sloped] node[align=center] {\gls*{dce}\\ \Cref{thm:ccdce:rtp:ms}} (D0);
-    \draw[->] (M) to[sloped] node[align=center] {\gls*{cf}\\ \Cref{thm:cccf:rtp:ms}} (C0);
-    \draw[->] (D0) to[sloped] node[align=center] {\gls*{cf}\\ \Cref{thm:cccf:rtp:ms}} (E);
-    \draw[->] (C0) to[sloped] node[align=center] {\gls*{dce}\\ \Cref{thm:ccdce:rtp:ms}} (E);
-    \draw[->] (E) to[sloped] node[align=center] {\gls*{scct}\\ \Cref{thm:ccscct:rtp:scct}} (O);
-    \draw[->] (O) to[sloped] node[align=center] {\gls*{ss}\\ \Cref{thm:ccspec:rtp:spec}} (Z);
-
-    % Sections
-    \node (sectms) at ($(S)+(1.25,2.5)$) {{\Cref{subsec:cs:tms}}};
-    \node (secsms) at ($(T)+(1.25,2.5)$) {{\Cref{subsec:cs:ms}}};
-    \node (secopts) at ($(M)+(2.0,2.5)$) {{\Cref{subsec:cs:opts}}};
-    \node (secscct) at ($(E)+(1.5,2.5)$) {{\Cref{subsec:cs:scct}}};
-    \node (secss) at ($(O)+(1.5,2.5)$) {{\Cref{subsec:cs:ss}}};
-
-    \draw[thick,dotted,->] ($(S) + (0.2,-0.27)$) to ($(S) - (-0.2,1)$) to node[align=center] {\gls*{ms}\\ \Cref{corr:ccab:rtp:ms}} ($(M) - (0.0,1)$) to (M);
-
-    \draw[thick,dotted,->] (S) to ($(S) - (0.0,2.5)$) to node[align=center,pos=.65] {\gls*{specms}\\ \Cref{thm:ccall:rtp:specms}} ($(Z) - (0.0,2.5)$) to (Z);
-
-    % \draw[thick,dotted,->] (S) to[bend right=33,sloped] node[align=center] {\gls*{ms}\\ \Cref{corr:ccab:rtp:ms}} (M);
-    \draw[thick,dotted,->] (M) to[bend right=0,sloped] node[align=center] {\gls*{ms}\\ \Cref{thm:cccfccdce:rtp:ms}} (E);
-%    \draw[thick,dotted,->] (S) to[out=-45,in=198,sloped] node[align=center] {\gls*{ms}+\gls*{scct}\\ \Cref{thm:ccall:rtp:msscct}} (O);
-    % \draw[thick,dotted,->] (S) to[out=-40,in=210,sloped] node[align=center,pos=.8] {\gls*{specms}\\ \Cref{thm:ccall:rtp:specms}} (Z);
-
-    \draw[dashed] ($(S)-(0.0,-0.5)$) -- ($(S)-(0.0,-2.5)$);
-    \draw[dashed] ($(T)-(0.0,-0.5)$) -- ($(T)-(0.0,-2.5)$);
-    \draw[dashed] ($(M)-(0.0,-0.5)$) -- ($(M)-(0.0,-2.5)$);
-    \draw[dashed] ($(E)-(0.0,-0.5)$) -- ($(E)-(0.0,-2.5)$);
-    \draw[dashed] ($(O)-(0.0,-0.5)$) -- ($(O)-(0.0,-2.5)$);
-    \draw[dashed] ($(Z)-(0.0,-0.5)$) -- ($(Z)-(0.0,-2.5)$);
-  \end{tikzpicture}
-  \vspace{-1em}
-  \caption{Visualisation of the optimising compilation pipeline that preserves \gls*{specms}. %
-    Vertices in the graph are the programming languages from earlier sections (\Cref{sec:casestud:defs}). %
-    Full edges are secure compilers passes.
-    Dotted edges are composition of passes and use the presented framework (\Cref{sec:sequential}) to indicate the property they preserve. %
-    The dashed lines partition the graph into the sections where the respective theorems are presented.
-  }\label{fig:pipeline}
-\end{figure*}
-The section demonstrates the power of the framework (\Cref{sec:sequential}) by composing these compilers for a secure and optimising compilation chain that robustly preserves \gls*{specms}.
-The first step in this chain is the compiler from $\src{L_{\tmssafe}}$ to $\trg{L}$ that robustly preserves just \gls*{tms} (\Cref{thm:cca:rtp:tms}).
-From here, another pass from $\trg{L}$ to $\irl{L_{\mssafe}}$ ensures that no out-of-bounds accesses can happen and, thus, programs at this point attain \gls*{sms} (\Cref{thm:ccb:rtp:sms}).
-Since these properties compose into \gls*{ms}, composing these passes yields a compiler that robustly preserves \gls*{ms} (\Cref{corr:ccab:rtp:ms}).
-Then, the section presents two optimisation passes, namely \gls*{cf} and \gls*{dce}, each of which robustly preserves \gls*{ms} (\Cref{thm:ccdce:rtp:ms,thm:cccf:rtp:ms}).
-These passes can be freely ordered in the compilation chain without compromising memory safety (\Cref{thm:cccfccdce:rtp:ms}).
-The next step in the chain ensures that code stays \gls*{scct} (\Cref{thm:ccscct:rtp:scct}) when compiled from $\Lms$ to $\Lscct$, which is done by switching on a constant-time mode for the computation.
-Lastly, by introducing barriers immediately after branches, speculative leaks via SPECTRE-PHT are prevented when compiling $\Lscct$ to $\Lspec$.
-The final result is that the whole compilation chain robustly preserves \gls*{specms} (\Cref{thm:ccall:rtp:specms}).
-
-\subsection{Robust Temporal Memory Safety Preservation}\label{subsec:cs:tms}
-
-The secure compiler from $\Ltms$ to $\Ltrg$ needs to ensure that when execution switches from context to component, the type signatures are respected.
-Thus, the compiler inserts the following dynamic typechecks before entering the body of a component-defined function (anything elided is a trivial identity function from source to target):
-
-\vspace{-1em}
-{
-\begin{gather}
-  \begin{align*}%{rcl}
-    %\cca(\src{x}) = &\ \trg{x} 
-    % \\
-  	%&
-  	%\qquad
-    %\cca(\src{\lbinop{\expr[_{1}]}{\expr[_{2}]}}) = &\ \lbinop[\trg]{\left[\cca(\src{\expr[_{1}]})\right]}{\left[\cca(\src{\expr[_{2}]})\right]} \\
-    %\cca(\src{n}) = &\ \trg{n} 
-    % \\
-    %&
-    \cca(\src{\lget{x}{\expr}}) = &\ \lget[\trg]{\trg{x}}{\left[\cca(\src{\expr})\right]} \\
-    \cca(\src{\ldelete{x}}) = &\ \ldelete[\trg]{\left[\cca(\src{x})\right]} 
-    \\
-    \cca(\src{\lfunction{g}{x:\natt\to\type[_{e}]}{\expr}})  =&
-  \end{align*}
-  \\
-  \begin{align*}
-%    \cca(\src{\lset{x}{\expr[_{1}]}{\expr[_{2}]}}) & = \lset[\trg]{\left[\cca(\src{x})\right]}{\left[\cca(\src{\expr[_{1}]})\right]}{\left[\cca(\src{\expr_{2}})\right]} \\
-% \\ 
-\lfunction[\trg]{\trg{g}}{\trg{x}}{\lifz[\trg]{\trg{\lhast{x}{\natt}}}{
-                                                            \left[\cca(\src{\expr})\right] %
-                                                                                                 }{\labort[\trg]}}
-  \end{align*}
-\end{gather}
-}
-
-\noindent Since $\trg{L}$ has no static typing, an attacker $\trg{\library_{\ctx}}$ can invoke a component function accepting a $\src{\natt}$ with $\trg{\lpair{17}{29}}$.
-With the dynamic check, the compiler ensures that execution aborts in such cases.
-%The compiler does not insert other checks and proceeds as the identity function (which in this paper amounts to a simple re-colouring of $\src{L_{\tmssafe}}$ to $\trg{L}$ expressions).
-
-Compiling the \texttt{strncpy} function from \Cref{sec:introduction} with $\cca$, the compiler would in this case ensure that the arguments that are evaluated in the compiled \texttt{strncpy} are valid.
-%This effectively has no influence on the resulting trace, so a corresponding cross-language trace relation $\sim^{\Ltms}_{\Ltrg}$ is just equality.
-
-$\cca$ is robustly preserving (\Cref{def:rtp}) \gls*{tms}:
-\begin{theorem}[$\cca$ secure w.r.t. \gls*{tms}]\label{thm:cca:rtp:tms}
-    $\rtp{\cca}{\tmssafe}$ % \Coqed
-\end{theorem}
-
-\subsection{Robust Spatial Memory Safety Preservation}\label{subsec:cs:ms}
-
-The spatial memory safety preserving compiler from $\Ltrg$ to $\Lms$ only inserts bounds-checks whenever reading from or writing to memory in order to enforce \gls*{sms}.
-These bounds checks need the bounds information, which the compiler keeps around by introducing a fresh identifier $\irl{x_{SIZE}}$ for each allocation that binds $\irl{x}$.
-Then, it is simply a matter of referring to that variable and ensuring that memory accesses are in the interval $[\irl{0},\irl{x_{SIZE}})$.
-When the check fails, the code aborts.
-
-\begin{nscenter}
-% \small
-  $$
-  \begin{array}{rcl}
-    \ccb(\trg{\lnew{x}{\expr[_{1}]}{\expr[_{2}]}}) & = 
-                                                   & \llet[\irl]{\irl{x_{SIZE}}}{\ccb(\trg{\expr[_{1}]})}{}
-    		\\&&
-    		\lnew[\irl]{\irl{x}}{\irl{x_{SIZE}}}{\ccb(\trg{\expr[_{2}]})}
-    		 \\
-  \ccb(\trg{\lget{x}{\expr}}) & = 
-                              & \llet[\irl]{\irl{x_{n}}}{\ccb(\trg{\expr})}{}
-  	\\&&
-  \lifz[\irl]{\irl{0\le x_{n}<x_{SIZE}}}{\\&&\irl{\lget{x}{x_{n}}}}{}
-  		\irl{\labort}
-  	  \\
-  \ccb(\trg{\lset{x}{\expr[_{1}]}{\expr[_{2}]}}) & = 
-                                                 & \llet[\irl]{\irl{x_{n}}}{\ccb(\trg{\expr[_{1}]})}{}
-  		\\&&
-  \lifz[\irl]{\irl{0\le x_{n}<x_{SIZE}}}{\\&&\lset[\irl]{\irl{x}}{\irl{x_{n}}}{}
-  		\ccb(\trg{\expr[_{2}]})
-  		}{\irl{\labort}} 
-  		% \\
-  \end{array}
-  $$
-\end{nscenter}
-
-\begin{exampleenv}[Instrumented \texttt{strncpy}]
-  Consider again \texttt{strncpy}, but instrumented for \gls*{sms}:
-    \begin{lstlisting}[language=c,basicstyle=\ttfamily, morekeywords={size_t}]
-void strncpy(size_t n, size_t dst_size, char *dst,
-             size_t src_size, char *src) {
-  for(size_t i = 0; i < src_size
-      && src[i] != '\0' && i < n; ++i) {
-    if(i < src_size && i < dst_size) {
-      dst[i] = src[i];
-    }
-  }
-}
-    \end{lstlisting}
-Consider context \texttt{strncpy(2, x, y)}, where \texttt{x} and \texttt{y} are pointers to valid regions of memory with allocated space for exactly two cells and do not contain the null-terminating character \texttt{'\textbackslash 0'}.
-% 
-Without the \gls*{sms} pass, the event $\ev{Use}\ \loc_{x}\ 2;\comp;\unlock$ would appear on the trace, but that indicates an out-of-bounds access! 
-Fortunately, with \gls*{sms} mitigation in place, that event does not appear during execution, since the bounds check prevents the condition \texttt{src[i] != '\textbackslash 0'} from executing.
-\end{exampleenv}
-
-Contrary to the previous compiler, $\ccb$ may change the trace of the original $\Ltrg$ program: if there is a memory access, it needs to be protected with a bounds check.
-% 
-The corresponding relation $\sim^{\Ltrg}_{\Lms} : \trg{\trace}\times\irl{\trace}$ that describes this semantic effect of the compiler is defined partially below.
-% 
-We omit action $\trg{Set}$, which is treated analogously, and any other event, which is related to its cross-language equivalent.
-% \footnote{Up to colours.}
-% \footnote{In the interest of space, the full details are in the technical report.}
-
-{
-\[
-  \typerulenolabel{xrel:sms:read}{\trg{n}\text{ in bounds}}{\trg{Get\ \loc\ n;\comp}\sim^{\Ltrg}_{\Lms}\irl{Get\ \loc\ n;\comp}}
-\]
-% \[
-%   \typerulenolabel{xrel:sms:panic}{\trg{n}\text{ out of bounds}}{\trg{Get\ \loc\ n;\comp}\sim^{\Ltrg}_{\Lms}\irl{\lightning}}
-% \]
-}
-
-% Note that neither $\Ltrg$ nor $\Lms$ expose branches on the trace.\MP{so what?}
-For simplicity, we elide the environment that this relation carries around in order to bind each location to its metadata (such as its size), and resolve the ``$\trg{n}$ in/out of bounds'' premise.
-% The bounds check mentioned in the premises of the rules above is left abstract for the sake of simplicity, the interested reader can find
-% since the precise definition needs some technicalities that only inhibit readability without any technical insights.
-% 
-We can now prove that compiler $\ccb$ robustly preserves \gls*{sms}.
-\begin{theorem}[$\ccb$ secure w.r.t. \gls*{sms}]\label{thm:ccb:rtp:sms}
-  $\rtpsim{\ccb}{\smssafe}{\sim^{\Ltrg}_{\Lms}}$ % \Coqed
-\end{theorem}
-
-At this point we can compose $\ccb$ with the previous compiler ($\cca$), but in order to do so, we need a trace relation from $\Ltms$ to $\Lms$.
-% 
-We can obtain this relation by composing the trace relation we just defined ($\sim^{\Ltrg}_{\Lms}$) with the one used by the previous compiler: $\sim^{\Ltms}_{\Ltrg} : \src{\trace}\times\trg{\trace}$.
-% 
-The latter has not been previously defined (nor has it been used in the related theorem) because that is just an equality relation, since the trace models of $\Ltms$ and of $\Ltrg$ are the same.
-% 
-Thus, we formally define $\sim^{\Ltms}_{\Lms}: \src{\trace} \times \irl{\trace}$ as the following composition: $\sim^{\Ltms}_{\Ltrg}\bullet\sim^{\Ltrg}_{\Lms}$.
-% 
-With this relation, \Cref{corr:ccab:rtp:ms} states that the composition of $\cca$ and $\ccb$ is secure w.r.t.
-\gls*{ms} and it follows from \Cref{thm:cca:rtp:tms,thm:ccb:rtp:sms} using \Cref{thm:rtpsim:sig}.
-% 
-\begin{corollary}[$\cca\circ\ccb$ secure w.r.t. \gls*{ms}]\label{corr:ccab:rtp:ms}
-  $\;$
-
-  \begin{nscenter}
-    $\rtpsim{\cca\circ\ccb}{\mssafe}{\sim^{\Ltms}_{\Lms}}$ % \Coqed
-  \end{nscenter}
-\end{corollary}
-% 
-This proof requires another precondition besides \Cref{thm:ccb:rtp:sms,thm:cca:rtp:tms}: $\sim^{\Ltms}_{\Ltrg}$ needs to be well-formed with respect to $\smssafe$.
-This follows trivially since $\sim^{\Ltms}_{\Ltrg}$ is an equality. 
-% 
-\begin{lemma}[$\sim^{\Ltms}_{\Ltrg}$ well-formed w.r.t. $\smssafe$]\label{lem:wf:ltmsltrg}
-$\;$
-
-  \begin{nscenter}
-  $\wfcsig{\sim^{\Ltms}_{\Ltrg}}{\smssafe}$
-  \end{nscenter}
-\end{lemma}
-
-\subsection{Optimising Compilers}\label{subsec:cs:opts}
-
-This section defines two optimising compiler passes from $\Lms$ to $\Lms$ which perform \gls*{dce} and \gls*{cf}, respectively.
-The \gls*{dce} pass applies a na\"ive rewrite rule on conditionals.
-The \gls*{cf} pass relies on an auxiliary function \texttt{mix} that uses a substitutions accumulator $\irl{\substlist}$ in order to rewrite constant binary operations, e.g., $\irl{{17}-1}$ to $\irl{16}$, and replace variables that are assigned to constants, e.g., $\irl{\llet{x}{7}{x}}$ to $\irl{7}$.
-
-\vspace{-1em}
-\begin{gather*}
-  \begin{align*}
-    \ccdce(\irl{\lifz{true}{\expr[_{1}]}{\expr[_{2}]}}) & = \ccdce(\irl{\expr[_{1}]}) &\\
-    \ccdce(\irl{\lifz{false}{\expr[_{1}]}{\expr[_{2}]}}) & = \ccdce(\irl{\expr[_{2}]}) &
-  \end{align*}
-  \\
-  % \begin{align*}
-  %   \ccdce(\irl{\lbinop{\expr[_{1}]}{\expr[_{2}]}}) & = \lbinop[\irl]{\ccdce(\irl{\expr[_{1}]})}{\ccdce(\irl{\expr[_{2}]})} &
-  % \end{align*}
-  % \\[0.25cm]
-  \begin{align*}
-    \cccf(\irl{\expr}) & = \partialeval{\irl{\expr}}{\irl{\hole{\cdot}}} &
-  \end{align*}
-  \\[0.125cm]
-  \begin{align*}
-   \partialeval{\irl{x}}{\irl{\substlist}} & = \irl{n} 
-   	\qquad\qquad \text{if } \irl{\subst{n}{x}}\in\irl{\substlist} \\
-   \partialeval{\irl{x}}{\irl{\substlist}} & = \irl{x} 
-   \qquad\qquad \text{if } \irl{\subst{n}{x}}\notin\irl{\substlist} \\
-   \partialeval{\irl{\lbinop{n}{m}}}{\irl{\substlist}} & = \irl{k} 
-   \qquad\qquad \text{if } \lbinop{\irl{n}}{\irl{m}}=k \\
-   %\partialeval{\irl{\lbinop{\expr[_{1}]}{\expr[_{2}]}}}{\irl{\substlist}} & = \lbinop[\irl]{\partialeval{\irl{\expr[_{1}]}}{\irl{\substlist}}}{\partialeval{\irl{\expr[_{2}]}}{\irl{\substlist}}} \\
-   \partialeval{\irl{\llet{x}{n}{\expr}}}{\irl{\substlist}} & = \partialeval{\irl{\expr}}{\irl{\subst{n}{x}\cdot\substlist}} 
-  % \\
-  % \partialeval{\irl{\lget{x}{\expr}}}{\irl{\substlist}} & = \lget[\irl]{\irl{x}}{\partialeval{\irl{\expr}}{\irl{\substlist}}}
-% \end{align*}
-\end{align*}
-% \\
-% \begin{align*}
-%   \partialeval{\irl{\llet{x}{\expr[_{1}]}{\expr[_{2}]}}}{\irl{\substlist}} & = \llet[\irl]{\irl{x}}{\partialeval{\irl{\expr[_{1}]}}{\irl{\substlist}}}{\\&\partialeval{\irl{\expr[_{2}]}}{\irl{\substlist}}} \\
-%   \partialeval{\irl{\lifz{\expr[_{1}]}{\expr[_{2}]}{\expr[_{3}]}}}{\irl{\substlist}} & = \lifz[\irl]{\partialeval{\irl{\expr[_{1}]}}{\irl{\substlist}}}{\\&\partialeval{\irl{\expr[_{2}]}}{\irl{\substlist}}}{\partialeval{\irl{\expr[_{3}]}}{\irl{\substlist}}} \\
-% \end{align*}
-\end{gather*}
-% \vspace{-3em}
-
-Note that both passes have no effect on the resulting trace of a program, up to $\emptyevent$-steps. 
-Because of this, both passes have equality as corresponding cross language trace relation. 
-Moreover, it is straightforward to prove both passes as secure (\Cref{def:rtp}) w.r.t. \gls*{ms}.
-
-\begin{theorem}[$\ccdce$ secure w.r.t. \gls*{ms}]\label{thm:ccdce:rtp:ms}
-  $\rtp{\ccdce}{\mssafe}$ %\Coqed
-\end{theorem}
-\begin{theorem}[$\cccf$ secure w.r.t. \gls*{ms}]\label{thm:cccf:rtp:ms}
-  $\rtp{\cccf}{\mssafe}$ %\Coqe
-\end{theorem}
-
-With both \Cref{thm:ccdce:rtp:ms,thm:cccf:rtp:ms} it follows from \Cref{corr:swappable} that the two passes can be interchanged arbitrarily:
-
-\begin{theorem}[$\cccf\circ\ccdce$ and $\cccf\circ\ccdce$ are secure w.r.t. \gls*{ms}]\label{thm:cccfccdce:rtp:ms}
-$\;$
-
-  \begin{nscenter}
-  \phantom{and~}\!\!$\rtp{\cccf\circ\ccdce}{\mssafe}$
-
-  and~$\rtp{\ccdce\circ\cccf}{\mssafe}$ % \Coqed
-  \end{nscenter}
-\end{theorem}
-
-\subsection{Robust Strict Cryptographic Constant Time Preservation}\label{subsec:cs:scct}
-
-This section defines a compiler $\ccscct$ from $\Lms$ to $\Lscct$ that robustly preserves \gls*{scct}. 
-Given the fact that $\Lscct$ provides a \gls*{cct}-mode that can be turned on or off, the compiler inserts wrapper code for function calls and function bodies to ensure that execution in the component always happen in \gls*{cct}-mode.
-This simple flag combines the effect of FaCT~\cite{cauligi2019fact}
-
-\vspace{-1em}
-\[
-\begin{array}{rcl}
-  \ccscct(\irl{\lfunction{g}{x}{\expr}}) & = & \lfunction[\obj]{\obj{g}}{\obj{x}}{\obj{\lwrdoit{ON};}\ccscct(\irl{\expr})} \\
-  \ccscct(\irl{\lcall{g}{\expr}}) & = & \lcall[\obj]{\obj{g}}{\ccscct(\irl{\expr})\obj{; \lwrdoit{ON}}} 
-    % \\
-    % \ccscct(\irl{\lbinop{\expr[_{1}]}{\expr[_{2}]}}) & = \lbinop[\obj]{\ccscct{\irl{\expr[_{1}]}}}{\ccscct{\irl{\expr[_{2}]}}} 
-    % \\
-\end{array}
-\]
-% \vspace{-2em}
-%
-The context can overwrite the flag and exit the mode, but upon invoking a function that is part of the component, the flag is set again.
-Because of this, the corresponding cross-language trace relation $\sim^{\Lms}_{\Lscct}$, only relates events without secrets:%
-
-\begin{center} 
-  \typerulenolabel{xrel:scct:noleak}{}{\irl{\preevent;\comp}\sim^{\Lms}_{\Lscct}\obj{\preevent;\comp;\unlock}}
-\end{center}
-
-The compiler is secure w.r.t. \gls*{scct}: %, similarly proven as in \Cref{subsec:cs:tms}.
-
-\begin{theorem}[$\ccscct$ secure w.r.t. \gls*{scct}]\label{thm:ccscct:rtp:scct}
-  \small
-  $\rtpsim{\ccscct}{\scctsafe}{\sim^{\Lms}_{\Lscct}}$ % \Coqed
-\end{theorem}
-
-\subsection{Robust Speculative Safety Preservation}\label{subsec:cs:ss}
-
-This section defines the final compilation pass $\ccspec$, which ensures that $\Lscct$ programs, which are assumed to be \gls*{ss}, stay \gls*{ss} at $\Lspec$-level. 
-To do so, $\ccspec$ inserts a speculation barrier after branches, which is sufficient to harden the program against speculative leaks, since SPECTRE-PHT~\cite{kocher2019spectre} is the only speculative leak modeled in the semantics of $\Lspec$.
-
-\vspace{-1em}
-\[
-\begin{array}{cl}
-  &\ccspec{(\obj{\lifz{\expr[_0]}{\expr[_1]}{\expr[_2]}})} = 
-  \\
-  &\qquad\qquad \lifz[\ird]{\ccspec{\left(\obj{\expr[_0]}\right)}}{\ird{\lbarrier;}\ccspec{\left(\obj{\expr[_1]}\right)} \\&\qquad\qquad}{ \ird{\lbarrier;}\ccspec{\left(\obj{\expr[_2]}\right)}} 
-\end{array}
-\]
-% \vspace{-2em}
-%
-
-Clearly, the corresponding cross-language trace relation $\sim^{\Lscct}_{\Lspec}$ has only one non-trivial case: for branches, only relate them where speculation is blocked by a barrier:
-
-\begin{center}
-  \typerulenolabel{xrel:spec:if}{}{\obj{Branch\ n}\sim^{\Lscct}_{\Lspec}\ird{Branch\ n}\cdot\ird{Spec}\cdot\ird{Barrier}\cdot\ird{Rlb}}
-\end{center}
-
-The base-event relation above scales to full events by ensuring the missing annotations ($\obj{\comp;\securitytag{}}$ and $\ird{\comp;\securitytag{}}$) are the same.
-% 
-With this relation, we prove that $\ccspec$ is secure with respect to \gls*{ss}.
-\begin{theorem}[$\ccspec$ secure w.r.t. \gls*{ss}]\label{thm:ccspec:rtp:spec}
-  \small$\rtpsim{\ccspec}{\sssafe}{\sim^{\Lscct}_{\Lspec}}$ % \Coqed
-\end{theorem}
-
-\subsection{Robust Preservation of Memory Safety, Strict Cryptographic Constant Time, and Speculative Safety}
-
-Finally, this subsection combines all previous results into one compilation chain to get that it preserves full \gls*{specms}.
-Let $\ccspecms$ be the compiler that is the composition of $\cca$, $\ccb$, $\cccf$, $\ccdce$, $\ccscct$, and $\ccspec$. 
-Let $\sim^{\Ltms}_{\Lspec}$ be the composition of $\sim^{\Ltms}_{\Lms}$, $\sim^{\Lms}_{\Lscct}$, and $\sim^{\Lscct}_{\Lspec}$.
-Then, the following corollary holds.
-
-\begin{corollary}[$\ccspecms$ secure w.r.t. \gls*{specms}]\label{thm:ccall:rtp:specms}
-  $\;$
-
-  \begin{nscenter}
-    $\rtpsim{\cc{\Ltms}{\Lspec}}{\mssafe\cap\scctsafe\cap\sssafe}{\sim^{\Ltms}_{\Lspec}}$ % \Coqed
-  \end{nscenter}
-\end{corollary}
-
-As with \Cref{corr:ccab:rtp:ms}, it is important to ensure that the respective cross language trace relations are well-formed (\Cref{def:wfc:sig:tracerel}).
-It is already known that $\sim^{\Ltms}_{\Lms}$ is well-formed with respect to $\mssafe$ (\Cref{lem:wf:ltmsltrg}).
-Next in the chain is $\sim^{\Lms}_{\Lscct}$, which has to be well-formed w.r.t. $\scctsafe$.
-This lemma holds, since a trace that was $\scctsafe$ is $\scctsafe$ even after applying $\sim^{\Lms}_{\Lscct}$: the relation enforces that $\Lscct$ traces related to $\Lms$ traces have no leaks of secrets whatsoever.
-
-\begin{lemma}[$\sim^{\Lms}_{\Lscct}$ well-formed w.r.t. $\scctsafe$]\label{lem:wf:lsmslscct}
-$\;$
-
-  \begin{nscenter}
-  $\wfcsig{\sim^{\Lms}_{\Lscct}}{\scctsafe}$
-  \end{nscenter}
-\end{lemma}
-
-The last relation is $\sim^{\Lscct}_{\Lspec}$ which needs to be well-formed w.r.t. $\sssafe$.
-Similarly to the previous relation, this holds, since $\sim^{\Lscct}_{\Lspec}$ only relates $\Lspec$ traces, which do not have speculative leaks, with $\Lscct$ traces.
-
-\begin{lemma}[$\sim^{\Lscct}_{\Lspec}$ well-formed w.r.t. $\sssafe$]\label{lem:wf:lscctlspec}
-  $\wfcsig{\sim^{\Lscct}_{\Lspec}}{\sssafe}$
-\end{lemma}
+\input{sections/compo.tex}
+\input{sections/prop.tex}
+\input{sections/langs.tex}
+\input{sections/csc.tex}
 

### 2025-02-13 icfp/sections/compo.tex
move

--- 2024-09-26 csf2/main.tex
@@
-% Fixing: Too many math alphabets used in version normal
-\newcommand\hmmax{0}
-\newcommand\bmmax{0}
-\documentclass[dvipsnames,conference]{IEEEtran}
-
-\usepackage[T1]{fontenc}
-\usepackage[scaled=.83]{beramono}
-
-\usepackage[colorinlistoftodos]{todonotes}
-\usepackage[inference]{semantic}
-\usepackage[switch]{lineno}
-\usepackage{halloweenmath}
-\usepackage{fontawesome5}
-\usepackage{listofitems}
-\usepackage{breakcites}
-\usepackage{glossaries}
-\usepackage{hyperref}
-\usepackage{cleveref}
-\usepackage{stmaryrd}
-\usepackage{marvosym}
-\usepackage{listings}
-\usepackage{amssymb}
-\usepackage{amsmath}
-\usepackage{nameref}
-\usepackage{amsthm}
-\usepackage{xspace}
-\usepackage{xfrac}
-\usepackage{tikz}
-\usepackage{soul}
-\usepackage{bm}
-\usepackage{paralist}
-
-\hypersetup{
-    colorlinks,
-    linkcolor={red!50!black},
-    citecolor={blue!50!black},
-    urlcolor={blue!80!black}
-}
-%\renewcommand\UrlFont{\color{blue}\rmfamily}
-%\newcommand{\url}[1]{\lstinline{#1}}
-
-\setul{0.95ex}{0.3ex}
-
-\input{cmds.tex}
-
-% allow page breaks inside multiline alignment displays
-\allowdisplaybreaks
-
-\Crefname{exampleenv}{Example}{Examples}
-
-\theoremstyle{definition}
-\newtheorem{exampleenv}{Example}[section]
-\newtheorem{lemma}{Lemma}[section]
-\newtheorem{theorem}{Theorem}[section]
-\newtheorem{corollary}{Corollary}[section]
-\newtheorem{definition}{Definition}[section]
-
-\loadglsentries{acronyms}
-\makeglossaries
-
-\AtBeginDocument{%
-  \providecommand\BibTeX{{%
-    \normalfont B\kern-0.5em{\scshape i\kern-0.25em b}\kern-0.8em\TeX}}}
-\makeatletter
-\newcommand{\linebreakand}{%
-  \end{@IEEEauthorhalign}
-  \hfill\mbox{}\par
-  \mbox{}\hfill\begin{@IEEEauthorhalign}
-}
-\makeatother
-
-\renewcommand\thelinenumber{\color{red}\arabic{linenumber}}
-\begin{document}
-\linenumbers
-
-\title{
-  Secure Composition 
-  \\ 
-  of Robust and Optimising Compilers
-  \vspace{-1.5em}
-}
-
-% \author{
-% Matthis Kruse\orcidID{0000-0003-4062-9666}
-% \and
-% Michael Backes
-% \and
-% Marco Patrignani
-% }
-% \authorrunning{M. Kruse et al.}
-
-\maketitle
-
-\begin{abstract}
-% \MPpost{
-%   this abstract fails in setting up the robustness argument.
-%   it'll be a good exercise for you to take inspiration from this, but then write something as concise but with a more direct focus on robust properties
-% }
-% context
-To ensure that secure applications do not leak their secrets, they are required to uphold several security properties such as spatial and temporal memory safety, cryptographic constant time, as well as speculative safety.
-% need
-Existing work shows how to enforce these properties individually, in an architecture-independent way, by using secure compiler passes that each focus on an individual property.
-% task
-Unfortunately, given two secure compiler passes that each preserve a possibly different security property, it is unclear what kind of security property is preserved by the composition of those secure compiler passes.
-%there is no way to tell what kind of security property will the composition of those secure compilers preserve.
-
-% object
-This paper is the first to study what security properties are preserved across the composition of different secure compiler passes.
-% findings
-Starting from a general theory of property composition for security-relevant properties (such as the aforementioned ones), this paper formalises a theory of composition of secure compilers.
-Then, it showcases this theory on a secure multi-pass compiler that preserves the aforementioned security-relevant properties.
-% conclusion
-Crucially, this paper derives the security of the multi-pass compiler from the composition of the security properties preserved by its individual passes, which include security-preserving as well as optimisation passes.
-% 
-From an engineering perspective, this is the desirable approach to building secure compilers.
-% \begin{center}\small\it
-% 	{This paper uses syntax highlighting accessible to both colourblind and black \& white readers.
-% 	% ~\cite{patrignani2020use}.
-% % 	% Specifically, it makes use of a $\src{blue}$, $\src{sans\text{-}serif}$ font for a $\src{source}$,
-% % 	% a $\trg{red}$, $\trg{bold}$ font for an $\trg{intermediate}$,
-% % 	% and a $\obj{green}$, $\obj{teletype}$ font for a $\obj{target}$ language.	
-% 	For a better experience, please print or view in colours.
-% 	}
-% \end{center}
-\end{abstract}
-
-% \begin{IEEEkeywords}
-%   memory-safety, cryptographic constant-time, secure compilation, robustness
-% \end{IEEEkeywords}
-
-\section{Introduction\pages{4}}\label{sec:introduction}
-
-% \gls*{ms} is a security property that ensures that there are no out-of-bounds accesses as well as no reads or writes to uninitialised memory.
-% Programs that do not attain memory safety, can easily have their private data be leaked to attackers or be corrupted~\cite{lemay2021ccc}.
-% Because of this, \gls*{ms} is a requirement for programs to attain an even stronger security property, namely \gls*{cct}.
-% Programs that attain \gls*{cct} are secure w.r.t.
-timing attacks, and this is the gold standard for secure applications, including cryptographic ones, since timing attacks can expose secrets via a simple performance analysis~\cite{kocher1996timing}.
-% Timing attacks have become more and more relevant recently, especially given the era of cloud-computing~\cite{aviram2010cloudtime,kumar2019cloudsecsurvey}, where programs share the same hardware but are supposed to run in a sandboxed environment that does not leak secrets.
-% Unfortunately, several attacks within such a cloud environment have been demonstrated as viable strategies to extract secrets~\cite{mehmet2015getoff,flowers2022zeroday,atya2019catchme,venkatanathan2015placevul}.
-
-% Even though mitigations for such attacks exist~\cite{bond2017vale,almeida2017jasmin}, some compilers may optimise these mitigations away~\cite{barthe2018sec}.
-% This is of course not what programmers expect and to ensure such mitigations persist in the compiled code, one can use a {\em secure} compilers.
-% More generally, secure compilers should preserve security properties even when the program runs in a hostile context such as the cloud one, i.e., where it interacts with other possibly malicious programs.
-% An example of such secure compilers are those that ascribe to the theory of robust compilation~\cite{abate2019jour,abate2021extacc,patrignani2021rsc}, which takes this interaction between components and potentially malicious contexts into account.
-
-% Unfortunately, it is unclear if and how robust compilation supports compositionality, and compilation passes are typically developed as a standalone module to be plugged into a compilation pipeline and composed with other passes.
-% Different kinds of compilation passes exist, e.g., there are usually several compilation passes that perform optimisations~\cite{lattner2004llvm,googlev8,androidstudio,kuepper2023cryptopt,manjikian1997fusion,wegman1991ccp} or source-code instrumentations~\cite{nagarakatte2009soft,nagarakatte2010cets,akritidis2009baggy,dhumbumroong2020boundwarden,jung2021pico,nam2019framer,shankaranarayana2023tailcheck,younan2010paricheck,zhou2023fatptrs,bond2017vale,almeida2017jasmin,kuepper2023cryptopt,cauligi2019fact} that enforce that the compiled program fulfils a property of interest.
-% Without a framework for composing secure compiler passes, it is not possible to enable separation of concerns, e.g., to have a secure compilation pass that ensures \gls*{ms} that is developed independently of another secure pass for \gls*{cct} as suggested before and as showcased in \Cref{ex:strncpy}.
-% \MPpost{
-%   this intro does a poor job at introducing the robust compilation argument.
-%   perhaps it'd be best to start from robust properties, which motivates the use of robust compilers?
-%   idk
-%   as for the abstract, this is a good starting point for you to make a better intro
-% }
-
-\gls*{ms} is a security property obtained by composing \gls*{sms}, which ensures array accesses are all within bounds, and \gls*{tms}, which ensures pointers are only used when they are valid~\cite{azevedo2018meaningsofms,jim2002cyclone,necula2005ccured,nagarakatte2010cets,nagarakatte2009soft,akritidis2009baggy,michael2023mswasm}.
-\gls*{cct} is a security property that ensures sensitive data is not leaked via timing side-channels~\cite{kocher1996timing} and \gls*{ss} is a security property that enforces the same but under a speculative semantics~\cite{guarnieri2018spectector,fabian2022automatic} that captures speculative execution attacks such as spectre~\cite{kocher2019spectre}.
-Together, \gls*{sms}, \gls*{tms}, \gls*{cct} and \gls*{ss}, yield \gls*{specms}, which is the gold standard of security properties for secure applications.
-Programs attaining \gls*{specms} do not leak sensitive data either through erroneous memory accesses, nor through timing side-channels, even under speculative execution.
-\Cref{ex:strncpy} below discusses how these security properties can be enforced by compiler passes~\cite{bond2017vale,almeida2017jasmin}, to ensure programmers need not be aware of the architectural details of where their code will run.
-
-\begin{exampleenv}[strncpy]\label{ex:strncpy}
-Consider the following \texttt{C} function \texttt{strncpy} that copies a null-terminated string \texttt{src} into \texttt{dst} up to a length of \texttt{n} characters.
-This function is subject to a subtle \gls*{sms} vulnerability: the bounds check \texttt{i < n} should happen {\it before} the access to memory location \texttt{src[i]}: otherwise
-% If \texttt{n} is at least as large as the length of the array \texttt{src}, 
-the memory location past the last element is leaked to an attacker.
-\begin{lstlisting}[language=c,basicstyle=\small\ttfamily,morekeywords={size_t}]
-void strncpy(size_t n, char *dst, char *src) {
-  for(size_t i = 0; src[i] != '\0' && i < n; ++i) {
-    dst[i] = src[i];
-  }
-}
-\end{lstlisting}
-% Assume that the length of the array pointed by \texttt{dst} is at least \texttt{n}, and \texttt{src} is terminated by \texttt{'\textbackslash 0'}.
-% Even under these assumptions, there is a 
-To prevent this vulnerability, one can use a compilation pass that enforces \gls*{sms}, such as Softbounds~\cite{nagarakatte2009soft} or BaggyBounds~\cite{akritidis2009baggy}.
-The most na\"ive solution in this case is to insert bounds-checks in front of every access to memory.
-
-Because of timing attacks, addressing \gls*{sms} is not enough to make \texttt{strncpy} secure.
-In fact, the loop can terminate early, as soon as the string-terminating character \texttt{'\textbackslash 0'} is encountered, thus making program execution time proportional to the length of the array pointed by \texttt{src}, and violating the \gls*{cct} property.
-% Often, programmers rewrite programs like these to have loops that run with a fixed number of iterations.
-Also in this case there exist compiler passes that can rewrite such programs into \gls*{cct} ones~\cite{cauligi2019fact}.
- % so that they are no longer vulnerable to timing attacks.
-%
-Finally, even with these precautions, code is not run in isolation, so a malicious attacker could supply code that interacts with \texttt{strncpy} and trigger a violation of either \gls*{ms} or \gls*{cct} by calling \texttt{strncpy} with an argument for \texttt{src} that points to uninitialised memory.
-This would, in turn, trigger a series of reads from uninitialised memory, which is an immediate \gls*{ms} violation with devastating real-world consequences~\cite{uninit-0,uninit-1,uninit-2,uninit-3,uninit-4}.
-% An easy fix to sidestep this is to use non-nullable pointers as parameters instead, e.g., as presented in Cyclone~\cite{jim2002cyclone}.
-\end{exampleenv}
-
-  % last sentence is before the example, which talks about compiler passes enforcing props
-Whether or not compiler passes enforce certain security properties, it is important for a secure compiler to consider partial programs that interact with potentially malicious code, since the latter may lead to, e.g., memory-safety issues in the considered partial program.
-Robust compilers~\cite{abate2019jour} are a form of secure compilers that preserve security properties even in the presence of arbitrary attackers interacting with compiled code.
-Thus, robust compilers can be used to prevent vulnerabilities resulting from uninitialised memory (as well as many other ones), e.g., by targeting capability-based languages such as CHERI~\cite{woodruff2014CHERI}, Arm Morello~\cite{arm-morello}, or MSWasm~\cite{michael2023mswasm}, where the compiler relies on capabilities to check that pointers are always initialised.
-
-% To ensure these compiler-preserved properties still hold even in the presence of arbitrary attackers interacting with compiled code, one can use secure compilers in the form of robust compilers~\cite{abate2019jour,abate2021extacc,patrignani2021rsc}.
-
-Unfortunately, given secure compiler passes that each preserve a possibly different security property, there is no way to tell what kind of security property will the composition of those secure compilers preserve.
-% Different kinds of compilation passes exist, e.g., there are usually several compilation passes that perform optimisations~\cite{lattner2004llvm,googlev8,androidstudio,kuepper2023cryptopt,manjikian1997fusion,wegman1991ccp} or source-code instrumentations~\cite{nagarakatte2009soft,nagarakatte2010cets,akritidis2009baggy,dhumbumroong2020boundwarden,jung2021pico,nam2019framer,shankaranarayana2023tailcheck,younan2010paricheck,zhou2023fatptrs,bond2017vale,almeida2017jasmin,kuepper2023cryptopt,cauligi2019fact} that enforce that the compiled program fulfils a property of interest.
-Worse, without a framework for composing secure compiler passes, it is not possible to enable separation of concerns, e.g., to have a secure compilation pass that ensures \gls*{ms} that is developed independently of another secure pass for \gls*{cct}, that is developed independently of other passes, such as optimisation ones.
-
-This paper introduces a framework for reasoning about the composition of secure and optimising compiler passes %akin to those of \Cref{ex:strncpy}\MP{there aren't really passes there} 
-and it showcases the power of this framework by instantiating it on a multi-pass compilation chain.
-To this end, this paper first discusses how to compose security properties, such as \gls*{tms} and \gls*{sms} into \gls*{ms}, and then adds \gls*{cct} as well as \gls*{ss} into the mix to obtain \gls*{specms}.
-%Then, this paper defines compiler composition and formalises that given two passes that securely preserve two (possibly distinct) properties, their composition securely preserves the composition of those properties.
-The paper then defines several secure compiler passes, where each is either preserving a different security property from the list above
-% (\gls*{tms}, \gls*{sms}, \gls*{scct}, \gls*{ss}) 
-or performing a security-preserving optimisation.
-% , (e.g., applying \gls*{cf} or \gls*{dce}).
-Finally, this paper shows that composing these secure compiler passes into a multi-pass compilation chain results in the end-to-end (robust) preservation of \gls*{specms}.
-Crucially, this paper derives the security of the multi-pass compiler from the composition of the security properties preserved by its individual passes.
-% Crucially, this end-to-end security preservation is obtained by composing the security property preserved by each individual pass, as dictated by the framework of this paper.
-This result showcases how the framework allows the kind of formal security reasoning that compiler writers already want (and already do), obtaining precise, compositional security reasoning while providing minimal (and modular) proof effort.
-%
-In summary, this paper makes the following contributions:
-% \begin{itemize}[leftmargin=0em,nosep,label=$\blacktriangleright$]
-\begin{asparaitem}[$\blacktriangleright$]
-  \item %
-        %\Cref{sec:sequential} studies different forms of secure compiler compositions.
-        This paper takes the secure compilation framework of Abate et al.~\cite{abate2021extacc} and extends it for compositionality. (\Cref{sec:sequential}).
-        %Concretely, it investigates upper and lower compositions, as well as the sequential composition.
-        %For this, this paper studies compiler composition as well as compilers with multiple input languages or multiple output ones, as used in existing compilation chains.
-        %
-        This paper proves that starting from two compilers that preserve two (possibly distinct) properties, their composition preserves roughly the intersection of those properties.
-        % For example, a compiler that is secure w.r.t. \gls*{tms} can be composed with a compiler that is secure w.r.t. \gls*{sms}, resulting in a compiler that is secure w.r.t. \gls*{ms}.
-        % 
-        % From this, the paper proves that the order of composition of sequential compiler passes is irrelevant for security.
-        % % 
-        % This is crucial for reordering individual optimisation passes and thus generating secure and efficient code.
-        %
-        %Finally, this paper extends parts of the compositionality framework to support different trace-models between compilation passes.
-        % \MP{this isn't really true, is it?. the order matters for the relations' lemmas}
-        Then, this paper identifies which conditions make the composition of secure compilers meaningful from a security perspective, and which other conditions allow passes to be swapped without losing security meaning.
-
-  \item %
-        This paper presents a case-study with five programming languages showcasing the previous contribution (\Cref{sec:casestud:defs,sec:casestud:rtp}).
-        To this end, it presents a compilation chain consisting of six passes that ultimately preserve \gls*{specms} by means of composing four secure passes that \emph{individually} preserve \gls*{tms}, \gls*{sms}, \gls*{scct}---an enforceable version of \gls*{cct}---, and \gls*{ss}, and two passes which are well-known optimisations \gls*{dce} and \gls*{cf}.
-        %Furthermore, the chain includes two optimisation passes: one performs \gls*{dce} and the other \gls*{cf}.
-        The formalisation of this case study showcases the power of the presented framework: the divide-and-conquer approach to software engineering is a viable strategy even for the development of secure compilers.
-%        This demonstrates that the usage of the presented framework is both simpler and more expressive compared to one monolithic security proof of the whole compilation chain.
-
-  \item %
-  		This paper provides formal proofs of the security properties preserved by each of the presented passes. 
-  		% 
-  		Additionally this paper analyses each pass in terms of their compatibility with each other (\Cref{sec:formalities}), proving how does each pass fulfill the conditions that make their composition meaningful from a security standpoint.
-        % In doing so, the paper provides formal insights into some technicalities of the concrete proofs necessary as technical setup to reap the benefits of the presented compositionality framework.
-
-  \item The key contributions of this paper are formalised in the Coq proof assistant and the paper indicates this with \CoqSymbol.
-\end{asparaitem}
-% 
-This paper starts by introducing relevant notions of security properties and secure compilation (\Cref{sec:background}),
-and discusses related work (\Cref{sec:relwork}) before concluding (\Cref{sec:concl}).
-
-% \contribution{Technical Report} 
-The omitted formal details, lemmas and proofs, as well as the Coq formalisation are available as supplementary material.
-% \MKin{cite/link}
-
-\section{Background: Properties and Secure Compilers\pages{1}}\label{sec:background}
-
-To introduce the security argument of this paper, this section defines (security) properties, their satisfaction, and their robust satisfaction (i.e., satisfaction in the presence of an active attacker; \Cref{subsec:bg:tprop}).
-Then, borrowing from existing work~\cite{abate2019jour,abate2021extacc}, this section introduces secure compilers as compilers that preserve robust property satisfaction (\Cref{subsec:bg:rtp}).
-
-\subsection{Properties and (Robust) Satisfaction}\label{subsec:bg:tprop}
-
-This paper employs the security model where programs are written in a language whose semantics emits events $\event$.
-Events include security-relevant actions (e.g., reading from and writing to memory, as detailed in \Cref{sec:compprop}) and the unobservable event $\emptyevent$.
-As programs execute, their emitted events are concatenated in traces $\trace$, which serve as the description of the behaviour of a program.%
-\footnote{
-Throughout the paper, sequences are indicated with an overbar (i.e., $\trace$), empty sequences with $\hole{\cdot}$, and concatenation of sequences $\trace[_{1}],\trace[_{2}]$ as $\trace[_{1}]\cdot\trace[_{2}]$.
-Prepending element $\event$ to a sequence $\trace$ uses the same notation: $\event\cdot\trace$.
-}
-
-Properties $\pi$ are sets of traces of admissible program behaviours, ascribing what said property considers valid.
-% 
-The set of all properties can be partitioned into different {\em classes} ($\class$), i.e., safety, liveness, and neither safety nor liveness~\cite{clarkson2008hyper}, so a class is simply a set of properties.
-% 
-The compositionality framework (\Cref{sec:sequential}) presented in this paper considers arbitrary classes, while the case-study (\Cref{sec:casestud:rtp}) fixes them to concrete instances of safety properties, since it is decidable whether a trace satisfies a safety property with just a finite trace (i.e., a \emph{prefix}).
-As an example, consider the trace:
-$$\ev{Dealloc\ \loc}\cdot\ev{Read\ \loc\ 1729}\cdot\dots$$
-which describes the interaction with a memory where the deallocation of an address $\loc$ precedes a read (of some value $1729$) at that address in memory.
-% 
-This program behaviour is insecure w.r.t.
-a canonical notion of (temporal) memory safety dictating no use-after-frees of pointers~\cite{nagarakatte2010cets,azevedo2018meaningsofms}, because it reads from a memory location that was freed already.
-The previous finite trace prefix is enough to decide that the trace does not satisfy \gls*{tms} and there is no way to append events to this prefix which would result in the trace being admissible.
-
-In the following, the execution of a whole program $\wholeprogvar$ that terminates in state $\runtimetermvar$ according to the language semantics and produces trace $\trace$ is written as $\progstepto{\wholeprogvar}{\runtimetermvar}{\trace}$.
-With this, we defined property satisfaction as follows:
-\bul{a whole program $\wholeprogvar$ satisfies a property $\pi$} iff \iul{$\wholeprogvar$ yields a trace $\trace$} such that \oul{$\trace$ satisfies $\pi$}.
-
-\begin{definition}[Property Satisfaction]\label{def:propsat}
-$\;$
-
-  \begin{nscenter}
-    \bul{$\sat{\wholeprogvar}{\pi}$}
-    $\isdef$
-    if \iul{$\forall\runtimetermvar\ \trace,\progstepto{\wholeprogvar}{\runtimetermvar}{\trace}$},
-    then \oul{$\trace\in\pi$}.
-  \end{nscenter}
-\end{definition}
-
-Property satisfaction is defined on whole programs, i.e., programs without missing definitions.
-Thus, from a security perspective, \Cref{def:propsat} considers only a passive attacker model, where the attacker observes the execution and, e.g., retrieves secrets from that.
-To consider a stronger model similarly to what existing work does~\cite{abate2019jour,abate2021extacc,maffeis2008code-carrying,gordon2003authenticity,fournet2007authorization,bengtson2011refine,backes2014uniontyps,michael2023mswasm,swasey2017robust,sammler2019benefits}, we extend the concept of satisfaction with {\em robustness}.
-Robust satisfaction considers partial programs $\progvar$, i.e., components with missing imports, which cannot run until said imports are fulfilled.
-To remedy this, {\em linking} takes two partial programs $\progvar[_{1}],\progvar[_{2}]$ and produces a whole program $\wholeprogvar$, i.e., $\link{\progvar[_{1}]}{\progvar[_{2}]}=\wholeprogvar$.
-As typically done in works that consider the execution of partial programs~\cite{abate2019jour,devriese2018parametricity,patrignani2021rsc,korashy2021capableptrs,strydonck2019lincap,devriese2017modular,bowman2015noninterference,ahmed2011equivcps,patterson2017linkingtyps},
-this paper assumes that whole programs are the result of linking partial programs referred to as {\em context} ($\ctx$) and {\em component} ($\comp$).
-The context is an arbitrary program and thus has the role of an {\em attacker} that can interact with the component by means of any features the programming language has, and the component is what is security-relevant.
-%In this work, the semantics of the programming language is expected to differentiate between the component and the context.
-With this, \Thmref{def:propsat} can be extended as follows: for \bul{a component $\progvar$ to robustly satisfy a property $\pi$}, take an \iul{attacker context $\contextvar$ and link it with $\progvar$}, \oul{the resulting whole program must satisfy $\pi$}.
-
-\begin{definition}[Robust Satisfaction]\label{def:proprsat}
-  $\;$
-
-  \begin{nscenter}
-  \bul{$\rsat{\progvar}{\pi}$}
-  $\isdef$ \iul{$\forall \contextvar\ \wholeprogvar$, if $\link{\contextvar}{\progvar}=\wholeprogvar$}, then \oul{$\sat{\wholeprogvar}{\pi}$}.
-  \end{nscenter}
-\end{definition}
-
-% \begin{exampleenv}[Double Free in Bluetooth Subsystem]
-%   Consider \texttt{CVE-2021-3564}~\cite{doublefree-bluetooth}, one of many submissions for a double-free vulnerability.
-%   The vulnerability arises due to a race condition where the context-level function \texttt{hci\_cmd\_work} was not expected to behave maliciously, since it resides in the same source-code repository where the vulnerability occurs.
-%   Nevertheless, the component-level code of \texttt{hci\_dev\_do\_open} is linked with \texttt{hci\_cmd\_work} and does not atomically check whether a pointer has been freed already:
-%   Therefore, \texttt{hci\_dev\_do\_open} does not satisfy the no-double-frees property robustly, since there is an implementation for \texttt{hci\_cmd\_work} that leads to a violation of that property when linked with \texttt{hci\_dev\_do\_open}.
-% \end{exampleenv}
-
-\subsection{Secure Compilers}\label{subsec:bg:rtp}
-
-A {\em compiler} ($\cc{\src{L}}{\trg{L}}$) translates syntactic descriptions of programs from a {\em source} ($\src{L}$) into a {\em target} ($\trg{L}$) programming language.
-This translation is considered {\em correct} if it is semantics-preserving.
-That is, for a whole program $\src{\wholeprogvar}$, the compiler should relate the $\src{L}$ semantics of $\src{\wholeprogvar}$ with the semantics of $\trg{L}$ of the compiled counterpart of $\src{p}$ in such a way that they are ``compatible''.
-Unfortunately, correct compilers may be insecure compilers~\cite{patrignani2019survey,kennedy2006secure.net,abadi1999protect,ahmed2018dagstuhl} and programs translated by insecure compilers can violate security properties that the programmer assumes to hold.
-This is why {\em robust preservation} is a good candidate as a compiler-level security property~\cite{abate2019jour}.
-
-This paper uses a general notion of robust preservation~\cite{abate2021extacc} that considers compilers that use languages with different trace models. 
-To this end, considering a source trace $\src{\trace}$ and a target trace $\trg{\trace}$, there is a relation ($\sim$) describing the effect of a corresponding compiler (see \Cref{sec:casestud:rtp}). 
-This relation induces the following two projection functions~\cite{abate2021extacc}: (1) the \emph{existential image} $\tau_\sim\left(\src{\pi}\right)$ and (2) the \emph{universal image} $\sigma_\sim\left(\trg{\pi}\right)$.
-These projections map source-level (resp. target-level) properties to target-level (resp. source-level) properties in a way that identifies the ``same'' property across languages. 
-The case study of this paper uses the universal image, since some considered properties, such as \gls*{ss}, are not definable in a higher-level language that, e.g., does not model speculation.
-% For self-containment, the universal image is defined as follows~\cite{abate2021extacc}:
-% 
-\begin{definition}[Universal Image]
-% [Existential and Universal Image]
-\label{def:universal:img}\label{def:sigma}
-  % \[
-  %   \tau_\sim\left(\src{\pi}\right) := 
-  %     \left\{ 
-  %       \trg{\trace} \mid \exists \src{\trace}\ldotp \src{\trace}\sim\trg{\trace}, \text{ and } \src{\trace}\in\src{\pi} 
-  %     \right\}
-  % \]
-  \[ 
-    \sigma_\sim\left(\trg{\pi}\right) := 
-      \left\{ 
-        \src{\trace} \mid \forall \trg{\trace}\ldotp \text{if }\src{\trace}\sim\trg{\trace}, \text{ then } \trg{\trace}\in\trg{\pi} 
-      \right\}
-  \]
-\end{definition}
-\noindent
-With this projection function, we define a more general version of robust preservation as follows~\cite{abate2021extacc}.
-% For a \bul{compiler $\cc{\src{L}}{\trg{L}}$ to robustly preserve a class of source properties $\src{\class}$}, if for any \rul{property $\src{\pi}$ of that class $\src{\class}$ and programs $\src{p}$ written in $\src{L}$} where \iul{$\src{\progvar}$ robustly satisfies $\src{\pi}$}, then \oul{the compilation of $\src{\progvar}$, $\cc{\src{L}}{\trg{L}}\left(\src{p}\right)$, must robustly satisfy $\tau_\sim\left(\src{\pi}\right)$}.
-% 
-% \begin{definition}[Robust Preservation with $\tau_\sim$]\label{def:rtp:tau}
-%   % $\;$\\
-%   % \vspace{-1em}
-%   % \begin{nscenter}
-%   % Compiler $\cc{\src{L}}{\trg{L}}$ robustly preserves $\class$, 
-%   \bul{$\rtptau{\cc{\src{L}}{\trg{L}}}{\src{\class}}{\sim}$}
-%   %, iff 
-%   $\isdef$
-%   \rul{$\forall \src{\pi}\in\src{\class}, \src{p}\in\src{L},$} if \iul{$\rsat{\src{\progvar}}{\src{\pi}}$}, then \oul{$\rsat{\cc{\src{L}}{\trg{L}}\left(\src{p}\right)}{\tau_\sim{\left(\src{\pi}\right)}}$}.
-%   % \end{nscenter}
-% \end{definition}
-% 
-A \bul{compiler $\cc{\src{L}}{\trg{L}}$ robustly preserves a class of target properties $\trg{\class}$}, if for any \rul{property $\trg{\pi}$ of class $\trg{\class}$ and programs $\src{p}$}, where \iul{$\src{\progvar}$ robustly satisfies $\sigma_\sim\left(\trg{\pi}\right)$}, \oul{the compilation of $\src{\progvar}$, we have that $\cc{\src{L}}{\trg{L}}\left(\src{p}\right)$ robustly satisfies $\trg{\pi}$}.
-
-\begin{definition}[Robust Preservation with $\sigma_\sim$]\label{def:rtp:sigma}
-  % $\;$\\
-  % \vspace{-1em}
-  % \begin{nscenter}
-  % Compiler $\cc{\src{L}}{\trg{L}}$ robustly preserves $\class$, 
-  \bul{$\rtpsig{\cc{\src{L}}{\trg{L}}}{\trg{\class}}{\sim}$}
-  %, iff 
-  $\isdef$
-  \rul{$\forall \trg{\pi}\in\trg{\class}, \src{p}\in\src{L},$} if \iul{$\rsat{\src{\progvar}}{\sigma_\sim\left(\trg{\pi}\right)}$}, then \oul{$\rsat{\cc{\src{L}}{\trg{L}}\left(\src{p}\right)}{\trg{\pi}}$}.
-  % \end{nscenter}
-\end{definition}
-Note that a class of properties $\class$ can represent just one property $\pi$ by lifting~\cite{clarkson2008hyper} that property to sets of properties, i.e., use the powerset of $\pi$ instead of $\pi$ itself.
-Because of this, this paper may write $\rtptau{\cc{\src{L}}{\trg{L}}}{\trg{\pi}}{\sim}$, even though $\trg{\pi}$ is a property and not a class.
-A similar construction can be used to the projection functions (see \Cref{def:universal:img}) by applying them to the lifted version of $\trg{\pi}$ instead of just $\trg{\pi}$.
-
-In case the trace model is the same for both source and target programs (and thus $\sim$ is equality), we obtain~\cite{abate2019jour}:
-\begin{definition}[Robust Preservation]\label{def:rtp}
-  % $\;$\\
-  % \vspace{-1em}
-  % \begin{nscenter}
-  % Compiler $\cc{\src{L}}{\trg{L}}$ robustly preserves $\class$, 
-  $\;$
-
-  {$\rtp{\cc{\src{L}}{\trg{L}}}{\class}$}
-  %, iff 
-  $\isdef$
-  {$\forall \pi\in\class, \src{p}\in\src{L},$} if {$\rsat{\src{\progvar}}{\pi}$}, then {$\rsat{\cc{\src{L}}{\trg{L}}\left(\src{p}\right)}{\pi}$}.
-  % \end{nscenter}
-\end{definition}
-
-Examples of compilers fulfilling \Cref{def:rtp} exist in the literature~\cite{korashy2022secureptrs,korashy2021capableptrs,abate2021extacc,abate2019jour,patrignani2021rsc}.
-For example, SecurePtrs~\cite{korashy2022secureptrs} gives a compiler that robustly preserves all safety properties for a C-like language to an assembly-like language. 
-% The mechanization has well over 20k lines in total, which demonstrates that proving a compiler secure with respect to \Cref{def:rtp} is a significant effort. \MP{move this line to later ?}
-% i don't see how adding a def of cct is beneficial
-As another example, even though it is not strictly satisfying \Cref{def:rtp}, the FaCT~\cite{cauligi2019fact} compiler preserves the \gls*{cct} property for a C-like language with constant-time primitives, e.g., \texttt{ctselect} for branching.
-%, whose precise definition can be found in \Cref{sec:cct}.
-Throughout this work, it is assumed that FaCT satisfies \Cref{def:rtp}.
-% While we were not able to find compilers in literature that provably fulfill \Cref{def:rtp}, we anticipate that 
- % would satisfy this hyperproperty.
-% FaCT proves constant-timeness by means of overapproximation, i.e., instead of directly proving the hyperproperty, they fall back to a property that overapproximates the corresponding hyperproperty.
-
-%\smallskip%MP: the underline compacts the sections too much
-
 \section{Secure Composition}\label{sec:sequential}
@@
 \end{corollary}
-
-\section{Security Properties Formalisation \& Composition\pages{1.5}}\label{sec:compprop}
-
-This section introduces trace properties of interest for this paper: \gls*{tms}, \gls*{sms}, \gls*{ms}, \gls*{scct}, and \gls*{ss}.
-These properties are of practical importance (as mentioned in \Cref{sec:introduction}) and also of interest in the case study presented later (\Cref{sec:casestud:defs,sec:casestud:rtp}). 
-This section presents all of them, despite the fact that they are inspired by existing work, in order to showcase all that is required for a formal proof of security for a realistic compilation toolchain.
-%\MP{ say something about this in the rebuttal: only the defs are borrowed, the universal images and all that work is new}
-%\MK{ universal images are not new. }
-% , none of the properties is novel, for self-containment of the paper, their definition is necessary for the case study.
-% For each of the presented properties, the technical report defines corresponding monitors, which refine a given property and have a key role in the proofs of this paper. 
-The technical report defines monitors for each of the presented properties.
-% 
-Monitors refine each property and have a key tole in the proofs of this paper.
-% managed to prove 
-%  if tms-monitor --[trace]-->* ∅  
-%  then trace ∈ tms
-% but not
-%  if trace ∈ tms
-%  then tms-monitor --[trace]-->* ∅
-
-\subsection{A Trace Model for Memory Safety}\label{subsec:basic:memsafety:tracemodel}
-
-For simple memory safety composed of temporal and spatial memory safety, the trace model defines events ($\event[_{\mssafe}]$) as either the empty event ($\emptyevent$), a crash ($\lightning$), or a base-event ($\preevent[_{\mssafe}]$).
-
-\vspace{-1.0em}
-\[
-  \begin{array}{rcll}
-    \text{(Base-Events)}&\preevent[_{\mssafe}] &:=& \ev{Alloc \loc\ n} \mid \ev{Dealloc \loc} \mid \ev{Use \loc\ n} \\
-    \text{(Events)}&\event[_{\mssafe}] &:=& \preevent[_{\mssafe}] \mid \emptyevent \mid \lightning \\ 
-  \end{array}
-\]
-
-Base-events describe the actual kind of event that happened.
-For the basic memory-safety properties, these are three variants:
-First, the allocation event ($\ev{Alloc\ \loc\ n}$) that fires whenever a program claims $n$ cells of memory and stores them at address $\loc$, where addresses are assumed to be unique.
-Second, deallocation ($\ev{Dealloc\ \loc}$) announces that the object at location $\loc$ is freed.
-Third, an event to describe reads from and writes to the $n$-th memory cell from address $\loc$ ($\ev{Use\ \loc\ n}$).
-
-\subsubsection{Temporal Memory Safety}
-
-\gls*{tms}~\cite{nagarakatte2010cets} is a safety property that describes that an unallocated object must not be (re-)used.
-
-\begin{definition}[\glsfirst*{tms}]\label{def:trace:tmsdef}
-  % \footnotesize
-  $$
-  \tmssafe:=\left\{\trace[_{\mssafe}] \left| \begin{array}{rcl}
-    \ev{Alloc\ \loc\ n}&\le_{\trace[_{\mssafe}]}&\ev{Dealloc\ \loc} \\
-    \ev{Use\ \loc\ n}&\le_{\trace[_{\mssafe}]}&\ev{Dealloc\ \loc} \\
-    \text{at most one }\ev{Dealloc\ \loc}&\text{in}&\trace[_{\mssafe}] \\
-    \text{at most one }\ev{Alloc\ \loc\ n}&\text{in}&\trace[_{\mssafe}] \\
-  \end{array}\right.\right\}
-  $$
-\end{definition}
-Hereby, the notation $\event[_{1}]\le_{\trace}\event[_{2}]$ means that if $\event[_{1}]$ is in $\trace$ and if $\event[_{2}]$ is in $\trace$, then $\event[_{1}]$ appears before $\event[_{2}]$.
-
-\subsubsection{Spatial Memory Safety}
-
-\gls*{sms}~\cite{nagarakatte2009soft} is a safety property that prohibits out-of-bounds accesses.
-
-\begin{definition}[\glsfirst*{sms}]\label{def:trace:smsdef}
-% \small
-  % \begin{nscenter}
-
-  \noindent
-  \[
-  \smssafe:=\left\{\trace[_{\mssafe}] \left|\begin{array}{rcl}
-      \text{If }\ev{Alloc\ \loc\ n}\le_{\trace_{\mssafe}}\ev{Use\ \loc\ m}, \text{ then }m<n
-  \end{array}\right.\right\}
-  \]
-  % \end{nscenter}
-\end{definition}
-
-\subsubsection{Memory Safety}
-
-In spirit of earlier work~\cite{nagarakatte2009soft,nagarakatte2010cets,jim2002cyclone,necula2005ccured,michael2023mswasm}, full \gls*{ms} is the intersection of \Cref{def:trace:tmsdef,def:trace:smsdef}.
-
-\begin{definition}[\glsfirst*{ms}]\label{def:trace:msdef}
-  $
-  \mssafe:=\tmssafe \cap \smssafe
-  $
-\end{definition}
-
-Note that \Cref{def:trace:msdef} ignores data isolation, so there may still be memory-safety issues introduced by side-channels.
-
-\subsection{A Trace Model for Memory Safety with Constant Time}\label{subsec:scct:tracemodel}
-
-To express Constant Time, we extend the memory safety trace model with a {\em security tag} ($\securitytag{}$) that indicates whether events contains sensitive information ($\lock$) or not ($\unlock$).
-
-\vspace{-.5em}
-\[
-  \begin{array}{rrcl}
-    (\text{Base-Events}) & \preevent[_{\ctsafe}] &:=& \preevent[_{\mssafe}] \mid \ev{Branch\ }n \mid \ev{Binop\ }n\\
-    (\text{Security Tags}) & \securitytag{} &:=& \lock \mid \unlock\\ 
-    (\text{Events}) & \event[_{\ctsafe}] &:=& \preevent[_{\ctsafe}];\securitytag{} \mid \emptyevent \mid \lightning \\ 
-  \end{array}
-\]
-
-For cryptographic code, there is a general guideline that secrets must not be visible on a trace~\cite{ctguidelines}, i.e., secrets should not be marked as $\unlock$.
-In turn, an instruction whose timing is data-dependent must not have a secret as an operand.
-Typical operations with data-dependent timing are branches and certain binary operations, such as division.%
-\footnote{
-	This is architecture-dependent, but division is an operation that serves as a classic example for a data-dependent timing instruction~\cite[p.~755]{arm-refman}.
-}
-Both operations are represented in the trace model by extending the set of base-events with branches ($\ev{Branch\ n}$) and binary operations ($\ev{Binop\ n}$).
-
-\subsubsection{Strict Cryptographic Constant Time}
-
-\gls*{cct} is a hypersafety property~\cite{barthe2018sec} and, thus, difficult to check with monitors.
-This is because, intuitively, hypersafety properties can relate multiple execution traces with each other, but monitors work on a single execution.
-It is a common trick to sidestep this issue by means of overapproximation: this section defines the property \gls*{scct}, a stricter variant of \gls*{cct} (inspired by earlier work~\cite{almeida2017jasmin}) that enforces the policy that no secret appears on a trace.
-Programs that satisfy \gls*{scct} also satisfy \gls*{cct}, but programs that satisfy \gls*{cct} may not satisfy \gls*{scct}.
-
-\begin{definition}[\glsfirst*{scct}]\label{def:trace:scctdef}
-% \begin{nscenter}
-
-  \noindent\[
-  \scctsafe:=\left\{\trace[_{\ctsafe}] \left|\begin{array}{l}
-      \trace[_{\ctsafe}]=\hole{\cdot} \text{ or } \\\exists\trace[_{\ctsafe}'],\trace[_{\ctsafe}]=\preevent[_{\ctsafe}];\unlock\cdot\trace[_{\ctsafe}'] \wedge \trace[_{\ctsafe}']\in\scctsafe
-    \end{array}\right.\right\}
-  \]
-% \end{nscenter}
-\end{definition}
-
-\gls*{scct} may appear overly strict, since it seems that secrets must not occur on a trace (since $\securitytag{}$ is forced to be $\unlock$). 
-However, this is considered standard practice in terms of coding guidelines~\cite{ctguidelines}.
-Moreover, programs that have been compiled with FaCT~\cite{cauligi2019fact} and run with a ``data independent timing mode''~\cite{arm-refman,intel-refman} enabled do not leak secrets (see \Cref{ex:lscct}).
-
-\subsubsection{\gls*{ms}, Strict Cryptographic Constant Time}\label{sec:msscct-rel}
-
-The combination of \gls*{ms} and \gls*{scct} is the intersection of these properties, \gls*{msscct}.
-However, \gls*{ms} uses a different trace model than \gls*{scct}, so intersecting them would trivially yield the empty set. 
-To remedy this issue, we introduce $\sim_{\ctsafe}: \preevent[_{\ctsafe}] \times \preevent[_{\mssafe}] $, a cross-language trace relation (whose key cases are presented below), that we use to intuitively unify the trace model in which the two properties are expressed:
-\[
-  \typerulenolabel{mscctrel:drop:tag}{}{\preevent[_{\mssafe}]\sim_{\ctsafe}\preevent[_{\mssafe}];\unlock{}}
-  \typerulenolabel{mscctrel:drop:crash}{}{\lightning\sim_{\ctsafe}\lightning}
-\]
-\[
-  \typerulenolabel{mscctrel:drop:branch}{}{\emptyevent\sim_{\ctsafe}\ev{Branch n};\securitytag{}}
-  \typerulenolabel{mscctrel:drop:binop}{}{\emptyevent\sim_{\ctsafe}\ev{Binop n};\securitytag{}}
-\]
-
-Essentially, $\sim_{\ctsafe}$ ignores both the new $\ev{Branch\ n}$ and $\ev{Binop\ n}$ base-events as it relates security-insensitive actions ($\unlock$) to their equivalent counterparts.
-Thus, \gls*{ms} traces trivially satisfy \gls*{scct}.
-% 
-The above relation is extended point-wise to traces, skipping the empty event $\emptyevent$ on either side, and it is now possible to define \gls*{msscct} using the universal image:
-
-\begin{definition}[\gls*{ms} and \gls*{scct}]\label{def:trace:msscctdef}
-  $
-  \msscctsafe:=\mssafe\cap\sigma_{\sim_{\ctsafe}}\left(\scctsafe\right)
-  $
-\end{definition}
-
-\subsection{Extending the Trace Model with Speculation}\label{subsec:msctss:tracemodel}
-
-So far, the considered trace models do not let us express speculative execution attacks such as Spectre~\cite{kocher2019spectre}. 
-For this, we extend the earlier trace model (see \Cref{subsec:scct:tracemodel}) so that the security tags ($\securitytag{}$) carry additional information about the kind of private data leakage, i.e., the type of speculative leak.
-Moreover, we add base-events signalling the beginning of a speculative execution ($\ev{Spec}$), a barrier ($\ev{Barrier}$) that signals that any speculative execution may not go past it, as well as a rollback event ($\ev{Rlb}$), which signals that execution resumes to where speculation started.
-
-\vspace{-1em}
-{
-\[
-  \begin{array}{rrcl}
-    (\text{Base-Events}) & \preevent[_{\mathghost}] &:=& \preevent[_{\ctsafe}] \\
-    (\text{Spectre Variants}) & vX &:=& \operatorname{NONE} \mid \operatorname{PHT} \\
-    (\text{Security Tags}) & \securitytag{} &:=& \lock_{vX} \mid \unlock\\ 
-    (\text{Events}) & \event[_{\mathghost}] &:=& \preevent[_{\mathghost}];\securitytag{} \mid \emptyevent \mid \lightning \mid \ev{Spec} \mid \ev{Rlb} \mid \ev{Barrier}\\ 
-  \end{array}
-\]
-}
-
-Even though the considered Spectre variants are just SPECTRE-PHT~\cite{kocher2019spectre}, NONE just describes secret data as in \gls*{scct} (see \Cref{subsec:scct:tracemodel}), the trace model is general enough to allow for potential future extension with different variants~\cite{kocher2019spectre,maisuradze2018ret2spec,horn2019zero}.
-%For sake of readability, this paper just uses the notation $\lock$ in place of $\lock_{\text{PHT}}$.
-
-\subsubsection{Speculative Safety}
-
-\gls*{ss}~\cite{patrignani2021exorcising}, similar to \gls*{scct}, is a sound overapproximation of a variant of noninterference.
-
-\begin{definition}[\glsfirst*{ss}]\label{def:trace:ss}
-  \noindent
-
-  \begin{nscenter}
-  $
-    \sssafe:=\left\{\trace[_{\mathghost}] \left|\begin{array}{l}
-      \trace[_{\mathghost}]=\hole{\cdot} \text{ or } \exists\trace[_{\mathghost}'].\\
-      \left(\trace[_{\mathghost}]=\preevent[_{\mathghost}];\unlock\cdot\trace[_{\mathghost}'] \text{or }\trace[_{\mathghost}]=\preevent[_{\mathghost}];\lock_{\text{NONE}}\cdot\trace[_{\mathghost}']\right)\\
-      \text{and }\ \trace[_{\mathghost}']\in\sssafe
-                                 \end{array}\right.\right\}
-  $ 
-  \end{nscenter}
-\end{definition}
-The technical setup so far leads to the above definition, where only locks annotated with $\text{SPECTRE-PHT}$ are disallowed to occur on the trace.
-That way, programs attaining \gls*{ss} do not necessarily attain \gls*{scct}.
-
-\subsubsection{Speculation Memory Safety}\label{sec:spec-ms-rel}
-
-As before, we need to relate the different trace models with each other, so that the memory safety property without speculation can be lifted to speculation. 
-To this end, let $\sim_{\mathghost}: \preevent[_{\mathghost}] \times \preevent[_{\ctsafe}]$ be a cross-language trace relation whose key cases are below.
-The intuition is that \gls*{ss} is trivially satisfied in \gls*{scct}, since speculation is inexpressible there, which amounts to dropping events $\ev{Spec}$, $\ev{Rlb}$, or $\ev{Barrier}$, as well as all base events tagged with $\lock_{\text{PHT}}$.
-
-\[
-  \typerulenolabel{specrel:drop:tag}{}{\preevent[_{\ctsafe}];\lock\sim_{\mathghost}\preevent[_{\ctsafe}];\lock_{\text{NONE}}}
-  \typerulenolabel{specrel:drop:spectag}{}{\emptyevent\sim_{\mathghost}\preevent[_{\mathghost}];\lock_{\text{PHT}}}
-\]
-\[
-  \typerulenolabel{specrel:drop:pub}{}{\preevent[_{\ctsafe}];\unlock\sim_{\mathghost}\preevent[_{\ctsafe}];\unlock}
-  \typerulenolabel{specrel:drop:crash}{}{\lightning\sim_{\mathghost}\lightning}
-\]
-\[
-  \typerulenolabel{specrel:drop:spec}{}{\emptyevent\sim_{\mathghost}\ev{Spec}}
-  \typerulenolabel{specrel:drop:rlb}{}{\emptyevent\sim_{\mathghost}\ev{Rlb}}
-  \typerulenolabel{specrel:drop:barrier}{}{\emptyevent\sim_{\mathghost}\ev{Barrier}}
-\]
-
-We conclude by defining the ultimate property of interest for secure compilers: \gls*{specms}.
-\begin{definition}[\gls*{specms}]\label{def:trace:specmsdef}
-  $
-  \specmssafe := \msscctsafe\cap\sigma_{\sim_{\mathghost}}\left(\sssafe\right)
-  $
-\end{definition}
-
-\section{Case Study: Language Formalisations\pages{2}}\label{sec:casestud:defs}
-
-This section defines programming languages $\Ltms$, $\Ltrg$, $\Lms$, $\Lscct$, and $\Lspec$, all of which share common elements (presented in \Cref{subsec:cs:defs}).
-$\Ltms$ is the only statically-typed language, and it exhibits the property that all well-typed programs are \gls*{tms} (\Cref{subsec:ltms}).
-However, not all $\Ltms$ programs are \gls*{sms}.
-That is, there are well-typed $\Ltms$ programs that perform an out-of-bounds access.
-Language $\Ltrg$ is untyped and does not provide any guarantees with regards to \gls*{ms} (\Cref{subsec:lsms}).
-$\Lms$ is exactly the same language as $\Ltrg$, but this paper still distinguishes the two for sake of readability (\Cref{subsec:lms}).
-All three languages --- so $\Ltms$, $\Ltrg$, and $\Lms$ --- assume \gls*{scct} to hold, since this is -- in an ideal world -- what the programmer would expect, too: it is the job of the compiler to preserve and (potentially) enforce \gls*{scct} security~\cite{cauligi2019fact,nagarakatte2010cets,nagarakatte2009soft,akritidis2009baggy}.
-% This paper considers \gls*{cct} as an architecture-dependent property\MPin{can we justify this}, so
-
-Such consideration is also backed up by architecture providing a data (operand) independent timing mode, such as processors by Arm~\cite[p.~543]{arm-refman} and Intel~\cite[p.~80]{intel-refman}.
-%In spirit of this, all languages besides $\Lscct$ are trivially \gls*{scct}. 
-This kind of processor feature is modelled in language $\Lscct$ (\Cref{subsec:lscct}), where programs have access to a ``\gls*{cct}-mode'' and can change the leakage of emitted events according to the value of this mode (either $\obj{ON}$ or $\obj{OFF}$).
-
-Finally, modern processors also employ speculative execution to achieve speedups---and unfortunately generate Spectre attacks~\cite{kocher2019spectre}---and this is the extension of $\Lspec$ (\Cref{subsec:lspec}).
- % extends $\Lscct$ to model speculative execution.
-%
-Thus, all previous languages trivially satisfy \gls*{ss}, since they do not support speculative execution at all.
-% All languages share large parts of syntactic elements, which are presented in the next subsection (\Cref{subsec:cs:defs}).
-% Subsequent subsections extend on this accordingly.
-
-\subsection{Shared Language Definitions}\label{subsec:cs:defs}
-
-All presented programming languages share a common fragment which is partially presented here and in full detail in the technical report. 
-%\MP{put omitted stuff in appendix + ref there (and add textual explanation)}
-
-% \vspace{-.5em}
-{
-  \renewcommand{\src}[1]{\mathsf{#1}}
- \small
-\[
-  \begin{array}{rrcl}
-    \text{(Base-Events)} & \src{\preevent} &:=& \src{\ev{Alloc \loc\ n}} \mid \src{\ev{Dealloc \loc}} \mid \src{\ev{Get \loc\ n}} \mid \src{\ev{Set \loc\ n}} \\
-    \text{(Control Tags)} & \src{\sandboxtag{}} &:=& \src{\ctx} \mid \src{\comp} \\
-    \text{(Events)} & \src{\event} &:=& \src{\preevent;\sandboxtag{}} \mid \src{\emptyevent} \mid \src{\lightning} \\ 
-    \text{(Values)} & \src{v} &:=& \src{n} \\
-    \text{(Expressions)} & \src{\expr} &:=& \src{x} \mid \src{n} \mid \src{\lbinop{\expr[_1]}{\expr[_2]}} \mid \src{\lifz{\expr[_1]}{\expr[_2]}{\expr[_3]}} \\ 
-                         &&&\mid \src{\llet{x}{\expr[_1]}{\expr[_2]}} \mid \src{\llet{x}{new\ \expr[_1]}{\expr[_2]}} \\
-                         &&&\mid \src{\ldelete{x}} \mid \src{\lget{x}{\expr}} \mid \src{\lset{x}{\expr[_1]}{\expr[]}} \mid \src{stuck}\\
-                         &&&\mid \src{\lcall{f}{\expr}} \mid \src{\lreturn{\expr}} \\
-    \text{(Functions)} & \src{F} &:=& (\src{x};\src{\expr}) \\
-    \text{(Libraries)} &&&\hspace{-3.8em} \src{\library_{\ctx}},\src{\library_{\comp}} : \src{Vars} \to \src{F} 
-    % \hspace{2em}
-    %  : \src{Vars} \to \src{F} 
-     \\
-    \text{(Heaps)} & \src{H} &:=& \src{\hole{\cdot}} \mid \src{n},\src{H}
-    % \hspace{0.5cm}    
-    \\
-    \text{(Pointer Maps)} & \src{\Delta} && \text{ omitted for simplicity}
-    \\
-    \text{(Memory States)} & \src{\memstate} &:=& \left(\src{H^{\ctx}};\src{H^{\comp}};\src{\Delta}\right)\\
-    \text{(Control States)} & \src{\cfstate} & & \text{ omitted for simplicity} 
-    \\
-    \text{(States)} &\src{\Omega} &:=& \left(\src{\cfstate};\src{\sandboxtag{}};\src{\memstate}\right)\\
-  \end{array}
-\]
-}
-{
-  \renewcommand{\src}[1]{\mathsf{#1}}
-The trace models of all languages are similar to those presented earlier (\Cref{sec:compprop}).
-One technical detail is the addition of a control tag, indicating who is to blame for an emitted action: context ($\ctx$) or component ($\comp$).
-This is a standard technique in secure compilation in order to rule out irrelevant context-level events. 
-For example, a context immediately deallocating an allocated memory region twice trivially violates memory safety, but the main interest in secure compilation is to preserve component-level security, and thus component events.
-Even though this tagging could be used for blame preservation~\cite{patrignani2023blame}, this is beyond the focus of this paper.
-Another key difference is that memory accesses are now explicitly modelled as reads ($\src{\ev{Read\ \loc\ n}}$) and writes ($\src{\ev{Write\ \loc\ n}}$) instead of just uses.
-
-All languages have at least numbers as values ($\src{v}$) and second class functions ($\src{F}$).
-Functions are modelled as pairs containing the name of one argument and the body of the function.
-Bodies are just ordinary expressions, which can be simple binary operations ($\src{\lbinop{\expr[_1]}{\expr[_2]}}$), conditionals ($\src{\lifz{\expr[_1]}{\expr[_2]}{\expr[_3]}}$), function calls ($\src{\lcall{f}{e}}$) and returns ($\src{\lreturn{\expr}}$), as well as C-like memory operations.
-Programs have sets of pre-determined functions called libraries and they are marked as being part of some component ($\src{\library_{\comp}}$) or context ($\src{\library_{\ctx}}$).
-
-For the operational semantics, the runtime state ($\src{\Omega}$) is a triple consisting of a control-flow state ($\src{\cfstate}$), a control tag ($\src{\sandboxtag{}}$), and a memory state ($\src{\memstate}$).
-The latter carries information about pointers that are kept ``alive'' in pointer maps ($\src{\Delta}$), so that the semantics does not get stuck when encountering, e.g., a double-free.
-The memory state also carries two heaps to model sandboxing between a context ($\src{H^{\ctx}}$) and a component ($\src{H^{\comp}}$).
-
-}
-
-\subsection{$\src{L_{\tmssafe}}$: A Temporal but Not Spatial Memory Safe Language}\label{subsec:ltms}
-
-$\src{L_{\tmssafe}}$ is the only statically-typed language in this case study and restricts functions ($\src{F}$) to the typing signature $\src{\natt\to\natt}$. 
-The type system of $\src{L_{\tmssafe}}$ is inspired by $L^{3}$~\cite{morrisett2005L3,scherer2018fabulous} and enforces that every well-typed $\src{L_{\tmssafe}}$ program satisfies \gls*{tms}.
- % (\Cref{thm:wt:tms}).
-
-\begin{theorem}[$\src{L_{\tmssafe}}$-programs are \gls*{tms}]\label{thm:wt:tms}
-$\rsat{\src{\library_{\comp}}}{\tmssafe}$ \Coqed
-  % $\;$ 
-% 
-  % \begin{nscenter}
-  %   \hfill For any component $\src{\library_{\comp}}$, $\rsat{\src{\library_{\comp}}}{\tmssafe}$ \Coqed
-  % \end{nscenter}
-\end{theorem}
-
-\subsection{$\Ltrg$: A Memory-Unsafe Language}\label{subsec:lsms}
-
-$\Ltrg$ extends the syntax presented earlier (\Cref{subsec:cs:defs}) with dynamic typechecks ($\trg{\lhast{\expr}{\type}}$), evaluating to $\trg{0}$ if the type matches with the shape of $\trg{\expr}$ and $\trg{1}$ otherwise.
-%The pointers carry poison tags, which signify a pointer being allocated ($\trg{\poisoned}$) or freed ($\trg{\poisonless}$). 
-Furthermore, the syntax of $\Ltrg$ is extended with a way to inspect whether a pointer is freed ($\trg{\lispoisoned{x}}$), evaluating to $\trg{0}$ if it is freed and to $\trg{1}$ otherwise.
-Functions may receive arguments that are not $\trg{\natt}$, values are extended with tuples, and expressions are also extended with pair projections.
-
-% \vspace{-1em}
-\[
-  \begin{array}{rrcl}
-    \text{(Values)} & \trg{v} &:=& \dots \mid \trg{\lpair{v_1}{v_2}} \\
-    \text{(Expressions)} & \trg{\expr} &:=& \dots \mid \trg{\lhast{\expr}{\tau}} \mid \trg{\lispoisoned{x}}
-  \end{array}
-\]
-
-No changes are done to the trace model, but $\Ltrg$ has no static typing, making double-free code patterns possible.
-
-\subsection{$\Lms$: Another Memory-Unsafe Language}\label{subsec:lms}
-$\Lms$ is exactly equal to $\Ltrg$ (\Cref{subsec:lsms}), but used to emphasize that this is code after applying $\cc{\Ltrg}{\Lms}$ (\Cref{subsec:cs:ms}).
-
-\subsection{$\Lscct$: A Memory-Unsafe Language with a Constant Time Mode}\label{subsec:lscct}
-
-$\Lscct$ extends $\Lms$ (\Cref{subsec:lms}) with a ,,constant-time mode''. 
-The activation of the mode can be checked ($\obj{\lrddoit{x}{\expr}}$), changed ($\obj{\lwrdoit{D}}$), and is stored in program states ($\obj{\Omega}$).
-At the beginning of program execution of $\Lscct$ programs, the mode is turned off ($\obj{OFF}$).
-If the mode is enabled (i.e., set to $\obj{ON}$), the intuition is that no secrets are leaked. 
-This models the real-world\footnote{As present in Intel~\cite[p.80]{intel-refman} and ARM~\cite[p.~543]{arm-refman} processors.} data-independent timing mode as well as the result of compiling a program with FaCT~\cite{cauligi2019fact}.
-For example, FaCT rewrite code that branches to use a constant-time selection primitive.
-To not obfuscate our formalisation unnecessarily by duplicating syntax, we simply added the ,,constant-time mode'' to $\Lscct$.
-
-% $\Lscct$ extends $\Lms$ (\Cref{subsec:lms}) with a way to read from ($\obj{\lrddoit{x}{\expr}}$) and to write to ($\obj{\lwrdoit{D}}$) a model-specific register that controls a constant-time mode. 
-% This language is a combination of the effect of the FaCT compiler~\cite{cauligi2019fact} as well as a data-independent timing mode, a feature that is present in both Arm~\cite[p.~543]{arm-refman} and Intel~\cite[p.~80]{intel-refman} processors.
-% \MP{ can we be more specific about the fact pass? }
-% To this end, states ($\obj{\Omega}$) are extended with the value of the register ($\obj{D}$), which is initially set to be not active ($\obj{OFF}$).
-% If the register is marked active ($\obj{ON}$), the intuition is that no secrets can appear on specification traces.
-% If the register is marked inactive, secrets may appear on traces.
-% The state of the register can be read and written to, e.g., \Cref{tr:e-wrdoit-off}.
-% For the other languages seen earlier, the mode is intuitively always-on, i.e., the mode of execution always uses constant-time mode.
-
-% \vspace{-.5em}
-\[
-  \begin{array}{rrcl}
-    \text{(Mode Values)} & \obj{D} &:=& \obj{ON} \mid \obj{OFF} \\
-    \text{(Security Tags)} & \obj{\securitytag{}} &:=& \obj{\lock} \mid \obj{\unlock} \\
-    \text{(Base-Events)} & \obj{\preevent} &:=& \dots \mid \obj{iGet\ \loc\ v} \mid \obj{iSet\ \loc\ v} \\
-                       &&& \mid\ \obj{Binop\ n} \mid \obj{Branch\ n} \\
-    \text{(Events)} & \obj{\event} &:=& \obj{\preevent;\sandboxtag{};\securitytag{}} \mid \obj{\emptyevent} \mid \obj{\lightning} \\
-    \text{(Expressions)} & \obj{\expr} &:=& \dots \mid \obj{\lrddoit{x}{\expr}} \mid \obj{\lwrdoit{D}} \\
-                         &&&\mid\ \obj{\llet{x^{\securitytag{}}}{\expr[_1]}{\expr[_2]}}\\
-    \text{(States)} & \obj{\Omega} & := & \left(\obj{\cfstate};\obj{\sandboxtag{}};\obj{D};\obj{\memstate}\right)\\
-  \end{array}
-\]
-\begin{center}
-\newcommand{\expreval}[5]{{#1}\triangleright\xspace {#2}\xrightarrow{#5}\ {#3}\triangleright\xspace {#4}\xspace}
-\newcommand{\exprevalo}[5]{\expreval{\obj{#1}}{\obj{#2}}{\obj{#3}}{\obj{#4}}{\obj{#5}}}
-  \typerule{$e-\obj{wrdoit}-\text{off}$}{
-  }{
-    \pstepto{\rtt{\obj{\cfstate;\sandboxtag{};D;\memstate}}{\obj{\lwrdoit{OFF}}}}{\rtt{\obj{\cfstate;\sandboxtag{};OFF;\memstate}}{\obj{0^{\securitytag{}}}}}{\obj{\emptyevent}}
-  }{e-wrdoit-off}
-
-  \typerule{$e-\obj{get}-\in-$noleak}{
-    \obj{\memstate}=\obj{H^{\ctx};H^{\comp};\Delta_1,x\mapsto(\loc;\sandboxtag{};\poison),\Delta_2} &
-    \obj{\loc}+\obj{n}\in\text{dom }\obj{H^{\sandboxtag{}}} \\
-    \obj{\securitytag{''}}=\obj{\securitytag{}\sqcap\securitytag{'}} &
-    \obj{\event} = \obj{{iGet}\ \loc\ n;\sandboxtag{};\securitytag{''}}
-  }{
-    \exprevalo{
-    	\cfstate;\sandboxtag{'};ON;\memstate
-    }{
-      	x^{\securitytag[n^{\securitytag{'}}] }
-    }{
-    	\cfstate;\sandboxtag{'};ON;\memstate
-    }{
-    	(H^{\sandboxtag{}}(\loc+n))^{\securitytag{''}}
-    }{
-      \event
-    }
-  }{to-e-get-in-noleak}
-\end{center}
-
-The language also adds user annotations $\obj{\securitytag{}}$ for the secrecy of variables, which can be either private (high secrecy) $\obj{\lock}$ or public (low secrecy) $\obj{\unlock}$.
-Security tags $\obj{\securitytag{}}$ are arranged in the usual secrecy lattice~\cite{zdancewicphd}, where $\obj{\unlock}\sqsubseteq\obj{\lock}$.
-% $\obj{\lock}\sqsubseteq\obj{\securitytag{}}$ and $\obj{\securitytag{}}\sqsubseteq\obj{\unlock}$.
-
-Memory accesses to secret data need to be present to reason about memory safety, even when execution is in constant-time mode, e.g, \Cref{tr:to-e-get-in-noleak}.
-$\Lscct$ extends base-events with $\obj{\ev{iGet}\ \loc\ \valueexpr}$ and $\obj{\ev{iSet}\ \loc\ \valueexpr}$ (for data $\obj{i}$ndependent get and set) to prevent secrets from leaking but still enable reasoning about memory safety.
-Due to the technical setup, the rule needs to check if the access is in bounds ($\obj{\loc}+\obj{n}\in\obj{H^t}$) and update the secrecy tag ($\obj{\securitytag{''}}=\obj{\securitytag{}\sqcap\securitytag{'}}$) with the least upper bound of the tags, according to the aforementioned lattice.
-The precise information carried by the event, e.g., location ($\obj{\loc}$) is taken from the pointer map, which carries information irrelevant to this rule ($\obj{\poison}$).
-
-Base-events include $\obj{\ev{Branch}\ n}$ and $\obj{\ev{Binop}\ n}$ that are emitted when evaluating a branch or certain binary expressions, such as division, respectively, whenever the constant-time mode is inactive.
-Events are extended with a security-tag ($\obj{\securitytag{}}$) to signal the secrecy of the involved data.
-
-The evaluation steps are amended to propagate the security-tag annotations $\obj{\securitytag{}}$.
-When the constant-time mode is inactive, base-events $\obj{\ev{Branch}\ n}$ and $\obj{\ev{Binop}\ n}$ are emitted for conditionals and binary operations, respectively.
-Otherwise, just like in the semantics of the earlier languages, $\obj{\emptyevent}$ is emitted for binary and branching operations.
-%To mark this mode as active or inactive, \Cref{tr:e-wrdoit} allows to access the register in $\obj{\statevar}$.
-
-\Cref{ex:lscct} illustrates the differences between $\Lscct$ and other languages.
-
-\begin{exampleenv}[$\Lscct$ with and without constant-time mode]\label{ex:lscct}
-  Consider again \Cref{ex:strncpy}, with a context copying the string \texttt{Hello World}, where everything is marked with a high security tag: $\lock$.
-  The top half of \Cref{fig:ex-cct} (titled $\Lscct$), describes the execution trace of the program, while the bottom side of the table (titled ``Specification''), describes the related specification trace (\Cref{subsec:msctss:tracemodel}).
-  Read in parallel from top to bottom, the figure shows parts of the execution trace. 
-  In each half, the left column (\textit{Active}) has constant-time mode $\obj{ON}$ and the right one (\textit{Inactive}) has it $\obj{OFF}$.
-
-  When the constant-time mode is off, the execution yields events in similar fashion to before (\Cref{subsec:ltms,subsec:lsms,subsec:lms}).
-  But, if it is turned on, then the branching event \colorbox{yellow}{does not fire} anymore and both reading and writing to memory is related to a specification trace with \colorbox{lightgray}{no exposed secrets}.
-\end{exampleenv}
-\begin{figure}[!htb]
-	\vspace{-1em}\small
-  $$
-  % {
-  \begin{array}{ccc}
-    & \Lscct &        \\
-    \text{Active} & \mid & \text{Inactive} \\\hline
-    \obj{\ev{Alloc}\ \loc_{x}\ 12;\comp;\unlock} & \mid & \obj{\ev{Alloc}\ \loc_{x}\ 12;\comp;\unlock} \\
-    \obj{\ev{Alloc}\ \loc_{y}\ 12;\comp;\unlock} & \mid & \obj{\ev{Alloc}\ \loc_{y}\ 12;\comp;\unlock} \\
-    \obj{{\ev{iGet}}\ \loc_{x}\ 0;\comp;}\text{\colorbox{lightgray}{$\obj{\lock}$}} & \mid & \obj{\ev{Get}\ \loc_{x}\ 0;\comp;\lock} \\
-    \text{\colorbox{yellow}{$\obj{\emptyevent}$}} & \mid & \obj{\ev{Branch}\ 0;\comp;\lock} \\
-    \obj{{\ev{iGet}}\ \loc_{x}\ 0;\comp;}\text{\colorbox{lightgray}{$\obj{\lock}$}} & \mid & \obj{\ev{Get}\ \loc_{x}\ 0;\comp;\lock} \\
-    \obj{{\ev{iSet}}\ \loc_{y}\ 0\ \mathtt{'H'};\comp;}\text{\colorbox{lightgray}{$\obj{\lock}$}} & \mid & \obj{\ev{Set}\ \loc_{y}\ 0\ \mathtt{'H'};\comp;\lock} \\
-    \obj{{\ev{iGet}}\ \loc_{x}\ 1;\comp;}\text{\colorbox{lightgray}{$\obj{\lock}$}} & \mid & \obj{\ev{Get}\ \loc_{x}\ 1;\comp;\lock} \\
-    \text{\colorbox{yellow}{$\obj{\emptyevent}$}} & \mid & \obj{\ev{Branch}\ 0;\comp;\lock} \\
-    \vdots & \mid & \vdots \\
-    \obj{{\ev{iGet}}\ \loc_{x}\ 12;\comp;}\text{\colorbox{lightgray}{$\obj{\lock}$}} & \mid & \obj{\ev{Get}\ \loc_{x}\ 12;\comp;\lock} \\
-    \text{\colorbox{yellow}{$\obj{\emptyevent}$}}& \mid & \obj{\ev{Branch}\ 1;\comp;\lock} \\
-    \obj{\ev{Dealloc}\ \loc_{y};\comp;\unlock} & \mid & \obj{\ev{Dealloc}\ \loc_{y};\comp;\lock} \\
-    \obj{{\ev{iGet}}\ \loc_{y}\ 6;\comp;}\text{\colorbox{lightgray}{$\obj{\lock}$}} & \mid & \obj{\ev{Get}\ \loc_{y}\ 6;\comp;\lock} \\
-  \end{array}
-  % }
-  $$
-  $$
-  % {
-  \begin{array}{ccc}
-          & \text{Specification} & \\
-  \text{Active} & \mid & \text{Inactive} \\\hline
-     \ev{Alloc}\ \loc_{x}\ 12;\unlock & \mid & \ev{Alloc}\ \loc_{x}\ 12;\unlock\\
-     \ev{Alloc}\ \loc_{y}\ 12;\unlock & \mid & \ev{Alloc}\ \loc_{x}\ 12;\unlock\\
-     \ev{Use}\ \loc_{x}\ 0;\text{\colorbox{lightgray}{$\unlock$}} & \mid & \ev{Use}\ \loc_{x}\ 0;\lock\\
-     \emptyevent & \mid & \ev{Branch}\ 0;\lock\\
-     \ev{Use}\ \loc_{x}\ 0;\text{\colorbox{lightgray}{$\unlock$}} & \mid & \ev{Use}\ \loc_{x}\ 0;\lock\\
-     \ev{Use}\ \loc_{y}\ 0;\text{\colorbox{lightgray}{$\unlock$}} & \mid & \ev{Use}\ \loc_{y}\ 0;\lock\\
-     \ev{Use}\ \loc_{x}\ 1;\text{\colorbox{lightgray}{$\unlock$}} & \mid & \ev{Use}\ \loc_{x}\ 1;\lock\\
-     \emptyevent & \mid & \ev{Branch}\ 0;\lock\\
-     \vdots & \mid & \vdots \\
-      \ev{Use}\ \loc_{x}\ 12;\text{\colorbox{lightgray}{$\unlock$}} & \mid & \ev{Use}\ \loc_{x}\ 12;\lock\\
-      \emptyevent & \mid & \ev{Branch}\ 1;\lock\\
-      \ev{Dealloc}\ \loc_{y};\unlock & \mid & \ev{Dealloc}\ \loc_{y};\unlock\\
-      \ev{Use}\ \loc_{y}\ 6;\text{\colorbox{lightgray}{$\unlock$}} & \mid & \ev{Use}\ \loc_{y}\ 6;\lock\\
-  \end{array}
-  % }
-  $$
-  \caption{Traces for \Cref{ex:lscct}.}
-  \label{fig:ex-cct}
-\end{figure}
-
-\subsection{$\Lspec$: A Memory-Unsafe Language with Speculation}\label{subsec:lspec}
-
-$\Lspec$ extends $\Lscct$ (\Cref{subsec:lscct}) with speculative dynamic semantics that is inspired by existing work~\cite{guarnieri2018spectector,fabian2022automatic}.
-After branching, speculation starts by pushing the current configuration into the speculation state ($\ird{S}$), and run subsequent code in a predetermined window (whose size is $\omega$) until a rollback of operational state is performed.
-The window is set inside the speculation state.
-Within such windows, data may leak, this is marked as high $\ird{\lock}$ and with an annotation that indicates the respective speculative execution variant. 
-In this paper, we just consider SPECTRE-PHT~\cite{kocher2019spectre}, whose starting point is \Cref{tr:e-ifz-true-spec}.
-% where $\omega$ is the size of the speculation window, i.e., a number of reduction steps to be taken under speculative semantics.
-Additionally, $\ird{\lock}$ may also carry a $\ird{NONE}$ annotation, to signal that this is a leak irrespective of speculation, i.e., as defined earlier (\Cref{subsec:lscct}).
-The semantics is kept general enough to allow for future extension to support different variants~\cite{kocher2019spectre,maisuradze2018ret2spec,horn2019zero}.
-
-In terms of language features, $\Lspec$ includes a new barrier operation $\ird{\lbarrier}$ that blocks speculative execution. 
-To facilitate speculative execution in a non-assembly-like language, a stack of operational state is used and speculation is active if that stack has a size greater than one.
-This is exploited in \Cref{tr:e-spec-eat} to consume any leftover speculation.
-
-\vspace{-.5em}
-\[
-  \begin{array}{rrcl}
-    \text{(Base-Events)} & \ird{\preevent} &:=& \dots \mid \ird{Spec} \mid \ird{Rlb} \mid \ird{Barrier}\\
-    \text{(Expressions)} & \ird{\expr} &:=& \dots \mid \ird{\lbarrier} \\
-    \text{(Speculation State)} & \ird{S} &:=& \ird{\Omega\triangleright\expr} \mid \ird{S},\left(\ird{\Omega};\ird{n};\ird{\expr}\right)
-  \end{array}
-\]
-
-\begin{center}
-\newcommand{\expreval}[5]{{#1}\triangleright\xspace {#2}\xrightarrow{#5}\ {#3}\triangleright\xspace {#4}\xspace}
-\newcommand{\exprevald}[5]{\expreval{\ird{#1}}{\ird{#2}}{\ird{#3}}{\ird{#4}}{\ird{#5}}}
-
-  %
-  \typerule{$e-\ird{ifz}-\ird{true}-\text{spec}$}{
-    \ird{S} = \rtt{(\ird{\cfstate['];\sandboxtag{};D;\memstate});\ird{S}}{\ird{\expr[_1]}},({\ird{\cfstate;\sandboxtag{};D;\memstate}};\omega;{\ird{\expr[_2]}})
-  }{
-    \pstepto{\rtt{\ird{\cfstate;\sandboxtag{};D;\memstate}}{\ird{\lifz{0^{\securitytag{}}}{\expr[_1]}{\expr[_2]}}}}{\ird{S}}{\ird{{Branch}\ 0;\sandboxtag{};\securitytag{}}\cdot\ird{{Spec};\sandboxtag{};\securitytag{}}}
-  }{e-ifz-true-spec}
-  %
-  \typerule{$e-\ird{spec}-\text{eat}$}{
-    \ird{n} > 0 &
-    \pstepto{\rtt{\ird{\Omega}}{\ird{\expr}}}{\rtt{\ird{\Omega'}}{\ird{\expr[']}}}{\ird{\event}}
-  }{
-    \ghoststepto{\ird{S},(\ird{\Omega};\ird{n};\ird{\expr})}{\ird{S},(\ird{\Omega'};\ird{n}-1;\ird{\expr'})}{\ird{\event}}
-  }{e-spec-eat}
-  % 
-  \typerule{$e-\ird{spec}-\text{eaten}$}{
-  }{
-    \ghoststepto{\ird{S},(\ird{\Omega};\ird{0};\ird{\expr})}{\ird{S}}{\ird{Rlb;\Omega.\sandboxtag{};\unlock}}
-  }{e-spec-eaten}
-  %
-  \typerule{$e-\ird{spec}-\text{barrier}$}{
-  }{
-    \ghoststepto{\ird{S},(\ird{\Omega};\ird{n};\ird{\lbarrier})}{\ird{S},(\ird{\Omega};\ird{0};\ird{0})}{\ird{Barrier;\Omega.\sandboxtag{};\unlock}}
-  }{e-spec-barrier}
-\end{center}
-
-The trace model is extended with $\ird{{Spec}}$, $\ird{{Rlb}}$, and $\ird{{Barrier}}$ events that signal the respective operational action. 
-A $\ird{\lbarrier}$ prevents any execution whatsoever besides $\ird{{Rlb}}$ when run in speculative execution mode and does nothing when run in normal mode, e.g., \Cref{tr:e-spec-barrier}.%
-\footnote{In the rule, the notation $\cfstate' = \cfstate[.\text{win} = 0]$ means that $\cfstate'$ is a copy of $\cfstate$ up to field $\text{win}$, which is set to $0$.}
-Lastly, $\ird{Rlb}$ is emitted when the speculation window is zero and the operational state is rolled back (see \Cref{tr:e-spec-eaten}).
-
-\section{Case Study: Composing Secure Compiler Passes and Optimisations \pages{7}}\label{sec:casestud:rtp}
-This section defines several secure compilers, each of which robustly preserves a different property of interest as depicted in \Cref{fig:pipeline}.
-\begin{figure*}[!h]
-  \centering
-  \begin{tikzpicture}
-    \node (S) {$\src{L_{\tmssafe}}$};
-    \node[right=2.0 of S] (T) {$\trg{L}$};
-    \node[right=2.0 of T] (M) {$\irl{L_{\mssafe}}$};
-    \node[below right=1.5 and 1.0 of M] (D0) {$\irl{L_{\mssafe}}$};
-    \node[above right=1.5 and 1.0 of M] (C0) {$\irl{L_{\mssafe}}$};
-    \node[right=3.0 of M] (E) {$\irl{L_{\mssafe}}$};
-    \node[right=2.0 of E] (O) {$\obj{L_{\scctsafe}}$};
-    \node[right=2.0 of O] (Z) {$\ird{L_{\mathghost}}$};
-
-    \draw[->] (S) to[sloped] node[align=center] (tmsedge) {\gls*{tms}\\ \Cref{thm:cca:rtp:tms}} (T);
-    \draw[->] (T) to[sloped] node[align=center] {\gls*{sms}\\ \Cref{thm:ccb:rtp:sms}} (M);
-    \draw[->] (M) to[sloped] node[align=center] {\gls*{dce}\\ \Cref{thm:ccdce:rtp:ms}} (D0);
-    \draw[->] (M) to[sloped] node[align=center] {\gls*{cf}\\ \Cref{thm:cccf:rtp:ms}} (C0);
-    \draw[->] (D0) to[sloped] node[align=center] {\gls*{cf}\\ \Cref{thm:cccf:rtp:ms}} (E);
-    \draw[->] (C0) to[sloped] node[align=center] {\gls*{dce}\\ \Cref{thm:ccdce:rtp:ms}} (E);
-    \draw[->] (E) to[sloped] node[align=center] {\gls*{scct}\\ \Cref{thm:ccscct:rtp:scct}} (O);
-    \draw[->] (O) to[sloped] node[align=center] {\gls*{ss}\\ \Cref{thm:ccspec:rtp:spec}} (Z);
-
-    % Sections
-    \node (sectms) at ($(S)+(1.25,2.5)$) {{\Cref{subsec:cs:tms}}};
-    \node (secsms) at ($(T)+(1.25,2.5)$) {{\Cref{subsec:cs:ms}}};
-    \node (secopts) at ($(M)+(2.0,2.5)$) {{\Cref{subsec:cs:opts}}};
-    \node (secscct) at ($(E)+(1.5,2.5)$) {{\Cref{subsec:cs:scct}}};
-    \node (secss) at ($(O)+(1.5,2.5)$) {{\Cref{subsec:cs:ss}}};
-
-    \draw[thick,dotted,->] ($(S) + (0.2,-0.27)$) to ($(S) - (-0.2,1)$) to node[align=center] {\gls*{ms}\\ \Cref{corr:ccab:rtp:ms}} ($(M) - (0.0,1)$) to (M);
-
-    \draw[thick,dotted,->] (S) to ($(S) - (0.0,2.5)$) to node[align=center,pos=.65] {\gls*{specms}\\ \Cref{thm:ccall:rtp:specms}} ($(Z) - (0.0,2.5)$) to (Z);
-
-    % \draw[thick,dotted,->] (S) to[bend right=33,sloped] node[align=center] {\gls*{ms}\\ \Cref{corr:ccab:rtp:ms}} (M);
-    \draw[thick,dotted,->] (M) to[bend right=0,sloped] node[align=center] {\gls*{ms}\\ \Cref{thm:cccfccdce:rtp:ms}} (E);
-%    \draw[thick,dotted,->] (S) to[out=-45,in=198,sloped] node[align=center] {\gls*{ms}+\gls*{scct}\\ \Cref{thm:ccall:rtp:msscct}} (O);
-    % \draw[thick,dotted,->] (S) to[out=-40,in=210,sloped] node[align=center,pos=.8] {\gls*{specms}\\ \Cref{thm:ccall:rtp:specms}} (Z);
-
-    \draw[dashed] ($(S)-(0.0,-0.5)$) -- ($(S)-(0.0,-2.5)$);
-    \draw[dashed] ($(T)-(0.0,-0.5)$) -- ($(T)-(0.0,-2.5)$);
-    \draw[dashed] ($(M)-(0.0,-0.5)$) -- ($(M)-(0.0,-2.5)$);
-    \draw[dashed] ($(E)-(0.0,-0.5)$) -- ($(E)-(0.0,-2.5)$);
-    \draw[dashed] ($(O)-(0.0,-0.5)$) -- ($(O)-(0.0,-2.5)$);
-    \draw[dashed] ($(Z)-(0.0,-0.5)$) -- ($(Z)-(0.0,-2.5)$);
-  \end{tikzpicture}
-  \vspace{-1em}
-  \caption{Visualisation of the optimising compilation pipeline that preserves \gls*{specms}. %
-    Vertices in the graph are the programming languages from earlier sections (\Cref{sec:casestud:defs}). %
-    Full edges are secure compilers passes.
-    Dotted edges are composition of passes and use the presented framework (\Cref{sec:sequential}) to indicate the property they preserve. %
-    The dashed lines partition the graph into the sections where the respective theorems are presented.
-  }\label{fig:pipeline}
-\end{figure*}
-The section demonstrates the power of the framework (\Cref{sec:sequential}) by composing these compilers for a secure and optimising compilation chain that robustly preserves \gls*{specms}.
-The first step in this chain is the compiler from $\src{L_{\tmssafe}}$ to $\trg{L}$ that robustly preserves just \gls*{tms} (\Cref{thm:cca:rtp:tms}).
-From here, another pass from $\trg{L}$ to $\irl{L_{\mssafe}}$ ensures that no out-of-bounds accesses can happen and, thus, programs at this point attain \gls*{sms} (\Cref{thm:ccb:rtp:sms}).
-Since these properties compose into \gls*{ms}, composing these passes yields a compiler that robustly preserves \gls*{ms} (\Cref{corr:ccab:rtp:ms}).
-Then, the section presents two optimisation passes, namely \gls*{cf} and \gls*{dce}, each of which robustly preserves \gls*{ms} (\Cref{thm:ccdce:rtp:ms,thm:cccf:rtp:ms}).
-These passes can be freely ordered in the compilation chain without compromising memory safety (\Cref{thm:cccfccdce:rtp:ms}).
-The next step in the chain ensures that code stays \gls*{scct} (\Cref{thm:ccscct:rtp:scct}) when compiled from $\Lms$ to $\Lscct$, which is done by switching on a constant-tame mode for the computation.
-Lastly, by introducing barriers immediately after branches, speculative leaks via SPECTRE-PHT are prevented when compiling $\Lscct$ to $\Lspec$.
-The final result is that the whole compilation chain robustly preserves \gls*{specms} (\Cref{thm:ccall:rtp:specms}).
-
-\subsection{Robust Temporal Memory Safety Preservation}\label{subsec:cs:tms}
-
-The secure compiler from $\Ltms$ to $\Ltrg$ needs to ensure that when execution switches from context to component, the type signatures are respected.
-Thus, the compiler inserts the following dynamic typechecks before entering the body of a component-defined function (anything elided is a trivial identity function from source to target):
-
-\vspace{-1em}
-{
-\begin{gather}
-  \begin{align*}%{rcl}
-    %\cca(\src{x}) = &\ \trg{x} 
-    % \\
-  	%&
-  	%\qquad
-    %\cca(\src{\lbinop{\expr[_{1}]}{\expr[_{2}]}}) = &\ \lbinop[\trg]{\left[\cca(\src{\expr[_{1}]})\right]}{\left[\cca(\src{\expr[_{2}]})\right]} \\
-    %\cca(\src{n}) = &\ \trg{n} 
-    % \\
-    %&
-    \cca(\src{\lget{x}{\expr}}) = &\ \lget[\trg]{\trg{x}}{\left[\cca(\src{\expr})\right]} \\
-    \cca(\src{\ldelete{x}}) = &\ \ldelete[\trg]{\left[\cca(\src{x})\right]} 
-    \\
-    \cca(\src{\lfunction{g}{x:\natt\to\type[_{e}]}{\expr}})  =&
-  \end{align*}
-  \\
-  \begin{align*}
-%    \cca(\src{\lset{x}{\expr[_{1}]}{\expr[_{2}]}}) & = \lset[\trg]{\left[\cca(\src{x})\right]}{\left[\cca(\src{\expr[_{1}]})\right]}{\left[\cca(\src{\expr_{2}})\right]} \\
-% \\ 
-\lfunction[\trg]{\trg{g}}{\trg{x}}{\lifz[\trg]{\trg{\lhast{x}{\natt}}}{
-                                                            \left[\cca(\src{\expr})\right] %
-                                                                                                 }{\labort[\trg]}}
-  \end{align*}
-\end{gather}
-}
-
-\noindent Since $\trg{L}$ has no static typing, an attacker $\trg{\library_{\ctx}}$ can invoke a component function accepting a $\src{\natt}$ with $\trg{\lpair{17}{29}}$.
-With the dynamic check, the compiler ensures that execution aborts in such cases.
-%The compiler does not insert other checks and proceeds as the identity function (which in this paper amounts to a simple re-colouring of $\src{L_{\tmssafe}}$ to $\trg{L}$ expressions).
-
-Compiling the \texttt{strncpy} function from \Cref{sec:introduction} with $\cca$, the compiler would in this case ensure that the arguments that are evaluated in the compiled \texttt{strncpy} are valid.
-%This effectively has no influence on the resulting trace, so a corresponding cross-language trace relation $\sim^{\Ltms}_{\Ltrg}$ is just equality.
-
-$\cca$ is robustly preserving (\Cref{def:rtp}) \gls*{tms}:
-\begin{theorem}[$\cca$ secure w.r.t. \gls*{tms}]\label{thm:cca:rtp:tms}
-    $\rtp{\cca}{\tmssafe}$ % \Coqed
-\end{theorem}
-
-\subsection{Robust Spatial Memory Safety Preservation}\label{subsec:cs:ms}
-
-The spatial memory safety preserving compiler from $\Ltrg$ to $\Lms$ only inserts bounds-checks whenever reading from or writing to memory in order to enforce \gls*{sms}.
-These bounds checks need the bounds information, which the compiler keeps around by introducing a fresh identifier $\irl{x_{SIZE}}$ for each allocation that binds $\irl{x}$.
-Then, it is simply a matter of referring to that variable and ensuring that memory accesses are in the interval $[\irl{0},\irl{x_{SIZE}})$.
-When the check fails, the code aborts.
-
-\begin{nscenter}
-% \small
-  $$
-  \begin{array}{rcl}
-    \ccb(\trg{\lnew{x}{\expr[_{1}]}{\expr[_{2}]}}) & = 
-                                                   & \llet[\irl]{\irl{x_{SIZE}}}{\ccb(\trg{\expr[_{1}]})}{}
-    		\\&&
-    		\lnew[\irl]{\irl{x}}{\irl{x_{SIZE}}}{\ccb(\trg{\expr[_{2}]})}
-    		 \\
-  \ccb(\trg{\lget{x}{\expr}}) & = 
-                              & \llet[\irl]{\irl{x_{n}}}{\ccb(\trg{\expr})}{}
-  	\\&&
-  \lifz[\irl]{\irl{0\le x_{n}<x_{SIZE}}}{\\&&\irl{\lget{x}{x_{n}}}}{}
-  		\irl{\labort}
-  	  \\
-  \ccb(\trg{\lset{x}{\expr[_{1}]}{\expr[_{2}]}}) & = 
-                                                 & \llet[\irl]{\irl{x_{n}}}{\ccb(\trg{\expr[_{1}]})}{}
-  		\\&&
-  \lifz[\irl]{\irl{0\le x_{n}<x_{SIZE}}}{\\&&\lset[\irl]{\irl{x}}{\irl{x_{n}}}{}
-  		\ccb(\trg{\expr[_{2}]})
-  		}{\irl{\labort}} 
-  		% \\
-  \end{array}
-  $$
-\end{nscenter}
-
-\begin{exampleenv}[Instrumented \texttt{strncpy}]
-  Consider again \texttt{strncpy}, but instrumented for \gls*{sms}:
-    \begin{lstlisting}[language=c,basicstyle=\ttfamily, morekeywords={size_t}]
-void strncpy(size_t n, size_t dst_size, char *dst,
-             size_t src_size, char *src) {
-  for(size_t i = 0; i < src_size
-      && src[i] != '\0' && i < n; ++i) {
-    if(i < src_size && i < dst_size) {
-      dst[i] = src[i];
-    }
-  }
-}
-    \end{lstlisting}
-Consider context \texttt{strncpy(2, x, y)}, where \texttt{x} and \texttt{y} are pointers to valid regions of memory with allocated space for exactly two cells and do not contain the null-terminating character \texttt{'\textbackslash 0'}.
-% 
-Without the \gls*{sms} pass, the event $\ev{Use}\ \loc_{x}\ 2;\comp;\unlock$ would appear on the trace, but that indicates an out-of-bounds access! 
-Fortunately, with \gls*{sms} mitigation in place, that event does not appear during execution, since the bounds check prevents the condition \texttt{src[i] != '\textbackslash 0'} from executing.
-\end{exampleenv}
-
-Contrary to the previous compiler, $\ccb$ may change the trace of the original $\Ltrg$ program: if there is a memory access, it needs to be protected with a bounds check.
-% 
-The corresponding relation $\sim^{\Ltrg}_{\Lms} : \trg{\trace}\times\irl{\trace}$ that describes this semantic effect of the compiler is defined partially below.
-% 
-We omit action $\trg{Set}$, which is treated analogously, and any other event, which is related to its cross-language equivalent.
-% \footnote{Up to colours.}
-% \footnote{In the interest of space, the full details are in the technical report.}
-
-{
-\[
-  \typerulenolabel{xrel:sms:read}{\trg{n}\text{ in bounds}}{\trg{Get\ \loc\ n;\comp}\sim^{\Ltrg}_{\Lms}\irl{Get\ \loc\ n;\comp}}
-\]
-% \[
-%   \typerulenolabel{xrel:sms:panic}{\trg{n}\text{ out of bounds}}{\trg{Get\ \loc\ n;\comp}\sim^{\Ltrg}_{\Lms}\irl{\lightning}}
-% \]
-}
-
-% Note that neither $\Ltrg$ nor $\Lms$ expose branches on the trace.\MP{so what?}
-For simplicity, we elide the environment that this relation carries around in order to bind each location to its metadata (such as its size), and resolve the ``$\trg{n}$ in/out of bounds'' premise.
-% The bounds check mentioned in the premises of the rules above is left abstract for the sake of simplicity, the interested reader can find
-% since the precise definition needs some technicalities that only inhibit readability without any technical insights.
-% 
-We can now prove that compiler $\ccb$ robustly preserves \gls*{sms}.
-\begin{theorem}[$\ccb$ secure w.r.t. \gls*{sms}]\label{thm:ccb:rtp:sms}
-  $\rtpsim{\ccb}{\smssafe}{\sim^{\Ltrg}_{\Lms}}$ % \Coqed
-\end{theorem}
-
-At this point we can compose $\ccb$ with the previous compiler ($\cca$), but in order to do so, we need a trace relation from $\Ltms$ to $\Lms$.
-% 
-We can obtain this relation by composing the trace relation we just defined ($\sim^{\Ltrg}_{\Lms}$) with the one used by the previous compiler: $\sim^{\Ltms}_{\Ltrg} : \src{\trace}\times\trg{\trace}$.
-% 
-The latter has not been previously defined (nor has it been used in the related theorem) because that is just an equality relation, since the trace models of $\Ltms$ and of $\Ltrg$ are the same.
-% 
-Thus, we formally define $\sim^{\Ltms}_{\Lms}: \src{\trace} \times \irl{\trace}$ as the following composition: $\sim^{\Ltms}_{\Ltrg}\bullet\sim^{\Ltrg}_{\Lms}$.
-% 
-With this relation, \Cref{corr:ccab:rtp:ms} states that the composition of $\cca$ and $\ccb$ is secure w.r.t.
-\gls*{ms} and it follows from \Cref{thm:cca:rtp:tms,thm:ccb:rtp:sms} using \Cref{thm:rtpsim:sig}.
-% 
-\begin{corollary}[$\cca\circ\ccb$ secure w.r.t. \gls*{ms}]\label{corr:ccab:rtp:ms}
-  $\;$
-
-  \begin{nscenter}
-    $\rtpsim{\cca\circ\ccb}{\mssafe}{\sim^{\Ltms}_{\Lms}}$ % \Coqed
-  \end{nscenter}
-\end{corollary}
-% 
-This proof requires another precondition besides \Cref{thm:ccb:rtp:sms,thm:cca:rtp:tms}: $\sim^{\Ltms}_{\Ltrg}$ needs to be well-formed with respect to $\smssafe$.
-This follows trivially since $\sim^{\Ltms}_{\Ltrg}$ is an equality. 
-% 
-\begin{lemma}[$\sim^{\Ltms}_{\Ltrg}$ well-formed w.r.t. $\smssafe$]\label{lem:wf:ltmsltrg}
-$\;$
-
-  \begin{nscenter}
-  $\wfcsig{\sim^{\Ltms}_{\Ltrg}}{\smssafe}$
-  \end{nscenter}
-\end{lemma}
-
-\subsection{Optimising Compilers}\label{subsec:cs:opts}
-
-This section defines two optimising compiler passes from $\Lms$ to $\Lms$ which perform \gls*{dce} and \gls*{cf}, respectively.
-The \gls*{dce} pass applies a na\"ive rewrite rule on conditionals.
-The \gls*{cf} pass relies on an auxiliary function \texttt{mix} that uses a substitutions accumulator $\irl{\substlist}$ in order to rewrite constant binary operations, e.g., $\irl{{17}-1}$ to $\irl{16}$, and replace variables that are assigned to constants, e.g., $\irl{\llet{x}{7}{x}}$ to $\irl{7}$.
-
-\vspace{-1em}
-\begin{gather*}
-  \begin{align*}
-    \ccdce(\irl{\lifz{true}{\expr[_{1}]}{\expr[_{2}]}}) & = \ccdce(\irl{\expr[_{1}]}) &\\
-    \ccdce(\irl{\lifz{false}{\expr[_{1}]}{\expr[_{2}]}}) & = \ccdce(\irl{\expr[_{2}]}) &
-  \end{align*}
-  \\
-  % \begin{align*}
-  %   \ccdce(\irl{\lbinop{\expr[_{1}]}{\expr[_{2}]}}) & = \lbinop[\irl]{\ccdce(\irl{\expr[_{1}]})}{\ccdce(\irl{\expr[_{2}]})} &
-  % \end{align*}
-  % \\[0.25cm]
-  \begin{align*}
-    \cccf(\irl{\expr}) & = \partialeval{\irl{\expr}}{\irl{\hole{\cdot}}} &
-  \end{align*}
-  \\[0.125cm]
-  \begin{align*}
-   \partialeval{\irl{x}}{\irl{\substlist}} & = \irl{n} 
-   	\qquad\qquad \text{if } \irl{\subst{n}{x}}\in\irl{\substlist} \\
-   \partialeval{\irl{x}}{\irl{\substlist}} & = \irl{x} 
-   \qquad\qquad \text{if } \irl{\subst{n}{x}}\notin\irl{\substlist} \\
-   \partialeval{\irl{\lbinop{n}{m}}}{\irl{\substlist}} & = \irl{k} 
-   \qquad\qquad \text{if } \lbinop{\irl{n}}{\irl{m}}=k \\
-   %\partialeval{\irl{\lbinop{\expr[_{1}]}{\expr[_{2}]}}}{\irl{\substlist}} & = \lbinop[\irl]{\partialeval{\irl{\expr[_{1}]}}{\irl{\substlist}}}{\partialeval{\irl{\expr[_{2}]}}{\irl{\substlist}}} \\
-   \partialeval{\irl{\llet{x}{n}{\expr}}}{\irl{\substlist}} & = \partialeval{\irl{\expr}}{\irl{\subst{n}{x}\cdot\substlist}} 
-  % \\
-  % \partialeval{\irl{\lget{x}{\expr}}}{\irl{\substlist}} & = \lget[\irl]{\irl{x}}{\partialeval{\irl{\expr}}{\irl{\substlist}}}
-% \end{align*}
-\end{align*}
-% \\
-% \begin{align*}
-%   \partialeval{\irl{\llet{x}{\expr[_{1}]}{\expr[_{2}]}}}{\irl{\substlist}} & = \llet[\irl]{\irl{x}}{\partialeval{\irl{\expr[_{1}]}}{\irl{\substlist}}}{\\&\partialeval{\irl{\expr[_{2}]}}{\irl{\substlist}}} \\
-%   \partialeval{\irl{\lifz{\expr[_{1}]}{\expr[_{2}]}{\expr[_{3}]}}}{\irl{\substlist}} & = \lifz[\irl]{\partialeval{\irl{\expr[_{1}]}}{\irl{\substlist}}}{\\&\partialeval{\irl{\expr[_{2}]}}{\irl{\substlist}}}{\partialeval{\irl{\expr[_{3}]}}{\irl{\substlist}}} \\
-% \end{align*}
-\end{gather*}
-% \vspace{-3em}
-
-Note that both passes have no effect on the resulting trace of a program, up to $\emptyevent$-steps. 
-Because of this, both passes have equality as corresponding cross language trace relation. 
-Moreover, it is straightforward to prove both passes as secure (\Cref{def:rtp}) w.r.t. \gls*{ms}.
-
-\begin{theorem}[$\ccdce$ secure w.r.t. \gls*{ms}]\label{thm:ccdce:rtp:ms}
-  $\rtp{\ccdce}{\mssafe}$ %\Coqed
-\end{theorem}
-\begin{theorem}[$\cccf$ secure w.r.t. \gls*{ms}]\label{thm:cccf:rtp:ms}
-  $\rtp{\cccf}{\mssafe}$ %\Coqe
-\end{theorem}
-
-With both \Cref{thm:ccdce:rtp:ms,thm:cccf:rtp:ms} it follows from \Cref{corr:swappable} that the two passes can be interchanged arbitrarily:
-
-\begin{theorem}[$\cccf\circ\ccdce$ and $\cccf\circ\ccdce$ are secure w.r.t. \gls*{ms}]\label{thm:cccfccdce:rtp:ms}
-$\;$
-
-  \begin{nscenter}
-  \phantom{and~}\!\!$\rtp{\cccf\circ\ccdce}{\mssafe}$
-
-  and~$\rtp{\ccdce\circ\cccf}{\mssafe}$ % \Coqed
-  \end{nscenter}
-\end{theorem}
-
-\subsection{Robust Strict Cryptographic Constant Time Preservation}\label{subsec:cs:scct}
-
-This section defines a compiler $\ccscct$ from $\Lms$ to $\Lscct$ that robustly preserves \gls*{scct}. 
-Given the fact that $\Lscct$ provides a \gls*{cct}-mode that can be turned on or off, the compiler inserts wrapper code for function calls and function bodies to ensure that execution in the component always happen in \gls*{cct}-mode.
-This simple flag combines the effect of FaCT~\cite{cauligi2019fact}
-
-\vspace{-1em}
-\[
-\begin{array}{rcl}
-  \ccscct(\irl{\lfunction{g}{x}{\expr}}) & = & \lfunction[\obj]{\obj{g}}{\obj{x}}{\obj{\lwrdoit{ON};}\ccscct(\irl{\expr})} \\
-  \ccscct(\irl{\lcall{g}{\expr}}) & = & \lcall[\obj]{\obj{g}}{\ccscct(\irl{\expr})\obj{; \lwrdoit{ON}}} 
-    % \\
-    % \ccscct(\irl{\lbinop{\expr[_{1}]}{\expr[_{2}]}}) & = \lbinop[\obj]{\ccscct{\irl{\expr[_{1}]}}}{\ccscct{\irl{\expr[_{2}]}}} 
-    % \\
-\end{array}
-\]
-% \vspace{-2em}
-%
-The context can overwrite the flag and exit the mode, but upon invoking a function that is part of the component, the flag is set again.
-Because of this, the corresponding cross-language trace relation $\sim^{\Lms}_{\Lscct}$, only relates events without secrets:%
-
-\begin{center} 
-  \typerulenolabel{xrel:scct:noleak}{}{\irl{\preevent;\comp}\sim^{\Lms}_{\Lscct}\obj{\preevent;\comp;\unlock}}
-\end{center}
-
-The compiler is secure w.r.t. \gls*{scct}: %, similarly proven as in \Cref{subsec:cs:tms}.
-
-\begin{theorem}[$\ccscct$ secure w.r.t. \gls*{scct}]\label{thm:ccscct:rtp:scct}
-  \small
-  $\rtpsim{\ccscct}{\scctsafe}{\sim^{\Lms}_{\Lscct}}$ % \Coqed
-\end{theorem}
-
-\subsection{Robust Speculative Safety Preservation}\label{subsec:cs:ss}
-
-This section defines the final compilation pass $\ccspec$, which ensures that $\Lscct$ programs, which are assumed to be \gls*{ss}, stay \gls*{ss} at $\Lspec$-level. 
-To do so, $\ccspec$ inserts a speculation barrier after branches, which is sufficient to harden the program against speculative leaks, since SPECTRE-PHT~\cite{kocher2019spectre} is the only speculative leak modeled in the semantics of $\Lspec$.
-
-\vspace{-1em}
-\[
-\begin{array}{cl}
-  &\ccspec{(\obj{\lifz{\expr[_0]}{\expr[_1]}{\expr[_2]}})} = 
-  \\
-  &\qquad\qquad \lifz[\ird]{\ccspec{\left(\obj{\expr[_0]}\right)}}{\ird{\lbarrier;}\ccspec{\left(\obj{\expr[_1]}\right)} \\&\qquad\qquad}{ \ird{\lbarrier;}\ccspec{\left(\obj{\expr[_2]}\right)}} 
-\end{array}
-\]
-% \vspace{-2em}
-%
-
-Clearly, the corresponding cross-language trace relation $\sim^{\Lscct}_{\Lspec}$ has only one non-trivial case: for branches, only relate them where speculation is blocked by a barrier:
-
-\begin{center}
-  \typerulenolabel{xrel:spec:if}{}{\obj{Branch\ n}\sim^{\Lscct}_{\Lspec}\ird{Branch\ n}\cdot\ird{Spec}\cdot\ird{Barrier}\cdot\ird{Rlb}}
-\end{center}
-
-The base-event relation above scales to full events by ensuring the missing annotations ($\obj{\comp;\securitytag{}}$ and $\ird{\comp;\securitytag{}}$) are the same.
-% 
-With this relation, we prove that $\ccspec$ is secure with respect to \gls*{ss}.
-\begin{theorem}[$\ccspec$ secure w.r.t. \gls*{ss}]\label{thm:ccspec:rtp:spec}
-  \small$\rtpsim{\ccspec}{\sssafe}{\sim^{\Lscct}_{\Lspec}}$ % \Coqed
-\end{theorem}
-
-\subsection{Robust Preservation of Memory Safety, Strict Cryptographic Constant Time, and Speculative Safety}
-
-Finally, this subsection combines all previous results into one compilation chain to get that it preserves full \gls*{specms}.
-Let $\ccspecms$ be the compiler that is the composition of $\cca$, $\ccb$, $\cccf$, $\ccdce$, $\ccscct$, and $\ccspec$. 
-Let $\sim^{\Ltms}_{\Lspec}$ be the composition of $\sim^{\Ltms}_{\Lms}$, $\sim^{\Lms}_{\Lscct}$, and $\sim^{\Lscct}_{\Lspec}$.
-Then, the following corollary holds.
-
-\begin{corollary}[$\ccspecms$ secure w.r.t. \gls*{specms}]\label{thm:ccall:rtp:specms}
-  $\;$
-
-  \begin{nscenter}
-    $\rtpsim{\cc{\Ltms}{\Lspec}}{\mssafe\cap\scctsafe\cap\sssafe}{\sim^{\Ltms}_{\Lspec}}$ % \Coqed
-  \end{nscenter}
-\end{corollary}
-
-As with \Cref{corr:ccab:rtp:ms}, it is important to ensure that the respective cross language trace relations are well-formed (\Cref{def:wfc:sig:tracerel}).
-It is already known that $\sim^{\Ltms}_{\Lms}$ is well-formed with respect to $\mssafe$ (\Cref{lem:wf:ltmsltrg}).
-Next in the chain is $\sim^{\Lms}_{\Lscct}$, which has to be well-formed w.r.t. $\scctsafe$.
-This lemma holds, since a trace that was $\scctsafe$ is $\scctsafe$ even after applying $\sim^{\Lms}_{\Lscct}$: the relation enforces that $\Lscct$ traces related to $\Lms$ traces have no leaks of secrets whatsoever.
-
-\begin{lemma}[$\sim^{\Lms}_{\Lscct}$ well-formed w.r.t. $\scctsafe$]\label{lem:wf:lsmslscct}
-$\;$
-
-  \begin{nscenter}
-  $\wfcsig{\sim^{\Lms}_{\Lscct}}{\scctsafe}$
-  \end{nscenter}
-\end{lemma}
-
-The last relation is $\sim^{\Lscct}_{\Lspec}$ which needs to be well-formed w.r.t. $\sssafe$.
-Similarly to the previous relation, this holds, since $\sim^{\Lscct}_{\Lspec}$ only relates $\Lspec$ traces, which do not have speculative leaks, with $\Lscct$ traces.
-
-\begin{lemma}[$\sim^{\Lscct}_{\Lspec}$ well-formed w.r.t. $\sssafe$]\label{lem:wf:lscctlspec}
-  $\wfcsig{\sim^{\Lscct}_{\Lspec}}{\sssafe}$
-\end{lemma}
-
-\section{Formal Insights}\label{sec:formalities}
-
-This section discusses how to connect each language-specific security property to the general security properties of \Cref{sec:compprop} (\Cref{subsec:formalities:maps}), and it demonstrates that the security property resulting from the universal image projection is faithful to the original property (\Cref{subsec:formalities:props}).
-% 
-Then, this section discusses why the order of compiler passes matters, and how does our framework help with identifying insecure compositions (\Cref{subsec:compatsecpasses}), and it gives additional technical insights on the secure compilation proofs (\Cref{subsec:seccompproofs}).
-
-\subsection{From Language Traces to General Ones}\label{subsec:formalities:maps}
-
-The previous theorems talk about preserving properties expressed in the trace model of the languages of each compiler.
-However, these trace models are not the same trace model we used to specify the properties of \Cref{sec:compprop} (indicated with \ev{L}), which serves as the ``ground truth'' for the meaning of our properties.
-% 
-To bridge this gap, the formal development requires specifying additional trace relations, from each of the language trace models to the \ev{L} one, that, for example, relate $\src{Get\ \loc\ n}$ and $\src{Set\ \loc\ n}$ to $\ev{Use\ \loc\ n}$ (and that induce a related universal image that we use in \Cref{subsec:formalities:props}).
-% 
-One key insight of these relations is that they all omit context-made actions for two reasons: (1) contexts (which are universally quantified) can trivially invalidate any property and (2) we are interested in the component upholding the properties.
-%
-
- % disregarded a glaring but minor technical issue: take $\tmssafe$ (\Cref{def:trace:tmsdef}) as an example.
-% It contains traces as specified in \Cref{subsec:basic:memsafety:tracemodel}.
-% However, these are different from $\Ltms$ traces (\Cref{subsec:ltms}) which means that \Cref{thm:wt:tms} is ``ill-typed''.
-% Similarly for the other theorems.
-% In the interest of reducing visual baggage, the notation left out yet another cross-language trace relation, translating language-specific traces into these non-language specific traces. 
-% In doing so, context-level actions are dropped and, e.g., both $\src{Get\ \loc\ n}$ and $\src{Set\ \loc\ n}$ are translated into $\ev{Use\ \loc\ n}$.
-
-\subsection{Security Properties and Their Meaning}\label{subsec:formalities:props}
-Each of the presented compilers use a cross-language trace relation, which is also used to translate the property from one language to the other one (via the existential or universal images).
-% 
-%Whenever such translations occur, it is important to check whether there has been a change of meaning.
-%Secure compiler engineers need to ensure that such translations of properties do not change meaning in a fundamental wrong way. 
-% lifting or lowering reasoning from target- to source-level or vice-versa,
-While the meaning of projected properties does change with a translation, the change should not allow for a flawed compilation pipeline.
-% 
-For example, we could be using a trace relation that translates a property in a language to a totally different one in another language.
-% 
-To raise the trust into the translation of properties, \Cref{thm:prop-rel-corr} states (in a general fashion) that each security property is faithfully translated using the universal image according to the cross-language trace relation induced by the compiler.
-
-\begin{theorem}[Properties Relation Correctness]\label{thm:prop-rel-corr}
-  \begin{align*}
-    \forall& \pi \in \{\tmssafe,\smssafe,\scctsafe,\sssafe\}, 
-    \\
-    \text{ for each }& \text{pair of languages } \src{L} \text{ and } \trg{L} \text{ used by the compilers},
-    \\
-    \text{ if }& 
-    % \tau_{\sim^{\trg{L}}_{\ev{L}}}(\pi) \sim^{\trg{L}}_{\ev{L}} \pi
-    \sigma_{\sim^{\trg{L}}_{\ev{L}}}(\pi) \sim^{\trg{L}}_{\ev{L}} \pi
-    \text{ and } 
-    % \sigma_{\sim^{\src{L}}_{\trg{L}}}(\tau_{\sim^{\trg{L}}_{\ev{L}}}(\pi)) \sim^{\src{L}}_{\trg{L}} \tau_{\sim^{\trg{L}}_{\ev{L}}}(\pi)
-    \sigma_{\sim^{\src{L}}_{\trg{L}}}(\sigma_{\sim^{\trg{L}}_{\ev{L}}}(\pi)) \sim^{\src{L}}_{\trg{L}} \sigma_{\sim^{\trg{L}}_{\ev{L}}}(\pi)
-    \\
-    \text{ then }& 
-    % \tau_{\sim^{\src{L}}_{\ev{L}}}(\pi) = \sigma_{\sim^{\src{L}}_{\trg{L}}}(\tau_{\sim^{\trg{L}}_{\ev{L}}}(\pi)) 
-    \sigma_{\sim^{\src{L}}_{\trg{L}}}(\sigma_{\sim^{\trg{L}}_{\ev{L}}}(\pi)) \sim^{\src{L}}_{\ev{L}} \pi
-  \end{align*}
-\end{theorem}
-The complexity of this theorem is that the relation in the conclusion cannot be obtained by composing the two relations in the premises.
-% 
-%% MK: the following does not say anything about the theorem, only about the syntactic compiler
-%What happens in practice is that compiler engineers are given the relations from each language to \ev{L} and they need to build a compiler that induces a relation $\sim^{\src{L}}_{\trg{L}}$ that respects this theorem.
-%
-
-We now informally argue why this theorem holds for the composition of all considered properties from \Cref{sec:compprop}.
-
-\paragraph{$\sigma_{\sim^{\Ltms}_{\Lms}}\left(\tmssafe\cap\smssafe\cap\scctsafe\cap\sssafe\right)$}
-
-For the four properties considered here, the trace models of $\Ltms$, $\Ltrg$, and $\Lms$ do not consider actions related to $\scctsafe$ and $\sssafe$, so these two properties are trivially translated correctly.
-% 
-We now discuss the remaining $\tmssafe$ and $\smssafe$ in the form of their intersection $\mssafe$.
-
-Recall that $\sim^{\Ltms}_{\Lms}$ is the composition of $\sim^{\Ltms}_{\Ltrg}$ and $\sim^{\Ltms}_{\Lms}$.
-% 
-Since $\sim^{\Ltms}_{\Ltrg}$ is an equality, this relation trivially preserves the meaning of translated properties: related traces are identical!
-
-Finally, let us consider a trace $\irl{\trace}\in\mssafe$ and understand what is that is related to via $\sim^{\Ltrg}_{\Lms}$.
-% 
-{\em All} traces $\trg{\trace}$ with $\trg{\trace}\sim^{\Ltrg}_{\Lms}\irl{\trace}$ are identical to $\irl{\trace}$ except for get and set actions, which require for in-bound accesses (as stated in \Cref{subsec:cs:ms}): this clearly respect $\mssafe$.
-%
-
-% In both cases, the translated event does not represent an out-of-bounds memory access.
-
-% \footnote{Both $\sim^{\Ltms}_{\Ltrg}$ and $\sim^{\Ltrg}_{\Lms}$ filter context actions.}.
-
-% Now, starting from $\mssafe$ expressed in $\Lms$, we check that its translation in $\Ltms$ is still $\mssafe$.
-% % 
-% We can see that the property resulting from the projection of $\Lms$ traces back to $\Ltms$ faithfully models $\mssafe$, which translates to both $\scctsafe$ and $\sssafe$, 
-% \MPin{ equiv  +  other rel only adds traces that fail }
-% since these properties trivially hold for $\Ltms$, $\Ltrg$, and $\Lms$ traces.
-
-\paragraph{$\sigma_{\sim^{\Lms}_{\Lscct}}\left(\tmssafe\cap\smssafe\cap\scctsafe\cap\sssafe\right)$}
-
-As before, the trace models of $\Lscct$ and $\Lms$ cannot express $\sssafe$, so that is trivially translated correctly.
-% 
-Also, the trace model of $\Lscct$ extends the one of $\Lms$ with respect to $\tmssafe$ and $\smssafe$ events, so the translation argument regarding those two properties is the same as before.
-% 
-Thus, we need to reason about whether $\scctsafe$ is translated correctly.
-
-By definition, $\sim^{\Lms}_{\Lscct}$ only relates $\obj{\unlock}$ events, which is also the same relation induced by $\sim_{\ctsafe}$ in \Cref{sec:msscct-rel}.
-% 
-This ensures that composing the relations only relates $\unlock$ events, and thus the property is translated correctly.
-
-% Consider a trace $\obj{\trace}\in\scctsafe$ and all $\irl{\trace}$ such that $\irl{\trace}\sim^{\Lms}_{\Lscct}\obj{\trace}$. 
-% In turn, $\obj{\trace}$ cannot contain any leaks of secret values, so it cannot violate $\scctsafe$.
-% Toggling data-independent timing mode has no impact on $\mssafe$ or $\sssafe$.
-
-\paragraph{$\sigma_{\sim^{\Lscct}_{\Lspec}}\left(\tmssafe\cap\smssafe\cap\scctsafe\cap\sssafe\right)$}
-The trace model of $\Lspec$ extends the one of $\Lscct$ with respect to $\tmssafe$, $\smssafe$, and $\scctsafe$, so the translation argument for those three properties is the same as before.
-% 
-Concerning $\sssafe$, $\sim^{\Lscct}_{\Lspec}$ relates $\Lspec$ speculation traces with $\Lscct$ branches with the $\lock_{\text{NONE}}$ tag, so the property is translated correctly, according to the relation defined in \Cref{sec:spec-ms-rel}.
- % introducing $\ird{barrier}$ never violates $\sssafe$ and has no effect on $\mssafe$ or $\scctsafe$.
-
-%\MPin{
-%	can this be made formal?
-%	what's the point?
-%}
-
-\subsection{Compatibility of Secure Compiler Passes}\label{subsec:compatsecpasses}
-
-% \Cref{thm:rtpsim:sig} is applicable, up to the well-formedness of $\sim_1$ w.r.t. $\trg{\class[_2]}$.
-Consider applying $\ccspec$ first and then $\ccb$ (albeit currently syntactically impossible).
-% , let us pretend that it is possible by just extending $\ccb$ for the constructs present in $\Lscct$ that are not present in $\Ltrg$.
-% Clearly, as evidenced earlier (\Cref{subsec:formalities:props}), $\ccb\cdot\ccspec$ is an acceptable composition that yields an interesting class of security properties, i.e., $\sigma_{\sim_{\irl{ms}}\bullet\sim_{\ird{\mathghost}}}\left(\mssafe\cap\scctsafe\cap\sssafe\right)$.
-In this case, $\ccb$ would insert new branches into the code that are not protected by a speculation barrier! 
-This concern is reflected in the proofs that establish that the security class resulting of the composition of the trace relations is meaningful.
-% , much similar to ensuring that a definition models what the formal methods expert intends to model. 
-For the $\ccspec\cdot\ccb$ case, the class is $\sigma_{\sim_{\ird{\mathghost}}\bullet\sim_{\irl{ms}}}\left(\mssafe\cap\scctsafe\cap\sssafe\right)$.
-Since $\ev{Use\ \loc\ n}\sim_{\irl{ms}}\irl{Branch\ n}\cdot\irl{Spec}\cdots$, where $\cdots$ does not contain a $\ev{{Barrier}}$ event, the resulting class is \emph{not} the original $\sssafe$ that is intended and it would break the corresponding \Cref{thm:prop-rel-corr}.
-% Because of this, it is crucial to analyse the precise shape of the class of properties after mapping a target-level property to a source-level property\footnote{Dito for mapping a source-level property to a target-level property.}.
-
-However, the composition is still technically possible and it is the job of the compiler engineers to ensure that the secure compilation pipeline happens in an order that ensures that the mapped security property is the intended one.
-
-\subsection{Secure Compilation Proofs}\label{subsec:seccompproofs}
-
-Our secure compilation proofs rely on backtranslations~\cite{abate2019jour,patrignani2021rsc}, which let one construct a source context starting from either target traces (aka trace-based backtranslations) or target contexts (aka context-based backtranslations).
-These backtranslations also require setting up cross-language relations between various language elements such as expressions and program states, so we leave these details for the technical reports.
-% , the usual strategy is to establish a backwards simulation between source components and their compiled target counterpart. 
-% For the universally quantified target context, a technique known as backtranslation is well-established, which constructs a source-level context either by backtranslating the target-level context, amounting to a ``decompilation'', or by constructing it using the target-level trace, so that the source-context emits parts of the whole execution trace that are related to the parts emitted by the target-level context.
-All backtranslations in the case-study are trace based except for those required by $\ccdce$ and $\cccf$, which are context-based (and they are an identity function). 
-% Since these optimisations have no significant semantic effect and source- and target-language are equal, the backtranslation amounts to an identity translation.
-
-\section{Related Work\pages{2}}\label{sec:relwork}
-
-This section discusses robust compilation, other secure compilation criteria and work related to the properties preserved in the case study.
-% (\Cref{subsec:relw:seccomprtp}) 
-% (\Cref{subsec:relw:seccompcrit}).
-% Since the case study of \Cref{sec:casestud:defs,sec:casestud:rtp} implements measures for preserving \gls*{ms}, \gls*{cct}, and \gls*{ss}, 
-% Then, this section discusses work related 
-% relevant related work as well 
-% (\Cref{subsec:relw:msmechs,subsec:relw:cctmechs,subsec:relw:ssmechs}).
-
-\paragraph*{Secure Compilation as Robust Preservation}\label{subsec:relw:seccomprtp}
-
-The robust preservation of properties as a compiler-level criterion has been analysed extensively~\cite{abate2019jour,patrignani2021rsc,abate2021extacc,patrignani2019survey} and thus we build on that framework.
-No existing work is concerned with composing secure compilers, however, existing work~\cite{abate2021extacc} sketches composition of trace-relating compiler correctness in a similar way to what has been presented here.
-%Parts of these works consider languages with different trace models and our technical setup can handle this. 
-The work relating robust preservation with universal composability~\cite{patrignani2022universal} is closest to what this paper presents.
-% 
-The authors demonstrate a similar compositionality theorem to ours (\Cref{sec:sequential}) but use it in the context of protocols.
-% 
-The work does not consider the generality to support different trace models or composition of compilers which robustly preserve different classes.
-% The authors demonstrate a similar compositionality theorem to what is presented here (\Cref{sec:sequential}) as well as in an earlier version of this work~\cite{kruse2022csc}.
-% However, they do not demonstrate the scalability of the approach by means of a case study.
-
-%\paragraph*{Composition of Secure Compilers}\label{subsec:relw:seccompcomp}
-
-There is work on lifting exploits for single compilers to the whole chain~\cite{paykin2019weird}.
-%Our framework does not consider exploits.
-%\MP{ a bit dry and short}
-While that work considers {\em in}secure compilation and composition thereof in terms of exploits, the composition they are interested in allows to lift an exploit for one compiler pass to the whole compilation chain.
-Our framework takes the opposite direction and provides compositionality results for secure compilers.
-%
-%
-%To the best of our knowledge, there does not exist published work discussing the compositionality of the robust preservation criterion, the gold-standard of secure compilation.
-
-\paragraph*{Other Secure Compilation Criteria}\label{subsec:relw:seccompcrit}
-
-While this paper focuses on the robust preservation framework~\cite{abate2019jour}, other secure compilation criteria exist.
-The survey on formal approaches to secure compilation~\cite{patrignani2019survey} discusses a broad spectrum already, while this section presents a very high-level overview.
-Fully abstract compilation~\cite{abadi1999fullabstraction} states that a compiler should preserve and reflect observational equivalence between source and target programs.
-Abate \emph{et al.}~\cite{abate2021faandrc} showed that fully abstract compilers robustly preserve program properties that are either trivial or meaningless.
-As a mitigation for this, the authors presented a categorical approach based on maps of distributive laws~\cite{watanabe2002modl}, which they call many maps of distributive laws.
-Maps of distributive laws have been investigated before as a possible secure compilation criterion~\cite{tsampas2020catsc}.
-Other approaches are extensions of the compiler correctness criterion as discussed in other work~\cite{patterson2019next700} or the introduction of opaque observations~\cite{vu2021reconciling} to reconcile compiler optimisations with security.
-Note that this work also presents secure compilers that are optimising, but contrary to the other~\cite{vu2021reconciling}, provides a formal account of these in the robust preservation framework.
-% Lastly, the authors of this paper have presented ongoing work~\cite{patrignani2023blame} on a weaker robust preservation criteria based on the concept blame.
-
-\paragraph*{Memory Safety Mechanisms}\label{subsec:relw:msmechs}
-
-Different mechanisms for enforcing memory safety exist that also consider the secure compilation domain, i.e., have an active attacker model.
-For example, the ``pointers as capabilities'' principle represents pointers as machine-level capabilities~\cite{korashy2021capableptrs}, which behave in a similar fashion to capabilities by means of linear typing~\cite{morrisett2005L3}.
-The approach of this paper also uses linear typing, but differs from $L^{3}$~\cite{morrisett2005L3} in the way that functions are not first-class.
-Moreover, this paper considers an active attacker, while the work on $L^{3}$ only discusses whole programs and, thus, has no active attacker model.
-The instrumentation to ensure memory safety that this paper presents is inspired by Softbounds~\cite{nagarakatte2009soft}.
-That work inserts bounds-checks in front of pointer-dereferences and, for this to work, inserts meta-data information on pointer creation.
-Softbounds also works in a more advanced setting with structured fields accesses and also introduces a table-lookup for pointers that are stored in memory.
-This paper only considers arrays of primitive data, i.e., there are no pointers to pointers or structures.
-Several other approaches to memory-safety exist in literature, specifically as compiler instrumentations~\cite{akritidis2009baggy,younan2010paricheck,jung2021pico,shankaranarayana2023tailcheck,dhumbumroong2020boundwarden,nam2019framer,zhou2023fatptrs}, hardware-extensions~\cite{kwon2013lowfat,saileshwar2022heapcheck,chen2023flexpointer,kim2023whistle}, or programming language extensions~\cite{elliott2018checkedc,li2022formalcheckedc,jim2002cyclone,elliott2015guilt,west2005cuckoo,weis2019fyr,benoit2019uniqueness}.
-What differentiates this work from them is that this work uses known, compiler-based approaches to ensure memory-safety as a means to investigate secure compiler compositions.
-This paper does not provide efficient memory-safety, but serves as a theoretical foundation for the secure compilation domain.
-
-To extend the languages in this paper with a less restricted form of pointer arithmetic, the region colouring memory safety monitor presented in earlier work~\cite{michael2023mswasm} can be used.
-The work presenting this monitor provides an approach for the robust preservation of memory safety compiling from C to WASM.
-However, they do not discuss composition of secure compilers but rather investigate an instance of a secure compiler.
-
-\paragraph*{Cryptographic Constant Time Mechanisms}\label{subsec:relw:cctmechs}
-
-The approach to preserving cryptographic constant time in this paper is high-level, where a programming language exposes a way to switch the semantics to a data (operand) independent timing mode.
-Since identifiers in $\Lscct$ are annotated with a secrecy tag, this approach is similar to others with information flow control.
-For example, Vale~\cite{bond2017vale} uses Dafny to ensure constant-time assembly code, while Jasmin~\cite{almeida2017jasmin} makes use of the Coq proof assistant to reject non-constant-time programs.
-CT-Wasm~\cite{watt2019ctwasm} enforces constant-timeness by means of a type system.
-Different to the approach of this paper, these approaches necessitate that the programmer writes \gls*{cct} code.
-An approach to allow programmers to write more high-level code is CryptOpt~\cite{kuepper2023cryptopt}, which generates efficient target-code by means of a randomised search.
-This paper abstracts over concrete mitigation strategies and simply assumes that there is a flag to switch to a cryptographic-constant time execution mode.
-This can be realised by employing the FaCT~\cite{cauligi2019fact} compiler, which translates common non-constant time code patterns to be constant-time, and the data (object) independent timing execution mode of modern processors.
-
-\paragraph*{Speculation Safety Mechanisms}\label{subsec:relw:ssmechs}
-
-This paper uses a taint-tracking mechanism inspired by existing work~\cite{guarnieri2018spectector,fabian2022automatic}.
-These taints are used to express absence of any speculative leaks in \gls*{ss}~\cite{guarnieri2018spectector}. 
-The semantics of $\Lspec$ hardcodes the kind of speculative leaks to just SPECTRE-PHT~\cite{kocher2019spectre}, but future work could use semantics composition~\cite{fabian2022automatic} to support more variants.
-Note that our framework composes compilers and not semantics.
-
-%This work has taken inspiration in the sematics of $\Lspec$ and in \Cref{def:trace:ss} from recent formalisations~\cite{guarnieri2018spectector,fabian2022automatic}.
-% \MPin{
-%  we take SS from exorcising.
-%  we do not take SNI from spectector because it is a hypersafety.
-%  other work defined similar props. 
-% }
-% A subsequent line of work~\cite{fabian2022automatic} shows how to modularly define and compose formal semantics that models different speculative execution variants. 
-% Recent work~\cite{fabian2024lift} demonstrates how to lift existing secure compilation proofs for security properties related to speculative execution to more powerful properties, i.e., properties defined for languages with a trace model that enables more vulnerabilities, such as combinations of different speculative execution exploits.
-% To this end, they define a lifting framework and demonstrate how a secure compilation proof for a weaker property can be lifted to a stronger one. 
-% Contrary to the compositionality framework presented in this paper, these works are only concerned with composing semantics of the respective programming languages, while in this paper, the secure compilers themselves are composed.
-
-\section{Conclusion\pages{$\sfrac{1}{2}$}}\label{sec:concl}
-This paper tackles the problem of understanding what kind of security properties a secure compiler preserves, when said compiler is the combination of compiler passes that preserve possibly different security properties.
-% 
-%For this, this paper first formalised security properties of interest and their composition.
-% 
-The paper proves that composing secure compilers that preserve certain properties results in a secure compiler that preserves the composition of these properties.
-% While the presented security property composition relied on monitors that check only trace-properties, the composition of secure compilers does not make any restriction towards the kind of property involved in the composition.
-% It is subject to future work to develop techniques to verify relational hyperproperties~\cite{abate2019jour}, while the composition of hyperproperties could be very similar to the composition of ordinary properties as presented in this paper, since hyperproperties can be checked with an automata-based model-checker~\cite{beutner23hyperltl}.
-% 
-Finally, this paper defines a multi-pass compiler and proves that it preserves \gls*{specms}.
-Crucially, this paper derives the security of the multi-pass compiler from the composition of the security properties preserved by its individual passes, which include security-preserving as well as optimisation passes.
-% For future work, it is interesting to look at more sophisticated optimisation passes that, e.g., reorder events that appear on traces.
-
-\newpage
-
-\bibliographystyle{IEEEtranS}
-\bibliography{main}
-
-%%
-%% If your work has an appendix, this is the place to put it.
-\appendix
-\subsection{Existential Image}\label{subsec:extimg}
-
-\begin{definition}[Existential Image]
-% [Existential and Universal Image]
-\label{def:existential:img}\label{def:tau}
-  \[
-    \tau_\sim\left(\src{\pi}\right) := 
-      \left\{ 
-        \trg{\trace} \mid \exists \src{\trace}\ldotp \src{\trace}\sim\trg{\trace}, \text{ and } \src{\trace}\in\src{\pi} 
-      \right\}
-  \]
-  %\[ 
-  %  \sigma_\sim\left(\trg{\pi}\right) := 
-  %    \left\{ 
-  %      \src{\trace} \mid \forall \trg{\trace}\ldotp \text{if }\src{\trace}\sim\trg{\trace}, \text{ then } \trg{\trace}\in\trg{\pi} 
-  %    \right\}
-  %\]
-\end{definition}
-
-\begin{definition}[Robust Preservation with $\tau_\sim$]\label{def:rtp:tau}
-  % $\;$\\
-  % \vspace{-1em}
-  % \begin{nscenter}
-  % Compiler $\cc{\src{L}}{\trg{L}}$ robustly preserves $\class$, 
-  {$\rtptau{\cc{\src{L}}{\trg{L}}}{\src{\class}}{\sim}$}
-  %, iff 
-  $\isdef$
-    {$\forall \src{\pi}\in\src{\class}, \src{p}\in\src{L},$} if {$\rsat{\src{\progvar}}{\src{\pi}}$}, then {$\rsat{\cc{\src{L}}{\trg{L}}\left(\src{p}\right)}{\tau_\sim\left(\src{\pi}\right)}$}.
-  % \end{nscenter}
-\end{definition}
-
-\begin{theorem}[Composition of Secure Compilers w.r.t. $\tau$]\label{thm:rtpsim:tau}
-  $\;$
-
-  If {$\rtptau{\cc{\src{L}}{\trg{L}}}{\src{\class[_{1}]}}{\sim_1}$}, {$\rtptau{\cc{\trg{L}}{\obj{L}}}{\tilde{\tau}_{\sim_1}\left(\src{\class[_2]}\right)}{\sim_2}$}, and {$\wfctau{\sim_1}{\src{\class[_2]}}$}, \\ then {$\rtptau{\cc{\src{L}}{\trg{L}}\circ\cc{\trg{L}}{\obj{L}}}{\src{\class[_{1}]}\cap\src{\class[_{2}]}}{\sim_1\bullet\sim_2}$}.
-\Coqed
-\end{theorem}
-
-\begin{corollary}[Swapping Secure Compiler Passes]\label{corr:swappable:tau}
-  $\;$
-
-  If {$\rtptau{\cc[_{1}]{\trg{L}}{\trg{L}}}{\src{\class[_{1}]}}{\sim_1}$ and $\rtptau{\cc[_{2}]{\trg{L}}{\trg{L}}}{\src{\class[_{2}]}}{\sim_2}$}, %
-  {$\wfctau{\sim_1}{\src{\class[_2]}}$ and $\wfctau{\sim_2}{\src{\class[_1]}}$}, %
-  and {$\tilde{\tau}_{\sim_1}\left(\src{\class[_2]}\right)=\src{\class[_2]}$ as well as $\tilde{\tau}_{\sim_2}\left(\src{\class[_1]}\right)=\src{\class[_1]}$},
-  then {$\rtptau{\cc[_{1}]{\trg{L}}{\trg{L}}\circ\cc[_{2}]{\trg{L}}{\trg{L}}}{\src{\class[_{1}]}\cap\src{\class[_{2}]}}{\sim_1\circ\sim_2}$ and $\rtptau{\cc[_{2}]{\trg{L}}{\trg{L}}\circ\cc[_{1}]{\trg{L}}{\trg{L}}}{\src{\class[_{2}]}\cap\src{\class[_{1}]}}{\sim_2\circ\sim_1}$}.
-\Coqed
-\end{corollary}
-
-\subsection{Secure Upper and Lower Composition}\label{sec:other-compos}
-Besides sequential composition, there are two other compositions, namely an {\em upper}, i.e., a compiler that takes multiple inputs and yields one output, and a {\em lower} composition, i.e., a compiler that takes one input and yields multiple outputs.
-We {define the upper composition $\cc{\src{L}+\obj{L}}{\trg{L}}$} as follows:
-Given a program \texttt{p}, its compiled counterpart is obtained by {plugging \texttt{p} into $\cc{\src{L}}{\trg{L}}$ if $\texttt{p}\in\src{L}$} or by {plugging \texttt{p} into $\cc{\obj{L}}{\trg{L}}$ if $\texttt{p}\in\obj{L}$}.
-\begin{definition}[Upper Composition]
-  \[
-    \text{{$\cc{\src{L}+\obj{L}}{\trg{L}}$}}\isdef
-  % Given \texttt{p}, yield
-  \lambda \texttt{p}\ldotp
-  \left\{\mbox{\begin{tabular}{c}
-    {{if $\texttt{p}\in\src{L}$, then $\cc{\src{L}}{\trg{L}}(\texttt{p})$}} \\
-    \mbox{{if $\texttt{p}\in\obj{L}$, then $\cc{\obj{L}}{\trg{L}}(\texttt{p})$}} \\
-  \end{tabular}}\right. 
-  \]
-\end{definition}
-
-Examples of this are present in industry:
-Consider the Java Virtual Machine bytecode $\trg{JVM BC}$, which is a popular target for programming language designers due to its high performance and relevance in industry.
-Compilers for several programming languages have it as their target language, some popular instances are $\src{Java}$ and $\obj{Kotlin}$.
-Technically speaking, they both compile to class files and $\obj{Kotlin}$ objects are considered to be the same as $\src{Java}$ objects at that point.
-Both languages can be used at the same time in one project~\cite{androidstudio}.
-A compiler that accepts both $\src{Java}$ and $\obj{Kotlin}$ code translating to the same target language or intermediate representation performs a kind of {\em upper} composition.
-Now, the following theorem tells us what happens if these are secure:
-Given {$\cc{\src{L}}{\trg{L}}$ robustly preserves $\class[_{1}]$} and {$\cc{\obj{L}}{\trg{L}}$ robustly preserves $\class[_{2}]$}, it follows that {their upper composition $\cc{\src{L}+\obj{L}}{\trg{L}}$ robustly preserves the intersection of classes $\class[_{1}]$ and $\class[_{2}]$}.
-
-\begin{theorem}[Upper Composition of Secure Compilers]\label{thm:urtp}
-  $\;$
-
-  If {$\rtp{\cc{\src{L}}{\trg{L}}}{\class[_{1}]}$} and {$\rtp{\cc{\obj{L}}{\trg{L}}}{\class[_{2}]}$}, then {$\rtp{\cc{\src{L}+\obj{L}}{\trg{L}}}{\class[_{1}]\cap\class[_{2}]}$}. %\Coqed
-\end{theorem}
-
-Dually, the {\em lower} composition is concerned about compilers that accept the same source but yield different target languages. %and the same theoretical results apply.
-{Define the lower composition $\cc{\src{L}}{\trg{L}+\obj{L}}$} as follows:
-Given a program $\src{p}$, its compiled counterpart is obtained by {plugging $\src{p}$ into $\cc{\src{L}}{\trg{L}}$} or by {plugging \texttt{p} into $\cc{\src{L}}{\obj{L}}$}, respectively, {based on the internal decision}.
-\begin{definition}[Lower Composition]
-  \[
-    \text{{$\cc{\src{L}}{\trg{L}+\obj{L}}$}}\isdef
-  % Given $L$ and $\src{p}$, yield
-  \lambda \src{p}, L\ldotp
-  \left\{\mbox{\begin{tabular}{c}
-    {{if $L=\trg{L}$, then} {$\cc{\src{L}}{\trg{L}}(\src{p})$}} \\
-    \mbox{{if $L=\obj{L}$, then} {$\cc{\src{L}}{\obj{L}}(\src{p})$}} \\
-  \end{tabular}}\right.\]
-\end{definition}
-
- Consider two compilers both accepting $\src{LLVM IR}$~\cite{lattner2004llvm} and one of them emits $\trg{x86\_64}$, while the other emits $\obj{ARMv8}$.
- It is intuitive that they are in some sense composed in the LLVM framework, but the decision of when to use one over the other is inherently {\em internal} to the formalisation effort of this kind of composition.
- For example, the user of this compiler provides an explicit flag that instructs to emit $\trg{x86\_64}$ or the framework itself detects the target platform via heuristics, such as supported instructions.
-
- The following theorem demonstrates what happens if the involved compilers are secure:
-Given {$\cc{\src{L}}{\trg{L}}$ robustly preserves $\class[_{1}]$} and {$\cc{\src{L}}{\obj{L}}$ robustly preserves $\class[_{2}]$}, it follows that {their lower composition $\cc{\src{L}}{\trg{L}+\obj{L}}$ robustly preserves the intersection of classes $\class[_{1}]$ and $\class[_{2}]$}.
-
-\begin{theorem}[Lower Composition of Secure Compilers]\label{thm:lrtp}
-  $\;$
-
-  If {$\rtp{\cc{\src{L}}{\trg{L}}}{\class[_{1}]}$} and {$\rtp{\cc{\obj{L}}{\trg{L}}}{\class[_{2}]}$}, then {$\rtp{\cc{\src{L}}{\trg{L}+\obj{L}}}{\class[_{1}]\cap\class[_{2}]}$}. % \Coqed
-\end{theorem}
-
-Either way, the theoretical results suggest that it is possible to always find a ``most-general'', secure compiler, given two secure compilers, that robustly preserves the least-upper bound of the classes involved in their compilation process.
-
-%Future work intends to use upper and lower composition as means to build a secure compiler incrementally
-%This follows current non-formal-methods based compiler engineering, such as LLVM~\cite{lattner2004llvm}, where languages such as \texttt{C++} and \texttt{Fortran} translate into LLVM IR and languages such as \texttt{x86} or \texttt{ARMv8} assembly variants translate out of LLVM IR.
-
-% no coq, because above definitions are non-formal. 
-%   How would you precisely define the lower compo compiler, given two compilers?
-
-%\section{Cryptographic Constant Time}\label{sec:cct}
-
-\end{document}
-\endinput

### 2025-02-13 icfp/sections/csc.tex
move

--- 2024-09-24 csf2/main.tex
@@
-% Fixing: Too many math alphabets used in version normal
-\newcommand\hmmax{0}
-\newcommand\bmmax{0}
-\documentclass[dvipsnames,conference]{IEEEtran}
-
-\usepackage[T1]{fontenc}
-\usepackage[scaled=.83]{beramono}
-
-\usepackage[colorinlistoftodos]{todonotes}
-\usepackage[inference]{semantic}
-\usepackage[switch]{lineno}
-\usepackage{halloweenmath}
-\usepackage{fontawesome5}
-\usepackage{listofitems}
-\usepackage{breakcites}
-\usepackage{glossaries}
-\usepackage{hyperref}
-\usepackage{cleveref}
-\usepackage{stmaryrd}
-\usepackage{marvosym}
-\usepackage{listings}
-\usepackage{amssymb}
-\usepackage{amsmath}
-\usepackage{nameref}
-\usepackage{amsthm}
-\usepackage{xspace}
-\usepackage{xfrac}
-\usepackage{tikz}
-\usepackage{soul}
-\usepackage{bm}
-\usepackage{paralist}
-
-\hypersetup{
-    colorlinks,
-    linkcolor={red!50!black},
-    citecolor={blue!50!black},
-    urlcolor={blue!80!black}
-}
-%\renewcommand\UrlFont{\color{blue}\rmfamily}
-%\newcommand{\url}[1]{\lstinline{#1}}
-
-\setul{0.95ex}{0.3ex}
-
-\input{cmds.tex}
-
-% allow page breaks inside multiline alignment displays
-\allowdisplaybreaks
-
-\Crefname{exampleenv}{Example}{Examples}
-
-\theoremstyle{definition}
-\newtheorem{exampleenv}{Example}[section]
-\newtheorem{lemma}{Lemma}[section]
-\newtheorem{theorem}{Theorem}[section]
-\newtheorem{corollary}{Corollary}[section]
-\newtheorem{definition}{Definition}[section]
-
-\loadglsentries{acronyms}
-\makeglossaries
-
-\AtBeginDocument{%
-  \providecommand\BibTeX{{%
-    \normalfont B\kern-0.5em{\scshape i\kern-0.25em b}\kern-0.8em\TeX}}}
-\makeatletter
-\newcommand{\linebreakand}{%
-  \end{@IEEEauthorhalign}
-  \hfill\mbox{}\par
-  \mbox{}\hfill\begin{@IEEEauthorhalign}
-}
-\makeatother
-
-\renewcommand\thelinenumber{\color{red}\arabic{linenumber}}
-\begin{document}
-\linenumbers
-
-\title{
-  Secure Composition 
-  \\ 
-  of Robust and Optimising Compilers
-  % \vspace{-1.5em}
-}
-
-% \author{
-% Matthis Kruse\orcidID{0000-0003-4062-9666}
-% \and
-% Michael Backes
-% \and
-% Marco Patrignani
-% }
-% \authorrunning{M. Kruse et al.}
-
-\maketitle
-
-\begin{abstract}
-% \MPpost{
-%   this abstract fails in setting up the robustness argument.
-%   it'll be a good exercise for you to take inspiration from this, but then write something as concise but with a more direct focus on robust properties
-% }
-% context
-To ensure that secure applications do not leak their secrets, they are required to uphold several security properties such as spatial and temporal memory safety, cryptographic constant time, as well as speculative safety.
-% need
-Existing work shows how to enforce these properties individually, in an architecture-independent way, by using secure compiler passes that each focus on an individual property.
-% task
-Unfortunately, given two secure compiler passes that each preserve a possibly different security property, it is unclear what kind of security property is preserved by the composition of those secure compiler passes.
-%there is no way to tell what kind of security property will the composition of those secure compilers preserve.
-
-% object
-This paper is the first to study what security properties are preserved across the composition of different secure compiler passes.
-% findings
-Starting from a general theory of property composition for security-relevant properties (such as the aforementioned ones), this paper formalises a theory of composition of secure compilers.
-Then, it showcases this theory on a secure multi-pass compiler that preserves the aforementioned security-relevant properties.
-% conclusion
-Crucially, this paper derives the security of the multi-pass compiler from the composition of the security properties preserved by its individual passes, which include security-preserving as well as optimisation passes.
-% 
-From an engineering perspective, this is the desirable approach to building secure compilers.
-\begin{center}\small\it
-	{This paper uses syntax highlighting accessible to both colourblind and black \& white readers.
-	% ~\cite{patrignani2020use}.
-% 	% Specifically, it makes use of a $\src{blue}$, $\src{sans\text{-}serif}$ font for a $\src{source}$,
-% 	% a $\trg{red}$, $\trg{bold}$ font for an $\trg{intermediate}$,
-% 	% and a $\obj{green}$, $\obj{teletype}$ font for a $\obj{target}$ language.	
-	For a better experience, please print or view in colours.
-	}
-\end{center}
-\end{abstract}
-
-\begin{IEEEkeywords}
-  memory-safety, cryptographic constant-time, secure compilation, robustness
-\end{IEEEkeywords}
-
-\section{Introduction\pages{4}}\label{sec:introduction}
-
-% \gls*{ms} is a security property that ensures that there are no out-of-bounds accesses as well as no reads or writes to uninitialised memory.
-% Programs that do not attain memory safety, can easily have their private data be leaked to attackers or be corrupted~\cite{lemay2021ccc}.
-% Because of this, \gls*{ms} is a requirement for programs to attain an even stronger security property, namely \gls*{cct}.
-% Programs that attain \gls*{cct} are secure w.r.t.
-timing attacks, and this is the gold standard for secure applications, including cryptographic ones, since timing attacks can expose secrets via a simple performance analysis~\cite{kocher1996timing}.
-% Timing attacks have become more and more relevant recently, especially given the era of cloud-computing~\cite{aviram2010cloudtime,kumar2019cloudsecsurvey}, where programs share the same hardware but are supposed to run in a sandboxed environment that does not leak secrets.
-% Unfortunately, several attacks within such a cloud environment have been demonstrated as viable strategies to extract secrets~\cite{mehmet2015getoff,flowers2022zeroday,atya2019catchme,venkatanathan2015placevul}.
-
-% Even though mitigations for such attacks exist~\cite{bond2017vale,almeida2017jasmin}, some compilers may optimise these mitigations away~\cite{barthe2018sec}.
-% This is of course not what programmers expect and to ensure such mitigations persist in the compiled code, one can use a {\em secure} compilers.
-% More generally, secure compilers should preserve security properties even when the program runs in a hostile context such as the cloud one, i.e., where it interacts with other possibly malicious programs.
-% An example of such secure compilers are those that ascribe to the theory of robust compilation~\cite{abate2019jour,abate2021extacc,patrignani2021rsc}, which takes this interaction between components and potentially malicious contexts into account.
-
-% Unfortunately, it is unclear if and how robust compilation supports compositionality, and compilation passes are typically developed as a standalone module to be plugged into a compilation pipeline and composed with other passes.
-% Different kinds of compilation passes exist, e.g., there are usually several compilation passes that perform optimisations~\cite{lattner2004llvm,googlev8,androidstudio,kuepper2023cryptopt,manjikian1997fusion,wegman1991ccp} or source-code instrumentations~\cite{nagarakatte2009soft,nagarakatte2010cets,akritidis2009baggy,dhumbumroong2020boundwarden,jung2021pico,nam2019framer,shankaranarayana2023tailcheck,younan2010paricheck,zhou2023fatptrs,bond2017vale,almeida2017jasmin,kuepper2023cryptopt,cauligi2019fact} that enforce that the compiled program fulfils a property of interest.
-% Without a framework for composing secure compiler passes, it is not possible to enable separation of concerns, e.g., to have a secure compilation pass that ensures \gls*{ms} that is developed independently of another secure pass for \gls*{cct} as suggested before and as showcased in \Cref{ex:strncpy}.
-% \MPpost{
-%   this intro does a poor job at introducing the robust compilation argument.
-%   perhaps it'd be best to start from robust properties, which motivates the use of robust compilers?
-%   idk
-%   as for the abstract, this is a good starting point for you to make a better intro
-% }
-
-\gls*{ms} is a security property obtained by composing \gls*{sms}, which ensures array accesses are all within bounds, and \gls*{tms}, which ensures pointers are only used when they are valid~\cite{azevedo2018meaningsofms,jim2002cyclone,necula2005ccured,nagarakatte2010cets,nagarakatte2009soft,akritidis2009baggy,michael2023mswasm}.
-\gls*{cct} is a security property that ensures sensitive data is not leaked via timing side-channels~\cite{kocher1996timing} and \gls*{ss} is a security property that enforces the same but under a speculative semantics~\cite{guarnieri2018spectector,fabian2022automatic} that captures speculative execution attacks such as spectre~\cite{kocher2019spectre}.
-Together, \gls*{sms}, \gls*{tms}, \gls*{cct} and \gls*{ss}, yield \gls*{specms}, which is the gold standard of security properties for secure applications.
-Programs attaining \gls*{specms} do not leak sensitive data either through erroneous memory accesses, nor through timing side-channels, even under speculative execution.
-\Cref{ex:strncpy} below discusses how these security properties can be enforced by compiler passes~\cite{bond2017vale,almeida2017jasmin}, to ensure programmers need not be aware of the architectural details of where their code will run.
-
-\begin{exampleenv}[strncpy]\label{ex:strncpy}
-Consider the following \texttt{C} function \texttt{strncpy} that copies a null-terminated string \texttt{src} into \texttt{dst} up to a length of \texttt{n} characters.
-This function is subject to a subtle \gls*{sms} vulnerability: the bounds check \texttt{i < n} should happen {\it before} the access to memory location \texttt{src[i]}: otherwise
-% If \texttt{n} is at least as large as the length of the array \texttt{src}, 
-the memory location past the last element is leaked to an attacker.
-\begin{lstlisting}[language=c,basicstyle=\small\ttfamily,morekeywords={size_t}]
-void strncpy(size_t n, char *dst, char *src) {
-  for(size_t i = 0; src[i] != '\0' && i < n; ++i) {
-    dst[i] = src[i];
-  }
-}
-\end{lstlisting}
-% Assume that the length of the array pointed by \texttt{dst} is at least \texttt{n}, and \texttt{src} is terminated by \texttt{'\textbackslash 0'}.
-% Even under these assumptions, there is a 
-To prevent this vulnerability, one can use a compilation pass that enforces \gls*{sms}, such as Softbounds~\cite{nagarakatte2009soft} or BaggyBounds~\cite{akritidis2009baggy}.
-The most na\"ive solution in this case is to insert bounds-checks in front of every access to memory.
-
-Because of timing attacks, addressing \gls*{sms} is not enough to make \texttt{strncpy} secure.
-In fact, the loop can terminate early, as soon as the string-terminating character \texttt{'\textbackslash 0'} is encountered, thus making program execution time proportional to the length of the array pointed by \texttt{src}, and violating the \gls*{cct} property.
-% Often, programmers rewrite programs like these to have loops that run with a fixed number of iterations.
-Also in this case there exist compiler passes that can rewrite such programs into \gls*{cct} ones~\cite{cauligi2019fact}.
- % so that they are no longer vulnerable to timing attacks.
-%
-Finally, even with these precautions, code is not run in isolation, so a malicious attacker could supply code that interacts with \texttt{strncpy} and trigger a violation of either \gls*{ms} or \gls*{cct} by calling \texttt{strncpy} with an argument for \texttt{src} that points to uninitialised memory.
-This would, in turn, trigger a series of reads from uninitialised memory, which is an immediate \gls*{ms} violation with devastating real-world consequences~\cite{uninit-0,uninit-1,uninit-2,uninit-3,uninit-4}.
-% An easy fix to sidestep this is to use non-nullable pointers as parameters instead, e.g., as presented in Cyclone~\cite{jim2002cyclone}.
-\end{exampleenv}
-
-  % last sentence is before the example, which talks about compiler passes enforcing props
-Whether or not compiler passes enforce certain security properties, it is important for a secure compiler to consider partial programs that interact with potentially malicious code, since the latter may lead to, e.g., memory-safety issues in the considered partial program.
-Robust compilers~\cite{abate2019jour} are a form of secure compilers that preserve security properties even in the presence of arbitrary attackers interacting with compiled code.
-Thus, robust compilers can be used to prevent vulnerabilities resulting from uninitialised memory (as well as many other ones), e.g., by targeting capability-based languages such as CHERI~\cite{woodruff2014CHERI}, Arm Morello~\cite{arm-morello}, or MSWasm~\cite{michael2023mswasm}, where the compiler relies on capabilities to check that pointers are always initialised.
-
-% To ensure these compiler-preserved properties still hold even in the presence of arbitrary attackers interacting with compiled code, one can use secure compilers in the form of robust compilers~\cite{abate2019jour,abate2021extacc,patrignani2021rsc}.
-
-Unfortunately, given secure compiler passes that each preserve a possibly different security property, there is no way to tell what kind of security property will the composition of those secure compilers preserve.
-% Different kinds of compilation passes exist, e.g., there are usually several compilation passes that perform optimisations~\cite{lattner2004llvm,googlev8,androidstudio,kuepper2023cryptopt,manjikian1997fusion,wegman1991ccp} or source-code instrumentations~\cite{nagarakatte2009soft,nagarakatte2010cets,akritidis2009baggy,dhumbumroong2020boundwarden,jung2021pico,nam2019framer,shankaranarayana2023tailcheck,younan2010paricheck,zhou2023fatptrs,bond2017vale,almeida2017jasmin,kuepper2023cryptopt,cauligi2019fact} that enforce that the compiled program fulfils a property of interest.
-Worse, without a framework for composing secure compiler passes, it is not possible to enable separation of concerns, e.g., to have a secure compilation pass that ensures \gls*{ms} that is developed independently of another secure pass for \gls*{cct}, that is developed independently of other passes, such as optimisation ones.
-
-This paper introduces a framework for reasoning about the composition of secure and optimising compiler passes %akin to those of \Cref{ex:strncpy}\MP{there aren't really passes there} 
-and it showcases the power of this framework by instantiating it on a multi-pass compilation chain.
-To this end, this paper first discusses how to compose security properties, such as \gls*{tms} and \gls*{sms} into \gls*{ms}, and then adds \gls*{cct} as well as \gls*{ss} into the mix to obtain \gls*{specms}.
-%Then, this paper defines compiler composition and formalises that given two passes that securely preserve two (possibly distinct) properties, their composition securely preserves the composition of those properties.
-The paper then defines several secure compiler passes, where each is either preserving a different security property from the list above
-% (\gls*{tms}, \gls*{sms}, \gls*{scct}, \gls*{ss}) 
-or performing a security-preserving optimisation.
-% , (e.g., applying \gls*{cf} or \gls*{dce}).
-Finally, this paper shows that composing these secure compiler passes into a multi-pass compilation chain results in the end-to-end (robust) preservation of \gls*{specms}.
-Crucially, this paper derives the security of the multi-pass compiler from the composition of the security properties preserved by its individual passes.
-% Crucially, this end-to-end security preservation is obtained by composing the security property preserved by each individual pass, as dictated by the framework of this paper.
-This result showcases how the framework allows the kind of formal security reasoning that compiler writers already want (and already do), obtaining precise, compositional security reasoning while providing minimal (and modular) proof effort.
-%
-In summary, this paper makes the following contributions:
-% \begin{itemize}[leftmargin=0em,nosep,label=$\blacktriangleright$]
-\begin{asparaitem}[$\blacktriangleright$]
-  \item %
-        %\Cref{sec:sequential} studies different forms of secure compiler compositions.
-        This paper takes the secure compilation framework of Abate et al.~\cite{abate2021extacc} and extends it for compositionality. (\Cref{sec:sequential}).
-        %Concretely, it investigates upper and lower compositions, as well as the sequential composition.
-        %For this, this paper studies compiler composition as well as compilers with multiple input languages or multiple output ones, as used in existing compilation chains.
-        %
-        This paper proves that starting from two compilers that preserve two (possibly distinct) properties, their composition preserves roughly the intersection of those properties.
-        % For example, a compiler that is secure w.r.t. \gls*{tms} can be composed with a compiler that is secure w.r.t. \gls*{sms}, resulting in a compiler that is secure w.r.t. \gls*{ms}.
-        % 
-        % From this, the paper proves that the order of composition of sequential compiler passes is irrelevant for security.
-        % % 
-        % This is crucial for reordering individual optimisation passes and thus generating secure and efficient code.
-        %
-        %Finally, this paper extends parts of the compositionality framework to support different trace-models between compilation passes.
-        % \MP{this isn't really true, is it?. the order matters for the relations' lemmas}
-        Then, this paper identifies which conditions make the composition of secure compilers meaningful from a security perspective, and which other conditions allow passes to be swapped without losing security meaning.
-
-  \item %
-        This paper presents a case-study with five programming languages showcasing the previous contribution (\Cref{sec:casestud:defs,sec:casestud:rtp}).
-        To this end, it presents a compilation chain consisting of six passes that ultimately preserve \gls*{specms} by means of composing four secure passes that \emph{individually} preserve \gls*{tms}, \gls*{sms}, \gls*{scct}---an enforceable version of \gls*{cct}---, and \gls*{ss}, and two passes which are well-known optimisations \gls*{dce} and \gls*{cf}.
-        %Furthermore, the chain includes two optimisation passes: one performs \gls*{dce} and the other \gls*{cf}.
-        The formalisation of this case study showcases the power of the presented framework: the divide-and-conquer approach to software engineering is a viable strategy even for the development of secure compilers.
-%        This demonstrates that the usage of the presented framework is both simpler and more expressive compared to one monolithic security proof of the whole compilation chain.
-
-  \item %
-  		This paper provides formal proofs of the security properties preserved by each of the presented passes. 
-  		% 
-  		Additionally this paper analyses each pass in terms of their compatibility with each other (\Cref{sec:formalities}), proving how does each pass fulfill the conditions that make their composition meaningful from a security standpoint.
-        % In doing so, the paper provides formal insights into some technicalities of the concrete proofs necessary as technical setup to reap the benefits of the presented compositionality framework.
-
-  \item The key contributions of this paper are formalised in the Coq proof assistant and the paper indicates this with \CoqSymbol.
-\end{asparaitem}
-% 
-This paper starts by introducing relevant notions of security properties and secure compilation (\Cref{sec:background}),
-and discusses related work (\Cref{sec:relwork}) before concluding (\Cref{sec:concl}).
-
-% \contribution{Technical Report} 
-The omitted formal details, lemmas and proofs, as well as the Coq formalisation are available as supplementary material.
-% \MKin{cite/link}
-
-\section{Background: Properties and Secure Compilers\pages{1}}\label{sec:background}
-
-To introduce the security argument of this paper, this section defines (security) properties, their satisfaction, and their robust satisfaction (i.e., satisfaction in the presence of an active attacker; \Cref{subsec:bg:tprop}).
-Then, borrowing from existing work~\cite{abate2019jour,abate2021extacc}, this section introduces secure compilers as compilers that preserve robust property satisfaction (\Cref{subsec:bg:rtp}).
-
-\subsection{Properties and (Robust) Satisfaction}\label{subsec:bg:tprop}
-
-This paper employs the security model where programs are written in a language whose semantics emits events $\event$.
-Events include security-relevant actions (e.g., reading from and writing to memory, as detailed in \Cref{sec:compprop}) and the unobservable event $\emptyevent$.
-As programs execute, their emitted events are concatenated in traces $\trace$, which serve as the description of the behaviour of a program.%
-\footnote{
-Throughout the paper, sequences are indicated with an overbar (i.e., $\trace$), empty sequences with $\hole{\cdot}$, and concatenation of sequences $\trace[_{1}],\trace[_{2}]$ as $\trace[_{1}]\cdot\trace[_{2}]$.
-Prepending element $\event$ to a sequence $\trace$ uses the same notation: $\event\cdot\trace$.
-}
-
-Properties $\pi$ are sets of traces of admissible program behaviours, ascribing what said property considers valid.
-% 
-The set of all properties can be partitioned into different {\em classes} ($\class$), i.e., safety, liveness, and neither safety nor liveness~\cite{clarkson2008hyper}, so a class is simply a set of properties.
-% 
-The compositionality framework (\Cref{sec:sequential}) presented in this paper considers arbitrary classes, while the case-study (\Cref{sec:casestud:rtp}) fixes them to concrete instances of safety properties, since it is decidable whether a trace satisfies a safety property with just a finite trace (i.e., a \emph{prefix}).
-As an example, consider the trace:
-$$\ev{Dealloc\ \loc}\cdot\ev{Read\ \loc\ 1729}\cdot\dots$$
-which describes the interaction with a memory where the deallocation of an address $\loc$ precedes a read (of some value $1729$) at that address in memory.
-% 
-This program behaviour is insecure w.r.t.
-a canonical notion of (temporal) memory safety dictating no use-after-frees of pointers~\cite{nagarakatte2010cets,azevedo2018meaningsofms}, because it reads from a memory location that was freed already.
-The previous finite trace prefix is enough to decide that the trace does not satisfy \gls*{tms} and there is no way to append events to this prefix which would result in the trace being admissible.
-
-In the following, the execution of a whole program $\wholeprogvar$ that terminates in state $\runtimetermvar$ according to the language semantics and produces trace $\trace$ is written as $\progstepto{\wholeprogvar}{\runtimetermvar}{\trace}$.
-With this, we defined property satisfaction as follows:
-\bul{a whole program $\wholeprogvar$ satisfies a property $\pi$} iff \iul{$\wholeprogvar$ yields a trace $\trace$} such that \oul{$\trace$ satisfies $\pi$}.
-
-\begin{definition}[Property Satisfaction]\label{def:propsat}
-$\;$
-
-  \begin{nscenter}
-    \bul{$\sat{\wholeprogvar}{\pi}$}
-    $\isdef$
-    if \iul{$\forall\runtimetermvar\ \trace,\progstepto{\wholeprogvar}{\runtimetermvar}{\trace}$},
-    then \oul{$\trace\in\pi$}.
-  \end{nscenter}
-\end{definition}
-
-Property satisfaction is defined on whole programs, i.e., programs without missing definitions.
-Thus, from a security perspective, \Cref{def:propsat} considers only a passive attacker model, where the attacker observes the execution and, e.g., retrieves secrets from that.
-To consider a stronger model similarly to what existing work does~\cite{abate2019jour,abate2021extacc,maffeis2008code-carrying,gordon2003authenticity,fournet2007authorization,bengtson2011refine,backes2014uniontyps,michael2023mswasm,swasey2017robust,sammler2019benefits}, we extend the concept of satisfaction with {\em robustness}.
-Robust satisfaction considers partial programs $\progvar$, i.e., components with missing imports, which cannot run until said imports are fulfilled.
-To remedy this, {\em linking} takes two partial programs $\progvar[_{1}],\progvar[_{2}]$ and produces a whole program $\wholeprogvar$, i.e., $\link{\progvar[_{1}]}{\progvar[_{2}]}=\wholeprogvar$.
-As typically done in works that consider the execution of partial programs~\cite{abate2019jour,devriese2018parametricity,patrignani2021rsc,korashy2021capableptrs,strydonck2019lincap,devriese2017modular,bowman2015noninterference,ahmed2011equivcps,patterson2017linkingtyps},
-this paper assumes that whole programs are the result of linking partial programs referred to as {\em context} ($\ctx$) and {\em component} ($\comp$).
-The context is an arbitrary program and thus has the role of an {\em attacker} that can interact with the component by means of any features the programming language has, and the component is what is security-relevant.
-%In this work, the semantics of the programming language is expected to differentiate between the component and the context.
-With this, \Thmref{def:propsat} can be extended as follows: for \bul{a component $\progvar$ to robustly satisfy a property $\pi$}, take an \iul{attacker context $\contextvar$ and link it with $\progvar$}, \oul{the resulting whole program must satisfy $\pi$}.
-
-\begin{definition}[Robust Satisfaction]\label{def:proprsat}
-  $\;$
-
-  \begin{nscenter}
-  \bul{$\rsat{\progvar}{\pi}$}
-  $\isdef$ \iul{$\forall \contextvar\ \wholeprogvar$, if $\link{\contextvar}{\progvar}=\wholeprogvar$}, then \oul{$\sat{\wholeprogvar}{\pi}$}.
-  \end{nscenter}
-\end{definition}
-
-% \begin{exampleenv}[Double Free in Bluetooth Subsystem]
-%   Consider \texttt{CVE-2021-3564}~\cite{doublefree-bluetooth}, one of many submissions for a double-free vulnerability.
-%   The vulnerability arises due to a race condition where the context-level function \texttt{hci\_cmd\_work} was not expected to behave maliciously, since it resides in the same source-code repository where the vulnerability occurs.
-%   Nevertheless, the component-level code of \texttt{hci\_dev\_do\_open} is linked with \texttt{hci\_cmd\_work} and does not atomically check whether a pointer has been freed already:
-%   Therefore, \texttt{hci\_dev\_do\_open} does not satisfy the no-double-frees property robustly, since there is an implementation for \texttt{hci\_cmd\_work} that leads to a violation of that property when linked with \texttt{hci\_dev\_do\_open}.
-% \end{exampleenv}
-
-\subsection{Secure Compilers}\label{subsec:bg:rtp}
-
-A {\em compiler} ($\cc{\src{L}}{\trg{L}}$) translates syntactic descriptions of programs from a {\em source} ($\src{L}$) into a {\em target} ($\trg{L}$) programming language.
-This translation is considered {\em correct} if it is semantics-preserving.
-That is, for a whole program $\src{\wholeprogvar}$, the compiler should relate the $\src{L}$ semantics of $\src{\wholeprogvar}$ with the semantics of $\trg{L}$ of the compiled counterpart of $\src{p}$ in such a way that they are ``compatible''.
-Unfortunately, correct compilers may be insecure compilers~\cite{patrignani2019survey,kennedy2006secure.net,abadi1999protect,ahmed2018dagstuhl} and programs translated by insecure compilers can violate security properties that the programmer assumes to hold.
-This is why {\em robust preservation} is a good candidate as a compiler-level security property~\cite{abate2019jour}.
-
-This paper uses a general notion of robust preservation~\cite{abate2021extacc} that considers compilers that use languages with different trace models. 
-To this end, considering a source trace $\src{\trace}$ and a target trace $\trg{\trace}$, there is a relation ($\sim$) describing the effect of a corresponding compiler (see \Cref{sec:casestud:rtp}). 
-This relation induces the following two projection functions~\cite{abate2021extacc}: (1) the \emph{existential image} $\tau_\sim\left(\src{\pi}\right)$ and (2) the \emph{universal image} $\sigma_\sim\left(\trg{\pi}\right)$.
-These projections map source-level (resp. target-level) properties to target-level (resp. source-level) properties in a way that identifies the ``same'' property across languages. 
-The case study of this paper uses the universal image, since some considered properties, such as \gls*{ss}, are not definable in a higher-level language that, e.g., does not model speculation.
-% For self-containment, the universal image is defined as follows~\cite{abate2021extacc}:
-% 
-\begin{definition}[Universal Image]
-% [Existential and Universal Image]
-\label{def:universal:img}\label{def:sigma}
-  % \[
-  %   \tau_\sim\left(\src{\pi}\right) := 
-  %     \left\{ 
-  %       \trg{\trace} \mid \exists \src{\trace}\ldotp \src{\trace}\sim\trg{\trace}, \text{ and } \src{\trace}\in\src{\pi} 
-  %     \right\}
-  % \]
-  \[ 
-    \sigma_\sim\left(\trg{\pi}\right) := 
-      \left\{ 
-        \src{\trace} \mid \forall \trg{\trace}\ldotp \text{if }\src{\trace}\sim\trg{\trace}, \text{ then } \trg{\trace}\in\trg{\pi} 
-      \right\}
-  \]
-\end{definition}
-\noindent
-With this projection function, we define a more general version of robust preservation as follows~\cite{abate2021extacc}.
-% For a \bul{compiler $\cc{\src{L}}{\trg{L}}$ to robustly preserve a class of source properties $\src{\class}$}, if for any \rul{property $\src{\pi}$ of that class $\src{\class}$ and programs $\src{p}$ written in $\src{L}$} where \iul{$\src{\progvar}$ robustly satisfies $\src{\pi}$}, then \oul{the compilation of $\src{\progvar}$, $\cc{\src{L}}{\trg{L}}\left(\src{p}\right)$, must robustly satisfy $\tau_\sim\left(\src{\pi}\right)$}.
-% 
-% \begin{definition}[Robust Preservation with $\tau_\sim$]\label{def:rtp:tau}
-%   % $\;$\\
-%   % \vspace{-1em}
-%   % \begin{nscenter}
-%   % Compiler $\cc{\src{L}}{\trg{L}}$ robustly preserves $\class$, 
-%   \bul{$\rtptau{\cc{\src{L}}{\trg{L}}}{\src{\class}}{\sim}$}
-%   %, iff 
-%   $\isdef$
-%   \rul{$\forall \src{\pi}\in\src{\class}, \src{p}\in\src{L},$} if \iul{$\rsat{\src{\progvar}}{\src{\pi}}$}, then \oul{$\rsat{\cc{\src{L}}{\trg{L}}\left(\src{p}\right)}{\tau_\sim{\left(\src{\pi}\right)}}$}.
-%   % \end{nscenter}
-% \end{definition}
-% 
-A \bul{compiler $\cc{\src{L}}{\trg{L}}$ robustly preserves a class of target properties $\trg{\class}$}, if for any \rul{property $\trg{\pi}$ of class $\trg{\class}$ and programs $\src{p}$}, where \iul{$\src{\progvar}$ robustly satisfies $\sigma_\sim\left(\trg{\pi}\right)$}, \oul{the compilation of $\src{\progvar}$, we have that $\cc{\src{L}}{\trg{L}}\left(\src{p}\right)$ robustly satisfies $\trg{\pi}$}.
-
-\begin{definition}[Robust Preservation with $\sigma_\sim$]\label{def:rtp:sigma}
-  % $\;$\\
-  % \vspace{-1em}
-  % \begin{nscenter}
-  % Compiler $\cc{\src{L}}{\trg{L}}$ robustly preserves $\class$, 
-  \bul{$\rtpsig{\cc{\src{L}}{\trg{L}}}{\trg{\class}}{\sim}$}
-  %, iff 
-  $\isdef$
-  \rul{$\forall \trg{\pi}\in\trg{\class}, \src{p}\in\src{L},$} if \iul{$\rsat{\src{\progvar}}{\sigma_\sim\left(\trg{\pi}\right)}$}, then \oul{$\rsat{\cc{\src{L}}{\trg{L}}\left(\src{p}\right)}{\trg{\pi}}$}.
-  % \end{nscenter}
-\end{definition}
-Note that a class of properties $\class$ can represent just one property $\pi$ by lifting~\cite{clarkson2008hyper} that property to sets of properties, i.e., use the powerset of $\pi$ instead of $\pi$ itself.
-Because of this, this paper may write $\rtptau{\cc{\src{L}}{\trg{L}}}{\trg{\pi}}{\sim}$, even though $\trg{\pi}$ is a property and not a class.
-A similar construction can be used to the projection functions (see \Cref{def:universal:img}) by applying them to the lifted version of $\trg{\pi}$ instead of just $\trg{\pi}$.
-
-In case the trace model is the same for both source and target programs (and thus $\sim$ is equality), we obtain~\cite{abate2019jour}:
-\begin{definition}[Robust Preservation]\label{def:rtp}
-  % $\;$\\
-  % \vspace{-1em}
-  % \begin{nscenter}
-  % Compiler $\cc{\src{L}}{\trg{L}}$ robustly preserves $\class$, 
-  $\;$
-
-  {$\rtp{\cc{\src{L}}{\trg{L}}}{\class}$}
-  %, iff 
-  $\isdef$
-  {$\forall \pi\in\class, \src{p}\in\src{L},$} if {$\rsat{\src{\progvar}}{\pi}$}, then {$\rsat{\cc{\src{L}}{\trg{L}}\left(\src{p}\right)}{\pi}$}.
-  % \end{nscenter}
-\end{definition}
-
-Examples of compilers fulfilling \Cref{def:rtp} exist in the literature~\cite{korashy2022secureptrs,korashy2021capableptrs,abate2021extacc,abate2019jour,patrignani2021rsc}.
-For example, SecurePtrs~\cite{korashy2022secureptrs} gives a compiler that robustly preserves all safety properties for a C-like language to an assembly-like language. 
-% The mechanization has well over 20k lines in total, which demonstrates that proving a compiler secure with respect to \Cref{def:rtp} is a significant effort. \MP{move this line to later ?}
-% i don't see how adding a def of cct is beneficial
-As another example, even though it is not strictly satisfying \Cref{def:rtp}, the FaCT~\cite{cauligi2019fact} compiler preserves the \gls*{cct} property for a C-like language with constant-time primitives, e.g., \texttt{ctselect} for branching.
-%, whose precise definition can be found in \Cref{sec:cct}.
-Throughout this work, it is assumed that FaCT satisfies \Cref{def:rtp}.
-% While we were not able to find compilers in literature that provably fulfill \Cref{def:rtp}, we anticipate that 
- % would satisfy this hyperproperty.
-% FaCT proves constant-timeness by means of overapproximation, i.e., instead of directly proving the hyperproperty, they fall back to a property that overapproximates the corresponding hyperproperty.
-
-%\smallskip%MP: the underline compacts the sections too much
-
-\section{Secure Composition}\label{sec:sequential}
-
-Notably, real-world compilation chains also perform a series of (sequential) passes whose main purpose is not necessarily to translate from one language to another, but to, e.g., optimise the code or enforce a certain property.
-Both examples can be seen in practice, e.g.,~\cite{nagarakatte2009soft,nagarakatte2010cets,akritidis2009baggy,wegman1991ccp,manjikian1997fusion} and many more.
-% \MPin{survey to cite?
-this is a bit informal MK: hard to find a survey on different compiler passes...
-I can't even find one just for compiler enforced memsafety (that looks good) or for compiler translations in general}.
-% 
-% Consider the compilation passes of LLVM~\cite{lattner2004llvm}, which are handled in the \texttt{PassManager} class.
-% By manipulating the internal state of a \texttt{PassManager} object, the schedule of compiler optimisation passes can be hardcoded.
-% That way, an order of optimisations can be found by hand that yields the best results.
-% \MP{what's the point of the previous 3 sentences?(commented)}
-Consider the following two LLVM optimisation passes: \gls*{cf}, which rewrites constant expressions to the constant they evaluate to, and \gls*{dce}, which removes dead code by rewriting conditional branches.
-The order in which \gls*{cf} and \gls*{dce} are performed influences the final result of the compilation (see \Cref{fig:cfdceex}).
-\begin{figure}[!ht]
-  \begin{center}
-    \begin{tikzpicture}[scale=0.62, every node/.style={transform shape}]
-      \node[minimum width=3cm,minimum height=2.5cm,draw=black,very thick,align=left] (progunopt) {\begin{lstlisting}[language=ml,basicstyle=\small\ttfamily]
-let a = true in
-if a then
-  print "a"
-else
-  print "b"
-\end{lstlisting}};
-      \node[minimum width=3cm,minimum height=2.5cm,draw=black,very thick,align=left,right=1.0 of progunopt] (progoptcf) {\begin{lstlisting}[language=ml,basicstyle=\large\ttfamily]
-if true then
-  print "a"
-else
-  print "b"
-\end{lstlisting}};
-      \node[minimum width=3cm,minimum height=2.5cm,draw=black,very thick,align=left,right=1.0 of progoptcf] (progoptcfdce) {\begin{lstlisting}[language=ml,basicstyle=\Large\ttfamily]
-print "a"
-\end{lstlisting}};% the following is a hack to get my syntax highlighting to work again \end{lstlisting}
-      % edges
-      \draw[->,very thick] (progunopt) edge[loop left,left] node {\gls*{dce}} (progunopt);
-      \draw[->,very thick] (progunopt) edge[sloped,above] node {\gls*{cf}} (progoptcf);
-      \draw[->,very thick] (progoptcf) edge[sloped,above] node {\gls*{dce}} (progoptcfdce);
-    \end{tikzpicture}
-  \end{center}
-  \caption{Example program where the level of optimisations differ for one pass of applying \gls*{cf} and \gls*{dce} in any order. %
-  Every edge is a compilation pass and the label on the edge states what the pass does, i.e., \gls*{cf} or \gls*{dce}. %
-  The source code in the nodes is a glorified compiler intermediate representation and the code gets more optimised towards the right hand side of the figure.}\label{fig:cfdceex}
-\end{figure}
-This {\em phase ordering problem} is well--known in literature and a practical solution is to simply perform a fixpoint iteration of the optimisation pipeline~\cite{click1995combining}.
-
-\smallskip
-
-To analyse the security of compilation passes and their interaction within a compilation pipeline, we rely on a few key notions: the definition of a trace relation, and the definition of when is a trace relation well-formed with respect to a class.
-%
-Consider traces $\src{\trace}$ and $\trg{\trace}$ as well as two trace relations $\sim_1$ and $\sim_2$. 
-% 
-The traces are related $\src{\trace}\sim_1\bullet\sim_2\trg{\trace}$ if there is another trace $\irl{\trace}$ such that $\src{\trace}\sim_1\irl{\trace}$ and $\irl{\trace}\sim_2\trg{\trace}$.
-% 
-\bul{A relation $\sim$ is well-formed w.r.t. a class of target-level properties $\trg{\class}$} iff \rul{the universal image preserves set membership}.
-\begin{definition}[Well-formedness of $\sim$ for a Class $\trg{\class}$]\label{def:wfc:sig:tracerel}
-  % $$
-
-  \begin{nscenter}
-  \noindent
-  \text{\bul{$\wfcsig{\sim}{\trg{\class}}$}} \isdef \text{\rul{$\forall \trg{\pi}\in\trg{\class}, \sigma_\sim(\trg{\pi})\in\sigma_\sim(\trg{\class})$}}
-  \end{nscenter}
-  % $$
-\end{definition}
-
-We can now state our main result: secure compilers in the robust compilation framework~\cite{abate2021extacc} compose {\em sequentially}. 
-%
-
-Let \bul{$\cc{\src{L}}{\trg{L}}$ robustly preserve the class $\sigma_{\sim_2}\left(\trg{\class[_1]}\right)$ under $\sim_1$} and let \rul{$\cc{\trg{L}}{\obj{L}}$ robustly preserve the class $\trg{\class[_2]}$ under $\sim_2$}.
-Then, \iul{when the cross-language relation $\sim_2$ is well-formed w.r.t.
-class $\trg{\class[_1]}$}, it follows that \oul{the composed compiler $\cc{\src{L}}{\trg{L}}\circ\cc{\trg{L}}{\obj{L}}$ robustly preserves the intersection of classes $\trg{\class[_1]}\cap\trg{\class[_2]}$ under $\sim_1\bullet\sim_2$}.
-% 
-%For space reasons, \Thmref{thm:urtp} and \Thmref{thm:lrtp} are not presented here, but can be found in the technical appendix ($\CoqSymbol$).
-% Note that it is not necessary to require $\sim_2$ to be well-formed w.r.t. $\src{\class[_1]}$, since the 
-% \MPin{why}
-\begin{theorem}[Composition of Secure Compilers w.r.t. $\sigma$]\label{thm:rtpsim:sig}
-  $\;$
-
-  If \bul{$\rtpsig{\cc{\src{L}}{\trg{L}}}{\tilde{\sigma}_{\sim_2}\left(\trg{\class[_{1}]}\right)}{\sim_1}$}, \rul{$\rtpsig{\cc{\trg{L}}{\obj{L}}}{\trg{\class[_2]}}{\sim_2}$}, and \iul{$\wfcsig{\sim_2}{\trg{\class[_1]}}$}, \\ then \oul{$\rtpsig{\cc{\src{L}}{\trg{L}}\circ\cc{\trg{L}}{\obj{L}}}{\trg{\class[_{1}]}\cap\trg{\class[_{2}]}}{\sim_1\bullet\sim_2}$}.
-\Coqed
-\end{theorem}
-
-Since the composition of secure compilers is again a secure compiler, the theorems generalise to a whole chain of $n$ secure compilers. 
-% 
-\Cref{thm:rtpsim:sig} can also be stated for the existential image $\tau_\sim\left(\src{\pi}\right)$, but in the interest of space that result has been moved to the appendix (\Cref{subsec:extimg}).
-% 
-Crucially, \Cref{thm:rtpsim:sig} also holds for \emph{classes of hyperproperties}, and thus, compilers that robustly preserve hyperproperties can be composed with each other as well as with compilers that robustly preserve properties.
-% 
-%Unfortunately, proofs of compilers that robustly preserve hyperproperties (including hypersafety) are scarce in the literature due to their complexity.
-%\MP{ so what ? }
-%  we don't discuss this further?
-%
-
-% It is furthermore possible to conflate the notion of classes and hyperproperties by intersecting all elements of a class of hyperproperties, to obtain one combined hyperproperty.
-% This idea has been discussed in other work~\cite{clarkson2008hyper}.
-%  MP: i think i stated the point of these 2 lines above
-
-If we take SecurePtrs and FaCT from \Cref{subsec:bg:rtp} and compose them according to \Cref{thm:rtpsim:sig}, we obtain a compiler that robustly preserves the intersection of safety properties and the \gls*{cct} hyperproperty.
-That is, for a source component that robustly satisfies any set of safety properties {\em and} \gls*{cct}, the compiled target component also robustly satisfies the same set of safety properties and \gls*{cct}.
-% The intersection simply uses the lifted version of the respective safety properties.
-%\MP{meaning?}
-
-Compiler engineers typically try to find an order of optimisations that yields well-optimised programs for either code size~\cite{cooper1999geneticphases} or performance~\cite{kulkarni2006exhaustivephase}.
-\Cref{corr:swappable} justifies that any such order of compilation passes is valid w.r.t. security, as long as the trace-relations have no effect on the respective classes.
-
-So, given two compilation passes \bul{$\cc[_{1}]{\trg{L}}{\trg{L}}$, $\cc[_{2}]{\trg{L}}{\trg{L}}$, both robustly preserving class $\trg{\class[_{1}]}$ or $\trg{\class[_{2}]}$, respectively}, \rul{their corresponding well-formed trace-relations}, \iul{and indifference of the classes with respect to these trace relations}, \oul{for any order of their composition, the composed compiler robustly preserves the intersection of $\trg{\class[_{1}]}$ and $\trg{\class[_{2}]}$}.
-% 
-% \MPin{
-% 	later on, say that optimisation passes remove elements from traces, so they should be fine w.r.t. the result with $\tau$.
-% 	also, yes, if $\tau$ adds we have a problem, if it removes we don't
-% \MKin{ Optimisations that remove elements from traces are not necessarily sound,
-%  e.g., remove all deallocations to "optimise" memory management by offloading
-%  the work to the OS.
-% }
-% }
-
-\begin{corollary}[Swapping Secure Compiler Passes]\label{corr:swappable}
-  $\;$
-
-  If \bul{$\rtpsig{\cc[_{1}]{\trg{L}}{\trg{L}}}{\trg{\class[_{1}]}}{\sim_1}$ and $\rtpsig{\cc[_{2}]{\trg{L}}{\trg{L}}}{\trg{\class[_{2}]}}{\sim_2}$}, %
-  \rul{$\wfcsig{\sim_1}{\trg{\class[_2]}}$ and $\wfcsig{\sim_2}{\trg{\class[_1]}}$}, %
-  and \iul{$\tilde{\sigma}_{\sim_2}\left(\trg{\class[_1]}\right)=\trg{\class[_1]}$ as well as $\tilde{\sigma}_{\sim_1}\left(\trg{\class[_2]}\right)=\trg{\class[_2]}$},
-  then \oul{$\rtpsig{\cc[_{1}]{\trg{L}}{\trg{L}}\circ\cc[_{2}]{\trg{L}}{\trg{L}}}{\trg{\class[_{1}]}\cap\trg{\class[_{2}]}}{\sim_1\circ\sim_2}$ and $\rtpsig{\cc[_{2}]{\trg{L}}{\trg{L}}\circ\cc[_{1}]{\trg{L}}{\trg{L}}}{\trg{\class[_{2}]}\cap\trg{\class[_{1}]}}{\sim_2\circ\sim_1}$}.
-\Coqed
-\end{corollary}
-
-Coming back to the example composing SecurePtrs with FaCT, it is likely the case that \Cref{corr:swappable} is not applicable.
-While first running SecurePtrs and then FaCT should be fine, the other direction has potential security hazards, since the SecurePtrs compiler does not account for cryptographic-constant time primitives, such as \texttt{ctselect}.
-
-\subsection{Secure Compiler Composition with Same Trace Models}
-When the cross-language trace relation is an equality, \Cref{thm:rtpsim:sig} collapses:
-Given \bul{$\cc{\src{L}}{\trg{L}}$ robustly preserves $\class[_{1}]$} and \rul{$\cc{\trg{L}}{\obj{L}}$ robustly preserves $\class[_{2}]$}, it follows that \oul{their sequential composition $\cc{\src{L}}{\trg{L}}\circ\cc{\trg{L}}{\obj{L}}$ robustly preserves the intersection of classes $\class[_{1}]$ and $\class[_{2}]$}.
-
-\begin{corollary}[Composition of Secure Compilers]\label{corr:rtp}
-  $\;$
-
-  If \bul{$\rtp{\cc{\src{L}}{\trg{L}}}{\class[_{1}]}$} and \rul{$\rtp{\cc{\trg{L}}{\obj{L}}}{\class[_{2}]}$}, then \oul{$\rtp{\cc{\src{L}}{\trg{L}}\circ\cc{\trg{L}}{\obj{L}}}{\class[_{1}]\cap\class[_{2}]}$}.
-\Coqed
-\end{corollary}
-
-\Cref{corr:rtp} provides an easy way to compose secure compilers without well-formedness of trace relations. 
-% 
-However, while \Cref{thm:rtpsim:sig} explicitly requires that $\sim_1$ is well-formed w.r.t. $\trg{\class[_2]}$, if $\sim_2$ is not well-formed w.r.t. $\trg{\class[_1]}$, care must be taken. 
-%Worst-case is that the resulting class $\sigma_{\sim_1\bullet\sim_2}\left(\trg{\class[_1]}\cap\trg{\class[_2]}\right)$ becomes empty, even though $\trg{\class[_1]}\cap\trg{\class[_2]}$ is not necessarily empty.
-This is further discussed in \Cref{subsec:compatsecpasses}.
-
-%If possible, optimisations are merged to enhance the analysis' result, since even when iterating to a fixpoint the final program may not be as optimised as it could be~\cite{click1995combining}.
-We can also obtain a specialised version of \Cref{corr:swappable}:
-%So, given two compilation passes \bul{$\cc[_{1}]{\trg{L}}{\trg{L}}$, $\cc[_{2}]{\trg{L}}{\trg{L}}$, both robustly preserving class $\class[_{1}]$ or $\class[_{2}]$, respectively}, \oul{for any order of their composition, the composed compiler robustly preserves the intersection of $\class[_{1}]$ and $\class[_{2}]$}.
-% 
-% \MPin{
-% 	later on, say that optimisation passes remove elements from traces, so they should be fine w.r.t. the result with $\tau$.
-% 	also, yes, if $\tau$ adds we have a problem, if it removes we don't
-% \MKin{ Optimisations that remove elements from traces are not necessarily sound,
-%  e.g., remove all deallocations to "optimise" memory management by offloading
-%  the work to the OS.
-% }
-% }
-
-\begin{corollary}[Swapping Secure Compiler Passes]\label{corr:swappable:one}
-  $\;$
-
-  If {$\rtp{\cc[_{1}]{\trg{L}}{\trg{L}}}{\class[_{1}]}$ and $\rtp{\cc[_{2}]{\trg{L}}{\trg{L}}}{\class[_{2}]}$}, then {$\rtp{\cc[_{1}]{\trg{L}}{\trg{L}}\circ\cc[_{2}]{\trg{L}}{\trg{L}}}{\class[_{1}]\cap\class[_{2}]}$ and $\rtp{\cc[_{2}]{\trg{L}}{\trg{L}}\circ\cc[_{1}]{\trg{L}}{\trg{L}}}{\class[_{2}]\cap\class[_{1}]}$}.
-\Coqed
-\end{corollary}
-
-\section{Security Properties Formalisation \& Composition\pages{1.5}}\label{sec:compprop}
-
-This section introduces trace properties of interest for this paper: \gls*{tms}, \gls*{sms}, \gls*{ms}, \gls*{scct}, and \gls*{ss}.
-These properties are of practical importance (as mentioned in \Cref{sec:introduction}) and also of interest in the case study presented later (\Cref{sec:casestud:defs,sec:casestud:rtp}). 
-This section presents all of them, despite the fact that they are inspired by existing work, in order to showcase all that is required for a formal proof of security for a realistic compilation toolchain.
-%\MP{ say something about this in the rebuttal: only the defs are borrowed, the universal images and all that work is new}
-%\MK{ universal images are not new. }
-% , none of the properties is novel, for self-containment of the paper, their definition is necessary for the case study.
-For each of the presented properties, the technical report defines corresponding monitors, which refine a given property and have a key role in the proofs of this paper. 
-% managed to prove 
-%  if tms-monitor --[trace]-->* ∅  
-%  then trace ∈ tms
-% but not
-%  if trace ∈ tms
-%  then tms-monitor --[trace]-->* ∅
-
-\subsection{A Trace Model for Memory Safety}\label{subsec:basic:memsafety:tracemodel}
-
-For simple memory safety composed of temporal and spatial memory safety, the trace model defines events ($\event[_{\mssafe}]$) as either the empty event ($\emptyevent$), a crash ($\lightning$), or a base-event ($\preevent[_{\mssafe}]$).
-
-\vspace{-1.0em}
-\[
-  \begin{array}{rcll}
-    \text{(Base-Events)}&\preevent[_{\mssafe}] &:=& \ev{Alloc \loc\ n} \mid \ev{Dealloc \loc} \mid \ev{Use \loc\ n} \\
-    \text{(Events)}&\event[_{\mssafe}] &:=& \preevent[_{\mssafe}] \mid \emptyevent \mid \lightning \\ 
-  \end{array}
-\]
-
-Base-events describe the actual kind of event that happened.
-For the basic memory-safety properties, these are three variants:
-First, the allocation event ($\ev{Alloc\ \loc\ n}$) that fires whenever a program claims $n$ cells of memory and stores them at address $\loc$, where addresses are assumed to be unique.
-Second, deallocation ($\ev{Dealloc\ \loc}$) announces that the object at location $\loc$ is freed.
-Third, an event to describe reads from and writes to the $n$-th memory cell from address $\loc$ ($\ev{Use\ \loc\ n}$).
-
-\subsubsection{Temporal Memory Safety}
-
-\gls*{tms}~\cite{nagarakatte2010cets} is a safety property that describes that an unallocated object must not be (re-)used.
-
-\begin{definition}[\glsfirst*{tms}]\label{def:trace:tmsdef}
-  % \footnotesize
-  $$
-  \tmssafe:=\left\{\trace[_{\mssafe}] \left| \begin{array}{rcl}
-    \ev{Alloc\ \loc\ n}&\le_{\trace[_{\mssafe}]}&\ev{Dealloc\ \loc} \\
-    \ev{Use\ \loc\ n}&\le_{\trace[_{\mssafe}]}&\ev{Dealloc\ \loc} \\
-    \text{at most one }\ev{Dealloc\ \loc}&\text{in}&\trace[_{\mssafe}] \\
-    \text{at most one }\ev{Alloc\ \loc\ n}&\text{in}&\trace[_{\mssafe}] \\
-  \end{array}\right.\right\}
-  $$
-\end{definition}
-Hereby, the notation $\event[_{1}]\le_{\trace}\event[_{2}]$ means that if $\event[_{1}]$ is in $\trace$ and if $\event[_{2}]$ is in $\trace$, then $\event[_{1}]$ appears before $\event[_{2}]$.
-
-\subsubsection{Spatial Memory Safety}
-
-\gls*{sms}~\cite{nagarakatte2009soft} is a safety property that prohibits out-of-bounds accesses.
-
-\begin{definition}[\glsfirst*{sms}]\label{def:trace:smsdef}
-% \small
-  % \begin{nscenter}
-
-  \noindent
-  \[
-  \smssafe:=\left\{\trace[_{\mssafe}] \left|\begin{array}{rcl}
-      \text{If }\ev{Alloc\ \loc\ n}\le_{\trace_{\mssafe}}\ev{Use\ \loc\ m}, \text{ then }m<n
-  \end{array}\right.\right\}
-  \]
-  % \end{nscenter}
-\end{definition}
-
-\subsubsection{Memory Safety}
-
-In spirit of earlier work~\cite{nagarakatte2009soft,nagarakatte2010cets,jim2002cyclone,necula2005ccured,michael2023mswasm}, full \gls*{ms} is the intersection of \Cref{def:trace:tmsdef,def:trace:smsdef}.
-
-\begin{definition}[\glsfirst*{ms}]\label{def:trace:msdef}
-  $
-  \mssafe:=\tmssafe \cap \smssafe
-  $
-\end{definition}
-
-Note that \Cref{def:trace:msdef} ignores data isolation, so there may still be memory-safety issues introduced by side-channels.
-
-\subsection{A Trace Model for Memory Safety with Constant Time}\label{subsec:scct:tracemodel}
-
-To express Constant Time, we extend the memory safety trace model with a {\em security tag} ($\securitytag{}$) that indicates whether events contains sensitive information ($\lock$) or not ($\unlock$).
-
-\vspace{-.5em}
-\[
-  \begin{array}{rrcl}
-    (\text{Base-Events}) & \preevent[_{\ctsafe}] &:=& \preevent[_{\mssafe}] \mid \ev{Branch\ }n \mid \ev{Binop\ }n\\
-    (\text{Security Tags}) & \securitytag{} &:=& \lock \mid \unlock\\ 
-    (\text{Events}) & \event[_{\ctsafe}] &:=& \preevent[_{\ctsafe}];\securitytag{} \mid \emptyevent \mid \lightning \\ 
-  \end{array}
-\]
-
-For cryptographic code, there is a general guideline that secrets must not be visible on a trace~\cite{ctguidelines}, i.e., secrets should not be marked as $\unlock$.
-In turn, an instruction whose timing is data-dependent must not have a secret as an operand.
-Typical operations with data-dependent timing are branches and certain binary operations, such as division.%
-\footnote{
-	This is architecture-dependent, but division is an operation that serves as a classic example for a data-dependent timing instruction~\cite[p.~755]{arm-refman}.
-}
-Both operations are represented in the trace model by extending the set of base-events with branches ($\ev{Branch\ n}$) and binary operations ($\ev{Binop\ n}$).
-
-\subsubsection{Strict Cryptographic Constant Time}
-
-\gls*{cct} is a hypersafety property~\cite{barthe2018sec} and, thus, difficult to check with monitors.
-This is because, intuitively, hypersafety properties can relate multiple execution traces with each other, but monitors work on a single execution.
-It is a common trick to sidestep this issue by means of overapproximation: this section defines the property \gls*{scct}, a stricter variant of \gls*{cct} (inspired by earlier work~\cite{almeida2017jasmin}) that enforces the policy that no secret appears on a trace.
-Programs that satisfy \gls*{scct} also satisfy \gls*{cct}, but programs that satisfy \gls*{cct} may not satisfy \gls*{scct}.
-
-\begin{definition}[\glsfirst*{scct}]\label{def:trace:scctdef}
-% \begin{nscenter}
-
-  \noindent\[
-  \scctsafe:=\left\{\trace[_{\ctsafe}] \left|\begin{array}{l}
-      \trace[_{\ctsafe}]=\hole{\cdot} \text{ or } \\\exists\trace[_{\ctsafe}'],\trace[_{\ctsafe}]=\preevent[_{\ctsafe}];\unlock\cdot\trace[_{\ctsafe}'] \wedge \trace[_{\ctsafe}']\in\scctsafe
-    \end{array}\right.\right\}
-  \]
-% \end{nscenter}
-\end{definition}
-
-\gls*{scct} may appear overly strict, since it seems that secrets must not occur on a trace (since $\securitytag{}$ is forced to be $\unlock$). 
-However, this is considered standard practice in terms of coding guidelines~\cite{ctguidelines}.
-Moreover, programs that have been compiled with FaCT~\cite{cauligi2019fact} and run with a ``data independent timing mode''~\cite{arm-refman,intel-refman} enabled do not leak secrets (see \Cref{ex:lscct}).
-
-\subsubsection{\gls*{ms}, Strict Cryptographic Constant Time}\label{sec:msscct-rel}
-
-The combination of \gls*{ms} and \gls*{scct} is the intersection of these properties, \gls*{msscct}.
-However, \gls*{ms} uses a different trace model than \gls*{scct}, so intersecting them would trivially yield the empty set. 
-To remedy this issue, we introduce $\sim_{\ctsafe}: \preevent[_{\ctsafe}] \times \preevent[_{\mssafe}] $, a cross-language trace relation (whose key cases are presented below), that we use to intuitively unify the trace model in which the two properties are expressed:
-\[
-  \typerulenolabel{mscctrel:drop:tag}{}{\preevent[_{\mssafe}]\sim_{\ctsafe}\preevent[_{\mssafe}];\unlock{}}
-  \typerulenolabel{mscctrel:drop:crash}{}{\lightning\sim_{\ctsafe}\lightning}
-\]
-\[
-  \typerulenolabel{mscctrel:drop:branch}{}{\emptyevent\sim_{\ctsafe}\ev{Branch n};\securitytag{}}
-  \typerulenolabel{mscctrel:drop:binop}{}{\emptyevent\sim_{\ctsafe}\ev{Binop n};\securitytag{}}
-\]
-
-Essentially, $\sim_{\ctsafe}$ ignores both the new $\ev{Branch\ n}$ and $\ev{Binop\ n}$ base-events as it relates security-insensitive actions ($\unlock$) to their equivalent counterparts.
-Thus, \gls*{ms} traces trivially satisfy \gls*{scct}.
-% 
-The above relation is extended point-wise to traces, skipping the empty event $\emptyevent$ on either side, and it is now possible to define \gls*{msscct} using the universal image:
-
-\begin{definition}[\gls*{ms} and \gls*{scct}]\label{def:trace:msscctdef}
-  $
-  \msscctsafe:=\mssafe\cap\sigma_{\sim_{\ctsafe}}\left(\scctsafe\right)
-  $
-\end{definition}
-
-\subsection{Extending the Trace Model with Speculation}\label{subsec:msctss:tracemodel}
-
-So far, the considered trace models do not let us express speculative execution attacks such as Spectre~\cite{kocher2019spectre}. 
-For this, we extend the earlier trace model (see \Cref{subsec:scct:tracemodel}) so that the security tags ($\securitytag{}$) carry additional information about the kind of private data leakage, i.e., the type of speculative leak.
-Moreover, we add base-events signalling the beginning of a speculative execution ($\ev{Spec}$), a barrier ($\ev{Barrier}$) that signals that any speculative execution may not go past it, as well as a rollback event ($\ev{Rlb}$), which signals that execution resumes to where speculation started.
-
-\vspace{-1em}
-{
-\[
-  \begin{array}{rrcl}
-    (\text{Base-Events}) & \preevent[_{\mathghost}] &:=& \preevent[_{\ctsafe}] \\
-    (\text{Spectre Variants}) & vX &:=& \operatorname{NONE} \mid \operatorname{PHT} \\
-    (\text{Security Tags}) & \securitytag{} &:=& \lock_{vX} \mid \unlock\\ 
-    (\text{Events}) & \event[_{\mathghost}] &:=& \preevent[_{\mathghost}];\securitytag{} \mid \emptyevent \mid \lightning \mid \ev{Spec} \mid \ev{Rlb} \mid \ev{Barrier}\\ 
-  \end{array}
-\]
-}
-
-Even though the considered Spectre variants are just SPECTRE-PHT~\cite{kocher2019spectre}, NONE just describes secret data as in \gls*{scct} (see \Cref{subsec:scct:tracemodel}), the trace model is general enough to allow for potential future extension with different variants~\cite{kocher2019spectre,maisuradze2018ret2spec,horn2019zero}.
-%For sake of readability, this paper just uses the notation $\lock$ in place of $\lock_{\text{PHT}}$.
-
-\subsubsection{Speculative Safety}
-
-\gls*{ss}~\cite{patrignani2021exorcising}, similar to \gls*{scct}, is a sound overapproximation of a variant of noninterference.
-
-\begin{definition}[\glsfirst*{ss}]\label{def:trace:ss}
-  \noindent
-
-  \begin{nscenter}
-  $
-    \sssafe:=\left\{\trace[_{\mathghost}] \left|\begin{array}{l}
-      \trace[_{\mathghost}]=\hole{\cdot} \text{ or } \exists\trace[_{\mathghost}'].\\
-      \left(\trace[_{\mathghost}]=\preevent[_{\mathghost}];\unlock\cdot\trace[_{\mathghost}'] \text{or }\trace[_{\mathghost}]=\preevent[_{\mathghost}];\lock_{\text{NONE}}\cdot\trace[_{\mathghost}']\right)\\
-      \text{and }\ \trace[_{\mathghost}']\in\sssafe
-                                 \end{array}\right.\right\}
-  $ 
-  \end{nscenter}
-\end{definition}
-The technical setup so far leads to the above definition, where only locks annotated with $\text{SPECTRE-PHT}$ are disallowed to occur on the trace.
-That way, programs attaining \gls*{ss} do not necessarily attain \gls*{scct}.
-
-\subsubsection{Speculation Memory Safety}\label{sec:spec-ms-rel}
-
-As before, we need to relate the different trace models with each other, so that the memory safety property without speculation can be lifted to speculation. 
-To this end, let $\sim_{\mathghost}: \preevent[_{\mathghost}] \times \preevent[_{\ctsafe}]$ be a cross-language trace relation whose key cases are below.
-The intuition is that \gls*{ss} is trivially satisfied in \gls*{scct}, since speculation is inexpressible there, which amounts to dropping events $\ev{Spec}$, $\ev{Rlb}$, or $\ev{Barrier}$, as well as all base events tagged with $\lock_{\text{PHT}}$.
-
-\[
-  \typerulenolabel{specrel:drop:tag}{}{\preevent[_{\ctsafe}];\lock\sim_{\mathghost}\preevent[_{\ctsafe}];\lock_{\text{NONE}}}
-  \typerulenolabel{specrel:drop:spectag}{}{\emptyevent\sim_{\mathghost}\preevent[_{\mathghost}];\lock_{\text{PHT}}}
-\]
-\[
-  \typerulenolabel{specrel:drop:pub}{}{\preevent[_{\ctsafe}];\unlock\sim_{\mathghost}\preevent[_{\ctsafe}];\unlock}
-  \typerulenolabel{specrel:drop:crash}{}{\lightning\sim_{\mathghost}\lightning}
-\]
-\[
-  \typerulenolabel{specrel:drop:spec}{}{\emptyevent\sim_{\mathghost}\ev{Spec}}
-  \typerulenolabel{specrel:drop:rlb}{}{\emptyevent\sim_{\mathghost}\ev{Rlb}}
-  \typerulenolabel{specrel:drop:barrier}{}{\emptyevent\sim_{\mathghost}\ev{Barrier}}
-\]
-
-We conclude by defining the ultimate property of interest for secure compilers: \gls*{specms}.
-\begin{definition}[\gls*{specms}]\label{def:trace:specmsdef}
-  $
-  \specmssafe := \msscctsafe\cap\sigma_{\sim_{\mathghost}}\left(\sssafe\right)
-  $
-\end{definition}
-
-\section{Case Study: Language Formalisations\pages{2}}\label{sec:casestud:defs}
-
-This section defines programming languages $\Ltms$, $\Ltrg$, $\Lms$, $\Lscct$, and $\Lspec$, all of which share common elements (presented in \Cref{subsec:cs:defs}).
-$\Ltms$ is the only statically-typed language, and it exhibits the property that all well-typed programs are \gls*{tms} (\Cref{subsec:ltms}).
-However, not all $\Ltms$ programs are \gls*{sms}.
-That is, there are well-typed $\Ltms$ programs that perform an out-of-bounds access.
-Language $\Ltrg$ is untyped and does not provide any guarantees with regards to \gls*{ms} (\Cref{subsec:lsms}).
-$\Lms$ is exactly the same language as $\Ltrg$, but this paper still distinguishes the two for sake of readability (\Cref{subsec:lms}).
-All three languages --- so $\Ltms$, $\Ltrg$, and $\Lms$ --- assume \gls*{scct} to hold, since this is -- in an ideal world -- what the programmer would expect, too: it is the job of the compiler to preserve and (potentially) enforce \gls*{scct} security~\cite{cauligi2019fact,nagarakatte2010cets,nagarakatte2009soft,akritidis2009baggy}.
-% This paper considers \gls*{cct} as an architecture-dependent property\MPin{can we justify this}, so
-
-Such consideration is also backed up by architecture providing a data (operand) independent timing mode, such as processors by Arm~\cite[p.~543]{arm-refman} and Intel~\cite[p.~80]{intel-refman}.
-%In spirit of this, all languages besides $\Lscct$ are trivially \gls*{scct}. 
-This kind of processor feature is modelled in language $\Lscct$ (\Cref{subsec:lscct}), where programs have access to a ``\gls*{cct}-mode'' and can change the leakage of emitted events according to the value of this mode (either $\obj{ON}$ or $\obj{OFF}$).
-
-Finally, modern processors also employ speculative execution to achieve speedups---and unfortunately generate Spectre attacks~\cite{kocher2019spectre}---and this is the extension of $\Lspec$ (\Cref{subsec:lspec}).
- % extends $\Lscct$ to model speculative execution.
-%
-Thus, all previous languages trivially satisfy \gls*{ss}, since they do not support speculative execution at all.
-% All languages share large parts of syntactic elements, which are presented in the next subsection (\Cref{subsec:cs:defs}).
-% Subsequent subsections extend on this accordingly.
-
-\subsection{Shared Language Definitions}\label{subsec:cs:defs}
-
-All presented programming languages share a common fragment which is partially presented here and in full detail in the technical report. 
-%\MP{put omitted stuff in appendix + ref there (and add textual explanation)}
-
-% \vspace{-.5em}
-{
-  \renewcommand{\src}[1]{\mathsf{#1}}
- \small
-\[
-  \begin{array}{rrcl}
-    \text{(Base-Events)} & \src{\preevent} &:=& \src{\ev{Alloc \loc\ n}} \mid \src{\ev{Dealloc \loc}} \mid \src{\ev{Get \loc\ n}} \mid \src{\ev{Set \loc\ n}} \\
-    \text{(Control Tags)} & \src{\sandboxtag{}} &:=& \src{\ctx} \mid \src{\comp} \\
-    \text{(Events)} & \src{\event} &:=& \src{\preevent;\sandboxtag{}} \mid \src{\emptyevent} \mid \src{\lightning} \\ 
-    \text{(Values)} & \src{v} &:=& \src{n} \\
-    \text{(Expressions)} & \src{\expr} &:=& \src{x} \mid \src{n} \mid \src{\lbinop{\expr[_1]}{\expr[_2]}} \mid \src{\lifz{\expr[_1]}{\expr[_2]}{\expr[_3]}} \\ 
-                         &&&\mid \src{\llet{x}{\expr[_1]}{\expr[_2]}} \mid \src{\llet{x}{new\ \expr[_1]}{\expr[_2]}} \\
-                         &&&\mid \src{\ldelete{x}} \mid \src{\lget{x}{\expr}} \mid \src{\lset{x}{\expr[_1]}{\expr[]}} \mid \src{stuck}\\
-                         &&&\mid \src{\lcall{f}{\expr}} \mid \src{\lreturn{\expr}} \\
-    \text{(Functions)} & \src{F} &:=& (\src{x};\src{\expr}) \\
-    \text{(Libraries)} &&&\hspace{-3.8em} \src{\library_{\ctx}},\src{\library_{\comp}} : \src{Vars} \to \src{F} 
-    % \hspace{2em}
-    %  : \src{Vars} \to \src{F} 
-     \\
-    \text{(Heaps)} & \src{H} &:=& \src{\hole{\cdot}} \mid \src{n},\src{H}
-    % \hspace{0.5cm}    
-    \\
-    \text{(Pointer Maps)} & \src{\Delta} && \text{ omitted for simplicity}
-    \\
-    \text{(Memory States)} & \src{\memstate} &:=& \left(\src{H^{\ctx}};\src{H^{\comp}};\src{\Delta}\right)\\
-    \text{(Control States)} & \src{\cfstate} & & \text{ omitted for simplicity} 
-    \\
-    \text{(States)} &\src{\Omega} &:=& \left(\src{\cfstate};\src{\sandboxtag{}};\src{\memstate}\right)\\
-  \end{array}
-\]
-}
-{
-  \renewcommand{\src}[1]{\mathsf{#1}}
-The trace models of all languages are similar to those presented earlier (\Cref{sec:compprop}).
-One technical detail is the addition of a control tag, indicating who is to blame for an emitted action: context ($\ctx$) or component ($\comp$).
-This is a standard technique in secure compilation in order to rule out irrelevant context-level events. 
-For example, a context immediately deallocating an allocated memory region twice trivially violates memory safety, but the main interest in secure compilation is to preserve component-level security, and thus component events.
-Even though this tagging could be used for blame preservation~\cite{patrignani2023blame}, this is beyond the focus of this paper.
-Another key difference is that memory accesses are now explicitly modelled as reads ($\src{\ev{Read\ \loc\ n}}$) and writes ($\src{\ev{Write\ \loc\ n}}$) instead of just uses.
-
-All languages have at least numbers as values ($\src{v}$) and second class functions ($\src{F}$).
-Functions are modelled as pairs containing the name of one argument and the body of the function.
-Bodies are just ordinary expressions, which can be simple binary operations ($\src{\lbinop{\expr[_1]}{\expr[_2]}}$), conditionals ($\src{\lifz{\expr[_1]}{\expr[_2]}{\expr[_3]}}$), function calls ($\src{\lcall{f}{e}}$) and returns ($\src{\lreturn{\expr}}$), as well as C-like memory operations.
-Programs have sets of pre-determined functions called libraries and they are marked as being part of some component ($\src{\library_{\comp}}$) or context ($\src{\library_{\ctx}}$).
-
-For the operational semantics, the runtime state ($\src{\Omega}$) is a triple consisting of a control-flow state ($\src{\cfstate}$), a control tag ($\src{\sandboxtag{}}$), and a memory state ($\src{\memstate}$).
-The latter carries information about pointers that are kept ``alive'' in pointer maps ($\src{\Delta}$), so that the semantics does not get stuck when encountering, e.g., a double-free.
-The memory state also carries two heaps to model sandboxing between a context ($\src{H^{\ctx}}$) and a component ($\src{H^{\comp}}$).
-
-}
-
-\subsection{$\src{L_{\tmssafe}}$: A Temporal but Not Spatial Memory Safe Language}\label{subsec:ltms}
-
-$\src{L_{\tmssafe}}$ is the only statically-typed language in this case study and restricts functions ($\src{F}$) to the typing signature $\src{\natt\to\natt}$. 
-The type system of $\src{L_{\tmssafe}}$ is inspired by $L^{3}$~\cite{morrisett2005L3,scherer2018fabulous} and enforces that every well-typed $\src{L_{\tmssafe}}$ program satisfies \gls*{tms}.
- % (\Cref{thm:wt:tms}).
-
-\begin{theorem}[$\src{L_{\tmssafe}}$-programs are \gls*{tms}]\label{thm:wt:tms}
-$\rsat{\src{\library_{\comp}}}{\tmssafe}$ \Coqed
-  % $\;$ 
-% 
-  % \begin{nscenter}
-  %   \hfill For any component $\src{\library_{\comp}}$, $\rsat{\src{\library_{\comp}}}{\tmssafe}$ \Coqed
-  % \end{nscenter}
-\end{theorem}
-
-\subsection{$\Ltrg$: A Memory-Unsafe Language}\label{subsec:lsms}
-
-$\Ltrg$ extends the syntax presented earlier (\Cref{subsec:cs:defs}) with dynamic typechecks ($\trg{\lhast{\expr}{\type}}$), evaluating to $\trg{0}$ if the type matches with the shape of $\trg{\expr}$ and $\trg{1}$ otherwise.
-%The pointers carry poison tags, which signify a pointer being allocated ($\trg{\poisoned}$) or freed ($\trg{\poisonless}$). 
-Furthermore, the syntax of $\Ltrg$ is extended with a way to inspect whether a pointer is freed ($\trg{\lispoisoned{x}}$), evaluating to $\trg{0}$ if it is freed and to $\trg{1}$ otherwise.
-Functions may receive arguments that are not $\trg{\natt}$, values are extended with tuples, and expressions are also extended with pair projections.
-
-% \vspace{-1em}
-\[
-  \begin{array}{rrcl}
-    \text{(Values)} & \trg{v} &:=& \dots \mid \trg{\lpair{v_1}{v_2}} \\
-    \text{(Expressions)} & \trg{\expr} &:=& \dots \mid \trg{\lhast{\expr}{\tau}} \mid \trg{\lispoisoned{x}}
-  \end{array}
-\]
-
-No changes are done to the trace model, but $\Ltrg$ has no static typing, making double-free code patterns possible.
-
-\subsection{$\Lms$: Another Memory-Unsafe Language}\label{subsec:lms}
-$\Lms$ is exactly equal to $\Ltrg$ (\Cref{subsec:lsms}), but used to emphasize that this is code after applying $\cc{\Ltrg}{\Lms}$ (\Cref{subsec:cs:ms}).
-
-\subsection{$\Lscct$: A Memory-Unsafe Language with a Constant Time Mode}\label{subsec:lscct}
-
-$\Lscct$ extends $\Lms$ (\Cref{subsec:lms}) with a ,,constant-time mode''. 
-The activation of the mode can be checked ($\obj{\lrddoit{x}{\expr}}$), changed ($\obj{\lwrdoit{D}}$), and is stored in program states ($\obj{\Omega}$).
-At the beginning of program execution of $\Lscct$ programs, the mode is turned off ($\obj{OFF}$).
-If the mode is enabled (i.e., set to $\obj{ON}$), the intuition is that no secrets are leaked. 
-This models the real-world\footnote{As present in Intel~\cite[p.80]{intel-refman} and ARM~\cite[p.~543]{arm-refman} processors.} data-independent timing mode as well as the result of compiling a program with FaCT~\cite{cauligi2019fact}.
-For example, FaCT rewrite code that branches to use a constant-time selection primitive.
-To not obfuscate our formalisation unnecessarily by duplicating syntax, we simply added the ,,constant-time mode'' to $\Lscct$.
-
-% $\Lscct$ extends $\Lms$ (\Cref{subsec:lms}) with a way to read from ($\obj{\lrddoit{x}{\expr}}$) and to write to ($\obj{\lwrdoit{D}}$) a model-specific register that controls a constant-time mode. 
-% This language is a combination of the effect of the FaCT compiler~\cite{cauligi2019fact} as well as a data-independent timing mode, a feature that is present in both Arm~\cite[p.~543]{arm-refman} and Intel~\cite[p.~80]{intel-refman} processors.
-% \MP{ can we be more specific about the fact pass? }
-% To this end, states ($\obj{\Omega}$) are extended with the value of the register ($\obj{D}$), which is initially set to be not active ($\obj{OFF}$).
-% If the register is marked active ($\obj{ON}$), the intuition is that no secrets can appear on specification traces.
-% If the register is marked inactive, secrets may appear on traces.
-% The state of the register can be read and written to, e.g., \Cref{tr:e-wrdoit-off}.
-% For the other languages seen earlier, the mode is intuitively always-on, i.e., the mode of execution always uses constant-time mode.
-
-% \vspace{-.5em}
-\[
-  \begin{array}{rrcl}
-    \text{(Mode Values)} & \obj{D} &:=& \obj{ON} \mid \obj{OFF} \\
-    \text{(Security Tags)} & \obj{\securitytag{}} &:=& \obj{\lock} \mid \obj{\unlock} \\
-    \text{(Base-Events)} & \obj{\preevent} &:=& \dots \mid \obj{iGet\ \loc\ v} \mid \obj{iSet\ \loc\ v} \\
-                       &&& \mid\ \obj{Binop\ n} \mid \obj{Branch\ n} \\
-    \text{(Events)} & \obj{\event} &:=& \obj{\preevent;\sandboxtag{};\securitytag{}} \mid \obj{\emptyevent} \mid \obj{\lightning} \\
-    \text{(Expressions)} & \obj{\expr} &:=& \dots \mid \obj{\lrddoit{x}{\expr}} \mid \obj{\lwrdoit{D}} \\
-                         &&&\mid\ \obj{\llet{x^{\securitytag{}}}{\expr[_1]}{\expr[_2]}}\\
-    \text{(States)} & \obj{\Omega} & := & \left(\obj{\cfstate};\obj{\sandboxtag{}};\obj{D};\obj{\memstate}\right)\\
-  \end{array}
-\]
-\begin{center}
-\newcommand{\expreval}[5]{{#1}\triangleright\xspace {#2}\xrightarrow{#5}\ {#3}\triangleright\xspace {#4}\xspace}
-\newcommand{\exprevalo}[5]{\expreval{\obj{#1}}{\obj{#2}}{\obj{#3}}{\obj{#4}}{\obj{#5}}}
-  \typerule{$e-\obj{wrdoit}-\text{off}$}{
-  }{
-    \pstepto{\rtt{\obj{\cfstate;\sandboxtag{};D;\memstate}}{\obj{\lwrdoit{OFF}}}}{\rtt{\obj{\cfstate;\sandboxtag{};OFF;\memstate}}{\obj{0^{\securitytag{}}}}}{\obj{\emptyevent}}
-  }{e-wrdoit-off}
-
-  \typerule{$e-\obj{get}-\in-$noleak}{
-    \obj{\memstate}=\obj{H^{\ctx};H^{\comp};\Delta_1,x\mapsto(\loc;\sandboxtag{};\poison),\Delta_2} &
-    \obj{\loc}+\obj{n}\in\text{dom }\obj{H^{\sandboxtag{}}} \\
-    \obj{\securitytag{''}}=\obj{\securitytag{}\sqcap\securitytag{'}} &
-    \obj{\event} = \obj{{iGet}\ \loc\ n;\sandboxtag{};\securitytag{''}}
-  }{
-    \exprevalo{
-    	\cfstate;\sandboxtag{'};ON;\memstate
-    }{
-      	x^{\securitytag[n^{\securitytag{'}}] }
-    }{
-    	\cfstate;\sandboxtag{'};ON;\memstate
-    }{
-    	(H^{\sandboxtag{}}(\loc+n))^{\securitytag{''}}
-    }{
-      \event
-    }
-  }{to-e-get-in-noleak}
-\end{center}
-
-The language also adds user annotations $\obj{\securitytag{}}$ for the secrecy of variables, which can be either private (high secrecy) $\obj{\lock}$ or public (low secrecy) $\obj{\unlock}$.
-Security tags $\obj{\securitytag{}}$ are arranged in the usual secrecy lattice~\cite{zdancewicphd}, where $\obj{\unlock}\sqsubseteq\obj{\lock}$.
-% $\obj{\lock}\sqsubseteq\obj{\securitytag{}}$ and $\obj{\securitytag{}}\sqsubseteq\obj{\unlock}$.
-
-Memory accesses to secret data need to be present to reason about memory safety, even when execution is in constant-time mode, e.g, \Cref{tr:to-e-get-in-noleak}.
-$\Lscct$ extends base-events with $\obj{\ev{iGet}\ \loc\ \valueexpr}$ and $\obj{\ev{iSet}\ \loc\ \valueexpr}$ (for data $\obj{i}$ndependent get and set) to prevent secrets from leaking but still enable reasoning about memory safety.
-Due to the technical setup, the rule needs to check if the access is in bounds ($\obj{\loc}+\obj{n}\in\obj{H^t}$) and update the secrecy tag ($\obj{\securitytag{''}}=\obj{\securitytag{}\sqcap\securitytag{'}}$) with the least upper bound of the tags, according to the aforementioned lattice.
-The precise information carried by the event, e.g., location ($\obj{\loc}$) is taken from the pointer map, which carries information irrelevant to this rule ($\obj{\poison}$).
-
-Base-events include $\obj{\ev{Branch}\ n}$ and $\obj{\ev{Binop}\ n}$ that are emitted when evaluating a branch or certain binary expressions, such as division, respectively, whenever the constant-time mode is inactive.
-Events are extended with a security-tag ($\obj{\securitytag{}}$) to signal the secrecy of the involved data.
-
-The evaluation steps are amended to propagate the security-tag annotations $\obj{\securitytag{}}$.
-When the constant-time mode is inactive, base-events $\obj{\ev{Branch}\ n}$ and $\obj{\ev{Binop}\ n}$ are emitted for conditionals and binary operations, respectively.
-Otherwise, just like in the semantics of the earlier languages, $\obj{\emptyevent}$ is emitted for binary and branching operations.
-%To mark this mode as active or inactive, \Cref{tr:e-wrdoit} allows to access the register in $\obj{\statevar}$.
-
-\Cref{ex:lscct} illustrates the differences between $\Lscct$ and other languages.
-
-\begin{exampleenv}[$\Lscct$ with and without constant-time mode]\label{ex:lscct}
-  Consider again \Cref{ex:strncpy}, with a context copying the string \texttt{Hello World}, where everything is marked with a high security tag: $\lock$.
-  The top half of \Cref{fig:ex-cct} (titled $\Lscct$), describes the execution trace of the program, while the bottom side of the table (titled ``Specification''), describes the related specification trace (\Cref{subsec:msctss:tracemodel}).
-  Read in parallel from top to bottom, the figure shows parts of the execution trace. 
-  In each half, the left column (\textit{Active}) has constant-time mode $\obj{ON}$ and the right one (\textit{Inactive}) has it $\obj{OFF}$.
-
-  When the constant-time mode is off, the execution yields events in similar fashion to before (\Cref{subsec:ltms,subsec:lsms,subsec:lms}).
-  But, if it is turned on, then the branching event \colorbox{yellow}{does not fire} anymore and both reading and writing to memory is related to a specification trace with \colorbox{lightgray}{no exposed secrets}.
-\end{exampleenv}
-\begin{figure}[!htb]
-	\vspace{-1em}\small
-  $$
-  % {
-  \begin{array}{ccc}
-    & \Lscct &        \\
-    \text{Active} & \mid & \text{Inactive} \\\hline
-    \obj{\ev{Alloc}\ \loc_{x}\ 12;\comp;\unlock} & \mid & \obj{\ev{Alloc}\ \loc_{x}\ 12;\comp;\unlock} \\
-    \obj{\ev{Alloc}\ \loc_{y}\ 12;\comp;\unlock} & \mid & \obj{\ev{Alloc}\ \loc_{y}\ 12;\comp;\unlock} \\
-    \obj{{\ev{iGet}}\ \loc_{x}\ 0;\comp;}\text{\colorbox{lightgray}{$\obj{\lock}$}} & \mid & \obj{\ev{Get}\ \loc_{x}\ 0;\comp;\lock} \\
-    \text{\colorbox{yellow}{$\obj{\emptyevent}$}} & \mid & \obj{\ev{Branch}\ 0;\comp;\lock} \\
-    \obj{{\ev{iGet}}\ \loc_{x}\ 0;\comp;}\text{\colorbox{lightgray}{$\obj{\lock}$}} & \mid & \obj{\ev{Get}\ \loc_{x}\ 0;\comp;\lock} \\
-    \obj{{\ev{iSet}}\ \loc_{y}\ 0\ \mathtt{'H'};\comp;}\text{\colorbox{lightgray}{$\obj{\lock}$}} & \mid & \obj{\ev{Set}\ \loc_{y}\ 0\ \mathtt{'H'};\comp;\lock} \\
-    \obj{{\ev{iGet}}\ \loc_{x}\ 1;\comp;}\text{\colorbox{lightgray}{$\obj{\lock}$}} & \mid & \obj{\ev{Get}\ \loc_{x}\ 1;\comp;\lock} \\
-    \text{\colorbox{yellow}{$\obj{\emptyevent}$}} & \mid & \obj{\ev{Branch}\ 0;\comp;\lock} \\
-    \vdots & \mid & \vdots \\
-    \obj{{\ev{iGet}}\ \loc_{x}\ 12;\comp;}\text{\colorbox{lightgray}{$\obj{\lock}$}} & \mid & \obj{\ev{Get}\ \loc_{x}\ 12;\comp;\lock} \\
-    \text{\colorbox{yellow}{$\obj{\emptyevent}$}}& \mid & \obj{\ev{Branch}\ 1;\comp;\lock} \\
-    \obj{\ev{Dealloc}\ \loc_{y};\comp;\unlock} & \mid & \obj{\ev{Dealloc}\ \loc_{y};\comp;\lock} \\
-    \obj{{\ev{iGet}}\ \loc_{y}\ 6;\comp;}\text{\colorbox{lightgray}{$\obj{\lock}$}} & \mid & \obj{\ev{Get}\ \loc_{y}\ 6;\comp;\lock} \\
-  \end{array}
-  % }
-  $$
-  $$
-  % {
-  \begin{array}{ccc}
-          & \text{Specification} & \\
-  \text{Active} & \mid & \text{Inactive} \\\hline
-     \ev{Alloc}\ \loc_{x}\ 12;\unlock & \mid & \ev{Alloc}\ \loc_{x}\ 12;\unlock\\
-     \ev{Alloc}\ \loc_{y}\ 12;\unlock & \mid & \ev{Alloc}\ \loc_{x}\ 12;\unlock\\
-     \ev{Use}\ \loc_{x}\ 0;\text{\colorbox{lightgray}{$\unlock$}} & \mid & \ev{Use}\ \loc_{x}\ 0;\lock\\
-     \emptyevent & \mid & \ev{Branch}\ 0;\lock\\
-     \ev{Use}\ \loc_{x}\ 0;\text{\colorbox{lightgray}{$\unlock$}} & \mid & \ev{Use}\ \loc_{x}\ 0;\lock\\
-     \ev{Use}\ \loc_{y}\ 0;\text{\colorbox{lightgray}{$\unlock$}} & \mid & \ev{Use}\ \loc_{y}\ 0;\lock\\
-     \ev{Use}\ \loc_{x}\ 1;\text{\colorbox{lightgray}{$\unlock$}} & \mid & \ev{Use}\ \loc_{x}\ 1;\lock\\
-     \emptyevent & \mid & \ev{Branch}\ 0;\lock\\
-     \vdots & \mid & \vdots \\
-      \ev{Use}\ \loc_{x}\ 12;\text{\colorbox{lightgray}{$\unlock$}} & \mid & \ev{Use}\ \loc_{x}\ 12;\lock\\
-      \emptyevent & \mid & \ev{Branch}\ 1;\lock\\
-      \ev{Dealloc}\ \loc_{y};\unlock & \mid & \ev{Dealloc}\ \loc_{y};\unlock\\
-      \ev{Use}\ \loc_{y}\ 6;\text{\colorbox{lightgray}{$\unlock$}} & \mid & \ev{Use}\ \loc_{y}\ 6;\lock\\
-  \end{array}
-  % }
-  $$
-  \caption{Traces for \Cref{ex:lscct}.}
-  \label{fig:ex-cct}
-\end{figure}
-
-\subsection{$\Lspec$: A Memory-Unsafe Language with Speculation}\label{subsec:lspec}
-
-$\Lspec$ extends $\Lscct$ (\Cref{subsec:lscct}) with speculative dynamic semantics that is inspired by existing work~\cite{guarnieri2018spectector,fabian2022automatic}.
-After branching, speculation starts by pushing the current configuration into the speculation state ($\ird{S}$), and run subsequent code in a predetermined window (whose size is $\omega$) until a rollback of operational state is performed.
-The window is set inside the speculation state.
-Within such windows, data may leak, this is marked as high $\ird{\lock}$ and with an annotation that indicates the respective speculative execution variant. 
-In this paper, we just consider SPECTRE-PHT~\cite{kocher2019spectre}, whose starting point is \Cref{tr:e-ifz-true-spec}.
-% where $\omega$ is the size of the speculation window, i.e., a number of reduction steps to be taken under speculative semantics.
-Additionally, $\ird{\lock}$ may also carry a $\ird{NONE}$ annotation, to signal that this is a leak irrespective of speculation, i.e., as defined earlier (\Cref{subsec:lscct}).
-The semantics is kept general enough to allow for future extension to support different variants~\cite{kocher2019spectre,maisuradze2018ret2spec,horn2019zero}.
-
-In terms of language features, $\Lspec$ includes a new barrier operation $\ird{\lbarrier}$ that blocks speculative execution. 
-To facilitate speculative execution in a non-assembly-like language, a stack of operational state is used and speculation is active if that stack has a size greater than one.
-This is exploited in \Cref{tr:e-spec-eat} to consume any leftover speculation.
-
-\vspace{-.5em}
-\[
-  \begin{array}{rrcl}
-    \text{(Base-Events)} & \ird{\preevent} &:=& \dots \mid \ird{Spec} \mid \ird{Rlb} \mid \ird{Barrier}\\
-    \text{(Expressions)} & \ird{\expr} &:=& \dots \mid \ird{\lbarrier} \\
-    \text{(Speculation State)} & \ird{S} &:=& \ird{\Omega\triangleright\expr} \mid \ird{S},\left(\ird{\Omega};\ird{n};\ird{\expr}\right)
-  \end{array}
-\]
-
-\begin{center}
-\newcommand{\expreval}[5]{{#1}\triangleright\xspace {#2}\xrightarrow{#5}\ {#3}\triangleright\xspace {#4}\xspace}
-\newcommand{\exprevald}[5]{\expreval{\ird{#1}}{\ird{#2}}{\ird{#3}}{\ird{#4}}{\ird{#5}}}
-
-  %
-  \typerule{$e-\ird{ifz}-\ird{true}-\text{spec}$}{
-    \ird{S} = \rtt{(\ird{\cfstate['];\sandboxtag{};D;\memstate});\ird{S}}{\ird{\expr[_1]}},({\ird{\cfstate;\sandboxtag{};D;\memstate}};\omega;{\ird{\expr[_2]}})
-  }{
-    \pstepto{\rtt{\ird{\cfstate;\sandboxtag{};D;\memstate}}{\ird{\lifz{0^{\securitytag{}}}{\expr[_1]}{\expr[_2]}}}}{\ird{S}}{\ird{{Branch}\ 0;\sandboxtag{};\securitytag{}}\cdot\ird{{Spec};\sandboxtag{};\securitytag{}}}
-  }{e-ifz-true-spec}
-  %
-  \typerule{$e-\ird{spec}-\text{eat}$}{
-    \ird{n} > 0 &
-    \pstepto{\rtt{\ird{\Omega}}{\ird{\expr}}}{\rtt{\ird{\Omega'}}{\ird{\expr[']}}}{\ird{\event}}
-  }{
-    \ghoststepto{\ird{S},(\ird{\Omega};\ird{n};\ird{\expr})}{\ird{S},(\ird{\Omega'};\ird{n}-1;\ird{\expr'})}{\ird{\event}}
-  }{e-spec-eat}
-  % 
-  \typerule{$e-\ird{spec}-\text{eaten}$}{
-  }{
-    \ghoststepto{\ird{S},(\ird{\Omega};\ird{0};\ird{\expr})}{\ird{S}}{\ird{Rlb;\Omega.\sandboxtag{};\unlock}}
-  }{e-spec-eaten}
-  %
-  \typerule{$e-\ird{spec}-\text{barrier}$}{
-  }{
-    \ghoststepto{\ird{S},(\ird{\Omega};\ird{n};\ird{\lbarrier})}{\ird{S},(\ird{\Omega};\ird{0};\ird{0})}{\ird{Barrier;\Omega.\sandboxtag{};\unlock}}
-  }{e-spec-barrier}
-\end{center}
-
-The trace model is extended with $\ird{{Spec}}$, $\ird{{Rlb}}$, and $\ird{{Barrier}}$ events that signal the respective operational action. 
-A $\ird{\lbarrier}$ prevents any execution whatsoever besides $\ird{{Rlb}}$ when run in speculative execution mode and does nothing when run in normal mode, e.g., \Cref{tr:e-spec-barrier}.%
-\footnote{In the rule, the notation $\cfstate' = \cfstate[.\text{win} = 0]$ means that $\cfstate'$ is a copy of $\cfstate$ up to field $\text{win}$, which is set to $0$.}
-Lastly, $\ird{Rlb}$ is emitted when the speculation window is zero and the operational state is rolled back (see \Cref{tr:e-spec-eaten}).
-
 \section{Case Study: Composing Secure Compiler Passes and Optimisations \pages{7}}\label{sec:casestud:rtp}
@@
 \end{lemma}
-
-\section{Formal Insights}\label{sec:formalities}
-
-This section first discusses how to connect each language-specific security property to the general security properties presented in \Cref{sec:compprop} (\Cref{subsec:formalities:maps}), then it demonstrates that the security property resulting from the universal image projection is faithful to the original property (\Cref{subsec:formalities:props}).
-Finally, this section discusses why the order of compiler passes matters, and how does our framework help with identifying insecure compositions (\Cref{subsec:compatsecpasses}), and it gives additional technical insights on the secure compilation proofs (\Cref{subsec:seccompproofs}).
-
-\subsection{From Language Traces to General Ones}\label{subsec:formalities:maps}
-
-The previous theorems talk about preserving properties expressed in the trace model of the languages of each compiler.
-However, these trace models are not the same trace model we used to specify the properties of \Cref{sec:compprop} (indicated with \ev{L}), which serves as the ``ground truth'' for the meaning of our properties.
-% 
-To bridge this gap, the formal development requires specifying additional trace relations, from each of the language trace models to the \ev{L} one, that, for example, relate $\src{Get\ \loc\ n}$ and $\src{Set\ \loc\ n}$ to $\ev{Use\ \loc\ n}$ (and that induce a related universal image that we use in \Cref{subsec:formalities:props}).
-% 
-One key insight of these relations is that they all omit context-made actions for two reasons: (1) contexts (which are universally quantified) can trivially invalidate any property and (2) we are interested in the component upholding the properties.
-%
-
- % disregarded a glaring but minor technical issue: take $\tmssafe$ (\Cref{def:trace:tmsdef}) as an example.
-% It contains traces as specified in \Cref{subsec:basic:memsafety:tracemodel}.
-% However, these are different from $\Ltms$ traces (\Cref{subsec:ltms}) which means that \Cref{thm:wt:tms} is ``ill-typed''.
-% Similarly for the other theorems.
-% In the interest of reducing visual baggage, the notation left out yet another cross-language trace relation, translating language-specific traces into these non-language specific traces. 
-% In doing so, context-level actions are dropped and, e.g., both $\src{Get\ \loc\ n}$ and $\src{Set\ \loc\ n}$ are translated into $\ev{Use\ \loc\ n}$.
-
-\subsection{Security Properties and Their Meaning}\label{subsec:formalities:props}
-Each of the presented compilers use a cross-language trace relation, which is also used to translate the property from one language to the other one (via the existential or universal images).
-% 
-%Whenever such translations occur, it is important to check whether there has been a change of meaning.
-%Secure compiler engineers need to ensure that such translations of properties do not change meaning in a fundamental wrong way. 
-% lifting or lowering reasoning from target- to source-level or vice-versa,
-While the meaning of projected properties does change with a translation, the change should not allow for a flawed compilation pipeline.
-% 
-For example, we could be using a trace relation that translates a property in a language to a totally different one in another language.
-% 
-To raise the trust into the translation of properties, \Cref{thm:prop-rel-corr} states (in a general fashion) that each security property is faithfully translated using the universal image according to the cross-language trace relation induced by the compiler.
-
-\begin{theorem}[Properties Relation Correctness]\label{thm:prop-rel-corr}
-  \begin{align*}
-    \forall& \pi \in \{\tmssafe,\smssafe,\scctsafe,\sssafe\}, 
-    \\
-    \text{ for each }& \text{pair of languages } \src{L} \text{ and } \trg{L} \text{ used by the compilers},
-    \\
-    \text{ if }& 
-    % \tau_{\sim^{\trg{L}}_{\ev{L}}}(\pi) \sim^{\trg{L}}_{\ev{L}} \pi
-    \sigma_{\sim^{\trg{L}}_{\ev{L}}}(\pi) \sim^{\trg{L}}_{\ev{L}} \pi
-    \text{ and } 
-    % \sigma_{\sim^{\src{L}}_{\trg{L}}}(\tau_{\sim^{\trg{L}}_{\ev{L}}}(\pi)) \sim^{\src{L}}_{\trg{L}} \tau_{\sim^{\trg{L}}_{\ev{L}}}(\pi)
-    \sigma_{\sim^{\src{L}}_{\trg{L}}}(\sigma_{\sim^{\trg{L}}_{\ev{L}}}(\pi)) \sim^{\src{L}}_{\trg{L}} \sigma_{\sim^{\trg{L}}_{\ev{L}}}(\pi)
-    \\
-    \text{ then }& 
-    % \tau_{\sim^{\src{L}}_{\ev{L}}}(\pi) = \sigma_{\sim^{\src{L}}_{\trg{L}}}(\tau_{\sim^{\trg{L}}_{\ev{L}}}(\pi)) 
-    \sigma_{\sim^{\src{L}}_{\trg{L}}}(\sigma_{\sim^{\trg{L}}_{\ev{L}}}(\pi)) \sim^{\src{L}}_{\ev{L}} \pi
-  \end{align*}
-\end{theorem}
-The complexity of this theorem is that the relation in the conclusion cannot be obtained by composing the two relations in the premises.
-% 
-%% MK: the following does not say anything about the theorem, only about the syntactic compiler
-%What happens in practice is that compiler engineers are given the relations from each language to \ev{L} and they need to build a compiler that induces a relation $\sim^{\src{L}}_{\trg{L}}$ that respects this theorem.
-%
-
-We now informally argue why this theorem holds for the composition of all considered properties from \Cref{sec:compprop}.
-
-\paragraph{$\sigma_{\sim^{\Ltms}_{\Lms}}\left(\tmssafe\cap\smssafe\cap\scctsafe\cap\sssafe\right)$}
-
-For the four properties considered here, the trace models of $\Ltms$, $\Ltrg$, and $\Lms$ do not consider actions related to $\scctsafe$ and $\sssafe$, so these two properties are trivially translated correctly.
-% 
-We now discuss the remaining $\tmssafe$ and $\smssafe$ in the form of their intersection $\mssafe$.
-
-Recall that $\sim^{\Ltms}_{\Lms}$ is the composition of $\sim^{\Ltms}_{\Ltrg}$ and $\sim^{\Ltms}_{\Lms}$.
-% 
-Since $\sim^{\Ltms}_{\Ltrg}$ is an equality, this relation trivially preserves the meaning of translated properties: related traces are identical!
-
-Finally, let us consider a trace $\irl{\trace}\in\mssafe$ and understand what is that is related to via $\sim^{\Ltrg}_{\Lms}$.
-% 
-{\em All} traces $\trg{\trace}$ with $\trg{\trace}\sim^{\Ltrg}_{\Lms}\irl{\trace}$ are identical to $\irl{\trace}$ except for get and set actions, which require for in-bound accesses (as stated in \Cref{subsec:cs:ms}): this clearly respect $\mssafe$.
-%
-
-% In both cases, the translated event does not represent an out-of-bounds memory access.
-
-% \footnote{Both $\sim^{\Ltms}_{\Ltrg}$ and $\sim^{\Ltrg}_{\Lms}$ filter context actions.}.
-
-% Now, starting from $\mssafe$ expressed in $\Lms$, we check that its translation in $\Ltms$ is still $\mssafe$.
-% % 
-% We can see that the property resulting from the projection of $\Lms$ traces back to $\Ltms$ faithfully models $\mssafe$, which translates to both $\scctsafe$ and $\sssafe$, 
-% \MPin{ equiv  +  other rel only adds traces that fail }
-% since these properties trivially hold for $\Ltms$, $\Ltrg$, and $\Lms$ traces.
-
-\paragraph{$\sigma_{\sim^{\Lms}_{\Lscct}}\left(\tmssafe\cap\smssafe\cap\scctsafe\cap\sssafe\right)$}
-
-As before, the trace models of $\Lscct$ and $\Lms$ cannot express $\sssafe$, so that is trivially translated correctly.
-% 
-Also, the trace model of $\Lscct$ extends the one of $\Lms$ with respect to $\tmssafe$ and $\smssafe$ events, so the translation argument regarding those two properties is the same as before.
-% 
-Thus, we need to reason about whether $\scctsafe$ is translated correctly.
-
-By definition, $\sim^{\Lms}_{\Lscct}$ only relates $\obj{\unlock}$ events, which is also the same relation induced by $\sim_{\ctsafe}$ in \Cref{sec:msscct-rel}.
-% 
-This ensures that composing the relations only relates $\unlock$ events, and thus the property is translated correctly.
-
-% Consider a trace $\obj{\trace}\in\scctsafe$ and all $\irl{\trace}$ such that $\irl{\trace}\sim^{\Lms}_{\Lscct}\obj{\trace}$. 
-% In turn, $\obj{\trace}$ cannot contain any leaks of secret values, so it cannot violate $\scctsafe$.
-% Toggling data-independent timing mode has no impact on $\mssafe$ or $\sssafe$.
-
-\paragraph{$\sigma_{\sim^{\Lscct}_{\Lspec}}\left(\tmssafe\cap\smssafe\cap\scctsafe\cap\sssafe\right)$}
-The trace model of $\Lspec$ extends the one of $\Lscct$ with respect to $\tmssafe$, $\smssafe$, and $\scctsafe$, so the translation argument for those three properties is the same as before.
-% 
-Concerning $\sssafe$, $\sim^{\Lscct}_{\Lspec}$ relates $\Lspec$ speculation traces with $\Lscct$ branches with the $\lock_{\text{NONE}}$ tag, so the property is translated correctly, according to the relation defined in \Cref{sec:spec-ms-rel}.
- % introducing $\ird{barrier}$ never violates $\sssafe$ and has no effect on $\mssafe$ or $\scctsafe$.
-
-%\MPin{
-%	can this be made formal?
-%	what's the point?
-%}
-
-\subsection{Compatibility of Secure Compiler Passes}\label{subsec:compatsecpasses}
-
-% \Cref{thm:rtpsim:sig} is applicable, up to the well-formedness of $\sim_1$ w.r.t. $\trg{\class[_2]}$.
-Consider applying $\ccspec$ first and then $\ccb$ (albeit currently syntactically impossible).
-% , let us pretend that it is possible by just extending $\ccb$ for the constructs present in $\Lscct$ that are not present in $\Ltrg$.
-% Clearly, as evidenced earlier (\Cref{subsec:formalities:props}), $\ccb\cdot\ccspec$ is an acceptable composition that yields an interesting class of security properties, i.e., $\sigma_{\sim_{\irl{ms}}\bullet\sim_{\ird{\mathghost}}}\left(\mssafe\cap\scctsafe\cap\sssafe\right)$.
-In this case, $\ccb$ would insert new branches into the code that are not protected by a speculation barrier! 
-This concern is reflected in the proofs that establish that the security class resulting of the composition of the trace relations is meaningful.
-% , much similar to ensuring that a definition models what the formal methods expert intends to model. 
-For the $\ccspec\cdot\ccb$ case, the class is $\sigma_{\sim_{\ird{\mathghost}}\bullet\sim_{\irl{ms}}}\left(\mssafe\cap\scctsafe\cap\sssafe\right)$.
-Since $\ev{Use\ \loc\ n}\sim_{\irl{ms}}\irl{Branch\ n}\cdot\irl{Spec}\cdots$, where $\cdots$ does not contain a $\ev{{Barrier}}$ event, the resulting class is \emph{not} the original $\sssafe$ that is intended and it would break the corresponding \Cref{thm:prop-rel-corr}.
-% Because of this, it is crucial to analyse the precise shape of the class of properties after mapping a target-level property to a source-level property\footnote{Dito for mapping a source-level property to a target-level property.}.
-
-However, the composition is still technically possible and it is the job of the compiler engineers to ensure that the secure compilation pipeline happens in an order that ensures that the mapped security property is the intended one.
-
-\subsection{Secure Compilation Proofs}\label{subsec:seccompproofs}
-
-Our secure compilation proofs rely on backtranslations~\cite{abate2019jour,patrignani2021rsc}, which let one construct a source context starting from either target traces (aka trace-based backtranslations) or target contexts (aka context-based backtranslations).
-These backtranslations also require setting up cross-language relations between various language elements such as expressions and program states, so we leave these details for the technical reports.
-% , the usual strategy is to establish a backwards simulation between source components and their compiled target counterpart. 
-% For the universally quantified target context, a technique known as backtranslation is well-established, which constructs a source-level context either by backtranslating the target-level context, amounting to a ``decompilation'', or by constructing it using the target-level trace, so that the source-context emits parts of the whole execution trace that are related to the parts emitted by the target-level context.
-All backtranslations in the case-study are trace based except for those required by $\ccdce$ and $\cccf$, which are context-based (and they are an identity function). 
-% Since these optimisations have no significant semantic effect and source- and target-language are equal, the backtranslation amounts to an identity translation.
-
-\section{Related Work\pages{2}}\label{sec:relwork}
-
-This section discusses robust compilation, other secure compilation criteria and work related to the properties preserved in the case study.
-% (\Cref{subsec:relw:seccomprtp}) 
-% (\Cref{subsec:relw:seccompcrit}).
-% Since the case study of \Cref{sec:casestud:defs,sec:casestud:rtp} implements measures for preserving \gls*{ms}, \gls*{cct}, and \gls*{ss}, 
-% Then, this section discusses work related 
-% relevant related work as well 
-% (\Cref{subsec:relw:msmechs,subsec:relw:cctmechs,subsec:relw:ssmechs}).
-
-\paragraph*{Secure Compilation as Robust Preservation}\label{subsec:relw:seccomprtp}
-
-The robust preservation of properties as a compiler-level criterion has been analysed extensively~\cite{abate2019jour,patrignani2021rsc,abate2021extacc,patrignani2019survey} and thus we build on that framework.
-No existing work is concerned with composing secure compilers, however, existing work~\cite{abate2021extacc} sketches composition of trace-relating compiler correctness in a similar way to what has been presented here.
-%Parts of these works consider languages with different trace models and our technical setup can handle this. 
-The work relating robust preservation with universal composability~\cite{patrignani2022universal} is closest to what this paper presents.
-The authors demonstrate a similar compositionality theorem to what is presented here (\Cref{sec:sequential}) but use it in the context of protocols.
-The work does not consider the generality to support different trace models or composition of compilers which robustly preserve different classes.
-% The authors demonstrate a similar compositionality theorem to what is presented here (\Cref{sec:sequential}) as well as in an earlier version of this work~\cite{kruse2022csc}.
-% However, they do not demonstrate the scalability of the approach by means of a case study.
-
-%\paragraph*{Composition of Secure Compilers}\label{subsec:relw:seccompcomp}
-
-There is work on lifting exploits for single compilers to the whole chain~\cite{paykin2019weird}.
-%Our framework does not consider exploits.
-%\MP{ a bit dry and short}
-While that work considers {\em in}secure compilation and composition thereof in terms of exploits, the composition they are interested in allows to lift an exploit for one compiler pass to the whole compilation chain.
-Our framework takes the opposite direction and provides compositionality results for secure compilers.
-%
-%
-%To the best of our knowledge, there does not exist published work discussing the compositionality of the robust preservation criterion, the gold-standard of secure compilation.
-
-\paragraph*{Other Secure Compilation Criteria}\label{subsec:relw:seccompcrit}
-
-While this paper focuses on the robust preservation framework~\cite{abate2019jour}, other secure compilation criteria exist.
-The survey on formal approaches to secure compilation~\cite{patrignani2019survey} discusses a broad spectrum already, while this section presents a very high-level overview.
-Fully abstract compilation~\cite{abadi1999fullabstraction} states that a compiler should preserve and reflect observational equivalence between source and target programs.
-Abate \emph{et al.}~\cite{abate2021faandrc} showed that fully abstract compilers robustly preserve program properties that are either trivial or meaningless.
-As a mitigation for this, the authors presented a categorical approach based on maps of distributive laws~\cite{watanabe2002modl}, which they call many maps of distributive laws.
-Maps of distributive laws have been investigated before as a possible secure compilation criterion~\cite{tsampas2020catsc}.
-Other approaches are extensions of the compiler correctness criterion as discussed in other work~\cite{patterson2019next700} or the introduction of opaque observations~\cite{vu2021reconciling} to reconcile compiler optimisations with security.
-Note that this work also presents secure compilers that are optimising, but contrary to the other~\cite{vu2021reconciling}, provides a formal account of these in the robust preservation framework.
-% Lastly, the authors of this paper have presented ongoing work~\cite{patrignani2023blame} on a weaker robust preservation criteria based on the concept blame.
-
-\paragraph*{Memory Safety Mechanisms}\label{subsec:relw:msmechs}
-
-Different mechanisms for enforcing memory safety exist that also consider the secure compilation domain, i.e., have an active attacker model.
-For example, the ``pointers as capabilities'' principle represents pointers as machine-level capabilities~\cite{korashy2021capableptrs}, which behave in a similar fashion to capabilities by means of linear typing~\cite{morrisett2005L3}.
-The approach of this paper also uses linear typing, but differs from $L^{3}$~\cite{morrisett2005L3} in the way that functions are not first-class.
-Moreover, this paper considers an active attacker, while the work on $L^{3}$ only discusses whole programs and, thus, has no active attacker model.
-The instrumentation to ensure memory safety that this paper presents is inspired by Softbounds~\cite{nagarakatte2009soft}.
-That work inserts bounds-checks in front of pointer-dereferences and, for this to work, inserts meta-data information on pointer creation.
-Softbounds also works in a more advanced setting with structured fields accesses and also introduces a table-lookup for pointers that are stored in memory.
-This paper only considers arrays of primitive data, i.e., there are no pointers to pointers or structures.
-Several other approaches to memory-safety exist in literature, specifically as compiler instrumentations~\cite{akritidis2009baggy,younan2010paricheck,jung2021pico,shankaranarayana2023tailcheck,dhumbumroong2020boundwarden,nam2019framer,zhou2023fatptrs}, hardware-extensions~\cite{kwon2013lowfat,saileshwar2022heapcheck,chen2023flexpointer,kim2023whistle}, or programming language extensions~\cite{elliott2018checkedc,li2022formalcheckedc,jim2002cyclone,elliott2015guilt,west2005cuckoo,weis2019fyr,benoit2019uniqueness}.
-What differentiates this work from them is that this work uses known, compiler-based approaches to ensure memory-safety as a means to investigate secure compiler compositions.
-This paper does not provide efficient memory-safety, but serves as a theoretical foundation for the secure compilation domain.
-
-To extend the languages in this paper with a less restricted form of pointer arithmetic, the region colouring memory safety monitor presented in earlier work~\cite{michael2023mswasm} can be used.
-The work presenting this monitor provides an approach for the robust preservation of memory safety compiling from C to WASM.
-However, they do not discuss composition of secure compilers but rather investigate an instance of a secure compiler.
-
-\paragraph*{Cryptographic Constant Time Mechanisms}\label{subsec:relw:cctmechs}
-
-The approach to preserving cryptographic constant time in this paper is high-level, where a programming language exposes a way to switch the semantics to a data (operand) independent timing mode.
-Since identifiers in $\Lscct$ are annotated with a secrecy tag, this approach is similar to others with information flow control.
-For example, Vale~\cite{bond2017vale} uses Dafny to ensure constant-time assembly code, while Jasmin~\cite{almeida2017jasmin} makes use of the Coq proof assistant to reject non-constant-time programs.
-CT-Wasm~\cite{watt2019ctwasm} enforces constant-timeness by means of a type system.
-Different to the approach of this paper, these approaches necessitate that the programmer writes \gls*{cct} code.
-An approach to allow programmers to write more high-level code is CryptOpt~\cite{kuepper2023cryptopt}, which generates efficient target-code by means of a randomised search.
-This paper abstracts over concrete mitigation strategies and simply assumes that there is a flag to switch to a cryptographic-constant time execution mode.
-This can be realised by employing the FaCT~\cite{cauligi2019fact} compiler, which translates common non-constant time code patterns to be constant-time, and the data (object) independent timing execution mode of modern processors.
-
-\paragraph*{Speculation Safety Mechanisms}\label{subsec:relw:ssmechs}
-
-This paper uses a taint-tracking mechanism inspired by existing work~\cite{guarnieri2018spectector,fabian2022automatic}.
-These taints are used to express absence of any speculative leaks in \gls*{ss}~\cite{guarnieri2018spectector}. 
-The semantics of $\Lspec$ hardcodes the kind of speculative leaks to just SPECTRE-PHT~\cite{kocher2019spectre}, but future work could use semantics composition~\cite{fabian2022automatic} to support more variants.
-Note that our framework composes compilers and not semantics.
-
-%This work has taken inspiration in the sematics of $\Lspec$ and in \Cref{def:trace:ss} from recent formalisations~\cite{guarnieri2018spectector,fabian2022automatic}.
-% \MPin{
-%  we take SS from exorcising.
-%  we do not take SNI from spectector because it is a hypersafety.
-%  other work defined similar props. 
-% }
-% A subsequent line of work~\cite{fabian2022automatic} shows how to modularly define and compose formal semantics that models different speculative execution variants. 
-% Recent work~\cite{fabian2024lift} demonstrates how to lift existing secure compilation proofs for security properties related to speculative execution to more powerful properties, i.e., properties defined for languages with a trace model that enables more vulnerabilities, such as combinations of different speculative execution exploits.
-% To this end, they define a lifting framework and demonstrate how a secure compilation proof for a weaker property can be lifted to a stronger one. 
-% Contrary to the compositionality framework presented in this paper, these works are only concerned with composing semantics of the respective programming languages, while in this paper, the secure compilers themselves are composed.
-
-\section{Conclusion\pages{$\sfrac{1}{2}$}}\label{sec:concl}
-This paper tackles the problem of understanding what kind of security properties a secure compiler preserves, when said compiler is the combination of compiler passes that preserve possibly different security properties.
-% 
-%For this, this paper first formalised security properties of interest and their composition.
-% 
-The paper proves that composing secure compilers that preserve certain properties results in a secure compiler that preserves the composition of these properties.
-% While the presented security property composition relied on monitors that check only trace-properties, the composition of secure compilers does not make any restriction towards the kind of property involved in the composition.
-% It is subject to future work to develop techniques to verify relational hyperproperties~\cite{abate2019jour}, while the composition of hyperproperties could be very similar to the composition of ordinary properties as presented in this paper, since hyperproperties can be checked with an automata-based model-checker~\cite{beutner23hyperltl}.
-% 
-Finally, this paper defines a multi-pass compiler and proves that it preserves \gls*{specms}.
-Crucially, this paper derives the security of the multi-pass compiler from the composition of the security properties preserved by its individual passes, which include security-preserving as well as optimisation passes.
-% For future work, it is interesting to look at more sophisticated optimisation passes that, e.g., reorder events that appear on traces.
-
-\newpage
-
-\bibliographystyle{IEEEtranS}
-\bibliography{main}
-
-%%
-%% If your work has an appendix, this is the place to put it.
-\appendix
-\subsection{Existential Image}\label{subsec:extimg}
-
-\begin{definition}[Existential Image]
-% [Existential and Universal Image]
-\label{def:existential:img}\label{def:tau}
-  \[
-    \tau_\sim\left(\src{\pi}\right) := 
-      \left\{ 
-        \trg{\trace} \mid \exists \src{\trace}\ldotp \src{\trace}\sim\trg{\trace}, \text{ and } \src{\trace}\in\src{\pi} 
-      \right\}
-  \]
-  %\[ 
-  %  \sigma_\sim\left(\trg{\pi}\right) := 
-  %    \left\{ 
-  %      \src{\trace} \mid \forall \trg{\trace}\ldotp \text{if }\src{\trace}\sim\trg{\trace}, \text{ then } \trg{\trace}\in\trg{\pi} 
-  %    \right\}
-  %\]
-\end{definition}
-
-\begin{definition}[Robust Preservation with $\tau_\sim$]\label{def:rtp:tau}
-  % $\;$\\
-  % \vspace{-1em}
-  % \begin{nscenter}
-  % Compiler $\cc{\src{L}}{\trg{L}}$ robustly preserves $\class$, 
-  {$\rtptau{\cc{\src{L}}{\trg{L}}}{\src{\class}}{\sim}$}
-  %, iff 
-  $\isdef$
-    {$\forall \src{\pi}\in\src{\class}, \src{p}\in\src{L},$} if {$\rsat{\src{\progvar}}{\src{\pi}}$}, then {$\rsat{\cc{\src{L}}{\trg{L}}\left(\src{p}\right)}{\tau_\sim\left(\src{\pi}\right)}$}.
-  % \end{nscenter}
-\end{definition}
-
-\begin{theorem}[Composition of Secure Compilers w.r.t. $\tau$]\label{thm:rtpsim:tau}
-  $\;$
-
-  If {$\rtptau{\cc{\src{L}}{\trg{L}}}{\src{\class[_{1}]}}{\sim_1}$}, {$\rtptau{\cc{\trg{L}}{\obj{L}}}{\tilde{\tau}_{\sim_1}\left(\src{\class[_2]}\right)}{\sim_2}$}, and {$\wfctau{\sim_1}{\src{\class[_2]}}$}, \\ then {$\rtptau{\cc{\src{L}}{\trg{L}}\circ\cc{\trg{L}}{\obj{L}}}{\src{\class[_{1}]}\cap\src{\class[_{2}]}}{\sim_1\bullet\sim_2}$}.
-\Coqed
-\end{theorem}
-
-\begin{corollary}[Swapping Secure Compiler Passes]\label{corr:swappable:tau}
-  $\;$
-
-  If {$\rtptau{\cc[_{1}]{\trg{L}}{\trg{L}}}{\src{\class[_{1}]}}{\sim_1}$ and $\rtptau{\cc[_{2}]{\trg{L}}{\trg{L}}}{\src{\class[_{2}]}}{\sim_2}$}, %
-  {$\wfctau{\sim_1}{\src{\class[_2]}}$ and $\wfctau{\sim_2}{\src{\class[_1]}}$}, %
-  and {$\tilde{\tau}_{\sim_1}\left(\src{\class[_2]}\right)=\src{\class[_2]}$ as well as $\tilde{\tau}_{\sim_2}\left(\src{\class[_1]}\right)=\src{\class[_1]}$},
-  then {$\rtptau{\cc[_{1}]{\trg{L}}{\trg{L}}\circ\cc[_{2}]{\trg{L}}{\trg{L}}}{\src{\class[_{1}]}\cap\src{\class[_{2}]}}{\sim_1\circ\sim_2}$ and $\rtptau{\cc[_{2}]{\trg{L}}{\trg{L}}\circ\cc[_{1}]{\trg{L}}{\trg{L}}}{\src{\class[_{2}]}\cap\src{\class[_{1}]}}{\sim_2\circ\sim_1}$}.
-\Coqed
-\end{corollary}
-
-\subsection{Secure Upper and Lower Composition}\label{sec:other-compos}
-Besides sequential composition, there are two other compositions, namely an {\em upper}, i.e., a compiler that takes multiple inputs and yields one output, and a {\em lower} composition, i.e., a compiler that takes one input and yields multiple outputs.
-We {define the upper composition $\cc{\src{L}+\obj{L}}{\trg{L}}$} as follows:
-Given a program \texttt{p}, its compiled counterpart is obtained by {plugging \texttt{p} into $\cc{\src{L}}{\trg{L}}$ if $\texttt{p}\in\src{L}$} or by {plugging \texttt{p} into $\cc{\obj{L}}{\trg{L}}$ if $\texttt{p}\in\obj{L}$}.
-\begin{definition}[Upper Composition]
-  \[
-    \text{{$\cc{\src{L}+\obj{L}}{\trg{L}}$}}\isdef
-  % Given \texttt{p}, yield
-  \lambda \texttt{p}\ldotp
-  \left\{\mbox{\begin{tabular}{c}
-    {{if $\texttt{p}\in\src{L}$, then $\cc{\src{L}}{\trg{L}}(\texttt{p})$}} \\
-    \mbox{{if $\texttt{p}\in\obj{L}$, then $\cc{\obj{L}}{\trg{L}}(\texttt{p})$}} \\
-  \end{tabular}}\right. 
-  \]
-\end{definition}
-
-Examples of this are present in industry:
-Consider the Java Virtual Machine bytecode $\trg{JVM BC}$, which is a popular target for programming language designers due to its high performance and relevance in industry.
-Compilers for several programming languages have it as their target language, some popular instances are $\src{Java}$ and $\obj{Kotlin}$.
-Technically speaking, they both compile to class files and $\obj{Kotlin}$ objects are considered to be the same as $\src{Java}$ objects at that point.
-Both languages can be used at the same time in one project~\cite{androidstudio}.
-A compiler that accepts both $\src{Java}$ and $\obj{Kotlin}$ code translating to the same target language or intermediate representation performs a kind of {\em upper} composition.
-Now, the following theorem tells us what happens if these are secure:
-Given {$\cc{\src{L}}{\trg{L}}$ robustly preserves $\class[_{1}]$} and {$\cc{\obj{L}}{\trg{L}}$ robustly preserves $\class[_{2}]$}, it follows that {their upper composition $\cc{\src{L}+\obj{L}}{\trg{L}}$ robustly preserves the intersection of classes $\class[_{1}]$ and $\class[_{2}]$}.
-
-\begin{theorem}[Upper Composition of Secure Compilers]\label{thm:urtp}
-  $\;$
-
-  If {$\rtp{\cc{\src{L}}{\trg{L}}}{\class[_{1}]}$} and {$\rtp{\cc{\obj{L}}{\trg{L}}}{\class[_{2}]}$}, then {$\rtp{\cc{\src{L}+\obj{L}}{\trg{L}}}{\class[_{1}]\cap\class[_{2}]}$}. %\Coqed
-\end{theorem}
-
-Dually, the {\em lower} composition is concerned about compilers that accept the same source but yield different target languages. %and the same theoretical results apply.
-{Define the lower composition $\cc{\src{L}}{\trg{L}+\obj{L}}$} as follows:
-Given a program $\src{p}$, its compiled counterpart is obtained by {plugging $\src{p}$ into $\cc{\src{L}}{\trg{L}}$} or by {plugging \texttt{p} into $\cc{\src{L}}{\obj{L}}$}, respectively, {based on the internal decision}.
-\begin{definition}[Lower Composition]
-  \[
-    \text{{$\cc{\src{L}}{\trg{L}+\obj{L}}$}}\isdef
-  % Given $L$ and $\src{p}$, yield
-  \lambda \src{p}, L\ldotp
-  \left\{\mbox{\begin{tabular}{c}
-    {{if $L=\trg{L}$, then} {$\cc{\src{L}}{\trg{L}}(\src{p})$}} \\
-    \mbox{{if $L=\obj{L}$, then} {$\cc{\src{L}}{\obj{L}}(\src{p})$}} \\
-  \end{tabular}}\right.\]
-\end{definition}
-
- Consider two compilers both accepting $\src{LLVM IR}$~\cite{lattner2004llvm} and one of them emits $\trg{x86\_64}$, while the other emits $\obj{ARMv8}$.
- It is intuitive that they are in some sense composed in the LLVM framework, but the decision of when to use one over the other is inherently {\em internal} to the formalisation effort of this kind of composition.
- For example, the user of this compiler provides an explicit flag that instructs to emit $\trg{x86\_64}$ or the framework itself detects the target platform via heuristics, such as supported instructions.
-
- The following theorem demonstrates what happens if the involved compilers are secure:
-Given {$\cc{\src{L}}{\trg{L}}$ robustly preserves $\class[_{1}]$} and {$\cc{\src{L}}{\obj{L}}$ robustly preserves $\class[_{2}]$}, it follows that {their lower composition $\cc{\src{L}}{\trg{L}+\obj{L}}$ robustly preserves the intersection of classes $\class[_{1}]$ and $\class[_{2}]$}.
-
-\begin{theorem}[Lower Composition of Secure Compilers]\label{thm:lrtp}
-  $\;$
-
-  If {$\rtp{\cc{\src{L}}{\trg{L}}}{\class[_{1}]}$} and {$\rtp{\cc{\obj{L}}{\trg{L}}}{\class[_{2}]}$}, then {$\rtp{\cc{\src{L}}{\trg{L}+\obj{L}}}{\class[_{1}]\cap\class[_{2}]}$}. % \Coqed
-\end{theorem}
-
-Either way, the theoretical results suggest that it is possible to always find a ``most-general'', secure compiler, given two secure compilers, that robustly preserves the least-upper bound of the classes involved in their compilation process.
-
-%Future work intends to use upper and lower composition as means to build a secure compiler incrementally
-%This follows current non-formal-methods based compiler engineering, such as LLVM~\cite{lattner2004llvm}, where languages such as \texttt{C++} and \texttt{Fortran} translate into LLVM IR and languages such as \texttt{x86} or \texttt{ARMv8} assembly variants translate out of LLVM IR.
-
-% no coq, because above definitions are non-formal. 
-%   How would you precisely define the lower compo compiler, given two compilers?
-
-%\section{Cryptographic Constant Time}\label{sec:cct}
-
-\end{document}
-\endinput

### 2025-02-13 icfp/sections/langs.tex
move

--- 2024-09-24 csf2/main.tex
@@
-% Fixing: Too many math alphabets used in version normal
-\newcommand\hmmax{0}
-\newcommand\bmmax{0}
-\documentclass[dvipsnames,conference]{IEEEtran}
-
-\usepackage[T1]{fontenc}
-\usepackage[scaled=.83]{beramono}
-
-\usepackage[colorinlistoftodos]{todonotes}
-\usepackage[inference]{semantic}
-\usepackage[switch]{lineno}
-\usepackage{halloweenmath}
-\usepackage{fontawesome5}
-\usepackage{listofitems}
-\usepackage{breakcites}
-\usepackage{glossaries}
-\usepackage{hyperref}
-\usepackage{cleveref}
-\usepackage{stmaryrd}
-\usepackage{marvosym}
-\usepackage{listings}
-\usepackage{amssymb}
-\usepackage{amsmath}
-\usepackage{nameref}
-\usepackage{amsthm}
-\usepackage{xspace}
-\usepackage{xfrac}
-\usepackage{tikz}
-\usepackage{soul}
-\usepackage{bm}
-\usepackage{paralist}
-
-\hypersetup{
-    colorlinks,
-    linkcolor={red!50!black},
-    citecolor={blue!50!black},
-    urlcolor={blue!80!black}
-}
-%\renewcommand\UrlFont{\color{blue}\rmfamily}
-%\newcommand{\url}[1]{\lstinline{#1}}
-
-\setul{0.95ex}{0.3ex}
-
-\input{cmds.tex}
-
-% allow page breaks inside multiline alignment displays
-\allowdisplaybreaks
-
-\Crefname{exampleenv}{Example}{Examples}
-
-\theoremstyle{definition}
-\newtheorem{exampleenv}{Example}[section]
-\newtheorem{lemma}{Lemma}[section]
-\newtheorem{theorem}{Theorem}[section]
-\newtheorem{corollary}{Corollary}[section]
-\newtheorem{definition}{Definition}[section]
-
-\loadglsentries{acronyms}
-\makeglossaries
-
-\AtBeginDocument{%
-  \providecommand\BibTeX{{%
-    \normalfont B\kern-0.5em{\scshape i\kern-0.25em b}\kern-0.8em\TeX}}}
-\makeatletter
-\newcommand{\linebreakand}{%
-  \end{@IEEEauthorhalign}
-  \hfill\mbox{}\par
-  \mbox{}\hfill\begin{@IEEEauthorhalign}
-}
-\makeatother
-
-\renewcommand\thelinenumber{\color{red}\arabic{linenumber}}
-\begin{document}
-\linenumbers
-
-\title{
-  Secure Composition 
-  \\ 
-  of Robust and Optimising Compilers
-  % \vspace{-1.5em}
-}
-
-% \author{
-% Matthis Kruse\orcidID{0000-0003-4062-9666}
-% \and
-% Michael Backes
-% \and
-% Marco Patrignani
-% }
-% \authorrunning{M. Kruse et al.}
-
-\maketitle
-
-\begin{abstract}
-% \MPpost{
-%   this abstract fails in setting up the robustness argument.
-%   it'll be a good exercise for you to take inspiration from this, but then write something as concise but with a more direct focus on robust properties
-% }
-% context
-To ensure that secure applications do not leak their secrets, they are required to uphold several security properties such as spatial and temporal memory safety, cryptographic constant time, as well as speculative safety.
-% need
-Existing work shows how to enforce these properties individually, in an architecture-independent way, by using secure compiler passes that each focus on an individual property.
-% task
-Unfortunately, given two secure compiler passes that each preserve a possibly different security property, it is unclear what kind of security property is preserved by the composition of those secure compiler passes.
-%there is no way to tell what kind of security property will the composition of those secure compilers preserve.
-
-% object
-This paper is the first to study what security properties are preserved across the composition of different secure compiler passes.
-% findings
-Starting from a general theory of property composition for security-relevant properties (such as the aforementioned ones), this paper formalises a theory of composition of secure compilers.
-Then, it showcases this theory on a secure multi-pass compiler that preserves the aforementioned security-relevant properties.
-% conclusion
-Crucially, this paper derives the security of the multi-pass compiler from the composition of the security properties preserved by its individual passes, which include security-preserving as well as optimisation passes.
-% 
-From an engineering perspective, this is the desirable approach to building secure compilers.
-\begin{center}\small\it
-	{This paper uses syntax highlighting accessible to both colourblind and black \& white readers.
-	% ~\cite{patrignani2020use}.
-% 	% Specifically, it makes use of a $\src{blue}$, $\src{sans\text{-}serif}$ font for a $\src{source}$,
-% 	% a $\trg{red}$, $\trg{bold}$ font for an $\trg{intermediate}$,
-% 	% and a $\obj{green}$, $\obj{teletype}$ font for a $\obj{target}$ language.	
-	For a better experience, please print or view in colours.
-	}
-\end{center}
-\end{abstract}
-
-\begin{IEEEkeywords}
-  memory-safety, cryptographic constant-time, secure compilation, robustness
-\end{IEEEkeywords}
-
-\section{Introduction\pages{4}}\label{sec:introduction}
-
-% \gls*{ms} is a security property that ensures that there are no out-of-bounds accesses as well as no reads or writes to uninitialised memory.
-% Programs that do not attain memory safety, can easily have their private data be leaked to attackers or be corrupted~\cite{lemay2021ccc}.
-% Because of this, \gls*{ms} is a requirement for programs to attain an even stronger security property, namely \gls*{cct}.
-% Programs that attain \gls*{cct} are secure w.r.t.
-timing attacks, and this is the gold standard for secure applications, including cryptographic ones, since timing attacks can expose secrets via a simple performance analysis~\cite{kocher1996timing}.
-% Timing attacks have become more and more relevant recently, especially given the era of cloud-computing~\cite{aviram2010cloudtime,kumar2019cloudsecsurvey}, where programs share the same hardware but are supposed to run in a sandboxed environment that does not leak secrets.
-% Unfortunately, several attacks within such a cloud environment have been demonstrated as viable strategies to extract secrets~\cite{mehmet2015getoff,flowers2022zeroday,atya2019catchme,venkatanathan2015placevul}.
-
-% Even though mitigations for such attacks exist~\cite{bond2017vale,almeida2017jasmin}, some compilers may optimise these mitigations away~\cite{barthe2018sec}.
-% This is of course not what programmers expect and to ensure such mitigations persist in the compiled code, one can use a {\em secure} compilers.
-% More generally, secure compilers should preserve security properties even when the program runs in a hostile context such as the cloud one, i.e., where it interacts with other possibly malicious programs.
-% An example of such secure compilers are those that ascribe to the theory of robust compilation~\cite{abate2019jour,abate2021extacc,patrignani2021rsc}, which takes this interaction between components and potentially malicious contexts into account.
-
-% Unfortunately, it is unclear if and how robust compilation supports compositionality, and compilation passes are typically developed as a standalone module to be plugged into a compilation pipeline and composed with other passes.
-% Different kinds of compilation passes exist, e.g., there are usually several compilation passes that perform optimisations~\cite{lattner2004llvm,googlev8,androidstudio,kuepper2023cryptopt,manjikian1997fusion,wegman1991ccp} or source-code instrumentations~\cite{nagarakatte2009soft,nagarakatte2010cets,akritidis2009baggy,dhumbumroong2020boundwarden,jung2021pico,nam2019framer,shankaranarayana2023tailcheck,younan2010paricheck,zhou2023fatptrs,bond2017vale,almeida2017jasmin,kuepper2023cryptopt,cauligi2019fact} that enforce that the compiled program fulfils a property of interest.
-% Without a framework for composing secure compiler passes, it is not possible to enable separation of concerns, e.g., to have a secure compilation pass that ensures \gls*{ms} that is developed independently of another secure pass for \gls*{cct} as suggested before and as showcased in \Cref{ex:strncpy}.
-% \MPpost{
-%   this intro does a poor job at introducing the robust compilation argument.
-%   perhaps it'd be best to start from robust properties, which motivates the use of robust compilers?
-%   idk
-%   as for the abstract, this is a good starting point for you to make a better intro
-% }
-
-\gls*{ms} is a security property obtained by composing \gls*{sms}, which ensures array accesses are all within bounds, and \gls*{tms}, which ensures pointers are only used when they are valid~\cite{azevedo2018meaningsofms,jim2002cyclone,necula2005ccured,nagarakatte2010cets,nagarakatte2009soft,akritidis2009baggy,michael2023mswasm}.
-\gls*{cct} is a security property that ensures sensitive data is not leaked via timing side-channels~\cite{kocher1996timing} and \gls*{ss} is a security property that enforces the same but under a speculative semantics~\cite{guarnieri2018spectector,fabian2022automatic} that captures speculative execution attacks such as spectre~\cite{kocher2019spectre}.
-Together, \gls*{sms}, \gls*{tms}, \gls*{cct} and \gls*{ss}, yield \gls*{specms}, which is the gold standard of security properties for secure applications.
-Programs attaining \gls*{specms} do not leak sensitive data either through erroneous memory accesses, nor through timing side-channels, even under speculative execution.
-\Cref{ex:strncpy} below discusses how these security properties can be enforced by compiler passes~\cite{bond2017vale,almeida2017jasmin}, to ensure programmers need not be aware of the architectural details of where their code will run.
-
-\begin{exampleenv}[strncpy]\label{ex:strncpy}
-Consider the following \texttt{C} function \texttt{strncpy} that copies a null-terminated string \texttt{src} into \texttt{dst} up to a length of \texttt{n} characters.
-This function is subject to a subtle \gls*{sms} vulnerability: the bounds check \texttt{i < n} should happen {\it before} the access to memory location \texttt{src[i]}: otherwise
-% If \texttt{n} is at least as large as the length of the array \texttt{src}, 
-the memory location past the last element is leaked to an attacker.
-\begin{lstlisting}[language=c,basicstyle=\small\ttfamily,morekeywords={size_t}]
-void strncpy(size_t n, char *dst, char *src) {
-  for(size_t i = 0; src[i] != '\0' && i < n; ++i) {
-    dst[i] = src[i];
-  }
-}
-\end{lstlisting}
-% Assume that the length of the array pointed by \texttt{dst} is at least \texttt{n}, and \texttt{src} is terminated by \texttt{'\textbackslash 0'}.
-% Even under these assumptions, there is a 
-To prevent this vulnerability, one can use a compilation pass that enforces \gls*{sms}, such as Softbounds~\cite{nagarakatte2009soft} or BaggyBounds~\cite{akritidis2009baggy}.
-The most na\"ive solution in this case is to insert bounds-checks in front of every access to memory.
-
-Because of timing attacks, addressing \gls*{sms} is not enough to make \texttt{strncpy} secure.
-In fact, the loop can terminate early, as soon as the string-terminating character \texttt{'\textbackslash 0'} is encountered, thus making program execution time proportional to the length of the array pointed by \texttt{src}, and violating the \gls*{cct} property.
-% Often, programmers rewrite programs like these to have loops that run with a fixed number of iterations.
-Also in this case there exist compiler passes that can rewrite such programs into \gls*{cct} ones~\cite{cauligi2019fact}.
- % so that they are no longer vulnerable to timing attacks.
-%
-Finally, even with these precautions, code is not run in isolation, so a malicious attacker could supply code that interacts with \texttt{strncpy} and trigger a violation of either \gls*{ms} or \gls*{cct} by calling \texttt{strncpy} with an argument for \texttt{src} that points to uninitialised memory.
-This would, in turn, trigger a series of reads from uninitialised memory, which is an immediate \gls*{ms} violation with devastating real-world consequences~\cite{uninit-0,uninit-1,uninit-2,uninit-3,uninit-4}.
-% An easy fix to sidestep this is to use non-nullable pointers as parameters instead, e.g., as presented in Cyclone~\cite{jim2002cyclone}.
-\end{exampleenv}
-
-  % last sentence is before the example, which talks about compiler passes enforcing props
-Whether or not compiler passes enforce certain security properties, it is important for a secure compiler to consider partial programs that interact with potentially malicious code, since the latter may lead to, e.g., memory-safety issues in the considered partial program.
-Robust compilers~\cite{abate2019jour} are a form of secure compilers that preserve security properties even in the presence of arbitrary attackers interacting with compiled code.
-Thus, robust compilers can be used to prevent vulnerabilities resulting from uninitialised memory (as well as many other ones), e.g., by targeting capability-based languages such as CHERI~\cite{woodruff2014CHERI}, Arm Morello~\cite{arm-morello}, or MSWasm~\cite{michael2023mswasm}, where the compiler relies on capabilities to check that pointers are always initialised.
-
-% To ensure these compiler-preserved properties still hold even in the presence of arbitrary attackers interacting with compiled code, one can use secure compilers in the form of robust compilers~\cite{abate2019jour,abate2021extacc,patrignani2021rsc}.
-
-Unfortunately, given secure compiler passes that each preserve a possibly different security property, there is no way to tell what kind of security property will the composition of those secure compilers preserve.
-% Different kinds of compilation passes exist, e.g., there are usually several compilation passes that perform optimisations~\cite{lattner2004llvm,googlev8,androidstudio,kuepper2023cryptopt,manjikian1997fusion,wegman1991ccp} or source-code instrumentations~\cite{nagarakatte2009soft,nagarakatte2010cets,akritidis2009baggy,dhumbumroong2020boundwarden,jung2021pico,nam2019framer,shankaranarayana2023tailcheck,younan2010paricheck,zhou2023fatptrs,bond2017vale,almeida2017jasmin,kuepper2023cryptopt,cauligi2019fact} that enforce that the compiled program fulfils a property of interest.
-Worse, without a framework for composing secure compiler passes, it is not possible to enable separation of concerns, e.g., to have a secure compilation pass that ensures \gls*{ms} that is developed independently of another secure pass for \gls*{cct}, that is developed independently of other passes, such as optimisation ones.
-
-This paper introduces a framework for reasoning about the composition of secure and optimising compiler passes %akin to those of \Cref{ex:strncpy}\MP{there aren't really passes there} 
-and it showcases the power of this framework by instantiating it on a multi-pass compilation chain.
-To this end, this paper first discusses how to compose security properties, such as \gls*{tms} and \gls*{sms} into \gls*{ms}, and then adds \gls*{cct} as well as \gls*{ss} into the mix to obtain \gls*{specms}.
-%Then, this paper defines compiler composition and formalises that given two passes that securely preserve two (possibly distinct) properties, their composition securely preserves the composition of those properties.
-The paper then defines several secure compiler passes, where each is either preserving a different security property from the list above
-% (\gls*{tms}, \gls*{sms}, \gls*{scct}, \gls*{ss}) 
-or performing a security-preserving optimisation.
-% , (e.g., applying \gls*{cf} or \gls*{dce}).
-Finally, this paper shows that composing these secure compiler passes into a multi-pass compilation chain results in the end-to-end (robust) preservation of \gls*{specms}.
-Crucially, this paper derives the security of the multi-pass compiler from the composition of the security properties preserved by its individual passes.
-% Crucially, this end-to-end security preservation is obtained by composing the security property preserved by each individual pass, as dictated by the framework of this paper.
-This result showcases how the framework allows the kind of formal security reasoning that compiler writers already want (and already do), obtaining precise, compositional security reasoning while providing minimal (and modular) proof effort.
-%
-In summary, this paper makes the following contributions:
-% \begin{itemize}[leftmargin=0em,nosep,label=$\blacktriangleright$]
-\begin{asparaitem}[$\blacktriangleright$]
-  \item %
-        %\Cref{sec:sequential} studies different forms of secure compiler compositions.
-        This paper takes the secure compilation framework of Abate et al.~\cite{abate2021extacc} and extends it for compositionality. (\Cref{sec:sequential}).
-        %Concretely, it investigates upper and lower compositions, as well as the sequential composition.
-        %For this, this paper studies compiler composition as well as compilers with multiple input languages or multiple output ones, as used in existing compilation chains.
-        %
-        This paper proves that starting from two compilers that preserve two (possibly distinct) properties, their composition preserves roughly the intersection of those properties.
-        % For example, a compiler that is secure w.r.t. \gls*{tms} can be composed with a compiler that is secure w.r.t. \gls*{sms}, resulting in a compiler that is secure w.r.t. \gls*{ms}.
-        % 
-        % From this, the paper proves that the order of composition of sequential compiler passes is irrelevant for security.
-        % % 
-        % This is crucial for reordering individual optimisation passes and thus generating secure and efficient code.
-        %
-        %Finally, this paper extends parts of the compositionality framework to support different trace-models between compilation passes.
-        % \MP{this isn't really true, is it?. the order matters for the relations' lemmas}
-        Then, this paper identifies which conditions make the composition of secure compilers meaningful from a security perspective, and which other conditions allow passes to be swapped without losing security meaning.
-
-  \item %
-        This paper presents a case-study with five programming languages showcasing the previous contribution (\Cref{sec:casestud:defs,sec:casestud:rtp}).
-        To this end, it presents a compilation chain consisting of six passes that ultimately preserve \gls*{specms} by means of composing four secure passes that \emph{individually} preserve \gls*{tms}, \gls*{sms}, \gls*{scct}---an enforceable version of \gls*{cct}---, and \gls*{ss}, and two passes which are well-known optimisations \gls*{dce} and \gls*{cf}.
-        %Furthermore, the chain includes two optimisation passes: one performs \gls*{dce} and the other \gls*{cf}.
-        The formalisation of this case study showcases the power of the presented framework: the divide-and-conquer approach to software engineering is a viable strategy even for the development of secure compilers.
-%        This demonstrates that the usage of the presented framework is both simpler and more expressive compared to one monolithic security proof of the whole compilation chain.
-
-  \item %
-  		This paper provides formal proofs of the security properties preserved by each of the presented passes. 
-  		% 
-  		Additionally this paper analyses each pass in terms of their compatibility with each other (\Cref{sec:formalities}), proving how does each pass fulfill the conditions that make their composition meaningful from a security standpoint.
-        % In doing so, the paper provides formal insights into some technicalities of the concrete proofs necessary as technical setup to reap the benefits of the presented compositionality framework.
-
-  \item The key contributions of this paper are formalised in the Coq proof assistant and the paper indicates this with \CoqSymbol.
-\end{asparaitem}
-% 
-This paper starts by introducing relevant notions of security properties and secure compilation (\Cref{sec:background}),
-and discusses related work (\Cref{sec:relwork}) before concluding (\Cref{sec:concl}).
-
-% \contribution{Technical Report} 
-The omitted formal details, lemmas and proofs, as well as the Coq formalisation are available as supplementary material.
-% \MKin{cite/link}
-
-\section{Background: Properties and Secure Compilers\pages{1}}\label{sec:background}
-
-To introduce the security argument of this paper, this section defines (security) properties, their satisfaction, and their robust satisfaction (i.e., satisfaction in the presence of an active attacker; \Cref{subsec:bg:tprop}).
-Then, borrowing from existing work~\cite{abate2019jour,abate2021extacc}, this section introduces secure compilers as compilers that preserve robust property satisfaction (\Cref{subsec:bg:rtp}).
-
-\subsection{Properties and (Robust) Satisfaction}\label{subsec:bg:tprop}
-
-This paper employs the security model where programs are written in a language whose semantics emits events $\event$.
-Events include security-relevant actions (e.g., reading from and writing to memory, as detailed in \Cref{sec:compprop}) and the unobservable event $\emptyevent$.
-As programs execute, their emitted events are concatenated in traces $\trace$, which serve as the description of the behaviour of a program.%
-\footnote{
-Throughout the paper, sequences are indicated with an overbar (i.e., $\trace$), empty sequences with $\hole{\cdot}$, and concatenation of sequences $\trace[_{1}],\trace[_{2}]$ as $\trace[_{1}]\cdot\trace[_{2}]$.
-Prepending element $\event$ to a sequence $\trace$ uses the same notation: $\event\cdot\trace$.
-}
-
-Properties $\pi$ are sets of traces of admissible program behaviours, ascribing what said property considers valid.
-% 
-The set of all properties can be partitioned into different {\em classes} ($\class$), i.e., safety, liveness, and neither safety nor liveness~\cite{clarkson2008hyper}, so a class is simply a set of properties.
-% 
-The compositionality framework (\Cref{sec:sequential}) presented in this paper considers arbitrary classes, while the case-study (\Cref{sec:casestud:rtp}) fixes them to concrete instances of safety properties, since it is decidable whether a trace satisfies a safety property with just a finite trace (i.e., a \emph{prefix}).
-As an example, consider the trace:
-$$\ev{Dealloc\ \loc}\cdot\ev{Read\ \loc\ 1729}\cdot\dots$$
-which describes the interaction with a memory where the deallocation of an address $\loc$ precedes a read (of some value $1729$) at that address in memory.
-% 
-This program behaviour is insecure w.r.t.
-a canonical notion of (temporal) memory safety dictating no use-after-frees of pointers~\cite{nagarakatte2010cets,azevedo2018meaningsofms}, because it reads from a memory location that was freed already.
-The previous finite trace prefix is enough to decide that the trace does not satisfy \gls*{tms} and there is no way to append events to this prefix which would result in the trace being admissible.
-
-In the following, the execution of a whole program $\wholeprogvar$ that terminates in state $\runtimetermvar$ according to the language semantics and produces trace $\trace$ is written as $\progstepto{\wholeprogvar}{\runtimetermvar}{\trace}$.
-With this, we defined property satisfaction as follows:
-\bul{a whole program $\wholeprogvar$ satisfies a property $\pi$} iff \iul{$\wholeprogvar$ yields a trace $\trace$} such that \oul{$\trace$ satisfies $\pi$}.
-
-\begin{definition}[Property Satisfaction]\label{def:propsat}
-$\;$
-
-  \begin{nscenter}
-    \bul{$\sat{\wholeprogvar}{\pi}$}
-    $\isdef$
-    if \iul{$\forall\runtimetermvar\ \trace,\progstepto{\wholeprogvar}{\runtimetermvar}{\trace}$},
-    then \oul{$\trace\in\pi$}.
-  \end{nscenter}
-\end{definition}
-
-Property satisfaction is defined on whole programs, i.e., programs without missing definitions.
-Thus, from a security perspective, \Cref{def:propsat} considers only a passive attacker model, where the attacker observes the execution and, e.g., retrieves secrets from that.
-To consider a stronger model similarly to what existing work does~\cite{abate2019jour,abate2021extacc,maffeis2008code-carrying,gordon2003authenticity,fournet2007authorization,bengtson2011refine,backes2014uniontyps,michael2023mswasm,swasey2017robust,sammler2019benefits}, we extend the concept of satisfaction with {\em robustness}.
-Robust satisfaction considers partial programs $\progvar$, i.e., components with missing imports, which cannot run until said imports are fulfilled.
-To remedy this, {\em linking} takes two partial programs $\progvar[_{1}],\progvar[_{2}]$ and produces a whole program $\wholeprogvar$, i.e., $\link{\progvar[_{1}]}{\progvar[_{2}]}=\wholeprogvar$.
-As typically done in works that consider the execution of partial programs~\cite{abate2019jour,devriese2018parametricity,patrignani2021rsc,korashy2021capableptrs,strydonck2019lincap,devriese2017modular,bowman2015noninterference,ahmed2011equivcps,patterson2017linkingtyps},
-this paper assumes that whole programs are the result of linking partial programs referred to as {\em context} ($\ctx$) and {\em component} ($\comp$).
-The context is an arbitrary program and thus has the role of an {\em attacker} that can interact with the component by means of any features the programming language has, and the component is what is security-relevant.
-%In this work, the semantics of the programming language is expected to differentiate between the component and the context.
-With this, \Thmref{def:propsat} can be extended as follows: for \bul{a component $\progvar$ to robustly satisfy a property $\pi$}, take an \iul{attacker context $\contextvar$ and link it with $\progvar$}, \oul{the resulting whole program must satisfy $\pi$}.
-
-\begin{definition}[Robust Satisfaction]\label{def:proprsat}
-  $\;$
-
-  \begin{nscenter}
-  \bul{$\rsat{\progvar}{\pi}$}
-  $\isdef$ \iul{$\forall \contextvar\ \wholeprogvar$, if $\link{\contextvar}{\progvar}=\wholeprogvar$}, then \oul{$\sat{\wholeprogvar}{\pi}$}.
-  \end{nscenter}
-\end{definition}
-
-% \begin{exampleenv}[Double Free in Bluetooth Subsystem]
-%   Consider \texttt{CVE-2021-3564}~\cite{doublefree-bluetooth}, one of many submissions for a double-free vulnerability.
-%   The vulnerability arises due to a race condition where the context-level function \texttt{hci\_cmd\_work} was not expected to behave maliciously, since it resides in the same source-code repository where the vulnerability occurs.
-%   Nevertheless, the component-level code of \texttt{hci\_dev\_do\_open} is linked with \texttt{hci\_cmd\_work} and does not atomically check whether a pointer has been freed already:
-%   Therefore, \texttt{hci\_dev\_do\_open} does not satisfy the no-double-frees property robustly, since there is an implementation for \texttt{hci\_cmd\_work} that leads to a violation of that property when linked with \texttt{hci\_dev\_do\_open}.
-% \end{exampleenv}
-
-\subsection{Secure Compilers}\label{subsec:bg:rtp}
-
-A {\em compiler} ($\cc{\src{L}}{\trg{L}}$) translates syntactic descriptions of programs from a {\em source} ($\src{L}$) into a {\em target} ($\trg{L}$) programming language.
-This translation is considered {\em correct} if it is semantics-preserving.
-That is, for a whole program $\src{\wholeprogvar}$, the compiler should relate the $\src{L}$ semantics of $\src{\wholeprogvar}$ with the semantics of $\trg{L}$ of the compiled counterpart of $\src{p}$ in such a way that they are ``compatible''.
-Unfortunately, correct compilers may be insecure compilers~\cite{patrignani2019survey,kennedy2006secure.net,abadi1999protect,ahmed2018dagstuhl} and programs translated by insecure compilers can violate security properties that the programmer assumes to hold.
-This is why {\em robust preservation} is a good candidate as a compiler-level security property~\cite{abate2019jour}.
-
-This paper uses a general notion of robust preservation~\cite{abate2021extacc} that considers compilers that use languages with different trace models. 
-To this end, considering a source trace $\src{\trace}$ and a target trace $\trg{\trace}$, there is a relation ($\sim$) describing the effect of a corresponding compiler (see \Cref{sec:casestud:rtp}). 
-This relation induces the following two projection functions~\cite{abate2021extacc}: (1) the \emph{existential image} $\tau_\sim\left(\src{\pi}\right)$ and (2) the \emph{universal image} $\sigma_\sim\left(\trg{\pi}\right)$.
-These projections map source-level (resp. target-level) properties to target-level (resp. source-level) properties in a way that identifies the ``same'' property across languages. 
-The case study of this paper uses the universal image, since some considered properties, such as \gls*{ss}, are not definable in a higher-level language that, e.g., does not model speculation.
-% For self-containment, the universal image is defined as follows~\cite{abate2021extacc}:
-% 
-\begin{definition}[Universal Image]
-% [Existential and Universal Image]
-\label{def:universal:img}\label{def:sigma}
-  % \[
-  %   \tau_\sim\left(\src{\pi}\right) := 
-  %     \left\{ 
-  %       \trg{\trace} \mid \exists \src{\trace}\ldotp \src{\trace}\sim\trg{\trace}, \text{ and } \src{\trace}\in\src{\pi} 
-  %     \right\}
-  % \]
-  \[ 
-    \sigma_\sim\left(\trg{\pi}\right) := 
-      \left\{ 
-        \src{\trace} \mid \forall \trg{\trace}\ldotp \text{if }\src{\trace}\sim\trg{\trace}, \text{ then } \trg{\trace}\in\trg{\pi} 
-      \right\}
-  \]
-\end{definition}
-\noindent
-With this projection function, we define a more general version of robust preservation as follows~\cite{abate2021extacc}.
-% For a \bul{compiler $\cc{\src{L}}{\trg{L}}$ to robustly preserve a class of source properties $\src{\class}$}, if for any \rul{property $\src{\pi}$ of that class $\src{\class}$ and programs $\src{p}$ written in $\src{L}$} where \iul{$\src{\progvar}$ robustly satisfies $\src{\pi}$}, then \oul{the compilation of $\src{\progvar}$, $\cc{\src{L}}{\trg{L}}\left(\src{p}\right)$, must robustly satisfy $\tau_\sim\left(\src{\pi}\right)$}.
-% 
-% \begin{definition}[Robust Preservation with $\tau_\sim$]\label{def:rtp:tau}
-%   % $\;$\\
-%   % \vspace{-1em}
-%   % \begin{nscenter}
-%   % Compiler $\cc{\src{L}}{\trg{L}}$ robustly preserves $\class$, 
-%   \bul{$\rtptau{\cc{\src{L}}{\trg{L}}}{\src{\class}}{\sim}$}
-%   %, iff 
-%   $\isdef$
-%   \rul{$\forall \src{\pi}\in\src{\class}, \src{p}\in\src{L},$} if \iul{$\rsat{\src{\progvar}}{\src{\pi}}$}, then \oul{$\rsat{\cc{\src{L}}{\trg{L}}\left(\src{p}\right)}{\tau_\sim{\left(\src{\pi}\right)}}$}.
-%   % \end{nscenter}
-% \end{definition}
-% 
-A \bul{compiler $\cc{\src{L}}{\trg{L}}$ robustly preserves a class of target properties $\trg{\class}$}, if for any \rul{property $\trg{\pi}$ of class $\trg{\class}$ and programs $\src{p}$}, where \iul{$\src{\progvar}$ robustly satisfies $\sigma_\sim\left(\trg{\pi}\right)$}, \oul{the compilation of $\src{\progvar}$, we have that $\cc{\src{L}}{\trg{L}}\left(\src{p}\right)$ robustly satisfies $\trg{\pi}$}.
-
-\begin{definition}[Robust Preservation with $\sigma_\sim$]\label{def:rtp:sigma}
-  % $\;$\\
-  % \vspace{-1em}
-  % \begin{nscenter}
-  % Compiler $\cc{\src{L}}{\trg{L}}$ robustly preserves $\class$, 
-  \bul{$\rtpsig{\cc{\src{L}}{\trg{L}}}{\trg{\class}}{\sim}$}
-  %, iff 
-  $\isdef$
-  \rul{$\forall \trg{\pi}\in\trg{\class}, \src{p}\in\src{L},$} if \iul{$\rsat{\src{\progvar}}{\sigma_\sim\left(\trg{\pi}\right)}$}, then \oul{$\rsat{\cc{\src{L}}{\trg{L}}\left(\src{p}\right)}{\trg{\pi}}$}.
-  % \end{nscenter}
-\end{definition}
-Note that a class of properties $\class$ can represent just one property $\pi$ by lifting~\cite{clarkson2008hyper} that property to sets of properties, i.e., use the powerset of $\pi$ instead of $\pi$ itself.
-Because of this, this paper may write $\rtptau{\cc{\src{L}}{\trg{L}}}{\trg{\pi}}{\sim}$, even though $\trg{\pi}$ is a property and not a class.
-A similar construction can be used to the projection functions (see \Cref{def:universal:img}) by applying them to the lifted version of $\trg{\pi}$ instead of just $\trg{\pi}$.
-
-In case the trace model is the same for both source and target programs (and thus $\sim$ is equality), we obtain~\cite{abate2019jour}:
-\begin{definition}[Robust Preservation]\label{def:rtp}
-  % $\;$\\
-  % \vspace{-1em}
-  % \begin{nscenter}
-  % Compiler $\cc{\src{L}}{\trg{L}}$ robustly preserves $\class$, 
-  $\;$
-
-  {$\rtp{\cc{\src{L}}{\trg{L}}}{\class}$}
-  %, iff 
-  $\isdef$
-  {$\forall \pi\in\class, \src{p}\in\src{L},$} if {$\rsat{\src{\progvar}}{\pi}$}, then {$\rsat{\cc{\src{L}}{\trg{L}}\left(\src{p}\right)}{\pi}$}.
-  % \end{nscenter}
-\end{definition}
-
-Examples of compilers fulfilling \Cref{def:rtp} exist in the literature~\cite{korashy2022secureptrs,korashy2021capableptrs,abate2021extacc,abate2019jour,patrignani2021rsc}.
-For example, SecurePtrs~\cite{korashy2022secureptrs} gives a compiler that robustly preserves all safety properties for a C-like language to an assembly-like language. 
-% The mechanization has well over 20k lines in total, which demonstrates that proving a compiler secure with respect to \Cref{def:rtp} is a significant effort. \MP{move this line to later ?}
-% i don't see how adding a def of cct is beneficial
-As another example, even though it is not strictly satisfying \Cref{def:rtp}, the FaCT~\cite{cauligi2019fact} compiler preserves the \gls*{cct} property for a C-like language with constant-time primitives, e.g., \texttt{ctselect} for branching.
-%, whose precise definition can be found in \Cref{sec:cct}.
-Throughout this work, it is assumed that FaCT satisfies \Cref{def:rtp}.
-% While we were not able to find compilers in literature that provably fulfill \Cref{def:rtp}, we anticipate that 
- % would satisfy this hyperproperty.
-% FaCT proves constant-timeness by means of overapproximation, i.e., instead of directly proving the hyperproperty, they fall back to a property that overapproximates the corresponding hyperproperty.
-
-%\smallskip%MP: the underline compacts the sections too much
-
-\section{Secure Composition}\label{sec:sequential}
-
-Notably, real-world compilation chains also perform a series of (sequential) passes whose main purpose is not necessarily to translate from one language to another, but to, e.g., optimise the code or enforce a certain property.
-Both examples can be seen in practice, e.g.,~\cite{nagarakatte2009soft,nagarakatte2010cets,akritidis2009baggy,wegman1991ccp,manjikian1997fusion} and many more.
-% \MPin{survey to cite?
-this is a bit informal MK: hard to find a survey on different compiler passes...
-I can't even find one just for compiler enforced memsafety (that looks good) or for compiler translations in general}.
-% 
-% Consider the compilation passes of LLVM~\cite{lattner2004llvm}, which are handled in the \texttt{PassManager} class.
-% By manipulating the internal state of a \texttt{PassManager} object, the schedule of compiler optimisation passes can be hardcoded.
-% That way, an order of optimisations can be found by hand that yields the best results.
-% \MP{what's the point of the previous 3 sentences?(commented)}
-Consider the following two LLVM optimisation passes: \gls*{cf}, which rewrites constant expressions to the constant they evaluate to, and \gls*{dce}, which removes dead code by rewriting conditional branches.
-The order in which \gls*{cf} and \gls*{dce} are performed influences the final result of the compilation (see \Cref{fig:cfdceex}).
-\begin{figure}[!ht]
-  \begin{center}
-    \begin{tikzpicture}[scale=0.62, every node/.style={transform shape}]
-      \node[minimum width=3cm,minimum height=2.5cm,draw=black,very thick,align=left] (progunopt) {\begin{lstlisting}[language=ml,basicstyle=\small\ttfamily]
-let a = true in
-if a then
-  print "a"
-else
-  print "b"
-\end{lstlisting}};
-      \node[minimum width=3cm,minimum height=2.5cm,draw=black,very thick,align=left,right=1.0 of progunopt] (progoptcf) {\begin{lstlisting}[language=ml,basicstyle=\large\ttfamily]
-if true then
-  print "a"
-else
-  print "b"
-\end{lstlisting}};
-      \node[minimum width=3cm,minimum height=2.5cm,draw=black,very thick,align=left,right=1.0 of progoptcf] (progoptcfdce) {\begin{lstlisting}[language=ml,basicstyle=\Large\ttfamily]
-print "a"
-\end{lstlisting}};% the following is a hack to get my syntax highlighting to work again \end{lstlisting}
-      % edges
-      \draw[->,very thick] (progunopt) edge[loop left,left] node {\gls*{dce}} (progunopt);
-      \draw[->,very thick] (progunopt) edge[sloped,above] node {\gls*{cf}} (progoptcf);
-      \draw[->,very thick] (progoptcf) edge[sloped,above] node {\gls*{dce}} (progoptcfdce);
-    \end{tikzpicture}
-  \end{center}
-  \caption{Example program where the level of optimisations differ for one pass of applying \gls*{cf} and \gls*{dce} in any order. %
-  Every edge is a compilation pass and the label on the edge states what the pass does, i.e., \gls*{cf} or \gls*{dce}. %
-  The source code in the nodes is a glorified compiler intermediate representation and the code gets more optimised towards the right hand side of the figure.}\label{fig:cfdceex}
-\end{figure}
-This {\em phase ordering problem} is well--known in literature and a practical solution is to simply perform a fixpoint iteration of the optimisation pipeline~\cite{click1995combining}.
-
-\smallskip
-
-To analyse the security of compilation passes and their interaction within a compilation pipeline, we rely on a few key notions: the definition of a trace relation, and the definition of when is a trace relation well-formed with respect to a class.
-%
-Consider traces $\src{\trace}$ and $\trg{\trace}$ as well as two trace relations $\sim_1$ and $\sim_2$. 
-% 
-The traces are related $\src{\trace}\sim_1\bullet\sim_2\trg{\trace}$ if there is another trace $\irl{\trace}$ such that $\src{\trace}\sim_1\irl{\trace}$ and $\irl{\trace}\sim_2\trg{\trace}$.
-% 
-\bul{A relation $\sim$ is well-formed w.r.t. a class of target-level properties $\trg{\class}$} iff \rul{the universal image preserves set membership}.
-\begin{definition}[Well-formedness of $\sim$ for a Class $\trg{\class}$]\label{def:wfc:sig:tracerel}
-  % $$
-
-  \begin{nscenter}
-  \noindent
-  \text{\bul{$\wfcsig{\sim}{\trg{\class}}$}} \isdef \text{\rul{$\forall \trg{\pi}\in\trg{\class}, \sigma_\sim(\trg{\pi})\in\sigma_\sim(\trg{\class})$}}
-  \end{nscenter}
-  % $$
-\end{definition}
-
-We can now state our main result: secure compilers in the robust compilation framework~\cite{abate2021extacc} compose {\em sequentially}. 
-%
-
-Let \bul{$\cc{\src{L}}{\trg{L}}$ robustly preserve the class $\sigma_{\sim_2}\left(\trg{\class[_1]}\right)$ under $\sim_1$} and let \rul{$\cc{\trg{L}}{\obj{L}}$ robustly preserve the class $\trg{\class[_2]}$ under $\sim_2$}.
-Then, \iul{when the cross-language relation $\sim_2$ is well-formed w.r.t.
-class $\trg{\class[_1]}$}, it follows that \oul{the composed compiler $\cc{\src{L}}{\trg{L}}\circ\cc{\trg{L}}{\obj{L}}$ robustly preserves the intersection of classes $\trg{\class[_1]}\cap\trg{\class[_2]}$ under $\sim_1\bullet\sim_2$}.
-% 
-%For space reasons, \Thmref{thm:urtp} and \Thmref{thm:lrtp} are not presented here, but can be found in the technical appendix ($\CoqSymbol$).
-% Note that it is not necessary to require $\sim_2$ to be well-formed w.r.t. $\src{\class[_1]}$, since the 
-% \MPin{why}
-\begin{theorem}[Composition of Secure Compilers w.r.t. $\sigma$]\label{thm:rtpsim:sig}
-  $\;$
-
-  If \bul{$\rtpsig{\cc{\src{L}}{\trg{L}}}{\tilde{\sigma}_{\sim_2}\left(\trg{\class[_{1}]}\right)}{\sim_1}$}, \rul{$\rtpsig{\cc{\trg{L}}{\obj{L}}}{\trg{\class[_2]}}{\sim_2}$}, and \iul{$\wfcsig{\sim_2}{\trg{\class[_1]}}$}, \\ then \oul{$\rtpsig{\cc{\src{L}}{\trg{L}}\circ\cc{\trg{L}}{\obj{L}}}{\trg{\class[_{1}]}\cap\trg{\class[_{2}]}}{\sim_1\bullet\sim_2}$}.
-\Coqed
-\end{theorem}
-
-Since the composition of secure compilers is again a secure compiler, the theorems generalise to a whole chain of $n$ secure compilers. 
-% 
-\Cref{thm:rtpsim:sig} can also be stated for the existential image $\tau_\sim\left(\src{\pi}\right)$, but in the interest of space that result has been moved to the appendix (\Cref{subsec:extimg}).
-% 
-Crucially, \Cref{thm:rtpsim:sig} also holds for \emph{classes of hyperproperties}, and thus, compilers that robustly preserve hyperproperties can be composed with each other as well as with compilers that robustly preserve properties.
-% 
-%Unfortunately, proofs of compilers that robustly preserve hyperproperties (including hypersafety) are scarce in the literature due to their complexity.
-%\MP{ so what ? }
-%  we don't discuss this further?
-%
-
-% It is furthermore possible to conflate the notion of classes and hyperproperties by intersecting all elements of a class of hyperproperties, to obtain one combined hyperproperty.
-% This idea has been discussed in other work~\cite{clarkson2008hyper}.
-%  MP: i think i stated the point of these 2 lines above
-
-If we take SecurePtrs and FaCT from \Cref{subsec:bg:rtp} and compose them according to \Cref{thm:rtpsim:sig}, we obtain a compiler that robustly preserves the intersection of safety properties and the \gls*{cct} hyperproperty.
-That is, for a source component that robustly satisfies any set of safety properties {\em and} \gls*{cct}, the compiled target component also robustly satisfies the same set of safety properties and \gls*{cct}.
-% The intersection simply uses the lifted version of the respective safety properties.
-%\MP{meaning?}
-
-Compiler engineers typically try to find an order of optimisations that yields well-optimised programs for either code size~\cite{cooper1999geneticphases} or performance~\cite{kulkarni2006exhaustivephase}.
-\Cref{corr:swappable} justifies that any such order of compilation passes is valid w.r.t. security, as long as the trace-relations have no effect on the respective classes.
-
-So, given two compilation passes \bul{$\cc[_{1}]{\trg{L}}{\trg{L}}$, $\cc[_{2}]{\trg{L}}{\trg{L}}$, both robustly preserving class $\trg{\class[_{1}]}$ or $\trg{\class[_{2}]}$, respectively}, \rul{their corresponding well-formed trace-relations}, \iul{and indifference of the classes with respect to these trace relations}, \oul{for any order of their composition, the composed compiler robustly preserves the intersection of $\trg{\class[_{1}]}$ and $\trg{\class[_{2}]}$}.
-% 
-% \MPin{
-% 	later on, say that optimisation passes remove elements from traces, so they should be fine w.r.t. the result with $\tau$.
-% 	also, yes, if $\tau$ adds we have a problem, if it removes we don't
-% \MKin{ Optimisations that remove elements from traces are not necessarily sound,
-%  e.g., remove all deallocations to "optimise" memory management by offloading
-%  the work to the OS.
-% }
-% }
-
-\begin{corollary}[Swapping Secure Compiler Passes]\label{corr:swappable}
-  $\;$
-
-  If \bul{$\rtpsig{\cc[_{1}]{\trg{L}}{\trg{L}}}{\trg{\class[_{1}]}}{\sim_1}$ and $\rtpsig{\cc[_{2}]{\trg{L}}{\trg{L}}}{\trg{\class[_{2}]}}{\sim_2}$}, %
-  \rul{$\wfcsig{\sim_1}{\trg{\class[_2]}}$ and $\wfcsig{\sim_2}{\trg{\class[_1]}}$}, %
-  and \iul{$\tilde{\sigma}_{\sim_2}\left(\trg{\class[_1]}\right)=\trg{\class[_1]}$ as well as $\tilde{\sigma}_{\sim_1}\left(\trg{\class[_2]}\right)=\trg{\class[_2]}$},
-  then \oul{$\rtpsig{\cc[_{1}]{\trg{L}}{\trg{L}}\circ\cc[_{2}]{\trg{L}}{\trg{L}}}{\trg{\class[_{1}]}\cap\trg{\class[_{2}]}}{\sim_1\circ\sim_2}$ and $\rtpsig{\cc[_{2}]{\trg{L}}{\trg{L}}\circ\cc[_{1}]{\trg{L}}{\trg{L}}}{\trg{\class[_{2}]}\cap\trg{\class[_{1}]}}{\sim_2\circ\sim_1}$}.
-\Coqed
-\end{corollary}
-
-Coming back to the example composing SecurePtrs with FaCT, it is likely the case that \Cref{corr:swappable} is not applicable.
-While first running SecurePtrs and then FaCT should be fine, the other direction has potential security hazards, since the SecurePtrs compiler does not account for cryptographic-constant time primitives, such as \texttt{ctselect}.
-
-\subsection{Secure Compiler Composition with Same Trace Models}
-When the cross-language trace relation is an equality, \Cref{thm:rtpsim:sig} collapses:
-Given \bul{$\cc{\src{L}}{\trg{L}}$ robustly preserves $\class[_{1}]$} and \rul{$\cc{\trg{L}}{\obj{L}}$ robustly preserves $\class[_{2}]$}, it follows that \oul{their sequential composition $\cc{\src{L}}{\trg{L}}\circ\cc{\trg{L}}{\obj{L}}$ robustly preserves the intersection of classes $\class[_{1}]$ and $\class[_{2}]$}.
-
-\begin{corollary}[Composition of Secure Compilers]\label{corr:rtp}
-  $\;$
-
-  If \bul{$\rtp{\cc{\src{L}}{\trg{L}}}{\class[_{1}]}$} and \rul{$\rtp{\cc{\trg{L}}{\obj{L}}}{\class[_{2}]}$}, then \oul{$\rtp{\cc{\src{L}}{\trg{L}}\circ\cc{\trg{L}}{\obj{L}}}{\class[_{1}]\cap\class[_{2}]}$}.
-\Coqed
-\end{corollary}
-
-\Cref{corr:rtp} provides an easy way to compose secure compilers without well-formedness of trace relations. 
-% 
-However, while \Cref{thm:rtpsim:sig} explicitly requires that $\sim_1$ is well-formed w.r.t. $\trg{\class[_2]}$, if $\sim_2$ is not well-formed w.r.t. $\trg{\class[_1]}$, care must be taken. 
-%Worst-case is that the resulting class $\sigma_{\sim_1\bullet\sim_2}\left(\trg{\class[_1]}\cap\trg{\class[_2]}\right)$ becomes empty, even though $\trg{\class[_1]}\cap\trg{\class[_2]}$ is not necessarily empty.
-This is further discussed in \Cref{subsec:compatsecpasses}.
-
-%If possible, optimisations are merged to enhance the analysis' result, since even when iterating to a fixpoint the final program may not be as optimised as it could be~\cite{click1995combining}.
-We can also obtain a specialised version of \Cref{corr:swappable}:
-%So, given two compilation passes \bul{$\cc[_{1}]{\trg{L}}{\trg{L}}$, $\cc[_{2}]{\trg{L}}{\trg{L}}$, both robustly preserving class $\class[_{1}]$ or $\class[_{2}]$, respectively}, \oul{for any order of their composition, the composed compiler robustly preserves the intersection of $\class[_{1}]$ and $\class[_{2}]$}.
-% 
-% \MPin{
-% 	later on, say that optimisation passes remove elements from traces, so they should be fine w.r.t. the result with $\tau$.
-% 	also, yes, if $\tau$ adds we have a problem, if it removes we don't
-% \MKin{ Optimisations that remove elements from traces are not necessarily sound,
-%  e.g., remove all deallocations to "optimise" memory management by offloading
-%  the work to the OS.
-% }
-% }
-
-\begin{corollary}[Swapping Secure Compiler Passes]\label{corr:swappable:one}
-  $\;$
-
-  If {$\rtp{\cc[_{1}]{\trg{L}}{\trg{L}}}{\class[_{1}]}$ and $\rtp{\cc[_{2}]{\trg{L}}{\trg{L}}}{\class[_{2}]}$}, then {$\rtp{\cc[_{1}]{\trg{L}}{\trg{L}}\circ\cc[_{2}]{\trg{L}}{\trg{L}}}{\class[_{1}]\cap\class[_{2}]}$ and $\rtp{\cc[_{2}]{\trg{L}}{\trg{L}}\circ\cc[_{1}]{\trg{L}}{\trg{L}}}{\class[_{2}]\cap\class[_{1}]}$}.
-\Coqed
-\end{corollary}
-
-\section{Security Properties Formalisation \& Composition\pages{1.5}}\label{sec:compprop}
-
-This section introduces trace properties of interest for this paper: \gls*{tms}, \gls*{sms}, \gls*{ms}, \gls*{scct}, and \gls*{ss}.
-These properties are of practical importance (as mentioned in \Cref{sec:introduction}) and also of interest in the case study presented later (\Cref{sec:casestud:defs,sec:casestud:rtp}). 
-This section presents all of them, despite the fact that they are inspired by existing work, in order to showcase all that is required for a formal proof of security for a realistic compilation toolchain.
-%\MP{ say something about this in the rebuttal: only the defs are borrowed, the universal images and all that work is new}
-%\MK{ universal images are not new. }
-% , none of the properties is novel, for self-containment of the paper, their definition is necessary for the case study.
-For each of the presented properties, the technical report defines corresponding monitors, which refine a given property and have a key role in the proofs of this paper. 
-% managed to prove 
-%  if tms-monitor --[trace]-->* ∅  
-%  then trace ∈ tms
-% but not
-%  if trace ∈ tms
-%  then tms-monitor --[trace]-->* ∅
-
-\subsection{A Trace Model for Memory Safety}\label{subsec:basic:memsafety:tracemodel}
-
-For simple memory safety composed of temporal and spatial memory safety, the trace model defines events ($\event[_{\mssafe}]$) as either the empty event ($\emptyevent$), a crash ($\lightning$), or a base-event ($\preevent[_{\mssafe}]$).
-
-\vspace{-1.0em}
-\[
-  \begin{array}{rcll}
-    \text{(Base-Events)}&\preevent[_{\mssafe}] &:=& \ev{Alloc \loc\ n} \mid \ev{Dealloc \loc} \mid \ev{Use \loc\ n} \\
-    \text{(Events)}&\event[_{\mssafe}] &:=& \preevent[_{\mssafe}] \mid \emptyevent \mid \lightning \\ 
-  \end{array}
-\]
-
-Base-events describe the actual kind of event that happened.
-For the basic memory-safety properties, these are three variants:
-First, the allocation event ($\ev{Alloc\ \loc\ n}$) that fires whenever a program claims $n$ cells of memory and stores them at address $\loc$, where addresses are assumed to be unique.
-Second, deallocation ($\ev{Dealloc\ \loc}$) announces that the object at location $\loc$ is freed.
-Third, an event to describe reads from and writes to the $n$-th memory cell from address $\loc$ ($\ev{Use\ \loc\ n}$).
-
-\subsubsection{Temporal Memory Safety}
-
-\gls*{tms}~\cite{nagarakatte2010cets} is a safety property that describes that an unallocated object must not be (re-)used.
-
-\begin{definition}[\glsfirst*{tms}]\label{def:trace:tmsdef}
-  % \footnotesize
-  $$
-  \tmssafe:=\left\{\trace[_{\mssafe}] \left| \begin{array}{rcl}
-    \ev{Alloc\ \loc\ n}&\le_{\trace[_{\mssafe}]}&\ev{Dealloc\ \loc} \\
-    \ev{Use\ \loc\ n}&\le_{\trace[_{\mssafe}]}&\ev{Dealloc\ \loc} \\
-    \text{at most one }\ev{Dealloc\ \loc}&\text{in}&\trace[_{\mssafe}] \\
-    \text{at most one }\ev{Alloc\ \loc\ n}&\text{in}&\trace[_{\mssafe}] \\
-  \end{array}\right.\right\}
-  $$
-\end{definition}
-Hereby, the notation $\event[_{1}]\le_{\trace}\event[_{2}]$ means that if $\event[_{1}]$ is in $\trace$ and if $\event[_{2}]$ is in $\trace$, then $\event[_{1}]$ appears before $\event[_{2}]$.
-
-\subsubsection{Spatial Memory Safety}
-
-\gls*{sms}~\cite{nagarakatte2009soft} is a safety property that prohibits out-of-bounds accesses.
-
-\begin{definition}[\glsfirst*{sms}]\label{def:trace:smsdef}
-% \small
-  % \begin{nscenter}
-
-  \noindent
-  \[
-  \smssafe:=\left\{\trace[_{\mssafe}] \left|\begin{array}{rcl}
-      \text{If }\ev{Alloc\ \loc\ n}\le_{\trace_{\mssafe}}\ev{Use\ \loc\ m}, \text{ then }m<n
-  \end{array}\right.\right\}
-  \]
-  % \end{nscenter}
-\end{definition}
-
-\subsubsection{Memory Safety}
-
-In spirit of earlier work~\cite{nagarakatte2009soft,nagarakatte2010cets,jim2002cyclone,necula2005ccured,michael2023mswasm}, full \gls*{ms} is the intersection of \Cref{def:trace:tmsdef,def:trace:smsdef}.
-
-\begin{definition}[\glsfirst*{ms}]\label{def:trace:msdef}
-  $
-  \mssafe:=\tmssafe \cap \smssafe
-  $
-\end{definition}
-
-Note that \Cref{def:trace:msdef} ignores data isolation, so there may still be memory-safety issues introduced by side-channels.
-
-\subsection{A Trace Model for Memory Safety with Constant Time}\label{subsec:scct:tracemodel}
-
-To express Constant Time, we extend the memory safety trace model with a {\em security tag} ($\securitytag{}$) that indicates whether events contains sensitive information ($\lock$) or not ($\unlock$).
-
-\vspace{-.5em}
-\[
-  \begin{array}{rrcl}
-    (\text{Base-Events}) & \preevent[_{\ctsafe}] &:=& \preevent[_{\mssafe}] \mid \ev{Branch\ }n \mid \ev{Binop\ }n\\
-    (\text{Security Tags}) & \securitytag{} &:=& \lock \mid \unlock\\ 
-    (\text{Events}) & \event[_{\ctsafe}] &:=& \preevent[_{\ctsafe}];\securitytag{} \mid \emptyevent \mid \lightning \\ 
-  \end{array}
-\]
-
-For cryptographic code, there is a general guideline that secrets must not be visible on a trace~\cite{ctguidelines}, i.e., secrets should not be marked as $\unlock$.
-In turn, an instruction whose timing is data-dependent must not have a secret as an operand.
-Typical operations with data-dependent timing are branches and certain binary operations, such as division.%
-\footnote{
-	This is architecture-dependent, but division is an operation that serves as a classic example for a data-dependent timing instruction~\cite[p.~755]{arm-refman}.
-}
-Both operations are represented in the trace model by extending the set of base-events with branches ($\ev{Branch\ n}$) and binary operations ($\ev{Binop\ n}$).
-
-\subsubsection{Strict Cryptographic Constant Time}
-
-\gls*{cct} is a hypersafety property~\cite{barthe2018sec} and, thus, difficult to check with monitors.
-This is because, intuitively, hypersafety properties can relate multiple execution traces with each other, but monitors work on a single execution.
-It is a common trick to sidestep this issue by means of overapproximation: this section defines the property \gls*{scct}, a stricter variant of \gls*{cct} (inspired by earlier work~\cite{almeida2017jasmin}) that enforces the policy that no secret appears on a trace.
-Programs that satisfy \gls*{scct} also satisfy \gls*{cct}, but programs that satisfy \gls*{cct} may not satisfy \gls*{scct}.
-
-\begin{definition}[\glsfirst*{scct}]\label{def:trace:scctdef}
-% \begin{nscenter}
-
-  \noindent\[
-  \scctsafe:=\left\{\trace[_{\ctsafe}] \left|\begin{array}{l}
-      \trace[_{\ctsafe}]=\hole{\cdot} \text{ or } \\\exists\trace[_{\ctsafe}'],\trace[_{\ctsafe}]=\preevent[_{\ctsafe}];\unlock\cdot\trace[_{\ctsafe}'] \wedge \trace[_{\ctsafe}']\in\scctsafe
-    \end{array}\right.\right\}
-  \]
-% \end{nscenter}
-\end{definition}
-
-\gls*{scct} may appear overly strict, since it seems that secrets must not occur on a trace (since $\securitytag{}$ is forced to be $\unlock$). 
-However, this is considered standard practice in terms of coding guidelines~\cite{ctguidelines}.
-Moreover, programs that have been compiled with FaCT~\cite{cauligi2019fact} and run with a ``data independent timing mode''~\cite{arm-refman,intel-refman} enabled do not leak secrets (see \Cref{ex:lscct}).
-
-\subsubsection{\gls*{ms}, Strict Cryptographic Constant Time}\label{sec:msscct-rel}
-
-The combination of \gls*{ms} and \gls*{scct} is the intersection of these properties, \gls*{msscct}.
-However, \gls*{ms} uses a different trace model than \gls*{scct}, so intersecting them would trivially yield the empty set. 
-To remedy this issue, we introduce $\sim_{\ctsafe}: \preevent[_{\ctsafe}] \times \preevent[_{\mssafe}] $, a cross-language trace relation (whose key cases are presented below), that we use to intuitively unify the trace model in which the two properties are expressed:
-\[
-  \typerulenolabel{mscctrel:drop:tag}{}{\preevent[_{\mssafe}]\sim_{\ctsafe}\preevent[_{\mssafe}];\unlock{}}
-  \typerulenolabel{mscctrel:drop:crash}{}{\lightning\sim_{\ctsafe}\lightning}
-\]
-\[
-  \typerulenolabel{mscctrel:drop:branch}{}{\emptyevent\sim_{\ctsafe}\ev{Branch n};\securitytag{}}
-  \typerulenolabel{mscctrel:drop:binop}{}{\emptyevent\sim_{\ctsafe}\ev{Binop n};\securitytag{}}
-\]
-
-Essentially, $\sim_{\ctsafe}$ ignores both the new $\ev{Branch\ n}$ and $\ev{Binop\ n}$ base-events as it relates security-insensitive actions ($\unlock$) to their equivalent counterparts.
-Thus, \gls*{ms} traces trivially satisfy \gls*{scct}.
-% 
-The above relation is extended point-wise to traces, skipping the empty event $\emptyevent$ on either side, and it is now possible to define \gls*{msscct} using the universal image:
-
-\begin{definition}[\gls*{ms} and \gls*{scct}]\label{def:trace:msscctdef}
-  $
-  \msscctsafe:=\mssafe\cap\sigma_{\sim_{\ctsafe}}\left(\scctsafe\right)
-  $
-\end{definition}
-
-\subsection{Extending the Trace Model with Speculation}\label{subsec:msctss:tracemodel}
-
-So far, the considered trace models do not let us express speculative execution attacks such as Spectre~\cite{kocher2019spectre}. 
-For this, we extend the earlier trace model (see \Cref{subsec:scct:tracemodel}) so that the security tags ($\securitytag{}$) carry additional information about the kind of private data leakage, i.e., the type of speculative leak.
-Moreover, we add base-events signalling the beginning of a speculative execution ($\ev{Spec}$), a barrier ($\ev{Barrier}$) that signals that any speculative execution may not go past it, as well as a rollback event ($\ev{Rlb}$), which signals that execution resumes to where speculation started.
-
-\vspace{-1em}
-{
-\[
-  \begin{array}{rrcl}
-    (\text{Base-Events}) & \preevent[_{\mathghost}] &:=& \preevent[_{\ctsafe}] \\
-    (\text{Spectre Variants}) & vX &:=& \operatorname{NONE} \mid \operatorname{PHT} \\
-    (\text{Security Tags}) & \securitytag{} &:=& \lock_{vX} \mid \unlock\\ 
-    (\text{Events}) & \event[_{\mathghost}] &:=& \preevent[_{\mathghost}];\securitytag{} \mid \emptyevent \mid \lightning \mid \ev{Spec} \mid \ev{Rlb} \mid \ev{Barrier}\\ 
-  \end{array}
-\]
-}
-
-Even though the considered Spectre variants are just SPECTRE-PHT~\cite{kocher2019spectre}, NONE just describes secret data as in \gls*{scct} (see \Cref{subsec:scct:tracemodel}), the trace model is general enough to allow for potential future extension with different variants~\cite{kocher2019spectre,maisuradze2018ret2spec,horn2019zero}.
-%For sake of readability, this paper just uses the notation $\lock$ in place of $\lock_{\text{PHT}}$.
-
-\subsubsection{Speculative Safety}
-
-\gls*{ss}~\cite{patrignani2021exorcising}, similar to \gls*{scct}, is a sound overapproximation of a variant of noninterference.
-
-\begin{definition}[\glsfirst*{ss}]\label{def:trace:ss}
-  \noindent
-
-  \begin{nscenter}
-  $
-    \sssafe:=\left\{\trace[_{\mathghost}] \left|\begin{array}{l}
-      \trace[_{\mathghost}]=\hole{\cdot} \text{ or } \exists\trace[_{\mathghost}'].\\
-      \left(\trace[_{\mathghost}]=\preevent[_{\mathghost}];\unlock\cdot\trace[_{\mathghost}'] \text{or }\trace[_{\mathghost}]=\preevent[_{\mathghost}];\lock_{\text{NONE}}\cdot\trace[_{\mathghost}']\right)\\
-      \text{and }\ \trace[_{\mathghost}']\in\sssafe
-                                 \end{array}\right.\right\}
-  $ 
-  \end{nscenter}
-\end{definition}
-The technical setup so far leads to the above definition, where only locks annotated with $\text{SPECTRE-PHT}$ are disallowed to occur on the trace.
-That way, programs attaining \gls*{ss} do not necessarily attain \gls*{scct}.
-
-\subsubsection{Speculation Memory Safety}\label{sec:spec-ms-rel}
-
-As before, we need to relate the different trace models with each other, so that the memory safety property without speculation can be lifted to speculation. 
-To this end, let $\sim_{\mathghost}: \preevent[_{\mathghost}] \times \preevent[_{\ctsafe}]$ be a cross-language trace relation whose key cases are below.
-The intuition is that \gls*{ss} is trivially satisfied in \gls*{scct}, since speculation is inexpressible there, which amounts to dropping events $\ev{Spec}$, $\ev{Rlb}$, or $\ev{Barrier}$, as well as all base events tagged with $\lock_{\text{PHT}}$.
-
-\[
-  \typerulenolabel{specrel:drop:tag}{}{\preevent[_{\ctsafe}];\lock\sim_{\mathghost}\preevent[_{\ctsafe}];\lock_{\text{NONE}}}
-  \typerulenolabel{specrel:drop:spectag}{}{\emptyevent\sim_{\mathghost}\preevent[_{\mathghost}];\lock_{\text{PHT}}}
-\]
-\[
-  \typerulenolabel{specrel:drop:pub}{}{\preevent[_{\ctsafe}];\unlock\sim_{\mathghost}\preevent[_{\ctsafe}];\unlock}
-  \typerulenolabel{specrel:drop:crash}{}{\lightning\sim_{\mathghost}\lightning}
-\]
-\[
-  \typerulenolabel{specrel:drop:spec}{}{\emptyevent\sim_{\mathghost}\ev{Spec}}
-  \typerulenolabel{specrel:drop:rlb}{}{\emptyevent\sim_{\mathghost}\ev{Rlb}}
-  \typerulenolabel{specrel:drop:barrier}{}{\emptyevent\sim_{\mathghost}\ev{Barrier}}
-\]
-
-We conclude by defining the ultimate property of interest for secure compilers: \gls*{specms}.
-\begin{definition}[\gls*{specms}]\label{def:trace:specmsdef}
-  $
-  \specmssafe := \msscctsafe\cap\sigma_{\sim_{\mathghost}}\left(\sssafe\right)
-  $
-\end{definition}
-
 \section{Case Study: Language Formalisations\pages{2}}\label{sec:casestud:defs}
@@
 Lastly, $\ird{Rlb}$ is emitted when the speculation window is zero and the operational state is rolled back (see \Cref{tr:e-spec-eaten}).
-
-\section{Case Study: Composing Secure Compiler Passes and Optimisations \pages{7}}\label{sec:casestud:rtp}
-This section defines several secure compilers, each of which robustly preserves a different property of interest as depicted in \Cref{fig:pipeline}.
-\begin{figure*}[!h]
-  \centering
-  \begin{tikzpicture}
-    \node (S) {$\src{L_{\tmssafe}}$};
-    \node[right=2.0 of S] (T) {$\trg{L}$};
-    \node[right=2.0 of T] (M) {$\irl{L_{\mssafe}}$};
-    \node[below right=1.5 and 1.0 of M] (D0) {$\irl{L_{\mssafe}}$};
-    \node[above right=1.5 and 1.0 of M] (C0) {$\irl{L_{\mssafe}}$};
-    \node[right=3.0 of M] (E) {$\irl{L_{\mssafe}}$};
-    \node[right=2.0 of E] (O) {$\obj{L_{\scctsafe}}$};
-    \node[right=2.0 of O] (Z) {$\ird{L_{\mathghost}}$};
-
-    \draw[->] (S) to[sloped] node[align=center] (tmsedge) {\gls*{tms}\\ \Cref{thm:cca:rtp:tms}} (T);
-    \draw[->] (T) to[sloped] node[align=center] {\gls*{sms}\\ \Cref{thm:ccb:rtp:sms}} (M);
-    \draw[->] (M) to[sloped] node[align=center] {\gls*{dce}\\ \Cref{thm:ccdce:rtp:ms}} (D0);
-    \draw[->] (M) to[sloped] node[align=center] {\gls*{cf}\\ \Cref{thm:cccf:rtp:ms}} (C0);
-    \draw[->] (D0) to[sloped] node[align=center] {\gls*{cf}\\ \Cref{thm:cccf:rtp:ms}} (E);
-    \draw[->] (C0) to[sloped] node[align=center] {\gls*{dce}\\ \Cref{thm:ccdce:rtp:ms}} (E);
-    \draw[->] (E) to[sloped] node[align=center] {\gls*{scct}\\ \Cref{thm:ccscct:rtp:scct}} (O);
-    \draw[->] (O) to[sloped] node[align=center] {\gls*{ss}\\ \Cref{thm:ccspec:rtp:spec}} (Z);
-
-    % Sections
-    \node (sectms) at ($(S)+(1.25,2.5)$) {{\Cref{subsec:cs:tms}}};
-    \node (secsms) at ($(T)+(1.25,2.5)$) {{\Cref{subsec:cs:ms}}};
-    \node (secopts) at ($(M)+(2.0,2.5)$) {{\Cref{subsec:cs:opts}}};
-    \node (secscct) at ($(E)+(1.5,2.5)$) {{\Cref{subsec:cs:scct}}};
-    \node (secss) at ($(O)+(1.5,2.5)$) {{\Cref{subsec:cs:ss}}};
-
-    \draw[thick,dotted,->] ($(S) + (0.2,-0.27)$) to ($(S) - (-0.2,1)$) to node[align=center] {\gls*{ms}\\ \Cref{corr:ccab:rtp:ms}} ($(M) - (0.0,1)$) to (M);
-
-    \draw[thick,dotted,->] (S) to ($(S) - (0.0,2.5)$) to node[align=center,pos=.65] {\gls*{specms}\\ \Cref{thm:ccall:rtp:specms}} ($(Z) - (0.0,2.5)$) to (Z);
-
-    % \draw[thick,dotted,->] (S) to[bend right=33,sloped] node[align=center] {\gls*{ms}\\ \Cref{corr:ccab:rtp:ms}} (M);
-    \draw[thick,dotted,->] (M) to[bend right=0,sloped] node[align=center] {\gls*{ms}\\ \Cref{thm:cccfccdce:rtp:ms}} (E);
-%    \draw[thick,dotted,->] (S) to[out=-45,in=198,sloped] node[align=center] {\gls*{ms}+\gls*{scct}\\ \Cref{thm:ccall:rtp:msscct}} (O);
-    % \draw[thick,dotted,->] (S) to[out=-40,in=210,sloped] node[align=center,pos=.8] {\gls*{specms}\\ \Cref{thm:ccall:rtp:specms}} (Z);
-
-    \draw[dashed] ($(S)-(0.0,-0.5)$) -- ($(S)-(0.0,-2.5)$);
-    \draw[dashed] ($(T)-(0.0,-0.5)$) -- ($(T)-(0.0,-2.5)$);
-    \draw[dashed] ($(M)-(0.0,-0.5)$) -- ($(M)-(0.0,-2.5)$);
-    \draw[dashed] ($(E)-(0.0,-0.5)$) -- ($(E)-(0.0,-2.5)$);
-    \draw[dashed] ($(O)-(0.0,-0.5)$) -- ($(O)-(0.0,-2.5)$);
-    \draw[dashed] ($(Z)-(0.0,-0.5)$) -- ($(Z)-(0.0,-2.5)$);
-  \end{tikzpicture}
-  \vspace{-1em}
-  \caption{Visualisation of the optimising compilation pipeline that preserves \gls*{specms}. %
-    Vertices in the graph are the programming languages from earlier sections (\Cref{sec:casestud:defs}). %
-    Full edges are secure compilers passes.
-    Dotted edges are composition of passes and use the presented framework (\Cref{sec:sequential}) to indicate the property they preserve. %
-    The dashed lines partition the graph into the sections where the respective theorems are presented.
-  }\label{fig:pipeline}
-\end{figure*}
-The section demonstrates the power of the framework (\Cref{sec:sequential}) by composing these compilers for a secure and optimising compilation chain that robustly preserves \gls*{specms}.
-The first step in this chain is the compiler from $\src{L_{\tmssafe}}$ to $\trg{L}$ that robustly preserves just \gls*{tms} (\Cref{thm:cca:rtp:tms}).
-From here, another pass from $\trg{L}$ to $\irl{L_{\mssafe}}$ ensures that no out-of-bounds accesses can happen and, thus, programs at this point attain \gls*{sms} (\Cref{thm:ccb:rtp:sms}).
-Since these properties compose into \gls*{ms}, composing these passes yields a compiler that robustly preserves \gls*{ms} (\Cref{corr:ccab:rtp:ms}).
-Then, the section presents two optimisation passes, namely \gls*{cf} and \gls*{dce}, each of which robustly preserves \gls*{ms} (\Cref{thm:ccdce:rtp:ms,thm:cccf:rtp:ms}).
-These passes can be freely ordered in the compilation chain without compromising memory safety (\Cref{thm:cccfccdce:rtp:ms}).
-The next step in the chain ensures that code stays \gls*{scct} (\Cref{thm:ccscct:rtp:scct}) when compiled from $\Lms$ to $\Lscct$, which is done by switching on a constant-tame mode for the computation.
-Lastly, by introducing barriers immediately after branches, speculative leaks via SPECTRE-PHT are prevented when compiling $\Lscct$ to $\Lspec$.
-The final result is that the whole compilation chain robustly preserves \gls*{specms} (\Cref{thm:ccall:rtp:specms}).
-
-\subsection{Robust Temporal Memory Safety Preservation}\label{subsec:cs:tms}
-
-The secure compiler from $\Ltms$ to $\Ltrg$ needs to ensure that when execution switches from context to component, the type signatures are respected.
-Thus, the compiler inserts the following dynamic typechecks before entering the body of a component-defined function (anything elided is a trivial identity function from source to target):
-
-\vspace{-1em}
-{
-\begin{gather}
-  \begin{align*}%{rcl}
-    %\cca(\src{x}) = &\ \trg{x} 
-    % \\
-  	%&
-  	%\qquad
-    %\cca(\src{\lbinop{\expr[_{1}]}{\expr[_{2}]}}) = &\ \lbinop[\trg]{\left[\cca(\src{\expr[_{1}]})\right]}{\left[\cca(\src{\expr[_{2}]})\right]} \\
-    %\cca(\src{n}) = &\ \trg{n} 
-    % \\
-    %&
-    \cca(\src{\lget{x}{\expr}}) = &\ \lget[\trg]{\trg{x}}{\left[\cca(\src{\expr})\right]} \\
-    \cca(\src{\ldelete{x}}) = &\ \ldelete[\trg]{\left[\cca(\src{x})\right]} 
-    \\
-    \cca(\src{\lfunction{g}{x:\natt\to\type[_{e}]}{\expr}})  =&
-  \end{align*}
-  \\
-  \begin{align*}
-%    \cca(\src{\lset{x}{\expr[_{1}]}{\expr[_{2}]}}) & = \lset[\trg]{\left[\cca(\src{x})\right]}{\left[\cca(\src{\expr[_{1}]})\right]}{\left[\cca(\src{\expr_{2}})\right]} \\
-% \\ 
-\lfunction[\trg]{\trg{g}}{\trg{x}}{\lifz[\trg]{\trg{\lhast{x}{\natt}}}{
-                                                            \left[\cca(\src{\expr})\right] %
-                                                                                                 }{\labort[\trg]}}
-  \end{align*}
-\end{gather}
-}
-
-\noindent Since $\trg{L}$ has no static typing, an attacker $\trg{\library_{\ctx}}$ can invoke a component function accepting a $\src{\natt}$ with $\trg{\lpair{17}{29}}$.
-With the dynamic check, the compiler ensures that execution aborts in such cases.
-%The compiler does not insert other checks and proceeds as the identity function (which in this paper amounts to a simple re-colouring of $\src{L_{\tmssafe}}$ to $\trg{L}$ expressions).
-
-Compiling the \texttt{strncpy} function from \Cref{sec:introduction} with $\cca$, the compiler would in this case ensure that the arguments that are evaluated in the compiled \texttt{strncpy} are valid.
-%This effectively has no influence on the resulting trace, so a corresponding cross-language trace relation $\sim^{\Ltms}_{\Ltrg}$ is just equality.
-
-$\cca$ is robustly preserving (\Cref{def:rtp}) \gls*{tms}:
-\begin{theorem}[$\cca$ secure w.r.t. \gls*{tms}]\label{thm:cca:rtp:tms}
-    $\rtp{\cca}{\tmssafe}$ % \Coqed
-\end{theorem}
-
-\subsection{Robust Spatial Memory Safety Preservation}\label{subsec:cs:ms}
-
-The spatial memory safety preserving compiler from $\Ltrg$ to $\Lms$ only inserts bounds-checks whenever reading from or writing to memory in order to enforce \gls*{sms}.
-These bounds checks need the bounds information, which the compiler keeps around by introducing a fresh identifier $\irl{x_{SIZE}}$ for each allocation that binds $\irl{x}$.
-Then, it is simply a matter of referring to that variable and ensuring that memory accesses are in the interval $[\irl{0},\irl{x_{SIZE}})$.
-When the check fails, the code aborts.
-
-\begin{nscenter}
-% \small
-  $$
-  \begin{array}{rcl}
-    \ccb(\trg{\lnew{x}{\expr[_{1}]}{\expr[_{2}]}}) & = 
-                                                   & \llet[\irl]{\irl{x_{SIZE}}}{\ccb(\trg{\expr[_{1}]})}{}
-    		\\&&
-    		\lnew[\irl]{\irl{x}}{\irl{x_{SIZE}}}{\ccb(\trg{\expr[_{2}]})}
-    		 \\
-  \ccb(\trg{\lget{x}{\expr}}) & = 
-                              & \llet[\irl]{\irl{x_{n}}}{\ccb(\trg{\expr})}{}
-  	\\&&
-  \lifz[\irl]{\irl{0\le x_{n}<x_{SIZE}}}{\\&&\irl{\lget{x}{x_{n}}}}{}
-  		\irl{\labort}
-  	  \\
-  \ccb(\trg{\lset{x}{\expr[_{1}]}{\expr[_{2}]}}) & = 
-                                                 & \llet[\irl]{\irl{x_{n}}}{\ccb(\trg{\expr[_{1}]})}{}
-  		\\&&
-  \lifz[\irl]{\irl{0\le x_{n}<x_{SIZE}}}{\\&&\lset[\irl]{\irl{x}}{\irl{x_{n}}}{}
-  		\ccb(\trg{\expr[_{2}]})
-  		}{\irl{\labort}} 
-  		% \\
-  \end{array}
-  $$
-\end{nscenter}
-
-\begin{exampleenv}[Instrumented \texttt{strncpy}]
-  Consider again \texttt{strncpy}, but instrumented for \gls*{sms}:
-    \begin{lstlisting}[language=c,basicstyle=\ttfamily, morekeywords={size_t}]
-void strncpy(size_t n, size_t dst_size, char *dst,
-             size_t src_size, char *src) {
-  for(size_t i = 0; i < src_size
-      && src[i] != '\0' && i < n; ++i) {
-    if(i < src_size && i < dst_size) {
-      dst[i] = src[i];
-    }
-  }
-}
-    \end{lstlisting}
-Consider context \texttt{strncpy(2, x, y)}, where \texttt{x} and \texttt{y} are pointers to valid regions of memory with allocated space for exactly two cells and do not contain the null-terminating character \texttt{'\textbackslash 0'}.
-% 
-Without the \gls*{sms} pass, the event $\ev{Use}\ \loc_{x}\ 2;\comp;\unlock$ would appear on the trace, but that indicates an out-of-bounds access! 
-Fortunately, with \gls*{sms} mitigation in place, that event does not appear during execution, since the bounds check prevents the condition \texttt{src[i] != '\textbackslash 0'} from executing.
-\end{exampleenv}
-
-Contrary to the previous compiler, $\ccb$ may change the trace of the original $\Ltrg$ program: if there is a memory access, it needs to be protected with a bounds check.
-% 
-The corresponding relation $\sim^{\Ltrg}_{\Lms} : \trg{\trace}\times\irl{\trace}$ that describes this semantic effect of the compiler is defined partially below.
-% 
-We omit action $\trg{Set}$, which is treated analogously, and any other event, which is related to its cross-language equivalent.
-% \footnote{Up to colours.}
-% \footnote{In the interest of space, the full details are in the technical report.}
-
-{
-\[
-  \typerulenolabel{xrel:sms:read}{\trg{n}\text{ in bounds}}{\trg{Get\ \loc\ n;\comp}\sim^{\Ltrg}_{\Lms}\irl{Get\ \loc\ n;\comp}}
-\]
-% \[
-%   \typerulenolabel{xrel:sms:panic}{\trg{n}\text{ out of bounds}}{\trg{Get\ \loc\ n;\comp}\sim^{\Ltrg}_{\Lms}\irl{\lightning}}
-% \]
-}
-
-% Note that neither $\Ltrg$ nor $\Lms$ expose branches on the trace.\MP{so what?}
-For simplicity, we elide the environment that this relation carries around in order to bind each location to its metadata (such as its size), and resolve the ``$\trg{n}$ in/out of bounds'' premise.
-% The bounds check mentioned in the premises of the rules above is left abstract for the sake of simplicity, the interested reader can find
-% since the precise definition needs some technicalities that only inhibit readability without any technical insights.
-% 
-We can now prove that compiler $\ccb$ robustly preserves \gls*{sms}.
-\begin{theorem}[$\ccb$ secure w.r.t. \gls*{sms}]\label{thm:ccb:rtp:sms}
-  $\rtpsim{\ccb}{\smssafe}{\sim^{\Ltrg}_{\Lms}}$ % \Coqed
-\end{theorem}
-
-At this point we can compose $\ccb$ with the previous compiler ($\cca$), but in order to do so, we need a trace relation from $\Ltms$ to $\Lms$.
-% 
-We can obtain this relation by composing the trace relation we just defined ($\sim^{\Ltrg}_{\Lms}$) with the one used by the previous compiler: $\sim^{\Ltms}_{\Ltrg} : \src{\trace}\times\trg{\trace}$.
-% 
-The latter has not been previously defined (nor has it been used in the related theorem) because that is just an equality relation, since the trace models of $\Ltms$ and of $\Ltrg$ are the same.
-% 
-Thus, we formally define $\sim^{\Ltms}_{\Lms}: \src{\trace} \times \irl{\trace}$ as the following composition: $\sim^{\Ltms}_{\Ltrg}\bullet\sim^{\Ltrg}_{\Lms}$.
-% 
-With this relation, \Cref{corr:ccab:rtp:ms} states that the composition of $\cca$ and $\ccb$ is secure w.r.t.
-\gls*{ms} and it follows from \Cref{thm:cca:rtp:tms,thm:ccb:rtp:sms} using \Cref{thm:rtpsim:sig}.
-% 
-\begin{corollary}[$\cca\circ\ccb$ secure w.r.t. \gls*{ms}]\label{corr:ccab:rtp:ms}
-  $\;$
-
-  \begin{nscenter}
-    $\rtpsim{\cca\circ\ccb}{\mssafe}{\sim^{\Ltms}_{\Lms}}$ % \Coqed
-  \end{nscenter}
-\end{corollary}
-% 
-This proof requires another precondition besides \Cref{thm:ccb:rtp:sms,thm:cca:rtp:tms}: $\sim^{\Ltms}_{\Ltrg}$ needs to be well-formed with respect to $\smssafe$.
-This follows trivially since $\sim^{\Ltms}_{\Ltrg}$ is an equality. 
-% 
-\begin{lemma}[$\sim^{\Ltms}_{\Ltrg}$ well-formed w.r.t. $\smssafe$]\label{lem:wf:ltmsltrg}
-$\;$
-
-  \begin{nscenter}
-  $\wfcsig{\sim^{\Ltms}_{\Ltrg}}{\smssafe}$
-  \end{nscenter}
-\end{lemma}
-
-\subsection{Optimising Compilers}\label{subsec:cs:opts}
-
-This section defines two optimising compiler passes from $\Lms$ to $\Lms$ which perform \gls*{dce} and \gls*{cf}, respectively.
-The \gls*{dce} pass applies a na\"ive rewrite rule on conditionals.
-The \gls*{cf} pass relies on an auxiliary function \texttt{mix} that uses a substitutions accumulator $\irl{\substlist}$ in order to rewrite constant binary operations, e.g., $\irl{{17}-1}$ to $\irl{16}$, and replace variables that are assigned to constants, e.g., $\irl{\llet{x}{7}{x}}$ to $\irl{7}$.
-
-\vspace{-1em}
-\begin{gather*}
-  \begin{align*}
-    \ccdce(\irl{\lifz{true}{\expr[_{1}]}{\expr[_{2}]}}) & = \ccdce(\irl{\expr[_{1}]}) &\\
-    \ccdce(\irl{\lifz{false}{\expr[_{1}]}{\expr[_{2}]}}) & = \ccdce(\irl{\expr[_{2}]}) &
-  \end{align*}
-  \\
-  % \begin{align*}
-  %   \ccdce(\irl{\lbinop{\expr[_{1}]}{\expr[_{2}]}}) & = \lbinop[\irl]{\ccdce(\irl{\expr[_{1}]})}{\ccdce(\irl{\expr[_{2}]})} &
-  % \end{align*}
-  % \\[0.25cm]
-  \begin{align*}
-    \cccf(\irl{\expr}) & = \partialeval{\irl{\expr}}{\irl{\hole{\cdot}}} &
-  \end{align*}
-  \\[0.125cm]
-  \begin{align*}
-   \partialeval{\irl{x}}{\irl{\substlist}} & = \irl{n} 
-   	\qquad\qquad \text{if } \irl{\subst{n}{x}}\in\irl{\substlist} \\
-   \partialeval{\irl{x}}{\irl{\substlist}} & = \irl{x} 
-   \qquad\qquad \text{if } \irl{\subst{n}{x}}\notin\irl{\substlist} \\
-   \partialeval{\irl{\lbinop{n}{m}}}{\irl{\substlist}} & = \irl{k} 
-   \qquad\qquad \text{if } \lbinop{\irl{n}}{\irl{m}}=k \\
-   %\partialeval{\irl{\lbinop{\expr[_{1}]}{\expr[_{2}]}}}{\irl{\substlist}} & = \lbinop[\irl]{\partialeval{\irl{\expr[_{1}]}}{\irl{\substlist}}}{\partialeval{\irl{\expr[_{2}]}}{\irl{\substlist}}} \\
-   \partialeval{\irl{\llet{x}{n}{\expr}}}{\irl{\substlist}} & = \partialeval{\irl{\expr}}{\irl{\subst{n}{x}\cdot\substlist}} 
-  % \\
-  % \partialeval{\irl{\lget{x}{\expr}}}{\irl{\substlist}} & = \lget[\irl]{\irl{x}}{\partialeval{\irl{\expr}}{\irl{\substlist}}}
-% \end{align*}
-\end{align*}
-% \\
-% \begin{align*}
-%   \partialeval{\irl{\llet{x}{\expr[_{1}]}{\expr[_{2}]}}}{\irl{\substlist}} & = \llet[\irl]{\irl{x}}{\partialeval{\irl{\expr[_{1}]}}{\irl{\substlist}}}{\\&\partialeval{\irl{\expr[_{2}]}}{\irl{\substlist}}} \\
-%   \partialeval{\irl{\lifz{\expr[_{1}]}{\expr[_{2}]}{\expr[_{3}]}}}{\irl{\substlist}} & = \lifz[\irl]{\partialeval{\irl{\expr[_{1}]}}{\irl{\substlist}}}{\\&\partialeval{\irl{\expr[_{2}]}}{\irl{\substlist}}}{\partialeval{\irl{\expr[_{3}]}}{\irl{\substlist}}} \\
-% \end{align*}
-\end{gather*}
-% \vspace{-3em}
-
-Note that both passes have no effect on the resulting trace of a program, up to $\emptyevent$-steps. 
-Because of this, both passes have equality as corresponding cross language trace relation. 
-Moreover, it is straightforward to prove both passes as secure (\Cref{def:rtp}) w.r.t. \gls*{ms}.
-
-\begin{theorem}[$\ccdce$ secure w.r.t. \gls*{ms}]\label{thm:ccdce:rtp:ms}
-  $\rtp{\ccdce}{\mssafe}$ %\Coqed
-\end{theorem}
-\begin{theorem}[$\cccf$ secure w.r.t. \gls*{ms}]\label{thm:cccf:rtp:ms}
-  $\rtp{\cccf}{\mssafe}$ %\Coqe
-\end{theorem}
-
-With both \Cref{thm:ccdce:rtp:ms,thm:cccf:rtp:ms} it follows from \Cref{corr:swappable} that the two passes can be interchanged arbitrarily:
-
-\begin{theorem}[$\cccf\circ\ccdce$ and $\cccf\circ\ccdce$ are secure w.r.t. \gls*{ms}]\label{thm:cccfccdce:rtp:ms}
-$\;$
-
-  \begin{nscenter}
-  \phantom{and~}\!\!$\rtp{\cccf\circ\ccdce}{\mssafe}$
-
-  and~$\rtp{\ccdce\circ\cccf}{\mssafe}$ % \Coqed
-  \end{nscenter}
-\end{theorem}
-
-\subsection{Robust Strict Cryptographic Constant Time Preservation}\label{subsec:cs:scct}
-
-This section defines a compiler $\ccscct$ from $\Lms$ to $\Lscct$ that robustly preserves \gls*{scct}. 
-Given the fact that $\Lscct$ provides a \gls*{cct}-mode that can be turned on or off, the compiler inserts wrapper code for function calls and function bodies to ensure that execution in the component always happen in \gls*{cct}-mode.
-This simple flag combines the effect of FaCT~\cite{cauligi2019fact}
-
-\vspace{-1em}
-\[
-\begin{array}{rcl}
-  \ccscct(\irl{\lfunction{g}{x}{\expr}}) & = & \lfunction[\obj]{\obj{g}}{\obj{x}}{\obj{\lwrdoit{ON};}\ccscct(\irl{\expr})} \\
-  \ccscct(\irl{\lcall{g}{\expr}}) & = & \lcall[\obj]{\obj{g}}{\ccscct(\irl{\expr})\obj{; \lwrdoit{ON}}} 
-    % \\
-    % \ccscct(\irl{\lbinop{\expr[_{1}]}{\expr[_{2}]}}) & = \lbinop[\obj]{\ccscct{\irl{\expr[_{1}]}}}{\ccscct{\irl{\expr[_{2}]}}} 
-    % \\
-\end{array}
-\]
-% \vspace{-2em}
-%
-The context can overwrite the flag and exit the mode, but upon invoking a function that is part of the component, the flag is set again.
-Because of this, the corresponding cross-language trace relation $\sim^{\Lms}_{\Lscct}$, only relates events without secrets:%
-
-\begin{center} 
-  \typerulenolabel{xrel:scct:noleak}{}{\irl{\preevent;\comp}\sim^{\Lms}_{\Lscct}\obj{\preevent;\comp;\unlock}}
-\end{center}
-
-The compiler is secure w.r.t. \gls*{scct}: %, similarly proven as in \Cref{subsec:cs:tms}.
-
-\begin{theorem}[$\ccscct$ secure w.r.t. \gls*{scct}]\label{thm:ccscct:rtp:scct}
-  \small
-  $\rtpsim{\ccscct}{\scctsafe}{\sim^{\Lms}_{\Lscct}}$ % \Coqed
-\end{theorem}
-
-\subsection{Robust Speculative Safety Preservation}\label{subsec:cs:ss}
-
-This section defines the final compilation pass $\ccspec$, which ensures that $\Lscct$ programs, which are assumed to be \gls*{ss}, stay \gls*{ss} at $\Lspec$-level. 
-To do so, $\ccspec$ inserts a speculation barrier after branches, which is sufficient to harden the program against speculative leaks, since SPECTRE-PHT~\cite{kocher2019spectre} is the only speculative leak modeled in the semantics of $\Lspec$.
-
-\vspace{-1em}
-\[
-\begin{array}{cl}
-  &\ccspec{(\obj{\lifz{\expr[_0]}{\expr[_1]}{\expr[_2]}})} = 
-  \\
-  &\qquad\qquad \lifz[\ird]{\ccspec{\left(\obj{\expr[_0]}\right)}}{\ird{\lbarrier;}\ccspec{\left(\obj{\expr[_1]}\right)} \\&\qquad\qquad}{ \ird{\lbarrier;}\ccspec{\left(\obj{\expr[_2]}\right)}} 
-\end{array}
-\]
-% \vspace{-2em}
-%
-
-Clearly, the corresponding cross-language trace relation $\sim^{\Lscct}_{\Lspec}$ has only one non-trivial case: for branches, only relate them where speculation is blocked by a barrier:
-
-\begin{center}
-  \typerulenolabel{xrel:spec:if}{}{\obj{Branch\ n}\sim^{\Lscct}_{\Lspec}\ird{Branch\ n}\cdot\ird{Spec}\cdot\ird{Barrier}\cdot\ird{Rlb}}
-\end{center}
-
-The base-event relation above scales to full events by ensuring the missing annotations ($\obj{\comp;\securitytag{}}$ and $\ird{\comp;\securitytag{}}$) are the same.
-% 
-With this relation, we prove that $\ccspec$ is secure with respect to \gls*{ss}.
-\begin{theorem}[$\ccspec$ secure w.r.t. \gls*{ss}]\label{thm:ccspec:rtp:spec}
-  \small$\rtpsim{\ccspec}{\sssafe}{\sim^{\Lscct}_{\Lspec}}$ % \Coqed
-\end{theorem}
-
-\subsection{Robust Preservation of Memory Safety, Strict Cryptographic Constant Time, and Speculative Safety}
-
-Finally, this subsection combines all previous results into one compilation chain to get that it preserves full \gls*{specms}.
-Let $\ccspecms$ be the compiler that is the composition of $\cca$, $\ccb$, $\cccf$, $\ccdce$, $\ccscct$, and $\ccspec$. 
-Let $\sim^{\Ltms}_{\Lspec}$ be the composition of $\sim^{\Ltms}_{\Lms}$, $\sim^{\Lms}_{\Lscct}$, and $\sim^{\Lscct}_{\Lspec}$.
-Then, the following corollary holds.
-
-\begin{corollary}[$\ccspecms$ secure w.r.t. \gls*{specms}]\label{thm:ccall:rtp:specms}
-  $\;$
-
-  \begin{nscenter}
-    $\rtpsim{\cc{\Ltms}{\Lspec}}{\mssafe\cap\scctsafe\cap\sssafe}{\sim^{\Ltms}_{\Lspec}}$ % \Coqed
-  \end{nscenter}
-\end{corollary}
-
-As with \Cref{corr:ccab:rtp:ms}, it is important to ensure that the respective cross language trace relations are well-formed (\Cref{def:wfc:sig:tracerel}).
-It is already known that $\sim^{\Ltms}_{\Lms}$ is well-formed with respect to $\mssafe$ (\Cref{lem:wf:ltmsltrg}).
-Next in the chain is $\sim^{\Lms}_{\Lscct}$, which has to be well-formed w.r.t. $\scctsafe$.
-This lemma holds, since a trace that was $\scctsafe$ is $\scctsafe$ even after applying $\sim^{\Lms}_{\Lscct}$: the relation enforces that $\Lscct$ traces related to $\Lms$ traces have no leaks of secrets whatsoever.
-
-\begin{lemma}[$\sim^{\Lms}_{\Lscct}$ well-formed w.r.t. $\scctsafe$]\label{lem:wf:lsmslscct}
-$\;$
-
-  \begin{nscenter}
-  $\wfcsig{\sim^{\Lms}_{\Lscct}}{\scctsafe}$
-  \end{nscenter}
-\end{lemma}
-
-The last relation is $\sim^{\Lscct}_{\Lspec}$ which needs to be well-formed w.r.t. $\sssafe$.
-Similarly to the previous relation, this holds, since $\sim^{\Lscct}_{\Lspec}$ only relates $\Lspec$ traces, which do not have speculative leaks, with $\Lscct$ traces.
-
-\begin{lemma}[$\sim^{\Lscct}_{\Lspec}$ well-formed w.r.t. $\sssafe$]\label{lem:wf:lscctlspec}
-  $\wfcsig{\sim^{\Lscct}_{\Lspec}}{\sssafe}$
-\end{lemma}
-
-\section{Formal Insights}\label{sec:formalities}
-
-This section first discusses how to connect each language-specific security property to the general security properties presented in \Cref{sec:compprop} (\Cref{subsec:formalities:maps}), then it demonstrates that the security property resulting from the universal image projection is faithful to the original property (\Cref{subsec:formalities:props}).
-Finally, this section discusses why the order of compiler passes matters, and how does our framework help with identifying insecure compositions (\Cref{subsec:compatsecpasses}), and it gives additional technical insights on the secure compilation proofs (\Cref{subsec:seccompproofs}).
-
-\subsection{From Language Traces to General Ones}\label{subsec:formalities:maps}
-
-The previous theorems talk about preserving properties expressed in the trace model of the languages of each compiler.
-However, these trace models are not the same trace model we used to specify the properties of \Cref{sec:compprop} (indicated with \ev{L}), which serves as the ``ground truth'' for the meaning of our properties.
-% 
-To bridge this gap, the formal development requires specifying additional trace relations, from each of the language trace models to the \ev{L} one, that, for example, relate $\src{Get\ \loc\ n}$ and $\src{Set\ \loc\ n}$ to $\ev{Use\ \loc\ n}$ (and that induce a related universal image that we use in \Cref{subsec:formalities:props}).
-% 
-One key insight of these relations is that they all omit context-made actions for two reasons: (1) contexts (which are universally quantified) can trivially invalidate any property and (2) we are interested in the component upholding the properties.
-%
-
- % disregarded a glaring but minor technical issue: take $\tmssafe$ (\Cref{def:trace:tmsdef}) as an example.
-% It contains traces as specified in \Cref{subsec:basic:memsafety:tracemodel}.
-% However, these are different from $\Ltms$ traces (\Cref{subsec:ltms}) which means that \Cref{thm:wt:tms} is ``ill-typed''.
-% Similarly for the other theorems.
-% In the interest of reducing visual baggage, the notation left out yet another cross-language trace relation, translating language-specific traces into these non-language specific traces. 
-% In doing so, context-level actions are dropped and, e.g., both $\src{Get\ \loc\ n}$ and $\src{Set\ \loc\ n}$ are translated into $\ev{Use\ \loc\ n}$.
-
-\subsection{Security Properties and Their Meaning}\label{subsec:formalities:props}
-Each of the presented compilers use a cross-language trace relation, which is also used to translate the property from one language to the other one (via the existential or universal images).
-% 
-%Whenever such translations occur, it is important to check whether there has been a change of meaning.
-%Secure compiler engineers need to ensure that such translations of properties do not change meaning in a fundamental wrong way. 
-% lifting or lowering reasoning from target- to source-level or vice-versa,
-While the meaning of projected properties does change with a translation, the change should not allow for a flawed compilation pipeline.
-% 
-For example, we could be using a trace relation that translates a property in a language to a totally different one in another language.
-% 
-To raise the trust into the translation of properties, \Cref{thm:prop-rel-corr} states (in a general fashion) that each security property is faithfully translated using the universal image according to the cross-language trace relation induced by the compiler.
-
-\begin{theorem}[Properties Relation Correctness]\label{thm:prop-rel-corr}
-  \begin{align*}
-    \forall& \pi \in \{\tmssafe,\smssafe,\scctsafe,\sssafe\}, 
-    \\
-    \text{ for each }& \text{pair of languages } \src{L} \text{ and } \trg{L} \text{ used by the compilers},
-    \\
-    \text{ if }& 
-    % \tau_{\sim^{\trg{L}}_{\ev{L}}}(\pi) \sim^{\trg{L}}_{\ev{L}} \pi
-    \sigma_{\sim^{\trg{L}}_{\ev{L}}}(\pi) \sim^{\trg{L}}_{\ev{L}} \pi
-    \text{ and } 
-    % \sigma_{\sim^{\src{L}}_{\trg{L}}}(\tau_{\sim^{\trg{L}}_{\ev{L}}}(\pi)) \sim^{\src{L}}_{\trg{L}} \tau_{\sim^{\trg{L}}_{\ev{L}}}(\pi)
-    \sigma_{\sim^{\src{L}}_{\trg{L}}}(\sigma_{\sim^{\trg{L}}_{\ev{L}}}(\pi)) \sim^{\src{L}}_{\trg{L}} \sigma_{\sim^{\trg{L}}_{\ev{L}}}(\pi)
-    \\
-    \text{ then }& 
-    % \tau_{\sim^{\src{L}}_{\ev{L}}}(\pi) = \sigma_{\sim^{\src{L}}_{\trg{L}}}(\tau_{\sim^{\trg{L}}_{\ev{L}}}(\pi)) 
-    \sigma_{\sim^{\src{L}}_{\trg{L}}}(\sigma_{\sim^{\trg{L}}_{\ev{L}}}(\pi)) \sim^{\src{L}}_{\ev{L}} \pi
-  \end{align*}
-\end{theorem}
-The complexity of this theorem is that the relation in the conclusion cannot be obtained by composing the two relations in the premises.
-% 
-%% MK: the following does not say anything about the theorem, only about the syntactic compiler
-%What happens in practice is that compiler engineers are given the relations from each language to \ev{L} and they need to build a compiler that induces a relation $\sim^{\src{L}}_{\trg{L}}$ that respects this theorem.
-%
-
-We now informally argue why this theorem holds for the composition of all considered properties from \Cref{sec:compprop}.
-
-\paragraph{$\sigma_{\sim^{\Ltms}_{\Lms}}\left(\tmssafe\cap\smssafe\cap\scctsafe\cap\sssafe\right)$}
-
-For the four properties considered here, the trace models of $\Ltms$, $\Ltrg$, and $\Lms$ do not consider actions related to $\scctsafe$ and $\sssafe$, so these two properties are trivially translated correctly.
-% 
-We now discuss the remaining $\tmssafe$ and $\smssafe$ in the form of their intersection $\mssafe$.
-
-Recall that $\sim^{\Ltms}_{\Lms}$ is the composition of $\sim^{\Ltms}_{\Ltrg}$ and $\sim^{\Ltms}_{\Lms}$.
-% 
-Since $\sim^{\Ltms}_{\Ltrg}$ is an equality, this relation trivially preserves the meaning of translated properties: related traces are identical!
-
-Finally, let us consider a trace $\irl{\trace}\in\mssafe$ and understand what is that is related to via $\sim^{\Ltrg}_{\Lms}$.
-% 
-{\em All} traces $\trg{\trace}$ with $\trg{\trace}\sim^{\Ltrg}_{\Lms}\irl{\trace}$ are identical to $\irl{\trace}$ except for get and set actions, which require for in-bound accesses (as stated in \Cref{subsec:cs:ms}): this clearly respect $\mssafe$.
-%
-
-% In both cases, the translated event does not represent an out-of-bounds memory access.
-
-% \footnote{Both $\sim^{\Ltms}_{\Ltrg}$ and $\sim^{\Ltrg}_{\Lms}$ filter context actions.}.
-
-% Now, starting from $\mssafe$ expressed in $\Lms$, we check that its translation in $\Ltms$ is still $\mssafe$.
-% % 
-% We can see that the property resulting from the projection of $\Lms$ traces back to $\Ltms$ faithfully models $\mssafe$, which translates to both $\scctsafe$ and $\sssafe$, 
-% \MPin{ equiv  +  other rel only adds traces that fail }
-% since these properties trivially hold for $\Ltms$, $\Ltrg$, and $\Lms$ traces.
-
-\paragraph{$\sigma_{\sim^{\Lms}_{\Lscct}}\left(\tmssafe\cap\smssafe\cap\scctsafe\cap\sssafe\right)$}
-
-As before, the trace models of $\Lscct$ and $\Lms$ cannot express $\sssafe$, so that is trivially translated correctly.
-% 
-Also, the trace model of $\Lscct$ extends the one of $\Lms$ with respect to $\tmssafe$ and $\smssafe$ events, so the translation argument regarding those two properties is the same as before.
-% 
-Thus, we need to reason about whether $\scctsafe$ is translated correctly.
-
-By definition, $\sim^{\Lms}_{\Lscct}$ only relates $\obj{\unlock}$ events, which is also the same relation induced by $\sim_{\ctsafe}$ in \Cref{sec:msscct-rel}.
-% 
-This ensures that composing the relations only relates $\unlock$ events, and thus the property is translated correctly.
-
-% Consider a trace $\obj{\trace}\in\scctsafe$ and all $\irl{\trace}$ such that $\irl{\trace}\sim^{\Lms}_{\Lscct}\obj{\trace}$. 
-% In turn, $\obj{\trace}$ cannot contain any leaks of secret values, so it cannot violate $\scctsafe$.
-% Toggling data-independent timing mode has no impact on $\mssafe$ or $\sssafe$.
-
-\paragraph{$\sigma_{\sim^{\Lscct}_{\Lspec}}\left(\tmssafe\cap\smssafe\cap\scctsafe\cap\sssafe\right)$}
-The trace model of $\Lspec$ extends the one of $\Lscct$ with respect to $\tmssafe$, $\smssafe$, and $\scctsafe$, so the translation argument for those three properties is the same as before.
-% 
-Concerning $\sssafe$, $\sim^{\Lscct}_{\Lspec}$ relates $\Lspec$ speculation traces with $\Lscct$ branches with the $\lock_{\text{NONE}}$ tag, so the property is translated correctly, according to the relation defined in \Cref{sec:spec-ms-rel}.
- % introducing $\ird{barrier}$ never violates $\sssafe$ and has no effect on $\mssafe$ or $\scctsafe$.
-
-%\MPin{
-%	can this be made formal?
-%	what's the point?
-%}
-
-\subsection{Compatibility of Secure Compiler Passes}\label{subsec:compatsecpasses}
-
-% \Cref{thm:rtpsim:sig} is applicable, up to the well-formedness of $\sim_1$ w.r.t. $\trg{\class[_2]}$.
-Consider applying $\ccspec$ first and then $\ccb$ (albeit currently syntactically impossible).
-% , let us pretend that it is possible by just extending $\ccb$ for the constructs present in $\Lscct$ that are not present in $\Ltrg$.
-% Clearly, as evidenced earlier (\Cref{subsec:formalities:props}), $\ccb\cdot\ccspec$ is an acceptable composition that yields an interesting class of security properties, i.e., $\sigma_{\sim_{\irl{ms}}\bullet\sim_{\ird{\mathghost}}}\left(\mssafe\cap\scctsafe\cap\sssafe\right)$.
-In this case, $\ccb$ would insert new branches into the code that are not protected by a speculation barrier! 
-This concern is reflected in the proofs that establish that the security class resulting of the composition of the trace relations is meaningful.
-% , much similar to ensuring that a definition models what the formal methods expert intends to model. 
-For the $\ccspec\cdot\ccb$ case, the class is $\sigma_{\sim_{\ird{\mathghost}}\bullet\sim_{\irl{ms}}}\left(\mssafe\cap\scctsafe\cap\sssafe\right)$.
-Since $\ev{Use\ \loc\ n}\sim_{\irl{ms}}\irl{Branch\ n}\cdot\irl{Spec}\cdots$, where $\cdots$ does not contain a $\ev{{Barrier}}$ event, the resulting class is \emph{not} the original $\sssafe$ that is intended and it would break the corresponding \Cref{thm:prop-rel-corr}.
-% Because of this, it is crucial to analyse the precise shape of the class of properties after mapping a target-level property to a source-level property\footnote{Dito for mapping a source-level property to a target-level property.}.
-
-However, the composition is still technically possible and it is the job of the compiler engineers to ensure that the secure compilation pipeline happens in an order that ensures that the mapped security property is the intended one.
-
-\subsection{Secure Compilation Proofs}\label{subsec:seccompproofs}
-
-Our secure compilation proofs rely on backtranslations~\cite{abate2019jour,patrignani2021rsc}, which let one construct a source context starting from either target traces (aka trace-based backtranslations) or target contexts (aka context-based backtranslations).
-These backtranslations also require setting up cross-language relations between various language elements such as expressions and program states, so we leave these details for the technical reports.
-% , the usual strategy is to establish a backwards simulation between source components and their compiled target counterpart. 
-% For the universally quantified target context, a technique known as backtranslation is well-established, which constructs a source-level context either by backtranslating the target-level context, amounting to a ``decompilation'', or by constructing it using the target-level trace, so that the source-context emits parts of the whole execution trace that are related to the parts emitted by the target-level context.
-All backtranslations in the case-study are trace based except for those required by $\ccdce$ and $\cccf$, which are context-based (and they are an identity function). 
-% Since these optimisations have no significant semantic effect and source- and target-language are equal, the backtranslation amounts to an identity translation.
-
-\section{Related Work\pages{2}}\label{sec:relwork}
-
-This section discusses robust compilation, other secure compilation criteria and work related to the properties preserved in the case study.
-% (\Cref{subsec:relw:seccomprtp}) 
-% (\Cref{subsec:relw:seccompcrit}).
-% Since the case study of \Cref{sec:casestud:defs,sec:casestud:rtp} implements measures for preserving \gls*{ms}, \gls*{cct}, and \gls*{ss}, 
-% Then, this section discusses work related 
-% relevant related work as well 
-% (\Cref{subsec:relw:msmechs,subsec:relw:cctmechs,subsec:relw:ssmechs}).
-
-\paragraph*{Secure Compilation as Robust Preservation}\label{subsec:relw:seccomprtp}
-
-The robust preservation of properties as a compiler-level criterion has been analysed extensively~\cite{abate2019jour,patrignani2021rsc,abate2021extacc,patrignani2019survey} and thus we build on that framework.
-No existing work is concerned with composing secure compilers, however, existing work~\cite{abate2021extacc} sketches composition of trace-relating compiler correctness in a similar way to what has been presented here.
-%Parts of these works consider languages with different trace models and our technical setup can handle this. 
-The work relating robust preservation with universal composability~\cite{patrignani2022universal} is closest to what this paper presents.
-The authors demonstrate a similar compositionality theorem to what is presented here (\Cref{sec:sequential}) but use it in the context of protocols.
-The work does not consider the generality to support different trace models or composition of compilers which robustly preserve different classes.
-% The authors demonstrate a similar compositionality theorem to what is presented here (\Cref{sec:sequential}) as well as in an earlier version of this work~\cite{kruse2022csc}.
-% However, they do not demonstrate the scalability of the approach by means of a case study.
-
-%\paragraph*{Composition of Secure Compilers}\label{subsec:relw:seccompcomp}
-
-There is work on lifting exploits for single compilers to the whole chain~\cite{paykin2019weird}.
-%Our framework does not consider exploits.
-%\MP{ a bit dry and short}
-While that work considers {\em in}secure compilation and composition thereof in terms of exploits, the composition they are interested in allows to lift an exploit for one compiler pass to the whole compilation chain.
-Our framework takes the opposite direction and provides compositionality results for secure compilers.
-%
-%
-%To the best of our knowledge, there does not exist published work discussing the compositionality of the robust preservation criterion, the gold-standard of secure compilation.
-
-\paragraph*{Other Secure Compilation Criteria}\label{subsec:relw:seccompcrit}
-
-While this paper focuses on the robust preservation framework~\cite{abate2019jour}, other secure compilation criteria exist.
-The survey on formal approaches to secure compilation~\cite{patrignani2019survey} discusses a broad spectrum already, while this section presents a very high-level overview.
-Fully abstract compilation~\cite{abadi1999fullabstraction} states that a compiler should preserve and reflect observational equivalence between source and target programs.
-Abate \emph{et al.}~\cite{abate2021faandrc} showed that fully abstract compilers robustly preserve program properties that are either trivial or meaningless.
-As a mitigation for this, the authors presented a categorical approach based on maps of distributive laws~\cite{watanabe2002modl}, which they call many maps of distributive laws.
-Maps of distributive laws have been investigated before as a possible secure compilation criterion~\cite{tsampas2020catsc}.
-Other approaches are extensions of the compiler correctness criterion as discussed in other work~\cite{patterson2019next700} or the introduction of opaque observations~\cite{vu2021reconciling} to reconcile compiler optimisations with security.
-Note that this work also presents secure compilers that are optimising, but contrary to the other~\cite{vu2021reconciling}, provides a formal account of these in the robust preservation framework.
-% Lastly, the authors of this paper have presented ongoing work~\cite{patrignani2023blame} on a weaker robust preservation criteria based on the concept blame.
-
-\paragraph*{Memory Safety Mechanisms}\label{subsec:relw:msmechs}
-
-Different mechanisms for enforcing memory safety exist that also consider the secure compilation domain, i.e., have an active attacker model.
-For example, the ``pointers as capabilities'' principle represents pointers as machine-level capabilities~\cite{korashy2021capableptrs}, which behave in a similar fashion to capabilities by means of linear typing~\cite{morrisett2005L3}.
-The approach of this paper also uses linear typing, but differs from $L^{3}$~\cite{morrisett2005L3} in the way that functions are not first-class.
-Moreover, this paper considers an active attacker, while the work on $L^{3}$ only discusses whole programs and, thus, has no active attacker model.
-The instrumentation to ensure memory safety that this paper presents is inspired by Softbounds~\cite{nagarakatte2009soft}.
-That work inserts bounds-checks in front of pointer-dereferences and, for this to work, inserts meta-data information on pointer creation.
-Softbounds also works in a more advanced setting with structured fields accesses and also introduces a table-lookup for pointers that are stored in memory.
-This paper only considers arrays of primitive data, i.e., there are no pointers to pointers or structures.
-Several other approaches to memory-safety exist in literature, specifically as compiler instrumentations~\cite{akritidis2009baggy,younan2010paricheck,jung2021pico,shankaranarayana2023tailcheck,dhumbumroong2020boundwarden,nam2019framer,zhou2023fatptrs}, hardware-extensions~\cite{kwon2013lowfat,saileshwar2022heapcheck,chen2023flexpointer,kim2023whistle}, or programming language extensions~\cite{elliott2018checkedc,li2022formalcheckedc,jim2002cyclone,elliott2015guilt,west2005cuckoo,weis2019fyr,benoit2019uniqueness}.
-What differentiates this work from them is that this work uses known, compiler-based approaches to ensure memory-safety as a means to investigate secure compiler compositions.
-This paper does not provide efficient memory-safety, but serves as a theoretical foundation for the secure compilation domain.
-
-To extend the languages in this paper with a less restricted form of pointer arithmetic, the region colouring memory safety monitor presented in earlier work~\cite{michael2023mswasm} can be used.
-The work presenting this monitor provides an approach for the robust preservation of memory safety compiling from C to WASM.
-However, they do not discuss composition of secure compilers but rather investigate an instance of a secure compiler.
-
-\paragraph*{Cryptographic Constant Time Mechanisms}\label{subsec:relw:cctmechs}
-
-The approach to preserving cryptographic constant time in this paper is high-level, where a programming language exposes a way to switch the semantics to a data (operand) independent timing mode.
-Since identifiers in $\Lscct$ are annotated with a secrecy tag, this approach is similar to others with information flow control.
-For example, Vale~\cite{bond2017vale} uses Dafny to ensure constant-time assembly code, while Jasmin~\cite{almeida2017jasmin} makes use of the Coq proof assistant to reject non-constant-time programs.
-CT-Wasm~\cite{watt2019ctwasm} enforces constant-timeness by means of a type system.
-Different to the approach of this paper, these approaches necessitate that the programmer writes \gls*{cct} code.
-An approach to allow programmers to write more high-level code is CryptOpt~\cite{kuepper2023cryptopt}, which generates efficient target-code by means of a randomised search.
-This paper abstracts over concrete mitigation strategies and simply assumes that there is a flag to switch to a cryptographic-constant time execution mode.
-This can be realised by employing the FaCT~\cite{cauligi2019fact} compiler, which translates common non-constant time code patterns to be constant-time, and the data (object) independent timing execution mode of modern processors.
-
-\paragraph*{Speculation Safety Mechanisms}\label{subsec:relw:ssmechs}
-
-This paper uses a taint-tracking mechanism inspired by existing work~\cite{guarnieri2018spectector,fabian2022automatic}.
-These taints are used to express absence of any speculative leaks in \gls*{ss}~\cite{guarnieri2018spectector}. 
-The semantics of $\Lspec$ hardcodes the kind of speculative leaks to just SPECTRE-PHT~\cite{kocher2019spectre}, but future work could use semantics composition~\cite{fabian2022automatic} to support more variants.
-Note that our framework composes compilers and not semantics.
-
-%This work has taken inspiration in the sematics of $\Lspec$ and in \Cref{def:trace:ss} from recent formalisations~\cite{guarnieri2018spectector,fabian2022automatic}.
-% \MPin{
-%  we take SS from exorcising.
-%  we do not take SNI from spectector because it is a hypersafety.
-%  other work defined similar props. 
-% }
-% A subsequent line of work~\cite{fabian2022automatic} shows how to modularly define and compose formal semantics that models different speculative execution variants. 
-% Recent work~\cite{fabian2024lift} demonstrates how to lift existing secure compilation proofs for security properties related to speculative execution to more powerful properties, i.e., properties defined for languages with a trace model that enables more vulnerabilities, such as combinations of different speculative execution exploits.
-% To this end, they define a lifting framework and demonstrate how a secure compilation proof for a weaker property can be lifted to a stronger one. 
-% Contrary to the compositionality framework presented in this paper, these works are only concerned with composing semantics of the respective programming languages, while in this paper, the secure compilers themselves are composed.
-
-\section{Conclusion\pages{$\sfrac{1}{2}$}}\label{sec:concl}
-This paper tackles the problem of understanding what kind of security properties a secure compiler preserves, when said compiler is the combination of compiler passes that preserve possibly different security properties.
-% 
-%For this, this paper first formalised security properties of interest and their composition.
-% 
-The paper proves that composing secure compilers that preserve certain properties results in a secure compiler that preserves the composition of these properties.
-% While the presented security property composition relied on monitors that check only trace-properties, the composition of secure compilers does not make any restriction towards the kind of property involved in the composition.
-% It is subject to future work to develop techniques to verify relational hyperproperties~\cite{abate2019jour}, while the composition of hyperproperties could be very similar to the composition of ordinary properties as presented in this paper, since hyperproperties can be checked with an automata-based model-checker~\cite{beutner23hyperltl}.
-% 
-Finally, this paper defines a multi-pass compiler and proves that it preserves \gls*{specms}.
-Crucially, this paper derives the security of the multi-pass compiler from the composition of the security properties preserved by its individual passes, which include security-preserving as well as optimisation passes.
-% For future work, it is interesting to look at more sophisticated optimisation passes that, e.g., reorder events that appear on traces.
-
-\newpage
-
-\bibliographystyle{IEEEtranS}
-\bibliography{main}
-
-%%
-%% If your work has an appendix, this is the place to put it.
-\appendix
-\subsection{Existential Image}\label{subsec:extimg}
-
-\begin{definition}[Existential Image]
-% [Existential and Universal Image]
-\label{def:existential:img}\label{def:tau}
-  \[
-    \tau_\sim\left(\src{\pi}\right) := 
-      \left\{ 
-        \trg{\trace} \mid \exists \src{\trace}\ldotp \src{\trace}\sim\trg{\trace}, \text{ and } \src{\trace}\in\src{\pi} 
-      \right\}
-  \]
-  %\[ 
-  %  \sigma_\sim\left(\trg{\pi}\right) := 
-  %    \left\{ 
-  %      \src{\trace} \mid \forall \trg{\trace}\ldotp \text{if }\src{\trace}\sim\trg{\trace}, \text{ then } \trg{\trace}\in\trg{\pi} 
-  %    \right\}
-  %\]
-\end{definition}
-
-\begin{definition}[Robust Preservation with $\tau_\sim$]\label{def:rtp:tau}
-  % $\;$\\
-  % \vspace{-1em}
-  % \begin{nscenter}
-  % Compiler $\cc{\src{L}}{\trg{L}}$ robustly preserves $\class$, 
-  {$\rtptau{\cc{\src{L}}{\trg{L}}}{\src{\class}}{\sim}$}
-  %, iff 
-  $\isdef$
-    {$\forall \src{\pi}\in\src{\class}, \src{p}\in\src{L},$} if {$\rsat{\src{\progvar}}{\src{\pi}}$}, then {$\rsat{\cc{\src{L}}{\trg{L}}\left(\src{p}\right)}{\tau_\sim\left(\src{\pi}\right)}$}.
-  % \end{nscenter}
-\end{definition}
-
-\begin{theorem}[Composition of Secure Compilers w.r.t. $\tau$]\label{thm:rtpsim:tau}
-  $\;$
-
-  If {$\rtptau{\cc{\src{L}}{\trg{L}}}{\src{\class[_{1}]}}{\sim_1}$}, {$\rtptau{\cc{\trg{L}}{\obj{L}}}{\tilde{\tau}_{\sim_1}\left(\src{\class[_2]}\right)}{\sim_2}$}, and {$\wfctau{\sim_1}{\src{\class[_2]}}$}, \\ then {$\rtptau{\cc{\src{L}}{\trg{L}}\circ\cc{\trg{L}}{\obj{L}}}{\src{\class[_{1}]}\cap\src{\class[_{2}]}}{\sim_1\bullet\sim_2}$}.
-\Coqed
-\end{theorem}
-
-\begin{corollary}[Swapping Secure Compiler Passes]\label{corr:swappable:tau}
-  $\;$
-
-  If {$\rtptau{\cc[_{1}]{\trg{L}}{\trg{L}}}{\src{\class[_{1}]}}{\sim_1}$ and $\rtptau{\cc[_{2}]{\trg{L}}{\trg{L}}}{\src{\class[_{2}]}}{\sim_2}$}, %
-  {$\wfctau{\sim_1}{\src{\class[_2]}}$ and $\wfctau{\sim_2}{\src{\class[_1]}}$}, %
-  and {$\tilde{\tau}_{\sim_1}\left(\src{\class[_2]}\right)=\src{\class[_2]}$ as well as $\tilde{\tau}_{\sim_2}\left(\src{\class[_1]}\right)=\src{\class[_1]}$},
-  then {$\rtptau{\cc[_{1}]{\trg{L}}{\trg{L}}\circ\cc[_{2}]{\trg{L}}{\trg{L}}}{\src{\class[_{1}]}\cap\src{\class[_{2}]}}{\sim_1\circ\sim_2}$ and $\rtptau{\cc[_{2}]{\trg{L}}{\trg{L}}\circ\cc[_{1}]{\trg{L}}{\trg{L}}}{\src{\class[_{2}]}\cap\src{\class[_{1}]}}{\sim_2\circ\sim_1}$}.
-\Coqed
-\end{corollary}
-
-\subsection{Secure Upper and Lower Composition}\label{sec:other-compos}
-Besides sequential composition, there are two other compositions, namely an {\em upper}, i.e., a compiler that takes multiple inputs and yields one output, and a {\em lower} composition, i.e., a compiler that takes one input and yields multiple outputs.
-We {define the upper composition $\cc{\src{L}+\obj{L}}{\trg{L}}$} as follows:
-Given a program \texttt{p}, its compiled counterpart is obtained by {plugging \texttt{p} into $\cc{\src{L}}{\trg{L}}$ if $\texttt{p}\in\src{L}$} or by {plugging \texttt{p} into $\cc{\obj{L}}{\trg{L}}$ if $\texttt{p}\in\obj{L}$}.
-\begin{definition}[Upper Composition]
-  \[
-    \text{{$\cc{\src{L}+\obj{L}}{\trg{L}}$}}\isdef
-  % Given \texttt{p}, yield
-  \lambda \texttt{p}\ldotp
-  \left\{\mbox{\begin{tabular}{c}
-    {{if $\texttt{p}\in\src{L}$, then $\cc{\src{L}}{\trg{L}}(\texttt{p})$}} \\
-    \mbox{{if $\texttt{p}\in\obj{L}$, then $\cc{\obj{L}}{\trg{L}}(\texttt{p})$}} \\
-  \end{tabular}}\right. 
-  \]
-\end{definition}
-
-Examples of this are present in industry:
-Consider the Java Virtual Machine bytecode $\trg{JVM BC}$, which is a popular target for programming language designers due to its high performance and relevance in industry.
-Compilers for several programming languages have it as their target language, some popular instances are $\src{Java}$ and $\obj{Kotlin}$.
-Technically speaking, they both compile to class files and $\obj{Kotlin}$ objects are considered to be the same as $\src{Java}$ objects at that point.
-Both languages can be used at the same time in one project~\cite{androidstudio}.
-A compiler that accepts both $\src{Java}$ and $\obj{Kotlin}$ code translating to the same target language or intermediate representation performs a kind of {\em upper} composition.
-Now, the following theorem tells us what happens if these are secure:
-Given {$\cc{\src{L}}{\trg{L}}$ robustly preserves $\class[_{1}]$} and {$\cc{\obj{L}}{\trg{L}}$ robustly preserves $\class[_{2}]$}, it follows that {their upper composition $\cc{\src{L}+\obj{L}}{\trg{L}}$ robustly preserves the intersection of classes $\class[_{1}]$ and $\class[_{2}]$}.
-
-\begin{theorem}[Upper Composition of Secure Compilers]\label{thm:urtp}
-  $\;$
-
-  If {$\rtp{\cc{\src{L}}{\trg{L}}}{\class[_{1}]}$} and {$\rtp{\cc{\obj{L}}{\trg{L}}}{\class[_{2}]}$}, then {$\rtp{\cc{\src{L}+\obj{L}}{\trg{L}}}{\class[_{1}]\cap\class[_{2}]}$}. %\Coqed
-\end{theorem}
-
-Dually, the {\em lower} composition is concerned about compilers that accept the same source but yield different target languages. %and the same theoretical results apply.
-{Define the lower composition $\cc{\src{L}}{\trg{L}+\obj{L}}$} as follows:
-Given a program $\src{p}$, its compiled counterpart is obtained by {plugging $\src{p}$ into $\cc{\src{L}}{\trg{L}}$} or by {plugging \texttt{p} into $\cc{\src{L}}{\obj{L}}$}, respectively, {based on the internal decision}.
-\begin{definition}[Lower Composition]
-  \[
-    \text{{$\cc{\src{L}}{\trg{L}+\obj{L}}$}}\isdef
-  % Given $L$ and $\src{p}$, yield
-  \lambda \src{p}, L\ldotp
-  \left\{\mbox{\begin{tabular}{c}
-    {{if $L=\trg{L}$, then} {$\cc{\src{L}}{\trg{L}}(\src{p})$}} \\
-    \mbox{{if $L=\obj{L}$, then} {$\cc{\src{L}}{\obj{L}}(\src{p})$}} \\
-  \end{tabular}}\right.\]
-\end{definition}
-
- Consider two compilers both accepting $\src{LLVM IR}$~\cite{lattner2004llvm} and one of them emits $\trg{x86\_64}$, while the other emits $\obj{ARMv8}$.
- It is intuitive that they are in some sense composed in the LLVM framework, but the decision of when to use one over the other is inherently {\em internal} to the formalisation effort of this kind of composition.
- For example, the user of this compiler provides an explicit flag that instructs to emit $\trg{x86\_64}$ or the framework itself detects the target platform via heuristics, such as supported instructions.
-
- The following theorem demonstrates what happens if the involved compilers are secure:
-Given {$\cc{\src{L}}{\trg{L}}$ robustly preserves $\class[_{1}]$} and {$\cc{\src{L}}{\obj{L}}$ robustly preserves $\class[_{2}]$}, it follows that {their lower composition $\cc{\src{L}}{\trg{L}+\obj{L}}$ robustly preserves the intersection of classes $\class[_{1}]$ and $\class[_{2}]$}.
-
-\begin{theorem}[Lower Composition of Secure Compilers]\label{thm:lrtp}
-  $\;$
-
-  If {$\rtp{\cc{\src{L}}{\trg{L}}}{\class[_{1}]}$} and {$\rtp{\cc{\obj{L}}{\trg{L}}}{\class[_{2}]}$}, then {$\rtp{\cc{\src{L}}{\trg{L}+\obj{L}}}{\class[_{1}]\cap\class[_{2}]}$}. % \Coqed
-\end{theorem}
-
-Either way, the theoretical results suggest that it is possible to always find a ``most-general'', secure compiler, given two secure compilers, that robustly preserves the least-upper bound of the classes involved in their compilation process.
-
-%Future work intends to use upper and lower composition as means to build a secure compiler incrementally
-%This follows current non-formal-methods based compiler engineering, such as LLVM~\cite{lattner2004llvm}, where languages such as \texttt{C++} and \texttt{Fortran} translate into LLVM IR and languages such as \texttt{x86} or \texttt{ARMv8} assembly variants translate out of LLVM IR.
-
-% no coq, because above definitions are non-formal. 
-%   How would you precisely define the lower compo compiler, given two compilers?
-
-%\section{Cryptographic Constant Time}\label{sec:cct}
-
-\end{document}
-\endinput

### 2025-02-13 icfp/sections/prop.tex
move

--- 2024-09-26 csf2/main.tex
@@
-% Fixing: Too many math alphabets used in version normal
-\newcommand\hmmax{0}
-\newcommand\bmmax{0}
-\documentclass[dvipsnames,conference]{IEEEtran}
-
-\usepackage[T1]{fontenc}
-\usepackage[scaled=.83]{beramono}
-
-\usepackage[colorinlistoftodos]{todonotes}
-\usepackage[inference]{semantic}
-\usepackage[switch]{lineno}
-\usepackage{halloweenmath}
-\usepackage{fontawesome5}
-\usepackage{listofitems}
-\usepackage{breakcites}
-\usepackage{glossaries}
-\usepackage{hyperref}
-\usepackage{cleveref}
-\usepackage{stmaryrd}
-\usepackage{marvosym}
-\usepackage{listings}
-\usepackage{amssymb}
-\usepackage{amsmath}
-\usepackage{nameref}
-\usepackage{amsthm}
-\usepackage{xspace}
-\usepackage{xfrac}
-\usepackage{tikz}
-\usepackage{soul}
-\usepackage{bm}
-\usepackage{paralist}
-
-\hypersetup{
-    colorlinks,
-    linkcolor={red!50!black},
-    citecolor={blue!50!black},
-    urlcolor={blue!80!black}
-}
-%\renewcommand\UrlFont{\color{blue}\rmfamily}
-%\newcommand{\url}[1]{\lstinline{#1}}
-
-\setul{0.95ex}{0.3ex}
-
-\input{cmds.tex}
-
-% allow page breaks inside multiline alignment displays
-\allowdisplaybreaks
-
-\Crefname{exampleenv}{Example}{Examples}
-
-\theoremstyle{definition}
-\newtheorem{exampleenv}{Example}[section]
-\newtheorem{lemma}{Lemma}[section]
-\newtheorem{theorem}{Theorem}[section]
-\newtheorem{corollary}{Corollary}[section]
-\newtheorem{definition}{Definition}[section]
-
-\loadglsentries{acronyms}
-\makeglossaries
-
-\AtBeginDocument{%
-  \providecommand\BibTeX{{%
-    \normalfont B\kern-0.5em{\scshape i\kern-0.25em b}\kern-0.8em\TeX}}}
-\makeatletter
-\newcommand{\linebreakand}{%
-  \end{@IEEEauthorhalign}
-  \hfill\mbox{}\par
-  \mbox{}\hfill\begin{@IEEEauthorhalign}
-}
-\makeatother
-
-\renewcommand\thelinenumber{\color{red}\arabic{linenumber}}
-\begin{document}
-\linenumbers
-
-\title{
-  Secure Composition 
-  \\ 
-  of Robust and Optimising Compilers
-  \vspace{-1.5em}
-}
-
-% \author{
-% Matthis Kruse\orcidID{0000-0003-4062-9666}
-% \and
-% Michael Backes
-% \and
-% Marco Patrignani
-% }
-% \authorrunning{M. Kruse et al.}
-
-\maketitle
-
-\begin{abstract}
-% \MPpost{
-%   this abstract fails in setting up the robustness argument.
-%   it'll be a good exercise for you to take inspiration from this, but then write something as concise but with a more direct focus on robust properties
-% }
-% context
-To ensure that secure applications do not leak their secrets, they are required to uphold several security properties such as spatial and temporal memory safety, cryptographic constant time, as well as speculative safety.
-% need
-Existing work shows how to enforce these properties individually, in an architecture-independent way, by using secure compiler passes that each focus on an individual property.
-% task
-Unfortunately, given two secure compiler passes that each preserve a possibly different security property, it is unclear what kind of security property is preserved by the composition of those secure compiler passes.
-%there is no way to tell what kind of security property will the composition of those secure compilers preserve.
-
-% object
-This paper is the first to study what security properties are preserved across the composition of different secure compiler passes.
-% findings
-Starting from a general theory of property composition for security-relevant properties (such as the aforementioned ones), this paper formalises a theory of composition of secure compilers.
-Then, it showcases this theory on a secure multi-pass compiler that preserves the aforementioned security-relevant properties.
-% conclusion
-Crucially, this paper derives the security of the multi-pass compiler from the composition of the security properties preserved by its individual passes, which include security-preserving as well as optimisation passes.
-% 
-From an engineering perspective, this is the desirable approach to building secure compilers.
-% \begin{center}\small\it
-% 	{This paper uses syntax highlighting accessible to both colourblind and black \& white readers.
-% 	% ~\cite{patrignani2020use}.
-% % 	% Specifically, it makes use of a $\src{blue}$, $\src{sans\text{-}serif}$ font for a $\src{source}$,
-% % 	% a $\trg{red}$, $\trg{bold}$ font for an $\trg{intermediate}$,
-% % 	% and a $\obj{green}$, $\obj{teletype}$ font for a $\obj{target}$ language.	
-% 	For a better experience, please print or view in colours.
-% 	}
-% \end{center}
-\end{abstract}
-
-% \begin{IEEEkeywords}
-%   memory-safety, cryptographic constant-time, secure compilation, robustness
-% \end{IEEEkeywords}
-
-\section{Introduction\pages{4}}\label{sec:introduction}
-
-% \gls*{ms} is a security property that ensures that there are no out-of-bounds accesses as well as no reads or writes to uninitialised memory.
-% Programs that do not attain memory safety, can easily have their private data be leaked to attackers or be corrupted~\cite{lemay2021ccc}.
-% Because of this, \gls*{ms} is a requirement for programs to attain an even stronger security property, namely \gls*{cct}.
-% Programs that attain \gls*{cct} are secure w.r.t.
-timing attacks, and this is the gold standard for secure applications, including cryptographic ones, since timing attacks can expose secrets via a simple performance analysis~\cite{kocher1996timing}.
-% Timing attacks have become more and more relevant recently, especially given the era of cloud-computing~\cite{aviram2010cloudtime,kumar2019cloudsecsurvey}, where programs share the same hardware but are supposed to run in a sandboxed environment that does not leak secrets.
-% Unfortunately, several attacks within such a cloud environment have been demonstrated as viable strategies to extract secrets~\cite{mehmet2015getoff,flowers2022zeroday,atya2019catchme,venkatanathan2015placevul}.
-
-% Even though mitigations for such attacks exist~\cite{bond2017vale,almeida2017jasmin}, some compilers may optimise these mitigations away~\cite{barthe2018sec}.
-% This is of course not what programmers expect and to ensure such mitigations persist in the compiled code, one can use a {\em secure} compilers.
-% More generally, secure compilers should preserve security properties even when the program runs in a hostile context such as the cloud one, i.e., where it interacts with other possibly malicious programs.
-% An example of such secure compilers are those that ascribe to the theory of robust compilation~\cite{abate2019jour,abate2021extacc,patrignani2021rsc}, which takes this interaction between components and potentially malicious contexts into account.
-
-% Unfortunately, it is unclear if and how robust compilation supports compositionality, and compilation passes are typically developed as a standalone module to be plugged into a compilation pipeline and composed with other passes.
-% Different kinds of compilation passes exist, e.g., there are usually several compilation passes that perform optimisations~\cite{lattner2004llvm,googlev8,androidstudio,kuepper2023cryptopt,manjikian1997fusion,wegman1991ccp} or source-code instrumentations~\cite{nagarakatte2009soft,nagarakatte2010cets,akritidis2009baggy,dhumbumroong2020boundwarden,jung2021pico,nam2019framer,shankaranarayana2023tailcheck,younan2010paricheck,zhou2023fatptrs,bond2017vale,almeida2017jasmin,kuepper2023cryptopt,cauligi2019fact} that enforce that the compiled program fulfils a property of interest.
-% Without a framework for composing secure compiler passes, it is not possible to enable separation of concerns, e.g., to have a secure compilation pass that ensures \gls*{ms} that is developed independently of another secure pass for \gls*{cct} as suggested before and as showcased in \Cref{ex:strncpy}.
-% \MPpost{
-%   this intro does a poor job at introducing the robust compilation argument.
-%   perhaps it'd be best to start from robust properties, which motivates the use of robust compilers?
-%   idk
-%   as for the abstract, this is a good starting point for you to make a better intro
-% }
-
-\gls*{ms} is a security property obtained by composing \gls*{sms}, which ensures array accesses are all within bounds, and \gls*{tms}, which ensures pointers are only used when they are valid~\cite{azevedo2018meaningsofms,jim2002cyclone,necula2005ccured,nagarakatte2010cets,nagarakatte2009soft,akritidis2009baggy,michael2023mswasm}.
-\gls*{cct} is a security property that ensures sensitive data is not leaked via timing side-channels~\cite{kocher1996timing} and \gls*{ss} is a security property that enforces the same but under a speculative semantics~\cite{guarnieri2018spectector,fabian2022automatic} that captures speculative execution attacks such as spectre~\cite{kocher2019spectre}.
-Together, \gls*{sms}, \gls*{tms}, \gls*{cct} and \gls*{ss}, yield \gls*{specms}, which is the gold standard of security properties for secure applications.
-Programs attaining \gls*{specms} do not leak sensitive data either through erroneous memory accesses, nor through timing side-channels, even under speculative execution.
-\Cref{ex:strncpy} below discusses how these security properties can be enforced by compiler passes~\cite{bond2017vale,almeida2017jasmin}, to ensure programmers need not be aware of the architectural details of where their code will run.
-
-\begin{exampleenv}[strncpy]\label{ex:strncpy}
-Consider the following \texttt{C} function \texttt{strncpy} that copies a null-terminated string \texttt{src} into \texttt{dst} up to a length of \texttt{n} characters.
-This function is subject to a subtle \gls*{sms} vulnerability: the bounds check \texttt{i < n} should happen {\it before} the access to memory location \texttt{src[i]}: otherwise
-% If \texttt{n} is at least as large as the length of the array \texttt{src}, 
-the memory location past the last element is leaked to an attacker.
-\begin{lstlisting}[language=c,basicstyle=\small\ttfamily,morekeywords={size_t}]
-void strncpy(size_t n, char *dst, char *src) {
-  for(size_t i = 0; src[i] != '\0' && i < n; ++i) {
-    dst[i] = src[i];
-  }
-}
-\end{lstlisting}
-% Assume that the length of the array pointed by \texttt{dst} is at least \texttt{n}, and \texttt{src} is terminated by \texttt{'\textbackslash 0'}.
-% Even under these assumptions, there is a 
-To prevent this vulnerability, one can use a compilation pass that enforces \gls*{sms}, such as Softbounds~\cite{nagarakatte2009soft} or BaggyBounds~\cite{akritidis2009baggy}.
-The most na\"ive solution in this case is to insert bounds-checks in front of every access to memory.
-
-Because of timing attacks, addressing \gls*{sms} is not enough to make \texttt{strncpy} secure.
-In fact, the loop can terminate early, as soon as the string-terminating character \texttt{'\textbackslash 0'} is encountered, thus making program execution time proportional to the length of the array pointed by \texttt{src}, and violating the \gls*{cct} property.
-% Often, programmers rewrite programs like these to have loops that run with a fixed number of iterations.
-Also in this case there exist compiler passes that can rewrite such programs into \gls*{cct} ones~\cite{cauligi2019fact}.
- % so that they are no longer vulnerable to timing attacks.
-%
-Finally, even with these precautions, code is not run in isolation, so a malicious attacker could supply code that interacts with \texttt{strncpy} and trigger a violation of either \gls*{ms} or \gls*{cct} by calling \texttt{strncpy} with an argument for \texttt{src} that points to uninitialised memory.
-This would, in turn, trigger a series of reads from uninitialised memory, which is an immediate \gls*{ms} violation with devastating real-world consequences~\cite{uninit-0,uninit-1,uninit-2,uninit-3,uninit-4}.
-% An easy fix to sidestep this is to use non-nullable pointers as parameters instead, e.g., as presented in Cyclone~\cite{jim2002cyclone}.
-\end{exampleenv}
-
-  % last sentence is before the example, which talks about compiler passes enforcing props
-Whether or not compiler passes enforce certain security properties, it is important for a secure compiler to consider partial programs that interact with potentially malicious code, since the latter may lead to, e.g., memory-safety issues in the considered partial program.
-Robust compilers~\cite{abate2019jour} are a form of secure compilers that preserve security properties even in the presence of arbitrary attackers interacting with compiled code.
-Thus, robust compilers can be used to prevent vulnerabilities resulting from uninitialised memory (as well as many other ones), e.g., by targeting capability-based languages such as CHERI~\cite{woodruff2014CHERI}, Arm Morello~\cite{arm-morello}, or MSWasm~\cite{michael2023mswasm}, where the compiler relies on capabilities to check that pointers are always initialised.
-
-% To ensure these compiler-preserved properties still hold even in the presence of arbitrary attackers interacting with compiled code, one can use secure compilers in the form of robust compilers~\cite{abate2019jour,abate2021extacc,patrignani2021rsc}.
-
-Unfortunately, given secure compiler passes that each preserve a possibly different security property, there is no way to tell what kind of security property will the composition of those secure compilers preserve.
-% Different kinds of compilation passes exist, e.g., there are usually several compilation passes that perform optimisations~\cite{lattner2004llvm,googlev8,androidstudio,kuepper2023cryptopt,manjikian1997fusion,wegman1991ccp} or source-code instrumentations~\cite{nagarakatte2009soft,nagarakatte2010cets,akritidis2009baggy,dhumbumroong2020boundwarden,jung2021pico,nam2019framer,shankaranarayana2023tailcheck,younan2010paricheck,zhou2023fatptrs,bond2017vale,almeida2017jasmin,kuepper2023cryptopt,cauligi2019fact} that enforce that the compiled program fulfils a property of interest.
-Worse, without a framework for composing secure compiler passes, it is not possible to enable separation of concerns, e.g., to have a secure compilation pass that ensures \gls*{ms} that is developed independently of another secure pass for \gls*{cct}, that is developed independently of other passes, such as optimisation ones.
-
-This paper introduces a framework for reasoning about the composition of secure and optimising compiler passes %akin to those of \Cref{ex:strncpy}\MP{there aren't really passes there} 
-and it showcases the power of this framework by instantiating it on a multi-pass compilation chain.
-To this end, this paper first discusses how to compose security properties, such as \gls*{tms} and \gls*{sms} into \gls*{ms}, and then adds \gls*{cct} as well as \gls*{ss} into the mix to obtain \gls*{specms}.
-%Then, this paper defines compiler composition and formalises that given two passes that securely preserve two (possibly distinct) properties, their composition securely preserves the composition of those properties.
-The paper then defines several secure compiler passes, where each is either preserving a different security property from the list above
-% (\gls*{tms}, \gls*{sms}, \gls*{scct}, \gls*{ss}) 
-or performing a security-preserving optimisation.
-% , (e.g., applying \gls*{cf} or \gls*{dce}).
-Finally, this paper shows that composing these secure compiler passes into a multi-pass compilation chain results in the end-to-end (robust) preservation of \gls*{specms}.
-Crucially, this paper derives the security of the multi-pass compiler from the composition of the security properties preserved by its individual passes.
-% Crucially, this end-to-end security preservation is obtained by composing the security property preserved by each individual pass, as dictated by the framework of this paper.
-This result showcases how the framework allows the kind of formal security reasoning that compiler writers already want (and already do), obtaining precise, compositional security reasoning while providing minimal (and modular) proof effort.
-%
-In summary, this paper makes the following contributions:
-% \begin{itemize}[leftmargin=0em,nosep,label=$\blacktriangleright$]
-\begin{asparaitem}[$\blacktriangleright$]
-  \item %
-        %\Cref{sec:sequential} studies different forms of secure compiler compositions.
-        This paper takes the secure compilation framework of Abate et al.~\cite{abate2021extacc} and extends it for compositionality. (\Cref{sec:sequential}).
-        %Concretely, it investigates upper and lower compositions, as well as the sequential composition.
-        %For this, this paper studies compiler composition as well as compilers with multiple input languages or multiple output ones, as used in existing compilation chains.
-        %
-        This paper proves that starting from two compilers that preserve two (possibly distinct) properties, their composition preserves roughly the intersection of those properties.
-        % For example, a compiler that is secure w.r.t. \gls*{tms} can be composed with a compiler that is secure w.r.t. \gls*{sms}, resulting in a compiler that is secure w.r.t. \gls*{ms}.
-        % 
-        % From this, the paper proves that the order of composition of sequential compiler passes is irrelevant for security.
-        % % 
-        % This is crucial for reordering individual optimisation passes and thus generating secure and efficient code.
-        %
-        %Finally, this paper extends parts of the compositionality framework to support different trace-models between compilation passes.
-        % \MP{this isn't really true, is it?. the order matters for the relations' lemmas}
-        Then, this paper identifies which conditions make the composition of secure compilers meaningful from a security perspective, and which other conditions allow passes to be swapped without losing security meaning.
-
-  \item %
-        This paper presents a case-study with five programming languages showcasing the previous contribution (\Cref{sec:casestud:defs,sec:casestud:rtp}).
-        To this end, it presents a compilation chain consisting of six passes that ultimately preserve \gls*{specms} by means of composing four secure passes that \emph{individually} preserve \gls*{tms}, \gls*{sms}, \gls*{scct}---an enforceable version of \gls*{cct}---, and \gls*{ss}, and two passes which are well-known optimisations \gls*{dce} and \gls*{cf}.
-        %Furthermore, the chain includes two optimisation passes: one performs \gls*{dce} and the other \gls*{cf}.
-        The formalisation of this case study showcases the power of the presented framework: the divide-and-conquer approach to software engineering is a viable strategy even for the development of secure compilers.
-%        This demonstrates that the usage of the presented framework is both simpler and more expressive compared to one monolithic security proof of the whole compilation chain.
-
-  \item %
-  		This paper provides formal proofs of the security properties preserved by each of the presented passes. 
-  		% 
-  		Additionally this paper analyses each pass in terms of their compatibility with each other (\Cref{sec:formalities}), proving how does each pass fulfill the conditions that make their composition meaningful from a security standpoint.
-        % In doing so, the paper provides formal insights into some technicalities of the concrete proofs necessary as technical setup to reap the benefits of the presented compositionality framework.
-
-  \item The key contributions of this paper are formalised in the Coq proof assistant and the paper indicates this with \CoqSymbol.
-\end{asparaitem}
-% 
-This paper starts by introducing relevant notions of security properties and secure compilation (\Cref{sec:background}),
-and discusses related work (\Cref{sec:relwork}) before concluding (\Cref{sec:concl}).
-
-% \contribution{Technical Report} 
-The omitted formal details, lemmas and proofs, as well as the Coq formalisation are available as supplementary material.
-% \MKin{cite/link}
-
-\section{Background: Properties and Secure Compilers\pages{1}}\label{sec:background}
-
-To introduce the security argument of this paper, this section defines (security) properties, their satisfaction, and their robust satisfaction (i.e., satisfaction in the presence of an active attacker; \Cref{subsec:bg:tprop}).
-Then, borrowing from existing work~\cite{abate2019jour,abate2021extacc}, this section introduces secure compilers as compilers that preserve robust property satisfaction (\Cref{subsec:bg:rtp}).
-
-\subsection{Properties and (Robust) Satisfaction}\label{subsec:bg:tprop}
-
-This paper employs the security model where programs are written in a language whose semantics emits events $\event$.
-Events include security-relevant actions (e.g., reading from and writing to memory, as detailed in \Cref{sec:compprop}) and the unobservable event $\emptyevent$.
-As programs execute, their emitted events are concatenated in traces $\trace$, which serve as the description of the behaviour of a program.%
-\footnote{
-Throughout the paper, sequences are indicated with an overbar (i.e., $\trace$), empty sequences with $\hole{\cdot}$, and concatenation of sequences $\trace[_{1}],\trace[_{2}]$ as $\trace[_{1}]\cdot\trace[_{2}]$.
-Prepending element $\event$ to a sequence $\trace$ uses the same notation: $\event\cdot\trace$.
-}
-
-Properties $\pi$ are sets of traces of admissible program behaviours, ascribing what said property considers valid.
-% 
-The set of all properties can be partitioned into different {\em classes} ($\class$), i.e., safety, liveness, and neither safety nor liveness~\cite{clarkson2008hyper}, so a class is simply a set of properties.
-% 
-The compositionality framework (\Cref{sec:sequential}) presented in this paper considers arbitrary classes, while the case-study (\Cref{sec:casestud:rtp}) fixes them to concrete instances of safety properties, since it is decidable whether a trace satisfies a safety property with just a finite trace (i.e., a \emph{prefix}).
-As an example, consider the trace:
-$$\ev{Dealloc\ \loc}\cdot\ev{Read\ \loc\ 1729}\cdot\dots$$
-which describes the interaction with a memory where the deallocation of an address $\loc$ precedes a read (of some value $1729$) at that address in memory.
-% 
-This program behaviour is insecure w.r.t.
-a canonical notion of (temporal) memory safety dictating no use-after-frees of pointers~\cite{nagarakatte2010cets,azevedo2018meaningsofms}, because it reads from a memory location that was freed already.
-The previous finite trace prefix is enough to decide that the trace does not satisfy \gls*{tms} and there is no way to append events to this prefix which would result in the trace being admissible.
-
-In the following, the execution of a whole program $\wholeprogvar$ that terminates in state $\runtimetermvar$ according to the language semantics and produces trace $\trace$ is written as $\progstepto{\wholeprogvar}{\runtimetermvar}{\trace}$.
-With this, we defined property satisfaction as follows:
-\bul{a whole program $\wholeprogvar$ satisfies a property $\pi$} iff \iul{$\wholeprogvar$ yields a trace $\trace$} such that \oul{$\trace$ satisfies $\pi$}.
-
-\begin{definition}[Property Satisfaction]\label{def:propsat}
-$\;$
-
-  \begin{nscenter}
-    \bul{$\sat{\wholeprogvar}{\pi}$}
-    $\isdef$
-    if \iul{$\forall\runtimetermvar\ \trace,\progstepto{\wholeprogvar}{\runtimetermvar}{\trace}$},
-    then \oul{$\trace\in\pi$}.
-  \end{nscenter}
-\end{definition}
-
-Property satisfaction is defined on whole programs, i.e., programs without missing definitions.
-Thus, from a security perspective, \Cref{def:propsat} considers only a passive attacker model, where the attacker observes the execution and, e.g., retrieves secrets from that.
-To consider a stronger model similarly to what existing work does~\cite{abate2019jour,abate2021extacc,maffeis2008code-carrying,gordon2003authenticity,fournet2007authorization,bengtson2011refine,backes2014uniontyps,michael2023mswasm,swasey2017robust,sammler2019benefits}, we extend the concept of satisfaction with {\em robustness}.
-Robust satisfaction considers partial programs $\progvar$, i.e., components with missing imports, which cannot run until said imports are fulfilled.
-To remedy this, {\em linking} takes two partial programs $\progvar[_{1}],\progvar[_{2}]$ and produces a whole program $\wholeprogvar$, i.e., $\link{\progvar[_{1}]}{\progvar[_{2}]}=\wholeprogvar$.
-As typically done in works that consider the execution of partial programs~\cite{abate2019jour,devriese2018parametricity,patrignani2021rsc,korashy2021capableptrs,strydonck2019lincap,devriese2017modular,bowman2015noninterference,ahmed2011equivcps,patterson2017linkingtyps},
-this paper assumes that whole programs are the result of linking partial programs referred to as {\em context} ($\ctx$) and {\em component} ($\comp$).
-The context is an arbitrary program and thus has the role of an {\em attacker} that can interact with the component by means of any features the programming language has, and the component is what is security-relevant.
-%In this work, the semantics of the programming language is expected to differentiate between the component and the context.
-With this, \Thmref{def:propsat} can be extended as follows: for \bul{a component $\progvar$ to robustly satisfy a property $\pi$}, take an \iul{attacker context $\contextvar$ and link it with $\progvar$}, \oul{the resulting whole program must satisfy $\pi$}.
-
-\begin{definition}[Robust Satisfaction]\label{def:proprsat}
-  $\;$
-
-  \begin{nscenter}
-  \bul{$\rsat{\progvar}{\pi}$}
-  $\isdef$ \iul{$\forall \contextvar\ \wholeprogvar$, if $\link{\contextvar}{\progvar}=\wholeprogvar$}, then \oul{$\sat{\wholeprogvar}{\pi}$}.
-  \end{nscenter}
-\end{definition}
-
-% \begin{exampleenv}[Double Free in Bluetooth Subsystem]
-%   Consider \texttt{CVE-2021-3564}~\cite{doublefree-bluetooth}, one of many submissions for a double-free vulnerability.
-%   The vulnerability arises due to a race condition where the context-level function \texttt{hci\_cmd\_work} was not expected to behave maliciously, since it resides in the same source-code repository where the vulnerability occurs.
-%   Nevertheless, the component-level code of \texttt{hci\_dev\_do\_open} is linked with \texttt{hci\_cmd\_work} and does not atomically check whether a pointer has been freed already:
-%   Therefore, \texttt{hci\_dev\_do\_open} does not satisfy the no-double-frees property robustly, since there is an implementation for \texttt{hci\_cmd\_work} that leads to a violation of that property when linked with \texttt{hci\_dev\_do\_open}.
-% \end{exampleenv}
-
-\subsection{Secure Compilers}\label{subsec:bg:rtp}
-
-A {\em compiler} ($\cc{\src{L}}{\trg{L}}$) translates syntactic descriptions of programs from a {\em source} ($\src{L}$) into a {\em target} ($\trg{L}$) programming language.
-This translation is considered {\em correct} if it is semantics-preserving.
-That is, for a whole program $\src{\wholeprogvar}$, the compiler should relate the $\src{L}$ semantics of $\src{\wholeprogvar}$ with the semantics of $\trg{L}$ of the compiled counterpart of $\src{p}$ in such a way that they are ``compatible''.
-Unfortunately, correct compilers may be insecure compilers~\cite{patrignani2019survey,kennedy2006secure.net,abadi1999protect,ahmed2018dagstuhl} and programs translated by insecure compilers can violate security properties that the programmer assumes to hold.
-This is why {\em robust preservation} is a good candidate as a compiler-level security property~\cite{abate2019jour}.
-
-This paper uses a general notion of robust preservation~\cite{abate2021extacc} that considers compilers that use languages with different trace models. 
-To this end, considering a source trace $\src{\trace}$ and a target trace $\trg{\trace}$, there is a relation ($\sim$) describing the effect of a corresponding compiler (see \Cref{sec:casestud:rtp}). 
-This relation induces the following two projection functions~\cite{abate2021extacc}: (1) the \emph{existential image} $\tau_\sim\left(\src{\pi}\right)$ and (2) the \emph{universal image} $\sigma_\sim\left(\trg{\pi}\right)$.
-These projections map source-level (resp. target-level) properties to target-level (resp. source-level) properties in a way that identifies the ``same'' property across languages. 
-The case study of this paper uses the universal image, since some considered properties, such as \gls*{ss}, are not definable in a higher-level language that, e.g., does not model speculation.
-% For self-containment, the universal image is defined as follows~\cite{abate2021extacc}:
-% 
-\begin{definition}[Universal Image]
-% [Existential and Universal Image]
-\label{def:universal:img}\label{def:sigma}
-  % \[
-  %   \tau_\sim\left(\src{\pi}\right) := 
-  %     \left\{ 
-  %       \trg{\trace} \mid \exists \src{\trace}\ldotp \src{\trace}\sim\trg{\trace}, \text{ and } \src{\trace}\in\src{\pi} 
-  %     \right\}
-  % \]
-  \[ 
-    \sigma_\sim\left(\trg{\pi}\right) := 
-      \left\{ 
-        \src{\trace} \mid \forall \trg{\trace}\ldotp \text{if }\src{\trace}\sim\trg{\trace}, \text{ then } \trg{\trace}\in\trg{\pi} 
-      \right\}
-  \]
-\end{definition}
-\noindent
-With this projection function, we define a more general version of robust preservation as follows~\cite{abate2021extacc}.
-% For a \bul{compiler $\cc{\src{L}}{\trg{L}}$ to robustly preserve a class of source properties $\src{\class}$}, if for any \rul{property $\src{\pi}$ of that class $\src{\class}$ and programs $\src{p}$ written in $\src{L}$} where \iul{$\src{\progvar}$ robustly satisfies $\src{\pi}$}, then \oul{the compilation of $\src{\progvar}$, $\cc{\src{L}}{\trg{L}}\left(\src{p}\right)$, must robustly satisfy $\tau_\sim\left(\src{\pi}\right)$}.
-% 
-% \begin{definition}[Robust Preservation with $\tau_\sim$]\label{def:rtp:tau}
-%   % $\;$\\
-%   % \vspace{-1em}
-%   % \begin{nscenter}
-%   % Compiler $\cc{\src{L}}{\trg{L}}$ robustly preserves $\class$, 
-%   \bul{$\rtptau{\cc{\src{L}}{\trg{L}}}{\src{\class}}{\sim}$}
-%   %, iff 
-%   $\isdef$
-%   \rul{$\forall \src{\pi}\in\src{\class}, \src{p}\in\src{L},$} if \iul{$\rsat{\src{\progvar}}{\src{\pi}}$}, then \oul{$\rsat{\cc{\src{L}}{\trg{L}}\left(\src{p}\right)}{\tau_\sim{\left(\src{\pi}\right)}}$}.
-%   % \end{nscenter}
-% \end{definition}
-% 
-A \bul{compiler $\cc{\src{L}}{\trg{L}}$ robustly preserves a class of target properties $\trg{\class}$}, if for any \rul{property $\trg{\pi}$ of class $\trg{\class}$ and programs $\src{p}$}, where \iul{$\src{\progvar}$ robustly satisfies $\sigma_\sim\left(\trg{\pi}\right)$}, \oul{the compilation of $\src{\progvar}$, we have that $\cc{\src{L}}{\trg{L}}\left(\src{p}\right)$ robustly satisfies $\trg{\pi}$}.
-
-\begin{definition}[Robust Preservation with $\sigma_\sim$]\label{def:rtp:sigma}
-  % $\;$\\
-  % \vspace{-1em}
-  % \begin{nscenter}
-  % Compiler $\cc{\src{L}}{\trg{L}}$ robustly preserves $\class$, 
-  \bul{$\rtpsig{\cc{\src{L}}{\trg{L}}}{\trg{\class}}{\sim}$}
-  %, iff 
-  $\isdef$
-  \rul{$\forall \trg{\pi}\in\trg{\class}, \src{p}\in\src{L},$} if \iul{$\rsat{\src{\progvar}}{\sigma_\sim\left(\trg{\pi}\right)}$}, then \oul{$\rsat{\cc{\src{L}}{\trg{L}}\left(\src{p}\right)}{\trg{\pi}}$}.
-  % \end{nscenter}
-\end{definition}
-Note that a class of properties $\class$ can represent just one property $\pi$ by lifting~\cite{clarkson2008hyper} that property to sets of properties, i.e., use the powerset of $\pi$ instead of $\pi$ itself.
-Because of this, this paper may write $\rtptau{\cc{\src{L}}{\trg{L}}}{\trg{\pi}}{\sim}$, even though $\trg{\pi}$ is a property and not a class.
-A similar construction can be used to the projection functions (see \Cref{def:universal:img}) by applying them to the lifted version of $\trg{\pi}$ instead of just $\trg{\pi}$.
-
-In case the trace model is the same for both source and target programs (and thus $\sim$ is equality), we obtain~\cite{abate2019jour}:
-\begin{definition}[Robust Preservation]\label{def:rtp}
-  % $\;$\\
-  % \vspace{-1em}
-  % \begin{nscenter}
-  % Compiler $\cc{\src{L}}{\trg{L}}$ robustly preserves $\class$, 
-  $\;$
-
-  {$\rtp{\cc{\src{L}}{\trg{L}}}{\class}$}
-  %, iff 
-  $\isdef$
-  {$\forall \pi\in\class, \src{p}\in\src{L},$} if {$\rsat{\src{\progvar}}{\pi}$}, then {$\rsat{\cc{\src{L}}{\trg{L}}\left(\src{p}\right)}{\pi}$}.
-  % \end{nscenter}
-\end{definition}
-
-Examples of compilers fulfilling \Cref{def:rtp} exist in the literature~\cite{korashy2022secureptrs,korashy2021capableptrs,abate2021extacc,abate2019jour,patrignani2021rsc}.
-For example, SecurePtrs~\cite{korashy2022secureptrs} gives a compiler that robustly preserves all safety properties for a C-like language to an assembly-like language. 
-% The mechanization has well over 20k lines in total, which demonstrates that proving a compiler secure with respect to \Cref{def:rtp} is a significant effort. \MP{move this line to later ?}
-% i don't see how adding a def of cct is beneficial
-As another example, even though it is not strictly satisfying \Cref{def:rtp}, the FaCT~\cite{cauligi2019fact} compiler preserves the \gls*{cct} property for a C-like language with constant-time primitives, e.g., \texttt{ctselect} for branching.
-%, whose precise definition can be found in \Cref{sec:cct}.
-Throughout this work, it is assumed that FaCT satisfies \Cref{def:rtp}.
-% While we were not able to find compilers in literature that provably fulfill \Cref{def:rtp}, we anticipate that 
- % would satisfy this hyperproperty.
-% FaCT proves constant-timeness by means of overapproximation, i.e., instead of directly proving the hyperproperty, they fall back to a property that overapproximates the corresponding hyperproperty.
-
-%\smallskip%MP: the underline compacts the sections too much
-
-\section{Secure Composition}\label{sec:sequential}
-
-Notably, real-world compilation chains also perform a series of (sequential) passes whose main purpose is not necessarily to translate from one language to another, but to, e.g., optimise the code or enforce a certain property.
-Both examples can be seen in practice, e.g.,~\cite{nagarakatte2009soft,nagarakatte2010cets,akritidis2009baggy,wegman1991ccp,manjikian1997fusion} and many more.
-% \MPin{survey to cite?
-this is a bit informal MK: hard to find a survey on different compiler passes...
-I can't even find one just for compiler enforced memsafety (that looks good) or for compiler translations in general}.
-% 
-% Consider the compilation passes of LLVM~\cite{lattner2004llvm}, which are handled in the \texttt{PassManager} class.
-% By manipulating the internal state of a \texttt{PassManager} object, the schedule of compiler optimisation passes can be hardcoded.
-% That way, an order of optimisations can be found by hand that yields the best results.
-% \MP{what's the point of the previous 3 sentences?(commented)}
-Consider the following two LLVM optimisation passes: \gls*{cf}, which rewrites constant expressions to the constant they evaluate to, and \gls*{dce}, which removes dead code by rewriting conditional branches.
-The order in which \gls*{cf} and \gls*{dce} are performed influences the final result of the compilation (see \Cref{fig:cfdceex}).
-\begin{figure}[!ht]
-  \begin{center}
-    \begin{tikzpicture}[scale=0.62, every node/.style={transform shape}]
-      \node[minimum width=3cm,minimum height=2.5cm,draw=black,very thick,align=left] (progunopt) {\begin{lstlisting}[language=ml,basicstyle=\small\ttfamily]
-let a = true in
-if a then
-  print "a"
-else
-  print "b"
-\end{lstlisting}};
-      \node[minimum width=3cm,minimum height=2.5cm,draw=black,very thick,align=left,right=1.0 of progunopt] (progoptcf) {\begin{lstlisting}[language=ml,basicstyle=\large\ttfamily]
-if true then
-  print "a"
-else
-  print "b"
-\end{lstlisting}};
-      \node[minimum width=3cm,minimum height=2.5cm,draw=black,very thick,align=left,right=1.0 of progoptcf] (progoptcfdce) {\begin{lstlisting}[language=ml,basicstyle=\Large\ttfamily]
-print "a"
-\end{lstlisting}};% the following is a hack to get my syntax highlighting to work again \end{lstlisting}
-      % edges
-      \draw[->,very thick] (progunopt) edge[loop left,left] node {\gls*{dce}} (progunopt);
-      \draw[->,very thick] (progunopt) edge[sloped,above] node {\gls*{cf}} (progoptcf);
-      \draw[->,very thick] (progoptcf) edge[sloped,above] node {\gls*{dce}} (progoptcfdce);
-    \end{tikzpicture}
-  \end{center}
-  \caption{Example program where the level of optimisations differ for one pass of applying \gls*{cf} and \gls*{dce} in any order. %
-  Every edge is a compilation pass and the label on the edge states what the pass does, i.e., \gls*{cf} or \gls*{dce}. %
-  The source code in the nodes is a glorified compiler intermediate representation and the code gets more optimised towards the right hand side of the figure.}\label{fig:cfdceex}
-\end{figure}
-This {\em phase ordering problem} is well--known in literature and a practical solution is to simply perform a fixpoint iteration of the optimisation pipeline~\cite{click1995combining}.
-
-\smallskip
-
-To analyse the security of compilation passes and their interaction within a compilation pipeline, we rely on a few key notions: the definition of a trace relation, and the definition of when is a trace relation well-formed with respect to a class.
-%
-Consider traces $\src{\trace}$ and $\trg{\trace}$ as well as two trace relations $\sim_1$ and $\sim_2$. 
-% 
-The traces are related $\src{\trace}\sim_1\bullet\sim_2\trg{\trace}$ if there is another trace $\irl{\trace}$ such that $\src{\trace}\sim_1\irl{\trace}$ and $\irl{\trace}\sim_2\trg{\trace}$.
-% 
-\bul{A relation $\sim$ is well-formed w.r.t. a class of target-level properties $\trg{\class}$} iff \rul{the universal image preserves set membership}.
-\begin{definition}[Well-formedness of $\sim$ for a Class $\trg{\class}$]\label{def:wfc:sig:tracerel}
-  % $$
-
-  \begin{nscenter}
-  \noindent
-  \text{\bul{$\wfcsig{\sim}{\trg{\class}}$}} \isdef \text{\rul{$\forall \trg{\pi}\in\trg{\class}, \sigma_\sim(\trg{\pi})\in\sigma_\sim(\trg{\class})$}}
-  \end{nscenter}
-  % $$
-\end{definition}
-
-We can now state our main result: secure compilers in the robust compilation framework~\cite{abate2021extacc} compose {\em sequentially}. 
-%
-
-Let \bul{$\cc{\src{L}}{\trg{L}}$ robustly preserve the class $\sigma_{\sim_2}\left(\trg{\class[_1]}\right)$ under $\sim_1$} and let \rul{$\cc{\trg{L}}{\obj{L}}$ robustly preserve the class $\trg{\class[_2]}$ under $\sim_2$}.
-Then, \iul{when the cross-language relation $\sim_2$ is well-formed w.r.t.
-class $\trg{\class[_1]}$}, it follows that \oul{the composed compiler $\cc{\src{L}}{\trg{L}}\circ\cc{\trg{L}}{\obj{L}}$ robustly preserves the intersection of classes $\trg{\class[_1]}\cap\trg{\class[_2]}$ under $\sim_1\bullet\sim_2$}.
-% 
-%For space reasons, \Thmref{thm:urtp} and \Thmref{thm:lrtp} are not presented here, but can be found in the technical appendix ($\CoqSymbol$).
-% Note that it is not necessary to require $\sim_2$ to be well-formed w.r.t. $\src{\class[_1]}$, since the 
-% \MPin{why}
-\begin{theorem}[Composition of Secure Compilers w.r.t. $\sigma$]\label{thm:rtpsim:sig}
-  $\;$
-
-  If \bul{$\rtpsig{\cc{\src{L}}{\trg{L}}}{\tilde{\sigma}_{\sim_2}\left(\trg{\class[_{1}]}\right)}{\sim_1}$}, \rul{$\rtpsig{\cc{\trg{L}}{\obj{L}}}{\trg{\class[_2]}}{\sim_2}$}, and \iul{$\wfcsig{\sim_2}{\trg{\class[_1]}}$}, \\ then \oul{$\rtpsig{\cc{\src{L}}{\trg{L}}\circ\cc{\trg{L}}{\obj{L}}}{\trg{\class[_{1}]}\cap\trg{\class[_{2}]}}{\sim_1\bullet\sim_2}$}.
-\Coqed
-\end{theorem}
-
-Since the composition of secure compilers is again a secure compiler, the theorems generalise to a whole chain of $n$ secure compilers. 
-% 
-\Cref{thm:rtpsim:sig} can also be stated for the existential image $\tau_\sim\left(\src{\pi}\right)$, but in the interest of space that result has been moved to the appendix.
- % (\Cref{subsec:extimg}).
-% 
-Crucially, \Cref{thm:rtpsim:sig} also holds for \emph{classes of hyperproperties}, and thus, compilers that robustly preserve hyperproperties can be composed with each other as well as with compilers that robustly preserve properties.
-% 
-%Unfortunately, proofs of compilers that robustly preserve hyperproperties (including hypersafety) are scarce in the literature due to their complexity.
-%\MP{ so what ? }
-%  we don't discuss this further?
-%
-
-% It is furthermore possible to conflate the notion of classes and hyperproperties by intersecting all elements of a class of hyperproperties, to obtain one combined hyperproperty.
-% This idea has been discussed in other work~\cite{clarkson2008hyper}.
-%  MP: i think i stated the point of these 2 lines above
-
-If we take SecurePtrs and FaCT from \Cref{subsec:bg:rtp} and compose them according to \Cref{thm:rtpsim:sig}, we obtain a compiler that robustly preserves the intersection of safety properties and the \gls*{cct} hyperproperty.
-That is, for a source component that robustly satisfies any set of safety properties {\em and} \gls*{cct}, the compiled target component also robustly satisfies the same set of safety properties and \gls*{cct}.
-% The intersection simply uses the lifted version of the respective safety properties.
-%\MP{meaning?}
-
-Compiler engineers typically try to find an order of optimisations that yields well-optimised programs for either code size~\cite{cooper1999geneticphases} or performance~\cite{kulkarni2006exhaustivephase}.
-\Cref{corr:swappable} justifies that any such order of compilation passes is valid w.r.t. security, as long as the trace-relations have no effect on the respective classes.
-
-So, given two compilation passes \bul{$\cc[_{1}]{\trg{L}}{\trg{L}}$, $\cc[_{2}]{\trg{L}}{\trg{L}}$, both robustly preserving class $\trg{\class[_{1}]}$ or $\trg{\class[_{2}]}$, respectively}, \rul{their corresponding well-formed trace-relations}, \iul{and indifference of the classes with respect to these trace relations}, \oul{for any order of their composition, the composed compiler robustly preserves the intersection of $\trg{\class[_{1}]}$ and $\trg{\class[_{2}]}$}.
-% 
-% \MPin{
-% 	later on, say that optimisation passes remove elements from traces, so they should be fine w.r.t. the result with $\tau$.
-% 	also, yes, if $\tau$ adds we have a problem, if it removes we don't
-% \MKin{ Optimisations that remove elements from traces are not necessarily sound,
-%  e.g., remove all deallocations to "optimise" memory management by offloading
-%  the work to the OS.
-% }
-% }
-
-\begin{corollary}[Swapping Secure Compiler Passes]\label{corr:swappable}
-  $\;$
-
-  If \bul{$\rtpsig{\cc[_{1}]{\trg{L}}{\trg{L}}}{\trg{\class[_{1}]}}{\sim_1}$ and $\rtpsig{\cc[_{2}]{\trg{L}}{\trg{L}}}{\trg{\class[_{2}]}}{\sim_2}$}, %
-  \rul{$\wfcsig{\sim_1}{\trg{\class[_2]}}$ and $\wfcsig{\sim_2}{\trg{\class[_1]}}$}, %
-  and \iul{$\tilde{\sigma}_{\sim_2}\left(\trg{\class[_1]}\right)=\trg{\class[_1]}$ as well as $\tilde{\sigma}_{\sim_1}\left(\trg{\class[_2]}\right)=\trg{\class[_2]}$},
-  then \oul{$\rtpsig{\cc[_{1}]{\trg{L}}{\trg{L}}\circ\cc[_{2}]{\trg{L}}{\trg{L}}}{\trg{\class[_{1}]}\cap\trg{\class[_{2}]}}{\sim_1\circ\sim_2}$ and $\rtpsig{\cc[_{2}]{\trg{L}}{\trg{L}}\circ\cc[_{1}]{\trg{L}}{\trg{L}}}{\trg{\class[_{2}]}\cap\trg{\class[_{1}]}}{\sim_2\circ\sim_1}$}.
-\Coqed
-\end{corollary}
-
-Coming back to the example composing SecurePtrs with FaCT, it is likely the case that \Cref{corr:swappable} is not applicable.
-While first running SecurePtrs and then FaCT should be fine, the other direction has potential security hazards, since the SecurePtrs compiler does not account for cryptographic-constant time primitives, such as \texttt{ctselect}.
-
-\subsection{Secure Compiler Composition with Same Trace Models}
-When the cross-language trace relation is an equality, \Cref{thm:rtpsim:sig} collapses:
-Given \bul{$\cc{\src{L}}{\trg{L}}$ robustly preserves $\class[_{1}]$} and \rul{$\cc{\trg{L}}{\obj{L}}$ robustly preserves $\class[_{2}]$}, it follows that \oul{their sequential composition $\cc{\src{L}}{\trg{L}}\circ\cc{\trg{L}}{\obj{L}}$ robustly preserves the intersection of classes $\class[_{1}]$ and $\class[_{2}]$}.
-
-\begin{corollary}[Composition of Secure Compilers]\label{corr:rtp}
-  $\;$
-
-  If \bul{$\rtp{\cc{\src{L}}{\trg{L}}}{\class[_{1}]}$} and \rul{$\rtp{\cc{\trg{L}}{\obj{L}}}{\class[_{2}]}$}, then \oul{$\rtp{\cc{\src{L}}{\trg{L}}\circ\cc{\trg{L}}{\obj{L}}}{\class[_{1}]\cap\class[_{2}]}$}.
-\Coqed
-\end{corollary}
-
-\Cref{corr:rtp} provides an easy way to compose secure compilers without well-formedness of trace relations. 
-% 
-However, while \Cref{thm:rtpsim:sig} explicitly requires that $\sim_1$ is well-formed w.r.t. $\trg{\class[_2]}$, if $\sim_2$ is not well-formed w.r.t. $\trg{\class[_1]}$, care must be taken. 
-%Worst-case is that the resulting class $\sigma_{\sim_1\bullet\sim_2}\left(\trg{\class[_1]}\cap\trg{\class[_2]}\right)$ becomes empty, even though $\trg{\class[_1]}\cap\trg{\class[_2]}$ is not necessarily empty.
-This is further discussed in \Cref{subsec:compatsecpasses}.
-
-%If possible, optimisations are merged to enhance the analysis' result, since even when iterating to a fixpoint the final program may not be as optimised as it could be~\cite{click1995combining}.
-We can also obtain a specialised version of \Cref{corr:swappable}:
-%So, given two compilation passes \bul{$\cc[_{1}]{\trg{L}}{\trg{L}}$, $\cc[_{2}]{\trg{L}}{\trg{L}}$, both robustly preserving class $\class[_{1}]$ or $\class[_{2}]$, respectively}, \oul{for any order of their composition, the composed compiler robustly preserves the intersection of $\class[_{1}]$ and $\class[_{2}]$}.
-% 
-% \MPin{
-% 	later on, say that optimisation passes remove elements from traces, so they should be fine w.r.t. the result with $\tau$.
-% 	also, yes, if $\tau$ adds we have a problem, if it removes we don't
-% \MKin{ Optimisations that remove elements from traces are not necessarily sound,
-%  e.g., remove all deallocations to "optimise" memory management by offloading
-%  the work to the OS.
-% }
-% }
-
-\begin{corollary}[Swapping Secure Compiler Passes]\label{corr:swappable:one}
-  $\;$
-
-  If {$\rtp{\cc[_{1}]{\trg{L}}{\trg{L}}}{\class[_{1}]}$ and $\rtp{\cc[_{2}]{\trg{L}}{\trg{L}}}{\class[_{2}]}$}, then {$\rtp{\cc[_{1}]{\trg{L}}{\trg{L}}\circ\cc[_{2}]{\trg{L}}{\trg{L}}}{\class[_{1}]\cap\class[_{2}]}$ and $\rtp{\cc[_{2}]{\trg{L}}{\trg{L}}\circ\cc[_{1}]{\trg{L}}{\trg{L}}}{\class[_{2}]\cap\class[_{1}]}$}.
-\Coqed
-\end{corollary}
-
 \section{Security Properties Formalisation \& Composition\pages{1.5}}\label{sec:compprop}
@@
 \end{definition}
-
-\section{Case Study: Language Formalisations\pages{2}}\label{sec:casestud:defs}
-
-This section defines programming languages $\Ltms$, $\Ltrg$, $\Lms$, $\Lscct$, and $\Lspec$, all of which share common elements (presented in \Cref{subsec:cs:defs}).
-$\Ltms$ is the only statically-typed language, and it exhibits the property that all well-typed programs are \gls*{tms} (\Cref{subsec:ltms}).
-However, not all $\Ltms$ programs are \gls*{sms}.
-That is, there are well-typed $\Ltms$ programs that perform an out-of-bounds access.
-Language $\Ltrg$ is untyped and does not provide any guarantees with regards to \gls*{ms} (\Cref{subsec:lsms}).
-$\Lms$ is exactly the same language as $\Ltrg$, but this paper still distinguishes the two for sake of readability (\Cref{subsec:lms}).
-All three languages --- so $\Ltms$, $\Ltrg$, and $\Lms$ --- assume \gls*{scct} to hold, since this is -- in an ideal world -- what the programmer would expect, too: it is the job of the compiler to preserve and (potentially) enforce \gls*{scct} security~\cite{cauligi2019fact,nagarakatte2010cets,nagarakatte2009soft,akritidis2009baggy}.
-% This paper considers \gls*{cct} as an architecture-dependent property\MPin{can we justify this}, so
-
-Such consideration is also backed up by architecture providing a data (operand) independent timing mode, such as processors by Arm~\cite[p.~543]{arm-refman} and Intel~\cite[p.~80]{intel-refman}.
-%In spirit of this, all languages besides $\Lscct$ are trivially \gls*{scct}. 
-This kind of processor feature is modelled in language $\Lscct$ (\Cref{subsec:lscct}), where programs have access to a ``\gls*{cct}-mode'' and can change the leakage of emitted events according to the value of this mode (either $\obj{ON}$ or $\obj{OFF}$).
-
-Finally, modern processors also employ speculative execution to achieve speedups---and unfortunately generate Spectre attacks~\cite{kocher2019spectre}---and this is the extension of $\Lspec$ (\Cref{subsec:lspec}).
- % extends $\Lscct$ to model speculative execution.
-%
-Thus, all previous languages trivially satisfy \gls*{ss}, since they do not support speculative execution at all.
-% All languages share large parts of syntactic elements, which are presented in the next subsection (\Cref{subsec:cs:defs}).
-% Subsequent subsections extend on this accordingly.
-
-\subsection{Shared Language Definitions}\label{subsec:cs:defs}
-
-All presented programming languages share a common fragment which is partially presented here and in full detail in the technical report. 
-%\MP{put omitted stuff in appendix + ref there (and add textual explanation)}
-
-% \vspace{-.5em}
-{
-  \renewcommand{\src}[1]{\mathsf{#1}}
- \small
-\[
-  \begin{array}{rrcl}
-    \text{(Base-Events)} & \src{\preevent} &:=& \src{\ev{Alloc \loc\ n}} \mid \src{\ev{Dealloc \loc}} \mid \src{\ev{Get \loc\ n}} \mid \src{\ev{Set \loc\ n}} \\
-    \text{(Control Tags)} & \src{\sandboxtag{}} &:=& \src{\ctx} \mid \src{\comp} \\
-    \text{(Events)} & \src{\event} &:=& \src{\preevent;\sandboxtag{}} \mid \src{\emptyevent} \mid \src{\lightning} \\ 
-    \text{(Values)} & \src{v} &:=& \src{n} \\
-    \text{(Expressions)} & \src{\expr} &:=& \src{x} \mid \src{n} \mid \src{\lbinop{\expr[_1]}{\expr[_2]}} \mid \src{\lifz{\expr[_1]}{\expr[_2]}{\expr[_3]}} \\ 
-                         &&&\mid \src{\llet{x}{\expr[_1]}{\expr[_2]}} \mid \src{\llet{x}{new\ \expr[_1]}{\expr[_2]}} \\
-                         &&&\mid \src{\ldelete{x}} \mid \src{\lget{x}{\expr}} \mid \src{\lset{x}{\expr[_1]}{\expr[]}} \mid \src{stuck}\\
-                         &&&\mid \src{\lcall{f}{\expr}} \mid \src{\lreturn{\expr}} \\
-    \text{(Functions)} & \src{F} &:=& (\src{x};\src{\expr}) \\
-    \text{(Libraries)} &&&\hspace{-3.8em} \src{\library_{\ctx}},\src{\library_{\comp}} : \src{Vars} \to \src{F} 
-    % \hspace{2em}
-    %  : \src{Vars} \to \src{F} 
-     \\
-    \text{(Heaps)} & \src{H} &:=& \src{\hole{\cdot}} \mid \src{n},\src{H}
-    % \hspace{0.5cm}    
-    \\
-    \text{(Pointer Maps)} & \src{\Delta} && \text{ omitted for simplicity}
-    \\
-    \text{(Memory States)} & \src{\memstate} &:=& \left(\src{H^{\ctx}};\src{H^{\comp}};\src{\Delta}\right)\\
-    \text{(Control States)} & \src{\cfstate} & & \text{ omitted for simplicity} 
-    \\
-    \text{(States)} &\src{\Omega} &:=& \left(\src{\cfstate};\src{\sandboxtag{}};\src{\memstate}\right)\\
-  \end{array}
-\]
-}
-{
-  \renewcommand{\src}[1]{\mathsf{#1}}
-The trace models of all languages are similar to those presented earlier (\Cref{sec:compprop}).
-One technical detail is the addition of a control tag, indicating who is to blame for an emitted action: context ($\ctx$) or component ($\comp$).
-This is a standard technique in secure compilation in order to rule out irrelevant context-level events. 
-For example, a context immediately deallocating an allocated memory region twice trivially violates memory safety, but the main interest in secure compilation is to preserve component-level security, and thus component events.
-Even though this tagging could be used for blame preservation~\cite{patrignani2023blame}, this is beyond the focus of this paper.
-Another key difference is that memory accesses are now explicitly modelled as reads ($\src{\ev{Read\ \loc\ n}}$) and writes ($\src{\ev{Write\ \loc\ n}}$) instead of just uses.
-
-All languages have at least numbers as values ($\src{v}$) and second class functions ($\src{F}$).
-Functions are modelled as pairs containing the name of one argument and the body of the function.
-Bodies are just ordinary expressions, which can be simple binary operations ($\src{\lbinop{\expr[_1]}{\expr[_2]}}$), conditionals ($\src{\lifz{\expr[_1]}{\expr[_2]}{\expr[_3]}}$), function calls ($\src{\lcall{f}{e}}$) and returns ($\src{\lreturn{\expr}}$), as well as C-like memory operations.
-Programs have sets of pre-determined functions called libraries and they are marked as being part of some component ($\src{\library_{\comp}}$) or context ($\src{\library_{\ctx}}$).
-
-For the operational semantics, the runtime state ($\src{\Omega}$) is a triple consisting of a control-flow state ($\src{\cfstate}$), a control tag ($\src{\sandboxtag{}}$), and a memory state ($\src{\memstate}$).
-The latter carries information about pointers that are kept ``alive'' in pointer maps ($\src{\Delta}$), so that the semantics does not get stuck when encountering, e.g., a double-free.
-The memory state also carries two heaps to model sandboxing between a context ($\src{H^{\ctx}}$) and a component ($\src{H^{\comp}}$).
-
-}
-
-\subsection{$\src{L_{\tmssafe}}$: A Temporal but Not Spatial Memory Safe Language}\label{subsec:ltms}
-
-$\src{L_{\tmssafe}}$ is the only statically-typed language in this case study and restricts functions ($\src{F}$) to the typing signature $\src{\natt\to\natt}$. 
-The type system of $\src{L_{\tmssafe}}$ is inspired by $L^{3}$~\cite{morrisett2005L3,scherer2018fabulous} and enforces that every well-typed $\src{L_{\tmssafe}}$ program satisfies \gls*{tms}.
- % (\Cref{thm:wt:tms}).
-
-\begin{theorem}[$\src{L_{\tmssafe}}$-programs are \gls*{tms}]\label{thm:wt:tms}
-$\rsat{\src{\library_{\comp}}}{\tmssafe}$ \Coqed
-  % $\;$ 
-% 
-  % \begin{nscenter}
-  %   \hfill For any component $\src{\library_{\comp}}$, $\rsat{\src{\library_{\comp}}}{\tmssafe}$ \Coqed
-  % \end{nscenter}
-\end{theorem}
-
-\subsection{$\Ltrg$: A Memory-Unsafe Language}\label{subsec:lsms}
-
-$\Ltrg$ extends the syntax presented earlier (\Cref{subsec:cs:defs}) with dynamic typechecks ($\trg{\lhast{\expr}{\type}}$), evaluating to $\trg{0}$ if the type matches with the shape of $\trg{\expr}$ and $\trg{1}$ otherwise.
-%The pointers carry poison tags, which signify a pointer being allocated ($\trg{\poisoned}$) or freed ($\trg{\poisonless}$). 
-Furthermore, the syntax of $\Ltrg$ is extended with a way to inspect whether a pointer is freed ($\trg{\lispoisoned{x}}$), evaluating to $\trg{0}$ if it is freed and to $\trg{1}$ otherwise.
-Functions may receive arguments that are not $\trg{\natt}$, values are extended with tuples, and expressions are also extended with pair projections.
-
-% \vspace{-1em}
-\[
-  \begin{array}{rrcl}
-    \text{(Values)} & \trg{v} &:=& \dots \mid \trg{\lpair{v_1}{v_2}} \\
-    \text{(Expressions)} & \trg{\expr} &:=& \dots \mid \trg{\lhast{\expr}{\tau}} \mid \trg{\lispoisoned{x}}
-  \end{array}
-\]
-
-No changes are done to the trace model, but $\Ltrg$ has no static typing, making double-free code patterns possible.
-
-\subsection{$\Lms$: Another Memory-Unsafe Language}\label{subsec:lms}
-$\Lms$ is exactly equal to $\Ltrg$ (\Cref{subsec:lsms}), but used to emphasize that this is code after applying $\cc{\Ltrg}{\Lms}$ (\Cref{subsec:cs:ms}).
-
-\subsection{$\Lscct$: A Memory-Unsafe Language with a Constant Time Mode}\label{subsec:lscct}
-
-$\Lscct$ extends $\Lms$ (\Cref{subsec:lms}) with a ,,constant-time mode''. 
-The activation of the mode can be checked ($\obj{\lrddoit{x}{\expr}}$), changed ($\obj{\lwrdoit{D}}$), and is stored in program states ($\obj{\Omega}$).
-At the beginning of program execution of $\Lscct$ programs, the mode is turned off ($\obj{OFF}$).
-If the mode is enabled (i.e., set to $\obj{ON}$), the intuition is that no secrets are leaked. 
-This models the real-world\footnote{As present in Intel~\cite[p.80]{intel-refman} and ARM~\cite[p.~543]{arm-refman} processors.} data-independent timing mode as well as the result of compiling a program with FaCT~\cite{cauligi2019fact}.
-For example, FaCT rewrite code that branches to use a constant-time selection primitive.
-To not obfuscate our formalisation unnecessarily by duplicating syntax, we simply added the ,,constant-time mode'' to $\Lscct$.
-
-% $\Lscct$ extends $\Lms$ (\Cref{subsec:lms}) with a way to read from ($\obj{\lrddoit{x}{\expr}}$) and to write to ($\obj{\lwrdoit{D}}$) a model-specific register that controls a constant-time mode. 
-% This language is a combination of the effect of the FaCT compiler~\cite{cauligi2019fact} as well as a data-independent timing mode, a feature that is present in both Arm~\cite[p.~543]{arm-refman} and Intel~\cite[p.~80]{intel-refman} processors.
-% \MP{ can we be more specific about the fact pass? }
-% To this end, states ($\obj{\Omega}$) are extended with the value of the register ($\obj{D}$), which is initially set to be not active ($\obj{OFF}$).
-% If the register is marked active ($\obj{ON}$), the intuition is that no secrets can appear on specification traces.
-% If the register is marked inactive, secrets may appear on traces.
-% The state of the register can be read and written to, e.g., \Cref{tr:e-wrdoit-off}.
-% For the other languages seen earlier, the mode is intuitively always-on, i.e., the mode of execution always uses constant-time mode.
-
-% \vspace{-.5em}
-\[
-  \begin{array}{rrcl}
-    \text{(Mode Values)} & \obj{D} &:=& \obj{ON} \mid \obj{OFF} \\
-    \text{(Security Tags)} & \obj{\securitytag{}} &:=& \obj{\lock} \mid \obj{\unlock} \\
-    \text{(Base-Events)} & \obj{\preevent} &:=& \dots \mid \obj{iGet\ \loc\ v} \mid \obj{iSet\ \loc\ v} \\
-                       &&& \mid\ \obj{Binop\ n} \mid \obj{Branch\ n} \\
-    \text{(Events)} & \obj{\event} &:=& \obj{\preevent;\sandboxtag{};\securitytag{}} \mid \obj{\emptyevent} \mid \obj{\lightning} \\
-    \text{(Expressions)} & \obj{\expr} &:=& \dots \mid \obj{\lrddoit{x}{\expr}} \mid \obj{\lwrdoit{D}} \\
-                         &&&\mid\ \obj{\llet{x^{\securitytag{}}}{\expr[_1]}{\expr[_2]}}\\
-    \text{(States)} & \obj{\Omega} & := & \left(\obj{\cfstate};\obj{\sandboxtag{}};\obj{D};\obj{\memstate}\right)\\
-  \end{array}
-\]
-\begin{center}
-\newcommand{\expreval}[5]{{#1}\triangleright\xspace {#2}\xrightarrow{#5}\ {#3}\triangleright\xspace {#4}\xspace}
-\newcommand{\exprevalo}[5]{\expreval{\obj{#1}}{\obj{#2}}{\obj{#3}}{\obj{#4}}{\obj{#5}}}
-  \typerule{$e-\obj{wrdoit}-\text{off}$}{
-  }{
-    \pstepto{\rtt{\obj{\cfstate;\sandboxtag{};D;\memstate}}{\obj{\lwrdoit{OFF}}}}{\rtt{\obj{\cfstate;\sandboxtag{};OFF;\memstate}}{\obj{0^{\securitytag{}}}}}{\obj{\emptyevent}}
-  }{e-wrdoit-off}
-
-  \typerule{$e-\obj{get}-\in-$noleak}{
-    \obj{\memstate}=\obj{H^{\ctx};H^{\comp};\Delta_1,x\mapsto(\loc;\sandboxtag{};\poison),\Delta_2} &
-    \obj{\loc}+\obj{n}\in\text{dom }\obj{H^{\sandboxtag{}}} \\
-    \obj{\securitytag{''}}=\obj{\securitytag{}\sqcap\securitytag{'}} &
-    \obj{\event} = \obj{{iGet}\ \loc\ n;\sandboxtag{};\securitytag{''}}
-  }{
-    \exprevalo{
-    	\cfstate;\sandboxtag{'};ON;\memstate
-    }{
-      	x^{\securitytag[n^{\securitytag{'}}] }
-    }{
-    	\cfstate;\sandboxtag{'};ON;\memstate
-    }{
-    	(H^{\sandboxtag{}}(\loc+n))^{\securitytag{''}}
-    }{
-      \event
-    }
-  }{to-e-get-in-noleak}
-\end{center}
-
-The language also adds user annotations $\obj{\securitytag{}}$ for the secrecy of variables, which can be either private (high secrecy) $\obj{\lock}$ or public (low secrecy) $\obj{\unlock}$.
-Security tags $\obj{\securitytag{}}$ are arranged in the usual secrecy lattice~\cite{zdancewicphd}, where $\obj{\unlock}\sqsubseteq\obj{\lock}$.
-% $\obj{\lock}\sqsubseteq\obj{\securitytag{}}$ and $\obj{\securitytag{}}\sqsubseteq\obj{\unlock}$.
-
-Memory accesses to secret data need to be present to reason about memory safety, even when execution is in constant-time mode, e.g, \Cref{tr:to-e-get-in-noleak}.
-$\Lscct$ extends base-events with $\obj{\ev{iGet}\ \loc\ \valueexpr}$ and $\obj{\ev{iSet}\ \loc\ \valueexpr}$ (for data $\obj{i}$ndependent get and set) to prevent secrets from leaking but still enable reasoning about memory safety.
-Due to the technical setup, the rule needs to check if the access is in bounds ($\obj{\loc}+\obj{n}\in\obj{H^t}$) and update the secrecy tag ($\obj{\securitytag{''}}=\obj{\securitytag{}\sqcap\securitytag{'}}$) with the least upper bound of the tags, according to the aforementioned lattice.
-The precise information carried by the event, e.g., location ($\obj{\loc}$) is taken from the pointer map, which carries information irrelevant to this rule ($\obj{\poison}$).
-
-Base-events include $\obj{\ev{Branch}\ n}$ and $\obj{\ev{Binop}\ n}$ that are emitted when evaluating a branch or certain binary expressions, such as division, respectively, whenever the constant-time mode is inactive.
-Events are extended with a security-tag ($\obj{\securitytag{}}$) to signal the secrecy of the involved data.
-
-The evaluation steps are amended to propagate the security-tag annotations $\obj{\securitytag{}}$.
-When the constant-time mode is inactive, base-events $\obj{\ev{Branch}\ n}$ and $\obj{\ev{Binop}\ n}$ are emitted for conditionals and binary operations, respectively.
-Otherwise, just like in the semantics of the earlier languages, $\obj{\emptyevent}$ is emitted for binary and branching operations.
-%To mark this mode as active or inactive, \Cref{tr:e-wrdoit} allows to access the register in $\obj{\statevar}$.
-
-\Cref{ex:lscct} illustrates the differences between $\Lscct$ and other languages.
-
-\begin{exampleenv}[$\Lscct$ with and without constant-time mode]\label{ex:lscct}
-  Consider again \Cref{ex:strncpy}, with a context copying the string \texttt{Hello World}, where everything is marked with a high security tag: $\lock$.
-  The top half of \Cref{fig:ex-cct} (titled $\Lscct$), describes the execution trace of the program, while the bottom side of the table (titled ``Specification''), describes the related specification trace (\Cref{subsec:msctss:tracemodel}).
-  Read in parallel from top to bottom, the figure shows parts of the execution trace. 
-  In each half, the left column (\textit{Active}) has constant-time mode $\obj{ON}$ and the right one (\textit{Inactive}) has it $\obj{OFF}$.
-
-  When the constant-time mode is off, the execution yields events in similar fashion to before (\Cref{subsec:ltms,subsec:lsms,subsec:lms}).
-  But, if it is turned on, then the branching event \colorbox{yellow}{does not fire} anymore and both reading and writing to memory is related to a specification trace with \colorbox{lightgray}{no exposed secrets}.
-\end{exampleenv}
-\begin{figure}[!htb]
-	\vspace{-1em}\small
-  $$
-  % {
-  \begin{array}{ccc}
-    & \Lscct &        \\
-    \text{Active} & \mid & \text{Inactive} \\\hline
-    \obj{\ev{Alloc}\ \loc_{x}\ 12;\comp;\unlock} & \mid & \obj{\ev{Alloc}\ \loc_{x}\ 12;\comp;\unlock} \\
-    \obj{\ev{Alloc}\ \loc_{y}\ 12;\comp;\unlock} & \mid & \obj{\ev{Alloc}\ \loc_{y}\ 12;\comp;\unlock} \\
-    \obj{{\ev{iGet}}\ \loc_{x}\ 0;\comp;}\text{\colorbox{lightgray}{$\obj{\lock}$}} & \mid & \obj{\ev{Get}\ \loc_{x}\ 0;\comp;\lock} \\
-    \text{\colorbox{yellow}{$\obj{\emptyevent}$}} & \mid & \obj{\ev{Branch}\ 0;\comp;\lock} \\
-    \obj{{\ev{iGet}}\ \loc_{x}\ 0;\comp;}\text{\colorbox{lightgray}{$\obj{\lock}$}} & \mid & \obj{\ev{Get}\ \loc_{x}\ 0;\comp;\lock} \\
-    \obj{{\ev{iSet}}\ \loc_{y}\ 0\ \mathtt{'H'};\comp;}\text{\colorbox{lightgray}{$\obj{\lock}$}} & \mid & \obj{\ev{Set}\ \loc_{y}\ 0\ \mathtt{'H'};\comp;\lock} \\
-    \obj{{\ev{iGet}}\ \loc_{x}\ 1;\comp;}\text{\colorbox{lightgray}{$\obj{\lock}$}} & \mid & \obj{\ev{Get}\ \loc_{x}\ 1;\comp;\lock} \\
-    \text{\colorbox{yellow}{$\obj{\emptyevent}$}} & \mid & \obj{\ev{Branch}\ 0;\comp;\lock} \\
-    \vdots & \mid & \vdots \\
-    \obj{{\ev{iGet}}\ \loc_{x}\ 12;\comp;}\text{\colorbox{lightgray}{$\obj{\lock}$}} & \mid & \obj{\ev{Get}\ \loc_{x}\ 12;\comp;\lock} \\
-    \text{\colorbox{yellow}{$\obj{\emptyevent}$}}& \mid & \obj{\ev{Branch}\ 1;\comp;\lock} \\
-    \obj{\ev{Dealloc}\ \loc_{y};\comp;\unlock} & \mid & \obj{\ev{Dealloc}\ \loc_{y};\comp;\lock} \\
-    \obj{{\ev{iGet}}\ \loc_{y}\ 6;\comp;}\text{\colorbox{lightgray}{$\obj{\lock}$}} & \mid & \obj{\ev{Get}\ \loc_{y}\ 6;\comp;\lock} \\
-  \end{array}
-  % }
-  $$
-  $$
-  % {
-  \begin{array}{ccc}
-          & \text{Specification} & \\
-  \text{Active} & \mid & \text{Inactive} \\\hline
-     \ev{Alloc}\ \loc_{x}\ 12;\unlock & \mid & \ev{Alloc}\ \loc_{x}\ 12;\unlock\\
-     \ev{Alloc}\ \loc_{y}\ 12;\unlock & \mid & \ev{Alloc}\ \loc_{x}\ 12;\unlock\\
-     \ev{Use}\ \loc_{x}\ 0;\text{\colorbox{lightgray}{$\unlock$}} & \mid & \ev{Use}\ \loc_{x}\ 0;\lock\\
-     \emptyevent & \mid & \ev{Branch}\ 0;\lock\\
-     \ev{Use}\ \loc_{x}\ 0;\text{\colorbox{lightgray}{$\unlock$}} & \mid & \ev{Use}\ \loc_{x}\ 0;\lock\\
-     \ev{Use}\ \loc_{y}\ 0;\text{\colorbox{lightgray}{$\unlock$}} & \mid & \ev{Use}\ \loc_{y}\ 0;\lock\\
-     \ev{Use}\ \loc_{x}\ 1;\text{\colorbox{lightgray}{$\unlock$}} & \mid & \ev{Use}\ \loc_{x}\ 1;\lock\\
-     \emptyevent & \mid & \ev{Branch}\ 0;\lock\\
-     \vdots & \mid & \vdots \\
-      \ev{Use}\ \loc_{x}\ 12;\text{\colorbox{lightgray}{$\unlock$}} & \mid & \ev{Use}\ \loc_{x}\ 12;\lock\\
-      \emptyevent & \mid & \ev{Branch}\ 1;\lock\\
-      \ev{Dealloc}\ \loc_{y};\unlock & \mid & \ev{Dealloc}\ \loc_{y};\unlock\\
-      \ev{Use}\ \loc_{y}\ 6;\text{\colorbox{lightgray}{$\unlock$}} & \mid & \ev{Use}\ \loc_{y}\ 6;\lock\\
-  \end{array}
-  % }
-  $$
-  \caption{Traces for \Cref{ex:lscct}.}
-  \label{fig:ex-cct}
-\end{figure}
-
-\subsection{$\Lspec$: A Memory-Unsafe Language with Speculation}\label{subsec:lspec}
-
-$\Lspec$ extends $\Lscct$ (\Cref{subsec:lscct}) with speculative dynamic semantics that is inspired by existing work~\cite{guarnieri2018spectector,fabian2022automatic}.
-After branching, speculation starts by pushing the current configuration into the speculation state ($\ird{S}$), and run subsequent code in a predetermined window (whose size is $\omega$) until a rollback of operational state is performed.
-The window is set inside the speculation state.
-Within such windows, data may leak, this is marked as high $\ird{\lock}$ and with an annotation that indicates the respective speculative execution variant. 
-In this paper, we just consider SPECTRE-PHT~\cite{kocher2019spectre}, whose starting point is \Cref{tr:e-ifz-true-spec}.
-% where $\omega$ is the size of the speculation window, i.e., a number of reduction steps to be taken under speculative semantics.
-Additionally, $\ird{\lock}$ may also carry a $\ird{NONE}$ annotation, to signal that this is a leak irrespective of speculation, i.e., as defined earlier (\Cref{subsec:lscct}).
-The semantics is kept general enough to allow for future extension to support different variants~\cite{kocher2019spectre,maisuradze2018ret2spec,horn2019zero}.
-
-In terms of language features, $\Lspec$ includes a new barrier operation $\ird{\lbarrier}$ that blocks speculative execution. 
-To facilitate speculative execution in a non-assembly-like language, a stack of operational state is used and speculation is active if that stack has a size greater than one.
-This is exploited in \Cref{tr:e-spec-eat} to consume any leftover speculation.
-
-\vspace{-.5em}
-\[
-  \begin{array}{rrcl}
-    \text{(Base-Events)} & \ird{\preevent} &:=& \dots \mid \ird{Spec} \mid \ird{Rlb} \mid \ird{Barrier}\\
-    \text{(Expressions)} & \ird{\expr} &:=& \dots \mid \ird{\lbarrier} \\
-    \text{(Speculation State)} & \ird{S} &:=& \ird{\Omega\triangleright\expr} \mid \ird{S},\left(\ird{\Omega};\ird{n};\ird{\expr}\right)
-  \end{array}
-\]
-
-\begin{center}
-\newcommand{\expreval}[5]{{#1}\triangleright\xspace {#2}\xrightarrow{#5}\ {#3}\triangleright\xspace {#4}\xspace}
-\newcommand{\exprevald}[5]{\expreval{\ird{#1}}{\ird{#2}}{\ird{#3}}{\ird{#4}}{\ird{#5}}}
-
-  %
-  \typerule{$e-\ird{ifz}-\ird{true}-\text{spec}$}{
-    \ird{S} = \rtt{(\ird{\cfstate['];\sandboxtag{};D;\memstate});\ird{S}}{\ird{\expr[_1]}},({\ird{\cfstate;\sandboxtag{};D;\memstate}};\omega;{\ird{\expr[_2]}})
-  }{
-    \pstepto{\rtt{\ird{\cfstate;\sandboxtag{};D;\memstate}}{\ird{\lifz{0^{\securitytag{}}}{\expr[_1]}{\expr[_2]}}}}{\ird{S}}{\ird{{Branch}\ 0;\sandboxtag{};\securitytag{}}\cdot\ird{{Spec};\sandboxtag{};\securitytag{}}}
-  }{e-ifz-true-spec}
-  %
-  \typerule{$e-\ird{spec}-\text{eat}$}{
-    \ird{n} > 0 &
-    \pstepto{\rtt{\ird{\Omega}}{\ird{\expr}}}{\rtt{\ird{\Omega'}}{\ird{\expr[']}}}{\ird{\event}}
-  }{
-    \ghoststepto{\ird{S},(\ird{\Omega};\ird{n};\ird{\expr})}{\ird{S},(\ird{\Omega'};\ird{n}-1;\ird{\expr'})}{\ird{\event}}
-  }{e-spec-eat}
-  % 
-  \typerule{$e-\ird{spec}-\text{eaten}$}{
-  }{
-    \ghoststepto{\ird{S},(\ird{\Omega};\ird{0};\ird{\expr})}{\ird{S}}{\ird{Rlb;\Omega.\sandboxtag{};\unlock}}
-  }{e-spec-eaten}
-  %
-  \typerule{$e-\ird{spec}-\text{barrier}$}{
-  }{
-    \ghoststepto{\ird{S},(\ird{\Omega};\ird{n};\ird{\lbarrier})}{\ird{S},(\ird{\Omega};\ird{0};\ird{0})}{\ird{Barrier;\Omega.\sandboxtag{};\unlock}}
-  }{e-spec-barrier}
-\end{center}
-
-The trace model is extended with $\ird{{Spec}}$, $\ird{{Rlb}}$, and $\ird{{Barrier}}$ events that signal the respective operational action. 
-A $\ird{\lbarrier}$ prevents any execution whatsoever besides $\ird{{Rlb}}$ when run in speculative execution mode and does nothing when run in normal mode, e.g., \Cref{tr:e-spec-barrier}.%
-\footnote{In the rule, the notation $\cfstate' = \cfstate[.\text{win} = 0]$ means that $\cfstate'$ is a copy of $\cfstate$ up to field $\text{win}$, which is set to $0$.}
-Lastly, $\ird{Rlb}$ is emitted when the speculation window is zero and the operational state is rolled back (see \Cref{tr:e-spec-eaten}).
-
-\section{Case Study: Composing Secure Compiler Passes and Optimisations \pages{7}}\label{sec:casestud:rtp}
-This section defines several secure compilers, each of which robustly preserves a different property of interest as depicted in \Cref{fig:pipeline}.
-\begin{figure*}[!h]
-  \centering
-  \begin{tikzpicture}
-    \node (S) {$\src{L_{\tmssafe}}$};
-    \node[right=2.0 of S] (T) {$\trg{L}$};
-    \node[right=2.0 of T] (M) {$\irl{L_{\mssafe}}$};
-    \node[below right=1.5 and 1.0 of M] (D0) {$\irl{L_{\mssafe}}$};
-    \node[above right=1.5 and 1.0 of M] (C0) {$\irl{L_{\mssafe}}$};
-    \node[right=3.0 of M] (E) {$\irl{L_{\mssafe}}$};
-    \node[right=2.0 of E] (O) {$\obj{L_{\scctsafe}}$};
-    \node[right=2.0 of O] (Z) {$\ird{L_{\mathghost}}$};
-
-    \draw[->] (S) to[sloped] node[align=center] (tmsedge) {\gls*{tms}\\ \Cref{thm:cca:rtp:tms}} (T);
-    \draw[->] (T) to[sloped] node[align=center] {\gls*{sms}\\ \Cref{thm:ccb:rtp:sms}} (M);
-    \draw[->] (M) to[sloped] node[align=center] {\gls*{dce}\\ \Cref{thm:ccdce:rtp:ms}} (D0);
-    \draw[->] (M) to[sloped] node[align=center] {\gls*{cf}\\ \Cref{thm:cccf:rtp:ms}} (C0);
-    \draw[->] (D0) to[sloped] node[align=center] {\gls*{cf}\\ \Cref{thm:cccf:rtp:ms}} (E);
-    \draw[->] (C0) to[sloped] node[align=center] {\gls*{dce}\\ \Cref{thm:ccdce:rtp:ms}} (E);
-    \draw[->] (E) to[sloped] node[align=center] {\gls*{scct}\\ \Cref{thm:ccscct:rtp:scct}} (O);
-    \draw[->] (O) to[sloped] node[align=center] {\gls*{ss}\\ \Cref{thm:ccspec:rtp:spec}} (Z);
-
-    % Sections
-    \node (sectms) at ($(S)+(1.25,2.5)$) {{\Cref{subsec:cs:tms}}};
-    \node (secsms) at ($(T)+(1.25,2.5)$) {{\Cref{subsec:cs:ms}}};
-    \node (secopts) at ($(M)+(2.0,2.5)$) {{\Cref{subsec:cs:opts}}};
-    \node (secscct) at ($(E)+(1.5,2.5)$) {{\Cref{subsec:cs:scct}}};
-    \node (secss) at ($(O)+(1.5,2.5)$) {{\Cref{subsec:cs:ss}}};
-
-    \draw[thick,dotted,->] ($(S) + (0.2,-0.27)$) to ($(S) - (-0.2,1)$) to node[align=center] {\gls*{ms}\\ \Cref{corr:ccab:rtp:ms}} ($(M) - (0.0,1)$) to (M);
-
-    \draw[thick,dotted,->] (S) to ($(S) - (0.0,2.5)$) to node[align=center,pos=.65] {\gls*{specms}\\ \Cref{thm:ccall:rtp:specms}} ($(Z) - (0.0,2.5)$) to (Z);
-
-    % \draw[thick,dotted,->] (S) to[bend right=33,sloped] node[align=center] {\gls*{ms}\\ \Cref{corr:ccab:rtp:ms}} (M);
-    \draw[thick,dotted,->] (M) to[bend right=0,sloped] node[align=center] {\gls*{ms}\\ \Cref{thm:cccfccdce:rtp:ms}} (E);
-%    \draw[thick,dotted,->] (S) to[out=-45,in=198,sloped] node[align=center] {\gls*{ms}+\gls*{scct}\\ \Cref{thm:ccall:rtp:msscct}} (O);
-    % \draw[thick,dotted,->] (S) to[out=-40,in=210,sloped] node[align=center,pos=.8] {\gls*{specms}\\ \Cref{thm:ccall:rtp:specms}} (Z);
-
-    \draw[dashed] ($(S)-(0.0,-0.5)$) -- ($(S)-(0.0,-2.5)$);
-    \draw[dashed] ($(T)-(0.0,-0.5)$) -- ($(T)-(0.0,-2.5)$);
-    \draw[dashed] ($(M)-(0.0,-0.5)$) -- ($(M)-(0.0,-2.5)$);
-    \draw[dashed] ($(E)-(0.0,-0.5)$) -- ($(E)-(0.0,-2.5)$);
-    \draw[dashed] ($(O)-(0.0,-0.5)$) -- ($(O)-(0.0,-2.5)$);
-    \draw[dashed] ($(Z)-(0.0,-0.5)$) -- ($(Z)-(0.0,-2.5)$);
-  \end{tikzpicture}
-  \vspace{-1em}
-  \caption{Visualisation of the optimising compilation pipeline that preserves \gls*{specms}. %
-    Vertices in the graph are the programming languages from earlier sections (\Cref{sec:casestud:defs}). %
-    Full edges are secure compilers passes.
-    Dotted edges are composition of passes and use the presented framework (\Cref{sec:sequential}) to indicate the property they preserve. %
-    The dashed lines partition the graph into the sections where the respective theorems are presented.
-  }\label{fig:pipeline}
-\end{figure*}
-The section demonstrates the power of the framework (\Cref{sec:sequential}) by composing these compilers for a secure and optimising compilation chain that robustly preserves \gls*{specms}.
-The first step in this chain is the compiler from $\src{L_{\tmssafe}}$ to $\trg{L}$ that robustly preserves just \gls*{tms} (\Cref{thm:cca:rtp:tms}).
-From here, another pass from $\trg{L}$ to $\irl{L_{\mssafe}}$ ensures that no out-of-bounds accesses can happen and, thus, programs at this point attain \gls*{sms} (\Cref{thm:ccb:rtp:sms}).
-Since these properties compose into \gls*{ms}, composing these passes yields a compiler that robustly preserves \gls*{ms} (\Cref{corr:ccab:rtp:ms}).
-Then, the section presents two optimisation passes, namely \gls*{cf} and \gls*{dce}, each of which robustly preserves \gls*{ms} (\Cref{thm:ccdce:rtp:ms,thm:cccf:rtp:ms}).
-These passes can be freely ordered in the compilation chain without compromising memory safety (\Cref{thm:cccfccdce:rtp:ms}).
-The next step in the chain ensures that code stays \gls*{scct} (\Cref{thm:ccscct:rtp:scct}) when compiled from $\Lms$ to $\Lscct$, which is done by switching on a constant-tame mode for the computation.
-Lastly, by introducing barriers immediately after branches, speculative leaks via SPECTRE-PHT are prevented when compiling $\Lscct$ to $\Lspec$.
-The final result is that the whole compilation chain robustly preserves \gls*{specms} (\Cref{thm:ccall:rtp:specms}).
-
-\subsection{Robust Temporal Memory Safety Preservation}\label{subsec:cs:tms}
-
-The secure compiler from $\Ltms$ to $\Ltrg$ needs to ensure that when execution switches from context to component, the type signatures are respected.
-Thus, the compiler inserts the following dynamic typechecks before entering the body of a component-defined function (anything elided is a trivial identity function from source to target):
-
-\vspace{-1em}
-{
-\begin{gather}
-  \begin{align*}%{rcl}
-    %\cca(\src{x}) = &\ \trg{x} 
-    % \\
-  	%&
-  	%\qquad
-    %\cca(\src{\lbinop{\expr[_{1}]}{\expr[_{2}]}}) = &\ \lbinop[\trg]{\left[\cca(\src{\expr[_{1}]})\right]}{\left[\cca(\src{\expr[_{2}]})\right]} \\
-    %\cca(\src{n}) = &\ \trg{n} 
-    % \\
-    %&
-    \cca(\src{\lget{x}{\expr}}) = &\ \lget[\trg]{\trg{x}}{\left[\cca(\src{\expr})\right]} \\
-    \cca(\src{\ldelete{x}}) = &\ \ldelete[\trg]{\left[\cca(\src{x})\right]} 
-    \\
-    \cca(\src{\lfunction{g}{x:\natt\to\type[_{e}]}{\expr}})  =&
-  \end{align*}
-  \\
-  \begin{align*}
-%    \cca(\src{\lset{x}{\expr[_{1}]}{\expr[_{2}]}}) & = \lset[\trg]{\left[\cca(\src{x})\right]}{\left[\cca(\src{\expr[_{1}]})\right]}{\left[\cca(\src{\expr_{2}})\right]} \\
-% \\ 
-\lfunction[\trg]{\trg{g}}{\trg{x}}{\lifz[\trg]{\trg{\lhast{x}{\natt}}}{
-                                                            \left[\cca(\src{\expr})\right] %
-                                                                                                 }{\labort[\trg]}}
-  \end{align*}
-\end{gather}
-}
-
-\noindent Since $\trg{L}$ has no static typing, an attacker $\trg{\library_{\ctx}}$ can invoke a component function accepting a $\src{\natt}$ with $\trg{\lpair{17}{29}}$.
-With the dynamic check, the compiler ensures that execution aborts in such cases.
-%The compiler does not insert other checks and proceeds as the identity function (which in this paper amounts to a simple re-colouring of $\src{L_{\tmssafe}}$ to $\trg{L}$ expressions).
-
-Compiling the \texttt{strncpy} function from \Cref{sec:introduction} with $\cca$, the compiler would in this case ensure that the arguments that are evaluated in the compiled \texttt{strncpy} are valid.
-%This effectively has no influence on the resulting trace, so a corresponding cross-language trace relation $\sim^{\Ltms}_{\Ltrg}$ is just equality.
-
-$\cca$ is robustly preserving (\Cref{def:rtp}) \gls*{tms}:
-\begin{theorem}[$\cca$ secure w.r.t. \gls*{tms}]\label{thm:cca:rtp:tms}
-    $\rtp{\cca}{\tmssafe}$ % \Coqed
-\end{theorem}
-
-\subsection{Robust Spatial Memory Safety Preservation}\label{subsec:cs:ms}
-
-The spatial memory safety preserving compiler from $\Ltrg$ to $\Lms$ only inserts bounds-checks whenever reading from or writing to memory in order to enforce \gls*{sms}.
-These bounds checks need the bounds information, which the compiler keeps around by introducing a fresh identifier $\irl{x_{SIZE}}$ for each allocation that binds $\irl{x}$.
-Then, it is simply a matter of referring to that variable and ensuring that memory accesses are in the interval $[\irl{0},\irl{x_{SIZE}})$.
-When the check fails, the code aborts.
-
-\begin{nscenter}
-% \small
-  $$
-  \begin{array}{rcl}
-    \ccb(\trg{\lnew{x}{\expr[_{1}]}{\expr[_{2}]}}) & = 
-                                                   & \llet[\irl]{\irl{x_{SIZE}}}{\ccb(\trg{\expr[_{1}]})}{}
-    		\\&&
-    		\lnew[\irl]{\irl{x}}{\irl{x_{SIZE}}}{\ccb(\trg{\expr[_{2}]})}
-    		 \\
-  \ccb(\trg{\lget{x}{\expr}}) & = 
-                              & \llet[\irl]{\irl{x_{n}}}{\ccb(\trg{\expr})}{}
-  	\\&&
-  \lifz[\irl]{\irl{0\le x_{n}<x_{SIZE}}}{\\&&\irl{\lget{x}{x_{n}}}}{}
-  		\irl{\labort}
-  	  \\
-  \ccb(\trg{\lset{x}{\expr[_{1}]}{\expr[_{2}]}}) & = 
-                                                 & \llet[\irl]{\irl{x_{n}}}{\ccb(\trg{\expr[_{1}]})}{}
-  		\\&&
-  \lifz[\irl]{\irl{0\le x_{n}<x_{SIZE}}}{\\&&\lset[\irl]{\irl{x}}{\irl{x_{n}}}{}
-  		\ccb(\trg{\expr[_{2}]})
-  		}{\irl{\labort}} 
-  		% \\
-  \end{array}
-  $$
-\end{nscenter}
-
-\begin{exampleenv}[Instrumented \texttt{strncpy}]
-  Consider again \texttt{strncpy}, but instrumented for \gls*{sms}:
-    \begin{lstlisting}[language=c,basicstyle=\ttfamily, morekeywords={size_t}]
-void strncpy(size_t n, size_t dst_size, char *dst,
-             size_t src_size, char *src) {
-  for(size_t i = 0; i < src_size
-      && src[i] != '\0' && i < n; ++i) {
-    if(i < src_size && i < dst_size) {
-      dst[i] = src[i];
-    }
-  }
-}
-    \end{lstlisting}
-Consider context \texttt{strncpy(2, x, y)}, where \texttt{x} and \texttt{y} are pointers to valid regions of memory with allocated space for exactly two cells and do not contain the null-terminating character \texttt{'\textbackslash 0'}.
-% 
-Without the \gls*{sms} pass, the event $\ev{Use}\ \loc_{x}\ 2;\comp;\unlock$ would appear on the trace, but that indicates an out-of-bounds access! 
-Fortunately, with \gls*{sms} mitigation in place, that event does not appear during execution, since the bounds check prevents the condition \texttt{src[i] != '\textbackslash 0'} from executing.
-\end{exampleenv}
-
-Contrary to the previous compiler, $\ccb$ may change the trace of the original $\Ltrg$ program: if there is a memory access, it needs to be protected with a bounds check.
-% 
-The corresponding relation $\sim^{\Ltrg}_{\Lms} : \trg{\trace}\times\irl{\trace}$ that describes this semantic effect of the compiler is defined partially below.
-% 
-We omit action $\trg{Set}$, which is treated analogously, and any other event, which is related to its cross-language equivalent.
-% \footnote{Up to colours.}
-% \footnote{In the interest of space, the full details are in the technical report.}
-
-{
-\[
-  \typerulenolabel{xrel:sms:read}{\trg{n}\text{ in bounds}}{\trg{Get\ \loc\ n;\comp}\sim^{\Ltrg}_{\Lms}\irl{Get\ \loc\ n;\comp}}
-\]
-% \[
-%   \typerulenolabel{xrel:sms:panic}{\trg{n}\text{ out of bounds}}{\trg{Get\ \loc\ n;\comp}\sim^{\Ltrg}_{\Lms}\irl{\lightning}}
-% \]
-}
-
-% Note that neither $\Ltrg$ nor $\Lms$ expose branches on the trace.\MP{so what?}
-For simplicity, we elide the environment that this relation carries around in order to bind each location to its metadata (such as its size), and resolve the ``$\trg{n}$ in/out of bounds'' premise.
-% The bounds check mentioned in the premises of the rules above is left abstract for the sake of simplicity, the interested reader can find
-% since the precise definition needs some technicalities that only inhibit readability without any technical insights.
-% 
-We can now prove that compiler $\ccb$ robustly preserves \gls*{sms}.
-\begin{theorem}[$\ccb$ secure w.r.t. \gls*{sms}]\label{thm:ccb:rtp:sms}
-  $\rtpsim{\ccb}{\smssafe}{\sim^{\Ltrg}_{\Lms}}$ % \Coqed
-\end{theorem}
-
-At this point we can compose $\ccb$ with the previous compiler ($\cca$), but in order to do so, we need a trace relation from $\Ltms$ to $\Lms$.
-% 
-We can obtain this relation by composing the trace relation we just defined ($\sim^{\Ltrg}_{\Lms}$) with the one used by the previous compiler: $\sim^{\Ltms}_{\Ltrg} : \src{\trace}\times\trg{\trace}$.
-% 
-The latter has not been previously defined (nor has it been used in the related theorem) because that is just an equality relation, since the trace models of $\Ltms$ and of $\Ltrg$ are the same.
-% 
-Thus, we formally define $\sim^{\Ltms}_{\Lms}: \src{\trace} \times \irl{\trace}$ as the following composition: $\sim^{\Ltms}_{\Ltrg}\bullet\sim^{\Ltrg}_{\Lms}$.
-% 
-With this relation, \Cref{corr:ccab:rtp:ms} states that the composition of $\cca$ and $\ccb$ is secure w.r.t.
-\gls*{ms} and it follows from \Cref{thm:cca:rtp:tms,thm:ccb:rtp:sms} using \Cref{thm:rtpsim:sig}.
-% 
-\begin{corollary}[$\cca\circ\ccb$ secure w.r.t. \gls*{ms}]\label{corr:ccab:rtp:ms}
-  $\;$
-
-  \begin{nscenter}
-    $\rtpsim{\cca\circ\ccb}{\mssafe}{\sim^{\Ltms}_{\Lms}}$ % \Coqed
-  \end{nscenter}
-\end{corollary}
-% 
-This proof requires another precondition besides \Cref{thm:ccb:rtp:sms,thm:cca:rtp:tms}: $\sim^{\Ltms}_{\Ltrg}$ needs to be well-formed with respect to $\smssafe$.
-This follows trivially since $\sim^{\Ltms}_{\Ltrg}$ is an equality. 
-% 
-\begin{lemma}[$\sim^{\Ltms}_{\Ltrg}$ well-formed w.r.t. $\smssafe$]\label{lem:wf:ltmsltrg}
-$\;$
-
-  \begin{nscenter}
-  $\wfcsig{\sim^{\Ltms}_{\Ltrg}}{\smssafe}$
-  \end{nscenter}
-\end{lemma}
-
-\subsection{Optimising Compilers}\label{subsec:cs:opts}
-
-This section defines two optimising compiler passes from $\Lms$ to $\Lms$ which perform \gls*{dce} and \gls*{cf}, respectively.
-The \gls*{dce} pass applies a na\"ive rewrite rule on conditionals.
-The \gls*{cf} pass relies on an auxiliary function \texttt{mix} that uses a substitutions accumulator $\irl{\substlist}$ in order to rewrite constant binary operations, e.g., $\irl{{17}-1}$ to $\irl{16}$, and replace variables that are assigned to constants, e.g., $\irl{\llet{x}{7}{x}}$ to $\irl{7}$.
-
-\vspace{-1em}
-\begin{gather*}
-  \begin{align*}
-    \ccdce(\irl{\lifz{true}{\expr[_{1}]}{\expr[_{2}]}}) & = \ccdce(\irl{\expr[_{1}]}) &\\
-    \ccdce(\irl{\lifz{false}{\expr[_{1}]}{\expr[_{2}]}}) & = \ccdce(\irl{\expr[_{2}]}) &
-  \end{align*}
-  \\
-  % \begin{align*}
-  %   \ccdce(\irl{\lbinop{\expr[_{1}]}{\expr[_{2}]}}) & = \lbinop[\irl]{\ccdce(\irl{\expr[_{1}]})}{\ccdce(\irl{\expr[_{2}]})} &
-  % \end{align*}
-  % \\[0.25cm]
-  \begin{align*}
-    \cccf(\irl{\expr}) & = \partialeval{\irl{\expr}}{\irl{\hole{\cdot}}} &
-  \end{align*}
-  \\[0.125cm]
-  \begin{align*}
-   \partialeval{\irl{x}}{\irl{\substlist}} & = \irl{n} 
-   	\qquad\qquad \text{if } \irl{\subst{n}{x}}\in\irl{\substlist} \\
-   \partialeval{\irl{x}}{\irl{\substlist}} & = \irl{x} 
-   \qquad\qquad \text{if } \irl{\subst{n}{x}}\notin\irl{\substlist} \\
-   \partialeval{\irl{\lbinop{n}{m}}}{\irl{\substlist}} & = \irl{k} 
-   \qquad\qquad \text{if } \lbinop{\irl{n}}{\irl{m}}=k \\
-   %\partialeval{\irl{\lbinop{\expr[_{1}]}{\expr[_{2}]}}}{\irl{\substlist}} & = \lbinop[\irl]{\partialeval{\irl{\expr[_{1}]}}{\irl{\substlist}}}{\partialeval{\irl{\expr[_{2}]}}{\irl{\substlist}}} \\
-   \partialeval{\irl{\llet{x}{n}{\expr}}}{\irl{\substlist}} & = \partialeval{\irl{\expr}}{\irl{\subst{n}{x}\cdot\substlist}} 
-  % \\
-  % \partialeval{\irl{\lget{x}{\expr}}}{\irl{\substlist}} & = \lget[\irl]{\irl{x}}{\partialeval{\irl{\expr}}{\irl{\substlist}}}
-% \end{align*}
-\end{align*}
-% \\
-% \begin{align*}
-%   \partialeval{\irl{\llet{x}{\expr[_{1}]}{\expr[_{2}]}}}{\irl{\substlist}} & = \llet[\irl]{\irl{x}}{\partialeval{\irl{\expr[_{1}]}}{\irl{\substlist}}}{\\&\partialeval{\irl{\expr[_{2}]}}{\irl{\substlist}}} \\
-%   \partialeval{\irl{\lifz{\expr[_{1}]}{\expr[_{2}]}{\expr[_{3}]}}}{\irl{\substlist}} & = \lifz[\irl]{\partialeval{\irl{\expr[_{1}]}}{\irl{\substlist}}}{\\&\partialeval{\irl{\expr[_{2}]}}{\irl{\substlist}}}{\partialeval{\irl{\expr[_{3}]}}{\irl{\substlist}}} \\
-% \end{align*}
-\end{gather*}
-% \vspace{-3em}
-
-Note that both passes have no effect on the resulting trace of a program, up to $\emptyevent$-steps. 
-Because of this, both passes have equality as corresponding cross language trace relation. 
-Moreover, it is straightforward to prove both passes as secure (\Cref{def:rtp}) w.r.t. \gls*{ms}.
-
-\begin{theorem}[$\ccdce$ secure w.r.t. \gls*{ms}]\label{thm:ccdce:rtp:ms}
-  $\rtp{\ccdce}{\mssafe}$ %\Coqed
-\end{theorem}
-\begin{theorem}[$\cccf$ secure w.r.t. \gls*{ms}]\label{thm:cccf:rtp:ms}
-  $\rtp{\cccf}{\mssafe}$ %\Coqe
-\end{theorem}
-
-With both \Cref{thm:ccdce:rtp:ms,thm:cccf:rtp:ms} it follows from \Cref{corr:swappable} that the two passes can be interchanged arbitrarily:
-
-\begin{theorem}[$\cccf\circ\ccdce$ and $\cccf\circ\ccdce$ are secure w.r.t. \gls*{ms}]\label{thm:cccfccdce:rtp:ms}
-$\;$
-
-  \begin{nscenter}
-  \phantom{and~}\!\!$\rtp{\cccf\circ\ccdce}{\mssafe}$
-
-  and~$\rtp{\ccdce\circ\cccf}{\mssafe}$ % \Coqed
-  \end{nscenter}
-\end{theorem}
-
-\subsection{Robust Strict Cryptographic Constant Time Preservation}\label{subsec:cs:scct}
-
-This section defines a compiler $\ccscct$ from $\Lms$ to $\Lscct$ that robustly preserves \gls*{scct}. 
-Given the fact that $\Lscct$ provides a \gls*{cct}-mode that can be turned on or off, the compiler inserts wrapper code for function calls and function bodies to ensure that execution in the component always happen in \gls*{cct}-mode.
-This simple flag combines the effect of FaCT~\cite{cauligi2019fact}
-
-\vspace{-1em}
-\[
-\begin{array}{rcl}
-  \ccscct(\irl{\lfunction{g}{x}{\expr}}) & = & \lfunction[\obj]{\obj{g}}{\obj{x}}{\obj{\lwrdoit{ON};}\ccscct(\irl{\expr})} \\
-  \ccscct(\irl{\lcall{g}{\expr}}) & = & \lcall[\obj]{\obj{g}}{\ccscct(\irl{\expr})\obj{; \lwrdoit{ON}}} 
-    % \\
-    % \ccscct(\irl{\lbinop{\expr[_{1}]}{\expr[_{2}]}}) & = \lbinop[\obj]{\ccscct{\irl{\expr[_{1}]}}}{\ccscct{\irl{\expr[_{2}]}}} 
-    % \\
-\end{array}
-\]
-% \vspace{-2em}
-%
-The context can overwrite the flag and exit the mode, but upon invoking a function that is part of the component, the flag is set again.
-Because of this, the corresponding cross-language trace relation $\sim^{\Lms}_{\Lscct}$, only relates events without secrets:%
-
-\begin{center} 
-  \typerulenolabel{xrel:scct:noleak}{}{\irl{\preevent;\comp}\sim^{\Lms}_{\Lscct}\obj{\preevent;\comp;\unlock}}
-\end{center}
-
-The compiler is secure w.r.t. \gls*{scct}: %, similarly proven as in \Cref{subsec:cs:tms}.
-
-\begin{theorem}[$\ccscct$ secure w.r.t. \gls*{scct}]\label{thm:ccscct:rtp:scct}
-  \small
-  $\rtpsim{\ccscct}{\scctsafe}{\sim^{\Lms}_{\Lscct}}$ % \Coqed
-\end{theorem}
-
-\subsection{Robust Speculative Safety Preservation}\label{subsec:cs:ss}
-
-This section defines the final compilation pass $\ccspec$, which ensures that $\Lscct$ programs, which are assumed to be \gls*{ss}, stay \gls*{ss} at $\Lspec$-level. 
-To do so, $\ccspec$ inserts a speculation barrier after branches, which is sufficient to harden the program against speculative leaks, since SPECTRE-PHT~\cite{kocher2019spectre} is the only speculative leak modeled in the semantics of $\Lspec$.
-
-\vspace{-1em}
-\[
-\begin{array}{cl}
-  &\ccspec{(\obj{\lifz{\expr[_0]}{\expr[_1]}{\expr[_2]}})} = 
-  \\
-  &\qquad\qquad \lifz[\ird]{\ccspec{\left(\obj{\expr[_0]}\right)}}{\ird{\lbarrier;}\ccspec{\left(\obj{\expr[_1]}\right)} \\&\qquad\qquad}{ \ird{\lbarrier;}\ccspec{\left(\obj{\expr[_2]}\right)}} 
-\end{array}
-\]
-% \vspace{-2em}
-%
-
-Clearly, the corresponding cross-language trace relation $\sim^{\Lscct}_{\Lspec}$ has only one non-trivial case: for branches, only relate them where speculation is blocked by a barrier:
-
-\begin{center}
-  \typerulenolabel{xrel:spec:if}{}{\obj{Branch\ n}\sim^{\Lscct}_{\Lspec}\ird{Branch\ n}\cdot\ird{Spec}\cdot\ird{Barrier}\cdot\ird{Rlb}}
-\end{center}
-
-The base-event relation above scales to full events by ensuring the missing annotations ($\obj{\comp;\securitytag{}}$ and $\ird{\comp;\securitytag{}}$) are the same.
-% 
-With this relation, we prove that $\ccspec$ is secure with respect to \gls*{ss}.
-\begin{theorem}[$\ccspec$ secure w.r.t. \gls*{ss}]\label{thm:ccspec:rtp:spec}
-  \small$\rtpsim{\ccspec}{\sssafe}{\sim^{\Lscct}_{\Lspec}}$ % \Coqed
-\end{theorem}
-
-\subsection{Robust Preservation of Memory Safety, Strict Cryptographic Constant Time, and Speculative Safety}
-
-Finally, this subsection combines all previous results into one compilation chain to get that it preserves full \gls*{specms}.
-Let $\ccspecms$ be the compiler that is the composition of $\cca$, $\ccb$, $\cccf$, $\ccdce$, $\ccscct$, and $\ccspec$. 
-Let $\sim^{\Ltms}_{\Lspec}$ be the composition of $\sim^{\Ltms}_{\Lms}$, $\sim^{\Lms}_{\Lscct}$, and $\sim^{\Lscct}_{\Lspec}$.
-Then, the following corollary holds.
-
-\begin{corollary}[$\ccspecms$ secure w.r.t. \gls*{specms}]\label{thm:ccall:rtp:specms}
-  $\;$
-
-  \begin{nscenter}
-    $\rtpsim{\cc{\Ltms}{\Lspec}}{\mssafe\cap\scctsafe\cap\sssafe}{\sim^{\Ltms}_{\Lspec}}$ % \Coqed
-  \end{nscenter}
-\end{corollary}
-
-As with \Cref{corr:ccab:rtp:ms}, it is important to ensure that the respective cross language trace relations are well-formed (\Cref{def:wfc:sig:tracerel}).
-It is already known that $\sim^{\Ltms}_{\Lms}$ is well-formed with respect to $\mssafe$ (\Cref{lem:wf:ltmsltrg}).
-Next in the chain is $\sim^{\Lms}_{\Lscct}$, which has to be well-formed w.r.t. $\scctsafe$.
-This lemma holds, since a trace that was $\scctsafe$ is $\scctsafe$ even after applying $\sim^{\Lms}_{\Lscct}$: the relation enforces that $\Lscct$ traces related to $\Lms$ traces have no leaks of secrets whatsoever.
-
-\begin{lemma}[$\sim^{\Lms}_{\Lscct}$ well-formed w.r.t. $\scctsafe$]\label{lem:wf:lsmslscct}
-$\;$
-
-  \begin{nscenter}
-  $\wfcsig{\sim^{\Lms}_{\Lscct}}{\scctsafe}$
-  \end{nscenter}
-\end{lemma}
-
-The last relation is $\sim^{\Lscct}_{\Lspec}$ which needs to be well-formed w.r.t. $\sssafe$.
-Similarly to the previous relation, this holds, since $\sim^{\Lscct}_{\Lspec}$ only relates $\Lspec$ traces, which do not have speculative leaks, with $\Lscct$ traces.
-
-\begin{lemma}[$\sim^{\Lscct}_{\Lspec}$ well-formed w.r.t. $\sssafe$]\label{lem:wf:lscctlspec}
-  $\wfcsig{\sim^{\Lscct}_{\Lspec}}{\sssafe}$
-\end{lemma}
-
-\section{Formal Insights}\label{sec:formalities}
-
-This section discusses how to connect each language-specific security property to the general security properties of \Cref{sec:compprop} (\Cref{subsec:formalities:maps}), and it demonstrates that the security property resulting from the universal image projection is faithful to the original property (\Cref{subsec:formalities:props}).
-% 
-Then, this section discusses why the order of compiler passes matters, and how does our framework help with identifying insecure compositions (\Cref{subsec:compatsecpasses}), and it gives additional technical insights on the secure compilation proofs (\Cref{subsec:seccompproofs}).
-
-\subsection{From Language Traces to General Ones}\label{subsec:formalities:maps}
-
-The previous theorems talk about preserving properties expressed in the trace model of the languages of each compiler.
-However, these trace models are not the same trace model we used to specify the properties of \Cref{sec:compprop} (indicated with \ev{L}), which serves as the ``ground truth'' for the meaning of our properties.
-% 
-To bridge this gap, the formal development requires specifying additional trace relations, from each of the language trace models to the \ev{L} one, that, for example, relate $\src{Get\ \loc\ n}$ and $\src{Set\ \loc\ n}$ to $\ev{Use\ \loc\ n}$ (and that induce a related universal image that we use in \Cref{subsec:formalities:props}).
-% 
-One key insight of these relations is that they all omit context-made actions for two reasons: (1) contexts (which are universally quantified) can trivially invalidate any property and (2) we are interested in the component upholding the properties.
-%
-
- % disregarded a glaring but minor technical issue: take $\tmssafe$ (\Cref{def:trace:tmsdef}) as an example.
-% It contains traces as specified in \Cref{subsec:basic:memsafety:tracemodel}.
-% However, these are different from $\Ltms$ traces (\Cref{subsec:ltms}) which means that \Cref{thm:wt:tms} is ``ill-typed''.
-% Similarly for the other theorems.
-% In the interest of reducing visual baggage, the notation left out yet another cross-language trace relation, translating language-specific traces into these non-language specific traces. 
-% In doing so, context-level actions are dropped and, e.g., both $\src{Get\ \loc\ n}$ and $\src{Set\ \loc\ n}$ are translated into $\ev{Use\ \loc\ n}$.
-
-\subsection{Security Properties and Their Meaning}\label{subsec:formalities:props}
-Each of the presented compilers use a cross-language trace relation, which is also used to translate the property from one language to the other one (via the existential or universal images).
-% 
-%Whenever such translations occur, it is important to check whether there has been a change of meaning.
-%Secure compiler engineers need to ensure that such translations of properties do not change meaning in a fundamental wrong way. 
-% lifting or lowering reasoning from target- to source-level or vice-versa,
-While the meaning of projected properties does change with a translation, the change should not allow for a flawed compilation pipeline.
-% 
-For example, we could be using a trace relation that translates a property in a language to a totally different one in another language.
-% 
-To raise the trust into the translation of properties, \Cref{thm:prop-rel-corr} states (in a general fashion) that each security property is faithfully translated using the universal image according to the cross-language trace relation induced by the compiler.
-
-\begin{theorem}[Properties Relation Correctness]\label{thm:prop-rel-corr}
-  \begin{align*}
-    \forall& \pi \in \{\tmssafe,\smssafe,\scctsafe,\sssafe\}, 
-    \\
-    \text{ for each }& \text{pair of languages } \src{L} \text{ and } \trg{L} \text{ used by the compilers},
-    \\
-    \text{ if }& 
-    % \tau_{\sim^{\trg{L}}_{\ev{L}}}(\pi) \sim^{\trg{L}}_{\ev{L}} \pi
-    \sigma_{\sim^{\trg{L}}_{\ev{L}}}(\pi) \sim^{\trg{L}}_{\ev{L}} \pi
-    \text{ and } 
-    % \sigma_{\sim^{\src{L}}_{\trg{L}}}(\tau_{\sim^{\trg{L}}_{\ev{L}}}(\pi)) \sim^{\src{L}}_{\trg{L}} \tau_{\sim^{\trg{L}}_{\ev{L}}}(\pi)
-    \sigma_{\sim^{\src{L}}_{\trg{L}}}(\sigma_{\sim^{\trg{L}}_{\ev{L}}}(\pi)) \sim^{\src{L}}_{\trg{L}} \sigma_{\sim^{\trg{L}}_{\ev{L}}}(\pi)
-    \\
-    \text{ then }& 
-    % \tau_{\sim^{\src{L}}_{\ev{L}}}(\pi) = \sigma_{\sim^{\src{L}}_{\trg{L}}}(\tau_{\sim^{\trg{L}}_{\ev{L}}}(\pi)) 
-    \sigma_{\sim^{\src{L}}_{\trg{L}}}(\sigma_{\sim^{\trg{L}}_{\ev{L}}}(\pi)) \sim^{\src{L}}_{\ev{L}} \pi
-  \end{align*}
-\end{theorem}
-The complexity of this theorem is that the relation in the conclusion cannot be obtained by composing the two relations in the premises.
-% 
-%% MK: the following does not say anything about the theorem, only about the syntactic compiler
-%What happens in practice is that compiler engineers are given the relations from each language to \ev{L} and they need to build a compiler that induces a relation $\sim^{\src{L}}_{\trg{L}}$ that respects this theorem.
-%
-
-We now informally argue why this theorem holds for the composition of all considered properties from \Cref{sec:compprop}.
-
-\paragraph{$\sigma_{\sim^{\Ltms}_{\Lms}}\left(\tmssafe\cap\smssafe\cap\scctsafe\cap\sssafe\right)$}
-
-For the four properties considered here, the trace models of $\Ltms$, $\Ltrg$, and $\Lms$ do not consider actions related to $\scctsafe$ and $\sssafe$, so these two properties are trivially translated correctly.
-% 
-We now discuss the remaining $\tmssafe$ and $\smssafe$ in the form of their intersection $\mssafe$.
-
-Recall that $\sim^{\Ltms}_{\Lms}$ is the composition of $\sim^{\Ltms}_{\Ltrg}$ and $\sim^{\Ltms}_{\Lms}$.
-% 
-Since $\sim^{\Ltms}_{\Ltrg}$ is an equality, this relation trivially preserves the meaning of translated properties: related traces are identical!
-
-Finally, let us consider a trace $\irl{\trace}\in\mssafe$ and understand what is that is related to via $\sim^{\Ltrg}_{\Lms}$.
-% 
-{\em All} traces $\trg{\trace}$ with $\trg{\trace}\sim^{\Ltrg}_{\Lms}\irl{\trace}$ are identical to $\irl{\trace}$ except for get and set actions, which require for in-bound accesses (as stated in \Cref{subsec:cs:ms}): this clearly respect $\mssafe$.
-%
-
-% In both cases, the translated event does not represent an out-of-bounds memory access.
-
-% \footnote{Both $\sim^{\Ltms}_{\Ltrg}$ and $\sim^{\Ltrg}_{\Lms}$ filter context actions.}.
-
-% Now, starting from $\mssafe$ expressed in $\Lms$, we check that its translation in $\Ltms$ is still $\mssafe$.
-% % 
-% We can see that the property resulting from the projection of $\Lms$ traces back to $\Ltms$ faithfully models $\mssafe$, which translates to both $\scctsafe$ and $\sssafe$, 
-% \MPin{ equiv  +  other rel only adds traces that fail }
-% since these properties trivially hold for $\Ltms$, $\Ltrg$, and $\Lms$ traces.
-
-\paragraph{$\sigma_{\sim^{\Lms}_{\Lscct}}\left(\tmssafe\cap\smssafe\cap\scctsafe\cap\sssafe\right)$}
-
-As before, the trace models of $\Lscct$ and $\Lms$ cannot express $\sssafe$, so that is trivially translated correctly.
-% 
-Also, the trace model of $\Lscct$ extends the one of $\Lms$ with respect to $\tmssafe$ and $\smssafe$ events, so the translation argument regarding those two properties is the same as before.
-% 
-Thus, we need to reason about whether $\scctsafe$ is translated correctly.
-
-By definition, $\sim^{\Lms}_{\Lscct}$ only relates $\obj{\unlock}$ events, which is also the same relation induced by $\sim_{\ctsafe}$ in \Cref{sec:msscct-rel}.
-% 
-This ensures that composing the relations only relates $\unlock$ events, and thus the property is translated correctly.
-
-% Consider a trace $\obj{\trace}\in\scctsafe$ and all $\irl{\trace}$ such that $\irl{\trace}\sim^{\Lms}_{\Lscct}\obj{\trace}$. 
-% In turn, $\obj{\trace}$ cannot contain any leaks of secret values, so it cannot violate $\scctsafe$.
-% Toggling data-independent timing mode has no impact on $\mssafe$ or $\sssafe$.
-
-\paragraph{$\sigma_{\sim^{\Lscct}_{\Lspec}}\left(\tmssafe\cap\smssafe\cap\scctsafe\cap\sssafe\right)$}
-The trace model of $\Lspec$ extends the one of $\Lscct$ with respect to $\tmssafe$, $\smssafe$, and $\scctsafe$, so the translation argument for those three properties is the same as before.
-% 
-Concerning $\sssafe$, $\sim^{\Lscct}_{\Lspec}$ relates $\Lspec$ speculation traces with $\Lscct$ branches with the $\lock_{\text{NONE}}$ tag, so the property is translated correctly, according to the relation defined in \Cref{sec:spec-ms-rel}.
- % introducing $\ird{barrier}$ never violates $\sssafe$ and has no effect on $\mssafe$ or $\scctsafe$.
-
-%\MPin{
-%	can this be made formal?
-%	what's the point?
-%}
-
-\subsection{Compatibility of Secure Compiler Passes}\label{subsec:compatsecpasses}
-
-% \Cref{thm:rtpsim:sig} is applicable, up to the well-formedness of $\sim_1$ w.r.t. $\trg{\class[_2]}$.
-Consider applying $\ccspec$ first and then $\ccb$ (albeit currently syntactically impossible).
-% , let us pretend that it is possible by just extending $\ccb$ for the constructs present in $\Lscct$ that are not present in $\Ltrg$.
-% Clearly, as evidenced earlier (\Cref{subsec:formalities:props}), $\ccb\cdot\ccspec$ is an acceptable composition that yields an interesting class of security properties, i.e., $\sigma_{\sim_{\irl{ms}}\bullet\sim_{\ird{\mathghost}}}\left(\mssafe\cap\scctsafe\cap\sssafe\right)$.
-In this case, $\ccb$ would insert new branches into the code that are not protected by a speculation barrier! 
-This concern is reflected in the proofs that establish that the security class resulting of the composition of the trace relations is meaningful.
-% , much similar to ensuring that a definition models what the formal methods expert intends to model. 
-For the $\ccspec\cdot\ccb$ case, the class is $\sigma_{\sim_{\ird{\mathghost}}\bullet\sim_{\irl{ms}}}\left(\mssafe\cap\scctsafe\cap\sssafe\right)$.
-Since $\ev{Use\ \loc\ n}\sim_{\irl{ms}}\irl{Branch\ n}\cdot\irl{Spec}\cdots$, where $\cdots$ does not contain a $\ev{{Barrier}}$ event, the resulting class is \emph{not} the original $\sssafe$ that is intended and it would break the corresponding \Cref{thm:prop-rel-corr}.
-% Because of this, it is crucial to analyse the precise shape of the class of properties after mapping a target-level property to a source-level property\footnote{Dito for mapping a source-level property to a target-level property.}.
-
-However, the composition is still technically possible and it is the job of the compiler engineers to ensure that the secure compilation pipeline happens in an order that ensures that the mapped security property is the intended one.
-
-\subsection{Secure Compilation Proofs}\label{subsec:seccompproofs}
-
-Our secure compilation proofs rely on backtranslations~\cite{abate2019jour,patrignani2021rsc}, which let one construct a source context starting from either target traces (aka trace-based backtranslations) or target contexts (aka context-based backtranslations).
-These backtranslations also require setting up cross-language relations between various language elements such as expressions and program states, so we leave these details for the technical reports.
-% , the usual strategy is to establish a backwards simulation between source components and their compiled target counterpart. 
-% For the universally quantified target context, a technique known as backtranslation is well-established, which constructs a source-level context either by backtranslating the target-level context, amounting to a ``decompilation'', or by constructing it using the target-level trace, so that the source-context emits parts of the whole execution trace that are related to the parts emitted by the target-level context.
-All backtranslations in the case-study are trace based except for those required by $\ccdce$ and $\cccf$, which are context-based (and they are an identity function). 
-% Since these optimisations have no significant semantic effect and source- and target-language are equal, the backtranslation amounts to an identity translation.
-
-\section{Related Work\pages{2}}\label{sec:relwork}
-
-This section discusses robust compilation, other secure compilation criteria and work related to the properties preserved in the case study.
-% (\Cref{subsec:relw:seccomprtp}) 
-% (\Cref{subsec:relw:seccompcrit}).
-% Since the case study of \Cref{sec:casestud:defs,sec:casestud:rtp} implements measures for preserving \gls*{ms}, \gls*{cct}, and \gls*{ss}, 
-% Then, this section discusses work related 
-% relevant related work as well 
-% (\Cref{subsec:relw:msmechs,subsec:relw:cctmechs,subsec:relw:ssmechs}).
-
-\paragraph*{Secure Compilation as Robust Preservation}\label{subsec:relw:seccomprtp}
-
-The robust preservation of properties as a compiler-level criterion has been analysed extensively~\cite{abate2019jour,patrignani2021rsc,abate2021extacc,patrignani2019survey} and thus we build on that framework.
-No existing work is concerned with composing secure compilers, however, existing work~\cite{abate2021extacc} sketches composition of trace-relating compiler correctness in a similar way to what has been presented here.
-%Parts of these works consider languages with different trace models and our technical setup can handle this. 
-The work relating robust preservation with universal composability~\cite{patrignani2022universal} is closest to what this paper presents.
-% 
-The authors demonstrate a similar compositionality theorem to ours (\Cref{sec:sequential}) but use it in the context of protocols.
-% 
-The work does not consider the generality to support different trace models or composition of compilers which robustly preserve different classes.
-% The authors demonstrate a similar compositionality theorem to what is presented here (\Cref{sec:sequential}) as well as in an earlier version of this work~\cite{kruse2022csc}.
-% However, they do not demonstrate the scalability of the approach by means of a case study.
-
-%\paragraph*{Composition of Secure Compilers}\label{subsec:relw:seccompcomp}
-
-There is work on lifting exploits for single compilers to the whole chain~\cite{paykin2019weird}.
-%Our framework does not consider exploits.
-%\MP{ a bit dry and short}
-While that work considers {\em in}secure compilation and composition thereof in terms of exploits, the composition they are interested in allows to lift an exploit for one compiler pass to the whole compilation chain.
-Our framework takes the opposite direction and provides compositionality results for secure compilers.
-%
-%
-%To the best of our knowledge, there does not exist published work discussing the compositionality of the robust preservation criterion, the gold-standard of secure compilation.
-
-\paragraph*{Other Secure Compilation Criteria}\label{subsec:relw:seccompcrit}
-
-While this paper focuses on the robust preservation framework~\cite{abate2019jour}, other secure compilation criteria exist.
-The survey on formal approaches to secure compilation~\cite{patrignani2019survey} discusses a broad spectrum already, while this section presents a very high-level overview.
-Fully abstract compilation~\cite{abadi1999fullabstraction} states that a compiler should preserve and reflect observational equivalence between source and target programs.
-Abate \emph{et al.}~\cite{abate2021faandrc} showed that fully abstract compilers robustly preserve program properties that are either trivial or meaningless.
-As a mitigation for this, the authors presented a categorical approach based on maps of distributive laws~\cite{watanabe2002modl}, which they call many maps of distributive laws.
-Maps of distributive laws have been investigated before as a possible secure compilation criterion~\cite{tsampas2020catsc}.
-Other approaches are extensions of the compiler correctness criterion as discussed in other work~\cite{patterson2019next700} or the introduction of opaque observations~\cite{vu2021reconciling} to reconcile compiler optimisations with security.
-Note that this work also presents secure compilers that are optimising, but contrary to the other~\cite{vu2021reconciling}, provides a formal account of these in the robust preservation framework.
-% Lastly, the authors of this paper have presented ongoing work~\cite{patrignani2023blame} on a weaker robust preservation criteria based on the concept blame.
-
-\paragraph*{Memory Safety Mechanisms}\label{subsec:relw:msmechs}
-
-Different mechanisms for enforcing memory safety exist that also consider the secure compilation domain, i.e., have an active attacker model.
-For example, the ``pointers as capabilities'' principle represents pointers as machine-level capabilities~\cite{korashy2021capableptrs}, which behave in a similar fashion to capabilities by means of linear typing~\cite{morrisett2005L3}.
-The approach of this paper also uses linear typing, but differs from $L^{3}$~\cite{morrisett2005L3} in the way that functions are not first-class.
-Moreover, this paper considers an active attacker, while the work on $L^{3}$ only discusses whole programs and, thus, has no active attacker model.
-The instrumentation to ensure memory safety that this paper presents is inspired by Softbounds~\cite{nagarakatte2009soft}.
-That work inserts bounds-checks in front of pointer-dereferences and, for this to work, inserts meta-data information on pointer creation.
-Softbounds also works in a more advanced setting with structured fields accesses and also introduces a table-lookup for pointers that are stored in memory.
-This paper only considers arrays of primitive data, i.e., there are no pointers to pointers or structures.
-Several other approaches to memory-safety exist in literature, specifically as compiler instrumentations~\cite{akritidis2009baggy,younan2010paricheck,jung2021pico,shankaranarayana2023tailcheck,dhumbumroong2020boundwarden,nam2019framer,zhou2023fatptrs}, hardware-extensions~\cite{kwon2013lowfat,saileshwar2022heapcheck,chen2023flexpointer,kim2023whistle}, or programming language extensions~\cite{elliott2018checkedc,li2022formalcheckedc,jim2002cyclone,elliott2015guilt,west2005cuckoo,weis2019fyr,benoit2019uniqueness}.
-What differentiates this work from them is that this work uses known, compiler-based approaches to ensure memory-safety as a means to investigate secure compiler compositions.
-This paper does not provide efficient memory-safety, but serves as a theoretical foundation for the secure compilation domain.
-
-To extend the languages in this paper with a less restricted form of pointer arithmetic, the region colouring memory safety monitor presented in earlier work~\cite{michael2023mswasm} can be used.
-The work presenting this monitor provides an approach for the robust preservation of memory safety compiling from C to WASM.
-However, they do not discuss composition of secure compilers but rather investigate an instance of a secure compiler.
-
-\paragraph*{Cryptographic Constant Time Mechanisms}\label{subsec:relw:cctmechs}
-
-The approach to preserving cryptographic constant time in this paper is high-level, where a programming language exposes a way to switch the semantics to a data (operand) independent timing mode.
-Since identifiers in $\Lscct$ are annotated with a secrecy tag, this approach is similar to others with information flow control.
-For example, Vale~\cite{bond2017vale} uses Dafny to ensure constant-time assembly code, while Jasmin~\cite{almeida2017jasmin} makes use of the Coq proof assistant to reject non-constant-time programs.
-CT-Wasm~\cite{watt2019ctwasm} enforces constant-timeness by means of a type system.
-Different to the approach of this paper, these approaches necessitate that the programmer writes \gls*{cct} code.
-An approach to allow programmers to write more high-level code is CryptOpt~\cite{kuepper2023cryptopt}, which generates efficient target-code by means of a randomised search.
-This paper abstracts over concrete mitigation strategies and simply assumes that there is a flag to switch to a cryptographic-constant time execution mode.
-This can be realised by employing the FaCT~\cite{cauligi2019fact} compiler, which translates common non-constant time code patterns to be constant-time, and the data (object) independent timing execution mode of modern processors.
-
-\paragraph*{Speculation Safety Mechanisms}\label{subsec:relw:ssmechs}
-
-This paper uses a taint-tracking mechanism inspired by existing work~\cite{guarnieri2018spectector,fabian2022automatic}.
-These taints are used to express absence of any speculative leaks in \gls*{ss}~\cite{guarnieri2018spectector}. 
-The semantics of $\Lspec$ hardcodes the kind of speculative leaks to just SPECTRE-PHT~\cite{kocher2019spectre}, but future work could use semantics composition~\cite{fabian2022automatic} to support more variants.
-Note that our framework composes compilers and not semantics.
-
-%This work has taken inspiration in the sematics of $\Lspec$ and in \Cref{def:trace:ss} from recent formalisations~\cite{guarnieri2018spectector,fabian2022automatic}.
-% \MPin{
-%  we take SS from exorcising.
-%  we do not take SNI from spectector because it is a hypersafety.
-%  other work defined similar props. 
-% }
-% A subsequent line of work~\cite{fabian2022automatic} shows how to modularly define and compose formal semantics that models different speculative execution variants. 
-% Recent work~\cite{fabian2024lift} demonstrates how to lift existing secure compilation proofs for security properties related to speculative execution to more powerful properties, i.e., properties defined for languages with a trace model that enables more vulnerabilities, such as combinations of different speculative execution exploits.
-% To this end, they define a lifting framework and demonstrate how a secure compilation proof for a weaker property can be lifted to a stronger one. 
-% Contrary to the compositionality framework presented in this paper, these works are only concerned with composing semantics of the respective programming languages, while in this paper, the secure compilers themselves are composed.
-
-\section{Conclusion\pages{$\sfrac{1}{2}$}}\label{sec:concl}
-This paper tackles the problem of understanding what kind of security properties a secure compiler preserves, when said compiler is the combination of compiler passes that preserve possibly different security properties.
-% 
-%For this, this paper first formalised security properties of interest and their composition.
-% 
-The paper proves that composing secure compilers that preserve certain properties results in a secure compiler that preserves the composition of these properties.
-% While the presented security property composition relied on monitors that check only trace-properties, the composition of secure compilers does not make any restriction towards the kind of property involved in the composition.
-% It is subject to future work to develop techniques to verify relational hyperproperties~\cite{abate2019jour}, while the composition of hyperproperties could be very similar to the composition of ordinary properties as presented in this paper, since hyperproperties can be checked with an automata-based model-checker~\cite{beutner23hyperltl}.
-% 
-Finally, this paper defines a multi-pass compiler and proves that it preserves \gls*{specms}.
-Crucially, this paper derives the security of the multi-pass compiler from the composition of the security properties preserved by its individual passes, which include security-preserving as well as optimisation passes.
-% For future work, it is interesting to look at more sophisticated optimisation passes that, e.g., reorder events that appear on traces.
-
-\newpage
-
-\bibliographystyle{IEEEtranS}
-\bibliography{main}
-
-%%
-%% If your work has an appendix, this is the place to put it.
-\appendix
-\subsection{Existential Image}\label{subsec:extimg}
-
-\begin{definition}[Existential Image]
-% [Existential and Universal Image]
-\label{def:existential:img}\label{def:tau}
-  \[
-    \tau_\sim\left(\src{\pi}\right) := 
-      \left\{ 
-        \trg{\trace} \mid \exists \src{\trace}\ldotp \src{\trace}\sim\trg{\trace}, \text{ and } \src{\trace}\in\src{\pi} 
-      \right\}
-  \]
-  %\[ 
-  %  \sigma_\sim\left(\trg{\pi}\right) := 
-  %    \left\{ 
-  %      \src{\trace} \mid \forall \trg{\trace}\ldotp \text{if }\src{\trace}\sim\trg{\trace}, \text{ then } \trg{\trace}\in\trg{\pi} 
-  %    \right\}
-  %\]
-\end{definition}
-
-\begin{definition}[Robust Preservation with $\tau_\sim$]\label{def:rtp:tau}
-  % $\;$\\
-  % \vspace{-1em}
-  % \begin{nscenter}
-  % Compiler $\cc{\src{L}}{\trg{L}}$ robustly preserves $\class$, 
-  {$\rtptau{\cc{\src{L}}{\trg{L}}}{\src{\class}}{\sim}$}
-  %, iff 
-  $\isdef$
-    {$\forall \src{\pi}\in\src{\class}, \src{p}\in\src{L},$} if {$\rsat{\src{\progvar}}{\src{\pi}}$}, then {$\rsat{\cc{\src{L}}{\trg{L}}\left(\src{p}\right)}{\tau_\sim\left(\src{\pi}\right)}$}.
-  % \end{nscenter}
-\end{definition}
-
-\begin{theorem}[Composition of Secure Compilers w.r.t. $\tau$]\label{thm:rtpsim:tau}
-  $\;$
-
-  If {$\rtptau{\cc{\src{L}}{\trg{L}}}{\src{\class[_{1}]}}{\sim_1}$}, {$\rtptau{\cc{\trg{L}}{\obj{L}}}{\tilde{\tau}_{\sim_1}\left(\src{\class[_2]}\right)}{\sim_2}$}, and {$\wfctau{\sim_1}{\src{\class[_2]}}$}, \\ then {$\rtptau{\cc{\src{L}}{\trg{L}}\circ\cc{\trg{L}}{\obj{L}}}{\src{\class[_{1}]}\cap\src{\class[_{2}]}}{\sim_1\bullet\sim_2}$}.
-\Coqed
-\end{theorem}
-
-\begin{corollary}[Swapping Secure Compiler Passes]\label{corr:swappable:tau}
-  $\;$
-
-  If {$\rtptau{\cc[_{1}]{\trg{L}}{\trg{L}}}{\src{\class[_{1}]}}{\sim_1}$ and $\rtptau{\cc[_{2}]{\trg{L}}{\trg{L}}}{\src{\class[_{2}]}}{\sim_2}$}, %
-  {$\wfctau{\sim_1}{\src{\class[_2]}}$ and $\wfctau{\sim_2}{\src{\class[_1]}}$}, %
-  and {$\tilde{\tau}_{\sim_1}\left(\src{\class[_2]}\right)=\src{\class[_2]}$ as well as $\tilde{\tau}_{\sim_2}\left(\src{\class[_1]}\right)=\src{\class[_1]}$},
-  then {$\rtptau{\cc[_{1}]{\trg{L}}{\trg{L}}\circ\cc[_{2}]{\trg{L}}{\trg{L}}}{\src{\class[_{1}]}\cap\src{\class[_{2}]}}{\sim_1\circ\sim_2}$ and $\rtptau{\cc[_{2}]{\trg{L}}{\trg{L}}\circ\cc[_{1}]{\trg{L}}{\trg{L}}}{\src{\class[_{2}]}\cap\src{\class[_{1}]}}{\sim_2\circ\sim_1}$}.
-\Coqed
-\end{corollary}
-
-\subsection{Secure Upper and Lower Composition}\label{sec:other-compos}
-Besides sequential composition, there are two other compositions, namely an {\em upper}, i.e., a compiler that takes multiple inputs and yields one output, and a {\em lower} composition, i.e., a compiler that takes one input and yields multiple outputs.
-We {define the upper composition $\cc{\src{L}+\obj{L}}{\trg{L}}$} as follows:
-Given a program \texttt{p}, its compiled counterpart is obtained by {plugging \texttt{p} into $\cc{\src{L}}{\trg{L}}$ if $\texttt{p}\in\src{L}$} or by {plugging \texttt{p} into $\cc{\obj{L}}{\trg{L}}$ if $\texttt{p}\in\obj{L}$}.
-\begin{definition}[Upper Composition]
-  \[
-    \text{{$\cc{\src{L}+\obj{L}}{\trg{L}}$}}\isdef
-  % Given \texttt{p}, yield
-  \lambda \texttt{p}\ldotp
-  \left\{\mbox{\begin{tabular}{c}
-    {{if $\texttt{p}\in\src{L}$, then $\cc{\src{L}}{\trg{L}}(\texttt{p})$}} \\
-    \mbox{{if $\texttt{p}\in\obj{L}$, then $\cc{\obj{L}}{\trg{L}}(\texttt{p})$}} \\
-  \end{tabular}}\right. 
-  \]
-\end{definition}
-
-Examples of this are present in industry:
-Consider the Java Virtual Machine bytecode $\trg{JVM BC}$, which is a popular target for programming language designers due to its high performance and relevance in industry.
-Compilers for several programming languages have it as their target language, some popular instances are $\src{Java}$ and $\obj{Kotlin}$.
-Technically speaking, they both compile to class files and $\obj{Kotlin}$ objects are considered to be the same as $\src{Java}$ objects at that point.
-Both languages can be used at the same time in one project~\cite{androidstudio}.
-A compiler that accepts both $\src{Java}$ and $\obj{Kotlin}$ code translating to the same target language or intermediate representation performs a kind of {\em upper} composition.
-Now, the following theorem tells us what happens if these are secure:
-Given {$\cc{\src{L}}{\trg{L}}$ robustly preserves $\class[_{1}]$} and {$\cc{\obj{L}}{\trg{L}}$ robustly preserves $\class[_{2}]$}, it follows that {their upper composition $\cc{\src{L}+\obj{L}}{\trg{L}}$ robustly preserves the intersection of classes $\class[_{1}]$ and $\class[_{2}]$}.
-
-\begin{theorem}[Upper Composition of Secure Compilers]\label{thm:urtp}
-  $\;$
-
-  If {$\rtp{\cc{\src{L}}{\trg{L}}}{\class[_{1}]}$} and {$\rtp{\cc{\obj{L}}{\trg{L}}}{\class[_{2}]}$}, then {$\rtp{\cc{\src{L}+\obj{L}}{\trg{L}}}{\class[_{1}]\cap\class[_{2}]}$}. %\Coqed
-\end{theorem}
-
-Dually, the {\em lower} composition is concerned about compilers that accept the same source but yield different target languages. %and the same theoretical results apply.
-{Define the lower composition $\cc{\src{L}}{\trg{L}+\obj{L}}$} as follows:
-Given a program $\src{p}$, its compiled counterpart is obtained by {plugging $\src{p}$ into $\cc{\src{L}}{\trg{L}}$} or by {plugging \texttt{p} into $\cc{\src{L}}{\obj{L}}$}, respectively, {based on the internal decision}.
-\begin{definition}[Lower Composition]
-  \[
-    \text{{$\cc{\src{L}}{\trg{L}+\obj{L}}$}}\isdef
-  % Given $L$ and $\src{p}$, yield
-  \lambda \src{p}, L\ldotp
-  \left\{\mbox{\begin{tabular}{c}
-    {{if $L=\trg{L}$, then} {$\cc{\src{L}}{\trg{L}}(\src{p})$}} \\
-    \mbox{{if $L=\obj{L}$, then} {$\cc{\src{L}}{\obj{L}}(\src{p})$}} \\
-  \end{tabular}}\right.\]
-\end{definition}
-
- Consider two compilers both accepting $\src{LLVM IR}$~\cite{lattner2004llvm} and one of them emits $\trg{x86\_64}$, while the other emits $\obj{ARMv8}$.
- It is intuitive that they are in some sense composed in the LLVM framework, but the decision of when to use one over the other is inherently {\em internal} to the formalisation effort of this kind of composition.
- For example, the user of this compiler provides an explicit flag that instructs to emit $\trg{x86\_64}$ or the framework itself detects the target platform via heuristics, such as supported instructions.
-
- The following theorem demonstrates what happens if the involved compilers are secure:
-Given {$\cc{\src{L}}{\trg{L}}$ robustly preserves $\class[_{1}]$} and {$\cc{\src{L}}{\obj{L}}$ robustly preserves $\class[_{2}]$}, it follows that {their lower composition $\cc{\src{L}}{\trg{L}+\obj{L}}$ robustly preserves the intersection of classes $\class[_{1}]$ and $\class[_{2}]$}.
-
-\begin{theorem}[Lower Composition of Secure Compilers]\label{thm:lrtp}
-  $\;$
-
-  If {$\rtp{\cc{\src{L}}{\trg{L}}}{\class[_{1}]}$} and {$\rtp{\cc{\obj{L}}{\trg{L}}}{\class[_{2}]}$}, then {$\rtp{\cc{\src{L}}{\trg{L}+\obj{L}}}{\class[_{1}]\cap\class[_{2}]}$}. % \Coqed
-\end{theorem}
-
-Either way, the theoretical results suggest that it is possible to always find a ``most-general'', secure compiler, given two secure compilers, that robustly preserves the least-upper bound of the classes involved in their compilation process.
-
-%Future work intends to use upper and lower composition as means to build a secure compiler incrementally
-%This follows current non-formal-methods based compiler engineering, such as LLVM~\cite{lattner2004llvm}, where languages such as \texttt{C++} and \texttt{Fortran} translate into LLVM IR and languages such as \texttt{x86} or \texttt{ARMv8} assembly variants translate out of LLVM IR.
-
-% no coq, because above definitions are non-formal. 
-%   How would you precisely define the lower compo compiler, given two compilers?
-
-%\section{Cryptographic Constant Time}\label{sec:cct}
-
-\end{document}
-\endinput
